### 2020-06-15 Code/Rmd/data_analysis_notes.Rmd
First commit

---
title: "Notes on data anlaysis"
output: html_notebook
---

The aim of this study is two-fold. First, we want to test whether word-forms are acquired earlier by bilinguals when their translation equivalent is form-similar. Second, we want to test if the difference in age of acquisition between both translation equivalents is smaller when they are form-similar.

In order to gather information about acquisition of vocabulary, many developmental studies rely on vocabulary inventories such as the McArthur-Bates Communicative Developmental Inventories (*CDI* from now on). These questionnaires are administered to parents or caretakers, which are presented with a printed list of words. For each word, parents must report whether they think the child understands and/or produces the word. Age of acquisition (AoA) is then defined as the age at which the proportion of children that are able to understand or produce a given word exceeds a given threshold. For instance, the word *door* may be considered to be aquired when at least 60% of the participants in our sample are reported to understand the word.

We find two reasons to consider this approach suboptimal. First, the value of the cut-off is arbitrary. Although most studies use thresholds between 50% and 60%, this varibility makes it difficult to build cumulative knowledge from the listerature when the un-aggregated data (not yet filtered by the cut-off) is available. Second, it is unlikely that the acquisition status of a word is well represented by a dicotomy between acquired and not acquired. Word learning is a multidomain process where many processes take place. It is more likely that the trajectory of acquisition of a given word is better defined by a curve that characterises the probability of a child being able to understand or produce a word across ages. Third, by considering AoA as a point estimate, other dimensions that characterise the trajectory of aquisition are being neglected. For instance, it seems reasonable to think that some words are acquired faster the others. This means that the developmental curve of some words is steeper that that of others. For instance, the effect of word frequency on AoA has been extensively: more frequent words are aquired earlier then less frequent words. We suggest that this effect could also be reflected on steeper acquisition trajectories.

Recent studies have explored the role of word frequency, as well as other well know predictors (e.g., concreteness) on age of acquisition by going beyond arbitraty cut-offs.

Description of @braginsky2019 and @jones2019.

Both studies have made substantial contributions to the study of the internal structure of developing lexicon by performing item-wise analysis. However, neither of them adresses the potential non-linearity of the developmental trajectory of words. By fitting a linear (multilevel) model, two main parameters of interest are estimated. Both parameters characterise the line that defines the effect of each predictor (e.g., age, frequency) on the outcome variable (e.g., proportion of children reported the understand the word, probability that a given child will be reported to undestand the word). A first parameter is the intercept, which provides information about the value of the outcome when all predictors are at baseline. A second set of parameters informs us about the contribution of each of the predictor or combination of predictors. This model assumes that *age* holds a linear relationship with the proportion of infants that have acquired the word. We consider that this assumed linear relation does not capture well enought the consitribution of *age* on the acquisition status of the word. Instead, we propose that this relationship is better defined by logarithmic curves.

Logarithmic curves are characterised by three parameters: 1) Upper asymptote, 2) inflexion point, and 3) steepness at inflexion point. First, although it seems reasonable to assume that, as age increases, the proportion of children that have acquired a given word increases as well. However, it is most likely that such word will never have been acquired by all children in the population. This asymptotic nature of the acquisition status of the words is capture by the first parameter indicated above. Second, previous studies have reported that vocabulary size (both receptive and productive) remains relatively stable during the first year and a half of life, and then grows exponentially until one year after, when it stabilises again. Later, vocabulary size keeps growing, but a slower pace. This means that, around 20 to 24 months of age, the developmental trajectory of most of the words that will be part of the vocabulary of the child during the next 3 years, will find an inflexion point, where the curve is steepest. If this is to be true, it could be reasonable to consider this point as the critical age at which the acquisition status of a given word is evolving faster. We argue that this age point reflects overlaps better with what most researchers mean by *word-acquistion*, compared to the tradicitonal cut-off approach. Third, some items may differ on how steep their acquisition trajectory is in this inflexion point. For instance, hight-frequency words may entail a steeper curve than low_frequency words, thus leading to high-frequency words to be acquired faster than low-frequency word once the inflexion point is reached.

We consider that logarithmic curves are more suitable to developmental trajectories of individual items, because they capture cross-item variability in age of acquisition and speed of acquisition, at the same time they account for non-linear trends in such trajectories, as well as for the existence of an upper asymptotic boundary in the proportion of infants that have acquired each word.

### Model specification

In this study, we model developmental tracjectories of individual items using logarithmic curves. We assume a common upper boundary for all items (i.e., 1). We aim to estimate the inflexion point (i.e., age at which the curve is steepest) and the steepness (i.e., how fast the proportion of infants that know the word increases) at the inflexion point. We will follow a multilevel Bayesian approach to estimate the mid point and steepness of each word. We assume that mid-points will be generated from a normal distribution with mean 7.5 (age bin corresponsing to 22 months of age) and standard deviation 1. This means that the inflexion point (i.e., age of maximum steepness), 68% of the items will be located within 1 stand deviation at 22 months. Our model will update collected data to update the estimated mid-point of the words, by using word frequency, cognate status, and language profile to predict the mean of the distribution. We will include by-item random intercepts, and by-item random slopes for cognate status and language profile. We will include a correlation parameter for each grouping variable (i.e. cognate status and language profile), but we won't allow correlation across grouping variables. We assume that the steepness of each item will be generated from a normal distribution with mean 1 and standard deviation 0.5. This means that for 68% of the items, the proportion of children that understand a given word will be close 1 standard deviation  far maximum from in a 100% at the age of maximum steepness. Our model can be described as follows:

$$Y_{ij} = \frac{1}{1 + e \exp{-k_i(Age-x_{0ij})}}$$
$$Y_{i} \sim Binomial(p, N)$$
$$k \sim Normal(1, 0.5)$$
$$x_{0ij} \sim N{\beta_0 + \beta_{0j} + (\beta_1 + \beta_{1j}) \times Profile_{ij} + (\beta_2 + \beta_{2j}) \times Cognateness_{ij}, \sigma^2}$$
$$\beta_0 \sim Normal(0, 1)$$
$$\beta_{0j} \sim Normal(0, \tau^2_{0})$$
$$\beta_{1j} \sim Normal(0, \tau^2_{1})$$
$$\beta_2 \sim Normal(0, 1)$$
$$\beta_{2j} \sim Normal(0, \tau^2_{2})$$

Where:

* $Y_{ij}$ is the proportion of children that have been reported to have acquired the item $i$ at age $j$; assumed to be distributed following a Gaussian distribution with mean 0 and variance $\sigma^2$.
* $k_i$ is the steepness of the curve at the inflexion point of item $i$, assumed to be distributed following a Gaussian distribution with mean 1 and standard deviation 0.5.
* $x_{0ij}$ is the age at which the inflexion point is located in item $i$ and cognate status $j$. This parameter is defined by a linear model the includes the following parameters:
     * $\beta_0$ is the intercept of the model, that is, the grand mean age of of infants when the curve of item $i$ is at its steepest point, when all other predictors are at baseline; it is assumed to be distributed following a Gaussian distribution with mean 7.5 (age bin corresponding at 22 months of age) and standard deviation 4.
     * $\beta_{0j}$ is the intercept of the model, that is, the grand mean age of of infants when the curve of item $i$ is at its steepest point, when all other predictors are at baseline; it is assumed to be distributed following a Gaussian distribution with mean 7.5 (age bin corresponding at 22 months of age) and standard deviation $\tau^2_{0}$.
     * $\beta_1$ is the slope of *language profile* on the outcome; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation 1.
     * $\beta_{1j}$ is the slope of *language profile* on cognate status $j$; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation $\tau^2_{1}$.
     * $\beta_2$ is the slope of *cognateness* on the outcome; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation 1.
     * $\beta_{2j}$ is the slope of *cognateness* on cognate status $j$; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation $\tau^2_{2}$.
     * $\beta_3$ is the slope of the interaction between *language profile* and *cognateness* on the outcome; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation 1.
     * $\beta_{3j}$ is the slope of the interaction term between *language profile* and *cognateness* on cognate status $j$; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation $\tau^2_{3}$.

We will run separate models for items in the most dominant and less dominant language, and for receptive and productive vocabulary.

### Setting priors

We will use the Bayesian framework to estimate the parameters of our model. This involves the use of the Bayes' rule to compute the probability ($\pi$) of all values our set of estimated parameters ($\theta$) can take. To do this, it combines prior information coming from scientific expertise (also known as priors, $\pi(\theta)$), and the likelihood of each value the parameter can take, given the observed data ($\pi(y|\theta)$)). The product of both prior and likelihood is then standardised using the marginal likelihood of the parameter sampling space ($\pi(y)$). The combination of both sources of information leads to what is known as the posterior distribution of each parameter ($\pi(\theta|y)$), which informs us of the likelihood of each value the parameters can take, given our prior knowledge about it, and the observed data. We can extract information from this posterior distribution in several ways. We will use probabilistic posterior sampling to approximate this distribution and characterise it.

$$\pi(\theta | y) = \frac{\pi(y | \theta)\pi(\theta)}{\pi(y)}$$

We need to set priors for $k_i$, $\sigma$, $\tau^2_{0}$, $\tau^2_{1}$, $\tau^2_{2}$, $\tau^2_{3}$, $\beta_{0j}$, $\beta_{1j}$, and $\beta_{2j}$ and $\beta_{3j}$. We will specify the following priors:

    * $k_i$: `normal(1, 0.5)`
    * $\beta_{1}$: `normal(0, 1)`
    * $\beta_{2}$: `normal(0, 1)`
    * $\beta_{3}$: `normal(0, 1)`
    * $\sigma$: `student_t(3, 0, 10)`
    * $\tau^2_{0}$: `student_t(3, 0, 10)`,
    * $\tau^2_{1}$: `student_t(3, 0, 10)`
    * $\tau^2_{2}$: `student_t(3, 0, 10)`
    * $\tau^2_{3}$: `student_t(3, 0, 10)`

### Model fitting

We run our model using the `brm()` function in the `brms` R package [@bruckner2018], which uses the Stan environment [@stan2019] in the back end to approximate the posterior distribution via Hamiltonian Monte Carlo (HMC).

### Comparing models
*
To analyse the effect of the interaction between *language profile* and *cognate status*, we will fit three models. A first minimal model will not include any of the predictors of interest. A second null model will include *language profile* as a population effect A third complete model will also include *cognate status*, and its interaction with *language profile* as population effects. All models will include *item* as grouping variable. We will specify, in each model, a maximal random effects structure that includes by-item intercepts, slopes, and correlation parameters [@barr2013a]. We will then compare the three models following the workflow suggested by @schad2019. The minimal model will be expanded to the null model, and then the null model be expanded to the complete model. At each stage, we will check whether the expanded model is more adequate than the previous. In case the expanded model proves more inadequate, we will fall back to the simpler, previous model.

### 2020-06-15 Manuscript/trajectories_manuscript.Rmd
First commit

---
title             : "The role of cross-linguistic lexical similarity on bilingual word acquisition"
shorttitle        : "Cross-linguistic similarity and  word acquisition"

author: 
  - name          : "Gonzalo García-Castro"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain"
    email         : "gonzalo.garciadecastro@upf.edu"
  - name          : "Daniela Avila-Varela"
    affiliation   : "1"
  - name          : "Núria Sebastian-Galles"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Previous literature on early vocabulary has been mostly committed to exploring how the receptive and productive vocabulary size changes with age, and its relationship with toddlers’ performance on language tasks (e.g. Fernald, Swingley, & Pinto, 2001; Fernald, Perfors, & Marchman, 2006). The content of the developing lexicon has remained relatively unexplored. More recently, studies have focused on the characterisation of the developmental trajectory of individual words, reporting an earlier age of acquisition for words with high frequency, concreteness, and phonological neighbourhood density (e.g. Braginsky, Yurovsky, Marchman, & Frank, 2019; Jones & Brandt, 2019). Such item-level analysis allows not only to predict the age of acquisition of specific words, but also to shed light on the cognitive processes that underlie early word learning. This approach is particularly interesting for investigating the potential impact of acquiring two languages simultaneously. A recent study by Floccia et al. (2018) compared vocabulary sizes of 24-month-old children learning British English, together with an additional language from a pool of 13 diverse languages. They found larger productive vocabulary sizes of toddlers learning two languages that are phonologically similar (see for a similar approach Bosch & Ramon-Casas, 2014). The mechanisms underlying this effect remain unknown. One possibility is that the similarity between the two languages speeds the acquisition of form-similar translation equivalents. The aim of this study is to perform an item-wise analysis on infant’s vocabulary contents. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish. We expect that phonological overlap between translation equivalents will predict earlier age of acquisition. If this effect is driven by the phonological overlap between translation equivalents, cognate pairs should be acquired closer in time than non-cognates. For instance, the translation of cat /gato/ in Spanish and /gat/ in Catalan should be learnt at approximately the same age. This should not necessarily happen for the translations of dog, /pero/ and /gos/. We analyse the time elapsed between the acquisition of a word in one language, and its translation in the other.
  We present preliminary data (data collection is ongoing) on receptive vocabulary of 90 monolingual and bilingual toddlers aged 18 to 36 months. We obtained parental responses to 230 pairs of Catalan-Spanish translation equivalents, resulting in a total of 41490 responses. Words that yielded a phonologically similar translation equivalent were more likely to be present in toddlers’ receptive vocabulary at all ages, and were acquired earlier, even when accounting for the effect of word frequency. Further analysis will explore the distance in time between the acquisition of pairs of translation equivalents using a more detailed measure of phonological overlap across translation equivalents and taking into account cross-individual and cross-item variability.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "word acquisition, vocabulary, bilingualism, lexical similarity"
wordcount         : "X"

bibliography      : ["../References/BiLexicon.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
keep_text         : false
class             : "man"

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  cache.extra = knitr::rand_seed,
  results = "asis",
  out.width = "60%",
  fig.align = "center"
)
```

```{r prepare, include=FALSE}
# load packages
library(papaja)
library(knitr)

library(magrittr)
library(tibble)
library(dplyr)
library(forcats)
library(stringr)
library(readxl)
library(tidyr)
library(lubridate)

library(ggplot2)
library(ggridges)
library(patchwork)
library(wesanderson)

library(truncnorm)
library(data.table)
library(here)

# load/create functions
source(here("R", "functions.R"))

# set params
age_bins <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "30-32")

```

```{r prepare_data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# prepared data
dat_items <- fread(here("Data", "04_prepared.csv"), na.strings = c("")) %>%
  as_tibble() %>%
  drop_na(language)

# load data
dat <- fread(here("Data", "02_merged.csv"), na.strings = "") %>%
  filter(source %in% c("formr1", "formr2", "formr_short"),
         lp %in% c("Monolingual", "Bilingual"),
         age_bin %in% age_bins) %>%
  as_tibble() %>%
  mutate_at(vars(date_birth, time_stamp), as_date)

# import data
wordbank <- fread(here("Data", "05_wordbank.csv"))
wordbank_priors <- fread(here("Data", "05_wordbank-priors.csv"))
mid_prior <- round(wordbank_priors$estimate[wordbank_priors$term=="mid"], 2)
mid_steep <- round(wordbank_priors$estimate[wordbank_priors$term=="steep"], 2)

# load results
bayes_factor <- readRDS(here("Results", "bayes_factor.rds"))
loo_comp <- readRDS(here("Results", "loo.rds"))
fit0_production <- readRDS(here("Results", "fit0_productive.rds"))
fit1_production <- readRDS(here("Results", "fit1_productive.rds"))
fit1_priors_production <- readRDS(here("Results", "fit1_prior_productive.rds"))

```

# Introduction

One of the main challenges bilingual infants and toddlers face is learning two distinct set of words (one for each language) that partially overlap in sound and meaning. Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals). When taking both languages into account, bilingual children between seem to know, at least, as many words as monolinguals do [@benzeev1977; @doyle1978; @rosenblum1983; @fernandez1992; @pearson1994; @bialystok2010; @blom2019]. Tough, not all bilinguals show asimilar developmental trajectory of lexical acquisition.

Recent studies have capitalised on the role of the similarity between the specific pair of language the infant is learning. @floccia2018 analysed vocabulary scores across of 24-month-old toddlers learning English and an additional language. Toddlers learning languages sharing a large amounth of cognates (e.g., English-German) showed larger vocabulary sizes than those learning less languages sharing less cognates (e.g., English-Chinese). @blom2019 extended these results to children aged three to 10 years. These results point to the possibility that the overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition. However, none of these studies provide an account for the mechanisms involved in this facilitatory effect of language driven by similarity across languages. The aim of this study is to explore two scenarios in which this effect may be taking place:

1) In a first scenario, it is the phonological overlap between language pairs, and not the amount of cognates they share, what boosts the vocabulary growthof bilingual toddlers. Languages sharing a lot of cognates tend to also share a lot of phonemic categories. It is possible that having to learn fewer phonemic categories makes it easier for toddlers to acquire words. If this is true, toddlers learning two languages with similar phonemic inventories should show larger vocabulary sizes than those learning languages that barely ovelap phonologically. Moreover, this facilitation effect should be reflected in the developmental trajectory of all the items included in a vocabulary inventory, independently of their cognate status.

2) In a second scenario, Words that are form similar It is the amount of cognates language pairs share that boosts vocabulary growth in bilingual toddlers. If this is true, toddlers may acquire cognates earlier

To our knowledge, there are no previous studies that have investigated the effect of cognateness on the acquisition trajectories of individual word forms. Braginsky et al. (2019) used data from MacArthur-Bates CDIs from multiple languages to explore what properties are associated with an earlier acquisition. The found that frequency, concreteness, and [...] significantly predicted an earlier age of acquisition. We plan to extend these analyses to the case of bilinguals.

We have developed a fully-online vocabulary inventory in Catalan and Spanish that is more exhaustive than the MacArthur-Bates CDI, including a larger number of items. All participants filled the questionnaire it both languages, thus providing responses to pairs of translation equivalents.

In this study, we examine the role of cognateness on age of acquisition by modelling the probability of parents reporting that the participant is able to understand, say and understand on each item in the questionnaire. We will test the effect the participants' language profile (monolingual vs. bilingual), and its cognate status on the developmetnal curves of each item.

We will analyse the developmental trajectories of individual words in monolinguals and bilinguals learning Catalan and Spanish (two phonologically close languages sharing a great deal of cognates), between 14 and 30 months of age. We will test whether cognate words are learnt earlier than non-cognate words, taking into account wether they are part of the most- o less-dominant language of the infant, as well as their frequency, concreteness, and number of syllables. We have deloped an on-line questionnaire that entails a short language exposure questionnaire, asks for some demographic information, and presents parent with a list of ~800 words in each language, for which parent report whether their child understands, says, or doesn't understand and say, in each item.

We will fit a Bayesian multilevel model on the item responses, modelling the probability of parents reporting that a given word is acquired by the infant, including language profile (monolingual vs. bilingual), item dominance (dominant language vs. non-dominant language), frequency, concreteness, and number of syllables. We will include *meaning* as grouping variable. We will fit a sigmoidal curve [see @mayor2008] on each item, defined by the steepness of the developmental curve, and the mid-point (the poinnt at which the developmental curve is steepest, which will be considered the point of acquisition). Priors will be derived from previous literature on the subject.

NOTES:

Evidence that bilingual lexica barely overlap [@leopold1948; @taeschner1978].

Studies with infants from 8 to 30 months of age show weaker evidence for this claim [@pearson1993; @dehoouwer2014], pointing to the posibility that differences between bilingual and monolingual vocabulary sizes may be dependent on maturational factors.

sugested that the developmental trajectory of lexical aquisition varies across two dimensions: the language profile of the infant, and the joint proporties of the specific language pair the infant is learning. Regarding the first, @pearson1994 found thant infants exposed to a more balanced

Of special interest is the fact than cognates (form-similar translation equivalents) tend to be overrepresented in the early bilingual vocabulary [@bosch2014].

The similarity between both languages may play a role in bilingual lexical aquisition.

## Hypotheses

Hypothesis 1
:  The boost effect of language similarity is driven by cognates being acquired earlier than non-cognates. We predict that age of acquisition of cognates will be ealier thant that of non-cognates. This effect should only be present in bilinguals.

Hypothesis 2
:  That this effect is driven by the word-forms in one language scaffolding the acquisition of their translation equivalents via parallel activation. We predict that the difference in time between the acquisition pairs of translation equivalents is shorter in cognates than in non-cognates. This effect will only be present in bilinguals.

# Method

## Participants

Data from `r dat$id_db %>% unique %>% length() %>% printnum()` participants from the different cities in Spain were invited to participate through social media. Families in the Metropolitan area of Barcelona were also recruited at birth. All families participated voluntarily. Data were collected between `r min(dat$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`, using the BiLexicon inventory. This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.

```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

dat %>%
  group_by(dominance, sex) %>%
  summarise(n = n(),
            age_mean = mean(age, na.rm = TRUE),
            age_sd = sd(age, na.rm = TRUE),
            doe_catalan_mean = mean(doe_catalan, na.rm = TRUE),
            doe_catalan_sd = sd(doe_catalan, na.rm = TRUE),
            doe_spanish_mean = mean(doe_spanish, na.rm = TRUE),
            doe_spanish_sd = sd(doe_spanish, na.rm = TRUE)) %>%
  ungroup() %>%
  kable(digits = 2,
        col.names = c("Dominance", "Sex", "N", "M_Age", "SD_Age", "M DOE Cat", "SD_DOE_Cat",  "M_DOE_Spa", "SD_DOE_Spa"),
        format.args = list(big.mark = ",", na.string = "-"),
        align = "c",
        caption = "Sample demographics and overall language profile. Dominance = Language to which the participant is exposed the most, N = sample size, M age = Mean age, SD age = standard deviation of age, M DOE Cat = mean degree of exposure to Catalan, SD DOE Cat = standard deviation of exposure to Catalan, M DOE Spa = mean degree of exposure to Spanish, SD DOE Spa = standard deviation of exposure to Spanish.")

```

* *Location*
* *Age*
* *Sex*
* *Language profile*: : Linguistic profiles were assessed using the Language Exposure Questionnaire (LEQ) [@bosch2001]. This questionnaire estimates de degree of exposure (DOE) to each language via a 10 minutes long interview with the parents. LEQs were administered inmediately before or after the testing session at the lab. We here considered only exposure to Catalan and Spanish. Participants with >10% exposure to a third language other than Catalan or Spanish were excluded.
* *SES*

### Exclusion criteria

* Language disorders

### Questionnaires

#### BiLexicon 1.0

#### BiLexicon 2.0

#### BiLexicon Short

This questionnaire was developed as a short version of the BiLexicon 1.0 and BiLexicon 2.0 inventories. Given the large amount of items to be answered to, we restricted the use of such inventories to participants that were to participate presentially in one of the studies run in the lab at that time. These participants only filled one of the questionnaires, and did so only once.

The aim of developing a short version of this questionnaire (BiLexicon Short) was to provde a more on-line friendly format to families that would participate remotely exclusively, or that would fill the questionnaire more than once (e.g., in longitudinal studies). We divided the items of each language in the pool into four versions named A, B, C, and D. Items in each category (e.g., animals, auxiliary words) were randomly assigned one of the versions. Thos categories that did not entail more than 16 items (thus resulting in less than four items after dividing it in ofur lists), were not divided (all of their items were included in all versions).

When filling the questionnaire for the first time, participants were randomly assigned one version. Each familiy filled only one of the versions, thus reducing the number of items to be answered. To ensure that longitudinal participants would provide longitudinal data in the questionnaire, we made sure that participants were always assigned the same version of this questionnaire. Words that were used as part of the trial lists in the experiment families would participate in were present across all versions.

### Item inclusion criteria

We gathered data from `r dat_items %>% distinct(language, meaning) %>% group_by(language) %>% summarise(n = n()) %>% pull(n) %>% min() %>% printnum(round = 0)` distinct translation equivalents across Spanish (`r dat_items %>% distinct(language, meaning) %>% group_by(language) %>% summarise(n = n()) %$%  n[2] %>% printnum()` word forms) and Catalan (`r dat_items %>% distinct(language, meaning) %>% group_by(language) %>% summarise(n = n()) %$% n[1] %>% printnum()` word forms). The unequal number of word-orms across both languages is due to that some tanslation equivalents did not entail a one-to-one-corespondence: some items had more than one translation equivalent in the other language (e.g. the Catalan word form *walking*, *anar*, can be translated as both *caminar* [to walk] or *ir* [to go]).

We included items from the following categories: Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles. We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.

We excluded items that enailed more than one word (e.g., *barrita de cereales* [cereal bar]).

### Item properties

#### Frequencies

We retrieved frequencies from SUBTLEX-ESP [@cuetos2009] and SUBTLEX-CAT [@ferre2013] for Spanish and Catalan words, respectively. Word frequency was calculated as the Zipf score of the word [@zipf1949; @vanheuven2014].

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

include_graphics(here("Figures", "pool_freq_cognateness.png"))

```

#### Cognate status

#### Demographics by item

```{r echo=FALSE, message=FALSE, warning=FALSE}

dat %>%
  distinct(item, id, language, lp, sex, age) %>%
  group_by(item, language, lp, sex) %>%
  summarise(mean_age = mean(age, na.rm = TRUE),
            sd_age = mean(age, na.rm = TRUE),
            min_age = min(age, na.rm = TRUE),
            max_age =  max(age, na.rm = TRUE),
            n = n()) %>%
  group_by(language, lp, sex) %>%
  summarise(mean_age = mean(mean_age, na.rm = TRUE),
            sd_age = mean(sd_age, na.rm = TRUE),
            min_age = min(min_age, na.rm = TRUE),
            max_age =  max(max_age, na.rm = TRUE),
            mean_n = mean(n, na.rm = TRUE),
            n = sum(n, na.rm = TRUE)) %>%
  kable(.,
        digits = 2,
        col.names = c("Language", "LP", "Sex", "M Age", "SD Age", "Min Age", "Max Age","M N", "N"),
        align = "c",
        caption = "Demographics of responses by item. Language = item language, LP = language profile of participant, Sex = sex of participant, M Age = mean age of responses, SD Age = mean SD of responses across items, Min Age = minimum age of responses across items, Max Age = maximum age of responses across items, M N = mean number of responses across items, N = total number of responses across items."
  )

```

## Data analysis

For testing the first hypothesis, we fit a non-linear model using logistic curves to model the probability of aquisition of words across ages. Logistic curves are characterised by three parameters: 1) an asymptote (maximum value of the curve in the Y-axis), 2) the steepness (how fast the curve grows), and 3) a mid-point (the point in the X-axis at which the steepness is highest). Figure 1 shows the elements of a simulated logarithmic curve:

```{r logcurve, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, fig.cap="Simulated logarithmic curve. The curve is characterised by three parameters: Asymptote (red), mid-point (blue), and steepness (green). This curve is the result of simulating learning curves of 1000 items. We set fixed values for the asymptote at 1. Mid-points were randomly generated from a normal distribution truncated at 0 (no negative age values are allowed), with mean 23 (*SD* = 0.5), and steepness values were generated from a normal ditribution truncated at 0 (we assume that the proportion of participants that have acquired each item can only grow), with mean 0.1 (*SD* = 1) The solid black line represents the average proportion of participants who have acquired the item, and the shaded ribbon indicates the standard error of the mean. We suggest that learning curves of individual word-forms, as well as the evolution of vocabulary size across ages, follows this trend. Our model will estimate the three of to adress our hypotheses."}

include_graphics(here("Figures", "05_illustration.png"))

```

One reason for using logistic curves to fit the data is that previous studies describe lexical development as a non-linear process, where between 20 and 24 months a vocabulary spurt happens. An analysis of the available data in Wordbank  [@frank2016] provides moderate support for this hypothesis. We fitted a logistic model using the proportion of infants that acquired each item as output, and age as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness). We included data from `r round(nrow(wordbank)/length(unique(wordbank$age)), 0)` items from `r wordbank$language %>% unique %>% length` languages^[Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish. All the other available languages were excluded, as no (or insufficient) data were available.], from infants between 16 and 30 months of age. Then we compared the fit of the logarithmic model agains a linear model, and found that the formr fitted the data slightly better (see Figure 2).

A second reason for using logistic curves is that the parameters that characterise a logistic curve map onto our hypotheses quite easily. Out first hypothesis concerns the mid-point parameter, which indicates the point at which the steepness of the curve is maximum. In our model, the value of this parameter can be interpreted as the age at which the proportion of participants that have acquired a given item is at maximum rate. The mathematical interpretation of this parameter is easily mapped into a definition of age of acquisition.

```{r wordbank, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Logarithmic model fitted on Wordbank data, compared to a linear model. The logarithmic model fits the data slightly better. The estimated values of the three parameters involved in the logarithmic model (asymptote, mid-point and steepness) are indicated."}

include_graphics(here("Figures", "05_wordbank.png"))

```

We built a Bayesian logistic model that estimates the value of the mid-point and steepness across items. Under our hypothesis, mid-points are generated from a linear model that incorporates language profile (monolingual vs. bilingual) and cognateness (non-cognate vs. cognate) as predictors, where cognateness predictes ealier mid-points for bilingual, but not monolingual participants. We wil also include frequency as a covariate in this linear model. Due to computational limitations, we set a constant value for the asymptote at `r round(wordbank_priors$estimate[wordbank_priors$term=="asym"], 2)`, according to our model fit on Wordbank data. Due to computational limitations, we did not run the same linear model to estimate steepness. Rather, this parameter was just included as an intercept. We compared how well this model fits the data *versus* how a null model that does not include *cognateness* as predictor does, by using Bayes factors and Pareto-smoothed importance sampling [PSIS; @vehtari206].

We used the information provided by Wordbank as priors to feed our Bayesian model. Accordingly, we set an informative prior for the mid-point $N(`r round(mid_prior, 2)`, 0)$ and the steepness $N(`r round(mid_steep, 2)`, 0)$ intercepts. We set weakly informative priors for the coefficients of frequency, language profile, cognateness, and their interaction at $N(0, 10)$. We also set a weakly informative prior for the standard deviation, $\sigma \sim t(3, 0, 2.5)$.

We fit both null and alternative models within the R environment [@rcoreteam2019; @rstudio2020] using the `brms` package [@brueckner2016], which relies on the probabilistic language Stan [@carpenter2013] to estimate posterior distributions. Stan uses Hamiltonian Montecarlo to sample the posterior. for model selection, we used both the `brms` and `loo` [@vehtari2016] packages. Data and model results will be processed and visualised using the R packages the `tidyverse` family of packages [@wickham2020] and the `tidybayes` package [@mskay2019].

# Results

## Model selection

We compared both models using two different criteria. First, we computed a Bayes Factor comparing the average posterior probability of the Null and the Extended models under the observed data. We obtained little evidence in favour of the Extended model over the Null model (*BF* = `r printnum(bayes_factor$bf`)). Second, we used an approximation to leave-one-out cross-valitation [Pareto-smoothed Importance Sampling, @vehtari2016] to compare how well each model was able to predict the data. Both models were very accurate when predicting the outcome: . Extended model's predictions were not substantially more accurate ($\hat{elpd}_{loo}^{Extended} - \hat{elpd}_{loo}^{Null}$ = `r printnum(loo_comp %>% as_tibble %$% elpd_diff[2]`, *SE* = `r printnum(loo_comp %>% as_tibble %$% se_diff[2])`).

## Posterior distibutions

```{r posteriors, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

include_graphics(here("Figures", "05_gca-coefficients.png"))

```

## Posterior correlations

```{r correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

include_graphics(here("Figures", "05_gca-correlations.png"))

```

## Prior predictive checks

Following @gabry2019, we performed prior predictive chackes to disgnose the appropriateness of our priors.

## Posterior predictive checks

## Model diagnosis

```{r convergence, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

include_graphics(here("Figures", "05_gca-convergence.png"))

```

### Auto-correlation

### Potential scale reduction factor ($\hat{R}$)

## Posterior predictive checks

Does our model generate similar observations to the ones in our data-set?

# Discussion

## Limitations

* Cross-sectional data for longitudinal claims
* Data is right censored    
* Inventory:
    - Phonological forms?
    - We can't say how frequently words are produced by the parents or the toddler.
    - Although instructed to ignore imitations, it is difficult to say whether a toddler has *really* acquired a word that produces, or she can only imitate it.
    - Responses in the questionnaire rely heavily on parental memory.
    - We don't know about the context at which words are heard or produced.
    - Classification system of items (e.g., *household items*) is adult centric. Children may use the words *pretty* to name jewellery [@bates1994].
* Mid and mid-upper class families are overrepresented in the sample.

# Appendix

## Appendix 1: Session info

```{r session_info, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
sessionInfo()
```

## Appendix 2: Model

Our model is incomplete, in that we have not included all parameters that may play a role in the real mechanisms that generate our observed data. Previous studies have found robust effects of sex (e.g., @frank2016), word length/number of phonemes (e.g., @frank2016), etc. Computational limitations have prevented us from specifying and fitting a more exhaustive model. Following the model-building strategy suggested by @schad2020, we have specified an intitial model that captures the fundamental structure of our phenomenon of interest and that can be computed with our available resources. Then, we have expanded such model incorporating a new predictor that adresses our hypotheses.

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>

\endgroup

```{r create_r-references, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

r_refs(file = "../References/BiLexicon.bib")

```

### 2020-06-15 Code/R/05_simulate.R

@@ -1,0 +1,219 @@
+#### 05_simulations: Simnulate data
+
+#### set up ##############################################
+
+# load packages
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(ggplot2)
+library(wesanderson)
+library(data.table)
+library(truncnorm)
+library(brms)
+library(modelr)
+library(tidybayes)
+library(here)
+
+# create/load functions
+source(here("R", "functions.R"))
+inv_logit <- function(x) 1 / (1 + exp(-x))
+
+# set params
+age_bins <- c("10-12", "12-14", "12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "30-32")
+
+#### generate data #######################################
+mid           <- 7.5
+mid_sd        <- 3
+steep         <- 0.5
+steep_sd      <- 0.3
+asym          <- 1
+b_cognateness <- 1
+n <- 100
+
+item <- factor(1:n)
+age <- factor(0:length(age_bin), ordered = TRUE)
+
+dat <- expand_grid(item, age) %>%
+  mutate(cognateness = rep(c(-0.5, 0.5), each = nrow(.)/2),
+         cognateness_label = ifelse(cognateness==-0.5, "Non-cognate", "Cognate"),
+         age = as.numeric(age)) %>%
+  group_by(item) %>%
+  mutate(mid = rtruncnorm(n = 1, mean = mid, sd = mid_sd, a = 0) + cognateness_slope*cognateness,
+         steep = rtruncnorm(n = 1, mean = steep, sd = steep_sd, a = 0),
+         proportion = 1 / (1 + exp((mid - age) * steep))) %>%
+  arrange(item, age, cognateness) %>%
+  as_tibble()
+
+ggplot(dat, aes(age, proportion, colour = cognateness_label, fill = cognateness_label)) +
+  geom_hline(yintercept = 0.50, linetype = "dotted") +
+  geom_density(aes(x = mid, y = stat(density)), alpha = 0.5, colour = NA, bw = 1) +
+  geom_rug(aes(x = mid), sides = "b", alpha = 0.5) +
+  geom_line(aes(group = item), alpha = 0.1) +
+  stat_summary(fun.data = "mean_se", geom = "ribbon", colour = NA, alpha = 0.5) +
+  stat_summary(fun = "mean", geom = "line") +
[…]

### 2020-06-15 Manuscript/trajectories_manuscript.tex
First commit

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,man,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
            pdfkeywords={word acquisition, vocabulary, bilingualism, lexical similarity},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Cross-linguistic similarity and  word acquisition}
\author{Gonzalo García-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Núria Sebastian-Galles\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra}
\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

Correspondence concerning this article should be addressed to Gonzalo García-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain. E-mail: gonzalo.garciadecastro@upf.edu}
\keywords{word acquisition, vocabulary, bilingualism, lexical similarity\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{The role of cross-linguistic lexical similarity on bilingual word acquisition}

\date{}

\abstract{
Previous literature on early vocabulary has been mostly committed to exploring how the receptive and productive vocabulary size changes with age, and its relationship with toddlers' performance on language tasks (e.g.~Fernald, Swingley, \& Pinto, 2001; Fernald, Perfors, \& Marchman, 2006). The content of the developing lexicon has remained relatively unexplored. More recently, studies have focused on the characterisation of the developmental trajectory of individual words, reporting an earlier age of acquisition for words with high frequency, concreteness, and phonological neighbourhood density (e.g.~Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019). Such item-level analysis allows not only to predict the age of acquisition of specific words, but also to shed light on the cognitive processes that underlie early word learning. This approach is particularly interesting for investigating the potential impact of acquiring two languages simultaneously. A recent study by Floccia et al. (2018) compared vocabulary sizes of 24-month-old children learning British English, together with an additional language from a pool of 13 diverse languages. They found larger productive vocabulary sizes of toddlers learning two languages that are phonologically similar (see for a similar approach Bosch \& Ramon-Casas, 2014). The mechanisms underlying this effect remain unknown. One possibility is that the similarity between the two languages speeds the acquisition of form-similar translation equivalents. The aim of this study is to perform an item-wise analysis on infant's vocabulary contents. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish. We expect that phonological overlap between translation equivalents will predict earlier age of acquisition. If this effect is driven by the phonological overlap between translation equivalents, cognate pairs should be acquired closer in time than non-cognates. For instance, the translation of cat /gato/ in Spanish and /gat/ in Catalan should be learnt at approximately the same age. This should not necessarily happen for the translations of dog, /pero/ and /gos/. We analyse the time elapsed between the acquisition of a word in one language, and its translation in the other.
We present preliminary data (data collection is ongoing) on receptive vocabulary of 90 monolingual and bilingual toddlers aged 18 to 36 months. We obtained parental responses to 230 pairs of Catalan-Spanish translation equivalents, resulting in a total of 41490 responses. Words that yielded a phonologically similar translation equivalent were more likely to be present in toddlers' receptive vocabulary at all ages, and were acquired earlier, even when accounting for the effect of word frequency. Further analysis will explore the distance in time between the acquisition of pairs of translation equivalents using a more detailed measure of phonological overlap across translation equivalents and taking into account cross-individual and cross-item variability.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

One of the main challenges bilingual infants and toddlers face is learning two distinct set of words (one for each language) that partially overlap in sound and meaning. Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals). When taking both languages into account, bilingual children between seem to know, at least, as many words as monolinguals do ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}). Tough, not all bilinguals show asimilar developmental trajectory of lexical acquisition.

Recent studies have capitalised on the role of the similarity between the specific pair of language the infant is learning. Floccia et al. (2018) analysed vocabulary scores across of 24-month-old toddlers learning English and an additional language. Toddlers learning languages sharing a large amounth of cognates (e.g., English-German) showed larger vocabulary sizes than those learning less languages sharing less cognates (e.g., English-Chinese). ({\textbf{???}}) extended these results to children aged three to 10 years. These results point to the possibility that the overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition. However, none of these studies provide an account for the mechanisms involved in this facilitatory effect of language driven by similarity across languages. The aim of this study is to explore two scenarios in which this effect may be taking place:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  In a first scenario, it is the phonological overlap between language pairs, and not the amount of cognates they share, what boosts the vocabulary growthof bilingual toddlers. Languages sharing a lot of cognates tend to also share a lot of phonemic categories. It is possible that having to learn fewer phonemic categories makes it easier for toddlers to acquire words. If this is true, toddlers learning two languages with similar phonemic inventories should show larger vocabulary sizes than those learning languages that barely ovelap phonologically. Moreover, this facilitation effect should be reflected in the developmental trajectory of all the items included in a vocabulary inventory, independently of their cognate status.
\item
  In a second scenario, Words that are form similar It is the amount of cognates language pairs share that boosts vocabulary growth in bilingual toddlers. If this is true, toddlers may acquire cognates earlier
\end{enumerate}

To our knowledge, there are no previous studies that have investigated the effect of cognateness on the acquisition trajectories of individual word forms. Braginsky et al. (2019) used data from MacArthur-Bates CDIs from multiple languages to explore what properties are associated with an earlier acquisition. The found that frequency, concreteness, and {[}\ldots{}{]} significantly predicted an earlier age of acquisition. We plan to extend these analyses to the case of bilinguals.

We have developed a fully-online vocabulary inventory in Catalan and Spanish that is more exhaustive than the MacArthur-Bates CDI, including a larger number of items. All participants filled the questionnaire it both languages, thus providing responses to pairs of translation equivalents.

In this study, we examine the role of cognateness on age of acquisition by modelling the probability of parents reporting that the participant is able to understand, say and understand on each item in the questionnaire. We will test the effect the participants' language profile (monolingual vs.~bilingual), and its cognate status on the developmetnal curves of each item.

We will analyse the developmental trajectories of individual words in monolinguals and bilinguals learning Catalan and Spanish (two phonologically close languages sharing a great deal of cognates), between 14 and 30 months of age. We will test whether cognate words are learnt earlier than non-cognate words, taking into account wether they are part of the most- o less-dominant language of the infant, as well as their frequency, concreteness, and number of syllables. We have deloped an on-line questionnaire that entails a short language exposure questionnaire, asks for some demographic information, and presents parent with a list of \textasciitilde{}800 words in each language, for which parent report whether their child understands, says, or doesn't understand and say, in each item.

We will fit a Bayesian multilevel model on the item responses, modelling the probability of parents reporting that a given word is acquired by the infant, including language profile (monolingual vs.~bilingual), item dominance (dominant language vs.~non-dominant language), frequency, concreteness, and number of syllables. We will include \emph{meaning} as grouping variable. We will fit a sigmoidal curve (see {\textbf{???}}) on each item, defined by the steepness of the developmental curve, and the mid-point (the poinnt at which the developmental curve is steepest, which will be considered the point of acquisition). Priors will be derived from previous literature on the subject.

NOTES:

Evidence that bilingual lexica barely overlap ({\textbf{???}}; {\textbf{???}}).

Studies with infants from 8 to 30 months of age show weaker evidence for this claim ({\textbf{???}}; {\textbf{???}}), pointing to the posibility that differences between bilingual and monolingual vocabulary sizes may be dependent on maturational factors.

sugested that the developmental trajectory of lexical aquisition varies across two dimensions: the language profile of the infant, and the joint proporties of the specific language pair the infant is learning. Regarding the first, ({\textbf{???}}) found thant infants exposed to a more balanced

Of special interest is the fact than cognates (form-similar translation equivalents) tend to be overrepresented in the early bilingual vocabulary (Bosch \& Ramon-Casas, 2014).

The similarity between both languages may play a role in bilingual lexical aquisition.

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

\begin{description}
\tightlist
\item[Hypothesis 1]
The boost effect of language similarity is driven by cognates being acquired earlier than non-cognates. We predict that age of acquisition of cognates will be ealier thant that of non-cognates. This effect should only be present in bilinguals.
\item[Hypothesis 2]
That this effect is driven by the word-forms in one language scaffolding the acquisition of their translation equivalents via parallel activation. We predict that the difference in time between the acquisition pairs of translation equivalents is shorter in cognates than in non-cognates. This effect will only be present in bilinguals.
\end{description}

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

Data from 235 participants from the different cities in Spain were invited to participate through social media. Families in the Metropolitan area of Barcelona were also recruited at birth. All families participated voluntarily. Data were collected between 28th October, 2019 and 31th May, 2020, using the BiLexicon inventory. This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.

\begin{table}

\caption{\label{tab:participants}Sample demographics and overall language profile. Dominance = Language to which the participant is exposed the most, N = sample size, M age = Mean age, SD age = standard deviation of age, M DOE Cat = mean degree of exposure to Catalan, SD DOE Cat = standard deviation of exposure to Catalan, M DOE Spa = mean degree of exposure to Spanish, SD DOE Spa = standard deviation of exposure to Spanish.}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c}
\hline
Dominance & Sex & N & M\_Age & SD\_Age & M DOE Cat & SD\_DOE\_Cat & M\_DOE\_Spa & SD\_DOE\_Spa\\
\hline
Catalan & Female & 126,526 & 20.96 & 4.59 & 73.94 & 18.48 & 23.52 & 18.29\\
\hline
Catalan & Male & 132,856 & 20.86 & 4.78 & 70.83 & 17.74 & 27.77 & 17.08\\
\hline
Spanish & Female & 66,476 & 21.22 & 4.76 & 17.38 & 14.74 & 81.14 & 15.60\\
\hline
Spanish & Male & 47,511 & 20.83 & 4.88 & 26.67 & 13.12 & 72.66 & 11.67\\
\hline
\end{tabular}
\end{table}

\begin{itemize}
\tightlist
\item
  \emph{Location}
\item
  \emph{Age}
\item
  \emph{Sex}
\item
  \emph{Language profile}: : Linguistic profiles were assessed using the Language Exposure Questionnaire (LEQ) ({\textbf{???}}). This questionnaire estimates de degree of exposure (DOE) to each language via a 10 minutes long interview with the parents. LEQs were administered inmediately before or after the testing session at the lab. We here considered only exposure to Catalan and Spanish. Participants with \textgreater{}10\% exposure to a third language other than Catalan or Spanish were excluded.
\item
  \emph{SES}
\end{itemize}

\hypertarget{exclusion-criteria}{%
\subsubsection{Exclusion criteria}\label{exclusion-criteria}}

\begin{itemize}
\tightlist
\item
  Language disorders
\end{itemize}

\hypertarget{questionnaires}{%
\subsubsection{Questionnaires}\label{questionnaires}}

\hypertarget{bilexicon-1.0}{%
\paragraph{BiLexicon 1.0}\label{bilexicon-1.0}}

\hypertarget{bilexicon-2.0}{%
\paragraph{BiLexicon 2.0}\label{bilexicon-2.0}}

\hypertarget{bilexicon-short}{%
\paragraph{BiLexicon Short}\label{bilexicon-short}}

This questionnaire was developed as a short version of the BiLexicon 1.0 and BiLexicon 2.0 inventories. Given the large amount of items to be answered to, we restricted the use of such inventories to participants that were to participate presentially in one of the studies run in the lab at that time. These participants only filled one of the questionnaires, and did so only once.

The aim of developing a short version of this questionnaire (BiLexicon Short) was to provde a more on-line friendly format to families that would participate remotely exclusively, or that would fill the questionnaire more than once (e.g., in longitudinal studies). We divided the items of each language in the pool into four versions named A, B, C, and D. Items in each category (e.g., animals, auxiliary words) were randomly assigned one of the versions. Thos categories that did not entail more than 16 items (thus resulting in less than four items after dividing it in ofur lists), were not divided (all of their items were included in all versions).

When filling the questionnaire for the first time, participants were randomly assigned one version. Each familiy filled only one of the versions, thus reducing the number of items to be answered. To ensure that longitudinal participants would provide longitudinal data in the questionnaire, we made sure that participants were always assigned the same version of this questionnaire. Words that were used as part of the trial lists in the experiment families would participate in were present across all versions.

\hypertarget{item-inclusion-criteria}{%
\subsubsection{Item inclusion criteria}\label{item-inclusion-criteria}}

We gathered data from 695 distinct translation equivalents across Spanish (695 word forms) and Catalan (703 word forms). The unequal number of word-orms across both languages is due to that some tanslation equivalents did not entail a one-to-one-corespondence: some items had more than one translation equivalent in the other language (e.g.~the Catalan word form \emph{walking}, \emph{anar}, can be translated as both \emph{caminar} {[}to walk{]} or \emph{ir} {[}to go{]}).

We included items from the following categories: Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles. We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.

We excluded items that enailed more than one word (e.g., \emph{barrita de cereales} {[}cereal bar{]}).

\hypertarget{item-properties}{%
\subsubsection{Item properties}\label{item-properties}}

\hypertarget{frequencies}{%
\paragraph{Frequencies}\label{frequencies}}

We retrieved frequencies from SUBTLEX-ESP ({\textbf{???}}) and SUBTLEX-CAT ({\textbf{???}}) for Spanish and Catalan words, respectively. Word frequency was calculated as the Zipf score of the word ({\textbf{???}}; {\textbf{???}}).

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/pool_freq_cognateness}

}

\caption{ }\label{fig:unnamed-chunk-1}
\end{figure}

\hypertarget{cognate-status}{%
\paragraph{Cognate status}\label{cognate-status}}

\hypertarget{demographics-by-item}{%
\paragraph{Demographics by item}\label{demographics-by-item}}

\begin{table}

\caption{\label{tab:unnamed-chunk-2}Demographics of responses by item. Language = item language, LP = language profile of participant, Sex = sex of participant, M Age = mean age of responses, SD Age = mean SD of responses across items, Min Age = minimum age of responses across items, Max Age = maximum age of responses across items, M N = mean number of responses across items, N = total number of responses across items.}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c}
\hline
Language & LP & Sex & M Age & SD Age & Min Age & Max Age & M N & N\\
\hline
Catalan & Bilingual & Female & 22.36 & 22.36 & 12.35 & 31.42 & 52.22 & 41620\\
\hline
Catalan & Bilingual & Male & 20.77 & 20.77 & 12.37 & 31.91 & 64.06 & 51057\\
\hline
Catalan & Monolingual & Female & 20.14 & 20.14 & 12.83 & 30.15 & 68.01 & 54204\\
\hline
Catalan & Monolingual & Male & 21.11 & 21.11 & 12.60 & 30.93 & 48.30 & 38495\\
\hline
Spanish & Bilingual & Female & 22.33 & 22.33 & 12.35 & 31.42 & 52.76 & 42209\\
\hline
Spanish & Bilingual & Male & 20.72 & 20.72 & 12.37 & 31.91 & 64.72 & 51778\\
\hline
Spanish & Monolingual & Female & 20.10 & 20.10 & 12.83 & 30.15 & 68.71 & 54969\\
\hline
Spanish & Monolingual & Male & 21.06 & 21.06 & 12.60 & 30.93 & 48.80 & 39037\\
\hline
\end{tabular}
\end{table}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

For testing the first hypothesis, we fit a non-linear model using logistic curves to model the probability of aquisition of words across ages. Logistic curves are characterised by three parameters: 1) an asymptote (maximum value of the curve in the Y-axis), 2) the steepness (how fast the curve grows), and 3) a mid-point (the point in the X-axis at which the steepness is highest). Figure 1 shows the elements of a simulated logarithmic curve:

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_illustration}

}

\caption{Simulated logarithmic curve. The curve is characterised by three parameters: Asymptote (red), mid-point (blue), and steepness (green). This curve is the result of simulating learning curves of 1000 items. We set fixed values for the asymptote at 1. Mid-points were randomly generated from a normal distribution truncated at 0 (no negative age values are allowed), with mean 23 (*SD* = 0.5), and steepness values were generated from a normal ditribution truncated at 0 (we assume that the proportion of participants that have acquired each item can only grow), with mean 0.1 (*SD* = 1) The solid black line represents the average proportion of participants who have acquired the item, and the shaded ribbon indicates the standard error of the mean. We suggest that learning curves of individual word-forms, as well as the evolution of vocabulary size across ages, follows this trend. Our model will estimate the three of to adress our hypotheses.}\label{fig:logcurve}
\end{figure}

One reason for using logistic curves to fit the data is that previous studies describe lexical development as a non-linear process, where between 20 and 24 months a vocabulary spurt happens. An analysis of the available data in Wordbank ({\textbf{???}}) provides moderate support for this hypothesis. We fitted a logistic model using the proportion of infants that acquired each item as output, and age as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness). We included data from 13119 items from 20 languages\footnote{Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish. All the other available languages were excluded, as no (or insufficient) data were available.}, from infants between 16 and 30 months of age. Then we compared the fit of the logarithmic model agains a linear model, and found that the formr fitted the data slightly better (see Figure 2).

A second reason for using logistic curves is that the parameters that characterise a logistic curve map onto our hypotheses quite easily. Out first hypothesis concerns the mid-point parameter, which indicates the point at which the steepness of the curve is maximum. In our model, the value of this parameter can be interpreted as the age at which the proportion of participants that have acquired a given item is at maximum rate. The mathematical interpretation of this parameter is easily mapped into a definition of age of acquisition.

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_wordbank}

}

\caption{Logarithmic model fitted on Wordbank data, compared to a linear model. The logarithmic model fits the data slightly better. The estimated values of the three parameters involved in the logarithmic model (asymptote, mid-point and steepness) are indicated.}\label{fig:wordbank}
\end{figure}

We built a Bayesian logistic model that estimates the value of the mid-point and steepness across items. Under our hypothesis, mid-points are generated from a linear model that incorporates language profile (monolingual vs.~bilingual) and cognateness (non-cognate vs.~cognate) as predictors, where cognateness predictes ealier mid-points for bilingual, but not monolingual participants. Due to computational limitations, we set a fixed value for the asymptote at 0.75, according to our model fit on Wordbank data. Due to computational limitations, we did not run the same linear model to estimate steepness. Rather, this parameter was just included as an intercept. We compared how well this model fits the data \emph{versus} how a null model that does not include \emph{cognateness} as predictor does, by using Bayes factors and Pareto-smoothed importance sampling (PSIS; {\textbf{???}}).

We used the information provided by Wordbank as priors to feed our Bayesian model. For mid-points, we specified a strong prior with a normal prior with mean = 23.08 and \emph{SD} = 1. For steepness, we also set a strong prior with a normal distribution with mean = 3.28 and SD = 2. We had little prior information about how cognateness impacts the mid-point of an item, and therefore we will set a weak prior.

We fit both null and alternative models within the R environment ({\textbf{???}}; {\textbf{???}}) using the \texttt{brms} package ({\textbf{???}}), which relies on the probabilistic language Stan ({\textbf{???}}). Data and model results will be processed and visualised using the R packages the \texttt{tidyverse} family of packages ({\textbf{???}}) and the \texttt{tidybayes} package ({\textbf{???}}).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{model-selection}{%
\subsection{Model selection}\label{model-selection}}

\hypertarget{posterior-distibutions}{%
\subsection{Posterior distibutions}\label{posterior-distibutions}}

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-coefficients}

}

\caption{ }\label{fig:posteriors}
\end{figure}

\hypertarget{posterior-correlations}{%
\subsection{Posterior correlations}\label{posterior-correlations}}

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-correlations}

}

\caption{ }\label{fig:correlations}
\end{figure}

\hypertarget{prior-predictive-checks}{%
\subsection{Prior predictive checks}\label{prior-predictive-checks}}

Following ({\textbf{???}}), we performed prior predictive chackes to disgnose the appropriateness of our priors.

\hypertarget{posterior-predictive-checks}{%
\subsection{Posterior predictive checks}\label{posterior-predictive-checks}}

\hypertarget{model-diagnosis}{%
\subsection{Model diagnosis}\label{model-diagnosis}}

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-convergence}

}

\caption{ }\label{fig:convergence}
\end{figure}

\hypertarget{auto-correlation}{%
\subsubsection{Auto-correlation}\label{auto-correlation}}

\hypertarget{potential-scale-reduction-factor-hatr}{%
\subsubsection{\texorpdfstring{Potential scale reduction factor (\(\hat{R}\))}{Potential scale reduction factor (\textbackslash{}hat\{R\})}}\label{potential-scale-reduction-factor-hatr}}

\hypertarget{posterior-predictive-checks-1}{%
\subsection{Posterior predictive checks}\label{posterior-predictive-checks-1}}

Does our model generate similar observations to the ones in our data-set?

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

\begin{itemize}
\tightlist
\item
  Cross-sectional data for longitudinal claims
\item
  Data is right censored\\
\item
  Inventory:

  \begin{itemize}
  \tightlist
  \item
    Phonological forms?
  \item
    We can't say how frequently words are produced by the parents or the toddler.
  \item
    Although instructed to ignore imitations, it is difficult to say whether a toddler has \emph{really} acquired a word that produces, or she can only imitate it.
  \item
    Responses in the questionnaire rely heavily on parental memory.
  \item
    We don't know about the context at which words are heard or produced.
  \item
    Classification system of items (e.g., \emph{household items}) is adult centric. Children may use the words \emph{pretty} to name jewellery (Bates et al., 1994).
  \end{itemize}
\item
  Mid and mid-upper class families are overrepresented in the sample.
\end{itemize}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{appendix-1-session-info}{%
\subsection{Appendix 1: Session info}\label{appendix-1-session-info}}

R version 3.6.3 (2020-02-29)
Platform: x86\_64-apple-darwin15.6.0 (64-bit)
Running under: macOS Catalina 10.15.4

Matrix products: default
BLAS: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRblas.0.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRlapack.dylib

locale:
{[}1{]} en\_US.UTF-8/en\_US.UTF-8/en\_US.UTF-8/C/en\_US.UTF-8/en\_US.UTF-8

attached base packages:
{[}1{]} stats graphics grDevices utils datasets methods base

other attached packages:
{[}1{]} rlang\_0.4.6 here\_0.1 data.table\_1.12.8\\
{[}4{]} truncnorm\_1.0-8 wesanderson\_0.3.6 patchwork\_1.0.0.9000
{[}7{]} ggridges\_0.5.2 ggplot2\_3.3.1 lubridate\_1.7.8\\
{[}10{]} tidyr\_1.0.3 readxl\_1.3.1 stringr\_1.4.0\\
{[}13{]} forcats\_0.5.0 dplyr\_1.0.0 tibble\_3.0.1\\
{[}16{]} magrittr\_1.5 knitr\_1.28 papaja\_0.1.0.9942

loaded via a namespace (and not attached):
{[}1{]} Rcpp\_1.0.4.6 pillar\_1.4.4 compiler\_3.6.3 cellranger\_1.1.0
{[}5{]} plyr\_1.8.6 tools\_3.6.3 digest\_0.6.25 evaluate\_0.14\\
{[}9{]} lifecycle\_0.2.0 gtable\_0.3.0 pkgconfig\_2.0.3 yaml\_2.2.1\\
{[}13{]} xfun\_0.14 withr\_2.2.0 generics\_0.0.2 vctrs\_0.3.0\\
{[}17{]} rprojroot\_1.3-2 grid\_3.6.3 tidyselect\_1.1.0 glue\_1.4.1\\
{[}21{]} R6\_2.4.1 rmarkdown\_2.1 bookdown\_0.18 purrr\_0.3.4\\
{[}25{]} backports\_1.1.7 scales\_1.1.1 ellipsis\_0.3.1 htmltools\_0.4.0
{[}29{]} colorspace\_1.4-1 stringi\_1.4.6 munsell\_0.5.0 crayon\_1.3.4

\hypertarget{appendix-2-model}{%
\subsection{Appendix 2: Model}\label{appendix-2-model}}

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-bates1994}{}%
Bates, E., Marchman, V., Thal, D., Fenson, L., Dale, P., Reznick, J. S., \ldots{} Hartung, J. (1994). Developmental and stylistic variation in the composition of early vocabulary. \emph{Journal of Child Language}, \emph{21}(1), 85--123. \url{https://doi.org/10.1017/S0305000900008680}

\leavevmode\hypertarget{ref-bosch2014}{}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in bilingual toddlers' expressive vocabulary: Does form similarity matter? \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\hypertarget{ref-floccia2018}{}%
Floccia, C., Sambrook, T. D., Luche, C. D., Kwok, R., Goslin, J., White, L., \ldots{} Plunkett, K. (2018). I: Introduction. \emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29. \url{https://doi.org/10.1111/mono.12348}

\endgroup

\end{document}

### 2020-06-15 Code/R/00_pool.R

@@ -1,0 +1,51 @@
+# 00_items: Import statistics of words in pool ##########
+# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
+# Center for Brain and Cognition, Pompeu Fabra University
+
+#### set up #############################################
+
+# load packages
+library(data.table)    # for importing and exporting data
+library(dplyr)         # for manipulating data
+library(tidyr)         # for reshaping datasets
+library(readxl)        # for importing Excel files
+library(stringr)       # for working with character strings
+library(ggplot2)       # for visualising data
+library(wesanderon)    # for nice colours
+library(here)          # for locating files
+
+# load/create functions
+source(here("Code", "R", "functions.R"))
+
+#### import data ########################################
+
+# import pool
+dat <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
+  mutate(frequency = as.numeric(frequency),
+         meaning = as.character(meaning),
+         cognate_rater1 = ifelse(cognate_rater1==1, "Cognate", "Non-cognate")) %>%
+  drop_na(language, cognate_rater1)
+
+#### visualise data #####################################
+
+dat_freqdiff <- dat %>%
+  select(item, meaning, language, frequency, cognate_rater1) %>%
+  pivot_wider(id_cols = c(cognate_rater1, meaning), names_from = language,
+              values_from = frequency,
+              names_sep = "_") %>%
+  unnest(cols = c(Catalan, Spanish)) %>%
+  rowwise() %>%
+  mutate(frequency_diff = abs(Catalan-Spanish)) %>%
+  pivot_longer(cols = c(Spanish, Catalan), names_to = "language", values_to = "frequency")
+
+ggplot(dat_freqdiff, aes(x = language, y = frequency, colour = frequency_diff, alpha = frequency_diff)) +
+  facet_wrap(~cognate_rater1) +
+  geom_violin(fill = "black", colour = NA, alpha = 0.4, na.rm = TRUE) +
+  geom_line(aes(group = meaning), na.rm = TRUE, size = 0.5) +
+  geom_boxplot(width = 0.05, na.rm = TRUE, colour = "black") +
+  labs(x = "Language", y = "Frequency (Zipf score)", colour = "Absolute\ndifference") +
+  scale_colour_gradientn(colours = wes_palette("Zissou1")) +
+  scale_alpha(guide = FALSE) +
+  theme_custom +
+  theme(legend.position = "right") +
+  ggsave(here("Figures", "00_pool_freq_cognateness.png"))

### 2020-06-15 Code/R/04_prepare.R

@@ -1,0 +1,63 @@
+#### 04_prepare: Prepare data for analysis ################
+# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
+# Center for Brain and Cognition, Universitat Pompeu Fabra
+
+#### set up ###############################################
+
+# load packages
+library(data.table) # for importing data
+library(dplyr)      # for manipulating data
+library(tidyr)      # for rehsaping datasets
+library(readxl)     # for importing Excel spreadsheets
+library(mice)       # for imputing data
+library(purrr)      # for working with lists
+library(here)       # for locating files
+
+# create functions
+source(here("Code", "R", "functions.R"))
+
+# set params
+age_bins <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "30-32")
+
+#### import data #########################################
+pool <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
+    mutate(include = as.logical(include)) %>%
+    filter(include) %>%
+    drop_na(cognate_rater1)
+
+dat_familiarity <- fread(here("Data", "03_familiarity.csv"), quote = '"', na.strings = c("", "NA"))
+
+#### aggregate data ######################################
+dat_aggregated <- dat_familiarity %>%
+    group_by(meaning, item, type, age_bin, language, item_dominance, frequency, lp, cognate_rater1) %>%
+    summarise(n = n(),
+              successes = sum(response, na.rm = TRUE),
+              failures = sum(n, -successes, na.rm = TRUE)) %>%
+    ungroup() %>%
+    mutate(proportion = successes/n) %>%
+    arrange(meaning, item_dominance, lp, age_bin)
+
+
+#### filter data ########################################
+dat_filtered <- dat_aggregated %>%
+    filter(item %in% pool$item,
+           lp != "Other",
+           age_bin %in% age_bins,
+           item_dominance == "L1") %>%
+    right_join(expand(., item, age_bin, item_dominance, lp),
+               by = c("item", "age_bin", "item_dominance", "lp")) %>%
+    select(-item_dominance)
+
+#### impute data ########################################
+dat_imputed <- dat_filtered 
+#mice(data = .,
+#    method = "norm",
+#     m = 5,
+#     maxit = 5, 
+#     seed = 888) %>%
+#complete() %>%
+#as_tibble() %>%
+#mutate_at(vars(n, successes, failures), round, 1)
+
[…]

### 2020-06-30 Code/Rmd/data_analysis_notes.Rmd
First version of the Introduction and Methods, and second version of the ICIS 2020 poster

@@
 In case the expanded model proves more inadequate, we will fall back to the simpler, previous model.
+
+## Limitations
+
+* Cross-sectional data for longitudinal claims
+* Data is right censored    
+* Inventory:
+    - Phonological forms?
+    - We can't say how frequently words are produced by the parents or the toddler.
+    - Although instructed to ignore imitations, it is difficult to say whether a toddler has *really* acquired a word that produces, or she can only imitate it.
+    - Responses in the questionnaire rely heavily on parental memory.
+    - We don't know about the context at which words are heard or produced.
+    - Classification system of items (e.g., *household items*) is adult centric. Children may use the words *pretty* to name jewellery [@bates1994].
+* Mid and mid-upper class families are overrepresented in the sample.

### 2020-06-30 Manuscript/trajectories_manuscript.Rmd
First version of the Introduction and Methods, and second version of the ICIS 2020 poster

@@
   - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
+    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
 
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
+  "Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
 
 abstract: |
-  Previous literature on early vocabulary has been mostly committed to exploring how the receptive and productive vocabulary size changes with age, and its relationship with toddlers’ performance on language tasks (e.g.
-Fernald, Swingley, & Pinto, 2001; Fernald, Perfors, & Marchman, 2006).
-The content of the developing lexicon has remained relatively unexplored.
-More recently, studies have focused on the characterisation of the developmental trajectory of individual words, reporting an earlier age of acquisition for words with high frequency, concreteness, and phonological neighbourhood density (e.g.
-Braginsky, Yurovsky, Marchman, & Frank, 2019; Jones & Brandt, 2019).
-Such item-level analysis allows not only to predict the age of acquisition of specific words, but also to shed light on the cognitive processes that underlie early word learning.
-This approach is particularly interesting for investigating the potential impact of acquiring two languages simultaneously.
-A recent study by Floccia et al.
-(2018) compared vocabulary sizes of 24-month-old children learning British English, together with an additional language from a pool of 13 diverse languages.
-They found larger productive vocabulary sizes of toddlers learning two languages that are phonologically similar (see for a similar approach Bosch & Ramon-Casas, 2014).
+ Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
+Floccia et al.
+(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
 The mechanisms underlying this effect remain unknown.
-One possibility is that the similarity between the two languages speeds the acquisition of form-similar translation equivalents.
-The aim of this study is to perform an item-wise analysis on infant’s vocabulary contents.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish.
-We expect that phonological overlap between translation equivalents will predict earlier age of acquisition.
-If this effect is driven by the phonological overlap between translation equivalents, cognate pairs should be acquired closer in time than non-cognates.
-For instance, the translation of cat /gato/ in Spanish and /gat/ in Catalan should be learnt at approximately the same age.
-This should not necessarily happen for the translations of dog, /pero/ and /gos/.
-We analyse the time elapsed between the acquisition of a word in one language, and its translation in the other.
-  We present preliminary data (data collection is ongoing) on receptive vocabulary of 90 monolingual and bilingual toddlers aged 18 to 36 months.
-We obtained parental responses to 230 pairs of Catalan-Spanish translation equivalents, resulting in a total of 41490 responses.
-Words that yielded a phonologically similar translation equivalent were more likely to be present in toddlers’ receptive vocabulary at all ages, and were acquired earlier, even when accounting for the effect of word frequency.
-Further analysis will explore the distance in time between the acquisition of pairs of translation equivalents using a more detailed measure of phonological overlap across translation equivalents and taking into account cross-individual and cross-item variability.
+We explore two compatible scenarios.
+First, we test whether cognates are learnt earlier than non-cognates.
+This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
+Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
+This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
+We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
 
@@
 
-keywords          : "word acquisition, vocabulary, bilingualism, lexical similarity"
+keywords          : "lexical acquisition, vocabulary, bilingualism"
 wordcount         : "X"
 
-bibliography      : ["../References/BiLexicon.bib"]
+bibliography      : ["../References/trajectories.bib"]
 
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            :
+  word_document:
+    reference_docx: trajectories_template.docx
+
 ---
@@
 	echo = FALSE,
+	fig.align = "center",
 	message = FALSE,
@@
 	cache.extra = knitr::rand_seed,
-  results = "asis",
-  out.width = "60%",
-  fig.align = "center"
+	out.width = "80%",
+	results = "asis"
 )
@@
 
-```{r prepare, include=FALSE}
+```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
 # load packages
@@
 library(magrittr)
-library(tibble)
+library(tidyverse)
 library(dplyr)
@@
 library(readxl)
-library(tidyr)
 library(lubridate)
+library(janitor)
 
-library(ggplot2)
 library(ggridges)
 library(patchwork)
-library(wesanderson)
+library(gt)
 
-library(truncnorm)
 library(data.table)
@@
 # load/create functions
-source(here("R", "functions.R"))
+source(here("Code", "R", "functions.R"))
 
 # set params
-age_bins <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "30-32")
+bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
 
-```
-
-```{r prepare_data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# prepared data
-dat_items <- fread(here("Data", "04_prepared.csv"), na.strings = c("")) %>%
-  as_tibble() %>%
-  drop_na(language)
-
-# load data
-dat <- fread(here("Data", "02_merged.csv"), na.strings = "") %>%
-  filter(source %in% c("formr1", "formr2", "formr_short"),
-         lp %in% c("Monolingual", "Bilingual"),
-         age_bin %in% age_bins) %>%
+# import data 
+dat_studies <- fread(here("Data", "00_studies.csv")) %>% distinct(q_version, language, q_items)
+dat_items <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
+  group_by() %>%
+  drop_na(language) %>%
+  mutate_at(vars(A, B, C, D, cognate_rater1, include, frequency), as.numeric) %>%
+  mutate_at(vars(A, B, C, D, cognate_rater1, include), as.logical)
+dat_participants <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
+  last() %>%
+  fread(na.strings = "") %>%
+  filter(age_bin %in% bins,
+         completed,
+         lp == "Bilingual")
+dat_raw <- fread(here("Data", "02_merged.csv"), na.strings = "") %>%
+  filter(lp %in% c("Bilingual"),
+         age_bin %in% bins) %>%
   as_tibble() %>%
@@
 
-# import data
-wordbank <- fread(here("Data", "05_wordbank.csv"))
-wordbank_priors <- fread(here("Data", "05_wordbank-priors.csv"))
-mid_prior <- round(wordbank_priors$estimate[wordbank_priors$term=="mid"], 2)
-mid_steep <- round(wordbank_priors$estimate[wordbank_priors$term=="steep"], 2)
-
-# load results
-bayes_factor <- readRDS(here("Results", "bayes_factor.rds"))
-loo_comp <- readRDS(here("Results", "loo.rds"))
-fit0_production <- readRDS(here("Results", "fit0_productive.rds"))
-fit1_production <- readRDS(here("Results", "fit1_productive.rds"))
-fit1_priors_production <- readRDS(here("Results", "fit1_prior_productive.rds"))
+dat_wordbank <- fread(here("Data", "05_wordbank.csv"))
+dat_wordbank_priors <- fread(here("Data", "05_priors-wordbank.csv"))
 
@@
 
-One of the main challenges bilingual infants and toddlers face is learning two distinct set of words (one for each language) that partially overlap in sound and meaning.
+Bilingual toddlers must learn two distinct set of words (one for each language) that partially overlap in sound and meaning.
 Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When taking both languages into account, bilingual children between seem to know, at least, as many words as monolinguals do [@benzeev1977; @doyle1978; @rosenblum1983; @fernandez1992; @pearson1994; @bialystok2010; @blom2019].
-Tough, not all bilinguals show asimilar developmental trajectory of lexical acquisition.
+When taking both languages into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev1977; @doyle1977some; @rosenblum1983word; @fernandez1992; @pearson1994; @bialystok2010; @hoff2012; @core2013; @hoff2014; @blom2019; but see @pearson1993; @houwer2014].
+Not all bilinguals show a similar developmental trajectory of lexical acquisition.
+Recent studies have capitalised on the role of the similarity between the specific pair of languages the infant is learning.
+One way of measuring language similarity is by estimating the amount of cognates (i.e., form-similar translation equivalents) they share.
+@floccia2018 explored the impact of this form of similarity on vocabulary sizes from a large sample of 24-month-old toddlers learning English and an additional language.
+Toddlers learning languages that shared a large amount of cognates (e.g., English-German) showed larger vocabulary sizes than those learning languages that shared fewer cognates (e.g., English-Spanish).
+@blom2019 extended these findings to children aged three to 10 years.
+These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
 
-Recent studies have capitalised on the role of the similarity between the specific pair of language the infant is learning.
-@floccia2018 analysed vocabulary scores across of 24-month-old toddlers learning English and an additional language.
-Toddlers learning languages sharing a large amounth of cognates (e.g., English-German) showed larger vocabulary sizes than those learning less languages sharing less cognates (e.g., English-Chinese).
-@blom2019 extended these results to children aged three to 10 years.
-These results point to the possibility that the overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
-However, none of these studies provide an account for the mechanisms involved in this facilitatory effect of language driven by similarity across languages.
-The aim of this study is to explore two scenarios in which this effect may be taking place:
+How does cognateness influence word learning?
+@floccia2018 point to the possibility that parallel activation during speech perception may play a role in this cognate facilitation effect.
+The parallel activation principle suggests that bilinguals activate both languages in parallel during speech perception and production, even when only one of them is being used.
+This hypothesis can only be understood under the more general account of language non-selectivity of lexical processing, which states that translation equivalents are co-activated at the lexical level [e.g., @de1991lexical; @costa2000; @dijkstra2005; @thierry2007; @hoshino2008; @yudes2010; @bobb2020; see @kroll2013 for review].
+A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model [@kroll1994; @kroll2010], the Bilingual Interactive Activation model [BIA; @dijkstra1998bia], and its successor BIA+ [@dijkstra2002architecture], the Bilingual Language Interaction Network for Comprehension of Speech model [BLINCS; @shook2013] and, more recently, Multilink [@dijkstra2019], which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
+However, these models do not address word learning during language acquistion.
 
-1) In a first scenario, it is the phonological overlap between language pairs, and not the amount of cognates they share, what boosts the vocabulary growthof bilingual toddlers.
-Languages sharing a lot of cognates tend to also share a lot of phonemic categories.
-It is possible that having to learn fewer phonemic categories makes it easier for toddlers to acquire words.
-If this is true, toddlers learning two languages with similar phonemic inventories should show larger vocabulary sizes than those learning languages that barely ovelap phonologically.
-Moreover, this facilitation effect should be reflected in the developmental trajectory of all the items included in a vocabulary inventory, independently of their cognate status.
+It is feasible that parallel activation plays a role in vocabulary acquisition during the first years of life.
+On top of Floccia et al.'s findings, other studies have provided experimental evidence of parallel activation during speech perception in bilingual toddlers [@vonholzen2012; @vonholzen2019].
+Yet, it is unclear how parallel activation influences word learning at these ages.
+What mechanisms are engaged, and how do they relate to vocabulary size?
+We contemplate two scenarios.
+First, it is possible that the reported "similarity boost" on vocabulary size is not directly related to the amount of cognates shared between both languages, but rather to the overlap across their correspondent phonological inventories.
+One of the measures of language similarity Floccia et al.
+computed is the mean phonological Levenshtein distance across translation equivalents.
+This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
+These scores were aggregated across pais of tanslation equivalents, obtaining a single similarity estimate of for each language pair.
+This measure embeds confounded information about the "degree of cognateness" and the amount of phonological overlap between both languages.
+Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, Enlgish-Dutch), with few exceptions (e.g., Spanish-Basque).
+In the light of these data, it is not possible to discern whether vocabulary size is increased by the ealier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
+We address this issue by testing two hypotesis.
 
-2) In a second scenario, Words that are form similar It is the amount of cognates language pairs share that boosts vocabulary growth in bilingual toddlers.
-If this is true, toddlers may acquire cognates earlier
+## Hypothesis 1: Cognates are acquired earlier than non-cognates
 
-To our knowledge, there are no previous studies that have investigated the effect of cognateness on the acquisition trajectories of individual word forms.
-Braginsky et al.
-(2019) used data from MacArthur-Bates CDIs from multiple languages to explore what properties are associated with an earlier acquisition.
-The found that frequency, concreteness, and [...] significantly predicted an earlier age of acquisition.
-We plan to extend these analyses to the case of bilinguals.
+First, we test whether the increase in vocabulary size reported by Floccia et al.
+for languages sharing a greater amount of cognates is indeed caused by cognates being acquired earlier.
+In line with this account, @bosch2014 found higher productive vocabulary scores in 18 month-old Catalan-Spanish bilinguals, compared to monolinguals, due to the higher presence of cognates their lexicon.
+We will adopt a similar approach to @braginsky2019, who modelled cross-sectional data from individual items to explore what properties were associated with an earlier acquisition.
+We will analyse the acquisition trajectories of a selection of `r dat_items %>% filter(include) %>% group_by(language) %>% summarise(n = n(), .groups = "drop") %>% pull(n) %>% max()` translation equivalents in a cross-section sample of Catalan/Spanish bilinguals aged 14 to 34 months of age.
+This approach allows to analyse how the proportion of infants that know each item changes across ages, providing a more suitable framework to put developmental hypotheses to test, instead of taking a "snapshot" of the state of the lexical of children at a single age.
 
-We have developed a fully-online vocabulary inventory in Catalan and Spanish that is more exhaustive than the MacArthur-Bates CDI, including a larger number of items.
-All participants filled the questionnaire it both languages, thus providing responses to pairs of translation equivalents.
+## Hypothesis 2: Cognates are acquired closer in times than non-cognates
 
-In this study, we examine the role of cognateness on age of acquisition by modelling the probability of parents reporting that the participant is able to understand, say and understand on each item in the questionnaire.
-We will test the effect the participants' language profile (monolingual vs.
-bilingual), and its cognate status on the developmetnal curves of each item.
-
-We will analyse the developmental trajectories of individual words in monolinguals and bilinguals learning Catalan and Spanish (two phonologically close languages sharing a great deal of cognates), between 14 and 30 months of age.
-We will test whether cognate words are learnt earlier than non-cognate words, taking into account wether they are part of the most- o less-dominant language of the infant, as well as their frequency, concreteness, and number of syllables.
-We have deloped an on-line questionnaire that entails a short language exposure questionnaire, asks for some demographic information, and presents parent with a list of ~800 words in each language, for which parent report whether their child understands, says, or doesn't understand and say, in each item.
-
-We will fit a Bayesian multilevel model on the item responses, modelling the probability of parents reporting that a given word is acquired by the infant, including language profile (monolingual vs.
-bilingual), item dominance (dominant language vs.
-non-dominant language), frequency, concreteness, and number of syllables.
-We will include *meaning* as grouping variable.
-We will fit a sigmoidal curve [see @mayor2008] on each item, defined by the steepness of the developmental curve, and the mid-point (the poinnt at which the developmental curve is steepest, which will be considered the point of acquisition).
-Priors will be derived from previous literature on the subject.
-
-NOTES:
-
-Evidence that bilingual lexica barely overlap [@leopold1948; @taeschner1978].
-
-Studies with infants from 8 to 30 months of age show weaker evidence for this claim [@pearson1993; @dehoouwer2014], pointing to the posibility that differences between bilingual and monolingual vocabulary sizes may be dependent on maturational factors.
-
-sugested that the developmental trajectory of lexical aquisition varies across two dimensions: the language profile of the infant, and the joint proporties of the specific language pair the infant is learning.
-Regarding the first, @pearson1994 found thant infants exposed to a more balanced
-
-Of special interest is the fact than cognates (form-similar translation equivalents) tend to be overrepresented in the early bilingual vocabulary [@bosch2014].
-
-The similarity between both languages may play a role in bilingual lexical aquisition.
-
-## Hypotheses
-
-Hypothesis 1
-:  The boost effect of language similarity is driven by cognates being acquired earlier than non-cognates.
-We predict that age of acquisition of cognates will be ealier thant that of non-cognates.
-This effect should only be present in bilinguals.
-
-Hypothesis 2
-:  That this effect is driven by the word-forms in one language scaffolding the acquisition of their translation equivalents via parallel activation.
-We predict that the difference in time between the acquisition pairs of translation equivalents is shorter in cognates than in non-cognates.
-This effect will only be present in bilinguals.
+One mechanism that could acccount for an ealier acquisition of cognates is that parallel activation makes word-forms acquired in one language available during the acquisition of word forms in the other language.
+One consequence of this situation is that translation equivalents scaffold the acquisition of each other.
+This would point to different mechanisms engaged in the acquisition of cognates vs.
+non-cognates.
+While the phonology of cognate translation equivalents overlaps, and thus could be co-activated, the phonology of non-cognates does it in a much lesser degree.
+Consider the case of a Catalan-Spanish bilingual toddler that has already acquired the Catalan words *porta* (/ˈpɔr.tə/, door) and *taula* (/ˈtaw.ɫə/, table), but not yet their correspondent translation equivalents in Spanish (*puerta*, /ˈpwer.ta/; *mesa* /ˈme.sa/).
+According to our hypothesis, hearing the word *puerta* in presence of a door should lead to the activation of *porta*, which should facilitate the acquisition of the former.
+On the contrary, when hearing *mesa* in presence of a table, it is unlikely that the phonology of *taula* will be activated.
+This mechanism should lead to cognate translation equivalents being learnt closer in time than non-cognate translation equivalents.
+To test this hypothesis, we will extract the age of acquisition of each word-form in the questionnaire (as estimated by the model used to test Hypothesis 1), and will calculate the difference in time between the acquisition of such word-forms and their translation equivalents.
+We will test how cognateness affects such outcome.
 
@@
 
-Data from `r dat$id_db %>% unique %>% length() %>% printnum()` participants from the different cities in Spain were invited to participate through social media.
-Families in the Metropolitan area of Barcelona were also recruited at birth.
+We collected data from `r dat_raw$id_db %>% unique %>% length() %>% printnum()` bilinguals from the Metropolitan Area of Barcelona, between `r min(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`.
 All families participated voluntarily.
-Data were collected between `r min(dat$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`, using the BiLexicon inventory.
 This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
 
-```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+We assessed toddlers' language background asking parents for an estimated proportion of exposure to each language.
+We defined participants as Spanish- or Catalan-dominant according to the language they were exposed the most to.
+Participants were considered bilingual - therefore included - if exposed 50-90% of the time to the non-dominant language.
+We excluded participants with >10% exposure to a third language.
+Fig.
+\@ref(fig:lpdist) shows how participants were distributed across age bins and language profiles.
 
-dat %>%
-  group_by(dominance, sex) %>%
-  summarise(n = n(),
-            age_mean = mean(age, na.rm = TRUE),
-            age_sd = sd(age, na.rm = TRUE),
-            doe_catalan_mean = mean(doe_catalan, na.rm = TRUE),
-            doe_catalan_sd = sd(doe_catalan, na.rm = TRUE),
-            doe_spanish_mean = mean(doe_spanish, na.rm = TRUE),
-            doe_spanish_sd = sd(doe_spanish, na.rm = TRUE)) %>%
+```{r lpdist, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+dat_participants %>%
+  filter(completed, lp %!in% "Other", age_bin %in% bins) %>%
+  ggplot(aes(x = age_bin, fill = dominance)) +
+  geom_bar() +
+  labs(x = "Age (months)", y = "Number of partcipants",
+       fill = "Dominant language") +
+  scale_y_continuous(labels = as.integer) +
+  scale_fill_brewer(palette = "Set1") +
+  theme_custom +
+  theme(legend.position = c(0.05, 0.85),
+        legend.direction = "vertical",
+        legend.justification = 0,
+        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))
+```
+
+## Questionnaire
+
+Parents completed an online questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
+We implemented the questionnaire in `formr` [@arslan2020], a free, open-source platform.
+Participants were randomly allocated into one of four list of items.
+Each list contained a distinct inventory of translation equivalents.
+We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/funcional categories:  Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
+We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
+Those categories that did not entail more than 16 items - thus resulting in less than four items after dividing it in four lists - were not divided: all of their items were included in the four lists.
+Another subset of items that were part of the trial lsits of some experiments in the lab were also included in all versions.
+
+Each item of the inventory presented a word-form, to which parents were instructed to report whether the toddlers was able to understand it, understand ans say it, or neither.
+The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates) .
+Items of one language were translation equivalents from the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+
+## Data analysis
+
+For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item  within age bins of 2 months.
+Our model assumes that this proportion follows a logistic curve accross ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the dominance of the language of the item (*item dominance*, L1 vs.
+L2), the cognate status of the word (*cognateness*, Non-cognate vs.
+Cognate), and the frequency of the word form in its correspondent language.
+Our prediction  is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words.
+
+Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
+Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item porta (door, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
+We retrieved frequencies from SUBTLEX-ESP [@cuetos2011] and SUBTLEX-CAT [@boada2019] for Spanish and Catalan words, respectively.
+Word frequency was calculated as the Zipf score of the word [@zipf1949; @heuven2014].
+Missing frequency scores were imputed using multiple imputation [@vanbuuren2011].
+Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
+
+```{r itemsinfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+dat_items %>%
+  drop_na(cognate_rater1) %>%
+  filter(include) %>%
+  select(A, B, C, D, cognate_rater1, frequency, language) %>% 
+  mutate(cognate_rater1 = ifelse(cognate_rater1, "Cognate", "Non-cognate")) %>%
+  pivot_longer(c(A, B, C, D), names_to = c("list"), values_to = "n") %>% 
+  filter(n) %>% 
+  group_by(list, language, cognate_rater1) %>% 
+  summarise(n = sum(n, na.rm = TRUE),
+            mean = mean(frequency, na.rm = TRUE),
+            sd = sd(frequency, na.rm = TRUE),
+            min = min(frequency, na.rm = TRUE),
+            max = max(frequency, na.rm = TRUE)) %>%
   ungroup() %>% 
-  kable(digits = 2,
-        col.names = c("Dominance", "Sex", "N", "M_Age", "SD_Age", "M DOE Cat", "SD_DOE_Cat",  "M_DOE_Spa", "SD_DOE_Spa"),
-        format.args = list(big.mark = ",", na.string = "-"),
-        align = "c",
-        caption = "Sample demographics and overall language profile.
-Dominance = Language to which the participant is exposed the most, N = sample size, M age = Mean age, SD age = standard deviation of age, M DOE Cat = mean degree of exposure to Catalan, SD DOE Cat = standard deviation of exposure to Catalan, M DOE Spa = mean degree of exposure to Spanish, SD DOE Spa = standard deviation of exposure to Spanish.")
+  rename(cognateness = cognate_rater1) %>%
+  relocate(list, language, cognateness, n, mean, sd, min, max) %>%
+  arrange(list, language, cognateness) %>%
+  mutate(list = paste("List", list)) %>% 
+  pivot_wider(names_from = cognateness, values_from = c(n, mean, sd, min, max)) %>%
+  clean_names() %>%
+  gt(groupname_col = c("language")) %>% 
+  fmt_number(columns = contains("cognate"), decimals = 2) %>%
+  fmt_number(columns = starts_with("n_"), decimals = 0) %>%
+  tab_header(title = md("**Questionnaire and item properties**"),
+             subtitle = "Participants were randomly allocated on of the the four lists, and completed both the Catalan and the Spanish versions.") %>% 
+  tab_spanner(label = "Frequency (Cognates)", columns = vars("n_cognate", "mean_cognate", "sd_cognate", "min_cognate", "max_cognate")) %>% 
+    tab_spanner(label = "Frequency (Non-cognates)", columns = contains("non_cognate")) %>% 
+  cols_label(list = md("**List**"), 
+             n_cognate = md("*N*"),
+             mean_cognate = md("*Mean*"),
+             sd_cognate = md("*SD*"),
+             min_cognate = md("*Min*"),
+             max_cognate = md("*Max*"),
+             n_non_cognate = md("*N*"),
+             mean_non_cognate = md("*Mean*"),
+             sd_non_cognate = md("*SD*"),
+             min_non_cognate = md("*Min*"),
+             max_non_cognate = md("*Max*")) %>%
+  gtsave(here("Results", "descriptives_items.png"))
 
@@
 
-* *Location*
-* *Age*
-* *Sex*
-* *Language profile*: : Linguistic profiles were assessed using the Language Exposure Questionnaire (LEQ) [@bosch2001].
-This questionnaire estimates de degree of exposure (DOE) to each language via a 10 minutes long interview with the parents.
-LEQs were administered inmediately before or after the testing session at the lab.
-We here considered only exposure to Catalan and Spanish.
-Participants with >10% exposure to a third language other than Catalan or Spanish were excluded.
-* *SES*
+We fit two separate logistic models on comprehensive and productive data, respectively.
+Logistic curves are characterised by three parameters[see @mahr2019 and @mahr2020 for an excellent introduction to this approach, in which we based this analysis]: (a) an *asymptote* (`asym`, upper boundary  of the proportion of infants that have acquired the item), (b) the *steepness* (`steep`, how fast the a given item is acquired by infants), and (c) a *mid-point* (`mid`, the age at which steepness is maximum).
+We work on the assumption that the trajectory of aquisition of items follows a sigmoidal shape [@mayor2011], and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
+This approach allows us to model and estimate the age of acquisition of items explicitely.
+Our model estimated the asymptote and stepness as grand means across items (i.e., intercepts).
+Mid-points were estimated by a linear zero-one-beta-regression model^[Zero-or-one-inflated beta regression models allow to account for overdispersion then using proportions as the output of a model accounting, while also allowing values to be at boundary (0 and 1).][@ospina2012] that included our predictors of interest: `age`, `dominance` (contrast-coded, L2 = -0.5, L2 = 0.5), `cognateness` (contrast-coded, Non-cognate = -0.5, Cognate = 0.5), their interaction, and `frequency` (standardised Zipf score).
+We also included a random intercept for each translation equivalent (`TE`).
+Appendix 1 shows a detailed description of the model.
 
-### Exclusion criteria
+We estimated the parameters of our model using the Bayesian framework.
+Bayesian models use the Bayes rule to combine prior knowlegde about the distribution of a given parameter, and the likelihood of each of the values it can take given tha observed data.
+We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our priors).
+In order to test the contribution of the predictor *cognateness* on the fit of the model,  we compared the fit of our (expanded) model to the fit of a (null) model that did not include *cognateness* as predictor [@schad2020] using leave-one-out cross validation.
+We fit both the null and alternative models using the R environment [@rcoreteam2013; @rstudioteam2015] and the `brms` package [@burkner2017], which relies on the probabilistic language Stan [@carpenter2017] to approximate posterior distributions.
+We used Pareto-smoothed importance sampling [PSIS; @vehtari2017; @vehtari2019] to compare the null and extended models.
+Data and model results were processed and visualised using the `tidyverse` family of R packages [@wickham2019] and the `tidybayes` package [@kay2020].
 
-* Language disorders
+To test Hypothesis 2, we run a Bayesian ANOVA using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included *cognateness* as predictor.
+Using the `BayesFactor` R package [@morey2018a], we computed a Bayes factor (*BF*) that compared the likelihood of a linear model that included *cognateness* as a predictor against the likelihood of a linear model that did not, under the light of our data.
 
-### Questionnaires
+# Results
 
-#### BiLexicon 1.0
+## Hypothesis 1
 
-#### BiLexicon 2.0
+### Comprehensive vocabulary
 
-#### BiLexicon Short
+### Productive vocabulary
 
-This questionnaire was developed as a short version of the BiLexicon 1.0 and BiLexicon 2.0 inventories.
-Given the large amount of items to be answered to, we restricted the use of such inventories to participants that were to participate presentially in one of the studies run in the lab at that time.
-These participants only filled one of the questionnaires, and did so only once.
+## Hypothesis 2
 
-The aim of developing a short version of this questionnaire (BiLexicon Short) was to provde a more on-line friendly format to families that would participate remotely exclusively, or that would fill the questionnaire more than once (e.g., in longitudinal studies).
-We divided the items of each language in the pool into four versions named A, B, C, and D.
-Items in each category (e.g., animals, auxiliary words) were randomly assigned one of the versions.
-Thos categories that did not entail more than 16 items (thus resulting in less than four items after dividing it in ofur lists), were not divided (all of their items were included in all versions).
+### Comprehensive vocabulary
 
-When filling the questionnaire for the first time, participants were randomly assigned one version.
-Each familiy filled only one of the versions, thus reducing the number of items to be answered.
-To ensure that longitudinal participants would provide longitudinal data in the questionnaire, we made sure that participants were always assigned the same version of this questionnaire.
-Words that were used as part of the trial lists in the experiment families would participate in were present across all versions.
+### Productive vocabulary
 
-### Item inclusion criteria
+# Discussion
 
-We gathered data from `r dat_items %>% distinct(language, meaning) %>% group_by(language) %>% summarise(n = n()) %>% pull(n) %>% min() %>% printnum(round = 0)` distinct translation equivalents across Spanish (`r dat_items %>% distinct(language, meaning) %>% group_by(language) %>% summarise(n = n()) %$%  n[2] %>% printnum()` word forms) and Catalan (`r dat_items %>% distinct(language, meaning) %>% group_by(language) %>% summarise(n = n()) %$% n[1] %>% printnum()` word forms).
-The unequal number of word-orms across both languages is due to that some tanslation equivalents did not entail a one-to-one-corespondence: some items had more than one translation equivalent in the other language (e.g.
-the Catalan word form *walking*, *anar*, can be translated as both *caminar* [to walk] or *ir* [to go]).
+# Appendix
 
-We included items from the following categories: Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
+## Appendix 1
 
-We excluded items that enailed more than one word (e.g., *barrita de cereales* [cereal bar]).
+Model formula and priors:
 
-### Item properties
+$$proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ \beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\\phi \sim Normal(1.5, 1)$$
 
-#### Frequencies
+## Appendix 2: Computing priors
 
-We retrieved frequencies from SUBTLEX-ESP [@cuetos2009] and SUBTLEX-CAT [@ferre2013] for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word [@zipf1949; @vanheuven2014].
+We fitted a logistic model using the proportion of infants that acquired each item as output, and age-bin (defined as we did with our data) as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
+We included data from `r nrow(dat_wordbank)/length(unique(dat_wordbank$age)) %>% printnum()` items from `r dat_wordbank$language %>% unique %>% length` languages^[Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
+All the other available languages were excluded, as no (or insufficient) data were available.], from monolingual infants aged 16 to 30 months of age.
+Table \@ref(tab:wordbank) shows the estimates of this model, which were rescaled and included as priors in our main model to test Hypothesis 1.
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r worbank, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-include_graphics(here("Figures", "pool_freq_cognateness.png"))
+dat_wordbank_priors %>%
+  rename(estimate_scaled = estimate) %>%
+  mutate(conf = paste0("[", round(conf_low, 2), ", ", round(conf_high, 2), "]"),
+         estimate = estimate_scaled*2+20,
+         term = c("Asymptote", "Mid-point", "Steepness")) %>% 
+  select(term, estimate, estimate_scaled, std_error, conf) %>%
+  gt(rowname_col = "term") %>% 
+  fmt_number(vars(estimate, estimate_scaled, std_error), decimals = 2) %>% 
+  cols_label(estimate = md("*Estimate<br>(months)*"),
+             estimate_scaled = md("*Estimate<br>(scaled)*"),
+             std_error = md("*SE*"),
+             conf = md("*95% CI*")) %>% 
+  cols_align(align = "c") %>% 
+  gtsave(here("Results", "wordbank.png"))
 
@@
 
-#### Cognate status
+We performed prior-predictive checks to analyse the feasibility of our priors. Figure \@ref(fig:priorchecks)
 
-#### Demographics by item
+```{r priorchecks, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-
-dat %>%
-  distinct(item, id, language, lp, sex, age) %>%
-  group_by(item, language, lp, sex) %>%
-  summarise(mean_age = mean(age, na.rm = TRUE),
-            sd_age = mean(age, na.rm = TRUE),
-            min_age = min(age, na.rm = TRUE),
-            max_age =  max(age, na.rm = TRUE),
-            n = n()) %>%
-  group_by(language, lp, sex) %>%
-  summarise(mean_age = mean(mean_age, na.rm = TRUE),
-            sd_age = mean(sd_age, na.rm = TRUE),
-            min_age = min(min_age, na.rm = TRUE),
-            max_age =  max(max_age, na.rm = TRUE),
-            mean_n = mean(n, na.rm = TRUE),
-            n = sum(n, na.rm = TRUE)) %>%
-  kable(.,
-        digits = 2,
-        col.names = c("Language", "LP", "Sex", "M Age", "SD Age", "Min Age", "Max Age","M N", "N"),
-        align = "c",
-        caption = "Demographics of responses by item.
-Language = item language, LP = language profile of participant, Sex = sex of participant, M Age = mean age of responses, SD Age = mean SD of responses across items, Min Age = minimum age of responses across items, Max Age = maximum age of responses across items, M N = mean number of responses across items, N = total number of responses across items."
-  )
+include_graphics(here("Figures", "06_simulate-priors.png"))
 
@@
 
-## Data analysis
+```{r priorstab, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-For testing the first hypothesis, we fit a non-linear model using logistic curves to model the probability of aquisition of words across ages.
-Logistic curves are characterised by three parameters: 1) an asymptote (maximum value of the curve in the Y-axis), 2) the steepness (how fast the curve grows), and 3) a mid-point (the point in the X-axis at which the steepness is highest).
-Figure 1 shows the elements of a simulated logarithmic curve:
-
-```{r logcurve, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, fig.cap="Simulated logarithmic curve.
-The curve is characterised by three parameters: Asymptote (red), mid-point (blue), and steepness (green).
-This curve is the result of simulating learning curves of 1000 items.
-We set fixed values for the asymptote at 1.
-Mid-points were randomly generated from a normal distribution truncated at 0 (no negative age values are allowed), with mean 23 (*SD* = 0.5), and steepness values were generated from a normal ditribution truncated at 0 (we assume that the proportion of participants that have acquired each item can only grow), with mean 0.1 (*SD* = 1) The solid black line represents the average proportion of participants who have acquired the item, and the shaded ribbon indicates the standard error of the mean.
-We suggest that learning curves of individual word-forms, as well as the evolution of vocabulary size across ages, follows this trend.
-Our model will estimate the three of to adress our hypotheses."}
-
-include_graphics(here("Figures", "05_illustration.png"))
+tribble(~param, ~coefficient, ~notation, ~prior,
+        "Asymptote", "Intercept", "$A$", paste0("$N(", dat_wordbank_priors[term=='asym'] %>% pull(estimate) %>% printnum(), ", 0.05)$"),
+        "Steepness", "Intercept", "$S$", paste0("$N(", dat_wordbank_priors[term=='steep'] %>% pull(estimate) %>% printnum(), ", 0.6)$"),
+        "Mid-point", "Intercept", "$\\beta_{0}$", paste0("$N(", dat_wordbank_priors[term=='mid'] %>% pull(estimate) %>% printnum(), ", 1)$"),
+        "Mid-point", "Random Intercepts", "$I_{i0}$", "N(0, $\\pi$)",
+        "Mid-point", "Slope", "$\\beta_{1}$", "$N(0, 5)$",
+        "Mid-point", "Slope", "$\\beta_{2}$", "$N(0, 5)$",
+        "Mid-point", "Slope", "$\\beta_{3}$", "$N(0, 5)$",
+        "Mid-point", "Slope", "$\\beta_{4}$", "$N(0, 5)$",
+        "Mid-point", "Intercept", "$\\varphi$", "$N(2, 1)$",
+        "Mid-point", "Intercept", "$COI$", "Beta(1, 1)", 
+        "Mid-point", "Intercept", "$ZOI$", "Beta(1, 1)",
+        "Mid-point", "Std. Deviation", "$\\sigma{ijk}$", "t(3, 0, 2.5)") %>%
+  kable(digits = 2, row.names = FALSE, col.names = c("**Adjusted parameter**", "**Coefficient**", "**Notation**", "**Distribution**"), align = "c", caption = "Prior distribution of the parameters in our models.
+We set specific informative priors for the asymptote, steepness, and mid-point of the curves (using data from Wordbank) and and weakly informative priors for the rest of the coefficients.", escape = TRUE, format = "latex")
 
@@
 
-One reason for using logistic curves to fit the data is that previous studies describe lexical development as a non-linear process, where between 20 and 24 months a vocabulary spurt happens.
-An analysis of the available data in Wordbank  [@frank2016] provides moderate support for this hypothesis.
-We fitted a logistic model using the proportion of infants that acquired each item as output, and age as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
-We included data from `r round(nrow(wordbank)/length(unique(wordbank$age)), 0)` items from `r wordbank$language %>% unique %>% length` languages^[Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
-All the other available languages were excluded, as no (or insufficient) data were available.], from infants between 16 and 30 months of age.
-Then we compared the fit of the logarithmic model agains a linear model, and found that the formr fitted the data slightly better (see Figure 2).
+## Appendix 3: Session info
 
-A second reason for using logistic curves is that the parameters that characterise a logistic curve map onto our hypotheses quite easily.
-Out first hypothesis concerns the mid-point parameter, which indicates the point at which the steepness of the curve is maximum.
-In our model, the value of this parameter can be interpreted as the age at which the proportion of participants that have acquired a given item is at maximum rate.
-The mathematical interpretation of this parameter is easily mapped into a definition of age of acquisition.
+```{r session_info, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-```{r wordbank, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Logarithmic model fitted on Wordbank data, compared to a linear model.
-The logarithmic model fits the data slightly better.
-The estimated values of the three parameters involved in the logarithmic model (asymptote, mid-point and steepness) are indicated."}
-
-include_graphics(here("Figures", "05_wordbank.png"))
+sessionInfo()
 
 ```
-
-We built a Bayesian logistic model that estimates the value of the mid-point and steepness across items.
-Under our hypothesis, mid-points are generated from a linear model that incorporates language profile (monolingual vs.
-bilingual) and cognateness (non-cognate vs.
-cognate) as predictors, where cognateness predictes ealier mid-points for bilingual, but not monolingual participants.
-We wil also include frequency as a covariate in this linear model.
-Due to computational limitations, we set a constant value for the asymptote at `r round(wordbank_priors$estimate[wordbank_priors$term=="asym"], 2)`, according to our model fit on Wordbank data.
-Due to computational limitations, we did not run the same linear model to estimate steepness.
-Rather, this parameter was just included as an intercept.
-We compared how well this model fits the data *versus* how a null model that does not include *cognateness* as predictor does, by using Bayes factors and Pareto-smoothed importance sampling [PSIS; @vehtari206].
-
-We used the information provided by Wordbank as priors to feed our Bayesian model.
-Accordingly, we set an informative prior for the mid-point $N(`r round(mid_prior, 2)`, 0)$ and the steepness $N(`r round(mid_steep, 2)`, 0)$ intercepts.
-We set weakly informative priors for the coefficients of frequency, language profile, cognateness, and their interaction at $N(0, 10)$.
-We also set a weakly informative prior for the standard deviation, $\sigma \sim t(3, 0, 2.5)$.
-
-We fit both null and alternative models within the R environment [@rcoreteam2019; @rstudio2020] using the `brms` package [@brueckner2016], which relies on the probabilistic language Stan [@carpenter2013] to estimate posterior distributions.
-Stan uses Hamiltonian Montecarlo to sample the posterior.
-for model selection, we used both the `brms` and `loo` [@vehtari2016] packages.
-Data and model results will be processed and visualised using the R packages the `tidyverse` family of packages [@wickham2020] and the `tidybayes` package [@mskay2019].
-
-# Results
-
-## Model selection
-
-We compared both models using two different criteria.
-First, we computed a Bayes Factor comparing the average posterior probability of the Null and the Extended models under the observed data.
-We obtained little evidence in favour of the Extended model over the Null model (*BF* = `r printnum(bayes_factor$bf`)).
-Second, we used an approximation to leave-one-out cross-valitation [Pareto-smoothed Importance Sampling, @vehtari2016] to compare how well each model was able to predict the data.
-Both models were very accurate when predicting the outcome: .
-Extended model's predictions were not substantially more accurate ($\hat{elpd}_{loo}^{Extended} - \hat{elpd}_{loo}^{Null}$ = `r printnum(loo_comp %>% as_tibble %$% elpd_diff[2]`, *SE* = `r printnum(loo_comp %>% as_tibble %$% se_diff[2])`).
-
-## Posterior distibutions
-
-```{r posteriors, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-include_graphics(here("Figures", "05_gca-coefficients.png"))
-
-```
-
-## Posterior correlations
-
-```{r correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-include_graphics(here("Figures", "05_gca-correlations.png"))
-
-```
-
-## Prior predictive checks
-
-Following @gabry2019, we performed prior predictive chackes to disgnose the appropriateness of our priors.
-
-## Posterior predictive checks
-
-## Model diagnosis
-
-```{r convergence, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-include_graphics(here("Figures", "05_gca-convergence.png"))
-
-```
-
-### Auto-correlation
-
-### Potential scale reduction factor ($\hat{R}$)
-
-## Posterior predictive checks
-
-Does our model generate similar observations to the ones in our data-set?
-
-# Discussion
-
-## Limitations
-
-* Cross-sectional data for longitudinal claims
-* Data is right censored    
-* Inventory:
-    - Phonological forms?
-    - We can't say how frequently words are produced by the parents or the toddler.
-    - Although instructed to ignore imitations, it is difficult to say whether a toddler has *really* acquired a word that produces, or she can only imitate it.
-    - Responses in the questionnaire rely heavily on parental memory.
-    - We don't know about the context at which words are heard or produced.
-    - Classification system of items (e.g., *household items*) is adult centric. Children may use the words *pretty* to name jewellery [@bates1994].
-* Mid and mid-upper class families are overrepresented in the sample.
-
-# Appendix
-
-## Appendix 1: Session info
-
-```{r session_info, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-sessionInfo()
-```
-
-## Appendix 2: Model
-
-Our model is incomplete, in that we have not included all parameters that may play a role in the real mechanisms that generate our observed data.
-Previous studies have found robust effects of sex (e.g., @frank2016), word length/number of phonemes (e.g., @frank2016), etc.
-Computational limitations have prevented us from specifying and fitting a more exhaustive model.
-Following the model-building strategy suggested by @schad2020, we have specified an intitial model that captures the fundamental structure of our phenomenon of interest and that can be computed with our available resources.
-Then, we have expanded such model incorporating a new predictor that adresses our hypotheses.
 
@@
 
-r_refs(file = "../References/BiLexicon.bib")
+r_refs(file = "../References/trajectories.bib")
 

### 2020-06-30 Manuscript/trajectories_manuscript_appendix.Rmd
First version of the Introduction and Methods, and second version of the ICIS 2020 poster

--- 2020-06-30 Manuscript/trajectories_manuscript.Rmd
@@
 ---
-title             : "The role of cross-linguistic lexical similarity on bilingual word acquisition"
-shorttitle        : "Cross-linguistic similarity and  word acquisition"
-
-author: 
-  - name          : "Gonzalo García-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Daniela Avila-Varela"
-    affiliation   : "1"
-  - name          : "Núria Sebastian-Galles"
-    affiliation   : "1"
-
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
-
-authornote: |
-  "Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
-
-abstract: |
- Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "lexical acquisition, vocabulary, bilingualism"
-wordcount         : "X"
-
-bibliography      : ["../References/trajectories.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-keep_text         : false
-class             : "man"
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  word_document:
-    reference_docx: trajectories_template.docx
-
+appendix: trajectories_manuscript_appendix.Rmd
+output: pdf_document
 ---
 
-```{r setup, include = FALSE}
-
-knitr::opts_chunk$set(
-	echo = FALSE,
-	fig.align = "center",
-	message = FALSE,
-	warning = FALSE,
-	cache.extra = knitr::rand_seed,
-	out.width = "80%",
-	results = "asis"
-)
-```
-
-```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# load packages
-library(papaja)
-library(knitr)
-
-library(magrittr)
-library(tidyverse)
-library(dplyr)
-library(forcats)
-library(stringr)
-library(readxl)
-library(lubridate)
-library(janitor)
-
-library(ggridges)
-library(patchwork)
-library(gt)
-
-library(data.table)
-library(here)
-
-# load/create functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
-
-# import data 
-dat_studies <- fread(here("Data", "00_studies.csv")) %>% distinct(q_version, language, q_items)
-dat_items <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
-  group_by() %>%
-  drop_na(language) %>%
-  mutate_at(vars(A, B, C, D, cognate_rater1, include, frequency), as.numeric) %>%
-  mutate_at(vars(A, B, C, D, cognate_rater1, include), as.logical)
-dat_participants <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
-  last() %>%
-  fread(na.strings = "") %>%
-  filter(age_bin %in% bins,
-         completed,
-         lp == "Bilingual")
-dat_raw <- fread(here("Data", "02_merged.csv"), na.strings = "") %>%
-  filter(lp %in% c("Bilingual"),
-         age_bin %in% bins) %>%
-  as_tibble() %>%
-  mutate_at(vars(date_birth, time_stamp), as_date)
+```{r include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 
@@
 dat_wordbank_priors <- fread(here("Data", "05_priors-wordbank.csv"))
-
 ```
-
-# Introduction
-
-Bilingual toddlers must learn two distinct set of words (one for each language) that partially overlap in sound and meaning.
-Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When taking both languages into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev1977; @doyle1977some; @rosenblum1983word; @fernandez1992; @pearson1994; @bialystok2010; @hoff2012; @core2013; @hoff2014; @blom2019; but see @pearson1993; @houwer2014].
-Not all bilinguals show a similar developmental trajectory of lexical acquisition.
-Recent studies have capitalised on the role of the similarity between the specific pair of languages the infant is learning.
-One way of measuring language similarity is by estimating the amount of cognates (i.e., form-similar translation equivalents) they share.
-@floccia2018 explored the impact of this form of similarity on vocabulary sizes from a large sample of 24-month-old toddlers learning English and an additional language.
-Toddlers learning languages that shared a large amount of cognates (e.g., English-German) showed larger vocabulary sizes than those learning languages that shared fewer cognates (e.g., English-Spanish).
-@blom2019 extended these findings to children aged three to 10 years.
-These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
-
-How does cognateness influence word learning?
-@floccia2018 point to the possibility that parallel activation during speech perception may play a role in this cognate facilitation effect.
-The parallel activation principle suggests that bilinguals activate both languages in parallel during speech perception and production, even when only one of them is being used.
-This hypothesis can only be understood under the more general account of language non-selectivity of lexical processing, which states that translation equivalents are co-activated at the lexical level [e.g., @de1991lexical; @costa2000; @dijkstra2005; @thierry2007; @hoshino2008; @yudes2010; @bobb2020; see @kroll2013 for review].
-A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model [@kroll1994; @kroll2010], the Bilingual Interactive Activation model [BIA; @dijkstra1998bia], and its successor BIA+ [@dijkstra2002architecture], the Bilingual Language Interaction Network for Comprehension of Speech model [BLINCS; @shook2013] and, more recently, Multilink [@dijkstra2019], which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
-However, these models do not address word learning during language acquistion.
-
-It is feasible that parallel activation plays a role in vocabulary acquisition during the first years of life.
-On top of Floccia et al.'s findings, other studies have provided experimental evidence of parallel activation during speech perception in bilingual toddlers [@vonholzen2012; @vonholzen2019].
-Yet, it is unclear how parallel activation influences word learning at these ages.
-What mechanisms are engaged, and how do they relate to vocabulary size?
-We contemplate two scenarios.
-First, it is possible that the reported "similarity boost" on vocabulary size is not directly related to the amount of cognates shared between both languages, but rather to the overlap across their correspondent phonological inventories.
-One of the measures of language similarity Floccia et al.
-computed is the mean phonological Levenshtein distance across translation equivalents.
-This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
-These scores were aggregated across pais of tanslation equivalents, obtaining a single similarity estimate of for each language pair.
-This measure embeds confounded information about the "degree of cognateness" and the amount of phonological overlap between both languages.
-Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, Enlgish-Dutch), with few exceptions (e.g., Spanish-Basque).
-In the light of these data, it is not possible to discern whether vocabulary size is increased by the ealier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
-We address this issue by testing two hypotesis.
-
-## Hypothesis 1: Cognates are acquired earlier than non-cognates
-
-First, we test whether the increase in vocabulary size reported by Floccia et al.
-for languages sharing a greater amount of cognates is indeed caused by cognates being acquired earlier.
-In line with this account, @bosch2014 found higher productive vocabulary scores in 18 month-old Catalan-Spanish bilinguals, compared to monolinguals, due to the higher presence of cognates their lexicon.
-We will adopt a similar approach to @braginsky2019, who modelled cross-sectional data from individual items to explore what properties were associated with an earlier acquisition.
-We will analyse the acquisition trajectories of a selection of `r dat_items %>% filter(include) %>% group_by(language) %>% summarise(n = n(), .groups = "drop") %>% pull(n) %>% max()` translation equivalents in a cross-section sample of Catalan/Spanish bilinguals aged 14 to 34 months of age.
-This approach allows to analyse how the proportion of infants that know each item changes across ages, providing a more suitable framework to put developmental hypotheses to test, instead of taking a "snapshot" of the state of the lexical of children at a single age.
-
-## Hypothesis 2: Cognates are acquired closer in times than non-cognates
-
-One mechanism that could acccount for an ealier acquisition of cognates is that parallel activation makes word-forms acquired in one language available during the acquisition of word forms in the other language.
-One consequence of this situation is that translation equivalents scaffold the acquisition of each other.
-This would point to different mechanisms engaged in the acquisition of cognates vs.
-non-cognates.
-While the phonology of cognate translation equivalents overlaps, and thus could be co-activated, the phonology of non-cognates does it in a much lesser degree.
-Consider the case of a Catalan-Spanish bilingual toddler that has already acquired the Catalan words *porta* (/ˈpɔr.tə/, door) and *taula* (/ˈtaw.ɫə/, table), but not yet their correspondent translation equivalents in Spanish (*puerta*, /ˈpwer.ta/; *mesa* /ˈme.sa/).
-According to our hypothesis, hearing the word *puerta* in presence of a door should lead to the activation of *porta*, which should facilitate the acquisition of the former.
-On the contrary, when hearing *mesa* in presence of a table, it is unlikely that the phonology of *taula* will be activated.
-This mechanism should lead to cognate translation equivalents being learnt closer in time than non-cognate translation equivalents.
-To test this hypothesis, we will extract the age of acquisition of each word-form in the questionnaire (as estimated by the model used to test Hypothesis 1), and will calculate the difference in time between the acquisition of such word-forms and their translation equivalents.
-We will test how cognateness affects such outcome.
-
-# Method
-
-## Participants
-
-We collected data from `r dat_raw$id_db %>% unique %>% length() %>% printnum()` bilinguals from the Metropolitan Area of Barcelona, between `r min(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`.
-All families participated voluntarily.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-
-We assessed toddlers' language background asking parents for an estimated proportion of exposure to each language.
-We defined participants as Spanish- or Catalan-dominant according to the language they were exposed the most to.
-Participants were considered bilingual - therefore included - if exposed 50-90% of the time to the non-dominant language.
-We excluded participants with >10% exposure to a third language.
-Fig.
-\@ref(fig:lpdist) shows how participants were distributed across age bins and language profiles.
-
-```{r lpdist, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_participants %>%
-  filter(completed, lp %!in% "Other", age_bin %in% bins) %>%
-  ggplot(aes(x = age_bin, fill = dominance)) +
-  geom_bar() +
-  labs(x = "Age (months)", y = "Number of partcipants",
-       fill = "Dominant language") +
-  scale_y_continuous(labels = as.integer) +
-  scale_fill_brewer(palette = "Set1") +
-  theme_custom +
-  theme(legend.position = c(0.05, 0.85),
-        legend.direction = "vertical",
-        legend.justification = 0,
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))
-```
-
-## Questionnaire
-
-Parents completed an online questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
-We implemented the questionnaire in `formr` [@arslan2020], a free, open-source platform.
-Participants were randomly allocated into one of four list of items.
-Each list contained a distinct inventory of translation equivalents.
-We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/funcional categories:  Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-Those categories that did not entail more than 16 items - thus resulting in less than four items after dividing it in four lists - were not divided: all of their items were included in the four lists.
-Another subset of items that were part of the trial lsits of some experiments in the lab were also included in all versions.
-
-Each item of the inventory presented a word-form, to which parents were instructed to report whether the toddlers was able to understand it, understand ans say it, or neither.
-The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates) .
-Items of one language were translation equivalents from the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-
-## Data analysis
-
-For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item  within age bins of 2 months.
-Our model assumes that this proportion follows a logistic curve accross ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the dominance of the language of the item (*item dominance*, L1 vs.
-L2), the cognate status of the word (*cognateness*, Non-cognate vs.
-Cognate), and the frequency of the word form in its correspondent language.
-Our prediction  is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words.
-
-Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
-Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item porta (door, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
-We retrieved frequencies from SUBTLEX-ESP [@cuetos2011] and SUBTLEX-CAT [@boada2019] for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word [@zipf1949; @heuven2014].
-Missing frequency scores were imputed using multiple imputation [@vanbuuren2011].
-Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
-
-```{r itemsinfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_items %>%
-  drop_na(cognate_rater1) %>%
-  filter(include) %>%
-  select(A, B, C, D, cognate_rater1, frequency, language) %>% 
-  mutate(cognate_rater1 = ifelse(cognate_rater1, "Cognate", "Non-cognate")) %>%
-  pivot_longer(c(A, B, C, D), names_to = c("list"), values_to = "n") %>% 
-  filter(n) %>% 
-  group_by(list, language, cognate_rater1) %>% 
-  summarise(n = sum(n, na.rm = TRUE),
-            mean = mean(frequency, na.rm = TRUE),
-            sd = sd(frequency, na.rm = TRUE),
-            min = min(frequency, na.rm = TRUE),
-            max = max(frequency, na.rm = TRUE)) %>%
-  ungroup() %>% 
-  rename(cognateness = cognate_rater1) %>%
-  relocate(list, language, cognateness, n, mean, sd, min, max) %>%
-  arrange(list, language, cognateness) %>%
-  mutate(list = paste("List", list)) %>% 
-  pivot_wider(names_from = cognateness, values_from = c(n, mean, sd, min, max)) %>%
-  clean_names() %>%
-  gt(groupname_col = c("language")) %>% 
-  fmt_number(columns = contains("cognate"), decimals = 2) %>%
-  fmt_number(columns = starts_with("n_"), decimals = 0) %>%
-  tab_header(title = md("**Questionnaire and item properties**"),
-             subtitle = "Participants were randomly allocated on of the the four lists, and completed both the Catalan and the Spanish versions.") %>% 
-  tab_spanner(label = "Frequency (Cognates)", columns = vars("n_cognate", "mean_cognate", "sd_cognate", "min_cognate", "max_cognate")) %>% 
-    tab_spanner(label = "Frequency (Non-cognates)", columns = contains("non_cognate")) %>% 
-  cols_label(list = md("**List**"), 
-             n_cognate = md("*N*"),
-             mean_cognate = md("*Mean*"),
-             sd_cognate = md("*SD*"),
-             min_cognate = md("*Min*"),
-             max_cognate = md("*Max*"),
-             n_non_cognate = md("*N*"),
-             mean_non_cognate = md("*Mean*"),
-             sd_non_cognate = md("*SD*"),
-             min_non_cognate = md("*Min*"),
-             max_non_cognate = md("*Max*")) %>%
-  gtsave(here("Results", "descriptives_items.png"))
-
-```
-
-We fit two separate logistic models on comprehensive and productive data, respectively.
-Logistic curves are characterised by three parameters[see @mahr2019 and @mahr2020 for an excellent introduction to this approach, in which we based this analysis]: (a) an *asymptote* (`asym`, upper boundary  of the proportion of infants that have acquired the item), (b) the *steepness* (`steep`, how fast the a given item is acquired by infants), and (c) a *mid-point* (`mid`, the age at which steepness is maximum).
-We work on the assumption that the trajectory of aquisition of items follows a sigmoidal shape [@mayor2011], and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
-This approach allows us to model and estimate the age of acquisition of items explicitely.
-Our model estimated the asymptote and stepness as grand means across items (i.e., intercepts).
-Mid-points were estimated by a linear zero-one-beta-regression model^[Zero-or-one-inflated beta regression models allow to account for overdispersion then using proportions as the output of a model accounting, while also allowing values to be at boundary (0 and 1).][@ospina2012] that included our predictors of interest: `age`, `dominance` (contrast-coded, L2 = -0.5, L2 = 0.5), `cognateness` (contrast-coded, Non-cognate = -0.5, Cognate = 0.5), their interaction, and `frequency` (standardised Zipf score).
-We also included a random intercept for each translation equivalent (`TE`).
-Appendix 1 shows a detailed description of the model.
-
-We estimated the parameters of our model using the Bayesian framework.
-Bayesian models use the Bayes rule to combine prior knowlegde about the distribution of a given parameter, and the likelihood of each of the values it can take given tha observed data.
-We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our priors).
-In order to test the contribution of the predictor *cognateness* on the fit of the model,  we compared the fit of our (expanded) model to the fit of a (null) model that did not include *cognateness* as predictor [@schad2020] using leave-one-out cross validation.
-We fit both the null and alternative models using the R environment [@rcoreteam2013; @rstudioteam2015] and the `brms` package [@burkner2017], which relies on the probabilistic language Stan [@carpenter2017] to approximate posterior distributions.
-We used Pareto-smoothed importance sampling [PSIS; @vehtari2017; @vehtari2019] to compare the null and extended models.
-Data and model results were processed and visualised using the `tidyverse` family of R packages [@wickham2019] and the `tidybayes` package [@kay2020].
-
-To test Hypothesis 2, we run a Bayesian ANOVA using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included *cognateness* as predictor.
-Using the `BayesFactor` R package [@morey2018a], we computed a Bayes factor (*BF*) that compared the likelihood of a linear model that included *cognateness* as a predictor against the likelihood of a linear model that did not, under the light of our data.
-
-# Results
-
-## Hypothesis 1
-
-### Comprehensive vocabulary
-
-### Productive vocabulary
-
-## Hypothesis 2
-
-### Comprehensive vocabulary
-
-### Productive vocabulary
-
-# Discussion
-
-# Appendix
-
-## Appendix 1
-
-Model formula and priors:
-
-$$proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ \beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\\phi \sim Normal(1.5, 1)$$
-
-## Appendix 2: Computing priors
 
@@
 All the other available languages were excluded, as no (or insufficient) data were available.], from monolingual infants aged 16 to 30 months of age.
-Table \@ref(tab:wordbank) shows the estimates of this model, which were rescaled and included as priors in our main model to test Hypothesis 1.
 
-```{r worbank, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_wordbank_priors %>%
-  rename(estimate_scaled = estimate) %>%
-  mutate(conf = paste0("[", round(conf_low, 2), ", ", round(conf_high, 2), "]"),
-         estimate = estimate_scaled*2+20,
-         term = c("Asymptote", "Mid-point", "Steepness")) %>% 
-  select(term, estimate, estimate_scaled, std_error, conf) %>%
-  gt(rowname_col = "term") %>% 
-  fmt_number(vars(estimate, estimate_scaled, std_error), decimals = 2) %>% 
-  cols_label(estimate = md("*Estimate<br>(months)*"),
-             estimate_scaled = md("*Estimate<br>(scaled)*"),
-             std_error = md("*SE*"),
-             conf = md("*95% CI*")) %>% 
-  cols_align(align = "c") %>% 
-  gtsave(here("Results", "wordbank.png"))
-
-```
-
-We performed prior-predictive checks to analyse the feasibility of our priors. Figure \@ref(fig:priorchecks)
-
-```{r priorchecks, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-include_graphics(here("Figures", "06_simulate-priors.png"))
-
-```
-
-```{r priorstab, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tribble(~param, ~coefficient, ~notation, ~prior,
-        "Asymptote", "Intercept", "$A$", paste0("$N(", dat_wordbank_priors[term=='asym'] %>% pull(estimate) %>% printnum(), ", 0.05)$"),
-        "Steepness", "Intercept", "$S$", paste0("$N(", dat_wordbank_priors[term=='steep'] %>% pull(estimate) %>% printnum(), ", 0.6)$"),
-        "Mid-point", "Intercept", "$\\beta_{0}$", paste0("$N(", dat_wordbank_priors[term=='mid'] %>% pull(estimate) %>% printnum(), ", 1)$"),
-        "Mid-point", "Random Intercepts", "$I_{i0}$", "N(0, $\\pi$)",
-        "Mid-point", "Slope", "$\\beta_{1}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{2}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{3}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{4}$", "$N(0, 5)$",
-        "Mid-point", "Intercept", "$\\varphi$", "$N(2, 1)$",
-        "Mid-point", "Intercept", "$COI$", "Beta(1, 1)", 
-        "Mid-point", "Intercept", "$ZOI$", "Beta(1, 1)",
-        "Mid-point", "Std. Deviation", "$\\sigma{ijk}$", "t(3, 0, 2.5)") %>%
-  kable(digits = 2, row.names = FALSE, col.names = c("**Adjusted parameter**", "**Coefficient**", "**Notation**", "**Distribution**"), align = "c", caption = "Prior distribution of the parameters in our models.
-We set specific informative priors for the asymptote, steepness, and mid-point of the curves (using data from Wordbank) and and weakly informative priors for the rest of the coefficients.", escape = TRUE, format = "latex")
-
-```
-
-## Appendix 3: Session info
-
-```{r session_info, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-sessionInfo()
-
-```
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-
-\endgroup
-
-```{r create_r-references, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-r_refs(file = "../References/trajectories.bib")
-
-```
+**Figure caption**: "Logarithmic model fitted on Wordbank data, compared to a linear model.
+The logarithmic model fits the data slightly better.
+The estimated values of the three parameters involved in the logarithmic model (asymptote, mid-point and steepness) are indicated."

### 2020-06-30 Code/R/03_familiarity.R

@@ -21,7 +21,7 @@ source(here("R", "functions.R"))
 
 # set params
 bins          <- c("< 10", "10-12", "12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "36-38", "38-40", "> 40")
-bins_interest <- c("18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32")
+bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
 breaks <- c(0, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 100)
 
 #### import data ############################################
@@ -36,6 +36,9 @@ dat <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactors = FA
   mutate(response = case_when(response == 1 ~ "no", response == 2 ~ "understands", response == 3 ~ "produces", TRUE ~ NA_character_),
          understands = response %in% c("understands", "produces"),
          produces = response %in% "produces",
+         doe = ifelse(dominance=="Spanish", doe_spanish, doe_catalan),
+         lp = ifelse(!between(doe, 50, 100, incbounds = TRUE), "Other", lp),
+         lp = factor(lp, levels = c("Monolingual", "Bilingual")),
          age_bin = factor(cut(age, breaks = breaks, labels = bins), levels = bins, ordered = TRUE)) %>%
   filter(lp %in% c("Monolingual", "Bilingual"),
          age_bin %in% bins_interest) %>%
@@ -43,19 +46,20 @@ dat <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactors = FA
   pivot_longer(c(understands, produces), names_to = "type", values_to = "response") %>%
   mutate(type = ifelse(type=="understands", "Comprehensive", "Productive")) %>%
   arrange(item, lp, item_dominance, age_bin) %>%
-  group_by(item, lp, age_bin, type) %>%
+  # aggregate data
+  group_by(item, lp, age_bin, item_dominance, type) %>%
   summarise(n = sum(!is.na(response)),
             successes = sum(response, na.rm = TRUE),
             proportion = mean(response, na.rm = TRUE),
             .groups = "drop") %>%
   ungroup() %>%
[…]

### 2020-06-30 Code/R/05_priors.R

@@ -1,0 +1,100 @@
+#### 05_wordbank: compute priors from Worbank data
+# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
+# Center for Brain and Cognition, Universitat Pompeu Fabra
+
+#### set up ##############################################
+
+# load packages
+library(data.table)
+library(tidyverse)
+library(janitor)
+library(ggridges)
+library(broom)
+library(modelr)
+library(patchwork)
+library(here)
+
+# load functions
+source(here("Code", "R", "functions.R"))
+
+# set params
+set.seed(888)
+bins <- c("< 16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "> 40")
+breaks <- c(0, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 100)
+bins_interest <- c("16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
+
+
+#### load data ############################################
+dat <- list.files(here("Data", "Wordbank"), full.names = TRUE) %>%
+  map(., fread) %>%
+  set_names(str_remove(list.files(here("Data", "Wordbank")), ".csv")) %>%
+  bind_rows(.id = "language") %>%
+  as_tibble() %>%
+  pivot_longer(6:ncol(.), names_to = "age", values_to = "proportion") %>%
+  mutate(age = as.numeric(age),
+         age_bin = factor(cut(age, breaks = breaks, labels = bins), levels = bins_interest),
+         age_bin = as.numeric(age_bin)-1) %>%
+  drop_na(age_bin)
+
+#### non-linear least squares ##############################
+fit <- nls(proportion ~ SSlogis(age_bin, asym, mid, steep), dat)
+s <- summary(fit)
+dat_fit <- expand_grid(age_bin = unique(dat$age_bin)) %>% mutate(fit = predict(fit, newdata = .))
+coefs <- tidy(fit, conf.int = 0.95) %>%
+  clean_names() %>%
+  mutate(estimate_scaled = estimate*2+16) %>%
+  relocate(term, estimate, estimate_scaled)
+
+#### visualise data #######################################
+
+# visualise model fit
+x_mid <- coefs$estimate[coefs$term=="mid"]
+y_mid <- dat_fit$fit[which.min(abs(dat_fit$age_bin-x_mid))]
+steepness <- coefs$estimate[coefs$term=="steep"]
[…]

### 2020-06-30 Code/R/06_priors.R

@@ -1,101 +1,0 @@
-#### 05_wordbank: compute priors from Worbank data
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up ##############################################
-
-# load packages
-library(data.table)
-library(dplyr)
-library(tibble)
-library(tidyr)
-library(purrr)
-library(stringr)
-library(janitor)
-library(ggplot2)
-library(ggridges)
-library(broom)
-library(modelr)
-library(wesanderson)
-library(patchwork)
-library(here)
-
-# load functions
-source(here("Code", "R", "functions.R"))
-inv_logit <- function(x) 1 / (1 + exp(-x))
-
-# set params
-set.seed(888)
-
-#### load data ############################################
-
-dat <- list.files(here("Data", "Wordbank"), full.names = TRUE) %>%
-  map(., fread) %>%
-  set_names(str_remove(list.files(here("Data", "Wordbank")), ".csv")) %>%
-  bind_rows(.id = "language") %>%
-  as_tibble() %>%
-  pivot_longer(6:ncol(.), names_to = "age", values_to = "proportion") %>%
-  mutate(age = as.numeric(age)) %>%
-  filter(between(age, 16, 30))
-
-#### linear least squares
-fit_l     <- lm(proportion ~ age, dat)
-s_l       <- summary(fit_l)
-dat_fit_l <- expand_grid(age = unique(dat$age)) %>% mutate(fit = predict(fit_l, newdata = .))
-coefs_l   <- tidy(fit_l, conf.int = 0.95) %>% clean_names()
-
-#### non-linear least squares ##############################
-fit_nl <- nls(proportion ~ SSlogis(age, asym, mid, steep), dat)
-s_nl <- summary(fit_nl)
-dat_fit_nl <- expand_grid(age = unique(dat$age)) %>% mutate(fit = predict(fit_nl, newdata = .))
-coefs_nl <- tidy(fit_nl, conf.int = 0.95) %>% clean_names()
-
-#### compare models
-dat_fit <- bind_rows(dat_fit_l, dat_fit_nl, .id = "type") %>%
-  mutate(type = ifelse(type==1, "Linear", "Non-linear"))
-residuals <- dat %>% 
-  drop_na() %>%
-  mutate(residual_l = (s_nl$residuals-mean(s_l$residuals))/sd(s_l$residuals),
[…]

### 2020-06-30 Code/R/07_gca-comp.R

@@ -1,0 +1,249 @@
+#### 05_gca-comp: Bayesian non-linear model ##############
+# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
+# Center for Brain and Cognition, Universitat Pompeu Fabra
+
+#### set up ##############################################
+
+# load packages
+library(tidyverse)
+library(data.table)
+library(truncnorm)
+library(brms)
+library(modelr)
+library(janitor)
+library(tidybayes)
+library(ggmcmc)
+library(mcmcplots) 
+library(here)
+
+# create/load functions
+source(here("Code", "R", "functions.R"))
+
+# set params
+set.seed(888)
+bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+options(contrasts = c("contr.sum", "contr.poly"))
+
+#### import data ##########################################
+dat <- fread(here("Data", "04_prepared.csv")) %>%
+    as_tibble() %>% 
+    filter(type=="Comprehensive",
+           lp=="Bilingual") %>%
+    select(item, meaning, age_bin, item_dominance, cognate, proportion, frequency) %>%
+    mutate(age_bin = as.numeric(factor(age_bin, levels = bins, ordered = TRUE)),
+           item_dominance = as.factor(item_dominance),
+           frequency = scale(frequency)[,1],
+           cognate = as.factor(cognate)) %>%
+    arrange(item, meaning, age_bin)
+
+contrasts(dat$item_dominance) <- contr.sum(c("L2", "L1"))/2
+contrasts(dat$cognate) <- contr.sum(c("Non-cognate", "Cognate"))/2
+
+dat_priors <- fread(here("Data", "05_priors-wordbank.csv")) %>%
+    mutate(estimate_scaled = (estimate_scaled-14)/2)
+
+#### fit models ###########################################
+
+# load in case models are already fitted
+fit_prior <- readRDS(here("Results", "comp_fit-prior.rds"))
+fit0 <- readRDS(here("Results", "comp_fit0.rds"))
+fit1 <- readRDS(here("Results", "comp_fit1.rds"))
+
+# model 0
+fit0 <- brm(formula = bf(proportion ~ inv_logit(asym) * inv(1 + exp((mid - age_bin) * exp(steep))),
+                         asym ~ 1,
+                         mid ~ 1 + item_dominance + frequency + (1 | meaning),
+                         steep ~ 1, 
+                         phi ~ 1,
+                         nl = TRUE,
[…]

### 2020-06-30 Code/R/functions.R

@@ -14,7 +14,6 @@ library(ggplot2)
 
 # logistic curve
 inv_logit <- function(x) 1 / (1 + exp(-x))
-log_curve <- function(x, u, k, mid){ {{ u }}/(1+exp(-{{ k }}*({{ x }} - {{ mid }} )))}
 
 # logit to probability
 logit2prob <- function(logit){

### 2020-06-30 Manuscript/trajectories_manuscript.tex
First version of the Introduction and Methods, and second version of the ICIS 2020 poster

@@
 \abstract{
-Previous literature on early vocabulary has been mostly committed to exploring how the receptive and productive vocabulary size changes with age, and its relationship with toddlers' performance on language tasks (e.g.~Fernald, Swingley, \& Pinto, 2001; Fernald, Perfors, \& Marchman, 2006).
-The content of the developing lexicon has remained relatively unexplored.
-More recently, studies have focused on the characterisation of the developmental trajectory of individual words, reporting an earlier age of acquisition for words with high frequency, concreteness, and phonological neighbourhood density (e.g.~Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019).
-Such item-level analysis allows not only to predict the age of acquisition of specific words, but also to shed light on the cognitive processes that underlie early word learning.
-This approach is particularly interesting for investigating the potential impact of acquiring two languages simultaneously.
-A recent study by Floccia et al.
-(2018) compared vocabulary sizes of 24-month-old children learning British English, together with an additional language from a pool of 13 diverse languages.
-They found larger productive vocabulary sizes of toddlers learning two languages that are phonologically similar (see for a similar approach Bosch \& Ramon-Casas, 2014).
+Bilinguals face the task of learning words from languages with overlapping phonologies.
+Floccia et al.
+(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages sharing large amount of cognates (e.g., English-Dutch).
 The mechanisms underlying this effect remain unknown.
-One possibility is that the similarity between the two languages speeds the acquisition of form-similar translation equivalents.
-The aim of this study is to perform an item-wise analysis on infant's vocabulary contents.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish.
-We expect that phonological overlap between translation equivalents will predict earlier age of acquisition.
-If this effect is driven by the phonological overlap between translation equivalents, cognate pairs should be acquired closer in time than non-cognates.
-For instance, the translation of cat /gato/ in Spanish and /gat/ in Catalan should be learnt at approximately the same age.
-This should not necessarily happen for the translations of dog, /pero/ and /gos/.
-We analyse the time elapsed between the acquisition of a word in one language, and its translation in the other.
-We present preliminary data (data collection is ongoing) on receptive vocabulary of 90 monolingual and bilingual toddlers aged 18 to 36 months.
-We obtained parental responses to 230 pairs of Catalan-Spanish translation equivalents, resulting in a total of 41490 responses.
-Words that yielded a phonologically similar translation equivalent were more likely to be present in toddlers' receptive vocabulary at all ages, and were acquired earlier, even when accounting for the effect of word frequency.
-Further analysis will explore the distance in time between the acquisition of pairs of translation equivalents using a more detailed measure of phonological overlap across translation equivalents and taking into account cross-individual and cross-item variability.
+We explore two compatible scenarios.
+First, cognates may be learnt earlier than non-cognates, boosting lexical acquisition at early ages.
+This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
+Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
+This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
+We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
 }
@@
 
-One of the main challenges bilingual infants and toddlers face is learning two distinct set of words (one for each language) that partially overlap in sound and meaning.
+Bilingual toddlers must learn two distinct set of words (one for each language) that partially overlap in sound and meaning.
 Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When taking both languages into account, bilingual children between seem to know, at least, as many words as monolinguals do ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
-Tough, not all bilinguals show asimilar developmental trajectory of lexical acquisition.
+When taking both languages into account, bilingual children seem to know, at least, as many words as monolinguals do (Ben-Zeev, 1977; Bialystok, Luk, Peets, \& Yang, 2010; Blom et al., 2019; Core, Hoff, Rumiche, \& Señor, 2013; Doyle \& others, 1977; Fernandez, Pearson, Umbel, Oiler, \& Molinet-Molina, 1992; Hoff et al., 2012; Hoff, Rumiche, Burridge, Ribot, \& Welsh, 2014; Pearson \& Fernández, 1994; Rosenblum \& Pinker, 1983; but see Pearson, Fernández, \& Oller, 1993; Houwer, Bornstein, \& Putnick, 2014).
+Not all bilinguals show a similar developmental trajectory of lexical acquisition.
+Recent studies have capitalised on the role of the similarity between the specific pair of languages the infant is learning.
+One way of measuring language similarity is by estimating the amount of cognates (i.e., form-similar translation equivalents) they share.
+Floccia et al.
+(2018) explored the impact of this form of similarity on vocabulary sizes from a large sample of 24-month-old toddlers learning English and an additional language.
+Toddlers learning languages that shared a large amount of cognates (e.g., English-German) showed larger vocabulary sizes than those learning languages that shared fewer cognates (e.g., English-Spanish).
+Blom et al.
+(2019) extended these findings to children aged three to 10 years.
+These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
 
-Recent studies have capitalised on the role of the similarity between the specific pair of language the infant is learning.
+How does cognateness influence word learning?
 Floccia et al.
-(2018) analysed vocabulary scores across of 24-month-old toddlers learning English and an additional language.
-Toddlers learning languages sharing a large amounth of cognates (e.g., English-German) showed larger vocabulary sizes than those learning less languages sharing less cognates (e.g., English-Chinese).
-({\textbf{???}}) extended these results to children aged three to 10 years.
-These results point to the possibility that the overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
-However, none of these studies provide an account for the mechanisms involved in this facilitatory effect of language driven by similarity across languages.
-The aim of this study is to explore two scenarios in which this effect may be taking place:
+(2018) point to the possibility that parallel activation during speech perception may play a role in this cognate facilitation effect.
+The parallel activation principle suggests that bilinguals activate both languages in parallel during speech perception and production, even when only one of them is being used.
+This hypothesis can only be understood under the more general account of language non-selectivity of lexical processing, which states that translation equivalents are co-activated at the lexical level (e.g., De Groot \& Nas, 1991; Bobb, Von Holzen, Mayor, Mani, \& Carreiras, 2020; Costa, Caramazza, \& Sebastian-Galles, 2000; Dijkstra, 2005; Hoshino \& Kroll, 2008; Thierry \& Wu, 2007; Yudes, Macizo, \& Bajo, 2010; see Kroll, Gullifer, \& Rossi, 2013 for review).
+A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model (Kroll, Hell, Tokowicz, \& Green, 2010; Kroll \& Stewart, 1994), the Bilingual Interactive Activation model (BIA; Dijkstra \& Van Heuven, 1998), and its successor BIA+ (Dijkstra \& Van Heuven, 2002), the Bilingual Language Interaction Network for Comprehension of Speech model (BLINCS; Shook \& Marian, 2013) and, more recently, Multilink (Dijkstra et al., 2019), which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
+However, these models do not address word learning during language acquistion.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  In a first scenario, it is the phonological overlap between language pairs, and not the amount of cognates they share, what boosts the vocabulary growthof bilingual toddlers.
-Languages sharing a lot of cognates tend to also share a lot of phonemic categories.
-It is possible that having to learn fewer phonemic categories makes it easier for toddlers to acquire words.
-If this is true, toddlers learning two languages with similar phonemic inventories should show larger vocabulary sizes than those learning languages that barely ovelap phonologically.
-Moreover, this facilitation effect should be reflected in the developmental trajectory of all the items included in a vocabulary inventory, independently of their cognate status.
-\item
-  In a second scenario, Words that are form similar It is the amount of cognates language pairs share that boosts vocabulary growth in bilingual toddlers.
-If this is true, toddlers may acquire cognates earlier
-\end{enumerate}
+It is feasible that parallel activation plays a role in vocabulary acquisition during the first years of life.
+On top of Floccia et al.'s findings, other studies have provided experimental evidence of parallel activation during speech perception in bilingual toddlers (Von Holzen, Fennell, \& Mani, 2019; Von Holzen \& Mani, 2012).
+Yet, it is unclear how parallel activation influences word learning at these ages.
+What mechanisms are engaged, and how do they relate to vocabulary size?
+We contemplate two scenarios.
+First, it is possible that the reported \enquote{similarity boost} on vocabulary size is not directly related to the amount of cognates shared between both languages, but rather to the overlap across their correspondent phonological inventories.
+One of the measures of language similarity Floccia et al.~computed is the mean phonological Levenshtein distance across translation equivalents.
+This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
+These scores were aggregated across pais of tanslation equivalents, obtaining a single similarity estimate of for each language pair.
+This measure embeds confounded information about the \enquote{degree of cognateness} and the amount of phonological overlap between both languages.
+Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, Enlgish-Dutch), with few exceptions (e.g., Spanish-Basque).
+In the light of these data, it is not possible to discern whether vocabulary size is increased by the ealier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
+We address this issue by testing two hypotesis.
 
-To our knowledge, there are no previous studies that have investigated the effect of cognateness on the acquisition trajectories of individual word forms.
-Braginsky et al.
-(2019) used data from MacArthur-Bates CDIs from multiple languages to explore what properties are associated with an earlier acquisition.
-The found that frequency, concreteness, and {[}\ldots{}{]} significantly predicted an earlier age of acquisition.
-We plan to extend these analyses to the case of bilinguals.
+\hypertarget{hypothesis-1-cognates-are-acquired-earlier-than-non-cognates}{%
+\subsection{Hypothesis 1: Cognates are acquired earlier than non-cognates}\label{hypothesis-1-cognates-are-acquired-earlier-than-non-cognates}}
 
-We have developed a fully-online vocabulary inventory in Catalan and Spanish that is more exhaustive than the MacArthur-Bates CDI, including a larger number of items.
-All participants filled the questionnaire it both languages, thus providing responses to pairs of translation equivalents.
+First, we test whether the increase in vocabulary size reported by Floccia et al.~for languages sharing a greater amount of cognates is indeed caused by cognates being acquired earlier.
+In line with this account, Bosch and Ramon-Casas (2014) found higher productive vocabulary scores in 18 month-old Catalan-Spanish bilinguals, compared to monolinguals, due to the higher presence of cognates their lexicon.
+We will adopt a similar approach to Braginsky, Yurovsky, Marchman, and Frank (2019), who modelled cross-sectional data from individual items to explore what properties were associated with an earlier acquisition.
+We will analyse the acquisition trajectories of a selection of 717 translation equivalents in a cross-section sample of Catalan/Spanish bilinguals aged 14 to 34 months of age.
+This approach allows to analyse how the proportion of infants that know each item changes across ages, providing a more suitable framework to put developmental hypotheses to test, instead of taking a \enquote{snapshot} of the state of the lexical of children at a single age.
 
-In this study, we examine the role of cognateness on age of acquisition by modelling the probability of parents reporting that the participant is able to understand, say and understand on each item in the questionnaire.
-We will test the effect the participants' language profile (monolingual vs.~bilingual), and its cognate status on the developmetnal curves of each item.
+\hypertarget{hypothesis-2-cognates-are-acquired-closer-in-times-than-non-cognates}{%
+\subsection{Hypothesis 2: Cognates are acquired closer in times than non-cognates}\label{hypothesis-2-cognates-are-acquired-closer-in-times-than-non-cognates}}
 
-We will analyse the developmental trajectories of individual words in monolinguals and bilinguals learning Catalan and Spanish (two phonologically close languages sharing a great deal of cognates), between 14 and 30 months of age.
-We will test whether cognate words are learnt earlier than non-cognate words, taking into account wether they are part of the most- o less-dominant language of the infant, as well as their frequency, concreteness, and number of syllables.
-We have deloped an on-line questionnaire that entails a short language exposure questionnaire, asks for some demographic information, and presents parent with a list of \textasciitilde{}800 words in each language, for which parent report whether their child understands, says, or doesn't understand and say, in each item.
-
-We will fit a Bayesian multilevel model on the item responses, modelling the probability of parents reporting that a given word is acquired by the infant, including language profile (monolingual vs.~bilingual), item dominance (dominant language vs.~non-dominant language), frequency, concreteness, and number of syllables.
-We will include \emph{meaning} as grouping variable.
-We will fit a sigmoidal curve (see {\textbf{???}}) on each item, defined by the steepness of the developmental curve, and the mid-point (the poinnt at which the developmental curve is steepest, which will be considered the point of acquisition).
-Priors will be derived from previous literature on the subject.
-
-NOTES:
-
-Evidence that bilingual lexica barely overlap ({\textbf{???}}; {\textbf{???}}).
-
-Studies with infants from 8 to 30 months of age show weaker evidence for this claim ({\textbf{???}}; {\textbf{???}}), pointing to the posibility that differences between bilingual and monolingual vocabulary sizes may be dependent on maturational factors.
-
-sugested that the developmental trajectory of lexical aquisition varies across two dimensions: the language profile of the infant, and the joint proporties of the specific language pair the infant is learning.
-Regarding the first, ({\textbf{???}}) found thant infants exposed to a more balanced
-
-Of special interest is the fact than cognates (form-similar translation equivalents) tend to be overrepresented in the early bilingual vocabulary (Bosch \& Ramon-Casas, 2014).
-
-The similarity between both languages may play a role in bilingual lexical aquisition.
-
-\hypertarget{hypotheses}{%
-\subsection{Hypotheses}\label{hypotheses}}
-
-\begin{description}
-\tightlist
-\item[Hypothesis 1]
-The boost effect of language similarity is driven by cognates being acquired earlier than non-cognates.
-We predict that age of acquisition of cognates will be ealier thant that of non-cognates.
-This effect should only be present in bilinguals.
-\item[Hypothesis 2]
-That this effect is driven by the word-forms in one language scaffolding the acquisition of their translation equivalents via parallel activation.
-We predict that the difference in time between the acquisition pairs of translation equivalents is shorter in cognates than in non-cognates.
-This effect will only be present in bilinguals.
-\end{description}
+One mechanism that could acccount for an ealier acquisition of cognates is that parallel activation makes word-forms acquired in one language available during the acquisition of word forms in the other language.
+One consequence of this situation is that translation equivalents scaffold the acquisition of each other.
+This would point to different mechanisms engaged in the acquisition of cognates vs.~non-cognates.
+While the phonology of cognate translation equivalents overlaps, and thus could be co-activated, the phonology of non-cognates does it in a much lesser degree.
+Consider the case of a Catalan-Spanish bilingual toddler that has already acquired the Catalan words \emph{porta} (/ˈpɔr.tə/, door) and \emph{taula} (/ˈtaw.ɫə/, table), but not yet their correspondent translation equivalents in Spanish (\emph{puerta}, /ˈpwer.ta/; \emph{mesa} /ˈme.sa/).
+According to our hypothesis, hearing the word \emph{puerta} in presence of a door should lead to the activation of \emph{porta}, which should facilitate the acquisition of the former.
+On the contrary, when hearing \emph{mesa} in presence of a table, it is unlikely that the phonology of \emph{taula} will be activated.
+This mechanism should lead to cognate translation equivalents being learnt closer in time than non-cognate translation equivalents.
+To test this hypothesis, we will extract the age of acquisition of each word-form in the questionnaire (as estimated by the model used to test Hypothesis 1), and will calculate the difference in time between the acquisition of such word-forms and their translation equivalents.
+We will test how cognateness affects such outcome.
 
@@
 
-Data from 235 participants from the different cities in Spain were invited to participate through social media.
-Families in the Metropolitan area of Barcelona were also recruited at birth.
+We collected data from 407 bilinguals from the Metropolitan Area of Barcelona, between 21th November, 2017 and 22th June, 2020.
 All families participated voluntarily.
-Data were collected between 28th October, 2019 and 31th May, 2020, using the BiLexicon inventory.
 This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
 
-\begin{table}
-
-\caption{\label{tab:participants}Sample demographics and overall language profile.
-Dominance = Language to which the participant is exposed the most, N = sample size, M age = Mean age, SD age = standard deviation of age, M DOE Cat = mean degree of exposure to Catalan, SD DOE Cat = standard deviation of exposure to Catalan, M DOE Spa = mean degree of exposure to Spanish, SD DOE Spa = standard deviation of exposure to Spanish.}
-\centering
-\begin{tabular}[t]{c|c|c|c|c|c|c|c|c}
-\hline
-Dominance & Sex & N & M\_Age & SD\_Age & M DOE Cat & SD\_DOE\_Cat & M\_DOE\_Spa & SD\_DOE\_Spa\\
-\hline
-Catalan & Female & 126,526 & 20.96 & 4.59 & 73.94 & 18.48 & 23.52 & 18.29\\
-\hline
-Catalan & Male & 132,856 & 20.86 & 4.78 & 70.83 & 17.74 & 27.77 & 17.08\\
-\hline
-Spanish & Female & 66,476 & 21.22 & 4.76 & 17.38 & 14.74 & 81.14 & 15.60\\
-\hline
-Spanish & Male & 47,511 & 20.83 & 4.88 & 26.67 & 13.12 & 72.66 & 11.67\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{itemize}
-\tightlist
-\item
-  \emph{Location}
-\item
-  \emph{Age}
-\item
-  \emph{Sex}
-\item
-  \emph{Language profile}: : Linguistic profiles were assessed using the Language Exposure Questionnaire (LEQ) ({\textbf{???}}).
-This questionnaire estimates de degree of exposure (DOE) to each language via a 10 minutes long interview with the parents.
-LEQs were administered inmediately before or after the testing session at the lab.
-We here considered only exposure to Catalan and Spanish.
-Participants with \textgreater{}10\% exposure to a third language other than Catalan or Spanish were excluded.
-\item
-  \emph{SES}
-\end{itemize}
-
-\hypertarget{exclusion-criteria}{%
-\subsubsection{Exclusion criteria}\label{exclusion-criteria}}
-
-\begin{itemize}
-\tightlist
-\item
-  Language disorders
-\end{itemize}
-
-\hypertarget{questionnaires}{%
-\subsubsection{Questionnaires}\label{questionnaires}}
-
-\hypertarget{bilexicon-1.0}{%
-\paragraph{BiLexicon 1.0}\label{bilexicon-1.0}}
-
-\hypertarget{bilexicon-2.0}{%
-\paragraph{BiLexicon 2.0}\label{bilexicon-2.0}}
-
-\hypertarget{bilexicon-short}{%
-\paragraph{BiLexicon Short}\label{bilexicon-short}}
-
-This questionnaire was developed as a short version of the BiLexicon 1.0 and BiLexicon 2.0 inventories.
-Given the large amount of items to be answered to, we restricted the use of such inventories to participants that were to participate presentially in one of the studies run in the lab at that time.
-These participants only filled one of the questionnaires, and did so only once.
-
-The aim of developing a short version of this questionnaire (BiLexicon Short) was to provde a more on-line friendly format to families that would participate remotely exclusively, or that would fill the questionnaire more than once (e.g., in longitudinal studies).
-We divided the items of each language in the pool into four versions named A, B, C, and D.
-Items in each category (e.g., animals, auxiliary words) were randomly assigned one of the versions.
-Thos categories that did not entail more than 16 items (thus resulting in less than four items after dividing it in ofur lists), were not divided (all of their items were included in all versions).
-
-When filling the questionnaire for the first time, participants were randomly assigned one version.
-Each familiy filled only one of the versions, thus reducing the number of items to be answered.
-To ensure that longitudinal participants would provide longitudinal data in the questionnaire, we made sure that participants were always assigned the same version of this questionnaire.
-Words that were used as part of the trial lists in the experiment families would participate in were present across all versions.
-
-\hypertarget{item-inclusion-criteria}{%
-\subsubsection{Item inclusion criteria}\label{item-inclusion-criteria}}
-
-We gathered data from 695 distinct translation equivalents across Spanish (695 word forms) and Catalan (703 word forms).
-The unequal number of word-orms across both languages is due to that some tanslation equivalents did not entail a one-to-one-corespondence: some items had more than one translation equivalent in the other language (e.g.~the Catalan word form \emph{walking}, \emph{anar}, can be translated as both \emph{caminar} {[}to walk{]} or \emph{ir} {[}to go{]}).
-
-We included items from the following categories: Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-
-We excluded items that enailed more than one word (e.g., \emph{barrita de cereales} {[}cereal bar{]}).
-
-\hypertarget{item-properties}{%
-\subsubsection{Item properties}\label{item-properties}}
-
-\hypertarget{frequencies}{%
-\paragraph{Frequencies}\label{frequencies}}
-
-We retrieved frequencies from SUBTLEX-ESP ({\textbf{???}}) and SUBTLEX-CAT ({\textbf{???}}) for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word ({\textbf{???}}; {\textbf{???}}).
+We assessed toddlers' language background asking parents for an estimated proportion of exposure to each language.
+We defined participants as Spanish- or Catalan-dominant according to the language they were exposed the most to.
+Participants were considered bilingual - therefore included - if exposed 50-90\% of the time to the non-dominant language.
+We excluded participants with \textgreater{}10\% exposure to a third language.
+Fig.
+\ref{fig:lpdist} shows how participants were distributed across age bins and language profiles.
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/pool_freq_cognateness}
+{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/lpdist-1}
 
@@
 
-\caption{ }\label{fig:unnamed-chunk-1}
+\caption{ }\label{fig:lpdist}
 \end{figure}
 
-\hypertarget{cognate-status}{%
-\paragraph{Cognate status}\label{cognate-status}}
+\hypertarget{questionnaire}{%
+\subsection{Questionnaire}\label{questionnaire}}
 
-\hypertarget{demographics-by-item}{%
-\paragraph{Demographics by item}\label{demographics-by-item}}
+Parents completed an online questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
+We implemented the questionnaire in \texttt{formr} (Arslan, Walther, \& Tata, 2020), a free, open-source platform.
+Participants were randomly allocated into one of four list of items.
+Each list contained a distinct inventory of translation equivalents.
+We included items from a diverse sample of 26 semantic/funcional categories: Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
+We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
+Those categories that did not entail more than 16 items - thus resulting in less than four items after dividing it in four lists - were not divided: all of their items were included in the four lists.
+Another subset of items that were part of the trial lsits of some experiments in the lab were also included in all versions.
 
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-2}Demographics of responses by item.
-Language = item language, LP = language profile of participant, Sex = sex of participant, M Age = mean age of responses, SD Age = mean SD of responses across items, Min Age = minimum age of responses across items, Max Age = maximum age of responses across items, M N = mean number of responses across items, N = total number of responses across items.}
-\centering
-\begin{tabular}[t]{c|c|c|c|c|c|c|c|c}
-\hline
-Language & LP & Sex & M Age & SD Age & Min Age & Max Age & M N & N\\
-\hline
-Catalan & Bilingual & Female & 22.36 & 22.36 & 12.35 & 31.42 & 52.22 & 41620\\
-\hline
-Catalan & Bilingual & Male & 20.77 & 20.77 & 12.37 & 31.91 & 64.06 & 51057\\
-\hline
-Catalan & Monolingual & Female & 20.14 & 20.14 & 12.83 & 30.15 & 68.01 & 54204\\
-\hline
-Catalan & Monolingual & Male & 21.11 & 21.11 & 12.60 & 30.93 & 48.30 & 38495\\
-\hline
-Spanish & Bilingual & Female & 22.33 & 22.33 & 12.35 & 31.42 & 52.76 & 42209\\
-\hline
-Spanish & Bilingual & Male & 20.72 & 20.72 & 12.37 & 31.91 & 64.72 & 51778\\
-\hline
-Spanish & Monolingual & Female & 20.10 & 20.10 & 12.83 & 30.15 & 68.71 & 54969\\
-\hline
-Spanish & Monolingual & Male & 21.06 & 21.06 & 12.60 & 30.93 & 48.80 & 39037\\
-\hline
-\end{tabular}
-\end{table}
+Each item of the inventory presented a word-form, to which parents were instructed to report whether the toddlers was able to understand it, understand ans say it, or neither.
+The Catalan inventory contained 796 items (582 cognates, 195 non-cognates) and the Spanish inventory contained 801 (584 cognates, 197 non-cognates) .
+Items of one language were translation equivalents from the other (e.g., whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the word \emph{perro} was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar} and \emph{terminar}), or merged them into the same one (e.g., Spanish \emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}).
+We excluded from the analysis multi-word items (e.g., \emph{barrita de cereales} {[}cereal bar{]}) and items that included more than one word-form (e.g., \emph{mono / mico}).
 
@@
 
-For testing the first hypothesis, we fit a non-linear model using logistic curves to model the probability of aquisition of words across ages.
-Logistic curves are characterised by three parameters: 1) an asymptote (maximum value of the curve in the Y-axis), 2) the steepness (how fast the curve grows), and 3) a mid-point (the point in the X-axis at which the steepness is highest).
-Figure 1 shows the elements of a simulated logarithmic curve:
+For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item within age bins of 2 months.
+Our model assumes that this proportion follows a logistic curve accross ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the dominance of the language of the item (\emph{item dominance}, L1 vs.~L2), the cognate status of the word (\emph{cognateness}, Non-cognate vs.~Cognate), and the frequency of the word form in its correspondent language.
+Our prediction is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words.
+
+Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
+Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item porta (door, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
+We retrieved frequencies from SUBTLEX-ESP (Cuetos, Glez-Nosti, Barbón, \& Brysbaert, 2011) and SUBTLEX-CAT (Boada, Guasch, Haro, Demestre, \& Ferré, 2019) for Spanish and Catalan words, respectively.
+Word frequency was calculated as the Zipf score of the word (Heuven, Mandera, Keuleers, \& Brysbaert, 2014; Zipf, 1949).
+Missing frequency scores were imputed using multiple imputation (Buuren \& Groothuis-Oudshoorn, 2011).
+Table \ref{tab:itemsinfo} shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_illustration}
+{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/itemsinfo-1}
 
@@
 
-\caption{Simulated logarithmic curve.
-The curve is characterised by three parameters: Asymptote (red), mid-point (blue), and steepness (green).
-This curve is the result of simulating learning curves of 1000 items.
-We set fixed values for the asymptote at 1.
-Mid-points were randomly generated from a normal distribution truncated at 0 (no negative age values are allowed), with mean 23 (*SD* = 0.5), and steepness values were generated from a normal ditribution truncated at 0 (we assume that the proportion of participants that have acquired each item can only grow), with mean 0.1 (*SD* = 1) The solid black line represents the average proportion of participants who have acquired the item, and the shaded ribbon indicates the standard error of the mean.
-We suggest that learning curves of individual word-forms, as well as the evolution of vocabulary size across ages, follows this trend.
-Our model will estimate the three of to adress our hypotheses.}\label{fig:logcurve}
+\caption{ }\label{fig:itemsinfo}
 \end{figure}
 
-One reason for using logistic curves to fit the data is that previous studies describe lexical development as a non-linear process, where between 20 and 24 months a vocabulary spurt happens.
-An analysis of the available data in Wordbank ({\textbf{???}}) provides moderate support for this hypothesis.
-We fitted a logistic model using the proportion of infants that acquired each item as output, and age as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
-We included data from 13119 items from 20 languages\footnote{Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
-All the other available languages were excluded, as no (or insufficient) data were available.}, from infants between 16 and 30 months of age.
-Then we compared the fit of the logarithmic model agains a linear model, and found that the formr fitted the data slightly better (see Figure 2).
+We fit two separate logistic models on comprehensive and productive data, respectively.
+Logistic curves are characterised by three parameters(see Mahr, 2019, and @mahr2020 for an excellent introduction to this approach, in which we based this analysis): (a) an \emph{asymptote} (upper boundary of the proportion of infants that have acquired the item), (b) the \emph{steepness} (how fast the a given item is acquired by infants), and (c) a \emph{mid-point} (the age at which steepness is maximum).
+We work on the assumption that the trajectory of aquisition of items follows a sigmoidal shape (Mayor \& Plunkett, 2011), and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
+This approach allows us to model and estimate the age of acquisition of items explicitely.
+Our model estimated the asymptote and stepness as grand means across items (i.e., intercepts).
+Mid-points were estimated by a linear zero-one-beta-regression model\footnote{Zero-or-one-inflated beta regression models allow to account for overdispersion then using proportions as the output of a model accounting, while also allowing values to be at boundary (0 and 1).}(Ospina \& Ferrari, 2012) that included our predictors of interest (\emph{age bin}, \emph{item dominance} (L1/L2, sum-coded), \emph{cognateness} (\emph{non-cognate/cognate, sum-coded), the interaction between }item dominance* and \emph{cognateness}, and frequency (standardised Zipf score){]}.
+We also included a random intercept for each translation equivalent.
 
-A second reason for using logistic curves is that the parameters that characterise a logistic curve map onto our hypotheses quite easily.
-Out first hypothesis concerns the mid-point parameter, which indicates the point at which the steepness of the curve is maximum.
-In our model, the value of this parameter can be interpreted as the age at which the proportion of participants that have acquired a given item is at maximum rate.
-The mathematical interpretation of this parameter is easily mapped into a definition of age of acquisition.
+\[y_{ijk} = \frac{A}{1 + e^{(M-Age) \times S}}\]
 
-\begin{figure}
+\[M = \beta_{0} + I_{i0} + \beta_{1} Dominance_{j} + \beta_{2} Cognateness_{ik} + \beta_{3} Dominance_{ij} \times Cognateness_{ik} + \beta_{4} Frequency + \sigma_{ijk}\]
 
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_wordbank}
+We estimated the parameters of our model using the Bayesian framework.
+Bayesian models use the Bayes rule to combine prior knowlegde about the distribution of a given parameter, and the likelihood of each of the values it can take given tha observed data.
+We relied on the available data on Wordbank (Frank, Braginsky, Yurovsky, \& Marchman, 2017) to generate our priors (see Appendix 1 for more details about our priors).
+Finally, to test the contribution of the \emph{cognateness} predictor to the estimation of mid-points, we compared the fit of the model with a reduced (null) model that did not include the \emph{cognateness} factor using Pareto-smoothed importance sampling (PSIS; Vehtari, Gelman, \& Gabry, 2017; Vehtari, Simpson, Gelman, Yao, \& Gabry, 2019).
+We fit both null and alternative models using the R environment (R Core Team, 2013; RStudio Team, 2015) and the \texttt{brms} package (Bürkner, 2017), which relies on the probabilistic language Stan (Carpenter et al., 2017) to estimate posterior distributions.
+Data and model results will be processed and visualised using the R packages the \texttt{tidyverse} family of packages (Wickham et al., 2019) and the \texttt{tidybayes} package (Kay, 2020).
 
-}
+To test Hypothesis 2, we run a Bayesian anova using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included \emph{cognateness} as predictor.
+Using the \texttt{BayesFactor} R package (Morey \& Rouder, 2018), we computed a Bayes factor (\emph{BF}) that compared the likelihood of a linear model that included \emph{cognateness} as a predictor against the likelihood of a linear model that did not, under the light of our data.
 
-\caption{Logarithmic model fitted on Wordbank data, compared to a linear model.
-The logarithmic model fits the data slightly better.
-The estimated values of the three parameters involved in the logarithmic model (asymptote, mid-point and steepness) are indicated.}\label{fig:wordbank}
-\end{figure}
+\begin{table}
 
-We built a Bayesian logistic model that estimates the value of the mid-point and steepness across items.
-Under our hypothesis, mid-points are generated from a linear model that incorporates language profile (monolingual vs.~bilingual) and cognateness (non-cognate vs.~cognate) as predictors, where cognateness predictes ealier mid-points for bilingual, but not monolingual participants.
-Due to computational limitations, we set a fixed value for the asymptote at 0.75, according to our model fit on Wordbank data.
-Due to computational limitations, we did not run the same linear model to estimate steepness.
-Rather, this parameter was just included as an intercept.
-We compared how well this model fits the data \emph{versus} how a null model that does not include \emph{cognateness} as predictor does, by using Bayes factors and Pareto-smoothed importance sampling (PSIS; {\textbf{???}}).
-
-We used the information provided by Wordbank as priors to feed our Bayesian model.
-For mid-points, we specified a strong prior with a normal prior with mean = 23.08 and \emph{SD} = 1.
-For steepness, we also set a strong prior with a normal distribution with mean = 3.28 and SD = 2.
-We had little prior information about how cognateness impacts the mid-point of an item, and therefore we will set a weak prior.
-
-We fit both null and alternative models within the R environment ({\textbf{???}}; {\textbf{???}}) using the \texttt{brms} package ({\textbf{???}}), which relies on the probabilistic language Stan ({\textbf{???}}).
-Data and model results will be processed and visualised using the R packages the \texttt{tidyverse} family of packages ({\textbf{???}}) and the \texttt{tidybayes} package ({\textbf{???}}).
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{model-selection}{%
-\subsection{Model selection}\label{model-selection}}
-
-\hypertarget{posterior-distibutions}{%
-\subsection{Posterior distibutions}\label{posterior-distibutions}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-coefficients}
-
-}
-
-\caption{ }\label{fig:posteriors}
-\end{figure}
-
-\hypertarget{posterior-correlations}{%
-\subsection{Posterior correlations}\label{posterior-correlations}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-correlations}
-
-}
-
-\caption{ }\label{fig:correlations}
-\end{figure}
-
-\hypertarget{prior-predictive-checks}{%
-\subsection{Prior predictive checks}\label{prior-predictive-checks}}
-
-Following ({\textbf{???}}), we performed prior predictive chackes to disgnose the appropriateness of our priors.
-
-\hypertarget{posterior-predictive-checks}{%
-\subsection{Posterior predictive checks}\label{posterior-predictive-checks}}
-
-\hypertarget{model-diagnosis}{%
-\subsection{Model diagnosis}\label{model-diagnosis}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-convergence}
-
-}
-
-\caption{ }\label{fig:convergence}
-\end{figure}
-
-\hypertarget{auto-correlation}{%
-\subsubsection{Auto-correlation}\label{auto-correlation}}
-
-\hypertarget{potential-scale-reduction-factor-hatr}{%
-\subsubsection{\texorpdfstring{Potential scale reduction factor (\(\hat{R}\))}{Potential scale reduction factor (\textbackslash{}hat\{R\})}}\label{potential-scale-reduction-factor-hatr}}
-
-\hypertarget{posterior-predictive-checks-1}{%
-\subsection{Posterior predictive checks}\label{posterior-predictive-checks-1}}
-
-Does our model generate similar observations to the ones in our data-set?
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-\begin{itemize}
-\tightlist
-\item
-  Cross-sectional data for longitudinal claims
-\item
-  Data is right censored\\
-\item
-  Inventory:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Phonological forms?
-  \item
-    We can't say how frequently words are produced by the parents or the toddler.
-  \item
-    Although instructed to ignore imitations, it is difficult to say whether a toddler has \emph{really} acquired a word that produces, or she can only imitate it.
-  \item
-    Responses in the questionnaire rely heavily on parental memory.
-  \item
-    We don't know about the context at which words are heard or produced.
-  \item
-    Classification system of items (e.g., \emph{household items}) is adult centric. Children may use the words \emph{pretty} to name jewellery (Bates et al., 1994).
-  \end{itemize}
-\item
-  Mid and mid-upper class families are overrepresented in the sample.
-\end{itemize}
+\caption{\label{tab:priorstab}Prior distribution of the parameters in our models.
+We set specific informative priors for the asymptote, steepness, and mid-point of the curves (using data from Wordbank) and and weakly informative priors for the rest of the coefficients.}
+\centering
+\begin{tabular}[t]{c|c|c|c}
+\hline
+**Adjusted parameter** & **Coefficient** & **Notation** & **Distribution**\\
+\hline
+Asymptote & Intercept & \$A\$ & \$N(0.79, 0.05)\$\\
+\hline
+Steepness & Intercept & \$S\$ & \$N(1.79, 0.6)\$\\
+\hline
+Mid-point & Intercept & \$\textbackslash{}beta\_\{0\}\$ & \$N(2.52, 1)\$\\
+\hline
+Mid-point & Random Intercepts & \$I\_\{i0\}\$ & N(0, \$\textbackslash{}pi\$)\\
+\hline
+Mid-point & Slope & \$\textbackslash{}beta\_\{1\}\$ & \$N(0, 5)\$\\
+\hline
+Mid-point & Slope & \$\textbackslash{}beta\_\{2\}\$ & \$N(0, 5)\$\\
+\hline
+Mid-point & Slope & \$\textbackslash{}beta\_\{3\}\$ & \$N(0, 5)\$\\
+\hline
+Mid-point & Slope & \$\textbackslash{}beta\_\{4\}\$ & \$N(0, 5)\$\\
+\hline
+Mid-point & Intercept & \$\textbackslash{}varphi\$ & \$N(2, 1)\$\\
+\hline
+Mid-point & Intercept & \$COI\$ & Beta(1, 1)\\
+\hline
+Mid-point & Intercept & \$ZOI\$ & Beta(1, 1)\\
+\hline
+Mid-point & Std. Deviation & \$\textbackslash{}sigma\{ijk\}\$ & t(3, 0, 2.5)\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 {[}1{]} rlang\_0.4.6 here\_0.1 data.table\_1.12.8\\
-{[}4{]} truncnorm\_1.0-8 wesanderson\_0.3.6 patchwork\_1.0.0.9000
-{[}7{]} ggridges\_0.5.2 ggplot2\_3.3.1 lubridate\_1.7.8\\
-{[}10{]} tidyr\_1.0.3 readxl\_1.3.1 stringr\_1.4.0\\
-{[}13{]} forcats\_0.5.0 dplyr\_1.0.0 tibble\_3.0.1\\
-{[}16{]} magrittr\_1.5 knitr\_1.28 papaja\_0.1.0.9942
+{[}4{]} gt\_0.2.1 patchwork\_1.0.0.9000 ggridges\_0.5.2\\
+{[}7{]} janitor\_2.0.1 lubridate\_1.7.9 readxl\_1.3.1\\
+{[}10{]} forcats\_0.5.0 stringr\_1.4.0 dplyr\_1.0.0\\
+{[}13{]} purrr\_0.3.4 readr\_1.3.1 tidyr\_1.1.0\\
+{[}16{]} tibble\_3.0.1 ggplot2\_3.3.2 tidyverse\_1.3.0\\
+{[}19{]} magrittr\_1.5 knitr\_1.28 papaja\_0.1.0.9942
 
 loaded via a namespace (and not attached):
-{[}1{]} Rcpp\_1.0.4.6 pillar\_1.4.4 compiler\_3.6.3 cellranger\_1.1.0
-{[}5{]} plyr\_1.8.6 tools\_3.6.3 digest\_0.6.25 evaluate\_0.14\\
-{[}9{]} lifecycle\_0.2.0 gtable\_0.3.0 pkgconfig\_2.0.3 yaml\_2.2.1\\
-{[}13{]} xfun\_0.14 withr\_2.2.0 generics\_0.0.2 vctrs\_0.3.0\\
-{[}17{]} rprojroot\_1.3-2 grid\_3.6.3 tidyselect\_1.1.0 glue\_1.4.1\\
-{[}21{]} R6\_2.4.1 rmarkdown\_2.1 bookdown\_0.18 purrr\_0.3.4\\
-{[}25{]} backports\_1.1.7 scales\_1.1.1 ellipsis\_0.3.1 htmltools\_0.4.0
-{[}29{]} colorspace\_1.4-1 stringi\_1.4.6 munsell\_0.5.0 crayon\_1.3.4
+{[}1{]} Rcpp\_1.0.4.6 lattice\_0.20-38 ps\_1.3.3 assertthat\_0.2.1\\
+{[}5{]} rprojroot\_1.3-2 digest\_0.6.25 R6\_2.4.1 cellranger\_1.1.0\\
+{[}9{]} plyr\_1.8.6 backports\_1.1.8 reprex\_0.3.0 evaluate\_0.14\\
+{[}13{]} httr\_1.4.1 pillar\_1.4.4 rstudioapi\_0.11 callr\_3.4.3.9000\\
+{[}17{]} blob\_1.2.1 checkmate\_2.0.0 rmarkdown\_2.3 labeling\_0.3\\
+{[}21{]} webshot\_0.5.2 munsell\_0.5.0 broom\_0.5.6 compiler\_3.6.3\\
+{[}25{]} modelr\_0.1.8 xfun\_0.15 pkgconfig\_2.0.3 htmltools\_0.5.0\\
+{[}29{]} tidyselect\_1.1.0 bookdown\_0.18 fansi\_0.4.1 crayon\_1.3.4\\
+{[}33{]} dbplyr\_1.4.4 withr\_2.2.0 commonmark\_1.7 grid\_3.6.3\\
+{[}37{]} nlme\_3.1-147 jsonlite\_1.6.1 gtable\_0.3.0 lifecycle\_0.2.0\\
+{[}41{]} DBI\_1.1.0 scales\_1.1.1 cli\_2.0.2 stringi\_1.4.6\\
+{[}45{]} farver\_2.0.3 fs\_1.4.1 snakecase\_0.11.0 xml2\_1.3.2\\
+{[}49{]} ellipsis\_0.3.1 generics\_0.0.2 vctrs\_0.3.1 RColorBrewer\_1.1-2
+{[}53{]} tools\_3.6.3 glue\_1.4.1 hms\_0.5.3 processx\_3.4.2\\
+{[}57{]} yaml\_2.2.1 colorspace\_1.4-1 rvest\_0.3.5 haven\_2.3.1\\
+{[}61{]} sass\_0.2.0
 
@@
 \subsection{Appendix 2: Model}\label{appendix-2-model}}
+
+Our model is incomplete, in that we have not included all parameters that may play a role in the real mechanisms that generate our observed data.
+Previous studies have found robust effects of sex (e.g., Frank et al.
+(2017)), word length/number of phonemes (e.g., Frank et al.
+(2017)), etc.
+Computational limitations have prevented us from specifying and fitting a more exhaustive model.
+Following the model-building strategy suggested by Schad, Betancourt, and Vasishth (2020), we have specified an intitial model that captures the fundamental structure of our phenomenon of interest and that can be computed with our available resources.
+Then, we have expanded such model incorporating a new predictor that adresses our hypotheses.
+
+\hypertarget{appendix-3-priors}{%
+\subsection{Appendix 3: Priors}\label{appendix-3-priors}}
+
+We fitted a logistic model using the proportion of infants that acquired each item as output, and age-bin (defined as we did with our data) as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
+We included data from 13119 items from 20 languages\footnote{Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
+All the other available languages were excluded, as no (or insufficient) data were available.}, from monolingual infants aged 16 to 30 months of age.
+Table \ref{tab:wordbank} shows the estimates of this model, which were rescaled and included as priors in our main model to test Hypothesis 1.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/worbank-1}
+
+}
+
+\caption{ }\label{fig:worbank}
+\end{figure}
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-bates1994}{}%
-Bates, E., Marchman, V., Thal, D., Fenson, L., Dale, P., Reznick, J.
-S., \ldots{} Hartung, J.
-(1994).
-Developmental and stylistic variation in the composition of early vocabulary.
-\emph{Journal of Child Language}, \emph{21}(1), 85--123.
-\url{https://doi.org/10.1017/S0305000900008680}
+\leavevmode\hypertarget{ref-arslan2020}{}%
+Arslan, R.
+C., Walther, M.
+P., \& Tata, C.
+S.
+(2020).
+Formr: A study framework allowing for automated feedback generation and complex longitudinal experience-sampling studies using R.
+\emph{Behavior Research Methods}, \emph{52}(1), 376--387.
+\url{https://doi.org/10.3758/s13428-019-01236-y}
+
+\leavevmode\hypertarget{ref-ben-zeev1977}{}%
+Ben-Zeev, S. (1977). The Influence of Bilingualism on Cognitive Strategy and Cognitive Development. \emph{Child Development}, \emph{48}(3), 1009--1018. \url{https://doi.org/10.2307/1128353}
+
+\leavevmode\hypertarget{ref-bialystok2010}{}%
+Bialystok, E., Luk, G., Peets, K.
+F., \& Yang, S.
+(2010).
+Receptive vocabulary differences in monolingual and bilingual children.
+\emph{Bilingualism: Language and Cognition}, \emph{13}(4), 525--531.
+\url{https://doi.org/10.1017/S1366728909990423}
+
+\leavevmode\hypertarget{ref-blom2019}{}%
+Blom, E., Boerma, T., Bosma, E., Cornips, L., Heuij, K.
+van den, \& Timmermeister, M.
+(2019).
+Cross-language distance influences receptive vocabulary outcomes of bilingual children.
+\emph{First Language}, 0142723719892794.
+\url{https://doi.org/10.1177/0142723719892794}
+
+\leavevmode\hypertarget{ref-boada2019}{}%
+Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P.
+(2019).
+SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for Catalan.
+\emph{Behavior Research Methods}.
+\url{https://doi.org/10.3758/s13428-019-01233-1}
+
+\leavevmode\hypertarget{ref-bobb2020}{}%
+Bobb, S.
+C., Von Holzen, K., Mayor, J., Mani, N., \& Carreiras, M.
+(2020).
+Co-activation of the L2 during L1 auditory processing: An ERP cross-modal priming study.
+\emph{Brain and Language}, \emph{203}, 104739.
+\url{https://doi.org/10.1016/j.bandl.2019.104739}
 
@@
 \url{https://doi.org/10.1177/0165025414532559}
+
+\leavevmode\hypertarget{ref-braginsky2019}{}%
+Braginsky, M., Yurovsky, D., Marchman, V.
+A., \& Frank, M.
+C.
+(2019).
+Consistency and Variability in Children's Word Learning Across Languages.
+\emph{Open Mind}, \emph{3}, 52--67.
+\url{https://doi.org/10.1162/opmi_a_00026}
+
+\leavevmode\hypertarget{ref-vanbuuren2011}{}%
+Buuren, S.
+van, \& Groothuis-Oudshoorn, K.
+(2011).
+\textless{}span class="nocase"\textgreater{}mice\textless{}/span\textgreater{}: Multivariate imputation by chained equations in r.
+\emph{Journal of Statistical Software}, \emph{45}(3), 1--67.
+Retrieved from \url{https://www.jstatsoft.org/v45/i03/}
+
+\leavevmode\hypertarget{ref-burkner2017}{}%
+Bürkner, P.-C. (2017). Brms: An R Package for Bayesian Multilevel Models Using Stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+
+\leavevmode\hypertarget{ref-carpenter2017}{}%
+Carpenter, B., Gelman, A., Hoffman, M.
+D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A.
+(2017).
+Stan: A Probabilistic Programming Language.
+\emph{Journal of Statistical Software}, \emph{76}(1), 1--32.
+\url{https://doi.org/10.18637/jss.v076.i01}
+
+\leavevmode\hypertarget{ref-core2013}{}%
+Core, C., Hoff, E., Rumiche, R., \& Señor, M.
+(2013).
+Total and Conceptual Vocabulary in Spanish--English Bilinguals From 22 to 30 Months: Implications for Assessment.
+\emph{Journal of Speech, Language, and Hearing Research}, \emph{56}(5), 1637--1649.
+\url{https://doi.org/10.1044/1092-4388(2013/11-0044)}
+
+\leavevmode\hypertarget{ref-costa2000}{}%
+Costa, A., Caramazza, A., \& Sebastian-Galles, N.
+(2000).
+The Cognate Facilitation Effect: Implications for Models of Lexical Access.
+\emph{Journal of Experimental Psychology}, \emph{26}(5), 1283--1296.
+\url{https://doi.org/10.1037/0278-7393.26.5.1283}
+
+\leavevmode\hypertarget{ref-cuetos2011}{}%
+Cuetos, F., Glez-Nosti, M., Barbón, A., \& Brysbaert, M. (2011). SUBTLEX-ESP: Spanish word frequencies based on film subtitles. \emph{Psicológica}, \emph{32}(2), 133--143.
+
+\leavevmode\hypertarget{ref-de1991lexical}{}%
+De Groot, A. M., \& Nas, G. L. (1991). Lexical representation of cognates and noncognates in compound bilinguals. \emph{Journal of Memory and Language}, \emph{30}(1), 90--123.
+
+\leavevmode\hypertarget{ref-dijkstra2002architecture}{}%
+Dijkstra, A., \& Van Heuven, W. J. (2002). The architecture of the bilingual word recognition system: From identification to decision.
+
+\leavevmode\hypertarget{ref-dijkstra2005}{}%
+Dijkstra, T. (2005). Bilingual Visual Word Recognition and Lexical Access. In \emph{Handbook of bilingualism: Psycholinguistic approaches} (pp. 179--201). New York, NY, US: Oxford University Press.
+
+\leavevmode\hypertarget{ref-dijkstra1998bia}{}%
+Dijkstra, T., \& Van Heuven, W. J. (1998). The BIA model and bilingual word recognition. \emph{Localist Connectionist Approaches to Human Cognition}, 189--225.
+
+\leavevmode\hypertarget{ref-dijkstra2019}{}%
+Dijkstra, T., Wahl, A., Buytenhuijs, F., Halem, N.
+V., Al-Jibouri, Z., Korte, M.
+D., \& Rekké, S.
+(2019).
+Multilink: A computational model for bilingual word recognition and word translation.
+\emph{Bilingualism: Language and Cognition}, \emph{22}(4), 657--679.
+\url{https://doi.org/10.1017/S1366728918000287}
+
+\leavevmode\hypertarget{ref-doyle1977some}{}%
+Doyle, A. B., \& others. (1977). Some issues in the assessment of linguistic consequences of early bilingualism. \emph{Working Papers on Bilingualism}, (14).
+
+\leavevmode\hypertarget{ref-fernandez1992}{}%
+Fernandez, M.
+C., Pearson, B.
+Z., Umbel, V.
+M., Oiler, D.
+K., \& Molinet-Molina, M.
+(1992).
+Bilingual Receptive Vocabulary in Hispanic Preschool Children.
+\emph{Hispanic Journal of Behavioral Sciences}, \emph{14}(2), 268--276.
+\url{https://doi.org/10.1177/07399863920142006}
 
@@
 
+\leavevmode\hypertarget{ref-frank2017}{}%
+Frank, M.
+C., Braginsky, M., Yurovsky, D., \& Marchman, V.
+A.
+(2017).
+Wordbank: An open repository for developmental vocabulary data.
+\emph{Journal of Child Language}, \emph{44}(3), 677--694.
+\url{https://doi.org/10.1017/S0305000916000209}
+
+\leavevmode\hypertarget{ref-heuven2014}{}%
+Heuven, W.
+J.
+B.
+van, Mandera, P., Keuleers, E., \& Brysbaert, M.
+(2014).
+Subtlex-UK: A New and Improved Word Frequency Database for British English: \emph{Quarterly Journal of Experimental Psychology}.
+Retrieved from \url{https://journals.sagepub.com/doi/10.1080/17470218.2013.850521}
+
+\leavevmode\hypertarget{ref-hoff2012}{}%
+Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M.
+(2012).
+Dual language exposure and early bilingual development*.
+\emph{Journal of Child Language}, \emph{39}(1), 1--27.
+\url{https://doi.org/10.1017/S0305000910000759}
+
+\leavevmode\hypertarget{ref-hoff2014}{}%
+Hoff, E., Rumiche, R., Burridge, A., Ribot, K.
+M., \& Welsh, S.
+N.
+(2014).
+Expressive vocabulary development in children from bilingual and monolingual homes: A longitudinal study from two to four years.
+\emph{Early Childhood Research Quarterly}, \emph{29}(4), 433--444.
+\url{https://doi.org/10.1016/j.ecresq.2014.04.012}
+
+\leavevmode\hypertarget{ref-hoshino2008}{}%
+Hoshino, N., \& Kroll, J.
+F.
+(2008).
+Cognate effects in picture naming: Does cross-language activation survive a change of script?
+\emph{Cognition}, \emph{106}(1), 501--511.
+\url{https://doi.org/10.1016/j.cognition.2007.02.001}
+
+\leavevmode\hypertarget{ref-houwer2014}{}%
+Houwer, A.
+D., Bornstein, M.
+H., \& Putnick, D.
+L.
+(2014).
+A bilingual--monolingual comparison of young children's vocabulary size: Evidence from comprehension and production.
+\emph{Applied Psycholinguistics}, \emph{35}(6), 1189--1211.
+\url{https://doi.org/10.1017/S0142716412000744}
+
+\leavevmode\hypertarget{ref-kay2020}{}%
+Kay, M.
+(2020).
+\emph{\textless{}span class="nocase"\textgreater{}tidybayes\textless{}/span\textgreater{}: Tidy data and geoms for Bayesian models} (manual).
+\url{https://doi.org/10.5281/zenodo.1308151}
+
+\leavevmode\hypertarget{ref-kroll2013}{}%
+Kroll, J.
+F., Gullifer, J.
+W., \& Rossi, E.
+(2013).
+The Multilingual Lexicon: The Cognitive and Neural Basis of Lexical Comprehension and Production in Two or More Languages.
+\emph{Annual Review of Applied Linguistics}, \emph{33}, 102--127.
+\url{https://doi.org/10.1017/S0267190513000111}
+
+\leavevmode\hypertarget{ref-kroll2010}{}%
+Kroll, J.
+F., Hell, J.
+G.
+V., Tokowicz, N., \& Green, D.
+W.
+(2010).
+The Revised Hierarchical Model: A critical review and assessment*.
+\emph{Bilingualism: Language and Cognition}, \emph{13}(3), 373--381.
+\url{https://doi.org/10.1017/S136672891000009X}
+
+\leavevmode\hypertarget{ref-kroll1994}{}%
+Kroll, J.
+F., \& Stewart, E.
+(1994).
+Category interference in translation and picture naming: Evidence for asymmetric connection between bilingual memory representations.
+\emph{Journal of Memory and Language}, \emph{33}(2), 149--174.
+\url{https://doi.org/10.1006/jmla.1994.1008}
+
+\leavevmode\hypertarget{ref-mahr2019}{}%
+Mahr, T. (2019). Anatomy of a logistic growth curve. \emph{Higher Order Functions}. Retrieved from \url{https://tjmahr.github.io/anatomy-of-a-logistic-growth-curve/}
+
+\leavevmode\hypertarget{ref-mahr2020}{}%
+Mahr, T. (2020). Bayes' theorem in three panels. \emph{Higher Order Functions}. Retrieved from \url{https://tjmahr.github.io/bayes-theorem-in-three-panels/}
+
+\leavevmode\hypertarget{ref-mayor2011}{}%
+Mayor, J., \& Plunkett, K.
+(2011).
+A statistical estimate of infant and toddler vocabulary size from CDI analysis.
+\emph{Developmental Science}, \emph{14}(4), 769--785.
+\url{https://doi.org/10.1111/j.1467-7687.2010.01024.x}
+
+\leavevmode\hypertarget{ref-morey2018a}{}%
+Morey, R. D., \& Rouder, J. N. (2018). \emph{BayesFactor: Computation of bayes factors for common designs} (manual). Retrieved from \url{https://CRAN.R-project.org/package=BayesFactor}
+
+\leavevmode\hypertarget{ref-ospina2012}{}%
+Ospina, R., \& Ferrari, S.
+L.
+P.
+(2012).
+A general class of zero-or-one inflated beta regression models.
+\emph{Computational Statistics \& Data Analysis}, \emph{56}(6), 1609--1623.
+\url{https://doi.org/10.1016/j.csda.2011.10.005}
+
+\leavevmode\hypertarget{ref-pearson1994}{}%
+Pearson, B.
+Z., \& Fernández, S.
+C.
+(1994).
+Patterns of Interaction in the Lexical Growth in Two Languages of Bilingual Infants and Toddlers.
+\emph{Language Learning}, \emph{44}(4), 617--653.
+\url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}
+
+\leavevmode\hypertarget{ref-pearson1993}{}%
+Pearson, B.
+Z., Fernández, S.
+C., \& Oller, D.
+K.
+(1993).
+Lexical Development in Bilingual Infants and Toddlers: Comparison to Monolingual Norms.
+\emph{Language Learning}, \emph{43}(1), 93--120.
+\url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}
+
+\leavevmode\hypertarget{ref-rcoreteam2013}{}%
+R Core Team. (2013). \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{http://www.R-project.org/}
+
+\leavevmode\hypertarget{ref-rosenblum1983word}{}%
+Rosenblum, T., \& Pinker, S.
+A.
+(1983).
+Word magic revisited: Monolingual and bilingual children's understanding of the word-object relationship.
+\emph{Child Development}, 773--780.
+\url{https://doi.org/10.2307/1130064}
+
+\leavevmode\hypertarget{ref-rstudioteam2015}{}%
+RStudio Team. (2015). \emph{RStudio: Integrated Development Environment for R}. Boston, MA: RStudio, Inc. Retrieved from \url{http://www.rstudio.com/}
+
+\leavevmode\hypertarget{ref-schad2020}{}%
+Schad, D.
+J., Betancourt, M., \& Vasishth, S.
+(2020).
+Toward a principled Bayesian workflow in cognitive science.
+\emph{arXiv:1904.12765 {[}Stat{]}}.
+Retrieved from \url{http://arxiv.org/abs/1904.12765}
+
+\leavevmode\hypertarget{ref-shook2013}{}%
+Shook, A., \& Marian, V.
+(2013).
+The Bilingual Language Interaction Network for Comprehension of Speech*.
+\emph{Bilingualism: Language and Cognition}, \emph{16}(2), 304--324.
+\url{https://doi.org/10.1017/S1366728912000466}
+
+\leavevmode\hypertarget{ref-thierry2007}{}%
+Thierry, G., \& Wu, Y.
+J.
+(2007).
+Brain potentials reveal unconscious translation during foreign-language comprehension.
+\emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535.
+\url{https://doi.org/10.1073/pnas.0609927104}
+
+\leavevmode\hypertarget{ref-vehtari2017}{}%
+Vehtari, A., Gelman, A., \& Gabry, J.
+(2017).
+Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC.
+\emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
+\url{https://doi.org/10.1007/s11222-016-9696-4}
+
+\leavevmode\hypertarget{ref-vehtari2019}{}%
+Vehtari, A., Simpson, D., Gelman, A., Yao, Y., \& Gabry, J. (2019). Pareto Smoothed Importance Sampling. \emph{arXiv:1507.02646 {[}Stat{]}}. Retrieved from \url{http://arxiv.org/abs/1507.02646}
+
+\leavevmode\hypertarget{ref-vonholzen2019}{}%
+Von Holzen, K., Fennell, C.
+T., \& Mani, N.
+(2019).
+The impact of cross-language phonological overlap on bilingual and monolingual toddlers' word recognition.
+\emph{Bilingualism: Language and Cognition}, \emph{22}(3), 476--499.
+\url{https://doi.org/10.1017/S1366728918000597}
+
+\leavevmode\hypertarget{ref-vonholzen2012}{}%
+Von Holzen, K., \& Mani, N.
+(2012).
+Language nonselective lexical access in bilingual toddlers.
+\emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.
+\url{https://doi.org/10.1016/j.jecp.2012.08.001}
+
+\leavevmode\hypertarget{ref-wickham2019}{}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
+D., François, R., \ldots{} Yutani, H.
+(2019).
+Welcome to the \textless{}span class="nocase"\textgreater{}tidyverse\textless{}/span\textgreater{}.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
+
+\leavevmode\hypertarget{ref-yudes2010}{}%
+Yudes, C., Macizo, P., \& Bajo, T. (2010). Cognate effects in bilingual language comprehension tasks. \emph{NeuroReport}, \emph{21}(7), 507--512. \url{https://doi.org/10.1097/WNR.0b013e328338b9e1}
+
+\leavevmode\hypertarget{ref-zipf1949}{}%
+Zipf, G. K. (1949). \emph{Human behavior and the principle of least effort}. Oxford, England: Addison-Wesley Press.
+
 \endgroup

### 2020-06-30 Manuscript/trajectories_manuscript_appendix.tex
First version of the Introduction and Methods, and second version of the ICIS 2020 poster

--- 2020-06-15 Manuscript/trajectories_manuscript.tex
@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[english,man,man,floatsintext]{apa6}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
-\hypersetup{
-            pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
-            pdfkeywords={word acquisition, vocabulary, bilingualism, lexical similarity},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
+\clearpage
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
+\efloat@restorefloats
 \makeatother
 
-% Manuscript styling
-\usepackage{upgreek}
-\captionsetup{font=singlespacing,justification=justified}
+\begin{appendix}
+\section{}
+We fitted a logistic model using the proportion of infants that acquired
+each item as output, and age-bin (Defined as we did with our data) as
+input, estimating the three parameters afore mentioned (asymptote,
+mid-point, and steepness). We included data from 13119 items from 20
+languages\footnote{Chinese (Cantonese and Mandarin), Croatian, Czech,
+  Danish, English (Australian, British, and American), French
+  (European), German, Hebrew, Italian, Korean, Latvian, Portuguese,
+  Russian, Slovak, Spanish (European and Mexican), and Turkish. All the
+  other available languages were excluded, as no (or insufficient) data
+  were available.}, from monolingual infants aged 16 to 30 months of
+age.
 
-% Table formatting
-\usepackage{longtable}
-\usepackage{lscape}
-% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
-\usepackage{multirow}		% Table styling
-\usepackage{tabularx}		% Control Column width
-\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
-\usepackage{threeparttablex}            % Lets threeparttable work with longtable
-
-% Create new environments so endfloat can handle them
-% \newenvironment{ltable}
-%   {\begin{landscape}\begin{center}\begin{threeparttable}}
-%   {\end{threeparttable}\end{center}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
-
-% Enables adjusting longtable caption width to table width
-% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
-\makeatletter
-\newcommand\LastLTentrywidth{1em}
-\newlength\longtablewidth
-\setlength{\longtablewidth}{1in}
-\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
-
-% \setlength{\parindent}{0.5in}
-% \setlength{\parskip}{0pt plus 0pt minus 0pt}
-
-% \usepackage{etoolbox}
-\makeatletter
-\patchcmd{\HyOrg@maketitle}
-  {\section{\normalfont\normalsize\abstractname}}
-  {\section*{\normalfont\normalsize\abstractname}}
-  {}{\typeout{Failed to patch abstract.}}
-\makeatother
-\shorttitle{Cross-linguistic similarity and  word acquisition}
-\author{Gonzalo García-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Núria Sebastian-Galles\textsuperscript{1}}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra}
-\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-Enter author note here.
-
-Correspondence concerning this article should be addressed to Gonzalo García-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain. E-mail: gonzalo.garciadecastro@upf.edu}
-\keywords{word acquisition, vocabulary, bilingualism, lexical similarity\newline\indent Word count: X}
-\usepackage{lineno}
-
-\linenumbers
-\usepackage{csquotes}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[shorthands=off,main=english]{babel}
-\else
-  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\fi
-
-\title{The role of cross-linguistic lexical similarity on bilingual word acquisition}
-
-\date{}
-
-\abstract{
-Previous literature on early vocabulary has been mostly committed to exploring how the receptive and productive vocabulary size changes with age, and its relationship with toddlers' performance on language tasks (e.g.~Fernald, Swingley, \& Pinto, 2001; Fernald, Perfors, \& Marchman, 2006).
-The content of the developing lexicon has remained relatively unexplored.
-More recently, studies have focused on the characterisation of the developmental trajectory of individual words, reporting an earlier age of acquisition for words with high frequency, concreteness, and phonological neighbourhood density (e.g.~Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019).
-Such item-level analysis allows not only to predict the age of acquisition of specific words, but also to shed light on the cognitive processes that underlie early word learning.
-This approach is particularly interesting for investigating the potential impact of acquiring two languages simultaneously.
-A recent study by Floccia et al.
-(2018) compared vocabulary sizes of 24-month-old children learning British English, together with an additional language from a pool of 13 diverse languages.
-They found larger productive vocabulary sizes of toddlers learning two languages that are phonologically similar (see for a similar approach Bosch \& Ramon-Casas, 2014).
-The mechanisms underlying this effect remain unknown.
-One possibility is that the similarity between the two languages speeds the acquisition of form-similar translation equivalents.
-The aim of this study is to perform an item-wise analysis on infant's vocabulary contents.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish.
-We expect that phonological overlap between translation equivalents will predict earlier age of acquisition.
-If this effect is driven by the phonological overlap between translation equivalents, cognate pairs should be acquired closer in time than non-cognates.
-For instance, the translation of cat /gato/ in Spanish and /gat/ in Catalan should be learnt at approximately the same age.
-This should not necessarily happen for the translations of dog, /pero/ and /gos/.
-We analyse the time elapsed between the acquisition of a word in one language, and its translation in the other.
-We present preliminary data (data collection is ongoing) on receptive vocabulary of 90 monolingual and bilingual toddlers aged 18 to 36 months.
-We obtained parental responses to 230 pairs of Catalan-Spanish translation equivalents, resulting in a total of 41490 responses.
-Words that yielded a phonologically similar translation equivalent were more likely to be present in toddlers' receptive vocabulary at all ages, and were acquired earlier, even when accounting for the effect of word frequency.
-Further analysis will explore the distance in time between the acquisition of pairs of translation equivalents using a more detailed measure of phonological overlap across translation equivalents and taking into account cross-individual and cross-item variability.
-}
-
-\begin{document}
-\maketitle
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-One of the main challenges bilingual infants and toddlers face is learning two distinct set of words (one for each language) that partially overlap in sound and meaning.
-Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When taking both languages into account, bilingual children between seem to know, at least, as many words as monolinguals do ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
-Tough, not all bilinguals show asimilar developmental trajectory of lexical acquisition.
-
-Recent studies have capitalised on the role of the similarity between the specific pair of language the infant is learning.
-Floccia et al.
-(2018) analysed vocabulary scores across of 24-month-old toddlers learning English and an additional language.
-Toddlers learning languages sharing a large amounth of cognates (e.g., English-German) showed larger vocabulary sizes than those learning less languages sharing less cognates (e.g., English-Chinese).
-({\textbf{???}}) extended these results to children aged three to 10 years.
-These results point to the possibility that the overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
-However, none of these studies provide an account for the mechanisms involved in this facilitatory effect of language driven by similarity across languages.
-The aim of this study is to explore two scenarios in which this effect may be taking place:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  In a first scenario, it is the phonological overlap between language pairs, and not the amount of cognates they share, what boosts the vocabulary growthof bilingual toddlers.
-Languages sharing a lot of cognates tend to also share a lot of phonemic categories.
-It is possible that having to learn fewer phonemic categories makes it easier for toddlers to acquire words.
-If this is true, toddlers learning two languages with similar phonemic inventories should show larger vocabulary sizes than those learning languages that barely ovelap phonologically.
-Moreover, this facilitation effect should be reflected in the developmental trajectory of all the items included in a vocabulary inventory, independently of their cognate status.
-\item
-  In a second scenario, Words that are form similar It is the amount of cognates language pairs share that boosts vocabulary growth in bilingual toddlers.
-If this is true, toddlers may acquire cognates earlier
-\end{enumerate}
-
-To our knowledge, there are no previous studies that have investigated the effect of cognateness on the acquisition trajectories of individual word forms.
-Braginsky et al.
-(2019) used data from MacArthur-Bates CDIs from multiple languages to explore what properties are associated with an earlier acquisition.
-The found that frequency, concreteness, and {[}\ldots{}{]} significantly predicted an earlier age of acquisition.
-We plan to extend these analyses to the case of bilinguals.
-
-We have developed a fully-online vocabulary inventory in Catalan and Spanish that is more exhaustive than the MacArthur-Bates CDI, including a larger number of items.
-All participants filled the questionnaire it both languages, thus providing responses to pairs of translation equivalents.
-
-In this study, we examine the role of cognateness on age of acquisition by modelling the probability of parents reporting that the participant is able to understand, say and understand on each item in the questionnaire.
-We will test the effect the participants' language profile (monolingual vs.~bilingual), and its cognate status on the developmetnal curves of each item.
-
-We will analyse the developmental trajectories of individual words in monolinguals and bilinguals learning Catalan and Spanish (two phonologically close languages sharing a great deal of cognates), between 14 and 30 months of age.
-We will test whether cognate words are learnt earlier than non-cognate words, taking into account wether they are part of the most- o less-dominant language of the infant, as well as their frequency, concreteness, and number of syllables.
-We have deloped an on-line questionnaire that entails a short language exposure questionnaire, asks for some demographic information, and presents parent with a list of \textasciitilde{}800 words in each language, for which parent report whether their child understands, says, or doesn't understand and say, in each item.
-
-We will fit a Bayesian multilevel model on the item responses, modelling the probability of parents reporting that a given word is acquired by the infant, including language profile (monolingual vs.~bilingual), item dominance (dominant language vs.~non-dominant language), frequency, concreteness, and number of syllables.
-We will include \emph{meaning} as grouping variable.
-We will fit a sigmoidal curve (see {\textbf{???}}) on each item, defined by the steepness of the developmental curve, and the mid-point (the poinnt at which the developmental curve is steepest, which will be considered the point of acquisition).
-Priors will be derived from previous literature on the subject.
-
-NOTES:
-
-Evidence that bilingual lexica barely overlap ({\textbf{???}}; {\textbf{???}}).
-
-Studies with infants from 8 to 30 months of age show weaker evidence for this claim ({\textbf{???}}; {\textbf{???}}), pointing to the posibility that differences between bilingual and monolingual vocabulary sizes may be dependent on maturational factors.
-
-sugested that the developmental trajectory of lexical aquisition varies across two dimensions: the language profile of the infant, and the joint proporties of the specific language pair the infant is learning.
-Regarding the first, ({\textbf{???}}) found thant infants exposed to a more balanced
-
-Of special interest is the fact than cognates (form-similar translation equivalents) tend to be overrepresented in the early bilingual vocabulary (Bosch \& Ramon-Casas, 2014).
-
-The similarity between both languages may play a role in bilingual lexical aquisition.
-
-\hypertarget{hypotheses}{%
-\subsection{Hypotheses}\label{hypotheses}}
-
-\begin{description}
-\tightlist
-\item[Hypothesis 1]
-The boost effect of language similarity is driven by cognates being acquired earlier than non-cognates.
-We predict that age of acquisition of cognates will be ealier thant that of non-cognates.
-This effect should only be present in bilinguals.
-\item[Hypothesis 2]
-That this effect is driven by the word-forms in one language scaffolding the acquisition of their translation equivalents via parallel activation.
-We predict that the difference in time between the acquisition pairs of translation equivalents is shorter in cognates than in non-cognates.
-This effect will only be present in bilinguals.
-\end{description}
-
-\hypertarget{method}{%
-\section{Method}\label{method}}
-
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
-
-Data from 235 participants from the different cities in Spain were invited to participate through social media.
-Families in the Metropolitan area of Barcelona were also recruited at birth.
-All families participated voluntarily.
-Data were collected between 28th October, 2019 and 31th May, 2020, using the BiLexicon inventory.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-
-\begin{table}
-
-\caption{\label{tab:participants}Sample demographics and overall language profile.
-Dominance = Language to which the participant is exposed the most, N = sample size, M age = Mean age, SD age = standard deviation of age, M DOE Cat = mean degree of exposure to Catalan, SD DOE Cat = standard deviation of exposure to Catalan, M DOE Spa = mean degree of exposure to Spanish, SD DOE Spa = standard deviation of exposure to Spanish.}
-\centering
-\begin{tabular}[t]{c|c|c|c|c|c|c|c|c}
-\hline
-Dominance & Sex & N & M\_Age & SD\_Age & M DOE Cat & SD\_DOE\_Cat & M\_DOE\_Spa & SD\_DOE\_Spa\\
-\hline
-Catalan & Female & 126,526 & 20.96 & 4.59 & 73.94 & 18.48 & 23.52 & 18.29\\
-\hline
-Catalan & Male & 132,856 & 20.86 & 4.78 & 70.83 & 17.74 & 27.77 & 17.08\\
-\hline
-Spanish & Female & 66,476 & 21.22 & 4.76 & 17.38 & 14.74 & 81.14 & 15.60\\
-\hline
-Spanish & Male & 47,511 & 20.83 & 4.88 & 26.67 & 13.12 & 72.66 & 11.67\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{itemize}
-\tightlist
-\item
-  \emph{Location}
-\item
-  \emph{Age}
-\item
-  \emph{Sex}
-\item
-  \emph{Language profile}: : Linguistic profiles were assessed using the Language Exposure Questionnaire (LEQ) ({\textbf{???}}).
-This questionnaire estimates de degree of exposure (DOE) to each language via a 10 minutes long interview with the parents.
-LEQs were administered inmediately before or after the testing session at the lab.
-We here considered only exposure to Catalan and Spanish.
-Participants with \textgreater{}10\% exposure to a third language other than Catalan or Spanish were excluded.
-\item
-  \emph{SES}
-\end{itemize}
-
-\hypertarget{exclusion-criteria}{%
-\subsubsection{Exclusion criteria}\label{exclusion-criteria}}
-
-\begin{itemize}
-\tightlist
-\item
-  Language disorders
-\end{itemize}
-
-\hypertarget{questionnaires}{%
-\subsubsection{Questionnaires}\label{questionnaires}}
-
-\hypertarget{bilexicon-1.0}{%
-\paragraph{BiLexicon 1.0}\label{bilexicon-1.0}}
-
-\hypertarget{bilexicon-2.0}{%
-\paragraph{BiLexicon 2.0}\label{bilexicon-2.0}}
-
-\hypertarget{bilexicon-short}{%
-\paragraph{BiLexicon Short}\label{bilexicon-short}}
-
-This questionnaire was developed as a short version of the BiLexicon 1.0 and BiLexicon 2.0 inventories.
-Given the large amount of items to be answered to, we restricted the use of such inventories to participants that were to participate presentially in one of the studies run in the lab at that time.
-These participants only filled one of the questionnaires, and did so only once.
-
-The aim of developing a short version of this questionnaire (BiLexicon Short) was to provde a more on-line friendly format to families that would participate remotely exclusively, or that would fill the questionnaire more than once (e.g., in longitudinal studies).
-We divided the items of each language in the pool into four versions named A, B, C, and D.
-Items in each category (e.g., animals, auxiliary words) were randomly assigned one of the versions.
-Thos categories that did not entail more than 16 items (thus resulting in less than four items after dividing it in ofur lists), were not divided (all of their items were included in all versions).
-
-When filling the questionnaire for the first time, participants were randomly assigned one version.
-Each familiy filled only one of the versions, thus reducing the number of items to be answered.
-To ensure that longitudinal participants would provide longitudinal data in the questionnaire, we made sure that participants were always assigned the same version of this questionnaire.
-Words that were used as part of the trial lists in the experiment families would participate in were present across all versions.
-
-\hypertarget{item-inclusion-criteria}{%
-\subsubsection{Item inclusion criteria}\label{item-inclusion-criteria}}
-
-We gathered data from 695 distinct translation equivalents across Spanish (695 word forms) and Catalan (703 word forms).
-The unequal number of word-orms across both languages is due to that some tanslation equivalents did not entail a one-to-one-corespondence: some items had more than one translation equivalent in the other language (e.g.~the Catalan word form \emph{walking}, \emph{anar}, can be translated as both \emph{caminar} {[}to walk{]} or \emph{ir} {[}to go{]}).
-
-We included items from the following categories: Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-
-We excluded items that enailed more than one word (e.g., \emph{barrita de cereales} {[}cereal bar{]}).
-
-\hypertarget{item-properties}{%
-\subsubsection{Item properties}\label{item-properties}}
-
-\hypertarget{frequencies}{%
-\paragraph{Frequencies}\label{frequencies}}
-
-We retrieved frequencies from SUBTLEX-ESP ({\textbf{???}}) and SUBTLEX-CAT ({\textbf{???}}) for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word ({\textbf{???}}; {\textbf{???}}).
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/pool_freq_cognateness}
-
-}
-
-\caption{ }\label{fig:unnamed-chunk-1}
-\end{figure}
-
-\hypertarget{cognate-status}{%
-\paragraph{Cognate status}\label{cognate-status}}
-
-\hypertarget{demographics-by-item}{%
-\paragraph{Demographics by item}\label{demographics-by-item}}
-
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-2}Demographics of responses by item.
-Language = item language, LP = language profile of participant, Sex = sex of participant, M Age = mean age of responses, SD Age = mean SD of responses across items, Min Age = minimum age of responses across items, Max Age = maximum age of responses across items, M N = mean number of responses across items, N = total number of responses across items.}
-\centering
-\begin{tabular}[t]{c|c|c|c|c|c|c|c|c}
-\hline
-Language & LP & Sex & M Age & SD Age & Min Age & Max Age & M N & N\\
-\hline
-Catalan & Bilingual & Female & 22.36 & 22.36 & 12.35 & 31.42 & 52.22 & 41620\\
-\hline
-Catalan & Bilingual & Male & 20.77 & 20.77 & 12.37 & 31.91 & 64.06 & 51057\\
-\hline
-Catalan & Monolingual & Female & 20.14 & 20.14 & 12.83 & 30.15 & 68.01 & 54204\\
-\hline
-Catalan & Monolingual & Male & 21.11 & 21.11 & 12.60 & 30.93 & 48.30 & 38495\\
-\hline
-Spanish & Bilingual & Female & 22.33 & 22.33 & 12.35 & 31.42 & 52.76 & 42209\\
-\hline
-Spanish & Bilingual & Male & 20.72 & 20.72 & 12.37 & 31.91 & 64.72 & 51778\\
-\hline
-Spanish & Monolingual & Female & 20.10 & 20.10 & 12.83 & 30.15 & 68.71 & 54969\\
-\hline
-Spanish & Monolingual & Male & 21.06 & 21.06 & 12.60 & 30.93 & 48.80 & 39037\\
-\hline
-\end{tabular}
-\end{table}
-
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
-
-For testing the first hypothesis, we fit a non-linear model using logistic curves to model the probability of aquisition of words across ages.
-Logistic curves are characterised by three parameters: 1) an asymptote (maximum value of the curve in the Y-axis), 2) the steepness (how fast the curve grows), and 3) a mid-point (the point in the X-axis at which the steepness is highest).
-Figure 1 shows the elements of a simulated logarithmic curve:
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_illustration}
-
-}
-
-\caption{Simulated logarithmic curve.
-The curve is characterised by three parameters: Asymptote (red), mid-point (blue), and steepness (green).
-This curve is the result of simulating learning curves of 1000 items.
-We set fixed values for the asymptote at 1.
-Mid-points were randomly generated from a normal distribution truncated at 0 (no negative age values are allowed), with mean 23 (*SD* = 0.5), and steepness values were generated from a normal ditribution truncated at 0 (we assume that the proportion of participants that have acquired each item can only grow), with mean 0.1 (*SD* = 1) The solid black line represents the average proportion of participants who have acquired the item, and the shaded ribbon indicates the standard error of the mean.
-We suggest that learning curves of individual word-forms, as well as the evolution of vocabulary size across ages, follows this trend.
-Our model will estimate the three of to adress our hypotheses.}\label{fig:logcurve}
-\end{figure}
-
-One reason for using logistic curves to fit the data is that previous studies describe lexical development as a non-linear process, where between 20 and 24 months a vocabulary spurt happens.
-An analysis of the available data in Wordbank ({\textbf{???}}) provides moderate support for this hypothesis.
-We fitted a logistic model using the proportion of infants that acquired each item as output, and age as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
-We included data from 13119 items from 20 languages\footnote{Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
-All the other available languages were excluded, as no (or insufficient) data were available.}, from infants between 16 and 30 months of age.
-Then we compared the fit of the logarithmic model agains a linear model, and found that the formr fitted the data slightly better (see Figure 2).
-
-A second reason for using logistic curves is that the parameters that characterise a logistic curve map onto our hypotheses quite easily.
-Out first hypothesis concerns the mid-point parameter, which indicates the point at which the steepness of the curve is maximum.
-In our model, the value of this parameter can be interpreted as the age at which the proportion of participants that have acquired a given item is at maximum rate.
-The mathematical interpretation of this parameter is easily mapped into a definition of age of acquisition.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_wordbank}
-
-}
-
-\caption{Logarithmic model fitted on Wordbank data, compared to a linear model.
-The logarithmic model fits the data slightly better.
-The estimated values of the three parameters involved in the logarithmic model (asymptote, mid-point and steepness) are indicated.}\label{fig:wordbank}
-\end{figure}
-
-We built a Bayesian logistic model that estimates the value of the mid-point and steepness across items.
-Under our hypothesis, mid-points are generated from a linear model that incorporates language profile (monolingual vs.~bilingual) and cognateness (non-cognate vs.~cognate) as predictors, where cognateness predictes ealier mid-points for bilingual, but not monolingual participants.
-Due to computational limitations, we set a fixed value for the asymptote at 0.75, according to our model fit on Wordbank data.
-Due to computational limitations, we did not run the same linear model to estimate steepness.
-Rather, this parameter was just included as an intercept.
-We compared how well this model fits the data \emph{versus} how a null model that does not include \emph{cognateness} as predictor does, by using Bayes factors and Pareto-smoothed importance sampling (PSIS; {\textbf{???}}).
-
-We used the information provided by Wordbank as priors to feed our Bayesian model.
-For mid-points, we specified a strong prior with a normal prior with mean = 23.08 and \emph{SD} = 1.
-For steepness, we also set a strong prior with a normal distribution with mean = 3.28 and SD = 2.
-We had little prior information about how cognateness impacts the mid-point of an item, and therefore we will set a weak prior.
-
-We fit both null and alternative models within the R environment ({\textbf{???}}; {\textbf{???}}) using the \texttt{brms} package ({\textbf{???}}), which relies on the probabilistic language Stan ({\textbf{???}}).
-Data and model results will be processed and visualised using the R packages the \texttt{tidyverse} family of packages ({\textbf{???}}) and the \texttt{tidybayes} package ({\textbf{???}}).
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{model-selection}{%
-\subsection{Model selection}\label{model-selection}}
-
-\hypertarget{posterior-distibutions}{%
-\subsection{Posterior distibutions}\label{posterior-distibutions}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-coefficients}
-
-}
-
-\caption{ }\label{fig:posteriors}
-\end{figure}
-
-\hypertarget{posterior-correlations}{%
-\subsection{Posterior correlations}\label{posterior-correlations}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-correlations}
-
-}
-
-\caption{ }\label{fig:correlations}
-\end{figure}
-
-\hypertarget{prior-predictive-checks}{%
-\subsection{Prior predictive checks}\label{prior-predictive-checks}}
-
-Following ({\textbf{???}}), we performed prior predictive chackes to disgnose the appropriateness of our priors.
-
-\hypertarget{posterior-predictive-checks}{%
-\subsection{Posterior predictive checks}\label{posterior-predictive-checks}}
-
-\hypertarget{model-diagnosis}{%
-\subsection{Model diagnosis}\label{model-diagnosis}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{/Users/GonzaloGGC/projects/BiLexicon/Figures/05_gca-convergence}
-
-}
-
-\caption{ }\label{fig:convergence}
-\end{figure}
-
-\hypertarget{auto-correlation}{%
-\subsubsection{Auto-correlation}\label{auto-correlation}}
-
-\hypertarget{potential-scale-reduction-factor-hatr}{%
-\subsubsection{\texorpdfstring{Potential scale reduction factor (\(\hat{R}\))}{Potential scale reduction factor (\textbackslash{}hat\{R\})}}\label{potential-scale-reduction-factor-hatr}}
-
-\hypertarget{posterior-predictive-checks-1}{%
-\subsection{Posterior predictive checks}\label{posterior-predictive-checks-1}}
-
-Does our model generate similar observations to the ones in our data-set?
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-\begin{itemize}
-\tightlist
-\item
-  Cross-sectional data for longitudinal claims
-\item
-  Data is right censored\\
-\item
-  Inventory:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Phonological forms?
-  \item
-    We can't say how frequently words are produced by the parents or the toddler.
-  \item
-    Although instructed to ignore imitations, it is difficult to say whether a toddler has \emph{really} acquired a word that produces, or she can only imitate it.
-  \item
-    Responses in the questionnaire rely heavily on parental memory.
-  \item
-    We don't know about the context at which words are heard or produced.
-  \item
-    Classification system of items (e.g., \emph{household items}) is adult centric. Children may use the words \emph{pretty} to name jewellery (Bates et al., 1994).
-  \end{itemize}
-\item
-  Mid and mid-upper class families are overrepresented in the sample.
-\end{itemize}
-
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\hypertarget{appendix-1-session-info}{%
-\subsection{Appendix 1: Session info}\label{appendix-1-session-info}}
-
-R version 3.6.3 (2020-02-29)
-Platform: x86\_64-apple-darwin15.6.0 (64-bit)
-Running under: macOS Catalina 10.15.4
-
-Matrix products: default
-BLAS: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRblas.0.dylib
-LAPACK: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRlapack.dylib
-
-locale:
-{[}1{]} en\_US.UTF-8/en\_US.UTF-8/en\_US.UTF-8/C/en\_US.UTF-8/en\_US.UTF-8
-
-attached base packages:
-{[}1{]} stats graphics grDevices utils datasets methods base
-
-other attached packages:
-{[}1{]} rlang\_0.4.6 here\_0.1 data.table\_1.12.8\\
-{[}4{]} truncnorm\_1.0-8 wesanderson\_0.3.6 patchwork\_1.0.0.9000
-{[}7{]} ggridges\_0.5.2 ggplot2\_3.3.1 lubridate\_1.7.8\\
-{[}10{]} tidyr\_1.0.3 readxl\_1.3.1 stringr\_1.4.0\\
-{[}13{]} forcats\_0.5.0 dplyr\_1.0.0 tibble\_3.0.1\\
-{[}16{]} magrittr\_1.5 knitr\_1.28 papaja\_0.1.0.9942
-
-loaded via a namespace (and not attached):
-{[}1{]} Rcpp\_1.0.4.6 pillar\_1.4.4 compiler\_3.6.3 cellranger\_1.1.0
-{[}5{]} plyr\_1.8.6 tools\_3.6.3 digest\_0.6.25 evaluate\_0.14\\
-{[}9{]} lifecycle\_0.2.0 gtable\_0.3.0 pkgconfig\_2.0.3 yaml\_2.2.1\\
-{[}13{]} xfun\_0.14 withr\_2.2.0 generics\_0.0.2 vctrs\_0.3.0\\
-{[}17{]} rprojroot\_1.3-2 grid\_3.6.3 tidyselect\_1.1.0 glue\_1.4.1\\
-{[}21{]} R6\_2.4.1 rmarkdown\_2.1 bookdown\_0.18 purrr\_0.3.4\\
-{[}25{]} backports\_1.1.7 scales\_1.1.1 ellipsis\_0.3.1 htmltools\_0.4.0
-{[}29{]} colorspace\_1.4-1 stringi\_1.4.6 munsell\_0.5.0 crayon\_1.3.4
-
-\hypertarget{appendix-2-model}{%
-\subsection{Appendix 2: Model}\label{appendix-2-model}}
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-bates1994}{}%
-Bates, E., Marchman, V., Thal, D., Fenson, L., Dale, P., Reznick, J.
-S., \ldots{} Hartung, J.
-(1994).
-Developmental and stylistic variation in the composition of early vocabulary.
-\emph{Journal of Child Language}, \emph{21}(1), 85--123.
-\url{https://doi.org/10.1017/S0305000900008680}
-
-\leavevmode\hypertarget{ref-bosch2014}{}%
-Bosch, L., \& Ramon-Casas, M.
-(2014).
-First translation equivalents in bilingual toddlers' expressive vocabulary: Does form similarity matter?
-\emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322.
-\url{https://doi.org/10.1177/0165025414532559}
-
-\leavevmode\hypertarget{ref-floccia2018}{}%
-Floccia, C., Sambrook, T.
-D., Luche, C.
-D., Kwok, R., Goslin, J., White, L., \ldots{} Plunkett, K.
-(2018).
-I: Introduction.
-\emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29.
-\url{https://doi.org/10.1111/mono.12348}
-
-\endgroup
-
-\end{document}
+\textbf{Figure caption}: ``Logarithmic model fitted on Wordbank data,
+compared to a linear model. The logarithmic model fits the data slightly
+better. The estimated values of the three parameters involved in the
+logarithmic model (asymptote, mid-point and steepness) are indicated.''
+\end{appendix}

### 2020-06-30 Code/R/01_lang-profiles.R

@@ -5,16 +5,12 @@
 #### set up ##############################################
 
 # load packages
-library(tibble)        # for nice data frames
+library(tidyverse)   
 library(googlesheets4) # for import Google spreadsheets
 library(magrittr)      # for using pipes
 library(data.table)    # for importing data
-library(dplyr)         # for manipulating data
-library(stringr)       # for working with character strings
 library(lubridate)     # for working with dates
-library(tidyr)         # for reshaping datasets
 library(readxl)        # for importing Excel spreadsheets
-library(ggplot2)       # for visualising data
 library(ggridges)      # for plotting densities
 library(ggalluvial)
 library(patchwork)     # for arranging plots
@@ -25,7 +21,8 @@ source(here("Code", "R", "functions.R"))
 
 # set params
 bins <- c("< 10", "10-12", "12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "36-38", "38-40", "> 40")
-bins_interest <- c("18-20", "20-22", "22-24", "24-26", "26-28", "28-30")
+bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+breaks <- c(11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33)
 
 ### import data #########################################
 pool <- read_xlsx(here("Data","01_pool.xlsx"))
@@ -44,7 +41,7 @@ dat <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactors = FA
     as_tibble() %>%
     mutate_at(vars(time_stamp), as_date) %>%
     mutate_at(vars(id_db), as.character) %>%
-    left_join(logs, by = c("id_db", "study", "time_stamp", "age", "sex", "dominance", "code", "version", "lp", "age_bin")) %>%
+    left_join(logs) %>%
     mutate(version = factor(version, levels = c("CBC", "BL-Short-A", "BL-Short-B", "BL-Short-C", "BL-Short-D", "BL-Long-1", "BL-Long-2", "DevLex"), ordered = TRUE)) %>%
     left_join(studies, by = c("version" = "q_version", "language")) %>%
     select(-c(id_exp, code)) %>%
@@ -64,8 +61,8 @@ dat <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactors = FA
[…]

### 2020-06-30 Code/R/07_analysis-prod.R

@@ -1,153 +1,0 @@
-#### 05_analysis-prod: Bayesian non-linear model ##############
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up ##############################################
-
-# load packages
-library(dplyr)
-library(tidyr)
-library(stringr)
-library(data.table)
-library(truncnorm)
-library(brms)
-library(modelr)
-library(janitor)
-library(tidybayes)
-library(ggplot2)
-library(ggmcmc)
-library(mcmcplots) 
-library(wesanderson)
-library(here)
-
-# create/load functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-set.seed(888)
-bins <- c("18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32")
-
-#### import data ##########################################
-dat <- fread(here("Data", "04_prepared.csv")) %>%
-    as_tibble() %>% 
-    arrange(meaning, age_bin, lp) %>%
-    select(meaning, type, age_bin, lp, successes, n, cognate, frequency) %>%
-    filter(type=="Productive")
-
-dat_priors <- fread(here("Data", "05_priors.csv")) %>%
-    as_tibble()
-
-#### priors ###############################################
-prior_asym  <- dat_priors$estimate[dat_priors$term=="asym"] %>% inv_logit_scaled()
-prior_mid   <- dat_priors$estimate[dat_priors$term=="mid"] 
-prior_steep <- dat_priors$estimate[dat_priors$term=="steep"] 
-
-#### fit models ###########################################
-
-# model 0
-formula0 <- bf(successes | trials(n) ~ 0.6789804 * inv(1 + exp((mid - age_bin) * exp(steep))),
-               steep ~ 1, mid ~ 1 + lp + frequency,
-               nl = TRUE, family = binomial(link = "logit"))
-
-priors0 <- c(prior(normal(6, 1), class = "b", nlpar = "mid", coef = "Intercept"),
-             prior(normal(3.28, 1.5),class = "b", nlpar = "steep", coef = "Intercept"),
-             prior(normal(0, 5), class = "b", nlpar = "mid", coef = "lp"))
-
-fit0 <- brm(formula = formula0,
-            data = dat,
-            prior = priors0,
-            save_model = here("Code", "Stan", "prod_fit0.stan"),
[…]

### 2020-06-30 Code/R/07_gca-prod.R

@@ -1,0 +1,197 @@
+#### 05_analysis-prod: Bayesian non-linear model ##############
+# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
+# Center for Brain and Cognition, Universitat Pompeu Fabra
+
+#### set up ##############################################
+
+# load packages
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(data.table)
+library(truncnorm)
+library(brms)
+library(modelr)
+library(janitor)
+library(tidybayes)
+library(ggplot2)
+library(ggmcmc)
+library(mcmcplots) 
+library(here)
+
+# create/load functions
+source(here("Code", "R", "functions.R"))
+
+# set params
+set.seed(888)
+bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+options(contrasts = c("contr.sum", "contr.poly"))
+
+#### import data ##########################################
+dat <- fread(here("Data", "04_prepared.csv")) %>%
+    as_tibble() %>% 
+    filter(type=="Productive",
+           lp=="Bilingual") %>%
+    select(item, meaning, age_bin, item_dominance, cognate, proportion, frequency) %>%
+    mutate(age_bin = as.numeric(factor(age_bin, levels = bins, ordered = TRUE)),
+           item_dominance = factor(item_dominance, levels = c("L2", "L1")),
+           frequency = scale(frequency)[,1],
+           cognate = factor(cognate, levels = c("Non-cognate", "Cognate"))) %>%
+    arrange(item, meaning, age_bin)
+
+dat_priors <- fread(here("Data", "05_priors-wordbank.csv"))
+
+#### fit models ###########################################
+
+# load in case models are already fitted
+fit0 <- readRDS(here("Results", "prod-L1_fit0.rds"))
+fit1 <- readRDS(here("Results", "prod-L1_fit1.rds"))
+
+# model 0
+fit0 <- brm(formula = bf(proportion ~ inv_logit(asym) * inv(1 + exp((mid - age_bin) * exp(steep))),
+                         asym ~ 1,
+                         mid ~ 1 + item_dominance + frequency + (1 | meaning),
+                         steep ~ 1, 
+                         phi ~ 1,
+                         nl = TRUE,
+                         family = zero_one_inflated_beta),
[…]

### 2020-06-30 Code/R/08_survival-comp.R

@@ -1,0 +1,105 @@
+#### 07_survival-comp: Survival analysis on comprehensive data ####
+# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
+# Center for Brain and Cognition, Universitat Pompeu Fabra
+
+#### set up #######################################################
+
+# load packages
+library(tidyverse)
+library(data.table)
+library(tidybayes)
+library(BayesFactor)
+library(ggridges)
+library(patchwork)
+library(see)
+library(here)
+
+# load functions
+source(here("Code", "R", "functions.R"))
+
+# set params
+set.seed(888)
+bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+
+#### import data ###################################################
+fit1 <- readRDS(here("Results", "comp_fit1.rds"))
+dat <- fread(here("Data", "04_prepared.csv")) %>%
+    as_tibble() %>% 
+    filter(type=="Comprehensive",
+           lp=="Bilingual") %>%
+    select(item, meaning, age_bin, item_dominance, cognate, proportion, frequency) %>%
+    mutate(age_bin = as.numeric(factor(age_bin, levels = bins, ordered = TRUE)),
+           item_dominance = factor(item_dominance, levels = c("L2", "L1")),
+           frequency = scale(frequency)[,1],
+           cognate = factor(cognate, levels = c("Non-cognate", "Cognate"))) %>%
+    arrange(item, meaning, age_bin) 
+
+#### extract midpoints ##############################################
+midpoints <- fit1 %>%
+    spread_draws(b_mid_Intercept, b_mid_item_dominance1, b_mid_frequency, r_meaning__mid[meaning, term]) %>% 
+    mutate(mid_L1 = b_mid_Intercept + r_meaning__mid + 0.5*b_mid_item_dominance1 + b_mid_frequency,
+           mid_L2 = b_mid_Intercept + r_meaning__mid + -0.5*b_mid_item_dominance1 + b_mid_frequency) %>% 
+    median_qi(mid_L1, mid_L2) %>%
+    select(-c(term, .interval, .point, .width)) %>%
+    ungroup() %>% 
+    right_join(dat, by = "meaning") %>%
+    rowwise() %>% 
+    mutate(mid_diff = mid_L2 - mid_L1) %>% 
+    ungroup() %>% 
[…]

### 2020-07-03 Code/R/07_gca-prod.R
Final revision of poster, second attempt on productive data

@@ -12,6 +12,8 @@ library(brms)
 library(modelr)
 library(janitor)
 library(tidybayes)
+library(ggmcmc)
+library(mcmcplots) 
 library(here)
 
 # create/load functions
@@ -19,7 +21,7 @@ source(here("Code", "R", "functions.R"))
 
 # set params
 set.seed(888)
-bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
 options(contrasts = c("contr.sum", "contr.poly"))
 
 #### import data ##########################################
@@ -28,7 +30,7 @@ dat <- fread(here("Data", "04_prepared.csv")) %>%
     filter(type=="Productive",
            lp=="Bilingual") %>%
     select(item, meaning, age_bin, item_dominance, cognate, proportion, frequency) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins, ordered = TRUE)),
+    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE)),
            item_dominance = as.factor(item_dominance),
            frequency = scale(frequency)[,1],
            cognate = as.factor(cognate)) %>%
@@ -43,7 +45,7 @@ dat_priors <- fread(here("Data", "05_priors-wordbank.csv")) %>%
 #### fit models ###########################################
 
 # load in case models are already fitted
-fit_prior <- readRDS(here("Results", "prod_fit-prior.rds"))
+fit_prior <- readRDS(here("Results", "comp_fit-prior.rds"))
 fit0 <- readRDS(here("Results", "prod_fit0.rds"))
 fit1 <- readRDS(here("Results", "prod_fit1.rds"))
 
@@ -56,12 +58,13 @@ fit0 <- brm(formula = bf(proportion ~ inv_logit(asym) * inv(1 + exp((mid - age_b
                          nl = TRUE,
                          family = zero_one_inflated_beta),
             prior = c(prior(normal(0.7857192, 0.1), nlpar = "asym", coef = "Intercept"),
-                      prior(normal(4.497546, 1), nlpar = "mid", coef = "Intercept"),
+                      prior(normal(5.669435, 1), nlpar = "mid", coef = "Intercept"),
                       prior(normal(1.7576520, 0.8), nlpar = "steep", coef = "Intercept"),
[…]

### 2020-07-03 Manuscript/trajectories_manuscript.Rmd
Final revision of poster, second attempt on productive data

@@
 library(tidyverse)
-library(dplyr)
-library(forcats)
-library(stringr)
 library(readxl)
@@
 
+library(broom)
 library(ggridges)
@@
 # set params
-bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
+bins <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
 
@@
   mutate_at(vars(date_birth, time_stamp), as_date)
-
 dat_wordbank <- fread(here("Data", "05_wordbank.csv"))
 dat_wordbank_priors <- fread(here("Data", "05_priors-wordbank.csv"))
+
+# import results
+comp_fit1 <- readRDS(here("Results", "comp_fit1.rds"))
+comp_selection <- readRDS(here("Results", "comp_selection.rds"))
 
@@
 
-## Hypothesis 1
+## Hypothesis 1: Are cognates acquired ealier?
 
 ### Comprehensive vocabulary
+
+Table \@ref(tab:resultscomp) shows the outcomes of the model.
+
+```{r resultscomp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+tidy(comp_fit1) %>% 
+    gt()
+
+```
 

### 2020-07-03 Code/R/00_logs.R

@@ -41,7 +41,7 @@ dat <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
     
 #### visualise dates #######################
 samples <- dat %>% 
-    filter(completed, lp %!in% "Other", age_bin %in% bins_interest) %>% 
+    filter(completed, lp %in% "Bilingual", age_bin %in% bins_interest) %>% 
     group_by(lp, dominance, age_bin) %>%
     summarise(n = n(), .groups = "drop") %>%
     group_by(age_bin, lp) %>%
@@ -54,10 +54,10 @@ dat %>%
     labs(x = "Age (months)", y = "Number of partcipants",
          fill = "Language of highest exposure") +
     scale_y_continuous(labels = as.integer) +
-    scale_fill_brewer(palette = "Set1") +
+    scale_fill_manual(values = c("#44546A", "orange")) +
     theme_custom +
     theme(legend.position = "top",
           panel.background = element_rect(fill = "white"),
           axis.text = element_text(colour = "black"),
           panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")) +
-    ggsave(here("Figures", "01_logs-distribution.png"))
+    ggsave(here("Figures", "01_logs-distribution.png"), height = 4, width = 5)

### 2020-07-07 Code/Rmd/model.Rmd
Add SI about the model

---
title: "Model and priors"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Our model specifies a logistic curve that fits our outcome measure. Out outcome measure is the porportion of participants that are reported to understand each word. This logistic curve is characterised by three parameters: *asymptote* ($asym$, the value at which the outcome *pleateaus*), *steepness*, ($steep$, how fast the curve grows), and *mid-point* ($mid-point$, the age at which the curve reaches its maximum steepness). We identify the *mid-point* of a word as its age of acquisition, and thus us the parameter of interest. Out hypothesis is that the language a word belongs to (non-dominant vs. dominant), together with its cognate status (non-cognate vs. cognate) and frequency, affects the position of its mid-point. We assume that the mid-point is generated from a zero-one beta regression model that includes the aforementioned predictors. A beta regression model helps us model data that include many values at boundary, as ours are. We also included random intercepts for each *translation equivalent* ($TE$). We set informed priors for *asymptotes*, *steepnesses*, and *mid-points* according to Wordbank data, and weak priors the the coefficients of the beta-regression model.

$$
\begin{aligned}proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\ mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ \beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\ \phi \sim Normal(1.5, 1)
\end{aligned}
$$

### 2020-07-14 Code/Rmd/model.Rmd
Avoid uniform priors for distributional parameters for faster and more reliable convergence.

@@
 $$
-\begin{aligned}proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\ mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ \beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\ \phi \sim Normal(1.5, 1)
+\begin{aligned}proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\ mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ TE_{0i} \sim Normal(0, \sigma_i) \\\beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\ \phi \sim Normal(1.5, 1)\\ \sigma_i \sim HalfCauchy(5)
 \end{aligned}

### 2020-07-14 Code/R/06_gca-comp.R

@@ -11,7 +11,6 @@ library(truncnorm)
 library(brms)
 library(modelr)
 library(janitor)
-library(rayshader)
 library(tidybayes)
 library(here)
 
@@ -26,16 +25,14 @@ options(contrasts = c("contr.sum", "contr.poly"))
 #### import data ##########################################
 dat <- fread(here("Data", "04_prepared.csv")) %>%
     as_tibble() %>% 
-    filter(type=="Comprehensive") %>%
-    select(item, meaning, age_bin, bilingualism, item_dominance, cognate, proportion, frequency) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE)),
-           item_dominance = as.factor(item_dominance),
+    filter(type=="Comprehensive",
+           item_dominance=="L2") %>%
+    select(item, meaning, category, age_bin, bilingualism, cognate, proportion) %>%
+    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1,
            bilingualism = scale(bilingualism)[,1],
-           frequency = scale(frequency)[,1],
            cognate = as.factor(cognate)) %>%
     arrange(item, meaning, age_bin)
 
-contrasts(dat$item_dominance) <- contr.sum(c("L2", "L1"))/2
 contrasts(dat$cognate) <- contr.sum(c("Non-cognate", "Cognate"))/2
 
 dat_priors <- fread(here("Data", "05_priors-wordbank.csv")) %>%
@@ -46,105 +43,78 @@ dat_priors <- fread(here("Data", "05_priors-wordbank.csv")) %>%
 # load in case models are already fitted
 fit_prior <- readRDS(here("Results", "comp_fit-prior.rds"))
 
+# set priors 
+priors <- c(prior(normal(0.7857192, 0.1), nlpar = "asym", coef = "Intercept"),
+            prior(normal(5.369435, 1), nlpar = "mid", coef = "Intercept"),
+            prior(normal(1.7576520, 0.8), nlpar = "steep", coef = "Intercept"),
+            prior(normal(0, 5), class = "b", nlpar = "mid"),
+            prior(normal(1.5, 1), dpar = "phi", class = "Intercept"),
+            prior(beta(10, 10), class = "coi"),
+            prior(beta(10, 10), class = "zoi"),
+            prior(HalfCauchy(0, 5), class  = "sd"))
 # model 0
[…]

### 2020-07-14 Code/R/functions.R

@@ -47,3 +47,20 @@ theme_custom <- theme(panel.background = element_rect(fill = "transparent"),
                       legend.background = element_rect(fill = "transparent"),
                       legend.key = element_rect(colour = "transparent", fill = "transparent")
 )
+
+pred_labels <- function(x){
+  require(dplyr)
+  case_when({{ x }} %in% "b_asym_Intercept" ~ "Asymptote (Intercept)",
+            {{ x }} %in% "b_steep_Intercept" ~ "Steepness (Intercept)",
+            {{ x }} %in% "b_mid_Intercept" ~ "Mid-point (Intercept)",
+            {{ x }} %in% "b_mid_bilingualism" ~ "Bilingualism (Slope)",
+            {{ x }} %in% "b_mid_cognate1" ~ "Cognateness (Slope)",
+            {{ x }} %in% "b_mid_bilingualism:cognate1" ~ "Bilingualism \U000D7 Cognateness",
+            {{ x }} %in% "sd_meaning__mid_Intercept" ~ "SD TE (Intercept)",
+            {{ x }} %in% "sd_meaning__mid_cognate1" ~ "SD TE-Cognateness (Slope)",
+            
+            {{ x }} %in% "phi_Intercept" ~ "\U03C6 (Intercept)",
+            {{ x }} %in% "b_phi_Intercept" ~ "\U03C6 (Slope)",
+            {{ x }} %in% "zoi" ~ "ZOI",
+            {{ x }} %in% "coi" ~ "COI")
+}

### 2020-08-11 Manuscript/trajectories_manuscript.Rmd
De-scale aysmptote parameter

@@
 library(magrittr)
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(tibble)
+library(ggplot2)
+library(patchwork)
 library(readxl)
@@
 library(ggridges)
+library(brms)
+library(modelr)
+library(tidybayes)
 library(patchwork)
@@
 dat_studies <- fread(here("Data", "00_studies.csv")) %>% distinct(q_version, language, q_items)
+
 dat_items <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
@@
     mutate_at(vars(A, B, C, D, cognate_rater1, include), as.logical)
-dat_participants <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
-  last() %>%
-  fread(na.strings = "") %>%
-  filter(age_bin %in% bins,
-         completed,
-         lp == "Bilingual")
+
+dat_participants <- fread(here("Data", "01_participants.csv"), na.strings = "")
 dat_raw <- fread(here("Data", "02_merged.csv"), na.strings = "") %>%
@@
     mutate_at(vars(date_birth, time_stamp), as_date)
+dat <- fread(here("Data", "preprocessed.csv"), na.strings = "", stringsAsFactors = TRUE)
 dat_wordbank <- fread(here("Data", "05_wordbank.csv"))
@@
 # import results
+comp_fit0 <- readRDS(here("Results", "comp_fit0.rds"))
 comp_fit1 <- readRDS(here("Results", "comp_fit1.rds"))
-comp_selection <- readRDS(here("Results", "comp_selection.rds"))
+comp_fit2 <- readRDS(here("Results", "comp_fit2.rds"))
+comp_selection <- readRDS(here("Results", "comp_loo.rds"))
+
+random <- readRDS(here("Results", "ranef.rds"))
+corr <- readRDS(here("Results", "corr.rds"))
 
@@
 
-dat_participants %>%
-  filter(completed, lp %!in% "Other", age_bin %in% bins) %>%
-  ggplot(aes(x = age_bin, fill = dominance)) +
-  geom_bar() +
-  labs(x = "Age (months)", y = "Number of partcipants",
-       fill = "Dominant language") +
-  scale_y_continuous(labels = as.integer) +
-  scale_fill_brewer(palette = "Set1") +
-  theme_custom +
-  theme(legend.position = c(0.05, 0.85),
-        legend.direction = "vertical",
-        legend.justification = 0,
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))
+participants %>%
+    mutate(bilingualism = cut_interval(bilingualism, n = 6,
+                                       labels = paste0(seq(0, 50, by = 10), "%"))) %>%
+    ggplot(aes(x = age_bin, y = bilingualism)) +
+    stat_bin_2d(colour = "white") +
+    scale_fill_viridis_c(option = "plasma") +
+    labs(x = "Age (months)", y = "L2 exposure", fill = "Participants") +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 270),
+          axis.text = element_text(colour = "black"))
+ggsave(here("Figures", "01_logs-distribution.png"), height = 4, width = 5)
 ```
@@
 
-For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item  within age bins of 2 months.
-Our model assumes that this proportion follows a logistic curve accross ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the dominance of the language of the item (*item dominance*, L1 vs.
-L2), the cognate status of the word (*cognateness*, Non-cognate vs.
-Cognate), and the frequency of the word form in its correspondent language.
-Our prediction  is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words.
+For testing Hypothesis 1, we modeled the proportion of infants that understood or produced each item in the non-dominant language within age bins of 2 months.
+Our model assumes that this proportion follows a logistic curve across ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the degree of exposure to the non-dominant language (*bilingualism*, 0-50%), the cognate status of the word (*cognateness*, Non-cognate vs.
+Cognate), and their interaction.
+Our prediction is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words, and that this effect will be larger for more balanced bilinguals.
 
 Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
-Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item porta (door, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
-We retrieved frequencies from SUBTLEX-ESP [@cuetos2011] and SUBTLEX-CAT [@boada2019] for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word [@zipf1949; @heuven2014].
-Missing frequency scores were imputed using multiple imputation [@vanbuuren2011].
+Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item *porta* (*door*, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
 Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
@@
 This approach allows us to model and estimate the age of acquisition of items explicitely.
-Our model estimated the asymptote and stepness as grand means across items (i.e., intercepts).
-Mid-points were estimated by a linear zero-one-beta-regression model^[Zero-or-one-inflated beta regression models allow to account for overdispersion then using proportions as the output of a model accounting, while also allowing values to be at boundary (0 and 1).][@ospina2012] that included our predictors of interest: `age`, `dominance` (contrast-coded, L2 = -0.5, L2 = 0.5), `cognateness` (contrast-coded, Non-cognate = -0.5, Cognate = 0.5), their interaction, and `frequency` (standardised Zipf score).
-We also included a random intercept for each translation equivalent (`TE`).
+Our model estimated the asymptote and steepness as grand means across items (i.e., intercepts).
+Mid-points were estimated by a linear regression model that included our predictors of interest: `age`, `bilingualism` (standardised), `cognateness` (contrast-coded, Non-cognate = -0.5, Cognate = 0.5).
+We also included random intercepts and slopes for each translation equivalent (`TE`) [@barr2013].
 Appendix 1 shows a detailed description of the model.
@@
 We estimated the parameters of our model using the Bayesian framework.
-Bayesian models use the Bayes rule to combine prior knowlegde about the distribution of a given parameter, and the likelihood of each of the values it can take given tha observed data.
+Bayesian models use the Bayes rule to combine prior knowledge about the distribution of a given parameter, and the likelihood of each of the values it can take given the data.
 We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our priors).
-In order to test the contribution of the predictor *cognateness* on the fit of the model,  we compared the fit of our (expanded) model to the fit of a (null) model that did not include *cognateness* as predictor [@schad2020] using leave-one-out cross validation.
+In order to test the contribution of each predictor on the fit of the model, we started fitting a null model that only included an global intercept and random intercepts for each TE, then fitted a model that also included *bilingualism* and its random slopes by TE, and finally a model that also included *cognateness* and its interaction with *bilingualism*, as well as random slopes for all the fixed effects.
+We compared how the fit of the model changed in every step [@schad2020] using leave-one-out cross validation.
 We fit both the null and alternative models using the R environment [@rcoreteam2013; @rstudioteam2015] and the `brms` package [@burkner2017], which relies on the probabilistic language Stan [@carpenter2017] to approximate posterior distributions.
@@
 
-### Comprehensive vocabulary
+```{r post}
+posterior <- comp_fit2 %>%
+    recover_types(dat) %>% 
+    gather_draws(`b_.*`, regex = TRUE) %>% 
+    mutate_at(vars(.variable, .chain), as.factor) %>%
+    filter(!str_detect(.variable, "Intercept"))
+
+ggplot(posterior, aes(.value, .variable)) +
+    stat_slab(fill = "black") +
+    stat_interval(.width = c(0.95, 0.89, 0.67, 0.5),
+                  point_interval = "median_hdi",
+                  position = position_nudge(y = -0.1),) +
+    geom_vline(xintercept = 0, linetype = "dashed") +
+    labs(x = "Value", y = "Coefficient", colour = "HDI", fill = "Type") +
+    scale_colour_brewer() +
+    theme_classic() +
+    theme(legend.position = "top",
+          panel.grid.major.y = element_line(colour = "grey"),
+          axis.text = element_text(colour = "black")) +
+    ggsave(here("Figures", "06_gca_comp-coefs.png"), width = 6)
+
+```
+
+```{r post_checks}
+posterior_check <- expand_grid(n = 1,
+                               age_bin = seq_range(dat$age_bin, n = 50),
+                               item_dominance = c("L2", "L1"),
+                               cognate = c("Non-cognate", "Cognate")) %>%
+    add_fitted_draws(model = comp_fit2, newdata = ., n = 100, scale = "response", re_formula = NA) %>% 
+    ungroup() %>% 
+    mutate(cognate = factor(cognate, levels = c("Non-cognate", "Cognate"), ordered = TRUE),
+           .draw = as.character(.draw))
+
+ggplot(posterior_check, aes(age_bin, .value, colour = cognate, fill = cognate)) +
+    facet_grid(~item_dominance) +
+    stat_lineribbon(.width = 89, colour = NA, alpha = 0.5) +
+    stat_summary(fun = "mean", geom = "line") +
+    labs(x = "Age (months)", y = "Proportion",
+         colour = "Cognateness", fill = "Cognateness", shape = "Cognateness") +
+    scale_fill_brewer(palette = "Set1") +
+    scale_colour_brewer(palette = "Set1") +
+    scale_x_continuous(breaks = seq(0, 10), labels = bins_interest) +
+    theme_classic() +
+    theme(legend.position = "top", 
+          axis.text = element_text(colour = "black"),
+          axis.text.x = element_text(angle = 270),
+          legend.title = element_blank(),
+          plot.caption = element_text(hjust = 0),
+          plot.caption.position = "plot",
+          plot.title.position = "plot") +
+    ggsave(here("Figures", "07_gca_comp_posterior-checks.png"))
+
+```
 
@@
 
-tidy(comp_fit1) %>% 
-    gt()
+summary(comp_fit2)$fixed %>% 
+    as.data.frame() %>% 
+    rownames_to_column("term") %>% 
+    select(1:5) %>% 
+    mutate_at(4:5, round, 2) %>% 
+    unite("ci", 4:5, sep = ", ") %>% 
+    gt(rowname_col = "term") %>% 
+    fmt_number(2:3) %>% 
+    cols_label(Estimate = "Estimate",
+               `Est.Error` = "SE",
+               ci = "95% CrI")
 
@@
 
-### Productive vocabulary
+```{r random_effects}
+ggplot(random, aes(r_intercept, item_dominance1)) +
+    geom_point(shape = 1, alpha = 0.5, size = 0.25) +
+    geom_abline(slope = corr$cor_meaning_mid_intercept_mid_item_dominance1, colour = "red") +
+    labs(x = "Intercept", y = "Dominance") +
+
+    ggplot(random, aes(r_intercept, cognate1)) +
+    geom_point(shape = 1, alpha = 0.5, size = 0.25) +
+    geom_abline(slope = corr$cor_meaning_mid_intercept_mid_cognate1, colour = "red") +
+    labs(x = "Intercept", y = "Cognateness slope") +
+
+    ggplot(random, aes(r_intercept, item_dominance1_cognate1)) +
+    geom_point(shape = 1, alpha = 0.5, size = 0.25) +
+    geom_abline(slope = corr$cor_meaning_mid_intercept_mid_item_dominance1_cognate1, colour = "red") +
+    labs(x = "Intercept", y = "Dominance x Cognateness") +
+
+    ggplot(random, aes(item_dominance1, cognate1)) +
+    geom_point(shape = 1, alpha = 0.5, size = 0.25) +
+    geom_abline(slope = corr$sd_meaning_mid_item_dominance1_cognate1, colour = "red") +
+    labs(x = "Dominance", y = "Cognateness") +
+
+    ggplot(random, aes(item_dominance1, item_dominance1_cognate1)) +
+    geom_point(shape = 1, alpha = 0.5, size = 0.25) +
+    geom_abline(slope = corr$sd_meaning_mid_item_dominance1_cognate1, colour = "red") +
+    labs(x = "Dominance", y = "Dominance x Cognateness") +
+
+    ggplot(random, aes(cognate1, item_dominance1_cognate1)) +
+    geom_point(shape = 1, alpha = 0.5, size = 0.25) +
+    geom_abline(slope = corr$sd_meaning_mid_item_dominance1_cognate1, colour = "red") +
+    labs(x = "Cognateness", y = "Dominance x Cognateness") +
+
+    plot_layout(ncol = 3) &
+    plot_annotation(tag_levels = "A") &
+    theme_classic() &
+    ggsave(here("Figures", "random_effects.png"), height = 4, width = 6)
+
+```
 

### 2020-08-11 Code/R/01_logs.R

@@ -5,7 +5,8 @@
 #### set up ################################
 
 # load packages
-library(tidyverse)
+library(dplyr)
+library(tidyr)
 library(lubridate)
 library(patchwork)
 library(data.table)
@@ -30,14 +31,12 @@ dat <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
     mutate_at(vars(age_bin), factor, levels = bins, ordered = TRUE) %>%
     mutate(version = factor(version, levels = c("CBC", "DevLex", "BL-Short-A", "BL-Short-B", "BL-Short-C", "BL-Short-D", "BL-Long-1", "BL-Long-2"), ordered = TRUE),
            doe = ifelse(dominance=="Spanish", doe_spanish, doe_catalan),
-           lp = ifelse(!between(doe, 50, 100, incbounds = TRUE), "Other", lp),
+           lp = ifelse(!data.table::between(doe, 50, 100, incbounds = TRUE), "Other", lp),
            lp = factor(lp, levels = c("Monolingual", "Bilingual"))) %>%
     rowwise() %>% 
-    mutate(bilingualism = ifelse(
-        dominance %in% "Spanish",
-        1-(abs((doe_spanish-doe_catalan)/(doe_spanish+doe_catalan))),
-        1-(abs((doe_catalan-doe_spanish)/(doe_spanish+doe_catalan))))
-    ) %>% 
+    mutate(bilingualism = ifelse(dominance %in% "Spanish",
+                                 1-(abs((doe_spanish-doe_catalan)/(doe_spanish+doe_catalan))),
+                                 1-(abs((doe_catalan-doe_spanish)/(doe_spanish+doe_catalan))))) %>% 
     drop_na(lp) %>% 
     filter(completed, age_bin %in% bins_interest) 
 

### 2020-08-11 Code/R/03_familiarity.R

@@ -5,7 +5,8 @@
 #### set up ##############################################
 
 # load packages
-library(tidyverse)        # for nice data frames
+library(dplyr)        # for nice data frames
+library(tidyr)
 library(magrittr)      # for using pipes
 library(data.table)    # for importing data
 library(lubridate)     # for working with dates
@@ -14,6 +15,7 @@ library(here)          # for locating files
 
 # load/create functions
 source(here("Code", "R", "functions.R"))
+
 
 # set params
 bins          <- c("< 10", "10-12", "12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "36-38", "38-40", "> 40")

### 2020-08-11 Code/R/analysis.R

@@ -1,0 +1,183 @@
+#### analysis ########################
+
+#### set up ##########################
+
+# load packages
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(ggplot2)
+library(data.table)
+library(brms)
+library(tidybayes)
+library(BayesFactor)
+library(janitor)
+library(here)
+# load functions
+
+# set params
+set.seed(888)
+bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+
+#### import data #####################
+
+dat <- fread(here("Data", "preprocessed.csv"), na.strings = "", stringsAsFactors = TRUE)
+
+contrasts(dat$cognate) <- contr.sum(c("Non-cognate", "Cognate"))/2
+contrasts(dat$item_dominance) <- contr.sum(c("L2", "L1"))/2
+
+dat_priors <- fread(here("Data", "05_priors-wordbank.csv")) %>%
+    mutate(estimate_scaled = (estimate_scaled-14)/2)
+
+inv_logit <- function(x) 1 / (1 + exp(-x))
+
+#### comprehension ###################
+# set priors 
+comp_priors <- c(prior(normal(0.7857192, 0.5), nlpar = "asym", coef = "Intercept"),
+                 prior(normal(-1.7576520, 0.5), nlpar = "steep", coef = "Intercept"),
+                 prior(normal(4.369435, 1), nlpar = "mid", coef = "Intercept"),
+                 prior(normal(0, 5), class = "b", nlpar = "mid"),
+                 prior(cauchy(0, 5), class  = "sd", nlpar = "mid"))
+
+# `model 0
+comp_fit0 <- brm(formula = bf(proportion ~ inv_logit(asym) * inv(1 + exp((mid - age_bin) * steep)),
+                              asym ~ 1,
+                              steep ~ 1,
+                              mid ~ 1 + (1 | meaning),
+                              nl = TRUE),
+                 prior = comp_priors,
+                 data = filter(dat, type=="Comprehensive"),
+                 chains = 1,
+                 iter = 2000,
+                 file = here("Results", "comp_fit0.rds"),
+                 save_model = here("Code", "Stan", "comp_fit0.stan"),
+                 control = list(adapt_delta = 0.95, max_treedepth = 15))
+
+# model 1
+comp_fit1 <- brm(formula = bf(proportion ~ inv_logit(asym) * inv(1 + exp((mid - age_bin) * steep)),
[…]

### 2020-08-11 Code/R/preprocessing.R

@@ -1,0 +1,118 @@
+#### preprocessing #########################
+
+#### set up ################################
+
+# load packages
+library(dplyr)
+library(tidyr)
+library(magrittr)
+library(stringr)
+library(purrr)
+library(lubridate)
+library(readxl)
+library(patchwork)
+library(data.table)
+library(mice)
+library(here)
+
+# load functions
+source(here("Code", "R", "functions.R"))
+
+# set params
+set.seed(888)
+bins <- c("< 10", "10-12", "12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "36-38", "38-40", "> 40")
+bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+breaks <- c(0, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 100)
+
+#### participant data ######################
+
+participants <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
+    .[!str_detect(., "summary")] %>%
+    last() %>%
+    fread(na.strings = "") %>%
+    as_tibble() %>%
+    mutate_at(vars(date_sent, time_stamp), as_date) %>%
+    mutate_at(vars(id_db), factor) %>%
+    mutate_at(vars(age_bin), factor, levels = bins, ordered = TRUE) %>%
+    mutate(version = factor(version, levels = c("CBC", "DevLex", "BL-Short-A", "BL-Short-B", "BL-Short-C", "BL-Short-D", "BL-Long-1", "BL-Long-2"), ordered = TRUE),
+           doe = ifelse(dominance=="Spanish", doe_spanish, doe_catalan),
+           lp = ifelse(!data.table::between(doe, 50, 100, incbounds = TRUE), "Other", lp),
+           lp = factor(lp, levels = c("Monolingual", "Bilingual"))) %>%
+    rowwise() %>% 
+    mutate(bilingualism = ifelse(dominance %in% "Spanish",
+                                 1-(abs((doe_spanish-doe_catalan)/(doe_spanish+doe_catalan))),
+                                 1-(abs((doe_catalan-doe_spanish)/(doe_spanish+doe_catalan))))) %>% 
+    drop_na(lp, bilingualism) %>% 
+    filter(completed, age_bin %in% bins_interest) 
+
+# export data
[…]

### 2020-08-11 Code/R/00_pool.R
De-scale aysmptote parameter

@@ -5,7 +5,10 @@
 #### set up #############################################
 
 # load packages
-library(tidyverse)    
+library(dplyr)
+library(tidyr)
+library(purrr)
+library(ggplot2)
 library(data.table)    # for importing and exporting data
 library(readxl)        # for importing Excel files
 library(here)          # for locating files
@@ -37,10 +40,10 @@ dat_freqdiff <- dat %>%
 ggplot(dat_freqdiff, aes(x = language, y = frequency, colour = frequency_diff, alpha = frequency_diff)) +
   facet_wrap(~cognate_rater1) +
   geom_violin(fill = "black", colour = NA, alpha = 0.4, na.rm = TRUE) +
-  geom_line(aes(group = meaning), na.rm = TRUE, size = 0.5) +
+  geom_line(aes(group = meaning), na.rm = TRUE, size = 0.5, alpha = 0.5) +
   geom_boxplot(width = 0.05, na.rm = TRUE, colour = "black") +
-  labs(x = "Language", y = "Frequency (Zipf score)", colour = "Absolute\ndifference") +
+  labs(x = "Language", y = "Frequency (Zipf score)", colour = "Absolute difference") +
   scale_alpha(guide = FALSE) +
   theme_custom +
-  theme(legend.position = "right") +
+  theme(legend.position = "top") +
   ggsave(here("Figures", "00_pool_freq_cognateness.png"))

### 2020-08-11 Code/R/02_vocabulary.R

@@ -5,7 +5,9 @@
 #### set up ##############################################
 
 # load packages
-library(tidyverse)       
+library(dplyr)
+library(tidyr)
+library(ggplot2)
 library(googlesheets4) # for import Google spreadsheets
 library(magrittr)      # for using pipes
 library(data.table)    # for importing data
@@ -25,8 +27,10 @@ breaks <- c(0, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 1
 
 #### import data #########################################
 pool <- read_xlsx(here("Data","01_pool.xlsx"))
-studies <- fread(here("Data", "00_studies.csv")) %>% distinct(q_version, language, q_items) 
-logs <- fread(here("Data", "01_participants.csv"))
+studies <- fread(here("Data", "00_studies.csv")) %>%
+  distinct(q_version, language, q_items)
+logs <- fread(here("Data", "01_participants.csv")) %>% 
+  mutate(time_stamp = as_date(time_stamp))
 
 dat <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactors = FALSE, na.strings = "") %>%
   as_tibble() %>%

### 2020-08-22 Manuscript/trajectories_manuscript.Rmd
Smoother predictions, source analysis from external R in manuscript, and plot random effects

--- 2020-06-30 Manuscript/trajectories_manuscript.Rmd
@@
 authornote: |
-  "Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
+  Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
 
@@
 output            :
-  word_document:
-    reference_docx: trajectories_template.docx
+  papaja::apa6_pdf:
+    latex_engine  : "xelatex"
 
@@
 
-```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r packages, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 library(knitr)
-
 library(magrittr)
-library(tidyverse)
 library(dplyr)
+library(tidyr)
+library(tibble)
+library(ggplot2)
 library(forcats)
-library(stringr)
 library(readxl)
@@
 library(janitor)
-
-library(ggridges)
-library(patchwork)
+library(nlme)
+library(broom.mixed)
+library(modelr)
 library(gt)
-
 library(data.table)
@@
 
-# load/create functions
-source(here("Code", "R", "functions.R"))
+```
+
+```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
 # set params
-bins <- c("14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
+set.seed(888)
+bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
+bins_expanded <- as.character(seq(1, 41, by = 2))
 
 # import data 
-dat_studies <- fread(here("Data", "00_studies.csv")) %>% distinct(q_version, language, q_items)
+dat_studies <- fread(here("Data", "00_studies.csv"), na.strings = c("", "NA")) %>%
+  distinct(q_version, language, q_items)
+
 dat_items <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
@@
     mutate_at(vars(A, B, C, D, cognate_rater1, include), as.logical)
-dat_participants <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
-  last() %>%
-  fread(na.strings = "") %>%
-  filter(age_bin %in% bins,
-         completed,
-         lp == "Bilingual")
-dat_raw <- fread(here("Data", "02_merged.csv"), na.strings = "") %>%
+
+dat_participants <- fread(here("Data", "participants.csv"), na.strings = c("", "NA"))
+
+dat_raw <- fread(here("Data", "raw.csv"), na.strings = c("", "NA")) %>%
     filter(lp %in% c("Bilingual"),
-         age_bin %in% bins) %>%
+           age_bin %in% bins_interest) %>%
     as_tibble() %>%
@@
 
-dat_wordbank <- fread(here("Data", "05_wordbank.csv"))
-dat_wordbank_priors <- fread(here("Data", "05_priors-wordbank.csv"))
+dat <- fread(here("Data", "preprocessed.csv"), na.string = c("", "NA")) %>% 
+    as_tibble() %>% 
+    select(item, type, te, category, age_bin, item_dominance, cognate, frequency, n, successes, proportion) %>%
+    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1) %>% 
+    mutate_at(vars(frequency), function(x) scale(x, center = TRUE, scale = FALSE)[,1]) %>% 
+    mutate_at(vars(cognate, item_dominance), as.factor) %>% 
+    arrange(type, item, te, age_bin)
+
+```
+
+```{r analyses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, cache=TRUE}
+
+# run the analyses (see analysis.R to check to code)
+# running this code chunk may take a few seconds
+source(here("Code", "R", "analysis.R"), local = knitr::knit_global())
 
@@
 @blom2019 extended these findings to children aged three to 10 years.
-These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
+These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical acquisition.
 
@@
 A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model [@kroll1994; @kroll2010], the Bilingual Interactive Activation model [BIA; @dijkstra1998bia], and its successor BIA+ [@dijkstra2002architecture], the Bilingual Language Interaction Network for Comprehension of Speech model [BLINCS; @shook2013] and, more recently, Multilink [@dijkstra2019], which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
-However, these models do not address word learning during language acquistion.
+However, these models do not address word learning during language acquisition.
 
@@
 This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
-These scores were aggregated across pais of tanslation equivalents, obtaining a single similarity estimate of for each language pair.
+These scores were aggregated across pairs of translation equivalents, obtaining a single similarity estimate of for each language pair.
 This measure embeds confounded information about the "degree of cognateness" and the amount of phonological overlap between both languages.
-Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, Enlgish-Dutch), with few exceptions (e.g., Spanish-Basque).
-In the light of these data, it is not possible to discern whether vocabulary size is increased by the ealier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
-We address this issue by testing two hypotesis.
+Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, English-Dutch), with few exceptions (e.g., Spanish-Basque).
+In the light of these data, it is not possible to discern whether vocabulary size is increased by the earlier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
+We address this issue by testing two hypothesis.
 
@@
 dat_participants %>%
-  filter(completed, lp %!in% "Other", age_bin %in% bins) %>%
-  ggplot(aes(x = age_bin, fill = dominance)) +
-  geom_bar() +
-  labs(x = "Age (months)", y = "Number of partcipants",
-       fill = "Dominant language") +
-  scale_y_continuous(labels = as.integer) +
-  scale_fill_brewer(palette = "Set1") +
-  theme_custom +
-  theme(legend.position = c(0.05, 0.85),
-        legend.direction = "vertical",
-        legend.justification = 0,
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))
+  filter(lp %in% "Bilingual") %>% 
+    mutate(bilingualism = cut_interval(bilingualism, n = 6,
+                                       labels = paste0(seq(0, 50, by = 10), "%"))) %>%
+    ggplot(aes(x = age_bin, y = bilingualism)) +
+    stat_bin_2d(colour = "white") +
+  scale_x_discrete(guide = guide_axis(n.dodge = 1)) +
+    scale_fill_viridis_c(option = "plasma") +
+    labs(x = "Age (months)", y = "L2 exposure", fill = "Participants") +
+    theme_classic() +
+    theme(axis.text = element_text(colour = "black"))
+ggsave(here("Figures", "participants.png"), height = 4, width = 5)
 ```
@@
 
-Parents completed an online questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
+Parents completed an on-line questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
 We implemented the questionnaire in `formr` [@arslan2020], a free, open-source platform.
@@
 
-For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item  within age bins of 2 months.
-Our model assumes that this proportion follows a logistic curve accross ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the dominance of the language of the item (*item dominance*, L1 vs.
-L2), the cognate status of the word (*cognateness*, Non-cognate vs.
-Cognate), and the frequency of the word form in its correspondent language.
-Our prediction  is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words.
+For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item in the non-dominant language within age bins of 2 months.
+Our model assumes that this proportion follows a logistic curve across ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the degree of exposure to the non-dominant language (*bilingualism*, 0-50%), the cognate status of the word (*cognateness*, Non-cognate vs.
+Cognate), and their interaction.
+Our prediction is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words, and that this effect will be larger for more balanced bilinguals.
 
 Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
-Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item porta (door, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
-We retrieved frequencies from SUBTLEX-ESP [@cuetos2011] and SUBTLEX-CAT [@boada2019] for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word [@zipf1949; @heuven2014].
-Missing frequency scores were imputed using multiple imputation [@vanbuuren2011].
+Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item *porta* (*door*, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
 Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
@@
               min = min(frequency, na.rm = TRUE),
-            max = max(frequency, na.rm = TRUE)) %>%
+              max = max(frequency, na.rm = TRUE),
+              .groups = "drop") %>%
     ungroup() %>% 
@@
                min_non_cognate = md("*Min*"),
-             max_non_cognate = md("*Max*")) %>%
-  gtsave(here("Results", "descriptives_items.png"))
+               max_non_cognate = md("*Max*"))
 
@@
 Logistic curves are characterised by three parameters[see @mahr2019 and @mahr2020 for an excellent introduction to this approach, in which we based this analysis]: (a) an *asymptote* (`asym`, upper boundary  of the proportion of infants that have acquired the item), (b) the *steepness* (`steep`, how fast the a given item is acquired by infants), and (c) a *mid-point* (`mid`, the age at which steepness is maximum).
-We work on the assumption that the trajectory of aquisition of items follows a sigmoidal shape [@mayor2011], and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
-This approach allows us to model and estimate the age of acquisition of items explicitely.
-Our model estimated the asymptote and stepness as grand means across items (i.e., intercepts).
-Mid-points were estimated by a linear zero-one-beta-regression model^[Zero-or-one-inflated beta regression models allow to account for overdispersion then using proportions as the output of a model accounting, while also allowing values to be at boundary (0 and 1).][@ospina2012] that included our predictors of interest: `age`, `dominance` (contrast-coded, L2 = -0.5, L2 = 0.5), `cognateness` (contrast-coded, Non-cognate = -0.5, Cognate = 0.5), their interaction, and `frequency` (standardised Zipf score).
-We also included a random intercept for each translation equivalent (`TE`).
+We work on the assumption that the trajectory of acquisition of items follows a logistic curve shape [@mayor2011], and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
+This approach allows us to model and estimate the age of acquisition of items explicitly.
+Our model estimated the asymptote and steepness as grand means across items (i.e., intercepts).
+Mid-points were estimated by a linear regression model that included our predictors of interest: `age`, `item_dominance` (sum-coded, L2 = -0.5, L1 = 0.5), `cognateness` (sum-coded, Non-cognate = -0.5, Cognate = 0.5).
+We also included random intercepts and `item dominance` slopes for each translation equivalent (`te`) [@barr2013].
 Appendix 1 shows a detailed description of the model.
@@
 We estimated the parameters of our model using the Bayesian framework.
-Bayesian models use the Bayes rule to combine prior knowlegde about the distribution of a given parameter, and the likelihood of each of the values it can take given tha observed data.
-We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our priors).
-In order to test the contribution of the predictor *cognateness* on the fit of the model,  we compared the fit of our (expanded) model to the fit of a (null) model that did not include *cognateness* as predictor [@schad2020] using leave-one-out cross validation.
+Bayesian models use the Bayes rule to combine prior knowledge about the distribution of a given parameter, and the likelihood of each of the values it can take given the data.
+We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our prior).
+In order to test the contribution of each predictor on the fit of the model, we started fitting a null model that only included an global intercept and random intercepts for each TE, then fitted a model that also included *bilingualism* and its random slopes by TE, and finally a model that also included *cognateness* and its interaction with *bilingualism*, as well as random slopes for all the fixed effects.
+We compared how the fit of the model changed in every step [@schad2020] using leave-one-out cross validation.
 We fit both the null and alternative models using the R environment [@rcoreteam2013; @rstudioteam2015] and the `brms` package [@burkner2017], which relies on the probabilistic language Stan [@carpenter2017] to approximate posterior distributions.
@@
 
-## Hypothesis 1
+## Hypothesis 1: Are cognates acquired ealier?
 
-### Comprehensive vocabulary
+#### Comprehensive data
 
-### Productive vocabulary
+```{r compselection, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+# compare models
+selection_comp %>% 
+  as.data.frame() %>% 
+  clean_names() %>% 
+  select(model, df, aic, bic, log_lik, l_ratio, p_value) %>% 
+  mutate(p_value = printp(p_value)) %>% 
+  gt() %>% 
+  fmt_number(columns = 1:2, decimals = 0) %>%
+  fmt_number(columns = 3:6) %>%
+  fmt_missing(columns = c(6:7), missing_text = "-") %>% 
+  cols_label(model = "Model",
+             df = md("*df*"),
+             aic = "AIC",
+             bic = "BIC",
+             log_lik = "Log. Likelihood",
+             l_ratio = "Likelihood ratio",
+             p_value = md("*p*"))
+
+```
+
+```{r companova, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+anova_comp %>% 
+  as.data.frame() %>% 
+  clean_names() %>% 
+  rownames_to_column("term") %>% 
+  mutate(p_value = printp(p_value)) %>% 
+  gt() %>% 
+  fmt_number(columns = 2, decimals = 0) %>%
+  fmt_number(columns = 3:4) %>%
+  cols_label(term = "Term",
+             num_df = md("Num. *df*"),
+             den_df = md("Den. *df*"),
+             f_value = md("*F*"),
+             p_value = md("*p*"))
+
+summary_comp <- summary(fit4_comp)
+
+```
+
+```{r comppredictions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+dat_comp %>% 
+    arrange(item, age_bin, item_dominance) %>% 
+    ggplot(aes(age_bin, proportion, colour = interaction(cognate, item_dominance, sep = " - "))) +
+    geom_line(data = comp_preds, aes(age_bin, proportion), size = 1) + 
+    stat_summary(fun = "mean", geom = "point", shape = 1, size = 3) +
+    labs(x = "Age (months)", y = "Proportion", colour = "Cognateness") +
+    scale_colour_brewer(palette = "Set1") +
+    scale_x_continuous(breaks = unique(dat_prod$age_bin), labels = bins_interest,
+                       guide = guide_axis(n.dodge = 2)) +
+    theme_bw() +
+    theme(text = element_text(size = 12),
+          legend.position = "right",
+          legend.title = element_blank(),
+          axis.text = element_text(colour = "black")) +
+  ggsave(here("Figures", "predictions_comp.png"))
+
+```
+
+### Productive data
+
+```{r prodselection, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+# compare models
+selection_prod %>% 
+  as.data.frame() %>% 
+  clean_names() %>% 
+  select(model, df, aic, bic, log_lik, l_ratio, p_value) %>% 
+  mutate(p_value = printp(p_value)) %>% 
+  gt() %>% 
+  fmt_number(columns = 1:2, decimals = 0) %>%
+  fmt_number(columns = 3:6) %>%
+  fmt_missing(columns = c(6:7), missing_text = "-") %>% 
+  cols_label(model = "Model",
+             df = md("*df*"),
+             aic = "AIC",
+             bic = "BIC",
+             log_lik = "Log. Likelihood",
+             l_ratio = "Likelihood ratio",
+             p_value = md("*p*"))
+
+```
+
+```{r prodanova, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+anova_prod %>% 
+  as.data.frame() %>% 
+  clean_names() %>% 
+  rownames_to_column("term") %>% 
+  mutate(p_value = printp(p_value)) %>% 
+  gt() %>% 
+  fmt_number(columns = 2, decimals = 0) %>%
+  fmt_number(columns = 3:4) %>%
+  cols_label(term = "Term",
+             num_df = md("Num. *df*"),
+             den_df = md("Den. *df*"),
+             f_value = md("*F*"),
+             p_value = md("*p*"))
+
+```
+
+```{r prodpredictions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+dat_prod %>% 
+    arrange(item, age_bin, item_dominance) %>% 
+    ggplot(aes(age_bin, proportion, colour = interaction(cognate, item_dominance, sep = " - "))) +
+    geom_line(data = prod_preds, aes(age_bin, proportion), size = 1) + 
+    stat_summary(fun = "mean", geom = "point", shape = 1, size = 3) +
+    labs(x = "Age (months)", y = "Proportion", colour = "Cognateness") +
+    scale_colour_brewer(palette = "Set1") +
+    scale_x_continuous(breaks = unique(dat_prod$age_bin), labels = bins_interest,
+                       guide = guide_axis(n.dodge = 2)) +
+    theme_bw() +
+    theme(text = element_text(size = 12),
+          legend.position = "right",
+          legend.title = element_blank(),
+          axis.text = element_text(colour = "black")) +
+  ggsave(here("Figures", "predictions_prod.png"))
+
+```
+
+```{r ranefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+ggplot(rancoefs, aes(xmid_intercept, xmid_item_dominance1, colour = cognate, fill = cognate)) +
+  facet_wrap(type~cognate, scales = "free") +
+  geom_point(size = 3, alpha = 0.5, shape = 1, colour = "grey") +
+  geom_smooth(method = "lm", formula = y~x) +
+  labs(x = "Mid-point intercept", y = "Item dominance slope on Mid-point", colour = "Cognateness") +
+  scale_colour_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  scale_x_continuous() + 
+  theme_bw() +
+  theme(legend.background = element_rect(fill = NA),
+        legend.title = element_blank(),
+        legend.key = element_rect(fill = NA),
+        legend.position = "none")
+
+```
 
@@
 
-### Comprehensive vocabulary
+```{r aoafig, echo=FALSE, message=TRUE, warning=FALSE, paged.print=FALSE}
 
-### Productive vocabulary
+ggplot(aoas, aes(xmid, fct_rev(item_dominance))) +
+    facet_grid(cognate~type) +
+    annotate(geom = "rect", xmin = 0, xmax = 10, ymin = -Inf, ymax = Inf, colour = NA, fill = "orange", alpha = 0.25)  +
+    #annotate(geom = "text", x = 5, y = 0, label = "Range of observed data", size = 7) +
+    geom_line(aes(group = te), colour = "black", alpha = 0.05) +
+    geom_violin(fill = "grey30", colour = "grey30", width = 0.5) +
+    geom_boxplot(width = 0.25, fill = "white", colour = "black", outlier.colour = NA) +
+    labs(x = "Age (months)", y = "Item dominance", fill = "Item dominance", colour = "Item dominance") +
+    theme_bw() +
+    scale_x_continuous(breaks = -5:15, labels = bins_expanded, guide = guide_axis(n.dodge = 1)) +
+    theme(legend.title = element_blank(),
+          legend.position = "none",
+          legend.background = element_rect(fill = NA),
+          axis.text = element_text(colour = "black"),
+          axis.text.x = element_text(size = 8),
+          axis.title.y = element_blank(),
+          panel.grid.major.y = element_blank(),
+          panel.grid.minor.y = element_blank()) +
+  ggsave(here("Figures", "aoa.png"))
+
+```
 
@@
 
-## Appendix 1
+## Appendix 1: Questionnaires
 
-Model formula and priors:
-
-$$proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ \beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\\phi \sim Normal(1.5, 1)$$
-
-## Appendix 2: Computing priors
-
-We fitted a logistic model using the proportion of infants that acquired each item as output, and age-bin (defined as we did with our data) as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
-We included data from `r nrow(dat_wordbank)/length(unique(dat_wordbank$age)) %>% printnum()` items from `r dat_wordbank$language %>% unique %>% length` languages^[Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
-All the other available languages were excluded, as no (or insufficient) data were available.], from monolingual infants aged 16 to 30 months of age.
-Table \@ref(tab:wordbank) shows the estimates of this model, which were rescaled and included as priors in our main model to test Hypothesis 1.
-
-```{r worbank, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_wordbank_priors %>%
-  rename(estimate_scaled = estimate) %>%
-  mutate(conf = paste0("[", round(conf_low, 2), ", ", round(conf_high, 2), "]"),
-         estimate = estimate_scaled*2+20,
-         term = c("Asymptote", "Mid-point", "Steepness")) %>% 
-  select(term, estimate, estimate_scaled, std_error, conf) %>%
-  gt(rowname_col = "term") %>% 
-  fmt_number(vars(estimate, estimate_scaled, std_error), decimals = 2) %>% 
-  cols_label(estimate = md("*Estimate<br>(months)*"),
-             estimate_scaled = md("*Estimate<br>(scaled)*"),
-             std_error = md("*SE*"),
-             conf = md("*95% CI*")) %>% 
-  cols_align(align = "c") %>% 
-  gtsave(here("Results", "wordbank.png"))
-
-```
-
-We performed prior-predictive checks to analyse the feasibility of our priors. Figure \@ref(fig:priorchecks)
-
-```{r priorchecks, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-include_graphics(here("Figures", "06_simulate-priors.png"))
-
-```
-
-```{r priorstab, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tribble(~param, ~coefficient, ~notation, ~prior,
-        "Asymptote", "Intercept", "$A$", paste0("$N(", dat_wordbank_priors[term=='asym'] %>% pull(estimate) %>% printnum(), ", 0.05)$"),
-        "Steepness", "Intercept", "$S$", paste0("$N(", dat_wordbank_priors[term=='steep'] %>% pull(estimate) %>% printnum(), ", 0.6)$"),
-        "Mid-point", "Intercept", "$\\beta_{0}$", paste0("$N(", dat_wordbank_priors[term=='mid'] %>% pull(estimate) %>% printnum(), ", 1)$"),
-        "Mid-point", "Random Intercepts", "$I_{i0}$", "N(0, $\\pi$)",
-        "Mid-point", "Slope", "$\\beta_{1}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{2}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{3}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{4}$", "$N(0, 5)$",
-        "Mid-point", "Intercept", "$\\varphi$", "$N(2, 1)$",
-        "Mid-point", "Intercept", "$COI$", "Beta(1, 1)", 
-        "Mid-point", "Intercept", "$ZOI$", "Beta(1, 1)",
-        "Mid-point", "Std. Deviation", "$\\sigma{ijk}$", "t(3, 0, 2.5)") %>%
-  kable(digits = 2, row.names = FALSE, col.names = c("**Adjusted parameter**", "**Coefficient**", "**Notation**", "**Distribution**"), align = "c", caption = "Prior distribution of the parameters in our models.
-We set specific informative priors for the asymptote, steepness, and mid-point of the curves (using data from Wordbank) and and weakly informative priors for the rest of the coefficients.", escape = TRUE, format = "latex")
-
-```
+## Appendix 2: Model details
 
@@
 
-```{r session_info, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r sessioninfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 

### 2020-08-22 Manuscript/trajectories_manuscript.tex
Smoother predictions, source analysis from external R in manuscript, and plot random effects

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[english,man,man,floatsintext]{apa6}
+\documentclass[
+  english,
+  man,man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
   pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
-            pdfkeywords={word acquisition, vocabulary, bilingualism, lexical similarity},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  pdfauthor={Gonzalo García-Castro1, Daniela Avila-Varela1, \& Núria Sebastian-Galles1},
+  pdflang={en-EN},
+  pdfkeywords={lexical acquisition, vocabulary, bilingualism},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 % Manuscript styling
@@
   {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
 \makeatother
 \shorttitle{Cross-linguistic similarity and  word acquisition}
-\author{Gonzalo García-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Núria Sebastian-Galles\textsuperscript{1}}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra}
-\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-Enter author note here.
-
-Correspondence concerning this article should be addressed to Gonzalo García-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain. E-mail: gonzalo.garciadecastro@upf.edu}
-\keywords{word acquisition, vocabulary, bilingualism, lexical similarity\newline\indent Word count: X}
+\keywords{lexical acquisition, vocabulary, bilingualism\newline\indent Word count: X}
 \usepackage{lineno}
@@
 \usepackage{csquotes}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[shorthands=off,main=english]{babel}
-\else
-  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
   \usepackage{polyglossia}
   \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
 \fi
@@
 \title{The role of cross-linguistic lexical similarity on bilingual word acquisition}
-
+\author{Gonzalo García-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Núria Sebastian-Galles\textsuperscript{1}}
 \date{}
 
+\authornote{
+
+Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
+
+Correspondence concerning this article should be addressed to Gonzalo García-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain.
+E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}
+
+}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain}
+
 \abstract{
-Bilinguals face the task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages sharing large amount of cognates (e.g., English-Dutch).
+Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
+Floccia et al.~(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
 The mechanisms underlying this effect remain unknown.
 We explore two compatible scenarios.
-First, cognates may be learnt earlier than non-cognates, boosting lexical acquisition at early ages.
+First, we test whether cognates are learnt earlier than non-cognates.
 This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
@@
 (2019) extended these findings to children aged three to 10 years.
-These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
+These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical acquisition.
 
@@
 A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model (Kroll, Hell, Tokowicz, \& Green, 2010; Kroll \& Stewart, 1994), the Bilingual Interactive Activation model (BIA; Dijkstra \& Van Heuven, 1998), and its successor BIA+ (Dijkstra \& Van Heuven, 2002), the Bilingual Language Interaction Network for Comprehension of Speech model (BLINCS; Shook \& Marian, 2013) and, more recently, Multilink (Dijkstra et al., 2019), which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
-However, these models do not address word learning during language acquistion.
+However, these models do not address word learning during language acquisition.
 
@@
 This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
-These scores were aggregated across pais of tanslation equivalents, obtaining a single similarity estimate of for each language pair.
+These scores were aggregated across pairs of translation equivalents, obtaining a single similarity estimate of for each language pair.
 This measure embeds confounded information about the \enquote{degree of cognateness} and the amount of phonological overlap between both languages.
-Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, Enlgish-Dutch), with few exceptions (e.g., Spanish-Basque).
-In the light of these data, it is not possible to discern whether vocabulary size is increased by the ealier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
-We address this issue by testing two hypotesis.
+Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, English-Dutch), with few exceptions (e.g., Spanish-Basque).
+In the light of these data, it is not possible to discern whether vocabulary size is increased by the earlier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
+We address this issue by testing two hypothesis.
 
@@
 
-We collected data from 407 bilinguals from the Metropolitan Area of Barcelona, between 21th November, 2017 and 22th June, 2020.
+We collected data from 437 bilinguals from the Metropolitan Area of Barcelona, between 21th noviembre, 2017 and 23th julio, 2020.
 All families participated voluntarily.
@@
 Participants were considered bilingual - therefore included - if exposed 50-90\% of the time to the non-dominant language.
-We excluded participants with \textgreater{}10\% exposure to a third language.
+We excluded participants with \textgreater10\% exposure to a third language.
 Fig.
@@
 
-Parents completed an online questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
+Parents completed an on-line questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
 We implemented the questionnaire in \texttt{formr} (Arslan, Walther, \& Tata, 2020), a free, open-source platform.
@@
 
-For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item within age bins of 2 months.
-Our model assumes that this proportion follows a logistic curve accross ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the dominance of the language of the item (\emph{item dominance}, L1 vs.~L2), the cognate status of the word (\emph{cognateness}, Non-cognate vs.~Cognate), and the frequency of the word form in its correspondent language.
-Our prediction is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words.
+For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item in the non-dominant language within age bins of 2 months.
+Our model assumes that this proportion follows a logistic curve across ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the degree of exposure to the non-dominant language (\emph{bilingualism}, 0-50\%), the cognate status of the word (\emph{cognateness}, Non-cognate vs.~Cognate), and their interaction.
+Our prediction is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words, and that this effect will be larger for more balanced bilinguals.
 
 Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
-Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item porta (door, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
-We retrieved frequencies from SUBTLEX-ESP (Cuetos, Glez-Nosti, Barbón, \& Brysbaert, 2011) and SUBTLEX-CAT (Boada, Guasch, Haro, Demestre, \& Ferré, 2019) for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word (Heuven, Mandera, Keuleers, \& Brysbaert, 2014; Zipf, 1949).
-Missing frequency scores were imputed using multiple imputation (Buuren \& Groothuis-Oudshoorn, 2011).
+Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item \emph{porta} (\emph{door}, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
 Table \ref{tab:itemsinfo} shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lcrrrrcrrrr}
+\caption*{
+\large \textbf{Questionnaire and item properties}\\ 
+\small Participants were randomly allocated on of the the four lists, and completed both the Catalan and the Spanish versions.\\ 
+} \\ 
+\toprule
+& \multicolumn{5}{c}{Frequency (Cognates)} & \multicolumn{5}{c}{Frequency (Non-cognates)} \\ 
+ \cmidrule(lr){2-6}\cmidrule(lr){7-11}
+\textbf{List} & \emph{N} & \emph{Mean} & \emph{SD} & \emph{Min} & \emph{Max} & \emph{N} & \emph{Mean} & \emph{SD} & \emph{Min} & \emph{Max} \\ 
+\midrule
+\multicolumn{1}{l}{Catalan} \\ 
+\midrule
+List A & $225$ & $4.47$ & $0.96$ & $1.25$ & $7.52$ & $100$ & $4.61$ & $1.06$ & $2.03$ & $7.44$ \\ 
+List B & $227$ & $4.49$ & $0.90$ & $0.86$ & $7.52$ & $94$ & $4.40$ & $0.97$ & $1.94$ & $6.94$ \\ 
+List C & $212$ & $4.46$ & $0.97$ & $1.67$ & $7.52$ & $106$ & $4.41$ & $1.00$ & $1.90$ & $6.94$ \\ 
+List D & $212$ & $4.43$ & $0.85$ & $1.51$ & $7.52$ & $103$ & $4.45$ & $0.94$ & $2.10$ & $6.94$ \\ 
+\midrule
+\multicolumn{1}{l}{Spanish} \\ 
+\midrule
+List A & $222$ & $4.49$ & $0.99$ & $1.68$ & $7.12$ & $98$ & $4.47$ & $0.99$ & $2.28$ & $7.53$ \\ 
+List B & $242$ & $4.35$ & $0.96$ & $1.68$ & $7.22$ & $95$ & $4.44$ & $0.93$ & $2.16$ & $7.53$ \\ 
+List C & $228$ & $4.55$ & $1.02$ & $1.68$ & $7.52$ & $104$ & $4.58$ & $0.99$ & $2.87$ & $7.53$ \\ 
+List D & $228$ & $4.41$ & $1.00$ & $1.68$ & $7.32$ & $106$ & $4.54$ & $1.05$ & $2.70$ & $7.53$ \\ 
+\bottomrule
+\end{longtable}
+
+We fit two separate logistic models on comprehensive and productive data, respectively.
+Logistic curves are characterised by three parameters(see Mahr, 2019, and @mahr2020 for an excellent introduction to this approach, in which we based this analysis): (a) an \emph{asymptote} (\texttt{asym}, upper boundary of the proportion of infants that have acquired the item), (b) the \emph{steepness} (\texttt{steep}, how fast the a given item is acquired by infants), and (c) a \emph{mid-point} (\texttt{mid}, the age at which steepness is maximum).
+We work on the assumption that the trajectory of acquisition of items follows a logistic curve shape (Mayor \& Plunkett, 2011), and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
+This approach allows us to model and estimate the age of acquisition of items explicitly.
+Our model estimated the asymptote and steepness as grand means across items (i.e., intercepts).
+Mid-points were estimated by a linear regression model that included our predictors of interest: \texttt{age}, \texttt{item\_dominance} (sum-coded, L2 = -0.5, L1 = 0.5), \texttt{cognateness} (sum-coded, Non-cognate = -0.5, Cognate = 0.5).
+We also included random intercepts and \texttt{item\ dominance} slopes for each translation equivalent (\texttt{te}) ({\textbf{???}}).
+Appendix 1 shows a detailed description of the model.
+
+We estimated the parameters of our model using the Bayesian framework.
+Bayesian models use the Bayes rule to combine prior knowledge about the distribution of a given parameter, and the likelihood of each of the values it can take given the data.
+We relied on the available data on Wordbank (Frank, Braginsky, Yurovsky, \& Marchman, 2017) to generate our priors (see Appendix 1 for more details about our prior).
+In order to test the contribution of each predictor on the fit of the model, we started fitting a null model that only included an global intercept and random intercepts for each TE, then fitted a model that also included \emph{bilingualism} and its random slopes by TE, and finally a model that also included \emph{cognateness} and its interaction with \emph{bilingualism}, as well as random slopes for all the fixed effects.
+We compared how the fit of the model changed in every step (Schad, Betancourt, \& Vasishth, 2020) using leave-one-out cross validation.
+We fit both the null and alternative models using the R environment (R Core Team, 2013; RStudio Team, 2015) and the \texttt{brms} package (Bürkner, 2017), which relies on the probabilistic language Stan (Carpenter et al., 2017) to approximate posterior distributions.
+We used Pareto-smoothed importance sampling (PSIS; Vehtari, Gelman, \& Gabry, 2017; Vehtari, Simpson, Gelman, Yao, \& Gabry, 2019) to compare the null and extended models.
+Data and model results were processed and visualised using the \texttt{tidyverse} family of R packages (Wickham et al., 2019) and the \texttt{tidybayes} package (Kay, 2020).
+
+To test Hypothesis 2, we run a Bayesian ANOVA using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included \emph{cognateness} as predictor.
+Using the \texttt{BayesFactor} R package (Morey \& Rouder, 2018), we computed a Bayes factor (\emph{BF}) that compared the likelihood of a linear model that included \emph{cognateness} as a predictor against the likelihood of a linear model that did not, under the light of our data.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{hypothesis-1-are-cognates-acquired-ealier}{%
+\subsection{Hypothesis 1: Are cognates acquired ealier?}\label{hypothesis-1-are-cognates-acquired-ealier}}
+
+\hypertarget{comprehensive-data}{%
+\paragraph{Comprehensive data}\label{comprehensive-data}}
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{crrrrrl}
+\toprule
+Model & \emph{df} & AIC & BIC & Log. Likelihood & Likelihood ratio & \emph{p} \\ 
+\midrule
+$1$ & $5$ & $53,715.17$ & $53,761.83$ & $-26,852.58$ & - &  \\ 
+$2$ & $6$ & $53,716.39$ & $53,772.38$ & $-26,852.20$ & $0.78$ & .378 \\ 
+$3$ & $9$ & $51,646.15$ & $51,730.14$ & $-25,814.08$ & $2,076.24$ & < .001 \\ 
+$4$ & $10$ & $51,702.84$ & $51,796.16$ & $-25,841.42$ & $54.69$ & < .001 \\ 
+$5$ & $11$ & $51,565.73$ & $51,668.38$ & $-25,771.86$ & $139.11$ & < .001 \\ 
+\bottomrule
+\end{longtable}
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lcrrl}
+\toprule
+Term & Num. \emph{df} & Den. \emph{df} & \emph{F} & \emph{p} \\ 
+\midrule
+Asym & $1$ & $82,832.00$ & $60,748.66$ & < .001 \\ 
+xmid.(Intercept) & $1$ & $82,832.00$ & $1,132.58$ & < .001 \\ 
+xmid.frequency & $1$ & $82,832.00$ & $0.16$ & .693 \\ 
+xmid.item\_dominance & $1$ & $82,832.00$ & $843.09$ & < .001 \\ 
+xmid.cognate & $1$ & $82,832.00$ & $3.00$ & .083 \\ 
+xmid.item\_dominance:cognate & $1$ & $82,832.00$ & $50.68$ & < .001 \\ 
+scal & $1$ & $82,832.00$ & $7,635.78$ & < .001 \\ 
+\bottomrule
+\end{longtable}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/itemsinfo-1}
+{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/comppredictions-1}
 
@@
 
-\caption{ }\label{fig:itemsinfo}
+\caption{ }\label{fig:comppredictions}
 \end{figure}
 
-We fit two separate logistic models on comprehensive and productive data, respectively.
-Logistic curves are characterised by three parameters(see Mahr, 2019, and @mahr2020 for an excellent introduction to this approach, in which we based this analysis): (a) an \emph{asymptote} (upper boundary of the proportion of infants that have acquired the item), (b) the \emph{steepness} (how fast the a given item is acquired by infants), and (c) a \emph{mid-point} (the age at which steepness is maximum).
-We work on the assumption that the trajectory of aquisition of items follows a sigmoidal shape (Mayor \& Plunkett, 2011), and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
-This approach allows us to model and estimate the age of acquisition of items explicitely.
-Our model estimated the asymptote and stepness as grand means across items (i.e., intercepts).
-Mid-points were estimated by a linear zero-one-beta-regression model\footnote{Zero-or-one-inflated beta regression models allow to account for overdispersion then using proportions as the output of a model accounting, while also allowing values to be at boundary (0 and 1).}(Ospina \& Ferrari, 2012) that included our predictors of interest (\emph{age bin}, \emph{item dominance} (L1/L2, sum-coded), \emph{cognateness} (\emph{non-cognate/cognate, sum-coded), the interaction between }item dominance* and \emph{cognateness}, and frequency (standardised Zipf score){]}.
-We also included a random intercept for each translation equivalent.
+\hypertarget{productive-data}{%
+\subsubsection{Productive data}\label{productive-data}}
 
-\[y_{ijk} = \frac{A}{1 + e^{(M-Age) \times S}}\]
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{crrrrrl}
+\toprule
+Model & \emph{df} & AIC & BIC & Log. Likelihood & Likelihood ratio & \emph{p} \\ 
+\midrule
+$1$ & $5$ & $23,516.51$ & $23,562.84$ & $-11,753.26$ & - &  \\ 
+$2$ & $6$ & $23,503.04$ & $23,558.64$ & $-11,745.52$ & $15.47$ & < .001 \\ 
+$3$ & $9$ & $21,344.08$ & $21,427.48$ & $-10,663.04$ & $2,164.96$ & < .001 \\ 
+$4$ & $9$ & $21,343.35$ & $21,426.74$ & $-10,662.67$ & - &  \\ 
+$5$ & $11$ & $21,291.22$ & $21,393.15$ & $-10,634.61$ & $56.13$ & < .001 \\ 
+\bottomrule
+\end{longtable}
 
-\[M = \beta_{0} + I_{i0} + \beta_{1} Dominance_{j} + \beta_{2} Cognateness_{ik} + \beta_{3} Dominance_{ij} \times Cognateness_{ik} + \beta_{4} Frequency + \sigma_{ijk}\]
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lcrrl}
+\toprule
+Term & Num. \emph{df} & Den. \emph{df} & \emph{F} & \emph{p} \\ 
+\midrule
+Asym & $1$ & $77,537.00$ & $5,719.12$ & < .001 \\ 
+xmid.(Intercept) & $1$ & $77,537.00$ & $3,585.35$ & < .001 \\ 
+xmid.frequency & $1$ & $77,537.00$ & $8.17$ & .004 \\ 
+xmid.item\_dominance & $1$ & $77,537.00$ & $655.13$ & < .001 \\ 
+xmid.cognate & $1$ & $77,537.00$ & $0.52$ & .473 \\ 
+xmid.item\_dominance:cognate & $1$ & $77,537.00$ & $36.83$ & < .001 \\ 
+scal & $1$ & $77,537.00$ & $4,870.09$ & < .001 \\ 
+\bottomrule
+\end{longtable}
 
-We estimated the parameters of our model using the Bayesian framework.
-Bayesian models use the Bayes rule to combine prior knowlegde about the distribution of a given parameter, and the likelihood of each of the values it can take given tha observed data.
-We relied on the available data on Wordbank (Frank, Braginsky, Yurovsky, \& Marchman, 2017) to generate our priors (see Appendix 1 for more details about our priors).
-Finally, to test the contribution of the \emph{cognateness} predictor to the estimation of mid-points, we compared the fit of the model with a reduced (null) model that did not include the \emph{cognateness} factor using Pareto-smoothed importance sampling (PSIS; Vehtari, Gelman, \& Gabry, 2017; Vehtari, Simpson, Gelman, Yao, \& Gabry, 2019).
-We fit both null and alternative models using the R environment (R Core Team, 2013; RStudio Team, 2015) and the \texttt{brms} package (Bürkner, 2017), which relies on the probabilistic language Stan (Carpenter et al., 2017) to estimate posterior distributions.
-Data and model results will be processed and visualised using the R packages the \texttt{tidyverse} family of packages (Wickham et al., 2019) and the \texttt{tidybayes} package (Kay, 2020).
+\begin{figure}
 
-To test Hypothesis 2, we run a Bayesian anova using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included \emph{cognateness} as predictor.
-Using the \texttt{BayesFactor} R package (Morey \& Rouder, 2018), we computed a Bayes factor (\emph{BF}) that compared the likelihood of a linear model that included \emph{cognateness} as a predictor against the likelihood of a linear model that did not, under the light of our data.
+{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/prodpredictions-1}
 
-\begin{table}
+}
 
-\caption{\label{tab:priorstab}Prior distribution of the parameters in our models.
-We set specific informative priors for the asymptote, steepness, and mid-point of the curves (using data from Wordbank) and and weakly informative priors for the rest of the coefficients.}
-\centering
-\begin{tabular}[t]{c|c|c|c}
-\hline
-**Adjusted parameter** & **Coefficient** & **Notation** & **Distribution**\\
-\hline
-Asymptote & Intercept & \$A\$ & \$N(0.79, 0.05)\$\\
-\hline
-Steepness & Intercept & \$S\$ & \$N(1.79, 0.6)\$\\
-\hline
-Mid-point & Intercept & \$\textbackslash{}beta\_\{0\}\$ & \$N(2.52, 1)\$\\
-\hline
-Mid-point & Random Intercepts & \$I\_\{i0\}\$ & N(0, \$\textbackslash{}pi\$)\\
-\hline
-Mid-point & Slope & \$\textbackslash{}beta\_\{1\}\$ & \$N(0, 5)\$\\
-\hline
-Mid-point & Slope & \$\textbackslash{}beta\_\{2\}\$ & \$N(0, 5)\$\\
-\hline
-Mid-point & Slope & \$\textbackslash{}beta\_\{3\}\$ & \$N(0, 5)\$\\
-\hline
-Mid-point & Slope & \$\textbackslash{}beta\_\{4\}\$ & \$N(0, 5)\$\\
-\hline
-Mid-point & Intercept & \$\textbackslash{}varphi\$ & \$N(2, 1)\$\\
-\hline
-Mid-point & Intercept & \$COI\$ & Beta(1, 1)\\
-\hline
-Mid-point & Intercept & \$ZOI\$ & Beta(1, 1)\\
-\hline
-Mid-point & Std. Deviation & \$\textbackslash{}sigma\{ijk\}\$ & t(3, 0, 2.5)\\
-\hline
-\end{tabular}
-\end{table}
+\caption{ }\label{fig:prodpredictions}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/ranefs-1}
+
+}
+
+\caption{ }\label{fig:ranefs}
+\end{figure}
+
+\hypertarget{hypothesis-2}{%
+\subsection{Hypothesis 2}\label{hypothesis-2}}
+
+\begin{verbatim}
+## Saving 6.5 x 4.5 in image
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/aoafig-1}
+
+}
+
+\caption{ }\label{fig:aoafig}
+\end{figure}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\hypertarget{appendix-1-session-info}{%
-\subsection{Appendix 1: Session info}\label{appendix-1-session-info}}
+\hypertarget{appendix-1-questionnaires}{%
+\subsection{Appendix 1: Questionnaires}\label{appendix-1-questionnaires}}
+
+\hypertarget{appendix-2-model-details}{%
+\subsection{Appendix 2: Model details}\label{appendix-2-model-details}}
+
+\hypertarget{appendix-3-session-info}{%
+\subsection{Appendix 3: Session info}\label{appendix-3-session-info}}
 
 R version 3.6.3 (2020-02-29)
-Platform: x86\_64-apple-darwin15.6.0 (64-bit)
-Running under: macOS Catalina 10.15.4
+Platform: x86\_64-pc-linux-gnu (64-bit)
+Running under: Ubuntu 20.04.1 LTS
 
 Matrix products: default
-BLAS: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRblas.0.dylib
-LAPACK: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRlapack.dylib
+BLAS: /usr/lib/x86\_64-linux-gnu/blas/libblas.so.3.9.0
+LAPACK: /usr/lib/x86\_64-linux-gnu/lapack/liblapack.so.3.9.0
 
 locale:
-{[}1{]} en\_US.UTF-8/en\_US.UTF-8/en\_US.UTF-8/C/en\_US.UTF-8/en\_US.UTF-8
+{[}1{]} LC\_CTYPE=en\_GB.UTF-8 LC\_NUMERIC=C\\
+{[}3{]} LC\_TIME=es\_ES.UTF-8 LC\_COLLATE=en\_GB.UTF-8\\
+{[}5{]} LC\_MONETARY=es\_ES.UTF-8 LC\_MESSAGES=en\_GB.UTF-8\\
+{[}7{]} LC\_PAPER=es\_ES.UTF-8 LC\_NAME=C\\
+{[}9{]} LC\_ADDRESS=C LC\_TELEPHONE=C\\
+{[}11{]} LC\_MEASUREMENT=es\_ES.UTF-8 LC\_IDENTIFICATION=C
 
@@
 other attached packages:
-{[}1{]} rlang\_0.4.6 here\_0.1 data.table\_1.12.8\\
-{[}4{]} gt\_0.2.1 patchwork\_1.0.0.9000 ggridges\_0.5.2\\
-{[}7{]} janitor\_2.0.1 lubridate\_1.7.9 readxl\_1.3.1\\
-{[}10{]} forcats\_0.5.0 stringr\_1.4.0 dplyr\_1.0.0\\
-{[}13{]} purrr\_0.3.4 readr\_1.3.1 tidyr\_1.1.0\\
-{[}16{]} tibble\_3.0.1 ggplot2\_3.3.2 tidyverse\_1.3.0\\
-{[}19{]} magrittr\_1.5 knitr\_1.28 papaja\_0.1.0.9942
+{[}1{]} rlang\_0.4.7 ggdist\_2.2.0 here\_0.1 data.table\_1.13.0
+{[}5{]} gt\_0.2.1 modelr\_0.1.8 broom.mixed\_0.2.6 nlme\_3.1-145\\
+{[}9{]} janitor\_2.0.1 lubridate\_1.7.9 readxl\_1.3.1 forcats\_0.5.0\\
+{[}13{]} ggplot2\_3.3.2 tibble\_3.0.3 tidyr\_1.1.0 dplyr\_1.0.1\\
+{[}17{]} magrittr\_1.5 knitr\_1.29 papaja\_0.1.0.9997
 
 loaded via a namespace (and not attached):
-{[}1{]} Rcpp\_1.0.4.6 lattice\_0.20-38 ps\_1.3.3 assertthat\_0.2.1\\
-{[}5{]} rprojroot\_1.3-2 digest\_0.6.25 R6\_2.4.1 cellranger\_1.1.0\\
-{[}9{]} plyr\_1.8.6 backports\_1.1.8 reprex\_0.3.0 evaluate\_0.14\\
-{[}13{]} httr\_1.4.1 pillar\_1.4.4 rstudioapi\_0.11 callr\_3.4.3.9000\\
-{[}17{]} blob\_1.2.1 checkmate\_2.0.0 rmarkdown\_2.3 labeling\_0.3\\
-{[}21{]} webshot\_0.5.2 munsell\_0.5.0 broom\_0.5.6 compiler\_3.6.3\\
-{[}25{]} modelr\_0.1.8 xfun\_0.15 pkgconfig\_2.0.3 htmltools\_0.5.0\\
-{[}29{]} tidyselect\_1.1.0 bookdown\_0.18 fansi\_0.4.1 crayon\_1.3.4\\
-{[}33{]} dbplyr\_1.4.4 withr\_2.2.0 commonmark\_1.7 grid\_3.6.3\\
-{[}37{]} nlme\_3.1-147 jsonlite\_1.6.1 gtable\_0.3.0 lifecycle\_0.2.0\\
-{[}41{]} DBI\_1.1.0 scales\_1.1.1 cli\_2.0.2 stringi\_1.4.6\\
-{[}45{]} farver\_2.0.3 fs\_1.4.1 snakecase\_0.11.0 xml2\_1.3.2\\
-{[}49{]} ellipsis\_0.3.1 generics\_0.0.2 vctrs\_0.3.1 RColorBrewer\_1.1-2
-{[}53{]} tools\_3.6.3 glue\_1.4.1 hms\_0.5.3 processx\_3.4.2\\
-{[}57{]} yaml\_2.2.1 colorspace\_1.4-1 rvest\_0.3.5 haven\_2.3.1\\
-{[}61{]} sass\_0.2.0
-
-\hypertarget{appendix-2-model}{%
-\subsection{Appendix 2: Model}\label{appendix-2-model}}
-
-Our model is incomplete, in that we have not included all parameters that may play a role in the real mechanisms that generate our observed data.
-Previous studies have found robust effects of sex (e.g., Frank et al.
-(2017)), word length/number of phonemes (e.g., Frank et al.
-(2017)), etc.
-Computational limitations have prevented us from specifying and fitting a more exhaustive model.
-Following the model-building strategy suggested by Schad, Betancourt, and Vasishth (2020), we have specified an intitial model that captures the fundamental structure of our phenomenon of interest and that can be computed with our available resources.
-Then, we have expanded such model incorporating a new predictor that adresses our hypotheses.
-
-\hypertarget{appendix-3-priors}{%
-\subsection{Appendix 3: Priors}\label{appendix-3-priors}}
-
-We fitted a logistic model using the proportion of infants that acquired each item as output, and age-bin (defined as we did with our data) as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
-We included data from 13119 items from 20 languages\footnote{Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
-All the other available languages were excluded, as no (or insufficient) data were available.}, from monolingual infants aged 16 to 30 months of age.
-Table \ref{tab:wordbank} shows the estimates of this model, which were rescaled and included as priors in our main model to test Hypothesis 1.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{trajectories_manuscript_files/figure-latex/worbank-1}
-
-}
-
-\caption{ }\label{fig:worbank}
-\end{figure}
+{[}1{]} tidyselect\_1.1.0 xfun\_0.16 TMB\_1.7.18\\
+{[}4{]} purrr\_0.3.4 reshape2\_1.4.4 splines\_3.6.3\\
+{[}7{]} lattice\_0.20-40 snakecase\_0.11.0 colorspace\_1.4-1\\
+{[}10{]} vctrs\_0.3.2 generics\_0.0.2 viridisLite\_0.3.0\\
+{[}13{]} htmltools\_0.5.0 mgcv\_1.8-31 yaml\_2.2.1\\
+{[}16{]} pillar\_1.4.6 glue\_1.4.1 withr\_2.2.0\\
+{[}19{]} RColorBrewer\_1.1-2 distributional\_0.1.0 lifecycle\_0.2.0\\
+{[}22{]} plyr\_1.8.6 stringr\_1.4.0 commonmark\_1.7\\
+{[}25{]} munsell\_0.5.0 gtable\_0.3.0 cellranger\_1.1.0\\
+{[}28{]} codetools\_0.2-16 coda\_0.19-3 evaluate\_0.14\\
+{[}31{]} labeling\_0.3 fansi\_0.4.1 broom\_0.7.0.9001\\
+{[}34{]} Rcpp\_1.0.5 checkmate\_2.0.0 scales\_1.1.1\\
+{[}37{]} backports\_1.1.8 farver\_2.0.3 digest\_0.6.25\\
+{[}40{]} stringi\_1.4.6 bookdown\_0.20 grid\_3.6.3\\
+{[}43{]} rprojroot\_1.3-2 cli\_2.0.2 tools\_3.6.3\\
+{[}46{]} crayon\_1.3.4 pkgconfig\_2.0.3 ellipsis\_0.3.1\\
+{[}49{]} Matrix\_1.2-18 assertthat\_0.2.1 rmarkdown\_2.3\\
+{[}52{]} R6\_2.4.1 compiler\_3.6.3
 
@@
 
-\leavevmode\hypertarget{ref-boada2019}{}%
-Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P.
-(2019).
-SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for Catalan.
-\emph{Behavior Research Methods}.
-\url{https://doi.org/10.3758/s13428-019-01233-1}
-
 \leavevmode\hypertarget{ref-bobb2020}{}%
@@
 \url{https://doi.org/10.1162/opmi_a_00026}
-
-\leavevmode\hypertarget{ref-vanbuuren2011}{}%
-Buuren, S.
-van, \& Groothuis-Oudshoorn, K.
-(2011).
-\textless{}span class="nocase"\textgreater{}mice\textless{}/span\textgreater{}: Multivariate imputation by chained equations in r.
-\emph{Journal of Statistical Software}, \emph{45}(3), 1--67.
-Retrieved from \url{https://www.jstatsoft.org/v45/i03/}
 
@@
 \url{https://doi.org/10.1037/0278-7393.26.5.1283}
-
-\leavevmode\hypertarget{ref-cuetos2011}{}%
-Cuetos, F., Glez-Nosti, M., Barbón, A., \& Brysbaert, M. (2011). SUBTLEX-ESP: Spanish word frequencies based on film subtitles. \emph{Psicológica}, \emph{32}(2), 133--143.
 
@@
 
-\leavevmode\hypertarget{ref-heuven2014}{}%
-Heuven, W.
-J.
-B.
-van, Mandera, P., Keuleers, E., \& Brysbaert, M.
-(2014).
-Subtlex-UK: A New and Improved Word Frequency Database for British English: \emph{Quarterly Journal of Experimental Psychology}.
-Retrieved from \url{https://journals.sagepub.com/doi/10.1080/17470218.2013.850521}
-
 \leavevmode\hypertarget{ref-hoff2012}{}%
@@
 \leavevmode\hypertarget{ref-kay2020}{}%
-Kay, M.
-(2020).
-\emph{\textless{}span class="nocase"\textgreater{}tidybayes\textless{}/span\textgreater{}: Tidy data and geoms for Bayesian models} (manual).
-\url{https://doi.org/10.5281/zenodo.1308151}
+Kay, M. (2020). \emph{\textless span class="nocase"\textgreater tidybayes\textless/span\textgreater: Tidy data and geoms for Bayesian models} (manual). \url{https://doi.org/10.5281/zenodo.1308151}
 
@@
 Morey, R. D., \& Rouder, J. N. (2018). \emph{BayesFactor: Computation of bayes factors for common designs} (manual). Retrieved from \url{https://CRAN.R-project.org/package=BayesFactor}
-
-\leavevmode\hypertarget{ref-ospina2012}{}%
-Ospina, R., \& Ferrari, S.
-L.
-P.
-(2012).
-A general class of zero-or-one inflated beta regression models.
-\emph{Computational Statistics \& Data Analysis}, \emph{56}(6), 1609--1623.
-\url{https://doi.org/10.1016/j.csda.2011.10.005}
 
@@
 (2019).
-Welcome to the \textless{}span class="nocase"\textgreater{}tidyverse\textless{}/span\textgreater{}.
+Welcome to the \textless span class="nocase"\textgreater tidyverse\textless/span\textgreater.
 \emph{Journal of Open Source Software}, \emph{4}(43), 1686.
@@
 
-\leavevmode\hypertarget{ref-zipf1949}{}%
-Zipf, G. K. (1949). \emph{Human behavior and the principle of least effort}. Oxford, England: Addison-Wesley Press.
-
 \endgroup

### 2020-08-22 Code/R/05_priors-prod.R

@@ -1,228 +1,0 @@
-#### 05_priors-prod: compute priors from Worbank data
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up ##############################################
-
-# load packages
-library(data.table)
-library(tidyverse)
-library(janitor)
-library(broom)
-library(modelr)
-library(brms)
-library(tidybayes)
-library(relayer)
-library(patchwork)
-library(here)
-
-# load functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-set.seed(888)
-bins_wordbank <- c("< 16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "> 40")
-breaks_wordbank <- c(0, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 100)
-bins_interest_wordbank <- c("16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-
-#### import data ############################################
-
-# wordbank data
-dat_wordbank <- list.files(here("Data", "Wordbank"), full.names = TRUE) %>%
-  map(., fread) %>%
-  set_names(str_remove(list.files(here("Data", "Wordbank")), ".csv")) %>%
-  bind_rows(.id = "language") %>%
-  as_tibble() %>%
-  pivot_longer(6:ncol(.), names_to = "age", values_to = "proportion") %>%
-  mutate(age = as.numeric(age),
-         age_bin = factor(cut(age, breaks = breaks_wordbank, labels = bins_wordbank), levels = bins_interest),
-         age_bin = as.numeric(age_bin)-2) %>%
-  drop_na(age_bin)
-
-# raw data
-dat <- fread(here("Data", "04_prepared.csv")) %>%
-  as_tibble() %>% 
-  filter(type=="Productive",
-         lp=="Bilingual") %>%
-  select(item, meaning, age_bin, item_dominance, cognate, proportion, frequency) %>%
-  mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE)),
-         item_dominance = as.factor(item_dominance),
[…]

### 2020-08-22 Code/R/03_familiarity.R
Smoother predictions, source analysis from external R in manuscript, and plot random effects

@@ -1,104 +1,0 @@
-#### 03_vocabulary: vocabulary sizes #####################
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up ##############################################
-
-# load packages
-library(dplyr)        # for nice data frames
-library(tidyr)
-library(magrittr)      # for using pipes
-library(data.table)    # for importing data
-library(lubridate)     # for working with dates
-library(readxl)        # for importing Excel spreadsheets
-library(here)          # for locating files
-
-# load/create functions
-source(here("Code", "R", "functions.R"))
-
-
-# set params
-bins          <- c("< 10", "10-12", "12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "36-38", "38-40", "> 40")
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-breaks <- c(0, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 100)
-
-#### import data ############################################
-pool <- read_xlsx(here("Data","01_pool.xlsx")) %>%
-  rename(te = meaning) %>% 
-  select(-c(survey, version, ipa, source, label, A, B, C, D, cognate_expert, cognate_rater2, agreement, comments)) %>% 
-  rename(cognate = cognate_rater1) %>%
-  mutate_at(vars(include, cognate), ~as.logical(as.numeric(.)))
-
-  
-
-dat <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactors = FALSE, na.strings = "") %>%
-  as_tibble() %>%
-  mutate_at(vars(time_stamp), as_date) %>%
-  mutate_at(vars(id_db), as.character) %>%
-  mutate(dominance = case_when(id_db %in% "54469" ~ "Spanish",
-                               id_db %in% "57157" ~ "Catalan",
-                               TRUE ~ dominance),
-         doe_catalan = case_when(id_db %in% "54469" ~ 0,
-                                 id_db %in% "57157" ~ 80,
-                                 id_db %in% "57046" ~ 50,
-                                 TRUE ~ as.numeric(doe_catalan)),
[…]

### 2020-08-22 Code/R/04_prepare.R

@@ -1,55 +1,0 @@
-#### 04_prepare: Prepare data for analysis ################
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up ###############################################
-
-# load packages
-library(data.table) # for importing data
-library(tibble)     # more more informative data frames
-library(dplyr)      # for manipulating data
-library(mice)       # for imputing data
-library(here)       # for locating files
-
-# create functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-
-#### import data #########################################
-dat_familiarity <- fread(here("Data", "03_familiarity.csv"), quote = '"', na.strings = c("", "NA")) %>% 
-    rowwise() %>%
-    mutate(prop = prod(successes, 1/n, na.rm = TRUE))  %>%
-    filter(class %in% c("noun", "verb"),
-           lp %in% "Bilingual") %>% 
-    ungroup()
-
-#### impute data ########################################
-dat_imputed <- dat_familiarity %>%
-    mice(data = .,
-         method = "norm",
-         m = 5,
-         maxit = 5, 
-         seed = 888) %>%
-    complete() %>%
-    as_tibble() %>%
-    ungroup()
-
-#### export data ########################################
-fwrite(dat_imputed, here("Data", "04_prepared.csv"), sep = ",", row.names = FALSE)
-
-#### visualise data #######################################
-ggplot(dat_imputed, aes(age_bin, proportion, colour = cognate, group = cognate)) +
-    facet_grid(item_dominance~type) +
-    geom_smooth(method = "lm", formula = y ~ log(x), se = FALSE, size = 0.75) +
-    stat_summary(fun = "mean", geom = "point", shape = 1,  na.rm = TRUE) +
-    labs(x = "Age (months)", y = "Proportion", colour = "Cognateness", fill = "Cognateness") +
-    scale_colour_manual(values = c("red", "blue")) +
-    scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
-    theme_bw() +
-    theme(legend.title = element_blank(),
-          legend.text = element_text(size = 7),
-          axis.text.x = element_text(size = 7),
-          legend.position = "top") +
-    ggsave(here("Figures", "03_familiarity.png"), height = 5)

### 2020-08-22 Code/R/05_priors-comp.R

@@ -1,255 +1,0 @@
-#### 05_priors-comp: compute priors from Worbank data ####
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up ##############################################
-
-# load packages
-library(data.table)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(purrr)
-library(stringr)
-library(janitor)
-library(broom)
-library(modelr)
-library(brms)
-library(tidybayes)
-library(patchwork)
-library(here)
-
-# load functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-set.seed(888)
-bins_wordbank <- c("< 16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "> 40")
-breaks_wordbank <- c(0, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 100)
-bins_interest_wordbank <- c("16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-
-#### import data ############################################
-
-# wordbank data
-dat_wordbank <- list.files(here("Data", "Wordbank"), full.names = TRUE) %>%
-  map(., fread) %>%
-  set_names(str_remove(list.files(here("Data", "Wordbank")), ".csv")) %>%
-  bind_rows(.id = "language") %>%
-  as_tibble() %>%
-  pivot_longer(6:ncol(.), names_to = "age", values_to = "proportion") %>%
-  mutate(age = as.numeric(age),
-         age_bin = factor(cut(age, breaks = breaks_wordbank, labels = bins_wordbank), levels = bins_interest),
-         age_bin = as.numeric(age_bin)-2) %>%
-  drop_na(age_bin)
-
-# raw data
-dat <- fread(here("Data", "04_prepared.csv")) %>%
-  as_tibble() %>% 
-  filter(type=="Comprehensive") %>%
-  select(item, meaning, age_bin, bilingualism, item_dominance, cognate, proportion, frequency) %>%
-  mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE)),
[…]

### 2020-08-22 Code/R/06_gca-comp.R

@@ -30,7 +30,8 @@ control <- list(adapt_delta = 0.95, max_treedepth = 15)
 #### import data ##########################################
 dat <- fread(here("Data", "04_prepared.csv")) %>%
     as_tibble() %>% 
-    filter(type=="Comprehensive") %>%
+    filter(type=="Comprehensive",
+           n > 4) %>%
     select(item, te, category, age_bin, item_dominance, cognate, frequency, n, successes, proportion) %>%
     mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1) %>% 
     mutate_at(vars(cognate, item_dominance), as.factor) %>% 
@@ -53,8 +54,7 @@ priors <- c(prior(normal(0.7631182, 0.05), nlpar = "asym", coef = "Intercept"),
 
 # model 0
 fit0 <- brm(formula = bf(proportion ~ asym*inv(1+exp((mid-age_bin)*steep)),
-                         asym ~ 1, steep ~ 1,
-                         mid ~ 1 + frequency + (1 | te),
+                         asym ~ 1, steep ~ 1, mid ~ 1 + frequency + (1 | te),
                          nl = TRUE),
             prior = priors, data = dat, chains = nchains, cores = ncores, iter = niter, control = control,
             file = here("Results", "comp_fit0.rds"),
@@ -63,8 +63,7 @@ fit0 <- brm(formula = bf(proportion ~ asym*inv(1+exp((mid-age_bin)*steep)),
 
 # model 1
 fit1 <- brm(formula = bf(proportion ~ asym * inv(1+exp((mid-age_bin)*steep)),
-                         asym ~ 1,
-                         steep ~ 1,
+                         asym ~ 1, steep ~ 1,
                          mid ~ 1 + frequency + item_dominance + (1 + item_dominance | te),
                          nl = TRUE),
             prior = c(priors, prior(lkj(2), class = "cor")),

### 2020-08-22 Code/R/06_gca-prod.R

@@ -1,207 +1,0 @@
-#### 06_gca-prod: Bayesian non-linear model ##############
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up ##############################################
-
-# load packages
-library(dplyr)
-library(tibble)
-library(tidyr)
-library(stringr)
-library(data.table)
-library(brms)
-library(modelr)
-library(janitor)
-library(tidybayes)
-library(here)
-
-# create/load functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-set.seed(888)
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-nchains <- 4
-ncores <- 4
-niter <- 2000
-control <- list(adapt_delta = 0.95, max_treedepth = 15)
-
-#### import data ##########################################
-dat <- fread(here("Data", "04_prepared.csv")) %>%
-    as_tibble() %>% 
-    filter(type=="Productive") %>%
-    select(item, te, category, age_bin, item_dominance, cognate, frequency, n, successes, proportion) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1) %>%
-    mutate_at(vars(cognate, item_dominance), as.factor) %>% 
-    mutate_at(vars(age_bin, frequency), function(x) scale(x, scale = FALSE, center = TRUE)[,1]) %>% 
-    arrange(item, te, age_bin)
-
-contrasts(dat$cognate) <- contr.sum(c("Non-cognate", "Cognate"))/2
-contrasts(dat$item_dominance) <- contr.sum(c("L2", "L1"))/2
-
-dat_priors <- fread(here("Data", "05_priors-wordbank.csv")) %>%
-    mutate(estimate_scaled = (estimate_scaled-14)/2)
-
-#### fit models ###########################################
-a <- expand_grid(item = paste("id", 1:25), age_bin = 1:11) %>%
-    group_by(item) %>%
-    mutate(asym =  rnorm(n = 1, mean = 0.7631182, sd = 0.05),
-           mid = rnorm(n = 1, mean = 5.3694351, sd = 1),
-           steep = rnorm(n = 1, mean = 1.6859966, sd = 0.5)) %>% 
-    rowwise() %>% 
[…]

### 2020-08-22 Code/R/07_anova_prod.R

@@ -1,107 +1,0 @@
-#### 07_anova-prod: Survival analysis on productive data ##########
-# Gonzalo García-Castro, gonzalo.garciadecastro@upf.edu
-# Center for Brain and Cognition, Universitat Pompeu Fabra
-
-#### set up #######################################################
-
-# load packages
-library(tidyverse)
-library(data.table)
-library(tidybayes)
-library(BayesFactor)
-library(ggridges)
-library(patchwork)
-library(see)
-library(here)
-
-# load functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-set.seed(888)
-bins <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-
-#### import data ###################################################
-fit1 <- readRDS(here("Results", "prod_fit1.rds"))
-dat <- fread(here("Data", "04_prepared.csv")) %>%
-    as_tibble() %>% 
-    filter(type=="Productive",
-           lp=="Bilingual") %>%
-    select(item, meaning, age_bin, item_dominance, cognate, proportion, frequency) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins, ordered = TRUE)),
-           item_dominance = factor(item_dominance, levels = c("L2", "L1")),
-           frequency = scale(frequency)[,1],
-           cognate = factor(cognate, levels = c("Non-cognate", "Cognate"))) %>%
-    arrange(item, meaning, age_bin) 
-
-#### extract midpoints ##############################################
-midpoints <- fit1 %>%
-    spread_draws(b_mid_Intercept, b_mid_item_dominance1, b_mid_frequency, r_meaning__mid[meaning, term]) %>% 
-    mutate(mid_L1 = b_mid_Intercept + r_meaning__mid + 0.5*b_mid_item_dominance1 + b_mid_frequency,
-           mid_L2 = b_mid_Intercept + r_meaning__mid + -0.5*b_mid_item_dominance1 + b_mid_frequency) %>% 
-    median_qi(mid_L1, mid_L2) %>%
-    select(-c(term, .interval, .point, .width)) %>%
-    ungroup() %>% 
-    right_join(dat, by = "meaning") %>%
-    rowwise() %>% 
-    mutate(mid_diff = mid_L2 - mid_L1) %>% 
-    ungroup() %>% 
[…]

### 2020-08-22 Code/R/preprocessing.R

@@ -1,6 +1,6 @@
-#### preprocessing #########################
+#### Preprocessing: import, clean, and tidy data #########################
 
-#### set up ################################
+#### set up ##############################################################
 
 # load packages
 library(dplyr)
@@ -8,6 +8,7 @@ library(tidyr)
 library(stringr)
 library(lubridate)
 library(readxl)
+library(janitor)
 library(data.table)
 library(mice)
 library(here)
@@ -52,12 +53,19 @@ participants <- list.files(here("Data", "Logs"), pattern = "logs", full.names = 
     filter(completed, age_bin %in% bins_interest) 
 
 # export data
-fwrite(participants, here("Data", "01_participants.csv"), sep = ",", dec = ".", row.names = FALSE)
+fwrite(participants, here("Data", "participants.csv"), sep = ",", dec = ".", row.names = FALSE)
 
-#### process responses #######################
-pool <- read_xlsx(here("Data","01_pool.xlsx")) %>% 
-    rename(te = meaning)
 
+#### items data ############################################################
+pool <- read_xlsx(here("Data", "01_pool.xlsx"), na = c("", "NA")) %>%
+    rename(te = meaning) %>% 
+    mutate(frequency = as.numeric(frequency),
+           te = as.character(te),
+           cognate_rater1 = ifelse(cognate_rater1==1, "Cognate", "Non-cognate")) %>%
+    drop_na(language, cognate_rater1)
+
+
+#### process responses #####################################################
 responses <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactors = FALSE, na.strings = "") %>%
     as_tibble() %>%
     mutate_at(vars(time_stamp), as_date) %>%
@@ -71,7 +79,6 @@ responses <- fread(here("Data", "02_merged.csv"), header = TRUE, stringsAsFactor
                                    TRUE ~ as.numeric(doe_catalan)),
            doe_spanish = case_when(id_db %in% "57046" ~ 50,
                                    TRUE ~ as.numeric(doe_spanish))) %>% 
-    mutate(version = factor(version, levels = c("CBC", "BL-Short-A", "BL-Short-B", "BL-Short-C", "BL-Short-D", "BL-Long-1", "BL-Long-2", "DevLex"), ordered = TRUE)) %>%
     drop_na(response) %>%
[…]

### 2020-08-22 gca_freq.R

@@ -1,233 +1,0 @@
-#### Analysis: Analyse vocabulary data ##########################################
-
-#### set up #####################################################################
-
-# load packages
-library(dplyr)
-library(tibble)
-library(tidyr)
-library(stringr)
-library(purrr)
-library(ggplot2)
-library(patchwork)
-library(ggdist)
-library(broom.mixed)
-library(data.table)
-library(nlme)
-library(lme4)
-library(bayestestR)
-library(BayesFactor)
-library(janitor)
-library(here)
-
-# create/load functions
-source(here("Code", "R", "functions.R"))
-
-# set params
-set.seed(888)
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-
-##### import data and set params #######################################################################################3
-dat_comp <- fread(here("Data", "04_prepared.csv")) %>%
-    as_tibble() %>% 
-    filter(type=="Comprehensive") %>% 
-    select(item, type, te, category, age_bin, item_dominance, cognate, frequency, n, successes, proportion, probability, logodds, weights) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1) %>% 
-    mutate_at(vars(cognate, item_dominance), as.factor) %>% 
-    arrange(item, te, age_bin) %>%
-    groupedData(proportion ~ 1|te, .)
-
-dat_prod <- fread(here("Data", "04_prepared.csv")) %>%
-    as_tibble() %>% 
-    filter(type=="Productive") %>% 
-    select(item, type, te, category, age_bin, item_dominance, cognate, frequency, n, successes, proportion, probability, logodds, weights) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1) %>% 
-    mutate_at(vars(cognate, item_dominance), as.factor) %>% 
-    arrange(item, te, age_bin) %>% 
-    groupedData(proportion ~ 1|te, .)
-
-contrasts(dat_comp$cognate) <- contr.sum(c("Non-cognate", "Cognate"))/2
-contrasts(dat_comp$item_dominance) <- contr.sum(c("L2", "L1"))/2
[…]

### 2020-11-28 Code/R/preprocessing.R
Ordinal regression in brms

@@ -1,74 +1,0 @@
-#### Preprocessing: import, clean, and tidy data #########################
-
-#### set up ##############################################################
-
-# load packages
-library(tidyverse)
-library(multilex)
-library(data.table)
-library(janitor)
-library(here)
-
-# set params
-set.seed(888)
-bins <- c("< 10", "10-12", "12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36", "36-38", "38-40", "> 40")
-bins_interest <- bins[-c(1, length(bins))]
-breaks <- c(0, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 100)
-
-#### import data ######################
-participants_raw <- ml_participants("gonzalo.garciadecastro@upf.edu")
-responses_raw <- ml_responses(participants = participants_raw,
-                              formr_email = "gonzalo.garciadecastro@upf.edu")
-logs_raw <- ml_logs(responses = responses_raw, participants = participants_raw) %>% 
-    filter(age_bin %in% bins_interest,
-           lp %in% c("Monolingual", "Bilingual")) 
-
-#### process responses #####################################################
-responses <- responses_raw %>% 
-    select(id, time, item, response, language) %>% 
-    left_join(logs_raw, by = c("id", "time")) %>% 
-    drop_na(response) %>% 
-    mutate(age_bin = factor(cut(age, breaks = breaks, labels = bins), levels = bins, ordered = TRUE),
-           item_dominance = ifelse(language==dominance, "L1", "L2"),
-           bilingualism = ifelse(dominance=="Catalan", doe_spanish, doe_catalan)) %>% 
-    select(id, age_bin, age, sex, dominance, bilingualism, time, item, language, item_dominance, response) %>% 
-    left_join(pool, by = c("item", "language")) %>% 
-    select(id, time, age_bin, age, sex, dominance, bilingualism, time, te, item, cognate, language, item_dominance, response) %>% 
-    mutate(cognate = case_when(cognate ~ "Cognate",
-                               !cognate ~ "Non-cognate",
-                               TRUE ~ NA_character_))
-
-fwrite(responses, here("Data", "responses.csv"), sep = ",", dec = ".", row.names = FALSE)
-
-#### vocabulary sizes ###########################################################################
-
-vocabulary <- responses %>% 
[…]

### 2020-11-28 Python/vocabulary.py

@@ -1,0 +1,7 @@
+from pandas import read_csv
+import os
+import pystan
+
+
+dat = read_csv('Data/vocabulary.csv')
+fit0_stan = pystan.StanModel(file = 'Stan/vocabulary0.stan')

### 2020-11-28 Code/R/vocabulary.R
Ordinal regression in brms

@@ -1,145 +1,0 @@
-#### vocabulary: Analyse vocabulary sizes ######################################
-
-#### set up ####################################################################
-
-# load packages
-library(tidyverse)
-library(data.table)
-library(brms)
-library(tidybayes)
-library(here)
-
-# set params
-set.seed(888)
-
-#### import data ###############################################################
-dat <- fread(here("Data", "vocabulary.csv")) %>% 
-    as_tibble() %>% 
-    mutate_at(vars(type, lp, item_dominance), as.factor) %>% 
-    mutate(age_center = age-mean(.$age)) %>% 
-    filter(type %in% "Comprehensive")
-
-contrasts(dat$lp) <- c(0.5, -0.5)
-contrasts(dat$item_dominance) <- c(0.5, -0.5)
-
-#### fit models ################################################################
-priors <- c(prior(normal(0, 1), class = "Intercept"),
-            prior(normal(0.5, 0.4), class = "b", coef = "age_center"))
-
-fitp <- brm(sum | trials(n) ~ age_center,
-            data = dat,
-            family = binomial("logit"),
-            prior = priors,
-            sample_prior = "only",
-            file = here("Results", "vocabulary_p.rds"),
-            cores = 4)
-
-fit0 <- update(fitp, sample_prior = "no",
-               newdata = dat, cores = 4,
-               save_all_pars = TRUE, file = here("Results", "vocabulary_0.rds"))
-fit1 <- update(fit0, . ~ . -age_center + age_center*item_dominance,
-               newdata = dat, cores = 4,
-               save_all_pars = TRUE, file = here("Results", "vocabulary_1.rds"))
-fit2 <- update(fit1, . ~ . -age_center*item_dominance + age_center*item_dominance*lp,
-               newdata = dat, cores = 4,
-               save_all_pars = TRUE, file = here("Results", "vocabulary_2.rds"))
-
-#### model comparison ##########################################################
-loo0 <- loo(fit0, moment_match = TRUE)
-loo1 <- loo(fit1, moment_match = TRUE)
-loo2 <- loo(fit2, moment_match = TRUE)
-loo_comp <- loo_compare(loo0, loo1, loo2)
-
-#### examine prior #############################################################
-prior_dist <- gather_draws(fitp, `b_.*`, regex = TRUE) %>% 
-    mutate(.chain = as.factor(.chain))
-
-# traceplots
[…]

### 2020-11-28 R/priors.R

@@ -34,7 +34,7 @@ dat <- list.files(here("Data", "Wordbank"), full.names = TRUE) %>%
     filter(num_item_id %in% sample(unique(.$num_item_id), size = 10))
 
 #### fit model #################################################################
-fit <- brm(successes | trials(n)  ~ age_bin + (1 + age_bin | language:num_item_id),
+fit <- brm(successes | trials(n)  ~ age_bin + (1 + age_bin | language),
            family = binomial("logit"),
            chains = 1,
            file = here("Results", "fitp_wordbank.rds"),
@@ -50,18 +50,21 @@ post_preds_fixed <- expand.grid(age_bin = seq(min(dat$age_bin), max(dat$age_bin)
                           n = 1) %>% 
     add_fitted_draws(fit, n = 10, re_formula = NA) 
 post_preds_group <- expand.grid(age_bin = seq(min(dat$age_bin), max(dat$age_bin), by = 0.1),
-                          num_item_id = unique(dat$num_item_id),
+                                language = unique(dat$language),
                           n = 1) %>% 
     left_join(distinct(dat, language, num_item_id)) %>% 
     add_fitted_draws(fit, n = 10) %>% 
     mutate(num_item_id = as.factor(num_item_id))
-ggplot(post_preds_group, aes(age_bin, .value)) +
-    facet_wrap(~language) +
-    geom_line(aes(group = interaction(.draw, num_item_id), colour = num_item_id),
-              size = 0.25, alpha = 0.5, show.legend = FALSE) +
+ggplot(post_preds_group, aes(age_bin, .value, colour = language)) +
+    stat_summary(data = post_preds_group, fun = "median", geom = "line", size = 0.5, alpha = 0.75) +
     stat_lineribbon(data = post_preds_fixed, .width = 0.95, alpha = 0.5, size = 0,
-                    colour = "black", show.legend = FALSE) +
-    stat_summary(data = post_preds_fixed, fun = "median", geom = "line", size = 1)
+                    colour = "black", fill = "black", show.legend = FALSE) +
+    stat_summary(data = post_preds_fixed, fun = "median", geom = "line", size = 1, colour = "black") +
+    labs(x = "Age (months)", y = "p(y|age)", colour = "Language") +
+    scale_color_viridis_d() +
+    theme_minimal() +
+    theme(axis.title = element_text(face = "bold"),
+          legend.title = element_text(face = "bold"))
 
[…]

### 2020-12-07 Rmd/02_results.Rmd
Prepare introduction

--- 2020-08-22 Manuscript/trajectories_manuscript.Rmd
@@
 ---
-title             : "The role of cross-linguistic lexical similarity on bilingual word acquisition"
-shorttitle        : "Cross-linguistic similarity and  word acquisition"
-
-author: 
-  - name          : "Gonzalo García-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Daniela Avila-Varela"
-    affiliation   : "1"
-  - name          : "Núria Sebastian-Galles"
-    affiliation   : "1"
-
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
-
-authornote: |
-  Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
-
-abstract: |
-  Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "lexical acquisition, vocabulary, bilingualism"
-wordcount         : "X"
-
-bibliography      : ["../References/trajectories.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-keep_text         : false
-class             : "man"
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  papaja::apa6_pdf:
-    latex_engine  : "xelatex"
-
+title: "Results"
+output: html_document
 ---
 
-```{r setup, include = FALSE}
-
-knitr::opts_chunk$set(
-    echo = FALSE,
-    fig.align = "center",
-    message = FALSE,
-    warning = FALSE,
-    cache.extra = knitr::rand_seed,
-    out.width = "80%",
-    results = "asis"
-)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
-
-```{r packages, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# load packages
-library(papaja)
-library(knitr)
-library(magrittr)
-library(dplyr)
-library(tidyr)
-library(tibble)
-library(ggplot2)
-library(forcats)
-library(readxl)
-library(lubridate)
-library(janitor)
-library(nlme)
-library(broom.mixed)
-library(modelr)
-library(gt)
-library(data.table)
-library(here)
-
-```
-
-```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# set params
-set.seed(888)
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-bins_expanded <- as.character(seq(1, 41, by = 2))
-
-# import data 
-dat_studies <- fread(here("Data", "00_studies.csv"), na.strings = c("", "NA")) %>%
-  distinct(q_version, language, q_items)
-
-dat_items <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
-    group_by() %>%
-    drop_na(language) %>%
-    mutate_at(vars(A, B, C, D, cognate_rater1, include, frequency), as.numeric) %>%
-    mutate_at(vars(A, B, C, D, cognate_rater1, include), as.logical)
-
-dat_participants <- fread(here("Data", "participants.csv"), na.strings = c("", "NA"))
-
-dat_raw <- fread(here("Data", "raw.csv"), na.strings = c("", "NA")) %>%
-    filter(lp %in% c("Bilingual"),
-           age_bin %in% bins_interest) %>%
-    as_tibble() %>%
-    mutate_at(vars(date_birth, time_stamp), as_date)
-
-dat <- fread(here("Data", "preprocessed.csv"), na.string = c("", "NA")) %>% 
-    as_tibble() %>% 
-    select(item, type, te, category, age_bin, item_dominance, cognate, frequency, n, successes, proportion) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1) %>% 
-    mutate_at(vars(frequency), function(x) scale(x, center = TRUE, scale = FALSE)[,1]) %>% 
-    mutate_at(vars(cognate, item_dominance), as.factor) %>% 
-    arrange(type, item, te, age_bin)
-
-```
-
-```{r analyses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, cache=TRUE}
-
-# run the analyses (see analysis.R to check to code)
-# running this code chunk may take a few seconds
-source(here("Code", "R", "analysis.R"), local = knitr::knit_global())
-
-```
-
-# Introduction
-
-Bilingual toddlers must learn two distinct set of words (one for each language) that partially overlap in sound and meaning.
-Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When taking both languages into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev1977; @doyle1977some; @rosenblum1983word; @fernandez1992; @pearson1994; @bialystok2010; @hoff2012; @core2013; @hoff2014; @blom2019; but see @pearson1993; @houwer2014].
-Not all bilinguals show a similar developmental trajectory of lexical acquisition.
-Recent studies have capitalised on the role of the similarity between the specific pair of languages the infant is learning.
-One way of measuring language similarity is by estimating the amount of cognates (i.e., form-similar translation equivalents) they share.
-@floccia2018 explored the impact of this form of similarity on vocabulary sizes from a large sample of 24-month-old toddlers learning English and an additional language.
-Toddlers learning languages that shared a large amount of cognates (e.g., English-German) showed larger vocabulary sizes than those learning languages that shared fewer cognates (e.g., English-Spanish).
-@blom2019 extended these findings to children aged three to 10 years.
-These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical acquisition.
-
-How does cognateness influence word learning?
-@floccia2018 point to the possibility that parallel activation during speech perception may play a role in this cognate facilitation effect.
-The parallel activation principle suggests that bilinguals activate both languages in parallel during speech perception and production, even when only one of them is being used.
-This hypothesis can only be understood under the more general account of language non-selectivity of lexical processing, which states that translation equivalents are co-activated at the lexical level [e.g., @de1991lexical; @costa2000; @dijkstra2005; @thierry2007; @hoshino2008; @yudes2010; @bobb2020; see @kroll2013 for review].
-A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model [@kroll1994; @kroll2010], the Bilingual Interactive Activation model [BIA; @dijkstra1998bia], and its successor BIA+ [@dijkstra2002architecture], the Bilingual Language Interaction Network for Comprehension of Speech model [BLINCS; @shook2013] and, more recently, Multilink [@dijkstra2019], which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
-However, these models do not address word learning during language acquisition.
-
-It is feasible that parallel activation plays a role in vocabulary acquisition during the first years of life.
-On top of Floccia et al.'s findings, other studies have provided experimental evidence of parallel activation during speech perception in bilingual toddlers [@vonholzen2012; @vonholzen2019].
-Yet, it is unclear how parallel activation influences word learning at these ages.
-What mechanisms are engaged, and how do they relate to vocabulary size?
-We contemplate two scenarios.
-First, it is possible that the reported "similarity boost" on vocabulary size is not directly related to the amount of cognates shared between both languages, but rather to the overlap across their correspondent phonological inventories.
-One of the measures of language similarity Floccia et al.
-computed is the mean phonological Levenshtein distance across translation equivalents.
-This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
-These scores were aggregated across pairs of translation equivalents, obtaining a single similarity estimate of for each language pair.
-This measure embeds confounded information about the "degree of cognateness" and the amount of phonological overlap between both languages.
-Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, English-Dutch), with few exceptions (e.g., Spanish-Basque).
-In the light of these data, it is not possible to discern whether vocabulary size is increased by the earlier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
-We address this issue by testing two hypothesis.
-
-## Hypothesis 1: Cognates are acquired earlier than non-cognates
-
-First, we test whether the increase in vocabulary size reported by Floccia et al.
-for languages sharing a greater amount of cognates is indeed caused by cognates being acquired earlier.
-In line with this account, @bosch2014 found higher productive vocabulary scores in 18 month-old Catalan-Spanish bilinguals, compared to monolinguals, due to the higher presence of cognates their lexicon.
-We will adopt a similar approach to @braginsky2019, who modelled cross-sectional data from individual items to explore what properties were associated with an earlier acquisition.
-We will analyse the acquisition trajectories of a selection of `r dat_items %>% filter(include) %>% group_by(language) %>% summarise(n = n(), .groups = "drop") %>% pull(n) %>% max()` translation equivalents in a cross-section sample of Catalan/Spanish bilinguals aged 14 to 34 months of age.
-This approach allows to analyse how the proportion of infants that know each item changes across ages, providing a more suitable framework to put developmental hypotheses to test, instead of taking a "snapshot" of the state of the lexical of children at a single age.
-
-## Hypothesis 2: Cognates are acquired closer in times than non-cognates
-
-One mechanism that could acccount for an ealier acquisition of cognates is that parallel activation makes word-forms acquired in one language available during the acquisition of word forms in the other language.
-One consequence of this situation is that translation equivalents scaffold the acquisition of each other.
-This would point to different mechanisms engaged in the acquisition of cognates vs.
-non-cognates.
-While the phonology of cognate translation equivalents overlaps, and thus could be co-activated, the phonology of non-cognates does it in a much lesser degree.
-Consider the case of a Catalan-Spanish bilingual toddler that has already acquired the Catalan words *porta* (/ˈpɔr.tə/, door) and *taula* (/ˈtaw.ɫə/, table), but not yet their correspondent translation equivalents in Spanish (*puerta*, /ˈpwer.ta/; *mesa* /ˈme.sa/).
-According to our hypothesis, hearing the word *puerta* in presence of a door should lead to the activation of *porta*, which should facilitate the acquisition of the former.
-On the contrary, when hearing *mesa* in presence of a table, it is unlikely that the phonology of *taula* will be activated.
-This mechanism should lead to cognate translation equivalents being learnt closer in time than non-cognate translation equivalents.
-To test this hypothesis, we will extract the age of acquisition of each word-form in the questionnaire (as estimated by the model used to test Hypothesis 1), and will calculate the difference in time between the acquisition of such word-forms and their translation equivalents.
-We will test how cognateness affects such outcome.
-
-# Method
-
-## Participants
-
-We collected data from `r dat_raw$id_db %>% unique %>% length() %>% printnum()` bilinguals from the Metropolitan Area of Barcelona, between `r min(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`.
-All families participated voluntarily.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-
-We assessed toddlers' language background asking parents for an estimated proportion of exposure to each language.
-We defined participants as Spanish- or Catalan-dominant according to the language they were exposed the most to.
-Participants were considered bilingual - therefore included - if exposed 50-90% of the time to the non-dominant language.
-We excluded participants with >10% exposure to a third language.
-Fig.
-\@ref(fig:lpdist) shows how participants were distributed across age bins and language profiles.
-
-```{r lpdist, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_participants %>%
-  filter(lp %in% "Bilingual") %>% 
-    mutate(bilingualism = cut_interval(bilingualism, n = 6,
-                                       labels = paste0(seq(0, 50, by = 10), "%"))) %>%
-    ggplot(aes(x = age_bin, y = bilingualism)) +
-    stat_bin_2d(colour = "white") +
-  scale_x_discrete(guide = guide_axis(n.dodge = 1)) +
-    scale_fill_viridis_c(option = "plasma") +
-    labs(x = "Age (months)", y = "L2 exposure", fill = "Participants") +
-    theme_classic() +
-    theme(axis.text = element_text(colour = "black"))
-ggsave(here("Figures", "participants.png"), height = 4, width = 5)
-```
-
-## Questionnaire
-
-Parents completed an on-line questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
-We implemented the questionnaire in `formr` [@arslan2020], a free, open-source platform.
-Participants were randomly allocated into one of four list of items.
-Each list contained a distinct inventory of translation equivalents.
-We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/funcional categories:  Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-Those categories that did not entail more than 16 items - thus resulting in less than four items after dividing it in four lists - were not divided: all of their items were included in the four lists.
-Another subset of items that were part of the trial lsits of some experiments in the lab were also included in all versions.
-
-Each item of the inventory presented a word-form, to which parents were instructed to report whether the toddlers was able to understand it, understand ans say it, or neither.
-The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates) .
-Items of one language were translation equivalents from the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-
-## Data analysis
-
-For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item in the non-dominant language within age bins of 2 months.
-Our model assumes that this proportion follows a logistic curve across ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the degree of exposure to the non-dominant language (*bilingualism*, 0-50%), the cognate status of the word (*cognateness*, Non-cognate vs.
-Cognate), and their interaction.
-Our prediction is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words, and that this effect will be larger for more balanced bilinguals.
-
-Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
-Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item *porta* (*door*, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
-Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
-
-```{r itemsinfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_items %>%
-    drop_na(cognate_rater1) %>%
-    filter(include) %>%
-    select(A, B, C, D, cognate_rater1, frequency, language) %>% 
-    mutate(cognate_rater1 = ifelse(cognate_rater1, "Cognate", "Non-cognate")) %>%
-    pivot_longer(c(A, B, C, D), names_to = c("list"), values_to = "n") %>% 
-    filter(n) %>% 
-    group_by(list, language, cognate_rater1) %>% 
-    summarise(n = sum(n, na.rm = TRUE),
-              mean = mean(frequency, na.rm = TRUE),
-              sd = sd(frequency, na.rm = TRUE),
-              min = min(frequency, na.rm = TRUE),
-              max = max(frequency, na.rm = TRUE),
-              .groups = "drop") %>%
-    ungroup() %>% 
-    rename(cognateness = cognate_rater1) %>%
-    relocate(list, language, cognateness, n, mean, sd, min, max) %>%
-    arrange(list, language, cognateness) %>%
-    mutate(list = paste("List", list)) %>% 
-    pivot_wider(names_from = cognateness, values_from = c(n, mean, sd, min, max)) %>%
-    clean_names() %>%
-    gt(groupname_col = c("language")) %>% 
-    fmt_number(columns = contains("cognate"), decimals = 2) %>%
-    fmt_number(columns = starts_with("n_"), decimals = 0) %>%
-    tab_header(title = md("**Questionnaire and item properties**"),
-               subtitle = "Participants were randomly allocated on of the the four lists, and completed both the Catalan and the Spanish versions.") %>% 
-    tab_spanner(label = "Frequency (Cognates)", columns = vars("n_cognate", "mean_cognate", "sd_cognate", "min_cognate", "max_cognate")) %>% 
-    tab_spanner(label = "Frequency (Non-cognates)", columns = contains("non_cognate")) %>% 
-    cols_label(list = md("**List**"), 
-               n_cognate = md("*N*"),
-               mean_cognate = md("*Mean*"),
-               sd_cognate = md("*SD*"),
-               min_cognate = md("*Min*"),
-               max_cognate = md("*Max*"),
-               n_non_cognate = md("*N*"),
-               mean_non_cognate = md("*Mean*"),
-               sd_non_cognate = md("*SD*"),
-               min_non_cognate = md("*Min*"),
-               max_non_cognate = md("*Max*"))
-
-```
-
-We fit two separate logistic models on comprehensive and productive data, respectively.
-Logistic curves are characterised by three parameters[see @mahr2019 and @mahr2020 for an excellent introduction to this approach, in which we based this analysis]: (a) an *asymptote* (`asym`, upper boundary  of the proportion of infants that have acquired the item), (b) the *steepness* (`steep`, how fast the a given item is acquired by infants), and (c) a *mid-point* (`mid`, the age at which steepness is maximum).
-We work on the assumption that the trajectory of acquisition of items follows a logistic curve shape [@mayor2011], and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
-This approach allows us to model and estimate the age of acquisition of items explicitly.
-Our model estimated the asymptote and steepness as grand means across items (i.e., intercepts).
-Mid-points were estimated by a linear regression model that included our predictors of interest: `age`, `item_dominance` (sum-coded, L2 = -0.5, L1 = 0.5), `cognateness` (sum-coded, Non-cognate = -0.5, Cognate = 0.5).
-We also included random intercepts and `item dominance` slopes for each translation equivalent (`te`) [@barr2013].
-Appendix 1 shows a detailed description of the model.
-
-We estimated the parameters of our model using the Bayesian framework.
-Bayesian models use the Bayes rule to combine prior knowledge about the distribution of a given parameter, and the likelihood of each of the values it can take given the data.
-We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our prior).
-In order to test the contribution of each predictor on the fit of the model, we started fitting a null model that only included an global intercept and random intercepts for each TE, then fitted a model that also included *bilingualism* and its random slopes by TE, and finally a model that also included *cognateness* and its interaction with *bilingualism*, as well as random slopes for all the fixed effects.
-We compared how the fit of the model changed in every step [@schad2020] using leave-one-out cross validation.
-We fit both the null and alternative models using the R environment [@rcoreteam2013; @rstudioteam2015] and the `brms` package [@burkner2017], which relies on the probabilistic language Stan [@carpenter2017] to approximate posterior distributions.
-We used Pareto-smoothed importance sampling [PSIS; @vehtari2017; @vehtari2019] to compare the null and extended models.
-Data and model results were processed and visualised using the `tidyverse` family of R packages [@wickham2019] and the `tidybayes` package [@kay2020].
-
-To test Hypothesis 2, we run a Bayesian ANOVA using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included *cognateness* as predictor.
-Using the `BayesFactor` R package [@morey2018a], we computed a Bayes factor (*BF*) that compared the likelihood of a linear model that included *cognateness* as a predictor against the likelihood of a linear model that did not, under the light of our data.
 
@@
 ```
-
-# Discussion
-
-# Appendix
-
-## Appendix 1: Questionnaires
-
-## Appendix 2: Model details
-
-## Appendix 3: Session info
-
-```{r sessioninfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-sessionInfo()
-
-```
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-
-\endgroup
-
-```{r create_r-references, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-r_refs(file = "../References/trajectories.bib")
-
-```

### 2020-12-07 Rmd/00_introduction.Rmd
Prepare introduction

---
title: "Introduction"
output: 
    pdf_document:
        latex_engine: "xelatex"
    word_document:
        reference_docx: "trajectories_reference.docx"
bibliography: "trajectories.bib"
csl: "apa6.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# load packages
library(knitr)
library(tidyverse)

```

Lexical acquisition is the process by which children form representations for new words in their lexicon. This involves two main events: the encoding of a (modality-specific) word form (e.g, phonology, orthography, signed form) and its association with the semantic features of its referent [@caramazza1997]. Lexical acquisition starts at early ages: infants as young as 6 months of age look longer to a named picture than to an unnamed picture presented next to it, demonstrating both (a) recognition of a word-form and (b) association with its (visual) referent [@bergleson2012]. 
At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered [@jusczyk1995], as opposed to when only part of the word is uttered. At 12 months, infants show sensitivity to both vowel and consonant mispronunciations [@mani2010a], looking shorter to named pictures when their label is mispronounced. Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones. One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon [@mcclelland2003].

Words are not learnt in isolation, but rather in contexts rich in speech tokens and referents. The structure of toddlers' lexicon reflects this high connectivity [@wojcik2018; @hills2010; @fourtassi2020]. At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing [@mani2010b; @mani2011]. The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming  [@styles2009; @arias-trejo2013; @wojcik2013], and show evidence of inhibitory links between lexical-semantic entries [@chow2016]. By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages [@chow2017]. Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words [@chow2017; @rama2013], and  the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers. This suggests that the emergence of rich lexical networks -- both at the phonological and the semantic levels -- grows as a function of the number of words acquired by the toddler.

The structure of the lexicon also impacts the order of acquisition of new words. Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills2009]. @borosky2016 and @borovsky2018 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired. This points to the structure of the lexicon facilitating the strategies children engage during word learning.

The case of bilingual children (here defined as those learning two languages from birth) presents an opportunity to study how lexical acquisition takes placed under a more complex environment, as they  face the challenge of learning two distinct sets of words -- one for each language -- that partially overlap in sound and meaning.

EARLY STAGES IN BILINGUAL LANGUAGE ACQUISITION

When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals). When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev_influence_1977; @doyle1977some; @rosenblum1983word; @fernandez1992bilingual; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].

The vocabulary size of

Previous studies have highlighted

Previous studies point to bilinguals' vocabulary size as a function of the degree

@floccia_i_2018, gathered parental vocabulary estimates [@hamilton2000infant] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages). They computed the average phonological similarity between translation equivalents (TEs from now on; e.g.,  *table* in English and *tafel* in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language. Using a similar measure of lexical similarity, @blom_cross-language_2019 extended these results to bilingual children aged three to 10, and reported a positive association between the lexical distance and comprehensive vocabulary size: bilinguals learning lexically close languages showed similar vocabulary sizes to those of monolinguals, but those learning the more distant languages showed lower vocabulary sizes.

These results suggest that linguistic distance between languages plays an important role during bilingual lexical acquisition, and that this effect operates within TEs. What mechanisms underlie this effect, and why they seem to operate differently across comprehension and production, or across the dominant and the non-dominant language, remain open issues. @floccia_i_2018 pointed to parallel activation as a candidate mechanism behind this facilitation effect. The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., *cat*, in English) its corresponding translation in the other language is activated too [e.g., *gato*, in Spanish. There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell1991mediated; @spivey1999cross; @costa2000lexical; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh2014one; @von_holzen_impact_2019; @bobb_co-activation_2020; see @costa2005facilitatory for review], and children [e.g. @von_holzen_language_2012; @poulin2013lexical; @jardak2019labels; @bosma2020cognate]. A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.

@floccia_i_2018 argue that the degree of phonological similarity between TEs (*cognateness* from now on) of both languages should lead to increased parallel activation of both languages during language exposure, facilitating the acquisition of words in both languages, and leading participants leading two lexically close languages to show larger vocabulary sizes than those learning two lexically distant languages. This account, however, neglects the fact that for parallel activation to take place and play a facilitation role, lexical representations of both pairs of the TE must have been already established in the lexicon, and their corresponding phonological forms must have been encoded. This is not the case during lexical acquisition, where at least one of the members of the pair has not being acquired yet. A more lenient definition of parallel activation could adjust to the context of lexical acquisition more easily: when only one of the members of the TE has been acquired, its lexical representation is activated when, in the presence of its referent, a phonologically similar label is uttered. Consider the case of a Catalan-Spanish bilingual that has already learned the word *gat* (Catalan for *cat*), but not *gato* (in Spanish). It is possible that, when presented with *gato* in the presence of a cat, she maps this novel label to the familiar word *gat*, and creates a lexical representation for *gato* as a synonym for *gat*. The case of a non-cognate is differing. Consider now the case of a toddler that has already learnt *gos* (Catalan for *dog*), but not *perro* (Spanish). In this case, it would not be possible to map both labels via phonology, give their lack of phonological similarity.

Under this account, cognates (i.e., phonologically similar TEs) would be acquired earlier than non-cognates, but this effect would only play a role once one of the members of the TE has been acquired. The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse. @schelletter2002effect's reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE. @bosch_first_2014, showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months. The low number of words included in the analyses of both studies, and the lack of adjustment for lexical frequency, limits severely the strength of any conclusions concerning the difference in age of acquisition of cognates and non-cognates. On the other hand @floccia_i_2018 worked with aggregated estimates of participants' vocabulary size, losing item-level information that could provide information about change in age of acquisition of TEs associated with their degree of phonological similarity. Therefore, their results can only be considered as compatible with cognateness playing a role on age of acquisition. In this study, we aimed at overcoming these pitfalls by using unaggregated data from individual responses, testing the role of cognateness explicitly and adjusting for lexical frequency.

It also follows from the fact that cognateness operates *within* TEs that the effect of cognateness can only play a role once one of the members of the TE has been acquired, and its phonological form is available during the acquisition of its translation. Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language [@floccia_i_2018]. @dijkstra2013 tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE. Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian. This suggests that infants that were mostly exposed to Dutch used the phonology of  dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
A study by @dijkstra2013 showed that the degree of phonological similarity between 
the performance of  was better for cognates than for non-cognates, only for those children

to have a larger compr Therefore, we predicted that the effect of cognateness should, on average, impact more strongly the acquisition of words in the non-dominant language, relative to words in the dominant language.

Finally, previous studies have reported that the properties that describe the form of a word play a stronger role in production than comprehension than in production. This is the case of the number of phonemes [@braginsky_consistency_2019], phonological neighbourhood density [@jones_children_2019]. These results converge with @floccia_i_2018's finding that the effect of the average cognateness between languages was larger in production than in comprehension^[Although @bosch_first_2014 found that participants in their sample produced cognates earlier than non-cognates, no analyses were conducted on comprehensive data, and therefore they cannot be taken as supporting evidence for a larger effect of cognateness on production than in comprehension].  Accordingly, we expected the role of cognateness to be more central in production than in comprehension.

In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.

### 2020-12-07 Rmd/01_method.Rmd
Prepare introduction

--- 2020-08-22 Manuscript/trajectories_manuscript.Rmd
@@
 ---
-title             : "The role of cross-linguistic lexical similarity on bilingual word acquisition"
-shorttitle        : "Cross-linguistic similarity and  word acquisition"
-
-author: 
-  - name          : "Gonzalo García-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Daniela Avila-Varela"
-    affiliation   : "1"
-  - name          : "Núria Sebastian-Galles"
-    affiliation   : "1"
-
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
-
-authornote: |
-  Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
-
-abstract: |
-  Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "lexical acquisition, vocabulary, bilingualism"
-wordcount         : "X"
-
-bibliography      : ["../References/trajectories.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-keep_text         : false
-class             : "man"
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  papaja::apa6_pdf:
-    latex_engine  : "xelatex"
-
+title: "Method"
+output: html_document
 ---
 
-```{r setup, include = FALSE}
-
-knitr::opts_chunk$set(
-    echo = FALSE,
-    fig.align = "center",
-    message = FALSE,
-    warning = FALSE,
-    cache.extra = knitr::rand_seed,
-    out.width = "80%",
-    results = "asis"
-)
-```
-
-```{r packages, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# load packages
-library(papaja)
-library(knitr)
-library(magrittr)
-library(dplyr)
-library(tidyr)
-library(tibble)
-library(ggplot2)
-library(forcats)
-library(readxl)
-library(lubridate)
-library(janitor)
-library(nlme)
-library(broom.mixed)
-library(modelr)
-library(gt)
-library(data.table)
-library(here)
-
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 
-# set params
-set.seed(888)
-bins_interest <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-bins_expanded <- as.character(seq(1, 41, by = 2))
-
-# import data 
-dat_studies <- fread(here("Data", "00_studies.csv"), na.strings = c("", "NA")) %>%
-  distinct(q_version, language, q_items)
-
-dat_items <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
-    group_by() %>%
-    drop_na(language) %>%
-    mutate_at(vars(A, B, C, D, cognate_rater1, include, frequency), as.numeric) %>%
-    mutate_at(vars(A, B, C, D, cognate_rater1, include), as.logical)
-
-dat_participants <- fread(here("Data", "participants.csv"), na.strings = c("", "NA"))
-
-dat_raw <- fread(here("Data", "raw.csv"), na.strings = c("", "NA")) %>%
-    filter(lp %in% c("Bilingual"),
-           age_bin %in% bins_interest) %>%
-    as_tibble() %>%
-    mutate_at(vars(date_birth, time_stamp), as_date)
-
-dat <- fread(here("Data", "preprocessed.csv"), na.string = c("", "NA")) %>% 
-    as_tibble() %>% 
-    select(item, type, te, category, age_bin, item_dominance, cognate, frequency, n, successes, proportion) %>%
-    mutate(age_bin = as.numeric(factor(age_bin, levels = bins_interest, ordered = TRUE))-1) %>% 
-    mutate_at(vars(frequency), function(x) scale(x, center = TRUE, scale = FALSE)[,1]) %>% 
-    mutate_at(vars(cognate, item_dominance), as.factor) %>% 
-    arrange(type, item, te, age_bin)
+library(tidyverse)
+library(papaja)
+library(gt)
+library(knitr)
 
 ```
-
-```{r analyses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, cache=TRUE}
-
-# run the analyses (see analysis.R to check to code)
-# running this code chunk may take a few seconds
-source(here("Code", "R", "analysis.R"), local = knitr::knit_global())
-
-```
-
-# Introduction
-
-Bilingual toddlers must learn two distinct set of words (one for each language) that partially overlap in sound and meaning.
-Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When taking both languages into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev1977; @doyle1977some; @rosenblum1983word; @fernandez1992; @pearson1994; @bialystok2010; @hoff2012; @core2013; @hoff2014; @blom2019; but see @pearson1993; @houwer2014].
-Not all bilinguals show a similar developmental trajectory of lexical acquisition.
-Recent studies have capitalised on the role of the similarity between the specific pair of languages the infant is learning.
-One way of measuring language similarity is by estimating the amount of cognates (i.e., form-similar translation equivalents) they share.
-@floccia2018 explored the impact of this form of similarity on vocabulary sizes from a large sample of 24-month-old toddlers learning English and an additional language.
-Toddlers learning languages that shared a large amount of cognates (e.g., English-German) showed larger vocabulary sizes than those learning languages that shared fewer cognates (e.g., English-Spanish).
-@blom2019 extended these findings to children aged three to 10 years.
-These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical acquisition.
-
-How does cognateness influence word learning?
-@floccia2018 point to the possibility that parallel activation during speech perception may play a role in this cognate facilitation effect.
-The parallel activation principle suggests that bilinguals activate both languages in parallel during speech perception and production, even when only one of them is being used.
-This hypothesis can only be understood under the more general account of language non-selectivity of lexical processing, which states that translation equivalents are co-activated at the lexical level [e.g., @de1991lexical; @costa2000; @dijkstra2005; @thierry2007; @hoshino2008; @yudes2010; @bobb2020; see @kroll2013 for review].
-A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model [@kroll1994; @kroll2010], the Bilingual Interactive Activation model [BIA; @dijkstra1998bia], and its successor BIA+ [@dijkstra2002architecture], the Bilingual Language Interaction Network for Comprehension of Speech model [BLINCS; @shook2013] and, more recently, Multilink [@dijkstra2019], which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
-However, these models do not address word learning during language acquisition.
-
-It is feasible that parallel activation plays a role in vocabulary acquisition during the first years of life.
-On top of Floccia et al.'s findings, other studies have provided experimental evidence of parallel activation during speech perception in bilingual toddlers [@vonholzen2012; @vonholzen2019].
-Yet, it is unclear how parallel activation influences word learning at these ages.
-What mechanisms are engaged, and how do they relate to vocabulary size?
-We contemplate two scenarios.
-First, it is possible that the reported "similarity boost" on vocabulary size is not directly related to the amount of cognates shared between both languages, but rather to the overlap across their correspondent phonological inventories.
-One of the measures of language similarity Floccia et al.
-computed is the mean phonological Levenshtein distance across translation equivalents.
-This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
-These scores were aggregated across pairs of translation equivalents, obtaining a single similarity estimate of for each language pair.
-This measure embeds confounded information about the "degree of cognateness" and the amount of phonological overlap between both languages.
-Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, English-Dutch), with few exceptions (e.g., Spanish-Basque).
-In the light of these data, it is not possible to discern whether vocabulary size is increased by the earlier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
-We address this issue by testing two hypothesis.
-
-## Hypothesis 1: Cognates are acquired earlier than non-cognates
-
-First, we test whether the increase in vocabulary size reported by Floccia et al.
-for languages sharing a greater amount of cognates is indeed caused by cognates being acquired earlier.
-In line with this account, @bosch2014 found higher productive vocabulary scores in 18 month-old Catalan-Spanish bilinguals, compared to monolinguals, due to the higher presence of cognates their lexicon.
-We will adopt a similar approach to @braginsky2019, who modelled cross-sectional data from individual items to explore what properties were associated with an earlier acquisition.
-We will analyse the acquisition trajectories of a selection of `r dat_items %>% filter(include) %>% group_by(language) %>% summarise(n = n(), .groups = "drop") %>% pull(n) %>% max()` translation equivalents in a cross-section sample of Catalan/Spanish bilinguals aged 14 to 34 months of age.
-This approach allows to analyse how the proportion of infants that know each item changes across ages, providing a more suitable framework to put developmental hypotheses to test, instead of taking a "snapshot" of the state of the lexical of children at a single age.
-
-## Hypothesis 2: Cognates are acquired closer in times than non-cognates
-
-One mechanism that could acccount for an ealier acquisition of cognates is that parallel activation makes word-forms acquired in one language available during the acquisition of word forms in the other language.
-One consequence of this situation is that translation equivalents scaffold the acquisition of each other.
-This would point to different mechanisms engaged in the acquisition of cognates vs.
-non-cognates.
-While the phonology of cognate translation equivalents overlaps, and thus could be co-activated, the phonology of non-cognates does it in a much lesser degree.
-Consider the case of a Catalan-Spanish bilingual toddler that has already acquired the Catalan words *porta* (/ˈpɔr.tə/, door) and *taula* (/ˈtaw.ɫə/, table), but not yet their correspondent translation equivalents in Spanish (*puerta*, /ˈpwer.ta/; *mesa* /ˈme.sa/).
-According to our hypothesis, hearing the word *puerta* in presence of a door should lead to the activation of *porta*, which should facilitate the acquisition of the former.
-On the contrary, when hearing *mesa* in presence of a table, it is unlikely that the phonology of *taula* will be activated.
-This mechanism should lead to cognate translation equivalents being learnt closer in time than non-cognate translation equivalents.
-To test this hypothesis, we will extract the age of acquisition of each word-form in the questionnaire (as estimated by the model used to test Hypothesis 1), and will calculate the difference in time between the acquisition of such word-forms and their translation equivalents.
-We will test how cognateness affects such outcome.
 
@@
 Each list contained a distinct inventory of translation equivalents.
-We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/funcional categories:  Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
+We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/functional categories:  Action words, adventures, animals, body parts, clothes, colour, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
 We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
 Those categories that did not entail more than 16 items - thus resulting in less than four items after dividing it in four lists - were not divided: all of their items were included in the four lists.
-Another subset of items that were part of the trial lsits of some experiments in the lab were also included in all versions.
+Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
 
@@
 Bayesian models use the Bayes rule to combine prior knowledge about the distribution of a given parameter, and the likelihood of each of the values it can take given the data.
-We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our prior).
+We relied on the available data on Wordbank [@frank2017] to generate our prior (see Appendix 1 for more details about our prior).
 In order to test the contribution of each predictor on the fit of the model, we started fitting a null model that only included an global intercept and random intercepts for each TE, then fitted a model that also included *bilingualism* and its random slopes by TE, and finally a model that also included *cognateness* and its interaction with *bilingualism*, as well as random slopes for all the fixed effects.
@@
 Using the `BayesFactor` R package [@morey2018a], we computed a Bayes factor (*BF*) that compared the likelihood of a linear model that included *cognateness* as a predictor against the likelihood of a linear model that did not, under the light of our data.
-
-# Results
-
-## Hypothesis 1: Are cognates acquired ealier?
-
-#### Comprehensive data
-
-```{r compselection, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# compare models
-selection_comp %>% 
-  as.data.frame() %>% 
-  clean_names() %>% 
-  select(model, df, aic, bic, log_lik, l_ratio, p_value) %>% 
-  mutate(p_value = printp(p_value)) %>% 
-  gt() %>% 
-  fmt_number(columns = 1:2, decimals = 0) %>%
-  fmt_number(columns = 3:6) %>%
-  fmt_missing(columns = c(6:7), missing_text = "-") %>% 
-  cols_label(model = "Model",
-             df = md("*df*"),
-             aic = "AIC",
-             bic = "BIC",
-             log_lik = "Log. Likelihood",
-             l_ratio = "Likelihood ratio",
-             p_value = md("*p*"))
-
-```
-
-```{r companova, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-anova_comp %>% 
-  as.data.frame() %>% 
-  clean_names() %>% 
-  rownames_to_column("term") %>% 
-  mutate(p_value = printp(p_value)) %>% 
-  gt() %>% 
-  fmt_number(columns = 2, decimals = 0) %>%
-  fmt_number(columns = 3:4) %>%
-  cols_label(term = "Term",
-             num_df = md("Num. *df*"),
-             den_df = md("Den. *df*"),
-             f_value = md("*F*"),
-             p_value = md("*p*"))
-
-summary_comp <- summary(fit4_comp)
-
-```
-
-```{r comppredictions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_comp %>% 
-    arrange(item, age_bin, item_dominance) %>% 
-    ggplot(aes(age_bin, proportion, colour = interaction(cognate, item_dominance, sep = " - "))) +
-    geom_line(data = comp_preds, aes(age_bin, proportion), size = 1) + 
-    stat_summary(fun = "mean", geom = "point", shape = 1, size = 3) +
-    labs(x = "Age (months)", y = "Proportion", colour = "Cognateness") +
-    scale_colour_brewer(palette = "Set1") +
-    scale_x_continuous(breaks = unique(dat_prod$age_bin), labels = bins_interest,
-                       guide = guide_axis(n.dodge = 2)) +
-    theme_bw() +
-    theme(text = element_text(size = 12),
-          legend.position = "right",
-          legend.title = element_blank(),
-          axis.text = element_text(colour = "black")) +
-  ggsave(here("Figures", "predictions_comp.png"))
-
-```
-
-### Productive data
-
-```{r prodselection, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# compare models
-selection_prod %>% 
-  as.data.frame() %>% 
-  clean_names() %>% 
-  select(model, df, aic, bic, log_lik, l_ratio, p_value) %>% 
-  mutate(p_value = printp(p_value)) %>% 
-  gt() %>% 
-  fmt_number(columns = 1:2, decimals = 0) %>%
-  fmt_number(columns = 3:6) %>%
-  fmt_missing(columns = c(6:7), missing_text = "-") %>% 
-  cols_label(model = "Model",
-             df = md("*df*"),
-             aic = "AIC",
-             bic = "BIC",
-             log_lik = "Log. Likelihood",
-             l_ratio = "Likelihood ratio",
-             p_value = md("*p*"))
-
-```
-
-```{r prodanova, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-anova_prod %>% 
-  as.data.frame() %>% 
-  clean_names() %>% 
-  rownames_to_column("term") %>% 
-  mutate(p_value = printp(p_value)) %>% 
-  gt() %>% 
-  fmt_number(columns = 2, decimals = 0) %>%
-  fmt_number(columns = 3:4) %>%
-  cols_label(term = "Term",
-             num_df = md("Num. *df*"),
-             den_df = md("Den. *df*"),
-             f_value = md("*F*"),
-             p_value = md("*p*"))
-
-```
-
-```{r prodpredictions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_prod %>% 
-    arrange(item, age_bin, item_dominance) %>% 
-    ggplot(aes(age_bin, proportion, colour = interaction(cognate, item_dominance, sep = " - "))) +
-    geom_line(data = prod_preds, aes(age_bin, proportion), size = 1) + 
-    stat_summary(fun = "mean", geom = "point", shape = 1, size = 3) +
-    labs(x = "Age (months)", y = "Proportion", colour = "Cognateness") +
-    scale_colour_brewer(palette = "Set1") +
-    scale_x_continuous(breaks = unique(dat_prod$age_bin), labels = bins_interest,
-                       guide = guide_axis(n.dodge = 2)) +
-    theme_bw() +
-    theme(text = element_text(size = 12),
-          legend.position = "right",
-          legend.title = element_blank(),
-          axis.text = element_text(colour = "black")) +
-  ggsave(here("Figures", "predictions_prod.png"))
-
-```
-
-```{r ranefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-ggplot(rancoefs, aes(xmid_intercept, xmid_item_dominance1, colour = cognate, fill = cognate)) +
-  facet_wrap(type~cognate, scales = "free") +
-  geom_point(size = 3, alpha = 0.5, shape = 1, colour = "grey") +
-  geom_smooth(method = "lm", formula = y~x) +
-  labs(x = "Mid-point intercept", y = "Item dominance slope on Mid-point", colour = "Cognateness") +
-  scale_colour_brewer(palette = "Set1") +
-  scale_fill_brewer(palette = "Set1") +
-  scale_x_continuous() + 
-  theme_bw() +
-  theme(legend.background = element_rect(fill = NA),
-        legend.title = element_blank(),
-        legend.key = element_rect(fill = NA),
-        legend.position = "none")
-
-```
-
-## Hypothesis 2
-
-```{r aoafig, echo=FALSE, message=TRUE, warning=FALSE, paged.print=FALSE}
-
-ggplot(aoas, aes(xmid, fct_rev(item_dominance))) +
-    facet_grid(cognate~type) +
-    annotate(geom = "rect", xmin = 0, xmax = 10, ymin = -Inf, ymax = Inf, colour = NA, fill = "orange", alpha = 0.25)  +
-    #annotate(geom = "text", x = 5, y = 0, label = "Range of observed data", size = 7) +
-    geom_line(aes(group = te), colour = "black", alpha = 0.05) +
-    geom_violin(fill = "grey30", colour = "grey30", width = 0.5) +
-    geom_boxplot(width = 0.25, fill = "white", colour = "black", outlier.colour = NA) +
-    labs(x = "Age (months)", y = "Item dominance", fill = "Item dominance", colour = "Item dominance") +
-    theme_bw() +
-    scale_x_continuous(breaks = -5:15, labels = bins_expanded, guide = guide_axis(n.dodge = 1)) +
-    theme(legend.title = element_blank(),
-          legend.position = "none",
-          legend.background = element_rect(fill = NA),
-          axis.text = element_text(colour = "black"),
-          axis.text.x = element_text(size = 8),
-          axis.title.y = element_blank(),
-          panel.grid.major.y = element_blank(),
-          panel.grid.minor.y = element_blank()) +
-  ggsave(here("Figures", "aoa.png"))
-
-```
-
-# Discussion
-
-# Appendix
-
-## Appendix 1: Questionnaires
-
-## Appendix 2: Model details
-
-## Appendix 3: Session info
-
-```{r sessioninfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-sessionInfo()
-
-```
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-
-\endgroup
-
-```{r create_r-references, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-r_refs(file = "../References/trajectories.bib")
-
-```

### 2020-12-07 Rmd/04_appendix_2.Rmd
Prepare introduction

--- 2020-06-30 Code/Rmd/data_analysis_notes.Rmd
@@
 ---
-title: "Notes on data anlaysis"
-output: html_notebook
+title: 'Appendix 2: Model'
+output: pdf_document
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+Our model specifies a logistic curve that fits our outcome measure.
+Out outcome measure is the porportion of participants that are reported to understand each word.
+This logistic curve is characterised by three parameters: *asymptote* ($asym$, the value at which the outcome *pleateaus*), *steepness*, ($steep$, how fast the curve grows), and *mid-point* ($mid-point$, the age at which the curve reaches its maximum steepness).
+We identify the *mid-point* of a word as its age of acquisition, and thus us the parameter of interest.
+Out hypothesis is that the language a word belongs to (non-dominant vs.
+dominant), together with its cognate status (non-cognate vs.
+cognate) and frequency, affects the position of its mid-point.
+We assume that the mid-point is generated from a zero-one beta regression model that includes the aforementioned predictors.
+A beta regression model helps us model data that include many values at boundary, as ours are.
+We also included random intercepts for each *translation equivalent* ($TE$).
+We set informed priors for *asymptotes*, *steepnesses*, and *mid-points* according to Wordbank data, and weak priors the the coefficients of the beta-regression model.
+
+$$
+\begin{aligned}proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\ mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ TE_{0i} \sim Normal(0, \sigma_i) \\\beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\ \phi \sim Normal(1.5, 1)\\ \sigma_i \sim HalfCauchy(5)
+\end{aligned}
+$$
 

### 2020-12-07 Rmd/trajectories_manuscript.Rmd
Prepare introduction

--- 2020-07-03 Manuscript/trajectories_manuscript.Rmd
@@
 library(knitr)
-
 library(magrittr)
@@
 library(janitor)
-
-library(broom)
-library(ggridges)
-library(patchwork)
 library(gt)
-
 library(data.table)
@@
 
-# load/create functions
-source(here("Code", "R", "functions.R"))
+```
 
-# set params
-bins <- c("12-14", "14-16", "16-18", "18-20", "20-22", "22-24", "24-26", "26-28", "28-30", "30-32", "32-34")
-
-# import data 
-dat_studies <- fread(here("Data", "00_studies.csv")) %>% distinct(q_version, language, q_items)
-dat_items <- read_xlsx(here("Data", "01_pool.xlsx")) %>%
-  group_by() %>%
-  drop_na(language) %>%
-  mutate_at(vars(A, B, C, D, cognate_rater1, include, frequency), as.numeric) %>%
-  mutate_at(vars(A, B, C, D, cognate_rater1, include), as.logical)
-dat_participants <- list.files(here("Data", "Logs"), pattern = "logs", full.names = TRUE) %>%
-  last() %>%
-  fread(na.strings = "") %>%
-  filter(age_bin %in% bins,
-         completed,
-         lp == "Bilingual")
-dat_raw <- fread(here("Data", "02_merged.csv"), na.strings = "") %>%
-  filter(lp %in% c("Bilingual"),
-         age_bin %in% bins) %>%
-  as_tibble() %>%
-  mutate_at(vars(date_birth, time_stamp), as_date)
-dat_wordbank <- fread(here("Data", "05_wordbank.csv"))
-dat_wordbank_priors <- fread(here("Data", "05_priors-wordbank.csv"))
-
-# import results
-comp_fit1 <- readRDS(here("Results", "comp_fit1.rds"))
-comp_selection <- readRDS(here("Results", "comp_selection.rds"))
+```{r, child= "00_introduction.Rmd"}
 
@@
 
-# Introduction
-
-Bilingual toddlers must learn two distinct set of words (one for each language) that partially overlap in sound and meaning.
-Previous studies have reported that bilingual children know fewer words than their monolingual counterparts when only language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When taking both languages into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev1977; @doyle1977some; @rosenblum1983word; @fernandez1992; @pearson1994; @bialystok2010; @hoff2012; @core2013; @hoff2014; @blom2019; but see @pearson1993; @houwer2014].
-Not all bilinguals show a similar developmental trajectory of lexical acquisition.
-Recent studies have capitalised on the role of the similarity between the specific pair of languages the infant is learning.
-One way of measuring language similarity is by estimating the amount of cognates (i.e., form-similar translation equivalents) they share.
-@floccia2018 explored the impact of this form of similarity on vocabulary sizes from a large sample of 24-month-old toddlers learning English and an additional language.
-Toddlers learning languages that shared a large amount of cognates (e.g., English-German) showed larger vocabulary sizes than those learning languages that shared fewer cognates (e.g., English-Spanish).
-@blom2019 extended these findings to children aged three to 10 years.
-These results suggest that the phonological overlap between the word inventories infants are acquiring impacts their trajectory of lexical aquisition.
-
-How does cognateness influence word learning?
-@floccia2018 point to the possibility that parallel activation during speech perception may play a role in this cognate facilitation effect.
-The parallel activation principle suggests that bilinguals activate both languages in parallel during speech perception and production, even when only one of them is being used.
-This hypothesis can only be understood under the more general account of language non-selectivity of lexical processing, which states that translation equivalents are co-activated at the lexical level [e.g., @de1991lexical; @costa2000; @dijkstra2005; @thierry2007; @hoshino2008; @yudes2010; @bobb2020; see @kroll2013 for review].
-A growing body of experimental evidence has been given shape to different formal models of bilingual lexical processing, like the Revised Hierarchical model [@kroll1994; @kroll2010], the Bilingual Interactive Activation model [BIA; @dijkstra1998bia], and its successor BIA+ [@dijkstra2002architecture], the Bilingual Language Interaction Network for Comprehension of Speech model [BLINCS; @shook2013] and, more recently, Multilink [@dijkstra2019], which represents an effort for integrating the former to account for language-non selective activation during both comprehension and production.
-However, these models do not address word learning during language acquistion.
-
-It is feasible that parallel activation plays a role in vocabulary acquisition during the first years of life.
-On top of Floccia et al.'s findings, other studies have provided experimental evidence of parallel activation during speech perception in bilingual toddlers [@vonholzen2012; @vonholzen2019].
-Yet, it is unclear how parallel activation influences word learning at these ages.
-What mechanisms are engaged, and how do they relate to vocabulary size?
-We contemplate two scenarios.
-First, it is possible that the reported "similarity boost" on vocabulary size is not directly related to the amount of cognates shared between both languages, but rather to the overlap across their correspondent phonological inventories.
-One of the measures of language similarity Floccia et al.
-computed is the mean phonological Levenshtein distance across translation equivalents.
-This involves the calculation of the number of edit operations (e.g., insertions, deletions or replacements) the phonological transcription of a word-form has to go through to become identical to that of its translation equivalent.
-These scores were aggregated across pais of tanslation equivalents, obtaining a single similarity estimate of for each language pair.
-This measure embeds confounded information about the "degree of cognateness" and the amount of phonological overlap between both languages.
-Languages that are phonologically closer tend to share more cognates than those that are phonologically more distant (e.g., Spanish-Italian, Enlgish-Dutch), with few exceptions (e.g., Spanish-Basque).
-In the light of these data, it is not possible to discern whether vocabulary size is increased by the ealier acquisition of cognates or by the overlap across the phonological systems of the languages of a bilingual.
-We address this issue by testing two hypotesis.
-
-## Hypothesis 1: Cognates are acquired earlier than non-cognates
-
-First, we test whether the increase in vocabulary size reported by Floccia et al.
-for languages sharing a greater amount of cognates is indeed caused by cognates being acquired earlier.
-In line with this account, @bosch2014 found higher productive vocabulary scores in 18 month-old Catalan-Spanish bilinguals, compared to monolinguals, due to the higher presence of cognates their lexicon.
-We will adopt a similar approach to @braginsky2019, who modelled cross-sectional data from individual items to explore what properties were associated with an earlier acquisition.
-We will analyse the acquisition trajectories of a selection of `r dat_items %>% filter(include) %>% group_by(language) %>% summarise(n = n(), .groups = "drop") %>% pull(n) %>% max()` translation equivalents in a cross-section sample of Catalan/Spanish bilinguals aged 14 to 34 months of age.
-This approach allows to analyse how the proportion of infants that know each item changes across ages, providing a more suitable framework to put developmental hypotheses to test, instead of taking a "snapshot" of the state of the lexical of children at a single age.
-
-## Hypothesis 2: Cognates are acquired closer in times than non-cognates
-
-One mechanism that could acccount for an ealier acquisition of cognates is that parallel activation makes word-forms acquired in one language available during the acquisition of word forms in the other language.
-One consequence of this situation is that translation equivalents scaffold the acquisition of each other.
-This would point to different mechanisms engaged in the acquisition of cognates vs.
-non-cognates.
-While the phonology of cognate translation equivalents overlaps, and thus could be co-activated, the phonology of non-cognates does it in a much lesser degree.
-Consider the case of a Catalan-Spanish bilingual toddler that has already acquired the Catalan words *porta* (/ˈpɔr.tə/, door) and *taula* (/ˈtaw.ɫə/, table), but not yet their correspondent translation equivalents in Spanish (*puerta*, /ˈpwer.ta/; *mesa* /ˈme.sa/).
-According to our hypothesis, hearing the word *puerta* in presence of a door should lead to the activation of *porta*, which should facilitate the acquisition of the former.
-On the contrary, when hearing *mesa* in presence of a table, it is unlikely that the phonology of *taula* will be activated.
-This mechanism should lead to cognate translation equivalents being learnt closer in time than non-cognate translation equivalents.
-To test this hypothesis, we will extract the age of acquisition of each word-form in the questionnaire (as estimated by the model used to test Hypothesis 1), and will calculate the difference in time between the acquisition of such word-forms and their translation equivalents.
-We will test how cognateness affects such outcome.
-
-# Method
-
-## Participants
-
-We collected data from `r dat_raw$id_db %>% unique %>% length() %>% printnum()` bilinguals from the Metropolitan Area of Barcelona, between `r min(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`.
-All families participated voluntarily.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-
-We assessed toddlers' language background asking parents for an estimated proportion of exposure to each language.
-We defined participants as Spanish- or Catalan-dominant according to the language they were exposed the most to.
-Participants were considered bilingual - therefore included - if exposed 50-90% of the time to the non-dominant language.
-We excluded participants with >10% exposure to a third language.
-Fig.
-\@ref(fig:lpdist) shows how participants were distributed across age bins and language profiles.
-
-```{r lpdist, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_participants %>%
-  filter(completed, lp %!in% "Other", age_bin %in% bins) %>%
-  ggplot(aes(x = age_bin, fill = dominance)) +
-  geom_bar() +
-  labs(x = "Age (months)", y = "Number of partcipants",
-       fill = "Dominant language") +
-  scale_y_continuous(labels = as.integer) +
-  scale_fill_brewer(palette = "Set1") +
-  theme_custom +
-  theme(legend.position = c(0.05, 0.85),
-        legend.direction = "vertical",
-        legend.justification = 0,
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))
-```
-
-## Questionnaire
-
-Parents completed an online questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
-We implemented the questionnaire in `formr` [@arslan2020], a free, open-source platform.
-Participants were randomly allocated into one of four list of items.
-Each list contained a distinct inventory of translation equivalents.
-We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/funcional categories:  Action words, adventures, animals, body parts, clothes, color, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-Those categories that did not entail more than 16 items - thus resulting in less than four items after dividing it in four lists - were not divided: all of their items were included in the four lists.
-Another subset of items that were part of the trial lsits of some experiments in the lab were also included in all versions.
-
-Each item of the inventory presented a word-form, to which parents were instructed to report whether the toddlers was able to understand it, understand ans say it, or neither.
-The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates) .
-Items of one language were translation equivalents from the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-
-## Data analysis
-
-For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item  within age bins of 2 months.
-Our model assumes that this proportion follows a logistic curve accross ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the dominance of the language of the item (*item dominance*, L1 vs.
-L2), the cognate status of the word (*cognateness*, Non-cognate vs.
-Cognate), and the frequency of the word form in its correspondent language.
-Our prediction  is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words.
-
-Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
-Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item porta (door, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
-We retrieved frequencies from SUBTLEX-ESP [@cuetos2011] and SUBTLEX-CAT [@boada2019] for Spanish and Catalan words, respectively.
-Word frequency was calculated as the Zipf score of the word [@zipf1949; @heuven2014].
-Missing frequency scores were imputed using multiple imputation [@vanbuuren2011].
-Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
-
-```{r itemsinfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_items %>%
-  drop_na(cognate_rater1) %>%
-  filter(include) %>%
-  select(A, B, C, D, cognate_rater1, frequency, language) %>% 
-  mutate(cognate_rater1 = ifelse(cognate_rater1, "Cognate", "Non-cognate")) %>%
-  pivot_longer(c(A, B, C, D), names_to = c("list"), values_to = "n") %>% 
-  filter(n) %>% 
-  group_by(list, language, cognate_rater1) %>% 
-  summarise(n = sum(n, na.rm = TRUE),
-            mean = mean(frequency, na.rm = TRUE),
-            sd = sd(frequency, na.rm = TRUE),
-            min = min(frequency, na.rm = TRUE),
-            max = max(frequency, na.rm = TRUE)) %>%
-  ungroup() %>% 
-  rename(cognateness = cognate_rater1) %>%
-  relocate(list, language, cognateness, n, mean, sd, min, max) %>%
-  arrange(list, language, cognateness) %>%
-  mutate(list = paste("List", list)) %>% 
-  pivot_wider(names_from = cognateness, values_from = c(n, mean, sd, min, max)) %>%
-  clean_names() %>%
-  gt(groupname_col = c("language")) %>% 
-  fmt_number(columns = contains("cognate"), decimals = 2) %>%
-  fmt_number(columns = starts_with("n_"), decimals = 0) %>%
-  tab_header(title = md("**Questionnaire and item properties**"),
-             subtitle = "Participants were randomly allocated on of the the four lists, and completed both the Catalan and the Spanish versions.") %>% 
-  tab_spanner(label = "Frequency (Cognates)", columns = vars("n_cognate", "mean_cognate", "sd_cognate", "min_cognate", "max_cognate")) %>% 
-    tab_spanner(label = "Frequency (Non-cognates)", columns = contains("non_cognate")) %>% 
-  cols_label(list = md("**List**"), 
-             n_cognate = md("*N*"),
-             mean_cognate = md("*Mean*"),
-             sd_cognate = md("*SD*"),
-             min_cognate = md("*Min*"),
-             max_cognate = md("*Max*"),
-             n_non_cognate = md("*N*"),
-             mean_non_cognate = md("*Mean*"),
-             sd_non_cognate = md("*SD*"),
-             min_non_cognate = md("*Min*"),
-             max_non_cognate = md("*Max*")) %>%
-  gtsave(here("Results", "descriptives_items.png"))
+```{r, child= "01_method.Rmd"}
 
@@
 
-We fit two separate logistic models on comprehensive and productive data, respectively.
-Logistic curves are characterised by three parameters[see @mahr2019 and @mahr2020 for an excellent introduction to this approach, in which we based this analysis]: (a) an *asymptote* (`asym`, upper boundary  of the proportion of infants that have acquired the item), (b) the *steepness* (`steep`, how fast the a given item is acquired by infants), and (c) a *mid-point* (`mid`, the age at which steepness is maximum).
-We work on the assumption that the trajectory of aquisition of items follows a sigmoidal shape [@mayor2011], and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
-This approach allows us to model and estimate the age of acquisition of items explicitely.
-Our model estimated the asymptote and stepness as grand means across items (i.e., intercepts).
-Mid-points were estimated by a linear zero-one-beta-regression model^[Zero-or-one-inflated beta regression models allow to account for overdispersion then using proportions as the output of a model accounting, while also allowing values to be at boundary (0 and 1).][@ospina2012] that included our predictors of interest: `age`, `dominance` (contrast-coded, L2 = -0.5, L2 = 0.5), `cognateness` (contrast-coded, Non-cognate = -0.5, Cognate = 0.5), their interaction, and `frequency` (standardised Zipf score).
-We also included a random intercept for each translation equivalent (`TE`).
-Appendix 1 shows a detailed description of the model.
-
-We estimated the parameters of our model using the Bayesian framework.
-Bayesian models use the Bayes rule to combine prior knowlegde about the distribution of a given parameter, and the likelihood of each of the values it can take given tha observed data.
-We relied on the available data on Wordbank [@frank2017] to generate our priors (see Appendix 1 for more details about our priors).
-In order to test the contribution of the predictor *cognateness* on the fit of the model,  we compared the fit of our (expanded) model to the fit of a (null) model that did not include *cognateness* as predictor [@schad2020] using leave-one-out cross validation.
-We fit both the null and alternative models using the R environment [@rcoreteam2013; @rstudioteam2015] and the `brms` package [@burkner2017], which relies on the probabilistic language Stan [@carpenter2017] to approximate posterior distributions.
-We used Pareto-smoothed importance sampling [PSIS; @vehtari2017; @vehtari2019] to compare the null and extended models.
-Data and model results were processed and visualised using the `tidyverse` family of R packages [@wickham2019] and the `tidybayes` package [@kay2020].
-
-To test Hypothesis 2, we run a Bayesian ANOVA using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included *cognateness* as predictor.
-Using the `BayesFactor` R package [@morey2018a], we computed a Bayes factor (*BF*) that compared the likelihood of a linear model that included *cognateness* as a predictor against the likelihood of a linear model that did not, under the light of our data.
-
-# Results
-
-## Hypothesis 1: Are cognates acquired ealier?
-
-### Comprehensive vocabulary
-
-Table \@ref(tab:resultscomp) shows the outcomes of the model.
-
-```{r resultscomp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tidy(comp_fit1) %>% 
-    gt()
+```{r, child= "02_results.Rmd"}
 
@@
 
-### Productive vocabulary
-
-## Hypothesis 2
-
-### Comprehensive vocabulary
-
-### Productive vocabulary
-
-# Discussion
-
-# Appendix
-
-## Appendix 1
-
-Model formula and priors:
-
-$$proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ \beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\\phi \sim Normal(1.5, 1)$$
-
-## Appendix 2: Computing priors
-
-We fitted a logistic model using the proportion of infants that acquired each item as output, and age-bin (defined as we did with our data) as input, estimating the three parameters afore mentioned (asymptote, mid-point, and steepness).
-We included data from `r nrow(dat_wordbank)/length(unique(dat_wordbank$age)) %>% printnum()` items from `r dat_wordbank$language %>% unique %>% length` languages^[Chinese (Cantonese and Mandarin), Croatian, Czech, Danish, English (Australian, British, and American), French (European), German, Hebrew, Italian, Korean, Latvian, Portuguese, Russian, Slovak, Spanish (European and Mexican), and Turkish.
-All the other available languages were excluded, as no (or insufficient) data were available.], from monolingual infants aged 16 to 30 months of age.
-Table \@ref(tab:wordbank) shows the estimates of this model, which were rescaled and included as priors in our main model to test Hypothesis 1.
-
-```{r worbank, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-dat_wordbank_priors %>%
-  rename(estimate_scaled = estimate) %>%
-  mutate(conf = paste0("[", round(conf_low, 2), ", ", round(conf_high, 2), "]"),
-         estimate = estimate_scaled*2+20,
-         term = c("Asymptote", "Mid-point", "Steepness")) %>% 
-  select(term, estimate, estimate_scaled, std_error, conf) %>%
-  gt(rowname_col = "term") %>% 
-  fmt_number(vars(estimate, estimate_scaled, std_error), decimals = 2) %>% 
-  cols_label(estimate = md("*Estimate<br>(months)*"),
-             estimate_scaled = md("*Estimate<br>(scaled)*"),
-             std_error = md("*SE*"),
-             conf = md("*95% CI*")) %>% 
-  cols_align(align = "c") %>% 
-  gtsave(here("Results", "wordbank.png"))
+```{r, child= "03_discussion.Rmd"}
 
@@
 
-We performed prior-predictive checks to analyse the feasibility of our priors. Figure \@ref(fig:priorchecks)
-
-```{r priorchecks, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-include_graphics(here("Figures", "06_simulate-priors.png"))
+```{r, child= "04_appendix_1.Rmd"}
 
@@
 
-```{r priorstab, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tribble(~param, ~coefficient, ~notation, ~prior,
-        "Asymptote", "Intercept", "$A$", paste0("$N(", dat_wordbank_priors[term=='asym'] %>% pull(estimate) %>% printnum(), ", 0.05)$"),
-        "Steepness", "Intercept", "$S$", paste0("$N(", dat_wordbank_priors[term=='steep'] %>% pull(estimate) %>% printnum(), ", 0.6)$"),
-        "Mid-point", "Intercept", "$\\beta_{0}$", paste0("$N(", dat_wordbank_priors[term=='mid'] %>% pull(estimate) %>% printnum(), ", 1)$"),
-        "Mid-point", "Random Intercepts", "$I_{i0}$", "N(0, $\\pi$)",
-        "Mid-point", "Slope", "$\\beta_{1}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{2}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{3}$", "$N(0, 5)$",
-        "Mid-point", "Slope", "$\\beta_{4}$", "$N(0, 5)$",
-        "Mid-point", "Intercept", "$\\varphi$", "$N(2, 1)$",
-        "Mid-point", "Intercept", "$COI$", "Beta(1, 1)", 
-        "Mid-point", "Intercept", "$ZOI$", "Beta(1, 1)",
-        "Mid-point", "Std. Deviation", "$\\sigma{ijk}$", "t(3, 0, 2.5)") %>%
-  kable(digits = 2, row.names = FALSE, col.names = c("**Adjusted parameter**", "**Coefficient**", "**Notation**", "**Distribution**"), align = "c", caption = "Prior distribution of the parameters in our models.
-We set specific informative priors for the asymptote, steepness, and mid-point of the curves (using data from Wordbank) and and weakly informative priors for the rest of the coefficients.", escape = TRUE, format = "latex")
+```{r, child= "04_appendix_2.Rmd"}
 
@@
 
-## Appendix 3: Session info
-
-```{r session_info, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-sessionInfo()
+```{r, child= "04_appendix_3.Rmd"}
 
@@
 
-r_refs(file = "../References/trajectories.bib")
+r_refs(file = "trajectories.bib")
 

### 2020-12-07 R/analyses.R

@@ -40,7 +40,6 @@ dat_vocab <- logs %>%
     ) %>% 
     mutate_at(vars(vocab_type), as.factor) 
 
-
 #### process responses ---------------------------------------------------------
 dat_responses <- responses %>% 
     select(id, time, item, response, language) %>% 
@@ -70,9 +69,12 @@ dat_responses <- responses %>%
     ungroup() %>% 
     drop_na(response, te, cognate) %>% 
     select(id, age, sex, dominance, lp, te, item, response, language, item_dominance, cognate, frequency, bilingualism) %>% 
-    mutate(age_centred = scale(age, center = TRUE, scale = FALSE)[,1]) %>%
+    #mutate(age_centred = scale(age, center = TRUE, scale = FALSE)[,1]) %>%
     mutate_at(vars(cognate, item_dominance), as.factor) %>%
-    mutate(response = factor(response, ordered = TRUE, labels = c("No", "Understands", "Understands & Says"))) %>% 
+    mutate(response = factor(
+        response, ordered = TRUE,
+        labels = c("No", "Understands", "Understands & Says"))
+    ) %>% 
     arrange(item, te) %>% 
     select(id, age, bilingualism, item, te, item_dominance, cognate, response) 
 
@@ -83,24 +85,25 @@ contrasts(dat_responses$item_dominance) <- c(0, 1)
 #### model fitting -------------------------------------------------------------
 priors <- c(
     prior(normal(0.5, 0.5), class = Intercept),
-    prior(normal(0.1, 0.15), class = b, coef = "age_centred"),
+    prior(normal(0.1, 0.15), class = b, coef = "age"),
     prior(normal(0, 0.5), class = b, coef = "item_dominance1"),
-    prior(normal(0, 0.5), class = b, coef = "age_centred:item_dominance1"),
+    prior(normal(0, 0.5), class = b, coef = "age:item_dominance1"),
     prior(normal(0, 0.5), class = b, coef = "cognateness1"),
-    prior(normal(0, 0.5), class = b, coef = "age_centred:cognateness1"),
+    prior(normal(0, 0.5), class = b, coef = "age:cognateness1"),
     prior(normal(0, 0.5), class = b, coef = "item_dominance1:cognateness1"),
-    prior(normal(0, 0.5), class = b, coef = "age_centred:item_dominance1:cognateness1")
[…]

### 2021-01-09 Rmd/01_method.Rmd
Summarised version (bernoulli takes too long)

@@
 
+dat_participants <- read.csv("Data/participants.csv")
+
 ```
@@
 We collected data from `r dat_raw$id_db %>% unique %>% length() %>% printnum()` bilinguals from the Metropolitan Area of Barcelona, between `r min(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`.
-All families participated voluntarily.
+All families gave informed consent before participating.
 This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
 
-We assessed toddlers' language background asking parents for an estimated proportion of exposure to each language.
-We defined participants as Spanish- or Catalan-dominant according to the language they were exposed the most to.
-Participants were considered bilingual - therefore included - if exposed 50-90% of the time to the non-dominant language.
+We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language.
 We excluded participants with >10% exposure to a third language.
 Fig.
-\@ref(fig:lpdist) shows how participants were distributed across age bins and language profiles.
+\@ref(fig:lpdist) illustrates the distribution of participants across language profiles and ages.
 
@@
   filter(lp %in% "Bilingual") %>% 
-    mutate(bilingualism = cut_interval(bilingualism, n = 6,
+  mutate(bilingualism = cut_interval(
+    bilingualism, n = 6,
     labels = paste0(seq(0, 50, by = 10), "%"))) %>%
@@
   theme(axis.text = element_text(colour = "black"))
-ggsave(here("Figures", "participants.png"), height = 4, width = 5)
 ```
@@
 
-Parents completed an on-line questionnaire that gathered information about their demographic status, language environment of the child, and that included two vocabulary inventories: one for Catalan and one for Spanish.
-We implemented the questionnaire in `formr` [@arslan2020], a free, open-source platform.
-Participants were randomly allocated into one of four list of items.
-Each list contained a distinct inventory of translation equivalents.
+We implemented an on-line questionnaire using `formr` [@arslan2020], divided in three forms: (1) language questionnaire, (2) demographic survey, and (3) Catalan and Spanish vocabulary checklists.
+Vocabulary checklists consisted in two lists of words (one in Catalan and one in Spanish).
+Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
 We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/functional categories: Action words, adventures, animals, body parts, clothes, colour, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
 We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-Those categories that did not entail more than 16 items - thus resulting in less than four items after dividing it in four lists - were not divided: all of their items were included in the four lists.
+The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates).
+
+For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default).
+Participants filled a long or a short version of the questionnaire.
+Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items.
+Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish).
+Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists.
 Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-
-Each item of the inventory presented a word-form, to which parents were instructed to report whether the toddlers was able to understand it, understand ans say it, or neither.
-The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates) .
-Items of one language were translation equivalents from the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
 
@@
 
-For testing Hypothesis 1, we modelled the proportion of infants that understood or produced each item in the non-dominant language within age bins of 2 months.
-Our model assumes that this proportion follows a logistic curve across ages, an that the age of acquisition of each item is generated from a linear model that adjusts for the degree of exposure to the non-dominant language (*bilingualism*, 0-50%), the cognate status of the word (*cognateness*, Non-cognate vs.
-Cognate), and their interaction.
-Our prediction is that the age of acquisition will be lower for cognate words, compared to that of non-cognate words, and that this effect will be larger for more balanced bilinguals.
+```{r n_responses}
+n_obs <- nrow(dat_responses)
+sum_obs <- dat_responses %>%
+  group_by(item) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  summarise(
+    mean = mean(n),
+    sd = sd(n),
+    min = min(n),
+    max = max(n)
+  )
 
-Items were classified into cognates and non-cognates by a trained Catalan-Spanish bilingual with broad experience in phonology and phonetics.
-Items were classified as L1 if the item was included in the inventory of the language the participant was dominant in, and as L2 if otherwise For instance, the item *porta* (*door*, in Catalan) was classified as L1 when answered by Catalan-dominant participants, and as L2 when answered by Spanish-dominant participants.
+```
+
+We gathered `r printnum(n_obs, digits = 0)` responses.
+Items received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$max, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`.
+We modelled the probability of participants' being reported to understand/produce a translation equivalent adjusting by:
+
+* $Age$ (in months), calculated as the number of days passed from birth date to the day the questionnaire was completed, divided by 30.
+* $Exposure$: composite score of the lexical frequency of the word form in Catalan or Spanish, as extracted from SUBTLEX-ESP [@cuetos2011] or SUBTLEX-CAT [@]
+
+We modelled comprehensive and productive data separately, following a similar procedure.
+First, we fit a base logistic regression model with a Bernoulli distribution that adjusted by the age of participants.
+Then we extended this model including the main effect of exposure
+To address our hypothesis, we modelled the probability of parents reporting that their child understood and/or produced translation equivalents, conditional to how much the child is exposed to it, and to the phonological similarity between both translation equivalents.
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+
 Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.

### 2021-01-11 R/analyses.R
Code responses

@@ -14,8 +14,14 @@ library(here)
 source(here("R", "utils.R"))
 
 #### import data ---------------------------------------------------------------
-responses <- read_csv(here("Data", "responses.csv"))
-responses_subset <- filter(responses, te %in% sample(responses$te, 25))
+responses <- read_csv(here("Data", "responses.csv")) %>% 
+  mutate_at(vars(age, frequency, doe), function(x) scale(x, scale = FALSE)[,1]) %>% 
+  mutate_at(vars(item_dominance, cognate, response), as.factor) 
+
+responses_subset <- filter(responses, te %in% sample(responses$te, 25)) %>% 
+  mutate_at(vars(age, frequency, doe), function(x) scale(x, scale = FALSE)[,1]) %>% 
+  mutate_at(vars(item_dominance, cognate), as.factor) %>% 
+  mutate(response = factor(response, ordered = TRUE))
 
 # summarise responses (for visualisation purposes)
 proportion <- responses %>% 
@@ -27,8 +33,10 @@ proportion <- responses %>%
   ) %>%
   select(-doe) %>% 
   pivot_longer(c(understands, says), names_to = ".category", values_to = ".value") %>% 
-  mutate(.category = str_to_sentence(.category),
-         .value = as.numeric(.value)) %>% 
+  mutate(
+    .category = str_to_sentence(.category),
+    .value = as.numeric(.value)
+  ) %>% 
   group_by(age, cognate, item_dominance, .category) %>%
   summarise(
     .value = sum(.value),
@@ -37,7 +45,7 @@ proportion <- responses %>%
   ) %>% 
   rowwise() %>% 
   mutate(.value = prop_adj(.value, n)) %>% 
-  ungroup()
+  ungroup() 
 
 
 # set sum contrasts (Schad et al., 2018, https://arxiv.org/abs/1807.10451)
@@ -139,19 +147,19 @@ fit_9 <- update(
 #### model comparison ----------------------------------------------------------
 
 # paretho-smooth leave-one-out cross-validation
-loo_0 <- loo(fit_0)
-loo_1 <- loo(fit_1)
-loo_2 <- loo(fit_2)
-loo_3 <- loo(fit_3)
-loo_4 <- loo(fit_4)
[…]

### 2021-01-11 R/preprocessing.R

@@ -26,21 +26,7 @@ items <- pool %>%  # from multilex::pool
         class %in% c("Noun")
     ) %>% 
     rename(frequency = frequency_zipf) %>% 
-    select(te, language, item, ipa, frequency, cognate) %>% 
-    pivot_wider(
-        id_cols = c(te, cognate),
-        names_from = language,
-        values_from = c(item, ipa, frequency)
-    ) %>% 
-    clean_names() %>% 
-    mutate(
-        similarity = stringsim(ipa_catalan, ipa_spanish, method = "lv")
-    ) %>% 
-    pivot_longer(
-        cols = matches("_catalan|_spanish"),
-        names_to = c(".value", "language"),
-        names_sep = "_",
-    ) %>% 
+    select(te, language, category, item, ipa, frequency, cognate) %>% 
     mutate(
         language = str_to_sentence(language),
         cognate = ifelse(cognate, "Cognate", "Non-Cognate")
@@ -60,8 +46,12 @@ participants <- l %>%
     group_by(id) %>% 
     filter(time==max(time)) %>% 
     ungroup() %>% 
-    select(id, time, age, dominance, doe_catalan, doe_spanish) %>% 
-    mutate_at(vars(matches("doe")), function(x) x*0.01) 
+    mutate(doe = case_when(
+        dominance=="Catalan" ~ doe_spanish,
+        dominance=="Spanish" ~ doe_catalan
+    )) %>% 
+    mutate_at(vars(matches("doe")), function(x) x*0.01) %>% 
+    select(id, time_stamp, time, age, sex, dominance, doe)
 
 write_csv(participants, here("Data", "participants.csv"))
 
@@ -82,25 +72,18 @@ responses <- expand_grid(
 ) %>%
     left_join(participants, by = "id") %>% 
     left_join(items, by = "item") %>% 
-    left_join(select(responses, id, item, response), by = c("id", "item")) %>% 
-    select(id, age, dominance, doe_catalan, doe_spanish, te, language, item, ipa, cognate, frequency, response) %>% 
+    left_join(select(r, id, item, response), by = c("id", "item")) %>% 
+    select(id, age, dominance, doe, te, language, item, ipa, cognate, frequency, response) %>% 
     drop_na() %>%
     distinct(id, te, age, item, .keep_all = TRUE) %>% 
     mutate(
-        doe = case_when(
-            dominance=="Catalan" ~ doe_spanish,
[…]

### 2021-01-11 Rmd/01_method.Rmd
Code responses

@@
 library(knitr)
+library(here)
 
-dat_participants <- read.csv("Data/participants.csv")
+participants <- read.csv(here("Data", "participants.csv"))
+items <- read.csv(here("Data", "items.csv"))
+vocabulary <- read.csv(here("Data", "vocabulary.csv"))
+responses <- read.csv(here("Data", "responses.csv"))
 
@@
 
-We collected data from `r dat_raw$id_db %>% unique %>% length() %>% printnum()` bilinguals from the Metropolitan Area of Barcelona, between `r min(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`.
+```{r participants}
+n_participants <- length(unique(participants$id))
+dates <- range(participants$time_stamp, na.rm = TRUE)
+sex <- participants %>% drop_na(sex) %>% count(sex) %>% pull(n) %>% set_names(c("female", "male"))
+
+```
+
+We collected data from `r printnum(n_participants)` bilinguals (`r sex[1]` female), from the Metropolitan Area of Barcelona, between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
 All families gave informed consent before participating.
@@
 Fig.
-\@ref(fig:lpdist) illustrates the distribution of participants across language profiles and ages.
+\@ref(tab:lps) illustrates the distribution of participants across language profiles and ages.
 
-```{r lpdist, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r lps, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-dat_participants %>%
-  filter(lp %in% "Bilingual") %>% 
-  mutate(bilingualism = cut_interval(
-    bilingualism, n = 6,
-    labels = paste0(seq(0, 50, by = 10), "%"))) %>%
-  ggplot(aes(x = age_bin, y = bilingualism)) +
-  stat_bin_2d(colour = "white") +
-  scale_x_discrete(guide = guide_axis(n.dodge = 1)) +
-  scale_fill_viridis_c(option = "plasma") +
-  labs(x = "Age (months)", y = "L2 exposure", fill = "Participants") +
-  theme_classic() +
-  theme(axis.text = element_text(colour = "black"))
+participants %>%
+  mutate(age = cut_width(age, width = 4, boundary = 0),
+         doe = round(doe, 1)) %>%
+  count(age, doe) %>% 
+  mutate(doe = scales::percent(doe)) %>% 
+  pivot_wider(names_from = doe, values_from = n) %>% 
+  gt() %>% 
+  tab_spanner(columns = 2:7, label = md("**Exposure to L2**")) %>% 
+  fmt_missing(columns = everything(), missing_text = 0) %>% 
+  cols_label(age = md("**Age (months)**"))
+
 ```
@@
 
-We implemented an on-line questionnaire using `formr` [@arslan2020], divided in three forms: (1) language questionnaire, (2) demographic survey, and (3) Catalan and Spanish vocabulary checklists.
-Vocabulary checklists consisted in two lists of words (one in Catalan and one in Spanish).
+```{r items}
+pool <- pool %>% 
+  drop_na(cognate)
+n_categories <- distinct(pool, category) %>% nrow()
+n_items <- length(unique(pool$item))
+n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
+n_item_cognate <- count(pool, language, cognate) %>% 
+  pull(n) %>% 
+  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
+
+```
+
+We implemented an on-line questionnaire using `formr` [@arslan2020], divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000] and consisted in two lists of words, one in Catalan and one in Spanish).
 Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
 When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/functional categories: Action words, adventures, animals, body parts, clothes, colour, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles.
+We included items from a diverse sample of `r n_categories` semantic/functional categorie (see Appendix 1).
 We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates).
+The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
 
@@
 Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
+Table 2 in Appendix 1 shows the distribution of items accross questionnaire versions.
 
@@
 ```{r n_responses}
-n_obs <- nrow(dat_responses)
-sum_obs <- dat_responses %>%
-  group_by(item) %>%
+n_obs <- nrow(responses)
+sum_obs <- responses %>%
+  group_by(te) %>%
   summarise(n = n(), .groups = "drop") %>%
@@
 
-We gathered `r printnum(n_obs, digits = 0)` responses.
-Items received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$max, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`.
+We gathered `r printnum(n_obs, big.mark = ",")` responses.
+Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
 We modelled the probability of participants' being reported to understand/produce a translation equivalent adjusting by:
@@
 
-dat_items %>%
-  drop_na(cognate_rater1) %>%
+pool %>%
   filter(include) %>%
-  select(A, B, C, D, cognate_rater1, frequency, language) %>% 
-  mutate(cognate_rater1 = ifelse(cognate_rater1, "Cognate", "Non-cognate")) %>%
-  pivot_longer(c(A, B, C, D), names_to = c("list"), values_to = "n") %>% 
-  filter(n) %>% 
-  group_by(list, language, cognate_rater1) %>% 
-  summarise(n = sum(n, na.rm = TRUE),
+  select(version, cognate, frequency = frequency_zipf, language) %>% 
+  mutate(cognate = ifelse(cognate, "Cognate", "Non-cognate")) %>%
+  group_by(version, language, cognate) %>% 
+  summarise(
+    n = sum(n, na.rm = TRUE),
     mean = mean(frequency, na.rm = TRUE),
@@
     max = max(frequency, na.rm = TRUE),
-            .groups = "drop") %>%
+    .groups = "drop"
+  ) %>%
   ungroup() %>% 
-  rename(cognateness = cognate_rater1) %>%
-  relocate(list, language, cognateness, n, mean, sd, min, max) %>%
-  arrange(list, language, cognateness) %>%
-  mutate(list = paste("List", list)) %>% 
-  pivot_wider(names_from = cognateness, values_from = c(n, mean, sd, min, max)) %>%
+  relocate(version, language, cognate, n, mean, sd, min, max) %>%
+  arrange(version, language, cognate) %>%
+  mutate(version = collapse(version, ",")) %>% 
+  pivot_wider(names_from = cognate, values_from = c(n, mean, sd, min, max)) %>%
   clean_names() %>%

### 2021-01-13 Rmd/04_appendix_2.Rmd
Merge manuscript sections

--- 2020-06-15 Code/Rmd/data_analysis_notes.Rmd
@@
 ---
-title: "Notes on data anlaysis"
-output: html_notebook
+title: 'Appendix 2: Model'
+output: pdf_document
 ---
 
-The aim of this study is two-fold.
-First, we want to test whether word-forms are acquired earlier by bilinguals when their translation equivalent is form-similar.
-Second, we want to test if the difference in age of acquisition between both translation equivalents is smaller when they are form-similar.
-
-In order to gather information about acquisition of vocabulary, many developmental studies rely on vocabulary inventories such as the McArthur-Bates Communicative Developmental Inventories (*CDI* from now on).
-These questionnaires are administered to parents or caretakers, which are presented with a printed list of words.
-For each word, parents must report whether they think the child understands and/or produces the word.
-Age of acquisition (AoA) is then defined as the age at which the proportion of children that are able to understand or produce a given word exceeds a given threshold.
-For instance, the word *door* may be considered to be aquired when at least 60% of the participants in our sample are reported to understand the word.
-
-We find two reasons to consider this approach suboptimal.
-First, the value of the cut-off is arbitrary.
-Although most studies use thresholds between 50% and 60%, this varibility makes it difficult to build cumulative knowledge from the listerature when the un-aggregated data (not yet filtered by the cut-off) is available.
-Second, it is unlikely that the acquisition status of a word is well represented by a dicotomy between acquired and not acquired.
-Word learning is a multidomain process where many processes take place.
-It is more likely that the trajectory of acquisition of a given word is better defined by a curve that characterises the probability of a child being able to understand or produce a word across ages.
-Third, by considering AoA as a point estimate, other dimensions that characterise the trajectory of aquisition are being neglected.
-For instance, it seems reasonable to think that some words are acquired faster the others.
-This means that the developmental curve of some words is steeper that that of others.
-For instance, the effect of word frequency on AoA has been extensively: more frequent words are aquired earlier then less frequent words.
-We suggest that this effect could also be reflected on steeper acquisition trajectories.
-
-Recent studies have explored the role of word frequency, as well as other well know predictors (e.g., concreteness) on age of acquisition by going beyond arbitraty cut-offs.
-
-Description of @braginsky2019 and @jones2019.
-
-Both studies have made substantial contributions to the study of the internal structure of developing lexicon by performing item-wise analysis.
-However, neither of them adresses the potential non-linearity of the developmental trajectory of words.
-By fitting a linear (multilevel) model, two main parameters of interest are estimated.
-Both parameters characterise the line that defines the effect of each predictor (e.g., age, frequency) on the outcome variable (e.g., proportion of children reported the understand the word, probability that a given child will be reported to undestand the word).
-A first parameter is the intercept, which provides information about the value of the outcome when all predictors are at baseline.
-A second set of parameters informs us about the contribution of each of the predictor or combination of predictors.
-This model assumes that *age* holds a linear relationship with the proportion of infants that have acquired the word.
-We consider that this assumed linear relation does not capture well enought the consitribution of *age* on the acquisition status of the word.
-Instead, we propose that this relationship is better defined by logarithmic curves.
-
-Logarithmic curves are characterised by three parameters: 1) Upper asymptote, 2) inflexion point, and 3) steepness at inflexion point.
-First, although it seems reasonable to assume that, as age increases, the proportion of children that have acquired a given word increases as well.
-However, it is most likely that such word will never have been acquired by all children in the population.
-This asymptotic nature of the acquisition status of the words is capture by the first parameter indicated above.
-Second, previous studies have reported that vocabulary size (both receptive and productive) remains relatively stable during the first year and a half of life, and then grows exponentially until one year after, when it stabilises again.
-Later, vocabulary size keeps growing, but a slower pace.
-This means that, around 20 to 24 months of age, the developmental trajectory of most of the words that will be part of the vocabulary of the child during the next 3 years, will find an inflexion point, where the curve is steepest.
-If this is to be true, it could be reasonable to consider this point as the critical age at which the acquisition status of a given word is evolving faster.
-We argue that this age point reflects overlaps better with what most researchers mean by *word-acquistion*, compared to the tradicitonal cut-off approach.
-Third, some items may differ on how steep their acquisition trajectory is in this inflexion point.
-For instance, hight-frequency words may entail a steeper curve than low_frequency words, thus leading to high-frequency words to be acquired faster than low-frequency word once the inflexion point is reached.
-
-We consider that logarithmic curves are more suitable to developmental trajectories of individual items, because they capture cross-item variability in age of acquisition and speed of acquisition, at the same time they account for non-linear trends in such trajectories, as well as for the existence of an upper asymptotic boundary in the proportion of infants that have acquired each word.
-
-### Model specification
-
-In this study, we model developmental tracjectories of individual items using logarithmic curves.
-We assume a common upper boundary for all items (i.e., 1).
-We aim to estimate the inflexion point (i.e., age at which the curve is steepest) and the steepness (i.e., how fast the proportion of infants that know the word increases) at the inflexion point.
-We will follow a multilevel Bayesian approach to estimate the mid point and steepness of each word.
-We assume that mid-points will be generated from a normal distribution with mean 7.5 (age bin corresponsing to 22 months of age) and standard deviation 1.
-This means that the inflexion point (i.e., age of maximum steepness), 68% of the items will be located within 1 stand deviation at 22 months.
-Our model will update collected data to update the estimated mid-point of the words, by using word frequency, cognate status, and language profile to predict the mean of the distribution.
-We will include by-item random intercepts, and by-item random slopes for cognate status and language profile.
-We will include a correlation parameter for each grouping variable (i.e.
-cognate status and language profile), but we won't allow correlation across grouping variables.
-We assume that the steepness of each item will be generated from a normal distribution with mean 1 and standard deviation 0.5.
-This means that for 68% of the items, the proportion of children that understand a given word will be close 1 standard deviation  far maximum from in a 100% at the age of maximum steepness.
-Our model can be described as follows:
-
-$$Y_{ij} = \frac{1}{1 + e \exp{-k_i(Age-x_{0ij})}}$$
-$$Y_{i} \sim Binomial(p, N)$$
-$$k \sim Normal(1, 0.5)$$
-$$x_{0ij} \sim N{\beta_0 + \beta_{0j} + (\beta_1 + \beta_{1j}) \times Profile_{ij} + (\beta_2 + \beta_{2j}) \times Cognateness_{ij}, \sigma^2}$$
-$$\beta_0 \sim Normal(0, 1)$$
-$$\beta_{0j} \sim Normal(0, \tau^2_{0})$$
-$$\beta_{1j} \sim Normal(0, \tau^2_{1})$$
-$$\beta_2 \sim Normal(0, 1)$$
-$$\beta_{2j} \sim Normal(0, \tau^2_{2})$$
-
-Where:
-
-* $Y_{ij}$ is the proportion of children that have been reported to have acquired the item $i$ at age $j$; assumed to be distributed following a Gaussian distribution with mean 0 and variance $\sigma^2$.
-* $k_i$ is the steepness of the curve at the inflexion point of item $i$, assumed to be distributed following a Gaussian distribution with mean 1 and standard deviation 0.5.
-* $x_{0ij}$ is the age at which the inflexion point is located in item $i$ and cognate status $j$. This parameter is defined by a linear model the includes the following parameters:
-     * $\beta_0$ is the intercept of the model, that is, the grand mean age of of infants when the curve of item $i$ is at its steepest point, when all other predictors are at baseline; it is assumed to be distributed following a Gaussian distribution with mean 7.5 (age bin corresponding at 22 months of age) and standard deviation 4.
-     * $\beta_{0j}$ is the intercept of the model, that is, the grand mean age of of infants when the curve of item $i$ is at its steepest point, when all other predictors are at baseline; it is assumed to be distributed following a Gaussian distribution with mean 7.5 (age bin corresponding at 22 months of age) and standard deviation $\tau^2_{0}$.
-     * $\beta_1$ is the slope of *language profile* on the outcome; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation 1.
-     * $\beta_{1j}$ is the slope of *language profile* on cognate status $j$; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation $\tau^2_{1}$.
-     * $\beta_2$ is the slope of *cognateness* on the outcome; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation 1.
-     * $\beta_{2j}$ is the slope of *cognateness* on cognate status $j$; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation $\tau^2_{2}$.
-     * $\beta_3$ is the slope of the interaction between *language profile* and *cognateness* on the outcome; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation 1.
-     * $\beta_{3j}$ is the slope of the interaction term between *language profile* and *cognateness* on cognate status $j$; it is assumed to be distributed following a Gaussian distribution with mean 0 and standard deviation $\tau^2_{3}$.
-
-We will run separate models for items in the most dominant and less dominant language, and for receptive and productive vocabulary.
-
-### Setting priors
-
-We will use the Bayesian framework to estimate the parameters of our model.
-This involves the use of the Bayes' rule to compute the probability ($\pi$) of all values our set of estimated parameters ($\theta$) can take.
-To do this, it combines prior information coming from scientific expertise (also known as priors, $\pi(\theta)$), and the likelihood of each value the parameter can take, given the observed data ($\pi(y|\theta)$)).
-The product of both prior and likelihood is then standardised using the marginal likelihood of the parameter sampling space ($\pi(y)$).
-The combination of both sources of information leads to what is known as the posterior distribution of each parameter ($\pi(\theta|y)$), which informs us of the likelihood of each value the parameters can take, given our prior knowledge about it, and the observed data.
-We can extract information from this posterior distribution in several ways.
-We will use probabilistic posterior sampling to approximate this distribution and characterise it.
-
-$$\pi(\theta | y) = \frac{\pi(y | \theta)\pi(\theta)}{\pi(y)}$$
-
-We need to set priors for $k_i$, $\sigma$, $\tau^2_{0}$, $\tau^2_{1}$, $\tau^2_{2}$, $\tau^2_{3}$, $\beta_{0j}$, $\beta_{1j}$, and $\beta_{2j}$ and $\beta_{3j}$. We will specify the following priors:
-
-    * $k_i$: `normal(1, 0.5)`
-    * $\beta_{1}$: `normal(0, 1)`
-    * $\beta_{2}$: `normal(0, 1)`
-    * $\beta_{3}$: `normal(0, 1)`
-    * $\sigma$: `student_t(3, 0, 10)`
-    * $\tau^2_{0}$: `student_t(3, 0, 10)`,
-    * $\tau^2_{1}$: `student_t(3, 0, 10)`
-    * $\tau^2_{2}$: `student_t(3, 0, 10)`
-    * $\tau^2_{3}$: `student_t(3, 0, 10)`
-
-### Model fitting
-
-We run our model using the `brm()` function in the `brms` R package [@bruckner2018], which uses the Stan environment [@stan2019] in the back end to approximate the posterior distribution via Hamiltonian Monte Carlo (HMC).
-
-### Comparing models
-*
-To analyse the effect of the interaction between *language profile* and *cognate status*, we will fit three models.
-A first minimal model will not include any of the predictors of interest.
-A second null model will include *language profile* as a population effect A third complete model will also include *cognate status*, and its interaction with *language profile* as population effects.
-All models will include *item* as grouping variable.
-We will specify, in each model, a maximal random effects structure that includes by-item intercepts, slopes, and correlation parameters [@barr2013a].
-We will then compare the three models following the workflow suggested by @schad2019.
-The minimal model will be expanded to the null model, and then the null model be expanded to the complete model.
-At each stage, we will check whether the expanded model is more adequate than the previous.
-In case the expanded model proves more inadequate, we will fall back to the simpler, previous model.
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```

### 2021-01-13 Rmd/trajectories_manuscript.Rmd

@@
 
-bibliography      : ["../References/trajectories.bib"]
+bibliography      : ["trajectories.bib"]
 
@@
 
-```{r, child= "00_introduction.Rmd"}
+# Introduction
+
+Learning words involves two key steps: (1) the encoding of a word form (e.g, /dɒɡ/)
+ and (2) its association with a referent (e.g., a DOG).
+There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side [@bergleson2012].
+At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered [@jusczyk1995], as opposed to when only part of the word is uttered.
+At 12 months, infants show sensitivity to both vowel and consonant mispronunciations [@mani2010a], looking shorter to named pictures when their label is mispronounced.
+Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones.
+One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon [@mcclelland2003].
+
+Words are not learnt in isolation, but in contexts rich in word tokens and referents.
+The structure of toddlers' lexicon reflects this high connectivity [@wojcik2018; @hills2010; @fourtassi2020].
+At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing [@mani2010b; @mani2011].
+The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming  [@styles2009; @arias-trejo2013; @wojcik2013], and show evidence of inhibitory links between lexical-semantic entries [@chow2016].
+By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages [@chow2017].
+Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words [@chow2017; @rama2013], and  the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers.
+This suggests that the emergence of rich lexical networks -- both at the phonological and the semantic levels -- grows as a function of the number of words acquired by the toddler.
+
+The structure of the lexicon also impacts the order of acquisition of new words.
+Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills2009].
+@borosky2016 and @borovsky2018 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
+This points to the structure of the lexicon facilitating the strategies children engage during word learning.
+
+The case of bilingual children (here defined as those learning two languages from birth) presents an opportunity to study how lexical acquisition takes placed under a more complex environment, as they  face the challenge of learning two distinct sets of words -- one for each language -- that partially overlap in sound and meaning.
+
+EARLY STAGES IN BILINGUAL LANGUAGE ACQUISITION
+
+When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
+When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev_influence_1977; @doyle1977some; @rosenblum1983word; @fernandez1992bilingual; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
+
+@floccia_i_2018, gathered parental vocabulary estimates [@hamilton2000infant] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
+They computed the average phonological similarity between translation equivalents (TEs from now on; e.g.,  *table* in English and *tafel* in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
+Using a similar measure of lexical similarity, @blom_cross-language_2019 extended these results to bilingual children aged three to 10, and reported a positive association between the lexical distance and comprehensive vocabulary size: bilinguals learning lexically close languages showed similar vocabulary sizes to those of monolinguals, but those learning the more distant languages showed lower vocabulary sizes.
+
+These results suggest that linguistic distance between languages plays an important role during bilingual lexical acquisition, and that this effect operates within TEs.
+What mechanisms underlie this effect, and why they seem to operate differently across comprehension and production, or across the dominant and the non-dominant language, remain open issues.
+@floccia_i_2018 pointed to parallel activation as a candidate mechanism behind this facilitation effect.
+The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., *cat*, in English) its corresponding translation in the other language is activated too [e.g., *gato*, in Spanish.
+There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell1991mediated; @spivey1999cross; @costa2000lexical; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh2014one; @von_holzen_impact_2019; @bobb_co-activation_2020; see @costa2005facilitatory for review], and children [e.g.
+@von_holzen_language_2012; @poulin2013lexical; @jardak2019labels; @bosma2020cognate].
+A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
+
+@floccia_i_2018 argue that the degree of phonological similarity between TEs (*cognateness* from now on) of both languages should lead to increased parallel activation of both languages during language exposure, facilitating the acquisition of words in both languages, and leading participants leading two lexically close languages to show larger vocabulary sizes than those learning two lexically distant languages.
+This account, however, neglects the fact that for parallel activation to take place and play a facilitation role, lexical representations of both pairs of the TE must have been already established in the lexicon, and their corresponding phonological forms must have been encoded.
+This is not the case during lexical acquisition, where at least one of the members of the pair has not being acquired yet.
+A more lenient definition of parallel activation could adjust to the context of lexical acquisition more easily: when only one of the members of the TE has been acquired, its lexical representation is activated when, in the presence of its referent, a phonologically similar label is uttered.
+Consider the case of a Catalan-Spanish bilingual that has already learned the word *gat* (Catalan for *cat*), but not *gato* (in Spanish).
+It is possible that, when presented with *gato* in the presence of a cat, she maps this novel label to the familiar word *gat*, and creates a lexical representation for *gato* as a synonym for *gat*.
+The case of a non-cognate is differing.
+Consider now the case of a toddler that has already learnt *gos* (Catalan for *dog*), but not *perro* (Spanish).
+In this case, it would not be possible to map both labels via phonology, give their lack of phonological similarity.
+
+Under this account, cognates (i.e., phonologically similar TEs) would be acquired earlier than non-cognates, but this effect would only play a role once one of the members of the TE has been acquired.
+The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse.
+@schelletter2002effect's reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
+@bosch_first_2014, showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months.
+The low number of words included in the analyses of both studies, and the lack of adjustment for lexical frequency, limits severely the strength of any conclusions concerning the difference in age of acquisition of cognates and non-cognates.
+On the other hand @floccia_i_2018 worked with aggregated estimates of participants' vocabulary size, losing item-level information that could provide information about change in age of acquisition of TEs associated with their degree of phonological similarity.
+Therefore, their results can only be considered as compatible with cognateness playing a role on age of acquisition.
+In this study, we aimed at overcoming these pitfalls by using unaggregated data from individual responses, testing the role of cognateness explicitly and adjusting for lexical frequency.
+
+It also follows from the fact that cognateness operates *within* TEs that the effect of cognateness can only play a role once one of the members of the TE has been acquired, and its phonological form is available during the acquisition of its translation.
+Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language [@floccia_i_2018].
+@dijkstra2013 tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE.
+Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian.
+This suggests that infants that were mostly exposed to Dutch used the phonology of  dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
+A study by @dijkstra2013 showed that the degree of phonological similarity between 
+the performance of  was better for cognates than for non-cognates, only for those children
+
+Finally, previous studies have reported that the properties that describe the form of a word play a stronger role in production than comprehension than in production.
+This is the case of the number of phonemes [@braginsky_consistency_2019], phonological neighbourhood density [@jones_children_2019].
+These results converge with @floccia_i_2018's finding that the effect of the average cognateness between languages was larger in production than in comprehension^[Although @bosch_first_2014 found that participants in their sample produced cognates earlier than non-cognates, no analyses were conducted on comprehensive data, and therefore they cannot be taken as supporting evidence for a larger effect of cognateness on production than in comprehension].
+Accordingly, we expected the role of cognateness to be more central in production than in comprehension.
+
+In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.
+
+```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+participants <- read.csv(here("Data", "participants.csv"))
+items <- read.csv(here("Data", "items.csv"))
+vocabulary <- read.csv(here("Data", "vocabulary.csv"))
+responses <- read.csv(here("Data", "responses.csv"))
 
@@
 
-```{r, child= "01_method.Rmd"}
+# Method
+
+## Participants
+
+```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+n_participants <- length(unique(participants$id))
+dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
+sex <- participants %>% drop_na(sex) %>% count(sex) %>% pull(n) %>% set_names(c("female", "male"))
 
@@
 
-```{r, child= "02_results.Rmd"}
+We collected data from `r printnum(n_participants)` bilinguals (`r sex[1]` female), from the Metropolitan Area of Barcelona, between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
+All families gave informed consent before participating.
+This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
+We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language.
+We excluded participants with >10% exposure to a third language.
+Fig.
+\@ref(tab:lps) illustrates the distribution of participants across language profiles and ages.
+
+## Questionnaire
+
+```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+pool <- multilex::pool %>% 
+  drop_na(cognate)
+n_categories <- distinct(pool, category) %>% nrow()
+n_items <- length(unique(pool$item))
+n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
+n_item_cognate <- count(pool, language, cognate) %>% 
+  pull(n) %>% 
+  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
+```
+
+We implemented an on-line questionnaire using `formr` [@arslan_formr_2020], divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton_infant_2000] and consisted in two lists of words, one in Catalan and one in Spanish).
+Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
+We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
+We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
+The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
+
+For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default).
+Participants filled a long or a short version of the questionnaire.
+Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items.
+Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish).
+Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
+Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
+
+## Data analysis
+
+```{r n_responses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+n_obs <- nrow(responses)
+sum_obs <- responses %>%
+  group_by(te) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  summarise(
+    mean = mean(n),
+    sd = sd(n),
+    min = min(n),
+    max = max(n)
+  )
 
@@
 
-```{r, child= "03_discussion.Rmd"}
+We gathered `r printnum(n_obs, big.mark = ",")` responses.
+Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
+We modelled the probability of responses ($y_i$) belonging to any of the three possible categories $k$--1 = *No* or say, 2 = *Understands*, and 3 = *Understands and Says*-- using an ordinal regression model with a cumulative logit link function [@agresti_analysis_2010; @burkner_ordinal_2019].
+This model estimates the probability of response $y_i$ to belong to category $k$ or lower.
+In our case, this will result in the estimation of the probability of $y_i \leq 1$,  $y_i \leq 2$, and  $y_i \leq 3$, where
 
-```
+$$
+p_k = Pr(y_i = k) = Pr(y_i \leq k) - Pr(y_i \leq k-1).
+$$
 
-```{r, child= "04_appendix_1.Rmd"}
+We include several predictors of interest to adjust this probability to participants' and items' (lexical frequency, dominance, cognateness) characteristics. These predictors are:
 
-```
+* Age of participant in months (`age`, $Age$) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30 and centred around the mean
+* Degree of bilingualism of participant (`bilingualism`, $Bilingualism$) computed as the percentage of exposure to a second language (Spanish for participants exposed to >50% to Catalan an *vice versa*) rounded to units and centered around the mean
+* Lexical frequency of item (`frequency`, $Frequency$) retrieved from SUBTLEX-CAT [@boada_subtlex-cat_2019] for Catalan words and from SUBTLEX-ESP [@cuetos_subtlex-esp_2011] for Spanish words, expressed as Zipf scores [@heuven_subtlex-uk_2014], and centred around the mean
+* Language the item belongs to (`dominance`, $Dominance$) labelled as `L1` if the item belongs to the dominant language (e.g., a Catalan item is labelled as `L1` for Catalan-dominant participants and as `L2` for Spanish-dominant participants), coded as `L2` = -0.5 and `L1` = 0.5
+* Cognateness of the item (`cognate`, $Cognate$), coded as `Non-cognate` = -0.5 and `Cognate` = 0.5
 
-```{r, child= "04_appendix_2.Rmd"}
+We adopted a Bayesian approach toward statistical inference, which allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model.
+Our extended model can be formalised as follows:
 
-```
+$$
+\begin{aligned}
+y_i \sim Ordered(p) \\
+log\frac{Pr(y_i \leq k)}{1-Pr(y_i \leq k)} &= \alpha_k - \dots \\
+& \beta_1 \times Age_i + \dots \\
+& \beta_2 \times Frequency_i + \dots \\
+& \beta_3 \times Dominance_i + \dots \\
+& \beta_4 \times Bilingualism_i + \dots \\
+& \beta_5 \times Cognate_i + \dots \\
+& \beta_6 \times (Dominance_i \times Bilingualism_i) + \dots \\
+& \beta_7 \times (Dominance_i \times Cognate_i) + \dots \\
+& \beta_8 \times (Bilingualism_i \times Cognate_i) + \dots \\
+& \beta_9 \times (Dominance_i \times Bilingualism_i \times Cognate_i), \\
+\alpha \sim Normal(0, 10) \\
+\beta_{0-9} \sim Normal(0, 10)
+\end{aligned}
+$$
 
-```{r, child= "04_appendix_3.Rmd"}
+Where:
 
-```
+* $k$ is one of the possible categories (coded as ) responses can take
+* $y_i$ is the `response` value of observation $i$
+* $\alpha_k$ is the mean probability of responses to belong to category $k$
+* $\beta_{1, \dots, 9}$ are the coefficients of the predictors
+
+We implemented the model using the R package `brms` [@burkner_brms_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 4,000 iterations (including 2,000 warm-up iterations per chain).
+We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal due to the large amount of observations introduced in the model.
+We report the 95% credible intervals (CrI) of the estimated posterior distribution of each parameter, which denotes a range of values we are 95% certain contain the true value, given our prior and observed data.
+We report analyses on the probability scale, having transformed the log-odds returned by the model using an inverse logit transformation.
+
+We then fit increasingly simpler models dropping predictors one by one in the following order: (1) `dominance:bilingualism:cognate`, (2) `bilingualism:cognate`, (3) `dominance:cognate:`, (4) `cognate`, (5) `dominance:bilingualism`, (6) `bilingualism`, (7) `dominance`, (8) `frequency`, and (9) `age`.
+We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017] and widely applicable information criterion (WAIC) [@watanabe_asymptotic_2010].
+We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95% credible intervals and testing whether this interval excluded 0.
+We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@lenth_emmeans_2020].
+Data processing and visualisation was done in R [@rcoreteam_r_2020] using the `tidyverse` family of packages [@wickham_welcome_2019], and posterior samples were extracted using the `tidybayes` R package [@kay_tidybayes_2020].
+
+# Results
+
+# Discussion
 

### 2021-01-13 Rmd/trajectories_manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
  pdfauthor={Gonzalo García-Castro1, Daniela Avila-Varela1, \& Núria Sebastian-Galles1},
  pdflang={en-EN},
  pdfkeywords={lexical acquisition, vocabulary, bilingualism},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Cross-linguistic similarity and  word acquisition}
\keywords{lexical acquisition, vocabulary, bilingualism\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{The role of cross-linguistic lexical similarity on bilingual word acquisition}
\author{Gonzalo García-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Núria Sebastian-Galles\textsuperscript{1}}
\date{}

\authornote{

Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain

Correspondence concerning this article should be addressed to Gonzalo García-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain. E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain}

\abstract{
Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al.~(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Learning words involves two key steps: (1) the encoding of a word form (e.g, /dɒɡ/)
and (2) its association with a referent (e.g., a DOG). There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side ({\textbf{???}}). At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered ({\textbf{???}}), as opposed to when only part of the word is uttered. At 12 months, infants show sensitivity to both vowel and consonant mispronunciations ({\textbf{???}}), looking shorter to named pictures when their label is mispronounced. Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones. One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon ({\textbf{???}}).

Words are not learnt in isolation, but in contexts rich in word tokens and referents. The structure of toddlers' lexicon reflects this high connectivity ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}). At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing ({\textbf{???}}; {\textbf{???}}). The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}), and show evidence of inhibitory links between lexical-semantic entries ({\textbf{???}}). By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages ({\textbf{???}}). Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words ({\textbf{???}}; {\textbf{???}}), and the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers. This suggests that the emergence of rich lexical networks -- both at the phonological and the semantic levels -- grows as a function of the number of words acquired by the toddler.

The structure of the lexicon also impacts the order of acquisition of new words. Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next ({\textbf{???}}). ({\textbf{???}}) and ({\textbf{???}}) showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired. This points to the structure of the lexicon facilitating the strategies children engage during word learning.

The case of bilingual children (here defined as those learning two languages from birth) presents an opportunity to study how lexical acquisition takes placed under a more complex environment, as they face the challenge of learning two distinct sets of words -- one for each language -- that partially overlap in sound and meaning.

EARLY STAGES IN BILINGUAL LANGUAGE ACQUISITION

When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals). When both languages are taken into account, bilingual children seem to know, at least, as many words as monolinguals do (Ben-Zeev, 1977; Bialystok, Luk, Peets, \& Yang, 2010; Blom et al., 2019; Core, Hoff, Rumiche, \& Señor, 2013; Doyle \& others, 1977; Fernandez, Pearson, Umbel, Oiler, \& Molinet-Molina, 1992; Hoff et al., 2012; Hoff, Rumiche, Burridge, Ribot, \& Welsh, 2014; Pearson \& Fernández, 1994; Rosenblum \& Pinker, 1983; but see Pearson, Fernández, \& Oller, 1993; Houwer, Bornstein, \& Putnick, 2014).

Floccia et al. (2018), gathered parental vocabulary estimates ({\textbf{???}}) from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages). They computed the average phonological similarity between translation equivalents (TEs from now on; e.g., \emph{table} in English and \emph{tafel} in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language. Using a similar measure of lexical similarity, Blom et al. (2019) extended these results to bilingual children aged three to 10, and reported a positive association between the lexical distance and comprehensive vocabulary size: bilinguals learning lexically close languages showed similar vocabulary sizes to those of monolinguals, but those learning the more distant languages showed lower vocabulary sizes.

These results suggest that linguistic distance between languages plays an important role during bilingual lexical acquisition, and that this effect operates within TEs. What mechanisms underlie this effect, and why they seem to operate differently across comprehension and production, or across the dominant and the non-dominant language, remain open issues. Floccia et al. (2018) pointed to parallel activation as a candidate mechanism behind this facilitation effect. The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., \emph{cat}, in English) its corresponding translation in the other language is activated too {[}e.g., \emph{gato}, in Spanish. There is a vast body of evidence supporting language non-selective lexical access in both adults (e.g., Dell \& O'Seaghdha, 1991; {\textbf{???}}; Bobb, Von Holzen, Mayor, Mani, \& Carreiras, 2020; Dijkstra, 2005; Hoshino \& Kroll, 2008; Kroll, Gullifer, \& Rossi, 2013; Singh, 2014; Spivey \& Marian, 1999; Thierry \& Wu, 2007; Von Holzen, Fennell, \& Mani, 2019; Yudes, Macizo, \& Bajo, 2010; see Costa, Santesteban, \& Caño, 2005 for review), and children (e.g.~Von Holzen \& Mani, 2012; Bosma \& Nota, 2020; Jardak \& Byers-Heinlein, 2019; Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013). A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.

Floccia et al. (2018) argue that the degree of phonological similarity between TEs (\emph{cognateness} from now on) of both languages should lead to increased parallel activation of both languages during language exposure, facilitating the acquisition of words in both languages, and leading participants leading two lexically close languages to show larger vocabulary sizes than those learning two lexically distant languages. This account, however, neglects the fact that for parallel activation to take place and play a facilitation role, lexical representations of both pairs of the TE must have been already established in the lexicon, and their corresponding phonological forms must have been encoded. This is not the case during lexical acquisition, where at least one of the members of the pair has not being acquired yet. A more lenient definition of parallel activation could adjust to the context of lexical acquisition more easily: when only one of the members of the TE has been acquired, its lexical representation is activated when, in the presence of its referent, a phonologically similar label is uttered. Consider the case of a Catalan-Spanish bilingual that has already learned the word \emph{gat} (Catalan for \emph{cat}), but not \emph{gato} (in Spanish). It is possible that, when presented with \emph{gato} in the presence of a cat, she maps this novel label to the familiar word \emph{gat}, and creates a lexical representation for \emph{gato} as a synonym for \emph{gat}. The case of a non-cognate is differing. Consider now the case of a toddler that has already learnt \emph{gos} (Catalan for \emph{dog}), but not \emph{perro} (Spanish). In this case, it would not be possible to map both labels via phonology, give their lack of phonological similarity.

Under this account, cognates (i.e., phonologically similar TEs) would be acquired earlier than non-cognates, but this effect would only play a role once one of the members of the TE has been acquired. The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse. Schelletter (2002)`s reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE. Bosch and Ramon-Casas (2014), showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months. The low number of words included in the analyses of both studies, and the lack of adjustment for lexical frequency, limits severely the strength of any conclusions concerning the difference in age of acquisition of cognates and non-cognates. On the other hand Floccia et al. (2018) worked with aggregated estimates of participants' vocabulary size, losing item-level information that could provide information about change in age of acquisition of TEs associated with their degree of phonological similarity. Therefore, their results can only be considered as compatible with cognateness playing a role on age of acquisition. In this study, we aimed at overcoming these pitfalls by using unaggregated data from individual responses, testing the role of cognateness explicitly and adjusting for lexical frequency.

It also follows from the fact that cognateness operates \emph{within} TEs that the effect of cognateness can only play a role once one of the members of the TE has been acquired, and its phonological form is available during the acquisition of its translation. Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language (Floccia et al., 2018). ({\textbf{???}}) tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE. Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian. This suggests that infants that were mostly exposed to Dutch used the phonology of dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
A study by ({\textbf{???}}) showed that the degree of phonological similarity between
the performance of was better for cognates than for non-cognates, only for those children

Finally, previous studies have reported that the properties that describe the form of a word play a stronger role in production than comprehension than in production. This is the case of the number of phonemes (Braginsky, Yurovsky, Marchman, \& Frank, 2019), phonological neighbourhood density (Jones \& Brandt, 2019). These results converge with Floccia et al. (2018)'s finding that the effect of the average cognateness between languages was larger in production than in comprehension\footnote{Although Bosch and Ramon-Casas (2014) found that participants in their sample produced cognates earlier than non-cognates, no analyses were conducted on comprehensive data, and therefore they cannot be taken as supporting evidence for a larger effect of cognateness on production than in comprehension}. Accordingly, we expected the role of cognateness to be more central in production than in comprehension.

In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We collected data from 349 bilinguals (167 female), from the Metropolitan Area of Barcelona, between 28th October, 2019 and 09th January, 2021. All families gave informed consent before participating. This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX. We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language. We excluded participants with \textgreater10\% exposure to a third language. Fig. \ref{tab:lps} illustrates the distribution of participants across language profiles and ages.

\hypertarget{questionnaire}{%
\subsection{Questionnaire}\label{questionnaire}}

We implemented an on-line questionnaire using \texttt{formr} (Arslan, Walther, \& Tata, 2020), divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists. Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (Hamilton, Plunkett, \& Schafer, 2000) and consisted in two lists of words, one in Catalan and one in Spanish). Items in one language were translation equivalents of the items in the other (e.g., whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the word \emph{perro} was included in the Spanish inventory), roughly following a one-to-one mapping. When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar} and \emph{terminar}), or merged them into the same one (e.g., Spanish \emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}). We included items from a diverse sample of 26 semantic/functional categories (see Appendix 1). We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines. The Catalan inventory contained 778 items (196 cognates, 582 non-cognates) and the Spanish inventory contained 781 (197 cognates, 584 non-cognates).

For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default). Participants filled a long or a short version of the questionnaire. Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items. Each list contained a different set of \textasciitilde400 translation equivalents (\textasciitilde400 in Catalan, \textasciitilde400 in Spanish). Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists. Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions. Table 2 in Appendix 1 shows the distribution of items across questionnaire versions. We excluded from the analysis multi-word items (e.g., \emph{barrita de cereales} {[}cereal bar{]}) and items that included more than one word-form (e.g., \emph{mono / mico}). Table \ref{tab:itemsinfo} shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We gathered 181587 responses. Translation equivalents received an average of 386.36 (\emph{SD} = 228.76, , Max = 698), both languages summed together. We modelled the probability of responses (\(y_i\)) to belong to any of the three possible categories (\(k\)): 1 = \emph{No} or say, 2 = \emph{Understands}, and 3 = \emph{Understands and Says}) using an ordinal regression model with a cumulative logit link function (Agresti, 2010; Bürkner \& Vuorre, 2019). This model estimates the probability of response \(y_i\) to belong to category \(k\) or lower. In our case, this will result in the estimation of the probability of \(y_i \leq 1\), \(y_i \leq 2\), and \(y_i \leq 3\), where

\[
p_k = Pr(y_i = k) = Pr(y_i \leq k) - Pr(y_i \leq k-1).
\]

We include several predictors of interest to adjust this probability to participants' and items' (lexical frequency, dominance, cognateness) characteristics. These predictors are:

\begin{itemize}
\tightlist
\item
  Age of participant in months (\texttt{age}, \(Age\)) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30 and centred around the mean
\item
  Degree of bilingualism of participant (\texttt{bilingualism}, \(Bilingualism\)) computed as the percentage of exposure to a second language (Spanish for participants exposed to \textgreater50\% to Catalan an \emph{vice versa}) rounded to units and centered around the mean
\item
  Lexical frequency of item (\texttt{frequency}, \(Frequency\)) retrieved from SUBTLEX-CAT (Boada, Guasch, Haro, Demestre, \& Ferré, 2019) for Catalan words and from SUBTLEX-ESP (Cuetos, Glez-Nosti, Barbón, \& Brysbaert, 2011) for Spanish words, expressed as Zipf scores (Heuven, Mandera, Keuleers, \& Brysbaert, 2014), and centred around the mean
\item
  Language the item belongs to (\texttt{dominance}, \(Dominance\)) labelled as \texttt{L1} if the item belongs to the dominant language (e.g., a Catalan item is labelled as \texttt{L1} for Catalan-dominant participants and as \texttt{L2} for Spanish-dominant participants), coded as \texttt{L2} = -0.5 and \texttt{L1} = 0.5
\item
  Cognateness of the item (\texttt{cognate}, \(Cognate\)), coded as \texttt{Non-cognate} = -0.5 and \texttt{Cognate} = 0.5
\end{itemize}

We adopted a Bayesian approach toward statistical inference, which allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model. Our extended model can be formalised as follows:

\[
\begin{aligned}
y_i \sim Ordered(p) \\
log\frac{Pr(y_i \leq k)}{1-Pr(y_i \leq k)} &= \alpha_k - \dots \\
& \beta_1 \times Age_i + \dots \\
& \beta_2 \times Frequency_i + \dots \\
& \beta_3 \times Dominance_i + \dots \\
& \beta_4 \times Bilingualism_i + \dots \\
& \beta_5 \times Cognate_i + \dots \\
& \beta_6 \times (Dominance_i \times Bilingualism_i) + \dots \\
& \beta_7 \times (Dominance_i \times Cognate_i) + \dots \\
& \beta_8 \times (Bilingualism_i \times Cognate_i) + \dots \\
& \beta_9 \times (Dominance_i \times Bilingualism_i \times Cognate_i), \\
\alpha \sim Normal(0, 10) \\
\beta_{0-9} \sim Normal(0, 10)
\end{aligned}
\]

Where:

\begin{itemize}
\tightlist
\item
  \(k\) is one of the possible categories (coded as ) responses can take
\item
  \(y_i\) is the \texttt{response} value of observation \(i\)
\item
  \(\alpha_k\) is the mean probability of responses to belong to category \(k\)
\item
  \(\beta_{1, \dots, 9}\) are the coefficients of the predictors
\end{itemize}

We implemented the model using the R package \texttt{brms} (Bürkner, 2017). This package uses the Stan probabilistic language (Carpenter et al., 2017) to approximate the posterior distribution of the model using Hamiltonian Monte Carlo. We implemented our extended model as \texttt{response\ \textasciitilde{}\ \ age\ +\ frequency\ +\ dominance*bilingualism*cognate}. We fit this model running four MCMC chains with 2,000 iterations and .80 delta parameter. We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal relative to that of the likelihood, due to the amount of observations introduced in the model. We then fit increasingly simpler models dropping predictors one by one in the following order: (1) \texttt{dominance:bilingualism:cognate}, (2) \texttt{bilingualism:cognate}, (3) \texttt{dominance:cognate:}, (4) \texttt{cognate}, (5) \texttt{dominance:bilingualism}, (6) \texttt{bilingualism}, (7) \texttt{dominance}, (8) \texttt{frequency}, and (9) \texttt{age}. We compared all models using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017) and widely applicable information criterion (WAIC) (Watanabe \& Opper, 2010). We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95\% credible intervals and testing whether this interval excluded 0. We performed follow-up tests on interactions whose 95\% credible interval excluded 0 by comparing the 95\% credible interval of their estimated marginal means using the \texttt{emmeans} package (Lenth, 2020). Data processing and visualisation was done in R (R Core Team, 2020) using the \texttt{tidyverse} family of packages (Wickham et al., 2019), and posterior samples were extracted using the \texttt{tidybayes} R package (Kay, 2020).

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-agresti_analysis_2010}{}%
Agresti, A. (2010). \emph{Analysis of ordinal categorical data} (Vol. 656). John Wiley \& Sons.

\leavevmode\hypertarget{ref-arslan_formr_2020}{}%
Arslan, R. C., Walther, M. P., \& Tata, C. S. (2020). Formr: A study framework allowing for automated feedback generation and complex longitudinal experience-sampling studies using R. \emph{Behavior Research Methods}, \emph{52}(1), 376--387. \url{https://doi.org/10.3758/s13428-019-01236-y}

\leavevmode\hypertarget{ref-ben-zeev_influence_1977}{}%
Ben-Zeev, S. (1977). The Influence of Bilingualism on Cognitive Strategy and Cognitive Development. \emph{Child Development}, \emph{48}(3), 1009--1018. \url{https://doi.org/10.2307/1128353}

\leavevmode\hypertarget{ref-bialystok_receptive_2010}{}%
Bialystok, E., Luk, G., Peets, K. F., \& Yang, S. (2010). Receptive vocabulary differences in monolingual and bilingual children. \emph{Bilingualism: Language and Cognition}, \emph{13}(4), 525--531. \url{https://doi.org/10.1017/S1366728909990423}

\leavevmode\hypertarget{ref-blom_cross-language_2019}{}%
Blom, E., Boerma, T., Bosma, E., Cornips, L., Heuij, K. van den, \& Timmermeister, M. (2019). Cross-language distance influences receptive vocabulary outcomes of bilingual children. \emph{First Language}, 0142723719892794. \url{https://doi.org/10.1177/0142723719892794}

\leavevmode\hypertarget{ref-boada_subtlex-cat_2019}{}%
Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P. (2019). SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for Catalan. \emph{Behavior Research Methods}. \url{https://doi.org/10.3758/s13428-019-01233-1}

\leavevmode\hypertarget{ref-bobb_co-activation_2020}{}%
Bobb, S. C., Von Holzen, K., Mayor, J., Mani, N., \& Carreiras, M. (2020). Co-activation of the L2 during L1 auditory processing: An ERP cross-modal priming study. \emph{Brain and Language}, \emph{203}, 104739. \url{https://doi.org/10.1016/j.bandl.2019.104739}

\leavevmode\hypertarget{ref-bosch_first_2014}{}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in bilingual toddlers' expressive vocabulary: Does form similarity matter? \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\hypertarget{ref-bosma2020cognate}{}%
Bosma, E., \& Nota, N. (2020). Cognate facilitation in frisian--dutch bilingual children's sentence reading: An eye-tracking study. \emph{Journal of Experimental Child Psychology}, \emph{189}, 104699.

\leavevmode\hypertarget{ref-braginsky_consistency_2019}{}%
Braginsky, M., Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and Variability in Children's Word Learning Across Languages. \emph{Open Mind}, \emph{3}, 52--67. \url{https://doi.org/10.1162/opmi_a_00026}

\leavevmode\hypertarget{ref-burkner_brms_2017}{}%
Bürkner, P.-C. (2017). Brms: An R Package for Bayesian Multilevel Models Using Stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}

\leavevmode\hypertarget{ref-burkner_ordinal_2019}{}%
Bürkner, P.-C., \& Vuorre, M. (2019). Ordinal regression models in psychology: A tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 77--101.

\leavevmode\hypertarget{ref-carpenter_stan_2017}{}%
Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A. (2017). Stan: A Probabilistic Programming Language. \emph{Journal of Statistical Software}, \emph{76}(1), 1--32. \url{https://doi.org/10.18637/jss.v076.i01}

\leavevmode\hypertarget{ref-core_total_2013}{}%
Core, C., Hoff, E., Rumiche, R., \& Señor, M. (2013). Total and Conceptual Vocabulary in Spanish--English Bilinguals From 22 to 30 Months: Implications for Assessment. \emph{Journal of Speech, Language, and Hearing Research}, \emph{56}(5), 1637--1649. \url{https://doi.org/10.1044/1092-4388(2013/11-0044)}

\leavevmode\hypertarget{ref-costa2005facilitatory}{}%
Costa, A., Santesteban, M., \& Caño, A. (2005). On the facilitatory effects of cognate words in bilingual speech production. \emph{Brain and Language}, \emph{94}(1), 94--103.

\leavevmode\hypertarget{ref-cuetos_subtlex-esp_2011}{}%
Cuetos, F., Glez-Nosti, M., Barbón, A., \& Brysbaert, M. (2011). SUBTLEX-ESP: Spanish word frequencies based on film subtitles. \emph{Psicológica}, \emph{32}(2), 133--143.

\leavevmode\hypertarget{ref-dell1991mediated}{}%
Dell, G. S., \& O'Seaghdha, P. G. (1991). Mediated and convergent lexical priming in language production: A comment on levelt et al (1991).

\leavevmode\hypertarget{ref-dijkstra_bilingual_2005}{}%
Dijkstra, T. (2005). Bilingual Visual Word Recognition and Lexical Access. In \emph{Handbook of bilingualism: Psycholinguistic approaches} (pp. 179--201). New York, NY, US: Oxford University Press.

\leavevmode\hypertarget{ref-doyle1977some}{}%
Doyle, A. B., \& others. (1977). Some issues in the assessment of linguistic consequences of early bilingualism. \emph{Working Papers on Bilingualism}, (14).

\leavevmode\hypertarget{ref-fernandez1992bilingual}{}%
Fernandez, M. C., Pearson, B. Z., Umbel, V. M., Oiler, D., \& Molinet-Molina, M. (1992). Bilingual receptive vocabulary in hispanic preschool children. \emph{Hispanic Journal of Behavioral Sciences}, \emph{14}(2), 268--276.

\leavevmode\hypertarget{ref-floccia_i_2018}{}%
Floccia, C., Sambrook, T. D., Luche, C. D., Kwok, R., Goslin, J., White, L., \ldots{} Plunkett, K. (2018). I: Introduction. \emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29. \url{https://doi.org/10.1111/mono.12348}

\leavevmode\hypertarget{ref-hamilton_infant_2000}{}%
Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary development assessed with a british communicative development inventory. \emph{Journal of Child Language}, \emph{27}(3), 689--705.

\leavevmode\hypertarget{ref-heuven_subtlex-uk_2014}{}%
Heuven, W. J. B. van, Mandera, P., Keuleers, E., \& Brysbaert, M. (2014). Subtlex-UK: A New and Improved Word Frequency Database for British English: \emph{Quarterly Journal of Experimental Psychology}. Retrieved from \url{https://journals.sagepub.com/doi/10.1080/17470218.2013.850521}

\leavevmode\hypertarget{ref-hoff_dual_2012}{}%
Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M. (2012). Dual language exposure and early bilingual development*. \emph{Journal of Child Language}, \emph{39}(1), 1--27. \url{https://doi.org/10.1017/S0305000910000759}

\leavevmode\hypertarget{ref-hoff_expressive_2014}{}%
Hoff, E., Rumiche, R., Burridge, A., Ribot, K. M., \& Welsh, S. N. (2014). Expressive vocabulary development in children from bilingual and monolingual homes: A longitudinal study from two to four years. \emph{Early Childhood Research Quarterly}, \emph{29}(4), 433--444. \url{https://doi.org/10.1016/j.ecresq.2014.04.012}

\leavevmode\hypertarget{ref-hoshino_cognate_2008}{}%
Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming: Does cross-language activation survive a change of script? \emph{Cognition}, \emph{106}(1), 501--511. \url{https://doi.org/10.1016/j.cognition.2007.02.001}

\leavevmode\hypertarget{ref-houwer_bilingualmonolingual_2014}{}%
Houwer, A. D., Bornstein, M. H., \& Putnick, D. L. (2014). A bilingual--monolingual comparison of young children's vocabulary size: Evidence from comprehension and production. \emph{Applied Psycholinguistics}, \emph{35}(6), 1189--1211. \url{https://doi.org/10.1017/S0142716412000744}

\leavevmode\hypertarget{ref-jardak2019labels}{}%
Jardak, A., \& Byers-Heinlein, K. (2019). Labels or concepts? The development of semantic networks in bilingual two-year-olds. \emph{Child Development}, \emph{90}(2), e212--e229.

\leavevmode\hypertarget{ref-jones_children_2019}{}%
Jones, S. D., \& Brandt, S. (2019). Do children really acquire dense neighbourhoods? \emph{Journal of Child Language}, \emph{46}(6), 1260--1273. \url{https://doi.org/10.1017/S0305000919000473}

\leavevmode\hypertarget{ref-kay_tidybayes_2020}{}%
Kay, M. (2020). \emph{tidybayes: Tidy data and geoms for Bayesian models}. \url{https://doi.org/10.5281/zenodo.1308151}

\leavevmode\hypertarget{ref-kroll_multilingual_2013}{}%
Kroll, J. F., Gullifer, J. W., \& Rossi, E. (2013). The Multilingual Lexicon: The Cognitive and Neural Basis of Lexical Comprehension and Production in Two or More Languages. \emph{Annual Review of Applied Linguistics}, \emph{33}, 102--127. \url{https://doi.org/10.1017/S0267190513000111}

\leavevmode\hypertarget{ref-lenth_emmeans_2020}{}%
Lenth, R. V. (2020). \emph{Emmeans: Estimated marginal means, aka least-squares means}. Retrieved from \url{https://CRAN.R-project.org/package=emmeans}

\leavevmode\hypertarget{ref-pearson_patterns_1994}{}%
Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of Interaction in the Lexical Growth in Two Languages of Bilingual Infants and Toddlers. \emph{Language Learning}, \emph{44}(4), 617--653. \url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}

\leavevmode\hypertarget{ref-pearson_lexical_1993}{}%
Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical Development in Bilingual Infants and Toddlers: Comparison to Monolingual Norms. \emph{Language Learning}, \emph{43}(1), 93--120. \url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}

\leavevmode\hypertarget{ref-poulin2013lexical}{}%
Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J. (2013). Lexical access and vocabulary development in very young bilinguals. \emph{International Journal of Bilingualism}, \emph{17}(1), 57--70.

\leavevmode\hypertarget{ref-rcoreteam_r_2020}{}%
R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-rosenblum1983word}{}%
Rosenblum, T., \& Pinker, S. A. (1983). Word magic revisited: Monolingual and bilingual children's understanding of the word-object relationship. \emph{Child Development}, 773--780. \url{https://doi.org/10.2307/1130064}

\leavevmode\hypertarget{ref-schelletter2002effect}{}%
Schelletter, C. (2002). The effect of form similarity on bilingual children's lexical development. \emph{Bilingualism: Language and Cognition}, \emph{5}(2), 93--107.

\leavevmode\hypertarget{ref-singh2014one}{}%
Singh, L. (2014). One world, two languages: Cross-language semantic priming in bilingual toddlers. \emph{Child Development}, \emph{85}(2), 755--766.

\leavevmode\hypertarget{ref-spivey1999cross}{}%
Spivey, M. J., \& Marian, V. (1999). Cross talk between native and second languages: Partial activation of an irrelevant lexicon. \emph{Psychological Science}, \emph{10}(3), 281--284.

\leavevmode\hypertarget{ref-thierry_brain_2007}{}%
Thierry, G., \& Wu, Y. J. (2007). Brain potentials reveal unconscious translation during foreign-language comprehension. \emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535. \url{https://doi.org/10.1073/pnas.0609927104}

\leavevmode\hypertarget{ref-vehtari_practical_2017}{}%
Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC. \emph{Statistics and Computing}, \emph{27}(5), 1413--1432. \url{https://doi.org/10.1007/s11222-016-9696-4}

\leavevmode\hypertarget{ref-von_holzen_impact_2019}{}%
Von Holzen, K., Fennell, C. T., \& Mani, N. (2019). The impact of cross-language phonological overlap on bilingual and monolingual toddlers' word recognition. \emph{Bilingualism: Language and Cognition}, \emph{22}(3), 476--499. \url{https://doi.org/10.1017/S1366728918000597}

\leavevmode\hypertarget{ref-von_holzen_language_2012}{}%
Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586. \url{https://doi.org/10.1016/j.jecp.2012.08.001}

\leavevmode\hypertarget{ref-watanabe_asymptotic_2010}{}%
Watanabe, S., \& Opper, M. (2010). Asymptotic equivalence of bayes cross validation and widely applicable information criterion in singular learning theory. \emph{Journal of Machine Learning Research}, \emph{11}(12).

\leavevmode\hypertarget{ref-wickham_welcome_2019}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R., \ldots{} Yutani, H. (2019). Welcome to the \textless span class="nocase"\textgreater tidyverse\textless/span\textgreater. \emph{Journal of Open Source Software}, \emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\leavevmode\hypertarget{ref-yudes_cognate_2010}{}%
Yudes, C., Macizo, P., \& Bajo, T. (2010). Cognate effects in bilingual language comprehension tasks. \emph{NeuroReport}, \emph{21}(7), 507--512. \url{https://doi.org/10.1097/WNR.0b013e328338b9e1}
\end{cslreferences}

\endgroup

\end{document}

### 2021-01-15 R/analyses.R
Separate vocab type

@@ -12,12 +12,12 @@ library(here)
 
 # load helper functions
 source(here("R", "utils.R"))
+options(mc.cores = 4)
 
 #### import data ---------------------------------------------------------------
 responses <- read_csv(here("Data", "responses.csv")) %>% 
-  mutate_at(vars(age, frequency, doe), function(x) scale(x, scale = FALSE)[,1]) %>% 
-  mutate_at(vars(item_dominance, cognate, response), as.factor) %>% 
-  mutate(response = factor(response, ordered = TRUE))
+  mutate_at(vars(age, frequency, bilingualism), function(x) scale(x)[,1]) %>% 
+  mutate_at(vars(dominance, cognate), as.factor)
 
 # responses_subset <- filter(responses, te %in% sample(responses$te, 25)) %>% 
 #   mutate_at(vars(age, frequency, doe), function(x) scale(x, scale = FALSE)[,1]) %>% 
@@ -26,21 +26,11 @@ responses <- read_csv(here("Data", "responses.csv")) %>%
 
 # summarise responses (for visualisation purposes)
 proportion <- responses %>% 
-  mutate(
-    understands = response %in% 2,
-    says = response %in% 3, 
-    frequency = cut_quantiles(frequency),
-    age = round(age)
-  ) %>%
-  select(-doe) %>% 
-  pivot_longer(c(understands, says), names_to = ".category", values_to = ".value") %>% 
-  mutate(
-    .category = str_to_sentence(.category),
-    .value = as.numeric(.value)
-  ) %>% 
-  group_by(age, cognate, item_dominance, .category) %>%
+  pivot_longer(c(understands, produces), names_to = "type", values_to = ".value") %>% 
+  mutate(type = str_to_sentence(type)) %>% 
+  group_by(age, frequency, bilingualism, cognate, dominance, type) %>%
   summarise(
-    .value = sum(.value),
+    .value = log((.value+0.5)/(n()-.value+0.5)),
     n = n(),
     .groups = "drop"
   ) %>% 
@@ -50,124 +40,162 @@ proportion <- responses %>%
 
 
 # set sum contrasts (Schad et al., 2018, https://arxiv.org/abs/1807.10451)
-contrasts(responses$item_dominance) <- contr.sum(2)/2
[…]

### 2021-01-15 Rmd/trajectories_manuscript.Rmd
Separate vocab type

@@
 author: 
-  - name          : "Gonzalo García-Castro"
+  - name          : "Gonzalo Garcia-Castro"
     affiliation   : "1"
@@
     affiliation   : "1"
-  - name          : "Núria Sebastian-Galles"
+  - name          : "Nuria Sebastian-Galles"
     affiliation   : "1"
@@
 library(lubridate)
+library(brms)
+library(tidybayes)
 library(janitor)
@@
 
+source(here("R", "utils.R"))
 ```
@@
 The structure of the lexicon also impacts the order of acquisition of new words.
-Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills2009].
+Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills_longitudinal_2009].
 @borosky2016 and @borovsky2018 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
@@
 When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev_influence_1977; @doyle1977some; @rosenblum1983word; @fernandez1992bilingual; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
+When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev_influence_1977; @doyle_some_1977; @rosenblum_word_1983; @fernandez_bilingual_1992; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
 
-@floccia_i_2018, gathered parental vocabulary estimates [@hamilton2000infant] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
+@floccia_i_2018, gathered parental vocabulary estimates [@hamilton_infant_2000] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
 They computed the average phonological similarity between translation equivalents (TEs from now on; e.g.,  *table* in English and *tafel* in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
@@
 The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., *cat*, in English) its corresponding translation in the other language is activated too [e.g., *gato*, in Spanish.
-There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell1991mediated; @spivey1999cross; @costa2000lexical; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh2014one; @von_holzen_impact_2019; @bobb_co-activation_2020; see @costa2005facilitatory for review], and children [e.g.
-@von_holzen_language_2012; @poulin2013lexical; @jardak2019labels; @bosma2020cognate].
+There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell_mediated_1991; @spivey_cross_1999; @costa_lexical_2000; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh_one_2014; @von_holzen_impact_2019; @bobb_co-activation_2020; see @costa_facilitatory_2005 for review], and children [e.g.
+@von_holzen_language_2012; @poulin_lexical_2013; @jardak_labels_2019; @bosma_cognate_2020].
 A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
@@
 The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse.
-@schelletter2002effect's reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
+@schelletter_effect_2002 reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
 @bosch_first_2014, showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months.
@@
 
-We then fit increasingly simpler models dropping predictors one by one in the following order: (1) `dominance:bilingualism:cognate`, (2) `bilingualism:cognate`, (3) `dominance:cognate:`, (4) `cognate`, (5) `dominance:bilingualism`, (6) `bilingualism`, (7) `dominance`, (8) `frequency`, and (9) `age`.
-We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017] and widely applicable information criterion (WAIC) [@watanabe_asymptotic_2010].
+We fit increasingly simpler models dropping predictors one by one in the following order: (1) `dominance:bilingualism:cognate`, (2) `bilingualism:cognate`, (3) `dominance:cognate:`, (4) `cognate`, (5) `dominance:bilingualism`, (6) `bilingualism`, (7) `dominance`, (8) `frequency`, and (9) `age`.
+We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017]^[Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson_leave_2019]].
 We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95% credible intervals and testing whether this interval excluded 0.
@@
 # Results
+
+```{r results}
+fit_9 <- readRDS(here("Results", "fit_9.rds"))
+loos <- readRDS(here("Results", "loos.rds"))
+
+```
+
+Table 2 summarises shows the outcome of the LOO-CV model comparison. The extended model was most favoured by the data.
+
+```{r loos, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+as.data.frame(loos$diffs) %>% 
+  mutate(
+    model = md(c(
+      "~ 1 (intercept only)",
+      "... + age",
+      "... + frequency",
+      "... + dominance",
+      "... + bilingualism",
+      "... + dominance:bilingualism",
+      "... + cognate",
+      "... + dominance:cognate",
+      "... + bilingualism:cognate",
+      "... + dominance:bilingualism:cognate"
+    ))) %>% 
+  select(model, elpd_loo, se_elpd_loo, p_loo, elpd_diff, se_diff) %>% 
+  gt() %>% 
+  tab_spanner(
+    label = md("**Comparison**"),
+    columns = c("elpd_diff", "se_diff")
+  ) %>% 
+  cols_label(
+    model = md("**Model**"),
+    elpd_loo = md("**ELPD**"),
+    se_elpd_loo = md("**SE**"),
+    p_loo = md("**p**"),
+    elpd_diff = md("*LOO<sub>diff</sup>*"),
+    se_diff = md("*SE<sub>diff</sup>*")
+  ) %>% 
+  fmt_number(2:5) %>% 
+  tab_footnote(
+    footnote = md("*ELPD*: theoretical expected log pointwise predictive density for a new dataset, as estimated using Bayesian LOO-CV"),
+    locations = cells_column_labels(columns = vars("elpd_loo"))
+  ) %>% 
+  tab_footnote(
+    footnote = md("*SE*: standard error of the ELPD"),
+    locations = cells_column_labels(columns = vars("se_elpd_loo"))
+  ) %>% 
+  tab_footnote(
+    footnote = md("*p*: effective number of parameters"),
+    locations = cells_column_labels(columns = vars("p_loo"))
+  ) %>% 
+  tab_options(footnotes.marks = "letters")
+
+```
+
+Table 3 shows the mean and 95% CrI of each regression coefficient of the best fitting model.
+
+```{r coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+cbind(as.data.frame(exp(fixef(fit_9))),
+      as.data.frame(inv_logit_scaled(fixef(fit_9)))) %>% 
+  rownames_to_column("coefficient") %>% 
+  mutate(coefficient = c(
+    "Intercept 1 (Understands)",
+    "Intercept 2 (Understands & Says)",
+    "Age",
+    "Frequency",
+    "Dominance",
+    "Bilingualism",
+    "Cognate",
+    "Dominance * Bilingualism",
+    "Dominance * Cognate",
+    "Bilingualism * Cognate",
+    "Dominance * Bilingualism * Cognate"
+  )) %>% 
+  clean_names() %>% 
+  gt() %>% 
+  tab_spanner(label = md("**Odds Ratio**"), columns = 2:5) %>% 
+  tab_spanner(label = md("**Probability**"), columns = 6:8) %>% 
+  fmt_number(2:9) %>% 
+  cols_merge_range(col_begin = "q2_5", col_end = "q97_5") %>% 
+  cols_merge_range(col_begin = "q2_5_1", col_end = "q97_5_1") %>% 
+  cols_label(
+    coefficient = md("**Coefficient**"),
+    estimate = md("*Mean*"),
+    est_error = md("*SE*"),
+    q2_5 = md("*95% CrI*"),
+    estimate_1 = md("*Mean*"),
+    est_error_1 = md("*SE*"),
+    q2_5_1 = md("*95% CrI*"),
+  ) %>% 
+  tab_footnote(
+    footnote = "SE: Standard Error of the mean",
+    locations = cells_column_labels(columns = vars("est_error"))
+  ) %>% 
+  tab_footnote(
+    footnote = "95% CrI: 95% credible interval",
+    locations = cells_column_labels(columns = vars("q2_5"))
+  )
+
+```
+
+```{r marginalmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6.5, fig.height=6.5, out.width="20cm", fig.cap="Mean posterior probability and 95\\% CrI of responses, conditional to Dominance, Bilingualism, and Cognate."}
+
+# summarise data
+proportion <- responses %>% 
+  mutate(
+    understands = response %in% 2,
+    says = response %in% 3, 
+    frequency = cut_quantiles(frequency),
+    age = round(age)
+  ) %>%
+  select(-doe) %>% 
+  pivot_longer(c(understands, says), names_to = ".category", values_to = ".value") %>% 
+  mutate(
+    .category = str_to_sentence(.category),
+    .value = as.numeric(.value)
+  ) %>% 
+  group_by(age, cognate, item_dominance, .category) %>%
+  summarise(
+    .value = sum(.value),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  rowwise() %>% 
+  mutate(.value = prop_adj(.value, n)) %>% 
+  ungroup()
+
+# get 20 posterior draws
+# values to get posterior draws for (set frequency at mean = 0)
+nd <- expand_grid(
+  age = seq(min(responses$age), max(responses$age), 1),
+  item_dominance = c("L1", "L2"),
+  doe = c(-1, 0, 1),
+  frequency = 0,
+  cognate = c("Cognate", "Non-Cognate")
+)
+
+post_preds <- add_fitted_draws(nd, fit_9, n = 20) %>% 
+  mutate(doe = doe %>%
+           as.character() %>%  
+           str_replace_all(c("0" = "Mean","1" = "1 SD")) %>% 
+           paste0("DOE = ", .) %>% 
+           factor(levels = c("DOE = -1 SD", "DOE = Mean","DOE = 1 SD"), ordered = TRUE)) %>% 
+  mutate(
+    .category = case_when(
+      .category==1 ~ "None",
+      .category==2 ~ "Understands",
+      .category==3 ~ "Says"
+    ),
+    .category = factor(.category, levels = c("None", "Understands", "Says"), ordered = TRUE)
+  )
+
+# visualise posterior predictions
+post_preds %>% 
+  filter(.category!="None") %>% 
+  ggplot(aes(
+    age, .value,
+    colour = interaction(item_dominance, cognate, sep = " - "),
+    fill = interaction(item_dominance, cognate, sep = " - ")
+  )) +
+  facet_grid(doe~.category) +
+  #geom_line(aes(group = interaction(item_dominance, cognate, doe, .draw)), size = 0.4) +
+  stat_lineribbon(.width = 0.95, colour = NA, alpha = 0.5) +
+  stat_summary(fun = "median", geom = "line", size = 0.75) +
+  geom_point(data = proportion, alpha = 0.5) +
+  labs(x = "Age (months)", y = "P(Y|X)") +
+  guides(colour = guide_legend(ncol = 2)) +
+  scale_color_brewer(palette = "Dark2") +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_custom() +
+  theme(
+    legend.position = "top",
+    legend.title = element_blank(),
+    legend.direction = "horizontal"
+  )
+
+```
+
+The age at which the average item changed
 

### 2021-01-15 Rmd/trajectories_manuscript.tex
Separate vocab type

@@
   pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
-  pdfauthor={Gonzalo García-Castro1, Daniela Avila-Varela1, \& Núria Sebastian-Galles1},
+  pdfauthor={Gonzalo Garcia-Castro1, Daniela Avila-Varela1, \& Nuria Sebastian-Galles1},
   pdflang={en-EN},
@@
 \title{The role of cross-linguistic lexical similarity on bilingual word acquisition}
-\author{Gonzalo García-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Núria Sebastian-Galles\textsuperscript{1}}
+\author{Gonzalo Garcia-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Nuria Sebastian-Galles\textsuperscript{1}}
 \date{}
@@
 
-Correspondence concerning this article should be addressed to Gonzalo García-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain.
+Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain.
 E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}
@@
 The structure of the lexicon also impacts the order of acquisition of new words.
-Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next ({\textbf{???}}).
+Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next (Hills, Maouene, Maouene, Sheya, \& Smith, 2009).
 ({\textbf{???}}) and ({\textbf{???}}) showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
@@
 When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When both languages are taken into account, bilingual children seem to know, at least, as many words as monolinguals do (Ben-Zeev, 1977; Bialystok, Luk, Peets, \& Yang, 2010; Blom et al., 2019; Core, Hoff, Rumiche, \& Señor, 2013; Doyle \& others, 1977; Fernandez, Pearson, Umbel, Oiler, \& Molinet-Molina, 1992; Hoff et al., 2012; Hoff, Rumiche, Burridge, Ribot, \& Welsh, 2014; Pearson \& Fernández, 1994; Rosenblum \& Pinker, 1983; but see Pearson, Fernández, \& Oller, 1993; Houwer, Bornstein, \& Putnick, 2014).
+When both languages are taken into account, bilingual children seem to know, at least, as many words as monolinguals do ({\textbf{???}}; {\textbf{???}}; Ben-Zeev, 1977; Bialystok, Luk, Peets, \& Yang, 2010; Blom et al., 2019; Core, Hoff, Rumiche, \& Señor, 2013; Fernandez, Pearson, Umbel, Oiler, \& Molinet-Molina, 1992; Hoff et al., 2012; Hoff, Rumiche, Burridge, Ribot, \& Welsh, 2014; Pearson \& Fernández, 1994; but see Pearson, Fernández, \& Oller, 1993; Houwer, Bornstein, \& Putnick, 2014).
 
 Floccia et al.
-(2018), gathered parental vocabulary estimates ({\textbf{???}}) from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
+(2018), gathered parental vocabulary estimates (Hamilton, Plunkett, \& Schafer, 2000) from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
 They computed the average phonological similarity between translation equivalents (TEs from now on; e.g., \emph{table} in English and \emph{tafel} in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
@@
 The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., \emph{cat}, in English) its corresponding translation in the other language is activated too {[}e.g., \emph{gato}, in Spanish.
-There is a vast body of evidence supporting language non-selective lexical access in both adults (e.g., Dell \& O'Seaghdha, 1991; {\textbf{???}}; Bobb, Von Holzen, Mayor, Mani, \& Carreiras, 2020; Dijkstra, 2005; Hoshino \& Kroll, 2008; Kroll, Gullifer, \& Rossi, 2013; Singh, 2014; Spivey \& Marian, 1999; Thierry \& Wu, 2007; Von Holzen, Fennell, \& Mani, 2019; Yudes, Macizo, \& Bajo, 2010; see Costa, Santesteban, \& Caño, 2005 for review), and children (e.g.~Von Holzen \& Mani, 2012; Bosma \& Nota, 2020; Jardak \& Byers-Heinlein, 2019; Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013).
+There is a vast body of evidence supporting language non-selective lexical access in both adults (e.g., Dell \& O'Seaghdha, 1991; {\textbf{???}}; Bobb, Von Holzen, Mayor, Mani, \& Carreiras, 2020; Costa \& Caramazza, 2000; Dijkstra, 2005; Hoshino \& Kroll, 2008; Kroll, Gullifer, \& Rossi, 2013; Singh, 2014; Thierry \& Wu, 2007; Von Holzen, Fennell, \& Mani, 2019; Yudes, Macizo, \& Bajo, 2010; see Costa, Santesteban, \& Caño, 2005 for review), and children (e.g.~Von Holzen \& Mani, 2012; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
 A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
@@
 The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse.
-Schelletter (2002)`s reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
+Schelletter (2002) reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
 Bosch and Ramon-Casas (2014), showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months.
@@
 We implemented an on-line questionnaire using \texttt{formr} (Arslan, Walther, \& Tata, 2020), divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (Hamilton, Plunkett, \& Schafer, 2000) and consisted in two lists of words, one in Catalan and one in Spanish).
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (Hamilton et al., 2000) and consisted in two lists of words, one in Catalan and one in Spanish).
 Items in one language were translation equivalents of the items in the other (e.g., whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the word \emph{perro} was included in the Spanish inventory), roughly following a one-to-one mapping.
@@
 Translation equivalents received an average of 386.36 (\emph{SD} = 228.76, , Max = 698), both languages summed together.
-We modelled the probability of responses (\(y_i\)) to belong to any of the three possible categories (\(k\)): 1 = \emph{No} or say, 2 = \emph{Understands}, and 3 = \emph{Understands and Says}) using an ordinal regression model with a cumulative logit link function (Agresti, 2010; Bürkner \& Vuorre, 2019).
+We modelled the probability of responses (\(y_i\)) belonging to any of the three possible categories \(k\)--1 = \emph{No} or say, 2 = \emph{Understands}, and 3 = \emph{Understands and Says}-- using an ordinal regression model with a cumulative logit link function (Agresti, 2010; Bürkner \& Vuorre, 2019).
 This model estimates the probability of response \(y_i\) to belong to category \(k\) or lower.
@@
 
-We implemented the model using the R package \texttt{brms} (Bürkner, 2017).
-This package uses the Stan probabilistic language (Carpenter et al., 2017) to approximate the posterior distribution of the model using Hamiltonian Monte Carlo.
-We implemented our extended model as \texttt{response\ \textasciitilde{}\ \ age\ +\ frequency\ +\ dominance*bilingualism*cognate}.
-We fit this model running four MCMC chains with 2,000 iterations and .80 delta parameter.
-We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal relative to that of the likelihood, due to the amount of observations introduced in the model.
-We then fit increasingly simpler models dropping predictors one by one in the following order: (1) \texttt{dominance:bilingualism:cognate}, (2) \texttt{bilingualism:cognate}, (3) \texttt{dominance:cognate:}, (4) \texttt{cognate}, (5) \texttt{dominance:bilingualism}, (6) \texttt{bilingualism}, (7) \texttt{dominance}, (8) \texttt{frequency}, and (9) \texttt{age}.
-We compared all models using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017) and widely applicable information criterion (WAIC) (Watanabe \& Opper, 2010).
+We implemented the model using the R package \texttt{brms} (Bürkner, 2017) as \texttt{response\ \textasciitilde{}\ \ age\ +\ frequency\ +\ dominance*bilingualism*cognate}, running four MCMC sampling chains with 4,000 iterations (including 2,000 warm-up iterations per chain).
+We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal due to the large amount of observations introduced in the model.
+We report the 95\% credible intervals (CrI) of the estimated posterior distribution of each parameter, which denotes a range of values we are 95\% certain contain the true value, given our prior and observed data.
+We report analyses on the probability scale, having transformed the log-odds returned by the model using an inverse logit transformation.
+
+We fit increasingly simpler models dropping predictors one by one in the following order: (1) \texttt{dominance:bilingualism:cognate}, (2) \texttt{bilingualism:cognate}, (3) \texttt{dominance:cognate:}, (4) \texttt{cognate}, (5) \texttt{dominance:bilingualism}, (6) \texttt{bilingualism}, (7) \texttt{dominance}, (8) \texttt{frequency}, and (9) \texttt{age}.
+We compared all models using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017)\footnote{Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1000 samples (Magnusson, Andersen, Jonasson, \& Vehtari, 2019)}.
 We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95\% credible intervals and testing whether this interval excluded 0.
@@
 Data processing and visualisation was done in R (R Core Team, 2020) using the \texttt{tidyverse} family of packages (Wickham et al., 2019), and posterior samples were extracted using the \texttt{tidybayes} R package (Kay, 2020).
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+Table 2 summarises shows the outcome of the LOO-CV model comparison. The extended model was most favoured by the data.
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{crrrrr}
+\toprule
+& & & & \multicolumn{2}{c}{\textbf{Comparison}} \\ 
+ \cmidrule(lr){5-6}
+\textbf{Model} & \textbf{ELPD}\textsuperscript{a} & \textbf{SE}\textsuperscript{b} & \textbf{p}\textsuperscript{c} & \emph{LOOdiff} & \emph{SEdiff} \\ 
+\midrule
+\textasciitilde  1 (intercept only) & $-157,272.37$ & $250.56$ & $12.00$ & $0.00$ & 0.0000 \\ 
+... + age & $-157,308.76$ & $250.47$ & $10.37$ & $36.39$ & 354.2836 \\ 
+... + frequency & $-157,318.93$ & $250.45$ & $9.93$ & $46.56$ & 354.2703 \\ 
+... + dominance & $-157,463.31$ & $250.25$ & $8.08$ & $190.94$ & 354.1293 \\ 
+... + bilingualism & $-157,472.54$ & $250.29$ & $7.67$ & $200.17$ & 354.1564 \\ 
+... + dominance:bilingualism & $-158,724.35$ & $246.22$ & $6.33$ & $1,451.98$ & 351.2863 \\ 
+... + cognate & $-159,063.42$ & $245.10$ & $5.09$ & $1,791.06$ & 350.5085 \\ 
+... + dominance:cognate & $-161,704.16$ & $237.62$ & $3.95$ & $4,431.79$ & 345.3184 \\ 
+... + bilingualism:cognate & $-165,570.87$ & $228.08$ & $3.13$ & $8,298.50$ & 338.8223 \\ 
+... + dominance:bilingualism:cognate & $-184,511.91$ & $172.86$ & $2.04$ & $27,239.54$ & 304.4033 \\ 
+\bottomrule
+\end{longtable}
+\vspace{-5mm}
+\begin{minipage}{\linewidth}
+\textsuperscript{a}ELPD: theoretical expected log pointwise predictive density for a new dataset, as estimated using Bayesian LOO-CV \\ 
+\textsuperscript{b}SE: standard error of the ELPD \\ 
+\textsuperscript{c}p: effective number of parameters \\ 
+\end{minipage}
+
+Table 3 shows the mean and 95\% CrI of each regression coefficient of the best fitting model.
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lrrrrrr}
+\toprule
+& \multicolumn{3}{c}{\textbf{Odds Ratio}} & \multicolumn{3}{c}{\textbf{Probability}} \\ 
+ \cmidrule(lr){2-4}\cmidrule(lr){5-7}
+\textbf{Coefficient} & \emph{Mean} & \emph{SE}\textsuperscript{1} & \emph{95\% CrI}\textsuperscript{2} & \emph{Mean} & \emph{SE} & \emph{95\% CrI} \\ 
+\midrule
+Intercept 1 (Understands) & $1.12$ & $1.01$ & $1.11$–$1.14$ & $0.53$ & $0.50$ & $0.53$–$0.53$ \\ 
+Intercept 2 (Understands \& Says) & $5.31$ & $1.01$ & $5.24$–$5.38$ & $0.84$ & $0.50$ & $0.84$–$0.84$ \\ 
+Age & $1.16$ & $1.00$ & $1.16$–$1.17$ & $0.54$ & $0.50$ & $0.54$–$0.54$ \\ 
+Frequency & $1.86$ & $1.01$ & $1.83$–$1.88$ & $0.65$ & $0.50$ & $0.65$–$0.65$ \\ 
+Dominance & $2.27$ & $1.01$ & $2.22$–$2.32$ & $0.69$ & $0.50$ & $0.69$–$0.70$ \\ 
+Bilingualism & $1.10$ & $1.00$ & $1.09$–$1.11$ & $0.52$ & $0.50$ & $0.52$–$0.53$ \\ 
+Cognate & $1.08$ & $1.01$ & $1.06$–$1.10$ & $0.52$ & $0.50$ & $0.51$–$0.52$ \\ 
+Dominance * Bilingualism & $0.73$ & $1.01$ & $0.73$–$0.74$ & $0.42$ & $0.50$ & $0.42$–$0.43$ \\ 
+Dominance * Cognate & $0.68$ & $1.02$ & $0.65$–$0.71$ & $0.40$ & $0.51$ & $0.39$–$0.42$ \\ 
+Bilingualism * Cognate & $0.97$ & $1.01$ & $0.96$–$0.98$ & $0.49$ & $0.50$ & $0.49$–$0.50$ \\ 
+Dominance * Bilingualism * Cognate & $1.13$ & $1.01$ & $1.10$–$1.15$ & $0.53$ & $0.50$ & $0.52$–$0.54$ \\ 
+\bottomrule
+\end{longtable}
+\vspace{-5mm}
+\begin{minipage}{\linewidth}
+\textsuperscript{1}SE: Standard Error of the mean \\ 
+\textsuperscript{2}95\% CrI: 95\% credible interval \\ 
+\end{minipage}
+
+\begin{figure}
+
+{\centering \includegraphics[width=20cm]{trajectories_manuscript_files/figure-latex/marginalmeans-1}
+
+}
+
+\caption{Mean posterior probability and 95\% CrI of responses, conditional to Dominance, Bilingualism, and Cognate.}\label{fig:marginalmeans}
+\end{figure}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\leavevmode\hypertarget{ref-bosma2020cognate}{}%
-Bosma, E., \& Nota, N. (2020). Cognate facilitation in frisian--dutch bilingual children's sentence reading: An eye-tracking study. \emph{Journal of Experimental Child Psychology}, \emph{189}, 104699.
-
 \leavevmode\hypertarget{ref-braginsky_consistency_2019}{}%
@@
 
-\leavevmode\hypertarget{ref-carpenter_stan_2017}{}%
-Carpenter, B., Gelman, A., Hoffman, M.
-D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A.
-(2017).
-Stan: A Probabilistic Programming Language.
-\emph{Journal of Statistical Software}, \emph{76}(1), 1--32.
-\url{https://doi.org/10.18637/jss.v076.i01}
-
 \leavevmode\hypertarget{ref-core_total_2013}{}%
@@
 
-\leavevmode\hypertarget{ref-costa2005facilitatory}{}%
+\leavevmode\hypertarget{ref-costa_lexical_2000}{}%
+Costa, A., \& Caramazza, A. (2000). Lexical access in speech production: The bilingual case. \emph{Psicológica}, \emph{21}(2), 403--437.
+
+\leavevmode\hypertarget{ref-costa_facilitatory_2005}{}%
 Costa, A., Santesteban, M., \& Caño, A. (2005). On the facilitatory effects of cognate words in bilingual speech production. \emph{Brain and Language}, \emph{94}(1), 94--103.
@@
 
-\leavevmode\hypertarget{ref-dell1991mediated}{}%
+\leavevmode\hypertarget{ref-dell_mediated_1991}{}%
 Dell, G. S., \& O'Seaghdha, P. G. (1991). Mediated and convergent lexical priming in language production: A comment on levelt et al (1991).
@@
 
-\leavevmode\hypertarget{ref-doyle1977some}{}%
-Doyle, A. B., \& others. (1977). Some issues in the assessment of linguistic consequences of early bilingualism. \emph{Working Papers on Bilingualism}, (14).
-
-\leavevmode\hypertarget{ref-fernandez1992bilingual}{}%
+\leavevmode\hypertarget{ref-fernandez_bilingual_1992}{}%
 Fernandez, M.
@@
 Z., Umbel, V.
-M., Oiler, D., \& Molinet-Molina, M.
+M., Oiler, D.
+K., \& Molinet-Molina, M.
 (1992).
-Bilingual receptive vocabulary in hispanic preschool children.
+Bilingual Receptive Vocabulary in Hispanic Preschool Children.
 \emph{Hispanic Journal of Behavioral Sciences}, \emph{14}(2), 268--276.
+\url{https://doi.org/10.1177/07399863920142006}
 
@@
 Retrieved from \url{https://journals.sagepub.com/doi/10.1080/17470218.2013.850521}
+
+\leavevmode\hypertarget{ref-hills_longitudinal_2009}{}%
+Hills, T.
+T., Maouene, M., Maouene, J., Sheya, A., \& Smith, L.
+(2009).
+Longitudinal Analysis of Early Semantic Networks Preferential Attachment or Preferential Acquisition?
+\emph{Psychological Science}, \emph{20}(6), 729--739.
+\url{https://doi.org/10.1111/j.1467-9280.2009.02365.x}
 
@@
 
-\leavevmode\hypertarget{ref-jardak2019labels}{}%
-Jardak, A., \& Byers-Heinlein, K. (2019). Labels or concepts? The development of semantic networks in bilingual two-year-olds. \emph{Child Development}, \emph{90}(2), e212--e229.
-
 \leavevmode\hypertarget{ref-jones_children_2019}{}%
@@
 Lenth, R. V. (2020). \emph{Emmeans: Estimated marginal means, aka least-squares means}. Retrieved from \url{https://CRAN.R-project.org/package=emmeans}
+
+\leavevmode\hypertarget{ref-magnusson_leave_2019}{}%
+Magnusson, M., Andersen, M., Jonasson, J., \& Vehtari, A.
+(2019).
+Bayesian leave-one-out cross-validation for large data.
+In K.
+Chaudhuri \& R.
+Salakhutdinov (Eds.), \emph{Proceedings of the 36th international conference on machine learning} (Vol.
+97, pp.
+4244--4253).
+Long Beach, California, USA: PMLR.
+Retrieved from \url{http://proceedings.mlr.press/v97/magnusson19a.html}
 
@@
 
-\leavevmode\hypertarget{ref-poulin2013lexical}{}%
-Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J.
-(2013).
-Lexical access and vocabulary development in very young bilinguals.
-\emph{International Journal of Bilingualism}, \emph{17}(1), 57--70.
-
 \leavevmode\hypertarget{ref-rcoreteam_r_2020}{}%
@@
 
-\leavevmode\hypertarget{ref-rosenblum1983word}{}%
-Rosenblum, T., \& Pinker, S.
-A.
-(1983).
-Word magic revisited: Monolingual and bilingual children's understanding of the word-object relationship.
-\emph{Child Development}, 773--780.
-\url{https://doi.org/10.2307/1130064}
+\leavevmode\hypertarget{ref-schelletter_effect_2002}{}%
+Schelletter, C.
+(2002).
+The effect of form similarity on bilingual children's lexical development.
+\emph{Bilingualism: Language and Cognition}, \emph{5}(2), 93--107.
+\url{https://doi.org/10.1017/S1366728902000214}
 
-\leavevmode\hypertarget{ref-schelletter2002effect}{}%
-Schelletter, C. (2002). The effect of form similarity on bilingual children's lexical development. \emph{Bilingualism: Language and Cognition}, \emph{5}(2), 93--107.
-
-\leavevmode\hypertarget{ref-singh2014one}{}%
+\leavevmode\hypertarget{ref-singh_one_2014}{}%
 Singh, L. (2014). One world, two languages: Cross-language semantic priming in bilingual toddlers. \emph{Child Development}, \emph{85}(2), 755--766.
-
-\leavevmode\hypertarget{ref-spivey1999cross}{}%
-Spivey, M. J., \& Marian, V. (1999). Cross talk between native and second languages: Partial activation of an irrelevant lexicon. \emph{Psychological Science}, \emph{10}(3), 281--284.
 
@@
 
-\leavevmode\hypertarget{ref-watanabe_asymptotic_2010}{}%
-Watanabe, S., \& Opper, M.
-(2010).
-Asymptotic equivalence of bayes cross validation and widely applicable information criterion in singular learning theory.
-\emph{Journal of Machine Learning Research}, \emph{11}(12).
-
 \leavevmode\hypertarget{ref-wickham_welcome_2019}{}%

### 2021-01-21 Rmd/trajectories_manuscript.Rmd
Move appendices to main Rmd

--- 2021-01-13 Rmd/trajectories_manuscript.Rmd
@@
 
+source(here("R", "analyses.R"))
 ```
@@
 and (2) its association with a referent (e.g., a DOG).
-There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side [@bergleson2012].
-At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered [@jusczyk1995], as opposed to when only part of the word is uttered.
-At 12 months, infants show sensitivity to both vowel and consonant mispronunciations [@mani2010a], looking shorter to named pictures when their label is mispronounced.
+There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side [@bergelson_at_2012].
+At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered [@jusczyk_infants_1995], as opposed to when only part of the word is uttered.
+At 12 months, infants show sensitivity to both vowel and consonant mispronunciations [@mani_twelve_2010], looking shorter to named pictures when their label is mispronounced.
 Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones.
-One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon [@mcclelland2003].
+One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon [@mcclelland_trace_1986].
 
 Words are not learnt in isolation, but in contexts rich in word tokens and referents.
-The structure of toddlers' lexicon reflects this high connectivity [@wojcik2018; @hills2010; @fourtassi2020].
-At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing [@mani2010b; @mani2011].
-The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming  [@styles2009; @arias-trejo2013; @wojcik2013], and show evidence of inhibitory links between lexical-semantic entries [@chow2016].
-By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages [@chow2017].
-Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words [@chow2017; @rama2013], and  the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers.
-This suggests that the emergence of rich lexical networks -- both at the phonological and the semantic levels -- grows as a function of the number of words acquired by the toddler.
+The structure of toddlers' lexicon reflects this high connectivity [@wojcik_development_2018; @hills_longitudinal_2009; @fourtassi_growth_2020].
+At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing [@mani_infant_2010; @mani_phonological_2011].
+The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming  [@styles_infants_2009; @arias-trejo_what_2013; @wojcik_ontogeny_2013], and show evidence of inhibitory links between lexical-semantic entries [@chow_backward_2016; chow_vocabulary_2019].
+By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages [@chow_spoken_2017].
+Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words [@chow_spoken_2017; @novogrodsky_age_2020], and  the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers.
+This suggests that the emergence of rich lexical network-- both at the phonological and the semantic levels--grows as a function of the number of words acquired by the toddler.
 
 The structure of the lexicon also impacts the order of acquisition of new words.
-Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills2009].
-@borosky2016 and @borovsky2018 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
+Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills_longitudinal_2009].
+@borovsky_lexical_2016 and @borovsky_slowing_2020 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
 This points to the structure of the lexicon facilitating the strategies children engage during word learning.
+
+@babineau_familiar_2020
 
@@
 When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev_influence_1977; @doyle1977some; @rosenblum1983word; @fernandez1992bilingual; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
+When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@leopold_study_1948; @ben-zeev_influence_1977; @doyle_some_1977; @rosenblum_word_1983; @fernandez_bilingual_1992; @umbel_measuring_1992; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
 
-@floccia_i_2018, gathered parental vocabulary estimates [@hamilton2000infant] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
+@floccia_i_2018, gathered parental vocabulary estimates [@hamilton_infant_2000] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
 They computed the average phonological similarity between translation equivalents (TEs from now on; e.g.,  *table* in English and *tafel* in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
@@
 The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., *cat*, in English) its corresponding translation in the other language is activated too [e.g., *gato*, in Spanish.
-There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell1991mediated; @spivey1999cross; @costa2000lexical; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh2014one; @von_holzen_impact_2019; @bobb_co-activation_2020; see @costa2005facilitatory for review], and children [e.g.
-@von_holzen_language_2012; @poulin2013lexical; @jardak2019labels; @bosma2020cognate].
+There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell_mediated_1991; @spivey_cross_1999; @costa_cognate_2000; @costa_lexical_2000; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh_one_2014; @von-holzen_impact_2019; @bobb_co-activation_2020; see @costa_facilitatory_2005 for review], and children [e.g.
+@von-holzen_language_2012; @poulin_lexical_2013; @jardak_labels_2019; @bosma_cognate_2020].
 A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
@@
 Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language [@floccia_i_2018].
-@dijkstra2013 tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE.
+@dijkstra_role_2016 tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE.
 Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian.
 This suggests that infants that were mostly exposed to Dutch used the phonology of  dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
-A study by @dijkstra2013 showed that the degree of phonological similarity between 
+A study by @dijkstra_role_2016 showed that the degree of phonological similarity between 
 the performance of  was better for cognates than for non-cognates, only for those children
@@
 In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.
+
+@taverna_early_2020
 
@@
 Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish).
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
 Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
@@
 We gathered `r printnum(n_obs, big.mark = ",")` responses.
-Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
-We modelled the probability of responses ($y_i$) belonging to any of the three possible categories $k$--1 = *No* or say, 2 = *Understands*, and 3 = *Understands and Says*-- using an ordinal regression model with a cumulative logit link function [@agresti_analysis_2010; @burkner_ordinal_2019].
-This model estimates the probability of response $y_i$ to belong to category $k$ or lower.
-In our case, this will result in the estimation of the probability of $y_i \leq 1$,  $y_i \leq 2$, and  $y_i \leq 3$, where
+Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, *Min* = `r printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
+We modelled the probability of children understanding or producing TEs separately using logistic regression model with a binomial distribution and a logit link function.
+[@agresti_analysis_2010].
+We included several predictors of interest to adjust this probability to participants' and items' properties of interest.
+These predictors are:
 
-$$
-p_k = Pr(y_i = k) = Pr(y_i \leq k) - Pr(y_i \leq k-1).
-$$
+* Age of participant in months (`age`, $Age$) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width.
+* Degree of bilingualism of participant (`bilingualism`, $Bilingualism$) computed as the percentage of exposure to a second language (Spanish for participants exposed to >50% to Catalan an *vice versa*), chunked into six groups of participants, with degree of bilingualism 0%, 10%, 20%, 30%, 40% and 50%)
+* Lexical frequency of item (`frequency`, $Frequency$) retrieved from SUBTLEX-CAT [@boada_subtlex-cat_2019] for Catalan words and from SUBTLEX-ESP [@cuetos_subtlex-esp_2011] for Spanish words, expressed as Zipf scores [@zipf_human_1949; @heuven_subtlex-uk_2014], and centred around the mean
+* Language the item belongs to (`dominance`, $Dominance$) labelled as `L1` if the item belongs to the dominant language (e.g., a Catalan item is labelled as `L1` for Catalan-dominant participants and as `L2` for Spanish-dominant participants)
+* Cognateness of the item (`cognate`, $Cognate$), as rated by a trained Spanish-Catalan bilingual linguist.
 
-We include several predictors of interest to adjust this probability to participants' and items' (lexical frequency, dominance, cognateness) characteristics. These predictors are:
+We aggregated the data across participants.
+We summed the number of participants that *understood*, or *understood and produced* each TE for each combination of the `age`, `bilingualism` levels.
+Numerical variables (`age`, `bilingualism`, and `frequency`) were standardized, and categorical predictors were sum-coded.
+`dominance` was coded as `L2` = -0.5 and `L1` = +0.5, and `cognate` was coded as `Non-cognate` = -0.5 and `Cognate` = +0.5.
+In summary, each TE provided four data points per age bin: L1-Cognate, L1-Non-cognate, L2-Cognate, and L2-Non-cognate.
 
-* Age of participant in months (`age`, $Age$) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30 and centred around the mean
-* Degree of bilingualism of participant (`bilingualism`, $Bilingualism$) computed as the percentage of exposure to a second language (Spanish for participants exposed to >50% to Catalan an *vice versa*) rounded to units and centered around the mean
-* Lexical frequency of item (`frequency`, $Frequency$) retrieved from SUBTLEX-CAT [@boada_subtlex-cat_2019] for Catalan words and from SUBTLEX-ESP [@cuetos_subtlex-esp_2011] for Spanish words, expressed as Zipf scores [@heuven_subtlex-uk_2014], and centred around the mean
-* Language the item belongs to (`dominance`, $Dominance$) labelled as `L1` if the item belongs to the dominant language (e.g., a Catalan item is labelled as `L1` for Catalan-dominant participants and as `L2` for Spanish-dominant participants), coded as `L2` = -0.5 and `L1` = 0.5
-* Cognateness of the item (`cognate`, $Cognate$), coded as `Non-cognate` = -0.5 and `Cognate` = 0.5
-
-We adopted a Bayesian approach toward statistical inference, which allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model.
+We adopted a Bayesian approach toward statistical inference.
+This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model [@mcelreath_statistical_2016].
+We first fit a base model that only adjusted for the age of participants and the lexical frequency of the items [@braginsky_consistency_2016: @jones_children_2018].
+We then introduced our predictors of interest in the following order: (1) `dominance`, (2) `bilingualism`, (3) `dominance` per `bilingualism` interaction, (4) `cognate`, (5) `dominance` per `cognate` interaction, (6) `bilingualism` per `cognate` interaction, and (7) `dominance` per `bilingualism` per `cognate` interaction.
 Our extended model can be formalised as follows:
@@
 \begin{aligned}
-y_i \sim Ordered(p) \\
-log\frac{Pr(y_i \leq k)}{1-Pr(y_i \leq k)} &= \alpha_k - \dots \\
+y_i \sim Binomial(n_i, p_i), \\
+logit(p_i) &=& \alpha - \dots \\
 & \beta_1 \times Age_i + \dots \\
@@
 & \beta_9 \times (Dominance_i \times Bilingualism_i \times Cognate_i), \\
-\alpha \sim Normal(0, 10) \\
-\beta_{0-9} \sim Normal(0, 10)
+logit(p_i) = log\frac{y_i}{n_i-y_i} \\
+\alpha \sim Normal(0, 1.5) \\
+\beta_{0-9} \sim Normal(0, 0.5)
 \end{aligned}
@@
 
-* $k$ is one of the possible categories (coded as ) responses can take
-* $y_i$ is the `response` value of observation $i$
-* $\alpha_k$ is the mean probability of responses to belong to category $k$
+* $y_i$ is the sum of positive responses to *Understands* or *Understands and Says* for translation equivalent $i$
+* $n_i$ is the total number of responses for translation equivalent $i$
+* $\alpha$ is the intercept, which corresponds to the probability of a positive response when all predictors equal 0
 * $\beta_{1, \dots, 9}$ are the coefficients of the predictors
 
-We implemented the model using the R package `brms` [@burkner_brms_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 4,000 iterations (including 2,000 warm-up iterations per chain).
-We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal due to the large amount of observations introduced in the model.
-We report the 95% credible intervals (CrI) of the estimated posterior distribution of each parameter, which denotes a range of values we are 95% certain contain the true value, given our prior and observed data.
-We report analyses on the probability scale, having transformed the log-odds returned by the model using an inverse logit transformation.
+We implemented this model using the R package `brms` [@R-brms_b; carpenter_stan_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
+Appendix 2 shows different diagnostics of chain convergence.
+We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017]^[Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson_leave_2019]].
+This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
+The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
+We compared the expected log posterior density of each each model to test which one fitted the data the best.
 
-We then fit increasingly simpler models dropping predictors one by one in the following order: (1) `dominance:bilingualism:cognate`, (2) `bilingualism:cognate`, (3) `dominance:cognate:`, (4) `cognate`, (5) `dominance:bilingualism`, (6) `bilingualism`, (7) `dominance`, (8) `frequency`, and (9) `age`.
-We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017] and widely applicable information criterion (WAIC) [@watanabe_asymptotic_2010].
-We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95% credible intervals and testing whether this interval excluded 0.
-We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@lenth_emmeans_2020].
-Data processing and visualisation was done in R [@rcoreteam_r_2020] using the `tidyverse` family of packages [@wickham_welcome_2019], and posterior samples were extracted using the `tidybayes` R package [@kay_tidybayes_2020].
+We then explored the posterior distribution of each parameter in the model that fitted the data the best computing the 95% credible intervals and testing whether this interval excluded 0.
+Credible intervals (CrI) indicate the range of values we are 95% certain contain the true value of the parameter, given our prior and observed data.
+We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@R-emmeans].
+Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], and posterior samples were extracted using the `tidybayes` R package [@R-tidybayes].
 
 # Results
+
+```{r results, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+load(here("Results", "comp_fits.RData"))
+load(here("Results", "prod_fits.RData"))
+load(here("Results", "loos.RData"))
+
+coef_comp <- as.data.frame(fixef(comp_7))
+```
+
+## Comprehension
+
+The extended model, including the three-way interaction `dominance1:bilingualism;cognate1` fitted comprehension data the best: it showed the smallest absolute expected log predictive density.
+This estimate is several times larger than its standard error indicating high predictive accuracy.
+Table 2 summarises the LOO scores of the fitted models.
+
+```{r loos, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+loos_diff <- bind_rows(
+  "Comprehension" = select(
+    as.data.frame(loos_comp$diffs),
+    one_of("elpd_loo", "se_elpd_loo", "p_loo", "elpd_diff", "se_diff")
+  ),
+  "Production" = select(
+    as.data.frame(loos_prod$diffs),
+    one_of("elpd_loo", "se_elpd_loo", "p_loo", "elpd_diff", "se_diff")
+  ),
+  .id = "type"
+) %>% 
+  mutate(
+    model = rep(rev(c(
+      "~ 1 + age + frequency",
+      "... + dominance",
+      "... + bilingualism",
+      "... + dominance:bilingualism",
+      "... + cognate",
+      "... + dominance:cognate",
+      "... + bilingualism:cognate",
+      "... + dominance:bilingualism:cognate"
+    )), 2)) %>% 
+  relocate(model) %>% 
+  group_by(type) %>% 
+  arrange(elpd_loo)
+
+gt(loos_diff) %>% 
+  tab_spanner(
+    label = md("**Comparison**"),
+    columns = c("elpd_diff", "se_diff")
+  ) %>% 
+  cols_label(
+    model = md("**Model**"),
+    elpd_loo = md("**ELPD**"),
+    se_elpd_loo = md("**SE**"),
+    p_loo = md("**p**"),
+    elpd_diff = md("*LOO<sub>diff</sup>*"),
+    se_diff = md("*SE<sub>diff</sup>*")
+  ) %>% 
+  fmt_number(2:5) %>% 
+  tab_footnote(
+    footnote = md("*ELPD*: theoretical expected log pointwise predictive density for a new dataset, as estimated using Bayesian LOO-CV"),
+    locations = cells_column_labels(columns = vars("elpd_loo"))
+  ) %>% 
+  tab_footnote(
+    footnote = md("*SE*: standard error of the ELPD"),
+    locations = cells_column_labels(columns = vars("se_elpd_loo"))
+  ) %>% 
+  tab_footnote(
+    footnote = md("*p*: effective number of parameters"),
+    locations = cells_column_labels(columns = vars("p_loo"))
+  ) %>% 
+  tab_options(footnotes.marks = "letters")
+
+```
+
+We report the mean of each coefficient's posterior distribution along with standard errors and 95% CrI in Table 3.
+For interpretability, we report de derivative of the logistic function^[The derivative is calculated as $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.
+This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.] for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable [@gelman_regression_2020].
+
+```{r coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+estimates <- bind_cols(
+  "Comprehension" =  rename_all(as.data.frame(inv_logit_scaled(fixef(comp_7))), ~paste0(., "_comp")),
+  "Production" =  rename_all(as.data.frame(inv_logit_scaled(fixef(prod_7))), ~paste0(., "_prod"))
+) %>% 
+  mutate(
+    coefficient = c(
+      "Intercept",
+      "Age",
+      "Frequency",
+      "Dominance",
+      "Bilingualism",
+      "Cognate",
+      "Dominance * Bilingualism",
+      "Dominance * Cognate",
+      "Bilingualism * Cognate",
+      "Dominance * Bilingualism * Cognate"
+    )
+  ) %>% 
+  clean_names() %>% 
+  relocate(coefficient)
+
+rownames(estimates) <- NULL
+
+gt(estimates) %>% 
+  tab_spanner(label = md("**Comprehension**"), columns = matches("_comp")) %>% 
+  tab_spanner(label = md("**Production**"), columns = matches("_prod")) %>% 
+  fmt_number(2:9) %>% 
+  cols_merge_range(col_begin = "q2_5_comp", col_end = "q97_5_comp") %>% 
+  cols_merge_range(col_begin = "q2_5_prod", col_end = "q97_5_prod") %>% 
+  cols_label(
+    coefficient = md("**Coefficient**"),
+    estimate_comp = md("*Mean*"),
+    est_error_comp = md("*SE*"),
+    q2_5_comp = md("*95% CrI*"),
+    estimate_prod = md("*Mean*"),
+    est_error_prod = md("*SE*"),
+    q2_5_prod = md("*95% CrI*"),
+  ) %>% 
+  tab_footnote(
+    footnote = "SE: Standard Error of the mean",
+    locations = cells_column_labels(columns = matches("est_error"))
+  ) %>% 
+  tab_footnote(
+    footnote = "95% CrI: 95% credible interval",
+    locations = cells_column_labels(columns = matches("q2_5"))
+  )
+
+```
+
+The intercept (`r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`%) indicates probability of participants understanding TEs at mean age (`r printnum(mean(responses$age_scaled))` months), mean frequency (`r printnum(mean(responses$frequency_scaled))` Zipf score), and mean L2 exposure (`r printnum(mean(responses$bilingualism_scaled))`%), across the levels of the `dominance` and `cognate` factors.
+An increase in one standard deviation increment in age (1 *SD* = `r printnum(sd(responses$age_scaled))` months) is associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability.
+A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency_scaled))` Zipf scores) is associated with a `r printnum(coef_comp["frequency", "Estimate"]*100/4)`% increment in comprehension probability.
+Items in the dominant language were `r printnum(coef_comp["dominance1", "Estimate"]*100/4)`% more likely to be understood than items in the non-dominant language.
+A one standard deviation increase in exposure to the non-dominant language (1 *SD* = `printnum(sd(responses$bilingualism_scaled))`%) is associated with a `r printnum(coef_comp["bilingualism", "Estimate"]*100/4)` increment in comprehension probability.
+Cognates were `r printnum(coef_comp["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates.
+The 95% credible interval of the three-way interaction (`dominance:bilingualism:cognate`) excluded 0, indicating that that the cognate status of the items changed the effect of dominance differently depending on the amount of exposure to the second language.
+Figure 1 shows the estimated marginal mean posterior probability of comprehension and production across ages, and the different levels of `dominance`, `bilingualism`, and `cognate`.
+
+Children exposed to their second language more than -1 SD away from the mean (`r printnum(mean(responses$bilingualism_scaled)-sd(responses$bilingualism_scaled))`%) were more likely to understand cognates than non-cognates in their non-dominance language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[6]*100)`%, 95% CrI = [`r printnum(deltas$.lower[6]*100)`-`r printnum(deltas$.upper[6]*100)`]), but not in their dominant language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[2]*100)`%, 95% CrI = [`r printnum(deltas$.lower[2]*100)`-`r printnum(deltas$.upper[2]*100)`]).
+This difference was smaller for participants with mean exposure to the non-dominant language (`r printnum(mean(responses$bilingualism_scaled))`%): in the non-dominant language cognates were slight more likely to be understood than non-cognates ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[12]*100)`%, 95% CrI = [`r printnum(deltas$.lower[12]*100)`-`r printnum(deltas$.upper[12]*100)`]), but not in the dominant language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[8]*100)`%, 95% CrI = [`r printnum(deltas$.lower[8]*100)`-`r printnum(deltas$.upper[8]*100)`]).
+Finally this effect was negligible for participants exposed to their second language more than -1 SD away from the mean (`r printnum(mean(responses$bilingualism_scaled)+sd(responses$bilingualism_scaled))`%).
+Cognates and non-cognates were equally likely to be understood in both the non-dominant ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[18]*100)`%, 95% CrI = [`r printnum(deltas$.lower[18]*100)`-`r printnum(deltas$.upper[18]*100)`]), and the dominant language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[14]*100)`%, 95% CrI = [`r printnum(deltas$.lower[14]*100)`-`r printnum(deltas$.upper[14]*100)`]).
+
+The difference in probability of comprehension between the dominant and the non-dominant languages differed between cognates and non-cognates and different levels of exposure to the non-dominant language.
+For children exposed.
+For children exposed to their non-dominant language less than 1 SD from the mean, difference in probability between the dominant and the non-dominant language was smaller for cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[4]*100)`%, 95% CrI = [`r printnum(deltas$.lower[4]*100)`-`r printnum(deltas$.upper[4]*100)`]) than for non-cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[1]*100)`%, 95% CrI = [`r printnum(deltas$.lower[1]*100)`-`r printnum(deltas$.upper[1]*100)`]).
+For children with mean exposure to the non-dominant language, this effect was smaller: the difference between the comprehension probability of items in L1 and L2 was slightly smaller for cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[10]*100)`%, 95% CrI = [`r printnum(deltas$.lower[10]*100)`-`r printnum(deltas$.upper[10]*100)`]) than for non-cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[7]*100)`%, 95% CrI = [`r printnum(deltas$.lower[7]*100)`-`r printnum(deltas$.upper[7]*100)`]).
+Finally this effect was negligible for children exposed more than 1 SD than the mean to their non-dominant language: the difference in comprehension probability between items in the dominant and the non-dominant language was roughly similar for cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[16]*100)`%, 95% CrI = [`r printnum(deltas$.lower[16]*100)`-`r printnum(deltas$.upper[16]*100)`]) and non-cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[13]*100)`%, 95% CrI = [`r printnum(deltas$.lower[13]*100)`-`r printnum(deltas$.upper[13]*100)`]).
+
+```{r marginalmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6.5, fig.height=6.5, out.width="20cm", fig.cap="Mean posterior probability and 95\\% CrI of responses, conditional to Dominance, Bilingualism, and Cognate."}
+
+post_preds %>% 
+  ggplot(aes(
+    age, .value,
+    colour = interaction(dominance, cognate, sep = " - "),
+    fill = interaction(dominance, cognate, sep = " - ")
+  )) +
+  facet_wrap(type~bilingualism) +
+  #geom_line(aes(group = interaction(item_dominance, cognate, doe, .draw)), size = 0.4) +
+  stat_lineribbon(.width = 0.95, colour = NA, alpha = 0.5) +
+  stat_summary(fun = "mean", geom = "line", size = 0.75) +
+  #geom_point(data = filter(proportion, type=="Understands"), alpha = 0.5) +
+  labs(x = "Age (months)", y = "P(Y|X)") +
+  guides(colour = guide_legend(ncol = 2)) +
+  scale_color_brewer(palette = "Dark2") +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_custom() +
+  theme(
+    legend.position = "top",
+    legend.title = element_blank(),
+    legend.direction = "horizontal"
+  )
+
+```
 

### 2021-01-21 Rmd/trajectories_manuscript.tex
Move appendices to main Rmd

@@
 
-We collected data from 349 bilinguals (167 female), from the Metropolitan Area of Barcelona, between 28th October, 2019 and 09th January, 2021.
+We collected data from 349 bilinguals (182 female), from the Metropolitan Area of Barcelona, between 28th October, 2019 and 09th January, 2021.
 All families gave informed consent before participating.
@@
 
-We gathered 181587 responses.
-Translation equivalents received an average of 386.36 (\emph{SD} = 228.76, , Max = 698), both languages summed together.
-We modelled the probability of responses (\(y_i\)) belonging to any of the three possible categories \(k\)--1 = \emph{No} or say, 2 = \emph{Understands}, and 3 = \emph{Understands and Says}-- using an ordinal regression model with a cumulative logit link function (Agresti, 2010; Bürkner \& Vuorre, 2019).
-This model estimates the probability of response \(y_i\) to belong to category \(k\) or lower.
-In our case, this will result in the estimation of the probability of \(y_i \leq 1\), \(y_i \leq 2\), and \(y_i \leq 3\), where
-
-\[
-p_k = Pr(y_i = k) = Pr(y_i \leq k) - Pr(y_i \leq k-1).
-\]
-
-We include several predictors of interest to adjust this probability to participants' and items' (lexical frequency, dominance, cognateness) characteristics. These predictors are:
+We gathered 89220 responses.
+Translation equivalents received an average of 189.83 (\emph{SD} = 67.27, \emph{Min} = 75, Max = 312), both languages summed together.
+We modelled the probability of children understanding or producing TEs separately using logistic regression model with a binomial distribution and a logit link function.
+(Agresti, 2010).
+We included several predictors of interest to adjust this probability to participants' and items' properties of interest.
+These predictors are:
 
@@
 \item
-  Age of participant in months (\texttt{age}, \(Age\)) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30 and centred around the mean
+  Age of participant in months (\texttt{age}, \(Age\)) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width.
 \item
-  Degree of bilingualism of participant (\texttt{bilingualism}, \(Bilingualism\)) computed as the percentage of exposure to a second language (Spanish for participants exposed to \textgreater50\% to Catalan an \emph{vice versa}) rounded to units and centered around the mean
+  Degree of bilingualism of participant (\texttt{bilingualism}, \(Bilingualism\)) computed as the percentage of exposure to a second language (Spanish for participants exposed to \textgreater50\% to Catalan an \emph{vice versa}), chunked into six groups of participants, with degree of bilingualism 0\%, 10\%, 20\%, 30\%, 40\% and 50\%)
 \item
@@
 \item
-  Language the item belongs to (\texttt{dominance}, \(Dominance\)) labelled as \texttt{L1} if the item belongs to the dominant language (e.g., a Catalan item is labelled as \texttt{L1} for Catalan-dominant participants and as \texttt{L2} for Spanish-dominant participants), coded as \texttt{L2} = -0.5 and \texttt{L1} = 0.5
+  Language the item belongs to (\texttt{dominance}, \(Dominance\)) labelled as \texttt{L1} if the item belongs to the dominant language (e.g., a Catalan item is labelled as \texttt{L1} for Catalan-dominant participants and as \texttt{L2} for Spanish-dominant participants)
 \item
-  Cognateness of the item (\texttt{cognate}, \(Cognate\)), coded as \texttt{Non-cognate} = -0.5 and \texttt{Cognate} = 0.5
+  Cognateness of the item (\texttt{cognate}, \(Cognate\)), as rated by a trained Spanish-Catalan bilingual linguist.
 \end{itemize}
 
-We adopted a Bayesian approach toward statistical inference, which allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model.
+We aggregated the data across participants.
+We summed the number of participants that \emph{understood}, or \emph{understood and produced} each TE for each combination of the \texttt{age}, \texttt{bilingualism} levels.
+Numerical variables (\texttt{age}, \texttt{bilingualism}, and \texttt{frequency}) were standardized, and categorical predictors were sum-coded.
+\texttt{dominance} was coded as \texttt{L2} = -0.5 and \texttt{L1} = +0.5, and \texttt{cognate} was coded as \texttt{Non-cognate} = -0.5 and \texttt{Cognate} = +0.5.
+In summary, each TE provided four data points per age bin: L1-Cognate, L1-Non-cognate, L2-Cognate, and L2-Non-cognate.
+
+We adopted a Bayesian approach toward statistical inference.
+This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model ({\textbf{???}}).
+We first fit a base model that only adjusted for the age of participants and the lexical frequency of the items ({\textbf{???}} @jones\_do\_2018).
+We then introduced our predictors of interest in the following order: (1) \texttt{dominance}, (2) \texttt{bilingualism}, (3) \texttt{dominance} per \texttt{bilingualism} interaction, (4) \texttt{cognate}, (5) \texttt{dominance} per \texttt{cognate} interaction, (6) \texttt{bilingualism} per \texttt{cognate} interaction, and (7) \texttt{dominance} per \texttt{bilingualism} per \texttt{cognate} interaction.
 Our extended model can be formalised as follows:
@@
 \item
-  \(k\) is one of the possible categories (coded as ) responses can take
+  \(y_i\) is the sum of positive responses to \emph{Understands} or \emph{Understands and Says} for translation equivalent \(i\)
 \item
-  \(y_i\) is the \texttt{response} value of observation \(i\)
+  \(n_i\) is the total number of responses for translation equivalent \(i\)
 \item
-  \(\alpha_k\) is the mean probability of responses to belong to category \(k\)
+  \(\alpha\) is the intercept, which corresponds to the probability of a positive response when all predictors equal 0
 \item
@@
 
-We implemented the model using the R package \texttt{brms} (Bürkner, 2017) as \texttt{response\ \textasciitilde{}\ \ age\ +\ frequency\ +\ dominance*bilingualism*cognate}, running four MCMC sampling chains with 4,000 iterations (including 2,000 warm-up iterations per chain).
-We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal due to the large amount of observations introduced in the model.
-We report the 95\% credible intervals (CrI) of the estimated posterior distribution of each parameter, which denotes a range of values we are 95\% certain contain the true value, given our prior and observed data.
-We report analyses on the probability scale, having transformed the log-odds returned by the model using an inverse logit transformation.
+We implemented this model using the R package \texttt{brms} (Bürkner, 2017) as \texttt{response\ \textasciitilde{}\ \ age\ +\ frequency\ +\ dominance*bilingualism*cognate}, running four MCMC sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
+Appendix 2 shows different diagnostics of chain convergence.
+We compared all models using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017)\footnote{Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples (Magnusson, Andersen, Jonasson, \& Vehtari, 2019)}.
+This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
+The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
+We compared the expected log posterior density of each each model to test which one fitted the data the best.
 
-We fit increasingly simpler models dropping predictors one by one in the following order: (1) \texttt{dominance:bilingualism:cognate}, (2) \texttt{bilingualism:cognate}, (3) \texttt{dominance:cognate:}, (4) \texttt{cognate}, (5) \texttt{dominance:bilingualism}, (6) \texttt{bilingualism}, (7) \texttt{dominance}, (8) \texttt{frequency}, and (9) \texttt{age}.
-We compared all models using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017)\footnote{Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1000 samples (Magnusson, Andersen, Jonasson, \& Vehtari, 2019)}.
-We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95\% credible intervals and testing whether this interval excluded 0.
+We then explored the posterior distribution of each parameter in the model that fitted the data the best computing the 95\% credible intervals and testing whether this interval excluded 0.
+Credible intervals (CrI) indicate the range of values we are 95\% certain contain the true value of the parameter, given our prior and observed data.
 We performed follow-up tests on interactions whose 95\% credible interval excluded 0 by comparing the 95\% credible interval of their estimated marginal means using the \texttt{emmeans} package (Lenth, 2020).
@@
 
-Table 2 summarises shows the outcome of the LOO-CV model comparison. The extended model was most favoured by the data.
+\hypertarget{comprehension}{%
+\subsection{Comprehension}\label{comprehension}}
+
+The extended model, including the three-way interaction \texttt{dominance1:bilingualism;cognate1} fitted comprehension data the best: it showed the smallest absolute expected log predictive density.
+This estimate is several times larger than its standard error indicating high predictive accuracy.
+Table 2 summarises the LOO scores of the fitted models.
 
 \captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{crrrrr}
+\begin{longtable}{lrrrrr}
 \toprule
@@
 \midrule
-\textasciitilde  1 (intercept only) & $-157,272.37$ & $250.56$ & $12.00$ & $0.00$ & 0.0000 \\ 
-... + age & $-157,308.76$ & $250.47$ & $10.37$ & $36.39$ & 354.2836 \\ 
-... + frequency & $-157,318.93$ & $250.45$ & $9.93$ & $46.56$ & 354.2703 \\ 
-... + dominance & $-157,463.31$ & $250.25$ & $8.08$ & $190.94$ & 354.1293 \\ 
-... + bilingualism & $-157,472.54$ & $250.29$ & $7.67$ & $200.17$ & 354.1564 \\ 
-... + dominance:bilingualism & $-158,724.35$ & $246.22$ & $6.33$ & $1,451.98$ & 351.2863 \\ 
-... + cognate & $-159,063.42$ & $245.10$ & $5.09$ & $1,791.06$ & 350.5085 \\ 
-... + dominance:cognate & $-161,704.16$ & $237.62$ & $3.95$ & $4,431.79$ & 345.3184 \\ 
-... + bilingualism:cognate & $-165,570.87$ & $228.08$ & $3.13$ & $8,298.50$ & 338.8223 \\ 
-... + dominance:bilingualism:cognate & $-184,511.91$ & $172.86$ & $2.04$ & $27,239.54$ & 304.4033 \\ 
+\multicolumn{1}{l}{Production} \\ 
+\midrule
+\textasciitilde  1 + age + frequency & $-73,655.36$ & $277.24$ & $5.93$ & 4577.29336 & 380.0306 \\ 
+... + dominance & $-71,344.56$ & $271.05$ & $7.19$ & 2266.48651 & 375.5426 \\ 
+... + bilingualism & $-70,706.33$ & $267.57$ & $9.23$ & 1628.26183 & 373.0344 \\ 
+... + dominance:bilingualism & $-69,246.49$ & $261.81$ & $11.57$ & 168.42565 & 368.9302 \\ 
+... + cognate & $-69,242.57$ & $261.70$ & $10.65$ & 164.49868 & 368.8513 \\ 
+... + dominance:cognate & $-69,143.92$ & $260.65$ & $13.29$ & 65.84950 & 368.1056 \\ 
+... + bilingualism:cognate & $-69,129.05$ & $260.58$ & $14.47$ & 50.98597 & 368.0542 \\ 
+... + dominance:bilingualism:cognate & $-69,078.07$ & $259.93$ & $15.72$ & 0.00000 & 0.0000 \\ 
+\midrule
+\multicolumn{1}{l}{Comprehension} \\ 
+\midrule
+\textasciitilde  1 + age + frequency & $-73,653.35$ & $277.26$ & $5.03$ & 4577.41755 & 380.0019 \\ 
+... + dominance & $-71,348.67$ & $271.01$ & $6.44$ & 2272.73124 & 375.4690 \\ 
+... + bilingualism & $-70,702.30$ & $267.45$ & $6.57$ & 1626.36697 & 372.9028 \\ 
+... + dominance:bilingualism & $-69,245.99$ & $261.71$ & $8.87$ & 170.05482 & 368.8094 \\ 
+... + cognate & $-69,240.75$ & $261.61$ & $9.43$ & 164.81145 & 368.7373 \\ 
+... + dominance:cognate & $-69,142.47$ & $260.63$ & $9.67$ & 66.53140 & 368.0479 \\ 
+... + bilingualism:cognate & $-69,134.13$ & $260.64$ & $15.77$ & 58.19037 & 368.0546 \\ 
+... + dominance:bilingualism:cognate & $-69,075.94$ & $259.86$ & $17.30$ & 0.00000 & 0.0000 \\ 
 \bottomrule
@@
 
-Table 3 shows the mean and 95\% CrI of each regression coefficient of the best fitting model.
+We report the mean of each coefficient's posterior distribution along with standard errors and 95\% CrI in Table 3.
+For interpretability, we report de derivative of the logistic function\footnote{The derivative is calculated as \(\hat{\beta_j}/4\), where \(\hat{\beta_j}\) is the estimated mean of the posterior distribution of coefficient \(j\).
+This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.} for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable ({\textbf{???}}).
 
@@
 \toprule
-& \multicolumn{3}{c}{\textbf{Odds Ratio}} & \multicolumn{3}{c}{\textbf{Probability}} \\ 
+& \multicolumn{3}{c}{\textbf{Comprehension}} & \multicolumn{3}{c}{\textbf{Production}} \\ 
  \cmidrule(lr){2-4}\cmidrule(lr){5-7}
-\textbf{Coefficient} & \emph{Mean} & \emph{SE}\textsuperscript{1} & \emph{95\% CrI}\textsuperscript{2} & \emph{Mean} & \emph{SE} & \emph{95\% CrI} \\ 
+\textbf{Coefficient} & \emph{Mean} & \emph{SE}\textsuperscript{1} & \emph{95\% CrI}\textsuperscript{2} & \emph{Mean} & \emph{SE}\textsuperscript{1} & \emph{95\% CrI}\textsuperscript{2} \\ 
 \midrule
-Intercept 1 (Understands) & $1.12$ & $1.01$ & $1.11$–$1.14$ & $0.53$ & $0.50$ & $0.53$–$0.53$ \\ 
-Intercept 2 (Understands \& Says) & $5.31$ & $1.01$ & $5.24$–$5.38$ & $0.84$ & $0.50$ & $0.84$–$0.84$ \\ 
-Age & $1.16$ & $1.00$ & $1.16$–$1.17$ & $0.54$ & $0.50$ & $0.54$–$0.54$ \\ 
-Frequency & $1.86$ & $1.01$ & $1.83$–$1.88$ & $0.65$ & $0.50$ & $0.65$–$0.65$ \\ 
-Dominance & $2.27$ & $1.01$ & $2.22$–$2.32$ & $0.69$ & $0.50$ & $0.69$–$0.70$ \\ 
-Bilingualism & $1.10$ & $1.00$ & $1.09$–$1.11$ & $0.52$ & $0.50$ & $0.52$–$0.53$ \\ 
-Cognate & $1.08$ & $1.01$ & $1.06$–$1.10$ & $0.52$ & $0.50$ & $0.51$–$0.52$ \\ 
-Dominance * Bilingualism & $0.73$ & $1.01$ & $0.73$–$0.74$ & $0.42$ & $0.50$ & $0.42$–$0.43$ \\ 
-Dominance * Cognate & $0.68$ & $1.02$ & $0.65$–$0.71$ & $0.40$ & $0.51$ & $0.39$–$0.42$ \\ 
-Bilingualism * Cognate & $0.97$ & $1.01$ & $0.96$–$0.98$ & $0.49$ & $0.50$ & $0.49$–$0.50$ \\ 
-Dominance * Bilingualism * Cognate & $1.13$ & $1.01$ & $1.10$–$1.15$ & $0.53$ & $0.50$ & $0.52$–$0.54$ \\ 
+Intercept & $0.48$ & $0.50$ & $0.48$–$0.48$ & $0.14$ & $0.50$ & $0.14$–$0.14$ \\ 
+Age & $0.76$ & $0.50$ & $0.75$–$0.76$ & $0.81$ & $0.50$ & $0.81$–$0.81$ \\ 
+Frequency & $0.61$ & $0.50$ & $0.61$–$0.62$ & $0.61$ & $0.50$ & $0.61$–$0.61$ \\ 
+Dominance & $0.68$ & $0.50$ & $0.68$–$0.69$ & $0.71$ & $0.50$ & $0.71$–$0.72$ \\ 
+Bilingualism & $0.55$ & $0.50$ & $0.55$–$0.56$ & $0.54$ & $0.50$ & $0.53$–$0.54$ \\ 
+Cognate & $0.51$ & $0.50$ & $0.50$–$0.51$ & $0.55$ & $0.50$ & $0.54$–$0.55$ \\ 
+Dominance * Bilingualism & $0.34$ & $0.50$ & $0.34$–$0.35$ & $0.37$ & $0.50$ & $0.37$–$0.38$ \\ 
+Dominance * Cognate & $0.42$ & $0.51$ & $0.41$–$0.43$ & $0.39$ & $0.51$ & $0.38$–$0.41$ \\ 
+Bilingualism * Cognate & $0.48$ & $0.50$ & $0.48$–$0.49$ & $0.49$ & $0.50$ & $0.48$–$0.50$ \\ 
+Dominance * Bilingualism * Cognate & $0.56$ & $0.51$ & $0.55$–$0.57$ & $0.56$ & $0.51$ & $0.54$–$0.57$ \\ 
 \bottomrule
@@
 \end{minipage}
+
+The intercept (47.98\%) indicates probability of participants understanding TEs at mean age (25.48 months), mean frequency (2.93 Zipf score), and mean L2 exposure (24.22\%), across the levels of the \texttt{dominance} and \texttt{cognate} factors.
+An increase in one standard deviation increment in age (1 \emph{SD} = 7.30 months) is associated with a 28.20\% increment in comprehension probability.
+A one standard deviation increase in lexical frequency (1 \emph{SD} = 1.41 Zipf scores) is associated with a 11.49\% increment in comprehension probability.
+Items in the dominant language were 19.23\% more likely to be understood than items in the non-dominant language.
+A one standard deviation increase in exposure to the non-dominant language (1 \emph{SD} = \texttt{printnum(sd(responses\$bilingualism\_scaled))}\%) is associated with a 5.48 increment in comprehension probability.
+Cognates were 0.62\% more like to be understood than non-cognates.
+The 95\% credible interval of the three-way interaction (\texttt{dominance:bilingualism:cognate}) excluded 0, indicating that that the cognate status of the items changed the effect of dominance differently depending on the amount of exposure to the second language.
+Figure 1 shows the estimated marginal mean posterior probability of comprehension and production across ages, and the different levels of \texttt{dominance}, \texttt{bilingualism}, and \texttt{cognate}.
+
+Children exposed to their second language more than -1 SD away from the mean (7.33\%) were more likely to understand cognates than non-cognates in their non-dominance language (\(M_{Cognate}-M_{Non-cognate}\) = 9.61\%, 95\% CrI = {[}8.33-10.87{]}), but not in their dominant language (\(M_{Cognate}-M_{Non-cognate}\) = -5.11\%, 95\% CrI = {[}-6.14--4.07{]}).
+This difference was smaller for participants with mean exposure to the non-dominant language (24.22\%): in the non-dominant language cognates were slight more likely to be understood than non-cognates (\(M_{Cognate}-M_{Non-cognate}\) = 4.83\%, 95\% CrI = {[}3.96-5.69{]}), but not in the dominant language (\(M_{Cognate}-M_{Non-cognate}\) = -3.58\%, 95\% CrI = {[}-4.36--2.81{]}).
+Finally this effect was negligible for participants exposed to their second language more than -1 SD away from the mean (41.11\%).
+Cognates and non-cognates were equally likely to be understood in both the non-dominant (\(M_{Cognate}-M_{Non-cognate}\) = 0.05\%, 95\% CrI = {[}-1.18-1.19{]}), and the dominant language (\(M_{Cognate}-M_{Non-cognate}\) = -2.05\%, 95\% CrI = {[}-3.20--0.91{]}).
+
+The difference in probability of comprehension between the dominant and the non-dominant languages differed between cognates and non-cognates and different levels of exposure to the non-dominant language.
+For children exposed.
+For children exposed to their non-dominant language less than 1 SD from the mean, difference in probability between the dominant and the non-dominant language was smaller for cognates (\(M_{Dominant}-M_{Non-dominant}\) = 27.97\%, 95\% CrI = {[}27.06-28.88{]}) than for non-cognates (\(M_{Dominant}-M_{Non-dominant}\) = 42.68\%, 95\% CrI = {[}41.31-44.03{]}).
+For children with mean exposure to the non-dominant language, this effect was smaller: the difference between the comprehension probability of items in L1 and L2 was slightly smaller for cognates (\(M_{Dominant}-M_{Non-dominant}\) = 15.03\%, 95\% CrI = {[}14.39-15.68{]}) than for non-cognates (\(M_{Dominant}-M_{Non-dominant}\) = 23.44\%, 95\% CrI = {[}22.48-24.37{]}).
+Finally this effect was negligible for children exposed more than 1 SD than the mean to their non-dominant language: the difference in comprehension probability between items in the dominant and the non-dominant language was roughly similar for cognates (\(M_{Dominant}-M_{Non-dominant}\) = 2.09\%, 95\% CrI = {[}1.19-3.03{]}) and non-cognates (\(M_{Dominant}-M_{Non-dominant}\) = 4.19\%, 95\% CrI = {[}2.81-5.58{]}).
 
@@
 Bürkner, P.-C. (2017). Brms: An R Package for Bayesian Multilevel Models Using Stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
-
-\leavevmode\hypertarget{ref-burkner_ordinal_2019}{}%
-Bürkner, P.-C., \& Vuorre, M. (2019). Ordinal regression models in psychology: A tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 77--101.
 

### 2021-05-17 Rmd/introduction.Rmd
First introduction draft

---
title: "Introduction"
author: "Gonzalo García-Castro"
date: "3/17/2021"
output:
    word_document:
        reference_docx: "trajectories_template.docx"
bibliography: "trajectories.bib"
csl: "apa6.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Before the end of their first year of life, infants are able to encode sound patterns [@jusczyk1995infants; @tincoff1999some] and associate them with their referents [@bergelson2012months]. During the second year, infants represent those word-forms with finer phonological detail [e.g. @werker2002infants], and from 18 to 24 months of age, they acquire new words at an increasingly faster rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension]. What words are acquired earlier at these ages provides rich information about the mechanisms and biases that constrain early vocabulary growth. For instance, @braginsky2019consistency modelled vocabulary data from a large sample of children learning a diverse pool of languages and found that, across languages, words were acquired earlier if they were nouns, were highly concrete, or appeared in short sentences on average. Despite this consistency in their trajectories of acquisition, children display large variability in the composition of their vocabulary, in part due to the diversity of their language input.

One of the causes of sources of this variability is the fact that a substantial proportion of children around the world learn to languages from birth. These children--to which we will refer as *bilinguals*--are a particular case of language learners: their linguistic input is divided into two languages. Taking into account that the relative amount of linguistic input the child receives in each language is a strong predictor of their vocabulary size [e.g., @cattani2014much; @thordardottir2006bilingual; @david2008individual; @pearson1997relation; @floccia2018introduction], one could predict that bilinguals know less words in each of their languages than their monolinguals peers. This is the case for some bilingual populations. For instance, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals that only learnt English. Other studies have provided converging evidence that bilinguals and monolinguals know a similar [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @junker2002expressive; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating] or even larger [@de2014bilingual; @aguila2008media] amount of words, only when summing up both languages together.

A recent study by @floccia2018introduction suggests that the size of the difference between bilinguals' and monolinguals' trajectory of vocabulary growth varies considerably, conditional to the similarity between the two languages bilinguals are learning. The authors collected vocabulary data from 24 month-old bilingual toddlers in the United Kingdom, who were learning English and an additional language. The additional language was one a pool of 13 diverse languages that differed in their phonological and grammatical similarity with English (e.g., Dutch, Italian, Mandarin). Among their many findings, the authors reported a positive association between language similarity and vocabulary size. Particularly, the average phonological overlap between translation equivalents (TEs) in both languages (i.e., cognateness) predicted larger productive vocabulary sizes. For instance, children learning English and Dutch, sharing 22.14% phonological overlap, were able to produce more words than children learning English and Mandarin, only sharing 1.97% phonological overlap. These results put Hoff et al. and others' results--obtained in participants relatively distant languages--in a scenario where the differences in vocabulary size between monolinguals and bilinguals are difficult to generalise to other bilingual populations whose languages are linguistically closer. In line with these findings, @aguila2008media reported that bilingual toddlers learning Catalan and Spanish--two lexically very close languages--showed larger total vocabulary sizes than monolinguals, following a different pattern than bilinguals learning English and Spanish in @hoff2012dual.

Floccia et al. pointed to parallel activation as the main mechanism behind their results. *Parallel activation* refers to the hypothesis that bilinguals activate both languages simultaneously during speech production or comprehension. This phenomenon is understood as the result of participants activating lexical representations in both languages, even when only using one of them [@thierry2007brain; @hoshino2008cognate; @bobb2020co]. One of the clearest examples of this parallel activation was provided by @costa2000cognate. In this study, Catalan-Spanish bilingual adults were asked to name pictures of common objects in Catalan, their dominant language. In half of the trials, the names of the objects were cognates (form-similar translation equivalents) in Catalan and Spanish (*sofà*-*sofá*, translations of *sofa*), whereas in the other half of the trials the names were non-cognates (*taula*-*mesa*, translations of *table*). Participants named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. Importantly, monolinguals did not show this effect. The authors suggested that, in cognate trials, the phonology of the Spanish word-forms was activated and facilitated the production of the Catalan word-form. More recently, several studies have provided evidence of similar cross-language activation in children [@von2012language; @poulin2013lexical; @jardak2019labels; @bosma2020cognate]. @floccia2018introduction pointed to bilingual toddlers' vocabulary size growing faster when both languages share a high proportion of cognate TEs, hypothesising that this leads to increased parallel activation of both words forms during speech processing. But do toddlers really know enough TEs for parallel activation to be a plausible mechanism for this effect?

Despite initial claims that bilinguals do not acquire TEs during their first years of life [@volterra1978acquisition], later studies reported that not only TEs are present at early stages of lexical development, but also represent an important proportion of the early bilingual [@pearson1993lexical; @de2014bilingual; @holowka2002semantic; @junker2002expressive; @legacy2016vocabulary; @bilson2015semantic] or even trilingual [@montanari2010translation] lexicon. If parallel activation plays a role during word acquisition at these ages, it could be predicted that cognate TEs are acquired earlier than non-cognate TEs, on the assumption that their acquisition is facilitated by their co-activation during speech processing. In line with this prediction, @bosch2014first found that cognates represented a larger proportion of the productive vocabulary of bilinguals than it would be expected from data in monolingual children [see @fabian2016investigating for similar results in Portuguese-English bilinguals]. Converging evidence was also provided by @schelletter2002effect, who reported the longitudinal single case of an English-German bilingual who learnt form-similar TEs earlier than form-dissimilar TEs. Interestingly, @gampe2018bilex reported experimental data in the same direction, by using a vocabulary assessment tool in which bilingual children learning Swiss German and an additional language, were asked to select a named picture from a set of eight pictures where the remaining seven were distractors. On average, children responded more accurately to cognate items than to non-cognate trials. In summary, there is evidence that cognate TEs may be acquired earlier.

The most common notion of parallel activation in the literature, however, does not quite satisfy the explanation needed for why TEs are acquired earlier when their forms are similar across languages. Parallel activation requires the existence of two already acquired representations, one in each language, that share meaning and are activated simultaneously during speech production and comprehension. This is not the case during the acquisition of TEs. First, if no members of the TE have been acquired yet, it would be impossible that form-similarity plays any role during the acquisition of any, as no representations are available for activation during speech processing. Even if one of the members of the TE has been acquired, it is not possible to activate a second member that has not been acquired! Finally, if both members have been acquired, parallel activation is possible, but it cannot account for the acquisition of the TE, which has already been completed. If parallel activation is to play a role during the acquisition of TEs, it must be assumed that the acquisition of a word form is gradual, in line with accumulator models accounts of vocabulary acquisition [e.g., @mcmurray2007defusing; @hidaka2013computational; @mollica2017data; @kachergis2021toward], and does not require a mature representation in the lexicon. This way, once one of the members of the TE has been acquired, the exposure to the acoustic form of the other member can activate the former, and their form similarity can facilitation the acquisition of the latter.

Within each TE, members that belong to the language of higher exposure are more likely to be acquired first, thus to facilitate the later acquisition of the member in the other language. This means that, if parallel activation plays a role in the cognate facilitation effect during bilingual word acquisition, this effect must be stronger in the language of lower exposure than in the language of higher exposure. Supporting this claim, the effect of cognateness reported by Floccia et al. only reached statistical significance in the additional language, not in English, the dominant language of participants. However, not this nor the other aforementioned studies tested  the effect of cognateness on the acquisition of TEs, but rather on the overall vocabulary size of participants. From this data it is not possible to draw any conclusions about the potential role of language dominance on the cognate facilitation effect. Another reason for not drawing conclusions from @floccia2018introduction et al.'s results about how cognateness of language dominance impact the age of acquisition of TEs is that their sample consisted on 24 month-old participants only, therefore overlooking the temporal dimension of word acquisition we are interested in.

In the present study we build onto previous studies of word acquisition to extend @floccia2018introduction's findings. We collected data from a sample of children of ages ranging from 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language. We hypothesised that TEs would be acquired earlier if both the Catalan and the Spanish word-forms were cognates, compared to when they are non-cognates. We also predicted that this effect would be larger for participants whose degree of exposure to the language a given word-form belongs to is lower, compard to when their exposure to that language is higer. The reaons behind this secodn prediction is that, for the acquistion of a word-form in one language to be facilitated by the phonological activation of its TE  in the other language, such TEs must have already been acquired. Given that words are more likely to be acquired in the language of higher exposure first, and in the language of lower exposure second, the facilitation effect should, on average, take place during the acquisition of the word-form in the language of lower exposure.

## References

### 2021-06-17 Rmd/introduction.Rmd
Add Nuria's comments (v2)

@@
   word_document:
-        reference_docx: "trajectories_template.docx"
-bibliography: "trajectories.bib"
-csl: "apa6.csl"
+    reference_docx: trajectories_template.docx
+  pdf_document:
+      latex_engine: xelatex
+bibliography: trajectories.bib
+csl: apa6.csl
 ---
@@
 Before the end of their first year of life, infants are able to encode sound patterns [@jusczyk1995infants; @tincoff1999some] and associate them with their referents [@bergelson2012months].
-During the second year, infants represent those word-forms with finer phonological detail [e.g.
-@werker2002infants], and from 18 to 24 months of age, they acquire new words at an increasingly faster rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension].
-What words are acquired earlier at these ages provides rich information about the mechanisms and biases that constrain early vocabulary growth.
-For instance, @braginsky2019consistency modelled vocabulary data from a large sample of children learning a diverse pool of languages and found that, across languages, words were acquired earlier if they were nouns, were highly concrete, or appeared in short sentences on average.
-Despite this consistency in their trajectories of acquisition, children display large variability in the composition of their vocabulary, in part due to the diversity of their language input.
+Later on, infants begin to represent those word-forms with finer phonological detail [e.g.
+@werker2002infants], and during the last half of their second year, they acquire new words at an increasingly fast rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension].
+Despite the considerable variation in children's trajectories of vocabulary growth, such trajectories seem quite consistent across languages [@braginsky2019consistency].
+Nonetheless, most of the literature on early word acquisition has been conducted on monolingual children, and does not address the problem of how bilinguals--who represent a substantial proportion of the populatio in most societies--acquire words at early ages.
 
-One of the causes of sources of this variability is the fact that a substantial proportion of children around the world learn to languages from birth.
-These children--to which we will refer as *bilinguals*--are a particular case of language learners: their linguistic input is divided into two languages.
-Taking into account that the relative amount of linguistic input the child receives in each language is a strong predictor of their vocabulary size [e.g., @cattani2014much; @thordardottir2006bilingual; @david2008individual; @pearson1997relation; @floccia2018introduction], one could predict that bilinguals know less words in each of their languages than their monolinguals peers.
-This is the case for some bilingual populations.
-For instance, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals that only learnt English.
-Other studies have provided converging evidence that bilinguals and monolinguals know a similar [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @junker2002expressive; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating] or even larger [@de2014bilingual; @aguila2008media] amount of words, only when summing up both languages together.
+There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
+For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals (who only learnt English).
+In contrast, both groups knew a similar total amount of words, when both English and Spanish vocabularies were counted together.
+Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating].
+While these studies have mostly relied on samples of bilingual children learning two relatively distant languages (as it is the case of English and Spanish) it is unclear whether children learning two more similar languages also know less words in each of their languages than monolinguals.
+What role could linguistic distance play during early vocabulary growth?
 
-A recent study by @floccia2018introduction suggests that the size of the difference between bilinguals' and monolinguals' trajectory of vocabulary growth varies considerably, conditional to the similarity between the two languages bilinguals are learning.
-The authors collected vocabulary data from 24 month-old bilingual toddlers in the United Kingdom, who were learning English and an additional language.
-The additional language was one a pool of 13 diverse languages that differed in their phonological and grammatical similarity with English (e.g., Dutch, Italian, Mandarin).
-Among their many findings, the authors reported a positive association between language similarity and vocabulary size.
-Particularly, the average phonological overlap between translation equivalents (TEs) in both languages (i.e., cognateness) predicted larger productive vocabulary sizes.
-For instance, children learning English and Dutch, sharing 22.14% phonological overlap, were able to produce more words than children learning English and Mandarin, only sharing 1.97% phonological overlap.
-These results put Hoff et al.
-and others' results--obtained in participants relatively distant languages--in a scenario where the differences in vocabulary size between monolinguals and bilinguals are difficult to generalise to other bilingual populations whose languages are linguistically closer.
-In line with these findings, @aguila2008media reported that bilingual toddlers learning Catalan and Spanish--two lexically very close languages--showed larger total vocabulary sizes than monolinguals, following a different pattern than bilinguals learning English and Spanish in @hoff2012dual.
+For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language.
+Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
+Particularly, when both languages are linguistically close, like Spanish and Catalan (both Roman languages), they are more likely to share a larger amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English--one Roman, the other Germanic.
+For instance, in the presence of a door, a Spanish-Catalan bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
+It could be the case that mapping two phonologically similar labels (cognates, like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
+If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
+
+A recent study by @floccia2018introduction provided evidence in line with this claim.
+The authors collected vocabulary data from 372 24-month-old bilingual toddlers in the United Kingdom who were learning English and an additional language.
+The additional language was one a pool of 13 languages that differed in their phonological similarity with respect to English, such as Dutch, Italian, or Mandarin.
+The authors defined phonological similarity as the average edit distance (length-adjusted Levenshtein distance) between translation equivalents in each language pair.
+The authors reported an increase in productive vocabulary size in the additional language associated with an increase in phonological overlap between the language pair.
+For example, English-Dutch bilinguals (22.14% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% phonological similarity) were able to produce in Mandarin.
 
 Floccia et al.
-pointed to parallel activation as the main mechanism behind their results.
-*Parallel activation* refers to the hypothesis that bilinguals activate both languages simultaneously during speech production or comprehension.
-This phenomenon is understood as the result of participants activating lexical representations in both languages, even when only using one of them [@thierry2007brain; @hoshino2008cognate; @bobb2020co].
-One of the clearest examples of this parallel activation was provided by @costa2000cognate.
-In this study, Catalan-Spanish bilingual adults were asked to name pictures of common objects in Catalan, their dominant language.
-In half of the trials, the names of the objects were cognates (form-similar translation equivalents) in Catalan and Spanish (*sofà*-*sofá*, translations of *sofa*), whereas in the other half of the trials the names were non-cognates (*taula*-*mesa*, translations of *table*).
-Participants named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
+pointed to *parallel activation* as the main mechanism underpinning their results.
+The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension.
+This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@thierry2007brain].
+One of the clearest examples of parallel activation was provided by @costa2000cognate.
+In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Catalan (their dominant language).
+In half of the trials, the object labels were cognates in Catalan and Spanish (*sofà*-*sofá*, translations of *sofa*), whereas in the other half of the trials labels were non-cognates (*taula*-*mesa*, translations of *table*).
+Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
 Importantly, monolinguals did not show this effect.
-The authors suggested that, in cognate trials, the phonology of the Spanish word-forms was activated and facilitated the production of the Catalan word-form.
-More recently, several studies have provided evidence of similar cross-language activation in children [@von2012language; @poulin2013lexical; @jardak2019labels; @bosma2020cognate].
-@floccia2018introduction pointed to bilingual toddlers' vocabulary size growing faster when both languages share a high proportion of cognate TEs, hypothesising that this leads to increased parallel activation of both words forms during speech processing.
-But do toddlers really know enough TEs for parallel activation to be a plausible mechanism for this effect?
+These results suggest that, for the bilinguals, Spanish phonology was activated during the production of Catalan words, facilitating the naming of cognate pictures.
+Several recent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
+In summary, the parallel activation account of bilingual word acquisition suggests that bilinguals' vocabulary size grows faster when both languages share a high proportion of cognate translation equivalents, as a result of increased activation of word forms in both languages during speech processing.
 
-Despite initial claims that bilinguals do not acquire TEs during their first years of life [@volterra1978acquisition], later studies reported that not only TEs are present at early stages of lexical development, but also represent an important proportion of the early bilingual [@pearson1993lexical; @de2014bilingual; @holowka2002semantic; @junker2002expressive; @legacy2016vocabulary; @bilson2015semantic] or even trilingual [@montanari2010translation] lexicon.
-If parallel activation plays a role during word acquisition at these ages, it could be predicted that cognate TEs are acquired earlier than non-cognate TEs, on the assumption that their acquisition is facilitated by their co-activation during speech processing.
-In line with this prediction, @bosch2014first found that cognates represented a larger proportion of the productive vocabulary of bilinguals than it would be expected from data in monolingual children [see @fabian2016investigating for similar results in Portuguese-English bilinguals].
-Converging evidence was also provided by @schelletter2002effect, who reported the longitudinal single case of an English-German bilingual who learnt form-similar TEs earlier than form-dissimilar TEs.
-Interestingly, @gampe2018bilex reported experimental data in the same direction, by using a vocabulary assessment tool in which bilingual children learning Swiss German and an additional language, were asked to select a named picture from a set of eight pictures where the remaining seven were distractors.
-On average, children responded more accurately to cognate items than to non-cognate trials.
-In summary, there is evidence that cognate TEs may be acquired earlier.
+If parallel activation plays a role during early word acquisition, it could be predicted that the age of acquisition of cognate translation equivalents is, on average, lower than that of non-cognates.
+However, the evidence supporting this claim is scarce.
+On the one hand, @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates, and @schelletter2002effect reported a longitudinal single case of an English-German bilingual who produced cognates earlier than non-cognates, on average.
+The low sample size in these studies make it challenging to draw strong conclusions about the effect of cognateness in vocabulary growth.
+On the other hand Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
+We identify two main reasons why we cannot conclude an earlier age of acquisition for cognates than for non cognate from Floccia et al.'s results.
+First, the response variable used was the proportion of words each participant understood and/or produced, from the list of lexical items in the vocabulary checklists.
+By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.
+non-cognates was lost.
+Second, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at this age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
+In summary, the evidence supporting an earlier age of acquisition of cognates vs.
+non-cognates presents some limitations that prevents drawing sound conclusions about this effect.
 
-The most common notion of parallel activation in the literature, however, does not quite satisfy the explanation needed for why TEs are acquired earlier when their forms are similar across languages.
-Parallel activation requires the existence of two already acquired representations, one in each language, that share meaning and are activated simultaneously during speech production and comprehension.
-This is not the case during the acquisition of TEs.
-First, if no members of the TE have been acquired yet, it would be impossible that form-similarity plays any role during the acquisition of any, as no representations are available for activation during speech processing.
-Even if one of the members of the TE has been acquired, it is not possible to activate a second member that has not been acquired!
-Finally, if both members have been acquired, parallel activation is possible, but it cannot account for the acquisition of the TE, which has already been completed.
-If parallel activation is to play a role during the acquisition of TEs, it must be assumed that the acquisition of a word form is gradual, in line with accumulator models accounts of vocabulary acquisition [e.g., @mcmurray2007defusing; @hidaka2013computational; @mollica2017data; @kachergis2021toward], and does not require a mature representation in the lexicon.
-This way, once one of the members of the TE has been acquired, the exposure to the acoustic form of the other member can activate the former, and their form similarity can facilitation the acquisition of the latter.
+Another worth-mentioning finding in Floccia et al.
+is that the increase in vocabulary size associated with lingusitic similarity was larger in the additional language vocabulary than in English vocabulary.
+Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
+Given that the relative amount of exposure to a given language is a strong predictor of children's vocabulary size in that language [e.g., @cattani2014much; @thordardottir2006bilingual], participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
+If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
 
-Within each TE, members that belong to the language of higher exposure are more likely to be acquired first, thus to facilitate the later acquisition of the member in the other language.
-This means that, if parallel activation plays a role in the cognate facilitation effect during bilingual word acquisition, this effect must be stronger in the language of lower exposure than in the language of higher exposure.
-Supporting this claim, the effect of cognateness reported by Floccia et al.
-only reached statistical significance in the additional language, not in English, the dominant language of participants.
-However, not this nor the other aforementioned studies tested  the effect of cognateness on the acquisition of TEs, but rather on the overall vocabulary size of participants.
-From this data it is not possible to draw any conclusions about the potential role of language dominance on the cognate facilitation effect.
-Another reason for not drawing conclusions from @floccia2018introduction et al.'s results about how cognateness of language dominance impact the age of acquisition of TEs is that their sample consisted on 24 month-old participants only, therefore overlooking the temporal dimension of word acquisition we are interested in.
-
-In the present study we build onto previous studies of word acquisition to extend @floccia2018introduction's findings.
-We collected data from a sample of children of ages ranging from 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We hypothesised that TEs would be acquired earlier if both the Catalan and the Spanish word-forms were cognates, compared to when they are non-cognates.
-We also predicted that this effect would be larger for participants whose degree of exposure to the language a given word-form belongs to is lower, compard to when their exposure to that language is higer.
-The reaons behind this secodn prediction is that, for the acquistion of a word-form in one language to be facilitated by the phonological activation of its TE  in the other language, such TEs must have already been acquired.
-Given that words are more likely to be acquired in the language of higher exposure first, and in the language of lower exposure second, the facilitation effect should, on average, take place during the acquisition of the word-form in the language of lower exposure.
+The aim of this study is to investigate the role of cognateness in bilingual word acquisition.
+We hypothesised that cognate words would be acquired earlier than non-cognate word, and that this difference would be larger in participants with lower exposure to the language such word belongs to.
+Using an *ad hoc* online vocabulary checklist, we collected data from a sample of bilinguals aged 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure tto the corresponding language, while adjusting for participants' age and the item's lexical frequency.
 

### 2021-06-17 Rmd/introduction.Rmd

@@
     word_document:
-    reference_docx: trajectories_template.docx
-  pdf_document:
-      latex_engine: xelatex
-bibliography: trajectories.bib
-csl: apa6.csl
+        reference_docx: "trajectories_template.docx"
+bibliography: "trajectories.bib"
+csl: "apa6.csl"
 ---
@@
 Later on, infants begin to represent those word-forms with finer phonological detail [e.g.
-@werker2002infants], and during the last half of their second year, they acquire new words at an increasingly fast rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension].
-Despite the considerable variation in children's trajectories of vocabulary growth, such trajectories seem quite consistent across languages [@braginsky2019consistency].
-Nonetheless, most of the literature on early word acquisition has been conducted on monolingual children, and does not address the problem of how bilinguals--who represent a substantial proportion of the populatio in most societies--acquire words at early ages.
+@werker2002infants], and from 18 to 24 months of age, they acquire new words at an increasingly fast rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension].
+The type of words acquired earlier at these ages provides  information about the mechanisms and biases that constrain early vocabulary growth.
+Recent works show that the words that are acquired  are nouns, highly concrete words, or words that appear in short sentences, an that this pattern is relatively consistent across languages [@braginsky2019consistency].
+Most of the literature on early word acquisition has been conducted on monolingual children.
+However, a substantial proportion of children around the world are exposed to two languages from birth (bilinguals, henceforth).
+Given that the relative amount of input a child receives in each language is a strong predictor of vocabulary size [e.g., @cattani2014much; @thordardottir2006bilingual; @david2008individual; @floccia2018introduction], generalising the aforementioned developmental patterns of word acquisition to bilinguals is challenging.
 
-There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
-For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals (who only learnt English).
-In contrast, both groups knew a similar total amount of words, when both English and Spanish vocabularies were counted together.
-Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating].
-While these studies have mostly relied on samples of bilingual children learning two relatively distant languages (as it is the case of English and Spanish) it is unclear whether children learning two more similar languages also know less words in each of their languages than monolinguals.
-What role could linguistic distance play during early vocabulary growth?
+There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when both languages are aggregated.
+For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals that only learnt English.
+In contrast, both groups knew a similar amount of total words, when words in both English and Spanish were counted together.
+Other studies have provided converging evidence that bilinguals and monolinguals know a similar number of words [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @junker2002expressive; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating] or even a larger amount of words [@de2014bilingual; @aguila2008media], only when the languages were aggregated.
+While these studies have mostly relied on samples of bilingual children learning two relatively distant languages (as it is the case of English and Spanish) it is unclear whether children learning two more similar languages follow this pattern too.
 
-For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language.
-Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
-Particularly, when both languages are linguistically close, like Spanish and Catalan (both Roman languages), they are more likely to share a larger amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English--one Roman, the other Germanic.
-For instance, in the presence of a door, a Spanish-Catalan bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
-It could be the case that mapping two phonologically similar labels (cognates, like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
-If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
-
-A recent study by @floccia2018introduction provided evidence in line with this claim.
-The authors collected vocabulary data from 372 24-month-old bilingual toddlers in the United Kingdom who were learning English and an additional language.
-The additional language was one a pool of 13 languages that differed in their phonological similarity with respect to English, such as Dutch, Italian, or Mandarin.
-The authors defined phonological similarity as the average edit distance (length-adjusted Levenshtein distance) between translation equivalents in each language pair.
-The authors reported an increase in productive vocabulary size in the additional language associated with an increase in phonological overlap between the language pair.
-For example, English-Dutch bilinguals (22.14% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% phonological similarity) were able to produce in Mandarin.
+A recent study by @floccia2018introduction highlighted language similarity as an important factor underlying the differences between monolingual and bilingual vocabulary sizes.
+The authors collected vocabulary data from 372 24-month-old bilingual toddlers in the UK who were learning English and an additional language.
+The additional language was one a pool of 13 languages that differed in their phonological and grammatical similarity with respect to English, such as Dutch, Italian, or Mandarin.
+To measure phonological similarity, the authors computed the average edit distance between translation equivalents in each language pair.
+Translation equivalents (TEs) are word-forms that belong to two different languages, but share meaning (or at least enough conceptual overlap to be considered synonyms).
+TEs that also overlap in form (e.g., *cat*-*gato*, in English and Spanish) are called *cognates*, whereas form-dissimilar TEs (e.g., *dog*-*perro*) are called *non-cognates*.
+In this study, the more cognates a pair of language shared, the higher their index of phonological similarity was.
+The authors reported a positive association between the index of phonological similarly and participants' vocabulary sizes.
+For example, children learning English and Dutch, sharing 22.14% phonological overlap, were able to produce more words than children learning English and Mandarin, only sharing 1.97% phonological overlap.
+Hoff et al.
+results, combined with other findings that show differences in vocabulary size across monolingual and bilingual children learning relatively distant languages are difficult to generalise to other bilingual populations whose languages are closer.
+For instance, @aguila2008media reported that bilingual toddlers learning Catalan and Spanish, two lexically very close languages, showed *larger* aggregated vocabulary sizes than monolinguals, following a different pattern than bilinguals learning English and Spanish in @hoff2012dual.
+In summary, bilinguals learning two languages that share more cognates show larger vocabulary sizes than those learning two languages that share less cognates.
+However, how cognateness impacts vocabulary acquisition is still unclear.
 
 Floccia et al.
-pointed to *parallel activation* as the main mechanism underpinning their results.
-The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension.
-This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@thierry2007brain].
-One of the clearest examples of parallel activation was provided by @costa2000cognate.
-In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Catalan (their dominant language).
-In half of the trials, the object labels were cognates in Catalan and Spanish (*sofà*-*sofá*, translations of *sofa*), whereas in the other half of the trials labels were non-cognates (*taula*-*mesa*, translations of *table*).
-Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
+pointed to parallel activation as the main mechanism underpinning their results.
+The *parallel activation* hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension.
+This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use [@thierry2007brain; @hoshino2008cognate; @bobb2020co].
+One of the clearest examples of parallel activation was provided by @costa2000cognate's study, in which Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Catalan, their dominant language.
+In half of the trials, the object labels were cognates in Catalan and Spanish (*sofà*-*sofá*, translations of *sofa*)--whereas in the other half of the trials labels were non-cognates (*taula*-*mesa*, translations of *table*).
+Bilingual named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
 Importantly, monolinguals did not show this effect.
-These results suggest that, for the bilinguals, Spanish phonology was activated during the production of Catalan words, facilitating the naming of cognate pictures.
-Several recent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
-In summary, the parallel activation account of bilingual word acquisition suggests that bilinguals' vocabulary size grows faster when both languages share a high proportion of cognate translation equivalents, as a result of increased activation of word forms in both languages during speech processing.
+These results suggested that, for the bilinguals, Spanish phonology was activated during cognate trials, facilitating the production of the Catalan word-forms.
+Several recent studies have also provided similar evidence in children [@von2012language; @poulin2013lexical; @jardak2019labels; @bosma2020cognate].
+These results, taken together with Floccia et al.'s, seem to point to the fact that bilinguals' vocabulary size tends to grow faster when both languages share a high proportion of cognate TEs, leading to a increased activation of word forms in both languages during speech processing.
+An interesting question is whether toddlers really know enough TEs for parallel activation to be a plausible mechanism for such effect.
 
-If parallel activation plays a role during early word acquisition, it could be predicted that the age of acquisition of cognate translation equivalents is, on average, lower than that of non-cognates.
-However, the evidence supporting this claim is scarce.
-On the one hand, @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates, and @schelletter2002effect reported a longitudinal single case of an English-German bilingual who produced cognates earlier than non-cognates, on average.
-The low sample size in these studies make it challenging to draw strong conclusions about the effect of cognateness in vocabulary growth.
-On the other hand Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
-We identify two main reasons why we cannot conclude an earlier age of acquisition for cognates than for non cognate from Floccia et al.'s results.
-First, the response variable used was the proportion of words each participant understood and/or produced, from the list of lexical items in the vocabulary checklists.
-By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.
-non-cognates was lost.
-Second, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at this age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
-In summary, the evidence supporting an earlier age of acquisition of cognates vs.
-non-cognates presents some limitations that prevents drawing sound conclusions about this effect.
+Despite initial claims that bilinguals do not acquire TEs during their first years of life [@volterra1978acquisition], later studies reported that not only TEs are present at early stages of lexical development, but also represent an important proportion of the early bilingual [@pearson1993lexical; @de2014bilingual; @holowka2002semantic; @junker2002expressive; @bilson2015semantic] or even trilingual [@montanari2010translation] lexicon.
+If parallel activation plays a role during word acquisition at these ages, it could be predicted that cognate TEs are acquired earlier than non-cognate TEs, based on the assumption that the acquisition of cognates is facilitated by the co-activation of their translation equivalent during speech processing.
+Indeed, cognates seem to repressent a larger proportion of bilinguals' productive vocabulary than what expected from monolinguals' data [@bosch2014first; see @fabian2016investigating for similar results in Portuguese-English bilinguals].
+Additionally, three-year-old children learning German and an additional language responded more accurately to cognate items than to non-cognate trials, in a task that required to select a named picture from a pool of pictures and distractors [@gampe2018bilex].
+Converging evidence was also provided by a longitudinal case study of an English-German bilingual who learnt form-similar TEs earlier than form-dissimilar TEs [@schelletter2002effect].
+In summary, there is evidence that during their first three years of life, children know many TEs, and seem to acquire cognate TEs earlier than non-cognate TEs.
 
-Another worth-mentioning finding in Floccia et al.
-is that the increase in vocabulary size associated with lingusitic similarity was larger in the additional language vocabulary than in English vocabulary.
-Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
-Given that the relative amount of exposure to a given language is a strong predictor of children's vocabulary size in that language [e.g., @cattani2014much; @thordardottir2006bilingual], participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
-If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
+The most common notion of parallel activation in the literature, however, does not quite satisfy the explanation needed for why TEs are acquired earlier when their forms are similar across languages.
+Parallel activation requires the existence of two already acquired representations, one in each language, that are activated simultaneously during language production and comprehension.
+This is not the case in the acquisition of TEs.
+First, if no members of the TE have been acquired yet, it would be impossible that form-similarity plays a role during the acquisition of TEs (as no representations are available for activation during speech processing).
+Second, even if one of the members of the TE has been acquired, it is not possible to activate a second member that has not yet been acquired.
+Third, if both members have been acquired, parallel activation is possible, but cannot account for the acquisition of the TE, which has already occurred.
+If parallel activation plays a role during the acquisition of TEs, it must be assumed that word-form acquisition is a gradual process that does not require mature lexical representations, in line with accumulator models accounts of vocabulary acquisition [e.g., @mcmurray2007defusing; @hidaka2013computational; @mollica2017data; @kachergis2021toward].
+This way, once one of another member of the TE has been acquired, the exposure to the acoustic form of another member can activate the former, and their form similarity would facilitate the acquisition of the latter.
 
-The aim of this study is to investigate the role of cognateness in bilingual word acquisition.
-We hypothesised that cognate words would be acquired earlier than non-cognate word, and that this difference would be larger in participants with lower exposure to the language such word belongs to.
-Using an *ad hoc* online vocabulary checklist, we collected data from a sample of bilinguals aged 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure tto the corresponding language, while adjusting for participants' age and the item's lexical frequency.
+Moreover, within each TE, members that belong to the language of higher exposure are more likely to be acquired first.
+This would facilitate the subsequent acquisition of the member in the other language (the language of lower exposure).
+If parallel activation plays a role in the facilitation effect of cognateness during word acquisition, such effect should be stronger in the language to which children are less exposed.
+In support to this claim, the effect of cognateness reported by Floccia et al.
+only reached statistical significance in the additional language, not in English, participant's dominant language.
+However, not this nor the other aforementioned studies tested  the effect of cognateness on the acquisition of TEs, rather on participants' aggregated vocabulary size.
+Thus, from this data it is not possible to draw any conclusions about the potential role of language dominance on the cognate facilitation effect.
+
+In the present study we build onto previous studies of word acquisition to extend @floccia2018introduction's findings using data from children at earlier and later ages, and modelling the acquisition trajectories of cognate and non-cognate TEs.
+We collected data from a sample of children of ages ranging from 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+We hypothesised that TEs would be acquired earlier if both the Catalan and the Spanish word-forms were cognates, compared to when they are non-cognates.
+We also predicted that this effect would be larger in participants whose exposure to the language a given word-form belongs to is lower, compared to when exposure to that language is higher.
+The reason behind this prediction is that TEs should have already been acquired in one of the languages in order to generate a facilitation effect.
+Since word-forms are more likely to be acquired in the language of higher exposure, the potential facilitation effect of cognatness should occur more frequently, and thus be stronger, in the acquisition of the word-form in the language of lower exposure.
 

### 2021-06-17 Rmd/introduction.Rmd
Restore master branch

--- 2021-06-17 Rmd/introduction.Rmd
@@
 Using an *ad hoc* online vocabulary checklist, we collected data from a sample of bilinguals aged 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure tto the corresponding language, while adjusting for participants' age and the item's lexical frequency.
+We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.
 

### 2021-06-17 R/analysis.R

@@ -7,217 +7,288 @@ library(tidyverse)
 library(brms)
 library(scales)
 library(tidybayes)
-library(emmeans)
-library(here)
 
 # load helper functions
-source(here("R", "utils.R"))
-options(mc.cores = 7, chains = 7, iter = 1000)
+source("R/utils.R")
 set.seed(888)
 
 #### import data ---------------------------------------------------------------
-responses <- readRDS(here("Data", "responses.rds")) %>% 
-  filter(between(age, 12, 36)) %>% 
-  mutate(
-    doe = doe*10,
-    age_scaled = age,
-    frequency_scaled = frequency,
-    doe_scaled = doe
-  ) %>% 
-  mutate_at(vars(frequency), function(x) scale(x)[,1]) %>% 
-  mutate_at(vars(age, doe), function(x) scale(x, scale = FALSE)[,1])
-
-
-# responses_subset <- filter(responses, te %in% sample(responses$te, 25)) %>% 
-#   mutate_at(vars(age, frequency, doe), function(x) scale(x, scale = FALSE)[,1]) %>% 
-#   mutate_at(vars(item_dominance, cognate), as.factor) %>% 
-#   mutate(response = factor(response, ordered = TRUE))
+proportions <- readRDS("Data/proportions.rds")
 
 # set sum contrasts (Schad et al., 2018, https://arxiv.org/abs/1807.10451)
-contrasts(responses$cognate) <- contr.sum(2)/2
+contrasts(proportions$cognate) <- contr.sum(2)/2
 
-#### model fitting -------------------------------------------------------------
-
-fit_0 <- brm(
-  response ~ age + frequency,
-  family = cratio(link = "logit"),
-  data = responses,
-  prior = prior(normal(0, 5), class = b),
-  chains = 4, iter = 1000,
-  file = here("Results", "fit_0.rds"),
-  save_model = here("Stan", "fit_0.stan"),
-  backend = "cmdstanr"
+## model fitting -------------------------------------------------------------
+formula_understands <- bf(
+  understands ~ inv_logit(asymlogit) * inv(1 + exp((mid - age_bin) * scale)),
+  mid ~ 1 + frequency + doe_bin*cognate + (1 | te),
+  scale ~ 1,
+  asymlogit ~ 1,
+  phi ~ 1,
+  nl = TRUE,
+  family = Beta(link = "identity")
 )
 
-fit_1 <- update(
-  fit_0, . ~ . + doe,
[…]

### 2021-06-17 R/preprocessing.R

@@ -1,29 +1,32 @@
-#### preprocessing -------------------------------------------------------------
-
-# set up -----------------------------------------------------------------------
+# set up -----
 
 # load packages
 library(multilex)
 library(janitor)
 library(tidyverse)
-library(here)
 
 # load helper functions
-source(here("R", "utils.R"))
+source("R/utils.R")
 
-# import data ------------------------------------------------------------------
+# import data ----
 ml_connect("gonzalo.garciadecastro@upf.edu")
 p <- ml_participants()
-r <- ml_responses(p, longitudinal = "first", update = FALSE)  
+r <- ml_responses(p, longitudinal = "first", update = FALSE) %>% 
+    mutate(postcode = as.integer(postcode)) %>% 
+    bind_rows(multilex:::devlex, multilex:::cbc, multilex:::formr1, multilex:::formr_short)
 l <- ml_logs(p, r)
 v <- ml_vocabulary(p, r, by = c("dominance", "language"), scale = "prop")
 
-# items ------------------------------------------------------------------------
+age_bin_labels <- c("10-12", "12-14", "14-16", "16-18", "18-20", "20-22",
+                    "22-24", "24-26", "26-28", "28-30", "30-32", "32-34", "34-36")
+doe_bin_labels <- paste0("DOE: ", c("0-25%", "25-50%", "50%-75%", "75-100%"))
+
+# items ----
 items <- multilex::pool %>%
     drop_na(cognate, ipa) %>% 
     filter(
         include,
-        class %in% c("Noun", "Verb", "Adjective")
+        class %in% c("Noun")
     ) %>% 
     rename(frequency = frequency_zipf) %>% 
     select(te, language, category, item, ipa, frequency, cognate) %>% 
@@ -31,50 +34,74 @@ items <- multilex::pool %>%
         language = str_to_sentence(language),
         cognate = ifelse(cognate, "Cognate", "Non-Cognate")
     ) %>% 
-    relocate(te, language, item, ipa, cognate, frequency)
+    relocate(te, language, item, cognate, frequency)
 
-# participants -----------------------------------------------------------------
+# participants ----
 participants <- l %>%
     rename(dominant_language = dominance) %>% 
     filter(
         completed,
-        !(version %in% c("DevLex", "CBC")),
         lp %in% c("Monolingual", "Bilingual"),
-        between(age, 12, 36),
[…]

### 2021-06-17 R/irt.R
Restore master branch

@@ -1,0 +1,141 @@
+# IRT
+
+# set up ----
+library(tidyverse)
+library(multilex)
+library(brms)
+library(tidybayes)
+
+# import data ----
+ml_connect("gonzalo.garciadecastro@upf.edu")
+p <- ml_participants()
+r <- ml_responses(p, longitudinal = "first", update = FALSE)
+l <- ml_logs(p, r)
+
+# items ----
+items <- multilex::pool %>%
+    drop_na(cognate, ipa) %>% 
+    filter(
+        include,
+        class %in% c("Noun")
+    ) %>% 
+    rename(frequency = frequency_zipf) %>% 
+    select(te, language, category, item, ipa, frequency, cognate) %>% 
+    mutate(
+        language = str_to_sentence(language),
+        cognate = ifelse(cognate, "Cognate", "Non-Cognate")
+    ) %>% 
+    relocate(te, language, item, cognate, frequency)
+
+# participants ----
+participants <- l %>%
+    rename(dominant_language = dominance) %>% 
+    filter(
+        completed,
+        lp %in% c("Monolingual", "Bilingual"),
+        between(age, 10, 36),
+        sum(doe_catalan+doe_spanish)>0.1
+    ) %>% 
+    pivot_longer(c(doe_catalan, doe_spanish), names_to = "language", values_to = "doe") %>% 
+    mutate_at(vars(matches("doe")), function(x) x*0.01) %>% 
+    filter(between(doe, 0, 1)) %>% 
+    mutate(
+        language = str_to_sentence(str_remove(language, "doe_"))
+    ) %>% 
+    select(id, time_stamp, age, lp, dominant_language, doe)
+
+# responses ----
+d <- expand_grid(
+    id = participants$id,
+    item = items$item
+) %>%
+    left_join(participants) %>% 
+    left_join(items) %>% 
+    left_join(select(r, id, item, response))  %>% 
+    drop_na() %>%
+    distinct(id, te, age, item, .keep_all = TRUE) %>% 
+    mutate(
+        doe = scale(doe, scale = FALSE)[,1],
+        age = scale(age, scale = FALSE)[,1],
+        understands = response %in% c(2, 3),
+        produces = response %in% 3,
+        dominance = ifelse(dominant_language==language, "L1", "L2")
+    ) %>% 
+    mutate_at(vars(cognate), as.factor) %>%
[…]

### 2021-06-17 R/simulations.R

@@ -9,67 +9,4 @@ bayesplot_theme_set(theme_minimal())
 inv_logit <- function(x) 1 / (1 + exp(-x))
 
 # create data ------------------------------------------------------------------
-N_te <- 100
-N_participant <- 50
-age <- c(10, 36)
-dominance <- c("L1", "L2")
-cognate <- c("Cognate", "Non-cognate")
-
-d <- full_join(
-    full_join(
-        data.frame(te = 1:N_te, cognate = sample(cognate, size = N_te, prob = c(0.25, 0.75), replace = TRUE)),
-        expand.grid(te = 1:N_te, dominance = dominance
-        )
-    ),
-    full_join(
-        expand.grid(participant = 1:N_participant, te = 1:N_te),
-        data.frame(participant = 1:N_participant, age = runif(N_participant, age[1], age[2]))
-    )
-) %>% 
-    mutate_at(vars(dominance, cognate), as.factor) %>% 
-    relocate(participant, age, te, dominance, cognate) %>% 
-    arrange(participant, te) 
-
-contrasts(d$dominance) <- c(0.5, -0.5)
-contrasts(d$cognate) <- c(0.5, -0.5)
-
-x <- model.matrix(~age + dominance*cognate, d)
-
-b <- c(
-    alpha = -5,
-    beta_age = 0.2,
-    beta_dominance = 2,
-    beta_cognate = 1,
-    beta_dominance_cognate = 1
-)
-
-mu <- (x %*% b)[,1]
-y <- rbinom(n = length(mu), size = 1, prob = inv_logit(mu))
-d$y <- y
-
-# plot data --------------------------------------------------------------------
-ggplot(d, aes(age, y, colour = interaction(dominance, cognate, sep = " - "))) +
-    facet_wrap(~dominance) +
-    geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
-    geom_jitter(height = 0.05, shape = 1, alpha = 0.5) +
-    labs(x = "Age (months)", y = "Pr(Y|X)", colour = "Dominance") +
-    theme_bw() +
-    theme(legend.title = element_blank())
-
-# fit model --------------------------------------------------------------------
-X <- model.matrix(~age + dominance*cognate, d)
-d_stan <- list(
-    N = nrow(d),
-    J = length(unique(d$te)),
-    te = as.numeric(as.factor(d$te)),
-    X = X,
-    K = ncol(X),
[…]

### 2021-07-12 Rmd/00_outline.Rmd

---
title: "Outline"
author: "Gonzalo García-Castro"
date: "5/6/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Aim: 4,000 words

# Introduction (1,250 words)

## Monolingual vocabulary acquisition (300 words)

* Participant-level predictors: age, sex, SES, quantity and quality of input
* Word-level predictors: grammatical class, frequency, length, phonological neighbourhood density

## Bilingual vocabulary acquisition (800 words)

* How bilinguals differ from monolinguals
* How the specific pair of languages they learn impacts bilingual vocabulary acquisition: Floccia et al. (2018)
* We suggest that cognates are, on average, acquired earlier than non-cognates.
* We suggest that this facilitation effect is due to infants activating a familiar word form when hearing its form-similar translation equivalent in presence of their shared referent.
* This effect can only take place once one of the forms of the translation equivalent in one language has been acquired. When acquiring the other form of the translation equivalent in the other language, form-similar should facilitate the process. Since words in the dominant language are more likely to be acquired first, the effect of cognateness should be larger in the non-dominant language (i.e., language of least exposure) than in the dominant language (language of most exposure).

## The present study (150 words)

* We compared the acquisition trajectories of cognate and non-cognate translation equivalents using vovcabulary data from bilinguals learning Catalan and Spanish.
* We hypothesised that (1) cognates are acquired earlier than non-cognates, and (2) that this effect is larger for items that belong to the language of least exposure

# Methods (750 words)

## Participants (250 words)

* Brief description of the sample (age, sex, location, SES, etc.)
* Ethics
* Language profile
* Sample sizes

## Vocabulary inventory (250 words)

* What the questionnaire looks like (online, list of words, similar to CDI, etc)
* Description of the items included (frequency, grammatical class)
* Randomisation (short versions)

## Data analysis (250 words)

* Dependent variable
* Predictors
* Model

# Results (1,000 words)

## Comprehension (500 words)

## Production (500 words)

# Discussion (1,000 words)

# Acknowledgements

# References

# Appendix

## Appendix 1: Participant information

## Appendix 2: Questionnaire information

## Appendix 3: Model

### 2021-07-12 Rmd/01_introduction.Rmd

--- 2021-06-17 Rmd/introduction.Rmd
@@
 Another worth-mentioning finding in Floccia et al.
-is that the increase in vocabulary size associated with lingusitic similarity was larger in the additional language vocabulary than in English vocabulary.
+is that the increase in vocabulary size associated with linguistic similarity was larger in the additional language vocabulary than in English vocabulary.
 Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.

### 2021-07-12 Rmd/03_results.Rmd

--- 2021-01-21 Rmd/trajectories_manuscript.Rmd
@@
 ---
-title             : "The role of cross-linguistic lexical similarity on bilingual word acquisition"
-shorttitle        : "Cross-linguistic similarity and  word acquisition"
-
-author: 
-  - name          : "Gonzalo Garcia-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Daniela Avila-Varela"
-    affiliation   : "1"
-  - name          : "Nuria Sebastian-Galles"
-    affiliation   : "1"
-
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
-
-authornote: |
-  Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
-
-abstract: |
-  Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "lexical acquisition, vocabulary, bilingualism"
-wordcount         : "X"
-
-bibliography      : ["trajectories.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-keep_text         : false
-class             : "man"
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  papaja::apa6_pdf:
-    latex_engine  : "xelatex"
-
+title: "Results"
+output:
+  word_document:
+    reference_docx: trajectories_template.docx
+  pdf_document:
+      latex_engine: xelatex
+bibliography: trajectories.bib
+csl: apa6.csl
 ---
 
-```{r setup, include = FALSE}
-
-knitr::opts_chunk$set(
-  echo = FALSE,
-  fig.align = "center",
-  message = FALSE,
-  warning = FALSE,
-  cache.extra = knitr::rand_seed,
-  out.width = "80%",
-  results = "asis"
-)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
-
-```{r packages, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# load packages
-library(papaja)
-library(knitr)
-library(magrittr)
-library(tidyverse)
-library(readxl)
-library(lubridate)
-library(brms)
-library(tidybayes)
-library(janitor)
-library(gt)
-library(data.table)
-library(here)
-
-source(here("R", "analyses.R"))
-```
-
-# Introduction
-
-Learning words involves two key steps: (1) the encoding of a word form (e.g, /dɒɡ/)
-and (2) its association with a referent (e.g., a DOG).
-There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side [@bergelson_at_2012].
-At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered [@jusczyk_infants_1995], as opposed to when only part of the word is uttered.
-At 12 months, infants show sensitivity to both vowel and consonant mispronunciations [@mani_twelve_2010], looking shorter to named pictures when their label is mispronounced.
-Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones.
-One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon [@mcclelland_trace_1986].
-
-Words are not learnt in isolation, but in contexts rich in word tokens and referents.
-The structure of toddlers' lexicon reflects this high connectivity [@wojcik_development_2018; @hills_longitudinal_2009; @fourtassi_growth_2020].
-At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing [@mani_infant_2010; @mani_phonological_2011].
-The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming  [@styles_infants_2009; @arias-trejo_what_2013; @wojcik_ontogeny_2013], and show evidence of inhibitory links between lexical-semantic entries [@chow_backward_2016; chow_vocabulary_2019].
-By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages [@chow_spoken_2017].
-Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words [@chow_spoken_2017; @novogrodsky_age_2020], and  the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers.
-This suggests that the emergence of rich lexical network-- both at the phonological and the semantic levels--grows as a function of the number of words acquired by the toddler.
-
-The structure of the lexicon also impacts the order of acquisition of new words.
-Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills_longitudinal_2009].
-@borovsky_lexical_2016 and @borovsky_slowing_2020 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
-This points to the structure of the lexicon facilitating the strategies children engage during word learning.
-
-@babineau_familiar_2020
-
-The case of bilingual children (here defined as those learning two languages from birth) presents an opportunity to study how lexical acquisition takes placed under a more complex environment, as they  face the challenge of learning two distinct sets of words -- one for each language -- that partially overlap in sound and meaning.
-
-EARLY STAGES IN BILINGUAL LANGUAGE ACQUISITION
-
-When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@leopold_study_1948; @ben-zeev_influence_1977; @doyle_some_1977; @rosenblum_word_1983; @fernandez_bilingual_1992; @umbel_measuring_1992; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
-
-@floccia_i_2018, gathered parental vocabulary estimates [@hamilton_infant_2000] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
-They computed the average phonological similarity between translation equivalents (TEs from now on; e.g.,  *table* in English and *tafel* in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
-Using a similar measure of lexical similarity, @blom_cross-language_2019 extended these results to bilingual children aged three to 10, and reported a positive association between the lexical distance and comprehensive vocabulary size: bilinguals learning lexically close languages showed similar vocabulary sizes to those of monolinguals, but those learning the more distant languages showed lower vocabulary sizes.
-
-These results suggest that linguistic distance between languages plays an important role during bilingual lexical acquisition, and that this effect operates within TEs.
-What mechanisms underlie this effect, and why they seem to operate differently across comprehension and production, or across the dominant and the non-dominant language, remain open issues.
-@floccia_i_2018 pointed to parallel activation as a candidate mechanism behind this facilitation effect.
-The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., *cat*, in English) its corresponding translation in the other language is activated too [e.g., *gato*, in Spanish.
-There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell_mediated_1991; @spivey_cross_1999; @costa_cognate_2000; @costa_lexical_2000; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh_one_2014; @von-holzen_impact_2019; @bobb_co-activation_2020; see @costa_facilitatory_2005 for review], and children [e.g.
-@von-holzen_language_2012; @poulin_lexical_2013; @jardak_labels_2019; @bosma_cognate_2020].
-A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
-
-@floccia_i_2018 argue that the degree of phonological similarity between TEs (*cognateness* from now on) of both languages should lead to increased parallel activation of both languages during language exposure, facilitating the acquisition of words in both languages, and leading participants leading two lexically close languages to show larger vocabulary sizes than those learning two lexically distant languages.
-This account, however, neglects the fact that for parallel activation to take place and play a facilitation role, lexical representations of both pairs of the TE must have been already established in the lexicon, and their corresponding phonological forms must have been encoded.
-This is not the case during lexical acquisition, where at least one of the members of the pair has not being acquired yet.
-A more lenient definition of parallel activation could adjust to the context of lexical acquisition more easily: when only one of the members of the TE has been acquired, its lexical representation is activated when, in the presence of its referent, a phonologically similar label is uttered.
-Consider the case of a Catalan-Spanish bilingual that has already learned the word *gat* (Catalan for *cat*), but not *gato* (in Spanish).
-It is possible that, when presented with *gato* in the presence of a cat, she maps this novel label to the familiar word *gat*, and creates a lexical representation for *gato* as a synonym for *gat*.
-The case of a non-cognate is differing.
-Consider now the case of a toddler that has already learnt *gos* (Catalan for *dog*), but not *perro* (Spanish).
-In this case, it would not be possible to map both labels via phonology, give their lack of phonological similarity.
-
-Under this account, cognates (i.e., phonologically similar TEs) would be acquired earlier than non-cognates, but this effect would only play a role once one of the members of the TE has been acquired.
-The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse.
-@schelletter_effect_2002 reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
-@bosch_first_2014, showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months.
-The low number of words included in the analyses of both studies, and the lack of adjustment for lexical frequency, limits severely the strength of any conclusions concerning the difference in age of acquisition of cognates and non-cognates.
-On the other hand @floccia_i_2018 worked with aggregated estimates of participants' vocabulary size, losing item-level information that could provide information about change in age of acquisition of TEs associated with their degree of phonological similarity.
-Therefore, their results can only be considered as compatible with cognateness playing a role on age of acquisition.
-In this study, we aimed at overcoming these pitfalls by using unaggregated data from individual responses, testing the role of cognateness explicitly and adjusting for lexical frequency.
-
-It also follows from the fact that cognateness operates *within* TEs that the effect of cognateness can only play a role once one of the members of the TE has been acquired, and its phonological form is available during the acquisition of its translation.
-Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language [@floccia_i_2018].
-@dijkstra_role_2016 tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE.
-Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian.
-This suggests that infants that were mostly exposed to Dutch used the phonology of  dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
-A study by @dijkstra_role_2016 showed that the degree of phonological similarity between 
-the performance of  was better for cognates than for non-cognates, only for those children
-
-Finally, previous studies have reported that the properties that describe the form of a word play a stronger role in production than comprehension than in production.
-This is the case of the number of phonemes [@braginsky_consistency_2019], phonological neighbourhood density [@jones_children_2019].
-These results converge with @floccia_i_2018's finding that the effect of the average cognateness between languages was larger in production than in comprehension^[Although @bosch_first_2014 found that participants in their sample produced cognates earlier than non-cognates, no analyses were conducted on comprehensive data, and therefore they cannot be taken as supporting evidence for a larger effect of cognateness on production than in comprehension].
-Accordingly, we expected the role of cognateness to be more central in production than in comprehension.
-
-In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.
-
-@taverna_early_2020
-
-```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-participants <- read.csv(here("Data", "participants.csv"))
-items <- read.csv(here("Data", "items.csv"))
-vocabulary <- read.csv(here("Data", "vocabulary.csv"))
-responses <- read.csv(here("Data", "responses.csv"))
-
-```
-
-# Method
-
-## Participants
-
-```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-n_participants <- length(unique(participants$id))
-dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
-sex <- participants %>% drop_na(sex) %>% count(sex) %>% pull(n) %>% set_names(c("female", "male"))
-
-```
-
-We collected data from `r printnum(n_participants)` bilinguals (`r sex[1]` female), from the Metropolitan Area of Barcelona, between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
-All families gave informed consent before participating.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language.
-We excluded participants with >10% exposure to a third language.
-Fig.
-\@ref(tab:lps) illustrates the distribution of participants across language profiles and ages.
-
-## Questionnaire
-
-```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-pool <- multilex::pool %>% 
-  drop_na(cognate)
-n_categories <- distinct(pool, category) %>% nrow()
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
-n_item_cognate <- count(pool, language, cognate) %>% 
-  pull(n) %>% 
-  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
-```
-
-We implemented an on-line questionnaire using `formr` [@arslan_formr_2020], divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton_infant_2000] and consisted in two lists of words, one in Catalan and one in Spanish).
-Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default).
-Participants filled a long or a short version of the questionnaire.
-Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items.
-Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish).
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
-
-## Data analysis
-
-```{r n_responses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-n_obs <- nrow(responses)
-sum_obs <- responses %>%
-  group_by(te) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  summarise(
-    mean = mean(n),
-    sd = sd(n),
-    min = min(n),
-    max = max(n)
-  )
-
-```
-
-We gathered `r printnum(n_obs, big.mark = ",")` responses.
-Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, *Min* = `r printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
-We modelled the probability of children understanding or producing TEs separately using logistic regression model with a binomial distribution and a logit link function.
-[@agresti_analysis_2010].
-We included several predictors of interest to adjust this probability to participants' and items' properties of interest.
-These predictors are:
-
-* Age of participant in months (`age`, $Age$) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width.
-* Degree of bilingualism of participant (`bilingualism`, $Bilingualism$) computed as the percentage of exposure to a second language (Spanish for participants exposed to >50% to Catalan an *vice versa*), chunked into six groups of participants, with degree of bilingualism 0%, 10%, 20%, 30%, 40% and 50%)
-* Lexical frequency of item (`frequency`, $Frequency$) retrieved from SUBTLEX-CAT [@boada_subtlex-cat_2019] for Catalan words and from SUBTLEX-ESP [@cuetos_subtlex-esp_2011] for Spanish words, expressed as Zipf scores [@zipf_human_1949; @heuven_subtlex-uk_2014], and centred around the mean
-* Language the item belongs to (`dominance`, $Dominance$) labelled as `L1` if the item belongs to the dominant language (e.g., a Catalan item is labelled as `L1` for Catalan-dominant participants and as `L2` for Spanish-dominant participants)
-* Cognateness of the item (`cognate`, $Cognate$), as rated by a trained Spanish-Catalan bilingual linguist.
-
-We aggregated the data across participants.
-We summed the number of participants that *understood*, or *understood and produced* each TE for each combination of the `age`, `bilingualism` levels.
-Numerical variables (`age`, `bilingualism`, and `frequency`) were standardized, and categorical predictors were sum-coded.
-`dominance` was coded as `L2` = -0.5 and `L1` = +0.5, and `cognate` was coded as `Non-cognate` = -0.5 and `Cognate` = +0.5.
-In summary, each TE provided four data points per age bin: L1-Cognate, L1-Non-cognate, L2-Cognate, and L2-Non-cognate.
-
-We adopted a Bayesian approach toward statistical inference.
-This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model [@mcelreath_statistical_2016].
-We first fit a base model that only adjusted for the age of participants and the lexical frequency of the items [@braginsky_consistency_2016: @jones_children_2018].
-We then introduced our predictors of interest in the following order: (1) `dominance`, (2) `bilingualism`, (3) `dominance` per `bilingualism` interaction, (4) `cognate`, (5) `dominance` per `cognate` interaction, (6) `bilingualism` per `cognate` interaction, and (7) `dominance` per `bilingualism` per `cognate` interaction.
-Our extended model can be formalised as follows:
-
-$$
-\begin{aligned}
-y_i \sim Binomial(n_i, p_i), \\
-logit(p_i) &=& \alpha - \dots \\
-& \beta_1 \times Age_i + \dots \\
-& \beta_2 \times Frequency_i + \dots \\
-& \beta_3 \times Dominance_i + \dots \\
-& \beta_4 \times Bilingualism_i + \dots \\
-& \beta_5 \times Cognate_i + \dots \\
-& \beta_6 \times (Dominance_i \times Bilingualism_i) + \dots \\
-& \beta_7 \times (Dominance_i \times Cognate_i) + \dots \\
-& \beta_8 \times (Bilingualism_i \times Cognate_i) + \dots \\
-& \beta_9 \times (Dominance_i \times Bilingualism_i \times Cognate_i), \\
-logit(p_i) = log\frac{y_i}{n_i-y_i} \\
-\alpha \sim Normal(0, 1.5) \\
-\beta_{0-9} \sim Normal(0, 0.5)
-\end{aligned}
-$$
-
-Where:
-
-* $y_i$ is the sum of positive responses to *Understands* or *Understands and Says* for translation equivalent $i$
-* $n_i$ is the total number of responses for translation equivalent $i$
-* $\alpha$ is the intercept, which corresponds to the probability of a positive response when all predictors equal 0
-* $\beta_{1, \dots, 9}$ are the coefficients of the predictors
-
-We implemented this model using the R package `brms` [@R-brms_b; carpenter_stan_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
-Appendix 2 shows different diagnostics of chain convergence.
-We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017]^[Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson_leave_2019]].
-This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
-The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
-We compared the expected log posterior density of each each model to test which one fitted the data the best.
-
-We then explored the posterior distribution of each parameter in the model that fitted the data the best computing the 95% credible intervals and testing whether this interval excluded 0.
-Credible intervals (CrI) indicate the range of values we are 95% certain contain the true value of the parameter, given our prior and observed data.
-We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@R-emmeans].
-Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], and posterior samples were extracted using the `tidybayes` R package [@R-tidybayes].
-
-# Results
 
@@
 ```
-
-# Discussion
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-
-\endgroup
-
-```{r create_r-references, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-r_refs(file = "trajectories.bib")
-
-```

### 2021-07-12 Rmd/05_appendix.Rmd

---
title: "Appendix"
output:
  word_document:
    reference_docx: trajectories_template.docx
  pdf_document:
      latex_engine: xelatex
bibliography: trajectories.bib
csl: apa6.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2021-07-12 Rmd/02_methods.Rmd

--- 2021-01-13 Rmd/trajectories_manuscript.Rmd
@@
 ---
-title             : "The role of cross-linguistic lexical similarity on bilingual word acquisition"
-shorttitle        : "Cross-linguistic similarity and  word acquisition"
-
-author: 
-  - name          : "Gonzalo García-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Daniela Avila-Varela"
-    affiliation   : "1"
-  - name          : "Núria Sebastian-Galles"
-    affiliation   : "1"
-
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
-
-authornote: |
-  Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
-
-abstract: |
-  Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "lexical acquisition, vocabulary, bilingualism"
-wordcount         : "X"
-
-bibliography      : ["trajectories.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-keep_text         : false
-class             : "man"
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  papaja::apa6_pdf:
-    latex_engine  : "xelatex"
-
+title: "Methods"
+output:
+  word_document:
+    reference_docx: trajectories_template.docx
+  pdf_document:
+      latex_engine: xelatex
+bibliography: trajectories.bib
+csl: apa6.csl
 ---
 
-```{r setup, include = FALSE}
-
-knitr::opts_chunk$set(
-    echo = FALSE,
-    fig.align = "center",
-    message = FALSE,
-    warning = FALSE,
-    cache.extra = knitr::rand_seed,
-    out.width = "80%",
-    results = "asis"
-)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
-
-```{r packages, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# load packages
-library(papaja)
-library(knitr)
-library(magrittr)
-library(tidyverse)
-library(readxl)
-library(lubridate)
-library(janitor)
-library(gt)
-library(data.table)
-library(here)
-
-```
-
-# Introduction
-
-Learning words involves two key steps: (1) the encoding of a word form (e.g, /dɒɡ/)
- and (2) its association with a referent (e.g., a DOG).
-There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side [@bergleson2012].
-At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered [@jusczyk1995], as opposed to when only part of the word is uttered.
-At 12 months, infants show sensitivity to both vowel and consonant mispronunciations [@mani2010a], looking shorter to named pictures when their label is mispronounced.
-Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones.
-One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon [@mcclelland2003].
-
-Words are not learnt in isolation, but in contexts rich in word tokens and referents.
-The structure of toddlers' lexicon reflects this high connectivity [@wojcik2018; @hills2010; @fourtassi2020].
-At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing [@mani2010b; @mani2011].
-The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming  [@styles2009; @arias-trejo2013; @wojcik2013], and show evidence of inhibitory links between lexical-semantic entries [@chow2016].
-By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages [@chow2017].
-Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words [@chow2017; @rama2013], and  the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers.
-This suggests that the emergence of rich lexical networks -- both at the phonological and the semantic levels -- grows as a function of the number of words acquired by the toddler.
-
-The structure of the lexicon also impacts the order of acquisition of new words.
-Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills2009].
-@borosky2016 and @borovsky2018 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
-This points to the structure of the lexicon facilitating the strategies children engage during word learning.
-
-The case of bilingual children (here defined as those learning two languages from birth) presents an opportunity to study how lexical acquisition takes placed under a more complex environment, as they  face the challenge of learning two distinct sets of words -- one for each language -- that partially overlap in sound and meaning.
-
-EARLY STAGES IN BILINGUAL LANGUAGE ACQUISITION
-
-When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev_influence_1977; @doyle1977some; @rosenblum1983word; @fernandez1992bilingual; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
-
-@floccia_i_2018, gathered parental vocabulary estimates [@hamilton2000infant] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
-They computed the average phonological similarity between translation equivalents (TEs from now on; e.g.,  *table* in English and *tafel* in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
-Using a similar measure of lexical similarity, @blom_cross-language_2019 extended these results to bilingual children aged three to 10, and reported a positive association between the lexical distance and comprehensive vocabulary size: bilinguals learning lexically close languages showed similar vocabulary sizes to those of monolinguals, but those learning the more distant languages showed lower vocabulary sizes.
-
-These results suggest that linguistic distance between languages plays an important role during bilingual lexical acquisition, and that this effect operates within TEs.
-What mechanisms underlie this effect, and why they seem to operate differently across comprehension and production, or across the dominant and the non-dominant language, remain open issues.
-@floccia_i_2018 pointed to parallel activation as a candidate mechanism behind this facilitation effect.
-The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., *cat*, in English) its corresponding translation in the other language is activated too [e.g., *gato*, in Spanish.
-There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell1991mediated; @spivey1999cross; @costa2000lexical; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh2014one; @von_holzen_impact_2019; @bobb_co-activation_2020; see @costa2005facilitatory for review], and children [e.g.
-@von_holzen_language_2012; @poulin2013lexical; @jardak2019labels; @bosma2020cognate].
-A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
-
-@floccia_i_2018 argue that the degree of phonological similarity between TEs (*cognateness* from now on) of both languages should lead to increased parallel activation of both languages during language exposure, facilitating the acquisition of words in both languages, and leading participants leading two lexically close languages to show larger vocabulary sizes than those learning two lexically distant languages.
-This account, however, neglects the fact that for parallel activation to take place and play a facilitation role, lexical representations of both pairs of the TE must have been already established in the lexicon, and their corresponding phonological forms must have been encoded.
-This is not the case during lexical acquisition, where at least one of the members of the pair has not being acquired yet.
-A more lenient definition of parallel activation could adjust to the context of lexical acquisition more easily: when only one of the members of the TE has been acquired, its lexical representation is activated when, in the presence of its referent, a phonologically similar label is uttered.
-Consider the case of a Catalan-Spanish bilingual that has already learned the word *gat* (Catalan for *cat*), but not *gato* (in Spanish).
-It is possible that, when presented with *gato* in the presence of a cat, she maps this novel label to the familiar word *gat*, and creates a lexical representation for *gato* as a synonym for *gat*.
-The case of a non-cognate is differing.
-Consider now the case of a toddler that has already learnt *gos* (Catalan for *dog*), but not *perro* (Spanish).
-In this case, it would not be possible to map both labels via phonology, give their lack of phonological similarity.
-
-Under this account, cognates (i.e., phonologically similar TEs) would be acquired earlier than non-cognates, but this effect would only play a role once one of the members of the TE has been acquired.
-The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse.
-@schelletter2002effect's reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
-@bosch_first_2014, showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months.
-The low number of words included in the analyses of both studies, and the lack of adjustment for lexical frequency, limits severely the strength of any conclusions concerning the difference in age of acquisition of cognates and non-cognates.
-On the other hand @floccia_i_2018 worked with aggregated estimates of participants' vocabulary size, losing item-level information that could provide information about change in age of acquisition of TEs associated with their degree of phonological similarity.
-Therefore, their results can only be considered as compatible with cognateness playing a role on age of acquisition.
-In this study, we aimed at overcoming these pitfalls by using unaggregated data from individual responses, testing the role of cognateness explicitly and adjusting for lexical frequency.
-
-It also follows from the fact that cognateness operates *within* TEs that the effect of cognateness can only play a role once one of the members of the TE has been acquired, and its phonological form is available during the acquisition of its translation.
-Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language [@floccia_i_2018].
-@dijkstra2013 tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE.
-Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian.
-This suggests that infants that were mostly exposed to Dutch used the phonology of  dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
-A study by @dijkstra2013 showed that the degree of phonological similarity between 
-the performance of  was better for cognates than for non-cognates, only for those children
-
-Finally, previous studies have reported that the properties that describe the form of a word play a stronger role in production than comprehension than in production.
-This is the case of the number of phonemes [@braginsky_consistency_2019], phonological neighbourhood density [@jones_children_2019].
-These results converge with @floccia_i_2018's finding that the effect of the average cognateness between languages was larger in production than in comprehension^[Although @bosch_first_2014 found that participants in their sample produced cognates earlier than non-cognates, no analyses were conducted on comprehensive data, and therefore they cannot be taken as supporting evidence for a larger effect of cognateness on production than in comprehension].
-Accordingly, we expected the role of cognateness to be more central in production than in comprehension.
-
-In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.
-
-```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-participants <- read.csv(here("Data", "participants.csv"))
-items <- read.csv(here("Data", "items.csv"))
-vocabulary <- read.csv(here("Data", "vocabulary.csv"))
-responses <- read.csv(here("Data", "responses.csv"))
-
-```
-
-# Method
 
@@
 
-* $k$ is one of the possible categories (coded as ) responses can take
-* $y_i$ is the `response` value of observation $i$
-* $\alpha_k$ is the mean probability of responses to belong to category $k$
+* $y_i$ is the sum of positive responses to *Understands* or *Understands and Says* for translation equivalent $i$
+* $n_i$ is the total number of responses for translation equivalent $i$
+* $\alpha$ is the intercept, which corresponds to the probability of a positive response when all predictors equal 0
 * $\beta_{1, \dots, 9}$ are the coefficients of the predictors
 
-We implemented the model using the R package `brms` [@burkner_brms_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 4,000 iterations (including 2,000 warm-up iterations per chain).
-We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal due to the large amount of observations introduced in the model.
-We report the 95% credible intervals (CrI) of the estimated posterior distribution of each parameter, which denotes a range of values we are 95% certain contain the true value, given our prior and observed data.
-We report analyses on the probability scale, having transformed the log-odds returned by the model using an inverse logit transformation.
+We implemented this model using the R package `brms` [@R-brms_b; carpenter_stan_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
+Appendix 2 shows different diagnostics of chain convergence.
+We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017]^[Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson_leave_2019]].
+This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
+The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
+We compared the expected log posterior density of each each model to test which one fitted the data the best.
 
-We then fit increasingly simpler models dropping predictors one by one in the following order: (1) `dominance:bilingualism:cognate`, (2) `bilingualism:cognate`, (3) `dominance:cognate:`, (4) `cognate`, (5) `dominance:bilingualism`, (6) `bilingualism`, (7) `dominance`, (8) `frequency`, and (9) `age`.
-We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017] and widely applicable information criterion (WAIC) [@watanabe_asymptotic_2010].
-We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95% credible intervals and testing whether this interval excluded 0.
-We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@lenth_emmeans_2020].
-Data processing and visualisation was done in R [@rcoreteam_r_2020] using the `tidyverse` family of packages [@wickham_welcome_2019], and posterior samples were extracted using the `tidybayes` R package [@kay_tidybayes_2020].
-
-# Results
-
-# Discussion
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-
-\endgroup
-
-```{r create_r-references, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-r_refs(file = "trajectories.bib")
-
-```
+We then explored the posterior distribution of each parameter in the model that fitted the data the best computing the 95% credible intervals and testing whether this interval excluded 0.
+Credible intervals (CrI) indicate the range of values we are 95% certain contain the true value of the parameter, given our prior and observed data.
+We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@R-emmeans].
+Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], and posterior samples were extracted using the `tidybayes` R package [@R-tidybayes]./

### 2021-07-12 Rmd/Old/notes.Rmd

--- 2020-06-30 Code/Rmd/data_analysis_notes.Rmd
@@
 ---
-title: "Notes on data anlaysis"
-output: html_notebook
+title: "Notes"
+author: "Gonzalo García-Castro"
+date: "1/11/2021"
+output: html_document
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+Our model specifies a logistic curve that fits our outcome measure.
+Out outcome measure is the porportion of participants that are reported to understand each word.
+This logistic curve is characterised by three parameters: *asymptote* ($asym$, the value at which the outcome *pleateaus*), *steepness*, ($steep$, how fast the curve grows), and *mid-point* ($mid-point$, the age at which the curve reaches its maximum steepness).
+We identify the *mid-point* of a word as its age of acquisition, and thus us the parameter of interest.
+Out hypothesis is that the language a word belongs to (non-dominant vs.
+dominant), together with its cognate status (non-cognate vs.
+cognate) and frequency, affects the position of its mid-point.
+We assume that the mid-point is generated from a zero-one beta regression model that includes the aforementioned predictors.
+A beta regression model helps us model data that include many values at boundary, as ours are.
+We also included random intercepts for each *translation equivalent* ($TE$).
+We set informed priors for *asymptotes*, *steepnesses*, and *mid-points* according to Wordbank data, and weak priors the the coefficients of the beta-regression model.
+
+$$
+\begin{aligned}proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\ mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ TE_{0i} \sim Normal(0, \sigma_i) \\\beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\ \phi \sim Normal(1.5, 1)\\ \sigma_i \sim HalfCauchy(5)
+\end{aligned}
+$$
 
@@
     - Classification system of items (e.g., *household items*) is adult centric. Children may use the words *pretty* to name jewellery [@bates1994].
-* Mid and mid-upper class families are overrepresented in the sample.
+* Mid and mid-upper class families are overrepresented in the sample.t the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+
+We fit two separate logistic models on comprehensive and productive data, respectively.
+Logistic curves are characterised by three parameters[see @mahr2019 and @mahr2020 for an excellent introduction to this approach, in which we based this analysis]: (a) an *asymptote* (`asym`, upper boundary  of the proportion of infants that have acquired the item), (b) the *steepness* (`steep`, how fast the a given item is acquired by infants), and (c) a *mid-point* (`mid`, the age at which steepness is maximum).
+We work on the assumption that the trajectory of acquisition of items follows a logistic curve shape [@mayor2011], and that the age of acquisition of a word is fairly captured by the position of the mid-point of its curve.
+This approach allows us to model and estimate the age of acquisition of items explicitly.
+Our model estimated the asymptote and steepness as grand means across items (i.e., intercepts).
+Mid-points were estimated by a linear regression model that included our predictors of interest: `age`, `item_dominance` (sum-coded, L2 = -0.5, L1 = 0.5), `cognateness` (sum-coded, Non-cognate = -0.5, Cognate = 0.5).
+We also included random intercepts and `item dominance` slopes for each translation equivalent (`te`) [@barr2013].
+Appendix 1 shows a detailed description of the model.
+
+We estimated the parameters of our model using the Bayesian framework.
+Bayesian models use the Bayes rule to combine prior knowledge about the distribution of a given parameter, and the likelihood of each of the values it can take given the data.
+We relied on the available data on Wordbank [@frank2017] to generate our prior (see Appendix 1 for more details about our prior).
+In order to test the contribution of each predictor on the fit of the model, we started fitting a null model that only included an global intercept and random intercepts for each TE, then fitted a model that also included *bilingualism* and its random slopes by TE, and finally a model that also included *cognateness* and its interaction with *bilingualism*, as well as random slopes for all the fixed effects.
+We compared how the fit of the model changed in every step [@schad2020] using leave-one-out cross validation.
+We fit both the null and alternative models using the R environment [@rcoreteam2013; @rstudioteam2015] and the `brms` package [@burkner2017], which relies on the probabilistic language Stan [@carpenter2017] to approximate posterior distributions.
+We used Pareto-smoothed importance sampling [PSIS; @vehtari2017; @vehtari2019] to compare the null and extended models.
+Data and model results were processed and visualised using the `tidyverse` family of R packages [@wickham2019] and the `tidybayes` package [@kay2020].
+
+To test Hypothesis 2, we run a Bayesian ANOVA using the difference in mid-points between the two word-forms of each translation equivalent (in Catalan and Spanish, respectively), that included *cognateness* as predictor.
+Using the `BayesFactor` R package [@morey2018a], we computed a Bayes factor (*BF*) that compared the likelihood of a linear model that included *cognateness* as a predictor against the likelihood of a linear model that did not, under the light of our data.

### 2021-07-12 Rmd/trajectories_manuscript.Rmd

--- 2021-01-13 Rmd/trajectories_manuscript.Rmd
@@
 
-Learning words involves two key steps: (1) the encoding of a word form (e.g, /dɒɡ/)
- and (2) its association with a referent (e.g., a DOG).
-There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side [@bergleson2012].
-At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered [@jusczyk1995], as opposed to when only part of the word is uttered.
-At 12 months, infants show sensitivity to both vowel and consonant mispronunciations [@mani2010a], looking shorter to named pictures when their label is mispronounced.
-Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones.
-One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon [@mcclelland2003].
-
-Words are not learnt in isolation, but in contexts rich in word tokens and referents.
-The structure of toddlers' lexicon reflects this high connectivity [@wojcik2018; @hills2010; @fourtassi2020].
-At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing [@mani2010b; @mani2011].
-The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming  [@styles2009; @arias-trejo2013; @wojcik2013], and show evidence of inhibitory links between lexical-semantic entries [@chow2016].
-By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages [@chow2017].
-Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words [@chow2017; @rama2013], and  the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers.
-This suggests that the emergence of rich lexical networks -- both at the phonological and the semantic levels -- grows as a function of the number of words acquired by the toddler.
-
-The structure of the lexicon also impacts the order of acquisition of new words.
-Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next [@hills2009].
-@borosky2016 and @borovsky2018 showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
-This points to the structure of the lexicon facilitating the strategies children engage during word learning.
-
-The case of bilingual children (here defined as those learning two languages from birth) presents an opportunity to study how lexical acquisition takes placed under a more complex environment, as they  face the challenge of learning two distinct sets of words -- one for each language -- that partially overlap in sound and meaning.
-
-EARLY STAGES IN BILINGUAL LANGUAGE ACQUISITION
-
-When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When both languages are taken into account, bilingual children  seem to know, at least, as many words as monolinguals do [@ben-zeev_influence_1977; @doyle1977some; @rosenblum1983word; @fernandez1992bilingual; @pearson_patterns_1994; @bialystok_receptive_2010; @hoff_dual_2012; @core_total_2013; @hoff_expressive_2014; @blom_cross-language_2019; but see @pearson_lexical_1993; @houwer_bilingualmonolingual_2014].
-
-@floccia_i_2018, gathered parental vocabulary estimates [@hamilton2000infant] from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
-They computed the average phonological similarity between translation equivalents (TEs from now on; e.g.,  *table* in English and *tafel* in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
-Using a similar measure of lexical similarity, @blom_cross-language_2019 extended these results to bilingual children aged three to 10, and reported a positive association between the lexical distance and comprehensive vocabulary size: bilinguals learning lexically close languages showed similar vocabulary sizes to those of monolinguals, but those learning the more distant languages showed lower vocabulary sizes.
-
-These results suggest that linguistic distance between languages plays an important role during bilingual lexical acquisition, and that this effect operates within TEs.
-What mechanisms underlie this effect, and why they seem to operate differently across comprehension and production, or across the dominant and the non-dominant language, remain open issues.
-@floccia_i_2018 pointed to parallel activation as a candidate mechanism behind this facilitation effect.
-The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., *cat*, in English) its corresponding translation in the other language is activated too [e.g., *gato*, in Spanish.
-There is a vast body of evidence supporting language non-selective lexical access in both adults [e.g., @dell1991mediated; @spivey1999cross; @costa2000lexical; @dijkstra_bilingual_2005; @thierry_brain_2007; @hoshino_cognate_2008; @yudes_cognate_2010; @kroll_multilingual_2013; @singh2014one; @von_holzen_impact_2019; @bobb_co-activation_2020; see @costa2005facilitatory for review], and children [e.g.
-@von_holzen_language_2012; @poulin2013lexical; @jardak2019labels; @bosma2020cognate].
-A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
-
-@floccia_i_2018 argue that the degree of phonological similarity between TEs (*cognateness* from now on) of both languages should lead to increased parallel activation of both languages during language exposure, facilitating the acquisition of words in both languages, and leading participants leading two lexically close languages to show larger vocabulary sizes than those learning two lexically distant languages.
-This account, however, neglects the fact that for parallel activation to take place and play a facilitation role, lexical representations of both pairs of the TE must have been already established in the lexicon, and their corresponding phonological forms must have been encoded.
-This is not the case during lexical acquisition, where at least one of the members of the pair has not being acquired yet.
-A more lenient definition of parallel activation could adjust to the context of lexical acquisition more easily: when only one of the members of the TE has been acquired, its lexical representation is activated when, in the presence of its referent, a phonologically similar label is uttered.
-Consider the case of a Catalan-Spanish bilingual that has already learned the word *gat* (Catalan for *cat*), but not *gato* (in Spanish).
-It is possible that, when presented with *gato* in the presence of a cat, she maps this novel label to the familiar word *gat*, and creates a lexical representation for *gato* as a synonym for *gat*.
-The case of a non-cognate is differing.
-Consider now the case of a toddler that has already learnt *gos* (Catalan for *dog*), but not *perro* (Spanish).
-In this case, it would not be possible to map both labels via phonology, give their lack of phonological similarity.
-
-Under this account, cognates (i.e., phonologically similar TEs) would be acquired earlier than non-cognates, but this effect would only play a role once one of the members of the TE has been acquired.
-The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse.
-@schelletter2002effect's reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
-@bosch_first_2014, showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months.
-The low number of words included in the analyses of both studies, and the lack of adjustment for lexical frequency, limits severely the strength of any conclusions concerning the difference in age of acquisition of cognates and non-cognates.
-On the other hand @floccia_i_2018 worked with aggregated estimates of participants' vocabulary size, losing item-level information that could provide information about change in age of acquisition of TEs associated with their degree of phonological similarity.
-Therefore, their results can only be considered as compatible with cognateness playing a role on age of acquisition.
-In this study, we aimed at overcoming these pitfalls by using unaggregated data from individual responses, testing the role of cognateness explicitly and adjusting for lexical frequency.
-
-It also follows from the fact that cognateness operates *within* TEs that the effect of cognateness can only play a role once one of the members of the TE has been acquired, and its phonological form is available during the acquisition of its translation.
-Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language [@floccia_i_2018].
-@dijkstra2013 tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE.
-Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian.
-This suggests that infants that were mostly exposed to Dutch used the phonology of  dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
-A study by @dijkstra2013 showed that the degree of phonological similarity between 
-the performance of  was better for cognates than for non-cognates, only for those children
-
-Finally, previous studies have reported that the properties that describe the form of a word play a stronger role in production than comprehension than in production.
-This is the case of the number of phonemes [@braginsky_consistency_2019], phonological neighbourhood density [@jones_children_2019].
-These results converge with @floccia_i_2018's finding that the effect of the average cognateness between languages was larger in production than in comprehension^[Although @bosch_first_2014 found that participants in their sample produced cognates earlier than non-cognates, no analyses were conducted on comprehensive data, and therefore they cannot be taken as supporting evidence for a larger effect of cognateness on production than in comprehension].
-Accordingly, we expected the role of cognateness to be more central in production than in comprehension.
-
-In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.
-
-```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-participants <- read.csv(here("Data", "participants.csv"))
-items <- read.csv(here("Data", "items.csv"))
-vocabulary <- read.csv(here("Data", "vocabulary.csv"))
-responses <- read.csv(here("Data", "responses.csv"))
+```{r introduction, child="01_introduction.Rmd"}
 
@@
 # Method
-
-## Participants
-
-```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-n_participants <- length(unique(participants$id))
-dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
-sex <- participants %>% drop_na(sex) %>% count(sex) %>% pull(n) %>% set_names(c("female", "male"))
-
-```
-
-We collected data from `r printnum(n_participants)` bilinguals (`r sex[1]` female), from the Metropolitan Area of Barcelona, between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
-All families gave informed consent before participating.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language.
-We excluded participants with >10% exposure to a third language.
-Fig.
-\@ref(tab:lps) illustrates the distribution of participants across language profiles and ages.
-
-## Questionnaire
-
-```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-pool <- multilex::pool %>% 
-  drop_na(cognate)
-n_categories <- distinct(pool, category) %>% nrow()
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
-n_item_cognate <- count(pool, language, cognate) %>% 
-  pull(n) %>% 
-  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
-```
-
-We implemented an on-line questionnaire using `formr` [@arslan_formr_2020], divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton_infant_2000] and consisted in two lists of words, one in Catalan and one in Spanish).
-Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default).
-Participants filled a long or a short version of the questionnaire.
-Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items.
-Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish).
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
-
-## Data analysis
-
-```{r n_responses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-n_obs <- nrow(responses)
-sum_obs <- responses %>%
-  group_by(te) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  summarise(
-    mean = mean(n),
-    sd = sd(n),
-    min = min(n),
-    max = max(n)
-  )
-
-```
-
-We gathered `r printnum(n_obs, big.mark = ",")` responses.
-Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
-We modelled the probability of responses ($y_i$) belonging to any of the three possible categories $k$--1 = *No* or say, 2 = *Understands*, and 3 = *Understands and Says*-- using an ordinal regression model with a cumulative logit link function [@agresti_analysis_2010; @burkner_ordinal_2019].
-This model estimates the probability of response $y_i$ to belong to category $k$ or lower.
-In our case, this will result in the estimation of the probability of $y_i \leq 1$,  $y_i \leq 2$, and  $y_i \leq 3$, where
-
-$$
-p_k = Pr(y_i = k) = Pr(y_i \leq k) - Pr(y_i \leq k-1).
-$$
-
-We include several predictors of interest to adjust this probability to participants' and items' (lexical frequency, dominance, cognateness) characteristics. These predictors are:
-
-* Age of participant in months (`age`, $Age$) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30 and centred around the mean
-* Degree of bilingualism of participant (`bilingualism`, $Bilingualism$) computed as the percentage of exposure to a second language (Spanish for participants exposed to >50% to Catalan an *vice versa*) rounded to units and centered around the mean
-* Lexical frequency of item (`frequency`, $Frequency$) retrieved from SUBTLEX-CAT [@boada_subtlex-cat_2019] for Catalan words and from SUBTLEX-ESP [@cuetos_subtlex-esp_2011] for Spanish words, expressed as Zipf scores [@heuven_subtlex-uk_2014], and centred around the mean
-* Language the item belongs to (`dominance`, $Dominance$) labelled as `L1` if the item belongs to the dominant language (e.g., a Catalan item is labelled as `L1` for Catalan-dominant participants and as `L2` for Spanish-dominant participants), coded as `L2` = -0.5 and `L1` = 0.5
-* Cognateness of the item (`cognate`, $Cognate$), coded as `Non-cognate` = -0.5 and `Cognate` = 0.5
-
-We adopted a Bayesian approach toward statistical inference, which allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model.
-Our extended model can be formalised as follows:
-
-$$
-\begin{aligned}
-y_i \sim Ordered(p) \\
-log\frac{Pr(y_i \leq k)}{1-Pr(y_i \leq k)} &= \alpha_k - \dots \\
-& \beta_1 \times Age_i + \dots \\
-& \beta_2 \times Frequency_i + \dots \\
-& \beta_3 \times Dominance_i + \dots \\
-& \beta_4 \times Bilingualism_i + \dots \\
-& \beta_5 \times Cognate_i + \dots \\
-& \beta_6 \times (Dominance_i \times Bilingualism_i) + \dots \\
-& \beta_7 \times (Dominance_i \times Cognate_i) + \dots \\
-& \beta_8 \times (Bilingualism_i \times Cognate_i) + \dots \\
-& \beta_9 \times (Dominance_i \times Bilingualism_i \times Cognate_i), \\
-\alpha \sim Normal(0, 10) \\
-\beta_{0-9} \sim Normal(0, 10)
-\end{aligned}
-$$
-
-Where:
-
-* $k$ is one of the possible categories (coded as ) responses can take
-* $y_i$ is the `response` value of observation $i$
-* $\alpha_k$ is the mean probability of responses to belong to category $k$
-* $\beta_{1, \dots, 9}$ are the coefficients of the predictors
-
-We implemented the model using the R package `brms` [@burkner_brms_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 4,000 iterations (including 2,000 warm-up iterations per chain).
-We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal due to the large amount of observations introduced in the model.
-We report the 95% credible intervals (CrI) of the estimated posterior distribution of each parameter, which denotes a range of values we are 95% certain contain the true value, given our prior and observed data.
-We report analyses on the probability scale, having transformed the log-odds returned by the model using an inverse logit transformation.
-
-We then fit increasingly simpler models dropping predictors one by one in the following order: (1) `dominance:bilingualism:cognate`, (2) `bilingualism:cognate`, (3) `dominance:cognate:`, (4) `cognate`, (5) `dominance:bilingualism`, (6) `bilingualism`, (7) `dominance`, (8) `frequency`, and (9) `age`.
-We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017] and widely applicable information criterion (WAIC) [@watanabe_asymptotic_2010].
-We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95% credible intervals and testing whether this interval excluded 0.
-We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@lenth_emmeans_2020].
-Data processing and visualisation was done in R [@rcoreteam_r_2020] using the `tidyverse` family of packages [@wickham_welcome_2019], and posterior samples were extracted using the `tidybayes` R package [@kay_tidybayes_2020].
 

### 2021-07-12 Rmd/01_introduction.Rmd
I hope I don't overwrite remote repo

--- 2021-06-17 Rmd/introduction.Rmd
@@
 csl: apa6.csl
+self_contained: true
 ---

### 2021-07-12 Rmd/02_methods.Rmd
I hope I don't overwrite remote repo

@@
 title: "Methods"
+author: "Gonzalo García-Castro"
+date: "6/22/2021"
 output:
@@
 
+# Methods
+
 ## Participants
 
-```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-n_participants <- length(unique(participants$id))
-dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
-sex <- participants %>% drop_na(sex) %>% count(sex) %>% pull(n) %>% set_names(c("female", "male"))
-
-```
-
-We collected data from `r printnum(n_participants)` bilinguals (`r sex[1]` female), from the Metropolitan Area of Barcelona, between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
-All families gave informed consent before participating.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language.
-We excluded participants with >10% exposure to a third language.
-Fig.
-\@ref(tab:lps) illustrates the distribution of participants across language profiles and ages.
-
-## Questionnaire
-
-```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-pool <- multilex::pool %>% 
-  drop_na(cognate)
-n_categories <- distinct(pool, category) %>% nrow()
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
-n_item_cognate <- count(pool, language, cognate) %>% 
-  pull(n) %>% 
-  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
-```
-
-We implemented an on-line questionnaire using `formr` [@arslan_formr_2020], divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton_infant_2000] and consisted in two lists of words, one in Catalan and one in Spanish).
-Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default).
-Participants filled a long or a short version of the questionnaire.
-Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items.
-Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish).
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
+## Vocabulary checklist
 
@@
 
-```{r n_responses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-n_obs <- nrow(responses)
-sum_obs <- responses %>%
-  group_by(te) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  summarise(
-    mean = mean(n),
-    sd = sd(n),
-    min = min(n),
-    max = max(n)
-  )
-
-```
-
-We gathered `r printnum(n_obs, big.mark = ",")` responses.
-Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, *Min* = `r printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
-We modelled the probability of children understanding or producing TEs separately using logistic regression model with a binomial distribution and a logit link function.
-[@agresti_analysis_2010].
-We included several predictors of interest to adjust this probability to participants' and items' properties of interest.
-These predictors are:
-
-* Age of participant in months (`age`, $Age$) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width.
-* Degree of bilingualism of participant (`bilingualism`, $Bilingualism$) computed as the percentage of exposure to a second language (Spanish for participants exposed to >50% to Catalan an *vice versa*), chunked into six groups of participants, with degree of bilingualism 0%, 10%, 20%, 30%, 40% and 50%)
-* Lexical frequency of item (`frequency`, $Frequency$) retrieved from SUBTLEX-CAT [@boada_subtlex-cat_2019] for Catalan words and from SUBTLEX-ESP [@cuetos_subtlex-esp_2011] for Spanish words, expressed as Zipf scores [@zipf_human_1949; @heuven_subtlex-uk_2014], and centred around the mean
-* Language the item belongs to (`dominance`, $Dominance$) labelled as `L1` if the item belongs to the dominant language (e.g., a Catalan item is labelled as `L1` for Catalan-dominant participants and as `L2` for Spanish-dominant participants)
-* Cognateness of the item (`cognate`, $Cognate$), as rated by a trained Spanish-Catalan bilingual linguist.
-
-We aggregated the data across participants.
-We summed the number of participants that *understood*, or *understood and produced* each TE for each combination of the `age`, `bilingualism` levels.
-Numerical variables (`age`, `bilingualism`, and `frequency`) were standardized, and categorical predictors were sum-coded.
-`dominance` was coded as `L2` = -0.5 and `L1` = +0.5, and `cognate` was coded as `Non-cognate` = -0.5 and `Cognate` = +0.5.
-In summary, each TE provided four data points per age bin: L1-Cognate, L1-Non-cognate, L2-Cognate, and L2-Non-cognate.
-
-We adopted a Bayesian approach toward statistical inference.
-This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model [@mcelreath_statistical_2016].
-We first fit a base model that only adjusted for the age of participants and the lexical frequency of the items [@braginsky_consistency_2016: @jones_children_2018].
-We then introduced our predictors of interest in the following order: (1) `dominance`, (2) `bilingualism`, (3) `dominance` per `bilingualism` interaction, (4) `cognate`, (5) `dominance` per `cognate` interaction, (6) `bilingualism` per `cognate` interaction, and (7) `dominance` per `bilingualism` per `cognate` interaction.
-Our extended model can be formalised as follows:
+$$
 
 $$
-\begin{aligned}
-y_i \sim Binomial(n_i, p_i), \\
-logit(p_i) &=& \alpha - \dots \\
-& \beta_1 \times Age_i + \dots \\
-& \beta_2 \times Frequency_i + \dots \\
-& \beta_3 \times Dominance_i + \dots \\
-& \beta_4 \times Bilingualism_i + \dots \\
-& \beta_5 \times Cognate_i + \dots \\
-& \beta_6 \times (Dominance_i \times Bilingualism_i) + \dots \\
-& \beta_7 \times (Dominance_i \times Cognate_i) + \dots \\
-& \beta_8 \times (Bilingualism_i \times Cognate_i) + \dots \\
-& \beta_9 \times (Dominance_i \times Bilingualism_i \times Cognate_i), \\
-logit(p_i) = log\frac{y_i}{n_i-y_i} \\
-\alpha \sim Normal(0, 1.5) \\
-\beta_{0-9} \sim Normal(0, 0.5)
-\end{aligned}
-$$
-
-Where:
-
-* $y_i$ is the sum of positive responses to *Understands* or *Understands and Says* for translation equivalent $i$
-* $n_i$ is the total number of responses for translation equivalent $i$
-* $\alpha$ is the intercept, which corresponds to the probability of a positive response when all predictors equal 0
-* $\beta_{1, \dots, 9}$ are the coefficients of the predictors
-
-We implemented this model using the R package `brms` [@R-brms_b; carpenter_stan_2017] as `response ~  age + frequency + dominance*bilingualism*cognate`, running four MCMC sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
-Appendix 2 shows different diagnostics of chain convergence.
-We compared all models using leave-one-out cross-validation (LOO-CV) [@vehtari_practical_2017]^[Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson_leave_2019]].
-This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
-The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
-We compared the expected log posterior density of each each model to test which one fitted the data the best.
-
-We then explored the posterior distribution of each parameter in the model that fitted the data the best computing the 95% credible intervals and testing whether this interval excluded 0.
-Credible intervals (CrI) indicate the range of values we are 95% certain contain the true value of the parameter, given our prior and observed data.
-We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@R-emmeans].
-Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], and posterior samples were extracted using the `tidybayes` R package [@R-tidybayes]./

### 2021-07-12 Rmd/report.Rmd
I hope I don't overwrite remote repo

---
title: "Trajectories"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: "trajectories.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "100%",
  results = "asis",
  dev.args = list(png = list(type = "cairo"))
)
options(knitr.kable.NA = '-',
        ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
        ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# load packages
library(tidyverse)
library(gt)
library(lubridate)
library(wesanderson)
library(brms)
library(htmltools)
library(tidybayes)
library(scales)
library(patchwork)
library(janitor)
library(here)

# set params
theme_custom <- function(){
  theme_minimal() +
    theme(
      panel.grid = element_line(colour = "grey", linetype = "dotted"),
      axis.line = element_line(colour = "black"),
      text = element_text(size = 12, colour = "black"),
      axis.text = element_text(colour = "black")
    )
}

theme_set(theme_custom())

# import data
items <- readRDS(here("Data", "items.rds"))
participants <- readRDS(here("Data", "participants.rds"))
responses <- readRDS(here("Data", "responses.rds"))

# import models
vocabulary <- readRDS(here("Data", "vocabulary.rds"))
vocabulary_long <- readRDS(here("Data", "vocabulary_long.rds"))
fit_prior <- readRDS(here("Results", "irt_2_prior.rds"))
fit <- readRDS(here("Results", "irt_2.rds"))

```

# Questionnaires

# Items {.tabset .tab-pills}

```{r items_list, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=4, out.width="100%"}
items %>% 
  pivot_wider(
    id_cols = c("te", "class", "cognate"),
    names_from = "language",
    values_from = c("label", "frequency", "ipa")
  ) %>% 
  clean_names() %>% 
  arrange(te) %>% 
  gt() %>% 
  tab_spanner(label = md("**Catalan**"), columns = contains("_catalan")) %>% 
  tab_spanner(label = md("**Spanish**"), columns = contains("_spanish")) %>% 
  cols_label(
    te = "TE ID",
    class = "Class",
    cognate = "Cognateness",
    label_catalan = "Label",
    label_spanish = "Label",
    ipa_catalan = "Phonology",
    ipa_spanish = "Phonology",
    frequency_catalan = "Frequency",
    frequency_spanish = "Frequency"
  ) %>% 
  fmt_number(columns = starts_with("frequency")) %>% 
  div(style = "height:600px; overflow-x: scroll; overflow-y: scroll", .)
```

<br>

## Lexical frequency

```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

by_cognate <- items %>% 
  drop_na(frequency) %>% 
  ggplot(aes(frequency, cognate, fill = cognate, colour = cognate)) +
  geom_jitter(height = 0.05, alpha = 0.1, shape = 1, stroke = 1) +
  stat_slab(position = position_nudge(y = 0.1), alpha = 0.5) +
  geom_boxplot(width = 0.05, colour = "black", fill = "white",
               position = position_nudge(y = 0.1), outlier.colour = NA) +
  labs(
    y = "Cognateness", 
    x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
    colour = "cognateness", 
    fill = "Cognateness"
  )

by_language <- items %>% 
  drop_na(frequency) %>% 
  ggplot(aes(frequency, language, fill = cognate, colour = cognate)) +
  geom_jitter(height = 0.05, alpha = 0.1, shape = 1, stroke = 1) +
  stat_slab(position = position_nudge(y = 0.1), alpha = 0.5) +
  geom_boxplot(width = 0.05, colour = "black", fill = "white",
               position = position_nudge(y = 0.1), outlier.colour = NA) +
  labs(y = "Cognateness", x = "Lexical frequency", colour = "cognateness", fill = "Cognateness")

by_class <- items %>% 
  drop_na(frequency, class) %>% 
  ggplot(aes(frequency, class, fill = cognate, colour = cognate)) +
  geom_jitter(height = 0.05, alpha = 0.1, shape = 1, stroke = 1) +
  stat_slab(position = position_nudge(y = 0.1), alpha = 0.5) +
  geom_boxplot(width = 0.05, colour = "black", fill = "white",
               position = position_nudge(y = 0.1), outlier.colour = NA) +
  labs(y = "Cognateness", x = "Lexical frequency", colour = "cognateness", fill = "Cognateness")

(by_cognate / by_language) + by_class +
  plot_layout(2) &
  theme_custom() &
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    panel.grid.major.y = element_blank()
  )

```

## Cognateness

```{r items_cognate, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
by_cognate <- items %>% 
  distinct(te, class, cognate) %>% 
  ggplot(aes(x = cognate, fill = class)) +
  geom_bar() +
  coord_flip() +
  labs(x = "Cognateness", y = "Number of items", fill = "Class")

by_class <- items %>% 
  distinct(te, class, cognate) %>% 
  ggplot(aes(x = class, fill = cognate)) +
  geom_bar(position = position_fill()) +
  labs(x = "Cognateness", y = "Proportion of items (%)", fill = "Cognateness") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent)

by_cognate / by_class +
  plot_layout() &
  theme_custom() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right",
    legend.title = element_blank()
  )

```

## {-}

# Participants {.tabset .tab-pills}

```{r participants_time, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
  group_by(time_stamp, lp, dominant_language) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  group_by(lp, dominant_language) %>% 
  mutate(n = cumsum(n)) %>% 
  ggplot(aes(time_stamp, n, colour = interaction(lp, dominant_language, sep = " - "), fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_line(size = 1) +
  labs(x = "Date", y = "Number of responses", colour = "LP - Dominant language") +
  scale_y_continuous(breaks = seq(0, 250, 50)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
  )

```

## Age

```{r participants_age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>% 
  mutate(age = round(age)) %>% 
  count(lp, dominant_language, age) %>% 
  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_col(position = position_fill(), colour = "white") +
  guides(colour = guide_legend(ncol = 2)) +
  coord_flip() +
  labs(y = "Proportion of the sample", x = "Age (months)", colour = "LP - Dominant language") +
  scale_y_continuous(labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
  ) +

  participants %>% 
  mutate(age = round(age)) %>% 
  count(lp, dominant_language, age) %>% 
  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_col(colour = "white") +
  labs(y = "Number of participants", x = "Age (months)") +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  coord_flip() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank()
  ) +

  plot_layout(nrow = 1, guides = "collect") &
  theme(
    legend.position = "top", 
    axis.text.y = element_text(size = 9),
    panel.grid.major = element_blank()
  )

```

## Language profile

```{r participants_lp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 11)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  )

```

## {-}

# Data analysis {.tabset .tab-pills}

We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age` by `item`, and their correlation parameter. We then extended this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope. Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`. The models were implemented in `brms` as:

* `understands/produces ~ 1 + age + frequency + (1 + age | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age + frequency + doe + (1 + age + doe | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age + frequency + doe*cognate + (1 + age + doe | item) + (1 + frequency + cognate| id)`

## Model equation

$$
\begin{aligned}
\log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
& (\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
& (\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
& (\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
& (\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
& (\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
& \varepsilon_{pi}
\end{aligned}
$$

## R code (`brms`)

```
understands ~
1 + age + frequency + lp*cognate +
(1 + age + lp | te) +
(1 + frequency + cognate | id),
family =  bernoulli("logit")
```

## Stan code

Stan code generated by `brms::stancode`:

```{stan stan_model, echo=TRUE, output.var="model"}
// generated with brms 2.15.0
functions {
/* compute correlated group-level effects
* Args: 
*   z: matrix of unscaled group-level effects
*   SD: vector of standard deviation parameters
*   L: cholesky factor correlation matrix
* Returns: 
*   matrix of scaled group-level effects
*/ 
matrix scale_r_cor(matrix z, vector SD, matrix L) {
// r is stored in another dimension order than z
return transpose(diag_pre_multiply(SD, L) * z);
}
}
data {
int<lower=1> N;  // total number of observations
int Y[N];  // response variable
int<lower=1> K;  // number of population-level effects
matrix[N, K] X;  // population-level design matrix
// data for group-level effects of ID 1
int<lower=1> N_1;  // number of grouping levels
int<lower=1> M_1;  // number of coefficients per level
int<lower=1> J_1[N];  // grouping indicator per observation
// group-level predictor values
vector[N] Z_1_1;
vector[N] Z_1_2;
vector[N] Z_1_3;
int<lower=1> NC_1;  // number of group-level correlations
// data for group-level effects of ID 2
int<lower=1> N_2;  // number of grouping levels
int<lower=1> M_2;  // number of coefficients per level
int<lower=1> J_2[N];  // grouping indicator per observation
// group-level predictor values
vector[N] Z_2_1;
vector[N] Z_2_2;
vector[N] Z_2_3;
int<lower=1> NC_2;  // number of group-level correlations
int prior_only;  // should the likelihood be ignored?
}
transformed data {
int Kc = K - 1;
matrix[N, Kc] Xc;  // centered version of X without an intercept
vector[Kc] means_X;  // column means of X before centering
for (i in 2:K) {
means_X[i - 1] = mean(X[, i]);
Xc[, i - 1] = X[, i] - means_X[i - 1];
}
}
parameters {
vector[Kc] b;  // population-level effects
real Intercept;  // temporary intercept for centered predictors
matrix[M_1, N_1] z_1;  // standardized group-level effects
cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
vector<lower=0>[M_2] sd_2;  // group-level standard deviations
matrix[M_2, N_2] z_2;  // standardized group-level effects
cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
}
transformed parameters {
vector<lower=0>[M_1] sd_1;  // group-level standard deviations
matrix[N_1, M_1] r_1;  // actual group-level effects
// using vectors speeds up indexing in loops
vector[N_1] r_1_1;
vector[N_1] r_1_2;
vector[N_1] r_1_3;
matrix[N_2, M_2] r_2;  // actual group-level effects
// using vectors speeds up indexing in loops
vector[N_2] r_2_1;
vector[N_2] r_2_2;
vector[N_2] r_2_3;
sd_1 = rep_vector(1, rows(sd_1));
// compute actual group-level effects
r_1 = scale_r_cor(z_1, sd_1, L_1);
r_1_1 = r_1[, 1];
r_1_2 = r_1[, 2];
r_1_3 = r_1[, 3];
// compute actual group-level effects
r_2 = scale_r_cor(z_2, sd_2, L_2);
r_2_1 = r_2[, 1];
r_2_2 = r_2[, 2];
r_2_3 = r_2[, 3];
}
model {
// likelihood including constants
if (!prior_only) {
// initialize linear predictor term
vector[N] mu = Intercept + rep_vector(0.0, N);
for (n in 1:N) {
// add more terms to the linear predictor
mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n];
}
target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
}
// priors including constants
target += normal_lpdf(b | 0, 3);
target += normal_lpdf(Intercept | 0, 3);
target += std_normal_lpdf(to_vector(z_1));
target += lkj_corr_cholesky_lpdf(L_1 | 5);
target += normal_lpdf(sd_2 | 0, 0.5)
- 3 * normal_lccdf(0 | 0, 0.5);
target += std_normal_lpdf(to_vector(z_2));
target += lkj_corr_cholesky_lpdf(L_2 | 5);
}
generated quantities {
// actual population-level intercept
real b_Intercept = Intercept - dot_product(means_X, b);
// compute group-level correlations
corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
vector<lower=-1,upper=1>[NC_1] cor_1;
// compute group-level correlations
corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
vector<lower=-1,upper=1>[NC_2] cor_2;
// extract upper diagonal of correlation matrix
for (k in 1:M_1) {
for (j in 1:(k - 1)) {
cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
}
}
// extract upper diagonal of correlation matrix
for (k in 1:M_2) {
for (j in 1:(k - 1)) {
cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
}
}
}
```

## {-}

# Results {.tabset .tab-pills}

## Vocabulary {.tabset .tab-pills}

### Vocabulary size {.tabset .tab-pills}

#### Comprehension

```{r vocabulary_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

vocabulary %>% 
  filter(type=="Understands", modality != "Te") %>% 
  ggplot(aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  vocabulary %>% 
  filter(type=="Understands", modality != "Te") %>% 
  ggplot(aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

  plot_layout(guides = "collect") &
  theme(
    legend.position = "top",
    legend.title = element_blank()
  ) &
  ggsave(here("Figures", "vocabulary_comprehension.png"))

```

#### Production

```{r vocabulary_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

vocabulary %>% 
  filter(type=="Produces", modality != "Te") %>% 
  ggplot(aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  vocabulary %>% 
  filter(type=="Understands", modality != "Te") %>% 
  ggplot(aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

  plot_layout(guides = "collect") &
  theme(
    legend.position = "top",
    legend.title = element_blank()
  ) &
  ggsave(here("Figures", "vocabulary_production.png"))

```

#### {-}

### Short vs. Long {.tabset .tab-pills}

#### Comprehension

```{r vocabulary_long_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary_long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  ggsave(here("Figures", "vocabulary_long_comprehension.png"))
```

#### Production

```{r vocabulary_long_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary_long %>% 
  filter(type=="produces") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  ggsave(here("Figures", "vocabulary_long_production.png"))
```

#### {-}

### {-}

## Prior-predictive checks

```{r ppc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = c("Monolingual", "Bilingual"),
  cognate = c("Cognate", "Non-Cognate"),
  frequency_center = c(-1, 0, 1)
)

m <- add_fitted_draws(nd, fit_prior, n = 20, re_formula = NA) %>% 
  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_grid(lp~frequency_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

```

## Model selection

We compared the performance of these models using Bayesian leave-one-out cross-validation (LOO) using the `loo` and `loo_compare` functions of the `brms` R package (dependent of the `LOO` R package). LOO consists in computing the average likelihood of each observation after estimating the model's parameters leave that same observation out of the data set. Although the `loo` function uses a particular algorithm that speeds up the computation of this criterion (pareto-smooth importance sampling, PSIS), the size of our data set lead us to rely on the computation of the same criterion using a sampling approach via de `loo_subsample` function.

## Fixed effects

```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fit)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  )

```

<br>

```{r fix_effets_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

post <- gather_draws(fit, `b_.*`, regex = TRUE) %>% 
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))

ggplot(post, aes(.value)) +
  facet_wrap(~.variable, scales = "free_x") +
  geom_vline(xintercept = 0, size = 1, colour = "grey") +
  stat_histinterval(fill = "lightblue") +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_custom() +
  theme(
    legend.position = "none",
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0),
    # axis.text = element_text(size = 7),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  ) +
  ggsave(here("Figures", "responses_fixed.png"))
```

## Random effects {.tabset .tab-pils}

### Participant

```{r random_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

re_id <- ranef(fit)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fit)[1])) %>% 
  left_join(distinct(fit$data, id, lp))

ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[1]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Participant average probability of comprehension",
       y = "Participant") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_custom() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

```

### Item

```{r re_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

re_te <- ranef(fit)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fit)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  left_join(distinct(fit$data, te, lp))

ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[2]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Item average probability of comprehension",
       y = "Item") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_custom() +
  theme(
    axis.text.y = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  )
```

### {-}

## Marginal means

```{r emmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = c("Monolingual", "Bilingual"),
  cognate = c("Cognate", "Non-Cognate"),
  frequency_center = c(-1,  0, 1)
)

m <- add_fitted_draws(nd, fit, re_formula = NA, n = 25) %>% 
  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_wrap(lp~frequency_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +

  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "right",
    legend.title = element_blank()
  )

```

### Area under the curve (AUC)

### Traceplots

```{r diagnostics_traceplots, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

gather_draws(fit, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain") +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )

```

## {-}

### 2021-07-12 Rmd/02_methods.Rmd

--- 2021-07-12 Rmd/02_methods.Rmd
@@
 We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@R-emmeans].
-Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], and posterior samples were extracted using the `tidybayes` R package [@R-tidybayes]./
+Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], and posterior samples were extracted using the `tidybayes` R package [@R-tidybayes].

### 2021-07-12 Rmd/01_introduction.Rmd

--- 2021-06-17 Rmd/introduction.Rmd
@@
+<<<<<<< HEAD:Rmd/01_introduction.Rmd
 ---
@@
 We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.
-
-## References

### 2021-07-14 Rmd/01_introduction.Rmd
First corrections to Núria's comments

--- 2021-07-12 Rmd/01_introduction.Rmd
@@
 ---
-title: "Introduction"
-author: "Gonzalo García-Castro"
-date: "3/17/2021"
-output:
-  word_document:
-    reference_docx: trajectories_template.docx
-  pdf_document:
-      latex_engine: xelatex
-bibliography: trajectories.bib
-csl: apa6.csl
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 
-Before the end of their first year of life, infants are able to encode sound patterns [@jusczyk1995infants; @tincoff1999some] and associate them with their referents [@bergelson2012months].
-Later on, infants begin to represent those word-forms with finer phonological detail [e.g.
-@werker2002infants], and during the last half of their second year, they acquire new words at an increasingly fast rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension].
+Before the end of their first year of life, infants start directing their gaze to some object when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months] and parental reports [e.g., @fenson1994variability].
+During the last half of their second year, they acquire new words at an increasingly fast rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension].
 Despite the considerable variation in children's trajectories of vocabulary growth, such trajectories seem quite consistent across languages [@braginsky2019consistency].
-Nonetheless, most of the literature on early word acquisition has been conducted on monolingual children, and does not address the problem of how bilinguals--who represent a substantial proportion of the populatio in most societies--acquire words at early ages.
+Nonetheless, most of the literature on early word acquisition has been conducted on monolingual children, and does not address the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
 
@@
 Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating].
-While these studies have mostly relied on samples of bilingual children learning two relatively distant languages (as it is the case of English and Spanish) it is unclear whether children learning two more similar languages also know less words in each of their languages than monolinguals.
+While these studies have mostly relied on samples of bilingual children learning two relatively distant languages (as it is the case of English and Spanish) it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
 What role could linguistic distance play during early vocabulary growth?
@@
 Particularly, when both languages are linguistically close, like Spanish and Catalan (both Roman languages), they are more likely to share a larger amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English--one Roman, the other Germanic.
-For instance, in the presence of a door, a Spanish-Catalan bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
+For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
 It could be the case that mapping two phonologically similar labels (cognates, like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
@@
 A recent study by @floccia2018introduction provided evidence in line with this claim.
-The authors collected vocabulary data from 372 24-month-old bilingual toddlers in the United Kingdom who were learning English and an additional language.
-The additional language was one a pool of 13 languages that differed in their phonological similarity with respect to English, such as Dutch, Italian, or Mandarin.
-The authors defined phonological similarity as the average edit distance (length-adjusted Levenshtein distance) between translation equivalents in each language pair.
-The authors reported an increase in productive vocabulary size in the additional language associated with an increase in phonological overlap between the language pair.
+The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers in the United Kingdom who were learning English and an additional language.
+The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
+The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English.
+Phonological similarity was measuring by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
+The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other [@levenshtein1966binary].
+Floccia et al.
+divided the resulting score by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
+Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
 For example, English-Dutch bilinguals (22.14% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% phonological similarity) were able to produce in Mandarin.
@@
 One of the clearest examples of parallel activation was provided by @costa2000cognate.
-In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Catalan (their dominant language).
-In half of the trials, the object labels were cognates in Catalan and Spanish (*sofà*-*sofá*, translations of *sofa*), whereas in the other half of the trials labels were non-cognates (*taula*-*mesa*, translations of *table*).
+In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish.
+In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trials labels were non-cognates (*mesa*-*taula*, translations of *table*).
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
-Importantly, monolinguals did not show this effect.
-These results suggest that, for the bilinguals, Spanish phonology was activated during the production of Catalan words, facilitating the naming of cognate pictures.
+Importantly, Spanish monolinguals did not show this effect.
+These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
 Several recent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
-In summary, the parallel activation account of bilingual word acquisition suggests that bilinguals' vocabulary size grows faster when both languages share a high proportion of cognate translation equivalents, as a result of increased activation of word forms in both languages during speech processing.
+Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition.
+Therefore, children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
+If cognates are the reason for this facilitation effect, it could be predicted that cognates are acquired earlier than non-cognates.
+Likely as it might seem, an earlier age of acquisition for cognates is not the only scenario where Floccia et al.'s results are account for.
+For example, languages that share more cognates might also share a larger overlap in their phonemic inventories, which could facilitate the acquisition of all words, regardless of their cognate status.
 
-If parallel activation plays a role during early word acquisition, it could be predicted that the age of acquisition of cognate translation equivalents is, on average, lower than that of non-cognates.
-However, the evidence supporting this claim is scarce.
-On the one hand, @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates, and @schelletter2002effect reported a longitudinal single case of an English-German bilingual who produced cognates earlier than non-cognates, on average.
+However, the evidence on an earlier age of acquisition for cognates compared to non-cognates is scarce.
+To the best of our knowledge only two studies have provided direct data on this issue.
+First, @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
+Second, @schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
 The low sample size in these studies make it challenging to draw strong conclusions about the effect of cognateness in vocabulary growth.
-On the other hand Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
+Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
+In their discussion the authors state the following (pp.
+70):
+
+> "This  finding  also  provides  support  to  the  proposal  that  the  cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life."
+
 We identify two main reasons why we cannot conclude an earlier age of acquisition for cognates than for non cognate from Floccia et al.'s results.
@@
 We hypothesised that cognate words would be acquired earlier than non-cognate word, and that this difference would be larger in participants with lower exposure to the language such word belongs to.
-Using an *ad hoc* online vocabulary checklist, we collected data from a sample of bilinguals aged 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+Using an designed online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of bilinguals aged 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
 We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.
-
-## References

### 2021-07-14 Rmd/trajectories_manuscript.tex
First corrections to Núria's comments

--- 2021-01-13 Rmd/trajectories_manuscript.tex
@@
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
+\usepackage{amsmath}
 \usepackage{ifxetex,ifluatex}
@@
   \usepackage{textcomp} % provide euro and other symbols
+  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \fi
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 \newlength{\cslhangindent}
 \setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
-  {\par}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
 \title{The role of cross-linguistic lexical similarity on bilingual word acquisition}
-\author{Gonzalo García-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Núria Sebastian-Galles\textsuperscript{1}}
+\author{Gonzalo Garcia-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Nuria Sebastian-Galles\textsuperscript{1}}
 \date{}
@@
 
-Learning words involves two key steps: (1) the encoding of a word form (e.g, /dɒɡ/)
-and (2) its association with a referent (e.g., a DOG).
-There is evidence that word learning occurs at early ages: six-month-old infants show a preference toward named pictures, relative to unnamed pictures displayed side-by-side ({\textbf{???}}).
-At 7.5 months, the phonological representations of familiar words seem to be quite detailed, as infants show evidence of recognition when the whole word is uttered ({\textbf{???}}), as opposed to when only part of the word is uttered.
-At 12 months, infants show sensitivity to both vowel and consonant mispronunciations ({\textbf{???}}), looking shorter to named pictures when their label is mispronounced.
-Remarkable as it is, the phonological specificity of infants' lexical entries is only the foundation of following developmental milestones.
-One of them, central to this study, is the emergence of excitatory and inhibitory links between word representations, which represents one of the essential characteristics of the adult lexicon ({\textbf{???}}).
+\hypertarget{introduction-1}{%
+\section{Introduction}\label{introduction-1}}
 
-Words are not learnt in isolation, but in contexts rich in word tokens and referents.
-The structure of toddlers' lexicon reflects this high connectivity ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
-At 18 months, infants' recognition of spoken words is sensitive to phonological priming effects, suggesting that their lexical entries are phonological linked, and co-activated during speech processing ({\textbf{???}}; {\textbf{???}}).
-The emergence of semantic links in the developing lexicon occurs somewhat later, at 24 months of age, when infants show sensitivity to semantic priming ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}), and show evidence of inhibitory links between lexical-semantic entries ({\textbf{???}}).
-By 24 months, toddlers' recognition of familiar words also seems to follow a hierarchical fashion, as revealed by their sensitivity to the phonology and semantic similarity of distractor pictures presented along the named picture: the interference of phonological distractors is stronger at earlier stages of spoken word recognition, that of semantic distractors is stronger at later stages ({\textbf{???}}).
-Vocabulary size predicts the emergence and the strength of these effects: 18 months-old toddlers are sensitive to semantic priming and interference too, provided they know a similar amount of words ({\textbf{???}}; {\textbf{???}}), and the strength of the phonological and semantic distractors during spoken word recognition grows stronger along the vocabulary size of the toddlers.
-This suggests that the emergence of rich lexical networks -- both at the phonological and the semantic levels -- grows as a function of the number of words acquired by the toddler.
+Before the end of their first year of life, infants start directing their gaze to some object when hearing their labels, according to both experimental data (Bergelson \& Swingley, 2012; Jusczyk \& Aslin, 1995; Tincoff \& Jusczyk, 1999) and parental reports (e.g., Fenson et al., 1994).
+During the last half of their second year, they acquire new words at an increasingly fast rate (Bergelson, 2020; Bloom, 2002; Fenson et al., 1994).
+Despite the considerable variation in children's trajectories of vocabulary growth, such trajectories seem quite consistent across languages (Braginsky, Yurovsky, Marchman, \& Frank, 2019).
+Nonetheless, most of the literature on early word acquisition has been conducted on monolingual children, and does not address the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
 
-The structure of the lexicon also impacts the order of acquisition of new words.
-Words that share a high degree of phonological and/or semantic overlap with words already acquired are more likely to be acquired next ({\textbf{???}}).
-({\textbf{???}}) and ({\textbf{???}}) showed that the connectivity of a given semantic category (i.e., animals) in a child's lexicon predicts a better performance in a disambiguation task, where participants are presented with a novel label in the presence of a familiar and a novel object: infants show stronger looking preference for the novel object if it belonged to a category for which many words had already been acquired.
-This points to the structure of the lexicon facilitating the strategies children engage during word learning.
+There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
+For example, Hoff et al.
+(2012) found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals (who only learnt English).
+In contrast, both groups knew a similar total amount of words, when both English and Spanish vocabularies were counted together.
+Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated (Fabian, 2016; Gonzalez-Barrero, Schott, \& Byers-Heinlein, 2020; Oller \& Eilers, 2002; J.
+L.
+Patterson, 2004; J.
+Patterson, Pearson, \& Goldstein, 2004; Pearson \& Fernández, 1994; Pearson, Fernández, \& Oller, 1993; Petitto et al., 2001; Smithson, Paradis, \& Nicoladis, 2014).
+While these studies have mostly relied on samples of bilingual children learning two relatively distant languages (as it is the case of English and Spanish) it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
+What role could linguistic distance play during early vocabulary growth?
 
-The case of bilingual children (here defined as those learning two languages from birth) presents an opportunity to study how lexical acquisition takes placed under a more complex environment, as they face the challenge of learning two distinct sets of words -- one for each language -- that partially overlap in sound and meaning.
+For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language.
+Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
+Particularly, when both languages are linguistically close, like Spanish and Catalan (both Roman languages), they are more likely to share a larger amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English--one Roman, the other Germanic.
+For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear \emph{puerta} and \emph{porta} (cognates), whereas a Spanish-English bilingual might hear \emph{puerta} and \emph{door} (non-cognates).
+It could be the case that mapping two phonologically similar labels (cognates, like \emph{puerta}-\emph{porta}) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like \emph{puerta} and \emph{door}).
+If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
 
-EARLY STAGES IN BILINGUAL LANGUAGE ACQUISITION
+A recent study by Floccia et al.
+(2018) provided evidence in line with this claim.
+The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers in the United Kingdom who were learning English and an additional language.
+The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
+The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English.
+Phonological similarity was measuring by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
+The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other (Levenshtein \& others, 1966).
+Floccia et al.~divided the resulting score by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
+Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
+For example, English-Dutch bilinguals (22.14\% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97\% phonological similarity) were able to produce in Mandarin.
 
-When compared to their monolingual peers, they know fewer words only when just on language is considered (e.g., English monolinguals know more words in English than English-Spanish bilinguals).
-When both languages are taken into account, bilingual children seem to know, at least, as many words as monolinguals do (Ben-Zeev, 1977; Bialystok, Luk, Peets, \& Yang, 2010; Blom et al., 2019; Core, Hoff, Rumiche, \& Señor, 2013; Doyle \& others, 1977; Fernandez, Pearson, Umbel, Oiler, \& Molinet-Molina, 1992; Hoff et al., 2012; Hoff, Rumiche, Burridge, Ribot, \& Welsh, 2014; Pearson \& Fernández, 1994; Rosenblum \& Pinker, 1983; but see Pearson, Fernández, \& Oller, 1993; Houwer, Bornstein, \& Putnick, 2014).
+Floccia et al.~pointed to \emph{parallel activation} as the main mechanism underpinning their results.
+The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension.
+This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production (Costa, Caramazza, \& Sebastian-Galles, 2000; Hoshino \& Kroll, 2008) or comprehension (Thierry \& Wu, 2007).
+One of the clearest examples of parallel activation was provided by Costa, Caramazza, and Sebastian-Galles (2000).
+In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish.
+In half of the trials, the object labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre}, translations of \emph{tree}), whereas in the other half of the trials labels were non-cognates (\emph{mesa}-\emph{taula}, translations of \emph{table}).
+Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
+Importantly, Spanish monolinguals did not show this effect.
+These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
+Several recent studies have also provided similar evidence in comprehension in children (Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013; e.g., Von Holzen \& Mani, 2012).
+Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition.
+Therefore, children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
+If cognates are the reason for this facilitation effect, it could be predicted that cognates are acquired earlier than non-cognates.
+Likely as it might seem, an earlier age of acquisition for cognates is not the only scenario where Floccia et al.'s results are account for.
+For example, languages that share more cognates might also share a larger overlap in their phonemic inventories, which could facilitate the acquisition of all words, regardless of their cognate status.
 
-Floccia et al.
-(2018), gathered parental vocabulary estimates ({\textbf{???}}) from a large sample of 24-month-old toddlers learning English and an additional language (from a pool of 13 linguistically diverse languages).
-They computed the average phonological similarity between translation equivalents (TEs from now on; e.g., \emph{table} in English and \emph{tafel} in Dutch) for each pair of languages, and found a positive association between this measure of phonological similarity and participants' productive vocabulary sizes in their additional language.
-Using a similar measure of lexical similarity, Blom et al.
-(2019) extended these results to bilingual children aged three to 10, and reported a positive association between the lexical distance and comprehensive vocabulary size: bilinguals learning lexically close languages showed similar vocabulary sizes to those of monolinguals, but those learning the more distant languages showed lower vocabulary sizes.
+However, the evidence on an earlier age of acquisition for cognates compared to non-cognates is scarce.
+To the best of our knowledge only two studies have provided direct data on this issue.
+First, Bosch and Ramon-Casas (2014) used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
+Second, Schelletter (2002) reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
+The low sample size in these studies make it challenging to draw strong conclusions about the effect of cognateness in vocabulary growth.
+Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
+In their discussion the authors state the following (pp.~70):
 
-These results suggest that linguistic distance between languages plays an important role during bilingual lexical acquisition, and that this effect operates within TEs.
-What mechanisms underlie this effect, and why they seem to operate differently across comprehension and production, or across the dominant and the non-dominant language, remain open issues.
-Floccia et al.
-(2018) pointed to parallel activation as a candidate mechanism behind this facilitation effect.
-The parallel activation principle suggests that bilinguals activate lexical representations in a language non-selective way: during the comprehension or production of a word in one language (e.g., \emph{cat}, in English) its corresponding translation in the other language is activated too {[}e.g., \emph{gato}, in Spanish.
-There is a vast body of evidence supporting language non-selective lexical access in both adults (e.g., Dell \& O'Seaghdha, 1991; {\textbf{???}}; Bobb, Von Holzen, Mayor, Mani, \& Carreiras, 2020; Dijkstra, 2005; Hoshino \& Kroll, 2008; Kroll, Gullifer, \& Rossi, 2013; Singh, 2014; Spivey \& Marian, 1999; Thierry \& Wu, 2007; Von Holzen, Fennell, \& Mani, 2019; Yudes, Macizo, \& Bajo, 2010; see Costa, Santesteban, \& Caño, 2005 for review), and children (e.g.~Von Holzen \& Mani, 2012; Bosma \& Nota, 2020; Jardak \& Byers-Heinlein, 2019; Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013).
-A critical implication of parallel activation is that the word form (i.e., phonology, orthography, etc.) of both translation equivalents are available to each other, impacting the comprehension or production dynamic of any of them.
+\begin{quote}
+``This finding also provides support to the proposal that the cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life.''
+\end{quote}
 
-Floccia et al.
-(2018) argue that the degree of phonological similarity between TEs (\emph{cognateness} from now on) of both languages should lead to increased parallel activation of both languages during language exposure, facilitating the acquisition of words in both languages, and leading participants leading two lexically close languages to show larger vocabulary sizes than those learning two lexically distant languages.
-This account, however, neglects the fact that for parallel activation to take place and play a facilitation role, lexical representations of both pairs of the TE must have been already established in the lexicon, and their corresponding phonological forms must have been encoded.
-This is not the case during lexical acquisition, where at least one of the members of the pair has not being acquired yet.
-A more lenient definition of parallel activation could adjust to the context of lexical acquisition more easily: when only one of the members of the TE has been acquired, its lexical representation is activated when, in the presence of its referent, a phonologically similar label is uttered.
-Consider the case of a Catalan-Spanish bilingual that has already learned the word \emph{gat} (Catalan for \emph{cat}), but not \emph{gato} (in Spanish).
-It is possible that, when presented with \emph{gato} in the presence of a cat, she maps this novel label to the familiar word \emph{gat}, and creates a lexical representation for \emph{gato} as a synonym for \emph{gat}.
-The case of a non-cognate is differing.
-Consider now the case of a toddler that has already learnt \emph{gos} (Catalan for \emph{dog}), but not \emph{perro} (Spanish).
-In this case, it would not be possible to map both labels via phonology, give their lack of phonological similarity.
+We identify two main reasons why we cannot conclude an earlier age of acquisition for cognates than for non cognate from Floccia et al.'s results.
+First, the response variable used was the proportion of words each participant understood and/or produced, from the list of lexical items in the vocabulary checklists.
+By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.~non-cognates was lost.
+Second, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at this age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
+In summary, the evidence supporting an earlier age of acquisition of cognates vs.~non-cognates presents some limitations that prevents drawing sound conclusions about this effect.
 
-Under this account, cognates (i.e., phonologically similar TEs) would be acquired earlier than non-cognates, but this effect would only play a role once one of the members of the TE has been acquired.
-The evidence supporting an earlier age of acquisition of cognates relative to non-cognates if sparse.
-Schelletter (2002)`s reported a longitudinal case-study of a German/English child who produced cognate TEs earlier than non-cognate TE.
-Bosch and Ramon-Casas (2014), showed converging results from a sample of 48 monolingual and bilingual infants aged 24 months.
-The low number of words included in the analyses of both studies, and the lack of adjustment for lexical frequency, limits severely the strength of any conclusions concerning the difference in age of acquisition of cognates and non-cognates.
-On the other hand Floccia et al.
-(2018) worked with aggregated estimates of participants' vocabulary size, losing item-level information that could provide information about change in age of acquisition of TEs associated with their degree of phonological similarity.
-Therefore, their results can only be considered as compatible with cognateness playing a role on age of acquisition.
-In this study, we aimed at overcoming these pitfalls by using unaggregated data from individual responses, testing the role of cognateness explicitly and adjusting for lexical frequency.
+Another worth-mentioning finding in Floccia et al.~is that the increase in vocabulary size associated with linguistic similarity was larger in the additional language vocabulary than in English vocabulary.
+Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
+Given that the relative amount of exposure to a given language is a strong predictor of children's vocabulary size in that language (e.g., Cattani et al., 2014; Thordardottir, Rothenberg, Rivard, \& Naves, 2006), participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
+If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
 
-It also follows from the fact that cognateness operates \emph{within} TEs that the effect of cognateness can only play a role once one of the members of the TE has been acquired, and its phonological form is available during the acquisition of its translation.
-Words that belong to the language of most exposure (dominant language) are likely to be acquired earlier than words in the non-dominant language (Floccia et al., 2018).
-({\textbf{???}}) tested Dutch-Frisian bilingual children aged 2.5 to four years in a comprehension task (PPVT-NL) that involved Frisian words with different degrees of phonological similarity with their Dutch TE.
-Infants with lower exposure to Frisian showed a better performance for words with high similarity, while no such benefit was found in children who were exposed mostly to Frisian.
-This suggests that infants that were mostly exposed to Dutch used the phonology of dutch words when processing Frisian words, as revealed by a better performance for cognates than for non-cognates.
-A study by ({\textbf{???}}) showed that the degree of phonological similarity between
-the performance of was better for cognates than for non-cognates, only for those children
+The aim of this study is to investigate the role of cognateness in bilingual word acquisition.
+We hypothesised that cognate words would be acquired earlier than non-cognate word, and that this difference would be larger in participants with lower exposure to the language such word belongs to.
+Using an designed online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of bilinguals aged 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.
 
-Finally, previous studies have reported that the properties that describe the form of a word play a stronger role in production than comprehension than in production.
-This is the case of the number of phonemes (Braginsky, Yurovsky, Marchman, \& Frank, 2019), phonological neighbourhood density (Jones \& Brandt, 2019).
-These results converge with Floccia et al.
-(2018)'s finding that the effect of the average cognateness between languages was larger in production than in comprehension\footnote{Although Bosch and Ramon-Casas (2014) found that participants in their sample produced cognates earlier than non-cognates, no analyses were conducted on comprehensive data, and therefore they cannot be taken as supporting evidence for a larger effect of cognateness on production than in comprehension}.
-Accordingly, we expected the role of cognateness to be more central in production than in comprehension.
-
-In summary, we investigated the effect of cognateness on the probability of acquisition of TEs, and predicted that (1) cognate TEs are acquired earlier than non-cognate TEs, that this effect will be larger in the dominant than in the non-dominant language, and larger in production than in comprehension.
+\hypertarget{references}{%
+\subsection{References}\label{references}}
 
@@
 
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-We collected data from 349 bilinguals (167 female), from the Metropolitan Area of Barcelona, between 28th October, 2019 and 09th January, 2021.
-All families gave informed consent before participating.
-This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language.
-We excluded participants with \textgreater10\% exposure to a third language.
-Fig.
-\ref{tab:lps} illustrates the distribution of participants across language profiles and ages.
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-\hypertarget{questionnaire}{%
-\subsection{Questionnaire}\label{questionnaire}}
-
-We implemented an on-line questionnaire using \texttt{formr} (Arslan, Walther, \& Tata, 2020), divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (Hamilton, Plunkett, \& Schafer, 2000) and consisted in two lists of words, one in Catalan and one in Spanish).
-Items in one language were translation equivalents of the items in the other (e.g., whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the word \emph{perro} was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar} and \emph{terminar}), or merged them into the same one (e.g., Spanish \emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}).
-We included items from a diverse sample of 26 semantic/functional categories (see Appendix 1).
-We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
-The Catalan inventory contained 778 items (196 cognates, 582 non-cognates) and the Spanish inventory contained 781 (197 cognates, 584 non-cognates).
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default).
-Participants filled a long or a short version of the questionnaire.
-Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items.
-Each list contained a different set of \textasciitilde400 translation equivalents (\textasciitilde400 in Catalan, \textasciitilde400 in Spanish).
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., \emph{barrita de cereales} {[}cereal bar{]}) and items that included more than one word-form (e.g., \emph{mono / mico}).
-Table \ref{tab:itemsinfo} shows the classification of items in cognates and non-cognates and their frequency scores across the four lists of the inventories.
-
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
-
-We gathered 181587 responses.
-Translation equivalents received an average of 386.36 (\emph{SD} = 228.76, , Max = 698), both languages summed together.
-We modelled the probability of responses (\(y_i\)) to belong to any of the three possible categories (\(k\)): 1 = \emph{No} or say, 2 = \emph{Understands}, and 3 = \emph{Understands and Says}) using an ordinal regression model with a cumulative logit link function (Agresti, 2010; Bürkner \& Vuorre, 2019).
-This model estimates the probability of response \(y_i\) to belong to category \(k\) or lower.
-In our case, this will result in the estimation of the probability of \(y_i \leq 1\), \(y_i \leq 2\), and \(y_i \leq 3\), where
-
-\[
-p_k = Pr(y_i = k) = Pr(y_i \leq k) - Pr(y_i \leq k-1).
-\]
-
-We include several predictors of interest to adjust this probability to participants' and items' (lexical frequency, dominance, cognateness) characteristics. These predictors are:
-
-\begin{itemize}
-\tightlist
-\item
-  Age of participant in months (\texttt{age}, \(Age\)) calculated as the difference in days between participants' birth date and questionnaire completion divided by 30 and centred around the mean
-\item
-  Degree of bilingualism of participant (\texttt{bilingualism}, \(Bilingualism\)) computed as the percentage of exposure to a second language (Spanish for participants exposed to \textgreater50\% to Catalan an \emph{vice versa}) rounded to units and centered around the mean
-\item
-  Lexical frequency of item (\texttt{frequency}, \(Frequency\)) retrieved from SUBTLEX-CAT (Boada, Guasch, Haro, Demestre, \& Ferré, 2019) for Catalan words and from SUBTLEX-ESP (Cuetos, Glez-Nosti, Barbón, \& Brysbaert, 2011) for Spanish words, expressed as Zipf scores (Heuven, Mandera, Keuleers, \& Brysbaert, 2014), and centred around the mean
-\item
-  Language the item belongs to (\texttt{dominance}, \(Dominance\)) labelled as \texttt{L1} if the item belongs to the dominant language (e.g., a Catalan item is labelled as \texttt{L1} for Catalan-dominant participants and as \texttt{L2} for Spanish-dominant participants), coded as \texttt{L2} = -0.5 and \texttt{L1} = 0.5
-\item
-  Cognateness of the item (\texttt{cognate}, \(Cognate\)), coded as \texttt{Non-cognate} = -0.5 and \texttt{Cognate} = 0.5
-\end{itemize}
-
-We adopted a Bayesian approach toward statistical inference, which allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model.
-Our extended model can be formalised as follows:
-
-\[
-\begin{aligned}
-y_i \sim Ordered(p) \\
-log\frac{Pr(y_i \leq k)}{1-Pr(y_i \leq k)} &= \alpha_k - \dots \\
-& \beta_1 \times Age_i + \dots \\
-& \beta_2 \times Frequency_i + \dots \\
-& \beta_3 \times Dominance_i + \dots \\
-& \beta_4 \times Bilingualism_i + \dots \\
-& \beta_5 \times Cognate_i + \dots \\
-& \beta_6 \times (Dominance_i \times Bilingualism_i) + \dots \\
-& \beta_7 \times (Dominance_i \times Cognate_i) + \dots \\
-& \beta_8 \times (Bilingualism_i \times Cognate_i) + \dots \\
-& \beta_9 \times (Dominance_i \times Bilingualism_i \times Cognate_i), \\
-\alpha \sim Normal(0, 10) \\
-\beta_{0-9} \sim Normal(0, 10)
-\end{aligned}
-\]
-
-Where:
-
-\begin{itemize}
-\tightlist
-\item
-  \(k\) is one of the possible categories (coded as ) responses can take
-\item
-  \(y_i\) is the \texttt{response} value of observation \(i\)
-\item
-  \(\alpha_k\) is the mean probability of responses to belong to category \(k\)
-\item
-  \(\beta_{1, \dots, 9}\) are the coefficients of the predictors
-\end{itemize}
-
-We implemented the model using the R package \texttt{brms} (Bürkner, 2017).
-This package uses the Stan probabilistic language (Carpenter et al., 2017) to approximate the posterior distribution of the model using Hamiltonian Monte Carlo.
-We implemented our extended model as \texttt{response\ \textasciitilde{}\ \ age\ +\ frequency\ +\ dominance*bilingualism*cognate}.
-We fit this model running four MCMC chains with 2,000 iterations and .80 delta parameter.
-We used weakly regularising prior for the parameters in our model, given that its weight in the estimation of the posterior will be marginal relative to that of the likelihood, due to the amount of observations introduced in the model.
-We then fit increasingly simpler models dropping predictors one by one in the following order: (1) \texttt{dominance:bilingualism:cognate}, (2) \texttt{bilingualism:cognate}, (3) \texttt{dominance:cognate:}, (4) \texttt{cognate}, (5) \texttt{dominance:bilingualism}, (6) \texttt{bilingualism}, (7) \texttt{dominance}, (8) \texttt{frequency}, and (9) \texttt{age}.
-We compared all models using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017) and widely applicable information criterion (WAIC) (Watanabe \& Opper, 2010).
-We then explored the posterior distribution of each parameter in the model that fitted the data the best, computing the 95\% credible intervals and testing whether this interval excluded 0.
-We performed follow-up tests on interactions whose 95\% credible interval excluded 0 by comparing the 95\% credible interval of their estimated marginal means using the \texttt{emmeans} package (Lenth, 2020).
-Data processing and visualisation was done in R (R Core Team, 2020) using the \texttt{tidyverse} family of packages (Wickham et al., 2019), and posterior samples were extracted using the \texttt{tidybayes} R package (Kay, 2020).
-
-\hypertarget{references}{%
-\section{References}\label{references}}
+\hypertarget{references-1}{%
+\section{References}\label{references-1}}
 
@@
 \hypertarget{refs}{}
-\begin{cslreferences}
-\leavevmode\hypertarget{ref-agresti_analysis_2010}{}%
-Agresti, A. (2010). \emph{Analysis of ordinal categorical data} (Vol. 656). John Wiley \& Sons.
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-bergelson2020comprehension}{}%
+Bergelson, E. (2020). The comprehension boost in early word learning: Older infants are better learners. \emph{Child Development Perspectives}, \emph{14}(3), 142--149.
 
-\leavevmode\hypertarget{ref-arslan_formr_2020}{}%
-Arslan, R.
-C., Walther, M.
-P., \& Tata, C.
-S.
-(2020).
-Formr: A study framework allowing for automated feedback generation and complex longitudinal experience-sampling studies using R.
-\emph{Behavior Research Methods}, \emph{52}(1), 376--387.
-\url{https://doi.org/10.3758/s13428-019-01236-y}
+\leavevmode\hypertarget{ref-bergelson2012months}{}%
+Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants know the meanings of many common nouns. \emph{Proceedings of the National Academy of Sciences}, \emph{109}(9), 3253--3258.
 
-\leavevmode\hypertarget{ref-ben-zeev_influence_1977}{}%
-Ben-Zeev, S. (1977). The Influence of Bilingualism on Cognitive Strategy and Cognitive Development. \emph{Child Development}, \emph{48}(3), 1009--1018. \url{https://doi.org/10.2307/1128353}
+\leavevmode\hypertarget{ref-bloom2002children}{}%
+Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT press.
 
-\leavevmode\hypertarget{ref-bialystok_receptive_2010}{}%
-Bialystok, E., Luk, G., Peets, K.
-F., \& Yang, S.
-(2010).
-Receptive vocabulary differences in monolingual and bilingual children.
-\emph{Bilingualism: Language and Cognition}, \emph{13}(4), 525--531.
-\url{https://doi.org/10.1017/S1366728909990423}
-
-\leavevmode\hypertarget{ref-blom_cross-language_2019}{}%
-Blom, E., Boerma, T., Bosma, E., Cornips, L., Heuij, K.
-van den, \& Timmermeister, M.
-(2019).
-Cross-language distance influences receptive vocabulary outcomes of bilingual children.
-\emph{First Language}, 0142723719892794.
-\url{https://doi.org/10.1177/0142723719892794}
-
-\leavevmode\hypertarget{ref-boada_subtlex-cat_2019}{}%
-Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P.
-(2019).
-SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for Catalan.
-\emph{Behavior Research Methods}.
-\url{https://doi.org/10.3758/s13428-019-01233-1}
-
-\leavevmode\hypertarget{ref-bobb_co-activation_2020}{}%
-Bobb, S.
-C., Von Holzen, K., Mayor, J., Mani, N., \& Carreiras, M.
-(2020).
-Co-activation of the L2 during L1 auditory processing: An ERP cross-modal priming study.
-\emph{Brain and Language}, \emph{203}, 104739.
-\url{https://doi.org/10.1016/j.bandl.2019.104739}
-
-\leavevmode\hypertarget{ref-bosch_first_2014}{}%
+\leavevmode\hypertarget{ref-bosch2014first}{}%
 Bosch, L., \& Ramon-Casas, M.
 (2014).
-First translation equivalents in bilingual toddlers' expressive vocabulary: Does form similarity matter?
+First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter?
 \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322.
@@
 
-\leavevmode\hypertarget{ref-bosma2020cognate}{}%
-Bosma, E., \& Nota, N. (2020). Cognate facilitation in frisian--dutch bilingual children's sentence reading: An eye-tracking study. \emph{Journal of Experimental Child Psychology}, \emph{189}, 104699.
+\leavevmode\hypertarget{ref-bosch2014first}{}%
+Bosch, L., \& Ramon-Casas, M.
+(2014).
+First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter?
+\emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322.
+\url{https://doi.org/10.1177/0165025414532559}
 
-\leavevmode\hypertarget{ref-braginsky_consistency_2019}{}%
-Braginsky, M., Yurovsky, D., Marchman, V.
-A., \& Frank, M.
-C.
-(2019).
-Consistency and Variability in Children's Word Learning Across Languages.
-\emph{Open Mind}, \emph{3}, 52--67.
-\url{https://doi.org/10.1162/opmi_a_00026}
+\leavevmode\hypertarget{ref-braginsky2019consistency}{}%
+Braginsky, M., Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and variability in children's word learning across languages. \emph{Open Mind}, \emph{3}, 52--67.
 
-\leavevmode\hypertarget{ref-burkner_brms_2017}{}%
-Bürkner, P.-C. (2017). Brms: An R Package for Bayesian Multilevel Models Using Stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+\leavevmode\hypertarget{ref-cattani2014much}{}%
+Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis, I., \& Floccia, C.
+(2014).
+How much exposure to english is necessary for a bilingual toddler to perform like a monolingual peer in language tests?
+\emph{International Journal of Language \& Communication Disorders}, \emph{49}(6), 649--671.
 
-\leavevmode\hypertarget{ref-burkner_ordinal_2019}{}%
-Bürkner, P.-C., \& Vuorre, M. (2019). Ordinal regression models in psychology: A tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 77--101.
+\leavevmode\hypertarget{ref-costa2000cognate}{}%
+Costa, A., Caramazza, A., \& Sebastian-Galles, N.
+(2000).
+The {Cognate} {Facilitation} {Effect}: {Implications} for {Models} of {Lexical} {Access}.
+\emph{Journal of Experimental Psychology}, \emph{26}(5), 1283--1296.
+\url{https://doi.org/10.1037/0278-7393.26.5.1283}
 
-\leavevmode\hypertarget{ref-carpenter_stan_2017}{}%
-Carpenter, B., Gelman, A., Hoffman, M.
-D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A.
-(2017).
-Stan: A Probabilistic Programming Language.
-\emph{Journal of Statistical Software}, \emph{76}(1), 1--32.
-\url{https://doi.org/10.18637/jss.v076.i01}
+\leavevmode\hypertarget{ref-costa2000cognate}{}%
+Costa, A., Caramazza, A., \& Sebastian-Galles, N.
+(2000).
+The {Cognate} {Facilitation} {Effect}: {Implications} for {Models} of {Lexical} {Access}.
+\emph{Journal of Experimental Psychology}, \emph{26}(5), 1283--1296.
+\url{https://doi.org/10.1037/0278-7393.26.5.1283}
 
-\leavevmode\hypertarget{ref-core_total_2013}{}%
-Core, C., Hoff, E., Rumiche, R., \& Señor, M.
-(2013).
-Total and Conceptual Vocabulary in Spanish--English Bilinguals From 22 to 30 Months: Implications for Assessment.
-\emph{Journal of Speech, Language, and Hearing Research}, \emph{56}(5), 1637--1649.
-\url{https://doi.org/10.1044/1092-4388(2013/11-0044)}
+\leavevmode\hypertarget{ref-fabian2016investigating}{}%
+Fabian, A. P. (2016). Investigating vocabulary abilities in bilingual portuguese-english-speaking children.
 
-\leavevmode\hypertarget{ref-costa2005facilitatory}{}%
-Costa, A., Santesteban, M., \& Caño, A. (2005). On the facilitatory effects of cognate words in bilingual speech production. \emph{Brain and Language}, \emph{94}(1), 94--103.
+\leavevmode\hypertarget{ref-fenson1994variability}{}%
+Fenson, L., Dale, P.
+S., Reznick, J.
+S., Bates, E., Thal, D.
+J., Pethick, S.
+J., \ldots{} Stiles, J.
+(1994).
+Variability in early communicative development.
+\emph{Monographs of the Society for Research in Child Development}, i--185.
 
-\leavevmode\hypertarget{ref-cuetos_subtlex-esp_2011}{}%
-Cuetos, F., Glez-Nosti, M., Barbón, A., \& Brysbaert, M. (2011). SUBTLEX-ESP: Spanish word frequencies based on film subtitles. \emph{Psicológica}, \emph{32}(2), 133--143.
+\leavevmode\hypertarget{ref-fenson1994variability}{}%
+Fenson, L., Dale, P.
+S., Reznick, J.
+S., Bates, E., Thal, D.
+J., Pethick, S.
+J., \ldots{} Stiles, J.
+(1994).
+Variability in early communicative development.
+\emph{Monographs of the Society for Research in Child Development}, i--185.
 
-\leavevmode\hypertarget{ref-dell1991mediated}{}%
-Dell, G. S., \& O'Seaghdha, P. G. (1991). Mediated and convergent lexical priming in language production: A comment on levelt et al (1991).
+\leavevmode\hypertarget{ref-floccia2018introduction}{}%
+Floccia, C., Sambrook, T.
+D., Delle Luche, C., Kwok, R., Goslin, J., White, L., \ldots{} others.
+(2018).
+I: INTRODUCTION.
+\emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29.
 
-\leavevmode\hypertarget{ref-dijkstra_bilingual_2005}{}%
-Dijkstra, T. (2005). Bilingual Visual Word Recognition and Lexical Access. In \emph{Handbook of bilingualism: Psycholinguistic approaches} (pp. 179--201). New York, NY, US: Oxford University Press.
+\leavevmode\hypertarget{ref-gonzalez2020bilingual}{}%
+Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020). Bilingual adjusted vocabulary: A developmentally-informed bilingual vocabulary measure.
 
-\leavevmode\hypertarget{ref-doyle1977some}{}%
-Doyle, A. B., \& others. (1977). Some issues in the assessment of linguistic consequences of early bilingualism. \emph{Working Papers on Bilingualism}, (14).
+\leavevmode\hypertarget{ref-hoff2012dual}{}%
+Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M. (2012). Dual language exposure and early bilingual development. \emph{Journal of Child Language}, \emph{39}(1), 1.
 
-\leavevmode\hypertarget{ref-fernandez1992bilingual}{}%
-Fernandez, M.
-C., Pearson, B.
-Z., Umbel, V.
-M., Oiler, D., \& Molinet-Molina, M.
-(1992).
-Bilingual receptive vocabulary in hispanic preschool children.
-\emph{Hispanic Journal of Behavioral Sciences}, \emph{14}(2), 268--276.
+\leavevmode\hypertarget{ref-hoshino2008cognate}{}%
+Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming: Does cross-language activation survive a change of script? \emph{Cognition}, \emph{106}(1), 501--511.
 
-\leavevmode\hypertarget{ref-floccia_i_2018}{}%
-Floccia, C., Sambrook, T.
-D., Luche, C.
-D., Kwok, R., Goslin, J., White, L., \ldots{} Plunkett, K.
-(2018).
-I: Introduction.
-\emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29.
-\url{https://doi.org/10.1111/mono.12348}
+\leavevmode\hypertarget{ref-jusczyk1995infants}{}%
+Jusczyk, P. W., \& Aslin, R. N. (1995). Infants′ detection of the sound patterns of words in fluent speech. \emph{Cognitive Psychology}, \emph{29}(1), 1--23.
 
-\leavevmode\hypertarget{ref-hamilton_infant_2000}{}%
-Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary development assessed with a british communicative development inventory. \emph{Journal of Child Language}, \emph{27}(3), 689--705.
+\leavevmode\hypertarget{ref-levenshtein1966binary}{}%
+Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. In \emph{Soviet physics doklady} (Vol. 10, pp. 707--710). Soviet Union.
 
-\leavevmode\hypertarget{ref-heuven_subtlex-uk_2014}{}%
-Heuven, W.
-J.
-B.
-van, Mandera, P., Keuleers, E., \& Brysbaert, M.
-(2014).
-Subtlex-UK: A New and Improved Word Frequency Database for British English: \emph{Quarterly Journal of Experimental Psychology}.
-Retrieved from \url{https://journals.sagepub.com/doi/10.1080/17470218.2013.850521}
+\leavevmode\hypertarget{ref-oller2002language}{}%
+Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in bilingual children} (Vol. 2). Multilingual Matters.
 
-\leavevmode\hypertarget{ref-hoff_dual_2012}{}%
-Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M.
-(2012).
-Dual language exposure and early bilingual development*.
-\emph{Journal of Child Language}, \emph{39}(1), 1--27.
-\url{https://doi.org/10.1017/S0305000910000759}
+\leavevmode\hypertarget{ref-patterson2004comparing}{}%
+Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers' expressive vocabulary size.
 
-\leavevmode\hypertarget{ref-hoff_expressive_2014}{}%
-Hoff, E., Rumiche, R., Burridge, A., Ribot, K.
-M., \& Welsh, S.
-N.
-(2014).
-Expressive vocabulary development in children from bilingual and monolingual homes: A longitudinal study from two to four years.
-\emph{Early Childhood Research Quarterly}, \emph{29}(4), 433--444.
-\url{https://doi.org/10.1016/j.ecresq.2014.04.012}
+\leavevmode\hypertarget{ref-patterson2004bilingual}{}%
+Patterson, J., Pearson, B., \& Goldstein, B. (2004). Bilingual language development and disorders in spanish--english speakers.
 
-\leavevmode\hypertarget{ref-hoshino_cognate_2008}{}%
-Hoshino, N., \& Kroll, J.
-F.
-(2008).
-Cognate effects in picture naming: Does cross-language activation survive a change of script?
-\emph{Cognition}, \emph{106}(1), 501--511.
-\url{https://doi.org/10.1016/j.cognition.2007.02.001}
+\leavevmode\hypertarget{ref-pearson1994patterns}{}%
+Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in the lexical growth in two languages of bilingual infants and toddlers. \emph{Language Learning}, \emph{44}(4), 617--653.
 
-\leavevmode\hypertarget{ref-houwer_bilingualmonolingual_2014}{}%
-Houwer, A.
-D., Bornstein, M.
-H., \& Putnick, D.
-L.
-(2014).
-A bilingual--monolingual comparison of young children's vocabulary size: Evidence from comprehension and production.
-\emph{Applied Psycholinguistics}, \emph{35}(6), 1189--1211.
-\url{https://doi.org/10.1017/S0142716412000744}
+\leavevmode\hypertarget{ref-pearson1993lexical}{}%
+Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical development in bilingual infants and toddlers: Comparison to monolingual norms. \emph{Language Learning}, \emph{43}(1), 93--120.
 
-\leavevmode\hypertarget{ref-jardak2019labels}{}%
-Jardak, A., \& Byers-Heinlein, K. (2019). Labels or concepts? The development of semantic networks in bilingual two-year-olds. \emph{Child Development}, \emph{90}(2), e212--e229.
-
-\leavevmode\hypertarget{ref-jones_children_2019}{}%
-Jones, S. D., \& Brandt, S. (2019). Do children really acquire dense neighbourhoods? \emph{Journal of Child Language}, \emph{46}(6), 1260--1273. \url{https://doi.org/10.1017/S0305000919000473}
-
-\leavevmode\hypertarget{ref-kay_tidybayes_2020}{}%
-Kay, M. (2020). \emph{tidybayes: Tidy data and geoms for Bayesian models}. \url{https://doi.org/10.5281/zenodo.1308151}
-
-\leavevmode\hypertarget{ref-kroll_multilingual_2013}{}%
-Kroll, J.
-F., Gullifer, J.
-W., \& Rossi, E.
-(2013).
-The Multilingual Lexicon: The Cognitive and Neural Basis of Lexical Comprehension and Production in Two or More Languages.
-\emph{Annual Review of Applied Linguistics}, \emph{33}, 102--127.
-\url{https://doi.org/10.1017/S0267190513000111}
-
-\leavevmode\hypertarget{ref-lenth_emmeans_2020}{}%
-Lenth, R. V. (2020). \emph{Emmeans: Estimated marginal means, aka least-squares means}. Retrieved from \url{https://CRAN.R-project.org/package=emmeans}
-
-\leavevmode\hypertarget{ref-pearson_patterns_1994}{}%
-Pearson, B.
-Z., \& Fernández, S.
-C.
-(1994).
-Patterns of Interaction in the Lexical Growth in Two Languages of Bilingual Infants and Toddlers.
-\emph{Language Learning}, \emph{44}(4), 617--653.
-\url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}
-
-\leavevmode\hypertarget{ref-pearson_lexical_1993}{}%
-Pearson, B.
-Z., Fernández, S.
-C., \& Oller, D.
-K.
-(1993).
-Lexical Development in Bilingual Infants and Toddlers: Comparison to Monolingual Norms.
-\emph{Language Learning}, \emph{43}(1), 93--120.
-\url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}
+\leavevmode\hypertarget{ref-petitto2001bilingual}{}%
+Petitto, L.
+A., Katerelos, M., Levy, B.
+G., Gauna, K., Tétreault, K., \& Ferraro, V.
+(2001).
+Bilingual signed and spoken language acquisition from birth: Implications for the mechanisms underlying early bilingual language acquisition.
+\emph{Journal of Child Language}, \emph{28}(2), 453.
 
@@
 
-\leavevmode\hypertarget{ref-rcoreteam_r_2020}{}%
-R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
+\leavevmode\hypertarget{ref-schelletter2002effect}{}%
+Schelletter, C.
+(2002).
+The effect of form similarity on bilingual children's lexical development.
+\emph{Bilingualism: Language and Cognition}, \emph{5}(2), 93--107.
+\url{https://doi.org/10.1017/S1366728902000214}
 
-\leavevmode\hypertarget{ref-rosenblum1983word}{}%
-Rosenblum, T., \& Pinker, S.
-A.
-(1983).
-Word magic revisited: Monolingual and bilingual children's understanding of the word-object relationship.
-\emph{Child Development}, 773--780.
-\url{https://doi.org/10.2307/1130064}
+\leavevmode\hypertarget{ref-smithson2014bilingualism}{}%
+Smithson, L., Paradis, J., \& Nicoladis, E.
+(2014).
+Bilingualism and receptive vocabulary achievement: Could sociocultural context make a difference.
+\emph{Bilingualism: Language and Cognition}, \emph{17}(4), 810--821.
 
-\leavevmode\hypertarget{ref-schelletter2002effect}{}%
-Schelletter, C. (2002). The effect of form similarity on bilingual children's lexical development. \emph{Bilingualism: Language and Cognition}, \emph{5}(2), 93--107.
-
-\leavevmode\hypertarget{ref-singh2014one}{}%
-Singh, L. (2014). One world, two languages: Cross-language semantic priming in bilingual toddlers. \emph{Child Development}, \emph{85}(2), 755--766.
-
-\leavevmode\hypertarget{ref-spivey1999cross}{}%
-Spivey, M. J., \& Marian, V. (1999). Cross talk between native and second languages: Partial activation of an irrelevant lexicon. \emph{Psychological Science}, \emph{10}(3), 281--284.
-
-\leavevmode\hypertarget{ref-thierry_brain_2007}{}%
+\leavevmode\hypertarget{ref-thierry2007brain}{}%
 Thierry, G., \& Wu, Y.
@@
 \emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535.
-\url{https://doi.org/10.1073/pnas.0609927104}
 
-\leavevmode\hypertarget{ref-vehtari_practical_2017}{}%
-Vehtari, A., Gelman, A., \& Gabry, J.
-(2017).
-Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC.
-\emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
-\url{https://doi.org/10.1007/s11222-016-9696-4}
+\leavevmode\hypertarget{ref-thordardottir2006bilingual}{}%
+Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R.
+(2006).
+Bilingual assessment: Can overall proficiency be estimated from separate measurement of two languages?
+\emph{Journal of Multilingual Communication Disorders}, \emph{4}(1), 1--21.
 
-\leavevmode\hypertarget{ref-von_holzen_impact_2019}{}%
-Von Holzen, K., Fennell, C.
-T., \& Mani, N.
-(2019).
-The impact of cross-language phonological overlap on bilingual and monolingual toddlers' word recognition.
-\emph{Bilingualism: Language and Cognition}, \emph{22}(3), 476--499.
-\url{https://doi.org/10.1017/S1366728918000597}
+\leavevmode\hypertarget{ref-thordardottir2006bilingual}{}%
+Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R.
+(2006).
+Bilingual assessment: Can overall proficiency be estimated from separate measurement of two languages?
+\emph{Journal of Multilingual Communication Disorders}, \emph{4}(1), 1--21.
 
-\leavevmode\hypertarget{ref-von_holzen_language_2012}{}%
-Von Holzen, K., \& Mani, N.
-(2012).
-Language nonselective lexical access in bilingual toddlers.
-\emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.
-\url{https://doi.org/10.1016/j.jecp.2012.08.001}
+\leavevmode\hypertarget{ref-tincoff1999some}{}%
+Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word comprehension in 6-month-olds. \emph{Psychological Science}, \emph{10}(2), 172--175.
 
-\leavevmode\hypertarget{ref-watanabe_asymptotic_2010}{}%
-Watanabe, S., \& Opper, M.
-(2010).
-Asymptotic equivalence of bayes cross validation and widely applicable information criterion in singular learning theory.
-\emph{Journal of Machine Learning Research}, \emph{11}(12).
+\leavevmode\hypertarget{ref-von2012language}{}%
+Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.
 
-\leavevmode\hypertarget{ref-wickham_welcome_2019}{}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
-D., François, R., \ldots{} Yutani, H.
-(2019).
-Welcome to the \textless span class="nocase"\textgreater tidyverse\textless/span\textgreater.
-\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
-\url{https://doi.org/10.21105/joss.01686}
+\leavevmode\hypertarget{ref-von2012language}{}%
+Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.
 
-\leavevmode\hypertarget{ref-yudes_cognate_2010}{}%
-Yudes, C., Macizo, P., \& Bajo, T. (2010). Cognate effects in bilingual language comprehension tasks. \emph{NeuroReport}, \emph{21}(7), 507--512. \url{https://doi.org/10.1097/WNR.0b013e328338b9e1}
-\end{cslreferences}
+\end{CSLReferences}
 

### 2021-09-15 Rmd/report.Rmd
Restore branch

@@
 # load packages
+library(targets)
 library(tidyverse)
@@
 # set params
-theme_custom <- function(){
-  theme_minimal() +
-    theme(
-      panel.grid = element_line(colour = "grey", linetype = "dotted"),
-      axis.line = element_line(colour = "black"),
-      text = element_text(size = 12, colour = "black"),
-      axis.text = element_text(colour = "black")
-    )
-}
-
 theme_set(theme_custom())
@@
 # import data
-items <- readRDS(here("Data", "items.rds"))
-participants <- readRDS(here("Data", "participants.rds"))
-responses <- readRDS(here("Data", "responses.rds"))
-
-# import models
-vocabulary <- readRDS(here("Data", "vocabulary.rds"))
-vocabulary_long <- readRDS(here("Data", "vocabulary_long.rds"))
-fit_prior <- readRDS(here("Results", "irt_2_prior.rds"))
-fit <- readRDS(here("Results", "irt_2.rds"))
+tar_load(items)
+tar_load(participants)
+tar_load(vocabulary)
+tar_load(responses)
+tar_load(fits)
 
@@
     legend.title = element_blank()
-  ) &
-  ggsave(here("Figures", "vocabulary_comprehension.png"))
+  )
 
@@
     legend.title = element_blank()
-  ) &
-  ggsave(here("Figures", "vocabulary_production.png"))
+  )
 
@@
     axis.ticks.y = element_blank()
-  ) +
-  ggsave(here("Figures", "vocabulary_long_comprehension.png"))
+  )
 ```
@@
     axis.ticks.y = element_blank()
-  ) +
-  ggsave(here("Figures", "vocabulary_long_production.png"))
+  )
 ```
@@
 ### {-}
+
+## Raw data
+
+```{r raw, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses %>%
+  group_by(te, cognate) %>%
+  summarise(
+    understands = mean(understands, na.rm = TRUE),
+    produces = mean(produces, na.rm = TRUE)
+  ) %>%
+  pivot_longer(
+    c(understands, produces),
+    names_to = "type",
+    values_to = "prop"
+  ) %>%
+  mutate(type = str_to_sentence(type)) %>% 
+  ggplot(aes(prop, reorder(te, -prop), colour = cognate)) +
+  facet_wrap(vars(type)) +
+  geom_vline(xintercept = 0.5, linetype = "dotted", size = 1) +
+  geom_point(alpha = 0.5, size = 2) +
+  labs(x = "Mean % of acquisition", y = "Translation equivalent ID", colour = "Type") +
+  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "top",
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank()
+  )
+```
+
+```{r raw_age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses %>%
+  mutate(age = as.factor(round(age))) %>% 
+  group_by(te, age, cognate) %>%
+  summarise(
+    understands = mean(understands, na.rm = TRUE),
+    produces = mean(produces, na.rm = TRUE)
+  ) %>%
+  pivot_longer(
+    c(understands, produces),
+    names_to = "type",
+    values_to = "prop"
+  ) %>%
+  mutate(type = str_to_sentence(type)) %>% 
+  ggplot(aes(age, prop, colour = cognate, fill = cognate, shape = cognate)) +
+  facet_wrap(vars(type)) +
+  geom_hline(yintercept = 0.5, linetype = "dotted", size = 1) +
+  geom_point(aes(group = interaction(age, cognate)), alpha = 0.5, size = 0.5) +
+  stat_summary(fun.data = mean_se, geom = "pointrange") +
+  labs(x = "Mean % of acquisition", y = "Translation equivalent ID",
+       colour = "Cognate", shape = "Cognate", fill = "Cognate") +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "top",
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    panel.grid = element_blank(),
+    legend.title = element_blank()
+  )
+```
 
@@
 
-m <- add_fitted_draws(nd, fit_prior, n = 20, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, fits$fit_2, n = 20, re_formula = NA) %>% 
   mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))
@@
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "responses_fixed.png"))
+  )
 ```

### 2021-09-15 R/03_responses.R

@@ -1,0 +1,44 @@
+# responses
+
+get_responses <- function(
+    participants,
+    items,
+    multilex_data
+){
+    responses <- expand_grid(
+        id = participants$id,
+        item = items$item
+    ) %>%
+        left_join(
+            participants %>% 
+                pivot_longer(
+                    c(doe_catalan, doe_spanish),
+                    names_to = "language",
+                    values_to = "doe"
+                ) %>% 
+                mutate(language = str_to_sentence(str_remove(language, "doe_"))) %>% 
+                select(id, time_stamp, age, lp, dominant_language, doe)
+        ) %>% 
+        left_join(items) %>% 
+        left_join(select(multilex_data$responses, id, item, response))  %>% 
+        drop_na() %>%
+        distinct(id, te, age, item, .keep_all = TRUE) %>% 
+        mutate(
+            doe = scale(doe, scale = FALSE)[,1],
+            age = scale(age, scale = FALSE)[,1],
+            understands = response>1,
+            produces = response>2,
+            dominance = ifelse(dominant_language==language, "L1", "L2")
+        ) %>% 
+        mutate_at(vars(cognate, lp), as.factor) %>%
+        mutate(frequency_center = scale(frequency)[,1]) %>% 
+        select(id, age, lp, te, language, item, frequency, frequency_center, doe, cognate, understands, produces) %>% 
+        arrange(id, te)
+    
+    # contrasts
+    contrasts(responses$cognate) <- cbind(c(0.25, -0.5, 0.25), c(0.5, 0, -0.5))
+    contrasts(responses$lp) <- c(0.5, -0.5)
+    
+    return(responses)
+    
+}

### 2021-09-15 R/utils.R

@@ -1,0 +1,77 @@
+# utils
+
+# custom ggplot theme
+theme_custom <- function(){
+    theme_minimal() +
+        theme(
+            panel.grid = element_line(colour = "grey", linetype = "dotted"),
+            axis.line = element_line(colour = "black"),
+            text = element_text(size = 12, colour = "black"),
+            axis.text = element_text(colour = "black")
+        )
+}
+
+
+# get multilex data ----
+get_credentials <- function(
+    google_email = "gonzalo.garciadecastro@upf.edu",
+    formr_email = "gonzalo.garciadecastro@upf.edu"
+){
+    ml_connect(
+        google_email,
+        formr_email,
+        formr_password = key_get("formr", "gonzalo.garciadecastro@upf.edu")
+    )
+}
+
+get_multilex <- function(
+    update = FALSE,
+    type = "understands"
+){
+    get_credentials()
+    
+    p <- ml_participants()
+    r <- ml_responses(p, update = update)
+    l <- ml_logs(p, r) %>% 
+        mutate(
+            age_group =  as.factor(
+                case_when(
+                    between(age, 19, 24) ~ 21,
+                    between(age, 24, 28) ~ 25,
+                    between(age, 28, 34) ~ 30
+                )),
+            age_group = paste0(age_group, " months")
+        ) %>% 
+        filter(age_group != "NA months") %>% 
+        select(id, time, time_stamp, age_group, age, lp, dominance, version,
+               completed, doe_catalan, doe_spanish, doe_others)
+    
+    v <- ml_vocabulary(p, r, scale = "prop", by = "id_exp") %>% 
+        filter(type==type)
+    
+    m <- list(
+        participants = p, responses = r, logs = l,
+        vocabulary = v, pool = multilex::pool
+    )
+    
+    return(m)
+}
+
+# adjusted proportion, SE, and CI ----
+# from Gelman, Hill & Vehtari (2020)
+prop_adj <- function(y, n) (y+2)/(n+4)
+prop_se_adj <- function(y, n) {
+    prop <- prop_adj(y, n)
+    sqrt(prop*(1-prop)/(n+4))
+}
+prop_ci_adj <- function(y, n, conf = 0.95){
+    prop <- prop_adj(y, n)
+    se <- sqrt(prop*(1-prop)/(n+4))
[…]

### 2021-09-15 Rmd/report.Rmd
mode report to Markdown

@@
 output:
-  html_document:
-    df_print: paged
-    toc: true
-    toc_depth: 4
-    toc_float: true
+  github_document:
+    pandoc_args: --webtex
 bibliography: "trajectories.bib"
@@
 knitr::opts_chunk$set(
-  echo = FALSE,
   message = TRUE,
+  echo = TRUE,
   warning = TRUE,
-  cache.extra = knitr::rand_seed,
+  dev.args = list(png = list(type = "cairo")),
+  include = FALSE,
   out.width = "100%",
   results = "asis",
-  dev.args = list(png = list(type = "cairo"))
+  dpi = 300
 )
-options(knitr.kable.NA = '-',
+options(
+  knitr.kable.NA = "-",
+  knitr.duplicate.label = "allow",
   ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
@@
 library(targets)
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(stringr)
 library(gt)
@@
 library(patchwork)
+library(tibble)
 library(janitor)
@@
 # set params
+tar_load_globals()
 theme_set(theme_custom())
@@
 labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
-participants %>% 
+a <- participants %>% 
   mutate(
@@
 
-vocabulary %>% 
-  filter(type=="Understands", modality != "Te") %>% 
-  ggplot(aes(age, prop, colour = lp, fill = lp)) +
+d <- vocabulary$wide %>% 
+  filter(type=="Understands", modality != "Te")
+
+ggplot(d, aes(age, prop, colour = lp, fill = lp)) +
   facet_wrap(~modality) +
@@
 
-  vocabulary %>% 
-  filter(type=="Understands", modality != "Te") %>% 
-  ggplot(aes(age, count, colour = lp, fill = lp)) +
+  ggplot(d, aes(age, count, colour = lp, fill = lp)) +
   facet_wrap(~modality) +
@@
 
-vocabulary %>% 
+vocabulary$wide %>% 
   filter(type=="Produces", modality != "Te") %>% 
@@
 
-  vocabulary %>% 
-  filter(type=="Understands", modality != "Te") %>% 
+  vocabulary$wide %>% 
+  filter(type=="Produces", modality != "Te") %>% 
   ggplot(aes(age, count, colour = lp, fill = lp)) +
@@
 ```{r vocabulary_long_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-vocabulary_long %>% 
+vocabulary$long %>% 
   filter(type=="understands") %>% 
@@
 ```{r vocabulary_long_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-vocabulary_long %>% 
+vocabulary$long %>% 
   filter(type=="produces") %>% 
@@
   age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = c("Monolingual", "Bilingual"),
-  cognate = c("Cognate", "Non-Cognate"),
+  lp = unique(participants$lp),
+  cognate = unique(items$cognate),
   frequency_center = c(-1, 0, 1)
@@
 ```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-summary(fit)$fixed %>% 
+summary(fits$fit_2)$fixed %>% 
   as.data.frame() %>%
@@
 
-post <- gather_draws(fit, `b_.*`, regex = TRUE) %>% 
+post <- gather_draws(fits$fit_2, `b_.*`, regex = TRUE) %>% 
   mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))
@@
 
-re_id <- ranef(fit)$id[,,1] %>% 
+re_id <- ranef(fits$fit_2)$id[,,1] %>% 
   as.data.frame() %>% 
@@
   as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fit)[1])) %>% 
-  left_join(distinct(fit$data, id, lp))
+  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
+  left_join(distinct(fits$fit_2$data, id, lp))
 
@@
 
-re_te <- ranef(fit)$te[,,1] %>% 
+re_te <- ranef(fits$fit_2)$te[,,1] %>% 
   as.data.frame() %>% 
@@
   as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fit)[1])) %>% 
+  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
   mutate(te = as.numeric(te)) %>% 
-  left_join(distinct(fit$data, te, lp))
+  left_join(distinct(fits$fit_2$data, te, lp))
 
@@
   age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = c("Monolingual", "Bilingual"),
-  cognate = c("Cognate", "Non-Cognate"),
+  lp = unique(participants$lp),
+  cognate =unique(items$cognate),
   frequency_center = c(-1,  0, 1)
@@
 
-m <- add_fitted_draws(nd, fit, re_formula = NA, n = 25) %>% 
+m <- add_fitted_draws(nd, fits$fit_2, re_formula = NA, n = 25) %>% 
   mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))
@@
 
-gather_draws(fit, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
+gather_draws(fits$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
   mutate(.chain = paste("Chain ", .chain)) %>% 

### 2021-09-15 Rmd/trajectories_manuscript.tex
mode report to Markdown

@@
   man,man,floatsintext]{apa6}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
   \usepackage{textcomp} % provide euro and other symbols
-  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \usepackage{csquotes}
-\ifxetex
+\ifXeTeX
   % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
@@
 \else
-  \usepackage[shorthands=off,main=english]{babel}
+  \usepackage[main=english]{babel}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 \fi
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
@@
 This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production (Costa, Caramazza, \& Sebastian-Galles, 2000; Hoshino \& Kroll, 2008) or comprehension (Thierry \& Wu, 2007).
-One of the clearest examples of parallel activation was provided by Costa, Caramazza, and Sebastian-Galles (2000).
+One of the clearest examples of parallel activation was provided by Costa et al.
+(2000).
 In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish.
@@
 
-\hypertarget{references}{%
-\subsection{References}\label{references}}
-
 \hypertarget{method}{%
@@
 
-\hypertarget{references-1}{%
-\section{References}\label{references-1}}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-bergelson2020comprehension}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bergelson2020comprehension}{}}%
 Bergelson, E. (2020). The comprehension boost in early word learning: Older infants are better learners. \emph{Child Development Perspectives}, \emph{14}(3), 142--149.
 
-\leavevmode\hypertarget{ref-bergelson2012months}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bergelson2012months}{}}%
 Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants know the meanings of many common nouns. \emph{Proceedings of the National Academy of Sciences}, \emph{109}(9), 3253--3258.
 
-\leavevmode\hypertarget{ref-bloom2002children}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bloom2002children}{}}%
 Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT press.
 
-\leavevmode\hypertarget{ref-bosch2014first}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bosch2014first}{}}%
 Bosch, L., \& Ramon-Casas, M.
@@
 
-\leavevmode\hypertarget{ref-bosch2014first}{}%
-Bosch, L., \& Ramon-Casas, M.
-(2014).
-First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter?
-\emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322.
-\url{https://doi.org/10.1177/0165025414532559}
-
-\leavevmode\hypertarget{ref-braginsky2019consistency}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-braginsky2019consistency}{}}%
 Braginsky, M., Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and variability in children's word learning across languages. \emph{Open Mind}, \emph{3}, 52--67.
 
-\leavevmode\hypertarget{ref-cattani2014much}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cattani2014much}{}}%
 Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis, I., \& Floccia, C.
@@
 
-\leavevmode\hypertarget{ref-costa2000cognate}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
 Costa, A., Caramazza, A., \& Sebastian-Galles, N.
@@
 
-\leavevmode\hypertarget{ref-costa2000cognate}{}%
-Costa, A., Caramazza, A., \& Sebastian-Galles, N.
-(2000).
-The {Cognate} {Facilitation} {Effect}: {Implications} for {Models} of {Lexical} {Access}.
-\emph{Journal of Experimental Psychology}, \emph{26}(5), 1283--1296.
-\url{https://doi.org/10.1037/0278-7393.26.5.1283}
-
-\leavevmode\hypertarget{ref-fabian2016investigating}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fabian2016investigating}{}}%
 Fabian, A. P. (2016). Investigating vocabulary abilities in bilingual portuguese-english-speaking children.
 
-\leavevmode\hypertarget{ref-fenson1994variability}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fenson1994variability}{}}%
 Fenson, L., Dale, P.
@@
 
-\leavevmode\hypertarget{ref-fenson1994variability}{}%
-Fenson, L., Dale, P.
-S., Reznick, J.
-S., Bates, E., Thal, D.
-J., Pethick, S.
-J., \ldots{} Stiles, J.
-(1994).
-Variability in early communicative development.
-\emph{Monographs of the Society for Research in Child Development}, i--185.
-
-\leavevmode\hypertarget{ref-floccia2018introduction}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
 Floccia, C., Sambrook, T.
@@
 
-\leavevmode\hypertarget{ref-gonzalez2020bilingual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gonzalez2020bilingual}{}}%
 Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020). Bilingual adjusted vocabulary: A developmentally-informed bilingual vocabulary measure.
 
-\leavevmode\hypertarget{ref-hoff2012dual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-hoff2012dual}{}}%
 Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M. (2012). Dual language exposure and early bilingual development. \emph{Journal of Child Language}, \emph{39}(1), 1.
 
-\leavevmode\hypertarget{ref-hoshino2008cognate}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-hoshino2008cognate}{}}%
 Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming: Does cross-language activation survive a change of script? \emph{Cognition}, \emph{106}(1), 501--511.
 
-\leavevmode\hypertarget{ref-jusczyk1995infants}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-jusczyk1995infants}{}}%
 Jusczyk, P. W., \& Aslin, R. N. (1995). Infants′ detection of the sound patterns of words in fluent speech. \emph{Cognitive Psychology}, \emph{29}(1), 1--23.
 
-\leavevmode\hypertarget{ref-levenshtein1966binary}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
 Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. In \emph{Soviet physics doklady} (Vol. 10, pp. 707--710). Soviet Union.
 
-\leavevmode\hypertarget{ref-oller2002language}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-oller2002language}{}}%
 Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in bilingual children} (Vol. 2). Multilingual Matters.
 
-\leavevmode\hypertarget{ref-patterson2004comparing}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-patterson2004comparing}{}}%
 Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers' expressive vocabulary size.
 
-\leavevmode\hypertarget{ref-patterson2004bilingual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-patterson2004bilingual}{}}%
 Patterson, J., Pearson, B., \& Goldstein, B. (2004). Bilingual language development and disorders in spanish--english speakers.
 
-\leavevmode\hypertarget{ref-pearson1994patterns}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pearson1994patterns}{}}%
 Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in the lexical growth in two languages of bilingual infants and toddlers. \emph{Language Learning}, \emph{44}(4), 617--653.
 
-\leavevmode\hypertarget{ref-pearson1993lexical}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pearson1993lexical}{}}%
 Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical development in bilingual infants and toddlers: Comparison to monolingual norms. \emph{Language Learning}, \emph{43}(1), 93--120.
 
-\leavevmode\hypertarget{ref-petitto2001bilingual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-petitto2001bilingual}{}}%
 Petitto, L.
@@
 
-\leavevmode\hypertarget{ref-poulin2013lexical}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-poulin2013lexical}{}}%
 Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J.
@@
 
-\leavevmode\hypertarget{ref-schelletter2002effect}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-schelletter2002effect}{}}%
 Schelletter, C.
@@
 
-\leavevmode\hypertarget{ref-smithson2014bilingualism}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-smithson2014bilingualism}{}}%
 Smithson, L., Paradis, J., \& Nicoladis, E.
@@
 
-\leavevmode\hypertarget{ref-thierry2007brain}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-thierry2007brain}{}}%
 Thierry, G., \& Wu, Y.
@@
 
-\leavevmode\hypertarget{ref-thordardottir2006bilingual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-thordardottir2006bilingual}{}}%
 Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R.
@@
 
-\leavevmode\hypertarget{ref-thordardottir2006bilingual}{}%
-Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R.
-(2006).
-Bilingual assessment: Can overall proficiency be estimated from separate measurement of two languages?
-\emph{Journal of Multilingual Communication Disorders}, \emph{4}(1), 1--21.
-
-\leavevmode\hypertarget{ref-tincoff1999some}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999some}{}}%
 Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word comprehension in 6-month-olds. \emph{Psychological Science}, \emph{10}(2), 172--175.
 
-\leavevmode\hypertarget{ref-von2012language}{}%
-Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.
-
-\leavevmode\hypertarget{ref-von2012language}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-von2012language}{}}%
 Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.

### 2021-09-15 R/03_irt.R
mode report to Markdown

@@ -1,90 +1,0 @@
-# IRT
-
-# set up ----
-library(tidyverse)
-library(brms) # for Bayesian models
-library(tidybayes) # for posterior draws
-library(job) # for running code in the background
-library(here) # for reproducible file paths
-
-# set params
-options(mc.cores = 4)
-
-# import data ----
-responses <- readRDS(here("Data", "responses.rds"))
-
-# models ----
-# set contrasts
-contrasts(responses$cognate) <- c(0.5, -0.5)
-contrasts(responses$lp) <- c(0.5, -0.5)
-
-# formula
-formulas <- list(
-    f_0 = bf(understands ~ 1 + age + frequency_center + (1 + age | te) + (1 + frequency_center | id), family = bernoulli("logit")),
-    f_1 = bf(understands ~ 1 + age + frequency_center + lp + (1 + age + lp | te) + (1 + frequency_center | id), family =  bernoulli("logit")),
-    f_2 = bf(understands ~ 1 + age + frequency_center + lp*cognate + (1 + age + lp | te) + (1 + frequency_center + cognate | id), family =  bernoulli("logit"))
-)
-
-# priors (derived from prior samples)
-priors <- c(
-    # model 0
-    prior(normal(0, 0.1), class = "Intercept"),
-    prior(normal(0.75, 0.1), class = "b", coef = "age"),
-    prior(normal(0, 0.1), class = "b", coef = "frequency_center"),
-    prior(normal(0.2, 0.1), class = "sd", group = "te"),
-    prior(constant(0.2), class = "sd", group = "id"),
-    # model 1
-    prior(lkj(7), class = "cor"),
-    prior(normal(0, 0.1), class = "b", coef = "lp1"),
-    # model 2
-    prior(normal(0, 0.1), class = "b", coef = "cognate1"),
-    prior(normal(0, 0.1), class = "b", coef = "lp1:cognate1")
-)
-
-# fit model
-job(
-    title = "Fit models",
-    fits = {
-        options(mc.cores = 4)
-        
-        fit_0 = brm(
-            formula = formulas$f_0, data = responses, prior = priors[1:6,],
-            chains = 4, iter = 2000, inits = 0,
-            backend = "cmdstanr", save_pars = save_pars(all = TRUE),
-            save_model = here("Stan", "irt_0.stan"), file = here("Results", "irt_0.rds")
-        ) 
-        
-        fit_1 = brm(
-            formula = formulas$f_1, data = responses, prior = priors[1:7,],
[…]

### 2021-09-15 Rmd/report.Rmd
fix gt tables

@@
 
-```{r items_list, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=4, out.width="100%"}
+```{r items_list, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=4, out.width="100%"}
 items %>% 
@@
   fmt_number(columns = starts_with("frequency")) %>% 
-  div(style = "height:600px; overflow-x: scroll; overflow-y: scroll", .)
+  div(style = "height:600px; overflow-x: scroll; overflow-y: scroll", .) %>% 
+  as_raw_html()
 ```
@@
     age = md("**Age<br>(months)**")
-  )
+  ) %>% 
+  as_raw_html()
 
@@
     columns = 2:4
-  )
+  ) %>% 
+  as_raw_html()
 

### 2021-09-15 Rmd/report.Rmd
mode report to Markdown

@@
   fmt_number(columns = starts_with("frequency")) %>% 
-  div(style = "height:600px; overflow-x: scroll; overflow-y: scroll", .) %>% 
-  as_raw_html()
+  as_raw_html() %>% 
+  div(style = "height:600px; overflow-x: scroll; overflow-y: scroll", .)
 ```

### 2021-09-15 Rmd/report.Rmd
Fix figures

@@
   dev.args = list(png = list(type = "cairo")),
-  include = FALSE,
+  include = TRUE,
   out.width = "100%",
@@
 
-```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r items_frequency, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-```{r items_cognate, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r items_cognate, message=FALSE, warning=FALSE, paged.print=FALSE}
 by_cognate <- items %>% 
@@
 
-```{r participants_time, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
 participants %>% 
@@
 
-```{r participants_age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants_age, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-```{r participants_lp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE}
 labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
@@
 \begin{aligned}
+\color{white}
 \log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
@@
 
-```{stan stan_model, echo=TRUE, output.var="model"}
+```{stan stan_model, output.var="model"}
 // generated with brms 2.15.0
@@
 
-```{r vocabulary_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r vocabulary_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-```{r vocabulary_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r vocabulary_production, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-```{r vocabulary_long_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r vocabulary_long_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}
 vocabulary$long %>% 
@@
 
-```{r vocabulary_long_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r vocabulary_long_production, message=FALSE, warning=FALSE, paged.print=FALSE}
 vocabulary$long %>% 
@@
 
-```{r raw, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
 responses %>%
@@
 
-```{r raw_age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r raw_age, message=FALSE, warning=FALSE, paged.print=FALSE}
 responses %>%
@@
 
-```{r ppc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}
+```{r ppc, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}
 
@@
 
-```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
 summary(fits$fit_2)$fixed %>% 
@@
 
-```{r fix_effets_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+```{r fix_effets_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
 The intercept has been transformed using the inverse logit to get the average probability of correct response.
@@
 
-```{r random_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r random_participant, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-```{r re_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r re_item, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-```{r emmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-```{r diagnostics_traceplots, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r diagnostics_traceplots, message=FALSE, warning=FALSE, paged.print=FALSE}
 

### 2021-09-15 Rmd/report.Rmd
Add GitHub dark theme

@@
 output:
+  html_document: default
   github_document:
@@
   message = TRUE,
-  echo = TRUE,
+  echo = FALSE,
   warning = TRUE,
@@
 tar_load_globals()
-theme_set(theme_custom())
+theme_set(theme_github())
 

### 2021-09-15 Rmd/report.Rmd
Add GitHub dark theme

@@
 output:
-  html_document: default
   github_document:
@@
   plot_layout(2) &
-  theme_custom() &
+  theme_github() &
   theme(
@@
   plot_layout() &
-  theme_custom() +
+  theme_github() +
   theme(
@@
   guides(colour = guide_legend(ncol = 2)) +
-  theme_custom() +
+  theme_github() +
   theme(
@@
   ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
-  geom_col(position = position_fill(), colour = "white") +
+  geom_col(position = position_fill(), colour = "#0D1117", size = 1) +
   guides(colour = guide_legend(ncol = 2)) +
@@
   scale_y_continuous(labels = scales::percent) +
-  theme_custom() +
+  theme_github() +
   theme(
@@
   ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
-  geom_col(colour = "white") +
+  geom_col(size = 1, colour = "#0D1117") +
   labs(y = "Number of participants", x = "Age (months)") +
   guides(colour = guide_legend(ncol = 2)) +
-  theme_custom() +
+  theme_github() +
   coord_flip() +
@@
   scale_x_continuous(labels = scales::percent) +
-  theme_bw() +
   theme(
@@
 vocabulary$long %>% 
-  filter(type=="produces") %>% 
+  filter(type=="understands") %>% 
   ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
@@
   scale_x_continuous(labels = scales::percent) +
-  theme_bw() +
+  theme_github() +
   theme(
+    panel.grid.major.y = element_blank(),
     legend.position = "top",
@@
   scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme_custom() +
+  theme_github() +
   theme(
@@
   scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme_custom() +
+  theme_github() +
   theme(
@@
   labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
-  theme_custom() +
+  theme_github() +
   theme(
@@
   scale_x_continuous(limits = c(0, 1)) +
-  theme_custom() +
+  theme_github() +
   theme(
@@
   labs(x = "Iteration", y = "Sample value", colour = "Chain") +
-  theme_custom() +
+  theme_github() +
   theme(

### 2021-09-15 Rmd/report.Rmd
Add GitHub dark theme

@@
 # Items {.tabset .tab-pills}
-
-```{r items_list, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=4, out.width="100%"}
-items %>% 
-  pivot_wider(
-    id_cols = c("te", "class", "cognate"),
-    names_from = "language",
-    values_from = c("label", "frequency", "ipa")
-  ) %>% 
-  clean_names() %>% 
-  arrange(te) %>% 
-  gt() %>% 
-  tab_spanner(label = md("**Catalan**"), columns = contains("_catalan")) %>% 
-  tab_spanner(label = md("**Spanish**"), columns = contains("_spanish")) %>% 
-  cols_label(
-    te = "TE ID",
-    class = "Class",
-    cognate = "Cognateness",
-    label_catalan = "Label",
-    label_spanish = "Label",
-    ipa_catalan = "Phonology",
-    ipa_spanish = "Phonology",
-    frequency_catalan = "Frequency",
-    frequency_spanish = "Frequency"
-  ) %>% 
-  fmt_number(columns = starts_with("frequency")) %>% 
-  as_raw_html() %>% 
-  div(style = "height:600px; overflow-x: scroll; overflow-y: scroll", .)
-```
-
-<br>
 

### 2021-09-15 Rmd/report.Rmd
Add GitHub dark theme

@@
   include = TRUE,
-  out.width = "100%",
+  out.width = "80%",
   results = "asis",
-  dpi = 300
+  dpi = 250
 )
@@
 \begin{aligned}
-\color{white}
-\log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
-& (\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
-& (\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
-& (\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
-& (\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
-& (\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
-& \varepsilon_{pi}
+\color{white}\log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
+& \color{white}(\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
+& \color{white}(\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
+& \color{white}(\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
+& \color{white}(\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
+& \color{white}(\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
+& \color{white}\varepsilon_{pi}
 \end{aligned}

### 2021-09-15 Rmd/report.Rmd
Add GitHub dark theme

@@
   warning = TRUE,
-  dev.args = list(png = list(type = "cairo")),
   include = TRUE,
   out.width = "80%",
-  results = "asis",
   dpi = 250
@@
 \begin{aligned}
-\color{white}\log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
+\color{white}\log(\frac{p}{n - q}) &= \color{white}(\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
 & \color{white}(\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\

### 2021-09-15 R/04_models.R
Add production PPCs

@@ -1,9 +1,9 @@
 # fit models
 
-fit_models <- function(
+fit_models_comp <- function(
     ... # arguments passed to brms::brm()
 ){
-
+    
     # formula
     formulas <- list(
         f_0 = bf(understands ~ 1 + age + frequency_center + (1 + age | te) + (1 + frequency_center | id), family = bernoulli("logit")),
@@ -32,17 +32,20 @@ fit_models <- function(
     # fit models
     fit_0 <- brm(
         formula = formulas$f_0, prior = priors[1:6,],
-        save_model = here("Stan", "irt_0.stan"), file = here("Results", "irt_0.rds"),
+        save_model = here("Stan", "irt_comp_0.stan"),
+        file = here("Results", "irt_comp_0.rds"),
         ...
     ) 
     fit_1 <- brm(
         formula = formulas$f_1, prior = priors[1:7,],
-        save_model = here("Stan", "irt_1.stan"), file = here("Results", "irt_1.rds"),
+        save_model = here("Stan", "irt_comp_1.stan"),
+        file = here("Results", "irt_comp_1.rds"),
         ...
     ) 
     fit_2 <- brm(
         formula = formulas$f_2, prior = priors,
-        save_model = here("Stan", "irt_2.stan"), file = here("Results", "irt_2.rds"),
+        save_model = here("Stan", "irt_comp_2.stan"),
+        file = here("Results", "irt_comp_2.rds"),
         ...
     ) 
     
@@ -59,3 +62,62 @@ compare_models <- function(
     loo <- loo_compare(map(fits, loo_subsample))
     saveRDS(loo, here("Results", "loo.rds"))
 }
+
+
+# fit models
+
+fit_models_prod <- function(
+    ... # arguments passed to brms::brm()
+){
+    
+    # formula
+    formulas <- list(
+        f_0 = bf(produces ~ 1 + age + frequency_center + (1 + age | te) + (1 + frequency_center | id), family = bernoulli("logit")),
+        f_1 = bf(produces ~ 1 + age + frequency_center + lp + (1 + age + lp | te) + (1 + frequency_center | id), family =  bernoulli("logit")),
+        f_2 = bf(produces ~ 1 + age + frequency_center + lp*cognate + (1 + age + lp | te) + (1 + frequency_center + cognate | id), family =  bernoulli("logit"))
+    )
+    
+    # priors (derived from prior samples)
[…]

### 2021-09-15 Rmd/report.Rmd
Add production PPCs

@@
 tar_load(responses)
-tar_load(fits)
+tar_load(fits_comp)
+tar_load(fits_prod)
 
@@
   plot_layout(guides = "collect") &
+  theme_github() +
   theme(
@@
   facet_grid(~scale) +
-  geom_point() +
+  geom_point(size = 3) +
   labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
@@
 
-## Prior-predictive checks
+## Prior-predictive checks (PPC)
 
-```{r ppc, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}
+### PPC: Comprehension
+
+```{r ppc_comp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}
 
@@
 
-m <- add_fitted_draws(nd, fits$fit_2, n = 20, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, fits_comp$fit_2, n = 20, re_formula = NA) %>% 
   mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))
@@
   scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
+  theme_github() +
+  theme(
+    legend.position = "top",
+    legend.title = element_blank()
+  )
+
+```
+
+### PPC: Production
+
+```{r ppc_prod, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}
+
+# emmeans
+nd <- expand.grid(
+  age = seq(min(responses$age), max(responses$age), length.out = 50),
+  lp = unique(participants$lp),
+  cognate = unique(items$cognate),
+  frequency_center = c(-1, 0, 1)
+)
+
+m <- add_fitted_draws(nd, fits_prod$fit_2, n = 20, re_formula = NA) %>% 
+  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))
+
+ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
+  facet_grid(lp~frequency_center) +
+  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
+  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
+  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
+  # stat_summary(fun = "mean", geom = "line", size = 1) +
+  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
+  labs(x = "Age (months)", y = "P(Production)", colour = "Cognateness", fill = "Cognateness") +
+  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
+  theme_github() +
   theme(
@@
 
+### Fixed effects: Compehension
+
 ```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
-summary(fits$fit_2)$fixed %>% 
+summary(fits_comp$fit_2)$fixed %>% 
   as.data.frame() %>%
@@
 
-```{r fix_effets_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+```{r fix_effets_comp_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
 The intercept has been transformed using the inverse logit to get the average probability of correct response.
@@
 
-post <- gather_draws(fits$fit_2, `b_.*`, regex = TRUE) %>% 
+post <- gather_draws(fits_comp$fit_2, `b_.*`, regex = TRUE) %>% 
   mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))
@@
   geom_vline(xintercept = 0, size = 1, colour = "grey") +
-  stat_histinterval(fill = "lightblue") +
+  stat_histinterval(fill = "white") +
+  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
+  theme_github() +
+  theme(
+    legend.position = "none",
+    plot.caption.position = "plot",
+    plot.caption = element_text(hjust = 0),
+    # axis.text = element_text(size = 7),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+  )
+```
+
+### Fixed effects: Production
+
+```{r fix_effets_prod_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+The intercept has been transformed using the inverse logit to get the average probability of correct response.
+The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
+
+post <- gather_draws(fits_prod$fit_2, `b_.*`, regex = TRUE) %>% 
+  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))
+
+ggplot(post, aes(.value)) +
+  facet_wrap(~.variable, scales = "free_x") +
+  geom_vline(xintercept = 0, size = 1, colour = "grey") +
+  stat_histinterval(fill = "white") +
   labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
@@
 
-### Participant
+### Random effects: Comprehension
 
-```{r random_participant, message=FALSE, warning=FALSE, paged.print=FALSE}
+#### Participant
 
-re_id <- ranef(fits$fit_2)$id[,,1] %>% 
+```{r random_participant_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+re_id <- ranef(fits_comp$fit_2)$id[,,1] %>% 
   as.data.frame() %>% 
@@
   as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
-  left_join(distinct(fits$fit_2$data, id, lp))
+  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
+  left_join(distinct(fits_comp$fit_2$data, id, lp))
 
@@
 
-### Item
+#### Item
 
-```{r re_item, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r re_item_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-re_te <- ranef(fits$fit_2)$te[,,1] %>% 
+re_te <- ranef(fits_comp$fit_2)$te[,,1] %>% 
   as.data.frame() %>% 
@@
   as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
+  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
   mutate(te = as.numeric(te)) %>% 
-  left_join(distinct(fits$fit_2$data, te, lp))
+  left_join(distinct(fits_comp$fit_2$data, te, lp))
+
+ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
+                 colour = wes_palettes$Darjeeling1[2]) +
+  geom_point(size = 1, shape = "|") +
+  geom_vline(xintercept = 0.5, linetype = "dashed") +
+  labs(x = "Item average probability of comprehension",
+       y = "Item") +
+  scale_color_brewer(palette = "Set1") +
+  scale_x_continuous(limits = c(0, 1)) +
+  theme_github() +
+  theme(
+    axis.text.y = element_blank(),
+    panel.grid = element_blank(),
+    axis.ticks = element_blank()
+  )
+```
+
+### Random effects: Production
+
+#### Participant
+
+```{r random_participant_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+re_id <- ranef(fits_prod$fit_2)$id[,,1] %>% 
+  as.data.frame() %>% 
+  rownames_to_column("id") %>% 
+  as_tibble() %>% 
+  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
+  left_join(distinct(fits_prod$fit_2$data, id, lp))
+
+ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
+                 colour = wes_palettes$Darjeeling1[1]) +
+  geom_point(size = 1, shape = "|") +
+  geom_vline(xintercept = 0.5, linetype = "dashed") +
+  labs(x = "Participant average probability of comprehension",
+       y = "Participant") +
+  scale_x_continuous(limits = c(0, 1)) +
+  theme_github() +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    panel.grid = element_blank()
+  )
+
+```
+
+#### Item
+
+```{r re_item_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+re_te <- ranef(fits_prod$fit_2)$te[,,1] %>% 
+  as.data.frame() %>% 
+  rownames_to_column("te") %>% 
+  as_tibble() %>% 
+  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
+  mutate(te = as.numeric(te)) %>% 
+  left_join(distinct(fits_prod$fit_2$data, te, lp))
 
@@
 
-```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE}
+### Marginal means: Comprehension
+
+```{r emmeans_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-m <- add_fitted_draws(nd, fits$fit_2, re_formula = NA, n = 25) %>% 
+m <- add_fitted_draws(nd, fits_comp$fit_2, re_formula = NA, n = 25) %>% 
   mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))
@@
   theme(
-    legend.position = "right",
+    legend.position = "top",
+    legend.title = element_blank()
+  )
+
+```
+
+## Marginal means: Production
+
+```{r emmeans_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+nd <- expand.grid(
+  age = seq(min(responses$age), max(responses$age), length.out = 50),
+  lp = unique(participants$lp),
+  cognate =unique(items$cognate),
+  frequency_center = c(-1,  0, 1)
+)
+
+m <- add_fitted_draws(nd, fits_prod$fit_2, re_formula = NA, n = 25) %>% 
+  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))
+
+ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
+  facet_wrap(lp~frequency_center) +
+  geom_line(aes(group = interaction(.draw, frequency_center, cognate)), alpha = 0.5) +
+
+  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
+  # stat_summary(fun = "mean", geom = "line", size = 1) +
+  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
+  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
+  scale_y_continuous(labels = scales::percent) +
+  theme_github() +
+  theme(
+    legend.position = "top",
     legend.title = element_blank()
@@
 
-```{r diagnostics_traceplots, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-gather_draws(fits$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
+```{r diagnostics_traceplots_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
+gather_draws(fits_comp$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
   mutate(.chain = paste("Chain ", .chain)) %>% 
@@
   geom_line(alpha = 0.5) +
-  labs(x = "Iteration", y = "Sample value", colour = "Chain") +
+  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Comprehension") +
+  theme_github() +
+  theme(
+    legend.position = "top",
+    legend.title = element_blank(),
+    panel.grid = element_blank()
+  )
+```
+
+```{r diagnostics_traceplots_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+gather_draws(fits_prod$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
+  mutate(.chain = paste("Chain ", .chain)) %>% 
+  ggplot(aes(.iteration, .value, colour = .chain)) +
+  facet_wrap(~.variable, scales = "free_y") +
+  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
+  geom_line(alpha = 0.5) +
+  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Production") +
   theme_github() +

### 2021-09-15 R/utils.R

@@ -20,9 +20,9 @@ theme_github <- function(){
             legend.title = element_text(colour = "white", size = 10),
             legend.text = element_text(colour = "white", size = 10),
             legend.background = element_rect(fill = "#0D1117"),
+            strip.text = element_text(size = 10),
             legend.key = element_rect(fill = "#0D1117", colour = "#0D1117"),
             strip.background = element_rect(fill = "#161B22"),
-            strip.text = element_text(colour = "white", size = 12),
             plot.background = element_rect(fill = "#0D1117"),
             panel.background = element_rect(fill = "#0D1117"),
             panel.border = element_rect(fill = "transparent", colour = "#0D1117")

### 2021-09-20 Rmd/report.Rmd
Add LOO

@@
 tar_load(fits_prod)
+tar_load(fits_comp_prior)
+tar_load(fits_prod_prior)
+tar_load(loo_comp)
+tar_load(loo_prod)
 
@@
 labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
-a <- participants %>% 
+participants %>% 
   mutate(
@@
       palette = c("white", "orange"),
-      domain = c(0, 11)
+      domain = c(0, 30)
     )
@@
 
+```
+
+## SES/parental education
+
+```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+participants %>%
+  mutate(edu_parent = fct_explicit_na(as.factor(edu_parent))) %>% 
+  count(lp, dominant_language, edu_parent) %>% 
+  right_join(
+    expand_grid(
+      lp = c("Monolingual", "Bilingual"),
+      dominant_language = c("Catalan", "Spanish"),
+      edu_parent = unique(participants$edu_parent)
+    )
+  ) %>% 
+  replace_na(list(n = 0)) %>% 
+  pivot_wider(names_from = edu_parent, values_from = n, values_fill = 0) %>%
+  mutate(n = rowSums(cbind(.[,3:8]))) %>% 
+  select(-"NA") %>% 
+  gt() %>% 
+  cols_label(
+    lp = "Lang. Profile",
+    dominant_language = "L1"
+  ) %>% 
+  summary_rows(
+    columns = 3:8,
+    fns = list(n = ~sum(.)),
+    formatter = fmt_number,
+    decimals = 0
+  ) %>% 
+  tab_style(
+    style = list(cell_borders(weight = px(2), sides = "left")),
+    locations = cells_body(columns = 10, rows = 1:4)
+  )
 ```
@@
 
-### Short vs. Long {.tabset .tab-pills}
-
-#### Comprehension
-
-```{r vocabulary_long_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}
-vocabulary$long %>% 
-  filter(type=="understands") %>% 
-  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
-  facet_grid(~scale) +
-  geom_point() +
-  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
-       colour = "Version", fill = "Version", shape = "Version") +
-  scale_x_continuous(labels = scales::percent) +
-  theme(
-    legend.position = "top",
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-```
-
-#### Production
-
-```{r vocabulary_long_production, message=FALSE, warning=FALSE, paged.print=FALSE}
-vocabulary$long %>% 
-  filter(type=="understands") %>% 
-  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
-  facet_grid(~scale) +
-  geom_point(size = 3) +
-  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
-       colour = "Version", fill = "Version", shape = "Version") +
-  scale_x_continuous(labels = scales::percent) +
-  theme_github() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    legend.position = "top",
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-```
-
 #### {-}
@@
     understands = mean(understands, na.rm = TRUE),
-    produces = mean(produces, na.rm = TRUE)
+    produces = mean(produces, na.rm = TRUE),
+    .groups = "drop"
   ) %>%
@@
   age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = unique(participants$lp),
+  lp = c("Monolingual", "Bilingual"),
   cognate = unique(items$cognate),
@@
 
-m <- add_fitted_draws(nd, fits_comp$fit_2, n = 20, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, fits_comp_prior$fit_2, n = 20, re_formula = NA) %>% 
   mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))
@@
 Although the `loo` function uses a particular algorithm that speeds up the computation of this criterion (pareto-smooth importance sampling, PSIS), the size of our data set lead us to rely on the computation of the same criterion using a sampling approach via de `loo_subsample` function.
+
+### Comprehension
+
+```{r loo_comp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+loo_comp %>% 
+  as_tibble() %>% 
+  rownames_to_column("model") %>% 
+  relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
+  mutate_all(as.numeric) %>% 
+  gt() %>% 
+  tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
+  tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
+  tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
+  tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
+  fmt_number(3:13) %>% 
+  cols_label(
+    model = "Model",
+    # elpd
+    elpd_loo = md("*ELPD*"),
+    se_elpd_loo = md("*SE*"),
+    subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
+    # p
+    p_loo = md("*p*"),
+    se_p_loo = md("SE"),
+    subsampling_se_p_loo = md("SE<sub>sub</sub>"),
+    # looic
+    looic = md("*LOO-IC*"),
+    se_looic = md("*SE*"),
+    subsampling_se_looic = md("SE<sub>sub</sub>"),
+    # diff
+    elpd_diff = md("*diff*"),
+    se_diff = md("*SE*"),
+    subsampling_se_diff = md("*SE<sub>sub</sub>*"),
+  ) %>% 
+  as_raw_html()
+```
+
+### Production
+
+```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+loo_comp %>% 
+  as_tibble() %>% 
+  rownames_to_column("model") %>% 
+  relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
+  mutate_all(as.numeric) %>% 
+  gt() %>% 
+  tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
+  tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
+  tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
+  tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
+  fmt_number(3:13) %>% 
+  cols_label(
+    model = "Model",
+    # elpd
+    elpd_loo = md("*ELPD*"),
+    se_elpd_loo = md("*SE*"),
+    subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
+    # p
+    p_loo = md("*p*"),
+    se_p_loo = md("SE"),
+    subsampling_se_p_loo = md("SE<sub>sub</sub>"),
+    # looic
+    looic = md("*LOO-IC*"),
+    se_looic = md("*SE*"),
+    subsampling_se_looic = md("SE<sub>sub</sub>"),
+    # diff
+    elpd_diff = md("*diff*"),
+    se_diff = md("*SE*"),
+    subsampling_se_diff = md("*SE<sub>sub</sub>*"),
+  ) %>% 
+  as_raw_html()
+```
 
@@
 ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
-                 colour = wes_palettes$Darjeeling1[1]) +
-  geom_point(size = 1, shape = "|") +
-  geom_vline(xintercept = 0.5, linetype = "dashed") +
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", height = 0) +
+  geom_point(size = 0.5, colour = "white") +
+  geom_vline(xintercept = 0.5, colour = "white") +
   labs(x = "Participant average probability of comprehension",
@@
     axis.ticks = element_blank(),
-    panel.grid = element_blank()
-  )
-
+    panel.grid = element_line(colour = "white", linetype = "dotted"),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    axis.line = element_line(colour = "white")
+  )
 ```
@@
 ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
-                 colour = wes_palettes$Darjeeling1[2]) +
-  geom_point(size = 1, shape = "|") +
-  geom_vline(xintercept = 0.5, linetype = "dashed") +
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
+  geom_point(size = 0.5, colour = "white") +
+  geom_vline(xintercept = 0.5, colour = "white") +
   labs(x = "Item average probability of comprehension",
@@
     axis.text.y = element_blank(),
-    panel.grid = element_blank(),
-    axis.ticks = element_blank()
+    axis.ticks = element_blank(),
+    panel.grid = element_line(colour = "white", linetype = "dotted"),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    axis.line = element_line(colour = "white")
   )
@@
 ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
-                 colour = wes_palettes$Darjeeling1[1]) +
-  geom_point(size = 1, shape = "|") +
-  geom_vline(xintercept = 0.5, linetype = "dashed") +
-  labs(x = "Participant average probability of comprehension",
-       y = "Participant") +
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
+  geom_point(size = 0.5, colour = "white") +
+  geom_vline(xintercept = 0.5, colour = "white") +
+  labs(x = "Item average probability of production",
+       y = "Item") +
+  scale_color_brewer(palette = "Set1") +
   scale_x_continuous(limits = c(0, 1)) +
@@
     axis.ticks = element_blank(),
-    panel.grid = element_blank()
+    panel.grid = element_line(colour = "white", linetype = "dotted"),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    axis.line = element_line(colour = "white")
   )
@@
 ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
-                 colour = wes_palettes$Darjeeling1[2]) +
-  geom_point(size = 1, shape = "|") +
-  geom_vline(xintercept = 0.5, linetype = "dashed") +
-  labs(x = "Item average probability of comprehension",
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
+  geom_point(size = 0.5, colour = "white") +
+  geom_vline(xintercept = 0.5, colour = "white") +
+  labs(x = "Item average probability of production",
        y = "Item") +
@@
   age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = unique(participants$lp),
-  cognate =unique(items$cognate),
+  lp = c("Monolingual", "Bilingual"),
+  cognate = unique(items$cognate),
   frequency_center = c(-1,  0, 1)

### 2021-09-20 Rmd/report.Rmd
Fix report GT tables

@@
     locations = cells_body(columns = 10, rows = 1:4)
-  )
+  ) %>%
+  as_raw_html()
 ```

### 2021-09-20 Rmd/report.Rmd
Change col name in SES table

--- 2021-09-20 Rmd/report.Rmd
@@
   replace_na(list(n = 0)) %>% 
-  pivot_wider(names_from = edu_parent, values_from = n, values_fill = 0) %>%
-  mutate(n = rowSums(cbind(.[,3:8]))) %>% 
+  rename(total = n) %>% 
+  pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
+  mutate(total = rowSums(cbind(.[,3:8]))) %>% 
+  rename(Total = total) %>% 
   select(-"NA") %>% 
@@
     columns = 3:8, 
-    fns = list(n = ~sum(.)),
+    fns = list(Total = ~sum(.)),
     formatter = fmt_number,
@@
 
-```{stan stan_model, output.var="model"}
+```{stan stan_model, echo=FALSE, output.var="model"}
 // generated with brms 2.15.0

### 2021-09-20 Rmd/report.Rmd
fix fixed effets plots

@@
 # Results {.tabset .tab-pills}
-
-## Vocabulary {.tabset .tab-pills}
-
-### Vocabulary size {.tabset .tab-pills}
-
-#### Comprehension
-
-```{r vocabulary_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-d <- vocabulary$wide %>% 
-  filter(type=="Understands", modality != "Te")
-
-ggplot(d, aes(age, prop, colour = lp, fill = lp)) +
-  facet_wrap(~modality) +
-  geom_point(size = 0.5, alpha = 0.21) +
-  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
-  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-
-  ggplot(d, aes(age, count, colour = lp, fill = lp)) +
-  facet_wrap(~modality) +
-  geom_point(size = 0.5, alpha = 0.21) +
-  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
-  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +
-
-  plot_layout(guides = "collect") &
-  theme_github() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-
-```
-
-#### Production
-
-```{r vocabulary_production, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-vocabulary$wide %>% 
-  filter(type=="Produces", modality != "Te") %>% 
-  ggplot(aes(age, prop, colour = lp, fill = lp)) +
-  facet_wrap(~modality) +
-  geom_point(size = 0.5, alpha = 0.21) +
-  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
-  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-
-  vocabulary$wide %>% 
-  filter(type=="Produces", modality != "Te") %>% 
-  ggplot(aes(age, count, colour = lp, fill = lp)) +
-  facet_wrap(~modality) +
-  geom_point(size = 0.5, alpha = 0.21) +
-  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
-  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +
-
-  plot_layout(guides = "collect") &
-  theme_github() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-
-```
-
-#### {-}
-
-#### {-}
-
-### {-}
 
@@
 ```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-loo_comp %>% 
+loo_prod %>% 
   as_tibble() %>% 
@@
 
+var_levels <- c("b_Intercept", "b_age", "b_frequency_center", "b_lp1", "b_cognate1", "b_cognate2", "b_lp1:cognate1", "b_lp1:cognate2")
+var_labels <- c("Intercept", "Age (+1 month)", "Frequency (+1 SD)", "Lang.
+Profile (Bil.
+- Mon.)", "Cognateness 1 (IDC+NIC - NC)", "Cognateness 2 (IDC -NIC)", "Lang.
+Profile \u00d7 Cognateness 1", "Lang.
+Profile \u00d7 Cognateness 2")
 post <- gather_draws(fits_comp$fit_2, `b_.*`, regex = TRUE) %>% 
-  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))
-
-ggplot(post, aes(.value)) +
-  facet_wrap(~.variable, scales = "free_x") +
-  geom_vline(xintercept = 0, size = 1, colour = "grey") +
-  stat_histinterval(fill = "white") +
-  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
+  mutate(
+    .value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4),
+    .variable = factor(.variable, levels = var_levels, labels = var_labels, ordered = TRUE)
+  )
+
+ggplot(post, aes(.value, fct_rev(.variable))) +
+  geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
+  stat_pointinterval(colour = "white") +
+  labs(x = "Comprehension posterior probability", y = "Variable", fill = "Variable") +
+  scale_x_continuous(labels = percent) +
   theme_github() +
@@
     plot.caption = element_text(hjust = 0),
+    axis.title.y = element_blank(),
     # axis.text = element_text(size = 7),
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+    panel.grid.major.y = element_blank()
   )
@@
 
+var_levels <- c("b_Intercept", "b_age", "b_frequency_center", "b_lp1", "b_cognate1", "b_cognate2", "b_lp1:cognate1", "b_lp1:cognate2")
+var_labels <- c("Intercept", "Age (+1 month)", "Frequency (+1 SD)", "Lang.
+Profile (Bil.
+- Mon.)", "Cognateness 1 (IDC+NIC - NC)", "Cognateness 2 (IDC -NIC)", "Lang.
+Profile \u00d7 Cognateness 1", "Lang.
+Profile \u00d7 Cognateness 2")
 post <- gather_draws(fits_prod$fit_2, `b_.*`, regex = TRUE) %>% 
-  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))
-
-ggplot(post, aes(.value)) +
-  facet_wrap(~.variable, scales = "free_x") +
-  geom_vline(xintercept = 0, size = 1, colour = "grey") +
-  stat_histinterval(fill = "white") +
-  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
+  mutate(
+    .value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4),
+    .variable = factor(.variable, levels = var_levels, labels = var_labels, ordered = TRUE)
+  )
+
+ggplot(post, aes(.value, fct_rev(.variable))) +
+  geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
+  stat_pointinterval(colour = "white") +
+  labs(x = "Production posterior probability", y = "Variable", fill = "Variable") +
+  scale_x_continuous(labels = percent) +
   theme_github() +

### 2021-09-24 Rmd/02_methods.Rmd
First corrections to Núria's comments
Add methods section and rename manuscript files

--- 2021-07-12 Rmd/02_methods.Rmd
@@
 ---
-title: "Methods"
-author: "Gonzalo García-Castro"
-date: "6/22/2021"
-output:
-  word_document:
-    reference_docx: trajectories_template.docx
-  pdf_document:
-      latex_engine: xelatex
-bibliography: trajectories.bib
-csl: apa6.csl
-self_contained: true
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 
+Data and materials are available at OSF (https://osf.io/hy984/), and code is available at GitHub (https://github.com/gongcastro/trajectories).
+
 ## Participants
 
-## Vocabulary checklist
+```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+tar_load_globals()
+tar_load(participants)
+tar_load(items)
+tar_load(responses)
+
+n_participants <- length(unique(participants$id))
+dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
+sex <- participants %>% drop_na(sex) %>% count(sex) %>% pull(n) %>% set_names(c("female", "male"))
+
+```
+
+We collected data from `r printnum(n_participants)` children (`r sex[1]` female) from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
+All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
+Families were contacted by e-mail or phone if their child were aged between 10 and 34 months, and had not previously reported to be exposed more than 10% of the time to a language other than Spanish or Catalan.
+Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate.
+Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to L2.
+
+```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Participant sample size by age and degree of exposure to a second language.
+Information about DoE was provided by families when filling the questionnaire.
+For illustration purposes this table, DoEs were binned into 10\\% wide bins, and ages (in months) were round to the nearest integer.
+Hyphens indicate that no participants from that specific combination of age filled the questionnaire."}
+labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
+participants %>% 
+  mutate(
+    age = round(age),
+    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
+    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
+  ) %>% 
+  count(age, doe_2) %>% 
+  arrange(doe_2) %>% 
+  mutate(doe_2 = fct_inorder(doe_2)) %>% 
+  pivot_wider(names_from = age, values_from = n, values_fill = NA) %>% 
+  gt() %>% 
+  tab_spanner(label = "Number of participants by age (months)", columns = 2:17) %>% 
+  fmt_missing(everything(), everything(), missing_text = "-") %>% 
+  cols_label(
+    doe_2 = md("**L2 DoE**")
+  ) %>% 
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels(everything())
+  ) %>% 
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners(everything())
+  )
+
+```
+
+We use the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain.
+Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r `percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
+Highest educational attainment was similar across participants with different DoE to a second language, and therefore, differences in comprehension or production between participants with different language profiles are unlikely to stem from a difference in SES [@fernald2013ses; see Table 2].
+
+```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Educational attainment.
+Proportion of highest parental educational attainment by DoE to the second language.
+For illustration purposes, DoEs were binned into 10\\% wide bins.
+The bottom row indicates the mean proportion of participant families for each educational attainment."}
+labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
+
+participants %>%
+  mutate(
+    age = round(age),
+    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
+    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
+  ) %>% 
+  group_by(doe_2) %>%
+  summarise(
+    n_no_education = length(edu_parent[edu_parent=="No education"])/n(),
+    n_primary = length(edu_parent[edu_parent=="Primary"])/n(),
+    n_secondary = length(edu_parent[edu_parent=="Secondary"])/n(),
+    n_complementary = length(edu_parent[edu_parent=="Complementary"])/n(),
+    n_vocational = length(edu_parent[edu_parent=="Vocational"])/n(),
+    n_university = length(edu_parent[edu_parent=="University"])/n(),
+    .groups = "drop"
+  ) %>% 
+  gt() %>% 
+  tab_spanner("Educational attainment", columns = 2:7) %>% 
+  cols_label(
+    doe_2 = md("**L2 DoE**"),
+    n_no_education = "No education",
+    n_primary = "Primary",
+    n_secondary = "Secondary",
+    n_complementary = "Complementary",
+    n_vocational = "Vocational",
+    n_university = "University"
+  ) %>% 
+  fmt_percent(2:7) %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  summary_rows(
+    columns = 2:7, 
+    fns = list(Mean = ~mean(.)),
+    formatter = fmt_percent,
+    decimals = 2
+  ) %>%  
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels(everything())
+  ) %>% 
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners(everything())
+  )
+```
+
+## Questionnaire
+
+```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+pool <- multilex::pool %>% 
+  drop_na(cognate)
+n_categories <- distinct(pool, category) %>% nrow()
+n_items <- length(unique(pool$item))
+n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
+n_item_cognate <- count(pool, language, cognate) %>% 
+  pull(n) %>% 
+  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
+```
+
+The questionnaire was implemented on-line using `formr` [@arslan2020formr].
+This questionnaire was divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*.
+We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
+We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines.
+The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
+
+For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default).
+Parents filled a long or a short version of the questionnaire.
+Those presented with the long version filled a list of 800 translation equivalents (800 items in Catalan, 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items.
+Each list contained a different set of ~400 translation equivalents (~400 items in Catalan, ~400 items in Spanish).
+Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
+Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+Table \@ref(tab:itemsinfo) shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
+
+```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Lexical frequencies.
+Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates.
+The bottom row indicates the grand mean, mean standard error and mean 95\\% confidence interval for lexical frequencies of identical cognates, non-identical cognates, and non-cognates."}
+items %>%
+  select(te, language, cognate, frequency) %>% 
+  left_join(select(multilex_data$pool, te, language, version)) %>% 
+  unnest(version) %>% 
+  group_by(language, version, cognate) %>% 
+  summarise(
+    frequency_mean = mean(frequency, na.rm = TRUE),
+    frequency_sd = sd(frequency, na.rm = TRUE),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    frequency_se = frequency_sd/sqrt(n),
+    frequency_ci_lower = frequency_mean + qnorm(0.025)*frequency_se,
+    frequency_ci_upper = frequency_mean + qnorm(0.975)*frequency_se
+  ) %>% 
+  pivot_wider(names_from = c("cognate"), values_from = c("n", matches("frequency"))) %>% 
+  clean_names() %>% 
+  relocate(
+    language, version,
+    n_identical_cognate,
+    frequency_mean_identical_cognate,
+    frequency_se_identical_cognate,
+    frequency_ci_lower_identical_cognate,
+    frequency_ci_upper_identical_cognate,
+    matches("_non_identical"),
+    matches("_non_cognate")
+  ) %>% 
+  select(-matches("sd")) %>% 
+  gt(groupname_col = "language") %>% 
+  fmt_number(matches("frequency")) %>% 
+  summary_rows(
+    columns = matches("frequency"), 
+    fns = list(Mean = ~mean(.)),
+    formatter = fmt_number,
+    decimals = 2
+  ) %>%
+  cols_merge(
+    c("frequency_ci_lower_identical_cognate", "frequency_ci_upper_identical_cognate"), 
+    pattern = "[{1}, {2}]") %>% 
+  cols_merge(
+    c("frequency_ci_lower_non_identical_cognate", "frequency_ci_upper_non_identical_cognate"),
+    pattern = "[{1}, {2}]"
+  ) %>% 
+  cols_merge(
+    c("frequency_ci_lower_non_cognate", "frequency_ci_upper_non_cognate"),
+    pattern = "[{1}, {2}]"
+  ) %>% 
+  tab_spanner("Identical cognates", columns = ends_with("_identical_cognate")) %>% 
+  tab_spanner("Non-identical cognates", columns = ends_with("_non_identical_cognate")) %>% 
+  tab_spanner("Non-cognates", columns = ends_with("_non_cognate")) %>% 
+  cols_width(
+    starts_with("frequency_ci_lower_") ~ px(100),
+    TRUE ~ px(50)
+  ) %>% 
+  cols_label(
+    version = "List",
+    n_identical_cognate = "N",
+    n_non_identical_cognate = "N",
+    n_non_cognate = "N",
+    frequency_mean_identical_cognate = md("Mean<sub>freq</sub>"),
+    frequency_mean_non_identical_cognate = md("Mean<sub>freq</sub>"),
+    frequency_mean_non_cognate = md("Mean<sub>freq</sub>"),
+    frequency_se_identical_cognate = md("SD<sub>freq</sub>"),
+    frequency_se_non_identical_cognate = md("SD<sub>freq</sub>"),
+    frequency_se_non_cognate = md("SD<sub>freq</sub>"),
+    frequency_ci_lower_identical_cognate = md("95% CI<sub>freq</sub>"),
+    frequency_ci_lower_non_identical_cognate = md("95% CI<sub>freq</sub>"),
+    frequency_ci_lower_non_cognate = md("95% CI<sub>freq</sub>")
+  ) %>%  
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels(everything())
+  ) %>% 
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners(everything())
+  )
+
+```
 
@@
 
-$$
+```{r n_responses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+n_obs <- nrow(responses)
+sum_obs <- responses %>%
+  group_by(te) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  summarise(
+    mean = mean(n),
+    sd = sd(n),
+    min = min(n),
+    max = max(n)
+  )
 
-$$
+```
+
+We gathered `r printnum(n_obs, big.mark = ",")` responses.
+Translation equivalents (TEs) received an average of `r printnum(sum_obs$mean, digits = 2)` responses (*SD* = `r printnum(sum_obs$sd, digits = 2)`, *Min* = `r printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
+To investigate our hypotheses, we modelled the probability of children understanding or producing TEs using a logistic regression model with a Bernoulli distribution and a logit link function [@agresti2003categorical].
+We included several predictors of interest as fixed effects to adjust this probability to participants' and items' characteristics, namely:
+
+* *Age* of participant in months (`age`), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
+* *Lexical frequency* of item (`frequency_center`), Relative lexical frequencies (frequency per million words, $fpmw$) were extracted from the CHILDES data base, using the `childesr` R package [@braginsky2018childesr], and were transformed to Zipf scores [@van2014subtlex; @zipf1949human] ($log10(fpmw)+3$).
+Missing frequencies were imputed by multiple imputation using the `mice` R package [van2011mice].
+The resulting scores were then centred around the mean.
+* *Degree of exposure to L2* of participant (`doe_center`), computed as the percentage of exposure to a second language (Spanish for participants exposed to >50% to Catalan an *vice versa*), chunked into six groups of participants (a proxy of degree of bilingualism), ranging from 0% to 50%, and centred around the mean.
+* *Cognateness* of the item, indicating whether the two forms of each TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist.
+We sum-coded two contrasts [@schad2020capitalize], one comparing IC (coded as +0.5) and NIC with NC (coded as -0.5) (`cognate1`), and the other comparing IC (coded as +0.5) with NIC (coded as -0.5) (`cognate2`).
+
+We first fit a base model that only included the main effects of `age` and `frequency_center`.
+[see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches] (Model 1).
+We then introduced our predictors of interest by adding `doe_center` (Model 2) and then `cognate` and its interaction with `doe` (Model 3).
+To account for the hierarchical structure of the data--some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE--every model included *Participants* (`id`) and *TEs* (`te`) as grouping variables (i.e.
+random effects), with all possible random intercepts and slopes where appropriate [@barr2013random; see Appendix for more details of the model].
+
+We adopted a Bayesian approach toward statistical inference.
+This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model [@mcelreath2018statistical].
+We fit the models using the R package `brms` [@burkner2017brms], an R interface to the probabilistic language Stan [@carpenter2017stan].
+We run four sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
+All R-hat values were smaller than 1.1, indicating that the chains converged successfully in all cases [@gelman1992inference; see Appendix for an extended description of model diagnostics].
+
+We then compared comprehension and production models separately using leave-one-out cross-validation (LOO-CV) [@vehtari2017practical]^[Due to the high computational cost associated with a large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson2019bayesian]].
+This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
+The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
+We compared the expected log posterior density (ELPD) of each each model to test which one fitted the data the best.
+
+We then explored the posterior distribution of each parameter in the comprehension and production models that fitted the data the best computing the 95% credible intervals and testing whether this interval excluded 0.
+Credible intervals (CrI) indicate the range of values we are 95% certain contain the true value of the parameter, given our prior and observed data.
+We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% credible interval of their estimated marginal means using the `emmeans` package [@R-emmeans].
+Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], and posterior samples and marginal means were extracted using the `tidybayes` R package [@R-tidybayes].

### 2021-09-24 Rmd/03_results.Rmd
Add methods section and rename manuscript files

@@
 ```{r results, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-load(here("Results", "comp_fits.RData"))
-load(here("Results", "prod_fits.RData"))
-load(here("Results", "loos.RData"))
-
-coef_comp <- as.data.frame(fixef(comp_7))
+tar_load(fits_comp)
+tar_load(fits_prod)
+tar_load(loo_comp)
+tar_load(loo_prod)
 ```
@@
 
-```{r loos, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-loos_diff <- bind_rows(
-  "Comprehension" = select(
-    as.data.frame(loos_comp$diffs),
-    one_of("elpd_loo", "se_elpd_loo", "p_loo", "elpd_diff", "se_diff")
-  ),
-  "Production" = select(
-    as.data.frame(loos_prod$diffs),
-    one_of("elpd_loo", "se_elpd_loo", "p_loo", "elpd_diff", "se_diff")
-  ),
-  .id = "type"
-) %>% 
-  mutate(
-    model = rep(rev(c(
-      "~ 1 + age + frequency",
-      "... + dominance",
-      "... + bilingualism",
-      "... + dominance:bilingualism",
-      "... + cognate",
-      "... + dominance:cognate",
-      "... + bilingualism:cognate",
-      "... + dominance:bilingualism:cognate"
-    )), 2)) %>% 
-  relocate(model) %>% 
-  group_by(type) %>% 
-  arrange(elpd_loo)
-
-gt(loos_diff) %>% 
-  tab_spanner(
-    label = md("**Comparison**"),
-    columns = c("elpd_diff", "se_diff")
-  ) %>% 
-  cols_label(
-    model = md("**Model**"),
-    elpd_loo = md("**ELPD**"),
-    se_elpd_loo = md("**SE**"),
-    p_loo = md("**p**"),
-    elpd_diff = md("*LOO<sub>diff</sup>*"),
-    se_diff = md("*SE<sub>diff</sup>*")
-  ) %>% 
-  fmt_number(2:5) %>% 
-  tab_footnote(
-    footnote = md("*ELPD*: theoretical expected log pointwise predictive density for a new dataset, as estimated using Bayesian LOO-CV"),
-    locations = cells_column_labels(columns = vars("elpd_loo"))
-  ) %>% 
-  tab_footnote(
-    footnote = md("*SE*: standard error of the ELPD"),
-    locations = cells_column_labels(columns = vars("se_elpd_loo"))
-  ) %>% 
-  tab_footnote(
-    footnote = md("*p*: effective number of parameters"),
-    locations = cells_column_labels(columns = vars("p_loo"))
-  ) %>% 
-  tab_options(footnotes.marks = "letters")
-
-```
-
 We report the mean of each coefficient's posterior distribution along with standard errors and 95% CrI in Table 3.
@@
 This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.] for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable [@gelman_regression_2020].
-
-```{r coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-estimates <- bind_cols(
-  "Comprehension" =  rename_all(as.data.frame(inv_logit_scaled(fixef(comp_7))), ~paste0(., "_comp")),
-  "Production" =  rename_all(as.data.frame(inv_logit_scaled(fixef(prod_7))), ~paste0(., "_prod"))
-) %>% 
-  mutate(
-    coefficient = c(
-      "Intercept",
-      "Age",
-      "Frequency",
-      "Dominance",
-      "Bilingualism",
-      "Cognate",
-      "Dominance * Bilingualism",
-      "Dominance * Cognate",
-      "Bilingualism * Cognate",
-      "Dominance * Bilingualism * Cognate"
-    )
-  ) %>% 
-  clean_names() %>% 
-  relocate(coefficient)
-
-rownames(estimates) <- NULL
-
-gt(estimates) %>% 
-  tab_spanner(label = md("**Comprehension**"), columns = matches("_comp")) %>% 
-  tab_spanner(label = md("**Production**"), columns = matches("_prod")) %>% 
-  fmt_number(2:9) %>% 
-  cols_merge_range(col_begin = "q2_5_comp", col_end = "q97_5_comp") %>% 
-  cols_merge_range(col_begin = "q2_5_prod", col_end = "q97_5_prod") %>% 
-  cols_label(
-    coefficient = md("**Coefficient**"),
-    estimate_comp = md("*Mean*"),
-    est_error_comp = md("*SE*"),
-    q2_5_comp = md("*95% CrI*"),
-    estimate_prod = md("*Mean*"),
-    est_error_prod = md("*SE*"),
-    q2_5_prod = md("*95% CrI*"),
-  ) %>% 
-  tab_footnote(
-    footnote = "SE: Standard Error of the mean",
-    locations = cells_column_labels(columns = matches("est_error"))
-  ) %>% 
-  tab_footnote(
-    footnote = "95% CrI: 95% credible interval",
-    locations = cells_column_labels(columns = matches("q2_5"))
-  )
-
-```
 
@@
 Finally this effect was negligible for children exposed more than 1 SD than the mean to their non-dominant language: the difference in comprehension probability between items in the dominant and the non-dominant language was roughly similar for cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[16]*100)`%, 95% CrI = [`r printnum(deltas$.lower[16]*100)`-`r printnum(deltas$.upper[16]*100)`]) and non-cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[13]*100)`%, 95% CrI = [`r printnum(deltas$.lower[13]*100)`-`r printnum(deltas$.upper[13]*100)`]).
-
-```{r marginalmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6.5, fig.height=6.5, out.width="20cm", fig.cap="Mean posterior probability and 95\\% CrI of responses, conditional to Dominance, Bilingualism, and Cognate."}
-
-post_preds %>% 
-  ggplot(aes(
-    age, .value,
-    colour = interaction(dominance, cognate, sep = " - "),
-    fill = interaction(dominance, cognate, sep = " - ")
-  )) +
-  facet_wrap(type~bilingualism) +
-  #geom_line(aes(group = interaction(item_dominance, cognate, doe, .draw)), size = 0.4) +
-  stat_lineribbon(.width = 0.95, colour = NA, alpha = 0.5) +
-  stat_summary(fun = "mean", geom = "line", size = 0.75) +
-  #geom_point(data = filter(proportion, type=="Understands"), alpha = 0.5) +
-  labs(x = "Age (months)", y = "P(Y|X)") +
-  guides(colour = guide_legend(ncol = 2)) +
-  scale_color_brewer(palette = "Dark2") +
-  scale_fill_brewer(palette = "Dark2") +
-  scale_y_continuous(limits = c(0, 1)) +
-  theme_custom() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank(),
-    legend.direction = "horizontal"
-  )
-
-```

### 2021-09-24 Rmd/Old/notes.Rmd
Add methods section and rename manuscript files

@@
 Age of acquisition (AoA) is then defined as the age at which the proportion of children that are able to understand or produce a given word exceeds a given threshold.
-For instance, the word *door* may be considered to be aquired when at least 60% of the participants in our sample are reported to understand the word.
+For instance, the word *door* may be considered to be acquired when at least 60% of the participants in our sample are reported to understand the word.
 
@@
 
-In this study, we model developmental tracjectories of individual items using logarithmic curves.
+In this study, we model developmental trajectories of individual items using logarithmic curves.
 We assume a common upper boundary for all items (i.e., 1).
@@
 We will follow a multilevel Bayesian approach to estimate the mid point and steepness of each word.
-We assume that mid-points will be generated from a normal distribution with mean 7.5 (age bin corresponsing to 22 months of age) and standard deviation 1.
+We assume that mid-points will be generated from a normal distribution with mean 7.5 (age bin corresponding to 22 months of age) and standard deviation 1.
 This means that the inflexion point (i.e., age of maximum steepness), 68% of the items will be located within 1 stand deviation at 22 months.

### 2021-09-24 Rmd/report.Rmd
Add methods section and rename manuscript files

@@
 participants %>% 
-  group_by(time_stamp, lp, dominant_language) %>% 
+  group_by(time_stamp, dominant_language) %>% 
   summarise(n = n(), .groups = "drop") %>% 
-  group_by(lp, dominant_language) %>% 
+  group_by(dominant_language) %>% 
   mutate(n = cumsum(n)) %>% 
-  ggplot(aes(time_stamp, n, colour = interaction(lp, dominant_language, sep = " - "), fill = interaction(lp, dominant_language, sep = " - "))) +
+  ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
   geom_line(size = 1) +
-  labs(x = "Date", y = "Number of responses", colour = "LP - Dominant language") +
+  labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
   scale_y_continuous(breaks = seq(0, 250, 50)) +
@@
 We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age` by `item`, and their correlation parameter.
-We then extended this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope.
+We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope.
 Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`.
@@
 nd <- expand.grid(
-  age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = c("Monolingual", "Bilingual"),
+  age = seq(min(responses$age), max(responses$age), length.out = 30),
+  doe_center = range(responses$doe_center),
   cognate = unique(items$cognate),
-  frequency_center = c(-1, 0, 1)
+  frequency_center = 0
 )
@@
 m <- add_fitted_draws(nd, fits_comp_prior$fit_2, n = 20, re_formula = NA) %>% 
-  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))
+  mutate(
+    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
+  )
 
 ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
-  facet_grid(lp~frequency_center) +
+  facet_grid(~doe_center) +
   geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
@@
 
-# emmeans
 nd <- expand.grid(
-  age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = c("Monolingual", "Bilingual"),
+  age = seq(min(responses$age), max(responses$age), length.out = 30),
+  doe_center = range(responses$doe_center),
   cognate = unique(items$cognate),
-  frequency_center = c(-1, 0, 1)
-)
-
-m <- add_fitted_draws(nd, fits_prod$fit_2, n = 20, re_formula = NA) %>% 
-  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))
+  frequency_center = 0
+)
+
+m <- add_fitted_draws(nd, fits_prod_prior$fit_2, n = 20, re_formula = NA) %>% 
+  mutate(
+    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
+  )
 
 ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
-  facet_grid(lp~frequency_center) +
+  facet_grid(~doe_center) +
   geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
@@
     footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
+    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
   ) %>% 
@@
 The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-
-var_levels <- c("b_Intercept", "b_age", "b_frequency_center", "b_lp1", "b_cognate1", "b_cognate2", "b_lp1:cognate1", "b_lp1:cognate2")
-var_labels <- c("Intercept", "Age (+1 month)", "Frequency (+1 SD)", "Lang.
-Profile (Bil.
-- Mon.)", "Cognateness 1 (IDC+NIC - NC)", "Cognateness 2 (IDC -NIC)", "Lang.
-Profile \u00d7 Cognateness 1", "Lang.
-Profile \u00d7 Cognateness 2")
+The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
+
+var_levels <- c("b_Intercept", "b_age", "b_frequency_center", "b_doe_center", "b_cognate1", "b_cognate2", "b_doe_center:cognate1", "b_doe_center:cognate2")
+var_labels <- c("Intercept", "Age (+1 month)", "Frequency (+1 SD)", "DoE (0% vs.
+100%)", "Cognateness 1 (IDC+NIC - NC)", "Cognateness 2 (IDC -NIC)", "DoE \u00d7 Cognateness 1", "DoE \u00d7 Cognateness 2")
 post <- gather_draws(fits_comp$fit_2, `b_.*`, regex = TRUE) %>% 
@@
   labs(x = "Comprehension posterior probability", y = "Variable", fill = "Variable") +
-  scale_x_continuous(labels = percent) +
+  scale_x_continuous(labels = percent, breaks = seq(-0.15, 2, 0.05), limits = c(-0.15, 0.2)) +
   theme_github() +
@@
 
-var_levels <- c("b_Intercept", "b_age", "b_frequency_center", "b_lp1", "b_cognate1", "b_cognate2", "b_lp1:cognate1", "b_lp1:cognate2")
-var_labels <- c("Intercept", "Age (+1 month)", "Frequency (+1 SD)", "Lang.
-Profile (Bil.
-- Mon.)", "Cognateness 1 (IDC+NIC - NC)", "Cognateness 2 (IDC -NIC)", "Lang.
-Profile \u00d7 Cognateness 1", "Lang.
-Profile \u00d7 Cognateness 2")
+var_levels <- c("b_Intercept", "b_age", "b_frequency_center", "b_doe_center", "b_cognate1", "b_cognate2", "b_doe_center:cognate1", "b_doe_center:cognate2")
+var_labels <- c("Intercept", "Age (+1 month)", "Frequency (+1 SD)", "DoE (0% vs.
+100%)", "Cognateness 1 (IDC+NIC - NC)", "Cognateness 2 (IDC -NIC)", "DoE \u00d7 Cognateness 1", "DoE \u00d7 Cognateness 2")
 post <- gather_draws(fits_prod$fit_2, `b_.*`, regex = TRUE) %>% 
@@
   stat_pointinterval(colour = "white") +
-  labs(x = "Production posterior probability", y = "Variable", fill = "Variable") +
-  scale_x_continuous(labels = percent) +
+  labs(x = "Comprehension posterior probability", y = "Variable", fill = "Variable") +
+  scale_x_continuous(labels = percent, breaks = seq(-0.15, 2, 0.05), limits = c(-0.15, 0.2)) +
   theme_github() +
@@
   mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
-  left_join(distinct(fits_comp$fit_2$data, id, lp))
-
-ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", height = 0) +
-  geom_point(size = 0.5, colour = "white") +
-  geom_vline(xintercept = 0.5, colour = "white") +
-  labs(x = "Participant average probability of comprehension",
-       y = "Participant") +
-  scale_x_continuous(limits = c(0, 1)) +
-  theme_github() +
-  theme(
+  left_join(distinct(fits_comp$fit_2$data, id, doe_center)) %>% 
+  clean_names()
+
+ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
+  geom_ribbon(alpha = 0.5, colour = NA, fill = "white") +
+  geom_line(aes(group = 1), size = 1, colour = "white") +
+  # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
+  # geom_point(size = 0.5) +
+  geom_vline(group = 1, xintercept = 0.5, colour = "white") +
+  labs(x = "Participant posterior probability of comprehension",
+       y = "Participant", colour = "Lang. Prof.") +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
+  theme_github() +
+  theme(
+    legend.position = "none",
     axis.text.y = element_blank(),
@@
   mutate(te = as.numeric(te)) %>% 
-  left_join(distinct(fits_comp$fit_2$data, te, lp))
-
-ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
-  geom_point(size = 0.5, colour = "white") +
+  left_join(distinct(fits_comp$fit_2$data, te, cognate)) %>% 
+  clean_names()
+
+ggplot(re_te, aes(estimate, reorder(te, estimate), xmin = q2_5, xmax = q97_5, 
+                  colour = cognate, fill = cognate)) +
+  facet_wrap(~cognate, scales = "free_y", ncol = 2) +
+  geom_ribbon(aes(group = cognate), alpha = 0.5, colour = NA) +
+  geom_line(aes(group = cognate), size = 1) +
+  # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
+  # geom_point(size = 0.5) +
   geom_vline(xintercept = 0.5, colour = "white") +
-  labs(x = "Item average probability of comprehension",
-       y = "Item") +
-  scale_color_brewer(palette = "Set1") +
-  scale_x_continuous(limits = c(0, 1)) +
-  theme_github() +
-  theme(
+  labs(x = "Item posterior probability of comprehension",
+       y = "Item", colour = "Cognate") +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
+  theme_github() +
+  theme(
+    legend.position = "none",
     axis.text.y = element_blank(),
@@
   )
+
 ```
@@
   mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
-  left_join(distinct(fits_prod$fit_2$data, id, lp))
-
-ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
-  geom_point(size = 0.5, colour = "white") +
+  left_join(distinct(fits_prod$fit_2$data, id)) %>% 
+  clean_names()
+
+ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
+  geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA, fill = "white") +
+  geom_line(aes(group = 1), size = 1, colour = "white") +
+  # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
+  # geom_point(size = 0.5) +
   geom_vline(xintercept = 0.5, colour = "white") +
-  labs(x = "Item average probability of production",
-       y = "Item") +
-  scale_color_brewer(palette = "Set1") +
-  scale_x_continuous(limits = c(0, 1)) +
-  theme_github() +
-  theme(
+  labs(x = "Participant posterior probability of production",
+       y = "Participant", colour = "Lang. Prof.") +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
+  theme_github() +
+  theme(
+    legend.position = "none",
     axis.text.y = element_blank(),
@@
   )
-
 ```
@@
   mutate(te = as.numeric(te)) %>% 
-  left_join(distinct(fits_prod$fit_2$data, te, lp))
-
-ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
-  geom_point(size = 0.5, colour = "white") +
+  left_join(distinct(fits_prod$fit_2$data, te, cognate)) %>% 
+  clean_names()
+
+ggplot(re_te, aes(estimate, reorder(te, estimate), xmin = q2_5, xmax = q97_5, 
+                  colour = cognate, fill = cognate)) +
+  facet_wrap(~cognate, scales = "free_y", ncol = 2) +
+  geom_ribbon(aes(group = cognate), alpha = 0.5, colour = NA) +
+  geom_line(aes(group = cognate), size = 1) +
+  # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
+  # geom_point(size = 0.5) +
   geom_vline(xintercept = 0.5, colour = "white") +
-  labs(x = "Item average probability of production",
-       y = "Item") +
-  scale_color_brewer(palette = "Set1") +
-  scale_x_continuous(limits = c(0, 1)) +
-  theme_github() +
-  theme(
+  labs(x = "Item posterior probability of production",
+       y = "Item", colour = "Cognate") +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
+  theme_github() +
+  theme(
+    legend.position = "none",
     axis.text.y = element_blank(),
@@
 nd <- expand.grid(
-  age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = c("Monolingual", "Bilingual"),
+  age = seq(min(responses$age), max(responses$age), length.out = 30),
+  doe_center = range(responses$doe_center),
   cognate = unique(items$cognate),
-  frequency_center = c(-1,  0, 1)
-)
-
-m <- add_fitted_draws(nd, fits_comp$fit_2, re_formula = NA, n = 25) %>% 
-  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))
+  frequency_center = 0
+)
+
+m <- add_fitted_draws(nd, fits_comp$fit_2, n = 20, re_formula = NA) %>% 
+  mutate(
+    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
+  )
 
 ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
-  facet_wrap(lp~frequency_center) +
-  geom_line(aes(group = interaction(.draw, frequency_center, cognate)), alpha = 0.5) +
-
+  facet_wrap(~doe_center) +
+  geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.25) +
   # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
-  # stat_summary(fun = "mean", geom = "line", size = 1) +
+  stat_summary(fun = "mean", geom = "line", size = 1) +  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
+  labs(x = "Age (months)", y = "Posterior probability of comprehension\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
+  scale_y_continuous(labels = scales::percent) +
+  theme_github() +
+  theme(
+    legend.position = "top",
+    legend.title = element_blank()
+  )
+
+```
+
+## Marginal means: Production
+
+```{r emmeans_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+nd <- expand.grid(
+  age = seq(min(responses$age), max(responses$age), length.out = 30),
+  doe_center = range(responses$doe_center),
+  cognate = unique(items$cognate),
+  frequency_center = 0
+)
+
+m <- add_fitted_draws(nd, fits_prod$fit_2, n = 20, re_formula = NA) %>% 
+  mutate(
+    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
+    )
+
+ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
+  facet_wrap(~doe_center) +
+  geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.1, size = 1) +
+  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
+  stat_summary(fun = "mean", geom = "line", size = 1) +
   geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
-  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
-  scale_y_continuous(labels = scales::percent) +
-  theme_github() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-
-```
-
-## Marginal means: Production
-
-```{r emmeans_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-nd <- expand.grid(
-  age = seq(min(responses$age), max(responses$age), length.out = 50),
-  lp = c("Monolingual", "Bilingual"),
-  cognate = unique(items$cognate),
-  frequency_center = c(-1,  0, 1)
-)
-
-m <- add_fitted_draws(nd, fits_prod$fit_2, re_formula = NA, n = 25) %>% 
-  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))
-
-ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
-  facet_wrap(lp~frequency_center) +
-  geom_line(aes(group = interaction(.draw, frequency_center, cognate)), alpha = 0.5) +
-
-  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
-  # stat_summary(fun = "mean", geom = "line", size = 1) +
-  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
-  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
+  labs(x = "Age (months)", y = "Posterior probability of production\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
   scale_y_continuous(labels = scales::percent) +

### 2021-09-24 Rmd/01_introduction.Rmd
Add methods section and rename manuscript files

@@
 
-Before the end of their first year of life, infants start directing their gaze to some object when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months] and parental reports [e.g., @fenson1994variability].
-During the last half of their second year, they acquire new words at an increasingly fast rate [@fenson1994variability; @bloom2002children; @bergelson2020comprehension].
-Despite the considerable variation in children's trajectories of vocabulary growth, such trajectories seem quite consistent across languages [@braginsky2019consistency].
-Nonetheless, most of the literature on early word acquisition has been conducted on monolingual children, and does not address the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
+```{r intro_participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+tar_load(participants)
+```
+
+Before the end of their first year of life, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental reports [e.g., @fenson2007macarthur].
+During the last half of their second year, they acquire new words at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002children; @bergelson2020comprehension].
+These early stages of lexical acquisition are characterised by substantial variation between children, reflected, for instance, on the variation on the number of words they know [i.e.
+vocabulary size; @fenson1994variability; @frank2021variability] or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, bates1994developmental].
+
+Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages.
+@tardif2008baby collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing.
+Since birth, these infants had been learning English, Cantonese and Mandarin, respectively.
+The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*).
+These results were later extended by @frank2021variability to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
+Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
 
 There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
-For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals (who only learnt English).
-In contrast, both groups knew a similar total amount of words, when both English and Spanish vocabularies were counted together.
+For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English.
+Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development.
 Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating].
-While these studies have mostly relied on samples of bilingual children learning two relatively distant languages (as it is the case of English and Spanish) it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
+While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
 What role could linguistic distance play during early vocabulary growth?
@@
 Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
-Particularly, when both languages are linguistically close, like Spanish and Catalan (both Roman languages), they are more likely to share a larger amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English--one Roman, the other Germanic.
+Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English (one Roman, the other Germanic).
 For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
-It could be the case that mapping two phonologically similar labels (cognates, like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
+It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
 If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
 
-A recent study by @floccia2018introduction provided evidence in line with this claim.
-The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers in the United Kingdom who were learning English and an additional language.
+@floccia2018introduction provided evidence in line with this claim.
+The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
 The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
 The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English.
-Phonological similarity was measuring by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
+Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
 The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other [@levenshtein1966binary].
-Floccia et al.
-divided the resulting score by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
+The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
 Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
@@
 Several recent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
-Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition.
-Therefore, children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
-If cognates are the reason for this facilitation effect, it could be predicted that cognates are acquired earlier than non-cognates.
-Likely as it might seem, an earlier age of acquisition for cognates is not the only scenario where Floccia et al.'s results are account for.
-For example, languages that share more cognates might also share a larger overlap in their phonemic inventories, which could facilitate the acquisition of all words, regardless of their cognate status.
+Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
+A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.
 
-However, the evidence on an earlier age of acquisition for cognates compared to non-cognates is scarce.
-To the best of our knowledge only two studies have provided direct data on this issue.
+In this study, we hypothesise that cognates are acquired earlier in age than non-cognates, which might explain why children learning languages that share many cognates show larger vocabulary sizes.
+Evidence supporting this claim is scarce.
+To the best of our knowledge, only two studies have provided direct data on this issue.
 First, @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
 Second, @schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
-The low sample size in these studies make it challenging to draw strong conclusions about the effect of cognateness in vocabulary growth.
+The low sample size in these studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
 Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
@@
 
-We identify two main reasons why we cannot conclude an earlier age of acquisition for cognates than for non cognate from Floccia et al.'s results.
-First, the response variable used was the proportion of words each participant understood and/or produced, from the list of lexical items in the vocabulary checklists.
+We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results.
+First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
 By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.
 non-cognates was lost.
-Second, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at this age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
-In summary, the evidence supporting an earlier age of acquisition of cognates vs.
+Second, all participants were aged ~24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
+In summary, the current evidence supporting an earlier age of acquisition of cognates vs.
 non-cognates presents some limitations that prevents drawing sound conclusions about this effect.
 
-Another worth-mentioning finding in Floccia et al.
-is that the increase in vocabulary size associated with linguistic similarity was larger in the additional language vocabulary than in English vocabulary.
+Our second hypothesis detaches from the fact that, for parallel activation to take place, and for cognateness to facilitate the acquisition of a word form in a given language, the child must be familiar with the (phonologically similar) translation of the word, in the other language.
+If the child is not previously familiar with one of the word forms, the acquisition of its translation in  language cannot be facilitated by cognateness.
+This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair.
+Given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014much; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+Therefore, we hypothesised that the effect of cognateness for words in the language of lower exposure would be larger than in the language of higher exposure.
+
+In line with this hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.
+reported was larger in the additional language vocabulary than in English vocabulary.
 Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
-Given that the relative amount of exposure to a given language is a strong predictor of children's vocabulary size in that language [e.g., @cattani2014much; @thordardottir2006bilingual], participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
+Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
 If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
+To test this hypothesis, we conditioned the effect of cognateness to the relative degree of exposure to the language.
 
-The aim of this study is to investigate the role of cognateness in bilingual word acquisition.
-We hypothesised that cognate words would be acquired earlier than non-cognate word, and that this difference would be larger in participants with lower exposure to the language such word belongs to.
-Using an designed online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of bilinguals aged 10 to 36 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+In summary, we investigated the role of cognateness in bilingual word acquisition (conditional to the degree of relative exposure to the language) as possible mechanism to account for the facilitation  of language similarity on bilingual children's vocabulary size.
+We hypothesised that cognate words would be acquired earlier than non-cognate words, and that this effect should be larger for words in the language with lower exposure.
+Using an online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
 We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.

### 2021-09-24 Rmd/05_appendix.Rmd
Add methods section and rename manuscript files

@@
 ```
+
+## Appendix: Model
+
+The likelihood of our extended model can be formalised as follows:
+
+$$
+\begin{aligned}
+y_{ip} &\sim Bernoulli(\mu_{ip}) \\
+\mu_{ip} &= (\beta_{0} + \beta_{0p} + \beta_{0i}) +(\beta_{1} + \beta_{1i}) \times Age_{pi} + (\beta_{2} + \beta_{2p}) \times Frequency_{pi} + ... \\
+&  (\beta_{3} + \beta_{3p})\times Cognate_{pi} + (\beta_{4} + \beta_{4i}) \times DoE_{pi} + ... \\
+& (\beta_{5} + \beta_{5i} + \beta_{5p}) \times (Cognate_{pi} \times DoE_{pi}) + \varepsilon_{pi}
+\end{aligned}
+$$
+
+Where:
+
+* $y_i$ is the probability of participant $p$ to understand translation equivalent $i$
+* $n_i$ is the total number of responses for translation equivalent $i$
+* $\beta_0$ is the intercept, which corresponds to the probability of a positive response when all predictors equal 0
+* $\beta_{1, \dots, 9}$ are the coefficients of the predictors
+
+$$
+\begin{aligned}
+\beta_0 &\sim Normal(0.5, 0.1) \\
+\beta_1 &\sim Normal(0.75, 0.1) \\
+\beta_{2-9} &\sim Normal(0, 0.1) \\
+\sigma_{0-9i} &\sim HalfNormal(0.2, \Sigma_{i}) \\
+
+\Sigma_i & =
+\begin{pmatrix} 
+\sigma_0i^2 & \dots & \rho_i \sigma_{0-9i} \\
+\vdots & \ddots & \vdots \\
+\rho_i \sigma_{0-9i} & \dots &\sigma_9i^2 \\
+\end{pmatrix}\\
+
+\Sigma_i & =
+\begin{pmatrix} 
+\sigma_0i^2 & \dots & \rho_i \sigma_{0-9i} \\
+\vdots & \ddots & \vdots \\
+\rho_i \sigma_{0-9i} & \dots &\sigma_9i^2 \\
+\end{pmatrix}\\
+
+\end{aligned}
+$$
+
+## Appendix: Stan code
+
+Stan code generated by `brms::stancode`:
+
+```{stan stan_model, echo=FALSE, output.var="model"}
+// generated with brms 2.15.0
+functions {
+/* compute correlated group-level effects
+* Args: 
+*   z: matrix of unscaled group-level effects
+*   SD: vector of standard deviation parameters
+*   L: cholesky factor correlation matrix
+* Returns: 
+*   matrix of scaled group-level effects
+*/ 
+matrix scale_r_cor(matrix z, vector SD, matrix L) {
+// r is stored in another dimension order than z
+return transpose(diag_pre_multiply(SD, L) * z);
+}
+}
+data {
+int<lower=1> N;  // total number of observations
+int Y[N];  // response variable
+int<lower=1> K;  // number of population-level effects
+matrix[N, K] X;  // population-level design matrix
+// data for group-level effects of ID 1
+int<lower=1> N_1;  // number of grouping levels
+int<lower=1> M_1;  // number of coefficients per level
+int<lower=1> J_1[N];  // grouping indicator per observation
+// group-level predictor values
+vector[N] Z_1_1;
+vector[N] Z_1_2;
+vector[N] Z_1_3;
+int<lower=1> NC_1;  // number of group-level correlations
+// data for group-level effects of ID 2
+int<lower=1> N_2;  // number of grouping levels
+int<lower=1> M_2;  // number of coefficients per level
+int<lower=1> J_2[N];  // grouping indicator per observation
+// group-level predictor values
+vector[N] Z_2_1;
+vector[N] Z_2_2;
+vector[N] Z_2_3;
+int<lower=1> NC_2;  // number of group-level correlations
+int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+int Kc = K - 1;
+matrix[N, Kc] Xc;  // centered version of X without an intercept
+vector[Kc] means_X;  // column means of X before centering
+for (i in 2:K) {
+means_X[i - 1] = mean(X[, i]);
+Xc[, i - 1] = X[, i] - means_X[i - 1];
+}
+}
+parameters {
+vector[Kc] b;  // population-level effects
+real Intercept;  // temporary intercept for centered predictors
+matrix[M_1, N_1] z_1;  // standardized group-level effects
+cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+matrix[M_2, N_2] z_2;  // standardized group-level effects
+cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+matrix[N_1, M_1] r_1;  // actual group-level effects
+// using vectors speeds up indexing in loops
+vector[N_1] r_1_1;
+vector[N_1] r_1_2;
+vector[N_1] r_1_3;
+matrix[N_2, M_2] r_2;  // actual group-level effects
+// using vectors speeds up indexing in loops
+vector[N_2] r_2_1;
+vector[N_2] r_2_2;
+vector[N_2] r_2_3;
+sd_1 = rep_vector(1, rows(sd_1));
+// compute actual group-level effects
+r_1 = scale_r_cor(z_1, sd_1, L_1);
+r_1_1 = r_1[, 1];
+r_1_2 = r_1[, 2];
+r_1_3 = r_1[, 3];
+// compute actual group-level effects
+r_2 = scale_r_cor(z_2, sd_2, L_2);
+r_2_1 = r_2[, 1];
+r_2_2 = r_2[, 2];
+r_2_3 = r_2[, 3];
+}
+model {
+// likelihood including constants
+if (!prior_only) {
+// initialize linear predictor term
+vector[N] mu = Intercept + rep_vector(0.0, N);
+for (n in 1:N) {
+// add more terms to the linear predictor
+mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n];
+}
+target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
+}
+// priors including constants
+target += normal_lpdf(b | 0, 3);
+target += normal_lpdf(Intercept | 0, 3);
+target += std_normal_lpdf(to_vector(z_1));
+target += lkj_corr_cholesky_lpdf(L_1 | 5);
+target += normal_lpdf(sd_2 | 0, 0.5)
+- 3 * normal_lccdf(0 | 0, 0.5);
+target += std_normal_lpdf(to_vector(z_2));
+target += lkj_corr_cholesky_lpdf(L_2 | 5);
+}
+generated quantities {
+// actual population-level intercept
+real b_Intercept = Intercept - dot_product(means_X, b);
+// compute group-level correlations
+corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+vector<lower=-1,upper=1>[NC_1] cor_1;
+// compute group-level correlations
+corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
+vector<lower=-1,upper=1>[NC_2] cor_2;
+// extract upper diagonal of correlation matrix
+for (k in 1:M_1) {
+for (j in 1:(k - 1)) {
+cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+}
+}
+// extract upper diagonal of correlation matrix
+for (k in 1:M_2) {
+for (j in 1:(k - 1)) {
+cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
+}
+}
+}
+```

### 2021-09-26 Rmd/01_introduction.Rmd
Merge into one manuscript

@@
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual; @fabian2016investigating].
+Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual].
 While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.

### 2021-09-26 Rmd/manuscript.Rmd
Add methods section and rename manuscript files
Merge into one manuscript

--- 2021-07-12 Rmd/trajectories_manuscript.Rmd
@@
   papaja::apa6_pdf:
-    latex_engine  : "xelatex"
+    latex_engine  : "pdflatex"
 
@@
 library(knitr)
-library(magrittr)
-library(tidyverse)
-library(readxl)
-library(lubridate)
-library(brms)
+library(here)
+library(dplyr)
+library(tidyr)
 library(tidybayes)
+library(scales)
 library(janitor)
-library(gt)
-library(data.table)
-library(here)
+library(stringr)
+library(ggplot2)
+library(patchwork)
 
-source(here("R", "analyses.R"))
+tar_load(multilex_data)
+tar_load(items)
+tar_load(participants)
+tar_load(responses)
+tar_load(loo_comp)
+tar_load(loo_prod)
+tar_load(fits_comp)
+tar_load(fits_prod)
+
+options(
+  knitr.kable.NA = "-",
+  knitr.duplicate.label = "allow",
+  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
+  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
+  ggplot2.continuous.fill = ggplot2::scale_color_gradient,
+  ggplot2.continuous.colour = ggplot2::scale_color_gradient
+)
+
+theme_set(theme_apa())
+
 ```
@@
 
-```{r introduction, child="01_introduction.Rmd"}
+Before the end of their first year of life, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental reports [e.g., @fenson2007macarthur].
+During the last half of their second year, they acquire new words at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002children; @bergelson2020comprehension].
+These early stages of lexical acquisition are characterised by substantial variation between children, reflected, for instance, on the variation on the number of words they know [i.e.
+vocabulary size; @fenson1994variability; @frank2021variability] or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, bates1994developmental].
+
+Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages.
+@tardif2008baby collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing.
+Since birth, these infants had been learning English, Cantonese and Mandarin, respectively.
+The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*).
+These results were later extended by @frank2021variability to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
+Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
+
+There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
+For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English.
+Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development.
+Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual].
+While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
+What role could linguistic distance play during early vocabulary growth?
+
+For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language.
+Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
+Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English (one Roman, the other Germanic).
+For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
+It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
+If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
+
+@floccia2018introduction provided evidence in line with this claim.
+The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
+The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
+The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English.
+Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
+The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other [@levenshtein1966binary].
+The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
+Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
+For example, English-Dutch bilinguals (22.14% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% phonological similarity) were able to produce in Mandarin.
+
+Floccia et al.
+pointed to *parallel activation* as the main mechanism underpinning their results.
+The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension.
+This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@thierry2007brain].
+One of the clearest examples of parallel activation was provided by @costa2000cognate.
+In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish.
+In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trials labels were non-cognates (*mesa*-*taula*, translations of *table*).
+Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
+Importantly, Spanish monolinguals did not show this effect.
+These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
+Several recent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
+Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
+A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.
+
+In this study, we hypothesise that cognates are acquired earlier in age than non-cognates, which might explain why children learning languages that share many cognates show larger vocabulary sizes.
+Evidence supporting this claim is scarce.
+To the best of our knowledge, only two studies have provided direct data on this issue.
+First, @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
+Second, @schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
+The low sample size in these studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
+Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
+In their discussion the authors state the following (pp.
+70):
+
+> "This  finding  also  provides  support  to  the  proposal  that  the  cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life."
+
+We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results.
+First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
+By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.
+non-cognates was lost.
+Second, all participants were aged ~24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
+In summary, the current evidence supporting an earlier age of acquisition of cognates vs.
+non-cognates presents some limitations that prevents drawing sound conclusions about this effect.
+
+Our second hypothesis detaches from the fact that, for parallel activation to take place, and for cognateness to facilitate the acquisition of a word form in a given language, the child must be familiar with the (phonologically similar) translation of the word, in the other language.
+If the child is not previously familiar with one of the word forms, the acquisition of its translation in  language cannot be facilitated by cognateness.
+This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair.
+Given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014much; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+Therefore, we hypothesised that the effect of cognateness for words in the language of lower exposure would be larger than in the language of higher exposure.
+
+In line with this hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.
+reported was larger in the additional language vocabulary than in English vocabulary.
+Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
+Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
+If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
+To test this hypothesis, we conditioned the effect of cognateness to the relative degree of exposure to the language.
+
+In summary, we investigated the role of cognateness in bilingual word acquisition (conditional to the degree of relative exposure to the language) as possible mechanism to account for the facilitation  of language similarity on bilingual children's vocabulary size.
+We hypothesised that cognate words would be acquired earlier than non-cognate words, and that this effect should be larger for words in the language with lower exposure.
+Using an online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.
+
+# Methods
+
+Data and materials are available at OSF (https://osf.io/hy984/), and code is available at GitHub (https://github.com/gongcastro/trajectories).
+
+## Participants
+
+```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+n_participants <- length(unique(participants$id))
+dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
 
@@
 
-# Method
+We collected data from `r printnum(n_participants)` children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
+All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
+Families were contacted by e-mail or phone if their child were aged between 10 and 34 months, and had not previously reported to be exposed more than 10% of the time to a language other than Spanish or Catalan.
+Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate.
+Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant languag (i.e., their degree of bilingualism).
+
+```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Participant sample size by age and degree of exposure to a second language.
+Information about DoE was provided by families when filling the questionnaire.
+For illustration purposes this table, DoEs were binned into 10\\% wide bins, and ages (in months) were round to the nearest integer.
+Hyphens indicate that no participants from that specific combination of age filled the questionnaire."}
+labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
+participants %>% 
+  mutate(
+    age = round(age),
+    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
+    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
+  ) %>% 
+  count(age, doe_2) %>% 
+  arrange(doe_2) %>% 
+  mutate(doe_2 = fct_inorder(doe_2)) %>% 
+  pivot_wider(names_from = age, values_from = n, values_fill = NA) %>% 
+  gt() %>% 
+  tab_spanner(label = "Number of participants by age (months)", columns = 2:17) %>% 
+  fmt_missing(everything(), everything(), missing_text = "-") %>% 
+  cols_label(
+    doe_2 = md("**L2 DoE**")
+  ) %>% 
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels(everything())
+  ) %>% 
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners(everything())
+  )
+
+```
+
+We use the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain.
+Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))` and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
+Highest educational attainment was similar across participants with different DoE to a second language, and therefore, differences in comprehension or production between participants with different language profiles are unlikely to stem from a difference in SES [@fernald2013ses] (see Table 2).
+
+```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Educational attainment.
+Proportion of highest parental educational attainment by DoE to the second language.
+For illustration purposes, DoEs were binned into 10\\% wide bins."}
+labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
+
+participants %>%
+  mutate(
+    age = round(age),
+    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
+    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
+  ) %>% 
+  group_by(doe_2) %>%
+  summarise(
+    n_no_education = length(edu_parent[edu_parent=="No education"])/n(),
+    n_primary = length(edu_parent[edu_parent=="Primary"])/n(),
+    n_secondary = length(edu_parent[edu_parent=="Secondary"])/n(),
+    n_complementary = length(edu_parent[edu_parent=="Complementary"])/n(),
+    n_vocational = length(edu_parent[edu_parent=="Vocational"])/n(),
+    n_university = length(edu_parent[edu_parent=="University"])/n(),
+    .groups = "drop"
+  ) %>% 
+  gt() %>% 
+  tab_spanner("Educational attainment", columns = 2:7) %>% 
+  cols_label(
+    doe_2 = md("**L2 DoE**"),
+    n_no_education = "No education",
+    n_primary = "Primary",
+    n_secondary = "Secondary",
+    n_complementary = "Complementary",
+    n_vocational = "Vocational",
+    n_university = "University"
+  ) %>% 
+  fmt_percent(2:7) %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels(everything())
+  ) %>% 
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners(everything())
+  )
+```
+
+## Questionnaire
+
+```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+pool <- multilex::pool %>% 
+  drop_na(cognate)
+n_categories <- distinct(pool, category) %>% nrow()
+n_items <- length(unique(pool$item))
+n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
+n_item_cognate <- count(pool, language, cognate) %>% 
+  pull(n) %>% 
+  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
+```
+
+The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
+Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*.
+We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
+For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
+
+For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
+Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
+These last families were randomly allocated into one of four different subsets of the complete list of items.
+These lists were carefully designed so that each contained a representative subsample of the items from the complete list.
+Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
+Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
+
+```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Lexical frequencies.
+Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
+items %>%
+  select(te, language, cognate, frequency) %>% 
+  left_join(select(multilex_data$pool, te, language, version)) %>% 
+  unnest(version) %>% 
+  group_by(language, version, cognate) %>% 
+  summarise(
+    frequency_mean = mean(frequency, na.rm = TRUE),
+    frequency_sd = sd(frequency, na.rm = TRUE),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    frequency_se = frequency_sd/sqrt(n),
+    frequency_ci_lower = frequency_mean + qnorm(0.025)*frequency_se,
+    frequency_ci_upper = frequency_mean + qnorm(0.975)*frequency_se
+  ) %>% 
+  pivot_wider(names_from = c("cognate"), values_from = c("n", matches("frequency"))) %>% 
+  clean_names() %>% 
+  mutate(
+    frequency_ci_identical_cognate = paste0(
+      "[", round(frequency_ci_lower_identical_cognate, 2), ", ", round(frequency_ci_lower_identical_cognate, 2), "]"
+    ),
+    frequency_ci_non_identical_cognate = paste0(
+      "[", round(frequency_ci_lower_non_identical_cognate, 2), ", ", round(frequency_ci_lower_non_identical_cognate, 2), "]"
+    ),
+    frequency_ci_non_cognate = paste0(
+      "[", round(frequency_ci_lower_non_cognate, 2), ", ", round(frequency_ci_lower_non_cognate, 2), "]"
+    ),
+  ) %>% 
+  relocate(
+    language, version,
+    n_identical_cognate,
+    frequency_mean_identical_cognate,
+    frequency_se_identical_cognate,
+    frequency_ci_identical_cognate,
+    matches("_non_identical"),
+    matches("_non_cognate")
+  ) %>% 
+  mutate(language = ifelse(language=="Catalan", "CAT", "SPA")) %>% 
+  select(-matches("sd|upper|lower")) %>% 
+  gt(groupname_col = "language") %>% 
+  fmt_number(matches("mean|se")) %>% 
+  tab_spanner("Identical cognates", columns = ends_with("_identical_cognate")) %>% 
+  tab_spanner("Non-identical cognates", columns = ends_with("_non_identical_cognate")) %>% 
+  tab_spanner("Non-cognates", columns = ends_with("_non_cognate")) %>% 
+  cols_label(
+    version = "List",
+    n_identical_cognate = md("N"),
+    n_non_identical_cognate = md("N"),
+    n_non_cognate = md("N"),
+    frequency_mean_identical_cognate = md("M"),
+    frequency_mean_non_identical_cognate = md("M"),
+    frequency_mean_non_cognate = md("M"),
+    frequency_se_identical_cognate = md("SD"),
+    frequency_se_non_identical_cognate = md("SD"),
+    frequency_se_non_cognate = md("SD"),
+    frequency_ci_identical_cognate = md("95% *CI"),
+    frequency_ci_non_identical_cognate = md("95% CI"),
+    frequency_ci_non_cognate = md("95% CI")
+  ) %>% 
+  cols_width(
+    starts_with("frequency_ci_") ~ px(100),
+    TRUE ~ px(40)
+  ) %>% 
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels(everything())
+  ) %>% 
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners(everything())
+  )
+
+```
+
+## Data analysis
+
+```{r n_responses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+n_obs <- nrow(responses)
+sum_obs <- responses %>%
+  group_by(te) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  summarise(
+    mean = mean(n),
+    sd = sd(n),
+    min = min(n),
+    max = max(n)
+  )
+
+```
+
+We gathered `r printnum(n_obs, big.mark = ",")` observations, with each observation corresponding to a single reponse of one participant to a single item.
+Translation equivalents (TEs) received an average of `r printnum(sum_obs$mean, digits = 2)` responses (*SD* = `r printnum(sum_obs$sd, digits = 2)`, *Min* = `r printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
+To investigate our hypotheses, we modelled the probability of children understanding or producing TEs using a logistic regression model with a Bernoulli distribution and a logit link function [@agresti2003categorical].
+We included several predictors of interest as fixed effects to adjust this probability to participants' and items' characteristics, namely:
+
+* *Age* of participant in months (`age`), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
+* *Lexical frequency* of item (`frequency_center`), Relative lexical frequencies (frequency per million words, $fpmw$) were extracted from the CHILDES data base, using the `childesr` R package [@braginsky2018childesr], and were transformed to Zipf scores [@van2014subtlex; @zipf1949human] ($log10(fpmw)+3$).
+Missing frequencies were imputed by multiple imputation using the `mice` R package [van2011mice].
+The resulting scores were then centred around the mean.
+* *Degree of exposure* to the language (`doe_center`), computed as the proportion of the time the child is exposed to the language the item belongs to.
+For instance, for a response to the item `perro` (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1).
+Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10% increase in exposure to the language of the item.
+* *Cognateness*, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist.
+We sum-coded two contrasts [@schad2020capitalize]: `cognate1`, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and `cognate2`, comparing IC (coded as +0.5) with NIC (coded as -0.5).
+
+We first fit a base model that only included the main effects of `age` and `frequency_center`  (Model 1) [see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches].
+We then entered our predictors of interest by adding `doe_center` (Model 2) and then `cognate` and its interaction with `doe_center` (Model 3).
+To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included *Participants* (`id`) and *TEs* (`te`) as grouping variables (i.e.
+random effects), with all possible random intercepts and slopes where appropriate [@barr2013random] (see Appendix for more details of the model).
+
+We adopted a Bayesian approach toward statistical inference.
+This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model [@mcelreath2018statistical].
+We fit the models using the R package `brms` [@burkner2017brms], an R interface to the probabilistic language Stan [@carpenter2017stan].
+We run four sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
+All R-hat values were smaller than 1.1, indicating that the chains converged successfully in all cases [@gelman1992inference] (see Appendix for an extended description of model diagnostics).
+
+We then compared comprehension and production models separately using leave-one-out cross-validation (LOO-CV) [@vehtari2017practical]^[Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson2019bayesian]].
+This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
+The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
+We compared the expected log posterior density (ELPD) of each each model to test which one fitted the data the best.
+
+We then explored the posterior distribution of each parameter in the comprehension and production models that fitted the data the best computing the 95% credible intervals and testing whether this interval excluded 0.
+Credible intervals (CrI) indicate the range of values we are 95% certain contain the true value of the parameter, conditional to the observed data.
+We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% CrI of their expected posterior marginal means using the `tidybayes` R package [@R-tidybayes].
+Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse].
 
 # Results
+
+```{r coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+coef_comp <- as.data.frame(fixef(fits_comp$fit_2))
+coef_prod <- as.data.frame(fixef(fits_prod$fit_2))
+
+```
+
+For both comprehension and production, the extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density.
+This estimate is several times larger than its standard error indicating high predictive accuracy.
+Table 3 summarises the LOO scores of the fitted models.
+In the next sections we report the fixed effects of the comprehension and production models separately.
+We report the mean of each coefficient's posterior distribution along with standard errors and 95% CrIs in Tables 4 and 5.
+For interpretability, we report the derivative of the logistic function for each coefficient^[The derivative is calculated as $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.
+This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.] for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable [@gelman2020regression].
+
+```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+l_comp <- loo_comp %>% 
+  as_tibble() %>% 
+  rownames_to_column("model") %>% 
+  select(-matches("p_loo|looic")) %>% 
+  relocate(model, matches("elpd_loo"), matches("diff")) %>%
+  mutate_all(as.numeric) %>% 
+  mutate(type = "Comprehension", model = paste0("Model ", model))
+
+l_prod <- loo_prod %>% 
+  as_tibble() %>% 
+  rownames_to_column("model") %>% 
+  select(-matches("p_loo|looic")) %>% 
+  relocate(model, matches("elpd_loo"), matches("diff")) %>%
+  mutate_all(as.numeric) %>% 
+  mutate(type = "Production", model = paste0("Model ", model))
+
+l <- bind_rows(l_comp, l_prod) %>% 
+  mutate_all(function(x) ifelse(x==0, NA, x))
+
+gt(l, groupname_col = "type") %>% 
+  tab_spanner(md("**ELPD</sub>**"), matches("elpd_loo")) %>% 
+  tab_spanner(md("**Diff.**"), matches("diff")) %>%
+  fmt_number(2:7) %>% 
+  fmt_missing(columns = everything(), missing_text = "-") %>% 
+  cols_label(
+    model = "Model",
+    # elpd
+    elpd_loo = md("*ELPD*"),
+    se_elpd_loo = md("*SE*"),
+    subsampling_se_elpd_loo = md("*SE* (sub)"),
+    # diff
+    elpd_diff = md("*diff*"),
+    se_diff = md("*SE*"),
+    subsampling_se_diff = md("*SE* (sub)"),
+  )
+```
+
+# Comprehension
+
+The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe_center))`%), and regardless of its cognate status (average acros the three levels).
+An increase in one standard deviation in age (1 *SD* = `r printnum(sd(responses$age))` months) was associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability.
+A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_comp["frequency", "Estimate"]*100/4)`% increment in comprehension probability.
+A 10% increase in relative language exposure made TEs `r printnum(coef_comp["doe_center", "Estimate"]*100/4)`% more likely to be understood.
+Cognates (identical and non-identical) were `r printnum(coef_comp["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates.
+The 95% credible interval of the `doe_center:cognate1` interaction did not exclude 0.
+Identical cognates were `r printnum(abs(coef_comp["cognate2", "Estimate"]*100/4))`% less like to be understood than non-identical cognates.
+The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero).
+Figure 1 summarises the posterior distribution of the fixed effects of the model.
+
+```{r fix_effets_comp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension.
+Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+The intercept has been transformed using the inverse logit to get the average probability of correct response.
+The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
+
+c_comp <- fixef(fits_comp$fit_2) %>% 
+  as.data.frame() %>%
+  rownames_to_column("term") %>% 
+  clean_names() %>% 
+  mutate(
+    estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
+    q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
+    q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
+    est_error = ifelse(term=="Intercept", inv_logit_scaled(est_error), est_error/4)
+  ) %>% 
+  rename_at(vars(-matches("term")), ~paste0("comp_", .)) %>% 
+  select(-one_of(c("rhat", "bulk_ess", "tail_ess")))
+
+c_prod <- fixef(fits_prod$fit_2) %>% 
+  as.data.frame() %>%
+  rownames_to_column("term") %>% 
+  clean_names() %>% 
+  mutate(
+    estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
+    q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
+    q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
+    est_error = ifelse(term=="Intercept", inv_logit_scaled(est_error), est_error/4)
+  ) %>% 
+  rename_at(vars(-matches("term")), ~paste0("prod_", .)) %>% 
+  select(-term, -one_of(c("term", "rhat", "bulk_ess", "tail_ess")))
+
+c <- bind_cols(c_comp, c_prod)
+
+gt(c) %>% 
+  fmt_percent(matches("estimate|error")) %>% 
+  fmt_percent(matches("_q")) %>% 
+  tab_spanner("Comprehension", matches("comp_")) %>% 
+  tab_spanner("Production", matches("prod_")) %>% 
+  cols_merge(vars("comp_q2_5", "comp_q97_5"), pattern = "[{1}, {2}]") %>% 
+  cols_merge(vars("prod_q2_5", "prod_q97_5"), pattern = "[{1}, {2}]") %>%
+  cols_label(
+    term = md("**Predictor**"),
+    comp_estimate = md("**Mean**"),
+    comp_est_error = md("**SE**"),
+    comp_q2_5 = md("**95\\% CrI**"),
+    comp_q2_5 = md("**95\\% CrI**"),
+    prod_q2_5 = md("**95\\% CrI**"),
+    prod_estimate = md("**Mean**"),
+    prod_est_error = md("**SE**"),
+    prod_q2_5 = md("**95\\% CrI**")
+  ) %>% 
+  tab_footnote(
+    footnote = "Transformed using the inverse logit to get the average probability of correct response",
+    locations = cells_body(columns = "term", rows = term=="Intercept")
+  ) %>% 
+  tab_footnote(
+    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
+    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
+  ) %>% 
+  cols_align(
+    align = c("left"),
+    columns = 2:4
+  ) 
+```
+
+## Production
+
+The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_prod["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe_center))`%), and regardless of its cognate status (average acros the three levels).
+An increase in one standard deviation in age (1 *SD* = `r printnum(sd(responses$age))` months) was associated with a `r printnum(coef_prod["age", "Estimate"]*100/4)`% increment in comprehension probability.
+A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_prod["frequency", "Estimate"]*100/4)`% increment in comprehension probability.
+A 10% increase in relative language exposure made TEs `r printnum(coef_prod["doe_center", "Estimate"]*100/4)`% more likely to be understood.
+Cognates (identical and non-identical) were `r printnum(coef_prod["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates.
+The 95% credible interval of the `doe_center:cognate1` interaction did not exclude 0.
+Identical cognates were `r printnum(abs(coef_comp["cognate2", "Estimate"]*100/4))`% less like to be understood than non-identical cognates.
+The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero).
+Figure 1 summarises the posterior distribution of the fixed effects of the model.
+
+```{r fix_effets_prod, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for production Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+The intercept has been transformed using the inverse logit to get the average probability of correct response.
+The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
+
+summary(fits_prod$fit_2)$fixed %>% 
+  as.data.frame() %>%
+  rownames_to_column("term") %>% 
+  mutate(
+    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
+    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
+    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
+    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
+  ) %>% 
+  gt() %>% 
+  fmt_percent(2:5) %>% 
+  fmt_number(6:6, decimals = 2) %>% 
+  fmt_number(7:8, decimals = 0) %>% 
+  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
+  cols_label(
+    term = md("**Predictor**"),
+    Estimate = md("**Mean**"),
+    Est.Error = md("**SEM**"),
+    "l-95% CI" = md("**95\\% CrI**"),
+    Rhat = md("**Rhat**"),
+    Bulk_ESS = md("**Bulk ESS**"),
+    Tail_ESS = md("**Tail ESS**")
+  ) %>% 
+  tab_footnote(
+    footnote = "Transformed using the inverse logit to get the average probability of correct response",
+    locations = cells_body(columns = "term", rows = term=="Intercept")
+  ) %>% 
+  tab_footnote(
+    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
+    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
+  ) %>% 
+  tab_footnote(
+    footnote = "ESS: Effective sample size",
+    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
+  ) %>% 
+  cols_align(
+    align = c("center"),
+    columns = 2:4
+  ) 
+```
+
+```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Expected marginal mean and 95% credible interval of the posterior probability of comprehension and production for identical cognates, non-identical cognates, and non cognates, plotted separately for three levels of degree of exposue to the language: 0% (no exposure, monolingual), 50% (exposed half of the time, balanced bilingual), and 100% (exposed all of the time, monolingual).
+Dots represent the mean, and error bars represent the range of values within the true values is contained with 50% (thick line) and 95% (thin line) cprobability, given the data."}
+
+nd <- expand.grid(
+  age = mean(responses$age),
+  doe_center = c(min(responses$doe_center), mean(responses$doe_center), max(responses$doe_center)),
+  cognate = unique(items$cognate),
+  frequency_center = 0
+)
+
+m_comp <- add_fitted_draws(nd, fits_comp$fit_2, n = 20, re_formula = NA) %>% 
+  mutate(
+    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%"))
+  )
+
+m_prod <- add_fitted_draws(nd, fits_prod$fit_2, n = 20, re_formula = NA) %>% 
+  mutate(
+    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%"))
+  )
+
+m <- bind_rows(m_comp, m_prod, .id = "type") %>% 
+  mutate(type = ifelse(type==1, "Comprehension", "Production"))
+
+ggplot(m, aes(cognate, .value, colour = cognate, fill = cognate, shape = doe_center)) +
+  facet_grid(~type, scales = "free_y") +
+  # geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.25) +
+  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
+  # stat_summary(fun = "mean", geom = "line", size = 1) + 
+  geom_hline(yintercept = 0.5, linetype = "dotted") +
+  stat_pointinterval(.width = 0.95, position = position_dodge(width = 0.5), size = 0.75, point_size = 2) +
+  scale_y_continuous(labels = scales::percent, limits = c(0.1, 0.9), breaks = seq(0, 1, 0.10)) +
+  labs(x = "Cognateness", y = "Posterior probability",
+       colour = "Cognateness", fill = "Cognateness", shape = "DoE") +
+  theme(
+    legend.position = "right",
+    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
+    axis.title.x = element_blank(),
+    legend.title = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+
+```
 
@@
 
-r_refs(file = "trajectories.bib")
+r_refs(file = "manuscript.bib")
 

### 2021-09-26 Rmd/manuscript.tex
Merge into one manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
  pdfauthor={Gonzalo Garcia-Castro1, Daniela Avila-Varela1, \& Nuria Sebastian-Galles1},
  pdflang={en-EN},
  pdfkeywords={lexical acquisition, vocabulary, bilingualism},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Cross-linguistic similarity and  word acquisition}
\keywords{lexical acquisition, vocabulary, bilingualism\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{The role of cross-linguistic lexical similarity on bilingual word acquisition}
\author{Gonzalo Garcia-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Nuria Sebastian-Galles\textsuperscript{1}}
\date{}

\authornote{

Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain

Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain. E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain}

\abstract{
Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al.~(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Before the end of their first year of life, infants start directing their gaze to some objects when hearing their labels, according to both experimental data (Bergelson \& Swingley, 2012, 2015; Jusczyk \& Aslin, 1995; Tincoff \& Jusczyk, 1999) and parental reports (e.g., Fenson \& others, 2007). During the last half of their second year, they acquire new words at an increasingly fast rate (Bergelson, 2020; Bloom, 2002; Fenson et al., 1994; Goldfield \& Reznick, 1990; Mayor \& Plunkett, 2011). These early stages of lexical acquisition are characterised by substantial variation between children, reflected, for instance, on the variation on the number of words they know {[}i.e.~vocabulary size; Fenson et al. (1994); Frank, Braginsky, Yurovsky, and Marchman (2021){]} or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words (e.g., Nelson, 1973, p. bates1994developmental).

Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages. Tardif et al. (2008) collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing. Since birth, these infants had been learning English, Cantonese and Mandarin, respectively. The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (\emph{daddy}, \emph{mommy}), social routines (\emph{bye}, \emph{uh-oh}) or animals (\emph{woof-woof}). These results were later extended by Frank, Braginsky, Yurovsky, and Marchman (2021) to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages. Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.

There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated. For example, Hoff et al. (2012) found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English. Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development. Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated (Gonzalez-Barrero, Schott, \& Byers-Heinlein, 2020; Oller \& Eilers, 2002; J. L. Patterson, 2004; J. Patterson, Pearson, \& Goldstein, 2004; Pearson \& Fernández, 1994; Pearson, Fernández, \& Oller, 1993; Petitto et al., 2001; Smithson, Paradis, \& Nicoladis, 2014). While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals. What role could linguistic distance play during early vocabulary growth?

For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language. Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees. Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English (one Roman, the other Germanic). For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear \emph{puerta} and \emph{porta} (cognates), whereas a Spanish-English bilingual might hear \emph{puerta} and \emph{door} (non-cognates). It could be the case that mapping two phonologically similar labels (cognates like \emph{puerta}-\emph{porta}) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like \emph{puerta} and \emph{door}). If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.

Floccia et al. (2018) provided evidence in line with this claim. The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language. The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English. Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions. The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other (Levenshtein \& others, 1966). The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes. Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair. For example, English-Dutch bilinguals (22.14\% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97\% phonological similarity) were able to produce in Mandarin.

Floccia et al.~pointed to \emph{parallel activation} as the main mechanism underpinning their results. The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension. This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production (Costa, Caramazza, \& Sebastian-Galles, 2000; Hoshino \& Kroll, 2008) or comprehension (Thierry \& Wu, 2007). One of the clearest examples of parallel activation was provided by Costa, Caramazza, and Sebastian-Galles (2000). In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish. In half of the trials, the object labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre}, translations of \emph{tree}), whereas in the other half of the trials labels were non-cognates (\emph{mesa}-\emph{taula}, translations of \emph{table}). Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. Importantly, Spanish monolinguals did not show this effect. These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures. Several recent studies have also provided similar evidence in comprehension in children (Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013; e.g., Von Holzen \& Mani, 2012). Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes. A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.

In this study, we hypothesise that cognates are acquired earlier in age than non-cognates, which might explain why children learning languages that share many cognates show larger vocabulary sizes. Evidence supporting this claim is scarce. To the best of our knowledge, only two studies have provided direct data on this issue. First, Bosch and Ramon-Casas (2014) used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates. Second, Schelletter (2002) reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average. The low sample size in these studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth. Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly. In their discussion the authors state the following (pp.~70):

\begin{quote}
``This finding also provides support to the proposal that the cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life.''
\end{quote}

We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results. First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists. By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.~non-cognates was lost. Second, all participants were aged \textasciitilde24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item. In summary, the current evidence supporting an earlier age of acquisition of cognates vs.~non-cognates presents some limitations that prevents drawing sound conclusions about this effect.

Our second hypothesis detaches from the fact that, for parallel activation to take place, and for cognateness to facilitate the acquisition of a word form in a given language, the child must be familiar with the (phonologically similar) translation of the word, in the other language. If the child is not previously familiar with one of the word forms, the acquisition of its translation in language cannot be facilitated by cognateness. This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair. Given that children are more likely to acquire words from languages to which they are exposed more often (Cattani et al., 2014; David \& Wei, 2008; Thordardottir, Rothenberg, Rivard, \& Naves, 2006), the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness. Therefore, we hypothesised that the effect of cognateness for words in the language of lower exposure would be larger than in the language of higher exposure.

In line with this hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.~reported was larger in the additional language vocabulary than in English vocabulary. Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language. Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language. If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form. To test this hypothesis, we conditioned the effect of cognateness to the relative degree of exposure to the language.

In summary, we investigated the role of cognateness in bilingual word acquisition (conditional to the degree of relative exposure to the language) as possible mechanism to account for the facilitation of language similarity on bilingual children's vocabulary size. We hypothesised that cognate words would be acquired earlier than non-cognate words, and that this effect should be larger for words in the language with lower exposure. Using an online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of children aged 19 to 33 months learning Catalan and/or Spanish, with varying degrees of exposure to each language. We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Data and materials are available at OSF (\url{https://osf.io/hy984/}), and code is available at GitHub (\url{https://github.com/gongcastro/trajectories}).

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We collected data from 289 children from the Metropolitan Area of Barcelona between 28th October, 2019 and 24th July, 2021. All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX. Families were contacted by e-mail or phone if their child were aged between 10 and 34 months, and had not previously reported to be exposed more than 10\% of the time to a language other than Spanish or Catalan. Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate. Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant languag (i.e., their degree of bilingualism).

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{ccccccccccccccccc}
\toprule
& \multicolumn{16}{c}{Number of participants by age (months)} \\ 
 \cmidrule(lr){2-17}
\textbf{L2 DoE} & 19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 & 28 & 29 & 30 & 31 & 32 & 33 & 34 \\ 
\midrule
0-10\% & 4 & 6 & 10 & 11 & 7 & 5 & 8 & 12 & 13 & 2 & 9 & 2 & 1 & 2 & 2 & - \\ 
10-20\% & - & 6 & 11 & 3 & 5 & 2 & 6 & 4 & 7 & 2 & 6 & - & 2 & 1 & - & - \\ 
20-30\% & - & 1 & 13 & 9 & 7 & 5 & 12 & 5 & 7 & 5 & 1 & - & 2 & - & - & 1 \\ 
30-40\% & 1 & 4 & 12 & 6 & 2 & 6 & 6 & 4 & 9 & 2 & 2 & 4 & 1 & - & - & - \\ 
40-50\% & 3 & 2 & 6 & 7 & 3 & 1 & 5 & 1 & - & 2 & 2 & 1 & - & 1 & 1 & - \\ 
\bottomrule
\end{longtable}

We use the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: \emph{No education}, \emph{Primary}, \emph{Secondary}, \emph{Complementary}, \emph{Vocational}, and \emph{University}, in line with the current educational system in Spain. Most families reported university studies (244, 0.84), followed by families were the highest educational attainment were vocational studies (51, 18\%), complementary studies (5, 2\%), secondary education (2, 1\%, 0\%), no formal education (2, 1\% and primary education (1, \textless1\%). Highest educational attainment was similar across participants with different DoE to a second language, and therefore, differences in comprehension or production between participants with different language profiles are unlikely to stem from a difference in SES (Fernald, Marchman, \& Weisleder, 2013) (see Table 2).

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{crrrrrr}
\toprule
& \multicolumn{6}{c}{Educational attainment} \\ 
 \cmidrule(lr){2-7}
\textbf{L2 DoE} & No education & Primary & Secondary & Complementary & Vocational & University \\ 
\midrule
0-10\% & $3.19\%$ & $3.19\%$ & $3.19\%$ & $7.45\%$ & $17.02\%$ & $81.91\%$ \\ 
10-20\% & $1.82\%$ & $1.82\%$ & $1.82\%$ & $1.82\%$ & $20.00\%$ & $81.82\%$ \\ 
20-30\% & $2.94\%$ & $1.47\%$ & $2.94\%$ & $1.47\%$ & $17.65\%$ & $80.88\%$ \\ 
30-40\% & $0.00\%$ & $0.00\%$ & $0.00\%$ & $0.00\%$ & $23.73\%$ & $76.27\%$ \\ 
40-50\% & $5.71\%$ & $5.71\%$ & $5.71\%$ & $5.71\%$ & $11.43\%$ & $80.00\%$ \\ 
\bottomrule
\end{longtable}

\hypertarget{questionnaire}{%
\subsection{Questionnaire}\label{questionnaire}}

The questionnaire was implemented on-line using the formR platform (Arslan, Walther, \& Tata, 2020), and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists. Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (Hamilton, Plunkett, \& Schafer, 2000) and consisted in two lists of words: one in Catalan and one in Spanish. The Catalan inventory contained 777 items (196 cognates, 581 non-cognates) and the Spanish inventory contained 782 (197 cognates, 585 non-cognates). Items in one language were translation equivalents of the items in the other language (e.g., whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the word \emph{perro} was included in the Spanish inventory), roughly following a one-to-one mapping. When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar} and \emph{terminar}), or merged them into a single items (e.g., Spanish \emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}. We included items from a diverse sample of 26 semantic/functional categories (see Appendix 1). For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.

For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand \emph{and} say it, or did not understand or say it (checked out by default). Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\textasciitilde400 translation equivalents, \textasciitilde400 items in Catalan, \textasciitilde400 items in Spanish). These last families were randomly allocated into one of four different subsets of the complete list of items. These lists were carefully designed so that each contained a representative subsample of the items from the complete list. Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists. Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions. Table 2 in Appendix 1 shows the distribution of items across questionnaire versions. We excluded from the analysis multi-word items (e.g., \emph{barrita de cereales} {[}cereal bar{]}) and items that included more than one word-form (e.g., \emph{mono / mico}). Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lcrrlcrrlcrrl}
\toprule
& \multicolumn{4}{c}{Identical cognates} & \multicolumn{4}{c}{Non-identical cognates} & \multicolumn{4}{c}{Non-cognates} \\ 
 \cmidrule(lr){2-5}\cmidrule(lr){6-9}\cmidrule(lr){10-13}
List & N & M & SD & 95\% *CI & N & M & SD & 95\% CI & N & M & SD & 95\% CI \\ 
\midrule
\multicolumn{1}{l}{CAT} \\ 
\midrule
A & 41 & $4.28$ & $0.10$ & [4.09, 4.09] & 137 & $4.49$ & $0.05$ & [4.39, 4.39] & 105 & $4.23$ & $0.06$ & [4.12, 4.12] \\ 
B & 44 & $4.46$ & $0.11$ & [4.25, 4.25] & 131 & $4.43$ & $0.06$ & [4.32, 4.32] & 101 & $4.24$ & $0.07$ & [4.11, 4.11] \\ 
C & 37 & $4.29$ & $0.11$ & [4.08, 4.08] & 135 & $4.41$ & $0.06$ & [4.3, 4.3] & 102 & $4.26$ & $0.06$ & [4.14, 4.14] \\ 
D & 38 & $4.25$ & $0.08$ & [4.08, 4.08] & 134 & $4.38$ & $0.06$ & [4.27, 4.27] & 101 & $4.27$ & $0.06$ & [4.15, 4.15] \\ 
\midrule
\multicolumn{1}{l}{SPA} \\ 
\midrule
A & 42 & $4.18$ & $0.11$ & [3.96, 3.96] & 137 & $4.40$ & $0.06$ & [4.28, 4.28] & 106 & $4.03$ & $0.07$ & [3.9, 3.9] \\ 
B & 48 & $4.29$ & $0.10$ & [4.09, 4.09] & 136 & $4.36$ & $0.07$ & [4.22, 4.22] & 105 & $4.07$ & $0.07$ & [3.93, 3.93] \\ 
C & 40 & $4.19$ & $0.13$ & [3.93, 3.93] & 139 & $4.32$ & $0.07$ & [4.19, 4.19] & 103 & $4.10$ & $0.07$ & [3.96, 3.96] \\ 
D & 42 & $4.17$ & $0.12$ & [3.94, 3.94] & 137 & $4.33$ & $0.06$ & [4.21, 4.21] & 105 & $4.09$ & $0.07$ & [3.96, 3.96] \\ 
\bottomrule
\end{longtable}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We gathered 192059 observations, with each observation corresponding to a single reponse of one participant to a single item. Translation equivalents (TEs) received an average of 305.83 responses (\emph{SD} = 189.00, \emph{Min} = 71, Max = 622), both languages summed together. To investigate our hypotheses, we modelled the probability of children understanding or producing TEs using a logistic regression model with a Bernoulli distribution and a logit link function (Agresti, 2003). We included several predictors of interest as fixed effects to adjust this probability to participants' and items' characteristics, namely:

\begin{itemize}
\tightlist
\item
  \emph{Age} of participant in months (\texttt{age}), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
\item
  \emph{Lexical frequency} of item (\texttt{frequency\_center}), Relative lexical frequencies (frequency per million words, \(fpmw\)) were extracted from the CHILDES data base, using the \texttt{childesr} R package (M. Braginsky, Sanchez, \& Yurovsky, 2018), and were transformed to Zipf scores (Van Heuven, Mandera, Keuleers, \& Brysbaert, 2014; Zipf, 1949) (\(log10(fpmw)+3\)). Missing frequencies were imputed by multiple imputation using the \texttt{mice} R package {[}van2011mice{]}. The resulting scores were then centred around the mean.
\item
  \emph{Degree of exposure} to the language (\texttt{doe\_center}), computed as the proportion of the time the child is exposed to the language the item belongs to. For instance, for a response to the item \texttt{perro} (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1). Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10\% increase in exposure to the language of the item.
\item
  \emph{Cognateness}, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist. We sum-coded two contrasts (Schad, Vasishth, Hohenstein, \& Kliegl, 2020): \texttt{cognate1}, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and \texttt{cognate2}, comparing IC (coded as +0.5) with NIC (coded as -0.5).
\end{itemize}

We first fit a base model that only included the main effects of \texttt{age} and \texttt{frequency\_center} (Model 1) (see Mika Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019; Kachergis, Marchman, \& Frank, 2021 for similar approaches). We then entered our predictors of interest by adding \texttt{doe\_center} (Model 2) and then \texttt{cognate} and its interaction with \texttt{doe\_center} (Model 3). To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included \emph{Participants} (\texttt{id}) and \emph{TEs} (\texttt{te}) as grouping variables (i.e.~random effects), with all possible random intercepts and slopes where appropriate (Barr, Levy, Scheepers, \& Tily, 2013) (see Appendix for more details of the model).

We adopted a Bayesian approach toward statistical inference. This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model (McElreath, 2018).
We fit the models using the R package \texttt{brms} (Bürkner, 2017), an R interface to the probabilistic language Stan (Carpenter et al., 2017). We run four sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain. All R-hat values were smaller than 1.1, indicating that the chains converged successfully in all cases (Gelman \& Rubin, 1992) (see Appendix for an extended description of model diagnostics).

We then compared comprehension and production models separately using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017)\footnote{Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples (Magnusson, Andersen, Jonasson, \& Vehtari, 2019)}. This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit. The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting. We compared the expected log posterior density (ELPD) of each each model to test which one fitted the data the best.

We then explored the posterior distribution of each parameter in the comprehension and production models that fitted the data the best computing the 95\% credible intervals and testing whether this interval excluded 0. Credible intervals (CrI) indicate the range of values we are 95\% certain contain the true value of the parameter, conditional to the observed data. We performed follow-up tests on interactions whose 95\% credible interval excluded 0 by comparing the 95\% CrI of their expected posterior marginal means using the \texttt{tidybayes} R package (Kay, 2020). Data processing and visualisation was done in R (R Core Team, 2020) using the \texttt{tidyverse} family of packages (Wickham et al., 2019).

\hypertarget{results}{%
\section{Results}\label{results}}

For both comprehension and production, the extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density. This estimate is several times larger than its standard error indicating high predictive accuracy. Table 3 summarises the LOO scores of the fitted models. In the next sections we report the fixed effects of the comprehension and production models separately. We report the mean of each coefficient's posterior distribution along with standard errors and 95\% CrIs in Tables 4 and 5. For interpretability, we report the derivative of the logistic function for each coefficient\footnote{The derivative is calculated as \(\hat{\beta_j}/4\), where \(\hat{\beta_j}\) is the estimated mean of the posterior distribution of coefficient \(j\). This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.} for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable (Gelman, Hill, \& Vehtari, 2020).

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lrrrrrrrrr}
\toprule
& \multicolumn{3}{c}{\textbf{LOOELPD}} & \multicolumn{3}{c}{\textbf{LOOIC}} & \multicolumn{3}{c}{\textbf{LOODiff.}} \\ 
 \cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
Model & \emph{ELPD} & \emph{SE} & \emph{SE} (sub) & \emph{LOO-IC} & \emph{SE} & \emph{SE} (sub) & \emph{diff} & \emph{SE} & \emph{SE} (sub) \\ 
\midrule
\multicolumn{1}{l}{Comprehension} \\ 
\midrule
Model 1 & $-83,789.09$ & $211.11$ & $114.66$ & $167,578.18$ & $422.21$ & $229.32$ & $0.00$ & $0.00$ & $0.00$ \\ 
Model 2 & $-233,348.92$ & $682.60$ & $3,067.85$ & $466,697.84$ & $1,365.20$ & $6,135.70$ & $149,559.83$ & $714.50$ & $3,069.99$ \\ 
Model 3 & $-242,729.62$ & $721.85$ & $3,523.44$ & $485,459.24$ & $1,443.70$ & $7,046.88$ & $158,940.53$ & $752.08$ & $3,525.30$ \\ 
Model 1 & $-59,417.81$ & $209.84$ & $103.31$ & $118,835.62$ & $419.69$ & $206.61$ & $0.00$ & $0.00$ & $0.00$ \\ 
Model 2 & $-60,421.29$ & $206.95$ & $80.42$ & $120,842.58$ & $413.91$ & $160.85$ & $1,003.48$ & $294.73$ & $130.92$ \\ 
Model 3 & $-60,481.42$ & $208.06$ & $105.51$ & $120,962.85$ & $416.12$ & $211.02$ & $1,063.61$ & $295.50$ & $147.67$ \\ 
\bottomrule
\end{longtable}

\hypertarget{comprehension}{%
\section{Comprehension}\label{comprehension}}

The average probability of participants understanding a TE was 53.47\% (intercept) when all predictors were equal to 0: mean participant age (0.00 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels). An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 2.93\% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 7.09\% increment in comprehension probability. A 10\% increase in relative language exposure made TEs -3.00\% more likely to be understood. Cognates (identical and non-identical) were 1.95\% more like to be understood than non-cognates. The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0. Identical cognates were 2.06\% less like to be understood than non-identical cognates. The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero). Figure 1 summarises the posterior distribution of the fixed effects of the model.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{llllrrr}
\toprule
& \multicolumn{3}{c}{Comprehension} & \multicolumn{3}{c}{Production} \\ 
 \cmidrule(lr){2-4}\cmidrule(lr){5-7}
\textbf{Predictor} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} \\ 
\midrule
Intercept\textsuperscript{1} & $53.47\%$ & $51.30\%$ & [$51.07\%$, $55.99\%$] & $16.24\%$ & $51.39\%$ & [$14.92\%$, $17.82\%$] \\ 
age\textsuperscript{2} & $2.93\%$ & $0.12\%$ & [$2.69\%$, $3.15\%$] & $5.57\%$ & $0.14\%$ & [$5.27\%$, $5.83\%$] \\ 
frequency\_center & $7.09\%$ & $0.51\%$ & [$6.09\%$, $8.00\%$] & $7.57\%$ & $0.61\%$ & [$6.34\%$, $8.77\%$] \\ 
doe\_center\textsuperscript{2} & $-3.00\%$ & $0.56\%$ & [$-4.12\%$, $-1.92\%$] & $-0.67\%$ & $0.57\%$ & [$-1.74\%$, $0.41\%$] \\ 
cognate1\textsuperscript{2} & $1.95\%$ & $2.19\%$ & [$-2.33\%$, $6.16\%$] & $2.91\%$ & $2.01\%$ & [$-1.15\%$, $6.87\%$] \\ 
cognate2 & $-2.06\%$ & $1.85\%$ & [$-5.71\%$, $1.49\%$] & $-0.46\%$ & $2.09\%$ & [$-4.88\%$, $3.30\%$] \\ 
doe\_center:cognate1\textsuperscript{2} & $0.60\%$ & $0.72\%$ & [$-0.72\%$, $2.13\%$] & $-0.72\%$ & $0.70\%$ & [$-2.05\%$, $0.62\%$] \\ 
doe\_center:cognate2 & $1.76\%$ & $0.68\%$ & [$0.47\%$, $3.10\%$] & $0.60\%$ & $0.72\%$ & [$-0.82\%$, $2.01\%$] \\ 
\bottomrule
\end{longtable}
\vspace{-5mm}
\begin{minipage}{\linewidth}
\textsuperscript{1}Transformed using the inverse logit to get the average probability of correct response \\ 
\textsuperscript{2}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
\end{minipage}

\hypertarget{production}{%
\subsection{Production}\label{production}}

The average probability of participants understanding a TE was 16.24\% (intercept) when all predictors were equal to 0: mean participant age (0.00 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels). An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 5.57\% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 7.57\% increment in comprehension probability. A 10\% increase in relative language exposure made TEs -0.67\% more likely to be understood. Cognates (identical and non-identical) were 2.91\% more like to be understood than non-cognates. The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0. Identical cognates were 2.06\% less like to be understood than non-identical cognates. The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero). Figure 1 summarises the posterior distribution of the fixed effects of the model.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lcccrrr}
\toprule
\textbf{Predictor} & \textbf{Mean} & \textbf{SEM} & \textbf{95\% CrI} & \textbf{Rhat} & \textbf{Bulk ESS}\textsuperscript{1} & \textbf{Tail ESS}\textsuperscript{1} \\ 
\midrule
Intercept\textsuperscript{2} & $16.24\%$ & $51.39\%$ & [$14.92\%$, $17.82\%$] & $1.05$ & $102$ & $201$ \\ 
age\textsuperscript{3} & $5.57\%$ & $0.14\%$ & [$5.27\%$, $5.83\%$] & $1.01$ & $540$ & $504$ \\ 
frequency\_center & $7.57\%$ & $0.61\%$ & [$6.34\%$, $8.77\%$] & $1.00$ & $445$ & $432$ \\ 
doe\_center\textsuperscript{3} & $-0.67\%$ & $0.57\%$ & [$-1.74\%$, $0.41\%$] & $1.00$ & $847$ & $849$ \\ 
cognate1\textsuperscript{3} & $2.91\%$ & $2.01\%$ & [$-1.15\%$, $6.87\%$] & $1.01$ & $260$ & $518$ \\ 
cognate2 & $-0.46\%$ & $2.09\%$ & [$-4.88\%$, $3.30\%$] & $1.02$ & $203$ & $312$ \\ 
doe\_center:cognate1\textsuperscript{3} & $-0.72\%$ & $0.70\%$ & [$-2.05\%$, $0.62\%$] & $1.01$ & $1,119$ & $838$ \\ 
doe\_center:cognate2 & $0.60\%$ & $0.72\%$ & [$-0.82\%$, $2.01\%$] & $1.00$ & $1,279$ & $743$ \\ 
\bottomrule
\end{longtable}
\vspace{-5mm}
\begin{minipage}{\linewidth}
\textsuperscript{1}ESS: Effective sample size \\ 
\textsuperscript{2}Transformed using the inverse logit to get the average probability of correct response \\ 
\textsuperscript{3}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
\end{minipage}

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=0.8\linewidth]{manuscript_files/figure-latex/emmeans-1}

\}

\textbackslash caption\{Expected marginal mean and 95\% credible interval of the posterior probability of comprehension and production for identical cognates, non-identical cognates, and non cognates, plotted separately for three levels of degree of exposue to the language: 0\% (no exposure, monolingual), 50\% (exposed half of the time, balanced bilingual), and 100\% (exposed all of the time, monolingual). Dots represent the mean, and error bars represent the range of values within the true values is contained with 50\% (thick line) and 95\% (thin line) cprobability, given the data.\}\label{fig:emmeans}
\textbackslash end\{figure\}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-agresti2003categorical}{}%
Agresti, A. (2003). \emph{Categorical data analysis} (Vol. 482). John Wiley \& Sons.

\leavevmode\hypertarget{ref-arslan2020formr}{}%
Arslan, R. C., Walther, M. P., \& Tata, C. S. (2020). Formr: A study framework allowing for automated feedback generation and complex longitudinal experience-sampling studies using r. \emph{Behavior Research Methods}, \emph{52}(1), 376--387.

\leavevmode\hypertarget{ref-barr2013random}{}%
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing: Keep it maximal. \emph{Journal of Memory and Language}, \emph{68}(3), 255--278.

\leavevmode\hypertarget{ref-bergelson2020comprehension}{}%
Bergelson, E. (2020). The comprehension boost in early word learning: Older infants are better learners. \emph{Child Development Perspectives}, \emph{14}(3), 142--149.

\leavevmode\hypertarget{ref-bergelson2012months}{}%
Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants know the meanings of many common nouns. \emph{Proceedings of the National Academy of Sciences}, \emph{109}(9), 3253--3258.

\leavevmode\hypertarget{ref-bergelson2015early}{}%
Bergelson, E., \& Swingley, D. (2015). Early word comprehension in infants: Replication and extension. \emph{Language Learning and Development}, \emph{11}(4), 369--380.

\leavevmode\hypertarget{ref-bloom2002children}{}%
Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT press.

\leavevmode\hypertarget{ref-bosch2014first}{}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter? \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\hypertarget{ref-bosch2014first}{}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter? \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\hypertarget{ref-braginsky2018childesr}{}%
Braginsky, M., Sanchez, A., \& Yurovsky, D. (2018). Childesr: Accessing the `CHILDES'database. \emph{R Package Version 0.1. 0}.

\leavevmode\hypertarget{ref-braginsky2019consistency}{}%
Braginsky, Mika, Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and variability in children's word learning across languages. \emph{Open Mind}, \emph{3}, 52--67.

\leavevmode\hypertarget{ref-burkner2017brms}{}%
Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.

\leavevmode\hypertarget{ref-carpenter2017stan}{}%
Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A. (2017). Stan: A probabilistic programming language. \emph{Journal of Statistical Software}, \emph{76}(1), 1--32.

\leavevmode\hypertarget{ref-cattani2014much}{}%
Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis, I., \& Floccia, C. (2014). How much exposure to english is necessary for a bilingual toddler to perform like a monolingual peer in language tests? \emph{International Journal of Language \& Communication Disorders}, \emph{49}(6), 649--671.

\leavevmode\hypertarget{ref-costa2000cognate}{}%
Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate facilitation effect: Implications for models of lexical access. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{26}(5), 1283.

\leavevmode\hypertarget{ref-david2008individual}{}%
David, A., \& Wei, L. (2008). Individual differences in the lexical development of french--english bilingual children. \emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.

\leavevmode\hypertarget{ref-david2008individual}{}%
David, A., \& Wei, L. (2008). Individual differences in the lexical development of french--english bilingual children. \emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.

\leavevmode\hypertarget{ref-fenson1994variability}{}%
Fenson, L., Dale, P. S., Reznick, J. S., Bates, E., Thal, D. J., Pethick, S. J., \ldots{} Stiles, J. (1994). Variability in early communicative development. \emph{Monographs of the Society for Research in Child Development}, i--185.

\leavevmode\hypertarget{ref-fenson2007macarthur}{}%
Fenson, L., \& others. (2007). \emph{MacArthur-bates communicative development inventories}. Paul H. Brookes Publishing Company Baltimore, MD.

\leavevmode\hypertarget{ref-fernald2013ses}{}%
Fernald, A., Marchman, V. A., \& Weisleder, A. (2013). SES differences in language processing skill and vocabulary are evident at 18 months. \emph{Developmental Science}, \emph{16}(2), 234--248.

\leavevmode\hypertarget{ref-floccia2018introduction}{}%
Floccia, C., Sambrook, T. D., Delle Luche, C., Kwok, R., Goslin, J., White, L., \ldots{} others. (2018). I: INTRODUCTION. \emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29.

\leavevmode\hypertarget{ref-frank2021variability}{}%
Frank, M. C., Braginsky, M., Yurovsky, D., \& Marchman, V. A. (2021). \emph{Variability and consistency in early language learning: The wordbank project}. MIT Press.

\leavevmode\hypertarget{ref-gelman2020regression}{}%
Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.

\leavevmode\hypertarget{ref-gelman1992inference}{}%
Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation using multiple sequences. \emph{Statistical Science}, \emph{7}(4), 457--472.

\leavevmode\hypertarget{ref-goldfield1990early}{}%
Goldfield, B. A., \& Reznick, J. S. (1990). Early lexical acquisition: Rate, content, and the vocabulary spurt. \emph{Journal of Child Language}, \emph{17}(1), 171--183.

\leavevmode\hypertarget{ref-gonzalez2020bilingual}{}%
Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020). Bilingual adjusted vocabulary: A developmentally-informed bilingual vocabulary measure.

\leavevmode\hypertarget{ref-hamilton2000infant}{}%
Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary development assessed with a british communicative development inventory. \emph{Journal of Child Language}, \emph{27}(3), 689--705.

\leavevmode\hypertarget{ref-hoff2012dual}{}%
Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M. (2012). Dual language exposure and early bilingual development. \emph{Journal of Child Language}, \emph{39}(1), 1.

\leavevmode\hypertarget{ref-hoshino2008cognate}{}%
Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming: Does cross-language activation survive a change of script? \emph{Cognition}, \emph{106}(1), 501--511.

\leavevmode\hypertarget{ref-jones2019children}{}%
Jones, S., \& Brandt, S. (2019). Do children really acquire dense neighbourhoods? \emph{Journal of Child Language}, \emph{46}(6), 1260--1273.

\leavevmode\hypertarget{ref-jusczyk1995infants}{}%
Jusczyk, P. W., \& Aslin, R. N. (1995). Infants' detection of the sound patterns of words in fluent speech. \emph{Cognitive Psychology}, \emph{29}(1), 1--23.

\leavevmode\hypertarget{ref-kachergis2021toward}{}%
Kachergis, G., Marchman, V. A., \& Frank, M. C. (2021). Toward a {``standard model''} of early language learning.

\leavevmode\hypertarget{ref-R-tidybayes}{}%
Kay, M. (2020). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}

\leavevmode\hypertarget{ref-levenshtein1966binary}{}%
Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. In \emph{Soviet physics doklady} (Vol. 10, pp. 707--710). Soviet Union.

\leavevmode\hypertarget{ref-magnusson2019bayesian}{}%
Magnusson, M., Andersen, M., Jonasson, J., \& Vehtari, A. (2019). Bayesian leave-one-out cross-validation for large data. In \emph{International conference on machine learning} (pp. 4244--4253). PMLR.

\leavevmode\hypertarget{ref-mayor2011statistical}{}%
Mayor, J., \& Plunkett, K. (2011). A statistical estimate of infant and toddler vocabulary size from CDI analysis. \emph{Developmental Science}, \emph{14}(4), 769--785.

\leavevmode\hypertarget{ref-mcelreath2018statistical}{}%
McElreath, R. (2018). \emph{Statistical rethinking: A bayesian course with examples in r and stan}. Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-nelson1973structure}{}%
Nelson, K. (1973). Structure and strategy in learning to talk. \emph{Monographs of the Society for Research in Child Development}, 1--135.

\leavevmode\hypertarget{ref-oller2002language}{}%
Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in bilingual children} (Vol. 2). Multilingual Matters.

\leavevmode\hypertarget{ref-patterson2004comparing}{}%
Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers' expressive vocabulary size.

\leavevmode\hypertarget{ref-patterson2004bilingual}{}%
Patterson, J., Pearson, B., \& Goldstein, B. (2004). Bilingual language development and disorders in spanish--english speakers.

\leavevmode\hypertarget{ref-pearson1994patterns}{}%
Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in the lexical growth in two languages of bilingual infants and toddlers. \emph{Language Learning}, \emph{44}(4), 617--653.

\leavevmode\hypertarget{ref-pearson1993lexical}{}%
Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical development in bilingual infants and toddlers: Comparison to monolingual norms. \emph{Language Learning}, \emph{43}(1), 93--120.

\leavevmode\hypertarget{ref-petitto2001bilingual}{}%
Petitto, L. A., Katerelos, M., Levy, B. G., Gauna, K., Tétreault, K., \& Ferraro, V. (2001). Bilingual signed and spoken language acquisition from birth: Implications for the mechanisms underlying early bilingual language acquisition. \emph{Journal of Child Language}, \emph{28}(2), 453--496.

\leavevmode\hypertarget{ref-poulin2013lexical}{}%
Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J. (2013). Lexical access and vocabulary development in very young bilinguals. \emph{International Journal of Bilingualism}, \emph{17}(1), 57--70.

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-schad2020capitalize}{}%
Schad, D. J., Vasishth, S., Hohenstein, S., \& Kliegl, R. (2020). How to capitalize on a priori contrasts in linear (mixed) models: A tutorial. \emph{Journal of Memory and Language}, \emph{110}, 104038.

\leavevmode\hypertarget{ref-schelletter2002effect}{}%
Schelletter, C. (2002). The effect of form similarity on bilingual children's lexical development. \emph{Bilingualism: Language and Cognition}, \emph{5}(2), 93--107. \url{https://doi.org/10.1017/S1366728902000214}

\leavevmode\hypertarget{ref-smithson2014bilingualism}{}%
Smithson, L., Paradis, J., \& Nicoladis, E. (2014). Bilingualism and receptive vocabulary achievement: Could sociocultural context make a difference? \emph{Bilingualism: Language and Cognition}, \emph{17}(4), 810--821.

\leavevmode\hypertarget{ref-tardif2008baby}{}%
Tardif, T., Fletcher, P., Liang, W., Zhang, Z., Kaciroti, N., \& Marchman, V. A. (2008). Baby's first 10 words. \emph{Developmental Psychology}, \emph{44}(4), 929.

\leavevmode\hypertarget{ref-tardif2008baby}{}%
Tardif, T., Fletcher, P., Liang, W., Zhang, Z., Kaciroti, N., \& Marchman, V. A. (2008). Baby's first 10 words. \emph{Developmental Psychology}, \emph{44}(4), 929.

\leavevmode\hypertarget{ref-thierry2007brain}{}%
Thierry, G., \& Wu, Y. J. (2007). Brain potentials reveal unconscious translation during foreign-language comprehension. \emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535.

\leavevmode\hypertarget{ref-thordardottir2006bilingual}{}%
Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R. (2006). Bilingual assessment: Can overall proficiency be estimated from separate measurement of two languages? \emph{Journal of Multilingual Communication Disorders}, \emph{4}(1), 1--21.

\leavevmode\hypertarget{ref-thordardottir2006bilingual}{}%
Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R. (2006). Bilingual assessment: Can overall proficiency be estimated from separate measurement of two languages? \emph{Journal of Multilingual Communication Disorders}, \emph{4}(1), 1--21.

\leavevmode\hypertarget{ref-tincoff1999some}{}%
Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word comprehension in 6-month-olds. \emph{Psychological Science}, \emph{10}(2), 172--175.

\leavevmode\hypertarget{ref-van2014subtlex}{}%
Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014). SUBTLEX-UK: A new and improved word frequency database for british english. \emph{Quarterly Journal of Experimental Psychology}, \emph{67}(6), 1176--1190.

\leavevmode\hypertarget{ref-vehtari2017practical}{}%
Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model evaluation using leave-one-out cross-validation and WAIC. \emph{Statistics and Computing}, \emph{27}(5), 1413--1432.

\leavevmode\hypertarget{ref-von2012language}{}%
Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.

\leavevmode\hypertarget{ref-R-tidyverse}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R., \ldots{} Yutani, H. (2019). Welcome to the {tidyverse}. \emph{Journal of Open Source Software}, \emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\leavevmode\hypertarget{ref-zipf1949human}{}%
Zipf, G. K. (1949). Human behavior and the principle of least effort: An introd. To human ecology.

\end{CSLReferences}

\endgroup

\end{document}

### 2021-09-26 R/01_participants.R

@@ -13,7 +13,7 @@ get_participants <- function(
             completed,
             lp %in% lps,
             between(age, age_range[1], age_range[2]),
-            sum(doe_catalan+doe_spanish)>other_threshold
+            sum(doe_catalan + doe_spanish) > other_threshold
         ) %>% 
         filter(
             between(doe_spanish, 0, 1),

### 2021-09-30 Rmd/manuscript.tex
Fix comprehension model and update report

@@
 The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (\emph{daddy}, \emph{mommy}), social routines (\emph{bye}, \emph{uh-oh}) or animals (\emph{woof-woof}).
-These results were later extended by Frank, Braginsky, Yurovsky, and Marchman (2021) to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
+These results were later extended by Frank et al.
+(2021) to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
 Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
@@
 \item
-  \emph{Age} of participant in months (\texttt{age}), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
+  \emph{Age} of participant in months (\texttt{age\_center}), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
 \item
@@
 
-We first fit a base model that only included the main effects of \texttt{age} and \texttt{frequency\_center} (Model 1) (see Mika Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019; Kachergis, Marchman, \& Frank, 2021 for similar approaches).
+We first fit a base model that only included the main effects of \texttt{age\_center} and \texttt{frequency\_center} (Model 1) (see Mika Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019; Kachergis, Marchman, \& Frank, 2021 for similar approaches).
 We then entered our predictors of interest by adding \texttt{doe\_center} (Model 2) and then \texttt{cognate} and its interaction with \texttt{doe\_center} (Model 3).
@@
 \captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{lrrrrrrrrr}
+\begin{longtable}{lrrrrrr}
 \toprule
-& \multicolumn{3}{c}{\textbf{LOOELPD}} & \multicolumn{3}{c}{\textbf{LOOIC}} & \multicolumn{3}{c}{\textbf{LOODiff.}} \\ 
- \cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
-Model & \emph{ELPD} & \emph{SE} & \emph{SE} (sub) & \emph{LOO-IC} & \emph{SE} & \emph{SE} (sub) & \emph{diff} & \emph{SE} & \emph{SE} (sub) \\ 
+& \multicolumn{3}{c}{\textbf{ELPD}} & \multicolumn{3}{c}{\textbf{Diff.}} \\ 
+ \cmidrule(lr){2-4}\cmidrule(lr){5-7}
+Model & \emph{ELPD} & \emph{SE} & \emph{SE} (sub) & \emph{diff} & \emph{SE} & \emph{SE} (sub) \\ 
 \midrule
@@
 \midrule
-Model 1 & $-83,789.09$ & $211.11$ & $114.66$ & $167,578.18$ & $422.21$ & $229.32$ & $0.00$ & $0.00$ & $0.00$ \\ 
-Model 2 & $-233,348.92$ & $682.60$ & $3,067.85$ & $466,697.84$ & $1,365.20$ & $6,135.70$ & $149,559.83$ & $714.50$ & $3,069.99$ \\ 
-Model 3 & $-242,729.62$ & $721.85$ & $3,523.44$ & $485,459.24$ & $1,443.70$ & $7,046.88$ & $158,940.53$ & $752.08$ & $3,525.30$ \\ 
-Model 1 & $-59,417.81$ & $209.84$ & $103.31$ & $118,835.62$ & $419.69$ & $206.61$ & $0.00$ & $0.00$ & $0.00$ \\ 
-Model 2 & $-60,421.29$ & $206.95$ & $80.42$ & $120,842.58$ & $413.91$ & $160.85$ & $1,003.48$ & $294.73$ & $130.92$ \\ 
-Model 3 & $-60,481.42$ & $208.06$ & $105.51$ & $120,962.85$ & $416.12$ & $211.02$ & $1,063.61$ & $295.50$ & $147.67$ \\ 
+Model 1 & $-233,647.32$ & $688.29$ & $3,175.16$ & - & - & - \\ 
+Model 2 & $-241,540.78$ & $716.13$ & $3,477.35$ & $7,893.47$ & $993.27$ & $4,708.89$ \\ 
+Model 3 & $-248,415.02$ & $746.19$ & $3,830.96$ & $14,767.71$ & $1,015.16$ & $4,975.74$ \\ 
+\midrule
+\multicolumn{1}{l}{Production} \\ 
+\midrule
+Model 1 & $-51,210.58$ & $198.89$ & $106.81$ & - & - & - \\ 
+Model 2 & $-55,243.99$ & $204.72$ & $124.16$ & $4,033.42$ & $285.42$ & $163.78$ \\ 
+Model 3 & $-60,438.69$ & $207.27$ & $87.25$ & $9,228.11$ & $287.26$ & $137.92$ \\ 
 \bottomrule
@@
 
-The average probability of participants understanding a TE was 53.47\% (intercept) when all predictors were equal to 0: mean participant age (0.00 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels).
-An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 2.93\% increment in comprehension probability.
-A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 7.09\% increment in comprehension probability.
-A 10\% increase in relative language exposure made TEs -3.00\% more likely to be understood.
-Cognates (identical and non-identical) were 1.95\% more like to be understood than non-cognates.
+The average probability of participants understanding a TE was 62.37\% (intercept) when all predictors were equal to 0: mean participant age (24.72 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average across the three levels).
+An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 18.75\% increment in comprehension probability.
+A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a -0.08\% increment in comprehension probability.
+A 10\% increase in relative language exposure made TEs -0.10\% more likely to be understood.
+Cognates (identical and non-identical) were -0.05\% more like to be understood than non-cognates.
 The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0.
-Identical cognates were 2.06\% less like to be understood than non-identical cognates.
+Identical cognates were 0.01\% less like to be understood than non-identical cognates.
 The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero).
@@
 \midrule
-Intercept\textsuperscript{1} & $53.47\%$ & $51.30\%$ & [$51.07\%$, $55.99\%$] & $16.24\%$ & $51.39\%$ & [$14.92\%$, $17.82\%$] \\ 
-age\textsuperscript{2} & $2.93\%$ & $0.12\%$ & [$2.69\%$, $3.15\%$] & $5.57\%$ & $0.14\%$ & [$5.27\%$, $5.83\%$] \\ 
-frequency\_center & $7.09\%$ & $0.51\%$ & [$6.09\%$, $8.00\%$] & $7.57\%$ & $0.61\%$ & [$6.34\%$, $8.77\%$] \\ 
-doe\_center\textsuperscript{2} & $-3.00\%$ & $0.56\%$ & [$-4.12\%$, $-1.92\%$] & $-0.67\%$ & $0.57\%$ & [$-1.74\%$, $0.41\%$] \\ 
-cognate1\textsuperscript{2} & $1.95\%$ & $2.19\%$ & [$-2.33\%$, $6.16\%$] & $2.91\%$ & $2.01\%$ & [$-1.15\%$, $6.87\%$] \\ 
-cognate2 & $-2.06\%$ & $1.85\%$ & [$-5.71\%$, $1.49\%$] & $-0.46\%$ & $2.09\%$ & [$-4.88\%$, $3.30\%$] \\ 
-doe\_center:cognate1\textsuperscript{2} & $0.60\%$ & $0.72\%$ & [$-0.72\%$, $2.13\%$] & $-0.72\%$ & $0.70\%$ & [$-2.05\%$, $0.62\%$] \\ 
-doe\_center:cognate2 & $1.76\%$ & $0.68\%$ & [$0.47\%$, $3.10\%$] & $0.60\%$ & $0.72\%$ & [$-0.82\%$, $2.01\%$] \\ 
+Intercept\textsuperscript{1} & $62.37\%$ & $52.59\%$ & [$57.48\%$, $66.97\%$] & $15.30\%$ & $51.25\%$ & [$14.08\%$, $16.57\%$] \\ 
+age\_center\textsuperscript{2} & $18.75\%$ & $2.44\%$ & [$13.79\%$, $23.38\%$] & $5.93\%$ & $0.13\%$ & [$5.66\%$, $6.19\%$] \\ 
+frequency\_center\textsuperscript{2} & $-0.08\%$ & $2.54\%$ & [$-5.32\%$, $4.87\%$] & $3.27\%$ & $0.59\%$ & [$2.13\%$, $4.49\%$] \\ 
+doe\_center\textsuperscript{2} & $-0.10\%$ & $2.71\%$ & [$-5.41\%$, $5.31\%$] & $14.41\%$ & $0.48\%$ & [$13.49\%$, $15.34\%$] \\ 
+cognate1\textsuperscript{2} & $-0.05\%$ & $2.50\%$ & [$-4.99\%$, $4.74\%$] & $2.95\%$ & $2.11\%$ & [$-1.01\%$, $7.16\%$] \\ 
+cognate2 & $-0.01\%$ & $2.57\%$ & [$-4.99\%$, $4.95\%$] & $-0.64\%$ & $2.04\%$ & [$-4.67\%$, $3.33\%$] \\ 
+doe\_center:cognate1\textsuperscript{2} & $-0.07\%$ & $2.47\%$ & [$-5.15\%$, $4.61\%$] & $-9.05\%$ & $0.83\%$ & [$-10.61\%$, $-7.28\%$] \\ 
+doe\_center:cognate2 & $-0.05\%$ & $2.44\%$ & [$-4.85\%$, $4.71\%$] & $-6.53\%$ & $0.73\%$ & [$-7.95\%$, $-5.11\%$] \\ 
 \bottomrule
@@
 
-The average probability of participants understanding a TE was 16.24\% (intercept) when all predictors were equal to 0: mean participant age (0.00 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels).
-An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 5.57\% increment in comprehension probability.
-A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 7.57\% increment in comprehension probability.
-A 10\% increase in relative language exposure made TEs -0.67\% more likely to be understood.
-Cognates (identical and non-identical) were 2.91\% more like to be understood than non-cognates.
+The average probability of participants understanding a TE was 15.30\% (intercept) when all predictors were equal to 0: mean participant age (24.72 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels).
+An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 5.93\% increment in comprehension probability.
+A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 3.27\% increment in comprehension probability.
+A 10\% increase in relative language exposure made TEs 14.41\% more likely to be understood.
+Cognates (identical and non-identical) were 2.95\% more like to be understood than non-cognates.
 The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0.
-Identical cognates were 2.06\% less like to be understood than non-identical cognates.
+Identical cognates were 0.01\% less like to be understood than non-identical cognates.
 The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero).
@@
 \midrule
-Intercept\textsuperscript{2} & $16.24\%$ & $51.39\%$ & [$14.92\%$, $17.82\%$] & $1.05$ & $102$ & $201$ \\ 
-age\textsuperscript{3} & $5.57\%$ & $0.14\%$ & [$5.27\%$, $5.83\%$] & $1.01$ & $540$ & $504$ \\ 
-frequency\_center & $7.57\%$ & $0.61\%$ & [$6.34\%$, $8.77\%$] & $1.00$ & $445$ & $432$ \\ 
-doe\_center\textsuperscript{3} & $-0.67\%$ & $0.57\%$ & [$-1.74\%$, $0.41\%$] & $1.00$ & $847$ & $849$ \\ 
-cognate1\textsuperscript{3} & $2.91\%$ & $2.01\%$ & [$-1.15\%$, $6.87\%$] & $1.01$ & $260$ & $518$ \\ 
-cognate2 & $-0.46\%$ & $2.09\%$ & [$-4.88\%$, $3.30\%$] & $1.02$ & $203$ & $312$ \\ 
-doe\_center:cognate1\textsuperscript{3} & $-0.72\%$ & $0.70\%$ & [$-2.05\%$, $0.62\%$] & $1.01$ & $1,119$ & $838$ \\ 
-doe\_center:cognate2 & $0.60\%$ & $0.72\%$ & [$-0.82\%$, $2.01\%$] & $1.00$ & $1,279$ & $743$ \\ 
+Intercept\textsuperscript{2} & $15.30\%$ & $51.25\%$ & [$14.08\%$, $16.57\%$] & $1.03$ & $108$ & $208$ \\ 
+age\_center\textsuperscript{3} & $5.93\%$ & $0.13\%$ & [$5.66\%$, $6.19\%$] & $1.00$ & $620$ & $654$ \\ 
+frequency\_center\textsuperscript{3} & $3.27\%$ & $0.59\%$ & [$2.13\%$, $4.49\%$] & $1.01$ & $603$ & $666$ \\ 
+doe\_center\textsuperscript{3} & $14.41\%$ & $0.48\%$ & [$13.49\%$, $15.34\%$] & $1.00$ & $675$ & $931$ \\ 
+cognate1\textsuperscript{3} & $2.95\%$ & $2.11\%$ & [$-1.01\%$, $7.16\%$] & $1.03$ & $249$ & $419$ \\ 
+cognate2 & $-0.64\%$ & $2.04\%$ & [$-4.67\%$, $3.33\%$] & $1.01$ & $307$ & $469$ \\ 
+doe\_center:cognate1\textsuperscript{3} & $-9.05\%$ & $0.83\%$ & [$-10.61\%$, $-7.28\%$] & $1.00$ & $1,121$ & $890$ \\ 
+doe\_center:cognate2 & $-6.53\%$ & $0.73\%$ & [$-7.95\%$, $-5.11\%$] & $1.00$ & $1,172$ & $790$ \\ 
 \bottomrule
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-agresti2003categorical}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-agresti2003categorical}{}}%
 Agresti, A. (2003). \emph{Categorical data analysis} (Vol. 482). John Wiley \& Sons.
 
-\leavevmode\hypertarget{ref-arslan2020formr}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%
 Arslan, R.
@@
 
-\leavevmode\hypertarget{ref-barr2013random}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-barr2013random}{}}%
 Barr, D.
@@
 
-\leavevmode\hypertarget{ref-bergelson2020comprehension}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bergelson2020comprehension}{}}%
 Bergelson, E. (2020). The comprehension boost in early word learning: Older infants are better learners. \emph{Child Development Perspectives}, \emph{14}(3), 142--149.
 
-\leavevmode\hypertarget{ref-bergelson2012months}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bergelson2012months}{}}%
 Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants know the meanings of many common nouns. \emph{Proceedings of the National Academy of Sciences}, \emph{109}(9), 3253--3258.
 
-\leavevmode\hypertarget{ref-bergelson2015early}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bergelson2015early}{}}%
 Bergelson, E., \& Swingley, D. (2015). Early word comprehension in infants: Replication and extension. \emph{Language Learning and Development}, \emph{11}(4), 369--380.
 
-\leavevmode\hypertarget{ref-bloom2002children}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bloom2002children}{}}%
 Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT press.
 
-\leavevmode\hypertarget{ref-bosch2014first}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bosch2014first}{}}%
 Bosch, L., \& Ramon-Casas, M.
@@
 
-\leavevmode\hypertarget{ref-bosch2014first}{}%
-Bosch, L., \& Ramon-Casas, M.
-(2014).
-First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter?
-\emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322.
-\url{https://doi.org/10.1177/0165025414532559}
-
-\leavevmode\hypertarget{ref-braginsky2018childesr}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-braginsky2018childesr}{}}%
 Braginsky, M., Sanchez, A., \& Yurovsky, D. (2018). Childesr: Accessing the `CHILDES'database. \emph{R Package Version 0.1. 0}.
 
-\leavevmode\hypertarget{ref-braginsky2019consistency}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-braginsky2019consistency}{}}%
 Braginsky, Mika, Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and variability in children's word learning across languages. \emph{Open Mind}, \emph{3}, 52--67.
 
-\leavevmode\hypertarget{ref-burkner2017brms}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
 Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
 
-\leavevmode\hypertarget{ref-carpenter2017stan}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
 Carpenter, B., Gelman, A., Hoffman, M.
@@
 
-\leavevmode\hypertarget{ref-cattani2014much}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cattani2014much}{}}%
 Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis, I., \& Floccia, C.
@@
 
-\leavevmode\hypertarget{ref-david2008individual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-david2008individual}{}}%
 David, A., \& Wei, L.
@@
 
-\leavevmode\hypertarget{ref-david2008individual}{}%
-David, A., \& Wei, L.
-(2008).
-Individual differences in the lexical development of french--english bilingual children.
-\emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.
-
-\leavevmode\hypertarget{ref-fenson1994variability}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fenson1994variability}{}}%
 Fenson, L., Dale, P.
@@
 
-\leavevmode\hypertarget{ref-fenson2007macarthur}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fenson2007macarthur}{}}%
 Fenson, L., \& others. (2007). \emph{MacArthur-bates communicative development inventories}. Paul H. Brookes Publishing Company Baltimore, MD.
 
-\leavevmode\hypertarget{ref-fernald2013ses}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fernald2013ses}{}}%
 Fernald, A., Marchman, V. A., \& Weisleder, A. (2013). SES differences in language processing skill and vocabulary are evident at 18 months. \emph{Developmental Science}, \emph{16}(2), 234--248.
 
-\leavevmode\hypertarget{ref-floccia2018introduction}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
 Floccia, C., Sambrook, T.
@@
 
-\leavevmode\hypertarget{ref-frank2021variability}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-frank2021variability}{}}%
 Frank, M. C., Braginsky, M., Yurovsky, D., \& Marchman, V. A. (2021). \emph{Variability and consistency in early language learning: The wordbank project}. MIT Press.
 
-\leavevmode\hypertarget{ref-gelman2020regression}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
 Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.
 
-\leavevmode\hypertarget{ref-gelman1992inference}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gelman1992inference}{}}%
 Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation using multiple sequences. \emph{Statistical Science}, \emph{7}(4), 457--472.
 
-\leavevmode\hypertarget{ref-goldfield1990early}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-goldfield1990early}{}}%
 Goldfield, B. A., \& Reznick, J. S. (1990). Early lexical acquisition: Rate, content, and the vocabulary spurt. \emph{Journal of Child Language}, \emph{17}(1), 171--183.
 
-\leavevmode\hypertarget{ref-gonzalez2020bilingual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gonzalez2020bilingual}{}}%
 Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020). Bilingual adjusted vocabulary: A developmentally-informed bilingual vocabulary measure.
 
-\leavevmode\hypertarget{ref-hamilton2000infant}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-hamilton2000infant}{}}%
 Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary development assessed with a british communicative development inventory. \emph{Journal of Child Language}, \emph{27}(3), 689--705.
 
-\leavevmode\hypertarget{ref-hoff2012dual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-hoff2012dual}{}}%
 Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M. (2012). Dual language exposure and early bilingual development. \emph{Journal of Child Language}, \emph{39}(1), 1.
 
-\leavevmode\hypertarget{ref-hoshino2008cognate}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-hoshino2008cognate}{}}%
 Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming: Does cross-language activation survive a change of script? \emph{Cognition}, \emph{106}(1), 501--511.
 
-\leavevmode\hypertarget{ref-jones2019children}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-jones2019children}{}}%
 Jones, S., \& Brandt, S. (2019). Do children really acquire dense neighbourhoods? \emph{Journal of Child Language}, \emph{46}(6), 1260--1273.
 
-\leavevmode\hypertarget{ref-jusczyk1995infants}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-jusczyk1995infants}{}}%
 Jusczyk, P. W., \& Aslin, R. N. (1995). Infants' detection of the sound patterns of words in fluent speech. \emph{Cognitive Psychology}, \emph{29}(1), 1--23.
 
-\leavevmode\hypertarget{ref-kachergis2021toward}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-kachergis2021toward}{}}%
 Kachergis, G., Marchman, V. A., \& Frank, M. C. (2021). Toward a {``standard model''} of early language learning.
 
-\leavevmode\hypertarget{ref-R-tidybayes}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-R-tidybayes}{}}%
 Kay, M. (2020). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
 
-\leavevmode\hypertarget{ref-levenshtein1966binary}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
 Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. In \emph{Soviet physics doklady} (Vol. 10, pp. 707--710). Soviet Union.
 
-\leavevmode\hypertarget{ref-magnusson2019bayesian}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-magnusson2019bayesian}{}}%
 Magnusson, M., Andersen, M., Jonasson, J., \& Vehtari, A. (2019). Bayesian leave-one-out cross-validation for large data. In \emph{International conference on machine learning} (pp. 4244--4253). PMLR.
 
-\leavevmode\hypertarget{ref-mayor2011statistical}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-mayor2011statistical}{}}%
 Mayor, J., \& Plunkett, K. (2011). A statistical estimate of infant and toddler vocabulary size from CDI analysis. \emph{Developmental Science}, \emph{14}(4), 769--785.
 
-\leavevmode\hypertarget{ref-mcelreath2018statistical}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-mcelreath2018statistical}{}}%
 McElreath, R. (2018). \emph{Statistical rethinking: A bayesian course with examples in r and stan}. Chapman; Hall/CRC.
 
-\leavevmode\hypertarget{ref-nelson1973structure}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-nelson1973structure}{}}%
 Nelson, K. (1973). Structure and strategy in learning to talk. \emph{Monographs of the Society for Research in Child Development}, 1--135.
 
-\leavevmode\hypertarget{ref-oller2002language}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-oller2002language}{}}%
 Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in bilingual children} (Vol. 2). Multilingual Matters.
 
-\leavevmode\hypertarget{ref-patterson2004comparing}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-patterson2004comparing}{}}%
 Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers' expressive vocabulary size.
 
-\leavevmode\hypertarget{ref-patterson2004bilingual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-patterson2004bilingual}{}}%
 Patterson, J., Pearson, B., \& Goldstein, B. (2004). Bilingual language development and disorders in spanish--english speakers.
 
-\leavevmode\hypertarget{ref-pearson1994patterns}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pearson1994patterns}{}}%
 Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in the lexical growth in two languages of bilingual infants and toddlers. \emph{Language Learning}, \emph{44}(4), 617--653.
 
-\leavevmode\hypertarget{ref-pearson1993lexical}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pearson1993lexical}{}}%
 Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical development in bilingual infants and toddlers: Comparison to monolingual norms. \emph{Language Learning}, \emph{43}(1), 93--120.
 
-\leavevmode\hypertarget{ref-petitto2001bilingual}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-petitto2001bilingual}{}}%
 Petitto, L.
@@
 
-\leavevmode\hypertarget{ref-R-base}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
 R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
 
-\leavevmode\hypertarget{ref-schad2020capitalize}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-schad2020capitalize}{}}%
 Schad, D.
@@
 
-\leavevmode\hypertarget{ref-tardif2008baby}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-tardif2008baby}{}}%
 Tardif, T., Fletcher, P., Liang, W., Zhang, Z., Kaciroti, N., \& Marchman, V. A. (2008). Baby's first 10 words. \emph{Developmental Psychology}, \emph{44}(4), 929.
 
-\leavevmode\hypertarget{ref-tardif2008baby}{}%
-Tardif, T., Fletcher, P., Liang, W., Zhang, Z., Kaciroti, N., \& Marchman, V. A. (2008). Baby's first 10 words. \emph{Developmental Psychology}, \emph{44}(4), 929.
-
-\leavevmode\hypertarget{ref-thierry2007brain}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-thierry2007brain}{}}%
 Thierry, G., \& Wu, Y.
@@
 
-\leavevmode\hypertarget{ref-thordardottir2006bilingual}{}%
-Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R.
-(2006).
-Bilingual assessment: Can overall proficiency be estimated from separate measurement of two languages?
-\emph{Journal of Multilingual Communication Disorders}, \emph{4}(1), 1--21.
-
-\leavevmode\hypertarget{ref-tincoff1999some}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999some}{}}%
 Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word comprehension in 6-month-olds. \emph{Psychological Science}, \emph{10}(2), 172--175.
 
-\leavevmode\hypertarget{ref-van2014subtlex}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-van2014subtlex}{}}%
 Van Heuven, W.
@@
 
-\leavevmode\hypertarget{ref-vehtari2017practical}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-vehtari2017practical}{}}%
 Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model evaluation using leave-one-out cross-validation and WAIC. \emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
 
-\leavevmode\hypertarget{ref-von2012language}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-von2012language}{}}%
 Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.
 
-\leavevmode\hypertarget{ref-R-tidyverse}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-R-tidyverse}{}}%
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
@@
 
-\leavevmode\hypertarget{ref-zipf1949human}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-zipf1949human}{}}%
 Zipf, G. K. (1949). Human behavior and the principle of least effort: An introd. To human ecology.

### 2021-09-30 Rmd/report.Rmd
Merge into one manuscript
Fix comprehension model and update report

@@
 
-We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age` by `item`, and their correlation parameter.
+We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_center` by `item`, and their correlation parameter.
 We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope.
@@
 
-* `understands/produces ~ 1 + age + frequency + (1 + age | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age + frequency + doe + (1 + age + doe | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age + frequency + doe*cognate + (1 + age + doe | item) + (1 + frequency + cognate| id)`
+* `understands/produces ~ 1 + age_center + frequency + (1 + age_center | item) + (1 + frequency | id)`
+* `understands/produces ~ 1 + age_center + frequency + doe + (1 + age_center + doe | item) + (1 + frequency | id)`
+* `understands/produces ~ 1 + age_center + frequency + doe*cognate + (1 + age_center + doe | item) + (1 + frequency + cognate| id)`
 
@@
 understands ~
-1 + age + frequency + lp*cognate +
-(1 + age + lp | te) +
+1 + age_center + frequency + lp*cognate +
+(1 + age_center + lp | te) +
 (1 + frequency + cognate | id),
@@
 
-```{stan stan_model, echo=FALSE, output.var="model"}
+```
 // generated with brms 2.15.0
@@
   )
-```
-
-## Prior-predictive checks (PPC)
-
-### PPC: Comprehension
-
-```{r ppc_comp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}
-
-# emmeans
-nd <- expand.grid(
-  age = seq(min(responses$age), max(responses$age), length.out = 30),
-  doe_center = range(responses$doe_center),
-  cognate = unique(items$cognate),
-  frequency_center = 0
-)
-
-m <- add_fitted_draws(nd, fits_comp_prior$fit_2, n = 20, re_formula = NA) %>% 
-  mutate(
-    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
-  )
-
-ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
-  facet_grid(~doe_center) +
-  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
-  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
-  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
-  # stat_summary(fun = "mean", geom = "line", size = 1) +
-  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
-  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  theme_github() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-
-```
-
-### PPC: Production
-
-```{r ppc_prod, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}
-
-nd <- expand.grid(
-  age = seq(min(responses$age), max(responses$age), length.out = 30),
-  doe_center = range(responses$doe_center),
-  cognate = unique(items$cognate),
-  frequency_center = 0
-)
-
-m <- add_fitted_draws(nd, fits_prod_prior$fit_2, n = 20, re_formula = NA) %>% 
-  mutate(
-    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
-  )
-
-ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
-  facet_grid(~doe_center) +
-  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
-  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
-  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
-  # stat_summary(fun = "mean", geom = "line", size = 1) +
-  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
-  labs(x = "Age (months)", y = "P(Production)", colour = "Cognateness", fill = "Cognateness") +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  theme_github() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-
 ```
@@
     footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
+    locations = cells_body(columns = "term", rows = term %in% c("age_center", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
   ) %>% 
@@
 nd <- expand.grid(
-  age = seq(min(responses$age), max(responses$age), length.out = 30),
+  age_center = seq(min(responses$age_center), max(responses$age_center), length.out = 30),
   doe_center = range(responses$doe_center),
@@
 
-ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
+ggplot(m, aes(age_center, .value, colour = cognate, fill = cognate)) +
   facet_wrap(~doe_center) +

### 2021-09-30 _targets.R

@@ -13,7 +13,7 @@ tar_option_set(
     packages = c(
         "dplyr", "tidyr", "stringr", "ggplot2", "tibble", "forcats", "multilex", "keyring",
         "readxl", "janitor", "mice", "here", "lubridate", "purrr", "scales",
-        "brms", "tidybayes", "gt", "patchwork", "wesanderson", "papaja"
+        "brms", "tidybayes", "gt", "patchwork", "wesanderson", "papaja", "knitr"
     )
 )
 options(tidyverse.quiet = TRUE)
@@ -80,36 +80,6 @@ list(
         )
     ),
     
-    # fit models (prior, comprehension)
-    tar_target(
-        fits_comp_prior,
-        fit_models_comp(
-            data = responses,
-            iter = 500,
-            cores = 4,
-            chains = 4,
-            inits = 0,
-            save_pars = save_pars(all = TRUE),
-            backend = "cmdstanr",
-            sample_prior = "only"
-        )
-    ),
-    
-    # fit models (prior, production)
-    tar_target(
-        fits_prod_prior,
-        fit_models_prod(
-            data = responses,
-            iter = 500,
-            cores = 4,
-            chains = 4,
-            inits = 0,
-            save_pars = save_pars(all = TRUE),
-            backend = "cmdstanr",
-            sample_prior = "only"
-        )
-    ),
-    
     # fit models (comprehension)
     tar_target(
         fits_comp,
@@ -149,7 +119,7 @@ list(
     ),
 
     # render report.Rmd
-    # tar_render(report, "Rmd/report.Rmd"),
+    tar_render(report, "Rmd/report.Rmd"),
     
     # render manuscript
     tar_render(manuscript, "Rmd/manuscript.Rmd")

### 2021-09-30 Rmd/manuscript.Rmd
Fix comprehension model and update report

@@
 output            :
-  papaja::apa6_pdf:
-    latex_engine  : "pdflatex"
+  word_document:
+    reference_docx: "manuscript_template.docx"
 
@@
   cache.extra = knitr::rand_seed,
-  out.width = "80%",
+  out.width = "100%",
+  dpi = 300,
   results = "asis"
@@
 Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate.
-Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant languag (i.e., their degree of bilingualism).
+Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant language (i.e., their degree of bilingualism).
 
-```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Participant sample size by age and degree of exposure to a second language.
+```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Participant sample size by age and degree of exposure to a second language.
 Information about DoE was provided by families when filling the questionnaire.
@@
 labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
-participants %>% 
+t <- participants %>% 
   mutate(
@@
 
+gtsave(t, here("Figures/participants_lp.png"))
+
 ```
@@
 
-```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Educational attainment.
+```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Educational attainment.
 Proportion of highest parental educational attainment by DoE to the second language.
@@
 
-participants %>%
+t <- participants %>%
   mutate(
@@
   )
+
+gtsave(t, here("Figures/participants_edu.png"))
+
 ```
@@
 
-```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Lexical frequencies.
+```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
 Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
-items %>%
+
+d <- items %>%
   select(te, language, cognate, frequency) %>% 
@@
   mutate(language = ifelse(language=="Catalan", "CAT", "SPA")) %>% 
-  select(-matches("sd|upper|lower")) %>% 
+  select(-matches("sd|upper|lower"))
+
+t <- d %>% 
   gt(groupname_col = "language") %>% 
@@
 
+gtsave(t, here("Figures/item_frequency.png"))
+
 ```
@@
 
-* *Age* of participant in months (`age`), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
+* *Age* of participant in months (`age_center`), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
 * *Lexical frequency* of item (`frequency_center`), Relative lexical frequencies (frequency per million words, $fpmw$) were extracted from the CHILDES data base, using the `childesr` R package [@braginsky2018childesr], and were transformed to Zipf scores [@van2014subtlex; @zipf1949human] ($log10(fpmw)+3$).
@@
 
-We first fit a base model that only included the main effects of `age` and `frequency_center`  (Model 1) [see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches].
+We first fit a base model that only included the main effects of `age_center` and `frequency_center`  (Model 1) [see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches].
 We then entered our predictors of interest by adding `doe_center` (Model 2) and then `cognate` and its interaction with `doe_center` (Model 3).
@@
 In the next sections we report the fixed effects of the comprehension and production models separately.
-We report the mean of each coefficient's posterior distribution along with standard errors and 95% CrIs in Tables 4 and 5.
+We report the mean of each coefficient's posterior distribution along with standard errors and 95% credible intervals (CrI) in Tables 4 and 5.
 For interpretability, we report the derivative of the logistic function for each coefficient^[The derivative is calculated as $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.
@@
 
-```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Leave-one-out cross-validation.
+The expected log predictive density (*ELPD*) indicates the sum of the likelihood of each observation, given the model when fitted without such obervation.
+Values close to 0 indicate that the model fits data better.
+Standard errors (*SE*) indicate the uncertainty around the computed ELPD in each model.
+A small *SE* (in realtion to the magnitde of the *ELPD* indicates less uncertainty around the ELPD estimate."}
 
@@
   as_tibble() %>% 
-  rownames_to_column("model") %>% 
   select(-matches("p_loo|looic")) %>% 
-  relocate(model, matches("elpd_loo"), matches("diff")) %>%
   mutate_all(as.numeric) %>% 
-  mutate(type = "Comprehension", model = paste0("Model ", model))
+  mutate(
+    type = "Comprehension",
+    model = paste0("Model ", 2:0)
+  ) %>% 
+  relocate(model, type, matches("elpd_loo"), matches("diff"))
 
@@
   as_tibble() %>% 
-  rownames_to_column("model") %>% 
   select(-matches("p_loo|looic")) %>% 
-  relocate(model, matches("elpd_loo"), matches("diff")) %>%
   mutate_all(as.numeric) %>% 
-  mutate(type = "Production", model = paste0("Model ", model))
+  mutate(
+    type = "Production",
+    model = paste0("Model ", 2:0)
+  ) %>% 
+  relocate(model, type, matches("elpd_loo"), matches("diff"))
 
@@
 
-gt(l, groupname_col = "type") %>% 
-  tab_spanner(md("**ELPD</sub>**"), matches("elpd_loo")) %>% 
+t <- gt(l, groupname_col = "type") %>% 
+  tab_spanner(md("**ELPD**"), matches("elpd_loo")) %>% 
   tab_spanner(md("**Diff.**"), matches("diff")) %>%
@@
   )
+
+gtsave(t, here("Figures/loos.png"))
+
 ```
 
-# Comprehension
+## Comprehension
 
-The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe_center))`%), and regardless of its cognate status (average acros the three levels).
-An increase in one standard deviation in age (1 *SD* = `r printnum(sd(responses$age))` months) was associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability.
+The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe)*10)`%), and regardless of its cognate status (average across the three levels).
+An increase in one month of age was associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability.
 A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_comp["frequency", "Estimate"]*100/4)`% increment in comprehension probability.
-A 10% increase in relative language exposure made TEs `r printnum(coef_comp["doe_center", "Estimate"]*100/4)`% more likely to be understood.
+A 10% increase in relative language exposure made TEs `r printnum(abs(coef_comp["doe_center", "Estimate"]*100/4))`% less likely to be understood.
 Cognates (identical and non-identical) were `r printnum(coef_comp["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates.
-The 95% credible interval of the `doe_center:cognate1` interaction did not exclude 0.
+The 95% credible interval of the `doe_center:cognate1` interaction excluded 0 and covered negative values, indicating that for every 10% increment in DoE, the difference in comprehension probability between cognates and non-cognates decreased in `r printnum(abs(coef_comp["doe_center:cognate1", "Estimate"]*100/4))`%.
 Identical cognates were `r printnum(abs(coef_comp["cognate2", "Estimate"]*100/4))`% less like to be understood than non-identical cognates.
-The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero).
+The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero), and covered only negative values, indicating that for every 10% increase in DoE, the difference in comprehension probability between identical cognates and non-identical cognates decreased in `r printnum(abs(coef_comp["doe_center:cognate2", "Estimate"]*100/4))`%.
 Figure 1 summarises the posterior distribution of the fixed effects of the model.
 
-```{r fix_effets_comp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension.
+```{r fix_effets, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension and production.
 Points and error bars in indicate the posterior means, and 50% and 95% CrI.
@@
   rename_at(vars(-matches("term")), ~paste0("comp_", .)) %>% 
-  select(-one_of(c("rhat", "bulk_ess", "tail_ess")))
+  select(-any_of(c("rhat", "bulk_ess", "tail_ess")))
 
@@
 
-gt(c) %>% 
+t <- gt(c) %>% 
   fmt_percent(matches("estimate|error")) %>% 
@@
     footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
+    locations = cells_body(columns = "term", rows = term %in% c("age_center", "frequency_center", "doe_center", "cognate1", "cognate2", "doe_center:cognate1"))
   ) %>% 
@@
   )
+gtsave(t, here("Figures/fix_effects.png"))
+
 ```
@@
 
-The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_prod["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe_center))`%), and regardless of its cognate status (average acros the three levels).
-An increase in one standard deviation in age (1 *SD* = `r printnum(sd(responses$age))` months) was associated with a `r printnum(coef_prod["age", "Estimate"]*100/4)`% increment in comprehension probability.
-A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_prod["frequency", "Estimate"]*100/4)`% increment in comprehension probability.
-A 10% increase in relative language exposure made TEs `r printnum(coef_prod["doe_center", "Estimate"]*100/4)`% more likely to be understood.
-Cognates (identical and non-identical) were `r printnum(coef_prod["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates.
-The 95% credible interval of the `doe_center:cognate1` interaction did not exclude 0.
-Identical cognates were `r printnum(abs(coef_comp["cognate2", "Estimate"]*100/4))`% less like to be understood than non-identical cognates.
-The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero).
-Figure 1 summarises the posterior distribution of the fixed effects of the model.
+The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_prod["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0.
+An increase in one month of age was associated with a `r printnum(coef_prod["age_center", "Estimate"]*100/4)`% increment in production probability.
+A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_prod["frequency", "Estimate"]*100/4)`% increment in production probability.
+A 10% increase in relative language degree of exposure  made TEs `r printnum(coef_prod["doe_center", "Estimate"]*100/4)`% less likely to be produced.
+Cognates (identical and non-identical) were `r printnum(coef_prod["cognate1", "Estimate"]*100/4)`% more like to be produced than non-cognates.
+The `doe_center:cognate1` interaction (whose posterior 95% CrI did not exclude 0) indicated that the size of this effect decreased in `r printnum(abs(coef_comp["cognate1", "Estimate"]*100/4))`% for every 10% increment in language degree of exposure.
+Identical cognates were `r printnum(abs(coef_prod["cognate2", "Estimate"]*100/4))`% less likely to be produced than non-identical cognates.
+The `doe_center:cognate1` interaction (whose posterior 95% CrI did not excdude 0 either) indicated that the size of this effect decreased in `r printnum(abs(coef_comp["doe_center:cognate2", "Estimate"]*100/4))`% for every 10% increment in language degree of exposure.
 
-```{r fix_effets_prod, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for production Points and error bars in indicate the posterior means, and 50% and 95% CrI.
-The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-
-summary(fits_prod$fit_2)$fixed %>% 
-  as.data.frame() %>%
-  rownames_to_column("term") %>% 
-  mutate(
-    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
-    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
-    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
-    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
-  ) %>% 
-  gt() %>% 
-  fmt_percent(2:5) %>% 
-  fmt_number(6:6, decimals = 2) %>% 
-  fmt_number(7:8, decimals = 0) %>% 
-  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
-  cols_label(
-    term = md("**Predictor**"),
-    Estimate = md("**Mean**"),
-    Est.Error = md("**SEM**"),
-    "l-95% CI" = md("**95\\% CrI**"),
-    Rhat = md("**Rhat**"),
-    Bulk_ESS = md("**Bulk ESS**"),
-    Tail_ESS = md("**Tail ESS**")
-  ) %>% 
-  tab_footnote(
-    footnote = "Transformed using the inverse logit to get the average probability of correct response",
-    locations = cells_body(columns = "term", rows = term=="Intercept")
-  ) %>% 
-  tab_footnote(
-    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
-  ) %>% 
-  tab_footnote(
-    footnote = "ESS: Effective sample size",
-    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
-  ) %>% 
-  cols_align(
-    align = c("center"),
-    columns = 2:4
-  ) 
-```
-
-```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Expected marginal mean and 95% credible interval of the posterior probability of comprehension and production for identical cognates, non-identical cognates, and non cognates, plotted separately for three levels of degree of exposue to the language: 0% (no exposure, monolingual), 50% (exposed half of the time, balanced bilingual), and 100% (exposed all of the time, monolingual).
+```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=6, fig.width=8, fig.cap="Expected marginal mean and 95% credible interval of the posterior probability of comprehension and production for identical cognates, non-identical cognates, and non cognates, plotted separately for three levels of degree of exposue to the language: 0% (no exposure, monolingual), 50% (exposed half of the time, balanced bilingual), and 100% (exposed all of the time, monolingual).
 Dots represent the mean, and error bars represent the range of values within the true values is contained with 50% (thick line) and 95% (thin line) cprobability, given the data."}
@@
 nd <- expand.grid(
-  age = mean(responses$age),
+  age_center = floor(seq(min(responses$age_center), max(responses$age_center), 0.5)),
   doe_center = c(min(responses$doe_center), mean(responses$doe_center), max(responses$doe_center)),
@@
 
-m_comp <- add_fitted_draws(nd, fits_comp$fit_2, n = 20, re_formula = NA) %>% 
+# comprehension
+m_comp <- epred_draws(fits_comp$fit_2, newdata = nd, ndraws = 50, re_formula = NA) %>% 
   mutate(
     frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%"))
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%")),
+    cognate = factor(cognate, levels = c("Identical cognate", "Non-identical cognate", "Non-cognate"))
+  )
+m_comp_ci <- mean_qi(m_comp)
+
+# production
+m_prod <- epred_draws(fits_prod$fit_2, newdata = nd, ndraws = 50, re_formula = NA) %>% 
+  mutate(
+    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
+    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%")),
+    cognate = factor(cognate, levels = c("Identical cognate", "Non-identical cognate", "Non-cognate"))
+  )
+m_prod_ci <- mean_qi(m_prod)
+
+# join datasets
+m <- bind_rows(m_comp, m_prod, .id = "type") %>% 
+  mutate(
+    type = ifelse(type==1, "Comprehension", "Production"),
+    age_center = age_center+mean(responses$age)
   )
 
-m_prod <- add_fitted_draws(nd, fits_prod$fit_2, n = 20, re_formula = NA) %>% 
+m_ci <- bind_rows(m_comp_ci, m_prod_ci, .id = "type") %>% 
   mutate(
-    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%"))
+    type = ifelse(type==1, "Comprehension", "Production"),
+    age_center = age_center+mean(responses$age)
   )
 
-m <- bind_rows(m_comp, m_prod, .id = "type") %>% 
-  mutate(type = ifelse(type==1, "Comprehension", "Production"))
+ggplot(m, aes(age_center, .epred, colour = cognate, fill = cognate, shape = cognate)) +
+  facet_grid(type~doe_center) +
+  # geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.25, size = 0.2,
+  #           show.legend = FALSE) +
+  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
+  geom_hline(yintercept = 0.5, colour = "grey") +
+  geom_vline(xintercept = mean(responses$age), colour = "grey") +
 
-ggplot(m, aes(cognate, .value, colour = cognate, fill = cognate, shape = doe_center)) +
-  facet_grid(~type, scales = "free_y") +
-  # geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.25) +
-  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
-  # stat_summary(fun = "mean", geom = "line", size = 1) + 
-  geom_hline(yintercept = 0.5, linetype = "dotted") +
-  stat_pointinterval(.width = 0.95, position = position_dodge(width = 0.5), size = 0.75, point_size = 2) +
-  scale_y_continuous(labels = scales::percent, limits = c(0.1, 0.9), breaks = seq(0, 1, 0.10)) +
-  labs(x = "Cognateness", y = "Posterior probability",
-       colour = "Cognateness", fill = "Cognateness", shape = "DoE") +
+  geom_errorbar(data = m_ci, aes(ymin = .lower, ymax = .upper), width = 0.5, size = 0.25) +
+  stat_summary(fun = "mean", geom = "line", size = 0.75) +
+  # stat_pointinterval(.width = 0.95, position = position_dodge(width = 0.5), size = 0.75, point_size = 2) +
+  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.25)) +
+  scale_x_continuous(labels = floor, breaks = floor(seq(min(responses$age), max(responses$age), by = 2))) +
+  labs(
+    x = "Age (months)", y = "Posterior probability",
+    colour = "Cognateness", fill = "Cognateness", shape = "DoE"
+  ) +
   theme(
-    legend.position = "right",
+    legend.position = "top",
+    axis.text.x = element_text(size = 8),
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-    axis.title.x = element_blank(),
     legend.title = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
   )

### 2022-08-02 docs/index.Rmd
Add manuscript comments

---
title: "Trajectories"
output:
  github_document:
    pandoc_args: --webtex
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  echo = FALSE,
  warning = TRUE,
  include = TRUE,
  out.width = "80%",
  dpi = 250
)
options(
  knitr.kable.NA = "-",
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.continuous.fill = ggplot2::scale_color_gradient,
  ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# set params
tar_load_globals()
tar_load_all()
theme_set(theme_custom())

```

# Questionnaires

# Items {.tabset .tab-pills}

## Lexical frequency

```{r items_frequency, message=FALSE, warning=FALSE, paged.print=FALSE}

by_language <- items %>% 
  drop_na(frequency) %>% 
  ggplot(aes(frequency, lv, colour = language)) +
  geom_point(alpha = 0.5, shape = 1, stroke = 1) +
  labs(
    y = "Levenshtein similarity", 
    x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
    colour = "Cognateness", 
    fill = "Cognateness"
  ) +
  scale_y_continuous(labels = percent)

by_class <- items %>% 
  drop_na(frequency, class) %>% 
  ggplot(aes(frequency, lv, fill = class, colour = class)) +
  facet_wrap(vars(class), ncol = 2) +
  geom_point(alpha = 0.25, shape = 1, stroke = 1, show.legend = FALSE) +
  labs(y = "Levenshtein similarity", x = "Lexical frequency", colour = "cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = percent)

(by_language + by_class) +
  plot_layout(2) &
  theme(
    legend.position = "top",
    panel.grid.major.y = element_blank()
  )

```

## {-}

# Participants {.tabset .tab-pills}

```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
  group_by(time_stamp, dominant_language) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  group_by(dominant_language) %>% 
  mutate(n = cumsum(n)) %>% 
  ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
  geom_line(size = 1) +
  labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
  scale_y_continuous(breaks = seq(0, 250, 50)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
  )

```

## Age

```{r participants_age, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>% 
  mutate(age = round(age)) %>% 
  count(lp, dominant_language, age) %>% 
  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_col(position = position_fill(), size = 0.5, colour = "white") +
  guides(colour = guide_legend(ncol = 2)) +
  coord_flip() +
  labs(y = "Proportion of the sample", x = "Age (months)", colour = "LP - Dominant language") +
  scale_y_continuous(labels = scales::percent) +
  theme(
    axis.text.y = element_text(size = 9),
    legend.position = "top",
    legend.title = element_blank(),
  ) +

  participants %>% 
  mutate(age = round(age)) %>% 
  count(lp, dominant_language, age) %>% 
  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_col(size = 0.5, colour = "white") +
  labs(y = "Number of participants", x = "Age (months)") +
  guides(colour = guide_legend(ncol = 2)) +
  coord_flip() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()

  ) +

  plot_layout(nrow = 1, guides = "collect") &
  theme(
    legend.position = "top", 
    panel.grid.major = element_blank()
  )

```

## Language profile

```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")

tbl <- participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0)

gt(tbl) %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, max(tbl[,2:ncol(tbl)]))
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  ) %>% 
  as_raw_html()

```

## SES/parental education

```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tbl <- participants %>%
  mutate(edu_parent = fct_explicit_na(as.factor(edu_parent))) %>% 
  count(lp, dominant_language, edu_parent) %>% 
  right_join(
    expand_grid(
      lp = c("Monolingual", "Bilingual"),
      dominant_language = c("Catalan", "Spanish"),
      edu_parent = unique(participants$edu_parent)
    )
  ) %>% 
  replace_na(list(n = 0)) %>% 
  rename(total = n) %>% 
  pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
  mutate(total = rowSums(cbind(.[,3:8]))) %>% 
  rename(Total = total) %>% 
  select(-"NA")

gt(tbl, groupname_col = "lp") %>% 
  cols_label(
    lp = "Lang. Profile",
    dominant_language = "L1"
  ) %>% 
  summary_rows(
    columns = 3:8, 
    fns = list(Total = ~sum(.)),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_style(
    style = list(cell_borders(weight = px(2), sides = "left")),
    locations = cells_body(columns = 10, rows = 1:4)
  ) %>% 
  as_raw_html()
```

## {-}

# Data analysis {.tabset .tab-pills}

We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_center` by `item`, and their correlation parameter. We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope. Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`. The models were implemented in `brms` as:

* `understands/produces ~ 1 + age_center + frequency + (1 + age_center | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age_center + frequency + doe + (1 + age_center + doe | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age_center + frequency + doe*cognate + (1 + age_center + doe | item) + (1 + frequency + cognate| id)`

## Model equation

$$
\begin{aligned}
\color{white}\log(\frac{p}{n - q}) &= \color{white}(\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
& \color{white}(\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
& \color{white}(\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
& \color{white}(\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
& \color{white}(\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
& \color{white}(\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
& \color{white}\varepsilon_{pi}
\end{aligned}
$$

## R code (`brms`)

```
understands ~
1 + age_center + frequency + lp*cognate +
(1 + age_center + lp | te) +
(1 + frequency + cognate | id),
family =  bernoulli("logit")
```

## Stan code

Stan code generated by `brms::stancode`:

```
// generated with brms 2.15.0
functions {
/* compute correlated group-level effects
* Args: 
*   z: matrix of unscaled group-level effects
*   SD: vector of standard deviation parameters
*   L: cholesky factor correlation matrix
* Returns: 
*   matrix of scaled group-level effects
*/ 
matrix scale_r_cor(matrix z, vector SD, matrix L) {
// r is stored in another dimension order than z
return transpose(diag_pre_multiply(SD, L) * z);
}
}
data {
int<lower=1> N;  // total number of observations
int Y[N];  // response variable
int<lower=1> K;  // number of population-level effects
matrix[N, K] X;  // population-level design matrix
// data for group-level effects of ID 1
int<lower=1> N_1;  // number of grouping levels
int<lower=1> M_1;  // number of coefficients per level
int<lower=1> J_1[N];  // grouping indicator per observation
// group-level predictor values
vector[N] Z_1_1;
vector[N] Z_1_2;
vector[N] Z_1_3;
int<lower=1> NC_1;  // number of group-level correlations
// data for group-level effects of ID 2
int<lower=1> N_2;  // number of grouping levels
int<lower=1> M_2;  // number of coefficients per level
int<lower=1> J_2[N];  // grouping indicator per observation
// group-level predictor values
vector[N] Z_2_1;
vector[N] Z_2_2;
vector[N] Z_2_3;
int<lower=1> NC_2;  // number of group-level correlations
int prior_only;  // should the likelihood be ignored?
}
transformed data {
int Kc = K - 1;
matrix[N, Kc] Xc;  // centered version of X without an intercept
vector[Kc] means_X;  // column means of X before centering
for (i in 2:K) {
means_X[i - 1] = mean(X[, i]);
Xc[, i - 1] = X[, i] - means_X[i - 1];
}
}
parameters {
vector[Kc] b;  // population-level effects
real Intercept;  // temporary intercept for centered predictors
matrix[M_1, N_1] z_1;  // standardized group-level effects
cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
vector<lower=0>[M_2] sd_2;  // group-level standard deviations
matrix[M_2, N_2] z_2;  // standardized group-level effects
cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
}
transformed parameters {
vector<lower=0>[M_1] sd_1;  // group-level standard deviations
matrix[N_1, M_1] r_1;  // actual group-level effects
// using vectors speeds up indexing in loops
vector[N_1] r_1_1;
vector[N_1] r_1_2;
vector[N_1] r_1_3;
matrix[N_2, M_2] r_2;  // actual group-level effects
// using vectors speeds up indexing in loops
vector[N_2] r_2_1;
vector[N_2] r_2_2;
vector[N_2] r_2_3;
sd_1 = rep_vector(1, rows(sd_1));
// compute actual group-level effects
r_1 = scale_r_cor(z_1, sd_1, L_1);
r_1_1 = r_1[, 1];
r_1_2 = r_1[, 2];
r_1_3 = r_1[, 3];
// compute actual group-level effects
r_2 = scale_r_cor(z_2, sd_2, L_2);
r_2_1 = r_2[, 1];
r_2_2 = r_2[, 2];
r_2_3 = r_2[, 3];
}
model {
// likelihood including constants
if (!prior_only) {
// initialize linear predictor term
vector[N] mu = Intercept + rep_vector(0.0, N);
for (n in 1:N) {
// add more terms to the linear predictor
mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n];
}
target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
}
// priors including constants
target += normal_lpdf(b | 0, 3);
target += normal_lpdf(Intercept | 0, 3);
target += std_normal_lpdf(to_vector(z_1));
target += lkj_corr_cholesky_lpdf(L_1 | 5);
target += normal_lpdf(sd_2 | 0, 0.5)
- 3 * normal_lccdf(0 | 0, 0.5);
target += std_normal_lpdf(to_vector(z_2));
target += lkj_corr_cholesky_lpdf(L_2 | 5);
}
generated quantities {
// actual population-level intercept
real b_Intercept = Intercept - dot_product(means_X, b);
// compute group-level correlations
corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
vector<lower=-1,upper=1>[NC_1] cor_1;
// compute group-level correlations
corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
vector<lower=-1,upper=1>[NC_2] cor_2;
// extract upper diagonal of correlation matrix
for (k in 1:M_1) {
for (j in 1:(k - 1)) {
cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
}
}
// extract upper diagonal of correlation matrix
for (k in 1:M_2) {
for (j in 1:(k - 1)) {
cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
}
}
}
```

## {-}

# Results {.tabset .tab-pills}

## Raw data

```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  mutate(lv = cut_interval(lv, 5)) %>% 
  group_by(te, lv) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(prop, reorder(te, -prop), colour = lv, fill = lv)) +
  facet_wrap(type~lv, scales = "free_y", ncol = 5) +
  geom_vline(xintercept = 0.5, linetype = "dashed", size = 1) +
  geom_point(alpha = 0.5, size = 2, shape = 1, stroke = 1) +
  geom_smooth(aes(group = lv)) +
  labs(
    x = "Mean % of acquisition", y = "Translation equivalent ID", 
    colour = "Levenshtein similarity", fill = "Levenshtein similarity"
  ) +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

```{r raw_age, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  mutate(
    lv = cut_interval(lv, 5),
    age = as.factor(round(age))
  ) %>% 
  group_by(te, age, lv) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(age, prop, colour = lv, fill = lv)) +
  facet_grid(type~lv) +
  geom_hline(yintercept = 0.5, size = 0.75) +
  stat_pointinterval(
    .width = c(0.5), size = 1
  ) +
  labs(
    x = "Age (months)", y = "Mean % of acquisition",
    colour = "Cognate", shape = "Cognate", fill = "Cognate"
  ) +
  scale_x_discrete(breaks = seq(10, 36, 2)) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 7),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_blank()
  )
```

## Model selection

We compared the performance of these models using Bayesian leave-one-out cross-validation (LOO) using the `loo` and `loo_compare` functions of the `brms` R package (dependent of the `LOO` R package). LOO consists in computing the average likelihood of each observation after estimating the model's parameters leave that same observation out of the data set. Although the `loo` function uses a particular algorithm that speeds up the computation of this criterion (pareto-smooth importance sampling, PSIS), the size of our data set lead us to rely on the computation of the same criterion using a sampling approach via de `loo_subsample` function.

### Comprehension

```{r loo_comp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
loo_comp %>% 
  as_tibble() %>% 
  rownames_to_column("model") %>% 
  relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
  mutate_all(as.numeric) %>% 
  gt() %>% 
  tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
  tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
  tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
  tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
  fmt_number(3:13) %>% 
  cols_label(
    model = "Model",
    # elpd
    elpd_loo = md("*ELPD*"),
    se_elpd_loo = md("*SE*"),
    subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
    # p
    p_loo = md("*p*"),
    se_p_loo = md("SE"),
    subsampling_se_p_loo = md("SE<sub>sub</sub>"),
    # looic
    looic = md("*LOO-IC*"),
    se_looic = md("*SE*"),
    subsampling_se_looic = md("SE<sub>sub</sub>"),
    # diff
    elpd_diff = md("*diff*"),
    se_diff = md("*SE*"),
    subsampling_se_diff = md("*SE<sub>sub</sub>*"),
  ) %>% 
  as_raw_html()
```

### Production

```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
loo_prod %>% 
  as_tibble() %>% 
  rownames_to_column("model") %>% 
  relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
  mutate_all(as.numeric) %>% 
  gt() %>% 
  tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
  tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
  tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
  tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
  fmt_number(3:13) %>% 
  cols_label(
    model = "Model",
    # elpd
    elpd_loo = md("*ELPD*"),
    se_elpd_loo = md("*SE*"),
    subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
    # p
    p_loo = md("*p*"),
    se_p_loo = md("SE"),
    subsampling_se_p_loo = md("SE<sub>sub</sub>"),
    # looic
    looic = md("*LOO-IC*"),
    se_looic = md("*SE*"),
    subsampling_se_looic = md("SE<sub>sub</sub>"),
    # diff
    elpd_diff = md("*diff*"),
    se_diff = md("*SE*"),
    subsampling_se_diff = md("*SE<sub>sub</sub>*"),
  ) %>% 
  as_raw_html()
```

## Fixed effects

### Fixed effects: Compehension

```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
str_repl <- c(
  "doe_center:lv_center" = "DoE \u00d7 Levenshtein",
  "Intercept" = "Intercept",
  "age_center" = "Age (+1 month)",
  "frequency_center" = "Frequency (+1 SD)",
  "doe_center" = "DoE (+10%)",
  "lv_center" = "Levenshtein (+1 SD)"
)

draws_fix <- fixef(fits_comp$fit_2) %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  clean_names() %>% 
  select(-est_error) %>% 
  mutate(
    estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
    q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
    q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
    term = str_replace_all(term, str_repl)
  )

gt(draws_fix) %>% 
  fmt_percent(2) %>% 
  fmt_number(3:4, decimals = 2) %>% 
  cols_merge(vars("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    estimate = md("**Mean**"),
    q2_5 = md("**95\\% CrI**"),
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("Age (+1 month)", "Frequency (+ 1 SD)",  "DoE (+10%)", "Levenshtein (+ 1 SD)", "doe_center:lv_center" = "DoE \u00d7 Levenshtein"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:3
  ) %>% 
  as_raw_html()

```

<br>

```{r fix_effets_comp_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

var_repl <- c(
  "b_doe_center:lv_center" = "DoE \u00d7 Levenshtein", 
  "b_Intercept" = "Intercept", 
  "b_age_center" = "Age (+1 month)", 
  "b_frequency_center" = "Frequency (+1 SD)",
  "b_doe_center" = "DoE (+10%)", 
  "b_lv_center" = "Levenshtein (+1 SD)", 
  "sd_te__Intercept" = "Intercept", 
  "sd_te__age_center" = "Age",
  "sd_te__doe_center:lv_center" = "DoE \u00d7 Levenshtein",
  "sd_te__doe_center" = "DoE",
  "sd_te__lv_center" = "Levenshtein"
)

post <- gather_draws(fits_comp$fit_2, `b_.*`, `sd_te_.*`, regex = TRUE, ndraws = 1000) %>% 
  mutate(
    .value = ifelse(grepl("Intercept|sd", .variable), inv_logit_scaled(.value), .value/4),
    .variable_name = str_replace_all(.variable, var_repl) 
  )

post_intercept <- post %>% 
  filter(.variable=="b_Intercept") %>% 
  ggplot(aes(.value, fct_rev(.variable_name))) +
  geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
  stat_slab(
    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
    size = 0.5,
    color = "#FF0000",
    fill = "#FF0000",
    position = position_nudge(y = 0.2)
  ) +
  stat_pointinterval(point_size = 2) +
  labs(
    x = "Comprehension posterior probability", 
    y = "Variable", 
    fill = "Variable",
    alpha = "CrI"
  ) +
  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
  scale_x_continuous(labels = percent, limits = c(0, 1))

post_fix <- post %>% 
  filter(!grepl("Intercept|sd", .variable)) %>% 
  ggplot(aes(.value, fct_rev(.variable_name))) +
  geom_vline(xintercept = 0, size = 0.5, linetype = "dashed") +
  stat_slab(
    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
    size = 0.5,
    color = "#00A08A",
    fill = "#00A08A",
    position = position_nudge(y = 0.1)
  ) +
  annotate(
    geom = "label",
    fill = "white",
    label.size = 0,
    # alpha = 0.5,
    x = -0.05, 
    y = 0.75,
    label = "Increases\nP(comprehension)",
    size = 3.5,
    hjust = 0
  ) +
  annotate(
    geom = "label",
    fill = "white",
    label.size = 0,
    # alpha = 0.5,
    x = 0.2, 
    y = 0.75,
    label = "Decreases\nP(comprehension)",
    size = 3.5,
    hjust = 1
  ) +
  annotate(
    geom = "segment",
    x = 0,
    xend = -0.05,
    y = 0.5,
    yend =  0.5,
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  annotate(
    geom = "segment",
    x = 0.15,
    xend = 0.2,
    y = 0.5,
    yend = 0.5,
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  stat_pointinterval(point_size = 2) +
  labs(
    x = "Comprehension posterior probability",
    y = "Variable", 
    fill = "Variable",
    alpha = "CrI"
  ) +
  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
  scale_x_continuous(labels = percent, breaks = seq(-0.15, 2, 0.05))

post_sd <- post %>% 
  filter(grepl("sd", .variable)) %>% 
  mutate(group = ifelse(grepl("id", .variable), "Participant", "TE")) %>% 
  ggplot(aes(.value, fct_rev(.variable_name))) +
  stat_slab(
    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
    size = 0.5,
    color = "#F98400",
    fill = "#F98400",
    position = position_nudge(y = 0.1)
  ) +
  stat_pointinterval(point_size = 2) +
  labs(
    x = "Comprehension posterior probability",
    y = "Variable", 
    fill = "Variable",
    alpha = "CrI"
  ) +
  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
  scale_x_continuous(labels = percent, breaks = seq(0.5, 1, 0.05)) +
  theme(
    axis.text.y = element_blank()
  )

(post_intercept / (post_fix | post_sd)) +
  plot_layout(heights = c(0.1, 0.9)) &
  theme(
    legend.position = c(0.9, 0.9),
    legend.direction = "horizontal",
    legend.justification = "right",
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0),
    axis.title.y = element_blank(),
    # axis.text = element_text(size = 7),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_blank()
  )
```

### Fixed effects: Production

```{r fix_effets_prod_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

var_repl <- c(
  "b_doe_center:lv_center" = "DoE \u00d7 Levenshtein", 
  "b_Intercept" = "Intercept", 
  "b_age_center" = "Age (+1 month)", 
  "b_frequency_center" = "Frequency (+1 SD)",
  "b_doe_center" = "DoE (+10%)", 
  "b_lv_center" = "Levenshtein (+1 SD)", 
  "sd_te__Intercept" = "Intercept", 
  "sd_te__age_center" = "Age",
  "sd_te__doe_center:lv_center" = "DoE \u00d7 Levenshtein",
  "sd_te__doe_center" = "DoE",
  "sd_te__lv_center" = "Levenshtein"
)

post <- gather_draws(fits_prod$fit_2, `b_.*`, `sd_te_.*`, regex = TRUE, ndraws = 1000) %>% 
  mutate(
    .value = ifelse(grepl("Intercept|sd", .variable), inv_logit_scaled(.value), .value/4),
    .variable_name = str_replace_all(.variable, var_repl) 
  )

post_intercept <- post %>% 
  filter(.variable=="b_Intercept") %>% 
  ggplot(aes(.value, fct_rev(.variable_name))) +
  geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
  stat_slab(
    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
    size = 0.5,
    color = "#FF0000",
    fill = "#FF0000",
    position = position_nudge(y = 0.2)
  ) +
  stat_pointinterval(point_size = 2) +
  labs(
    x = "Comprehension posterior probability", 
    y = "Variable", 
    fill = "Variable",
    alpha = "CrI"
  ) +
  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
  scale_x_continuous(labels = percent, limits = c(0, 1))

post_fix <- post %>% 
  filter(!grepl("Intercept|sd", .variable)) %>% 
  ggplot(aes(.value, fct_rev(.variable_name))) +
  geom_vline(xintercept = 0, size = 0.5, linetype = "dashed") +
  stat_slab(
    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
    size = 0.5,
    color = "#00A08A",
    fill = "#00A08A",
    position = position_nudge(y = 0.1)
  ) +
  annotate(
    geom = "label",
    fill = "white",
    label.size = 0,
    # alpha = 0.5,
    x = -0.05, 
    y = 0.75,
    label = "Increases\nP(comprehension)",
    size = 3.5,
    hjust = 0
  ) +
  annotate(
    geom = "label",
    fill = "white",
    label.size = 0,
    # alpha = 0.5,
    x = 0.2, 
    y = 0.75,
    label = "Decreases\nP(comprehension)",
    size = 3.5,
    hjust = 1
  ) +
  annotate(
    geom = "segment",
    x = 0,
    xend = -0.05,
    y = 0.5,
    yend =  0.5,
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  annotate(
    geom = "segment",
    x = 0.15,
    xend = 0.2,
    y = 0.5,
    yend = 0.5,
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  stat_pointinterval(point_size = 2) +
  labs(
    x = "Comprehension posterior probability",
    y = "Variable", 
    fill = "Variable",
    alpha = "CrI"
  ) +
  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
  scale_x_continuous(labels = percent, breaks = seq(-0.15, 2, 0.05))

post_sd <- post %>% 
  filter(grepl("sd", .variable)) %>% 
  mutate(group = ifelse(grepl("id", .variable), "Participant", "TE")) %>% 
  ggplot(aes(.value, fct_rev(.variable_name))) +
  stat_slab(
    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
    size = 0.5,
    color = "#F98400",
    fill = "#F98400",
    position = position_nudge(y = 0.1)
  ) +
  stat_pointinterval(point_size = 2) +
  labs(
    x = "Comprehension posterior probability",
    y = "Variable", 
    fill = "Variable",
    alpha = "CrI"
  ) +
  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
  scale_x_continuous(labels = percent, breaks = seq(0.5, 1, 0.05)) +
  theme(
    axis.text.y = element_blank()
  )

(post_intercept / (post_fix | post_sd)) +
  plot_layout(heights = c(0.1, 0.9)) &
  theme(
    legend.position = c(0.9, 0.9),
    legend.direction = "horizontal",
    legend.justification = "right",
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0),
    axis.title.y = element_blank(),
    # axis.text = element_text(size = 7),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_blank()
  )
```

## Random effects {.tabset .tab-pils}

### Random effects: Comprehension

#### Participant

```{r random_participant_comp, message=FALSE, warning=FALSE, paged.print=FALSE}

re_id <- ranef(fits_comp$fit_2)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
  left_join(distinct(fits_comp$fit_2$data, id)) %>% 
  clean_names()

ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
  geom_ribbon(aes(group = 1), alpha = 0.5) +
  geom_line(aes(group = 1), size = 1) +
  labs(
    x = "Participant posterior probability of comprehension",
    y = "Participant", 
    colour = "Lang. Prof."
  ) +
  scale_x_continuous(limits = c(0, 1), labels = percent) +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )
```

#### Item

```{r re_item_comp, message=FALSE, warning=FALSE, paged.print=FALSE}

re_te <- ranef(fits_comp$fit_2)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  clean_names()

ggplot(re_te, aes(estimate, reorder(te, estimate), xmin = q2_5, xmax = q97_5)) +
  geom_ribbon(aes(group = 1), alpha = 0.5) +
  geom_line(aes(group = 1), size = 1) +
  labs(
    x = "Participant posterior probability of comprehension",
    y = "TE", 
    colour = "Lang. Prof."
  ) +
  scale_x_continuous(limits = c(0, 1), labels = percent) +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )
```

### Random effects: Production

#### Participant

```{r random_participant_prod, message=FALSE, warning=FALSE, paged.print=FALSE}

re_id <- ranef(fits_prod$fit_2)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
  left_join(distinct(fits_prod$fit_2$data, id)) %>% 
  clean_names()

ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
  geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
  geom_line(aes(group = 1), size = 1) +
  # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
  # geom_point(size = 0.5) +
  geom_vline(xintercept = 0.5, colour = "white") +
  labs(x = "Participant posterior probability of production",
       y = "Participant", colour = "Lang. Prof.") +
  scale_x_continuous(limits = c(0, 1), labels = percent) +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )
```

#### Item

```{r re_item_prod, message=FALSE, warning=FALSE, paged.print=FALSE}

re_te <- ranef(fits_prod$fit_2)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  clean_names()

ggplot(re_te, aes(estimate, reorder(te, estimate), xmin = q2_5, xmax = q97_5)) +
  geom_ribbon(aes(group = 1), alpha = 0.5) +
  geom_line(aes(group = 1), size = 1) +
  labs(
    x = "Participant posterior probability of comprehension",
    y = "TE", 
    colour = "Lang. Prof."
  ) +
  scale_x_continuous(limits = c(0, 1), labels = percent) +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )

```

### {-}

## Marginal means

### Marginal means: Comprehension

```{r emmeans_comp, message=FALSE, warning=FALSE, paged.print=FALSE}

nd <- expand.grid(
  age_center = seq(min(responses$age_center), max(responses$age_center), length.out = 30),
  doe_center = range(responses$doe_center),
  lv_center = c(range(responses$lv_center), 0),
  frequency_center = 0
)

m <- add_epred_draws(nd, fits_comp$fit_2, ndraws = 50, re_formula = NA) %>% 
  mutate(
    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%")),
    lv_center = factor(lv_center, levels = unique(nd$lv_center), labels = c("LV: 0%", "LV: 50%", "LV: 100%"))

  )

ggplot(m, aes(age_center + mean(responses$age, na.rm = TRUE), .epred, colour = lv_center, fill = lv_center)) +
  facet_wrap(~doe_center) +
  geom_line(aes(group = interaction(.draw, lv_center), colour = lv_center), alpha = 0.25, size = 1) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  # stat_summary(fun = "mean", geom = "line", size = 1) + 
  geom_hline(yintercept = 0.5, size = 1, linetype = "dashed") +
  labs(
    x = "Age (months)", 
    y = "Posterior probability of comprehension\n(Linear prediction)",
    colour = "Cognateness", 
    fill = "Cognateness"
  ) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

```

## Marginal means: Production

```{r emmeans_prod, message=FALSE, warning=FALSE, paged.print=FALSE}

nd <- expand.grid(
  age_center = seq(min(responses$age_center), max(responses$age_center), length.out = 30),
  doe_center = range(responses$doe_center),
  lv_center = c(range(responses$lv_center), 0),
  frequency_center = 0
)

m <- add_epred_draws(nd, fits_prod$fit_2, ndraws = 50, re_formula = NA) %>% 
  mutate(
    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%")),
    lv_center = factor(lv_center, levels = unique(nd$lv_center), labels = c("LV: 0%", "LV: 50%", "LV: 100%"))
  )

ggplot(m, aes(age_center + mean(responses$age, na.rm = TRUE), .epred, colour = lv_center, fill = lv_center)) +
  facet_wrap(~doe_center) +
  geom_line(aes(group = interaction(.draw, lv_center), colour = lv_center), alpha = 0.25, size = 1) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "Posterior probability of production\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

```

### Area under the curve (AUC)

```{r confusion_matrix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# comprehension
preds <- predict(fits_comp$fit_2, ndraws = 50) 
preds <- ifelse(preds[, 1] > 0.5, 1, 0)
confusion_mat_comp <- table(preds, pull(responses, understands))
#correct classification rate
rate_comp <- sum(diag(confusion_mat_comp))/sum(confusion_mat_comp)
confusion_mat_comp <- confusion_mat_comp %>% 
  as.data.frame() 
colnames(confusion_mat_comp) <- c("pred", "value", "freq")

# production
preds <- predict(fits_prod$fit_2, ndraws = 50) 
preds <- ifelse(preds[, 1] > 0.5, 1, 0)
confusion_mat_prod <- table(preds, pull(responses, produces))
#correct classification rate
rate_prod <- sum(diag(confusion_mat_prod))/sum(confusion_mat_prod)
confusion_mat_prod <- confusion_mat_prod %>% 
  as.data.frame() 
colnames(confusion_mat_prod) <- c("pred", "value", "freq")
```

The comprehension model classified `r percent(round(rate_comp, 2))` of the responses correctly and the production model classified `r percent(round(rate_prod, 2))` of the responses correctly.

```{r confusion, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
bind_rows(
  comp = confusion_mat_comp, 
  prod = confusion_mat_prod, 
  .id = "type"
) %>% 
  mutate(pred = pred==1) %>% 
  pivot_wider(names_from = c(type, value), values_from = freq) %>% 
  clean_names() %>% 
  gt() %>% 
  tab_spanner("Observed comprehension", matches("comp")) %>% 
  tab_spanner("Observed production", matches("prod")) %>% 
  cols_label(
    pred = "Predicted",
    comp_false = "FALSE",
    comp_true = "TRUE",
    prod_false = "FALSE",
    prod_true = "TRUE"
  ) %>% 
  fmt_number(2:5, decimals = 0)

```

### Traceplots

```{r diagnostics_traceplots_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
gather_draws(fits_comp$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE, ndraws = 1000) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Comprehension") +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )
```

```{r diagnostics_traceplots_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
gather_draws(fits_prod$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE, ndraws = 1000) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Production") +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )

```

## {-}

### 2022-08-22 docs/index.Rmd
Add index.Rmd
Clean manuscript folder

@@
 ```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
 # set params
-tar_load_globals()
 tar_load_all()
 theme_set(theme_custom())
-
 ```
@@
     geom_point(alpha = 0.25, shape = 1, stroke = 1, show.legend = FALSE) +
-  labs(y = "Levenshtein similarity", x = "Lexical frequency", colour = "cognateness", fill = "Cognateness") +
+    labs(
+        y = "Levenshtein similarity",
+        x = "Lexical frequency", 
+        colour = "cognateness", 
+        fill = "Cognateness"
+    ) +
     scale_y_continuous(labels = percent)

### 2022-08-22 manuscript/manuscript.Rmd
Add manuscript comments
Clean manuscript folder

--- 2021-09-30 Rmd/manuscript.Rmd
@@
     affiliation   : "1"
+  - name          : "Ignacio Castillejo"
+    affiliation   : "2"
   - name          : "Nuria Sebastian-Galles"
@@
     institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
+  - id            : "2"
+    institution   : "Universidad Autónoma de Madrid"
 

### 2022-08-27 docs/index.Rmd
Update index

--- 2022-08-02 docs/index.Rmd
@@
     knitr.duplicate.label = "allow",
-  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
-  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
+    ggplot2.discrete.fill = ggsci::pal_d3()(4),
+    ggplot2.discrete.colour = ggsci::pal_d3()(4),
     ggplot2.continuous.fill = ggplot2::scale_color_gradient,
@@
 
+The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
+Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*.
+We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
+For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
+
+For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
+Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
+These last families were randomly allocated into one of four different subsets of the complete list of items.
+These lists were carefully designed so that each contained a representative subsample of the items from the complete list.
+Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
+Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
+
 # Items {.tabset .tab-pills}
 
+```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
+Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
+
+items %>%
+    left_join(select(multilex_data$pool, te, language)) %>% 
+    unnest(list) %>% 
+    group_by(language, list) %>% 
+    summarise_at(
+        vars(freq, n_phon, lv), 
+        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
+    ) %>% 
+    select(-c(n_phon_n, lv_n)) %>% 
+    rename(n = freq_n) %>% 
+    gt(groupname_col = "language") %>% 
+    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
+    fmt_percent(matches("lv_")) %>% 
+    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
+    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
+    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
+    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
+    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
+    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
+    cols_label(
+        list = "",
+        n = md("*N*"),
+        freq_mean = md("Mean"),
+        freq_sd = md("*SD*"),
+        freq_min = md("Range"),
+        n_phon_mean = md("Mean"),
+        n_phon_sd = md("*SD*"),
+        n_phon_min = md("Range"),
+        lv_mean = md("Mean"),
+        lv_sd = md("*SD*"),
+        lv_min = md("Range")
+    ) %>% 
+    tab_style(
+        cell_text(style = "italic"),
+        cells_column_labels(everything())
+    ) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_spanners(everything())
+    )
+
+```
+
 ## Lexical frequency
 
-```{r items_frequency, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-by_language <- items %>% 
-  drop_na(frequency) %>% 
-  ggplot(aes(frequency, lv, colour = language)) +
-  geom_point(alpha = 0.5, shape = 1, stroke = 1) +
+```{r items-frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+
+items %>% 
+    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
+    ggplot() +
+    aes(
+        x = freq,
+        fill = language,
+        colour = language,
+        shape = language
+    ) +
+    stat_slab(
+        slab_colour = "white",
+        slab_size = 1,
+        alpha = 0.5,
+    ) +
+    stat_pointinterval(
+        aes(y = -0.1),
+        position = position_dodge(width = 0.1),
+        point_size = 3
+    ) +
+    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
+    labs(
+        x = "Frequency (Zipf)",
+        y = "Probability density",
+        fill = "Language",
+        shape = "Language",
+        colour = "Language"
+    ) +
+    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
+    theme(
+        legend.position = "top",
+        axis.title = element_blank()
+    )
+
+```
+
+## # Phonemes
+
+```{r items-nphon, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+
+items %>%
+    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
+    count(language, n_phon) %>% 
+    mutate(prop = n/sum(.$n)) %>% 
+    ggplot() +
+    aes(
+        x = n_phon,
+        y = prop,
+        fill = language,
+        colour = language,
+        shape = language
+    ) +
+    geom_col(
+        position = position_dodge()
+    ) +
+    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
+    labs(
+        x = "Frequency (Zipf)",
+        y = "Probability density",
+        fill = "Language",
+        shape = "Language",
+        colour = "Language"
+    ) +
+    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
+    theme(
+        legend.position = "top",
+        axis.title = element_blank()
+    )
+
+```
+
+## Levenshtein
+
+```{r items-lv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+
+items %>% 
+    distinct(te, .keep_all = TRUE) %>% 
+    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
+    count(lv_cut) %>% 
+    mutate(prop = n/sum(.$n)) %>% 
+    ggplot() +
+    aes(x = lv_cut, y = prop) +
+    geom_col(fill = ggsci::pal_d3()(4)[1]) +
+    labs(
+        x = "Levenshtein",
+        y = "Probability density",
+    ) +
+    theme(
+        legend.position = "top",
+        axis.title = element_blank()
+    )
+
+```
+
+## Frequency by # Phonemes
+
+```{r items-frequency-phonemes, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+items %>% 
+    left_join(select(pool, item, language)) %>% 
+    drop_na(freq) %>% 
+    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
+    facet_wrap(~language) +
+    geom_point(size = 1, alpha = 0.25) +
+    geom_smooth(method = "lm", formula = "y ~ x") +
+    labs(
+        x = "# Phonemes", 
+        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
+        colour = "Language", 
+        fill = "Language"
+    ) +
+    scale_x_continuous(breaks = seq(0, 16, 1)) +
+    theme(
+        legend.position = "none",
+        panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank()
+    )
+
+```
+
+## Frequency by Levenshtein
+
+```{r items-frequency-levenshtein, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+items %>% 
+    left_join(select(pool, item, language)) %>% 
+    drop_na(freq) %>% 
+    ggplot(aes(freq, lv, colour = language, fill = language)) +
+    facet_wrap(~language) +
+    geom_point(size = 1, alpha = 0.5) +
+    geom_smooth(method = "lm", formula = "y ~ x") +
     labs(
@@
         x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
-    colour = "Cognateness", 
-    fill = "Cognateness"
+        colour = "Language", 
+        fill = "Language"
     ) +
-  scale_y_continuous(labels = percent)
-
-by_class <- items %>% 
-  drop_na(frequency, class) %>% 
-  ggplot(aes(frequency, lv, fill = class, colour = class)) +
-  facet_wrap(vars(class), ncol = 2) +
-  geom_point(alpha = 0.25, shape = 1, stroke = 1, show.legend = FALSE) +
-  labs(y = "Levenshtein similarity", x = "Lexical frequency", colour = "cognateness", fill = "Cognateness") +
-  scale_y_continuous(labels = percent)
-
-(by_language + by_class) +
-  plot_layout(2) &
+    scale_y_continuous(labels = percent) +
     theme(
-    legend.position = "top",
+        legend.position = "none",
         panel.grid.major.y = element_blank()
@@
 
+```
+
+## # Phonemes by Levenshtein
+
+```{r items-phonemes-lv, message=FALSE, warning=FALSE, paged.print=FALSE}
+items %>% 
+    left_join(select(pool, item, language)) %>% 
+    drop_na(freq) %>% 
+    ggplot(aes(n_phon, lv, colour = language, fill = language)) +
+    facet_wrap(~language) +
+    geom_point(size = 1, alpha = 0.25) +
+    geom_smooth(method = "lm", formula = "y ~ x") +
+    labs(
+        x = "# Phonemes", 
+        y = "Levenshtein similarity", 
+        colour = "Language", 
+        fill = "Language"
+    ) +
+    scale_y_continuous(labels = percent) +
+    scale_x_continuous(breaks = seq(0, 16, 1)) +
+    theme(
+        legend.position = "none",
+        panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank()
+    ) 
 ```
@@
 participants %>% 
+    mutate(dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish")) %>% 
     group_by(time_stamp, dominant_language) %>% 
@@
     ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
+    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
+    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
+             angle = 90, vjust = 1.5, hjust = 1) +
     geom_line(size = 1) +
     labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
-  scale_y_continuous(breaks = seq(0, 250, 50)) +
+    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
     guides(colour = guide_legend(ncol = 2)) +
@@
 
-```{r participants_age, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+```{r participants-age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 participants %>% 
-  mutate(age = round(age)) %>% 
-  count(lp, dominant_language, age) %>% 
-  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
-  geom_col(position = position_fill(), size = 0.5, colour = "white") +
-  guides(colour = guide_legend(ncol = 2)) +
-  coord_flip() +
-  labs(y = "Proportion of the sample", x = "Age (months)", colour = "LP - Dominant language") +
-  scale_y_continuous(labels = scales::percent) +
+    mutate(age = floor(age)) %>% 
+    count(age) %>% 
+    ggplot() +
+    aes(x = age, y = n) +
+    geom_col() +
+    geom_text(aes(label = n), size = 3.5, vjust = -1) +
+    labs(
+        x = "Age (months)",
+        y = "# participants"
+    ) +
+    scale_x_continuous(breaks = seq(0, 40, 2)) +
+    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
     theme(
-    axis.text.y = element_text(size = 9),
-    legend.position = "top",
-    legend.title = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()
+
+    )
+```
+
+## Language profile
+
+```{r participants-lp, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+participants %>% 
+    mutate(
+        dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish"),
+        age = round(age),
+        id = paste(id, "(", time, ")"),
+        rank = rank(doe_catalan)
+    ) %>% 
+    pivot_longer(
+        starts_with("doe_"), 
+        names_to = "language",
+        values_to = "doe",
+        names_prefix = "doe_",
+        names_transform = list(language = str_to_sentence)
+    ) %>% 
+    ggplot() +
+    aes(x = doe, y = reorder(id, -rank), fill = language) +
+    geom_col(width = 0.75, position = position_fill(), colour = "white", size = 0.1) +
+    labs(
+        x = "Degree of exposure to Catalan",
+        y = "Participant",
+        colour = "Language",
+        fill = "Language"
     ) +
-
+    scale_x_continuous(labels = percent) +
+    theme(
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.y = element_blank()
+    )
+
+```
+
+## SES/parental education
+
+```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 participants %>%
-  mutate(age = round(age)) %>% 
-  count(lp, dominant_language, age) %>% 
-  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
-  geom_col(size = 0.5, colour = "white") +
-  labs(y = "Number of participants", x = "Age (months)") +
-  guides(colour = guide_legend(ncol = 2)) +
-  coord_flip() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank(),
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-
-  ) +
-
-  plot_layout(nrow = 1, guides = "collect") &
-  theme(
-    legend.position = "top", 
-    panel.grid.major = element_blank()
-  )
-
-```
-
-## Language profile
-
-```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE}
-labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
-
-tbl <- participants %>% 
+    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
     mutate(
-    age = round(age),
-    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
-    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
+        edu_parent = fct_explicit_na(as.factor(edu_parent)),
+        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
     ) %>% 
-  count(age, doe_2) %>% 
-  arrange(doe_2) %>% 
-  mutate(doe_2 = fct_inorder(doe_2)) %>% 
-  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0)
-
-gt(tbl) %>% 
-  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
-  data_color(
-    columns = 2:6,
-    colors = scales::col_numeric(
-      palette = c("white", "orange"),
-      domain = c(0, max(tbl[,2:ncol(tbl)]))
-    )
-  ) %>% 
-  cols_label(
-    age = md("**Age<br>(months)**")
-  ) %>% 
-  as_raw_html()
-
-```
-
-## SES/parental education
-
-```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-tbl <- participants %>%
-  mutate(edu_parent = fct_explicit_na(as.factor(edu_parent))) %>% 
-  count(lp, dominant_language, edu_parent) %>% 
+    count(doe_catalan, edu_parent) %>% 
     right_join(
         expand_grid(
-      lp = c("Monolingual", "Bilingual"),
-      dominant_language = c("Catalan", "Spanish"),
-      edu_parent = unique(participants$edu_parent)
+            doe_catalan = levels(.$doe_catalan),
+            edu_parent = unique(multilex_data$logs$edu_parent)
         )
     ) %>% 
-  replace_na(list(n = 0)) %>% 
     rename(total = n) %>% 
@@
     rename(Total = total) %>% 
-  select(-"NA")
-
-gt(tbl, groupname_col = "lp") %>% 
+    select(-"NA") %>% 
+    gt() %>% 
     cols_label(
-    lp = "Lang. Profile",
-    dominant_language = "L1"
+        doe_catalan = "DoE Catalan"
     ) %>% 
     summary_rows(
-    columns = 3:8, 
-    fns = list(Total = ~sum(.)),
+        columns = 2:7, 
+        fns = list(Total = ~sum(., na.rm = TRUE)),
         formatter = fmt_number,
@@
     ) %>% 
+    fmt_missing(columns = 3:8, missing_text = "--") %>% 
+    tab_spanner("Educational attainment", 3:8) %>% 
     tab_style(
-    style = list(cell_borders(weight = px(2), sides = "left")),
-    locations = cells_body(columns = 10, rows = 1:4)
+        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
+        locations = cells_body(columns = 9, rows = 1:10)
     ) %>% 
-  as_raw_html()
+    tab_style(
+        cell_text(style = "italic"),
+        cells_column_labels(columns = 2:7)
+    ) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_labels(columns = c(1, 8))
+    ) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_spanners(spanners = "Educational attainment")
+    )
 ```
@@
 
-We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_center` by `item`, and their correlation parameter.
+We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_std` by `item`, and their correlation parameter.
 We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope.
@@
 
-* `understands/produces ~ 1 + age_center + frequency + (1 + age_center | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age_center + frequency + doe + (1 + age_center + doe | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age_center + frequency + doe*cognate + (1 + age_center + doe | item) + (1 + frequency + cognate| id)`
+* `understands/produces ~ 1 + age_std + frequency + (1 + age_std | item) + (1 + frequency | id)`
+* `understands/produces ~ 1 + age_std + frequency + doe + (1 + age_std + doe | item) + (1 + frequency | id)`
+* `understands/produces ~ 1 + age_std + frequency + doe*cognate + (1 + age_std + doe | item) + (1 + frequency + cognate| id)`
 
@@
 
-## R code (`brms`)
-
-```
-understands ~
-1 + age_center + frequency + lp*cognate +
-(1 + age_center + lp | te) +
-(1 + frequency + cognate | id),
-family =  bernoulli("logit")
-```
-
-## Stan code
+## R code (`brms`) {.tabset}
+
+### Model 0
+
+```{r}
+print(model_fit_0$formula)
+```
+
+### Model 1
+
+```{r}
+print(model_fit_1$formula)
+```
+
+### Model 2
+
+```{r}
+print(model_fit_2$formula)
+```
+
+### Model 3
+
+```{r}
+print(model_fit_3$formula)
+```
+
+### Model 4
+
+```{r}
+print(model_fit_4$formula)
+```
+
+### Model 5
+
+```{r}
+print(model_fit_5$formula)
+```
+
+### Model 6
+
+```{r}
+print(model_fit_6$formula)
+```
+
+### Model 7
+
+```{r}
+print(model_fit_7$formula)
+```
+
+### {-}
+
+## Stan code {.tabset}
 
@@
 
-```
-// generated with brms 2.15.0
+### Model 0
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
+functions {
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+}
+parameters {
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  vector[N_1] z_1[M_1];  // standardized group-level effects
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  vector[N_2] z_2[M_2];  // standardized group-level effects
+}
+transformed parameters {
+  real<lower=0> disc = 1;  // discrimination parameters
+  vector[N_1] r_1_1;  // actual group-level effects
+  vector[N_2] r_2_1;  // actual group-level effects
+  r_1_1 = (sd_1[1] * (z_1[1]));
+  r_2_1 = (sd_2[1] * (z_2[1]));
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = rep_vector(0.0, N);
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_2_1[J_2[n]] * Z_2_1[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
+  }
+  // priors including constants
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 1 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(z_1[1]);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 1 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(z_2[1]);
+}
+generated quantities {
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept;
+  // additionally sample draws from priors
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_sd_2 = normal_rng(1,0.1);
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+```
+
+### Model 1
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
 functions {
@@
   }
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
 }
@@
   int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
+  int<lower=1> K;  // number of population-level effects
+  matrix[N, K] X;  // population-level design matrix
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  vector[N] Z_1_2;
+  int<lower=1> NC_1;  // number of group-level correlations
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  vector[N] Z_2_2;
+  int<lower=1> NC_2;  // number of group-level correlations
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+  int Kc = K;
+  matrix[N, Kc] Xc;  // centered version of X
+  vector[Kc] means_X;  // column means of X before centering
+  for (i in 1:K) {
+    means_X[i] = mean(X[, i]);
+    Xc[, i] = X[, i] - means_X[i];
+  }
+}
+parameters {
+  vector[Kc] b;  // population-level effects
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  matrix[M_1, N_1] z_1;  // standardized group-level effects
+  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  matrix[M_2, N_2] z_2;  // standardized group-level effects
+  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+  real<lower=0> disc = 1;  // discrimination parameters
+  matrix[N_1, M_1] r_1;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_1] r_1_1;
+  vector[N_1] r_1_2;
+  matrix[N_2, M_2] r_2;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_2] r_2_1;
+  vector[N_2] r_2_2;
+  // compute actual group-level effects
+  r_1 = scale_r_cor(z_1, sd_1, L_1);
+  r_1_1 = r_1[, 1];
+  r_1_2 = r_1[, 2];
+  // compute actual group-level effects
+  r_2 = scale_r_cor(z_2, sd_2, L_2);
+  r_2_1 = r_2[, 1];
+  r_2_2 = r_2[, 2];
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = Xc * b;
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
+  }
+  // priors including constants
+  target += normal_lpdf(b[1] | 1, 0.1);
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 2 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_1));
+  target += lkj_corr_cholesky_lpdf(L_1 | 2);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 2 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_2));
+  target += lkj_corr_cholesky_lpdf(L_2 | 2);
+}
+generated quantities {
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
+  // compute group-level correlations
+  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+  vector<lower=-1,upper=1>[NC_1] cor_1;
+  // compute group-level correlations
+  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
+  vector<lower=-1,upper=1>[NC_2] cor_2;
+  // additionally sample draws from priors
+  real prior_b_1 = normal_rng(1,0.1);
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
+  real prior_sd_2 = normal_rng(1,0.1);
+  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_1) {
+    for (j in 1:(k - 1)) {
+      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+    }
+  }
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_2) {
+    for (j in 1:(k - 1)) {
+      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
+    }
+  }
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+```
+
+### Model 2
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
+functions {
+ /* compute correlated group-level effects
+  * Args: 
+  *   z: matrix of unscaled group-level effects
+  *   SD: vector of standard deviation parameters
+  *   L: cholesky factor correlation matrix
+  * Returns: 
+  *   matrix of scaled group-level effects
+  */ 
+  matrix scale_r_cor(matrix z, vector SD, matrix L) {
+    // r is stored in another dimension order than z
+    return transpose(diag_pre_multiply(SD, L) * z);
+  }
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
   int<lower=1> K;  // number of population-level effects
@@
   vector[N] Z_2_2;
-vector[N] Z_2_3;
   int<lower=1> NC_2;  // number of group-level correlations
@@
 transformed data {
-int Kc = K - 1;
-matrix[N, Kc] Xc;  // centered version of X without an intercept
+  int Kc = K;
+  matrix[N, Kc] Xc;  // centered version of X
   vector[Kc] means_X;  // column means of X before centering
-for (i in 2:K) {
-means_X[i - 1] = mean(X[, i]);
-Xc[, i - 1] = X[, i] - means_X[i - 1];
+  for (i in 1:K) {
+    means_X[i] = mean(X[, i]);
+    Xc[, i] = X[, i] - means_X[i];
   }
@@
   vector[Kc] b;  // population-level effects
-real Intercept;  // temporary intercept for centered predictors
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
   matrix[M_1, N_1] z_1;  // standardized group-level effects
@@
 transformed parameters {
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  real<lower=0> disc = 1;  // discrimination parameters
   matrix[N_1, M_1] r_1;  // actual group-level effects
@@
   vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-sd_1 = rep_vector(1, rows(sd_1));
   // compute actual group-level effects
@@
   r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
 }
@@
     // initialize linear predictor term
-vector[N] mu = Intercept + rep_vector(0.0, N);
+    vector[N] mu = Xc * b;
     for (n in 1:N) {
       // add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n];
-}
-target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
   }
   // priors including constants
-target += normal_lpdf(b | 0, 3);
-target += normal_lpdf(Intercept | 0, 3);
+  target += normal_lpdf(b[1] | 1, 0.1);
+  target += normal_lpdf(b[2] | 0, 0.1);
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 3 * normal_lccdf(0 | 1, 0.1);
   target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 5);
-target += normal_lpdf(sd_2 | 0, 0.5)
-- 3 * normal_lccdf(0 | 0, 0.5);
+  target += lkj_corr_cholesky_lpdf(L_1 | 2);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 2 * normal_lccdf(0 | 1, 0.1);
   target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 5);
+  target += lkj_corr_cholesky_lpdf(L_2 | 2);
 }
 generated quantities {
-// actual population-level intercept
-real b_Intercept = Intercept - dot_product(means_X, b);
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
   // compute group-level correlations
@@
   vector<lower=-1,upper=1>[NC_2] cor_2;
+  // additionally sample draws from priors
+  real prior_b_1 = normal_rng(1,0.1);
+  real prior_b_2 = normal_rng(0,0.1);
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
+  real prior_sd_2 = normal_rng(1,0.1);
+  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
   // extract upper diagonal of correlation matrix
@@
   }
-}
-```
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+```
+
+### Model 3
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
+functions {
+ /* compute correlated group-level effects
+  * Args: 
+  *   z: matrix of unscaled group-level effects
+  *   SD: vector of standard deviation parameters
+  *   L: cholesky factor correlation matrix
+  * Returns: 
+  *   matrix of scaled group-level effects
+  */ 
+  matrix scale_r_cor(matrix z, vector SD, matrix L) {
+    // r is stored in another dimension order than z
+    return transpose(diag_pre_multiply(SD, L) * z);
+  }
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
+  int<lower=1> K;  // number of population-level effects
+  matrix[N, K] X;  // population-level design matrix
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  vector[N] Z_1_2;
+  vector[N] Z_1_3;
+  vector[N] Z_1_4;
+  int<lower=1> NC_1;  // number of group-level correlations
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  vector[N] Z_2_2;
+  vector[N] Z_2_3;
+  vector[N] Z_2_4;
+  int<lower=1> NC_2;  // number of group-level correlations
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+  int Kc = K;
+  matrix[N, Kc] Xc;  // centered version of X
+  vector[Kc] means_X;  // column means of X before centering
+  for (i in 1:K) {
+    means_X[i] = mean(X[, i]);
+    Xc[, i] = X[, i] - means_X[i];
+  }
+}
+parameters {
+  vector[Kc] b;  // population-level effects
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  matrix[M_1, N_1] z_1;  // standardized group-level effects
+  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  matrix[M_2, N_2] z_2;  // standardized group-level effects
+  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+  real<lower=0> disc = 1;  // discrimination parameters
+  matrix[N_1, M_1] r_1;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_1] r_1_1;
+  vector[N_1] r_1_2;
+  vector[N_1] r_1_3;
+  vector[N_1] r_1_4;
+  matrix[N_2, M_2] r_2;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_2] r_2_1;
+  vector[N_2] r_2_2;
+  vector[N_2] r_2_3;
+  vector[N_2] r_2_4;
+  // compute actual group-level effects
+  r_1 = scale_r_cor(z_1, sd_1, L_1);
+  r_1_1 = r_1[, 1];
+  r_1_2 = r_1[, 2];
+  r_1_3 = r_1[, 3];
+  r_1_4 = r_1[, 4];
+  // compute actual group-level effects
+  r_2 = scale_r_cor(z_2, sd_2, L_2);
+  r_2_1 = r_2[, 1];
+  r_2_2 = r_2[, 2];
+  r_2_3 = r_2[, 3];
+  r_2_4 = r_2[, 4];
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = Xc * b;
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
+  }
+  // priors including constants
+  target += normal_lpdf(b[1] | 1, 0.1);
+  target += normal_lpdf(b[2] | 0, 0.1);
+  target += normal_lpdf(b[3] | 0, 0.1);
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 4 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_1));
+  target += lkj_corr_cholesky_lpdf(L_1 | 2);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 4 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_2));
+  target += lkj_corr_cholesky_lpdf(L_2 | 2);
+}
+generated quantities {
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
+  // compute group-level correlations
+  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+  vector<lower=-1,upper=1>[NC_1] cor_1;
+  // compute group-level correlations
+  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
+  vector<lower=-1,upper=1>[NC_2] cor_2;
+  // additionally sample draws from priors
+  real prior_b_1 = normal_rng(1,0.1);
+  real prior_b_2 = normal_rng(0,0.1);
+  real prior_b_3 = normal_rng(0,0.1);
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
+  real prior_sd_2 = normal_rng(1,0.1);
+  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_1) {
+    for (j in 1:(k - 1)) {
+      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+    }
+  }
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_2) {
+    for (j in 1:(k - 1)) {
+      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
+    }
+  }
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+
+```
+
+### Model 4
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
+functions {
+ /* compute correlated group-level effects
+  * Args: 
+  *   z: matrix of unscaled group-level effects
+  *   SD: vector of standard deviation parameters
+  *   L: cholesky factor correlation matrix
+  * Returns: 
+  *   matrix of scaled group-level effects
+  */ 
+  matrix scale_r_cor(matrix z, vector SD, matrix L) {
+    // r is stored in another dimension order than z
+    return transpose(diag_pre_multiply(SD, L) * z);
+  }
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
+  int<lower=1> K;  // number of population-level effects
+  matrix[N, K] X;  // population-level design matrix
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  vector[N] Z_1_2;
+  vector[N] Z_1_3;
+  vector[N] Z_1_4;
+  vector[N] Z_1_5;
+  int<lower=1> NC_1;  // number of group-level correlations
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  vector[N] Z_2_2;
+  vector[N] Z_2_3;
+  vector[N] Z_2_4;
+  vector[N] Z_2_5;
+  int<lower=1> NC_2;  // number of group-level correlations
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+  int Kc = K;
+  matrix[N, Kc] Xc;  // centered version of X
+  vector[Kc] means_X;  // column means of X before centering
+  for (i in 1:K) {
+    means_X[i] = mean(X[, i]);
+    Xc[, i] = X[, i] - means_X[i];
+  }
+}
+parameters {
+  vector[Kc] b;  // population-level effects
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  matrix[M_1, N_1] z_1;  // standardized group-level effects
+  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  matrix[M_2, N_2] z_2;  // standardized group-level effects
+  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+  real<lower=0> disc = 1;  // discrimination parameters
+  matrix[N_1, M_1] r_1;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_1] r_1_1;
+  vector[N_1] r_1_2;
+  vector[N_1] r_1_3;
+  vector[N_1] r_1_4;
+  vector[N_1] r_1_5;
+  matrix[N_2, M_2] r_2;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_2] r_2_1;
+  vector[N_2] r_2_2;
+  vector[N_2] r_2_3;
+  vector[N_2] r_2_4;
+  vector[N_2] r_2_5;
+  // compute actual group-level effects
+  r_1 = scale_r_cor(z_1, sd_1, L_1);
+  r_1_1 = r_1[, 1];
+  r_1_2 = r_1[, 2];
+  r_1_3 = r_1[, 3];
+  r_1_4 = r_1[, 4];
+  r_1_5 = r_1[, 5];
+  // compute actual group-level effects
+  r_2 = scale_r_cor(z_2, sd_2, L_2);
+  r_2_1 = r_2[, 1];
+  r_2_2 = r_2[, 2];
+  r_2_3 = r_2[, 3];
+  r_2_4 = r_2[, 4];
+  r_2_5 = r_2[, 5];
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = Xc * b;
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
+  }
+  // priors including constants
+  target += normal_lpdf(b[1] | 1, 0.1);
+  target += normal_lpdf(b[2] | 0, 0.1);
+  target += normal_lpdf(b[3] | 0, 0.1);
+  target += normal_lpdf(b[4] | 0, 0.1);
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 5 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_1));
+  target += lkj_corr_cholesky_lpdf(L_1 | 2);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 5 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_2));
+  target += lkj_corr_cholesky_lpdf(L_2 | 2);
+}
+generated quantities {
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
+  // compute group-level correlations
+  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+  vector<lower=-1,upper=1>[NC_1] cor_1;
+  // compute group-level correlations
+  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
+  vector<lower=-1,upper=1>[NC_2] cor_2;
+  // additionally sample draws from priors
+  real prior_b_1 = normal_rng(1,0.1);
+  real prior_b_2 = normal_rng(0,0.1);
+  real prior_b_3 = normal_rng(0,0.1);
+  real prior_b_4 = normal_rng(0,0.1);
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
+  real prior_sd_2 = normal_rng(1,0.1);
+  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_1) {
+    for (j in 1:(k - 1)) {
+      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+    }
+  }
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_2) {
+    for (j in 1:(k - 1)) {
+      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
+    }
+  }
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+
+```
+
+### Model 5
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
+functions {
+ /* compute correlated group-level effects
+  * Args: 
+  *   z: matrix of unscaled group-level effects
+  *   SD: vector of standard deviation parameters
+  *   L: cholesky factor correlation matrix
+  * Returns: 
+  *   matrix of scaled group-level effects
+  */ 
+  matrix scale_r_cor(matrix z, vector SD, matrix L) {
+    // r is stored in another dimension order than z
+    return transpose(diag_pre_multiply(SD, L) * z);
+  }
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
+  int<lower=1> K;  // number of population-level effects
+  matrix[N, K] X;  // population-level design matrix
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  vector[N] Z_1_2;
+  vector[N] Z_1_3;
+  vector[N] Z_1_4;
+  vector[N] Z_1_5;
+  vector[N] Z_1_6;
+  int<lower=1> NC_1;  // number of group-level correlations
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  vector[N] Z_2_2;
+  vector[N] Z_2_3;
+  vector[N] Z_2_4;
+  vector[N] Z_2_5;
+  int<lower=1> NC_2;  // number of group-level correlations
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+  int Kc = K;
+  matrix[N, Kc] Xc;  // centered version of X
+  vector[Kc] means_X;  // column means of X before centering
+  for (i in 1:K) {
+    means_X[i] = mean(X[, i]);
+    Xc[, i] = X[, i] - means_X[i];
+  }
+}
+parameters {
+  vector[Kc] b;  // population-level effects
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  matrix[M_1, N_1] z_1;  // standardized group-level effects
+  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  matrix[M_2, N_2] z_2;  // standardized group-level effects
+  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+  real<lower=0> disc = 1;  // discrimination parameters
+  matrix[N_1, M_1] r_1;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_1] r_1_1;
+  vector[N_1] r_1_2;
+  vector[N_1] r_1_3;
+  vector[N_1] r_1_4;
+  vector[N_1] r_1_5;
+  vector[N_1] r_1_6;
+  matrix[N_2, M_2] r_2;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_2] r_2_1;
+  vector[N_2] r_2_2;
+  vector[N_2] r_2_3;
+  vector[N_2] r_2_4;
+  vector[N_2] r_2_5;
+  // compute actual group-level effects
+  r_1 = scale_r_cor(z_1, sd_1, L_1);
+  r_1_1 = r_1[, 1];
+  r_1_2 = r_1[, 2];
+  r_1_3 = r_1[, 3];
+  r_1_4 = r_1[, 4];
+  r_1_5 = r_1[, 5];
+  r_1_6 = r_1[, 6];
+  // compute actual group-level effects
+  r_2 = scale_r_cor(z_2, sd_2, L_2);
+  r_2_1 = r_2[, 1];
+  r_2_2 = r_2[, 2];
+  r_2_3 = r_2[, 3];
+  r_2_4 = r_2[, 4];
+  r_2_5 = r_2[, 5];
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = Xc * b;
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
+  }
+  // priors including constants
+  target += normal_lpdf(b[1] | 1, 0.1);
+  target += normal_lpdf(b[2] | 0, 0.1);
+  target += normal_lpdf(b[3] | 0, 0.1);
+  target += normal_lpdf(b[4] | 0, 0.1);
+  target += normal_lpdf(b[5] | 0, 0.1);
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 6 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_1));
+  target += lkj_corr_cholesky_lpdf(L_1 | 2);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 5 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_2));
+  target += lkj_corr_cholesky_lpdf(L_2 | 2);
+}
+generated quantities {
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
+  // compute group-level correlations
+  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+  vector<lower=-1,upper=1>[NC_1] cor_1;
+  // compute group-level correlations
+  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
+  vector<lower=-1,upper=1>[NC_2] cor_2;
+  // additionally sample draws from priors
+  real prior_b_1 = normal_rng(1,0.1);
+  real prior_b_2 = normal_rng(0,0.1);
+  real prior_b_3 = normal_rng(0,0.1);
+  real prior_b_4 = normal_rng(0,0.1);
+  real prior_b_5 = normal_rng(0,0.1);
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
+  real prior_sd_2 = normal_rng(1,0.1);
+  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_1) {
+    for (j in 1:(k - 1)) {
+      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+    }
+  }
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_2) {
+    for (j in 1:(k - 1)) {
+      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
+    }
+  }
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+```
+
+### Model 6
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
+functions {
+ /* compute correlated group-level effects
+  * Args: 
+  *   z: matrix of unscaled group-level effects
+  *   SD: vector of standard deviation parameters
+  *   L: cholesky factor correlation matrix
+  * Returns: 
+  *   matrix of scaled group-level effects
+  */ 
+  matrix scale_r_cor(matrix z, vector SD, matrix L) {
+    // r is stored in another dimension order than z
+    return transpose(diag_pre_multiply(SD, L) * z);
+  }
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
+  int<lower=1> K;  // number of population-level effects
+  matrix[N, K] X;  // population-level design matrix
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  vector[N] Z_1_2;
+  vector[N] Z_1_3;
+  vector[N] Z_1_4;
+  vector[N] Z_1_5;
+  vector[N] Z_1_6;
+  vector[N] Z_1_7;
+  int<lower=1> NC_1;  // number of group-level correlations
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  vector[N] Z_2_2;
+  vector[N] Z_2_3;
+  vector[N] Z_2_4;
+  vector[N] Z_2_5;
+  int<lower=1> NC_2;  // number of group-level correlations
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+  int Kc = K;
+  matrix[N, Kc] Xc;  // centered version of X
+  vector[Kc] means_X;  // column means of X before centering
+  for (i in 1:K) {
+    means_X[i] = mean(X[, i]);
+    Xc[, i] = X[, i] - means_X[i];
+  }
+}
+parameters {
+  vector[Kc] b;  // population-level effects
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  matrix[M_1, N_1] z_1;  // standardized group-level effects
+  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  matrix[M_2, N_2] z_2;  // standardized group-level effects
+  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+  real<lower=0> disc = 1;  // discrimination parameters
+  matrix[N_1, M_1] r_1;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_1] r_1_1;
+  vector[N_1] r_1_2;
+  vector[N_1] r_1_3;
+  vector[N_1] r_1_4;
+  vector[N_1] r_1_5;
+  vector[N_1] r_1_6;
+  vector[N_1] r_1_7;
+  matrix[N_2, M_2] r_2;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_2] r_2_1;
+  vector[N_2] r_2_2;
+  vector[N_2] r_2_3;
+  vector[N_2] r_2_4;
+  vector[N_2] r_2_5;
+  // compute actual group-level effects
+  r_1 = scale_r_cor(z_1, sd_1, L_1);
+  r_1_1 = r_1[, 1];
+  r_1_2 = r_1[, 2];
+  r_1_3 = r_1[, 3];
+  r_1_4 = r_1[, 4];
+  r_1_5 = r_1[, 5];
+  r_1_6 = r_1[, 6];
+  r_1_7 = r_1[, 7];
+  // compute actual group-level effects
+  r_2 = scale_r_cor(z_2, sd_2, L_2);
+  r_2_1 = r_2[, 1];
+  r_2_2 = r_2[, 2];
+  r_2_3 = r_2[, 3];
+  r_2_4 = r_2[, 4];
+  r_2_5 = r_2[, 5];
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = Xc * b;
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
+  }
+  // priors including constants
+  target += normal_lpdf(b[1] | 1, 0.1);
+  target += normal_lpdf(b[2] | 0, 0.1);
+  target += normal_lpdf(b[3] | 0, 0.1);
+  target += normal_lpdf(b[4] | 0, 0.1);
+  target += normal_lpdf(b[5] | 0, 0.1);
+  target += normal_lpdf(b[6] | 0, 0.1);
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 7 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_1));
+  target += lkj_corr_cholesky_lpdf(L_1 | 2);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 5 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_2));
+  target += lkj_corr_cholesky_lpdf(L_2 | 2);
+}
+generated quantities {
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
+  // compute group-level correlations
+  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+  vector<lower=-1,upper=1>[NC_1] cor_1;
+  // compute group-level correlations
+  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
+  vector<lower=-1,upper=1>[NC_2] cor_2;
+  // additionally sample draws from priors
+  real prior_b_1 = normal_rng(1,0.1);
+  real prior_b_2 = normal_rng(0,0.1);
+  real prior_b_3 = normal_rng(0,0.1);
+  real prior_b_4 = normal_rng(0,0.1);
+  real prior_b_5 = normal_rng(0,0.1);
+  real prior_b_6 = normal_rng(0,0.1);
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
+  real prior_sd_2 = normal_rng(1,0.1);
+  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_1) {
+    for (j in 1:(k - 1)) {
+      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+    }
+  }
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_2) {
+    for (j in 1:(k - 1)) {
+      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
+    }
+  }
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+```
+
+### Model 7
+
+```{stan, eval = FALSE, tidy = FALSE}
+// generated with brms 2.16.1
+functions {
+ /* compute correlated group-level effects
+  * Args: 
+  *   z: matrix of unscaled group-level effects
+  *   SD: vector of standard deviation parameters
+  *   L: cholesky factor correlation matrix
+  * Returns: 
+  *   matrix of scaled group-level effects
+  */ 
+  matrix scale_r_cor(matrix z, vector SD, matrix L) {
+    // r is stored in another dimension order than z
+    return transpose(diag_pre_multiply(SD, L) * z);
+  }
+  /* cratio-logit log-PDF for a single response
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: ordinal thresholds
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
+     int nthres = num_elements(thres);
+     vector[nthres + 1] p;
+     vector[nthres] q;
+     int k = 1;
+     while (k <= min(y, nthres)) {
+       q[k] = log_inv_logit(disc * (mu - thres[k]));
+       p[k] = log1m_exp(q[k]);
+       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
+       k += 1;
+     }
+     if (y == nthres + 1) {
+       p[nthres + 1] = sum(q);
+     }
+     return p[y];
+   }
+  /* cratio-logit log-PDF for a single response and merged thresholds
+   * Args:
+   *   y: response category
+   *   mu: latent mean parameter
+   *   disc: discrimination parameter
+   *   thres: vector of merged ordinal thresholds
+   *   j: start and end index for the applid threshold within 'thres'
+   * Returns:
+   *   a scalar to be added to the log posterior
+   */
+   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
+     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
+   }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=2> nthres;  // number of thresholds
+  int<lower=1> K;  // number of population-level effects
+  matrix[N, K] X;  // population-level design matrix
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  vector[N] Z_1_2;
+  vector[N] Z_1_3;
+  vector[N] Z_1_4;
+  vector[N] Z_1_5;
+  vector[N] Z_1_6;
+  vector[N] Z_1_7;
+  int<lower=1> NC_1;  // number of group-level correlations
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  vector[N] Z_2_2;
+  vector[N] Z_2_3;
+  vector[N] Z_2_4;
+  vector[N] Z_2_5;
+  int<lower=1> NC_2;  // number of group-level correlations
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+  int Kc = K;
+  matrix[N, Kc] Xc;  // centered version of X
+  vector[Kc] means_X;  // column means of X before centering
+  for (i in 1:K) {
+    means_X[i] = mean(X[, i]);
+    Xc[, i] = X[, i] - means_X[i];
+  }
+}
+parameters {
+  vector[Kc] b;  // population-level effects
+  vector[nthres] Intercept;  // temporary thresholds for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  matrix[M_1, N_1] z_1;  // standardized group-level effects
+  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  matrix[M_2, N_2] z_2;  // standardized group-level effects
+  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+  real<lower=0> disc = 1;  // discrimination parameters
+  matrix[N_1, M_1] r_1;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_1] r_1_1;
+  vector[N_1] r_1_2;
+  vector[N_1] r_1_3;
+  vector[N_1] r_1_4;
+  vector[N_1] r_1_5;
+  vector[N_1] r_1_6;
+  vector[N_1] r_1_7;
+  matrix[N_2, M_2] r_2;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_2] r_2_1;
+  vector[N_2] r_2_2;
+  vector[N_2] r_2_3;
+  vector[N_2] r_2_4;
+  vector[N_2] r_2_5;
+  // compute actual group-level effects
+  r_1 = scale_r_cor(z_1, sd_1, L_1);
+  r_1_1 = r_1[, 1];
+  r_1_2 = r_1[, 2];
+  r_1_3 = r_1[, 3];
+  r_1_4 = r_1[, 4];
+  r_1_5 = r_1[, 5];
+  r_1_6 = r_1[, 6];
+  r_1_7 = r_1[, 7];
+  // compute actual group-level effects
+  r_2 = scale_r_cor(z_2, sd_2, L_2);
+  r_2_1 = r_2[, 1];
+  r_2_2 = r_2[, 2];
+  r_2_3 = r_2[, 3];
+  r_2_4 = r_2[, 4];
+  r_2_5 = r_2[, 5];
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = Xc * b;
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
+    }
+    for (n in 1:N) {
+      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
+    }
+  }
+  // priors including constants
+  target += normal_lpdf(b[1] | 1, 0.1);
+  target += normal_lpdf(b[2] | 0, 0.1);
+  target += normal_lpdf(b[3] | 0, 0.1);
+  target += normal_lpdf(b[4] | 0, 0.1);
+  target += normal_lpdf(b[5] | 0, 0.1);
+  target += normal_lpdf(b[6] | 0, 0.1);
+  target += normal_lpdf(b[7] | 0, 0.1);
+  target += normal_lpdf(b[8] | 0, 0.1);
+  target += normal_lpdf(b[9] | 0, 0.1);
+  target += normal_lpdf(Intercept | -0.25, 0.1);
+  target += normal_lpdf(sd_1 | 1, 0.1)
+    - 7 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_1));
+  target += lkj_corr_cholesky_lpdf(L_1 | 2);
+  target += normal_lpdf(sd_2 | 1, 0.1)
+    - 5 * normal_lccdf(0 | 1, 0.1);
+  target += std_normal_lpdf(to_vector(z_2));
+  target += lkj_corr_cholesky_lpdf(L_2 | 2);
+}
+generated quantities {
+  // compute actual thresholds
+  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
+  // compute group-level correlations
+  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+  vector<lower=-1,upper=1>[NC_1] cor_1;
+  // compute group-level correlations
+  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
+  vector<lower=-1,upper=1>[NC_2] cor_2;
+  // additionally sample draws from priors
+  real prior_b_1 = normal_rng(1,0.1);
+  real prior_b_2 = normal_rng(0,0.1);
+  real prior_b_3 = normal_rng(0,0.1);
+  real prior_b_4 = normal_rng(0,0.1);
+  real prior_b_5 = normal_rng(0,0.1);
+  real prior_b_6 = normal_rng(0,0.1);
+  real prior_b_7 = normal_rng(0,0.1);
+  real prior_b_8 = normal_rng(0,0.1);
+  real prior_b_9 = normal_rng(0,0.1);
+  real prior_Intercept = normal_rng(-0.25,0.1);
+  real prior_sd_1 = normal_rng(1,0.1);
+  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
+  real prior_sd_2 = normal_rng(1,0.1);
+  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_1) {
+    for (j in 1:(k - 1)) {
+      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+    }
+  }
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_2) {
+    for (j in 1:(k - 1)) {
+      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
+    }
+  }
+  // use rejection sampling for truncated priors
+  while (prior_sd_1 < 0) {
+    prior_sd_1 = normal_rng(1,0.1);
+  }
+  while (prior_sd_2 < 0) {
+    prior_sd_2 = normal_rng(1,0.1);
+  }
+}
+```
+
+### {-}
 
@@
 ```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses %>%
-  mutate(lv = cut_interval(lv, 5)) %>% 
-  group_by(te, lv) %>%
-  summarise(
-    understands = mean(understands, na.rm = TRUE),
-    produces = mean(produces, na.rm = TRUE),
-    .groups = "drop"
+df %>%
+    mutate(
+        lv = cut_interval(lv, 5),
+        age = floor(age)
     ) %>% 
-  pivot_longer(
-    c(understands, produces),
-    names_to = "type",
-    values_to = "prop"
-  ) %>%
-  mutate(type = str_to_sentence(type)) %>% 
-  ggplot(aes(prop, reorder(te, -prop), colour = lv, fill = lv)) +
-  facet_wrap(type~lv, scales = "free_y", ncol = 5) +
-  geom_vline(xintercept = 0.5, linetype = "dashed", size = 1) +
-  geom_point(alpha = 0.5, size = 2, shape = 1, stroke = 1) +
-  geom_smooth(aes(group = lv)) +
+    count(lv, age, response) %>%
+    mutate(prop = n/sum(.$n)) %>% 
+    ggplot(aes(age, prop, colour = response, fill = response)) +
+    facet_wrap(~lv, nrow = 1) +
+    geom_col(position = position_fill()) +
+    geom_hline(yintercept = 0.5, colour = "black", size = 0.5) +
     labs(
-    x = "Mean % of acquisition", y = "Translation equivalent ID", 
-    colour = "Levenshtein similarity", fill = "Levenshtein similarity"
+        x = "Age (months)",
+        y = "% responses", 
+        colour = "Response",
+        fill = "Response"
     ) +
-  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
+    scale_fill_d3() +
+    scale_colour_d3() +
+    scale_x_continuous(breaks = seq(0, 40, 2)) +
+    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
     theme(
@@
 
-```{r raw_age, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses %>%
-  mutate(
-    lv = cut_interval(lv, 5),
-    age = as.factor(round(age))
-  ) %>% 
-  group_by(te, age, lv) %>%
-  summarise(
-    understands = mean(understands, na.rm = TRUE),
-    produces = mean(produces, na.rm = TRUE),
-    .groups = "drop"
-  ) %>%
-  pivot_longer(
-    c(understands, produces),
-    names_to = "type",
-    values_to = "prop"
-  ) %>%
-  mutate(type = str_to_sentence(type)) %>% 
-  ggplot(aes(age, prop, colour = lv, fill = lv)) +
-  facet_grid(type~lv) +
-  geom_hline(yintercept = 0.5, size = 0.75) +
-  stat_pointinterval(
-    .width = c(0.5), size = 1
-  ) +
-  labs(
-    x = "Age (months)", y = "Mean % of acquisition",
-    colour = "Cognate", shape = "Cognate", fill = "Cognate"
-  ) +
-  scale_x_discrete(breaks = seq(10, 36, 2)) +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme(
-    legend.position = "none",
-    axis.text.x = element_text(size = 7),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    panel.grid = element_blank(),
-    legend.title = element_blank()
-  )
-```
-
 ## Model selection
@@
 
-### Comprehension
-
-```{r loo_comp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r loos, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, eval = FALSE}
 loo_comp %>% 
@@
 
-### Production
-
-```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-loo_prod %>% 
-  as_tibble() %>% 
-  rownames_to_column("model") %>% 
-  relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
-  mutate_all(as.numeric) %>% 
-  gt() %>% 
-  tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
-  tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
-  tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
-  tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
-  fmt_number(3:13) %>% 
-  cols_label(
-    model = "Model",
-    # elpd
-    elpd_loo = md("*ELPD*"),
-    se_elpd_loo = md("*SE*"),
-    subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
-    # p
-    p_loo = md("*p*"),
-    se_p_loo = md("SE"),
-    subsampling_se_p_loo = md("SE<sub>sub</sub>"),
-    # looic
-    looic = md("*LOO-IC*"),
-    se_looic = md("*SE*"),
-    subsampling_se_looic = md("SE<sub>sub</sub>"),
-    # diff
-    elpd_diff = md("*diff*"),
-    se_diff = md("*SE*"),
-    subsampling_se_diff = md("*SE<sub>sub</sub>*"),
-  ) %>% 
-  as_raw_html()
-```
-
 ## Fixed effects
-
-### Fixed effects: Compehension
 
@@
 str_repl <- c(
-  "doe_center:lv_center" = "DoE \u00d7 Levenshtein",
+    "age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein",
+    "age_std:doe_std" = "Age \u00d7 DoE",
+    "age_std:lv_std" = "Age \u00d7 Levenshtein",
+    "doe_std:lv_std" = "DoE \u00d7 Levenshtein",
     "Intercept" = "Intercept",
-  "age_center" = "Age (+1 month)",
-  "frequency_center" = "Frequency (+1 SD)",
-  "doe_center" = "DoE (+10%)",
-  "lv_center" = "Levenshtein (+1 SD)"
-)
-
-draws_fix <- fixef(fits_comp$fit_2) %>% 
+    "age_std" = "Age (+1 month)",
+    "freq_std" = "Frequency (+1 SD)",
+    "n_phon_std" = "# Phonemes (+1 SD)",
+    "doe_std" = "DoE (+10%)",
+    "lv_std" = "Levenshtein (+1 SD)"
+)
+
+draws_fix <- fixef(model_fit_7) %>% 
     as.data.frame() %>%
@@
         footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("Age (+1 month)", "Frequency (+ 1 SD)",  "DoE (+10%)", "Levenshtein (+ 1 SD)", "doe_center:lv_center" = "DoE \u00d7 Levenshtein"))
+        locations = cells_body(columns = "term", rows = term %in% c("Age (+1 month)", "Frequency (+ 1 SD)",  "DoE (+10%)", "Levenshtein (+ 1 SD)", "doe_std:lv_std" = "DoE \u00d7 Levenshtein"))
     ) %>% 
@@
         columns = 2:3
-  ) %>% 
-  as_raw_html()
+    )
 
@@
 
-```{r fix_effets_comp_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+```{r fix_effects-plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
 The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
+The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
 
 var_repl <- c(
-  "b_doe_center:lv_center" = "DoE \u00d7 Levenshtein", 
+    "b_Intercept[1]" = "Intercept (Comprehension)",
+    "b_Intercept[2]" = "Intercept (Production)",
+    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein", 
+    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
+    "b_age_std:doe_std" = "Age \u00d7 DoE",
+    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein", 
     "b_Intercept" = "Intercept", 
-  "b_age_center" = "Age (+1 month)", 
-  "b_frequency_center" = "Frequency (+1 SD)",
-  "b_doe_center" = "DoE (+10%)", 
-  "b_lv_center" = "Levenshtein (+1 SD)", 
-  "sd_te__Intercept" = "Intercept", 
-  "sd_te__age_center" = "Age",
-  "sd_te__doe_center:lv_center" = "DoE \u00d7 Levenshtein",
-  "sd_te__doe_center" = "DoE",
-  "sd_te__lv_center" = "Levenshtein"
-)
-
-post <- gather_draws(fits_comp$fit_2, `b_.*`, `sd_te_.*`, regex = TRUE, ndraws = 1000) %>% 
+    "b_age_std" = "Age (+1 month)", 
+    "b_freq_std" = "Frequency (+1 SD)",
+    "b_n_phon_std" = "# Phonemes (+1 SD)",
+    "b_doe_std" = "DoE (+1 SD)", 
+    "b_lv_std" = "Levenshtein (+1 SD)"
+)
+
+post <- gather_draws(model_fit_7, `b_.*`, regex = TRUE) %>% 
     mutate(
         .value = ifelse(grepl("Intercept|sd", .variable), inv_logit_scaled(.value), .value/4),
-    .variable_name = str_replace_all(.variable, var_repl) 
+        .variable_name = factor(.variable, levels = names(var_repl), labels = var_repl)
     )
@@
 post_intercept <- post %>% 
-  filter(.variable=="b_Intercept") %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
+    filter(.variable %in% c("b_Intercept[1]", "b_Intercept[2]")) %>% 
+    ggplot(aes(.value, .variable_name)) +
     geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
@@
     labs(
-    x = "Comprehension posterior probability", 
+        x = "Posterior probability of acquisition", 
         y = "Variable", 
@@
     filter(!grepl("Intercept|sd", .variable)) %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
+    ggplot(aes(.value, .variable_name)) +
     geom_vline(xintercept = 0, size = 0.5, linetype = "dashed") +
@@
     ) +
-  annotate(
-    geom = "label",
-    fill = "white",
-    label.size = 0,
-    # alpha = 0.5,
-    x = -0.05, 
-    y = 0.75,
-    label = "Increases\nP(comprehension)",
-    size = 3.5,
-    hjust = 0
-  ) +
-  annotate(
-    geom = "label",
-    fill = "white",
-    label.size = 0,
-    # alpha = 0.5,
-    x = 0.2, 
-    y = 0.75,
-    label = "Decreases\nP(comprehension)",
-    size = 3.5,
-    hjust = 1
-  ) +
-  annotate(
-    geom = "segment",
-    x = 0,
-    xend = -0.05,
-    y = 0.5,
-    yend =  0.5,
-    arrow = arrow(length = unit(0.2, "cm"))
-  ) +
-  annotate(
-    geom = "segment",
-    x = 0.15,
-    xend = 0.2,
-    y = 0.5,
-    yend = 0.5,
-    arrow = arrow(length = unit(0.2, "cm"))
-  ) +
     stat_pointinterval(point_size = 2) +
     labs(
-    x = "Comprehension posterior probability",
+        x = "Posterior probability of acquisition",
         y = "Variable", 
@@
     scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, breaks = seq(-0.15, 2, 0.05))
-
-post_sd <- post %>% 
-  filter(grepl("sd", .variable)) %>% 
-  mutate(group = ifelse(grepl("id", .variable), "Participant", "TE")) %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
-  stat_slab(
-    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
-    size = 0.5,
-    color = "#F98400",
-    fill = "#F98400",
-    position = position_nudge(y = 0.1)
-  ) +
-  stat_pointinterval(point_size = 2) +
-  labs(
-    x = "Comprehension posterior probability",
-    y = "Variable", 
-    fill = "Variable",
-    alpha = "CrI"
-  ) +
-  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, breaks = seq(0.5, 1, 0.05)) +
+    scale_x_continuous(labels = percent, breaks = seq(-0.2, 2.5, 0.1))
+
+(post_intercept / post_fix) +
+    plot_layout(heights = c(0.2, 0.8)) &
     theme(
-    axis.text.y = element_blank()
-  )
-
-(post_intercept / (post_fix | post_sd)) +
-  plot_layout(heights = c(0.1, 0.9)) &
-  theme(
-    legend.position = c(0.9, 0.9),
+        legend.position = "top",
         legend.direction = "horizontal",
@@
 
-### Fixed effects: Production
-
-```{r fix_effets_prod_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
-The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-
-var_repl <- c(
-  "b_doe_center:lv_center" = "DoE \u00d7 Levenshtein", 
-  "b_Intercept" = "Intercept", 
-  "b_age_center" = "Age (+1 month)", 
-  "b_frequency_center" = "Frequency (+1 SD)",
-  "b_doe_center" = "DoE (+10%)", 
-  "b_lv_center" = "Levenshtein (+1 SD)", 
-  "sd_te__Intercept" = "Intercept", 
-  "sd_te__age_center" = "Age",
-  "sd_te__doe_center:lv_center" = "DoE \u00d7 Levenshtein",
-  "sd_te__doe_center" = "DoE",
-  "sd_te__lv_center" = "Levenshtein"
-)
-
-post <- gather_draws(fits_prod$fit_2, `b_.*`, `sd_te_.*`, regex = TRUE, ndraws = 1000) %>% 
-  mutate(
-    .value = ifelse(grepl("Intercept|sd", .variable), inv_logit_scaled(.value), .value/4),
-    .variable_name = str_replace_all(.variable, var_repl) 
-  )
-
-post_intercept <- post %>% 
-  filter(.variable=="b_Intercept") %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
-  geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
-  stat_slab(
-    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
-    size = 0.5,
-    color = "#FF0000",
-    fill = "#FF0000",
-    position = position_nudge(y = 0.2)
-  ) +
-  stat_pointinterval(point_size = 2) +
-  labs(
-    x = "Comprehension posterior probability", 
-    y = "Variable", 
-    fill = "Variable",
-    alpha = "CrI"
-  ) +
-  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, limits = c(0, 1))
-
-post_fix <- post %>% 
-  filter(!grepl("Intercept|sd", .variable)) %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
-  geom_vline(xintercept = 0, size = 0.5, linetype = "dashed") +
-  stat_slab(
-    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
-    size = 0.5,
-    color = "#00A08A",
-    fill = "#00A08A",
-    position = position_nudge(y = 0.1)
-  ) +
-  annotate(
-    geom = "label",
-    fill = "white",
-    label.size = 0,
-    # alpha = 0.5,
-    x = -0.05, 
-    y = 0.75,
-    label = "Increases\nP(comprehension)",
-    size = 3.5,
-    hjust = 0
-  ) +
-  annotate(
-    geom = "label",
-    fill = "white",
-    label.size = 0,
-    # alpha = 0.5,
-    x = 0.2, 
-    y = 0.75,
-    label = "Decreases\nP(comprehension)",
-    size = 3.5,
-    hjust = 1
-  ) +
-  annotate(
-    geom = "segment",
-    x = 0,
-    xend = -0.05,
-    y = 0.5,
-    yend =  0.5,
-    arrow = arrow(length = unit(0.2, "cm"))
-  ) +
-  annotate(
-    geom = "segment",
-    x = 0.15,
-    xend = 0.2,
-    y = 0.5,
-    yend = 0.5,
-    arrow = arrow(length = unit(0.2, "cm"))
-  ) +
-  stat_pointinterval(point_size = 2) +
-  labs(
-    x = "Comprehension posterior probability",
-    y = "Variable", 
-    fill = "Variable",
-    alpha = "CrI"
-  ) +
-  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, breaks = seq(-0.15, 2, 0.05))
-
-post_sd <- post %>% 
-  filter(grepl("sd", .variable)) %>% 
-  mutate(group = ifelse(grepl("id", .variable), "Participant", "TE")) %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
-  stat_slab(
-    aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
-    size = 0.5,
-    color = "#F98400",
-    fill = "#F98400",
-    position = position_nudge(y = 0.1)
-  ) +
-  stat_pointinterval(point_size = 2) +
-  labs(
-    x = "Comprehension posterior probability",
-    y = "Variable", 
-    fill = "Variable",
-    alpha = "CrI"
-  ) +
-  scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, breaks = seq(0.5, 1, 0.05)) +
-  theme(
-    axis.text.y = element_blank()
-  )
-
-(post_intercept / (post_fix | post_sd)) +
-  plot_layout(heights = c(0.1, 0.9)) &
-  theme(
-    legend.position = c(0.9, 0.9),
-    legend.direction = "horizontal",
-    legend.justification = "right",
-    plot.caption.position = "plot",
-    plot.caption = element_text(hjust = 0),
-    axis.title.y = element_blank(),
-    # axis.text = element_text(size = 7),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-    panel.grid.major.y = element_blank()
-  )
-```
-
 ## Random effects {.tabset .tab-pils}
 
-### Random effects: Comprehension
-
-#### Participant
-
-```{r random_participant_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-re_id <- ranef(fits_comp$fit_2)$id[,,1] %>% 
+### Participant
+
+```{r random-participant, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+re_id <- ranef(model_fit_7)$id[,,1] %>% 
     as.data.frame() %>% 
@@
     as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
-  left_join(distinct(fits_comp$fit_2$data, id)) %>% 
+    mutate_if(is.numeric, ~inv_logit_scaled(. + fixef(model_fit_7)[1])) %>% 
+    left_join(distinct(model_fit_7$data, id)) %>% 
     clean_names()
@@
 ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
-  geom_ribbon(aes(group = 1), alpha = 0.5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
     geom_line(aes(group = 1), size = 1) +
+    # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
+    # geom_point(size = 0.5) +
+    geom_vline(xintercept = 0.5, colour = "white") +
     labs(
-    x = "Participant posterior probability of comprehension",
+        x = "Posterior probability of acquisition",
         y = "Participant", 
@@
         axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    panel.grid = element_line(colour = "white", linetype = "dotted"),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
         panel.grid.major.y = element_blank(),
-    panel.grid.minor.y = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    axis.line = element_line(colour = "white")
-  )
-```
-
-#### Item
-
-```{r re_item_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-re_te <- ranef(fits_comp$fit_2)$te[,,1] %>% 
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+### Item
+
+```{r re-item, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+re_te <- ranef(model_fit_7)$te[,,1] %>% 
     as.data.frame() %>% 
@@
     as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
+    mutate_if(is.numeric, ~inv_logit_scaled(. + fixef(model_fit_7)[1])) %>% 
     mutate(te = as.numeric(te)) %>% 
@@
     labs(
-    x = "Participant posterior probability of comprehension",
+        x = "Posterior probability of acquisition",
         y = "TE", 
@@
         axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    panel.grid = element_line(colour = "white", linetype = "dotted"),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
         panel.grid.major.y = element_blank(),
-    panel.grid.minor.y = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    axis.line = element_line(colour = "white")
-  )
-```
-
-### Random effects: Production
-
-#### Participant
-
-```{r random_participant_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-re_id <- ranef(fits_prod$fit_2)$id[,,1] %>% 
-  as.data.frame() %>% 
-  rownames_to_column("id") %>% 
-  as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
-  left_join(distinct(fits_prod$fit_2$data, id)) %>% 
-  clean_names()
-
-ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
-  geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
-  geom_line(aes(group = 1), size = 1) +
-  # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
-  # geom_point(size = 0.5) +
-  geom_vline(xintercept = 0.5, colour = "white") +
-  labs(x = "Participant posterior probability of production",
-       y = "Participant", colour = "Lang. Prof.") +
-  scale_x_continuous(limits = c(0, 1), labels = percent) +
-  theme(
-    legend.position = "none",
-    axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    panel.grid = element_line(colour = "white", linetype = "dotted"),
-    panel.grid.major.y = element_blank(),
-    panel.grid.minor.y = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    axis.line = element_line(colour = "white")
-  )
-```
-
-#### Item
-
-```{r re_item_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-re_te <- ranef(fits_prod$fit_2)$te[,,1] %>% 
-  as.data.frame() %>% 
-  rownames_to_column("te") %>% 
-  as_tibble() %>% 
-  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
-  mutate(te = as.numeric(te)) %>% 
-  clean_names()
-
-ggplot(re_te, aes(estimate, reorder(te, estimate), xmin = q2_5, xmax = q97_5)) +
-  geom_ribbon(aes(group = 1), alpha = 0.5) +
-  geom_line(aes(group = 1), size = 1) +
-  labs(
-    x = "Participant posterior probability of comprehension",
-    y = "TE", 
-    colour = "Lang. Prof."
+        panel.grid.minor.y = element_blank()
+    )
+
+```
+
+### {-}
+
+## Marginal means
+
+### Marginal means: Comprehension
+
+```{r emmeans_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+nd <- expand.grid(
+    n_phon_std = 0,
+    freq_std = 0,
+    age_std = seq(-4, 4, length.out = 30),
+    doe_std = range(df$doe_std),
+    lv_std = c(range(df$lv_std), 0)
+)
+
+m <- add_epred_draws(nd, model_fit_7, ndraws = 50, re_formula = NA) %>% 
+    mutate(
+        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
+        doe_std = factor(
+            doe_std, 
+            levels = unique(nd$doe_std), 
+            labels = c("DoE: 0%", "DoE: 100%")
+        ),
+        lv_std = factor(
+            lv_std, 
+            levels = unique(nd$lv_std)[c(1, 3, 2)],
+            labels = c("0%", "50%", "100%")
+        )
+    ) %>% 
+    filter(.category != "No") %>% 
+    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
+    mutate(Understands = Understands + `Understands and Says`) %>% 
+    pivot_longer(
+        c(Understands, `Understands and Says`),
+        names_to = ".category",
+        values_to = ".epred"
+    )
+
+ggplot(m, aes(age_std, .epred, colour = lv_std, fill = lv_std)) +
+    facet_grid(.category~doe_std) +
+    geom_line(
+        aes(
+            group = interaction(.draw, lv_std, doe_std), 
+            colour = lv_std),
+        alpha = 0.1, 
+        size = 1
     ) +
-  scale_x_continuous(limits = c(0, 1), labels = percent) +
-  theme(
-    legend.position = "none",
-    axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    panel.grid = element_line(colour = "white", linetype = "dotted"),
-    panel.grid.major.y = element_blank(),
-    panel.grid.minor.y = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    axis.line = element_line(colour = "white")
-  )
-
-```
-
-### {-}
-
-## Marginal means
-
-### Marginal means: Comprehension
-
-```{r emmeans_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-nd <- expand.grid(
-  age_center = seq(min(responses$age_center), max(responses$age_center), length.out = 30),
-  doe_center = range(responses$doe_center),
-  lv_center = c(range(responses$lv_center), 0),
-  frequency_center = 0
-)
-
-m <- add_epred_draws(nd, fits_comp$fit_2, ndraws = 50, re_formula = NA) %>% 
-  mutate(
-    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%")),
-    lv_center = factor(lv_center, levels = unique(nd$lv_center), labels = c("LV: 0%", "LV: 50%", "LV: 100%"))
-
-  )
-
-ggplot(m, aes(age_center + mean(responses$age, na.rm = TRUE), .epred, colour = lv_center, fill = lv_center)) +
-  facet_wrap(~doe_center) +
-  geom_line(aes(group = interaction(.draw, lv_center), colour = lv_center), alpha = 0.25, size = 1) +
     # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
-  # stat_summary(fun = "mean", geom = "line", size = 1) + 
-  geom_hline(yintercept = 0.5, size = 1, linetype = "dashed") +
+    stat_summary(fun = "mean", geom = "line", size = 1) + 
+    geom_hline(yintercept = 0.5, size = 0.5, linetype = "dashed") +
+    scale_color_d3() +
     labs(
         x = "Age (months)", 
-    y = "Posterior probability of comprehension\n(Linear prediction)",
-    colour = "Cognateness", 
-    fill = "Cognateness"
+        y = "Posterior probability of acquisition\n(Linear prediction)",
+        colour = "Phonological similarity (Levenshtein distance)", 
+        fill = "Phonological similarity (Levenshtein distance)"
     ) +
@@
     theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-
-```
-
-## Marginal means: Production
-
-```{r emmeans_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-nd <- expand.grid(
-  age_center = seq(min(responses$age_center), max(responses$age_center), length.out = 30),
-  doe_center = range(responses$doe_center),
-  lv_center = c(range(responses$lv_center), 0),
-  frequency_center = 0
-)
-
-m <- add_epred_draws(nd, fits_prod$fit_2, ndraws = 50, re_formula = NA) %>% 
-  mutate(
-    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%")),
-    lv_center = factor(lv_center, levels = unique(nd$lv_center), labels = c("LV: 0%", "LV: 50%", "LV: 100%"))
-  )
-
-ggplot(m, aes(age_center + mean(responses$age, na.rm = TRUE), .epred, colour = lv_center, fill = lv_center)) +
-  facet_wrap(~doe_center) +
-  geom_line(aes(group = interaction(.draw, lv_center), colour = lv_center), alpha = 0.25, size = 1) +
-  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
-  stat_summary(fun = "mean", geom = "line", size = 1) +
-  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
-  labs(x = "Age (months)", y = "Posterior probability of production\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
-  scale_y_continuous(labels = percent, limits = c(0, 1)) +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-
-```
-
-### Area under the curve (AUC)
-
-```{r confusion_matrix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# comprehension
-preds <- predict(fits_comp$fit_2, ndraws = 50) 
-preds <- ifelse(preds[, 1] > 0.5, 1, 0)
-confusion_mat_comp <- table(preds, pull(responses, understands))
-#correct classification rate
-rate_comp <- sum(diag(confusion_mat_comp))/sum(confusion_mat_comp)
-confusion_mat_comp <- confusion_mat_comp %>% 
-  as.data.frame() 
-colnames(confusion_mat_comp) <- c("pred", "value", "freq")
-
-# production
-preds <- predict(fits_prod$fit_2, ndraws = 50) 
-preds <- ifelse(preds[, 1] > 0.5, 1, 0)
-confusion_mat_prod <- table(preds, pull(responses, produces))
-#correct classification rate
-rate_prod <- sum(diag(confusion_mat_prod))/sum(confusion_mat_prod)
-confusion_mat_prod <- confusion_mat_prod %>% 
-  as.data.frame() 
-colnames(confusion_mat_prod) <- c("pred", "value", "freq")
-```
-
-The comprehension model classified `r percent(round(rate_comp, 2))` of the responses correctly and the production model classified `r percent(round(rate_prod, 2))` of the responses correctly.
-
-```{r confusion, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-bind_rows(
-  comp = confusion_mat_comp, 
-  prod = confusion_mat_prod, 
-  .id = "type"
+        legend.position = "top"
+    )
+
+```
+
+### Traceplots
+
+```{r diagnostics_traceplots_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
+gather_draws(
+    model_fit_7, 
+    `b_.*`, `sd_te__.*`, `cor_te__.*`, 
+    regex = TRUE, 
+    ndraws = 1000
 ) %>% 
-  mutate(pred = pred==1) %>% 
-  pivot_wider(names_from = c(type, value), values_from = freq) %>% 
-  clean_names() %>% 
-  gt() %>% 
-  tab_spanner("Observed comprehension", matches("comp")) %>% 
-  tab_spanner("Observed production", matches("prod")) %>% 
-  cols_label(
-    pred = "Predicted",
-    comp_false = "FALSE",
-    comp_true = "TRUE",
-    prod_false = "FALSE",
-    prod_true = "TRUE"
-  ) %>% 
-  fmt_number(2:5, decimals = 0)
-
-```
-
-### Traceplots
-
-```{r diagnostics_traceplots_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
-gather_draws(fits_comp$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE, ndraws = 1000) %>% 
     mutate(.chain = paste("Chain ", .chain)) %>% 
@@
     facet_wrap(~.variable, scales = "free_y") +
-  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
-  geom_line(alpha = 0.5) +
-  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Comprehension") +
-  theme_github() +
+    annotate(
+        geom = "rect", 
+        colour = NA, 
+        xmin = 0,
+        xmax = 500, 
+        ymin = -Inf, 
+        ymax = Inf, 
+        alpha = 0.25, 
+    ) +
+    geom_line() +
+    labs(
+        x = "Iteration",
+        y = "Sample value",
+        colour = "Chain"
+    ) +
+    scale_color_d3() +
     theme(
@@
 
-```{r diagnostics_traceplots_prod, message=FALSE, warning=FALSE, paged.print=FALSE}
-gather_draws(fits_prod$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE, ndraws = 1000) %>% 
-  mutate(.chain = paste("Chain ", .chain)) %>% 
-  ggplot(aes(.iteration, .value, colour = .chain)) +
-  facet_wrap(~.variable, scales = "free_y") +
-  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
-  geom_line(alpha = 0.5) +
-  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Production") +
-  theme_github() +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank(),
-    panel.grid = element_blank()
-  )
-
-```
-
 ## {-}

### 2022-08-29 docs/index.Rmd
Update report

@@
 title: "Trajectories"
+authors:
+    - Gonzalo Garcia-Castro^[Universitat Pompeu Fabra, gonzalo.garciadecastro@upf.edu]
+    - Daniela Avila-Varela^[Universitat Pompeu Fabra]
+    - Ignacio Castillejo^[Universidad Autónoma de Madrid]
+    - Núria Sebastian-Galles^[Universitat Pompeu Fabra]
 output:
+    html_document:
+        code_folding: hide
+        number_sections: true
+        toc: true
     github_document:
         pandoc_args: --webtex
-bibliography: "references.bib"
+        number_sections: true
+        toc: true
+bibliography: "../references.bib"
 ---
@@
     message = TRUE,
-    echo = FALSE,
+    echo = TRUE,
     warning = TRUE,
@@
 
-```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
-# set params
+```{r prepare, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
+# set params and load target contents
 tar_load_globals()
@@
 
+conflict_prefer("filter", "dplyr")
+conflict_prefer("select", "dplyr")
+
 ```
@@
 # Questionnaires
+
+```{r items, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
+pool <- multilex::pool %>% 
+    drop_na(cognate)
+n_categories <- distinct(pool, category) %>% nrow()
+n_items <- length(unique(pool$item))
+n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
+n_item_cognate <- count(pool, language, cognate) %>% 
+    pull(n) %>% 
+    set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
+```
 
@@
 
-```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
+```{r items_frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
 Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
 
+items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))
+
 items %>%
-    left_join(select(multilex_data$pool, te, language)) %>% 
     unnest(list) %>% 
@@
 
-```{r items-frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r items-frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
 
 items %>% 
-    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
     ggplot() +
@@
         legend.position = "top",
-        axis.title = element_blank()
+        legend.title = element_blank()
     )
@@
 
-```{r items-nphon, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r items-nphon, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
 
 items %>%
-    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
     count(language, n_phon) %>% 
@@
     geom_col(
-        position = position_dodge()
+        position = position_dodge(width = 0.95)
     ) +
@@
     labs(
-        x = "Frequency (Zipf)",
+        x = "# Phonemes",
         y = "Probability density",
@@
     ) +
-    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
+    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
+    scale_x_continuous(breaks = seq(0, 20, 1)) +
     theme(
         legend.position = "top",
-        axis.title = element_blank()
+        axis.ticks.x = element_line(),
+        legend.title = element_blank()
     )
@@
 
-```{r items-lv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r items-lv, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
 
@@
     geom_col(fill = ggsci::pal_d3()(4)[1]) +
+    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
     labs(
@@
     theme(
-        legend.position = "top",
-        axis.title = element_blank()
+        legend.position = "top"
     )
@@
 ```{r items-frequency-phonemes, message=FALSE, warning=FALSE, paged.print=FALSE}
-
 items %>% 
-    left_join(select(pool, item, language)) %>% 
     drop_na(freq) %>% 
@@
     ) 
-
 ```
@@
 ```{r items-frequency-levenshtein, message=FALSE, warning=FALSE, paged.print=FALSE}
-
 items %>% 
@@
     )
-
 ```
@@
 ```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
+participants <- left_join(
+    participants, 
+    select(multilex_data$logs, id, time, edu_parent, dominant_language = dominance)
+)
+
 participants %>% 
-    mutate(dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish")) %>% 
     group_by(time_stamp, dominant_language) %>% 
@@
 
-```{r participants-age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants-age, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
 participants %>% 
@@
     aes(x = age, y = n) +
-    geom_col() +
+    geom_col(fill = pal_d3()(3)[3]) +
     geom_text(aes(label = n), size = 3.5, vjust = -1) +
@@
     mutate(
-        dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish"),
         age = round(age),
@@
     aes(x = doe, y = reorder(id, -rank), fill = language) +
-    geom_col(width = 0.75, position = position_fill(), colour = "white", size = 0.1) +
+    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
+    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
     labs(
-        x = "Degree of exposure to Catalan",
+        x = "Degree of Exposure (DoE)",
         y = "Participant",
@@
         axis.ticks.y = element_blank(),
-        panel.grid.major.y = element_blank()
+        panel.grid.major.y = element_blank(),
+        axis.title = element_blank()
     )
@@
 
-```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants_edu, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
 participants %>%
-    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
     mutate(
@@
     pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
-    mutate(total = rowSums(cbind(.[,3:8]))) %>% 
+    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
     rename(Total = total) %>% 
@@
 Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`.
-The models were implemented in `brms` as:
-
-* `understands/produces ~ 1 + age_std + frequency + (1 + age_std | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age_std + frequency + doe + (1 + age_std + doe | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age_std + frequency + doe*cognate + (1 + age_std + doe | item) + (1 + frequency + cognate| id)`
 
 ## Model equation
+
+$$\begin{aligned}
+g[\mathrm{Pr}(y_{ij} \le a_s)] = \tau_s - \nu_{ij}
+\end{aligned}$$
+
+$s$ is the s-th response category, $a_1 < a_2 < a_3$ are ordered response categories, $\mathrm{Pr}(y_i \le a_3) = 1$ and $\tau_s$ are threshold parameters where $\tau_1 < \tau_2 < \tau_3$.
 
@@
 \begin{aligned}
-\color{white}\log(\frac{p}{n - q}) &= \color{white}(\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
-& \color{white}(\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
-& \color{white}(\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
-& \color{white}(\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
-& \color{white}(\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
-& \color{white}(\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
-& \color{white}\varepsilon_{pi}
+
+\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} +
+\beta_6\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_7\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_8\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + 
+\beta_9{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) +\\
+
+&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{TE})} \mathrm{Freq}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{5j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) +\\
+
+&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{ID})} \mathrm{Freq}_{j} +   \theta_{3j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{5j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{6j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) +
+\theta_{7j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}) +
+\theta_{8j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j})
+
 \end{aligned}
@@
 
+## Model prior
+
+$$\begin{aligned}
+\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
+\beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
+\beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
+\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]
+\end{aligned}
+$$
+
+In addition, we specified a $LKJ$ prior ($\sim LKJ(2)$ for the correlations between group-level coefficients. We implemented the model prior following {brms} syntax as:
+
+```r
+c(
+prior(normal(-0.25, 0.1), class = "Intercept"),
+prior(normal(1, 0.1), class = "sd", group = "te"),
+prior(normal(1, 0.1), class = "sd", group = "id"),
+prior(lkj(2), class = "cor"),
+prior(normal(1, 0.1), class = "b", coef = "age_std"),
+prior(normal(0, 0.1), class = "b", coef = "freq_std"),
+prior(normal(0, 0.1), class = "b", coef = "n_phon_std"),
+prior(normal(0, 0.1), class = "b", coef = "doe_std"),
+prior(normal(0, 0.1), class = "b", coef = "lv_std"),
+prior(normal(0, 0.1), class = "b", coef = "doe_std:lv_std"),
+prior(normal(0, 0.1), class = "b", coef = "age_std:doe_std"),
+prior(normal(0, 0.1), class = "b", coef = "age_std:lv_std"),
+prior(normal(0, 0.1), class = "b", coef = "age_std:doe_std:lv_std")
+)
+```
+
+## Model settings
+
+* OS: `r sessionInfo()$running`, `r `sessionInfo()$platform`
+* R version: `r sessionInfo()$R.version$version.string`
+* Algorithm: Hamiltonian Montecarlo (No U-turn Sampler, *NUTS*)
+* Engine: Stan (brms interface, `r sessionInfo()$otherPkgs$brms$Version`)
+* brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
+* Chains: `r dim(model_fit_7$fit)[2]`
+* Cores: `r dim(model_fit_7$fit)[2]`
+* Iterations: `r dim(model_fit_7$fit)[1]` (`r dim(model_fit_7$fit)[1]/2`)
+
+More details in Appendix: Session Info.
+
 ## R code (`brms`) {.tabset}
@@
 
-```{stan, eval = FALSE, tidy = FALSE}
+```
 // generated with brms 2.16.1
@@
 
-```{r loos, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, eval = FALSE}
-loo_comp %>% 
+```{r loos, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
+model_loos %>% 
     as_tibble() %>% 
     rownames_to_column("model") %>% 
-    relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
+    relocate(
+        model, 
+        matches("elpd_loo"),
+        matches("p_loo"), 
+        matches("looic"), 
+        matches("diff")
+    ) %>%
     mutate_all(as.numeric) %>% 
@@
         subsampling_se_diff = md("*SE<sub>sub</sub>*"),
-    ) %>% 
-    as_raw_html()
+    ) 
 ```
@@
 The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-
 var_repl <- c(
@@
     mutate(
-        .value = ifelse(grepl("Intercept|sd", .variable), inv_logit_scaled(.value), .value/4),
-        .variable_name = factor(.variable, levels = names(var_repl), labels = var_repl)
+        .value = ifelse(
+            grepl("Intercept|sd", .variable),
+            inv_logit_scaled(.value), 
+            .value/4
+        ),
+        .variable_name = factor(
+            .variable, 
+            levels = names(var_repl),
+            labels = var_repl
+        )
     )
@@
 ```{r random-participant, message=FALSE, warning=FALSE, paged.print=FALSE}
-
 re_id <- ranef(model_fit_7)$id[,,1] %>% 
-    as.data.frame() %>% 
+    as_tibble() %>% 
     rownames_to_column("id") %>% 
-    as_tibble() %>% 
-    mutate_if(is.numeric, ~inv_logit_scaled(. + fixef(model_fit_7)[1])) %>% 
     left_join(distinct(model_fit_7$data, id)) %>% 
-    clean_names()
-
-ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
+    clean_names() %>% 
+    mutate(
+        across(
+            .cols = where(is.numeric),
+            .fns = list(
+                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_7)[1]),
+                Production = ~inv_logit_scaled(. + fixef(model_fit_7)[2])
+            ), .names = "{.col}__{.fn}"
+        )
+    ) %>% 
+    select(-c(estimate:q97_5)) %>% 
+    pivot_longer(
+        where(is.numeric), 
+        names_to = c(".value", "type"),
+        names_sep = "__"
+    )
+
+re_id %>% 
+    ggplot() +
+    aes(
+        x = estimate, 
+        y = reorder(id, estimate), 
+        xmin = q2_5, 
+        xmax = q97_5,
+        fill = type,
+        colour = type
+    ) +
+    facet_wrap(~type) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
     geom_line(aes(group = 1), size = 1) +
-    # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
-    # geom_point(size = 0.5) +
-    geom_vline(xintercept = 0.5, colour = "white") +
+    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
     labs(
@@
         y = "Participant", 
-        colour = "Lang. Prof."
+        colour = "Type"
     ) +
@@
 ```{r re-item, message=FALSE, warning=FALSE, paged.print=FALSE}
-
 re_te <- ranef(model_fit_7)$te[,,1] %>% 
-    as.data.frame() %>% 
+    as_tibble() %>% 
     rownames_to_column("te") %>% 
-    as_tibble() %>% 
-    mutate_if(is.numeric, ~inv_logit_scaled(. + fixef(model_fit_7)[1])) %>% 
-    mutate(te = as.numeric(te)) %>% 
-    clean_names()
-
-ggplot(re_te, aes(estimate, reorder(te, estimate), xmin = q2_5, xmax = q97_5)) +
-    geom_ribbon(aes(group = 1), alpha = 0.5) +
+    clean_names() %>% 
+    mutate(
+        across(
+            .cols = where(is.numeric),
+            .fns = list(
+                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_7)[1]),
+                Production = ~inv_logit_scaled(. + fixef(model_fit_7)[2])
+            ), .names = "{.col}__{.fn}"
+        )
+    ) %>% 
+    select(-c(estimate:q97_5)) %>% 
+    pivot_longer(
+        where(is.numeric), 
+        names_to = c(".value", "type"),
+        names_sep = "__"
+    ) %>% 
+    mutate(te = as.numeric(te))
+
+re_te %>% 
+    ggplot() +
+    aes(
+        x = estimate, 
+        y = reorder(te, estimate), 
+        xmin = q2_5,
+        xmax = q97_5,
+        colour = type,
+        fill = type
+    ) +
+    facet_wrap(~type) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
     geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
     labs(
@@
         y = "TE", 
-        colour = "Lang. Prof."
+        colour = "Type"
     ) +
@@
 
-### Marginal means: Comprehension
-
-```{r emmeans_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+```{r marginal, message=FALSE, warning=FALSE, paged.print=FALSE}
 nd <- expand.grid(
@@
 
-```{r diagnostics_traceplots_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r diagnostics-traceplots, message=FALSE, warning=FALSE, paged.print=FALSE}
 gather_draws(
@@
         xmin = 0,
-        xmax = 500, 
+        xmax = dim(model_fit_7$fit)[1]/2, 
         ymin = -Inf, 
@@
 
+## Gelman-Rubin diagnostic (R-hat)
+
+```{r diagnostics-rhat, message=FALSE, warning=FALSE, paged.print=FALSE}
+mcmc_rhat_data(rhat(model_fit_7)) %>% 
+    ggplot() +
+    aes(x = value, fill = description, colour = description) +
+    geom_histogram(
+        bins = 100, 
+        na.rm = TRUE, 
+        fill = pal_d3()(1)[1], 
+        colour = "white"
+    ) +
+    geom_rug(alpha = 0.1) +
+    labs(
+        x = "R-hat",
+        y = "Number of samples",
+        colour = "Description",
+        fill = "Description"
+    ) +
+    theme(
+        legend.position = "top"
+    )
+```
+
+## Efficient sample size
+
+```{r diagnostics-neff, message=FALSE, warning=FALSE, paged.print=FALSE}
+mcmc_neff_data(neff_ratio(model_fit_7)) %>% 
+    ggplot() +
+    aes(x = value, fill = description, colour = description) +
+    geom_histogram(
+        bins = 100, 
+        na.rm = TRUE, 
+        colour = "white"
+    ) +
+    geom_rug(alpha = 0.01) +
+    geom_vline(xintercept = 1, colour = "black", size = 0.75) +
+    labs(
+        x = "Effective sample size ratio",
+        y = "Number of samples",
+        colour = "Rating",
+        fill = "Rating"
+    ) +
+    theme(
+        legend.position = "top"
+    )
+```
+
 ## {-}
+
+# Appendix {.tabset}
+
+## Session info
+
+```{r}
+sessionInfo()
+```
+
+## {-}

### 2022-08-29 _targets_packages.R

@@ -1,5 +1,6 @@
 # Generated by targets::tar_renv(): do not edit by hand
 library(arrow)
+library(bayesplot)
 library(brms)
 library(bs4Dash)
 library(clustermq)
@@ -24,6 +25,7 @@ library(pingr)
 library(purrr)
 library(readxl)
 library(rlang)
+library(rmarkdown)
 library(rstudioapi)
 library(scales)
 library(shiny)

### 2022-09-07 docs/index.qmd
Report to Quarto

--- 2022-08-29 docs/index.Rmd
@@
 title: "Trajectories"
-authors:
-    - Gonzalo Garcia-Castro^[Universitat Pompeu Fabra, gonzalo.garciadecastro@upf.edu]
-    - Daniela Avila-Varela^[Universitat Pompeu Fabra]
-    - Ignacio Castillejo^[Universidad Autónoma de Madrid]
-    - Núria Sebastian-Galles^[Universitat Pompeu Fabra]
-output:
-    html_document:
-        code_folding: hide
-        number_sections: true
+subtitle: "Lab notes"
+date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
+format:
+    html:
+        abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
+Floccia et al.
+(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
+The mechanisms underlying this effect remain unknown.
+We explore two compatible scenarios.
+First, we test whether cognates are learnt earlier than non-cognates.
+This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
+Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
+This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
+We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
         toc: true
-    github_document:
-        pandoc_args: --webtex
-        number_sections: true
-        toc: true
-bibliography: "../references.bib"
+        smooth-scroll: true
+        link-external-newwindow: true
+        number-sections: true
+        number-offset: 0
+        code-fold: true
+        code-overflow: scroll
+        code-line-numbers: true
+        code-copy: hover
+        fig-dpi: 500
+        reference-location: margin
+        execute-dir: "C:/Users/U155880/Documents/trajectories/"
+bibliography: "references.bib"
+csl: "apa6.csl"
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-    message = TRUE,
-    echo = TRUE,
-    warning = TRUE,
-    include = TRUE,
-    out.width = "80%",
-    dpi = 250
-)
+```{r}
+#| label: setup
+#| include: false
 options(
@@
 )
-
-```
-
-```{r prepare, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```
+
+```{r}
+#| label: prepare
+#| echo: true
+#| message: false
+#| warning: false
 # set params and load target contents
-tar_load_globals()
-tar_load_all()
+targets::tar_load_globals()
+targets::tar_load_everything()
+
+# set custom ggplot2 theme
 theme_set(theme_custom())
 
+# resolve namespace conflicts
 conflict_prefer("filter", "dplyr")
 conflict_prefer("select", "dplyr")
-
 ```
@@
 
-```{r items, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: questionnaire
+#| echo: true
+#| message: false
+#| warning: false
 pool <- multilex::pool %>% 
@@
 
-# Items {.tabset .tab-pills}
-
-```{r items_frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
-Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
-
+# Items
+
+```{r}
+#| label: items-summary
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Lexical frequencies.
+Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
 items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))
@@
 
-```{r items-frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
-
+```{r}
+#| label: items-frequency
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 items %>% 
@@
 
-```{r items-nphon, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
-
+```{r}
+#| label: items-nphon
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 items %>%
@@
 
-```{r items-lv, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
-
+```{r}
+#| label: items-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 items %>% 
@@
 
-```{r items-frequency-phonemes, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: items-frequency-phonemes
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 items %>% 
@@
 
-```{r items-frequency-levenshtein, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: items-frequency-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 items %>% 
@@
 
-```{r items-phonemes-lv, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: items-nphon-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 items %>% 
@@
 
-## {-}
-
-# Participants {.tabset .tab-pills}
-
-```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
+# Participants
+
+```{r}
+#| label: participants-time
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 participants <- left_join(
@@
 
-```{r participants-age, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: participants-age
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 participants %>% 
@@
     aes(x = age, y = n) +
-    geom_col(fill = pal_d3()(3)[3]) +
+    geom_col(fill = pal_d3()(3)[1]) +
     geom_text(aes(label = n), size = 3.5, vjust = -1) +
@@
         panel.grid.minor.x = element_blank()
-
     )
@@
 
-```{r participants-lp, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+```{r}
+#| label: participants-lp
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 participants %>% 
@@
 
-```{r participants_edu, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: participants-edu
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 participants %>%
@@
 
-## {-}
-
-# Data analysis {.tabset .tab-pills}
+# Data analysis
 
@@
 
+:::{.column-body-outset}
 $$\begin{aligned}
@@
 \end{aligned}$$
+:::
 
@@
 
+:::{.column-body-outset}
 $$
@@
 $$
+:::
 
@@
 
-## R code (`brms`) {.tabset}
+## R code (`brms`)
+
+::: {.panel-tabset}
 
@@
 
-### {-}
-
-## Stan code {.tabset}
-
-Stan code generated by `brms::stancode`:
+:::
+
+## Stan code
+
+Stan code generated by `brms::stancode`.
+
+::: {.panel-tabset}
 
@@
 
-### {-}
-
-## {-}
-
-# Results {.tabset .tab-pills}
+:::
+
+# Results
 
@@
 
-```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: results-raw
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 df %>%
@@
 
-```{r loos, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
+```{r}
+#| label: results-loos
+#| echo: true
+#| message: false
+#| warning: false
+#| eval: false
+#| fig-width: 7
 model_loos %>% 
@@
 
-```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: results-fixed-table
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 str_repl <- c(
@@
 
-```{r fix_effects-plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+```{r}
+#| label: results-fixed
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Points and error bars in indicate the posterior means, and 50% and 95% CrI.
 The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
+The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."
 var_repl <- c(
@@
 
-## Random effects {.tabset .tab-pils}
+## Random effects
 
@@
 
-```{r random-participant, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: results-random-id
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 re_id <- ranef(model_fit_7)$id[,,1] %>% 
@@
 
-```{r re-item, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: results-random-te
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 re_te <- ranef(model_fit_7)$te[,,1] %>% 
@@
 
-### {-}
-
 ## Marginal means
 
-```{r marginal, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: results-marginal
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 nd <- expand.grid(
@@
     )
-
-```
+```
+
+## Model diagnostics
 
@@
 
-```{r diagnostics-traceplots, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: diagnostics-traceplots
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 gather_draws(
@@
 
-## Gelman-Rubin diagnostic (R-hat)
-
-```{r diagnostics-rhat, message=FALSE, warning=FALSE, paged.print=FALSE}
+### Gelman-Rubin diagnostic (R-hat)
+
+```{r}
+#| label: diagnostics-rhat
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 mcmc_rhat_data(rhat(model_fit_7)) %>% 
@@
 
-## Efficient sample size
-
-```{r diagnostics-neff, message=FALSE, warning=FALSE, paged.print=FALSE}
+### Effective sample size
+
+```{r}
+#| label: diagnostics-neff
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
 mcmc_neff_data(neff_ratio(model_fit_7)) %>% 
@@
 
-## {-}
-
-# Appendix {.tabset}
+# Appendix
 
@@
 ```
-
-## {-}

### 2022-09-07 index.qmd
Report to Quarto

--- 2022-09-07 docs/index.qmd
@@
         execute-dir: "C:/Users/U155880/Documents/trajectories/"
-bibliography: "references.bib"
-csl: "apa6.csl"
+bibliography: "manuscript/references.bib"
+csl: "manuscript/apa6.csl"
 ---
@@
 # set params and load target contents
-targets::tar_load_globals()
-targets::tar_load_everything()
+
+library(bayesplot)
+library(brms) 
+library(conflicted)
+library(dplyr) 
+library(forcats)
+library(ggplot2) 
+library(ggsci)
+library(gt) 
+library(here) 
+library(janitor)
+library(knitr)
+library(lubridate)
+library(multilex)
+library(patchwork) 
+library(papaja) 
+library(purrr)
+library(scales) 
+library(stringdist)
+library(stringr)
+library(targets)
+library(tibble)
+library(tidybayes) 
+library(tidyr)
+
+# load R objects
+tar_load_everything()
 
@@
 The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*.
+Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* \[dog\] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* \[*to finish*\] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* \[*monkey*\] and Catalan *mono/mico*.
 We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
@@
 For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
+Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\~400 translation equivalents, \~400 items in Catalan, \~400 items in Spanish).
 These last families were randomly allocated into one of four different subsets of the complete list of items.
@@
 Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* \[cereal bar\]) and items that included more than one word-form (e.g., *mono / mico*).
 Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
@@
 
-## # Phonemes
+## \# Phonemes
 
@@
 
-## Frequency by # Phonemes
+## Frequency by \# Phonemes
 
@@
 
-## # Phonemes by Levenshtein
+## \# Phonemes by Levenshtein
 
@@
 
+```{r}
+#| label: participants-edu-plot
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+participants %>%
+    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
+    mutate(
+        doe_2 = pmin(doe_catalan, doe_spanish),
+        doe_2 = cut(doe_2, seq(0, 0.5, 0.1), include.lowest = TRUE)
+    ) %>% 
+    count(doe_2, edu_parent) %>% 
+    ggplot() +
+    aes(
+        x = doe_2,
+        y = n,
+        fill = edu_parent
+    ) +
+    geom_col(position = position_fill()) +
+    scale_fill_manual(values = sort(pal_d3()(6), decreasing = TRUE)) +
+    labs(
+        x = "Exposure to second language (L2 DoE)",
+        y = "Proportion of participants",
+        fill = "Highest parental educational attainment"
+    ) +
+    scale_y_continuous(labels = percent) +
+    theme(
+        legend.position = "none",
+        panel.grid = element_blank()
+    )
+```
+
 # Data analysis
@@
 
-:::{.column-body-outset}
+::: column-body-outset
 $$\begin{aligned}
@@
 
-:::{.column-body-outset}
+::: column-body-outset
+
 $$
-\begin{aligned}
-
-\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} +
+\begin{aligned}\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} +
 \beta_6\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_7\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_8\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + 
@@
 \theta_{7j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}) +
-\theta_{8j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j})
-
-\end{aligned}
-$$
+\theta_{8j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j})\end{aligned}$$
 :::
@@
 
-:::{.column-body-outset}
-$$\begin{aligned}
-\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
+::: column-body-outset
+$$\begin{aligned}\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
 \beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
 \beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
-\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]
-\end{aligned}
-$$
+\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]\end{aligned}$$
 :::
@@
 
-* OS: `r sessionInfo()$running`, `r `sessionInfo()$platform`
-* R version: `r sessionInfo()$R.version$version.string`
-* Algorithm: Hamiltonian Montecarlo (No U-turn Sampler, *NUTS*)
-* Engine: Stan (brms interface, `r sessionInfo()$otherPkgs$brms$Version`)
-* brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
-* Chains: `r dim(model_fit_7$fit)[2]`
-* Cores: `r dim(model_fit_7$fit)[2]`
-* Iterations: `r dim(model_fit_7$fit)[1]` (`r dim(model_fit_7$fit)[1]/2`)
+-   OS: `r sessionInfo()$running`, `r`sessionInfo()\$platform\`
+-   R version: `r sessionInfo()$R.version$version.string`
+-   Algorithm: Hamiltonian Montecarlo (No U-turn Sampler, *NUTS*)
+-   Engine: Stan (brms interface, `r sessionInfo()$otherPkgs$brms$Version`)
+-   brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
+-   Chains: `r dim(model_fit_7$fit)[2]`
+-   Cores: `r dim(model_fit_7$fit)[2]`
+-   Iterations: `r dim(model_fit_7$fit)[1]` (`r dim(model_fit_7$fit)[1]/2`)
 
@@
 
-::: {.panel-tabset}
-
+::: panel-tabset
 ### Model 0
@@
 ```
-
 :::
@@
 }
-
 ```
@@
 #| fig-width: 7
+#| fig-height: 4
 df %>%
@@
         legend.position = "top",
-        axis.text.y = element_blank(),
+        axis.text.y = element_text(size = 7),
+        axis.text.x = element_text(size = 7),
         axis.ticks.y = element_blank(),
-        panel.grid.minor.x = element_blank(),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
+        panel.grid = element_blank()
     )
@@
 #| warning: false
-#| fig-width: 7
+#| fig-width: 12
+#| fig-height: 12
 gather_draws(
@@
         legend.title = element_blank(),
-        panel.grid = element_blank()
+        panel.grid = element_blank(),
+        axis.text.x = element_text(size = 7)
     )
@@
 #| fig-width: 7
-mcmc_rhat_data(rhat(model_fit_7)) %>% 
+#| eval: false
+mcmc_rhat_data(model_rhats$model_fit_7) %>% 
     ggplot() +
@@
 #| fig-width: 7
-mcmc_neff_data(neff_ratio(model_fit_7)) %>% 
+#| eval: false
+mcmc_neff_data(model_neffs$model_fit_7) %>% 
     ggplot() +

### 2022-09-07 index.qmd
Report to Quarto

@@
 
-$$
-\begin{aligned}\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} +
+$$\begin{aligned}\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} +
 \beta_6\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_7\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_8\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + 
@@
 #| warning: false
-#| fig-width: 7
-#| fig-height: 4
+#| fig-width: 15
+#| fig-height: 10
 df %>%

### 2022-09-13 index.qmd
Update report

--- 2022-09-07 index.qmd
@@
 date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
-format:
-    html:
+authors:
+  - name: Gonzalo Garcia-Castro
+    orcid: 0000-0002-8553-4209
+    email: gonzalo.garciadecastro@upf.edu
+    corresponding: true
+    affiliations:
+      - name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+        address: Ramon Trias fargas 25-27
+        city: Barcelona
+        state: Spain
+        postal-code: 08005
+  - name: Daniela S. Ávila-Varela
+    orcid: 0000-0002-3518-8117
+    affiliations:
+      - name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+  - name: Ignacio Castillejo
+    orcid: 0000-0001-7445-0416
+    affiliations:
+      - name: Universidad Atónoma de Madrid
+        department: Departamento de Psicología
+  - name: Nuria Sebastian-Galles
+    orcid: 0000-0001-6938-2498
+    affiliations:
+      - name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
 abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
@@
 We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
+format:
+    html:
         toc: true
@@
 library(ggplot2) 
+library(marginaleffects)
 library(ggsci)
@@
 
-We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_std` by `item`, and their correlation parameter.
-We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope.
-Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and random slopes for `cognate` by `id`.
+## Model comparison
+
+We initially fitted a base model (`model_fit_0`) that only included the predictors $Age$ (`age_std`), $Frequency$ (`freq_std`), $Phonemes$ (`n_phon_std`), and $DoE$ (`doe_std`) as nuisance parameters, along with random intercepts and slopes of all predictors by $Participant$ (`id`) and $Item$ (`te`), and all correlation parameters.
+We then expanded this model by adding the main effect of $Levenshtein$ (`lv_std`) (`model_fit_1`) and its random slope by $Participant$.
+Then, we added the $DoE \times Levenshtein$ (`doe_std:lv_std`) two-way interaction (`model_fit_2`) and its random slope by $Participant$.
+Finally, we added the $Age \times Levenshtein$ (`age_std:doe_std:lv_std`) three-way interaction with its random slope by $Participant$, along with a random slope for the $Age \times DoE$ (`age_std:doe_std`) two-way interaction.
 
@@
 
-$$
-\begin{aligned}\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} +
-\beta_6\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_7\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_8\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + 
-\beta_9{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) +\\
-
-&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{TE})} \mathrm{Freq}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{5j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) +\\
-
-&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{ID})} \mathrm{Freq}_{j} +   \theta_{3j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{5j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{6j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) +
-\theta_{7j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}) +
-\theta_{8j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j})\end{aligned}$$
+$$\begin{aligned}
+\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} + & [\mbox{Main effects}]\\
+&\beta_6\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_7\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_8\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Two-way interactions}]\\
+&\beta_9{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Three-way interaction}]\\
+
+&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{TE})} \mathrm{Freq}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{5j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) + & [\mbox{Item-level effects}]\\
+
+&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{ID})} \mathrm{Freq}_{j} +   \theta_{3j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{5j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{6j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) & [\mbox{Participant-level effects}]
+\end{aligned}$$
 :::
@@
 
-In addition, we specified a $LKJ$ prior ($\sim LKJ(2)$ for the correlations between group-level coefficients. We implemented the model prior following {brms} syntax as:
+In addition, we specified a $LKJ$ prior--$\sim LKJ(2)$ for the correlations between group-level coefficients. We implemented the model prior following {brms} syntax as:
 
@@
 
+### Prior predictions
+
+```{r}
+#| label: prior-predictions
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+nd <- expand.grid(
+    n_phon_std = 0,
+    freq_std = 0,
+    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    doe_std = 0,
+    lv_std = 0
+)
+
+m <- add_epred_draws(nd, model_fit_3_prior, ndraws = NULL, re_formula = NA) %>% 
+    filter(.category != "No") %>% 
+    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
+    mutate(Understands = Understands + `Understands and Says`) %>% 
+    pivot_longer(
+        c(Understands, `Understands and Says`),
+        names_to = ".category",
+        values_to = ".epred"
+    ) %>% 
+    mutate(
+        .category = factor(
+            .category,
+            levels = c("Understands and Says", "Understands"),
+            labels = c("Production\n(Understands and Says))", "Comprehension\n(Understands)")
+        )
+    )
+
+m %>% 
+    ggplot() +
+    aes(
+        x = age_std, 
+        y = .epred, 
+    ) +
+    facet_wrap(~.category) +
+    # geom_line(
+    #     aes(group = interaction(.draw)),
+    #     alpha = 0.05,
+    #     size = 1
+    # ) +
+    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
+    stat_summary(fun = "mean", geom = "line", size = 0.75) +
+    scale_x_continuous(
+        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 36, 4),
+        sec.axis = dup_axis(
+            name = "Age (years; months)",
+            labels = months_to_years(
+                rescale_variable(
+                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+                    mean(df$age),
+                    sd(df$age)
+                )
+            )
+        )
+    ) +
+    scale_fill_brewer(
+        palette = "YlOrRd",
+        labels = c("95%", "89%", "67%", "50%")
+    ) +
+    labs(
+        x = "Age (months)", 
+        y = "Probability of acquisition\n(Linear prediction)",
+        colour = "Credible interval", 
+        fill = "Credible interval"
+    ) +
+    scale_y_continuous(labels = percent, limits = c(0, 1)) +
+    theme(
+        legend.position = "top",
+        axis.ticks.x = element_line(colour = "black"),
+        panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
+        strip.background = element_rect(fill = "grey", colour = NA),
+        plot.background = element_rect(fill = "white", colour = NA)
+    )
+```
+
 ## Model settings
 
--   OS: `r sessionInfo()$running`, `r`sessionInfo()\$platform\`
+-   OS: `r sessionInfo()$running`, `r sessionInfo()$platform`
 -   R version: `r sessionInfo()$R.version$version.string`
@@
 -   Cores: `r dim(model_fit_7$fit)[2]`
--   Iterations: `r dim(model_fit_7$fit)[1]` (`r dim(model_fit_7$fit)[1]/2`)
+-   Iterations: `r format(dim(model_fit_7$fit)[1], big.mark = ",")` (`r format(dim(model_fit_7$fit)[1]/2, big.mark = ",")`)
 
@@
 ::: panel-tabset
+
 ### Model 0
@@
 ```
-
-### Model 4
-
-```{r}
-print(model_fit_4$formula)
-```
-
-### Model 5
-
-```{r}
-print(model_fit_5$formula)
-```
-
-### Model 6
-
-```{r}
-print(model_fit_6$formula)
-```
-
-### Model 7
-
-```{r}
-print(model_fit_7$formula)
-```
 :::
@@
 ::: panel-tabset
+
 ### Model 0
-
-```
-// generated with brms 2.16.1
-functions {
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-}
-parameters {
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-vector[N_1] z_1[M_1];  // standardized group-level effects
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-vector[N_2] z_2[M_2];  // standardized group-level effects
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-vector[N_1] r_1_1;  // actual group-level effects
-vector[N_2] r_2_1;  // actual group-level effects
-r_1_1 = (sd_1[1] * (z_1[1]));
-r_2_1 = (sd_2[1] * (z_2[1]));
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = rep_vector(0.0, N);
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_2_1[J_2[n]] * Z_2_1[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 1 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(z_1[1]);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 1 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(z_2[1]);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept;
-// additionally sample draws from priors
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_sd_2 = normal_rng(1,0.1);
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 1
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 2 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 2 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 2
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 3 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 2 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 3
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 4 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 4 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 4
 
@@
 
-### Model 5
+### Model 1
 
@@
 
-### Model 6
+### Model 2
 
@@
 
-### Model 7
+### Model 3
 
@@
 #| fig-width: 7
-#| fig-height: 4
 df %>%
     mutate(
-        lv = cut_interval(lv, 5),
-        age = floor(age)
-    ) %>% 
-    count(lv, age, response) %>%
-    mutate(prop = n/sum(.$n)) %>% 
-    ggplot(aes(age, prop, colour = response, fill = response)) +
-    facet_wrap(~lv, nrow = 1) +
-    geom_col(position = position_fill()) +
-    geom_hline(yintercept = 0.5, colour = "black", size = 0.5) +
+        lv = cut_interval(lv*100, 5) %>% 
+            gsub("([0-9.]+)", "\\1%", .) %>% 
+            factor(levels = c("[0%,20%]", "(20%,40%]", "(40%,60%]", "(60%,80%]", "(80%,100%]")),
+        doe = cut_interval(doe*100, 3) %>% 
+            gsub("([0-9.]+)", "\\1%", .) %>% 
+            factor(levels = c("[0%,33.3%]", "(33.3%,66.7%]", "(66.7%,100%]")) %>% 
+            paste0("Degree of Exposure: ", .) %>% 
+            factor(
+                levels = paste0(
+                    "Degree of Exposure: ", 
+                    c("[0%,33.3%]", "(33.3%,66.7%]", "(66.7%,100%]")
+                )
+            ),
+        age = floor(age),
+        Understands = response %in% c("Understands", "Understands and Says"),
+        `Understands and Says` = response %in% c("Understands and Says")
+    ) %>% 
+    group_by(te, lv, age, doe) %>%
+    summarise(
+        n = n(),
+        Understands = mean(Understands),
+        `Understands and Says` = mean(`Understands and Says`),
+        .groups = "drop"
+    ) %>% 
+    pivot_longer(
+        c(Understands, `Understands and Says`), 
+        names_to = ".category", 
+        values_to = "prop"
+    ) %>% 
+    group_by(lv, age, doe, .category) %>%
+    summarise(
+        n = sum(n),
+        prop = mean(prop),
+        .groups = "drop"
+    ) %>% 
+    mutate(
+        prop_se = sd(prop)/sqrt(n),
+        prop_lower = prop-prop_se,
+        prop_upper = prop+prop_se,
+        prop_lower = ifelse(prop_lower < 0, 0, prop_lower),
+        prop_upper = ifelse(prop_upper >1, 1, prop_upper)
+    ) %>% 
+    ggplot() +
+    aes(age, prop, colour = lv, fill = lv, shape = lv) +
+    facet_grid(.category~doe) +
+    geom_hline(yintercept = 0.5, colour = "grey", size = 0.5) +
+    geom_line(size = 1, alpha = 0.75) +
+    # geom_point(size = 2) +
     labs(
         x = "Age (months)",
-        y = "% responses", 
-        colour = "Response",
-        fill = "Response"
-    ) +
-    scale_fill_d3() +
-    scale_colour_d3() +
-    scale_x_continuous(breaks = seq(0, 40, 2)) +
-    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+        y = "Proportion of responses", 
+        colour = "Phonological similarity (Levenshtein)",
+        fill = "Phonological similarity (Levenshtein)",
+        shape = "Phonological similarity (Levenshtein)",
+        linetype = "Phonological similarity (Levenshtein)"
+    ) +
+    scale_color_brewer() +
+    # scale_colour_manual(values = wesanderson::wes_palette("Zissou1", n = 5, type = "continuous")[1:5]) +
+    scale_x_continuous(breaks = seq(0, 40, 4)) +
+    scale_y_continuous(limits = c(0, 1), labels = percent) +
     theme(
         legend.position = "top",
-        axis.text.y = element_text(size = 7),
-        axis.text.x = element_text(size = 7),
         axis.ticks.y = element_blank(),
-        panel.grid = element_blank()
-    )
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
+        plot.background = element_rect(fill = "white", colour = NA),
+        strip.background = element_rect(fill = "grey", colour = "grey")
+    )
+
+ggsave("plot.png", height = 6, width = 9)
 ```
@@
 model_loos %>% 
-    as_tibble() %>% 
+    loo_compare() %>% 
+    as.data.frame() %>%   
     rownames_to_column("model") %>% 
@@
     ) %>%
-    mutate_all(as.numeric) %>% 
     gt() %>% 
@@
     tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
-    fmt_number(3:13) %>% 
+    fmt_number(2:9) %>% 
     cols_label(
@@
         se_elpd_loo = md("*SE*"),
-        subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
         # p
@@
         se_p_loo = md("SE"),
-        subsampling_se_p_loo = md("SE<sub>sub</sub>"),
         # looic
@@
         se_looic = md("*SE*"),
-        subsampling_se_looic = md("SE<sub>sub</sub>"),
         # diff
         elpd_diff = md("*diff*"),
-        se_diff = md("*SE*"),
-        subsampling_se_diff = md("*SE<sub>sub</sub>*"),
-    ) 
-```
+        se_diff = md("*SE*")
+    ) %>% 
+    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)"))
+```
+
+### Leave-one-out diagnostics (Pareto-*k* diagnostics)
+
+::: panel-tabset
+
+#### Model 0
+
+```{r}
+#| label: loo-diagnostics-0
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+get_loo_plot <- function(x){
+    x$pointwise[,5] %>% 
+        data.frame(x = 1:length(.), y = .) %>% 
+        ggplot(aes(x, y)) +
+        geom_point(
+            size = 0.5, alpha = 0.1, colour = pal_d3()(3)[2], 
+            shape = 1
+        ) +
+        geom_hline(yintercept = 0.5, linetype = "dotted") +
+        labs(
+            x = "Data point",
+            y = "Pareto-k value",
+            title = "PSIS diagnostic plot"
+        ) +
+        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
+        scale_y_continuous(
+            limits = c(-0.6, 0.6)
+        )
+}
+
+get_loo_plot(model_loos$model_fit_0)
+```
+
+#### Model 1
+
+```{r}
+#| label: loo-diagnostics-1
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+get_loo_plot(model_loos$model_fit_1)
+```
+
+#### Model 2
+
+```{r}
+#| label: loo-diagnostics-2
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+get_loo_plot(model_loos$model_fit_2)
+```
+
+#### Model 3
+
+```{r}
+#| label: loo-diagnostics-3
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+get_loo_plot(model_loos$model_fit_3)
+```
+
+:::
 
@@
 str_repl <- c(
+    "Intercept\\[1\\]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
+    "Intercept\\[2\\]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
     "age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein",
@@
     "doe_std:lv_std" = "DoE \u00d7 Levenshtein",
-    "Intercept" = "Intercept",
-    "age_std" = "Age (+1 month)",
-    "freq_std" = "Frequency (+1 SD)",
-    "n_phon_std" = "# Phonemes (+1 SD)",
-    "doe_std" = "DoE (+10%)",
-    "lv_std" = "Levenshtein (+1 SD)"
-)
-
-draws_fix <- fixef(model_fit_7) %>% 
+    "age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
+    "freq_std" = paste0("Frequency (+1 SD, ",  round(sd(df$freq), 2), " Zipf)"),
+    "n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
+    "doe_std" = paste0("DoE (+1 SD, ",  percent(sd(df$doe)), ")"),
+    "lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")")
+)
+
+intercept_footnote <- paste0(
+    "Probability of comprehension or production at the average age of the sample (", round(mean(df$age), 2), " months), transformed to the probability scale using the inverse logit."
+)
+
+slopes_footnote <- "Transformed using the divide-by-four rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."
+
+draws_fix <- fixef(model_fit_3) %>% 
     as.data.frame() %>%
@@
     mutate(
-        estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
-        q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
-        q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
-        term = str_replace_all(term, str_repl)
-    )
-
-gt(draws_fix) %>% 
-    fmt_percent(2) %>% 
-    fmt_number(3:4, decimals = 2) %>% 
-    cols_merge(vars("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
+        estimate_p = ifelse(grepl("Intercept", term), plogis(estimate), estimate/4),
+        q2_5_p = ifelse(grepl("Intercept", term), plogis(q2_5), q2_5/4),
+        q97_5_p = ifelse(grepl("Intercept", term), plogis(q97_5), q97_5/4),
+        term = str_replace_all(term, str_repl),
+        group = ifelse(grepl("Intercept", term), "Intercepts", "Slopes")
+    )
+
+gt(draws_fix, rowname_col = "term") %>% 
+    fmt_number(2:7, decimals = 2) %>% 
+    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
+    cols_merge(c("q2_5_p", "q97_5_p"), pattern = "[{1}, {2}]") %>% 
+    tab_spanner("Logit scale", c(estimate, q2_5, q97_5)) %>% 
+    tab_spanner("Probability scale", c(estimate_p, q2_5_p, q97_5_p)) %>% 
+    tab_row_group(label = "Intercepts", rows = 1:2) %>% 
+    tab_row_group(label = "Slopes", rows = 3:nrow(draws_fix)) %>% 
+    row_group_order(groups = c("Intercepts", "Slopes")) %>% 
     cols_label(
         term = md("**Predictor**"),
-        estimate = md("**Mean**"),
-        q2_5 = md("**95\\% CrI**"),
+        estimate = "Mean",
+        q2_5 = md("95\\% CrI"),
+        estimate_p = "Mean",
+        q2_5_p = md("95\\% CrI"),
     ) %>% 
     tab_footnote(
-        footnote = "Transformed using the inverse logit to get the average probability of correct response",
-        locations = cells_body(columns = "term", rows = term=="Intercept")
+        footnote = intercept_footnote,
+        locations = cells_body(
+            columns = "term", 
+            rows = grepl("Intercept", term)
+        )
     ) %>% 
     tab_footnote(
-        footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-        locations = cells_body(columns = "term", rows = term %in% c("Age (+1 month)", "Frequency (+ 1 SD)",  "DoE (+10%)", "Levenshtein (+ 1 SD)", "doe_std:lv_std" = "DoE \u00d7 Levenshtein"))
+        footnote = slopes_footnote,
+        locations = cells_body(
+            columns = "term", 
+            rows = term %in% str_repl[-c(1, 2)]
+        )
     ) %>% 
     cols_align(
-        align = c("center"),
-        columns = 2:3
-    )
+        align = c("right"),
+        columns = 2:7
+    ) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_spanners(everything())
+    ) %>% 
+    tab_style(
+        list(cell_text(style = "italic"), cell_fill(color = "grey")),
+        cells_row_groups()
+    ) %>% 
+    cols_hide(group)
 ```
@@
 The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."
-var_repl <- c(
-    "b_Intercept[1]" = "Intercept (Comprehension)",
-    "b_Intercept[2]" = "Intercept (Production)",
-    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein", 
+str_repl <- c(
+    "b_Intercept\\[1\\]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
+    "b_Intercept\\[2\\]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
+    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
+    "b_freq_std" = paste0("Frequency (+1 SD, ",  round(sd(df$freq), 2), " Zipf)"),
+    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
+    "b_doe_std" = paste0("DoE (+1 SD, ",  percent(sd(df$doe)), ")"),
+    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
+    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein",
+    "b_age_std:doe_std" = "Age \u00d7 DoE",
     "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:doe_std" = "Age \u00d7 DoE",
-    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein", 
-    "b_Intercept" = "Intercept", 
-    "b_age_std" = "Age (+1 month)", 
-    "b_freq_std" = "Frequency (+1 SD)",
-    "b_n_phon_std" = "# Phonemes (+1 SD)",
-    "b_doe_std" = "DoE (+1 SD)", 
-    "b_lv_std" = "Levenshtein (+1 SD)"
-)
-
-post <- gather_draws(model_fit_7, `b_.*`, regex = TRUE) %>% 
+    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein"
+)
+
+post <- gather_draws(model_fit_3, `b_.*`, regex = TRUE) %>% 
     mutate(
@@
             .variable, 
-            levels = names(var_repl),
-            labels = var_repl
-        )
-    )
-
-post_intercept <- post %>% 
-    filter(.variable %in% c("b_Intercept[1]", "b_Intercept[2]")) %>% 
-    ggplot(aes(.value, .variable_name)) +
-    geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
+            levels = names(str_repl),
+            labels = str_repl,
+            ordered = TRUE
+        )
+    )
+
+post %>%
+    filter(!grepl("Intercept|sd", .variable)) %>%
+    ggplot(aes(.value, fct_rev(.variable_name))) +
+    geom_vline(xintercept = 0, size = 0.5, colour = "grey") +
     stat_slab(
-        aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
         size = 0.5,
-        color = "#FF0000",
-        fill = "#FF0000",
-        position = position_nudge(y = 0.2)
-    ) +
-    stat_pointinterval(point_size = 2) +
+        color = pal_d3()(2)[2],
+        fill = pal_d3()(2)[2],
+        position = position_nudge(y = 0.1)
+    ) +
+    geom_errorbar(
+        data = post %>% 
+            filter(!grepl("Intercept|sd", .variable)) %>%
+            mean_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>% 
+            mutate(
+                .variable_name = factor(
+                    .variable, 
+                    levels = names(str_repl),
+                    labels = str_repl,
+                    ordered = TRUE
+                )
+            ),
+        aes(xmin = .lower, xmax = .upper, x = .value),
+        width = 0.15
+    ) +
+    geom_text(
+        data = post %>% 
+            filter(!grepl("Intercept|sd", .variable)) %>%
+            mean_hdi(
+                .exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")
+            ) %>% 
+            mutate(
+                .variable_name = factor(
+                    .variable, 
+                    levels = names(str_repl),
+                    labels = str_repl,
+                    ordered = TRUE
+                )
+            ),
+        aes(
+            label = paste0(
+                percent(.value, accuracy = 0.1),
+                " [", percent(.lower, accuracy = 0.1), ", ",
+                percent(.upper, accuracy = 0.1), "]"
+            )
+        ),
+        position = position_nudge(y = -0.2),
+        size = 3
+    ) +
     labs(
@@
     ) +
-    scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-    scale_x_continuous(labels = percent, limits = c(0, 1))
-
-post_fix <- post %>% 
-    filter(!grepl("Intercept|sd", .variable)) %>% 
-    ggplot(aes(.value, .variable_name)) +
-    geom_vline(xintercept = 0, size = 0.5, linetype = "dashed") +
-    stat_slab(
-        aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
-        size = 0.5,
-        color = "#00A08A",
-        fill = "#00A08A",
-        position = position_nudge(y = 0.1)
-    ) +
-    stat_pointinterval(point_size = 2) +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "Variable", 
-        fill = "Variable",
-        alpha = "CrI"
-    ) +
-    scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-    scale_x_continuous(labels = percent, breaks = seq(-0.2, 2.5, 0.1))
-
-(post_intercept / post_fix) +
-    plot_layout(heights = c(0.2, 0.8)) &
+    scale_x_continuous(labels = percent, breaks = seq(-0.2, 2.5, 0.1)) +
     theme(
-        legend.position = "top",
-        legend.direction = "horizontal",
-        legend.justification = "right",
-        plot.caption.position = "plot",
-        plot.caption = element_text(hjust = 0),
         axis.title.y = element_blank(),
-        # axis.text = element_text(size = 7),
         panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
         panel.grid.major.y = element_blank()
+
     )
@@
 #| fig-width: 7
-re_id <- ranef(model_fit_7)$id[,,1] %>% 
+re_id <- ranef(model_fit_3)$id[,,1] %>% 
     as_tibble() %>% 
@@
 #| fig-width: 7
-re_te <- ranef(model_fit_7)$te[,,1] %>% 
+re_te <- ranef(model_fit_3)$te[,,1] %>% 
     as_tibble() %>% 
@@
 #| fig-width: 7
-nd <- expand.grid(
+nd <- datagrid(
+    model = model_fit_3,
     n_phon_std = 0,
     freq_std = 0,
-    age_std = seq(-4, 4, length.out = 30),
-    doe_std = range(df$doe_std),
-    lv_std = c(range(df$lv_std), 0)
-)
-
-m <- add_epred_draws(nd, model_fit_7, ndraws = 50, re_formula = NA) %>% 
+    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe)),
+    lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
+)
+
+m <- add_epred_draws(nd, model_fit_3, ndraws = NULL, re_formula = NA) %>% 
     mutate(
@@
             levels = unique(nd$doe_std), 
-            labels = c("DoE: 0%", "DoE: 100%")
+            labels = paste0("DoE: ", c(10, 50, 90), "%")
         ),
@@
             lv_std, 
-            levels = unique(nd$lv_std)[c(1, 3, 2)],
-            labels = c("0%", "50%", "100%")
+            levels = unique(nd$lv_std),
+            labels = paste0(c(0, 50, 100), "%")
         )
@@
         values_to = ".epred"
-    )
-
-ggplot(m, aes(age_std, .epred, colour = lv_std, fill = lv_std)) +
+    ) %>% 
+    mutate(
+        .category = str_replace_all(
+            .category,
+            c(
+                "Understands and Says" = "Production\n(Understands and Says))",
+                "Understands" = "Comprehension\n(Understands)"
+            )
+        )
+    )
+
+d <- df %>%
+    mutate(
+        doe_std = case_when(
+            between(doe, 0.00, 0.20) ~ "DoE: 10%",
+            between(doe, 0.40, 0.60) ~ "DoE: 50%",
+            between(doe, 0.80, 1.00) ~ "DoE: 90%"
+        ) %>% 
+            factor(., levels = paste0("DoE: ", percent(c(0.1, 0.5, 0.9)))),
+        lv_std = case_when(
+            between(lv, 0.00, 0.20) ~ "0%",
+            between(lv, 0.40, 0.60) ~ "50%",
+            between(lv, 0.80, 1.00) ~ "100%"
+        ) %>% 
+            factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
+        age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
+            cut(
+                .,
+                breaks = ((seq(10, 38, 4)-mean(df$age))/(sd(df$age))),
+                labels = ((seq(10, 34, 4)-mean(df$age))/(sd(df$age))),
+                include.lowest = TRUE
+            ) %>% 
+            as.character() %>% 
+            as.numeric(),
+        Understands = response %in% c("Understands", "Understands and Says"),
+        `Understands and Says` = response %in% "Understands and Says"
+    ) %>% 
+    drop_na(doe_std, lv_std) %>% 
+    group_by(te, age_std, doe_std, lv_std) %>% 
+    summarise(
+        Understands = mean(Understands),
+        `Understands and Says` = mean(`Understands and Says`),
+        .groups = "drop"
+    ) %>% 
+    pivot_longer(
+        c(Understands, `Understands and Says`),
+        names_to = ".category",
+        values_to = ".epred"
+    ) %>% 
+    mutate(
+        .category = str_replace_all(
+            .category,
+            c(
+                "Understands and Says" = "Production\n(Understands and Says))",
+                "Understands" = "Comprehension\n(Understands)"
+            )
+        )
+    )
+# group_by(age_std, doe_std, lv_std, .category) %>% 
+# summarise(
+#     .epred_mean = mean(.epred),
+#     .epred_sd = sd(.epred),
+#     n = n(),
+#     .groups = "drop"
+# ) %>% 
+# mutate(
+#     .epred_se = .epred_sd/n,
+#     .epred_lower = .epred_mean-.epred_se,
+#     .epred_upper = .epred_mean+.epred_se,
+#     .category = str_replace_all(
+#         .category, 
+#         c(
+#             "Understands" = "Comprehension\n(Understands)",
+#             "Understands and Says" = "Production\n(Understands and Says))"
+#         )
+#     )
+# ) %>% 
+# select(.category, age_std, doe_std, lv_std, .epred = .epred_mean, .epred_lower, .epred_upper, n)
+
+m %>% 
+    ggplot() +
+    aes(
+        x = age_std, 
+        y = .epred, 
+        colour = lv_std, 
+        fill = lv_std,
+    ) +
     facet_grid(.category~doe_std) +
-    geom_line(
-        aes(
-            group = interaction(.draw, lv_std, doe_std), 
-            colour = lv_std),
-        alpha = 0.1, 
-        size = 1
-    ) +
-    # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
+    # geom_line(
+    #     aes(group = interaction(.draw, lv_std, doe_std), colour = lv_std),
+    #     alpha = 0.05, 
+    #     size = 1
+    # ) +
+    # stat_lineribbon(size = 1, alpha = 0.5, .width = 0.95, colour = NA) +
+    geom_point(
+        data = d,
+        position = position_dodge(width = 0.25),
+        size = 0.5,
+        alpha = 0.5
+    ) +
     stat_summary(fun = "mean", geom = "line", size = 1) +
-    geom_hline(yintercept = 0.5, size = 0.5, linetype = "dashed") +
+    # geom_line(
+    #     data = d,
+    #     position = position_dodge(width = 0.25)
+    # ) +
+    # geom_errorbar(
+    #     data = d,
+    #     aes(
+    #         ymin = .epred_lower,
+    #         ymax = .epred_upper
+    #     ),
+    #     width = 0,
+#     position = position_dodge(width = 0.25)
+# ) +
+
+scale_x_continuous(
+    breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+    labels = seq(10, 36, 4)
+    # sec.axis = dup_axis(
+    #     name = "Age (years; months)",
+    #     labels = months_to_years(
+    #         rescale_variable(
+    #             (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+    #             mean(df$age), 
+    #             sd(df$age)
+    #         )
+    #     )
+    # )
+) +
     scale_color_d3() +
@@
     theme(
-        legend.position = "top"
-    )
+        legend.position = "top",
+        axis.ticks.x = element_line(colour = "black"),
+        panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
+        strip.background = element_rect(fill = "grey", colour = NA),
+        plot.background = element_rect(fill = "white", colour = NA)
+    )
+
+ggsave("plot.png", height = 9, width = 12)
 ```
@@
 gather_draws(
-    model_fit_7, 
+    model_fit_3, 
     `b_.*`, `sd_te__.*`, `cor_te__.*`, 
@@
         xmin = 0,
-        xmax = dim(model_fit_7$fit)[1]/2, 
+        xmax = dim(model_fit_3$fit)[1]/2, 
         ymin = -Inf, 
@@
 #| eval: false
-mcmc_rhat_data(model_rhats$model_fit_7) %>% 
+mcmc_rhat_data(model_rhats$model_fit_3) %>% 
     ggplot() +
-    aes(x = value, fill = description, colour = description) +
+    aes(x = value) +
     geom_histogram(
@@
     ) +
-    geom_rug(alpha = 0.1) +
+    geom_rug(alpha = 0.01, colour = pal_d3()(1)[1]) +
     labs(
@@
 #| eval: false
-mcmc_neff_data(model_neffs$model_fit_7) %>% 
+mcmc_neff_data(model_neffs$model_fit_3) %>% 
     ggplot() +

### 2022-09-13 manuscript/manuscript.Rmd
Update report

@@
 
-bibliography      : ["trajectories.bib"]
+bibliography      : ["references.bib"]
 
@@
     out.width = "100%",
-  dpi = 300,
+    dpi = 800,
     results = "asis"
@@
 # load packages
+library(bayesplot)
+library(brms) 
+library(conflicted)
+library(dplyr) 
+library(forcats)
+library(ggplot2) 
+library(ggsci)
+library(gt) 
+library(here) 
+library(janitor)
+library(knitr)
+library(lubridate)
+library(multilex)
+library(patchwork) 
 library(papaja) 
-library(knitr)
-library(here)
-library(dplyr)
+library(purrr)
+library(quarto)
+library(rmarkdown)
+library(scales) 
+library(stringdist)
+library(stringr)
+library(tidybayes) 
 library(tidyr)
-library(tidybayes)
-library(scales)
-library(janitor)
-library(stringr)
-library(ggplot2)
-library(patchwork)
 
-tar_load(multilex_data)
-tar_load(items)
-tar_load(participants)
-tar_load(responses)
-tar_load(loo_comp)
-tar_load(loo_prod)
-tar_load(fits_comp)
-tar_load(fits_prod)
+# load and process objects
+tar_load_everything()
+participants <- participants %>% 
+    left_join(select(multilex_data$logs, id, time, edu_parent))
+items <- items %>% 
+    left_join(select(pool, item, language, category, class))
 
+# set global options
 options(
@@
 
-theme_set(theme_apa())
+# set custom ggplot2 theme
+theme_set(theme_custom())
 
+# resolve namespace conflicts
+conflict_prefer("filter", "dplyr")
+conflict_prefer("select", "dplyr")
 ```
@@
 ```{r participants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
 n_participants <- length(unique(participants$id))
 dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
-
 ```
@@
 
-```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Participant sample size by age and degree of exposure to a second language.
+```{r participants-lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Participant sample size by age and degree of exposure to a second language.
 Information about DoE was provided by families when filling the questionnaire.
@@
 Hyphens indicate that no participants from that specific combination of age filled the questionnaire."}
-labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
-t <- participants %>% 
-  mutate(
-    age = round(age),
-    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
-    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
+participants %>% 
+    mutate_at(
+        vars(starts_with("doe_")), 
+        ~floor(.*10)/10
     ) %>% 
-  count(age, doe_2) %>% 
-  arrange(doe_2) %>% 
-  mutate(doe_2 = fct_inorder(doe_2)) %>% 
-  pivot_wider(names_from = age, values_from = n, values_fill = NA) %>% 
+    mutate(age = cut_width(floor(age), 4)) %>% 
+    count(age, doe_catalan) %>% 
+    arrange(-doe_catalan) %>% 
+    pivot_wider(names_from = age, values_from = n) %>% 
     gt() %>% 
-  tab_spanner(label = "Number of participants by age (months)", columns = 2:17) %>% 
-  fmt_missing(everything(), everything(), missing_text = "-") %>% 
+    fmt_missing(everything(), everything(), missing_text = "--") %>% 
     cols_label(
-    doe_2 = md("**L2 DoE**")
+        doe_catalan = "DoE Catalan"
+    ) %>%
+    fmt_percent("doe_catalan", decimals = 0) %>% 
+    tab_spanner("Age (months)", 2:8)
+```
+
+We use the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain.
+Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))` and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
+Highest parental educational attainment was similar across participants with different DoE to a second language (see Appendix).
+
+## Questionnaire
+
+```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+n_categories <- nrow(distinct(multilex_data$pool, category))
+n_te <- length(unique(multilex_data$pool$te))
+n_items <- length(unique(pool$item))
+n_item_language <- count(pool, language) %>%
+    pull(n) %>% 
+    set_names(c("catalan", "spanish"))
+```
+
+The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]`.
+Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+In case two translation equivalents were possible for a given word, both were inluded as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
+We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
+For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
+
+For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
+Some families filled a long version of the vocabulary checklists (`r format(n_te, big.mark = ",")` translation equivalents; `r format(n_item_language["catalan"], big.mark = ",")` items in Catalan, `r format(n_item_language["spanish"], big.mark = ",")` items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
+These last families were randomly allocated into one of four different subsets of the complete list of items.
+These lists were designed so that each contained a representative sub-sample of the items from the complete list.
+Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
+Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
+
+To analyse the data, we considered responses to items corresponding to nouns, verbs, and adjectives [@fourtassi2020].
+We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]), items that included more than one word-form (e.g., *mono / mico*), and words with missing lexical frequency scores.
+
+```{r item-summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
+Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
+items %>%
+    unnest(list) %>% 
+    group_by(language, list) %>% 
+    summarise_at(
+        vars(freq, n_phon, lv), 
+        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
+    ) %>% 
+    select(-c(n_phon_n, lv_n)) %>% 
+    rename(n = freq_n) %>% 
+    gt(groupname_col = "language") %>% 
+    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
+    fmt_percent(matches("lv_")) %>% 
+    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
+    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
+    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
+    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
+    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
+    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
+    cols_label(
+        list = "",
+        n = md("*N*"),
+        freq_mean = md("Mean"),
+        freq_sd = md("*SD*"),
+        freq_min = md("Range"),
+        n_phon_mean = md("Mean"),
+        n_phon_sd = md("*SD*"),
+        n_phon_min = md("Range"),
+        lv_mean = md("Mean"),
+        lv_sd = md("*SD*"),
+        lv_min = md("Range")
     ) %>% 
@@
     )
-
-gtsave(t, here("Figures/participants_lp.png"))
-
-```
-
-We use the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain.
-Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))` and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
-Highest educational attainment was similar across participants with different DoE to a second language, and therefore, differences in comprehension or production between participants with different language profiles are unlikely to stem from a difference in SES [@fernald2013ses] (see Table 2).
-
-```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Educational attainment.
-Proportion of highest parental educational attainment by DoE to the second language.
-For illustration purposes, DoEs were binned into 10\\% wide bins."}
-labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
-
-t <- participants %>%
-  mutate(
-    age = round(age),
-    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
-    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
-  ) %>% 
-  group_by(doe_2) %>%
-  summarise(
-    n_no_education = length(edu_parent[edu_parent=="No education"])/n(),
-    n_primary = length(edu_parent[edu_parent=="Primary"])/n(),
-    n_secondary = length(edu_parent[edu_parent=="Secondary"])/n(),
-    n_complementary = length(edu_parent[edu_parent=="Complementary"])/n(),
-    n_vocational = length(edu_parent[edu_parent=="Vocational"])/n(),
-    n_university = length(edu_parent[edu_parent=="University"])/n(),
-    .groups = "drop"
-  ) %>% 
-  gt() %>% 
-  tab_spanner("Educational attainment", columns = 2:7) %>% 
-  cols_label(
-    doe_2 = md("**L2 DoE**"),
-    n_no_education = "No education",
-    n_primary = "Primary",
-    n_secondary = "Secondary",
-    n_complementary = "Complementary",
-    n_vocational = "Vocational",
-    n_university = "University"
-  ) %>% 
-  fmt_percent(2:7) %>% 
-  fmt_missing(everything(), missing_text = "-") %>% 
-  tab_style(
-    cell_text(style = "italic"),
-    cells_column_labels(everything())
-  ) %>% 
-  tab_style(
-    cell_text(weight = "bold"),
-    cells_column_spanners(everything())
-  )
-
-gtsave(t, here("Figures/participants_edu.png"))
-
-```
-
-## Questionnaire
-
-```{r items, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-pool <- multilex::pool %>% 
-  drop_na(cognate)
-n_categories <- distinct(pool, category) %>% nrow()
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
-n_item_cognate <- count(pool, language, cognate) %>% 
-  pull(n) %>% 
-  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
-```
-
-The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*.
-We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
-For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
-These last families were randomly allocated into one of four different subsets of the complete list of items.
-These lists were carefully designed so that each contained a representative subsample of the items from the complete list.
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
-
-```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
-Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
-
-d <- items %>%
-  select(te, language, cognate, frequency) %>% 
-  left_join(select(multilex_data$pool, te, language, version)) %>% 
-  unnest(version) %>% 
-  group_by(language, version, cognate) %>% 
-  summarise(
-    frequency_mean = mean(frequency, na.rm = TRUE),
-    frequency_sd = sd(frequency, na.rm = TRUE),
-    n = n(),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    frequency_se = frequency_sd/sqrt(n),
-    frequency_ci_lower = frequency_mean + qnorm(0.025)*frequency_se,
-    frequency_ci_upper = frequency_mean + qnorm(0.975)*frequency_se
-  ) %>% 
-  pivot_wider(names_from = c("cognate"), values_from = c("n", matches("frequency"))) %>% 
-  clean_names() %>% 
-  mutate(
-    frequency_ci_identical_cognate = paste0(
-      "[", round(frequency_ci_lower_identical_cognate, 2), ", ", round(frequency_ci_lower_identical_cognate, 2), "]"
-    ),
-    frequency_ci_non_identical_cognate = paste0(
-      "[", round(frequency_ci_lower_non_identical_cognate, 2), ", ", round(frequency_ci_lower_non_identical_cognate, 2), "]"
-    ),
-    frequency_ci_non_cognate = paste0(
-      "[", round(frequency_ci_lower_non_cognate, 2), ", ", round(frequency_ci_lower_non_cognate, 2), "]"
-    ),
-  ) %>% 
-  relocate(
-    language, version,
-    n_identical_cognate,
-    frequency_mean_identical_cognate,
-    frequency_se_identical_cognate,
-    frequency_ci_identical_cognate,
-    matches("_non_identical"),
-    matches("_non_cognate")
-  ) %>% 
-  mutate(language = ifelse(language=="Catalan", "CAT", "SPA")) %>% 
-  select(-matches("sd|upper|lower"))
-
-t <- d %>% 
-  gt(groupname_col = "language") %>% 
-  fmt_number(matches("mean|se")) %>% 
-  tab_spanner("Identical cognates", columns = ends_with("_identical_cognate")) %>% 
-  tab_spanner("Non-identical cognates", columns = ends_with("_non_identical_cognate")) %>% 
-  tab_spanner("Non-cognates", columns = ends_with("_non_cognate")) %>% 
-  cols_label(
-    version = "List",
-    n_identical_cognate = md("N"),
-    n_non_identical_cognate = md("N"),
-    n_non_cognate = md("N"),
-    frequency_mean_identical_cognate = md("M"),
-    frequency_mean_non_identical_cognate = md("M"),
-    frequency_mean_non_cognate = md("M"),
-    frequency_se_identical_cognate = md("SD"),
-    frequency_se_non_identical_cognate = md("SD"),
-    frequency_se_non_cognate = md("SD"),
-    frequency_ci_identical_cognate = md("95% *CI"),
-    frequency_ci_non_identical_cognate = md("95% CI"),
-    frequency_ci_non_cognate = md("95% CI")
-  ) %>% 
-  cols_width(
-    starts_with("frequency_ci_") ~ px(100),
-    TRUE ~ px(40)
-  ) %>% 
-  tab_style(
-    cell_text(style = "italic"),
-    cells_column_labels(everything())
-  ) %>% 
-  tab_style(
-    cell_text(weight = "bold"),
-    cells_column_spanners(everything())
-  )
-
-gtsave(t, here("Figures/item_frequency.png"))
-
 ```
@@
 ```{r n_responses, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-n_obs <- nrow(responses)
-sum_obs <- responses %>%
+n_obs <- nrow(df)
+sum_obs <- df %>%
     group_by(te) %>%
@@
         mean = mean(n),
+        median = median(n),
         sd = sd(n),
@@
     )
-
 ```
 
-We gathered `r printnum(n_obs, big.mark = ",")` observations, with each observation corresponding to a single reponse of one participant to a single item.
-Translation equivalents (TEs) received an average of `r printnum(sum_obs$mean, digits = 2)` responses (*SD* = `r printnum(sum_obs$sd, digits = 2)`, *Min* = `r printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
-To investigate our hypotheses, we modelled the probability of children understanding or producing TEs using a logistic regression model with a Bernoulli distribution and a logit link function [@agresti2003categorical].
-We included several predictors of interest as fixed effects to adjust this probability to participants' and items' characteristics, namely:
+We gathered `r printnum(n_obs, big.mark = ",")` observations, with each observation corresponding to a single response of one participant to a single item.
+Translation equivalents (TEs) received a median of `r printnum(sum_obs$median, digits = 2)` responses (*Min* = `r printnum(sum_obs$min, digits = 2)`, *Max* = `r printnum(sum_obs$max, digits = 2)`), both languages summed together.
 
-* *Age* of participant in months (`age_center`), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
-* *Lexical frequency* of item (`frequency_center`), Relative lexical frequencies (frequency per million words, $fpmw$) were extracted from the CHILDES data base, using the `childesr` R package [@braginsky2018childesr], and were transformed to Zipf scores [@van2014subtlex; @zipf1949human] ($log10(fpmw)+3$).
-Missing frequencies were imputed by multiple imputation using the `mice` R package [van2011mice].
-The resulting scores were then centred around the mean.
-* *Degree of exposure* to the language (`doe_center`), computed as the proportion of the time the child is exposed to the language the item belongs to.
-For instance, for a response to the item `perro` (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1).
-Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10% increase in exposure to the language of the item.
-* *Cognateness*, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist.
-We sum-coded two contrasts [@schad2020capitalize]: `cognate1`, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and `cognate2`, comparing IC (coded as +0.5) with NIC (coded as -0.5).
+We modelled the probability of answering each response category (*No* < *Understands* < *Understands and Says*) using a mixed effects model.
+This allowed us to simultaneously estimate the impact of item- and participant-level predictors on the probability of word comprehension and production.
+We included the following predictors: age of the participant in months, calculated as the difference in months between participants' birth date and questionnaire completion ($Age$), word lexical frequency, extracted from the CHILDES database using the `childesr` R package [[@braginsky2018childesr]] and expressed in Zipf scores, ($log10(fpmw)+3$) [@van2014subtlex; @zipf1949human] ($Frequency$), number of phonemes included in the IPA phonological transcription of the word-form ($Phonemes$), participant's degree of exposure to the language the item belongs to ($DoE$)^[For instance, for a response to the item `perro` (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1).], and the Levenshtein distance between the IPA transcriptions of the translation equivalent in Catalan and in Spanish, as a measure of their phonological similarity ($Levenshtein$) [@floccia2018methods].
 
-We first fit a base model that only included the main effects of `age_center` and `frequency_center`  (Model 1) [see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches].
-We then entered our predictors of interest by adding `doe_center` (Model 2) and then `cognate` and its interaction with `doe_center` (Model 3).
-To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included *Participants* (`id`) and *TEs* (`te`) as grouping variables (i.e.
-random effects), with all possible random intercepts and slopes where appropriate [@barr2013random] (see Appendix for more details of the model).
+We included $Age$, $Frequency$, $Phonemes$, $DoE$, and $Levenshtein$ as fixed effects.
+We also included a three-way interaction between $Age$, $DoE$, and $Levenshtein$ and its corresponding two-way interactions.
+We added participants and items as crossed random effects including all intercepts and slopes when appropriate [@barr2013maximal].
+This model structure, including word- and participant-level properties to estimate vocabulary acquisition trajectories was discussed by @kachergis2021toward in the context of explanatory item response models.
+Our model differs from their proposal in that  1) while they modelled global trajectories exclusively, we modelled individual trajectories as well, 2) our outcome variable is ordinal in contrast to the binary outcome used in their study, 3) we propose a Bayesian approach.
+Our motivation for a Bayesian approach towards statistical inference is that (1) it allows us to incorporate previous domain knowledge to improve estimation accuracy, and (2) to quantify the uncertainty associated with the estimated parameters in our model [@mcelreath2018statistical].
 
-We adopted a Bayesian approach toward statistical inference.
-This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model [@mcelreath2018statistical].
-We fit the models using the R package `brms` [@burkner2017brms], an R interface to the probabilistic language Stan [@carpenter2017stan].
-We run four sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
-All R-hat values were smaller than 1.1, indicating that the chains converged successfully in all cases [@gelman1992inference] (see Appendix for an extended description of model diagnostics).
+The Bayesian approach implies using Bayes’ rule to assign a probability to possible parameter values (i.e., posterior distribution), based on previous knowledge about such probability distribution (prior distribution) and our data (likelihood).
+We used Markov chain Monte Carlo (MCMC) to approximate the posterior distribution via sampling.
+We run `r dim(model_fit_7$fit)[2]` sampling chains with `r dim(model_fit_7$fit)[1]` iterations each, including `r dim(model_fit_7$fit)[1]/2` warm-up iterations per chain.
+All $\hat{R}$ values were smaller than 1.1, indicating that the chains converged successfully in all cases [@gelman1992inference] (see Appendix for an extended description of model structure and diagnostics).
 
-We then compared comprehension and production models separately using leave-one-out cross-validation (LOO-CV) [@vehtari2017practical]^[Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson2019bayesian]].
-This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
-The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
-We compared the expected log posterior density (ELPD) of each each model to test which one fitted the data the best.
+We compared a base model–only including fixed and random intercepts–to increasingly more complex models, introducing predictors in the following order: $Age$, $Frequency$, $Phonemes$, $DoE$, $Levenshtein$, two-, and three-way interactions [see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches].
+We compared the predictive validity of the models using leave-one-out cross-validation (LOO-CV), and selected the model with the smallest expected log posterior density (ELPD) [@vehtari2017practical]^[Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples [@magnusson2019bayesian]].
+Finally, we explored the posterior distribution of each parameter in the selected model.
+We report the mean and 95% credible interval of the posterior distribution of the parameters of interest.
+We performed follow-up tests on interactions whose 95% credible interval excluded 0 by comparing the 95% CrI of their expected posterior marginal means using the `tidybayes` R package [@R-tidybayes].
+Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse].
 
@@
 ```{r coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+coefs <- as.data.frame(fixef(model_fit_7)) %>% 
+    clean_names()
 
-coef_comp <- as.data.frame(fixef(fits_comp$fit_2))
-coef_prod <- as.data.frame(fixef(fits_prod$fit_2))
-
+draws_fixed <- gather_draws(model_fit_7, `b_.*`, regex = TRUE)
 ```
 
-For both comprehension and production, the extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density.
+Table 2 shows the summary of the posterior distribution of the fixed coefficients of the model.
+
+```{r results-fixed, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+str_repl <- c(
+    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein",
+    "b_age_std:doe_std" = "Age \u00d7 DoE",
+    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
+    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein",
+    "b_lv_std" = "Levenshtein (+1 SD)",
+    "b_doe_std" = "DoE (+10%)",
+    "b_n_phon_std" = "# Phonemes (+1 SD)",
+    "b_freq_std" = "Frequency (+1 SD)",
+    "b_age_std" = "Age (+1 month)",
+    "b_Intercept[1]" = "Intercept (Production)",
+    "b_Intercept[2]" = "Intercept (Comprehension)"
+)
+
+draws_fixed %>%
+    mutate(
+        .value = ifelse(
+            grepl("Intercept", .variable),
+            inv_logit_scaled(.value), 
+            .value/4
+        ),
+        .variable = factor(.variable, levels = names(str_repl), labels = str_repl)
+    ) %>% 
+    mean_qi(.width = c(0.95, 0.89)) %>% 
+    pivot_wider(names_from = .width, values_from = c(.value, .lower, .upper)) %>% 
+    select(-c(.interval, .point, .value_0.89)) %>% 
+    arrange(desc(.variable)) %>% 
+    gt() %>% 
+    fmt_percent(2:6) %>% 
+    cols_merge(vars(".lower_0.95", ".upper_0.95"), pattern = "[{1}, {2}]") %>% 
+    cols_merge(vars(".lower_0.89", ".upper_0.89"), pattern = "[{1}, {2}]") %>% 
+    cols_label(
+        .variable = md("Predictor"),
+        .value_0.95 = md("Mean"),
+        .lower_0.95 = md("95\\% *CrI*"),
+        .lower_0.89 = md("89\\% *CrI*")
+    ) %>% 
+    tab_footnote(
+        footnote = "Transformed using the inverse logit to get the average probability of correct response",
+        locations = cells_body(
+            columns = ".variable", 
+            rows = grepl("Intercept", .variable)
+        ) 
+    ) %>% 
+    tab_footnote(
+        footnote = "Transformed using the divide-by-four rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
+        locations = cells_body(
+            columns = ".variable",
+            rows = !grepl("Intercept", .variable)
+        )
+    )  %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_labels(columns = everything())
+    ) %>% 
+    tab_header(title = "Model fixed coefficients", subtitle = "Posterior distribution") %>% 
+    tab_style(
+        cell_text(align = "left"),
+        list(cells_body(), cells_column_labels(columns = everything()))
+    )
+```
+
+The extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density.
 This estimate is several times larger than its standard error indicating high predictive accuracy.
@@
 Standard errors (*SE*) indicate the uncertainty around the computed ELPD in each model.
-A small *SE* (in realtion to the magnitde of the *ELPD* indicates less uncertainty around the ELPD estimate."}
-
+A small *SE* (in realtion to the magnitde of the *ELPD* indicates less uncertainty around the ELPD estimate.", eval=FALSE}
 l_comp <- loo_comp %>% 
@@
     )
-
-gtsave(t, here("Figures/loos.png"))
-
 ```
@@
 
-```{r fix_effets, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension and production.
+```{r results-fixed, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension and production.
 Points and error bars in indicate the posterior means, and 50% and 95% CrI.
@@
 The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-
 c_comp <- fixef(fits_comp$fit_2) %>% 
@@
 gtsave(t, here("Figures/fix_effects.png"))
-
 ```
@@
 Dots represent the mean, and error bars represent the range of values within the true values is contained with 50% (thick line) and 95% (thin line) cprobability, given the data."}
-
 nd <- expand.grid(

### 2022-09-13 R/02_models.R

@@ -7,7 +7,6 @@ fit_model <- function(name, ...){
     # see R/utils.R
     fit <- brm(
         ...,
-        sample_prior = "yes", # samples prior for faster computation of Bayes Factors and LOO
         iter = 4000,
         chains = 2,
         init = 0.5, # where to initialise MCMCs
@@ -25,34 +24,36 @@ fit_model <- function(name, ...){
     return(fit)
 }
 
-
-#' Compare brms models
+#' Extract log-likelihood of brmsfit object or list of brmsfit objects via \code{loo::log_lik}
 #' @param x A list of \code{brmsfit} objects that will be compared
-#' @param criterion A character string indicating the criterion to use to compare models. For now, it can only take "loo" as value, which invokes Leave-One-Out Cross-Validation (LOO).
-compare_models <- function(x, criterion = "loo", ...) {
-    
-    if (criterion=="loo") {
-        # models are too large, compute LOO from subsamples (500 by default)
-        loos <- lapply(x, loo_subsample, ...)
-        saveRDS(loos, here("results", paste0("model_", criterion, ".rds")))
-        
-    } else {
-        # future versions of this function might allow other criteria like WAIC
-        stop("Criterion must be one of: 'loo'")
-        
+#' @param ... Arguments to be passed to \code{log_lik}
+get_log_lik <- function(x, ...){
+    # make sure x is a brms fit object or a list of thereof
+    stopifnot("x must be a brmsfit object or a list of brmsfit objects" = all(sapply(x, is.brmsfit)))
+    # if x is a single brmsfit object, make it a list (of one element)
+    if (is.brmsfit(x)) x <- list(x)
+    n_models <- length(x) 
+    log_liks <- vector(mode = "list", length = n_models) # pre-allocate vectors
+    for (i in 1:n_models){
+        log_liks[[i]] <- log_lik(x[[i]], ...)  # compute log-likelihood
+        message(paste0(i, "/", n_models, " log-likelihoods computed"))
     }
+    if (!is.null(names(x))) names(log_liks) <- names(x) # keep names if any
+    return(log_liks)
+}
+
+
+#' Compute leave-one-out cross-validation of brmsfit object or list of brmsfit objects via \code{loo::loo}
+#' @param x A list of log-likelihoods extracted from brmsfit objects using \code{loo:log_lik}
[…]

### 2022-09-13 _targets.R

@@ -30,7 +30,8 @@ tar_option_set(
         "multilex",
         "patchwork", 
         "papaja", 
-        "purrr", 
+        "purrr",
+        "quarto",
         "readxl", 
         "rmarkdown",
         "rlang",
@@ -51,7 +52,8 @@ options(
     mc.cores = 4,
     brms.backend = "cmdstanr",
     tidyverse.quiet = TRUE,
-    knitr.duplicate.label = "allow"
+    knitr.duplicate.label = "allow",
+    loo.cores = 1
 )
 
 list(
@@ -175,68 +177,6 @@ list(
         model_fit_0,
         fit_model(
             name = "fit_0",
-            formula =  bf(
-                response ~ 1 +
-                    (1 | id) +
-                    (1 | te),
-                family = cratio(link = "logit") # cumulative, continuation ratio
-            ), 
-            data = df,
-            prior = model_prior[1:3, ]
-        )
-    ),
-    
-    # add age main effect
-    tar_target(
-        model_fit_1,
-        fit_model(
-            name = "fit_1",
-            formula = bf(
-                response ~ age_std +
-                    (1 + age_std | id) +
-                    (1 + age_std | te),
-                family = cratio(link = "logit") # cumulative, continuation ratio
-            ), 
-            data = df,
-            prior = model_prior[1:5, ]
-        )
-    ),
-    
-    # add age main effect
-    tar_target(
-        model_fit_2,
-        fit_model(
-            name = "fit_2",
-            formula = bf(
-                response ~ age_std + freq_std + 
-                    (1 + age_std + freq_std | id) +
-                    (1 + age_std | te),
-                family = cratio(link = "logit") # cumulative, continuation ratio
-            ), 
-            data = df,
-            prior = model_prior[1:6, ]
-        )
-    ),
-    
-    # add dominance main effect
-    tar_target(
-        model_fit_3,
-        fit_model(
-            name = "fit_3",
-            formula = bf(
-                response ~ age_std + freq_std + n_phon_std +
-                    (1 + age_std + freq_std + n_phon_std | id) +
-                    (1 + age_std + freq_std + n_phon_std | te),
-                family = cratio(link = "logit") # cumulative, continuation ratio
-            ), 
-            data = df,
-            prior = model_prior[1:7, ]
-        )
-    ),
-    tar_target(
-        model_fit_4,
-        fit_model(
-            name = "fit_4",
             formula = bf(
                 response ~ age_std + freq_std + n_phon_std + doe_std +
[…]

### 2022-09-16 index.qmd
Updte report

@@
         toc: true
-        smooth-scroll: true
         link-external-newwindow: true
@@
         code-copy: hover
-        fig-dpi: 500
+        fig-dpi: 1000
         reference-location: margin
         execute-dir: "C:/Users/U155880/Documents/trajectories/"
-bibliography: "manuscript/references.bib"
+    pdf:
+        toc: true
+        number-sections: true
+        colorlinks: true
+        fig-dpi: 1000
+bibliography: "references.bib"
 csl: "manuscript/apa6.csl"
@@
 library(brms) 
+library(bayestestR)
 library(conflicted)
@@
 library(ggplot2) 
-library(marginaleffects)
 library(ggsci)
@@
 ```{r}
-#| label: items-summary
+#| label: tbl-items-summary
 #| echo: true
@@
 #| fig-cap: "Lexical frequencies.
-Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
+Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
 items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))
@@
 ```{r}
-#| label: items-frequency
+#| label: fig-items-freq
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Distribution of lexical frequency across items"
 items %>% 
@@
         slab_size = 1,
-        alpha = 0.5,
+        alpha = 0.75,
     ) +
@@
 ```{r}
-#| label: items-nphon
+#| label: fig-items-nphon
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Distribution of number of phonemes across items"
 items %>%
@@
 ```{r}
-#| label: items-lv
+#| label: fig-items-lv
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Distribution of Levenshtein similarities across items"
 items %>% 
@@
 ```{r}
-#| label: items-frequency-phonemes
+#| label: fig-items-frequency-phonemes
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
 items %>% 
@@
 ```{r}
-#| label: items-frequency-lv
+#| label: fig-items-frequency-lv
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Association betwen lexical frequency and phonological similarity"
 items %>% 
@@
 ```{r}
-#| label: items-nphon-lv
+#| label: fig-items-nphon-lv
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Association betwen number of phonemes and phonological similarity"
 items %>% 
@@
 ```{r}
-#| label: participants-time
+#| label: fig-participants-time
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Data at which responses were collected"
 participants <- left_join(
@@
 ```{r}
-#| label: participants-age
+#| label: fig-participants-age
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Distribution of participants' ages"
 participants %>% 
@@
 ```{r}
-#| label: participants-lp
+#| label: fig-participants-lp
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
 participants %>% 
@@
 ```{r}
-#| label: participants-edu
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
+#| label: tbl-participants-edu
+#| tbl-cap: Participant exposure to the language (*DoE*) and socio-economic status (*SES*)
+#| echo: true
+#| message: false
+#| warning: false
 participants %>%
@@
     ) %>% 
-    fmt_missing(columns = 3:8, missing_text = "--") %>% 
+    sub_missing(columns = 2:8, missing_text = "--") %>% 
     tab_spanner("Educational attainment", 3:8) %>% 
@@
 ```{r}
-#| label: participants-edu-plot
+#| label: fig-participants-edu
 #| echo: true
@@
 ```{r}
-#| label: prior-predictions
+#| label: fig-prior-predictions
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Expected prior predictions."
 nd <- expand.grid(
@@
     scale_fill_brewer(
-        palette = "YlOrRd",
+        palette = "Reds",
         labels = c("95%", "89%", "67%", "50%")
@@
 ```{r}
-#| label: results-raw
+#| label: fig-results-raw
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Observed proportion of comprehension and production responses across ages.
+Colours indicate different values of phonological similarity (Levenshtein similarity, binned).
+Results for different degrees of exposure (binned) are plotted separately."
 df %>%
@@
     ) +
-    scale_color_brewer() +
+    scale_color_brewer(palette = "Reds") +
     # scale_colour_manual(values = wesanderson::wes_palette("Zissou1", n = 5, type = "continuous")[1:5]) +
@@
     )
-
-ggsave("plot.png", height = 6, width = 9)
 ```
@@
 
-We compared the performance of these models using Bayesian leave-one-out cross-validation (LOO) using the `loo` and `loo_compare` functions of the `brms` R package (dependent of the `LOO` R package).
-LOO consists in computing the average likelihood of each observation after estimating the model's parameters leave that same observation out of the data set.
-Although the `loo` function uses a particular algorithm that speeds up the computation of this criterion (pareto-smooth importance sampling, PSIS), the size of our data set lead us to rely on the computation of the same criterion using a sampling approach via de `loo_subsample` function.
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO) via the `loo` R package.
+LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
+Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_3$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model as:
 
@@
 #| eval: false
+#  loo(log_lik(model_fit_3, ndraws = 750, re_formula = NA))
+```
+
+@tbl-results-loos shows the outputs of the *LOO-CV*.
+The extended model (`model_fit_3`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference.
+This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.
+
+```{r}
+#| label: tbl-results-loos
+#| echo: true
+#| message: false
+#| warning: false
 #| fig-width: 7
+#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
 model_loos %>% 
@@
     gt() %>% 
-    tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
-    tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
-    tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
-    tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
+    tab_header(
+        title = "",
+        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation.
+The likelihood of each data point given a model fitted without including that same observation was estimated.
+The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
+*ELPD* values closer to zero indicate better predictive acuracy.
+We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
+The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_3`)")
+    ) %>% 
+    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
+    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>% 
+    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>% 
+    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>% 
     fmt_number(2:9) %>% 
@@
         p_loo = md("*p*"),
-        se_p_loo = md("SE"),
+        se_p_loo = md("*SE*"),
         # looic
@@
     ) %>% 
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)"))
+    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_spanners()
+    ) %>% 
+    tab_style(
+        cell_text(align = "left"),
+        cells_title(groups = "subtitle")
+    )
 ```
@@
 ```{r}
-#| label: loo-diagnostics-0
+#| label: fig-loo-diagnostics-0
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
 get_loo_plot <- function(x){
@@
 ```{r}
-#| label: loo-diagnostics-1
+#| label: fig-loo-diagnostics-1
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
 get_loo_plot(model_loos$model_fit_1)
@@
 ```{r}
-#| label: loo-diagnostics-2
+#| label: fig-loo-diagnostics-2
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
 get_loo_plot(model_loos$model_fit_2)
@@
 ```{r}
-#| label: loo-diagnostics-3
+#| label: fig-loo-diagnostics-3
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
 get_loo_plot(model_loos$model_fit_3)
@@
 ```{r}
-#| label: results-fixed-table
-#| echo: true
-#| message: false
-#| warning: false
+#| label: fig-results-rope
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-height: 4
 #| fig-width: 7
+rope_coefs %>% 
+    ggplot(aes(estimate, y = study)) +
+    geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
+    geom_vline(
+        xintercept = median(rope_coefs$estimate),
+        size = 1,
+        colour = pal_d3()(4)[4]
+    ) +
+    stat_dots(
+        dotsize = 0.5, 
+        binwidth = 0.03,
+        fill = "grey50",
+        colour = "grey50"
+    ) +
+    stat_summary(
+        fun = median,
+        geom = "point",
+        position = position_nudge(y = -0.125),
+        size = 6, 
+        shape = 17,
+        colour = pal_d3()(4)[4]
+    ) +
+    annotate(
+        geom = "text",
+        label = paste0(
+            "Grand median of cross-study
+             smallest effect size of interest\n",
+            "Median = ", round(median(rope_coefs$estimate), 3)
+        ),
+        x = median(rope_coefs$estimate)-0.2,
+        y = 4,
+        hjust = 1,
+        vjust = 1.2,
+        size = 4.5,
+        colour = pal_d3()(4)[4]
+    ) +
+    annotate(
+        geom = "curve",
+        x = median(rope_coefs$estimate)-0.2,
+        xend = median(rope_coefs$estimate)-0.01,
+        size = 1,
+        y = 3.6,
+        yend = 3.5,
+        colour = pal_d3()(4)[4],
+        arrow = arrow(length = unit(0.5, "cm"))
+    ) +
+    labs(
+        x = "Standard regression coefficient (logit scale)\n(Excluding intercepts and predictor age)",
+        y = "Study"
+    ) +
+    guides(fill = guide_legend(ncol = 1)) +
+    scale_x_continuous(breaks = seq(-0.5, 0.5, 0.1)) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        legend.justification = 0,
+        axis.title.y = element_blank(),
+        panel.grid.major.y = element_blank()
+    )
+```
+
+```{r}
+#| label: results-posterior-fixed
+#| echo: true
+#| message: false
+#| warning: false
+post_draws <- gather_draws(model_fit_3, `b_.*`, regex = TRUE)
+
+# tidy predictor names
 str_repl <- c(
-    "Intercept\\[1\\]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
-    "Intercept\\[2\\]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
-    "age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein",
-    "age_std:doe_std" = "Age \u00d7 DoE",
-    "age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "doe_std:lv_std" = "DoE \u00d7 Levenshtein",
-    "age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
-    "freq_std" = paste0("Frequency (+1 SD, ",  round(sd(df$freq), 2), " Zipf)"),
-    "n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
-    "doe_std" = paste0("DoE (+1 SD, ",  percent(sd(df$doe)), ")"),
-    "lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")")
-)
-
-intercept_footnote <- paste0(
-    "Probability of comprehension or production at the average age of the sample (", round(mean(df$age), 2), " months), transformed to the probability scale using the inverse logit."
-)
-
-slopes_footnote <- "Transformed using the divide-by-four rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."
-
-draws_fix <- fixef(model_fit_3) %>% 
-    as.data.frame() %>%
-    rownames_to_column("term") %>% 
-    clean_names() %>% 
-    select(-est_error) %>% 
-    mutate(
-        estimate_p = ifelse(grepl("Intercept", term), plogis(estimate), estimate/4),
-        q2_5_p = ifelse(grepl("Intercept", term), plogis(q2_5), q2_5/4),
-        q97_5_p = ifelse(grepl("Intercept", term), plogis(q97_5), q97_5/4),
-        term = str_replace_all(term, str_repl),
-        group = ifelse(grepl("Intercept", term), "Intercepts", "Slopes")
-    )
-
-gt(draws_fix, rowname_col = "term") %>% 
-    fmt_number(2:7, decimals = 2) %>% 
-    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
-    cols_merge(c("q2_5_p", "q97_5_p"), pattern = "[{1}, {2}]") %>% 
-    tab_spanner("Logit scale", c(estimate, q2_5, q97_5)) %>% 
-    tab_spanner("Probability scale", c(estimate_p, q2_5_p, q97_5_p)) %>% 
-    tab_row_group(label = "Intercepts", rows = 1:2) %>% 
-    tab_row_group(label = "Slopes", rows = 3:nrow(draws_fix)) %>% 
-    row_group_order(groups = c("Intercepts", "Slopes")) %>% 
-    cols_label(
-        term = md("**Predictor**"),
-        estimate = "Mean",
-        q2_5 = md("95\\% CrI"),
-        estimate_p = "Mean",
-        q2_5_p = md("95\\% CrI"),
-    ) %>% 
-    tab_footnote(
-        footnote = intercept_footnote,
-        locations = cells_body(
-            columns = "term", 
-            rows = grepl("Intercept", term)
-        )
-    ) %>% 
-    tab_footnote(
-        footnote = slopes_footnote,
-        locations = cells_body(
-            columns = "term", 
-            rows = term %in% str_repl[-c(1, 2)]
-        )
-    ) %>% 
-    cols_align(
-        align = c("right"),
-        columns = 2:7
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(everything())
-    ) %>% 
-    tab_style(
-        list(cell_text(style = "italic"), cell_fill(color = "grey")),
-        cells_row_groups()
-    ) %>% 
-    cols_hide(group)
-```
-
-<br>
-
-```{r}
-#| label: results-fixed
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Points and error bars in indicate the posterior means, and 50% and 95% CrI.
-The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."
-str_repl <- c(
-    "b_Intercept\\[1\\]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
-    "b_Intercept\\[2\\]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
+    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
+    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
     "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
@@
 
-post <- gather_draws(model_fit_3, `b_.*`, regex = TRUE) %>% 
+```
+
+```{r}
+#| label: tbl-results-fixed
+#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+intercept_footnote <- paste0(
+    "Probability of comprehension or production at the grand age of the sample (", 
+    round(mean(df$age), 2), " months)"
+)
+
+# summarise posterior draws
+posterior_description %>% 
     mutate(
-        .value = ifelse(
-            grepl("Intercept|sd", .variable),
-            inv_logit_scaled(.value), 
-            .value/4
-        ),
-        .variable_name = factor(
-            .variable, 
+        parameter = factor(
+            parameter, 
             levels = names(str_repl), 
-            labels = str_repl,
-            ordered = TRUE
-        )
-    )
-
-post %>%
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    geom_vline(xintercept = 0, size = 0.5, colour = "grey") +
-    stat_slab(
-        size = 0.5,
-        color = pal_d3()(2)[2],
-        fill = pal_d3()(2)[2],
-        position = position_nudge(y = 0.1)
-    ) +
-    geom_errorbar(
-        data = post %>% 
-            filter(!grepl("Intercept|sd", .variable)) %>%
-            mean_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>% 
+            labels = str_repl
+        )
+    ) %>%  
+    select(parameter, median, ci_high, ci_low, rope_percentage, rhat, ess) %>% 
+    gt(rowname_col = "parameter") %>% 
+    fmt_number(c(1:4, 6:7)) %>% 
+    fmt_percent(5) %>% 
+    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>% 
+    cols_label(
+        median = "Median",
+        rhat = "R-hat",
+        ess = "ESS",
+        ci_low = md("95% *HDI*"),
+        rope_percentage = md("*ROPE* prob.")
+    ) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_labels(columns = 1:7)
+    )
+post_draws_list <- split(posterior_description, posterior_description$parameter)
+
+```
+
+<br>
+
+We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
+We decided to draw this interval around zero so that it covered values smaller than those considered as negligible by similar previous studies: after consulting the regression tables of [ADD STUDIES], we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: 3%.
+consequently, we defined our *ROPE* from -3% to +3%.
+We report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
+This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
+Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the -3%--3% *ROPE* interval we previously defined.
+For those coefficients that did not exclude completely the *ROPE* we also report a Bayes Factor indicating how many times is such coefficient more likely to be bigger than the upper limit of the *ROPE*, or smaller than the lower limit of the *ROPE*, given our data.
+
+For interpretability, we report the estimated regression coefficients transformed from the probability scale.
+The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
+This means that one logit difference is not necessarily translated to a unique value in the probability scale.
+For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
+The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
+We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
+derivative) of the logistic curve at the age at which most participants were acquiring a given word.
+This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
+
+@fig-results-fixed shows the outputs of the model.
+**Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r printnum(post_draws_list[["b_age_std"]]$median, digits = 2)`, 95% *HDI* = [`r printnum(post_draws_list[["b_age_std"]]$ci_low, digits = 2)`, `r printnum(post_draws_list[["b_age_std"]]$ci_high, digits = 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month.
+
+**Lexical frequency** had a small, positive, but inconclusive on the probability of acquisition substantially, ($\beta$ = `r percent(post_draws_list[["b_freq_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_freq_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_freq_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_freq_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples overlapped with the *ROPE.*
+
+The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]).
+For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.
+
+The **degree of exposure to the language** ($DoE$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std"]]$ci_high, accuracy = 0.01)`]).
+All of the posterior samples of his regression excluded the *ROPE*.
+The impact of this predictor on the probability of acquisition was positive: for every 10% increase in exposure to the language the word belongs to, the participant was `r percent(post_draws_list[["b_doe_std"]]$median/(10*sd(df$doe)), accuracy = 0.01)` more likely to acquire it.
+
+The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
+
+However, the 95% *HDI* of the regression coefficient of the $Doe \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_doe_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the language it belonged to.
+Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
+This effect was negligible when exposure to both languages was balanced (e.g.
+50-50%).
+Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.
+
+[TO DO: emmeans tests]
+
+The 95% *HDI* of the regression coefficient of the $Age \times DoE$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:doe_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+This shows that the effect of the degree of exposure on the speed at which participants acquired word across ages was inconclusive.
+Give our data, it is difficult to tell whether this *DoE* increased or decreased the slope of the words' acquisition trajectories.
+
+The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
+
+Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_Std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of the degree of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
+
+```{r}
+#| label: fig-results-fixed
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 9
+#| fig-heigh: 7
+#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model.
+Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis).
+Intervals show the 95% highest density interval (*HDI*) of each distribution.
+This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
+the mean and HDI limits are indicated below each distribution.
+Coefficients were transformed from the logit scale to the probability scale for interpretability.
+We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e.
+the derivative of the logistic function in its mid-point."
+post <- post_draws %>% 
     mutate(
@@
         )
+    )
+
+post %>%
+    filter(!grepl("Intercept|sd", .variable)) %>%
+    ggplot(aes(.value, fct_rev(.variable_name))) +
+    annotate(
+        geom = "rect",
+        ymin = -Inf,
+        ymax = Inf,
+        xmin = rope_interval["lower"],
+        xmax = rope_interval["upper"],
+        colour = NA,
+        alpha = 0.2,
+        fill = pal_d3()(3)[2],
+    ) +
+    geom_vline(
+        xintercept = 0,
+        size = 1,
+        colour = pal_d3()(3)[2]
+    ) +
+    stat_slab(
+        aes(
+            fill = stat(abs(x) < rope_interval["upper"]),
+            colour = stat(abs(x) < rope_interval["upper"]),
             ),
-        aes(xmin = .lower, xmax = .upper, x = .value),
-        width = 0.15
-    ) +
-    geom_text(
+        size = 1,
+        # fill = pal_d3()(2)[2],
+        position = position_nudge(y = 0.15)
+    ) +
+    geom_errorbar(
         data = post %>% 
             filter(!grepl("Intercept|sd", .variable)) %>%
-            mean_hdi(
-                .exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")
+            median_hdi(
+                .exclude = c(
+                    ".chain", 
+                    ".iteration",
+                    ".draw",
+                    ".row", 
+                    ".variable_name"
+                )
             ) %>% 
@@
             ),
+        aes(xmin = .lower, xmax = .upper, x = .value),
+        width = 0.15
+    ) +
+    geom_text(
+        data = post %>% 
+            filter(!grepl("Intercept|sd", .variable)) %>%
+            mean_hdi(
+                .exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")
+            ) %>% 
+            mutate(
+                .variable_name = factor(
+                    .variable, 
+                    levels = names(str_repl),
+                    labels = str_repl,
+                    ordered = TRUE
+                )
+            ),
         aes(
             label = paste0(
-                percent(.value, accuracy = 0.1),
-                " [", percent(.lower, accuracy = 0.1), ", ",
-                percent(.upper, accuracy = 0.1), "]"
+                printnum(.value, digits = 2),
+                " [", printnum(.lower, digits = 2), ", ",
+                printnum(.upper, digits = 2), "]"
             )
@@
     ) +
+    annotate(
+        geom = "text",
+        label = "ROPE",
+        x = 0.3,
+        y = 10,
+        vjust = 1,
+        size = 5,
+        colour = pal_d3()(2)[2]
+    ) +
+    annotate(
+        geom = "curve",
+        arrow = arrow(length = unit(0.2, "cm")),
+        x = 0.3,
+        y = 9.5,
+        xend = 0.1,
+        yend = 9,
+        size = 1,
+        curvature = -0.25,
+        colour = pal_d3()(2)[2]
+    ) +
     labs(
-        x = "Posterior probability of acquisition",
+        x = "Logistic regression coefficient estimate",
         y = "Variable", 
-        fill = "Variable",
-        alpha = "CrI"
-    ) +
-    scale_x_continuous(labels = percent, breaks = seq(-0.2, 2.5, 0.1)) +
+        fill = "Overlaps with ROPE",
+        colour = "Overlaps with ROPE"
+    ) +
+    scale_x_continuous(breaks = seq(-0.5, 2, 0.2)) +
+    scale_fill_manual(
+        values = c(pal_d3()(1)[1], "grey"),
+        labels = c("No", "Yes")
+    ) +
+    scale_colour_manual(
+        values = c(pal_d3()(1)[1], "grey"),
+        labels = c("No", "Yes")
+    ) +
     theme(
+        legend.position = "top",
+        legend.justification = c(1, 1),
         axis.title.y = element_blank(),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.major.y = element_blank()
-
+        panel.grid.major.x = element_line(
+            colour = "grey85",
+            linetype = "dotted"
+        ),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.x = element_blank()
+    )
+```
+
+## Marginal means
+
+```{r}
+#| label: fig-results-marginal
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Expected posterior predictions.
+Posterior poreditions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
+These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model.
+Therefore, we generated 4,000 predictions for a series of combination of levels of interest.
+Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels.
+The X-axis indicates the age (in months) for which the prediction is generated.
+The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*).
+Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
+Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
+Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
+nd <- expand.grid(
+    n_phon_std = 0,
+    freq_std = 0,
+    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe)),
+    lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
+)
+
+m <- add_epred_draws(nd, model_fit_3, ndraws = NULL, re_formula = NA) %>% 
+    mutate(
+        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
+        doe_std = factor(
+            doe_std, 
+            levels = unique(nd$doe_std), 
+            labels = paste0("DoE: ", c(10, 50, 90), "%")
+        ),
+        lv_std = factor(
+            lv_std, 
+            levels = unique(nd$lv_std),
+            labels = paste0(c(0, 50, 100), "%")
+        )
+    ) %>% 
+    filter(.category != "No") %>% 
+    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
+    mutate(Understands = Understands + `Understands and Says`) %>% 
+    pivot_longer(
+        c(Understands, `Understands and Says`),
+        names_to = ".category",
+        values_to = ".epred"
+    ) %>% 
+    mutate(
+        .category = str_replace_all(
+            .category,
+            c(
+                "Understands and Says" = "Production\n(Understands and Says))",
+                "Understands" = "Comprehension\n(Understands)"
+            )
+        )
+    )
+
+d <- df %>%
+    mutate(
+        doe_std = case_when(
+            between(doe, 0.00, 0.20) ~ "DoE: 10%",
+            between(doe, 0.40, 0.60) ~ "DoE: 50%",
+            between(doe, 0.80, 1.00) ~ "DoE: 90%"
+        ) %>% 
+            factor(., levels = paste0("DoE: ", percent(c(0.1, 0.5, 0.9)))),
+        lv_std = case_when(
+            between(lv, 0.00, 0.20) ~ "0%",
+            between(lv, 0.40, 0.60) ~ "50%",
+            between(lv, 0.80, 1.00) ~ "100%"
+        ) %>% 
+            factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
+        age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
+            cut(
+                .,
+                breaks = ((seq(10, 38, 2)-mean(df$age))/(sd(df$age))),
+                labels = ((seq(10, 36, 2)-mean(df$age))/(sd(df$age))),
+                include.lowest = TRUE
+            ) %>% 
+            as.character() %>% 
+            as.numeric(),
+        Understands = response %in% c("Understands", "Understands and Says"),
+        `Understands and Says` = response %in% "Understands and Says"
+    ) %>% 
+    drop_na(doe_std, lv_std) %>% 
+    group_by(te, age_std, doe_std, lv_std) %>% 
+    summarise(
+        Understands = mean(Understands),
+        `Understands and Says` = mean(`Understands and Says`),
+        .groups = "drop"
+    ) %>% 
+    pivot_longer(
+        c(Understands, `Understands and Says`),
+        names_to = ".category",
+        values_to = ".epred"
+    ) %>% 
+    mutate(
+        .category = str_replace_all(
+            .category,
+            c(
+                "Understands and Says" = "Production\n(Understands and Says))",
+                "Understands" = "Comprehension\n(Understands)"
+            )
+        )
+    )
+
+m %>% 
+    ggplot() +
+    aes(
+        x = age_std, 
+        y = .epred, 
+        colour = lv_std, 
+        fill = lv_std,
+    ) +
+    facet_grid(.category~doe_std) +
+    stat_lineribbon(
+        size = 1,
+        alpha = 0.5,
+        .width = 0.95, 
+        colour = NA,
+        point_interval = mean_hdi
+    ) +
+    stat_summary(
+        data = d,
+        fun.data = "mean_se",
+        geom = "pointrange",
+        size = 0.25
+    ) +
+    # geom_point(
+    #     data = d,
+    #     position = position_dodge(width = 0.25),
+    #     size = 0.5,
+    #     alpha = 0.5
+    # ) +
+    stat_summary(fun = "mean", geom = "line", size = 1) +
+    scale_x_continuous(
+        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 36, 4)
+    ) +
+    scale_color_d3() +
+    labs(
+        x = "Age (months)", 
+        y = "Posterior probability of acquisition\n(Linear prediction)",
+        colour = "Phonological similarity (Levenshtein distance)", 
+        fill = "Phonological similarity (Levenshtein distance)"
+    ) +
+    scale_y_continuous(labels = percent, limits = c(0, 1)) +
+    theme(
+        legend.position = "top",
+        axis.ticks.x = element_line(colour = "black"),
+        panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
+        strip.background = element_rect(fill = "grey", colour = NA),
+        plot.background = element_rect(fill = "white", colour = NA)
     )
@@
 ```{r}
-#| label: results-random-id
+#| label: fig-results-random-id
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Posterior probability of acquisition by participant"
 re_id <- ranef(model_fit_3)$id[,,1] %>% 
@@
 ```{r}
-#| label: results-random-te
+#| label: fig-results-random-te
 #| echo: true
@@
 #| fig-width: 7
+#| fig-cap: "Posterior probability of acquisition by translation equivalent"
 re_te <- ranef(model_fit_3)$te[,,1] %>% 
@@
 
-## Marginal means
-
-```{r}
-#| label: results-marginal
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-nd <- datagrid(
-    model = model_fit_3,
-    n_phon_std = 0,
-    freq_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
-    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe)),
-    lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
-)
-
-m <- add_epred_draws(nd, model_fit_3, ndraws = NULL, re_formula = NA) %>% 
-    mutate(
-        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
-        doe_std = factor(
-            doe_std, 
-            levels = unique(nd$doe_std), 
-            labels = paste0("DoE: ", c(10, 50, 90), "%")
-        ),
-        lv_std = factor(
-            lv_std, 
-            levels = unique(nd$lv_std),
-            labels = paste0(c(0, 50, 100), "%")
-        )
-    ) %>% 
-    filter(.category != "No") %>% 
-    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
-    mutate(
-        .category = str_replace_all(
-            .category,
-            c(
-                "Understands and Says" = "Production\n(Understands and Says))",
-                "Understands" = "Comprehension\n(Understands)"
-            )
-        )
-    )
-
-d <- df %>%
-    mutate(
-        doe_std = case_when(
-            between(doe, 0.00, 0.20) ~ "DoE: 10%",
-            between(doe, 0.40, 0.60) ~ "DoE: 50%",
-            between(doe, 0.80, 1.00) ~ "DoE: 90%"
-        ) %>% 
-            factor(., levels = paste0("DoE: ", percent(c(0.1, 0.5, 0.9)))),
-        lv_std = case_when(
-            between(lv, 0.00, 0.20) ~ "0%",
-            between(lv, 0.40, 0.60) ~ "50%",
-            between(lv, 0.80, 1.00) ~ "100%"
-        ) %>% 
-            factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
-        age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
-            cut(
-                .,
-                breaks = ((seq(10, 38, 4)-mean(df$age))/(sd(df$age))),
-                labels = ((seq(10, 34, 4)-mean(df$age))/(sd(df$age))),
-                include.lowest = TRUE
-            ) %>% 
-            as.character() %>% 
-            as.numeric(),
-        Understands = response %in% c("Understands", "Understands and Says"),
-        `Understands and Says` = response %in% "Understands and Says"
-    ) %>% 
-    drop_na(doe_std, lv_std) %>% 
-    group_by(te, age_std, doe_std, lv_std) %>% 
-    summarise(
-        Understands = mean(Understands),
-        `Understands and Says` = mean(`Understands and Says`),
-        .groups = "drop"
-    ) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
-    mutate(
-        .category = str_replace_all(
-            .category,
-            c(
-                "Understands and Says" = "Production\n(Understands and Says))",
-                "Understands" = "Comprehension\n(Understands)"
-            )
-        )
-    )
-# group_by(age_std, doe_std, lv_std, .category) %>% 
-# summarise(
-#     .epred_mean = mean(.epred),
-#     .epred_sd = sd(.epred),
-#     n = n(),
-#     .groups = "drop"
+## Model diagnostics
+
+### Traceplots
+
+```{r}
+#| label: fig-diagnostics-traceplots
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 12
+# gather_draws(
+#     model_fit_3, 
+#     `b_.*`, `sd_te__.*`, `cor_te__.*`, 
+#     regex = TRUE, 
+#     ndraws = 1000
 # ) %>% 
-# mutate(
-#     .epred_se = .epred_sd/n,
-#     .epred_lower = .epred_mean-.epred_se,
-#     .epred_upper = .epred_mean+.epred_se,
-#     .category = str_replace_all(
-#         .category, 
-#         c(
-#             "Understands" = "Comprehension\n(Understands)",
-#             "Understands and Says" = "Production\n(Understands and Says))"
+#     mutate(.chain = paste("Chain ", .chain)) %>% 
+#     ggplot(aes(.iteration, .value, colour = .chain)) +
+#     facet_wrap(~.variable, scales = "free_y") +
+#     annotate(
+#         geom = "rect", 
+#         colour = NA, 
+#         xmin = 0,
+#         xmax = dim(model_fit_3$fit)[1]/2, 
+#         ymin = -Inf, 
+#         ymax = Inf, 
+#         alpha = 0.25, 
+#     ) +
+#     geom_line() +
+#     labs(
+#         x = "Iteration",
+#         y = "Sample value",
+#         colour = "Chain"
+#     ) +
+#     scale_color_d3() +
+#     theme(
+#         legend.position = "top",
+#         legend.title = element_blank(),
+#         panel.grid = element_blank(),
+#         axis.text.x = element_text(size = 7)
 #     )
-#     )
-# ) %>% 
-# select(.category, age_std, doe_std, lv_std, .epred = .epred_mean, .epred_lower, .epred_upper, n)
-
-m %>% 
-    ggplot() +
-    aes(
-        x = age_std, 
-        y = .epred, 
-        colour = lv_std, 
-        fill = lv_std,
-    ) +
-    facet_grid(.category~doe_std) +
-    # geom_line(
-    #     aes(group = interaction(.draw, lv_std, doe_std), colour = lv_std),
-    #     alpha = 0.05, 
-    #     size = 1
-    # ) +
-    # stat_lineribbon(size = 1, alpha = 0.5, .width = 0.95, colour = NA) +
-    geom_point(
-        data = d,
-        position = position_dodge(width = 0.25),
-        size = 0.5,
-        alpha = 0.5
-    ) +
-    stat_summary(fun = "mean", geom = "line", size = 1) +
-    # geom_line(
-    #     data = d,
-    #     position = position_dodge(width = 0.25)
-    # ) +
-    # geom_errorbar(
-    #     data = d,
-    #     aes(
-    #         ymin = .epred_lower,
-    #         ymax = .epred_upper
-    #     ),
-    #     width = 0,
-#     position = position_dodge(width = 0.25)
-# ) +
-
-scale_x_continuous(
-    breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-    labels = seq(10, 36, 4)
-    # sec.axis = dup_axis(
-    #     name = "Age (years; months)",
-    #     labels = months_to_years(
-    #         rescale_variable(
-    #             (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-    #             mean(df$age), 
-    #             sd(df$age)
-    #         )
-    #     )
-    # )
-) +
-    scale_color_d3() +
-    labs(
-        x = "Age (months)", 
-        y = "Posterior probability of acquisition\n(Linear prediction)",
-        colour = "Phonological similarity (Levenshtein distance)", 
-        fill = "Phonological similarity (Levenshtein distance)"
-    ) +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
-    theme(
-        legend.position = "top",
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        strip.background = element_rect(fill = "grey", colour = NA),
-        plot.background = element_rect(fill = "white", colour = NA)
-    )
-
-ggsave("plot.png", height = 9, width = 12)
-```
-
-## Model diagnostics
-
-### Traceplots
-
-```{r}
-#| label: diagnostics-traceplots
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 12
-#| fig-height: 12
-gather_draws(
-    model_fit_3, 
-    `b_.*`, `sd_te__.*`, `cor_te__.*`, 
-    regex = TRUE, 
-    ndraws = 1000
-) %>% 
-    mutate(.chain = paste("Chain ", .chain)) %>% 
-    ggplot(aes(.iteration, .value, colour = .chain)) +
-    facet_wrap(~.variable, scales = "free_y") +
-    annotate(
-        geom = "rect", 
-        colour = NA, 
-        xmin = 0,
-        xmax = dim(model_fit_3$fit)[1]/2, 
-        ymin = -Inf, 
-        ymax = Inf, 
-        alpha = 0.25, 
-    ) +
-    geom_line() +
-    labs(
-        x = "Iteration",
-        y = "Sample value",
-        colour = "Chain"
-    ) +
-    scale_color_d3() +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank(),
-        axis.text.x = element_text(size = 7)
-    )
 ```
@@
 ```{r}
-#| label: diagnostics-rhat
+#| label: fig-diagnostics-rhat
 #| echo: true
@@
 ```{r}
-#| label: diagnostics-neff
+#| label: fig-diagnostics-neff
 #| echo: true

### 2022-09-23 index.qmd
Fix random slopes and update manuscript

@@
 # set params and load target contents
-
 library(bayesplot)
@@
 
-# Questionnaires
+# Methods
+
+Data and materials are available at OSF (https://osf.io/hy984/), and code is available at GitHub (https://github.com/gongcastro/trajectories).
+
+## Questionnaires
 
@@
 
-# Items
+## Items
 
@@
 
-## Lexical frequency
+### Lexical frequency
 
@@
 
-## \# Phonemes
+### \# Phonemes
 
@@
 
-## Levenshtein
+### Levenshtein
 
@@
 
-## Frequency by \# Phonemes
+### Frequency by \# Phonemes
 
@@
 
-## Frequency by Levenshtein
+### Frequency by Levenshtein
 
@@
 
-## \# Phonemes by Levenshtein
+### \# Phonemes by Levenshtein
 
@@
 
-# Participants
+## Participants
 
@@
 
-## Age
+### Age
 
@@
 
-## Language profile
+### Language profile
 
@@
 
-## SES/parental education
+### SES/parental education
 
@@
     summary_rows(
-        columns = 2:7, 
+        columns = 1:7, 
         fns = list(Total = ~sum(., na.rm = TRUE)),
@@
 
-# Data analysis
-
-## Model comparison
+## Data analysis
+
+### Model comparison
 
@@
 
-## Model equation
+### Model equation
+
+```{mermaid}
+//| label: fig-causal-diagram
+//| fig-cap: Causal diagram
+flowchart LR
+  B(Age) --> A("P(Acquisition)")
+  C(Exposure) --> A
+  D(Frequency) --> A
+  E(Phonemes) --> A
+  F(Levenshtein) --> C
+  F(Levenshtein) --> A
+  A("P(Acquisition)") --> G[Understands/Says response]
+```
 
@@
 
-## Model prior
+### Model prior
 
@@
 
-### Prior predictions
+#### Prior predictions
 
@@
 
-## Model settings
+### Model settings
 
@@
 
-## R code (`brms`)
+### R code (`brms`)
 
@@
 
-### Model 0
+#### Model 0
 
@@
 
-### Model 1
+#### Model 1
 
@@
 
-### Model 2
+#### Model 2
 
@@
 
-### Model 3
+#### Model 3
 
@@
 ```
+
+#### Model 4
+
+```{r}
+print(model_fit_4$formula)
+```
 :::
 
-## Stan code
-
-Stan code generated by `brms::stancode`.
-
-::: panel-tabset
-
-### Model 0
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 1
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-vector[N] Z_1_6;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-vector[N_1] r_1_6;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-r_1_6 = r_1[, 6];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(b[5] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 6 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_b_5 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 2
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-vector[N] Z_1_6;
-vector[N] Z_1_7;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-vector[N_1] r_1_6;
-vector[N_1] r_1_7;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-r_1_6 = r_1[, 6];
-r_1_7 = r_1[, 7];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(b[5] | 0, 0.1);
-target += normal_lpdf(b[6] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 7 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_b_5 = normal_rng(0,0.1);
-real prior_b_6 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 3
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-vector[N] Z_1_6;
-vector[N] Z_1_7;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-vector[N_1] r_1_6;
-vector[N_1] r_1_7;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-r_1_6 = r_1[, 6];
-r_1_7 = r_1[, 7];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(b[5] | 0, 0.1);
-target += normal_lpdf(b[6] | 0, 0.1);
-target += normal_lpdf(b[7] | 0, 0.1);
-target += normal_lpdf(b[8] | 0, 0.1);
-target += normal_lpdf(b[9] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 7 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_b_5 = normal_rng(0,0.1);
-real prior_b_6 = normal_rng(0,0.1);
-real prior_b_7 = normal_rng(0,0.1);
-real prior_b_8 = normal_rng(0,0.1);
-real prior_b_9 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-:::
-
 # Results
@@
 #| warning: false
-#| fig-width: 7
+#| fig-width: 9
+#| fig-height: 7
 #| fig-cap: "Observed proportion of comprehension and production responses across ages.
@@
 
+#### Model 4
+
+```{r}
+#| label: fig-loo-diagnostics-4
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
+get_loo_plot(model_loos$model_fit_4)
+```
+
 :::
@@
 rope_coefs %>% 
+    mutate(study = str_replace(study, "Braginsky, Yurovsky, Marchman and Frank", "Braginsky et al."),
+           estimate = abs(estimate)) %>% 
     ggplot(aes(estimate, y = study)) +
@@
              smallest effect size of interest\n",
-            "Median = ", round(median(rope_coefs$estimate), 3)
+            "Median = ", round(median(abs(rope_coefs$estimate)), 3)
         ),
-        x = median(rope_coefs$estimate)-0.2,
+        x = median(abs(rope_coefs$estimate)-0.05),
         y = 4,
@@
         vjust = 1.2,
-        size = 4.5,
+        size = 3.5,
         colour = pal_d3()(4)[4]
     ) +
-    annotate(
-        geom = "curve",
-        x = median(rope_coefs$estimate)-0.2,
-        xend = median(rope_coefs$estimate)-0.01,
-        size = 1,
-        y = 3.6,
-        yend = 3.5,
-        colour = pal_d3()(4)[4],
-        arrow = arrow(length = unit(0.5, "cm"))
-    ) +
-    labs(
-        x = "Standard regression coefficient (logit scale)\n(Excluding intercepts and predictor age)",
+    labs(
+        x = "Absolute standard regression coefficient (logit scale)\n(Excluding intercepts and predictor age)",
         y = "Study"
@@
 
-### Participant
-
-```{r}
-#| label: fig-results-random-id
+### Participant-level effects
+
+::: panel-tabset
+
+#### Intercept
+
+```{r}
+#| label: fig-results-random-id-intercept
 #| echo: true
@@
 
-### Item
+#### Age slope
+
+```{r}
+#| label: fig-results-random-id-age
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_age <- ranef(model_fit_3)$id[,,"age_std"]
+
+re_age %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_3$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Frequency slope
+
+```{r}
+#| label: fig-results-random-id-freq
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of lexical frequency by participant"
+re_freq <- ranef(model_fit_3)$id[,,"freq_std"]
+
+re_freq %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_3$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### # Phonemes slope
+
+```{r}
+#| label: fig-results-random-id-nphon
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the number of phonemes by participant"
+re_n_phon <- ranef(model_fit_3)$id[,,"n_phon_std"]
+
+re_n_phon %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_3$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### DoE slope
+
+```{r}
+#| label: fig-results-random-id-doe
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by language degree of exposure"
+# re_doe <- ranef(model_fit_3)$id[,,"doe_std"] 
+# 
+# re_doe %>% 
+#     as_tibble() %>% 
+#     clean_names() %>% 
+#     mutate(id = unique(model_fit_3$data$id)) %>% 
+#     relocate(id) %>% 
+#     ggplot() +
+#     aes(
+#         x = estimate,
+#         y = reorder(id, estimate),
+#         xmin = q2_5,
+#         xmax = q97_5
+#     ) +
+#     geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+#     geom_line(aes(group = 1), size = 1) +
+#     geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+#     labs(
+#         x = "Posterior random slope estimate",
+#         y = "Participant", 
+#     ) +
+#     theme(
+#         legend.position = "none",
+#         axis.text.y = element_blank(),
+#         axis.ticks.y = element_blank(),
+#         panel.grid = element_line(linetype = "dotted"),
+#         panel.grid.major.y = element_blank(),
+#         panel.grid.minor.y = element_blank()
+#     )
+```
+
+#### Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of phonological similarity by participant"
+# re_lv <- ranef(model_fit_3)$id[,,"lv_std"]
+#
+# re_lv %>%
+#     as_tibble() %>%
+#     clean_names() %>%
+#     mutate(id = unique(model_fit_3$data$id)) %>%
+#     relocate(id) %>%
+#     ggplot() +
+#     aes(
+#         x = estimate,
+#         y = reorder(id, estimate),
+#         xmin = q2_5,
+#         xmax = q97_5
+#     ) +
+#     geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+#     geom_line(aes(group = 1), size = 1) +
+#     geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+#     labs(
+#         x = "Posterior random slope estimate",
+#         y = "Participant",
+#     ) +
+#     theme(
+#         legend.position = "none",
+#         axis.text.y = element_blank(),
+#         axis.ticks.y = element_blank(),
+#         panel.grid = element_line(linetype = "dotted"),
+#         panel.grid.major.y = element_blank(),
+#         panel.grid.minor.y = element_blank()
+#     )
+```
+
+#### Doe-by-Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-doe-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+# re_doe_lv <- ranef(model_fit_3)$id[,,"doe_std_lv_std"]
+# 
+# re_lv %>%
+#     as_tibble() %>%
+#     clean_names() %>%
+#     mutate(id = unique(model_fit_3$data$id)) %>%
+#     relocate(id) %>%
+#     ggplot() +
+#     aes(
+#         x = estimate,
+#         y = reorder(id, estimate),
+#         xmin = q2_5,
+#         xmax = q97_5
+#     ) +
+#     geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+#     geom_line(aes(group = 1), size = 1) +
+#     geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+#     labs(
+#         x = "Posterior random slope estimate",
+#         y = "Participant",
+#     ) +
+#     theme(
+#         legend.position = "none",
+#         axis.text.y = element_blank(),
+#         axis.ticks.y = element_blank(),
+#         panel.grid = element_line(linetype = "dotted"),
+#         panel.grid.major.y = element_blank(),
+#         panel.grid.minor.y = element_blank()
+#     )
+```
+
+#### Age-by-DoE slope
+
+```{r}
+#| label: fig-results-random-id-age-doe
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
+re_age_doe <- ranef(model_fit_3)$id[,,"age_std:doe_std"]
+
+re_age_doe %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_3$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-age-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
+re_age_lv <- ranef(model_fit_3)$id[,,"age_std:lv_std"]
+
+re_age_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_3$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-DoE-by-Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-age-doe-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-language degree of exposure-by-phonological similarity interaction by participant"
+re_age_doe_lv <- ranef(model_fit_3)$id[,,"age_std:doe_std:lv_std"]
+
+re_age_doe_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_3$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Correlations
+
+```{r}
+#| label: fig-results-random-id-correlations
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior correlations between random effects by participant"
+posterior_description_re %>% 
+    select(parameter, median) %>% 
+    filter(str_detect(parameter, "cor_id__")) %>% 
+    mutate(parameter = str_remove_all(parameter, "cor_id__")) %>%  
+    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
+    mutate(
+        across(
+            c(var_1, var_2), 
+            ~factor(
+                .,
+                levels = c("Intercept", "age_std", "freq_std", "n_phon_std", "doe_std", "lv_std", "age_std:doe_std", "age_std:lv_std"),
+                labels = c("Intercept", "Age", "Frequency", "Phonemes", "DoE", "Levenshtein", "Age \u00d7 DoE", "Age \u00d7 Levenshtein")
+            )
+        )
+    ) %>%
+    filter(var_1 != var_2) %>% 
+    ggplot() +
+    aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median)) +
+    geom_point(shape = 15, size = 30) +
+    geom_text(colour = "black") +
+    scale_color_distiller(palette = "RdYlBu") +
+    labs(
+        colour = "Correlation",
+        size = "Correlation"
+    ) +
+    theme(
+        axis.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+:::
+
+### Item-level effects
+
+::: panel-tabset
+
+#### Intercept
 
@@
 #| fig-cap: "Posterior probability of acquisition by translation equivalent"
-re_te <- ranef(model_fit_3)$te[,,1] %>% 
+re_te <- ranef(model_fit_3)$te[,,"Intercept"] %>% 
     as_tibble() %>% 
@@
 ```
+
+#### Age slope
+
+```{r}
+#| label: fig-results-random-te-age
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by translation equivalent"
+re_age <- ranef(model_fit_3)$te[,,"age_std"]
+
+re_age %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_3$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Frequency slope
+
+```{r}
+#| label: fig-results-random-te-freq
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of lexical frequency by translation equivalent"
+re_freq <- ranef(model_fit_3)$te[,,"freq_std"]
+
+re_freq %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_3$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### # Phonemes slope
+
+```{r}
+#| label: fig-results-random-te-nphon
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
+re_n_phon <- ranef(model_fit_3)$te[,,"n_phon_std"]
+
+re_n_phon %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_3$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### DoE slope
+
+```{r}
+#| label: fig-results-random-te-doe
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+# re_doe <- ranef(model_fit_3)$te[,,"doe_std"]
+# 
+# re_doe %>% 
+#     as_tibble() %>% 
+#     clean_names() %>% 
+#     mutate(te = unique(model_fit_3$data$te)) %>% 
+#     relocate(te) %>% 
+#     ggplot() +
+#     aes(
+#         x = estimate,
+#         y = reorder(te, estimate),
+#         xmin = q2_5,
+#         xmax = q97_5
+#     ) +
+#     geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+#     geom_line(aes(group = 1), size = 1) +
+#     geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+#     labs(
+#         x = "Posterior random slope estimate",
+#         y = "Participant", 
+#     ) +
+#     theme(
+#         legend.position = "none",
+#         axis.text.y = element_blank(),
+#         axis.ticks.y = element_blank(),
+#         panel.grid = element_line(linetype = "dotted"),
+#         panel.grid.major.y = element_blank(),
+#         panel.grid.minor.y = element_blank()
+#     )
+```
+
+#### Age-by-DoE slope
+
+```{r}
+#| label: fig-results-random-te-age-doe
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_age_doe <- ranef(model_fit_3)$te[,,"age_std:doe_std"]
+
+re_age_doe %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_3$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Correlations
+
+```{r}
+#| label: fig-results-random-te-correlations
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior correlations between random effects by translation equivalent"
+posterior_description_re %>% 
+    select(parameter, median) %>% 
+    filter(str_detect(parameter, "cor_te__")) %>% 
+    mutate(parameter = str_remove_all(parameter, "cor_te__")) %>%  
+    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
+    mutate(
+        across(
+            c(var_1, var_2), 
+            ~factor(
+                .,
+                levels = c("Intercept", "age_std", "freq_std", "n_phon_std", "doe_std", "lv_std", "age_std:doe_std", "age_std:lv_std"),
+                labels = c("Intercept", "Age", "Frequency", "Phonemes", "DoE", "Levenshtein", "Age \u00d7 DoE", "Age \u00d7 Levenshtein")
+            )
+        )
+    ) %>%
+    filter(var_1 != var_2) %>% 
+    ggplot() +
+    aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median)) +
+    geom_point(shape = 15, size = 30) +
+    geom_text(colour = "black") +
+    scale_color_distiller(palette = "RdYlBu") +
+    labs(
+        colour = "Correlation",
+        size = "Correlation"
+    ) +
+    theme(
+        axis.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+:::
 
@@
 #         ymax = Inf, 
-#         alpha = 0.25, 
+#         alpha = 0.25
 #     ) +
@@
 
+# Discussion
+
+## Summary of the study
+
+## Summary of the results
+
+## Results in contrast with the literature
+
+## Mechanistic explanations
+
+## Limitations
+
+## Further steps
+
 # Appendix

### 2022-09-23 manuscript.qmd
Fix random slopes and update manuscript

---
title: "The role of cross-linguistic lexical similarity on bilingual word acquisition"
editor: source
author:
    - name: Gonzalo Garcia-Castro
      corresponding: true
      affiliations:
        - name: Center for Brain and Cognition, Universitat Pompeu Fabra
    - name: Daniela S. Ávila-Varela
      affiliations:
        - name: Center for Brain and Cognition, Universitat Pompeu Fabra
    - name: Ignacio Castillejo
      affiliations:
        - name: Departamento de Psicología, Universidad Autónoma de Madrid
    - name: Núria Sebastian-Galles
      affiliations:
        - name: Center for Brain and Cognition, Universitat Pompeu Fabra

pagetitle: Cross-linguistic similarity and  word acquisition
abstract: |
    Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 10 to 36 months.
thanks: |
    This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award]. Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019). The authors declare no conflicts of interest with regard to the funding source of this study. We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and Ege E. Özer for their help-ful feedback. We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support. We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants. We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises. We also thank all families and infants who participated in the experiments. Data collection was half-way when the COVID-19 pandemic started. We would like to pay special tribute to the families that collaborated with us under these difficult circumstances.
keywords: lexical acquisition, vocabulary, bilingualism
toc: false
number-sections: true
fig-dpi: 1000
csl: apa.csl
bibliography: manuscript/references.bib
google-scholar: true
format:
  pdf:
    pdf-engine: xelatex
    geometry:
      - top=30mm
      - left=20mm
      - heightrounded
    keep-tex: true
    linestretch: 1.25
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
    echo = FALSE,
    fig.align = "center",
    message = FALSE,
    warning = FALSE,
    cache.extra = knitr::rand_seed,
    out.width = "100%",
    dpi = 800,
    results = "asis"
)
```

```{r}
#| label: packages
#| echo: false
#| message: false
#| warning: false
#| include: false
# load packages
library(brms) 
library(conflicted)
library(dplyr) 
library(forcats)
library(ggplot2) 
library(ggsci)
library(gt) 
library(here) 
library(janitor)
library(knitr)
library(lubridate)
library(multilex)
library(patchwork) 
library(papaja) 
library(purrr)
library(scales) 
library(stringr)
library(tidybayes) 
library(tidyr) 
library(tibble)

# load and process objects
tar_load(participants)
tar_load(items)
tar_load(multilex_data)
tar_load(model_log_liks)
tar_load(model_loos)
tar_load(model_fit_1)
tar_load(rope_coefs)
tar_load(rope_interval)
tar_load(df)
tar_load(posterior_description)

participants <- participants %>% 
    left_join(select(multilex_data$logs, id, time, edu_parent))
items <- items %>% 
    left_join(select(pool, item, language, category, class))
# set global options
options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    gt.html_tag_check = FALSE,
    ggplot2.discrete.fill = ggsci::pal_d3()(4),
    ggplot2.discrete.colour = ggsci::pal_d3()(4),
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

# set custom ggplot2 theme
theme_set(theme_custom())

# resolve namespace conflicts
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

# Introduction {#introduction}

The foundations of word learning are in place early in age: before the end of their first year of life infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental reports [e.g., @fenson2007macarthur]. During the last half of their second year, they acquire new words at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002children; @bergelson2020comprehension]. These early stages of lexical acquisition are characterised by substantial variation across children, reflected, for instance, on the variation on the number of words they know [i.e. vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, bates1994developmental].

Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages. @tardif2008baby collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing. Since birth, these infants had been learning English, Cantonese and Mandarin, respectively. The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*). These results were later extended by @frank2021variability to a wider diversity languages, also reporting that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages. Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.

There is evidence that bilinguals know, on average, less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated. For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English. Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development. Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual]. While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals. What role could linguistic distance play during early vocabulary growth?

For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language. Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees. Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English (one Roman, the other Germanic). For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates). It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*). If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.

@floccia2018introduction provided evidence in line with this claim. The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language. The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English. Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions. The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other [@levenshtein1966binary]. The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes. Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair. For example, English-Dutch bilinguals (22.14% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% phonological similarity) were able to produce in Mandarin.

Floccia et al. pointed to *parallel activation* as the main mechanism underpinning their results. The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension, and that this phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain]. One of the clearest examples of parallel activation was provided by @costa2000cognate. In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish. In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trials labels were non-cognates (*mesa*-*taula*, translations of *table*). Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. Importantly, Spanish monolinguals--unfamiliar with the Catalan translations of the Spanish words they uttered--did not show this effect. These results suggest that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures. Several subsequent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical]. Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes. A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.

In this study, we hypothesise that cognates are acquired earlier in age than non-cognates, which might explain why children learning languages that share many cognates show larger vocabulary sizes. Evidence supporting this claim is scarce. @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates. @schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average. The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth. On the other hand, Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly. In their discussion the authors state the following (pp. 70):

> "This  finding  also  provides  support  to  the  proposal  that  the  cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life."

We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results. First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists. By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs. non-cognates was lost. Second, all participants were aged ~24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.

More recently, @mitchell2022cognates addressed this issue overcoming some of these pitfalls. Using a larger, longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages. They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates. The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty). These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar: word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.

As @mitchell2022cognates note in their discussion, two competing hypotheses remain untested. The facilitative effect of cognateness can be explained by, at least, two alternative, arguably incompatible, mechanisms. On the one hand, cognateness might operate as a facilitator of the acquisition of both forms of a translation pair, with every exposure to either form activating the other, and therefore strengthening the association between both and their referent. If this is the case, then cognateness should increase the probability of acquisition of the forms in both languages equally.

Another possibility is that cognateness plays a role in the acquisition of a translation pair only when at least one of the word-forms has been acquired. This account departs from the premise that, if the child is not previously familiar with one of the word forms, the acquisition of its translation in the other language cannot be facilitated by cognateness. This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair. Given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014much; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.

In line with this second hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al. reported was larger in the additional language vocabulary than in English vocabulary. Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language. Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.  If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.

In this study, we investigated the role of cognateness on the acquisition of translation equivalent, and whether such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair. Using an online vocabulary checklist--designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language. We then adopted an Item Response Theory (*IRT*) approach to model the probability of participants being reported by their parents to understand or understand *and* say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for other indices of word difficulty--lexical frequency and length (phonemes)-- and participant skill--age [see @kachergis2022toward for a similar approach].

# Methods {#methods}

Data and materials are available at OSF [https://osf.io/hy984/](https://osf.io/hy984/), and code is available at GitHub [https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories).

## Participants {#participants}

```{r}
#| label: participants
#| echo: false
#| message: false
#| warning: false
#| include: false
# load packages
n_participants <- length(unique(participants$id))
dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
```

We collected data from `r printnum(n_participants)` children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`. All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX. Families from the participant database of the Laboratori de Recerca en Infància of the Universitat Pompeu Fabra were contacted by e-mail or phone if their child were aged between `r floor(min(participants$age))` and `r floor(max(participants$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan. Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate. @tbl-participants-lp summarises the distribution of participants across ages and degrees of exposure (DoE) to Catalan.

```{r}
#| label: tbl-participants-lp
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan."
#| tbl-subcap: "Information about DoE was provided by families before filling the questionnaire. A 100% indicated that the participant was exclusively exposed to Catalan, and never to Spanish. A 0% indicates that the participant was not exposed to Catalan ever, and rather most of the time to Spanish. A 50% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100%, 0%, and 50% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants %>%
    mutate_at(
        vars(starts_with("doe_")),
        ~floor(.*10)/10
    ) %>%
    mutate(
        age = cut(
            floor(age),
            c(10, 14, 18, 22, 26, 30, 34, 36),
            include.lowest = TRUE
        ),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%", "25-50%", "50-75%", "75-100%"),
            include.lowest = TRUE
        ) %>% 
            as.character()
    ) %>%
    count(age, doe_catalan) %>%
    arrange(desc(doe_catalan)) %>%
    pivot_wider(names_from = age, values_from = n) %>%
    gt() %>%
    summary_rows(
        columns = 2:8,
        fns = list("Total" = ~sum(., na.rm = TRUE)),
        formatter = fmt_integer
    ) %>%
    sub_missing(everything(), everything(), missing_text = "--") %>%
    cols_label(doe_catalan = "DoE Catalan") %>%
    tab_spanner("Age (months)", 2:8) %>%
    tab_footnote(
        "This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10% of the time",
        cells_column_labels(columns = doe_catalan)
    ) %>%
    tab_style(
        cell_text(align = "left"),
        list(
            cells_body(doe_catalan),
            cells_column_labels(doe_catalan)
        )
    )
```

We used the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain. Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))` and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).

## Questionnaire {#questionnaire}

```{r}
#| label: items
#| echo: false
#| message: false
#| warning: false
#| include: false
n_categories <- nrow(distinct(multilex_data$pool, category))
n_te <- length(unique(multilex_data$pool$te))
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>%
    pull(n) %>% 
    set_names(c("catalan", "spanish"))
n_missing_freq <- multilex::pool %>% 
    filter(
        include,
        class %in% c("Noun", "Verb", "Adjective")
    ) %>% pull(frequency_zipf) %>% 
    is.na() %>%
    table() %>% 
    {.[2] / .[1]}
```

The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists. Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish. The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]`. Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping. In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*). We included items from a diverse sample of `r n_categories` semantic/functional categories.

For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default). Some families filled a long version of the vocabulary checklists (`r format(n_te, big.mark = ",")` translation equivalents; `r format(n_item_language["catalan"], big.mark = ",")` items in Catalan, `r format(n_item_language["spanish"], big.mark = ",")` items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish). These last families were randomly allocated into one of four different subsets of the complete list of items. These lists were designed so that each contained a representative sub-sample of the items from the complete list. Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists. Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions. For the analyses in the present study, we considered responses to words corresponding to nouns, verbs, and adjectives [@fourtassi2020growth]. We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).

We extracted four variables of interest for each word-form: (1) Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human]. This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points. Lexical frequencies were extracted from an adult subtitle-based database: SUBTLEX-CAT [@boada2020subtlex] for Catalan words and SUBTLEX-ESP [@cuetos2012subtlex] for Spanish words. Using child-based estimates of lexical frequency like CHILDES was not possible due to the low number of Catalan participants and tokens in the available corpora. Responses to words with missing lexical frequency score were excluded from analyses (`r percent(n_missing_freq, accuracy = 0.01)` of the total number of items). (2) Word length ($Phonemes$): number of phonemes in the phonological transcription of the word-form in International Phonological association format. (3) Phonological similarity ($Levenshtein$): normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary]. This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1. This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches]. (4) Participant age: number of months elapsed between participants' birth date and questionnaire completion. Finally, (5) participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to. For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%. @tbl-item-summary shows a summary of the distribution of the word-level predictors across languages in and questionnaire versions.

```{r}
#| label: tbl-item-summary
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Lexical properties of the word-forms and translationpairs included in the vocabulary checklist and our analyses. We report the number of items, mean (*M*), standard deviation (*SD*), and range of the lexical frequency (epxressed in Zipf scores), length (number of phonemes) of the word forms, and the phonological similarity (normalised Levenshtein similarity) with their translation equivalent. Statistics are reported separately for Catalan and Spanish, and for the different vocabulary checklists that participants were allocated to (A, B, C, D)."
items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))

items %>%
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq, n_phon, lv), 
        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_n) %>% 
    gt(groupname_col = "language") %>%
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
    fmt_percent(matches("lv_")) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_mean = md("Mean"),
        freq_sd = md("*SD*"),
        freq_min = md("Range"),
        n_phon_mean = md("Mean"),
        n_phon_sd = md("*SD*"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean"),
        lv_sd = md("*SD*"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    )
```

## Data analysis {#analysis}

```{r}
#| label: sample-sizes
#| echo: false
#| message: false
#| warning: false
#| include: false
n_obs <- nrow(df)
sum_obs <- df %>%
    mutate(te = factor(te)) %>% 
    count(te) %>% 
    summarise(across(n, lst(median, min, max), .names = "{.fn}"))
n_participants <- df %>%
    distinct(id, age) %>% 
    count(id) %>% 
    pull(n) %>% 
    table()

n_items <- ifelse(
    grepl("cat_", distinct(df, item)$item),
    "Catalan",
    "Spanish"
) %>% 
    table()

n_te <- length(unique(df$te))
```

We gathered `r format(n_obs, big.mark = ",")` observations, with each observation corresponding to a single response of one participant to a given item in the questionnaire). These observations correspond to `r sum(n_participants)` distinct participants (of which `r n_participants[1]` participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times), responding to `r format(sum(n_items), big.mark = ",")` distinct items (`r n_items["Catalan"]` in Catalan, `r n_items["Spanish"]` in Spanish, `r n_te` translation equivalents). Translation equivalents (TEs) received a median of `r printnum(sum_obs$median, digits = 0)` responses (*Min* = `r printnum(sum_obs$min, digits = 0)`, *Max* = `r printnum(sum_obs$max, digits = 0)`), both languages summed together.

We modelled the probability of answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model. This allowed us to simultaneously estimate the impact of item- and participant-level predictors on the probability of word comprehension and production. We included $Age$, $Frequency$, $Phonemes$, $DoE$, and $Levenshtein$ as fixed effects. We also included the $Age \times DoE \times Levenshtein$ three-way interaction, and its corresponding two-way interactions. We added participants and items as crossed random effects including all intercepts and slopes when appropriate [@barr2013random].

This model structure, including word- and participant-level properties to estimate vocabulary acquisition trajectories was discussed by @kachergis2022toward in the context of explanatory item response models. Our model differs from their proposal in that  1) while they modelled global trajectories exclusively, we modelled individual trajectories as well, 2) our outcome variable is ordinal in contrast to the binary outcome used in their study, 3) we propose a Bayesian approach. Our motivation for a Bayesian approach towards statistical inference is that (1) it allows us to incorporate previous domain knowledge to improve estimation accuracy, and (2) to quantify the uncertainty associated with the estimated parameters in our model [@mcelreath2018statistical].

The Bayesian approach implies using Bayes’ rule to assign a probability to possible parameter values (i.e., posterior distribution), based on previous knowledge about such probability distribution (prior distribution) and our data (likelihood). We used Markov chain Monte Carlo (MCMC) to approximate the posterior distribution via sampling. We run 2 sampling chains with 4,000 iterations each, including 2,000 warm-up iterations per chain. All $\hat{R}$ values were smaller than 1.1, indicating that the chains converged successfully in all cases [@gelman1992inference].

We compared a base model–-only including the intercept, $Age$, $Frequency$, $Phonemes$, and $DoE$ as fixed effects, and their corresponding random effects-–to increasingly more complex models, introducing predictors in the following order: $Age \times DoE$, $Levenshtein$, $DoE \times Levenshtein$, and finally $Age \times Levenshtein$ and $Age \times DoE \times Levenshtein$. We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO) [@vehtari2017practical]: we computed the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set. Due to computational constraints, we performed LOO based on a random sub-sample of `r dim(model_log_liks$model_fit_1)[1]` samples (out of `r format(dim(model_fit_1$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model. and selected the model with the smallest expected log posterior density (ELPD)^[Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with `r format(dim(model_log_liks$model_fit_1)[1], big.mark = ",")` posterior samples [@magnusson2019bayesian]].

Finally, we explored the posterior distribution of each parameter in the selected model. Following @kruschke2018bayesian, We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian]. We defined our *ROPE* from -`r round(median(abs(rope_coefs$estimate)), 3)` to +`r round(median(abs(rope_coefs$estimate)), 3)` [see @#appendix-rope for justification of the *ROPE*]. In the results section, we report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*). This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data. We also report the proportion of the posterior samples of that coefficient that fell within the *ROPE*. A 90% overlap between the 95% *HDI* of the coefficients' posterior distribution and the *ROPE* indicates that, given the data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero, that is, negligible.

Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@loo], and `tidybayes` [@tidybayes] R packages.

# Results {#results}

## Model comparison and selection {#comparison}

@tbl-results-loos shows the outputs of the *LOO-CV*. The extended model (`model_fit_1`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference. This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.

```{r}
#| label: tbl-results-loos
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
model_loos %>% 
    loo_compare() %>% 
    as.data.frame() %>%   
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    gt() %>% 
    tab_header(
        title = "",
        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation. The likelihood of each data point given a model fitted without including that same observation was estimated. The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model. *ELPD* values closer to zero indicate better predictive acuracy. We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*). The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_1`)")
    ) %>% 
    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>% 
    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>% 
    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>% 
    fmt_number(2:9) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("*SE*"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*")
    ) %>% 
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners()
    ) %>% 
    tab_style(
        cell_text(align = "left"),
        cells_title(groups = "subtitle")
    )
```

## Regression coefficients {#coefficients}

```{r}
#| label: tbl-results-fixed
#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
#| echo: false
#| message: false
#| warning: false
post_draws <- gather_draws(model_fit_1, `b_.*`, regex = TRUE)

# tidy predictor names
str_repl <- c(
    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
    "b_freq_std" = paste0("Frequency (+1 SD, ",  round(sd(df$freq), 2), " Zipf)"),
    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
    "b_doe_std" = paste0("DoE (+1 SD, ",  percent(sd(df$doe)), ")"),
    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein",
    "b_age_std:doe_std" = "Age \u00d7 DoE",
    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein"
)

post_draws_list <- split(posterior_description, posterior_description$parameter)

# summarise posterior draws
posterior_description %>% 
    mutate(
        parameter = factor(
            parameter,
            levels = names(str_repl),
            labels = str_repl
        ) %>% 
            as.character()
    ) %>%
    select(parameter, median, ci_high, ci_low, rope_percentage) %>%  
    gt() %>% 
    fmt_number(2:4) %>%
    fmt_percent(5) %>%
    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
    cols_label(
        parameter = "Parameter",
        median = "Median",
        ci_low = md("95% *HDI*"),
        rope_percentage = md("*ROPE* prob.")
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = 1:5)
    )
```

@tbl-results-fixed shows the summary of the posterior distribution of the fixed regression coefficients of the model. For interpretability, we supplement the summaries of each posterior regression coefficient with its transformation to the probability scale. The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly. This means that one logit difference is not necessarily translated to a unique value in the probability scale. For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months. The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial. We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e. derivative) of the logistic curve at the age at which most participants were acquiring a given word. This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].

$Age$ was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r printnum(post_draws_list[["b_age_std"]]$median, digits = 2)`, 95% *HDI* = [`r printnum(post_draws_list[["b_age_std"]]$ci_low, digits = 2)`, `r printnum(post_draws_list[["b_age_std"]]$ci_high, digits = 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month. Lexical frequency ($Frequency$) had a small, positive, but inconclusive on the probability of acquisition substantially, ($\beta$ = `r percent(post_draws_list[["b_freq_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_freq_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_freq_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_freq_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples overlapped with the *ROPE.* The 95% *HDI* of the regression coefficient of the number of phonemes ($Phonemes$) did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*. This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]). For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.

```{r}
#| label: fig-results-fixed
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-heigh: 7
#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model. Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis). Intervals show the 95% highest density interval (*HDI*) of each distribution. This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data. the mean and HDI limits are indicated below each distribution. Coefficients were transformed from the logit scale to the probability scale for interpretability. We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e. the derivative of the logistic function in its mid-point."
post_draws %>%
    mutate(
        .variable_name = factor(
            .variable, 
            levels = names(str_repl),
            labels = str_repl,
            ordered = TRUE
        )
    ) %>% 
    filter(!grepl("Intercept|sd", .variable)) %>%
    ggplot(aes(.value, fct_rev(.variable_name))) +
    annotate(
        geom = "rect",
        ymin = -Inf,
        ymax = Inf,
        xmin = rope_interval["lower"],
        xmax = rope_interval["upper"],
        colour = NA,
        alpha = 0.5,
        fill = "grey",
    ) +
    geom_vline(
        xintercept = 0,
        size = 1,
        colour = "grey"
    ) +
    stat_slab(
        aes(
            fill = stat(abs(x) < rope_interval["upper"]),
            colour = stat(abs(x) < rope_interval["upper"]),
        ),
        size = 0.25,
        # fill = pal_d3()(2)[2],
        position = position_nudge(y = 0.15)
    ) +
    geom_errorbar(
        data = post_draws %>% 
            filter(!grepl("Intercept|sd", .variable)) %>%
            median_hdi(
                .exclude = c(
                    ".chain", 
                    ".iteration",
                    ".draw",
                    ".row", 
                    ".variable_name"
                )
            ) %>% 
            mutate(
                .variable_name = factor(
                    .variable, 
                    levels = names(str_repl),
                    labels = str_repl,
                    ordered = TRUE
                )
            ),
        aes(xmin = .lower, xmax = .upper, x = .value),
        width = 0.15
    ) +
    geom_point(
        data = post_draws %>% 
            filter(!grepl("Intercept|sd", .variable)) %>%
            median_hdi(
                .exclude = c(
                    ".chain", 
                    ".iteration",
                    ".draw",
                    ".row", 
                    ".variable_name"
                )
            ) %>% 
            mutate(
                .variable_name = factor(
                    .variable, 
                    levels = names(str_repl),
                    labels = str_repl,
                    ordered = TRUE
                )
            ),
        size = 2
    ) +
    geom_text(
        data = post_draws %>% 
            filter(!grepl("Intercept|sd", .variable)) %>%
            mean_hdi(
                .exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")
            ) %>% 
            mutate(
                .variable_name = factor(
                    .variable, 
                    levels = names(str_repl),
                    labels = str_repl,
                    ordered = TRUE
                )
            ),
        aes(
            label = paste0(
                printnum(.value, digits = 2),
                " [", printnum(.lower, digits = 2), ", ",
                printnum(.upper, digits = 2), "]"
            )
        ),
        position = position_nudge(y = -0.25),
        size = 3
    ) +

    labs(
        x = "Coefficient estimate (logit scale)",
        y = "Variable", 
        fill = "Overlaps with ROPE",
        colour = "Overlaps with ROPE"
    ) +
    scale_x_continuous(breaks = seq(-0.5, 2, 0.2)) +
    scale_fill_manual(
        values = c("dodgerblue", "#b3d9ff"),
        labels = c("No", "Yes")
    ) +
    scale_colour_manual(
        values = c("dodgerblue", "#b3d9ff"),
        labels = c("No", "Yes")
    ) +
    theme(
        legend.position = "top",
        legend.justification = c(1, 1),
        axis.title.y = element_blank(),
        panel.grid.major.x = element_line(
            colour = "grey85",
            linetype = "dotted"
        ),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank()
    )

# annotate(
#     geom = "text",
#     label = "ROPE",
#     x = 0.3,
#     y = 10,
#     vjust = 1,
#     size = 5,
#     colour = pal_d3()(2)[2]
# ) +
# annotate(
#     geom = "curve",
#     arrow = arrow(length = unit(0.2, "cm")),
#     x = 0.3,
#     y = 9.5,
#     xend = 0.1,
#     yend = 9,
#     size = 1,
#     curvature = -0.25,
#     colour = pal_d3()(2)[2]
# ) +
```

The degree of exposure to the language ($DoE$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std"]]$ci_high, accuracy = 0.01)`]). All of the posterior samples of his regression excluded the *ROPE*. The impact of this predictor on the probability of acquisition was positive: for every 10% increase in exposure to the language the word belongs to, the participant was `r percent(post_draws_list[["b_doe_std"]]$median/(10*sd(df$doe)), accuracy = 0.01)` more likely to acquire it. The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*. However, the 95% *HDI* of the regression coefficient of the $Doe \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_doe_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the language it belonged to. Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially. This effect was negligible when exposure to both languages was balanced (e.g. 50-50%). Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.

The 95% *HDI* of the regression coefficient of the $Age \times DoE$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:doe_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*. This shows that the effect of the degree of exposure on the speed at which participants acquired word across ages was inconclusive. Give our data, it is difficult to tell whether this *DoE* increased or decreased the slope of the words' acquisition trajectories. The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages. Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_Std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of the degree of exposure to the language on the effect of phonological similarity did not differ substantially across ages.

```{r}
#| label: fig-results-marginal
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected posterior predictions. Posterior poreditions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient. These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model. Therefore, we generated 4,000 predictions for a series of combination of levels of interest. Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels. The X-axis indicates the age (in months) for which the prediction is generated. The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*). Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents). Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual). Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
nd <- expand.grid(
    n_phon_std = 0,
    freq_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe))
    # lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
)

m <- add_epred_draws(nd, model_fit_1, ndraws = NULL, re_formula = NA) %>% 
    mutate(
        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
        doe_std = factor(
            doe_std, 
            levels = unique(nd$doe_std), 
            labels = paste0("DoE: ", c(10, 50, 90), "%")
        )
        # lv_std = factor(
        #     lv_std, 
        #     levels = unique(nd$lv_std),
        #     labels = paste0(c(0, 50, 100), "%")
        # )
    ) %>% 
    filter(.category != "No") %>% 
    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(
        c(Understands, `Understands and Says`),
        names_to = ".category",
        values_to = ".epred"
    ) %>% 
    mutate(
        .category = str_replace_all(
            .category,
            c(
                "Understands and Says" = "Production\n(Understands and Says))",
                "Understands" = "Comprehension\n(Understands)"
            )
        )
    )

d <- df %>%
    mutate(
        doe_std = case_when(
            between(doe, 0.00, 0.20) ~ "DoE: 10%",
            between(doe, 0.40, 0.60) ~ "DoE: 50%",
            between(doe, 0.80, 1.00) ~ "DoE: 90%"
        ) %>% 
            factor(., levels = paste0("DoE: ", percent(c(0.1, 0.5, 0.9)))),
        # lv_std = case_when(
        #     between(lv, 0.00, 0.20) ~ "0%",
        #     between(lv, 0.40, 0.60) ~ "50%",
        #     between(lv, 0.80, 1.00) ~ "100%"
        # ) %>% 
        #     factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
        age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
            cut(
                .,
                breaks = ((seq(10, 38, 2)-mean(df$age))/(sd(df$age))),
                labels = ((seq(10, 36, 2)-mean(df$age))/(sd(df$age))),
                include.lowest = TRUE
            ) %>% 
            as.character() %>% 
            as.numeric(),
        Understands = response %in% c("Understands", "Understands and Says"),
        `Understands and Says` = response %in% "Understands and Says"
    ) %>% 
    drop_na(doe_std) %>% 
    group_by(te, age_std, doe_std) %>% 
    summarise(
        Understands = mean(Understands),
        `Understands and Says` = mean(`Understands and Says`),
        .groups = "drop"
    ) %>% 
    pivot_longer(
        c(Understands, `Understands and Says`),
        names_to = ".category",
        values_to = ".epred"
    ) %>% 
    mutate(
        .category = str_replace_all(
            .category,
            c(
                "Understands and Says" = "Production\n(Understands and Says))",
                "Understands" = "Comprehension\n(Understands)"
            )
        )
    )

m %>% 
    ggplot() +
    aes(
        x = age_std, 
        y = .epred, 
        colour = doe_std,
        fill = doe_std
    ) +
    facet_grid(~.category) +
    stat_lineribbon(
        size = 1,
        alpha = 0.5,
        .width = 0.95, 
        colour = NA,
        point_interval = mean_hdi
    ) +
    stat_summary(
        data = d,
        fun.data = "mean_se",
        geom = "pointrange",
        size = 0.25
    ) +
    # geom_point(
    #     data = d,
    #     position = position_dodge(width = 0.25),
    #     size = 0.5,
    #     alpha = 0.5
    # ) +
    stat_summary(fun = "mean", geom = "line", size = 1) +
    scale_x_continuous(
        breaks = (seq(10, 36, 2)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 2)
    ) +
    scale_color_d3() +
    labs(
        x = "Age (months)", 
        y = "Posterior probability of acquisition\n(Linear prediction)",
        colour = "Degree of exposure",
        fill = "Degree of exposure"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

# Discussion {#discussion}

## Summary of the study

This study explored how word acquisition was affected by cognateness in bilingual toddlers. We collected vocabulary data from children learning Catalan and/or Spanish in differing degrees of exposure to each, and then measured the probability of each word being acquired at different ages as a function of (1) participants' exposure to the language it belongs to, and (2) its phonological overlap with its translation equivalent in the other language. We build a multilevel Bayesian model that estimated the effect of both predictors and their interaction, adjusting for other variables known to be relevant during the acquisition of words: participants' age, lexical frequency, and length in phonemes.

## Summary of the results

## Results in contrast with the literature

## Mechanistic explanations

## Limitations

## Further steps

# References {#references}

::: {#refs}
:::

#  Appendix {#appendix}

## Appendix 1: Words {#appendix-words}

```{r}
#| lab: tbl-appendix-words
#| tbl-cap: "Words included in the analysis"
#| echo: false
#| message: false
#| warning: false
#| eval: false
items %>% 
    left_join(distinct(multilex_data$pool, item, te, label)) %>% 
    mutate(
        language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish"),
        label = paste0(label, " /", ipa_flat, "/"),
        te = as.integer(te),
        n_phon = as.integer(n_phon)
    ) %>% 
    select(language, te, label, freq, n_phon, lv) %>% 
    pivot_wider(
        id_cols = c(te,lv),
        names_from = language,
        values_from = c(label, freq, n_phon), 
        values_fn = first
    ) %>% 
    clean_names() %>% 
    relocate(te, ends_with("catalan"), ends_with("spanish"), lv) %>% 
    arrange(label_catalan) %>% 
    select(-te) %>% 
    gt() %>% 
    fmt_number(starts_with("freq")) %>% 
    fmt_integer(starts_with("n_phon")) %>% 
    fmt_percent(lv) %>% 
    tab_spanner("Catalan", ends_with("catalan")) %>%
    tab_spanner("Spanish", ends_with("spanish")) %>%
    cols_label(
        label_catalan = "Word-form /IPA/",
        freq_catalan = "Frequency",
        n_phon_catalan = "Phonemes",
        label_spanish = "Word-form /IPA/",
        freq_spanish = "Frequency",
        n_phon_spanish = "Phonemes",
        lv = "Levenshtein"
    ) 
```

## Appendix 2: *ROPE* specification {#appendix-words}

We decided to draw the *ROPE* interval around zero so that it covered values smaller the absolute median of those relevant by similar previous studies. After consulting the regression tables of @mitchell2022cognates, @braginsky2019consistency, and @laing2022phonological, we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: `r round(median(abs(rope_coefs$estimate)), 3)`, in the logit scale.

### 2022-09-23 manuscript.tex
Fix random slopes and update manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=30mm,left=20mm,heightrounded]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
  pdfauthor={Gonzalo Garcia-Castro; Daniela S. Ávila-Varela; Ignacio Castillejo; Núria Sebastian-Galles},
  pdfkeywords={lexical acquisition, vocabulary, bilingualism},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The role of cross-linguistic lexical similarity on bilingual word
acquisition\thanks{This research was supported by grants from the
Spanish Ministerio de Ciencia, Innovación y Universidades
(PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government
{[}ICREA (Catalan Institution for Research and Advanced Studies)
Academia 2019 award{]}. Gonzalo Garcia-Castro was supported by a
fellowship of the Spanish Ministerio de Ciencia, Innovación y
Universidades (FPI 2019). The authors declare no conflicts of interest
with regard to the funding source of this study. We are grateful to
Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and
Ege E. Özer for their help-ful feedback. We thank Xavier Mayoral, Silvia
Blanch, and Cristina Cuadrado for their technical support. We also thank
Cristina Dominguez and Katia Pistrin for their efforts in recruiting
infants. We would like to thank the clinics Quirón and Sagrada Familia
that allowed us to recruit participants in their premises. We also thank
all families and infants who participated in the experiments. Data
collection was half-way when the COVID-19 pandemic started. We would
like to pay special tribute to the families that collaborated with us
under these difficult circumstances.}}
\author{Gonzalo Garcia-Castro \and Daniela S. Ávila-Varela \and Ignacio
Castillejo \and Núria Sebastian-Galles}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Bilinguals face the challenging task of learning words from languages
with overlapping phonologies. Floccia et al.~(2018) reported larger
vocabulary sizes for 24-month-old bilinguals that were learning
languages that shared a greater amount of cognates (e.g.,
English-Dutch). The mechanisms underlying this effect remain unknown. We
explore two compatible scenarios. First, we test whether cognates are
learnt earlier than non-cognates. This would account for the difference
in vocabulary size associated to the amount of shared cognates across
languages. Second, we explore the possibility that the word-forms of one
language interact with those form the other language, scaffolding the
acquisition of their translation equivalents when their phonologies
overlap. This mechanism, in line with the parallel activation account of
bilingual speech perception, would provide a plausible explanation to
why cognates are acquired ealier by bilinguals. We developed an online
tool to collect parental reports of receptive and productive
vocabularies from children learning Catalan and/or Spanish, and present
data on receptive and productive vocabulary of bilingual toddlers aged
10 to 36 months.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi

\setstretch{1.25}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The foundations of word learning are in place early in age: before the
end of their first year of life infants start directing their gaze to
some objects when hearing their labels, according to both experimental
data (Bergelson \& Swingley, 2012, 2015; Jusczyk \& Aslin, 1995; Tincoff
\& Jusczyk, 1999) and parental reports (e.g., Fenson et al., 2007).
During the last half of their second year, they acquire new words at an
increasingly fast rate (Bergelson, 2020; Bloom, 2002; Fenson et al.,
1994; Goldfield \& Reznick, 1990; Mayor \& Plunkett, 2011). These early
stages of lexical acquisition are characterised by substantial variation
across children, reflected, for instance, on the variation on the number
of words they know (i.e.~vocabulary size, Fenson et al., 1994; Frank et
al., 2021) or on the proportion of those words that fall into the
category of nouns, as opposed to verbs, adjectives, or function words
(e.g., Nelson, 1973, p. bates1994developmental).

Despite this variability, children's trajectories of vocabulary growth
seem quite stable across languages. Tardif et al. (2008) collected data
about the first ten words acquired by 10 to 16 month-old infants living
in the United States, Hong Kong, and Beijing. Since birth, these infants
had been learning English, Cantonese and Mandarin, respectively. The
authors found a common pattern across the three groups: their first ten
words referred to roughly the same concepts, namely relatives/caretakers
(\emph{daddy}, \emph{mommy}), social routines (\emph{bye}, \emph{uh-oh})
or animals (\emph{woof-woof}). These results were later extended by
Frank et al. (2021) to a wider diversity languages, also reporting that
such cross-language commonalities are stronger at earlier stages of
lexical acquisition, compared to later stages. Most of the literature on
early word acquisition, however, has been conducted on monolingual
children, and neglects the problem of how bilinguals--who represent a
substantial proportion of the population in most societies--acquire
words at early ages.

There is evidence that bilinguals know, on average, less words in each
of their languages than monolinguals, but also that both groups know a
similar amount of words when the two languages are aggregated. For
example, Hoff et al. (2012) found that English-Spanish bilingual
toddlers in South Florida knew less words in English than monolinguals,
who only learnt English. Both groups knew a similar total amount of
words when both English and Spanish vocabularies were counted together,
pointing to the importance of collecting data on both languages when
assessing bilinguals' communicative development. Other studies have
provided converging evidence that bilinguals know a similar--or even
larger--number of words than monolinguals, only when the languages are
aggregated (Gonzalez-Barrero et al., 2020; Oller \& Eilers, 2002; J.
Patterson et al., 2004; J. L. Patterson, 2004; Pearson et al., 1993;
Pearson \& Fernández, 1994; Petitto et al., 2001; Smithson et al.,
2014). While these studies have mostly relied on samples of bilingual
children learning two relatively distant languages, as it is the case of
English and Spanish, it is unclear whether children learning
typologically more similar languages also know less words in each of
their languages than monolinguals. What role could linguistic distance
play during early vocabulary growth?

For a given set of concepts, bilingual children may be exposed to two
distinct sets of word-forms--one in each language. Depending on the
linguistic distance between both languages, the two sets of words may
overlap in varying degrees. Particularly, when both languages are
typologically close, like Spanish and Catalan (both Roman languages),
they are more likely to share a large amount of cognates (i.e.,
form-similar translation equivalents) than two linguistically distant
languages, like Spanish and English (one Roman, the other Germanic). For
instance, in the presence of a door, a Spanish-Italian (or a
Spanish-Catalan) bilingual might hear \emph{puerta} and \emph{porta}
(cognates), whereas a Spanish-English bilingual might hear \emph{puerta}
and \emph{door} (non-cognates). It could be the case that mapping two
phonologically similar labels (cognates like \emph{puerta}-\emph{porta})
onto the same referent is easier than doing the same with two
phonologically dissimilar labels (non-cognates, like \emph{puerta} and
\emph{door}). If cognates are easier to acquire than non-cognates,
bilinguals learning a pair of languages that share a high proportion of
cognates should benefit more often from this facilitation effect than
those learning a pair of languages with a lower proportion of cognates,
and should therefore show larger vocabulary sizes.

Floccia et al. (2018) provided evidence in line with this claim. The
authors collected vocabulary data on word comprehension and production
from 372 24-month-old bilingual toddlers living in the United Kingdom
who were learning English and an additional language. The additional
language was one a pool of 13 typologically diverse languages: Bengali,
Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian,
Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The authors
calculated the average phonological similarity between the words in each
of these additional languages and their translation equivalents in
English. Phonological similarity was measured by computing the
Levenshtein distance between each cross-language pair of phonological
transcriptions. The Levenshtein distance is a metric that computes the
edit distance between two strings by counting the smallest number of
insertions, deletions and substitutions one of the strings has to go
through to become identical to the other (Levenshtein et al., 1966). The
resulting scores were then divided by the length of the longest string
to bound the similarity scores between 0 and 1, and then entered this
variable as a predictor as they modelled participants' vocabulary sizes.
Among other findings, the authors reported an increase in productive
vocabulary size in the additional language associated with an increase
in the average phonological similarity between the translation
equivalents of each language pair. For example, English-Dutch bilinguals
(22.14\% phonological similarity), were able to produce more Dutch words
than English-Mandarin bilinguals (1.97\% phonological similarity) were
able to produce in Mandarin.

Floccia et al.~pointed to \emph{parallel activation} as the main
mechanism underpinning their results. The parallel activation hypothesis
suggests that bilinguals activate both languages simultaneously during
speech production or comprehension, and that this phenomenon is the
result of the activation of lexical representations in both languages,
even when only one is in use during production (Costa et al., 2000;
Hoshino \& Kroll, 2008) or comprehension (Spivey \& Marian, 1999;
Thierry \& Wu, 2007). One of the clearest examples of parallel
activation was provided by Costa et al. (2000). In this study,
Catalan-Spanish monolingual and bilingual adults were asked to name
pictures of common objects in Spanish. In half of the trials, the object
labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre},
translations of \emph{tree}), whereas in the other half of the trials
labels were non-cognates (\emph{mesa}-\emph{taula}, translations of
\emph{table}). Bilinguals named cognate pictures faster than non-cognate
pictures, even after adjusting for the lexical frequency of the items.
Importantly, Spanish monolinguals--unfamiliar with the Catalan
translations of the Spanish words they uttered--did not show this
effect. These results suggest that bilinguals' Catalan phonology was
activated during the production of Spanish words, facilitating the
naming of cognate pictures. Several subsequent studies have also
provided similar evidence in comprehension in children (Poulin-Dubois et
al., 2013; e.g., Von Holzen \& Mani, 2012). Parallel activation is
therefore a plausible mechanism to account for Floccia et al.'s results:
cognates increase the amount of activation in both languages,
facilitating word acquisition, and ultimately leading to children
learning language pairs with a larger proportion of cognates are
predicted to show larger vocabulary sizes. A missing piece in this
account is how increase parallel activation might facilitate the
acquisition new words.

In this study, we hypothesise that cognates are acquired earlier in age
than non-cognates, which might explain why children learning languages
that share many cognates show larger vocabulary sizes. Evidence
supporting this claim is scarce. Bosch \& Ramon-Casas (2014) used
vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish
bilinguals aged 18 months, and found that cognates represented a larger
proportion of participant's vocabulary than non-cognates. Schelletter
(2002) reported a longitudinal single case of one English-German
bilingual who produced cognates earlier than non-cognates, on average.
The low sample size in these two studies makes it challenging to draw
strong conclusions about the effect of cognateness on vocabulary growth.
On the other hand, Floccia et al.'s estimates are statically more
reliable given their (much larger) sample size, but their study was not
aimed at testing the effect of cognateness on age of acquisition
directly. In their discussion the authors state the following (pp.~70):

\begin{quote}
``This finding also provides support to the proposal that the cognate
advantage is due to cognates being acquired before non-cognates in early
childhood (Costa et al., 2016), leading to an ease of processing later
in life.''
\end{quote}

We identify two main reasons why an earlier age of acquisition for
cognates than for non-cognates is an unwarranted conclusion from Floccia
et al.'s results. First, the response variable used was the proportion
of words each participant understood and/or produced (i.e., vocabulary
size), from the list of lexical items in the vocabulary checklists. By
aggregating the responses from all items into a single datum per child,
information about the acquisition status of cognates vs.~non-cognates
was lost. Second, all participants were aged \textasciitilde24 months,
meaning that even if the unaggregated responses to individual items were
included as response variable, the possible effect of cognateness could
only be interpreted as an increase or decrease in the likelihood of
participants at such age to have acquired each item, and not as an
increase or decrease in the age of acquisition of such item.

More recently, Mitchell et al. (2022) addressed this issue overcoming
some of these pitfalls. Using a larger, longitudinal sample of 47
16-to-30 month-old French-English bilinguals living in Canada, the
authors collected data on expressive vocabulary data in both languages.
They created two lists of translation equivalents: one made of 131
cognates, and one made of 406 non-cognates. The proportion of
translation equivalents that children were reportedly able to produce
was higher in the cognate lists than in the non-cognate list across
ages, even when both lists were matched by semantic category (furniture,
animals, food were similarity represented in both lists) and age of
acquisition norms (an index of word difficulty). These findings shed
some light on the ongoing exploration of why (if at all) bilinguals'
vocabulary size grows faster when both languages are phonologically more
similar: word-forms sharing more phonemes with their translation
equivalents (i.e., cognates) seem to be acquired faster.

As Mitchell et al. (2022) note in their discussion, two competing
hypotheses remain untested. The facilitative effect of cognateness can
be explained by, at least, two alternative, arguably incompatible,
mechanisms. On the one hand, cognateness might operate as a facilitator
of the acquisition of both forms of a translation pair, with every
exposure to either form activating the other, and therefore
strengthening the association between both and their referent. If this
is the case, then cognateness should increase the probability of
acquisition of the forms in both languages equally.

Another possibility is that cognateness plays a role in the acquisition
of a translation pair only when at least one of the word-forms has been
acquired. This account departs from the premise that, if the child is
not previously familiar with one of the word forms, the acquisition of
its translation in the other language cannot be facilitated by
cognateness. This implies that, if cognateness facilitates the
acquisition of a word form, it can only do so after the child has
acquired one of the word forms of the translation pair. Given that
children are more likely to acquire words from languages to which they
are exposed more often (Cattani et al., 2014; David \& Wei, 2008;
Thordardottir et al., 2006), the acquisition of words in the language of
lower exposure should, on average, be more susceptible to the effect of
cognateness.

In line with this second hypothesis, the vocabulary size associated with
linguistic similarity that Floccia et al.~reported was larger in the
additional language vocabulary than in English vocabulary. Most
participants in their sample were English-dominant, meaning that their
relative amount of exposure to English was larger than in the additional
language. Therefore, participants may have, on average, learnt the
English word-form of translation equivalents earlier than the word-form
in the additional language. If this is the case, then the acquisition of
English words by English-dominant participants would rarely benefit from
their cognate status (the other word-form is not available yet), while
the acquisition of words in the additional language would benefit from
their phonological similarity with the (available) English form.

In this study, we investigated the role of cognateness on the
acquisition of translation equivalent, and whether such effect is
conditional to the amount of time the child is exposed to the
corresponding language of each member of the translation pair. Using an
online vocabulary checklist--designed specifically for this study--we
collected data from a sample of children aged 10 to 35 months learning
Catalan and/or Spanish, with varying degrees of exposure to each
language. We then adopted an Item Response Theory (\emph{IRT}) approach
to model the probability of participants being reported by their parents
to understand or understand \emph{and} say each word in the checklist,
conditional to its cognate status in Catalan and Spanish and
participants' degree of exposure to the corresponding language, while
adjusting for other indices of word difficulty--lexical frequency and
length (phonemes)-- and participant skill--age (see Kachergis et al.,
2022 for a similar approach).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Data and materials are available at OSF \url{https://osf.io/hy984/}, and
code is available at GitHub
\url{https://github.com/gongcastro/trajectories}.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We collected data from 422 children from the Metropolitan Area of
Barcelona between 28th October, 2019 and 12th July, 2022. All families
gave informed consent before participating and this study was approved
by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from
Hospital del Mar (Barcelona, Spain), code XXXXXXXXX. Families from the
participant database of the Laboratori de Recerca en Infància of the
Universitat Pompeu Fabra were contacted by e-mail or phone if their
child were aged between 10 and 35 months, and had not been reported to
be exposed more than 10\% of the time to a language other than Spanish
or Catalan. Upon consent, families were sent a link to the questionnaire
via e-mail, which they filled from a computer, laptop, or mobile device
in a browser within the two week following the invitation to
participate. Table~\ref{tbl-participants-lp} summarises the distribution
of participants across ages and degrees of exposure (DoE) to Catalan.

\begin{table}

\caption{\label{tbl-participants-lp}Participant sample size by age and
degree of exposure to Catalan.}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-participants-lp-1}Information about DoE was provided by families before filling the
questionnaire. A 100\textbackslash\% indicated that the participant was exclusively
exposed to Catalan, and never to Spanish. A 0\textbackslash\% indicates that the
participant was not exposed to Catalan ever, and rather most of the time
to Spanish. A 50\textbackslash\% indicates that the participant was exposed to Catalan
and Spanish approximately half of the time each. The 100\textbackslash\%, 0\textbackslash\%, and
50\textbackslash\% would be traditionaly classified as Catalan monolingual, Spanish
monolingual, and Catalan-Spanish bilingual, respectively. For
illustration purposes this table, DoEs were binned into 25\textbackslash\%-wide bins,
and ages (in months) binned into 4 month-wide bins. Hyphens indicate
that no participants from that specific combination of age and DoE
filled the questionnaire. }

{\centering

\setlength{\LTpost}{0mm}
\begin{longtable}{l|rrrrrrrr}
\tabularnewline

\toprule
\multicolumn{1}{l}{} &  & \multicolumn{7}{c}{Age (months)} \\ 
\cmidrule(lr){3-9}
\multicolumn{1}{l}{} & DoE Catalan\textsuperscript{1} & [10,14] & (14,18] & (18,22] & (22,26] & (26,30] & (30,34] & (34,36] \\ 
\midrule
 & 75-100\% & 22 & 25 & 36 & 39 & 21 & 8 & – \\ 
 & 50-75\% & 12 & 13 & 27 & 38 & 21 & 6 & 8 \\ 
 & 25-50\% & 17 & 16 & 47 & 29 & 24 & 4 & 11 \\ 
 & 0-25\% & 15 & 12 & 21 & 19 & 9 & – & – \\ 
\midrule 
\midrule 
Total & — & $66$ & $66$ & $131$ & $125$ & $75$ & $18$ & $19$ \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
\textsuperscript{1}This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10\% of the time\\
\end{minipage}

}

\end{minipage}%

\end{table}

We used the highest educational attainment of parents or caretakers as a
proxy of participants' socio-economic status (SES), which families
self-reported in the questionnaire by filling two items asking for the
educational attainment of each parent or caretaker, with the following
available options: \emph{No education}, \emph{Primary},
\emph{Secondary}, \emph{Complementary}, \emph{Vocational}, and
\emph{University}, in line with the current educational system in Spain.
Most families reported university studies (403, 0.9549763), followed by
families were the highest educational attainment were vocational studies
(75, 18\%), complementary studies (6, 1\%), secondary education (7, 2\%,
0\%), no formal education (2, 0\% and primary education (1,
\textless1\%).

\hypertarget{questionnaire}{%
\subsection{Questionnaire}\label{questionnaire}}

The questionnaire was implemented on-line using the formR platform
(Arslan et al., 2020), and was structured in three blocks: a (1)
language questionnaire, a (2) demographic survey, and a (3) Catalan and
a Spanish vocabulary checklists. Vocabulary checklists followed a
similar structure as the Oxford Communicative Developmental Inventory
(Hamilton et al., 2000) and consisted in two lists of words: one in
Catalan and one in Spanish. The Catalan inventory contained 798 items
and the Spanish inventory contained 802. Items in one language were
translation equivalents of the items in the other language (e.g.,
whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the
word \emph{perro} was included in the Spanish inventory), roughly
following a one-to-one mapping. In case two translation equivalents were
possible for a given word, both were included as separate items (e.g.,
Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar}
and \emph{terminar}), or merged them into a single item (e.g., Spanish
\emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}). We
included items from a diverse sample of 27 semantic/functional
categories.

For each word in the vocabulary checklists, we asked parents to report
whether their child was able to understand it, understand \emph{and} say
it, or did not understand or say it (checked out by default). Some
families filled a long version of the vocabulary checklists (803
translation equivalents; 798 items in Catalan, 802 items in Spanish),
while others filled a shorter version (\textasciitilde400 translation
equivalents, \textasciitilde400 items in Catalan, \textasciitilde400
items in Spanish). These last families were randomly allocated into one
of four different subsets of the complete list of items. These lists
were designed so that each contained a representative sub-sample of the
items from the complete list. Semantic/functional categories with less
than 16 items--thus resulting in less than four items after dividing it
in four lists--were not divided in the short version of the
questionnaire: all of their items were included in the four lists.
Another subset of items that were part of the trial lists of some
experiments in the lab were also included in all versions. For the
analyses in the present study, we considered responses to words
corresponding to nouns, verbs, and adjectives (Fourtassi et al., 2020).
We excluded multi-word items (e.g., \emph{barrita de cereales} {[}cereal
bar{]}) and items that included more than one word-form (e.g.,
\emph{mono / mico}).

We extracted four variables of interest for each word-form: (1) Word
lexical frequency (\(Frequency\)): lexical frequency of the word in its
corresponding language, expressed as Zipf scores (Van Heuven et al.,
2014; Zipf, 1949). This variable ranges from 0 to 7, and follows and
approximates a normal distribution, with most values in a corpus ranging
from 3 to 5 points. Lexical frequencies were extracted from an adult
subtitle-based database: SUBTLEX-CAT (Boada et al., 2020) for Catalan
words and SUBTLEX-ESP (Cuetos et al., 2012) for Spanish words. Using
child-based estimates of lexical frequency like CHILDES was not possible
due to the low number of Catalan participants and tokens in the
available corpora. Responses to words with missing lexical frequency
score were excluded from analyses (1.22\% of the total number of items).
(2) Word length (\(Phonemes\)): number of phonemes in the phonological
transcription of the word-form in International Phonological association
format. (3) Phonological similarity (\(Levenshtein\)): normalised
Levenshtein similarity between the word-form and its translation in the
other language (Levenshtein et al., 1966). This score is calculated by
first calculating the Levenshtein distance between the two
transcriptions (number of insertions, deletions or replacements needed
for the shortest transcription to become identical to the longer
transcription), then dividing the resulting value by the length of the
longest transcription, and finally subtracting this value from 1. This
results in a proportion that indicates how much the two phonological
transcriptions of the translation equivalent are similar to each other,
ranging from 0\% (no similarity at all) to 100\% (both transcriptions
are identical) (see Floccia et al., 2018; Fourtassi et al., 2020; Laing,
2022 for similar approaches). (4) Participant age: number of months
elapsed between participants' birth date and questionnaire completion.
Finally, (5) participant degree of exposure to the words' language
(\(DoE\)): percentage of exposure to the language the word belonged to.
For example, for a participant with 90\% exposure to Catalan, and 10\%
to Spanish, the DoE of the Catalan word \emph{taula} would be 90\%, and
the DoE for the Spanish word \emph{mesa} would be 10\%.
Table~\ref{tbl-item-summary} shows a summary of the distribution of the
word-level predictors across languages in and questionnaire versions.

\hypertarget{tbl-item-summary}{}
\begin{longtable}{lrrrrrrrrrr}
\caption{\label{tbl-item-summary}Lexical properties of the word-forms and translationpairs included in
the vocabulary checklist and our analyses. We report the number of
items, mean (M), standard deviation (SD), and range of the lexical
frequency (epxressed in Zipf scores), length (number of phonemes) of the
word forms, and the phonological similarity (normalised Levenshtein
similarity) with their translation equivalent. Statistics are reported
separately for Catalan and Spanish, and for the different vocabulary
checklists that participants were allocated to (A, B, C, D). }\tabularnewline

\toprule
 &  & \multicolumn{3}{c}{Frequency (Zipf)} & \multicolumn{3}{c}{\# Phonemes} & \multicolumn{3}{c}{Phon. similarity (Levenshtein)} \\ 
\cmidrule(lr){3-5} \cmidrule(lr){6-8} \cmidrule(lr){9-11}
 & \emph{N} & Mean & \emph{SD} & Range & Mean & \emph{SD} & Range & Mean & \emph{SD} & Range \\ 
\midrule
\multicolumn{11}{l}{Catalan} \\ 
\midrule
A & 278 & $4.39$ & $0.63$ & $2.61$–$6.52$ & $5.28$ & $1.73$ & 2–14 & $37.32\%$ & $26.94\%$ & $0.00\%$–$100.00\%$ \\ 
B & 274 & $4.39$ & $0.69$ & $1.90$–$6.31$ & $5.20$ & $1.77$ & 2–14 & $35.96\%$ & $24.74\%$ & $0.00\%$–$100.00\%$ \\ 
C & 270 & $4.36$ & $0.65$ & $2.49$–$6.54$ & $5.46$ & $1.88$ & 2–14 & $36.51\%$ & $26.11\%$ & $0.00\%$–$100.00\%$ \\ 
D & 269 & $4.38$ & $0.61$ & $2.10$–$6.10$ & $5.29$ & $1.70$ & 2–14 & $36.33\%$ & $26.40\%$ & $0.00\%$–$100.00\%$ \\ 
\midrule
\multicolumn{11}{l}{Spanish} \\ 
\midrule
A & 277 & $4.25$ & $0.74$ & $2.28$–$6.76$ & $5.66$ & $1.63$ & 3–14 & $37.33\%$ & $26.89\%$ & $0.00\%$–$100.00\%$ \\ 
B & 281 & $4.28$ & $0.79$ & $2.16$–$6.14$ & $5.67$ & $1.68$ & 2–14 & $36.18\%$ & $24.95\%$ & $0.00\%$–$100.00\%$ \\ 
C & 278 & $4.24$ & $0.77$ & $1.68$–$6.26$ & $5.79$ & $1.79$ & 3–14 & $36.53\%$ & $26.24\%$ & $0.00\%$–$100.00\%$ \\ 
D & 273 & $4.24$ & $0.72$ & $1.98$–$6.14$ & $5.72$ & $1.68$ & 2–14 & $36.59\%$ & $26.55\%$ & $0.00\%$–$100.00\%$ \\ 
\bottomrule
\end{longtable}

\hypertarget{analysis}{%
\subsection{Data analysis}\label{analysis}}

We gathered 258,095 observations, with each observation corresponding to
a single response of one participant to a given item in the
questionnaire). These observations correspond to 395 distinct
participants (of which 343 participated once, 33 twice, 14 three times,
and 5 four times), responding to 1199 distinct items (597 in Catalan,
602 in Spanish, 602 translation equivalents). Translation equivalents
(TEs) received a median of 236 responses (\emph{Min} = 118, \emph{Max} =
942), both languages summed together.

We modelled the probability of answering each response category
(\emph{No} \textless{} \emph{Understands} \textless{} \emph{Understands
and Says}) using a Bayesian, mixed-effects ordinal regression model.
This allowed us to simultaneously estimate the impact of item- and
participant-level predictors on the probability of word comprehension
and production. We included \(Age\), \(Frequency\), \(Phonemes\),
\(DoE\), and \(Levenshtein\) as fixed effects. We also included the
\(Age \times DoE \times Levenshtein\) three-way interaction, and its
corresponding two-way interactions. We added participants and items as
crossed random effects including all intercepts and slopes when
appropriate (Barr et al., 2013).

This model structure, including word- and participant-level properties
to estimate vocabulary acquisition trajectories was discussed by
Kachergis et al. (2022) in the context of explanatory item response
models. Our model differs from their proposal in that 1) while they
modelled global trajectories exclusively, we modelled individual
trajectories as well, 2) our outcome variable is ordinal in contrast to
the binary outcome used in their study, 3) we propose a Bayesian
approach. Our motivation for a Bayesian approach towards statistical
inference is that (1) it allows us to incorporate previous domain
knowledge to improve estimation accuracy, and (2) to quantify the
uncertainty associated with the estimated parameters in our model
(McElreath, 2018).

The Bayesian approach implies using Bayes' rule to assign a probability
to possible parameter values (i.e., posterior distribution), based on
previous knowledge about such probability distribution (prior
distribution) and our data (likelihood). We used Markov chain Monte
Carlo (MCMC) to approximate the posterior distribution via sampling. We
run 2 sampling chains with 4,000 iterations each, including 2,000
warm-up iterations per chain. All \(\hat{R}\) values were smaller than
1.1, indicating that the chains converged successfully in all cases
(Gelman \& Rubin, 1992).

We compared a base model---only including the intercept, \(Age\),
\(Frequency\), \(Phonemes\), and \(DoE\) as fixed effects, and their
corresponding random effects-\/--to increasingly more complex models,
introducing predictors in the following order: \(Age \times DoE\),
\(Levenshtein\), \(DoE \times Levenshtein\), and finally
\(Age \times Levenshtein\) and \(Age \times DoE \times Levenshtein\). We
compared the predictive performance of the models using Bayesian
leave-one-out cross-validation (LOO) (Vehtari et al., 2017): we computed
the average likelihood of each observation after estimating the model's
parameters leaving that same observation out of the data set. Due to
computational constraints, we performed LOO based on a random sub-sample
of 700 samples (out of 2,000) of the posterior distribution of the fixed
effects of the model. and selected the model with the smallest expected
log posterior density (ELPD)\footnote{Due to the high computational cost
  associated with our large dataset, we used a sub-sampling approach for
  performing Bayesian LOO with 700 posterior samples (Magnusson et al.,
  2019)}.

Finally, we explored the posterior distribution of each parameter in the
selected model. Following Kruschke \& Liddell (2018), We specified a
region of practical equivalence (\emph{ROPE}) around the values of each
regression coefficient that we considered equivalent to zero (Kruschke
\& Liddell, 2018). We defined our \emph{ROPE} from -0.18 to +0.18 {[}see
@\#appendix-rope for justification of the \emph{ROPE}{]}. In the results
section, we report the mean of each coefficient's marginal posterior
distribution along with their 95\% highest density intervals
(\emph{HDI}). This interval indicates the narrowest range of values that
contain 95\% of the samples of the posterior distribution, and therefore
the true value of the coefficient with 95\% confidence, given our data.
We also report the proportion of the posterior samples of that
coefficient that fell within the \emph{ROPE}. A 90\% overlap between the
95\% \emph{HDI} of the coefficients' posterior distribution and the
\emph{ROPE} indicates that, given the data, there is a 90\% probability
that the true value of the coefficient falls within the \emph{ROPE}, and
is therefore equivalent to zero, that is, negligible.

Data processing and visualisation was done in R (R Core Team, 2020)
using the \texttt{tidyverse} family of packages (Wickham et al., 2019),
Bayesian modelling was done using the \texttt{brms} (Bürkner, 2017;
Carpenter et al., 2017), \texttt{loo} (Yao et al., 2017), and
\texttt{tidybayes} (Kay, 2021) R packages.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{comparison}{%
\subsection{Model comparison and selection}\label{comparison}}

Table~\ref{tbl-results-loos} shows the outputs of the \emph{LOO-CV}. The
extended model (\texttt{model\_fit\_1}), which included the three-way
interaction between \(Age\), \(DoE\), and \(Levenshtein\) showed the
best predictive performance, with a difference in \emph{ELPD} compared
to the other models several times larger than the standard error of such
difference. This indicates that, given the data, the predictions of the
extended model are confidently more accurate than those of the other
models.

\hypertarget{tbl-results-loos}{}
\setlength{\LTpost}{0mm}
\begin{longtable}{lrrrrrrrr}
\caption{\label{tbl-results-loos}Bayesian leave-one-out cross validation (LOO-CV) }\tabularnewline

\caption*{
{\large } \\ 
{\small The predictive accuracy of the models was compared using leave-one-out cross-validation. The likelihood of each data point given a model fitted without including that same observation was estimated. The resulting likelihoods of each model were summarised as its expected log-predicted density (\emph{ELPD}), which penalises the complexity of each model. \emph{ELPD} values closer to zero indicate better predictive acuracy. We report the \emph{ELPD}, effective number of parameters (\emph{p}), and information criterion (\emph{IC}) of each model along their associated standard errors (\emph{SE}). The last column indicates the difference in \emph{ELPD} between each model and the model with the best predictive accuracy (\texttt{model\_fit\_1})}
} \\ 
\toprule
 & \multicolumn{2}{c}{LOOELPD} & \multicolumn{2}{c}{LOOp} & \multicolumn{2}{c}{LOOIC} & \multicolumn{2}{c}{LOOdiff} \\ 
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
Model & \emph{ELPD} & \emph{SE} & \emph{p} & \emph{SE} & \emph{LOO-IC} & \emph{SE} & \emph{diff} & \emph{SE} \\ 
\midrule
model\_fit\_3 & $-244,590.04$ & $388.23$ & $1,688.79$ & $5.39$ & $489,180.08$ & $776.46$ & $0.00$ & $0.00$ \\ 
model\_fit\_0 & $-247,891.36$ & $353.08$ & $1,260.47$ & $3.24$ & $495,782.71$ & $706.16$ & $-3,301.32$ & $161.73$ \\ 
model\_fit\_2 & $-248,274.01$ & $353.95$ & $1,489.46$ & $3.65$ & $496,548.01$ & $707.91$ & $-3,683.97$ & $152.45$ \\ 
model\_fit\_1 & $-249,630.81$ & $358.40$ & $1,649.76$ & $4.03$ & $499,261.61$ & $716.79$ & $-5,040.77$ & $170.07$ \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Pareto-\emph{k} estimates of all models were acceptable (\emph{k} \textless{} 0.5)\\
\end{minipage}

\hypertarget{coefficients}{%
\subsection{Regression coefficients}\label{coefficients}}

\hypertarget{tbl-results-fixed}{}
\begin{longtable}{lrrr}
\caption{\label{tbl-results-fixed}Summary of the estimated posterior distribution of fixed regression
coefficients }\tabularnewline

\toprule
Parameter & Median & 95\% \emph{HDI} & \emph{ROPE} prob. \\ 
\midrule
Intercept (Comprehension and Production at 22 months) & $-0.85$ & [$-0.97$, $-0.72$] & $0.00\%$ \\ 
Intercept (Comprehension at 22 months) & $1.55$ & [$1.44$, $1.69$] & $0.00\%$ \\ 
Age (+1 SD, 5.72 months) & $1.32$ & [$1.17$, $1.47$] & $0.00\%$ \\ 
Frequency (+1 SD, 0.7 Zipf) & $0.16$ & [$0.07$, $0.24$] & $32.11\%$ \\ 
Phonemes (+1 SD, 1.75 phonemes) & $-0.20$ & [$-0.28$, $-0.13$] & $2.79\%$ \\ 
DoE (+1 SD, 30\%) & $0.89$ & [$0.87$, $0.91$] & $0.00\%$ \\ 
Levenshtein (+1 SD, 26\%) & $0.03$ & [$-0.09$, $0.13$] & $100.00\%$ \\ 
DoE × Levenshtein & $-0.20$ & [$-0.22$, $-0.19$] & $0.00\%$ \\ 
Age × DoE & $0.13$ & [$-0.02$, $0.28$] & $57.42\%$ \\ 
Age × Levenshtein & $0.07$ & [$0.03$, $0.11$] & $100.00\%$ \\ 
Age × DoE × Levenshtein & $-0.06$ & [$-0.09$, $-0.02$] & $100.00\%$ \\ 
\bottomrule
\end{longtable}

Table~\ref{tbl-results-fixed} shows the summary of the posterior
distribution of the fixed regression coefficients of the model. For
interpretability, we supplement the summaries of each posterior
regression coefficient with its transformation to the probability scale.
The resulting values correspond to the maximum difference in probability
of acquisition (\emph{Comprehension} or \emph{Comprehension and
Production}) that corresponds to a one standard deviation change in each
predictor\footnote{The logit and probability scales related
  non-linearly. This means that one logit difference is not necessarily
  translated to a unique value in the probability scale. For example,
  the probability of acquisition of a given word might increase in 5\%
  when age increases from 22 to 23 months, the probability of
  acquisition of the same word might only increase in 0.2\% when age
  increases from 34 to 36 months. The linear growth of the probability
  of acquisition differs along the logistic curve, and therefore
  deciding the age point at which to report the estimates of the
  regression coefficients in the probability scale is not trivial. We
  followed the suggestion by Gelman et al. (2020) and report the maximum
  value of such coefficient, which corresponds to the linear growth
  (i.e.~derivative) of the logistic curve at the age at which most
  participants were acquiring a given word. This value can be
  approximated by dividing the coefficient in the logit scale by four:
  \(\hat{\beta_j}/4\), where \(\hat{\beta_j}\) is the estimated mean of
  the posterior distribution of coefficient \(j\).}.

\(Age\) was the predictor with the largest effect on the probability of
acquisition (\(\beta\) = 1.32, 95\% \emph{HDI} = {[}1.17, 1.47{]}): at
the steepest point of the acquisition trajectory of an average
translation equivalent, the probability of acquisition increased 5.77\%
every month. Lexical frequency (\(Frequency\)) had a small, positive,
but inconclusive on the probability of acquisition substantially,
(\(\beta\) = 15.85\%, 95\% \emph{HDI} = {[}7.12\%, 23.62\%{]}): 32.11\%
of its posterior samples overlapped with the \emph{ROPE.} The 95\%
\emph{HDI} of the regression coefficient of the number of phonemes
(\(Phonemes\)) did exclude the \emph{ROPE}, although not by much: 2.79\%
of its posterior samples fell within the \emph{ROPE}. This suggests that
the number of phonemes in the word form had a small effect on the
probability of acquisition (\(\beta\) = -20.28\%, 95\% \emph{HDI} =
{[}-28.12\%, -12.95\%{]}). For every increase in phoneme length, the
probability of acquisition decreased in -2.90\%.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-results-fixed-1.pdf}

}

\caption{\label{fig-results-fixed}Marginal posterior distribution of the
regression coefficients of the fixed effects in the extended model.
Distributions show the estimated likelihood density of each value in the
parameter space (X-axis) of each coefficient (Y-axis). Intervals show
the 95\% highest density interval (\emph{HDI}) of each distribution.
This interval is the narrowest range of values that contains 95\% of the
distribution, and therefore is 95\% likely to contain the true value,
given the data. the mean and HDI limits are indicated below each
distribution. Coefficients were transformed from the logit scale to the
probability scale for interpretability. We used the divide-by-four rule
to get the maximum change in probability of correct response associated
with a unit increase in this variable, i.e.~the derivative of the
logistic function in its mid-point.}

\end{figure}

The degree of exposure to the language (\(DoE\)) had a strong effect on
the probability of acquisition (\(\beta\) = 89.26\%, 95\% \emph{HDI} =
{[}87.10\%, 91.34\%{]}). All of the posterior samples of his regression
excluded the \emph{ROPE}. The impact of this predictor on the
probability of acquisition was positive: for every 10\% increase in
exposure to the language the word belongs to, the participant was
29.76\% more likely to acquire it. The effect of phonological similarity
by itself (as indicated by the regression coefficient of the main effect
of \(Levenshtein\)) was equivalent to zero (\(\beta\) = 2.76\%, 95\%
\emph{HDI} = {[}-8.61\%, 12.88\%{]}), as 100.00\% of the posterior
samples of its regression coefficient fell into the \emph{ROPE}.
However, the 95\% \emph{HDI} of the regression coefficient of the
\(Doe \times Levenshtein\) interaction excluded the \emph{ROPE} entirely
(\(\beta\) = -20.20\%, 95\% \emph{HDI} = {[}-21.68\%, -18.83\%{]}),
suggesting that the effect of phonological similarity on a word's
probability of acquisition changed depending on participants' exposure
to the language it belonged to. Follow up analyses on this interaction
(see Figure~\ref{fig-results-marginal}) showed that, when exposure to
the language was low (e.g., 10\%), phonological similarity increased the
probability of acquisition substantially. This effect was negligible
when exposure to both languages was balanced (e.g.~50-50\%). Finally,
when exposure to the language was high (e.g., 90\%), this effect was
very close to zero, although with different direction, with phonological
less similar translation equivalents being more likely to be acquired
than phonologically more similar ones.

The 95\% \emph{HDI} of the regression coefficient of the
\(Age \times DoE\) interaction did not exclude the \emph{ROPE}
(\(\beta\) = 12.63\%, 95\% \emph{HDI} = {[}-1.50\%, 28.48\%{]}): 57.42\%
of its posterior samples fell within the \emph{ROPE}. This shows that
the effect of the degree of exposure on the speed at which participants
acquired word across ages was inconclusive. Give our data, it is
difficult to tell whether this \emph{DoE} increased or decreased the
slope of the words' acquisition trajectories. The 95\% \emph{HDI} of the
regression coefficient of the \(Age \times Levenshtein\) interaction
overlapped completely with the \emph{ROPE} (\(\beta\) = 6.59\%, 95\%
\emph{HDI} = {[}2.66\%, 10.84\%{]}), indicating that phonological
similarity did not speed-up the acquisition trajectories of the words
across ages. Finally, the 95\% \emph{HDI} of the regression coefficient
of the \(Age \times DoE \times Levenshtein\) interaction also overlapped
completely with the \emph{ROPE} (\(\beta\) = , 95\% \emph{HDI} =
{[}-9.32\%, -2.33\%{]}), suggesting that the impact of the degree of
exposure to the language on the effect of phonological similarity did
not differ substantially across ages.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-results-marginal-1.pdf}

}

\caption{\label{fig-results-marginal}Expected posterior predictions.
Posterior poreditions for the extended model were generated by computing
the probability acquisition that resulted from feeding the equation of
the regression model with values of the posterior distribution of each
coefficient. These values correspond to the 4,000 samples of the
posterior distribution that were drawn during the MCMC estimation of the
model. Therefore, we generated 4,000 predictions for a series of
combination of levels of interest. Each line and interval and in the
figure summarises the mean and 95\% \emph{HDI} of each combination of
levels. The X-axis indicates the age (in months) for which the
prediction is generated. The Y-axis indicates the predicted probability
of acquisition (\emph{Comprehension} or \emph{Comprehension and
Production}). Different colours indicate different levels of
phonological similarity (as indicated by the Levenshtein similarity
between pairs of translation equivalents). Finally, predictions are
presented separately for different degrees of exposure (10\%): little
exposure to the language (10\% DoE, traditionally classified as
monolingual), balanced exposure to both languages (50\%, traditionally
classified as bilingual), and high exposure to the language (90\%,
traditionally classified as monolingual). Predictions for
\emph{Comprehension} are show on top and predictions for
\emph{Comprehension and Production} are shown on the bottom.}

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{summary-of-the-study}{%
\subsection{Summary of the study}\label{summary-of-the-study}}

This study explored how word acquisition was affected by cognateness in
bilingual toddlers. We collected vocabulary data from children learning
Catalan and/or Spanish in differing degrees of exposure to each, and
then measured the probability of each word being acquired at different
ages as a function of (1) participants' exposure to the language it
belongs to, and (2) its phonological overlap with its translation
equivalent in the other language. We build a multilevel Bayesian model
that estimated the effect of both predictors and their interaction,
adjusting for other variables known to be relevant during the
acquisition of words: participants' age, lexical frequency, and length
in phonemes.

\hypertarget{summary-of-the-results}{%
\subsection{Summary of the results}\label{summary-of-the-results}}

\hypertarget{results-in-contrast-with-the-literature}{%
\subsection{Results in contrast with the
literature}\label{results-in-contrast-with-the-literature}}

\hypertarget{mechanistic-explanations}{%
\subsection{Mechanistic explanations}\label{mechanistic-explanations}}

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

\hypertarget{further-steps}{%
\subsection{Further steps}\label{further-steps}}

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%
Arslan, R. C., Walther, M. P., \& Tata, C. S. (2020). Formr: A study
framework allowing for automated feedback generation and complex
longitudinal experience-sampling studies using r. \emph{Behavior
Research Methods}, \emph{52}(1), 376--387.

\leavevmode\vadjust pre{\hypertarget{ref-barr2013random}{}}%
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random
effects structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3), 255--278.

\leavevmode\vadjust pre{\hypertarget{ref-bergelson2020comprehension}{}}%
Bergelson, E. (2020). The comprehension boost in early word learning:
Older infants are better learners. \emph{Child Development
Perspectives}, \emph{14}(3), 142--149.

\leavevmode\vadjust pre{\hypertarget{ref-bergelson2012months}{}}%
Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants
know the meanings of many common nouns. \emph{Proceedings of the
National Academy of Sciences}, \emph{109}(9), 3253--3258.

\leavevmode\vadjust pre{\hypertarget{ref-bergelson2015early}{}}%
Bergelson, E., \& Swingley, D. (2015). Early word comprehension in
infants: Replication and extension. \emph{Language Learning and
Development}, \emph{11}(4), 369--380.

\leavevmode\vadjust pre{\hypertarget{ref-bloom2002children}{}}%
Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT
press.

\leavevmode\vadjust pre{\hypertarget{ref-boada2020subtlex}{}}%
Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P. (2020).
SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for
catalan. \emph{Behavior Research Methods}, \emph{52}(1), 360--375.

\leavevmode\vadjust pre{\hypertarget{ref-bosch2014first}{}}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in
bilingual toddlers' expressive vocabulary: {Does} form similarity
matter? \emph{International Journal of Behavioral Development},
\emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\vadjust pre{\hypertarget{ref-braginsky2019consistency}{}}%
Braginsky, M., Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019).
Consistency and variability in children's word learning across
languages. \emph{Open Mind}, \emph{3}, 52--67.

\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.

\leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B.,
Betancourt, M., Brubaker, M., Guo, J., Li, P., \& Riddell, A. (2017).
Stan: A probabilistic programming language. \emph{Journal of Statistical
Software}, \emph{76}(1), 1--32.

\leavevmode\vadjust pre{\hypertarget{ref-cattani2014much}{}}%
Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis,
I., \& Floccia, C. (2014). How much exposure to english is necessary for
a bilingual toddler to perform like a monolingual peer in language
tests? \emph{International Journal of Language \& Communication
Disorders}, \emph{49}(6), 649--671.

\leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate
facilitation effect: Implications for models of lexical access.
\emph{Journal of Experimental Psychology: Learning, Memory, and
Cognition}, \emph{26}(5), 1283.

\leavevmode\vadjust pre{\hypertarget{ref-cuetos2012subtlex}{}}%
Cuetos, F., Glez-Nosti, M., Barbón, A., \& Brysbaert, M. (2012).
SUBTLEX-ESP: Spanish word frequencies based on film subtitles.
\emph{Psicol{ó}gica}, \emph{33}(2), 133--143.

\leavevmode\vadjust pre{\hypertarget{ref-david2008individual}{}}%
David, A., \& Wei, L. (2008). Individual differences in the lexical
development of french--english bilingual children. \emph{International
Journal of Bilingual Education and Bilingualism}, \emph{11}(5),
598--618.

\leavevmode\vadjust pre{\hypertarget{ref-fenson2007macarthur}{}}%
Fenson, L. et al. (2007). \emph{MacArthur-bates communicative
development inventories}. Paul H. Brookes Publishing Company Baltimore,
MD.

\leavevmode\vadjust pre{\hypertarget{ref-fenson1994variability}{}}%
Fenson, L., Dale, P. S., Reznick, J. S., Bates, E., Thal, D. J.,
Pethick, S. J., Tomasello, M., Mervis, C. B., \& Stiles, J. (1994).
Variability in early communicative development. \emph{Monographs of the
Society for Research in Child Development}, i--185.

\leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
Floccia, C., Sambrook, T. D., Delle Luche, C., Kwok, R., Goslin, J.,
White, L., Cattani, A., Sullivan, E., Abbot-Smith, K., Krott, A., et al.
(2018). I: INTRODUCTION. \emph{Monographs of the Society for Research in
Child Development}, \emph{83}(1), 7--29.

\leavevmode\vadjust pre{\hypertarget{ref-fourtassi2020growth}{}}%
Fourtassi, A., Bian, Y., \& Frank, M. C. (2020). The growth of
children's semantic and phonological networks: Insight from 10
languages. \emph{Cognitive Science}, \emph{44}(7), e12847.

\leavevmode\vadjust pre{\hypertarget{ref-frank2021variability}{}}%
Frank, M. C., Braginsky, M., Yurovsky, D., \& Marchman, V. A. (2021).
\emph{Variability and consistency in early language learning: The
wordbank project}. MIT Press.

\leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other
stories}. Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-gelman1992inference}{}}%
Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation
using multiple sequences. \emph{Statistical Science}, \emph{7}(4),
457--472.

\leavevmode\vadjust pre{\hypertarget{ref-goldfield1990early}{}}%
Goldfield, B. A., \& Reznick, J. S. (1990). Early lexical acquisition:
Rate, content, and the vocabulary spurt. \emph{Journal of Child
Language}, \emph{17}(1), 171--183.

\leavevmode\vadjust pre{\hypertarget{ref-gonzalez2020bilingual}{}}%
Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020).
\emph{Bilingual adjusted vocabulary: A developmentally-informed
bilingual vocabulary measure}.

\leavevmode\vadjust pre{\hypertarget{ref-hamilton2000infant}{}}%
Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary
development assessed with a british communicative development inventory.
\emph{Journal of Child Language}, \emph{27}(3), 689--705.

\leavevmode\vadjust pre{\hypertarget{ref-hoff2012dual}{}}%
Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M.
(2012). Dual language exposure and early bilingual development.
\emph{Journal of Child Language}, \emph{39}(1), 1.

\leavevmode\vadjust pre{\hypertarget{ref-hoshino2008cognate}{}}%
Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming:
Does cross-language activation survive a change of script?
\emph{Cognition}, \emph{106}(1), 501--511.

\leavevmode\vadjust pre{\hypertarget{ref-jusczyk1995infants}{}}%
Jusczyk, P. W., \& Aslin, R. N. (1995). Infants' detection of the sound
patterns of words in fluent speech. \emph{Cognitive Psychology},
\emph{29}(1), 1--23.

\leavevmode\vadjust pre{\hypertarget{ref-kachergis2022toward}{}}%
Kachergis, G., Marchman, V. A., \& Frank, M. C. (2022). Toward a
{``standard model''} of early language learning. \emph{Current
Directions in Psychological Science}, \emph{31}(1), 20--27.

\leavevmode\vadjust pre{\hypertarget{ref-tidybayes}{}}%
Kay, M. (2021). \emph{{tidybayes}: Tidy data and geoms for {Bayesian}
models}. \url{https://doi.org/10.5281/zenodo.1308151}

\leavevmode\vadjust pre{\hypertarget{ref-kruschke2018bayesian}{}}%
Kruschke, J., \& Liddell, T. (2018). The bayesian new statistics:
Hypothesis testing, estimation, meta-analysis, and planning from a
bayesian perspective. \emph{Psychonomic Bulletin and Review}, \emph{25},
178--206. \url{https://doi.org/10.3758/s13423-016-1221-4}

\leavevmode\vadjust pre{\hypertarget{ref-laing2022phonological}{}}%
Laing, C. E. (2022). \emph{Phonological networks and systematicity in
early lexical acquisition}.

\leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
Levenshtein, V. I. et al. (1966). Binary codes capable of correcting
deletions, insertions, and reversals. \emph{Soviet Physics Doklady},
\emph{10}, 707--710.

\leavevmode\vadjust pre{\hypertarget{ref-magnusson2019bayesian}{}}%
Magnusson, M., Andersen, M., Jonasson, J., \& Vehtari, A. (2019).
Bayesian leave-one-out cross-validation for large data.
\emph{International Conference on Machine Learning}, 4244--4253.

\leavevmode\vadjust pre{\hypertarget{ref-mayor2011statistical}{}}%
Mayor, J., \& Plunkett, K. (2011). A statistical estimate of infant and
toddler vocabulary size from CDI analysis. \emph{Developmental Science},
\emph{14}(4), 769--785.

\leavevmode\vadjust pre{\hypertarget{ref-mcelreath2018statistical}{}}%
McElreath, R. (2018). \emph{Statistical rethinking: A bayesian course
with examples in r and stan}. Chapman; Hall/CRC.

\leavevmode\vadjust pre{\hypertarget{ref-mitchell2022cognates}{}}%
Mitchell, L., Tsui, R. K., \& Byers-Heinlein, K. (2022). \emph{Cognates
are advantaged in early bilingual expressive vocabulary development}.

\leavevmode\vadjust pre{\hypertarget{ref-nelson1973structure}{}}%
Nelson, K. (1973). Structure and strategy in learning to talk.
\emph{Monographs of the Society for Research in Child Development},
1--135.

\leavevmode\vadjust pre{\hypertarget{ref-oller2002language}{}}%
Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in
bilingual children} (Vol. 2). Multilingual Matters.

\leavevmode\vadjust pre{\hypertarget{ref-patterson2004comparing}{}}%
Patterson, J. L. (2004). \emph{Comparing bilingual and monolingual
toddlers' expressive vocabulary size}.

\leavevmode\vadjust pre{\hypertarget{ref-patterson2004bilingual}{}}%
Patterson, J., Pearson, B., \& Goldstein, B. (2004). \emph{Bilingual
language development and disorders in spanish--english speakers}.

\leavevmode\vadjust pre{\hypertarget{ref-pearson1994patterns}{}}%
Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in
the lexical growth in two languages of bilingual infants and toddlers.
\emph{Language Learning}, \emph{44}(4), 617--653.

\leavevmode\vadjust pre{\hypertarget{ref-pearson1993lexical}{}}%
Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical
development in bilingual infants and toddlers: Comparison to monolingual
norms. \emph{Language Learning}, \emph{43}(1), 93--120.

\leavevmode\vadjust pre{\hypertarget{ref-petitto2001bilingual}{}}%
Petitto, L. A., Katerelos, M., Levy, B. G., Gauna, K., Tétreault, K., \&
Ferraro, V. (2001). Bilingual signed and spoken language acquisition
from birth: Implications for the mechanisms underlying early bilingual
language acquisition. \emph{Journal of Child Language}, \emph{28}(2),
453--496.

\leavevmode\vadjust pre{\hypertarget{ref-poulin2013lexical}{}}%
Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J.
(2013). Lexical access and vocabulary development in very young
bilinguals. \emph{International Journal of Bilingualism}, \emph{17}(1),
57--70.

\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
R Core Team. (2020). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-schelletter2002effect}{}}%
Schelletter, C. (2002). The effect of form similarity on bilingual
children's lexical development. \emph{Bilingualism: Language and
Cognition}, \emph{5}(2), 93--107.
\url{https://doi.org/10.1017/S1366728902000214}

\leavevmode\vadjust pre{\hypertarget{ref-smithson2014bilingualism}{}}%
Smithson, L., Paradis, J., \& Nicoladis, E. (2014). Bilingualism and
receptive vocabulary achievement: Could sociocultural context make a
difference? \emph{Bilingualism: Language and Cognition}, \emph{17}(4),
810--821.

\leavevmode\vadjust pre{\hypertarget{ref-spivey1999cross}{}}%
Spivey, M. J., \& Marian, V. (1999). Cross talk between native and
second languages: Partial activation of an irrelevant lexicon.
\emph{Psychological Science}, \emph{10}(3), 281--284.

\leavevmode\vadjust pre{\hypertarget{ref-tardif2008baby}{}}%
Tardif, T., Fletcher, P., Liang, W., Zhang, Z., Kaciroti, N., \&
Marchman, V. A. (2008). Baby's first 10 words. \emph{Developmental
Psychology}, \emph{44}(4), 929.

\leavevmode\vadjust pre{\hypertarget{ref-thierry2007brain}{}}%
Thierry, G., \& Wu, Y. J. (2007). Brain potentials reveal unconscious
translation during foreign-language comprehension. \emph{Proceedings of
the National Academy of Sciences}, \emph{104}(30), 12530--12535.

\leavevmode\vadjust pre{\hypertarget{ref-thordardottir2006bilingual}{}}%
Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R. (2006).
Bilingual assessment: Can overall proficiency be estimated from separate
measurement of two languages? \emph{Journal of Multilingual
Communication Disorders}, \emph{4}(1), 1--21.

\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999some}{}}%
Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word
comprehension in 6-month-olds. \emph{Psychological Science},
\emph{10}(2), 172--175.

\leavevmode\vadjust pre{\hypertarget{ref-van2014subtlex}{}}%
Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
SUBTLEX-UK: A new and improved word frequency database for british
english. \emph{Quarterly Journal of Experimental Psychology},
\emph{67}(6), 1176--1190.

\leavevmode\vadjust pre{\hypertarget{ref-vehtari2017practical}{}}%
Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model
evaluation using leave-one-out cross-validation and WAIC.
\emph{Statistics and Computing}, \emph{27}(5), 1413--1432.

\leavevmode\vadjust pre{\hypertarget{ref-von2012language}{}}%
Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access
in bilingual toddlers. \emph{Journal of Experimental Child Psychology},
\emph{113}(4), 569--586.

\leavevmode\vadjust pre{\hypertarget{ref-R-tidyverse}{}}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\leavevmode\vadjust pre{\hypertarget{ref-loo}{}}%
Yao, Y., Vehtari, A., Simpson, D., \& Gelman, A. (2017). Using stacking
to average bayesian predictive distributions. \emph{Bayesian Analysis}.
\url{https://doi.org/10.1214/17-BA1091}

\leavevmode\vadjust pre{\hypertarget{ref-zipf1949human}{}}%
Zipf, G. K. (1949). \emph{Human behavior and the principle of least
effort: An introd. To human ecology}.

\end{CSLReferences}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{appendix-words}{%
\subsection{Appendix 1: Words}\label{appendix-words}}

\hypertarget{appendix-words}{%
\subsection{\texorpdfstring{Appendix 2: \emph{ROPE}
specification}{Appendix 2: ROPE specification}}\label{appendix-words}}

We decided to draw the \emph{ROPE} interval around zero so that it
covered values smaller the absolute median of those relevant by similar
previous studies. After consulting the regression tables of Mitchell et
al. (2022), Braginsky et al. (2019), and Laing (2022), we selected the
largest regression coefficient that was interpreted as conclusive
evidence of the presence of an effect, as indicated by the authors, and
transformed it to the probability scale: 0.18, in the logit scale.

\end{document}

### 2022-09-23 manuscript/manuscript.Rmd
Fix random slopes and update manuscript

@@
 ---
-title             : "The role of cross-linguistic lexical similarity on bilingual word acquisition"
-shorttitle        : "Cross-linguistic similarity and  word acquisition"
-
+title: "The role of cross-linguistic lexical similarity on bilingual word acquisition"
 author:
-  - name          : "Gonzalo Garcia-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Daniela Avila-Varela"
-    affiliation   : "1"
-  - name          : "Ignacio Castillejo"
-    affiliation   : "2"
-  - name          : "Nuria Sebastian-Galles"
-    affiliation   : "1"
-
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain"
-  - id            : "2"
-    institution   : "Universidad Autónoma de Madrid"
-
+- name: Gonzalo Garcia-Castro
+  affiliation: '1'
+  corresponding: yes
+  address: Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain
+  email: gonzalo.garciadecastro@upf.edu
+- name: Daniela Avila-Varela
+  affiliation: '1'
+- name: Ignacio Castillejo
+  affiliation: '2'
+- name: Nuria Sebastian-Galles
+  affiliation: '1'
+shorttitle: Cross-linguistic similarity and  word acquisition
+output:
+  pdf_document: default
+  word_document:
+    reference_docx: manuscript_template.docx
 authornote: |
   Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
-
 abstract: |
@@
 We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
-
   <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "lexical acquisition, vocabulary, bilingualism"
-wordcount         : "X"
-
-bibliography      : ["references.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-keep_text         : false
-class             : "man"
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  word_document:
-    reference_docx: "manuscript_template.docx"
-
+keywords: lexical acquisition, vocabulary, bilingualism
+wordcount: X
+bibliography: references.bib
+floatsintext: yes
+figurelist: no
+tablelist: no
+footnotelist: no
+linenumbers: yes
+mask: no
+draft: no
+keep_text: no
+class: man
+documentclass: apa6
+classoption: man
+affiliation:
+- id: '1'
+  institution: Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona,
+    Spain
+- id: '2'
+  institution: Universidad Autónoma de Madrid
 ---
@@
 # load packages
-library(bayesplot)
 library(brms) 
@@
 library(purrr)
-library(quarto)
-library(rmarkdown)
 library(scales) 
-library(stringdist)
 library(stringr)
@@
 tar_load_everything()
+
 participants <- participants %>% 
@@
 
-Before the end of their first year of life, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental reports [e.g., @fenson2007macarthur].
+The foundations of word learning are in place early in age: before the end of their first year of life, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental reports [e.g., @fenson2007macarthur].
 During the last half of their second year, they acquire new words at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002children; @bergelson2020comprehension].
-These early stages of lexical acquisition are characterised by substantial variation between children, reflected, for instance, on the variation on the number of words they know [i.e.
+These early stages of lexical acquisition are characterised by substantial variation across children, reflected, for instance, on the variation on the number of words they know [i.e.
 vocabulary size; @fenson1994variability; @frank2021variability] or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, bates1994developmental].
@@
 The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*).
-These results were later extended by @frank2021variability to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
+These results were later extended by @frank2021variability to a wider diversity languages, also reporting that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
 Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
 
-There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
+There is evidence that bilinguals know, on average, less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
 For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English.
@@
 pointed to *parallel activation* as the main mechanism underpinning their results.
-The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension.
-This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@thierry2007brain].
+The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension, and that this phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain].
 One of the clearest examples of parallel activation was provided by @costa2000cognate.
@@
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
-Importantly, Spanish monolinguals did not show this effect.
-These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-Several recent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
+Importantly, Spanish monolinguals--unfamiliar with the Catalan translations of the Spanish words they uttered--did not show this effect.
+These results suggest that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
+Several subsequent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
 Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
@@
 Evidence supporting this claim is scarce.
-To the best of our knowledge, only two studies have provided direct data on this issue.
-First, @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
-Second, @schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
-The low sample size in these studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
+@bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
+@schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
+The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
+On the other hand, Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
 In their discussion the authors state the following (pp.
@@
 Second, all participants were aged ~24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
-In summary, the current evidence supporting an earlier age of acquisition of cognates vs.
-non-cognates presents some limitations that prevents drawing sound conclusions about this effect.
 
-Our second hypothesis detaches from the fact that, for parallel activation to take place, and for cognateness to facilitate the acquisition of a word form in a given language, the child must be familiar with the (phonologically similar) translation of the word, in the other language.
-If the child is not previously familiar with one of the word forms, the acquisition of its translation in  language cannot be facilitated by cognateness.
+More recently, @mitchell2022cognates addressed this issue overcoming some of these pitfalls.
+Using a larger, longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages.
+They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
+The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty).
+These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar: word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
+
+As @mitchell2022cognates note in their discussion, two competing hypotheses remain untested.
+The facilitative effect of cognateness can be explained by, at least, two alternative, arguably incompatible, mechanisms.
+On the one hand, cognateness might operate as a facilitator of the acquisition of both forms of a translation pair, with every exposure to either form activating the other, and therefore strengthening the association between both and their referent.
+If this is the case, then cognateness should increase the probability of acquisition of the forms in both languages equally.
+
+Another possibility is that cognateness plays a role in the acquisition of a translation pair only when at least one of the word-forms has been acquired.
+This account departs from the premise that, if the child is not previously familiar with one of the word forms, the acquisition of its translation in the other language cannot be facilitated by cognateness.
 This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair.
 Given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014much; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
-Therefore, we hypothesised that the effect of cognateness for words in the language of lower exposure would be larger than in the language of higher exposure.
 
-In line with this hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.
+In line with this second hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.
 reported was larger in the additional language vocabulary than in English vocabulary.
@@
 If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
-To test this hypothesis, we conditioned the effect of cognateness to the relative degree of exposure to the language.
 
-In summary, we investigated the role of cognateness in bilingual word acquisition (conditional to the degree of relative exposure to the language) as possible mechanism to account for the facilitation  of language similarity on bilingual children's vocabulary size.
-We hypothesised that cognate words would be acquired earlier than non-cognate words, and that this effect should be larger for words in the language with lower exposure.
-Using an online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.
+In this study, we investigated the role of cognateness on the acquisition of translation equivalent, and whether such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair.
+Using an online vocabulary checklist--designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+We then adopted an Item Response Theory (*IRT*) approach to model the probability of participants being reported by their parents to understand or understand *and* say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for other indices of word difficulty--lexical frequency and length (phonemes)-- and participant skill--age [see @kachergis2021towards for a similar approach].
 
@@
 All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-Families were contacted by e-mail or phone if their child were aged between 10 and 34 months, and had not previously reported to be exposed more than 10% of the time to a language other than Spanish or Catalan.
+Families from the participant database of the Laboratori de Recerca en Infància of the Universitat Pompeu Fabra were contacted by e-mail or phone if their child were aged between `r floor(min(participants$age))` and `r floor(max(participants$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan.
 Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate.
-Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant language (i.e., their degree of bilingualism).
+Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to Catalan.
 
-```{r participants-lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Participant sample size by age and degree of exposure to a second language.
-Information about DoE was provided by families when filling the questionnaire.
-For illustration purposes this table, DoEs were binned into 10\\% wide bins, and ages (in months) were round to the nearest integer.
-Hyphens indicate that no participants from that specific combination of age filled the questionnaire."}
+```{r tab-participants-lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Participant sample size by age and degree of exposure to Catalan.
+Information about DoE was provided by families before filling the questionnaire.
+A 100% indicated that the participant was exclusively exposed to Catalan, and never to Spanish.
+A 0% indicates that the participant was not exposed to Catalan ever, and rather most of the time to Spanish.
+A 50% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
+The 100%, 0%, and 50% would be traditionaly classified as Catlaan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
+For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins.
+Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."}
 participants %>% 
@@
     ) %>% 
-    mutate(age = cut_width(floor(age), 4)) %>% 
+    mutate(
+        age = cut(
+            floor(age),
+            c(10, 14, 18, 22, 26, 30, 34, 36),
+            include.lowest = TRUE
+        ),
+        doe_catalan = cut(
+            doe_catalan, 
+            c(0, 0.25, 0.5, 0.75, 1),
+            labels = c("0-25%", "25-50%", "50-75%", "75-100%"),
+            include.lowest = TRUE
+        )
+    ) %>% 
     count(age, doe_catalan) %>% 
-    arrange(-doe_catalan) %>% 
+    arrange(desc(doe_catalan)) %>% 
     pivot_wider(names_from = age, values_from = n) %>% 
     gt() %>% 
-    fmt_missing(everything(), everything(), missing_text = "--") %>% 
-    cols_label(
-        doe_catalan = "DoE Catalan"
+    summary_rows(
+        columns = 2:8, 
+        fns = list("Total" = ~sum(., na.rm = TRUE)),
+        formatter = fmt_integer
     ) %>% 
-    fmt_percent("doe_catalan", decimals = 0) %>% 
-    tab_spanner("Age (months)", 2:8)
+    sub_missing(everything(), everything(), missing_text = "--") %>% 
+    cols_label(doe_catalan = "DoE Catalan") %>%
+    tab_spanner("Age (months)", 2:8) %>% 
+    tab_footnote(
+        "This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10% of the time", 
+        cells_column_labels(columns = doe_catalan)
+    ) %>% 
+    tab_style(
+        cell_text(align = "left"),
+        list(
+            cells_body(doe_catalan),
+            cells_column_labels(doe_catalan)
+        )
+    )
 ```
@@
 Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-In case two translation equivalents were possible for a given word, both were inluded as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
+In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
 We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
@@
 
-```{r item-summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
-Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
+```{r tab-item-summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical properties of the word.forms and trasnlation pairs included in the vocabulary checklist and our analyses.
+We report the number of items, mean (*M*), standard deviation (*SD*), and range of the lexical frequency (epxressed in Zipf scores), length (number of phonemes) of the word forms, and the phonological similarity (normalised Levenshtein similarity) with their translation equivalent.
+Statistics are reported separately for Catalan and Spanish, and for the different vocabulary checklists that participants were allocated to (A, B, C, D)."}
+items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))
+
 items %>%
@@
     tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
-    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
+    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
     cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
@@
 
-```{r coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-coefs <- as.data.frame(fixef(model_fit_7)) %>% 
-    clean_names()
+# Discussion
 
-draws_fixed <- gather_draws(model_fit_7, `b_.*`, regex = TRUE)
-```
+## Summary of the study
 
-Table 2 shows the summary of the posterior distribution of the fixed coefficients of the model.
+This study explored whether word acquisition was affected by cognateness in bilingual toddlers, and how (if at all) this effect interacted with language exposure.
+We collected vocabulary data from children learning Catalan and/or Spanish in differing degrees of exposure to each, and then measured the probability of each word being acquired at different ages as a function of (1) participants' exposure to the language it belongs to, and (2) its phonological overlap with its translation equivalent in the other language.
+We build a multilevel Bayesian model that estimated the effect of both predictors and their interaction, adjusting for other variables known to be relevant during the acquisition of words: participants' age, lexical frequency, and length in phonemes.
+We tested the hypotheses that since, words in the language of higher exposure are more likely to be acquired earlier than words in the language of lower exposure, phonological similarity between translation equivalents (i.e., cognateness) should play a more important role during the acquisition of words in the later.
+We therefore predicted a larger cognateness effect as the exposure to language decreased.
 
-```{r results-fixed, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-str_repl <- c(
-    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein",
-    "b_age_std:doe_std" = "Age \u00d7 DoE",
-    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein",
-    "b_lv_std" = "Levenshtein (+1 SD)",
-    "b_doe_std" = "DoE (+10%)",
-    "b_n_phon_std" = "# Phonemes (+1 SD)",
-    "b_freq_std" = "Frequency (+1 SD)",
-    "b_age_std" = "Age (+1 month)",
-    "b_Intercept[1]" = "Intercept (Production)",
-    "b_Intercept[2]" = "Intercept (Comprehension)"
-)
+## Summary of the results
 
-draws_fixed %>%
-    mutate(
-        .value = ifelse(
-            grepl("Intercept", .variable),
-            inv_logit_scaled(.value), 
-            .value/4
-        ),
-        .variable = factor(.variable, levels = names(str_repl), labels = str_repl)
-    ) %>% 
-    mean_qi(.width = c(0.95, 0.89)) %>% 
-    pivot_wider(names_from = .width, values_from = c(.value, .lower, .upper)) %>% 
-    select(-c(.interval, .point, .value_0.89)) %>% 
-    arrange(desc(.variable)) %>% 
-    gt() %>% 
-    fmt_percent(2:6) %>% 
-    cols_merge(vars(".lower_0.95", ".upper_0.95"), pattern = "[{1}, {2}]") %>% 
-    cols_merge(vars(".lower_0.89", ".upper_0.89"), pattern = "[{1}, {2}]") %>% 
-    cols_label(
-        .variable = md("Predictor"),
-        .value_0.95 = md("Mean"),
-        .lower_0.95 = md("95\\% *CrI*"),
-        .lower_0.89 = md("89\\% *CrI*")
-    ) %>% 
-    tab_footnote(
-        footnote = "Transformed using the inverse logit to get the average probability of correct response",
-        locations = cells_body(
-            columns = ".variable", 
-            rows = grepl("Intercept", .variable)
-        ) 
-    ) %>% 
-    tab_footnote(
-        footnote = "Transformed using the divide-by-four rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-        locations = cells_body(
-            columns = ".variable",
-            rows = !grepl("Intercept", .variable)
-        )
-    )  %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = everything())
-    ) %>% 
-    tab_header(title = "Model fixed coefficients", subtitle = "Posterior distribution") %>% 
-    tab_style(
-        cell_text(align = "left"),
-        list(cells_body(), cells_column_labels(columns = everything()))
-    )
-```
+Adopting a Bayesian IRT modelling, we found conclusive evidence in favour of a facilitatory effect of cognateness, conditional to the degree of exposure to the language: phonological similarity across translation equivalents increased the probability of comprehension and production only when exposure to the language was low.
+We did not find evidence in favour of this interaction behaving differently across ages, although the uncertainty associated with this estimate (the regression coefficient of the $Age \times DoE \times Levenshtein$) was large, indicating that we cannot confidently state whether younger or elder children were any more sentitive to the effect of language exposure and congnateness.
 
-The extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density.
-This estimate is several times larger than its standard error indicating high predictive accuracy.
-Table 3 summarises the LOO scores of the fitted models.
-In the next sections we report the fixed effects of the comprehension and production models separately.
-We report the mean of each coefficient's posterior distribution along with standard errors and 95% credible intervals (CrI) in Tables 4 and 5.
-For interpretability, we report the derivative of the logistic function for each coefficient^[The derivative is calculated as $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.
-This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.] for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable [@gelman2020regression].
+## Results in contrast with the literature
 
-```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Leave-one-out cross-validation.
-The expected log predictive density (*ELPD*) indicates the sum of the likelihood of each observation, given the model when fitted without such obervation.
-Values close to 0 indicate that the model fits data better.
-Standard errors (*SE*) indicate the uncertainty around the computed ELPD in each model.
-A small *SE* (in realtion to the magnitde of the *ELPD* indicates less uncertainty around the ELPD estimate.", eval=FALSE}
-l_comp <- loo_comp %>% 
-    as_tibble() %>% 
-    select(-matches("p_loo|looic")) %>% 
-    mutate_all(as.numeric) %>% 
-    mutate(
-        type = "Comprehension",
-        model = paste0("Model ", 2:0)
-    ) %>% 
-    relocate(model, type, matches("elpd_loo"), matches("diff"))
+## Mechanistic explanations
 
-l_prod <- loo_prod %>% 
-    as_tibble() %>% 
-    select(-matches("p_loo|looic")) %>% 
-    mutate_all(as.numeric) %>% 
-    mutate(
-        type = "Production",
-        model = paste0("Model ", 2:0)
-    ) %>% 
-    relocate(model, type, matches("elpd_loo"), matches("diff"))
+## Limitations
 
-l <- bind_rows(l_comp, l_prod) %>% 
-    mutate_all(function(x) ifelse(x==0, NA, x))
-
-t <- gt(l, groupname_col = "type") %>% 
-    tab_spanner(md("**ELPD**"), matches("elpd_loo")) %>% 
-    tab_spanner(md("**Diff.**"), matches("diff")) %>%
-    fmt_number(2:7) %>% 
-    fmt_missing(columns = everything(), missing_text = "-") %>% 
-    cols_label(
-        model = "Model",
-        # elpd
-        elpd_loo = md("*ELPD*"),
-        se_elpd_loo = md("*SE*"),
-        subsampling_se_elpd_loo = md("*SE* (sub)"),
-        # diff
-        elpd_diff = md("*diff*"),
-        se_diff = md("*SE*"),
-        subsampling_se_diff = md("*SE* (sub)"),
-    )
-```
-
-## Comprehension
-
-The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe)*10)`%), and regardless of its cognate status (average across the three levels).
-An increase in one month of age was associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability.
-A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_comp["frequency", "Estimate"]*100/4)`% increment in comprehension probability.
-A 10% increase in relative language exposure made TEs `r printnum(abs(coef_comp["doe_center", "Estimate"]*100/4))`% less likely to be understood.
-Cognates (identical and non-identical) were `r printnum(coef_comp["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates.
-The 95% credible interval of the `doe_center:cognate1` interaction excluded 0 and covered negative values, indicating that for every 10% increment in DoE, the difference in comprehension probability between cognates and non-cognates decreased in `r printnum(abs(coef_comp["doe_center:cognate1", "Estimate"]*100/4))`%.
-Identical cognates were `r printnum(abs(coef_comp["cognate2", "Estimate"]*100/4))`% less like to be understood than non-identical cognates.
-The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero), and covered only negative values, indicating that for every 10% increase in DoE, the difference in comprehension probability between identical cognates and non-identical cognates decreased in `r printnum(abs(coef_comp["doe_center:cognate2", "Estimate"]*100/4))`%.
-Figure 1 summarises the posterior distribution of the fixed effects of the model.
-
-```{r results-fixed, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension and production.
-Points and error bars in indicate the posterior means, and 50% and 95% CrI.
-The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-c_comp <- fixef(fits_comp$fit_2) %>% 
-    as.data.frame() %>%
-    rownames_to_column("term") %>% 
-    clean_names() %>% 
-    mutate(
-        estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
-        q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
-        q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
-        est_error = ifelse(term=="Intercept", inv_logit_scaled(est_error), est_error/4)
-    ) %>% 
-    rename_at(vars(-matches("term")), ~paste0("comp_", .)) %>% 
-    select(-any_of(c("rhat", "bulk_ess", "tail_ess")))
-
-c_prod <- fixef(fits_prod$fit_2) %>% 
-    as.data.frame() %>%
-    rownames_to_column("term") %>% 
-    clean_names() %>% 
-    mutate(
-        estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
-        q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
-        q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
-        est_error = ifelse(term=="Intercept", inv_logit_scaled(est_error), est_error/4)
-    ) %>% 
-    rename_at(vars(-matches("term")), ~paste0("prod_", .)) %>% 
-    select(-term, -one_of(c("term", "rhat", "bulk_ess", "tail_ess")))
-
-c <- bind_cols(c_comp, c_prod)
-
-t <- gt(c) %>% 
-    fmt_percent(matches("estimate|error")) %>% 
-    fmt_percent(matches("_q")) %>% 
-    tab_spanner("Comprehension", matches("comp_")) %>% 
-    tab_spanner("Production", matches("prod_")) %>% 
-    cols_merge(vars("comp_q2_5", "comp_q97_5"), pattern = "[{1}, {2}]") %>% 
-    cols_merge(vars("prod_q2_5", "prod_q97_5"), pattern = "[{1}, {2}]") %>%
-    cols_label(
-        term = md("**Predictor**"),
-        comp_estimate = md("**Mean**"),
-        comp_est_error = md("**SE**"),
-        comp_q2_5 = md("**95\\% CrI**"),
-        comp_q2_5 = md("**95\\% CrI**"),
-        prod_q2_5 = md("**95\\% CrI**"),
-        prod_estimate = md("**Mean**"),
-        prod_est_error = md("**SE**"),
-        prod_q2_5 = md("**95\\% CrI**")
-    ) %>% 
-    tab_footnote(
-        footnote = "Transformed using the inverse logit to get the average probability of correct response",
-        locations = cells_body(columns = "term", rows = term=="Intercept")
-    ) %>% 
-    tab_footnote(
-        footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-        locations = cells_body(columns = "term", rows = term %in% c("age_center", "frequency_center", "doe_center", "cognate1", "cognate2", "doe_center:cognate1"))
-    ) %>% 
-    cols_align(
-        align = c("left"),
-        columns = 2:4
-    )
-gtsave(t, here("Figures/fix_effects.png"))
-```
-
-## Production
-
-The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_prod["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0.
-An increase in one month of age was associated with a `r printnum(coef_prod["age_center", "Estimate"]*100/4)`% increment in production probability.
-A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_prod["frequency", "Estimate"]*100/4)`% increment in production probability.
-A 10% increase in relative language degree of exposure  made TEs `r printnum(coef_prod["doe_center", "Estimate"]*100/4)`% less likely to be produced.
-Cognates (identical and non-identical) were `r printnum(coef_prod["cognate1", "Estimate"]*100/4)`% more like to be produced than non-cognates.
-The `doe_center:cognate1` interaction (whose posterior 95% CrI did not exclude 0) indicated that the size of this effect decreased in `r printnum(abs(coef_comp["cognate1", "Estimate"]*100/4))`% for every 10% increment in language degree of exposure.
-Identical cognates were `r printnum(abs(coef_prod["cognate2", "Estimate"]*100/4))`% less likely to be produced than non-identical cognates.
-The `doe_center:cognate1` interaction (whose posterior 95% CrI did not excdude 0 either) indicated that the size of this effect decreased in `r printnum(abs(coef_comp["doe_center:cognate2", "Estimate"]*100/4))`% for every 10% increment in language degree of exposure.
-
-```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=6, fig.width=8, fig.cap="Expected marginal mean and 95% credible interval of the posterior probability of comprehension and production for identical cognates, non-identical cognates, and non cognates, plotted separately for three levels of degree of exposue to the language: 0% (no exposure, monolingual), 50% (exposed half of the time, balanced bilingual), and 100% (exposed all of the time, monolingual).
-Dots represent the mean, and error bars represent the range of values within the true values is contained with 50% (thick line) and 95% (thin line) cprobability, given the data."}
-nd <- expand.grid(
-    age_center = floor(seq(min(responses$age_center), max(responses$age_center), 0.5)),
-    doe_center = c(min(responses$doe_center), mean(responses$doe_center), max(responses$doe_center)),
-    cognate = unique(items$cognate),
-    frequency_center = 0
-)
-
-# comprehension
-m_comp <- epred_draws(fits_comp$fit_2, newdata = nd, ndraws = 50, re_formula = NA) %>% 
-    mutate(
-        frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-        doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%")),
-        cognate = factor(cognate, levels = c("Identical cognate", "Non-identical cognate", "Non-cognate"))
-    )
-m_comp_ci <- mean_qi(m_comp)
-
-# production
-m_prod <- epred_draws(fits_prod$fit_2, newdata = nd, ndraws = 50, re_formula = NA) %>% 
-    mutate(
-        frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
-        doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "Doe: 50%", "DoE: 100%")),
-        cognate = factor(cognate, levels = c("Identical cognate", "Non-identical cognate", "Non-cognate"))
-    )
-m_prod_ci <- mean_qi(m_prod)
-
-# join datasets
-m <- bind_rows(m_comp, m_prod, .id = "type") %>% 
-    mutate(
-        type = ifelse(type==1, "Comprehension", "Production"),
-        age_center = age_center+mean(responses$age)
-    )
-
-m_ci <- bind_rows(m_comp_ci, m_prod_ci, .id = "type") %>% 
-    mutate(
-        type = ifelse(type==1, "Comprehension", "Production"),
-        age_center = age_center+mean(responses$age)
-    )
-
-ggplot(m, aes(age_center, .epred, colour = cognate, fill = cognate, shape = cognate)) +
-    facet_grid(type~doe_center) +
-    # geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.25, size = 0.2,
-    #           show.legend = FALSE) +
-    # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
-    geom_hline(yintercept = 0.5, colour = "grey") +
-    geom_vline(xintercept = mean(responses$age), colour = "grey") +
-
-    geom_errorbar(data = m_ci, aes(ymin = .lower, ymax = .upper), width = 0.5, size = 0.25) +
-    stat_summary(fun = "mean", geom = "line", size = 0.75) +
-    # stat_pointinterval(.width = 0.95, position = position_dodge(width = 0.5), size = 0.75, point_size = 2) +
-    scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.25)) +
-    scale_x_continuous(labels = floor, breaks = floor(seq(min(responses$age), max(responses$age), by = 2))) +
-    labs(
-        x = "Age (months)", y = "Posterior probability",
-        colour = "Cognateness", fill = "Cognateness", shape = "DoE"
-    ) +
-    theme(
-        legend.position = "top",
-        axis.text.x = element_text(size = 8),
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-        legend.title = element_blank(),
-    )
-
-```
-
-# Discussion
+## Further steps
 

### 2022-11-17 docs/childes-freqs.qmd
Update lab notes

---
title: "Cross/language CHILDES frequencies"
format: html
code-fold: true
---

```{r}
#| label: setup
#| echo: true
library(childesr)
library(dplyr)
library(ggplot2)
library(purrr)
library(patchwork)
library(tidyr)
library(janitor)
library(ggdist)
library(gt)
library(arrow)
library(readr)
```

```{r}
#| label: wordbank
#| echo: true
# function to retrieve frequencies
wordbank <- map_df(
    c(spa = "data/wordbank_spa.csv", eng = "data/wordbank_eng.csv"),
    read_csv,
    .id = "language"
    ) %>% 
    clean_names() %>% 
    select(language, age, definition, uni_lemma, ipa, length, freq)

wordbank %>%
    drop_na() %>% 
    distinct(uni_lemma, definition, language, freq) %>% 
    arrange(uni_lemma) %>% 
    pivot_wider(
        id_cols = uni_lemma,
        names_from = language,
        values_from = c(freq, definition),
        values_fn = first
        ) %>% 
    drop_na() %>% 
    ggplot(aes(freq_eng, freq_spa)) +
    geom_point()
```

```{r}
#| label: retrieve-freqs
#| echo: true
# function to retrieve frequencies

get_childes_data <- function(word, collection, ...){

    roles <- c("Mother", "Father", "Investigator", "Sibling", "Sister", "Grandmother", "Adult", "Friend", "Brother", "Visitor", "Relative", "Grandfather", "Teacher", "Student")

    counts <- get_types(type = word, collection = collection, role = roles, ...)

    speaker_ids <- distinct(counts, collection_id, corpus_id, transcript_id, speaker_id)

    speakers <- speaker_ids %>%
        left_join(get_speaker_statistics(collection = collection)) %>%
        select(collection_id, corpus_id, transcript_id, speaker_id, num_tokens)

    count_estimates <- counts %>%
        left_join(speakers) %>%
        mutate(
            id = as.character(id),
            age_months = target_child_age,
            age_bin = as.integer(floor(age_months / 6) * 6),
            token = tolower(gloss)
        ) %>%
        group_by(age_bin, token, target_child_id, transcript_id) %>%
        summarise(
            transcript_count = sum(count),
            transcript_num_tokens = sum(num_tokens),
            .groups = "drop"
        ) %>%
        group_by(age_bin, token) %>%
        summarise(
            freq_count = mean(transcript_count), 
            total_count = mean(transcript_num_tokens),
            n_children = length(unique(target_child_id)),
            .groups = "drop"
        ) %>%
        mutate(
            freq_million = freq_count / total_count * 1e6,
            freq_zipf = log10(freq_million)+3
        )

    return(count_estimates)

}

# params
tokens <- list(
    spa = c("casa", "puerta", "mesa", "gato", "caja", "lampara"),
    cat = c("casa", "porta", "taula", "gat", "caixa", "llum"),
    eng = c("house", "door", "table", "cat", "box", "lamp")
)

freqs <- map2_df(
    collections, 
    corpora, 
    ~get_frequencies(collection = .x, corpus = .y, age = c(10, 36)), 
    .id = "language"
)
```

# How many children, transcriptions, and tokens are available in each language?

```{r}
#| label: num-children
#| echo: true
collections <- list(
    spa = "Spanish",
    cat = "Romance",
    eng = c("Eng-NA", "Eng-UK")
)

corpora <- list(
    spa = NULL,
    cat = c("EsteveGibert/Prieto", "GRERLI", "Jordina", "Julia", "Mireia/Eva/Pascual", "Serra/Solé"),
    eng = NULL
)

counts <- map2_df(
    collections, 
    corpora, 
    ~get_types(collection = .x, role = roles, corpus = .y),
    .id = "language"
)

counts_to_transpose <- counts %>% 
    drop_na(target_child_id) %>% 
    group_by(language) %>% 
    summarise(
        n_children = length(unique(target_child_id)),
        n_transcripts = length(unique(transcript_id)),
        n_distinct_words = length(unique(gloss)),
        n_word_instances = sum(count),
        median_word_instances = median(count),
        max_word_instances = max(count)
    )

as_tibble(cbind(nms = names(counts_to_transpose), t(counts_to_transpose))) %>% 
    set_names(.[1,]) %>% 
    rename(variable = language) %>% 
    slice(-1) %>% 
    mutate(across(cat:spa, ~as.numeric(str_trim(.)))) %>% 
    gt() %>% 
    fmt_integer(2:4)

```

ggplot(freqs, aes(age_bin, freq_zipf)) +
    geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) +
    scale_y_continuous(limits  = c(1, 7))

tes <- tibble(
    language = c("cat", "spa", "eng"),
    house = c("casa", "casa", "house")
)

freqs %>% 
    filter(
        (language=="cat" & token=="casa") |
            (language=="spa" & token=="casa") |
            (language=="eng" & token=="house")
    ) %>% 
    ggplot(aes(age_bin, freq_zipf, colour = language)) +
    geom_point(size = 4, shape = 1, stroke = 2) +
    scale_y_continuous(limits = c(1, 7))

group_by(age_bin, language) %>% 
    summarise(total_count = sum(total_count)) %>% 
    ggplot(aes(age_bin, total_count, fill = language)) +
    geom_col(position = position_dodge(), width = 3) +
    geom_text(
        aes(label = round(total_count), y = total_count + 100),
        position = position_dodge(width = 1),
        angle = 90,
        nudge_y = 100
    )

freqs %>% 
    ggplot() +
    aes(
        x = age_bin,
        y = freq_zipf,
        colour = language
    ) +
    facet_wrap(~token) +
    stat_summary(fun.data = "mean_se", geom = "pointrange") +
    facet_wrap(~token, scales = "free_y") +
    scale_x_continuous(
        breaks = seq(10, 36, 4),
        name = "Age (months)"
    ) +
    scale_y_continuous(name = "Frequency (per million words)")

```

### 2022-11-17 manuscript.qmd
Update lab notes

@@
 pagetitle: Cross-linguistic similarity and  word acquisition
-abstract: |
-    Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 10 to 36 months.
 thanks: |
@@
     linestretch: 1.25
+  docx:
+    reference-doc: manuscript/manuscript_template.docx
+    linestretch: 1.25
 ---
@@
 tar_load(model_loos)
+tar_load(model_fit_0)
 tar_load(model_fit_1)
+tar_load(model_fit_2)
+tar_load(model_fit_4)
 tar_load(rope_coefs)
@@
 
-# Introduction {#introduction}
-
-The foundations of word learning are in place early in age: before the end of their first year of life infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental reports [e.g., @fenson2007macarthur].
-During the last half of their second year, they acquire new words at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002children; @bergelson2020comprehension].
-These early stages of lexical acquisition are characterised by substantial variation across children, reflected, for instance, on the variation on the number of words they know [i.e.
-vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, bates1994developmental].
+# Introduction {#sec-introduction}
+
+## Word acquisition during toddlerhood: the effect of language experience
+
+The foundations of word learning are in place early in age.
+Before the end of their first year of life infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthur].
+During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002children; @bergelson2020comprehension].
+These early stages of lexical acquisition are characterised by substantial variation across children.
+This is reflected, for instance, on the variability of the number of words they know [i.e.
+vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of acquired words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, @bates1994developmental].
 
@@
 
-There is evidence that bilinguals know, on average, less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
+Previous findings point to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the two languages are aggregated.
 For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English.
-Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development.
+Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, highlighting the importance of collecting data on both languages when assessing bilinguals' communicative development.
 Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual].
@@
 
+## Language distance as a mediator of bilingual vocabulary growth
+
 For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language.
 Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
-Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English (one Roman, the other Germanic).
+Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (one Romance, the other Germanic).
 For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
@@
 
+## Lexical non-selectiviy as a candidate acount: the role of parallel activation
+
 Floccia et al.
-pointed to *parallel activation* as the main mechanism underpinning their results.
+pointed to *parallel activation* as the main phenomenon underpinning their results.
 The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension, and that this phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain].
@@
 Several subsequent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
-Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
-A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.
-
-In this study, we hypothesise that cognates are acquired earlier in age than non-cognates, which might explain why children learning languages that share many cognates show larger vocabulary sizes.
-Evidence supporting this claim is scarce.
+Parallel activation is therefore a plausible explanation for Floccia et al.'s results: cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
+This account is is line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words, already acquired or not [e.g., @hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
+Importantly, bilinguals seem to be more likely to acquire words for which their translation equivalent has been already learnt in the other language, suggesting that semantic similarity also facilitates word acquisition across languages [@bilson2015semantic].
+
+If phonological similarity also plays a cross-language facilitation role during word acquisition, cognate translation equivalents--which share both semantic and phonological similarity--should be acquired earlier that non-cognate translation equivalents--which share semantic similarity, but not phonological similarity.
+Evidence supporting and earlier age of acquisition for cognates is scarce.
 @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
@@
 
-As @mitchell2022cognates note in their discussion, two competing hypotheses remain untested.
-The facilitative effect of cognateness can be explained by, at least, two alternative, arguably incompatible, mechanisms.
-On the one hand, cognateness might operate as a facilitator of the acquisition of both forms of a translation pair, with every exposure to either form activating the other, and therefore strengthening the association between both and their referent.
-If this is the case, then cognateness should increase the probability of acquisition of the forms in both languages equally.
+## Two competing mechanisms for a cognate facilitation during word acquisition
+
+The facilitative effect of cognateness on word acquisition can be explained by at least two arguably incompatible mechanisms.
+On the one hand, and in line with @mitchell2022cognates, cognateness might operate at early stages of word acquisition, before stable representations of any of the word-forms of the translation equivalent have been formed.
+In this scenario, recognising any of the two word-forms as familiar might lead to the activation of its (cognate) translation in the other language via their phonological similarity.
+Given that both word-forms are likely to be heard in the presence of the same referent, the association between each form and such reference could be facilitated by the exposure to either of them.
+
+Other words might be activated as well for the same reasons (similar sounding words in either language but with different meaning), but only the acquisition of translation equivalents might be f
+
+very exposure to a word-form activates its (phonologically similar) translation, strengthening the association between both word-forms and their referent, even when only the form in one of the languages is heard most of the times.
+
+For instance, hearing *flower* in English might activate *floeur* in French, and *vice versa*.
+Repeated exposure to the English or the French words, and therefore to similar -sounding words in the presence of  eventually leads to an earlier acquisition for both of them, as compared to non-cognate translation equivalents like *cow* and *vache*, which do not activate each other via phonological similarity.
+If this is the case, then cognateness should increase the probability of acquisition of the word-forms in both languages equally.
 
 Another possibility is that cognateness plays a role in the acquisition of a translation pair only when at least one of the word-forms has been acquired.
-This account departs from the premise that, if the child is not previously familiar with one of the word forms, the acquisition of its translation in the other language cannot be facilitated by cognateness.
+This account departs from the premise that phonological similarity operates at the lexical level to facilitate word learning across languages: if the child has still not formed a lexical representation with sufficient phonological detail for one of the word forms, the acquisition of its translation in the other language cannot be facilitated by cognateness.
 This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair.
@@
 
-In line with this second hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.
-reported was larger in the additional language vocabulary than in English vocabulary.
+In line with this second hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al.
+reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure.
 Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
@@
 
-In this study, we investigated the role of cognateness on the acquisition of translation equivalent, and whether such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair.
+## The present study
+
+In this study, we investigated the role of cognateness on the acquisition of translation equivalents, and if such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair.
 Using an online vocabulary checklist--designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We then adopted an Item Response Theory (*IRT*) approach to model the probability of participants being reported by their parents to understand or understand *and* say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for other indices of word difficulty--lexical frequency and length (phonemes)-- and participant skill--age [see @kachergis2022toward for a similar approach].
-
-# Methods {#methods}
-
-Data and materials are available at OSF [https://osf.io/hy984/](https://osf.io/hy984/), and code is available at GitHub [https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories).
-
-## Participants {#participants}
+We then adopted an Bayesian item response theory (*IRT*) approach to model the probability of participants being reported by their parents to understand or understand *and* say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for other indices of word difficulty--lexical frequency [@goodman2008does] and length (phonemes)-- and participant skill--age [see @kachergis2022toward for a similar approach].
+
+# Methods {#sec-methods}
+
+## Participants {#sec-participants}
 
@@
 
-## Questionnaire {#questionnaire}
+## Questionnaire {#sec-questionnaire}
 
@@
 
-The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+To collect vocabulary data from participants, we created an *ad hoc* questionnaire, MultiLex [@multilex].
+This questionnaire was inspired in the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthur] and its adaptations to other languages, and was implemented on-line using the formR platform [@arslan2020formr].
+This questionnaire is structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
 Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
@@
 For the analyses in the present study, we considered responses to words corresponding to nouns, verbs, and adjectives [@fourtassi2020growth].
-We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-
-We extracted four variables of interest for each word-form: (1) Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
+We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono/mico*).
+
+## Variables of interest
+
+We extracted four variables of interest for each word-form.
+(1) Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
 This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points.
@@
 #| warning: false
-#| tbl-cap: "Lexical properties of the word-forms and translationpairs included in the vocabulary checklist and our analyses.
-We report the number of items, mean (*M*), standard deviation (*SD*), and range of the lexical frequency (epxressed in Zipf scores), length (number of phonemes) of the word forms, and the phonological similarity (normalised Levenshtein similarity) with their translation equivalent.
+#| tbl-cap: "Lexical properties of the word-forms and translation pairs included in the vocabulary checklist and our analyses.
+We report the number of items, mean, standard deviation (*SD*), and range of the lexical frequency (epxressed in Zipf scores), length (number of phonemes) of the word forms, and the phonological similarity (normalised Levenshtein similarity) with their translation equivalent.
 Statistics are reported separately for Catalan and Spanish, and for the different vocabulary checklists that participants were allocated to (A, B, C, D)."
@@
     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
-    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
+    tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>% 
     tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
@@
 
-## Data analysis {#analysis}
+## Data analysis {#sec-analysis}
 
@@
 n_te <- length(unique(df$te))
+
+rope_interval <- c(lower = -0.1, upper = 0.1)
 ```
@@
 
-We modelled the probability of answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model.
-This allowed us to simultaneously estimate the impact of item- and participant-level predictors on the probability of word comprehension and production.
-We included $Age$, $Frequency$, $Phonemes$, $DoE$, and $Levenshtein$ as fixed effects.
-We also included the $Age \times DoE \times Levenshtein$ three-way interaction, and its corresponding two-way interactions.
-We added participants and items as crossed random effects including all intercepts and slopes when appropriate [@barr2013random].
-
-This model structure, including word- and participant-level properties to estimate vocabulary acquisition trajectories was discussed by @kachergis2022toward in the context of explanatory item response models.
-Our model differs from their proposal in that  1) while they modelled global trajectories exclusively, we modelled individual trajectories as well, 2) our outcome variable is ordinal in contrast to the binary outcome used in their study, 3) we propose a Bayesian approach.
-Our motivation for a Bayesian approach towards statistical inference is that (1) it allows us to incorporate previous domain knowledge to improve estimation accuracy, and (2) to quantify the uncertainty associated with the estimated parameters in our model [@mcelreath2018statistical].
-
-The Bayesian approach implies using Bayes’ rule to assign a probability to possible parameter values (i.e., posterior distribution), based on previous knowledge about such probability distribution (prior distribution) and our data (likelihood).
-We used Markov chain Monte Carlo (MCMC) to approximate the posterior distribution via sampling.
-We run 2 sampling chains with 4,000 iterations each, including 2,000 warm-up iterations per chain.
-All $\hat{R}$ values were smaller than 1.1, indicating that the chains converged successfully in all cases [@gelman1992inference].
-
-We compared a base model–-only including the intercept, $Age$, $Frequency$, $Phonemes$, and $DoE$ as fixed effects, and their corresponding random effects-–to increasingly more complex models, introducing predictors in the following order: $Age \times DoE$, $Levenshtein$, $DoE \times Levenshtein$, and finally $Age \times Levenshtein$ and $Age \times DoE \times Levenshtein$.
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO) [@vehtari2017practical]: we computed the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random sub-sample of `r dim(model_log_liks$model_fit_1)[1]` samples (out of `r format(dim(model_fit_1$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
-and selected the model with the smallest expected log posterior density (ELPD)^[Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with `r format(dim(model_log_liks$model_fit_1)[1], big.mark = ",")` posterior samples [@magnusson2019bayesian]].
-
-Finally, we explored the posterior distribution of each parameter in the selected model.
-Following @kruschke2018bayesian, We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
-We defined our *ROPE* from -`r round(median(abs(rope_coefs$estimate)), 3)` to +`r round(median(abs(rope_coefs$estimate)), 3)` [see @#appendix-rope for justification of the *ROPE*].
-In the results section, we report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
-This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
-We also report the proportion of the posterior samples of that coefficient that fell within the *ROPE*.
-A 90% overlap between the 95% *HDI* of the coefficients' posterior distribution and the *ROPE* indicates that, given the data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero, that is, negligible.
+We modelled the probability of answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_0$fit)[2]` sampling chains with `r format(dim(model_fit_0$fit)[1], big.mark = ",")` iterations each).
+This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$, $Frequency$, $Phonemes$, $DoE$, and $Levenshtein$.
+We also included participant and item random intercepts and slopes when appropriate [@barr2013random] [see @sec-appendix-words for details on the model structure, implementation, and diagnostics].
+We compared a baseline model $\mathcal{M}_{0}$, which only included the intercept, $Age$, $Frequency$, $Phonemes$, and $DoE$ as predictors, with increasingly more complex models, introducing predictors in the following order:
+$Age \times DoE$ ($\mathcal{M}_{1}$), $Levenshtein$ ($\mathcal{M}_{2}$), $DoE \times Levenshtein$ ($\mathcal{M}_{3}$), and finally $Age \times Levenshtein$ and $Age \times DoE \times Levenshtein$ ($\mathcal{M}_{4}$).
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO, @vehtari2017practical]^[Due to computational constraints, we performed LOO based on a random sub-sample of `r format(dim(model_log_liks$model_fit_0)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2019bayesian]], and selected for interpretation the one showing the best performance.
+
+We assessed the practical relevance of the estimated regression coefficients of the selected model by specifying a region of practical equivalence [*ROPE*, @kruschke2018bayesian] from `r rope_interval["lower"]` to +`r rope_interval["upper"]` [see @sec-appendix-rope].
+This region indicates the range of values that we considered as equivalent to zero.
+The degree of overlap between the posterior distribution of a regression coefficient with the *ROPE* can be interpreted as support for the true parameter of the coefficient being zero.
+In the results section, we report the 95% highest density interval (*HDI*) of the posterior distribution of each regression coefficient.
+This is interval contains the true value of this coefficient with 95% probability, given our data.
+We also report the proportion of posterior samples in the 95% *HDI* that overlap with the *ROPE*, which indicates the probability that the true value of the regression coefficient falls into the *ROPE*, and should therefore be considered equivalent to zero^[For example, a 90% overlap between the *HDI* of the coefficients’ posterior distribution and the *ROPE* indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero].
 
 Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@loo], and `tidybayes` [@tidybayes] R packages.
-
-# Results {#results}
-
-## Model comparison and selection {#comparison}
-
-@tbl-results-loos shows the outputs of the *LOO-CV*.
-The extended model (`model_fit_1`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference.
-This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.
+Data and materials are available at OSF [https://osf.io/hy984/](https://osf.io/hy984/), and code is available at GitHub [https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories).
+
+# Results {#sec-results}
+
+## Model comparison and selection {#sec-comparison}
+
+The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the other models several times larger than the standard error of such difference [see @tbl-results-loos].
+This indicates that, given the data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models.
 
@@
 #| warning: false
-#| fig-width: 7
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
+#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
+The predictive accuracy of the models was compared using leave-one-out cross-validation.
+The likelihood of each data point given a model fitted without including that same observation was estimated.
+The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
+*ELPD* values closer to zero indicate better predictive acuracy.
+We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
+The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
 model_loos %>% 
@@
     rownames_to_column("model") %>% 
+    mutate(model = str_remove(model, "model_fit_")) %>% 
     relocate(
@@
     gt() %>% 
-    tab_header(
-        title = "",
-        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
-The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_1`)")
-    ) %>% 
     tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
@@
 
-## Regression coefficients {#coefficients}
-
-```{r}
-#| label: tbl-results-fixed
-#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
-#| echo: false
-#| message: false
-#| warning: false
-post_draws <- gather_draws(model_fit_1, `b_.*`, regex = TRUE)
+## Regression coefficients {#sec-coefficients}
+
+```{r}
+#| label: results-coefs
+#| echo: false
+#| message: false
+#| warning: false
+post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)
 
@@
 
+posterior_description <- bayestestR::describe_posterior(
+    model_fit_4,
+    test = "rope",
+    rope_range = rope_interval, 
+    ci_method = "HDI"
+) %>% 
+    as_tibble() %>% 
+    clean_names()
+
 post_draws_list <- split(posterior_description, posterior_description$parameter)
-
+```
+
+@tbl-results-fixed shows the summary of the posterior distribution of the fixed regression coefficients of the selected model.
+For interpretability, we supplement the summaries of each posterior regression coefficient with its transformation to the probability scale.
+The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
+This means that one logit difference is not necessarily translated to a unique value in the probability scale.
+For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
+The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
+We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
+derivative) of the logistic curve at the age at which most participants were acquiring a given word.
+This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
+
+```{r}
+#| label: tbl-results-fixed
+#| tbl-cap: "Posterior distribution of regression coefficients.
+*Median*: median of the posterior distribution in the logit scale.
+*95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
+*ROPE Prob.*: percentage of the posterior samples that fall within the region o practical equivalence (*ROPE*, [-0.1, 0.1] in the logit scale); posterior probability of the true value of the parameter being equivalent to zero."
+#| echo: false
+#| message: false
+#| warning: false
 # summarise posterior draws
 posterior_description %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
     mutate(
@@
         cells_column_labels(columns = 1:5)
-    )
-```
-
-@tbl-results-fixed shows the summary of the posterior distribution of the fixed regression coefficients of the model.
-For interpretability, we supplement the summaries of each posterior regression coefficient with its transformation to the probability scale.
-The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
-This means that one logit difference is not necessarily translated to a unique value in the probability scale.
-For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
-The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
-We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
-derivative) of the logistic curve at the age at which most participants were acquiring a given word.
-This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
+    ) %>% 
+    tab_style(
+        cell_text(align = "left"),
+        cells_title(groups = "subtitle")
+    )
+```
 
@@
 #| echo: false
+#| eval: false
 #| message: false
@@
         ymax = Inf,
-        xmin = rope_interval["lower"],
-        xmax = rope_interval["upper"],
+        xmin = -0.10,
+        xmax = 0.10,
+        # xmin = rope_interval["lower"],
+        # xmax = rope_interval["upper"],
         colour = NA,
@@
         aes(
-            fill = stat(abs(x) < rope_interval["upper"]),
-            colour = stat(abs(x) < rope_interval["upper"]),
+            # fill = stat(abs(x) < rope_interval["upper"]),
+            # colour = stat(abs(x) < rope_interval["upper"]),
+            fill = stat(abs(x) < 0.10),
+            colour = stat(abs(x) < 0.10)       
         ),
@@
     ) +
-    scale_x_continuous(breaks = seq(-0.5, 2, 0.2)) +
+    scale_x_continuous(breaks = seq(-0.4, 2, 0.2)) +
     scale_fill_manual(
@@
 ```
-
 The degree of exposure to the language ($DoE$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std"]]$ci_high, accuracy = 0.01)`]).
@@
 #| warning: false
-#| fig-width: 7
+#| fig-width: 8
+#| fig-height: 5
 #| fig-cap: "Expected posterior predictions.
@@
 Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
-Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
+Finally, predictions are presented separately for different degrees of exposure (*DoE*): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
 Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
@@
     age_std = (10:36-mean(df$age))/(sd(df$age)),
-    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe))
-    # lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
-)
-
-m <- add_epred_draws(nd, model_fit_1, ndraws = NULL, re_formula = NA) %>% 
+    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe)),
+    lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
+)
+
+m <- add_epred_draws(nd, model_fit_4, ndraws = NULL, re_formula = NA) %>% 
     mutate(
@@
             labels = paste0("DoE: ", c(10, 50, 90), "%")
-        )
-        # lv_std = factor(
-        #     lv_std, 
-        #     levels = unique(nd$lv_std),
-        #     labels = paste0(c(0, 50, 100), "%")
-        # )
+        ),
+        lv_std = factor(
+            lv_std,
+            levels = unique(nd$lv_std),
+            labels = paste0(c(0, 50, 100), "%")
+        )
     ) %>% 
@@
     mutate(
-        .category = str_replace_all(
+        .category = factor(
             .category,
-            c(
-                "Understands and Says" = "Production\n(Understands and Says))",
-                "Understands" = "Comprehension\n(Understands)"
+            levels = c(
+                "Understands",
+                "Understands and Says"
+            ),
+            labels = c(
+                "Comprehension\n(Understands)",
+                "Production\n(Understands and Says)"
             )
@@
             factor(., levels = paste0("DoE: ", percent(c(0.1, 0.5, 0.9)))),
-        # lv_std = case_when(
-        #     between(lv, 0.00, 0.20) ~ "0%",
-        #     between(lv, 0.40, 0.60) ~ "50%",
-        #     between(lv, 0.80, 1.00) ~ "100%"
-        # ) %>% 
-        #     factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
+        lv_std = case_when(
+            between(lv, 0.00, 0.20) ~ "0%",
+            between(lv, 0.40, 0.60) ~ "50%",
+            between(lv, 0.80, 1.00) ~ "100%"
+        ) %>%
+            factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
         age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
@@
     drop_na(doe_std) %>% 
-    group_by(te, age_std, doe_std) %>% 
+    group_by(te, age_std, doe_std, lv_std) %>% 
     summarise(
@@
         y = .epred, 
-        colour = doe_std,
-        fill = doe_std
-    ) +
-    facet_grid(~.category) +
+        colour = lv_std,
+        fill = lv_std
+    ) +
+    facet_grid(.category~doe_std) +
     stat_lineribbon(
@@
     ) +
-    stat_summary(
-        data = d,
-        fun.data = "mean_se",
-        geom = "pointrange",
-        size = 0.25
-    ) +
+    # stat_summary(
+    #     data = d,
+    #     fun.data = "mean_se",
+    #     geom = "pointrange",
+    #     size = 0.25
+    # ) +
     # geom_point(
@@
     scale_x_continuous(
-        breaks = (seq(10, 36, 2)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 2)
+        breaks = (seq(8, 38, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(8, 38, 4)
     ) +
     scale_color_d3() +
+    scale_fill_d3() +
     labs(
         x = "Age (months)", 
-        y = "Posterior probability of acquisition\n(Linear prediction)",
-        colour = "Degree of exposure",
-        fill = "Degree of exposure"
+        y = "Probability of acquisition",
+        colour = "Levenshtein (Phonological similarity)",
+        fill = "Levenshtein (Phonological similarity)"
     ) +
@@
 
-# Discussion {#discussion}
+# Discussion {#sec-discussion}
 
@@
 
-This study explored how word acquisition was affected by cognateness in bilingual toddlers.
-We collected vocabulary data from children learning Catalan and/or Spanish in differing degrees of exposure to each, and then measured the probability of each word being acquired at different ages as a function of (1) participants' exposure to the language it belongs to, and (2) its phonological overlap with its translation equivalent in the other language.
-We build a multilevel Bayesian model that estimated the effect of both predictors and their interaction, adjusting for other variables known to be relevant during the acquisition of words: participants' age, lexical frequency, and length in phonemes.
-
-## Summary of the results
-
-## Results in contrast with the literature
+* This study explored how word acquisition is affected by cognateness in bilingual toddlers.
+* We collected vocabulary data from children learning Catalan and/or Spanish in differing degrees of exposure to each, and then measured the probability of each word being acquired at different ages as a function of (1) participants' exposure to the language it belongs to, and (2) its phonological similarity with its translation equivalent in the other language.
+* Using item response theory (*IRT*), we built an Bayesian model that estimated the effect of both predictors and their interaction, adjusting for other variables known to be relevant for the acquisition of words: participants' age, lexical frequency, and word length in phonemes.
+
+## Summary of results and contrast with the literature
+
+* We replicated effects reported in previous studies, such as age being the most determinant predictor of word acquisition, followed by exposure to the language, that frequent words are acquired earlier, and that word length predicts a later age of acquisition.
+* Regarding our hypotheses, we found strong evidence for a two-way interaction between language exposure and phonological similarity: participants were more likely to acquire word-forms from cognate translation equivalents than from non-cognate translation equivalents, but only when their exposure to the language of such word-form was low.
+* Our results converge with previous findings indicating that phonological similarity between translation equivalents plays a facilitation role in word acquisition, but suggest that this effect is mostly, or only present in participants' language of lower exposure.
 
@@
 
+* These findings support the hypothesis that the cognate facilitation effect on word acquisition is more likely to play such role *after* one of the word-forms of the translation equivalent has been acquired in one of the languages, as opposed to taking place before any of the forms has been acquired, by promoting their acquisition though cross-language activation at the sub-lexical, phonological level.
+* In line with the first hypothesis, words in the language of higher exposure are likely to be acquired earlier than those in the language of lower exposure, and also showed a larger effect facilitation effect of cognateness.
+* This suggests that cognateness might have played a larger role during the acquisition of words from the language of lower exposure than from the language of higher exposure in bilinguals
+
 ## Limitations
 
+* These results come from a particular population of bilingual toddlers that learn two very similar languages--Catalan and Spanish--and are therefore exposed to many cognates.
+Our results might not generalise to other groups of bilinguals learning two less similar languages.
+* [To be discussed]
+
 ## Further steps
-
-# References {#references}
+ * [To be discussed]
+
+# References {#sec-references}
 
@@
 
-#  Appendix {#appendix}
-
-## Appendix 1: Words {#appendix-words}
+#  Appendix {#sec-appendix}
+
+## Appendix 1: Words {#sec-appendix-words}
 
@@
 
-## Appendix 2: *ROPE* specification {#appendix-words}
-
-We decided to draw the *ROPE* interval around zero so that it covered values smaller the absolute median of those relevant by similar previous studies.
-After consulting the regression tables of @mitchell2022cognates, @braginsky2019consistency, and @laing2022phonological, we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: `r round(median(abs(rope_coefs$estimate)), 3)`, in the logit scale.
+## Appendix 2: Model details {#sec-appendix-model}
+
+## Appendix 3: *ROPE* specification {#sec-appendix-rope}

### 2022-11-17 manuscript.tex
Update lab notes

@@
 \maketitle
-\begin{abstract}
-Bilinguals face the challenging task of learning words from languages
-with overlapping phonologies. Floccia et al.~(2018) reported larger
-vocabulary sizes for 24-month-old bilinguals that were learning
-languages that shared a greater amount of cognates (e.g.,
-English-Dutch). The mechanisms underlying this effect remain unknown. We
-explore two compatible scenarios. First, we test whether cognates are
-learnt earlier than non-cognates. This would account for the difference
-in vocabulary size associated to the amount of shared cognates across
-languages. Second, we explore the possibility that the word-forms of one
-language interact with those form the other language, scaffolding the
-acquisition of their translation equivalents when their phonologies
-overlap. This mechanism, in line with the parallel activation account of
-bilingual speech perception, would provide a plausible explanation to
-why cognates are acquired ealier by bilinguals. We developed an online
-tool to collect parental reports of receptive and productive
-vocabularies from children learning Catalan and/or Spanish, and present
-data on receptive and productive vocabulary of bilingual toddlers aged
-10 to 36 months.
-\end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
 \setstretch{1.25}
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\hypertarget{sec-introduction}{%
+\section{Introduction}\label{sec-introduction}}
+
+\hypertarget{word-acquisition-during-toddlerhood-the-effect-of-language-experience}{%
+\subsection{Word acquisition during toddlerhood: the effect of language
+experience}\label{word-acquisition-during-toddlerhood-the-effect-of-language-experience}}
 
@@
 category of nouns, as opposed to verbs, adjectives, or function words
-(e.g., Nelson, 1973, p. bates1994developmental).
+Bates et al. (1994).
 
@@
 
-There is evidence that bilinguals know, on average, less words in each
-of their languages than monolinguals, but also that both groups know a
-similar amount of words when the two languages are aggregated. For
-example, Hoff et al. (2012) found that English-Spanish bilingual
-toddlers in South Florida knew less words in English than monolinguals,
-who only learnt English. Both groups knew a similar total amount of
-words when both English and Spanish vocabularies were counted together,
-pointing to the importance of collecting data on both languages when
-assessing bilinguals' communicative development. Other studies have
-provided converging evidence that bilinguals know a similar--or even
-larger--number of words than monolinguals, only when the languages are
-aggregated (Gonzalez-Barrero et al., 2020; Oller \& Eilers, 2002; J.
-Patterson et al., 2004; J. L. Patterson, 2004; Pearson et al., 1993;
-Pearson \& Fernández, 1994; Petitto et al., 2001; Smithson et al.,
-2014). While these studies have mostly relied on samples of bilingual
-children learning two relatively distant languages, as it is the case of
-English and Spanish, it is unclear whether children learning
+Previous findings point to bilingual toddlers knowing, on average, less
+words in each of their languages than monolinguals, but also to both
+groups knowing a similar amount of words when the two languages are
+aggregated. For example, Hoff et al. (2012) found that English-Spanish
+bilingual toddlers in South Florida knew less words in English than
+monolinguals, who only learnt English. Both groups knew a similar total
+amount of words when both English and Spanish vocabularies were counted
+together, highlighting the importance of collecting data on both
+languages when assessing bilinguals' communicative development. Other
+studies have provided converging evidence that bilinguals know a
+similar--or even larger--number of words than monolinguals, only when
+the languages are aggregated (Gonzalez-Barrero et al., 2020; Oller \&
+Eilers, 2002; J. Patterson et al., 2004; J. L. Patterson, 2004; Pearson
+et al., 1993; Pearson \& Fernández, 1994; Petitto et al., 2001; Smithson
+et al., 2014). While these studies have mostly relied on samples of
+bilingual children learning two relatively distant languages, as it is
+the case of English and Spanish, it is unclear whether children learning
 typologically more similar languages also know less words in each of
@@
 play during early vocabulary growth?
+
+\hypertarget{language-distance-as-a-mediator-of-bilingual-vocabulary-growth}{%
+\subsection{Language distance as a mediator of bilingual vocabulary
+growth}\label{language-distance-as-a-mediator-of-bilingual-vocabulary-growth}}
 
@@
 
+\hypertarget{a-candidate-account-lexical-non-selectiviy-and-parallel-activation}{%
+\subsection{A candidate account: lexical non-selectiviy and parallel
+activation}\label{a-candidate-account-lexical-non-selectiviy-and-parallel-activation}}
+
 Floccia et al.~pointed to \emph{parallel activation} as the main
-mechanism underpinning their results. The parallel activation hypothesis
-suggests that bilinguals activate both languages simultaneously during
-speech production or comprehension, and that this phenomenon is the
-result of the activation of lexical representations in both languages,
-even when only one is in use during production (Costa et al., 2000;
-Hoshino \& Kroll, 2008) or comprehension (Spivey \& Marian, 1999;
-Thierry \& Wu, 2007). One of the clearest examples of parallel
+phenomenon underpinning their results. The parallel activation
+hypothesis suggests that bilinguals activate both languages
+simultaneously during speech production or comprehension, and that this
+phenomenon is the result of the activation of lexical representations in
+both languages, even when only one is in use during production (Costa et
+al., 2000; Hoshino \& Kroll, 2008) or comprehension (Spivey \& Marian,
+1999; Thierry \& Wu, 2007). One of the clearest examples of parallel
 activation was provided by Costa et al. (2000). In this study,
@@
 al., 2013; e.g., Von Holzen \& Mani, 2012). Parallel activation is
-therefore a plausible mechanism to account for Floccia et al.'s results:
-cognates increase the amount of activation in both languages,
+therefore a plausible explanation for Floccia et al.'s results:
+cognateness increases the amount of cross-language activation,
 facilitating word acquisition, and ultimately leading to children
 learning language pairs with a larger proportion of cognates are
-predicted to show larger vocabulary sizes. A missing piece in this
-account is how increase parallel activation might facilitate the
-acquisition new words.
-
-In this study, we hypothesise that cognates are acquired earlier in age
-than non-cognates, which might explain why children learning languages
-that share many cognates show larger vocabulary sizes. Evidence
-supporting this claim is scarce. Bosch \& Ramon-Casas (2014) used
-vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish
-bilinguals aged 18 months, and found that cognates represented a larger
-proportion of participant's vocabulary than non-cognates. Schelletter
-(2002) reported a longitudinal single case of one English-German
-bilingual who produced cognates earlier than non-cognates, on average.
-The low sample size in these two studies makes it challenging to draw
-strong conclusions about the effect of cognateness on vocabulary growth.
-On the other hand, Floccia et al.'s estimates are statically more
-reliable given their (much larger) sample size, but their study was not
-aimed at testing the effect of cognateness on age of acquisition
-directly. In their discussion the authors state the following (pp.~70):
+predicted to show larger vocabulary sizes. This account is is line with
+previous studies suggesting that the acquisition of new words is
+facilitated by their phonological or semantic similarity with other
+words, already acquired or not (Fourtassi et al., 2020; e.g., Hills et
+al., 2009; Jones \& Brandt, 2019; Laing, 2022). Importantly, bilinguals
+seem to be more likely to acquire words for which their translation
+equivalent has been already learnt in the other language, suggesting
+that semantic similarity also facilitates word acquisition across
+languages (Bilson et al., 2015).
+
+If phonological similarity also plays a cross-language facilitation role
+during word acquisition, cognate translation equivalents--which share
+both semantic and phonological similarity--should be acquired earlier
+that non-cognate translation equivalents--which share semantic
+similarity, but not phonological similarity. Evidence supporting and
+earlier age of acquisition for cognates is scarce. Bosch \& Ramon-Casas
+(2014) used vocabulary parental reports (152 lexical items) from 48
+Catalan-Spanish bilinguals aged 18 months, and found that cognates
+represented a larger proportion of participant's vocabulary than
+non-cognates. Schelletter (2002) reported a longitudinal single case of
+one English-German bilingual who produced cognates earlier than
+non-cognates, on average. The low sample size in these two studies makes
+it challenging to draw strong conclusions about the effect of
+cognateness on vocabulary growth. On the other hand, Floccia et al.'s
+estimates are statically more reliable given their (much larger) sample
+size, but their study was not aimed at testing the effect of cognateness
+on age of acquisition directly. In their discussion the authors state
+the following (pp.~70):
 
@@
 
-As Mitchell et al. (2022) note in their discussion, two competing
-hypotheses remain untested. The facilitative effect of cognateness can
-be explained by, at least, two alternative, arguably incompatible,
-mechanisms. On the one hand, cognateness might operate as a facilitator
-of the acquisition of both forms of a translation pair, with every
-exposure to either form activating the other, and therefore
-strengthening the association between both and their referent. If this
-is the case, then cognateness should increase the probability of
-acquisition of the forms in both languages equally.
+\hypertarget{two-competing-mechanisms-for-a-cognate-facilitation-during-word-acquisition}{%
+\subsection{Two competing mechanisms for a cognate facilitation during
+word
+acquisition}\label{two-competing-mechanisms-for-a-cognate-facilitation-during-word-acquisition}}
+
+As Mitchell et al. (2022) note in their discussion, the facilitative
+effect of cognateness on word acquisition can be explained by at least
+two alternative but arguably incompatible mechanisms. On the one hand,
+cognateness might operate at the pre-lexical level as a bidirectional
+facilitator of the acquisition of the two word-forms of a translation
+pair: every exposure to either word-form activates the other, and
+therefore strengthens the association between both word-forms and their
+referent with every exposure. For instance, hearing \emph{flower} in
+English might activate \emph{floeur} in French, and \emph{vice versa}.
+Repeated exposure to the English or the French words eventually leads to
+an earlier acquisition for both of them, as compared to non-cognate
+translation equivalents like \emph{cow} and \emph{vache}, which do not
+activate each other via phonological similarity. If this is the case,
+then cognateness should increase the probability of acquisition of the
+word-forms in both languages equally.
 
@@
 of a translation pair only when at least one of the word-forms has been
-acquired. This account departs from the premise that, if the child is
-not previously familiar with one of the word forms, the acquisition of
-its translation in the other language cannot be facilitated by
-cognateness. This implies that, if cognateness facilitates the
-acquisition of a word form, it can only do so after the child has
-acquired one of the word forms of the translation pair. Given that
-children are more likely to acquire words from languages to which they
-are exposed more often (Cattani et al., 2014; David \& Wei, 2008;
+acquired. This account departs from the premise that phonological
+similarity operates at the lexical level to facilitate word learning
+across languages: if the child has still not formed a lexical
+representation with sufficient phonological detail for one of the word
+forms, the acquisition of its translation in the other language cannot
+be facilitated by cognateness. This implies that, if cognateness
+facilitates the acquisition of a word form, it can only do so after the
+child has acquired one of the word forms of the translation pair. Given
+that children are more likely to acquire words from languages to which
+they are exposed more often (Cattani et al., 2014; David \& Wei, 2008;
 Thordardottir et al., 2006), the acquisition of words in the language of
@@
 
-In line with this second hypothesis, the vocabulary size associated with
-linguistic similarity that Floccia et al.~reported was larger in the
-additional language vocabulary than in English vocabulary. Most
+In line with this second hypothesis, the size of the effect of
+linguistic similarity on vocabulary size that Floccia et al.~reported
+was larger in the additional language (language of lower exposure)
+vocabulary than in English vocabulary (language of higher exposure. Most
 participants in their sample were English-dominant, meaning that their
@@
 
+\hypertarget{the-present-study}{%
+\subsection{The present study}\label{the-present-study}}
+
 In this study, we investigated the role of cognateness on the
-acquisition of translation equivalent, and whether such effect is
+acquisition of translation equivalents, and if such effect is
 conditional to the amount of time the child is exposed to the
@@
 Catalan and/or Spanish, with varying degrees of exposure to each
-language. We then adopted an Item Response Theory (\emph{IRT}) approach
-to model the probability of participants being reported by their parents
-to understand or understand \emph{and} say each word in the checklist,
-conditional to its cognate status in Catalan and Spanish and
-participants' degree of exposure to the corresponding language, while
-adjusting for other indices of word difficulty--lexical frequency and
-length (phonemes)-- and participant skill--age (see Kachergis et al.,
-2022 for a similar approach).
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-Data and materials are available at OSF \url{https://osf.io/hy984/}, and
-code is available at GitHub
-\url{https://github.com/gongcastro/trajectories}.
-
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
+language. We then adopted an Bayesian item response theory (\emph{IRT})
+approach to model the probability of participants being reported by
+their parents to understand or understand \emph{and} say each word in
+the checklist, conditional to its cognate status in Catalan and Spanish
+and participants' degree of exposure to the corresponding language,
+while adjusting for other indices of word difficulty--lexical frequency
+(Goodman et al., 2008) and length (phonemes)-- and participant
+skill--age (see Kachergis et al., 2022 for a similar approach).
+
+\hypertarget{sec-methods}{%
+\section{Methods}\label{sec-methods}}
+
+\hypertarget{sec-participants}{%
+\subsection{Participants}\label{sec-participants}}
 
@@
 
-\hypertarget{questionnaire}{%
-\subsection{Questionnaire}\label{questionnaire}}
-
-The questionnaire was implemented on-line using the formR platform
-(Arslan et al., 2020), and was structured in three blocks: a (1)
-language questionnaire, a (2) demographic survey, and a (3) Catalan and
-a Spanish vocabulary checklists. Vocabulary checklists followed a
-similar structure as the Oxford Communicative Developmental Inventory
-(Hamilton et al., 2000) and consisted in two lists of words: one in
-Catalan and one in Spanish. The Catalan inventory contained 798 items
-and the Spanish inventory contained 802. Items in one language were
-translation equivalents of the items in the other language (e.g.,
-whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the
-word \emph{perro} was included in the Spanish inventory), roughly
-following a one-to-one mapping. In case two translation equivalents were
-possible for a given word, both were included as separate items (e.g.,
-Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar}
-and \emph{terminar}), or merged them into a single item (e.g., Spanish
+\hypertarget{sec-questionnaire}{%
+\subsection{Questionnaire}\label{sec-questionnaire}}
+
+To collect vocabulary data from participants, we created an \emph{ad
+hoc} questionnaire, MultiLex (Garcia-Castro, 2022). This questionnaire
+was inspired in the MacArthur-Bates Communicative Development Inventory
+(Fenson et al., 2007) and its adaptations to other languages, and was
+implemented on-line using the formR platform (Arslan et al., 2020). This
+questionnaire is structured in three blocks: a (1) language
+questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish
+vocabulary checklists. Vocabulary checklists followed a similar
+structure as the Oxford Communicative Developmental Inventory (Hamilton
+et al., 2000) and consisted in two lists of words: one in Catalan and
+one in Spanish. The Catalan inventory contained 798 items and the
+Spanish inventory contained 802. Items in one language were translation
+equivalents of the items in the other language (e.g., whenever
+\emph{gos} {[}dog{]} was included in the Catalan inventory, the word
+\emph{perro} was included in the Spanish inventory), roughly following a
+one-to-one mapping. In case two translation equivalents were possible
+for a given word, both were included as separate items (e.g., Catalan
+\emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar} and
+\emph{terminar}), or merged them into a single item (e.g., Spanish
 \emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}). We
@@
 bar{]}) and items that included more than one word-form (e.g.,
-\emph{mono / mico}).
-
-We extracted four variables of interest for each word-form: (1) Word
+\emph{mono/mico}).
+
+\hypertarget{variables-of-interest}{%
+\subsection{Variables of interest}\label{variables-of-interest}}
+
+We extracted four variables of interest for each word-form. (1) Word
 lexical frequency (\(Frequency\)): lexical frequency of the word in its
@@
 \begin{longtable}{lrrrrrrrrrr}
-\caption{\label{tbl-item-summary}Lexical properties of the word-forms and translationpairs included in
+\caption{\label{tbl-item-summary}Lexical properties of the word-forms and translation pairs included in
 the vocabulary checklist and our analyses. We report the number of
-items, mean (M), standard deviation (SD), and range of the lexical
-frequency (epxressed in Zipf scores), length (number of phonemes) of the
-word forms, and the phonological similarity (normalised Levenshtein
+items, mean, standard deviation (SD), and range of the lexical frequency
+(epxressed in Zipf scores), length (number of phonemes) of the word
+forms, and the phonological similarity (normalised Levenshtein
 similarity) with their translation equivalent. Statistics are reported
@@
 \toprule
- &  & \multicolumn{3}{c}{Frequency (Zipf)} & \multicolumn{3}{c}{\# Phonemes} & \multicolumn{3}{c}{Phon. similarity (Levenshtein)} \\ 
+ &  & \multicolumn{3}{c}{Frequency (Zipf)} & \multicolumn{3}{c}{Length (Phonemes)} & \multicolumn{3}{c}{Phon. similarity (Levenshtein)} \\ 
 \cmidrule(lr){3-5} \cmidrule(lr){6-8} \cmidrule(lr){9-11}
@@
 
-\hypertarget{analysis}{%
-\subsection{Data analysis}\label{analysis}}
+\hypertarget{sec-analysis}{%
+\subsection{Data analysis}\label{sec-analysis}}
 
@@
 participants (of which 343 participated once, 33 twice, 14 three times,
-and 5 four times), responding to 1199 distinct items (597 in Catalan,
+and 5 four times), responding to 1,199 distinct items (597 in Catalan,
 602 in Spanish, 602 translation equivalents). Translation equivalents
@@
 (\emph{No} \textless{} \emph{Understands} \textless{} \emph{Understands
-and Says}) using a Bayesian, mixed-effects ordinal regression model.
-This allowed us to simultaneously estimate the impact of item- and
-participant-level predictors on the probability of word comprehension
-and production. We included \(Age\), \(Frequency\), \(Phonemes\),
-\(DoE\), and \(Levenshtein\) as fixed effects. We also included the
-\(Age \times DoE \times Levenshtein\) three-way interaction, and its
-corresponding two-way interactions. We added participants and items as
-crossed random effects including all intercepts and slopes when
-appropriate (Barr et al., 2013).
-
-This model structure, including word- and participant-level properties
-to estimate vocabulary acquisition trajectories was discussed by
-Kachergis et al. (2022) in the context of explanatory item response
-models. Our model differs from their proposal in that 1) while they
-modelled global trajectories exclusively, we modelled individual
-trajectories as well, 2) our outcome variable is ordinal in contrast to
-the binary outcome used in their study, 3) we propose a Bayesian
-approach. Our motivation for a Bayesian approach towards statistical
-inference is that (1) it allows us to incorporate previous domain
-knowledge to improve estimation accuracy, and (2) to quantify the
-uncertainty associated with the estimated parameters in our model
-(McElreath, 2018).
-
-The Bayesian approach implies using Bayes' rule to assign a probability
-to possible parameter values (i.e., posterior distribution), based on
-previous knowledge about such probability distribution (prior
-distribution) and our data (likelihood). We used Markov chain Monte
-Carlo (MCMC) to approximate the posterior distribution via sampling. We
-run 2 sampling chains with 4,000 iterations each, including 2,000
-warm-up iterations per chain. All \(\hat{R}\) values were smaller than
-1.1, indicating that the chains converged successfully in all cases
-(Gelman \& Rubin, 1992).
-
-We compared a base model---only including the intercept, \(Age\),
-\(Frequency\), \(Phonemes\), and \(DoE\) as fixed effects, and their
-corresponding random effects-\/--to increasingly more complex models,
-introducing predictors in the following order: \(Age \times DoE\),
-\(Levenshtein\), \(DoE \times Levenshtein\), and finally
-\(Age \times Levenshtein\) and \(Age \times DoE \times Levenshtein\). We
-compared the predictive performance of the models using Bayesian
-leave-one-out cross-validation (LOO) (Vehtari et al., 2017): we computed
-the average likelihood of each observation after estimating the model's
-parameters leaving that same observation out of the data set. Due to
-computational constraints, we performed LOO based on a random sub-sample
-of 700 samples (out of 2,000) of the posterior distribution of the fixed
-effects of the model. and selected the model with the smallest expected
-log posterior density (ELPD)\footnote{Due to the high computational cost
-  associated with our large dataset, we used a sub-sampling approach for
-  performing Bayesian LOO with 700 posterior samples (Magnusson et al.,
-  2019)}.
-
-Finally, we explored the posterior distribution of each parameter in the
-selected model. Following Kruschke \& Liddell (2018), We specified a
-region of practical equivalence (\emph{ROPE}) around the values of each
-regression coefficient that we considered equivalent to zero (Kruschke
-\& Liddell, 2018). We defined our \emph{ROPE} from -0.18 to +0.18 {[}see
-@\#appendix-rope for justification of the \emph{ROPE}{]}. In the results
-section, we report the mean of each coefficient's marginal posterior
-distribution along with their 95\% highest density intervals
-(\emph{HDI}). This interval indicates the narrowest range of values that
-contain 95\% of the samples of the posterior distribution, and therefore
-the true value of the coefficient with 95\% confidence, given our data.
-We also report the proportion of the posterior samples of that
-coefficient that fell within the \emph{ROPE}. A 90\% overlap between the
-95\% \emph{HDI} of the coefficients' posterior distribution and the
-\emph{ROPE} indicates that, given the data, there is a 90\% probability
-that the true value of the coefficient falls within the \emph{ROPE}, and
-is therefore equivalent to zero, that is, negligible.
+and Says}) using a Bayesian, mixed-effects ordinal regression model (2
+sampling chains with 2,000 iterations each). This model allowed us to
+estimate both item and participant word-acquisition trajectories, as a
+function of different fixed effects from \(Age\), \(Frequency\),
+\(Phonemes\), \(DoE\), and \(Levenshtein\). We also included participant
+and item random intercepts and slopes when appropriate (Barr et al.,
+2013) see~\ref{sec-appendix-words}. We compared a baseline model
+\(\mathcal{M}_{0}\), which only included the intercept, \(Age\),
+\(Frequency\), \(Phonemes\), and \(DoE\) as predictors, with
+increasingly more complex models, introducing predictors in the
+following order: \(Age \times DoE\) (\(\mathcal{M}_{1}\)),
+\(Levenshtein\) (\(\mathcal{M}_{2}\)), \(DoE \times Levenshtein\)
+(\(\mathcal{M}_{3}\)), and finally \(Age \times Levenshtein\) and
+\(Age \times DoE \times Levenshtein\) (\(\mathcal{M}_{4}\)). We compared
+the predictive performance of the models using Bayesian leave-one-out
+cross-validation (LOO, Vehtari et al., 2017)\footnote{Due to
+  computational constraints, we performed LOO based on a random
+  sub-sample of 700 samples of the posterior distribution of the fixed
+  effects (Magnusson et al., 2019)}, and selected for interpretation the
+one showing the best performance.
+
+We assessed the practical relevance of the estimated regression
+coefficients of the selected model by specifying a region of practical
+equivalence (\emph{ROPE}, Kruschke \& Liddell, 2018) from -0.1 to +0.1
+see~\ref{sec-appendix-rope}. This region indicates the range of values
+that we considered as equivalent to zero. The degree of overlap between
+the posterior distribution of a regression coefficient with the
+\emph{ROPE} can be interpreted as support for the true parameter of the
+coefficient being zero. In the results section, we report the 95\%
+highest density interval (\emph{HDI}) of the posterior distribution of
+each regression coefficient. This is interval contains the true value of
+this coefficient with 95\% probability, given our data. We also report
+the proportion of posterior samples in the 95\% \emph{HDI} that overlap
+with the \emph{ROPE}, which indicates the probability that the true
+value of the regression coefficient falls into the \emph{ROPE}, and
+should therefore be considered equivalent to zero\footnote{For example,
+  a 90\% overlap between the \emph{HDI} of the coefficients' posterior
+  distribution and the \emph{ROPE} indicates that, given our data, there
+  is a 90\% probability that the true value of the coefficient falls
+  within the \emph{ROPE}, and is therefore equivalent to zero}.
 
@@
 Carpenter et al., 2017), \texttt{loo} (Yao et al., 2017), and
-\texttt{tidybayes} (Kay, 2021) R packages.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{comparison}{%
-\subsection{Model comparison and selection}\label{comparison}}
-
-Table~\ref{tbl-results-loos} shows the outputs of the \emph{LOO-CV}. The
-extended model (\texttt{model\_fit\_1}), which included the three-way
+\texttt{tidybayes} (Kay, 2021) R packages. Data and materials are
+available at OSF \url{https://osf.io/hy984/}, and code is available at
+GitHub \url{https://github.com/gongcastro/trajectories}.
+
+\hypertarget{sec-results}{%
+\section{Results}\label{sec-results}}
+
+\hypertarget{sec-comparison}{%
+\subsection{Model comparison and selection}\label{sec-comparison}}
+
+The extended model (\(\mathcal{M}_{4}\)), which included the three-way
 interaction between \(Age\), \(DoE\), and \(Levenshtein\) showed the
-best predictive performance, with a difference in \emph{ELPD} compared
-to the other models several times larger than the standard error of such
-difference. This indicates that, given the data, the predictions of the
-extended model are confidently more accurate than those of the other
-models.
+best predictive performance, with a difference in expected log-predicted
+density (\emph{ELPD}) compared to the other models several times larger
+than the standard error of such difference see~\ref{tbl-results-loos}.
+This indicates that, given the data, the predictions of the
+\(\mathcal{M}_{4}\) are confidently more accurate than those of the
+other models.
 
@@
 \setlength{\LTpost}{0mm}
-\begin{longtable}{lrrrrrrrr}
-\caption{\label{tbl-results-loos}Bayesian leave-one-out cross validation (LOO-CV) }\tabularnewline
-
-\caption*{
-{\large } \\ 
-{\small The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (\emph{ELPD}), which penalises the complexity of each model.
-\emph{ELPD} values closer to zero indicate better predictive acuracy.
-We report the \emph{ELPD}, effective number of parameters (\emph{p}), and information criterion (\emph{IC}) of each model along their associated standard errors (\emph{SE}).
-The last column indicates the difference in \emph{ELPD} between each model and the model with the best predictive accuracy (\texttt{model\_fit\_1})}
-} \\ 
+\begin{longtable}{rrrrrrrrr}
+\caption{\label{tbl-results-loos}Bayesian leave-one-out cross validation (LOO-CV). The predictive
+accuracy of the models was compared using leave-one-out
+cross-validation. The likelihood of each data point given a model fitted
+without including that same observation was estimated. The resulting
+likelihoods of each model were summarised as its expected log-predicted
+density (ELPD), which penalises the complexity of each model. ELPD
+values closer to zero indicate better predictive acuracy. We report the
+ELPD, effective number of parameters (p), and information criterion (IC)
+of each model along their associated standard errors (SE). The last
+column indicates the difference in ELPD between each model and the model
+with the best predictive accuracy (model 3). }\tabularnewline
+
 \toprule
@@
 \midrule
-model\_fit\_3 & $-244,590.04$ & $388.23$ & $1,688.79$ & $5.39$ & $489,180.08$ & $776.46$ & $0.00$ & $0.00$ \\ 
-model\_fit\_0 & $-247,891.36$ & $353.08$ & $1,260.47$ & $3.24$ & $495,782.71$ & $706.16$ & $-3,301.32$ & $161.73$ \\ 
-model\_fit\_2 & $-248,274.01$ & $353.95$ & $1,489.46$ & $3.65$ & $496,548.01$ & $707.91$ & $-3,683.97$ & $152.45$ \\ 
-model\_fit\_1 & $-249,630.81$ & $358.40$ & $1,649.76$ & $4.03$ & $499,261.61$ & $716.79$ & $-5,040.77$ & $170.07$ \\ 
+4 & $-252,321.17$ & $357.98$ & $2,059.83$ & $5.72$ & $504,642.34$ & $715.96$ & $0.00$ & $0.00$ \\ 
+0 & $-252,336.70$ & $356.17$ & $1,252.77$ & $3.17$ & $504,673.41$ & $712.35$ & $-15.53$ & $61.09$ \\ 
+1 & $-252,346.99$ & $359.27$ & $1,720.27$ & $5.16$ & $504,693.98$ & $718.54$ & $-25.82$ & $52.40$ \\ 
+3 & $-252,480.77$ & $361.35$ & $1,952.24$ & $5.38$ & $504,961.53$ & $722.70$ & $-159.60$ & $32.20$ \\ 
+2 & $-253,895.47$ & $363.30$ & $2,086.79$ & $5.73$ & $507,790.94$ & $726.60$ & $-1,574.30$ & $66.98$ \\ 
 \bottomrule
@@
 
-\hypertarget{coefficients}{%
-\subsection{Regression coefficients}\label{coefficients}}
-
-\hypertarget{tbl-results-fixed}{}
-\begin{longtable}{lrrr}
-\caption{\label{tbl-results-fixed}Summary of the estimated posterior distribution of fixed regression
-coefficients }\tabularnewline
-
-\toprule
-Parameter & Median & 95\% \emph{HDI} & \emph{ROPE} prob. \\ 
-\midrule
-Intercept (Comprehension and Production at 22 months) & $-0.85$ & [$-0.97$, $-0.72$] & $0.00\%$ \\ 
-Intercept (Comprehension at 22 months) & $1.55$ & [$1.44$, $1.69$] & $0.00\%$ \\ 
-Age (+1 SD, 5.72 months) & $1.32$ & [$1.17$, $1.47$] & $0.00\%$ \\ 
-Frequency (+1 SD, 0.7 Zipf) & $0.16$ & [$0.07$, $0.24$] & $32.11\%$ \\ 
-Phonemes (+1 SD, 1.75 phonemes) & $-0.20$ & [$-0.28$, $-0.13$] & $2.79\%$ \\ 
-DoE (+1 SD, 30\%) & $0.89$ & [$0.87$, $0.91$] & $0.00\%$ \\ 
-Levenshtein (+1 SD, 26\%) & $0.03$ & [$-0.09$, $0.13$] & $100.00\%$ \\ 
-DoE × Levenshtein & $-0.20$ & [$-0.22$, $-0.19$] & $0.00\%$ \\ 
-Age × DoE & $0.13$ & [$-0.02$, $0.28$] & $57.42\%$ \\ 
-Age × Levenshtein & $0.07$ & [$0.03$, $0.11$] & $100.00\%$ \\ 
-Age × DoE × Levenshtein & $-0.06$ & [$-0.09$, $-0.02$] & $100.00\%$ \\ 
-\bottomrule
-\end{longtable}
+\hypertarget{sec-coefficients}{%
+\subsection{Regression coefficients}\label{sec-coefficients}}
 
 Table~\ref{tbl-results-fixed} shows the summary of the posterior
-distribution of the fixed regression coefficients of the model. For
-interpretability, we supplement the summaries of each posterior
+distribution of the fixed regression coefficients of the selected model.
+For interpretability, we supplement the summaries of each posterior
 regression coefficient with its transformation to the probability scale.
@@
 
+\hypertarget{tbl-results-fixed}{}
+\begin{longtable}{lrrr}
+\caption{\label{tbl-results-fixed}Posterior distribution of regression coefficients. Median: median of the
+posterior distribution in the logit scale. 95\% HDI: 95\% highest
+density interval (HDI) of the distribution; narrowest range of values
+that contains 95\% of the distribution, and therefore is 95\% likely to
+contain the true value, given the data. ROPE Prob.: percentage of the
+posterior samples that fall within the region o practical equivalence
+(ROPE, {[}-0.1, 0.1{]} in the logit scale); posterior probability of the
+true value of the parameter being equivalent to zero. }\tabularnewline
+
+\toprule
+Parameter & Median & 95\% \emph{HDI} & \emph{ROPE} prob. \\ 
+\midrule
+Intercept (Comprehension and Production at 22 months) & $-0.92$ & [$-1.04$, $-0.80$] & $0.00\%$ \\ 
+Intercept (Comprehension at 22 months) & $1.55$ & [$1.43$, $1.67$] & $0.00\%$ \\ 
+Age (+1 SD, 5.72 months) & $1.34$ & [$1.19$, $1.48$] & $0.00\%$ \\ 
+DoE (+1 SD, 30\%) & $0.40$ & [$0.27$, $0.51$] & $0.00\%$ \\ 
+Levenshtein (+1 SD, 26\%) & $0.03$ & [$-0.09$, $0.13$] & $92.39\%$ \\ 
+Frequency (+1 SD, 0.7 Zipf) & $0.15$ & [$0.08$, $0.23$] & $7.03\%$ \\ 
+Phonemes (+1 SD, 1.75 phonemes) & $-0.15$ & [$-0.23$, $-0.08$] & $6.89\%$ \\ 
+Age × DoE & $0.12$ & [$-0.01$, $0.24$] & $35.16\%$ \\ 
+Age × Levenshtein & $0.05$ & [$0.01$, $0.09$] & $100.00\%$ \\ 
+DoE × Levenshtein & $-0.14$ & [$-0.18$, $-0.11$] & $0.00\%$ \\ 
+Age × DoE × Levenshtein & $-0.05$ & [$-0.08$, $-0.01$] & $100.00\%$ \\ 
+\bottomrule
+\end{longtable}
+
 \(Age\) was the predictor with the largest effect on the probability of
-acquisition (\(\beta\) = 1.32, 95\% \emph{HDI} = {[}1.17, 1.47{]}): at
+acquisition (\(\beta\) = 1.34, 95\% \emph{HDI} = {[}1.19, 1.48{]}): at
 the steepest point of the acquisition trajectory of an average
-translation equivalent, the probability of acquisition increased 5.77\%
+translation equivalent, the probability of acquisition increased 5.84\%
 every month. Lexical frequency (\(Frequency\)) had a small, positive,
 but inconclusive on the probability of acquisition substantially,
-(\(\beta\) = 15.85\%, 95\% \emph{HDI} = {[}7.12\%, 23.62\%{]}): 32.11\%
+(\(\beta\) = 15.41\%, 95\% \emph{HDI} = {[}7.59\%, 23.08\%{]}): 7.03\%
 of its posterior samples overlapped with the \emph{ROPE.} The 95\%
 \emph{HDI} of the regression coefficient of the number of phonemes
-(\(Phonemes\)) did exclude the \emph{ROPE}, although not by much: 2.79\%
+(\(Phonemes\)) did exclude the \emph{ROPE}, although not by much: 6.89\%
 of its posterior samples fell within the \emph{ROPE}. This suggests that
 the number of phonemes in the word form had a small effect on the
-probability of acquisition (\(\beta\) = -20.28\%, 95\% \emph{HDI} =
-{[}-28.12\%, -12.95\%{]}). For every increase in phoneme length, the
-probability of acquisition decreased in -2.90\%.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-results-fixed-1.pdf}
-
-}
-
-\caption{\label{fig-results-fixed}Marginal posterior distribution of the
-regression coefficients of the fixed effects in the extended model.
-Distributions show the estimated likelihood density of each value in the
-parameter space (X-axis) of each coefficient (Y-axis). Intervals show
-the 95\% highest density interval (\emph{HDI}) of each distribution.
-This interval is the narrowest range of values that contains 95\% of the
-distribution, and therefore is 95\% likely to contain the true value,
-given the data. the mean and HDI limits are indicated below each
-distribution. Coefficients were transformed from the logit scale to the
-probability scale for interpretability. We used the divide-by-four rule
-to get the maximum change in probability of correct response associated
-with a unit increase in this variable, i.e.~the derivative of the
-logistic function in its mid-point.}
-
-\end{figure}
+probability of acquisition (\(\beta\) = -15.10\%, 95\% \emph{HDI} =
+{[}-22.70\%, -7.95\%{]}). For every increase in phoneme length, the
+probability of acquisition decreased in -2.16\%.
 
 The degree of exposure to the language (\(DoE\)) had a strong effect on
-the probability of acquisition (\(\beta\) = 89.26\%, 95\% \emph{HDI} =
-{[}87.10\%, 91.34\%{]}). All of the posterior samples of his regression
+the probability of acquisition (\(\beta\) = 40.00\%, 95\% \emph{HDI} =
+{[}27.03\%, 50.78\%{]}). All of the posterior samples of his regression
 excluded the \emph{ROPE}. The impact of this predictor on the
@@
 exposure to the language the word belongs to, the participant was
-29.76\% more likely to acquire it. The effect of phonological similarity
+13.34\% more likely to acquire it. The effect of phonological similarity
 by itself (as indicated by the regression coefficient of the main effect
-of \(Levenshtein\)) was equivalent to zero (\(\beta\) = 2.76\%, 95\%
-\emph{HDI} = {[}-8.61\%, 12.88\%{]}), as 100.00\% of the posterior
+of \(Levenshtein\)) was equivalent to zero (\(\beta\) = 3.12\%, 95\%
+\emph{HDI} = {[}-8.89\%, 13.19\%{]}), as 92.39\% of the posterior
 samples of its regression coefficient fell into the \emph{ROPE}.
@@
 \(Doe \times Levenshtein\) interaction excluded the \emph{ROPE} entirely
-(\(\beta\) = -20.20\%, 95\% \emph{HDI} = {[}-21.68\%, -18.83\%{]}),
+(\(\beta\) = -14.46\%, 95\% \emph{HDI} = {[}-17.79\%, -11.10\%{]}),
 suggesting that the effect of phonological similarity on a word's
@@
 \(Age \times DoE\) interaction did not exclude the \emph{ROPE}
-(\(\beta\) = 12.63\%, 95\% \emph{HDI} = {[}-1.50\%, 28.48\%{]}): 57.42\%
+(\(\beta\) = 12.32\%, 95\% \emph{HDI} = {[}-1.48\%, 24.19\%{]}): 35.16\%
 of its posterior samples fell within the \emph{ROPE}. This shows that
@@
 regression coefficient of the \(Age \times Levenshtein\) interaction
-overlapped completely with the \emph{ROPE} (\(\beta\) = 6.59\%, 95\%
-\emph{HDI} = {[}2.66\%, 10.84\%{]}), indicating that phonological
+overlapped completely with the \emph{ROPE} (\(\beta\) = 4.76\%, 95\%
+\emph{HDI} = {[}0.83\%, 8.77\%{]}), indicating that phonological
 similarity did not speed-up the acquisition trajectories of the words
@@
 completely with the \emph{ROPE} (\(\beta\) = , 95\% \emph{HDI} =
-{[}-9.32\%, -2.33\%{]}), suggesting that the impact of the degree of
+{[}-7.56\%, -1.23\%{]}), suggesting that the impact of the degree of
 exposure to the language on the effect of phonological similarity did
@@
 between pairs of translation equivalents). Finally, predictions are
-presented separately for different degrees of exposure (10\%): little
-exposure to the language (10\% DoE, traditionally classified as
+presented separately for different degrees of exposure (\emph{DoE}):
+little exposure to the language (10\% DoE, traditionally classified as
 monolingual), balanced exposure to both languages (50\%, traditionally
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\hypertarget{sec-discussion}{%
+\section{Discussion}\label{sec-discussion}}
 
@@
 
-This study explored how word acquisition was affected by cognateness in
-bilingual toddlers. We collected vocabulary data from children learning
-Catalan and/or Spanish in differing degrees of exposure to each, and
-then measured the probability of each word being acquired at different
-ages as a function of (1) participants' exposure to the language it
-belongs to, and (2) its phonological overlap with its translation
-equivalent in the other language. We build a multilevel Bayesian model
+\begin{itemize}
+\tightlist
+\item
+  This study explored how word acquisition is affected by cognateness in
+  bilingual toddlers.
+\item
+  We collected vocabulary data from children learning Catalan and/or
+  Spanish in differing degrees of exposure to each, and then measured
+  the probability of each word being acquired at different ages as a
+  function of (1) participants' exposure to the language it belongs to,
+  and (2) its phonological similarity with its translation equivalent in
+  the other language.
+\item
+  Using item response theory (\emph{IRT}), we built an Bayesian model
   that estimated the effect of both predictors and their interaction,
-adjusting for other variables known to be relevant during the
-acquisition of words: participants' age, lexical frequency, and length
-in phonemes.
-
-\hypertarget{summary-of-the-results}{%
-\subsection{Summary of the results}\label{summary-of-the-results}}
-
-\hypertarget{results-in-contrast-with-the-literature}{%
-\subsection{Results in contrast with the
-literature}\label{results-in-contrast-with-the-literature}}
+  adjusting for other variables known to be relevant for the acquisition
+  of words: participants' age, lexical frequency, and word length in
+  phonemes.
+\end{itemize}
+
+\hypertarget{summary-of-results-and-contrast-with-the-literature}{%
+\subsection{Summary of results and contrast with the
+literature}\label{summary-of-results-and-contrast-with-the-literature}}
+
+\begin{itemize}
+\tightlist
+\item
+  We replicated effects reported in previous studies, such as age being
+  the most determinant predictor of word acquisition, followed by
+  exposure to the language, that frequent words are acquired earlier,
+  and that word length predicts a later age of acquisition.
+\item
+  Regarding our hypotheses, we found strong evidence for a two-way
+  interaction between language exposure and phonological similarity:
+  participants were more likely to acquire word-forms from cognate
+  translation equivalents than from non-cognate translation equivalents,
+  but only when their exposure to the language of such word-form was
+  low.
+\item
+  Our results converge with previous findings indicating that
+  phonological similarity between translation equivalents plays a
+  facilitation role in word acquisition, but suggest that this effect is
+  mostly, or only present in participants' language of lower exposure.
+\end{itemize}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  These findings support the hypothesis that the cognate facilitation
+  effect on word acquisition is more likely to play such role
+  \emph{after} one of the word-forms of the translation equivalent has
+  been acquired in one of the languages, as opposed to taking place
+  before any of the forms has been acquired, by promoting their
+  acquisition though cross-language activation at the sub-lexical,
+  phonological level.
+\item
+  In line with the first hypothesis, words in the language of higher
+  exposure are likely to be acquired earlier than those in the language
+  of lower exposure, and also showed a larger effect facilitation effect
+  of cognateness.
+\item
+  This suggests that cognateness might have played a larger role during
+  the acquisition of words from the language of lower exposure than from
+  the language of higher exposure in bilinguals
+\end{itemize}
+
 \hypertarget{limitations}{%
@@
 
+\begin{itemize}
+\tightlist
+\item
+  These results come from a particular population of bilingual toddlers
+  that learn two very similar languages--Catalan and Spanish--and are
+  therefore exposed to many cognates. Our results might not generalise
+  to other groups of bilinguals learning two less similar languages.
+\item
+  {[}To be discussed{]}
+\end{itemize}
+
 \hypertarget{further-steps}{%
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\begin{itemize}
+\tightlist
+\item
+  {[}To be discussed{]}
+\end{itemize}
+
+\hypertarget{sec-references}{%
+\section{References}\label{sec-references}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bates1994developmental}{}}%
+Bates, E., Marchman, V., Thal, D., Fenson, L., Dale, P., Reznick, J. S.,
+Reilly, J., \& Hartung, J. (1994). Developmental and stylistic variation
+in the composition of early vocabulary. \emph{Journal of Child
+Language}, \emph{21}(1), 85--123.
+
 \leavevmode\vadjust pre{\hypertarget{ref-bergelson2020comprehension}{}}%
@@
 Development}, \emph{11}(4), 369--380.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bilson2015semantic}{}}%
+Bilson, S., Yoshida, H., Tran, C. D., Woods, E. A., \& Hills, T. T.
+(2015). Semantic facilitation in bilingual first language acquisition.
+\emph{Cognition}, \emph{140}, 122--134.
 
@@
 \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}
-
-\leavevmode\vadjust pre{\hypertarget{ref-braginsky2019consistency}{}}%
-Braginsky, M., Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019).
-Consistency and variability in children's word learning across
-languages. \emph{Open Mind}, \emph{3}, 52--67.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-multilex}{}}%
+Garcia-Castro, G. (2022). \emph{Multilex: A r package for multilingual
+lexical assessment using online surveys}.
+\url{https://github.com/gongcastro/multilex}
+
 \leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
@@
 stories}. Cambridge University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gelman1992inference}{}}%
-Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation
-using multiple sequences. \emph{Statistical Science}, \emph{7}(4),
-457--472.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-goodman2008does}{}}%
+Goodman, J. C., Dale, P. S., \& Li, P. (2008). Does frequency count?
+Parental input and the acquisition of vocabulary. \emph{Journal of Child
+Language}, \emph{35}(3), 515--531.
+
 \leavevmode\vadjust pre{\hypertarget{ref-hamilton2000infant}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-hills2009longitudinal}{}}%
+Hills, T. T., Maouene, M., Maouene, J., Sheya, A., \& Smith, L. (2009).
+Longitudinal analysis of early semantic networks: Preferential
+attachment or preferential acquisition? \emph{Psychological Science},
+\emph{20}(6), 729--739.
+
 \leavevmode\vadjust pre{\hypertarget{ref-hoff2012dual}{}}%
@@
 \emph{Cognition}, \emph{106}(1), 501--511.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jones2019children}{}}%
+Jones, S., \& Brandt, S. (2019). Do children really acquire dense
+neighbourhoods? \emph{Journal of Child Language}, \emph{46}(6),
+1260--1273.
 
@@
 \emph{14}(4), 769--785.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcelreath2018statistical}{}}%
-McElreath, R. (2018). \emph{Statistical rethinking: A bayesian course
-with examples in r and stan}. Chapman; Hall/CRC.
 
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\hypertarget{appendix-words}{%
-\subsection{Appendix 1: Words}\label{appendix-words}}
-
-\hypertarget{appendix-words}{%
-\subsection{\texorpdfstring{Appendix 2: \emph{ROPE}
-specification}{Appendix 2: ROPE specification}}\label{appendix-words}}
-
-We decided to draw the \emph{ROPE} interval around zero so that it
-covered values smaller the absolute median of those relevant by similar
-previous studies. After consulting the regression tables of Mitchell et
-al. (2022), Braginsky et al. (2019), and Laing (2022), we selected the
-largest regression coefficient that was interpreted as conclusive
-evidence of the presence of an effect, as indicated by the authors, and
-transformed it to the probability scale: 0.18, in the logit scale.
+\hypertarget{sec-appendix}{%
+\section{Appendix}\label{sec-appendix}}
+
+\hypertarget{sec-appendix-words}{%
+\subsection{Appendix 1: Words}\label{sec-appendix-words}}
+
+\hypertarget{sec-appendix-model}{%
+\subsection{Appendix 2: Model details}\label{sec-appendix-model}}
+
+\hypertarget{sec-appendix-rope}{%
+\subsection{\texorpdfstring{Appendix 3: \emph{ROPE}
+specification}{Appendix 3: ROPE specification}}\label{sec-appendix-rope}}
 

### 2022-11-17 R/roc-curve.R

@@ -1,0 +1,35 @@
+library(brms)
+library(tidybayes)
+library(ggplot2)
+
+fit_0 <- brm(
+    rating ~ treat + period + carry + (1 | subject), 
+    data = inhaler, 
+    family = cumulative(), 
+    prior = set_prior("normal(0,5)"),
+    backend = "cmdstanr"
+)
+
+
+fit_1 <- brm(
+    rating ~ treat + period + (1 | subject), 
+    data = inhaler, 
+    family = cumulative(), 
+    prior = set_prior("normal(0,5)"),
+    backend = "cmdstanr"
+)
+
+
+fit_1$data %>% 
+    # sample_n(10) %>% 
+    add_predicted_draws(fit_1) %>% 
+    mutate(
+        correct = .prediction==rating,
+        rating = as.factor(rating)
+    ) %>% 
+    count(rating, treat, .prediction) %>% 
+    ggplot(aes(.prediction, n, fill = .prediction)) +
+    facet_grid(rating~treat) +
+    geom_col() +
+    theme_custom() +
+    scale_fill_d3()

### 2022-11-17 _targets.R

@@ -16,6 +16,8 @@ tar_option_set(
         "bayesplot",
         "bayestestR",
         "brms", 
+        "bvqdev",
+        "childesr",
         "conflicted",
         "dplyr", 
         "forcats",
@@ -28,7 +30,6 @@ tar_option_set(
         "knitr",
         "lubridate",
         "mice", 
-        "multilex",
         "patchwork", 
         "papaja", 
         "purrr",
@@ -59,7 +60,7 @@ options(
 
 list(
     
-    # resolve namespace conflicts ----------------------------------------------
+    # resolve namespace conflicts ##############################################
     tar_target(
         resolve_conficts,
         {
@@ -76,71 +77,32 @@ list(
         }
     ),
     
-    # import data --------------------------------------------------------------
-    tar_target(
-        credentials,
-        get_credentials() # see R/utils.R
-    ),
+    # import data ##############################################################
+    tar_target(bvq_data, get_bvq(update = TRUE)), # see R/00_importing.R
     
-    tar_target(
-        multilex_data,
-        # see R/00_importing.R
-        get_multilex(
-            update = TRUE, 
-            type = c("understands", "produces")
-        )
-    ),
+    # get CHILDES frequencies
+    tar_target(childes, get_childes_frequencies(age_range = c(12, 32))), # see R/utils.R
     
     # items
-    tar_target(
-        items,
-        # see R/01_preprocessing.R
-        get_items(
-            multilex_data = multilex_data,
-        )
-    ),
+    tar_target(items, get_items(bvq_data = bvq_data, childes = childes)), # see R/01_preprocessing.R
     
     # participants
     tar_target(
         participants,
         # see R/01_preprocessing.R
         get_participants(
-            multilex_data,
-            age = c(10, 36), 
+            bvq_data,
+            age = c(12, 32), 
             lp = c("Monolingual", "Bilingual"), 
             other_threshold = 0.1
         )
     ),
-    
-    # vocabulary
-    tar_target(
-        vocabulary,
-        # see R/01_preprocessing.R
-        get_vocabulary(
-            multilex_data = multilex_data,
-            type = c("understands", "produces")
-        )
-    ),
-    
+
     # responses
-    tar_target(
-        responses,
-        # see R/01_preprocessing.R
-        get_responses(
-            multilex_data = multilex_data
-        )
-    ),
[…]

### 2022-11-17 index.qmd
Update lab notes

--- 2022-09-16 index.qmd
@@
 #| include: false
+tar_load_globals() # load packages and options
+tar_load_everything() # load targets as R objects
+theme_set(theme_custom()) # set custom ggplot theme
+
+# resolve namespace conflicts
+conflict_prefer("filter", "dplyr")
+conflict_prefer("select", "dplyr")
 options(
@@
 
-```{r}
-#| label: prepare
-#| echo: true
-#| message: false
-#| warning: false
-# set params and load target contents
-
-library(bayesplot)
-library(brms) 
-library(bayestestR)
-library(conflicted)
-library(dplyr) 
-library(forcats)
-library(ggplot2) 
-library(ggsci)
-library(gt) 
-library(here) 
-library(janitor)
-library(knitr)
-library(lubridate)
-library(multilex)
-library(patchwork) 
-library(papaja) 
-library(purrr)
-library(scales) 
-library(stringdist)
-library(stringr)
-library(targets)
-library(tibble)
-library(tidybayes) 
-library(tidyr)
-
-# load R objects
-tar_load_everything()
-
-# set custom ggplot2 theme
-theme_set(theme_custom())
-
-# resolve namespace conflicts
-conflict_prefer("filter", "dplyr")
-conflict_prefer("select", "dplyr")
-```
-
-# Questionnaires
+# Methods
+
+Data and materials are available at OSF (https://osf.io/hy984/), and code is available at GitHub (https://github.com/gongcastro/trajectories).
+
+## Questionnaires
 
@@
 #| warning: false
-pool <- multilex::pool %>% 
-    drop_na(cognate)
+pool <- mutate(bvqdev::pool, language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish"))
 n_categories <- distinct(pool, category) %>% nrow()
 n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
-n_item_cognate <- count(pool, language, cognate) %>% 
+n_item_language <- count(pool, language) %>% 
     pull(n) %>%
-    set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
+    set_names(c("catalan", "spanish"))
 ```
@@
 Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
+The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]` items.
 Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* \[dog\] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
@@
 Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
-items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))
+items <- left_join(items, select(pool, te, item, language, class, category))
 
@@
     summarise_at(
-        vars(freq, n_phon, lv), 
-        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
+        vars(freq_zipf, n_phon, lv), 
+        list(
+            n = ~sum(!is.na(.)),
+            mean = ~mean(., na.rm = TRUE),
+            sd = ~sd(., na.rm = TRUE),
+            min = ~min(., na.rm = TRUE),
+            max = ~max(., na.rm = TRUE)
+        )
     ) %>% 
     select(-c(n_phon_n, lv_n)) %>% 
-    rename(n = freq_n) %>% 
+    rename(n = freq_zipf_n) %>% 
     gt(groupname_col = "language") %>% 
@@
     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
-    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
-    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
-    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
-    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
-    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
+    tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>% 
+    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
+    cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>% 
+    cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>% 
+    cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>% 
+    cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>% 
+    cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>% 
+    cols_merge_range(col_begin = lv_min, col_end = lv_max) %>% 
     cols_label(
@@
         n = md("*N*"),
-        freq_mean = md("Mean"),
-        freq_sd = md("*SD*"),
-        freq_min = md("Range"),
-        n_phon_mean = md("Mean"),
-        n_phon_sd = md("*SD*"),
+        freq_zipf_mean = md("Mean (SD)"),
+        freq_zipf_min = md("Range"),
+        n_phon_mean = md("Mean (SD)"),
         n_phon_min = md("Range"),
-        lv_mean = md("Mean"),
-        lv_sd = md("*SD*"),
+        lv_mean = md("Mean (SD)"),
         lv_min = md("Range")
@@
         cells_column_spanners(everything())
-    )
-```
-
-## Lexical frequency
+    ) %>% 
+    summary_rows(
+        fns = list("Grand mean" = ~mean(.)),
+        columns = ends_with("_mean")
+    ) %>% 
+    summary_rows(
+        fns = list("N" = ~sum(.)),
+        columns = n,
+        formatter = fmt_integer
+    )
+```
+
+### Lexical frequency
 
@@
     aes(
-        x = freq,
+        x = language,
+        y = freq_zipf,
         fill = language,
-        colour = language,
-        shape = language
-    ) +
-    stat_slab(
-        slab_colour = "white",
-        slab_size = 1,
-        alpha = 0.75,
-    ) +
+        colour = language
+    ) +
+    stat_slab(slab_colour = "white", slab_size = 1, alpha = 0.5, scale = 0.5) +
+    stat_dots(side = "left", scale = 0.5, position = position_nudge(x = -0.05)) + 
     stat_pointinterval(
-        aes(y = -0.1),
         position = position_dodge(width = 0.1),
-        point_size = 3
-    ) +
-    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
+        point_size = 2.5,
+        colour = "black"
+    ) +
     labs(
@@
     ) +
-    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
-    theme(
-        legend.position = "top",
+    theme(
+        legend.position = "none",
         legend.title = element_blank()
@@
     count(language, n_phon) %>% 
-    mutate(prop = n/sum(.$n)) %>% 
     ggplot() +
@@
         x = n_phon,
-        y = prop,
+        y = n,
         fill = language,
@@
     ) +
-    geom_col(
-        position = position_dodge(width = 0.95)
-    ) +
-    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
+    facet_wrap(~language, ncol = 1) +
+    geom_col() +
     labs(
         x = "# Phonemes",
-        y = "Probability density",
+        y = "Number of words",
         fill = "Language",
@@
     ) +
-    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
-    scale_x_continuous(breaks = seq(0, 20, 1)) +
-    theme(
-        legend.position = "top",
+    scale_x_continuous(breaks = 1:15) +
+    theme(
+        legend.position = "none",
         axis.ticks.x = element_line(),
@@
     distinct(te, .keep_all = TRUE) %>% 
-    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
-    count(lv_cut) %>% 
-    mutate(prop = n/sum(.$n)) %>% 
-    ggplot() +
-    aes(x = lv_cut, y = prop) +
-    geom_col(fill = ggsci::pal_d3()(4)[1]) +
-    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
-    labs(
-        x = "Levenshtein",
-        y = "Probability density",
-    ) +
+    ggplot(aes(x = lv)) +
+    geom_dots(fill = ggsci::pal_d3()(4)[1], colour = ggsci::pal_d3()(4)[1]) +
+    labs(x = "Levenshtein distance", y = "Probability density") +
+    scale_x_continuous(labels = percent) +
     theme(
@@
 items %>% 
-    drop_na(freq) %>% 
-    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
+    drop_na(freq_zipf) %>% 
+    count(n_phon, freq_zipf, language) %>% 
+    ggplot(aes(n_phon, freq_zipf, colour = language, fill = language)) +
     facet_wrap(~language) +
-    geom_point(size = 1, alpha = 0.25) +
+    geom_point(size = 2, alpha = 0.25) +
     geom_smooth(method = "lm", formula = "y ~ x") +
@@
     ) +
+    coord_equal(ratio = 10) +
     scale_x_continuous(breaks = seq(0, 16, 1)) +
@@
     left_join(select(pool, item, language)) %>% 
-    drop_na(freq) %>% 
-    ggplot(aes(freq, lv, colour = language, fill = language)) +
+    drop_na(freq_zipf) %>% 
+    count(freq_zipf, lv, language) %>% 
+    ggplot(aes(freq_zipf, lv, colour = language, fill = language, size = n)) +
     facet_wrap(~language) +
-    geom_point(size = 1, alpha = 0.5) +
+    geom_point(size = 2, alpha = 0.25) +
     geom_smooth(method = "lm", formula = "y ~ x") +
@@
     left_join(select(pool, item, language)) %>% 
-    drop_na(freq) %>% 
-    ggplot(aes(n_phon, lv, colour = language, fill = language)) +
+    drop_na(freq_zipf) %>% 
+    count(n_phon, lv, language) %>% 
+    ggplot(aes(n_phon, lv, colour = language, fill = language, size = n)) +
     facet_wrap(~language) +
-    geom_point(size = 1, alpha = 0.25) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
+    geom_point(shape = 1, stroke = 1) +
+    geom_smooth(method = "lm", formula = "y ~ x", size = 1) +
     labs(
@@
 #| fig-cap: "Data at which responses were collected"
-participants <- left_join(
-    participants, 
-    select(multilex_data$logs, id, time, edu_parent, dominant_language = dominance)
-)
-
 participants %>% 
-    group_by(time_stamp, dominant_language) %>% 
+    group_by(time_stamp) %>% 
     summarise(n = n(), .groups = "drop") %>% 
-    group_by(dominant_language) %>% 
     mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
+    ggplot(aes(time_stamp, n)) +
     geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
@@
              angle = 90, vjust = 1.5, hjust = 1) +
-    geom_line(size = 1) +
+    geom_line(size = 2, colour = pal_d3()(1)[1]) +
     labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
     scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
+    scale_x_date(
+        date_breaks = "3 month", 
+        date_labels = "%B %Y"
+    ) +
     guides(colour = guide_legend(ncol = 2)) +
@@
         legend.title = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle = 20, hjust = 1),
         axis.title.x = element_blank(),
@@
     theme(
+        legend.position = "top",
         axis.text.y = element_blank(),
         axis.ticks.y = element_blank(),
-        panel.grid.major.y = element_blank(),
-        axis.title = element_blank()
-    )
-```
-
-## SES/parental education
+        panel.grid.major.y = element_blank()
+    )
+```
+
+### SES/parental education
 
@@
 participants %>%
+    left_join(bvq_data$logs) %>% 
     mutate(
         edu_parent = fct_explicit_na(as.factor(edu_parent)),
-        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
+        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE)
     ) %>% 
@@
             doe_catalan = levels(.$doe_catalan),
-            edu_parent = unique(multilex_data$logs$edu_parent)
+            edu_parent = unique(bvq_data$logs$edu_parent)
         )
@@
     gt() %>% 
-    cols_label(
-        doe_catalan = "DoE Catalan"
-    ) %>% 
     summary_rows(
-        columns = 2:7, 
+        columns = Complementary:Primary, 
         fns = list(Total = ~sum(., na.rm = TRUE)),
-        formatter = fmt_number,
-        decimals = 0
-    ) %>% 
+        formatter = fmt_integer
+    ) %>% 
+    cols_label(doe_catalan = "DoE Catalan") %>%
     sub_missing(columns = 2:8, missing_text = "--") %>% 
-    tab_spanner("Educational attainment", 3:8) %>% 
+    tab_spanner("Educational attainment", Complementary:Primary) %>% 
     tab_style(
         style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
-        locations = cells_body(columns = 9, rows = 1:10)
+        locations = cells_body(columns = 9, rows = 1:5)
     ) %>% 
@@
 
-```{r}
-#| label: fig-participants-edu
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-participants %>%
-    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
-    mutate(
-        doe_2 = pmin(doe_catalan, doe_spanish),
-        doe_2 = cut(doe_2, seq(0, 0.5, 0.1), include.lowest = TRUE)
-    ) %>% 
-    count(doe_2, edu_parent) %>% 
-    ggplot() +
-    aes(
-        x = doe_2,
-        y = n,
-        fill = edu_parent
-    ) +
-    geom_col(position = position_fill()) +
-    scale_fill_manual(values = sort(pal_d3()(6), decreasing = TRUE)) +
-    labs(
-        x = "Exposure to second language (L2 DoE)",
-        y = "Proportion of participants",
-        fill = "Highest parental educational attainment"
-    ) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "none",
-        panel.grid = element_blank()
-    )
-```
-
-# Data analysis
-
-## Model comparison
-
-We initially fitted a base model (`model_fit_0`) that only included the predictors $Age$ (`age_std`), $Frequency$ (`freq_std`), $Phonemes$ (`n_phon_std`), and $DoE$ (`doe_std`) as nuisance parameters, along with random intercepts and slopes of all predictors by $Participant$ (`id`) and $Item$ (`te`), and all correlation parameters.
-We then expanded this model by adding the main effect of $Levenshtein$ (`lv_std`) (`model_fit_1`) and its random slope by $Participant$.
-Then, we added the $DoE \times Levenshtein$ (`doe_std:lv_std`) two-way interaction (`model_fit_2`) and its random slope by $Participant$.
-Finally, we added the $Age \times Levenshtein$ (`age_std:doe_std:lv_std`) three-way interaction with its random slope by $Participant$, along with a random slope for the $Age \times DoE$ (`age_std:doe_std`) two-way interaction.
-
-## Model equation
+## Exposure-weighted lexical frequency
+
+We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
+This measure is the product of the lexical frequency of the word (Zipf score) and the child's degree of exposure to the corresponding language (a proportion):
+
+$$
+\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
+$$
+
+For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)`:
+
+$$
+\begin{aligned}
+\textbf{Exposure}_{\textbf{cavall, Catalan}} &= \textbf{Frequency}_{cavall} \times \textbf{DoE}_{Catalan} = 6.14 \times 0.8 = 4.91 \\
+\textbf{Exposure}_{\textbf{caballo, Spanish}} &= \textbf{Frequency}_{caballo} \times \textbf{DoE}_{Spanish} = 6.14 \times 0.2 = 1.23
+\end{aligned}
+$$
+
+```{r}
+#| label: fig-exposure-distribution
+df %>% 
+    mutate(language = ifelse(grepl("cat_", item), "Catalan", "Spanish")) %>% 
+    ggplot(aes(exposure, fill = language)) +
+    facet_wrap(~language) +
+    geom_histogram(bins = 10, colour = "white", size = 1, show.legend = FALSE) +
+    labs(x = "Exposure (Frequency \u00d7 DoE)", y = "Number of observations") +
+
+    df %>% 
+    mutate(language = ifelse(grepl("cat_", item), "Catalan", "Spanish")) %>% 
+    ggplot(aes(x = freq, y = exposure, colour = language)) +
+    facet_wrap(~language) + 
+    geom_point(size = 0.1, alpha = 0.5, show.legend = FALSE) +
+    labs(x = "Frequency (Zipf score)", y = "Exposure") +
+    plot_layout(ncol = 1)
+```
+
+## Data analysis
+
+### Model comparison
+
+We initially fitted a base model (`model_fit_0`) that only included the predictors $Age$ (`age_std`), $Phonemes$ (`n_phon_std`), along with random intercepts and slopes of all predictors by $Participant$ (`id`) and $Item$ (`te`), and all correlation parameters.
+We then expanded this model by adding the main effect of $Exposure$ (`exposure`) (`model_fit_1`), and then $Levenshtein$ (`lv_std`) (`model_fit_2`) and their corresponding random slopes by $Participant$.
+Then, we added the $Exposure \times Levenshtein$ (`exposure_std:lv_std`) two-way interaction (`model_fit_3`) and its random slope by $Participant$.
+Finally, we added the $Age \times Exposure \times Levenshtein$ (`age_std:exposure_std:lv_std`) three-way interaction with its random slope by $Participant$, along with a random slope for the $Age \times Exposure$ (`age_std:exposure_std`) two-way interaction.
+
+### Model equation
 
@@
 $$\begin{aligned}
-\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2 \mathrm{Freq}_{j} + \beta_3\mathrm{Phon}_{j} + \beta_4\mathrm{Exposure}_{ij} + \beta_5 \mathrm{Levenshtein}_{j} + & [\mbox{Main effects}]\\
-&\beta_6\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_7\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_8\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Two-way interactions}]\\
-&\beta_9{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Three-way interaction}]\\
-
-&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{TE})} \mathrm{Freq}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{5j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) + & [\mbox{Item-level effects}]\\
-
-&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +  \theta_{2j}^{(\mathrm{ID})} \mathrm{Freq}_{j} +   \theta_{3j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{5j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{6j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) & [\mbox{Participant-level effects}]
+\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2\mathrm{Phon}_{j} + \beta_3\mathrm{Exposure}_{ij} + \beta_4 \mathrm{Levenshtein}_{j} + & [\mbox{Main effects}]\\
+&\beta_5\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_6\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_7\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Two-way interactions}]\\
+&\beta_8{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Three-way interaction}]\\
+
+&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} + \theta_{2j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{4j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) + & [\mbox{Item-level effects}]\\
+
+&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +   \theta_{2j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{5j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) & [\mbox{Participant-level effects}]
 \end{aligned}$$
@@
 
-In addition, we specified a $LKJ$ prior--$\sim LKJ(2)$ for the correlations between group-level coefficients. We implemented the model prior following {brms} syntax as:
-
-```r
-c(
-prior(normal(-0.25, 0.1), class = "Intercept"),
-prior(normal(1, 0.1), class = "sd", group = "te"),
-prior(normal(1, 0.1), class = "sd", group = "id"),
-prior(lkj(2), class = "cor"),
-prior(normal(1, 0.1), class = "b", coef = "age_std"),
-prior(normal(0, 0.1), class = "b", coef = "freq_std"),
-prior(normal(0, 0.1), class = "b", coef = "n_phon_std"),
-prior(normal(0, 0.1), class = "b", coef = "doe_std"),
-prior(normal(0, 0.1), class = "b", coef = "lv_std"),
-prior(normal(0, 0.1), class = "b", coef = "doe_std:lv_std"),
-prior(normal(0, 0.1), class = "b", coef = "age_std:doe_std"),
-prior(normal(0, 0.1), class = "b", coef = "age_std:lv_std"),
-prior(normal(0, 0.1), class = "b", coef = "age_std:doe_std:lv_std")
-)
-```
-
-### Prior predictions
-
-```{r}
-#| label: fig-prior-predictions
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected prior predictions."
+In addition, we specified a $LKJ$ prior--$\sim LKJ(2)$ for the correlations between group-level coefficients.
+
+#### Prior predictions
+
+::: {.panel-tabset}
+
+##### Prior-predicted mean
+
+```{r}
+#| label: fig-prior-mean
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Expected prior predictions"
 nd <- expand.grid(
     n_phon_std = 0,
-    freq_std = 0,
     age_std = (10:36-mean(df$age))/(sd(df$age)),
-    doe_std = 0,
+    exposure_std = 0,
     lv_std = 0
@@
 
-m <- add_epred_draws(nd, model_fit_3_prior, ndraws = NULL, re_formula = NA) %>% 
+m <- add_epred_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA) %>% 
     filter(.category != "No") %>% 
@@
 m %>% 
-    ggplot() +
-    aes(
-        x = age_std, 
-        y = .epred, 
-    ) +
+    ggplot(aes(x = age_std, y = .epred)) +
     facet_wrap(~.category) +
-    # geom_line(
-    #     aes(group = interaction(.draw)),
-    #     alpha = 0.05,
-    #     size = 1
-    # ) +
     stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
@@
     ) +
-    scale_fill_brewer(
-        palette = "Reds",
-        labels = c("95%", "89%", "67%", "50%")
-    ) +
+    scale_fill_brewer(palette = "Reds", labels = c("95%", "89%", "67%", "50%")) +
     labs(
         x = "Age (months)", 
-        y = "Probability of acquisition\n(Linear prediction)",
+        y = "Probability of acquisition",
         colour = "Credible interval", 
@@
 
-## Model settings
+##### Prior-predictions
+
+```{r}
+#| label: fig-prior-predictions
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Expected prior predictions"
+nd <- expand.grid(
+    n_phon_std = 0,
+    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    exposure_std = 0,
+    lv_std = 0
+)
+
+m <- add_predicted_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA)
+
+m %>% 
+    count(age_std, .prediction) %>% 
+    pivot_wider(names_from = .prediction, values_from = n) %>% 
+    mutate(Understands = Understands + `Understands and Says`) %>% 
+    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
+    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
+    geom_col(position = position_fill()) +
+    scale_x_continuous(
+        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 36, 4),
+        sec.axis = dup_axis(
+            name = "Age (years; months)",
+            labels = months_to_years(
+                rescale_variable(
+                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+                    mean(df$age),
+                    sd(df$age)
+                )
+            )
+        )
+    ) +
+    labs(
+        x = "Age (months)", 
+        y = "Probability of acquisition",
+        colour = "Category", 
+        fill = "Category"
+    ) +
+    scale_y_continuous(labels = percent) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        axis.ticks.x = element_line(colour = "black"),
+        panel.grid = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
+        strip.background = element_rect(fill = "grey", colour = NA),
+        plot.background = element_rect(fill = "white", colour = NA)
+    )
+```
+
+### Model settings
 
@@
 -   brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
--   Chains: `r dim(model_fit_7$fit)[2]`
--   Cores: `r dim(model_fit_7$fit)[2]`
--   Iterations: `r format(dim(model_fit_7$fit)[1], big.mark = ",")` (`r format(dim(model_fit_7$fit)[1]/2, big.mark = ",")`)
+-   Chains: `r dim(model_fit_4$fit)[2]`
+-   Cores: `r dim(model_fit_4$fit)[2]`
+-   Iterations: `r format(dim(model_fit_4$fit)[1], big.mark = ",")` (`r format(dim(model_fit_4$fit)[1]/2, big.mark = ",")`)
 
@@
 ```
+
+#### Model 4
+
+```{r}
+print(model_fit_4$formula)
+```
 :::
 
-## Stan code
-
-Stan code generated by `brms::stancode`.
-
-::: panel-tabset
-
-### Model 0
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 1
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-vector[N] Z_1_6;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-vector[N_1] r_1_6;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-r_1_6 = r_1[, 6];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(b[5] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 6 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_b_5 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 2
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-vector[N] Z_1_6;
-vector[N] Z_1_7;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-vector[N_1] r_1_6;
-vector[N_1] r_1_7;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-r_1_6 = r_1[, 6];
-r_1_7 = r_1[, 7];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(b[5] | 0, 0.1);
-target += normal_lpdf(b[6] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 7 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_b_5 = normal_rng(0,0.1);
-real prior_b_6 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-
-### Model 3
-
-```
-// generated with brms 2.16.1
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-/* cratio-logit log-PDF for a single response
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: ordinal thresholds
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-int nthres = num_elements(thres);
-vector[nthres + 1] p;
-vector[nthres] q;
-int k = 1;
-while (k <= min(y, nthres)) {
-q[k] = log_inv_logit(disc * (mu - thres[k]));
-p[k] = log1m_exp(q[k]);
-for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-k += 1;
-}
-if (y == nthres + 1) {
-p[nthres + 1] = sum(q);
-}
-return p[y];
-}
-/* cratio-logit log-PDF for a single response and merged thresholds
-* Args:
-*   y: response category
-*   mu: latent mean parameter
-*   disc: discrimination parameter
-*   thres: vector of merged ordinal thresholds
-*   j: start and end index for the applid threshold within 'thres'
-* Returns:
-*   a scalar to be added to the log posterior
-*/
-real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=2> nthres;  // number of thresholds
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-vector[N] Z_1_5;
-vector[N] Z_1_6;
-vector[N] Z_1_7;
-int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
-vector[N] Z_2_2;
-vector[N] Z_2_3;
-vector[N] Z_2_4;
-vector[N] Z_2_5;
-int<lower=1> NC_2;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K;
-matrix[N, Kc] Xc;  // centered version of X
-vector[Kc] means_X;  // column means of X before centering
-for (i in 1:K) {
-means_X[i] = mean(X[, i]);
-Xc[, i] = X[, i] - means_X[i];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-vector[nthres] Intercept;  // temporary thresholds for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-matrix[M_2, N_2] z_2;  // standardized group-level effects
-cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-real<lower=0> disc = 1;  // discrimination parameters
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-vector[N_1] r_1_5;
-vector[N_1] r_1_6;
-vector[N_1] r_1_7;
-matrix[N_2, M_2] r_2;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_2] r_2_1;
-vector[N_2] r_2_2;
-vector[N_2] r_2_3;
-vector[N_2] r_2_4;
-vector[N_2] r_2_5;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-r_1_5 = r_1[, 5];
-r_1_6 = r_1[, 6];
-r_1_7 = r_1[, 7];
-// compute actual group-level effects
-r_2 = scale_r_cor(z_2, sd_2, L_2);
-r_2_1 = r_2[, 1];
-r_2_2 = r_2[, 2];
-r_2_3 = r_2[, 3];
-r_2_4 = r_2[, 4];
-r_2_5 = r_2[, 5];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Xc * b;
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-}
-for (n in 1:N) {
-target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-}
-}
-// priors including constants
-target += normal_lpdf(b[1] | 1, 0.1);
-target += normal_lpdf(b[2] | 0, 0.1);
-target += normal_lpdf(b[3] | 0, 0.1);
-target += normal_lpdf(b[4] | 0, 0.1);
-target += normal_lpdf(b[5] | 0, 0.1);
-target += normal_lpdf(b[6] | 0, 0.1);
-target += normal_lpdf(b[7] | 0, 0.1);
-target += normal_lpdf(b[8] | 0, 0.1);
-target += normal_lpdf(b[9] | 0, 0.1);
-target += normal_lpdf(Intercept | -0.25, 0.1);
-target += normal_lpdf(sd_1 | 1, 0.1)
-- 7 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 2);
-target += normal_lpdf(sd_2 | 1, 0.1)
-- 5 * normal_lccdf(0 | 1, 0.1);
-target += std_normal_lpdf(to_vector(z_2));
-target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-// compute actual thresholds
-vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// compute group-level correlations
-corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-vector<lower=-1,upper=1>[NC_2] cor_2;
-// additionally sample draws from priors
-real prior_b_1 = normal_rng(1,0.1);
-real prior_b_2 = normal_rng(0,0.1);
-real prior_b_3 = normal_rng(0,0.1);
-real prior_b_4 = normal_rng(0,0.1);
-real prior_b_5 = normal_rng(0,0.1);
-real prior_b_6 = normal_rng(0,0.1);
-real prior_b_7 = normal_rng(0,0.1);
-real prior_b_8 = normal_rng(0,0.1);
-real prior_b_9 = normal_rng(0,0.1);
-real prior_Intercept = normal_rng(-0.25,0.1);
-real prior_sd_1 = normal_rng(1,0.1);
-real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-real prior_sd_2 = normal_rng(1,0.1);
-real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-// extract upper diagonal of correlation matrix
-for (k in 1:M_2) {
-for (j in 1:(k - 1)) {
-cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-}
-}
-// use rejection sampling for truncated priors
-while (prior_sd_1 < 0) {
-prior_sd_1 = normal_rng(1,0.1);
-}
-while (prior_sd_2 < 0) {
-prior_sd_2 = normal_rng(1,0.1);
-}
-}
-```
-:::
-
 # Results
-
-## Raw data
-
-```{r}
-#| label: fig-results-raw
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Observed proportion of comprehension and production responses across ages.
-Colours indicate different values of phonological similarity (Levenshtein similarity, binned).
-Results for different degrees of exposure (binned) are plotted separately."
-df %>%
-    mutate(
-        lv = cut_interval(lv*100, 5) %>% 
-            gsub("([0-9.]+)", "\\1%", .) %>% 
-            factor(levels = c("[0%,20%]", "(20%,40%]", "(40%,60%]", "(60%,80%]", "(80%,100%]")),
-        doe = cut_interval(doe*100, 3) %>% 
-            gsub("([0-9.]+)", "\\1%", .) %>% 
-            factor(levels = c("[0%,33.3%]", "(33.3%,66.7%]", "(66.7%,100%]")) %>% 
-            paste0("Degree of Exposure: ", .) %>% 
-            factor(
-                levels = paste0(
-                    "Degree of Exposure: ", 
-                    c("[0%,33.3%]", "(33.3%,66.7%]", "(66.7%,100%]")
-                )
-            ),
-        age = floor(age),
-        Understands = response %in% c("Understands", "Understands and Says"),
-        `Understands and Says` = response %in% c("Understands and Says")
-    ) %>% 
-    group_by(te, lv, age, doe) %>%
-    summarise(
-        n = n(),
-        Understands = mean(Understands),
-        `Understands and Says` = mean(`Understands and Says`),
-        .groups = "drop"
-    ) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`), 
-        names_to = ".category", 
-        values_to = "prop"
-    ) %>% 
-    group_by(lv, age, doe, .category) %>%
-    summarise(
-        n = sum(n),
-        prop = mean(prop),
-        .groups = "drop"
-    ) %>% 
-    mutate(
-        prop_se = sd(prop)/sqrt(n),
-        prop_lower = prop-prop_se,
-        prop_upper = prop+prop_se,
-        prop_lower = ifelse(prop_lower < 0, 0, prop_lower),
-        prop_upper = ifelse(prop_upper >1, 1, prop_upper)
-    ) %>% 
-    ggplot() +
-    aes(age, prop, colour = lv, fill = lv, shape = lv) +
-    facet_grid(.category~doe) +
-    geom_hline(yintercept = 0.5, colour = "grey", size = 0.5) +
-    geom_line(size = 1, alpha = 0.75) +
-    # geom_point(size = 2) +
-    labs(
-        x = "Age (months)",
-        y = "Proportion of responses", 
-        colour = "Phonological similarity (Levenshtein)",
-        fill = "Phonological similarity (Levenshtein)",
-        shape = "Phonological similarity (Levenshtein)",
-        linetype = "Phonological similarity (Levenshtein)"
-    ) +
-    scale_color_brewer(palette = "Reds") +
-    # scale_colour_manual(values = wesanderson::wes_palette("Zissou1", n = 5, type = "continuous")[1:5]) +
-    scale_x_continuous(breaks = seq(0, 40, 4)) +
-    scale_y_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "top",
-        axis.ticks.y = element_blank(),
-        panel.grid.minor.x = element_blank(),
-        panel.grid.minor.y = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        plot.background = element_rect(fill = "white", colour = NA),
-        strip.background = element_rect(fill = "grey", colour = "grey")
-    )
-```
 
@@
 LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_3$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model as:
-
-```{r}
-#| label: results-loos
-#| echo: true
-#| message: false
-#| warning: false
-#| eval: false
-#  loo(log_lik(model_fit_3, ndraws = 750, re_formula = NA))
-```
+Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model as:
 
@@
 ```{r}
-#| label: fig-results-rope
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-height: 4
-#| fig-width: 7
-rope_coefs %>% 
-    ggplot(aes(estimate, y = study)) +
-    geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
-    geom_vline(
-        xintercept = median(rope_coefs$estimate),
-        size = 1,
-        colour = pal_d3()(4)[4]
-    ) +
-    stat_dots(
-        dotsize = 0.5, 
-        binwidth = 0.03,
-        fill = "grey50",
-        colour = "grey50"
-    ) +
-    stat_summary(
-        fun = median,
-        geom = "point",
-        position = position_nudge(y = -0.125),
-        size = 6, 
-        shape = 17,
-        colour = pal_d3()(4)[4]
-    ) +
-    annotate(
-        geom = "text",
-        label = paste0(
-            "Grand median of cross-study
-             smallest effect size of interest\n",
-            "Median = ", round(median(rope_coefs$estimate), 3)
-        ),
-        x = median(rope_coefs$estimate)-0.2,
-        y = 4,
-        hjust = 1,
-        vjust = 1.2,
-        size = 4.5,
-        colour = pal_d3()(4)[4]
-    ) +
-    annotate(
-        geom = "curve",
-        x = median(rope_coefs$estimate)-0.2,
-        xend = median(rope_coefs$estimate)-0.01,
-        size = 1,
-        y = 3.6,
-        yend = 3.5,
-        colour = pal_d3()(4)[4],
-        arrow = arrow(length = unit(0.5, "cm"))
-    ) +
-    labs(
-        x = "Standard regression coefficient (logit scale)\n(Excluding intercepts and predictor age)",
-        y = "Study"
-    ) +
-    guides(fill = guide_legend(ncol = 1)) +
-    scale_x_continuous(breaks = seq(-0.5, 0.5, 0.1)) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        legend.justification = 0,
-        axis.title.y = element_blank(),
-        panel.grid.major.y = element_blank()
-    )
-```
-
-```{r}
 #| label: results-posterior-fixed
@@
 #| warning: false
-post_draws <- gather_draws(model_fit_3, `b_.*`, regex = TRUE)
+post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)
 
@@
     "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
-    "b_freq_std" = paste0("Frequency (+1 SD, ",  round(sd(df$freq), 2), " Zipf)"),
     "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
-    "b_doe_std" = paste0("DoE (+1 SD, ",  percent(sd(df$doe)), ")"),
+    "b_exposure_std" = paste0("Exposure (+1 SD, ",  percent(sd(df$exposure)), ")"),
     "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
-    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein",
-    "b_age_std:doe_std" = "Age \u00d7 DoE",
+    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
+    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
     "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein"
+    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
 )
@@
 
-<br>
-
 We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
@@
 
-**Lexical frequency** had a small, positive, but inconclusive on the probability of acquisition substantially, ($\beta$ = `r percent(post_draws_list[["b_freq_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_freq_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_freq_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_freq_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples overlapped with the *ROPE.*
-
 The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
@@
 
-The **degree of exposure to the language** ($DoE$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std"]]$ci_high, accuracy = 0.01)`]).
+**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]).
 All of the posterior samples of his regression excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every 10% increase in exposure to the language the word belongs to, the participant was `r percent(post_draws_list[["b_doe_std"]]$median/(10*sd(df$doe)), accuracy = 0.01)` more likely to acquire it.
+The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/(sd(df$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.
 
@@
 
-However, the 95% *HDI* of the regression coefficient of the $Doe \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_doe_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the language it belonged to.
+However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word.
 Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
@@
 
-The 95% *HDI* of the regression coefficient of the $Age \times DoE$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:doe_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of the degree of exposure on the speed at which participants acquired word across ages was inconclusive.
-Give our data, it is difficult to tell whether this *DoE* increased or decreased the slope of the words' acquisition trajectories.
+The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+This shows that the effect of exposure on the speed at which participants acquired word across ages was inconclusive.
+Give our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.
 
@@
 
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_Std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of the degree of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
+Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
 
@@
     ) +
-    annotate(
-        geom = "text",
-        label = "ROPE",
-        x = 0.3,
-        y = 10,
-        vjust = 1,
-        size = 5,
-        colour = pal_d3()(2)[2]
-    ) +
-    annotate(
-        geom = "curve",
-        arrow = arrow(length = unit(0.2, "cm")),
-        x = 0.3,
-        y = 9.5,
-        xend = 0.1,
-        yend = 9,
-        size = 1,
-        curvature = -0.25,
-        colour = pal_d3()(2)[2]
-    ) +
+    # annotate(
+    #     geom = "text",
+    #     label = "ROPE",
+    #     x = 0.3,
+    #     y = 10,
+    #     vjust = 1,
+    #     size = 5,
+    #     colour = pal_d3()(2)[2]
+    # ) +
+    # annotate(
+    #     geom = "curve",
+#     arrow = arrow(length = unit(0.2, "cm")),
+#     x = 0.3,
+#     y = 9.5,
+#     xend = 0.1,
+#     yend = 9,
+#     size = 1,
+#     curvature = -0.25,
+#     colour = pal_d3()(2)[2]
+# ) +
 labs(
@@
     scale_fill_manual(
-        values = c(pal_d3()(1)[1], "grey"),
+        values = c("dodgerblue", "grey"),
         labels = c("No", "Yes")
@@
     scale_colour_manual(
-        values = c(pal_d3()(1)[1], "grey"),
+        values = c("dodgerblue", "grey"),
         labels = c("No", "Yes")
@@
 #| fig-cap: "Expected posterior predictions.
-Posterior poreditions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
+Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
 These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model.
@@
     age_std = (10:36-mean(df$age))/(sd(df$age)),
-    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe)),
-    lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
-)
-
-m <- add_epred_draws(nd, model_fit_3, ndraws = NULL, re_formula = NA) %>% 
+    exposure_std = c(min(df$exposure_std, na.rm = TRUE), 0, max(df$exposure_std, na.rm = TRUE)),
+    lv_std = (c(0, 1)-mean(df$lv))/(sd(df$lv))
+)
+
+m <- add_epred_draws(nd, model_fit_4, ndraws = 50, re_formula = NA) %>% 
     mutate(
         freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
-        doe_std = factor(
-            doe_std, 
-            levels = unique(nd$doe_std), 
-            labels = paste0("DoE: ", c(10, 50, 90), "%")
+        exposure_std = factor(
+            exposure_std, 
+            levels = unique(nd$exposure_std), 
+            labels = c("Minimum exposure", "Mean exposure", "Maximum exposure")
         ),
-        lv_std = factor(
-            lv_std, 
-            levels = unique(nd$lv_std),
-            labels = paste0(c(0, 50, 100), "%")
+        lv_std = factor(lv_std, levels = unique(nd$lv_std), labels = paste0(c(0, 100), "% similarity")
         )
@@
     mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
+    pivot_longer(c(Understands, `Understands and Says`), names_to = ".category", values_to = ".epred") %>% 
     mutate(
-        .category = str_replace_all(
+        .category = factor(
             .category,
-            c(
-                "Understands and Says" = "Production\n(Understands and Says))",
-                "Understands" = "Comprehension\n(Understands)"
-            )
+            levels = c("Understands", "Understands and Says"),
+            labels = c("Comprehension\n(Understands)", "Production\n(Understands and Says")
         )
@@
     mutate(
-        doe_std = case_when(
-            between(doe, 0.00, 0.20) ~ "DoE: 10%",
-            between(doe, 0.40, 0.60) ~ "DoE: 50%",
-            between(doe, 0.80, 1.00) ~ "DoE: 90%"
-        ) %>% 
-            factor(., levels = paste0("DoE: ", percent(c(0.1, 0.5, 0.9)))),
         lv_std = case_when(
-            between(lv, 0.00, 0.20) ~ "0%",
-            between(lv, 0.40, 0.60) ~ "50%",
-            between(lv, 0.80, 1.00) ~ "100%"
-        ) %>% 
-            factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
+            between(lv, 0.00, 0.10) ~ "0% similarity",
+            between(lv, 0.90, 1.00) ~ "100% similarity"
+        ),
+        exposure_std = case_when(
+            between(exposure_std, min(df$exposure_std, na.rm = TRUE), min(df$exposure_std, na.rm = TRUE)+0.5) ~ "Minimum exposure",
+            between(exposure_std, mean(df$exposure_std, na.rm = TRUE)-0.25, mean(df$exposure_std, na.rm = TRUE)+0.25) ~ "Mean exposure",
+            between(exposure_std, max(df$exposure_std, na.rm = TRUE)-0.5, max(df$exposure_std, na.rm = TRUE)) ~ "Maximum exposure"
+        ),
         age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
@@
     ) %>% 
-    drop_na(doe_std, lv_std) %>% 
-    group_by(te, age_std, doe_std, lv_std) %>% 
+    drop_na(exposure_std, lv_std) %>% 
+    group_by(te, age_std, exposure_std, lv_std) %>% 
     summarise(
@@
     mutate(
-        .category = str_replace_all(
+        .category = factor(
             .category,
-            c(
-                "Understands and Says" = "Production\n(Understands and Says))",
-                "Understands" = "Comprehension\n(Understands)"
-            )
+            levels = c("Understands", "Understands and Says"),
+            labels = c("Comprehension\n(Understands)", "Production\n(Understands and Says")
         )
@@
     ) +
-    facet_grid(.category~doe_std) +
-    stat_lineribbon(
-        size = 1,
-        alpha = 0.5,
-        .width = 0.95, 
-        colour = NA,
-        point_interval = mean_hdi
-    ) +
-    stat_summary(
-        data = d,
-        fun.data = "mean_se",
-        geom = "pointrange",
-        size = 0.25
-    ) +
-    # geom_point(
-    #     data = d,
-    #     position = position_dodge(width = 0.25),
-    #     size = 0.5,
-    #     alpha = 0.5
+    facet_grid(.category~exposure_std) +
+    # stat_lineribbon(
+    #     size = 1,
+    #     alpha = 0.5,
+    #     .width = 0.95, 
+    #     colour = NA,
+    #     point_interval = mean_hdi
     # ) +
+    geom_line(aes(group = interaction(.draw, lv_std)), alpha = 0.25) +
     stat_summary(fun = "mean", geom = "line", size = 1) +
@@
 #| fig-cap: "Posterior probability of acquisition by participant"
-re_id <- ranef(model_fit_3)$id[,,1] %>% 
+re_id <- ranef(model_fit_4)$id[,,1] %>% 
     as_tibble() %>% 
     rownames_to_column("id") %>% 
-    left_join(distinct(model_fit_7$data, id)) %>% 
+    left_join(distinct(model_fit_4$data, id)) %>% 
     clean_names() %>% 
@@
             .fns = list(
-                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_7)[1]),
-                Production = ~inv_logit_scaled(. + fixef(model_fit_7)[2])
+                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
+                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
             ), .names = "{.col}__{.fn}"
@@
 
-### Item
+#### Age slope
+
+```{r}
+#| label: fig-results-random-id-age
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_age <- ranef(model_fit_4)$id[,,"age_std"]
+
+re_age %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_4$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### # Phonemes slope
+
+```{r}
+#| label: fig-results-random-id-nphon
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the number of phonemes by participant"
+re_n_phon <- ranef(model_fit_4)$id[,,"n_phon_std"]
+
+re_n_phon %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_4$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Exposure slope
+
+```{r}
+#| label: fig-results-random-id-exposure
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of lexical frequency by participant"
+re_freq <- ranef(model_fit_4)$id[,,"exposure_std"]
+
+re_freq %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_4$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of phonological similarity by participant"
+re_lv <- ranef(model_fit_4)$id[,,"lv_std"]
+
+re_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Doe-by-Levenshtein slope
+
+```{r}
+# | label: fig-results-random-id-exposure-lv
+# | echo: true
+# | message: false
+# | warning: false
+# | fig-width: 7
+# | fig-cap: "Posterior random slope of age by participant"
+re_exposure_lv <- ranef(model_fit_4)$id[,,"exposure_std:lv_std"]
+
+re_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-DoE slope
+
+```{r}
+#| label: fig-results-random-id-age-doe
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
+re_age_exposure <- ranef(model_fit_4)$id[,,"age_std:exposure_std"]
+
+re_age_exposure %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-age-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
+re_age_lv <- ranef(model_fit_4)$id[,,"age_std:lv_std"]
+
+re_age_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-DoE-by-Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-age-exposure-lv
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
+re_age_exposure_lv <- ranef(model_fit_4)$id[,,"age_std:exposure_std:lv_std"]
+
+re_age_exposure_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_3$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(id, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Correlations
+
+```{r}
+#| label: fig-results-random-id-correlations
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior correlations between random effects by participant"
+posterior_description_re %>% 
+    select(parameter, median) %>% 
+    filter(str_detect(parameter, "cor_id__")) %>% 
+    mutate(parameter = str_remove_all(parameter, "cor_id__")) %>%  
+    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
+    mutate(
+        across(
+            c(var_1, var_2), 
+            ~factor(
+                .,
+                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std", "exposure_std:lv_std", "age_std:exposure_std:lv_std"),
+                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein", "Exposure \u00d7 Levenshtein", "Age \u00d7 Exposure \u00d7 Levenshtein")
+            )
+        )
+    ) %>%
+    filter(var_1 != var_2) %>% 
+    ggplot(aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median))) +
+    geom_point(shape = 15, size = 20) +
+    geom_text(colour = "black") +
+    scale_color_distiller(palette = "RdYlBu") +
+    labs(colour = "Correlation", size = "Correlation") +
+    theme(
+        axis.title = element_blank(),
+        axis.text.x = element_text(angle = 35, hjust = 1),
+        panel.grid = element_blank()
+    )
+```
+:::
+
+### Item-level effects
+
+::: panel-tabset
+
+#### Intercept
 
@@
 #| fig-cap: "Posterior probability of acquisition by translation equivalent"
-re_te <- ranef(model_fit_3)$te[,,1] %>% 
+re_te <- ranef(model_fit_4)$te[,,"Intercept"] %>% 
     as_tibble() %>% 
@@
 
+#### Age slope
+
+```{r}
+#| label: fig-results-random-te-age
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by translation equivalent"
+re_age <- ranef(model_fit_4)$te[,,"age_std"]
+
+re_age %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_4$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### # Phonemes slope
+
+```{r}
+#| label: fig-results-random-te-nphon
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
+re_n_phon <- ranef(model_fit_3)$te[,,"n_phon_std"]
+
+re_n_phon %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_3$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Exosure slope
+
+```{r}
+#| label: fig-results-random-te-exposure
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_doe <- ranef(model_fit_4)$te[,,"exposure_std"]
+
+re_doe %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(te = unique(model_fit_4$data$te)) %>%
+    relocate(te) %>%
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-exposure slope
+
+```{r}
+#| label: fig-results-random-te-age-exposure
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_age_doe <- ranef(model_fit_4)$te[,,"age_std:exposure_std"]
+
+re_age_doe %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_4$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(
+        x = estimate,
+        y = reorder(te, estimate),
+        xmin = q2_5,
+        xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Correlations
+
+```{r}
+#| label: fig-results-random-te-correlations
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior correlations between random effects by translation equivalent"
+posterior_description_re %>% 
+    select(parameter, median) %>% 
+    filter(str_detect(parameter, "cor_te__")) %>% 
+    mutate(parameter = str_remove_all(parameter, "cor_te__")) %>%  
+    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
+    mutate(
+        across(
+            c(var_1, var_2), 
+            ~factor(
+                .,
+                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std"),
+                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein")
+            )
+        )
+    ) %>%
+    filter(var_1 != var_2) %>% 
+    ggplot() +
+    aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median)) +
+    geom_point(shape = 15, size = 40) +
+    geom_text(colour = "black") +
+    scale_color_distiller(palette = "RdYlBu") +
+    labs(
+        colour = "Correlation",
+        size = "Correlation"
+    ) +
+    theme(
+        axis.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+:::
+
 ## Model diagnostics
@@
 ```{r}
-#| label: fig-diagnostics-traceplots
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 12
-# gather_draws(
-#     model_fit_3, 
-#     `b_.*`, `sd_te__.*`, `cor_te__.*`, 
-#     regex = TRUE, 
-#     ndraws = 1000
-# ) %>% 
-#     mutate(.chain = paste("Chain ", .chain)) %>% 
-#     ggplot(aes(.iteration, .value, colour = .chain)) +
-#     facet_wrap(~.variable, scales = "free_y") +
-#     annotate(
-#         geom = "rect", 
-#         colour = NA, 
-#         xmin = 0,
-#         xmax = dim(model_fit_3$fit)[1]/2, 
-#         ymin = -Inf, 
-#         ymax = Inf, 
-#         alpha = 0.25, 
-#     ) +
-#     geom_line() +
-#     labs(
-#         x = "Iteration",
-#         y = "Sample value",
-#         colour = "Chain"
-#     ) +
-#     scale_color_d3() +
-#     theme(
-#         legend.position = "top",
-#         legend.title = element_blank(),
-#         panel.grid = element_blank(),
-#         axis.text.x = element_text(size = 7)
-#     )
+# | label: fig-diagnostics-traceplots
+# | echo: true
+# | message: false
+# | warning: false
+# | fig-width: 12
+gather_draws(model_fit_4, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
+    mutate(.chain = paste("Chain ", .chain)) %>%
+    ggplot(aes(.iteration, .value, colour = .chain)) +
+    facet_wrap(~.variable, scales = "free_y") +
+    annotate(
+        geom = "rect",
+        colour = NA,
+        xmin = 0,
+        xmax = dim(model_fit_4$fit)[1]/2,
+        ymin = -Inf,
+        ymax = Inf,
+        alpha = 0.25
+    ) +
+    geom_line() +
+    labs(
+        x = "Iteration",
+        y = "Sample value",
+        colour = "Chain"
+    ) +
+    scale_color_d3() +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid = element_blank(),
+        axis.text.x = element_text(size = 7)
+    )
 ```
@@
 #| eval: false
-mcmc_rhat_data(model_rhats$model_fit_3) %>% 
+mcmc_rhat_data(model_rhats$model_fit_4) %>% 
     ggplot() +
@@
 #| eval: false
-mcmc_neff_data(model_neffs$model_fit_3) %>% 
+mcmc_neff_data(model_neffs$model_fit_4) %>% 
     ggplot() +

### 2022-11-17 manuscript/manuscript.Rmd
Update lab notes

@@
 Adopting a Bayesian IRT modelling, we found conclusive evidence in favour of a facilitatory effect of cognateness, conditional to the degree of exposure to the language: phonological similarity across translation equivalents increased the probability of comprehension and production only when exposure to the language was low.
-We did not find evidence in favour of this interaction behaving differently across ages, although the uncertainty associated with this estimate (the regression coefficient of the $Age \times DoE \times Levenshtein$) was large, indicating that we cannot confidently state whether younger or elder children were any more sentitive to the effect of language exposure and congnateness.
+We did not find evidence in favour of this interaction behaving differently across ages, although the uncertainty associated with this estimate (the regression coefficient of the $Age \times DoE \times Levenshtein$) was large, indicating that we cannot confidently state whether younger or elder children were any more sensitive to the effect of language exposure and cognateness.
 
 ## Results in contrast with the literature
+
+Our result
 

### 2022-11-17 R/00_importing.R

@@ -1,76 +1,33 @@
-
-#' Get Multilex credentials
-#' @examples ml_connect()
-get_credentials <- function(){
-    ml_connect()
-}
-
-#' Get multilex data
-#' @param update A logical value. Should Multilex data be updated and new questionnaire responses be fetched?
-#' @param type A character vector indicating for what type of measure should vocabulary and prevalence data be computed? It takes "understands" and/or "produces"
-#' @returns A named list of data frames containing questionnaire responses, participant data, item data, prevalence data and vocabualary data
-get_multilex <- function(
-        update = FALSE,
-        type = c("understands", "produces")
-){
-    get_credentials() # see ?get_credentials
+#' Get BVQ data
+#' @param ... Arguments to be passed to \code{bvq_responses}
+#' @returns A named list of data frames containing questionnaire responses, participant data, and item data from BVQ
+get_bvq <- function(...){
     
     # get multilex participant data
-    p <- ml_participants() 
+    p <- bvq_participants() 
     # get multilex questionnaire responses
-    r <- ml_responses(p, update = update)
+    r <- bvq_responses(p, ...)
+    
     # merge participant data with questionnaire responses
-    l <- ml_logs(p, r) %>% 
-        mutate_at(
-            vars(starts_with("edu_")),
-            function(x){
-                factor(
-                    x,
-                    levels = c(
-                        "noeducation",
-                        "primary", 
-                        "secondary",
-                        "complementary", 
-                        "vocational",
-                        "university"
-                    ),
-                    ordered = TRUE
-                ) %>% 
-                    as.numeric()
-            }
-        ) %>% 
+    edu_levels <- c("noeducation", "primary", "secondary", "complementary", "vocational", "university") 
+    edu_labels <- c("No education", "Primary", "Secondary", "Complementary", "Vocational", "University")
+    
+    l <- bvq_logs(p, r) %>% 
         mutate(
+            across(starts_with("edu_"), ~as.numeric(factor(.x, levels = edu_levels, ordered = TRUE))),
             # get maximum educational attainment of parents
             edu_parent = apply(cbind(edu_parent1, edu_parent2), 1, max, na.rm = FALSE),
             # recode it as factor
-            edu_parent = factor(
-                edu_parent,
-                levels = 1:6,
-                labels = c(
-                    "No education",
-                    "Primary", 
-                    "Secondary",
[…]

### 2022-11-17 R/01_preprocessing.R

@@ -1,55 +1,39 @@
-
 #' Get item data
-#' @param multilex_data A named list resulting from calling \code{get_multilex}
+#' 
+#' @param bvq_data A named list resulting from calling \code{get_bvq}
 #' @param class A character vector indicating the word classes to be included in the resulting dataset. Takes "Adjective", "Noun" and/or "Verb" as values.
-get_items <- function(
-        multilex_data, 
-        class = c("Adjective", "Noun", "Verb")
-){
+get_items <- function(bvq_data, childes, class = "Noun") {
     
     # compute normalised Levenshtein distance (see ?stringdist::`stringdist-package`)
     # number of edit operations needed to make two strings identical
-    # when applied to phonological transcriptions, it provides an approximation of phonological similarity between two words
-    lv_similarities <- multilex_data$pool %>% 
-        pivot_wider(
-            id_cols = te, 
-            names_from = language,
-            values_from = ipa_flat
-        ) %>% 
+    # when applied to phonological transcriptions, it provides an approximation of phonological similarity between two word-forms
+    lv_similarities <- bvq_data$pool %>% 
+        mutate(language = ifelse(grepl("cat_", item), "Catalan", "Spanish")) %>% 
+        pivot_wider(id_cols = te, names_from = language, values_from = sampa) %>% 
         mutate(
             # make sure strings are coded as UTF-8 before computing LVs
-            lv = stringsim(enc2utf8(Catalan), enc2utf8(Spanish)),
-            lv_dist = as.integer(stringdist(enc2utf8(Catalan), enc2utf8(Spanish)))
+            lv = stringsim(Catalan, Spanish),
+            lv_dist = as.integer(stringdist(Catalan, Spanish))
         ) %>% 
         select(te, lv, lv_dist)
     
-    # see ?multilex::pool
-    single_te <- multilex_data$pool 
-    
     # merge datasets
-    items <- multilex_data$pool %>% 
+    items <- bvq_data$pool %>% 
         rename(list = version) %>% 
         left_join(lv_similarities) %>% # add LVs
         # drop items with missing observations in these variables
-        drop_na(cognate, lv, frequency_zipf, list) %>% 
-        mutate(n_phon = nchar(enc2utf8(ipa_flat))) %>% # number of phonemes
-        # only content words
-        filter(
-            class %in% .env$class, 
+        mutate(n_phon = nchar(sampa_flat)) %>% # number of phoneme
+        # get only translation equivalents with at least one item in each language
[…]

### 2022-11-17 R/02_models.R

@@ -7,18 +7,18 @@ fit_model <- function(name, ...){
     # see R/utils.R
     fit <- brm(
         ...,
-        iter = 4000,
-        chains = 2,
+        iter = 1000,
+        chains = 4,
         init = 0.5, # where to initialise MCMCs
         seed = 888, # for reproducibility
         backend = "cmdstanr", # for faster, less problematic compilation in C++
-        file = here("results", paste0(name, ".rds")), # save model as file
+        file = here("results", name), # save model as file
         # file_refit = "always", # should model be refitted or loaded from file?
         control = list(
             adapt_delta = 0.9, # for better convergence of MCMCs
             max_treedepth = 15
         ),
-        save_model = here("stan", paste0(name, ".stan")) # save Stan code
+        save_model = here("stan", name) # save Stan code
     )
     
     return(fit)
@@ -56,4 +56,6 @@ get_loo <- function(x, ...){
     if (!is.null(names(x))) names(loos) <- names(x) # keep names if any
     return(loos)
 }
+
+#' 
 

### 2022-11-30 index.qmd
Add marginal predictions and effects using the marginaleffects package

@@
 This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
+Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
 This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
@@
 tar_load_everything() # load targets as R objects
-theme_set(theme_custom()) # set custom ggplot theme
 
@@
 conflict_prefer("select", "dplyr")
+
+# set ggplot theme and colour palette
+theme_set(theme_custom()) # set custom ggplot theme
+
 options(
@@
     knitr.duplicate.label = "allow",
+    ggplot2.ordinal.fill = ggsci::pal_d3()(4),
+    ggplot2.ordinal.colour = ggsci::pal_d3()(4),
     ggplot2.discrete.fill = ggsci::pal_d3()(4),
@@
 
-Data and materials are available at OSF (https://osf.io/hy984/), and code is available at GitHub (https://github.com/gongcastro/trajectories).
+Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
 
@@
 #| warning: false
-pool <- mutate(bvqdev::pool, language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish"))
-n_categories <- distinct(pool, category) %>% nrow()
+data("pool", package = "bvqdev")
+n_categories <- distinct(pool, semantic_category) %>% nrow()
 n_items <- length(unique(pool$item))
@@
 The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted of two lists of words: one in Catalan and one in Spanish.
 The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]` items.
@@
 Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\~400 translation equivalents, \~400 items in Catalan, \~400 items in Spanish).
-These last families were randomly allocated into one of four different subsets of the complete list of items.
-These lists were carefully designed so that each contained a representative subsample of the items from the complete list.
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+These last families were randomly assigned to one of four different subsets of the complete list of items.
+These lists were designed so that each contained a representative subsample of the items from the complete list.
+Semantic/functional categories with less than 16 items were not divided in the short version of the questionnaire to preserve a minimum of four items per list for each category: all of their items were included in the four lists.
 Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
@@
 
+## Participants
+
+```{r}
+#| label: fig-participants-time
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 10 
+#| fig-cap: "Date at which responses were collected"
+participants %>% 
+    group_by(time_stamp) %>% 
+    summarise(n = n(), .groups = "drop") %>% 
+    mutate(n = cumsum(n)) %>% 
+    ggplot(aes(time_stamp, n)) +
+    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
+    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
+             angle = 90, vjust = 1.5, hjust = 1) +
+    geom_line(size = 1, colour = pal_d3()(1)[1]) +
+    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
+    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle = 20, hjust = 1),
+        axis.title.x = element_blank(),
+    ) +
+
+    participants %>% 
+    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
+    group_by(time_stamp, age) %>% 
+    summarise(n = n(), .groups = "drop") %>% 
+    group_by(age) %>% 
+    mutate(n = cumsum(n)) %>% 
+    ggplot(aes(time_stamp, n, colour = age)) +
+    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
+    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
+             angle = 90, vjust = 1.5, hjust = 1) +
+    geom_line(size = 1) +
+    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
+    scale_y_continuous(breaks = seq(0, 500, 100)) +
+    scale_color_d3() +
+
+    plot_layout(ncol = 1) &
+
+    scale_x_date(date_breaks = "3 month", date_labels = "%B %Y") &
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle = 20, hjust = 1),
+        axis.title.x = element_blank(),
+    )
+```
+
+### Age
+
+Number of months elapsed between participants' birth date and questionnaire completion.
+
+```{r}
+#| label: fig-participants-age
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Distribution of participants' ages"
+participants %>% 
+    mutate(age = floor(age)) %>% 
+    count(age) %>% 
+    ggplot() +
+    aes(x = age, y = n) +
+    geom_col(fill = pal_d3()(3)[1]) +
+    geom_text(aes(label = n), size = 3.5, vjust = -1) +
+    labs(
+        x = "Age (months)",
+        y = "# participants"
+    ) +
+    scale_x_continuous(breaks = seq(0, 40, 2)) +
+    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
+    theme(
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()
+    )
+```
+
+### Language exposure (Degree of Exposure)
+
+Participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to.
+For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.
+
+```{r}
+#| label: fig-participants-lp
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
+participants %>% 
+    pivot_longer(c(doe_catalan, doe_spanish),
+                 names_to = "language", values_to = "doe", 
+                 names_transform = ~str_to_sentence(str_remove(., "doe_"))) %>% 
+    mutate(doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
+           id = paste0(id, " (", time, ")")
+    ) %>% 
+    count(language, doe) %>% 
+    ggplot(aes(doe, y = n, label = n, fill = language)) +
+    geom_col(colour = "white", size = 1) +
+    geom_text(position = position_nudge(y = 5)) +
+    facet_wrap(~language, ncol = 1) + 
+    labs(
+        x = "Degree of exposure to Catalan",
+        y = "Number of participants",
+        colour = "Language",
+        fill = "Language"
+    ) +
+    theme(
+        legend.position = "top",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor = element_blank()
+    )
+```
+
+```{r}
+#| label: fig-participants-lp
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 15
+#| fig-height: 8
+#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
+participants %>% 
+    mutate(
+        dominance = ifelse(doe_catalan >= doe_spanish,
+                           "Dominant in Catalan",
+                           "Dominant in Spanish"),
+        id = paste0(id, " (", time, ")"),
+        doe_other = 1 - (doe_spanish + doe_catalan)
+    ) %>% 
+    mutate(rank = rank(doe_catalan)) %>% 
+    pivot_longer(
+        starts_with("doe_"), 
+        names_to = "language",
+        values_to = "doe", 
+        names_transform = ~str_to_sentence(str_remove(., "doe_"))
+    ) %>% 
+    mutate(language = factor(language, levels = c("Catalan", "Spanish", "Other"))) %>%
+    ggplot(aes(reorder(id, desc(rank)), doe, colour = language, fill = language)) +
+    facet_wrap(~dominance, ncol = 1, scales = "free_x") +
+    geom_line(aes(group = id), colour = "grey") +
+    geom_point(size = 0.5) +
+    labs(x = "Participant", y = "Degree of exposure (DoE)", 
+         fill = "Language", colour = "Language") +
+    scale_y_continuous(labels = percent) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+### SES/parental education
+
+Maximum educational attainment of the participant parent(s).
+
+```{r}
+#| label: tbl-participants-edu
+#| tbl-cap: Participant exposure to the language (*DoE*) and socio-economic status (*SES*)
+#| echo: true
+#| message: false
+#| warning: false
+participants %>%
+    left_join(bvq_data$logs) %>% 
+    mutate(
+        edu_parent = fct_explicit_na(as.factor(edu_parent)),
+        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE)
+    ) %>% 
+    count(doe_catalan, edu_parent) %>% 
+    right_join(
+        expand_grid(
+            doe_catalan = levels(.$doe_catalan),
+            edu_parent = unique(bvq_data$logs$edu_parent)
+        )
+    ) %>% 
+    rename(total = n) %>% 
+    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
+    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
+    rename(Total = total) %>% 
+    select(-"NA") %>% 
+    gt() %>% 
+    summary_rows(
+        columns = Complementary:Primary, 
+        fns = list(Total = ~sum(., na.rm = TRUE)),
+        formatter = fmt_integer
+    ) %>% 
+    cols_label(doe_catalan = "DoE Catalan") %>%
+    sub_missing(columns = 2:8, missing_text = "--") %>% 
+    tab_spanner("Educational attainment", Secondary:Primary) %>% 
+    tab_style(
+        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
+        locations = cells_body(columns = 9, rows = 1:5)
+    ) %>% 
+    tab_style(
+        cell_text(style = "italic"),
+        cells_column_labels(columns = 2:7)
+    ) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_labels(columns = c(1, 8))
+    ) %>% 
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_spanners(spanners = "Educational attainment")
+    )
+```
+
 ## Items
@@
 Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
-items <- left_join(items, select(pool, te, item, language, class, category))
+items <- left_join(items, select(pool, te, item, language, class, semantic_category))
 
@@
     rename(n = freq_zipf_n) %>% 
-    gt(groupname_col = "language") %>% 
+    gt(groupname_col = "language", rowname_col = "list") %>% 
     fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
@@
 
+Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
+This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points.
+Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [{childesr}](https://langcog.github.io/childes-db-website/) R package.
+Available words in the corpora were then mapped to their Catalan and Spanish translations.
+Responses to words with missing lexical frequency score were excluded from analyses of the total number of items).
+
 ```{r}
@@
 ### \# Phonemes
+
+Word length ($Phonemes$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.
 
@@
     geom_col() +
+    geom_text(aes(label = n), position = position_nudge(y = 5), colour = "black") +
     labs(
@@
 
-### Levenshtein
+### Phonological similarity (Levenshtein)
+
+Phonological similarity between translation equivalents was computed as the normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary].
+This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1.
+This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
 
@@
     distinct(te, .keep_all = TRUE) %>% 
-    ggplot(aes(x = lv)) +
-    geom_dots(fill = ggsci::pal_d3()(4)[1], colour = ggsci::pal_d3()(4)[1]) +
+    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
+    count(lv) %>% 
+    mutate(prop = n/sum(n)) %>% 
+    ggplot(aes(x = lv, n)) +
+    geom_col(
+        fill = ggsci::pal_d3()(4)[1],
+        colour = "white"
+    ) +
+    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
+              position = position_nudge(y = 5)) +
     labs(x = "Levenshtein distance", y = "Probability density") +
-    scale_x_continuous(labels = percent) +
-    theme(
+    theme(
+        panel.grid.major.x = element_blank(),
         legend.position = "top"
@@
 #| warning: false
-#| fig-width: 7
+#| fig-width: 8
+#| fig-height: 4
 #| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
@@
     ) +
-    coord_equal(ratio = 10) +
     scale_x_continuous(breaks = seq(0, 16, 1)) +
@@
 
-## Participants
-
-```{r}
-#| label: fig-participants-time
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Data at which responses were collected"
-participants %>% 
-    group_by(time_stamp) %>% 
-    summarise(n = n(), .groups = "drop") %>% 
-    mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n)) +
-    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
-             angle = 90, vjust = 1.5, hjust = 1) +
-    geom_line(size = 2, colour = pal_d3()(1)[1]) +
-    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
-    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
-    scale_x_date(
-        date_breaks = "3 month", 
-        date_labels = "%B %Y"
-    ) +
-    guides(colour = guide_legend(ncol = 2)) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 20, hjust = 1),
-        axis.title.x = element_blank(),
-    )
-```
-
-### Age
-
-```{r}
-#| label: fig-participants-age
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Distribution of participants' ages"
-participants %>% 
-    mutate(age = floor(age)) %>% 
-    count(age) %>% 
-    ggplot() +
-    aes(x = age, y = n) +
-    geom_col(fill = pal_d3()(3)[1]) +
-    geom_text(aes(label = n), size = 3.5, vjust = -1) +
-    labs(
-        x = "Age (months)",
-        y = "# participants"
-    ) +
-    scale_x_continuous(breaks = seq(0, 40, 2)) +
-    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
-    theme(
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
-
-### Language profile
-
-```{r}
-#| label: fig-participants-lp
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
-participants %>% 
-    mutate(
-        age = round(age),
-        id = paste(id, "(", time, ")"),
-        rank = rank(doe_catalan)
-    ) %>% 
-    pivot_longer(
-        starts_with("doe_"), 
-        names_to = "language",
-        values_to = "doe",
-        names_prefix = "doe_",
-        names_transform = list(language = str_to_sentence)
-    ) %>% 
-    ggplot() +
-    aes(x = doe, y = reorder(id, -rank), fill = language) +
-    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
-    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
-    labs(
-        x = "Degree of Exposure (DoE)",
-        y = "Participant",
-        colour = "Language",
-        fill = "Language"
-    ) +
-    scale_x_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid.major.y = element_blank()
-    )
-```
-
-### SES/parental education
-
-```{r}
-#| label: tbl-participants-edu
-#| tbl-cap: Participant exposure to the language (*DoE*) and socio-economic status (*SES*)
-#| echo: true
-#| message: false
-#| warning: false
-participants %>%
-    left_join(bvq_data$logs) %>% 
-    mutate(
-        edu_parent = fct_explicit_na(as.factor(edu_parent)),
-        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE)
-    ) %>% 
-    count(doe_catalan, edu_parent) %>% 
-    right_join(
-        expand_grid(
-            doe_catalan = levels(.$doe_catalan),
-            edu_parent = unique(bvq_data$logs$edu_parent)
-        )
-    ) %>% 
-    rename(total = n) %>% 
-    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
-    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
-    rename(Total = total) %>% 
-    select(-"NA") %>% 
-    gt() %>% 
-    summary_rows(
-        columns = Complementary:Primary, 
-        fns = list(Total = ~sum(., na.rm = TRUE)),
-        formatter = fmt_integer
-    ) %>% 
-    cols_label(doe_catalan = "DoE Catalan") %>%
-    sub_missing(columns = 2:8, missing_text = "--") %>% 
-    tab_spanner("Educational attainment", Complementary:Primary) %>% 
-    tab_style(
-        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
-        locations = cells_body(columns = 9, rows = 1:5)
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(columns = 2:7)
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = c(1, 8))
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(spanners = "Educational attainment")
-    )
-```
-
-## Exposure-weighted lexical frequency
+### Language exposure-weighted lexical frequency
 
 We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
-This measure is the product of the lexical frequency of the word (Zipf score) and the child's degree of exposure to the corresponding language (a proportion):
+the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion)
 
@@
 #| label: fig-exposure-distribution
+#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
+#| warnings: false
+#| fig-height: 7
+#| fig-width: 13
 df %>% 
-    mutate(language = ifelse(grepl("cat_", item), "Catalan", "Spanish")) %>% 
-    ggplot(aes(exposure, fill = language)) +
+    mutate(
+        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
+        exposure = cut(exposure, breaks = seq(0, 9, 1), include.lowest = TRUE)
+    ) %>% 
+    count(exposure, language) %>% 
+    ggplot(aes(exposure, n, fill = language)) +
     facet_wrap(~language) +
-    geom_histogram(bins = 10, colour = "white", size = 1, show.legend = FALSE) +
-    labs(x = "Exposure (Frequency \u00d7 DoE)", y = "Number of observations") +
+    geom_col(colour = "white") +
+    labs(
+        x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
+        y = "Number of observations",
+        fill ="Language"
+    ) +
 
     df %>% 
-    mutate(language = ifelse(grepl("cat_", item), "Catalan", "Spanish")) %>% 
-    ggplot(aes(x = freq, y = exposure, colour = language)) +
+    mutate(
+        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
+        exposure = cut(exposure, breaks = seq(0, 9, 1), include.lowest = TRUE),
+        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
+    ) %>% 
+    count(exposure, freq, language) %>%    
+    ggplot(aes(x = freq, y = exposure, colour = language, size = n)) +
     facet_wrap(~language) + 
-    geom_point(size = 0.1, alpha = 0.5, show.legend = FALSE) +
-    labs(x = "Frequency (Zipf score)", y = "Exposure") +
-    plot_layout(ncol = 1)
+    geom_point() +
+    labs(
+        x = "Frequency (Zipf score)",
+        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
+        colour = "Language",
+        fill = "Language",
+        size = "Number of observations"
+    ) +
+    theme(panel.grid = element_blank()) +
+
+    plot_layout(ncol = 1) &
+    guides(fill = "none", colour = "none") &
+    theme(panel.grid.major.x = element_blank(),
+          panel.grid.minor.y = element_blank())
 ```
@@
 ### Model comparison
+
+We conducted a series of nested Bayesian ordinal multilevel models to sequentially test the effects of the independent variables (number of phonemes, exposure, age, and Levenshtein) over the three levels of word learning (no, production, comprehension).
+In all these models participants’ and items’ individual differences were controlled by estimating different parameters for each one (different intercepts and/or slopes).
+This is similar to the control conducted by ANCOVAs.
+Thus, independent variable effects account for collective differences in word learning (unexplained by participants’ or items’ individual differences).
 
@@
 Finally, we added the $Age \times Exposure \times Levenshtein$ (`age_std:exposure_std:lv_std`) three-way interaction with its random slope by $Participant$, along with a random slope for the $Age \times Exposure$ (`age_std:exposure_std`) two-way interaction.
+
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO), a relative fit index, via the `loo` R package.
+LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
+Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
 
@@
 
+## Descriptive statistics
+
+::: panel-tabset
+
+### By number of phonemes
+
+```{r}
+#| lab: fig-responses-age-phonemes
+#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
+df %>% 
+    mutate(
+        age = floor(age),
+        n_phon = cut_interval(n_phon, n = 4)
+    ) %>%
+    count(age, n_phon, response) %>% 
+    ggplot(aes(age, n, fill = response)) +
+    facet_wrap(~n_phon) +
+    geom_col(position = position_fill()) +
+    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
+    scale_fill_d3() +
+    scale_y_continuous(labels = percent) +
+    theme_custom() +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+
+### By exposure
+
+```{r}
+#| lab: fig-responses-age-exposure
+#| fig-cap: "Proportion of responses to each category by age and exposure."
+df %>% 
+    mutate(
+        age = floor(age),
+        exposure = cut_interval(exposure, n = 4)
+    ) %>%
+    count(age, exposure, response) %>% 
+    ggplot(aes(age, n, fill = response)) +
+    facet_wrap(~exposure) +
+    geom_col(position = position_fill()) +
+    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
+    scale_fill_d3() +
+    scale_y_continuous(labels = percent) +
+    theme_custom() +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+
+### By Levenshtein
+
+```{r}
+#| lab: fig-responses-age-lv
+#| fig-cap: "Proportion of responses to each category by age and Levenhstein similarity"
+df %>% 
+    mutate(
+        age = floor(age),
+        lv = cut_interval(lv, n = 4)
+    ) %>%
+    count(age, lv, response) %>% 
+    ggplot(aes(age, n, fill = response)) +
+    facet_wrap(~lv) +
+    geom_col(position = position_fill()) +
+    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
+    scale_fill_d3() +
+    scale_y_continuous(labels = percent) +
+    theme_custom() +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+:::
+
 ## Model selection
-
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO) via the `loo` R package.
-LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model as:
 
@@
 posterior_description %>% 
-    mutate(
-        parameter = factor(
-            parameter, 
-            levels = names(str_repl), 
-            labels = str_repl
-        )
-    ) %>%  
+    mutate(parameter = factor(parameter, levels = names(str_repl), labels = str_repl)) %>%  
     select(parameter, median, ci_high, ci_low, rope_percentage, rhat, ess) %>% 
@@
     )
+
 post_draws_list <- split(posterior_description, posterior_description$parameter)
-
 ```
@@
 The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of exposure on the speed at which participants acquired word across ages was inconclusive.
-Give our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.
+This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
+Given our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.
 
@@
 
-post %>%
+post_summary <- post %>% 
     filter(!grepl("Intercept|sd", .variable)) %>%
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    annotate(
-        geom = "rect",
-        ymin = -Inf,
-        ymax = Inf,
-        xmin = rope_interval["lower"],
-        xmax = rope_interval["upper"],
-        colour = NA,
-        alpha = 0.2,
-        fill = pal_d3()(3)[2],
-    ) +
-    geom_vline(
-        xintercept = 0,
-        size = 1,
-        colour = pal_d3()(3)[2]
-    ) +
-    stat_slab(
-        aes(
-            fill = stat(abs(x) < rope_interval["upper"]),
-            colour = stat(abs(x) < rope_interval["upper"]),
-        ),
-        size = 1,
-        # fill = pal_d3()(2)[2],
-        position = position_nudge(y = 0.15)
-    ) +
-    geom_errorbar(
-        data = post %>% 
-            filter(!grepl("Intercept|sd", .variable)) %>%
-            median_hdi(
-                .exclude = c(
-                    ".chain", 
-                    ".iteration",
-                    ".draw",
-                    ".row", 
-                    ".variable_name"
-                )
-            ) %>% 
+    median_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>% 
     mutate(
@@
         )
-            ),
+    )
+
+post %>%
+    filter(!grepl("Intercept|sd", .variable)) %>%
+    ggplot(aes(.value, fct_rev(.variable_name))) +
+    annotate(geom = "rect", ymin = -Inf, ymax = Inf,
+             xmin = rope_interval["lower"], xmax = rope_interval["upper"],
+             colour = NA, alpha = 0.2, fill = pal_d3()(3)[2],) +
+    geom_vline(xintercept = 0, size = 1, colour = pal_d3()(3)[2]) +
+    stat_slab(
+        aes(fill = stat(abs(x) < rope_interval["upper"]),
+            colour = stat(abs(x) < rope_interval["upper"]),),
+        size = 1, position = position_nudge(y = 0.15)
+    ) +
+    geom_errorbar(
+        data = post_summary,
         aes(xmin = .lower, xmax = .upper, x = .value),
@@
     geom_text(
-        data = post %>% 
-            filter(!grepl("Intercept|sd", .variable)) %>%
-            mean_hdi(
-                .exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")
-            ) %>% 
-            mutate(
-                .variable_name = factor(
-                    .variable, 
-                    levels = names(str_repl),
-                    labels = str_repl,
-                    ordered = TRUE
-                )
+        data = post_summary,
+        aes(
+            x = .value, 
+            y = .variable_name,
+            label = paste0(
+                round(.value, 1),
+                " [", round(.lower, 1), ", ", round(.upper, 1), "]")
         ),
-        aes(
-            label = paste0(
-                printnum(.value, digits = 2),
-                " [", printnum(.lower, digits = 2), ", ",
-                printnum(.upper, digits = 2), "]"
-            )
-        ),
-        position = position_nudge(y = -0.2),
-        size = 3
-    ) +
-    # annotate(
-    #     geom = "text",
-    #     label = "ROPE",
-    #     x = 0.3,
-    #     y = 10,
-    #     vjust = 1,
-    #     size = 5,
-    #     colour = pal_d3()(2)[2]
-    # ) +
-    # annotate(
-    #     geom = "curve",
-#     arrow = arrow(length = unit(0.2, "cm")),
-#     x = 0.3,
-#     y = 9.5,
-#     xend = 0.1,
-#     yend = 9,
-#     size = 1,
-#     curvature = -0.25,
-#     colour = pal_d3()(2)[2]
-# ) +
-labs(
-    x = "Logistic regression coefficient estimate",
+        colour = "black", position = position_nudge(y = -0.3), size = 3
+    ) +
+    labs(
+        x = "Logistic regression coefficient estimate (Logit scale)",
         y = "Variable", 
@@
     ) +
-    scale_x_continuous(breaks = seq(-0.5, 2, 0.2)) +
+    scale_x_continuous(breaks = seq(-2, 2, 0.2)) +
     scale_fill_manual(
@@
 
-## Marginal means
+## Marginal predictions
 
@@
 Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
-nd <- expand.grid(
-    n_phon_std = 0,
-    freq_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
-    exposure_std = c(min(df$exposure_std, na.rm = TRUE), 0, max(df$exposure_std, na.rm = TRUE)),
-    lv_std = (c(0, 1)-mean(df$lv))/(sd(df$lv))
-)
-
-m <- add_epred_draws(nd, model_fit_4, ndraws = 50, re_formula = NA) %>% 
+
+nd <- datagrid(
+    model = model_fit_4,
+    age_std = seq_range(model_fit_4$data$age_std, n = 100),
+    lv_std = (seq(0, 1, 0.5)-mean(df$lv))/(sd(df$lv)),
+    exposure_std = c(-1, 0, 1),
+    te = NA,
+    id = NA
+)
+
+marg_epreds <- predictions(
+    model_fit_4,
+    newdata = nd,
+    ndraws = 25, 
+    re_formula = NA
+)
+
+marg_epreds %>% 
+    posteriordraws() %>% 
+    as_tibble() %>% 
+    select(drawid, group, draw, age_std, lv_std, exposure_std) %>%
+    filter(group != "No") %>% 
+    pivot_wider(
+        id_cols = c(drawid, age_std, lv_std, exposure_std), 
+        names_from = group, 
+        values_from = draw
+    ) %>% 
+    mutate(`Understands` = `Understands and Says` + `Understands`) %>% 
+    pivot_longer(c(`Understands`, `Understands and Says`), names_to = "group", values_to = "draw") %>% 
     mutate(
-        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
         exposure_std = factor(
@@
             levels = unique(nd$exposure_std), 
-            labels = c("Minimum exposure", "Mean exposure", "Maximum exposure")
+            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
         ),
-        lv_std = factor(lv_std, levels = unique(nd$lv_std), labels = paste0(c(0, 100), "% similarity")
-        )
-    ) %>% 
-    filter(.category != "No") %>% 
-    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(c(Understands, `Understands and Says`), names_to = ".category", values_to = ".epred") %>% 
-    mutate(
-        .category = factor(
-            .category,
-            levels = c("Understands", "Understands and Says"),
-            labels = c("Comprehension\n(Understands)", "Production\n(Understands and Says")
-        )
-    )
-
-d <- df %>%
-    mutate(
-        lv_std = case_when(
-            between(lv, 0.00, 0.10) ~ "0% similarity",
-            between(lv, 0.90, 1.00) ~ "100% similarity"
-        ),
-        exposure_std = case_when(
-            between(exposure_std, min(df$exposure_std, na.rm = TRUE), min(df$exposure_std, na.rm = TRUE)+0.5) ~ "Minimum exposure",
-            between(exposure_std, mean(df$exposure_std, na.rm = TRUE)-0.25, mean(df$exposure_std, na.rm = TRUE)+0.25) ~ "Mean exposure",
-            between(exposure_std, max(df$exposure_std, na.rm = TRUE)-0.5, max(df$exposure_std, na.rm = TRUE)) ~ "Maximum exposure"
-        ),
-        age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
-            cut(
-                .,
-                breaks = ((seq(10, 38, 2)-mean(df$age))/(sd(df$age))),
-                labels = ((seq(10, 36, 2)-mean(df$age))/(sd(df$age))),
-                include.lowest = TRUE
-            ) %>% 
-            as.character() %>% 
-            as.numeric(),
-        Understands = response %in% c("Understands", "Understands and Says"),
-        `Understands and Says` = response %in% "Understands and Says"
-    ) %>% 
-    drop_na(exposure_std, lv_std) %>% 
-    group_by(te, age_std, exposure_std, lv_std) %>% 
-    summarise(
-        Understands = mean(Understands),
-        `Understands and Says` = mean(`Understands and Says`),
-        .groups = "drop"
-    ) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
-    mutate(
-        .category = factor(
-            .category,
-            levels = c("Understands", "Understands and Says"),
-            labels = c("Comprehension\n(Understands)", "Production\n(Understands and Says")
-        )
-    )
-
-m %>% 
-    ggplot() +
-    aes(
-        x = age_std, 
-        y = .epred, 
-        colour = lv_std, 
-        fill = lv_std,
-    ) +
-    facet_grid(.category~exposure_std) +
-    # stat_lineribbon(
-    #     size = 1,
-    #     alpha = 0.5,
-    #     .width = 0.95, 
-    #     colour = NA,
-    #     point_interval = mean_hdi
-    # ) +
-    geom_line(aes(group = interaction(.draw, lv_std)), alpha = 0.25) +
-    stat_summary(fun = "mean", geom = "line", size = 1) +
+        lv_std = factor(
+            lv_std,
+            levels = unique(nd$lv_std), 
+            labels = paste0(c(0, 50, 100), "% similarity")
+        )
+    ) %>% 
+    ggplot(aes(age_std, draw, colour = lv_std, fill = lv_std)) +
+    facet_grid(group~exposure_std) +
+    geom_line(aes(group = interaction(lv_std, drawid)), 
+              size = 0.75, alpha = 0.15) +
+    stat_summary(fun = mean, geom = "line", size = 1) +
+    labs(
+        x = "Age (months)",
+        y = "Probability of acquisition",
+        colour = "Phonological similarity (Levenshtein)",
+        fill = "Phonological similarity (Levenshtein)"
+    ) +
+    scale_y_continuous(
+        breaks = seq(0, 1, 0.25),
+        limits = c(0, 1),
+        labels = percent
+    ) +
     scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 4)
-    ) +
-    scale_color_d3() +
-    labs(
-        x = "Age (months)", 
-        y = "Posterior probability of acquisition\n(Linear prediction)",
-        colour = "Phonological similarity (Levenshtein distance)", 
-        fill = "Phonological similarity (Levenshtein distance)"
-    ) +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
+        breaks = (seq(10, 32, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 32, 4)
+    ) +
     theme(
         legend.position = "top",
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        strip.background = element_rect(fill = "grey", colour = NA),
-        plot.background = element_rect(fill = "white", colour = NA)
-    )
+        panel.grid = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
+    )
+
+```
+
+## Marginal effects
+
+In order to examine the impact of the word exposure index on the effect of phonological similarity we used the [`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects/index.html) R package to compute the average marginal effect of the Levenshtein distance on the probability of acquisition (comprehension or production) for the range of observed values of the word exposure index in our dataset.
+
+```{r}
+#| label: fig-marginal-effect
+#| fig-cap: "Average marginal effect of phonological similarity (Levenshtein distance) on the probability of comprehension and production across the range of values of exposure.
+The observed range of values of the word exposure index is indicated in the X-axis, and the estimated average marginal effect of phonological similarity (Levenshtein distance) is indicated in the Y-axis.
+This quantity is generated by averaging the derivative of the regression coefficient of the predictor Levenshtein across the range of values of all predictors but Exposure.
+We represent the uncertainty associated with the resulting estimate by generating multiple values of this marginal effect by drawing multiple samples from the posterior distribution of the model coefficients, calculating the average marginal effect for each posterior draw, and plotting the result as a thin grey line.
+Closer lines indicate lower uncertainty over the true value of the average marginal effect of Levenshein, while more distant lines indicate higher uncertainty."
+nd <- datagrid(
+    model = model_fit_4,
+    n_phon_std = 0,
+    lv_std = 0,
+    exposure_std = seq_range(model_fit_4$data$exposure_std, 10),
+    age_std = 0,
+    id = NA,
+    te = NA
+)
+
+marg_effects <- marginaleffects(
+    model_fit_4,
+    newdata = nd, 
+    re_formula = NA,
+    ndraws = 25,
+    vcov = FALSE,
+    type = "link"
+)
+
+marg_effects %>% 
+    posteriordraws() %>%
+    as_tibble() %>% 
+    filter(term=="lv_std") %>% 
+    mutate(draw = draw/4) %>% 
+    ggplot(aes(exposure_std, draw)) +
+    geom_hline(yintercept = 0, size = 1, colour = "grey") +
+    geom_line(aes(group = drawid), alpha = 0.5, size = 0.75, colour = "grey") +
+    stat_summary(fun = median, geom = "line", size = 1) +
+    labs(
+        x = "Exposure index",
+        y = "Marginal effect of phonological similarity\n(Levenshtein distance)"
+    ) +
+    scale_x_continuous(
+        labels = label_number(style_positive = "plus", suffix = " SD"),
+    ) +
+    scale_y_continuous(
+        labels = label_percent(style_positive = "plus", ),
+        breaks = seq(-1, 1, 0.05)
+    ) +
+    theme(
+        legend.position = "top",
+        panel.grid = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
+    )
+
 ```

### 2022-11-30 R/marginal-effects.R
Add marginal predictions and effects using the marginaleffects package

@@ -1,0 +1,38 @@
+library(marginaleffects)
+
+nd <- datagrid(
+    model = model_fit_4,
+    n_phon_std = 0,
+    lv_std = 0,
+    exposure_std = c(-1, 0, 1),
+    age_std = 0,
+    id = NA,
+    te = NA
+)
+
+emmeans <- marginaleffects(
+    model_fit_4,
+    newdata = nd, 
+    re_formula = NA,
+    ndraws = 25,
+    vcov = FALSE, 
+)
+
+emmeans %>% 
+    posteriordraws() %>%
+    as_tibble() %>% 
+    mutate(
+        exposure_std = factor(
+            exposure_std, 
+            levels = unique(nd$exposure_std), 
+            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
+        ),
+        # lv_std = factor(
+        #     lv_std,
+        #     levels = unique(nd$lv_std), 
+        #     labels = paste0(c(0, 50, 100), "% similarity")
+        # )
+    ) %>% 
+    ggplot(aes(exposure_std, dydx, colour = lv_std)) +
+    facet_grid(group~term) +
+    geom_line(aes(group = interaction(drawid, lv_std)))

### 2022-12-01 index.qmd
Replace colour palettes with achromatic-safe colours

@@
 theme_set(theme_custom()) # set custom ggplot theme
+clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
 
@@
     knitr.duplicate.label = "allow",
-    ggplot2.ordinal.fill = ggsci::pal_d3()(4),
-    ggplot2.ordinal.colour = ggsci::pal_d3()(4),
-    ggplot2.discrete.fill = ggsci::pal_d3()(4),
-    ggplot2.discrete.colour = ggsci::pal_d3()(4),
+    ggplot2.ordinal.fill = clrs,
+    ggplot2.ordinal.colour = clrs,
+    ggplot2.discrete.fill = clrs,
+    ggplot2.discrete.colour = clrs,
     ggplot2.continuous.fill = ggplot2::scale_color_gradient,
@@
 ```
+
+::: {.callout-tip}
+Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
+:::
 
@@
              angle = 90, vjust = 1.5, hjust = 1) +
-    geom_line(size = 1, colour = pal_d3()(1)[1]) +
+    geom_line(size = 1, colour = clrs[1]) +
     labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
@@
     geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
-             angle = 90, vjust = 1.5, hjust = 1) +
     geom_line(size = 1) +
     labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
-    scale_y_continuous(breaks = seq(0, 500, 100)) +
-    scale_color_d3() +
+    scale_y_continuous(breaks = seq(0, 500, 50)) +
+    scale_colour_manual(values = clrs) +
 
@@
     aes(x = age, y = n) +
-    geom_col(fill = pal_d3()(3)[1]) +
+    geom_col(fill = clrs[1]) +
     geom_text(aes(label = n), size = 3.5, vjust = -1) +
@@
 participants %>% 
-    pivot_longer(c(doe_catalan, doe_spanish),
+    pivot_longer(
+        c(doe_catalan, doe_spanish),
         names_to = "language", values_to = "doe", 
-                 names_transform = ~str_to_sentence(str_remove(., "doe_"))) %>% 
-    mutate(doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
+        names_transform = ~str_to_sentence(str_remove(., "doe_"))
+    ) %>% 
+    mutate(
+        doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
         id = paste0(id, " (", time, ")")
@@
     ) +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
     theme(
@@
 
-```{r}
-#| label: fig-participants-lp
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 15
-#| fig-height: 8
-#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
-participants %>% 
-    mutate(
-        dominance = ifelse(doe_catalan >= doe_spanish,
-                           "Dominant in Catalan",
-                           "Dominant in Spanish"),
-        id = paste0(id, " (", time, ")"),
-        doe_other = 1 - (doe_spanish + doe_catalan)
-    ) %>% 
-    mutate(rank = rank(doe_catalan)) %>% 
-    pivot_longer(
-        starts_with("doe_"), 
-        names_to = "language",
-        values_to = "doe", 
-        names_transform = ~str_to_sentence(str_remove(., "doe_"))
-    ) %>% 
-    mutate(language = factor(language, levels = c("Catalan", "Spanish", "Other"))) %>%
-    ggplot(aes(reorder(id, desc(rank)), doe, colour = language, fill = language)) +
-    facet_wrap(~dominance, ncol = 1, scales = "free_x") +
-    geom_line(aes(group = id), colour = "grey") +
-    geom_point(size = 0.5) +
-    labs(x = "Participant", y = "Degree of exposure (DoE)", 
-         fill = "Language", colour = "Language") +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        axis.text.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid = element_blank()
-    )
-```
 ### SES/parental education
@@
 Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
-items <- left_join(items, select(pool, te, item, language, class, semantic_category))
+items <- left_join(items, select(bvqdev::pool, te, item, language, class, semantic_category))
 
@@
     gt(groupname_col = "language", rowname_col = "list") %>% 
-    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
-    fmt_percent(matches("lv_")) %>% 
+    fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>% 
+    fmt_percent(matches("lv_"), decimals = 1) %>% 
     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
@@
         n = md("*N*"),
-        freq_zipf_mean = md("Mean (SD)"),
+        freq_zipf_mean = md("Mean ± SD"),
         freq_zipf_min = md("Range"),
-        n_phon_mean = md("Mean (SD)"),
+        n_phon_mean = md("Mean ± SD"),
         n_phon_min = md("Range"),
-        lv_mean = md("Mean (SD)"),
+        lv_mean = md("Mean ± SD"),
         lv_min = md("Range")
@@
 #| warning: false
-#| fig-width: 7
+#| fig-width: 9
 #| fig-cap: "Distribution of Levenshtein similarities across items"
@@
     ggplot(aes(x = lv, n)) +
-    geom_col(
-        fill = ggsci::pal_d3()(4)[1],
-        colour = "white"
-    ) +
+    geom_col(fill = clrs[1], colour = "white") +
     geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
-              position = position_nudge(y = 5)) +
+              position = position_nudge(y = 3), size = 3.5) +
     labs(x = "Levenshtein distance", y = "Probability density") +
@@
 items %>% 
-    left_join(select(pool, item, language)) %>% 
+    left_join(select(bvqdev::pool, item, language)) %>% 
     drop_na(freq_zipf) %>% 
@@
 
-We conducted a series of nested Bayesian ordinal multilevel models to sequentially test the effects of the independent variables (number of phonemes, exposure, age, and Levenshtein) over the three levels of word learning (no, production, comprehension).
+We extracted four variables of interest for each observation (describe in previous sections):
+
+* Word lexical frequency ($Frequency$)
+* Word length ($Phonemes$)
+* Phonological similarity ($Levenshtein$)
+* Participant age ($Age$)
+* Word exposure index ($Exposure$)
+
+We fitted a series of nested Bayesian ordinal multilevel models to sequentially test the effects of the independent variables (number of phonemes, exposure, age, and Levenshtein) over the three levels of word learning (no, production, comprehension).
 In all these models participants’ and items’ individual differences were controlled by estimating different parameters for each one (different intercepts and/or slopes).
@@
 
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO), a relative fit index, via the `loo` R package.
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package.
 LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
@@
 
-#### Prior predictions
-
 ::: {.panel-tabset}
 
-##### Prior-predicted mean
+#### Prior-predicted mean
 
@@
         breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 4),
-        sec.axis = dup_axis(
-            name = "Age (years; months)",
-            labels = months_to_years(
-                rescale_variable(
-                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-                    mean(df$age),
-                    sd(df$age)
-                )
-            )
-        )
-    ) +
-    scale_fill_brewer(palette = "Reds", labels = c("95%", "89%", "67%", "50%")) +
+        labels = seq(10, 36, 4)
+    ) +
+    scale_fill_manual(
+        values = rev(clrs), 
+        labels = c("95%", "89%", "67%", "50%")
+    ) +
     labs(
@@
     theme(
-        legend.position = "top",
         axis.ticks.x = element_line(colour = "black"),
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        strip.background = element_rect(fill = "grey", colour = NA),
-        plot.background = element_rect(fill = "white", colour = NA)
-    )
-```
-
-##### Prior-predictions
+        panel.grid = element_blank()
+    )
+```
+
+#### Prior-predictions
 
@@
     geom_col(position = position_fill()) +
-    scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 4),
-        sec.axis = dup_axis(
-            name = "Age (years; months)",
-            labels = months_to_years(
-                rescale_variable(
-                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-                    mean(df$age),
-                    sd(df$age)
-                )
-            )
-        )
-    ) +
     labs(
@@
     ) +
+    scale_x_continuous(
+        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 36, 4)
+    ) +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
     scale_y_continuous(labels = percent) +
@@
 ```
+
+:::
 
@@
     labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_d3() +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
     scale_y_continuous(labels = percent) +
@@
 #| fig-width: 7
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
+#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
+This procedure involves estimating the likelihood of each observation given a model fitted without such observation.
+The resulting likelihoods of each model were summed to compute the expected log-predicted density (*ELPD*) of the model, and correcting it for the number of parameters in the model to penalised for its the complexity.
+*ELPD* values closer to zero indicate better predictive acuracy.
+We report the *ELPD*, effective number of parameters (*p*), and information criterion (*LOO-IC*) of each model along their associated standard errors in paratheses.
+The first column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_4`)."
 model_loos %>% 
@@
         model, 
+        matches("diff"),
         matches("elpd_loo"),
         matches("p_loo"), 
-        matches("looic"), 
-        matches("diff")
+        matches("looic")
     ) %>%
     gt() %>% 
-    tab_header(
-        title = "",
-        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
-The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_3`)")
-    ) %>% 
-    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
-    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>% 
-    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>% 
-    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>% 
-    fmt_number(2:9) %>% 
+    fmt_number(2:9, decimals = 1) %>% 
+    cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>% 
+    cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>% 
+    cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>% 
+    cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>% 
     cols_label(
         model = "Model",
-        # elpd
         elpd_loo = md("*ELPD*"),
-        se_elpd_loo = md("*SE*"),
-        # p
         p_loo = md("*p*"),
-        se_p_loo = md("*SE*"),
-        # looic
         looic = md("*LOO-IC*"),
-        se_looic = md("*SE*"),
-        # diff
-        elpd_diff = md("*diff*"),
-        se_diff = md("*SE*")
+        elpd_diff = md("Difference")
     ) %>% 
@@
         cell_text(weight = "bold"),
-        cells_column_spanners()
+        cells_column_labels()
     ) %>% 
     tab_style(
-        cell_text(align = "left"),
-        cells_title(groups = "subtitle")
+        cell_fill(color = "grey", alpha = 0.5),
+        cells_body(matches("diff"))
     )
@@
         geom_point(
-            size = 0.5, alpha = 0.1, colour = pal_d3()(3)[2], 
+            size = 0.5, alpha = 0.1, colour = clrs[5], 
             shape = 1
@@
     mutate(parameter = factor(parameter, levels = names(str_repl), labels = str_repl)) %>%  
-    select(parameter, median, ci_high, ci_low, rope_percentage, rhat, ess) %>% 
+    select(parameter, median, ci_high, ci_low, rope_percentage) %>% 
     gt(rowname_col = "parameter") %>% 
-    fmt_number(c(1:4, 6:7)) %>% 
-    fmt_percent(5) %>% 
+    fmt_number(c(1:4), decimals = 2) %>% 
+    fmt_percent(5, decimals = 2) %>% 
     cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>% 
@@
         median = "Median",
-        rhat = "R-hat",
-        ess = "ESS",
         ci_low = md("95% *HDI*"),
@@
         cell_text(weight = "bold"),
-        cells_column_labels(columns = 1:7)
+        cells_column_labels(columns = 1:5)
     )
@@
              xmin = rope_interval["lower"], xmax = rope_interval["upper"],
-             colour = NA, alpha = 0.2, fill = pal_d3()(3)[2],) +
-    geom_vline(xintercept = 0, size = 1, colour = pal_d3()(3)[2]) +
+             colour = NA, alpha = 0.2, fill = clrs[5],) +
+    geom_vline(xintercept = 0, size = 1, colour = clrs[5]) +
     stat_slab(
@@
     scale_fill_manual(
-        values = c("dodgerblue", "grey"),
+        values = clrs[c(1, 4)],
         labels = c("No", "Yes")
@@
     scale_colour_manual(
-        values = c("dodgerblue", "grey"),
+        values = clrs[c(1, 4)],
         labels = c("No", "Yes")
@@
     newdata = nd,
-    ndraws = 25, 
+    ndraws = NULL, 
     re_formula = NA
@@
     facet_grid(group~exposure_std) +
-    geom_line(aes(group = interaction(lv_std, drawid)), 
-              size = 0.75, alpha = 0.15) +
-    stat_summary(fun = mean, geom = "line", size = 1) +
+    stat_summary(fun.data = median_hdi, geom = "ribbon",
+                 alpha = 0.5, colour = NA) +
+    # geom_line(aes(group = interaction(lv_std, drawid)), 
+    #           size = 0.75, alpha = 0.15) +
+    stat_summary(aes(linetype = lv_std), fun = median, geom = "line", size = 1) +
     labs(
@@
         y = "Probability of acquisition",
+        linetype = "Phonological similarity (Levenshtein)",
         colour = "Phonological similarity (Levenshtein)",
@@
     ) +
+    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
     scale_y_continuous(
@@
     re_formula = NA,
-    ndraws = 25,
+    ndraws = NULL,
     vcov = FALSE,
@@
     geom_hline(yintercept = 0, size = 1, colour = "grey") +
-    geom_line(aes(group = drawid), alpha = 0.5, size = 0.75, colour = "grey") +
-    stat_summary(fun = median, geom = "line", size = 1) +
+    # geom_line(aes(group = drawid), alpha = 0.25, size = 0.75, colour = clrs[1]) +
+    stat_summary(fun.data = median_hdi, geom = "ribbon",
+                 colour = NA, alpha = 0.5, fill = clrs[1]) +
+    stat_summary(fun = median, geom = "line", size = 1, color = clrs[1]) +
     labs(
@@
     ggplot() +
-    aes(
-        x = estimate, 
-        y = reorder(id, estimate), 
-        xmin = q2_5, 
-        xmax = q97_5,
-        fill = type,
-        colour = type
-    ) +
+    aes(x = estimate, y = reorder(id, estimate), 
+        xmin = q2_5, xmax = q97_5,
+        fill = type, colour = type) +
     facet_wrap(~type) +
@@
     ggplot() +
-    aes(
-        x = estimate,
-        y = reorder(id, estimate),
-        xmin = q2_5,
-        xmax = q97_5
-    ) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
     geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
@@
     aes(
-        x = estimate,
-        y = reorder(id, estimate),
-        xmin = q2_5,
-        xmax = q97_5
-    ) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
+        x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
     geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
@@
     filter(var_1 != var_2) %>% 
-    ggplot(aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median))) +
+    ggplot(aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1))) +
     geom_point(shape = 15, size = 20) +
     geom_text(colour = "black") +
-    scale_color_distiller(palette = "RdYlBu") +
+    scale_colour_gradientn(colours = clrs) +
     labs(colour = "Correlation", size = "Correlation") +
@@
     ggplot() +
-    aes(
-        x = estimate, 
-        y = reorder(te, estimate), 
-        xmin = q2_5,
-        xmax = q97_5,
-        colour = type,
-        fill = type
-    ) +
+    aes(x = estimate, y = reorder(te, estimate), 
+        xmin = q2_5, xmax = q97_5,
+        colour = type, fill = type) +
     facet_wrap(~type) +
@@
     ggplot() +
-    aes(
-        x = estimate,
-        y = reorder(te, estimate),
-        xmin = q2_5,
-        xmax = q97_5
-    ) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
+    aes(x = estimate, y = reorder(te, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
     geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
@@
     ggplot() +
-    aes(
-        x = estimate,
-        y = reorder(te, estimate),
-        xmin = q2_5,
-        xmax = q97_5
-    ) +
+    aes(x = estimate, y = reorder(te, estimate),
+        xmin = q2_5, xmax = q97_5) +
     geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
@@
     ggplot() +
-    aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median)) +
+    aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1)) +
     geom_point(shape = 15, size = 40) +
     geom_text(colour = "black") +
-    scale_color_distiller(palette = "RdYlBu") +
+    scale_colour_gradientn(colours = clrs) +
     labs(
@@
 
+### Simulation-based calibration
+
+```{r}
+#| label: fig-simulation-based-calibration
+#| fig-cap: "Posterior predictive checks.
+We simulated 500 datasets from our model by drawing random samples from its posterior distribution.
+We then computed the proportion of 'No', 'Understands', and 'Understands and Says' responses in each dataset.
+These simulated proportions are depicted in the plot as solid horizontal lines.
+Their variation indicate the uncertainy of the model about what proportion of each type of response would be observe in a newly collected dataset.
+We also plotted the proportion of each type of response in the observed dataset (with a ± 5% interval for comparison)."
+predicted_dfs <- add_predicted_draws(
+    object = model_fit_4, 
+    newdata = df,
+    re_formula = NULL,
+    ndraws = 500,
+    seed = 888
+)
+
+observed_prop <- df %>% 
+    count(response) %>%
+    mutate(prop = n/sum(n)) %>% 
+    rename(.prediction = response)
+
+predicted_dfs %>% 
+    ungroup() %>% 
+    select(.prediction, .draw) %>% 
+    count(.draw, .prediction) %>%
+    group_by(.draw) %>% 
+    mutate(
+        prop = n/sum(n),
+        .prediction = factor(
+            .prediction,
+            levels = levels(df$response), 
+            ordered = TRUE
+        )
+    ) %>% 
+    ggplot(aes(.draw, prop, fill = .prediction, colour = .prediction)) +
+    geom_line(aes(group = .prediction)) +
+    geom_rect(
+        data = observed_prop,
+        aes(xmin = 0, xmax = max(predicted_dfs$.draw),
+            ymin = prop-0.025, ymax = prop+0.025,
+            colour = .prediction, fill = .prediction),
+        alpha = 0.25, colour = NA, inherit.aes = FALSE
+    ) +
+    # geom_segment(
+    #     data = observed_prop,
+    #     aes(x = -Inf, xend = Inf, y = prop, yend = prop,
+    #         colour = .prediction),
+    #     size = 1, inherit.aes = FALSE
+    # ) +
+    labs(
+        x = "Iteration",
+        y = "Proportion of responses",
+        colour = "Response",
+        fill = "Response"
+    ) +
+    scale_colour_d3() +
+    scale_fill_d3() +
+    scale_x_continuous(labels = label_number(big.mark = ","), 
+                       breaks = seq(0, max(predicted_dfs$.draw), 100)) +
+    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
+    theme(
+        legend.position = "top",
+        panel.grid = element_blank()
+    )
+
+```
+
 ### Traceplots
@@
 ```{r}
-# | label: fig-diagnostics-traceplots
-# | echo: true
-# | message: false
-# | warning: false
-# | fig-width: 12
+#| label: fig-diagnostics-traceplots
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 12
 gather_draws(model_fit_4, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
@@
     facet_wrap(~.variable, scales = "free_y") +
-    annotate(
-        geom = "rect",
-        colour = NA,
-        xmin = 0,
-        xmax = dim(model_fit_4$fit)[1]/2,
-        ymin = -Inf,
-        ymax = Inf,
-        alpha = 0.25
-    ) +
+    annotate(geom = "rect", colour = NA,
+             xmin = 0, xmax = dim(model_fit_4$fit)[1]/2,
+             ymin = -Inf, ymax = Inf,
+             alpha = 0.25) +
     geom_line() +
@@
     ) +
-    scale_color_d3() +
+    scale_colour_manual(values = clrs[c(1, 3, 4, 5)]) +
     theme(
@@
     aes(x = value) +
-    geom_histogram(
-        bins = 100, 
-        na.rm = TRUE, 
-        fill = pal_d3()(1)[1], 
-        colour = "white"
-    ) +
-    geom_rug(alpha = 0.01, colour = pal_d3()(1)[1]) +
+    geom_histogram(bins = 100, na.rm = TRUE, fill = clrs[1], colour = "white") +
+    geom_rug(colour = clrs[1]) +
     labs(
@@
     aes(x = value, fill = description, colour = description) +
-    geom_histogram(
-        bins = 100, 
-        na.rm = TRUE, 
-        colour = "white"
-    ) +
-    geom_rug(alpha = 0.01) +
+    geom_histogram(bins = 100, na.rm = TRUE, colour = "white") +
+    geom_rug() +
     geom_vline(xintercept = 1, colour = "black", size = 0.75) +
@@
     ) +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
+    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
     theme(
@@
 ```{r}
-sessionInfo()
-```
+sessioninfo::session_info()
+```

### 2022-12-01 index.qmd
Replace colour palettes with achromatic-safe colours

@@
 #| fig-width: 7
-#| fig-cap: "Lexical frequencies.
+#| tbl-cap: "Lexical frequencies.
 Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
@@
     theme(
+        legend.position = "top",
         axis.ticks.x = element_line(colour = "black"),
@@
     scale_y_continuous(labels = percent) +
-    theme_custom() +
     theme(
@@
     labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_d3() +
     scale_y_continuous(labels = percent) +
-    theme_custom() +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
     theme(
@@
 Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.
-
-[TO DO: emmeans tests]
 
@@
     lv_std = 0,
-    exposure_std = seq_range(model_fit_4$data$exposure_std, 10),
+    exposure_std = seq(-1.5, 1.5, length.out = 100),
     age_std = 0,
@@
     labs(
-        x = "Exposure index",
+        x = "Word exposure index",
         y = "Marginal effect of phonological similarity\n(Levenshtein distance)"
@@
         labels = label_number(style_positive = "plus", suffix = " SD"),
+        breaks = seq(-2, 2, 0.5)
     ) +

### 2022-12-02 index.qmd
Fix colours in report

@@
     ) +
-    scale_colour_d3() +
-    scale_fill_d3() +
+    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
     scale_x_continuous(labels = label_number(big.mark = ","), 

### 2022-12-09 dag.tex
Update model formula and add DAG

\begin{tikzpicture}

  % nodes
  \node (age) at (0,0) {$Age_{i, j}$};
  \node (phonemes) at (3,4) {$Phonemes_{i, j}$};
  \node (exposure) at (3,2) {$Exposure_{i, j}$};
  \node (cognateness) at (3,-2) {$Cognateness_{i}$};
  \node [ellipse, draw](p_acquisition) at (7, 0) {$P(Acquisition)$};
  \node (response) at (11, 0) {$Response$};
  % arrows
  \draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (exposure);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (cognateness);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (cognateness);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (phonemes) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (cognateness) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (p_acquisition) -- (response);

\end{tikzpicture}

### 2022-12-09 index.qmd
Update model formula and add DAG

--- 2022-12-01 index.qmd
@@
     affiliations:
-      - name: Universidad Atónoma de Madrid
+      - name: Universidad Autónoma de Madrid
         department: Departamento de Psicología
@@
 
+# Methods
+
 ::: {.callout-tip}
@@
 :::
-
-# Methods
-
-Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
 
@@
     ) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
     theme(
@@
 Available words in the corpora were then mapped to their Catalan and Spanish translations.
-Responses to words with missing lexical frequency score were excluded from analyses of the total number of items).
+Therefore, the Catalan and Spanish word-forms of the same translation equivalent had the same value for lexical frequency.
+Responses to words with missing lexical frequencies were excluded from analyses of the total number of items).
 
@@
 #| fig-width: 7
-#| fig-cap: "Distribution of lexical frequency across items"
+#| fig-height: 4
+#| fig-cap: "Distribution of lexical frequency across items.
+Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations.
+Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency.
+The black dot and interval indicate the mean and standard deviation."
+
+mean_sd <- function(x, ...) {
+    m <- mean(x, ...)
+    s <- sd(x, ...)
+    data.frame(mean = m, ymin = m-s, ymax = m+s)
+}
+
 items %>% 
-    ggplot() +
-    aes(
-        x = language,
-        y = freq_zipf,
-        fill = language,
-        colour = language
-    ) +
-    stat_slab(slab_colour = "white", slab_size = 1, alpha = 0.5, scale = 0.5) +
-    stat_dots(side = "left", scale = 0.5, position = position_nudge(x = -0.05)) + 
-    stat_pointinterval(
-        position = position_dodge(width = 0.1),
-        point_size = 2.5,
-        colour = "black"
-    ) +
-    labs(
-        x = "Frequency (Zipf)",
-        y = "Probability density",
-        fill = "Language",
-        shape = "Language",
-        colour = "Language"
-    ) +
+    distinct(te, .keep_all = TRUE) %>% 
+    ggplot(aes(x = freq_zipf)) +
+    geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
+    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
     theme(
         legend.position = "none",
+        panel.grid.minor = element_blank(),
         legend.title = element_blank()
@@
     facet_wrap(~language, ncol = 1) +
-    geom_col() +
+    geom_col(colour = "white") +
     geom_text(aes(label = n), position = position_nudge(y = 5), colour = "black") +
@@
     scale_x_continuous(breaks = 1:15) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
     theme(
         legend.position = "none",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor = element_blank(),
         axis.ticks.x = element_line(),
@@
     ggplot(aes(x = lv, n)) +
-    geom_col(fill = clrs[1], colour = "white") +
+    geom_col(fill = clrs[4], colour = "white") +
     geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
               position = position_nudge(y = 3), size = 3.5) +
-    labs(x = "Levenshtein distance", y = "Probability density") +
+    labs(x = "Levenshtein distance", y = "Number of words") +
     theme(
@@
     drop_na(freq_zipf) %>% 
-    count(n_phon, freq_zipf, language) %>% 
-    ggplot(aes(n_phon, freq_zipf, colour = language, fill = language)) +
-    facet_wrap(~language) +
-    geom_point(size = 2, alpha = 0.25) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
+    distinct(te, .keep_all = TRUE) %>% 
+    count(n_phon, freq_zipf) %>% 
+    ggplot(aes(n_phon, freq_zipf)) +
+    geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
+               position = position_jitter(width = 0.1), seed = 888) +
+    geom_smooth(method = "lm", formula = "y ~ x" ,
+                colour = clrs[1], fill = clrs[1]) +
     labs(
         x = "# Phonemes", 
-        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
+        y = "Lexical frequency (Zipf score)", 
         colour = "Language", 
@@
 items %>% 
-    left_join(select(bvqdev::pool, item, language)) %>% 
+    distinct(te, .keep_all = TRUE) %>% 
     drop_na(freq_zipf) %>% 
-    count(freq_zipf, lv, language) %>% 
-    ggplot(aes(freq_zipf, lv, colour = language, fill = language, size = n)) +
-    facet_wrap(~language) +
-    geom_point(size = 2, alpha = 0.25) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
+    count(freq_zipf, lv) %>% 
+    ggplot(aes(freq_zipf, lv, size = n)) +
+    geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
+    geom_smooth(method = "lm", formula = "y ~ x", colour = clrs[1], fill = clrs[1]) +
     labs(
         y = "Levenshtein similarity", 
-        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
+        x = "Lexical frequency (Zipf score)", 
         colour = "Language", 
@@
 #| fig-width: 7
-#| fig-cap: "Association betwen number of phonemes and phonological similarity"
+#| fig-cap: "Association between number of phonemes and phonological similarity"
 items %>% 
-    left_join(select(bvqdev::pool, item, language)) %>% 
+    distinct(te, .keep_all = TRUE) %>% 
     drop_na(freq_zipf) %>% 
-    count(n_phon, lv, language) %>% 
-    ggplot(aes(n_phon, lv, colour = language, fill = language, size = n)) +
-    facet_wrap(~language) +
-    geom_point(shape = 1, stroke = 1) +
-    geom_smooth(method = "lm", formula = "y ~ x", size = 1) +
+    count(n_phon, lv) %>% 
+    ggplot(aes(n_phon, lv, size = n)) +
+    geom_point(shape = 1, stroke = 1, colour = clrs[4]) +
+    geom_smooth(method = "lm", formula = "y ~ x", size = 1,
+                colour = clrs[1], fill = clrs[1]) +
     labs(
@@
 We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
-the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion)
+the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).
 
@@
 \textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
-$$
-
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)`:
+$$ {#eq-exposure}
+
+For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):
 
@@
 \end{aligned}
-$$
+$$ {#eq-exposure-example}
 
@@
 
+## DAG
+
+```{r tikz-settings, include=FALSE}
+# tikz stuff
+# Necessary for using dvisvgm on macOS
+# See https://www.andrewheiss.com/blog/2021/08/27/tikz-knitr-html-svg-fun/
+font_opts <- list(dvisvgm.opts = "--font-format=woff")
+```
+
+```{tikz dag, engine.opts=font_opts, echo=FALSE}
+#| fig-cap: "Directed acyclic graph indicating the hypothesised causal relationship between the variables of interest.
+*i* and *j* indicate the variables that are affected by individual differences at the level of participants and translation equivalents, respectively.
+Unobserved variables are encircled, and observed variables are not.
+We hypothesised that the number of phonemes of the word, the age of the participant, the amount of exposure to the word, and the Levenshtein distance all affect participants' familiarity with the word (i.e., the probability of acquisition), and therefore the probability with which each response category is generated.
+We further hypothesised that age influences the effect of word exposure and Levenshtein distance on familiarity, and that word exposure influences the effect of Levenshtein distance on familiarity."
+#| fig-align: center
+#| fig-ext: svg
+#| out-width: 100%
+\usetikzlibrary{positioning}
+\usetikzlibrary{shapes.geometric}
+\begin{tikzpicture}
+% nodes
+\node (age) at (0,0) {$\text{Age}_{i, j}$};
+\node (phonemes) at (3,4) {$\text{Phonemes}_{i, j}$};
+\node (exposure) at (3,2) {$\text{Exposure}_{i, j}$};
+\node (levenshtein) at (3,-2) {$\text{Levenshtein}_{i}$};
+\node [ellipse, draw](familiarity) at (7, 0) {$Familiarity$};
+\node (response) at (11, 0) {$\text{Response}$};
+% arrows
+\draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (familiarity);
+\draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (exposure);
+\draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (levenshtein);
+\draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (familiarity);
+\draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (levenshtein);
+\draw [->, shorten >= 4pt, shorten <= 4pt] (phonemes) -- (familiarity);
+\draw [->, shorten >= 4pt, shorten <= 4pt] (levenshtein) -- (familiarity);
+\draw [->, shorten >= 4pt, shorten <= 4pt] (familiarity) -- (response);
+
+\end{tikzpicture}
+```
+
 ## Data analysis
 
-### Model comparison
-
-We extracted four variables of interest for each observation (describe in previous sections):
-
-* Word lexical frequency ($Frequency$)
-* Word length ($Phonemes$)
-* Phonological similarity ($Levenshtein$)
-* Participant age ($Age$)
-* Word exposure index ($Exposure$)
-
-We fitted a series of nested Bayesian ordinal multilevel models to sequentially test the effects of the independent variables (number of phonemes, exposure, age, and Levenshtein) over the three levels of word learning (no, production, comprehension).
-In all these models participants’ and items’ individual differences were controlled by estimating different parameters for each one (different intercepts and/or slopes).
-This is similar to the control conducted by ANCOVAs.
-Thus, independent variable effects account for collective differences in word learning (unexplained by participants’ or items’ individual differences).
-
-We initially fitted a base model (`model_fit_0`) that only included the predictors $Age$ (`age_std`), $Phonemes$ (`n_phon_std`), along with random intercepts and slopes of all predictors by $Participant$ (`id`) and $Item$ (`te`), and all correlation parameters.
-We then expanded this model by adding the main effect of $Exposure$ (`exposure`) (`model_fit_1`), and then $Levenshtein$ (`lv_std`) (`model_fit_2`) and their corresponding random slopes by $Participant$.
-Then, we added the $Exposure \times Levenshtein$ (`exposure_std:lv_std`) two-way interaction (`model_fit_3`) and its random slope by $Participant$.
-Finally, we added the $Age \times Exposure \times Levenshtein$ (`age_std:exposure_std:lv_std`) three-way interaction with its random slope by $Participant$, along with a random slope for the $Age \times Exposure$ (`age_std:exposure_std`) two-way interaction.
-
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package.
-LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
-
-### Model equation
-
-::: column-body-outset
-$$\begin{aligned}
-g[\mathrm{Pr}(y_{ij} \le a_s)] = \tau_s - \nu_{ij}
-\end{aligned}$$
-:::
-
-$s$ is the s-th response category, $a_1 < a_2 < a_3$ are ordered response categories, $\mathrm{Pr}(y_i \le a_3) = 1$ and $\tau_s$ are threshold parameters where $\tau_1 < \tau_2 < \tau_3$.
-
-::: column-body-outset
-
-$$\begin{aligned}
-\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2\mathrm{Phon}_{j} + \beta_3\mathrm{Exposure}_{ij} + \beta_4 \mathrm{Levenshtein}_{j} + & [\mbox{Main effects}]\\
-&\beta_5\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_6\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_7\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Two-way interactions}]\\
-&\beta_8{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Three-way interaction}]\\
-
-&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} + \theta_{2j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{4j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) + & [\mbox{Item-level effects}]\\
-
-&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +   \theta_{2j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{5j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) & [\mbox{Participant-level effects}]
-\end{aligned}$$
-:::
-
-### Model prior
-
-::: column-body-outset
-$$\begin{aligned}\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
-\beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
-\beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
-\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]\end{aligned}$$
-:::
-
-In addition, we specified a $LKJ$ prior--$\sim LKJ(2)$ for the correlations between group-level coefficients.
+We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function.
+The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood (see @fig-likelihood for a graphical representation).
+
+$$
+\begin{aligned}
+\textbf{Likelihood:} \\
+y_{ij} &\sim \text{Cumulative}(p_{k})
+\end{aligned}
+$$ {#eq-likelihood}
+
+where:
+
+- $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$)
+- $i$ is the participant index
+- $j$ is the translation equivalent (TE) index
+- $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$)) between two response categories in the latent distribution
+$p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.
+
+```{r}
+#| label: fig-likelihood
+#| fig-cap: "Graphical representation of the assumed latent distribution that generates observed responses in the questionnaire.
+Cumulative ordinal regression assumes that observed categorical responses (*No*, *Understands* and *Understands and Says* are generated from a common latent (i.e., unobserved) distribution.
+For threotical reasons, we refer to the dimension captured by such distribution as *Familiarity*.
+*Familiarity* describes a given participant's familiarity with a given word.
+When familiarity is low (below some threshold 1), the reulting response is *No*.
+When familiarity is higher than such threshold 1, but lower than some threshold 2, the resulting result is *Understands* (i.e., comprehension).
+When familiarity exceeds threshold 2, the resulting response is *Understands and Says* (i.e., comprehension and production).
+These thresholds are estimated in the cumulative regression model as intercepts 1, and 2.
+The rest of the coefficients in the regression model shift the response thresholds, making comprehension or production more likely or less likely.
+See @burkner2019ordinal for a detailed description of cumulative ordinal regression and its implementation."
+#| fig-width: 7
+#| fig-height: 3.5
+dist_df = tribble(~mean, ~sd, 0,   1)
+
+responses <- data.frame(
+    ymin = c(-Inf, -1.5, 1.5),
+    ymax = c(-1.5, 1.5, Inf),
+    xmin = -Inf,
+    xmax = Inf,
+    .category = str_wrap(c("No", "Understands", "Understands and Says"), width = 12)
+)
+
+responses %>%
+    ggplot(aes(ymin = ymin, ymax = ymax, xmin = xmin, xmax = xmax, fill = .category)) +
+    geom_rect(alpha = 0.5) +
+    stat_dist_slab(
+        data = dist_df, 
+        aes(dist = "norm", arg1 = mean, arg2 = sd),
+        position = "dodge", inherit.aes = FALSE,
+        fill = NA, colour = "black"
+    ) +
+    geom_hline(yintercept = c(-1.5, 1.5), linetype = "dotted", size = 1) +
+    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
+              vjust = 1, hjust = 0.5) +
+    annotate(
+        geom = "text",
+        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
+        y = c(-1.5, 2), parse = TRUE,
+        x = 0.5, angle = 90, vjust = -0.5
+    ) +
+    labs(y = "Hypothetical response-generating latent distribution",
+         x = "Response probability",
+         fill = "Response type",
+         colour = "Response type") +
+    coord_flip() +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+    scale_x_continuous(labels = percent) +
+    scale_y_continuous(limits = c(-4, 4)) +
+    theme(panel.grid = element_blank(),
+          legend.position = "none")
+```
+
+To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
+We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset--each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
+For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].
+
+$$
+\begin{aligned}
+\textbf{Linear model:} \\
+logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
+& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
+& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Phonemes}_{ij} + & \\
+& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
+& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
+& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
+& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
+& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
+\end{aligned}
+$$ {#eq-linear}
+
+where:
+
+- $i$ and $j$ index the participant and translation equivalent (TE)
+- $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$
+- $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively
+- $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest
+- $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively
+
+We used the Bayesian framework to estimate the parameters in our model.
+This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical].
+This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations.
+We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].
+
+$$
+\begin{aligned}
+\\
+\textbf{Prior:} \\
+\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
+\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
+\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
+u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
+w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
+&&\mbox{[Participant-level coefficient variability]} \\ \\
+\Bigg(\begin{smallmatrix}
+u_{k_{0}} \\ 
+u_{1_{i}} \\ 
+\vdots \\ 
+u_{8_{i}} 
+\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
+\Bigg(\Bigg(\begin{smallmatrix}0 \\
+0 \\ 
+\vdots \\
+0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
+\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
+\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
+\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
+\vdots & \vdots & \vdots & \vdots \\
+\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
+\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
+\rho_{u} &\sim LKJcorr(2) \\
+\\
+&&\mbox{[TE-level coefficient variability]} \\ \\
+\Bigg(\begin{smallmatrix}
+w_{k_{0}}\\ 
+w_{1_{j}} \\ 
+\vdots \\ 
+w_{3_{j}} 
+\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
+0\\ 
+0 \\ 
+\vdots \\
+0 
+\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
+\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
+\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
+\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
+\vdots & \vdots & \vdots & \vdots \\
+\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
+\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
+\rho_{w_{0-3}} &\sim LKJcorr(2)
+\end{aligned}
+$$ {#eq-prior}
+
+where:
+
+- $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively
+- $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively
+- $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].
 
@@
     ) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
     scale_y_continuous(labels = percent) +
@@
 
+### Model comparison
+
+To test of goodness of fit of our model, we compared its predictive accuracy of this extended model--which included all predictors of interest--to four simpler models that included a subset of our predictors of interest: a base model that only included the main effects of $Age$, $Phonemes$, and $Exposure$ ($\mathcal{M}_{0}$), a model that also included the $Age \times Exposure$ interaction ($\mathcal{M}_{1}$), a model that also included the main effect of $Levenshtein$ ($\mathcal{M}_{2}$), a model that also included the $Age \times Levenshtein$ interaction ($\mathcal{M}_{3}$).
+Only the extended model included the $Age \times Exposure \times Levenshtein$ interaction.
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package.
+LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
+Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
+
 ### Model settings
@@
 -   R version: `r sessionInfo()$R.version$version.string`
--   Algorithm: Hamiltonian Montecarlo (No U-turn Sampler, *NUTS*)
--   Engine: Stan (brms interface, `r sessionInfo()$otherPkgs$brms$Version`)
--   brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
+-   Algorithm: Hamiltonian Monte Carlo (No U-turn Sampler)
+-   Engine: Stan (`brms` interface, `r sessionInfo()$otherPkgs$brms$Version`)
+-   `brms` backend: CmdStanR (`cmdstanr` `r sessionInfo()$loadedOnly$cmdstanr$Version`)
 -   Chains: `r dim(model_fit_4$fit)[2]`
@@
     labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
     scale_y_continuous(labels = percent) +
@@
 
-### Leave-one-out diagnostics (Pareto-*k* diagnostics)
-
-::: panel-tabset
-
-#### Model 0
-
-```{r}
-#| label: fig-loo-diagnostics-0
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
-get_loo_plot <- function(x){
-    x$pointwise[,5] %>% 
-        data.frame(x = 1:length(.), y = .) %>% 
-        ggplot(aes(x, y)) +
-        geom_point(
-            size = 0.5, alpha = 0.1, colour = clrs[5], 
-            shape = 1
-        ) +
-        geom_hline(yintercept = 0.5, linetype = "dotted") +
-        labs(
-            x = "Data point",
-            y = "Pareto-k value",
-            title = "PSIS diagnostic plot"
-        ) +
-        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
-        scale_y_continuous(
-            limits = c(-0.6, 0.6)
-        )
-}
-
-get_loo_plot(model_loos$model_fit_0)
-```
-
-#### Model 1
-
-```{r}
-#| label: fig-loo-diagnostics-1
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
-get_loo_plot(model_loos$model_fit_1)
-```
-
-#### Model 2
-
-```{r}
-#| label: fig-loo-diagnostics-2
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
-get_loo_plot(model_loos$model_fit_2)
-```
-
-#### Model 3
-
-```{r}
-#| label: fig-loo-diagnostics-3
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
-get_loo_plot(model_loos$model_fit_3)
-```
-
-#### Model 4
-
-```{r}
-#| label: fig-loo-diagnostics-4
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
-get_loo_plot(model_loos$model_fit_4)
-```
-
-:::
-
 ## Fixed effects
@@
 
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
 This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
@@
     ) +
+    geom_point(data = post_summary, aes(x = .value), size = 1.5) +
     geom_text(
@@
     newdata = nd,
-    ndraws = NULL, 
+    ndraws = 25, 
     re_formula = NA
@@
     facet_grid(group~exposure_std) +
-    stat_summary(fun.data = median_hdi, geom = "ribbon",
-                 alpha = 0.5, colour = NA) +
-    # geom_line(aes(group = interaction(lv_std, drawid)), 
-    #           size = 0.75, alpha = 0.15) +
+    # stat_summary(fun.data = median_hdi, geom = "ribbon",
+    #              alpha = 0.5, colour = NA) +
+    geom_line(aes(group = interaction(lv_std, drawid)),
+              size = 0.75, alpha = 0.15) +
     stat_summary(aes(linetype = lv_std), fun = median, geom = "line", size = 1) +
@@
     ) +
-    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
+    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
     scale_y_continuous(
@@
     re_formula = NA,
-    ndraws = NULL,
+    ndraws = 50,
     vcov = FALSE,
@@
     geom_hline(yintercept = 0, size = 1, colour = "grey") +
-    # geom_line(aes(group = drawid), alpha = 0.25, size = 0.75, colour = clrs[1]) +
-    stat_summary(fun.data = median_hdi, geom = "ribbon",
-                 colour = NA, alpha = 0.5, fill = clrs[1]) +
+    geom_line(aes(group = drawid), alpha = 0.25, size = 0.75, colour = clrs[4]) +
+    # stat_summary(fun.data = median_hdi, geom = "ribbon",
+    #              colour = NA, alpha = 0.5, fill = clrs[1]) +
     stat_summary(fun = median, geom = "line", size = 1, color = clrs[1]) +
@@
     ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
     scale_x_continuous(limits = c(0, 1), labels = percent) +
@@
     ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
     theme(
@@
     ) +
-    scale_colour_d3() +
-    scale_fill_d3() +
+    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
     scale_x_continuous(labels = label_number(big.mark = ","), 
@@
 
+### Leave-one-out diagnostics (Pareto-*k* diagnostics)
+
+::: panel-tabset
+
+#### Model 0
+
+```{r}
+#| label: fig-loo-diagnostics-0
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
+get_loo_plot <- function(x){
+    x$pointwise[,5] %>% 
+        data.frame(x = 1:length(.), y = .) %>% 
+        ggplot(aes(x, y)) +
+        geom_point(
+            size = 0.5, alpha = 0.1, colour = clrs[4], 
+            shape = 1
+        ) +
+        geom_hline(yintercept = 0.5, linetype = "dotted") +
+        labs(
+            x = "Data point",
+            y = "Pareto-k value",
+            title = "PSIS diagnostic plot"
+        ) +
+        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
+        scale_y_continuous(
+            limits = c(-0.6, 0.6)
+        )
+}
+
+get_loo_plot(model_loos$model_fit_0)
+```
+
+#### Model 1
+
+```{r}
+#| label: fig-loo-diagnostics-1
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
+get_loo_plot(model_loos$model_fit_1)
+```
+
+#### Model 2
+
+```{r}
+#| label: fig-loo-diagnostics-2
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
+get_loo_plot(model_loos$model_fit_2)
+```
+
+#### Model 3
+
+```{r}
+#| label: fig-loo-diagnostics-3
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
+get_loo_plot(model_loos$model_fit_3)
+```
+
+#### Model 4
+
+```{r}
+#| label: fig-loo-diagnostics-4
+#| echo: true
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
+get_loo_plot(model_loos$model_fit_4)
+```
+
+:::
+
 ### Gelman-Rubin diagnostic (R-hat)

### 2023-01-17 manuscript/manuscript.qmd
Add figures to manuscript, start discussion, add model and predictor appendix

---
title: "The role of cross-linguistic lexical similarity on bilingual word acquisition"
editor: source
author:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    affiliations:
      - ref: upf
    corresponding: true
  - name: Daniela S. Ávila-Varela
    affiliations:
      - ref: upf
  - name: Ignacio Castillejo
    affiliations:
      - ref: uam
  - name: Núria Sebastian-Galles
    affiliations:
      - ref: upf
affiliations:
  - id: upf
    name: Center for Brain and Cognition, Universitat Pompeu Fabra
  - id: uam
    name: Departamento de Psicología, Universidad Autónoma de Madrid

pagetitle: Cross-linguistic similarity and  word acquisition

thanks: "This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award]. Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019). The authors declare no conflicts of interest with regard to the funding source of this study. We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and Ege E. Özer for their help-ful feedback. We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support. We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants. We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises. We also thank all families and infants who participated in the experiments. Data collection was half-way when the COVID-19 pandemic started. We would like to pay special tribute to the families that collaborated with us under these difficult circumstances."
keywords: lexical acquisition, vocabulary, bilingualism
toc: false
fig-dpi: 1000
csl: "apa7.csl"
bibliography: "references.bib"
format:
  pdf:
    pdf-engine: lualatex
    linestretch: 1.5
    number-sections: true
    whitespace: small
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load packages
library(conflicted)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(purrr)
library(gt)
library(forcats)
library(stringr)
library(tibble)
library(tidybayes)
library(janitor)
library(brms)
library(papaja)
library(scales)
library(stringdist)
library(marginaleffects)
library(lubridate)
library(bvqdev)

# load and process objects
# load targets as R objects
tar_load(participants)
tar_load(df)
tar_load(items)
tar_load(bvq_data)
tar_load(model_fit_4)
tar_load(model_log_liks)
tar_load(model_fit_4_prior)
tar_load(model_loos)
tar_load(rope_interval)
tar_load(posterior_description)

participants <- left_join(participants, select(bvq_data$logs, id, time, edu_parent))
items <- left_join(items, select(bvq_data$pool, item, language, semantic_category, class))

# set ggplot theme and colour palette
theme_set(theme_custom()) # set custom ggplot theme
clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
        ggplot2.continuous.colour = ggplot2::scale_color_gradient)

# resolve namespace conflicts
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

# Introduction {#sec-introduction}

## Word acquisition during toddlerhood: the effect of language experience

The foundations of word learning are in place early in age. Before the end of their first year of life infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthurbates]. During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension]. These early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e. vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of acquired words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, @bates1994developmental]. Despite this variability, children's trajectories of early vocabulary growth seem quite stable across languages. @tardif2008baby collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing. Since birth, these infants had been learning English, Cantonese and Mandarin, respectively. The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*). These results were later extended by @frank2021variability to a more diverse set of languages, and found that such cross-language commonalities are stronger at earlier stages of lexical acquisition.

To date, most studies on early word acquisition have relied exclusively on data from monolingual children while neglecting, to a considerable extent, that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent]. Previous findings pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the two languages are taken into account. @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English. Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development. Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual]. While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals. What role could linguistic distance play during early vocabulary growth?

## Language distance as a mediator of bilingual vocabulary growth

For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language. Depending on the linguistic distance between both languages, the two sets of words may overlap in form in varying degrees. When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic, respectively). For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates). It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*). If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.

@floccia2018introduction provided evidence in line with this claim. The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language. The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of *cognateness* between each pair of languages. Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions. The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other [@levenshtein1966binary]. The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes. Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair. For example, English-Dutch bilinguals (22.14% cognateness), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% cognateness) were able to produce in Mandarin.

## Lexical non-selectiviy as a candidate account: the role of parallel activation

Floccia et al. pointed to *parallel activation* as a phenomenon that might potentially underpin their results. The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension, and that this phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain]. One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate. In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish. In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trials labels were non-cognates (*mesa*-*taula*, translations of *table*). Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. Importantly, Spanish monolinguals--unfamiliar with the Catalan translations of the Spanish words they uttered--did not show this effect. These results suggest that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures. Several subsequent studies have also provided similar evidence in comprehension in children [e.g., @vonholzen2012language; @poulin-dubois2013lexical]. Parallel activation is therefore a plausible explanation for Floccia et al.'s results: cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes. This account is is line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words, already acquired or not [e.g., @hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children]. Importantly, bilinguals seem to be more likely to acquire words for which their translation equivalent has been already learnt in the other language, suggesting that semantic similarity also facilitates word acquisition across languages [@bilson2015semantic].

If phonological similarity also plays a cross-language facilitation role during word acquisition, cognate translation equivalents--which share both semantic and phonological similarity--should be acquired, on average, earlier that non-cognate translation equivalents--which share semantic similarity, but not phonological similarity. Evidence supporting and earlier age of acquisition for cognates is, to date, scarce. @bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates. @schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average. The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth. On the other hand, Floccia et al.'s estimates are statically more reliable given their (remarkably larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly. In their discussion the authors state the following (pp. 70):

> "This  finding  also  provides  support  to  the  proposal  that  the  cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life."

We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results. First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists. By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs. non-cognates was no longer available. Second, all participants were aged ~24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.

More recently, @mitchell2022cognates addressed this issue overcoming some of these pitfalls. Using a larger, longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages. They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates. The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty). These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar: word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.

## Two competing mechanisms for a cognate facilitation during word acquisition

We argue that the facilitative effect of cognateness on word acquisition can be explained by at least two mechanisms, illustrated in @fig-hypotheses. On the one hand, and in line with @mitchell2022cognates, the effect of cognateness might take place during early stages of word acquisition, before stable representations of any of the two word-forms of the translation equivalent has been formed. In this scenario, the auditory recognition any of the two word-forms might lead to the activation of its (cognate) translation in the other language via their phonological similarity. Since that both word-forms are likely to be heard in the presence of the same referent, the association between one of the word-forms and its referent could be strengthened after the child is exposed to that same word-form, or to its (similar-sounding) phonologically translation.

For instance, in the presence of a tree, an English-Spanish bilingual child might hear *flower* in some occasions, and *flor* in some others. Given the phonological similarity between *flower* and *flor*, it is possible that both word-forms are activated in the child's lexicon every time any of them is encountered. This would lead to an increase in the association strength between each word form and their common referent every time one of them is encountered, leading to an earlier age of acquisition. The case of a non-cognate translation equivalent, such as *cow* and *vaca* would be different. Although both word-forms are also encountered in the presence of the same referent, their low phonological similarity would not guarantee such facilitation when the child hears either of them, thus leading to a later age of acquisition for both. This account departs from the premise that a word-form can potentially activate its translation equivalent via phonology before acquisition, that is, before the association between such word-form and its referent is consolidated. This would allow both word-forms to mutually boost their acquisition after only few exposures. In this scenario, the effect of cognateness on acquisition should taken place early in age, and should be symmetric for both languages: the facilitation should occur in both word-forms simultaneously.

An alternative account is that cognateness plays a role in the acquisition of a translation pair only when at least one of the word-forms has been acquired, that is, when then association between one of the word-forms and their common referent is consolidated. In this scenario, one of the word-forms can only activate the other via phonological similarity after its representation in the child's lexicon is robust enough. This way, *flower* and *flor* would not benefit from their phonological similarity until wither *flower* or *flor* is acquired. Only after, for instance, *flower* is acquired, the acquisition of *flor* would be boosted by its similarity to *flower*. This account predicts that cognateness has an asymmetric effect on cognate word acquisition: it should be observed only on the acquisition of the word-form of the translation equivalent that was acquired last. As a consequence, and given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.

![Schematic representation of the hypothesised cognate facilitation effect on word acquisition. To the left, the acquisition of a cognate translation equivalent (*flower*-*flor*) is affected by the phonological similarity between both word forms. When the child hears *flower*, activation spreads across languages to activate *flor* too. As a result, after exposure to *flower* not only the association between *flower* and its referent is strenghtened, but also that of *flower* and its common referent. To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*) does not undergo the such facilitation. Both word-forms are fairly disimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](hypotheses.jpg){#fig-hypotheses}

In line with this second hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al. reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure). Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language. Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.  Under the second hypothesis, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.

## The present study

In this study, we investigated the role of cognateness on the acquisition of translation equivalents, and if such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair. Using an online vocabulary checklist--designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language. We adopted an Bayesian item response theory (*IRT*) approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to the word-form's cognate status, and the participant's frequency of exposure to the word-form (a composite measure of lexical frequency and language degree of exposure), while adjusting for other indices of word difficulty---lexical frequency [@goodman2008does] and length (phonemes)---and participant skill---age [see @kachergis2022standard for a similar approach].

# Methods {#sec-methods}

All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/hy984/)) and GitHub ([https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories)) repositories. This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.

## Questionnaire {#sec-questionnaire}

```{r}
#| label: items
#| echo: false
#| message: false
#| warning: false
#| include: false
n_categories <- nrow(distinct(bvq_data$pool, semantic_category))

n_te <- length(unique(bvq_data$pool$te))

n_items <- length(unique(bvq_data$pool$item))

n_item_language <- count(bvq_data$pool, language) %>%
    pull(n) %>%
    set_names(c("catalan", "spanish"))
```

To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire [@garcia-castro2022multilex]. This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formR platform [@arslan2020formr]. This questionnaire is structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists. Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish. The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]`. Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping. In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*). We included items from a diverse sample of `r n_categories` semantic/functional categories.

For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default). Some families filled a long version of the vocabulary checklists (`r format(n_te, big.mark = ",")` translation equivalents; `r format(n_item_language["catalan"], big.mark = ",")` items in Catalan, `r format(n_item_language["spanish"], big.mark = ",")` items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish). These last families were randomly allocated into one of four different subsets of the complete list of items. These lists were designed so that each contained a representative sub-sample of the items from the complete list. Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists. Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions. For the analyses in the present study, we considered responses to words corresponding to nouns, verbs, and adjectives [@fourtassi2020growth]. We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono/mico*).

## Participants {#sec-participants}

```{r}
#| label: participants
#| echo: false
#| message: false
#| warning: false
#| include: false
# load packages
n_participants <- df %>%
    distinct(id, age) %>%
    count(id) %>%
    pull(n) %>%
    table()
n_responses <- nrow(distinct(df, id, age))
dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
```

We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`. `r n_participants[1]` of them participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times. Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(participants$age))` and `r floor(max(participants$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants-lp for a summary of the distribution of participants across ages and degrees of exposure to Catalan). All families gave informed consent before participating. Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate. Filling the questionnaire took 30 minutes approximately. After completion, families were rewarded with a small present.

```{r}
#| label: tbl-participants-lp
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about DoE was provided by families before filling the questionnaire. A 100% indicated that the participant was exclusively exposed to Catalan, and never to Spanish. A 0% indicates that the participant was not exposed to Catalan ever, and rather most of the time to Spanish. A 50% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100%, 0%, and 50% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants %>%
    filter(id %in% df$id) %>% 
    mutate_at(vars(starts_with("doe_")), function(x) {
        floor(x * 10) / 10
    }) %>%
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%", "25-50%", "50-75%", "75-100%"),
            include.lowest = TRUE
        ) %>%
            as.character()
    ) %>%
    count(age, doe_catalan) %>%
    arrange(desc(doe_catalan)) %>%
    pivot_wider(names_from = age, values_from = n) %>%
    gt() %>%
    summary_rows(
        columns = 2:7,
        fns = list("Total" = ~ sum(., na.rm = TRUE)),
        formatter = fmt_integer
    ) %>%
    sub_missing(everything(), everything(), missing_text = "--") %>%
    cols_label(doe_catalan = "DoE Catalan") %>%
    tab_spanner("Age (months)", 2:8) %>%
    tab_footnote(
        "This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10% of the time",
        cells_column_labels(columns = doe_catalan)
    ) %>%
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

We used the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain. Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))` and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).

## Variables of interest

Each observation in our data was matches with its corresponding values in four variables of interest that we used to model participants' responses:

(1) Participant age ($Age$): number of months elapsed between participants' birth date and questionnaire completion.

(2) Word length ($Phonemes$): number of phonemes in the phonological transcription of the word-form in Speech Assessment Methods Phonetic Alphabet (SAMPA) format.

(3) Word exposure index ($Exposure$): this is a composite measure that captures the frequency with which the participant is exposed to a given word from a given language (Catalan or Spanish), weighted by the child's amount of exposure to that language. This measure was calculated by multiplying the word-form's lexical frequency (extracted from the CHILDES corpora, @macwhinney2000childes; @sanchez2019childes) with the participant's degree of exposure to the language the word belongs to (see @sec-appendix-predictors for more details).

(4) Phonological similarity ($Levenshtein$): normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary]. This score is computed by counting the number of insertions, deletions and replacements needed both transcriptions to become identical, then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one. This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches] (see @sec-appendix-predictors for more details).

## Data analysis {#sec-analysis}

```{r}
#| label: sample-sizes
#| echo: false
#| message: false
#| warning: false
#| include: false
n_obs <- nrow(df)
sum_obs <- df %>%
    mutate(te = factor(te)) %>%
    count(te) %>%
    summarise(across(n, lst(median, min, max), .names = "{.fn}"))

n_items <- ifelse(grepl("cat_", distinct(df, item)$item),
                  "Catalan",
                  "Spanish") %>%
    table()

n_te <- length(unique(df$te))

rope_interval <- c(lower = -0.1, upper = 0.1)

r_info <- sessionInfo()
r_version <- paste(r_info$R.version$major, r_info$R.version$minor, sep = ".")
```

Our dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one of `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items["Catalan"]` in Catalan, `r n_items["Spanish"]` in Spanish, making up `r n_te` translation equivalents in total. Translation equivalents (TEs) received a median of `r printnum(sum_obs$median, digits = 0)` responses (*Min* = `r printnum(sum_obs$min, digits = 0)`, *Max* = `r printnum(sum_obs$max, digits = 0)`), both languages summed together.

We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_4$fit)[2]` sampling chains with `r format(dim(model_fit_4$fit)[1], big.mark = ",")` iterations each). This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$, $Phonemes$, $Exposure$, and $Levenshtein$. All predictors were standardised before entering the model for an easier interpretation of their regression coefficients. We also included participant and item random intercepts and slopes when appropriate [@barr2013random]. We compared a baseline model $\mathcal{M}_{0}$, which only included the intercept, $Age$, $Phonemes$, and $Exposure$ as predictors, with increasingly more complex models, introducing predictors in the following order: $Age \times Exposure$ ($\mathcal{M}_{1}$), $Levenshtein$ ($\mathcal{M}_{2}$), $Exposure \times Levenshtein$ ($\mathcal{M}_{3}$), and finally $Age \times Levenshtein$ and $Age \times Exposure \times Levenshtein$ ($\mathcal{M}_{4}$). We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO, @vehtari2017practical]^[Due to computational constraints, we performed LOO based on a random sub-sample of `r format(dim(model_log_liks$model_fit_4)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]], and selected for interpretation the one showing the best performance. More details about the models can be found in @sec-appendix-model.

We assessed the practical relevance of the estimated regression coefficients of the selected model following @kruschke2018bayesian. First, we specified a region of practical equivalence (*ROPE*) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`. This region indicates the range of values that we considered equivalent to zero. The degree of overlap between the posterior distribution of a regression coefficient with the *ROPE* can be interpreted as support for the true parameter of the coefficient being zero or equivalent. Then we computed the 95% highest density interval (*HDI*) of the posterior distribution of each regression coefficient. This is interval contains the true value of this coefficient with 95% probability, given the data. Finally, we calculated the proportion of posterior samples in the 95% *HDI* that fall into the *ROPE*, which indicates the probability that the true value of the regression coefficient falls into the *ROPE*, and should therefore be considered equivalent to zero. For example, a 90% overlap between the *HDI* of the coefficients’ posterior distribution and the *ROPE* indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero.

Data processing and visualisation was done in R [@rcoreteam2013language, `r r_version`] using the Tidyverse family of packages [@wickham2019welcome]. Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@vehtari2022loo], `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.

# Results {#sec-results}

```{r}
#| label: results-values
#| echo: false
loo_diff <- loo_compare(model_loos)[2, 1]
loo_diff_se <- loo_compare(model_loos)[2, 2]

post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)

# tidy predictor names
str_repl <- c(
    "b_Intercept[1]" = paste0("Comprehension and Production"),
    "b_Intercept[2]" = paste0("Comprehension"),
    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
    "b_exposure_std" = paste0("Exposure (+1 SD, ",  round(sd(df$exposure), 2), ")"),
    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
)

post_draws_list <- split(posterior_description, posterior_description$parameter)
```

The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the second best-performing model ($\mathcal{M}_{3}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`. Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see @sec-appendix-loo for all model comparisons). @tbl-results-fixed shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$. The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor.

```{r}
#| label: tbl-results-fixed
#| tbl-cap: "Posterior distribution of regression coefficients. *Median*: median of the posterior distribution in the logit scale. *95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE, [-0.1, +0.1]). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
posterior_description %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(.variable_name = factor(parameter,
                              levels = names(str_repl),
                              labels = str_repl) %>%
               as.character(),
           type = ifelse(str_detect(parameter, "Intercept"), 
                         paste0("Intercepts (at ", round(mean(df$age, 2)), " months)"),
                         "Slopes"),
           parameter = ifelse(str_detect(parameter, "Intercept"), 
                              str_remove_all(parameter, "Intercept \\(|\\)"),
                              parameter)) %>%
    select(.variable_name, median, ci_high, ci_low, rope_percentage, type) %>%
    gt(groupname_col = "type", rowname_col = "parameter") %>%
    fmt_number(2:4) %>%
    fmt_percent(5) %>%
    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
    cols_label(
        .variable_name = "Parameter",
        median = "Median",
        ci_low = md("95% *HDI*"),
        rope_percentage = md("*ROPE* prob.")
    ) %>%
    tab_style(cell_text(weight = "bold"),
              cells_column_labels(columns = 1:5)) %>%
    tab_style(cell_text(align = "left"),
              cells_title(groups = "subtitle")) %>% 
    tab_style(cell_text(size = "small"), 
              list(cells_body(), cells_row_groups(), cells_stub()))
```

```{r}
#| label: fig-results-fixed
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 4
#| eval: false
#| fig-cap: "Posterior distribution of the regression coefficients of the fixed effects in the extended model. Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis). Intervals show the 95% highest density interval (HDI) of each distribution. This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data. The mean and HDI limits are indicated below each distribution. On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE, [-1, +1]). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
post <- post_draws %>%
    mutate(.variable_name = factor(
        .variable,
        levels = names(str_repl),
        labels = str_repl,
        ordered = TRUE
    ))

post_summary <- post %>%
    filter(!grepl("Intercept|sd", .variable)) %>%
    median_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>%
    mutate(.variable_name = factor(
        .variable,
        levels = names(str_repl),
        labels = str_repl,
        ordered = TRUE
    ))

post_pd <- posterior_description %>% 
    mutate(.variable_name = factor(
        parameter,
        levels = names(str_repl),
        labels = str_repl,
        ordered = TRUE
    )) %>% 
    filter(!grepl("Intercept|sd", parameter))

post %>%
    filter(!grepl("Intercept|sd", .variable)) %>%
    ggplot(aes(.value, fct_rev(.variable_name))) +
    annotate(
        geom = "rect",
        ymin = -Inf,
        ymax = Inf,
        xmin = rope_interval["lower"],
        xmax = rope_interval["upper"],
        colour = NA,
        alpha = 0.2,
        fill = clrs[5],
    ) +
    geom_vline(xintercept = 0,
               size = 1,
               colour = clrs[5]) +
    stat_slab(
        aes(
            fill = stat(abs(x) < rope_interval["upper"]),
            colour = stat(abs(x) < rope_interval["upper"]),
        ),
        size = 1,
        position = position_nudge(y = 0.15),
        scale = 0.60
    ) +
    geom_errorbar(data = post_summary,
                  aes(xmin = .lower, xmax = .upper, x = .value),
                  width = 0.15) +
    geom_point(data = post_summary, aes(x = .value), size = 1.5) +
    geom_text(
        data = post_summary,
        aes(x = .value,
            y = fct_rev(.variable_name),
            label = paste0(
                round(.value, 1),
                " [",
                round(.lower, 1),
                ", ",
                round(.upper, 1),
                "]"
            )),
        colour = "black", size = 3,
        position = position_nudge(y = -0.3),
    ) +
    geom_label(
        data = post_pd,
        aes(x = -0.75,
            y = .variable_name,
            label = percent(rope_percentage)),
        hjust = 0, colour = "black", size = 3, alpha = 0.5,
        fill = clrs[4], label.r = unit(0, "lines"), label.size = 0
    ) +
    labs(
        x = "Logistic regression coefficient estimate (Logit scale)",
        y = "Variable",
        fill = "Overlaps with ROPE",
        colour = "Overlaps with ROPE"
    ) +
    scale_x_continuous(breaks = seq(-2, 2, 0.2)) +
    scale_fill_manual(values = clrs[c(1, 4)],
                      labels = c("No", "Yes")) +
    scale_colour_manual(values = clrs[c(1, 4)],
                        labels = c("No", "Yes")) +
    theme(
        legend.position = "top",
        legend.justification = c(1, 1),
        axis.title.y = element_blank(),
        panel.grid.major.x = element_line(colour = "grey85",
                                          linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```
The word exposure index ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]). All of the posterior samples of this regression coefficient excluded the *ROPE*. The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/sd(df$exposure), accuracy = 0.01)` more likely to acquire it.

The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.

However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to word. Follow-up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to word was low (e.g., -1 SD), phonological similarity increased the probability of acquisition substantially. This effect was negligible when for words with median or high exposure (+1 SD).

The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*. This shows that the effect of the word exposure index on the speed at which participants acquired words across ages was inconclusive: given our data, it is difficult to tell whether $Exposure$ increased or decreased the slope of the words' acquisition trajectories.

The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.

Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of the word exposure index on the effect of phonological similarity did not differ substantially across ages.

```{r}
#| label: fig-results-marginal
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 5
#| fig-cap: "Expected posterior predictions. Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient. These values correspond to samples of the posterior distribution that were drawn during the MCMC estimation of the model. We generated 25 posterior predictions for a series of combination of levels of interest. Each thin line and corresponds to one prediction for one set of predictor values. Thicker lines indicate the median posterior prediction. The X-axis indicates the age (in months) for which the prediction is generated. The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*). Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents). Finally, predictions are presented separately for different degrees of word exposure index: little exposure to the word (-1 SD), median exposure, and high exposure (+1 SD). Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
nd <- datagrid(
    model = model_fit_4,
    age_std = seq_range(model_fit_4$data$age_std, n = 100),
    lv_std = (seq(0, 1, 0.5)-mean(df$lv))/(sd(df$lv)),
    exposure_std = c(-1, 0, 1),
    te = NA,
    id = NA
)

marg_epreds <- predictions(
    model_fit_4,
    newdata = nd,
    ndraws = 25, 
    re_formula = NA
)

marg_epreds %>% 
    posteriordraws() %>% 
    as_tibble() %>% 
    select(drawid, group, draw, age_std, lv_std, exposure_std) %>%
    filter(group != "No") %>% 
    pivot_wider(
        id_cols = c(drawid, age_std, lv_std, exposure_std), 
        names_from = group, 
        values_from = draw
    ) %>% 
    mutate(`Understands` = `Understands and Says` + `Understands`) %>% 
    pivot_longer(c(`Understands`, `Understands and Says`), names_to = "group", values_to = "draw") %>% 
    mutate(
        exposure_std = factor(
            exposure_std, 
            levels = unique(nd$exposure_std), 
            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
        ),
        lv_std = factor(
            lv_std,
            levels = unique(nd$lv_std), 
            labels = paste0(c(0, 50, 100), "% similarity")
        )
    ) %>% 
    ggplot(aes(age_std, draw, colour = lv_std, fill = lv_std)) +
    facet_grid(group~exposure_std) +
    # stat_summary(fun.data = median_hdi, geom = "ribbon",
    #              alpha = 0.5, colour = NA) +
    geom_line(aes(group = interaction(lv_std, drawid)),
              linewidth = 0.75, alpha = 0.15) +
    stat_summary(aes(linetype = lv_std), fun = median, geom = "line", linewidth = 1) +
    labs(
        x = "Age (months)",
        y = "Probability of acquisition",
        linetype = "Phonological similarity (Levenshtein)",
        colour = "Phonological similarity (Levenshtein)",
        fill = "Phonological similarity (Levenshtein)"
    ) +
    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(
        breaks = seq(0, 1, 0.25),
        limits = c(0, 1),
        labels = percent
    ) +
    scale_x_continuous(
        breaks = (seq(10, 32, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 32, 4)
    ) +
    theme(
        legend.position = "top",
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
    )
```

# Discussion {#sec-discussion}

In this study, we built on previous studies to investigate the role of cognateness on word acquisition. We contrasted two hypotheses regarding the stage of the word acquisition process at which the possible (facilitative) effect of cognateness might take place. First, we considered the possibility that cognateness boosts the acquisition of translation equivalents at early stages, before a stable, robust representation of either of the word-forms involved has been formed in the child's lexicon. This account predicts that the effect of cognateness would take place early in age, and would be symmetric across languages: one word-form boosts the acquisition of the other via their phonological similarity. If this hypotheses were supported by data, we would expect to observe and early difference in probability of comprehension and/or production between more similar and less similar translation equivalents (i.e., an effect of $Levenshtein$ in interaction with $Age$), with this difference being equivalent for the word-for to which the child is exposed to more frequently, and for the word-form to which the while is exposed less frequently (no interaction effect between $Exposure$ and $Levenshtein$).

We also considered an alternative hypothesis: that the facilitation effect of cognateness takes place at later stages of word acquisition, when at least one of the word-forms of the translation equivalent hold a detailed, robust representation in the child's lexicon. Only then, the phonological similarity between the acquired word-form and its yet-to-be-acquired translation would boost the acquisition of the latter. This account predicts a larger effect of cognateness for word-forms to which the child is exposed to less frequently, and therefore are likely to be acquired later than their translation. This effect of cognateness would also be expected to take place at later ages. Data backing this account would support a model in which the effect of $Levenshtein$ interacts with $Exposure$, and in which possibly this interaction, or at least the effect of $Levenshtein$ interacts with age.

To test these hypotheses, we used an *ad-hoc* vocabulary checklist to collect parental reports of word acquisition status from children learning Catalan and/or Spanish, who were exposed to varying degrees of exposure to either language. We then used item response theory (*IRT*) to estimate the probability of each word-form being acquired by a child---either as comprehension or production---as a function of the child's $Age$, their $Exposure$ to the word-form (operationalised as a composite measure of lexical frequency and language exposure), and the phonological similarity of the word-form with its translation in the other language ($Levenshtein$). We also added the number of $Phonemes$ in the word-form as a nuance predictor the increase the predictive performance of the model.

As anticipated, words with fewer phonemes or to which the child was exposed to more frequently were more likely to be acquired [e.g., @cattani2014how; @floccia2018iii; @braginsky2019consistency]. We also found a facilitation effect of cognateness, providing converging evidence with previous studies [@floccia2018iii; @mitchell2022cognates; @bosch2014first; @schelletter2002effect]. Critically, this effect was mediated by exposure: only words to which children were exposed to less frequently benefited from their cognate status. These results favour the account that a word-form benefits from its phonological similarity with its translation only after such translation has been acquired first.

We expected to find such effect at earlier rather than later ages, but the size of the interaction effect between $Age$, $Exposure$ and $Levenshtein$ failed to exclude a considerable range of null values.

This study has several limitations. One of them is the fact that the languages involved are typologically very close. Catalan and Spanish are both Romance languages, share a considerable amount of cognates, and very often co-exist in the same social spaces (families, schools, working place, etc.). Although there are no populations of bilinguals with identical sociolinguistic characteristics, and therefore it would be pointless to compare the sample in this study to any standard, the generalisability of our results should be taken with a grain of salt. It is possible, for instance, that the closeness of the two languages, together with the frequent co-occurrence of translation equivalents in the same contexts leads our participants to engage cognateness as a facilitator of word acquisition, but that children learning a pair of languages that share a fewer amount of cognates do not benefit from such mechanism. Another consequence of the large amount of cognates that Catalan-Spanish bilingual children are exposed to is that they learn at early stages of word acquisition that words tend to sound similar in both languages, and therefore be more keen on assuming that similar sounding words are semantically related, while children exposed to a fewer amount of cognates might not make the same assumption, even when the word-forms they hear in the presence of the same referents are very similar at the phonological level.

## Limitations

* These results come from a particular population of bilingual toddlers that learn two very similar languages--Catalan and Spanish--and are therefore exposed to many cognates. Our results might not generalise to other groups of bilinguals learning two less similar languages.
* [To be discussed]

## Further steps
* [To be discussed]

# References {#sec-references}

::: {#refs}
:::

#  Appendix {#sec-appendix}

## Appendix 1: predictors {#sec-appendix-predictors}

### Word exposure index

Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database [@macwhinney2000childes; @sanchez2019childes]. Using the corresponding lexical frequencies from the Catalan and Spanish corpora was not possible due to the low number of Catalan participants and tokens available in those languages. Available words in the English corpora were mapped to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human]. Responses to words with missing lexical frequencies were excluded from analyses of the total number of items. Participant degree of exposure to the words' language were calculated as the percentage of exposure to the language, as reported by caretakers in the language profile section of the questionnaire. For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.

We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to. the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).

$$
\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
$$ {#eq-exposure}

For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)`.

### Levenshtein similarity

Generally, a pair of words from two different languages that share meaning---translation equivalents---are considered as *cognates* if they share etymological origin (e.g., *table* and *taula*, in English and Catalan). The fact that two words share etymological origin frequently leads to them also being form-similar, as reflected by their overlapping orthographic or phonological form. This makes many cognates perceptually similar. In psycholinguistics, the term *cognateness* is often used to refer to the form-similarity that a pair of translation equivalents share, which has been found to impact how bilinguals process such word-forms (e.g., @costa2000cognate; @spivey1999cross). Whether two form-similar translation equivalents are etymologically related or not is arguably tangential to the question of how cognateness affect language processing. For instance, the translation pair *much* and *mucho*, in English and Spanish have the same meaning, and are considerably similar at the orthographical and phonological level. Nonetheless, they come from different Proto-Indo-European roots [@campbell2007glossary]. This does not keep bilinguals from processing such kind of word-forms word forms differently than other translation equivalents with no form-similarity like *dog* and *perro*, in English and Spanish. In the scope of this experiment, we consider any form-similar translation equivalent as a *cognate* pair, regardless of whether both word-forms share etymological origin.

Our study capitalises in phonology, therefore we operationalised *cognateness* as the phonological similarity between a pair of translation equivalents. Our dataset contained `r max(n_items)` translation equivalents, each consisting in two word-forms: Catalan and Spanish (see @tbl-translation-levenshtein for an example). Measuring the phonological similarity between two word forms is non trial. While orthographic similarity can be calculated using string similarity/distance measures on the written word-forms, phonology poses additional challenges. A first, bottom-up approach to to the task might be considering acoustic similarity as a proxy to phonological similarity. This may involve registering audio recordings of a talker reading each word-form aloud, and then comparing their spectrograms,  cochleograms, or formant tracks [e.g., @heeringa2003norwegian]. While this is possible, this method computes acoustic similarity, as opposed to phonological similarity. This is not optimal, since a purely acoustic measure of similarity ignores how the actual signal is perceived by a actual listener. Since approximately 12 months of age, humans perceive the acoustic sounds in speech as phonemes, according to the phonology of their native language(s) [@werker1984crosslanguage; @kuhl1992linguistic]. This means that two listeners may not perceive the same linguistically relevant acoustic signal identically. This is why phonological similarity needs to be computed from discrete units that are meaningful for language perception.

A way to achieve this is to use phonological transcriptions of the word forms. Phonological transcriptions are symbolic, visual representations of the sounds in speech that capture some minimal  set of its features following some standard. The International Phonetic Alphabet (*IPA*) is one instance of such standards, providing a set of symbols that correspond to specific phonemes, as defined by their position across a series of dimensions. Using IPA transcriptions of word forms, one can use the same similarity/distance metric as in the case of orthographic word-forms. This method has been successfully used to measure the pairwise similarity between phonological word-forms from the  same language [e.g., @fourtassi2020growth] and across languages [@heeringa2004measuring; @floccia2018ii], and also between orthographic word-forms [@schepens2012distributions].

Multiple algorithms have been created to compute the distance between two strings of characters. The Levenshtein distance being one of them. The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary, see @eq-levenshtein]. For instance, /kaza/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /kasa/ (/z/ is replaced by /s/). Therefore, the Levenshtein distance between /kaza/ and /kasa/ is one. for an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), used phonological transcriptions in Speech Assessment Methods Phonetic Alphabet (SAMPA) format, as opposed to IPA, for its computer-friendly set of symbols.

$$
\text{lev}(x, y) =  
\begin{cases}
\max(i, j) & \text{if} \ \min(i, j)=0\\
\min \begin{cases}
\ \text{lev}_{x, y}(i-1, j) + 1 \\
\ \text{lev}_{x, y}(i, j-1) + 1 \\
\ \text{lev}_{x, y}(i-1, j-1) + 1_{a_i \neq b_j} \\
\end{cases}
\ &\text{otherwise}
\end{cases}
$${#eq-levenshtein}

Here, $a$ and $b$ are the character strings corresponding to the phonological transcriptions of two word-forms belonging to the same translation equivalent, where each character corresponds to one phoneme expressed as a symbol from the SAMPA alphabet. $i$ and $j$ are the length (i.e., number of phonemes) of the $s$ and $t$ strings respectively.

To measure the phonological similarity between the translation equivalents in our study, we computed a normalised versions of the Levenshtein distance between each pair of word-forms that accounts for the lenght of the strings. The rationale behind this correction is that longer word-forms are more likely to differ from their counterpart, compared to shorter word forms, and that this tendency does not correspond necessarily to a larger distance in terms of perception. This normalisation is achieved by dividing the Levenshtein distance by the length of the longest string, which leads to a distance metric that ranges from 0 to 1. This can be interpreted as a proportion of characters in the longest string that need to be edited in order for that string to become identical to the shorter string. A 0% normalised Levenshtein distance indicates that both word-forms are identical. A 50% normalised Levenshtein distance indicates that half of the phonemes in the longest word-form must be edited for both word-forms to become identical. A 100% normalised Levenshtein distance indicates that the two word-forms are completely different, as *all* phonemes in the longest word-form must be changed for it to become identical to the other.

Finally, since we are interested in *cognateness*, in order to easy the interpretation of the analyses we subtracted the normalised Levenshtein distance from one, so that the metric indicated the *similarity* between the each pair of word-forms, instead of their distance. @eq-levenshtein-similarity shows the formula of this metric.

$$
1-\frac{\text{lev}(a, b)}{\max{\{i, j})\}}
$${#eq-levenshtein-similarity}

The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure (see @tbl-translation-levenshtein for more examples):

| Catalan 	      | Spanish   	      | Levenshtein  |       
|---------	      |---------	      |-----         |
| *porta* (pOrt5) | *puerta* (pwe4ta) |   0.50 (3)   |
| *taula* (tawl5) | *mesa* (mesa)     |   0.00 (5)   |
| *cotxe* (kOtS5) | *coche* (kotSe)   |   0.40 (3)   |

: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study  {#tbl-translation-levenshtein}

## Appendix 2: Model comparisons {#sec-appendix-loo}

```{r}
#| label: tbl-results-loos
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV). The predictive accuracy of the models was compared using leave-one-out cross-validation. The likelihood of each data point given a model fitted without including that same observation was estimated. The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model. *ELPD* values closer to zero indicate better predictive acuracy. We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*). The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
model_loos %>%
    loo_compare() %>%
    as.data.frame() %>%
    rownames_to_column("model") %>%
    mutate(model = str_remove(model, "model_fit_")) %>%
    relocate(model,
             matches("elpd_loo"),
             matches("p_loo"),
             matches("looic"),
             matches("diff")) %>%
    gt() %>%
    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>%
    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>%
    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>%
    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>%
    fmt_number(2:9) %>%
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("*SE*"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*")
    ) %>%
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>%
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) %>%
    tab_style(cell_text(align = "left"),
              cells_title(groups = "subtitle"))
```

## Appendix 3: Model details {#sec-appendix-model}

We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function. The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood.

$$
\begin{aligned}
\textbf{Likelihood:} \\
y_{ij} &\sim \text{Cumulative}(p_{k})
\end{aligned}
$$ {#eq-likelihood}

where:

- $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$)
- $i$ is the participant index
- $j$ is the translation equivalent (TE) index
- $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$)) between two response categories in the latent distribution
$p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.

To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$. We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset--each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression]. For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].

$$
\begin{aligned}
\textbf{Linear model:} \\
logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Phonemes}_{ij} + & \\
& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
\end{aligned}
$$ {#eq-linear}

where:

- $i$ and $j$ index the participant and translation equivalent (TE)
- $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$
- $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively
- $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest
- $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively

We used the Bayesian framework to estimate the parameters in our model. This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical]. This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations. We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].

$$
\begin{aligned}
\\
\textbf{Prior:} \\
\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
&&\mbox{[Participant-level coefficient variability]} \\ \\
\Bigg(\begin{smallmatrix}
u_{k_{0}} \\ 
u_{1_{i}} \\ 
\vdots \\ 
u_{8_{i}} 
\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
\Bigg(\Bigg(\begin{smallmatrix}0 \\
0 \\ 
\vdots \\
0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
\vdots & \vdots & \vdots & \vdots \\
\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
\rho_{u} &\sim LKJcorr(2) \\
\\
&&\mbox{[TE-level coefficient variability]} \\ \\
\Bigg(\begin{smallmatrix}
w_{k_{0}}\\ 
w_{1_{j}} \\ 
\vdots \\ 
w_{3_{j}} 
\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
0\\ 
0 \\ 
\vdots \\
0 
\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
\vdots & \vdots & \vdots & \vdots \\
\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
\rho_{w_{0-3}} &\sim LKJcorr(2)
\end{aligned}
$$ {#eq-prior}

where:

- $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively
- $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively
- $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].

```{r}
#| label: fig-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 5
#| fig-cap: "Expected prior predictions."
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    exposure_std = 0,
    lv_std = 0
)

m <- add_epred_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA) %>% 
    filter(.category != "No") %>% 
    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(
        c(Understands, `Understands and Says`),
        names_to = ".category",
        values_to = ".epred"
    ) %>% 
    mutate(
        .category = factor(
            .category,
            levels = c("Understands and Says", "Understands"),
            labels = c("Production\n(Understands and Says)", "Comprehension\n(Understands)")
        )
    )

m %>% 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    stat_lineribbon(linewidth = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", linewidth = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(
        values = rev(clrs), 
        labels = c("95%", "89%", "67%", "50%")
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank()
    )
```

## Appendix 4: do results change when including frequency and language exposure as separate predictors?

### 2023-01-17 docs/index.qmd
Update lab notes
Add figures to manuscript, start discussion, add model and predictor appendix

--- 2022-08-27 docs/index.Rmd
@@
 title: "Trajectories"
-output:
-  github_document:
-    pandoc_args: --webtex
+subtitle: "Lab notes"
+date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
+authors:
+  - name: Gonzalo Garcia-Castro
+    orcid: 0000-0002-8553-4209
+    email: gonzalo.garciadecastro@upf.edu
+    corresponding: true
+    affiliations:
+      - name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+        address: Ramon Trias fargas 25-27
+        city: Barcelona
+        state: Spain
+        postal-code: 08005
+  - name: Daniela S. Ávila-Varela
+    orcid: 0000-0002-3518-8117
+    affiliations:
+      - name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+  - name: Ignacio Castillejo
+    orcid: 0000-0001-7445-0416
+    affiliations:
+      - name: Universidad Autónoma de Madrid
+        department: Departamento de Psicología
+  - name: Nuria Sebastian-Galles
+    orcid: 0000-0001-6938-2498
+    affiliations:
+      - name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
+Floccia et al.
+(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
+The mechanisms underlying this effect remain unknown.
+We explore two compatible scenarios.
+First, we test whether cognates are learnt earlier than non-cognates.
+This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
+Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
+This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
+We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
+format:
+    html:
+        code-copy: hover
+        reference-location: margin
+        code-overflow: scroll
+    pdf: default
+link-external-newwindow: true
+code-fold: true
+code-line-numbers: true
+fig-dpi: 1000
+warning: false
+echo: false
+toc: true
+colorlinks: true
+number-sections: true
 bibliography: "references.bib"
+csl: "apa7.csl"
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-    message = TRUE,
-    echo = FALSE,
-    warning = TRUE,
-    include = TRUE,
-    out.width = "80%",
-    dpi = 250
-)
-options(
-    knitr.kable.NA = "-",
-    knitr.duplicate.label = "allow",
-    ggplot2.discrete.fill = ggsci::pal_d3()(4),
-    ggplot2.discrete.colour = ggsci::pal_d3()(4),
+```{r}
+#| label: setup
+#| include: false
+
+# import packages
+library(conflicted)
+library(ggplot2)
+library(patchwork)
+library(dplyr)
+library(tidyr)
+library(purrr)
+library(gt)
+library(forcats)
+library(stringr)
+library(tibble)
+library(tidybayes)
+library(janitor)
+library(brms)
+library(scales)
+library(marginaleffects)
+library(lubridate)
+library(bvqdev)
+
+# load targets as R objects
+tar_load(participants)
+tar_load(df)
+tar_load(items)
+tar_load(bvq_data)
+tar_load(model_fit_0)
+tar_load(model_fit_1)
+tar_load(model_fit_2)
+tar_load(model_fit_3)
+tar_load(model_fit_4)
+tar_load(model_fit_4_prior)
+tar_load(model_fit_prior)
+tar_load(model_loos)
+tar_load(rope_interval)
+tar_load(posterior_description)
+tar_load(posterior_description_re)
+
+# resolve namespace conflicts
+conflict_prefer("filter", "dplyr")
+conflict_prefer("select", "dplyr")
+
+# set ggplot theme and colour palette
+theme_set(theme_custom()) # set custom ggplot theme
+clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+
+options(ggplot2.ordinal.fill = clrs,
+        ggplot2.ordinal.colour = clrs,
+        ggplot2.discrete.fill = clrs,
+        ggplot2.discrete.colour = clrs,
         ggplot2.continuous.fill = ggplot2::scale_color_gradient,
-    ggplot2.continuous.colour = ggplot2::scale_color_gradient
-)
-
-```
-
-```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
-# set params
-tar_load_globals()
-tar_load_all()
-theme_set(theme_custom())
-
-```
-
-# Questionnaires
+        ggplot2.continuous.colour = ggplot2::scale_color_gradient)
+```
+
+# Methods
+
+::: {.callout-tip}
+Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
+:::
+
+## Questionnaires
+
+```{r}
+#| label: questionnaire
+#| echo: false
+#| message: false
+#| warning: false
+data("pool", package = "bvqdev")
+n_categories <- distinct(pool, semantic_category) %>% nrow()
+n_items <- length(unique(pool$item))
+n_item_language <- count(pool, language) %>% 
+    pull(n) %>%
+    set_names(c("catalan", "spanish"))
+```
 
 The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted of two lists of words: one in Catalan and one in Spanish.
+The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]` items.
+Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* \[dog\] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* \[*to finish*\] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* \[*monkey*\] and Catalan *mono/mico*.
 We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
@@
 For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
-These last families were randomly allocated into one of four different subsets of the complete list of items.
-These lists were carefully designed so that each contained a representative subsample of the items from the complete list.
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\~400 translation equivalents, \~400 items in Catalan, \~400 items in Spanish).
+These last families were randomly assigned to one of four different subsets of the complete list of items.
+These lists were designed so that each contained a representative subsample of the items from the complete list.
+Semantic/functional categories with less than 16 items were not divided in the short version of the questionnaire to preserve a minimum of four items per list for each category: all of their items were included in the four lists.
 Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
 Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
+We excluded from the analysis multi-word items (e.g., *barrita de cereales* \[cereal bar\]) and items that included more than one word-form (e.g., *mono / mico*).
 Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
 
-# Items {.tabset .tab-pills}
-
-```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies.
-Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
-
-items %>%
-    left_join(select(multilex_data$pool, te, language)) %>% 
-    unnest(list) %>% 
-    group_by(language, list) %>% 
-    summarise_at(
-        vars(freq, n_phon, lv), 
-        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
-    ) %>% 
-    select(-c(n_phon_n, lv_n)) %>% 
-    rename(n = freq_n) %>% 
-    gt(groupname_col = "language") %>% 
-    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
-    fmt_percent(matches("lv_")) %>% 
-    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
-    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
-    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
-    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
-    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
-    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
-    cols_label(
-        list = "",
-        n = md("*N*"),
-        freq_mean = md("Mean"),
-        freq_sd = md("*SD*"),
-        freq_min = md("Range"),
-        n_phon_mean = md("Mean"),
-        n_phon_sd = md("*SD*"),
-        n_phon_min = md("Range"),
-        lv_mean = md("Mean"),
-        lv_sd = md("*SD*"),
-        lv_min = md("Range")
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(everything())
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(everything())
-    )
-
-```
-
-## Lexical frequency
-
-```{r items-frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-
-items %>% 
-    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
-    ggplot() +
-    aes(
-        x = freq,
-        fill = language,
-        colour = language,
-        shape = language
-    ) +
-    stat_slab(
-        slab_colour = "white",
-        slab_size = 1,
-        alpha = 0.5,
-    ) +
-    stat_pointinterval(
-        aes(y = -0.1),
-        position = position_dodge(width = 0.1),
-        point_size = 3
-    ) +
-    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
-    labs(
-        x = "Frequency (Zipf)",
-        y = "Probability density",
-        fill = "Language",
-        shape = "Language",
-        colour = "Language"
-    ) +
-    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
-    theme(
-        legend.position = "top",
-        axis.title = element_blank()
-    )
-
-```
-
-## # Phonemes
-
-```{r items-nphon, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-
-items %>%
-    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>% 
-    count(language, n_phon) %>% 
-    mutate(prop = n/sum(.$n)) %>% 
-    ggplot() +
-    aes(
-        x = n_phon,
-        y = prop,
-        fill = language,
-        colour = language,
-        shape = language
-    ) +
-    geom_col(
-        position = position_dodge()
-    ) +
-    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
-    labs(
-        x = "Frequency (Zipf)",
-        y = "Probability density",
-        fill = "Language",
-        shape = "Language",
-        colour = "Language"
-    ) +
-    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
-    theme(
-        legend.position = "top",
-        axis.title = element_blank()
-    )
-
-```
-
-## Levenshtein
-
-```{r items-lv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
-
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
-    count(lv_cut) %>% 
-    mutate(prop = n/sum(.$n)) %>% 
-    ggplot() +
-    aes(x = lv_cut, y = prop) +
-    geom_col(fill = ggsci::pal_d3()(4)[1]) +
-    labs(
-        x = "Levenshtein",
-        y = "Probability density",
-    ) +
-    theme(
-        legend.position = "top",
-        axis.title = element_blank()
-    )
-
-```
-
-## Frequency by # Phonemes
-
-```{r items-frequency-phonemes, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-items %>% 
-    left_join(select(pool, item, language)) %>% 
-    drop_na(freq) %>% 
-    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
-    facet_wrap(~language) +
-    geom_point(size = 1, alpha = 0.25) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
-    labs(
-        x = "# Phonemes", 
-        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
+## Participants
+
+```{r}
+#| label: fig-participants-time
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 10 
+#| fig-cap: "Date at which responses were collected"
+participants %>% 
+    group_by(time_stamp) %>% 
+    summarise(n = n(), .groups = "drop") %>% 
+    mutate(n = cumsum(n)) %>% 
+    ggplot(aes(time_stamp, n)) +
+    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
+    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
+             angle = 90, vjust = 1.5, hjust = 1) +
+    geom_line(size = 1, colour = clrs[1]) +
+    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
+    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
+    theme(legend.position = "top",
+          legend.title = element_blank(),
           panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    )
-
-```
-
-## Frequency by Levenshtein
-
-```{r items-frequency-levenshtein, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-items %>% 
-    left_join(select(pool, item, language)) %>% 
-    drop_na(freq) %>% 
-    ggplot(aes(freq, lv, colour = language, fill = language)) +
-    facet_wrap(~language) +
-    geom_point(size = 1, alpha = 0.5) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
-    labs(
-        y = "Levenshtein similarity", 
-        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "none",
-        panel.grid.major.y = element_blank()
-    )
-
-```
-
-## # Phonemes by Levenshtein
-
-```{r items-phonemes-lv, message=FALSE, warning=FALSE, paged.print=FALSE}
-items %>% 
-    left_join(select(pool, item, language)) %>% 
-    drop_na(freq) %>% 
-    ggplot(aes(n_phon, lv, colour = language, fill = language)) +
-    facet_wrap(~language) +
-    geom_point(size = 1, alpha = 0.25) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
-    labs(
-        x = "# Phonemes", 
-        y = "Levenshtein similarity", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_y_continuous(labels = percent) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    ) 
-```
-
-## {-}
-
-# Participants {.tabset .tab-pills}
-
-```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
+          axis.text.x = element_text(angle = 20, hjust = 1),
+          axis.title.x = element_blank()) +
+
     participants %>% 
-    mutate(dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish")) %>% 
-    group_by(time_stamp, dominant_language) %>% 
+    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
+    group_by(time_stamp, age) %>% 
     summarise(n = n(), .groups = "drop") %>% 
-    group_by(dominant_language) %>% 
+    group_by(age) %>% 
     mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
+    ggplot(aes(time_stamp, n, colour = age)) +
     geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
-             angle = 90, vjust = 1.5, hjust = 1) +
     geom_line(size = 1) +
-    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
-    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
-    guides(colour = guide_legend(ncol = 2)) +
+    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
+    scale_y_continuous(breaks = seq(0, 500, 50)) +
+    scale_colour_manual(values = clrs) +
+
+    plot_layout(ncol = 1) &
+
+    scale_x_date(date_breaks = "3 month", date_labels = "%B %Y") &
     theme(
@@
         legend.title = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle = 20, hjust = 1),
         axis.title.x = element_blank(),
     )
-
-```
-
-## Age
-
-```{r participants-age, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```
+
+### Age
+
+Number of months elapsed between participants' birth date and questionnaire completion.
+
+```{r}
+#| label: fig-participants-age
+#| fig-width: 7
+#| fig-cap: "Distribution of participants' ages"
 participants %>% 
@@
     aes(x = age, y = n) +
-    geom_col() +
+    geom_col(fill = clrs[1]) +
     geom_text(aes(label = n), size = 3.5, vjust = -1) +
@@
         panel.grid.minor.x = element_blank()
-
-    )
-```
-
-## Language profile
-
-```{r participants-lp, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+    )
+```
+
+### Language exposure (Degree of Exposure)
+
+Participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to.
+For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.
+
+```{r}
+#| label: fig-participants-lp
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
 participants %>% 
+    pivot_longer(
+        c(doe_catalan, doe_spanish),
+        names_to = "language", values_to = "doe", 
+        names_transform = ~str_to_sentence(str_remove(., "doe_"))
+    ) %>% 
     mutate(
-        dominant_language = ifelse(doe_catalan >= doe_spanish, "Catalan", "Spanish"),
-        age = round(age),
-        id = paste(id, "(", time, ")"),
-        rank = rank(doe_catalan)
+        doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
+        id = paste0(id, " (", time, ")")
     ) %>% 
-    pivot_longer(
-        starts_with("doe_"), 
-        names_to = "language",
-        values_to = "doe",
-        names_prefix = "doe_",
-        names_transform = list(language = str_to_sentence)
-    ) %>% 
-    ggplot() +
-    aes(x = doe, y = reorder(id, -rank), fill = language) +
-    geom_col(width = 0.75, position = position_fill(), colour = "white", size = 0.1) +
+    count(language, doe) %>% 
+    ggplot(aes(doe, y = n, label = n, fill = language)) +
+    geom_col(colour = "white", size = 1) +
+    geom_text(position = position_nudge(y = 5)) +
+    facet_wrap(~language, ncol = 1) + 
     labs(
         x = "Degree of exposure to Catalan",
-        y = "Participant",
+        y = "Number of participants",
         colour = "Language",
@@
     ) +
-    scale_x_continuous(labels = percent) +
-    theme(
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "top",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor = element_blank()
+    )
+```
+
+### SES/parental education
+
+<!-- # ```{r} -->
+<!-- #| label: tbl-participants-edu -->
+<!-- #|  tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)." -->
+<!-- # participants %>% -->
+<!-- #     left_join(bvq_data$logs) %>% -->
+<!-- #     mutate( -->
+<!-- #         edu_parent = fct_explicit_na(as.factor(edu_parent)), -->
+<!-- #         doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE) -->
+<!-- #     ) %>% -->
+<!-- #     count(doe_catalan, edu_parent) %>% -->
+<!-- #     right_join( -->
+<!-- #         expand_grid( -->
+<!-- #             doe_catalan = levels(.$doe_catalan), -->
+<!-- #             edu_parent = unique(bvq_data$logs$edu_parent) -->
+<!-- #         ) -->
+<!-- #     ) %>% -->
+<!-- #     rename(total = n) %>% -->
+<!-- #     pivot_wider(names_from = edu_parent, -->
+<!-- #                 values_from = total,  -->
+<!-- #                 values_fill = 0) %>% -->
+<!-- #     mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% -->
+<!-- #     rename(Total = total) %>% -->
+<!-- #     select(-"NA") %>%  -->
+<!-- #     gt()  -->
+<!-- # summary_rows( -->
+<!-- #     columns = Complementary:Primary, -->
+<!-- #     fns = list(Total = ~sum(., na.rm = TRUE)), -->
+<!-- #     formatter = fmt_integer -->
+<!-- # ) %>% -->
+<!-- # cols_label(doe_catalan = "DoE Catalan") %>% -->
+<!-- # sub_missing(columns = 2:8, missing_text = "--") %>% -->
+<!-- # tab_spanner("Educational attainment", Secondary:Primary) %>% -->
+<!-- # tab_style( -->
+<!-- #     style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")), -->
+<!-- #     locations = cells_body(columns = 9, rows = 1:5) -->
+<!-- # ) %>% -->
+<!-- # tab_style( -->
+<!-- #     cell_text(style = "italic"), -->
+<!-- #     cells_column_labels(columns = 2:7) -->
+<!-- # ) %>% -->
+<!-- # tab_style( -->
+<!-- #     cell_text(weight = "bold"), -->
+<!-- #     cells_column_labels(columns = c(1, 8)) -->
+<!-- # ) %>% -->
+<!-- # tab_style( -->
+<!-- #     cell_text(weight = "bold"), -->
+<!-- #     cells_column_spanners(spanners = "Educational attainment") -->
+<!-- # ) -->
+<!-- ``` -->
+
+## Items
+
+```{r}
+#| label: tbl-items-summary
+#| tbl-cap: "Lexical frequencies.
+Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
+# items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))
+# 
+# items %>%
+#     unnest(list) %>%
+#     group_by(language, list) %>%
+#     summarise_at(
+#         vars(freq_zipf, n_phon, lv),
+#         list(
+#             n = ~sum(!is.na(.)),
+#             mean = ~mean(., na.rm = TRUE),
+#             sd = ~sd(., na.rm = TRUE),
+#             min = ~min(., na.rm = TRUE),
+#             max = ~max(., na.rm = TRUE)
+#         )
+#     ) %>%
+#     select(-c(n_phon_n, lv_n)) %>%
+#     rename(n = freq_zipf_n) %>%
+#     gt(groupname_col = "language", rowname_col = "list") %>%
+#     fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>%
+#     fmt_percent(matches("lv_"), decimals = 1) %>%
+#     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>%
+#     tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>%
+#     tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>%
+#     cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>%
+#     cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>%
+#     cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>%
+#     cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>%
+#     cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>%
+#     cols_merge_range(col_begin = lv_min, col_end = lv_max) %>%
+#     cols_label(
+#         list = "",
+#         n = md("*N*"),
+#         freq_zipf_mean = md("Mean ± SD"),
+#         freq_zipf_min = md("Range"),
+#         n_phon_mean = md("Mean ± SD"),
+#         n_phon_min = md("Range"),
+#         lv_mean = md("Mean ± SD"),
+#         lv_min = md("Range")
+#     ) %>%
+#     tab_style(
+#         cell_text(style = "italic"),
+#         cells_column_labels(everything())
+#     ) %>%
+#     tab_style(
+#         cell_text(weight = "bold"),
+#         cells_column_spanners(everything())
+#     ) %>%
+#     summary_rows(
+#         fns = list("Grand mean" = ~mean(.)),
+#         columns = ends_with("_mean")
+#     ) %>%
+#     summary_rows(
+#         fns = list("N" = ~sum(.)),
+#         columns = n,
+#         formatter = fmt_integer
+#     )
+```
+
+### Lexical frequency
+
+Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
+This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points.
+Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [{childesr}](https://langcog.github.io/childes-db-website/) R package.
+Available words in the corpora were then mapped to their Catalan and Spanish translations.
+Therefore, the Catalan and Spanish word-forms of the same translation equivalent had the same value for lexical frequency.
+Responses to words with missing lexical frequencies were excluded from analyses of the total number of items).
+
+```{r}
+#| label: fig-items-freq
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 4
+#| fig-cap: "Distribution of lexical frequency across items.
+Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations.
+Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency.
+The black dot and interval indicate the mean and standard deviation."
+mean_sd <- function(x, ...) {
+    m <- mean(x, ...)
+    s <- sd(x, ...)
+    data.frame(mean = m, ymin = m-s, ymax = m+s)
+}
+
+items %>% 
+    distinct(te, .keep_all = TRUE) %>% 
+    ggplot(aes(x = freq_zipf)) +
+    geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
+    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
+    theme(
+        legend.position = "none",
+        panel.grid.minor = element_blank(),
+        legend.title = element_blank()
+    )
+```
+
+### \# Phonemes
+
+Word length ($Phonemes$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.
+
+```{r}
+#| label: fig-items-nphon
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Distribution of number of phonemes across items"
+items %>%
+    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>%
+    count(language, n_phon) %>%
+    ggplot() +
+    aes(
+        x = n_phon,
+        y = n,
+        fill = language,
+        colour = language,
+        shape = language
+    ) +
+    facet_wrap(~language, ncol = 1) +
+    geom_col(colour = "white") +
+    geom_text(aes(label = n), position = position_nudge(y = 5), colour = "black") +
+    labs(
+        x = "# Phonemes",
+        y = "Number of words",
+        fill = "Language",
+        shape = "Language",
+        colour = "Language"
+    ) +
+    scale_x_continuous(breaks = 1:15) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.ticks.x = element_line(),
+        legend.title = element_blank()
+    )
+```
+
+### Phonological similarity (Levenshtein)
+
+Phonological similarity between translation equivalents was computed as the normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary].
+This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1.
+This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018vocabulary; @fourtassi2020growth; @laing2022phonological for similar approaches].
+
+```{r}
+#| label: fig-items-lv
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 9
+#| fig-cap: "Distribution of Levenshtein similarities across items"
+items %>% 
+    distinct(te, .keep_all = TRUE) %>% 
+    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
+    count(lv) %>% 
+    mutate(prop = n/sum(n)) %>% 
+    ggplot(aes(x = lv, n)) +
+    geom_col(fill = clrs[4], colour = "white") +
+    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
+              position = position_nudge(y = 3), size = 3.5) +
+    labs(x = "Levenshtein distance", y = "Number of words") +
+    theme(
+        panel.grid.major.x = element_blank(),
+        legend.position = "top"
+    )
+```
+
+### Frequency by \# Phonemes
+
+```{r}
+#| label: fig-items-frequency-phonemes
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 8
+#| fig-height: 4
+#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
+items %>% 
+    drop_na(freq_zipf) %>% 
+    distinct(te, .keep_all = TRUE) %>% 
+    count(n_phon, freq_zipf) %>% 
+    ggplot(aes(n_phon, freq_zipf)) +
+    geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
+               position = position_jitter(width = 0.1), seed = 888) +
+    geom_smooth(method = "lm", formula = "y ~ x" ,
+                colour = clrs[1], fill = clrs[1]) +
+    labs(
+        x = "# Phonemes", 
+        y = "Lexical frequency (Zipf score)", 
+        colour = "Language", 
+        fill = "Language"
+    ) +
+    scale_x_continuous(breaks = seq(0, 16, 1)) +
+    theme(
+        legend.position = "none",
+        panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank()
+    ) 
+```
+
+### Frequency by Levenshtein
+
+```{r}
+#| label: fig-items-frequency-lv
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Association betwen lexical frequency and phonological similarity"
+items %>% 
+    distinct(te, .keep_all = TRUE) %>% 
+    drop_na(freq_zipf) %>% 
+    count(freq_zipf, lv) %>% 
+    ggplot(aes(freq_zipf, lv, size = n)) +
+    geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
+    geom_smooth(method = "lm", formula = "y ~ x", colour = clrs[1], fill = clrs[1]) +
+    labs(
+        y = "Levenshtein similarity", 
+        x = "Lexical frequency (Zipf score)", 
+        colour = "Language", 
+        fill = "Language"
+    ) +
+    scale_y_continuous(labels = percent) +
+    theme(
+        legend.position = "none",
         panel.grid.major.y = element_blank()
     )
-
-```
-
-## SES/parental education
-
-```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-participants %>%
-    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
-    mutate(
-        edu_parent = fct_explicit_na(as.factor(edu_parent)),
-        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
-    ) %>% 
-    count(doe_catalan, edu_parent) %>% 
-    right_join(
-        expand_grid(
-            doe_catalan = levels(.$doe_catalan),
-            edu_parent = unique(multilex_data$logs$edu_parent)
-        )
-    ) %>% 
-    rename(total = n) %>% 
-    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
-    mutate(total = rowSums(cbind(.[,3:8]))) %>% 
-    rename(Total = total) %>% 
-    select(-"NA") %>% 
-    gt() %>% 
-    cols_label(
-        doe_catalan = "DoE Catalan"
-    ) %>% 
-    summary_rows(
-        columns = 2:7, 
-        fns = list(Total = ~sum(., na.rm = TRUE)),
-        formatter = fmt_number,
-        decimals = 0
-    ) %>% 
-    fmt_missing(columns = 3:8, missing_text = "--") %>% 
-    tab_spanner("Educational attainment", 3:8) %>% 
-    tab_style(
-        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
-        locations = cells_body(columns = 9, rows = 1:10)
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(columns = 2:7)
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = c(1, 8))
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(spanners = "Educational attainment")
-    )
-```
-
-## {-}
-
-# Data analysis {.tabset .tab-pills}
-
-We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_std` by `item`, and their correlation parameter.
-We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope.
-Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`.
-The models were implemented in `brms` as:
-
-* `understands/produces ~ 1 + age_std + frequency + (1 + age_std | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age_std + frequency + doe + (1 + age_std + doe | item) + (1 + frequency | id)`
-* `understands/produces ~ 1 + age_std + frequency + doe*cognate + (1 + age_std + doe | item) + (1 + frequency + cognate| id)`
-
-## Model equation
+```
+
+### \# Phonemes by Levenshtein
+
+```{r}
+#| label: fig-items-nphon-lv
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Association between number of phonemes and phonological similarity"
+items %>% 
+    distinct(te, .keep_all = TRUE) %>% 
+    drop_na(freq_zipf) %>% 
+    count(n_phon, lv) %>% 
+    ggplot(aes(n_phon, lv, size = n)) +
+    geom_point(shape = 1, stroke = 1, colour = clrs[4]) +
+    geom_smooth(method = "lm", formula = "y ~ x", size = 1,
+                colour = clrs[1], fill = clrs[1]) +
+    labs(
+        x = "# Phonemes", 
+        y = "Levenshtein similarity", 
+        colour = "Language", 
+        fill = "Language"
+    ) +
+    scale_y_continuous(labels = percent) +
+    scale_x_continuous(breaks = seq(0, 16, 1)) +
+    theme(
+        legend.position = "none",
+        panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank()
+    ) 
+```
+
+### Word exposure index by lexical frequency
+
+We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
+the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).
+
+$$
+\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
+$$ {#eq-exposure}
+
+For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):
 
@@
 \begin{aligned}
-\color{white}\log(\frac{p}{n - q}) &= \color{white}(\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
-& \color{white}(\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
-& \color{white}(\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
-& \color{white}(\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
-& \color{white}(\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
-& \color{white}(\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
-& \color{white}\varepsilon_{pi}
+\textbf{Exposure}_{\textbf{cavall, Catalan}} &= \textbf{Frequency}_{cavall} \times \textbf{DoE}_{Catalan} = 6.14 \times 0.8 = 4.91 \\
+\textbf{Exposure}_{\textbf{caballo, Spanish}} &= \textbf{Frequency}_{caballo} \times \textbf{DoE}_{Spanish} = 6.14 \times 0.2 = 1.23
 \end{aligned}
-$$
-
-## R code (`brms`) {.tabset}
-
-### Model 0
-
-```{r}
-print(model_fit_0$formula)
-```
-
-### Model 1
-
-```{r}
-print(model_fit_1$formula)
-```
-
-### Model 2
-
-```{r}
-print(model_fit_2$formula)
-```
-
-### Model 3
-
-```{r}
-print(model_fit_3$formula)
-```
-
-### Model 4
-
-```{r}
-print(model_fit_4$formula)
-```
-
-### Model 5
-
-```{r}
-print(model_fit_5$formula)
-```
-
-### Model 6
-
-```{r}
-print(model_fit_6$formula)
-```
-
-### Model 7
-
-```{r}
-print(model_fit_7$formula)
-```
-
-### {-}
-
-## Stan code {.tabset}
-
-Stan code generated by `brms::stancode`:
-
-### Model 0
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-}
-parameters {
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  vector[N_1] z_1[M_1];  // standardized group-level effects
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  vector[N_2] z_2[M_2];  // standardized group-level effects
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  vector[N_1] r_1_1;  // actual group-level effects
-  vector[N_2] r_2_1;  // actual group-level effects
-  r_1_1 = (sd_1[1] * (z_1[1]));
-  r_2_1 = (sd_2[1] * (z_2[1]));
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = rep_vector(0.0, N);
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_2_1[J_2[n]] * Z_2_1[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 1 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(z_1[1]);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 1 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(z_2[1]);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept;
-  // additionally sample draws from priors
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_sd_2 = normal_rng(1,0.1);
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-```
-
-### Model 1
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
- /* compute correlated group-level effects
-  * Args: 
-  *   z: matrix of unscaled group-level effects
-  *   SD: vector of standard deviation parameters
-  *   L: cholesky factor correlation matrix
-  * Returns: 
-  *   matrix of scaled group-level effects
-  */ 
-  matrix scale_r_cor(matrix z, vector SD, matrix L) {
-    // r is stored in another dimension order than z
-    return transpose(diag_pre_multiply(SD, L) * z);
-  }
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  int<lower=1> K;  // number of population-level effects
-  matrix[N, K] X;  // population-level design matrix
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  vector[N] Z_1_2;
-  int<lower=1> NC_1;  // number of group-level correlations
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  vector[N] Z_2_2;
-  int<lower=1> NC_2;  // number of group-level correlations
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-  int Kc = K;
-  matrix[N, Kc] Xc;  // centered version of X
-  vector[Kc] means_X;  // column means of X before centering
-  for (i in 1:K) {
-    means_X[i] = mean(X[, i]);
-    Xc[, i] = X[, i] - means_X[i];
-  }
-}
-parameters {
-  vector[Kc] b;  // population-level effects
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  matrix[M_1, N_1] z_1;  // standardized group-level effects
-  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  matrix[M_2, N_2] z_2;  // standardized group-level effects
-  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  matrix[N_1, M_1] r_1;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_1] r_1_1;
-  vector[N_1] r_1_2;
-  matrix[N_2, M_2] r_2;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_2] r_2_1;
-  vector[N_2] r_2_2;
-  // compute actual group-level effects
-  r_1 = scale_r_cor(z_1, sd_1, L_1);
-  r_1_1 = r_1[, 1];
-  r_1_2 = r_1[, 2];
-  // compute actual group-level effects
-  r_2 = scale_r_cor(z_2, sd_2, L_2);
-  r_2_1 = r_2[, 1];
-  r_2_2 = r_2[, 2];
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = Xc * b;
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(b[1] | 1, 0.1);
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 2 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_1));
-  target += lkj_corr_cholesky_lpdf(L_1 | 2);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 2 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_2));
-  target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-  // compute group-level correlations
-  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-  vector<lower=-1,upper=1>[NC_1] cor_1;
-  // compute group-level correlations
-  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-  vector<lower=-1,upper=1>[NC_2] cor_2;
-  // additionally sample draws from priors
-  real prior_b_1 = normal_rng(1,0.1);
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-  real prior_sd_2 = normal_rng(1,0.1);
-  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_1) {
-    for (j in 1:(k - 1)) {
-      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-    }
-  }
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_2) {
-    for (j in 1:(k - 1)) {
-      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-    }
-  }
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-```
-
-### Model 2
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
- /* compute correlated group-level effects
-  * Args: 
-  *   z: matrix of unscaled group-level effects
-  *   SD: vector of standard deviation parameters
-  *   L: cholesky factor correlation matrix
-  * Returns: 
-  *   matrix of scaled group-level effects
-  */ 
-  matrix scale_r_cor(matrix z, vector SD, matrix L) {
-    // r is stored in another dimension order than z
-    return transpose(diag_pre_multiply(SD, L) * z);
-  }
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  int<lower=1> K;  // number of population-level effects
-  matrix[N, K] X;  // population-level design matrix
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  vector[N] Z_1_2;
-  vector[N] Z_1_3;
-  int<lower=1> NC_1;  // number of group-level correlations
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  vector[N] Z_2_2;
-  int<lower=1> NC_2;  // number of group-level correlations
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-  int Kc = K;
-  matrix[N, Kc] Xc;  // centered version of X
-  vector[Kc] means_X;  // column means of X before centering
-  for (i in 1:K) {
-    means_X[i] = mean(X[, i]);
-    Xc[, i] = X[, i] - means_X[i];
-  }
-}
-parameters {
-  vector[Kc] b;  // population-level effects
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  matrix[M_1, N_1] z_1;  // standardized group-level effects
-  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  matrix[M_2, N_2] z_2;  // standardized group-level effects
-  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  matrix[N_1, M_1] r_1;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_1] r_1_1;
-  vector[N_1] r_1_2;
-  vector[N_1] r_1_3;
-  matrix[N_2, M_2] r_2;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_2] r_2_1;
-  vector[N_2] r_2_2;
-  // compute actual group-level effects
-  r_1 = scale_r_cor(z_1, sd_1, L_1);
-  r_1_1 = r_1[, 1];
-  r_1_2 = r_1[, 2];
-  r_1_3 = r_1[, 3];
-  // compute actual group-level effects
-  r_2 = scale_r_cor(z_2, sd_2, L_2);
-  r_2_1 = r_2[, 1];
-  r_2_2 = r_2[, 2];
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = Xc * b;
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(b[1] | 1, 0.1);
-  target += normal_lpdf(b[2] | 0, 0.1);
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 3 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_1));
-  target += lkj_corr_cholesky_lpdf(L_1 | 2);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 2 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_2));
-  target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-  // compute group-level correlations
-  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-  vector<lower=-1,upper=1>[NC_1] cor_1;
-  // compute group-level correlations
-  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-  vector<lower=-1,upper=1>[NC_2] cor_2;
-  // additionally sample draws from priors
-  real prior_b_1 = normal_rng(1,0.1);
-  real prior_b_2 = normal_rng(0,0.1);
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-  real prior_sd_2 = normal_rng(1,0.1);
-  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_1) {
-    for (j in 1:(k - 1)) {
-      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-    }
-  }
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_2) {
-    for (j in 1:(k - 1)) {
-      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-    }
-  }
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-```
-
-### Model 3
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
- /* compute correlated group-level effects
-  * Args: 
-  *   z: matrix of unscaled group-level effects
-  *   SD: vector of standard deviation parameters
-  *   L: cholesky factor correlation matrix
-  * Returns: 
-  *   matrix of scaled group-level effects
-  */ 
-  matrix scale_r_cor(matrix z, vector SD, matrix L) {
-    // r is stored in another dimension order than z
-    return transpose(diag_pre_multiply(SD, L) * z);
-  }
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  int<lower=1> K;  // number of population-level effects
-  matrix[N, K] X;  // population-level design matrix
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  vector[N] Z_1_2;
-  vector[N] Z_1_3;
-  vector[N] Z_1_4;
-  int<lower=1> NC_1;  // number of group-level correlations
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  vector[N] Z_2_2;
-  vector[N] Z_2_3;
-  vector[N] Z_2_4;
-  int<lower=1> NC_2;  // number of group-level correlations
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-  int Kc = K;
-  matrix[N, Kc] Xc;  // centered version of X
-  vector[Kc] means_X;  // column means of X before centering
-  for (i in 1:K) {
-    means_X[i] = mean(X[, i]);
-    Xc[, i] = X[, i] - means_X[i];
-  }
-}
-parameters {
-  vector[Kc] b;  // population-level effects
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  matrix[M_1, N_1] z_1;  // standardized group-level effects
-  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  matrix[M_2, N_2] z_2;  // standardized group-level effects
-  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  matrix[N_1, M_1] r_1;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_1] r_1_1;
-  vector[N_1] r_1_2;
-  vector[N_1] r_1_3;
-  vector[N_1] r_1_4;
-  matrix[N_2, M_2] r_2;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_2] r_2_1;
-  vector[N_2] r_2_2;
-  vector[N_2] r_2_3;
-  vector[N_2] r_2_4;
-  // compute actual group-level effects
-  r_1 = scale_r_cor(z_1, sd_1, L_1);
-  r_1_1 = r_1[, 1];
-  r_1_2 = r_1[, 2];
-  r_1_3 = r_1[, 3];
-  r_1_4 = r_1[, 4];
-  // compute actual group-level effects
-  r_2 = scale_r_cor(z_2, sd_2, L_2);
-  r_2_1 = r_2[, 1];
-  r_2_2 = r_2[, 2];
-  r_2_3 = r_2[, 3];
-  r_2_4 = r_2[, 4];
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = Xc * b;
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(b[1] | 1, 0.1);
-  target += normal_lpdf(b[2] | 0, 0.1);
-  target += normal_lpdf(b[3] | 0, 0.1);
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 4 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_1));
-  target += lkj_corr_cholesky_lpdf(L_1 | 2);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 4 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_2));
-  target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-  // compute group-level correlations
-  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-  vector<lower=-1,upper=1>[NC_1] cor_1;
-  // compute group-level correlations
-  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-  vector<lower=-1,upper=1>[NC_2] cor_2;
-  // additionally sample draws from priors
-  real prior_b_1 = normal_rng(1,0.1);
-  real prior_b_2 = normal_rng(0,0.1);
-  real prior_b_3 = normal_rng(0,0.1);
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-  real prior_sd_2 = normal_rng(1,0.1);
-  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_1) {
-    for (j in 1:(k - 1)) {
-      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-    }
-  }
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_2) {
-    for (j in 1:(k - 1)) {
-      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-    }
-  }
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-
-```
-
-### Model 4
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
- /* compute correlated group-level effects
-  * Args: 
-  *   z: matrix of unscaled group-level effects
-  *   SD: vector of standard deviation parameters
-  *   L: cholesky factor correlation matrix
-  * Returns: 
-  *   matrix of scaled group-level effects
-  */ 
-  matrix scale_r_cor(matrix z, vector SD, matrix L) {
-    // r is stored in another dimension order than z
-    return transpose(diag_pre_multiply(SD, L) * z);
-  }
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  int<lower=1> K;  // number of population-level effects
-  matrix[N, K] X;  // population-level design matrix
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  vector[N] Z_1_2;
-  vector[N] Z_1_3;
-  vector[N] Z_1_4;
-  vector[N] Z_1_5;
-  int<lower=1> NC_1;  // number of group-level correlations
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  vector[N] Z_2_2;
-  vector[N] Z_2_3;
-  vector[N] Z_2_4;
-  vector[N] Z_2_5;
-  int<lower=1> NC_2;  // number of group-level correlations
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-  int Kc = K;
-  matrix[N, Kc] Xc;  // centered version of X
-  vector[Kc] means_X;  // column means of X before centering
-  for (i in 1:K) {
-    means_X[i] = mean(X[, i]);
-    Xc[, i] = X[, i] - means_X[i];
-  }
-}
-parameters {
-  vector[Kc] b;  // population-level effects
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  matrix[M_1, N_1] z_1;  // standardized group-level effects
-  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  matrix[M_2, N_2] z_2;  // standardized group-level effects
-  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  matrix[N_1, M_1] r_1;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_1] r_1_1;
-  vector[N_1] r_1_2;
-  vector[N_1] r_1_3;
-  vector[N_1] r_1_4;
-  vector[N_1] r_1_5;
-  matrix[N_2, M_2] r_2;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_2] r_2_1;
-  vector[N_2] r_2_2;
-  vector[N_2] r_2_3;
-  vector[N_2] r_2_4;
-  vector[N_2] r_2_5;
-  // compute actual group-level effects
-  r_1 = scale_r_cor(z_1, sd_1, L_1);
-  r_1_1 = r_1[, 1];
-  r_1_2 = r_1[, 2];
-  r_1_3 = r_1[, 3];
-  r_1_4 = r_1[, 4];
-  r_1_5 = r_1[, 5];
-  // compute actual group-level effects
-  r_2 = scale_r_cor(z_2, sd_2, L_2);
-  r_2_1 = r_2[, 1];
-  r_2_2 = r_2[, 2];
-  r_2_3 = r_2[, 3];
-  r_2_4 = r_2[, 4];
-  r_2_5 = r_2[, 5];
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = Xc * b;
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(b[1] | 1, 0.1);
-  target += normal_lpdf(b[2] | 0, 0.1);
-  target += normal_lpdf(b[3] | 0, 0.1);
-  target += normal_lpdf(b[4] | 0, 0.1);
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 5 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_1));
-  target += lkj_corr_cholesky_lpdf(L_1 | 2);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 5 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_2));
-  target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-  // compute group-level correlations
-  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-  vector<lower=-1,upper=1>[NC_1] cor_1;
-  // compute group-level correlations
-  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-  vector<lower=-1,upper=1>[NC_2] cor_2;
-  // additionally sample draws from priors
-  real prior_b_1 = normal_rng(1,0.1);
-  real prior_b_2 = normal_rng(0,0.1);
-  real prior_b_3 = normal_rng(0,0.1);
-  real prior_b_4 = normal_rng(0,0.1);
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-  real prior_sd_2 = normal_rng(1,0.1);
-  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_1) {
-    for (j in 1:(k - 1)) {
-      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-    }
-  }
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_2) {
-    for (j in 1:(k - 1)) {
-      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-    }
-  }
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-
-```
-
-### Model 5
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
- /* compute correlated group-level effects
-  * Args: 
-  *   z: matrix of unscaled group-level effects
-  *   SD: vector of standard deviation parameters
-  *   L: cholesky factor correlation matrix
-  * Returns: 
-  *   matrix of scaled group-level effects
-  */ 
-  matrix scale_r_cor(matrix z, vector SD, matrix L) {
-    // r is stored in another dimension order than z
-    return transpose(diag_pre_multiply(SD, L) * z);
-  }
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  int<lower=1> K;  // number of population-level effects
-  matrix[N, K] X;  // population-level design matrix
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  vector[N] Z_1_2;
-  vector[N] Z_1_3;
-  vector[N] Z_1_4;
-  vector[N] Z_1_5;
-  vector[N] Z_1_6;
-  int<lower=1> NC_1;  // number of group-level correlations
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  vector[N] Z_2_2;
-  vector[N] Z_2_3;
-  vector[N] Z_2_4;
-  vector[N] Z_2_5;
-  int<lower=1> NC_2;  // number of group-level correlations
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-  int Kc = K;
-  matrix[N, Kc] Xc;  // centered version of X
-  vector[Kc] means_X;  // column means of X before centering
-  for (i in 1:K) {
-    means_X[i] = mean(X[, i]);
-    Xc[, i] = X[, i] - means_X[i];
-  }
-}
-parameters {
-  vector[Kc] b;  // population-level effects
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  matrix[M_1, N_1] z_1;  // standardized group-level effects
-  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  matrix[M_2, N_2] z_2;  // standardized group-level effects
-  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  matrix[N_1, M_1] r_1;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_1] r_1_1;
-  vector[N_1] r_1_2;
-  vector[N_1] r_1_3;
-  vector[N_1] r_1_4;
-  vector[N_1] r_1_5;
-  vector[N_1] r_1_6;
-  matrix[N_2, M_2] r_2;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_2] r_2_1;
-  vector[N_2] r_2_2;
-  vector[N_2] r_2_3;
-  vector[N_2] r_2_4;
-  vector[N_2] r_2_5;
-  // compute actual group-level effects
-  r_1 = scale_r_cor(z_1, sd_1, L_1);
-  r_1_1 = r_1[, 1];
-  r_1_2 = r_1[, 2];
-  r_1_3 = r_1[, 3];
-  r_1_4 = r_1[, 4];
-  r_1_5 = r_1[, 5];
-  r_1_6 = r_1[, 6];
-  // compute actual group-level effects
-  r_2 = scale_r_cor(z_2, sd_2, L_2);
-  r_2_1 = r_2[, 1];
-  r_2_2 = r_2[, 2];
-  r_2_3 = r_2[, 3];
-  r_2_4 = r_2[, 4];
-  r_2_5 = r_2[, 5];
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = Xc * b;
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(b[1] | 1, 0.1);
-  target += normal_lpdf(b[2] | 0, 0.1);
-  target += normal_lpdf(b[3] | 0, 0.1);
-  target += normal_lpdf(b[4] | 0, 0.1);
-  target += normal_lpdf(b[5] | 0, 0.1);
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 6 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_1));
-  target += lkj_corr_cholesky_lpdf(L_1 | 2);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 5 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_2));
-  target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-  // compute group-level correlations
-  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-  vector<lower=-1,upper=1>[NC_1] cor_1;
-  // compute group-level correlations
-  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-  vector<lower=-1,upper=1>[NC_2] cor_2;
-  // additionally sample draws from priors
-  real prior_b_1 = normal_rng(1,0.1);
-  real prior_b_2 = normal_rng(0,0.1);
-  real prior_b_3 = normal_rng(0,0.1);
-  real prior_b_4 = normal_rng(0,0.1);
-  real prior_b_5 = normal_rng(0,0.1);
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-  real prior_sd_2 = normal_rng(1,0.1);
-  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_1) {
-    for (j in 1:(k - 1)) {
-      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-    }
-  }
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_2) {
-    for (j in 1:(k - 1)) {
-      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-    }
-  }
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-```
-
-### Model 6
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
- /* compute correlated group-level effects
-  * Args: 
-  *   z: matrix of unscaled group-level effects
-  *   SD: vector of standard deviation parameters
-  *   L: cholesky factor correlation matrix
-  * Returns: 
-  *   matrix of scaled group-level effects
-  */ 
-  matrix scale_r_cor(matrix z, vector SD, matrix L) {
-    // r is stored in another dimension order than z
-    return transpose(diag_pre_multiply(SD, L) * z);
-  }
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  int<lower=1> K;  // number of population-level effects
-  matrix[N, K] X;  // population-level design matrix
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  vector[N] Z_1_2;
-  vector[N] Z_1_3;
-  vector[N] Z_1_4;
-  vector[N] Z_1_5;
-  vector[N] Z_1_6;
-  vector[N] Z_1_7;
-  int<lower=1> NC_1;  // number of group-level correlations
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  vector[N] Z_2_2;
-  vector[N] Z_2_3;
-  vector[N] Z_2_4;
-  vector[N] Z_2_5;
-  int<lower=1> NC_2;  // number of group-level correlations
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-  int Kc = K;
-  matrix[N, Kc] Xc;  // centered version of X
-  vector[Kc] means_X;  // column means of X before centering
-  for (i in 1:K) {
-    means_X[i] = mean(X[, i]);
-    Xc[, i] = X[, i] - means_X[i];
-  }
-}
-parameters {
-  vector[Kc] b;  // population-level effects
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  matrix[M_1, N_1] z_1;  // standardized group-level effects
-  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  matrix[M_2, N_2] z_2;  // standardized group-level effects
-  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  matrix[N_1, M_1] r_1;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_1] r_1_1;
-  vector[N_1] r_1_2;
-  vector[N_1] r_1_3;
-  vector[N_1] r_1_4;
-  vector[N_1] r_1_5;
-  vector[N_1] r_1_6;
-  vector[N_1] r_1_7;
-  matrix[N_2, M_2] r_2;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_2] r_2_1;
-  vector[N_2] r_2_2;
-  vector[N_2] r_2_3;
-  vector[N_2] r_2_4;
-  vector[N_2] r_2_5;
-  // compute actual group-level effects
-  r_1 = scale_r_cor(z_1, sd_1, L_1);
-  r_1_1 = r_1[, 1];
-  r_1_2 = r_1[, 2];
-  r_1_3 = r_1[, 3];
-  r_1_4 = r_1[, 4];
-  r_1_5 = r_1[, 5];
-  r_1_6 = r_1[, 6];
-  r_1_7 = r_1[, 7];
-  // compute actual group-level effects
-  r_2 = scale_r_cor(z_2, sd_2, L_2);
-  r_2_1 = r_2[, 1];
-  r_2_2 = r_2[, 2];
-  r_2_3 = r_2[, 3];
-  r_2_4 = r_2[, 4];
-  r_2_5 = r_2[, 5];
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = Xc * b;
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(b[1] | 1, 0.1);
-  target += normal_lpdf(b[2] | 0, 0.1);
-  target += normal_lpdf(b[3] | 0, 0.1);
-  target += normal_lpdf(b[4] | 0, 0.1);
-  target += normal_lpdf(b[5] | 0, 0.1);
-  target += normal_lpdf(b[6] | 0, 0.1);
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 7 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_1));
-  target += lkj_corr_cholesky_lpdf(L_1 | 2);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 5 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_2));
-  target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-  // compute group-level correlations
-  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-  vector<lower=-1,upper=1>[NC_1] cor_1;
-  // compute group-level correlations
-  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-  vector<lower=-1,upper=1>[NC_2] cor_2;
-  // additionally sample draws from priors
-  real prior_b_1 = normal_rng(1,0.1);
-  real prior_b_2 = normal_rng(0,0.1);
-  real prior_b_3 = normal_rng(0,0.1);
-  real prior_b_4 = normal_rng(0,0.1);
-  real prior_b_5 = normal_rng(0,0.1);
-  real prior_b_6 = normal_rng(0,0.1);
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-  real prior_sd_2 = normal_rng(1,0.1);
-  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_1) {
-    for (j in 1:(k - 1)) {
-      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-    }
-  }
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_2) {
-    for (j in 1:(k - 1)) {
-      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-    }
-  }
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-```
-
-### Model 7
-
-```{stan, eval = FALSE, tidy = FALSE}
-// generated with brms 2.16.1
-functions {
- /* compute correlated group-level effects
-  * Args: 
-  *   z: matrix of unscaled group-level effects
-  *   SD: vector of standard deviation parameters
-  *   L: cholesky factor correlation matrix
-  * Returns: 
-  *   matrix of scaled group-level effects
-  */ 
-  matrix scale_r_cor(matrix z, vector SD, matrix L) {
-    // r is stored in another dimension order than z
-    return transpose(diag_pre_multiply(SD, L) * z);
-  }
-  /* cratio-logit log-PDF for a single response
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: ordinal thresholds
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_lpmf(int y, real mu, real disc, vector thres) {
-     int nthres = num_elements(thres);
-     vector[nthres + 1] p;
-     vector[nthres] q;
-     int k = 1;
-     while (k <= min(y, nthres)) {
-       q[k] = log_inv_logit(disc * (mu - thres[k]));
-       p[k] = log1m_exp(q[k]);
-       for (kk in 1:(k - 1)) p[k] = p[k] + q[kk];
-       k += 1;
-     }
-     if (y == nthres + 1) {
-       p[nthres + 1] = sum(q);
-     }
-     return p[y];
-   }
-  /* cratio-logit log-PDF for a single response and merged thresholds
-   * Args:
-   *   y: response category
-   *   mu: latent mean parameter
-   *   disc: discrimination parameter
-   *   thres: vector of merged ordinal thresholds
-   *   j: start and end index for the applid threshold within 'thres'
-   * Returns:
-   *   a scalar to be added to the log posterior
-   */
-   real cratio_logit_merged_lpmf(int y, real mu, real disc, vector thres, int[] j) {
-     return cratio_logit_lpmf(y | mu, disc, thres[j[1]:j[2]]);
-   }
-}
-data {
-  int<lower=1> N;  // total number of observations
-  int Y[N];  // response variable
-  int<lower=2> nthres;  // number of thresholds
-  int<lower=1> K;  // number of population-level effects
-  matrix[N, K] X;  // population-level design matrix
-  // data for group-level effects of ID 1
-  int<lower=1> N_1;  // number of grouping levels
-  int<lower=1> M_1;  // number of coefficients per level
-  int<lower=1> J_1[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_1_1;
-  vector[N] Z_1_2;
-  vector[N] Z_1_3;
-  vector[N] Z_1_4;
-  vector[N] Z_1_5;
-  vector[N] Z_1_6;
-  vector[N] Z_1_7;
-  int<lower=1> NC_1;  // number of group-level correlations
-  // data for group-level effects of ID 2
-  int<lower=1> N_2;  // number of grouping levels
-  int<lower=1> M_2;  // number of coefficients per level
-  int<lower=1> J_2[N];  // grouping indicator per observation
-  // group-level predictor values
-  vector[N] Z_2_1;
-  vector[N] Z_2_2;
-  vector[N] Z_2_3;
-  vector[N] Z_2_4;
-  vector[N] Z_2_5;
-  int<lower=1> NC_2;  // number of group-level correlations
-  int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-  int Kc = K;
-  matrix[N, Kc] Xc;  // centered version of X
-  vector[Kc] means_X;  // column means of X before centering
-  for (i in 1:K) {
-    means_X[i] = mean(X[, i]);
-    Xc[, i] = X[, i] - means_X[i];
-  }
-}
-parameters {
-  vector[Kc] b;  // population-level effects
-  vector[nthres] Intercept;  // temporary thresholds for centered predictors
-  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-  matrix[M_1, N_1] z_1;  // standardized group-level effects
-  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-  matrix[M_2, N_2] z_2;  // standardized group-level effects
-  cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-  real<lower=0> disc = 1;  // discrimination parameters
-  matrix[N_1, M_1] r_1;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_1] r_1_1;
-  vector[N_1] r_1_2;
-  vector[N_1] r_1_3;
-  vector[N_1] r_1_4;
-  vector[N_1] r_1_5;
-  vector[N_1] r_1_6;
-  vector[N_1] r_1_7;
-  matrix[N_2, M_2] r_2;  // actual group-level effects
-  // using vectors speeds up indexing in loops
-  vector[N_2] r_2_1;
-  vector[N_2] r_2_2;
-  vector[N_2] r_2_3;
-  vector[N_2] r_2_4;
-  vector[N_2] r_2_5;
-  // compute actual group-level effects
-  r_1 = scale_r_cor(z_1, sd_1, L_1);
-  r_1_1 = r_1[, 1];
-  r_1_2 = r_1[, 2];
-  r_1_3 = r_1[, 3];
-  r_1_4 = r_1[, 4];
-  r_1_5 = r_1[, 5];
-  r_1_6 = r_1[, 6];
-  r_1_7 = r_1[, 7];
-  // compute actual group-level effects
-  r_2 = scale_r_cor(z_2, sd_2, L_2);
-  r_2_1 = r_2[, 1];
-  r_2_2 = r_2[, 2];
-  r_2_3 = r_2[, 3];
-  r_2_4 = r_2[, 4];
-  r_2_5 = r_2[, 5];
-}
-model {
-  // likelihood including constants
-  if (!prior_only) {
-    // initialize linear predictor term
-    vector[N] mu = Xc * b;
-    for (n in 1:N) {
-      // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n] + r_2_4[J_2[n]] * Z_2_4[n] + r_2_5[J_2[n]] * Z_2_5[n];
-    }
-    for (n in 1:N) {
-      target += cratio_logit_lpmf(Y[n] | mu[n], disc, Intercept);
-    }
-  }
-  // priors including constants
-  target += normal_lpdf(b[1] | 1, 0.1);
-  target += normal_lpdf(b[2] | 0, 0.1);
-  target += normal_lpdf(b[3] | 0, 0.1);
-  target += normal_lpdf(b[4] | 0, 0.1);
-  target += normal_lpdf(b[5] | 0, 0.1);
-  target += normal_lpdf(b[6] | 0, 0.1);
-  target += normal_lpdf(b[7] | 0, 0.1);
-  target += normal_lpdf(b[8] | 0, 0.1);
-  target += normal_lpdf(b[9] | 0, 0.1);
-  target += normal_lpdf(Intercept | -0.25, 0.1);
-  target += normal_lpdf(sd_1 | 1, 0.1)
-    - 7 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_1));
-  target += lkj_corr_cholesky_lpdf(L_1 | 2);
-  target += normal_lpdf(sd_2 | 1, 0.1)
-    - 5 * normal_lccdf(0 | 1, 0.1);
-  target += std_normal_lpdf(to_vector(z_2));
-  target += lkj_corr_cholesky_lpdf(L_2 | 2);
-}
-generated quantities {
-  // compute actual thresholds
-  vector[nthres] b_Intercept = Intercept + dot_product(means_X, b);
-  // compute group-level correlations
-  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-  vector<lower=-1,upper=1>[NC_1] cor_1;
-  // compute group-level correlations
-  corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
-  vector<lower=-1,upper=1>[NC_2] cor_2;
-  // additionally sample draws from priors
-  real prior_b_1 = normal_rng(1,0.1);
-  real prior_b_2 = normal_rng(0,0.1);
-  real prior_b_3 = normal_rng(0,0.1);
-  real prior_b_4 = normal_rng(0,0.1);
-  real prior_b_5 = normal_rng(0,0.1);
-  real prior_b_6 = normal_rng(0,0.1);
-  real prior_b_7 = normal_rng(0,0.1);
-  real prior_b_8 = normal_rng(0,0.1);
-  real prior_b_9 = normal_rng(0,0.1);
-  real prior_Intercept = normal_rng(-0.25,0.1);
-  real prior_sd_1 = normal_rng(1,0.1);
-  real prior_cor_1 = lkj_corr_rng(M_1,2)[1, 2];
-  real prior_sd_2 = normal_rng(1,0.1);
-  real prior_cor_2 = lkj_corr_rng(M_2,2)[1, 2];
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_1) {
-    for (j in 1:(k - 1)) {
-      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-    }
-  }
-  // extract upper diagonal of correlation matrix
-  for (k in 1:M_2) {
-    for (j in 1:(k - 1)) {
-      cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
-    }
-  }
-  // use rejection sampling for truncated priors
-  while (prior_sd_1 < 0) {
-    prior_sd_1 = normal_rng(1,0.1);
-  }
-  while (prior_sd_2 < 0) {
-    prior_sd_2 = normal_rng(1,0.1);
-  }
-}
-```
-
-### {-}
-
-## {-}
-
-# Results {.tabset .tab-pills}
-
-## Raw data
-
-```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
+$$ {#eq-exposure-example}
+
+```{r}
+#| label: fig-exposure-distribution
+#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
+#| warnings: false
+#| fig-height: 7
+#| fig-width: 13
 df %>%
     mutate(
-        lv = cut_interval(lv, 5),
-        age = floor(age)
+        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
+        exposure = cut(
+            exposure,
+            breaks = seq(0, 9, 1),
+            include.lowest = TRUE
+        )
     ) %>%
-    count(lv, age, response) %>%
-    mutate(prop = n/sum(.$n)) %>% 
-    ggplot(aes(age, prop, colour = response, fill = response)) +
-    facet_wrap(~lv, nrow = 1) +
-    geom_col(position = position_fill()) +
-    geom_hline(yintercept = 0.5, colour = "black", size = 0.5) +
-    labs(
-        x = "Age (months)",
-        y = "% responses", 
-        colour = "Response",
-        fill = "Response"
-    ) +
-    scale_fill_d3() +
-    scale_colour_d3() +
-    scale_x_continuous(breaks = seq(0, 40, 2)) +
-    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-    theme(
-        legend.position = "top",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid.minor.x = element_blank(),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-## Model selection
-
-We compared the performance of these models using Bayesian leave-one-out cross-validation (LOO) using the `loo` and `loo_compare` functions of the `brms` R package (dependent of the `LOO` R package).
-LOO consists in computing the average likelihood of each observation after estimating the model's parameters leave that same observation out of the data set.
-Although the `loo` function uses a particular algorithm that speeds up the computation of this criterion (pareto-smooth importance sampling, PSIS), the size of our data set lead us to rely on the computation of the same criterion using a sampling approach via de `loo_subsample` function.
-
-```{r loos, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, eval = FALSE}
-loo_comp %>% 
-    as_tibble() %>% 
-    rownames_to_column("model") %>% 
-    relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
-    mutate_all(as.numeric) %>% 
-    gt() %>% 
-    tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
-    tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
-    tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
-    tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
-    fmt_number(3:13) %>% 
-    cols_label(
-        model = "Model",
-        # elpd
-        elpd_loo = md("*ELPD*"),
-        se_elpd_loo = md("*SE*"),
-        subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
-        # p
-        p_loo = md("*p*"),
-        se_p_loo = md("SE"),
-        subsampling_se_p_loo = md("SE<sub>sub</sub>"),
-        # looic
-        looic = md("*LOO-IC*"),
-        se_looic = md("*SE*"),
-        subsampling_se_looic = md("SE<sub>sub</sub>"),
-        # diff
-        elpd_diff = md("*diff*"),
-        se_diff = md("*SE*"),
-        subsampling_se_diff = md("*SE<sub>sub</sub>*"),
+    count(exposure, language) %>%
+    ggplot(aes(exposure, n, fill = language)) +
+    facet_wrap( ~ language) +
+    geom_col(colour = "white") +
+    labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
+         y = "Number of observations",
+         fill = "Language") +
+
+    df %>%
+    mutate(
+        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
+        exposure = cut(
+            exposure,
+            breaks = seq(0, 9, 1),
+            include.lowest = TRUE
+        ),
+        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
     ) %>%
-    as_raw_html()
-```
-
-## Fixed effects
-
-```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
-str_repl <- c(
-    "age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein",
-    "age_std:doe_std" = "Age \u00d7 DoE",
-    "age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "doe_std:lv_std" = "DoE \u00d7 Levenshtein",
-    "Intercept" = "Intercept",
-    "age_std" = "Age (+1 month)",
-    "freq_std" = "Frequency (+1 SD)",
-    "n_phon_std" = "# Phonemes (+1 SD)",
-    "doe_std" = "DoE (+10%)",
-    "lv_std" = "Levenshtein (+1 SD)"
-)
-
-draws_fix <- fixef(model_fit_7) %>% 
-    as.data.frame() %>%
-    rownames_to_column("term") %>% 
-    clean_names() %>% 
-    select(-est_error) %>% 
-    mutate(
-        estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
-        q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
-        q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
-        term = str_replace_all(term, str_repl)
-    )
-
-gt(draws_fix) %>% 
-    fmt_percent(2) %>% 
-    fmt_number(3:4, decimals = 2) %>% 
-    cols_merge(vars("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
-    cols_label(
-        term = md("**Predictor**"),
-        estimate = md("**Mean**"),
-        q2_5 = md("**95\\% CrI**"),
-    ) %>% 
-    tab_footnote(
-        footnote = "Transformed using the inverse logit to get the average probability of correct response",
-        locations = cells_body(columns = "term", rows = term=="Intercept")
-    ) %>% 
-    tab_footnote(
-        footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-        locations = cells_body(columns = "term", rows = term %in% c("Age (+1 month)", "Frequency (+ 1 SD)",  "DoE (+10%)", "Levenshtein (+ 1 SD)", "doe_std:lv_std" = "DoE \u00d7 Levenshtein"))
-    ) %>% 
-    cols_align(
-        align = c("center"),
-        columns = 2:3
-    )
-
-```
-
-<br>
-
-```{r fix_effects-plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
-The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-
-var_repl <- c(
-    "b_Intercept[1]" = "Intercept (Comprehension)",
-    "b_Intercept[2]" = "Intercept (Production)",
-    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein", 
-    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:doe_std" = "Age \u00d7 DoE",
-    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein", 
-    "b_Intercept" = "Intercept", 
-    "b_age_std" = "Age (+1 month)", 
-    "b_freq_std" = "Frequency (+1 SD)",
-    "b_n_phon_std" = "# Phonemes (+1 SD)",
-    "b_doe_std" = "DoE (+1 SD)", 
-    "b_lv_std" = "Levenshtein (+1 SD)"
-)
-
-post <- gather_draws(model_fit_7, `b_.*`, regex = TRUE) %>% 
-    mutate(
-        .value = ifelse(grepl("Intercept|sd", .variable), inv_logit_scaled(.value), .value/4),
-        .variable_name = factor(.variable, levels = names(var_repl), labels = var_repl)
-    )
-
-post_intercept <- post %>% 
-    filter(.variable %in% c("b_Intercept[1]", "b_Intercept[2]")) %>% 
-    ggplot(aes(.value, .variable_name)) +
-    geom_vline(xintercept = 0, size = 0.75, colour = "grey") +
-    stat_slab(
-        aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
-        size = 0.5,
-        color = "#FF0000",
-        fill = "#FF0000",
-        position = position_nudge(y = 0.2)
-    ) +
-    stat_pointinterval(point_size = 2) +
-    labs(
-        x = "Posterior probability of acquisition", 
-        y = "Variable", 
-        fill = "Variable",
-        alpha = "CrI"
-    ) +
-    scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-    scale_x_continuous(labels = percent, limits = c(0, 1))
-
-post_fix <- post %>% 
-    filter(!grepl("Intercept|sd", .variable)) %>% 
-    ggplot(aes(.value, .variable_name)) +
-    geom_vline(xintercept = 0, size = 0.5, linetype = "dashed") +
-    stat_slab(
-        aes(alpha = stat(cut_cdf_qi(cdf, .width = c(.95, .8, .5), labels = percent_format()))),
-        size = 0.5,
-        color = "#00A08A",
-        fill = "#00A08A",
-        position = position_nudge(y = 0.1)
-    ) +
-    stat_pointinterval(point_size = 2) +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "Variable", 
-        fill = "Variable",
-        alpha = "CrI"
-    ) +
-    scale_alpha_discrete(range = c(1, 0.4), na.translate = FALSE) +
-    scale_x_continuous(labels = percent, breaks = seq(-0.2, 2.5, 0.1))
-
-(post_intercept / post_fix) +
-    plot_layout(heights = c(0.2, 0.8)) &
-    theme(
-        legend.position = "top",
-        legend.direction = "horizontal",
-        legend.justification = "right",
-        plot.caption.position = "plot",
-        plot.caption = element_text(hjust = 0),
-        axis.title.y = element_blank(),
-        # axis.text = element_text(size = 7),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.major.y = element_blank()
-    )
-```
-
-## Random effects {.tabset .tab-pils}
-
-### Participant
-
-```{r random-participant, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-re_id <- ranef(model_fit_7)$id[,,1] %>% 
-    as.data.frame() %>% 
-    rownames_to_column("id") %>% 
-    as_tibble() %>% 
-    mutate_if(is.numeric, ~inv_logit_scaled(. + fixef(model_fit_7)[1])) %>% 
-    left_join(distinct(model_fit_7$data, id)) %>% 
-    clean_names()
-
-ggplot(re_id, aes(estimate, reorder(id, estimate), xmin = q2_5, xmax = q97_5)) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    # geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, height = 0) +
-    # geom_point(size = 0.5) +
-    geom_vline(xintercept = 0.5, colour = "white") +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "Participant", 
-        colour = "Lang. Prof."
-    ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-### Item
-
-```{r re-item, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-re_te <- ranef(model_fit_7)$te[,,1] %>% 
-    as.data.frame() %>% 
-    rownames_to_column("te") %>% 
-    as_tibble() %>% 
-    mutate_if(is.numeric, ~inv_logit_scaled(. + fixef(model_fit_7)[1])) %>% 
-    mutate(te = as.numeric(te)) %>% 
-    clean_names()
-
-ggplot(re_te, aes(estimate, reorder(te, estimate), xmin = q2_5, xmax = q97_5)) +
-    geom_ribbon(aes(group = 1), alpha = 0.5) +
-    geom_line(aes(group = 1), size = 1) +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "TE", 
-        colour = "Lang. Prof."
-    ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-
-```
-
-### {-}
-
-## Marginal means
-
-### Marginal means: Comprehension
-
-```{r emmeans_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+    count(exposure, freq, language) %>%
+    ggplot(aes(
+        x = freq,
+        y = exposure,
+        colour = language,
+        size = n
+    )) +
+    facet_wrap( ~ language) +
+    geom_point() +
+    labs(
+        x = "Frequency (Zipf score)",
+        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
+        colour = "Language",
+        fill = "Language",
+        size = "Number of observations"
+    ) +
+    theme(panel.grid = element_blank()) +
+    plot_layout(ncol = 1) &
+    scale_colour_manual(values = clrs[c(1, 4)]) &
+    scale_fill_manual(values = clrs[c(1, 4)]) &
+    guides(fill = "none", colour = "none") &
+    theme(panel.grid.major.x = element_blank(),
+          panel.grid.minor.y = element_blank())
+```
+
+## Data analysis
+
+We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function.
+The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood (see @fig-likelihood for a graphical representation).
+
+$$
+\begin{aligned}
+\textbf{Likelihood:} \\
+y_{ij} &\sim \text{Cumulative}(p_{k})
+\end{aligned}
+$$ {#eq-likelihood}
+
+where:
+
+- $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$)
+- $i$ is the participant index
+- $j$ is the translation equivalent (TE) index
+- $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$)) between two response categories in the latent distribution
+$p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.
+
+```{r}
+#| label: fig-likelihood
+#| fig-cap: "Graphical representation of the assumed latent distribution that generates observed responses in the questionnaire.
+Cumulative ordinal regression assumes that observed categorical responses (*No*, *Understands* and *Understands and Says* are generated from a common latent (i.e., unobserved) distribution.
+For threotical reasons, we refer to the dimension captured by such distribution as *Familiarity*.
+*Familiarity* describes a given participant's familiarity with a given word.
+When familiarity is low (below some threshold 1), the reulting response is *No*.
+When familiarity is higher than such threshold 1, but lower than some threshold 2, the resulting result is *Understands* (i.e., comprehension).
+When familiarity exceeds threshold 2, the resulting response is *Understands and Says* (i.e., comprehension and production).
+These thresholds are estimated in the cumulative regression model as intercepts 1, and 2.
+The rest of the coefficients in the regression model shift the response thresholds, making comprehension or production more likely or less likely.
+See @burkner2019ordinal for a detailed description of cumulative ordinal regression and its implementation."
+#| fig-width: 7
+#| fig-height: 3.5
+dist_df = tribble(~mean, ~sd, 0,   1)
+
+responses <- data.frame(
+    ymin = c(-Inf, -1.5, 1.5),
+    ymax = c(-1.5, 1.5, Inf),
+    xmin = -Inf,
+    xmax = Inf,
+    .category = str_wrap(c("No", "Understands", "Understands and Says"), width = 12)
+)
+
+responses %>%
+    ggplot(aes(ymin = ymin, ymax = ymax, xmin = xmin, xmax = xmax, fill = .category)) +
+    geom_rect(alpha = 0.5) +
+    stat_dist_slab(
+        data = dist_df, 
+        aes(dist = "norm", arg1 = mean, arg2 = sd),
+        position = "dodge", inherit.aes = FALSE,
+        fill = NA, colour = "black"
+    ) +
+    geom_hline(yintercept = c(-1.5, 1.5), linetype = "dotted", size = 1) +
+    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
+              vjust = 1, hjust = 0.5) +
+    annotate(
+        geom = "text",
+        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
+        y = c(-1.5, 2), parse = TRUE,
+        x = 0.5, angle = 90, vjust = -0.5
+    ) +
+    labs(y = "Hypothetical response-generating latent distribution",
+         x = "Response probability",
+         fill = "Response type",
+         colour = "Response type") +
+    coord_flip() +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+    scale_x_continuous(labels = percent) +
+    scale_y_continuous(limits = c(-4, 4)) +
+    theme(panel.grid = element_blank(),
+          legend.position = "none")
+```
+
+To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
+We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset--each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
+For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].
+
+$$
+\begin{aligned}
+\textbf{Linear model:} \\
+logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
+& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
+& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Phonemes}_{ij} + & \\
+& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
+& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
+& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
+& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
+& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
+\end{aligned}
+$$ {#eq-linear}
+
+where:
+
+- $i$ and $j$ index the participant and translation equivalent (TE)
+- $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$
+- $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively
+- $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest
+- $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively
+
+We used the Bayesian framework to estimate the parameters in our model.
+This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical].
+This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations.
+We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].
+
+$$
+\begin{aligned}
+\\
+\textbf{Prior:} \\
+\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
+\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
+\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
+u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
+w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
+&&\mbox{[Participant-level coefficient variability]} \\ \\
+\Bigg(\begin{smallmatrix}
+u_{k_{0}} \\ 
+u_{1_{i}} \\ 
+\vdots \\ 
+u_{8_{i}} 
+\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
+\Bigg(\Bigg(\begin{smallmatrix}0 \\
+0 \\ 
+\vdots \\
+0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
+\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
+\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
+\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
+\vdots & \vdots & \vdots & \vdots \\
+\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
+\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
+\rho_{u} &\sim LKJcorr(2) \\
+\\
+&&\mbox{[TE-level coefficient variability]} \\ \\
+\Bigg(\begin{smallmatrix}
+w_{k_{0}}\\ 
+w_{1_{j}} \\ 
+\vdots \\ 
+w_{3_{j}} 
+\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
+0\\ 
+0 \\ 
+\vdots \\
+0 
+\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
+\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
+\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
+\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
+\vdots & \vdots & \vdots & \vdots \\
+\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
+\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
+\rho_{w_{0-3}} &\sim LKJcorr(2)
+\end{aligned}
+$$ {#eq-prior}
+
+where:
+
+- $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively
+- $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively
+- $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].
+
+::: {.panel-tabset}
+
+#### Prior-predicted mean
+
+```{r}
+#| label: fig-prior-mean
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Expected prior predictions"
 nd <- expand.grid(
     n_phon_std = 0,
-    freq_std = 0,
-    age_std = seq(-4, 4, length.out = 30),
-    doe_std = range(df$doe_std),
-    lv_std = c(range(df$lv_std), 0)
-)
-
-m <- add_epred_draws(nd, model_fit_7, ndraws = 50, re_formula = NA) %>% 
-    mutate(
-        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
-        doe_std = factor(
-            doe_std, 
-            levels = unique(nd$doe_std), 
-            labels = c("DoE: 0%", "DoE: 100%")
-        ),
-        lv_std = factor(
-            lv_std, 
-            levels = unique(nd$lv_std)[c(1, 3, 2)],
-            labels = c("0%", "50%", "100%")
-        )
-    ) %>% 
+    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    exposure_std = 0,
+    lv_std = 0
+)
+
+m <- add_epred_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA) %>% 
     filter(.category != "No") %>% 
@@
         values_to = ".epred"
-    )
-
-ggplot(m, aes(age_std, .epred, colour = lv_std, fill = lv_std)) +
-    facet_grid(.category~doe_std) +
-    geom_line(
+    ) %>% 
+    mutate(
+        .category = factor(
+            .category,
+            levels = c("Understands and Says", "Understands"),
+            labels = c("Production\n(Understands and Says))", "Comprehension\n(Understands)")
+        )
+    )
+
+m %>% 
+    ggplot(aes(x = age_std, y = .epred)) +
+    facet_wrap(~.category) +
+    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
+    stat_summary(fun = "mean", geom = "line", size = 0.75) +
+    scale_x_continuous(
+        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 36, 4)
+    ) +
+    scale_fill_manual(
+        values = rev(clrs), 
+        labels = c("95%", "89%", "67%", "50%")
+    ) +
+    labs(
+        x = "Age (months)", 
+        y = "Probability of acquisition",
+        colour = "Credible interval", 
+        fill = "Credible interval"
+    ) +
+    scale_y_continuous(labels = percent, limits = c(0, 1)) +
+    theme(
+        legend.position = "top",
+        axis.ticks.x = element_line(colour = "black"),
+        panel.grid = element_blank()
+    )
+```
+
+#### Prior-predictions
+
+```{r}
+#| label: fig-prior-predictions
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Expected prior predictions"
+nd <- expand.grid(
+    n_phon_std = 0,
+    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    exposure_std = 0,
+    lv_std = 0
+)
+
+m <- add_predicted_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA)
+
+m %>% 
+    count(age_std, .prediction) %>% 
+    pivot_wider(names_from = .prediction, values_from = n) %>% 
+    mutate(Understands = Understands + `Understands and Says`) %>% 
+    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
+    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
+    geom_col(position = position_fill()) +
+    labs(
+        x = "Age (months)", 
+        y = "Probability of acquisition",
+        colour = "Category", 
+        fill = "Category"
+    ) +
+    scale_x_continuous(
+        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 36, 4)
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+    scale_y_continuous(labels = percent) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        axis.ticks.x = element_line(colour = "black"),
+        panel.grid = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
+        strip.background = element_rect(fill = "grey", colour = NA),
+        plot.background = element_rect(fill = "white", colour = NA)
+    )
+```
+
+:::
+
+### Model comparison
+
+To test of goodness of fit of our model, we compared its predictive accuracy of this extended model--which included all predictors of interest--to four simpler models that included a subset of our predictors of interest: a base model that only included the main effects of $Age$, $Phonemes$, and $Exposure$ ($\mathcal{M}_{0}$), a model that also included the $Age \times Exposure$ interaction ($\mathcal{M}_{1}$), a model that also included the main effect of $Levenshtein$ ($\mathcal{M}_{2}$), a model that also included the $Age \times Levenshtein$ interaction ($\mathcal{M}_{3}$).
+Only the extended model included the $Age \times Exposure \times Levenshtein$ interaction.
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package.
+LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
+Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
+
+### Model settings
+
+-   OS: `r sessionInfo()$running`, `r sessionInfo()$platform`
+-   R version: `r sessionInfo()$R.version$version.string`
+-   Algorithm: Hamiltonian Monte Carlo (No U-turn Sampler)
+-   Engine: Stan (`brms` interface, `r sessionInfo()$otherPkgs$brms$Version`)
+-   `brms` backend: CmdStanR (`cmdstanr` `r sessionInfo()$loadedOnly$cmdstanr$Version`)
+-   Chains: `r dim(model_fit_4$fit)[2]`
+-   Cores: `r dim(model_fit_4$fit)[2]`
+-   Iterations: `r format(dim(model_fit_4$fit)[1], big.mark = ",")` (`r format(dim(model_fit_4$fit)[1]/2, big.mark = ",")`)
+
+More details in Appendix: Session Info.
+
+### R code (`brms`)
+
+::: panel-tabset
+
+#### Model 0
+
+```{r}
+print(model_fit_0$formula)
+```
+
+#### Model 1
+
+```{r}
+print(model_fit_1$formula)
+```
+
+#### Model 2
+
+```{r}
+print(model_fit_2$formula)
+```
+
+#### Model 3
+
+```{r}
+print(model_fit_3$formula)
+```
+
+#### Model 4
+
+```{r}
+print(model_fit_4$formula)
+```
+:::
+
+# Results
+
+## Descriptive statistics
+
+::: panel-tabset
+
+### By number of phonemes
+
+```{r}
+#| lab: fig-responses-age-phonemes
+#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
+df %>% 
+    mutate(
+        age = floor(age),
+        n_phon = cut_interval(n_phon, n = 4)
+    ) %>%
+    count(age, n_phon, response) %>% 
+    ggplot(aes(age, n, fill = response)) +
+    facet_wrap(~n_phon) +
+    geom_col(position = position_fill()) +
+    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+    scale_y_continuous(labels = percent) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+
+### By exposure
+
+```{r}
+#| lab: fig-responses-age-exposure
+#| fig-cap: "Proportion of responses to each category by age and exposure."
+df %>% 
+    mutate(
+        age = floor(age),
+        exposure = cut_interval(exposure, n = 4)
+    ) %>%
+    count(age, exposure, response) %>% 
+    ggplot(aes(age, n, fill = response)) +
+    facet_wrap(~exposure) +
+    geom_col(position = position_fill()) +
+    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+    scale_y_continuous(labels = percent) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+
+### By Levenshtein
+
+```{r}
+#| lab: fig-responses-age-lv
+#| fig-cap: "Proportion of responses to each category by age and Levenhstein similarity"
+df %>%
+    mutate(
+        age = floor(age),
+        exposure = cut_interval(exposure, n = 4, labels = paste0("Q", 1:4)),
+        lv = cut_interval(lv, n = 4)
+    ) %>%
+    group_by(te, lv, exposure, response) %>%
+    count(lv, exposure, response) %>%
+    pivot_wider(names_from = response,
+                values_from = n,
+                values_fill = 0) %>%
+    mutate(
+        `Understands` = `Understands` + `Understands and Says`,
+        n_total = sum(No, `Understands`, `Understands and Says`)
+    ) %>%
+    pivot_longer(
+        c(`No`, `Understands`, `Understands and Says`),
+        names_to = "response",
+        values_to = "n"
+    ) %>%
+    filter(response != "No") %>%
+    mutate(prop = n / n_total) %>%
+    # group_by(age, lv, exposure, response) %>%
+    # summarise(prop_mean = mean(prop), .groups = "drop") %>%
+    ggplot(aes(prop, exposure, colour = lv, fill = lv)) +
+    facet_grid(~ response) +
+    stat_slab(alpha = 0.5, size = 0.2, trim = FALSE) +
+    geom_point(size = 1, alpha = 0.5, position = position_nudge(y = -0.1)) +
+    labs(x = "Proportion of responses", 
+         y = "Exposure",
+         fill = "Levenshtein", 
+         colour = "Levenshtein") +
+    scale_x_continuous(labels = percent) +
+    scale_colour_manual(values = clrs[c(1, 4, 3, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 3, 5)]) +
+
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+:::
+
+## Model selection
+
+@tbl-results-loos shows the outputs of the *LOO-CV*.
+The extended model (`model_fit_3`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference.
+This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.
+
+```{r}
+#| label: tbl-results-loos
+#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
+This procedure involves estimating the likelihood of each observation given a model fitted without such observation.
+The resulting likelihoods of each model were summed to compute the expected log-predicted density (ELPD) of the model, and correcting it for the number of parameters in the model to penalised for its the complexity.
+ELPD values closer to zero indicate better predictive acuracy.
+We report the ELPD, effective number of parameters (p), and information criterion (LOO-IC) of each model along their associated standard errors in paratheses.
+The first column indicates the difference in ELPD between each model and the model with the best predictive accuracy (model 4)."
+model_loos %>% 
+    loo_compare() %>% 
+    as.data.frame() %>%   
+    rownames_to_column("model") %>% 
+    relocate(
+        model, 
+        matches("diff"),
+        matches("elpd_loo"),
+        matches("p_loo"), 
+        matches("looic")
+    ) %>%
+    mutate(model = str_extract_all(model, "\\d")) %>% 
+    gt() 
+# fmt_number(2:9, decimals = 1) %>% 
+# cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>% 
+# cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>% 
+# cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>% 
+# cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>% 
+# cols_label(
+#     model = "Model",
+#     elpd_loo = md("*ELPD*"),
+#     p_loo = md("*p*"),
+#     looic = md("*LOO-IC*"),
+#     elpd_diff = md("Difference")
+# ) %>% 
+# tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
+# tab_style(
+#     cell_text(weight = "bold"),
+#     cells_column_labels()
+# ) %>% 
+# tab_style(
+#     cell_fill(color = "grey", alpha = 0.5),
+#     cells_body(matches("diff"))
+# ) %>% 
+# tab_style(
+#     cell_fill(color = clrs[4], alpha = 0.5),
+#     cells_body(rows = 1)
+# )
+```
+
+## Fixed effects
+
+```{r}
+#| label: results-posterior-fixed
+post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)
+
+# tidy predictor names
+str_repl <- c(
+    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
+    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
+    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
+    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
+    "b_exposure_std" = paste0("Exposure (+1 SD, ",  percent(sd(df$exposure)), ")"),
+    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
+    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
+    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
+    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
+    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
+)
+```
+
+```{r}
+#| label: tbl-results-fixed
+#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
+#| fig-width: 7
+intercept_footnote <- paste0(
+    "Probability of comprehension or production at the grand age of the sample (",
+    round(mean(df$age), 2),
+    " months)"
+)
+
+# summarise posterior draws
+posterior_description %>%
+    mutate(parameter = factor(parameter, levels = names(str_repl), labels = str_repl)) %>%
+    select(parameter, median, ci_high, ci_low, rope_percentage) %>%
+    gt(rowname_col = "parameter") %>%
+    fmt_number(c(2:4), decimals = 2) %>%
+    fmt_percent(5, decimals = 2) %>%
+    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
+    cols_label(median = "Median",
+               ci_low = "95% HDI",
+               rope_percentage = "ROPE prob.") %>%
+    tab_style(cell_text(weight = "bold"),
+              cells_column_labels(columns = 1:5)) %>%
+    tab_style(cell_text(align = "left"),
+              cells_stub())
+
+post_draws_list <-
+    split(posterior_description, posterior_description$parameter)
+```
+
+We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
+We decided to draw this interval around zero so that it covered values smaller than those considered as negligible by similar previous studies: after consulting the regression tables of [ADD STUDIES], we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: 3%.
+consequently, we defined our *ROPE* from -3% to +3%.
+We report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
+This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
+Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the -3%--3% *ROPE* interval we previously defined.
+For those coefficients that did not exclude completely the *ROPE* we also report a Bayes Factor indicating how many times is such coefficient more likely to be bigger than the upper limit of the *ROPE*, or smaller than the lower limit of the *ROPE*, given our data.
+
+For interpretability, we report the estimated regression coefficients transformed from the probability scale.
+The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
+This means that one logit difference is not necessarily translated to a unique value in the probability scale.
+For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
+The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
+We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
+derivative) of the logistic curve at the age at which most participants were acquiring a given word.
+This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
+
+@fig-results-fixed shows the outputs of the model.
+**Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$median, 2)`, 95% *HDI* = [`r round(post_draws_list[["b_age_std"]]$ci_low, 2)`, `r round(post_draws_list[["b_age_std"]]$ci_high, 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month.
+
+The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]).
+For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.
+
+**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]).
+All of the posterior samples of his regression excluded the *ROPE*.
+The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/(sd(df$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.
+
+The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
+
+However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word.
+Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
+This effect was negligible when exposure to both languages was balanced (e.g.
+50-50%).
+Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.
+
+The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
+Given our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.
+
+The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
+
+Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
+
+```{r}
+#| label: fig-results-fixed
+#| fig-width: 9
+#| fig-heigh: 7
+#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model.
+Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis).
+Intervals show the 95% highest density interval (*HDI*) of each distribution.
+This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
+the mean and HDI limits are indicated below each distribution.
+Coefficients were transformed from the logit scale to the probability scale for interpretability.
+We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e.
+the derivative of the logistic function in its mid-point."
+post <- post_draws %>% 
+    mutate(
+        .variable_name = factor(
+            .variable, 
+            levels = names(str_repl),
+            labels = str_repl,
+            ordered = TRUE
+        )
+    )
+
+post_summary <- post %>% 
+    filter(!grepl("Intercept|sd", .variable)) %>%
+    median_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>% 
+    mutate(
+        .variable_name = factor(
+            .variable, 
+            levels = names(str_repl),
+            labels = str_repl,
+            ordered = TRUE
+        )
+    )
+
+post %>%
+    filter(!grepl("Intercept|sd", .variable)) %>%
+    ggplot(aes(.value, fct_rev(.variable_name))) +
+    annotate(geom = "rect", ymin = -Inf, ymax = Inf,
+             xmin = rope_interval["lower"], xmax = rope_interval["upper"],
+             colour = NA, alpha = 0.2, fill = clrs[5],) +
+    geom_vline(xintercept = 0, size = 1, colour = clrs[5]) +
+    stat_slab(
+        aes(fill = stat(abs(x) < rope_interval["upper"]),
+            colour = stat(abs(x) < rope_interval["upper"]),),
+        size = 1, position = position_nudge(y = 0.15)
+    ) +
+    geom_errorbar(
+        data = post_summary,
+        aes(xmin = .lower, xmax = .upper, x = .value),
+        width = 0.15
+    ) +
+    geom_point(data = post_summary, aes(x = .value), size = 1.5) +
+    geom_text(
+        data = post_summary,
         aes(
-            group = interaction(.draw, lv_std, doe_std), 
-            colour = lv_std),
-        alpha = 0.1, 
-        size = 1
-    ) +
-    # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
-    stat_summary(fun = "mean", geom = "line", size = 1) + 
-    geom_hline(yintercept = 0.5, size = 0.5, linetype = "dashed") +
-    scale_color_d3() +
+            x = .value, 
+            y = .variable_name,
+            label = paste0(
+                round(.value, 1),
+                " [", round(.lower, 1), ", ", round(.upper, 1), "]")
+        ),
+        colour = "black", position = position_nudge(y = -0.3), size = 3
+    ) +
+    labs(
+        x = "Logistic regression coefficient estimate (Logit scale)",
+        y = "Variable", 
+        fill = "Overlaps with ROPE",
+        colour = "Overlaps with ROPE"
+    ) +
+    scale_x_continuous(breaks = seq(-2, 2, 0.2)) +
+    scale_fill_manual(
+        values = clrs[c(1, 4)],
+        labels = c("No", "Yes")
+    ) +
+    scale_colour_manual(
+        values = clrs[c(1, 4)],
+        labels = c("No", "Yes")
+    ) +
+    theme(
+        legend.position = "top",
+        legend.justification = c(1, 1),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(
+            colour = "grey85",
+            linetype = "dotted"
+        ),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.x = element_blank()
+    )
+```
+
+## Marginal predictions
+
+```{r}
+#| label: fig-results-marginal
+#| fig-width: 7
+#| fig-cap: "Expected posterior predictions.
+Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
+These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model.
+Therefore, we generated 4,000 predictions for a series of combination of levels of interest.
+Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels.
+The X-axis indicates the age (in months) for which the prediction is generated.
+The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*).
+Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
+Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
+Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
+
+nd <- datagrid(
+    model = model_fit_4,
+    age_std = seq_range(model_fit_4$data$age_std, n = 100),
+    lv_std = (seq(0, 1, 0.5)-mean(df$lv))/(sd(df$lv)),
+    exposure_std = c(-1, 0, 1),
+    te = NA,
+    id = NA
+)
+
+marg_epreds <- predictions(
+    model_fit_4,
+    newdata = nd,
+    ndraws = 25, 
+    re_formula = NA
+)
+
+marg_epreds %>% 
+    posteriordraws() %>% 
+    as_tibble() %>% 
+    select(drawid, group, draw, age_std, lv_std, exposure_std) %>%
+    filter(group != "No") %>% 
+    pivot_wider(
+        id_cols = c(drawid, age_std, lv_std, exposure_std), 
+        names_from = group, 
+        values_from = draw
+    ) %>% 
+    mutate(`Understands` = `Understands and Says` + `Understands`) %>% 
+    pivot_longer(c(`Understands`, `Understands and Says`), names_to = "group", values_to = "draw") %>% 
+    mutate(
+        exposure_std = factor(
+            exposure_std, 
+            levels = unique(nd$exposure_std), 
+            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
+        ),
+        lv_std = factor(
+            lv_std,
+            levels = unique(nd$lv_std), 
+            labels = paste0(c(0, 50, 100), "% similarity")
+        )
+    ) %>% 
+    ggplot(aes(age_std, draw, colour = lv_std, fill = lv_std)) +
+    facet_grid(group~exposure_std) +
+    # stat_summary(fun.data = median_hdi, geom = "ribbon",
+    #              alpha = 0.5, colour = NA) +
+    geom_line(aes(group = interaction(lv_std, drawid)),
+              size = 0.75, alpha = 0.15) +
+    stat_summary(aes(linetype = lv_std), fun = median, geom = "line", size = 1) +
     labs(
         x = "Age (months)",
-        y = "Posterior probability of acquisition\n(Linear prediction)",
-        colour = "Phonological similarity (Levenshtein distance)", 
-        fill = "Phonological similarity (Levenshtein distance)"
-    ) +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
-    theme(
-        legend.position = "top"
+        y = "Probability of acquisition",
+        linetype = "Phonological similarity (Levenshtein)",
+        colour = "Phonological similarity (Levenshtein)",
+        fill = "Phonological similarity (Levenshtein)"
+    ) +
+    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+    scale_y_continuous(
+        breaks = seq(0, 1, 0.25),
+        limits = c(0, 1),
+        labels = percent
+    ) +
+    scale_x_continuous(
+        breaks = (seq(10, 32, 4)-mean(df$age))/(sd(df$age)),
+        labels = seq(10, 32, 4)
+    ) +
+    theme(
+        legend.position = "top",
+        panel.grid = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
+    )
+
+```
+
+## Marginal effects
+
+In order to examine the impact of the word exposure index on the effect of phonological similarity we used the [`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects/index.html) R package to compute the average marginal effect of the Levenshtein distance on the probability of acquisition (comprehension or production) for the range of observed values of the word exposure index in our dataset.
+
+```{r}
+#| label: fig-marginal-effect
+#| fig-cap: "Average marginal effect of phonological similarity (Levenshtein distance) on the probability of comprehension and production across the range of values of exposure.
+The observed range of values of the word exposure index is indicated in the X-axis, and the estimated average marginal effect of phonological similarity (Levenshtein distance) is indicated in the Y-axis.
+This quantity is generated by averaging the derivative of the regression coefficient of the predictor Levenshtein across the range of values of all predictors but Exposure.
+We represent the uncertainty associated with the resulting estimate by generating multiple values of this marginal effect by drawing multiple samples from the posterior distribution of the model coefficients, calculating the average marginal effect for each posterior draw, and plotting the result as a thin grey line.
+Closer lines indicate lower uncertainty over the true value of the average marginal effect of Levenshein, while more distant lines indicate higher uncertainty."
+nd <- datagrid(
+    model = model_fit_4,
+    n_phon_std = 0,
+    lv_std = 0,
+    exposure_std = seq(-1.5, 1.5, length.out = 100),
+    age_std = 0,
+    id = NA,
+    te = NA
+)
+
+marg_effects <- marginaleffects(
+    model_fit_4,
+    newdata = nd, 
+    re_formula = NA,
+    ndraws = 50,
+    vcov = FALSE,
+    type = "link"
+)
+
+marg_effects %>% 
+    posteriordraws() %>%
+    as_tibble() %>% 
+    filter(term=="lv_std") %>% 
+    mutate(draw = draw/4) %>% 
+    ggplot(aes(exposure_std, draw)) +
+    geom_hline(yintercept = 0, size = 1, colour = "grey") +
+    geom_line(aes(group = drawid), alpha = 0.25, size = 0.75, colour = clrs[4]) +
+    # stat_summary(fun.data = median_hdi, geom = "ribbon",
+    #              colour = NA, alpha = 0.5, fill = clrs[1]) +
+    stat_summary(fun = median, geom = "line", size = 1, color = clrs[1]) +
+    labs(
+        x = "Word exposure index",
+        y = "Marginal effect of phonological similarity\n(Levenshtein distance)"
+    ) +
+    scale_x_continuous(
+        labels = label_number(style_positive = "plus", suffix = " SD"),
+        breaks = seq(-2, 2, 0.5)
+    ) +
+    scale_y_continuous(
+        labels = label_percent(style_positive = "plus", ),
+        breaks = seq(-1, 1, 0.05)
+    ) +
+    theme(
+        legend.position = "top",
+        panel.grid = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
+    )
+
+```
+
+## Random effects
+
+### Participant-level effects
+
+::: panel-tabset
+
+#### Intercept
+
+```{r}
+#| label: fig-results-random-id-intercept
+#| fig-width: 7
+#| fig-cap: "Posterior probability of acquisition by participant"
+re_id <- ranef(model_fit_4)$id[,,1] %>% 
+    as_tibble() %>% 
+    rownames_to_column("id") %>% 
+    left_join(distinct(model_fit_4$data, id)) %>% 
+    clean_names() %>% 
+    mutate(
+        across(
+            .cols = where(is.numeric),
+            .fns = list(
+                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
+                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
+            ), .names = "{.col}__{.fn}"
+        )
+    ) %>% 
+    select(-c(estimate:q97_5)) %>% 
+    pivot_longer(
+        where(is.numeric), 
+        names_to = c(".value", "type"),
+        names_sep = "__"
+    )
+
+re_id %>% 
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate), 
+        xmin = q2_5, xmax = q97_5,
+        fill = type, colour = type) +
+    facet_wrap(~type) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior probability of acquisition",
+        y = "Participant", 
+        colour = "Type"
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age slope
+
+```{r}
+#| label: fig-results-random-id-age
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_age <- ranef(model_fit_4)$id[,,"age_std"]
+
+re_age %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_4$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### # Phonemes slope
+
+```{r}
+#| label: fig-results-random-id-nphon
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the number of phonemes by participant"
+re_n_phon <- ranef(model_fit_4)$id[,,"n_phon_std"]
+
+re_n_phon %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_4$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(
+        x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5
+    ) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Exposure slope
+
+```{r}
+#| label: fig-results-random-id-exposure
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of lexical frequency by participant"
+re_freq <- ranef(model_fit_4)$id[,,"exposure_std"]
+
+re_freq %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(id = unique(model_fit_4$data$id)) %>% 
+    relocate(id) %>% 
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-lv
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of phonological similarity by participant"
+re_lv <- ranef(model_fit_4)$id[,,"lv_std"]
+
+re_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Doe-by-Levenshtein slope
+
+```{r}
+# | label: fig-results-random-id-exposure-lv
+# | echo: true
+# | message: false
+# | warning: false
+# | fig-width: 7
+# | fig-cap: "Posterior random slope of age by participant"
+re_exposure_lv <- ranef(model_fit_4)$id[,,"exposure_std:lv_std"]
+
+re_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-DoE slope
+
+```{r}
+#| label: fig-results-random-id-age-doe
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
+re_age_exposure <- ranef(model_fit_4)$id[,,"age_std:exposure_std"]
+
+re_age_exposure %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-age-lv
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
+re_age_lv <- ranef(model_fit_4)$id[,,"age_std:lv_std"]
+
+re_age_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_4$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-DoE-by-Levenshtein slope
+
+```{r}
+#| label: fig-results-random-id-age-exposure-lv
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
+re_age_exposure_lv <- ranef(model_fit_4)$id[,,"age_std:exposure_std:lv_std"]
+
+re_age_exposure_lv %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(id = unique(model_fit_3$data$id)) %>%
+    relocate(id) %>%
+    ggplot() +
+    aes(x = estimate, y = reorder(id, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Correlations
+
+```{r}
+#| label: fig-results-random-id-correlations
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 7
+#| fig-cap: "Posterior correlations between random effects by participant"
+posterior_description_re %>% 
+    select(parameter, median) %>% 
+    filter(str_detect(parameter, "cor_id__")) %>% 
+    mutate(parameter = str_remove_all(parameter, "cor_id__")) %>%  
+    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
+    mutate(
+        across(
+            c(var_1, var_2), 
+            ~factor(
+                .,
+                levels = c("Intercept", "age_std", "exposure_std", "lv_std", "n_phon_std", "age_std:exposure_std", "age_std:lv_std", "exposure_std:lv_std", "age_std:exposure_std:lv_std"),
+                labels = c("Intercept", "Age", "Exposure","Levenshtein", "Phonemes", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein", "Exposure \u00d7 Levenshtein", "Age \u00d7 Exposure \u00d7 Levenshtein")
+            )
+        )
+    ) %>%
+    filter(var_1 != var_2) %>% 
+    ggplot(aes(x = var_1, y = fct_rev(var_2), 
+               colour = median, label = percent(median, accuracy = 0.1))) +
+    geom_point(shape = 15, size = 20) +
+    geom_text(colour = "black") +
+    scale_colour_gradient2(low = clrs[1], 
+                           mid = "grey", 
+                           high = clrs[4],
+                           midpoint = 0) +
+    labs(colour = "Correlation", size = "Correlation") +
+    theme(
+        axis.title = element_blank(),
+        axis.text.x = element_text(angle = 35, hjust = 1),
+        panel.grid = element_blank()
+    )
+```
+:::
+
+### Item-level effects
+
+::: panel-tabset
+
+#### Intercept
+
+```{r}
+#| label: fig-results-random-te
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior probability of acquisition by translation equivalent"
+re_te <- ranef(model_fit_4)$te[,,"Intercept"] %>% 
+    as_tibble() %>% 
+    rownames_to_column("te") %>% 
+    clean_names() %>% 
+    mutate(
+        across(
+            .cols = where(is.numeric),
+            .fns = list(
+                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
+                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
+            ), .names = "{.col}__{.fn}"
+        )
+    ) %>% 
+    select(-c(estimate:q97_5)) %>% 
+    pivot_longer(
+        where(is.numeric), 
+        names_to = c(".value", "type"),
+        names_sep = "__"
+    ) %>% 
+    mutate(te = as.numeric(te))
+
+re_te %>% 
+    ggplot() +
+    aes(x = estimate, y = reorder(te, estimate), 
+        xmin = q2_5, xmax = q97_5,
+        colour = type, fill = type) +
+    facet_wrap(~type) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
+    geom_line(aes(group = 1), size = 1) +
+    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior probability of acquisition",
+        y = "TE", 
+        colour = "Type"
+    ) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age slope
+
+```{r}
+#| label: fig-results-random-te-age
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by translation equivalent"
+re_age <- ranef(model_fit_4)$te[,,"age_std"]
+
+re_age %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_4$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(x = estimate, y = reorder(te, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### # Phonemes slope
+
+```{r}
+#| label: fig-results-random-te-nphon
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
+re_n_phon <- ranef(model_fit_3)$te[,,"n_phon_std"]
+
+re_n_phon %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_3$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(x = estimate, y = reorder(te, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Exosure slope
+
+```{r}
+#| label: fig-results-random-te-exposure
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_doe <- ranef(model_fit_4)$te[,,"exposure_std"]
+
+re_doe %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(te = unique(model_fit_4$data$te)) %>%
+    relocate(te) %>%
+    ggplot() +
+    aes(x = estimate, y = reorder(te, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant",
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Age-by-exposure slope
+
+```{r}
+#| label: fig-results-random-te-age-exposure
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Posterior random slope of age by participant"
+re_age_doe <- ranef(model_fit_4)$te[,,"age_std:exposure_std"]
+
+re_age_doe %>% 
+    as_tibble() %>% 
+    clean_names() %>% 
+    mutate(te = unique(model_fit_4$data$te)) %>% 
+    relocate(te) %>% 
+    ggplot() +
+    aes(x = estimate, y = reorder(te, estimate),
+        xmin = q2_5, xmax = q97_5) +
+    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
+    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
+    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
+    labs(
+        x = "Posterior random slope estimate",
+        y = "Participant", 
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid = element_line(linetype = "dotted"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()
+    )
+```
+
+#### Correlations
+
+```{r}
+#| label: fig-results-random-te-correlations
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 7
+#| fig-cap: "Posterior correlations between random effects by translation equivalent"
+posterior_description_re %>% 
+    select(parameter, median) %>% 
+    filter(str_detect(parameter, "cor_te__")) %>% 
+    mutate(parameter = str_remove_all(parameter, "cor_te__")) %>%  
+    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
+    mutate(
+        across(
+            c(var_1, var_2), 
+            ~factor(
+                .,
+                levels = c("Intercept", "age_std", "exposure_std", "n_phon_std", "lv_std", "age_std:exposure_std", "age_std:lv_std"),
+                labels = c("Intercept", "Age", "Exposure", "Phonemes", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein")
+            )
+        )
+    ) %>%
+    filter(var_1 != var_2) %>% 
+    ggplot() +
+    aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1)) +
+    geom_point(shape = 15, size = 40) +
+    geom_text(colour = "black") +
+    scale_colour_gradient2(low = clrs[1], 
+                           mid = "grey", 
+                           high = clrs[4],
+                           midpoint = 0) +  
+    labs(
+        colour = "Correlation",
+        size = "Correlation"
+    ) +
+    theme(
+        axis.title = element_blank(),
+        panel.grid = element_blank()
+    )
+```
+:::
+
+## Model diagnostics
+
+### Simulation-based calibration
+
+```{r}
+#| label: fig-simulation-based-calibration
+#| fig-cap: "Posterior predictive checks.
+We simulated 500 datasets from our model by drawing random samples from its posterior distribution.
+We then computed the proportion of 'No', 'Understands', and 'Understands and Says' responses in each dataset.
+These simulated proportions are depicted in the plot as solid horizontal lines.
+Their variation indicate the uncertainy of the model about what proportion of each type of response would be observe in a newly collected dataset.
+We also plotted the proportion of each type of response in the observed dataset (with a ± 5% interval for comparison)."
+predicted_dfs <- add_predicted_draws(
+    object = model_fit_4, 
+    newdata = df,
+    re_formula = NULL,
+    ndraws = 500,
+    seed = 888
+)
+
+observed_prop <- df %>% 
+    count(response) %>%
+    mutate(prop = n/sum(n)) %>% 
+    rename(.prediction = response)
+
+predicted_dfs %>% 
+    ungroup() %>% 
+    select(.prediction, .draw) %>% 
+    count(.draw, .prediction) %>%
+    group_by(.draw) %>% 
+    mutate(
+        prop = n/sum(n),
+        .prediction = factor(
+            .prediction,
+            levels = levels(df$response), 
+            ordered = TRUE
+        )
+    ) %>% 
+    ggplot(aes(.draw, prop, fill = .prediction, colour = .prediction)) +
+    geom_line(aes(group = .prediction)) +
+    geom_rect(
+        data = observed_prop,
+        aes(xmin = 0, xmax = max(predicted_dfs$.draw),
+            ymin = prop-0.025, ymax = prop+0.025,
+            colour = .prediction, fill = .prediction),
+        alpha = 0.25, colour = NA, inherit.aes = FALSE
+    ) +
+    # geom_segment(
+    #     data = observed_prop,
+    #     aes(x = -Inf, xend = Inf, y = prop, yend = prop,
+    #         colour = .prediction),
+    #     size = 1, inherit.aes = FALSE
+    # ) +
+    labs(
+        x = "Iteration",
+        y = "Proportion of responses",
+        colour = "Response",
+        fill = "Response"
+    ) +
+    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+    scale_x_continuous(labels = label_number(big.mark = ","), 
+                       breaks = seq(0, max(predicted_dfs$.draw), 100)) +
+    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
+    theme(
+        legend.position = "top",
+        panel.grid = element_blank()
     )
@@
 
-```{r diagnostics_traceplots_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
-gather_draws(
-    model_fit_7, 
-    `b_.*`, `sd_te__.*`, `cor_te__.*`, 
-    regex = TRUE, 
-    ndraws = 1000
-) %>% 
+```{r}
+#| label: fig-diagnostics-traceplots
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 12
+gather_draws(model_fit_4, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
     mutate(.chain = paste("Chain ", .chain)) %>%
@@
     facet_wrap(~.variable, scales = "free_y") +
-    annotate(
-        geom = "rect", 
-        colour = NA, 
-        xmin = 0,
-        xmax = 500, 
-        ymin = -Inf, 
-        ymax = Inf, 
-        alpha = 0.25, 
-    ) +
+    annotate(geom = "rect", colour = NA,
+             xmin = 0, xmax = dim(model_fit_4$fit)[1]/2,
+             ymin = -Inf, ymax = Inf,
+             alpha = 0.25) +
     geom_line() +
@@
         legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-
-## {-}
+        panel.grid = element_blank(),
+        axis.text.x = element_text(size = 7)
+    )
+```
+
+### Leave-one-out diagnostics (Pareto-*k* diagnostics)
+
+::: panel-tabset
+
+#### Model 0
+
+```{r}
+#| label: fig-loo-diagnostics-0
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
+get_loo_plot <- function(x){
+    x$pointwise[,5] %>% 
+        data.frame(x = 1:length(.), y = .) %>% 
+        ggplot(aes(x, y)) +
+        geom_point(
+            size = 0.5, alpha = 0.1, colour = clrs[4], 
+            shape = 1
+        ) +
+        geom_hline(yintercept = 0.5, linetype = "dotted") +
+        labs(
+            x = "Data point",
+            y = "Pareto-k value",
+            title = "PSIS diagnostic plot"
+        ) +
+        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
+        scale_y_continuous(
+            limits = c(-0.6, 0.6)
+        )
+}
+
+get_loo_plot(model_loos$model_fit_0)
+```
+
+#### Model 1
+
+```{r}
+#| label: fig-loo-diagnostics-1
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
+get_loo_plot(model_loos$model_fit_1)
+```
+
+#### Model 2
+
+```{r}
+#| label: fig-loo-diagnostics-2
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
+get_loo_plot(model_loos$model_fit_2)
+```
+
+#### Model 3
+
+```{r}
+#| label: fig-loo-diagnostics-3
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
+get_loo_plot(model_loos$model_fit_3)
+```
+
+#### Model 4
+
+```{r}
+#| label: fig-loo-diagnostics-4
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
+get_loo_plot(model_loos$model_fit_4)
+```
+
+:::
+
+### Gelman-Rubin diagnostic (R-hat)
+
+```{r}
+#| label: fig-diagnostics-rhat
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| eval: false
+mcmc_rhat_data(model_rhats$model_fit_4) %>% 
+    ggplot() +
+    aes(x = value) +
+    geom_histogram(bins = 100, na.rm = TRUE, fill = clrs[1], colour = "white") +
+    geom_rug(colour = clrs[1]) +
+    labs(
+        x = "R-hat",
+        y = "Number of samples",
+        colour = "Description",
+        fill = "Description"
+    ) +
+    theme(
+        legend.position = "top"
+    )
+```
+
+### Effective sample size
+
+```{r}
+#| label: fig-diagnostics-neff
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| eval: false
+mcmc_neff_data(model_neffs$model_fit_4) %>% 
+    ggplot() +
+    aes(x = value, fill = description, colour = description) +
+    geom_histogram(bins = 100, na.rm = TRUE, colour = "white") +
+    geom_rug() +
+    geom_vline(xintercept = 1, colour = "black", size = 0.75) +
+    labs(
+        x = "Effective sample size ratio",
+        y = "Number of samples",
+        colour = "Rating",
+        fill = "Rating"
+    ) +
+    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
+    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
+    theme(
+        legend.position = "top"
+    )
+```
+
+# Discussion
+
+## Summary of the study
+
+## Summary of the results
+
+## Results in contrast with the literature
+
+## Mechanistic explanations
+
+## Limitations
+
+## Further steps
+
+# Appendix
+
+## Session info
+
+```{r}
+sessioninfo::session_info()
+```

### 2023-01-18 manuscript/manuscript.qmd
New draft version, extend discussion, update references, add appendix, change title, fix cross-references

@@
 ---
-title: "The role of cross-linguistic lexical similarity on bilingual word acquisition"
+title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual word acquisition"
 editor: source
@@
     pdf-engine: lualatex
-    linestretch: 1.5
     number-sections: true
     whitespace: small
+    linestretch: 1.5
 ---
@@
 These early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e.
-vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of acquired words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, @bates1994developmental].
-Despite this variability, children's trajectories of early vocabulary growth seem quite stable across languages.
+vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of acquired words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure; @bates1994developmental].
+Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
 @tardif2008baby collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing.
-Since birth, these infants had been learning English, Cantonese and Mandarin, respectively.
+Since birth, these infants had been learning English, Cantonese, and Mandarin, respectively.
 The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*).
@@
 
-To date, most studies on early word acquisition have relied exclusively on data from monolingual children while neglecting, to a considerable extent, that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent].
+Despite the recent increase in language representation in the literature, most studies on early word acquisition have relied exclusively on data from monolingual children while neglecting, to a considerable extent, that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent].
 Previous findings pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the two languages are taken into account.
@@
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
+Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
 While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
@@
 Depending on the linguistic distance between both languages, the two sets of words may overlap in form in varying degrees.
-When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic, respectively).
+When both languages are typologically close like Spanish and Catalan (both Romance languages) they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic, respectively).
 For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
@@
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
-Importantly, Spanish monolinguals--unfamiliar with the Catalan translations of the Spanish words they uttered--did not show this effect.
+Importantly, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, did not show this effect.
 These results suggest that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-Several subsequent studies have also provided similar evidence in comprehension in children [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
+
+Subsequent studies have also provided similar evidence in comprehension in children [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
 Parallel activation is therefore a plausible explanation for Floccia et al.'s results: cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
-This account is is line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words, already acquired or not [e.g., @hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
+This account is is line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
 Importantly, bilinguals seem to be more likely to acquire words for which their translation equivalent has been already learnt in the other language, suggesting that semantic similarity also facilitates word acquisition across languages [@bilson2015semantic].
 
-If phonological similarity also plays a cross-language facilitation role during word acquisition, cognate translation equivalents--which share both semantic and phonological similarity--should be acquired, on average, earlier that non-cognate translation equivalents--which share semantic similarity, but not phonological similarity.
+If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity, but not phonological similarity.
 Evidence supporting and earlier age of acquisition for cognates is, to date, scarce.
@@
 On the one hand, and in line with @mitchell2022cognates, the effect of cognateness might take place during early stages of word acquisition, before stable representations of any of the two word-forms of the translation equivalent has been formed.
-In this scenario, the auditory recognition any of the two word-forms might lead to the activation of its (cognate) translation in the other language via their phonological similarity.
+In this scenario, the auditory recognition of any of the two word-forms might lead to the activation of its (cognate) translation in the other language via their phonological similarity.
 Since that both word-forms are likely to be heard in the presence of the same referent, the association between one of the word-forms and its referent could be strengthened after the child is exposed to that same word-form, or to its (similar-sounding) phonologically translation.
@@
 Given the phonological similarity between *flower* and *flor*, it is possible that both word-forms are activated in the child's lexicon every time any of them is encountered.
-This would lead to an increase in the association strength between each word form and their common referent every time one of them is encountered, leading to an earlier age of acquisition.
+This would lead to an increase in the association strength between each word-form and their common referent every time one of them is encountered, leading to an earlier age of acquisition.
 The case of a non-cognate translation equivalent, such as *cow* and *vaca* would be different.
@@
 Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
-Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
+Therefore, participants may have on average learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
 Under the second hypothesis, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
@@
 In this study, we investigated the role of cognateness on the acquisition of translation equivalents, and if such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair.
-Using an online vocabulary checklist--designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We adopted an Bayesian item response theory (*IRT*) approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to the word-form's cognate status, and the participant's frequency of exposure to the word-form (a composite measure of lexical frequency and language degree of exposure), while adjusting for other indices of word difficulty---lexical frequency [@goodman2008does] and length (phonemes)---and participant skill---age [see @kachergis2022standard for a similar approach].
+Using an online vocabulary checklist designed specifically for this study we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
+We adopted an Bayesian item response theory (*IRT*) approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to the word-form's cognate status, and the participant's frequency of exposure to the word-form (a composite measure of lexical frequency and language degree of exposure), while adjusting for other indices of word difficulty like lexical frequency [@goodman2008does] and length (phonemes), and participant skill like age [see @kachergis2022standard for a similar approach].
 
@@
 
-To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire [@garcia-castro2022multilex].
+To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire.
 This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formR platform [@arslan2020formr].
-This questionnaire is structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
+This questionnaire is structured in three blocks: (1) a language questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
 Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]`.
+The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
 Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
+In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
 We included items from a diverse sample of `r n_categories` semantic/functional categories.
@@
 For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (`r format(n_te, big.mark = ",")` translation equivalents; `r format(n_item_language["catalan"], big.mark = ",")` items in Catalan, `r format(n_item_language["spanish"], big.mark = ",")` items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
-These last families were randomly allocated into one of four different subsets of the complete list of items.
+In total, each response to the questionnaire provided data for 400 translation equivalents, that is, for 800 words: 400 in Catalan and 400 in Spanish.
+Participants were randomly allocated into one of four different subsets of the complete list of items.
 These lists were designed so that each contained a representative sub-sample of the items from the complete list.
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Semantic/functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
 Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-For the analyses in the present study, we considered responses to words corresponding to nouns, verbs, and adjectives [@fourtassi2020growth].
+For the analyses in the present study, we considered responses nouns [@fourtassi2020growth].
 We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono/mico*).
@@
 We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
-`r n_participants[1]` of them participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
-Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(participants$age))` and `r floor(max(participants$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants-lp for a summary of the distribution of participants across ages and degrees of exposure to Catalan).
+`r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
+Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(df$age))` and `r ceiling(max(df$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan [see @tbl-participants for a summary of the distribution of participants across ages and degrees of exposure to Catalan].
 All families gave informed consent before participating.
@@
 ```{r}
-#| label: tbl-participants-lp
+#| label: tbl-participants
 #| echo: false
@@
 We used the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain.
-Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))` and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
+Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r percent(table(participants$edu_parent)["University"]/length(unique(participants$id)))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))`) and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
 
@@
 
-Each observation in our data was matches with its corresponding values in four variables of interest that we used to model participants' responses:
+Each observation in the resulting dataset was matched with its corresponding values in four variables of interest that we used to model participants' responses:
 
@@
 
-Our dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one of `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items["Catalan"]` in Catalan, `r n_items["Spanish"]` in Spanish, making up `r n_te` translation equivalents in total.
+Our final dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one of `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items["Catalan"]` in Catalan, `r n_items["Spanish"]` in Spanish, making up `r n_te` translation equivalents in total.
 Translation equivalents (TEs) received a median of `r printnum(sum_obs$median, digits = 0)` responses (*Min* = `r printnum(sum_obs$min, digits = 0)`, *Max* = `r printnum(sum_obs$max, digits = 0)`), both languages summed together.
@@
 This is interval contains the true value of this coefficient with 95% probability, given the data.
-Finally, we calculated the proportion of posterior samples in the 95% *HDI* that fall into the *ROPE*, which indicates the probability that the true value of the regression coefficient falls into the *ROPE*, and should therefore be considered equivalent to zero.
-For example, a 90% overlap between the *HDI* of the coefficients’ posterior distribution and the *ROPE* indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero.
+Finally, we calculated the proportion of posterior samples in the 95% *HDI* that fell into the *ROPE*, which indicates the probability that the true value of the regression coefficient falls into the *ROPE*, and should therefore be considered equivalent to zero.
+For example, a 90% overlap between the *HDI* of a coefficient's posterior distribution and the *ROPE* indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero.
 
@@
 Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see @sec-appendix-loo for all model comparisons).
-@tbl-results-fixed shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$.
+@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$.
 The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor.
@@
 ```{r}
-#| label: tbl-results-fixed
+#| label: tbl-coefs
 #| tbl-cap: "Posterior distribution of regression coefficients.
@@
 
-```{r}
-#| label: fig-results-fixed
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 4
-#| eval: false
-#| fig-cap: "Posterior distribution of the regression coefficients of the fixed effects in the extended model.
-Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis).
-Intervals show the 95% highest density interval (HDI) of each distribution.
-This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-The mean and HDI limits are indicated below each distribution.
-On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE, [-1, +1]).
-This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
-Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
-post <- post_draws %>%
-    mutate(.variable_name = factor(
-        .variable,
-        levels = names(str_repl),
-        labels = str_repl,
-        ordered = TRUE
-    ))
-
-post_summary <- post %>%
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    median_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>%
-    mutate(.variable_name = factor(
-        .variable,
-        levels = names(str_repl),
-        labels = str_repl,
-        ordered = TRUE
-    ))
-
-post_pd <- posterior_description %>% 
-    mutate(.variable_name = factor(
-        parameter,
-        levels = names(str_repl),
-        labels = str_repl,
-        ordered = TRUE
-    )) %>% 
-    filter(!grepl("Intercept|sd", parameter))
-
-post %>%
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    annotate(
-        geom = "rect",
-        ymin = -Inf,
-        ymax = Inf,
-        xmin = rope_interval["lower"],
-        xmax = rope_interval["upper"],
-        colour = NA,
-        alpha = 0.2,
-        fill = clrs[5],
-    ) +
-    geom_vline(xintercept = 0,
-               size = 1,
-               colour = clrs[5]) +
-    stat_slab(
-        aes(
-            fill = stat(abs(x) < rope_interval["upper"]),
-            colour = stat(abs(x) < rope_interval["upper"]),
-        ),
-        size = 1,
-        position = position_nudge(y = 0.15),
-        scale = 0.60
-    ) +
-    geom_errorbar(data = post_summary,
-                  aes(xmin = .lower, xmax = .upper, x = .value),
-                  width = 0.15) +
-    geom_point(data = post_summary, aes(x = .value), size = 1.5) +
-    geom_text(
-        data = post_summary,
-        aes(x = .value,
-            y = fct_rev(.variable_name),
-            label = paste0(
-                round(.value, 1),
-                " [",
-                round(.lower, 1),
-                ", ",
-                round(.upper, 1),
-                "]"
-            )),
-        colour = "black", size = 3,
-        position = position_nudge(y = -0.3),
-    ) +
-    geom_label(
-        data = post_pd,
-        aes(x = -0.75,
-            y = .variable_name,
-            label = percent(rope_percentage)),
-        hjust = 0, colour = "black", size = 3, alpha = 0.5,
-        fill = clrs[4], label.r = unit(0, "lines"), label.size = 0
-    ) +
-    labs(
-        x = "Logistic regression coefficient estimate (Logit scale)",
-        y = "Variable",
-        fill = "Overlaps with ROPE",
-        colour = "Overlaps with ROPE"
-    ) +
-    scale_x_continuous(breaks = seq(-2, 2, 0.2)) +
-    scale_fill_manual(values = clrs[c(1, 4)],
-                      labels = c("No", "Yes")) +
-    scale_colour_manual(values = clrs[c(1, 4)],
-                        labels = c("No", "Yes")) +
-    theme(
-        legend.position = "top",
-        legend.justification = c(1, 1),
-        axis.title.y = element_blank(),
-        panel.grid.major.x = element_line(colour = "grey85",
-                                          linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
 The word exposure index ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]).
@@
 
-In this study, we built on previous studies to investigate the role of cognateness on word acquisition.
+## Summary of the study
+
+In this study, we built on previous literature to investigate the role of cognateness on word acquisition.
 We contrasted two hypotheses regarding the stage of the word acquisition process at which the possible (facilitative) effect of cognateness might take place.
-First, we considered the possibility that cognateness boosts the acquisition of translation equivalents at early stages, before a stable, robust representation of either of the word-forms involved has been formed in the child's lexicon.
+First, we considered the possibility that cognateness boosts acquisition at early stages, before a stable, robust representation of either of the word-forms involved in the translation equivalent has been formed in the child's lexicon.
 This account predicts that the effect of cognateness would take place early in age, and would be symmetric across languages: one word-form boosts the acquisition of the other via their phonological similarity.
@@
 
-As anticipated, words with fewer phonemes or to which the child was exposed to more frequently were more likely to be acquired [e.g., @cattani2014how; @floccia2018iii; @braginsky2019consistency].
-We also found a facilitation effect of cognateness, providing converging evidence with previous studies [@floccia2018iii; @mitchell2022cognates; @bosch2014first; @schelletter2002effect].
+## Our findings
+
+Our results are, to a large extent, in line with previous studies.
+As it could be expected, we found that shorter word-forms were acquired at early ages than longer word-forms.
+For instance, $sol$ (Catalan for *sun*) was acquired earlier than *cocodril* (Catalan for *crocodile*).
+Shorter word forms consist on fewer phonemes and, frequently, fewer syllables.
+This makes them simpler to encode as phonological representations in the lexicon, and therefore easier to process.
+Ultimately, these words are easier to produce, and therefore to be reported as acquired in parental vocabulary checklists.
+Previous studies found word length to be more strongly associated with receptive vocabulary than with expressive vocabulary [e.g., @braginsky2019consistency; @jones2019children].
+
+According to previous literature, we also anticipated that words to which children were more frequently exposed would be acquired earlier [e.g., @braginsky2019consistency].
+In line with accumulator models of word acquisition, the age of acquisition of a word is a function of the child's rate of exposure to the word, which conspires with child- and word-level characteristics to the formation of a representation of the word in the child's lexicon [@mollica2017how; @hidaka2013computational].
+The amount of learning instances that a child is given to learn a word is given, to a large extent, by the lexical frequency of such word in the child's language.
+Words with higher lexical frequency are more likely to be uttered in the presence of the child, while words with lower lexical frequency appear less often.
+In the case of bilinguals, the exposure rate to a given word not only depends on lexical frequency, but also on the child's amount of exposure to the language such word belongs to [@cattani2014how; @floccia2018iii].
+For instance, a Spanish dominant bilingual child who is exposed to Catalan only 10% of the time will rarely hear the Catalan word *casa*, despite its high lexical frequency.
+We estimated the exposure rate of the words in our list by calculating a language exposure-weighted measure of lexical frequency, which adjusts the lexical frequency of the word by the child's amount of exposure to the language such word belongs to.
+We found a strong association between this measure of exposure rate and word's age of acquisition.
+Interestingly, this predictor had a stronger role during earlier ages: younger infants benefited from higher exposure rates at earlier ages, suggesting that early word acquisition is supported by mere exposure to word forms more strongly than later word acquisition.
+
+Finally, we found a facilitation effect of cognateness, providing converging evidence with previous studies [@floccia2018iii; @mitchell2022cognates; @bosch2014first; @schelletter2002effect].
 Critically, this effect was mediated by exposure: only words to which children were exposed to less frequently benefited from their cognate status.
 These results favour the account that a word-form benefits from its phonological similarity with its translation only after such translation has been acquired first.
-
-We expected to find such effect at earlier rather than later ages, but the size of the interaction effect between $Age$, $Exposure$ and $Levenshtein$ failed to exclude a considerable range of null values.
+On the one hand, our results provide converging evidence for @floccia2018iii, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample.
+Most children in their sample were English-dominant, and therefore their exposure to English was higher than to the additional language.
+However, it was in the additional language vocabulary where they found a facilitation effect of cognateness.
+Bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese).
+Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018iii's sample.
+
+## How our findings relate to previous literature
+
+In line with the hypothesis that received more support from our data, we predicted an interaction effect between participants' age, cognateness, and word exposure, since we expected cognateness to play a more important role at later stages of word learning, when at least one of the members of the translation equivalent has been acquired, and therefore its counterpart in the other language can benefit from its cognate status.
+We found evidence supporting that the true value of this interaction is close to zero.
+Finding such an effect would provide further evidence that cognateness benefits word acquisition only after the translation of such word is acquired.
+However, the absence of such interaction does not preclude a more careful endorsement
+of this hypothesis.
+It is possible that, even if the cognateness effect plays a role if the acquisition of the word-form with the least exposure of the translation equivalent, age itself might not modulate the size of this effect: cognateness plays a similar role across ages.
+
+## Limitations
 
@@
 Catalan and Spanish are both Romance languages, share a considerable amount of cognates, and very often co-exist in the same social spaces (families, schools, working place, etc.).
-Although there are no populations of bilinguals with identical sociolinguistic characteristics, and therefore it would be pointless to compare the sample in this study to any standard, the generalisability of our results should be taken with a grain of salt.
-It is possible, for instance, that the closeness of the two languages, together with the frequent co-occurrence of translation equivalents in the same contexts leads our participants to engage cognateness as a facilitator of word acquisition, but that children learning a pair of languages that share a fewer amount of cognates do not benefit from such mechanism.
-Another consequence of the large amount of cognates that Catalan-Spanish bilingual children are exposed to is that they learn at early stages of word acquisition that words tend to sound similar in both languages, and therefore be more keen on assuming that similar sounding words are semantically related, while children exposed to a fewer amount of cognates might not make the same assumption, even when the word-forms they hear in the presence of the same referents are very similar at the phonological level.
-
-## Limitations
-
-* These results come from a particular population of bilingual toddlers that learn two very similar languages--Catalan and Spanish--and are therefore exposed to many cognates.
-Our results might not generalise to other groups of bilinguals learning two less similar languages.
-* [To be discussed]
-
-## Further steps
-* [To be discussed]
+Although there are no populations of bilinguals with identical sociolinguistic characteristics, and therefore it would be pointless to compare the sample in this study to any "standard" population of bilinguals, the generalisability of our results should be taken with a grain of salt.
+It is possible, for instance, that the large amount of cognates that Catalan-Spanish bilingual children are exposed leads them to learn at early stages of word acquisition that words tend to sound similar in both languages, and therefore to be more keen on assuming that similar sounding words are semantically related.
+Children exposed to a pair of languages sharing a fewer amount of cognates might not make the same assumption, even when the word-forms they hear in the presence of the same referents are very similar at the phonological level.
+
+Another limitation is the predominantly cross-sectional design of our data set.
+We collected vocabulary data from children aged 12 to 32 months, a critical period in development in which the foundations of a lexicon get established.
+We estimated word acquisition curves by modelling the probability of comprehension and production across ages.
+However, most participants provided data at only one age point.
+Effectively, we are making longitudinal inferences based on cross-sectional data, and drawing conclusions about the development of a single participant across multiple age points using data from different participants at individual age points.
+For this reason, and also because of the observational nature of our study, causal inferences derived from our analyses are unwarranted [@kraemer2000how; @kesmodel2018crosssectional].
+One counterpoint to this shortcoming is that, while data our data is cross-sectional if participants are considered as the observational unit, data about word-forms/translation equivalents is longitudinal.
+The same translation equivalent is responded to at multiple ages.
+Since translation equivalents are our main observational unit of interest over which we primarily conduct statistical inference (e.g., estimating the size of the cognateness-effect), this provides some further justification for our conclusions.
+
+## Future steps
+
+Future studies addressing this subject may benefit for a more adequate design oriented towards making causal inferences.
+One such design can be partially satisfied by (1) collecting a more substantial longitudinal dataset, in which the same participant provides data points for each translation equivalent at multiple ages, and (2) testing the role of cognateness and word exposure in an experimental context, in which such variables can be orthogonally manipulated or randomised.
+This second possibility would also allow to explore an interpretation of the results obtained in the context of *word learning*, which is not possible in the present study.
+Another consideration for future investigations is the involvement of a more diverse range of language pairs as in @braginsky2019consistency or @floccia2018introduction to improve the generalisability of the resulting inferences across more populations of bilinguals.
+
+## Conclusions
+In summary, our study has replicated previous studies and provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
+We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word's phonological similarity with its translation in the other language.
+We suggest that this effect might be explained by the need of a robust lexical representation of one of the translations for cognateness to play a role.
 
@@
 To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
-We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset--each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
+We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
 For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].

### 2023-02-01 docs/index.qmd
Update manuscript

@@
         code-overflow: scroll
-    pdf: default
 link-external-newwindow: true
@@
 fig-dpi: 1000
+page-layout: full
 warning: false
@@
 library(janitor)
+library(glue)
 library(brms)
@@
 tar_load(participants)
-tar_load(df)
+tar_load(responses)
 tar_load(items)
@@
 tar_load(model_loos)
-tar_load(rope_interval)
-tar_load(posterior_description)
-tar_load(posterior_description_re)
+tar_load(posterior_draws)
+tar_load(posterior_draws_re)
 
@@
         ggplot2.continuous.colour = ggplot2::scale_color_gradient)
+
+rope_interval <- c(lower = -0.1, upper = +0.1)
 ```
@@
 #     summarise_at(
-#         vars(freq_zipf, n_phon, lv),
+#         vars(freq, n_phon, lv),
 #         list(
@@
 #     select(-c(n_phon_n, lv_n)) %>%
-#     rename(n = freq_zipf_n) %>%
+#     rename(n = freq_n) %>%
 #     gt(groupname_col = "language", rowname_col = "list") %>%
@@
 #     tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>%
-#     cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>%
+#     cols_merge_uncert(col_val = freq_mean, col_uncert = freq_sd) %>%
 #     cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>%
 #     cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>%
-#     cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>%
+#     cols_merge_range(col_begin = freq_min, col_end = freq_max) %>%
 #     cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>%
@@
 #         n = md("*N*"),
-#         freq_zipf_mean = md("Mean ± SD"),
-#         freq_zipf_min = md("Range"),
+#         freq_mean = md("Mean ± SD"),
+#         freq_min = md("Range"),
 #         n_phon_mean = md("Mean ± SD"),
@@
     distinct(te, .keep_all = TRUE) %>% 
-    ggplot(aes(x = freq_zipf)) +
+    ggplot(aes(x = freq)) +
     geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
@@
 items %>% 
-    drop_na(freq_zipf) %>% 
+    drop_na(freq) %>% 
     distinct(te, .keep_all = TRUE) %>% 
-    count(n_phon, freq_zipf) %>% 
-    ggplot(aes(n_phon, freq_zipf)) +
+    count(n_phon, freq) %>% 
+    ggplot(aes(n_phon, freq)) +
     geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
@@
     distinct(te, .keep_all = TRUE) %>% 
-    drop_na(freq_zipf) %>% 
-    count(freq_zipf, lv) %>% 
-    ggplot(aes(freq_zipf, lv, size = n)) +
+    drop_na(freq) %>% 
+    count(freq, lv) %>% 
+    ggplot(aes(freq, lv, size = n)) +
     geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
@@
     distinct(te, .keep_all = TRUE) %>% 
-    drop_na(freq_zipf) %>% 
+    drop_na(freq) %>% 
     count(n_phon, lv) %>% 
@@
 
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):
+For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`) would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):
 
@@
 #| fig-width: 13
-df %>%
+responses %>%
     mutate(
@@
 
-    df %>%
+    responses %>%
     mutate(
@@
 #| fig-height: 3.5
-dist_df = tribble(~mean, ~sd, 0,   1)
-
-responses <- data.frame(
+dist_df <- tribble(~mean, ~sd, 0,   1)
+
+thresholds <- data.frame(
     ymin = c(-Inf, -1.5, 1.5),
@@
     xmax = Inf,
-    .category = str_wrap(c("No", "Understands", "Understands and Says"), width = 12)
-)
-
-responses %>%
-    ggplot(aes(ymin = ymin, ymax = ymax, xmin = xmin, xmax = xmax, fill = .category)) +
+    .category = str_wrap(c(
+        "No", 
+        "Understands",
+        "Understands and Says"
+    ),
+    width = 12)
+)
+
+thresholds %>%
+    ggplot(aes(ymin = ymin, ymax = ymax,
+               xmin = xmin, xmax = xmax,
+               fill = .category)) +
     geom_rect(alpha = 0.5) +
@@
     ) +
-    geom_hline(yintercept = c(-1.5, 1.5), linetype = "dotted", size = 1) +
+    geom_hline(yintercept = c(-1.5, 1.5), 
+               linetype = "dotted",
+               linewidth = 1) +
     geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
@@
     n_phon_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
     exposure_std = 0,
@@
 
-m <- add_epred_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA) %>% 
+m <- add_epred_draws(nd, 
+                     model_fit_4_prior, 
+                     ndraws = NULL,
+                     re_formula = NA) %>% 
     filter(.category != "No") %>% 
@@
     scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
         labels = seq(10, 36, 4)
@@
 #| fig-cap: "Proportion of responses to each category by age and number of phonemes"
-df %>% 
+responses %>% 
     mutate(
@@
 #| fig-cap: "Proportion of responses to each category by age and exposure."
-df %>% 
+responses %>% 
     mutate(
@@
 #| fig-cap: "Proportion of responses to each category by age and Levenhstein similarity"
-df %>%
+responses %>%
     mutate(
@@
     mutate(model = str_extract_all(model, "\\d")) %>% 
-    gt() 
-# fmt_number(2:9, decimals = 1) %>% 
-# cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>% 
-# cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>% 
-# cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>% 
-# cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>% 
-# cols_label(
-#     model = "Model",
-#     elpd_loo = md("*ELPD*"),
-#     p_loo = md("*p*"),
-#     looic = md("*LOO-IC*"),
-#     elpd_diff = md("Difference")
-# ) %>% 
-# tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
-# tab_style(
-#     cell_text(weight = "bold"),
-#     cells_column_labels()
-# ) %>% 
-# tab_style(
-#     cell_fill(color = "grey", alpha = 0.5),
-#     cells_body(matches("diff"))
-# ) %>% 
-# tab_style(
-#     cell_fill(color = clrs[4], alpha = 0.5),
-#     cells_body(rows = 1)
-# )
+    gt() %>% 
+    fmt_number(2:9, decimals = 1) %>%
+    cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>%
+    cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>%
+    cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>%
+    cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>%
+    cols_label(
+        model = "Model",
+        elpd_loo = md("*ELPD*"),
+        p_loo = md("*p*"),
+        looic = md("*LOO-IC*"),
+        elpd_diff = md("Difference")
+    ) %>%
+    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>%
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_labels()
+    ) %>%
+    tab_style(
+        cell_fill(color = "grey", alpha = 0.5),
+        cells_body(matches("diff"))
+    ) %>%
+    tab_style(
+        cell_fill(color = clrs[4], alpha = 0.5),
+        cells_body(rows = 1)
+    )
 ```
@@
 ## Fixed effects
-
-```{r}
-#| label: results-posterior-fixed
-post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)
-
-# tidy predictor names
-str_repl <- c(
-    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
-    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
-    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
-    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
-    "b_exposure_std" = paste0("Exposure (+1 SD, ",  percent(sd(df$exposure)), ")"),
-    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
-    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
-    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
-    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
-)
-```
 
@@
 #| fig-width: 7
-intercept_footnote <- paste0(
-    "Probability of comprehension or production at the grand age of the sample (",
-    round(mean(df$age), 2),
-    " months)"
-)
+intercept_footnote <- glue("Probability of comprehension or production at the grand age of the sample ({round(mean(responses$age), 2)} months")
 
 # summarise posterior draws
-posterior_description %>%
-    mutate(parameter = factor(parameter, levels = names(str_repl), labels = str_repl)) %>%
-    select(parameter, median, ci_high, ci_low, rope_percentage) %>%
-    gt(rowname_col = "parameter") %>%
-    fmt_number(c(2:4), decimals = 2) %>%
-    fmt_percent(5, decimals = 2) %>%
-    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
-    cols_label(median = "Median",
-               ci_low = "95% HDI",
-               rope_percentage = "ROPE prob.") %>%
+posterior_draws$summary %>%
+    gt(rowname_col = ".variable_name", 
+       groupname_col = ".type") %>%
+    cols_hide(".variable") %>% 
+    fmt_number(.median:.lower, decimals = 2) %>%
+    fmt_percent(.rope_overlap, decimals = 2) %>%
+    cols_merge(c(.lower, .upper), pattern = "[{1}, {2}]") %>%
+    cols_label(.median = "Median",
+               .lower = "95% HDI",
+               .rope_overlap = "ROPE prob.") %>%
     tab_style(cell_text(weight = "bold"),
-              cells_column_labels(columns = 1:5)) %>%
+              cells_column_labels()) %>%
     tab_style(cell_text(align = "left"),
@@
 
-post_draws_list <-
-    split(posterior_description, posterior_description$parameter)
+post_draws_list <- split(posterior_draws$summary, 
+                         posterior_draws$summary$.variable)
 ```
@@
 We decided to draw this interval around zero so that it covered values smaller than those considered as negligible by similar previous studies: after consulting the regression tables of [ADD STUDIES], we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: 3%.
-consequently, we defined our *ROPE* from -3% to +3%.
+consequently, we defined our *ROPE* from -0.1% to +0.1%.
 We report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
 This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
-Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the -3%--3% *ROPE* interval we previously defined.
+Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the *ROPE* interval we previously defined.
 For those coefficients that did not exclude completely the *ROPE* we also report a Bayes Factor indicating how many times is such coefficient more likely to be bigger than the upper limit of the *ROPE*, or smaller than the lower limit of the *ROPE*, given our data.
@@
 @fig-results-fixed shows the outputs of the model.
-**Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$median, 2)`, 95% *HDI* = [`r round(post_draws_list[["b_age_std"]]$ci_low, 2)`, `r round(post_draws_list[["b_age_std"]]$ci_high, 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month.
+**Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 2)`, 95% *HDI* = [`r round(post_draws_list[["b_age_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std"]]$.upper, 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$.median/(4*sd(responses$age)), accuracy = 0.01)` every month.
 
 The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]).
-For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.
-
-**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]).
+This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$.upper, accuracy = 0.01)`]).
+For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$.median/(4*sd(responses$n_phon)), accuracy = 0.01)`.
+
+**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$.upper, accuracy = 0.01)`]).
 All of the posterior samples of his regression excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/(sd(df$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.
-
-The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
-
-However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word.
+The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/(sd(responses$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.
+
+The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
+
+However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word.
 Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
@@
 
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
 This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
@@
 
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
+The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$.upper, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
+
+Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
 
@@
 the derivative of the logistic function in its mid-point."
-post <- post_draws %>% 
-    mutate(
-        .variable_name = factor(
-            .variable, 
-            levels = names(str_repl),
-            labels = str_repl,
-            ordered = TRUE
-        )
-    )
-
-post_summary <- post %>% 
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    median_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>% 
-    mutate(
-        .variable_name = factor(
-            .variable, 
-            levels = names(str_repl),
-            labels = str_repl,
-            ordered = TRUE
-        )
-    )
-
-post %>%
-    filter(!grepl("Intercept|sd", .variable)) %>%
+
+posterior_draws$draws %>%
     ggplot(aes(.value, fct_rev(.variable_name))) +
     annotate(geom = "rect", ymin = -Inf, ymax = Inf,
-             xmin = rope_interval["lower"], xmax = rope_interval["upper"],
+             xmin = rope_interval["lower"],
+             xmax = rope_interval["upper"],
              colour = NA, alpha = 0.2, fill = clrs[5],) +
@@
             colour = stat(abs(x) < rope_interval["upper"]),),
-        size = 1, position = position_nudge(y = 0.15)
+        size = 1, 
+        position = position_nudge(y = 0.15)
     ) +
     geom_errorbar(
-        data = post_summary,
-        aes(xmin = .lower, xmax = .upper, x = .value),
+        data = posterior_draws$summary,
+        aes(xmin = .lower, 
+            xmax = .upper, 
+            x = .median),
         width = 0.15
     ) +
-    geom_point(data = post_summary, aes(x = .value), size = 1.5) +
+    geom_point(data = posterior_draws$summary, 
+               aes(x = .median),
+               size = 1.5) +
     geom_text(
-        data = post_summary,
-        aes(
-            x = .value, 
+        data = posterior_draws$summary,
+        aes(x = .median, 
             y = .variable_name,
-            label = paste0(
-                round(.value, 1),
-                " [", round(.lower, 1), ", ", round(.upper, 1), "]")
-        ),
-        colour = "black", position = position_nudge(y = -0.3), size = 3
+            label = glue(
+                "{round(.median, 1)}",
+                "[{round(.lower, 1)}, {round(.upper, 1)}]",
+                .sep = " ")),
+        colour = "black",
+        position = position_nudge(y = -0.3), size = 3
     ) +
@@
 
-nd <- datagrid(
-    model = model_fit_4,
-    age_std = seq_range(model_fit_4$data$age_std, n = 100),
-    lv_std = (seq(0, 1, 0.5)-mean(df$lv))/(sd(df$lv)),
-    exposure_std = c(-1, 0, 1),
-    te = NA,
-    id = NA
-)
-
-marg_epreds <- predictions(
-    model_fit_4,
-    newdata = nd,
-    ndraws = 25, 
-    re_formula = NA
-)
-
-marg_epreds %>% 
-    posteriordraws() %>% 
-    as_tibble() %>% 
-    select(drawid, group, draw, age_std, lv_std, exposure_std) %>%
-    filter(group != "No") %>% 
-    pivot_wider(
-        id_cols = c(drawid, age_std, lv_std, exposure_std), 
-        names_from = group, 
-        values_from = draw
-    ) %>% 
-    mutate(`Understands` = `Understands and Says` + `Understands`) %>% 
-    pivot_longer(c(`Understands`, `Understands and Says`), names_to = "group", values_to = "draw") %>% 
-    mutate(
-        exposure_std = factor(
-            exposure_std, 
-            levels = unique(nd$exposure_std), 
-            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
-        ),
-        lv_std = factor(
-            lv_std,
-            levels = unique(nd$lv_std), 
-            labels = paste0(c(0, 50, 100), "% similarity")
-        )
-    ) %>% 
-    ggplot(aes(age_std, draw, colour = lv_std, fill = lv_std)) +
-    facet_grid(group~exposure_std) +
+marginal_effects_epreds %>% 
+    ggplot(aes(age, .value, colour = lv, fill = lv)) +
+    facet_grid(.category~exposure) +
     # stat_summary(fun.data = median_hdi, geom = "ribbon",
     #              alpha = 0.5, colour = NA) +
-    geom_line(aes(group = interaction(lv_std, drawid)),
-              size = 0.75, alpha = 0.15) +
-    stat_summary(aes(linetype = lv_std), fun = median, geom = "line", size = 1) +
+    geom_line(aes(group = interaction(lv, .draw)),
+              size = 0.75,
+              alpha = 0.15) +
+    stat_summary(aes(linetype = lv), 
+                 fun = median, 
+                 geom = "line",
+                 size = 1) +
     labs(
@@
     ) +
-    scale_x_continuous(
-        breaks = (seq(10, 32, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 32, 4)
-    ) +
+    scale_x_continuous(breaks = seq(10, 32, 4)) +
     theme(
         legend.position = "top",
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
+        panel.grid = element_line(colour = "grey", 
+                                  linetype = "dotted"),
+        panel.border = element_rect(fill = NA, 
+                                    colour = "black",
+                                    linewidth = 0.75)
     )
@@
     geom_hline(yintercept = 0, size = 1, colour = "grey") +
-    geom_line(aes(group = drawid), alpha = 0.25, size = 0.75, colour = clrs[4]) +
+    geom_line(aes(group = drawid), 
+              alpha = 1, 
+              size = 0.75,
+              colour = "grey") +
     # stat_summary(fun.data = median_hdi, geom = "ribbon",
     #              colour = NA, alpha = 0.5, fill = clrs[1]) +
-    stat_summary(fun = median, geom = "line", size = 1, color = clrs[1]) +
+    stat_summary(fun = mean, 
+                 geom = "line", 
+                 size = 1, 
+                 color = clrs[1]) +
     labs(
@@
     rownames_to_column("id") %>%
+    mutate(id = as.integer(id)) %>% 
     left_join(distinct(model_fit_4$data, id)) %>% 
@@
         across(
-            .cols = where(is.numeric),
+            .cols = estimate:q97_5,
             .fns = list(
-                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
-                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
-            ), .names = "{.col}__{.fn}"
+                Comprehension = ~plogis(. + fixef(model_fit_4)[1]),
+                Production = ~plogis(. + fixef(model_fit_4)[2])
+            ), 
+            .names = "{.col}__{.fn}"
         )
     ) %>% 
-    select(-c(estimate:q97_5)) %>% 
+    select(id, ends_with("Comprehension"), ends_with("Production")) %>% 
     pivot_longer(
-        where(is.numeric), 
+        -id, 
         names_to = c(".value", "type"),
@@
 
-#### Correlations
-
-```{r}
-#| label: fig-results-random-id-correlations
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 7
-#| fig-cap: "Posterior correlations between random effects by participant"
-posterior_description_re %>% 
-    select(parameter, median) %>% 
-    filter(str_detect(parameter, "cor_id__")) %>% 
-    mutate(parameter = str_remove_all(parameter, "cor_id__")) %>%  
-    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
-    mutate(
-        across(
-            c(var_1, var_2), 
-            ~factor(
-                .,
-                levels = c("Intercept", "age_std", "exposure_std", "lv_std", "n_phon_std", "age_std:exposure_std", "age_std:lv_std", "exposure_std:lv_std", "age_std:exposure_std:lv_std"),
-                labels = c("Intercept", "Age", "Exposure","Levenshtein", "Phonemes", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein", "Exposure \u00d7 Levenshtein", "Age \u00d7 Exposure \u00d7 Levenshtein")
-            )
-        )
-    ) %>%
-    filter(var_1 != var_2) %>% 
-    ggplot(aes(x = var_1, y = fct_rev(var_2), 
-               colour = median, label = percent(median, accuracy = 0.1))) +
-    geom_point(shape = 15, size = 20) +
-    geom_text(colour = "black") +
-    scale_colour_gradient2(low = clrs[1], 
-                           mid = "grey", 
-                           high = clrs[4],
-                           midpoint = 0) +
-    labs(colour = "Correlation", size = "Correlation") +
-    theme(
-        axis.title = element_blank(),
-        axis.text.x = element_text(angle = 35, hjust = 1),
-        panel.grid = element_blank()
-    )
-```
 :::
@@
     ggplot() +
-    aes(x = estimate, y = reorder(te, estimate), 
-        xmin = q2_5, xmax = q97_5,
-        colour = type, fill = type) +
+    aes(x = estimate,
+        y = reorder(te, estimate), 
+        xmin = q2_5,
+        xmax = q97_5,
+        colour = type, 
+        fill = type) +
     facet_wrap(~type) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
+    geom_ribbon(aes(group = 1), 
+                alpha = 0.5, 
+                colour = NA) +
     geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior probability of acquisition",
+    geom_vline(xintercept = 0.5, 
+               colour = "black", 
+               linetype = "dotted") +
+    labs(x = "Posterior probability of acquisition",
          y = "TE", 
-        colour = "Type"
-    ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
+         colour = "Type") +
+    scale_x_continuous(limits = c(0, 1),
+                       labels = percent) +
+    scale_fill_manual(values = clrs[c(1, 4)]) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
     theme(
@@
 
-#### Correlations
-
-```{r}
-#| label: fig-results-random-te-correlations
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 7
-#| fig-cap: "Posterior correlations between random effects by translation equivalent"
-posterior_description_re %>% 
-    select(parameter, median) %>% 
-    filter(str_detect(parameter, "cor_te__")) %>% 
-    mutate(parameter = str_remove_all(parameter, "cor_te__")) %>%  
-    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
-    mutate(
-        across(
-            c(var_1, var_2), 
-            ~factor(
-                .,
-                levels = c("Intercept", "age_std", "exposure_std", "n_phon_std", "lv_std", "age_std:exposure_std", "age_std:lv_std"),
-                labels = c("Intercept", "Age", "Exposure", "Phonemes", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein")
-            )
-        )
-    ) %>%
-    filter(var_1 != var_2) %>% 
-    ggplot() +
-    aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1)) +
-    geom_point(shape = 15, size = 40) +
-    geom_text(colour = "black") +
-    scale_colour_gradient2(low = clrs[1], 
-                           mid = "grey", 
-                           high = clrs[4],
-                           midpoint = 0) +  
-    labs(
-        colour = "Correlation",
-        size = "Correlation"
-    ) +
-    theme(
-        axis.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
 :::
@@
 ## Model diagnostics
-
-### Simulation-based calibration
-
-```{r}
-#| label: fig-simulation-based-calibration
-#| fig-cap: "Posterior predictive checks.
-We simulated 500 datasets from our model by drawing random samples from its posterior distribution.
-We then computed the proportion of 'No', 'Understands', and 'Understands and Says' responses in each dataset.
-These simulated proportions are depicted in the plot as solid horizontal lines.
-Their variation indicate the uncertainy of the model about what proportion of each type of response would be observe in a newly collected dataset.
-We also plotted the proportion of each type of response in the observed dataset (with a ± 5% interval for comparison)."
-predicted_dfs <- add_predicted_draws(
-    object = model_fit_4, 
-    newdata = df,
-    re_formula = NULL,
-    ndraws = 500,
-    seed = 888
-)
-
-observed_prop <- df %>% 
-    count(response) %>%
-    mutate(prop = n/sum(n)) %>% 
-    rename(.prediction = response)
-
-predicted_dfs %>% 
-    ungroup() %>% 
-    select(.prediction, .draw) %>% 
-    count(.draw, .prediction) %>%
-    group_by(.draw) %>% 
-    mutate(
-        prop = n/sum(n),
-        .prediction = factor(
-            .prediction,
-            levels = levels(df$response), 
-            ordered = TRUE
-        )
-    ) %>% 
-    ggplot(aes(.draw, prop, fill = .prediction, colour = .prediction)) +
-    geom_line(aes(group = .prediction)) +
-    geom_rect(
-        data = observed_prop,
-        aes(xmin = 0, xmax = max(predicted_dfs$.draw),
-            ymin = prop-0.025, ymax = prop+0.025,
-            colour = .prediction, fill = .prediction),
-        alpha = 0.25, colour = NA, inherit.aes = FALSE
-    ) +
-    # geom_segment(
-    #     data = observed_prop,
-    #     aes(x = -Inf, xend = Inf, y = prop, yend = prop,
-    #         colour = .prediction),
-    #     size = 1, inherit.aes = FALSE
-    # ) +
-    labs(
-        x = "Iteration",
-        y = "Proportion of responses",
-        colour = "Response",
-        fill = "Response"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_x_continuous(labels = label_number(big.mark = ","), 
-                       breaks = seq(0, max(predicted_dfs$.draw), 100)) +
-    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank()
-    )
-
-```
 
@@
 get_loo_plot <- function(x){
+    custom_breaks<- seq(0, length(x$pointwise[,5]), 50000)
+
     x$pointwise[,5] %>% 
@@
         geom_point(
-            size = 0.5, alpha = 0.1, colour = clrs[4], 
+            size = 0.5, 
+            alpha = 0.1,
+            colour = clrs[4], 
             shape = 1
         ) +
-        geom_hline(yintercept = 0.5, linetype = "dotted") +
-        labs(
-            x = "Data point",
+        geom_hline(yintercept = 0.5,
+                   linetype = "dotted") +
+        labs(x = "Data point",
              y = "Pareto-k value",
-            title = "PSIS diagnostic plot"
-        ) +
-        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
-        scale_y_continuous(
-            limits = c(-0.6, 0.6)
-        )
+             title = "PSIS diagnostic plot") +
+        scale_x_continuous(breaks = custom_breaks) +
+        scale_y_continuous(limits = c(-0.6, 0.6))
 }

### 2023-02-01 manuscript/manuscript.qmd
Update manuscript

@@
 fig-dpi: 1000
-csl: "apa7.csl"
-bibliography: "references.bib"
+csl: "../assets/apa7.csl"
+bibliography: "../assets/references.bib"
 format:
   pdf:
-    pdf-engine: lualatex
-    number-sections: true
+    pdf-engine: xelatex
     whitespace: small
     linestretch: 1.5
+    include-in-header: 
+      text: '\usepackage{lineno}\linenumbers'
 ---
@@
 # load packages
-library(conflicted)
 library(ggplot2)
 library(patchwork)
+library(lubridate)
+library(brms)
+library(glue)
 library(dplyr)
@@
 library(gt)
-library(forcats)
 library(stringr)
@@
 library(janitor)
-library(brms)
-library(papaja)
 library(scales)
-library(stringdist)
-library(marginaleffects)
-library(lubridate)
 library(bvqdev)
@@
 tar_load(model_loos)
-tar_load(rope_interval)
-tar_load(posterior_description)
-
-participants <- left_join(participants, select(bvq_data$logs, id, time, edu_parent))
-items <- left_join(items, select(bvq_data$pool, item, language, semantic_category, class))
+tar_load(posterior_draws)
+tar_load(marginal_effects_epreds)
+
+items <- bvq_data$pool %>% 
+    mutate(item = str_remove(item, "cat_|spa_")) %>% 
+    select(item, language, semantic_category, class) %>% 
+    right_join(items, by = c("item", "language"))
 
@@
         ggplot2.discrete.colour = clrs,
-        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
-        ggplot2.continuous.colour = ggplot2::scale_color_gradient)
+        ggplot2.continuous.fill = scale_color_gradient,
+        ggplot2.continuous.colour = scale_color_gradient)
 
 # resolve namespace conflicts
-conflict_prefer("filter", "dplyr")
-conflict_prefer("select", "dplyr")
+# conflict_prefer("filter", "dplyr")
+# conflict_prefer("select", "dplyr")
 ```
@@
 The foundations of word learning are in place early in age.
-Before the end of their first year of life infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthurbates].
+From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthurbates].
 During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
-These early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e.
+Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e.
 vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of acquired words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure; @bates1994developmental].
 Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
-@tardif2008baby collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing.
+@tardif2008baby collected data about the first ten words acquired by 10-- to 16--month-old infants living in the United States, Hong Kong, and Beijing.
 Since birth, these infants had been learning English, Cantonese, and Mandarin, respectively.
@@
 
-Despite the recent increase in language representation in the literature, most studies on early word acquisition have relied exclusively on data from monolingual children while neglecting, to a considerable extent, that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent].
-Previous findings pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the two languages are taken into account.
-@hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English.
+Despite the recent increase in the representativeness of the samples in the literature, most studies on early word acquisition have relied exclusively on data from monolingual children while neglecting, to a considerable extent, that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent].
+Previous findings pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
+@hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals do.
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
+Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
+While these studies have mostly relied on samples of bilingual children learning two typologically relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
 What role could linguistic distance play during early vocabulary growth?
@@
 Depending on the linguistic distance between both languages, the two sets of words may overlap in form in varying degrees.
-When both languages are typologically close like Spanish and Catalan (both Romance languages) they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic, respectively).
+When both languages are typologically close like Spanish and Catalan (both Romance languages) they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic languages, respectively).
 For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
 It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
-If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
+If cognates are easier to acquire than non-cognates, bilinguals learning two languages that share a high proportion of cognates should benefit more often from this facilitation effect than bilinguals learning two languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
 
@@
 Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
-The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other [@levenshtein1966binary].
-The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
+The Levenshtein distance is a metric that computes the distance between two strings of characters by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
+The resulting scores are then divided by the length of the longest string to normalise the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
 Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
@@
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
-In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish.
+In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish.
 In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trials labels were non-cognates (*mesa*-*taula*, translations of *table*).
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
-Importantly, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, did not show this effect.
+Importantly, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, showed equivalent naming times for the two types of stimuli.
 These results suggest that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
 
-Subsequent studies have also provided similar evidence in comprehension in children [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
+This account is especially plausible in the light of (1) the evidence pointing to bilingual toddlers' lexicon consisting of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022translation; @pearson1995crosslanguage], and that (2) there is experimental evidence of parallel activation during word comprehension in children and toddlers [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
 Parallel activation is therefore a plausible explanation for Floccia et al.'s results: cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
-This account is is line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
-Importantly, bilinguals seem to be more likely to acquire words for which their translation equivalent has been already learnt in the other language, suggesting that semantic similarity also facilitates word acquisition across languages [@bilson2015semantic].
-
-If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity, but not phonological similarity.
-Evidence supporting and earlier age of acquisition for cognates is, to date, scarce.
-@bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
-@schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
+This account is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
+If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity but not phonological similarity.
+
+Evidence supporting an earlier age of acquisition for cognates is, to date, scarce nonetheless.
+@bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
+@schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual produced cognates earlier than non-cognates, on average.
 The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-On the other hand, Floccia et al.'s estimates are statically more reliable given their (remarkably larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
-In their discussion the authors state the following (pp.
-70):
-
-> "This  finding  also  provides  support  to  the  proposal  that  the  cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life."
-
-We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results.
-First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
+On the other hand, Floccia et al.'s estimates are statically more reliable given their (remarkably larger) sample size.
+Their results suggest a cognateness effect on vocabulary size, but their study was not aimed at testing the effect of cognateness on age of acquisition at the word-level.
+Their response variable was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
 By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.
 non-cognates was no longer available.
-Second, all participants were aged ~24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
+Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
 
 More recently, @mitchell2022cognates addressed this issue overcoming some of these pitfalls.
-Using a larger, longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages.
+Using a longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages.
 They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
@@
 These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar: word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
+Parallel activation remains a plausible explanation for this effect.
+When a child is exposed to a word form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition.
+The precise underpinnings of this effect are still unclear.
+
+## A mechanistic account for the cognate facilitation in word acquisition
+
+In this study, we suggest a mechanistic account for the cognate facilitation effect.
+We build this account from the assumption that bilinguals use fundamentally similar mechanisms than monolinguals to learn words.
+In both cases, learning a word requires at least two events to take place.
+First, a representation of the word-form, encoding its segmental information to sufficient detail, must be stored.
+
+This can be accomplished via bottom-up mechanism since early ages.
+Even if words are embedded in a continuous stream of speech sounds, in which there are no clear boundaries between words, infants can segment discrete word-forms.
+For instance, they may rely on their familiarity with the typical stress pattern [@jusczyk1995infants], the phonotactic constraints [@friederici1993phonotactic], or the distributional information of their native language [@saffran1996statistical] to detect word boundaries.
+A second event for word learning to take place is the association between the familiar word form and its referent.
+This task is far from trivial for many reasons.
+Referents are rarely labelled ostensively: children are exposed to most instances of a word in referentially ambiguous contexts or in the absence of their referent.
+Many words are also presented along remarkably variable exemplars of their referent, and even refer to abstract referents such as routines or play interactions (e.g., *uh-oh*; @casey2021moving).
+Infants nonetheless need few learning instances to create new word-referent associations.
+The extent to which they do so relying on bottom-up mechanisms such as cross-situational learning [@yu2007rapid], or on constraints and biases that guide their hypotheses about how words and referents are related [@golinkoff1994early], remains a matter of debate.
+Either way, there is evidence of word-referent links as early as six months of age [@bergelson2012months; @bergelson2015early].
+From a connectionist perspective, a word is said to be acquired when the association between its word-form and its conceptualised referent (i.e.
+semantic representation) has reached some level of associative strength.
+As infants accumulate experience with word-referent pairings, the association strength between both increases, and this leads to observable events such as infants' showing a stronger preference for a named object compared to unnamed competitors, or their parents being more likely to report such word as acquired in a vocabulary checklist.
+
+Bilinguals deal with a different linguistic landscape than monolinguals, but arguably use similar mechanisms and do so under common constraints [@curtin2011bilingual].
+Bilingual word acquisition entails similar events as described above, with the exception that many referents are
+
+they seem to reach their developmental goals of word learning at similar ages than monolinguals. Bilinguals store words
 
 ## Two competing mechanisms for a cognate facilitation during word acquisition
+
+We argue that the cognateness facilitation effect takes place
 
@@
 To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*) does not undergo the such facilitation.
-Both word-forms are fairly disimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](hypotheses.jpg){#fig-hypotheses}
+Both word-forms are fairly disimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](img/hypotheses.jpg){#fig-hypotheses}
 
@@
 
-n_item_language <- count(bvq_data$pool, language) %>%
-    pull(n) %>%
+n_item_language <- bvq_data$pool %>% 
+    count(language) |>
+    pull(n) |>
     set_names(c("catalan", "spanish"))
@@
 # load packages
-n_participants <- df %>%
-    distinct(id, age) %>%
-    count(id) %>%
-    pull(n) %>%
+n_participants <- responses |>
+    distinct(id, age) |>
+    count(id) |>
+    pull(n) |>
     table()
-n_responses <- nrow(distinct(df, id, age))
-dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
+
+n_responses <- nrow(distinct(responses, id, age))
+
+dates <- participants$time_stamp %>% 
+    range(na.rm = TRUE) %>% 
+    as_date()
 ```
@@
 `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
-Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(df$age))` and `r ceiling(max(df$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan [see @tbl-participants for a summary of the distribution of participants across ages and degrees of exposure to Catalan].
+Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan [see @tbl-participants for a summary of the distribution of participants across ages and degrees of exposure to Catalan].
 All families gave informed consent before participating.
@@
 Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
-participants %>%
-    filter(id %in% df$id) %>% 
-    mutate_at(vars(starts_with("doe_")), function(x) {
-        floor(x * 10) / 10
-    }) %>%
+participants |>
+    mutate_at(vars(starts_with("doe_")), 
+              (\(x) floor(x * 10) / 10)) %>% 
     mutate(
@@
             c(0, 0.25, 0.5, 0.75, 1),
-            labels = c("0-25%", "25-50%", "50-75%", "75-100%"),
+            labels = c("0-25%",
+                       "25-50%", 
+                       "50-75%",
+                       "75-100%"),
             include.lowest = TRUE
-        ) %>%
+        ) |>
             as.character()
-    ) %>%
-    count(age, doe_catalan) %>%
-    arrange(desc(doe_catalan)) %>%
-    pivot_wider(names_from = age, values_from = n) %>%
-    gt() %>%
+    ) |>
+    count(age, doe_catalan) |>
+    arrange(desc(doe_catalan)) |>
+    pivot_wider(names_from = age, values_from = n) |>
+    gt() |>
     summary_rows(
@@
         formatter = fmt_integer
-    ) %>%
-    sub_missing(everything(), everything(), missing_text = "--") %>%
-    cols_label(doe_catalan = "DoE Catalan") %>%
-    tab_spanner("Age (months)", 2:8) %>%
+    ) |>
+    sub_missing(everything(), everything(), missing_text = "--") |>
+    cols_label(doe_catalan = "DoE Catalan") |>
+    tab_spanner("Age (months)", 2:8) |>
     tab_footnote(
@@
         cells_column_labels(columns = doe_catalan)
-    ) %>%
+    ) |>
     tab_style(cell_text(align = "left"),
@@
 #| include: false
-n_obs <- nrow(df)
-sum_obs <- df %>%
-    mutate(te = factor(te)) %>%
-    count(te) %>%
+n_obs <- nrow(responses)
+
+sum_obs <- responses |>
+    mutate(te = factor(te)) |>
+    count(te) |>
     summarise(across(n, lst(median, min, max), .names = "{.fn}"))
 
-n_items <- ifelse(grepl("cat_", distinct(df, item)$item),
-                  "Catalan",
-                  "Spanish") %>%
-    table()
-
-n_te <- length(unique(df$te))
+n_items <- distinct(responses, language, item) |> 
+    count(language) |> 
+    group_split(language) |> 
+    set_names(c("Catalan", "Spanish")) |>
+    map_int("n")
+
+n_te <- length(unique(responses$te))
 
@@
 r_info <- sessionInfo()
-r_version <- paste(r_info$R.version$major, r_info$R.version$minor, sep = ".")
+
+r_version <- glue(r_info$R.version$major, 
+                  r_info$R.version$minor, 
+                  .sep = ".")
 ```
 
-Our final dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one of `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items["Catalan"]` in Catalan, `r n_items["Spanish"]` in Spanish, making up `r n_te` translation equivalents in total.
-Translation equivalents (TEs) received a median of `r printnum(sum_obs$median, digits = 0)` responses (*Min* = `r printnum(sum_obs$min, digits = 0)`, *Max* = `r printnum(sum_obs$max, digits = 0)`), both languages summed together.
+Our final dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one of `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items[["Catalan"]]` in Catalan, `r n_items[["Spanish"]]` in Spanish, making up `r n_te` translation equivalents in total.
+Translation equivalents (TEs) received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`), both languages summed together.
 
@@
 
-post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)
-
-# tidy predictor names
-str_repl <- c(
-    "b_Intercept[1]" = paste0("Comprehension and Production"),
-    "b_Intercept[2]" = paste0("Comprehension"),
-    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
-    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
-    "b_exposure_std" = paste0("Exposure (+1 SD, ",  round(sd(df$exposure), 2), ")"),
-    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
-    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
-    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
-    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
-)
-
-post_draws_list <- split(posterior_description, posterior_description$parameter)
+post_draws_list <- split(posterior_draws$summary, posterior_draws$summary$.variable)
 ```
@@
 # summarise posterior draws
-posterior_description %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(.variable_name = factor(parameter,
-                              levels = names(str_repl),
-                              labels = str_repl) %>%
-               as.character(),
-           type = ifelse(str_detect(parameter, "Intercept"), 
-                         paste0("Intercepts (at ", round(mean(df$age, 2)), " months)"),
-                         "Slopes"),
-           parameter = ifelse(str_detect(parameter, "Intercept"), 
-                              str_remove_all(parameter, "Intercept \\(|\\)"),
-                              parameter)) %>%
-    select(.variable_name, median, ci_high, ci_low, rope_percentage, type) %>%
-    gt(groupname_col = "type", rowname_col = "parameter") %>%
-    fmt_number(2:4) %>%
-    fmt_percent(5) %>%
-    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
+posterior_draws$summary |>
+    select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
+    gt(groupname_col = ".type", rowname_col = ".variable_name") |>
+    fmt_number(2:4) |>
+    fmt_percent(5) |>
+    cols_merge(c(.lower, .upper), pattern = "[{1}, {2}]") |>
     cols_label(
         .variable_name = "Parameter",
-        median = "Median",
-        ci_low = md("95% *HDI*"),
-        rope_percentage = md("*ROPE* prob.")
-    ) %>%
+        .median = "Median",
+        .lower = md("95% *HDI*"),
+        .rope_overlap = md("*ROPE* prob.")
+    ) |>
     tab_style(cell_text(weight = "bold"),
-              cells_column_labels(columns = 1:5)) %>%
+              cells_column_labels(columns = 1:5)) |>
     tab_style(cell_text(align = "left"),
-              cells_title(groups = "subtitle")) %>% 
+              cells_title(groups = "subtitle")) |> 
     tab_style(cell_text(size = "small"), 
@@
 
-The word exposure index ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]).
+The word exposure index ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$.upper, accuracy = 0.01)`]).
 All of the posterior samples of this regression coefficient excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/sd(df$exposure), accuracy = 0.01)` more likely to acquire it.
-
-The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
-
-However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to word.
+The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), accuracy = 0.01)` more likely to acquire it.
+
+The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$.rope_overlap, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
+
+However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to word.
 Follow-up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to word was low (e.g., -1 SD), phonological similarity increased the probability of acquisition substantially.
@@
 
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
 This shows that the effect of the word exposure index on the speed at which participants acquired words across ages was inconclusive: given our data, it is difficult to tell whether $Exposure$ increased or decreased the slope of the words' acquisition trajectories.
 
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of the word exposure index on the effect of phonological similarity did not differ substantially across ages.
+The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$.upper, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
+
+Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the impact of the word exposure index on the effect of phonological similarity did not differ substantially across ages.
 
@@
 Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
-nd <- datagrid(
-    model = model_fit_4,
-    age_std = seq_range(model_fit_4$data$age_std, n = 100),
-    lv_std = (seq(0, 1, 0.5)-mean(df$lv))/(sd(df$lv)),
-    exposure_std = c(-1, 0, 1),
-    te = NA,
-    id = NA
-)
-
-marg_epreds <- predictions(
-    model_fit_4,
-    newdata = nd,
-    ndraws = 25, 
-    re_formula = NA
-)
-
-marg_epreds %>% 
-    posteriordraws() %>% 
-    as_tibble() %>% 
-    select(drawid, group, draw, age_std, lv_std, exposure_std) %>%
-    filter(group != "No") %>% 
-    pivot_wider(
-        id_cols = c(drawid, age_std, lv_std, exposure_std), 
-        names_from = group, 
-        values_from = draw
-    ) %>% 
-    mutate(`Understands` = `Understands and Says` + `Understands`) %>% 
-    pivot_longer(c(`Understands`, `Understands and Says`), names_to = "group", values_to = "draw") %>% 
-    mutate(
-        exposure_std = factor(
-            exposure_std, 
-            levels = unique(nd$exposure_std), 
-            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
-        ),
-        lv_std = factor(
-            lv_std,
-            levels = unique(nd$lv_std), 
-            labels = paste0(c(0, 50, 100), "% similarity")
-        )
-    ) %>% 
-    ggplot(aes(age_std, draw, colour = lv_std, fill = lv_std)) +
-    facet_grid(group~exposure_std) +
+marginal_effects_epreds |> 
+    ggplot(aes(age, .value, colour = lv, fill = lv)) +
+    facet_grid(.category~exposure) +
     # stat_summary(fun.data = median_hdi, geom = "ribbon",
     #              alpha = 0.5, colour = NA) +
-    geom_line(aes(group = interaction(lv_std, drawid)),
+    geom_line(aes(group = interaction(lv, .draw)),
               linewidth = 0.75, alpha = 0.15) +
-    stat_summary(aes(linetype = lv_std), fun = median, geom = "line", linewidth = 1) +
+    stat_summary(aes(linetype = lv), fun = median, 
+                 geom = "line", linewidth = 1) +
     labs(
@@
     ) +
-    scale_x_continuous(
-        breaks = (seq(10, 32, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 32, 4)
-    ) +
+    scale_x_continuous(breaks = seq(10, 32, 4)) +
     theme(
@@
         panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
+        panel.border = element_rect(fill = NA, 
+                                    colour = "black", 
+                                    linewidth = 0.75)
     )
@@
 
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)`.
+For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`) would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.
 
@@
 The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
-model_loos %>%
-    loo_compare() %>%
-    as.data.frame() %>%
-    rownames_to_column("model") %>%
-    mutate(model = str_remove(model, "model_fit_")) %>%
+model_loos |>
+    loo_compare() |>
+    as.data.frame() |>
+    rownames_to_column("model") |>
+    mutate(model = str_remove(model, "model_fit_")) |>
     relocate(model,
@@
              matches("looic"),
-             matches("diff")) %>%
-    gt() %>%
-    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>%
-    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>%
-    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>%
-    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>%
-    fmt_number(2:9) %>%
+             matches("diff")) |>
+    gt() |>
+    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) |>
+    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) |>
+    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) |>
+    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) |>
+    fmt_number(2:9) |>
     cols_label(
@@
         se_diff = md("*SE*")
-    ) %>%
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>%
+    ) |>
+    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
     tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) %>%
+              cells_column_spanners()) |>
     tab_style(cell_text(align = "left"),
@@
 
-m <- add_epred_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA) %>% 
-    filter(.category != "No") %>% 
-    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
+m <- add_epred_draws(nd, 
+                     model_fit_4_prior,
+                     ndraws = NULL, 
+                     re_formula = NA) |> 
+    filter(.category != "No") |> 
+    pivot_wider(names_from = ".category", values_from = ".epred") |> 
+    mutate(Understands = Understands + `Understands and Says`) |> 
     pivot_longer(
@@
         values_to = ".epred"
-    ) %>% 
+    ) |> 
     mutate(
@@
 
-m %>% 
+m |> 
     ggplot(aes(x = age_std, y = .epred)) +
     facet_wrap(~.category) +
-    stat_lineribbon(linewidth = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
+    stat_lineribbon(linewidth = 1, 
+                    colour = NA,
+                    .width = c(0.95, 0.89, 0.67, 0.5)) +
     stat_summary(fun = "mean", geom = "line", linewidth = 0.75) +
     scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        breaks = scale(seq(10, 36, 4), 
+                       mean(responses$age), 
+                       sd(responses$age))[, 1],
         labels = seq(10, 36, 4)

### 2023-02-01 _targets.R

@@ -1,29 +1,33 @@
 library(targets)
 library(tarchetypes)
+library(conflicted)
+library(cli)
 
 # load R functions -------------------------------------------------------------
-source("R/utils.R")
-source("R/00_importing.R")
-source("R/01_preprocessing.R")
-source("R/02_models.R")
+function_paths <- list.files("R",
+                             pattern = ".R",
+                             full.names = TRUE)
+invisible(lapply(function_paths, source))
+
+test_paths <- list.files("tests/testthat",
+                         pattern = ".R",
+                         full.names = TRUE)
+invisible(lapply(test_paths, source))
 
 # list package dependencies ----------------------------------------------------
-
 tar_option_set(
     packages = c(
         "arrow",
-        "bayesplot",
         "bayestestR",
         "brms",
         "bvqdev",
         "childesr",
+        "cli",
         "conflicted",
         "dplyr",
-        "forcats",
         "ggplot2",
-        "ggsci",
+        "glue",
         "gt",
-        "here",
         "janitor",
         "keyring",
         "knitr",
@@ -31,11 +35,9 @@ tar_option_set(
         "marginaleffects",
         "mice",
         "patchwork",
-        "papaja",
         "purrr",
         "quarto",
         "readxl",
-        "rmarkdown",
         "rlang",
         "scales",
         "stringdist",
@@ -44,22 +46,11 @@ tar_option_set(
         "tibble",
         "tidybayes",
         "tidyr",
-        "usethis",
-        "wesanderson"
+        "usethis"
     )
 )
 
-# resolve conflicts
-conflicted::conflict_prefer("last_warnings", "rlang")
-conflicted::conflict_prefer("filter", "dplyr")
-conflicted::conflict_prefer("between", "dplyr")
-conflicted::conflict_prefer("timestamp", "utils")
-conflicted::conflict_prefer("ar", "brms")
-conflicted::conflict_prefer("chisq.test", "stats")
-conflicted::conflict_prefer("discard", "scales")
-conflicted::conflict_prefer("duration", "lubridate")
-conflicted::conflict_prefer("fisher.test", "stats")
-conflicted::conflict_prefer("lag", "dplyr")
+resolve_conflicts()
 
 # define global options --------------------------------------------------------
 options(
@@ -72,36 +63,23 @@ options(
 
 list(
     ## resolve namespace conflicts ---------------------------------------------
-    tar_target(resolve_conficts,
-               {
[…]

### 2023-02-01 _targets_packages.R
Update manuscript

@@ -1,34 +1,31 @@
 # Generated by targets::tar_renv(): do not edit by hand
 library(arrow)
-library(bayesplot)
 library(bayestestR)
 library(brms)
 library(bs4Dash)
 library(bvqdev)
 library(childesr)
+library(cli)
 library(clustermq)
 library(conflicted)
 library(dplyr)
-library(forcats)
 library(future)
 library(ggplot2)
-library(ggsci)
+library(glue)
 library(gt)
-library(here)
 library(janitor)
 library(keyring)
 library(knitr)
 library(lubridate)
+library(marginaleffects)
 library(markdown)
 library(mice)
-library(papaja)
 library(patchwork)
 library(pingr)
 library(purrr)
 library(quarto)
 library(readxl)
 library(rlang)
-library(rmarkdown)
 library(rstudioapi)
 library(scales)
 library(shiny)
@@ -42,4 +39,3 @@ library(tidybayes)
 library(tidyr)
 library(usethis)
 library(visNetwork)
-library(wesanderson)

### 2023-02-02 docs/index.qmd
Documents functions

@@
 
-The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
 This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$.upper, accuracy = 0.01)`]).
@@
 All of the posterior samples of his regression excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/(sd(responses$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.
-
-The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
+The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/(4*sd(responses$exposure)), accuracy = 0.01)` more likely to acquire it.
+
+The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$.rope_overlap, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
 
@@
 
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
+The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
 This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
@@
 
-## Marginal predictions
+## Marginal effects
 

### 2023-02-02 R/utils.R

@@ -283,3 +283,7 @@ save_files <- function(x,
     saveRDS(x, file_paths[grepl(".rds", file_paths)])
     cli_alert_success("Saved to {.path {folder}}")
 }
+
+remove_nul <- function() {
+    system("rename \\.\\C:\\Users\\U155880\\Documents\\trajectories\\nul. deletefile.txt")
+}

### 2023-02-16 manuscript/manuscript.qmd
Address comments

@@
     orcid: 0000-0002-8553-4209
-    email: gonzalo.garciadecastro@upf.edu
     affiliations:
-      - ref: upf
+      - Center for Brain and Cognition, Universitat Pompeu Fabra
     corresponding: true
@@
     affiliations:
-      - ref: upf
+      - Center for Brain and Cognition, Universitat Pompeu Fabra
   - name: Ignacio Castillejo
     affiliations:
-      - ref: uam
+      - Departamento de Psicología, Universidad Autónoma de Madrid
   - name: Núria Sebastian-Galles
     affiliations:
-      - ref: upf
-affiliations:
-  - id: upf
-    name: Center for Brain and Cognition, Universitat Pompeu Fabra
-  - id: uam
-    name: Departamento de Psicología, Universidad Autónoma de Madrid
+      - Center for Brain and Cognition, Universitat Pompeu Fabra
+
+affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
 
@@
 
-thanks: "This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
+abstract: |
+    Recent studies have reported larger vocabulary sizes for bilingual toddlers learning languages that shared a greater amount of cognates (e.g., English-Dutch).
+This effect was only observed in the language to which toddlers were exposed the least, and has been suggested to be the result of a cognateness facilitation effect during word acquisition.
+The specific underpinnings of tyhis effect, and its relationship with the degree of exposure to the language are unclear.
+In this study, we present vocabulary data from 436 Catalan-Spanish bilinguals and monolinguals aged 12 to 34.
+We used Bayesian Item Response Theory to estimate participants' probability of acquisition of 604 words, conditional to their word's cognate status and lexical frequency, and the child's age and degree of exposure to each language.
+We found converging evidence for an earlier age of acquisition for cognate words, mediated by word frequeny and language exposure.
+Low-frequency words, and words from the language of least exposure benefitted more strongly by their cognate status than high-frequency words.
+Our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition and processing of words in the other language.
+
+thanks: |
+    This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
 Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
@@
 Data collection was half-way when the COVID-19 pandemic started.
-We would like to pay special tribute to the families that collaborated with us under these difficult circumstances."
-keywords: lexical acquisition, vocabulary, bilingualism
+We would like to pay special tribute to the families that collaborated with us under these difficult circumstances.
+
+keywords: lexical acquisition, vocabulary, bilingualism, item response theory, bayesian
 toc: false
+citation: true
 fig-dpi: 1000
+
 csl: "../assets/apa7.csl"
 bibliography: "../assets/references.bib"
+
 format:
@@
     linestretch: 1.5
-    include-in-header: 
-      text: '\usepackage{lineno}\linenumbers'
+
 ---
@@
 library(patchwork)
+library(ggtext)
 library(lubridate)
@@
 
-# load and process objects
 # load targets as R objects
@@
 tar_load(items)
+tar_load(childes)
 tar_load(bvq_data)
+
 tar_load(model_fit_4)
+tar_load(posterior_draws)
+
+tar_load(model_fit_4_prior)
+tar_load(marginal_effects_epreds)
+tar_load(posterior_draws_doe)
+
 tar_load(model_log_liks)
-tar_load(model_fit_4_prior)
 tar_load(model_loos)
-tar_load(posterior_draws)
-tar_load(marginal_effects_epreds)
-
-items <- bvq_data$pool %>% 
-    mutate(item = str_remove(item, "cat_|spa_")) %>% 
-    select(item, language, semantic_category, class) %>% 
+
+tar_load(model_rhats)
+tar_load(model_neffs)
+
+items <- bvq_data$pool  |>  
+    mutate(item = str_remove(item, "cat_|spa_")) |> 
+    select(item, language, semantic_category, class) |> 
     right_join(items, by = c("item", "language"))
@@
 The foundations of word learning are in place early in age.
-From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthurbates].
+From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [@fenson2007macarthurbates, @samuelson2021precision].
 During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
 Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e.
-vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of acquired words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure; @bates1994developmental].
+vocabulary size, @fenson1994variability; @frank2021variability].
 Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
@@
 
-Despite the recent increase in the representativeness of the samples in the literature, most studies on early word acquisition have relied exclusively on data from monolingual children while neglecting, to a considerable extent, that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent].
-Previous findings pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
+Developmental psychology has increased the representativeness of its samples considerable in recent years, but most literature on early word acquisition relies extensively on data from monolingual children, and neglects to a considerable extent that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent].
+Previous research about bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
 @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals do.
@@
 Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-While these studies have mostly relied on samples of bilingual children learning two typologically relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
+These studies rely on bilingual samples whose languages are relatively distant in typology, as it is the case of English and Spanish.
+It is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
 What role could linguistic distance play during early vocabulary growth?
@@
 For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language.
-Depending on the linguistic distance between both languages, the two sets of words may overlap in form in varying degrees.
-When both languages are typologically close like Spanish and Catalan (both Romance languages) they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic languages, respectively).
+Depending on the similarity between both languages, the two sets of words may overlap in form in varying degrees.
+When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic languages, respectively).
 For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
@@
 Floccia et al.
-pointed to *parallel activation* as a phenomenon that might potentially underpin their results.
-The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension, and that this phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain].
+pointed to *parallel activation* as the underpinning mechanism behind their results.
+The parallel activation hypothesis stems from the language non-selective account of lexical access, and suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 
-This account is especially plausible in the light of (1) the evidence pointing to bilingual toddlers' lexicon consisting of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022translation; @pearson1995crosslanguage], and that (2) there is experimental evidence of parallel activation during word comprehension in children and toddlers [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
-Parallel activation is therefore a plausible explanation for Floccia et al.'s results: cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
+Regarding word acquisition, the parallel activation account is a plausible account for Floccia et al.'s results in the light of (1) the evidence pointing to bilingual toddlers' lexicon consisting of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage], and that (2) there is experimental evidence of parallel activation during word comprehension in children and toddlers [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
+It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
 This account is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
@@
 
-Evidence supporting an earlier age of acquisition for cognates is, to date, scarce nonetheless.
+Evidence supporting an earlier age of acquisition for cognates is nonetheless scarce to date.
 @bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
@@
 The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-On the other hand, Floccia et al.'s estimates are statically more reliable given their (remarkably larger) sample size.
-Their results suggest a cognateness effect on vocabulary size, but their study was not aimed at testing the effect of cognateness on age of acquisition at the word-level.
-Their response variable was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
-By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.
+On the other hand, Floccia et al.'s estimates are statiscally more reliable given their remarkably larger sample size.
+But although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level.* The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
+By aggregating the responses from all words into a single datum per child, information about the acquisition status of cognates vs.
 non-cognates was no longer available.
@@
 
-More recently, @mitchell2022cognates addressed this issue overcoming some of these pitfalls.
+More recently, @mitchell2022cognates addressed this issue providing a more fine grained measure of a  cognateness facilitation effect in vocabulary acquisition.
 Using a longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages.
@@
 
-## A mechanistic account for the cognate facilitation in word acquisition
-
-In this study, we suggest a mechanistic account for the cognate facilitation effect.
-We build this account from the assumption that bilinguals use fundamentally similar mechanisms than monolinguals to learn words.
-In both cases, learning a word requires at least two events to take place.
-First, a representation of the word-form, encoding its segmental information to sufficient detail, must be stored.
-
-This can be accomplished via bottom-up mechanism since early ages.
-Even if words are embedded in a continuous stream of speech sounds, in which there are no clear boundaries between words, infants can segment discrete word-forms.
-For instance, they may rely on their familiarity with the typical stress pattern [@jusczyk1995infants], the phonotactic constraints [@friederici1993phonotactic], or the distributional information of their native language [@saffran1996statistical] to detect word boundaries.
-A second event for word learning to take place is the association between the familiar word form and its referent.
-This task is far from trivial for many reasons.
-Referents are rarely labelled ostensively: children are exposed to most instances of a word in referentially ambiguous contexts or in the absence of their referent.
-Many words are also presented along remarkably variable exemplars of their referent, and even refer to abstract referents such as routines or play interactions (e.g., *uh-oh*; @casey2021moving).
-Infants nonetheless need few learning instances to create new word-referent associations.
-The extent to which they do so relying on bottom-up mechanisms such as cross-situational learning [@yu2007rapid], or on constraints and biases that guide their hypotheses about how words and referents are related [@golinkoff1994early], remains a matter of debate.
-Either way, there is evidence of word-referent links as early as six months of age [@bergelson2012months; @bergelson2015early].
-From a connectionist perspective, a word is said to be acquired when the association between its word-form and its conceptualised referent (i.e.
-semantic representation) has reached some level of associative strength.
-As infants accumulate experience with word-referent pairings, the association strength between both increases, and this leads to observable events such as infants' showing a stronger preference for a named object compared to unnamed competitors, or their parents being more likely to report such word as acquired in a vocabulary checklist.
-
-Bilinguals deal with a different linguistic landscape than monolinguals, but arguably use similar mechanisms and do so under common constraints [@curtin2011bilingual].
-Bilingual word acquisition entails similar events as described above, with the exception that many referents are
-
-they seem to reach their developmental goals of word learning at similar ages than monolinguals. Bilinguals store words
-
-## Two competing mechanisms for a cognate facilitation during word acquisition
-
-We argue that the cognateness facilitation effect takes place
-
-We argue that the facilitative effect of cognateness on word acquisition can be explained by at least two mechanisms, illustrated in @fig-hypotheses.
-On the one hand, and in line with @mitchell2022cognates, the effect of cognateness might take place during early stages of word acquisition, before stable representations of any of the two word-forms of the translation equivalent has been formed.
-In this scenario, the auditory recognition of any of the two word-forms might lead to the activation of its (cognate) translation in the other language via their phonological similarity.
-Since that both word-forms are likely to be heard in the presence of the same referent, the association between one of the word-forms and its referent could be strengthened after the child is exposed to that same word-form, or to its (similar-sounding) phonologically translation.
-
-For instance, in the presence of a tree, an English-Spanish bilingual child might hear *flower* in some occasions, and *flor* in some others.
-Given the phonological similarity between *flower* and *flor*, it is possible that both word-forms are activated in the child's lexicon every time any of them is encountered.
-This would lead to an increase in the association strength between each word-form and their common referent every time one of them is encountered, leading to an earlier age of acquisition.
-The case of a non-cognate translation equivalent, such as *cow* and *vaca* would be different.
-Although both word-forms are also encountered in the presence of the same referent, their low phonological similarity would not guarantee such facilitation when the child hears either of them, thus leading to a later age of acquisition for both.
-This account departs from the premise that a word-form can potentially activate its translation equivalent via phonology before acquisition, that is, before the association between such word-form and its referent is consolidated.
-This would allow both word-forms to mutually boost their acquisition after only few exposures.
-In this scenario, the effect of cognateness on acquisition should taken place early in age, and should be symmetric for both languages: the facilitation should occur in both word-forms simultaneously.
-
-An alternative account is that cognateness plays a role in the acquisition of a translation pair only when at least one of the word-forms has been acquired, that is, when then association between one of the word-forms and their common referent is consolidated.
-In this scenario, one of the word-forms can only activate the other via phonological similarity after its representation in the child's lexicon is robust enough.
-This way, *flower* and *flor* would not benefit from their phonological similarity until wither *flower* or *flor* is acquired.
-Only after, for instance, *flower* is acquired, the acquisition of *flor* would be boosted by its similarity to *flower*.
-This account predicts that cognateness has an asymmetric effect on cognate word acquisition: it should be observed only on the acquisition of the word-form of the translation equivalent that was acquired last.
-As a consequence, and given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+In this study, we provide an account for the cognate facilitation on word acquisition rooted in accumulator models of language acquisition [@hidaka2013computational; @mollica2017how].
+We build on the assumption that word acquisition is a function of mainly, but not exclusively, the amount of experience a child has accrued with the word-referent pairing [see @kachergis2022standard for review].
+As infants accumulate linguistic input, they also accumulate opportunities to associate familiar word-forms to their corresponding referents.
+Words with high lexical frequency like *flower* [`r floor(childes[childes$token=="flower",]$freq_million)` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as *headphones* (`r floor(childes[childes$token=="headphones",]$freq_million)` times per million words) [e.g., @goodman2008does]: infants might accumulate learning experiences with *flower* faster than with *headphones*.
+In the case of bilinguals, the amount of exposures to a word-form an infants receives is mediated by their degree of exposure to their languages.
+Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time might need longer to accumulate the same experience with *flower* and *headphones* than an English monolingual does (who listens 100% of the time to English).
+
+We suggest that the cognateness effect proposed by @floccia2018results and @mitchell2022cognates facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired.
+We hypothesise that both forms in the translation equivalent (e.g., *flower* and *flor*, for an English-Spanish bilingual) follow their acquisition trajectory in a similar way they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to.
+When one of the word-forms is acquired, the acquisition trajectory of the other is impacted, so that the number of needed learning instances left for acquisition are reduced.
+This facilitation would occur as the result of the semantic and phonological overlap between both words.
+When the child has acquired *flower*, and therefore its form-referent association has been formed, future presentations of the similar sounding translation *flor* in the presence of the same reference should lead to an increased likelihood of acquisition.
+In the case of a non-cognate translation equivalent, such facilitation would be unlikely.
+For instance, even if the child has already acquired the word *dog*, hearing its dissimilar Spanish translation *perro* would hardly lead to an increased probability of acquisition.
+In line with this account, only the later-acquired form of the translation equivalent would benefit from its cognate status.
+Given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age of acquisition for cognates than for non-cognates.
 
 ![Schematic representation of the hypothesised cognate facilitation effect on word acquisition.
-To the left, the acquisition of a cognate translation equivalent (*flower*-*flor*) is affected by the phonological similarity between both word forms.
+To the left, the acquisition of a cognate translation equivalent (*flower*-*flor*, which share four phonemes) is affected by the phonological similarity between both word forms.
 When the child hears *flower*, activation spreads across languages to activate *flor* too.
 As a result, after exposure to *flower* not only the association between *flower* and its referent is strenghtened, but also that of *flower* and its common referent.
-To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*) does not undergo the such facilitation.
-Both word-forms are fairly disimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](img/hypotheses.jpg){#fig-hypotheses}
-
-In line with this second hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al.
+To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*, which do not share any phoneme) does not undergo the such facilitation.
+Both word-forms are fairly dissimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](img/hypotheses.jpg){#fig-hypotheses}
+
+In line with this hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al.
 reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure).
@@
 Therefore, participants may have on average learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
-Under the second hypothesis, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
+Under this hypothesis, because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
 
@@
 
-In this study, we investigated the role of cognateness on the acquisition of translation equivalents, and if such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair.
-Using an online vocabulary checklist designed specifically for this study we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We adopted an Bayesian item response theory (*IRT*) approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to the word-form's cognate status, and the participant's frequency of exposure to the word-form (a composite measure of lexical frequency and language degree of exposure), while adjusting for other indices of word difficulty like lexical frequency [@goodman2008does] and length (phonemes), and participant skill like age [see @kachergis2022standard for a similar approach].
+In this study, we tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a words *only* once its translation has been already acquired.
+Using an online vocabulary checklist designed specifically for this study, we collected vocabulary data on production and comprehension from a sample of bilingual children.
+Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees.
+We adopted a Bayesian item response theory [*IRT*, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
+
+We operationalised participants' exposure rate to a word-form using a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it when listening to speech), and  the child's amount of exposure to the language such word belongs to.
+This measure was calculated by multiplying the word-form's lexical frequency with the participant's degree of exposure to the language the word belongs to.
+Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childes] (see Appendix 1 for more details).
+Participant's degree of exposure to Catalan and Spanish were reported by families in a language exposure questionnaire they filled before the vocabulary checklist, in which they provided a proportion of the time that the child was estimated to listen to any language they were spoken to regularly.
+The resulting measure is a exposure-weighted lexical frequency measure, which was included in analyses as a proxy of how often bilingual participants were exposed to each word-form.
+
+We defined *cognateness* as a continuous measure of phonological similarity between translation equivalents.
+We quantified the phonological similarity between each pair of word-forms by computing the inverse of the normalised Levenshtein distance between their SAMPA phonological transcriptions [@levenshtein1966binary].
+This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
+This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
+For instance, this measure of phonological similarity returns 0% for *ocell* (`uceL`) and *pájaro* (`paxa4o`), Catalan and Spanish for 'bird', and returns 50% for *lluna* (`Lun5`, /ˈʎu.nə/) and *luna* (`luna`, /ˈlu.na/), Catalan and Spanish for 'moon' (see Appendix 1 for more details).
 
@@
 
-n_item_language <- bvq_data$pool %>% 
+n_item_language <- bvq_data$pool |> 
     count(language) |>
@@
     set_names(c("catalan", "spanish"))
+
+n_items_version <- bvq_data$pool |>
+    unnest(version) |> 
+    count(version, language) |> 
+    pivot_wider(names_from = language,
+                values_from = n) |> 
+    clean_names() |> 
+    mutate(version = tolower(version))
+
 ```
@@
 This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formR platform [@arslan2020formr].
-This questionnaire is structured in three blocks: (1) a language questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
+This questionnaire is structured in three blocks: (1) a language exposure questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
 Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
+Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for 'dog').
 In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
@@
 
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-In total, each response to the questionnaire provided data for 400 translation equivalents, that is, for 800 words: 400 in Catalan and 400 in Spanish.
-Participants were randomly allocated into one of four different subsets of the complete list of items.
-These lists were designed so that each contained a representative sub-sample of the items from the complete list.
+For each word included in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
+Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items.
+Each subset contained a random but representative sub-sample of the items from the complete list.
 Semantic/functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
 Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-For the analyses in the present study, we considered responses nouns [@fourtassi2020growth].
-We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono/mico*).
+The resulting reduced vocabulary checklists contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words.
+Therefore, each response to the questionnaire provided data for a minimum of `r min(n_items_version$catalan, n_items_version$spanish)` and a maximum of `r max(n_items_version$catalan, n_items_version$spanish)` translation equivalents.
+Participants were randomly allocated into one of four subset, and always the same subset every time they participated.
 
@@
 
+diff_responses <- responses |>
+    distinct(id, age) |>
+    left_join(distinct(participants, id, time_stamp, age),
+              by = c("id", "age")) |> 
+    group_by(id) |> 
+    arrange(id, time_stamp) |> 
+    mutate(diff_lag = time_stamp-lag(time_stamp, default = NA)) |> 
+    summarise(diff_min = min(diff_lag, na.rm = TRUE),
+              diff_max = max(diff_lag, na.rm = TRUE),
+              .groups = "drop") |> 
+    filter(!(diff_min %in% c(-Inf, Inf))) |>
+    summarise(diff_min = min(diff_min, na.rm = TRUE),
+              diff_max = max(diff_max, na.rm = TRUE))
+
 n_responses <- nrow(distinct(responses, id, age))
 
-dates <- participants$time_stamp %>% 
-    range(na.rm = TRUE) %>% 
+dates <- participants$time_stamp |> 
+    range(na.rm = TRUE) |> 
     as_date()
@@
 
-We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
-`r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
-Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan [see @tbl-participants for a summary of the distribution of participants across ages and degrees of exposure to Catalan].
+We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between the `r format(dates[1], "%dth of %B, %Y")` and the `r format(dates[2], "%dth of %B, %Y")`: `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
+Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`.
+Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample).
 All families gave informed consent before participating.
-Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate.
+Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser.
 Filling the questionnaire took 30 minutes approximately.
@@
     mutate_at(vars(starts_with("doe_")), 
-              (\(x) floor(x * 10) / 10)) %>% 
+              (\(x) floor(x * 10) / 10)) |> 
     mutate(
@@
 
-We used the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain.
+We used the highest self-reported educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker.
+The following options were given: *sense escolaritzar/sin escolarizar* [no education], *educació primària/educación primaria* [primary school], *ducació secundària/educación secundaria* [secondary school], *batxillerat/bachillerato* [complementary studies/high school], *cicles formatius/ciclos formativos* [vocational training], and *educació universitària/educación universitaria* [university degree], in line with the current educational system in Spain.
 Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r percent(table(participants$edu_parent)["University"]/length(unique(participants$id)))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))`) and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
-
-## Variables of interest
-
-Each observation in the resulting dataset was matched with its corresponding values in four variables of interest that we used to model participants' responses:
-
-(1) Participant age ($Age$): number of months elapsed between participants' birth date and questionnaire completion.
-
-(2) Word length ($Phonemes$): number of phonemes in the phonological transcription of the word-form in Speech Assessment Methods Phonetic Alphabet (SAMPA) format.
-
-(3) Word exposure index ($Exposure$): this is a composite measure that captures the frequency with which the participant is exposed to a given word from a given language (Catalan or Spanish), weighted by the child's amount of exposure to that language.
-This measure was calculated by multiplying the word-form's lexical frequency (extracted from the CHILDES corpora, @macwhinney2000childes; @sanchez2019childes) with the participant's degree of exposure to the language the word belongs to (see @sec-appendix-predictors for more details).
-
-(4) Phonological similarity ($Levenshtein$): normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary].
-This score is computed by counting the number of insertions, deletions and replacements needed both transcriptions to become identical, then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
-This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches] (see @sec-appendix-predictors for more details).
 
@@
     count(te) |>
-    summarise(across(n, lst(median, min, max), .names = "{.fn}"))
+    summarise(across(n, lst(median, min, max), 
+                     .names = "{.fn}"))
 
@@
 
-rope_interval <- c(lower = -0.1, upper = 0.1)
+rope_interval <- percent(c(lower = -0.1/4, upper = 0.1/4), accuracy = 0.1)
 
@@
 
-Our final dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one of `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items[["Catalan"]]` in Catalan, `r n_items[["Spanish"]]` in Spanish, making up `r n_te` translation equivalents in total.
-Translation equivalents (TEs) received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`), both languages summed together.
+For the analyses in the present study, we considered responses to nouns.
+[@fourtassi2020growth].
+We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono/mico* [monkey]).
+This resulting in `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items[["Catalan"]]` in Catalan, `r n_items[["Spanish"]]` in Spanish, making up `r n_te` translation equivalents in total.
+The final dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one word.
+Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`), both languages summed together.
 
 We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_4$fit)[2]` sampling chains with `r format(dim(model_fit_4$fit)[1], big.mark = ",")` iterations each).
-This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$, $Phonemes$, $Exposure$, and $Levenshtein$.
-All predictors were standardised before entering the model for an easier interpretation of their regression coefficients.
-We also included participant and item random intercepts and slopes when appropriate [@barr2013random].
-We compared a baseline model $\mathcal{M}_{0}$, which only included the intercept, $Age$, $Phonemes$, and $Exposure$ as predictors, with increasingly more complex models, introducing predictors in the following order: $Age \times Exposure$ ($\mathcal{M}_{1}$), $Levenshtein$ ($\mathcal{M}_{2}$), $Exposure \times Levenshtein$ ($\mathcal{M}_{3}$), and finally $Age \times Levenshtein$ and $Age \times Exposure \times Levenshtein$ ($\mathcal{M}_{4}$).
+This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$ (number of months elapsed between participants' birth date and questionnaire completion), $Length$ (number of phonemes in the SAMPA
+phonological transcription of the word-form), $Exposure$ (composite measure of lexical frequency and language degree of exposure), and $Levenshtein$ (phonological similarity between translation equivalents).
+For an easier interpretation of the regression coefficients, the values of all predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
+We also included participant and item random intercepts and slopes when appropriate [@barr2013random]
+
+We compared a baseline model $\mathcal{M}_{0}$, which only included the intercept, $Age$, $Length$, and $Exposure$ as predictors, with increasingly more complex models, introducing predictors in the following order: $Age \times Exposure$ ($\mathcal{M}_{1}$), $Levenshtein$ ($\mathcal{M}_{2}$), $Exposure \times Levenshtein$ ($\mathcal{M}_{3}$), and finally $Age \times Levenshtein$ and $Age \times Exposure \times Levenshtein$ ($\mathcal{M}_{4}$).
 We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO, @vehtari2017practical]^[Due to computational constraints, we performed LOO based on a random sub-sample of `r format(dim(model_log_liks$model_fit_4)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]], and selected for interpretation the one showing the best performance.
-More details about the models can be found in @sec-appendix-model.
+More details about the models can be found in Appendix 3.
 
@@
 
-The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the second best-performing model ($\mathcal{M}_{3}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`.
-Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see @sec-appendix-loo for all model comparisons).
-@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$.
-The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor.
+All models showed adequate convergence diagnostics and little evidence of multicollinearity between predictors (see Appendix 4 for details).
+The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $Exposure$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the second best-performing model ($\mathcal{M}_{3}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`.
+Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see Appendix 3 for all model comparisons).
+@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$, and their degree of overlap with the ROPE.
+For interpretability, we report the estimated regression coefficients transformed to the probability scale.
+The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales relate non-linearly.
+This means that one logit difference is not necessarily translated to a unique value in the probability scale.
+For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 30 to 32 months.
+The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients in the probability scale is not trivial.
+Following @gelman2020regression, we report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
+derivative) of the logistic curve at the age at which most participants were acquiring a given word.
+This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
 
@@
 #| label: tbl-coefs
-#| tbl-cap: "Posterior distribution of regression coefficients.
+#| tbl-cap: "Posterior distribution of regression coefficients in the probability scale.
 *Median*: median of the posterior distribution in the logit scale.
 *95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE, [-0.1, +0.1]).
+On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE).
 This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
@@
     select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
-    gt(groupname_col = ".type", rowname_col = ".variable_name") |>
-    fmt_number(2:4) |>
-    fmt_percent(5) |>
-    cols_merge(c(.lower, .upper), pattern = "[{1}, {2}]") |>
+    mutate(
+        across(.median:.upper,
+               function(x) {
+                   ifelse(
+                       .type=="Intercepts (at 22 months)",
+                       plogis(x),
+                       x/4
+                   )
+               }
+        )) |> 
+    gt(groupname_col = ".type", 
+       rowname_col = ".variable_name") |>
+    fmt_percent(c(.median, .rope_overlap)) |>
+    fmt_number(.lower:.upper, scale_by = 100) |> 
+    cols_merge(c(.lower, .upper),
+               pattern = "[{1}, {2}]") |>
     cols_label(
@@
     tab_style(cell_text(size = "small"), 
-              list(cells_body(), cells_row_groups(), cells_stub()))
+              list(cells_body(),
+                   cells_row_groups(),
+                   cells_stub())) |> 
+    tab_style(cell_text(style = "italic"),
+              cells_row_groups())
 ```
@@
 Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
-Finally, predictions are presented separately for different degrees of word exposure index: little exposure to the word (-1 SD), median exposure, and high exposure (+1 SD).
-Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
+Predictions are presented separately for different degrees of word exposure index: little exposure to the word (-1 SD), median exposure, and high exposure (+1 SD).
+Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom.
+In-sample predictions lie inside the grey rectangles."
+
+clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")
+
 marginal_effects_epreds |> 
+    mutate(
+        age = rescale_variable(age_std, mean(responses$age), sd(responses$age)),
+        exposure = factor(
+            exposure_std,
+            levels = unique(exposure_std),
+            labels = paste0("Exposure: ", c("-1 SD", "mean", "+1 SD")),
+            ordered = TRUE
+        ),
+        lv = factor(
+            lv_std,
+            levels = unique(lv_std),
+            labels = paste0(c(0, 50, 100), "% similarity"),
+            ordered = TRUE
+        )
+    ) |> 
     ggplot(aes(age, .value, colour = lv, fill = lv)) +
@@
     #              alpha = 0.5, colour = NA) +
+    annotate(geom = "rect",
+             xmin = min(responses$age),
+             xmax = max(responses$age),
+             ymin = -Inf,
+             ymax = Inf,
+             fill = "grey90",
+             colour = NA) +
+    geom_vline(xintercept = mean(responses$age), colour = "grey",
+               linewidth = 1) +
+    geom_hline(yintercept = 0.5,
+               colour = "grey",
+               linewidth = 1) +
     geom_line(aes(group = interaction(lv, .draw)),
               linewidth = 0.75, alpha = 0.15) +
-    stat_summary(aes(linetype = lv), fun = median, 
-                 geom = "line", linewidth = 1) +
+    stat_summary(fun = median, geom = "line", linewidth = 1) +
     labs(
@@
         colour = "Phonological similarity (Levenshtein)",
-        fill = "Phonological similarity (Levenshtein)"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
+        fill = "Phonological similarity (Levenshtein)",
+        subtitle = glue("**Phonological similarity**<br> <span style = 'color:{clrs_reorder[5]};'>0% similarity</span>   <span style = 'color:{clrs_reorder[4]};'>50% similarity</span>   <span style = 'color:{clrs_reorder[1]};'>100% similarity</span>")) +
+    scale_colour_manual(values = clrs[c(5, 4, 1)]) +
+    scale_fill_manual(values = clrs[c(5, 4, 1)]) +
     scale_y_continuous(
@@
     ) +
-    scale_x_continuous(breaks = seq(10, 32, 4)) +
+    scale_x_continuous(breaks = seq(5, 50, 5),
+                       limits = c(10, 50)) +
     theme(
-        legend.position = "top",
+        legend.position = "none",
         panel.grid = element_blank(),
@@
                                     colour = "black", 
-                                    linewidth = 0.75)
-    )
+                                    linewidth = 0.75),
+        plot.subtitle = element_textbox_simple(
+            fill = "grey90",
+            padding = margin(8, 8, 8, 8),
+            margin = margin(0, 0, 10, 0)
+        ))
 ```
@@
 We then used item response theory (*IRT*) to estimate the probability of each word-form being acquired by a child---either as comprehension or production---as a function of the child's $Age$, their $Exposure$ to the word-form (operationalised as a composite measure of lexical frequency and language exposure), and the phonological similarity of the word-form with its translation in the other language ($Levenshtein$).
-We also added the number of $Phonemes$ in the word-form as a nuance predictor the increase the predictive performance of the model.
+We also added the $Length$  of the word-form as a nuance predictor the increase the predictive performance of the model.
 
@@
 For instance, $sol$ (Catalan for *sun*) was acquired earlier than *cocodril* (Catalan for *crocodile*).
-Shorter word forms consist on fewer phonemes and, frequently, fewer syllables.
+Shorter word-forms consist on fewer phonemes and, frequently, fewer syllables.
 This makes them simpler to encode as phonological representations in the lexicon, and therefore easier to process.
@@
 Words with higher lexical frequency are more likely to be uttered in the presence of the child, while words with lower lexical frequency appear less often.
-In the case of bilinguals, the exposure rate to a given word not only depends on lexical frequency, but also on the child's amount of exposure to the language such word belongs to [@cattani2014how; @floccia2018iii].
+In the case of bilinguals, the exposure rate to a given word not only depends on lexical frequency, but also on the child's amount of exposure to the language such word belongs to [@cattani2014how; @floccia2018results].
 For instance, a Spanish dominant bilingual child who is exposed to Catalan only 10% of the time will rarely hear the Catalan word *casa*, despite its high lexical frequency.
@@
 
-Finally, we found a facilitation effect of cognateness, providing converging evidence with previous studies [@floccia2018iii; @mitchell2022cognates; @bosch2014first; @schelletter2002effect].
+Finally, we found a facilitation effect of cognateness, providing converging evidence with previous studies [@floccia2018results; @mitchell2022cognates; @bosch2014first; @schelletter2002effect].
 Critically, this effect was mediated by exposure: only words to which children were exposed to less frequently benefited from their cognate status.
 These results favour the account that a word-form benefits from its phonological similarity with its translation only after such translation has been acquired first.
-On the one hand, our results provide converging evidence for @floccia2018iii, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample.
+On the one hand, our results provide converging evidence for @floccia2018results, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample.
 Most children in their sample were English-dominant, and therefore their exposure to English was higher than to the additional language.
@@
 Bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese).
-Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018iii's sample.
+Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018results's sample.
 
@@
 ::: {#refs}
+
 :::
 
-#  Appendix {#sec-appendix}
-
-## Appendix 1: predictors {#sec-appendix-predictors}
+# Appendix 1: Model predictors {#sec-appendix-predictors .appendix}
 
@@
 
-## Appendix 2: Model comparisons {#sec-appendix-loo}
-
-```{r}
-#| label: tbl-results-loos
-#| echo: false
-#| message: false
-#| warning: false
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
-The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
-The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
-model_loos |>
-    loo_compare() |>
-    as.data.frame() |>
-    rownames_to_column("model") |>
-    mutate(model = str_remove(model, "model_fit_")) |>
-    relocate(model,
-             matches("elpd_loo"),
-             matches("p_loo"),
-             matches("looic"),
-             matches("diff")) |>
-    gt() |>
-    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) |>
-    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) |>
-    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) |>
-    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) |>
-    fmt_number(2:9) |>
-    cols_label(
-        model = "Model",
-        # elpd
-        elpd_loo = md("*ELPD*"),
-        se_elpd_loo = md("*SE*"),
-        # p
-        p_loo = md("*p*"),
-        se_p_loo = md("*SE*"),
-        # looic
-        looic = md("*LOO-IC*"),
-        se_looic = md("*SE*"),
-        # diff
-        elpd_diff = md("*diff*"),
-        se_diff = md("*SE*")
-    ) |>
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) |>
-    tab_style(cell_text(align = "left"),
-              cells_title(groups = "subtitle"))
-```
-
-## Appendix 3: Model details {#sec-appendix-model}
+# Appendix 2: Model details {#sec-appendix-model .appendix}
 
@@
 
-To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
+To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), word-form length ($Length$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
 We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
@@
 & (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
-& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Phonemes}_{ij} + & \\
+& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Length}_{ij} + & \\
 & (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
@@
 
-## Appendix 4: do results change when including frequency and language exposure as separate predictors?
+# Appendix 3: model comparisons {.appendix}
+
+```{r}
+#| label: tbl-results-loos
+#| echo: false
+#| message: false
+#| warning: false
+#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
+The predictive accuracy of the models was compared using leave-one-out cross-validation.
+The likelihood of each data point given a model fitted without including that same observation was estimated.
+The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
+*ELPD* values closer to zero indicate better predictive acuracy.
+We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
+The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
+model_loos |>
+    loo_compare() |>
+    as.data.frame() |>
+    rownames_to_column("model") |>
+    mutate(model = str_remove(model, "model_fit_")) |>
+    relocate(model,
+             matches("elpd_loo"),
+             matches("p_loo"),
+             matches("looic"),
+             matches("diff")) |>
+    gt() |>
+    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) |>
+    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) |>
+    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) |>
+    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) |>
+    fmt_number(2:9) |>
+    cols_label(
+        model = "Model",
+        # elpd
+        elpd_loo = md("*ELPD*"),
+        se_elpd_loo = md("*SE*"),
+        # p
+        p_loo = md("*p*"),
+        se_p_loo = md("*SE*"),
+        # looic
+        looic = md("*LOO-IC*"),
+        se_looic = md("*SE*"),
+        # diff
+        elpd_diff = md("*diff*"),
+        se_diff = md("*SE*")
+    ) |>
+    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |>
+    tab_style(cell_text(align = "left"),
+              cells_title(groups = "subtitle"))
+```
+
+# Appendix 4: Convergence diagnostics {#sec-appendix-diagnostics .appendix}
+
+## Gelman-Rubin diagnostic
+
+```{r}
+#| label: fig-model-rhats
+#| fig-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-height: 3
+#| fig-width: 8
+map_df(model_rhats, bind_rows, .id = "model") |> 
+    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
+    pivot_longer(-model, 
+                 names_to = ".variable",
+                 values_to = ".rhat") |>
+    filter(str_detect(.variable, "b_|sd_")) |> 
+    drop_na() |> 
+    ggplot(aes(.rhat, colour = model, fill = model)) +
+    facet_wrap(~model, nrow = 1) +
+    annotate(geom = "rect",
+             fill = "grey80",
+             xmin = 1,
+             xmax = 1.01,
+             ymin = -Inf,
+             ymax = Inf) +
+    geom_histogram(bins = 10, colour = "white") +
+    geom_rug(alpha = 0.5) +
+    labs(x = "Gelman Rubin diagnostic (R-hat)",
+         y = "Number of MCMC samples") +
+    theme(legend.position = "none")
+```
+
+## Effective sample size
+
+```{r}
+#| label: fig-model-ess
+#| fig-cap: "Ratio of effective sample size in the MCMC chains in the Bayesian models."
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-height: 3
+#| fig-width: 8
+map_df(model_neffs, bind_rows, .id = "model") |> 
+    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
+    pivot_longer(-model, 
+                 names_to = ".variable",
+                 values_to = ".rhat") |>
+    filter(str_detect(.variable, "b_|sd_")) |> 
+    drop_na() |> 
+    ggplot(aes(.rhat, colour = model, fill = model)) +
+    facet_wrap(~model, nrow = 1) +
+    geom_vline(xintercept = 0.1, colour = "grey", linewidth = 0.75) +
+    geom_histogram(bins = 10, colour = "white") +
+    geom_rug(alpha = 0.5) +
+    labs(x = "Gelman Rubin diagnostic (R-hat)",
+         y = "Number of MCMC samples") +
+    scale_x_continuous(labels = percent) +
+    theme(legend.position = "none")
+```
+
+# Appendix 5: Frequency and language exposure as separate predictors {#sec-appendix-doe  .appendix}
+
+```{r}
+#| label: fig-coefs-doe
+#| fig-cap: "Posterior distribution of regression coefficients in the probability scale.
+*Median*: median of the posterior distribution in the logit scale.
+*95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
+On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE).
+This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
+Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-height: 4
+#| fig-width: 10
+# summarise posterior draws
+bind_rows("Exposure model" = posterior_draws$summary,
+          "DoE-Frequency model" = posterior_draws_doe,
+          .id = "model") |> 
+    mutate(
+        across(c(.median, .lower, .upper),
+               function(x) {
+                   ifelse(
+                       .type=="Intercepts (at 22 months)",
+                       plogis(x),
+                       x/4
+                   )
+               }
+        )) |>
+    filter(.type == "Slopes") |> 
+    ggplot(aes(.median, .variable_name, xmin = .lower, xmax = .upper,
+               colour = model)) +
+    facet_wrap(~model, scales = "free") +
+    annotate(geom = "rect",
+             ymin = -Inf,
+             ymax = Inf,
+             xmin = -0.1/4,
+             xmax = +0.1/4,
+             fill = "grey70") +
+    geom_errorbar(linewidth = 1, width = 0.25) +
+    geom_point(size = 2.5) +
+    labs(x = "Median and 95% HDI of the posterior distribution",
+         y = "Predictor",
+         colour = "Model") +
+    scale_x_continuous(labels = percent) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(legend.position = "none",
+          axis.title.y = element_blank())
+```

### 2023-02-16 manuscript/img/diagram.R

@@ -1,0 +1,102 @@
+logistic <- function(x, slope, mid, upper) {
+    upper / (1 + exp(-slope * (x - mid)))
+}
+
+
+x <- seq(6, 30, 0.1)
+slope <- c(0.75)
+mid <- c(20, 24, )
+upper <- c(1)
+
+logistic_data <- expand_grid(x, slope, mid, upper) %>% 
+    mutate(y = logistic(x, slope, mid, upper),
+           values = paste0("slope = ", round(slope, 2),
+                           ", mid = ", round(mid, 2)))
+
+ggplot(logistic_data, aes(x, y, colour = values)) +
+    facet_wrap(~upper, labeller = labeller(upper = ~paste0("upper = ", .))) +
+    geom_line(size = 1.25) +
+    labs(colour = "Parameter values") +
+    guides(colour = guide_legend(ncol = 2)) +
+    scale_color_manual(values = clrs[c(1, 4, 5)]) +
+    theme(panel.grid = element_line(linetype = "dotted", colour = "grey"),
+          panel.grid.minor = element_blank(),
+          legend.position = "top",
+          legend.title = element_blank(),
+          legend.justification = c("left", "top"))
+
+
+age <- seq(6, 32, 1)
+language <- c("English", "Spanish", "English", "Spanish")
+te <- c("flower-flor", "flower-flor", "cow-vaca", "cow-vaca")
+word <- c("flower", "flor", "cow", "vaca")
+doe <- c(0.80, 0.20, 0.80, 0.20)
+lambda <- c(5, 10, 5, 10)
+
+thres <- 30
+
+elis_df <- tibble(te, word, language, doe, lambda, lambda_wgt = lambda*doe) |> 
+    expand_grid(age) |> 
+    mutate(eli = rpois(n(), lambda_wgt)) |> 
+    group_by(te, word) |> 
+    mutate(eli_cum = cumsum(eli)) |> 
+    group_by(te) |> 
+    mutate(aoa_te = age[which.min(abs(eli_cum-thres))]) |> 
+    group_by(word) |> 
+    mutate(
+        aoa = age[which.min(abs(eli_cum-thres))],
+        eli_cum_corrected = ifelse(
+            (aoa > aoa_te) & (age >= aoa_te) & (word=="flor"),
+            eli_cum*1.5,
+            eli_cum
+        ),
+        aoa_corrected = ifelse(aoa > aoa_te,
[…]

### 2023-02-16 docs/index.qmd
Update report (still not functional) @

@@
 number-sections: true
-bibliography: "references.bib"
-csl: "apa7.csl"
+csl: "../assets/apa7.csl"
+bibliography: "../assets/references.bib"
 ---
@@
 tar_load(posterior_draws_re)
+tar_load(marginal_effects_epreds)
 

### 2023-02-16 manuscript/manuscript.qmd
Fix Appendix

@@
 ---
-title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual word acquisition"
+title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual lexical acquisition"
 editor: source
@@
 abstract: |
-    Recent studies have reported larger vocabulary sizes for bilingual toddlers learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-This effect was only observed in the language to which toddlers were exposed the least, and has been suggested to be the result of a cognateness facilitation effect during word acquisition.
-The specific underpinnings of tyhis effect, and its relationship with the degree of exposure to the language are unclear.
-In this study, we present vocabulary data from 436 Catalan-Spanish bilinguals and monolinguals aged 12 to 34.
-We used Bayesian Item Response Theory to estimate participants' probability of acquisition of 604 words, conditional to their word's cognate status and lexical frequency, and the child's age and degree of exposure to each language.
-We found converging evidence for an earlier age of acquisition for cognate words, mediated by word frequeny and language exposure.
-Low-frequency words, and words from the language of least exposure benefitted more strongly by their cognate status than high-frequency words.
-Our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition and processing of words in the other language.
+    One of the most prominent features of the bilingual lexicon is the availability of word-form representations from one language to those from the other language.
+One instance of such phenomenon is that fact that the form overlap between a word-form and its translation (i.e., cognateness) impacts how bilinguals process it.
+Recent studies have suggested that cognateness facilitates vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and who also acquire cognates earlier in age than non-cognates.
+The specific mechanisms underpinning such facilitation effect are unclear.
+In this study, we present a model of bilingual early lexical acquisition in which cognateness interacts with lexical frequency and language exposure to facilitate the acquisition of low-frequency words.
+We tested this model against vocabulary data from 436 Catalan-Spanish bilinguals aged 12 to 34 months.
+We used Bayesian Item Response Theory to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
+We found converging evidence for an earlier age of acquisition for cognate words, and for such effect being mediated by lexical frequency and language exposure.
+Low-frequency words, and words from the language of least exposure benefited more strongly by their cognate status than high-frequency words.
+Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition and processing of words in the other language.
 
@@
     linestretch: 1.5
+  docx: 
+    reference-doc: "../assets/templates/template.docx"
 
@@
 In this study, we provide an account for the cognate facilitation on word acquisition rooted in accumulator models of language acquisition [@hidaka2013computational; @mollica2017how].
-We build on the assumption that word acquisition is a function of mainly, but not exclusively, the amount of experience a child has accrued with the word-referent pairing [see @kachergis2022standard for review].
+We build on the assumption that word acquisition is a function of mainly---but not exclusively---the amount of experience a child has accrued with the corresponding word-referent pairing [see @kachergis2022standard for review].
 As infants accumulate linguistic input, they also accumulate opportunities to associate familiar word-forms to their corresponding referents.
-Words with high lexical frequency like *flower* [`r floor(childes[childes$token=="flower",]$freq_million)` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as *headphones* (`r floor(childes[childes$token=="headphones",]$freq_million)` times per million words) [e.g., @goodman2008does]: infants might accumulate learning experiences with *flower* faster than with *headphones*.
+Words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'headphones' (`r format(floor(childes[childes$token=="headphones",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does].
+Infants might accumulate learning experiences with 'flower' faster than with 'headphones'.
 In the case of bilinguals, the amount of exposures to a word-form an infants receives is mediated by their degree of exposure to their languages.
-Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time might need longer to accumulate the same experience with *flower* and *headphones* than an English monolingual does (who listens 100% of the time to English).
+Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time might need longer to accumulate the same experience with 'flower' and 'headphones' than an English monolingual does (who listens 100% of the time to English).
 
@@
 #| echo: false
+
+rhats_df <- map_df(model_rhats, bind_rows, .id = "model") |> 
+    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
+    pivot_longer(-model, 
+                 names_to = ".variable",
+                 values_to = ".rhat") |>
+    filter(str_detect(.variable, "b_|sd_")) |> 
+    drop_na()
+
 loo_diff <- loo_compare(model_loos)[2, 1]
@@
         fill = "Phonological similarity (Levenshtein)",
-        subtitle = glue("**Phonological similarity**<br> <span style = 'color:{clrs_reorder[5]};'>0% similarity</span>   <span style = 'color:{clrs_reorder[4]};'>50% similarity</span>   <span style = 'color:{clrs_reorder[1]};'>100% similarity</span>")) +
-    scale_colour_manual(values = clrs[c(5, 4, 1)]) +
-    scale_fill_manual(values = clrs[c(5, 4, 1)]) +
+        subtitle = glue("**Phonological similarity**<br> <span style = 'color:{clrs_reorder[5]};'>0% similarity</span>   <span style = 'color:{clrs_reorder[4]};'>50% similarity</span>   <span style = 'color:{clrs_reorder[3]};'>100% similarity</span>")) +
+    scale_colour_manual(values = clrs_reorder[c(5, 4, 3)]) +
+    scale_fill_manual(values = clrs_reorder[c(5, 4, 3)]) +
     scale_y_continuous(
@@
 
-# Appendix 1: Model predictors {#sec-appendix-predictors .appendix}
+{{< pagebreak >}}
+
+# Appendix 1: model predictors {#sec-appendix-predictors .appendix}
 
@@
 Nonetheless, they come from different Proto-Indo-European roots [@campbell2007glossary].
-This does not keep bilinguals from processing such kind of word-forms word forms differently than other translation equivalents with no form-similarity like *dog* and *perro*, in English and Spanish.
+This does not keep bilinguals from processing such kind of word-forms word forms differently than other translation equivalents with no form-similarity like 'dog' and *perro*, in English and Spanish.
 In the scope of this experiment, we consider any form-similar translation equivalent as a *cognate* pair, regardless of whether both word-forms share etymological origin.
@@
 
-# Appendix 2: Model details {#sec-appendix-model .appendix}
+{{< pagebreak >}}
+
+# Appendix 2: model details {#sec-appendix-model .appendix}
 
@@
 
+{{< pagebreak >}}
+
 # Appendix 3: model comparisons {.appendix}
@@
 
-# Appendix 4: Convergence diagnostics {#sec-appendix-diagnostics .appendix}
-
-## Gelman-Rubin diagnostic
+{{< pagebreak >}}
+
+# Appendix 4: convergence diagnostics {.appendix}
+
+We used Stan [@carpenter2017stan] as the probabilistic language behind the estimation of our Bayesian models in this study, with `brms` as its R interface [@burkner2017brms].
+This language implements the Markov Chain Monte-Carlo (MCMC) algorithm to explore the posterior distribution of the model.
+Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined.
+We run `r dim(model_fit_4$fit)[2]` MCMC chains, each `r dim(model_fit_4$fit)[1]` iterations long each.
+The correct performance of this algorithm is critical to the quality of the statistical evidence to which the outcomes of the model lead.
+
+One way to diagnose the behaviour of MCMC is to inspect whether the different MCMC chains (if more than one) have converged to a similar region of the posterior.
+The Gelman-Rubin diagnostic [$\hat{R}$ or R-hat @gelman1992inference] provides a measure of chain convergence by comparing the variance within each chain *versus* the variance between each chain.
+Both are expected to be identical when chains have perfectly converged, so that $\hat{R} = 1$.
+Values lower than 1.01 are recommended, while values higher than 1.05 indicate that chains might have trouble converging and therefore the estimated parameters must be taken with caution.
+@fig-model-rhats-neffs (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference.
+Most values are lower than 1.01, and never higher than 1.05, which provides evidence of successful MCMC convergence.
+
+Another diagnostic of good MCMC converge is the ratio of effective sample size to total sample size ($N_{eff}/N$), which indicates the proportion of samples in the chain that resulted from a non-divergent transition.
+Values closer to 1 are ideal, as they indicate that all posterior samples from the MCMC were used to estimate the posterior distribution of the parameter.
+Values larger than 0.1 are recommended.
+@fig-model-rhats-neffs (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models.
+Most values are larger than 0.1, although model 0 ($\mathcal{M}_0$) accumulates most effective sample sizes close to 0.1.
 
 ```{r}
-#| label: fig-model-rhats
+#| label: fig-model-rhats-neffs
 #| fig-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
@@
 #| warning: false
-#| fig-height: 3
-#| fig-width: 8
+#| fig-height: 6
+#| fig-width: 9
 map_df(model_rhats, bind_rows, .id = "model") |> 
@@
     facet_wrap(~model, nrow = 1) +
-    annotate(geom = "rect",
-             fill = "grey80",
-             xmin = 1,
-             xmax = 1.01,
-             ymin = -Inf,
-             ymax = Inf) +
-    geom_histogram(bins = 10, colour = "white") +
+    geom_vline(xintercept = 1.01, colour = "grey", linewidth = 0.75) +
+    geom_histogram(bins = 12, colour = "white") +
     geom_rug(alpha = 0.5) +
     labs(x = "Gelman Rubin diagnostic (R-hat)",
-         y = "Number of MCMC samples") +
-    theme(legend.position = "none")
-```
-
-## Effective sample size
-
-```{r}
-#| label: fig-model-ess
-#| fig-cap: "Ratio of effective sample size in the MCMC chains in the Bayesian models."
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-height: 3
-#| fig-width: 8
+         y = "MCMC samples",
+         title = "Gelman-Rubin (R-hat) diagnostic") +
+
     map_df(model_neffs, bind_rows, .id = "model") |> 
@@
     labs(x = "Gelman Rubin diagnostic (R-hat)",
-         y = "Number of MCMC samples") +
+         y = "MCMC samples",
+          title = "Ratio of effective sample size") +
     scale_x_continuous(labels = percent) +
-    theme(legend.position = "none")
+
+    plot_layout(ncol = 1) &
+    plot_annotation(tag_levels = "A") &
+    theme(legend.position = "none",
+          panel.grid = element_blank())
 ```
 
-# Appendix 5: Frequency and language exposure as separate predictors {#sec-appendix-doe  .appendix}
+# Appendix 5: frequency and language exposure as separate predictors {.appendix}
 

### 2023-02-26 manuscript/manuscript.qmd
Add appendices

--- 2023-02-16 manuscript/manuscript.qmd
@@
 ---
-title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual word acquisition"
-editor: source
-author:
-  - name: Gonzalo Garcia-Castro
-    orcid: 0000-0002-8553-4209
-    affiliations:
-      - Center for Brain and Cognition, Universitat Pompeu Fabra
-    corresponding: true
-  - name: Daniela S. Ávila-Varela
-    affiliations:
-      - Center for Brain and Cognition, Universitat Pompeu Fabra
-  - name: Ignacio Castillejo
-    affiliations:
-      - Departamento de Psicología, Universidad Autónoma de Madrid
-  - name: Núria Sebastian-Galles
-    affiliations:
-      - Center for Brain and Cognition, Universitat Pompeu Fabra
-
-affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
+title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual lexical acquisition"
 
@@
 abstract: |
-    Recent studies have reported larger vocabulary sizes for bilingual toddlers learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-This effect was only observed in the language to which toddlers were exposed the least, and has been suggested to be the result of a cognateness facilitation effect during word acquisition.
-The specific underpinnings of tyhis effect, and its relationship with the degree of exposure to the language are unclear.
-In this study, we present vocabulary data from 436 Catalan-Spanish bilinguals and monolinguals aged 12 to 34.
-We used Bayesian Item Response Theory to estimate participants' probability of acquisition of 604 words, conditional to their word's cognate status and lexical frequency, and the child's age and degree of exposure to each language.
-We found converging evidence for an earlier age of acquisition for cognate words, mediated by word frequeny and language exposure.
-Low-frequency words, and words from the language of least exposure benefitted more strongly by their cognate status than high-frequency words.
-Our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition and processing of words in the other language.
+    A prominent feature of the bilingual mental lexicon is that the activation of representations spreads in a language non-selective fashion.
+An instance of such non-selectivity is embodied by cognateness (i.e., form-similarity between translation equivalents, such as *flower*-*flor* in English and Spanish), which impacts lexical access and processing in bilinguals.
+Recent studies have suggested that cognateness facilitates vocabulary acquisition in bilingual toddlers---who show larger vocabulary sizes when their languages share many cognates---and an earlier age-of-acquisition for cognates than for non-cognates.
+The specific mechanisms underpinning such facilitation are unclear.
+We present an account of bilingual early lexical acquisition in which cognateness interacts with lexical frequency and language exposure to facilitate the acquisition of low-frequency words.
+We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals aged 12 to 34 months.
+We used a Bayesian Exploratory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
+We found converging evidence for an earlier age-of-acquisition for cognate words, and for such effect being mediated by lexical frequency and language exposure.
+Low-frequency words, and words from the language of least exposure were more strongly benefitted by their cognate status than high-frequency words.
+Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
 
@@
 keywords: lexical acquisition, vocabulary, bilingualism, item response theory, bayesian
-toc: false
-citation: true
-fig-dpi: 1000
-
-csl: "../assets/apa7.csl"
-bibliography: "../assets/references.bib"
-
-format:
-  pdf:
-    pdf-engine: xelatex
-    whitespace: small
-    linestretch: 1.5
-
 ---
@@
 library(patchwork)
+library(bayesplot)
 library(ggtext)
@@
 tar_load(model_neffs)
+tar_load(model_mcmc_pairs)
+tar_load(model_ppcs)
 
@@
 theme_set(theme_custom()) # set custom ggplot theme
+bayesplot_theme_set(
+    theme_custom()  +
+        theme(panel.grid = element_blank(),
+              axis.text = element_text(size = 5),
+              strip.text = element_text(size = 5),
+              plot.title = element_text(size = 8),
+              plot.subtitle = element_text(size = 8))
+)
 clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
+color_scheme_set(rep(clrs[4], 6))
 options(ggplot2.ordinal.fill = clrs,
@@
 # Introduction {#sec-introduction}
-
-## Word acquisition during toddlerhood: the effect of language experience
 
@@
 
-Developmental psychology has increased the representativeness of its samples considerable in recent years, but most literature on early word acquisition relies extensively on data from monolingual children, and neglects to a considerable extent that a substantial proportion of the world population acquires more than one language since early ages [@grosjean2021extent].
+Most literature on early word acquisition still relies extensively on data from monolingual children, and is oblivious to substantial proportion of the world population acquiring more than one language since early ages [@grosjean2021extent].
 Previous research about bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
-@hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals do.
+@hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals did.
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
@@
 
-## Language distance as a mediator of bilingual vocabulary growth
-
 For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language.
 Depending on the similarity between both languages, the two sets of words may overlap in form in varying degrees.
-When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically distant languages, like Spanish and English (Romance and Germanic languages, respectively).
+When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like Spanish and English (Romance and Germanic languages, respectively).
 For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
@@
 The Levenshtein distance is a metric that computes the distance between two strings of characters by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
-The resulting scores are then divided by the length of the longest string to normalise the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
-Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
-For example, English-Dutch bilinguals (22.14% cognateness), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% cognateness) were able to produce in Mandarin.
-
-## Lexical non-selectiviy as a candidate account: the role of parallel activation
+For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/kæt/-/\textipa{kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /æ/ by /\textipa{A}/).
+The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{kæt}/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/.
+Levenshtein distance scores are then divided by the length of the longest string to normalise the similarity scores between 0 and 1.
+Following the previous examples: 1/3 (0.33) for /kæt/-/\textipa{kAt}/, and 3/4 (0.75) for /kæt/-/\textipa{Gata}/.
+Finally, this variable was included as a predictor of participants' vocabulary sizes.
+Among other findings, Floccia and co-workers reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
+For example, English-Dutch bilinguals (22.14% mean cognateness), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% mean cognateness) were able to produce in Mandarin.
 
@@
 pointed to *parallel activation* as the underpinning mechanism behind their results.
-The parallel activation hypothesis stems from the language non-selective account of lexical access, and suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain].
+The parallel activation hypothesis stems from the language non-selective account of lexical access, and suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain], even in fully monolingual environments.
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
 In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish.
-In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trials labels were non-cognates (*mesa*-*taula*, translations of *table*).
+In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of *table*), importantly such distinction was only relevant for bilinguals.
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
-Importantly, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, showed equivalent naming times for the two types of stimuli.
-These results suggest that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
+In contrast, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, showed equivalent naming times for the two types of stimuli.
+These results showed that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
 
 Regarding word acquisition, the parallel activation account is a plausible account for Floccia et al.'s results in the light of (1) the evidence pointing to bilingual toddlers' lexicon consisting of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage], and that (2) there is experimental evidence of parallel activation during word comprehension in children and toddlers [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
-It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
-This account is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
+It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning languages with a larger proportion of cognates (such as children learning English and Dutch) showing larger vocabulary sizes.
+This account in which cognateness underlies the facilitation effect of typological language distance is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
 If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity but not phonological similarity.
 
-Evidence supporting an earlier age of acquisition for cognates is nonetheless scarce to date.
+Evidence supporting an earlier age of acquisition for cognates is scarce.
 @bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
@@
 The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-On the other hand, Floccia et al.'s estimates are statiscally more reliable given their remarkably larger sample size.
-But although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level.* The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
+On the other hand, Floccia et al.'s estimates are statistically more reliable given their remarkably larger sample size (372 participants).
+Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level.* The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
 By aggregating the responses from all words into a single datum per child, information about the acquisition status of cognates vs.
@@
 The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty).
-These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar: word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
+These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar.
+Word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
 Parallel activation remains a plausible explanation for this effect.
@@
 
-In this study, we provide an account for the cognate facilitation on word acquisition rooted in accumulator models of language acquisition [@hidaka2013computational; @mollica2017how].
-We build on the assumption that word acquisition is a function of mainly, but not exclusively, the amount of experience a child has accrued with the word-referent pairing [see @kachergis2022standard for review].
+In the present investigation we approach the cognate facilitation on word learning using the conceptual framework of accumulator models of language acquisition [@hidaka2013computational; @mollica2017how].
+We build on the assumption that word acquisition is a function of mainly---but not exclusively---the amount of experience a child has accrued with the corresponding word-referent pairing [see @kachergis2022standard for review].
 As infants accumulate linguistic input, they also accumulate opportunities to associate familiar word-forms to their corresponding referents.
-Words with high lexical frequency like *flower* [`r floor(childes[childes$token=="flower",]$freq_million)` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as *headphones* (`r floor(childes[childes$token=="headphones",]$freq_million)` times per million words) [e.g., @goodman2008does]: infants might accumulate learning experiences with *flower* faster than with *headphones*.
-In the case of bilinguals, the amount of exposures to a word-form an infants receives is mediated by their degree of exposure to their languages.
-Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time might need longer to accumulate the same experience with *flower* and *headphones* than an English monolingual does (who listens 100% of the time to English).
+Words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'headphones' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does].
+Infants might accumulate learning experiences with 'flower' faster than with 'taxi'.
+In the case of bilinguals, the amount of exposures to a word-form an infants receives is also mediated by their degree of exposure to their languages.
+Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual would (who listens 100% of the time to English).
 
@@
 We hypothesise that both forms in the translation equivalent (e.g., *flower* and *flor*, for an English-Spanish bilingual) follow their acquisition trajectory in a similar way they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to.
-When one of the word-forms is acquired, the acquisition trajectory of the other is impacted, so that the number of needed learning instances left for acquisition are reduced.
+When one of the word-forms is sufficiently acquired, the acquisition trajectory of the other is impacted, so that the number of needed learning instances left for acquisition are reduced.
 This facilitation would occur as the result of the semantic and phonological overlap between both words.
-When the child has acquired *flower*, and therefore its form-referent association has been formed, future presentations of the similar sounding translation *flor* in the presence of the same reference should lead to an increased likelihood of acquisition.
+When the child has sufficient lexical knowledge of *flower*, and therefore there is some for-referent association in place, future presentations of the similar sounding translation *flor* in the presence of the same reference should lead to an increased likelihood of acquisition.
 In the case of a non-cognate translation equivalent, such facilitation would be unlikely.
@@
 In line with this account, only the later-acquired form of the translation equivalent would benefit from its cognate status.
-Given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
 For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age of acquisition for cognates than for non-cognates.
@@
 When the child hears *flower*, activation spreads across languages to activate *flor* too.
-As a result, after exposure to *flower* not only the association between *flower* and its referent is strenghtened, but also that of *flower* and its common referent.
+As a result, after exposure to *flower* not only the association between *flower* and its referent is strengthened, but also that of *flower* and its common referent.
 To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*, which do not share any phoneme) does not undergo the such facilitation.
@@
 Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
-Therefore, participants may have on average learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
+Therefore, participants may have on average learned the English word-form of translation equivalents earlier than the word-form in the additional language.
 Under this hypothesis, because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
 
-## The present study
-
-In this study, we tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a words *only* once its translation has been already acquired.
-Using an online vocabulary checklist designed specifically for this study, we collected vocabulary data on production and comprehension from a sample of bilingual children.
+We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a words *only* once its translation has been already acquired.
+We collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
 Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees.
-We adopted a Bayesian item response theory [*IRT*, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
-
-We operationalised participants' exposure rate to a word-form using a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it when listening to speech), and  the child's amount of exposure to the language such word belongs to.
-This measure was calculated by multiplying the word-form's lexical frequency with the participant's degree of exposure to the language the word belongs to.
+We adopted a Bayesian explanatory item response theory  [*IRT*, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
+
+We operationalised participants' exposure rate to a word-form using a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it), and child's amount of exposure to the language such word belongs to.
+This measure was calculated by multiplying the word-form's lexical frequency with participant's degree of exposure to the language the word belongs to.
 Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childes] (see Appendix 1 for more details).
-Participant's degree of exposure to Catalan and Spanish were reported by families in a language exposure questionnaire they filled before the vocabulary checklist, in which they provided a proportion of the time that the child was estimated to listen to any language they were spoken to regularly.
+Participants' degree of exposure were reported by caregivers in a language exposure questionnaire they filled before the vocabulary checklist.
+Caregivers provided a proportion of the time the child was estimated to listen to any language they were spoken to regularly (described in @sec-methods).
 The resulting measure is a exposure-weighted lexical frequency measure, which was included in analyses as a proxy of how often bilingual participants were exposed to each word-form.
@@
 We defined *cognateness* as a continuous measure of phonological similarity between translation equivalents.
-We quantified the phonological similarity between each pair of word-forms by computing the inverse of the normalised Levenshtein distance between their SAMPA phonological transcriptions [@levenshtein1966binary].
+We quantified the phonological similarity between each pair of word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonological transcriptions [@levenshtein1966binary; @wells1995computer].
 This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
 This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-For instance, this measure of phonological similarity returns 0% for *ocell* (`uceL`) and *pájaro* (`paxa4o`), Catalan and Spanish for 'bird', and returns 50% for *lluna* (`Lun5`, /ˈʎu.nə/) and *luna* (`luna`, /ˈlu.na/), Catalan and Spanish for 'moon' (see Appendix 1 for more details).
+For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]`) and *pájaro* (`[paxa4o]`), Catalan and Spanish for 'bird', and returns 50% for *lluna* (`[Lun5]`) and *luna* (`[luna]`), Catalan and Spanish for 'moon' (see Appendix 1 for more details).
 
@@
 Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+both lists included items from a diverse sample of `r n_categories` semantic/functional categories.
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
 Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for 'dog').
-In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We included items from a diverse sample of `r n_categories` semantic/functional categories.
+
+```{r}
+#| label: tbl-items
+#| tbl-cap: "Summary of the items included in the final analyses."
+#| eval: true
+#| echo: false
+#| warning: false
+items |> 
+    arrange(language, list) |> 
+    distinct(te, .keep_all = TRUE) |> 
+    mutate(a = grepl("A", list),
+           b = grepl("B", list),
+           c = grepl("C", list),
+           d = grepl("D", list)) |> 
+    group_by(semantic_category) |> 
+    summarise(across(a:d, sum), 
+              items = list(meaning),
+              .groups = "drop") |> 
+    mutate(total = rowSums(across(a:d)),
+           items = map(items, function(x) head(x, 2))) |>  
+    arrange(desc(total)) |> 
+    relocate(a:d, total, items) |> 
+    gt(rowname_col = "semantic_category") |> 
+    fmt_percent(matches("_prop"), decimals = 1) |> 
+    cols_label(a = "List A",
+               b = "List B",
+               c = "List C",
+               d = "List D",
+               total = "Total",
+               items = "Examples") |> 
+    tab_style(cell_text(size = "small"),
+              list(cells_body(),
+                   cells_stub(),
+                   cells_column_labels(),
+                   cells_column_spanners(),
+                   cells_row_groups())) |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels()) |> 
+    tab_options(data_row.padding = px(2),
+                row_group.padding = px(1.2)) |> 
+    tab_style(cell_text(size = "x-small", align = "left"),
+              cells_body(items))
+
+```
 
@@
 Semantic/functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
+Items that were part of the trial lists of some ongoing experiments in the lab were also included in all versions.
 The resulting reduced vocabulary checklists contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words.
-Therefore, each response to the questionnaire provided data for a minimum of `r min(n_items_version$catalan, n_items_version$spanish)` and a maximum of `r max(n_items_version$catalan, n_items_version$spanish)` translation equivalents.
-Participants were randomly allocated into one of four subset, and always the same subset every time they participated.
+Participants were randomly allocated into one of four subset.
+Each response to the questionnaire provided data for a minimum of `r min(n_items_version$catalan, n_items_version$spanish)` and a maximum of `r max(n_items_version$catalan, n_items_version$spanish)` translation equivalents.
 
@@
     as_date()
+
+participants_edu <- table(participants$edu_parent)
 ```
@@
 We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between the `r format(dates[1], "%dth of %B, %Y")` and the `r format(dates[2], "%dth of %B, %Y")`: `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
-Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`.
+Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`, and were always allocated to the same questionnaire list (A, B, C, or D).
 Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample).
-All families gave informed consent before participating.
-Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser.
+All families provided informed consent before participating.
+Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device.
 Filling the questionnaire took 30 minutes approximately.
-After completion, families were rewarded with a small present.
+After completion, families were rewarded with a token of appreciation.
 
@@
 #| tbl-cap: "Participant sample size by age and degree of exposure to Catalan.
-Information about DoE was provided by families before filling the questionnaire.
-A 100% indicated that the participant was exclusively exposed to Catalan, and never to Spanish.
-A 0% indicates that the participant was not exposed to Catalan ever, and rather most of the time to Spanish.
+Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
+In this table, a 100% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
+A 0% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
 A 50% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
@@
 
-We used the highest self-reported educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker.
-The following options were given: *sense escolaritzar/sin escolarizar* [no education], *educació primària/educación primaria* [primary school], *ducació secundària/educación secundaria* [secondary school], *batxillerat/bachillerato* [complementary studies/high school], *cicles formatius/ciclos formativos* [vocational training], and *educació universitària/educación universitaria* [university degree], in line with the current educational system in Spain.
-Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r percent(table(participants$edu_parent)["University"]/length(unique(participants$id)))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))`) and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
+We used the highest self-reported educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES)This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: sense escolaritzar/sin escolarizar [no education], educació primària/educación primaria [primary school], educació secundària/educación secundaria [secondary school], batxillerat/bachillerato [complementary studies/high school], cicles formatius/ciclos formativos [vocational training], and educació universitària/educación universitaria [university degree].
+Most families reported university studies (`r participants_edu["University"]`, `r percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
 
@@
 
-n_items <- distinct(responses, language, item) |> 
+# attrition
+n_items <- bvq_data$pool
+
+n_items_nouns <- n_items |> 
+    filter(class=="Noun")
+
+n_items_freq <- n_items_nouns |> 
+    left_join(childes,
+              by = c("childes_lemma" = "token")) |> 
+    drop_na(freq_zipf, wordbank_lemma)
+
+n_items_lemma <- n_items_freq |> 
+    filter(n_lemmas == 1)
+
+n_items_multi <- n_items_lemma |> 
+    filter(!is_multiword) |> 
+    filter(include, !is.na(wordbank_lemma)) |> 
+    drop_na(sampa) |> 
+    distinct(te, language)
+
+n_items_duplicated <- n_items_multi |> 
+    add_count(te, name = "n_te") |>
+    filter(n_te == 2)
+
+n_items_final <- n_items_duplicated |> 
     count(language) |> 
@@
                   .sep = ".")
-```
-
-For the analyses in the present study, we considered responses to nouns.
-[@fourtassi2020growth].
-We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono/mico* [monkey]).
-This resulting in `r format(sum(n_items), big.mark = ",")` distinct words: `r n_items[["Catalan"]]` in Catalan, `r n_items[["Spanish"]]` in Spanish, making up `r n_te` translation equivalents in total.
-The final dataset consisted on  `r format(n_obs, big.mark = ",")` observations, each observation corresponding to a single response of one participant to one word.
-Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`), both languages summed together.
+
+```
+
+From the responses to the `r format(nrow(n_items), big.mark = ",")` items included in the questionnaire, we restricted the analyses to responses to nouns (`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded) [@fourtassi2020growth].
+We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
+Finally, we removed items without a translation in the other language (*n* = `r nrow(n_items_multi)-nrow(n_items_duplicated)`).
+This resulted in a final list of `r sum(n_items_final)` items, corresponding to words `r n_items_final["Catalan"]` Catalan words and their `r n_items_final["Spanish"]` Spanish translations (`r n_te` translation equivalents).
+After collecting participants responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item.
+Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
 
 We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_4$fit)[2]` sampling chains with `r format(dim(model_fit_4$fit)[1], big.mark = ",")` iterations each).
-This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$ (number of months elapsed between participants' birth date and questionnaire completion), $Length$ (number of phonemes in the SAMPA
+This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$ (number of months elapsed between participants' birth date and questionnaire completion), $Length$ (number of phonemes in the X-SAMPA
 phonological transcription of the word-form), $Exposure$ (composite measure of lexical frequency and language degree of exposure), and $Levenshtein$ (phonological similarity between translation equivalents).
@@
 #| echo: false
+rhats_df <- map_df(model_rhats, bind_rows, .id = "model") |> 
+    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
+    pivot_longer(-model, 
+                 names_to = ".variable",
+                 values_to = ".rhat") |>
+    filter(str_detect(.variable, "b_|sd_")) |> 
+    drop_na()
+
 loo_diff <- loo_compare(model_loos)[2, 1]
@@
 The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $Exposure$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the second best-performing model ($\mathcal{M}_{3}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`.
-Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see Appendix 3 for all model comparisons).
+Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see @tbl-results-loos for all model comparisons).
+
+```{r}
+#| label: tbl-results-loos
+#| echo: false
+#| message: false
+#| warning: false
+#| eval: true
+#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
+The predictive accuracy of the models was compared using leave-one-out cross-validation.
+The likelihood of each data point given a model fitted without including that same observation was estimated.
+The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
+*ELPD* values closer to zero indicate better predictive acuracy.
+We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
+The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
+model_loos |>
+    loo_compare() |>
+    as.data.frame() |>
+    rownames_to_column("model") |>
+    mutate(model = str_remove(model, "model_fit_")) |>
+    relocate(model,
+             matches("elpd_loo"),
+             matches("diff"),
+             matches("p_loo"),
+             matches("looic")) |>
+    gt() |>
+    fmt_number(2:9, decimals = 1) |>
+    cols_merge_uncert(elpd_loo, se_elpd_loo) |> 
+    cols_merge_uncert(p_loo, se_p_loo) |> 
+    cols_merge_uncert(looic, se_looic) |> 
+    cols_merge_uncert(elpd_diff, se_diff) |> 
+    cols_label(
+        model = "Model",
+        elpd_loo = "ELPD ± SE",
+        p_loo = "p ± SE",
+        looic = "IC ± SE",
+        elpd_diff = "ELPD diff. ± SE"
+    ) |>
+    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
+    tab_style(cell_text(weight = "bold"),
+              cells_column_labels()) |>
+    tab_style(cell_text(align = "left"),
+              cells_title(groups = "subtitle"))
+```
+
 @tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$, and their degree of overlap with the ROPE.
@@
 #| fig-height: 5
+#| eval: true
 #| fig-cap: "Expected posterior predictions.
@@
 The X-axis indicates the age (in months) for which the prediction is generated.
-The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*).
+The Y-axis indicates the predicted probability of acquisition.
 Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
-Predictions are presented separately for different degrees of word exposure index: little exposure to the word (-1 SD), median exposure, and high exposure (+1 SD).
+Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
 Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom.
@@
     mutate(
-        age = rescale_variable(age_std, mean(responses$age), sd(responses$age)),
+        age = rescale_variable(age_std, 
+                               mean(responses$age), 
+                               sd(responses$age)),
         exposure = factor(
@@
             levels = unique(exposure_std),
-            labels = paste0("Exposure: ", c("-1 SD", "mean", "+1 SD")),
+            labels = c("Low exposure (-1 SD)",
+                       "Mean exposure",
+                       "High exposure (+1 SD)"),
             ordered = TRUE
@@
             ordered = TRUE
-        )
+        ),
+        .category = ifelse(.category=="Understands",
+                           "Comprehension",
+                           "Production")
     ) |> 
@@
              colour = NA) +
+    annotate(geom = "text",
+             x = min(responses$age),
+             y = 1,
+             hjust = -0.075,
+             vjust = 1,
+             label = "Observed\nage range",
+             colour = "grey40",
+             size = 2) +
     geom_vline(xintercept = mean(responses$age), colour = "grey",
@@
         subtitle = glue("**Phonological similarity**<br> <span style = 'color:{clrs_reorder[5]};'>0% similarity</span>   <span style = 'color:{clrs_reorder[4]};'>50% similarity</span>   <span style = 'color:{clrs_reorder[1]};'>100% similarity</span>")) +
-    scale_colour_manual(values = clrs[c(5, 4, 1)]) +
-    scale_fill_manual(values = clrs[c(5, 4, 1)]) +
+    scale_colour_manual(values = clrs_reorder[c(5, 4, 1)]) +
+    scale_fill_manual(values = clrs_reorder[c(5, 4, 1)]) +
     scale_y_continuous(
@@
 
-## Summary of the study
-
-In this study, we built on previous literature to investigate the role of cognateness on word acquisition.
-We contrasted two hypotheses regarding the stage of the word acquisition process at which the possible (facilitative) effect of cognateness might take place.
+In this study, we investigated the role of cognateness on word acquisition.
+We contrasted two hypotheses regarding the word acquisition process at which the possible (facilitative) effect of cognateness might take place.
 First, we considered the possibility that cognateness boosts acquisition at early stages, before a stable, robust representation of either of the word-forms involved in the translation equivalent has been formed in the child's lexicon.
@@
 We also added the $Length$  of the word-form as a nuance predictor the increase the predictive performance of the model.
-
-## Our findings
 
@@
 
-## How our findings relate to previous literature
-
 In line with the hypothesis that received more support from our data, we predicted an interaction effect between participants' age, cognateness, and word exposure, since we expected cognateness to play a more important role at later stages of word learning, when at least one of the members of the translation equivalent has been acquired, and therefore its counterpart in the other language can benefit from its cognate status.
@@
 It is possible that, even if the cognateness effect plays a role if the acquisition of the word-form with the least exposure of the translation equivalent, age itself might not modulate the size of this effect: cognateness plays a similar role across ages.
-
-## Limitations
 
@@
 
-## Future steps
-
 Future studies addressing this subject may benefit for a more adequate design oriented towards making causal inferences.
@@
 ## Conclusions
+
 In summary, our study has replicated previous studies and provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
@@
 
-# Appendix 1: Model predictors {#sec-appendix-predictors .appendix}
-
-### Word exposure index
-
-Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database [@macwhinney2000childes; @sanchez2019childes].
-Using the corresponding lexical frequencies from the Catalan and Spanish corpora was not possible due to the low number of Catalan participants and tokens available in those languages.
-Available words in the English corpora were mapped to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human].
-Responses to words with missing lexical frequencies were excluded from analyses of the total number of items.
-Participant degree of exposure to the words' language were calculated as the percentage of exposure to the language, as reported by caretakers in the language profile section of the questionnaire.
-For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.
-
-We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
-the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).
-
-$$
-\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
-$$ {#eq-exposure}
-
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`) would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.
-
-### Levenshtein similarity
-
-Generally, a pair of words from two different languages that share meaning---translation equivalents---are considered as *cognates* if they share etymological origin (e.g., *table* and *taula*, in English and Catalan).
-The fact that two words share etymological origin frequently leads to them also being form-similar, as reflected by their overlapping orthographic or phonological form.
-This makes many cognates perceptually similar.
-In psycholinguistics, the term *cognateness* is often used to refer to the form-similarity that a pair of translation equivalents share, which has been found to impact how bilinguals process such word-forms (e.g., @costa2000cognate; @spivey1999cross).
-Whether two form-similar translation equivalents are etymologically related or not is arguably tangential to the question of how cognateness affect language processing.
-For instance, the translation pair *much* and *mucho*, in English and Spanish have the same meaning, and are considerably similar at the orthographical and phonological level.
-Nonetheless, they come from different Proto-Indo-European roots [@campbell2007glossary].
-This does not keep bilinguals from processing such kind of word-forms word forms differently than other translation equivalents with no form-similarity like *dog* and *perro*, in English and Spanish.
-In the scope of this experiment, we consider any form-similar translation equivalent as a *cognate* pair, regardless of whether both word-forms share etymological origin.
-
-Our study capitalises in phonology, therefore we operationalised *cognateness* as the phonological similarity between a pair of translation equivalents.
-Our dataset contained `r max(n_items)` translation equivalents, each consisting in two word-forms: Catalan and Spanish (see @tbl-translation-levenshtein for an example).
-Measuring the phonological similarity between two word forms is non trial.
-While orthographic similarity can be calculated using string similarity/distance measures on the written word-forms, phonology poses additional challenges.
-A first, bottom-up approach to to the task might be considering acoustic similarity as a proxy to phonological similarity.
-This may involve registering audio recordings of a talker reading each word-form aloud, and then comparing their spectrograms,  cochleograms, or formant tracks [e.g., @heeringa2003norwegian].
-While this is possible, this method computes acoustic similarity, as opposed to phonological similarity.
-This is not optimal, since a purely acoustic measure of similarity ignores how the actual signal is perceived by a actual listener.
-Since approximately 12 months of age, humans perceive the acoustic sounds in speech as phonemes, according to the phonology of their native language(s) [@werker1984crosslanguage; @kuhl1992linguistic].
-This means that two listeners may not perceive the same linguistically relevant acoustic signal identically.
-This is why phonological similarity needs to be computed from discrete units that are meaningful for language perception.
-
-A way to achieve this is to use phonological transcriptions of the word forms.
-Phonological transcriptions are symbolic, visual representations of the sounds in speech that capture some minimal  set of its features following some standard.
-The International Phonetic Alphabet (*IPA*) is one instance of such standards, providing a set of symbols that correspond to specific phonemes, as defined by their position across a series of dimensions.
-Using IPA transcriptions of word forms, one can use the same similarity/distance metric as in the case of orthographic word-forms.
-This method has been successfully used to measure the pairwise similarity between phonological word-forms from the  same language [e.g., @fourtassi2020growth] and across languages [@heeringa2004measuring; @floccia2018ii], and also between orthographic word-forms [@schepens2012distributions].
-
-Multiple algorithms have been created to compute the distance between two strings of characters.
-The Levenshtein distance being one of them.
-The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary, see @eq-levenshtein].
-For instance, /kaza/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /kasa/ (/z/ is replaced by /s/).
-Therefore, the Levenshtein distance between /kaza/ and /kasa/ is one.
-for an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), used phonological transcriptions in Speech Assessment Methods Phonetic Alphabet (SAMPA) format, as opposed to IPA, for its computer-friendly set of symbols.
-
-$$
-\text{lev}(x, y) =  
-\begin{cases}
-\max(i, j) & \text{if} \ \min(i, j)=0\\
-\min \begin{cases}
-\ \text{lev}_{x, y}(i-1, j) + 1 \\
-\ \text{lev}_{x, y}(i, j-1) + 1 \\
-\ \text{lev}_{x, y}(i-1, j-1) + 1_{a_i \neq b_j} \\
-\end{cases}
-\ &\text{otherwise}
-\end{cases}
-$${#eq-levenshtein}
-
-Here, $a$ and $b$ are the character strings corresponding to the phonological transcriptions of two word-forms belonging to the same translation equivalent, where each character corresponds to one phoneme expressed as a symbol from the SAMPA alphabet.
-$i$ and $j$ are the length (i.e., number of phonemes) of the $s$ and $t$ strings respectively.
-
-To measure the phonological similarity between the translation equivalents in our study, we computed a normalised versions of the Levenshtein distance between each pair of word-forms that accounts for the lenght of the strings.
-The rationale behind this correction is that longer word-forms are more likely to differ from their counterpart, compared to shorter word forms, and that this tendency does not correspond necessarily to a larger distance in terms of perception.
-This normalisation is achieved by dividing the Levenshtein distance by the length of the longest string, which leads to a distance metric that ranges from 0 to 1.
-This can be interpreted as a proportion of characters in the longest string that need to be edited in order for that string to become identical to the shorter string.
-A 0% normalised Levenshtein distance indicates that both word-forms are identical.
-A 50% normalised Levenshtein distance indicates that half of the phonemes in the longest word-form must be edited for both word-forms to become identical.
-A 100% normalised Levenshtein distance indicates that the two word-forms are completely different, as *all* phonemes in the longest word-form must be changed for it to become identical to the other.
-
-Finally, since we are interested in *cognateness*, in order to easy the interpretation of the analyses we subtracted the normalised Levenshtein distance from one, so that the metric indicated the *similarity* between the each pair of word-forms, instead of their distance.
-@eq-levenshtein-similarity shows the formula of this metric.
-
-$$
-1-\frac{\text{lev}(a, b)}{\max{\{i, j})\}}
-$${#eq-levenshtein-similarity}
-
-The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure (see @tbl-translation-levenshtein for more examples):
-
-| Catalan 	      | Spanish   	      | Levenshtein  |       
-|---------	      |---------	      |-----         |
-| *porta* (pOrt5) | *puerta* (pwe4ta) |   0.50 (3)   |
-| *taula* (tawl5) | *mesa* (mesa)     |   0.00 (5)   |
-| *cotxe* (kOtS5) | *coche* (kotSe)   |   0.40 (3)   |
-
-: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study  {#tbl-translation-levenshtein}
-
-# Appendix 2: Model details {#sec-appendix-model .appendix}
-
-We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function.
-The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood.
-
-$$
-\begin{aligned}
-\textbf{Likelihood:} \\
-y_{ij} &\sim \text{Cumulative}(p_{k})
-\end{aligned}
-$$ {#eq-likelihood}
-
-where:
-
-- $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$)
-- $i$ is the participant index
-- $j$ is the translation equivalent (TE) index
-- $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$)) between two response categories in the latent distribution
-$p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.
-
-To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), word-form length ($Length$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
-We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
-For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].
-
-$$
-\begin{aligned}
-\textbf{Linear model:} \\
-logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
-& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
-& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Length}_{ij} + & \\
-& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
-& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
-& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
-& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
-& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
-& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
-\end{aligned}
-$$ {#eq-linear}
-
-where:
-
-- $i$ and $j$ index the participant and translation equivalent (TE)
-- $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$
-- $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively
-- $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest
-- $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively
-
-We used the Bayesian framework to estimate the parameters in our model.
-This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical].
-This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations.
-We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].
-
-$$
-\begin{aligned}
-\\
-\textbf{Prior:} \\
-\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
-\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
-\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
-u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
-w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
-&&\mbox{[Participant-level coefficient variability]} \\ \\
-\Bigg(\begin{smallmatrix}
-u_{k_{0}} \\ 
-u_{1_{i}} \\ 
-\vdots \\ 
-u_{8_{i}} 
-\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
-\Bigg(\Bigg(\begin{smallmatrix}0 \\
-0 \\ 
-\vdots \\
-0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
-\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
-\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
-\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
-\vdots & \vdots & \vdots & \vdots \\
-\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
-\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
-\rho_{u} &\sim LKJcorr(2) \\
-\\
-&&\mbox{[TE-level coefficient variability]} \\ \\
-\Bigg(\begin{smallmatrix}
-w_{k_{0}}\\ 
-w_{1_{j}} \\ 
-\vdots \\ 
-w_{3_{j}} 
-\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
-0\\ 
-0 \\ 
-\vdots \\
-0 
-\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
-\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
-\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
-\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
-\vdots & \vdots & \vdots & \vdots \\
-\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
-\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
-\rho_{w_{0-3}} &\sim LKJcorr(2)
-\end{aligned}
-$$ {#eq-prior}
-
-where:
-
-- $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively
-- $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively
-- $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].
-
-```{r}
-#| label: fig-prior-mean
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 5
-#| fig-cap: "Expected prior predictions."
-nd <- expand.grid(
-    n_phon_std = 0,
-    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
-    exposure_std = 0,
-    lv_std = 0
-)
-
-m <- add_epred_draws(nd, 
-                     model_fit_4_prior,
-                     ndraws = NULL, 
-                     re_formula = NA) |> 
-    filter(.category != "No") |> 
-    pivot_wider(names_from = ".category", values_from = ".epred") |> 
-    mutate(Understands = Understands + `Understands and Says`) |> 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) |> 
-    mutate(
-        .category = factor(
-            .category,
-            levels = c("Understands and Says", "Understands"),
-            labels = c("Production\n(Understands and Says)", "Comprehension\n(Understands)")
-        )
-    )
-
-m |> 
-    ggplot(aes(x = age_std, y = .epred)) +
-    facet_wrap(~.category) +
-    stat_lineribbon(linewidth = 1, 
-                    colour = NA,
-                    .width = c(0.95, 0.89, 0.67, 0.5)) +
-    stat_summary(fun = "mean", geom = "line", linewidth = 0.75) +
-    scale_x_continuous(
-        breaks = scale(seq(10, 36, 4), 
-                       mean(responses$age), 
-                       sd(responses$age))[, 1],
-        labels = seq(10, 36, 4)
-    ) +
-    scale_fill_manual(
-        values = rev(clrs), 
-        labels = c("95%", "89%", "67%", "50%")
-    ) +
-    labs(
-        x = "Age (months)", 
-        y = "Probability of acquisition",
-        colour = "Credible interval", 
-        fill = "Credible interval"
-    ) +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
-    theme(
-        legend.position = "top",
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid = element_blank()
-    )
-```
-
-# Appendix 3: model comparisons {.appendix}
-
-```{r}
-#| label: tbl-results-loos
-#| echo: false
-#| message: false
-#| warning: false
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
-The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
-The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
-model_loos |>
-    loo_compare() |>
-    as.data.frame() |>
-    rownames_to_column("model") |>
-    mutate(model = str_remove(model, "model_fit_")) |>
-    relocate(model,
-             matches("elpd_loo"),
-             matches("p_loo"),
-             matches("looic"),
-             matches("diff")) |>
-    gt() |>
-    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) |>
-    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) |>
-    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) |>
-    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) |>
-    fmt_number(2:9) |>
-    cols_label(
-        model = "Model",
-        # elpd
-        elpd_loo = md("*ELPD*"),
-        se_elpd_loo = md("*SE*"),
-        # p
-        p_loo = md("*p*"),
-        se_p_loo = md("*SE*"),
-        # looic
-        looic = md("*LOO-IC*"),
-        se_looic = md("*SE*"),
-        # diff
-        elpd_diff = md("*diff*"),
-        se_diff = md("*SE*")
-    ) |>
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) |>
-    tab_style(cell_text(align = "left"),
-              cells_title(groups = "subtitle"))
-```
-
-# Appendix 4: Convergence diagnostics {#sec-appendix-diagnostics .appendix}
-
-## Gelman-Rubin diagnostic
-
-```{r}
-#| label: fig-model-rhats
-#| fig-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-height: 3
-#| fig-width: 8
-map_df(model_rhats, bind_rows, .id = "model") |> 
-    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
-    pivot_longer(-model, 
-                 names_to = ".variable",
-                 values_to = ".rhat") |>
-    filter(str_detect(.variable, "b_|sd_")) |> 
-    drop_na() |> 
-    ggplot(aes(.rhat, colour = model, fill = model)) +
-    facet_wrap(~model, nrow = 1) +
-    annotate(geom = "rect",
-             fill = "grey80",
-             xmin = 1,
-             xmax = 1.01,
-             ymin = -Inf,
-             ymax = Inf) +
-    geom_histogram(bins = 10, colour = "white") +
-    geom_rug(alpha = 0.5) +
-    labs(x = "Gelman Rubin diagnostic (R-hat)",
-         y = "Number of MCMC samples") +
-    theme(legend.position = "none")
-```
-
-## Effective sample size
-
-```{r}
-#| label: fig-model-ess
-#| fig-cap: "Ratio of effective sample size in the MCMC chains in the Bayesian models."
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-height: 3
-#| fig-width: 8
-map_df(model_neffs, bind_rows, .id = "model") |> 
-    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
-    pivot_longer(-model, 
-                 names_to = ".variable",
-                 values_to = ".rhat") |>
-    filter(str_detect(.variable, "b_|sd_")) |> 
-    drop_na() |> 
-    ggplot(aes(.rhat, colour = model, fill = model)) +
-    facet_wrap(~model, nrow = 1) +
-    geom_vline(xintercept = 0.1, colour = "grey", linewidth = 0.75) +
-    geom_histogram(bins = 10, colour = "white") +
-    geom_rug(alpha = 0.5) +
-    labs(x = "Gelman Rubin diagnostic (R-hat)",
-         y = "Number of MCMC samples") +
-    scale_x_continuous(labels = percent) +
-    theme(legend.position = "none")
-```
-
-# Appendix 5: Frequency and language exposure as separate predictors {#sec-appendix-doe  .appendix}
-
-```{r}
-#| label: fig-coefs-doe
-#| fig-cap: "Posterior distribution of regression coefficients in the probability scale.
-*Median*: median of the posterior distribution in the logit scale.
-*95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE).
-This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
-Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-height: 4
-#| fig-width: 10
-# summarise posterior draws
-bind_rows("Exposure model" = posterior_draws$summary,
-          "DoE-Frequency model" = posterior_draws_doe,
-          .id = "model") |> 
-    mutate(
-        across(c(.median, .lower, .upper),
-               function(x) {
-                   ifelse(
-                       .type=="Intercepts (at 22 months)",
-                       plogis(x),
-                       x/4
-                   )
-               }
-        )) |>
-    filter(.type == "Slopes") |> 
-    ggplot(aes(.median, .variable_name, xmin = .lower, xmax = .upper,
-               colour = model)) +
-    facet_wrap(~model, scales = "free") +
-    annotate(geom = "rect",
-             ymin = -Inf,
-             ymax = Inf,
-             xmin = -0.1/4,
-             xmax = +0.1/4,
-             fill = "grey70") +
-    geom_errorbar(linewidth = 1, width = 0.25) +
-    geom_point(size = 2.5) +
-    labs(x = "Median and 95% HDI of the posterior distribution",
-         y = "Predictor",
-         colour = "Model") +
-    scale_x_continuous(labels = percent) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(legend.position = "none",
-          axis.title.y = element_blank())
-```
+{{< pagebreak >}}
+
+# Appendix 1: model predictors {.appendix}
+
+{{< include _model-predictors.qmd >}}
+
+{{< pagebreak >}}
+
+# Appendix 2: model details {.appendix}
+
+{{< include _model-details.qmd >}}
+
+{{< pagebreak >}}
+
+# Appendix 3: convergence diagnostics {.appendix}
+
+{{< include _model-diagnostics.qmd >}}
+
+{{< pagebreak >}}
+
+# Appendix 4: frequency and language exposure as separate predictors {.appendix}
+
+{{< include _model-doe.qmd >}}
+
+{{< pagebreak >}}

### 2023-02-26 R/utils.R

@@ -295,5 +295,13 @@ save_files <- function(x,
 }
 
 remove_nul <- function() {
-    
+    paths <- c("manuscript", "docs")
+    cur_path <- gsub("/", "\\\\", getwd())
+    nul_path <- glue("{cur_path}\\{paths}\\NUL.")
+    file.exists(nul_path)
+    cmd1 <- glue("rename \\\\.\\{nul_path} delete.txt")
+    cmd2 <- glue("del \\\\.\\{nul_path}\\delete.txt")
+    lapply(cmd1, shell)
+    lapply(cmd2, shell)
+    shell(cmd2)
 }

### 2023-02-26 manuscript/_model-details.qmd
Add appendices

--- 2022-12-09 index.qmd
@@
----
-title: "Trajectories"
-subtitle: "Lab notes"
-date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
-authors:
-  - name: Gonzalo Garcia-Castro
-    orcid: 0000-0002-8553-4209
-    email: gonzalo.garciadecastro@upf.edu
-    corresponding: true
-    affiliations:
-      - name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-        address: Ramon Trias fargas 25-27
-        city: Barcelona
-        state: Spain
-        postal-code: 08005
-  - name: Daniela S. Ávila-Varela
-    orcid: 0000-0002-3518-8117
-    affiliations:
-      - name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-  - name: Ignacio Castillejo
-    orcid: 0000-0001-7445-0416
-    affiliations:
-      - name: Universidad Autónoma de Madrid
-        department: Departamento de Psicología
-  - name: Nuria Sebastian-Galles
-    orcid: 0000-0001-6938-2498
-    affiliations:
-      - name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
-format:
-    html:
-        toc: true
-        link-external-newwindow: true
-        number-sections: true
-        number-offset: 0
-        code-fold: true
-        code-overflow: scroll
-        code-line-numbers: true
-        code-copy: hover
-        fig-dpi: 1000
-        reference-location: margin
-        execute-dir: "C:/Users/U155880/Documents/trajectories/"
-    pdf:
-        toc: true
-        number-sections: true
-        colorlinks: true
-        fig-dpi: 1000
-bibliography: "references.bib"
-csl: "manuscript/apa6.csl"
----
-
-```{r}
-#| label: setup
-#| include: false
-tar_load_globals() # load packages and options
-tar_load_everything() # load targets as R objects
-
-# resolve namespace conflicts
-conflict_prefer("filter", "dplyr")
-conflict_prefer("select", "dplyr")
-
-# set ggplot theme and colour palette
-theme_set(theme_custom()) # set custom ggplot theme
-clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
-options(
-    knitr.kable.NA = "-",
-    knitr.duplicate.label = "allow",
-    ggplot2.ordinal.fill = clrs,
-    ggplot2.ordinal.colour = clrs,
-    ggplot2.discrete.fill = clrs,
-    ggplot2.discrete.colour = clrs,
-    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
-    ggplot2.continuous.colour = ggplot2::scale_color_gradient
-)
-```
-
-# Methods
-
-::: {.callout-tip}
-Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
-:::
-
-## Questionnaires
-
-```{r}
-#| label: questionnaire
-#| echo: true
-#| message: false
-#| warning: false
-data("pool", package = "bvqdev")
-n_categories <- distinct(pool, semantic_category) %>% nrow()
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% 
-    pull(n) %>%
-    set_names(c("catalan", "spanish"))
-```
-
-The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted of two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]` items.
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* \[dog\] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* \[*to finish*\] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* \[*monkey*\] and Catalan *mono/mico*.
-We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
-For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\~400 translation equivalents, \~400 items in Catalan, \~400 items in Spanish).
-These last families were randomly assigned to one of four different subsets of the complete list of items.
-These lists were designed so that each contained a representative subsample of the items from the complete list.
-Semantic/functional categories with less than 16 items were not divided in the short version of the questionnaire to preserve a minimum of four items per list for each category: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* \[cereal bar\]) and items that included more than one word-form (e.g., *mono / mico*).
-Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
-
-## Participants
-
-```{r}
-#| label: fig-participants-time
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 10 
-#| fig-cap: "Date at which responses were collected"
-participants %>% 
-    group_by(time_stamp) %>% 
-    summarise(n = n(), .groups = "drop") %>% 
-    mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n)) +
-    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
-             angle = 90, vjust = 1.5, hjust = 1) +
-    geom_line(size = 1, colour = clrs[1]) +
-    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
-    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 20, hjust = 1),
-        axis.title.x = element_blank(),
-    ) +
-
-    participants %>% 
-    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
-    group_by(time_stamp, age) %>% 
-    summarise(n = n(), .groups = "drop") %>% 
-    group_by(age) %>% 
-    mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n, colour = age)) +
-    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    geom_line(size = 1) +
-    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
-    scale_y_continuous(breaks = seq(0, 500, 50)) +
-    scale_colour_manual(values = clrs) +
-
-    plot_layout(ncol = 1) &
-
-    scale_x_date(date_breaks = "3 month", date_labels = "%B %Y") &
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 20, hjust = 1),
-        axis.title.x = element_blank(),
-    )
-```
-
-### Age
-
-Number of months elapsed between participants' birth date and questionnaire completion.
-
-```{r}
-#| label: fig-participants-age
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Distribution of participants' ages"
-participants %>% 
-    mutate(age = floor(age)) %>% 
-    count(age) %>% 
-    ggplot() +
-    aes(x = age, y = n) +
-    geom_col(fill = clrs[1]) +
-    geom_text(aes(label = n), size = 3.5, vjust = -1) +
-    labs(
-        x = "Age (months)",
-        y = "# participants"
-    ) +
-    scale_x_continuous(breaks = seq(0, 40, 2)) +
-    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
-    theme(
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
-
-### Language exposure (Degree of Exposure)
-
-Participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to.
-For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.
-
-```{r}
-#| label: fig-participants-lp
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
-participants %>% 
-    pivot_longer(
-        c(doe_catalan, doe_spanish),
-        names_to = "language", values_to = "doe", 
-        names_transform = ~str_to_sentence(str_remove(., "doe_"))
-    ) %>% 
-    mutate(
-        doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
-        id = paste0(id, " (", time, ")")
-    ) %>% 
-    count(language, doe) %>% 
-    ggplot(aes(doe, y = n, label = n, fill = language)) +
-    geom_col(colour = "white", size = 1) +
-    geom_text(position = position_nudge(y = 5)) +
-    facet_wrap(~language, ncol = 1) + 
-    labs(
-        x = "Degree of exposure to Catalan",
-        y = "Number of participants",
-        colour = "Language",
-        fill = "Language"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor = element_blank()
-    )
-```
-
-### SES/parental education
-
-Maximum educational attainment of the participant parent(s).
-
-```{r}
-#| label: tbl-participants-edu
-#| tbl-cap: Participant exposure to the language (*DoE*) and socio-economic status (*SES*)
-#| echo: true
-#| message: false
-#| warning: false
-participants %>%
-    left_join(bvq_data$logs) %>% 
-    mutate(
-        edu_parent = fct_explicit_na(as.factor(edu_parent)),
-        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE)
-    ) %>% 
-    count(doe_catalan, edu_parent) %>% 
-    right_join(
-        expand_grid(
-            doe_catalan = levels(.$doe_catalan),
-            edu_parent = unique(bvq_data$logs$edu_parent)
-        )
-    ) %>% 
-    rename(total = n) %>% 
-    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
-    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
-    rename(Total = total) %>% 
-    select(-"NA") %>% 
-    gt() %>% 
-    summary_rows(
-        columns = Complementary:Primary, 
-        fns = list(Total = ~sum(., na.rm = TRUE)),
-        formatter = fmt_integer
-    ) %>% 
-    cols_label(doe_catalan = "DoE Catalan") %>%
-    sub_missing(columns = 2:8, missing_text = "--") %>% 
-    tab_spanner("Educational attainment", Secondary:Primary) %>% 
-    tab_style(
-        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
-        locations = cells_body(columns = 9, rows = 1:5)
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(columns = 2:7)
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = c(1, 8))
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(spanners = "Educational attainment")
-    )
-```
-
-## Items
-
-```{r}
-#| label: tbl-items-summary
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| tbl-cap: "Lexical frequencies.
-Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
-items <- left_join(items, select(bvqdev::pool, te, item, language, class, semantic_category))
-
-items %>%
-    unnest(list) %>% 
-    group_by(language, list) %>% 
-    summarise_at(
-        vars(freq_zipf, n_phon, lv), 
-        list(
-            n = ~sum(!is.na(.)),
-            mean = ~mean(., na.rm = TRUE),
-            sd = ~sd(., na.rm = TRUE),
-            min = ~min(., na.rm = TRUE),
-            max = ~max(., na.rm = TRUE)
-        )
-    ) %>% 
-    select(-c(n_phon_n, lv_n)) %>% 
-    rename(n = freq_zipf_n) %>% 
-    gt(groupname_col = "language", rowname_col = "list") %>% 
-    fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>% 
-    fmt_percent(matches("lv_"), decimals = 1) %>% 
-    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
-    tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>% 
-    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
-    cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>% 
-    cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>% 
-    cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>% 
-    cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>% 
-    cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>% 
-    cols_merge_range(col_begin = lv_min, col_end = lv_max) %>% 
-    cols_label(
-        list = "",
-        n = md("*N*"),
-        freq_zipf_mean = md("Mean ± SD"),
-        freq_zipf_min = md("Range"),
-        n_phon_mean = md("Mean ± SD"),
-        n_phon_min = md("Range"),
-        lv_mean = md("Mean ± SD"),
-        lv_min = md("Range")
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(everything())
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(everything())
-    ) %>% 
-    summary_rows(
-        fns = list("Grand mean" = ~mean(.)),
-        columns = ends_with("_mean")
-    ) %>% 
-    summary_rows(
-        fns = list("N" = ~sum(.)),
-        columns = n,
-        formatter = fmt_integer
-    )
-```
-
-### Lexical frequency
-
-Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
-This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points.
-Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [{childesr}](https://langcog.github.io/childes-db-website/) R package.
-Available words in the corpora were then mapped to their Catalan and Spanish translations.
-Therefore, the Catalan and Spanish word-forms of the same translation equivalent had the same value for lexical frequency.
-Responses to words with missing lexical frequencies were excluded from analyses of the total number of items).
-
-```{r}
-#| label: fig-items-freq
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 4
-#| fig-cap: "Distribution of lexical frequency across items.
-Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations.
-Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency.
-The black dot and interval indicate the mean and standard deviation."
-
-mean_sd <- function(x, ...) {
-    m <- mean(x, ...)
-    s <- sd(x, ...)
-    data.frame(mean = m, ymin = m-s, ymax = m+s)
-}
-
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    ggplot(aes(x = freq_zipf)) +
-    geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
-    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        legend.title = element_blank()
-    )
-```
-
-### \# Phonemes
-
-Word length ($Phonemes$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.
-
-```{r}
-#| label: fig-items-nphon
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Distribution of number of phonemes across items"
-items %>%
-    count(language, n_phon) %>% 
-    ggplot() +
-    aes(
-        x = n_phon,
-        y = n,
-        fill = language,
-        colour = language,
-        shape = language
-    ) +
-    facet_wrap(~language, ncol = 1) +
-    geom_col(colour = "white") +
-    geom_text(aes(label = n), position = position_nudge(y = 5), colour = "black") +
-    labs(
-        x = "# Phonemes",
-        y = "Number of words",
-        fill = "Language",
-        shape = "Language",
-        colour = "Language"
-    ) +
-    scale_x_continuous(breaks = 1:15) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.ticks.x = element_line(),
-        legend.title = element_blank()
-    )
-```
-
-### Phonological similarity (Levenshtein)
-
-Phonological similarity between translation equivalents was computed as the normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary].
-This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1.
-This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-
-```{r}
-#| label: fig-items-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 9
-#| fig-cap: "Distribution of Levenshtein similarities across items"
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
-    count(lv) %>% 
-    mutate(prop = n/sum(n)) %>% 
-    ggplot(aes(x = lv, n)) +
-    geom_col(fill = clrs[4], colour = "white") +
-    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
-              position = position_nudge(y = 3), size = 3.5) +
-    labs(x = "Levenshtein distance", y = "Number of words") +
-    theme(
-        panel.grid.major.x = element_blank(),
-        legend.position = "top"
-    )
-```
-
-### Frequency by \# Phonemes
-
-```{r}
-#| label: fig-items-frequency-phonemes
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 4
-#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
-items %>% 
-    drop_na(freq_zipf) %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    count(n_phon, freq_zipf) %>% 
-    ggplot(aes(n_phon, freq_zipf)) +
-    geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
-               position = position_jitter(width = 0.1), seed = 888) +
-    geom_smooth(method = "lm", formula = "y ~ x" ,
-                colour = clrs[1], fill = clrs[1]) +
-    labs(
-        x = "# Phonemes", 
-        y = "Lexical frequency (Zipf score)", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    ) 
-```
-
-### Frequency by Levenshtein
-
-```{r}
-#| label: fig-items-frequency-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Association betwen lexical frequency and phonological similarity"
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    drop_na(freq_zipf) %>% 
-    count(freq_zipf, lv) %>% 
-    ggplot(aes(freq_zipf, lv, size = n)) +
-    geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
-    geom_smooth(method = "lm", formula = "y ~ x", colour = clrs[1], fill = clrs[1]) +
-    labs(
-        y = "Levenshtein similarity", 
-        x = "Lexical frequency (Zipf score)", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "none",
-        panel.grid.major.y = element_blank()
-    )
-```
-
-### \# Phonemes by Levenshtein
-
-```{r}
-#| label: fig-items-nphon-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Association between number of phonemes and phonological similarity"
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    drop_na(freq_zipf) %>% 
-    count(n_phon, lv) %>% 
-    ggplot(aes(n_phon, lv, size = n)) +
-    geom_point(shape = 1, stroke = 1, colour = clrs[4]) +
-    geom_smooth(method = "lm", formula = "y ~ x", size = 1,
-                colour = clrs[1], fill = clrs[1]) +
-    labs(
-        x = "# Phonemes", 
-        y = "Levenshtein similarity", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_y_continuous(labels = percent) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    ) 
-```
-
-### Language exposure-weighted lexical frequency
-
-We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
-the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).
-
-$$
-\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
-$$ {#eq-exposure}
-
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):
-
-$$
-\begin{aligned}
-\textbf{Exposure}_{\textbf{cavall, Catalan}} &= \textbf{Frequency}_{cavall} \times \textbf{DoE}_{Catalan} = 6.14 \times 0.8 = 4.91 \\
-\textbf{Exposure}_{\textbf{caballo, Spanish}} &= \textbf{Frequency}_{caballo} \times \textbf{DoE}_{Spanish} = 6.14 \times 0.2 = 1.23
-\end{aligned}
-$$ {#eq-exposure-example}
-
-```{r}
-#| label: fig-exposure-distribution
-#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
-#| warnings: false
-#| fig-height: 7
-#| fig-width: 13
-df %>% 
-    mutate(
-        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
-        exposure = cut(exposure, breaks = seq(0, 9, 1), include.lowest = TRUE)
-    ) %>% 
-    count(exposure, language) %>% 
-    ggplot(aes(exposure, n, fill = language)) +
-    facet_wrap(~language) +
-    geom_col(colour = "white") +
-    labs(
-        x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
-        y = "Number of observations",
-        fill ="Language"
-    ) +
-
-    df %>% 
-    mutate(
-        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
-        exposure = cut(exposure, breaks = seq(0, 9, 1), include.lowest = TRUE),
-        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
-    ) %>% 
-    count(exposure, freq, language) %>%    
-    ggplot(aes(x = freq, y = exposure, colour = language, size = n)) +
-    facet_wrap(~language) + 
-    geom_point() +
-    labs(
-        x = "Frequency (Zipf score)",
-        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
-        colour = "Language",
-        fill = "Language",
-        size = "Number of observations"
-    ) +
-    theme(panel.grid = element_blank()) +
-
-    plot_layout(ncol = 1) &
-    guides(fill = "none", colour = "none") &
-    theme(panel.grid.major.x = element_blank(),
-          panel.grid.minor.y = element_blank())
-```
-
-## DAG
-
-```{r tikz-settings, include=FALSE}
-# tikz stuff
-# Necessary for using dvisvgm on macOS
-# See https://www.andrewheiss.com/blog/2021/08/27/tikz-knitr-html-svg-fun/
-font_opts <- list(dvisvgm.opts = "--font-format=woff")
-```
-
-```{tikz dag, engine.opts=font_opts, echo=FALSE}
-#| fig-cap: "Directed acyclic graph indicating the hypothesised causal relationship between the variables of interest.
-*i* and *j* indicate the variables that are affected by individual differences at the level of participants and translation equivalents, respectively.
-Unobserved variables are encircled, and observed variables are not.
-We hypothesised that the number of phonemes of the word, the age of the participant, the amount of exposure to the word, and the Levenshtein distance all affect participants' familiarity with the word (i.e., the probability of acquisition), and therefore the probability with which each response category is generated.
-We further hypothesised that age influences the effect of word exposure and Levenshtein distance on familiarity, and that word exposure influences the effect of Levenshtein distance on familiarity."
-#| fig-align: center
-#| fig-ext: svg
-#| out-width: 100%
-\usetikzlibrary{positioning}
-\usetikzlibrary{shapes.geometric}
-\begin{tikzpicture}
-% nodes
-\node (age) at (0,0) {$\text{Age}_{i, j}$};
-\node (phonemes) at (3,4) {$\text{Phonemes}_{i, j}$};
-\node (exposure) at (3,2) {$\text{Exposure}_{i, j}$};
-\node (levenshtein) at (3,-2) {$\text{Levenshtein}_{i}$};
-\node [ellipse, draw](familiarity) at (7, 0) {$Familiarity$};
-\node (response) at (11, 0) {$\text{Response}$};
-% arrows
-\draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (familiarity);
-\draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (exposure);
-\draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (levenshtein);
-\draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (familiarity);
-\draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (levenshtein);
-\draw [->, shorten >= 4pt, shorten <= 4pt] (phonemes) -- (familiarity);
-\draw [->, shorten >= 4pt, shorten <= 4pt] (levenshtein) -- (familiarity);
-\draw [->, shorten >= 4pt, shorten <= 4pt] (familiarity) -- (response);
-
-\end{tikzpicture}
-```
-
-## Data analysis
-
 We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function.
-The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood (see @fig-likelihood for a graphical representation).
+The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood.
 
@@
 
-```{r}
-#| label: fig-likelihood
-#| fig-cap: "Graphical representation of the assumed latent distribution that generates observed responses in the questionnaire.
-Cumulative ordinal regression assumes that observed categorical responses (*No*, *Understands* and *Understands and Says* are generated from a common latent (i.e., unobserved) distribution.
-For threotical reasons, we refer to the dimension captured by such distribution as *Familiarity*.
-*Familiarity* describes a given participant's familiarity with a given word.
-When familiarity is low (below some threshold 1), the reulting response is *No*.
-When familiarity is higher than such threshold 1, but lower than some threshold 2, the resulting result is *Understands* (i.e., comprehension).
-When familiarity exceeds threshold 2, the resulting response is *Understands and Says* (i.e., comprehension and production).
-These thresholds are estimated in the cumulative regression model as intercepts 1, and 2.
-The rest of the coefficients in the regression model shift the response thresholds, making comprehension or production more likely or less likely.
-See @burkner2019ordinal for a detailed description of cumulative ordinal regression and its implementation."
-#| fig-width: 7
-#| fig-height: 3.5
-dist_df = tribble(~mean, ~sd, 0,   1)
-
-responses <- data.frame(
-    ymin = c(-Inf, -1.5, 1.5),
-    ymax = c(-1.5, 1.5, Inf),
-    xmin = -Inf,
-    xmax = Inf,
-    .category = str_wrap(c("No", "Understands", "Understands and Says"), width = 12)
-)
-
-responses %>%
-    ggplot(aes(ymin = ymin, ymax = ymax, xmin = xmin, xmax = xmax, fill = .category)) +
-    geom_rect(alpha = 0.5) +
-    stat_dist_slab(
-        data = dist_df, 
-        aes(dist = "norm", arg1 = mean, arg2 = sd),
-        position = "dodge", inherit.aes = FALSE,
-        fill = NA, colour = "black"
-    ) +
-    geom_hline(yintercept = c(-1.5, 1.5), linetype = "dotted", size = 1) +
-    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
-              vjust = 1, hjust = 0.5) +
-    annotate(
-        geom = "text",
-        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
-        y = c(-1.5, 2), parse = TRUE,
-        x = 0.5, angle = 90, vjust = -0.5
-    ) +
-    labs(y = "Hypothetical response-generating latent distribution",
-         x = "Response probability",
-         fill = "Response type",
-         colour = "Response type") +
-    coord_flip() +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_x_continuous(labels = percent) +
-    scale_y_continuous(limits = c(-4, 4)) +
-    theme(panel.grid = element_blank(),
-          legend.position = "none")
-```
-
-To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
-We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset--each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
+To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), word-form length ($Length$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
+We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
 For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].
@@
 
-::: {.panel-tabset}
-
-#### Prior-predicted mean
-
 ```{r}
 #| label: fig-prior-mean
-#| echo: true
+#| echo: false
 #| message: false
 #| warning: false
+#| eval: false
 #| fig-width: 7
-#| fig-cap: "Expected prior predictions"
+#| fig-height: 5
+#| fig-cap: "Expected prior predictions."
 nd <- expand.grid(
     n_phon_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
+    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
     exposure_std = 0,
@@
             levels = c("Understands and Says", "Understands"),
-            labels = c("Production\n(Understands and Says))", "Comprehension\n(Understands)")
+            labels = c("Production\n(Understands and Says)", "Comprehension\n(Understands)")
         )
@@
 
-m %>% 
+m |> 
     ggplot(aes(x = age_std, y = .epred)) +
     facet_wrap(~.category) +
-    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
-    stat_summary(fun = "mean", geom = "line", size = 0.75) +
+    stat_lineribbon(linewidth = 1, 
+                    colour = NA,
+                    .width = c(0.95, 0.89, 0.67, 0.5)) +
+    stat_summary(fun = "mean", geom = "line", linewidth = 0.75) +
     scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
+        breaks = scale(seq(10, 36, 4), 
+                       mean(responses$age), 
+                       sd(responses$age))[, 1],
         labels = seq(10, 36, 4)
@@
 
-#### Prior-predictions
-
-```{r}
-#| label: fig-prior-predictions
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected prior predictions"
-nd <- expand.grid(
-    n_phon_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
-    exposure_std = 0,
-    lv_std = 0
-)
-
-m <- add_predicted_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA)
-
-m %>% 
-    count(age_std, .prediction) %>% 
-    pivot_wider(names_from = .prediction, values_from = n) %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
-    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
-    geom_col(position = position_fill()) +
-    labs(
-        x = "Age (months)", 
-        y = "Probability of acquisition",
-        colour = "Category", 
-        fill = "Category"
-    ) +
-    scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 4)
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        strip.background = element_rect(fill = "grey", colour = NA),
-        plot.background = element_rect(fill = "white", colour = NA)
-    )
-```
-
-:::
-
-### Model comparison
-
-To test of goodness of fit of our model, we compared its predictive accuracy of this extended model--which included all predictors of interest--to four simpler models that included a subset of our predictors of interest: a base model that only included the main effects of $Age$, $Phonemes$, and $Exposure$ ($\mathcal{M}_{0}$), a model that also included the $Age \times Exposure$ interaction ($\mathcal{M}_{1}$), a model that also included the main effect of $Levenshtein$ ($\mathcal{M}_{2}$), a model that also included the $Age \times Levenshtein$ interaction ($\mathcal{M}_{3}$).
-Only the extended model included the $Age \times Exposure \times Levenshtein$ interaction.
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package.
-LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
-
-### Model settings
-
--   OS: `r sessionInfo()$running`, `r sessionInfo()$platform`
--   R version: `r sessionInfo()$R.version$version.string`
--   Algorithm: Hamiltonian Monte Carlo (No U-turn Sampler)
--   Engine: Stan (`brms` interface, `r sessionInfo()$otherPkgs$brms$Version`)
--   `brms` backend: CmdStanR (`cmdstanr` `r sessionInfo()$loadedOnly$cmdstanr$Version`)
--   Chains: `r dim(model_fit_4$fit)[2]`
--   Cores: `r dim(model_fit_4$fit)[2]`
--   Iterations: `r format(dim(model_fit_4$fit)[1], big.mark = ",")` (`r format(dim(model_fit_4$fit)[1]/2, big.mark = ",")`)
-
-More details in Appendix: Session Info.
-
-### R code (`brms`)
-
-::: panel-tabset
-
-#### Model 0
-
-```{r}
-print(model_fit_0$formula)
-```
-
-#### Model 1
-
-```{r}
-print(model_fit_1$formula)
-```
-
-#### Model 2
-
-```{r}
-print(model_fit_2$formula)
-```
-
-#### Model 3
-
-```{r}
-print(model_fit_3$formula)
-```
-
-#### Model 4
-
-```{r}
-print(model_fit_4$formula)
-```
-:::
-
-# Results
-
-## Descriptive statistics
-
-::: panel-tabset
-
-### By number of phonemes
-
-```{r}
-#| lab: fig-responses-age-phonemes
-#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
-df %>% 
-    mutate(
-        age = floor(age),
-        n_phon = cut_interval(n_phon, n = 4)
-    ) %>%
-    count(age, n_phon, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~n_phon) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-
-### By exposure
-
-```{r}
-#| lab: fig-responses-age-exposure
-#| fig-cap: "Proportion of responses to each category by age and exposure."
-df %>% 
-    mutate(
-        age = floor(age),
-        exposure = cut_interval(exposure, n = 4)
-    ) %>%
-    count(age, exposure, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~exposure) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-
-### By Levenshtein
-
-```{r}
-#| lab: fig-responses-age-lv
-#| fig-cap: "Proportion of responses to each category by age and Levenhstein similarity"
-df %>% 
-    mutate(
-        age = floor(age),
-        lv = cut_interval(lv, n = 4)
-    ) %>%
-    count(age, lv, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~lv) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_y_continuous(labels = percent) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-:::
-
-## Model selection
-
-@tbl-results-loos shows the outputs of the *LOO-CV*.
-The extended model (`model_fit_3`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference.
-This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.
-
-```{r}
-#| label: tbl-results-loos
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
-This procedure involves estimating the likelihood of each observation given a model fitted without such observation.
-The resulting likelihoods of each model were summed to compute the expected log-predicted density (*ELPD*) of the model, and correcting it for the number of parameters in the model to penalised for its the complexity.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*LOO-IC*) of each model along their associated standard errors in paratheses.
-The first column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_4`)."
-model_loos %>% 
-    loo_compare() %>% 
-    as.data.frame() %>%   
-    rownames_to_column("model") %>% 
-    relocate(
-        model, 
-        matches("diff"),
-        matches("elpd_loo"),
-        matches("p_loo"), 
-        matches("looic")
-    ) %>%
-    gt() %>% 
-    fmt_number(2:9, decimals = 1) %>% 
-    cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>% 
-    cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>% 
-    cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>% 
-    cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>% 
-    cols_label(
-        model = "Model",
-        elpd_loo = md("*ELPD*"),
-        p_loo = md("*p*"),
-        looic = md("*LOO-IC*"),
-        elpd_diff = md("Difference")
-    ) %>% 
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels()
-    ) %>% 
-    tab_style(
-        cell_fill(color = "grey", alpha = 0.5),
-        cells_body(matches("diff"))
-    )
-```
-
-## Fixed effects
-
-```{r}
-#| label: results-posterior-fixed
-#| echo: true
-#| message: false
-#| warning: false
-post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)
-
-# tidy predictor names
-str_repl <- c(
-    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
-    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
-    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
-    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
-    "b_exposure_std" = paste0("Exposure (+1 SD, ",  percent(sd(df$exposure)), ")"),
-    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
-    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
-    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
-    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
-)
-
-```
-
-```{r}
-#| label: tbl-results-fixed
-#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-intercept_footnote <- paste0(
-    "Probability of comprehension or production at the grand age of the sample (", 
-    round(mean(df$age), 2), " months)"
-)
-
-# summarise posterior draws
-posterior_description %>% 
-    mutate(parameter = factor(parameter, levels = names(str_repl), labels = str_repl)) %>%  
-    select(parameter, median, ci_high, ci_low, rope_percentage) %>% 
-    gt(rowname_col = "parameter") %>% 
-    fmt_number(c(1:4), decimals = 2) %>% 
-    fmt_percent(5, decimals = 2) %>% 
-    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>% 
-    cols_label(
-        median = "Median",
-        ci_low = md("95% *HDI*"),
-        rope_percentage = md("*ROPE* prob.")
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = 1:5)
-    )
-
-post_draws_list <- split(posterior_description, posterior_description$parameter)
-```
-
-We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
-We decided to draw this interval around zero so that it covered values smaller than those considered as negligible by similar previous studies: after consulting the regression tables of [ADD STUDIES], we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: 3%.
-consequently, we defined our *ROPE* from -3% to +3%.
-We report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
-This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
-Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the -3%--3% *ROPE* interval we previously defined.
-For those coefficients that did not exclude completely the *ROPE* we also report a Bayes Factor indicating how many times is such coefficient more likely to be bigger than the upper limit of the *ROPE*, or smaller than the lower limit of the *ROPE*, given our data.
-
-For interpretability, we report the estimated regression coefficients transformed from the probability scale.
-The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
-This means that one logit difference is not necessarily translated to a unique value in the probability scale.
-For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
-The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
-We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
-derivative) of the logistic curve at the age at which most participants were acquiring a given word.
-This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
-
-@fig-results-fixed shows the outputs of the model.
-**Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r printnum(post_draws_list[["b_age_std"]]$median, digits = 2)`, 95% *HDI* = [`r printnum(post_draws_list[["b_age_std"]]$ci_low, digits = 2)`, `r printnum(post_draws_list[["b_age_std"]]$ci_high, digits = 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month.
-
-The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]).
-For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.
-
-**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]).
-All of the posterior samples of his regression excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/(sd(df$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.
-
-The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
-
-However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word.
-Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
-This effect was negligible when exposure to both languages was balanced (e.g.
-50-50%).
-Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.
-
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
-Given our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.
-
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
-
-```{r}
-#| label: fig-results-fixed
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 9
-#| fig-heigh: 7
-#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model.
-Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis).
-Intervals show the 95% highest density interval (*HDI*) of each distribution.
-This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-the mean and HDI limits are indicated below each distribution.
-Coefficients were transformed from the logit scale to the probability scale for interpretability.
-We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e.
-the derivative of the logistic function in its mid-point."
-post <- post_draws %>% 
-    mutate(
-        .variable_name = factor(
-            .variable, 
-            levels = names(str_repl),
-            labels = str_repl,
-            ordered = TRUE
-        )
-    )
-
-post_summary <- post %>% 
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    median_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>% 
-    mutate(
-        .variable_name = factor(
-            .variable, 
-            levels = names(str_repl),
-            labels = str_repl,
-            ordered = TRUE
-        )
-    )
-
-post %>%
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    annotate(geom = "rect", ymin = -Inf, ymax = Inf,
-             xmin = rope_interval["lower"], xmax = rope_interval["upper"],
-             colour = NA, alpha = 0.2, fill = clrs[5],) +
-    geom_vline(xintercept = 0, size = 1, colour = clrs[5]) +
-    stat_slab(
-        aes(fill = stat(abs(x) < rope_interval["upper"]),
-            colour = stat(abs(x) < rope_interval["upper"]),),
-        size = 1, position = position_nudge(y = 0.15)
-    ) +
-    geom_errorbar(
-        data = post_summary,
-        aes(xmin = .lower, xmax = .upper, x = .value),
-        width = 0.15
-    ) +
-    geom_point(data = post_summary, aes(x = .value), size = 1.5) +
-    geom_text(
-        data = post_summary,
-        aes(
-            x = .value, 
-            y = .variable_name,
-            label = paste0(
-                round(.value, 1),
-                " [", round(.lower, 1), ", ", round(.upper, 1), "]")
-        ),
-        colour = "black", position = position_nudge(y = -0.3), size = 3
-    ) +
-    labs(
-        x = "Logistic regression coefficient estimate (Logit scale)",
-        y = "Variable", 
-        fill = "Overlaps with ROPE",
-        colour = "Overlaps with ROPE"
-    ) +
-    scale_x_continuous(breaks = seq(-2, 2, 0.2)) +
-    scale_fill_manual(
-        values = clrs[c(1, 4)],
-        labels = c("No", "Yes")
-    ) +
-    scale_colour_manual(
-        values = clrs[c(1, 4)],
-        labels = c("No", "Yes")
-    ) +
-    theme(
-        legend.position = "top",
-        legend.justification = c(1, 1),
-        axis.title.y = element_blank(),
-        panel.grid.major.x = element_line(
-            colour = "grey85",
-            linetype = "dotted"
-        ),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
-
-## Marginal predictions
-
-```{r}
-#| label: fig-results-marginal
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected posterior predictions.
-Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
-These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model.
-Therefore, we generated 4,000 predictions for a series of combination of levels of interest.
-Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels.
-The X-axis indicates the age (in months) for which the prediction is generated.
-The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*).
-Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
-Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
-Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
-
-nd <- datagrid(
-    model = model_fit_4,
-    age_std = seq_range(model_fit_4$data$age_std, n = 100),
-    lv_std = (seq(0, 1, 0.5)-mean(df$lv))/(sd(df$lv)),
-    exposure_std = c(-1, 0, 1),
-    te = NA,
-    id = NA
-)
-
-marg_epreds <- predictions(
-    model_fit_4,
-    newdata = nd,
-    ndraws = 25, 
-    re_formula = NA
-)
-
-marg_epreds %>% 
-    posteriordraws() %>% 
-    as_tibble() %>% 
-    select(drawid, group, draw, age_std, lv_std, exposure_std) %>%
-    filter(group != "No") %>% 
-    pivot_wider(
-        id_cols = c(drawid, age_std, lv_std, exposure_std), 
-        names_from = group, 
-        values_from = draw
-    ) %>% 
-    mutate(`Understands` = `Understands and Says` + `Understands`) %>% 
-    pivot_longer(c(`Understands`, `Understands and Says`), names_to = "group", values_to = "draw") %>% 
-    mutate(
-        exposure_std = factor(
-            exposure_std, 
-            levels = unique(nd$exposure_std), 
-            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
-        ),
-        lv_std = factor(
-            lv_std,
-            levels = unique(nd$lv_std), 
-            labels = paste0(c(0, 50, 100), "% similarity")
-        )
-    ) %>% 
-    ggplot(aes(age_std, draw, colour = lv_std, fill = lv_std)) +
-    facet_grid(group~exposure_std) +
-    # stat_summary(fun.data = median_hdi, geom = "ribbon",
-    #              alpha = 0.5, colour = NA) +
-    geom_line(aes(group = interaction(lv_std, drawid)),
-              size = 0.75, alpha = 0.15) +
-    stat_summary(aes(linetype = lv_std), fun = median, geom = "line", size = 1) +
-    labs(
-        x = "Age (months)",
-        y = "Probability of acquisition",
-        linetype = "Phonological similarity (Levenshtein)",
-        colour = "Phonological similarity (Levenshtein)",
-        fill = "Phonological similarity (Levenshtein)"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(
-        breaks = seq(0, 1, 0.25),
-        limits = c(0, 1),
-        labels = percent
-    ) +
-    scale_x_continuous(
-        breaks = (seq(10, 32, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 32, 4)
-    ) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
-    )
-
-```
-
-## Marginal effects
-
-In order to examine the impact of the word exposure index on the effect of phonological similarity we used the [`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects/index.html) R package to compute the average marginal effect of the Levenshtein distance on the probability of acquisition (comprehension or production) for the range of observed values of the word exposure index in our dataset.
-
-```{r}
-#| label: fig-marginal-effect
-#| fig-cap: "Average marginal effect of phonological similarity (Levenshtein distance) on the probability of comprehension and production across the range of values of exposure.
-The observed range of values of the word exposure index is indicated in the X-axis, and the estimated average marginal effect of phonological similarity (Levenshtein distance) is indicated in the Y-axis.
-This quantity is generated by averaging the derivative of the regression coefficient of the predictor Levenshtein across the range of values of all predictors but Exposure.
-We represent the uncertainty associated with the resulting estimate by generating multiple values of this marginal effect by drawing multiple samples from the posterior distribution of the model coefficients, calculating the average marginal effect for each posterior draw, and plotting the result as a thin grey line.
-Closer lines indicate lower uncertainty over the true value of the average marginal effect of Levenshein, while more distant lines indicate higher uncertainty."
-nd <- datagrid(
-    model = model_fit_4,
-    n_phon_std = 0,
-    lv_std = 0,
-    exposure_std = seq(-1.5, 1.5, length.out = 100),
-    age_std = 0,
-    id = NA,
-    te = NA
-)
-
-marg_effects <- marginaleffects(
-    model_fit_4,
-    newdata = nd, 
-    re_formula = NA,
-    ndraws = 50,
-    vcov = FALSE,
-    type = "link"
-)
-
-marg_effects %>% 
-    posteriordraws() %>%
-    as_tibble() %>% 
-    filter(term=="lv_std") %>% 
-    mutate(draw = draw/4) %>% 
-    ggplot(aes(exposure_std, draw)) +
-    geom_hline(yintercept = 0, size = 1, colour = "grey") +
-    geom_line(aes(group = drawid), alpha = 0.25, size = 0.75, colour = clrs[4]) +
-    # stat_summary(fun.data = median_hdi, geom = "ribbon",
-    #              colour = NA, alpha = 0.5, fill = clrs[1]) +
-    stat_summary(fun = median, geom = "line", size = 1, color = clrs[1]) +
-    labs(
-        x = "Word exposure index",
-        y = "Marginal effect of phonological similarity\n(Levenshtein distance)"
-    ) +
-    scale_x_continuous(
-        labels = label_number(style_positive = "plus", suffix = " SD"),
-        breaks = seq(-2, 2, 0.5)
-    ) +
-    scale_y_continuous(
-        labels = label_percent(style_positive = "plus", ),
-        breaks = seq(-1, 1, 0.05)
-    ) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
-    )
-
-```
-
-## Random effects
-
-### Participant-level effects
-
-::: panel-tabset
-
-#### Intercept
-
-```{r}
-#| label: fig-results-random-id-intercept
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior probability of acquisition by participant"
-re_id <- ranef(model_fit_4)$id[,,1] %>% 
-    as_tibble() %>% 
-    rownames_to_column("id") %>% 
-    left_join(distinct(model_fit_4$data, id)) %>% 
-    clean_names() %>% 
-    mutate(
-        across(
-            .cols = where(is.numeric),
-            .fns = list(
-                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
-                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
-            ), .names = "{.col}__{.fn}"
-        )
-    ) %>% 
-    select(-c(estimate:q97_5)) %>% 
-    pivot_longer(
-        where(is.numeric), 
-        names_to = c(".value", "type"),
-        names_sep = "__"
-    )
-
-re_id %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate), 
-        xmin = q2_5, xmax = q97_5,
-        fill = type, colour = type) +
-    facet_wrap(~type) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "Participant", 
-        colour = "Type"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age slope
-
-```{r}
-#| label: fig-results-random-id-age
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_age <- ranef(model_fit_4)$id[,,"age_std"]
-
-re_age %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(id = unique(model_fit_4$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### # Phonemes slope
-
-```{r}
-#| label: fig-results-random-id-nphon
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the number of phonemes by participant"
-re_n_phon <- ranef(model_fit_4)$id[,,"n_phon_std"]
-
-re_n_phon %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(id = unique(model_fit_4$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(
-        x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5
-    ) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Exposure slope
-
-```{r}
-#| label: fig-results-random-id-exposure
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of lexical frequency by participant"
-re_freq <- ranef(model_fit_4)$id[,,"exposure_std"]
-
-re_freq %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(id = unique(model_fit_4$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of phonological similarity by participant"
-re_lv <- ranef(model_fit_4)$id[,,"lv_std"]
-
-re_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Doe-by-Levenshtein slope
-
-```{r}
-# | label: fig-results-random-id-exposure-lv
-# | echo: true
-# | message: false
-# | warning: false
-# | fig-width: 7
-# | fig-cap: "Posterior random slope of age by participant"
-re_exposure_lv <- ranef(model_fit_4)$id[,,"exposure_std:lv_std"]
-
-re_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-DoE slope
-
-```{r}
-#| label: fig-results-random-id-age-doe
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
-re_age_exposure <- ranef(model_fit_4)$id[,,"age_std:exposure_std"]
-
-re_age_exposure %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-age-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
-re_age_lv <- ranef(model_fit_4)$id[,,"age_std:lv_std"]
-
-re_age_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-DoE-by-Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-age-exposure-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
-re_age_exposure_lv <- ranef(model_fit_4)$id[,,"age_std:exposure_std:lv_std"]
-
-re_age_exposure_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_3$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Correlations
-
-```{r}
-#| label: fig-results-random-id-correlations
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior correlations between random effects by participant"
-posterior_description_re %>% 
-    select(parameter, median) %>% 
-    filter(str_detect(parameter, "cor_id__")) %>% 
-    mutate(parameter = str_remove_all(parameter, "cor_id__")) %>%  
-    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
-    mutate(
-        across(
-            c(var_1, var_2), 
-            ~factor(
-                .,
-                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std", "exposure_std:lv_std", "age_std:exposure_std:lv_std"),
-                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein", "Exposure \u00d7 Levenshtein", "Age \u00d7 Exposure \u00d7 Levenshtein")
-            )
-        )
-    ) %>%
-    filter(var_1 != var_2) %>% 
-    ggplot(aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1))) +
-    geom_point(shape = 15, size = 20) +
-    geom_text(colour = "black") +
-    scale_colour_gradientn(colours = clrs) +
-    labs(colour = "Correlation", size = "Correlation") +
-    theme(
-        axis.title = element_blank(),
-        axis.text.x = element_text(angle = 35, hjust = 1),
-        panel.grid = element_blank()
-    )
-```
-:::
-
-### Item-level effects
-
-::: panel-tabset
-
-#### Intercept
-
-```{r}
-#| label: fig-results-random-te
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior probability of acquisition by translation equivalent"
-re_te <- ranef(model_fit_4)$te[,,"Intercept"] %>% 
-    as_tibble() %>% 
-    rownames_to_column("te") %>% 
-    clean_names() %>% 
-    mutate(
-        across(
-            .cols = where(is.numeric),
-            .fns = list(
-                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
-                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
-            ), .names = "{.col}__{.fn}"
-        )
-    ) %>% 
-    select(-c(estimate:q97_5)) %>% 
-    pivot_longer(
-        where(is.numeric), 
-        names_to = c(".value", "type"),
-        names_sep = "__"
-    ) %>% 
-    mutate(te = as.numeric(te))
-
-re_te %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate), 
-        xmin = q2_5, xmax = q97_5,
-        colour = type, fill = type) +
-    facet_wrap(~type) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "TE", 
-        colour = "Type"
-    ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age slope
-
-```{r}
-#| label: fig-results-random-te-age
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by translation equivalent"
-re_age <- ranef(model_fit_4)$te[,,"age_std"]
-
-re_age %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_4$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### # Phonemes slope
-
-```{r}
-#| label: fig-results-random-te-nphon
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
-re_n_phon <- ranef(model_fit_3)$te[,,"n_phon_std"]
-
-re_n_phon %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_3$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Exosure slope
-
-```{r}
-#| label: fig-results-random-te-exposure
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_doe <- ranef(model_fit_4)$te[,,"exposure_std"]
-
-re_doe %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(te = unique(model_fit_4$data$te)) %>%
-    relocate(te) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-exposure slope
-
-```{r}
-#| label: fig-results-random-te-age-exposure
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_age_doe <- ranef(model_fit_4)$te[,,"age_std:exposure_std"]
-
-re_age_doe %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_4$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Correlations
-
-```{r}
-#| label: fig-results-random-te-correlations
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior correlations between random effects by translation equivalent"
-posterior_description_re %>% 
-    select(parameter, median) %>% 
-    filter(str_detect(parameter, "cor_te__")) %>% 
-    mutate(parameter = str_remove_all(parameter, "cor_te__")) %>%  
-    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
-    mutate(
-        across(
-            c(var_1, var_2), 
-            ~factor(
-                .,
-                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std"),
-                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein")
-            )
-        )
-    ) %>%
-    filter(var_1 != var_2) %>% 
-    ggplot() +
-    aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1)) +
-    geom_point(shape = 15, size = 40) +
-    geom_text(colour = "black") +
-    scale_colour_gradientn(colours = clrs) +
-    labs(
-        colour = "Correlation",
-        size = "Correlation"
-    ) +
-    theme(
-        axis.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-:::
-
-## Model diagnostics
-
-### Simulation-based calibration
-
-```{r}
-#| label: fig-simulation-based-calibration
-#| fig-cap: "Posterior predictive checks.
-We simulated 500 datasets from our model by drawing random samples from its posterior distribution.
-We then computed the proportion of 'No', 'Understands', and 'Understands and Says' responses in each dataset.
-These simulated proportions are depicted in the plot as solid horizontal lines.
-Their variation indicate the uncertainy of the model about what proportion of each type of response would be observe in a newly collected dataset.
-We also plotted the proportion of each type of response in the observed dataset (with a ± 5% interval for comparison)."
-predicted_dfs <- add_predicted_draws(
-    object = model_fit_4, 
-    newdata = df,
-    re_formula = NULL,
-    ndraws = 500,
-    seed = 888
-)
-
-observed_prop <- df %>% 
-    count(response) %>%
-    mutate(prop = n/sum(n)) %>% 
-    rename(.prediction = response)
-
-predicted_dfs %>% 
-    ungroup() %>% 
-    select(.prediction, .draw) %>% 
-    count(.draw, .prediction) %>%
-    group_by(.draw) %>% 
-    mutate(
-        prop = n/sum(n),
-        .prediction = factor(
-            .prediction,
-            levels = levels(df$response), 
-            ordered = TRUE
-        )
-    ) %>% 
-    ggplot(aes(.draw, prop, fill = .prediction, colour = .prediction)) +
-    geom_line(aes(group = .prediction)) +
-    geom_rect(
-        data = observed_prop,
-        aes(xmin = 0, xmax = max(predicted_dfs$.draw),
-            ymin = prop-0.025, ymax = prop+0.025,
-            colour = .prediction, fill = .prediction),
-        alpha = 0.25, colour = NA, inherit.aes = FALSE
-    ) +
-    # geom_segment(
-    #     data = observed_prop,
-    #     aes(x = -Inf, xend = Inf, y = prop, yend = prop,
-    #         colour = .prediction),
-    #     size = 1, inherit.aes = FALSE
-    # ) +
-    labs(
-        x = "Iteration",
-        y = "Proportion of responses",
-        colour = "Response",
-        fill = "Response"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_x_continuous(labels = label_number(big.mark = ","), 
-                       breaks = seq(0, max(predicted_dfs$.draw), 100)) +
-    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank()
-    )
-
-```
-
-### Traceplots
-
-```{r}
-#| label: fig-diagnostics-traceplots
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 12
-gather_draws(model_fit_4, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
-    mutate(.chain = paste("Chain ", .chain)) %>%
-    ggplot(aes(.iteration, .value, colour = .chain)) +
-    facet_wrap(~.variable, scales = "free_y") +
-    annotate(geom = "rect", colour = NA,
-             xmin = 0, xmax = dim(model_fit_4$fit)[1]/2,
-             ymin = -Inf, ymax = Inf,
-             alpha = 0.25) +
-    geom_line() +
-    labs(
-        x = "Iteration",
-        y = "Sample value",
-        colour = "Chain"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 3, 4, 5)]) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank(),
-        axis.text.x = element_text(size = 7)
-    )
-```
-
-### Leave-one-out diagnostics (Pareto-*k* diagnostics)
-
-::: panel-tabset
-
-#### Model 0
-
-```{r}
-#| label: fig-loo-diagnostics-0
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
-get_loo_plot <- function(x){
-    x$pointwise[,5] %>% 
-        data.frame(x = 1:length(.), y = .) %>% 
-        ggplot(aes(x, y)) +
-        geom_point(
-            size = 0.5, alpha = 0.1, colour = clrs[4], 
-            shape = 1
-        ) +
-        geom_hline(yintercept = 0.5, linetype = "dotted") +
-        labs(
-            x = "Data point",
-            y = "Pareto-k value",
-            title = "PSIS diagnostic plot"
-        ) +
-        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
-        scale_y_continuous(
-            limits = c(-0.6, 0.6)
-        )
-}
-
-get_loo_plot(model_loos$model_fit_0)
-```
-
-#### Model 1
-
-```{r}
-#| label: fig-loo-diagnostics-1
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
-get_loo_plot(model_loos$model_fit_1)
-```
-
-#### Model 2
-
-```{r}
-#| label: fig-loo-diagnostics-2
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
-get_loo_plot(model_loos$model_fit_2)
-```
-
-#### Model 3
-
-```{r}
-#| label: fig-loo-diagnostics-3
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
-get_loo_plot(model_loos$model_fit_3)
-```
-
-#### Model 4
-
-```{r}
-#| label: fig-loo-diagnostics-4
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
-get_loo_plot(model_loos$model_fit_4)
-```
-
-:::
-
-### Gelman-Rubin diagnostic (R-hat)
-
-```{r}
-#| label: fig-diagnostics-rhat
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| eval: false
-mcmc_rhat_data(model_rhats$model_fit_4) %>% 
-    ggplot() +
-    aes(x = value) +
-    geom_histogram(bins = 100, na.rm = TRUE, fill = clrs[1], colour = "white") +
-    geom_rug(colour = clrs[1]) +
-    labs(
-        x = "R-hat",
-        y = "Number of samples",
-        colour = "Description",
-        fill = "Description"
-    ) +
-    theme(
-        legend.position = "top"
-    )
-```
-
-### Effective sample size
-
-```{r}
-#| label: fig-diagnostics-neff
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| eval: false
-mcmc_neff_data(model_neffs$model_fit_4) %>% 
-    ggplot() +
-    aes(x = value, fill = description, colour = description) +
-    geom_histogram(bins = 100, na.rm = TRUE, colour = "white") +
-    geom_rug() +
-    geom_vline(xintercept = 1, colour = "black", size = 0.75) +
-    labs(
-        x = "Effective sample size ratio",
-        y = "Number of samples",
-        colour = "Rating",
-        fill = "Rating"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
-    theme(
-        legend.position = "top"
-    )
-```
-
-# Discussion
-
-## Summary of the study
-
-## Summary of the results
-
-## Results in contrast with the literature
-
-## Mechanistic explanations
-
-## Limitations
-
-## Further steps
-
-# Appendix
-
-## Session info
-
-```{r}
-sessioninfo::session_info()
-```
+{{< pagebreak >}}

### 2023-02-26 manuscript/_model-diagnostics.qmd
Add appendices

We used Stan [@carpenter2017stan] as the probabilistic language behind the estimation of our Bayesian models in this study, with `brms` as its R interface [@burkner2017brms]. This language implements the Markov Chain Monte-Carlo (MCMC) algorithm to explore the posterior distribution of the model. Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined. We run `r dim(model_fit_4$fit)[2]` MCMC chains, each `r dim(model_fit_4$fit)[1]` iterations long each. The correct performance of this algorithm is critical to the quality of the statistical evidence to which the outcomes of the model lead.

One way to diagnose the behaviour of MCMC is to inspect whether the different MCMC chains (if more than one) have converged to a similar region of the posterior. The Gelman-Rubin diagnostic [$\hat{R}$ or R-hat @gelman1992inference] provides a measure of chain convergence by comparing the variance within each chain *versus* the variance between each chain. Both are expected to be identical when chains have perfectly converged, so that $\hat{R} = 1$. Values lower than 1.01 are recommended, while values higher than 1.05 indicate that chains might have trouble converging and therefore the estimated parameters must be taken with caution. @fig-model-rhats-neffs (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference. Most values are lower than 1.01, and never higher than 1.05, which provides evidence of successful MCMC convergence.

Another diagnostic of good MCMC converge is the ratio of effective sample size to total sample size ($N_{eff}/N$), which indicates the proportion of samples in the chain that resulted from a non-divergent transition. Values closer to 1 are ideal, as they indicate that all posterior samples from the MCMC were used to estimate the posterior distribution of the parameter. Values larger than 0.1 are recommended. @fig-model-rhats-neffs (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models. Most values are larger than 0.1, although model 0 ($\mathcal{M}_0$) accumulates most effective sample sizes close to 0.1.

```{r}
#| label: fig-model-rhats-neffs
#| fig-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6
#| fig-width: 9
map_df(model_rhats, bind_rows, .id = "model") |> 
    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
    pivot_longer(-model, 
                 names_to = ".variable",
                 values_to = ".rhat") |>
    filter(str_detect(.variable, "b_|sd_")) |> 
    drop_na() |> 
    ggplot(aes(.rhat, colour = model, fill = model)) +
    facet_wrap(~model, nrow = 1) +
    geom_vline(xintercept = 1.01, colour = "grey", linewidth = 0.75) +
    geom_histogram(bins = 12, colour = "white") +
    geom_rug(alpha = 0.5) +
    labs(x = "Gelman Rubin diagnostic (R-hat)",
         y = "MCMC samples",
         title = "Gelman-Rubin (R-hat) diagnostic") +

    map_df(model_neffs, bind_rows, .id = "model") |> 
    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
    pivot_longer(-model, 
                 names_to = ".variable",
                 values_to = ".rhat") |>
    filter(str_detect(.variable, "b_|sd_")) |> 
    drop_na() |> 
    ggplot(aes(.rhat, colour = model, fill = model)) +
    facet_wrap(~model, nrow = 1) +
    geom_vline(xintercept = 0.1, colour = "grey", linewidth = 0.75) +
    geom_histogram(bins = 10, colour = "white") +
    geom_rug(alpha = 0.5) +
    labs(x = "Gelman Rubin diagnostic (R-hat)",
         y = "MCMC samples",
         title = "Ratio of effective sample size") +
    scale_x_continuous(labels = percent) +

    plot_layout(ncol = 1) &
    plot_annotation(tag_levels = "A") &
    theme(legend.position = "none",
          panel.grid = element_blank())
```

## Posterior draws and bi-variate scatterplots

```{r}
#| label: fig-model-pairs
#| fig-cap: "Marginal distribution and bi-variate scatterplot of posterior samples for the fixed regression coefficients in Model 4."
#| fig-height: 12
#| fig-width: 14
#| echo: false
#| warning: false
#| message: false
titles <- c("Comprehension",
            "Production",
            "Age",
            "Phonemes",
            "Exposure",
            "Levenshtein",
            "Age × Exposure",
            "Age × Levenshtein",
            "Exposure × Levenshtein",
            "Age × Exposure × Levenshtein")

posterior <- as.array(model_fit_4)
np <- nuts_params(model_fit_4) 
mcmc_pairs(posterior,
           np = np, 
           pars = c("b_Intercept[1]",
                    "b_Intercept[2]",
                    "b_age_std",
                    "b_n_phon_std",
                    "b_exposure_std",
                    "b_lv_std",
                    "b_age_std:exposure_std",
                    "b_age_std:lv_std",
                    "b_exposure_std:lv_std",
                    "b_age_std:exposure_std:lv_std"),
           grid_args = list(subtitles = titles),
           off_diag_args = list(size = 0.5,
                                alpha = 1/10))
```

## Posterior-predictive checks

```{r}
#| label: fig-model-ppc
#| fig-cap: "Model posterior predictive checks (PPC). Bars indicate the observed proportion of responses to each category (*No*, *Understands*, and *Understands and Says*). Blue dots and error bars represent the mean proportion of responses simulated from the posterior for each category, and its 95% interval."
#| eval: true
#| echo: false
#| warning: false
#| message: false
y_int <- as.integer(as.integer(responses$response))
color_scheme_set(c(clrs[c(4, 4, 3, 2, 1)], "white"))

ppc_bars(y_int, 
         model_ppcs,
         prob = 0.95,
         linewidth = 1,
         size = 1.5,
         fatten = 2,
         freq = FALSE) +
     scale_x_continuous(breaks = 1:3,
                        labels = c("No",
                                   "Understands", 
                                   "Understands and Says")) +
    theme(legend.position = "none",
          axis.text = element_text(size = 11))
```

### 2023-02-26 manuscript/_model-doe.qmd
Add appendices

```{r}
#| label: fig-coefs-doe
#| fig-cap: "Posterior distribution of regression coefficients in the probability scale. *Median*: median of the posterior distribution in the logit scale. *95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
#| fig-height: 4
#| fig-width: 10
# summarise posterior draws
bind_rows("Exposure model" = posterior_draws$summary,
          "DoE-Frequency model" = posterior_draws_doe,
          .id = "model") |> 
    mutate(
        across(c(.median, .lower, .upper),
               function(x) {
                   ifelse(
                       .type=="Intercepts (at 22 months)",
                       plogis(x),
                       x/4
                   )
               }
        )) |>
    filter(.type == "Slopes") |> 
    ggplot(aes(.median, .variable_name, xmin = .lower, xmax = .upper,
               colour = model)) +
    facet_wrap(~model, scales = "free") +
    annotate(geom = "rect",
             ymin = -Inf,
             ymax = Inf,
             xmin = -0.1/4,
             xmax = +0.1/4,
             fill = "grey70") +
    geom_errorbar(linewidth = 1, width = 0.25) +
    geom_point(size = 2.5) +
    labs(x = "Median and 95% HDI of the posterior distribution",
         y = "Predictor",
         colour = "Model") +
    scale_x_continuous(labels = percent) +
    scale_colour_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          axis.title.y = element_blank())
```

### 2023-02-26 manuscript/_model-predictors.qmd
Add appendices

## Word exposure index

Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database [@macwhinney2000childes; @sanchez2019childes]. Using the corresponding lexical frequencies from the Catalan and Spanish corpora was not possible due to the low number of Catalan participants and tokens available in those languages. Available words in the English corpora were mapped to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human]. Responses to words with missing lexical frequencies were excluded from analyses of the total number of items. Participant degree of exposure to the words' language were calculated as the percentage of exposure to the language, as reported by caretakers in the language profile section of the questionnaire. For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.

We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to. the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).

$$
\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
$$ {#eq-exposure}

For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`) would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.

## Levenshtein similarity

Generally, a pair of words from two different languages that share meaning---translation equivalents---are considered as *cognates* if they share etymological origin (e.g., *table* and *taula*, in English and Catalan). The fact that two words share etymological origin frequently leads to them also being form-similar, as reflected by their overlapping orthographic or phonological form. This makes many cognates perceptually similar. In psycholinguistics, the term *cognateness* is often used to refer to the form-similarity that a pair of translation equivalents share, which has been found to impact how bilinguals process such word-forms (e.g., @costa2000cognate; @spivey1999cross). Whether two form-similar translation equivalents are etymologically related or not is arguably tangential to the question of how cognateness affect language processing. For instance, the translation pair *much* and *mucho*, in English and Spanish have the same meaning, and are considerably similar at the orthographical and phonological level. Nonetheless, they come from different Proto-Indo-European roots [@campbell2007glossary]. This does not keep bilinguals from processing such kind of word-forms word forms differently than other translation equivalents with no form-similarity like 'dog' and *perro*, in English and Spanish. In the scope of this experiment, we consider any form-similar translation equivalent as a *cognate* pair, regardless of whether both word-forms share etymological origin.

Our study capitalises in phonology, therefore we operationalised *cognateness* as the phonological similarity between a pair of translation equivalents. Our dataset contained `r max(n_items_final)` translation equivalents, each consisting in two word-forms: Catalan and Spanish (see @tbl-translation-levenshtein for an example). Measuring the phonological similarity between two word forms is non trial. While orthographic similarity can be calculated using string similarity/distance measures on the written word-forms, phonology poses additional challenges. A first, bottom-up approach to to the task might be considering acoustic similarity as a proxy to phonological similarity. This may involve registering audio recordings of a talker reading each word-form aloud, and then comparing their spectrograms,  cochleograms, or formant tracks [e.g., @heeringa2003norwegian]. While this is possible, this method computes acoustic similarity, as opposed to phonological similarity. This is not optimal, since a purely acoustic measure of similarity ignores how the actual signal is perceived by a actual listener. Since approximately 12 months of age, humans perceive the acoustic sounds in speech as phonemes, according to the phonology of their native language(s) [@werker1984crosslanguage; @kuhl1992linguistic]. This means that two listeners may not perceive the same linguistically relevant acoustic signal identically. This is why phonological similarity needs to be computed from discrete units that are meaningful for language perception.

A way to achieve this is to use phonological transcriptions of the word forms. Phonological transcriptions are symbolic, visual representations of the sounds in speech that capture some minimal  set of its features following some standard. The International Phonetic Alphabet (IPA) is one instance of such standards, providing a set of symbols that correspond to specific phonemes, as defined by their position across a series of dimensions. Using IPA transcriptions of word forms, one can use the same similarity/distance metric as in the case of orthographic word-forms. This method has been successfully used to measure the pairwise similarity between phonological word-forms from the  same language [e.g., @fourtassi2020growth] and across languages [@heeringa2004measuring; @floccia2018methods], and also between orthographic word-forms [@schepens2012distributions].

Multiple algorithms have been created to compute the distance between two strings of characters. The Levenshtein distance being one of them. The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary, see @eq-levenshtein]. For instance, /kaza/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /kasa/ (/z/ is replaced by /s/). Therefore, the Levenshtein distance between /kaza/ and /kasa/ is one. For an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions in X-SAMPA format , as opposed to IPA, for its computer-friendly set of symbols [@wells1995computer].

$$
\text{lev}(x, y) =  
\begin{cases}
\max(i, j) & \text{if} \ \min(i, j)=0\\
\min \begin{cases}
\ \text{lev}_{x, y}(i-1, j) + 1 \\
\ \text{lev}_{x, y}(i, j-1) + 1 \\
\ \text{lev}_{x, y}(i-1, j-1) + 1_{a_i \neq b_j} \\
\end{cases}
\ &\text{otherwise}
\end{cases}
$${#eq-levenshtein}

Here, $a$ and $b$ are the character strings corresponding to the phonological transcriptions of two word-forms belonging to the same translation equivalent, where each character corresponds to one phoneme expressed as a symbol from the SAMPA alphabet. $i$ and $j$ are the length (i.e., number of phonemes) of the $s$ and $t$ strings respectively.

To measure the phonological similarity between the translation equivalents in our study, we computed a normalised versions of the Levenshtein distance between each pair of word-forms that accounts for the lenght of the strings. The rationale behind this correction is that longer word-forms are more likely to differ from their counterpart, compared to shorter word forms, and that this tendency does not correspond necessarily to a larger distance in terms of perception. This normalisation is achieved by dividing the Levenshtein distance by the length of the longest string, which leads to a distance metric that ranges from 0 to 1. This can be interpreted as a proportion of characters in the longest string that need to be edited in order for that string to become identical to the shorter string. A 0% normalised Levenshtein distance indicates that both word-forms are identical. A 50% normalised Levenshtein distance indicates that half of the phonemes in the longest word-form must be edited for both word-forms to become identical. A 100% normalised Levenshtein distance indicates that the two word-forms are completely different, as *all* phonemes in the longest word-form must be changed for it to become identical to the other.

Finally, since we are interested in *cognateness*, in order to easy the interpretation of the analyses we subtracted the normalised Levenshtein distance from one, so that the metric indicated the *similarity* between the each pair of word-forms, instead of their distance. @eq-levenshtein-similarity shows the formula of this metric.

$$
1-\frac{\text{lev}(a, b)}{\max{\{i, j}\}}
$${#eq-levenshtein-similarity}

The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure (see @tbl-translation-levenshtein for more examples):

| Catalan 	      | Spanish   	      | Levenshtein  |       
|---------	      |---------	      |-----         |
| *porta* [pOrt5] | *puerta* [pwe4ta] |   0.50 (3)   |
| *taula* [tawl5] | *mesa* [mesa]     |   0.00 (5)   |
| *cotxe* [kOtS5] | *coche* [kotSe]   |   0.40 (3)   |

: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study {#tbl-translation-levenshtein}

### 2023-03-29 manuscript/appendix/_model-details.qmd
New version of the manuscript, with revised Discussion

--- 2023-02-26 manuscript/_model-details.qmd
@@
 
-To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), word-form length ($Length$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
+To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($\text{Age}$), word-form length ($\text{Length}$), word exposure index ($\text{Exposure}$), and of phonological similarity ($\text{Cognateness}$), the two-way interactions $\text{Age} \times \text{Exposure}$, $\text{Age} \times \text{Cognateness}$, and $\text{Exposure} \times text{Leveshtein}$, and the three-way interaction $\text{Age} \times \text{Exposure} \times \text{Leveshtein}$.
 We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
@@
 & (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
-& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) · \text{Cognateness}_{ij} + & \\
 & (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
-& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
-& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
-& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
+& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
+& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
+& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
 \end{aligned}
@@
 This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations.
-We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].
+We used a weakly informative prior for our parameters, with the exception of the main effect of $\text{Age}$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].
 

### 2023-03-29 manuscript/appendix/_model-diagnostics.qmd
New version of the manuscript, with revised Discussion

--- 2023-02-26 manuscript/_model-diagnostics.qmd
@@
 Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined.
-We run `r dim(model_fit_4$fit)[2]` MCMC chains, each `r dim(model_fit_4$fit)[1]` iterations long each.
+We run `r dim(model_fit_3$fit)[2]` MCMC chains, each `r dim(model_fit_3$fit)[1]` iterations long each.
 The correct performance of this algorithm is critical to the quality of the statistical evidence to which the outcomes of the model lead.
@@
 #| warning: false
-#| fig-height: 6
+#| fig-height: 4
 #| fig-width: 9
@@
 
-## Posterior draws and bi-variate scatterplots
+## Posterior draws and bi-variate scatterplots {.unnumbered}
 
@@
 #| message: false
+clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+
+color_scheme_set(rep(clrs[4], 6))
+
 titles <- c("Comprehension",
@@
             "Age",
-            "Phonemes",
+            "Length",
             "Exposure",
-            "Levenshtein",
+            "Cognateness",
             "Age × Exposure",
-            "Age × Levenshtein",
-            "Exposure × Levenshtein",
-            "Age × Exposure × Levenshtein")
+            "Age × Cognateness",
+            "Exposure × Cognateness")
 
-posterior <- as.array(model_fit_4)
-np <- nuts_params(model_fit_4) 
+posterior <- as.array(model_fit_3)
+np <- nuts_params(model_fit_3) 
 mcmc_pairs(posterior,
@@
                     "b_age_std:exposure_std",
-                    "b_age_std:lv_std",
-                    "b_exposure_std:lv_std",
-                    "b_age_std:exposure_std:lv_std"),
+                    "b_exposure_std:lv_std"),
            grid_args = list(subtitles = titles),
+           diag_args = list(fill = clrs[1]),
            off_diag_args = list(size = 0.5,
@@
 
-## Posterior-predictive checks
+## Posterior-predictive checks {.unnumbered}
 
@@
 y_int <- as.integer(as.integer(responses$response))
-color_scheme_set(c(clrs[c(4, 4, 3, 2, 1)], "white"))
+color_scheme_set(c(clrs[c(4, 4, 2, 2, 2)], "white"))
 

### 2023-03-29 manuscript/manuscript.qmd
New version of the manuscript, with revised Discussion

@@
 ---
-title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual lexical acquisition"
+title: "Lexical frequency mediates the facilitation effect of cognateness in bilingual lexical acquisition"
 
 pagetitle: Cross-linguistic similarity and word acquisition
-
-abstract: |
-    A prominent feature of the bilingual mental lexicon is that the activation of representations spreads in a language non-selective fashion.
-An instance of such non-selectivity is embodied by cognateness (i.e., form-similarity between translation equivalents, such as *flower*-*flor* in English and Spanish), which impacts lexical access and processing in bilinguals.
-Recent studies have suggested that cognateness facilitates vocabulary acquisition in bilingual toddlers---who show larger vocabulary sizes when their languages share many cognates---and an earlier age-of-acquisition for cognates than for non-cognates.
-The specific mechanisms underpinning such facilitation are unclear.
-We present an account of bilingual early lexical acquisition in which cognateness interacts with lexical frequency and language exposure to facilitate the acquisition of low-frequency words.
-We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals aged 12 to 34 months.
-We used a Bayesian Exploratory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
-We found converging evidence for an earlier age-of-acquisition for cognate words, and for such effect being mediated by lexical frequency and language exposure.
-Low-frequency words, and words from the language of least exposure were more strongly benefitted by their cognate status than high-frequency words.
-Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
-
-thanks: |
-    This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
-Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
-The authors declare no conflicts of interest with regard to the funding source of this study.
-We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and Ege E.
-Özer for their help-ful feedback.
-We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support.
-We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
-We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises.
-We also thank all families and infants who participated in the experiments.
-Data collection was half-way when the COVID-19 pandemic started.
-We would like to pay special tribute to the families that collaborated with us under these difficult circumstances.
 
@@
 
-tar_load(model_fit_4)
+tar_load(model_fit_3)
 tar_load(posterior_draws)
@@
 clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-color_scheme_set(rep(clrs[4], 6))
+color_scheme_set(rep(clrs[1], 6))
 options(ggplot2.ordinal.fill = clrs,
@@
 
-# Introduction {#sec-introduction}
+# Introduction[^acknowledgements] {#sec-introduction}
+
+[^acknowledgements]: This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
+Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
+The authors declare no conflicts of interest with regard to the funding source of this study.
+We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, Ege E.
+Özer, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
+We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support.
+We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
+We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises.
+We also thank all families and infants who participated in the experiments.
+Data collection was half-way when the COVID-19 pandemic started.
+We would like to pay special tribute to the families that collaborated with us under these difficult circumstances.
 
 The foundations of word learning are in place early in age.
-From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [@fenson2007macarthurbates, @samuelson2021precision].
+From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [@fenson2007macarthurbates; @samuelson2021precision].
 During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
-Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e.
-vocabulary size, @fenson1994variability; @frank2021variability].
+Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e., vocabulary size, @fenson1994variability; @frank2021variability].
 Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
@@
 Since birth, these infants had been learning English, Cantonese, and Mandarin, respectively.
-The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*).
+The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caregivers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animal sounds (*woof-woof*).
 These results were later extended by @frank2021variability to a more diverse set of languages, and found that such cross-language commonalities are stronger at earlier stages of lexical acquisition.
@@
 Most literature on early word acquisition still relies extensively on data from monolingual children, and is oblivious to substantial proportion of the world population acquiring more than one language since early ages [@grosjean2021extent].
-Previous research about bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
-@hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals did.
+Previous research on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
+@hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did.
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
 Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-These studies rely on bilingual samples whose languages are relatively distant in typology, as it is the case of English and Spanish.
-It is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
-What role could linguistic distance play during early vocabulary growth?
+These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical or the phonological level.
+Little is known about vocabulary acquisition in bilinguals learning a pair of typologically closer languages (e.g., when both are Romance languages).
 
@@
 When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like Spanish and English (Romance and Germanic languages, respectively).
-For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
+For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might be exposed to the words *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
 It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
@@
 For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/kæt/-/\textipa{kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /æ/ by /\textipa{A}/).
-The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{kæt}/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/.
+The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /kæt/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/.
 Levenshtein distance scores are then divided by the length of the longest string to normalise the similarity scores between 0 and 1.
@@
 Finally, this variable was included as a predictor of participants' vocabulary sizes.
-Among other findings, Floccia and co-workers reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
+Among other findings, Floccia and co-workers reported an increase in vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
 For example, English-Dutch bilinguals (22.14% mean cognateness), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% mean cognateness) were able to produce in Mandarin.
@@
 pointed to *parallel activation* as the underpinning mechanism behind their results.
-The parallel activation hypothesis stems from the language non-selective account of lexical access, and suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain], even in fully monolingual environments.
+The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain], even in fully monolingual environments.
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 
-Regarding word acquisition, the parallel activation account is a plausible account for Floccia et al.'s results in the light of (1) the evidence pointing to bilingual toddlers' lexicon consisting of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage], and that (2) there is experimental evidence of parallel activation during word comprehension in children and toddlers [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
-It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning languages with a larger proportion of cognates (such as children learning English and Dutch) showing larger vocabulary sizes.
+Regarding word acquisition, the parallel activation account is a plausible explanation for Floccia et al.'s results in the light of two observations.
+First, bilingual toddlers' lexicon seems to consist of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage].
+Second, there is experimental evidence of parallel activation during word comprehension in children and toddlers [@vonholzen2012language; @poulin-dubois2013lexical; @bosma2020cognate].
+It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition.
+This could ultimately lead to children who are learning languages with a larger proportion of cognates (e.g., English and Dutch) showing larger vocabulary sizes.
 This account in which cognateness underlies the facilitation effect of typological language distance is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
-If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity but not phonological similarity.
+If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity (e.g., are taxonomically related) but not phonological similarity.
 
@@
 On the other hand, Floccia et al.'s estimates are statistically more reliable given their remarkably larger sample size (372 participants).
-Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level.* The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
-By aggregating the responses from all words into a single datum per child, information about the acquisition status of cognates vs.
-non-cognates was no longer available.
-Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
+Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level*.
+The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
+By aggregating the responses from all words into a single datum per child, it was no longer possible test the effect of cognateness on word acquisition at the word-form level.
+Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as the response variable, age-sensitive claims about the effect of cognateness on word acquisition should be taken with caution.
 
@@
 
-In the present investigation we approach the cognate facilitation on word learning using the conceptual framework of accumulator models of language acquisition [@hidaka2013computational; @mollica2017how].
-We build on the assumption that word acquisition is a function of mainly---but not exclusively---the amount of experience a child has accrued with the corresponding word-referent pairing [see @kachergis2022standard for review].
-As infants accumulate linguistic input, they also accumulate opportunities to associate familiar word-forms to their corresponding referents.
-Words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'headphones' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does].
-Infants might accumulate learning experiences with 'flower' faster than with 'taxi'.
-In the case of bilinguals, the amount of exposures to a word-form an infants receives is also mediated by their degree of exposure to their languages.
+In the present investigation we approach the cognate facilitation on word learning by extending the conceptual framework of accumulator models of language acquisition to the bilingual case.
+Accumulator models formalise the widespread assumption that a child's vocabulary growth rate is a function of mainly---but not exclusively---the number of learning instances that a child has accrued with a particular word [see @kachergis2022standard for review].
+A learning instance of a word consists, for example, in hearing such word-form (*flower*) in the presence of its referent (a flower).
+As infants receive linguistic input, they also accumulate opportunities to learn words until a threshold is crossed, an the word can be considered to be acquired.
+In line with this account, the frequency with which a child encounters a particular word-form in their speech input is a critical determiner of the age at which such word will be acquired.
+The more frequently [@hidaka2013computational; @mollica2017how].
+For instance, words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'taxi' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does]: infants might accumulate learning experiences with 'flower' faster than with 'taxi'.
+
+Previous accounts on vocabulary growth based on accumulator models are a convenient framework to study bilingual vocabulary growth, as bilinguals share fundamentally similar learning mechanisms with monolinguals.
+Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilingual].
+One of such divergences is the fact that bilinguals' frequency of exposure to a word-form is mediated by their degree of exposure to their languages.
 Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual would (who listens 100% of the time to English).
@@
 We suggest that the cognateness effect proposed by @floccia2018results and @mitchell2022cognates facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired.
-We hypothesise that both forms in the translation equivalent (e.g., *flower* and *flor*, for an English-Spanish bilingual) follow their acquisition trajectory in a similar way they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to.
-When one of the word-forms is sufficiently acquired, the acquisition trajectory of the other is impacted, so that the number of needed learning instances left for acquisition are reduced.
+We hypothesise that both forms in the translation equivalent (e.g., *flower* and *flor*, for an English-Spanish bilingual) follow their acquisition trajectory as they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to.
+When one of the word-forms is sufficiently acquired, the number of learning instances left needed for the acquisition of the other word-form is reduced.
 This facilitation would occur as the result of the semantic and phonological overlap between both words.
@@
 ![Schematic representation of the hypothesised cognate facilitation effect on word acquisition.
-To the left, the acquisition of a cognate translation equivalent (*flower*-*flor*, which share four phonemes) is affected by the phonological similarity between both word forms.
-When the child hears *flower*, activation spreads across languages to activate *flor* too.
-As a result, after exposure to *flower* not only the association between *flower* and its referent is strengthened, but also that of *flower* and its common referent.
-To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*, which do not share any phoneme) does not undergo the such facilitation.
-Both word-forms are fairly dissimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](img/hypotheses.jpg){#fig-hypotheses}
+We show a theoretical account of the acquisition of a Catalan-Spanish cognate translation equivalent (*gat*-*gato*, *cat*, whose word-forms share three phonemes) and of a non-cognate translation equivalent (*gos*-*perro*, *dog*, whose word-forms only share one phoneme).
+We assume thecase of a Catalan-Spanish bilingual child who is exposed 65% of the time to Catalan, and 35% of the time to Spanish.
+On the top panels, we illustrate the cumulative sum of learning instances (Y-axis) that a child accumulates across ages (X-axis) for the Catalan word-form (/gat/ or /gos/) and its Spanish translation (/perro/ or /gato/).
+We assume an arbitrary threshold of learning instances at which the child is considered to have acquired a word (solid grey line).
+Points and triangles show the age-of-acquisition at which the amount of learning instances reaches the threshold.
+We show the predicted age-of-acquisition under two hyothesis: (1) a cognateness facilitation effect that lowers the acquisition threshold after one of the word-forms (in this case Catalan /gat/) has been acquired, and (2) no cognateness facilitation, for which the acquisition of the second word-form occurs unconditionally of the acquisition of the first.
+On the bottom, we should the predicted probability of acquisition for the Catalan and Spanish words under each hypothesis, as it would be expected to observe in a vocabulary questionnaire like the one in the present study.
+We assume that such probability starts at 0% at birth, reaches 50% (and grows the fastest) at the age of acquisition, and reaches 100% at some points in later ages.](img/diagram.png){#fig-hypotheses}
 
@@
 Therefore, participants may have on average learned the English word-form of translation equivalents earlier than the word-form in the additional language.
-Under this hypothesis, because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
+Because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants might then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
 
-We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a words *only* once its translation has been already acquired.
+We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a word *only* once its translation has been already acquired.
 We collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
 Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees.
-We adopted a Bayesian explanatory item response theory  [*IRT*, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
+We adopted a Bayesian explanatory item response theory  [IRT, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
 
@@
 We defined *cognateness* as a continuous measure of phonological similarity between translation equivalents.
-We quantified the phonological similarity between each pair of word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonological transcriptions [@levenshtein1966binary; @wells1995computer].
+We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonological transcriptions [@levenshtein1966binary; @wells1995computer].
 This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
 This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]`) and *pájaro* (`[paxa4o]`), Catalan and Spanish for 'bird', and returns 50% for *lluna* (`[Lun5]`) and *luna* (`[luna]`), Catalan and Spanish for 'moon' (see Appendix 1 for more details).
+For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]` and *pájaro* `[paxa4o]`, Catalan and Spanish for 'bird', and returns 50% for *lluna* `[Lun5]` and *luna* `[luna]`, Catalan and Spanish for 'moon' (see Appendix 1 for more details).
 
@@
 The resulting reduced vocabulary checklists contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words.
-Participants were randomly allocated into one of four subset.
+Participants were randomly allocated into one of the four subsets.
 Each response to the questionnaire provided data for a minimum of `r min(n_items_version$catalan, n_items_version$spanish)` and a maximum of `r max(n_items_version$catalan, n_items_version$spanish)` translation equivalents.
@@
     sub_missing(everything(), everything(), missing_text = "--") |>
-    cols_label(doe_catalan = "DoE Catalan") |>
+    cols_label(doe_catalan = "Catalan exposure") |>
     tab_spanner("Age (months)", 2:8) |>
@@
 
-We used the highest self-reported educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES)This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: sense escolaritzar/sin escolarizar [no education], educació primària/educación primaria [primary school], educació secundària/educación secundaria [secondary school], batxillerat/bachillerato [complementary studies/high school], cicles formatius/ciclos formativos [vocational training], and educació universitària/educación universitaria [university degree].
+We used the highest self-reported educational attainment of parents or caregivers as a proxy of participants' socio-economic status (SES).
+This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: 'sense escolaritzar/sin escolarizar' [no education], 'educació primària/educación primaria' [primary school], 'educació secundària/educación secundaria' [secondary school], 'batxillerat/bachillerato' [complementary studies/high school], 'cicles formatius/ciclos formativos' [vocational training], and 'educació universitària/educación universitaria' [university degree].
 Most families reported university studies (`r participants_edu["University"]`, `r percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
@@
 This resulted in a final list of `r sum(n_items_final)` items, corresponding to words `r n_items_final["Catalan"]` Catalan words and their `r n_items_final["Spanish"]` Spanish translations (`r n_te` translation equivalents).
-After collecting participants responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item.
+After collecting participants' responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item.
 Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
 
-We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_4$fit)[2]` sampling chains with `r format(dim(model_fit_4$fit)[1], big.mark = ",")` iterations each).
-This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$ (number of months elapsed between participants' birth date and questionnaire completion), $Length$ (number of phonemes in the X-SAMPA
-phonological transcription of the word-form), $Exposure$ (composite measure of lexical frequency and language degree of exposure), and $Levenshtein$ (phonological similarity between translation equivalents).
+We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_3$fit)[2]` sampling chains with `r format(dim(model_fit_3$fit)[1], big.mark = ",")` iterations each).
+This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $\text{Age}$ (number of months elapsed between participants' birth date and questionnaire completion), $\text{Length}$ (number of phonemes in the X-SAMPA
+phonological transcription of the word-form), $\text{Exposure}$ (composite measure of lexical frequency and language degree of exposure), and $\text{Cognateness}$ (a measure of cognateness, defined as the phonological similarity between translation equivalents).
 For an easier interpretation of the regression coefficients, the values of all predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
-We also included participant and item random intercepts and slopes when appropriate [@barr2013random]
+We also included participant and item random intercepts and slopes when appropriate [@barr2013random].
 
-We compared a baseline model $\mathcal{M}_{0}$, which only included the intercept, $Age$, $Length$, and $Exposure$ as predictors, with increasingly more complex models, introducing predictors in the following order: $Age \times Exposure$ ($\mathcal{M}_{1}$), $Levenshtein$ ($\mathcal{M}_{2}$), $Exposure \times Levenshtein$ ($\mathcal{M}_{3}$), and finally $Age \times Levenshtein$ and $Age \times Exposure \times Levenshtein$ ($\mathcal{M}_{4}$).
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO, @vehtari2017practical]^[Due to computational constraints, we performed LOO based on a random sub-sample of `r format(dim(model_log_liks$model_fit_4)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]], and selected for interpretation the one showing the best performance.
-More details about the models can be found in Appendix 3.
+We compared a baseline model $\mathcal{M}_{0}$---which only included the intercept and the main effects of $\text{Age}$, $\text{Length}$, and $\text{Exposure}$ as predictors---with increasingly more complex models, introducing predictors in the following order: $\text{Age} \times \text{Exposure}$ ($\mathcal{M}_{1}$), $\text{Cognateness}$ ($\mathcal{M}_{2}$), $\text{Exposure} \times \text{Cognateness}$ ($\mathcal{M}_{3}$), and finally $\text{Age} \times \text{Cognateness}$ and $\text{Age} \times \text{Exposure} \times \text{Cognateness}$ ($\mathcal{M}_{4}$).
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO-CV, @vehtari2017practical]^[Due to computational constraints, we performed LOO-CV based on a random sub-sample of `r format(dim(model_log_liks$model_fit_3)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]].
+LOO-CV involves computing the likelihood of each data point, given a model  fitted excluding such data point.
+The likelihoods generated by the model for each of the data points in the data set are summarised together in a measure of predictive performance: the expected log-predicted density (ELPD).
+This is measure penalises the complexity of the model ELPD values closer to zero indicate better predictive accuracy.
+We compared the ELPD of each model, and selected for interpretation the one showing the best performance (see Appendix 3 for more details about the models).
 
 We assessed the practical relevance of the estimated regression coefficients of the selected model following @kruschke2018bayesian.
-First, we specified a region of practical equivalence (*ROPE*) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`.
+First, we specified a region of practical equivalence () from `r rope_interval["lower"]` to +`r rope_interval["upper"]`, in the probability scale.
 This region indicates the range of values that we considered equivalent to zero.
-The degree of overlap between the posterior distribution of a regression coefficient with the *ROPE* can be interpreted as support for the true parameter of the coefficient being zero or equivalent.
-Then we computed the 95% highest density interval (*HDI*) of the posterior distribution of each regression coefficient.
-This is interval contains the true value of this coefficient with 95% probability, given the data.
-Finally, we calculated the proportion of posterior samples in the 95% *HDI* that fell into the *ROPE*, which indicates the probability that the true value of the regression coefficient falls into the *ROPE*, and should therefore be considered equivalent to zero.
-For example, a 90% overlap between the *HDI* of a coefficient's posterior distribution and the *ROPE* indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero.
+The degree of overlap between the posterior distribution of a regression coefficient with the  can be interpreted as support for the true parameter of the coefficient being zero or equivalent.
+We then computed the 95% highest density interval (HDI) of the posterior distribution of each regression coefficient.
+This interval contains the true value of this coefficient with 95% probability, given the data.
+Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the , which indicates the probability that the true value of the regression coefficient falls into the , and should therefore be considered equivalent to zero.
+For example, a 90% overlap between the HDI of a coefficient's posterior distribution and the  indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the , and is therefore equivalent to zero.
 
-Data processing and visualisation was done in R [@rcoreteam2013language, `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
+Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
 Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@vehtari2022loo], `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
@@
 All models showed adequate convergence diagnostics and little evidence of multicollinearity between predictors (see Appendix 4 for details).
-The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $Exposure$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the second best-performing model ($\mathcal{M}_{3}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`.
-Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see @tbl-results-loos for all model comparisons).
+$\mathcal{M}_{3}$, which did *not* include the $\text{Age} \times \text{Exposure} \times \text{Cognateness}$ three-way interaction, showed the best predictive performance with a difference in expected log-predicted density (ELPD) compared to the second best-performing model ($\mathcal{M}_{4}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`.
+Given our data, the predictions of the $\mathcal{M}_{3}$ are confidently more accurate than those of the other models (see @tbl-results-loos for all model comparisons).
+Both $\mathcal{M}_3$ and $\mathcal{M}_4$ performed considerably better than the rest of the models, which did not include the $\text{Exposure} \times \text{Cognateness}$ two-way interaction.
+This points to the effect of $\text{Cognateness}$ being strongly modulated by $\text{Exposure}$, and that this interaction remains relatively stable across ages.
 
@@
 #| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
-The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
-The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
+We report the ELPD of each model along their associated standard errors (*SE*).
+The last column indicates the difference in ELPD between each model and the model with the best predictive accuracy (Model 3)."
 model_loos |>
@@
     rownames_to_column("model") |>
-    mutate(model = str_remove(model, "model_fit_")) |>
-    relocate(model,
-             matches("elpd_loo"),
-             matches("diff"),
-             matches("p_loo"),
-             matches("looic")) |>
-    gt() |>
-    fmt_number(2:9, decimals = 1) |>
+    mutate(model = paste0("Model ", str_remove(model, "model_fit_"))) |>
+    select(model, elpd_loo, se_elpd_loo, elpd_diff, se_diff) |> 
+    gt(rowname_col = "model") |>
+    fmt_number(2:5, decimals = 2, sep_mark = ",") |>
     cols_merge_uncert(elpd_loo, se_elpd_loo) |> 
-    cols_merge_uncert(p_loo, se_p_loo) |> 
-    cols_merge_uncert(looic, se_looic) |> 
     cols_merge_uncert(elpd_diff, se_diff) |> 
@@
         elpd_loo = "ELPD ± SE",
-        p_loo = "p ± SE",
-        looic = "IC ± SE",
-        elpd_diff = "ELPD diff. ± SE"
+        elpd_diff = "Diff. ± SE"
     ) |>
     tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
-    tab_style(cell_text(weight = "bold"),
+    tab_style(cell_text(weight = "bold", style = "italic"),
               cells_column_labels()) |>
@@
 
-@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$, and their degree of overlap with the ROPE.
+@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{3}$, and their degree of overlap with the ROPE.
 For interpretability, we report the estimated regression coefficients transformed to the probability scale.
@@
 #| label: tbl-coefs
-#| tbl-cap: "Posterior distribution of regression coefficients in the probability scale.
+#| tbl-cap: "Posterior distribution of regression coefficients Model 3.
 *Median*: median of the posterior distribution in the logit scale.
-*95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
+*95% HDI*: 95% highest density interval (HDI) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
 On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE).
@@
         .median = "Median",
-        .lower = md("95% *HDI*"),
-        .rope_overlap = md("*ROPE* prob.")
+        .lower = md("95% HDI"),
+        .rope_overlap = md(" prob.")
     ) |>
@@
 
-The word exposure index ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$.upper, accuracy = 0.01)`]).
-All of the posterior samples of this regression coefficient excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), accuracy = 0.01)` more likely to acquire it.
+The main effect of $\text{Age}$ showed the strongest association with the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_age_std"]]$.median/4, accuracy = 0.01)`, 95% HDI = [`r percent(post_draws_list[["b_age_std"]]$.lower/4, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std"]]$.upper/4, accuracy = 0.01)`]), with all of its posterior samples falling out of the ROPE.
+A one-month increment in age increased a maximum of `r percent(post_draws_list[["b_age_std"]]$.median/sd(responses$age)/4, accuracy = 0.01)` the probability of acquisition.
+Similarly, the word exposure index ($\text{Exposure}$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$.median/4, accuracy = 0.01)`, 95% HDI = [`r percent(post_draws_list[["b_exposure_std"]]$.lower/4, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$.upper/4, accuracy = 0.01)`]).
+All of the posterior samples of this regression coefficient excluded the ROPE.
+The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure)/4, accuracy = 0.01)` more likely to acquire it.
+Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$.median/4, accuracy = 0.01)`, 95% HDI = [`r percent(post_draws_list[["b_n_phon_std"]]$.lower/4, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$.upper/4, accuracy = 0.01)`]).
+For every phoneme in the word-form, participants were `r percent((post_draws_list[["b_n_phon_std"]]$.median/4)/sd(responses$n_phon), accuracy = 0.01)` less likely to know it.
+The 95% HDI of the regression coefficient of the $\text{Age} \times \text{Exposure}$ interaction did also excluded the ROPE ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median/4, accuracy = 0.01)`, 95% HDI = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower/4, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper/4, accuracy = 0.01)`]), showing that the effect of the word exposure index differed across ages: older children were more sensitive likely to acquire words with higher exposre rate than younger children.
 
-The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$.rope_overlap, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
-
-However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to word.
-Follow-up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to word was low (e.g., -1 SD), phonological similarity increased the probability of acquisition substantially.
-This effect was negligible when for words with median or high exposure (+1 SD).
-
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of the word exposure index on the speed at which participants acquired words across ages was inconclusive: given our data, it is difficult to tell whether $Exposure$ increased or decreased the slope of the words' acquisition trajectories.
-
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$.upper, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the impact of the word exposure index on the effect of phonological similarity did not differ substantially across ages.
+The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median/4, accuracy = 0.01)`, 95% HDI = [`r percent(post_draws_list[["b_lv_std"]]$.lower/4, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper/4, accuracy = 0.01)`]).
+For every 10% increment in cognateness, the acquisition of a word increased in `r percent(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), accuracy = 0.01)`.
+The effect of $\text{Cognateness}$ interacted with that of $\text{Exposure}$: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.median/4, accuracy = 0.01)`, 95% HDI = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$.lower/4, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.upper/4, accuracy = 0.01)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
+Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
+This effect was negligible when for words with median or high exposure (+1 SD) (see @fig-results-marginal).
 
@@
 #| fig-width: 7
-#| fig-height: 5
+#| fig-height: 4
 #| eval: true
-#| fig-cap: "Expected posterior predictions.
-Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
-These values correspond to samples of the posterior distribution that were drawn during the MCMC estimation of the model.
-We generated 25 posterior predictions for a series of combination of levels of interest.
-Each thin line and corresponds to one prediction for one set of predictor values.
-Thicker lines indicate the median posterior prediction.
-The X-axis indicates the age (in months) for which the prediction is generated.
-The Y-axis indicates the predicted probability of acquisition.
-Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
+#| fig-cap: "Posterior marginal effects of Model 3.
+Thin lines correspond to 50 mean predictions generated from 50 MCMC samples of the posterior distribution of the model.
+Thicker lines indicate the median of those predictions.
+Different colours indicate different levels of phonological similarity (as indicated by the Cognateness similarity between pairs of translation equivalents).
 Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
@@
             levels = unique(lv_std),
-            labels = paste0(c(0, 50, 100), "% similarity"),
+            labels = paste0(c(0, 50, 100), "%"),
             ordered = TRUE
@@
     facet_grid(.category~exposure) +
-    # stat_summary(fun.data = median_hdi, geom = "ribbon",
-    #              alpha = 0.5, colour = NA) +
     annotate(geom = "rect",
@@
              y = 1,
-             hjust = -0.075,
+             hjust = 1,
              vjust = 1,
              label = "Observed\nage range",
-             colour = "grey40",
+             colour = "grey30",
              size = 2) +
-    geom_vline(xintercept = mean(responses$age), colour = "grey",
-               linewidth = 1) +
+    geom_vline(xintercept = mean(responses$age), 
+               colour = "grey",
+               linewidth = 0.75) +
     geom_hline(yintercept = 0.5,
                colour = "grey",
-               linewidth = 1) +
+               linewidth = 0.75) +
     geom_line(aes(group = interaction(lv, .draw)),
-              linewidth = 0.75, alpha = 0.15) +
-    stat_summary(fun = median, geom = "line", linewidth = 1) +
+              linewidth = 0.5,
+              alpha = 1/20) +
+    stat_summary(fun = median, geom = "line", linewidth = 0.75) +
     labs(
@@
         y = "Probability of acquisition",
-        linetype = "Phonological similarity (Levenshtein)",
-        colour = "Phonological similarity (Levenshtein)",
-        fill = "Phonological similarity (Levenshtein)",
-        subtitle = glue("**Phonological similarity**<br> <span style = 'color:{clrs_reorder[5]};'>0% similarity</span>   <span style = 'color:{clrs_reorder[4]};'>50% similarity</span>   <span style = 'color:{clrs_reorder[1]};'>100% similarity</span>")) +
-    scale_colour_manual(values = clrs_reorder[c(5, 4, 1)]) +
-    scale_fill_manual(values = clrs_reorder[c(5, 4, 1)]) +
-    scale_y_continuous(
-        breaks = seq(0, 1, 0.25),
+        linetype = "Cognateness (phonological similarity)",
+        colour = "Cognateness (phonological similarity)",
+        fill = "Cognateness (phonological similarity)") +
+    scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_fill_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_y_continuous(breaks = seq(0, 1, 0.25),
                        limits = c(0, 1),
-        labels = percent
-    ) +
-    scale_x_continuous(breaks = seq(5, 50, 5),
-                       limits = c(10, 50)) +
+                       labels = percent) +
+    scale_x_continuous(breaks = seq(0, 50, 5),
+                       limits = c(0, 50)) +
     theme(
-        legend.position = "none",
+        legend.position = "top",
+        legend.key.size = unit(1, "cm"),
+        legend.key.height = unit(1, "cm"),
+        axis.text.x = element_text(size = 9),
         panel.grid = element_blank(),
@@
             margin = margin(0, 0, 10, 0)
-        ))
+        )) +
+    guides(colour = guide_legend(override.aes = list(linewidth = 2)))
 ```
@@
 
-In this study, we investigated the role of cognateness on word acquisition.
-We contrasted two hypotheses regarding the word acquisition process at which the possible (facilitative) effect of cognateness might take place.
-First, we considered the possibility that cognateness boosts acquisition at early stages, before a stable, robust representation of either of the word-forms involved in the translation equivalent has been formed in the child's lexicon.
-This account predicts that the effect of cognateness would take place early in age, and would be symmetric across languages: one word-form boosts the acquisition of the other via their phonological similarity.
-If this hypotheses were supported by data, we would expect to observe and early difference in probability of comprehension and/or production between more similar and less similar translation equivalents (i.e., an effect of $Levenshtein$ in interaction with $Age$), with this difference being equivalent for the word-for to which the child is exposed to more frequently, and for the word-form to which the while is exposed less frequently (no interaction effect between $Exposure$ and $Levenshtein$).
+In the present study, we explored the cognateness facilitation effect on bilingual word acquisition, and found such effect to be mediated by lexical frequency and degree of language exposure.
+Phonologically similar translation equivalents were on average acquired earlier in age compared to phonologically dissimilar translation equivalents, but this effect was only observed for word-forms to which participants were exposed less often, i.e., words with low lexical frequency and/or words from the language of lower daily exposure.
+Our results are relevant in the light of two previous findings.
+First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (which shared many cognates, e.g., English-German) showed larger vocabulary sizes than those learning typologically more distance languages (which shared fewer cognates, e.g.,  English-Mandarin).
+Second, @mitchell2022cognates found an earlier age-of-acquisition for cognates, compared to non-cognates.
+The outcomes of both studies pointed to cognateness facilitating word acquisition, but the underpinnings of such effect were unclear.
 
-We also considered an alternative hypothesis: that the facilitation effect of cognateness takes place at later stages of word acquisition, when at least one of the word-forms of the translation equivalent hold a detailed, robust representation in the child's lexicon.
-Only then, the phonological similarity between the acquired word-form and its yet-to-be-acquired translation would boost the acquisition of the latter.
-This account predicts a larger effect of cognateness for word-forms to which the child is exposed to less frequently, and therefore are likely to be acquired later than their translation.
-This effect of cognateness would also be expected to take place at later ages.
-Data backing this account would support a model in which the effect of $Levenshtein$ interacts with $Exposure$, and in which possibly this interaction, or at least the effect of $Levenshtein$ interacts with age.
+In this study, we capitalised on accumulator models of language acquisition to propose a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
+We suggested that words benefit from their cognate status only *after* their translation equivalent has been acquired: after the child acquires a word-form, the number of learning instances they need to acquire its translation decreases.
+We further hypothesised that such facilitation is proportional to the phonological similarity between both word-forms.
+If two translation equivalents sharing a fair deal of phonological similarity (e.g., /\textipa{"pOR.t@}/-/\textipa{"pweR.ta}/, `r percent(stringdist::stringsim("pORt@", "pweRta"))` Levenshtein similarity), learning of one of them should facilitate the acquisition of the other more strongly, compared to when both word-forms share only few phonemes (e.g., /\textipa{k@"Di.R@}/-/\textipa{"si.\textlambda a}/ , `r percent(stringdist::stringsim("DiR@", "siLa"))`).
 
-To test these hypotheses, we used an *ad-hoc* vocabulary checklist to collect parental reports of word acquisition status from children learning Catalan and/or Spanish, who were exposed to varying degrees of exposure to either language.
-We then used item response theory (*IRT*) to estimate the probability of each word-form being acquired by a child---either as comprehension or production---as a function of the child's $Age$, their $Exposure$ to the word-form (operationalised as a composite measure of lexical frequency and language exposure), and the phonological similarity of the word-form with its translation in the other language ($Levenshtein$).
-We also added the $Length$  of the word-form as a nuance predictor the increase the predictive performance of the model.
+Under such scenario, we predicted that words from the language of lower exposure would benefit more often from their cognate status than those from the language of higher exposure.
+The rationale behind such prediction was that children are less frequently exposed to word-forms in the former language than to words in the latter [@cattani2014how; @thordardottir2006bilingual].
+Words in the language of lower exposure are therefore more likely to be acquired later in age than their translations in the other language.
+For instance, an infant exposed to Catalan and Spanish 65% and 35% of the time, respectively, is more likely to learn the Spanish word-form /\textipa{"ga.to}/ later in age than its Catalan translation /\textipa{gat}/.
+If this is the case, /\textipa{"ga.to}/ is more likely to receive facilitation from the acquisition of /\textipa{gat}/ than *vice versa* (see @fig-hypotheses for a graphic illustration).
+In line with these predictions, we found that cognateness facilitated word acquisition more strongly in words from the language of lower exposure.
+This asymmetric facilitation effect provides support for the hypothesis that words benefit from their cognate status *after* their translation has been acquired.
+Overall, our results provide support for an interactive account of bilingual vocabulary learning in which word representations from one language contribute to the acquisition of words in the other.
 
-Our results are, to a large extent, in line with previous studies.
-As it could be expected, we found that shorter word-forms were acquired at early ages than longer word-forms.
-For instance, $sol$ (Catalan for *sun*) was acquired earlier than *cocodril* (Catalan for *crocodile*).
-Shorter word-forms consist on fewer phonemes and, frequently, fewer syllables.
-This makes them simpler to encode as phonological representations in the lexicon, and therefore easier to process.
-Ultimately, these words are easier to produce, and therefore to be reported as acquired in parental vocabulary checklists.
-Previous studies found word length to be more strongly associated with receptive vocabulary than with expressive vocabulary [e.g., @braginsky2019consistency; @jones2019children].
+These findings are, to a large extent, in line with previous literature on word acquisition.
+First, words-forms with fewer phonemes like /\textipa{sO\textltilde}/ (Catalan for *sun*) were on average more likely to be acquired than word-forms with more phonemes like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*).
+These results converge with previous studies found word length to be more strongly associated with receptive vocabulary than with expressive vocabulary [e.g., @braginsky2019consistency; @jones2019children].
+This facilitation is most likely due to the fact that shorter word-forms are generally simpler to encode as phonological representations in the lexicon than longer word-forms.
+Shorter word-forms are easier to produce, and therefore to be reported as acquired in parental vocabulary checklists.
 
-According to previous literature, we also anticipated that words to which children were more frequently exposed would be acquired earlier [e.g., @braginsky2019consistency].
-In line with accumulator models of word acquisition, the age of acquisition of a word is a function of the child's rate of exposure to the word, which conspires with child- and word-level characteristics to the formation of a representation of the word in the child's lexicon [@mollica2017how; @hidaka2013computational].
-The amount of learning instances that a child is given to learn a word is given, to a large extent, by the lexical frequency of such word in the child's language.
+Our participants were also more likely to know words to which they were more frequently exposed to, replicating previously reported effects of lexical frequency [e.g., @braginsky2019consistency], and bilingual language exposure [e.g., @cattani2014how; @thordardottir2006bilingual].
+In line with accumulator models of word acquisition, the age-of-acquisition of a word is results, to a lager extent, from the child's rate of exposure to such word, which conspires with child- and word-level factors in the process of forming a representation of the word in the child's lexicon [@mollica2017how; @hidaka2013computational].
+The amount of learning instances that a child is given to learn a word is a function of the lexical frequency of such word.
 Words with higher lexical frequency are more likely to be uttered in the presence of the child, while words with lower lexical frequency appear less often.
@@
 On the one hand, our results provide converging evidence for @floccia2018results, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample.
-Most children in their sample were English-dominant, and therefore their exposure to English was higher than to the additional language.
-However, it was in the additional language vocabulary where they found a facilitation effect of cognateness.
-Bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese).
+Most children in their sample were English-dominant, and therefore their exposure to English was higher than their exposure to the additional language.
+However, it was in the vocabulary of the additional language where the authors found a facilitation effect of cognateness: bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese).
 Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018results's sample.
 
-In line with the hypothesis that received more support from our data, we predicted an interaction effect between participants' age, cognateness, and word exposure, since we expected cognateness to play a more important role at later stages of word learning, when at least one of the members of the translation equivalent has been acquired, and therefore its counterpart in the other language can benefit from its cognate status.
-We found evidence supporting that the true value of this interaction is close to zero.
-Finding such an effect would provide further evidence that cognateness benefits word acquisition only after the translation of such word is acquired.
-However, the absence of such interaction does not preclude a more careful endorsement
-of this hypothesis.
-It is possible that, even if the cognateness effect plays a role if the acquisition of the word-form with the least exposure of the translation equivalent, age itself might not modulate the size of this effect: cognateness plays a similar role across ages.
+Overall , our findings parallel cognates' significant role in many models of bilingual language processing.
+The importance of cognates in these model stems in the fact that cognates provide evidence for a language non-selective account of lexical processing, in which bilinguals activate unselected lexical representations in both languages, even in fully monolingual situations [@thierry2007brain; @christoffels2007bilingual].
+For instance, naming pictures whose labels are cognates leads to shorter latencies than when labels are non-cognates, which suggests that phonological similarity across translations facilitates production [@costa2000cognate; @hoshino2008cognate].
+When acting as distractors in naming tasks, cognates exert larger interference effects than non-cognate distractors [@hoshino2011language; @colome2010which].
+Both findings suggest that the non-selected translation of the target in @costa2000cognate, or of the distractor in @colome2010which was activated, and that when such word-form was phonologically related to its translation, the facilitation or interference effect was stronger.
 
-This study has several limitations.
-One of them is the fact that the languages involved are typologically very close.
+Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extend to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
+For instance, @thierry2007brain found neurophysiological evidence of Chinese-English bilinguals activating both languages in a word comprehension task.
+In this experiment, participants were presented with couples of words in English, and were asked to respond if the words were related in meaning.
+Unbeknownst to participants, the translations to Chinese of half of such word pairs shared a character.
+Participants would be presented (visually in a first experiment and auditorily in a second experiment) with an English word pair 'train-ham', which are semantically unrelated, but share a character in their Chinese translations (*火*車 [Huo Che]-*火*腿 [Huo Tui]).
+Bilinguals' brain signal differed across words pairs with and without translation overlap (e.g., 'apple-table', 平果 [Ping Guo]–桌子 [Zhuo Zi]).
+This suggests that participants were sensitive to the form-similarity of the Chinese translation of the word-pairs they were presented, despite the whole experimental task taking place in English.
+
+Bilingual toddlers and children---still in the process to developing a mature lexicon---also show evidence of parallel activation during comprehension [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic].
+Whether---and if so, how---the language non-selectivity shapes the early bilingual lexicon is still unclear.
+Previous work has pointed to an interactive account of lexical acquisition, in which toddlers' lexical representations in one language influence the acquisition of words in the other language.
+The acquisition of translation equivalents embodies a notable divergence between monolingual and bilingual developmental trajectories.
+For instance, monolingual word acquisition benefits from the semantic relatedness between the words the child learns [@fourtassi2020growth; @hills2009longitudinal], but is inhibited during the acquisition of words from the same language with the *equivalent* meaning, i.e.
+synonyms.
+This learning heuristic, called *mutual exclusivity*, is one of many that play a relevant role in models of monolingual word learning [@golinkoff1994early], acting as a constrainer of the hypothesis space for all possible word-referent associations: if a referent already holds a label, any label uttered in the presence of such referent is associated with a different referent [e.g., @markman1990constraints].
+Although this heuristic seems useful in the monolingual case (children are rarely exposed to synonyms), it would come with a huge cost for bilinguals, how are frequently exposed to more than one label for the same referent, one in each language.
+It was initially predicted that bilinguals would follow *mutual exclusivity*, and therefore avoid learning translation equivalents [@volterra1978acquisition].
+
+This prediction has been robustly contested by two main sources of evidence.
+First, experimental work suggests that bilinguals may not apply *mutual exclusivity* during word learning, or at least not as consistently monolinguals do [@byers-heinlein2009monolingual; @byers-heinlein2013lexicon].
+Second, detailed analyses of bilingual children's vocabulary contents shows that bilingual children do acquire translation equivalents from early ages [e.g., @pearson1995crosslanguage].
+In fact, translation equivalents may even be overrepresented in the early bilingual lexicon.
+For instance, @bilson2015semantic found that bilingual toddlers produced a higher proportion of translation equivalents than it would be expected if they learned two independent lexica simultaneously, which suggests that bilingual children preferentially acquired words for which they already knew a label in the other language.
+In the same line, @bosch2014first reported a similar overrepresentation of translation equivalents in Catalan-Spanish bilinguals: a high proportion of acquired Catalan words held an also acquired translation in Spanish.
+Finally, a recent investigation by @tsui2022area presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
+Overall, these findings suggest that bilingual children benefit from a semantic facilitation effect driven by the acquisition of translation equivalents.
+This is a substantial divergence from monolingual trajectories of lexical acquisition.
+Our study provides further evidence for a pivotal role of translation equivalents, showing that, especially in the case of cognates, translation equivalents, are not only part of bilinguals' vocabulary contents, but are also acquired close in age, depending primarily on children's degree of exposure to each language.
+
+The divergent trajectories of bilingual lexical acquisition go beyond semantics.
+As highlighted before, the phonological overlap between the languages being learned seems to impact vocabulary growth in bilinguals.
+Bilinguals learning languages that share many cognates show larger vocabulary sizes than those learning two languages sharing fewer cognates [@floccia2018introduction; @blom2020crosslanguage; @bosma2019longitudinal].
+This finding can be explained to a large extent by the fact that cognates seem to be acquired at earlier ages than non-cognates [@mitchell2022cognates].
+This phonological facilitation effect mimics previous findings in monolinguals, who seem to preferentially learn phonologically similar words.
+For instance, @jones2019children found that English monolingual toddlers show a facilitation effect when acquiring words from dense phonological neighbourhoods: a larger number of similar-sounding words was associated with an increase in likelihood of a word form being part of a toddler's productive vocabulary.
+A network analysis by @fourtassi2020growth found cross-linguistic evidence from monolingual children for a phonological (and semantic) preferential acquisition of new words, in which words with higher phonological relatedness to other words were more likely to be acquired.
+Again, these findings contrast with those in bilingual children in that the facilitation effect or preferential acquisition mechanisms described are based on phonological similarity *without* translation equivalence.
+The amount of similar-sounding words in the child input or their acquired vocabulary predicts a larger probability of acquisition of new words with *different* meaning.
+In the case of bilinguals, there is evidence of such phonological facilitation at the level of translation equivalents, as shown by an earlier acquisition of cognates than non-cognates in English-French [@mitchell2022cognates] and in the Catalan-Spanish bilinguals in the present study.
+Evidence of a cross-language facilitation effect between non-translation equivalents is to date absent.
+
+We consider three alternative explanations for our results.
+Regarding the cognate facilitation effect, it is possible that when hearing cognate translation equivalents, bilingual children accept both word forms as acceptable variations of the *same* word.
+This contrasts with monolinguals' high sensitivity to mispronunciations during word recognition [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological].
+It might be the case that every exposure to either of the word-forms of the translation equivalent is counted as a learning instance for both.
+This would impact the acquisition mechanism in such way that any cognate effect is not explained by a reduction of the number of learning instances needed to acquired a word (after its translation has been acquired), but by a faster increase in cumulative learning instances across ages, leading to a similar outcome: an earlier age-of-acquisition for cognates.
+Supporting such alternative explanation, @ramon-casas2009vowel examined 18-26 months-old Catalan-Spanish bilingual toddlers' sensitivity to mispronunciations in a word-recognition task and found that, contrary to their Catalan monolingual peers, bilinguals seemed to recognise mispronounced words, disregarding vowel replacements involving the Catalan vowel contrast /e/-/\textepsilon/.
+The same population of bilingual infants shows sensitivity to such contrast before 12 months of age in phoneme discrimination tasks [@bosch2003simultaneous]---which do not involve word-referent associations---suggesting that bilingual infants dismiss subtle changes in phonological forms when presented with their referents.
+Using the same procedure, @ramon2010non tested bilingual infants' recognition of mispronunciations involving the /e/-/\textepsilon/ contrast embedded in exclusively *non-cognate* word-forms.
+The materials in this study contrasted with those in @ramon-casas2009vowel in that the latter included almost 60% of cognate words in their stimuli list.
+@ramon2010non hypothesised that bilinguals' frequent exposure to cognates---which often involve vowel replacements between Catalan and Spanish word-forms---may have led bilinguals to be less alert to such changes when occurring within similar word-forms.
+With this new set of stimuli, including non-cognate words only, bilinguals showed monolingual-level sensitivity to mispronunciations involving the /e/-/\textepsilon/ contrast.
+Overall, these results suggest, at face value, that that bilinguals may to some degree encode similar-sounding word-forms with shared referent as into a common lexical representation, especially when the word-forms involved are cognates.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| out-width: 60%
+lv_distr <- items |> 
+    mutate(lv_bin = cut(lv, 
+                        breaks = seq(0, 1, 0.1),
+                        include.lowest = TRUE,
+                        labels = c("0-10%", "10-20%", "20-30%", "30-40%",
+                                   "40-50%", "50-60%", "60-70%", "70-80%",
+                                   "80-90%", "90-100%"))) |>
+    add_count(name = "n_total") |> 
+    count(lv_bin, n_total) |> 
+    mutate(prop = n/n_total)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+E_e <- bvq_data$pool |>
+    mutate(item = gsub("cat_|spa_", "", item),
+           sampa = gsub("\\.", "", sampa)) |>
+    filter(te %in% items$te, 
+           item %in% items$item) |>
+    select(te, language, sampa) |> 
+    pivot_wider(names_from = language,
+                values_from = sampa, 
+                id_cols = te,
+                names_repair = janitor::make_clean_names) |>
+    filter(grepl("E", catalan), 
+           grepl("e", spanish)) |>
+    left_join(distinct(items, te, lv),
+              by = "te")
+```
+
+We identify three main reasons why bilingual children are unlikely to conflate similar sounding translations into the same lexical representations.
+First, when @ramon-casas2009vowel probed Catalan-Spanish bilinguals using the /e/-/i/ contrast---present in both Catalan *and* Spanish---in the same task, bilinguals showed monolingual-level sensitivity to mispronunciations, suggesting that the conflation of variations of the same word-form into a similar lexical representation may be specific to some language-specific contrasts, and perhaps associated with the acoustic distance between the phonemes involved.
+The /e/-/\textepsilon/ contrast in which Catalan-Spanish bilinguals failed to mispronunciations, is particularly problematic this regard: both vowels differ in only one acoustic feature (/e/ is closed-mid, while /\textepsilon/ is open-mid) and is only present in Catalan.
+Our results are unlikely to be explained by the possibility that Catalan-Spanish bilingual toddlers form common lexical representations for cognate words.
+Second, few phoneme substitutions that separate Catalan word-forms from their cognate Spanish translations involve the /e/-/\textepsilon/ contrast.
+Rather, due to vowel reduction, Spanish and Catalan cognates predominantly involve the /a/-/\textschwa/, /e/-/\textschwa/, /\textepsilon/ and /o/-/u/ contrasts [@vigario2009phonetics].
+Only `r percent(nrow(E_e)/length(unique(items$te)))` translation equivalents in our questionnaire involved a /e/-/\textepsilon/ change between the Catalan and the Spanish items, only one of such was a minimal pair: /\textipa{tREn}/-/\textipa{tRen}/ [*train*].
+
+Second, few words in our study differ only in one phoneme substitution.
+Only `r percent(lv_distr$prop[nrow(lv_distr)], accuracy = 0.01)` of the translation equivalents included in our analyses shared between 90 and 100% phonological similarity (measured as the Levenshtein similarity between the X-SAMPA transcriptions of the Catalan and Spanish form), and only `r percent(lv_distr$prop[nrow(lv_distr)-1], accuracy = 0.01)` shared between 80 and 90% similarity.
+This shows that the presence of identical and near-identical cognates in our study is very low, and therefore most translation equivalents differed in more than one phoneme.
+In summary, even if any cognate involved the /a/-/\textepsilon/ substitution (or any phonetically close contrast that bilingual might disregard at the lexical level), such is very likely to have occurred in the context of a sufficiently distinct word-form to be distinguished and treated and relevant at the lexical level.
+Third, and perhaps more critically, bilinguals are able to learn new minimal pairs involving the /e/-/\textepsilon/ contrast, suggesting that their tolerance towards cognate mispronunciations might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@ramon-casas2017minimalpair].
+
+Another alternative explanation of our results is is that the likelihood of two translation equivalents being present in a child's lexicon results not as a function of such phonological facilitation effect, but as a reporting bias by caregivers when filling the questionnaire.
+In  this scenario, parents assume that because the child knows a word in one language, the child must also know the word in the other language, especially when both words are similar-sounding (i.e., cognates).
+Although there is little possibility to test this account, it is worth noting that parents were explicitly instructed *not* to rely on their responses to the words in one languages when responding to words in the other language.
+The potential presence of responder biases is pervasive across many measurement instruments that rely on caregivers' reports of the developmental status of the child.
+The outcomes of such instruments, like the one used in the present study, are susceptible to biases induced by the child themselves (i.e., assigned gender, temperament), their caregivers (e.g., time spent with the child, preconceptions about language acquisition), or even the researchers (e.g., the items included might over-rely on canonical realisations of words), which push against evidence of validity of the questionnaire as an instrument of the developmental status of the child.
+The findings in this investigation should be weighted according to the aforementioned limitations.
+
+Our results should also be taken with caution if generalised to other bilingual populations, given the particular typological similarity and sociolinguistic status of the languages being learned by our participants.
 Catalan and Spanish are both Romance languages, share a considerable amount of cognates, and very often co-exist in the same social spaces (families, schools, working place, etc.).
@@
 
-Another limitation is the predominantly cross-sectional design of our data set.
+Another limitation of the present study is the predominantly cross-sectional design of our data set.
 We collected vocabulary data from children aged 12 to 32 months, a critical period in development in which the foundations of a lexicon get established.
@@
 
-## Conclusions
-
 In summary, our study has replicated previous studies and provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
@@
 
-# Appendix 1: model predictors {.appendix}
+# Appendix 1: model predictors {.appendix .unnumbered}
 
-{{< include _model-predictors.qmd >}}
+{{< include appendix/_model-predictors.qmd >}}
 
@@
 
-# Appendix 2: model details {.appendix}
+# Appendix 2: model details {.appendix .unnumbered}
 
-{{< include _model-details.qmd >}}
+{{< include appendix/_model-details.qmd >}}
 
@@
 
-# Appendix 3: convergence diagnostics {.appendix}
+# Appendix 3: convergence diagnostics {.appendix .unnumbered}
 
-{{< include _model-diagnostics.qmd >}}
+{{< include appendix/_model-diagnostics.qmd >}}
 
@@
 
-# Appendix 4: frequency and language exposure as separate predictors {.appendix}
+# Appendix 4: frequency and language exposure as separate predictors {.appendix .unnumbered}
 
-{{< include _model-doe.qmd >}}
+{{< include appendix/_model-doe.qmd >}}
 

### 2023-03-29 R/models.R

@@ -1,23 +1,21 @@
 
-#' Specify and sample a brms model
+#' Specify and sample a [brms::brm()] model
 #' @param name A character string indicating the name to be assigned to the model
-#' @param ... Arguments to be passed to \code{brms::brm()}
+#' @param ... Arguments to be passed to [brms::brm()]
 fit_model <- function(name, ...) {
     # we run the model in the background as an RStudio job to keep the console free
     # see R/utils.R
-    fit <- brm(
-        ...,
-        iter = 2000,
-        chains = 4,
-        init = 0.5,
-        seed = 888,
-        backend = "cmdstanr",
-        file = glue("results/fits/{name}.rds"),
-        file_refit = "on_change",
-        control = list(adapt_delta = 0.9,
-                       max_treedepth = 15),
-        save_model = glue("stan/{name}.stan")
-    )
+    fit <- brm(...,
+               iter = 2000,
+               chains = 4,
+               init = 0.5,
+               seed = 888,
+               backend = "cmdstanr",
+               file = glue("results/fits/{name}.rds"),
+               file_refit = "on_change",
+               control = list(adapt_delta = 0.9,
+                              max_treedepth = 15),
+               save_model = glue("stan/{name}.stan"))
     
     return(fit)
 }
@@ -38,8 +36,7 @@ get_log_lik <- function(x, ...) {
     
     # pre-allocate vectors
     n_models <- length(x)
-    log_liks <- vector(mode = "list",
-                       length = n_models)
+    log_liks <- vector(mode = "list", length = n_models)
     
     # compute likelihoods
     cli_progress_bar("Computing likelihoods", total = n_models)
@@ -63,8 +60,7 @@ get_log_lik <- function(x, ...) {
 get_loo <- function(x, ...) {
     # pre-allocate vector
     n_models <- length(x)
-    loos <- vector(mode = "list",
-                   length = n_models)
+    loos <- vector(mode = "list", length = n_models)
     
     # compute loo
     cli_progress_bar("Computing LOOs", total = n_models)
@@ -97,26 +93,21 @@ get_posterior_draws <- function(model,
         "b_Intercept[1]" = "Comprehension and Production",
         "b_Intercept[2]" = "Comprehension",
         "b_age_std" = glue("Age (+1 SD, {round(sd(data$age), 2)}, months)"),
[…]

### 2023-03-29 _targets.R

@@ -27,6 +27,7 @@ tar_option_set(
         "conflicted",
         "dplyr",
         "ggplot2",
+        "ggtext",
         "glue",
         "gt",
         "janitor",
@@ -238,19 +239,19 @@ list(
     ## describe models ---------------------------------------------------------
     tar_target(
         posterior_draws,
-        get_posterior_draws(model_fit_4, data = responses)
+        get_posterior_draws(model_fit_3, data = responses)
     ),
     
-    # tar_target(posterior_draws_re,
-    #            get_posterior_draws_re(model_fit_4)),
+    tar_target(posterior_draws_re,
+               get_posterior_draws_re(model_fit_3)),
     
     ## marginal effects --------------------------------------------------------
     tar_target(
         marginal_effects_epreds,
         posterior_predictions(
-            model = model_fit_4,
+            model = model_fit_3,
             responses, 
-            age_std = scale(seq(12, 50),
+            age_std = scale(seq(0, 50),
                             mean(responses$age),
                             sd(responses$age)),
             exposure_std = c(-1, 0, 1),
@@ -337,7 +338,7 @@ list(
     tar_target(
         model_ppcs,
         {
-            yrep_char <- posterior_predict(model_fit_4, ndraws = 50)
+            yrep_char <- posterior_predict(model_fit_3, ndraws = 50)
             sapply(data.frame(yrep_char, stringsAsFactors = TRUE), as.integer)
         }
     ),
@@ -393,11 +394,11 @@ list(
                 "b_n_phon_std" = glue("Phonemes (+1 SD, {round(sd(responses$n_phon), 2)} phonemes)"),
                 "b_doe_std" = glue("DoE (+1 SD, {round(sd(responses$doe_std), 2)})"),
                 "b_freq_std" = glue("Frequency (+1 SD, {round(sd(responses$freq_std), 2)})"),
-                "b_lv_std" = glue("Levenshtein (+1 SD, {percent(sd(responses$lv))})"),
-                "b_doe_std:lv_std" = "Exposure \u00d7 Levenshtein",
+                "b_lv_std" = glue("Cognateness (+1 SD, {percent(sd(responses$lv))})"),
+                "b_doe_std:lv_std" = "Exposure \u00d7 Cognateness",
                 "b_age_std:doe_std" = "Age \u00d7 DoE",
-                "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
[…]

### 2023-03-29 docs/syllable-frequency.qmd
New version of the manuscript, with revised Discussion

---
title: "Syllable frequency by cognateness"
format: html
warning: false
echo: false
code-fold: true
message: false
---

```{r setup}
tar_load_globals()
tar_load(childes)
tar_load(bvq_data)
tar_load(items)

library(dplyr)
library(tidyr)
library(ggplot2)
library(gt)
library(janitor)
library(tidytext)

theme_set(theme_custom()) # set custom ggplot theme

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

duplicated_te <- bvq_data$pool$te[duplicated(bvq_data$pool$te)]

pool_tmp <- bvq_data$pool |>
    # drop items with missing observations in these variables
    drop_na(sampa, wordbank_lemma) |>
    filter(n_lemmas == 1,
           # exclude items with more than two lemmas
           !is_multiword,
           # exclude multi-word items
           include,
           # exclude problematic items (e.g., multi-word items)
           te %in% duplicated_te) |>
    add_count(te, name = "n_te") |> # get only items with one translation in each language
    filter(n_te == 2) |>
    distinct(language, te, .keep_all = TRUE) |> 
    select(language, te, item, label, childes_lemma, ipa, sampa) |> 
    mutate(syll = strsplit(ipa, "\\.|ˈ"),
           syll = map(syll, function(x) x[!(x=="")]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_"))

```

We define syllable frequency as the rate of appearance of individual syllables in the word-forms included in the Barcelona Vocabulary Questionnaire (BVQ). Each item corresponds to a Catalan or Spanish word, and has an associated phonological transcription in International Phonological Alphabet (IPA) format. These transcriptions are syllabified: syllables are separated by the `.` or the `ˈ` (marking that the upcoming syllable after is) symbols. Some examples:

```{r items-table}
#| label: tbl-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and  their syllabified IPA transcriptions in Catalan and Spanish
items |> 
    distinct(te, item, language, meaning, ipa) |> 
    left_join(distinct(pool_tmp, te, item, label)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, ipa),
                id_cols = c(te, meaning),
                names_repair = make_clean_names,
                values_fn = first) |> 
    sample_n(20) |> 
    gt() |> 
    cols_hide(c(te, item_catalan, item_spanish)) |> 
    tab_spanner("Catalan", matches("catalan")) |> 
    tab_spanner("Spanish", matches("spanish")) |> 
    cols_label(meaning = "Translation equivalent",
               label_catalan = "Item",
               ipa_catalan = "IPA",
               label_spanish = "Item",
               ipa_spanish = "IPA") |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels())
```

# Number of syllables

```{r syl-number}
#| label: fig-syl-number
#| fig-cap: Distribution of the number of syllables in Catalan and Spanish
pool_tmp |> 
    group_by(language) |> 
    add_count(language, name = "n_lang") |> 
    ungroup() |> 
    count(language, n_lang, n_syll) |> 
    mutate(prop = n/n_lang) |> 
    ggplot(aes(n_syll, n, fill = language)) +
    facet_wrap(~ language) +
    geom_col(colour = "white") +
    geom_text(aes(label = percent(prop)),
              size = 3,
              position = position_nudge(y = 15),
              colour = "black") +
    labs(x = "Number of syllables",
         y = "Number of words",
         fill = "Language") +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          panel.grid.major.x = element_blank())
```

# Lexical frequencies

We extracted lexical frequencies from the English corpora in the CHILDES database. Using the Catalan and Spanish corpora was not possible due to the low number of children and tokens included in the corpora.

```{r lex-freq}
#| label: fig-lexical-freq
#| fig-cap: "Distribution of lexical frequencies from CHILDES."
freq_df <- pool_tmp |>
    left_join(childes, by = c("childes_lemma" = "token")) |> 
    select(te, language, ipa, freq_million) |> 
    drop_na(freq_million) |> 
    rename(freq_lex = freq_million)

freq_df |> 
    distinct(te, freq_lex) |> 
    ggplot(aes(freq_lex)) +
    geom_histogram(colour = "white",
                   fill = clrs[1],
                   bins = 30) +
    geom_rug(colour = clrs[4],
             alpha = 0.5) +
    labs(x = "Lexical frequency (CHILDES)",
         y = "Counts") +
    scale_x_continuous(trans = "log10",
                       labels = \(x) format(x, big.mark = ","))
```

# Weighting syllable counts by lexical frequency

Every exposure to a word-form also counts as a exposure to each of the syllables that make up such word. Every time a child hears the word *casa* [house], they are exposed to the syllables *ca* and *sa*. Syllables that appear embedded in words with higher lexical frequency will also be more frequent. To compute the relative frequency of each syllable in Catalan and Spanish (i.e., how many times the syllables appears in every million words in Catalan or Spanish speech), we summed the relative lexical frequency in CHILDES of every word that contains such syllable in the corresponding language.

```{r syl-freq}
#| label: fig-syl-counts
#| fig-cap: "Distribution of apositional syllable frequencies in Spanish and Catalan"
#| fig-height: 5
#| fig-width: 9
syll <- items |> 
    left_join(distinct(pool_tmp, te, childes_lemma)) |> 
    left_join(select(childes, token, freq_lex = freq_million),
              by = c("childes_lemma" = "token")) |> 
    select(language, item, te, meaning, ipa, lv, freq_lex) |> 
    mutate(syll = strsplit(ipa, "\\.|ˈ"),
           syll = map(syll, function(x) x[!(x=="")]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_")) |> 
    unnest_longer(syll) |> 
    group_by(language, syll) |> 
    summarise(freq_syll = sum(freq_lex),
              .groups = "drop") |> 
    arrange(language, desc(freq_syll))

syll |> 
    add_count(language, name = "n_lang") |> 
    mutate(language = paste0(language, " (n = ", n_lang, ")")) |> 
    ggplot(aes(freq_syll,
               fill = language, 
               colour = language)) +
    facet_wrap(~language) +
    geom_histogram(colour = "white",
                   position = position_nudge(y = 0.1)) +
    geom_rug(alpha = 1/4) +
    labs(x = "Relative frequency (counts per million)",
         y = "Number of syllables",
         fill = "Language",
         colour = "Language") +
    scale_x_continuous(trans = "log10",
                       labels = \(x) format(x, big.mark = ",")) +
    scale_colour_manual(values = clrs[c(1, 4)]) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(legend.title = element_blank(),
          legend.position = "none")
```

```{r top-syll-freq}
#| label: fig-top-syll-freq
#| fig-cap: "Top 30 most frequent syllables"
syll |> 
    group_by(language) |> 
    top_n(30, freq_syll) |> 
    ggplot(aes(reorder_within(syll, desc(freq_syll), language),
               freq_syll,
               fill = language)) +
    facet_wrap(~ language,
               scales = "free_x",
               ncol = 1) +
    geom_col(colour = "white") +
    labs(x = "Syllable",
         y = "Relative frequency\n(counts per million)",
         colour = "Language",
         fill = "Language") +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    scale_x_reordered() +
    scale_y_continuous(labels = \(x) format(x, big.mark = ",")) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 10),
          panel.grid.major.x = element_blank())
```

# Positional syllable frequency

```{r pos-syll-freq}
#| label: fig-pos-freq
#| fig-cap: "Distribution of positional syllable frequencies, split by language and syllable position (1, 2, or 3)."

syll_pos <- items |> 
    left_join(distinct(pool_tmp, te, childes_lemma)) |> 
    left_join(select(childes, token, freq_lex = freq_million),
              by = c("childes_lemma" = "token")) |> 
    select(language, item, te, meaning, ipa, lv, freq_lex) |> 
    mutate(syll = strsplit(ipa, "\\.|ˈ"),
           syll = map(syll, function(x) x[!(x=="")]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_")) |> 
    unnest_longer(syll, indices_to = "position") |> 
    group_by(language, position, syll) |> 
    summarise(freq_syll_pos = sum(freq_lex),
              .groups = "drop") |> 
    mutate(position = factor(position, ordered = TRUE)) |> 
    filter(position < 4) |> 
    arrange(language, desc(freq_syll_pos))

ggplot(syll_pos, aes(position,
                     freq_syll_pos,
                     fill = language, 
                     colour = language)) +
    facet_wrap(~language) +
    geom_point(alpha = 1/4,
               size = 2,
               shape = 1,
               stroke = 1,
               position = position_jitter(width = 0.1)) +
    stat_summary(fun.data = mean_se,
                 geom = "errorbar",
                 width = 0.2,
                 colour = "black",
                 linewidth = 0.75) +
    stat_summary(fun = mean,
                 geom = "point",
                 size = 1.5,
                 colour = "black") +
    labs(x = "Syllable position",
         y = "Positional relative frequency\n(counts per million)",
         fill = "Language",
         colour = "Language") +
    scale_y_continuous(trans = "log10",
                       labels = \(x) format(x, big.mark = ",")) +
    scale_colour_manual(values = clrs[c(1, 4)]) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(legend.title = element_blank(),
          legend.position = "none",
          panel.grid.major.x = element_blank())
```

```{r top-pos-syll-freq}
#| label: fig-top-pos-syll-freq
#| fig-cap: "Top 20 most frequent syllables by position."
#| fig-width: 13
#| fig-height: 4
syll_pos |> 
    group_by(language, position) |> 
    top_n(20, freq_syll_pos) |> 
    ggplot(aes(reorder_within(syll, 
                              desc(freq_syll_pos),
                              list(language, position)),
               freq_syll_pos,
               fill = language)) +
    facet_wrap(language ~ position,
               scales = "free_x",
               ncol = 3,
               labeller = \(x) label_value(x, multi_line = FALSE)) +
    geom_col(colour = "white") +
    labs(x = "Syllable",
         y = "Relative frequency\n(counts per million)",
         colour = "Language",
         fill = "Language") +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    scale_x_reordered() +
    scale_y_continuous(labels = \(x) format(x, big.mark = ",")) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 8),
          panel.grid.major.x = element_blank())
```

# Are cognates made of more frequent syllables?

```{r cognate-freq}

lv_similarities <- pool_tmp |>
    mutate(sampa = str_remove_all(sampa, "\\.")) |>
    pivot_wider(
        id_cols = te,
        names_from = language,
        values_from = sampa,
        names_repair = make_clean_names
    ) |>
    # make sure strings are coded as UTF-8 before computing LVs
    mutate(lv = stringsim(catalan, spanish)) |>
    distinct(te, lv)

syll_lv <- pool_tmp |> 
    left_join(lv_similarities) |> 
    unnest_longer(syll, indices_to = "position") |> 
    left_join(syll, by = c("syll", "language")) |> 
    group_by(te, lv, n_syll, language) |> 
    summarise(freq_syll_sum = sum(freq_syll),
              .groups = "drop") |> 
    drop_na(freq_syll_sum)

syll_lv |> 
    add_count(language, name = "n_lang") |> 
    mutate(language = paste0(language, " (n = ", n_lang, ")"),
           n_syll = paste0(n_syll, " syllables")) |> 
    ggplot(aes(lv, freq_syll_sum)) +
    facet_grid(language ~ n_syll) +
    geom_point(shape = 1,
               stroke = 1,
               colour = clrs[4],
               size = 2,
               alpha = 1/2) +
    geom_smooth(colour = clrs[1],
                fill = clrs[1]) +
    labs(x = "Levenshtein distance (cognatness)",
         y = "Additive syllable frequency\n(Counts per million)",
         colour = "Language",
         fill = "Language") +
    scale_colour_manual(values = clrs[c(1, 4)]) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    scale_y_continuous(trans = "log10",
                       labels = \(x) format(x,
                                            big.mark = ",",
                                            scientific = FALSE)) +
    scale_x_continuous(labels = percent) +
    theme(panel.grid.minor = element_blank(),
          legend.position = "none",
          axis.text = element_text(size = 9))

```

### 2023-03-29 manuscript/appendix/_model-predictors.qmd
New version of the manuscript, with revised Discussion

--- 2023-02-26 manuscript/_model-predictors.qmd
@@
-## Word exposure index
+## Word exposure index  {.unnumbered}
 
@@
 
-We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
+We created a predictor ($\text{Exposure}$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
 the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).
@@
 $$
-\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
+\text{Exposure}_{ij} = \text{Frequency}_i \times \text{DoE}_{j}
 $$ {#eq-exposure}
@@
 
-## Levenshtein similarity
+## Levenshtein similarity {.unnumbered}
 

### 2023-03-29 R/figures.R

@@ -1,0 +1,269 @@
+library(ggplot2)
+library(patchwork)
+library(tibble)
+library(dplyr)
+library(tidyr)
+library(purrr)
+library(scales)
+library(bvqdev)
+library(glue)
+library(ggtext)
+library(magick)
+
+resolve_conflicts()
+
+# set options ------------------------------------------------------------------
+clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+
+options(ggplot2.ordinal.fill = clrs,
+        ggplot2.ordinal.colour = clrs,
+        ggplot2.discrete.fill = clrs,
+        ggplot2.discrete.colour = clrs,
+        ggplot2.continuous.fill = scale_color_gradient,
+        ggplot2.continuous.colour = scale_color_gradient)
+
+set.seed(888)
+
+# set parameters ---------------------------------------------------------------
+
+generate_aoa <- function(conditions = c("**Cognate**", "**Non-cognate**"),
+                         items = c("gat /gat/",
+                                   "gato /gato/",
+                                   "gos /gos/", 
+                                   "perro /ˈpe.ro/"),
+                         age_step = 0.5,
+                         age_range = c(10, 34),
+                         freq_month = 50,
+                         freq_month_sd = 10,
+                         thres = 500,
+                         doe_l1 = 0.65,
+                         thres_diff = 100) 
+{
+    
+    item_df <- tibble(te = rep(conditions, each = 2),
+                      item = items,
+                      doe = c(L1 = doe_l1, 
+                              Spanish = 1-doe_l1, 
+                              L1 = doe_l1, 
+                              Spanish = 1-doe_l1))
+    
+    eli_df <- expand_grid(item_df,
+                          age = seq(age_range[1], age_range[2], age_step),
+                          n_month = freq_month*age_step,
+                          threshold = thres,
+                          hypothesis = c("No cognateness effect",
+                                         "Cognateness effect")) |> 
+        mutate(n_month_wgt = ceiling(n_month*doe),
+               eli = n_month_wgt + rnorm(n(), 0, freq_month_sd),
+               eli = round(ifelse(eli < 0 , 0, eli), 2)) |> 
+        group_by(item, doe) |> 
+        mutate(eli_cum = cumsum(eli),
[…]

### 2023-03-29 R/responses.R

@@ -1,9 +1,9 @@
 
 #' Prepare data for analyses
 #' 
-#' @param bvq_data A named list resulting from calling \code{get_bvq}
-#' @param items A data frame resulting from calling \code{get_items}
-#' @param participants A data frame resulting from calling \code{get_participants}
+#' @param bvq_data A named list resulting from calling [get_bvq()].
+#' @param items A data frame resulting from calling [get_items()].
+#' @param participants A data frame resulting from calling [get_participants()].
 get_responses <- function(bvq_data, items, participants) {
     # merge all datasets
     responses_tmp <- bvq_data$responses |>
@@ -23,12 +23,12 @@ get_responses <- function(bvq_data, items, participants) {
         reduce(inner_join) |>
         mutate(
             # code responses as factor
-            response = factor(
-                response,
-                levels = c(1, 2, 3),
-                labels = c("No", "Understands", "Understands and Says"),
-                ordered = TRUE
-            ),
+            response = factor(response,
+                              levels = c(1, 2, 3),
+                              labels = c("No",
+                                         "Understands",
+                                         "Understands and Says"),
+                              ordered = TRUE),
             # does should have the value of the corresponding language
             doe = ifelse(language == "Catalan", doe_catalan, doe_spanish),
             # standardise numeric predictors
@@ -41,27 +41,9 @@ get_responses <- function(bvq_data, items, participants) {
             exposure_std = scale(exposure)[, 1],
         ) |>
         # get only relevant variables
-        select(
-            id,
-            time,
-            age,
-            age_std,
-            te,
-            language,
-            meaning,
-            item,
-            response,
-            lv,
-            lv_std,
-            freq,
-            freq_std,
-            n_phon,
-            n_phon_std,
-            doe,
-            doe_std,
-            exposure,
-            exposure_std
-        ) |>
+        select(id, time, age, age_std, te, language, meaning, item, response,
+               lv, lv_std, freq, freq_std, n_phon, n_phon_std, doe, doe_std,
+               exposure, exposure_std) |>
         # reorder rows
         arrange(id, te, language)
     

### 2023-03-29 R/utils.R

@@ -1,5 +1,3 @@
-# utils
-
 # run the targets workflow
 make <- function() {
     job::job({
@@ -8,6 +6,13 @@ make <- function() {
     },
     import = NULL,
     title = "Trajectories")
+    
+    to_delete <- c("manuscript/QTDublinIrish.otf")
+    for(i in to_delete) {
+        if (file.exists(i)) {
+            file.remove(i)
+        }
+    }
 }
 
 # remove targets products
@@ -47,12 +52,11 @@ resolve_conflicts <- function() {
 #' Custom ggplot2 theme
 theme_custom <- function() {
     theme_minimal() +
-        theme(
-            panel.grid = element_line(colour = "grey", linetype = "dotted"),
-            axis.line = element_line(colour = "black"),
-            text = element_text(size = 12, colour = "black"),
-            axis.text = element_text(colour = "black")
-        )
+        theme(panel.grid = element_line(colour = "grey",
+                                        linetype = "dotted"),
+              axis.line = element_line(colour = "black"),
+              text = element_text(size = 12, colour = "black"),
+              axis.text = element_text(colour = "black"))
 }
 
 #' Transform months to years and months
@@ -118,8 +122,7 @@ prop_adj_se <- function(y, n) {
 prop_adj_ci <- function(y, n, .width = 0.95, limit) {
     prop <- (y + 2) / (n + 4)
     se <- sqrt(prop * (1 - prop) / (n + 4))
-    ci <-
-        prop + qnorm(c((1 - .width) / 2, (1 - (1 - .width) / 2))) * se
+    ci <- prop + qnorm(c((1 - .width) / 2, (1 - (1 - .width) / 2))) * se
     ci[1] <- ifelse(ci[1] < 0, 0, ci[1]) # truncate at 0
     ci[2] <- ifelse(ci[2] > 1, 1, ci[2]) # truncate at 1
     
@@ -159,82 +162,82 @@ get_childes_frequencies <- function(collection = "Eng-NA",
     
     suppressMessages({
         
-    roles <- c(
-        "Mother",
-        "Father",
-        "Investigator",
-        "Sibling",
-        "Sister",
-        "Grandmother",
-        "Adult",
-        "Friend",
-        "Brother",
-        "Visitor",
-        "Relative",
-        "Grandfather",
-        "Teacher",
-        "Student"
-    )
-    
-    counts <- get_types(collection = collection, role = roles, ...)
-    
-    speaker_ids <- distinct(counts,
-                            collection_id,
-                            corpus_id,
[…]

### 2023-03-29 manuscript/img/diagram.R

@@ -1,11 +1,102 @@
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(patchwork)
+
+conflict_prefer("filter", "dplyr")
+
+clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+
 logistic <- function(x, slope, mid, upper) {
     upper / (1 + exp(-slope * (x - mid)))
 }
 
 
-x <- seq(6, 30, 0.1)
+
+token <- c("porc", "gat")
+it <- items |>
+    filter(item %in% token) |> 
+    select(item, te, meaning, ipa, lv, freq)
+n_month <- 24
+age <- seq(34)
+doe <- c(0.65, 0.35)
+
+
+elis_df <- it |> 
+    expand_grid(age, doe) |> 
+    mutate(exposure = freq*doe,
+           exposure_std = scale(exposure,
+                                mean(responses$exposure),
+                                sd(responses$exposure))[,1],
+           lambda = n_month/freq,
+           lambda_wgt = lambda * doe,
+           eli = lambda_wgt + runif(n = n(),
+                                    min = -lambda_wgt*0.50,
+                                    max = lambda_wgt*0.50)) |> 
+    group_by(te, item, doe) |> 
+    mutate(eli_cum = cumsum(eli)) |> 
+    arrange(item, doe, age) |> 
+    group_by(te, doe) |> 
+    mutate(aoa = age[which.min(abs(eli_cum-n_month))],
+           prob = logistic(age, 1, aoa, 1)) 
+
+
+
+ggplot(elis_df, aes(age, eli_cum, colour = as.factor(doe))) +
+    facet_wrap(~meaning) +
+    geom_step() +
+    geom_point(data = distinct(elis_df, item, doe, aoa),
+               aes(x = aoa,
+                   y = 0)) +
+    ggplot(elis_df, aes(age, prob, colour = as.factor(doe))) +
+    facet_wrap(~meaning) +
+    geom_hline(yintercept = 0.5) + 
+    geom_line(aes(group = as.factor(doe)),
+                 linewidth = 1) + 
+    geom_point(data = distinct(elis_df, te, meaning, doe, aoa),
+               aes(x = aoa,
+                   y = 0),
+               size = 5) +
+    
+    plot_layout(ncol = 1) &
+    scale_colour_manual(values = clrs[c(1, 4)])
+
+post <- posterior_predictions_re(
+    model = model_fit_4,
[…]

### 2023-04-17 appendix/_model-diagnostics.qmd
Fix comments and use apaquarto extension

We used Stan [@carpenter2017stan] as the probabilistic language behind the estimation of our Bayesian models in this study, with `brms` as its R interface [@burkner2017brms]. This language implements the Markov Chain Monte-Carlo (MCMC) algorithm to explore the posterior distribution of the model. Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined. We run `r dim(model_fit_3$fit)[2]` MCMC chains, each `r dim(model_fit_3$fit)[1]` iterations long each. The correct performance of this algorithm is critical to the quality of the statistical evidence to which the outcomes of the model lead.

One way to diagnose the behaviour of MCMC is to inspect whether the different MCMC chains (if more than one) have converged to a similar region of the posterior. The Gelman-Rubin diagnostic [$\hat{R}$ or R-hat @gelman1992inference] provides a measure of chain convergence by comparing the variance within each chain *versus* the variance between each chain. Both are expected to be identical when chains have perfectly converged, so that $\hat{R} = 1$. Values lower than 1.01 are recommended, while values higher than 1.05 indicate that chains might have trouble converging and therefore the estimated parameters must be taken with caution. {apafg-model-rhats-neffs} (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference. Most values are lower than 1.01, and never higher than 1.05, which provides evidence of successful MCMC convergence.

Another diagnostic of good MCMC converge is the ratio of effective sample size to total sample size ($N_{eff}/N$), which indicates the proportion of samples in the chain that resulted from a non-divergent transition. Values closer to 1 are ideal, as they indicate that all posterior samples from the MCMC were used to estimate the posterior distribution of the parameter. Values larger than 0.1 are recommended. {apafg-model-rhats-neffs} (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models. Most values are larger than 0.1, although model 0 ($\mathcal{M}_0$) accumulates most effective sample sizes close to 0.1.

```{r apafg-model-rhats-neffs}
#| label: apafg-model-rhats-neffs
#| apa-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6
#| fig-width: 9
#| out-width: 100%
map_df(model_rhats, bind_rows, .id = "model") |> 
    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
    pivot_longer(-model, 
                 names_to = ".variable",
                 values_to = ".rhat") |>
    drop_na() |> 
    ggplot(aes(.rhat)) +
    facet_wrap(~model, nrow = 1) +
    geom_vline(xintercept = 1.01, linetype = "dashed") +
    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
    labs(x = "Gelman Rubin diagnostic (R-hat)",
         y = "MCMC samples",
         title = "Gelman-Rubin (R-hat) diagnostic") +

    map_df(model_neffs, bind_rows, .id = "model") |> 
    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
    pivot_longer(-model, 
                 names_to = ".variable",
                 values_to = ".rhat") |>
    # filter(str_detect(.variable, "b_|sd_")) |> 
    drop_na() |> 
    ggplot(aes(.rhat)) +
    facet_wrap(~model, nrow = 1) +
    geom_vline(xintercept = 1.01, linetype = "dashed") +
    geom_histogram(binwidth = 0.1, colour = "white", fill = "grey60") +
    labs(x = "Gelman Rubin diagnostic (R-hat)",
         y = "MCMC samples",
         title = "Ratio of effective sample size") +
    plot_layout(ncol = 1) &
    plot_annotation(tag_levels = "A") &
    theme(legend.position = "none",
          panel.grid = element_blank())
```

## Posterior draws and bi-variate scatterplots {.unnumbered}

```{r}
#| label: fig-model-pairs
#| fig-cap: "Marginal distribution and bi-variate scatterplot of posterior samples for the fixed regression coefficients in Model 3."
#| fig-height: 8
#| fig-width: 10
#| out-width: 100%
#| out-height: 100%
#| echo: false
#| warning: false
#| message: false
color_scheme_set(rep("grey60", 6))
titles <- c("Comprehension",
            "Production",
            "Age",
            "Length",
            "Exposure",
            "Cognateness",
            "Age × Exposure",
            "Age × Cognateness",
            "Exposure × Cognateness")
posterior <- as.array(model_fit_3)
np <- nuts_params(model_fit_3) 
mcmc_pairs(posterior,
           np = np, 
           pars = c("b_Intercept[1]",
                    "b_Intercept[2]",
                    "b_age_std",
                    "b_n_phon_std",
                    "b_exposure_std",
                    "b_lv_std",
                    "b_age_std:exposure_std",
                    "b_exposure_std:lv_std"),
           draws = c(1, 0),
           grid_args = list(titles = titles),
           off_diag_args = list(size = 0.5,
                                alpha = 1/10))
```

## Posterior-predictive checks {.unnumbered}

```{r apafg-model-ppc}
#| label: apafg-model-ppc
#| apa-cap: "Model posterior predictive checks (PPC). Bars indicate the observed proportion of responses to each category (No, Understands, and Understands and Says). Blue dots and error bars represent the mean proportion of responses simulated from the posterior for each category, and its 95\\% interval."
#| eval: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 6
#| fig-height: 3
#| out-width: 50%
#| out-height: 50%
y_int <- as.integer(as.integer(responses$response))
color_scheme_set(c(rep("grey60", 2), rep("black", 4)))

ppc_bars(y_int, 
         model_ppcs,
         prob = 0.95,
         linewidth = 1,
         size = 1.25,
         fatten = 2,
         freq = FALSE) +
     scale_x_continuous(breaks = 1:3,
                        labels = c("No",
                                   "Understands", 
                                   "Understands and Says")) +
    theme(legend.position = "none",
          axis.text = element_text(size = 11))
```

### 2023-04-17 manuscript.qmd
Fix comments and use apaquarto extension

--- 2022-09-23 manuscript.qmd
@@
 ---
-title: "The role of cross-linguistic lexical similarity on bilingual word acquisition"
-editor: source
+title: "Lexical frequency mediates the facilitation effect of cognateness in bilingual lexical acquisition"
+
+shorttitle: Frequency and cognateness in lexical acquisition
+
 author:
@@
     corresponding: true
+    orcid: 0000-0002-8553-4209
+    email: gonzalo.garciadecastro@upf.edu
     affiliations:
-        - name: Center for Brain and Cognition, Universitat Pompeu Fabra
-    - name: Daniela S. Ávila-Varela
+      - id: cbc
+        name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+        address: Ramon Trias Fargas 25-27
+        city: Barcelona
+        region: Spain
+        postal-code: "08005"
+
+  - name: Daniela S. Avila-Varela
+    orcid: 0000-0002-3518-8117
     affiliations: 
-        - name: Center for Brain and Cognition, Universitat Pompeu Fabra
+      - ref: cbc
+      - id: ul
+        name: Universidade de Lisboa
+        department: Facultade de Letras
+        address: Alameda da Universidade
+        city: Lisboa
+        region: Portugal
+        postal-code: "1600-214"
+
   - name: Ignacio Castillejo
+    orcid: 0000-0001-7445-0416
     affiliations: 
-        - name: Departamento de Psicología, Universidad Autónoma de Madrid
+      - id: uam 
+        name: Universidad Autónoma de Madrid
+        department: Departamento de Psicología
+        address: Iván Pavlov, 6
+        city: Madrid
+        region: Spain
+        postal-code: "28049"
+
   - name: Núria Sebastian-Galles
+    orcid: 0000-0001-6938-2498
     affiliations: 
-        - name: Center for Brain and Cognition, Universitat Pompeu Fabra
-
-pagetitle: Cross-linguistic similarity and  word acquisition
-abstract: |
-    Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 10 to 36 months.
-thanks: |
+      - ref: cbc
+
+csl: "assets/apa7.csl"
+bibliography: "assets/references.bib"
+
+editor: source
+
+format:
+  apaquarto-docx: default
+  apaquarto-pdf:
+    pdf-engine: xelatex
+    documentmode: man
+    fontsize: 12pt
+    keep-tex: false
+    floatsintext: true
+    blank-lines-above-title: 2
+    header-includes:
+      - \usepackage{tipa}
+      - \usepackage[pinyin]{babel}
+      - \usepackage{CJKutf8}
+      - \usepackage{fontspec}
+      - \babelprovide[script=CJK, language={Chinese Simplified}]{chinese-simplified}
+      - \babelfont[chinese-simplified]{rm}[Ligatures = Common]{Noto Serif CJK SC}
+      - \babelfont[chinese-simplified]{sf}[Ligatures = Common]{Noto Sans CJK SC}
+
+    include-in-header:
+      - text: '\usepackage{lineno}\linenumbers'
+
+author-note:
+  blank-lines-above-author-note: 3
+  disclosures:
+    conflict-of-interest: |
+        The authors declare no conflicts of interest with regard to the funding source of this study.
+    financial-support: |
         This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
 Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
-The authors declare no conflicts of interest with regard to the funding source of this study.
-We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and Ege E.
-Özer for their help-ful feedback.
+    gratitude: |
+        We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, Ege E.
+Özer, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
 We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support.
 We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
-We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises.
 We also thank all families and infants who participated in the experiments.
-Data collection was half-way when the COVID-19 pandemic started.
-We would like to pay special tribute to the families that collaborated with us under these difficult circumstances.
-keywords: lexical acquisition, vocabulary, bilingualism
-toc: false
-number-sections: true
+
+abstract: |
+    Language non-selectivity is prominent feature of bilingual lexical processing.
+An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates.
+The specific mechanisms behind such facilitation are unclear.
+We present an account of early bilingual vocabulary growth in which cognateness interacts with lexical frequency and language exposure to facilitate the word acquisition.
+We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals aged 12 to 34 months.
+We used Bayesian Exploratory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
+We found converging evidence for an earlier age-of-acquisition for cognate words, and for such effect being mediated by lexical frequency and language exposure.
+Low-frequency words, and words from the language of least exposure were more strongly benefitted by their cognate status than high-frequency words.
+Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
+
+keywords: [lexical acquisition, vocabulary, bilingualism, item response theory, bayesian]
+
 fig-dpi: 1000
-csl: apa.csl
-bibliography: manuscript/references.bib
-google-scholar: true
-format:
-  pdf:
-    pdf-engine: xelatex
-    geometry:
-      - top=30mm
-      - left=20mm
-      - heightrounded
-    keep-tex: true
-    linestretch: 1.25
 ---
+{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}
+
+```{r setup}
+#| label: setup
+#| echo: false
+#| message: false
+#| warning: false
+# load packages ----------------------------------------------------------------
+tar_load_globals()
+
+# load objects -----------------------------------------------------------------
+tar_load_everything()
+
+items <- bvq_data$pool  |>  
+    mutate(item = str_remove(item, "cat_|spa_")) |> 
+    select(item, language, semantic_category, class) |> 
+    right_join(items, by = c("item", "language"))
+
+# set ggplot theme and colour palette ------------------------------------------
+
+theme_set(theme_custom())
+
+bayesplot_theme_set(theme_custom()  +
+                        theme(panel.grid = element_blank(),
+                              axis.text = element_text(size = 5),
+                              strip.text = element_text(size = 5),
+                              plot.title = element_text(size = 8),
+                              plot.subtitle = element_text(size = 8)))
+
+clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+
+# set params -------------------------------------------------------------------
+
+color_scheme_set(rep(clrs[1], 6))
+options(ggplot2.ordinal.fill = clrs,
+        ggplot2.ordinal.colour = clrs,
+        ggplot2.discrete.fill = clrs,
+        ggplot2.discrete.colour = clrs,
+        ggplot2.continuous.fill = scale_color_gradient,
+        ggplot2.continuous.colour = scale_color_gradient,
+        knitr.graphics.error = FALSE) 
+```
+
+# Introduction
+
+The foundations of word learning are in place early in age.
+From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [@fenson2007macarthurbates; @samuelson2021precision].
+During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
+Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e., vocabulary size, @fenson1994variability; @frank2021variability].
+Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
+@tardif2008baby collected data about the first ten words acquired by 10-- to 16--month-old infants living in the United States, Hong Kong, and Beijing.
+Since birth, these infants had been learning English, Cantonese, and Mandarin, respectively.
+The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caregivers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animal sounds (*woof-woof*).
+These results were later extended by @frank2021variability to a more diverse set of languages, and found that such cross-language commonalities are stronger at earlier stages of lexical acquisition.
+
+Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to substantial proportion of the world population acquiring more than one language since early ages [@grosjean2021extent].
+Previous research on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
+@hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did.
+Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
+Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
+These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical or the phonological level.
+Little is known about vocabulary acquisition in bilinguals learning a pair of typologically closer languages (e.g., when both are Romance languages).
+
+For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language.
+Depending on the similarity between both languages, the two sets of words may overlap in form in varying degrees.
+When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like Spanish and English (Romance and Germanic languages, respectively).
+For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might be exposed to the words *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
+It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
+If cognates are easier to acquire than non-cognates, bilinguals learning two languages that share a high proportion of cognates should benefit more often from this facilitation effect than bilinguals learning two languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
+
+@floccia2018introduction provided evidence in line with this claim.
+The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
+The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
+The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of *cognateness* between each pair of languages.
+Phonological overlap was measured by computing the Levenshtein distance between each cross-language pair of phonetic transcriptions.
+The Levenshtein distance is a metric that computes the distance between two strings of characters/phonemes by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
+For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/kæt/-/\textipa{kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /æ/ by /\textipa{A}/).
+The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /kæt/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/.
+
+Levenshtein distance scores are divided by the length of the longest string to normalise the similarity scores between 0 and 1.
+Finally, the result is subtracted from one, to get the complementary, and therefore a measure of phonological similarity, instad of dissimilarity.
+Following the previous examples: $1/4 = 0.25$ for /kæt/-/\textipa{kAt}/, and $3/4 = 0.75$ for /kæt/-/\textipa{Gata}/.
+Finally, this variable was included as a predictor of participants' vocabulary sizes.
+Among other findings, Floccia and co-workers reported an increase in vocabulary size in the additional language (i.e., not English) associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
+For example, English-Dutch bilinguals (0.2214 mean phonological overlap), were able to produce more Dutch words than English-Mandarin bilinguals (0.0197 mean phonological overlap) were able to produce in Mandarin.
+
+Floccia et al.
+pointed to *parallel activation* as the underpinning mechanism behind their results.
+The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain], even in fully monolingual environments.
+One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
+In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish.
+In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of *table*), importantly such distinction was only relevant for bilinguals.
+Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
+In contrast, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, showed equivalent naming times for the two types of stimuli.
+These results showed that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
+
+Regarding word acquisition, the parallel activation account is a plausible explanation for Floccia et al.'s results in the light of two observations.
+First, bilingual toddlers' lexicon seems to consist of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage].
+Second, there is experimental evidence of parallel activation during word comprehension in children and toddlers [@vonholzen2012language; @poulin-dubois2013lexical; @bosma2020cognate].
+It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition.
+This could ultimately lead to children who are learning languages with a larger proportion of cognates (e.g., English and Dutch) showing larger vocabulary sizes.
+This account in which cognateness underlies the facilitation effect of typological language distance is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
+If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity (e.g., are taxonomically related) but not phonological similarity.
+
+Evidence supporting an earlier age of acquisition for cognates is scarce.
+@bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
+@schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual produced cognates earlier than non-cognates, on average.
+The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
+Floccia et al.'s estimates are statistically more reliable given their remarkably larger sample size (372 participants).
+Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level*.
+The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
+By aggregating the responses from all words into a single datum per child, it was no longer possible to test the effect of cognateness on word acquisition at the word-form level.
+Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as the response variable, age-sensitive claims about the effect of cognateness on word acquisition should be taken with caution [a limitation also present in the @bosch2014first study].
+
+More recently, @mitchell2022cognates addressed this issue providing a more fine grained measure of a  cognateness facilitation effect in vocabulary acquisition.
+Using a longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages.
+They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
+The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty).
+These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size seems to grow faster when both languages are more similar at the lexical level.
+Word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
+Parallel activation remains a plausible explanation for this effect.
+When a child is exposed to a word form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition.
+The precise underpinnings of this effect are still unclear.
+
+In the present investigation we approach the cognate facilitation on word learning by extending the conceptual framework of accumulator models of language acquisition to the bilingual case.
+Accumulator models formalise the widespread assumption that a child's vocabulary growth rate is a function of mainly---but not exclusively---the number of learning instances that a child has accrued with a particular word [see @kachergis2022standard for review].
+A learning instance of a word consists, for example, in hearing such word-form (*flower*) in the presence of its referent (a flower).
+As infants receive linguistic input, they also accumulate opportunities to learn words until a threshold is crossed, an the word can be considered to be acquired.
+In line with this account, the frequency with which a child encounters a particular word-form in their speech input is a critical determiner of the age at which such word will be acquired.
+The more frequently [@hidaka2013computational; @mollica2017how].
+For instance, words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'taxi' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does]: infants might accumulate learning experiences with 'flower' faster than with 'taxi'.
+
+Previous accounts on vocabulary growth based on accumulator models are a convenient framework to study bilingual vocabulary growth, as bilinguals share fundamentally similar learning mechanisms with monolinguals.
+Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilingual].
+One of such divergences is the fact that bilinguals' frequency of exposure to a word-form is mediated by their degree of exposure to their languages.
+Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual would (who listens 100% of the time to English).
+
+We suggest that the cognateness effect proposed by Floccia et al.
+and Mitchell et al.
+facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired.
+We hypothesise that both forms in the translation equivalent (e.g., *gat* and *gato* [Catalan and Spanish for 'cat'], for a Catalan-Spanish bilingual) follow their acquisition trajectory as they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to.
+When one of the word-forms is sufficiently acquired, the number of learning instances left needed for the acquisition of the other word-form is reduced.
+
+This facilitation would occur as the result of the semantic and phonological overlap between both words.
+When the child has sufficient lexical knowledge of *gat*, and therefore there is some for-referent association in place, future presentations of the similar sounding translation *gato* in the presence of the same reference should lead to an increased likelihood of acquisition.
+In the case of a non-cognate translation equivalent, such facilitation would be unlikely.
+For instance, even if the child has already acquired the word *gos*, hearing its dissimilar Spanish translation *perro* would hardly lead to an increased probability of acquisition.
+In line with this account, only the later-acquired form of the translation equivalent would benefit from its cognate status.
+Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age of acquisition for cognates than for non-cognates.
+
+{apafg-hypotheses} illustrates graphically a theoretical account of the acquisition of a Catalan-Spanish cognate translation equivalent, *gat*-*gato* [cat] whose word-forms share three phonemes, and of a non-cognate translation equivalent, *gos*-*perro* [dog] whose word-forms only share one phoneme.
+We assume the case of a Catalan-Spanish bilingual child who is exposed 65% of the time to Catalan, and 35% of the time to Spanish.
+On the top panels, we illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/gat/ or /gos/) and its Spanish translation (/perro/ or /gato/).
+We assume an arbitrary threshold of learning instances at which the child is considered to have acquired a word (solid grey line).
+Points and triangles show the age-of-acquisition at which the amount of learning instances reaches the threshold.
+We show the predicted age-of-acquisition under two hypothesis: (1) a cognateness facilitation effect that lowers the acquisition threshold after one of the word-forms (in this case Catalan /gat/) has been acquired, and (2) no cognateness facilitation, for which the acquisition of the second word-form occurs unconditionally of the acquisition of the first.
+On the bottom, we show the predicted probability of acquisition for the Catalan and Spanish words under each hypothesis, as it would be expected to observe in a vocabulary questionnaire like the one in the present study.
+We assume that such probability starts at 0% at birth, reaches 50% (and grows the fastest) at the age of acquisition, and reaches 100% at some points in later ages.
+
+```{r apafg-hypotheses}
+#| label: apafg-hypotheses
+#| apa-cap: "Schematic representation of the hypothesised cognate facilitation effect on word acquisition."
+#| out-width: 100%
+knitr::include_graphics(normalizePath("assets/img/diagram.png"))
+```
+
+In line with this hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al.
+reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure).
+Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
+Therefore, participants may have on average learned the English word-form of translation equivalents earlier than the word-form in the additional language.
+Because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants might then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
+
+We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a word *only* once its translation has been already acquired.
+We collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
+Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees.
+We adopted a Bayesian explanatory item response theory  [IRT, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
+
+To operationalise participants' exposure rate to a word-form, we used a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it), and child's amount of exposure to the language such word belongs to.
+This measure was calculated by multiplying the word-form's lexical frequency with participant's degree of exposure to the language the word belongs to.
+Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childes] (see Appendix A for more details).
+Participants' degree of exposure were reported by caregivers in a language exposure questionnaire they filled before the vocabulary checklist.
+Caregivers provided a proportion of the time the child was estimated to listen to any language they were spoken to regularly (described in the methods section).
+The resulting measure is a exposure-weighted lexical frequency measure, which was included in analyses as a proxy of how often bilingual participants were exposed to each word-form.
+
+*Cognateness* was defined as a continuous measure of phonological similarity between translation equivalents.
+We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computer].
+This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
+This results in a proportion that indicates how much the two phonetic transcriptions of the translation equivalent are similar to each other, ranging from zero (no similarity at all) to one (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
+For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]` and *pájaro* `[paxa4o]`, Catalan and Spanish for 'bird', and returns 50% for *lluna* `[Lun5]` and *luna* `[luna]`, Catalan and Spanish for 'moon' (see Appendix A for more details).
+
+# Methods {#sec-methods}
+
+All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/hy984/)) and GitHub ([https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories)) repositories.
+This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.
+
+## Questionnaire {#sec-questionnaire}
 
 ```{r}
-#| label: setup
+#| label: items
+#| echo: false
+#| message: false
+#| warning: false
 #| include: false
-knitr::opts_chunk$set(
-    echo = FALSE,
-    fig.align = "center",
-    message = FALSE,
-    warning = FALSE,
-    cache.extra = knitr::rand_seed,
-    out.width = "100%",
-    dpi = 800,
-    results = "asis"
-)
-```
-
-```{r}
-#| label: packages
+n_categories <- nrow(distinct(bvq_data$pool, semantic_category))
+
+n_te <- length(unique(bvq_data$pool$te))
+
+n_items <- length(unique(bvq_data$pool$item))
+
+n_item_language <- bvq_data$pool |> 
+    count(language) |>
+    pull(n) |>
+    set_names(c("catalan", "spanish"))
+
+n_items_version <- bvq_data$pool |>
+    unnest(version) |> 
+    count(version, language) |> 
+    pivot_wider(names_from = language,
+                values_from = n) |> 
+    clean_names() |> 
+    mutate(version = tolower(version))
+
+```
+
+To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire.
+This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formR platform [@arslan2020formr].
+This questionnaire is structured in three blocks: (1) a language exposure questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+both lists included items from a diverse sample of `r n_categories` semantic/functional categories.
+The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
+Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for 'dog') (see {apatb-items} for a summary of the questionnaire items).
+
+```{r apatb-items}
+#| label: apatb-items
+#| apa-cap: "Summary of the items included in the final analyses."
+#| eval: true
+#| echo: false
+#| warning: false
+items |> 
+    arrange(language, list) |> 
+    distinct(te, .keep_all = TRUE) |> 
+    mutate(a = grepl("A", list),
+           b = grepl("B", list),
+           c = grepl("C", list),
+           d = grepl("D", list)) |> 
+    group_by(semantic_category) |> 
+    summarise(across(a:d, sum), 
+              items = list(meaning),
+              .groups = "drop") |> 
+    mutate(total = rowSums(across(a:d)),
+           items = map(items, function(x) head(x, 2))) |>  
+    arrange(desc(total)) |> 
+    relocate(a:d, total, items) |> 
+    gt(rowname_col = "semantic_category") |> 
+    fmt_percent(matches("_prop"), decimals = 1) |> 
+    cols_label(a = "List A",
+               b = "List B",
+               c = "List C",
+               d = "List D",
+               total = "Total",
+               items = "Examples") |> 
+    tab_style(cell_text(size = "small"),
+              list(cells_body(),
+                   cells_stub(),
+                   cells_column_labels(),
+                   cells_column_spanners(),
+                   cells_row_groups())) |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels()) |> 
+    tab_options(data_row.padding = px(2),
+                row_group.padding = px(1.2)) |> 
+    tab_style(cell_text(size = "x-small", align = "left"),
+              cells_body(items))
+
+```
+
+For each word included in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
+Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items.
+Each subset contained a random but representative sub-sample of the items from the complete list.
+Semantic/functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Items that were part of the trial lists of some ongoing experiments in the lab were also included in all versions.
+The resulting reduced vocabulary checklists contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words.
+Participants were randomly allocated into one of the four subsets.
+Each response to the questionnaire provided data for a minimum of `r min(n_items_version$catalan, n_items_version$spanish)` and a maximum of `r max(n_items_version$catalan, n_items_version$spanish)` translation equivalents.
+
+## Participants {#sec-participants}
+
+```{r participants}
+#| label: participants
 #| echo: false
@@
 # load packages
-library(brms) 
-library(conflicted)
-library(dplyr) 
-library(forcats)
-library(ggplot2) 
-library(ggsci)
-library(gt) 
-library(here) 
-library(janitor)
-library(knitr)
-library(lubridate)
-library(multilex)
-library(patchwork) 
-library(papaja) 
-library(purrr)
-library(scales) 
-library(stringr)
-library(tidybayes) 
-library(tidyr) 
-library(tibble)
-
-# load and process objects
-tar_load(participants)
-tar_load(items)
-tar_load(multilex_data)
-tar_load(model_log_liks)
-tar_load(model_loos)
-tar_load(model_fit_1)
-tar_load(rope_coefs)
-tar_load(rope_interval)
-tar_load(df)
-tar_load(posterior_description)
-
-participants <- participants %>% 
-    left_join(select(multilex_data$logs, id, time, edu_parent))
-items <- items %>% 
-    left_join(select(pool, item, language, category, class))
-# set global options
-options(
-    knitr.kable.NA = "-",
-    knitr.duplicate.label = "allow",
-    gt.html_tag_check = FALSE,
-    ggplot2.discrete.fill = ggsci::pal_d3()(4),
-    ggplot2.discrete.colour = ggsci::pal_d3()(4),
-    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
-    ggplot2.continuous.colour = ggplot2::scale_color_gradient
-)
-
-# set custom ggplot2 theme
-theme_set(theme_custom())
-
-# resolve namespace conflicts
-conflict_prefer("filter", "dplyr")
-conflict_prefer("select", "dplyr")
-```
-
-# Introduction {#introduction}
-
-The foundations of word learning are in place early in age: before the end of their first year of life infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999some; @bergelson2012months; @bergelson2015early] and parental reports [e.g., @fenson2007macarthur].
-During the last half of their second year, they acquire new words at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002children; @bergelson2020comprehension].
-These early stages of lexical acquisition are characterised by substantial variation across children, reflected, for instance, on the variation on the number of words they know [i.e.
-vocabulary size, @fenson1994variability; @frank2021variability] or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words [e.g., @nelson1973structure, bates1994developmental].
-
-Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages.
-@tardif2008baby collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing.
-Since birth, these infants had been learning English, Cantonese and Mandarin, respectively.
-The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*).
-These results were later extended by @frank2021variability to a wider diversity languages, also reporting that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
-Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
-
-There is evidence that bilinguals know, on average, less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
-For example, @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English.
-Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez2020bilingual].
-While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
-What role could linguistic distance play during early vocabulary growth?
-
-For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language.
-Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
-Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English (one Roman, the other Germanic).
-For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
-It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
-If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
-
-@floccia2018introduction provided evidence in line with this claim.
-The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
-The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
-The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English.
-Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
-The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other [@levenshtein1966binary].
-The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
-Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
-For example, English-Dutch bilinguals (22.14% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% phonological similarity) were able to produce in Mandarin.
-
-Floccia et al.
-pointed to *parallel activation* as the main mechanism underpinning their results.
-The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension, and that this phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain].
-One of the clearest examples of parallel activation was provided by @costa2000cognate.
-In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish.
-In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trials labels were non-cognates (*mesa*-*taula*, translations of *table*).
-Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
-Importantly, Spanish monolinguals--unfamiliar with the Catalan translations of the Spanish words they uttered--did not show this effect.
-These results suggest that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-Several subsequent studies have also provided similar evidence in comprehension in children [e.g., @von2012language; @poulin2013lexical].
-Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
-A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.
-
-In this study, we hypothesise that cognates are acquired earlier in age than non-cognates, which might explain why children learning languages that share many cognates show larger vocabulary sizes.
-Evidence supporting this claim is scarce.
-@bosch2014first used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
-@schelletter2002effect reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
-The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-On the other hand, Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
-In their discussion the authors state the following (pp.
-70):
-
-> "This  finding  also  provides  support  to  the  proposal  that  the  cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life."
-
-We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results.
-First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
-By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.
-non-cognates was lost.
-Second, all participants were aged ~24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
-
-More recently, @mitchell2022cognates addressed this issue overcoming some of these pitfalls.
-Using a larger, longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages.
-They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
-The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty).
-These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar: word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
-
-As @mitchell2022cognates note in their discussion, two competing hypotheses remain untested.
-The facilitative effect of cognateness can be explained by, at least, two alternative, arguably incompatible, mechanisms.
-On the one hand, cognateness might operate as a facilitator of the acquisition of both forms of a translation pair, with every exposure to either form activating the other, and therefore strengthening the association between both and their referent.
-If this is the case, then cognateness should increase the probability of acquisition of the forms in both languages equally.
-
-Another possibility is that cognateness plays a role in the acquisition of a translation pair only when at least one of the word-forms has been acquired.
-This account departs from the premise that, if the child is not previously familiar with one of the word forms, the acquisition of its translation in the other language cannot be facilitated by cognateness.
-This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair.
-Given that children are more likely to acquire words from languages to which they are exposed more often [@david2008individual; @cattani2014much; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
-
-In line with this second hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.
-reported was larger in the additional language vocabulary than in English vocabulary.
-Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
-Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
-If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
-
-In this study, we investigated the role of cognateness on the acquisition of translation equivalent, and whether such effect is conditional to the amount of time the child is exposed to the corresponding language of each member of the translation pair.
-Using an online vocabulary checklist--designed specifically for this study--we collected data from a sample of children aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We then adopted an Item Response Theory (*IRT*) approach to model the probability of participants being reported by their parents to understand or understand *and* say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for other indices of word difficulty--lexical frequency and length (phonemes)-- and participant skill--age [see @kachergis2022toward for a similar approach].
-
-# Methods {#methods}
-
-Data and materials are available at OSF [https://osf.io/hy984/](https://osf.io/hy984/), and code is available at GitHub [https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories).
-
-## Participants {#participants}
-
-```{r}
-#| label: participants
+n_participants <- responses |>
+    distinct(id, age) |>
+    count(id) |>
+    pull(n) |>
+    table()
+
+diff_responses <- responses |>
+    distinct(id, age) |>
+    left_join(distinct(participants, id, time_stamp, age),
+              by = c("id", "age")) |> 
+    group_by(id) |> 
+    arrange(id, time_stamp) |> 
+    mutate(diff_lag = time_stamp-dplyr::lag(time_stamp, default = NA)) |> 
+    summarise(diff_min = min(diff_lag, na.rm = TRUE),
+              diff_max = max(diff_lag, na.rm = TRUE),
+              .groups = "drop") |> 
+    filter(!(diff_min %in% c(-Inf, Inf))) |>
+    summarise(diff_min = min(diff_min, na.rm = TRUE),
+              diff_max = max(diff_max, na.rm = TRUE))
+
+n_responses <- nrow(distinct(responses, id, age))
+
+dates <- participants$time_stamp |> 
+    range(na.rm = TRUE) |> 
+    as_date()
+
+n_trilingual <- participants |> 
+    select(starts_with("doe_")) |> 
+    mutate(doe_total = rowSums(across(everything())),
+           is_trilingual = doe_total < 1) |> 
+    add_count(name = "n_total") |> 
+    count(is_trilingual, n_total) |> 
+    mutate(prop = n/n_total)
+
+participants_edu <- table(participants$edu_parent)
+```
+
+We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between the `r format(dates[1], "%dth of %B, %Y")` and the `r format(dates[2], "%dth of %B, %Y")`: `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
+Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`, and were always allocated to the same questionnaire list (A, B, C, or D).
+Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see {apatb-participants} for a more detailed description of the sample).
+In total, `r n_trilingual$n[2]` participants (`r percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed to a third language other than Catalan and Spanish.
+All families provided informed consent before participating.
+Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device.
+Filling the questionnaire took 30 minutes approximately.
+After completion, families were rewarded with a token of appreciation.
+
+```{r apatb-participants}
+#| label: apatb-participants
+#| echo: false
+#| message: false
+#| warning: false
+#| apa-cap: "Participant sample size by age and degree of exposure to Catalan."
+#| apa-note: "Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
+In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
+A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
+A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
+The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
+For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins.
+Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
+participants |>
+    mutate_at(vars(starts_with("doe_")), 
+              (\(x) floor(x * 10) / 10)) |> 
+    mutate(
+        age = cut(floor(age),
+                  c(10, 14, 18, 22, 26, 30, 34, 36),
+                  include.lowest = TRUE),
+        doe_catalan = cut(
+            doe_catalan,
+            c(0, 0.25, 0.5, 0.75, 1),
+            labels = c("0-25%",
+                       "25-50%", 
+                       "50-75%",
+                       "75-100%"),
+            include.lowest = TRUE
+        ) |>
+            as.character()
+    ) |>
+    count(age, doe_catalan) |>
+    arrange(desc(doe_catalan)) |>
+    pivot_wider(names_from = age, values_from = n) |>
+    gt() |>
+    summary_rows(
+        columns = 2:7,
+        fns = list("Total" = ~ sum(., na.rm = TRUE)),
+        formatter = fmt_integer
+    ) |>
+    sub_missing(everything(), everything(), missing_text = "--") |>
+    cols_label(doe_catalan = "Catalan exposure") |>
+    tab_spanner("Age (months)", 2:8) |>
+    tab_style(cell_text(align = "left"),
+              list(cells_body(doe_catalan),
+                   cells_column_labels(doe_catalan)))
+```
+
+We used the highest self-reported educational attainment of parents or caregivers as a proxy of participants' socio-economic status (SES).
+This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: 'sense escolaritzar/sin escolarizar' [no education], 'educació primària/educación primaria' [primary school], 'educació secundària/educación secundaria' [secondary school], 'batxillerat/bachillerato' [complementary studies/high school], 'cicles formatius/ciclos formativos' [vocational training], and 'educació universitària/educación universitaria' [university degree].
+Most families reported university studies (`r participants_edu["University"]`, `r percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
+
+## Data analysis {#sec-analysis}
+
+```{r sample-sizes}
+#| label: sample-sizes
 #| echo: false
@@
 #| include: false
-# load packages
-n_participants <- length(unique(participants$id))
-dates <- as_date(range(participants$time_stamp, na.rm = TRUE))
-```
-
-We collected data from `r printnum(n_participants)` children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`.
-All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-Families from the participant database of the Laboratori de Recerca en Infància of the Universitat Pompeu Fabra were contacted by e-mail or phone if their child were aged between `r floor(min(participants$age))` and `r floor(max(participants$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan.
-Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate.
-@tbl-participants-lp summarises the distribution of participants across ages and degrees of exposure (DoE) to Catalan.
-
-```{r}
-#| label: tbl-participants-lp
+n_obs <- nrow(responses)
+
+sum_obs <- responses |>
+    mutate(te = factor(te)) |>
+    count(te) |>
+    summarise(across(n, lst(median, min, max), 
+                     .names = "{.fn}"))
+
+# attrition
+n_items <- bvq_data$pool
+
+n_items_nouns <- n_items |> 
+    filter(class=="Noun")
+
+n_items_freq <- n_items_nouns |> 
+    left_join(childes,
+              by = c("childes_lemma" = "token")) |> 
+    drop_na(freq_zipf, wordbank_lemma)
+
+n_items_lemma <- n_items_freq |> 
+    filter(n_lemmas == 1)
+
+n_items_multi <- n_items_lemma |> 
+    filter(!is_multiword) |> 
+    filter(include, !is.na(wordbank_lemma)) |> 
+    drop_na(sampa) |> 
+    distinct(te, language)
+
+n_items_duplicated <- n_items_multi |> 
+    add_count(te, name = "n_te") |>
+    filter(n_te == 2)
+
+n_items_final <- n_items_duplicated |> 
+    count(language) |> 
+    group_split(language) |> 
+    set_names(c("Catalan", "Spanish")) |>
+    map_int("n")
+
+n_te <- length(unique(responses$te))
+
+rope_interval <- percent(c(lower = -0.1/4, upper = 0.1/4), accuracy = 0.1)
+
+r_info <- sessionInfo()
+
+r_version <- glue(r_info$R.version$major, 
+                  r_info$R.version$minor, 
+                  .sep = ".")
+```
+
+From the responses to the `r format(nrow(n_items), big.mark = ",")` items included in the questionnaire, we restricted the analyses to responses to nouns (`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded) [@fourtassi2020growth].
+We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
+Finally, we removed items without a translation in the other language (*n* = `r nrow(n_items_multi)-nrow(n_items_duplicated)`).
+This resulted in a final list of `r sum(n_items_final)` items, corresponding to words `r n_items_final["Catalan"]` Catalan words and their `r n_items_final["Spanish"]` Spanish translations (`r n_te` translation equivalents).
+After collecting participants' responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item.
+Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
+
+We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_3$fit)[2]` sampling chains with `r format(dim(model_fit_3$fit)[1], big.mark = ",")` iterations each).
+This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $\text{Age}$ (number of months elapsed between participants' birth date and questionnaire completion), $\text{Length}$ (number of phonemes in the X-SAMPA phonological transcription of the word-form), $\text{Exposure}$ (composite measure of lexical frequency and language degree of exposure), and $\text{Cognateness}$ (a measure of cognateness, defined as the phonological similarity between translation equivalents).
+For an easier interpretation of the regression coefficients, the values of all predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
+Participant- and item-level random intercepts and slopes were included when appropriate, according to the structure of the data [@barr2013random].
+We specified a weakly informative prior around the parameters of the model (see Appendix C for details).
+
+We built a baseline model $\mathcal{M}_{0}$ that only included the intercept and the main effects of $\text{Age}$, $\text{Length}$, and $\text{Exposure}$ as predictors, and compared it against increasingly more complex models, which introduced predictors in the following order: $\text{Age} \times \text{Exposure}$ ($\mathcal{M}_{1}$), $\text{Cognateness}$ ($\mathcal{M}_{2}$), $\text{Exposure} \times \text{Cognateness}$ ($\mathcal{M}_{3}$), and finally $\text{Age} \times \text{Cognateness}$ and $\text{Age} \times \text{Exposure} \times \text{Cognateness}$ ($\mathcal{M}_{4}$).
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO-CV, @vehtari2017practical]^[Due to computational constraints, we performed LOO-CV based on a random sub-sample of `r format(dim(model_log_liks$model_fit_3)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]].
+LOO-CV involves computing the likelihood of each data point, given a model  fitted excluding such data point.
+The likelihoods generated by the model for each of the data points are summarised together in a measure of predictive performance: the expected log-predicted density (ELPD).
+This measure penalises the complexity of the model ELPD.
+ELPD values closer to zero indicate better predictive accuracy.
+We compared the ELPD of each model, and selected for interpretation the one showing the best performance (see Appendix C for more details about the models).
+
+We assessed the practical relevance of the estimated regression coefficients of the selected model following @kruschke2018bayesian.
+First, we specified a region of practical equivalence (ROPE) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`, in the probability scale.
+This region indicates the range of values that we considered equivalent to zero.
+The degree of overlap between the posterior distribution of a regression coefficient with the  can be interpreted as support for the true parameter of the coefficient being zero or equivalent.
+We then computed the 95% highest density interval (HDI) of the posterior distribution of each regression coefficient.
+This interval contains the true value of this coefficient with 95% probability, given the data.
+Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, which indicates the probability that the true value of the regression coefficient falls into the ROPE, and should therefore be considered equivalent to zero.
+For example, a 90% overlap between the HDI of a coefficient's posterior distribution and the  indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the ROPE, and is therefore equivalent to zero.
+
+Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
+Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@vehtari2022loo], `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
+
+# Results {#sec-results}
+
+```{r results-values}
+#| label: results-values
+#| echo: false
+rhats_df <- map_df(model_rhats, bind_rows, .id = "model") |> 
+    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
+    pivot_longer(-model, 
+                 names_to = ".variable",
+                 values_to = ".rhat") |>
+    filter(str_detect(.variable, "b_|sd_")) |> 
+    drop_na()
+
+loo_diff <- loo_compare(model_loos)[2, 1]
+loo_diff_se <- loo_compare(model_loos)[2, 2]
+
+post_draws_list <- posterior_draws$summary |> 
+    mutate(across(.median:.lower,
+                  \(x) ifelse(grepl("Intercept", .type), plogis(x), x/4))) |> 
+    split(posterior_draws$summary$.variable)
+```
+
+All models showed adequate convergence diagnostics and little evidence of multicollinearity between predictors (see Appendix D for details).
+$\mathcal{M}_{3}$, which did *not* include the $\text{Age} \times \text{Exposure} \times \text{Cognateness}$ three-way interaction, showed the best predictive performance with a difference in expected log-predicted density (ELPD) compared to the second best-performing model ($\mathcal{M}_{4}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`.
+Given our data, the predictions of the $\mathcal{M}_{3}$ are confidently more accurate than those of the other models (see {apatb-results-loos} for all model comparisons).
+Both $\mathcal{M}_3$ and $\mathcal{M}_4$ performed considerably better than the rest of the models, which did not include the $\text{Exposure} \times \text{Cognateness}$ two-way interaction.
+This points to the effect of $\text{Cognateness}$ being strongly modulated by $\text{Exposure}$, and that this interaction remains relatively stable across ages.
+
+```{r apatb-loos}
+#| label: apatb-results-loos
 #| echo: false
@@
 #| warning: false
-#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan."
-#| tbl-subcap: "Information about DoE was provided by families before filling the questionnaire.
-A 100% indicated that the participant was exclusively exposed to Catalan, and never to Spanish.
-A 0% indicates that the participant was not exposed to Catalan ever, and rather most of the time to Spanish.
-A 50% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
-The 100%, 0%, and 50% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
-For illustration purposes this table, DoEs were binned into 25%-wide bins, and ages (in months) binned into 4 month-wide bins.
-Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
-participants %>%
-    mutate_at(
-        vars(starts_with("doe_")),
-        ~floor(.*10)/10
-    ) %>%
-    mutate(
-        age = cut(
-            floor(age),
-            c(10, 14, 18, 22, 26, 30, 34, 36),
-            include.lowest = TRUE
-        ),
-        doe_catalan = cut(
-            doe_catalan,
-            c(0, 0.25, 0.5, 0.75, 1),
-            labels = c("0-25%", "25-50%", "50-75%", "75-100%"),
-            include.lowest = TRUE
-        ) %>% 
-            as.character()
-    ) %>%
-    count(age, doe_catalan) %>%
-    arrange(desc(doe_catalan)) %>%
-    pivot_wider(names_from = age, values_from = n) %>%
-    gt() %>%
-    summary_rows(
-        columns = 2:8,
-        fns = list("Total" = ~sum(., na.rm = TRUE)),
-        formatter = fmt_integer
-    ) %>%
-    sub_missing(everything(), everything(), missing_text = "--") %>%
-    cols_label(doe_catalan = "DoE Catalan") %>%
-    tab_spanner("Age (months)", 2:8) %>%
-    tab_footnote(
-        "This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10% of the time",
-        cells_column_labels(columns = doe_catalan)
-    ) %>%
-    tab_style(
-        cell_text(align = "left"),
-        list(
-            cells_body(doe_catalan),
-            cells_column_labels(doe_catalan)
-        )
-    )
-```
-
-We used the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: *No education*, *Primary*, *Secondary*, *Complementary*, *Vocational*, and *University*, in line with the current educational system in Spain.
-Most families reported university studies (`r table(participants$edu_parent)["University"]`, `r table(participants$edu_parent)["University"]/length(unique(participants$id))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(table(participants$edu_parent)["Vocational"]/length(unique(participants$id)))`), complementary studies (`r table(participants$edu_parent)["Complementary"]`,  `r percent(table(participants$edu_parent)["Complementary"]/length(unique(participants$id)))`), secondary education (`r table(participants$edu_parent)["Secondary"]`, `r percent(table(participants$edu_parent)["Secondary"]/length(unique(participants$id)))`, `r percent(table(participants$edu_parent)["Primary"]/length(unique(participants$id)))`), no formal education (`r table(participants$edu_parent)["No education"]`, `r percent(table(participants$edu_parent)["No education"]/length(unique(participants$id)))` and primary education (`r table(participants$edu_parent)["Primary"]`, <1%).
-
-## Questionnaire {#questionnaire}
-
-```{r}
-#| label: items
+#| eval: true
+#| apa-cap: "Bayesian leave-one-out cross validation (LOO-CV)."
+#| apa-note: "We report the ELPD of each model along their associated standard errors (SE).
+The last column indicates the difference in ELPD between each model and the model with the best predictive accuracy (Model 3).
+Pareto-k estimates of all models were acceptable (k < 0.5)."
+model_loos |>
+    loo_compare() |>
+    as.data.frame() |>
+    rownames_to_column("model") |>
+    mutate(model = paste0("Model ", str_remove(model, "model_fit_"))) |>
+    select(model, elpd_loo, se_elpd_loo, elpd_diff, se_diff) |> 
+    gt(rowname_col = "model") |>
+    fmt_number(2:5, decimals = 2, sep_mark = ",") |>
+    cols_merge_uncert(elpd_loo, se_elpd_loo) |> 
+    cols_merge_uncert(elpd_diff, se_diff) |> 
+    cols_label(
+        model = "Model",
+        elpd_loo = "ELPD $$\\pm$$ SE",
+        elpd_diff = "Diff. $$\\pm$$ SE"
+    ) |>
+    tab_style(cell_text(weight = "bold", style = "italic"),
+              cells_column_labels()) |>
+    tab_style(cell_text(align = "left"),
+              cells_title(groups = "subtitle"))
+```
+
+{apatb-coefs} shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{3}$, and their degree of overlap with the ROPE.
+For interpretability, we report the estimated regression coefficients transformed to the probability scale.
+The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales relate non-linearly.
+This means that one logit difference is not necessarily translated to a unique value in the probability scale.
+For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 30 to 32 months.
+The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients in the probability scale is not trivial.
+Following @gelman2020regression, we report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
+derivative) of the logistic curve at the age at which most participants were acquiring a given word.
+This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
+
+```{r apatb-coefs}
+#| label: apatb-coefs
+#| apa-cap: "Posterior distribution of regression coefficients Model 3."
+#| apa-note: "Median: median of the posterior distribution in the probability scale.
+95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.
+On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE).
+This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
+Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
 #| echo: false
@@
 #| warning: false
-#| include: false
-n_categories <- nrow(distinct(multilex_data$pool, category))
-n_te <- length(unique(multilex_data$pool$te))
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>%
-    pull(n) %>% 
-    set_names(c("catalan", "spanish"))
-n_missing_freq <- multilex::pool %>% 
-    filter(
-        include,
-        class %in% c("Noun", "Verb", "Adjective")
-    ) %>% pull(frequency_zipf) %>% 
-    is.na() %>%
-    table() %>% 
-    {.[2] / .[1]}
-```
-
-The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]`.
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-In case two translation equivalents were possible for a given word, both were included as separate items (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into a single item (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*).
-We included items from a diverse sample of `r n_categories` semantic/functional categories.
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (`r format(n_te, big.mark = ",")` translation equivalents; `r format(n_item_language["catalan"], big.mark = ",")` items in Catalan, `r format(n_item_language["spanish"], big.mark = ",")` items in Spanish), while others filled a shorter version (~400 translation equivalents, ~400 items in Catalan, ~400 items in Spanish).
-These last families were randomly allocated into one of four different subsets of the complete list of items.
-These lists were designed so that each contained a representative sub-sample of the items from the complete list.
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-For the analyses in the present study, we considered responses to words corresponding to nouns, verbs, and adjectives [@fourtassi2020growth].
-We excluded multi-word items (e.g., *barrita de cereales* [cereal bar]) and items that included more than one word-form (e.g., *mono / mico*).
-
-We extracted four variables of interest for each word-form: (1) Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
-This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points.
-Lexical frequencies were extracted from an adult subtitle-based database: SUBTLEX-CAT [@boada2020subtlex] for Catalan words and SUBTLEX-ESP [@cuetos2012subtlex] for Spanish words.
-Using child-based estimates of lexical frequency like CHILDES was not possible due to the low number of Catalan participants and tokens in the available corpora.
-Responses to words with missing lexical frequency score were excluded from analyses (`r percent(n_missing_freq, accuracy = 0.01)` of the total number of items).
-(2) Word length ($Phonemes$): number of phonemes in the phonological transcription of the word-form in International Phonological association format.
-(3) Phonological similarity ($Levenshtein$): normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary].
-This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1.
-This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-(4) Participant age: number of months elapsed between participants' birth date and questionnaire completion.
-Finally, (5) participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to.
-For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.
-@tbl-item-summary shows a summary of the distribution of the word-level predictors across languages in and questionnaire versions.
-
-```{r}
-#| label: tbl-item-summary
+# summarise posterior draws
+intercept_label <- "Intercepts (at 22 months)"
+
+posterior_draws$summary |>
+    select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
+    mutate(across(.median:.upper,
+                  \(x) ifelse(.type==intercept_label, plogis(x), x/4))) |> 
+    gt(groupname_col = ".type", 
+       rowname_col = ".variable_name") |>
+    fmt_percent(.rope_overlap) |> 
+    fmt_number(c(.median:.upper)) |>
+    cols_merge(c(.lower, .upper),
+               pattern = "[{1}, {2}]") |>
+    cols_label(
+        .variable_name = "Parameter",
+        .median = "Median",
+        .lower = md("95% HDI"),
+        .rope_overlap = md("ROPE prob.")
+    ) |>
+    tab_style(cell_text(weight = "bold"),
+              cells_column_labels(columns = 1:5)) |>
+    tab_style(cell_text(align = "left"),
+              cells_title(groups = "subtitle")) |> 
+    tab_style(cell_text(size = "small"), 
+              list(cells_body(),
+                   cells_row_groups(),
+                   cells_stub())) |> 
+    tab_style(cell_text(style = "italic"),
+              cells_row_groups())
+```
+
+The main effect of $\text{Age}$ showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std"]]$.upper, 2)`]), with all of its posterior samples falling out of the ROPE.
+A one-month increment in age increased a maximum of `r round(post_draws_list[["b_age_std"]]$.median/sd(responses$age), 2)` the probability of acquisition.
+Similarly, the word exposure index ($\text{Exposure}$) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 2)`]).
+All of the posterior samples of this regression coefficient excluded the ROPE.
+The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), 2)` more likely to acquire it.
+Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r round(post_draws_list[["b_n_phon_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_phon_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_phon_std"]]$.upper, 2)`]).
+For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.median)/sd(responses$n_phon), 2)` less likely to know it.
+The 95% HDI of the regression coefficient of the $\text{Age} \times \text{Exposure}$ interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 2)`]), showing that the effect of the word exposure index differed across ages: older children were more sensitive likely to acquire words with higher exposure rate than younger children.
+
+The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).
+For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), 2)`.
+The effect of $\text{Cognateness}$ interacted with that of $\text{Exposure}$: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 2)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
+Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
+This effect was negligible when for words with median or high exposure (+1 SD) (see {apafg-results-marginal}).
+
+```{r apafg-results-marginal}
+#| label: apafg-results-marginal
 #| echo: false
@@
 #| warning: false
-#| tbl-cap: "Lexical properties of the word-forms and translationpairs included in the vocabulary checklist and our analyses.
-We report the number of items, mean (*M*), standard deviation (*SD*), and range of the lexical frequency (epxressed in Zipf scores), length (number of phonemes) of the word forms, and the phonological similarity (normalised Levenshtein similarity) with their translation equivalent.
-Statistics are reported separately for Catalan and Spanish, and for the different vocabulary checklists that participants were allocated to (A, B, C, D)."
-items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))
-
-items %>%
-    unnest(list) %>% 
-    group_by(language, list) %>% 
-    summarise_at(
-        vars(freq, n_phon, lv), 
-        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
-    ) %>% 
-    select(-c(n_phon_n, lv_n)) %>% 
-    rename(n = freq_n) %>% 
-    gt(groupname_col = "language") %>%
-    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
-    fmt_percent(matches("lv_")) %>% 
-    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
-    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
-    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
-    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
-    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
-    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
-    cols_label(
-        list = "",
-        n = md("*N*"),
-        freq_mean = md("Mean"),
-        freq_sd = md("*SD*"),
-        freq_min = md("Range"),
-        n_phon_mean = md("Mean"),
-        n_phon_sd = md("*SD*"),
-        n_phon_min = md("Range"),
-        lv_mean = md("Mean"),
-        lv_sd = md("*SD*"),
-        lv_min = md("Range")
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(everything())
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(everything())
-    )
-```
-
-## Data analysis {#analysis}
-
-```{r}
-#| label: sample-sizes
-#| echo: false
-#| message: false
-#| warning: false
-#| include: false
-n_obs <- nrow(df)
-sum_obs <- df %>%
-    mutate(te = factor(te)) %>% 
-    count(te) %>% 
-    summarise(across(n, lst(median, min, max), .names = "{.fn}"))
-n_participants <- df %>%
-    distinct(id, age) %>% 
-    count(id) %>% 
-    pull(n) %>% 
-    table()
-
-n_items <- ifelse(
-    grepl("cat_", distinct(df, item)$item),
-    "Catalan",
-    "Spanish"
-) %>% 
-    table()
-
-n_te <- length(unique(df$te))
-```
-
-We gathered `r format(n_obs, big.mark = ",")` observations, with each observation corresponding to a single response of one participant to a given item in the questionnaire).
-These observations correspond to `r sum(n_participants)` distinct participants (of which `r n_participants[1]` participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times), responding to `r format(sum(n_items), big.mark = ",")` distinct items (`r n_items["Catalan"]` in Catalan, `r n_items["Spanish"]` in Spanish, `r n_te` translation equivalents).
-Translation equivalents (TEs) received a median of `r printnum(sum_obs$median, digits = 0)` responses (*Min* = `r printnum(sum_obs$min, digits = 0)`, *Max* = `r printnum(sum_obs$max, digits = 0)`), both languages summed together.
-
-We modelled the probability of answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model.
-This allowed us to simultaneously estimate the impact of item- and participant-level predictors on the probability of word comprehension and production.
-We included $Age$, $Frequency$, $Phonemes$, $DoE$, and $Levenshtein$ as fixed effects.
-We also included the $Age \times DoE \times Levenshtein$ three-way interaction, and its corresponding two-way interactions.
-We added participants and items as crossed random effects including all intercepts and slopes when appropriate [@barr2013random].
-
-This model structure, including word- and participant-level properties to estimate vocabulary acquisition trajectories was discussed by @kachergis2022toward in the context of explanatory item response models.
-Our model differs from their proposal in that  1) while they modelled global trajectories exclusively, we modelled individual trajectories as well, 2) our outcome variable is ordinal in contrast to the binary outcome used in their study, 3) we propose a Bayesian approach.
-Our motivation for a Bayesian approach towards statistical inference is that (1) it allows us to incorporate previous domain knowledge to improve estimation accuracy, and (2) to quantify the uncertainty associated with the estimated parameters in our model [@mcelreath2018statistical].
-
-The Bayesian approach implies using Bayes’ rule to assign a probability to possible parameter values (i.e., posterior distribution), based on previous knowledge about such probability distribution (prior distribution) and our data (likelihood).
-We used Markov chain Monte Carlo (MCMC) to approximate the posterior distribution via sampling.
-We run 2 sampling chains with 4,000 iterations each, including 2,000 warm-up iterations per chain.
-All $\hat{R}$ values were smaller than 1.1, indicating that the chains converged successfully in all cases [@gelman1992inference].
-
-We compared a base model–-only including the intercept, $Age$, $Frequency$, $Phonemes$, and $DoE$ as fixed effects, and their corresponding random effects-–to increasingly more complex models, introducing predictors in the following order: $Age \times DoE$, $Levenshtein$, $DoE \times Levenshtein$, and finally $Age \times Levenshtein$ and $Age \times DoE \times Levenshtein$.
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO) [@vehtari2017practical]: we computed the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random sub-sample of `r dim(model_log_liks$model_fit_1)[1]` samples (out of `r format(dim(model_fit_1$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
-and selected the model with the smallest expected log posterior density (ELPD)^[Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with `r format(dim(model_log_liks$model_fit_1)[1], big.mark = ",")` posterior samples [@magnusson2019bayesian]].
-
-Finally, we explored the posterior distribution of each parameter in the selected model.
-Following @kruschke2018bayesian, We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
-We defined our *ROPE* from -`r round(median(abs(rope_coefs$estimate)), 3)` to +`r round(median(abs(rope_coefs$estimate)), 3)` [see @#appendix-rope for justification of the *ROPE*].
-In the results section, we report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
-This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
-We also report the proportion of the posterior samples of that coefficient that fell within the *ROPE*.
-A 90% overlap between the 95% *HDI* of the coefficients' posterior distribution and the *ROPE* indicates that, given the data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero, that is, negligible.
-
-Data processing and visualisation was done in R [@R-base] using the `tidyverse` family of packages [@R-tidyverse], Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@loo], and `tidybayes` [@tidybayes] R packages.
-
-# Results {#results}
-
-## Model comparison and selection {#comparison}
-
-@tbl-results-loos shows the outputs of the *LOO-CV*.
-The extended model (`model_fit_1`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference.
-This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.
-
-```{r}
-#| label: tbl-results-loos
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
-model_loos %>% 
-    loo_compare() %>% 
-    as.data.frame() %>%   
-    rownames_to_column("model") %>% 
-    relocate(
-        model, 
-        matches("elpd_loo"),
-        matches("p_loo"), 
-        matches("looic"), 
-        matches("diff")
-    ) %>%
-    gt() %>% 
-    tab_header(
-        title = "",
-        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
-The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_1`)")
-    ) %>% 
-    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
-    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>% 
-    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>% 
-    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>% 
-    fmt_number(2:9) %>% 
-    cols_label(
-        model = "Model",
-        # elpd
-        elpd_loo = md("*ELPD*"),
-        se_elpd_loo = md("*SE*"),
-        # p
-        p_loo = md("*p*"),
-        se_p_loo = md("*SE*"),
-        # looic
-        looic = md("*LOO-IC*"),
-        se_looic = md("*SE*"),
-        # diff
-        elpd_diff = md("*diff*"),
-        se_diff = md("*SE*")
-    ) %>% 
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners()
-    ) %>% 
-    tab_style(
-        cell_text(align = "left"),
-        cells_title(groups = "subtitle")
-    )
-```
-
-## Regression coefficients {#coefficients}
-
-```{r}
-#| label: tbl-results-fixed
-#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
-#| echo: false
-#| message: false
-#| warning: false
-post_draws <- gather_draws(model_fit_1, `b_.*`, regex = TRUE)
-
-# tidy predictor names
-str_repl <- c(
-    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
-    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
-    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
-    "b_freq_std" = paste0("Frequency (+1 SD, ",  round(sd(df$freq), 2), " Zipf)"),
-    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
-    "b_doe_std" = paste0("DoE (+1 SD, ",  percent(sd(df$doe)), ")"),
-    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
-    "b_doe_std:lv_std" = "DoE \u00d7 Levenshtein",
-    "b_age_std:doe_std" = "Age \u00d7 DoE",
-    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:doe_std:lv_std" = "Age \u00d7 DoE \u00d7 Levenshtein"
-)
-
-post_draws_list <- split(posterior_description, posterior_description$parameter)
-
-# summarise posterior draws
-posterior_description %>% 
-    mutate(
-        parameter = factor(
-            parameter,
-            levels = names(str_repl),
-            labels = str_repl
-        ) %>% 
-            as.character()
-    ) %>%
-    select(parameter, median, ci_high, ci_low, rope_percentage) %>%  
-    gt() %>% 
-    fmt_number(2:4) %>%
-    fmt_percent(5) %>%
-    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
-    cols_label(
-        parameter = "Parameter",
-        median = "Median",
-        ci_low = md("95% *HDI*"),
-        rope_percentage = md("*ROPE* prob.")
-    ) %>%
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = 1:5)
-    )
-```
-
-@tbl-results-fixed shows the summary of the posterior distribution of the fixed regression coefficients of the model.
-For interpretability, we supplement the summaries of each posterior regression coefficient with its transformation to the probability scale.
-The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
-This means that one logit difference is not necessarily translated to a unique value in the probability scale.
-For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
-The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
-We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
-derivative) of the logistic curve at the age at which most participants were acquiring a given word.
-This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
-
-$Age$ was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r printnum(post_draws_list[["b_age_std"]]$median, digits = 2)`, 95% *HDI* = [`r printnum(post_draws_list[["b_age_std"]]$ci_low, digits = 2)`, `r printnum(post_draws_list[["b_age_std"]]$ci_high, digits = 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month.
-Lexical frequency ($Frequency$) had a small, positive, but inconclusive on the probability of acquisition substantially, ($\beta$ = `r percent(post_draws_list[["b_freq_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_freq_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_freq_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_freq_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples overlapped with the *ROPE.* The 95% *HDI* of the regression coefficient of the number of phonemes ($Phonemes$) did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]).
-For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.
-
-```{r}
-#| label: fig-results-fixed
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 9
-#| fig-heigh: 7
-#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model.
-Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis).
-Intervals show the 95% highest density interval (*HDI*) of each distribution.
-This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-the mean and HDI limits are indicated below each distribution.
-Coefficients were transformed from the logit scale to the probability scale for interpretability.
-We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e.
-the derivative of the logistic function in its mid-point."
-post_draws %>%
-    mutate(
-        .variable_name = factor(
-            .variable, 
-            levels = names(str_repl),
-            labels = str_repl,
-            ordered = TRUE
-        )
-    ) %>% 
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    annotate(
-        geom = "rect",
+#| fig-width: 8
+#| fig-height: 6
+#| eval: true
+#| apa-cap: "Posterior marginal effects of Model 3.
+Thin lines correspond to 50 mean predictions generated from 50 MCMC samples of the posterior distribution of the model.
+Thicker lines indicate the median of those predictions.
+Different colours indicate different levels of phonological similarity (as indicated by the Cognateness similarity between pairs of translation equivalents).
+Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
+Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom.
+In-sample predictions lie inside the grey rectangles."
+
+clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")
+
+marginal_effects_epreds |> 
+    mutate(age = rescale_variable(age_std, 
+                                  mean(responses$age), 
+                                  sd(responses$age)),
+           exposure = factor(exposure_std,
+                             levels = unique(exposure_std),
+                             labels = c("Low exposure (-1 SD)",
+                                        "Mean exposure",
+                                        "High exposure (+1 SD)"),
+                             ordered = TRUE),
+           lv = factor(lv_std,
+                       levels = unique(lv_std),
+                       labels = paste0(c(0, 50, 100), "%"),
+                       ordered = TRUE),
+           .category = ifelse(.category=="Understands",
+                              "Comprehension",
+                              "Production")) |> 
+    ggplot(aes(age, .value, colour = lv, fill = lv)) +
+    facet_grid(.category~exposure) +
+    annotate(geom = "rect",
+             xmin = min(responses$age),
+             xmax = max(responses$age),
              ymin = -Inf,
              ymax = Inf,
-        xmin = rope_interval["lower"],
-        xmax = rope_interval["upper"],
-        colour = NA,
-        alpha = 0.5,
-        fill = "grey",
-    ) +
-    geom_vline(
-        xintercept = 0,
-        size = 1,
-        colour = "grey"
-    ) +
-    stat_slab(
-        aes(
-            fill = stat(abs(x) < rope_interval["upper"]),
-            colour = stat(abs(x) < rope_interval["upper"]),
-        ),
-        size = 0.25,
-        # fill = pal_d3()(2)[2],
-        position = position_nudge(y = 0.15)
-    ) +
-    geom_errorbar(
-        data = post_draws %>% 
-            filter(!grepl("Intercept|sd", .variable)) %>%
-            median_hdi(
-                .exclude = c(
-                    ".chain", 
-                    ".iteration",
-                    ".draw",
-                    ".row", 
-                    ".variable_name"
-                )
-            ) %>% 
-            mutate(
-                .variable_name = factor(
-                    .variable, 
-                    levels = names(str_repl),
-                    labels = str_repl,
-                    ordered = TRUE
-                )
-            ),
-        aes(xmin = .lower, xmax = .upper, x = .value),
-        width = 0.15
-    ) +
-    geom_point(
-        data = post_draws %>% 
-            filter(!grepl("Intercept|sd", .variable)) %>%
-            median_hdi(
-                .exclude = c(
-                    ".chain", 
-                    ".iteration",
-                    ".draw",
-                    ".row", 
-                    ".variable_name"
-                )
-            ) %>% 
-            mutate(
-                .variable_name = factor(
-                    .variable, 
-                    levels = names(str_repl),
-                    labels = str_repl,
-                    ordered = TRUE
-                )
-            ),
-        size = 2
-    ) +
-    geom_text(
-        data = post_draws %>% 
-            filter(!grepl("Intercept|sd", .variable)) %>%
-            mean_hdi(
-                .exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")
-            ) %>% 
-            mutate(
-                .variable_name = factor(
-                    .variable, 
-                    levels = names(str_repl),
-                    labels = str_repl,
-                    ordered = TRUE
-                )
-            ),
-        aes(
-            label = paste0(
-                printnum(.value, digits = 2),
-                " [", printnum(.lower, digits = 2), ", ",
-                printnum(.upper, digits = 2), "]"
-            )
-        ),
-        position = position_nudge(y = -0.25),
-        size = 3
-    ) +
-
-    labs(
-        x = "Coefficient estimate (logit scale)",
-        y = "Variable", 
-        fill = "Overlaps with ROPE",
-        colour = "Overlaps with ROPE"
-    ) +
-    scale_x_continuous(breaks = seq(-0.5, 2, 0.2)) +
-    scale_fill_manual(
-        values = c("dodgerblue", "#b3d9ff"),
-        labels = c("No", "Yes")
-    ) +
-    scale_colour_manual(
-        values = c("dodgerblue", "#b3d9ff"),
-        labels = c("No", "Yes")
-    ) +
-    theme(
-        legend.position = "top",
-        legend.justification = c(1, 1),
-        axis.title.y = element_blank(),
-        panel.grid.major.x = element_line(
-            colour = "grey85",
-            linetype = "dotted"
-        ),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-
-# annotate(
-#     geom = "text",
-#     label = "ROPE",
-#     x = 0.3,
-#     y = 10,
-#     vjust = 1,
-#     size = 5,
-#     colour = pal_d3()(2)[2]
-# ) +
-# annotate(
-#     geom = "curve",
-#     arrow = arrow(length = unit(0.2, "cm")),
-#     x = 0.3,
-#     y = 9.5,
-#     xend = 0.1,
-#     yend = 9,
-#     size = 1,
-#     curvature = -0.25,
-#     colour = pal_d3()(2)[2]
-# ) +
-```
-
-The degree of exposure to the language ($DoE$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std"]]$ci_high, accuracy = 0.01)`]).
-All of the posterior samples of his regression excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every 10% increase in exposure to the language the word belongs to, the participant was `r percent(post_draws_list[["b_doe_std"]]$median/(10*sd(df$doe)), accuracy = 0.01)` more likely to acquire it.
-The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
-However, the 95% *HDI* of the regression coefficient of the $Doe \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_doe_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the language it belonged to.
-Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
-This effect was negligible when exposure to both languages was balanced (e.g.
-50-50%).
-Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.
-
-The 95% *HDI* of the regression coefficient of the $Age \times DoE$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:doe_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of the degree of exposure on the speed at which participants acquired word across ages was inconclusive.
-Give our data, it is difficult to tell whether this *DoE* increased or decreased the slope of the words' acquisition trajectories.
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:doe_Std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:doe_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of the degree of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
-
-```{r}
-#| label: fig-results-marginal
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected posterior predictions.
-Posterior poreditions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
-These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model.
-Therefore, we generated 4,000 predictions for a series of combination of levels of interest.
-Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels.
-The X-axis indicates the age (in months) for which the prediction is generated.
-The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*).
-Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
-Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
-Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
-nd <- expand.grid(
-    n_phon_std = 0,
-    freq_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
-    doe_std = (c(0.1, 0.5, 0.9)-mean(df$doe))/(sd(df$doe))
-    # lv_std = (c(0, 0.5, 1)-mean(df$lv))/(sd(df$lv))
-)
-
-m <- add_epred_draws(nd, model_fit_1, ndraws = NULL, re_formula = NA) %>% 
-    mutate(
-        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
-        doe_std = factor(
-            doe_std, 
-            levels = unique(nd$doe_std), 
-            labels = paste0("DoE: ", c(10, 50, 90), "%")
-        )
-        # lv_std = factor(
-        #     lv_std, 
-        #     levels = unique(nd$lv_std),
-        #     labels = paste0(c(0, 50, 100), "%")
-        # )
-    ) %>% 
-    filter(.category != "No") %>% 
-    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
-    mutate(
-        .category = str_replace_all(
-            .category,
-            c(
-                "Understands and Says" = "Production\n(Understands and Says))",
-                "Understands" = "Comprehension\n(Understands)"
-            )
-        )
-    )
-
-d <- df %>%
-    mutate(
-        doe_std = case_when(
-            between(doe, 0.00, 0.20) ~ "DoE: 10%",
-            between(doe, 0.40, 0.60) ~ "DoE: 50%",
-            between(doe, 0.80, 1.00) ~ "DoE: 90%"
-        ) %>% 
-            factor(., levels = paste0("DoE: ", percent(c(0.1, 0.5, 0.9)))),
-        # lv_std = case_when(
-        #     between(lv, 0.00, 0.20) ~ "0%",
-        #     between(lv, 0.40, 0.60) ~ "50%",
-        #     between(lv, 0.80, 1.00) ~ "100%"
-        # ) %>% 
-        #     factor(., levels = paste0(percent(seq(0, 1, 0.5)))),
-        age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
-            cut(
-                .,
-                breaks = ((seq(10, 38, 2)-mean(df$age))/(sd(df$age))),
-                labels = ((seq(10, 36, 2)-mean(df$age))/(sd(df$age))),
-                include.lowest = TRUE
-            ) %>% 
-            as.character() %>% 
-            as.numeric(),
-        Understands = response %in% c("Understands", "Understands and Says"),
-        `Understands and Says` = response %in% "Understands and Says"
-    ) %>% 
-    drop_na(doe_std) %>% 
-    group_by(te, age_std, doe_std) %>% 
-    summarise(
-        Understands = mean(Understands),
-        `Understands and Says` = mean(`Understands and Says`),
-        .groups = "drop"
-    ) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
-    mutate(
-        .category = str_replace_all(
-            .category,
-            c(
-                "Understands and Says" = "Production\n(Understands and Says))",
-                "Understands" = "Comprehension\n(Understands)"
-            )
-        )
-    )
-
-m %>% 
-    ggplot() +
-    aes(
-        x = age_std, 
-        y = .epred, 
-        colour = doe_std,
-        fill = doe_std
-    ) +
-    facet_grid(~.category) +
-    stat_lineribbon(
-        size = 1,
-        alpha = 0.5,
-        .width = 0.95, 
-        colour = NA,
-        point_interval = mean_hdi
-    ) +
-    stat_summary(
-        data = d,
-        fun.data = "mean_se",
-        geom = "pointrange",
-        size = 0.25
-    ) +
-    # geom_point(
-    #     data = d,
-    #     position = position_dodge(width = 0.25),
-    #     size = 0.5,
-    #     alpha = 0.5
-    # ) +
-    stat_summary(fun = "mean", geom = "line", size = 1) +
-    scale_x_continuous(
-        breaks = (seq(10, 36, 2)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 2)
-    ) +
-    scale_color_d3() +
+             fill = "grey95",
+             colour = NA) +
+    annotate(geom = "text",
+             x = mean(responses$age),
+             y = 0,
+             hjust = -0.1,
+             vjust = 0,
+             label = "Observed age range",
+             size = 2) +
+    geom_vline(xintercept = mean(responses$age), 
+               linetype = "dotted") +
+    geom_hline(yintercept = 0.5,
+               linetype = "dotted") +
+    geom_line(aes(group = interaction(lv, .draw)),
+              linewidth = 0.5,
+              alpha = 1/10) +
+    stat_summary(fun = median, geom = "line", linewidth = 0.75) +
     labs(
         x = "Age (months)",
-        y = "Posterior probability of acquisition\n(Linear prediction)",
-        colour = "Degree of exposure",
-        fill = "Degree of exposure"
-    ) +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
-    theme(
-        legend.position = "top",
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        strip.background = element_rect(fill = "grey", colour = NA),
-        plot.background = element_rect(fill = "white", colour = NA)
-    )
-```
-
-# Discussion {#discussion}
-
-## Summary of the study
-
-This study explored how word acquisition was affected by cognateness in bilingual toddlers.
-We collected vocabulary data from children learning Catalan and/or Spanish in differing degrees of exposure to each, and then measured the probability of each word being acquired at different ages as a function of (1) participants' exposure to the language it belongs to, and (2) its phonological overlap with its translation equivalent in the other language.
-We build a multilevel Bayesian model that estimated the effect of both predictors and their interaction, adjusting for other variables known to be relevant during the acquisition of words: participants' age, lexical frequency, and length in phonemes.
-
-## Summary of the results
-
-## Results in contrast with the literature
-
-## Mechanistic explanations
-
-## Limitations
-
-## Further steps
-
-# References {#references}
+        y = "Probability of acquisition",
+        linetype = "Cognateness (phonological similarity)",
+        colour = "Cognateness (phonological similarity)",
+        fill = "Cognateness (phonological similarity)") +
+    scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_fill_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_y_continuous(breaks = seq(0, 1, 0.25),
+                       limits = c(0, 1)) +
+    scale_x_continuous(breaks = seq(0, 50, 5)) +
+    theme(legend.position = "top",
+          legend.key.size = unit(1, "cm"),
+          legend.key.height = unit(1, "cm"),
+          axis.text.x = element_text(size = 9),
+          panel.grid = element_blank(),
+          panel.border = element_rect(fill = NA, 
+                                      colour = "black", 
+                                      linewidth = 0.75),
+          plot.subtitle = element_textbox_simple(fill = "grey90",
+                                                 padding = margin(8, 8, 8, 8),
+                                                 margin = margin(0, 0, 10, 0))) +
+    guides(colour = guide_legend(override.aes = list(linewidth = 2)))
+```
+
+# Discussion {#sec-discussion}
+
+The present study explored the cognateness facilitation effect on bilingual word acquisition.
+We found such effect to be mediated by lexical frequency and degree of language exposure: phonologically similar translation equivalents were on average acquired earlier in age compared to phonologically dissimilar translation equivalents, but this effect was only observed for word-forms to which participants were exposed less often, i.e., words with low lexical frequency and/or words from the language of lower daily exposure.
+Our results are relevant in the light of two previous findings.
+First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (which shared many cognates, e.g., English-German) showed larger vocabulary sizes than those learning typologically more distance languages (which shared fewer cognates, e.g.,  English-Mandarin).
+Second, @mitchell2022cognates found an earlier age-of-acquisition for cognates, compared to non-cognates.
+The outcomes of both studies pointed to cognateness facilitating word acquisition, but the underpinnings of such effect were unclear.
+
+In this study, we capitalised on accumulator models of language acquisition to propose a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
+We suggested that words benefit from their cognate status only *after* their translation equivalent has been acquired: after the child acquires a word-form, the number of learning instances they need to acquire its translation decreases.
+We further hypothesised that such facilitation is proportional to the phonological similarity between both word-forms.
+If two translation equivalents sharing a fair deal of phonological similarity (e.g., /\textipa{"pOR.t@}/-/\textipa{"pweR.ta}/, `r percent(stringdist::stringsim("pORt@", "pweRta"))` Levenshtein similarity), learning of one of them should facilitate the acquisition of the other more strongly, compared to when both word-forms share only few phonemes (e.g., /\textipa{k@"Di.R@}/-/\textipa{"si.\textlambda a}/ , `r percent(stringdist::stringsim("DiR@", "siLa"))`).
+
+Under such scenario, we predicted that words from the language of lower exposure would benefit more often from their cognate status than those from the language of higher exposure.
+The rationale behind such prediction was that children are less frequently exposed to word-forms in the former language than to words in the latter [@cattani2014how; @thordardottir2006bilingual].
+Words in the language of lower exposure are therefore more likely to be acquired later in age than their translations in the other language.
+For instance, an infant exposed to Catalan and Spanish 65% and 35% of the time, respectively, is more likely to learn the Spanish word-form /\textipa{"ga.to}/ later in age than its Catalan translation /\textipa{gat}/.
+If this is the case, /\textipa{"ga.to}/ is more likely to receive facilitation from the acquisition of /\textipa{gat}/ than *vice versa* (see {apafg-hypotheses} for a graphic illustration).
+In line with these predictions, we found that cognateness facilitated word acquisition more strongly in words from the language of lower exposure.
+This asymmetric facilitation effect provides support for the hypothesis that words benefit from their cognate status *after* their translation has been acquired.
+Overall, our results provide support for an interactive account of bilingual vocabulary learning in which word representations from one language contribute to the acquisition of words in the other.
+
+These findings are, to a large extent, in line with previous literature on word acquisition.
+First, words-forms with fewer phonemes like /\textipa{sO\textltilde}/ (Catalan for *sun*) were on average more likely to be acquired than word-forms with more phonemes like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*).
+These results converge with previous studies found word length to be more strongly associated with receptive vocabulary than with expressive vocabulary [e.g., @braginsky2019consistency; @jones2019children].
+This facilitation is most likely due to the fact that shorter word-forms are generally simpler to encode as phonological representations in the lexicon than longer word-forms.
+Shorter word-forms are easier to produce, and therefore to be reported as acquired in parental vocabulary checklists.
+
+Our participants were also more likely to know words to which they were more frequently exposed to, replicating previously reported effects of lexical frequency [e.g., @braginsky2019consistency], and bilingual language exposure [e.g., @cattani2014how; @thordardottir2006bilingual].
+This strong effect of word frequency is a key component of accumulator models of word acquisition, which suggest that the age-of-acquisition of a word is, to a large extent, a function of the child's rate of exposure to such word, which conspires with child- and word-level factors in the process of forming a representation of the word in the child's lexicon [@mollica2017how; @hidaka2013computational].
+Our results provide ground for the central role of word frequency in accumulator models to be extended to the bilingual case, in the form of a composite measure of word rate of exposure that accounted for both lexical frequency and language exposure [@cattani2014how; @floccia2018results].
+Interestingly, our findings pointed to this predictor playing a stronger role during earlier ages: younger infants benefited from higher exposure rates at earlier ages, suggesting that early word acquisition is supported by mere exposure to word forms more strongly than later word acquisition.
+
+Finally, we found a frequency-mediated facilitation effect of cognateness, adding some nuance to the findings from previous studies [@floccia2018results; @mitchell2022cognates; @bosch2014first; @schelletter2002effect].
+Only words to which children were exposed to less frequently benefited from their cognate status.
+These results favour the account that a word-form benefits from its phonological similarity with its translation only after such translation has been acquired first.
+On the one hand, our results provide converging evidence for @floccia2018results, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample.
+Most children in their sample were English-dominant, and therefore their exposure to English was higher than their exposure to the additional language.
+However, it was in the vocabulary of the additional language where the authors found a facilitation effect of cognateness: bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese).
+Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018results's sample.
+
+Overall, our findings parallel cognates' significant role in many models of bilingual language processing.
+The importance of cognates in these model stems in the fact that cognates provide evidence for a language non-selective account of lexical processing, in which bilinguals activate unselected lexical representations in both languages, even in fully monolingual situations [@thierry2007brain; @christoffels2007bilingual].
+As mentioned in the introduction, naming pictures whose labels are cognates leads to shorter latencies than when labels are non-cognates, which suggests that phonological similarity across translations facilitates production [@costa2000cognate; @hoshino2008cognate].
+When acting as distractors in naming tasks, cognates exert larger interference effects than non-cognate distractors [@hoshino2011language; @colome2010which].
+Both findings suggest that the non-selected translation of the target in @costa2000cognate, or of the distractor in @colome2010which was activated, and that when such word-form was phonologically related to its translation, the facilitation or interference effect was stronger.
+
+Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extend to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
+For instance, @thierry2007brain found neurophysiological evidence of Chinese-English bilinguals activating both languages in a word comprehension task.
+In this experiment, participants were presented with couples of words in English, and were asked to respond if the words were related in meaning.
+Unbeknownst to participants, the translations to Chinese of half of such word pairs shared a character.
+Participants would be presented (visually in a first experiment and auditorily in a second experiment) with an English word pair 'train-ham', which are semantically unrelated, but share a character in their Chinese translations (\foreignlanguage{chinese-simplified}{火車} [Huo Che]-\foreignlanguage{chinese-simplified}{火腿} [Huo Tui]).
+Bilinguals' brain signal differed across words pairs with and without translation overlap (e.g., 'apple-table', \foreignlanguage{chinese-simplified}{平果} [Ping Guo]–\foreignlanguage{chinese-simplified}{桌子} [Zhuo Zi]).
+This suggests that participants were sensitive to the form-similarity of the Chinese translation of the word-pairs they were presented, despite the whole experimental task taking place in English.
+
+Bilingual toddlers and children---still in the process to developing a mature lexicon---also show evidence of parallel activation during comprehension [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic].
+Whether---and if so, how---the language non-selectivity shapes the early bilingual lexicon is still unclear.
+Previous work has pointed to an interactive account of lexical acquisition, in which toddlers' lexical representations in one language influence the acquisition of words in the other language.
+The acquisition of translation equivalents embodies a notable divergence between monolingual and bilingual developmental trajectories.
+For instance, monolingual word acquisition benefits from the semantic relatedness between the words the child learns [@fourtassi2020growth; @hills2009longitudinal], but is inhibited during the acquisition of words from the same language with the *equivalent* meaning, i.e.
+synonyms.
+This learning heuristic, called *mutual exclusivity*, is one of many that play a relevant role in models of monolingual word learning [@golinkoff1994early], acting as a constrainer of the hypothesis space for all possible word-referent associations: if a referent already holds a label, any label uttered in the presence of such referent is associated with a different referent [e.g., @markman1990constraints].
+Although this heuristic seems useful in the monolingual case (children are rarely exposed to synonyms), it would come with a huge cost for bilinguals, how are frequently exposed to more than one label for the same referent, one in each language.
+It was initially predicted that bilinguals would follow *mutual exclusivity*, and therefore avoid learning translation equivalents [@volterra1978acquisition].
+
+This prediction has been robustly contested by two main sources of evidence.
+First, experimental work suggests that bilinguals may not apply *mutual exclusivity* during word learning, or at least not as consistently monolinguals do [@byers-heinlein2009monolingual; @byers-heinlein2013lexicon].
+Second, detailed analyses of bilingual children's vocabulary contents shows that bilingual children do acquire translation equivalents from early ages [e.g., @pearson1995crosslanguage].
+In fact, translation equivalents may even be overrepresented in the early bilingual lexicon.
+For instance, @bilson2015semantic found that bilingual toddlers produced a higher proportion of translation equivalents than it would be expected if they learned two independent lexica simultaneously, which suggests that bilingual children preferentially acquired words for which they already knew a label in the other language.
+In the same line, @bosch2014first reported a similar overrepresentation of translation equivalents in Catalan-Spanish bilinguals: a high proportion of acquired Catalan words held an also acquired translation in Spanish.
+Finally, a recent investigation by @tsui2022area presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
+Overall, these findings suggest that bilingual children benefit from a semantic facilitation effect driven by the acquisition of translation equivalents.
+This is a substantial divergence from monolingual trajectories of lexical acquisition.
+Our study provides further evidence for a pivotal role of translation equivalents, showing that, especially in the case of cognates, translation equivalents, are not only part of bilinguals' vocabulary contents, but are also acquired close in age, depending primarily on children's degree of exposure to each language.
+
+The divergent trajectories of bilingual lexical acquisition go beyond semantics.
+As highlighted before, the phonological overlap between the languages being learned seems to impact vocabulary growth in bilinguals.
+Bilinguals learning languages that share many cognates show larger vocabulary sizes than those learning two languages sharing fewer cognates [@floccia2018introduction; @blom2020crosslanguage; @bosma2019longitudinal].
+This finding can be explained to a large extent by the fact that cognates seem to be acquired at earlier ages than non-cognates [@mitchell2022cognates].
+This phonological facilitation effect mimics previous findings in monolinguals, who seem to preferentially learn phonologically similar words.
+For instance, @jones2019children found that English monolingual toddlers show a facilitation effect when acquiring words from dense phonological neighbourhoods: a larger number of similar-sounding words was associated with an increase in likelihood of a word form being part of a toddler's productive vocabulary.
+A network analysis by @fourtassi2020growth found evidence in favour of a phonological (and semantic) preferential acquisition of new words in monolingual children: words with higher phonological relatedness to other words were more likely to be acquired.
+Again, these findings contrast with those in bilingual children in that the facilitation effect or preferential acquisition mechanisms described are based on phonological similarity *without* translation equivalence.
+The amount of similar-sounding words in the child input or their acquired vocabulary predicts a larger probability of acquisition of new words with *different* meaning.
+In the case of bilinguals, there is evidence of such phonological facilitation at the level of translation equivalents, as shown by an earlier acquisition of cognates than non-cognates in English-French [@mitchell2022cognates] and in the Catalan-Spanish bilinguals in the present study.
+Evidence of a cross-language facilitation effect between non-translation equivalents is to date absent.
+
+We consider three alternative explanations for our results.
+Regarding the cognate facilitation effect, it is possible that when hearing cognate translation equivalents, bilingual children accept both word forms as acceptable variations of the *same* word.
+This contrasts with monolinguals' high sensitivity to mispronunciations during word recognition [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological].
+It might be the case that every exposure to either of the word-forms of the translation equivalent is counted as a learning instance for both.
+This would impact the acquisition mechanism in such way that any cognate effect is not explained by a reduction of the number of learning instances needed to acquired a word (after its translation has been acquired), but by a faster increase in cumulative learning instances across ages, leading to a similar outcome: an earlier age-of-acquisition for cognates.
+Supporting such alternative explanation, @ramon-casas2009vowel examined 18-26 months-old Catalan-Spanish bilingual toddlers' sensitivity to mispronunciations in a word-recognition task and found that, contrary to their Catalan monolingual peers, bilinguals seemed to recognise mispronounced words, disregarding vowel replacements involving the Catalan vowel contrast /e/-/\textepsilon/.
+The same population of bilingual infants shows sensitivity to such contrast before 12 months of age in phoneme discrimination tasks [@bosch2003simultaneous]---which do not involve word-referent associations---suggesting that bilingual infants dismiss subtle changes in phonological forms when presented with their referents.
+Using the same procedure, @ramon2010non tested bilingual infants' recognition of mispronunciations involving the /e/-/\textepsilon/ contrast embedded in exclusively *non-cognate* word-forms.
+The materials in this study contrasted with those in @ramon-casas2009vowel in that the latter included almost 60% of cognate words in their stimuli list.
+@ramon2010non hypothesised that bilinguals' frequent exposure to cognates---which often involve vowel replacements between Catalan and Spanish word-forms---may have led bilinguals to be less alert to such changes when occurring within similar word-forms.
+With this new set of stimuli, including non-cognate words only, bilinguals showed monolingual-level sensitivity to mispronunciations involving the /e/-/\textepsilon/ contrast.
+Overall, these results suggest, at face value, that that bilinguals may to some degree encode similar-sounding word-forms with shared referent as into a common lexical representation, especially when the word-forms involved are cognates.
+
+```{r discussion-examples}
+#| echo: false
+#| warning: false
+#| message: false
+#| out-width: 60%
+lv_distr <- items |> 
+    mutate(lv_bin = cut(lv, 
+                        breaks = seq(0, 1, 0.1),
+                        include.lowest = TRUE,
+                        labels = c("0-10%", "10-20%", "20-30%", "30-40%",
+                                   "40-50%", "50-60%", "60-70%", "70-80%",
+                                   "80-90%", "90-100%"))) |>
+    add_count(name = "n_total") |> 
+    count(lv_bin, n_total) |> 
+    mutate(prop = n/n_total)
+```
+
+```{r discussion-pool}
+#| echo: false
+#| warning: false
+#| message: false
+E_e <- bvq_data$pool |>
+    mutate(item = gsub("cat_|spa_", "", item),
+           sampa = gsub("\\.", "", sampa)) |>
+    filter(te %in% items$te, 
+           item %in% items$item) |>
+    select(te, language, sampa) |> 
+    pivot_wider(names_from = language,
+                values_from = sampa, 
+                id_cols = te,
+                names_repair = janitor::make_clean_names) |>
+    filter(grepl("E", catalan), 
+           grepl("e", spanish)) |>
+    left_join(distinct(items, te, lv),
+              by = "te")
+```
+
+We identify three main reasons why bilingual children are unlikely to conflate similar sounding translations into the same lexical representations.
+First, when @ramon-casas2009vowel probed Catalan-Spanish bilinguals using the /e/-/i/ contrast---present in both Catalan *and* Spanish---in the same task, bilinguals showed monolingual-level sensitivity to mispronunciations, suggesting that the conflation of variations of the same word-form into a similar lexical representation may be specific to some language-specific contrasts, and perhaps associated with the acoustic distance between the phonemes involved.
+The /e/-/\textepsilon/ contrast in which Catalan-Spanish bilinguals failed to mispronunciations, is particularly problematic this regard: both vowels differ in only one acoustic feature (/e/ is closed-mid, while /\textepsilon/ is open-mid) and is only present in Catalan.
+Our results are unlikely to be explained by the possibility that Catalan-Spanish bilingual toddlers form common lexical representations for cognate words.
+Second, few phoneme substitutions that separate Catalan word-forms from their cognate Spanish translations involve the /e/-/\textepsilon/ contrast.
+Rather, due to vowel reduction, Spanish and Catalan cognates predominantly involve the /a/-/\textschwa/, /e/-/\textschwa/, /\textepsilon/ and /o/-/u/ contrasts [@vigario2009phonetics].
+Only `r percent(nrow(E_e)/length(unique(items$te)))` translation equivalents in our questionnaire involved a /e/-/\textepsilon/ change between the Catalan and the Spanish items, only one of such was a minimal pair: /\textipa{tREn}/-/\textipa{tRen}/ [*train*].
+
+Second, few words in our study differ only in one phoneme substitution.
+Only `r percent(lv_distr$prop[nrow(lv_distr)], accuracy = 0.01)` of the translation equivalents included in our analyses shared between 90 and 100% phonological similarity (measured as the Levenshtein similarity between the X-SAMPA transcriptions of the Catalan and Spanish form), and only `r percent(lv_distr$prop[nrow(lv_distr)-1], accuracy = 0.01)` shared between 80 and 90% similarity.
+This shows that the presence of identical and near-identical cognates in our study is very low, and therefore most translation equivalents differed in more than one phoneme.
+In summary, even if any cognate involved the /a/-/\textepsilon/ substitution (or any phonetically close contrast that bilingual might disregard at the lexical level), such is very likely to have occurred in the context of a sufficiently distinct word-form to be distinguished and treated and relevant at the lexical level.
+Third, and perhaps more critically, bilinguals are able to learn new minimal pairs involving the /e/-/\textepsilon/ contrast, suggesting that their tolerance towards cognate mispronunciations might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@ramon-casas2017minimalpair].
+
+It cannot be neglected a potential bias by caregivers when filling the questionnaire, in which the likelihood of two translation equivalents being present in a child's lexicon results not as a function of such phonological facilitation effect, but by caregivers assuming that because the child knows a word in one language, the child must also know the word in the other language.
+This bias might be especially likely when both words are similar-sounding (i.e., cognates).
+Although there is little possibility to test this account, it is worth noting that parents were explicitly instructed *not* to rely on their responses to the words in one languages when responding to words in the other language.
+The potential presence of responder biases is pervasive across many measurement instruments that rely on caregivers' reports of the developmental status of the child.
+The outcomes of such instruments, like the one used in the present study, are susceptible to biases induced by the child themselves (i.e., assigned gender, temperament), their caregivers (e.g., time spent with the child, preconceptions about language acquisition), or even the researchers (e.g., the items included might over-rely on canonical realisations of words), which push against evidence of validity of the questionnaire as an instrument of the developmental status of the child.
+The findings in this investigation should be weighted according to the aforementioned limitations.
+
+Our results should also be taken with caution if generalised to other bilingual populations, given the particular typological similarity and sociolinguistic status of the languages being learned by our participants.
+Catalan and Spanish are both Romance languages, share a considerable amount of cognates, and very often co-exist in the same social spaces (families, schools, working place, etc.).
+Although there are no populations of bilinguals with identical sociolinguistic characteristics, and therefore it would be pointless to compare the sample in this study to any "standard" population of bilinguals, the generalisability of our results should be taken with a grain of salt.
+It is possible, for instance, that the large amount of cognates that Catalan-Spanish bilingual children are exposed leads them to learn at early stages of word acquisition that words tend to sound similar in both languages, and therefore to be more keen on assuming that similar sounding words are semantically related.
+Children exposed to a pair of languages sharing a fewer amount of cognates might not make the same assumption, even when the word-forms they hear in the presence of the same referents are very similar at the phonological level.
+
+Another limitation of the present study is the predominantly cross-sectional design of our data set.
+We collected vocabulary data from children aged 12 to 32 months, a critical period in development in which the foundations of a lexicon get established.
+We estimated word acquisition curves by modelling the probability of comprehension and production across ages.
+However, most participants provided data at only one age point.
+Effectively, we are making longitudinal inferences based on cross-sectional data, and drawing conclusions about the development of a single participant across multiple age points using data from different participants at individual age points.
+For this reason, and also because of the observational nature of our study, causal inferences derived from our analyses are unwarranted [@kraemer2000how; @kesmodel2018crosssectional].
+One counterpoint to this shortcoming is that, while data our data is cross-sectional if participants are considered as the observational unit, data about word-forms/translation equivalents is longitudinal.
+The same translation equivalent is responded to at multiple ages.
+Since translation equivalents are our main observational unit of interest over which we primarily conduct statistical inference (e.g., estimating the size of the cognateness-effect), this provides some further justification for our conclusions.
+
+Future studies addressing this subject may benefit for a more adequate design oriented towards making causal inferences.
+One such design can be partially satisfied by (1) collecting a more substantial longitudinal dataset, in which the same participant provides data points for each translation equivalent at multiple ages, and (2) testing the role of cognateness and word exposure in an experimental context, in which such variables can be orthogonally manipulated or randomised.
+This second possibility would also allow to explore an interpretation of the results obtained in the context of *word learning*, which is not possible in the present study.
+Another consideration for future investigations is the involvement of a more diverse range of language pairs as in @braginsky2019consistency or @floccia2018introduction to improve the generalisability of the resulting inferences across more populations of bilinguals.
+
+In summary, our study has replicated previous studies and provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
+We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word's phonological similarity with its translation in the other language.
+We suggest that this effect might be explained by the need of a robust lexical representation of one of the translations for cognateness to play a role.
+
+# References {#sec-references}
 
 ::: {#refs}
+
 :::
 
-#  Appendix {#appendix}
-
-## Appendix 1: Words {#appendix-words}
-
-```{r}
-#| lab: tbl-appendix-words
-#| tbl-cap: "Words included in the analysis"
-#| echo: false
-#| message: false
-#| warning: false
-#| eval: false
-items %>% 
-    left_join(distinct(multilex_data$pool, item, te, label)) %>% 
-    mutate(
-        language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish"),
-        label = paste0(label, " /", ipa_flat, "/"),
-        te = as.integer(te),
-        n_phon = as.integer(n_phon)
-    ) %>% 
-    select(language, te, label, freq, n_phon, lv) %>% 
-    pivot_wider(
-        id_cols = c(te,lv),
-        names_from = language,
-        values_from = c(label, freq, n_phon), 
-        values_fn = first
-    ) %>% 
-    clean_names() %>% 
-    relocate(te, ends_with("catalan"), ends_with("spanish"), lv) %>% 
-    arrange(label_catalan) %>% 
-    select(-te) %>% 
-    gt() %>% 
-    fmt_number(starts_with("freq")) %>% 
-    fmt_integer(starts_with("n_phon")) %>% 
-    fmt_percent(lv) %>% 
-    tab_spanner("Catalan", ends_with("catalan")) %>%
-    tab_spanner("Spanish", ends_with("spanish")) %>%
-    cols_label(
-        label_catalan = "Word-form /IPA/",
-        freq_catalan = "Frequency",
-        n_phon_catalan = "Phonemes",
-        label_spanish = "Word-form /IPA/",
-        freq_spanish = "Frequency",
-        n_phon_spanish = "Phonemes",
-        lv = "Levenshtein"
-    ) 
-```
-
-## Appendix 2: *ROPE* specification {#appendix-words}
-
-We decided to draw the *ROPE* interval around zero so that it covered values smaller the absolute median of those relevant by similar previous studies.
-After consulting the regression tables of @mitchell2022cognates, @braginsky2019consistency, and @laing2022phonological, we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: `r round(median(abs(rope_coefs$estimate)), 3)`, in the logit scale.
+{{< pagebreak >}}
+
+# Appendix A: model predictors {.appendix .unnumbered}
+
+{{< include appendix/_model-predictors.qmd >}}
+
+{{< pagebreak >}}
+
+# Appendix B: model details {.appendix .unnumbered}
+
+{{< include appendix/_model-details.qmd >}}
+
+{{< pagebreak >}}
+
+# Appendix C: convergence diagnostics {.appendix .unnumbered}
+
+{{< include appendix/_model-diagnostics.qmd >}}
+
+{{< pagebreak >}}
+
+# Appendix D: frequency and language exposure as separate predictors {.appendix .unnumbered}
+
+{{< include appendix/_model-doe.qmd >}}
+
+{{< pagebreak >}}

### 2023-04-17 manuscript.tex
Fix comments and use apaquarto extension

--- 2022-09-23 manuscript.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
 \documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
-
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{setspace}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
+  man,
+  floatsintext,
+  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue,biblatex]{apa7}
+
+% TODO: Add custom LaTeX header directives here
 \makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+	{-.5em}%
+	{\normalfont\normalsize\bfseries\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
+	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+	{-\z@\relax}%
+	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
 \makeatother
-\usepackage{xcolor}
-\usepackage[top=30mm,left=20mm,heightrounded]{geometry}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
 \usepackage{amsmath}
@@
 \usepackage{caption}
-\usepackage{longtable}
-\KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
+\usepackage{longtable}\usepackage{lineno}\linenumbers\usepackage{tipa}\usepackage[pinyin]{babel}\usepackage{CJKutf8}\usepackage{fontspec}\babelprovide[script=CJK, language={Chinese Simplified}]{chinese-simplified}\babelfont[chinese-simplified]{rm}[Ligatures = Common]{Noto Serif CJK SC}\babelfont[chinese-simplified]{sf}[Ligatures = Common]{Noto Sans CJK SC}\makeatletter\makeatother\makeatletter\makeatother\makeatletter\@ifpackageloaded{caption}{}{\usepackage{caption}}\AtBeginDocument{%
 \ifdefined\contentsname
@@
 \fi
-}
-\@ifpackageloaded{float}{}{\usepackage{float}}
+}\@ifpackageloaded{float}{}{\usepackage{float}}
 \floatstyle{ruled}
 \@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
+\floatname{codelisting}{Listing}\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}\makeatother\makeatletter\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}\makeatother\makeatletter\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}\makeatother\makeatletter\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}\makeatother\makeatletter\makeatother\makeatletter\makeatother
+
+\usepackage[style=apa,backend=biber]{biblatex}
+\addbibresource{assets/references.bib}
+
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
-  pdfauthor={Gonzalo Garcia-Castro; Daniela S. Ávila-Varela; Ignacio Castillejo; Núria Sebastian-Galles},
-  pdfkeywords={lexical acquisition, vocabulary, bilingualism},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-
-\title{The role of cross-linguistic lexical similarity on bilingual word
-acquisition\thanks{This research was supported by grants from the
-Spanish Ministerio de Ciencia, Innovación y Universidades
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+\title{Lexical frequency mediates the facilitation effect of cognateness
+in bilingual lexical acquisition}
+\shorttitle{Frequency and cognateness in lexical acquisition}
+
+\authorsnames[{1},{1,2},{3},{1}]{
+Gonzalo Garcia-Castro,Daniela S. Avila-Varela,Ignacio Castillejo,Núria
+Sebastian-Galles
+}
+
+\authorsaffiliations{
+{Center for Brain and Cognition, Universitat Pompeu Fabra},{Facultade de
+Letras, Universidade de Lisboa},{Departamento de Psicología, Universidad
+Autónoma de Madrid}}
+
+\date{}
+\abstract{Language non-selectivity is prominent feature of bilingual
+lexical processing. An instance of such non-selectivity is embodied by
+cognateness, which recent studies have suggested to facilitate
+vocabulary acquisition in bilingual toddlers, who show larger vocabulary
+sizes when their languages share many cognates, and to be acquired at
+earlier ages than non-cognates. The specific mechanisms behind such
+facilitation are unclear. We present an account of early bilingual
+vocabulary growth in which cognateness interacts with lexical frequency
+and language exposure to facilitate the word acquisition. We evaluated
+this model against vocabulary data from 436 Catalan-Spanish bilinguals
+aged 12 to 34 months. We used Bayesian Exploratory Item Response Models
+to estimate participants' probability of acquisition of 604 words,
+conditional to the cognate status and lexical frequency of the
+word-form, and the age and degree of exposure to each language of the
+toddler. We found converging evidence for an earlier age-of-acquisition
+for cognate words, and for such effect being mediated by lexical
+frequency and language exposure. Low-frequency words, and words from the
+language of least exposure were more strongly benefitted by their
+cognate status than high-frequency words. Overall, our findings support
+an interactive account of bilingual vocabulary acquisition in which the
+lexical representations in one language interact with the acquisition of
+words in the other language.}
+\addbibresource{assets/references.bib}
+
+\keywords{lexical acquisition, vocabulary, bilingualism, item response
+theory, bayesian}
+
+\authornote{\par{\addORCIDlink{Gonzalo
+Garcia-Castro}{0000-0002-8553-4209}}\par{\addORCIDlink{Daniela S.
+Avila-Varela}{0000-0002-3518-8117}}\par{\addORCIDlink{Ignacio
+Castillejo}{0000-0001-7445-0416}}\par{\addORCIDlink{Núria
+Sebastian-Galles}{0000-0001-6938-2498}}
+\par{ }
+\par{   The authors declare no conflicts of interest with regard to the
+funding source of this study. This research was supported by grants from
+the Spanish Ministerio de Ciencia, Innovación y Universidades
 (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government
@@
 fellowship of the Spanish Ministerio de Ciencia, Innovación y
-Universidades (FPI 2019). The authors declare no conflicts of interest
-with regard to the funding source of this study. We are grateful to
-Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and
-Ege E. Özer for their help-ful feedback. We thank Xavier Mayoral, Silvia
-Blanch, and Cristina Cuadrado for their technical support. We also thank
-Cristina Dominguez and Katia Pistrin for their efforts in recruiting
-infants. We would like to thank the clinics Quirón and Sagrada Familia
-that allowed us to recruit participants in their premises. We also thank
-all families and infants who participated in the experiments. Data
-collection was half-way when the COVID-19 pandemic started. We would
-like to pay special tribute to the families that collaborated with us
-under these difficult circumstances.}}
-\author{Gonzalo Garcia-Castro \and Daniela S. Ávila-Varela \and Ignacio
-Castillejo \and Núria Sebastian-Galles}
-\date{}
+Universidades (FPI 2019). We are grateful to Chiara Santolin, Alicia
+Franco-Martínez, Cristina Rodríguez-Prada, Ege E. Özer, and the rest of
+the Speech Acquisition and Perception research group for their helpful
+feedback. We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado
+for their technical support. We also thank Cristina Dominguez and Katia
+Pistrin for their efforts in recruiting infants. We also thank all
+families and infants who participated in the experiments. }
+\par{Correspondence concerning this article should be addressed to Gonzalo
+Garcia-Castro, Center for Brain and Cognition, Universitat Pompeu
+Fabra, Ramon Trias Fargas
+25-27, Barcelona, Spain 8005, Email: gonzalo.garciadecastro@upf.edu}}
 
@@
 \maketitle
-\begin{abstract}
-Bilinguals face the challenging task of learning words from languages
-with overlapping phonologies. Floccia et al.~(2018) reported larger
-vocabulary sizes for 24-month-old bilinguals that were learning
-languages that shared a greater amount of cognates (e.g.,
-English-Dutch). The mechanisms underlying this effect remain unknown. We
-explore two compatible scenarios. First, we test whether cognates are
-learnt earlier than non-cognates. This would account for the difference
-in vocabulary size associated to the amount of shared cognates across
-languages. Second, we explore the possibility that the word-forms of one
-language interact with those form the other language, scaffolding the
-acquisition of their translation equivalents when their phonologies
-overlap. This mechanism, in line with the parallel activation account of
-bilingual speech perception, would provide a plausible explanation to
-why cognates are acquired ealier by bilinguals. We developed an online
-tool to collect parental reports of receptive and productive
-vocabularies from children learning Catalan and/or Spanish, and present
-data on receptive and productive vocabulary of bilingual toddlers aged
-10 to 36 months.
-\end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi
-
-\setstretch{1.25}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, breakable, enhanced]}{\end{tcolorbox}}\fi
 \hypertarget{introduction}{%
@@
 
-The foundations of word learning are in place early in age: before the
-end of their first year of life infants start directing their gaze to
-some objects when hearing their labels, according to both experimental
-data (Bergelson \& Swingley, 2012, 2015; Jusczyk \& Aslin, 1995; Tincoff
-\& Jusczyk, 1999) and parental reports (e.g., Fenson et al., 2007).
-During the last half of their second year, they acquire new words at an
-increasingly fast rate (Bergelson, 2020; Bloom, 2002; Fenson et al.,
-1994; Goldfield \& Reznick, 1990; Mayor \& Plunkett, 2011). These early
-stages of lexical acquisition are characterised by substantial variation
-across children, reflected, for instance, on the variation on the number
-of words they know (i.e.~vocabulary size, Fenson et al., 1994; Frank et
-al., 2021) or on the proportion of those words that fall into the
-category of nouns, as opposed to verbs, adjectives, or function words
-(e.g., Nelson, 1973, p. bates1994developmental).
-
-Despite this variability, children's trajectories of vocabulary growth
-seem quite stable across languages. Tardif et al. (2008) collected data
-about the first ten words acquired by 10 to 16 month-old infants living
-in the United States, Hong Kong, and Beijing. Since birth, these infants
-had been learning English, Cantonese and Mandarin, respectively. The
-authors found a common pattern across the three groups: their first ten
-words referred to roughly the same concepts, namely relatives/caretakers
-(\emph{daddy}, \emph{mommy}), social routines (\emph{bye}, \emph{uh-oh})
-or animals (\emph{woof-woof}). These results were later extended by
-Frank et al. (2021) to a wider diversity languages, also reporting that
-such cross-language commonalities are stronger at earlier stages of
-lexical acquisition, compared to later stages. Most of the literature on
-early word acquisition, however, has been conducted on monolingual
-children, and neglects the problem of how bilinguals--who represent a
-substantial proportion of the population in most societies--acquire
-words at early ages.
-
-There is evidence that bilinguals know, on average, less words in each
-of their languages than monolinguals, but also that both groups know a
-similar amount of words when the two languages are aggregated. For
-example, Hoff et al. (2012) found that English-Spanish bilingual
-toddlers in South Florida knew less words in English than monolinguals,
-who only learnt English. Both groups knew a similar total amount of
-words when both English and Spanish vocabularies were counted together,
-pointing to the importance of collecting data on both languages when
-assessing bilinguals' communicative development. Other studies have
-provided converging evidence that bilinguals know a similar--or even
-larger--number of words than monolinguals, only when the languages are
-aggregated (Gonzalez-Barrero et al., 2020; Oller \& Eilers, 2002; J.
-Patterson et al., 2004; J. L. Patterson, 2004; Pearson et al., 1993;
+The foundations of word learning are in place early in age. From six
+months of age, infants start directing their gaze to some objects when
+hearing their labels, according to both experimental data (Bergelson \&
+Swingley, 2012, 2015; Jusczyk \& Aslin, 1995; Tincoff \& Jusczyk, 1999)
+and parental vocabulary reports (Fenson et al., 2007; Samuelson, 2021).
+During the last half of their second year, new word forms are acquired
+at an increasingly fast rate (Bergelson, 2020; Bloom, 2002; Fenson et
+al., 1994; Goldfield \& Reznick, 1990; Mayor \& Plunkett, 2011). Such
+early stages of lexical acquisition are characterised by substantial
+variation across children reflected, for instance, on the variability of
+the number of words they know (i.e., vocabulary size, Fenson et al.,
+1994; Frank et al., 2021). Despite this variability, trajectories of
+early vocabulary growth seem quite stable across languages. Tardif et
+al. (2008) collected data about the first ten words acquired by 10-- to
+16--month-old infants living in the United States, Hong Kong, and
+Beijing. Since birth, these infants had been learning English,
+Cantonese, and Mandarin, respectively. The authors found a common
+pattern across the three groups: their first ten words referred to
+roughly the same concepts, namely relatives/caregivers (\emph{daddy},
+\emph{mommy}), social routines (\emph{bye}, \emph{uh-oh}) or animal
+sounds (\emph{woof-woof}). These results were later extended by Frank et
+al. (2021) to a more diverse set of languages, and found that such
+cross-language commonalities are stronger at earlier stages of lexical
+acquisition.
+
+Most research on early word acquisition relies extensively on data from
+monolingual children, and is oblivious to substantial proportion of the
+world population acquiring more than one language since early ages
+(Grosjean, 2021). Previous research on bilingual vocabulary acquisition
+pointed to bilingual toddlers knowing, on average, less words in each of
+their languages than their monolinguals peers, and also to both groups
+knowing a similar amount of words when the bilinguals' two languages are
+taken into account. Hoff et al. (2012) found that English-Spanish
+bilingual toddlers in South Florida (United States) knew less words in
+English than monolinguals did. Both groups knew a similar total amount
+of words when both English and Spanish vocabularies were counted
+together, which substantiates the importance of collecting data on both
+languages when assessing bilinguals' communicative development. Other
+studies have provided converging evidence that bilinguals know a similar
+or even larger number of words than monolinguals when the two languages
+are aggregated (Gonzalez-Barrero et al., 2020; Oller \& Eilers, 2002;
+Patterson, 2004; Patterson \& Pearson, 2004; Pearson et al., 1993;
 Pearson \& Fernández, 1994; Petitto et al., 2001; Smithson et al.,
-2014). While these studies have mostly relied on samples of bilingual
-children learning two relatively distant languages, as it is the case of
-English and Spanish, it is unclear whether children learning
-typologically more similar languages also know less words in each of
-their languages than monolinguals. What role could linguistic distance
-play during early vocabulary growth?
+2014). These studies rely on bilingual samples whose languages are
+relatively distant in typology: English is a Germanic language, while
+Spanish is a Romance language, and both languages differ substantially
+at the grammatical or the phonological level. Little is known about
+vocabulary acquisition in bilinguals learning a pair of typologically
+closer languages (e.g., when both are Romance languages).
 
 For a given set of concepts, bilingual children may be exposed to two
-distinct sets of word-forms--one in each language. Depending on the
-linguistic distance between both languages, the two sets of words may
-overlap in varying degrees. Particularly, when both languages are
-typologically close, like Spanish and Catalan (both Roman languages),
-they are more likely to share a large amount of cognates (i.e.,
-form-similar translation equivalents) than two linguistically distant
-languages, like Spanish and English (one Roman, the other Germanic). For
+distinct sets of word-forms: one in each language. Depending on the
+similarity between both languages, the two sets of words may overlap in
+form in varying degrees. When both languages are typologically close,
+like Spanish and Catalan (both Romance languages), they are more likely
+to share a large amount of cognates (i.e., form-similar translation
+equivalents) than two typologically more distant languages, like Spanish
+and English (Romance and Germanic languages, respectively). For
 instance, in the presence of a door, a Spanish-Italian (or a
-Spanish-Catalan) bilingual might hear \emph{puerta} and \emph{porta}
-(cognates), whereas a Spanish-English bilingual might hear \emph{puerta}
-and \emph{door} (non-cognates). It could be the case that mapping two
-phonologically similar labels (cognates like \emph{puerta}-\emph{porta})
-onto the same referent is easier than doing the same with two
-phonologically dissimilar labels (non-cognates, like \emph{puerta} and
-\emph{door}). If cognates are easier to acquire than non-cognates,
-bilinguals learning a pair of languages that share a high proportion of
-cognates should benefit more often from this facilitation effect than
-those learning a pair of languages with a lower proportion of cognates,
-and should therefore show larger vocabulary sizes.
-
-Floccia et al. (2018) provided evidence in line with this claim. The
-authors collected vocabulary data on word comprehension and production
-from 372 24-month-old bilingual toddlers living in the United Kingdom
-who were learning English and an additional language. The additional
-language was one a pool of 13 typologically diverse languages: Bengali,
-Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian,
-Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The authors
-calculated the average phonological similarity between the words in each
-of these additional languages and their translation equivalents in
-English. Phonological similarity was measured by computing the
-Levenshtein distance between each cross-language pair of phonological
-transcriptions. The Levenshtein distance is a metric that computes the
-edit distance between two strings by counting the smallest number of
-insertions, deletions and substitutions one of the strings has to go
-through to become identical to the other (Levenshtein et al., 1966). The
-resulting scores were then divided by the length of the longest string
-to bound the similarity scores between 0 and 1, and then entered this
-variable as a predictor as they modelled participants' vocabulary sizes.
-Among other findings, the authors reported an increase in productive
-vocabulary size in the additional language associated with an increase
-in the average phonological similarity between the translation
+Spanish-Catalan) bilingual might be exposed to the words \emph{puerta}
+and \emph{porta} (cognates), whereas a Spanish-English bilingual might
+hear \emph{puerta} and \emph{door} (non-cognates). It could be the case
+that mapping two phonologically similar labels (cognates like
+\emph{puerta}-\emph{porta}) onto the same referent is easier than doing
+the same with two phonologically dissimilar labels (non-cognates, like
+\emph{puerta} and \emph{door}). If cognates are easier to acquire than
+non-cognates, bilinguals learning two languages that share a high
+proportion of cognates should benefit more often from this facilitation
+effect than bilinguals learning two languages with a lower proportion of
+cognates, and should therefore show larger vocabulary sizes.
+
+Floccia, Sambrook, Delle Luche, et al. (2018) provided evidence in line
+with this claim. The authors collected vocabulary data on word
+comprehension and production from 372 24-month-old bilingual toddlers
+living in the United Kingdom who were learning English and an additional
+language. The additional language was one of 13 typologically diverse
+languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek,
+Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and
+Welsh. The authors calculated the average phonological overlap between
+the words in each of these additional languages and their translation
+equivalents in English, which was taken as a proxy of the degree of
+\emph{cognateness} between each pair of languages. Phonological overlap
+was measured by computing the Levenshtein distance between each
+cross-language pair of phonetic transcriptions. The Levenshtein distance
+is a metric that computes the distance between two strings of
+characters/phonemes by counting the smallest number of insertions,
+deletions and substitutions one of the strings has to undergo to become
+identical to the other (Levenshtein, 1966). For instance, the
+Levenshtein distance for the translation equivalent
+\emph{cat}-\emph{kat} (/kæt/-/\textipa{kAt}/) in English and Dutch would
+be one: a single edition would be needed to make both phonological
+translation identical (replacing /æ/ by /\textipa{A}/). The same
+translation equivalent in English and Greek
+(\emph{cat}-\emph{\(\gamma\acute{\alpha}\tau\alpha\)},
+/kæt/-/\textipa{Gata}/) would score three, as it requires at least three
+editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/
+and adding another /a/.
+
+Levenshtein distance scores are divided by the length of the longest
+string to normalise the similarity scores between 0 and 1. Finally, the
+result is subtracted from one, to get the complementary, and therefore a
+measure of phonological similarity, instad of dissimilarity. Following
+the previous examples: \(1/4 = 0.25\) for /kæt/-/\textipa{kAt}/, and
+\(3/4 = 0.75\) for /kæt/-/\textipa{Gata}/. Finally, this variable was
+included as a predictor of participants' vocabulary sizes. Among other
+findings, Floccia and co-workers reported an increase in vocabulary size
+in the additional language (i.e., not English) associated with an
+increase in the average phonological similarity between the translation
 equivalents of each language pair. For example, English-Dutch bilinguals
-(22.14\% phonological similarity), were able to produce more Dutch words
-than English-Mandarin bilinguals (1.97\% phonological similarity) were
-able to produce in Mandarin.
-
-Floccia et al.~pointed to \emph{parallel activation} as the main
-mechanism underpinning their results. The parallel activation hypothesis
+(0.2214 mean phonological overlap), were able to produce more Dutch
+words than English-Mandarin bilinguals (0.0197 mean phonological
+overlap) were able to produce in Mandarin.
+
+Floccia et al.~pointed to \emph{parallel activation} as the underpinning
+mechanism behind their results. The parallel activation hypothesis stems
+from the language non-selective account of lexical access, which
 suggests that bilinguals activate both languages simultaneously during
-speech production or comprehension, and that this phenomenon is the
-result of the activation of lexical representations in both languages,
-even when only one is in use during production (Costa et al., 2000;
-Hoshino \& Kroll, 2008) or comprehension (Spivey \& Marian, 1999;
-Thierry \& Wu, 2007). One of the clearest examples of parallel
-activation was provided by Costa et al. (2000). In this study,
-Catalan-Spanish monolingual and bilingual adults were asked to name
-pictures of common objects in Spanish. In half of the trials, the object
-labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre},
-translations of \emph{tree}), whereas in the other half of the trials
-labels were non-cognates (\emph{mesa}-\emph{taula}, translations of
-\emph{table}). Bilinguals named cognate pictures faster than non-cognate
-pictures, even after adjusting for the lexical frequency of the items.
-Importantly, Spanish monolinguals--unfamiliar with the Catalan
-translations of the Spanish words they uttered--did not show this
-effect. These results suggest that bilinguals' Catalan phonology was
-activated during the production of Spanish words, facilitating the
-naming of cognate pictures. Several subsequent studies have also
-provided similar evidence in comprehension in children (Poulin-Dubois et
-al., 2013; e.g., Von Holzen \& Mani, 2012). Parallel activation is
-therefore a plausible mechanism to account for Floccia et al.'s results:
-cognates increase the amount of activation in both languages,
-facilitating word acquisition, and ultimately leading to children
-learning language pairs with a larger proportion of cognates are
-predicted to show larger vocabulary sizes. A missing piece in this
-account is how increase parallel activation might facilitate the
-acquisition new words.
-
-In this study, we hypothesise that cognates are acquired earlier in age
-than non-cognates, which might explain why children learning languages
-that share many cognates show larger vocabulary sizes. Evidence
-supporting this claim is scarce. Bosch \& Ramon-Casas (2014) used
-vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish
-bilinguals aged 18 months, and found that cognates represented a larger
-proportion of participant's vocabulary than non-cognates. Schelletter
-(2002) reported a longitudinal single case of one English-German
-bilingual who produced cognates earlier than non-cognates, on average.
-The low sample size in these two studies makes it challenging to draw
-strong conclusions about the effect of cognateness on vocabulary growth.
-On the other hand, Floccia et al.'s estimates are statically more
-reliable given their (much larger) sample size, but their study was not
-aimed at testing the effect of cognateness on age of acquisition
-directly. In their discussion the authors state the following (pp.~70):
-
-\begin{quote}
-``This finding also provides support to the proposal that the cognate
-advantage is due to cognates being acquired before non-cognates in early
-childhood (Costa et al., 2016), leading to an ease of processing later
-in life.''
-\end{quote}
-
-We identify two main reasons why an earlier age of acquisition for
-cognates than for non-cognates is an unwarranted conclusion from Floccia
-et al.'s results. First, the response variable used was the proportion
-of words each participant understood and/or produced (i.e., vocabulary
-size), from the list of lexical items in the vocabulary checklists. By
-aggregating the responses from all items into a single datum per child,
-information about the acquisition status of cognates vs.~non-cognates
-was lost. Second, all participants were aged \textasciitilde24 months,
-meaning that even if the unaggregated responses to individual items were
-included as response variable, the possible effect of cognateness could
-only be interpreted as an increase or decrease in the likelihood of
-participants at such age to have acquired each item, and not as an
-increase or decrease in the age of acquisition of such item.
-
-More recently, Mitchell et al. (2022) addressed this issue overcoming
-some of these pitfalls. Using a larger, longitudinal sample of 47
-16-to-30 month-old French-English bilinguals living in Canada, the
-authors collected data on expressive vocabulary data in both languages.
-They created two lists of translation equivalents: one made of 131
-cognates, and one made of 406 non-cognates. The proportion of
-translation equivalents that children were reportedly able to produce
-was higher in the cognate lists than in the non-cognate list across
-ages, even when both lists were matched by semantic category (furniture,
-animals, food were similarity represented in both lists) and age of
-acquisition norms (an index of word difficulty). These findings shed
-some light on the ongoing exploration of why (if at all) bilinguals'
-vocabulary size grows faster when both languages are phonologically more
-similar: word-forms sharing more phonemes with their translation
-equivalents (i.e., cognates) seem to be acquired faster.
-
-As Mitchell et al. (2022) note in their discussion, two competing
-hypotheses remain untested. The facilitative effect of cognateness can
-be explained by, at least, two alternative, arguably incompatible,
-mechanisms. On the one hand, cognateness might operate as a facilitator
-of the acquisition of both forms of a translation pair, with every
-exposure to either form activating the other, and therefore
-strengthening the association between both and their referent. If this
-is the case, then cognateness should increase the probability of
-acquisition of the forms in both languages equally.
-
-Another possibility is that cognateness plays a role in the acquisition
-of a translation pair only when at least one of the word-forms has been
-acquired. This account departs from the premise that, if the child is
-not previously familiar with one of the word forms, the acquisition of
-its translation in the other language cannot be facilitated by
-cognateness. This implies that, if cognateness facilitates the
-acquisition of a word form, it can only do so after the child has
-acquired one of the word forms of the translation pair. Given that
-children are more likely to acquire words from languages to which they
-are exposed more often (Cattani et al., 2014; David \& Wei, 2008;
+speech production (Costa et al., 2000; Hoshino \& Kroll, 2008) or
+comprehension (Spivey \& Marian, 1999; Thierry \& Wu, 2007), even in
+fully monolingual environments. One of the clearest pieces of evidence
+of parallel activation was provided by Costa et al. (2000). In this
+study, Spanish monolinguals and Catalan-Spanish bilingual adults were
+asked to name pictures of common objects in Spanish. In half of the
+trials, the object labels were cognates in Spanish and Catalan
+(\emph{árbol}-\emph{arbre}, translations of \emph{tree}), whereas in the
+other half of the trial labels were non-cognates
+(\emph{mesa}-\emph{taula}, translations of \emph{table}), importantly
+such distinction was only relevant for bilinguals. Bilinguals named
+cognate pictures faster than non-cognate pictures, even after adjusting
+for the lexical frequency of the items. In contrast, Spanish
+monolinguals, who were unfamiliar with the Catalan translations of the
+Spanish words they uttered, showed equivalent naming times for the two
+types of stimuli. These results showed that bilinguals' Catalan
+phonology was activated during the production of Spanish words,
+facilitating the naming of cognate pictures.
+
+Regarding word acquisition, the parallel activation account is a
+plausible explanation for Floccia et al.'s results in the light of two
+observations. First, bilingual toddlers' lexicon seems to consist of a
+large proportion of translation equivalents: the presence of a word in
+their lexicon is a good predictor of its translation equivalent also
+being present (Bilson et al., 2015; Pearson et al., 1995; Tsui et al.,
+2022a). Second, there is experimental evidence of parallel activation
+during word comprehension in children and toddlers (Bosma \& Nota, 2020;
+Poulin-Dubois et al., 2013; Von Holzen \& Mani, 2012). It is possible
+that cognateness increases the amount of cross-language activation,
+facilitating word acquisition. This could ultimately lead to children
+who are learning languages with a larger proportion of cognates (e.g.,
+English and Dutch) showing larger vocabulary sizes. This account in
+which cognateness underlies the facilitation effect of typological
+language distance is also in line with previous studies suggesting that
+the acquisition of new words is facilitated by their phonological or
+semantic similarity with other words (Fourtassi et al., 2020; Hills et
+al., 2009; Jones \& Brandt, 2019; Laing, 2022). If phonological
+similarity plays a cross-language facilitation role during word
+acquisition, cognate translation equivalents---which share both semantic
+and phonological similarity---should be acquired, on average, earlier
+that non-cognate translation equivalents---which share semantic
+similarity (e.g., are taxonomically related) but not phonological
+similarity.
+
+Evidence supporting an earlier age of acquisition for cognates is
+scarce. Bosch and Ramon-Casas (2014) collected parental reports of
+expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months,
+and found that cognates represented a larger proportion of participant's
+vocabulary than non-cognates. Schelletter (2002) provided converging
+evidence from a longitudinal single-case study, in which an
+English-German bilingual produced cognates earlier than non-cognates, on
+average. The low sample size in these two studies makes it challenging
+to draw strong conclusions about the effect of cognateness on vocabulary
+growth. Floccia et al.'s estimates are statistically more reliable given
+their remarkably larger sample size (372 participants). Although their
+results suggest a cognateness effect on vocabulary size, their study was
+not aimed at testing the effect of cognateness on age of acquisition at
+the \emph{word-level}. The response variable in this study was the
+proportion of words each participant understood and/or produced (i.e.,
+vocabulary size), from the list of lexical items in the vocabulary
+checklists. By aggregating the responses from all words into a single
+datum per child, it was no longer possible to test the effect of
+cognateness on word acquisition at the word-form level. Also, all
+participants were aged 24 months, meaning that even if the unaggregated
+responses to individual items were included as the response variable,
+age-sensitive claims about the effect of cognateness on word acquisition
+should be taken with caution (a limitation also present in the Bosch \&
+Ramon-Casas, 2014 study).
+
+More recently, Mitchell et al. (2022) addressed this issue providing a
+more fine grained measure of a cognateness facilitation effect in
+vocabulary acquisition. Using a longitudinal sample of 47 16-to-30
+month-old French-English bilinguals living in Canada, the authors
+collected data on expressive vocabulary data in both languages. They
+created two lists of translation equivalents: one made of 131 cognates,
+and one made of 406 non-cognates. The proportion of translation
+equivalents that children were reportedly able to produce was higher in
+the cognate lists than in the non-cognate list across ages, even when
+both lists were matched by semantic category (furniture, animals, food
+were similarity represented in both lists) and age of acquisition norms
+(an index of word difficulty). These findings shed some light on the
+ongoing exploration of why (if at all) bilinguals' vocabulary size seems
+to grow faster when both languages are more similar at the lexical
+level. Word-forms sharing more phonemes with their translation
+equivalents (i.e., cognates) seem to be acquired faster. Parallel
+activation remains a plausible explanation for this effect. When a child
+is exposed to a word form, its (phonologically similar) translation
+equivalent is activated, and this cross-language activation increases
+the chances of acquisition. The precise underpinnings of this effect are
+still unclear.
+
+In the present investigation we approach the cognate facilitation on
+word learning by extending the conceptual framework of accumulator
+models of language acquisition to the bilingual case. Accumulator models
+formalise the widespread assumption that a child's vocabulary growth
+rate is a function of mainly---but not exclusively---the number of
+learning instances that a child has accrued with a particular word (see
+Kachergis et al., 2022 for review). A learning instance of a word
+consists, for example, in hearing such word-form (\emph{flower}) in the
+presence of its referent (a flower). As infants receive linguistic
+input, they also accumulate opportunities to learn words until a
+threshold is crossed, an the word can be considered to be acquired. In
+line with this account, the frequency with which a child encounters a
+particular word-form in their speech input is a critical determiner of
+the age at which such word will be acquired. The more frequently
+(Hidaka, 2013; Mollica \& Piantadosi, 2017). For instance, words with
+high lexical frequency like `flower' (1,077 times per million words in
+the CHILDES corpus, MacWhinney, 2000) are more likely to be acquired
+earlier than words with low lexical frequency such as `taxi' (581 times
+per million words) (e.g., Goodman et al., 2008): infants might
+accumulate learning experiences with `flower' faster than with `taxi'.
+
+Previous accounts on vocabulary growth based on accumulator models are a
+convenient framework to study bilingual vocabulary growth, as bilinguals
+share fundamentally similar learning mechanisms with monolinguals.
+Developmental divergences between both groups can be understood as the
+result of the the negotiation between such mechanisms and the particular
+demands of a dual language input (Curtin et al., 2011; e.g., Werker \&
+Curtin, 2005). One of such divergences is the fact that bilinguals'
+frequency of exposure to a word-form is mediated by their degree of
+exposure to their languages. Assuming that bilinguals and monolinguals
+are exposed to an equivalent amount of linguistic input, a
+Spanish-English bilingual who is exposed to Spanish 80\% of the time and
+to English 20\% of the time would need more time to accumulate the same
+experience with `flower' and `taxi' than an English monolingual would
+(who listens 100\% of the time to English).
+
+We suggest that the cognateness effect proposed by Floccia et al.~and
+Mitchell et al.~facilitates word acquisition by reducing the number of
+learning instances needed by a word to be acquired, and that this effect
+takes place \emph{only} once the (cognate) translation of such word has
+been acquired. We hypothesise that both forms in the translation
+equivalent (e.g., \emph{gat} and \emph{gato} {[}Catalan and Spanish for
+`cat'{]}, for a Catalan-Spanish bilingual) follow their acquisition
+trajectory as they would for a monolingual child: accumulating learning
+instances, which occur a a rate proportional to the lexical frequency of
+the word-form and the amount of exposure to the language they belong to.
+When one of the word-forms is sufficiently acquired, the number of
+learning instances left needed for the acquisition of the other
+word-form is reduced.
+
+This facilitation would occur as the result of the semantic and
+phonological overlap between both words. When the child has sufficient
+lexical knowledge of \emph{gat}, and therefore there is some
+for-referent association in place, future presentations of the similar
+sounding translation \emph{gato} in the presence of the same reference
+should lead to an increased likelihood of acquisition. In the case of a
+non-cognate translation equivalent, such facilitation would be unlikely.
+For instance, even if the child has already acquired the word
+\emph{gos}, hearing its dissimilar Spanish translation \emph{perro}
+would hardly lead to an increased probability of acquisition. In line
+with this account, only the later-acquired form of the translation
+equivalent would benefit from its cognate status. Given that children
+are more likely to acquire words from languages to which they are more
+often exposed to (Cattani et al., 2014; David \& Wei, 2008;
 Thordardottir et al., 2006), the acquisition of words in the language of
 lower exposure should, on average, be more susceptible to the effect of
-cognateness.
-
-In line with this second hypothesis, the vocabulary size associated with
-linguistic similarity that Floccia et al.~reported was larger in the
-additional language vocabulary than in English vocabulary. Most
-participants in their sample were English-dominant, meaning that their
-relative amount of exposure to English was larger than in the additional
-language. Therefore, participants may have, on average, learnt the
-English word-form of translation equivalents earlier than the word-form
-in the additional language. If this is the case, then the acquisition of
-English words by English-dominant participants would rarely benefit from
-their cognate status (the other word-form is not available yet), while
-the acquisition of words in the additional language would benefit from
-their phonological similarity with the (available) English form.
-
-In this study, we investigated the role of cognateness on the
-acquisition of translation equivalent, and whether such effect is
-conditional to the amount of time the child is exposed to the
-corresponding language of each member of the translation pair. Using an
-online vocabulary checklist--designed specifically for this study--we
-collected data from a sample of children aged 10 to 35 months learning
-Catalan and/or Spanish, with varying degrees of exposure to each
-language. We then adopted an Item Response Theory (\emph{IRT}) approach
-to model the probability of participants being reported by their parents
-to understand or understand \emph{and} say each word in the checklist,
-conditional to its cognate status in Catalan and Spanish and
-participants' degree of exposure to the corresponding language, while
-adjusting for other indices of word difficulty--lexical frequency and
-length (phonemes)-- and participant skill--age (see Kachergis et al.,
-2022 for a similar approach).
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-Data and materials are available at OSF \url{https://osf.io/hy984/}, and
-code is available at GitHub
-\url{https://github.com/gongcastro/trajectories}.
-
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
-
-We collected data from 422 children from the Metropolitan Area of
-Barcelona between 28th October, 2019 and 12th July, 2022. All families
-gave informed consent before participating and this study was approved
-by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from
-Hospital del Mar (Barcelona, Spain), code XXXXXXXXX. Families from the
-participant database of the Laboratori de Recerca en Infància of the
-Universitat Pompeu Fabra were contacted by e-mail or phone if their
-child were aged between 10 and 35 months, and had not been reported to
-be exposed more than 10\% of the time to a language other than Spanish
-or Catalan. Upon consent, families were sent a link to the questionnaire
-via e-mail, which they filled from a computer, laptop, or mobile device
-in a browser within the two week following the invitation to
-participate. Table~\ref{tbl-participants-lp} summarises the distribution
-of participants across ages and degrees of exposure (DoE) to Catalan.
+cognateness. For instance, if the previous child is exposed 80\% of the
+time to English, and 20\% of the time to Spanish, words in Spanish
+(especially low-frequency ones) would show a cognateness facilitation
+effect, with an earlier age of acquisition for cognates than for
+non-cognates.
+
+Figure~\ref{apafg-hypotheses} illustrates graphically a theoretical
+account of the acquisition of a Catalan-Spanish cognate translation
+equivalent, \emph{gat}-\emph{gato} {[}cat{]} whose word-forms share
+three phonemes, and of a non-cognate translation equivalent,
+\emph{gos}-\emph{perro} {[}dog{]} whose word-forms only share one
+phoneme. We assume the case of a Catalan-Spanish bilingual child who is
+exposed 65\% of the time to Catalan, and 35\% of the time to Spanish. On
+the top panels, we illustrate the cumulative sum of learning instances
+that a child accumulates across ages for the Catalan word-form (/gat/ or
+/gos/) and its Spanish translation (/perro/ or /gato/). We assume an
+arbitrary threshold of learning instances at which the child is
+considered to have acquired a word (solid grey line). Points and
+triangles show the age-of-acquisition at which the amount of learning
+instances reaches the threshold. We show the predicted
+age-of-acquisition under two hypothesis: (1) a cognateness facilitation
+effect that lowers the acquisition threshold after one of the word-forms
+(in this case Catalan /gat/) has been acquired, and (2) no cognateness
+facilitation, for which the acquisition of the second word-form occurs
+unconditionally of the acquisition of the first. On the bottom, we show
+the predicted probability of acquisition for the Catalan and Spanish
+words under each hypothesis, as it would be expected to observe in a
+vocabulary questionnaire like the one in the present study. We assume
+that such probability starts at 0\% at birth, reaches 50\% (and grows
+the fastest) at the age of acquisition, and reaches 100\% at some points
+in later ages.
+
+\begin{figure}[h!]
+\caption{Schematic representation of the hypothesised cognate facilitation effect on word acquisition.}
+\label{apafg-hypotheses}
+\includegraphics[width=1\linewidth]{assets/img/diagram.png}
+
+\end{figure}
+
+In line with this hypothesis, the size of the effect of linguistic
+similarity on vocabulary size that Floccia et al.~reported was larger in
+the additional language (language of lower exposure) vocabulary than in
+English vocabulary (language of higher exposure). Most participants in
+their sample were English-dominant, meaning that their relative amount
+of exposure to English was larger than in the additional language.
+Therefore, participants may have on average learned the English
+word-form of translation equivalents earlier than the word-form in the
+additional language. Because of children's imbalance in exposure to
+words in English and in the additional language, the acquisition of
+English words by English-dominant participants might then benefit less
+frequently from their cognate status (the other word-form is unlikely to
+be available yet), while the acquisition of words in the additional
+language would benefit more frequently from their phonological
+similarity with the (more likely available) English translation.
+
+We tested the hypothesis that the cognate effect on word acquisition is
+conditional to a child's exposure rate to the word, which would support
+the hypothesis that cognateness plays a role in the acquisition of a
+word \emph{only} once its translation has been already acquired. We
+collected vocabulary data on production and comprehension from a large
+sample of bilingual Catalan-Spanish children using an online vocabulary
+checklist designed for the present study. Participants were aged 12 to
+31 months, and were exposed to Catalan and/or Spanish to varying
+degrees. We adopted a Bayesian explanatory item response theory (IRT,
+see Kachergis et al., 2022 for a similar approach) approach to model the
+probability of a given participant being reported to \emph{understand}
+or \emph{understand and say} each word-form in the checklist,
+conditional to participants age and rate of exposure to the word-form,
+and the cognate status of the word form, while adjusting for the the
+number of phonemes in the word form (an indicator of word acquisition
+difficulty).
+
+To operationalise participants' exposure rate to a word-form, we used a
+composite measure that captures the lexical frequency of the word-form
+(an approximation of how often a participant is exposed to it), and
+child's amount of exposure to the language such word belongs to. This
+measure was calculated by multiplying the word-form's lexical frequency
+with participant's degree of exposure to the language the word belongs
+to. Lexical frequencies were extracted from the CHILDES corpora
+(MacWhinney, 2000; Sanchez et al., 2019) (see Appendix A for more
+details). Participants' degree of exposure were reported by caregivers
+in a language exposure questionnaire they filled before the vocabulary
+checklist. Caregivers provided a proportion of the time the child was
+estimated to listen to any language they were spoken to regularly
+(described in the methods section). The resulting measure is a
+exposure-weighted lexical frequency measure, which was included in
+analyses as a proxy of how often bilingual participants were exposed to
+each word-form.
+
+\emph{Cognateness} was defined as a continuous measure of phonological
+similarity between translation equivalents. We quantified the
+phonological similarity between each pair of Catalan-Spanish word-forms
+by computing the inverse of the normalised Levenshtein distance between
+their X-SAMPA phonetic transcriptions (Levenshtein, 1966; Wells, 1995).
+This score is computed by counting the number of insertions, deletions
+and replacements needed by both transcriptions to become identical,
+dividing the resulting value by the length of the longest transcription,
+and finally subtracting this value from one. This results in a
+proportion that indicates how much the two phonetic transcriptions of
+the translation equivalent are similar to each other, ranging from zero
+(no similarity at all) to one (both transcriptions are identical) (see
+Floccia, Sambrook, Delle Luche, et al., 2018; Fourtassi et al., 2020;
+Laing, 2022 for similar approaches). For instance, this measure of
+phonological similarity returns 0\% for \emph{ocell} \texttt{{[}uceL{]}}
+and \emph{pájaro} \texttt{{[}paxa4o{]}}, Catalan and Spanish for `bird',
+and returns 50\% for \emph{lluna} \texttt{{[}Lun5{]}} and \emph{luna}
+\texttt{{[}luna{]}}, Catalan and Spanish for `moon' (see Appendix A for
+more details).
+
+\hypertarget{sec-methods}{%
+\section{Methods}\label{sec-methods}}
+
+All materials, data, and reproducible code can be found at the OSF
+(\url{https://osf.io/hy984/}) and GitHub
+(\url{https://github.com/gongcastro/trajectories}) repositories. This
+study was conducted according to guidelines laid down in the Declaration
+of Helsinki, and was approved by the Drug Research Ethical Committee
+(CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.
+
+\hypertarget{sec-questionnaire}{%
+\subsection{Questionnaire}\label{sec-questionnaire}}
+
+To collect vocabulary data from participants, we created an \emph{ad
+hoc} questionnaire: the Barcelona Vocabulary Questionnaire. This
+questionnaire was inspired by the MacArthur-Bates Communicative
+Development Inventory (Fenson et al., 2007) and its adaptations to other
+languages, and was implemented on-line using the formR platform (Arslan
+et al., 2020). This questionnaire is structured in three blocks: (1) a
+language exposure questionnaire, (2) a demographic survey, and (3) two
+vocabulary checklists. Vocabulary checklists followed a similar
+structure as the Oxford Communicative Developmental Inventory (Hamilton
+et al., 2000) and consisted in two lists of words: one in Catalan and
+one in Spanish. both lists included items from a diverse sample of 26
+semantic/functional categories. The Catalan checklist contained 793
+items and the Spanish checklist contained 797. Items in one language
+were translation equivalents of the items in the other language, roughly
+following a one-to-one mapping (e.g., the same participant responded to
+both \emph{gos} and \emph{perro}, Catalan and Spanish for `dog') (see
+Table~\ref{apatb-items} for a summary of the questionnaire items).
 
 \begin{table}
-
-\caption{\label{tbl-participants-lp}Participant sample size by age and
-degree of exposure to Catalan.}\begin{minipage}[t]{\linewidth}
-\subcaption{\label{tbl-participants-lp-1}Information about DoE was provided by families before filling the
-questionnaire. A 100\textbackslash\% indicated that the participant was exclusively
-exposed to Catalan, and never to Spanish. A 0\textbackslash\% indicates that the
-participant was not exposed to Catalan ever, and rather most of the time
-to Spanish. A 50\textbackslash\% indicates that the participant was exposed to Catalan
-and Spanish approximately half of the time each. The 100\textbackslash\%, 0\textbackslash\%, and
-50\textbackslash\% would be traditionaly classified as Catalan monolingual, Spanish
-monolingual, and Catalan-Spanish bilingual, respectively. For
-illustration purposes this table, DoEs were binned into 25\textbackslash\%-wide bins,
-and ages (in months) binned into 4 month-wide bins. Hyphens indicate
-that no participants from that specific combination of age and DoE
-filled the questionnaire. }
-
-{\centering
-
-\setlength{\LTpost}{0mm}
-\begin{longtable}{l|rrrrrrrr}
-\tabularnewline
-
+\caption{Summary of the items included in the final analyses.}
+\label{apatb-items}
+
+\begin{longtable*}{l|rrrrrc}
 \toprule
-\multicolumn{1}{l}{} &  & \multicolumn{7}{c}{Age (months)} \\ 
-\cmidrule(lr){3-9}
-\multicolumn{1}{l}{} & DoE Catalan\textsuperscript{1} & [10,14] & (14,18] & (18,22] & (22,26] & (26,30] & (30,34] & (34,36] \\ 
+\multicolumn{1}{l}{} & List A & List B & List C & List D & Total & Examples \\ 
 \midrule
- & 75-100\% & 22 & 25 & 36 & 39 & 21 & 8 & – \\ 
- & 50-75\% & 12 & 13 & 27 & 38 & 21 & 6 & 8 \\ 
- & 25-50\% & 17 & 16 & 47 & 29 & 24 & 4 & 11 \\ 
- & 0-25\% & 15 & 12 & 21 & 19 & 9 & – & – \\ 
+Household items & 31 & 26 & 30 & 25 & 112 & bottle, bed \\ 
+Food and drink & 29 & 26 & 23 & 27 & 105 & water (beverage), sandwich \\ 
+Animals & 26 & 23 & 19 & 25 & 93 & bee, owl \\ 
+Outside & 14 & 13 & 13 & 15 & 55 & tree, ground \\ 
+Body parts & 14 & 12 & 11 & 11 & 48 & tummy, mouth \\ 
+Toys & 11 & 11 & 12 & 13 & 47 & box, balloon \\ 
+Clothes & 12 & 12 & 10 & 10 & 44 & bib, button \\ 
+Vehicles & 9 & 10 & 11 & 10 & 40 & airplane, boat \\ 
+Colours & 6 & 6 & 6 & 6 & 24 & yellow, blue \\ 
+People & 7 & 4 & 6 & 6 & 23 & child, grandma \\ 
+Furniture and rooms & 4 & 4 & 4 & 4 & 16 & bath, kitchen \\ 
+Time & 2 & 2 & 2 & 2 & 8 & day, night \\ 
+Adventures & 1 & 1 & 1 & 1 & 4 & witch \\ 
+Parts of things & 1 & 1 & 1 & 1 & 4 & wheel \\ 
+\bottomrule
+\end{longtable*}
+
+\end{table}
+
+For each word included in the vocabulary checklists, we asked parents to
+report whether their child was able to understand it, understand
+\emph{and} say it, or did not understand or say it (checked out by
+default). Given the large number of words in the vocabulary checklists,
+we created four different subsets of the complete list of items. Each
+subset contained a random but representative sub-sample of the items
+from the complete list. Semantic/functional categories with less than 16
+items---thus resulting in less than four items after dividing it in four
+lists---were not divided in the short version of the questionnaire: all
+of their items were included in the four lists. Items that were part of
+the trial lists of some ongoing experiments in the lab were also
+included in all versions. The resulting reduced vocabulary checklists
+contained between 343 and 349 Catalan words, and between 349 and 371
+Spanish words. Participants were randomly allocated into one of the four
+subsets. Each response to the questionnaire provided data for a minimum
+of 343 and a maximum of 371 translation equivalents.
+
+\hypertarget{sec-participants}{%
+\subsection{Participants}\label{sec-participants}}
+
+We collected 436 responses to the questionnaire from 367 distinct
+children from the Metropolitan Area of Barcelona between the 30th of
+March, 2020 and the 31th of October, 2022: 314 of those participants
+participated once, 41 twice, 8 three times, and 4 four times. Recurrent
+participants provided responses with a minimum of 25 days between
+responses, and a maximum of 527, and were always allocated to the same
+questionnaire list (A, B, C, or D). Participants were part of the
+database of the Laboratori de Recerca en Infància (Universitat Pompeu
+Fabra), and were contacted by e-mail or phone if their child were aged
+between 12 and 32 months, and had not been reported to be exposed more
+than 10\% of the time to a language other than Spanish or Catalan (see
+Table~\ref{apatb-participants} for a more detailed description of the
+sample). In total, 70 participants (16.06\%) participants were reported
+to be exposed to a third language other than Catalan and Spanish. All
+families provided informed consent before participating. Upon consent,
+families were sent a link to the questionnaire via e-mail, which they
+filled from a computer, laptop, or mobile device. Filling the
+questionnaire took 30 minutes approximately. After completion, families
+were rewarded with a token of appreciation.
+
+\begin{table}
+\caption{Participant sample size by age and degree of exposure to Catalan.}
+\label{apatb-participants}
+
+\begin{longtable*}{l|rrrrrrr}
+\toprule
+\multicolumn{1}{l}{} &  & \multicolumn{6}{c}{Age (months)} \\ 
+\cmidrule(lr){3-8}
+\multicolumn{1}{l}{} & Catalan exposure & [10,14] & (14,18] & (18,22] & (22,26] & (26,30] & (30,34] \\ 
+\midrule
+ & 75-100\% & 17 & 23 & 36 & 39 & 20 & 6 \\ 
+ & 50-75\% & 8 & 13 & 27 & 38 & 21 & 1 \\ 
+ & 25-50\% & 10 & 16 & 48 & 28 & 20 & – \\ 
+ & 0-25\% & 6 & 12 & 21 & 16 & 9 & 1 \\ 
 \midrule 
 \midrule 
-Total & — & $66$ & $66$ & $131$ & $125$ & $75$ & $18$ & $19$ \\ 
+Total & — & $41$ & $64$ & $132$ & $121$ & $70$ & $8$ \\ 
 \bottomrule
-\end{longtable}
-\begin{minipage}{\linewidth}
-\textsuperscript{1}This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10\% of the time\\
-\end{minipage}
-
-}
-
-\end{minipage}%
+\end{longtable*}
+
+\tablenote{Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
+In this table, a 100\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
+A 0\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
+A 50\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
+The 100\%, 0\%, and 50\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
+For illustration purposes this table, DoEs were binned into 25\%-wide bins, and ages (in months) binned into 4 month-wide bins.
+Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire.}
 
@@
 
-We used the highest educational attainment of parents or caretakers as a
-proxy of participants' socio-economic status (SES), which families
-self-reported in the questionnaire by filling two items asking for the
-educational attainment of each parent or caretaker, with the following
-available options: \emph{No education}, \emph{Primary},
-\emph{Secondary}, \emph{Complementary}, \emph{Vocational}, and
-\emph{University}, in line with the current educational system in Spain.
-Most families reported university studies (403, 0.9549763), followed by
-families were the highest educational attainment were vocational studies
-(75, 18\%), complementary studies (6, 1\%), secondary education (7, 2\%,
-0\%), no formal education (2, 0\% and primary education (1,
-\textless1\%).
-
-\hypertarget{questionnaire}{%
-\subsection{Questionnaire}\label{questionnaire}}
-
-The questionnaire was implemented on-line using the formR platform
-(Arslan et al., 2020), and was structured in three blocks: a (1)
-language questionnaire, a (2) demographic survey, and a (3) Catalan and
-a Spanish vocabulary checklists. Vocabulary checklists followed a
-similar structure as the Oxford Communicative Developmental Inventory
-(Hamilton et al., 2000) and consisted in two lists of words: one in
-Catalan and one in Spanish. The Catalan inventory contained 798 items
-and the Spanish inventory contained 802. Items in one language were
-translation equivalents of the items in the other language (e.g.,
-whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the
-word \emph{perro} was included in the Spanish inventory), roughly
-following a one-to-one mapping. In case two translation equivalents were
-possible for a given word, both were included as separate items (e.g.,
-Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar}
-and \emph{terminar}), or merged them into a single item (e.g., Spanish
-\emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}). We
-included items from a diverse sample of 27 semantic/functional
-categories.
-
-For each word in the vocabulary checklists, we asked parents to report
-whether their child was able to understand it, understand \emph{and} say
-it, or did not understand or say it (checked out by default). Some
-families filled a long version of the vocabulary checklists (803
-translation equivalents; 798 items in Catalan, 802 items in Spanish),
-while others filled a shorter version (\textasciitilde400 translation
-equivalents, \textasciitilde400 items in Catalan, \textasciitilde400
-items in Spanish). These last families were randomly allocated into one
-of four different subsets of the complete list of items. These lists
-were designed so that each contained a representative sub-sample of the
-items from the complete list. Semantic/functional categories with less
-than 16 items--thus resulting in less than four items after dividing it
-in four lists--were not divided in the short version of the
-questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some
-experiments in the lab were also included in all versions. For the
-analyses in the present study, we considered responses to words
-corresponding to nouns, verbs, and adjectives (Fourtassi et al., 2020).
-We excluded multi-word items (e.g., \emph{barrita de cereales} {[}cereal
-bar{]}) and items that included more than one word-form (e.g.,
-\emph{mono / mico}).
-
-We extracted four variables of interest for each word-form: (1) Word
-lexical frequency (\(Frequency\)): lexical frequency of the word in its
-corresponding language, expressed as Zipf scores (Van Heuven et al.,
-2014; Zipf, 1949). This variable ranges from 0 to 7, and follows and
-approximates a normal distribution, with most values in a corpus ranging
-from 3 to 5 points. Lexical frequencies were extracted from an adult
-subtitle-based database: SUBTLEX-CAT (Boada et al., 2020) for Catalan
-words and SUBTLEX-ESP (Cuetos et al., 2012) for Spanish words. Using
-child-based estimates of lexical frequency like CHILDES was not possible
-due to the low number of Catalan participants and tokens in the
-available corpora. Responses to words with missing lexical frequency
-score were excluded from analyses (1.22\% of the total number of items).
-(2) Word length (\(Phonemes\)): number of phonemes in the phonological
-transcription of the word-form in International Phonological association
-format. (3) Phonological similarity (\(Levenshtein\)): normalised
-Levenshtein similarity between the word-form and its translation in the
-other language (Levenshtein et al., 1966). This score is calculated by
-first calculating the Levenshtein distance between the two
-transcriptions (number of insertions, deletions or replacements needed
-for the shortest transcription to become identical to the longer
-transcription), then dividing the resulting value by the length of the
-longest transcription, and finally subtracting this value from 1. This
-results in a proportion that indicates how much the two phonological
-transcriptions of the translation equivalent are similar to each other,
-ranging from 0\% (no similarity at all) to 100\% (both transcriptions
-are identical) (see Floccia et al., 2018; Fourtassi et al., 2020; Laing,
-2022 for similar approaches). (4) Participant age: number of months
-elapsed between participants' birth date and questionnaire completion.
-Finally, (5) participant degree of exposure to the words' language
-(\(DoE\)): percentage of exposure to the language the word belonged to.
-For example, for a participant with 90\% exposure to Catalan, and 10\%
-to Spanish, the DoE of the Catalan word \emph{taula} would be 90\%, and
-the DoE for the Spanish word \emph{mesa} would be 10\%.
-Table~\ref{tbl-item-summary} shows a summary of the distribution of the
-word-level predictors across languages in and questionnaire versions.
-
-\hypertarget{tbl-item-summary}{}
-\begin{longtable}{lrrrrrrrrrr}
-\caption{\label{tbl-item-summary}Lexical properties of the word-forms and translationpairs included in
-the vocabulary checklist and our analyses. We report the number of
-items, mean (M), standard deviation (SD), and range of the lexical
-frequency (epxressed in Zipf scores), length (number of phonemes) of the
-word forms, and the phonological similarity (normalised Levenshtein
-similarity) with their translation equivalent. Statistics are reported
-separately for Catalan and Spanish, and for the different vocabulary
-checklists that participants were allocated to (A, B, C, D). }\tabularnewline
-
+We used the highest self-reported educational attainment of parents or
+caregivers as a proxy of participants' socio-economic status (SES). This
+information was provided by each parent or caregiver by selecting one of
+six possible alternatives in line with the current educational system in
+Spain: `sense escolaritzar/sin escolarizar' {[}no education{]},
+`educació primària/educación primaria' {[}primary school{]}, `educació
+secundària/educación secundaria' {[}secondary school{]},
+`batxillerat/bachillerato' {[}complementary studies/high school{]},
+`cicles formatius/ciclos formativos' {[}vocational training{]}, and
+`educació universitària/educación universitaria' {[}university
+degree{]}. Most families reported university studies (355, 82\%),
+followed by families were the highest educational attainment were
+vocational studies (61, 14\%), complementary studies (6, 1\%), secondary
+education (8, 2\%), no formal education (2, \textless1\%) and primary
+education (1, \textless1\%).
+
+\hypertarget{sec-analysis}{%
+\subsection{Data analysis}\label{sec-analysis}}
+
+From the responses to the 1,590 items included in the questionnaire, we
+restricted the analyses to responses to nouns (628 items corresponding
+to other grammatical classes were excluded) (Fourtassi et al., 2020). We
+then excluded items with missing lexical frequency scores (\emph{n} =
+269), items that included more than one lemma (e.g., \emph{mono/mico}
+{[}monkey{]}, \emph{n} = 48), multi-word items or phrases (e.g.,
+\emph{barrita de cereales} {[}cereal bar{]}, \emph{n} = 9). Finally, we
+removed items without a translation in the other language (\emph{n} =
+32). This resulted in a final list of 604 items, corresponding to words
+302 Catalan words and their 302 Spanish translations (302 translation
+equivalents). After collecting participants' responses, the final
+dataset consisted of 138,050 observations, each corresponding to a
+single response of one participant to one item. Each translation
+equivalent received a median of 232 responses (\emph{Min} = 106,
+\emph{Max} = 872) from participants, both languages pooled together.
+
+We modelled the probability of participants answering each response
+category (\emph{No} \textless{} \emph{Understands} \textless{}
+\emph{Understands and Says}) using a Bayesian, mixed-effects ordinal
+regression model (4 sampling chains with 1,000 iterations each). This
+model allowed us to estimate both item and participant word-acquisition
+trajectories, as a function of different fixed effects from
+\(\text{Age}\) (number of months elapsed between participants' birth
+date and questionnaire completion), \(\text{Length}\) (number of
+phonemes in the X-SAMPA phonological transcription of the word-form),
+\(\text{Exposure}\) (composite measure of lexical frequency and language
+degree of exposure), and \(\text{Cognateness}\) (a measure of
+cognateness, defined as the phonological similarity between translation
+equivalents). For an easier interpretation of the regression
+coefficients, the values of all predictors were standardised before
+entering the model by subtracting the mean of the predictor from each
+value and dividing the result by the standard deviation of the
+predictor. Participant- and item-level random intercepts and slopes were
+included when appropriate, according to the structure of the data (Barr
+et al., 2013). We specified a weakly informative prior around the
+parameters of the model (see Appendix C for details).
+
+We built a baseline model \(\mathcal{M}_{0}\) that only included the
+intercept and the main effects of \(\text{Age}\), \(\text{Length}\), and
+\(\text{Exposure}\) as predictors, and compared it against increasingly
+more complex models, which introduced predictors in the following order:
+\(\text{Age} \times \text{Exposure}\) (\(\mathcal{M}_{1}\)),
+\(\text{Cognateness}\) (\(\mathcal{M}_{2}\)),
+\(\text{Exposure} \times \text{Cognateness}\) (\(\mathcal{M}_{3}\)), and
+finally \(\text{Age} \times \text{Cognateness}\) and
+\(\text{Age} \times \text{Exposure} \times \text{Cognateness}\)
+(\(\mathcal{M}_{4}\)). We compared the predictive performance of the
+models using Bayesian leave-one-out cross-validation (LOO-CV, Vehtari et
+al., 2017)\footnote{Due to computational constraints, we performed
+  LOO-CV based on a random sub-sample of 700 samples of the posterior
+  distribution of the fixed effects (Magnusson et al., 2020)}. LOO-CV
+involves computing the likelihood of each data point, given a model
+fitted excluding such data point. The likelihoods generated by the model
+for each of the data points are summarised together in a measure of
+predictive performance: the expected log-predicted density (ELPD). This
+measure penalises the complexity of the model ELPD. ELPD values closer
+to zero indicate better predictive accuracy. We compared the ELPD of
+each model, and selected for interpretation the one showing the best
+performance (see Appendix C for more details about the models).
+
+We assessed the practical relevance of the estimated regression
+coefficients of the selected model following Kruschke and Liddell
+(2018). First, we specified a region of practical equivalence (ROPE)
+from -2.5\% to +2.5\%, in the probability scale. This region indicates
+the range of values that we considered equivalent to zero. The degree of
+overlap between the posterior distribution of a regression coefficient
+with the can be interpreted as support for the true parameter of the
+coefficient being zero or equivalent. We then computed the 95\% highest
+density interval (HDI) of the posterior distribution of each regression
+coefficient. This interval contains the true value of this coefficient
+with 95\% probability, given the data. Finally, we calculated the
+proportion of posterior samples in the 95\% HDI that fell into the ROPE,
+which indicates the probability that the true value of the regression
+coefficient falls into the ROPE, and should therefore be considered
+equivalent to zero. For example, a 90\% overlap between the HDI of a
+coefficient's posterior distribution and the indicates that, given our
+data, there is a 90\% probability that the true value of the coefficient
+falls within the ROPE, and is therefore equivalent to zero.
+
+Data processing and visualisation was done in R (R Core Team, 2013,
+version 4.2.2) using the Tidyverse family of packages (Wickham et al.,
+2019). Bayesian modelling was done using the \texttt{brms} (Bürkner,
+2017; Carpenter et al., 2017), \texttt{loo} (Vehtari et al., 2022),
+\texttt{tidybayes} (Kay, 2021) and \texttt{marginaleffects}
+(Arel-Bundock, 2022) R packages.
+
+\hypertarget{sec-results}{%
+\section{Results}\label{sec-results}}
+
+All models showed adequate convergence diagnostics and little evidence
+of multicollinearity between predictors (see Appendix D for details).
+\(\mathcal{M}_{3}\), which did \emph{not} include the
+\(\text{Age} \times \text{Exposure} \times \text{Cognateness}\)
+three-way interaction, showed the best predictive performance with a
+difference in expected log-predicted density (ELPD) compared to the
+second best-performing model (\(\mathcal{M}_{4}\)) of -109.17 \(\pm\)
+40.05. Given our data, the predictions of the \(\mathcal{M}_{3}\) are
+confidently more accurate than those of the other models (see
+Table~\ref{apatb-results-loos} for all model comparisons). Both
+\(\mathcal{M}_3\) and \(\mathcal{M}_4\) performed considerably better
+than the rest of the models, which did not include the
+\(\text{Exposure} \times \text{Cognateness}\) two-way interaction. This
+points to the effect of \(\text{Cognateness}\) being strongly modulated
+by \(\text{Exposure}\), and that this interaction remains relatively
+stable across ages.
+
+\begin{table}
+\caption{Bayesian leave-one-out cross validation (LOO-CV).}
+\label{apatb-results-loos}
+
+\begin{longtable*}{l|rr}
 \toprule
- &  & \multicolumn{3}{c}{Frequency (Zipf)} & \multicolumn{3}{c}{\# Phonemes} & \multicolumn{3}{c}{Phon. similarity (Levenshtein)} \\ 
-\cmidrule(lr){3-5} \cmidrule(lr){6-8} \cmidrule(lr){9-11}
- & \emph{N} & Mean & \emph{SD} & Range & Mean & \emph{SD} & Range & Mean & \emph{SD} & Range \\ 
+\multicolumn{1}{l}{} & ELPD \$\$\textbackslash{}pm\$\$ SE & Diff. \$\$\textbackslash{}pm\$\$ SE \\ 
 \midrule
-\multicolumn{11}{l}{Catalan} \\ 
-\midrule
-A & 278 & $4.39$ & $0.63$ & $2.61$–$6.52$ & $5.28$ & $1.73$ & 2–14 & $37.32\%$ & $26.94\%$ & $0.00\%$–$100.00\%$ \\ 
-B & 274 & $4.39$ & $0.69$ & $1.90$–$6.31$ & $5.20$ & $1.77$ & 2–14 & $35.96\%$ & $24.74\%$ & $0.00\%$–$100.00\%$ \\ 
-C & 270 & $4.36$ & $0.65$ & $2.49$–$6.54$ & $5.46$ & $1.88$ & 2–14 & $36.51\%$ & $26.11\%$ & $0.00\%$–$100.00\%$ \\ 
-D & 269 & $4.38$ & $0.61$ & $2.10$–$6.10$ & $5.29$ & $1.70$ & 2–14 & $36.33\%$ & $26.40\%$ & $0.00\%$–$100.00\%$ \\ 
-\midrule
-\multicolumn{11}{l}{Spanish} \\ 
-\midrule
-A & 277 & $4.25$ & $0.74$ & $2.28$–$6.76$ & $5.66$ & $1.63$ & 3–14 & $37.33\%$ & $26.89\%$ & $0.00\%$–$100.00\%$ \\ 
-B & 281 & $4.28$ & $0.79$ & $2.16$–$6.14$ & $5.67$ & $1.68$ & 2–14 & $36.18\%$ & $24.95\%$ & $0.00\%$–$100.00\%$ \\ 
-C & 278 & $4.24$ & $0.77$ & $1.68$–$6.26$ & $5.79$ & $1.79$ & 3–14 & $36.53\%$ & $26.24\%$ & $0.00\%$–$100.00\%$ \\ 
-D & 273 & $4.24$ & $0.72$ & $1.98$–$6.14$ & $5.72$ & $1.68$ & 2–14 & $36.59\%$ & $26.55\%$ & $0.00\%$–$100.00\%$ \\ 
+Model 3 & $-142,586.09$ ± $317.78$ & $0.00$ ± $0.00$ \\ 
+Model 4 & $-142,695.26$ ± $319.90$ & $-109.17$ ± $40.05$ \\ 
+Model 1 & $-143,112.89$ ± $315.28$ & $-526.80$ ± $63.02$ \\ 
+Model 2 & $-143,152.04$ ± $321.28$ & $-565.95$ ± $33.74$ \\ 
+Model 0 & $-144,709.07$ ± $317.31$ & $-2,122.98$ ± $83.59$ \\ 
 \bottomrule
-\end{longtable}
-
-\hypertarget{analysis}{%
-\subsection{Data analysis}\label{analysis}}
-
-We gathered 258,095 observations, with each observation corresponding to
-a single response of one participant to a given item in the
-questionnaire). These observations correspond to 395 distinct
-participants (of which 343 participated once, 33 twice, 14 three times,
-and 5 four times), responding to 1199 distinct items (597 in Catalan,
-602 in Spanish, 602 translation equivalents). Translation equivalents
-(TEs) received a median of 236 responses (\emph{Min} = 118, \emph{Max} =
-942), both languages summed together.
-
-We modelled the probability of answering each response category
-(\emph{No} \textless{} \emph{Understands} \textless{} \emph{Understands
-and Says}) using a Bayesian, mixed-effects ordinal regression model.
-This allowed us to simultaneously estimate the impact of item- and
-participant-level predictors on the probability of word comprehension
-and production. We included \(Age\), \(Frequency\), \(Phonemes\),
-\(DoE\), and \(Levenshtein\) as fixed effects. We also included the
-\(Age \times DoE \times Levenshtein\) three-way interaction, and its
-corresponding two-way interactions. We added participants and items as
-crossed random effects including all intercepts and slopes when
-appropriate (Barr et al., 2013).
-
-This model structure, including word- and participant-level properties
-to estimate vocabulary acquisition trajectories was discussed by
-Kachergis et al. (2022) in the context of explanatory item response
-models. Our model differs from their proposal in that 1) while they
-modelled global trajectories exclusively, we modelled individual
-trajectories as well, 2) our outcome variable is ordinal in contrast to
-the binary outcome used in their study, 3) we propose a Bayesian
-approach. Our motivation for a Bayesian approach towards statistical
-inference is that (1) it allows us to incorporate previous domain
-knowledge to improve estimation accuracy, and (2) to quantify the
-uncertainty associated with the estimated parameters in our model
-(McElreath, 2018).
-
-The Bayesian approach implies using Bayes' rule to assign a probability
-to possible parameter values (i.e., posterior distribution), based on
-previous knowledge about such probability distribution (prior
-distribution) and our data (likelihood). We used Markov chain Monte
-Carlo (MCMC) to approximate the posterior distribution via sampling. We
-run 2 sampling chains with 4,000 iterations each, including 2,000
-warm-up iterations per chain. All \(\hat{R}\) values were smaller than
-1.1, indicating that the chains converged successfully in all cases
-(Gelman \& Rubin, 1992).
-
-We compared a base model---only including the intercept, \(Age\),
-\(Frequency\), \(Phonemes\), and \(DoE\) as fixed effects, and their
-corresponding random effects-\/--to increasingly more complex models,
-introducing predictors in the following order: \(Age \times DoE\),
-\(Levenshtein\), \(DoE \times Levenshtein\), and finally
-\(Age \times Levenshtein\) and \(Age \times DoE \times Levenshtein\). We
-compared the predictive performance of the models using Bayesian
-leave-one-out cross-validation (LOO) (Vehtari et al., 2017): we computed
-the average likelihood of each observation after estimating the model's
-parameters leaving that same observation out of the data set. Due to
-computational constraints, we performed LOO based on a random sub-sample
-of 700 samples (out of 2,000) of the posterior distribution of the fixed
-effects of the model. and selected the model with the smallest expected
-log posterior density (ELPD)\footnote{Due to the high computational cost
-  associated with our large dataset, we used a sub-sampling approach for
-  performing Bayesian LOO with 700 posterior samples (Magnusson et al.,
-  2019)}.
-
-Finally, we explored the posterior distribution of each parameter in the
-selected model. Following Kruschke \& Liddell (2018), We specified a
-region of practical equivalence (\emph{ROPE}) around the values of each
-regression coefficient that we considered equivalent to zero (Kruschke
-\& Liddell, 2018). We defined our \emph{ROPE} from -0.18 to +0.18 {[}see
-@\#appendix-rope for justification of the \emph{ROPE}{]}. In the results
-section, we report the mean of each coefficient's marginal posterior
-distribution along with their 95\% highest density intervals
-(\emph{HDI}). This interval indicates the narrowest range of values that
-contain 95\% of the samples of the posterior distribution, and therefore
-the true value of the coefficient with 95\% confidence, given our data.
-We also report the proportion of the posterior samples of that
-coefficient that fell within the \emph{ROPE}. A 90\% overlap between the
-95\% \emph{HDI} of the coefficients' posterior distribution and the
-\emph{ROPE} indicates that, given the data, there is a 90\% probability
-that the true value of the coefficient falls within the \emph{ROPE}, and
-is therefore equivalent to zero, that is, negligible.
-
-Data processing and visualisation was done in R (R Core Team, 2020)
-using the \texttt{tidyverse} family of packages (Wickham et al., 2019),
-Bayesian modelling was done using the \texttt{brms} (Bürkner, 2017;
-Carpenter et al., 2017), \texttt{loo} (Yao et al., 2017), and
-\texttt{tidybayes} (Kay, 2021) R packages.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{comparison}{%
-\subsection{Model comparison and selection}\label{comparison}}
-
-Table~\ref{tbl-results-loos} shows the outputs of the \emph{LOO-CV}. The
-extended model (\texttt{model\_fit\_1}), which included the three-way
-interaction between \(Age\), \(DoE\), and \(Levenshtein\) showed the
-best predictive performance, with a difference in \emph{ELPD} compared
-to the other models several times larger than the standard error of such
-difference. This indicates that, given the data, the predictions of the
-extended model are confidently more accurate than those of the other
-models.
-
-\hypertarget{tbl-results-loos}{}
-\setlength{\LTpost}{0mm}
-\begin{longtable}{lrrrrrrrr}
-\caption{\label{tbl-results-loos}Bayesian leave-one-out cross validation (LOO-CV) }\tabularnewline
-
-\caption*{
-{\large } \\ 
-{\small The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (\emph{ELPD}), which penalises the complexity of each model.
-\emph{ELPD} values closer to zero indicate better predictive acuracy.
-We report the \emph{ELPD}, effective number of parameters (\emph{p}), and information criterion (\emph{IC}) of each model along their associated standard errors (\emph{SE}).
-The last column indicates the difference in \emph{ELPD} between each model and the model with the best predictive accuracy (\texttt{model\_fit\_1})}
-} \\ 
-\toprule
- & \multicolumn{2}{c}{LOOELPD} & \multicolumn{2}{c}{LOOp} & \multicolumn{2}{c}{LOOIC} & \multicolumn{2}{c}{LOOdiff} \\ 
-\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
-Model & \emph{ELPD} & \emph{SE} & \emph{p} & \emph{SE} & \emph{LOO-IC} & \emph{SE} & \emph{diff} & \emph{SE} \\ 
-\midrule
-model\_fit\_3 & $-244,590.04$ & $388.23$ & $1,688.79$ & $5.39$ & $489,180.08$ & $776.46$ & $0.00$ & $0.00$ \\ 
-model\_fit\_0 & $-247,891.36$ & $353.08$ & $1,260.47$ & $3.24$ & $495,782.71$ & $706.16$ & $-3,301.32$ & $161.73$ \\ 
-model\_fit\_2 & $-248,274.01$ & $353.95$ & $1,489.46$ & $3.65$ & $496,548.01$ & $707.91$ & $-3,683.97$ & $152.45$ \\ 
-model\_fit\_1 & $-249,630.81$ & $358.40$ & $1,649.76$ & $4.03$ & $499,261.61$ & $716.79$ & $-5,040.77$ & $170.07$ \\ 
-\bottomrule
-\end{longtable}
-\begin{minipage}{\linewidth}
-Pareto-\emph{k} estimates of all models were acceptable (\emph{k} \textless{} 0.5)\\
-\end{minipage}
-
-\hypertarget{coefficients}{%
-\subsection{Regression coefficients}\label{coefficients}}
-
-\hypertarget{tbl-results-fixed}{}
-\begin{longtable}{lrrr}
-\caption{\label{tbl-results-fixed}Summary of the estimated posterior distribution of fixed regression
-coefficients }\tabularnewline
-
-\toprule
-Parameter & Median & 95\% \emph{HDI} & \emph{ROPE} prob. \\ 
-\midrule
-Intercept (Comprehension and Production at 22 months) & $-0.85$ & [$-0.97$, $-0.72$] & $0.00\%$ \\ 
-Intercept (Comprehension at 22 months) & $1.55$ & [$1.44$, $1.69$] & $0.00\%$ \\ 
-Age (+1 SD, 5.72 months) & $1.32$ & [$1.17$, $1.47$] & $0.00\%$ \\ 
-Frequency (+1 SD, 0.7 Zipf) & $0.16$ & [$0.07$, $0.24$] & $32.11\%$ \\ 
-Phonemes (+1 SD, 1.75 phonemes) & $-0.20$ & [$-0.28$, $-0.13$] & $2.79\%$ \\ 
-DoE (+1 SD, 30\%) & $0.89$ & [$0.87$, $0.91$] & $0.00\%$ \\ 
-Levenshtein (+1 SD, 26\%) & $0.03$ & [$-0.09$, $0.13$] & $100.00\%$ \\ 
-DoE × Levenshtein & $-0.20$ & [$-0.22$, $-0.19$] & $0.00\%$ \\ 
-Age × DoE & $0.13$ & [$-0.02$, $0.28$] & $57.42\%$ \\ 
-Age × Levenshtein & $0.07$ & [$0.03$, $0.11$] & $100.00\%$ \\ 
-Age × DoE × Levenshtein & $-0.06$ & [$-0.09$, $-0.02$] & $100.00\%$ \\ 
-\bottomrule
-\end{longtable}
-
-Table~\ref{tbl-results-fixed} shows the summary of the posterior
-distribution of the fixed regression coefficients of the model. For
-interpretability, we supplement the summaries of each posterior
-regression coefficient with its transformation to the probability scale.
+\end{longtable*}
+
+\tablenote{We report the ELPD of each model along their associated standard errors (SE).
+The last column indicates the difference in ELPD between each model and the model with the best predictive accuracy (Model 3).
+Pareto-k estimates of all models were acceptable (k < 0.5).}
+
+\end{table}
+
+Table~\ref{apatb-coefs} shows the summary of the posterior distribution
+of the fixed regression coefficients of \(\mathcal{M}_{3}\), and their
+degree of overlap with the ROPE. For interpretability, we report the
+estimated regression coefficients transformed to the probability scale.
 The resulting values correspond to the maximum difference in probability
@@
 Production}) that corresponds to a one standard deviation change in each
-predictor\footnote{The logit and probability scales related
-  non-linearly. This means that one logit difference is not necessarily
-  translated to a unique value in the probability scale. For example,
-  the probability of acquisition of a given word might increase in 5\%
-  when age increases from 22 to 23 months, the probability of
-  acquisition of the same word might only increase in 0.2\% when age
-  increases from 34 to 36 months. The linear growth of the probability
-  of acquisition differs along the logistic curve, and therefore
-  deciding the age point at which to report the estimates of the
-  regression coefficients in the probability scale is not trivial. We
-  followed the suggestion by Gelman et al. (2020) and report the maximum
-  value of such coefficient, which corresponds to the linear growth
-  (i.e.~derivative) of the logistic curve at the age at which most
-  participants were acquiring a given word. This value can be
+predictor\footnote{The logit and probability scales relate non-linearly.
+  This means that one logit difference is not necessarily translated to
+  a unique value in the probability scale. For example, the probability
+  of acquisition of a given word might increase in 5\% when age
+  increases from 22 to 23 months, the probability of acquisition of the
+  same word might only increase in 0.2\% when age increases from 30 to
+  32 months. The linear growth of the probability of acquisition differs
+  along the logistic curve, and therefore deciding the age point at
+  which to report the estimates of the regression coefficients in the
+  probability scale is not trivial. Following Gelman et al. (2020), we
+  report the maximum value of such coefficient, which corresponds to the
+  linear growth (i.e.~derivative) of the logistic curve at the age at
+  which most participants were acquiring a given word. This value can be
   approximated by dividing the coefficient in the logit scale by four:
@@
 
-\(Age\) was the predictor with the largest effect on the probability of
-acquisition (\(\beta\) = 1.32, 95\% \emph{HDI} = {[}1.17, 1.47{]}): at
-the steepest point of the acquisition trajectory of an average
-translation equivalent, the probability of acquisition increased 5.77\%
-every month. Lexical frequency (\(Frequency\)) had a small, positive,
-but inconclusive on the probability of acquisition substantially,
-(\(\beta\) = 15.85\%, 95\% \emph{HDI} = {[}7.12\%, 23.62\%{]}): 32.11\%
-of its posterior samples overlapped with the \emph{ROPE.} The 95\%
-\emph{HDI} of the regression coefficient of the number of phonemes
-(\(Phonemes\)) did exclude the \emph{ROPE}, although not by much: 2.79\%
-of its posterior samples fell within the \emph{ROPE}. This suggests that
-the number of phonemes in the word form had a small effect on the
-probability of acquisition (\(\beta\) = -20.28\%, 95\% \emph{HDI} =
-{[}-28.12\%, -12.95\%{]}). For every increase in phoneme length, the
-probability of acquisition decreased in -2.90\%.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-results-fixed-1.pdf}
-
-}
-
-\caption{\label{fig-results-fixed}Marginal posterior distribution of the
-regression coefficients of the fixed effects in the extended model.
-Distributions show the estimated likelihood density of each value in the
-parameter space (X-axis) of each coefficient (Y-axis). Intervals show
-the 95\% highest density interval (\emph{HDI}) of each distribution.
-This interval is the narrowest range of values that contains 95\% of the
-distribution, and therefore is 95\% likely to contain the true value,
-given the data. the mean and HDI limits are indicated below each
-distribution. Coefficients were transformed from the logit scale to the
-probability scale for interpretability. We used the divide-by-four rule
-to get the maximum change in probability of correct response associated
-with a unit increase in this variable, i.e.~the derivative of the
-logistic function in its mid-point.}
+\begin{table}
+\caption{Posterior distribution of regression coefficients Model 3.}
+\label{apatb-coefs}
+
+\begin{longtable*}{l|rrr}
+\toprule
+\multicolumn{1}{l}{} & Median & 95\% HDI & ROPE prob. \\ 
+\midrule
+\multicolumn{4}{l}{Intercepts (at 22 months)} \\ 
+\midrule
+Comprehension and Production & $0.21$ & [$0.19$, $0.23$] & $0.00\%$ \\ 
+Comprehension & $0.84$ & [$0.82$, $0.85$] & $0.00\%$ \\ 
+\midrule
+\multicolumn{4}{l}{Slopes} \\ 
+\midrule
+Age (+1 SD, 4.81, months) & $0.39$ & [$0.34$, $0.44$] & $0.00\%$ \\ 
+Exposure (+1 SD, 1.81) & $0.20$ & [$0.17$, $0.23$] & $0.00\%$ \\ 
+Length (+1 SD, 1.56 phonemes) & $-0.08$ & [$-0.10$, $-0.05$] & $0.00\%$ \\ 
+Cognateness (+1 SD, 26\%) & $0.06$ & [$0.02$, $0.10$] & $0.42\%$ \\ 
+Age × Exposure & $0.06$ & [$0.03$, $0.10$] & $0.00\%$ \\ 
+Exposure × Cognateness & $-0.04$ & [$-0.05$, $-0.03$] & $0.00\%$ \\ 
+\bottomrule
+\end{longtable*}
+
+\tablenote{Median: median of the posterior distribution in the probability scale.
+95\% HDI: 95\% highest density interval of the distribution; narrowest range of values that contains 95\% of the distribution, and therefore is 95\% likely to contain the true value, given the data.
+On top of each distribution, we indicate the proportion of posterior samples in the 95\% HDI that fall into the region of practical equivalence (ROPE).
+This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
+Lower values indicate that the true value is unlikely to be zero of equivalent to zero.}
+
+\end{table}
+
+The main effect of \(\text{Age}\) showed the strongest association with
+the probability of acquisition (\(\beta\) = 0.39, 95\% HDI = {[}0.34,
+0.44{]}), with all of its posterior samples falling out of the ROPE. A
+one-month increment in age increased a maximum of 0.08 the probability
+of acquisition. Similarly, the word exposure index (\(\text{Exposure}\))
+had a strong effect on the probability of acquisition (\(\beta\) = 0.2,
+95\% HDI = {[}0.17, 0.23{]}). All of the posterior samples of this
+regression coefficient excluded the ROPE. The impact of this predictor
+on the probability of acquisition was positive: for every standard
+deviation increase in exposure, the participant was 0.11 more likely to
+acquire it. Word-form length also showed a significant association with
+probability of acquisition (\(\beta\) = -0.08, 95\% HDI = {[}-0.1,
+-0.05{]}). For every phoneme in the word-form, participants were -0.05
+less likely to know it. The 95\% HDI of the regression coefficient of
+the \(\text{Age} \times \text{Exposure}\) interaction also excluded the
+ROPE (\(\beta\) = 0.06, 95\% HDI = {[}0.03, 0.1{]}), showing that the
+effect of the word exposure index differed across ages: older children
+were more sensitive likely to acquire words with higher exposure rate
+than younger children.
+
+The posterior distribution of the main effect of cognateness excluded
+the ROPE completely (\(\beta\) = 0.06, 95\% HDI = {[}0.02, 0.1{]}). For
+every 10\% increment in cognateness, the acquisition of a word increased
+in 0.01. The effect of \(\text{Cognateness}\) interacted with that of
+\(\text{Exposure}\): the 95\% HDI of the regression coefficient of
+interaction excluded the ROPE entirely (\(\beta\) = -0.04, 95\% HDI =
+{[}-0.05, -0.03{]}), suggesting that the effect of cognateness on a
+word's probability of acquisition changed depending on participants'
+exposure to word. Follow-up analyses on this interaction showed that
+when exposure rate was low (e.g., -1 SD), cognateness increased the
+probability of acquisition substantially. This effect was negligible
+when for words with median or high exposure (+1 SD) (see
+Figure~\ref{apafg-results-marginal}).
+
+\begin{figure}[h!]
+\caption{Posterior marginal effects of Model 3.
+Thin lines correspond to 50 mean predictions generated from 50 MCMC samples of the posterior distribution of the model.
+Thicker lines indicate the median of those predictions.
+Different colours indicate different levels of phonological similarity (as indicated by the Cognateness similarity between pairs of translation equivalents).
+Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
+Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom.
+In-sample predictions lie inside the grey rectangles.}
+\label{apafg-results-marginal}
+\includegraphics[width=8 in]{manuscript_files/figure-pdf/apafg-results-marginal-1.pdf}
 
@@
 
-The degree of exposure to the language (\(DoE\)) had a strong effect on
-the probability of acquisition (\(\beta\) = 89.26\%, 95\% \emph{HDI} =
-{[}87.10\%, 91.34\%{]}). All of the posterior samples of his regression
-excluded the \emph{ROPE}. The impact of this predictor on the
-probability of acquisition was positive: for every 10\% increase in
-exposure to the language the word belongs to, the participant was
-29.76\% more likely to acquire it. The effect of phonological similarity
-by itself (as indicated by the regression coefficient of the main effect
-of \(Levenshtein\)) was equivalent to zero (\(\beta\) = 2.76\%, 95\%
-\emph{HDI} = {[}-8.61\%, 12.88\%{]}), as 100.00\% of the posterior
-samples of its regression coefficient fell into the \emph{ROPE}.
-However, the 95\% \emph{HDI} of the regression coefficient of the
-\(Doe \times Levenshtein\) interaction excluded the \emph{ROPE} entirely
-(\(\beta\) = -20.20\%, 95\% \emph{HDI} = {[}-21.68\%, -18.83\%{]}),
-suggesting that the effect of phonological similarity on a word's
-probability of acquisition changed depending on participants' exposure
-to the language it belonged to. Follow up analyses on this interaction
-(see Figure~\ref{fig-results-marginal}) showed that, when exposure to
-the language was low (e.g., 10\%), phonological similarity increased the
-probability of acquisition substantially. This effect was negligible
-when exposure to both languages was balanced (e.g.~50-50\%). Finally,
-when exposure to the language was high (e.g., 90\%), this effect was
-very close to zero, although with different direction, with phonological
-less similar translation equivalents being more likely to be acquired
-than phonologically more similar ones.
-
-The 95\% \emph{HDI} of the regression coefficient of the
-\(Age \times DoE\) interaction did not exclude the \emph{ROPE}
-(\(\beta\) = 12.63\%, 95\% \emph{HDI} = {[}-1.50\%, 28.48\%{]}): 57.42\%
-of its posterior samples fell within the \emph{ROPE}. This shows that
-the effect of the degree of exposure on the speed at which participants
-acquired word across ages was inconclusive. Give our data, it is
-difficult to tell whether this \emph{DoE} increased or decreased the
-slope of the words' acquisition trajectories. The 95\% \emph{HDI} of the
-regression coefficient of the \(Age \times Levenshtein\) interaction
-overlapped completely with the \emph{ROPE} (\(\beta\) = 6.59\%, 95\%
-\emph{HDI} = {[}2.66\%, 10.84\%{]}), indicating that phonological
-similarity did not speed-up the acquisition trajectories of the words
-across ages. Finally, the 95\% \emph{HDI} of the regression coefficient
-of the \(Age \times DoE \times Levenshtein\) interaction also overlapped
-completely with the \emph{ROPE} (\(\beta\) = , 95\% \emph{HDI} =
-{[}-9.32\%, -2.33\%{]}), suggesting that the impact of the degree of
-exposure to the language on the effect of phonological similarity did
-not differ substantially across ages.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-results-marginal-1.pdf}
-
-}
-
-\caption{\label{fig-results-marginal}Expected posterior predictions.
-Posterior poreditions for the extended model were generated by computing
-the probability acquisition that resulted from feeding the equation of
-the regression model with values of the posterior distribution of each
-coefficient. These values correspond to the 4,000 samples of the
-posterior distribution that were drawn during the MCMC estimation of the
-model. Therefore, we generated 4,000 predictions for a series of
-combination of levels of interest. Each line and interval and in the
-figure summarises the mean and 95\% \emph{HDI} of each combination of
-levels. The X-axis indicates the age (in months) for which the
-prediction is generated. The Y-axis indicates the predicted probability
-of acquisition (\emph{Comprehension} or \emph{Comprehension and
-Production}). Different colours indicate different levels of
-phonological similarity (as indicated by the Levenshtein similarity
-between pairs of translation equivalents). Finally, predictions are
-presented separately for different degrees of exposure (10\%): little
-exposure to the language (10\% DoE, traditionally classified as
-monolingual), balanced exposure to both languages (50\%, traditionally
-classified as bilingual), and high exposure to the language (90\%,
-traditionally classified as monolingual). Predictions for
-\emph{Comprehension} are show on top and predictions for
-\emph{Comprehension and Production} are shown on the bottom.}
-
-\end{figure}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\hypertarget{summary-of-the-study}{%
-\subsection{Summary of the study}\label{summary-of-the-study}}
-
-This study explored how word acquisition was affected by cognateness in
-bilingual toddlers. We collected vocabulary data from children learning
-Catalan and/or Spanish in differing degrees of exposure to each, and
-then measured the probability of each word being acquired at different
-ages as a function of (1) participants' exposure to the language it
-belongs to, and (2) its phonological overlap with its translation
-equivalent in the other language. We build a multilevel Bayesian model
-that estimated the effect of both predictors and their interaction,
-adjusting for other variables known to be relevant during the
-acquisition of words: participants' age, lexical frequency, and length
-in phonemes.
-
-\hypertarget{summary-of-the-results}{%
-\subsection{Summary of the results}\label{summary-of-the-results}}
-
-\hypertarget{results-in-contrast-with-the-literature}{%
-\subsection{Results in contrast with the
-literature}\label{results-in-contrast-with-the-literature}}
-
-\hypertarget{mechanistic-explanations}{%
-\subsection{Mechanistic explanations}\label{mechanistic-explanations}}
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-\hypertarget{further-steps}{%
-\subsection{Further steps}\label{further-steps}}
-
-\hypertarget{references}{%
-\section{References}\label{references}}
+\hypertarget{sec-discussion}{%
+\section{Discussion}\label{sec-discussion}}
+
+The present study explored the cognateness facilitation effect on
+bilingual word acquisition. We found such effect to be mediated by
+lexical frequency and degree of language exposure: phonologically
+similar translation equivalents were on average acquired earlier in age
+compared to phonologically dissimilar translation equivalents, but this
+effect was only observed for word-forms to which participants were
+exposed less often, i.e., words with low lexical frequency and/or words
+from the language of lower daily exposure. Our results are relevant in
+the light of two previous findings. First, Floccia, Sambrook, Delle
+Luche, et al. (2018) reported that bilingual toddlers learning two
+typologically close languages (which shared many cognates, e.g.,
+English-German) showed larger vocabulary sizes than those learning
+typologically more distance languages (which shared fewer cognates,
+e.g., English-Mandarin). Second, Mitchell et al. (2022) found an earlier
+age-of-acquisition for cognates, compared to non-cognates. The outcomes
+of both studies pointed to cognateness facilitating word acquisition,
+but the underpinnings of such effect were unclear.
+
+In this study, we capitalised on accumulator models of language
+acquisition to propose a theoretical account of bilingual word learning,
+in which cognateness interacts with lexical frequency and language
+exposure to boost the acquisition of translation equivalents. We
+suggested that words benefit from their cognate status only \emph{after}
+their translation equivalent has been acquired: after the child acquires
+a word-form, the number of learning instances they need to acquire its
+translation decreases. We further hypothesised that such facilitation is
+proportional to the phonological similarity between both word-forms. If
+two translation equivalents sharing a fair deal of phonological
+similarity (e.g., /\textipa{"pOR.t@}/-/\textipa{"pweR.ta}/, 50\%
+Levenshtein similarity), learning of one of them should facilitate the
+acquisition of the other more strongly, compared to when both word-forms
+share only few phonemes (e.g.,
+/\textipa{k@"Di.R@}/-/\textipa{"si.\textlambda a}/ , 25\%).
+
+Under such scenario, we predicted that words from the language of lower
+exposure would benefit more often from their cognate status than those
+from the language of higher exposure. The rationale behind such
+prediction was that children are less frequently exposed to word-forms
+in the former language than to words in the latter (Cattani et al.,
+2014; Thordardottir et al., 2006). Words in the language of lower
+exposure are therefore more likely to be acquired later in age than
+their translations in the other language. For instance, an infant
+exposed to Catalan and Spanish 65\% and 35\% of the time, respectively,
+is more likely to learn the Spanish word-form /\textipa{"ga.to}/ later
+in age than its Catalan translation /\textipa{gat}/. If this is the
+case, /\textipa{"ga.to}/ is more likely to receive facilitation from the
+acquisition of /\textipa{gat}/ than \emph{vice versa} (see
+Figure~\ref{apafg-hypotheses} for a graphic illustration). In line with
+these predictions, we found that cognateness facilitated word
+acquisition more strongly in words from the language of lower exposure.
+This asymmetric facilitation effect provides support for the hypothesis
+that words benefit from their cognate status \emph{after} their
+translation has been acquired. Overall, our results provide support for
+an interactive account of bilingual vocabulary learning in which word
+representations from one language contribute to the acquisition of words
+in the other.
+
+These findings are, to a large extent, in line with previous literature
+on word acquisition. First, words-forms with fewer phonemes like
+/\textipa{sO\textltilde}/ (Catalan for \emph{sun}) were on average more
+likely to be acquired than word-forms with more phonemes like
+/\textipa{ku.ku"DRi\textltilde}/ (Catalan for \emph{crocodile}). These
+results converge with previous studies found word length to be more
+strongly associated with receptive vocabulary than with expressive
+vocabulary (e.g., Braginsky et al., 2019; Jones \& Brandt, 2019). This
+facilitation is most likely due to the fact that shorter word-forms are
+generally simpler to encode as phonological representations in the
+lexicon than longer word-forms. Shorter word-forms are easier to
+produce, and therefore to be reported as acquired in parental vocabulary
+checklists.
+
+Our participants were also more likely to know words to which they were
+more frequently exposed to, replicating previously reported effects of
+lexical frequency (e.g., Braginsky et al., 2019), and bilingual language
+exposure (e.g., Cattani et al., 2014; Thordardottir et al., 2006). This
+strong effect of word frequency is a key component of accumulator models
+of word acquisition, which suggest that the age-of-acquisition of a word
+is, to a large extent, a function of the child's rate of exposure to
+such word, which conspires with child- and word-level factors in the
+process of forming a representation of the word in the child's lexicon
+(Hidaka, 2013; Mollica \& Piantadosi, 2017). Our results provide ground
+for the central role of word frequency in accumulator models to be
+extended to the bilingual case, in the form of a composite measure of
+word rate of exposure that accounted for both lexical frequency and
+language exposure (Cattani et al., 2014; Floccia, Sambrook, Luche, et
+al., 2018b). Interestingly, our findings pointed to this predictor
+playing a stronger role during earlier ages: younger infants benefited
+from higher exposure rates at earlier ages, suggesting that early word
+acquisition is supported by mere exposure to word forms more strongly
+than later word acquisition.
+
+Finally, we found a frequency-mediated facilitation effect of
+cognateness, adding some nuance to the findings from previous studies
+(Bosch \& Ramon-Casas, 2014; Floccia, Sambrook, Luche, et al., 2018b;
+Mitchell et al., 2022; Schelletter, 2002). Only words to which children
+were exposed to less frequently benefited from their cognate status.
+These results favour the account that a word-form benefits from its
+phonological similarity with its translation only after such translation
+has been acquired first. On the one hand, our results provide converging
+evidence for Floccia, Sambrook, Luche, et al. (2018b), and arguably
+account for the language distance facilitation effect they found in the
+additional (non-English) language of their bilingual sample. Most
+children in their sample were English-dominant, and therefore their
+exposure to English was higher than their exposure to the additional
+language. However, it was in the vocabulary of the additional language
+where the authors found a facilitation effect of cognateness: bilingual
+children learning two languages that shared many cognateness (e.g.,
+English-German) had larger vocabularies in their additional language
+than those bilingual who were learning two languages that shared less
+cognates (e.g., English-Chinese). Given our results, we argue that it is
+language exposure that mediates the relationship between cognateness and
+vocabulary size in Floccia, Sambrook, Luche, et al. (2018b)'s sample.
+
+Overall, our findings parallel cognates' significant role in many models
+of bilingual language processing. The importance of cognates in these
+model stems in the fact that cognates provide evidence for a language
+non-selective account of lexical processing, in which bilinguals
+activate unselected lexical representations in both languages, even in
+fully monolingual situations (Christoffels et al., 2007; Thierry \& Wu,
+2007). As mentioned in the introduction, naming pictures whose labels
+are cognates leads to shorter latencies than when labels are
+non-cognates, which suggests that phonological similarity across
+translations facilitates production (Costa et al., 2000; Hoshino \&
+Kroll, 2008). When acting as distractors in naming tasks, cognates exert
+larger interference effects than non-cognate distractors (Colomé \&
+Miozzo, 2010; Hoshino \& Thierry, 2011). Both findings suggest that the
+non-selected translation of the target in Costa et al. (2000), or of the
+distractor in Colomé and Miozzo (2010) was activated, and that when such
+word-form was phonologically related to its translation, the
+facilitation or interference effect was stronger.
+
+Evidence supporting the language-non selective account of lexical access
+is not restricted to production or the auditory modality, but rather
+extend to comprehension and translation, and to the visual modality
+(reading and signing) (Gimeno-Martínez et al., 2021; Hell \& Groot,
+2008; Morford et al., 2011; Schwartz et al., 2007; Schwartz \& Kroll,
+2006; Spivey \& Marian, 1999; see Kroll et al., 2013 for review). For
+instance, Thierry and Wu (2007) found neurophysiological evidence of
+Chinese-English bilinguals activating both languages in a word
+comprehension task. In this experiment, participants were presented with
+couples of words in English, and were asked to respond if the words were
+related in meaning. Unbeknownst to participants, the translations to
+Chinese of half of such word pairs shared a character. Participants
+would be presented (visually in a first experiment and auditorily in a
+second experiment) with an English word pair `train-ham', which are
+semantically unrelated, but share a character in their Chinese
+translations (\foreignlanguage{chinese-simplified}{火車} {[}Huo
+Che{]}-\foreignlanguage{chinese-simplified}{火腿} {[}Huo Tui{]}).
+Bilinguals' brain signal differed across words pairs with and without
+translation overlap (e.g., `apple-table',
+\foreignlanguage{chinese-simplified}{平果} {[}Ping
+Guo{]}--\foreignlanguage{chinese-simplified}{桌子} {[}Zhuo Zi{]}). This
+suggests that participants were sensitive to the form-similarity of the
+Chinese translation of the word-pairs they were presented, despite the
+whole experimental task taking place in English.
+
+Bilingual toddlers and children---still in the process to developing a
+mature lexicon---also show evidence of parallel activation during
+comprehension (Bosma \& Nota, 2020; Poarch \& Hell, 2012; Schröter \&
+Schroeder, 2016; Von Holzen et al., 2019; Von Holzen \& Mani, 2012).
+Whether---and if so, how---the language non-selectivity shapes the early
+bilingual lexicon is still unclear. Previous work has pointed to an
+interactive account of lexical acquisition, in which toddlers' lexical
+representations in one language influence the acquisition of words in
+the other language. The acquisition of translation equivalents embodies
+a notable divergence between monolingual and bilingual developmental
+trajectories. For instance, monolingual word acquisition benefits from
+the semantic relatedness between the words the child learns (Fourtassi
+et al., 2020; Hills et al., 2009), but is inhibited during the
+acquisition of words from the same language with the \emph{equivalent}
+meaning, i.e.~synonyms. This learning heuristic, called \emph{mutual
+exclusivity}, is one of many that play a relevant role in models of
+monolingual word learning (Golinkoff et al., 1994), acting as a
+constrainer of the hypothesis space for all possible word-referent
+associations: if a referent already holds a label, any label uttered in
+the presence of such referent is associated with a different referent
+(e.g., Markman, 1990). Although this heuristic seems useful in the
+monolingual case (children are rarely exposed to synonyms), it would
+come with a huge cost for bilinguals, how are frequently exposed to more
+than one label for the same referent, one in each language. It was
+initially predicted that bilinguals would follow \emph{mutual
+exclusivity}, and therefore avoid learning translation equivalents
+(Volterra \& Taeschner, 1978).
+
+This prediction has been robustly contested by two main sources of
+evidence. First, experimental work suggests that bilinguals may not
+apply \emph{mutual exclusivity} during word learning, or at least not as
+consistently monolinguals do (Byers-Heinlein \& Werker, 2009, 2013).
+Second, detailed analyses of bilingual children's vocabulary contents
+shows that bilingual children do acquire translation equivalents from
+early ages (e.g., Pearson et al., 1995). In fact, translation
+equivalents may even be overrepresented in the early bilingual lexicon.
+For instance, Bilson et al. (2015) found that bilingual toddlers
+produced a higher proportion of translation equivalents than it would be
+expected if they learned two independent lexica simultaneously, which
+suggests that bilingual children preferentially acquired words for which
+they already knew a label in the other language. In the same line, Bosch
+and Ramon-Casas (2014) reported a similar overrepresentation of
+translation equivalents in Catalan-Spanish bilinguals: a high proportion
+of acquired Catalan words held an also acquired translation in Spanish.
+Finally, a recent investigation by Tsui et al. (2022b) presented a
+quantitative model of bilingual vocabulary acquisition which showed a
+consistent preferential acquisition of translation equivalents by
+English-French bilinguals, with such preference being stronger at early
+stages of vocabulary growth. Overall, these findings suggest that
+bilingual children benefit from a semantic facilitation effect driven by
+the acquisition of translation equivalents. This is a substantial
+divergence from monolingual trajectories of lexical acquisition. Our
+study provides further evidence for a pivotal role of translation
+equivalents, showing that, especially in the case of cognates,
+translation equivalents, are not only part of bilinguals' vocabulary
+contents, but are also acquired close in age, depending primarily on
+children's degree of exposure to each language.
+
+The divergent trajectories of bilingual lexical acquisition go beyond
+semantics. As highlighted before, the phonological overlap between the
+languages being learned seems to impact vocabulary growth in bilinguals.
+Bilinguals learning languages that share many cognates show larger
+vocabulary sizes than those learning two languages sharing fewer
+cognates (Blom et al., 2020; Bosma et al., 2019; Floccia, Sambrook,
+Delle Luche, et al., 2018). This finding can be explained to a large
+extent by the fact that cognates seem to be acquired at earlier ages
+than non-cognates (Mitchell et al., 2022). This phonological
+facilitation effect mimics previous findings in monolinguals, who seem
+to preferentially learn phonologically similar words. For instance,
+Jones and Brandt (2019) found that English monolingual toddlers show a
+facilitation effect when acquiring words from dense phonological
+neighbourhoods: a larger number of similar-sounding words was associated
+with an increase in likelihood of a word form being part of a toddler's
+productive vocabulary. A network analysis by Fourtassi et al. (2020)
+found evidence in favour of a phonological (and semantic) preferential
+acquisition of new words in monolingual children: words with higher
+phonological relatedness to other words were more likely to be acquired.
+Again, these findings contrast with those in bilingual children in that
+the facilitation effect or preferential acquisition mechanisms described
+are based on phonological similarity \emph{without} translation
+equivalence. The amount of similar-sounding words in the child input or
+their acquired vocabulary predicts a larger probability of acquisition
+of new words with \emph{different} meaning. In the case of bilinguals,
+there is evidence of such phonological facilitation at the level of
+translation equivalents, as shown by an earlier acquisition of cognates
+than non-cognates in English-French (Mitchell et al., 2022) and in the
+Catalan-Spanish bilinguals in the present study. Evidence of a
+cross-language facilitation effect between non-translation equivalents
+is to date absent.
+
+We consider three alternative explanations for our results. Regarding
+the cognate facilitation effect, it is possible that when hearing
+cognate translation equivalents, bilingual children accept both word
+forms as acceptable variations of the \emph{same} word. This contrasts
+with monolinguals' high sensitivity to mispronunciations during word
+recognition (Bailey \& Plunkett, 2002; Mani \& Plunkett, 2011; Swingley,
+2005; Swingley \& Aslin, 2000; Tamási et al., 2017). It might be the
+case that every exposure to either of the word-forms of the translation
+equivalent is counted as a learning instance for both. This would impact
+the acquisition mechanism in such way that any cognate effect is not
+explained by a reduction of the number of learning instances needed to
+acquired a word (after its translation has been acquired), but by a
+faster increase in cumulative learning instances across ages, leading to
+a similar outcome: an earlier age-of-acquisition for cognates.
+Supporting such alternative explanation, Ramon-Casas et al. (2009)
+examined 18-26 months-old Catalan-Spanish bilingual toddlers'
+sensitivity to mispronunciations in a word-recognition task and found
+that, contrary to their Catalan monolingual peers, bilinguals seemed to
+recognise mispronounced words, disregarding vowel replacements involving
+the Catalan vowel contrast /e/-/\textepsilon/. The same population of
+bilingual infants shows sensitivity to such contrast before 12 months of
+age in phoneme discrimination tasks (Bosch \& Sebastián-Gallés,
+2003)---which do not involve word-referent associations---suggesting
+that bilingual infants dismiss subtle changes in phonological forms when
+presented with their referents. Using the same procedure, Ramon-Casas
+and Bosch (2010) tested bilingual infants' recognition of
+mispronunciations involving the /e/-/\textepsilon/ contrast embedded in
+exclusively \emph{non-cognate} word-forms. The materials in this study
+contrasted with those in Ramon-Casas et al. (2009) in that the latter
+included almost 60\% of cognate words in their stimuli list. Ramon-Casas
+and Bosch (2010) hypothesised that bilinguals' frequent exposure to
+cognates---which often involve vowel replacements between Catalan and
+Spanish word-forms---may have led bilinguals to be less alert to such
+changes when occurring within similar word-forms. With this new set of
+stimuli, including non-cognate words only, bilinguals showed
+monolingual-level sensitivity to mispronunciations involving the
+/e/-/\textepsilon/ contrast. Overall, these results suggest, at face
+value, that that bilinguals may to some degree encode similar-sounding
+word-forms with shared referent as into a common lexical representation,
+especially when the word-forms involved are cognates.
+
+We identify three main reasons why bilingual children are unlikely to
+conflate similar sounding translations into the same lexical
+representations. First, when Ramon-Casas et al. (2009) probed
+Catalan-Spanish bilinguals using the /e/-/i/ contrast---present in both
+Catalan \emph{and} Spanish---in the same task, bilinguals showed
+monolingual-level sensitivity to mispronunciations, suggesting that the
+conflation of variations of the same word-form into a similar lexical
+representation may be specific to some language-specific contrasts, and
+perhaps associated with the acoustic distance between the phonemes
+involved. The /e/-/\textepsilon/ contrast in which Catalan-Spanish
+bilinguals failed to mispronunciations, is particularly problematic this
+regard: both vowels differ in only one acoustic feature (/e/ is
+closed-mid, while /\textepsilon/ is open-mid) and is only present in
+Catalan. Our results are unlikely to be explained by the possibility
+that Catalan-Spanish bilingual toddlers form common lexical
+representations for cognate words. Second, few phoneme substitutions
+that separate Catalan word-forms from their cognate Spanish translations
+involve the /e/-/\textepsilon/ contrast. Rather, due to vowel reduction,
+Spanish and Catalan cognates predominantly involve the /a/-/\textschwa/,
+/e/-/\textschwa/, /\textepsilon/ and /o/-/u/ contrasts (Vigário et al.,
+2009). Only 10\% translation equivalents in our questionnaire involved a
+/e/-/\textepsilon/ change between the Catalan and the Spanish items,
+only one of such was a minimal pair: /\textipa{tREn}/-/\textipa{tRen}/
+{[}\emph{train}{]}.
+
+Second, few words in our study differ only in one phoneme substitution.
+Only 1.32\% of the translation equivalents included in our analyses
+shared between 90 and 100\% phonological similarity (measured as the
+Levenshtein similarity between the X-SAMPA transcriptions of the Catalan
+and Spanish form), and only 1.99\% shared between 80 and 90\%
+similarity. This shows that the presence of identical and near-identical
+cognates in our study is very low, and therefore most translation
+equivalents differed in more than one phoneme. In summary, even if any
+cognate involved the /a/-/\textepsilon/ substitution (or any
+phonetically close contrast that bilingual might disregard at the
+lexical level), such is very likely to have occurred in the context of a
+sufficiently distinct word-form to be distinguished and treated and
+relevant at the lexical level. Third, and perhaps more critically,
+bilinguals are able to learn new minimal pairs involving the
+/e/-/\textepsilon/ contrast, suggesting that their tolerance towards
+cognate mispronunciations might be restricted to familiar word
+recognition, while such subtle variations in the word-forms are
+considered relevant during word learning (Ramon-Casas et al., 2017).
+
+It cannot be neglected a potential bias by caregivers when filling the
+questionnaire, in which the likelihood of two translation equivalents
+being present in a child's lexicon results not as a function of such
+phonological facilitation effect, but by caregivers assuming that
+because the child knows a word in one language, the child must also know
+the word in the other language. This bias might be especially likely
+when both words are similar-sounding (i.e., cognates). Although there is
+little possibility to test this account, it is worth noting that parents
+were explicitly instructed \emph{not} to rely on their responses to the
+words in one languages when responding to words in the other language.
+The potential presence of responder biases is pervasive across many
+measurement instruments that rely on caregivers' reports of the
+developmental status of the child. The outcomes of such instruments,
+like the one used in the present study, are susceptible to biases
+induced by the child themselves (i.e., assigned gender, temperament),
+their caregivers (e.g., time spent with the child, preconceptions about
+language acquisition), or even the researchers (e.g., the items included
+might over-rely on canonical realisations of words), which push against
+evidence of validity of the questionnaire as an instrument of the
+developmental status of the child. The findings in this investigation
+should be weighted according to the aforementioned limitations.
+
+Our results should also be taken with caution if generalised to other
+bilingual populations, given the particular typological similarity and
+sociolinguistic status of the languages being learned by our
+participants. Catalan and Spanish are both Romance languages, share a
+considerable amount of cognates, and very often co-exist in the same
+social spaces (families, schools, working place, etc.). Although there
+are no populations of bilinguals with identical sociolinguistic
+characteristics, and therefore it would be pointless to compare the
+sample in this study to any ``standard'' population of bilinguals, the
+generalisability of our results should be taken with a grain of salt. It
+is possible, for instance, that the large amount of cognates that
+Catalan-Spanish bilingual children are exposed leads them to learn at
+early stages of word acquisition that words tend to sound similar in
+both languages, and therefore to be more keen on assuming that similar
+sounding words are semantically related. Children exposed to a pair of
+languages sharing a fewer amount of cognates might not make the same
+assumption, even when the word-forms they hear in the presence of the
+same referents are very similar at the phonological level.
+
+Another limitation of the present study is the predominantly
+cross-sectional design of our data set. We collected vocabulary data
+from children aged 12 to 32 months, a critical period in development in
+which the foundations of a lexicon get established. We estimated word
+acquisition curves by modelling the probability of comprehension and
+production across ages. However, most participants provided data at only
+one age point. Effectively, we are making longitudinal inferences based
+on cross-sectional data, and drawing conclusions about the development
+of a single participant across multiple age points using data from
+different participants at individual age points. For this reason, and
+also because of the observational nature of our study, causal inferences
+derived from our analyses are unwarranted (Kesmodel, 2018; Kraemer et
+al., 2000). One counterpoint to this shortcoming is that, while data our
+data is cross-sectional if participants are considered as the
+observational unit, data about word-forms/translation equivalents is
+longitudinal. The same translation equivalent is responded to at
+multiple ages. Since translation equivalents are our main observational
+unit of interest over which we primarily conduct statistical inference
+(e.g., estimating the size of the cognateness-effect), this provides
+some further justification for our conclusions.
+
+Future studies addressing this subject may benefit for a more adequate
+design oriented towards making causal inferences. One such design can be
+partially satisfied by (1) collecting a more substantial longitudinal
+dataset, in which the same participant provides data points for each
+translation equivalent at multiple ages, and (2) testing the role of
+cognateness and word exposure in an experimental context, in which such
+variables can be orthogonally manipulated or randomised. This second
+possibility would also allow to explore an interpretation of the results
+obtained in the context of \emph{word learning}, which is not possible
+in the present study. Another consideration for future investigations is
+the involvement of a more diverse range of language pairs as in
+Braginsky et al. (2019) or Floccia, Sambrook, Delle Luche, et al. (2018)
+to improve the generalisability of the resulting inferences across more
+populations of bilinguals.
+
+In summary, our study has replicated previous studies and provided novel
+insights about word acquisition in bilingual contexts, and how the
+presence of cognates in the children's linguistic input impacts the
+early formation of the lexicon. We found that during the acquisition of
+low frequency words, bilingual children seem to benefit more strongly
+from the word's phonological similarity with its translation in the
+other language. We suggest that this effect might be explained by the
+need of a robust lexical representation of one of the translations for
+cognateness to play a role.
+
+\hypertarget{sec-references}{%
+\section{References}\label{sec-references}}
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-arel-bundock2022marginaleffects}{}}%
+Arel-Bundock, V. (2022). \emph{Marginaleffects: Marginal effects,
+marginal means, predictions, and contrasts} {[}Manual{]}.
+\url{https://CRAN.R-project.org/package=marginaleffects}
+
 \leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%
@@
 Research Methods}, \emph{52}(1), 376--387.
+\url{https://doi.org/10.3758/s13428-019-01236-y}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bailey2002phonological}{}}%
+Bailey, T. M., \& Plunkett, K. (2002). Phonological specificity in early
+words. \emph{Cognitive Development}, \emph{17}(2), 1265--1282.
+\url{https://doi.org/10.1016/S0885-2014(02)00116-8}
 
@@
 Perspectives}, \emph{14}(3), 142--149.
+\url{https://doi.org/10.1111/cdep.12373}
 
@@
 National Academy of Sciences}, \emph{109}(9), 3253--3258.
+\url{https://doi.org/10.1073/pnas.1113380109}
 
@@
 Development}, \emph{11}(4), 369--380.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bloom2002children}{}}%
-Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT
-press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-boada2020subtlex}{}}%
-Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P. (2020).
-SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for
-catalan. \emph{Behavior Research Methods}, \emph{52}(1), 360--375.
+\url{https://doi.org/10.1080/15475441.2014.979387}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bilson2015semantic}{}}%
+Bilson, S., Yoshida, H., Tran, C. D., Woods, E. A., \& Hills, T. T.
+(2015). Semantic facilitation in bilingual first language acquisition.
+\emph{Cognition}, \emph{140}, 122--134.
+
+\leavevmode\vadjust pre{\hypertarget{ref-blom2020crosslanguage}{}}%
+Blom, E., Boerma, T., Bosma, E., Cornips, L., Heuij, K. van den, \&
+Timmermeister, M. (2020). Cross-language distance influences receptive
+vocabulary outcomes of bilingual children. \emph{First Language},
+\emph{40}(2), 151--171. \url{https://doi.org/10.1177/0142723719892794}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bloom2002how}{}}%
+Bloom, P. (2002). \emph{How children learn the meanings of words}. {MIT}
+Press.
 
@@
 Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in
-bilingual toddlers' expressive vocabulary: {Does} form similarity
-matter? \emph{International Journal of Behavioral Development},
-\emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}
+bilingual toddlers' expressive vocabulary: Does form similarity matter?
+\emph{International Journal of Behavioral Development}, \emph{38}(4),
+317--322. \url{https://doi.org/10.1177/0165025414532559}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosch2003simultaneous}{}}%
+Bosch, L., \& Sebastián-Gallés, N. (2003). Simultaneous bilingualism and
+the perception of a language-specific vowel contrast in the first year
+of life. \emph{Language and Speech}, \emph{46}(2), 217--243.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosma2019longitudinal}{}}%
+Bosma, E., Blom, E., Hoekstra, E., \& Versloot, A. (2019). A
+longitudinal study on the gradual cognate facilitation effect in
+bilingual children's frisian receptive vocabulary. \emph{International
+Journal of Bilingual Education and Bilingualism}, \emph{22}(4),
+371--385. \url{https://doi.org/10.1080/13670050.2016.1254152}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosma2020cognate}{}}%
+Bosma, E., \& Nota, N. (2020). Cognate facilitation in frisian--dutch
+bilingual children's sentence reading: An eye-tracking study.
+\emph{Journal of Experimental Child Psychology}, \emph{189}, 104699.
 
@@
 languages. \emph{Open Mind}, \emph{3}, 52--67.
+\url{https://doi.org/10.1162/opmi_a_00026}
 
@@
 Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
-using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+using stan. \emph{Journal of Statistical Software}, \emph{80}, 1--28.
+\url{https://doi.org/10.18637/jss.v080.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2019ordinal}{}}%
+Bürkner, P.-C., \& Vuorre, M. (2019). Ordinal regression models in
+psychology: A tutorial. \emph{Advances in Methods and Practices in
+Psychological Science}, \emph{2}(1), 77--101.
+
+\leavevmode\vadjust pre{\hypertarget{ref-byers-heinlein2009monolingual}{}}%
+Byers-Heinlein, K., \& Werker, J. F. (2009). Monolingual, bilingual,
+trilingual: Infants' language experience influences the development of a
+word-learning heuristic. \emph{Developmental Science}, \emph{12}(5),
+815--823. \url{https://doi.org/10.1111/j.1467-7687.2009.00902.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-byers-heinlein2013lexicon}{}}%
+Byers-Heinlein, K., \& Werker, J. F. (2013). Lexicon structure and the
+disambiguation of novel words: Evidence from bilingual infants.
+\emph{Cognition}, \emph{128}(3), 407--416.
+\url{https://doi.org/10.1016/j.cognition.2013.05.010}
+
+\leavevmode\vadjust pre{\hypertarget{ref-campbell2007glossary}{}}%
+Campbell, L., \& Mixco, M. J. (2007). \emph{A glossary of historical
+linguistics}. Edinburgh University Press.
 
@@
 Betancourt, M., Brubaker, M., Guo, J., Li, P., \& Riddell, A. (2017).
-Stan: A probabilistic programming language. \emph{Journal of Statistical
-Software}, \emph{76}(1), 1--32.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cattani2014much}{}}%
+Stan : A probabilistic programming language. \emph{Journal of
+Statistical Software}, \emph{76}(1).
+\url{https://doi.org/10.18637/jss.v076.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cattani2014how}{}}%
 Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis,
@@
 Disorders}, \emph{49}(6), 649--671.
+\url{https://doi.org/10.1111/1460-6984.12082}
+
+\leavevmode\vadjust pre{\hypertarget{ref-christoffels2007bilingual}{}}%
+Christoffels, I. K., Firk, C., \& Schiller, N. O. (2007). Bilingual
+language control: An event-related brain potential study. \emph{Brain
+Research}, \emph{1147}, 192--208.
+\url{https://doi.org/10.1016/j.brainres.2007.01.137}
+
+\leavevmode\vadjust pre{\hypertarget{ref-colome2010which}{}}%
+Colomé, A., \& Miozzo, M. (2010). Which words are activated during
+bilingual word production? \emph{Journal of Experimental Psychology.
+Learning, Memory, and Cognition}, \emph{36}(1), 96--109.
+\url{https://doi.org/10.1037/a0017677}
 
@@
 \emph{Journal of Experimental Psychology: Learning, Memory, and
-Cognition}, \emph{26}(5), 1283.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cuetos2012subtlex}{}}%
-Cuetos, F., Glez-Nosti, M., Barbón, A., \& Brysbaert, M. (2012).
-SUBTLEX-ESP: Spanish word frequencies based on film subtitles.
-\emph{Psicol{ó}gica}, \emph{33}(2), 133--143.
+Cognition}, \emph{26}, 1283--1296.
+\url{https://doi.org/10.1037/0278-7393.26.5.1283}
+
+\leavevmode\vadjust pre{\hypertarget{ref-curtin2011bilingual}{}}%
+Curtin, S., Byers-Heinlein, K., \& Werker, J. F. (2011). Bilingual
+beginnings as a lens for theory development: {PRIMIR} in focus.
+\emph{Journal of Phonetics}, \emph{39}(4), 492--504.
+\url{https://doi.org/10.1016/j.wocn.2010.12.002}
 
@@
 Journal of Bilingual Education and Bilingualism}, \emph{11}(5),
-598--618.
-
-\leavevmode\vadjust pre{\hypertarget{ref-fenson2007macarthur}{}}%
-Fenson, L. et al. (2007). \emph{MacArthur-bates communicative
+598--618. \url{https://doi.org/10.1080/13670050802149200}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fenson2007macarthurbates}{}}%
+Fenson, L. et al. (2007). \emph{{MacArthur}-bates communicative
 development inventories}. Paul H. Brookes Publishing Company Baltimore,
-MD.
+{MD}.
 
@@
 Variability in early communicative development. \emph{Monographs of the
-Society for Research in Child Development}, i--185.
+Society for Research in Child Development}, \emph{59}(5), i--185.
+\url{https://doi.org/10.2307/1166093}
 
@@
 Floccia, C., Sambrook, T. D., Delle Luche, C., Kwok, R., Goslin, J.,
-White, L., Cattani, A., Sullivan, E., Abbot-Smith, K., Krott, A., et al.
-(2018). I: INTRODUCTION. \emph{Monographs of the Society for Research in
-Child Development}, \emph{83}(1), 7--29.
+White, L., Cattani, A., Sullivan, E., Abbot‐Smith, K., Krott, A., Mills,
+D., Rowland, C., Gervain, J., \& Plunkett, K. (2018). I: introduction.
+\emph{Monographs of the Society for Research in Child Development},
+\emph{83}(1), 7--29. \url{https://doi.org/10.1111/mono.12348}
+
+\leavevmode\vadjust pre{\hypertarget{ref-floccia2018methods}{}}%
+Floccia, C., Sambrook, T. D., Luche, C. D., Kwok, R., Goslin, J., White,
+L., Cattani, A., Sullivan, E., Abbot‐Smith, K., Krott, A., Mills, D.,
+Rowland, C., Gervain, J., \& Plunkett, K. (2018a). {II}: methods.
+\emph{Monographs of the Society for Research in Child Development},
+\emph{83}(1), 30--42.
+https://doi.org/\url{https://doi.org/10.1111/mono.12349}
+
+\leavevmode\vadjust pre{\hypertarget{ref-floccia2018results}{}}%
+Floccia, C., Sambrook, T. D., Luche, C. D., Kwok, R., Goslin, J., White,
+L., Cattani, A., Sullivan, E., Abbot‐Smith, K., Krott, A., Mills, D.,
+Rowland, C., Gervain, J., \& Plunkett, K. (2018b). {III}: Analyses and
+results for study 1: Estimating the effect of linguistic distance on
+vocabulary development. \emph{Monographs of the Society for Research in
+Child Development}, \emph{83}(1), 43--60.
+https://doi.org/\url{https://doi.org/10.1111/mono.12350}
 
@@
 languages. \emph{Cognitive Science}, \emph{44}(7), e12847.
+\url{https://doi.org/10.1111/cogs.12847}
 
@@
 \emph{Variability and consistency in early language learning: The
-wordbank project}. MIT Press.
+wordbank project}. {MIT} Press.
 
@@
 using multiple sequences. \emph{Statistical Science}, \emph{7}(4),
-457--472.
+457--472. \url{https://www.jstor.org/stable/2246093}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gimeno-martinez2021crosslinguistic}{}}%
+Gimeno-Martínez, M., Mädebach, A., \& Baus, C. (2021). Cross-linguistic
+interactions across modalities: Effects of the oral language on sign
+production. \emph{Bilingualism: Language and Cognition}, \emph{24}(4),
+779--790. \url{https://doi.org/10.1017/S1366728921000171}
 
@@
 Goldfield, B. A., \& Reznick, J. S. (1990). Early lexical acquisition:
-Rate, content, and the vocabulary spurt. \emph{Journal of Child
+Rate, content, and the vocabulary spurt*. \emph{Journal of Child
 Language}, \emph{17}(1), 171--183.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gonzalez2020bilingual}{}}%
+\url{https://doi.org/10.1017/S0305000900013167}
+
+\leavevmode\vadjust pre{\hypertarget{ref-golinkoff1994early}{}}%
+Golinkoff, R. M., Mervis, C. B., \& Hirsh-Pasek, K. (1994). Early object
+labels: The case for a developmental lexical principles
+framework{[}*{]}. \emph{Journal of Child Language}, \emph{21}(1),
+125--155. \url{https://doi.org/10.1017/S0305000900008692}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gonzalez-barrero2020bilingual}{}}%
 Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020).
 \emph{Bilingual adjusted vocabulary: A developmentally-informed
-bilingual vocabulary measure}.
+bilingual vocabulary measure}. {PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/x7s4u}
+
+\leavevmode\vadjust pre{\hypertarget{ref-goodman2008does}{}}%
+Goodman, J. C., Dale, P. S., \& Li, P. (2008). Does frequency count?
+Parental input and the acquisition of vocabulary. \emph{Journal of Child
+Language}, \emph{35}(3), 515--531.
+
+\leavevmode\vadjust pre{\hypertarget{ref-grosjean2021extent}{}}%
+Grosjean, F. (2021). The extent of bilingualism. \emph{Life as a
+Bilingual}, 27--39.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-heeringa2004measuring}{}}%
+Heeringa, W. (2004). \emph{Measuring dialect pronunciation differences
+using levenshtein distance} {[}PhD thesis{]}. University Library
+Groningen{]}{[}Host{]}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-heeringa2003norwegian}{}}%
+Heeringa, W., \& Gooskens, C. (2003). Norwegian dialects examined
+perceptually and acoustically. \emph{Computers and the Humanities},
+\emph{37}(3), 293--315. \url{https://doi.org/10.1023/A:1025087115665}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanhell2008sentence}{}}%
+Hell, J. G. van, \& Groot, A. M. B. de. (2008). Sentence context
+modulates visual word recognition and translation in bilinguals.
+\emph{Acta Psychologica}, \emph{128}(3), 431--451.
+\url{https://doi.org/10.1016/j.actpsy.2008.03.010}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hidaka2013computational}{}}%
+Hidaka, S. (2013). A computational model associating learning process,
+word attributes, and age of acquisition. \emph{{PLOS} {ONE}},
+\emph{8}(11), e76242. \url{https://doi.org/10.1371/journal.pone.0076242}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hills2009longitudinal}{}}%
+Hills, T. T., Maouene, M., Maouene, J., Sheya, A., \& Smith, L. (2009).
+Longitudinal analysis of early semantic networks preferential attachment
+or preferential acquisition? \emph{Psychological Science}, \emph{20}(6),
+729--739. \url{https://doi.org/10.1111/j.1467-9280.2009.02365.x}
+
 \leavevmode\vadjust pre{\hypertarget{ref-hoff2012dual}{}}%
 Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M.
-(2012). Dual language exposure and early bilingual development.
-\emph{Journal of Child Language}, \emph{39}(1), 1.
+(2012). Dual language exposure and early bilingual development*.
+\emph{Journal of Child Language}, \emph{39}(1), 1--27.
+\url{https://doi.org/10.1017/S0305000910000759}
 
@@
 \emph{Cognition}, \emph{106}(1), 501--511.
+\url{https://doi.org/10.1016/j.cognition.2007.02.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hoshino2011language}{}}%
+Hoshino, N., \& Thierry, G. (2011). Language selection in bilingual word
+production: Electrophysiological evidence for cross-language
+competition. \emph{Brain Research}, \emph{1371}, 100--109.
+\url{https://doi.org/10.1016/j.brainres.2010.11.053}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jones2019children}{}}%
+Jones, S. D., \& Brandt, S. (2019). Do children really acquire dense
+neighbourhoods? \emph{Journal of Child Language}, \emph{46}(6),
+1260--1273. \url{https://doi.org/10.1017/S0305000919000473}
 
 \leavevmode\vadjust pre{\hypertarget{ref-jusczyk1995infants}{}}%
-Jusczyk, P. W., \& Aslin, R. N. (1995). Infants' detection of the sound
+Jusczyk, P. W., \& Aslin, R. N. (1995). Infants′ detection of the sound
 patterns of words in fluent speech. \emph{Cognitive Psychology},
-\emph{29}(1), 1--23.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kachergis2022toward}{}}%
+\emph{29}(1), 1--23. \url{https://doi.org/10.1006/cogp.1995.1010}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kachergis2022standard}{}}%
 Kachergis, G., Marchman, V. A., \& Frank, M. C. (2022). Toward a
@@
 Directions in Psychological Science}, \emph{31}(1), 20--27.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tidybayes}{}}%
-Kay, M. (2021). \emph{{tidybayes}: Tidy data and geoms for {Bayesian}
+\url{https://doi.org/10.1177/09637214211057836}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kay2021tidybayes}{}}%
+Kay, M. (2021). \emph{Tidybayes: Tidy data and geoms for bayesian
 models}. \url{https://doi.org/10.5281/zenodo.1308151}
 
+\leavevmode\vadjust pre{\hypertarget{ref-kesmodel2018crosssectional}{}}%
+Kesmodel, U. S. (2018). Cross-sectional studies -- what are they good
+for? \emph{Acta Obstetricia Et Gynecologica Scandinavica}, \emph{97}(4),
+388--393. \url{https://doi.org/10.1111/aogs.13331}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kraemer2000how}{}}%
+Kraemer, H. C., Yesavage, J. A., Taylor, J. L., \& Kupfer, D. (2000).
+How can we learn about developmental processes from cross-sectional
+studies, or can we? \emph{American Journal of Psychiatry},
+\emph{157}(2), 163--171.
+\url{https://doi.org/10.1176/appi.ajp.157.2.163}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroll2013multilingual}{}}%
+Kroll, J. F., Gullifer, J. W., \& Rossi, E. (2013). The multilingual
+lexicon: The cognitive and neural basis of lexical comprehension and
+production in two or more languages. \emph{Annual Review of Applied
+Linguistics}, \emph{33}, 102--127.
+\url{https://doi.org/10.1017/S0267190513000111}
+
 \leavevmode\vadjust pre{\hypertarget{ref-kruschke2018bayesian}{}}%
-Kruschke, J., \& Liddell, T. (2018). The bayesian new statistics:
+Kruschke, J. K., \& Liddell, T. M. (2018). The bayesian new statistics:
 Hypothesis testing, estimation, meta-analysis, and planning from a
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kuhl1992linguistic}{}}%
+Kuhl, P. K., Williams, K. A., Lacerda, F., Stevens, K. N., \& Lindblom,
+B. (1992). Linguistic experience alters phonetic perception in infants
+by 6 months of age. \emph{Science}, \emph{255}(5044), 606--608.
+\url{https://doi.org/10.1126/science.1736364}
+
 \leavevmode\vadjust pre{\hypertarget{ref-laing2022phonological}{}}%
 Laing, C. E. (2022). \emph{Phonological networks and systematicity in
-early lexical acquisition}.
+early lexical acquisition}. {PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/z8pyg}
 
 \leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
-Levenshtein, V. I. et al. (1966). Binary codes capable of correcting
-deletions, insertions, and reversals. \emph{Soviet Physics Doklady},
-\emph{10}, 707--710.
-
-\leavevmode\vadjust pre{\hypertarget{ref-magnusson2019bayesian}{}}%
-Magnusson, M., Andersen, M., Jonasson, J., \& Vehtari, A. (2019).
-Bayesian leave-one-out cross-validation for large data.
-\emph{International Conference on Machine Learning}, 4244--4253.
+Levenshtein, V. I. (1966). Binary codes capable of correcting deletions,
+insertions, and reversals. \emph{Soviet Physics Doklady}, \emph{10},
+707--710.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lewandowski2009generating}{}}%
+Lewandowski, D., Kurowicka, D., \& Joe, H. (2009). Generating random
+correlation matrices based on vines and extended onion method.
+\emph{Journal of Multivariate Analysis}, \emph{100}(9), 1989--2001.
+
+\leavevmode\vadjust pre{\hypertarget{ref-van-der-loo2014stringdist}{}}%
+Loo, M. P. J. van der. (2014). The stringdist package for approximate
+string matching. \emph{The R Journal}, \emph{6}(1), 111--122.
+\url{https://doi.org/10.32614/RJ-2014-011}
+
+\leavevmode\vadjust pre{\hypertarget{ref-macwhinney2000childes}{}}%
+MacWhinney, B. (2000). \emph{The {CHILDES} project: The database} (Vol.
+2). Psychology Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-magnusson2020leaveoneout}{}}%
+Magnusson, M., Vehtari, A., Jonasson, J., \& Andersen, M. (2020).
+Leave-one-out cross-validation for bayesian model comparison in large
+data. \emph{Proceedings of the Twenty Third International Conference on
+Artificial Intelligence and Statistics}, 341--351.
+\url{https://proceedings.mlr.press/v108/magnusson20a.html}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mani2011does}{}}%
+Mani, N., \& Plunkett, K. (2011). Does size matter? Subsegmental cues to
+vowel mispronunciation detection*. \emph{Journal of Child Language},
+\emph{38}(3), 606--627. \url{https://doi.org/10.1017/S0305000910000243}
+
+\leavevmode\vadjust pre{\hypertarget{ref-markman1990constraints}{}}%
+Markman, E. M. (1990). Constraints children place on word meanings.
+\emph{Cognitive Science}, \emph{14}(1), 57--77.
+\url{https://doi.org/10.1207/s15516709cog1401_4}
 
@@
 Mayor, J., \& Plunkett, K. (2011). A statistical estimate of infant and
-toddler vocabulary size from CDI analysis. \emph{Developmental Science},
-\emph{14}(4), 769--785.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcelreath2018statistical}{}}%
-McElreath, R. (2018). \emph{Statistical rethinking: A bayesian course
-with examples in r and stan}. Chapman; Hall/CRC.
+toddler vocabulary size from {CDI} analysis. \emph{Developmental
+Science}, \emph{14}(4), 769--785.
+\url{https://doi.org/10.1111/j.1467-7687.2010.01024.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcelreath2020statistical}{}}%
+McElreath, R. (2020). \emph{Statistical rethinking: A bayesian course
+with examples in r and stan}. Chapman; Hall/{CRC}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-mitchell2022cognates}{}}%
-Mitchell, L., Tsui, R. K., \& Byers-Heinlein, K. (2022). \emph{Cognates
-are advantaged in early bilingual expressive vocabulary development}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-nelson1973structure}{}}%
-Nelson, K. (1973). Structure and strategy in learning to talk.
-\emph{Monographs of the Society for Research in Child Development},
-1--135.
+Mitchell, L., Tsui, R. K. Y., \& Byers-Heinlein, K. (2022).
+\emph{Cognates are advantaged in early bilingual expressive vocabulary
+development}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/daktp}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mollica2017how}{}}%
+Mollica, F., \& Piantadosi, S. T. (2017). How data drive early word
+learning: A cross-linguistic waiting time analysis. \emph{Open Mind},
+\emph{1}(2), 67--77. \url{https://doi.org/10.1162/OPMI_a_00006}
+
+\leavevmode\vadjust pre{\hypertarget{ref-morford2011when}{}}%
+Morford, J. P., Wilkinson, E., Villwock, A., Piñar, P., \& Kroll, J. F.
+(2011). When deaf signers read english: Do written words activate their
+sign translations? \emph{Cognition}, \emph{118}(2), 286--292.
+\url{https://doi.org/10.1016/j.cognition.2010.11.006}
 
@@
 Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in
-bilingual children} (Vol. 2). Multilingual Matters.
+bilingual children}. Multilingual Matters.
 
 \leavevmode\vadjust pre{\hypertarget{ref-patterson2004comparing}{}}%
-Patterson, J. L. (2004). \emph{Comparing bilingual and monolingual
-toddlers' expressive vocabulary size}.
+Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers'
+expressive vocabulary size. \emph{Journal of Speech, Language, and
+Hearing Research}, \emph{47}(5), 1213--1215.
+\url{https://doi.org/10.1044/1092-4388(2004/089)}
 
 \leavevmode\vadjust pre{\hypertarget{ref-patterson2004bilingual}{}}%
-Patterson, J., Pearson, B., \& Goldstein, B. (2004). \emph{Bilingual
-language development and disorders in spanish--english speakers}.
+Patterson, J. L., \& Pearson, B. Z. (2004). Bilingual lexical
+development: Influences, contexts, and processes. In \emph{Bilingual
+language development and disorders in spanish-english speakers} (pp.
+77--104). Paul H. Brookes Publishing Co.
 
@@
 \emph{Language Learning}, \emph{44}(4), 617--653.
+\url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}
 
@@
 norms. \emph{Language Learning}, \emph{43}(1), 93--120.
+\url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-pearson1995crosslanguage}{}}%
+Pearson, B. Z., Fernández, S., \& Oller, D. K. (1995). Cross-language
+synonyms in the lexicons of bilingual infants: One language or two?*.
+\emph{Journal of Child Language}, \emph{22}(2), 345--368.
+\url{https://doi.org/10.1017/S030500090000982X}
 
@@
 language acquisition. \emph{Journal of Child Language}, \emph{28}(2),
-453--496.
-
-\leavevmode\vadjust pre{\hypertarget{ref-poulin2013lexical}{}}%
+453--496. \url{https://doi.org/10.1017/S0305000901004718}
+
+\leavevmode\vadjust pre{\hypertarget{ref-poarch2012crosslanguage}{}}%
+Poarch, G. J., \& Hell, J. G. van. (2012). Cross-language activation in
+children's speech production: Evidence from second language learners,
+bilinguals, and trilinguals. \emph{Journal of Experimental Child
+Psychology}, \emph{111}(3), 419--438.
+\url{https://doi.org/10.1016/j.jecp.2011.09.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-poulin-dubois2013lexical}{}}%
 Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
-R Core Team. (2020). \emph{R: A language and environment for statistical
+\leavevmode\vadjust pre{\hypertarget{ref-rcoreteam2013language}{}}%
+R Core Team. (2013). \emph{R: A language and environment for statistical
 computing}. R Foundation for Statistical Computing.
-\url{https://www.R-project.org/}
+\url{http://www.R-project.org/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon2010non}{}}%
+Ramon-Casas, M., \& Bosch, L. (2010). Are non-cognate words
+phonologically better specified than cognates in the early lexicon of
+bilingual children. \emph{Proceedings of the 4th Conference on
+Laboratory Approaches to Spanish Phonology}, 31--36.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2017minimalpair}{}}%
+Ramon-Casas, M., Fennell, C. T., \& Bosch, L. (2017). Minimal-pair word
+learning by bilingual toddlers: The catalan /e/-/ɛ/ contrast revisited*.
+\emph{Bilingualism: Language and Cognition}, \emph{20}(3), 649--656.
+\url{https://doi.org/10.1017/S1366728916001115}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2009vowel}{}}%
+Ramon-Casas, M., Swingley, D., Sebastián-Gallés, N., \& Bosch, L.
+(2009). Vowel categorization during word recognition in bilingual
+toddlers. \emph{Cognitive Psychology}, \emph{59}(1), 96--121.
+\url{https://doi.org/10.1016/j.cogpsych.2009.02.002}
+
+\leavevmode\vadjust pre{\hypertarget{ref-samuelson2021precision}{}}%
+Samuelson, L. K. (2021). Toward a precision science of word learning:
+Understanding individual vocabulary pathways. \emph{Child Development
+Perspectives}, \emph{15}(2), 117--124.
+\url{https://doi.org/10.1111/cdep.12408}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sanchez2019childes}{}}%
+Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
+D., \& Frank, M. C. (2019). Childes-db: A flexible and reproducible
+interface to the child language data exchange system. \emph{Behavior
+Research Methods}, \emph{51}(4), 1928--1941.
 
@@
 Cognition}, \emph{5}(2), 93--107.
-\url{https://doi.org/10.1017/S1366728902000214}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schepens2012distributions}{}}%
+Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of
+cognates in europe as based on levenshtein distance. \emph{Bilingualism:
+Language and Cognition}, \emph{15}(1), 157--166.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schroter2016orthographic}{}}%
+Schröter, P., \& Schroeder, S. (2016). Orthographic processing in
+balanced bilingual children: Cross-language evidence from cognates and
+false friends. \emph{Journal of Experimental Child Psychology},
+\emph{141}, 239--246. \url{https://doi.org/10.1016/j.jecp.2015.09.005}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schwartz2006bilingual}{}}%
+Schwartz, A. I., \& Kroll, J. F. (2006). Bilingual lexical activation in
+sentence context. \emph{Journal of Memory and Language}, \emph{55}(2),
+197--212. \url{https://doi.org/10.1016/j.jml.2006.03.004}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schwartz2007reading}{}}%
+Schwartz, A. I., Kroll, J. F., \& Diaz, M. (2007). Reading words in
+spanish and english: Mapping orthography to phonology in two languages.
+\emph{Language and Cognitive Processes}, \emph{22}(1), 106--129.
+\url{https://doi.org/10.1080/01690960500463920}
 
@@
 difference? \emph{Bilingualism: Language and Cognition}, \emph{17}(4),
-810--821.
+810--821. \url{https://doi.org/10.1017/S1366728913000813}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-swingley200511montholds}{}}%
+Swingley, D. (2005). 11-month-olds' knowledge of how familiar words
+sound. \emph{Developmental Science}, \emph{8}(5), 432--443.
+\url{https://doi.org/10.1111/j.1467-7687.2005.00432.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-swingley2000spoken}{}}%
+Swingley, D., \& Aslin, R. N. (2000). Spoken word recognition and
+lexical representation in very young children. \emph{Cognition},
+\emph{76}(2), 147--166.
+\url{https://doi.org/10.1016/S0010-0277(00)00081-0}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tamasi2017pupillometry}{}}%
+Tamási, K., McKean, C., Gafos, A., Fritzsche, T., \& Höhle, B. (2017).
+Pupillometry registers toddlers' sensitivity to degrees of
+mispronunciation. \emph{Journal of Experimental Child Psychology},
+\emph{153}, 140--148. \url{https://doi.org/10.1016/j.jecp.2016.07.014}
+
 \leavevmode\vadjust pre{\hypertarget{ref-tardif2008baby}{}}%
@@
 Marchman, V. A. (2008). Baby's first 10 words. \emph{Developmental
-Psychology}, \emph{44}(4), 929.
+Psychology}, \emph{44}, 929--938.
+\url{https://doi.org/10.1037/0012-1649.44.4.929}
 
@@
 the National Academy of Sciences}, \emph{104}(30), 12530--12535.
+\url{https://doi.org/10.1073/pnas.0609927104}
 
@@
 Communication Disorders}, \emph{4}(1), 1--21.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999some}{}}%
+\url{https://doi.org/10.1080/14769670500215647}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999beginnings}{}}%
 Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word
 comprehension in 6-month-olds. \emph{Psychological Science},
-\emph{10}(2), 172--175.
-
-\leavevmode\vadjust pre{\hypertarget{ref-van2014subtlex}{}}%
+\emph{10}(2), 172--175. \url{https://doi.org/10.1111/1467-9280.00127}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tsui2022are}{}}%
+Tsui, R. K.-Y., Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein,
+K. (2022a). Are translation equivalents special? Evidence from
+simulations and empirical data from bilingual infants. \emph{Cognition},
+\emph{225}, 105084.
+\url{https://doi.org/10.1016/j.cognition.2022.105084}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tsui2022area}{}}%
+Tsui, R. K.-Y., Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein,
+K. (2022b). Are translation equivalents special? Evidence from
+simulations and empirical data from bilingual infants. \emph{Cognition},
+\emph{225}, 105084.
+\url{https://doi.org/10.1016/j.cognition.2022.105084}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanheuven2014subtlexuk}{}}%
 Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
-SUBTLEX-UK: A new and improved word frequency database for british
+{SUBTLEX}-{UK}: A new and improved word frequency database for british
 english. \emph{Quarterly Journal of Experimental Psychology},
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-vehtari2022loo}{}}%
+Vehtari, A., Gabry, J., Magnusson, M., Yao, Y., Bürkner, P.-C.,
+Paananen, T., \& Gelman, A. (2022). \emph{Loo: Efficient leave-one-out
+cross-validation and {WAIC} for bayesian models}.
+\url{https://mc-stan.org/loo/}
+
 \leavevmode\vadjust pre{\hypertarget{ref-vehtari2017practical}{}}%
 Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model
-evaluation using leave-one-out cross-validation and WAIC.
+evaluation using leave-one-out cross-validation and {WAIC}.
 \emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
-
-\leavevmode\vadjust pre{\hypertarget{ref-von2012language}{}}%
+\url{https://doi.org/10.1007/s11222-016-9696-4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vigario2009phonetics}{}}%
+Vigário, M., Frota, S., \& Freitas, M. J. (2009). \emph{Phonetics and
+phonology: Interactions and interrelations}. John Benjamins Publishing.
+
+\leavevmode\vadjust pre{\hypertarget{ref-volterra1978acquisition}{}}%
+Volterra, V., \& Taeschner, T. (1978). The acquisition and development
+of language by bilingual children*. \emph{Journal of Child Language},
+\emph{5}(2), 311--326. \url{https://doi.org/10.1017/S0305000900007492}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonholzen2019impact}{}}%
+Von Holzen, K., Fennell, C. T., \& Mani, N. (2019). The impact of
+cross-language phonological overlap on bilingual and monolingual
+toddlers' word recognition. \emph{Bilingualism: Language and Cognition},
+\emph{22}(3), 476--499. \url{https://doi.org/10.1017/S1366728918000597}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonholzen2012language}{}}%
 Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access
@@
 \emph{113}(4), 569--586.
-
-\leavevmode\vadjust pre{\hypertarget{ref-R-tidyverse}{}}%
+\url{https://doi.org/10.1016/j.jecp.2012.08.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wells1995computer}{}}%
+Wells, J. C. (1995). Computer-coding the {IPA}: A proposed extension of
+{SAMPA}. \emph{Revised Draft}, \emph{4}(28), 1995.
+
+\leavevmode\vadjust pre{\hypertarget{ref-werker2005primir}{}}%
+Werker, J. F., \& Curtin, S. (2005). {PRIMIR}: A developmental framework
+of infant speech processing. \emph{Language Learning and Development},
+\emph{1}(2), 197--234.
+\url{https://doi.org/10.1080/15475441.2005.9684216}
+
+\leavevmode\vadjust pre{\hypertarget{ref-werker1984crosslanguage}{}}%
+Werker, J. F., \& Tees, R. C. (1984). Cross-language speech perception:
+Evidence for perceptual reorganization during the first year of life.
+\emph{Infant Behavior and Development}, \emph{7}(1), 49--63.
+\url{https://doi.org/10.1016/S0163-6383(84)80022-3}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham2019welcome}{}}%
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
@@
 Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
-Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
-\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
-
-\leavevmode\vadjust pre{\hypertarget{ref-loo}{}}%
-Yao, Y., Vehtari, A., Simpson, D., \& Gelman, A. (2017). Using stacking
-to average bayesian predictive distributions. \emph{Bayesian Analysis}.
-\url{https://doi.org/10.1214/17-BA1091}
+Welcome to the {\textless{}}span
+class="nocase"{\textgreater{}}tidyverse{\textless{}}/span{\textgreater{}}.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
 
@@
 Zipf, G. K. (1949). \emph{Human behavior and the principle of least
-effort: An introd. To human ecology}.
+effort}. Addison-Wesley Press.
 
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\hypertarget{appendix-words}{%
-\subsection{Appendix 1: Words}\label{appendix-words}}
-
-\hypertarget{appendix-words}{%
-\subsection{\texorpdfstring{Appendix 2: \emph{ROPE}
-specification}{Appendix 2: ROPE specification}}\label{appendix-words}}
-
-We decided to draw the \emph{ROPE} interval around zero so that it
-covered values smaller the absolute median of those relevant by similar
-previous studies. After consulting the regression tables of Mitchell et
-al. (2022), Braginsky et al. (2019), and Laing (2022), we selected the
-largest regression coefficient that was interpreted as conclusive
-evidence of the presence of an effect, as indicated by the authors, and
-transformed it to the probability scale: 0.18, in the logit scale.
+\newpage{}
+
+\hypertarget{appendix-a-model-predictors}{%
+\section*{Appendix A: model
+predictors}\label{appendix-a-model-predictors}}
+\addcontentsline{toc}{section}{Appendix A: model predictors}
+
+\hypertarget{word-exposure-index}{%
+\subsection*{Word exposure index}\label{word-exposure-index}}
+\addcontentsline{toc}{subsection}{Word exposure index}
+
+Lexical frequencies were extracted from the English corpora of the
+\href{https://childes.talkbank.org/}{CHILDES} database (MacWhinney,
+2000; Sanchez et al., 2019). Using the corresponding lexical frequencies
+from the Catalan and Spanish corpora was not possible due to the low
+number of Catalan participants and tokens available in those languages.
+Available words in the English corpora were mapped to their Catalan and
+Spanish translations (see Fourtassi et al., 2020 for a similar
+approach), and transformed to Zipf scores (Van Heuven et al., 2014;
+Zipf, 1949). Responses to words with missing lexical frequencies were
+excluded from analyses of the total number of items. Participant degree
+of exposure to the words' language were calculated as the percentage of
+exposure to the language, as reported by caretakers in the language
+profile section of the questionnaire. For example, for a participant
+with 90\% exposure to Catalan, and 10\% to Spanish, the DoE of the
+Catalan word \emph{taula} would be 90\%, and the DoE for the Spanish
+word \emph{mesa} would be 10\%.
+
+We created a predictor (\(\text{Exposure}\)) to account for the exposure
+rate of each child to each of the words their parents responded to,
+weighted by the child's exposure to the language the word belongs to.
+the exposure of the \(i\)-th child to the \(j\)-th word measure is the
+product of the lexical frequency of the word (Zipf score) and the
+child's degree of exposure to the corresponding language (a proportion)
+(see Equation~\ref{eq-exposure}).
+
+\begin{equation}\protect\hypertarget{eq-exposure}{}{
+\text{Exposure}_{ij} = \text{Frequency}_i \times \text{DoE}_{j}
+}\label{eq-exposure}\end{equation}
+
+For instance, for a child who is reportedly exposure to Catalan 80\% of
+the time, and to Spanish 20\% of the time, the expected exposure to the
+word \emph{cavall} (horse, in Catalan, with a lexical frequency of )
+would be , while that of its translation to Spanish \emph{caballo} would
+be .
+
+\hypertarget{levenshtein-similarity}{%
+\subsection*{Levenshtein similarity}\label{levenshtein-similarity}}
+\addcontentsline{toc}{subsection}{Levenshtein similarity}
+
+Generally, a pair of words from two different languages that share
+meaning---translation equivalents---are considered as \emph{cognates} if
+they share etymological origin (e.g., \emph{table} and \emph{taula}, in
+English and Catalan). The fact that two words share etymological origin
+frequently leads to them also being form-similar, as reflected by their
+overlapping orthographic or phonological form. This makes many cognates
+perceptually similar. In psycholinguistics, the term \emph{cognateness}
+is often used to refer to the form-similarity that a pair of translation
+equivalents share, which has been found to impact how bilinguals process
+such word-forms (e.g., Costa et al. (2000); Spivey and Marian (1999)).
+Whether two form-similar translation equivalents are etymologically
+related or not is arguably tangential to the question of how cognateness
+affect language processing. For instance, the translation pair
+\emph{much} and \emph{mucho}, in English and Spanish have the same
+meaning, and are considerably similar at the orthographical and
+phonological level. Nonetheless, they come from different
+Proto-Indo-European roots (Campbell \& Mixco, 2007). This does not keep
+bilinguals from processing such kind of word-forms word forms
+differently than other translation equivalents with no form-similarity
+like `dog' and \emph{perro}, in English and Spanish. In the scope of
+this experiment, we consider any form-similar translation equivalent as
+a \emph{cognate} pair, regardless of whether both word-forms share
+etymological origin.
+
+Our study capitalises in phonology, therefore we operationalised
+\emph{cognateness} as the phonological similarity between a pair of
+translation equivalents. Our dataset contained 302 translation
+equivalents, each consisting in two word-forms: Catalan and Spanish (see
+Table~\ref{apatb-translation-levenshtein} for an example). Measuring the
+phonological similarity between two word forms is non trial. While
+orthographic similarity can be calculated using string
+similarity/distance measures on the written word-forms, phonology poses
+additional challenges. A first, bottom-up approach to to the task might
+be considering acoustic similarity as a proxy to phonological
+similarity. This may involve registering audio recordings of a talker
+reading each word-form aloud, and then comparing their spectrograms,
+cochleograms, or formant tracks (e.g., Heeringa \& Gooskens, 2003).
+While this is possible, this method computes acoustic similarity, as
+opposed to phonological similarity. This is not optimal, since a purely
+acoustic measure of similarity ignores how the actual signal is
+perceived by a actual listener. Since approximately 12 months of age,
+humans perceive the acoustic sounds in speech as phonemes, according to
+the phonology of their native language(s) (Kuhl et al., 1992; Werker \&
+Tees, 1984). This means that two listeners may not perceive the same
+linguistically relevant acoustic signal identically. This is why
+phonological similarity needs to be computed from discrete units that
+are meaningful for language perception.
+
+A way to achieve this is to use phonological transcriptions of the word
+forms. Phonological transcriptions are symbolic, visual representations
+of the sounds in speech that capture some minimal set of its features
+following some standard. The International Phonetic Alphabet (IPA) is
+one instance of such standards, providing a set of symbols that
+correspond to specific phonemes, as defined by their position across a
+series of dimensions. Using IPA transcriptions of word forms, one can
+use the same similarity/distance metric as in the case of orthographic
+word-forms. This method has been successfully used to measure the
+pairwise similarity between phonological word-forms from the same
+language (e.g., Fourtassi et al., 2020) and across languages (Floccia,
+Sambrook, Luche, et al., 2018a; Heeringa, 2004), and also between
+orthographic word-forms (Schepens et al., 2012).
+
+Multiple algorithms have been created to compute the distance between
+two strings of characters. The Levenshtein distance being one of them.
+The this algorithm calculates the minimum number of edit operations
+(additions, deletions, and substitutions) that one string must go
+through to become identical to the other string (Levenshtein, 1966) (see
+Equation~\ref{eq-levenshtein}). For instance, /kaza/ (\emph{house}, in
+Catalan), would need to go through one substitution to become identical
+to its translation equivalent in Spanish /kasa/ (/z/ is replaced by
+/s/). Therefore, the Levenshtein distance between /kaza/ and /kasa/ is
+one. For an easier handling of non-ASCII characters (fairly prevalent in
+Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions
+in X-SAMPA format , as opposed to IPA, for its computer-friendly set of
+symbols (Wells, 1995).
+
+\begin{equation}\protect\hypertarget{eq-levenshtein}{}{
+\text{lev}(x, y) =  
+\begin{cases}
+\max(i, j) & \text{if} \ \min(i, j)=0\\
+\min \begin{cases}
+\ \text{lev}_{x, y}(i-1, j) + 1 \\
+\ \text{lev}_{x, y}(i, j-1) + 1 \\
+\ \text{lev}_{x, y}(i-1, j-1) + 1_{a_i \neq b_j} \\
+\end{cases}
+\ &\text{otherwise}
+\end{cases}
+}\label{eq-levenshtein}\end{equation}
+
+Here, \(a\) and \(b\) are the character strings corresponding to the
+phonological transcriptions of two word-forms belonging to the same
+translation equivalent, where each character corresponds to one phoneme
+expressed as a symbol from the SAMPA alphabet. \(i\) and \(j\) are the
+length (i.e., number of phonemes) of the \(s\) and \(t\) strings
+respectively.
+
+To measure the phonological similarity between the translation
+equivalents in our study, we computed a normalised versions of the
+Levenshtein distance between each pair of word-forms that accounts for
+the lenght of the strings. The rationale behind this correction is that
+longer word-forms are more likely to differ from their counterpart,
+compared to shorter word forms, and that this tendency does not
+correspond necessarily to a larger distance in terms of perception. This
+normalisation is achieved by dividing the Levenshtein distance by the
+length of the longest string, which leads to a distance metric that
+ranges from 0 to 1. This can be interpreted as a proportion of
+characters in the longest string that need to be edited in order for
+that string to become identical to the shorter string. A 0\% normalised
+Levenshtein distance indicates that both word-forms are identical. A
+50\% normalised Levenshtein distance indicates that half of the phonemes
+in the longest word-form must be edited for both word-forms to become
+identical. A 100\% normalised Levenshtein distance indicates that the
+two word-forms are completely different, as \emph{all} phonemes in the
+longest word-form must be changed for it to become identical to the
+other.
+
+Finally, since we are interested in \emph{cognateness}, in order to easy
+the interpretation of the analyses we subtracted the normalised
+Levenshtein distance from one, so that the metric indicated the
+\emph{similarity} between the each pair of word-forms, instead of their
+distance. Equation~\ref{eq-levenshtein-similarity} shows the formula of
+this metric.
+
+\begin{equation}\protect\hypertarget{eq-levenshtein-similarity}{}{
+1-\frac{\text{lev}(a, b)}{\max{\{i, j}\}}
+}\label{eq-levenshtein-similarity}\end{equation}
+
+The R package \texttt{stringdist} (Loo, 2014) offers the
+\texttt{stringsim()} function to compute this normalised Levenshtein
+similarity measure, see Table~\ref{apatb-translation-levenshtein} for
+more examples:
+
+\begin{table}
+\caption{Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study}
+\label{apatb-translation-levenshtein}
+
+\begin{longtable*}{llr}
+\toprule
+Catalan & Spanish & Levenshtein \\ 
+\midrule
+porta [pOrt5] & puerta [pwe4ta] & $0.50$ \\ 
+taula [tawl5] & mesa [mesa] & $0.00$ \\ 
+cotxe [kOtS5] & coche [kotSe] & $0.40$ \\ 
+\bottomrule
+\end{longtable*}
+
+\end{table}
+
+\newpage{}
+
+\hypertarget{appendix-b-model-details}{%
+\section*{Appendix B: model details}\label{appendix-b-model-details}}
+\addcontentsline{toc}{section}{Appendix B: model details}
+
+We used multilevel ordinal regression to model the cumulative
+probability of a \emph{No} response, a \emph{Understands} response, or a
+\emph{Understands and Says} response (Bürkner \& Vuorre, 2019) using the
+\emph{logit} link function. The likelihood of the cumulative probability
+distribution of the responses is defined by
+Equation~\ref{eq-likelihood}.
+
+\begin{equation}\protect\hypertarget{eq-likelihood}{}{
+\begin{aligned}
+\textbf{Likelihood:} \\
+y_{ij} &\sim \text{Cumulative}(p_{k})
+\end{aligned}
+}\label{eq-likelihood}\end{equation}
+
+where: \(y\) is an observed response
+(\(y \in \{\text{No, Understands, Understands and Says}\}\)), \(i\) is
+the participant index, \(j\) is the translation equivalent (TE) index,
+\(p_{k}\) is a probability (\(p \in (0, 1)\)) that indicates the
+threshold \(k\) (\(k \in (1, 2)\)) between two response categories in
+the latent distribution \(p_{k}\) is then estimated using a logit
+regression model as indicated in Equation~\ref{eq-linear}.
+
+To test our hypotheses, we included several predictors in the regression
+model as fixed effects: the main effects of participant age
+(\(\text{Age}\)), word-form length (\(\text{Length}\)), word exposure
+index (\(\text{Exposure}\)), and of phonological similarity
+(\(\text{Cognateness}\)), the two-way interactions
+\(\text{Age} \times \text{Exposure}\),
+\(\text{Age} \times \text{Cognateness}\), and
+\(\text{Exposure} \times \text{Cognateness}\), and the three-way
+interaction
+\(\text{Age} \times \text{Exposure} \times \text{Cognateness}\). We also
+included crossed random effects for participants and translation
+equivalents to account for the repeated measures in our dataset---each
+participant provided responses to multiple translation equivalents, and
+each translation equivalent was responded to by multiple participants
+(Gelman et al., 2020). For both grouping variables, we included random
+intercepts, random slopes, and correlation parameters for all predictors
+were repeated measures were observed in our dataset (Barr et al., 2013),
+see (Equation~\ref{eq-linear}).
+
+\begin{equation}\protect\hypertarget{eq-linear}{}{
+\begin{aligned}
+\textbf{Linear model:} \\
+\text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
+& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + & \\
+& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + & \\
+& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{ij} + & \\
+& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
+& (\beta_{6} + u_{6_{i}}) \cdot (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
+& (\beta_{7} + u_{7_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
+& (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij})
+\end{aligned}
+}\label{eq-linear}\end{equation}
+
+where \(i\) and \(j\) index the participant and translation equivalent
+(TE), \(\beta_{0_k}\) is the fixed coefficient of the regression model
+for the intercept of threshold \(k\), \(u_{0_{i}}\) and \(w_{0_{j}}\)
+are the by-participant and by-TE adjustments for \(\beta_{0_{k}}\)
+(i.e., random intercepts), respectively, \(\beta_{1-8}\) are the fixed
+coefficients of the regression model for the predictors of interest,
+\(u_{1-8_{i}}\) and \(w_{1-3_{j}}\) are the by-participant and by-TE
+adjustments for\(\beta_{1-8}\) (i.e., random slopes), respectively.
+
+We used the Bayesian framework to estimate the parameters in our model.
+This involves using the Bayes theorem to compute a distribution
+(\emph{posterior distribution}) that describes what values of each
+parameter in the model are more likely given the data
+(\emph{likelihood}, see Equation~\ref{eq-likelihood}), and previous
+knowledge about such distribution (\emph{prior distribution}, see
+Equation~\ref{eq-prior}) (McElreath, 2020). This posterior distribution
+not only informs about the most likely values of our regression
+coefficients of interest, but also about the uncertainty around such
+estimations. We used a weakly informative prior for our parameters, with
+the exception of the main effect of \(\text{Age}\), for which we
+specified a strongly informative prior based on previous literature
+about how age affects the acquisition of words (see
+Equation~\ref{eq-prior}).
+
+\begin{equation}\protect\hypertarget{eq-prior}{}{
+\begin{aligned}
+\\
+\textbf{Prior:} \\
+\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
+\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
+\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
+u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
+w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
+&&\mbox{[Participant-level coefficient variability]} \\ \\
+\Bigg(\begin{smallmatrix}
+u_{k_{0}} \\ 
+u_{1_{i}} \\ 
+\vdots \\ 
+u_{8_{i}} 
+\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
+\Bigg(\Bigg(\begin{smallmatrix}0 \\
+0 \\ 
+\vdots \\
+0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
+\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
+\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
+\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
+\vdots & \vdots & \vdots & \vdots \\
+\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
+\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
+\rho_{u} &\sim LKJcorr(2) \\
+\\
+&&\mbox{[TE-level coefficient variability]} \\ \\
+\Bigg(\begin{smallmatrix}
+w_{k_{0}}\\ 
+w_{1_{j}} \\ 
+\vdots \\ 
+w_{3_{j}} 
+\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
+0\\ 
+0 \\ 
+\vdots \\
+0 
+\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
+\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
+\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
+\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
+\vdots & \vdots & \vdots & \vdots \\
+\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
+\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
+\rho_{w_{0-3}} &\sim LKJcorr(2)
+\end{aligned}
+}\label{eq-prior}\end{equation}
+
+where: \(\rho_{u_{0-8}}\) and \(\rho_{w_{0-3}}\) indicate the
+correlation parameters between the by-participant and by-TE adjustments,
+respectively, \(\sigma_{u_{0-8}}^2\) and \(\sigma_{w_{0-3}}^2\) indicate
+the variance of the by-participant and by-TE variance of the
+adjustments, respectively, and \(\mathcal{N}\) indicates a normal
+distribution, \(\mathcal{N}_{+}\) indicates a truncated normal
+distribution with only positive values, and \(LKJcorr\) indicates a
+\href{https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html}{LKJ
+correlation distribution} (Lewandowski et al., 2009).
+
+\begin{figure}[h!]
+\caption{Expected prior-predicted mean.}
+\label{apafg-prior-mean}
+\includegraphics[width=7 in]{manuscript_files/figure-pdf/apafg-prior-mean-1.pdf}
+
+\end{figure}
+
+\newpage{}
+
+\newpage{}
+
+\hypertarget{appendix-c-convergence-diagnostics}{%
+\section*{Appendix C: convergence
+diagnostics}\label{appendix-c-convergence-diagnostics}}
+\addcontentsline{toc}{section}{Appendix C: convergence diagnostics}
+
+We used Stan (Carpenter et al., 2017) as the probabilistic language
+behind the estimation of our Bayesian models in this study, with
+\texttt{brms} as its R interface (Bürkner, 2017). This language
+implements the Markov Chain Monte-Carlo (MCMC) algorithm to explore the
+posterior distribution of the model. Broadly, this algorithm is used to
+iteratively sample the joint sampling space of the parameters to be
+estimated in the model, and compute, for each value sampled, its
+likelihood under some probability distribution previously defined. We
+run 4 MCMC chains, each 1000 iterations long each. The correct
+performance of this algorithm is critical to the quality of the
+statistical evidence to which the outcomes of the model lead.
+
+One way to diagnose the behaviour of MCMC is to inspect whether the
+different MCMC chains (if more than one) have converged to a similar
+region of the posterior. The Gelman-Rubin diagnostic (\(\hat{R}\) or
+R-hat Gelman \& Rubin, 1992) provides a measure of chain convergence by
+comparing the variance within each chain \emph{versus} the variance
+between each chain. Both are expected to be identical when chains have
+perfectly converged, so that \(\hat{R} = 1\). Values lower than 1.01 are
+recommended, while values higher than 1.05 indicate that chains might
+have trouble converging and therefore the estimated parameters must be
+taken with caution. Figure~\ref{apafg-model-rhats-neffs} (A) shows the
+distribution of \(\hat{R}\) values for the coefficients of the fixed
+effect of our models, which we used for statistical inference. Most
+values are lower than 1.01, and never higher than 1.05, which provides
+evidence of successful MCMC convergence.
+
+Another diagnostic of good MCMC converge is the ratio of effective
+sample size to total sample size (\(N_{eff}/N\)), which indicates the
+proportion of samples in the chain that resulted from a non-divergent
+transition. Values closer to 1 are ideal, as they indicate that all
+posterior samples from the MCMC were used to estimate the posterior
+distribution of the parameter. Values larger than 0.1 are recommended.
+Figure~\ref{apafg-model-rhats-neffs} (B) shows the distribution of the
+effective sample sizes of the coefficients of the fixed effects in our
+models. Most values are larger than 0.1, although model 0
+(\(\mathcal{M}_0\)) accumulates most effective sample sizes close to
+0.1.
+
+\begin{figure}[h!]
+\caption{Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models.}
+\label{apafg-model-rhats-neffs}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-pdf/apafg-model-rhats-neffs-1.pdf}
+
+\end{figure}
+
+\hypertarget{posterior-draws-and-bi-variate-scatterplots}{%
+\subsection*{Posterior draws and bi-variate
+scatterplots}\label{posterior-draws-and-bi-variate-scatterplots}}
+\addcontentsline{toc}{subsection}{Posterior draws and bi-variate
+scatterplots}
+
+\begin{figure}[h!]
+\caption{}
+\label{fig-model-pairs}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-pdf/fig-model-pairs-1.pdf}
+
+\end{figure}
+
+\hypertarget{posterior-predictive-checks}{%
+\subsection*{Posterior-predictive
+checks}\label{posterior-predictive-checks}}
+\addcontentsline{toc}{subsection}{Posterior-predictive checks}
+
+\begin{figure}[h!]
+\caption{Model posterior predictive checks (PPC).
+Bars indicate the observed proportion of responses to each category (No, Understands, and Understands and Says).
+Blue dots and error bars represent the mean proportion of responses simulated from the posterior for each category, and its 95\% interval.}
+\label{apafg-model-ppc}
+\includegraphics[width=0.5\linewidth]{manuscript_files/figure-pdf/apafg-model-ppc-1.pdf}
+
+\end{figure}
+
+\newpage{}
+
+\hypertarget{appendix-d-frequency-and-language-exposure-as-separate-predictors}{%
+\section*{Appendix D: frequency and language exposure as separate
+predictors}\label{appendix-d-frequency-and-language-exposure-as-separate-predictors}}
+\addcontentsline{toc}{section}{Appendix D: frequency and language
+exposure as separate predictors}
+
+\begin{figure}[h!]
+\caption{Posterior distribution of regression coefficients in the probability scale}
+\label{apafg-coefs-doe}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-pdf/apafg-coefs-doe-1.pdf}
+
+\figurenote{Median: median of the posterior distribution in the logit scale.
+95\% HDI: 95\% highest density interval of the distribution; narrowest range of values that contains 95\% of the distribution, and therefore is 95\% likely to contain the true value, given the data.
+On top of each distribution, we indicate the proportion of posterior samples in the 95\% HDI that fall into the region of practical equivalence (ROPE).
+This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
+Lower values indicate that the true value is unlikely to be zero of equivalent to zero.}
+
+\end{figure}
+
+\newpage{}
 

### 2023-04-17 appendix/_model-doe.qmd
Fix comments and use apaquarto extension

```{r apafg-coefs-doe}
#| label: apafg-coefs-doe
#| apa-cap: "Posterior distribution of regression coefficients in the probability scale"
#| apa-note: "Median: median of the posterior distribution in the logit scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
#| out-width: 100%
#| fig-height: 3
#| fig-width: 12
# summarise posterior draws
bind_rows("Exposure model" = posterior_draws$summary,
          "DoE-Frequency model" = posterior_draws_doe,
          .id = "model") |> 
    mutate(
        across(c(.median, .lower, .upper),
               function(x) {
                   ifelse(
                       .type=="Intercepts (at 22 months)",
                       plogis(x),
                       x/4
                   )
               }
        )) |>
    filter(.type == "Slopes") |> 
    ggplot(aes(.median, .variable_name,
               xmin = .lower, xmax = .upper)) +
    facet_wrap(~model, scales = "free") +
    annotate(geom = "rect",
             ymin = -Inf,
             ymax = Inf,
             xmin = -0.1/4,
             xmax = +0.1/4,
             fill = "grey70") +
    geom_errorbar(linewidth = 1, width = 0.25) +
    geom_point(size = 2.5) +
    labs(x = "Median and 95% HDI of the posterior distribution",
         y = "Predictor",
         colour = "Model") +
    scale_x_continuous(labels = percent) +
    scale_colour_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          axis.title.y = element_blank(),
          panel.grid.major.y = element_blank())
```

### 2023-04-17 R/figures.R

@@ -1,269 +1,262 @@
-library(ggplot2)
-library(patchwork)
-library(tibble)
-library(dplyr)
-library(tidyr)
-library(purrr)
-library(scales)
-library(bvqdev)
-library(glue)
-library(ggtext)
-library(magick)
-
-resolve_conflicts()
-
-# set options ------------------------------------------------------------------
-clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
-options(ggplot2.ordinal.fill = clrs,
-        ggplot2.ordinal.colour = clrs,
-        ggplot2.discrete.fill = clrs,
-        ggplot2.discrete.colour = clrs,
-        ggplot2.continuous.fill = scale_color_gradient,
-        ggplot2.continuous.colour = scale_color_gradient)
-
-set.seed(888)
-
-# set parameters ---------------------------------------------------------------
-
-generate_aoa <- function(conditions = c("**Cognate**", "**Non-cognate**"),
-                         items = c("gat /gat/",
-                                   "gato /gato/",
-                                   "gos /gos/", 
-                                   "perro /ˈpe.ro/"),
-                         age_step = 0.5,
-                         age_range = c(10, 34),
-                         freq_month = 50,
-                         freq_month_sd = 10,
-                         thres = 500,
-                         doe_l1 = 0.65,
-                         thres_diff = 100) 
-{
+generate_figures <- function() {
+    resolve_conflicts()
     
-    item_df <- tibble(te = rep(conditions, each = 2),
-                      item = items,
-                      doe = c(L1 = doe_l1, 
-                              Spanish = 1-doe_l1, 
-                              L1 = doe_l1, 
-                              Spanish = 1-doe_l1))
+    # set options ------------------------------------------------------------------
+    clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
     
-    eli_df <- expand_grid(item_df,
-                          age = seq(age_range[1], age_range[2], age_step),
-                          n_month = freq_month*age_step,
-                          threshold = thres,
-                          hypothesis = c("No cognateness effect",
-                                         "Cognateness effect")) |> 
-        mutate(n_month_wgt = ceiling(n_month*doe),
-               eli = n_month_wgt + rnorm(n(), 0, freq_month_sd),
[…]

### 2023-04-28 _appendix/_model-details.qmd
Add appendix for syllable frequencies and new revisions

--- 2023-02-26 manuscript/_model-details.qmd
@@
 
-where:
+where: $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$), $i$ is the participant index, $j$ is the translation equivalent (TE) index, $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$) between two response categories in the latent distribution $p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.
 
-- $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$)
-- $i$ is the participant index
-- $j$ is the translation equivalent (TE) index
-- $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$)) between two response categories in the latent distribution
-$p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.
-
-To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), word-form length ($Length$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
+To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($\text{Age}$), word-form length ($\text{Length}$), word exposure index ($\text{Exposure}$), and of phonological similarity ($\text{Cognateness}$), the two-way interactions $\text{Age} \times \text{Exposure}$, $\text{Age} \times \text{Cognateness}$, and $\text{Exposure} \times \text{Cognateness}$, and the three-way interaction $\text{Age} \times \text{Exposure} \times \text{Cognateness}$.
 We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
-For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].
+For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random], see (@eq-linear).
 
@@
 \textbf{Linear model:} \\
-logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
-& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
-& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Length}_{ij} + & \\
-& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
-& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
-& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
-& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
-& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
-& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
+\text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
+& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + & \\
+& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + & \\
+& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{ij} + & \\
+& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
+& (\beta_{6} + u_{6_{i}}) \cdot (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
+& (\beta_{7} + u_{7_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
+& (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij})
 \end{aligned}
@@
 
-where:
-
-- $i$ and $j$ index the participant and translation equivalent (TE)
-- $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$
-- $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively
-- $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest
-- $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively
+where $i$ and $j$ index the participant and translation equivalent (TE), $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$, $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively, $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest, $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively.
 
@@
 This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations.
-We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].
+We used a weakly informative prior for our parameters, with the exception of the main effect of $\text{Age}$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words (see @eq-prior).
 
@@
 
-where:
+where: $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively, $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively, and $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].
 
-- $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively
-- $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively
-- $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].
-
-```{r}
-#| label: fig-prior-mean
+```{r apafg}
+#| label: apafg-prior-mean
 #| echo: false
@@
 #| warning: false
-#| eval: false
 #| fig-width: 7
 #| fig-height: 5
-#| fig-cap: "Expected prior predictions."
-nd <- expand.grid(
-    n_phon_std = 0,
+#| apa-cap: "Expected prior-predicted mean."
+nd <- expand.grid(n_phon_std = 0,
                   age_std = (10:36-mean(responses$age))/(sd(responses$age)),
                   exposure_std = 0,
-    lv_std = 0
-)
+                  lv_std = 0)
 
@@
     mutate(Understands = Understands + `Understands and Says`) |> 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
+    pivot_longer(c(Understands, `Understands and Says`),
                  names_to = ".category",
-        values_to = ".epred"
-    ) |> 
-    mutate(
-        .category = factor(
-            .category,
-            levels = c("Understands and Says", "Understands"),
-            labels = c("Production\n(Understands and Says)", "Comprehension\n(Understands)")
-        )
-    )
+                 values_to = ".epred") |> 
+    mutate(.category = factor(.category,
+                              levels = c("Understands and Says",
+                                         "Understands"),
+                              labels = c("Production\n(Understands and Says)",
+                                         "Comprehension\n(Understands)")))
 
@@
     facet_wrap(~.category) +
-    stat_lineribbon(linewidth = 1, 
+    stat_lineribbon(size = 0.75, 
                     colour = NA,
                     .width = c(0.95, 0.89, 0.67, 0.5)) +
-    stat_summary(fun = "mean", geom = "line", linewidth = 0.75) +
-    scale_x_continuous(
-        breaks = scale(seq(10, 36, 4), 
+    stat_summary(fun = "mean", 
+                 geom = "line",
+                 linewidth = 0.75) +
+    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
                                       mean(responses$age), 
                                       sd(responses$age))[, 1],
-        labels = seq(10, 36, 4)
-    ) +
-    scale_fill_manual(
-        values = rev(clrs), 
-        labels = c("95%", "89%", "67%", "50%")
-    ) +
-    labs(
-        x = "Age (months)", 
+                       labels = seq(10, 36, 4)) +
+    scale_fill_manual(values = rev(clrs), 
+                      labels = c("95%", "89%", "67%", "50%")) +
+    labs(x = "Age (months)", 
          y = "Probability of acquisition",
          colour = "Credible interval", 
-        fill = "Credible interval"
-    ) +
+         fill = "Credible interval") +
     scale_y_continuous(labels = percent, limits = c(0, 1)) +
-    theme(
-        legend.position = "top",
+    theme(legend.position = "top",
           axis.ticks.x = element_line(colour = "black"),
-        panel.grid = element_blank()
-    )
+          panel.grid = element_blank())
 ```

### 2023-04-28 _appendix/_syllable-frequency.qmd
Add appendix for syllable frequencies and new revisions

````{r}
duplicated_te <- bvq_data$pool$te[duplicated(bvq_data$pool$te)]

pool_tmp <- bvq_data$pool |>
    # drop items with missing observations in these variables
    drop_na(sampa, wordbank_lemma) |>
    mutate(xsampa = ipa(ipa)) |> 
    filter(n_lemmas == 1,
           # exclude items with more than two lemmas
           !is_multiword,
           # exclude multi-word items
           include,
           # exclude problematic items (e.g., multi-word items)
           te %in% duplicated_te) |>
    add_count(te, name = "n_te") |> # get only items with one translation in each language
    filter(n_te == 2) |>
    distinct(language, te, .keep_all = TRUE) |> 
    select(language, te, item, label, childes_lemma, xsampa) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_"))

items <- items |> 
    mutate(xsampa = ipa(ipa)) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_"))
```

We define syllable frequency as the rate of appearance of individual syllables in the word-forms included in the Barcelona Vocabulary Questionnaire (BVQ). Each item corresponds to a Catalan or Spanish word, and has an associated phonological transcription in X-SAMPA format. These transcriptions are syllabified. Some examples:

```{r items-table}
#| label: tbl-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
items |> 
    distinct(te, item, language, meaning, xsampa, n_syll) |> 
    left_join(distinct(pool_tmp, te, language, item, label),
              by = join_by(te, item, language)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, xsampa, n_syll),
                id_cols = c(te, meaning),
                names_repair = make_clean_names,
                values_fn = first) |> 
    sample_n(15) |> 
    gt() |> 
    cols_hide(c(te, item_catalan, item_spanish)) |> 
    tab_spanner("Catalan", matches("catalan")) |> 
    tab_spanner("Spanish", matches("spanish")) |> 
    cols_label(meaning = "Translation",
               label_catalan = "Item",
               xsampa_catalan = "X-SAMPA",
               n_syll_catalan = "Syllables",
               label_spanish = "Item",
               xsampa_spanish = "X-SAMPA",
               n_syll_spanish = "Syllables") |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels())
```

Most Catalan and Spanish words had two syllables, with Spanish words having three and four syllables more often than Catalan words. Less than 1% of the words included in the analyses presented in the main body of the manuscripts had five syllables. No words had more than five syllables (see {apafg-syl-number}). We extracted lexical frequencies from the English corpora in the CHILDES database. Using the Catalan and Spanish corpora was not possible due to the low number of children and tokens included in the corpora.

```{r apafg-syl-number}
#| label: apafg-syl-number
#| apa-cap: Distribution of the number of syllables in Catalan and Spanish
#| fig-width: 7
#| fig-height: 3
pool_tmp |> 
    group_by(language) |> 
    add_count(language, name = "n_lang") |> 
    ungroup() |> 
    count(language, n_lang, n_syll) |> 
    mutate(prop = n/n_lang) |> 
    ggplot(aes(n_syll, n)) +
    facet_wrap(~ language) +
    geom_col(colour = "white", fill = "grey70") +
    geom_text(aes(label = percent(prop)),
              size = 3,
              position = position_nudge(y = 15),
              colour = "black") +
    labs(x = "Number of syllables",
         y = "Number of words",
         fill = "Language") +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
```

We now present how syllable frequencies were calculated. Every exposure to a word-form also counts as a exposure to each of the syllables that make up such word. Every time a child hears the word *casa* [house], they are exposed to the syllables *ca* and *sa*. Syllables that appear embedded in words with higher lexical frequency will also be more frequent. To compute the relative frequency of each syllable in Catalan and Spanish (i.e., how many times the syllables appears in every million words in Catalan or Spanish speech), we summed the relative lexical frequency in CHILDES of every word that contains such syllable in the corresponding language. {apafg-syl-freq} shows the distribution of frequencies across syllables in Catalan and Spanish. In the log10 scale, syllable frequencies in Catalan and Spanish followed a slightly asymmetric distribution, with most syllables scoring around 1,000 counts per million, and a longer tail to the right of the distribution.

```{r apafg-syl-freq}
#| label: apafg-syl-freq
#| apa-cap: "Distribution of apositional syllable frequencies in Spanish and Catalan"
#| fig-height: 3.5
#| fig-width: 7
syll <- items |> 
    left_join(distinct(pool_tmp, te, childes_lemma),
              by = join_by(te)) |> 
    left_join(select(childes, token, freq_lex = freq_million),
              by = c("childes_lemma" = "token")) |> 
    select(language, item, te, meaning, xsampa, lv, freq_lex, syll) |> 
    unnest_longer(syll) |> 
    group_by(language, syll) |> 
    summarise(freq_syll = sum(freq_lex),
              .groups = "drop") |> 
    arrange(language, desc(freq_syll))

syll |> 
    add_count(language, name = "n_lang") |> 
    mutate(language = paste0(language, " (n = ", n_lang, ")")) |> 
    ggplot(aes(freq_syll)) +
    facet_wrap(~language) +
    geom_histogram(colour = "white",
                   bins = 20,
                   fill = "grey70",
                   position = position_nudge(y = 0.1)) +
    geom_rug(alpha = 1/10) +
    labs(x = "Relative frequency (counts per million)",
         y = "Counts",
         fill = "Language",
         colour = "Language") +
    scale_x_continuous(trans = "log10",
                       labels = \(x) format(x, big.mark = ",")) +
    theme(legend.title = element_blank(),
          panel.grid = element_blank(),
          legend.position = "none")
```

{apafg-top-syll-freq} shows the top twenty most frequent syllables in Spanish. In both languages, most of the most frequent syllables followed a consonant-vowel (CV) structured. In Catalan, the most common nucleus in the most frequent syllables was /\textschwa/ (most likely due to vowel reduction in unstressed syllables), followed by /u/. In Spanish, the most common nucleus in the most frequent syllables were /a/ and /o/.

```{r apafg-top-syll-freq}
#| label: apafg-top-syll-freq
#| apa-cap: "Top 20 most frequent syllables"
#| fig-width: 8
#| fig-height: 3
#| out-width: 80%
syll |> 
    group_by(language) |> 
    top_n(20, freq_syll) |> 
    ggplot(aes(reorder_within(syll, desc(freq_syll), language),
               freq_syll)) +
    facet_wrap(~ language,
               scales = "free_x",
               ncol = 1) +
    geom_col(colour = "white",
             fill = "grey70") +
    labs(x = "Syllable",
         y = "Relative frequency\n(counts per million)",
         colour = "Language",
         fill = "Language") +
    scale_x_reordered() +
    scale_y_continuous(labels = \(x) format(x, big.mark = ",")) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 10),
          panel.grid = element_blank())
```

To estimate the association between word-level syllabic frequency and cognateness while controlling for the number of syllables in the word, as words are expected to necessarily increase the syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ (Levenshtein similarity between a word and its translation equivalent) as predictors. We added translation equivalent-level random effects for the intercept and the main effect of $Syllables$ (some translation pairs had a different number of syllables in each language). We used a Gaussian distribution to model syllabic frequency scores after standardising this variable and the predictors. We used a weakly informative prior for all parameters involved in the model (see @eq-model-syllables for a formal equation of this model and its prior).

$$
\begin{aligned}
\text{Syllable frequency} &\sim \mathcal{N}(\mu, \sigma)\\
\mu &= (\beta_0 + u_{0_{i}}) + (\beta_1 + u_{1_{i}}) \text{Syllables} + \beta_2 \text{Cognateness} \\
\beta_{0-3} &\sim \mathcal{N}(0, 10) \\
u_{{0-1}_{i}} &\sim \mathcal{N}(0, \sigma_{u_i}) \\
\sigma_y & \sim \text{Exponential}(2)\\
\Big(\begin{smallmatrix}
u_{0_{i}} \\ 
u_{1_{i}} \\ 
\end{smallmatrix}\Big) &\sim \mathcal{N} 
\Big(\Big(\begin{smallmatrix}0 \\
0\end{smallmatrix}\Big), \Sigma_{u}\Big) \\
\Sigma_{u} &= \Big(\begin{smallmatrix} \\
\sigma_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{u_{1}} \\ 
\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \sigma_{u_{1}} \end{smallmatrix}\Big) \\
\sigma_{u_{0-1}} &\sim \mathcal{N_{+}}(1, 0.1) \\
\rho_{u} &\sim LKJcorr(2) \\
\end{aligned}
$$ {#eq-model-syllables}

```{r syllables-posterior}
rope_interval <- c(lower = -0.5, upper = 0.5)
# tidy predictor names
str_repl <- c(
    "b_Intercept" = "Intercept",
    "b_n_syll_std" = glue("Syllables (+1 SD, {round(sd(syllables_data$n_syll), 3)})"),
    "b_lv_std" = glue("Cognateness (+1 SD, {percent(sd(syllables_data$lv), accuracy = 0.01)})")
)

post_syllables <- model_fit_syllables |>
    describe_posterior(test = "rope",
                       rope_range = rope_interval,
                       ci_method = "HDI") |>
    as_tibble() |>
    clean_names() |>
    mutate(.variable_name = factor(
        parameter,
        levels = names(str_repl),
        labels = str_repl) |>
            as.character(),
        parameter = ifelse(str_detect(parameter, "Intercept"),
                           str_remove_all(parameter, "Intercept \\(|\\)"),
                           parameter)) |>
    select(.variable = parameter,
           .variable_name,
           .median = median,
           .upper = ci_high,
           .lower = ci_low,
           .rope_overlap = rope_percentage) |>
    ungroup()

post_draws_list <- post_syllables |> 
    split(post_syllables$.variable)

```

We fit this model running `r dim(model_fit_syllables$fit)[2]` sampling chains with `r format(dim(model_fit_syllables$fit)[1], big.mark = ",")` iterations each. {apatb-coefs} shows a summary of the posterior distribution of the fixed effects in the model. As expected, words with more syllables scored higher in syllabic frequency: all posterior draws for the regression coefficient of the main effect of this predictor fell outside the ROPE defined between -0.5 and +0.5 ($\beta$ = `r round(post_draws_list[["b_n_syll_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_syll_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_syll_std"]]$.upper, 2)`]). Keeping the number of syllables constant, the effect of cognateness was negligible: all of the posterior distributions of this predictor fell within the ROPE, providing evidence that the true value of the increment in syllabic frequency for every increase in cognateness is equivalent to zero ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).

```{r apatb-syllable-coefs}
#| label: apatb-syllable-coefs
#| apa-cap: "Posterior distribution of regression coefficients."
#| apa-note: "Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
post_syllables |> 
    select(.variable_name, .median, .lower, .upper, .rope_overlap) |>
    gt(rowname_col = ".variable_name") |>
    fmt_percent(.rope_overlap, decimals = 2) |> 
    fmt_number(c(.median:.upper), decimals = 2) |>
    cols_merge(c(.lower, .upper),
               pattern = "[{1}, {2}]") |>
    cols_label(
        .variable_name = "Parameter",
        .median = "Median",
        .lower = md("95% HDI"),
        .rope_overlap = md("ROPE prob.")
    ) |>
    tab_style(cell_text(weight = "bold"),
              cells_column_labels(columns = 1:5)) |>
    tab_style(cell_text(align = "left"),
              cells_title(groups = "subtitle")) |> 
    tab_style(cell_text(size = "small"), 
              list(cells_body(),
                   cells_row_groups(),
                   cells_stub())) |> 
    tab_style(cell_text(style = "italic"),
              cells_row_groups())
```

{apafg-syllable-marginal} shows the median posterior-predicted syllabic frequencies for words with one to four syllables, for the whole range of cognateness values. Overall, cognate words' syllabic frequency is equivalent to that of non-cognates. This suggests that the cognate facilitation effect in word acquisition reported in the present study is not the result from an association between cognateness and higher syllabic frequencies.

```{r apafg-syllable-marginal}
#| label: apafg-syllable-marginal
#| apa-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
#| fig-width: 7
#| fig-height: 5
#| out-width: 80%
#| echo: false
#| message: false
#| warning: false
expand_grid(lv_std = scale(seq(0, 1, 0.1),
                           center = mean(syllables_data$lv),
                           scale = sd(syllables_data$lv))[, 1],
            n_syll_std = scale(seq(1, 4, 1),
                           center = mean(syllables_data$n_syll),
                           scale = sd(syllables_data$n_syll))[, 1]) |> 
    add_epred_draws(model_fit_syllables, ndraws = 50,
                    re_formula = NA) |> 
    mutate(lv_std = rescale_variable(lv_std,
                                     mean = mean(syllables_data$lv),
                                     sd = sd(syllables_data$lv)),
           n_syll_std = rescale_variable(n_syll_std,
                                     mean = mean(syllables_data$n_syll),
                                     sd = sd(syllables_data$n_syll)),
           n_syll_std = as.factor(n_syll_std)) |> 
    ggplot(aes(lv_std, .epred, colour = n_syll_std)) +
    geom_point(data = filter(syllables_data, n_syll < 5),
               aes(x = lv,
                   y = freq_syll,
                   colour = as.factor(n_syll)),
               shape = 1,
               stroke = 1,
               alpha = 1/2) +
    geom_line(aes(group = interaction(.draw, n_syll_std)),
              alpha = 1/10) +
    labs(x = "Cognateness (Levenshtein similarity)",
         y = "Syllabic frequency",
         colour = "Number of syllables") +
    scale_color_manual(values = clrs) +
    scale_y_continuous(n.breaks = 8) +
    theme(panel.grid = element_blank())
```

### 2023-04-28 docs/index.qmd
Add appendix for syllable frequencies and new revisions

@@
     aes(x = age, y = n) +
-    geom_col(fill = clrs[1]) +
+    geom_col(fill = "grey60") +
     geom_text(aes(label = n), size = 3.5, vjust = -1) +
-    labs(
-        x = "Age (months)",
-        y = "# participants"
-    ) +
+    labs(x = "Age (months)",
+         y = "# participants") +
     scale_x_continuous(breaks = seq(0, 40, 2)) +
-    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
+    scale_y_continuous(limits = c(0, 55),
+                       breaks = seq(0, 55, 10)) +
     theme(
@@
 
-```{r}
-#| label: fig-participants-lp
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
+```{r apatb-participants}
+#| label: tab-participants
+#| echo: false
+#| message: false
+#| warning: false
+#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan.
+Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
+In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
+A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
+A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
+The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
+For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins.
+Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
+participants |>
+    mutate_at(vars(starts_with("doe_")), 
+              (\(x) floor(x * 10) / 10)) |> 
+    mutate(
+        age = cut(floor(age),
+                  c(10, 14, 18, 22, 26, 30, 34, 36),
+                  include.lowest = TRUE),
+        doe_catalan = cut(
+            doe_catalan,
+            c(0, 0.25, 0.5, 0.75, 1),
+            labels = c("0-25%",
+                       "25-50%", 
+                       "50-75%",
+                       "75-100%"),
+            include.lowest = TRUE
+        ) |>
+            as.character()
+    ) |>
+    count(age, doe_catalan) |>
+    arrange(desc(doe_catalan)) |>
+    pivot_wider(names_from = age, values_from = n) |>
+    gt() |>
+    summary_rows(
+        columns = 2:7,
+        fns = list("Total" = ~ sum(., na.rm = TRUE)),
+        formatter = fmt_integer
+    ) |>
+    sub_missing(everything(), everything(), missing_text = "--") |>
+    cols_label(doe_catalan = "Catalan exposure") |>
+    tab_spanner("Age (months)", 2:8) |>
+    tab_style(cell_text(align = "left"),
+              list(cells_body(doe_catalan),
+                   cells_column_labels(doe_catalan)))
+```
+
+### SES/parental education
+
+```{r}
+#| label: tbl-participants-edu
+#|  tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)."
 participants %>%
-    pivot_longer(
-        c(doe_catalan, doe_spanish),
-        names_to = "language", values_to = "doe", 
-        names_transform = ~str_to_sentence(str_remove(., "doe_"))
+    mutate(id = as.character(id)) |> 
+    left_join(bvq_data$logs) %>%
+    mutate(
+        edu_parent = edu_parent |> 
+            as.factor() |> 
+            fct_na_value_to_level(),
+        doe_catalan = cut(doe_catalan,
+                          breaks = seq(0, 1, 0.2),
+                          include.lowest = TRUE)
     ) %>%
-    mutate(
-        doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
-        id = paste0(id, " (", time, ")")
+    count(doe_catalan, edu_parent) %>%
+    right_join(
+        expand_grid(
+            doe_catalan = levels(.$doe_catalan),
+            edu_parent = unique(bvq_data$logs$edu_parent)
+        )
     ) %>%
-    count(language, doe) %>% 
-    ggplot(aes(doe, y = n, label = n, fill = language)) +
-    geom_col(colour = "white", size = 1) +
-    geom_text(position = position_nudge(y = 5)) +
-    facet_wrap(~language, ncol = 1) + 
-    labs(
-        x = "Degree of exposure to Catalan",
-        y = "Number of participants",
-        colour = "Language",
-        fill = "Language"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor = element_blank()
-    )
-```
-
-### SES/parental education
-
-<!-- # ```{r} -->
-<!-- #| label: tbl-participants-edu -->
-<!-- #|  tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)." -->
-<!-- # participants %>% -->
-<!-- #     left_join(bvq_data$logs) %>% -->
-<!-- #     mutate( -->
-<!-- #         edu_parent = fct_explicit_na(as.factor(edu_parent)), -->
-<!-- #         doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE) -->
-<!-- #     ) %>% -->
-<!-- #     count(doe_catalan, edu_parent) %>% -->
-<!-- #     right_join( -->
-<!-- #         expand_grid( -->
-<!-- #             doe_catalan = levels(.$doe_catalan), -->
-<!-- #             edu_parent = unique(bvq_data$logs$edu_parent) -->
-<!-- #         ) -->
-<!-- #     ) %>% -->
-<!-- #     rename(total = n) %>% -->
-<!-- #     pivot_wider(names_from = edu_parent, -->
-<!-- #                 values_from = total,  -->
-<!-- #                 values_fill = 0) %>% -->
-<!-- #     mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% -->
-<!-- #     rename(Total = total) %>% -->
-<!-- #     select(-"NA") %>%  -->
-<!-- #     gt()  -->
-<!-- # summary_rows( -->
-<!-- #     columns = Complementary:Primary, -->
-<!-- #     fns = list(Total = ~sum(., na.rm = TRUE)), -->
-<!-- #     formatter = fmt_integer -->
-<!-- # ) %>% -->
-<!-- # cols_label(doe_catalan = "DoE Catalan") %>% -->
-<!-- # sub_missing(columns = 2:8, missing_text = "--") %>% -->
-<!-- # tab_spanner("Educational attainment", Secondary:Primary) %>% -->
-<!-- # tab_style( -->
-<!-- #     style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")), -->
-<!-- #     locations = cells_body(columns = 9, rows = 1:5) -->
-<!-- # ) %>% -->
-<!-- # tab_style( -->
-<!-- #     cell_text(style = "italic"), -->
-<!-- #     cells_column_labels(columns = 2:7) -->
-<!-- # ) %>% -->
-<!-- # tab_style( -->
-<!-- #     cell_text(weight = "bold"), -->
-<!-- #     cells_column_labels(columns = c(1, 8)) -->
-<!-- # ) %>% -->
-<!-- # tab_style( -->
-<!-- #     cell_text(weight = "bold"), -->
-<!-- #     cells_column_spanners(spanners = "Educational attainment") -->
-<!-- # ) -->
-<!-- ``` -->
+    rename(total = n) %>%
+    pivot_wider(names_from = edu_parent,
+                values_from = total,
+                values_fill = 0) %>%
+    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>%
+    rename(Total = total) %>%
+    select(-"NA") %>%
+    gt() |> 
+    summary_rows(columns = Complementary:Primary,
+                 fns = list(Total = ~sum(., na.rm = TRUE)),
+                 formatter = fmt_integer) %>%
+    cols_label(doe_catalan = "DoE Catalan") %>%
+    sub_missing(columns = everything(),
+                missing_text = "--") %>%
+    tab_spanner("Educational attainment", 
+                Secondary:Primary) %>%
+    tab_style(
+        style = list(cell_borders(weight = px(2),
+                                  sides = "left", 
+                                  color = "grey50")),
+        locations = cells_body(columns = 9,
+                               rows = 1:5)
+    ) %>%
+    tab_style(
+        cell_text(style = "italic"),
+        cells_column_labels(columns = 2:7)
+    ) %>%
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_labels(columns = c(1, 8))
+    ) %>%
+    tab_style(
+        cell_text(weight = "bold"),
+        cells_column_spanners(spanners = "Educational attainment")
+    )
+```
 
@@
 Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
-# items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))
-# 
-# items %>%
-#     unnest(list) %>%
-#     group_by(language, list) %>%
-#     summarise_at(
-#         vars(freq, n_phon, lv),
-#         list(
-#             n = ~sum(!is.na(.)),
-#             mean = ~mean(., na.rm = TRUE),
-#             sd = ~sd(., na.rm = TRUE),
-#             min = ~min(., na.rm = TRUE),
-#             max = ~max(., na.rm = TRUE)
-#         )
-#     ) %>%
-#     select(-c(n_phon_n, lv_n)) %>%
-#     rename(n = freq_n) %>%
-#     gt(groupname_col = "language", rowname_col = "list") %>%
-#     fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>%
-#     fmt_percent(matches("lv_"), decimals = 1) %>%
-#     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>%
-#     tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>%
-#     tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>%
-#     cols_merge_uncert(col_val = freq_mean, col_uncert = freq_sd) %>%
-#     cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>%
-#     cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>%
-#     cols_merge_range(col_begin = freq_min, col_end = freq_max) %>%
-#     cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>%
-#     cols_merge_range(col_begin = lv_min, col_end = lv_max) %>%
-#     cols_label(
-#         list = "",
-#         n = md("*N*"),
-#         freq_mean = md("Mean ± SD"),
-#         freq_min = md("Range"),
-#         n_phon_mean = md("Mean ± SD"),
-#         n_phon_min = md("Range"),
-#         lv_mean = md("Mean ± SD"),
-#         lv_min = md("Range")
-#     ) %>%
-#     tab_style(
-#         cell_text(style = "italic"),
-#         cells_column_labels(everything())
-#     ) %>%
-#     tab_style(
-#         cell_text(weight = "bold"),
-#         cells_column_spanners(everything())
-#     ) %>%
-#     summary_rows(
-#         fns = list("Grand mean" = ~mean(.)),
-#         columns = ends_with("_mean")
-#     ) %>%
-#     summary_rows(
-#         fns = list("N" = ~sum(.)),
-#         columns = n,
-#         formatter = fmt_integer
-#     )
+items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))
+
+items %>%
+    unnest(list) %>%
+    summarise(n = n(),
+              across(c(freq, n_phon, lv),
+                     lst(mean, sd, min, max)),
+              .by = c(language, list)) |> 
+    gt(groupname_col = "language", 
+       rowname_col = "list") %>%
+    fmt_number(-c(list, n),
+               decimals = 1) %>%
+    tab_spanner("Frequency (Zipf)",
+                columns = starts_with("freq_")) %>%
+    tab_spanner("Length (Phonemes)",
+                columns = starts_with("n_phon_")) %>%
+    tab_spanner("Levenshtein (cognateness)", 
+                columns = starts_with("lv_")) %>%
+    cols_merge_uncert(col_val = freq_mean,
+                      col_uncert = freq_sd) %>%
+    cols_merge_uncert(col_val = n_phon_mean,
+                      col_uncert = n_phon_sd) %>%
+    cols_merge_uncert(col_val = lv_mean, 
+                      col_uncert = lv_sd) %>%
+    cols_merge_range(col_begin = freq_min, 
+                     col_end = freq_max) %>%
+    cols_merge_range(col_begin = n_phon_min, 
+                     col_end = n_phon_max) %>%
+    cols_merge_range(col_begin = lv_min,
+                     col_end = lv_max) %>%
+    cols_label(list = "",
+               n = md("*N*"),
+               freq_mean = md("Mean ± SD"),
+               freq_min = md("Range"),
+               n_phon_mean = md("Mean ± SD"),
+               n_phon_min = md("Range"),
+               lv_mean = md("Mean ± SD"),
+               lv_min = md("Range")) %>%
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels(everything())) %>%
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners(everything())) %>%
+    summary_rows(fns = list("Grand mean" = ~mean(.)),
+                 columns = ends_with("_mean")) %>%
+    summary_rows(fns = list("N" = ~sum(.)),
+                 columns = n,
+                 formatter = fmt_integer)
 ```
@@
 #| warning: false
-#| fig-width: 7
-#| fig-height: 4
+#| fig-width: 6
+#| fig-height: 3
 #| fig-cap: "Distribution of lexical frequency across items.
@@
     ggplot(aes(x = freq)) +
-    geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
+    geom_histogram(fill = "grey60",
+                   colour = "white",
+                   bins = 30)  +
     labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
@@
 
-### \# Phonemes
-
-Word length ($Phonemes$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.
+### Word length
+
+Word length ($Length$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.
 
@@
 items %>%
-    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>%
     count(language, n_phon) %>%
-    ggplot() +
-    aes(
-        x = n_phon,
-        y = n,
-        fill = language,
-        colour = language,
-        shape = language
-    ) +
+    ggplot(aes(x = n_phon, y = n)) +
     facet_wrap(~language, ncol = 1) +
-    geom_col(colour = "white") +
-    geom_text(aes(label = n), position = position_nudge(y = 5), colour = "black") +
-    labs(
-        x = "# Phonemes",
+    geom_col(fill = "grey60", colour = "white") +
+    geom_text(aes(label = n),
+              position = position_nudge(y = 5),
+              colour = "black") +
+    labs(x = "Phonemes",
          y = "Number of words",
@@
          shape = "Language",
-        colour = "Language"
-    ) +
+         colour = "Language") +
     scale_x_continuous(breaks = 1:15) +
     scale_fill_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
+    theme(legend.position = "none",
           panel.grid.major.x = element_blank(),
@@
           axis.ticks.x = element_line(),
-        legend.title = element_blank()
-    )
-```
-
-### Phonological similarity (Levenshtein)
+          legend.title = element_blank())
+```
+
+### Levenshtein (cognateness)
 
@@
     count(lv) %>% 
-    mutate(prop = n/sum(n)) %>% 
+    mutate(prop = n/sum(n),
+           labels = paste0(
+               n, " (",
+               percent(prop, accuracy = 0.1), 
+               ")")) %>% 
     ggplot(aes(x = lv, n)) +
-    geom_col(fill = clrs[4], colour = "white") +
-    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
-              position = position_nudge(y = 3), size = 3.5) +
+    geom_col(fill = "grey60", colour = "white") +
+    geom_text(aes(label = labels),
+              position = position_nudge(y = 3), 
+              size = 3.5) +
     labs(x = "Levenshtein distance", y = "Number of words") +
@@
 
-### Frequency by \# Phonemes
+### Frequency by word length
 
@@
     ggplot(aes(n_phon, freq)) +
-    geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
-               position = position_jitter(width = 0.1), seed = 888) +
-    geom_smooth(method = "lm", formula = "y ~ x" ,
-                colour = clrs[1], fill = clrs[1]) +
+    geom_point(size = 2,
+               alpha = 0.5,
+               colour = "grey60", 
+               shape = 1, 
+               stroke = 1,
+               position = position_jitter(width = 0.1)) +
+    geom_smooth(method = "lm",
+                formula = "y ~ x" ,
+                colour = "black") +
     labs(
@@
     count(freq, lv) %>% 
-    ggplot(aes(freq, lv, size = n)) +
-    geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
-    geom_smooth(method = "lm", formula = "y ~ x", colour = clrs[1], fill = clrs[1]) +
-    labs(
-        y = "Levenshtein similarity", 
+    ggplot(aes(freq, lv)) +
+    geom_point(size = 2,
+               shape = 1, 
+               colour = "grey60",
+               stroke = 1) +
+    geom_smooth(method = "lm", 
+                formula = "y ~ x", 
+                colour ="black") +
+    labs(y = "Levenshtein similarity", 
          x = "Lexical frequency (Zipf score)", 
          colour = "Language", 
-        fill = "Language"
-    ) +
+         fill = "Language") +
     scale_y_continuous(labels = percent) +
@@
 
-### \# Phonemes by Levenshtein
+### Word length by Levenshtein
 
@@
     ggplot(aes(n_phon, lv, size = n)) +
-    geom_point(shape = 1, stroke = 1, colour = clrs[4]) +
-    geom_smooth(method = "lm", formula = "y ~ x", size = 1,
-                colour = clrs[1], fill = clrs[1]) +
-    labs(
-        x = "# Phonemes", 
+    geom_point(shape = 1,
+               stroke = 1,
+               colour = "grey60") +
+    geom_smooth(method = "lm",
+                formula = "y ~ x",
+                size = 1,
+                colour = "black") +
+    labs(x = "# Phonemes", 
          y = "Levenshtein similarity", 
          colour = "Language", 
-        fill = "Language"
-    ) +
+         fill = "Language") +
     scale_y_continuous(labels = percent) +
     scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
+    theme(legend.position = "none",
           panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    ) 
+          panel.grid.major.x = element_blank()) 
 ```
@@
     mutate(
-        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
-        exposure = cut(
-            exposure,
+        exposure = cut(exposure,
             breaks = seq(0, 9, 1),
-            include.lowest = TRUE
-        )
-    ) %>%
+            include.lowest = TRUE)) %>%
     count(exposure, language) %>%
-    ggplot(aes(exposure, n, fill = language)) +
+    ggplot(aes(exposure, n)) +
     facet_wrap( ~ language) +
-    geom_col(colour = "white") +
+    geom_col(colour = "white",
+             fill = "grey60") +
     labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
@@
     responses %>%
-    mutate(
-        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
-        exposure = cut(
-            exposure,
+    mutate(exposure = cut(exposure,
             breaks = seq(0, 9, 1),
-            include.lowest = TRUE
-        ),
-        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
-    ) %>%
+            include.lowest = TRUE),
+        freq = cut(freq, 
+                   breaks = seq(0, 9, 0.25), 
+                   include.lowest = TRUE)) %>%
     count(exposure, freq, language) %>%
-    ggplot(aes(
-        x = freq,
+    ggplot(aes(x = freq,
                y = exposure,
-        colour = language,
-        size = n
-    )) +
+               size = n)) +
     facet_wrap( ~ language) +
-    geom_point() +
-    labs(
-        x = "Frequency (Zipf score)",
-        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
+    geom_point(colour = "grey60",
+               alpha = 1/2) +
+    labs(x = "Frequency (Zipf score)",
+         y = "Exposure-weighted frequency\\
+         (Frequency \u00d7 DoE)",
          colour = "Language",
          fill = "Language",
-        size = "Number of observations"
-    ) +
+         size = "Number of observations") +
     theme(panel.grid = element_blank()) +
@@
     theme(panel.grid.major.x = element_blank(),
-          panel.grid.minor.y = element_blank())
+          panel.grid.minor.y = element_blank(),
+          legend.position = "none")
 ```

### 2023-04-28 manuscript.qmd
Add appendix for syllable frequencies and new revisions

@@
 format:
-  apaquarto-docx: default
+  # apaquarto-docx: default
   apaquarto-pdf:
@@
       - \usepackage{tipa}
-      - \usepackage[pinyin]{babel}
-      - \usepackage{CJKutf8}
-      - \usepackage{fontspec}
-      - \babelprovide[script=CJK, language={Chinese Simplified}]{chinese-simplified}
-      - \babelfont[chinese-simplified]{rm}[Ligatures = Common]{Noto Serif CJK SC}
-      - \babelfont[chinese-simplified]{sf}[Ligatures = Common]{Noto Sans CJK SC}
-
     include-in-header:
@@
     gratitude: |
-        We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, Ege E.
-Özer, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
-We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support.
-We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
+        We are grateful to Chiara Santolin, Ege E.
+Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
+We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support, and Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
 We also thank all families and infants who participated in the experiments.
@@
 color_scheme_set(rep(clrs[1], 6))
+
 options(ggplot2.ordinal.fill = clrs,
@@
 The foundations of word learning are in place early in age.
-From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [@fenson2007macarthurbates; @samuelson2021precision].
-During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
-Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e., vocabulary size, @fenson1994variability; @frank2021variability].
+From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthurbates; @samuelson2021precision].
+During the last half of the second year of life, new word-forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
+Such early stages of lexical acquisition are characterised by substantial heterogeneity across different populations of language learners, reflected for instance on (the variability of) quantitative indices of lexical development, like vocabulary size [i.e., vocabulary size, @fenson1994variability; @frank2021variability].
 Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
@@
 
-Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to substantial proportion of the world population acquiring more than one language since early ages [@grosjean2021extent].
+Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
 Previous research on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
@@
 For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language.
+These cross-language word pairs that share meaning are referred to as *translation equivalents*.
+Throughout the present study, we make the assumption that bilinguals form common semantic representations for both languages (e.g., the concept of a *chair* is shared for the English and Spanish words 'chair' and 'silla')^[The referential contexts in which bilingual children hear (and learn) both labels may not be identical.
+For instance, the English word *finger* refers to eight parts of the body (excluding thumbs and toes), whereas its Spanish translation *dedos* refers to twenty body parts (fingers, thumbs, and toes).
+These discrepancies are reflected, for instance, in adults' performance in lexical decision tasks [@chaouch-orozco2023are].
+The fact that some translation pairs share more semantic features than others does not, however, inform whether bilinguals form common or separate semantic representations for translation equivalents.
+What English speakers understand for 'fingers' might not be identical to what Spanish speakers understand for 'dedos'.
+But when an English speaker learns Spanish, it seems implausible that they would form a separate semantic representation for what 'dedo' conveys at the conceptual level, compared to the possibility that they associate the new label to the semantic representation they already had formed for 'fingers'.
+Therefore, the lack of complete semantic alignment is not necessarily an argument against the existence of common semantic representations for translation equivalents.
+For simplicity, and in line with previous (formal and verbal) models of bilingual lexical processing [e.g., @shook2013bilingual; @dijkstra2019multilink; @dijkstra1998bia; @kroll2010revised], we assume common semantic representations for translation pairs.].
 Depending on the similarity between both languages, the two sets of words may overlap in form in varying degrees.
@@
 
-@floccia2018introduction provided evidence in line with this claim.
+@floccia2018introduction provided evidence for such a facilitation effect of typological distance and vocabulary size.
 The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
@@
 The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /kæt/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/.
-
 Levenshtein distance scores are divided by the length of the longest string to normalise the similarity scores between 0 and 1.
-Finally, the result is subtracted from one, to get the complementary, and therefore a measure of phonological similarity, instad of dissimilarity.
+Finally, the result is subtracted from one, to get the complementary, and therefore a measure of phonological similarity, instead of dissimilarity.
 Following the previous examples: $1/4 = 0.25$ for /kæt/-/\textipa{kAt}/, and $3/4 = 0.75$ for /kæt/-/\textipa{Gata}/.
@@
 For example, English-Dutch bilinguals (0.2214 mean phonological overlap), were able to produce more Dutch words than English-Mandarin bilinguals (0.0197 mean phonological overlap) were able to produce in Mandarin.
+More recently, @blom2020crosslanguage and @bosma2019longitudinal reported similar results, providing converging evidence of for facilitatory effect of language distance on vocabulary size estimates.
 
@@
 pointed to *parallel activation* as the underpinning mechanism behind their results.
-The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain], even in fully monolingual environments.
+The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during speech production or comprehension, even in fully monolingual environments.
+Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extend to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @hoshino2008cognate; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 
-Regarding word acquisition, the parallel activation account is a plausible explanation for Floccia et al.'s results in the light of two observations.
-First, bilingual toddlers' lexicon seems to consist of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage].
-Second, there is experimental evidence of parallel activation during word comprehension in children and toddlers [@vonholzen2012language; @poulin-dubois2013lexical; @bosma2020cognate].
-It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition.
-This could ultimately lead to children who are learning languages with a larger proportion of cognates (e.g., English and Dutch) showing larger vocabulary sizes.
-This account in which cognateness underlies the facilitation effect of typological language distance is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
-If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity (e.g., are taxonomically related) but not phonological similarity.
+Bilingual toddlers and children---still in the process to developing a mature lexicon---also show evidence of parallel activation during comprehension [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic; @poulin-dubois2013lexical].
+Whether---and if so, how---the language non-selectivity shapes the early bilingual lexicon is still unclear.
+Previous work has pointed to an interactive account of lexical acquisition, in which toddlers' lexical representations in one language influence the acquisition of words in the other language.
+This cross-language facilitation is, in principle, possible since early ages: the bilingual infant lexicon consists in a substantial proportion of  translation equivalents: the presence of a word in the lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage; @pearson1995crosslanguage].
+In fact, translation equivalents may even be overrepresented in the early bilingual lexicon.
+@bilson2015semantic found that bilingual toddlers produced a higher proportion of translation equivalents than it would be expected if they learned two independent lexica simultaneously, which suggests that bilingual children preferentially acquired words for which they already knew a label in the other language.
+In the same line, @bosch2014first reported a similar overrepresentation of translation equivalents in Catalan-Spanish bilinguals: a high proportion of acquired Catalan words held an also acquired translation in Spanish.
+Finally, a recent investigation by @tsui2022area presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
+Overall, these findings suggest that bilingual children benefit from a semantic facilitation effect driven by the acquisition of translation equivalents.
+
+It is possible that cognateness increases the amount of cross-language activation, facilitating early word acquisition.
+This could ultimately lead to children who are learning languages with a larger proportion of cognates (e.g., English and Dutch) showing larger vocabulary sizes, as found by Floccia et al.
+There is evidence of for phonological and semantic similarity facilitating word acquisition in monolinguals: infants are more likely to acquired new words that are semantically or phonologically more connected to other words, than words with poorer connectivity in either of those dimensions, such as words belonging to sparse phonological neighbourhoods [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children; @storkel2004children].
+If phonological similarity plays a cross-language facilitation role during bilingual word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity (e.g., are taxonomically related) but not phonological similarity.
 
@@
 Parallel activation remains a plausible explanation for this effect.
-When a child is exposed to a word form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition.
+When a child is exposed to a word-form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition.
 The precise underpinnings of this effect are still unclear.
@@
 In line with this account, only the later-acquired form of the translation equivalent would benefit from its cognate status.
-Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual; @hoff2012dual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
 For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age of acquisition for cognates than for non-cognates.
@@
 Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees.
-We adopted a Bayesian explanatory item response theory  [IRT, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
+We adopted a Bayesian explanatory item response theory  [IRT, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word-form, while adjusting for the the number of phonemes in the word-form (an indicator of word acquisition difficulty).
 
@@
 For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]` and *pájaro* `[paxa4o]`, Catalan and Spanish for 'bird', and returns 50% for *lluna* `[Lun5]` and *luna* `[luna]`, Catalan and Spanish for 'moon' (see Appendix A for more details).
+
+Finally, we included word length as a nuance predictor in our analyses, motivated by previous studies reporting a strong association between word length and vocabulary size estimates [e.g., @braginsky2019consistency; @jones2019children].
+Sorter word-forms like /\textipa{sO\textltilde}/ (Catalan for *sun*) are generally simpler to encode as phonological representations in the lexicon than longer word-forms like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*), and are therefore more likely to be participant a child's vocabulary.
 
@@
     mutate(across(.median:.lower,
-                  \(x) ifelse(grepl("Intercept", .type), plogis(x), x/4))) |> 
+                  \(x) ifelse(grepl("Intercept", .type),
+                              plogis(x), x/4))) |> 
     split(posterior_draws$summary$.variable)
@@
         model = "Model",
-        elpd_loo = "ELPD $$\\pm$$ SE",
-        elpd_diff = "Diff. $$\\pm$$ SE"
+        elpd_loo = "ELPD (SE)",
+        elpd_diff = "ELPD difference (SE)"
     ) |>
@@
        rowname_col = ".variable_name") |>
-    fmt_percent(.rope_overlap) |> 
-    fmt_number(c(.median:.upper)) |>
+    fmt_percent(.rope_overlap, decimals = 3) |> 
+    fmt_number(c(.median:.upper), decimals = 3) |>
     cols_merge(c(.lower, .upper),
@@
 #| fig-height: 6
+#| out-width: 80%
 #| eval: true
@@
     stat_summary(fun = median, geom = "line", linewidth = 0.75) +
-    labs(
-        x = "Age (months)",
+    labs(x = "Age (months)",
          y = "Probability of acquisition",
@@
          fill = "Cognateness (phonological similarity)") +
+    guides(colour = guide_legend(override.aes = list(linewidth = 2))) +
     scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
@@
                                       colour = "black", 
-                                      linewidth = 0.75),
-          plot.subtitle = element_textbox_simple(fill = "grey90",
-                                                 padding = margin(8, 8, 8, 8),
-                                                 margin = margin(0, 0, 10, 0))) +
-    guides(colour = guide_legend(override.aes = list(linewidth = 2)))
+                                      linewidth = 0.75))
 ```
 
+To rule out the possibility that cognateness facilitation effect we found was due to cognateness comprising more frequent syllables than non-cognates---and therefore not because of their cognate status itself---, we compared the syllabic frequency of cognates and non-cognates included in our data analyses.
+To calculate syllable frequency, we first extracted all syllables embedded in the selected words.
+For each syllable, we summed the lexical frequency of all the words in which such syllable appeared.
+The resulting value provides an estimate of the number of times the syllable appears in child-directed speech, embedded within different words.
+Finally, for each word, we summed the frequency of its syllables, as an estimate of the syllabic frequency of the word.
+We fit a Bayesian model with cognateness (Levenshtein similarity) as response variable, and the main effects of syllable frequency and number of syllables (to control for the fact words with more syllables are more likely to score higher in syllabic frequency) as predictors.
+This model provided strong evidence for an association between cognateness and syllabic frequency being equivalent to zero (see Appendix E).
+
 # Discussion {#sec-discussion}
-
-The present study explored the cognateness facilitation effect on bilingual word acquisition.
-We found such effect to be mediated by lexical frequency and degree of language exposure: phonologically similar translation equivalents were on average acquired earlier in age compared to phonologically dissimilar translation equivalents, but this effect was only observed for word-forms to which participants were exposed less often, i.e., words with low lexical frequency and/or words from the language of lower daily exposure.
-Our results are relevant in the light of two previous findings.
-First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (which shared many cognates, e.g., English-German) showed larger vocabulary sizes than those learning typologically more distance languages (which shared fewer cognates, e.g.,  English-Mandarin).
-Second, @mitchell2022cognates found an earlier age-of-acquisition for cognates, compared to non-cognates.
-The outcomes of both studies pointed to cognateness facilitating word acquisition, but the underpinnings of such effect were unclear.
-
-In this study, we capitalised on accumulator models of language acquisition to propose a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
-We suggested that words benefit from their cognate status only *after* their translation equivalent has been acquired: after the child acquires a word-form, the number of learning instances they need to acquire its translation decreases.
-We further hypothesised that such facilitation is proportional to the phonological similarity between both word-forms.
-If two translation equivalents sharing a fair deal of phonological similarity (e.g., /\textipa{"pOR.t@}/-/\textipa{"pweR.ta}/, `r percent(stringdist::stringsim("pORt@", "pweRta"))` Levenshtein similarity), learning of one of them should facilitate the acquisition of the other more strongly, compared to when both word-forms share only few phonemes (e.g., /\textipa{k@"Di.R@}/-/\textipa{"si.\textlambda a}/ , `r percent(stringdist::stringsim("DiR@", "siLa"))`).
-
-Under such scenario, we predicted that words from the language of lower exposure would benefit more often from their cognate status than those from the language of higher exposure.
-The rationale behind such prediction was that children are less frequently exposed to word-forms in the former language than to words in the latter [@cattani2014how; @thordardottir2006bilingual].
-Words in the language of lower exposure are therefore more likely to be acquired later in age than their translations in the other language.
-For instance, an infant exposed to Catalan and Spanish 65% and 35% of the time, respectively, is more likely to learn the Spanish word-form /\textipa{"ga.to}/ later in age than its Catalan translation /\textipa{gat}/.
-If this is the case, /\textipa{"ga.to}/ is more likely to receive facilitation from the acquisition of /\textipa{gat}/ than *vice versa* (see {apafg-hypotheses} for a graphic illustration).
-In line with these predictions, we found that cognateness facilitated word acquisition more strongly in words from the language of lower exposure.
-This asymmetric facilitation effect provides support for the hypothesis that words benefit from their cognate status *after* their translation has been acquired.
-Overall, our results provide support for an interactive account of bilingual vocabulary learning in which word representations from one language contribute to the acquisition of words in the other.
-
-These findings are, to a large extent, in line with previous literature on word acquisition.
-First, words-forms with fewer phonemes like /\textipa{sO\textltilde}/ (Catalan for *sun*) were on average more likely to be acquired than word-forms with more phonemes like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*).
-These results converge with previous studies found word length to be more strongly associated with receptive vocabulary than with expressive vocabulary [e.g., @braginsky2019consistency; @jones2019children].
-This facilitation is most likely due to the fact that shorter word-forms are generally simpler to encode as phonological representations in the lexicon than longer word-forms.
-Shorter word-forms are easier to produce, and therefore to be reported as acquired in parental vocabulary checklists.
-
-Our participants were also more likely to know words to which they were more frequently exposed to, replicating previously reported effects of lexical frequency [e.g., @braginsky2019consistency], and bilingual language exposure [e.g., @cattani2014how; @thordardottir2006bilingual].
-This strong effect of word frequency is a key component of accumulator models of word acquisition, which suggest that the age-of-acquisition of a word is, to a large extent, a function of the child's rate of exposure to such word, which conspires with child- and word-level factors in the process of forming a representation of the word in the child's lexicon [@mollica2017how; @hidaka2013computational].
-Our results provide ground for the central role of word frequency in accumulator models to be extended to the bilingual case, in the form of a composite measure of word rate of exposure that accounted for both lexical frequency and language exposure [@cattani2014how; @floccia2018results].
-Interestingly, our findings pointed to this predictor playing a stronger role during earlier ages: younger infants benefited from higher exposure rates at earlier ages, suggesting that early word acquisition is supported by mere exposure to word forms more strongly than later word acquisition.
-
-Finally, we found a frequency-mediated facilitation effect of cognateness, adding some nuance to the findings from previous studies [@floccia2018results; @mitchell2022cognates; @bosch2014first; @schelletter2002effect].
-Only words to which children were exposed to less frequently benefited from their cognate status.
-These results favour the account that a word-form benefits from its phonological similarity with its translation only after such translation has been acquired first.
-On the one hand, our results provide converging evidence for @floccia2018results, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample.
-Most children in their sample were English-dominant, and therefore their exposure to English was higher than their exposure to the additional language.
-However, it was in the vocabulary of the additional language where the authors found a facilitation effect of cognateness: bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese).
-Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018results's sample.
-
-Overall, our findings parallel cognates' significant role in many models of bilingual language processing.
-The importance of cognates in these model stems in the fact that cognates provide evidence for a language non-selective account of lexical processing, in which bilinguals activate unselected lexical representations in both languages, even in fully monolingual situations [@thierry2007brain; @christoffels2007bilingual].
-As mentioned in the introduction, naming pictures whose labels are cognates leads to shorter latencies than when labels are non-cognates, which suggests that phonological similarity across translations facilitates production [@costa2000cognate; @hoshino2008cognate].
-When acting as distractors in naming tasks, cognates exert larger interference effects than non-cognate distractors [@hoshino2011language; @colome2010which].
-Both findings suggest that the non-selected translation of the target in @costa2000cognate, or of the distractor in @colome2010which was activated, and that when such word-form was phonologically related to its translation, the facilitation or interference effect was stronger.
-
-Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extend to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
-For instance, @thierry2007brain found neurophysiological evidence of Chinese-English bilinguals activating both languages in a word comprehension task.
-In this experiment, participants were presented with couples of words in English, and were asked to respond if the words were related in meaning.
-Unbeknownst to participants, the translations to Chinese of half of such word pairs shared a character.
-Participants would be presented (visually in a first experiment and auditorily in a second experiment) with an English word pair 'train-ham', which are semantically unrelated, but share a character in their Chinese translations (\foreignlanguage{chinese-simplified}{火車} [Huo Che]-\foreignlanguage{chinese-simplified}{火腿} [Huo Tui]).
-Bilinguals' brain signal differed across words pairs with and without translation overlap (e.g., 'apple-table', \foreignlanguage{chinese-simplified}{平果} [Ping Guo]–\foreignlanguage{chinese-simplified}{桌子} [Zhuo Zi]).
-This suggests that participants were sensitive to the form-similarity of the Chinese translation of the word-pairs they were presented, despite the whole experimental task taking place in English.
-
-Bilingual toddlers and children---still in the process to developing a mature lexicon---also show evidence of parallel activation during comprehension [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic].
-Whether---and if so, how---the language non-selectivity shapes the early bilingual lexicon is still unclear.
-Previous work has pointed to an interactive account of lexical acquisition, in which toddlers' lexical representations in one language influence the acquisition of words in the other language.
-The acquisition of translation equivalents embodies a notable divergence between monolingual and bilingual developmental trajectories.
-For instance, monolingual word acquisition benefits from the semantic relatedness between the words the child learns [@fourtassi2020growth; @hills2009longitudinal], but is inhibited during the acquisition of words from the same language with the *equivalent* meaning, i.e.
-synonyms.
-This learning heuristic, called *mutual exclusivity*, is one of many that play a relevant role in models of monolingual word learning [@golinkoff1994early], acting as a constrainer of the hypothesis space for all possible word-referent associations: if a referent already holds a label, any label uttered in the presence of such referent is associated with a different referent [e.g., @markman1990constraints].
-Although this heuristic seems useful in the monolingual case (children are rarely exposed to synonyms), it would come with a huge cost for bilinguals, how are frequently exposed to more than one label for the same referent, one in each language.
-It was initially predicted that bilinguals would follow *mutual exclusivity*, and therefore avoid learning translation equivalents [@volterra1978acquisition].
-
-This prediction has been robustly contested by two main sources of evidence.
-First, experimental work suggests that bilinguals may not apply *mutual exclusivity* during word learning, or at least not as consistently monolinguals do [@byers-heinlein2009monolingual; @byers-heinlein2013lexicon].
-Second, detailed analyses of bilingual children's vocabulary contents shows that bilingual children do acquire translation equivalents from early ages [e.g., @pearson1995crosslanguage].
-In fact, translation equivalents may even be overrepresented in the early bilingual lexicon.
-For instance, @bilson2015semantic found that bilingual toddlers produced a higher proportion of translation equivalents than it would be expected if they learned two independent lexica simultaneously, which suggests that bilingual children preferentially acquired words for which they already knew a label in the other language.
-In the same line, @bosch2014first reported a similar overrepresentation of translation equivalents in Catalan-Spanish bilinguals: a high proportion of acquired Catalan words held an also acquired translation in Spanish.
-Finally, a recent investigation by @tsui2022area presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
-Overall, these findings suggest that bilingual children benefit from a semantic facilitation effect driven by the acquisition of translation equivalents.
-This is a substantial divergence from monolingual trajectories of lexical acquisition.
-Our study provides further evidence for a pivotal role of translation equivalents, showing that, especially in the case of cognates, translation equivalents, are not only part of bilinguals' vocabulary contents, but are also acquired close in age, depending primarily on children's degree of exposure to each language.
-
-The divergent trajectories of bilingual lexical acquisition go beyond semantics.
-As highlighted before, the phonological overlap between the languages being learned seems to impact vocabulary growth in bilinguals.
-Bilinguals learning languages that share many cognates show larger vocabulary sizes than those learning two languages sharing fewer cognates [@floccia2018introduction; @blom2020crosslanguage; @bosma2019longitudinal].
-This finding can be explained to a large extent by the fact that cognates seem to be acquired at earlier ages than non-cognates [@mitchell2022cognates].
-This phonological facilitation effect mimics previous findings in monolinguals, who seem to preferentially learn phonologically similar words.
-For instance, @jones2019children found that English monolingual toddlers show a facilitation effect when acquiring words from dense phonological neighbourhoods: a larger number of similar-sounding words was associated with an increase in likelihood of a word form being part of a toddler's productive vocabulary.
-A network analysis by @fourtassi2020growth found evidence in favour of a phonological (and semantic) preferential acquisition of new words in monolingual children: words with higher phonological relatedness to other words were more likely to be acquired.
-Again, these findings contrast with those in bilingual children in that the facilitation effect or preferential acquisition mechanisms described are based on phonological similarity *without* translation equivalence.
-The amount of similar-sounding words in the child input or their acquired vocabulary predicts a larger probability of acquisition of new words with *different* meaning.
-In the case of bilinguals, there is evidence of such phonological facilitation at the level of translation equivalents, as shown by an earlier acquisition of cognates than non-cognates in English-French [@mitchell2022cognates] and in the Catalan-Spanish bilinguals in the present study.
-Evidence of a cross-language facilitation effect between non-translation equivalents is to date absent.
-
-We consider three alternative explanations for our results.
-Regarding the cognate facilitation effect, it is possible that when hearing cognate translation equivalents, bilingual children accept both word forms as acceptable variations of the *same* word.
-This contrasts with monolinguals' high sensitivity to mispronunciations during word recognition [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological].
-It might be the case that every exposure to either of the word-forms of the translation equivalent is counted as a learning instance for both.
-This would impact the acquisition mechanism in such way that any cognate effect is not explained by a reduction of the number of learning instances needed to acquired a word (after its translation has been acquired), but by a faster increase in cumulative learning instances across ages, leading to a similar outcome: an earlier age-of-acquisition for cognates.
-Supporting such alternative explanation, @ramon-casas2009vowel examined 18-26 months-old Catalan-Spanish bilingual toddlers' sensitivity to mispronunciations in a word-recognition task and found that, contrary to their Catalan monolingual peers, bilinguals seemed to recognise mispronounced words, disregarding vowel replacements involving the Catalan vowel contrast /e/-/\textepsilon/.
-The same population of bilingual infants shows sensitivity to such contrast before 12 months of age in phoneme discrimination tasks [@bosch2003simultaneous]---which do not involve word-referent associations---suggesting that bilingual infants dismiss subtle changes in phonological forms when presented with their referents.
-Using the same procedure, @ramon2010non tested bilingual infants' recognition of mispronunciations involving the /e/-/\textepsilon/ contrast embedded in exclusively *non-cognate* word-forms.
-The materials in this study contrasted with those in @ramon-casas2009vowel in that the latter included almost 60% of cognate words in their stimuli list.
-@ramon2010non hypothesised that bilinguals' frequent exposure to cognates---which often involve vowel replacements between Catalan and Spanish word-forms---may have led bilinguals to be less alert to such changes when occurring within similar word-forms.
-With this new set of stimuli, including non-cognate words only, bilinguals showed monolingual-level sensitivity to mispronunciations involving the /e/-/\textepsilon/ contrast.
-Overall, these results suggest, at face value, that that bilinguals may to some degree encode similar-sounding word-forms with shared referent as into a common lexical representation, especially when the word-forms involved are cognates.
 
@@
     mutate(prop = n/n_total)
-```
 
-```{r discussion-pool}
-#| echo: false
-#| warning: false
-#| message: false
 E_e <- bvq_data$pool |>
@@
 
-We identify three main reasons why bilingual children are unlikely to conflate similar sounding translations into the same lexical representations.
-First, when @ramon-casas2009vowel probed Catalan-Spanish bilinguals using the /e/-/i/ contrast---present in both Catalan *and* Spanish---in the same task, bilinguals showed monolingual-level sensitivity to mispronunciations, suggesting that the conflation of variations of the same word-form into a similar lexical representation may be specific to some language-specific contrasts, and perhaps associated with the acoustic distance between the phonemes involved.
-The /e/-/\textepsilon/ contrast in which Catalan-Spanish bilinguals failed to mispronunciations, is particularly problematic this regard: both vowels differ in only one acoustic feature (/e/ is closed-mid, while /\textepsilon/ is open-mid) and is only present in Catalan.
-Our results are unlikely to be explained by the possibility that Catalan-Spanish bilingual toddlers form common lexical representations for cognate words.
-Second, few phoneme substitutions that separate Catalan word-forms from their cognate Spanish translations involve the /e/-/\textepsilon/ contrast.
-Rather, due to vowel reduction, Spanish and Catalan cognates predominantly involve the /a/-/\textschwa/, /e/-/\textschwa/, /\textepsilon/ and /o/-/u/ contrasts [@vigario2009phonetics].
-Only `r percent(nrow(E_e)/length(unique(items$te)))` translation equivalents in our questionnaire involved a /e/-/\textepsilon/ change between the Catalan and the Spanish items, only one of such was a minimal pair: /\textipa{tREn}/-/\textipa{tRen}/ [*train*].
+The present study explored the cognateness facilitation effect on bilingual word acquisition.
+We found such effect to be mediated by lexical frequency and degree of language exposure: phonologically similar translation equivalents were on average acquired earlier in age compared to phonologically dissimilar translation equivalents, but this effect was only observed for word-forms to which participants were exposed less often, i.e., words with low lexical frequency and/or words from the language of lower daily exposure.
+In addition, we provided convergent evidence of a facilitation effect of word length and lexical frequency on word acquisition.
+Our results are relevant in the light of two previous findings.
+First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (which shared many cognates, e.g., English-German) showed larger vocabulary sizes than those learning typologically more distance languages (which shared fewer cognates, e.g.,  English-Mandarin).
+Second, @mitchell2022cognates found an earlier age-of-acquisition for cognates, compared to non-cognates.
+The outcomes of both studies pointed to cognateness facilitating word acquisition, but the underpinnings of such effect were unclear.
 
-Second, few words in our study differ only in one phoneme substitution.
-Only `r percent(lv_distr$prop[nrow(lv_distr)], accuracy = 0.01)` of the translation equivalents included in our analyses shared between 90 and 100% phonological similarity (measured as the Levenshtein similarity between the X-SAMPA transcriptions of the Catalan and Spanish form), and only `r percent(lv_distr$prop[nrow(lv_distr)-1], accuracy = 0.01)` shared between 80 and 90% similarity.
+In this study, we capitalised on accumulator models of language acquisition to put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
+We hypothesised that the acquisition of a word benefits from cognateness only *after* its translation equivalent has been acquired.
+We suggested that this effect results from a reduction of the number of learning instances needed for the word to be acquired.
+We further hypothesised that such facilitation is proportional to the phonological similarity between both word-forms.
+If two translation equivalents sharing a fair deal of phonological similarity (e.g., /\textipa{"pOR.t@}/-/\textipa{"pweR.ta}/, `r percent(stringdist::stringsim("pORt@", "pweRta"))` Levenshtein similarity), learning of one of them should facilitate the acquisition of the other more strongly, compared to when both word-forms share only few phonemes (e.g., /\textipa{k@"Di.R@}/-/\textipa{"si.\textlambda a}/ , `r percent(stringdist::stringsim("DiR@", "siLa"))`).
+
+Under such scenario, we predicted that words from the language of lower exposure would benefit more often from their cognate status than those from the language of higher exposure.
+The rationale behind such prediction was that children are less frequently exposed to word-forms in the former language than to words in the latter [@cattani2014how; @thordardottir2006bilingual].
+Words in the language of lower exposure should therefore be more likely to be acquired later in age than their translations in the other language.
+For instance, an infant exposed to Catalan and Spanish 65% and 35% of the time, respectively, should be more likely to learn the Spanish word-form /\textipa{"ga.to}/ later in age than its Catalan translation /\textipa{gat}/.
+If this was the case, /\textipa{"ga.to}/ should be more likely to receive facilitation from the acquisition of /\textipa{gat}/ than *vice versa* (see {apafg-hypotheses} for a graphic illustration).
+In line with these predictions, we found that cognateness facilitated word acquisition more strongly in words from the language of lower exposure.
+Only words to which children were exposed to less frequently benefited from their cognate status.
+This asymmetric facilitation effect provides support for the hypothesis that words benefit from their cognate status *after* their translation has been acquired, and more generally, for an interactive account of bilingual vocabulary learning in which word representations from one language contribute to the acquisition of words in the other.
+
+Overall, our results add further nuance to the findings from previous studies reporting a cognateness facilitation effect, and provide new insights into the mechanisms underpinning such effect.
+In particular, we argue that the account we provide in the present study is a plausible explanation for the language distance facilitation effect reported by Floccia et al.
+In their study, bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional (non-English) language than those bilinguals who were learning two languages that shared less cognates (e.g., English-Chinese).
+Critically, the additional language was consistently also the language of lower exposure for most children, while English was the language of higher exposure.
+We argue that this asymmetric facilitation may be the result of the same mechanism that we hypothesise to have generated the asymmetric facilitation observed in the present study: cognateness facilitates the acquisition of a word (and ultimately, vocabulary growth) *only after* its translation has been acquired.
+Given that words in the language of higher exposure are more likely to be acquired first, words in the language of lower exposure (especially those of lower lexical frequency) may be subject of the cognateness facilitation effect more often.
+
+Lexical frequency plays a central role in this model of word bilingual word acquisition, given that together with age and language exposure, it accounts for the number of times a child is exposed to some word-form.
+Consistently, our composite measure of lexical frequency (which accounts for both lexical frequency and language exposure) was the second strongest predictor of word acquisition, only after age.
+These results parallel previous work on language acquisition, in which lexical frequency effects are pervasive and play a capital role in many models of lexical acquisition [@ambridge2015ubiquity].
+There is robust evidence that, when keeping relevant variables constant, high-frequency words are learnt earlier in age than low-frequency words [e.g., @goodman2008does; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
+The outputs of our regression models parallel these findings.
+Given two words with identical scores in all the predictors we considered (e.g., word length), the with higher a higher score in the composite measure of lexical frequency (which corrects for dual language exposure) is predicted a higher probability of acquisition.
+Interestingly, the interaction between this predictor and participants' age was substantial, suggesting that this predictor plays a stronger role at later ages, with older infants benefiting more strongly from lexical frequency than younger infants.
+This points  to early word acquisition being supported by mere exposure to word-forms more strongly than later word acquisition.
+@braginsky2019consistency, found a similar interaction in their monolingual sample: consistently, infants learning one of a wide range of languages showed larger lexical frequency effects at earlier ages.
+Whether this effect results from an extant developmental process (and if so, what mechanisms underlie it), or from methodological artefacts (e.g., frequency measures being more representative of a child's linguistic input at older ages) is a matter that should be addressed in future work, in order to further explore the interaction of this predictor and the effect of cognateness in word acquisition.
+
+By explaining the cognate facilitation effect in terms of parallel activation, we work under the assumption that, when learning new words, bilinguals form two separate lexical representations for each translation equivalent---one in each language.
+Without this assumption, one might consider the possibility that bilingual children merge cognate translation equivalents into a shared lexical representation.
+Because bilingual children are exposed to such similar sounding word-forms in equivalent referential contexts (e.g., hearing *gat* and *gato* in the presence of a cat), it is possible that they classify both are as acceptable variations of the *same* word, therefore treating them as a single lexical item.
+This could explain a cognate advantage in word acquisition without the need for parallel activation playing a relevant role.
+This account would not necessarily imply that bilinguals are insensitive to the phonological mismatches between cognate word-forms.
+There is strong evidence that both monolingual and bilingual infants are highly sensitive to subtle mispronunciations of word-forms during word recognition [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels].
+Rather, the assimilation of both word-forms in the lexicon might occur at a higher level, where both word-forms are considered *referentially* equivalent, so that every exposure to either of the (distinguishable) word-forms of the translation equivalent is counted as a learning instance for both.
+This would impact the acquisition mechanism in such way that any cognate effect is not explained by a reduction of the number of learning instances needed to acquired a word (through parallel activation, after its translation has been acquired), but by a faster increase in cumulative learning instances across ages, leading to a similar outcome: an earlier age-of-acquisition for cognates.
+
+Previous work points in a different direction, with bilingual toddlers considering cognates as separate words.
+For instance, @ramon-casas2009vowel showed that 18-26 months-old bilingual toddlers reject mispronounced labels in word recognition tasks, although only when such mispronunciation involves a contrast that is phonemic in both languages.
+In their study, Catalan-Spanish bilinguals treated the words /pe\textesh/ and its mispronunciated form /p\textepsilon\textesh/, which involved the /e-\textepsilon/ contrast, only present in Catalan as different, but not the words /pe\textesh/ and /pi\textesh/, which involved the the /e-i/ contrast, present in both languages.
+Moreover, a later study by @ramon2010non showed that bilinguals treat the language-specific /e-\textepsilon/ contrast differently when embedded within *non-cognate* word-forms.
+These results suggest that, except for acoustically close, language-specific /e-\textepsilon/ contrast, bilinguals treat similar-sounding words differently in word recognition tasks, especially when both word-forms are separated by more than one phoneme.
+
+In the light of these findings, the possibility that bilinguals keep similar-sounding translation equivalents as separate lexical representations becomes more plausible.
+This is especially true for the case of the Catalan-Spanish bilinguals in our sample: there are few exemplars of Catalan-Spanish translation equivalents involving a /e-\textepsilon/ substitution.
+Only `r percent(nrow(E_e)/length(unique(items$te)))` translation equivalents in our questionnaire involved a /e-\textepsilon/ change between the Catalan and the Spanish items, only one of such was a minimal pair: /\textipa{tREN}/-/\textipa{tReN}/ [*train*].
+Moreover, few words in our study differ only in one phoneme substitution.
+Only `r percent(lv_distr$prop[nrow(lv_distr)], accuracy = 0.01)` of the translation equivalents included in our analyses shared between 90% and 100% phonological similarity (measured as the Levenshtein similarity between the X-SAMPA transcriptions of the Catalan and Spanish form), and only `r percent(lv_distr$prop[nrow(lv_distr)-1], accuracy = 0.01)` shared between 80 and 90% similarity.
 This shows that the presence of identical and near-identical cognates in our study is very low, and therefore most translation equivalents differed in more than one phoneme.
-In summary, even if any cognate involved the /a/-/\textepsilon/ substitution (or any phonetically close contrast that bilingual might disregard at the lexical level), such is very likely to have occurred in the context of a sufficiently distinct word-form to be distinguished and treated and relevant at the lexical level.
-Third, and perhaps more critically, bilinguals are able to learn new minimal pairs involving the /e/-/\textepsilon/ contrast, suggesting that their tolerance towards cognate mispronunciations might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@ramon-casas2017minimalpair].
+Finally, and perhaps more critically, there is evidence beyond word recognition that bilinguals treat minimal pairs differently at the referential level.
+For instance, @ramon-casas2017minimalpair found that bilingual toddlers are able to learn new minimal pairs involving the /e-\textepsilon/ contrast.This suggests that toddlers' tolerance towards cognate mispronunciations at the referential level might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@mattock2010first; @havy2016phonetic].
 
-It cannot be neglected a potential bias by caregivers when filling the questionnaire, in which the likelihood of two translation equivalents being present in a child's lexicon results not as a function of such phonological facilitation effect, but by caregivers assuming that because the child knows a word in one language, the child must also know the word in the other language.
-This bias might be especially likely when both words are similar-sounding (i.e., cognates).
-Although there is little possibility to test this account, it is worth noting that parents were explicitly instructed *not* to rely on their responses to the words in one languages when responding to words in the other language.
-The potential presence of responder biases is pervasive across many measurement instruments that rely on caregivers' reports of the developmental status of the child.
-The outcomes of such instruments, like the one used in the present study, are susceptible to biases induced by the child themselves (i.e., assigned gender, temperament), their caregivers (e.g., time spent with the child, preconceptions about language acquisition), or even the researchers (e.g., the items included might over-rely on canonical realisations of words), which push against evidence of validity of the questionnaire as an instrument of the developmental status of the child.
-The findings in this investigation should be weighted according to the aforementioned limitations.
+Our study shares similar methodological limitations with previous work using parental vocabulary reports provided.
+Such reports can be subject to measurement error induced by caregivers, who may sometimes overestimate or underestimate participants' true probability of acquisition of words.
+For instance, although in the present study caregivers were explicitly instructed *not* to rely on their responses to the words in one languages when responding to words in the other language, it is a possibility that caregivers were biased to assume---at least to some extent---that because the child knows a word in one language, the child must also know the word in the other language (especially when both words are similar-sounding, i.e., cognates).
+Production estimates may be more prompt to such biases, in part because of the slower pace at which their articulatory abilities develop, compared to their word recognition abilities.
+For instance, even at 47 months of age few (typically developing) English monolinguals reach 100% intelligibility scores in word production, as rated by English native adults [@hustad2020development; @hustad2021speech].
+This gap between comprehension and production is even larger in the less dominant language of bilingual children [@giguere2022bilingual].
+For this reason, caregivers may be more uncertain about what words can be counted as "acquired" in this modality.
+Despite these potential biases, vocabulary checklist filled by parents show strong evidence of concurrent validity with other estimates of vocabulary size or lexical processing [@killing2008move; @feldman2005concurrent; @gillen2021tapping; but see @houston-price2007discrepancy].
 
-Our results should also be taken with caution if generalised to other bilingual populations, given the particular typological similarity and sociolinguistic status of the languages being learned by our participants.
-Catalan and Spanish are both Romance languages, share a considerable amount of cognates, and very often co-exist in the same social spaces (families, schools, working place, etc.).
-Although there are no populations of bilinguals with identical sociolinguistic characteristics, and therefore it would be pointless to compare the sample in this study to any "standard" population of bilinguals, the generalisability of our results should be taken with a grain of salt.
-It is possible, for instance, that the large amount of cognates that Catalan-Spanish bilingual children are exposed leads them to learn at early stages of word acquisition that words tend to sound similar in both languages, and therefore to be more keen on assuming that similar sounding words are semantically related.
-Children exposed to a pair of languages sharing a fewer amount of cognates might not make the same assumption, even when the word-forms they hear in the presence of the same referents are very similar at the phonological level.
-
-Another limitation of the present study is the predominantly cross-sectional design of our data set.
-We collected vocabulary data from children aged 12 to 32 months, a critical period in development in which the foundations of a lexicon get established.
-We estimated word acquisition curves by modelling the probability of comprehension and production across ages.
-However, most participants provided data at only one age point.
-Effectively, we are making longitudinal inferences based on cross-sectional data, and drawing conclusions about the development of a single participant across multiple age points using data from different participants at individual age points.
-For this reason, and also because of the observational nature of our study, causal inferences derived from our analyses are unwarranted [@kraemer2000how; @kesmodel2018crosssectional].
-One counterpoint to this shortcoming is that, while data our data is cross-sectional if participants are considered as the observational unit, data about word-forms/translation equivalents is longitudinal.
-The same translation equivalent is responded to at multiple ages.
-Since translation equivalents are our main observational unit of interest over which we primarily conduct statistical inference (e.g., estimating the size of the cognateness-effect), this provides some further justification for our conclusions.
-
-Future studies addressing this subject may benefit for a more adequate design oriented towards making causal inferences.
-One such design can be partially satisfied by (1) collecting a more substantial longitudinal dataset, in which the same participant provides data points for each translation equivalent at multiple ages, and (2) testing the role of cognateness and word exposure in an experimental context, in which such variables can be orthogonally manipulated or randomised.
-This second possibility would also allow to explore an interpretation of the results obtained in the context of *word learning*, which is not possible in the present study.
-Another consideration for future investigations is the involvement of a more diverse range of language pairs as in @braginsky2019consistency or @floccia2018introduction to improve the generalisability of the resulting inferences across more populations of bilinguals.
+The present study contributes a specific data point into the complex landscape of bilingualism research.
+Bilinguals are a remarkably heterogeneous population that can hardly be comprehensively characterised [@sebastian-galles2020bilingual].
+Bilinguals differ across multiple dimensions, spanning from language-specific factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Arabic, or high like the case of Spanish and Greek), to higher-level factors like the socio-political situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others one of the languages may be less dominant than the other).
+This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.
+Our sample, although homogeneous (e.g., similar parental educational level across), represents a fairly particular population of bilinguals.
+The languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages.
+Both languages are used in fairly similar context and are known by the majority of the population: in 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@de2013els].
+In addition, and as commented in the introduction, are both Romance languages and share a considerable amount of cognates.
+Extending our analyses to other bilingual populations, learning typologically more distant languages, and whose languages tend to be used in different contexts (e.g., heritage languages) should be a natural future step for the present investigation.
 
@@
 
-{{< include appendix/_model-predictors.qmd >}}
+{{< include _appendix/_model-predictors.qmd >}}
 
@@
 
-{{< include appendix/_model-details.qmd >}}
+{{< include _appendix/_model-details.qmd >}}
 
@@
 
-{{< include appendix/_model-diagnostics.qmd >}}
+{{< include _appendix/_model-diagnostics.qmd >}}
 
@@
 
-{{< include appendix/_model-doe.qmd >}}
+{{< include _appendix/_model-doe.qmd >}}
 
 {{< pagebreak >}}
+
+# Appendix E: frequency and language exposure as separate predictors {.appendix .unnumbered}
+
+{{< include _appendix/_syllable-frequency.qmd >}}
+
+{{< pagebreak >}}

### 2023-04-28 _appendix/_model-predictors.qmd
Add appendix for syllable frequencies and new revisions

--- 2023-03-29 manuscript/appendix/_model-predictors.qmd
@@
 Our study capitalises in phonology, therefore we operationalised *cognateness* as the phonological similarity between a pair of translation equivalents.
-Our dataset contained `r max(n_items_final)` translation equivalents, each consisting in two word-forms: Catalan and Spanish (see @tbl-translation-levenshtein for an example).
+Our dataset contained `r max(n_items_final)` translation equivalents, each consisting in two word-forms: Catalan and Spanish (see {apatb-translation-levenshtein} for an example).
 Measuring the phonological similarity between two word forms is non trial.
@@
 The Levenshtein distance being one of them.
-The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary, see @eq-levenshtein].
+The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary] (see @eq-levenshtein).
 For instance, /kaza/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /kasa/ (/z/ is replaced by /s/).
@@
 
-The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure (see @tbl-translation-levenshtein for more examples):
+The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure, see {apatb-translation-levenshtein} for more examples:
 
-| Catalan 	      | Spanish   	      | Levenshtein  |       
-|---------	      |---------	      |-----         |
-| *porta* [pOrt5] | *puerta* [pwe4ta] |   0.50 (3)   |
-| *taula* [tawl5] | *mesa* [mesa]     |   0.00 (5)   |
-| *cotxe* [kOtS5] | *coche* [kotSe]   |   0.40 (3)   |
-
-: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study {#tbl-translation-levenshtein}
+```{r apatb-translation-levenshtein}
+#| label: apatb-translation-levenshtein
+#| apa-cap: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study
+tribble(~Catalan, ~Spanish, ~Levenshtein,
+        "porta [pOrt5]", "puerta [pwe4ta]", 0.50,
+        "taula [tawl5]", "mesa [mesa]", 0.00,
+        "cotxe [kOtS5]", "coche [kotSe]", 0.40) |> 
+    gt() |> 
+    fmt_number(is.numeric)
+```

### 2023-04-28 R/items.R

@@ -4,10 +4,13 @@
 #' @param childes A data frame with lexical frequencies extracted from CHILDES, as returned by the [get_childes_frequencies()]
 #' @param class A character vector indicating the word classes to be included in the resulting dataset. Takes `"Adjective"`, `"Noun"` and/or `"Verb"` as values.
 get_items <- function(bvq_data, childes, class = "Noun") {
+    
     classes_available <- c("Noun", "Verb", "Adjective")
+    
     if (!(class %in% classes_available)) {
         cli_abort("class must be one of {classes_available}")
     }
+    
     classes <- class
     
     # find TEs that have one word-form in each language
@@ -24,22 +27,39 @@ get_items <- function(bvq_data, childes, class = "Noun") {
                # exclude problematic items (e.g., multi-word items)
                te %in% duplicated_te,
                # get only translation equivalents with at least one item in each language
-               class %in% classes) |>
+               class %in% classes) |> 
         add_count(te, name = "n_te") |> # get only items with one translation in each language
         filter(n_te == 2) |>
-        distinct(language, te, .keep_all = TRUE)
+        distinct(language, te, .keep_all = TRUE) |> 
+        mutate(xsampa = ipa(ipa),
+               xsampa_flat = flatten_xsampa(xsampa),
+               syll = syllabify_xsampa(xsampa),
+               n_syll = map_int(syll, length),
+               item = str_remove(item, "cat_|spa_")) 
+    
+    syll_freq <- pool_tmp |> 
+        left_join(childes, by = c("childes_lemma" = "token")) |> 
+        replace_na(list(freq_million = 1)) |>
+        unnest_longer(syll) |> 
+        summarise(freq_syll = sum(freq_million),
+                  .by = c(language, syll)) |> 
+        mutate(freq_syll = log10(freq_syll) + 3)
+    
+    syllables <- pool_tmp |> 
+        unnest_longer(syll, indices_to = "position") |> 
+        left_join(syll_freq, by = c("syll", "language")) |> 
+        summarise(freq_syll_sum = sum(freq_syll),
+                  .by = c(te, n_syll, language))
     
     # compute normalised Levenshtein distance (see ?stringdist::`stringdist-package`)
     # number of edit operations needed to make two strings identical
[…]

### 2023-05-01 R/bvq.R
Upgrade bvq package and change targets workflow

@@ -1,67 +1,0 @@
-#' Get BVQ data
-#' @param ... Arguments to be passed to \code{bvq_responses}
-#' @returns A named list of data frames containing questionnaire responses, participant data, and item data from BVQ
-get_bvq <- function(...) {
-    # get multilex participant data
-    p <- bvq_participants()
-    # get multilex questionnaire responses
-    r <- bvq_responses(p, ...)
-    
-    # merge participant data with questionnaire responses
-    edu_levels <- c(
-        "noeducation",
-        "primary",
-        "secondary",
-        "complementary",
-        "vocational",
-        "university"
-    )
-    edu_labels <- c(
-        "No education",
-        "Primary",
-        "Secondary",
-        "Complementary",
-        "Vocational",
-        "University"
-    )
-    
-    l <- bvq_logs(p, r) |>
-        mutate(id_bvq = id) |> 
-        mutate(
-            across(starts_with("edu_"), ~ as.numeric(
-                factor(.x, levels = edu_levels, ordered = TRUE)
-            )),
-            # get maximum educational attainment of parents
-            edu_parent = apply(cbind(edu_parent1, edu_parent2), 1, max, na.rm = FALSE),
-            # recode it as factor
-            edu_parent = factor(edu_parent, levels = 1:6, labels = edu_labels)
-        ) |>
-        select(
-            id,
-            id_bvq,
-            time,
-            time_stamp,
-            age,
-            lp,
-            dominance,
-            version,
-            completed,
-            doe_catalan,
-            doe_spanish,
-            doe_others,
-            edu_parent
-        )
-    
-    pool <- bvqdev::pool |>
-        mutate(sampa_flat = flatten_sampa(sampa))
-    
-    # get list of all relevant datasets
-    bvq_data <- list(
-        participants = p,
-        responses = r,
-        logs = l,
-        pool = pool
-    )
-    
-    return(bvq_data)
-}

### 2023-05-01 R/participants.R

@@ -12,8 +12,7 @@ get_participants <- function(bvq_data,
                              other_threshold = 0.1) {
     
     participants <- bvq_data$logs |>
-        filter(
-            completed,
+        filter(completed,
             # get only short versions of the questionnaire
             str_detect(version, "Lockdown|Short"),
             # get only data from complete questionnaire responses
@@ -22,18 +21,14 @@ get_participants <- function(bvq_data,
             lp %in% .env$lp,
             between(age, .env$age[1], .env$age[2]),
             sum(doe_catalan + doe_spanish) > .env$other_threshold,
-            id_bvq != "bilexicon_1699",
-            # exclude participants (duplicated entry)
             # make sure that degrees of exposure are between 0 and 1
             between(doe_spanish, 0, 1),
             between(doe_catalan, 0, 1),
-            between(doe_others, 0, 1)
-        ) |>
+            between(doe_others, 0, 1)) |>
         mutate(time = as.integer(time)) |> 
-        # see ?multilex::get_longitudinal
+        # see ?bvq::get_longitudinal
         get_longitudinal(longitudinal = longitudinal) |>
-        mutate(id = as.integer(as.factor(id_bvq))) |>  # make ID shorter
-        select(id, id_bvq, time, time_stamp, list = version,
+        select(id, time, time_stamp = date_finished, list = version,
                age, lp, doe_catalan, doe_spanish, edu_parent) |> 
     arrange(id)
     

### 2023-05-01 _appendix/_model-diagnostics.qmd
Add appendix for syllable frequencies and new revisions
Upgrade bvq package and change targets workflow

--- 2023-04-17 appendix/_model-diagnostics.qmd
@@
 
+{{< pagebreak >}}
+
 ## Posterior draws and bi-variate scatterplots {.unnumbered}
@@
 
+{{< pagebreak >}}
+
 ## Posterior-predictive checks {.unnumbered}
@@
 #| fig-height: 3
-#| out-width: 50%
-#| out-height: 50%
+#| out-width: 80%
+#| out-height: 80%
 y_int <- as.integer(as.integer(responses$response))

### 2023-05-01 _targets.R

@@ -16,87 +16,79 @@ test_paths <- list.files("tests/testthat",
 invisible(lapply(test_paths, source))
 
 # list package dependencies ----------------------------------------------------
-tar_option_set(
-    packages = c(
-        "arrow",
-        "bayesplot",
-        "bayestestR",
-        "brms",
-        "bvqdev",
-        "childesr",
-        "cli",
-        "conflicted",
-        "dplyr",
-        "ggplot2",
-        "ggtext",
-        "glue",
-        "gt",
-        "ipa",
-        "janitor",
-        "keyring",
-        "knitr",
-        "lubridate",
-        "marginaleffects",
-        "mice",
-        "patchwork",
-        "purrr",
-        "quarto",
-        "readxl",
-        "rlang",
-        "scales",
-        "stringdist",
-        "stringr",
-        "testthat",
-        "tibble",
-        "tidybayes",
-        "tidyr",
-        "tidytext",
-        "usethis"
-    )
-)
+tar_option_set(packages = c("arrow",
+                            "bayesplot",
+                            "bayestestR",
+                            "brms",
+                            "bvq",
+                            "childesr",
+                            "cli",
+                            "conflicted",
+                            "dplyr",
+                            "ggplot2",
+                            "ggtext",
+                            "glue",
+                            "gt",
+                            "ipa",
+                            "janitor",
+                            "keyring",
+                            "knitr",
+                            "lubridate",
+                            "marginaleffects",
+                            "mice",
+                            "patchwork",
+                            "purrr",
+                            "quarto",
+                            "readxl",
+                            "rlang",
+                            "scales",
+                            "stringdist",
+                            "stringr",
+                            "testthat",
+                            "tibble",
+                            "tidybayes",
+                            "tidyr",
+                            "tidytext",
+                            "usethis"))
 
 resolve_conflicts()
 
 # define global options --------------------------------------------------------
-options(
-    mc.cores = 4,
-    brms.backend = "cmdstanr",
-    tidyverse.quiet = TRUE,
-    knitr.duplicate.label = "allow",
-    loo.cores = 1,
-    knitr.graphics.error = FALSE
-)
+options(mc.cores = 4,
+        brms.backend = "cmdstanr",
+        tidyverse.quiet = TRUE,
+        knitr.duplicate.label = "allow",
+        loo.cores = 1,
+        knitr.graphics.error = FALSE)
 
 list(
[…]

### 2023-05-08 manuscript.qmd
Fix phonological transcriptions and render manuscript

@@
 ---
-title: "Lexical frequency mediates the facilitation effect of cognateness in bilingual lexical acquisition"
+title: "Cognate beginnings to bilingual lexical growth"
 
-shorttitle: Frequency and cognateness in lexical acquisition
+shorttitle: "Cognate beginnings to bilingual lexical growth"
 
@@
 # set ggplot theme and colour palette ------------------------------------------
+theme_set(
+    theme_minimal() +
+        theme(
+            panel.grid = element_line(colour = "grey",
+                                      linetype = "dotted"),
+            axis.line = element_line(colour = "black"),
+            text = element_text(size = 12, colour = "black"),
+            axis.text = element_text(colour = "black")
+        )
+)
 
-theme_set(theme_custom())
-
-bayesplot_theme_set(theme_custom()  +
-                        theme(panel.grid = element_blank(),
+bayesplot_theme_set(
+        theme(
+            axis.line = element_line(colour = "black"),
+            text = element_text(size = 12, colour = "black"),
+            panel.grid = element_blank(),
             axis.text = element_text(size = 5),
@@
             plot.title = element_text(size = 8),
-                              plot.subtitle = element_text(size = 8)))
+            plot.subtitle = element_text(size = 8)
+        )
+)
 
 clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+color_scheme_set(rep(clrs[1], 6))
 
 # set params -------------------------------------------------------------------
-
-color_scheme_set(rep(clrs[1], 6))
 
@@
 ```
-
-# Introduction
 
@@
 Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical or the phonological level.
+These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical and the phonological level.
 Little is known about vocabulary acquisition in bilinguals learning a pair of typologically closer languages (e.g., when both are Romance languages).
 
-For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language.
-These cross-language word pairs that share meaning are referred to as *translation equivalents*.
-Throughout the present study, we make the assumption that bilinguals form common semantic representations for both languages (e.g., the concept of a *chair* is shared for the English and Spanish words 'chair' and 'silla')^[The referential contexts in which bilingual children hear (and learn) both labels may not be identical.
+One way typological distance may impact bilingual trajectories of language acquisition is the presence (or absence) of *cognates* in their linguistic input.
+Quite often, bilingual children encounter more than one label for the same referent: one in each language.
+For instance, an English-Spanish bilingual might hear both *dog* and *perro* in the presence of a dog We will refer to such cross-language synonyms as *translation equivalents*.^[Throughout the present study, we will make the assumption that bilinguals form a common semantic representations for each translation equivalent (e.g., the concept 'dog' is shared for the English and Spanish words *dog* and *perro*).
+The referential contexts in which bilingual children hear (and learn) both labels may not be identical.
 For instance, the English word *finger* refers to eight parts of the body (excluding thumbs and toes), whereas its Spanish translation *dedos* refers to twenty body parts (fingers, thumbs, and toes).
@@
 For simplicity, and in line with previous (formal and verbal) models of bilingual lexical processing [e.g., @shook2013bilingual; @dijkstra2019multilink; @dijkstra1998bia; @kroll2010revised], we assume common semantic representations for translation pairs.].
-Depending on the similarity between both languages, the two sets of words may overlap in form in varying degrees.
-When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like Spanish and English (Romance and Germanic languages, respectively).
-For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might be exposed to the words *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
-It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
+Sometimes, the word-forms involved in a translation equivalent may be phonologically similar.
+Such form-similar translation equivalents are referred to as cognates.
+While the previously mentioned translation pair *dog* and *perro* are non-cognates (both word-forms hardly share any phonemes), the translation pair *cat* and *gato* can be considered cognates.
+The amount of cognates shared by a pair of languages is dependent on their typological distance.
+When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like English and Spanish (Romance and Germanic languages, respectively).
+Therefore, bilingual children learning two typologically close languages may be exposed to a larger amount of cognates than those learning two typologically more distant languages.
+For instance, in the presence of a door, a Catalan-Spanish bilingual might be exposed to the words *porta* and *puerta* (cognate translation equivalent with the common meaning 'door'), whereas an English-Spanish bilingual might hear *door* and *puerta* (non-cognate translation equivalent).
+
+It could be the case that mapping two phonologically similar labels (cognates like *porta*-*puerta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *door* and *puerta*).
 If cognates are easier to acquire than non-cognates, bilinguals learning two languages that share a high proportion of cognates should benefit more often from this facilitation effect than bilinguals learning two languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
-
 @floccia2018introduction provided evidence for such a facilitation effect of typological distance and vocabulary size.
@@
 The Levenshtein distance is a metric that computes the distance between two strings of characters/phonemes by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
-For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/kæt/-/\textipa{kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /æ/ by /\textipa{A}/).
-The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /kæt/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/.
+For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/\textipa{kæt}/-/\textipa{kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /\textipa{æ}/ by /\textipa{A}/).
+The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{kæt}/-/\textipa{"Ga.ta}/) would score three, as it requires at least three editions: substituting /\textipa{k}/ for /\textipa{G}/, substituting /\textipa{\ae}/ for /\textipa{a}/ and adding another /\textipa{a}/.
 Levenshtein distance scores are divided by the length of the longest string to normalise the similarity scores between 0 and 1.
 Finally, the result is subtracted from one, to get the complementary, and therefore a measure of phonological similarity, instead of dissimilarity.
-Following the previous examples: $1/4 = 0.25$ for /kæt/-/\textipa{kAt}/, and $3/4 = 0.75$ for /kæt/-/\textipa{Gata}/.
+Following the previous examples: $1/4 = 0.25$ for /\textipa{kæt}/-/\textipa{kAt}/, and $3/4 = 0.75$ for /\textipa{kæt}/-/\textipa{"Ga.ta}/.
 Finally, this variable was included as a predictor of participants' vocabulary sizes.
@@
 The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during speech production or comprehension, even in fully monolingual environments.
-Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extend to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @hoshino2008cognate; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
+Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extends to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @hoshino2008cognate; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 In the same line, @bosch2014first reported a similar overrepresentation of translation equivalents in Catalan-Spanish bilinguals: a high proportion of acquired Catalan words held an also acquired translation in Spanish.
-Finally, a recent investigation by @tsui2022area presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
+Finally, a recent investigation by @tsui2022are presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
 Overall, these findings suggest that bilingual children benefit from a semantic facilitation effect driven by the acquisition of translation equivalents.
 
-It is possible that cognateness increases the amount of cross-language activation, facilitating early word acquisition.
+In this scenario, where translation equivalents are pervasive in the bilingual lexicon since early ages, it is possible that cognateness increases the amount of cross-language activation, boosting early word acquisition.
 This could ultimately lead to children who are learning languages with a larger proportion of cognates (e.g., English and Dutch) showing larger vocabulary sizes, as found by Floccia et al.
-There is evidence of for phonological and semantic similarity facilitating word acquisition in monolinguals: infants are more likely to acquired new words that are semantically or phonologically more connected to other words, than words with poorer connectivity in either of those dimensions, such as words belonging to sparse phonological neighbourhoods [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children; @storkel2004children].
+There is evidence of phonological and semantic similarity facilitating word acquisition in monolinguals: infants are more likely to acquired new words that are semantically or phonologically more connected to other words, than words with poorer connectivity in either of those dimensions, such as words belonging to sparse phonological neighbourhoods [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children; @storkel2004children].
 If phonological similarity plays a cross-language facilitation role during bilingual word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity (e.g., are taxonomically related) but not phonological similarity.
@@
 As infants receive linguistic input, they also accumulate opportunities to learn words until a threshold is crossed, an the word can be considered to be acquired.
-In line with this account, the frequency with which a child encounters a particular word-form in their speech input is a critical determiner of the age at which such word will be acquired.
-The more frequently [@hidaka2013computational; @mollica2017how].
-For instance, words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'taxi' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does]: infants might accumulate learning experiences with 'flower' faster than with 'taxi'.
-
-Previous accounts on vocabulary growth based on accumulator models are a convenient framework to study bilingual vocabulary growth, as bilinguals share fundamentally similar learning mechanisms with monolinguals.
+The frequency with which a child encounters a particular word-form in their speech input is therefore a critical determiner of the age at which such word will be acquired, in part because they accumulate learning experiences with high-frequency words faster than with lower-frequency words [@hidaka2013computational; @mollica2017how].
+Following this account, words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are predicted to be acquired earlier in age than words with low lexical frequency such as 'taxi' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words).
+We argue that previous accounts on vocabulary growth based on accumulator models are a convenient framework to study bilingual vocabulary growth.
+Bilinguals share fundamentally similar learning mechanisms with their monolinguals peers.
 Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilingual].
 One of such divergences is the fact that bilinguals' frequency of exposure to a word-form is mediated by their degree of exposure to their languages.
-Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual would (who listens 100% of the time to English).
+Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual (who listens 100% of the time to English).
+would need.
+In the present study, we incorporate this feature of the bilingual linguistic input to an account of the cognateness effect on word acquisition.
 
@@
 facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired.
-We hypothesise that both forms in the translation equivalent (e.g., *gat* and *gato* [Catalan and Spanish for 'cat'], for a Catalan-Spanish bilingual) follow their acquisition trajectory as they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to.
-When one of the word-forms is sufficiently acquired, the number of learning instances left needed for the acquisition of the other word-form is reduced.
+We hypothesise that both forms in the translation equivalent (e.g., *gat* and *gato* [Catalan and Spanish for 'cat'], for a Catalan-Spanish bilingual) follow their acquisition trajectory as they would for a monolingual child.
+As the infant receives speech input, both word-forms accumulate learning instances, which occur a a rate proportional to the lexical frequency of each word-form and the amount of exposure to the language the word-form belongs to.
+When one of the word-forms has accumulated a certain amount of learning instances, it is acquired.
+After such event takes places, we hypothesise that the number of learning instances left needed for the acquisition of the other word-form (yet to be acquired) is reduced, resulting in an earlier age of acquisition that it would be expected if the acquisition of both word-forms was independent.
+We further suggest that this facilitation is proportional to the amount of phonological similarity (i.e., cognateness) between both word-forms, in such way that the facilitation is larger within similar-sounding translation equivalents, and negligible for form-dissimilar translation equivalents.
 
-This facilitation would occur as the result of the semantic and phonological overlap between both words.
-When the child has sufficient lexical knowledge of *gat*, and therefore there is some for-referent association in place, future presentations of the similar sounding translation *gato* in the presence of the same reference should lead to an increased likelihood of acquisition.
+Consider the example of the Catalan-Spanish cognate translation equivalent *gat*-*gato* [cat] (whose word-forms share three phonemes), and of the non-cognate translation equivalent *gos*-*perro* [dog] whose word-forms only share one phoneme.
+When the child has sufficient lexical knowledge of *gat*, and therefore there is some form-referent association in place, future presentations of the similar-sounding translation *gato* in the presence of the same reference should lead to an increased likelihood of acquisition.
 In the case of a non-cognate translation equivalent, such facilitation would be unlikely.
@@
 
-{apafg-hypotheses} illustrates graphically a theoretical account of the acquisition of a Catalan-Spanish cognate translation equivalent, *gat*-*gato* [cat] whose word-forms share three phonemes, and of a non-cognate translation equivalent, *gos*-*perro* [dog] whose word-forms only share one phoneme.
+{apafg-hypotheses} provides a graphic, theoretical illustration of this account.
 We assume the case of a Catalan-Spanish bilingual child who is exposed 65% of the time to Catalan, and 35% of the time to Spanish.
-On the top panels, we illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/gat/ or /gos/) and its Spanish translation (/perro/ or /gato/).
-We assume an arbitrary threshold of learning instances at which the child is considered to have acquired a word (solid grey line).
+On the top panels, we illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/\textipa{gat}/ or /\textipa{gos}/) and its Spanish translation (/\textipa{"pe.ro}/ or /\textipa{"ga.to}/).
+We assume an arbitrary, theoretical threshold for the learning instances at which the child is considered to have acquired a word (solid grey line).
 Points and triangles show the age-of-acquisition at which the amount of learning instances reaches the threshold.
-We show the predicted age-of-acquisition under two hypothesis: (1) a cognateness facilitation effect that lowers the acquisition threshold after one of the word-forms (in this case Catalan /gat/) has been acquired, and (2) no cognateness facilitation, for which the acquisition of the second word-form occurs unconditionally of the acquisition of the first.
+We show the predicted age-of-acquisition under two hypothesis: (1) a cognateness facilitation effect that lowers the acquisition threshold after one of the word-forms (in this case Catalan /\textipa{gat}/) has been acquired, and (2) no cognateness facilitation, for which the acquisition of the second word-form occurs unconditionally of the acquisition of the first.
 On the bottom, we show the predicted probability of acquisition for the Catalan and Spanish words under each hypothesis, as it would be expected to observe in a vocabulary questionnaire like the one in the present study.
@@
 #| out-width: 100%
+#| fig-width: 6
+#| fig-height: 5
 knitr::include_graphics(normalizePath("assets/img/diagram.png"))
@@
 
-In line with this hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al.
+This hypothesis provides a convenient account for the results in Floccia et al.
+The size of the effect of linguistic similarity on vocabulary size that Floccia et al.
 reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure).
@@
 *Cognateness* was defined as a continuous measure of phonological similarity between translation equivalents.
-We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computer].
+We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-XSAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computer].
 This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
 This results in a proportion that indicates how much the two phonetic transcriptions of the translation equivalent are similar to each other, ranging from zero (no similarity at all) to one (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]` and *pájaro* `[paxa4o]`, Catalan and Spanish for 'bird', and returns 50% for *lluna* `[Lun5]` and *luna* `[luna]`, Catalan and Spanish for 'moon' (see Appendix A for more details).
+For instance, this measure of phonological similarity returns 0% for *ocell* /`uceL`/ and *pájaro* /`paxa4o]`/, Catalan and Spanish for 'bird', and returns 50% for *lluna* /`Lun5`/ and *luna* /`luna`/, Catalan and Spanish for 'moon' (see Appendix A for more details).
 
@@
 
-```{r}
+```{r methods-data}
 #| label: items
@@
 We used the highest self-reported educational attainment of parents or caregivers as a proxy of participants' socio-economic status (SES).
-This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: 'sense escolaritzar/sin escolarizar' [no education], 'educació primària/educación primaria' [primary school], 'educació secundària/educación secundaria' [secondary school], 'batxillerat/bachillerato' [complementary studies/high school], 'cicles formatius/ciclos formativos' [vocational training], and 'educació universitària/educación universitaria' [university degree].
+This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: *sense escolaritzar/sin escolarizar* [no education], *educació primària/educación primaria* [primary school], *educació secundària/educación secundaria* [secondary school], *batxillerat/bachillerato* [complementary studies/high school], *cicles formatius/ciclos formativos* [vocational training], and *educació universitària/educación universitaria* [university degree].
 Most families reported university studies (`r participants_edu["University"]`, `r percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
@@
     filter(include, !is.na(wordbank_lemma)) |> 
-    drop_na(sampa) |> 
+    drop_na(xsampa) |> 
     distinct(te, language)
@@
 #| apa-cap: "Bayesian leave-one-out cross validation (LOO-CV)."
-#| apa-note: "We report the ELPD of each model along their associated standard errors (SE).
-The last column indicates the difference in ELPD between each model and the model with the best predictive accuracy (Model 3).
+#| apa-note: "ELPD (*SD*): expected log-predictive density and standard error.
+Diff.
+(*SD*): diference in ELPD between each model and the model with the best predictive accuracy (Model 3) and standard error of the difference.
 Pareto-k estimates of all models were acceptable (k < 0.5)."
@@
     fmt_number(2:5, decimals = 2, sep_mark = ",") |>
-    cols_merge_uncert(elpd_loo, se_elpd_loo) |> 
-    cols_merge_uncert(elpd_diff, se_diff) |> 
+    cols_merge(c(elpd_loo, se_elpd_loo), pattern = "{1} ({2})") |> 
+    cols_merge(c(elpd_diff, se_diff), pattern = "{1} ({2})") |> 
     cols_label(
         model = "Model",
-        elpd_loo = "ELPD (SE)",
-        elpd_diff = "ELPD difference (SE)"
+        elpd_loo = "ELPD (SD)",
+        elpd_diff = "Diff. (SD)"
     ) |>
@@
 #| warning: false
+#| out-width: 80%
 # summarise posterior draws
@@
 
-The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).
-For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), 2)`.
-The effect of $\text{Cognateness}$ interacted with that of $\text{Exposure}$: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 2)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
-Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
-This effect was negligible when for words with median or high exposure (+1 SD) (see {apafg-results-marginal}).
-
 ```{r apafg-results-marginal}
@@
 #| warning: false
-#| fig-width: 8
-#| fig-height: 6
-#| out-width: 80%
+#| fig-width: 7
+#| fig-height: 4.5
+#| out-width: 100%
 #| eval: true
 #| apa-cap: "Posterior marginal effects of Model 3.
-Thin lines correspond to 50 mean predictions generated from 50 MCMC samples of the posterior distribution of the model.
+Thin lines correspond to 50 mean predictions.
 Thicker lines indicate the median of those predictions.
-Different colours indicate different levels of phonological similarity (as indicated by the Cognateness similarity between pairs of translation equivalents).
+Different colours indicate different levels of cognateness (phonological similarity).
 Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
@@
 In-sample predictions lie inside the grey rectangles."
-
 clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")
@@
                              levels = unique(exposure_std),
-                             labels = c("Low exposure (-1 SD)",
+                             labels = c("Low exposure (-1 *SD*)",
                                         "Mean exposure",
-                                        "High exposure (+1 SD)"),
+                                        "High exposure (+1 *SD*)"),
                              ordered = TRUE),
@@
 
-To rule out the possibility that cognateness facilitation effect we found was due to cognateness comprising more frequent syllables than non-cognates---and therefore not because of their cognate status itself---, we compared the syllabic frequency of cognates and non-cognates included in our data analyses.
+The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).
+For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), 2)`.
+The effect of $\text{Cognateness}$ interacted with that of $\text{Exposure}$: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 2)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
+Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 *SD*), cognateness increased the probability of acquisition substantially.
+This effect was negligible when for words with median or high exposure (+1 *SD*) (see {apafg-results-marginal}).
+
+To rule out the possibility that cognateness facilitation effect we found was due to cognateness comprising more frequent syllables than non-cognates---and therefore not because of their cognate status itself---, we compared the syllabic frequency of cognates and non-cognates included in our analyses.
 To calculate syllable frequency, we first extracted all syllables embedded in the selected words.
 For each syllable, we summed the lexical frequency of all the words in which such syllable appeared.
-The resulting value provides an estimate of the number of times the syllable appears in child-directed speech, embedded within different words.
+The resulting value provided an estimate of the number of times the syllable appears in child-directed speech, embedded within different words.
 Finally, for each word, we summed the frequency of its syllables, as an estimate of the syllabic frequency of the word.
 We fit a Bayesian model with cognateness (Levenshtein similarity) as response variable, and the main effects of syllable frequency and number of syllables (to control for the fact words with more syllables are more likely to score higher in syllabic frequency) as predictors.
-This model provided strong evidence for an association between cognateness and syllabic frequency being equivalent to zero (see Appendix E).
+This model provided strong evidence for the association between cognateness and syllabic frequency being equivalent to zero (see Appendix E).
 
@@
     mutate(item = gsub("cat_|spa_", "", item),
-           sampa = gsub("\\.", "", sampa)) |>
+           xsampa = gsub("\\.", "", xsampa)) |>
     filter(te %in% items$te, 
            item %in% items$item) |>
-    select(te, language, sampa) |> 
+    select(te, language, xsampa) |> 
     pivot_wider(names_from = language,
-                values_from = sampa, 
+                values_from = xsampa, 
                 id_cols = te,
@@
 The present study explored the cognateness facilitation effect on bilingual word acquisition.
-We found such effect to be mediated by lexical frequency and degree of language exposure: phonologically similar translation equivalents were on average acquired earlier in age compared to phonologically dissimilar translation equivalents, but this effect was only observed for word-forms to which participants were exposed less often, i.e., words with low lexical frequency and/or words from the language of lower daily exposure.
+We found such effect to be mediated by lexical frequency and degree of language exposure.
+Phonologically similar translation equivalents were on average acquired earlier in age than phonologically dissimilar translation equivalents, but only word-forms to which participants were exposed less often showed this facilitation effect (i.e., words with low lexical frequency and/or words from the language of lower daily exposure).
 In addition, we provided convergent evidence of a facilitation effect of word length and lexical frequency on word acquisition.
-Our results are relevant in the light of two previous findings.
-First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (which shared many cognates, e.g., English-German) showed larger vocabulary sizes than those learning typologically more distance languages (which shared fewer cognates, e.g.,  English-Mandarin).
+Our results are particularly relevant in the light of two previous findings.
+First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (which shared many cognates, e.g., English-German) showed larger vocabulary sizes than those learning typologically more distant languages (which shared fewer cognates, e.g., English-Mandarin).
 Second, @mitchell2022cognates found an earlier age-of-acquisition for cognates, compared to non-cognates.
 The outcomes of both studies pointed to cognateness facilitating word acquisition, but the underpinnings of such effect were unclear.
+In this study, we capitalised on accumulator models of language acquisition to put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
 
-In this study, we capitalised on accumulator models of language acquisition to put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
 We hypothesised that the acquisition of a word benefits from cognateness only *after* its translation equivalent has been acquired.
@@
 We further hypothesised that such facilitation is proportional to the phonological similarity between both word-forms.
-If two translation equivalents sharing a fair deal of phonological similarity (e.g., /\textipa{"pOR.t@}/-/\textipa{"pweR.ta}/, `r percent(stringdist::stringsim("pORt@", "pweRta"))` Levenshtein similarity), learning of one of them should facilitate the acquisition of the other more strongly, compared to when both word-forms share only few phonemes (e.g., /\textipa{k@"Di.R@}/-/\textipa{"si.\textlambda a}/ , `r percent(stringdist::stringsim("DiR@", "siLa"))`).
+If two translation equivalents sharing a fair deal of phonological similarity (e.g., /\textipa{"pOR.t@}/-/\textipa{"pweR.ta}/, `r round(stringdist::stringsim("pORt@", "pweRta"), 2)` Levenshtein similarity), learning of one of them should facilitate the acquisition of the other more strongly, compared to when both word-forms share only few phonemes (e.g., /\textipa{k@"Di.R@}/-/\textipa{"si.\textlambda a}/ , `r round(stringdist::stringsim("DiR@", "siLa"), 2)`).
 
@@
 For instance, an infant exposed to Catalan and Spanish 65% and 35% of the time, respectively, should be more likely to learn the Spanish word-form /\textipa{"ga.to}/ later in age than its Catalan translation /\textipa{gat}/.
-If this was the case, /\textipa{"ga.to}/ should be more likely to receive facilitation from the acquisition of /\textipa{gat}/ than *vice versa* (see {apafg-hypotheses} for a graphic illustration).
+If this was the case, /\textipa{"ga.to}/ should be more likely to receive facilitation from the acquisition of /\textipa{gat}/ than vice versa (see {apafg-hypotheses} for a graphic illustration).
 In line with these predictions, we found that cognateness facilitated word acquisition more strongly in words from the language of lower exposure.
@@
 
-Overall, our results add further nuance to the findings from previous studies reporting a cognateness facilitation effect, and provide new insights into the mechanisms underpinning such effect.
-In particular, we argue that the account we provide in the present study is a plausible explanation for the language distance facilitation effect reported by Floccia et al.
+Overall, our results add further nuance to the previously reported cognateness facilitation effect, and provide new insights into the mechanisms underpinning such effect.
+In particular, we argue that the account we provide in the present study is a plausible explanation for the language distance facilitation reported by Floccia et al.
 In their study, bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional (non-English) language than those bilinguals who were learning two languages that shared less cognates (e.g., English-Chinese).
@@
 
-Lexical frequency plays a central role in this model of word bilingual word acquisition, given that together with age and language exposure, it accounts for the number of times a child is exposed to some word-form.
-Consistently, our composite measure of lexical frequency (which accounts for both lexical frequency and language exposure) was the second strongest predictor of word acquisition, only after age.
-These results parallel previous work on language acquisition, in which lexical frequency effects are pervasive and play a capital role in many models of lexical acquisition [@ambridge2015ubiquity].
-There is robust evidence that, when keeping relevant variables constant, high-frequency words are learnt earlier in age than low-frequency words [e.g., @goodman2008does; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
+Lexical frequency plays a central role in this model of word bilingual word acquisition.
+Together with age and amount of language exposure, lexical frequency accounts for the number of times a child is exposed to some word-form.
+Consistently, our composite measure of lexical frequency (which considers both lexical frequency and language exposure) was the second strongest predictor of word acquisition in our results, only after age.
+These outcomes mirror previous work on language acquisition, in which lexical frequency effects are pervasive and play a capital role in many models of lexical acquisition [@ambridge2015ubiquity].
+There is robust evidence that, when keeping relevant variables constant, high-frequency words are learnt earlier in age than low-frequency words [e.g.; @goodman2008does; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
 The outputs of our regression models parallel these findings.
-Given two words with identical scores in all the predictors we considered (e.g., word length), the with higher a higher score in the composite measure of lexical frequency (which corrects for dual language exposure) is predicted a higher probability of acquisition.
+Given two words with identical scores in all the predictors we considered (e.g., word length), that with a higher score in the composite measure of lexical frequency (which corrects for dual language exposure) is predicted a higher probability of acquisition.
 Interestingly, the interaction between this predictor and participants' age was substantial, suggesting that this predictor plays a stronger role at later ages, with older infants benefiting more strongly from lexical frequency than younger infants.
 This points to early word acquisition being supported by mere exposure to word-forms more strongly than later word acquisition.
-@braginsky2019consistency, found a similar interaction in their monolingual sample: consistently, infants learning one of a wide range of languages showed larger lexical frequency effects at earlier ages.
-Whether this effect results from an extant developmental process (and if so, what mechanisms underlie it), or from methodological artefacts (e.g., frequency measures being more representative of a child's linguistic input at older ages) is a matter that should be addressed in future work, in order to further explore the interaction of this predictor and the effect of cognateness in word acquisition.
+@braginsky2019consistency found a similar interaction in their monolingual sample: consistently, infants learning one of a wide range of languages showed larger lexical frequency effects at earlier ages.
+Whether this effect results from an extant developmental process (and if so, what mechanisms underlie it), or from methodological artefacts (e.g., frequency measures being more representative of a child's linguistic input at older ages) is a matter that should be addressed in future work.
 
@@
 This could explain a cognate advantage in word acquisition without the need for parallel activation playing a relevant role.
-This account would not necessarily imply that bilinguals are insensitive to the phonological mismatches between cognate word-forms.
-There is strong evidence that both monolingual and bilingual infants are highly sensitive to subtle mispronunciations of word-forms during word recognition [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels].
+This account would not necessarily imply that bilinguals are insensitive to the phonological mismatches between cognate word-forms: there is strong evidence that both monolingual and bilingual infants are highly sensitive to subtle mispronunciations of word-forms during word recognition [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels].
 Rather, the assimilation of both word-forms in the lexicon might occur at a higher level, where both word-forms are considered *referentially* equivalent, so that every exposure to either of the (distinguishable) word-forms of the translation equivalent is counted as a learning instance for both.
-This would impact the acquisition mechanism in such way that any cognate effect is not explained by a reduction of the number of learning instances needed to acquired a word (through parallel activation, after its translation has been acquired), but by a faster increase in cumulative learning instances across ages, leading to a similar outcome: an earlier age-of-acquisition for cognates.
+This would impact the acquisition mechanism in such way that any cognate effect is not explained by a reduction of the number of learning instances needed to acquire a word (through parallel activation, after its translation has been acquired), but by a faster increase in cumulative learning instances across ages, leading to a similar outcome: an earlier age-of-acquisition for cognates.
 
@@
 For instance, @ramon-casas2009vowel showed that 18-26 months-old bilingual toddlers reject mispronounced labels in word recognition tasks, although only when such mispronunciation involves a contrast that is phonemic in both languages.
-In their study, Catalan-Spanish bilinguals treated the words /pe\textesh/ and its mispronunciated form /p\textepsilon\textesh/, which involved the /e-\textepsilon/ contrast, only present in Catalan as different, but not the words /pe\textesh/ and /pi\textesh/, which involved the the /e-i/ contrast, present in both languages.
-Moreover, a later study by @ramon2010non showed that bilinguals treat the language-specific /e-\textepsilon/ contrast differently when embedded within *non-cognate* word-forms.
-These results suggest that, except for acoustically close, language-specific /e-\textepsilon/ contrast, bilinguals treat similar-sounding words differently in word recognition tasks, especially when both word-forms are separated by more than one phoneme.
+In their study, Catalan-Spanish bilinguals treated differently the words /\textipa{pe\textesh}/ and its mispronunciated form /\textipa{pE\textesh}/, which involved the /\textipa{e}-\textipa{E}/ contrast (only present in Catalan), but treated similarly the words /\textipa{pe\textesh}/ and /\textipa{pi\textesh}/, which involved the /\textipa{e}-\textipa{i}/ contrast (present in both languages).
+Moreover, a later study by @ramon2010non showed that bilinguals treat the language-specific /\textipa{e}-\textipa{E}/ contrast differently when embedded within *non-cognate* word-forms.
+These results suggest that, except for the acoustically close, language-specific /\textipa{e}-\textipa{E}/ contrast, bilinguals treat similar-sounding words differently in word recognition tasks, especially when both word-forms are separated by more than one phoneme.
+In the light of these findings, the possibility that bilinguals merge cognates into shared lexical representations becomes more implausible.
 
-In the light of these findings, the possibility that bilinguals keep similar-sounding translation equivalents as separate lexical representations becomes more plausible.
-This is especially true for the case of the Catalan-Spanish bilinguals in our sample: there are few exemplars of Catalan-Spanish translation equivalents involving a /e-\textepsilon/ substitution.
-Only `r percent(nrow(E_e)/length(unique(items$te)))` translation equivalents in our questionnaire involved a /e-\textepsilon/ change between the Catalan and the Spanish items, only one of such was a minimal pair: /\textipa{tREN}/-/\textipa{tReN}/ [*train*].
+This is especially true for the case of the Catalan-Spanish bilinguals in our sample: there are few exemplars of Catalan-Spanish translation equivalents involving a /\textipa{e}-\textipa{E}/ substitution.
+Only `r percent(nrow(E_e)/length(unique(items$te)))` of the translation equivalents in our questionnaire involved a /\textipa{e}-\textipa{E}/ change between the Catalan and the Spanish items, and only one of them was a minimal pair: /\textipa{tREN}/-/\textipa{tReN}/ [*train*].
 Moreover, few words in our study differ only in one phoneme substitution.
 Only `r percent(lv_distr$prop[nrow(lv_distr)], accuracy = 0.01)` of the translation equivalents included in our analyses shared between 90% and 100% phonological similarity (measured as the Levenshtein similarity between the X-SAMPA transcriptions of the Catalan and Spanish form), and only `r percent(lv_distr$prop[nrow(lv_distr)-1], accuracy = 0.01)` shared between 80 and 90% similarity.
-This shows that the presence of identical and near-identical cognates in our study is very low, and therefore most translation equivalents differed in more than one phoneme.
+This shows that the presence of identical and near-identical cognates in our study was very low, and therefore most translation equivalents differed in more than one phoneme.
 Finally, and perhaps more critically, there is evidence beyond word recognition that bilinguals treat minimal pairs differently at the referential level.
-For instance, @ramon-casas2017minimalpair found that bilingual toddlers are able to learn new minimal pairs involving the /e-\textepsilon/ contrast.This suggests that toddlers' tolerance towards cognate mispronunciations at the referential level might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@mattock2010first; @havy2016phonetic].
+For instance, @ramon-casas2017minimalpair found that bilingual toddlers are able to learn new minimal pairs involving the /\textipa{e}-\textipa{E}/ contrast.
+This suggests that toddlers' tolerance towards cognate mispronunciations at the referential level might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@mattock2010first; @havy2016phonetic].
 
-Our study shares similar methodological limitations with previous work using parental vocabulary reports provided.
-Such reports can be subject to measurement error induced by caregivers, who may sometimes overestimate or underestimate participants' true probability of acquisition of words.
-For instance, although in the present study caregivers were explicitly instructed *not* to rely on their responses to the words in one languages when responding to words in the other language, it is a possibility that caregivers were biased to assume---at least to some extent---that because the child knows a word in one language, the child must also know the word in the other language (especially when both words are similar-sounding, i.e., cognates).
+Our study shares similar methodological limitations with previous work using vocabulary reports provided by caregivers.
+Such reports can be subject to measurement error induced by caregivers who may sometimes overestimate or underestimate participants' true probability of acquisition of words [e.g., @houston-price2007discrepancy].
+For instance, although in the present study caregivers were explicitly instructed *not* to rely on their responses to Catalan words when responding to Spanish (and vice versa), it is possible that some caregivers assumed---at least to some extent---that because the child knew a word in one language, the child should also know the word in the other language (especially when both words are similar-sounding, i.e., cognates).
 Production estimates may be more prompt to such biases, in part because of the slower pace at which their articulatory abilities develop, compared to their word recognition abilities.
-For instance, even at 47 months of age few (typically developing) English monolinguals reach 100% intelligibility scores in word production, as rated by English native adults [@hustad2020development; @hustad2021speech].
+For instance, even at 47 months of age, few (typically developing) English monolinguals reach 100% intelligibility scores in word production, as rated by English native adults [@hustad2020development; @hustad2021speech].
 This gap between comprehension and production is even larger in the less dominant language of bilingual children [@giguere2022bilingual].
-For this reason, caregivers may be more uncertain about what words can be counted as "acquired" in this modality.
+For this reason, caregivers may be more uncertain about what words can be counted as *acquired* in this modality.
 Despite these potential biases, vocabulary checklist filled by parents show strong evidence of concurrent validity with other estimates of vocabulary size or lexical processing [@killing2008move; @feldman2005concurrent; @gillen2021tapping; but see @houston-price2007discrepancy].
@@
 The present study contributes a specific data point into the complex landscape of bilingualism research.
-Bilinguals are a remarkably heterogeneous population that can hardly be comprehensively characterised [@sebastian-galles2020bilingual].
-Bilinguals differ across multiple dimensions, spanning from language-specific factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Arabic, or high like the case of Spanish and Greek), to higher-level factors like the socio-political situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others one of the languages may be less dominant than the other).
+Bilinguals are a remarkably heterogeneous population that can hardly be satisfactorily characterised in a comprehensive way [@sebastian-galles2020bilingual].
+Bilinguals differ across multiple dimensions, spanning from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Arabic, or high like the case of Spanish and Greek), to higher-level factors like the socio-political situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages may be less dominant than the other).
 This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.
-Our sample, although homogeneous (e.g., similar parental educational level across), represents a fairly particular population of bilinguals.
-The languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages.
-Both languages are used in fairly similar context and are known by the majority of the population: in 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@de2013els].
+Our sample, although homogeneous (e.g., similar parental educational level across), represents a fairly particular population of bilinguals: the languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages, both languages are used in fairly similar contexts, and both languages are known by the majority of the population.
+In 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@de2013els].
 In addition, and as commented in the introduction, are both Romance languages and share a considerable amount of cognates.
@@
 
-In summary, our study has replicated previous studies and provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
+In summary, our study provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
 We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word's phonological similarity with its translation in the other language.

### 2023-05-08 _targets_packages.R

@@ -4,7 +4,7 @@ library(bayesplot)
 library(bayestestR)
 library(brms)
 library(bs4Dash)
-library(bvqdev)
+library(bvq)
 library(childesr)
 library(cli)
 library(clustermq)
@@ -40,5 +40,6 @@ library(testthat)
 library(tibble)
 library(tidybayes)
 library(tidyr)
+library(tidytext)
 library(usethis)
 library(visNetwork)

### 2023-05-08 _appendix/_syllable-frequency.qmd
Fix phonological transcriptions and render manuscript

@@
-````{r}
+```{r prepare-syllables}
+#| label: prepare-syllables
+#| echo: false
+#| warning: false
+#| message: false
 duplicated_te <- bvq_data$pool$te[duplicated(bvq_data$pool$te)]
@@
     # drop items with missing observations in these variables
-    drop_na(sampa, wordbank_lemma) |>
-    mutate(xsampa = ipa(ipa)) |> 
+    drop_na(xsampa, wordbank_lemma) |>
     filter(n_lemmas == 1,
@@
 #| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
+#| echo: false
+#| warning: false
+#| message: false
 items |> 
@@
 #| fig-height: 3
+#| echo: false
+#| warning: false
+#| message: false
 pool_tmp |> 
@@
 #| fig-width: 7
+#| echo: false
+#| warning: false
+#| message: false
 syll <- items |> 
@@
 #| out-width: 80%
+#| echo: false
+#| warning: false
+#| message: false
 syll |> 
@@
 ```{r syllables-posterior}
+#| label: syllables-posterior
+#| echo: false
+#| warning: false
+#| message: false
 rope_interval <- c(lower = -0.5, upper = 0.5)

### 2023-05-08 src/helpers.R

@@ -1,0 +1,51 @@
+# run the targets workflow
+make <- function() {
+    job::job({
+        targets::tar_make()
+        job::export("none")  # return nothing
+    },
+    import = NULL,
+    title = "Trajectories")
+    
+    to_delete <- c("manuscript/QTDublinIrish.otf")
+    for(i in to_delete) {
+        if (file.exists(i)) {
+            file.remove(i)
+        }
+    }
+}
+
+# remove targets products
+unmake <- function(keep_fits = TRUE) {
+    path <- "results/fit.rds"
+    tar_destroy(ask = FALSE)
+    
+    if (!keep_fits) {
+        filenames <-
+            list.files("results", pattern = "fit", full.names = TRUE)
+        if (length(filenames > 0)) {
+            lapply(filenames, file.remove)
+        }
+    }
+    
+    usethis::ui_done("Removed project outputs!")
+}
+
+#' Resolve NAMESPACE conflicts
+#'
+resolve_conflicts <- function() {
+    suppressMessages({
+        conflict_prefer("last_warnings", "rlang")
+        conflict_prefer("filter", "dplyr")
+        conflict_prefer("lag", "dplyr")
+        conflict_prefer("between", "dplyr")
+        conflict_prefer("timestamp", "utils")
+        conflict_prefer("ar", "brms")
+        conflict_prefer("chisq.test", "stats")
+        conflict_prefer("discard", "scales")
+        conflict_prefer("duration", "lubridate")
+        conflict_prefer("fisher.test", "stats")
+        conflict_prefer("lag", "dplyr")
+    })
+    cli_alert_success("Resolved conflicts")
+}

### 2023-05-19 manuscript/_appendix/_model-diagnostics.qmd

--- 2023-05-01 _appendix/_model-diagnostics.qmd
@@
 Values lower than 1.01 are recommended, while values higher than 1.05 indicate that chains might have trouble converging and therefore the estimated parameters must be taken with caution.
-{apafg-model-rhats-neffs} (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference.
+@tbl-rhats-neffs (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference.
 Most values are lower than 1.01, and never higher than 1.05, which provides evidence of successful MCMC convergence.
@@
 Values larger than 0.1 are recommended.
-{apafg-model-rhats-neffs} (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models.
+@tbl-rhats-neffs (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models.
 Most values are larger than 0.1, although model 0 ($\mathcal{M}_0$) accumulates most effective sample sizes close to 0.1.
 
-```{r apafg-model-rhats-neffs}
-#| label: apafg-model-rhats-neffs
-#| apa-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
+```{r tbl-rhats-neffs}
+#| label: tbl-rhats-neffs
+#| tbl-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
 #| echo: false
@@
 
-```{r apafg-model-ppc}
-#| label: apafg-model-ppc
-#| apa-cap: "Model posterior predictive checks (PPC).
+```{r tbl-ppc}
+#| label: tbl-ppc
+#| tbl-cap: "Model posterior predictive checks (PPC).
 Bars indicate the observed proportion of responses to each category (No, Understands, and Understands and Says).

### 2023-05-19 manuscript/_appendix/_syllable-frequency.qmd

--- 2023-05-08 _appendix/_syllable-frequency.qmd
@@
 
-```{r items-table}
-#| label: tbl-items
+```{r tbl-items-table}
+#| label: tbl-syll-items
 #| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
@@
 Less than 1% of the words included in the analyses presented in the main body of the manuscripts had five syllables.
-No words had more than five syllables (see {apafg-syl-number}).
+No words had more than five syllables (see @fig-syll-number).
 We extracted lexical frequencies from the English corpora in the CHILDES database.
@@
 
-```{r apafg-syl-number}
-#| label: apafg-syl-number
-#| apa-cap: Distribution of the number of syllables in Catalan and Spanish
+```{r fig-syll-number}
+#| label: fig-syll-number
+#| fig-cap: Distribution of the number of syllables in Catalan and Spanish
 #| fig-width: 7
@@
 To compute the relative frequency of each syllable in Catalan and Spanish (i.e., how many times the syllables appears in every million words in Catalan or Spanish speech), we summed the relative lexical frequency in CHILDES of every word that contains such syllable in the corresponding language.
-{apafg-syl-freq} shows the distribution of frequencies across syllables in Catalan and Spanish.
+@fig-syll-freq shows the distribution of frequencies across syllables in Catalan and Spanish.
 In the log10 scale, syllable frequencies in Catalan and Spanish followed a slightly asymmetric distribution, with most syllables scoring around 1,000 counts per million, and a longer tail to the right of the distribution.
 
-```{r apafg-syl-freq}
-#| label: apafg-syl-freq
-#| apa-cap: "Distribution of apositional syllable frequencies in Spanish and Catalan"
+```{r fig-syll-freq}
+#| label: fig-syll-freq
+#| fig-cap: "Distribution of apositional syllable frequencies in Spanish and Catalan"
 #| fig-height: 3.5
@@
 
-{apafg-top-syll-freq} shows the top twenty most frequent syllables in Spanish.
+@fig-top-syll-freq shows the top twenty most frequent syllables in Spanish.
 In both languages, most of the most frequent syllables followed a consonant-vowel (CV) structured.
@@
 
-```{r apafg-top-syll-freq}
-#| label: apafg-top-syll-freq
-#| apa-cap: "Top 20 most frequent syllables"
+```{r fig-top-syll-freq}
+#| label: fig-top-syll-freq
+#| fig-cap: "Top 20 most frequent syllables"
 #| fig-width: 8
@@
 We fit this model running `r dim(model_fit_syllables$fit)[2]` sampling chains with `r format(dim(model_fit_syllables$fit)[1], big.mark = ",")` iterations each.
-{apatb-coefs} shows a summary of the posterior distribution of the fixed effects in the model.
+@tbl-syll-coefs shows a summary of the posterior distribution of the fixed effects in the model.
 As expected, words with more syllables scored higher in syllabic frequency: all posterior draws for the regression coefficient of the main effect of this predictor fell outside the ROPE defined between -0.5 and +0.5 ($\beta$ = `r round(post_draws_list[["b_n_syll_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_syll_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_syll_std"]]$.upper, 2)`]).
@@
 
-```{r apatb-syllable-coefs}
-#| label: apatb-syllable-coefs
-#| apa-cap: "Posterior distribution of regression coefficients."
-#| apa-note: "Median: median of the posterior distribution in the probability scale.
+```{r tbl-syll-coefs}
+#| label: tbl-syll-coefs
+#| tbl-cap: "Posterior distribution of regression coefficients."
+#| tbl-note: "Median: median of the posterior distribution in the probability scale.
 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.
@@
 
-{apafg-syllable-marginal} shows the median posterior-predicted syllabic frequencies for words with one to four syllables, for the whole range of cognateness values.
+@tbl-syll-marginal shows the median posterior-predicted syllabic frequencies for words with one to four syllables, for the whole range of cognateness values.
 Overall, cognate words' syllabic frequency is equivalent to that of non-cognates.
@@
 
-```{r apafg-syllable-marginal}
-#| label: apafg-syllable-marginal
-#| apa-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
+```{r tbl-syll-marginal}
+#| label: tbl-syll-marginal
+#| tbl-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
 #| fig-width: 7

### 2023-05-19 manuscript/appendix.qmd

---
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

format:
  pdf:
    documentclass: article
    papersize: letter
    fontsize: 12pt
    linestretch: 1.5
    date-format: full
    template-partials:
    - partials/title.tex
    - partials/before-body.tex
    include-in-header: include-in-header.tex
    header-includes: \usepackage{tipa}
    geometry:
    - top=1in
    - bottom=1in
    - left=1in
    - right=1in
    colorlinks: true
    linkcolor: DarkSlateBlue
    urlcolor: DarkSlateBlue
    citecolor: DarkSlateBlue
    link-citations: true
    biblio-title: References
    biblio-config: false
    cite-method: biblatex
    biblatex-chicago: false
    biblio-style: apa
    biblatexoptions: backend=biber
    block-headings: false
    indent: true
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

library(targets)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

tar_load_globals()

tar_load(c(bvq_data, childes, items, participants, responses))
tar_load(c(starts_with("model_"), model_loos))
tar_load(c(posterior_draws, marginal_effects_epreds))

# process objects
items <- bvq_data$pool  |>  
    mutate(item = str_remove(item, "cat_|spa_")) |> 
    select(item, language, semantic_category, class) |> 
    right_join(items, by = c("item", "language"))

# set ggplot theme and colour palette ------------------------------------------

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

color_scheme_set(rep(clrs[1], 6))

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(colour = "grey",
                                    linetype = "dotted"),
          axis.line = element_line(colour = "black"),
          text = element_text(size = 12, colour = "black"),
          axis.text = element_text(colour = "black"))

theme_set(my_theme)

bayesplot_theme_set(my_theme)

set.seed(888)

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = scale_color_gradient,
        ggplot2.continuous.colour = scale_color_gradient,
        knitr.graphics.error = FALSE) 
```

# Appendix A: model predictors {.appendix .unnumbered}

{{< include _appendix/_model-predictors.qmd >}}

{{< pagebreak >}}

# Appendix B: model details {.appendix .unnumbered}

{{< include _appendix/_model-details.qmd >}}

{{< pagebreak >}}

# Appendix C: convergence diagnostics {.appendix .unnumbered}

{{< include _appendix/_model-diagnostics.qmd >}}

{{< pagebreak >}}

# Appendix D: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_model-doe.qmd >}}

{{< pagebreak >}}

# Appendix E: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_syllable-frequency.qmd >}}

{{< pagebreak >}}

### 2023-05-19 manuscript/manuscript.qmd

--- 2023-02-26 manuscript/manuscript.qmd
@@
 ---
-title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual lexical acquisition"
+title: "Cognate beginnings to bilingual lexical growth"
+short-title: "Cognate beginnings"
+author:
+  - name: Gonzalo Garcia-Castro
+    corresponding: true
+    orcid: 0000-0002-8553-4209
+    email: gonzalo.garciadecastro@upf.edu
+    affiliations:
+      - id: cbc
+        name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+        address: Ramon Trias Fargas 25-27
+        city: Barcelona
+        region: Spain
+        postal-code: "08005"
+  - name: Daniela S. Avila-Varela
+    orcid: 0000-0002-3518-8117
+    affiliations: 
+      - ref: cbc
+      - id: ul
+        name: Universidade de Lisboa
+        department: Facultade de Letras
+        address: Alameda da Universidade
+        city: Lisboa
+        region: Portugal
+        postal-code: "1600-214"
+  - name: Ignacio Castillejo
+    orcid: 0000-0001-7445-0416
+    affiliations: 
+      - id: uam 
+        name: Universidad Autónoma de Madrid
+        department: Departamento de Psicología
+        address: Iván Pavlov, 6
+        city: Madrid
+        region: Spain
+        postal-code: "28049"
+  - name: Núria Sebastian-Galles
+    orcid: 0000-0001-6938-2498
+    affiliations: 
+      - ref: cbc
+date: "May 10th, 2023"
+code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
+keywords:
+  - lexical acquisition
+  - vocabulary
+  - bilingualism
+  - item response theory
+  - bayesian
+csl: "../assets/apa7.csl"
+bibliography: "../assets/references.bib"
+link-citations: true
+editor: source
 
-pagetitle: Cross-linguistic similarity and  word acquisition
+format:
+  docx:
+    reference-doc: "templates/reference.docx"
+    knitr:
+      opts_chunk:
+        dev: "png"
+        dev-args:
+          - "cairo-png"
+        dpi: 300
+        out.width: "80%"
+  pdf:
+    documentclass: article
+    papersize: letter
+    fontsize: 12pt
+    linestretch: 1.5
+    date-format: full
+    template-partials:
+    - templates/title.tex
+    - templates/before-body.tex
+    include-in-header: include-in-header.tex
+    header-includes: \usepackage{tipa}
+    geometry:
+    - top=1in
+    - bottom=1in
+    - left=1in
+    - right=1in
+    colorlinks: true
+    linkcolor: DarkSlateBlue
+    urlcolor: DarkSlateBlue
+    citecolor: DarkSlateBlue
+    link-citations: true
+    biblio-title: References
+    biblio-config: false
+    cite-method: biblatex
+    biblatex-chicago: false
+    biblio-style: apa
+    biblatexoptions: backend=biber
+    block-headings: false
+    indent: true
+
+author-note:
+  disclosures:
+    conflict-of-interest: |
+        The authors declare no conflicts of interest with regard to the funding source of this study.
+    financial-support: |
+        This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
+Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
+    gratitude: |
+        We are grateful to Chiara Santolin, Ege E.
+Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
+We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support, and Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
+We also thank all families and infants who participated in the experiments.
 
 abstract: |
-    A prominent feature of the bilingual mental lexicon is that the activation of representations spreads in a language non-selective fashion.
-An instance of such non-selectivity is embodied by cognateness (i.e., form-similarity between translation equivalents, such as *flower*-*flor* in English and Spanish), which impacts lexical access and processing in bilinguals.
-Recent studies have suggested that cognateness facilitates vocabulary acquisition in bilingual toddlers---who show larger vocabulary sizes when their languages share many cognates---and an earlier age-of-acquisition for cognates than for non-cognates.
-The specific mechanisms underpinning such facilitation are unclear.
-We present an account of bilingual early lexical acquisition in which cognateness interacts with lexical frequency and language exposure to facilitate the acquisition of low-frequency words.
-We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals aged 12 to 34 months.
-We used a Bayesian Exploratory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
-We found converging evidence for an earlier age-of-acquisition for cognate words, and for such effect being mediated by lexical frequency and language exposure.
-Low-frequency words, and words from the language of least exposure were more strongly benefitted by their cognate status than high-frequency words.
+    Language non-selectivity is prominent feature of bilingual lexical processing.
+An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates.
+The specific mechanisms behind such facilitation are unclear.
+We present an account of early bilingual vocabulary growth in which cognateness interacts with lexical frequency and language exposure to facilitate the word acquisition.
+We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals toddlers.
+We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
+We found cognateness facilitation effect only in low-frequency words.
 Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
 
-thanks: |
-    This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
-Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
-The authors declare no conflicts of interest with regard to the funding source of this study.
-We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and Ege E.
-Özer for their help-ful feedback.
-We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support.
-We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
-We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises.
-We also thank all families and infants who participated in the experiments.
-Data collection was half-way when the COVID-19 pandemic started.
-We would like to pay special tribute to the families that collaborated with us under these difficult circumstances.
-
-keywords: lexical acquisition, vocabulary, bilingualism, item response theory, bayesian
+fig-dpi: 1000
 ---
 
-```{r}
+```{r setup}
 #| label: setup
@@
 #| warning: false
-# load packages
-library(ggplot2)
-library(patchwork)
-library(bayesplot)
-library(ggtext)
-library(lubridate)
-library(brms)
-library(glue)
-library(dplyr)
-library(tidyr)
-library(purrr)
-library(gt)
-library(stringr)
-library(tibble)
-library(tidybayes)
-library(janitor)
-library(scales)
-library(bvqdev)
+# load objects -----------------------------------------------------------------
 
-# load targets as R objects
-tar_load(participants)
-tar_load(responses)
-tar_load(items)
-tar_load(childes)
-tar_load(bvq_data)
+library(targets)
 
-tar_load(model_fit_4)
-tar_load(posterior_draws)
+tar_config_set(store = here::here('_targets'),
+               script = here::here('_targets.R'))
 
-tar_load(model_fit_4_prior)
-tar_load(marginal_effects_epreds)
-tar_load(posterior_draws_doe)
+tar_load_globals()
 
-tar_load(model_log_liks)
-tar_load(model_loos)
+tar_load(c(bvq_data, childes, items, participants, responses))
+tar_load(c(starts_with("model_"), model_loos))
+tar_load(c(posterior_draws, marginal_effects_epreds))
 
-tar_load(model_rhats)
-tar_load(model_neffs)
-tar_load(model_mcmc_pairs)
-tar_load(model_ppcs)
-
+# process objects
 items <- bvq_data$pool  |>  
@@
 
-# set ggplot theme and colour palette
-theme_set(theme_custom()) # set custom ggplot theme
-bayesplot_theme_set(
-    theme_custom()  +
-        theme(panel.grid = element_blank(),
-              axis.text = element_text(size = 5),
-              strip.text = element_text(size = 5),
-              plot.title = element_text(size = 8),
-              plot.subtitle = element_text(size = 8))
-)
+# set ggplot theme and colour palette ------------------------------------------
+
 clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-color_scheme_set(rep(clrs[4], 6))
+
+color_scheme_set(rep(clrs[1], 6))
+
+my_theme <- theme_minimal() +
+    theme(panel.grid = element_line(colour = "grey",
+                                    linetype = "dotted"),
+          axis.line = element_line(colour = "black"),
+          text = element_text(size = 12, colour = "black"),
+          axis.text = element_text(colour = "black"))
+
+theme_set(my_theme)
+
+bayesplot_theme_set(my_theme)
+
+set.seed(888)
+
 options(ggplot2.ordinal.fill = clrs,
@@
         ggplot2.continuous.fill = scale_color_gradient,
-        ggplot2.continuous.colour = scale_color_gradient)
-
-# resolve namespace conflicts
-# conflict_prefer("filter", "dplyr")
-# conflict_prefer("select", "dplyr")
+        ggplot2.continuous.colour = scale_color_gradient,
+        knitr.graphics.error = FALSE) 
 ```
 
-# Introduction {#sec-introduction}
-
 The foundations of word learning are in place early in age.
-From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [@fenson2007macarthurbates, @samuelson2021precision].
-During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
-Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e.
-vocabulary size, @fenson1994variability; @frank2021variability].
+From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthurbates; @samuelson2021precision].
+During the last half of the second year of life, new word-forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
+Such early stages of lexical acquisition are characterised by substantial heterogeneity across different populations of language learners, reflected for instance on (the variability of) quantitative indices of lexical development, like vocabulary size [i.e., vocabulary size, @fenson1994variability; @frank2021variability].
 Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
@@
 Since birth, these infants had been learning English, Cantonese, and Mandarin, respectively.
-The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*).
+The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caregivers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animal sounds (*woof-woof*).
 These results were later extended by @frank2021variability to a more diverse set of languages, and found that such cross-language commonalities are stronger at earlier stages of lexical acquisition.
 
-Most literature on early word acquisition still relies extensively on data from monolingual children, and is oblivious to substantial proportion of the world population acquiring more than one language since early ages [@grosjean2021extent].
-Previous research about bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
-@hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals did.
+Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
+Previous research on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
+@hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did.
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
 Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-These studies rely on bilingual samples whose languages are relatively distant in typology, as it is the case of English and Spanish.
-It is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
-What role could linguistic distance play during early vocabulary growth?
+These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical and the phonological level.
+Little is known about vocabulary acquisition in bilinguals learning a pair of typologically closer languages (e.g., when both are Romance languages).
 
-For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language.
-Depending on the similarity between both languages, the two sets of words may overlap in form in varying degrees.
-When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like Spanish and English (Romance and Germanic languages, respectively).
-For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates).
-It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*).
+One way typological distance may impact bilingual trajectories of language acquisition is the presence (or absence) of *cognates* in their linguistic input.
+Quite often, bilingual children encounter more than one label for the same referent: one in each language.
+For instance, an English-Spanish bilingual might hear both *dog* and *perro* in the presence of a dog We will refer to such cross-language synonyms as *translation equivalents*.^[Throughout the present study, we will make the assumption that bilinguals form a common semantic representations for each translation equivalent (e.g., the concept 'dog' is shared for the English and Spanish words *dog* and *perro*).
+The referential contexts in which bilingual children hear (and learn) both labels may not be identical.
+For instance, the English word *finger* refers to eight parts of the body (excluding thumbs and toes), whereas its Spanish translation *dedos* refers to twenty body parts (fingers, thumbs, and toes).
+These discrepancies are reflected, for instance, in adults' performance in lexical decision tasks [@chaouch-orozco2023are].
+The fact that some translation pairs share more semantic features than others does not, however, inform whether bilinguals form common or separate semantic representations for translation equivalents.
+What English speakers understand for 'fingers' might not be identical to what Spanish speakers understand for 'dedos'.
+But when an English speaker learns Spanish, it seems implausible that they would form a separate semantic representation for what 'dedo' conveys at the conceptual level, compared to the possibility that they associate the new label to the semantic representation they already had formed for 'fingers'.
+Therefore, the lack of complete semantic alignment is not necessarily an argument against the existence of common semantic representations for translation equivalents.
+For simplicity, and in line with previous (formal and verbal) models of bilingual lexical processing [e.g., @shook2013bilingual; @dijkstra2019multilink; @dijkstra1998bia; @kroll2010revised], we assume common semantic representations for translation pairs.].
+
+Sometimes, the word-forms involved in a translation equivalent may be phonologically similar.
+Such form-similar translation equivalents are referred to as cognates.
+While the previously mentioned translation pair *dog* and *perro* are non-cognates (both word-forms hardly share any phonemes), the translation pair *cat* and *gato* can be considered cognates.
+The amount of cognates shared by a pair of languages is dependent on their typological distance.
+When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like English and Spanish (Romance and Germanic languages, respectively).
+Therefore, bilingual children learning two typologically close languages may be exposed to a larger amount of cognates than those learning two typologically more distant languages.
+For instance, in the presence of a door, a Catalan-Spanish bilingual might be exposed to the words *porta* and *puerta* (cognate translation equivalent with the common meaning 'door'), whereas an English-Spanish bilingual might hear *door* and *puerta* (non-cognate translation equivalent).
+
+It could be the case that mapping two phonologically similar labels (cognates like *porta*-*puerta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *door* and *puerta*).
 If cognates are easier to acquire than non-cognates, bilinguals learning two languages that share a high proportion of cognates should benefit more often from this facilitation effect than bilinguals learning two languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
-
-@floccia2018introduction provided evidence in line with this claim.
+@floccia2018introduction provided evidence for such a facilitation effect of typological distance and vocabulary size.
 The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
 The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
-The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of *cognateness* between each pair of languages.
-Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
-The Levenshtein distance is a metric that computes the distance between two strings of characters by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
-For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/kæt/-/\textipa{kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /æ/ by /\textipa{A}/).
-The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{kæt}/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/.
-Levenshtein distance scores are then divided by the length of the longest string to normalise the similarity scores between 0 and 1.
-Following the previous examples: 1/3 (0.33) for /kæt/-/\textipa{kAt}/, and 3/4 (0.75) for /kæt/-/\textipa{Gata}/.
+
+The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of *cognateness* between each pair of languages.
+Phonological overlap was measured by computing the Levenshtein distance between each cross-language pair of phonetic transcriptions.
+The Levenshtein distance is a metric that computes the distance between two strings of characters/phonemes by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
+For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/\textipa{"kæt}/-/\textipa{"kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /\textipa{æ}/ by /\textipa{A}/).
+The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{"kæt}/-/\textipa{"Ga.ta}/) would score three, as it requires at least three editions: substituting /\textipa{k}/ for /\textipa{G}/, substituting /\textipa{\ae}/ for /\textipa{a}/ and adding another /\textipa{a}/.
+Levenshtein distance scores are divided by the length of the longest string to normalise the similarity scores between 0 and 1.
+Finally, the result is subtracted from one, to get the complementary, and therefore a measure of phonological similarity, instead of dissimilarity.
+Following the previous examples: $1/4 = 0.25$ for /\textipa{"kæt}/-/\textipa{"kAt}/, and $3/4 = 0.75$ for /\textipa{"kæt}/-/\textipa{"Ga.ta}/.
 Finally, this variable was included as a predictor of participants' vocabulary sizes.
-Among other findings, Floccia and co-workers reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
-For example, English-Dutch bilinguals (22.14% mean cognateness), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% mean cognateness) were able to produce in Mandarin.
+Among other findings, Floccia and co-workers reported an increase in vocabulary size in the additional language (i.e., not English) associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
+For example, English-Dutch bilinguals (0.2214 mean phonological overlap), were able to produce more Dutch words than English-Mandarin bilinguals (0.0197 mean phonological overlap) were able to produce in Mandarin.
+More recently, @blom2020crosslanguage and @bosma2019longitudinal reported similar results, providing converging evidence of for facilitatory effect of language distance on vocabulary size estimates.
 
@@
 pointed to *parallel activation* as the underpinning mechanism behind their results.
-The parallel activation hypothesis stems from the language non-selective account of lexical access, and suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain], even in fully monolingual environments.
+The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during speech production or comprehension, even in fully monolingual environments.
+Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extends to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @hoshino2008cognate; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 
-Regarding word acquisition, the parallel activation account is a plausible account for Floccia et al.'s results in the light of (1) the evidence pointing to bilingual toddlers' lexicon consisting of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage], and that (2) there is experimental evidence of parallel activation during word comprehension in children and toddlers [e.g., @vonholzen2012language; @poulin-dubois2013lexical].
-It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning languages with a larger proportion of cognates (such as children learning English and Dutch) showing larger vocabulary sizes.
-This account in which cognateness underlies the facilitation effect of typological language distance is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children].
-If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity but not phonological similarity.
+Bilingual toddlers and children---still in the process to developing a mature lexicon---also show evidence of parallel activation during comprehension [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic; @poulin-dubois2013lexical].
+Whether---and if so, how---the language non-selectivity shapes the early bilingual lexicon is still unclear.
+Previous work has pointed to an interactive account of lexical acquisition, in which toddlers' lexical representations in one language influence the acquisition of words in the other language.
+This cross-language facilitation is, in principle, possible since early ages: the bilingual infant lexicon consists in a substantial proportion of  translation equivalents: the presence of a word in the lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage; @pearson1995crosslanguage].
+In fact, translation equivalents may even be overrepresented in the early bilingual lexicon.
+@bilson2015semantic found that bilingual toddlers produced a higher proportion of translation equivalents than it would be expected if they learned two independent lexica simultaneously, which suggests that bilingual children preferentially acquired words for which they already knew a label in the other language.
+In the same line, @bosch2014first reported a similar overrepresentation of translation equivalents in Catalan-Spanish bilinguals: a high proportion of acquired Catalan words held an also acquired translation in Spanish.
+Finally, a recent investigation by @tsui2022are presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
+Overall, these findings suggest that bilingual children benefit from a semantic facilitation effect driven by the acquisition of translation equivalents.
 
-Evidence supporting an earlier age of acquisition for cognates is scarce.
+If translation equivalents are pervasive in the bilingual lexicon since early ages, it is possible that cognateness increases the amount of cross-language activation, boosting early word acquisition.
+This could ultimately lead to children who are learning languages with a larger proportion of cognates (e.g., English and Dutch) showing larger vocabulary sizes, as found by Floccia et al.
+There is evidence of phonological and semantic similarity facilitating word acquisition in monolinguals: infants are more likely to acquired new words that are semantically or phonologically more connected to other words, than words with poorer connectivity in either of those dimensions, such as words belonging to sparse phonological neighbourhoods [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children; @storkel2004children].
+If phonological similarity plays a cross-language facilitation role during bilingual word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity (e.g., are taxonomically related) but not phonological similarity.
+
+Evidence supporting an earlier age-of-acquisition for cognates is scarce.
 @bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
@@
 The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-On the other hand, Floccia et al.'s estimates are statistically more reliable given their remarkably larger sample size (372 participants).
-Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level.* The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
-By aggregating the responses from all words into a single datum per child, information about the acquisition status of cognates vs.
-non-cognates was no longer available.
-Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
+Floccia et al.'s estimates are statistically more reliable given their remarkably larger sample size (372 participants).
+Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age-of-acquisition at the *word-level*.
+The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
+By aggregating the responses from all words into a single datum per child, it was no longer possible to test the effect of cognateness on word acquisition at the word-form level.
+Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as the response variable, age-sensitive claims about the effect of cognateness on word acquisition should be taken with caution [a limitation also present in the @bosch2014first study].
 
@@
 They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
-The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty).
-These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar.
+The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age-of-acquisition norms (an index of word difficulty).
+These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size seems to grow faster when both languages are more similar at the lexical level.
 Word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
 Parallel activation remains a plausible explanation for this effect.
-When a child is exposed to a word form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition.
+When a child is exposed to a word-form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition.
 The precise underpinnings of this effect are still unclear.
 
-In the present investigation we approach the cognate facilitation on word learning using the conceptual framework of accumulator models of language acquisition [@hidaka2013computational; @mollica2017how].
-We build on the assumption that word acquisition is a function of mainly---but not exclusively---the amount of experience a child has accrued with the corresponding word-referent pairing [see @kachergis2022standard for review].
-As infants accumulate linguistic input, they also accumulate opportunities to associate familiar word-forms to their corresponding referents.
-Words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'headphones' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does].
-Infants might accumulate learning experiences with 'flower' faster than with 'taxi'.
-In the case of bilinguals, the amount of exposures to a word-form an infants receives is also mediated by their degree of exposure to their languages.
-Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual would (who listens 100% of the time to English).
+In the present investigation we approach the cognate facilitation on word learning by extending the conceptual framework of accumulator models of language acquisition to the bilingual case.
+Accumulator models formalise the widespread assumption that a child's vocabulary growth rate is a function of mainly---but not exclusively---the number of learning instances that a child has accrued with a particular word [see @kachergis2022standard for review].
+A learning instance of a word consists, for example, in hearing such word-form (*flower*) in the presence of its referent (a flower).
+We devise word acquisition as a continuous process in which a child accumulates experience with a given word-form and its referential context until the number of cumulative learning instances meets some theoretical threshold.
+After this, we consider the lexical representation of the word as *consolidated*, which leads to a non-linear increase in the probability of caregivers reporting such word as acquired in a vocabulary checklist like the one in the present study.
+As infants receive linguistic input, they also accumulate opportunities to learn words until a threshold is crossed, an the word can be considered to be acquired.
+The rate at which monolingual children encounter learning instances with some word-form is strongly determined by age: as time passes children are exposed to a larger amount of linguistic input, and therefore accumulate more learning experiences with all words.
 
-We suggest that the cognateness effect proposed by @floccia2018results and @mitchell2022cognates facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired.
-We hypothesise that both forms in the translation equivalent (e.g., *flower* and *flor*, for an English-Spanish bilingual) follow their acquisition trajectory in a similar way they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to.
-When one of the word-forms is sufficiently acquired, the acquisition trajectory of the other is impacted, so that the number of needed learning instances left for acquisition are reduced.
-This facilitation would occur as the result of the semantic and phonological overlap between both words.
-When the child has sufficient lexical knowledge of *flower*, and therefore there is some for-referent association in place, future presentations of the similar sounding translation *flor* in the presence of the same reference should lead to an increased likelihood of acquisition.
+Not all words occur in the linguistic input with identical frequency, so children accumulate experience with some word-forms at faster rates than with other word-forms.
+This results in frequent words being acquired earlier in age than less frequent words [e.g., @ambridge2015ubiquity; @goodman2008does; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
+For this reason, lexical frequency plays a central role in most models of lexical growth, especially in the case of accumulator models [@hidaka2013computational; @mollica2017how].
+Following this account, words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are predicted to be acquired earlier in age than words with low lexical frequency such as 'taxi' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words).
+We argue that previous accounts on vocabulary growth based on accumulator models are a convenient framework to study bilingual vocabulary growth.
+Bilinguals share fundamentally similar learning mechanisms with their monolinguals peers.
+Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilingual].
+One of such divergences is the fact that bilinguals' frequency of exposure to a word-form is mediated by their degree of exposure to their languages.
+Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual (who listens 100% of the time to English).
+would need.
+In the present study, we incorporate this feature of the bilingual linguistic input to an account of the cognateness effect on word acquisition.
+
+We suggest that the cognateness effect proposed by Floccia et al.
+and Mitchell et al.
+facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired.
+We hypothesise that both forms in the translation equivalent (e.g., *gat* and *gato* [Catalan and Spanish for 'cat'], for a Catalan-Spanish bilingual) follow their acquisition trajectory as they would for a monolingual child.
+As the infant receives speech input, both word-forms accumulate learning instances, which occur a a rate proportional to the lexical frequency of each word-form and the amount of exposure to the language the word-form belongs to.
+When one of the word-forms has accumulated a certain amount of learning instances, it is acquired.
+After such event takes places, we hypothesise that the number of learning instances left needed for the acquisition of the other word-form (yet to be acquired) is reduced, resulting in an earlier age-of-acquisition that it would be expected if the acquisition of both word-forms was independent.
+We further suggest that this facilitation is proportional to the amount of phonological similarity (i.e., cognateness) between both word-forms, in such way that the facilitation is larger within similar-sounding translation equivalents, and negligible for form-dissimilar translation equivalents.
+
+Consider the example of the Catalan-Spanish cognate translation equivalent *gat*-*gato* [cat] (whose word-forms share three phonemes), and of the non-cognate translation equivalent *gos*-*perro* [dog] whose word-forms only share one phoneme.
+When the child has sufficient lexical knowledge of *gat*, and therefore there is some form-referent association in place, future presentations of the similar-sounding translation *gato* in the presence of the same reference should lead to an increased likelihood of acquisition.
 In the case of a non-cognate translation equivalent, such facilitation would be unlikely.
-For instance, even if the child has already acquired the word *dog*, hearing its dissimilar Spanish translation *perro* would hardly lead to an increased probability of acquisition.
+For instance, even if the child has already acquired the word *gos*, hearing its dissimilar Spanish translation *perro* would hardly lead to an increased probability of acquisition.
 In line with this account, only the later-acquired form of the translation equivalent would benefit from its cognate status.
-Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
-For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age of acquisition for cognates than for non-cognates.
+Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual; @hoff2012dual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
+For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age-of-acquisition for cognates than for non-cognates.
 
-![Schematic representation of the hypothesised cognate facilitation effect on word acquisition.
-To the left, the acquisition of a cognate translation equivalent (*flower*-*flor*, which share four phonemes) is affected by the phonological similarity between both word forms.
-When the child hears *flower*, activation spreads across languages to activate *flor* too.
-As a result, after exposure to *flower* not only the association between *flower* and its referent is strengthened, but also that of *flower* and its common referent.
-To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*, which do not share any phoneme) does not undergo the such facilitation.
-Both word-forms are fairly dissimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](img/hypotheses.jpg){#fig-hypotheses}
+@fig-hypothesis provides a graphic, theoretical illustration of this account.
+We assume the case of a Catalan-Spanish bilingual child who is exposed 65% of the time to Catalan, and 35% of the time to Spanish.
+On the top panels, we illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/\textipa{gat}/ or /\textipa{gos}/) and its Spanish translation (/\textipa{"pe.ro}/ or /\textipa{"ga.to}/).
+We assume an arbitrary, theoretical threshold for the learning instances at which the child is considered to have acquired a word (solid grey line).
+Points and triangles show the age-of-acquisition at which the amount of learning instances reaches the threshold.
+We show the predicted age-of-acquisition under two hypothesis: (1) a cognateness facilitation effect that lowers the acquisition threshold after one of the word-forms (in this case Catalan /\textipa{gat}/) has been acquired, and (2) no cognateness facilitation, for which the acquisition of the second word-form occurs unconditionally of the acquisition of the first.
+On the bottom, we show the predicted probability of acquisition for the Catalan and Spanish words under each hypothesis, as it would be expected to observe in a vocabulary questionnaire like the one in the present study.
+We assume that such probability starts at 0% at birth, reaches 50% (and grows the fastest) at the age-of-acquisition, and reaches 100% at some points in later ages.
 
-In line with this hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al.
+```{r fig-hypotheses}
+#| label: fig-hypotheses
+#| fig-cap: "Schematic representation of the hypothesised cognate facilitation effect on word acquisition."
+#| echo: FALSE
+#| out-width: 100%
+#| fig-width: 6
+#| fig-height: 5
+knitr::include_graphics(normalizePath("assets/img/diagram.png"))
+```
+
+This hypothesis provides a convenient account for the results in Floccia et al.
+The size of the effect of linguistic similarity on vocabulary size that Floccia et al.
 reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure).
@@
 Therefore, participants may have on average learned the English word-form of translation equivalents earlier than the word-form in the additional language.
-Under this hypothesis, because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
+Because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants might then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
 
-We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a words *only* once its translation has been already acquired.
+We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a word *only* once its translation has been already acquired.
 We collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
 Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees.
-We adopted a Bayesian explanatory item response theory  [*IRT*, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).
+We adopted a Bayesian explanatory item response theory  [IRT, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word-form, while adjusting for the the number of phonemes in the word-form (an indicator of word acquisition difficulty).
 
-We operationalised participants' exposure rate to a word-form using a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it), and child's amount of exposure to the language such word belongs to.
+To operationalise participants' exposure rate to a word-form, we used a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it), and child's amount of exposure to the language such word belongs to.
 This measure was calculated by multiplying the word-form's lexical frequency with participant's degree of exposure to the language the word belongs to.
-Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childes] (see Appendix 1 for more details).
+Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childes] (see Appendix A for more details).
 Participants' degree of exposure were reported by caregivers in a language exposure questionnaire they filled before the vocabulary checklist.
-Caregivers provided a proportion of the time the child was estimated to listen to any language they were spoken to regularly (described in @sec-methods).
+Caregivers provided a proportion of the time the child was estimated to listen to any language they were spoken to regularly (described in the methods section).
 The resulting measure is a exposure-weighted lexical frequency measure, which was included in analyses as a proxy of how often bilingual participants were exposed to each word-form.
 
-We defined *cognateness* as a continuous measure of phonological similarity between translation equivalents.
-We quantified the phonological similarity between each pair of word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonological transcriptions [@levenshtein1966binary; @wells1995computer].
+*Cognateness* was defined as a continuous measure of phonological similarity between translation equivalents.
+We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-XSAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computer].
 This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
-This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]`) and *pájaro* (`[paxa4o]`), Catalan and Spanish for 'bird', and returns 50% for *lluna* (`[Lun5]`) and *luna* (`[luna]`), Catalan and Spanish for 'moon' (see Appendix 1 for more details).
+This results in a proportion that indicates how much the two phonetic transcriptions of the translation equivalent are similar to each other, ranging from zero (no similarity at all) to one (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
+For instance, this measure of phonological similarity returns 0% for *ocell* /`uceL`/ and *pájaro* /`paxa4o`/, Catalan and Spanish for 'bird', and returns 50% for *lluna* /`Lun5`/ and *luna* /`luna`/, Catalan and Spanish for 'moon' (see Appendix A for more details).
+
+Finally, we included word length as a nuance predictor in our analyses, motivated by previous studies reporting a strong association between word length and vocabulary size estimates [e.g., @braginsky2019consistency; @jones2019children].
+Sorter word-forms like /\textipa{sO\textltilde}/ (Catalan for *sun*) are generally simpler to encode as phonological representations in the lexicon than longer word-forms like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*), and are therefore more likely to be participant a child's vocabulary.
 
@@
 To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire.
-This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formR platform [@arslan2020formr].
+This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formr platform [@arslan2020formr].
 This questionnaire is structured in three blocks: (1) a language exposure questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
@@
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
-Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for 'dog').
+Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for 'dog') (see @tbl-items for a summary of the questionnaire items).
 
-```{r}
+```{r tbl-items}
 #| label: tbl-items
@@
 
-```{r}
+```{r participants}
 #| label: participants
@@
     arrange(id, time_stamp) |> 
-    mutate(diff_lag = time_stamp-lag(time_stamp, default = NA)) |> 
+    mutate(diff_lag = time_stamp-dplyr::lag(time_stamp, default = NA)) |> 
     summarise(diff_min = min(diff_lag, na.rm = TRUE),
@@
 
+n_trilingual <- participants |> 
+    select(starts_with("doe_")) |> 
+    mutate(doe_total = rowSums(across(everything())),
+           is_trilingual = doe_total < 1) |> 
+    add_count(name = "n_total") |> 
+    count(is_trilingual, n_total) |> 
+    mutate(prop = n/n_total)
+
 participants_edu <- table(participants$edu_parent)
@@
 Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`, and were always allocated to the same questionnaire list (A, B, C, or D).
-Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample).
+Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants) for a more detailed description of the sample).
+In total, `r n_trilingual$n[2]` participants (`r percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed to a third language other than Catalan and Spanish.
 All families provided informed consent before participating.
@@
 
-```{r}
+```{r tbl-participants}
 #| label: tbl-participants
@@
 #| warning: false
-#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan.
-Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
-In this table, a 100% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
-A 0% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
-A 50% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
-The 100%, 0%, and 50% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
-For illustration purposes this table, DoEs were binned into 25%-wide bins, and ages (in months) binned into 4 month-wide bins.
+#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan."
+#| tbl-note: "Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
+In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
+A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
+A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
+The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
+For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins.
 Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
@@
     sub_missing(everything(), everything(), missing_text = "--") |>
-    cols_label(doe_catalan = "DoE Catalan") |>
+    cols_label(doe_catalan = "Catalan exposure") |>
     tab_spanner("Age (months)", 2:8) |>
-    tab_footnote(
-        "This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10% of the time",
-        cells_column_labels(columns = doe_catalan)
-    ) |>
     tab_style(cell_text(align = "left"),
@@
 
-We used the highest self-reported educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES)This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: sense escolaritzar/sin escolarizar [no education], educació primària/educación primaria [primary school], educació secundària/educación secundaria [secondary school], batxillerat/bachillerato [complementary studies/high school], cicles formatius/ciclos formativos [vocational training], and educació universitària/educación universitaria [university degree].
+We used the highest self-reported educational attainment of parents or caregivers as a proxy of participants' socio-economic status (SES).
+This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: *sense escolaritzar/sin escolarizar* [no education], *educació primària/educación primaria* [primary school], *educació secundària/educación secundaria* [secondary school], *batxillerat/bachillerato* [complementary studies/high school], *cicles formatius/ciclos formativos* [vocational training], and *educació universitària/educación universitaria* [university degree].
 Most families reported university studies (`r participants_edu["University"]`, `r percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
@@
 
-```{r}
+```{r sample-sizes}
 #| label: sample-sizes
@@
                   .sep = ".")
-
 ```
@@
 This resulted in a final list of `r sum(n_items_final)` items, corresponding to words `r n_items_final["Catalan"]` Catalan words and their `r n_items_final["Spanish"]` Spanish translations (`r n_te` translation equivalents).
-After collecting participants responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item.
+After collecting participants' responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item.
 Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
 
-We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_4$fit)[2]` sampling chains with `r format(dim(model_fit_4$fit)[1], big.mark = ",")` iterations each).
-This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$ (number of months elapsed between participants' birth date and questionnaire completion), $Length$ (number of phonemes in the X-SAMPA
-phonological transcription of the word-form), $Exposure$ (composite measure of lexical frequency and language degree of exposure), and $Levenshtein$ (phonological similarity between translation equivalents).
-For an easier interpretation of the regression coefficients, the values of all predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
-We also included participant and item random intercepts and slopes when appropriate [@barr2013random]
+We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_0$fit)[2]` sampling chains with `r format(dim(model_fit_0$fit)[1], big.mark = ",")` iterations each).
+This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $\text{Age}$ (number of months elapsed between participants' birth date and questionnaire completion), $\text{Length}$ (number of phonemes in the X-SAMPA phonological transcription of the word-form), $\text{Exposure}$ (composite measure of lexical frequency and language degree of exposure), and $\text{Cognateness}$ (a measure of cognateness, defined as the phonological similarity between translation equivalents).
+Phonological transcriptions were manually generated.
+Catalan word-forms were transcribed to central Catalan phonology, and Spanish translations were transcribed to  Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
+Participant- and item-level random intercepts and slopes were included when appropriate, according to the structure of the data [@barr2013random].
+We specified a weakly informative prior around the parameters of the model (see Appendix C for details).
 
-We compared a baseline model $\mathcal{M}_{0}$, which only included the intercept, $Age$, $Length$, and $Exposure$ as predictors, with increasingly more complex models, introducing predictors in the following order: $Age \times Exposure$ ($\mathcal{M}_{1}$), $Levenshtein$ ($\mathcal{M}_{2}$), $Exposure \times Levenshtein$ ($\mathcal{M}_{3}$), and finally $Age \times Levenshtein$ and $Age \times Exposure \times Levenshtein$ ($\mathcal{M}_{4}$).
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO, @vehtari2017practical]^[Due to computational constraints, we performed LOO based on a random sub-sample of `r format(dim(model_log_liks$model_fit_4)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]], and selected for interpretation the one showing the best performance.
-More details about the models can be found in Appendix 3.
+$$
+\begin{aligned}
+&\textbf{Response}~k~\textbf{to word}~i~ \textbf{by participant}~j \\
+\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}), \\ &\text{where}~k \in \{\text{No, Understands, Understands and Says}\}\\
+&\textbf{Distribution parameters} \\
+\theta_{k_{ij}} &= (\beta_{0_{k}} + u_{0_{k_{i}}} + w_{0_{k_{j}}}) + (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + \\
+& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + (\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{i} + (\beta_{5} + u_{5_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{i}) & \\
+u_{0-5_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}); ~w_{0-3_{j}} \sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) \\
+&\textbf{Prior} \\
+\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.5); ~\beta_{1-5} \sim \mathcal{N}(0, 1) \\
+\sigma_{u_{0-5}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2)
+\end{aligned}
+$$
+
+We built a base model $\mathcal{M}_{0}$ that only included the intercept and the main effects of $\text{Age}$, $\text{Length}$, and $\text{Exposure}$ as predictors, and compared it against increasingly more complex models.
+$\mathcal{M}_{1}$ extended the base model by adding the main effect of $\text{Cognateness}$, and $\mathcal{M}_{2}$ added the two-way interaction $\text{Exposure} \times \text{Cognateness}$.
+We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO-CV, @vehtari2017practical]^[Due to computational constraints, we performed LOO-CV based on a random sub-sample of `r format(dim(model_log_liks$model_fit_0)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]].
+LOO-CV involves computing the likelihood of each data point, given a model  fitted excluding such data point.
+The likelihoods generated by the model for each of the data points are summarised together in a measure of predictive performance: the expected log-predicted density (ELPD).
+This measure penalises the complexity of the model ELPD.
+ELPD values closer to zero indicate better predictive accuracy.
+We compared the ELPD of each model, and selected for interpretation the one showing the best performance (see Appendix C for more details about the models).
 
 We assessed the practical relevance of the estimated regression coefficients of the selected model following @kruschke2018bayesian.
-First, we specified a region of practical equivalence (*ROPE*) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`.
+First, we specified a region of practical equivalence (ROPE) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`, in the probability scale.
 This region indicates the range of values that we considered equivalent to zero.
-The degree of overlap between the posterior distribution of a regression coefficient with the *ROPE* can be interpreted as support for the true parameter of the coefficient being zero or equivalent.
-Then we computed the 95% highest density interval (*HDI*) of the posterior distribution of each regression coefficient.
-This is interval contains the true value of this coefficient with 95% probability, given the data.
-Finally, we calculated the proportion of posterior samples in the 95% *HDI* that fell into the *ROPE*, which indicates the probability that the true value of the regression coefficient falls into the *ROPE*, and should therefore be considered equivalent to zero.
-For example, a 90% overlap between the *HDI* of a coefficient's posterior distribution and the *ROPE* indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero.
+The degree of overlap between the posterior distribution of a regression coefficient with the  can be interpreted as support for the true parameter of the coefficient being zero or equivalent.
+We then computed the 95% highest density interval (HDI) of the posterior distribution of each regression coefficient.
+This interval contains the true value of this coefficient with 95% probability, given the data.
+Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, which indicates the probability that the true value of the regression coefficient falls into the ROPE, and should therefore be considered equivalent to zero.
+For example, a 90% overlap between the HDI of a coefficient's posterior distribution and the  indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the ROPE, and is therefore equivalent to zero.
 
-Data processing and visualisation was done in R [@rcoreteam2013language, `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
+Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
 Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@vehtari2022loo], `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
@@
 
-```{r}
+```{r results-values}
 #| label: results-values
@@
 
-loo_diff <- loo_compare(model_loos)[2, 1]
-loo_diff_se <- loo_compare(model_loos)[2, 2]
+loo_diff <- loo_compare(model_loos)[, 1]
+loo_diff_se <- loo_compare(model_loos)[, 2]
 
-post_draws_list <- split(posterior_draws$summary, posterior_draws$summary$.variable)
+post_draws_list <- posterior_draws$summary |> 
+    mutate(across(.median:.lower,
+                  \(x) ifelse(grepl("Intercept", .type),
+                              plogis(x), x/4))) |> 
+    split(posterior_draws$summary$.variable)
 ```
 
-All models showed adequate convergence diagnostics and little evidence of multicollinearity between predictors (see Appendix 4 for details).
-The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $Exposure$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the second best-performing model ($\mathcal{M}_{3}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`.
-Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see @tbl-results-loos for all model comparisons).
+All models showed adequate convergence diagnostics and little evidence of multicollinearity between predictors (see Appendix D for details).
+$\mathcal{M}_{2}$, which included the $\text{Exposure} \times \text{Cognateness}$ two-way interaction, showed the best predictive performance with a difference in expected log-predicted density (ELPD) compared to the second best-performing model ($\mathcal{M}_{1}$) of `r round(loo_diff[2], 2)` $\pm$ `r round(loo_diff_se[2], 2)`.
+$\mathcal{M_1}$ performed substantially better than $\mathcal{M_0}$, with a difference of `r round(loo_diff[3], 2)` $\pm$ `r round(loo_diff_se[3], 2)` This points to the effect of $\text{Cognateness}$ being strongly modulated by $\text{Exposure}$, and that this interaction remains relatively stable across ages.
 
-```{r}
-#| label: tbl-results-loos
+```{r tbl-results-loos}
+#| label: tbl-loos
 #| echo: false
@@
 #| eval: true
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
-The predictive accuracy of the models was compared using leave-one-out cross-validation.
-The likelihood of each data point given a model fitted without including that same observation was estimated.
-The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*).
-The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
+#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)."
+#| tbl-note: "ELPD (*SD*): expected log-predictive density and standard error.
+Diff.
+(*SD*): diference in ELPD between each model and the model with the best predictive accuracy (Model 3) and standard error of the difference.
+Pareto-k estimates of all models were acceptable (k < 0.5)."
 model_loos |>
@@
     rownames_to_column("model") |>
-    mutate(model = str_remove(model, "model_fit_")) |>
-    relocate(model,
-             matches("elpd_loo"),
-             matches("diff"),
-             matches("p_loo"),
-             matches("looic")) |>
-    gt() |>
-    fmt_number(2:9, decimals = 1) |>
-    cols_merge_uncert(elpd_loo, se_elpd_loo) |> 
-    cols_merge_uncert(p_loo, se_p_loo) |> 
-    cols_merge_uncert(looic, se_looic) |> 
-    cols_merge_uncert(elpd_diff, se_diff) |> 
+    mutate(model = paste0("Model ", str_remove(model, "model_fit_"))) |>
+    select(model, elpd_loo, se_elpd_loo, elpd_diff, se_diff) |> 
+    gt(rowname_col = "model") |>
+    fmt_number(2:5, decimals = 2, sep_mark = ",") |>
+    cols_merge(c(elpd_loo, se_elpd_loo), pattern = "{1} ({2})") |> 
+    cols_merge(c(elpd_diff, se_diff), pattern = "{1} ({2})") |> 
     cols_label(
         model = "Model",
-        elpd_loo = "ELPD ± SE",
-        p_loo = "p ± SE",
-        looic = "IC ± SE",
-        elpd_diff = "ELPD diff. ± SE"
+        elpd_loo = "ELPD (SD)",
+        elpd_diff = "Diff. (SD)"
     ) |>
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
-    tab_style(cell_text(weight = "bold"),
+    tab_style(cell_text(weight = "bold", style = "italic"),
               cells_column_labels()) |>
@@
 
-@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$, and their degree of overlap with the ROPE.
+@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{2}$, and their degree of overlap with the ROPE.
 For interpretability, we report the estimated regression coefficients transformed to the probability scale.
@@
 
-```{r}
+```{r tbl-coefs}
 #| label: tbl-coefs
-#| tbl-cap: "Posterior distribution of regression coefficients in the probability scale.
-*Median*: median of the posterior distribution in the logit scale.
-*95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE).
+#| tbl-cap: "Posterior distribution of regression coefficients Model 3."
+#| tbl-note: "Median: median of the posterior distribution in the probability scale.
+95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.
+On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE).
 This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
@@
 #| warning: false
+#| out-width: 80%
 # summarise posterior draws
+intercept_label <- "Intercepts (at 22 months)"
+
 posterior_draws$summary |>
     select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
-    mutate(
-        across(.median:.upper,
-               function(x) {
-                   ifelse(
-                       .type=="Intercepts (at 22 months)",
-                       plogis(x),
-                       x/4
-                   )
-               }
-        )) |> 
+    mutate(across(.median:.upper,
+                  \(x) ifelse(.type==intercept_label, plogis(x), x/4))) |> 
     gt(groupname_col = ".type", 
        rowname_col = ".variable_name") |>
-    fmt_percent(c(.median, .rope_overlap)) |>
-    fmt_number(.lower:.upper, scale_by = 100) |> 
+    fmt_percent(.rope_overlap, decimals = 3) |> 
+    fmt_number(c(.median:.upper), decimals = 3) |>
     cols_merge(c(.lower, .upper),
@@
         .median = "Median",
-        .lower = md("95% *HDI*"),
-        .rope_overlap = md("*ROPE* prob.")
+        .lower = md("95% HDI"),
+        .rope_overlap = md("ROPE prob.")
     ) |>
@@
 
-The word exposure index ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$.upper, accuracy = 0.01)`]).
-All of the posterior samples of this regression coefficient excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), accuracy = 0.01)` more likely to acquire it.
+The main effect of $\text{Age}$ showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std"]]$.upper, 2)`]), with all of its posterior samples falling out of the ROPE.
+A one-month increment in age increased a maximum of `r round(post_draws_list[["b_age_std"]]$.median/sd(responses$age), 2)` the probability of acquisition.
+Similarly, the word exposure index ($\text{Exposure}$) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 2)`]).
+All of the posterior samples of this regression coefficient excluded the ROPE.
+The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), 2)` more likely to acquire it.
+Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r round(post_draws_list[["b_n_phon_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_phon_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_phon_std"]]$.upper, 2)`]).
+For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.median)/sd(responses$n_phon), 2)` less likely to know it.
+The 95% HDI of the regression coefficient of the $\text{Age} \times \text{Exposure}$ interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 2)`]), showing that the effect of the word exposure index differed across ages: older children were more sensitive likely to acquire words with higher exposure rate than younger children.
 
-The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$.rope_overlap, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
+The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).
+For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), 2)`.
+The effect of $\text{Cognateness}$ interacted with that of $\text{Exposure}$: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 2)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
+Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
+This effect was negligible when for words with median or high exposure (+1 SD) (see @fig-marginal).
 
-However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to word.
-Follow-up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to word was low (e.g., -1 SD), phonological similarity increased the probability of acquisition substantially.
-This effect was negligible when for words with median or high exposure (+1 SD).
-
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of the word exposure index on the speed at which participants acquired words across ages was inconclusive: given our data, it is difficult to tell whether $Exposure$ increased or decreased the slope of the words' acquisition trajectories.
-
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$.upper, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the impact of the word exposure index on the effect of phonological similarity did not differ substantially across ages.
-
-```{r}
-#| label: fig-results-marginal
+```{r fig-marginal}
+#| label: fig-marginal
 #| echo: false
@@
 #| fig-width: 7
-#| fig-height: 5
+#| fig-height: 4.5
+#| out-width: 100%
 #| eval: true
-#| fig-cap: "Expected posterior predictions.
-Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
-These values correspond to samples of the posterior distribution that were drawn during the MCMC estimation of the model.
-We generated 25 posterior predictions for a series of combination of levels of interest.
-Each thin line and corresponds to one prediction for one set of predictor values.
-Thicker lines indicate the median posterior prediction.
-The X-axis indicates the age (in months) for which the prediction is generated.
-The Y-axis indicates the predicted probability of acquisition.
-Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
+#| fig-cap: "Posterior marginal effects of Model 3.
+Thin lines correspond to 50 mean predictions.
+Thicker lines indicate the median of those predictions.
+Different colours indicate different levels of cognateness (phonological similarity).
 Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
-Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom.
+Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom.
 In-sample predictions lie inside the grey rectangles."
-
 clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")
@@
 marginal_effects_epreds |> 
-    mutate(
-        age = rescale_variable(age_std, 
+    mutate(age = rescale_variable(age_std, 
                                   mean(responses$age), 
                                   sd(responses$age)),
-        exposure = factor(
-            exposure_std,
+           exposure = factor(exposure_std,
                              levels = unique(exposure_std),
@@
                                         "High exposure (+1 SD)"),
-            ordered = TRUE
-        ),
-        lv = factor(
-            lv_std,
+                             ordered = TRUE),
+           lv = factor(lv_std,
                        levels = unique(lv_std),
-            labels = paste0(c(0, 50, 100), "% similarity"),
-            ordered = TRUE
-        ),
+                       labels = paste0(c(0, 50, 100), "%"),
+                       ordered = TRUE),
            .category = ifelse(.category=="Understands",
                               "Comprehension",
-                           "Production")
-    ) |> 
+                              "Production")) |> 
     ggplot(aes(age, .value, colour = lv, fill = lv)) +
     facet_grid(.category~exposure) +
-    # stat_summary(fun.data = median_hdi, geom = "ribbon",
-    #              alpha = 0.5, colour = NA) +
     annotate(geom = "rect",
@@
              ymax = Inf,
-             fill = "grey90",
+             fill = "grey95",
              colour = NA) +
     annotate(geom = "text",
-             x = min(responses$age),
-             y = 1,
-             hjust = -0.075,
-             vjust = 1,
-             label = "Observed\nage range",
-             colour = "grey40",
+             x = mean(responses$age),
+             y = 0,
+             hjust = -0.1,
+             vjust = 0,
+             label = "Observed age range",
              size = 2) +
-    geom_vline(xintercept = mean(responses$age), colour = "grey",
-               linewidth = 1) +
+    geom_vline(xintercept = mean(responses$age), 
+               linetype = "dotted") +
     geom_hline(yintercept = 0.5,
-               colour = "grey",
-               linewidth = 1) +
+               linetype = "dotted") +
     geom_line(aes(group = interaction(lv, .draw)),
-              linewidth = 0.75, alpha = 0.15) +
-    stat_summary(fun = median, geom = "line", linewidth = 1) +
-    labs(
-        x = "Age (months)",
+              linewidth = 0.5,
+              alpha = 1/10) +
+    stat_summary(fun = median, geom = "line", linewidth = 0.75) +
+    labs(x = "Age (months)",
          y = "Probability of acquisition",
-        linetype = "Phonological similarity (Levenshtein)",
-        colour = "Phonological similarity (Levenshtein)",
-        fill = "Phonological similarity (Levenshtein)",
-        subtitle = glue("**Phonological similarity**<br> <span style = 'color:{clrs_reorder[5]};'>0% similarity</span>   <span style = 'color:{clrs_reorder[4]};'>50% similarity</span>   <span style = 'color:{clrs_reorder[1]};'>100% similarity</span>")) +
-    scale_colour_manual(values = clrs_reorder[c(5, 4, 1)]) +
-    scale_fill_manual(values = clrs_reorder[c(5, 4, 1)]) +
-    scale_y_continuous(
-        breaks = seq(0, 1, 0.25),
-        limits = c(0, 1),
-        labels = percent
-    ) +
-    scale_x_continuous(breaks = seq(5, 50, 5),
-                       limits = c(10, 50)) +
-    theme(
-        legend.position = "none",
+         linetype = "Cognateness (phonological similarity)",
+         colour = "Cognateness (phonological similarity)",
+         fill = "Cognateness (phonological similarity)") +
+    guides(colour = guide_legend(override.aes = list(linewidth = 2))) +
+    scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_fill_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_y_continuous(breaks = seq(0, 1, 0.25),
+                       limits = c(0, 1)) +
+    scale_x_continuous(breaks = seq(0, 50, 5)) +
+    theme(legend.position = "top",
+          legend.key.size = unit(1, "cm"),
+          legend.key.height = unit(1, "cm"),
+          axis.text.x = element_text(size = 9),
           panel.grid = element_blank(),
@@
                                       colour = "black", 
-                                    linewidth = 0.75),
-        plot.subtitle = element_textbox_simple(
-            fill = "grey90",
-            padding = margin(8, 8, 8, 8),
-            margin = margin(0, 0, 10, 0)
-        ))
+                                      linewidth = 0.75))
 ```
+
+To rule out the possibility that cognateness facilitation effect we found was due to cognateness comprising more frequent syllables than non-cognates---and therefore not because of their cognate status itself---, we compared the syllabic frequency of cognates and non-cognates included in our analyses.
+To calculate syllable frequency, we first extracted all syllables embedded in the selected words.
+For each syllable, we summed the lexical frequency of all the words in which such syllable appeared.
+The resulting value provided an estimate of the number of times the syllable appears in child-directed speech, embedded within different words.
+Finally, for each word, we summed the frequency of its syllables, as an estimate of the syllabic frequency of the word.
+We fit a Bayesian model with cognateness (Levenshtein similarity) as response variable, and the main effects of syllable frequency and number of syllables (to control for the fact words with more syllables are more likely to score higher in syllabic frequency) as predictors.
+This model provided strong evidence for the association between cognateness and syllabic frequency being equivalent to zero (see Appendix E).
 
@@
 
-In this study, we investigated the role of cognateness on word acquisition.
-We contrasted two hypotheses regarding the word acquisition process at which the possible (facilitative) effect of cognateness might take place.
-First, we considered the possibility that cognateness boosts acquisition at early stages, before a stable, robust representation of either of the word-forms involved in the translation equivalent has been formed in the child's lexicon.
-This account predicts that the effect of cognateness would take place early in age, and would be symmetric across languages: one word-form boosts the acquisition of the other via their phonological similarity.
-If this hypotheses were supported by data, we would expect to observe and early difference in probability of comprehension and/or production between more similar and less similar translation equivalents (i.e., an effect of $Levenshtein$ in interaction with $Age$), with this difference being equivalent for the word-for to which the child is exposed to more frequently, and for the word-form to which the while is exposed less frequently (no interaction effect between $Exposure$ and $Levenshtein$).
+```{r discussion-examples}
+#| echo: false
+#| warning: false
+#| message: false
+#| out-width: 60%
+lv_distr <- items |> 
+    mutate(lv_bin = cut(lv, 
+                        breaks = seq(0, 1, 0.1),
+                        include.lowest = TRUE,
+                        labels = c("0-10%", "10-20%", "20-30%", "30-40%",
+                                   "40-50%", "50-60%", "60-70%", "70-80%",
+                                   "80-90%", "90-100%"))) |>
+    add_count(name = "n_total") |> 
+    count(lv_bin, n_total) |> 
+    mutate(prop = n/n_total)
 
-We also considered an alternative hypothesis: that the facilitation effect of cognateness takes place at later stages of word acquisition, when at least one of the word-forms of the translation equivalent hold a detailed, robust representation in the child's lexicon.
-Only then, the phonological similarity between the acquired word-form and its yet-to-be-acquired translation would boost the acquisition of the latter.
-This account predicts a larger effect of cognateness for word-forms to which the child is exposed to less frequently, and therefore are likely to be acquired later than their translation.
-This effect of cognateness would also be expected to take place at later ages.
-Data backing this account would support a model in which the effect of $Levenshtein$ interacts with $Exposure$, and in which possibly this interaction, or at least the effect of $Levenshtein$ interacts with age.
+E_e <- bvq_data$pool |>
+    mutate(item = gsub("cat_|spa_", "", item),
+           xsampa = gsub("\\.", "", xsampa)) |>
+    filter(te %in% items$te, 
+           item %in% items$item) |>
+    select(te, language, xsampa) |> 
+    pivot_wider(names_from = language,
+                values_from = xsampa, 
+                id_cols = te,
+                names_repair = janitor::make_clean_names) |>
+    filter(grepl("E", catalan), 
+           grepl("e", spanish)) |>
+    left_join(distinct(items, te, lv),
+              by = "te")
+```
 
-To test these hypotheses, we used an *ad-hoc* vocabulary checklist to collect parental reports of word acquisition status from children learning Catalan and/or Spanish, who were exposed to varying degrees of exposure to either language.
-We then used item response theory (*IRT*) to estimate the probability of each word-form being acquired by a child---either as comprehension or production---as a function of the child's $Age$, their $Exposure$ to the word-form (operationalised as a composite measure of lexical frequency and language exposure), and the phonological similarity of the word-form with its translation in the other language ($Levenshtein$).
-We also added the $Length$  of the word-form as a nuance predictor the increase the predictive performance of the model.
+This study investigated the impact of cognateness on the early bilingual lexicon.
+We used Bayesian Item Response Theory to model the acquisition trajectories of a large sample of Catalan and Spanish words, estimating the effect of cognateness on the probability of acquisition.
+This model corrected for participants' age, word-form length (number of phonemes), and a novel measure of participants' exposure rate to each word.
+Exposure rates were calculated as a language exposure-weighted lexical frequency score in which each word-form's lexical frequency was corrected by the degree to which the participant was exposed to each language.
+Overall, we found that cognates (i.e., phonologically similar translation equivalents) were acquired earlier than non-cognates.
+This effect was mediated by exposure rate: low-exposure word-forms benefited from their cognate status, whereas high-exposure word-forms did not.
+Capitalising on accumulator models of language acquisition, we provide a theoretical account of bilingual lexical acquisition in which parallel activation plays a central role during the consolidation of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
 
-Our results are, to a large extent, in line with previous studies.
-As it could be expected, we found that shorter word-forms were acquired at early ages than longer word-forms.
-For instance, $sol$ (Catalan for *sun*) was acquired earlier than *cocodril* (Catalan for *crocodile*).
-Shorter word-forms consist on fewer phonemes and, frequently, fewer syllables.
-This makes them simpler to encode as phonological representations in the lexicon, and therefore easier to process.
-Ultimately, these words are easier to produce, and therefore to be reported as acquired in parental vocabulary checklists.
-Previous studies found word length to be more strongly associated with receptive vocabulary than with expressive vocabulary [e.g., @braginsky2019consistency; @jones2019children].
+This investigation is particularly relevant in the light of two previous findings.
+First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (e.g.
+shared many cognates, like., English-German) showed larger vocabulary sizes than those learning typologically distant languages (e.g.
+shared fewer cognates, like English-Mandarin).
+Second, Mitchell et al.
+found an earlier age-of-acquisition for cognates, compared to non-cognates.
+The outcomes of both studies pointed to cognateness facilitating word acquisition through parallel activation, that is, the co-activation of lexical representations from the two languages during word comprehension and production.
+But the underpinnings of such effect were unclear: while parallel activation has been extensively described in experimental studies, current paradigms of bilingual word acquisition and word learning are, to a large extent, dissociated from the mechanisms proposed by the former.
+Accumulator models of language acquisition provide a convenient theoretical framework to narrow this gap.
 
-According to previous literature, we also anticipated that words to which children were more frequently exposed would be acquired earlier [e.g., @braginsky2019consistency].
-In line with accumulator models of word acquisition, the age of acquisition of a word is a function of the child's rate of exposure to the word, which conspires with child- and word-level characteristics to the formation of a representation of the word in the child's lexicon [@mollica2017how; @hidaka2013computational].
-The amount of learning instances that a child is given to learn a word is given, to a large extent, by the lexical frequency of such word in the child's language.
-Words with higher lexical frequency are more likely to be uttered in the presence of the child, while words with lower lexical frequency appear less often.
-In the case of bilinguals, the exposure rate to a given word not only depends on lexical frequency, but also on the child's amount of exposure to the language such word belongs to [@cattani2014how; @floccia2018results].
-For instance, a Spanish dominant bilingual child who is exposed to Catalan only 10% of the time will rarely hear the Catalan word *casa*, despite its high lexical frequency.
-We estimated the exposure rate of the words in our list by calculating a language exposure-weighted measure of lexical frequency, which adjusts the lexical frequency of the word by the child's amount of exposure to the language such word belongs to.
-We found a strong association between this measure of exposure rate and word's age of acquisition.
-Interestingly, this predictor had a stronger role during earlier ages: younger infants benefited from higher exposure rates at earlier ages, suggesting that early word acquisition is supported by mere exposure to word forms more strongly than later word acquisition.
+Following the rationale of accumulator models, we suggest that parallel activation increases the rate at which bilingual children encounter learning instances for cognate words, but not for non-cognate words.
+We hypothesised that when a bilingual child is exposed to a word-form in presence of its referent, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
+This cross-language activation occurs at the phonological level, and the amount of co-activation that spreads from the spoken word to its translation is proportional to the amount of phonological similarity between both word-forms.
+Cognates would receive more activation from their translation than non-cognates, leading children to accumulate learning instances with cognate words at a faster rate than with non-cognate words.
+As a result, lexical representations of cognate words would consolidate at earlier ages than those of non-cognate words.
 
-Finally, we found a facilitation effect of cognateness, providing converging evidence with previous studies [@floccia2018results; @mitchell2022cognates; @bosch2014first; @schelletter2002effect].
-Critically, this effect was mediated by exposure: only words to which children were exposed to less frequently benefited from their cognate status.
-These results favour the account that a word-form benefits from its phonological similarity with its translation only after such translation has been acquired first.
-On the one hand, our results provide converging evidence for @floccia2018results, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample.
-Most children in their sample were English-dominant, and therefore their exposure to English was higher than to the additional language.
-However, it was in the additional language vocabulary where they found a facilitation effect of cognateness.
-Bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese).
-Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018results's sample.
+Consider a Catalan-Spanish bilingual child who is exposed to the Catalan word-form /\textipa{"gat}/ in the presence of a cat.
+This represents a learning instance, in which the child has the opportunity to strengthen the association between the word-form and its referent.
+Because the Spanish translation of this word (/\textipa{"ga.to}/) is phonologically very similar to the Catalan form (both word-forms share `r format(stringdist::stringsim("gat", "gato"), nsmall = 2)` Levenshtein similarity), it might receive some degree of co-activation.
+Both members of the translation pair would be activated to some extent in the presence of their common referent, and therefore the child would accumulate a learning instance with both word-forms.
+The case of the non-cognate translation equivalent for *dog* (/\textipa{"gOs}/-/\textipa{"pe.ro}/,  sharing `r format(stringdist::stringsim("gOs", "pero"), nsmall = 2)` Levenshtein similarity), would be different.
+Since both words hardly share any phonological similarity, a bilingual child would only accumulate learning instances for the word they encounter in speech, but not for the other.
+For instance, hearing /\textipa{"gOs}/ would be unlikely to activate /\textipa{"pe.ro}/ through phonology.
+Each word would accumulate learning instances independently, without any cross-language facilitation taking place.
 
-In line with the hypothesis that received more support from our data, we predicted an interaction effect between participants' age, cognateness, and word exposure, since we expected cognateness to play a more important role at later stages of word learning, when at least one of the members of the translation equivalent has been acquired, and therefore its counterpart in the other language can benefit from its cognate status.
-We found evidence supporting that the true value of this interaction is close to zero.
-Finding such an effect would provide further evidence that cognateness benefits word acquisition only after the translation of such word is acquired.
-However, the absence of such interaction does not preclude a more careful endorsement
-of this hypothesis.
-It is possible that, even if the cognateness effect plays a role if the acquisition of the word-form with the least exposure of the translation equivalent, age itself might not modulate the size of this effect: cognateness plays a similar role across ages.
+This account of bilingual lexical acquisition addresses a critical subject in bilingualism research: do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
+In the context of lexical acquisition, the former scenario predicts that every learning instance for a given word-form contributes to the consolidation of the representation of such word in the lexicon, while the consolidation of its translation remains unaffected by such experience.
+In the latter scenario, a learning instance to the same word-form would contribute not only to the consolidation of the representation of such word, but also, to some extent, to the consolidation of its translation.
+Our findings provide strong support for an account of bilingual vocabulary growth in which the experience and learning outcomes accumulated by the child in one language impact those in the other language.
+Such a facilitatory cross-language mechanism might be an important piece in the puzzle of bilingual language acquisition.
+In particular, it may shed some light on why bilingual infants do not show relevant delays in language acquisition milestones compared to their monolingual peers, while receiving a lower quantity of speech input in each of their languages.
+Our results provide some insights into this issue: infants benefitted more strongly from the cognateness facilitation effect when acquiring words from the language of lower exposure than in the language of higher exposure.
 
-This study has several limitations.
-One of them is the fact that the languages involved are typologically very close.
-Catalan and Spanish are both Romance languages, share a considerable amount of cognates, and very often co-exist in the same social spaces (families, schools, working place, etc.).
-Although there are no populations of bilinguals with identical sociolinguistic characteristics, and therefore it would be pointless to compare the sample in this study to any "standard" population of bilinguals, the generalisability of our results should be taken with a grain of salt.
-It is possible, for instance, that the large amount of cognates that Catalan-Spanish bilingual children are exposed leads them to learn at early stages of word acquisition that words tend to sound similar in both languages, and therefore to be more keen on assuming that similar sounding words are semantically related.
-Children exposed to a pair of languages sharing a fewer amount of cognates might not make the same assumption, even when the word-forms they hear in the presence of the same referents are very similar at the phonological level.
+We suggest that this asymmetry is the result of children being exposed to words in the higher-exposure language more often than to words in the lower-exposure language.
+A bilingual child's frequency of exposure to a given word-form is mostly determined by two factors: the word-form's lexical frequency and the child's amount of language exposure to the language such word belongs to.
+A dual linguistic input means lower exposure to each of the languages, unless one makes the---arguably implausible---assumption that bilinguals are exposed to larger quantities of linguistic input.
+In this scenario, bilinguals encounter each word on lower rates than their monolingual peers.
+Because of this difference in exposure, words lower-exposure language might receive activation from their translation in the higher-exposure language more often than words from the higher-exposure language receive activation from their translation in the lower-exposure language.
+As a result, the cognateness facilitation effect should be stronger in words from the lower-exposure language.
+For instance, in the case of a Catalan-dominant, Catalan-Spanish child, the Catalan word /\textipa{"gat}/ would be learned earlier than its Spanish translation /\textipa{"ga.to}/.
+Therefore, /\textipa{"ga.to}/ should be more likely to receive facilitation from /\textipa{"'pe.ro'gat}/ than vice versa (see @fig-hypotheses in the introduction for a graphic illustration).
 
-Another limitation is the predominantly cross-sectional design of our data set.
-We collected vocabulary data from children aged 12 to 32 months, a critical period in development in which the foundations of a lexicon get established.
-We estimated word acquisition curves by modelling the probability of comprehension and production across ages.
-However, most participants provided data at only one age point.
-Effectively, we are making longitudinal inferences based on cross-sectional data, and drawing conclusions about the development of a single participant across multiple age points using data from different participants at individual age points.
-For this reason, and also because of the observational nature of our study, causal inferences derived from our analyses are unwarranted [@kraemer2000how; @kesmodel2018crosssectional].
-One counterpoint to this shortcoming is that, while data our data is cross-sectional if participants are considered as the observational unit, data about word-forms/translation equivalents is longitudinal.
-The same translation equivalent is responded to at multiple ages.
-Since translation equivalents are our main observational unit of interest over which we primarily conduct statistical inference (e.g., estimating the size of the cognateness-effect), this provides some further justification for our conclusions.
+We tested this hypothesis by combining language exposure and lexical frequency into a composite variable---, which we coined exposure rate---to obtain an estimate of how often each participant encountered each word.
+We introduced the exposure rate as a predictor in our model, and found a substantial interaction between this predictor and cognateness.
+In line with our predictions, only words-forms to which children were exposed to less frequently benefited from their cognate status.
+This mechanism might be extended to provide a plausible explanation for the language similarity facilitation reported by Floccia et al.
+The authors observed a facilitation in the additional (non-English) language: children learning two typologically close languages knew more words in the additional language than those learning two typologically more distant languages.
+In their sample, the additional language was consistently also the lower-exposure language for most children, while English was the higher-exposure language.
+Given that words in English were more likely to be acquired first, higher phonological overlap for words in the language of lower exposure (especially those of lower lexical frequency) would facilitate vocabulary growth for languages sharing more cognates with English.
 
-Future studies addressing this subject may benefit for a more adequate design oriented towards making causal inferences.
-One such design can be partially satisfied by (1) collecting a more substantial longitudinal dataset, in which the same participant provides data points for each translation equivalent at multiple ages, and (2) testing the role of cognateness and word exposure in an experimental context, in which such variables can be orthogonally manipulated or randomised.
-This second possibility would also allow to explore an interpretation of the results obtained in the context of *word learning*, which is not possible in the present study.
-Another consideration for future investigations is the involvement of a more diverse range of language pairs as in @braginsky2019consistency or @floccia2018introduction to improve the generalisability of the resulting inferences across more populations of bilinguals.
+It might be argued that our results reflect the fact that cognates are represented in the initial bilingual lexicon as the *same* lexical entry.
+Because bilingual children are exposed to such similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the presence of a cat), it is possible that they classify both are as acceptable variations of the same word, therefore treating them as a single lexical item, without parallel activation taking place.
+This could explain a cognate advantage in word acquisition without the need for parallel activation playing a relevant role.
+This account would not necessarily imply that bilinguals are insensitive to the phonological mismatches between cognate word-forms.
+On the contrary, there is strong evidence that both monolingual and bilingual infants are highly sensitive to subtle mispronunciations of word-forms [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels].
+The assimilation of both word-forms in the lexicon might occur at a higher level, where both word-forms are considered referentially equivalent, so that every exposure to either of the (distinguishable) word-forms of the translation equivalent is counted as a learning instance for both.
+This would lead to a faster increase in cumulative learning instances across ages, leading to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
+This mechanism could potentially explain the earlier age-of-acquisition effect of cognates found in the present study, without the need of parallel activation playing any relevant role.
 
-## Conclusions
+Previous work points in a different direction, with bilingual toddlers considering cognates as separate words.
+For instance, @ramon-casas2009vowel showed that 18-26 months-old bilingual toddlers reject mispronounced labels in word recognition tasks, although only when such mispronunciation involves a contrast that is phonemic in both languages.
+In their study, Catalan-Spanish bilinguals treated similarly the words /\textipa{"pe\textesh}/ and its mispronunciated form /\textipa{"pE\textesh}/, which involved the /\textipa{e}-\textipa{E}/ contrast (only present in Catalan), but treated differently the words /\textipa{pe\textesh}/ and /\textipa{pi\textesh}/, which involved the /\textipa{e}-\textipa{i}/ contrast (present in both languages).
+Moreover, a later study by @ramon2010non showed that bilinguals treat the language-specific /\textipa{e}-\textipa{E}/ contrast differently when embedded within *non-cognate* word-forms.
+These results suggest that, except for the acoustically close, language-specific /\textipa{e}-\textipa{E}/ contrast, bilinguals treat similar-sounding words differently in word recognition tasks, especially when both word-forms are separated by more than one phoneme.
+In the light of these findings, the possibility that bilinguals merge cognates into shared lexical representations becomes more implausible.
 
-In summary, our study has replicated previous studies and provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
+This is especially true for the case of the Catalan-Spanish bilinguals in our sample: there are few exemplars of Catalan-Spanish translation equivalents involving a /\textipa{e}-\textipa{E}/ substitution.
+Only `r percent(nrow(E_e)/length(unique(items$te)))` of the translation equivalents in our questionnaire involved a /\textipa{e}-\textipa{E}/ change between the Catalan and the Spanish items, and only one of them was a minimal pair: /\textipa{"tREN}/-/\textipa{"tReN}/ [*train*].
+Moreover, few words in our study differ only in one phoneme substitution.
+Only `r percent(lv_distr$prop[nrow(lv_distr)], accuracy = 0.01)` of the translation equivalents included in our analyses shared between 90% and 100% phonological similarity (measured as the Levenshtein similarity between the X-SAMPA transcriptions of the Catalan and Spanish form), and only `r percent(lv_distr$prop[nrow(lv_distr)-1], accuracy = 0.01)` shared between 80 and 90% similarity.
+This shows that the presence of identical and near-identical cognates in our study was very low, and therefore most translation equivalents differed in more than one phoneme.
+Finally, and perhaps more critically, there is evidence beyond word recognition that bilinguals treat minimal pairs differently at the referential level.
+For instance, @ramon-casas2017minimalpair found that bilingual toddlers are able to learn new minimal pairs involving the /\textipa{e}-\textipa{E}/ contrast.
+This suggests that toddlers' tolerance towards cognate mispronunciations at the referential level might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@mattock2010first; @havy2016phonetic].
+
+Our study shares similar methodological limitations with previous work using vocabulary reports provided by caregivers.
+Such reports can be subject to measurement error induced by caregivers who may sometimes overestimate or underestimate participants' true probability of acquisition of words [e.g., @houston-price2007discrepancy].
+For instance, although in the present study caregivers were explicitly instructed *not* to rely on their responses to Catalan words when responding to Spanish (and vice versa), it is possible that some caregivers assumed---at least to some extent---that because the child knew a word in one language, the child should also know the word in the other language (especially when both words are similar-sounding, i.e., cognates).
+Production estimates may be more prompt to such biases, in part because of the slower pace at which their articulatory abilities develop, compared to their word recognition abilities.
+For instance, even at 47 months of age, few (typically developing) English monolinguals reach 100% intelligibility scores in word production, as rated by English native adults [@hustad2020development; @hustad2021speech].
+This gap between comprehension and production is even larger in the less dominant language of bilingual children [@giguere2022bilingual].
+For this reason, caregivers may be more uncertain about what words can be counted as *acquired* in this modality.
+Despite these potential biases, vocabulary checklist filled by parents show strong evidence of concurrent validity with other estimates of vocabulary size or lexical processing [@killing2008move; @feldman2005concurrent; @gillen2021tapping; but see @houston-price2007discrepancy].
+
+The present study contributes a specific data point into the complex landscape of bilingualism research.
+Bilinguals are a remarkably heterogeneous population that can hardly be satisfactorily characterised in a comprehensive way [@sebastian-galles2020bilingual].
+Bilinguals differ across multiple dimensions.
+these differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Arabic, or high like the case of Spanish and Greek), to higher-level factors like the socio-political situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages may be less dominant than the other).
+This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.
+Our sample, although homogeneous (e.g., similar parental educational level across), represents a fairly particular population of bilinguals: the languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages, both languages are used in fairly similar contexts, and both languages are known by the majority of the population.
+In 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@de2013els].
+In addition, and as commented in the introduction, Catalan and Spanish are both Romance languages and share a considerable amount of cognates.
+Extending our analyses to other bilingual populations learning typologically more distant languages, and whose languages tend to be used in more distinct contexts (e.g., heritage languages) should be a natural future step for the present investigation.
+
+In summary, our study provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
 We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word's phonological similarity with its translation in the other language.
-We suggest that this effect might be explained by the need of a robust lexical representation of one of the translations for cognateness to play a role.
+Capitalising on accumulator models of language acquisition we, put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
 
@@
 :::
-
-{{< pagebreak >}}
-
-# Appendix 1: model predictors {.appendix}
-
-{{< include _model-predictors.qmd >}}
-
-{{< pagebreak >}}
-
-# Appendix 2: model details {.appendix}
-
-{{< include _model-details.qmd >}}
-
-{{< pagebreak >}}
-
-# Appendix 3: convergence diagnostics {.appendix}
-
-{{< include _model-diagnostics.qmd >}}
-
-{{< pagebreak >}}
-
-# Appendix 4: frequency and language exposure as separate predictors {.appendix}
-
-{{< include _model-doe.qmd >}}
-
-{{< pagebreak >}}

### 2023-05-19 manuscript/_appendix/_model-doe.qmd

--- 2023-04-17 appendix/_model-doe.qmd
@@
-```{r apafg-coefs-doe}
-#| label: apafg-coefs-doe
-#| apa-cap: "Posterior distribution of regression coefficients in the probability scale"
-#| apa-note: "Median: median of the posterior distribution in the logit scale.
+```{r fig-coefs-doe}
+#| label: fig-coefs-doe
+#| fig-cap: "Posterior distribution of regression coefficients in the probability scale"
+#| fig-note: "Median: median of the posterior distribution in the logit scale.
 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.

### 2023-05-19 manuscript/_appendix/_model-predictors.qmd

--- 2023-04-28 _appendix/_model-predictors.qmd
@@
 Our study capitalises in phonology, therefore we operationalised *cognateness* as the phonological similarity between a pair of translation equivalents.
-Our dataset contained `r max(n_items_final)` translation equivalents, each consisting in two word-forms: Catalan and Spanish (see {apatb-translation-levenshtein} for an example).
+Our dataset contained `r max(n_items_final)` translation equivalents, each consisting in two word-forms: Catalan and Spanish (see @tbl-translation-levenshtein for an example).
 Measuring the phonological similarity between two word forms is non trial.
@@
 
-The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure, see {apatb-translation-levenshtein} for more examples:
+The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure, see @tbl-translation-levenshtein for more examples:
 
-```{r apatb-translation-levenshtein}
-#| label: apatb-translation-levenshtein
-#| apa-cap: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study
+```{r tbl-translation-levenshtein}
+#| label: tbl-translation-levenshtein
+#| tbl-cap: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study
 tribble(~Catalan, ~Spanish, ~Levenshtein,

### 2023-05-23 manuscript/_appendix/_model-diagnostics.qmd
Add AoA estimator

@@
 Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined.
-We run `r dim(model_fit_3$fit)[2]` MCMC chains, each `r dim(model_fit_3$fit)[1]` iterations long each.
+We run `r dim(model_fit$fit)[2]` MCMC chains, each `r dim(model_fit$fit)[1]` iterations long each.
 The correct performance of this algorithm is critical to the quality of the statistical evidence to which the outcomes of the model lead.
@@
 #| label: tbl-rhats-neffs
-#| tbl-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
+#| tbl-cap: "MCMC convergence diagnostic of all parameters in the model"
 #| echo: false
@@
 #| out-width: 100%
-map_df(model_rhats, bind_rows, .id = "model") |> 
-    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
-    pivot_longer(-model, 
-                 names_to = ".variable",
-                 values_to = ".rhat") |>
-    drop_na() |> 
+model_convergence |> 
     ggplot(aes(.rhat)) +
-    facet_wrap(~model, nrow = 1) +
     geom_vline(xintercept = 1.01, linetype = "dashed") +
@@
 
-    map_df(model_neffs, bind_rows, .id = "model") |> 
-    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
-    pivot_longer(-model, 
-                 names_to = ".variable",
-                 values_to = ".rhat") |>
-    # filter(str_detect(.variable, "b_|sd_")) |> 
-    drop_na() |> 
-    ggplot(aes(.rhat)) +
-    facet_wrap(~model, nrow = 1) +
+    model_convergence |> 
+    ggplot(aes(.neff)) +
     geom_vline(xintercept = 1.01, linetype = "dashed") +
     geom_histogram(binwidth = 0.1, colour = "white", fill = "grey60") +
-    labs(x = "Gelman Rubin diagnostic (R-hat)",
+    labs(x = "Effective sample size",
          y = "MCMC samples",
          title = "Ratio of effective sample size") +
+
     plot_layout(ncol = 1) &
@@
 color_scheme_set(rep("grey60", 6))
+
 titles <- c("Comprehension",
@@
             "Age × Cognateness",
-            "Exposure × Cognateness")
-posterior <- as.array(model_fit_3)
-np <- nuts_params(model_fit_3) 
+            "Exposure × Cognateness",
+            "Age × Exposure × Cognateness")
+
+posterior <- as.array(model_fit)
+np <- nuts_params(model_fit) 
 mcmc_pairs(posterior,
@@
                     "b_age_std:exposure_std",
-                    "b_exposure_std:lv_std"),
+                    "b_exposure_std:lv_std",
+                    "b_age_std:exposure_std:lv_std"),
            draws = c(1, 0),

### 2023-05-23 manuscript/_appendix/_model-predictors.qmd
Add AoA estimator

@@
+```{r sample-sizes}
+#| label: sample-sizes
+#| echo: false
+#| message: false
+#| warning: false
+#| include: false
+tar_load(c(responses, bvq_data, ))
+n_obs <- nrow(responses)
+
+sum_obs <- responses |>
+    mutate(te = factor(te)) |>
+    count(te) |>
+    summarise(across(n, lst(median, min, max), 
+                     .names = "{.fn}"))
+
+# attrition
+n_items <- bvq_data$pool
+
+n_items_nouns <- n_items |> 
+    dplyr::filter(class=="Noun")
+
+n_items_freq <- n_items_nouns |> 
+    left_join(childes,
+              by = c("childes_lemma" = "token")) |> 
+    drop_na(freq_zipf, wordbank_lemma)
+
+n_items_lemma <- n_items_freq |> 
+    dplyr::filter(n_lemmas == 1)
+
+n_items_multi <- n_items_lemma |> 
+    dplyr::filter(!is_multiword) |> 
+    dplyr::filter(include, !is.na(wordbank_lemma)) |> 
+    drop_na(xsampa) |> 
+    distinct(te, language)
+
+n_items_duplicated <- n_items_multi |> 
+    add_count(te, name = "n_te") |>
+    dplyr::filter(n_te == 2)
+
+n_items_final <- n_items_duplicated |> 
+    count(language) |> 
+    group_split(language) |> 
+    set_names(c("Catalan", "Spanish")) |>
+    map_int("n")
+
+n_te <- length(unique(responses$te))
+```
+
 ## Word exposure index  {.unnumbered}
@@
 The this algorithm calculates the minimum number of edit operations (additions, deletions, and substitutions) that one string must go through to become identical to the other string [@levenshtein1966binary] (see @eq-levenshtein).
-For instance, /kaza/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /kasa/ (/z/ is replaced by /s/).
-Therefore, the Levenshtein distance between /kaza/ and /kasa/ is one.
-For an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions in X-SAMPA format , as opposed to IPA, for its computer-friendly set of symbols [@wells1995computer].
+For instance, /\textipa{"ka.za}/ (*house*, in Catalan), would need to go through one substitution to become identical to its translation equivalent in Spanish /\textipa{"ka.sa}/ (/z/ is replaced by /s/).
+Therefore, the Levenshtein distance between /\textipa{"ka.za}/ and /\textipa{"ka.sa}/ is one.
+For an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions in X-SAMPA format, as opposed to IPA, for its computer-friendly set of symbols [@wells1995computer].
 

### 2023-05-23 R/responses.R

@@ -50,7 +50,9 @@ get_responses <- function(bvq_data, items, participants) {
         arrange(id, te, language)
     
     # export data
-    save_files(responses, folder = "data")
+    save_files(responses,
+               formats = "csv",
+               folder = "data")
     
     return(responses)
     

### 2023-05-23 src/figures.R

@@ -1,54 +1,55 @@
 generate_eli <- function(conditions = c("**Cognate**", "**Non-cognate**"),
                          items = c("gat /gat/",
                                    "gato /gato/",
-                                   "gos /gos/", 
+                                   "gos /gos/",
                                    "perro /ˈpe.ro/"),
                          age = seq(10, 34, length.out = 100),
                          freq_month = 1,
                          threshold = 500,
-                         l1_doe = 0.65) 
+                         l1_doe = 0.65)
 {
     
     item_df <- tibble(te = conditions,
                       l1.item = items[c(1, 3)],
-                      l2.item = items[c(2, 4)]) |> 
+                      l2.item = items[c(2, 4)]) |>
         mutate(lv = c(stringsim("gat", "gato"),
-                      stringsim("gos", "pero"))) |> 
+                      stringsim("gos", "pero"))) |>
         mutate(l1.doe = l1_doe, l2.doe = 1-l1_doe)
     
     eli <- expand_grid(item_df,
-                       age = age) |> 
+                       age = age) |>
         mutate(n_month = rpois(n(), freq_month) * age,
                l1.h0_eli = n_month * l1.doe,
                l2.h0_eli = n_month * l2.doe,
                l1.h1_eli = l1.h0_eli + (lv * l2.h0_eli),
                l2.h1_eli = l2.h0_eli + (lv * l1.h0_eli),
                across(matches("h0|h1"),
-                      \(x) ifelse(x < 0 , 0, x))) |> 
+                      \(x) ifelse(x < 0 , 0, x))) |>
         mutate(across(matches("h0|h1"), cumsum),
-               .by = te) |> 
-        select(te, lv, age, matches("l1|l2")) |> 
+               .by = te) |>
+        select(te, lv, age, matches("l1|l2")) |>
         pivot_longer(-c(te, lv, age),
                      names_to = c("language", ".value"),
-                     names_pattern = "(.+)\\.(.+)") |> 
-        mutate(language = toupper(language)) |> 
+                     names_pattern = "(.+)\\.(.+)") |>
+        mutate(language = toupper(language)) |>
         pivot_longer(matches("h0|h1"),
                      names_to = c("hypothesis", ".value"),
                      names_pattern = "(.+)_(.+)",
-                     names_transform = toupper) |> 
-        rename_with(tolower) |> 
[…]

### 2023-05-23 src/helpers.R

@@ -1,4 +1,4 @@
-# run the targets workflow
+#' Run the targets workflow as a RStudio job using [jobs::job()]
 make <- function() {
     job::job({
         targets::tar_make()
@@ -15,7 +15,8 @@ make <- function() {
     }
 }
 
-# remove targets products
+#' Remove targets products using
+#' 
 unmake <- function(keep_fits = TRUE) {
     path <- "results/fit.rds"
     tar_destroy(ask = FALSE)
@@ -31,7 +32,8 @@ unmake <- function(keep_fits = TRUE) {
     if (interactive()) cli_alert_success("Removed project outputs!")
 }
 
-
+#' Print welcome message in console
+#' 
 welcome_message <- function() {
     if (interactive()) {
         id <- cli_status("")
@@ -66,4 +68,3 @@ welcome_message <- function() {
         if (interactive()) cli_status_clear(id)
     }
 }
-

### 2023-05-23 manuscript/_appendix/_model-details.qmd
Add AoA estimator

--- 2023-04-28 _appendix/_model-details.qmd
@@
 
-m <- add_epred_draws(nd, 
-                     model_fit_4_prior,
-                     ndraws = NULL, 
+prior_epreds <- add_epred_draws(nd, 
+                                model_fit_prior,
+                                ndraws = 50, 
                                 re_formula = NA) |> 
@@
     mutate(.category = factor(.category,
-                              levels = c("Understands and Says",
-                                         "Understands"),
-                              labels = c("Production\n(Understands and Says)",
-                                         "Comprehension\n(Understands)")))
+                              levels = c("Understands",
+                                         "Understands and Says"),
+                              labels = c("Comprehension\n(Understands)",
+                                         "Production\n(Understands and Says)")))
 
-m |> 
+prior_epreds |> 
     ggplot(aes(x = age_std, y = .epred)) +
     facet_wrap(~.category) +
-    stat_lineribbon(size = 0.75, 
-                    colour = NA,
-                    .width = c(0.95, 0.89, 0.67, 0.5)) +
-    stat_summary(fun = "mean", 
-                 geom = "line",
-                 linewidth = 0.75) +
+    geom_line(linewidth = 3/4,
+              alpha = 1/2,
+              aes(group = .draw),
+              colour = "grey50") + 
     scale_x_continuous(breaks = scale(seq(10, 36, 4), 

### 2023-05-23 manuscript/manuscript.qmd
Add AoA estimator

@@
 tar_load(c(bvq_data, childes, items, participants, responses))
-tar_load(c(starts_with("model_"), model_loos))
-tar_load(c(posterior_draws, marginal_effects_epreds))
+tar_load(c(model_fit, model_summary, model_epreds, model_convergence))
 
@@
 
-@fig-hypothesis provides a graphic, theoretical illustration of this account.
+@fig-hypotheses provides a graphic, theoretical illustration of this account.
 We assume the case of a Catalan-Spanish bilingual child who is exposed 65% of the time to Catalan, and 35% of the time to Spanish.
@@
 
+For each word included in the questionnaire, we manually generated a broad phonological transcription in X-SAMPA format.
+Catalan word-forms were transcribed to Central Catalan phonology, and Spanish translations were transcribed to Castilian Spanish phonology.
+For each translation equivalent, we calculated the phonological similarity between the Catalan and the Spanish word-form using the same Levenshtein similarity measure used in @floccia2018introduction.
+
 ## Participants {#sec-participants}
@@
 
+suppressWarnings({
     diff_responses <- responses |>
@@
                   .groups = "drop") |> 
-    filter(!(diff_min %in% c(-Inf, Inf))) |>
+        dplyr::filter(!(diff_min %in% c(-Inf, Inf))) |>
         summarise(diff_min = min(diff_min, na.rm = TRUE),
                   diff_max = max(diff_max, na.rm = TRUE))
+})
 
@@
     gt() |>
-    summary_rows(
-        columns = 2:7,
-        fns = list("Total" = ~ sum(., na.rm = TRUE)),
-        formatter = fmt_integer
-    ) |>
     sub_missing(everything(), everything(), missing_text = "--") |>
     cols_label(doe_catalan = "Catalan exposure") |>
-    tab_spanner("Age (months)", 2:8) |>
+    tab_spanner("Age (months)", 2:7) |>
     tab_style(cell_text(align = "left"),
@@
 n_items_nouns <- n_items |> 
-    filter(class=="Noun")
+    dplyr::filter(class=="Noun")
 
@@
 n_items_lemma <- n_items_freq |> 
-    filter(n_lemmas == 1)
+    dplyr::filter(n_lemmas == 1)
 
 n_items_multi <- n_items_lemma |> 
-    filter(!is_multiword) |> 
-    filter(include, !is.na(wordbank_lemma)) |> 
+    dplyr::filter(!is_multiword) |> 
+    dplyr::filter(include, !is.na(wordbank_lemma)) |> 
     drop_na(xsampa) |> 
@@
     add_count(te, name = "n_te") |>
-    filter(n_te == 2)
+    dplyr::filter(n_te == 2)
 
@@
 
-We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_0$fit)[2]` sampling chains with `r format(dim(model_fit_0$fit)[1], big.mark = ",")` iterations each).
-This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $\text{Age}$ (number of months elapsed between participants' birth date and questionnaire completion), $\text{Length}$ (number of phonemes in the X-SAMPA phonological transcription of the word-form), $\text{Exposure}$ (composite measure of lexical frequency and language degree of exposure), and $\text{Cognateness}$ (a measure of cognateness, defined as the phonological similarity between translation equivalents).
-Phonological transcriptions were manually generated.
-Catalan word-forms were transcribed to central Catalan phonology, and Spanish translations were transcribed to  Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
-Participant- and item-level random intercepts and slopes were included when appropriate, according to the structure of the data [@barr2013random].
+We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model.
+This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest.
+We added $\text{Age}$ (number of months elapsed between participants' birth date and questionnaire completion), $\text{Length}$ (number of phonemes in the X-SAMPA phonological transcription of the word-form), $\text{Exposure}$ (composite measure of lexical frequency and language degree of exposure), and $\text{Cognateness}$ (a measure of cognateness, defined as the phonological similarity between translation equivalents) as population-level effects, together with their two- and three-way interactions.
+Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
+Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data [@barr2013random; see @eq-linear].
 We specified a weakly informative prior around the parameters of the model (see Appendix C for details).
@@
 \begin{aligned}
-&\textbf{Response}~k~\textbf{to word}~i~ \textbf{by participant}~j \\
+&\textbf{Response}~(k)~\textbf{to word}~i~ \textbf{by participant}~j \\
 \text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}), \\ &\text{where}~k \in \{\text{No, Understands, Understands and Says}\}\\
 &\textbf{Distribution parameters} \\
-\theta_{k_{ij}} &= (\beta_{0_{k}} + u_{0_{k_{i}}} + w_{0_{k_{j}}}) + (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + \\
-& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + (\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
-& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{i} + (\beta_{5} + u_{5_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{i}) & \\
-u_{0-5_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}); ~w_{0-3_{j}} \sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) \\
+\theta_{k_{ij}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + \\
+&(\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + 
+(\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{ij} + (\beta_{5} + u_{5_{i}} + w_{3_{j}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
+& (\beta_{6} + u_{6_{i}}) \cdot (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
+& (\beta_{7} + u_{7_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
+& (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij}) \\
 &\textbf{Prior} \\
 \beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.5); ~\beta_{1-5} \sim \mathcal{N}(0, 1) \\
-\sigma_{u_{0-5}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2)
+\sigma_{u_{0-8}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2)
 \end{aligned}
-$$
-
-We built a base model $\mathcal{M}_{0}$ that only included the intercept and the main effects of $\text{Age}$, $\text{Length}$, and $\text{Exposure}$ as predictors, and compared it against increasingly more complex models.
-$\mathcal{M}_{1}$ extended the base model by adding the main effect of $\text{Cognateness}$, and $\mathcal{M}_{2}$ added the two-way interaction $\text{Exposure} \times \text{Cognateness}$.
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO-CV, @vehtari2017practical]^[Due to computational constraints, we performed LOO-CV based on a random sub-sample of `r format(dim(model_log_liks$model_fit_0)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]].
-LOO-CV involves computing the likelihood of each data point, given a model  fitted excluding such data point.
-The likelihoods generated by the model for each of the data points are summarised together in a measure of predictive performance: the expected log-predicted density (ELPD).
-This measure penalises the complexity of the model ELPD.
-ELPD values closer to zero indicate better predictive accuracy.
-We compared the ELPD of each model, and selected for interpretation the one showing the best performance (see Appendix C for more details about the models).
-
-We assessed the practical relevance of the estimated regression coefficients of the selected model following @kruschke2018bayesian.
+$$ {#eq-linear}
+
+We assessed the practical relevance of the estimated regression coefficients of the model model following @kruschke2018bayesian.
 First, we specified a region of practical equivalence (ROPE) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`, in the probability scale.
@@
 Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
-Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@vehtari2022loo], `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
+We implemented the model using was done using the `brms` [@burkner2017brms], a R interface to the Stan probabilistic language [@carpenter2017stan].
+We ran `r nchains(model_fit)` MCMC  chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
+Model posterior draws and predictions were handled using the `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
 
@@
 #| echo: false
-rhats_df <- map_df(model_rhats, bind_rows, .id = "model") |> 
-    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
-    pivot_longer(-model, 
-                 names_to = ".variable",
-                 values_to = ".rhat") |>
-    filter(str_detect(.variable, "b_|sd_")) |> 
+rhats_df <- model_convergence |> 
+    dplyr::filter(str_detect(.variable, "b_|sd_")) |> 
     drop_na()
 
-loo_diff <- loo_compare(model_loos)[, 1]
-loo_diff_se <- loo_compare(model_loos)[, 2]
-
-post_draws_list <- posterior_draws$summary |> 
+post_draws_list <- model_summary |> 
     mutate(across(.median:.lower,
@@
                               plogis(x), x/4))) |> 
-    split(posterior_draws$summary$.variable)
-```
-
-All models showed adequate convergence diagnostics and little evidence of multicollinearity between predictors (see Appendix D for details).
-$\mathcal{M}_{2}$, which included the $\text{Exposure} \times \text{Cognateness}$ two-way interaction, showed the best predictive performance with a difference in expected log-predicted density (ELPD) compared to the second best-performing model ($\mathcal{M}_{1}$) of `r round(loo_diff[2], 2)` $\pm$ `r round(loo_diff_se[2], 2)`.
-$\mathcal{M_1}$ performed substantially better than $\mathcal{M_0}$, with a difference of `r round(loo_diff[3], 2)` $\pm$ `r round(loo_diff_se[3], 2)` This points to the effect of $\text{Cognateness}$ being strongly modulated by $\text{Exposure}$, and that this interaction remains relatively stable across ages.
-
-```{r tbl-results-loos}
-#| label: tbl-loos
-#| echo: false
-#| message: false
-#| warning: false
-#| eval: true
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)."
-#| tbl-note: "ELPD (*SD*): expected log-predictive density and standard error.
-Diff.
-(*SD*): diference in ELPD between each model and the model with the best predictive accuracy (Model 3) and standard error of the difference.
-Pareto-k estimates of all models were acceptable (k < 0.5)."
-model_loos |>
-    loo_compare() |>
-    as.data.frame() |>
-    rownames_to_column("model") |>
-    mutate(model = paste0("Model ", str_remove(model, "model_fit_"))) |>
-    select(model, elpd_loo, se_elpd_loo, elpd_diff, se_diff) |> 
-    gt(rowname_col = "model") |>
-    fmt_number(2:5, decimals = 2, sep_mark = ",") |>
-    cols_merge(c(elpd_loo, se_elpd_loo), pattern = "{1} ({2})") |> 
-    cols_merge(c(elpd_diff, se_diff), pattern = "{1} ({2})") |> 
-    cols_label(
-        model = "Model",
-        elpd_loo = "ELPD (SD)",
-        elpd_diff = "Diff. (SD)"
-    ) |>
-    tab_style(cell_text(weight = "bold", style = "italic"),
-              cells_column_labels()) |>
-    tab_style(cell_text(align = "left"),
-              cells_title(groups = "subtitle"))
-```
-
-@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{2}$, and their degree of overlap with the ROPE.
+    split(model_summary$.variable)
+```
+
+All parameters in the model showed adequate MCMC convergence diagnostics and little evidence of multicollinearity (see Appendix D for details).
+@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients, and their degree of overlap with the ROPE.
 For interpretability, we report the estimated regression coefficients transformed to the probability scale.
@@
 This means that one logit difference is not necessarily translated to a unique value in the probability scale.
-For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 30 to 32 months.
+For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 31 to 32 months.
 The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients in the probability scale is not trivial.
@@
 #| label: tbl-coefs
-#| tbl-cap: "Posterior distribution of regression coefficients Model 3."
+#| tbl-cap: "Posterior distribution of regression coefficients."
 #| tbl-note: "Median: median of the posterior distribution in the probability scale.
@@
 
-posterior_draws$summary |>
+model_summary |>
     select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
@@
 #| eval: true
-#| fig-cap: "Posterior marginal effects of Model 3.
+#| fig-cap: "Posterior marginal effects.
 Thin lines correspond to 50 mean predictions.
@@
 
-marginal_effects_epreds |> 
-    mutate(age = rescale_variable(age_std, 
-                                  mean(responses$age), 
-                                  sd(responses$age)),
+model_epreds |> 
+    mutate(age = (age_std * sd(responses$age)) + mean(responses$age),
            exposure = factor(exposure_std,
@@
            xsampa = gsub("\\.", "", xsampa)) |>
-    filter(te %in% items$te, 
+    dplyr::filter(te %in% items$te, 
                   item %in% items$item) |>
@@
                 names_repair = janitor::make_clean_names) |>
-    filter(grepl("E", catalan), 
+    dplyr::filter(grepl("E", catalan), 
                   grepl("e", spanish)) |>

### 2023-05-23 R/participants.R

@@ -12,7 +12,7 @@ get_participants <- function(bvq_data,
                              other_threshold = 0.1) {
     
     participants <- bvq_data$logs |>
-        filter(completed,
+        dplyr::filter(completed,
             # get only short versions of the questionnaire
             str_detect(version, "Lockdown|Short"),
             # get only data from complete questionnaire responses
@@ -33,7 +33,9 @@ get_participants <- function(bvq_data,
     arrange(id)
     
     # export data
-    save_files(participants, folder = "data")
+    save_files(participants,
+               formats = "csv",
+               folder = "data")
     
     return(participants)
 }

### 2023-05-24 manuscript/_appendix/_model-details.qmd
Add new corrections to discussion and reorganise project structure

@@
 ```{r apafg}
-#| label: apafg-prior-mean
+#| label: fig-prior-mean
 #| echo: false
@@
 #| fig-width: 7
-#| fig-height: 5
-#| apa-cap: "Expected prior-predicted mean."
+#| fig-height: 3
+#| out-width: 90%
+#| fig-cap: "Expected prior-predicted mean."
 nd <- expand.grid(n_phon_std = 0,
@@
     ggplot(aes(x = age_std, y = .epred)) +
-    facet_wrap(~.category) +
+    facet_wrap(~ .category) +
     geom_line(linewidth = 3/4,
@@
               colour = "grey50") + 
+    labs(x = "Age (months)", 
+         y = "Probability of acquisition",
+         colour = "Credible interval", 
+         fill = "Credible interval") +
     scale_x_continuous(breaks = scale(seq(10, 36, 4), 
@@
                        labels = seq(10, 36, 4)) +
-    scale_fill_manual(values = rev(clrs), 
-                      labels = c("95%", "89%", "67%", "50%")) +
-    labs(x = "Age (months)", 
-         y = "Probability of acquisition",
-         colour = "Credible interval", 
-         fill = "Credible interval") +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
     theme(legend.position = "top",

### 2023-05-24 manuscript/_appendix/_model-doe.qmd
Add new corrections to discussion and reorganise project structure

@@
 #| out-width: 100%
-#| fig-height: 3
-#| fig-width: 12
+#| fig-height: 4
+#| fig-width: 7
 # summarise posterior draws
-bind_rows("Exposure model" = posterior_draws$summary,
-          "DoE-Frequency model" = posterior_draws_doe,
+bind_rows("Exposure model" = model_summary,
+          "DoE-Frequency model" = posterior_doe_summary,
           .id = "model") |> 
@@
     ggplot(aes(.median, .variable_name,
-               xmin = .lower, xmax = .upper)) +
-    facet_wrap(~model, scales = "free") +
+               xmin = .lower, xmax = .upper,
+               colour = model,
+               shape = model)) +
     annotate(geom = "rect",
@@
              xmax = +0.1/4,
-             fill = "grey70") +
-    geom_errorbar(linewidth = 1, width = 0.25) +
-    geom_point(size = 2.5) +
+             fill = "grey90") +
+    geom_vline(xintercept = 0, 
+               linetype = "dashed",
+               linewidth = 3/4) +
+    geom_errorbar(linewidth = 3/4, 
+                  position = position_dodge(width = 0.5),
+                  width = 0.2) +
+    geom_point(size = 2,
+               position = position_dodge(width = 0.5)) +
     labs(x = "Median and 95% HDI of the posterior distribution",
@@
          colour = "Model") +
-    scale_x_continuous(labels = percent) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
+    scale_colour_grey(start = 0, end = 0.6) +
     theme(legend.position = "none",

### 2023-05-24 manuscript/_appendix/_model-predictors.qmd
Add new corrections to discussion and reorganise project structure

@@
 #| include: false
-tar_load(c(responses, bvq_data, ))
+tar_load(c(responses, bvq_data))
 n_obs <- nrow(responses)
@@
 
-## Word exposure index  {.unnumbered}
+### Word exposure index  {.unnumbered}
 
-Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database [@macwhinney2000childes; @sanchez2019childes].
+Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database [@macwhinney2000childes; @sanchez2019childesdb].
 Using the corresponding lexical frequencies from the Catalan and Spanish corpora was not possible due to the low number of Catalan participants and tokens available in those languages.
@@
 
-## Levenshtein similarity {.unnumbered}
+### Levenshtein similarity {.unnumbered}
 
@@
 This makes many cognates perceptually similar.
-In psycholinguistics, the term *cognateness* is often used to refer to the form-similarity that a pair of translation equivalents share, which has been found to impact how bilinguals process such word-forms (e.g., @costa2000cognate; @spivey1999cross).
+In psycholinguistics, the term *cognateness* is often used to refer to the form-similarity that a pair of translation equivalents share, which has been found to impact how bilinguals process such word-forms [e.g., @costa2000cognate; @spivey1999cross].
 Whether two form-similar translation equivalents are etymologically related or not is arguably tangential to the question of how cognateness affect language processing.
@@
 Using IPA transcriptions of word forms, one can use the same similarity/distance metric as in the case of orthographic word-forms.
-This method has been successfully used to measure the pairwise similarity between phonological word-forms from the  same language [e.g., @fourtassi2020growth] and across languages [@heeringa2004measuring; @floccia2018methods], and also between orthographic word-forms [@schepens2012distributions].
+This method has been successfully used to measure the pairwise similarity between phonological word-forms from the  same language [e.g., @fourtassi2020growth] and across languages [@heeringa2004measuring; @floccia2018ii], and also between orthographic word-forms [@schepens2012distributions].
 
@@
 Therefore, the Levenshtein distance between /\textipa{"ka.za}/ and /\textipa{"ka.sa}/ is one.
-For an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions in X-SAMPA format, as opposed to IPA, for its computer-friendly set of symbols [@wells1995computer].
+For an easier handling of non-ASCII characters (fairly prevalent in Catalan and Spanish, e.g., á, ü, ñ), we used phonological transcriptions in X-SAMPA format, as opposed to IPA, for its computer-friendly set of symbols [@wells1995computercoding].
 
@@
 
-To measure the phonological similarity between the translation equivalents in our study, we computed a normalised versions of the Levenshtein distance between each pair of word-forms that accounts for the lenght of the strings.
+To measure the phonological similarity between the translation equivalents in our study, we computed a normalised versions of the Levenshtein distance between each pair of word-forms that accounts for the length of the strings.
 The rationale behind this correction is that longer word-forms are more likely to differ from their counterpart, compared to shorter word forms, and that this tendency does not correspond necessarily to a larger distance in terms of perception.
@@
 
-The R package `stringdist` [@van-der-loo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure, see @tbl-translation-levenshtein for more examples:
+The R package `stringdist` [@vanderloo2014stringdist] offers the `stringsim()` function to compute this normalised Levenshtein similarity measure, see @tbl-translation-levenshtein for more examples:
 
@@
 #| tbl-cap: Normalised Levenshtein similarity computed for three exemplars of translation pairs in our study
+#| echo: false
 tribble(~Catalan, ~Spanish, ~Levenshtein,
-        "porta [pOrt5]", "puerta [pwe4ta]", 0.50,
-        "taula [tawl5]", "mesa [mesa]", 0.00,
-        "cotxe [kOtS5]", "coche [kotSe]", 0.40) |> 
+        "porta /pOrt5/", "puerta /pwe4ta/", 0.50,
+        "taula /tawl5/", "mesa /mesa/", 0.00,
+        "cotxe /kOtS5/", "coche /kotSe/", 0.40) |> 
     gt() |> 

### 2023-05-24 manuscript/_appendix/_syllable-frequency.qmd
Add new corrections to discussion and reorganise project structure

@@
 items <- items |> 
-    mutate(xsampa = ipa(ipa)) |> 
+    mutate(xsampa = ipa::ipa(ipa)) |> 
     mutate(syll = strsplit(xsampa, 
@@
                 id_cols = c(te, meaning),
-                names_repair = make_clean_names,
+                names_repair = janitor::make_clean_names,
                 values_fn = first) |> 
@@
     geom_col(colour = "white", fill = "grey70") +
-    geom_text(aes(label = percent(prop)),
+    geom_text(aes(label = scales::percent(prop)),
               size = 3,
@@
     top_n(20, freq_syll) |> 
-    ggplot(aes(reorder_within(syll, desc(freq_syll), language),
+    ggplot(aes(tidytext::reorder_within(syll, desc(freq_syll), language),
                freq_syll)) +
@@
          fill = "Language") +
-    scale_x_reordered() +
+    tidytext::scale_x_reordered() +
     scale_y_continuous(labels = \(x) format(x, big.mark = ",")) +
@@
     "b_Intercept" = "Intercept",
-    "b_n_syll_std" = glue("Syllables (+1 SD, {round(sd(syllables_data$n_syll), 3)})"),
-    "b_lv_std" = glue("Cognateness (+1 SD, {percent(sd(syllables_data$lv), accuracy = 0.01)})")
+    "b_n_syll_std" = glue::glue("Syllables (+1 SD, {round(sd(syllables_data$n_syll), 3)})"),
+    "b_lv_std" = glue::glue("Cognateness (+1 SD, {round(sd(syllables_data$lv), 2)})")
 )
@@
 post_syllables <- model_fit_syllables |>
-    describe_posterior(test = "rope",
+    bayestestR::describe_posterior(test = "rope",
                                     rope_range = rope_interval,
@@
     as_tibble() |>
-    clean_names() |>
+    janitor::clean_names() |>
     mutate(.variable_name = factor(
@@
             as.character(),
-        parameter = ifelse(str_detect(parameter, "Intercept"),
-                           str_remove_all(parameter, "Intercept \\(|\\)"),
+        parameter = ifelse(grepl("Intercept", parameter),
+                           stringr::str_remove_all(parameter, "Intercept \\(|\\)"),
                            parameter)) |>
@@
 
-@tbl-syll-marginal shows the median posterior-predicted syllabic frequencies for words with one to four syllables, for the whole range of cognateness values.
+@fig-syll-marginal shows the median posterior-predicted syllabic frequencies for words with one to four syllables, for the whole range of cognateness values.
 Overall, cognate words' syllabic frequency is equivalent to that of non-cognates.
@@
 
-```{r tbl-syll-marginal}
-#| label: tbl-syll-marginal
-#| tbl-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
+```{r fig-syll-marginal}
+#| label: fig-syll-marginal
+#| fig-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
 #| fig-width: 7
@@
                     re_formula = NA) |> 
-    mutate(lv_std = rescale_variable(lv_std,
-                                     mean = mean(syllables_data$lv),
-                                     sd = sd(syllables_data$lv)),
-           n_syll_std = rescale_variable(n_syll_std,
-                                         mean = mean(syllables_data$n_syll),
-                                         sd = sd(syllables_data$n_syll)),
+    mutate(
+        lv_std = (lv_std * sd(syllables_data$lv)) + mean(syllables_data$lv),
+        n_syll_std = (n_syll_std * sd(syllables_data$n_syll)) + mean(syllables_data$n_syll),
         n_syll_std = as.factor(n_syll_std)) |> 
-    ggplot(aes(lv_std, .epred, colour = n_syll_std)) +
+    ggplot(aes(lv_std, .epred, 
+               colour = n_syll_std)) +
     geom_point(data = filter(syllables_data, n_syll < 5),
@@
                    y = freq_syll,
-                   colour = as.factor(n_syll)),
-               shape = 1,
-               stroke = 1,
+                   colour = as.factor(n_syll),
+                   shape = as.factor(n_syll)),
                alpha = 1/2) + 
@@
          y = "Syllabic frequency",
+         shape = "Number of syllables",
          colour = "Number of syllables") +
-    scale_color_manual(values = clrs) +
+    scale_colour_grey(start = 0, end = 0.7) +
     scale_y_continuous(n.breaks = 8) +
-    theme(panel.grid = element_blank())
+    theme(panel.grid = element_blank(),
+          legend.position = "top")
 ```

### 2023-05-24 manuscript/appendix.qmd
Add new corrections to discussion and reorganise project structure

@@
 ---
+title: "Appendix"
 csl: "../assets/apa7.csl"
@@
 editor: source
+
+fig-align: center
+fig-dpi: 800
+echo: false
+warning: false
+message: false
 
@@
     linestretch: 1.5
-    date-format: full
-    template-partials:
-    - partials/title.tex
-    - partials/before-body.tex
-    include-in-header: include-in-header.tex
-    header-includes: \usepackage{tipa}
+    header-includes:
+    - \usepackage{tipa}
     geometry:
@@
     colorlinks: true
-    linkcolor: DarkSlateBlue
-    urlcolor: DarkSlateBlue
-    citecolor: DarkSlateBlue
+    linkcolor: DodgerBlue
+    urlcolor: DodgerBlue
+    citecolor: DodgerBlue
     link-citations: true
-    biblio-title: References
     biblio-config: false
-    cite-method: biblatex
-    biblatex-chicago: false
-    biblio-style: apa
-    biblatexoptions: backend=biber
     block-headings: false
@@
 
-library(targets)
+library(patchwork)
+library(gt)
+library(ggplot2)
 
-tar_config_set(store = here::here('_targets'),
-               script = here::here('_targets.R'))
+targets::tar_config_set(store = here::here("_targets"),
+                        script = here::here("_targets.R"))
 
-tar_load_globals()
+targets::tar_load_globals()
 
-tar_load(c(bvq_data, childes, items, participants, responses))
-tar_load(c(starts_with("model_"), model_loos))
-tar_load(c(posterior_draws, marginal_effects_epreds))
+targets::tar_load(c(bvq_data, childes, items, 
+                    participants, responses))
+
+targets::tar_load(c(model_fit, model_summary, model_epreds,
+                    model_convergence, model_fit_prior, 
+                    model_ppcs, posterior_doe_summary,
+                    posterior_doe_draws))
+
+targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))
 
@@
 
-clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
-color_scheme_set(rep(clrs[1], 6))
-
-my_theme <- theme_minimal() +
+my_theme <- function() {
+    theme_ggdist() +
         theme(panel.grid = element_line(colour = "grey",
@@
               axis.text = element_text(colour = "black"))
+}
 
-theme_set(my_theme)
+theme_set(my_theme())
 
-bayesplot_theme_set(my_theme)
+bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))
 
@@
 
-options(ggplot2.ordinal.fill = clrs,
-        ggplot2.ordinal.colour = clrs,
-        ggplot2.discrete.fill = clrs,
-        ggplot2.discrete.colour = clrs,
-        ggplot2.continuous.fill = scale_color_gradient,
-        ggplot2.continuous.colour = scale_color_gradient,
-        knitr.graphics.error = FALSE) 
+options(knitr.graphics.error = FALSE)
 ```
 
-# Appendix A: model predictors {.appendix .unnumbered}
+## Appendix A: model predictors {.appendix .unnumbered}
 
@@
 
-# Appendix B: model details {.appendix .unnumbered}
+## Appendix B: model details {.appendix .unnumbered}
 
@@
 
-# Appendix C: convergence diagnostics {.appendix .unnumbered}
+## Appendix C: convergence diagnostics {.appendix .unnumbered}
 
@@
 
-# Appendix D: frequency and language exposure as separate predictors {.appendix .unnumbered}
+## Appendix D: frequency and language exposure as separate predictors {.appendix .unnumbered}
 
@@
 
-# Appendix E: frequency and language exposure as separate predictors {.appendix .unnumbered}
+## Appendix E: frequency and language exposure as separate predictors {.appendix .unnumbered}
 
@@
 {{< pagebreak >}}
+
+## References

### 2023-05-24 manuscript/manuscript.qmd
Add new corrections to discussion and reorganise project structure

@@
 ---
-title: "Cognate beginnings to bilingual lexical growth"
+title: "Cognate beginnings to bilingual lexical acquisition"
 short-title: "Cognate beginnings"
@@
   pdf:
-    documentclass: article
     papersize: letter
@@
     - templates/before-body.tex
-    include-in-header: include-in-header.tex
-    header-includes: \usepackage{tipa}
+    include-in-header:
+    - templates/include-in-header.tex
+    header-includes: 
+    - \usepackage{tipa}
+    - \usepackage{nowidow}
     geometry:
@@
     colorlinks: true
-    linkcolor: DarkSlateBlue
-    urlcolor: DarkSlateBlue
-    citecolor: DarkSlateBlue
+    linkcolor: DodgerBlue
+    urlcolor: DodgerBlue
+    citecolor: DodgerBlue
     link-citations: true
-    biblio-title: References
     biblio-config: false
@@
 
-library(targets)
-
-tar_config_set(store = here::here('_targets'),
+targets::tar_config_set(store = here::here('_targets'),
                         script = here::here('_targets.R'))
 
-tar_load_globals()
+targets::tar_load_globals()
 
-tar_load(c(bvq_data, childes, items, participants, responses))
-tar_load(c(model_fit, model_summary, model_epreds, model_convergence))
+targets::tar_load(c(bvq_data, childes, items,
+                    participants, responses))
+
+targets::tar_load(c(model_fit, model_summary, 
+                    model_epreds, model_convergence))
+
+library(gt)
+library(ggplot2)
+library(patchwork)
 
@@
 items <- bvq_data$pool  |>  
-    mutate(item = str_remove(item, "cat_|spa_")) |> 
+    mutate(item = gsub("cat_|spa_", "", item)) |> 
     select(item, language, semantic_category, class) |> 
@@
 clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
-color_scheme_set(rep(clrs[1], 6))
 
@@
 theme_set(my_theme)
-
-bayesplot_theme_set(my_theme)
 
@@
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
+Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilinguala].
 These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical and the phonological level.
@@
 Not all words occur in the linguistic input with identical frequency, so children accumulate experience with some word-forms at faster rates than with other word-forms.
-This results in frequent words being acquired earlier in age than less frequent words [e.g., @ambridge2015ubiquity; @goodman2008does; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
+This results in frequent words being acquired earlier in age than less frequent words [e.g., @ambridge2015ubiquity; @goodman2008doesa; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
 For this reason, lexical frequency plays a central role in most models of lexical growth, especially in the case of accumulator models [@hidaka2013computational; @mollica2017how].
 Following this account, words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are predicted to be acquired earlier in age than words with low lexical frequency such as 'taxi' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words).
+
 We argue that previous accounts on vocabulary growth based on accumulator models are a convenient framework to study bilingual vocabulary growth.
 Bilinguals share fundamentally similar learning mechanisms with their monolinguals peers.
-Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilingual].
+Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilinguala].
 One of such divergences is the fact that bilinguals' frequency of exposure to a word-form is mediated by their degree of exposure to their languages.
-Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual (who listens 100% of the time to English).
-would need.
+Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual (who listens 100% of the time to English) would need.
+
 In the present study, we incorporate this feature of the bilingual linguistic input to an account of the cognateness effect on word acquisition.
+We suggest that when a bilingual child is exposed to a given word, they not only activate its corresponding lexical representation, but also that of its translation equivalent.
+This co-activation is proportional to the form similarity between both word-forms.
+When this parallel activation is in place, each exposure to one of the word-forms may count as a learning instance for both members of the translation equivalent, in a degree determined by their cognateness.
+Cognate translation equivalents, for which stronger co-activation is expected, accumulate learning instances faster than non-cognates, for which weaker or no co-activation is expected.
+Finally,  when the child has accumulator some threshold number of learning instances with any of these word-forms, its lexical representation reaches some degree of consolidation that leads to a non-linear change in the child's observed learning outputs, like the probability of such word being reported as "acquired" by caregivers, or by stronger naming effects in experimental tasks using the Visual World Paradigm [e.g., @swingley2002recognition].
+Because , lexical representations of cognate words accumulate learning instances faster than those of non-cognate words, the former would consolidate earlier in age.
 
-We suggest that the cognateness effect proposed by Floccia et al.
-and Mitchell et al.
-facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired.
-We hypothesise that both forms in the translation equivalent (e.g., *gat* and *gato* [Catalan and Spanish for 'cat'], for a Catalan-Spanish bilingual) follow their acquisition trajectory as they would for a monolingual child.
-As the infant receives speech input, both word-forms accumulate learning instances, which occur a a rate proportional to the lexical frequency of each word-form and the amount of exposure to the language the word-form belongs to.
-When one of the word-forms has accumulated a certain amount of learning instances, it is acquired.
-After such event takes places, we hypothesise that the number of learning instances left needed for the acquisition of the other word-form (yet to be acquired) is reduced, resulting in an earlier age-of-acquisition that it would be expected if the acquisition of both word-forms was independent.
-We further suggest that this facilitation is proportional to the amount of phonological similarity (i.e., cognateness) between both word-forms, in such way that the facilitation is larger within similar-sounding translation equivalents, and negligible for form-dissimilar translation equivalents.
-
-Consider the example of the Catalan-Spanish cognate translation equivalent *gat*-*gato* [cat] (whose word-forms share three phonemes), and of the non-cognate translation equivalent *gos*-*perro* [dog] whose word-forms only share one phoneme.
-When the child has sufficient lexical knowledge of *gat*, and therefore there is some form-referent association in place, future presentations of the similar-sounding translation *gato* in the presence of the same reference should lead to an increased likelihood of acquisition.
-In the case of a non-cognate translation equivalent, such facilitation would be unlikely.
-For instance, even if the child has already acquired the word *gos*, hearing its dissimilar Spanish translation *perro* would hardly lead to an increased probability of acquisition.
-In line with this account, only the later-acquired form of the translation equivalent would benefit from its cognate status.
-Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual; @hoff2012dual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
-For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age-of-acquisition for cognates than for non-cognates.
+Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [cat], whose word-forms share three phonemes.
+When the child is exposed to /\textipa{"gat}/, they might co-activate /\textipa{"ga.to}/ in parallel, due to their phonological similarity.
+If this exposure to the word form occurs in a convenient referential context (e.g., in the presence of a cat), this might count as a learning instance for both active word-forms.
+Consider now the case of the non-cognate translation equivalent /\textipa{"gos}-\textipa{"pe.ro}/.
+Because of their low phonological similarity, the amount of co-activate of both words might be low, which might leads to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for  /\textipa{"pe.ro}/.
+While /\textipa{"gat}-\textipa{"ga.to}/ are predicted to benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ are not.
+This facilitation effect may not be equivalent for /\textipa{"gat}-\textipa{"ga.to}/.
+If the child receives a larger amount of Catalan input than Spanish input, they will encounter the Catalan form /\textipa{"gat}/ more frequently than the Spanish form /\textipa{"ga.to}/.
+If parallel activation is in place, the former will activate the latter more often than vice versa.
+Ultimately, /\textipa{"ga.to}/ should benefit more strongly from its cognate status than /\textipa{"gat}/.
+For instance, if the previous child is exposed 80% of the time to Catalan, and 20% of the time to Spanish, words in Spanish should, on average, show a stronger cognateness facilitation effect than words in Catalan.
 
@@
 We assume the case of a Catalan-Spanish bilingual child who is exposed 65% of the time to Catalan, and 35% of the time to Spanish.
-On the top panels, we illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/\textipa{gat}/ or /\textipa{gos}/) and its Spanish translation (/\textipa{"pe.ro}/ or /\textipa{"ga.to}/).
+We illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/\textipa{gat}/ or /\textipa{gos}/) and its Spanish translation (/\textipa{"pe.ro}/ or /\textipa{"ga.to}/).
 We assume an arbitrary, theoretical threshold for the learning instances at which the child is considered to have acquired a word (solid grey line).
 Points and triangles show the age-of-acquisition at which the amount of learning instances reaches the threshold.
-We show the predicted age-of-acquisition under two hypothesis: (1) a cognateness facilitation effect that lowers the acquisition threshold after one of the word-forms (in this case Catalan /\textipa{gat}/) has been acquired, and (2) no cognateness facilitation, for which the acquisition of the second word-form occurs unconditionally of the acquisition of the first.
-On the bottom, we show the predicted probability of acquisition for the Catalan and Spanish words under each hypothesis, as it would be expected to observe in a vocabulary questionnaire like the one in the present study.
-We assume that such probability starts at 0% at birth, reaches 50% (and grows the fastest) at the age-of-acquisition, and reaches 100% at some points in later ages.
+We show the predicted age-of-acquisition under two hypothesis: (1) a cognateness facilitation effect through parallel activation, in which both word-forms co-activate each other accumulating learning instances faster, and (2) no cognateness facilitation, for which both word-forms accumulate learning instances independently.
 
@@
 This measure was calculated by multiplying the word-form's lexical frequency with participant's degree of exposure to the language the word belongs to.
-Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childes] (see Appendix A for more details).
+Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childesdb] (see Appendix A for more details).
 Participants' degree of exposure were reported by caregivers in a language exposure questionnaire they filled before the vocabulary checklist.
@@
 *Cognateness* was defined as a continuous measure of phonological similarity between translation equivalents.
-We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-XSAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computer].
+We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-XSAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computercoding].
 This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
@@
                 values_from = n) |> 
-    clean_names() |> 
+    janitor::clean_names() |> 
     mutate(version = tolower(version))
@@
 
-To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire.
+To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire [@garcia-castro2023bvq].
 This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formr platform [@arslan2020formr].
@@
     range(na.rm = TRUE) |> 
-    as_date()
+    lubridate::as_date()
 
@@
 Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants) for a more detailed description of the sample).
-In total, `r n_trilingual$n[2]` participants (`r percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed to a third language other than Catalan and Spanish.
+In total, `r n_trilingual$n[2]` participants (`r scales::percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed to a third language other than Catalan and Spanish.
 All families provided informed consent before participating.
@@
 This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: *sense escolaritzar/sin escolarizar* [no education], *educació primària/educación primaria* [primary school], *educació secundària/educación secundaria* [secondary school], *batxillerat/bachillerato* [complementary studies/high school], *cicles formatius/ciclos formativos* [vocational training], and *educació universitària/educación universitaria* [university degree].
-Most families reported university studies (`r participants_edu["University"]`, `r percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
+Most families reported university studies (`r participants_edu["University"]`, `r scales::percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r scales::percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r scales::percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r scales::percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
 
@@
 
-rope_interval <- percent(c(lower = -0.1/4, upper = 0.1/4), accuracy = 0.1)
+rope_interval <- scales::percent(c(lower = -0.1/4,
+                                   upper = 0.1/4),
+                                 accuracy = 0.1)
 
@@
 
-r_version <- glue(r_info$R.version$major, 
+r_version <- glue::glue(r_info$R.version$major, 
                         r_info$R.version$minor, 
@@
 rhats_df <- model_convergence |> 
-    dplyr::filter(str_detect(.variable, "b_|sd_")) |> 
+    dplyr::filter(grepl("b_|sd_", .variable)) |> 
     drop_na()
@@
         .lower = md("95% HDI"),
-        .rope_overlap = md("ROPE prob.")
+        .rope_overlap = md("*p*(H0)")
     ) |>
@@
 This effect was mediated by exposure rate: low-exposure word-forms benefited from their cognate status, whereas high-exposure word-forms did not.
-Capitalising on accumulator models of language acquisition, we provide a theoretical account of bilingual lexical acquisition in which parallel activation plays a central role during the consolidation of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
+Capitalising on accumulator models of language acquisition, we provided a theoretical account of bilingual lexical acquisition.
+In this account, parallel activation of the two languages plays a central role during the consolidation of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
 
@@
 shared fewer cognates, like English-Mandarin).
-Second, Mitchell et al.
-found an earlier age-of-acquisition for cognates, compared to non-cognates.
-The outcomes of both studies pointed to cognateness facilitating word acquisition through parallel activation, that is, the co-activation of lexical representations from the two languages during word comprehension and production.
-But the underpinnings of such effect were unclear: while parallel activation has been extensively described in experimental studies, current paradigms of bilingual word acquisition and word learning are, to a large extent, dissociated from the mechanisms proposed by the former.
+Second, @mitchell2022cognates found an earlier age-of-acquisition for cognates, compared to non-cognates.
+The outcomes of both studies pointed to cognateness facilitating word acquisition through parallel activation.
+But the underpinnings of such effect were unclear: while parallel activation has been extensively described in experimental studies, current paradigms of bilingual word acquisition and word learning are, to a large extent, dissociated from the mechanisms proposed by previous work on word processing.
 Accumulator models of language acquisition provide a convenient theoretical framework to narrow this gap.
@@
 Consider a Catalan-Spanish bilingual child who is exposed to the Catalan word-form /\textipa{"gat}/ in the presence of a cat.
-This represents a learning instance, in which the child has the opportunity to strengthen the association between the word-form and its referent.
+This may represent a learning instance, in which the child has the opportunity to strengthen the association between the word-form and its referent.
 Because the Spanish translation of this word (/\textipa{"ga.to}/) is phonologically very similar to the Catalan form (both word-forms share `r format(stringdist::stringsim("gat", "gato"), nsmall = 2)` Levenshtein similarity), it might receive some degree of co-activation.
@@
 Such a facilitatory cross-language mechanism might be an important piece in the puzzle of bilingual language acquisition.
-In particular, it may shed some light on why bilingual infants do not show relevant delays in language acquisition milestones compared to their monolingual peers, while receiving a lower quantity of speech input in each of their languages.
-Our results provide some insights into this issue: infants benefitted more strongly from the cognateness facilitation effect when acquiring words from the language of lower exposure than in the language of higher exposure.
+In particular, it may shed some light on why bilingual infants do not show relevant delays in language acquisition milestones compared to their monolingual peers, while receiving a reduced quantity of speech input in each of their languages.
+Our results provide some insights into this issue: infants benefited more strongly from the cognateness facilitation effect when acquiring words from the language of lower exposure than in the language of higher exposure.
 
-We suggest that this asymmetry is the result of children being exposed to words in the higher-exposure language more often than to words in the lower-exposure language.
+We suggest that this asymmetry is the result of children' unbalanced exposure to their languages.
 A bilingual child's frequency of exposure to a given word-form is mostly determined by two factors: the word-form's lexical frequency and the child's amount of language exposure to the language such word belongs to.
-A dual linguistic input means lower exposure to each of the languages, unless one makes the---arguably implausible---assumption that bilinguals are exposed to larger quantities of linguistic input.
-In this scenario, bilinguals encounter each word on lower rates than their monolingual peers.
+A dual linguistic input means lower exposure to each of the languages, unless one makes the---arguably implausible---assumption that bilinguals are exposed on average to twice the amount of linguistic input than monolinguals.
 Because of this difference in exposure, words lower-exposure language might receive activation from their translation in the higher-exposure language more often than words from the higher-exposure language receive activation from their translation in the lower-exposure language.
 As a result, the cognateness facilitation effect should be stronger in words from the lower-exposure language.
-For instance, in the case of a Catalan-dominant, Catalan-Spanish child, the Catalan word /\textipa{"gat}/ would be learned earlier than its Spanish translation /\textipa{"ga.to}/.
-Therefore, /\textipa{"ga.to}/ should be more likely to receive facilitation from /\textipa{"'pe.ro'gat}/ than vice versa (see @fig-hypotheses in the introduction for a graphic illustration).
 
-We tested this hypothesis by combining language exposure and lexical frequency into a composite variable---, which we coined exposure rate---to obtain an estimate of how often each participant encountered each word.
-We introduced the exposure rate as a predictor in our model, and found a substantial interaction between this predictor and cognateness.
-In line with our predictions, only words-forms to which children were exposed to less frequently benefited from their cognate status.
 This mechanism might be extended to provide a plausible explanation for the language similarity facilitation reported by Floccia et al.
@@
 It might be argued that our results reflect the fact that cognates are represented in the initial bilingual lexicon as the *same* lexical entry.
-Because bilingual children are exposed to such similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the presence of a cat), it is possible that they classify both are as acceptable variations of the same word, therefore treating them as a single lexical item, without parallel activation taking place.
-This could explain a cognate advantage in word acquisition without the need for parallel activation playing a relevant role.
-This account would not necessarily imply that bilinguals are insensitive to the phonological mismatches between cognate word-forms.
-On the contrary, there is strong evidence that both monolingual and bilingual infants are highly sensitive to subtle mispronunciations of word-forms [@swingley200511montholds; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels].
-The assimilation of both word-forms in the lexicon might occur at a higher level, where both word-forms are considered referentially equivalent, so that every exposure to either of the (distinguishable) word-forms of the translation equivalent is counted as a learning instance for both.
-This would lead to a faster increase in cumulative learning instances across ages, leading to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
+Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the presence of a cat), it is possible that they classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
+This would lead to a faster increase in cumulative learning instances, and to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
 This mechanism could potentially explain the earlier age-of-acquisition effect of cognates found in the present study, without the need of parallel activation playing any relevant role.
-
-Previous work points in a different direction, with bilingual toddlers considering cognates as separate words.
-For instance, @ramon-casas2009vowel showed that 18-26 months-old bilingual toddlers reject mispronounced labels in word recognition tasks, although only when such mispronunciation involves a contrast that is phonemic in both languages.
-In their study, Catalan-Spanish bilinguals treated similarly the words /\textipa{"pe\textesh}/ and its mispronunciated form /\textipa{"pE\textesh}/, which involved the /\textipa{e}-\textipa{E}/ contrast (only present in Catalan), but treated differently the words /\textipa{pe\textesh}/ and /\textipa{pi\textesh}/, which involved the /\textipa{e}-\textipa{i}/ contrast (present in both languages).
-Moreover, a later study by @ramon2010non showed that bilinguals treat the language-specific /\textipa{e}-\textipa{E}/ contrast differently when embedded within *non-cognate* word-forms.
-These results suggest that, except for the acoustically close, language-specific /\textipa{e}-\textipa{E}/ contrast, bilinguals treat similar-sounding words differently in word recognition tasks, especially when both word-forms are separated by more than one phoneme.
-In the light of these findings, the possibility that bilinguals merge cognates into shared lexical representations becomes more implausible.
-
-This is especially true for the case of the Catalan-Spanish bilinguals in our sample: there are few exemplars of Catalan-Spanish translation equivalents involving a /\textipa{e}-\textipa{E}/ substitution.
-Only `r percent(nrow(E_e)/length(unique(items$te)))` of the translation equivalents in our questionnaire involved a /\textipa{e}-\textipa{E}/ change between the Catalan and the Spanish items, and only one of them was a minimal pair: /\textipa{"tREN}/-/\textipa{"tReN}/ [*train*].
-Moreover, few words in our study differ only in one phoneme substitution.
-Only `r percent(lv_distr$prop[nrow(lv_distr)], accuracy = 0.01)` of the translation equivalents included in our analyses shared between 90% and 100% phonological similarity (measured as the Levenshtein similarity between the X-SAMPA transcriptions of the Catalan and Spanish form), and only `r percent(lv_distr$prop[nrow(lv_distr)-1], accuracy = 0.01)` shared between 80 and 90% similarity.
-This shows that the presence of identical and near-identical cognates in our study was very low, and therefore most translation equivalents differed in more than one phoneme.
-Finally, and perhaps more critically, there is evidence beyond word recognition that bilinguals treat minimal pairs differently at the referential level.
-For instance, @ramon-casas2017minimalpair found that bilingual toddlers are able to learn new minimal pairs involving the /\textipa{e}-\textipa{E}/ contrast.
-This suggests that toddlers' tolerance towards cognate mispronunciations at the referential level might be restricted to familiar word recognition, while such subtle variations in the word-forms are considered relevant during word learning [@mattock2010first; @havy2016phonetic].
+Previous work points in a different direction: bilingual toddlers show monolingual-level sensitivity slight phonetic changes in a word-form, according to their performance in word recognition tasks [@swingley200511montholds; @ramon-casas2017minimalpair; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels; @ramon-casas2010are; @ramon-casas2009vowel], and word learning [@ramon-casas2017minimalpair; @mattock2010first; @havy2016phonetic].
+Overall, there is compelling evidence that bilingual toddlers form distinct lexical representations for even near-identical cognates.
 
@@
 Such reports can be subject to measurement error induced by caregivers who may sometimes overestimate or underestimate participants' true probability of acquisition of words [e.g., @houston-price2007discrepancy].
-For instance, although in the present study caregivers were explicitly instructed *not* to rely on their responses to Catalan words when responding to Spanish (and vice versa), it is possible that some caregivers assumed---at least to some extent---that because the child knew a word in one language, the child should also know the word in the other language (especially when both words are similar-sounding, i.e., cognates).
+For instance, although in the present study caregivers were explicitly instructed *not* to rely on their responses to Catalan words when responding to Spanish (and vice versa), it is possible that some caregivers assumed---at least to some extent---that because the child knew a word in one language, the child should also know the word in the other language.
+This bias would especially affect similar-sounding words, i.e., cognates.
 Production estimates may be more prompt to such biases, in part because of the slower pace at which their articulatory abilities develop, compared to their word recognition abilities.
@@
 Bilinguals differ across multiple dimensions.
-these differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Arabic, or high like the case of Spanish and Greek), to higher-level factors like the socio-political situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages may be less dominant than the other).
+these differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to higher-level factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
 This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.
 Our sample, although homogeneous (e.g., similar parental educational level across), represents a fairly particular population of bilinguals: the languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages, both languages are used in fairly similar contexts, and both languages are known by the majority of the population.
-In 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@de2013els].
+In 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@2018els].
 In addition, and as commented in the introduction, Catalan and Spanish are both Romance languages and share a considerable amount of cognates.
@@
 Capitalising on accumulator models of language acquisition we, put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
-
-# References {#sec-references}
-
-::: {#refs}
-
-:::

### 2023-05-24 R/models.R

@@ -11,11 +11,10 @@ fit_model <- function(name, ...)
                seed = 888,
                threads = threading(4),
                backend = "cmdstanr",
-               file = glue("results/fits/{name}.rds"),
-               file_refit = "on_change",
+               file = glue::glue("results/fits/{name}.rds"),
                control = list(adapt_delta = 0.9,
                               max_treedepth = 15),
-               save_model = glue("stan/{name}.stan"))
+               save_model = glue::glue("stan/{name}.stan"))
     
     return(fit)
 }
@@ -29,10 +28,10 @@ get_vars_dict <- function(responses)
     c(
         "b_Intercept[1]" = "Comprehension and Production",
         "b_Intercept[2]" = "Comprehension",
-        "b_age_std" = glue("Age (+1 SD, {round(sd(responses$age), 2)}, months)"),
-        "b_n_phon_std" = glue("Length (+1 SD, {round(sd(responses$n_phon), 2)} phonemes)"),
-        "b_exposure_std" = glue("Exposure (+1 SD, {round(sd(responses$exposure), 2)})"),
-        "b_lv_std" = glue("Cognateness (+1 SD, {percent(sd(responses$lv), accuracy = 0.01)})"),
+        "b_age_std" = glue::glue("Age (+1 SD, {round(sd(responses$age), 2)}, months)"),
+        "b_n_phon_std" = glue::glue("Length (+1 SD, {round(sd(responses$n_phon), 2)} phonemes)"),
+        "b_exposure_std" = glue::glue("Exposure (+1 SD, {round(sd(responses$exposure), 2)})"),
+        "b_lv_std" = glue::glue("Cognateness (+1 SD, {round(sd(responses$lv), 2)})"),
         "b_exposure_std:lv_std" = "Exposure \u00d7 Cognateness",
         "b_age_std:exposure_std" = "Age \u00d7 Exposure",
         "b_age_std:lv_std" = "Age \u00d7 Cognateness",
@@ -47,19 +46,6 @@ get_vars_dict <- function(responses)
 #' @param ... Arguments to be passed to [tidybayes::gather_draws()].
 get_posterior_draws <- function(model, data, var_dict, ...)
 {
-    # tidy predictor names
-    str_repl <- c(
-        "b_Intercept[1]" = "Comprehension and Production",
-        "b_Intercept[2]" = "Comprehension",
[…]

### 2023-05-24 manuscript/_appendix/_model-diagnostics.qmd
Add new corrections to discussion and reorganise project structure

@@
 Values lower than 1.01 are recommended, while values higher than 1.05 indicate that chains might have trouble converging and therefore the estimated parameters must be taken with caution.
-@tbl-rhats-neffs (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference.
+@fig-rhats-neffs (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference.
 Most values are lower than 1.01, and never higher than 1.05, which provides evidence of successful MCMC convergence.
@@
 Values larger than 0.1 are recommended.
-@tbl-rhats-neffs (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models.
+@fig-rhats-neffs (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models.
 Most values are larger than 0.1, although model 0 ($\mathcal{M}_0$) accumulates most effective sample sizes close to 0.1.
 
-```{r tbl-rhats-neffs}
-#| label: tbl-rhats-neffs
-#| tbl-cap: "MCMC convergence diagnostic of all parameters in the model"
+```{r fig-rhats-neffs}
+#| label: fig-rhats-neffs
+#| fig-cap: "MCMC convergence diagnostic of all parameters in the model. A: distribution of the Gelman-Rubin (R-hat) scores. B: distribution of the ratio of effective sa"
 #| echo: false
@@
 #| warning: false
-#| fig-height: 6
-#| fig-width: 9
+#| fig-height: 3
+#| fig-width: 6
 #| out-width: 100%
@@
     ggplot(aes(.rhat)) +
+    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
     geom_vline(xintercept = 1.01, linetype = "dashed") +
-    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
-    labs(x = "Gelman Rubin diagnostic (R-hat)",
-         y = "MCMC samples",
-         title = "Gelman-Rubin (R-hat) diagnostic") +
+    labs(x = "R-hat",
+         y = "MCMC samples") +
 
@@
     ggplot(aes(.neff)) +
+    geom_histogram(binwidth = 0.1, colour = "white", fill = "grey60") +
     geom_vline(xintercept = 1.01, linetype = "dashed") +
-    geom_histogram(binwidth = 0.1, colour = "white", fill = "grey60") +
-    labs(x = "Effective sample size",
-         y = "MCMC samples",
-         title = "Ratio of effective sample size") +
+    labs(x = "Effective sample size ratio",
+         y = "MCMC samples") +
 
@@
 
-## Posterior draws and bi-variate scatterplots {.unnumbered}
+### Posterior draws and bi-variate scatterplots {.unnumbered}
 
@@
 #| message: false
-color_scheme_set(rep("grey60", 6))
-
 titles <- c("Comprehension",
@@
 
-posterior <- as.array(model_fit)
-np <- nuts_params(model_fit) 
-mcmc_pairs(posterior,
-           np = np, 
-           pars = c("b_Intercept[1]",
-                    "b_Intercept[2]",
-                    "b_age_std",
-                    "b_n_phon_std",
-                    "b_exposure_std",
-                    "b_lv_std",
-                    "b_age_std:exposure_std",
-                    "b_exposure_std:lv_std",
-                    "b_age_std:exposure_std:lv_std"),
-           draws = c(1, 0),
-           grid_args = list(titles = titles),
-           off_diag_args = list(size = 0.5,
-                                alpha = 1/10))
+posterior <- as_draws_df(model_fit) |> 
+    select(`b_Intercept[1]`:`b_age_std:exposure_std:lv_std`)
+
+my_diag <- function(data, mapping, ...) {
+    ggplot(data = data, mapping = mapping) + 
+        stat_slab(colour = "white",
+                  fill = "grey15")
+}
+
+my_upper <- function(data, mapping, ...) {
+    ggplot(data = data, mapping = mapping) + 
+        stat_density_2d(aes(fill = ..density..),
+                        geom = "raster",
+                        contour = FALSE) +
+        scale_fill_distiller(palette = "Greys",
+                             direction = 1) 
+}
+
+posterior %>%
+    GGally::ggpairs(
+        upper = list(continuous = my_upper),
+        diag = list(continuous = my_diag),
+        lower = NULL,
+        columnLabels = c("Intercept\n(Comprehension)",
+                         "Intercept\n(Production)",
+                         "Age", "Length", "Exposure",
+                         "Levenshtein", "Age × Exposure",
+                         "Age × Levenshtein", "Exposure × Leveshtein",
+                         "Age × Exposure ×\nLevenshtein"),
+        axisLabels = "none"
+    ) +
+    theme(strip.text = element_text(size = 5))
+
 ```
@@
 
-## Posterior-predictive checks {.unnumbered}
+### Posterior-predictive checks {.unnumbered}
 
 ```{r tbl-ppc}
-#| label: tbl-ppc
-#| tbl-cap: "Model posterior predictive checks (PPC).
+#| label: fig-ppc
+#| fig-cap: "Model posterior predictive checks (PPC).
 Bars indicate the observed proportion of responses to each category (No, Understands, and Understands and Says).
@@
 y_int <- as.integer(as.integer(responses$response))
-color_scheme_set(c(rep("grey60", 2), rep("black", 4)))
+bayesplot::color_scheme_set(c(rep("grey60", 2), rep("black", 4)))
 
-ppc_bars(y_int, 
+bayesplot::ppc_bars(
+    y_int, 
     model_ppcs,
@@
     fatten = 2,
-         freq = FALSE) +
+    freq = FALSE
+) +
     scale_x_continuous(breaks = 1:3,

### 2023-05-24 R/participants.R

@@ -12,9 +12,10 @@ get_participants <- function(bvq_data,
                              other_threshold = 0.1) {
     
     participants <- bvq_data$logs |>
-        dplyr::filter(completed,
+        dplyr::filter(
+            completed,
             # get only short versions of the questionnaire
-            str_detect(version, "Lockdown|Short"),
+            grepl("Lockdown|Short", version),
             # get only data from complete questionnaire responses
             # rlang::.env makes sure we use the objects provided in the arguments
             # of the function, and not variables in the piped data frame
@@ -24,13 +25,14 @@ get_participants <- function(bvq_data,
             # make sure that degrees of exposure are between 0 and 1
             between(doe_spanish, 0, 1),
             between(doe_catalan, 0, 1),
-            between(doe_others, 0, 1)) |>
+            between(doe_others, 0, 1)
+        ) |>
         mutate(time = as.integer(time)) |> 
         # see ?bvq::get_longitudinal
-        get_longitudinal(longitudinal = longitudinal) |>
+        bvq::get_longitudinal(longitudinal = longitudinal) |>
         select(id, time, time_stamp = date_finished, list = version,
                age, lp, doe_catalan, doe_spanish, edu_parent) |> 
-    arrange(id)
+        arrange(id)
     
     # export data
     save_files(participants,

### 2023-05-24 R/predictions.R

@@ -1,7 +1,9 @@
 #' Generate dataset with posterior predictions from a brmsfit model
 #'
 #' @param model A brmsfit object
+#' 
 #' @param data Data with which the model was fist, containing the original variables
+#' 
 #' @param group Group for which posterior predictions will be generated:
 #' * `NULL` (default): posterior predictions will be generated for fixed effect.
 #' * `"te"`: posterior predictions will be generated for translation equivalents.
@@ -17,7 +19,7 @@ generate_newdata <- function(model,
     # some checks
     if (!is.null(group)) {
         if (!(group %in% c("te", "id")) | length(group) > 1) {
-            cli_abort("group must be one of 'te' or 'id")
+            cli::cli_abort("group must be one of 'te' or 'id")
         }
     }
     
@@ -26,24 +28,32 @@ generate_newdata <- function(model,
         levels_in_data <- levels %in% data[[group]]
         if (!all(levels_in_data)) {
             missing_levels <- paste0(levels[!levels_in_data], collapse = ", ")
-            cli_abort("Level {missing_levels} in {.field {group}} is missing")
+            cli::cli_abort("Level {missing_levels} in {.field {group}} is missing")
         }
     }
     
     # data frame with prediction combinations
     if (is.null(group)) {
-        nd <- datagrid(model = model, te = NA, id = NA, ...)
+        nd <- marginaleffects::datagrid(model = model, te = NA, id = NA, ...)
     } else {
         # expand predictor levels in data frame to generate predictions
         if (group == "te") {
             # if group is "te", generate predictions for each level of `te`
-            nd <- datagrid(model = model, te = levels, id = NA, ...) |>
+            nd <- marginaleffects::datagrid(
+                model = model, 
+                te = levels, 
+                id = NA, 
+                ...) |>
                 select(-lv_std) |>
                 left_join(distinct(data, te, lv_std))
         }
         if (group == "id") {
             # expand predictor levels in data frame to generate predictions
-            nd <- datagrid(model = model, id = levels, te = NA, ...) |>
+            nd <- marginaleffects::datagrid(
+                model = model, 
+                id = levels,
+                te = NA, ...
+            ) |>
                 select(-age_std) |>
                 left_join(distinct(data, id, age_std))
         }
@@ -63,14 +73,16 @@ posterior_predictions <- function(model, data, ...) {
     newdata <- generate_newdata(model, data, ...)
     
     # use marginaleffects to get posterior means
[…]

### 2023-05-24 src/bvq.R

@@ -1,5 +1,5 @@
 #' Get BVQ data
-#' @param ... Arguments to be passed to \code{bvq_responses}
+#' @param ... Arguments to be passed to [bvq::bvq_responses()]
 #' @returns A named list of data frames containing questionnaire responses, participant data, and item data from BVQ
 get_bvq <- function(...) {
     # get bvq participant data

### 2023-05-24 src/helpers.R

@@ -35,36 +35,34 @@ unmake <- function(keep_fits = TRUE) {
 #' Print welcome message in console
 #' 
 welcome_message <- function() {
-    if (interactive()) {
-        id <- cli_status("")
-        cli({
-            cli_h1("trajectories")
-            cli_text("This project works with the {.pkg renv} and 
+    id <- cli_status("")
+    cli({
+        cli_h1("trajectories")
+        cli_text("This project works with the {.pkg renv} and 
                      {.pkg targets} R packages.")
-            cli_text()
-            cli_li("{.pkg renv} manages package dependencies so that you can
+        cli_text()
+        cli_li("{.pkg renv} manages package dependencies so that you can
                    install the packages needed to run the code in this repository
                    in their correct versions, without touching the packages that
                    you already have installed in your machine. This ensures that
                    the code and R packages that you use in your other projects is
                    not affected by running the code in this project.")
-            cli_text()
-            cli_li("{.pkg targets} manages reproducible workflows. The file 
+        cli_text()
+        cli_li("{.pkg targets} manages reproducible workflows. The file 
                    {.path _targets.R} contains the commands to be run in order to get the 
                    outcomes and byproducts of the code in this repository. These
                    functions are defined in the scripts in {.path R/}. See the
                    instructions below to run the workflow.")
-            
-            cli_h2("Instructions")
-            cli_ol(c("Run {.code renv::restore()}", 
-                     "Restart the R session",  
-                     "Run {.code tar_make()}"))
-            cli_h2("Repository info")
-            cli_ul(c("Lab notes: {.url gongcastro.github.io/trajectories}",
-                     "URL: {.url https://github.com/gongcastro/trajectories}",
-                     "OSF: {.url https://osf.io/hy984/}"
-            ))
-        })
-        if (interactive()) cli_status_clear(id)
-    }
+        
+        cli_h2("Instructions")
+        cli_ol(c("Run {.code renv::restore()}", 
+                 "Restart the R session",  
+                 "Run {.code tar_make()}"))
+        cli_h2("Repository info")
+        cli_ul(c("Lab notes: {.url gongcastro.github.io/trajectories}",
+                 "URL: {.url https://github.com/gongcastro/trajectories}",
+                 "OSF: {.url https://osf.io/hy984/}"
+        ))
+    })
+    cli_status_clear(id)
 }

### 2023-05-25 src/helpers.R
Rename files

@@ -5,7 +5,7 @@ make <- function() {
         job::export("none")  # return nothing
     },
     import = NULL,
-    title = "Trajectories")
+    title = "cognate-beginnings")
     
     to_delete <- c("manuscript/QTDublinIrish.otf")
     for(i in to_delete) {
@@ -37,7 +37,7 @@ unmake <- function(keep_fits = TRUE) {
 welcome_message <- function() {
     id <- cli_status("")
     cli({
-        cli_h1("trajectories")
+        cli_h1("cognate-beginnings")
         cli_text("This project works with the {.pkg renv} and 
                      {.pkg targets} R packages.")
         cli_text()
@@ -59,8 +59,8 @@ welcome_message <- function() {
                  "Restart the R session",  
                  "Run {.code tar_make()}"))
         cli_h2("Repository info")
-        cli_ul(c("Lab notes: {.url gongcastro.github.io/trajectories}",
-                 "URL: {.url https://github.com/gongcastro/trajectories}",
+        cli_ul(c("Lab notes: {.url gongcastro.github.io/cognate-beginnings}",
+                 "URL: {.url https://github.com/gongcastro/cognate-beginnings}",
                  "OSF: {.url https://osf.io/hy984/}"
         ))
     })

### 2023-05-25 manuscript/manuscript.qmd
Rename files

@@
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
-Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data [@barr2013random; see @eq-linear].
+Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data [see @eq-linear, @barr2013random].
 We specified a weakly informative prior around the parameters of the model (see Appendix C for details).

### 2023-05-25 manuscript/_appendix/_model-details.qmd
Make site

@@
 
-```{r apafg}
+```{r fig-prior-mean}
 #| label: fig-prior-mean

### 2023-05-25 manuscript/manuscript.qmd
Make site

@@
 format:
-  docx:
-    reference-doc: "templates/reference.docx"
-    knitr:
-      opts_chunk:
-        dev: "png"
-        dev-args:
-          - "cairo-png"
-        dpi: 300
-        out.width: "80%"
   pdf:
@@
 Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilinguala].
+Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
 These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical and the phonological level.
@@
 Not all words occur in the linguistic input with identical frequency, so children accumulate experience with some word-forms at faster rates than with other word-forms.
-This results in frequent words being acquired earlier in age than less frequent words [e.g., @ambridge2015ubiquity; @goodman2008doesa; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
+This results in frequent words being acquired earlier in age than less frequent words [e.g., @ambridge2015ubiquity; @goodman2008does; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
 For this reason, lexical frequency plays a central role in most models of lexical growth, especially in the case of accumulator models [@hidaka2013computational; @mollica2017how].
@@
 Bilinguals share fundamentally similar learning mechanisms with their monolinguals peers.
-Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilinguala].
+Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilingual].
 One of such divergences is the fact that bilinguals' frequency of exposure to a word-form is mediated by their degree of exposure to their languages.
@@
 #| fig-cap: "Schematic representation of the hypothesised cognate facilitation effect on word acquisition."
-#| echo: FALSE
+#| fig-width: 8
+#| fig-height: 7
 #| out-width: 100%
-#| fig-width: 6
-#| fig-height: 5
-knitr::include_graphics(normalizePath("assets/img/diagram.png"))
+#| out-height: 100%
+#| echo: false
+#| fig-align: center
+threshold <- 300
+
+# no parallel activation
+eli_df <- generate_eli(
+    threshold = threshold,
+    age = seq(1, 50, length.out = 100),
+    freq_month = 50,
+    freq_beta = 1,
+    l1_doe = 0.60,
+    parallel_beta = 0,
+    conditions = c(
+        glue::glue("**Cognate**: /<i style='color:{clrs[1]}'>ˈgat</i>-<i style='color:{clrs[4]}'>ˈga.to</i>/ [cat]"),
+        glue::glue("**Non-cognate**: /<i style='color:{clrs[1]}'>ˈgos</i>-<i style='color:{clrs[4]}'>ˈpe.ro</i>/ [dog]")
+    )
+) |>
+    filter(hypothesis=="H0") |> 
+    mutate(language = ifelse(language=="L1", "Catalan form", "Spanish form"))
+
+aoa_df <- generate_aoa(eli_df, threshold = threshold) |>
+    mutate(across(aoa, lst(min, max)), .by = c(te, language)) |>
+    filter(hypothesis=="H0")
+# 
+# img <- c(cat = "images/diagram_cat.png",
+#          dog = "images/diagram_dog.png") |>
+#     map(magick::image_read) |>
+#     map(\(x) magick::image_ggplot(x, interpolate = FALSE))
+
+plot_h0 <- eli_df |>
+    ggplot(aes(age, eli,
+               colour = language)) +
+    facet_grid(~ te) +
+    geom_segment(data = aoa_df,
+                 aes(x = aoa,
+                     xend = aoa,
+                     y = 0,
+                     yend = threshold),
+                 linewidth = 3/4) +
+    geom_hline(yintercept = aoa_df$threshold,
+               linewidth = 3/4,
+               colour = "grey80") +
+    annotate(geom = "label",
+             label = glue::glue("Threshold: {threshold}"),
+             label.r = unit(0, "lines"),
+             label.size = 0,
+             fill = "grey80",
+             x = 0,
+             y = threshold,
+             hjust = 0) +
+    geom_line(linewidth = 1) +
+    geom_label(data = aoa_df,
+               aes(x = aoa,
+                   y = threshold,
+                   label = round(aoa, 2),
+                   fill = language),
+               size = 3,
+               label.r = unit(0, "lines"),
+               label.size = 0,
+               colour = "white",
+               na.rm = TRUE,
+               show.legend = FALSE
+    ) +
+    labs(x = "Age (months)",
+         y = "Cumulative\nlearning instances",
+         colour = "",
+         shape = "Hypothesis",
+         fill = "",
+         title = "H0: no parallel activation",
+         linetype = "Hypothesis") +
+    scale_x_continuous(breaks = seq(min(eli_df$age),
+                                    max(eli_df$age), 4)) +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    theme(legend.position = "bottom",
+          legend.key.width = unit(1.5, "cm"),
+          legend.margin = margin(c(0, 0, 0, 0)),
+          legend.justification = c(0, 1),
+          legend.title = element_text(size = 9),
+          legend.text= element_text(size = 8))
+
+# with parallel activation
+eli_df <- generate_eli(
+    threshold = threshold,
+    age = seq(0, 50, by = 1),
+    l1_doe = 0.60,
+    freq_month = 50,
+    freq_beta = 1,
+    conditions = c(
+        glue::glue("**Cognate**: /<i style='color:{clrs[1]}'>ˈgat</i>-<i style='color:{clrs[4]}'>ˈga.to</i>/ [cat]"),
+        glue::glue("**Non-cognate**: /<i style='color:{clrs[1]}'>ˈgos</i>-<i style='color:{clrs[4]}'>ˈpe.ro</i>/ [dog]")
+    )
+) |>
+    filter(hypothesis=="H1") |> 
+    mutate(language = ifelse(language=="L1", "Catalan form", "Spanish form"))
+
+aoa_df <- generate_aoa(eli_df, threshold = 250) |>
+    mutate(across(aoa, lst(min, max)), .by = c(te, language)) |>
+    filter(hypothesis=="H1")
+
+plot_h1 <- eli_df |>
+    ggplot(aes(age, eli,
+               colour = language)) +
+    facet_grid(~ te) +
+    geom_segment(data = aoa_df,
+                 aes(x = aoa,
+                     xend = aoa,
+                     y = 0,
+                     yend = threshold),
+                 linewidth = 3/4) +
+    geom_hline(yintercept = aoa_df$threshold,
+               linewidth = 3/4,
+               colour = "grey80") +
+    annotate(geom = "label",
+             label = glue::glue("Threshold: {threshold}"),
+             label.r = unit(0, "lines"),
+             label.size = 0,
+             fill = "grey80",
+             x = 0,
+             y = threshold,
+             hjust = 0) +
+    geom_line(linewidth = 1) +
+    geom_label(data = aoa_df,
+               aes(x = aoa,
+                   y = threshold,
+                   label = round(aoa, 2),
+                   fill = language),
+               size = 3,
+               label.r = unit(0, "lines"),
+               label.size = 0,
+               colour = "white",
+               na.rm = TRUE,
+               show.legend = FALSE
+    ) +
+    labs(x = "Age (months)",
+         y = "Cumulative\nlearning instances",
+         colour = "",
+         shape = "Hypothesis",
+         fill = "",
+         title = "H1: parallel activation",
+         linetype = "Hypothesis") +
+    scale_colour_manual(values = clrs[c(1, 4)]) +
+    scale_x_continuous(breaks = seq(min(eli_df$age),
+                                    max(eli_df$age), 4)) +
+    theme(legend.position = "none")
+
+plot_h0 + plot_h1 +
+    plot_layout(ncol = 1, guides = "collect") &
+    plot_annotation(tag_levels = "A") +
+    scale_colour_manual(values = clrs[c(1, 4)]) &
+    scale_fill_manual(values = clrs[c(1, 4)]) &
+    scale_y_continuous(labels = function(x) format(x, big.mark = ",")) &
+    guides(colour = guide_legend(override.aes = list(linewidth = 2))) &
+    theme_ggdist() &
+    theme(legend.position = "top", 
+          legend.text = element_text(size = 10),
+          legend.key.width = unit(2, "cm"),
+                    legend.key.height = unit(0.5, "cm"),
+          strip.background = element_rect(fill = "grey90", colour = "grey90"),
+          strip.text = ggtext::element_markdown(size = 10),
+          panel.grid.major.y = element_line(colour = "grey",
+                                            linetype = "dotted"),
+          plot.background = element_rect(fill = "white",
+                                         colour = NA))
+
 ```
@@
 To operationalise participants' exposure rate to a word-form, we used a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it), and child's amount of exposure to the language such word belongs to.
-This measure was calculated by multiplying the word-form's lexical frequency with participant's degree of exposure to the language the word belongs to.
+This measure was calculated by multiplying the word-form's Zipf-transformed lexical frequency to participant's degree of exposure to the language the word belongs to.
 Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childesdb] (see Appendix A for more details).
@@
 
-Finally, we included word length as a nuance predictor in our analyses, motivated by previous studies reporting a strong association between word length and vocabulary size estimates [e.g., @braginsky2019consistency; @jones2019children].
-Sorter word-forms like /\textipa{sO\textltilde}/ (Catalan for *sun*) are generally simpler to encode as phonological representations in the lexicon than longer word-forms like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*), and are therefore more likely to be participant a child's vocabulary.
+Finally, we included word length (numbwe of phonemes) as a nuance predictor in our analyses, motivated by previous studies reporting a strong association between word length and vocabulary size estimates [e.g., @braginsky2019consistency; @jones2019children].
+Shorter word-forms like /\textipa{sO\textltilde}/ (Catalan for *sun*) are generally simpler to encode as phonological representations in the lexicon than longer word-forms like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*), and are therefore more likely to be participant a child's vocabulary.
 
@@
 
-All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/hy984/)) and GitHub ([https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories)) repositories.
+All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/hy984/)) and GitHub ([https://github.com/gongcastro/cognate-beginnings](https://github.com/gongcastro/cognate-beginnings)) repositories.
 This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.
@@
 This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest.
-We added $\text{Age}$ (number of months elapsed between participants' birth date and questionnaire completion), $\text{Length}$ (number of phonemes in the X-SAMPA phonological transcription of the word-form), $\text{Exposure}$ (composite measure of lexical frequency and language degree of exposure), and $\text{Cognateness}$ (a measure of cognateness, defined as the phonological similarity between translation equivalents) as population-level effects, together with their two- and three-way interactions.
+We added *Age* (number of months elapsed between participants' birth date and questionnaire completion), *Length* (number of phonemes in the X-SAMPA phonological transcription of the word-form), *Exposure* (composite measure of Zipf-transformed lexical frequency and language degree of exposure), and *Cognateness* (a measure of cognateness, defined as the phonological similarity between translation equivalents) as population-level effects, together with their two- and three-way interactions.
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
@@
 
-It might be argued that our results reflect the fact that cognates are represented in the initial bilingual lexicon as the *same* lexical entry.
+It might be argued that our results reflect the fact that cognate translation equivalents are represented in the initial bilingual lexicon as the *same* lexical entry.
 Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the presence of a cat), it is possible that they classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
@@
 This mechanism could potentially explain the earlier age-of-acquisition effect of cognates found in the present study, without the need of parallel activation playing any relevant role.
-Previous work points in a different direction: bilingual toddlers show monolingual-level sensitivity slight phonetic changes in a word-form, according to their performance in word recognition tasks [@swingley200511montholds; @ramon-casas2017minimalpair; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels; @ramon-casas2010are; @ramon-casas2009vowel], and word learning [@ramon-casas2017minimalpair; @mattock2010first; @havy2016phonetic].
-Overall, there is compelling evidence that bilingual toddlers form distinct lexical representations for even near-identical cognates.
+@mitchell2022cognates discuss this possibility as a possible explanation of the cognate facilitation effect, in which bilinguals only need to map one word-form to the referent in the case of cognates, while mapping two distinct word-forms in the case of non-cognates.
+Previous work on mispronunciation perception and minimal pair learning points in a different direction.
+Bilingual toddlers show monolingual-level sensitivity slight phonetic changes in a word-form, according to their performance in word recognition tasks [@swingley200511montholds; @ramon-casas2017minimalpair; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels; @ramon-casas2010are; @ramon-casas2009vowel].
+These fine abilities to differentiate between similar-sounding word-forms are reflected in word learning, as bilinguals seem to be able to map minimal pairs to distinct referents [@ramon-casas2017minimalpair; @mattock2010first; @havy2016phonetic].
+Overall, it seems that bilinguals consider small differences in the phonological forms of words as relevant at the lexical level.
+We argue that this shows evidence that bilingual toddlers likely form distinct lexical representations for even near-identical cognates.
 

### 2023-05-25 R/figures.R

@@ -1,0 +1,67 @@
+#' Generate cumulative learning instances
+#' 
+generate_eli <- function(conditions = c("**Cognate**", "**Non-cognate**"),
+                         items = c("gat /gat/",
+                                   "gato /gato/",
+                                   "gos /gos/",
+                                   "perro /ˈpe.ro/"),
+                         age = 0:34,
+                         freq_month = 5,
+                         freq_beta = 0.1,
+                         parallel_beta = 1,
+                         threshold = 500,
+                         l1_doe = 0.65)
+{
+    
+    freq_adj <- freq_month / (length(age)-1)
+    
+    item_df <- tibble(te = conditions,
+                      l1.item = items[c(1, 3)],
+                      l2.item = items[c(2, 4)]) |>
+        mutate(lv = c(stringdist::stringsim("gat", "gato"),
+                      stringdist::stringsim("gos", "pero"))) |>
+        mutate(l1.doe = l1_doe, l2.doe = 1-l1_doe)
+    
+    eli <- expand_grid(item_df,
+                       age = age) |>
+        mutate(n_month = rpois(n(), freq_adj) * age,
+               l1.h0_eli = n_month * l1.doe,
+               l2.h0_eli = n_month * l2.doe,
+               l1.h1_eli = l1.h0_eli + parallel_beta*(lv * l2.h0_eli),
+               l2.h1_eli = l2.h0_eli + parallel_beta*(lv * l1.h0_eli),
+               across(matches("h0|h1"),
+                      \(x) ifelse(x < 0 , 0, x))) |>
+        mutate(across(matches("h0|h1"), cumsum),
+               .by = te) |>
+        select(te, lv, age, matches("l1|l2")) |>
+        pivot_longer(-c(te, lv, age),
+                     names_to = c("language", ".value"),
+                     names_pattern = "(.+)\\.(.+)") |>
+        mutate(language = toupper(language)) |>
+        pivot_longer(matches("h0|h1"),
+                     names_to = c("hypothesis", ".value"),
+                     names_pattern = "(.+)_(.+)",
+                     names_transform = toupper) |>
+        rename_with(tolower) |>
+        arrange(te, language, hypothesis, age)
+    
+    return(eli)
+}
+
+#' Generate simulated ages-of-acquisition from cumulative learning instances
+#' 
+generate_aoa <- function(eli, threshold = 500) {
+    
+    aoa <- eli |>
[…]

### 2023-05-25 docs/index.qmd
Make site

@@
 ---
-title: "Trajectories"
+title: "Cognate beginnings to bilingual lexical acquisition"
 subtitle: "Lab notes"
-date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
-authors:
+short-title: "Cognate beginnings"
+author:
   - name: Gonzalo Garcia-Castro
+    corresponding: true
     orcid: 0000-0002-8553-4209
     email: gonzalo.garciadecastro@upf.edu
-    corresponding: true
     affiliations:
-      - name: Universitat Pompeu Fabra
+      - id: cbc
+        name: Universitat Pompeu Fabra
         department: Center for Brain and Cognition
-        address: Ramon Trias fargas 25-27
+        address: Ramon Trias Fargas 25-27
         city: Barcelona
-        state: Spain
-        postal-code: 08005
-  - name: Daniela S. Ávila-Varela
+        region: Spain
+        postal-code: "08005"
+  - name: Daniela S. Avila-Varela
     orcid: 0000-0002-3518-8117
     affiliations: 
-      - name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
+      - ref: cbc
+      - id: ul
+        name: Universidade de Lisboa
+        department: Facultade de Letras
+        address: Alameda da Universidade
+        city: Lisboa
+        region: Portugal
+        postal-code: "1600-214"
   - name: Ignacio Castillejo
@@
     affiliations: 
-      - name: Universidad Autónoma de Madrid
+      - id: uam 
+        name: Universidad Autónoma de Madrid
         department: Departamento de Psicología
-  - name: Nuria Sebastian-Galles
+        address: Iván Pavlov, 6
+        city: Madrid
+        region: Spain
+        postal-code: "28049"
+  - name: Núria Sebastian-Galles
     orcid: 0000-0001-6938-2498
     affiliations: 
-      - name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
-format:
-    html:
-        code-copy: hover
-        reference-location: margin
-        code-overflow: scroll
-link-external-newwindow: true
-code-fold: true
-code-line-numbers: true
-fig-dpi: 1000
-page-layout: full
-warning: false
-echo: false
-toc: true
-colorlinks: true
-number-sections: true
+      - ref: cbc
+date: "May 10th, 2023"
+code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
+keywords:
+  - lexical acquisition
+  - vocabulary
+  - bilingualism
+  - item response theory
+  - bayesian
 csl: "../assets/apa7.csl"
 bibliography: "../assets/references.bib"
+link-citations: true
+editor: source
+
+format:
+  html: default
+
+abstract: |
+    Language non-selectivity is prominent feature of bilingual lexical processing.
+An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates.
+The specific mechanisms behind such facilitation are unclear.
+We present an account of early bilingual vocabulary growth in which cognateness interacts with lexical frequency and language exposure to facilitate the word acquisition.
+We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals toddlers.
+We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
+We found cognateness facilitation effect only in low-frequency words.
+Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
+
+fig-dpi: 1000
+
 ---
 
-```{r}
+```{r setup}
 #| label: setup
-#| include: false
-
-# import packages
-library(conflicted)
+#| echo: false
+#| message: false
+#| warning: false
+# load objects -----------------------------------------------------------------
+
+library(patchwork)
+library(gt)
 library(ggplot2)
-library(patchwork)
-library(dplyr)
-library(tidyr)
-library(purrr)
-library(gt)
-library(forcats)
-library(stringr)
-library(tibble)
-library(tidybayes)
-library(janitor)
-library(glue)
-library(brms)
-library(scales)
-library(marginaleffects)
-library(lubridate)
-library(bvqdev)
-
-# load targets as R objects
-tar_load(participants)
-tar_load(responses)
-tar_load(items)
-tar_load(bvq_data)
-tar_load(model_fit_0)
-tar_load(model_fit_1)
-tar_load(model_fit_2)
-tar_load(model_fit_3)
-tar_load(model_fit_4)
-tar_load(model_fit_4_prior)
-tar_load(model_fit_prior)
-tar_load(model_loos)
-tar_load(posterior_draws)
-tar_load(posterior_draws_re)
-tar_load(marginal_effects_epreds)
-
-# resolve namespace conflicts
-conflict_prefer("filter", "dplyr")
-conflict_prefer("select", "dplyr")
-
-# set ggplot theme and colour palette
-theme_set(theme_custom()) # set custom ggplot theme
+
+targets::tar_config_set(store = here::here("_targets"),
+                        script = here::here("_targets.R"))
+
+targets::tar_load_globals()
+
+targets::tar_load(c(bvq_data, childes, items, 
+                    participants, responses))
+
+targets::tar_load(c(model_fit, model_summary, model_epreds,
+                    model_convergence, model_fit_prior, 
+                    model_ppcs, posterior_doe_summary,
+                    model_draws, model_summary,
+                    posterior_doe_draws))
+
+targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))
+
+# process objects
+items <- bvq_data$pool  |>  
+    mutate(item = str_remove(item, "cat_|spa_")) |> 
+    select(item, language, semantic_category, class) |> 
+    right_join(items, by = c("item", "language"))
+
+rope_interval <- c(lower = -0.1, upper = 0.1)
+
+# set ggplot theme and colour palette ------------------------------------------
+
+my_theme <- theme_minimal() +
+    theme(panel.grid = element_line(colour = "grey",
+                                    linetype = "dotted"),
+          axis.line = element_line(colour = "black"),
+          text = element_text(size = 12, colour = "black"),
+          axis.text = element_text(colour = "black"))
+
+theme_set(my_theme)
+
+bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))
+
 clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+
+set.seed(888)
 
@@
         ggplot2.discrete.colour = clrs,
-        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
-        ggplot2.continuous.colour = ggplot2::scale_color_gradient)
-
-rope_interval <- c(lower = -0.1, upper = +0.1)
+        ggplot2.continuous.fill = scale_color_gradient,
+        ggplot2.continuous.colour = scale_color_gradient,
+        knitr.graphics.error = FALSE) 
 ```
@@
 #| warning: false
-data("pool", package = "bvqdev")
+data("pool", package = "bvq")
 n_categories <- distinct(pool, semantic_category) %>% nrow()
@@
     mutate(id = as.character(id)) |> 
-    left_join(bvq_data$logs) %>%
+    left_join(bvq_data$logs,
+              by = join_by(id, time, age, lp, doe_catalan, doe_spanish, edu_parent)) %>%
     mutate(
@@
             edu_parent = unique(bvq_data$logs$edu_parent)
-        )
+        ),
+        by = join_by(doe_catalan, edu_parent)
     ) %>%
@@
     gt() |> 
-    summary_rows(columns = Complementary:Primary,
-                 fns = list(Total = ~sum(., na.rm = TRUE)),
-                 formatter = fmt_integer) %>%
     cols_label(doe_catalan = "DoE Catalan") %>%
@@
                                   color = "grey50")),
-        locations = cells_body(columns = 9,
+        locations = cells_body(columns = 8,
                                rows = 1:5)
@@
     tab_style(cell_text(weight = "bold"),
-              cells_column_spanners(everything())) %>%
-    summary_rows(fns = list("Grand mean" = ~mean(.)),
-                 columns = ends_with("_mean")) %>%
-    summary_rows(fns = list("N" = ~sum(.)),
-                 columns = n,
-                 formatter = fmt_integer)
+              cells_column_spanners(everything()))
 ```
@@
                n, " (",
-               percent(prop, accuracy = 0.1), 
+               scales::percent(prop, accuracy = 0.1), 
                ")")) %>% 
@@
          fill = "Language") +
-    scale_y_continuous(labels = percent) +
+    scale_y_continuous(labels = scales::percent) +
     theme(
@@
          fill = "Language") +
-    scale_y_continuous(labels = percent) +
+    scale_y_continuous(labels = scales::percent) +
     scale_x_continuous(breaks = seq(0, 16, 1)) +
@@
     scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_x_continuous(labels = percent) +
+    scale_x_continuous(labels = scales::percent) +
     scale_y_continuous(limits = c(-4, 4)) +
@@
 
-```{r}
+```{r fig_prior-mean}
 #| label: fig-prior-mean
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 3
+#| out-width: 90%
+#| fig-cap: "Expected prior-predicted mean."
+nd <- expand.grid(n_phon_std = 0,
+                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
+                  exposure_std = 0,
+                  lv_std = 0)
+
+prior_epreds <- add_epred_draws(nd, 
+                                model_fit_prior,
+                                ndraws = 50, 
+                                re_formula = NA) |> 
+    filter(.category != "No") |> 
+    pivot_wider(names_from = ".category", values_from = ".epred") |> 
+    mutate(Understands = Understands + `Understands and Says`) |> 
+    pivot_longer(c(Understands, `Understands and Says`),
+                 names_to = ".category",
+                 values_to = ".epred") |> 
+    mutate(.category = factor(.category,
+                              levels = c("Understands",
+                                         "Understands and Says"),
+                              labels = c("Comprehension\n(Understands)",
+                                         "Production\n(Understands and Says)")))
+
+prior_epreds |> 
+    ggplot(aes(x = age_std, y = .epred)) +
+    facet_wrap(~ .category) +
+    geom_line(linewidth = 3/4,
+              alpha = 1/2,
+              aes(group = .draw),
+              colour = "grey50") + 
+    labs(x = "Age (months)", 
+         y = "Probability of acquisition",
+         colour = "Credible interval", 
+         fill = "Credible interval") +
+    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
+                                      mean(responses$age), 
+                                      sd(responses$age))[, 1],
+                       labels = seq(10, 36, 4)) +
+    theme(legend.position = "top",
+          axis.ticks.x = element_line(colour = "black"),
+          panel.grid = element_blank())
+```
+
+#### Prior-predictions
+
+```{r}
+#| label: fig-prior-predictions
 #| echo: false
@@
 
-m <- add_epred_draws(nd, 
-                     model_fit_4_prior, 
-                     ndraws = NULL,
-                     re_formula = NA) %>% 
-    filter(.category != "No") %>% 
-    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
-    mutate(
-        .category = factor(
-            .category,
-            levels = c("Understands and Says", "Understands"),
-            labels = c("Production\n(Understands and Says))", "Comprehension\n(Understands)")
-        )
-    )
-
-m %>% 
-    ggplot(aes(x = age_std, y = .epred)) +
-    facet_wrap(~.category) +
-    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
-    stat_summary(fun = "mean", geom = "line", size = 0.75) +
-    scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
-        labels = seq(10, 36, 4)
-    ) +
-    scale_fill_manual(
-        values = rev(clrs), 
-        labels = c("95%", "89%", "67%", "50%")
-    ) +
-    labs(
-        x = "Age (months)", 
-        y = "Probability of acquisition",
-        colour = "Credible interval", 
-        fill = "Credible interval"
-    ) +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
-    theme(
-        legend.position = "top",
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid = element_blank()
-    )
-```
-
-#### Prior-predictions
-
-```{r}
-#| label: fig-prior-predictions
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected prior predictions"
-nd <- expand.grid(
-    n_phon_std = 0,
-    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
-    exposure_std = 0,
-    lv_std = 0
-)
-
-m <- add_predicted_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA)
+m <- add_predicted_draws(nd, model_fit_prior, ndraws = NULL, re_formula = NA)
 
@@
     mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
+    pivot_longer(c(No, Understands, `Understands and Says`),
+                 names_to = ".prediction", 
+                 values_to = "n") %>%
     ggplot(aes(x = age_std, y = n, fill = .prediction)) +
@@
     scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(labels = percent) +
+    scale_y_continuous(labels = scales::percent) +
     theme(
@@
 
-### Model comparison
-
-To test of goodness of fit of our model, we compared its predictive accuracy of this extended model--which included all predictors of interest--to four simpler models that included a subset of our predictors of interest: a base model that only included the main effects of $Age$, $Phonemes$, and $Exposure$ ($\mathcal{M}_{0}$), a model that also included the $Age \times Exposure$ interaction ($\mathcal{M}_{1}$), a model that also included the main effect of $Levenshtein$ ($\mathcal{M}_{2}$), a model that also included the $Age \times Levenshtein$ interaction ($\mathcal{M}_{3}$).
-Only the extended model included the $Age \times Exposure \times Levenshtein$ interaction.
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package.
-LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
-
 ### Model settings
@@
 -   `brms` backend: CmdStanR (`cmdstanr` `r sessionInfo()$loadedOnly$cmdstanr$Version`)
--   Chains: `r dim(model_fit_4$fit)[2]`
--   Cores: `r dim(model_fit_4$fit)[2]`
--   Iterations: `r format(dim(model_fit_4$fit)[1], big.mark = ",")` (`r format(dim(model_fit_4$fit)[1]/2, big.mark = ",")`)
+-   Chains: `r dim(model_fit$fit)[2]`
+-   Cores: `r dim(model_fit$fit)[2]`
+-   Iterations: `r format(dim(model_fit$fit)[1], big.mark = ",")` (`r format(dim(model_fit$fit)[1]/2, big.mark = ",")`)
 
@@
 
-::: panel-tabset
-
-#### Model 0
-
-```{r}
-print(model_fit_0$formula)
-```
-
-#### Model 1
-
-```{r}
-print(model_fit_1$formula)
-```
-
-#### Model 2
-
-```{r}
-print(model_fit_2$formula)
-```
-
-#### Model 3
-
-```{r}
-print(model_fit_3$formula)
-```
-
-#### Model 4
-
-```{r}
-print(model_fit_4$formula)
+```{r}
+print(model_fit$formula)
 ```
@@
     # summarise(prop_mean = mean(prop), .groups = "drop") %>%
-    ggplot(aes(prop, exposure, colour = lv, fill = lv)) +
-    facet_grid(~ response) +
-    stat_slab(alpha = 0.5, size = 0.2, trim = FALSE) +
-    geom_point(size = 1, alpha = 0.5, position = position_nudge(y = -0.1)) +
-    labs(x = "Proportion of responses", 
-         y = "Exposure",
+    ggplot(aes(exposure, prop, colour = lv, fill = lv)) +
+    facet_wrap(~ response, ncol = 1) +
+    geom_hline(yintercept = 0.5, linetype = "dashed") +
+    stat_dots(aes(group = lv),
+              position = position_dodge(width = 1),
+              layout = "swarm",
+              side = "both",
+              colour = "grey80",
+              fill = "grey80"
+    ) +
+    stat_summary(geom = "errorbar",
+                 fun.data = "mean_se",
+                 position = position_dodge(width = 1),
+                 width = 0.25) +
+    stat_summary(geom = "point",
+                 fun = "mean",
+                 position = position_dodge(width = 1),
+                 size = 2) +
+    labs(y = "Proportion of responses", 
+         x = "Exposure",
          fill = "Levenshtein", 
          colour = "Levenshtein") +
-    scale_x_continuous(labels = percent) +
     scale_colour_manual(values = clrs[c(1, 4, 3, 5)]) +
@@
 
-## Model selection
-
-@tbl-results-loos shows the outputs of the *LOO-CV*.
-The extended model (`model_fit_3`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference.
-This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.
-
-```{r}
-#| label: tbl-results-loos
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
-This procedure involves estimating the likelihood of each observation given a model fitted without such observation.
-The resulting likelihoods of each model were summed to compute the expected log-predicted density (ELPD) of the model, and correcting it for the number of parameters in the model to penalised for its the complexity.
-ELPD values closer to zero indicate better predictive acuracy.
-We report the ELPD, effective number of parameters (p), and information criterion (LOO-IC) of each model along their associated standard errors in paratheses.
-The first column indicates the difference in ELPD between each model and the model with the best predictive accuracy (model 4)."
-model_loos %>% 
-    loo_compare() %>% 
-    as.data.frame() %>%   
-    rownames_to_column("model") %>% 
-    relocate(
-        model, 
-        matches("diff"),
-        matches("elpd_loo"),
-        matches("p_loo"), 
-        matches("looic")
-    ) %>%
-    mutate(model = str_extract_all(model, "\\d")) %>% 
-    gt() %>% 
-    fmt_number(2:9, decimals = 1) %>%
-    cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>%
-    cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>%
-    cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>%
-    cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>%
+## Fixed effects
+
+```{r tbl-coefs}
+#| label: tbl-coefs
+#| tbl-cap: "Posterior distribution of regression coefficients."
+#| tbl-note: "Median: median of the posterior distribution in the probability scale.
+95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.
+On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE).
+This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
+Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
+#| echo: false
+#| message: false
+#| warning: false
+#| out-width: 80%
+# summarise posterior draws
+intercept_label <- "Intercepts (at 22 months)"
+
+model_summary |>
+    select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
+    mutate(across(.median:.upper,
+                  \(x) ifelse(.type==intercept_label, plogis(x), x/4))) |> 
+    gt(groupname_col = ".type", 
+       rowname_col = ".variable_name") |>
+    fmt_percent(.rope_overlap, decimals = 3) |> 
+    fmt_number(c(.median:.upper), decimals = 3) |>
+    cols_merge(c(.lower, .upper),
+               pattern = "[{1}, {2}]") |>
     cols_label(
-        model = "Model",
-        elpd_loo = md("*ELPD*"),
-        p_loo = md("*p*"),
-        looic = md("*LOO-IC*"),
-        elpd_diff = md("Difference")
-    ) %>%
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>%
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels()
-    ) %>%
-    tab_style(
-        cell_fill(color = "grey", alpha = 0.5),
-        cells_body(matches("diff"))
-    ) %>%
-    tab_style(
-        cell_fill(color = clrs[4], alpha = 0.5),
-        cells_body(rows = 1)
-    )
-```
-
-## Fixed effects
-
-```{r}
-#| label: tbl-results-fixed
-#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
-#| fig-width: 7
-intercept_footnote <- glue("Probability of comprehension or production at the grand age of the sample ({round(mean(responses$age), 2)} months")
-
-# summarise posterior draws
-posterior_draws$summary %>%
-    gt(rowname_col = ".variable_name", 
-       groupname_col = ".type") %>%
-    cols_hide(".variable") %>% 
-    fmt_number(.median:.lower, decimals = 2) %>%
-    fmt_percent(.rope_overlap, decimals = 2) %>%
-    cols_merge(c(.lower, .upper), pattern = "[{1}, {2}]") %>%
-    cols_label(.median = "Median",
-               .lower = "95% HDI",
-               .rope_overlap = "ROPE prob.") %>%
+        .variable_name = "Parameter",
+        .median = "Median",
+        .lower = md("95% HDI"),
+        .rope_overlap = md("*p*(H0)")
+    ) |>
     tab_style(cell_text(weight = "bold"),
-              cells_column_labels()) %>%
+              cells_column_labels(columns = 1:5)) |>
     tab_style(cell_text(align = "left"),
-              cells_stub())
-
-post_draws_list <- split(posterior_draws$summary, 
-                         posterior_draws$summary$.variable)
-```
-
-We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
-We decided to draw this interval around zero so that it covered values smaller than those considered as negligible by similar previous studies: after consulting the regression tables of [ADD STUDIES], we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: 3%.
-consequently, we defined our *ROPE* from -0.1% to +0.1%.
-We report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
-This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
-Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the *ROPE* interval we previously defined.
-For those coefficients that did not exclude completely the *ROPE* we also report a Bayes Factor indicating how many times is such coefficient more likely to be bigger than the upper limit of the *ROPE*, or smaller than the lower limit of the *ROPE*, given our data.
-
-For interpretability, we report the estimated regression coefficients transformed from the probability scale.
-The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
-This means that one logit difference is not necessarily translated to a unique value in the probability scale.
-For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
-The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
-We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
-derivative) of the logistic curve at the age at which most participants were acquiring a given word.
-This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
-
-@fig-results-fixed shows the outputs of the model.
-**Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 2)`, 95% *HDI* = [`r round(post_draws_list[["b_age_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std"]]$.upper, 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$.median/(4*sd(responses$age)), accuracy = 0.01)` every month.
-
-The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$.upper, accuracy = 0.01)`]).
-For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$.median/(4*sd(responses$n_phon)), accuracy = 0.01)`.
-
-**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$.upper, accuracy = 0.01)`]).
-All of the posterior samples of his regression excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/(4*sd(responses$exposure)), accuracy = 0.01)` more likely to acquire it.
-
-The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$.rope_overlap, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
-
-However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word.
-Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
-This effect was negligible when exposure to both languages was balanced (e.g.
-50-50%).
-Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.
-
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
-Given our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.
-
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$.upper, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
+              cells_title(groups = "subtitle")) |> 
+    tab_style(cell_text(size = "small"), 
+              list(cells_body(),
+                   cells_row_groups(),
+                   cells_stub())) |> 
+    tab_style(cell_text(style = "italic"),
+              cells_row_groups())
+```
 
@@
 
-posterior_draws$draws %>%
-    ggplot(aes(.value, fct_rev(.variable_name))) +
+model_draws %>%
+    ggplot(aes(.value, .variable_name)) +
     annotate(geom = "rect", ymin = -Inf, ymax = Inf,
@@
     geom_errorbar(
-        data = posterior_draws$summary,
+        data = model_summary,
         aes(xmin = .lower, 
@@
     ) +
-    geom_point(data = posterior_draws$summary, 
+    geom_point(data = model_summary, 
                aes(x = .median),
@@
     geom_text(
-        data = posterior_draws$summary,
+        data = model_summary,
         aes(x = .median, 
             y = .variable_name,
-            label = glue(
+            label = glue::glue(
                 "{round(.median, 1)}",
@@
     ) +
-    scale_x_continuous(breaks = seq(-2, 2, 0.2)) +
+    scale_x_continuous(breaks = seq(-3, 3, 0.2)) +
     scale_fill_manual(
@@
 ## Marginal effects
-
-```{r}
-#| label: fig-results-marginal
-#| fig-width: 7
-#| fig-cap: "Expected posterior predictions.
-Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
-These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model.
-Therefore, we generated 4,000 predictions for a series of combination of levels of interest.
-Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels.
-The X-axis indicates the age (in months) for which the prediction is generated.
-The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*).
-Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
-Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
-Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
-
-marginal_effects_epreds %>% 
+```{r fig-marginal}
+#| label: fig-marginal
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 4.5
+#| out-width: 100%
+#| eval: true
+#| fig-cap: "Posterior marginal effects.
+Thin lines correspond to 50 mean predictions.
+Thicker lines indicate the median of those predictions.
+Different colours indicate different levels of cognateness (phonological similarity).
+Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
+Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom.
+In-sample predictions lie inside the grey rectangles."
+clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")
+
+model_epreds |> 
+    mutate(age = (age_std * sd(responses$age)) + mean(responses$age),
+           exposure = factor(exposure_std,
+                             levels = unique(exposure_std),
+                             labels = c("Low exposure (-1 SD)",
+                                        "Mean exposure",
+                                        "High exposure (+1 SD)"),
+                             ordered = TRUE),
+           lv = factor(lv_std,
+                       levels = unique(lv_std),
+                       labels = paste0(c(0, 50, 100), "%"),
+                       ordered = TRUE),
+           .category = ifelse(.category=="Understands",
+                              "Comprehension",
+                              "Production")) |> 
     ggplot(aes(age, .value, colour = lv, fill = lv)) +
     facet_grid(.category~exposure) +
-    # stat_summary(fun.data = median_hdi, geom = "ribbon",
-    #              alpha = 0.5, colour = NA) +
+    annotate(geom = "rect",
+             xmin = min(responses$age),
+             xmax = max(responses$age),
+             ymin = -Inf,
+             ymax = Inf,
+             fill = "grey95",
+             colour = NA) +
+    annotate(geom = "text",
+             x = mean(responses$age),
+             y = 0,
+             hjust = -0.1,
+             vjust = 0,
+             label = "Observed age range",
+             size = 2) +
+    geom_vline(xintercept = mean(responses$age), 
+               linetype = "dotted") +
+    geom_hline(yintercept = 0.5,
+               linetype = "dotted") +
     geom_line(aes(group = interaction(lv, .draw)),
-              size = 0.75,
-              alpha = 0.15) +
-    stat_summary(aes(linetype = lv), 
-                 fun = median, 
-                 geom = "line",
-                 size = 1) +
-    labs(
-        x = "Age (months)",
+              linewidth = 0.5,
+              alpha = 1/10) +
+    stat_summary(fun = median, geom = "line", linewidth = 0.75) +
+    labs(x = "Age (months)",
          y = "Probability of acquisition",
-        linetype = "Phonological similarity (Levenshtein)",
-        colour = "Phonological similarity (Levenshtein)",
-        fill = "Phonological similarity (Levenshtein)"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(
-        breaks = seq(0, 1, 0.25),
-        limits = c(0, 1),
-        labels = percent
-    ) +
-    scale_x_continuous(breaks = seq(10, 32, 4)) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_line(colour = "grey", 
-                                  linetype = "dotted"),
+         linetype = "Cognateness (phonological similarity)",
+         colour = "Cognateness (phonological similarity)",
+         fill = "Cognateness (phonological similarity)") +
+    guides(colour = guide_legend(override.aes = list(linewidth = 2))) +
+    scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_fill_manual(values = clrs_reorder[c(4, 5, 1)]) +
+    scale_y_continuous(breaks = seq(0, 1, 0.25),
+                       limits = c(0, 1)) +
+    scale_x_continuous(breaks = seq(0, 50, 5)) +
+    theme(legend.position = "top",
+          legend.key.size = unit(1, "cm"),
+          legend.key.height = unit(1, "cm"),
+          axis.text.x = element_text(size = 9),
+          panel.grid = element_blank(),
           panel.border = element_rect(fill = NA, 
                                       colour = "black", 
-                                    linewidth = 0.75)
-    )
-
+                                      linewidth = 0.75))
 ```
@@
 Closer lines indicate lower uncertainty over the true value of the average marginal effect of Levenshein, while more distant lines indicate higher uncertainty."
-nd <- datagrid(
-    model = model_fit_4,
+nd <- marginaleffects::datagrid(
+    model = model_fit,
     n_phon_std = 0,
@@
 
-marg_effects <- marginaleffects(
-    model_fit_4,
+marg_effects <- marginaleffects::marginaleffects(
+    model_fit,
     newdata = nd, 
@@
 marg_effects %>% 
-    posteriordraws() %>%
+    marginaleffects::posteriordraws() %>%
     as_tibble() %>% 
@@
     scale_x_continuous(
-        labels = label_number(style_positive = "plus", suffix = " SD"),
+        labels = scales::label_number(style_positive = "plus", suffix = " SD"),
         breaks = seq(-2, 2, 0.5)
@@
     scale_y_continuous(
-        labels = label_percent(style_positive = "plus", ),
+        labels = scales::label_percent(style_positive = "plus", ),
         breaks = seq(-1, 1, 0.05)
@@
 #| fig-cap: "Posterior probability of acquisition by participant"
-re_id <- ranef(model_fit_4)$id[,,1] %>% 
+re_id <- ranef(model_fit)$id[,,1] %>% 
     as_tibble() %>% 
     rownames_to_column("id") %>%
-    mutate(id = as.integer(id)) %>% 
-    left_join(distinct(model_fit_4$data, id)) %>% 
-    clean_names() %>% 
+    left_join(distinct(model_fit$data, id),
+              by = join_by(id)) %>% 
+    janitor::clean_names() %>% 
     mutate(
@@
             .fns = list(
-                Comprehension = ~plogis(. + fixef(model_fit_4)[1]),
-                Production = ~plogis(. + fixef(model_fit_4)[2])
+                Comprehension = ~plogis(. + fixef(model_fit)[1]),
+                Production = ~plogis(. + fixef(model_fit)[2])
             ), 
@@
     scale_colour_manual(values = clrs[c(1, 4)]) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
     theme(
@@
 #| fig-cap: "Posterior random slope of age by participant"
-re_age <- ranef(model_fit_4)$id[,,"age_std"]
+re_age <- ranef(model_fit)$id[,,"age_std"]
 
@@
     as_tibble() %>% 
-    clean_names() %>% 
-    mutate(id = unique(model_fit_4$data$id)) %>% 
+    janitor::clean_names() %>% 
+    mutate(id = unique(model_fit$data$id)) %>% 
     relocate(id) %>% 
@@
 #| fig-cap: "Posterior random slope of the number of phonemes by participant"
-re_n_phon <- ranef(model_fit_4)$id[,,"n_phon_std"]
+re_n_phon <- ranef(model_fit)$id[,,"n_phon_std"]
 
@@
 #| fig-cap: "Posterior random slope of lexical frequency by participant"
-re_freq <- ranef(model_fit_4)$id[,,"exposure_std"]
+re_freq <- ranef(model_fit)$id[,,"exposure_std"]
 
@@
 #| fig-cap: "Posterior random slope of phonological similarity by participant"
-re_lv <- ranef(model_fit_4)$id[,,"lv_std"]
+re_lv <- ranef(model_fit)$id[,,"lv_std"]
 
@@
     as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
+    janitor::clean_names() %>%
+    mutate(id = unique(model_fit$data$id)) %>%
     relocate(id) %>%
@@
 # | fig-cap: "Posterior random slope of age by participant"
-re_exposure_lv <- ranef(model_fit_4)$id[,,"exposure_std:lv_std"]
+re_exposure_lv <- ranef(model_fit)$id[,,"exposure_std:lv_std"]
 
@@
 #| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
-re_age_exposure <- ranef(model_fit_4)$id[,,"age_std:exposure_std"]
+re_age_exposure <- ranef(model_fit)$id[,,"age_std:exposure_std"]
 
@@
 #| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
-re_age_lv <- ranef(model_fit_4)$id[,,"age_std:lv_std"]
+re_age_lv <- ranef(model_fit)$id[,,"age_std:lv_std"]
 
@@
 #| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
-re_age_exposure_lv <- ranef(model_fit_4)$id[,,"age_std:exposure_std:lv_std"]
+re_age_exposure_lv <- ranef(model_fit)$id[,,"age_std:exposure_std:lv_std"]
 
@@
     as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_3$data$id)) %>%
+    janitor::clean_names() %>%
+    mutate(id = unique(model_fit$data$id)) %>%
     relocate(id) %>%
@@
 #| fig-cap: "Posterior probability of acquisition by translation equivalent"
-re_te <- ranef(model_fit_4)$te[,,"Intercept"] %>% 
+re_te <- ranef(model_fit)$te[,,"Intercept"] %>% 
     as_tibble() %>% 
     rownames_to_column("te") %>% 
-    clean_names() %>% 
+    janitor::clean_names() %>% 
     mutate(
@@
             .fns = list(
-                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
-                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
+                Comprehension = ~plogis(. + fixef(model_fit)[1]),
+                Production = ~plogis(. + fixef(model_fit)[2])
             ), .names = "{.col}__{.fn}"
@@
     scale_x_continuous(limits = c(0, 1),
-                       labels = percent) +
+                       labels = scales::percent) +
     scale_fill_manual(values = clrs[c(1, 4)]) +
@@
 #| fig-cap: "Posterior random slope of age by translation equivalent"
-re_age <- ranef(model_fit_4)$te[,,"age_std"]
+re_age <- ranef(model_fit)$te[,,"age_std"]
 
@@
     as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_4$data$te)) %>% 
+    janitor::clean_names() %>% 
+    mutate(te = unique(model_fit$data$te)) %>% 
     relocate(te) %>% 
@@
 #| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
-re_n_phon <- ranef(model_fit_3)$te[,,"n_phon_std"]
+re_n_phon <- ranef(model_fit)$te[,,"n_phon_std"]
 
@@
     as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_3$data$te)) %>% 
+    janitor::clean_names() %>% 
+    mutate(te = unique(model_fit$data$te)) %>% 
     relocate(te) %>% 
@@
 #| fig-cap: "Posterior random slope of age by participant"
-re_doe <- ranef(model_fit_4)$te[,,"exposure_std"]
+re_doe <- ranef(model_fit)$te[,,"exposure_std"]
 
@@
     as_tibble() %>%
-    clean_names() %>%
-    mutate(te = unique(model_fit_4$data$te)) %>%
+    janitor::clean_names() %>%
+    mutate(te = unique(model_fit$data$te)) %>%
     relocate(te) %>%
@@
 #| fig-cap: "Posterior random slope of age by participant"
-re_age_doe <- ranef(model_fit_4)$te[,,"age_std:exposure_std"]
+re_age_doe <- ranef(model_fit)$te[,,"age_std:exposure_std"]
 
@@
 #| fig-width: 12
-gather_draws(model_fit_4, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
+gather_draws(model_fit, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
     mutate(.chain = paste("Chain ", .chain)) %>%
@@
     annotate(geom = "rect", colour = NA,
-             xmin = 0, xmax = dim(model_fit_4$fit)[1]/2,
+             xmin = 0, xmax = dim(model_fit$fit)[1]/2,
              ymin = -Inf, ymax = Inf,
@@
 
-### Leave-one-out diagnostics (Pareto-*k* diagnostics)
-
-::: panel-tabset
-
-#### Model 0
-
-```{r}
-#| label: fig-loo-diagnostics-0
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
-get_loo_plot <- function(x){
-    custom_breaks<- seq(0, length(x$pointwise[,5]), 50000)
-
-    x$pointwise[,5] %>% 
-        data.frame(x = 1:length(.), y = .) %>% 
-        ggplot(aes(x, y)) +
-        geom_point(
-            size = 0.5, 
-            alpha = 0.1,
-            colour = clrs[4], 
-            shape = 1
-        ) +
-        geom_hline(yintercept = 0.5,
-                   linetype = "dotted") +
-        labs(x = "Data point",
-             y = "Pareto-k value",
-             title = "PSIS diagnostic plot") +
-        scale_x_continuous(breaks = custom_breaks) +
-        scale_y_continuous(limits = c(-0.6, 0.6))
-}
-
-get_loo_plot(model_loos$model_fit_0)
-```
-
-#### Model 1
-
-```{r}
-#| label: fig-loo-diagnostics-1
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
-get_loo_plot(model_loos$model_fit_1)
-```
-
-#### Model 2
-
-```{r}
-#| label: fig-loo-diagnostics-2
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
-get_loo_plot(model_loos$model_fit_2)
-```
-
-#### Model 3
-
-```{r}
-#| label: fig-loo-diagnostics-3
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
-get_loo_plot(model_loos$model_fit_3)
-```
-
-#### Model 4
-
-```{r}
-#| label: fig-loo-diagnostics-4
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
-get_loo_plot(model_loos$model_fit_4)
-```
-
-:::
-
-### Gelman-Rubin diagnostic (R-hat)
-
-```{r}
-#| label: fig-diagnostics-rhat
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| eval: false
-mcmc_rhat_data(model_rhats$model_fit_4) %>% 
-    ggplot() +
-    aes(x = value) +
-    geom_histogram(bins = 100, na.rm = TRUE, fill = clrs[1], colour = "white") +
-    geom_rug(colour = clrs[1]) +
-    labs(
-        x = "R-hat",
-        y = "Number of samples",
-        colour = "Description",
-        fill = "Description"
-    ) +
-    theme(
-        legend.position = "top"
-    )
-```
-
-### Effective sample size
-
-```{r}
-#| label: fig-diagnostics-neff
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| eval: false
-mcmc_neff_data(model_neffs$model_fit_4) %>% 
-    ggplot() +
-    aes(x = value, fill = description, colour = description) +
-    geom_histogram(bins = 100, na.rm = TRUE, colour = "white") +
-    geom_rug() +
-    geom_vline(xintercept = 1, colour = "black", size = 0.75) +
-    labs(
-        x = "Effective sample size ratio",
-        y = "Number of samples",
-        colour = "Rating",
-        fill = "Rating"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
-    theme(
-        legend.position = "top"
-    )
-```
-
-# Discussion
-
-## Summary of the study
-
-## Summary of the results
-
-## Results in contrast with the literature
-
-## Mechanistic explanations
-
-## Limitations
-
-## Further steps
+### Momdel convergence
+
+```{r fig-rhats-neffs}
+#| label: fig-rhats-neffs
+#| fig-cap: "MCMC convergence diagnostic of all parameters in the model. A: distribution of the Gelman-Rubin (R-hat) scores. B: distribution of the ratio of effective sa"
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-height: 3
+#| fig-width: 6
+#| out-width: 100%
+model_convergence |> 
+    ggplot(aes(.rhat)) +
+    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
+    geom_vline(xintercept = 1.01, linetype = "dashed") +
+    labs(x = "R-hat",
+         y = "MCMC samples") +
+
+    model_convergence |> 
+    ggplot(aes(.neff)) +
+    geom_histogram(binwidth = 0.1, colour = "white", fill = "grey60") +
+    geom_vline(xintercept = 1.01, linetype = "dashed") +
+    labs(x = "Effective sample size ratio",
+         y = "MCMC samples") +
+
+    plot_layout(ncol = 1) &
+    plot_annotation(tag_levels = "A") &
+    theme(legend.position = "none",
+          panel.grid = element_blank())
+```
 

### 2023-05-25 docs/index.qmd
Fix README

@@
 
+code-fold: true
+warning: false
+
 format:
@@
 #| label: setup
-#| echo: false
 #| message: false
@@
 #| label: questionnaire
-#| echo: false
 #| message: false
@@
 #| label: fig-participants-time
-#| echo: false
 #| message: false
@@
 #| fig-width: 7
+#| fig-height: 3
 #| fig-cap: "Distribution of participants' ages"
@@
 
-```{r apatb-participants}
+```{r fig-participants}
 #| label: tab-participants
-#| echo: false
 #| message: false
@@
 #| label: fig-items-freq
-#| echo: false
 #| message: false
@@
 #| label: fig-items-nphon
-#| echo: false
 #| message: false
@@
 #| label: fig-items-lv
-#| echo: false
 #| message: false
@@
 #| label: fig-items-frequency-phonemes
-#| echo: false
 #| message: false
@@
 #| label: fig-items-frequency-lv
-#| echo: false
 #| message: false
@@
 #| label: fig-items-nphon-lv
-#| echo: false
 #| message: false
@@
 #| label: fig-prior-mean
-#| echo: false
 #| message: false
@@
 #| label: fig-prior-predictions
-#| echo: false
 #| message: false
@@
 Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
-#| echo: false
 #| message: false
@@
 ## Marginal effects
+
 ```{r fig-marginal}
 #| label: fig-marginal
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-id-age
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-id-nphon
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-id-exposure
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-id-lv
-#| echo: false
 #| message: false
@@
 # | label: fig-results-random-id-exposure-lv
-# | echo: true
 # | message: false
@@
 #| label: fig-results-random-id-age-doe
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-id-age-lv
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-id-age-exposure-lv
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-te
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-te-age
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-te-nphon
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-te-exposure
-#| echo: false
 #| message: false
@@
 #| label: fig-results-random-te-age-exposure
-#| echo: false
 #| message: false
@@
 #| label: fig-diagnostics-traceplots
-#| echo: false
 #| message: false
@@
 
-### Momdel convergence
+### Model convergence
 
@@
 #| fig-cap: "MCMC convergence diagnostic of all parameters in the model. A: distribution of the Gelman-Rubin (R-hat) scores. B: distribution of the ratio of effective sa"
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-height: 3
-#| fig-width: 6
+#| message: false
+#| warning: false
+#| fig-height: 4
+#| fig-width: 7
 #| out-width: 100%
@@
 ```{r}
-sessioninfo::session_info()
-```
+sessionInfo()
+```

### 2023-05-26 manuscript/manuscript.qmd
Fix typos

@@
 abstract: |
-    Language non-selectivity is prominent feature of bilingual lexical processing.
-An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates.
+    Language non-selectivity is a prominent feature of bilingual lexical processing.
+An instance of such non-selectivity is embodied by cognateness.
+Recent studies have pointed to cognateness as a facilitator of vocabulary growth, and to cognates being acquired at earlier ages than non-cognates.
 The specific mechanisms behind such facilitation are unclear.
-We present an account of early bilingual vocabulary growth in which cognateness interacts with lexical frequency and language exposure to facilitate the word acquisition.
-We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals toddlers.
-We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
-We found cognateness facilitation effect only in low-frequency words.
+Building on accumulator models of language acquisition, we suggest that cognateness increases the number of cumulative learning instances that a bilingual child encounters, thanks to cross-language activation.
+In line with previous studies, this model predicted a cognate facilitation, but only for words to which children are exposed less often.
+We tested these predictions against vocabulary data from 366 Catalan-Spanish bilinguals toddlers.
+We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the participant.
+We found a cognateness facilitation effect only in low-frequency words or words from the language of lower exposure.
 Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
@@
 Quite often, bilingual children encounter more than one label for the same referent: one in each language.
-For instance, an English-Spanish bilingual might hear both *dog* and *perro* in the presence of a dog We will refer to such cross-language synonyms as *translation equivalents*.^[Throughout the present study, we will make the assumption that bilinguals form a common semantic representations for each translation equivalent (e.g., the concept 'dog' is shared for the English and Spanish words *dog* and *perro*).
+For instance, an English-Spanish bilingual might hear both *dog* and *perro* in the presence of a dog We will refer to such cross-language synonyms as *translation equivalents*.^[Throughout the present study, we make the assumption that bilinguals form a common semantic representations for each translation equivalent (e.g., the concept 'dog' is shared for the English and Spanish words *dog* and *perro*).
 The referential contexts in which bilingual children hear (and learn) both labels may not be identical.
-For instance, the English word *finger* refers to eight parts of the body (excluding thumbs and toes), whereas its Spanish translation *dedos* refers to twenty body parts (fingers, thumbs, and toes).
+For instance, the English word *finger* refers to eight body parts (excluding thumbs and toes), whereas its Spanish translation *dedos* refers to twenty body parts (fingers, thumbs, and toes).
 These discrepancies are reflected, for instance, in adults' performance in lexical decision tasks [@chaouch-orozco2023are].
@@
 For simplicity, and in line with previous (formal and verbal) models of bilingual lexical processing [e.g., @shook2013bilingual; @dijkstra2019multilink; @dijkstra1998bia; @kroll2010revised], we assume common semantic representations for translation pairs.].
-
 Sometimes, the word-forms involved in a translation equivalent may be phonologically similar.
-Such form-similar translation equivalents are referred to as cognates.
+Such form-similar translation equivalents are referred to as *cognates*.
 While the previously mentioned translation pair *dog* and *perro* are non-cognates (both word-forms hardly share any phonemes), the translation pair *cat* and *gato* can be considered cognates.
@@
 
-The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of *cognateness* between each pair of languages.
+The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of cognateness between each pair of languages.
 Phonological overlap was measured by computing the Levenshtein distance between each cross-language pair of phonetic transcriptions.
 The Levenshtein distance is a metric that computes the distance between two strings of characters/phonemes by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
-For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/\textipa{"kæt}/-/\textipa{"kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /\textipa{æ}/ by /\textipa{A}/).
-The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{"kæt}/-/\textipa{"Ga.ta}/) would score three, as it requires at least three editions: substituting /\textipa{k}/ for /\textipa{G}/, substituting /\textipa{\ae}/ for /\textipa{a}/ and adding another /\textipa{a}/.
+For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/\textipa{"kæt}-\textipa{"kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /\textipa{æ}-\textipa{A}/).
+The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{"kæt}-\textipa{"Ga.ta}/) would score three, as it requires at least three editions: substituting /\textipa{k}/ for /\textipa{G}/, substituting /\textipa{\ae}/ for /\textipa{a}/ and adding another /\textipa{a}/.
 Levenshtein distance scores are divided by the length of the longest string to normalise the similarity scores between 0 and 1.
 Finally, the result is subtracted from one, to get the complementary, and therefore a measure of phonological similarity, instead of dissimilarity.
-Following the previous examples: $1/4 = 0.25$ for /\textipa{"kæt}/-/\textipa{"kAt}/, and $3/4 = 0.75$ for /\textipa{"kæt}/-/\textipa{"Ga.ta}/.
+Following the previous examples: $1/4 = 0.25$ for /\textipa{"kæt}-\textipa{"kAt}/, and $3/4 = 0.75$ for /\textipa{"kæt}-\textipa{"Ga.ta}/.
 Finally, this variable was included as a predictor of participants' vocabulary sizes.
@@
 
-Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [cat], whose word-forms share three phonemes.
+Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [cat], whose word-forms share three phonemes (0.75 Levenshtein similarity).
 When the child is exposed to /\textipa{"gat}/, they might co-activate /\textipa{"ga.to}/ in parallel, due to their phonological similarity.
@@
 Consider now the case of the non-cognate translation equivalent /\textipa{"gos}-\textipa{"pe.ro}/.
-Because of their low phonological similarity, the amount of co-activate of both words might be low, which might leads to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for  /\textipa{"pe.ro}/.
+Because of their low phonological similarity (0% Levenshtein similarity), the amount of co-activate of both words might be low, which might leads to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for  /\textipa{"pe.ro}/.
 While /\textipa{"gat}-\textipa{"ga.to}/ are predicted to benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ are not.
@@
 Ultimately, /\textipa{"ga.to}/ should benefit more strongly from its cognate status than /\textipa{"gat}/.
-For instance, if the previous child is exposed 80% of the time to Catalan, and 20% of the time to Spanish, words in Spanish should, on average, show a stronger cognateness facilitation effect than words in Catalan.
+For instance, if the previous child is exposed 80% of the time to Catalan, and 20% of the time to Spanish, words in Spanish should, on average, show an even stronger cognateness facilitation effect than words in Catalan.
 
 @fig-hypotheses provides a graphic, theoretical illustration of this account.
-We assume the case of a Catalan-Spanish bilingual child who is exposed 65% of the time to Catalan, and 35% of the time to Spanish.
-We illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/\textipa{gat}/ or /\textipa{gos}/) and its Spanish translation (/\textipa{"pe.ro}/ or /\textipa{"ga.to}/).
+We assume the case of a Catalan-Spanish bilingual child who is exposed 60% of the time to Catalan, and 40% of the time to Spanish.
+We also assume that the child is exposed to either word around 50 times per month, following a $\text{Poisson}(\lambda = 50)$ distribution.
+We illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/\textipa{"gat}/ or /\textipa{"gos}/) and its Spanish translation (/\textipa{"pe.ro}/ or /\textipa{"ga.to}/).
 We assume an arbitrary, theoretical threshold for the learning instances at which the child is considered to have acquired a word (solid grey line).
@@
              y = threshold,
+             size = 2,
              hjust = 0) +
@@
          linetype = "Hypothesis") +
-    scale_x_continuous(breaks = seq(min(eli_df$age),
-                                    max(eli_df$age), 4)) +
     scale_colour_manual(values = clrs[c(1, 4)]) +
@@
 
-aoa_df <- generate_aoa(eli_df, threshold = 250) |>
+aoa_df <- generate_aoa(eli_df, threshold = threshold) |>
     mutate(across(aoa, lst(min, max)), .by = c(te, language)) |>
@@
                  linewidth = 3/4) +
-    geom_hline(yintercept = aoa_df$threshold,
+    geom_hline(yintercept = threshold,
                linewidth = 3/4,
@@
     scale_colour_manual(values = clrs[c(1, 4)]) +
-    scale_x_continuous(breaks = seq(min(eli_df$age),
-                                    max(eli_df$age), 4)) +
     theme(legend.position = "none")
@@
     scale_fill_manual(values = clrs[c(1, 4)]) &
-    scale_y_continuous(labels = function(x) format(x, big.mark = ",")) &
+    scale_y_continuous(labels = function(x) format(x, big.mark = ","),
+                       limits = c(0, 1500)) &
+    scale_x_continuous(labels = function(x) format(x, big.mark = ","),
+                       limits = c(0, 50)) &
     guides(colour = guide_legend(override.aes = list(linewidth = 2))) &
@@
 
-We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a word *only* once its translation has been already acquired.
+We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word.
 We collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
@@
 *Cognateness* was defined as a continuous measure of phonological similarity between translation equivalents.
-We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-XSAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computercoding].
+We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computercoding].
 This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
@@
                         .sep = ".")
+stan_version <- cmdstanr::cmdstan_version()
+
 ```
 
-From the responses to the `r format(nrow(n_items), big.mark = ",")` items included in the questionnaire, we restricted the analyses to responses to nouns (`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded) [@fourtassi2020growth].
+We collected `r format(nrow(n_items), big.mark = ",")` items responses.
+We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, @fourtassi2020growth].
 We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
@@
 
-We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model.
+We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, multilevel ordinal regression model.
 This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest.
@@
 &\textbf{Response}~(k)~\textbf{to word}~i~ \textbf{by participant}~j \\
-\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}), \\ &\text{where}~k \in \{\text{No, Understands, Understands and Says}\}\\
+\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}), \\ 
+&\text{where}~k \in \{\text{No} \rightarrow \text{Understands}, \text{Understands} \rightarrow \text{Understands and Says}\}\\
 &\textbf{Distribution parameters} \\
@@
 Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
-We implemented the model using was done using the `brms` [@burkner2017brms], a R interface to the Stan probabilistic language [@carpenter2017stan].
-We ran `r nchains(model_fit)` MCMC  chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
+We implemented the model using was done using the `brms` [@burkner2017brms], a R interface to the Stan probabilistic language [@carpenter2017stan; `r stan_version`].
+We ran four MCMC  chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
 Model posterior draws and predictions were handled using the `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
@@
     geom_line(aes(group = interaction(lv, .draw)),
-              linewidth = 0.5,
-              alpha = 1/10) +
+              linewidth = 3/4,
+              alpha = 1/20) +
     stat_summary(fun = median, geom = "line", linewidth = 0.75) +

### 2023-06-13 manuscript/manuscript.qmd
Purge dependencies

@@
     indent: true
-
+  docx:
+    knitr:
+      opts_chunk:
+        dev: "png"
+        dev-args:
+          - "cairo-png"
+        dpi: 300
+        out.width: "80%"
 author-note:
@@
 abstract: |
-    Language non-selectivity is a prominent feature of bilingual lexical processing.
-An instance of such non-selectivity is embodied by cognateness.
-Recent studies have pointed to cognateness as a facilitator of vocabulary growth, and to cognates being acquired at earlier ages than non-cognates.
-The specific mechanisms behind such facilitation are unclear.
-Building on accumulator models of language acquisition, we suggest that cognateness increases the number of cumulative learning instances that a bilingual child encounters, thanks to cross-language activation.
-In line with previous studies, this model predicted a cognate facilitation, but only for words to which children are exposed less often.
+    Recent work on bilingual language acquisition points to cognate words being acquired earlier in age than non-cognates.
+Cross-language activation has been proposed as a candidate mechanism behind this facilitation effect, but the
+
+       but the specific dynamics of lexical Building on accumulator models of language acquisition, we present an account of word acquisition in which the co-activation of cognate translation equivalents increases the rate at which of bilingual children encounter learning instances for words in both languages.
 We tested these predictions against vocabulary data from 366 Catalan-Spanish bilinguals toddlers.
@@
 Consider now the case of the non-cognate translation equivalent /\textipa{"gos}-\textipa{"pe.ro}/.
-Because of their low phonological similarity (0% Levenshtein similarity), the amount of co-activate of both words might be low, which might leads to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for  /\textipa{"pe.ro}/.
+Because of their low phonological similarity (0% Levenshtein similarity), the amount of co-activation of both words might be low, which might leads to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for  /\textipa{"pe.ro}/.
 While /\textipa{"gat}-\textipa{"ga.to}/ are predicted to benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ are not.
-This facilitation effect may not be equivalent for /\textipa{"gat}-\textipa{"ga.to}/.
 If the child receives a larger amount of Catalan input than Spanish input, they will encounter the Catalan form /\textipa{"gat}/ more frequently than the Spanish form /\textipa{"ga.to}/.
@@
 Bilinguals differ across multiple dimensions.
-these differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to higher-level factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
+These differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to higher-level factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
 This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.

### 2023-06-15 manuscript/manuscript.qmd
Dockerise repository

@@
     Recent work on bilingual language acquisition points to cognate words being acquired earlier in age than non-cognates.
-Cross-language activation has been proposed as a candidate mechanism behind this facilitation effect, but the
-
-       but the specific dynamics of lexical Building on accumulator models of language acquisition, we present an account of word acquisition in which the co-activation of cognate translation equivalents increases the rate at which of bilingual children encounter learning instances for words in both languages.
+Cross-language activation has been proposed as a candidate mechanism behind this facilitation effect, but the specific dynamics of lexical Building on accumulator models of language acquisition, we present an account of word acquisition in which the co-activation of cognate translation equivalents increases the rate at which of bilingual children encounter learning instances for words in both languages.
 We tested these predictions against vocabulary data from 366 Catalan-Spanish bilinguals toddlers.
@@
 The foundations of word learning are in place early in age.
-From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [e.g., @fenson2007macarthurbates; @samuelson2021precision].
-During the last half of the second year of life, new word-forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension].
-Such early stages of lexical acquisition are characterised by substantial heterogeneity across different populations of language learners, reflected for instance on (the variability of) quantitative indices of lexical development, like vocabulary size [i.e., vocabulary size, @fenson1994variability; @frank2021variability].
-Despite this variability, trajectories of early vocabulary growth seem quite stable across languages.
-@tardif2008baby collected data about the first ten words acquired by 10-- to 16--month-old infants living in the United States, Hong Kong, and Beijing.
-Since birth, these infants had been learning English, Cantonese, and Mandarin, respectively.
-The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caregivers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animal sounds (*woof-woof*).
-These results were later extended by @frank2021variability to a more diverse set of languages, and found that such cross-language commonalities are stronger at earlier stages of lexical acquisition.
+From six months of age, infants start directing their gaze to some objects when hearing their labels, according to experimental data [@tincoff1999beginnings; @bergelson2012months; @bergelson2015early].
+Shortly after, caregivers start reporting some words as acquired by their infant in vocabulary checklists [e.g., @fenson2007macarthurbates; @samuelson2021precision].
+Most research on early word acquisition relies extensively on data from monolingual children and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
+Previous work on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and to both groups knowing a similar amount of words, if not more words, when the bilinguals’ two languages are pooled together.
+@hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did.
+Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together.
+Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
+A more detailed analysis of the words in bilinguals’ lexicon show some interesting patterns.
 
-Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
-Previous research on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account.
-@hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did.
-Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-These studies rely on bilingual samples whose languages are relatively distant in typology: English is a Germanic language, while Spanish is a Romance language, and both languages differ substantially at the grammatical and the phonological level.
-Little is known about vocabulary acquisition in bilinguals learning a pair of typologically closer languages (e.g., when both are Romance languages).
+One important observation of studies on bilinguals’ early vocabulary acquisition is that cognate words are easier to acquire than non-cognate words.
+Cognate words are translation equivalents that are phonologically similar (or share some type of form-similarity).
+For instance, the Spanish translation equivalent of cat is gato, a cognate word; the translation equivalent of dog is perro, a non-cognate word.
+The differences in the percentages of cognate and non-cognate words between two languages is related to historic reasons: languages typologically close (like Dutch and English or Italian and Spanish) share more cognates than languages typologically distant (like English and Chinese, or Urdu and Spanish).
+The conclusion that cognate words are easier to learn is based on two types of evidence: studies investigating vocabulary sizes in children learning language pairs with different percentages of cognates (that is, differing in their typological distance) and studies comparing the number of cognate and non-cognate words children know in a specific language pair.
 
-One way typological distance may impact bilingual trajectories of language acquisition is the presence (or absence) of *cognates* in their linguistic input.
-Quite often, bilingual children encounter more than one label for the same referent: one in each language.
-For instance, an English-Spanish bilingual might hear both *dog* and *perro* in the presence of a dog We will refer to such cross-language synonyms as *translation equivalents*.^[Throughout the present study, we make the assumption that bilinguals form a common semantic representations for each translation equivalent (e.g., the concept 'dog' is shared for the English and Spanish words *dog* and *perro*).
-The referential contexts in which bilingual children hear (and learn) both labels may not be identical.
-For instance, the English word *finger* refers to eight body parts (excluding thumbs and toes), whereas its Spanish translation *dedos* refers to twenty body parts (fingers, thumbs, and toes).
-These discrepancies are reflected, for instance, in adults' performance in lexical decision tasks [@chaouch-orozco2023are].
-The fact that some translation pairs share more semantic features than others does not, however, inform whether bilinguals form common or separate semantic representations for translation equivalents.
-What English speakers understand for 'fingers' might not be identical to what Spanish speakers understand for 'dedos'.
-But when an English speaker learns Spanish, it seems implausible that they would form a separate semantic representation for what 'dedo' conveys at the conceptual level, compared to the possibility that they associate the new label to the semantic representation they already had formed for 'fingers'.
-Therefore, the lack of complete semantic alignment is not necessarily an argument against the existence of common semantic representations for translation equivalents.
-For simplicity, and in line with previous (formal and verbal) models of bilingual lexical processing [e.g., @shook2013bilingual; @dijkstra2019multilink; @dijkstra1998bia; @kroll2010revised], we assume common semantic representations for translation pairs.].
-Sometimes, the word-forms involved in a translation equivalent may be phonologically similar.
-Such form-similar translation equivalents are referred to as *cognates*.
-While the previously mentioned translation pair *dog* and *perro* are non-cognates (both word-forms hardly share any phonemes), the translation pair *cat* and *gato* can be considered cognates.
-The amount of cognates shared by a pair of languages is dependent on their typological distance.
-When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like English and Spanish (Romance and Germanic languages, respectively).
-Therefore, bilingual children learning two typologically close languages may be exposed to a larger amount of cognates than those learning two typologically more distant languages.
-For instance, in the presence of a door, a Catalan-Spanish bilingual might be exposed to the words *porta* and *puerta* (cognate translation equivalent with the common meaning 'door'), whereas an English-Spanish bilingual might hear *door* and *puerta* (non-cognate translation equivalent).
-
-It could be the case that mapping two phonologically similar labels (cognates like *porta*-*puerta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *door* and *puerta*).
-If cognates are easier to acquire than non-cognates, bilinguals learning two languages that share a high proportion of cognates should benefit more often from this facilitation effect than bilinguals learning two languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
-@floccia2018introduction provided evidence for such a facilitation effect of typological distance and vocabulary size.
+@floccia2018introduction published an impressive study comparing vocabularies of children learning several language pairs differing in their percentage of cognates.
 The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
 The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
+The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of cognateness between each pair of languages.
+Floccia and co-workers reported an increase in vocabulary size in the additional language (i.e., not English) associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
+For example, English-Dutch bilinguals (languages with a high phonological overlap), were able to produce more Dutch words than English-Mandarin bilinguals (languages with a low phonological overlap) were able to produce in Mandarin.
+@blom2020crosslanguage, @bosma2019longitudinal, and @gampe2021does reported similar results, providing converging evidence of a facilitatory effect of language distance on vocabulary size.
 
-The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of cognateness between each pair of languages.
-Phonological overlap was measured by computing the Levenshtein distance between each cross-language pair of phonetic transcriptions.
-The Levenshtein distance is a metric that computes the distance between two strings of characters/phonemes by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary].
-For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/\textipa{"kæt}-\textipa{"kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /\textipa{æ}-\textipa{A}/).
-The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{"kæt}-\textipa{"Ga.ta}/) would score three, as it requires at least three editions: substituting /\textipa{k}/ for /\textipa{G}/, substituting /\textipa{\ae}/ for /\textipa{a}/ and adding another /\textipa{a}/.
-Levenshtein distance scores are divided by the length of the longest string to normalise the similarity scores between 0 and 1.
-Finally, the result is subtracted from one, to get the complementary, and therefore a measure of phonological similarity, instead of dissimilarity.
-Following the previous examples: $1/4 = 0.25$ for /\textipa{"kæt}-\textipa{"kAt}/, and $3/4 = 0.75$ for /\textipa{"kæt}-\textipa{"Ga.ta}/.
-Finally, this variable was included as a predictor of participants' vocabulary sizes.
-Among other findings, Floccia and co-workers reported an increase in vocabulary size in the additional language (i.e., not English) associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
-For example, English-Dutch bilinguals (0.2214 mean phonological overlap), were able to produce more Dutch words than English-Mandarin bilinguals (0.0197 mean phonological overlap) were able to produce in Mandarin.
-More recently, @blom2020crosslanguage and @bosma2019longitudinal reported similar results, providing converging evidence of for facilitatory effect of language distance on vocabulary size estimates.
+A second set of studies concluded that cognates are overrepresented in bilinguals’ early lexicon.
+@bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months and found that cognates represented a larger proportion of participant’s vocabulary than non-cognates.
+@schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual produced cognates earlier than non-cognates, on average.
+@mitchell2022cognates addressed this issue in a larger sample longitudinal study.
+These authors collected expressive vocabulary data of 47 16- to 30-month-old French-English bilinguals living in Canada, in both languages.
+They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
+The proportion of words that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age-of-acquisition norms (an index of word difficulty).
+Taken together, the results of these two lines of research provide a strong support to the hypothesis that phonological similarity (as reflected in cognateness) plays a facilitation role in bilingual word acquisition.
 
-Floccia et al.
-pointed to *parallel activation* as the underpinning mechanism behind their results.
-The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during speech production or comprehension, even in fully monolingual environments.
-Evidence supporting the language-non selective account of lexical access is not restricted to production or the auditory modality, but rather extends to comprehension and translation, and to the visual modality (reading and signing) [@spivey1999cross; @hoshino2008cognate; @schwartz2007reading; @schwartz2006bilingual; @morford2011when; @vanhell2008sentence; @gimeno-martinez2021crosslinguistic; see @kroll2013multilingual for review].
+Parallel activation of bilinguals’ lexicons has been proposed as the underlying mechanism for such facilitatory effect [e.g., @floccia2018introduction; @mitchell2022cognates].
+The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during language processing, even in fully monolingual environments.
+Evidence supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; see @kroll2017bilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
 In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish.
-In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of *table*), importantly such distinction was only relevant for bilinguals.
+In half of the trials, the object labels were cognates in Spanish and Catalan (árbol-arbre, translations of tree), whereas in the other half of the trial labels were non-cognates (mesa-taula, translations of table), obviously, such distinction was only relevant for bilinguals.
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
 In contrast, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, showed equivalent naming times for the two types of stimuli.
-These results showed that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
+The authors interpreted the difference between cognates and non-cognates in bilinguals as reflecting the additional phonological activation that cognate words would receive from their translation equivalents (due to language non-selective activation of bilinguals’ lexicons).
+These results showed that bilinguals’ Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
+Evidence of parallel activation has been reported in bilingual toddlers and children [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic; @poulin-dubois2013lexical].
 
-Bilingual toddlers and children---still in the process to developing a mature lexicon---also show evidence of parallel activation during comprehension [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic; @poulin-dubois2013lexical].
-Whether---and if so, how---the language non-selectivity shapes the early bilingual lexicon is still unclear.
-Previous work has pointed to an interactive account of lexical acquisition, in which toddlers' lexical representations in one language influence the acquisition of words in the other language.
-This cross-language facilitation is, in principle, possible since early ages: the bilingual infant lexicon consists in a substantial proportion of  translation equivalents: the presence of a word in the lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage; @pearson1995crosslanguage].
-In fact, translation equivalents may even be overrepresented in the early bilingual lexicon.
-@bilson2015semantic found that bilingual toddlers produced a higher proportion of translation equivalents than it would be expected if they learned two independent lexica simultaneously, which suggests that bilingual children preferentially acquired words for which they already knew a label in the other language.
-In the same line, @bosch2014first reported a similar overrepresentation of translation equivalents in Catalan-Spanish bilinguals: a high proportion of acquired Catalan words held an also acquired translation in Spanish.
-Finally, a recent investigation by @tsui2022are presented a quantitative model of bilingual vocabulary acquisition which showed a consistent preferential acquisition of translation equivalents by English-French bilinguals, with such preference being stronger at early stages of vocabulary growth.
-Overall, these findings suggest that bilingual children benefit from a semantic facilitation effect driven by the acquisition of translation equivalents.
+Although there is converging evidence on the role of parallel activation, the previous studies do not address its influence on the learning trajectories of words.
+Results are aggregated across words, and provide no information about the specific dynamics of how parallel activation influences word learning.
+This is the goal of the present research.
 
-If translation equivalents are pervasive in the bilingual lexicon since early ages, it is possible that cognateness increases the amount of cross-language activation, boosting early word acquisition.
-This could ultimately lead to children who are learning languages with a larger proportion of cognates (e.g., English and Dutch) showing larger vocabulary sizes, as found by Floccia et al.
-There is evidence of phonological and semantic similarity facilitating word acquisition in monolinguals: infants are more likely to acquired new words that are semantically or phonologically more connected to other words, than words with poorer connectivity in either of those dimensions, such as words belonging to sparse phonological neighbourhoods [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children; @storkel2004children].
-If phonological similarity plays a cross-language facilitation role during bilingual word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity (e.g., are taxonomically related) but not phonological similarity.
+We propose an account in which a learning instance for a word may also represent a learning instance for its translation equivalent, only if such translation equivalent is co-activated.
+The probability of this co-activation is proportional to the phonological similarity between the two translation equivalents, and given that cognates share higher phonological similarity than non-cognates, the former may be co-activated more frequently than the latter.
+This should lead to a faster accumulation of learning instances for cognates, compared to non-cognates.
+Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across both languages: words from the lower-exposure language would receive additional learning instances through parallel activation more often than words from the higher-exposure language.
+Therefore, the acquisition of words from the lower-exposure language would benefit more strongly from their cognate status than words from the higher-exposure language.
 
-Evidence supporting an earlier age-of-acquisition for cognates is scarce.
-@bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
-@schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual produced cognates earlier than non-cognates, on average.
-The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-Floccia et al.'s estimates are statistically more reliable given their remarkably larger sample size (372 participants).
-Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age-of-acquisition at the *word-level*.
-The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
-By aggregating the responses from all words into a single datum per child, it was no longer possible to test the effect of cognateness on word acquisition at the word-form level.
-Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as the response variable, age-sensitive claims about the effect of cognateness on word acquisition should be taken with caution [a limitation also present in the @bosch2014first study].
+Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [cat], that  are phonologically very similar.
+When the child is exposed to /\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in parallel.
+Therefore, this exposure will count as a learning instance for both co-activated forms.
+However, for the case of the non-cognate translation equivalent /\textipa{"gos}/, the low phonological similarity will result in a weak activation of /\textipa{"pe.ro}/ resulting in such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for /\textipa{"pe.ro}/.
+While /\textipa{"gat}-\textipa{"ga.to}/ will benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not.
+If the child receives linguistic input from one of the languages more often than from the other, this effect might affect each form of the cognate translation equivalent differently.
+For instance, if the child receives a larger amount of Catalan input than Spanish input, they will encounter the Catalan form /\textipa{"gat}/ more frequently than the Spanish form /\textipa{"ga.to}/.
+Through parallel activation, /\textipa{"gat}/ will activate /\textipa{"ga.to}/ more often than vice versa.
+Ultimately, /\textipa{"ga.to}/ will benefit more strongly from its cognate status than /\textipa{"gat}/, as it receives additional learning instances from its translation equivalent more often than /\textipa{"gat}/.
 
-More recently, @mitchell2022cognates addressed this issue providing a more fine grained measure of a  cognateness facilitation effect in vocabulary acquisition.
-Using a longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages.
-They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
-The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age-of-acquisition norms (an index of word difficulty).
-These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size seems to grow faster when both languages are more similar at the lexical level.
-Word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster.
-Parallel activation remains a plausible explanation for this effect.
-When a child is exposed to a word-form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition.
-The precise underpinnings of this effect are still unclear.
-
-In the present investigation we approach the cognate facilitation on word learning by extending the conceptual framework of accumulator models of language acquisition to the bilingual case.
-Accumulator models formalise the widespread assumption that a child's vocabulary growth rate is a function of mainly---but not exclusively---the number of learning instances that a child has accrued with a particular word [see @kachergis2022standard for review].
-A learning instance of a word consists, for example, in hearing such word-form (*flower*) in the presence of its referent (a flower).
-We devise word acquisition as a continuous process in which a child accumulates experience with a given word-form and its referential context until the number of cumulative learning instances meets some theoretical threshold.
-After this, we consider the lexical representation of the word as *consolidated*, which leads to a non-linear increase in the probability of caregivers reporting such word as acquired in a vocabulary checklist like the one in the present study.
-As infants receive linguistic input, they also accumulate opportunities to learn words until a threshold is crossed, an the word can be considered to be acquired.
-The rate at which monolingual children encounter learning instances with some word-form is strongly determined by age: as time passes children are exposed to a larger amount of linguistic input, and therefore accumulate more learning experiences with all words.
-
-Not all words occur in the linguistic input with identical frequency, so children accumulate experience with some word-forms at faster rates than with other word-forms.
-This results in frequent words being acquired earlier in age than less frequent words [e.g., @ambridge2015ubiquity; @goodman2008does; @braginsky2019consistency; @swingley2018quantitative; @hansen2017what; @verhagen2022determinants].
-For this reason, lexical frequency plays a central role in most models of lexical growth, especially in the case of accumulator models [@hidaka2013computational; @mollica2017how].
-Following this account, words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are predicted to be acquired earlier in age than words with low lexical frequency such as 'taxi' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words).
-
-We argue that previous accounts on vocabulary growth based on accumulator models are a convenient framework to study bilingual vocabulary growth.
-Bilinguals share fundamentally similar learning mechanisms with their monolinguals peers.
-Developmental divergences between both groups can be understood as the result of the the negotiation between such mechanisms and the particular demands of a dual language input [e.g., @werker2005primir; @curtin2011bilingual].
-One of such divergences is the fact that bilinguals' frequency of exposure to a word-form is mediated by their degree of exposure to their languages.
-Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual (who listens 100% of the time to English) would need.
-
-In the present study, we incorporate this feature of the bilingual linguistic input to an account of the cognateness effect on word acquisition.
-We suggest that when a bilingual child is exposed to a given word, they not only activate its corresponding lexical representation, but also that of its translation equivalent.
-This co-activation is proportional to the form similarity between both word-forms.
-When this parallel activation is in place, each exposure to one of the word-forms may count as a learning instance for both members of the translation equivalent, in a degree determined by their cognateness.
-Cognate translation equivalents, for which stronger co-activation is expected, accumulate learning instances faster than non-cognates, for which weaker or no co-activation is expected.
-Finally,  when the child has accumulator some threshold number of learning instances with any of these word-forms, its lexical representation reaches some degree of consolidation that leads to a non-linear change in the child's observed learning outputs, like the probability of such word being reported as "acquired" by caregivers, or by stronger naming effects in experimental tasks using the Visual World Paradigm [e.g., @swingley2002recognition].
-Because , lexical representations of cognate words accumulate learning instances faster than those of non-cognate words, the former would consolidate earlier in age.
-
-Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [cat], whose word-forms share three phonemes (0.75 Levenshtein similarity).
-When the child is exposed to /\textipa{"gat}/, they might co-activate /\textipa{"ga.to}/ in parallel, due to their phonological similarity.
-If this exposure to the word form occurs in a convenient referential context (e.g., in the presence of a cat), this might count as a learning instance for both active word-forms.
-Consider now the case of the non-cognate translation equivalent /\textipa{"gos}-\textipa{"pe.ro}/.
-Because of their low phonological similarity (0% Levenshtein similarity), the amount of co-activation of both words might be low, which might leads to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for  /\textipa{"pe.ro}/.
-While /\textipa{"gat}-\textipa{"ga.to}/ are predicted to benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ are not.
-If the child receives a larger amount of Catalan input than Spanish input, they will encounter the Catalan form /\textipa{"gat}/ more frequently than the Spanish form /\textipa{"ga.to}/.
-If parallel activation is in place, the former will activate the latter more often than vice versa.
-Ultimately, /\textipa{"ga.to}/ should benefit more strongly from its cognate status than /\textipa{"gat}/.
-For instance, if the previous child is exposed 80% of the time to Catalan, and 20% of the time to Spanish, words in Spanish should, on average, show an even stronger cognateness facilitation effect than words in Catalan.
-
-@fig-hypotheses provides a graphic, theoretical illustration of this account.
-We assume the case of a Catalan-Spanish bilingual child who is exposed 60% of the time to Catalan, and 40% of the time to Spanish.
-We also assume that the child is exposed to either word around 50 times per month, following a $\text{Poisson}(\lambda = 50)$ distribution.
-We illustrate the cumulative sum of learning instances that a child accumulates across ages for the Catalan word-form (/\textipa{"gat}/ or /\textipa{"gos}/) and its Spanish translation (/\textipa{"pe.ro}/ or /\textipa{"ga.to}/).
-We assume an arbitrary, theoretical threshold for the learning instances at which the child is considered to have acquired a word (solid grey line).
-Points and triangles show the age-of-acquisition at which the amount of learning instances reaches the threshold.
-We show the predicted age-of-acquisition under two hypothesis: (1) a cognateness facilitation effect through parallel activation, in which both word-forms co-activate each other accumulating learning instances faster, and (2) no cognateness facilitation, for which both word-forms accumulate learning instances independently.
-
-```{r fig-hypotheses}
-#| label: fig-hypotheses
-#| fig-cap: "Schematic representation of the hypothesised cognate facilitation effect on word acquisition."
-#| fig-width: 8
-#| fig-height: 7
-#| out-width: 100%
-#| out-height: 100%
-#| echo: false
-#| fig-align: center
-threshold <- 300
-
-# no parallel activation
-eli_df <- generate_eli(
-    threshold = threshold,
-    age = seq(1, 50, length.out = 100),
-    freq_month = 50,
-    freq_beta = 1,
-    l1_doe = 0.60,
-    parallel_beta = 0,
-    conditions = c(
-        glue::glue("**Cognate**: /<i style='color:{clrs[1]}'>ˈgat</i>-<i style='color:{clrs[4]}'>ˈga.to</i>/ [cat]"),
-        glue::glue("**Non-cognate**: /<i style='color:{clrs[1]}'>ˈgos</i>-<i style='color:{clrs[4]}'>ˈpe.ro</i>/ [dog]")
-    )
-) |>
-    filter(hypothesis=="H0") |> 
-    mutate(language = ifelse(language=="L1", "Catalan form", "Spanish form"))
-
-aoa_df <- generate_aoa(eli_df, threshold = threshold) |>
-    mutate(across(aoa, lst(min, max)), .by = c(te, language)) |>
-    filter(hypothesis=="H0")
-# 
-# img <- c(cat = "images/diagram_cat.png",
-#          dog = "images/diagram_dog.png") |>
-#     map(magick::image_read) |>
-#     map(\(x) magick::image_ggplot(x, interpolate = FALSE))
-
-plot_h0 <- eli_df |>
-    ggplot(aes(age, eli,
-               colour = language)) +
-    facet_grid(~ te) +
-    geom_segment(data = aoa_df,
-                 aes(x = aoa,
-                     xend = aoa,
-                     y = 0,
-                     yend = threshold),
-                 linewidth = 3/4) +
-    geom_hline(yintercept = aoa_df$threshold,
-               linewidth = 3/4,
-               colour = "grey80") +
-    annotate(geom = "label",
-             label = glue::glue("Threshold: {threshold}"),
-             label.r = unit(0, "lines"),
-             label.size = 0,
-             fill = "grey80",
-             x = 0,
-             y = threshold,
-             size = 2,
-             hjust = 0) +
-    geom_line(linewidth = 1) +
-    geom_label(data = aoa_df,
-               aes(x = aoa,
-                   y = threshold,
-                   label = round(aoa, 2),
-                   fill = language),
-               size = 3,
-               label.r = unit(0, "lines"),
-               label.size = 0,
-               colour = "white",
-               na.rm = TRUE,
-               show.legend = FALSE
-    ) +
-    labs(x = "Age (months)",
-         y = "Cumulative\nlearning instances",
-         colour = "",
-         shape = "Hypothesis",
-         fill = "",
-         title = "H0: no parallel activation",
-         linetype = "Hypothesis") +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(legend.position = "bottom",
-          legend.key.width = unit(1.5, "cm"),
-          legend.margin = margin(c(0, 0, 0, 0)),
-          legend.justification = c(0, 1),
-          legend.title = element_text(size = 9),
-          legend.text= element_text(size = 8))
-
-# with parallel activation
-eli_df <- generate_eli(
-    threshold = threshold,
-    age = seq(0, 50, by = 1),
-    l1_doe = 0.60,
-    freq_month = 50,
-    freq_beta = 1,
-    conditions = c(
-        glue::glue("**Cognate**: /<i style='color:{clrs[1]}'>ˈgat</i>-<i style='color:{clrs[4]}'>ˈga.to</i>/ [cat]"),
-        glue::glue("**Non-cognate**: /<i style='color:{clrs[1]}'>ˈgos</i>-<i style='color:{clrs[4]}'>ˈpe.ro</i>/ [dog]")
-    )
-) |>
-    filter(hypothesis=="H1") |> 
-    mutate(language = ifelse(language=="L1", "Catalan form", "Spanish form"))
-
-aoa_df <- generate_aoa(eli_df, threshold = threshold) |>
-    mutate(across(aoa, lst(min, max)), .by = c(te, language)) |>
-    filter(hypothesis=="H1")
-
-plot_h1 <- eli_df |>
-    ggplot(aes(age, eli,
-               colour = language)) +
-    facet_grid(~ te) +
-    geom_segment(data = aoa_df,
-                 aes(x = aoa,
-                     xend = aoa,
-                     y = 0,
-                     yend = threshold),
-                 linewidth = 3/4) +
-    geom_hline(yintercept = threshold,
-               linewidth = 3/4,
-               colour = "grey80") +
-    annotate(geom = "label",
-             label = glue::glue("Threshold: {threshold}"),
-             label.r = unit(0, "lines"),
-             label.size = 0,
-             fill = "grey80",
-             x = 0,
-             y = threshold,
-             size = 2,
-             hjust = 0) +
-    geom_line(linewidth = 1) +
-    geom_label(data = aoa_df,
-               aes(x = aoa,
-                   y = threshold,
-                   label = round(aoa, 2),
-                   fill = language),
-               size = 3,
-               label.r = unit(0, "lines"),
-               label.size = 0,
-               colour = "white",
-               na.rm = TRUE,
-               show.legend = FALSE
-    ) +
-    labs(x = "Age (months)",
-         y = "Cumulative\nlearning instances",
-         colour = "",
-         shape = "Hypothesis",
-         fill = "",
-         title = "H1: parallel activation",
-         linetype = "Hypothesis") +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(legend.position = "none")
-
-plot_h0 + plot_h1 +
-    plot_layout(ncol = 1, guides = "collect") &
-    plot_annotation(tag_levels = "A") +
-    scale_colour_manual(values = clrs[c(1, 4)]) &
-    scale_fill_manual(values = clrs[c(1, 4)]) &
-    scale_y_continuous(labels = function(x) format(x, big.mark = ","),
-                       limits = c(0, 1500)) &
-    scale_x_continuous(labels = function(x) format(x, big.mark = ","),
-                       limits = c(0, 50)) &
-    guides(colour = guide_legend(override.aes = list(linewidth = 2))) &
-    theme_ggdist() &
-    theme(legend.position = "top", 
-          legend.text = element_text(size = 10),
-          legend.key.width = unit(2, "cm"),
-          legend.key.height = unit(0.5, "cm"),
-          strip.background = element_rect(fill = "grey90", colour = "grey90"),
-          strip.text = ggtext::element_markdown(size = 10),
-          panel.grid.major.y = element_line(colour = "grey",
-                                            linetype = "dotted"),
-          plot.background = element_rect(fill = "white",
-                                         colour = NA))
-
-```
-
-This hypothesis provides a convenient account for the results in Floccia et al.
-The size of the effect of linguistic similarity on vocabulary size that Floccia et al.
-reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure).
-Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
-Therefore, participants may have on average learned the English word-form of translation equivalents earlier than the word-form in the additional language.
-Because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants might then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.
-
-We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word.
-We collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
-Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees.
-We adopted a Bayesian explanatory item response theory  [IRT, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word-form, while adjusting for the the number of phonemes in the word-form (an indicator of word acquisition difficulty).
-
-To operationalise participants' exposure rate to a word-form, we used a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it), and child's amount of exposure to the language such word belongs to.
-This measure was calculated by multiplying the word-form's Zipf-transformed lexical frequency to participant's degree of exposure to the language the word belongs to.
-Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childesdb] (see Appendix A for more details).
-Participants' degree of exposure were reported by caregivers in a language exposure questionnaire they filled before the vocabulary checklist.
-Caregivers provided a proportion of the time the child was estimated to listen to any language they were spoken to regularly (described in the methods section).
-The resulting measure is a exposure-weighted lexical frequency measure, which was included in analyses as a proxy of how often bilingual participants were exposed to each word-form.
-
-*Cognateness* was defined as a continuous measure of phonological similarity between translation equivalents.
-We quantified the phonological similarity between each pair of Catalan-Spanish word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonetic transcriptions [@levenshtein1966binary; @wells1995computercoding].
-This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one.
-This results in a proportion that indicates how much the two phonetic transcriptions of the translation equivalent are similar to each other, ranging from zero (no similarity at all) to one (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-For instance, this measure of phonological similarity returns 0% for *ocell* /`uceL`/ and *pájaro* /`paxa4o`/, Catalan and Spanish for 'bird', and returns 50% for *lluna* /`Lun5`/ and *luna* /`luna`/, Catalan and Spanish for 'moon' (see Appendix A for more details).
-
-Finally, we included word length (numbwe of phonemes) as a nuance predictor in our analyses, motivated by previous studies reporting a strong association between word length and vocabulary size estimates [e.g., @braginsky2019consistency; @jones2019children].
-Shorter word-forms like /\textipa{sO\textltilde}/ (Catalan for *sun*) are generally simpler to encode as phonological representations in the lexicon than longer word-forms like /\textipa{ku.ku"DRi\textltilde}/ (Catalan for *crocodile*), and are therefore more likely to be participant a child's vocabulary.
+To test these predictions, we collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
+We adopted a Bayesian explanatory item response theory approach (IRT, see @kachergis2022standard, for a similar approach) to model the probability of acquisition of `r nrow(distinct(items, language, item))` Catalan and Spanish nouns included in the vocabulary checklist.
+Words were considered as acquired if caregivers reported such word to be understood (comprehension) or understood and said (production) by their child.
+We estimated the impact of several predictors of interest on the probability of acquisition.
+We included participants’ age and rate of exposure to the word-form, and the cognate status of the word-form.
+As described in the methods section, rate of exposure was a composite measure taking into account participant’ language exposure and word’s lexical frequency).
+We predicted an interaction between cognate status and word exposure rate  in which the probability of comprehension is higher for low-exposure cognate words, but not for high-exposure words cognate words.
 
@@
 For each word included in the questionnaire, we manually generated a broad phonological transcription in X-SAMPA format.
-Catalan word-forms were transcribed to Central Catalan phonology, and Spanish translations were transcribed to Castilian Spanish phonology.
+Catalan word-forms were transcribed to Central Catalan phonology, and Spanish word-forms were transcribed to Castilian Spanish phonology.
 For each translation equivalent, we calculated the phonological similarity between the Catalan and the Spanish word-form using the same Levenshtein similarity measure used in @floccia2018introduction.
@@
 
-Following the rationale of accumulator models, we suggest that parallel activation increases the rate at which bilingual children encounter learning instances for cognate words, but not for non-cognate words.
+We suggest that parallel activation increases the rate at which bilingual children encounter learning instances for cognate words, but not for non-cognate words.
 We hypothesised that when a bilingual child is exposed to a word-form in presence of its referent, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
@@
 
-Consider a Catalan-Spanish bilingual child who is exposed to the Catalan word-form /\textipa{"gat}/ in the presence of a cat.
-This may represent a learning instance, in which the child has the opportunity to strengthen the association between the word-form and its referent.
-Because the Spanish translation of this word (/\textipa{"ga.to}/) is phonologically very similar to the Catalan form (both word-forms share `r format(stringdist::stringsim("gat", "gato"), nsmall = 2)` Levenshtein similarity), it might receive some degree of co-activation.
-Both members of the translation pair would be activated to some extent in the presence of their common referent, and therefore the child would accumulate a learning instance with both word-forms.
-The case of the non-cognate translation equivalent for *dog* (/\textipa{"gOs}/-/\textipa{"pe.ro}/, sharing `r format(stringdist::stringsim("gOs", "pero"), nsmall = 2)` Levenshtein similarity), would be different.
-Since both words hardly share any phonological similarity, a bilingual child would only accumulate learning instances for the word they encounter in speech, but not for the other.
-For instance, hearing /\textipa{"gOs}/ would be unlikely to activate /\textipa{"pe.ro}/ through phonology.
-Each word would accumulate learning instances independently, without any cross-language facilitation taking place.
+This account is in line with accumulator models of language acquisition.
 
@@
 It might be argued that our results reflect the fact that cognate translation equivalents are represented in the initial bilingual lexicon as the *same* lexical entry.
-Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the presence of a cat), it is possible that they classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
+Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}}/ and /\textipa{"ga.to}}/ in the presence of a cat), it is possible that they classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
 This would lead to a faster increase in cumulative learning instances, and to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).

### 2023-06-16 manuscript/manuscript.qmd
Fix symlinks

@@
 editor: source
+lang: en
 
@@
 We also thank all families and infants who participated in the experiments.
-
-abstract: |
-    Recent work on bilingual language acquisition points to cognate words being acquired earlier in age than non-cognates.
-Cross-language activation has been proposed as a candidate mechanism behind this facilitation effect, but the specific dynamics of lexical Building on accumulator models of language acquisition, we present an account of word acquisition in which the co-activation of cognate translation equivalents increases the rate at which of bilingual children encounter learning instances for words in both languages.
-We tested these predictions against vocabulary data from 366 Catalan-Spanish bilinguals toddlers.
-We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the participant.
-We found a cognateness facilitation effect only in low-frequency words or words from the language of lower exposure.
-Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
 
@@
 The foundations of word learning are in place early in age.
-From six months of age, infants start directing their gaze to some objects when hearing their labels, according to experimental data [@tincoff1999beginnings; @bergelson2012months; @bergelson2015early].
-Shortly after, caregivers start reporting some words as acquired by their infant in vocabulary checklists [e.g., @fenson2007macarthurbates; @samuelson2021precision].
-Most research on early word acquisition relies extensively on data from monolingual children and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
-Previous work on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and to both groups knowing a similar amount of words, if not more words, when the bilinguals’ two languages are pooled together.
+At six months of age, infants start directing their gaze to some objects when hearing their labels [@tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and, shortly after, caregivers start reporting some words as acquired by their infant in vocabulary checklists [e.g., @fenson2007macarthurbates; @samuelson2021precision].
+Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
+Previous work on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and to both groups knowing a similar amount of words---if not more words---when the bilinguals’ two languages are pooled together.
 @hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did.
@@
 Cognate words are translation equivalents that are phonologically similar (or share some type of form-similarity).
-For instance, the Spanish translation equivalent of cat is gato, a cognate word; the translation equivalent of dog is perro, a non-cognate word.
+For instance, the Spanish translation equivalent of cat is *gato*, a cognate word; the translation equivalent of dog is *perro*, a non-cognate word.
 The differences in the percentages of cognate and non-cognate words between two languages is related to historic reasons: languages typologically close (like Dutch and English or Italian and Spanish) share more cognates than languages typologically distant (like English and Chinese, or Urdu and Spanish).
@@
 In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish.
-In half of the trials, the object labels were cognates in Spanish and Catalan (árbol-arbre, translations of tree), whereas in the other half of the trial labels were non-cognates (mesa-taula, translations of table), obviously, such distinction was only relevant for bilinguals.
+In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of tree), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of table), obviously, such distinction was only relevant for bilinguals.
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
@@
 To test these predictions, we collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
-We adopted a Bayesian explanatory item response theory approach (IRT, see @kachergis2022standard, for a similar approach) to model the probability of acquisition of `r nrow(distinct(items, language, item))` Catalan and Spanish nouns included in the vocabulary checklist.
+We adopted a Bayesian explanatory item response theory approach [IRT, see @kachergis2022standard, for a similar approach] to model the probability of acquisition of `r nrow(distinct(items, language, item))` Catalan and Spanish nouns included in the vocabulary checklist.
 Words were considered as acquired if caregivers reported such word to be understood (comprehension) or understood and said (production) by their child.
-We estimated the impact of several predictors of interest on the probability of acquisition.
-We included participants’ age and rate of exposure to the word-form, and the cognate status of the word-form.
-As described in the methods section, rate of exposure was a composite measure taking into account participant’ language exposure and word’s lexical frequency).
+We estimated the impact of several predictors of interest on the probability of acquisition, including participants’ age and rate of exposure to the word-form, and the cognate status of the word-form.
+As described in the methods section, rate of exposure was a composite measure taking into account participant’ language exposure and word’s lexical frequency.
 We predicted an interaction between cognate status and word exposure rate  in which the probability of comprehension is higher for low-exposure cognate words, but not for high-exposure words cognate words.
@@
 Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
-both lists included items from a diverse sample of `r n_categories` semantic/functional categories.
+Both lists included items from a diverse sample of `r n_categories` semantic/functional categories.
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
-Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for 'dog') (see @tbl-items for a summary of the questionnaire items).
+Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for dog) (see @tbl-items for a summary of the questionnaire items).
 
@@
 Items that were part of the trial lists of some ongoing experiments in the lab were also included in all versions.
-The resulting reduced vocabulary checklists contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words.
+The resulting reduced list contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words.
 Participants were randomly allocated into one of the four subsets.
-Each response to the questionnaire provided data for a minimum of `r min(n_items_version$catalan, n_items_version$spanish)` and a maximum of `r max(n_items_version$catalan, n_items_version$spanish)` translation equivalents.
 
-For each word included in the questionnaire, we manually generated a broad phonological transcription in X-SAMPA format.
+For each word included in the questionnaire, we manually generated a broad phonological transcription in X-SAMPA format [@wells1995computercoding].
 Catalan word-forms were transcribed to Central Catalan phonology, and Spanish word-forms were transcribed to Castilian Spanish phonology.
-For each translation equivalent, we calculated the phonological similarity between the Catalan and the Spanish word-form using the same Levenshtein similarity measure used in @floccia2018introduction.
+For each translation equivalent, we calculated the phonological similarity between the Catalan and the Spanish word-form using the same Levenshtein similarity measure used in @floccia2018introduction [@heeringa2003norwegian; @schepens2012distributions; see also @fourtassi2020growth; @laing2022phonological for a similar approach].
 
@@
 Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
+ftn <- "For participants exposed to Catalan and Spanish exclusively this proportion is complementary to the degree of exposue to Spanish.
+For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10%."
 participants |>
@@
               list(cells_body(doe_catalan),
-                   cells_column_labels(doe_catalan)))
+                   cells_column_labels(doe_catalan))) |>
+    tab_footnote(ftn, location = cells_column_labels(doe_catalan))
 ```
@@
 This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: *sense escolaritzar/sin escolarizar* [no education], *educació primària/educación primaria* [primary school], *educació secundària/educación secundaria* [secondary school], *batxillerat/bachillerato* [complementary studies/high school], *cicles formatius/ciclos formativos* [vocational training], and *educació universitària/educación universitaria* [university degree].
-Most families reported university studies (`r participants_edu["University"]`, `r scales::percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r scales::percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r scales::percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r scales::percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).
+Most families reported university studies (`r participants_edu["University"]`, `r scales::percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r scales::percent(participants_edu["Vocational"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r scales::percent(participants_edu["Secondary"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r scales::percent(participants_edu["Complementary"]/sum(participants_edu))`), primary education (`r participants_edu["Primary"]`, <1%), and no formal education (`r participants_edu["No education"]`, <1%).
 
@@
                         .sep = ".")
-stan_version <- cmdstanr::cmdstan_version()
+stan_version <- "2.32.1"
 
@@
 We collected `r format(nrow(n_items), big.mark = ",")` items responses.
-We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, @fourtassi2020growth].
+We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, see @fourtassi2020growth for a similar approach].
 We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
@@
 
-We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, multilevel ordinal regression model.
+We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, multilevel, ordinal regression model.
 This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest.
@@
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
-Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data [see @eq-linear, @barr2013random].
+Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data (see @eq-linear) [@barr2013random].
 We specified a weakly informative prior around the parameters of the model (see Appendix C for details).
@@
 This region indicates the range of values that we considered equivalent to zero.
-The degree of overlap between the posterior distribution of a regression coefficient with the  can be interpreted as support for the true parameter of the coefficient being zero or equivalent.
-We then computed the 95% highest density interval (HDI) of the posterior distribution of each regression coefficient.
+We then summarised the posterior distribution of each regression coefficient with the 95% highest density interval (HDI).
 This interval contains the true value of this coefficient with 95% probability, given the data.
-Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, which indicates the probability that the true value of the regression coefficient falls into the ROPE, and should therefore be considered equivalent to zero.
-For example, a 90% overlap between the HDI of a coefficient's posterior distribution and the  indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the ROPE, and is therefore equivalent to zero.
+Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE ($p(\text{ROPE})$)), which indicates the probability that the true value of the regression coefficient falls into the ROPE (and therefore should be considered equivalent to zero).
+For example, a 80% overlap between the HDI of a coefficient's posterior distribution and the  indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero.
 
 Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
-We implemented the model using was done using the `brms` [@burkner2017brms], a R interface to the Stan probabilistic language [@carpenter2017stan; `r stan_version`].
+We implemented the model using was done using the `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
 We ran four MCMC  chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
@@
 intercept_label <- "Intercepts (at 22 months)"
-
+ftn <- "Proportion of overlap between the ROPE and the 95% HDI of the coefficient, interpreted as the probability of the true value of the coefficient being equivalent to zero."
 model_summary |>
@@
         .lower = md("95% HDI"),
-        .rope_overlap = md("*p*(H0)")
+        .rope_overlap = md("*p*(ROPE)")
     ) |>
@@
 This account is in line with accumulator models of language acquisition.
+Accumulator models devise word acquisition as a continuous process in which a child accumulates information for each word until they reach some theoretical threshold, from which the child is considered to have acquired a word.
+The age-of-acquisition of a word is determined by the age at which the a child reaches the threshold of cumulative learning instances, and the rate at which a child accumulates learning instances with a word is a function of child-level properties (e.g., ability, amount of quantitative language exposure) and word-level properties (e.g., lexical frequency) [@hidaka2013computational].
+In the present study, extended this account to the bilingual case, suggesting that the cognate facilitation effect on bilingual word acquisition is the result of a faster accumulation of learning instances driven by parallel activation.
+Accumulator models have been successfully applied to the study of vocabulary growth and word acquisition in previous work [@mcmurray2007defusing; @hidaka2013computational; @mollica2017how], but present some limitations that should also be considered in the context of the present study [see @bergelson2020comprehension].
+Overall, accumulator models are a convenient tool to unify compatible accounts of language acquisition under a coherent, formal theoretical framework [@kachergis2022standard], and present an opportunity for investigating bilingual language aquisition.
 
-This account of bilingual lexical acquisition addresses a critical subject in bilingualism research: do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
+In particular, the present study addressed a critical subject in bilingualism research: do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
 In the context of lexical acquisition, the former scenario predicts that every learning instance for a given word-form contributes to the consolidation of the representation of such word in the lexicon, while the consolidation of its translation remains unaffected by such experience.

### 2023-06-18 manuscript/manuscript.qmd
New version of introduction and discussion, n-th attempt to run Docker

@@
 The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
-For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins.
+For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4-month-wide bins.
 Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
@@
 #| tbl-note: "Median: median of the posterior distribution in the probability scale.
-95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.
-On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE).
-This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
-Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
+95\\% HDI: 95\\% highest density interval of the distribution.
+p(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero."
 #| echo: false
@@
 #| fig-cap: "Posterior marginal effects.
-Thin lines correspond to 50 mean predictions.
-Thicker lines indicate the median of those predictions.
+Lines correspond to 100 posterior-predicted medians.
 Different colours indicate different levels of cognateness (phonological similarity).
 Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
-Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom.
+Predictions for Comprehension are shown on top and predictions for Comprehension and Production are shown on the bottom.
 In-sample predictions lie inside the grey rectangles."
-clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")
+clrs_reorder <- c("#58508d",
+                  "#003f5c",
+                  "#122A36", 
+                  "#ffa600", 
+                  "#ff6361", 
+                  "#bc5090")
 
@@
                               "Production")) |> 
-    ggplot(aes(age, .value, colour = lv, fill = lv)) +
+    ggplot(aes(age, .value, 
+               colour = lv, 
+               fill = lv)) +
     facet_grid(.category~exposure) +
@@
     geom_line(aes(group = interaction(lv, .draw)),
-              linewidth = 3/4,
-              alpha = 1/20) +
-    stat_summary(fun = median, geom = "line", linewidth = 0.75) +
+              linewidth = 2/4,
+              alpha = 1/5) +
     labs(x = "Age (months)",
@@
 
-We suggest that parallel activation increases the rate at which bilingual children encounter learning instances for cognate words, but not for non-cognate words.
+Accumulator models devise word acquisition as a continuous process in which the child gathers information about words by accumulating learning instances with such words.
+When the number of cumulative learning instances for a word reaches some theoretical threshold, the child is considered to have acquired such word.
+The rate at which a child accumulates learning instances with a word is a function of child-level properties (e.g., ability, amount of quantitative language exposure) and word-level properties (e.g., lexical frequency) [@hidaka2013computational].
+Through statistical inference, formalised accumulator models provide meaningful information about parameters of interest like the aforementioned predictors [@mollica2017how; @kachergis2022standard], and allow to generate quantitative predictions about age-of-acquisition and vocabulary growth under compating theoretical accounts [@mcmurray2007defusing; @hidaka2013computational].
+Capistalising on accumulator models, we extended this account to the bilingual case, suggesting that the cognate facilitation effect on bilingual word acquisition is the result of a faster accumulation of learning instances driven by parallel activation.
+We suggested that parallel activation increases the rate at which bilingual children encounter learning instances for cognate words, but not for non-cognate words.
 We hypothesised that when a bilingual child is exposed to a word-form in presence of its referent, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
@@
 
-This account is in line with accumulator models of language acquisition.
-Accumulator models devise word acquisition as a continuous process in which a child accumulates information for each word until they reach some theoretical threshold, from which the child is considered to have acquired a word.
-The age-of-acquisition of a word is determined by the age at which the a child reaches the threshold of cumulative learning instances, and the rate at which a child accumulates learning instances with a word is a function of child-level properties (e.g., ability, amount of quantitative language exposure) and word-level properties (e.g., lexical frequency) [@hidaka2013computational].
-In the present study, extended this account to the bilingual case, suggesting that the cognate facilitation effect on bilingual word acquisition is the result of a faster accumulation of learning instances driven by parallel activation.
-Accumulator models have been successfully applied to the study of vocabulary growth and word acquisition in previous work [@mcmurray2007defusing; @hidaka2013computational; @mollica2017how], but present some limitations that should also be considered in the context of the present study [see @bergelson2020comprehension].
-Overall, accumulator models are a convenient tool to unify compatible accounts of language acquisition under a coherent, formal theoretical framework [@kachergis2022standard], and present an opportunity for investigating bilingual language aquisition.
-
-In particular, the present study addressed a critical subject in bilingualism research: do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
+These predictions address a critical subject in bilingualism research: do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
 In the context of lexical acquisition, the former scenario predicts that every learning instance for a given word-form contributes to the consolidation of the representation of such word in the lexicon, while the consolidation of its translation remains unaffected by such experience.

### 2023-06-18 R/predictions.R

@@ -68,7 +68,7 @@ generate_newdata <- function(model,
 #' @param model A brmsfit object
 #' @param data Data with the desired combination of levels of the predictors for which to generate posterior predictions, as generated by [generate_newdata()]
 #' @inheritParams marginaleffects::datagrid
-posterior_predictions <- function(model, data, ...) {
+posterior_predictions <- function(model, data, ndraws = NULL, ...) {
     # generate data for predictions
     newdata <- generate_newdata(model, data, ...)
     
@@ -78,7 +78,7 @@ posterior_predictions <- function(model, data, ...) {
         newdata = newdata,
         re_formula = NA,
         vcov = FALSE,
-        ndraws = 50
+        ndraws = ndraws
     ) |>
         marginaleffects::posteriordraws() |> # so that each draw gets a row
         as_tibble() |>

### 2023-06-27 docs/index.qmd
Fix figure and add new comments

@@
         postal-code: "28049"
-  - name: Núria Sebastian-Galles
+  - name: Nuria Sebastian-Galles
     orcid: 0000-0001-6938-2498
@@
 abstract: |
-    Language non-selectivity is prominent feature of bilingual lexical processing.
+    Language non-selectivity is prominent feature of the bilingual lexicon.
 An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates.
@@
 
-# Methods
-
 ::: {.callout-tip}
@@
 :::
-
-## Questionnaires
-
-```{r}
-#| label: questionnaire
-#| message: false
-#| warning: false
-data("pool", package = "bvq")
-n_categories <- distinct(pool, semantic_category) %>% nrow()
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% 
-    pull(n) %>%
-    set_names(c("catalan", "spanish"))
-```
-
-The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted of two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]` items.
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* \[dog\] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* \[*to finish*\] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* \[*monkey*\] and Catalan *mono/mico*.
-We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
-For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\~400 translation equivalents, \~400 items in Catalan, \~400 items in Spanish).
-These last families were randomly assigned to one of four different subsets of the complete list of items.
-These lists were designed so that each contained a representative subsample of the items from the complete list.
-Semantic/functional categories with less than 16 items were not divided in the short version of the questionnaire to preserve a minimum of four items per list for each category: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* \[cereal bar\]) and items that included more than one word-form (e.g., *mono / mico*).
-Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
 
@@
 Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
-This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points.
-Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [{childesr}](https://langcog.github.io/childes-db-website/) R package.
+Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [`childesr`](https://langcog.github.io/childes-db-website/) R package.
 Available words in the corpora were then mapped to their Catalan and Spanish translations.
-Therefore, the Catalan and Spanish word-forms of the same translation equivalent had the same value for lexical frequency.
-Responses to words with missing lexical frequencies were excluded from analyses of the total number of items).
+Responses to words with missing lexical frequencies were excluded from analyses.
 

### 2023-06-27 manuscript/manuscript.qmd
Fix figure and add new comments

@@
 Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
-Previous work on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and to both groups knowing a similar amount of words---if not more words---when the bilinguals’ two languages are pooled together.
-@hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did.
-Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together.
+Previous work on bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, and to both groups knowing a similar number of words---if not more words---when the bilinguals’ two languages are pooled together.
+@hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did, but both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together.
 Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
@@
 
-One important observation of studies on bilinguals’ early vocabulary acquisition is that cognate words are easier to acquire than non-cognate words.
+One important observation of studies on bilinguals' early vocabulary acquisition is that cognate words are easier to acquire than non-cognate words.
 Cognate words are translation equivalents that are phonologically similar (or share some type of form-similarity).
-For instance, the Spanish translation equivalent of cat is *gato*, a cognate word; the translation equivalent of dog is *perro*, a non-cognate word.
+For instance, the Spanish translation equivalent of *cat* is *gato*, a cognate word; the translation equivalent of *dog* is *perro*, a non-cognate word.
 The differences in the percentages of cognate and non-cognate words between two languages is related to historic reasons: languages typologically close (like Dutch and English or Italian and Spanish) share more cognates than languages typologically distant (like English and Chinese, or Urdu and Spanish).
@@
 @bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months and found that cognates represented a larger proportion of participant’s vocabulary than non-cognates.
-@schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual produced cognates earlier than non-cognates, on average.
+@schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual child produced cognates earlier than non-cognates, on average.
 @mitchell2022cognates addressed this issue in a larger sample longitudinal study.
@@
 Parallel activation of bilinguals’ lexicons has been proposed as the underlying mechanism for such facilitatory effect [e.g., @floccia2018introduction; @mitchell2022cognates].
-The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during language processing, even in fully monolingual environments.
-Evidence supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; see @kroll2017bilingual for review].
+The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during language processing, even in fully monolingual contexts Evidence with adult bilinguals supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; see @kroll2017bilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
 In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish.
-In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of tree), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of table), obviously, such distinction was only relevant for bilinguals.
+In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of *table*), obviously, such distinction was only relevant for bilinguals.
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
@@
 
-Although there is converging evidence on the role of parallel activation, the previous studies do not address its influence on the learning trajectories of words.
-Results are aggregated across words, and provide no information about the specific dynamics of how parallel activation influences word learning.
+Although there is a consensus on the role of parallel activation in bilinguals' lexical processing and acquisition, previous studies do not address its influence on the learning trajectories of words.
+Results are aggregated across words and provide no information about the specific dynamics of how parallel activation influences word learning.
 This is the goal of the present research.
 
-We propose an account in which a learning instance for a word may also represent a learning instance for its translation equivalent, only if such translation equivalent is co-activated.
-The probability of this co-activation is proportional to the phonological similarity between the two translation equivalents, and given that cognates share higher phonological similarity than non-cognates, the former may be co-activated more frequently than the latter.
+We propose an account in which a learning instance for a word may also represent a learning instance for its translation equivalent, to the extent that such translation equivalent is co-activated.
+The probability of this co-activation is proportional to the phonological similarity between the two translation equivalents; given that cognates share higher phonological similarity than non-cognates, the former should be co-activated more frequently than the latter.
 This should lead to a faster accumulation of learning instances for cognates, compared to non-cognates.
-Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across both languages: words from the lower-exposure language would receive additional learning instances through parallel activation more often than words from the higher-exposure language.
+Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across languages: words from the lower-exposure language would receive additional activation through parallel activation more often than words from the higher-exposure language.
 Therefore, the acquisition of words from the lower-exposure language would benefit more strongly from their cognate status than words from the higher-exposure language.
@@
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
-Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for dog) (see @tbl-items for a summary of the questionnaire items).
+Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping^[Although for some of the included words in Catalan did not have a clear translation or had more than one possible translation in Spanish, and vice versa, therefore the unequal number of words included in the Catalan and Spanish lists].
+(e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for *dog*) (see @tbl-items for a summary of the questionnaire items).
 
@@
 Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items.
-Each subset contained a random but representative sub-sample of the items from the complete list.
+Each subset contained a random but representative sub-sample of the items from the complete list (see @tbl-items).
 Semantic/functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
@@
 
-For each word included in the questionnaire, we manually generated a broad phonological transcription in X-SAMPA format [@wells1995computercoding].
+We manually generated a broad phonological transcription of every word included in the vocabulary checklists in X-SAMPA format [@wells1995computercoding].
 Catalan word-forms were transcribed to Central Catalan phonology, and Spanish word-forms were transcribed to Castilian Spanish phonology.
-For each translation equivalent, we calculated the phonological similarity between the Catalan and the Spanish word-form using the same Levenshtein similarity measure used in @floccia2018introduction [@heeringa2003norwegian; @schepens2012distributions; see also @fourtassi2020growth; @laing2022phonological for a similar approach].
+For each translation equivalent, we calculated the phonological similarity between the Catalan and the Spanish word-form using the Levenshtein distance, which measures the edit distance between each pair of phoneme strings [@levenshtein1966binary].
+The output indicates the number of editions (insertions, deletions, or substitutions) that one string of phonemes/characters must go through to become identical to the other.
+We then divided this value by the length of the longest phoneme string and subtracted the outcome from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
+This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different, and one indicates that the two word-forms are identical [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
 
@@
 Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`, and were always allocated to the same questionnaire list (A, B, C, or D).
-Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants) for a more detailed description of the sample).
+Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra) and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants) for a more detailed description of the sample).
 In total, `r n_trilingual$n[2]` participants (`r scales::percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed to a third language other than Catalan and Spanish.
@@
 This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest.
-We added *Age* (number of months elapsed between participants' birth date and questionnaire completion), *Length* (number of phonemes in the X-SAMPA phonological transcription of the word-form), *Exposure* (composite measure of Zipf-transformed lexical frequency and language degree of exposure), and *Cognateness* (a measure of cognateness, defined as the phonological similarity between translation equivalents) as population-level effects, together with their two- and three-way interactions.
+(1) *Age*, calculated as the number of months elapsed between participants' birth date and questionnaire completion).
+(2) *Length*, the number of phonemes in the X-SAMPA phonological transcription of the word-form.
+(3) *Exposure*, a language exposure-weighted lexical frequency.
+To account for the fact that bilinguals' exposure to a given word-form is a function of the word-forms lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
+First, we extracted the child-directed lexical frequency of each word from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
+Using the corresponding lexical frequencies directly from Catalan and Spanish was not possible due to the low number of Catalan participants and tokens available in their corresponding CHILDES corpora, so they were extracted from the English corpora instead.
+We then mapped the lexical frequencies of the English words to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed them to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human].
+We then multiplied the resulting lexical frequencies by the reported degree of exposure of the child to Catalan or Spanish.
+For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure score to the Catalan word *cavall* [*horse*]---with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`---would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.
+(4) *Cognateness*, defined as the phonological similarity between translation equivalents, and calculated as detailed in the *Questionnaire* section.
+These predictors were included as main effects in the model, together with their two-way and three-way interactions.
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
@@
 Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
-We implemented the model using was done using the `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
+We implemented the model using `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
 We ran four MCMC  chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
@@
 
-The main effect of $\text{Age}$ showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std"]]$.upper, 2)`]), with all of its posterior samples falling out of the ROPE.
+The main effect of $\text{Age}$ showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std"]]$.upper, 2)`]), with all posterior samples falling out of the ROPE.
 A one-month increment in age increased a maximum of `r round(post_draws_list[["b_age_std"]]$.median/sd(responses$age), 2)` the probability of acquisition.
@@
 Different colours indicate different levels of cognateness (phonological similarity).
+Black lines indicate the median of the posterior predictions.
 Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
@@
     geom_vline(xintercept = mean(responses$age), 
-               linetype = "dotted") +
+               colour = "grey") +
     geom_hline(yintercept = 0.5,
-               linetype = "dotted") +
+               colour = "grey") +
     geom_line(aes(group = interaction(lv, .draw)),
@@
               alpha = 1/5) +
+    stat_summary(geom = "line", fun = mean,
+                 linewidth = 2/4,
+                 colour = "black",
+                 aes(linetype = lv,
+                     group = interaction(lv, exposure))) +
     labs(x = "Age (months)",
@@
          fill = "Cognateness (phonological similarity)") +
-    guides(colour = guide_legend(override.aes = list(linewidth = 2))) +
+    guides(colour = guide_legend(override.aes = list(linewidth = 2,
+                                                     alpha = 1)),
+           linetype = guide_none()) +
     scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
@@
 This effect was mediated by exposure rate: low-exposure word-forms benefited from their cognate status, whereas high-exposure word-forms did not.
-Capitalising on accumulator models of language acquisition, we provided a theoretical account of bilingual lexical acquisition.
+Capitalising on accumulator models of language acquisition, we provide a theoretical account of bilingual lexical acquisition.
 In this account, parallel activation of the two languages plays a central role during the consolidation of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
 
-This investigation is particularly relevant in the light of two previous findings.
+The present investigation is particularly relevant in the light of two previous findings.
 First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (e.g.
@@
 But the underpinnings of such effect were unclear: while parallel activation has been extensively described in experimental studies, current paradigms of bilingual word acquisition and word learning are, to a large extent, dissociated from the mechanisms proposed by previous work on word processing.
-Accumulator models of language acquisition provide a convenient theoretical framework to narrow this gap.
+Accumulator models of language acquisition may provide a convenient theoretical framework to narrow this gap.
 
@@
 The rate at which a child accumulates learning instances with a word is a function of child-level properties (e.g., ability, amount of quantitative language exposure) and word-level properties (e.g., lexical frequency) [@hidaka2013computational].
-Through statistical inference, formalised accumulator models provide meaningful information about parameters of interest like the aforementioned predictors [@mollica2017how; @kachergis2022standard], and allow to generate quantitative predictions about age-of-acquisition and vocabulary growth under compating theoretical accounts [@mcmurray2007defusing; @hidaka2013computational].
-Capistalising on accumulator models, we extended this account to the bilingual case, suggesting that the cognate facilitation effect on bilingual word acquisition is the result of a faster accumulation of learning instances driven by parallel activation.
-We suggested that parallel activation increases the rate at which bilingual children encounter learning instances for cognate words, but not for non-cognate words.
-We hypothesised that when a bilingual child is exposed to a word-form in presence of its referent, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
+Through statistical inference, formalised accumulator models provide meaningful information about parameters of interest like the aforementioned predictors [@mollica2017how; @kachergis2022standard], and allow to generate quantitative predictions about age-of-acquisition and vocabulary growth under competing theoretical accounts [@mcmurray2007defusing; @hidaka2013computational].
+Capistalising on accumulator models, we extended this account to the bilingual case, suggesting that the cognate facilitation effect on bilingual word acquisition is the result of cognate words being activated more strongly by their translation than non-cognates, therefore accumulating learning instances at a faster rate.
+We hypothesised that when a bilingual child is exposed to a word-form, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
 This cross-language activation occurs at the phonological level, and the amount of co-activation that spreads from the spoken word to its translation is proportional to the amount of phonological similarity between both word-forms.
@@
 It might be argued that our results reflect the fact that cognate translation equivalents are represented in the initial bilingual lexicon as the *same* lexical entry.
-Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}}/ and /\textipa{"ga.to}}/ in the presence of a cat), it is possible that they classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
+Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}}/ and /\textipa{"ga.to}}/ in the presence of a cat), it is possible that infants classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
 This would lead to a faster increase in cumulative learning instances, and to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
@@
 @mitchell2022cognates discuss this possibility as a possible explanation of the cognate facilitation effect, in which bilinguals only need to map one word-form to the referent in the case of cognates, while mapping two distinct word-forms in the case of non-cognates.
-Previous work on mispronunciation perception and minimal pair learning points in a different direction.
-Bilingual toddlers show monolingual-level sensitivity slight phonetic changes in a word-form, according to their performance in word recognition tasks [@swingley200511montholds; @ramon-casas2017minimalpair; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels; @ramon-casas2010are; @ramon-casas2009vowel].
-These fine abilities to differentiate between similar-sounding word-forms are reflected in word learning, as bilinguals seem to be able to map minimal pairs to distinct referents [@ramon-casas2017minimalpair; @mattock2010first; @havy2016phonetic].
+However, previous work on mispronunciation perception and learning of minimal pair words points in a different direction.
+Bilingual toddlers show monolingual-level sensitivity to slight phonetic changes in a word-form, according to their performance in word recognition tasks [@swingley200511montholds; @ramon-casas2017minimalpair; @swingley2000spoken; @tamasi2017pupillometry; @mani2011does; @bailey2002phonological; @wewalaarachchi2017vowels; @ramon-casas2010are; @ramon-casas2009vowel].
+The ability to differentiate between similar-sounding word-forms is also reflected in word learning, as bilinguals seem to be able to map minimal pairs to distinct referents [@ramon-casas2017minimalpair; @mattock2010first; @havy2016phonetic].
 Overall, it seems that bilinguals consider small differences in the phonological forms of words as relevant at the lexical level.
@@
 Such reports can be subject to measurement error induced by caregivers who may sometimes overestimate or underestimate participants' true probability of acquisition of words [e.g., @houston-price2007discrepancy].
-For instance, although in the present study caregivers were explicitly instructed *not* to rely on their responses to Catalan words when responding to Spanish (and vice versa), it is possible that some caregivers assumed---at least to some extent---that because the child knew a word in one language, the child should also know the word in the other language.
+In the case of bilingual research additional biases may be in place.
+Although in the present study caregivers were explicitly instructed *not* to rely on their responses to Catalan words when responding to Spanish (and vice versa), it is possible that some caregivers assumed---at least to some extent---that because the child knew a word in one language, the child should also know the word in the other language.
 This bias would especially affect similar-sounding words, i.e., cognates.
-Production estimates may be more prompt to such biases, in part because of the slower pace at which their articulatory abilities develop, compared to their word recognition abilities.
-For instance, even at 47 months of age, few (typically developing) English monolinguals reach 100% intelligibility scores in word production, as rated by English native adults [@hustad2020development; @hustad2021speech].
+Production estimates may be more prompt to such biases, in part because of the slower pace at which infants’ articulatory abilities develop, compared to their word recognition abilities [@hustad2020development; @hustad2021speech].
 This gap between comprehension and production is even larger in the less dominant language of bilingual children [@giguere2022bilingual].
 For this reason, caregivers may be more uncertain about what words can be counted as *acquired* in this modality.
-Despite these potential biases, vocabulary checklist filled by parents show strong evidence of concurrent validity with other estimates of vocabulary size or lexical processing [@killing2008move; @feldman2005concurrent; @gillen2021tapping; but see @houston-price2007discrepancy].
+Despite such potential biases, vocabulary checklist filled by parents show strong evidence of concurrent validity with other estimates of vocabulary size or lexical processing [@killing2008move; @feldman2005concurrent; @gillen2021tapping; but see @houston-price2007discrepancy].
 
-The present study contributes a specific data point into the complex landscape of bilingualism research.
-Bilinguals are a remarkably heterogeneous population that can hardly be satisfactorily characterised in a comprehensive way [@sebastian-galles2020bilingual].
+The present study contributes with a specific data point to the complex landscape of bilingualism research.
+Bilinguals are a remarkably heterogeneous population difficult to be satisfactorily characterised in a comprehensive way [@sebastian-galles2020bilingual].
 Bilinguals differ across multiple dimensions.
-These differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to higher-level factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
+Such differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to higher-level factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
 This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.
-Our sample, although homogeneous (e.g., similar parental educational level across), represents a fairly particular population of bilinguals: the languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages, both languages are used in fairly similar contexts, and both languages are known by the majority of the population.
+Our sample, although homogeneous (e.g., similar parental educational level across), represents a particular bilingual sociolinguistic environment: the languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages, both languages are used in fairly similar contexts, and both languages are known by the majority of the population.
 In 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@2018els].
-In addition, and as commented in the introduction, Catalan and Spanish are both Romance languages and share a considerable amount of cognates.
+In addition, Catalan and Spanish are Romance languages and share a considerable amount of cognates.
 Extending our analyses to other bilingual populations learning typologically more distant languages, and whose languages tend to be used in more distinct contexts (e.g., heritage languages) should be a natural future step for the present investigation.
 
-In summary, our study provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
+To conclude, our study provides novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
 We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word's phonological similarity with its translation in the other language.
-Capitalising on accumulator models of language acquisition we, put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
+Capitalising on accumulator models of language acquisition we put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.

### 2023-06-28 manuscript/_appendix/_model-doe.qmd
Update PDF format

@@
+A a robustness check, we fit a model similar to the one described in the main manuscript, but including lexical frequency and language degree of exposure as separate predictors, instead of the composite measure *Exposure*.
+Language degree of exposure (*DoE*) was included in interaction with *Age* and *Cognateness*, while lexical frequency (*Frequency*) was included as a main effect.
+
 ```{r fig-coefs-doe}

### 2023-06-28 manuscript/_appendix/_syllable-frequency.qmd
Update PDF format

@@
 
-@fig-top-syll-freq shows the top twenty most frequent syllables in Spanish.
-In both languages, most of the most frequent syllables followed a consonant-vowel (CV) structured.
-In Catalan, the most common nucleus in the most frequent syllables was /\textschwa/ (most likely due to vowel reduction in unstressed syllables), followed by /u/.
-In Spanish, the most common nucleus in the most frequent syllables were /a/ and /o/.
-
-```{r fig-top-syll-freq}
-#| label: fig-top-syll-freq
-#| fig-cap: "Top 20 most frequent syllables"
-#| fig-width: 8
-#| fig-height: 3
-#| out-width: 80%
-#| echo: false
-#| warning: false
-#| message: false
-syll |> 
-    group_by(language) |> 
-    top_n(20, freq_syll) |> 
-    ggplot(aes(tidytext::reorder_within(syll, desc(freq_syll), language),
-               freq_syll)) +
-    facet_wrap(~ language,
-               scales = "free_x",
-               ncol = 1) +
-    geom_col(colour = "white",
-             fill = "grey70") +
-    labs(x = "Syllable",
-         y = "Relative frequency\n(counts per million)",
-         colour = "Language",
-         fill = "Language") +
-    tidytext::scale_x_reordered() +
-    scale_y_continuous(labels = \(x) format(x, big.mark = ",")) +
-    theme(legend.position = "none",
-          axis.title.x = element_blank(),
-          axis.text.x = element_text(size = 10),
-          panel.grid = element_blank())
-```
-
-To estimate the association between word-level syllabic frequency and cognateness while controlling for the number of syllables in the word, as words are expected to necessarily increase the syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ (Levenshtein similarity between a word and its translation equivalent) as predictors.
+To estimate the association between word-level syllabic frequency and cognateness, while controlling for the number of syllables in the word, as words are expected to necessarily increase the syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ (Levenshtein similarity between a word and its translation equivalent) as predictors.
 We added translation equivalent-level random effects for the intercept and the main effect of $Syllables$ (some translation pairs had a different number of syllables in each language).

### 2023-06-28 manuscript/manuscript.qmd
Update PDF format

@@
         department: Center for Brain and Cognition
-        address: Ramon Trias Fargas 25-27
-        city: Barcelona
-        region: Spain
-        postal-code: "08005"
+        city: Barcelona, 08005
   - name: Daniela S. Avila-Varela
     orcid: 0000-0002-3518-8117
+    email: avila.varela.daniela@gmail.com
     affiliations: 
@@
         department: Facultade de Letras
-        address: Alameda da Universidade
-        city: Lisboa
-        region: Portugal
-        postal-code: "1600-214"
+        city: Lisboa, 1600-214
   - name: Ignacio Castillejo
     orcid: 0000-0001-7445-0416
+    email: jignaciocastillejo@gmail.com
     affiliations: 
@@
         department: Departamento de Psicología
-        address: Iván Pavlov, 6
-        city: Madrid
-        region: Spain
-        postal-code: "28049"
-  - name: Núria Sebastian-Galles
+        city: Madrid, 28049
+  - name: Nuria Sebastian-Galles
     orcid: 0000-0001-6938-2498
+    email: nuria.sebastian@upf.edu
     affiliations: 
       - ref: cbc
-date: "May 10th, 2023"
-code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
 keywords:
-  - lexical acquisition
+  - cognate
+  - word acquisition
   - vocabulary
@@
 format:
-  pdf:
-    papersize: letter
-    fontsize: 12pt
-    linestretch: 1.5
-    date-format: full
-    template-partials:
-    - templates/title.tex
-    - templates/before-body.tex
+  arxiv-pdf:
+    keep-tex: true  
+    linenumbers: true
+    doublespacing: true
+    runninghead: "Cognate beginnings to bilingual lexical acquistion"
     include-in-header: 
-    - templates/include-in-header.tex
-    header-includes: 
-    - \usepackage{tipa}
-    - \usepackage{nowidow}
-    geometry:
-    - top=1in
-    - bottom=1in
-    - left=1in
-    - right=1in
-    colorlinks: true
-    linkcolor: DodgerBlue
-    urlcolor: DodgerBlue
-    citecolor: DodgerBlue
-    link-citations: true
-    biblio-config: false
-    cite-method: biblatex
-    biblatex-chicago: false
-    biblio-style: apa
-    biblatexoptions: backend=biber
-    block-headings: false
-    indent: true
-  docx:
-    knitr:
-      opts_chunk:
-        dev: "png"
-        dev-args:
-          - "cairo-png"
-        dpi: 300
-        out.width: "80%"
-author-note:
-  disclosures:
-    conflict-of-interest: |
-        The authors declare no conflicts of interest with regard to the funding source of this study.
-    financial-support: |
-        This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
-Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
-    gratitude: |
-        We are grateful to Chiara Santolin, Ege E.
-Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
-We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support, and Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
-We also thank all families and infants who participated in the experiments.
+      text: |
+        \usepackage{tipa}
+        \usepackage{booktabs}
+abstract: |
+    Bilingual infants show equivalent developmental trajectories of lexical acquisition compared to their monolingual peers.
+This is remarkable, given the increased complexity of their linguistic input.
+Recent studies suggested that bilingual vocabulary growth is boosted by the number of cognates shared by the pair of languages being learned, and that this facilitation effect is driven by a stronger parallel activation of cognates during linguistic exposure, compared to non-cognates.
+The mechanisms behind this facilitation are still unclear.
+In this study, we capitalise on accumulator models of language acquisition to propose an account of bilingual lexical acquisition in which parallel activation increases the rate at which children accumulate learning instances for words in both languages, even in fully monolingual situations.
+Under this hypothesis, we predicted a stronger cognate facilitation for words to which children were exposed less frequently (low-exposure words), as they are co-activated by their translation more often than high-exposure words do.
+We developed an extensive online vocabulary checklist, the Barcelona Vocabulary questionnaire (BVQ) to collect vocabulary data from 366 Catalan-Spanish bilingual toddlers aged 12 to 32 months.
+We then used Bayesian explanatory item response theory to model the acquisition trajectories of 302 Catalan/Spanish translation equivalents.
+We found an interaction between exposure and cognateness, which pointed to cognateness facilitating the aquisition of low-exposure words, but not of mean exposure or high-exposure words.
+Overall, our findings suggest that cognateness plays a key role in bilingual lexical acquisition, and provides evidence for a frequency-mediated facilitation effect driven by parallel activation.
 
@@
 
-clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
 my_theme <- theme_minimal() +
@@
 set.seed(888)
-
-options(ggplot2.ordinal.fill = clrs,
-        ggplot2.ordinal.colour = clrs,
-        ggplot2.discrete.fill = clrs,
-        ggplot2.discrete.colour = clrs,
-        ggplot2.continuous.fill = scale_color_gradient,
-        ggplot2.continuous.colour = scale_color_gradient,
-        knitr.graphics.error = FALSE) 
 ```
 
+# Introduction[^author-note]
+
+[^author-note]: The authors declare no conflicts of interest with regard to the funding source of this study.
+This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
+Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
+We are grateful to Chiara Santolin, Ege E.
+Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
+We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support, and Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
+We also thank all families and infants who participated in the experiments.
+
 The foundations of word learning are in place early in age.
-At six months of age, infants start directing their gaze to some objects when hearing their labels [@tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and, shortly after, caregivers start reporting some words as acquired by their infant in vocabulary checklists [e.g., @fenson2007macarthurbates; @samuelson2021precision].
+At six months, infants start directing their gaze to some objects when hearing their labels [@tincoff1999beginnings; @bergelson2012months; @bergelson2015early], and shortly after caregivers start reporting some words as acquired by their infant in vocabulary checklists [e.g., @fenson2007macarthurbates; @samuelson2021precision].
 Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
@@
 
-We propose an account in which a learning instance for a word may also represent a learning instance for its translation equivalent, to the extent that such translation equivalent is co-activated.
-The probability of this co-activation is proportional to the phonological similarity between the two translation equivalents; given that cognates share higher phonological similarity than non-cognates, the former should be co-activated more frequently than the latter.
+We propose an account in which a learning instance for a word may also represent a learning instance for its translation equivalent, to the extent that such translation equivalent is co-activated^[We use the term *learning instance* in the fashion of accumulator models of language acquisition; as an exposure to a word-form that constitutes an opportunity for the child to accumulate information about the word.
+We do not make strong assumptions about whether a learning instance is a discrete or a continuous unit of accumulation of information.
+We rather consider that a learning instance of a word is so to the degree of the strength of activation of its phonological representation.
+This activation may be the result of the infant being exposed to the word-form, or the result of activation spreading through phonological or semantic links across lexicl representation, as in the case of parallel activation.].
+The strength of this co-activation is proportional to the phonological similarity between the two translation equivalents; given that cognates share higher phonological similarity than non-cognates, the former should be co-activated more frequently than the latter.
 This should lead to a faster accumulation of learning instances for cognates, compared to non-cognates.
@@
 
-To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire [@garcia-castro2023bvq].
+To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire (BVQ) [@garcia-castro2023bvq].
 This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formr platform [@arslan2020formr].
 This questionnaire is structured in three blocks: (1) a language exposure questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (OCDI) [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
 Both lists included items from a diverse sample of `r n_categories` semantic/functional categories.
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
-Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping^[Although for some of the included words in Catalan did not have a clear translation or had more than one possible translation in Spanish, and vice versa, therefore the unequal number of words included in the Catalan and Spanish lists].
-(e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for *dog*) (see @tbl-items for a summary of the questionnaire items).
+Items in one language were translation equivalents of the items in the other language (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for *dog*), roughly following a one-to-one mapping^[Although for some of the included words in Catalan did not have a clear translation or had more than one possible translation in Spanish, and vice versa, therefore the unequal number of words included in the Catalan and Spanish lists]  (see @tbl-items for a summary of the questionnaire items).
 
@@
 items |> 
+    distinct(te, .keep_all = TRUE) |> 
     arrange(language, list) |> 
-    distinct(te, .keep_all = TRUE) |> 
     mutate(a = grepl("A", list),
@@
            d = grepl("D", list)) |> 
-    group_by(semantic_category) |> 
     summarise(across(a:d, sum), 
               items = list(meaning),
-              .groups = "drop") |> 
+              .by = semantic_category) |> 
     mutate(total = rowSums(across(a:d)),
-           items = map(items, function(x) head(x, 2))) |>  
+           items = map(items, function(x) tail(x, 2))) |>  
     arrange(desc(total)) |> 
     relocate(a:d, total, items) |> 
-    gt(rowname_col = "semantic_category") |> 
+    mutate(group = "Semantic category") |> 
+    select(-total) |> 
+    gt(groupname_col = "group",
+       rowname_col = "semantic_category") |> 
     fmt_percent(matches("_prop"), decimals = 1) |> 
@@
                d = "List D",
-               total = "Total",
                items = "Examples") |> 
+    summary_rows(groups = "Semantic category",
+                 columns = a:d,
+                 fns = list("Total" = ~sum(., na.rm = TRUE))
+    ) |> 
     tab_style(cell_text(size = "small"),
               list(cells_body(),
+                   cells_summary(),
                    cells_stub(),
@@
                    cells_column_spanners(),
+                   cells_stub_summary(),
                    cells_row_groups())) |> 
@@
 
-We manually generated a broad phonological transcription of every word included in the vocabulary checklists in X-SAMPA format [@wells1995computercoding].
+To compute predictors of interest, we manually generated a broad phonological transcription of every word included in the vocabulary checklists in X-SAMPA format [@wells1995computercoding].
 Catalan word-forms were transcribed to Central Catalan phonology, and Spanish word-forms were transcribed to Castilian Spanish phonology.
-For each translation equivalent, we calculated the phonological similarity between the Catalan and the Spanish word-form using the Levenshtein distance, which measures the edit distance between each pair of phoneme strings [@levenshtein1966binary].
-The output indicates the number of editions (insertions, deletions, or substitutions) that one string of phonemes/characters must go through to become identical to the other.
-We then divided this value by the length of the longest phoneme string and subtracted the outcome from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
-This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different, and one indicates that the two word-forms are identical [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
 
@@
     gt() |>
-    sub_missing(everything(), everything(), missing_text = "--") |>
+    sub_missing(everything(), everything(), missing_text = "-") |>
     cols_label(doe_catalan = "Catalan exposure") |>
@@
 
-We collected `r format(nrow(n_items), big.mark = ",")` items responses.
+We collected responses to `r format(nrow(n_items), big.mark = ",")` items.
 We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, see @fourtassi2020growth for a similar approach].
@@
 We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, multilevel, ordinal regression model.
-This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest.
-(1) *Age*, calculated as the number of months elapsed between participants' birth date and questionnaire completion).
-(2) *Length*, the number of phonemes in the X-SAMPA phonological transcription of the word-form.
-(3) *Exposure*, a language exposure-weighted lexical frequency.
-To account for the fact that bilinguals' exposure to a given word-form is a function of the word-forms lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
+This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest: *Age* (number of months elapsed between participants' birth date and questionnaire completion), *Length* (number of phonemes in the X-SAMPA phonological transcription of the word-form), *Exposure* (a language exposure-weighted lexical frequency), and *Cognateness* (defined as the phonological similarity between translation equivalentsFor each translation equivalent).
+We added these variables as main effects, together with the two-way and three-way interactions between *Age*, *Exposure*, and *Cognateness*.
+
+We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-forms lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
 First, we extracted the child-directed lexical frequency of each word from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
@@
 For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure score to the Catalan word *cavall* [*horse*]---with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`---would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.
-(4) *Cognateness*, defined as the phonological similarity between translation equivalents, and calculated as detailed in the *Questionnaire* section.
-These predictors were included as main effects in the model, together with their two-way and three-way interactions.
+
+We defined *Cognateness* as the phonological similarity between each word-form and its translation.
+For each translation equivalent, we used the `stringdist` [@vanderloo2014stringdist] R package to calculate the Levenshtein distance between the Catalan and the Spanish phonological transcriptions of the word-forms.
+The Levenshtein distance measures the number of editions (insertions, deletions, or substitutions) that one string of phonemes/characters must go through to become identical to the other [@levenshtein1966binary].
+We divided the Levnehstein distance of each translation equivalent by the length of the longest word-form to correct for word length (longer strings are likely to show a larger number of mismatches).
+Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
+This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /'taw.l5-'me.sa/, *table*), and one indicates that the two word-forms are identical (e.g., /'mar-'mar/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
+
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
-Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data (see @eq-linear) [@barr2013random].
-We specified a weakly informative prior around the parameters of the model (see Appendix C for details).
+Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data [@barr2013random].
+We specified a weakly informative prior around the parameters of the model.
+@eq-linear shows a detailed description of the model.
 
@@
 &\textbf{Response}~(k)~\textbf{to word}~i~ \textbf{by participant}~j \\
-\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}), \\ 
+\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}) \\ 
 &\text{where}~k \in \{\text{No} \rightarrow \text{Understands}, \text{Understands} \rightarrow \text{Understands and Says}\}\\
@@
 & (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij}) \\
+&\text{where:}\\
+& u_{1-8_{i}}~\text{: participant-level adjustments} \\
+& w_{1-3_{j}}~\text{: TE-level adjustments} \\
 &\textbf{Prior} \\
 \beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.5); ~\beta_{1-5} \sim \mathcal{N}(0, 1) \\
-\sigma_{u_{0-8}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2)
+\sigma_{u_{0-8}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2) \\
+&\text{where }~\rho_{u_{0-8}, w_{0-3}}~\text{are the correlations between group-level adjustments}\\
 \end{aligned}
@@
 This interval contains the true value of this coefficient with 95% probability, given the data.
-Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE ($p(\text{ROPE})$)), which indicates the probability that the true value of the regression coefficient falls into the ROPE (and therefore should be considered equivalent to zero).
+Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, noted as $p(\text{ROPE})$, which indicates the probability that the true value of the regression coefficient falls into the ROPE (and therefore should be considered equivalent to zero).
 For example, a 80% overlap between the HDI of a coefficient's posterior distribution and the  indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero.
@@
 # Results {#sec-results}
+
+\renewcommand{\arraystretch}{1.25}
 
@@
 
-All parameters in the model showed adequate MCMC convergence diagnostics and little evidence of multicollinearity (see Appendix D for details).
+Model posterior showed adequate MCMC convergence diagnostics for all parameters ($\hat{R} \le 1.06$, $0.022 \le N_{eff}/N < 1.1001$) and little evidence of multicollinearity.
 @tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients, and their degree of overlap with the ROPE.
-For interpretability, we report the estimated regression coefficients transformed to the probability scale.
-The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales relate non-linearly.
+For interpretability, we report the estimated regression coefficients transformed to the probability scale^[The logit and probability scales relate non-linearly.
 This means that one logit difference is not necessarily translated to a unique value in the probability scale.
@@
 This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
+The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor.
 
@@
 intercept_label <- "Intercepts (at 22 months)"
-ftn <- "Proportion of overlap between the ROPE and the 95% HDI of the coefficient, interpreted as the probability of the true value of the coefficient being equivalent to zero."
 model_summary |>
     select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
-    mutate(across(.median:.upper,
-                  \(x) ifelse(.type==intercept_label, plogis(x), x/4))) |> 
-    gt(groupname_col = ".type", 
-       rowname_col = ".variable_name") |>
-    fmt_percent(.rope_overlap, decimals = 3) |> 
-    fmt_number(c(.median:.upper), decimals = 3) |>
-    cols_merge(c(.lower, .upper),
-               pattern = "[{1}, {2}]") |>
-    cols_label(
-        .variable_name = "Parameter",
-        .median = "Median",
-        .lower = md("95% HDI"),
-        .rope_overlap = md("*p*(ROPE)")
+    mutate(
+        across(.median:.upper,
+               \(x) ifelse(.type==intercept_label, plogis(x), x/4)),
+        .variable_name = gsub("×", '$\\\\times$', .variable_name),
+        .hdi = paste0("[", round(.lower, 3),
+                      ", ", round(.upper, 3), "]")) |> 
+    select(.variable_name, .median, .hdi, .rope_overlap) |> 
+    knitr::kable(format = "latex", 
+                 escape = FALSE,
+                 col.names = c("",
+                               "Median",
+                               "95\\% HDI",
+                               "$p(\\text{ROPE})$"),
+                 digits = 4,
+                 booktabs = TRUE,
+                 linesep = ""
     ) |> 
-    tab_style(cell_text(weight = "bold"),
-              cells_column_labels(columns = 1:5)) |>
-    tab_style(cell_text(align = "left"),
-              cells_title(groups = "subtitle")) |> 
-    tab_style(cell_text(size = "small"), 
-              list(cells_body(),
-                   cells_row_groups(),
-                   cells_stub())) |> 
-    tab_style(cell_text(style = "italic"),
-              cells_row_groups())
+    kableExtra::kable_styling() |> 
+    kableExtra::pack_rows("Intercepts (at 22 months)", start_row = 1, end_row = 2) |> 
+    kableExtra::pack_rows("Slopes (upper bound)", start_row = 3, end_row = 10)
+
 ```
 
-The main effect of $\text{Age}$ showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std"]]$.upper, 2)`]), with all posterior samples falling out of the ROPE.
+The main effect of *Age* showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std"]]$.upper, 3)`]), with all posterior samples falling out of the ROPE.
 A one-month increment in age increased a maximum of `r round(post_draws_list[["b_age_std"]]$.median/sd(responses$age), 2)` the probability of acquisition.
-Similarly, the word exposure index ($\text{Exposure}$) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 2)`]).
+Similarly, the word exposure index (*Exposure*) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 3)`]).
 All of the posterior samples of this regression coefficient excluded the ROPE.
-The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), 2)` more likely to acquire it.
-Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r round(post_draws_list[["b_n_phon_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_phon_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_phon_std"]]$.upper, 2)`]).
-For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.median)/sd(responses$n_phon), 2)` less likely to know it.
-The 95% HDI of the regression coefficient of the $\text{Age} \times \text{Exposure}$ interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 2)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 2)`]), showing that the effect of the word exposure index differed across ages: older children were more sensitive likely to acquire words with higher exposure rate than younger children.
+The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), 3)` more likely to acquire it.
+Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r round(post_draws_list[["b_n_phon_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_n_phon_std"]]$.lower, 3)`, `r round(post_draws_list[["b_n_phon_std"]]$.upper, 3)`]).
+For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.median)/sd(responses$n_phon), 3)` less likely to know it.
+The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word exposure index differed across ages: older children were more sensitive likely to acquire words with higher exposure rate than younger children.
 
-The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).
-For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), 2)`.
-The effect of $\text{Cognateness}$ interacted with that of $\text{Exposure}$: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 2)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
+The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 3)`]).
+For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), 3)`.
+The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
 Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
@@
 Different colours indicate different levels of cognateness (phonological similarity).
-Black lines indicate the median of the posterior predictions.
 Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
@@
 In-sample predictions lie inside the grey rectangles."
-clrs_reorder <- c("#58508d",
-                  "#003f5c",
-                  "#122A36", 
-                  "#ffa600", 
-                  "#ff6361", 
-                  "#bc5090")
-
 model_epreds |> 
@@
              ymax = Inf,
-             fill = "grey95",
-             colour = NA) +
+             fill = "grey98",
+             linetype = "dotted") +
     annotate(geom = "text",
-             x = mean(responses$age),
-             y = 0,
-             hjust = -0.1,
-             vjust = 0,
-             label = "Observed age range",
+             x = min(responses$age),
+             y = 1,
+             hjust = 0,
+             vjust = 0.9,
+             label = "Observed\nage range",
              size = 2) +
     geom_vline(xintercept = mean(responses$age), 
-               colour = "grey") +
+               linewidth = 1/4,
+               linetype = "dotted") +
     geom_hline(yintercept = 0.5, 
-               colour = "grey") +
+               linewidth = 1/4,
+               linetype = "dotted") +
     geom_line(aes(group = interaction(lv, .draw)),
               linewidth = 2/4,
-              alpha = 1/5) +
-    stat_summary(geom = "line", fun = mean,
-                 linewidth = 2/4,
-                 colour = "black",
-                 aes(linetype = lv,
-                     group = interaction(lv, exposure))) +
+              alpha = 1/2) +
+    # stat_summary(geom = "line", fun = mean,
+    #              linewidth = 2/4,
+    #              colour = "black",
+    #              aes(linetype = lv,
+    #                  group = interaction(lv, exposure))) +
     labs(x = "Age (months)",
          y = "Probability of acquisition",
-         linetype = "Cognateness (phonological similarity)",
          colour = "Cognateness (phonological similarity)",
-         fill = "Cognateness (phonological similarity)") +
-    guides(colour = guide_legend(override.aes = list(linewidth = 2,
-                                                     alpha = 1)),
-           linetype = guide_none()) +
-    scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
-    scale_fill_manual(values = clrs_reorder[c(4, 5, 1)]) +
+         fill = "Cognateness (phonological similarity)\n") +
+    scale_color_grey(start = 0.9, end = 0.2,
+                     guide = guide_legend(
+                         override.aes = list(
+                             linewidth = 5,
+                             alpha = 1
+                         )
+                     )) +
     scale_y_continuous(breaks = seq(0, 1, 0.25),
@@
     theme(legend.position = "top",
+          legend.box = "vertical",
+          legend.title.align = 1,
+          legend.justification = "right",
+          legend.direction = "horizontal",
           legend.key.size = unit(1, "cm"),
@@
 
+An additional analysis including lexical frequency and language exposure as separate predictors (instead of the composite *Exposure* measure) showed equivalent results (see Appendix A).
 To rule out the possibility that cognateness facilitation effect we found was due to cognateness comprising more frequent syllables than non-cognates---and therefore not because of their cognate status itself---, we compared the syllabic frequency of cognates and non-cognates included in our analyses.
@@
 Finally, for each word, we summed the frequency of its syllables, as an estimate of the syllabic frequency of the word.
-We fit a Bayesian model with cognateness (Levenshtein similarity) as response variable, and the main effects of syllable frequency and number of syllables (to control for the fact words with more syllables are more likely to score higher in syllabic frequency) as predictors.
-This model provided strong evidence for the association between cognateness and syllabic frequency being equivalent to zero (see Appendix E).
+We fit a Bayesian model with *Cognateness* as response variable, and the main effects of syllable frequency and number of syllables (to control for the fact words with more syllables are more likely to score higher in syllabic frequency) as predictors.
+This model provided strong evidence for the association between cognateness and syllabic frequency being equivalent to zero (see Appendix B).
 

### 2023-06-28 manuscript/manuscript.tex
Add manuscript comments
Update PDF format

--- 2021-09-30 Rmd/manuscript.tex
@@
 \PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
 %
 \documentclass[
-  english,
-  man,man,floatsintext]{apa6}
+]{article}
+
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+  \setmainfont[]{Latin Modern Roman}
+  \setmathfont[]{Latin Modern Math}
 \fi
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={The role of cross-linguistic lexical similarity on bilingual word acquisition},
-  pdfauthor={Gonzalo Garcia-Castro1, Daniela Avila-Varela1, \& Nuria Sebastian-Galles1},
-  pdflang={en-EN},
-  pdfkeywords={lexical acquisition, vocabulary, bilingualism},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 \makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-% Manuscript styling
-\usepackage{upgreek}
-\captionsetup{font=singlespacing,justification=justified}
-
-% Table formatting
-\usepackage{longtable}
-\usepackage{lscape}
-% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
-\usepackage{multirow}		% Table styling
-\usepackage{tabularx}		% Control Column width
-\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
-\usepackage{threeparttablex}            % Lets threeparttable work with longtable
-
-% Create new environments so endfloat can handle them
-% \newenvironment{ltable}
-%   {\begin{landscape}\centering\begin{threeparttable}}
-%   {\end{threeparttable}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
-
-% Enables adjusting longtable caption width to table width
-% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
-\makeatletter
-\newcommand\LastLTentrywidth{1em}
-\newlength\longtablewidth
-\setlength{\longtablewidth}{1in}
-\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
-
-% \setlength{\parindent}{0.5in}
-% \setlength{\parskip}{0pt plus 0pt minus 0pt}
-
-% \usepackage{etoolbox}
-\makeatletter
-\patchcmd{\HyOrg@maketitle}
-  {\section{\normalfont\normalsize\abstractname}}
-  {\section*{\normalfont\normalsize\abstractname}}
-  {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
-  {\section{\protect\normalfont{\@title}}}
-  {\section*{\protect\normalfont{\@title}}}
-  {}{\typeout{Failed to patch title.}}
-\makeatother
-\shorttitle{Cross-linguistic similarity and  word acquisition}
-\keywords{lexical acquisition, vocabulary, bilingualism\newline\indent Word count: X}
-\usepackage{lineno}
-
-\linenumbers
-\usepackage{csquotes}
-\ifXeTeX
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
 \newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
@@
   % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
   % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
+  \setlength{\parskip}{#2\cslentryspacingunit}
  }%
@@
 
-\title{The role of cross-linguistic lexical similarity on bilingual word acquisition}
-\author{Gonzalo Garcia-Castro\textsuperscript{1}, Daniela Avila-Varela\textsuperscript{1}, \& Nuria Sebastian-Galles\textsuperscript{1}}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{tipa}
+\usepackage{booktabs}
+\usepackage{arxiv}
+\usepackage{orcidlink}
+\usepackage{amsmath}
+\usepackage[T1]{fontenc}
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Cognate beginnings to bilingual lexical acquisition},
+  pdfauthor={Gonzalo Garcia-Castro; Daniela S. Avila-Varela; Ignacio Castillejo; Nuria Sebastian-Galles},
+  pdflang={en},
+  pdfkeywords={cognate, word acquisition, vocabulary, bilingualism, item
+response theory, bayesian},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
+
+\usepackage{lineno}
+\linenumbers
+\usepackage{setspace}
+\doublespacing
+\newcommand{\runninghead}{A Preprint }
+\renewcommand{\runninghead}{Cognate beginnings to bilingual lexical
+acquistion }
+\title{Cognate beginnings to bilingual lexical acquisition}
+\author{
+\textbf{Gonzalo Garcia-Castro}~\orcidlink{0000-0002-8553-4209}\\Center
+for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
+08005\\\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}\\\\\\
+\textbf{Daniela S. Avila-Varela}~\orcidlink{0000-0002-3518-8117}\\Center
+for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
+08005\\\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}\\\\\\
+\textbf{Ignacio
+Castillejo}~\orcidlink{0000-0001-7445-0416}\\Departamento de
+Psicología\\Universidad Autónoma de Madrid\\Madrid,
+28049\\\href{mailto:jignaciocastillejo@gmail.com}{jignaciocastillejo@gmail.com}\\\\\\
+\textbf{Nuria Sebastian-Galles}~\orcidlink{0000-0001-6938-2498}\\Center
+for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
+08005\\\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}
 \date{}
-
-\authornote{
-
-Campus de Ciutadella, Universitat Pompeu Fabra, 08005, Barcelona, Spain
-
-Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona, Spain.
-E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}
-
-}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra, Barcelona, Spain}
-
-\abstract{
-Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.~(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months.
-}
-
 \begin{document}
 \maketitle
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Before the end of their first year of life, infants start directing their gaze to some objects when hearing their labels, according to both experimental data (Bergelson \& Swingley, 2012, 2015; Jusczyk \& Aslin, 1995; Tincoff \& Jusczyk, 1999) and parental reports (e.g., Fenson \& others, 2007).
-During the last half of their second year, they acquire new words at an increasingly fast rate (Bergelson, 2020; Bloom, 2002; Fenson et al., 1994; Goldfield \& Reznick, 1990; Mayor \& Plunkett, 2011).
-These early stages of lexical acquisition are characterised by substantial variation between children, reflected, for instance, on the variation on the number of words they know {[}i.e.~vocabulary size; Fenson et al.
-(1994); Frank, Braginsky, Yurovsky, and Marchman (2021){]} or on the proportion of those words that fall into the category of nouns, as opposed to verbs, adjectives, or function words (e.g., Nelson, 1973, p.
-bates1994developmental).
-
-Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages.
-Tardif et al.
-(2008) collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing.
-Since birth, these infants had been learning English, Cantonese and Mandarin, respectively.
-The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (\emph{daddy}, \emph{mommy}), social routines (\emph{bye}, \emph{uh-oh}) or animals (\emph{woof-woof}).
-These results were later extended by Frank et al.
-(2021) to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages.
-Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.
-
-There is evidence that bilinguals know less words in each of their languages than monolinguals, but also that both groups know a similar amount of words when the two languages are aggregated.
-For example, Hoff et al.
-(2012) found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals, who only learnt English.
-Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, pointing to the importance of collecting data on both languages when assessing bilinguals' communicative development.
-Other studies have provided converging evidence that bilinguals know a similar--or even larger--number of words than monolinguals, only when the languages are aggregated (Gonzalez-Barrero, Schott, \& Byers-Heinlein, 2020; Oller \& Eilers, 2002; J.
-L.
-Patterson, 2004; J.
-Patterson, Pearson, \& Goldstein, 2004; Pearson \& Fernández, 1994; Pearson, Fernández, \& Oller, 1993; Petitto et al., 2001; Smithson, Paradis, \& Nicoladis, 2014).
-While these studies have mostly relied on samples of bilingual children learning two relatively distant languages, as it is the case of English and Spanish, it is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals.
-What role could linguistic distance play during early vocabulary growth?
-
-For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms--one in each language.
-Depending on the linguistic distance between both languages, the two sets of words may overlap in varying degrees.
-Particularly, when both languages are typologically close, like Spanish and Catalan (both Roman languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two linguistically distant languages, like Spanish and English (one Roman, the other Germanic).
-For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear \emph{puerta} and \emph{porta} (cognates), whereas a Spanish-English bilingual might hear \emph{puerta} and \emph{door} (non-cognates).
-It could be the case that mapping two phonologically similar labels (cognates like \emph{puerta}-\emph{porta}) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like \emph{puerta} and \emph{door}).
-If cognates are easier to acquire than non-cognates, bilinguals learning a pair of languages that share a high proportion of cognates should benefit more often from this facilitation effect than those learning a pair of languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.
-
-Floccia et al.
-(2018) provided evidence in line with this claim.
-The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
-The additional language was one a pool of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
-The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English.
-Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions.
-The Levenshtein distance is a metric that computes the edit distance between two strings by counting the smallest number of insertions, deletions and substitutions one of the strings has to go through to become identical to the other (Levenshtein \& others, 1966).
-The resulting scores were then divided by the length of the longest string to bound the similarity scores between 0 and 1, and then entered this variable as a predictor as they modelled participants' vocabulary sizes.
-Among other findings, the authors reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
-For example, English-Dutch bilinguals (22.14\% phonological similarity), were able to produce more Dutch words than English-Mandarin bilinguals (1.97\% phonological similarity) were able to produce in Mandarin.
-
-Floccia et al.~pointed to \emph{parallel activation} as the main mechanism underpinning their results.
-The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension.
-This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production (Costa, Caramazza, \& Sebastian-Galles, 2000; Hoshino \& Kroll, 2008) or comprehension (Thierry \& Wu, 2007).
-One of the clearest examples of parallel activation was provided by Costa et al.
-(2000).
-In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish.
-In half of the trials, the object labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre}, translations of \emph{tree}), whereas in the other half of the trials labels were non-cognates (\emph{mesa}-\emph{taula}, translations of \emph{table}).
-Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
-Importantly, Spanish monolinguals did not show this effect.
-These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-Several recent studies have also provided similar evidence in comprehension in children (Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013; e.g., Von Holzen \& Mani, 2012).
-Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes.
-A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.
-
-In this study, we hypothesise that cognates are acquired earlier in age than non-cognates, which might explain why children learning languages that share many cognates show larger vocabulary sizes.
-Evidence supporting this claim is scarce.
-To the best of our knowledge, only two studies have provided direct data on this issue.
-First, Bosch and Ramon-Casas (2014) used vocabulary parental reports (152 lexical items) from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates.
-Second, Schelletter (2002) reported a longitudinal single case of one English-German bilingual who produced cognates earlier than non-cognates, on average.
-The low sample size in these studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth.
-Floccia et al.'s estimates are statically more reliable given their (much larger) sample size, but their study was not aimed at testing the effect of cognateness on age of acquisition directly.
-In their discussion the authors state the following (pp.~70):
-
-\begin{quote}
-``This finding also provides support to the proposal that the cognate advantage is due to cognates being acquired before non-cognates in early childhood (Costa et al., 2016), leading to an ease of processing later in life.''
-\end{quote}
-
-We identify two main reasons why an earlier age of acquisition for cognates than for non-cognates is an unwarranted conclusion from Floccia et al.'s results.
-First, the response variable used was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists.
-By aggregating the responses from all items into a single datum per child, information about the acquisition status of cognates vs.~non-cognates was lost.
-Second, all participants were aged \textasciitilde24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.
-In summary, the current evidence supporting an earlier age of acquisition of cognates vs.~non-cognates presents some limitations that prevents drawing sound conclusions about this effect.
-
-Our second hypothesis detaches from the fact that, for parallel activation to take place, and for cognateness to facilitate the acquisition of a word form in a given language, the child must be familiar with the (phonologically similar) translation of the word, in the other language.
-If the child is not previously familiar with one of the word forms, the acquisition of its translation in language cannot be facilitated by cognateness.
-This implies that, if cognateness facilitates the acquisition of a word form, it can only do so after the child has acquired one of the word forms of the translation pair.
-Given that children are more likely to acquire words from languages to which they are exposed more often (Cattani et al., 2014; David \& Wei, 2008; Thordardottir, Rothenberg, Rivard, \& Naves, 2006), the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness.
-Therefore, we hypothesised that the effect of cognateness for words in the language of lower exposure would be larger than in the language of higher exposure.
-
-In line with this hypothesis, the vocabulary size associated with linguistic similarity that Floccia et al.~reported was larger in the additional language vocabulary than in English vocabulary.
-Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language.
-Therefore, participants may have, on average, learnt the English word-form of translation equivalents earlier than the word-form in the additional language.
-If this is the case, then the acquisition of English words by English-dominant participants would rarely benefit from their cognate status (the other word-form is not available yet), while the acquisition of words in the additional language would benefit from their phonological similarity with the (available) English form.
-To test this hypothesis, we conditioned the effect of cognateness to the relative degree of exposure to the language.
-
-In summary, we investigated the role of cognateness in bilingual word acquisition (conditional to the degree of relative exposure to the language) as possible mechanism to account for the facilitation of language similarity on bilingual children's vocabulary size.
-We hypothesised that cognate words would be acquired earlier than non-cognate words, and that this effect should be larger for words in the language with lower exposure.
-Using an online vocabulary checklist--carefully designed specifically for this study--we collected data from a sample of children aged 19 to 33 months learning Catalan and/or Spanish, with varying degrees of exposure to each language.
-We then modelled the probability of participants being reported by their parents to understand and/or say each word in the checklist, conditional to its cognate status in Catalan and Spanish and participants' degree of exposure to the corresponding language, while adjusting for participants' age and the item's lexical frequency.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-Data and materials are available at OSF (\url{https://osf.io/hy984/}), and code is available at GitHub (\url{https://github.com/gongcastro/trajectories}).
-
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
-
-We collected data from 289 children from the Metropolitan Area of Barcelona between 28th October, 2019 and 24th July, 2021.
-All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.
-Families were contacted by e-mail or phone if their child were aged between 10 and 34 months, and had not previously reported to be exposed more than 10\% of the time to a language other than Spanish or Catalan.
-Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate.
-Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant languag (i.e., their degree of bilingualism).
-
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{ccccccccccccccccc}
+\begin{abstract}
+Bilingual infants show equivalent developmental trajectories of lexical
+acquisition compared to their monolingual peers. This is remarkable,
+given the increased complexity of their linguistic input. Recent studies
+suggested that bilingual vocabulary growth is boosted by the number of
+cognates shared by the pair of languages being learned, and that this
+facilitation effect is driven by a stronger parallel activation of
+cognates during linguistic exposure, compared to non-cognates. The
+mechanisms behind this facilitation are still unclear. In this study, we
+capitalise on accumulator models of language acquisition to propose an
+account of bilingual lexical acquisition in which parallel activation
+increases the rate at which children accumulate learning instances for
+words in both languages, even in fully monolingual situations. Under
+this hypothesis, we predicted a stronger cognate facilitation for words
+to which children were exposed less frequently (low-exposure words), as
+they are co-activated by their translation more often than high-exposure
+words do. We developed an extensive online vocabulary checklist, the
+Barcelona Vocabulary questionnaire (BVQ) to collect vocabulary data from
+366 Catalan-Spanish bilingual toddlers aged 12 to 32 months. We then
+used Bayesian explanatory item response theory to model the acquisition
+trajectories of 302 Catalan/Spanish translation equivalents. We found an
+interaction between exposure and cognateness, which pointed to
+cognateness facilitating the aquisition of low-exposure words, but not
+of mean exposure or high-exposure words. Overall, our findings suggest
+that cognateness plays a key role in bilingual lexical acquisition, and
+provides evidence for a frequency-mediated facilitation effect driven by
+parallel activation.
+\end{abstract}
+{\bfseries \emph Keywords}
+\def\sep{\textbullet\ }
+cognate \sep word
+acquisition \sep vocabulary \sep bilingualism \sep item response
+theory \sep 
+bayesian
+
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, frame hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+
+\hypertarget{introductionauthor-note}{%
+\section[Introduction]{\texorpdfstring{Introduction\footnote{The authors
+  declare no conflicts of interest with regard to the funding source of
+  this study. This research was supported by grants from the Spanish
+  Ministerio de Ciencia, Innovación y Universidades
+  (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government
+  {[}ICREA (Catalan Institution for Research and Advanced Studies)
+  Academia 2019 award{]}. Gonzalo Garcia-Castro was supported by a
+  fellowship of the Spanish Ministerio de Ciencia, Innovación y
+  Universidades (FPI 2019). We are grateful to Chiara Santolin, Ege E.
+  Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest
+  of the Speech Acquisition and Perception research group for their
+  helpful feedback. We thank Xavier Mayoral, Silvia Blanch, and Cristina
+  Cuadrado for their technical support, and Cristina Dominguez and Katia
+  Pistrin for their efforts in recruiting infants. We also thank all
+  families and infants who participated in the experiments.}}{Introduction}}\label{introductionauthor-note}}
+
+The foundations of word learning are in place early in age. At six
+months, infants start directing their gaze to some objects when hearing
+their labels (\protect\hyperlink{ref-bergelson2012months}{Bergelson \&
+Swingley, 2012}, \protect\hyperlink{ref-bergelson2015early}{2015};
+\protect\hyperlink{ref-tincoff1999beginnings}{Tincoff \& Jusczyk,
+1999}), and shortly after caregivers start reporting some words as
+acquired by their infant in vocabulary checklists (e.g.,
+\protect\hyperlink{ref-fenson2007macarthurbates}{Fenson et al., 2007};
+\protect\hyperlink{ref-samuelson2021precision}{Samuelson, 2021}). Most
+research on early word acquisition relies extensively on data from
+monolingual children, and is oblivious to the fact that a substantial
+proportion of the world population acquires more than one language from
+early ages (\protect\hyperlink{ref-grosjean2021extent}{Grosjean, 2021}).
+Previous work on bilingual vocabulary acquisition pointed to bilingual
+toddlers knowing, on average, less words in each of their languages than
+their monolinguals peers, and to both groups knowing a similar number of
+words---if not more words---when the bilinguals' two languages are
+pooled together. Hoff et al.
+(\protect\hyperlink{ref-hoff2012dual}{2012}) found that English-Spanish
+bilingual toddlers in South Florida (United States) knew less words in
+English than monolinguals did, but both groups knew a similar total
+amount of words when both English and Spanish vocabularies were counted
+together. Other studies have provided converging evidence that
+bilinguals know a similar or even larger number of words than
+monolinguals when the two languages are aggregated
+(\protect\hyperlink{ref-gonzalez-barrero2020bilingual}{Gonzalez-Barrero
+et al., 2020}; \protect\hyperlink{ref-oller2002language}{Oller \&
+Eilers, 2002}; \protect\hyperlink{ref-patterson2004comparing}{Patterson,
+2004}; \protect\hyperlink{ref-patterson2004bilingual}{Patterson \&
+Pearson, 2004}; \protect\hyperlink{ref-pearson1993lexical}{Pearson et
+al., 1993}; \protect\hyperlink{ref-pearson1994patterns}{Pearson \&
+Fernández, 1994}; \protect\hyperlink{ref-petitto2001bilingual}{Petitto
+et al., 2001}; \protect\hyperlink{ref-smithson2014bilingualism}{Smithson
+et al., 2014}). A more detailed analysis of the words in bilinguals'
+lexicon show some interesting patterns.
+
+One important observation of studies on bilinguals' early vocabulary
+acquisition is that cognate words are easier to acquire than non-cognate
+words. Cognate words are translation equivalents that are phonologically
+similar (or share some type of form-similarity). For instance, the
+Spanish translation equivalent of \emph{cat} is \emph{gato}, a cognate
+word; the translation equivalent of \emph{dog} is \emph{perro}, a
+non-cognate word. The differences in the percentages of cognate and
+non-cognate words between two languages is related to historic reasons:
+languages typologically close (like Dutch and English or Italian and
+Spanish) share more cognates than languages typologically distant (like
+English and Chinese, or Urdu and Spanish). The conclusion that cognate
+words are easier to learn is based on two types of evidence: studies
+investigating vocabulary sizes in children learning language pairs with
+different percentages of cognates (that is, differing in their
+typological distance) and studies comparing the number of cognate and
+non-cognate words children know in a specific language pair.
+
+Floccia et al. (\protect\hyperlink{ref-floccia2018introduction}{2018})
+published an impressive study comparing vocabularies of children
+learning several language pairs differing in their percentage of
+cognates. The authors collected vocabulary data on word comprehension
+and production from 372 24-month-old bilingual toddlers living in the
+United Kingdom who were learning English and an additional language. The
+additional language was one of 13 typologically diverse languages:
+Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu,
+Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The
+authors calculated the average phonological overlap between the words in
+each of these additional languages and their translation equivalents in
+English, which was taken as a proxy of the degree of cognateness between
+each pair of languages. Floccia and co-workers reported an increase in
+vocabulary size in the additional language (i.e., not English)
+associated with an increase in the average phonological similarity
+between the translation equivalents of each language pair. For example,
+English-Dutch bilinguals (languages with a high phonological overlap),
+were able to produce more Dutch words than English-Mandarin bilinguals
+(languages with a low phonological overlap) were able to produce in
+Mandarin. Blom et al.
+(\protect\hyperlink{ref-blom2020crosslanguage}{2020}), Bosma et al.
+(\protect\hyperlink{ref-bosma2019longitudinal}{2019}), and Gampe et al.
+(\protect\hyperlink{ref-gampe2021does}{2021}) reported similar results,
+providing converging evidence of a facilitatory effect of language
+distance on vocabulary size.
+
+A second set of studies concluded that cognates are overrepresented in
+bilinguals' early lexicon. Bosch \& Ramon-Casas
+(\protect\hyperlink{ref-bosch2014first}{2014}) collected parental
+reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged
+18 months and found that cognates represented a larger proportion of
+participant's vocabulary than non-cognates. Schelletter
+(\protect\hyperlink{ref-schelletter2002effect}{2002}) provided
+converging evidence from a longitudinal single-case study, in which an
+English-German bilingual child produced cognates earlier than
+non-cognates, on average. Mitchell et al.
+(\protect\hyperlink{ref-mitchell2022cognates}{2022}) addressed this
+issue in a larger sample longitudinal study. These authors collected
+expressive vocabulary data of 47 16- to 30-month-old French-English
+bilinguals living in Canada, in both languages. They created two lists
+of translation equivalents: one made of 131 cognates, and one made of
+406 non-cognates. The proportion of words that children were reportedly
+able to produce was higher in the cognate lists than in the non-cognate
+list across ages, even when both lists were matched by semantic category
+(furniture, animals, food were similarity represented in both lists) and
+age-of-acquisition norms (an index of word difficulty). Taken together,
+the results of these two lines of research provide a strong support to
+the hypothesis that phonological similarity (as reflected in
+cognateness) plays a facilitation role in bilingual word acquisition.
+
+Parallel activation of bilinguals' lexicons has been proposed as the
+underlying mechanism for such facilitatory effect (e.g.,
+\protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
+\protect\hyperlink{ref-mitchell2022cognates}{Mitchell et al., 2022}).
+The parallel activation hypothesis stems from the language non-selective
+account of lexical access, which suggests that bilinguals activate both
+languages simultaneously during language processing, even in fully
+monolingual contexts Evidence with adult bilinguals supporting the
+language-non selective account of lexical access has been reported for
+language comprehension and production, across the auditory and visual
+(reading and signing) modalities
+(\protect\hyperlink{ref-gimeno-martinez2021crosslinguistic}{Gimeno-Martínez
+et al., 2021}; \protect\hyperlink{ref-vanhell2008sentence}{Hell \&
+Groot, 2008}; \protect\hyperlink{ref-hoshino2008cognate}{Hoshino \&
+Kroll, 2008}; \protect\hyperlink{ref-morford2011when}{Morford et al.,
+2011}; \protect\hyperlink{ref-schwartz2007reading}{Schwartz et al.,
+2007}; \protect\hyperlink{ref-spivey1999cross}{Spivey \& Marian, 1999};
+see \protect\hyperlink{ref-kroll2017bilingual}{Kroll \& Ma, 2017} for
+review). One of the clearest pieces of evidence of parallel activation
+was provided by Costa et al.
+(\protect\hyperlink{ref-costa2000cognate}{2000}). In this study, Spanish
+monolinguals and Catalan-Spanish bilingual adults were asked to name
+pictures of common objects in Spanish. In half of the trials, the object
+labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre},
+translations of \emph{tree}), whereas in the other half of the trial
+labels were non-cognates (\emph{mesa}-\emph{taula}, translations of
+\emph{table}), obviously, such distinction was only relevant for
+bilinguals. Bilinguals named cognate pictures faster than non-cognate
+pictures, even after adjusting for the lexical frequency of the items.
+In contrast, Spanish monolinguals, who were unfamiliar with the Catalan
+translations of the Spanish words they uttered, showed equivalent naming
+times for the two types of stimuli. The authors interpreted the
+difference between cognates and non-cognates in bilinguals as reflecting
+the additional phonological activation that cognate words would receive
+from their translation equivalents (due to language non-selective
+activation of bilinguals' lexicons). These results showed that
+bilinguals' Catalan phonology was activated during the production of
+Spanish words, facilitating the naming of cognate pictures. Evidence of
+parallel activation has been reported in bilingual toddlers and children
+(\protect\hyperlink{ref-bosma2020cognate}{Bosma \& Nota, 2020};
+\protect\hyperlink{ref-poarch2012crosslanguage}{Poarch \& Hell, 2012};
+\protect\hyperlink{ref-poulin-dubois2013lexical}{Poulin-Dubois et al.,
+2013}; \protect\hyperlink{ref-schroter2016orthographic}{Schröter \&
+Schroeder, 2016}; \protect\hyperlink{ref-vonholzen2019impact}{Von Holzen
+et al., 2019}; \protect\hyperlink{ref-vonholzen2012language}{Von Holzen
+\& Mani, 2012}).
+
+Although there is a consensus on the role of parallel activation in
+bilinguals' lexical processing and acquisition, previous studies do not
+address its influence on the learning trajectories of words. Results are
+aggregated across words and provide no information about the specific
+dynamics of how parallel activation influences word learning. This is
+the goal of the present research.
+
+We propose an account in which a learning instance for a word may also
+represent a learning instance for its translation equivalent, to the
+extent that such translation equivalent is co-activated\footnote{We use
+  the term \emph{learning instance} in the fashion of accumulator models
+  of language acquisition; as an exposure to a word-form that
+  constitutes an opportunity for the child to accumulate information
+  about the word. We do not make strong assumptions about whether a
+  learning instance is a discrete or a continuous unit of accumulation
+  of information. We rather consider that a learning instance of a word
+  is so to the degree of the strength of activation of its phonological
+  representation. This activation may be the result of the infant being
+  exposed to the word-form, or the result of activation spreading
+  through phonological or semantic links across lexicl representation,
+  as in the case of parallel activation.}. The strength of this
+co-activation is proportional to the phonological similarity between the
+two translation equivalents; given that cognates share higher
+phonological similarity than non-cognates, the former should be
+co-activated more frequently than the latter. This should lead to a
+faster accumulation of learning instances for cognates, compared to
+non-cognates. Parallel activation would allow bilingual children to
+accumulate learning instances for words in both languages even during
+fully monolingual situations, but the impact of this mechanism would be
+asymmetric across languages: words from the lower-exposure language
+would receive additional activation through parallel activation more
+often than words from the higher-exposure language. Therefore, the
+acquisition of words from the lower-exposure language would benefit more
+strongly from their cognate status than words from the higher-exposure
+language.
+
+Consider the example of the Catalan-Spanish cognate translation
+equivalent /\textipa{"gat}-\textipa{"ga.to}/ {[}cat{]}, that are
+phonologically very similar. When the child is exposed to
+/\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in
+parallel. Therefore, this exposure will count as a learning instance for
+both co-activated forms. However, for the case of the non-cognate
+translation equivalent /\textipa{"gos}/, the low phonological similarity
+will result in a weak activation of /\textipa{"pe.ro}/ resulting in such
+exposure counting as a learning instance for /\textipa{"gos}/ (which the
+child was exposed to), but not for /\textipa{"pe.ro}/. While
+/\textipa{"gat}-\textipa{"ga.to}/ will benefit from parallel activation,
+/\textipa{"gos}-\textipa{"pe.ro}/ will not. If the child receives
+linguistic input from one of the languages more often than from the
+other, this effect might affect each form of the cognate translation
+equivalent differently. For instance, if the child receives a larger
+amount of Catalan input than Spanish input, they will encounter the
+Catalan form /\textipa{"gat}/ more frequently than the Spanish form
+/\textipa{"ga.to}/. Through parallel activation, /\textipa{"gat}/ will
+activate /\textipa{"ga.to}/ more often than vice versa. Ultimately,
+/\textipa{"ga.to}/ will benefit more strongly from its cognate status
+than /\textipa{"gat}/, as it receives additional learning instances from
+its translation equivalent more often than /\textipa{"gat}/.
+
+To test these predictions, we collected vocabulary data on production
+and comprehension from a large sample of bilingual Catalan-Spanish
+children using an online vocabulary checklist designed for the present
+study. We adopted a Bayesian explanatory item response theory approach
+(IRT, see \protect\hyperlink{ref-kachergis2022standard}{Kachergis et
+al., 2022}, for a similar approach) to model the probability of
+acquisition of 604 Catalan and Spanish nouns included in the vocabulary
+checklist. Words were considered as acquired if caregivers reported such
+word to be understood (comprehension) or understood and said
+(production) by their child. We estimated the impact of several
+predictors of interest on the probability of acquisition, including
+participants' age and rate of exposure to the word-form, and the cognate
+status of the word-form. As described in the methods section, rate of
+exposure was a composite measure taking into account participant'
+language exposure and word's lexical frequency. We predicted an
+interaction between cognate status and word exposure rate in which the
+probability of comprehension is higher for low-exposure cognate words,
+but not for high-exposure words cognate words.
+
+\hypertarget{sec-methods}{%
+\section{Methods}\label{sec-methods}}
+
+All materials, data, and reproducible code can be found at the OSF
+(\url{https://osf.io/hy984/}) and GitHub
+(\url{https://github.com/gongcastro/cognate-beginnings}) repositories.
+This study was conducted according to guidelines laid down in the
+Declaration of Helsinki, and was approved by the Drug Research Ethical
+Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.
+
+\hypertarget{sec-questionnaire}{%
+\subsection{Questionnaire}\label{sec-questionnaire}}
+
+To collect vocabulary data from participants, we created an \emph{ad
+hoc} questionnaire: the Barcelona Vocabulary Questionnaire (BVQ)
+(\protect\hyperlink{ref-garcia-castro2023bvq}{Garcia-Castro et al.,
+2023}). This questionnaire was inspired by the MacArthur-Bates
+Communicative Development Inventory
+(\protect\hyperlink{ref-fenson2007macarthurbates}{Fenson et al., 2007})
+and its adaptations to other languages, and was implemented on-line
+using the formr platform (\protect\hyperlink{ref-arslan2020formr}{Arslan
+et al., 2020}). This questionnaire is structured in three blocks: (1) a
+language exposure questionnaire, (2) a demographic survey, and (3) two
+vocabulary checklists. Vocabulary checklists followed a similar
+structure as the Oxford Communicative Developmental Inventory (OCDI)
+(\protect\hyperlink{ref-hamilton2000infant}{Hamilton et al., 2000}) and
+consisted in two lists of words: one in Catalan and one in Spanish. Both
+lists included items from a diverse sample of 26 semantic/functional
+categories. The Catalan checklist contained 793 items and the Spanish
+checklist contained 797. Items in one language were translation
+equivalents of the items in the other language (e.g., the same
+participant responded to both \emph{gos} and \emph{perro}, Catalan and
+Spanish for \emph{dog}), roughly following a one-to-one
+mapping\footnote{Although for some of the included words in Catalan did
+  not have a clear translation or had more than one possible translation
+  in Spanish, and vice versa, therefore the unequal number of words
+  included in the Catalan and Spanish lists} (see Table~\ref{tbl-items}
+for a summary of the questionnaire items).
+
+\hypertarget{tbl-items}{}
+\begin{longtable}{l|rrrrc}
+\caption{\label{tbl-items}Summary of the items included in the final analyses. }\tabularnewline
+
 \toprule
-& \multicolumn{16}{c}{Number of participants by age (months)} \\ 
- \cmidrule(lr){2-17}
-\textbf{L2 DoE} & 19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 & 28 & 29 & 30 & 31 & 32 & 33 & 34 \\ 
+\multicolumn{1}{l}{} & List A & List B & List C & List D & Examples \\ 
 \midrule
-0-10\% & 4 & 6 & 10 & 11 & 7 & 5 & 8 & 12 & 13 & 2 & 9 & 2 & 1 & 2 & 2 & - \\ 
-10-20\% & - & 6 & 11 & 3 & 5 & 2 & 6 & 4 & 7 & 2 & 6 & - & 2 & 1 & - & - \\ 
-20-30\% & - & 1 & 13 & 9 & 7 & 5 & 12 & 5 & 7 & 5 & 1 & - & 2 & - & - & 1 \\ 
-30-40\% & 1 & 4 & 12 & 6 & 2 & 6 & 6 & 4 & 9 & 2 & 2 & 4 & 1 & - & - & - \\ 
-40-50\% & 3 & 2 & 6 & 7 & 3 & 1 & 5 & 1 & - & 2 & 2 & 1 & - & 1 & 1 & - \\ 
+\multicolumn{6}{l}{Semantic category} \\ 
+\midrule
+Household items & 31 & 26 & 30 & 25 & clock, video \\ 
+Food and drink & 29 & 26 & 23 & 27 & sausage, yogurt \\ 
+Animals & 26 & 23 & 19 & 25 & panther, tiger \\ 
+Outside & 14 & 13 & 13 & 15 & farm, stone \\ 
+Body parts & 14 & 12 & 11 & 11 & face, finger \\ 
+Toys & 11 & 11 & 12 & 13 & piano, racket \\ 
+Clothes & 12 & 12 & 10 & 10 & zipper, sandal \\ 
+Vehicles & 9 & 10 & 11 & 10 & helicopter, tractor \\ 
+Colours & 6 & 6 & 6 & 6 & red, green \\ 
+People & 7 & 4 & 6 & 6 & police, babysitter \\ 
+Furniture and rooms & 4 & 4 & 4 & 4 & corridor, terrace \\ 
+Time & 2 & 2 & 2 & 2 & day, night \\ 
+Adventures & 1 & 1 & 1 & 1 & witch \\ 
+Parts of things & 1 & 1 & 1 & 1 & wheel \\ 
+\midrule 
+Total & 167 & 151 & 149 & 156 & — \\ 
 \bottomrule
@@
 
-We use the highest educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES), which families self-reported in the questionnaire by filling two items asking for the educational attainment of each parent or caretaker, with the following available options: \emph{No education}, \emph{Primary}, \emph{Secondary}, \emph{Complementary}, \emph{Vocational}, and \emph{University}, in line with the current educational system in Spain.
-Most families reported university studies (244, 0.84), followed by families were the highest educational attainment were vocational studies (51, 18\%), complementary studies (5, 2\%), secondary education (2, 1\%, 0\%), no formal education (2, 1\% and primary education (1, \textless1\%).
-Highest educational attainment was similar across participants with different DoE to a second language, and therefore, differences in comprehension or production between participants with different language profiles are unlikely to stem from a difference in SES (Fernald, Marchman, \& Weisleder, 2013) (see Table 2).
-
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{crrrrrr}
+For each word included in the vocabulary checklists, we asked parents to
+report whether their child was able to understand it, understand
+\emph{and} say it, or did not understand or say it (checked out by
+default). Given the large number of words in the vocabulary checklists,
+we created four different subsets of the complete list of items. Each
+subset contained a random but representative sub-sample of the items
+from the complete list (see Table~\ref{tbl-items}). Semantic/functional
+categories with less than 16 items---thus resulting in less than four
+items after dividing it in four lists---were not divided in the short
+version of the questionnaire: all of their items were included in the
+four lists. Items that were part of the trial lists of some ongoing
+experiments in the lab were also included in all versions. The resulting
+reduced list contained between 343 and 349 Catalan words, and between
+349 and 371 Spanish words. Participants were randomly allocated into one
+of the four subsets.
+
+To compute predictors of interest, we manually generated a broad
+phonological transcription of every word included in the vocabulary
+checklists in X-SAMPA format
+(\protect\hyperlink{ref-wells1995computercoding}{Wells, 1995}). Catalan
+word-forms were transcribed to Central Catalan phonology, and Spanish
+word-forms were transcribed to Castilian Spanish phonology.
+
+\hypertarget{sec-participants}{%
+\subsection{Participants}\label{sec-participants}}
+
+We collected 436 responses to the questionnaire from 366 distinct
+children from the Metropolitan Area of Barcelona between the 30th of
+March, 2020 and the 31th of October, 2022: 312 of those participants
+participated once, 42 twice, 8 three times, and 4 four times. Recurrent
+participants provided responses with a minimum of 25 days between
+responses, and a maximum of 527, and were always allocated to the same
+questionnaire list (A, B, C, or D). Participants were part of the
+database of the Laboratori de Recerca en Infància (Universitat Pompeu
+Fabra) and were contacted by e-mail or phone if their child was aged
+between 12 and 32 months, and had not been reported to be exposed more
+than 10\% of the time to a language other than Spanish or Catalan (see
+Table~\ref{tbl-participants}) for a more detailed description of the
+sample). In total, 70 participants (16.06\%) participants were reported
+to be exposed to a third language other than Catalan and Spanish. All
+families provided informed consent before participating. Upon consent,
+families were sent a link to the questionnaire via e-mail, which they
+filled from a computer, laptop, or mobile device. Filling the
+questionnaire took 30 minutes approximately. After completion, families
+were rewarded with a token of appreciation.
+
+\hypertarget{tbl-participants}{}
+\setlength{\LTpost}{0mm}
+\begin{longtable}{rrrrrrr}
+\caption{\label{tbl-participants}Participant sample size by age and degree of exposure to Catalan. }\tabularnewline
+
 \toprule
-& \multicolumn{6}{c}{Educational attainment} \\ 
+ & \multicolumn{6}{c}{Age (months)} \\ 
 \cmidrule(lr){2-7}
-\textbf{L2 DoE} & No education & Primary & Secondary & Complementary & Vocational & University \\ 
+Catalan exposure\textsuperscript{\textit{1}} & [10,14] & (14,18] & (18,22] & (22,26] & (26,30] & (30,34] \\ 
 \midrule
-0-10\% & $3.19\%$ & $3.19\%$ & $3.19\%$ & $7.45\%$ & $17.02\%$ & $81.91\%$ \\ 
-10-20\% & $1.82\%$ & $1.82\%$ & $1.82\%$ & $1.82\%$ & $20.00\%$ & $81.82\%$ \\ 
-20-30\% & $2.94\%$ & $1.47\%$ & $2.94\%$ & $1.47\%$ & $17.65\%$ & $80.88\%$ \\ 
-30-40\% & $0.00\%$ & $0.00\%$ & $0.00\%$ & $0.00\%$ & $23.73\%$ & $76.27\%$ \\ 
-40-50\% & $5.71\%$ & $5.71\%$ & $5.71\%$ & $5.71\%$ & $11.43\%$ & $80.00\%$ \\ 
+75-100\% & 18 & 23 & 36 & 38 & 20 & 7 \\ 
+50-75\% & 8 & 13 & 30 & 41 & 18 & 1 \\ 
+25-50\% & 10 & 17 & 45 & 29 & 17 & - \\ 
+0-25\% & 7 & 11 & 21 & 17 & 8 & 1 \\ 
 \bottomrule
 \end{longtable}
-
-\hypertarget{questionnaire}{%
-\subsection{Questionnaire}\label{questionnaire}}
-
-The questionnaire was implemented on-line using the formR platform (Arslan, Walther, \& Tata, 2020), and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (Hamilton, Plunkett, \& Schafer, 2000) and consisted in two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained 777 items (196 cognates, 581 non-cognates) and the Spanish inventory contained 782 (197 cognates, 585 non-cognates).
-Items in one language were translation equivalents of the items in the other language (e.g., whenever \emph{gos} {[}dog{]} was included in the Catalan inventory, the word \emph{perro} was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan \emph{acabar} {[}\emph{to finish}{]} and Spanish \emph{acabar} and \emph{terminar}), or merged them into a single items (e.g., Spanish \emph{mono} {[}\emph{monkey}{]} and Catalan \emph{mono/mico}.
-We included items from a diverse sample of 26 semantic/functional categories (see Appendix 1).
-For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand \emph{and} say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\textasciitilde400 translation equivalents, \textasciitilde400 items in Catalan, \textasciitilde400 items in Spanish).
-These last families were randomly allocated into one of four different subsets of the complete list of items.
-These lists were carefully designed so that each contained a representative subsample of the items from the complete list.
-Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., \emph{barrita de cereales} {[}cereal bar{]}) and items that included more than one word-form (e.g., \emph{mono / mico}).
-Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
-
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{lcrrlcrrlcrrl}
+\begin{minipage}{\linewidth}
+\textsuperscript{\textit{1}}For participants exposed to Catalan and Spanish exclusively this proportion is complementary to the degree of exposue to Spanish.
+For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10\%.\\
+\end{minipage}
+
+We used the highest self-reported educational attainment of parents or
+caregivers as a proxy of participants' socio-economic status (SES). This
+information was provided by each parent or caregiver by selecting one of
+six possible alternatives in line with the current educational system in
+Spain: \emph{sense escolaritzar/sin escolarizar} {[}no education{]},
+\emph{educació primària/educación primaria} {[}primary school{]},
+\emph{educació secundària/educación secundaria} {[}secondary school{]},
+\emph{batxillerat/bachillerato} {[}complementary studies/high school{]},
+\emph{cicles formatius/ciclos formativos} {[}vocational training{]}, and
+\emph{educació universitària/educación universitaria} {[}university
+degree{]}. Most families reported university studies (356, 82\%),
+followed by families were the highest educational attainment were
+vocational studies (59, 14\%), secondary education (8, 2\%),
+complementary studies (6, 1\%), primary education (1, \textless1\%), and
+no formal education (2, \textless1\%).
+
+\hypertarget{sec-analysis}{%
+\subsection{Data analysis}\label{sec-analysis}}
+
+We collected responses to 1,590 items. We restricted the analyses to
+responses to nouns (628 items corresponding to other grammatical classes
+were excluded, see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
+et al., 2020} for a similar approach). We then excluded items with
+missing lexical frequency scores (\emph{n} = 269), items that included
+more than one lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} =
+48), multi-word items or phrases (e.g., \emph{barrita de cereales}
+{[}cereal bar{]}, \emph{n} = 9). Finally, we removed items without a
+translation in the other language (\emph{n} = 32). This resulted in a
+final list of 604 items, corresponding to words 302 Catalan words and
+their 302 Spanish translations (302 translation equivalents). After
+collecting participants' responses, the final dataset consisted of
+138,078 observations, each corresponding to a single response of one
+participant to one item. Each translation equivalent received a median
+of 234 responses (\emph{Min} = 106, \emph{Max} = 872) from participants,
+both languages pooled together.
+
+We modelled the probability of participants answering each response
+category (\emph{No} \textless{} \emph{Understands} \textless{}
+\emph{Understands and Says}) using a Bayesian, multilevel, ordinal
+regression model. This model allowed us to estimate both item and
+participant word-acquisition trajectories, while estimating the effect
+of our variables of interest: \emph{Age} (number of months elapsed
+between participants' birth date and questionnaire completion),
+\emph{Length} (number of phonemes in the X-SAMPA phonological
+transcription of the word-form), \emph{Exposure} (a language
+exposure-weighted lexical frequency), and \emph{Cognateness} (defined as
+the phonological similarity between translation equivalentsFor each
+translation equivalent). We added these variables as main effects,
+together with the two-way and three-way interactions between \emph{Age},
+\emph{Exposure}, and \emph{Cognateness}.
+
+We developed the \emph{Exposure} predictor to account for the fact that
+bilinguals' exposure to a given word-form is not only a function of the
+word-forms lexical frequency, but also of the quantitative input they
+receive from the language such word-form belongs to, we expressed
+lexical frequencies as the product between both variables. First, we
+extracted the child-directed lexical frequency of each word from the
+CHILDES database
+(\protect\hyperlink{ref-macwhinney2000childes}{MacWhinney, 2000};
+\protect\hyperlink{ref-sanchez2019childesdb}{Sanchez et al., 2019}).
+Using the corresponding lexical frequencies directly from Catalan and
+Spanish was not possible due to the low number of Catalan participants
+and tokens available in their corresponding CHILDES corpora, so they
+were extracted from the English corpora instead. We then mapped the
+lexical frequencies of the English words to their Catalan and Spanish
+translations (see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
+et al., 2020} for a similar approach), and transformed them to Zipf
+scores (\protect\hyperlink{ref-vanheuven2014subtlexuk}{Van Heuven et
+al., 2014}; \protect\hyperlink{ref-zipf1949human}{Zipf, 1949}). We then
+multiplied the resulting lexical frequencies by the reported degree of
+exposure of the child to Catalan or Spanish. For instance, for a child
+who is reportedly exposure to Catalan 80\% of the time, and to Spanish
+20\% of the time, the expected exposure score to the Catalan word
+\emph{cavall} {[}\emph{horse}{]}---with a lexical frequency of ---would
+be , while that of its translation to Spanish \emph{caballo} would be .
+
+We defined \emph{Cognateness} as the phonological similarity between
+each word-form and its translation. For each translation equivalent, we
+used the \texttt{stringdist}
+(\protect\hyperlink{ref-vanderloo2014stringdist}{Loo, 2014}) R package
+to calculate the Levenshtein distance between the Catalan and the
+Spanish phonological transcriptions of the word-forms. The Levenshtein
+distance measures the number of editions (insertions, deletions, or
+substitutions) that one string of phonemes/characters must go through to
+become identical to the other
+(\protect\hyperlink{ref-levenshtein1966binary}{Levenshtein, 1966}). We
+divided the Levnehstein distance of each translation equivalent by the
+length of the longest word-form to correct for word length (longer
+strings are likely to show a larger number of mismatches). Finally, we
+subtracted the result from one so that it could be interpreted in terms
+of phonological similarity, instead of phonological distance. This led
+to a distance metric that ranged from zero to one, where zero indicates
+that both word-forms are completely different (e.g., /'taw.l5-'me.sa/,
+\emph{table}), and one indicates that the two word-forms are identical
+(e.g., /'mar-'mar/, \emph{sea}) (see
+\protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
+\protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020};
+\protect\hyperlink{ref-heeringa2003norwegian}{Heeringa \& Gooskens,
+2003}; \protect\hyperlink{ref-laing2022phonological}{Laing, 2022} for
+similar approaches;
+\protect\hyperlink{ref-schepens2012distributions}{Schepens et al.,
+2012}).
+
+Predictors were standardised before entering the model by subtracting
+the mean of the predictor from each value and dividing the result by the
+standard deviation of the predictor. Participant-level and item-level
+random intercepts and slopes were included where appropriate, according
+to the structure of the data
+(\protect\hyperlink{ref-barr2013random}{Barr et al., 2013}). We
+specified a weakly informative prior around the parameters of the model.
+Equation~\ref{eq-linear} shows a detailed description of the model.
+
+\begin{equation}\protect\hypertarget{eq-linear}{}{
+\begin{aligned}
+&\textbf{Response}~(k)~\textbf{to word}~i~ \textbf{by participant}~j \\
+\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}) \\ 
+&\text{where}~k \in \{\text{No} \rightarrow \text{Understands}, \text{Understands} \rightarrow \text{Understands and Says}\}\\
+&\textbf{Distribution parameters} \\
+\theta_{k_{ij}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + \\
+&(\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + 
+(\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
+& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{ij} + (\beta_{5} + u_{5_{i}} + w_{3_{j}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
+& (\beta_{6} + u_{6_{i}}) \cdot (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
+& (\beta_{7} + u_{7_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
+& (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij}) \\
+&\text{where:}\\
+& u_{1-8_{i}}~\text{: participant-level adjustments} \\
+& w_{1-3_{j}}~\text{: TE-level adjustments} \\
+&\textbf{Prior} \\
+\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.5); ~\beta_{1-5} \sim \mathcal{N}(0, 1) \\
+\sigma_{u_{0-8}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2) \\
+&\text{where }~\rho_{u_{0-8}, w_{0-3}}~\text{are the correlations between group-level adjustments}\\
+\end{aligned}
+}\label{eq-linear}\end{equation}
+
+We assessed the practical relevance of the estimated regression
+coefficients of the model model following Kruschke \& Liddell
+(\protect\hyperlink{ref-kruschke2018bayesian}{2018}). First, we
+specified a region of practical equivalence (ROPE) from -2.5\% to
++2.5\%, in the probability scale. This region indicates the range of
+values that we considered equivalent to zero. We then summarised the
+posterior distribution of each regression coefficient with the 95\%
+highest density interval (HDI). This interval contains the true value of
+this coefficient with 95\% probability, given the data. Finally, we
+calculated the proportion of posterior samples in the 95\% HDI that fell
+into the ROPE, noted as \(p(\text{ROPE})\), which indicates the
+probability that the true value of the regression coefficient falls into
+the ROPE (and therefore should be considered equivalent to zero). For
+example, a 80\% overlap between the HDI of a coefficient's posterior
+distribution and the indicates that, given our data, there is a 80\%
+probability that the true value of the coefficient falls within the
+ROPE, and can therefore be considered equivalent to zero.
+
+Data processing and visualisation was done in R
+(\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
+version 4.2.2) using the Tidyverse family of packages
+(\protect\hyperlink{ref-wickham2019welcome}{Wickham et al., 2019}). We
+implemented the model using \texttt{brms}
+(\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), a R interface
+to the Stan probabilistic language (2.32.1)
+(\protect\hyperlink{ref-carpenter2017stan}{Carpenter et al., 2017}). We
+ran four MCMC chains with 1,000 iterations each and an additional 1,000
+warm-up iterations per chain. Model posterior draws and predictions were
+handled using the \texttt{tidybayes}
+(\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) and
+\texttt{marginaleffects}
+(\protect\hyperlink{ref-arel-bundock2022marginaleffects}{Arel-Bundock,
+2022}) R packages.
+
+\hypertarget{sec-results}{%
+\section{Results}\label{sec-results}}
+
+\renewcommand{\arraystretch}{1.25}
+
+Model posterior showed adequate MCMC convergence diagnostics for all
+parameters (\(\hat{R} \le 1.06\), \(0.022 \le N_{eff}/N < 1.1001\)) and
+little evidence of multicollinearity. Table~\ref{tbl-coefs} shows the
+summary of the posterior distribution of the fixed regression
+coefficients, and their degree of overlap with the ROPE. For
+interpretability, we report the estimated regression coefficients
+transformed to the probability scale\footnote{The logit and probability
+  scales relate non-linearly. This means that one logit difference is
+  not necessarily translated to a unique value in the probability scale.
+  For example, the probability of acquisition of a given word might
+  increase in 5\% when age increases from 22 to 23 months, the
+  probability of acquisition of the same word might only increase in
+  0.2\% when age increases from 31 to 32 months. The linear growth of
+  the probability of acquisition differs along the logistic curve, and
+  therefore deciding the age point at which to report the estimates of
+  the regression coefficients in the probability scale is not trivial.
+  Following Gelman et al.
+  (\protect\hyperlink{ref-gelman2020regression}{2020}), we report the
+  maximum value of such coefficient, which corresponds to the linear
+  growth (i.e.~derivative) of the logistic curve at the age at which
+  most participants were acquiring a given word. This value can be
+  approximated by dividing the coefficient in the logit scale by four:
+  \(\hat{\beta_j}/4\), where \(\hat{\beta_j}\) is the estimated mean of
+  the posterior distribution of coefficient \(j\).}. The resulting
+values correspond to the maximum difference in probability of
+acquisition (\emph{Comprehension} or \emph{Comprehension and
+Production}) that corresponds to a one standard deviation change in each
+predictor.
+
+\hypertarget{tbl-coefs}{}
+\begin{table}
+\caption{\label{tbl-coefs}Posterior distribution of regression coefficients. }\tabularnewline
+
+\centering
+\begin{tabular}{lrlr}
 \toprule
-& \multicolumn{4}{c}{Identical cognates} & \multicolumn{4}{c}{Non-identical cognates} & \multicolumn{4}{c}{Non-cognates} \\ 
- \cmidrule(lr){2-5}\cmidrule(lr){6-9}\cmidrule(lr){10-13}
-List & N & M & SD & 95\% *CI & N & M & SD & 95\% CI & N & M & SD & 95\% CI \\ 
+ & Median & 95\% HDI & $p(\text{ROPE})$\\
 \midrule
-\multicolumn{1}{l}{CAT} \\ 
-\midrule
-A & 41 & $4.28$ & $0.10$ & [4.09, 4.09] & 137 & $4.49$ & $0.05$ & [4.39, 4.39] & 105 & $4.23$ & $0.06$ & [4.12, 4.12] \\ 
-B & 44 & $4.46$ & $0.11$ & [4.25, 4.25] & 131 & $4.43$ & $0.06$ & [4.32, 4.32] & 101 & $4.24$ & $0.07$ & [4.11, 4.11] \\ 
-C & 37 & $4.29$ & $0.11$ & [4.08, 4.08] & 135 & $4.41$ & $0.06$ & [4.3, 4.3] & 102 & $4.26$ & $0.06$ & [4.14, 4.14] \\ 
-D & 38 & $4.25$ & $0.08$ & [4.08, 4.08] & 134 & $4.38$ & $0.06$ & [4.27, 4.27] & 101 & $4.27$ & $0.06$ & [4.15, 4.15] \\ 
-\midrule
-\multicolumn{1}{l}{SPA} \\ 
-\midrule
-A & 42 & $4.18$ & $0.11$ & [3.96, 3.96] & 137 & $4.40$ & $0.06$ & [4.28, 4.28] & 106 & $4.03$ & $0.07$ & [3.9, 3.9] \\ 
-B & 48 & $4.29$ & $0.10$ & [4.09, 4.09] & 136 & $4.36$ & $0.07$ & [4.22, 4.22] & 105 & $4.07$ & $0.07$ & [3.93, 3.93] \\ 
-C & 40 & $4.19$ & $0.13$ & [3.93, 3.93] & 139 & $4.32$ & $0.07$ & [4.19, 4.19] & 103 & $4.10$ & $0.07$ & [3.96, 3.96] \\ 
-D & 42 & $4.17$ & $0.12$ & [3.94, 3.94] & 137 & $4.33$ & $0.06$ & [4.21, 4.21] & 105 & $4.09$ & $0.07$ & [3.96, 3.96] \\ 
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Intercepts (at 22 months)}}\\
+\hspace{1em}Comprehension and Production & 0.4378 & {}[0.379, 0.496] & 0.0884\\
+\hspace{1em}Comprehension & 0.9363 & {}[0.92, 0.949] & 0.0000\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Slopes (upper bound)}}\\
+\hspace{1em}Age (+1 SD, 4.87, months) & 0.4049 & {}[0.357, 0.451] & 0.0000\\
+\hspace{1em}Exposure (+1 SD, 1.81) & 0.2333 & {}[0.201, 0.268] & 0.0000\\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.0584 & {}[0.014, 0.104] & 0.0371\\
+\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.0620 & {}[-0.086, -0.036] & 0.0000\\
+\hspace{1em}Age $\times$ Exposure & 0.0714 & {}[0.039, 0.104] & 0.0000\\
+\hspace{1em}Age $\times$ Cognateness & 0.0141 & {}[0, 0.026] & 0.9847\\
+\hspace{1em}Exposure $\times$ Cognateness & -0.0569 & {}[-0.069, -0.046] & 0.0000\\
+\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.0180 & {}[-0.027, -0.01] & 0.9747\\
 \bottomrule
-\end{longtable}
-
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
-
-We gathered 192059 observations, with each observation corresponding to a single reponse of one participant to a single item.
-Translation equivalents (TEs) received an average of 305.83 responses (\emph{SD} = 189.00, \emph{Min} = 71, Max = 622), both languages summed together.
-To investigate our hypotheses, we modelled the probability of children understanding or producing TEs using a logistic regression model with a Bernoulli distribution and a logit link function (Agresti, 2003).
-We included several predictors of interest as fixed effects to adjust this probability to participants' and items' characteristics, namely:
-
-\begin{itemize}
-\tightlist
-\item
-  \emph{Age} of participant in months (\texttt{age\_center}), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
-\item
-  \emph{Lexical frequency} of item (\texttt{frequency\_center}), Relative lexical frequencies (frequency per million words, \(fpmw\)) were extracted from the CHILDES data base, using the \texttt{childesr} R package (M.
-Braginsky, Sanchez, \& Yurovsky, 2018), and were transformed to Zipf scores (Van Heuven, Mandera, Keuleers, \& Brysbaert, 2014; Zipf, 1949) (\(log10(fpmw)+3\)).
-Missing frequencies were imputed by multiple imputation using the \texttt{mice} R package {[}van2011mice{]}.
-The resulting scores were then centred around the mean.
-\item
-  \emph{Degree of exposure} to the language (\texttt{doe\_center}), computed as the proportion of the time the child is exposed to the language the item belongs to.
-For instance, for a response to the item \texttt{perro} (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1).
-Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10\% increase in exposure to the language of the item.
-\item
-  \emph{Cognateness}, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist.
-We sum-coded two contrasts (Schad, Vasishth, Hohenstein, \& Kliegl, 2020): \texttt{cognate1}, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and \texttt{cognate2}, comparing IC (coded as +0.5) with NIC (coded as -0.5).
-\end{itemize}
-
-We first fit a base model that only included the main effects of \texttt{age\_center} and \texttt{frequency\_center} (Model 1) (see Mika Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019; Kachergis, Marchman, \& Frank, 2021 for similar approaches).
-We then entered our predictors of interest by adding \texttt{doe\_center} (Model 2) and then \texttt{cognate} and its interaction with \texttt{doe\_center} (Model 3).
-To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included \emph{Participants} (\texttt{id}) and \emph{TEs} (\texttt{te}) as grouping variables (i.e.~random effects), with all possible random intercepts and slopes where appropriate (Barr, Levy, Scheepers, \& Tily, 2013) (see Appendix for more details of the model).
-
-We adopted a Bayesian approach toward statistical inference.
-This approach allows to (1) incorporate previous domain knowledge into the inference process implementing the Bayes theorem, and (2) to quantify the uncertainty associated to the estimated parameters in our model (McElreath, 2018).
-We fit the models using the R package \texttt{brms} (Bürkner, 2017), an R interface to the probabilistic language Stan (Carpenter et al., 2017).
-We run four sampling chains with 1,000 iterations each, including 500 warm-up iterations per chain.
-All R-hat values were smaller than 1.1, indicating that the chains converged successfully in all cases (Gelman \& Rubin, 1992) (see Appendix for an extended description of model diagnostics).
-
-We then compared comprehension and production models separately using leave-one-out cross-validation (LOO-CV) (Vehtari, Gelman, \& Gabry, 2017)\footnote{Due to the high computational cost associated with our large dataset, we used a sub-sampling approach for performing Bayesian LOO with 1,000 samples (Magnusson, Andersen, Jonasson, \& Vehtari, 2019)}.
-This method computes, for a given model, the sum of the log scores of the posterior predictive distributions that result from removing one data-point at a time, providing an estimate of the model fit.
-The LOO estimates are adjusted for the number of parameters estimated in the model, therefore accounting for overfitting.
-We compared the expected log posterior density (ELPD) of each each model to test which one fitted the data the best.
-
-We then explored the posterior distribution of each parameter in the comprehension and production models that fitted the data the best computing the 95\% credible intervals and testing whether this interval excluded 0.
-Credible intervals (CrI) indicate the range of values we are 95\% certain contain the true value of the parameter, conditional to the observed data.
-We performed follow-up tests on interactions whose 95\% credible interval excluded 0 by comparing the 95\% CrI of their expected posterior marginal means using the \texttt{tidybayes} R package (Kay, 2020).
-Data processing and visualisation was done in R (R Core Team, 2020) using the \texttt{tidyverse} family of packages (Wickham et al., 2019).
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-For both comprehension and production, the extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density.
-This estimate is several times larger than its standard error indicating high predictive accuracy.
-Table 3 summarises the LOO scores of the fitted models.
-In the next sections we report the fixed effects of the comprehension and production models separately.
-We report the mean of each coefficient's posterior distribution along with standard errors and 95\% CrIs in Tables 4 and 5.
-For interpretability, we report the derivative of the logistic function for each coefficient\footnote{The derivative is calculated as \(\hat{\beta_j}/4\), where \(\hat{\beta_j}\) is the estimated mean of the posterior distribution of coefficient \(j\).
-This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.} for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable (Gelman, Hill, \& Vehtari, 2020).
-
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{lrrrrrr}
-\toprule
-& \multicolumn{3}{c}{\textbf{ELPD}} & \multicolumn{3}{c}{\textbf{Diff.}} \\ 
- \cmidrule(lr){2-4}\cmidrule(lr){5-7}
-Model & \emph{ELPD} & \emph{SE} & \emph{SE} (sub) & \emph{diff} & \emph{SE} & \emph{SE} (sub) \\ 
-\midrule
-\multicolumn{1}{l}{Comprehension} \\ 
-\midrule
-Model 1 & $-233,647.32$ & $688.29$ & $3,175.16$ & - & - & - \\ 
-Model 2 & $-241,540.78$ & $716.13$ & $3,477.35$ & $7,893.47$ & $993.27$ & $4,708.89$ \\ 
-Model 3 & $-248,415.02$ & $746.19$ & $3,830.96$ & $14,767.71$ & $1,015.16$ & $4,975.74$ \\ 
-\midrule
-\multicolumn{1}{l}{Production} \\ 
-\midrule
-Model 1 & $-51,210.58$ & $198.89$ & $106.81$ & - & - & - \\ 
-Model 2 & $-55,243.99$ & $204.72$ & $124.16$ & $4,033.42$ & $285.42$ & $163.78$ \\ 
-Model 3 & $-60,438.69$ & $207.27$ & $87.25$ & $9,228.11$ & $287.26$ & $137.92$ \\ 
-\bottomrule
-\end{longtable}
-
-\hypertarget{comprehension}{%
-\section{Comprehension}\label{comprehension}}
-
-The average probability of participants understanding a TE was 62.37\% (intercept) when all predictors were equal to 0: mean participant age (24.72 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average across the three levels).
-An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 18.75\% increment in comprehension probability.
-A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a -0.08\% increment in comprehension probability.
-A 10\% increase in relative language exposure made TEs -0.10\% more likely to be understood.
-Cognates (identical and non-identical) were -0.05\% more like to be understood than non-cognates.
-The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0.
-Identical cognates were 0.01\% less like to be understood than non-identical cognates.
-The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero).
-Figure 1 summarises the posterior distribution of the fixed effects of the model.
-
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{llllrrr}
-\toprule
-& \multicolumn{3}{c}{Comprehension} & \multicolumn{3}{c}{Production} \\ 
- \cmidrule(lr){2-4}\cmidrule(lr){5-7}
-\textbf{Predictor} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} \\ 
-\midrule
-Intercept\textsuperscript{1} & $62.37\%$ & $52.59\%$ & [$57.48\%$, $66.97\%$] & $15.30\%$ & $51.25\%$ & [$14.08\%$, $16.57\%$] \\ 
-age\_center\textsuperscript{2} & $18.75\%$ & $2.44\%$ & [$13.79\%$, $23.38\%$] & $5.93\%$ & $0.13\%$ & [$5.66\%$, $6.19\%$] \\ 
-frequency\_center\textsuperscript{2} & $-0.08\%$ & $2.54\%$ & [$-5.32\%$, $4.87\%$] & $3.27\%$ & $0.59\%$ & [$2.13\%$, $4.49\%$] \\ 
-doe\_center\textsuperscript{2} & $-0.10\%$ & $2.71\%$ & [$-5.41\%$, $5.31\%$] & $14.41\%$ & $0.48\%$ & [$13.49\%$, $15.34\%$] \\ 
-cognate1\textsuperscript{2} & $-0.05\%$ & $2.50\%$ & [$-4.99\%$, $4.74\%$] & $2.95\%$ & $2.11\%$ & [$-1.01\%$, $7.16\%$] \\ 
-cognate2 & $-0.01\%$ & $2.57\%$ & [$-4.99\%$, $4.95\%$] & $-0.64\%$ & $2.04\%$ & [$-4.67\%$, $3.33\%$] \\ 
-doe\_center:cognate1\textsuperscript{2} & $-0.07\%$ & $2.47\%$ & [$-5.15\%$, $4.61\%$] & $-9.05\%$ & $0.83\%$ & [$-10.61\%$, $-7.28\%$] \\ 
-doe\_center:cognate2 & $-0.05\%$ & $2.44\%$ & [$-4.85\%$, $4.71\%$] & $-6.53\%$ & $0.73\%$ & [$-7.95\%$, $-5.11\%$] \\ 
-\bottomrule
-\end{longtable}
-\vspace{-5mm}
-\begin{minipage}{\linewidth}
-\textsuperscript{1}Transformed using the inverse logit to get the average probability of correct response \\ 
-\textsuperscript{2}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
-\end{minipage}
-
-\hypertarget{production}{%
-\subsection{Production}\label{production}}
-
-The average probability of participants understanding a TE was 15.30\% (intercept) when all predictors were equal to 0: mean participant age (24.72 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels).
-An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 5.93\% increment in comprehension probability.
-A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 3.27\% increment in comprehension probability.
-A 10\% increase in relative language exposure made TEs 14.41\% more likely to be understood.
-Cognates (identical and non-identical) were 2.95\% more like to be understood than non-cognates.
-The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0.
-Identical cognates were 0.01\% less like to be understood than non-identical cognates.
-The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero).
-Figure 1 summarises the posterior distribution of the fixed effects of the model.
-
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{lcccrrr}
-\toprule
-\textbf{Predictor} & \textbf{Mean} & \textbf{SEM} & \textbf{95\% CrI} & \textbf{Rhat} & \textbf{Bulk ESS}\textsuperscript{1} & \textbf{Tail ESS}\textsuperscript{1} \\ 
-\midrule
-Intercept\textsuperscript{2} & $15.30\%$ & $51.25\%$ & [$14.08\%$, $16.57\%$] & $1.03$ & $108$ & $208$ \\ 
-age\_center\textsuperscript{3} & $5.93\%$ & $0.13\%$ & [$5.66\%$, $6.19\%$] & $1.00$ & $620$ & $654$ \\ 
-frequency\_center\textsuperscript{3} & $3.27\%$ & $0.59\%$ & [$2.13\%$, $4.49\%$] & $1.01$ & $603$ & $666$ \\ 
-doe\_center\textsuperscript{3} & $14.41\%$ & $0.48\%$ & [$13.49\%$, $15.34\%$] & $1.00$ & $675$ & $931$ \\ 
-cognate1\textsuperscript{3} & $2.95\%$ & $2.11\%$ & [$-1.01\%$, $7.16\%$] & $1.03$ & $249$ & $419$ \\ 
-cognate2 & $-0.64\%$ & $2.04\%$ & [$-4.67\%$, $3.33\%$] & $1.01$ & $307$ & $469$ \\ 
-doe\_center:cognate1\textsuperscript{3} & $-9.05\%$ & $0.83\%$ & [$-10.61\%$, $-7.28\%$] & $1.00$ & $1,121$ & $890$ \\ 
-doe\_center:cognate2 & $-6.53\%$ & $0.73\%$ & [$-7.95\%$, $-5.11\%$] & $1.00$ & $1,172$ & $790$ \\ 
-\bottomrule
-\end{longtable}
-\vspace{-5mm}
-\begin{minipage}{\linewidth}
-\textsuperscript{1}ESS: Effective sample size \\ 
-\textsuperscript{2}Transformed using the inverse logit to get the average probability of correct response \\ 
-\textsuperscript{3}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
-\end{minipage}
-
-\textbackslash begin\{figure\}
-
-\{\centering \includegraphics[width=0.8\linewidth]{manuscript_files/figure-latex/emmeans-1}
-
-\}
-
-\textbackslash caption\{Expected marginal mean and 95\% credible interval of the posterior probability of comprehension and production for identical cognates, non-identical cognates, and non cognates, plotted separately for three levels of degree of exposue to the language: 0\% (no exposure, monolingual), 50\% (exposed half of the time, balanced bilingual), and 100\% (exposed all of the time, monolingual).
-Dots represent the mean, and error bars represent the range of values within the true values is contained with 50\% (thick line) and 95\% (thin line) cprobability, given the data.\}\label{fig:emmeans}
-\textbackslash end\{figure\}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+\end{tabular}
+\end{table}
+
+The main effect of \emph{Age} showed the strongest association with the
+probability of acquisition (\(\beta\) = 0.405, 95\% HDI = {[}0.357,
+0.451{]}), with all posterior samples falling out of the ROPE. A
+one-month increment in age increased a maximum of 0.08 the probability
+of acquisition. Similarly, the word exposure index (\emph{Exposure}) had
+a strong effect on the probability of acquisition (\(\beta\) = 0.233,
+95\% HDI = {[}0.201, 0.268{]}). All of the posterior samples of this
+regression coefficient excluded the ROPE. The impact of this predictor
+on the probability of acquisition was positive: for every standard
+deviation increase in exposure, the participant was 0.129 more likely to
+acquire it. Word-form length also showed a significant association with
+probability of acquisition (\(\beta\) = -0.062, 95\% HDI = {[}-0.086,
+-0.036{]}). For every phoneme in the word-form, participants were -0.04
+less likely to know it. The 95\% HDI of the regression coefficient of
+the \emph{Age} \(\times\) \emph{Exposure} interaction also excluded the
+ROPE (\(\beta\) = 0.071, 95\% HDI = {[}0.039, 0.104{]}), showing that
+the effect of the word exposure index differed across ages: older
+children were more sensitive likely to acquire words with higher
+exposure rate than younger children.
+
+The posterior distribution of the main effect of cognateness excluded
+the ROPE completely (\(\beta\) = 0.058, 95\% HDI = {[}0.014, 0.104{]}).
+For every 10\% increment in cognateness, the acquisition of a word
+increased in 0.006. The effect of \emph{Cognateness} interacted with
+that of \emph{Exposure}: the 95\% HDI of the regression coefficient of
+interaction excluded the ROPE entirely (\(\beta\) = -0.06, 95\% HDI =
+{[}-0.069, -0.046{]}), suggesting that the effect of cognateness on a
+word's probability of acquisition changed depending on participants'
+exposure to word. Follow-up analyses on this interaction showed that
+when exposure rate was low (e.g., -1 SD), cognateness increased the
+probability of acquisition substantially. This effect was negligible
+when for words with median or high exposure (+1 SD) (see
+Figure~\ref{fig-marginal}).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-marginal-1.pdf}
+
+}
+
+\caption{\label{fig-marginal}Posterior marginal effects. Lines
+correspond to 100 posterior-predicted medians. Different colours
+indicate different levels of cognateness (phonological similarity).
+Predictions are presented separately for different degrees of word
+exposure index: little exposure to the word, mean exposure, and high
+exposure. Predictions for Comprehension are shown on top and predictions
+for Comprehension and Production are shown on the bottom. In-sample
+predictions lie inside the grey rectangles.}
+
+\end{figure}
+
+An additional analysis including lexical frequency and language exposure
+as separate predictors (instead of the composite \emph{Exposure}
+measure) showed equivalent results (see Appendix A). To rule out the
+possibility that cognateness facilitation effect we found was due to
+cognateness comprising more frequent syllables than non-cognates---and
+therefore not because of their cognate status itself---, we compared the
+syllabic frequency of cognates and non-cognates included in our
+analyses. To calculate syllable frequency, we first extracted all
+syllables embedded in the selected words. For each syllable, we summed
+the lexical frequency of all the words in which such syllable appeared.
+The resulting value provided an estimate of the number of times the
+syllable appears in child-directed speech, embedded within different
+words. Finally, for each word, we summed the frequency of its syllables,
+as an estimate of the syllabic frequency of the word. We fit a Bayesian
+model with \emph{Cognateness} as response variable, and the main effects
+of syllable frequency and number of syllables (to control for the fact
+words with more syllables are more likely to score higher in syllabic
+frequency) as predictors. This model provided strong evidence for the
+association between cognateness and syllabic frequency being equivalent
+to zero (see Appendix B).
+
+\hypertarget{sec-discussion}{%
+\section{Discussion}\label{sec-discussion}}
+
+This study investigated the impact of cognateness on the early bilingual
+lexicon. We used Bayesian Item Response Theory to model the acquisition
+trajectories of a large sample of Catalan and Spanish words, estimating
+the effect of cognateness on the probability of acquisition. This model
+corrected for participants' age, word-form length (number of phonemes),
+and a novel measure of participants' exposure rate to each word.
+Exposure rates were calculated as a language exposure-weighted lexical
+frequency score in which each word-form's lexical frequency was
+corrected by the degree to which the participant was exposed to each
+language. Overall, we found that cognates (i.e., phonologically similar
+translation equivalents) were acquired earlier than non-cognates. This
+effect was mediated by exposure rate: low-exposure word-forms benefited
+from their cognate status, whereas high-exposure word-forms did not.
+Capitalising on accumulator models of language acquisition, we provide a
+theoretical account of bilingual lexical acquisition. In this account,
+parallel activation of the two languages plays a central role during the
+consolidation of early representations in the bilingual lexicon, and in
+which the dynamics of co-activation between translation equivalents
+results in an earlier age-of-acquisition.
+
+The present investigation is particularly relevant in the light of two
+previous findings. First, Floccia et al.
+(\protect\hyperlink{ref-floccia2018introduction}{2018}) reported that
+bilingual toddlers learning two typologically close languages
+(e.g.~shared many cognates, like., English-German) showed larger
+vocabulary sizes than those learning typologically distant languages
+(e.g.~shared fewer cognates, like English-Mandarin). Second, Mitchell et
+al. (\protect\hyperlink{ref-mitchell2022cognates}{2022}) found an
+earlier age-of-acquisition for cognates, compared to non-cognates. The
+outcomes of both studies pointed to cognateness facilitating word
+acquisition through parallel activation. But the underpinnings of such
+effect were unclear: while parallel activation has been extensively
+described in experimental studies, current paradigms of bilingual word
+acquisition and word learning are, to a large extent, dissociated from
+the mechanisms proposed by previous work on word processing. Accumulator
+models of language acquisition may provide a convenient theoretical
+framework to narrow this gap.
+
+Accumulator models devise word acquisition as a continuous process in
+which the child gathers information about words by accumulating learning
+instances with such words. When the number of cumulative learning
+instances for a word reaches some theoretical threshold, the child is
+considered to have acquired such word. The rate at which a child
+accumulates learning instances with a word is a function of child-level
+properties (e.g., ability, amount of quantitative language exposure) and
+word-level properties (e.g., lexical frequency)
+(\protect\hyperlink{ref-hidaka2013computational}{Hidaka, 2013}). Through
+statistical inference, formalised accumulator models provide meaningful
+information about parameters of interest like the aforementioned
+predictors (\protect\hyperlink{ref-kachergis2022standard}{Kachergis et
+al., 2022}; \protect\hyperlink{ref-mollica2017how}{Mollica \&
+Piantadosi, 2017}), and allow to generate quantitative predictions about
+age-of-acquisition and vocabulary growth under competing theoretical
+accounts (\protect\hyperlink{ref-hidaka2013computational}{Hidaka, 2013};
+\protect\hyperlink{ref-mcmurray2007defusing}{McMurray, 2007}).
+Capistalising on accumulator models, we extended this account to the
+bilingual case, suggesting that the cognate facilitation effect on
+bilingual word acquisition is the result of cognate words being
+activated more strongly by their translation than non-cognates,
+therefore accumulating learning instances at a faster rate. We
+hypothesised that when a bilingual child is exposed to a word-form, they
+activate not only its corresponding lexical representation, but also the
+lexical representation of its translation. This cross-language
+activation occurs at the phonological level, and the amount of
+co-activation that spreads from the spoken word to its translation is
+proportional to the amount of phonological similarity between both
+word-forms. Cognates would receive more activation from their
+translation than non-cognates, leading children to accumulate learning
+instances with cognate words at a faster rate than with non-cognate
+words. As a result, lexical representations of cognate words would
+consolidate at earlier ages than those of non-cognate words.
+
+These predictions address a critical subject in bilingualism research:
+do bilingual infants accumulate learning experiences in both languages
+independently, or does exposure to one language impact the acquisition
+trajectory of the other language? In the context of lexical acquisition,
+the former scenario predicts that every learning instance for a given
+word-form contributes to the consolidation of the representation of such
+word in the lexicon, while the consolidation of its translation remains
+unaffected by such experience. In the latter scenario, a learning
+instance to the same word-form would contribute not only to the
+consolidation of the representation of such word, but also, to some
+extent, to the consolidation of its translation. Our findings provide
+strong support for an account of bilingual vocabulary growth in which
+the experience and learning outcomes accumulated by the child in one
+language impact those in the other language. Such a facilitatory
+cross-language mechanism might be an important piece in the puzzle of
+bilingual language acquisition. In particular, it may shed some light on
+why bilingual infants do not show relevant delays in language
+acquisition milestones compared to their monolingual peers, while
+receiving a reduced quantity of speech input in each of their languages.
+Our results provide some insights into this issue: infants benefited
+more strongly from the cognateness facilitation effect when acquiring
+words from the language of lower exposure than in the language of higher
+exposure.
+
+We suggest that this asymmetry is the result of children' unbalanced
+exposure to their languages. A bilingual child's frequency of exposure
+to a given word-form is mostly determined by two factors: the
+word-form's lexical frequency and the child's amount of language
+exposure to the language such word belongs to. A dual linguistic input
+means lower exposure to each of the languages, unless one makes
+the---arguably implausible---assumption that bilinguals are exposed on
+average to twice the amount of linguistic input than monolinguals.
+Because of this difference in exposure, words lower-exposure language
+might receive activation from their translation in the higher-exposure
+language more often than words from the higher-exposure language receive
+activation from their translation in the lower-exposure language. As a
+result, the cognateness facilitation effect should be stronger in words
+from the lower-exposure language.
+
+This mechanism might be extended to provide a plausible explanation for
+the language similarity facilitation reported by Floccia et al.~The
+authors observed a facilitation in the additional (non-English)
+language: children learning two typologically close languages knew more
+words in the additional language than those learning two typologically
+more distant languages. In their sample, the additional language was
+consistently also the lower-exposure language for most children, while
+English was the higher-exposure language. Given that words in English
+were more likely to be acquired first, higher phonological overlap for
+words in the language of lower exposure (especially those of lower
+lexical frequency) would facilitate vocabulary growth for languages
+sharing more cognates with English.
+
+It might be argued that our results reflect the fact that cognate
+translation equivalents are represented in the initial bilingual lexicon
+as the \emph{same} lexical entry. Because cognates correspond to similar
+sounding word-forms in equivalent referential contexts (e.g., hearing
+/\textipa{"gat}\}/ and /\textipa{"ga.to}\}/ in the presence of a cat),
+it is possible that infants classify both are as acceptable variations
+of the same word, therefore treating them as a single lexical item. This
+would lead to a faster increase in cumulative learning instances, and to
+earlier ages of acquisition for cognate translation equivalents (for
+which listening to each word-form contributes to the acquisition of its
+shared representation), compared to non-cognates (for which listening to
+each word-form contributes to the acquisition of a separate
+representation). This mechanism could potentially explain the earlier
+age-of-acquisition effect of cognates found in the present study,
+without the need of parallel activation playing any relevant role.
+Mitchell et al. (\protect\hyperlink{ref-mitchell2022cognates}{2022})
+discuss this possibility as a possible explanation of the cognate
+facilitation effect, in which bilinguals only need to map one word-form
+to the referent in the case of cognates, while mapping two distinct
+word-forms in the case of non-cognates. However, previous work on
+mispronunciation perception and learning of minimal pair words points in
+a different direction. Bilingual toddlers show monolingual-level
+sensitivity to slight phonetic changes in a word-form, according to
+their performance in word recognition tasks
+(\protect\hyperlink{ref-bailey2002phonological}{Bailey \& Plunkett,
+2002}; \protect\hyperlink{ref-mani2011does}{Mani \& Plunkett, 2011};
+\protect\hyperlink{ref-ramon-casas2009vowel}{Ramon-Casas et al., 2009},
+\protect\hyperlink{ref-ramon-casas2017minimalpair}{2017};
+\protect\hyperlink{ref-ramon-casas2010are}{Ramon-Casas \& Bosch, 2010};
+\protect\hyperlink{ref-swingley200511montholds}{Swingley, 2005};
+\protect\hyperlink{ref-swingley2000spoken}{Swingley \& Aslin, 2000};
+\protect\hyperlink{ref-tamasi2017pupillometry}{Tamási et al., 2017};
+\protect\hyperlink{ref-wewalaarachchi2017vowels}{Wewalaarachchi et al.,
+2017}). The ability to differentiate between similar-sounding word-forms
+is also reflected in word learning, as bilinguals seem to be able to map
+minimal pairs to distinct referents
+(\protect\hyperlink{ref-havy2016phonetic}{Havy et al., 2016};
+\protect\hyperlink{ref-mattock2010first}{Mattock et al., 2010};
+\protect\hyperlink{ref-ramon-casas2017minimalpair}{Ramon-Casas et al.,
+2017}). Overall, it seems that bilinguals consider small differences in
+the phonological forms of words as relevant at the lexical level. We
+argue that this shows evidence that bilingual toddlers likely form
+distinct lexical representations for even near-identical cognates.
+
+Our study shares similar methodological limitations with previous work
+using vocabulary reports provided by caregivers. Such reports can be
+subject to measurement error induced by caregivers who may sometimes
+overestimate or underestimate participants' true probability of
+acquisition of words (e.g.,
+\protect\hyperlink{ref-houston-price2007discrepancy}{Houston-Price et
+al., 2007}). In the case of bilingual research additional biases may be
+in place. Although in the present study caregivers were explicitly
+instructed \emph{not} to rely on their responses to Catalan words when
+responding to Spanish (and vice versa), it is possible that some
+caregivers assumed---at least to some extent---that because the child
+knew a word in one language, the child should also know the word in the
+other language. This bias would especially affect similar-sounding
+words, i.e., cognates. Production estimates may be more prompt to such
+biases, in part because of the slower pace at which infants'
+articulatory abilities develop, compared to their word recognition
+abilities (\protect\hyperlink{ref-hustad2020development}{Hustad et al.,
+2020}, \protect\hyperlink{ref-hustad2021speech}{2021}). This gap between
+comprehension and production is even larger in the less dominant
+language of bilingual children
+(\protect\hyperlink{ref-giguere2022bilingual}{Giguere \& Hoff, 2022}).
+For this reason, caregivers may be more uncertain about what words can
+be counted as \emph{acquired} in this modality. Despite such potential
+biases, vocabulary checklist filled by parents show strong evidence of
+concurrent validity with other estimates of vocabulary size or lexical
+processing (\protect\hyperlink{ref-feldman2005concurrent}{Feldman et
+al., 2005}; \protect\hyperlink{ref-gillen2021tapping}{Gillen et al.,
+2021}; \protect\hyperlink{ref-killing2008move}{Killing \& Bishop, 2008};
+but see
+\protect\hyperlink{ref-houston-price2007discrepancy}{Houston-Price et
+al., 2007}).
+
+The present study contributes with a specific data point to the complex
+landscape of bilingualism research. Bilinguals are a remarkably
+heterogeneous population difficult to be satisfactorily characterised in
+a comprehensive way
+(\protect\hyperlink{ref-sebastian-galles2020bilingual}{Sebastian-Galles
+\& Santolin, 2020}). Bilinguals differ across multiple dimensions. Such
+differences span from exclusively linguistic factors such as the amount
+of overlap between the phonemic inventories of the two languages being
+learned (e.g., low, like the case of English and Mandarin, or high like
+the case of Spanish and Greek), to higher-level factors like the
+socio-linguistic situation in which the two languages co-exist (e.g., in
+some regions both languages are co-official and used in similar
+contexts, while in others, one of the languages hardly has any societal
+presence, i.e., heritage languages). This diversity of situations in
+which bilingual toddlers acquire language calls for special
+consideration of the generalisability of results in bilingualism
+research. Our sample, although homogeneous (e.g., similar parental
+educational level across), represents a particular bilingual
+sociolinguistic environment: the languages involved in the present
+investigation, Catalan and Spanish, co-exist in Catalonia as official
+languages, both languages are used in fairly similar contexts, and both
+languages are known by the majority of the population. In 2018, more
+than 81.2\% of a representative sample of 8,780 adults aged 15 years or
+older living in Catalonia reported being able to speak Catalan, and more
+than 99.5\% of the same population reported being able to speak Spanish
+(\protect\hyperlink{ref-2018els}{\emph{Els Usos Lingüístics de La
+Població de Catalunya}, 2018}). In addition, Catalan and Spanish are
+Romance languages and share a considerable amount of cognates. Extending
+our analyses to other bilingual populations learning typologically more
+distant languages, and whose languages tend to be used in more distinct
+contexts (e.g., heritage languages) should be a natural future step for
+the present investigation.
+
+To conclude, our study provides novel insights about word acquisition in
+bilingual contexts, and how the presence of cognates in the children's
+linguistic input impacts the early formation of the lexicon. We found
+that during the acquisition of low frequency words, bilingual children
+seem to benefit more strongly from the word's phonological similarity
+with its translation in the other language. Capitalising on accumulator
+models of language acquisition we put forward a theoretical account of
+bilingual word learning, in which cognateness interacts with lexical
+frequency and language exposure to boost the acquisition of translation
+equivalents.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-agresti2003categorical}{}}%
-Agresti, A. (2003). \emph{Categorical data analysis} (Vol. 482). John Wiley \& Sons.
+\leavevmode\vadjust pre{\hypertarget{ref-arel-bundock2022marginaleffects}{}}%
+Arel-Bundock, V. (2022). \emph{Marginaleffects: Marginal effects,
+marginal means, predictions, and contrasts}.
+\url{https://CRAN.R-project.org/package=marginaleffects}
 
 \leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%
-Arslan, R.
-C., Walther, M.
-P., \& Tata, C.
-S.
-(2020).
-Formr: A study framework allowing for automated feedback generation and complex longitudinal experience-sampling studies using r.
-\emph{Behavior Research Methods}, \emph{52}(1), 376--387.
+Arslan, R. C., Walther, M. P., \& Tata, C. S. (2020). Formr: A study
+framework allowing for automated feedback generation and complex
+longitudinal experience-sampling studies using r. \emph{Behavior
+Research Methods}, \emph{52}(1), 376--387.
+\url{https://doi.org/10.3758/s13428-019-01236-y}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bailey2002phonological}{}}%
+Bailey, T. M., \& Plunkett, K. (2002). Phonological specificity in early
+words. \emph{Cognitive Development}, \emph{17}(2), 1265--1282.
+\url{https://doi.org/10.1016/S0885-2014(02)00116-8}
 
 \leavevmode\vadjust pre{\hypertarget{ref-barr2013random}{}}%
-Barr, D.
-J., Levy, R., Scheepers, C., \& Tily, H.
-J.
-(2013).
-Random effects structure for confirmatory hypothesis testing: Keep it maximal.
+Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random
+effects structure for confirmatory hypothesis testing: Keep it maximal.
 \emph{Journal of Memory and Language}, \emph{68}(3), 255--278.
 
-\leavevmode\vadjust pre{\hypertarget{ref-bergelson2020comprehension}{}}%
-Bergelson, E. (2020). The comprehension boost in early word learning: Older infants are better learners. \emph{Child Development Perspectives}, \emph{14}(3), 142--149.
-
 \leavevmode\vadjust pre{\hypertarget{ref-bergelson2012months}{}}%
-Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants know the meanings of many common nouns. \emph{Proceedings of the National Academy of Sciences}, \emph{109}(9), 3253--3258.
+Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants
+know the meanings of many common nouns. \emph{Proceedings of the
+National Academy of Sciences}, \emph{109}(9), 3253--3258.
+\url{https://doi.org/10.1073/pnas.1113380109}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bergelson2015early}{}}%
-Bergelson, E., \& Swingley, D. (2015). Early word comprehension in infants: Replication and extension. \emph{Language Learning and Development}, \emph{11}(4), 369--380.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bloom2002children}{}}%
-Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT press.
+Bergelson, E., \& Swingley, D. (2015). Early word comprehension in
+infants: Replication and extension. \emph{Language Learning and
+Development}, \emph{11}(4), 369--380.
+\url{https://doi.org/10.1080/15475441.2014.979387}
+
+\leavevmode\vadjust pre{\hypertarget{ref-blom2020crosslanguage}{}}%
+Blom, E., Boerma, T., Bosma, E., Cornips, L., Heuij, K. van den, \&
+Timmermeister, M. (2020). Cross-language distance influences receptive
+vocabulary outcomes of bilingual children. \emph{First Language},
+\emph{40}(2), 151--171. \url{https://doi.org/10.1177/0142723719892794}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bosch2014first}{}}%
-Bosch, L., \& Ramon-Casas, M.
-(2014).
-First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter?
-\emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322.
-\url{https://doi.org/10.1177/0165025414532559}
-
-\leavevmode\vadjust pre{\hypertarget{ref-braginsky2018childesr}{}}%
-Braginsky, M., Sanchez, A., \& Yurovsky, D. (2018). Childesr: Accessing the `CHILDES'database. \emph{R Package Version 0.1. 0}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-braginsky2019consistency}{}}%
-Braginsky, Mika, Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and variability in children's word learning across languages. \emph{Open Mind}, \emph{3}, 52--67.
+Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in
+bilingual toddlers' expressive vocabulary: Does form similarity matter?
+\emph{International Journal of Behavioral Development}, \emph{38}(4),
+317--322. \url{https://doi.org/10.1177/0165025414532559}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosma2019longitudinal}{}}%
+Bosma, E., Blom, E., Hoekstra, E., \& Versloot, A. (2019). A
+longitudinal study on the gradual cognate facilitation effect in
+bilingual children's frisian receptive vocabulary. \emph{International
+Journal of Bilingual Education and Bilingualism}, \emph{22}(4),
+371--385. \url{https://doi.org/10.1080/13670050.2016.1254152}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosma2020cognate}{}}%
+Bosma, E., \& Nota, N. (2020). Cognate facilitation in frisian--dutch
+bilingual children's sentence reading: An eye-tracking study.
+\emph{Journal of Experimental Child Psychology}, \emph{189}, 104699.
 
 \leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
-Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
+using stan. \emph{Journal of Statistical Software}, \emph{80}, 1--28.
+\url{https://doi.org/10.18637/jss.v080.i01}
 
 \leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
-Carpenter, B., Gelman, A., Hoffman, M.
-D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A.
-(2017).
-Stan: A probabilistic programming language.
-\emph{Journal of Statistical Software}, \emph{76}(1), 1--32.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cattani2014much}{}}%
-Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis, I., \& Floccia, C.
-(2014).
-How much exposure to english is necessary for a bilingual toddler to perform like a monolingual peer in language tests?
-\emph{International Journal of Language \& Communication Disorders}, \emph{49}(6), 649--671.
+Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B.,
+Betancourt, M., Brubaker, M., Guo, J., Li, P., \& Riddell, A. (2017).
+Stan : A probabilistic programming language. \emph{Journal of
+Statistical Software}, \emph{76}(1).
+\url{https://doi.org/10.18637/jss.v076.i01}
 
 \leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
-Costa, A., Caramazza, A., \& Sebastian-Galles, N.
-(2000).
-The cognate facilitation effect: Implications for models of lexical access.
-\emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{26}(5), 1283.
-
-\leavevmode\vadjust pre{\hypertarget{ref-david2008individual}{}}%
-David, A., \& Wei, L.
-(2008).
-Individual differences in the lexical development of french--english bilingual children.
-\emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.
-
-\leavevmode\vadjust pre{\hypertarget{ref-fenson1994variability}{}}%
-Fenson, L., Dale, P.
-S., Reznick, J.
-S., Bates, E., Thal, D.
-J., Pethick, S.
-J., \ldots{} Stiles, J.
-(1994).
-Variability in early communicative development.
-\emph{Monographs of the Society for Research in Child Development}, i--185.
-
-\leavevmode\vadjust pre{\hypertarget{ref-fenson2007macarthur}{}}%
-Fenson, L., \& others. (2007). \emph{MacArthur-bates communicative development inventories}. Paul H. Brookes Publishing Company Baltimore, MD.
-
-\leavevmode\vadjust pre{\hypertarget{ref-fernald2013ses}{}}%
-Fernald, A., Marchman, V. A., \& Weisleder, A. (2013). SES differences in language processing skill and vocabulary are evident at 18 months. \emph{Developmental Science}, \emph{16}(2), 234--248.
+Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate
+facilitation effect: Implications for models of lexical access.
+\emph{Journal of Experimental Psychology: Learning, Memory, and
+Cognition}, \emph{26}, 1283--1296.
+\url{https://doi.org/10.1037/0278-7393.26.5.1283}
+
+\leavevmode\vadjust pre{\hypertarget{ref-2018els}{}}%
+\emph{Els usos lingüístics de la població de catalunya}. (2018).
+Generalitat de Catalunya.
+\url{https://llengua.gencat.cat/web/.content/documents/dadesestudis/altres/arxius/dossier-eulp-2018.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-feldman2005concurrent}{}}%
+Feldman, H. M., Dale, P. S., Campbell, T. F., Colborn, D. K.,
+Kurs-Lasky, M., Rockette, H. E., \& Paradise, J. L. (2005). Concurrent
+and predictive validity of parent reports of child language at ages 2
+and 3 years. \emph{Child Development}, \emph{76}(4), 856--868.
+\url{https://doi.org/10.1111/j.1467-8624.2005.00882.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fenson2007macarthurbates}{}}%
+Fenson, L. et al. (2007). \emph{{MacArthur}-bates communicative
+development inventories}. Paul H. Brookes Publishing Company Baltimore,
+{MD}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
-Floccia, C., Sambrook, T.
-D., Delle Luche, C., Kwok, R., Goslin, J., White, L., \ldots{} others.
-(2018).
-I: INTRODUCTION.
-\emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29.
-
-\leavevmode\vadjust pre{\hypertarget{ref-frank2021variability}{}}%
-Frank, M. C., Braginsky, M., Yurovsky, D., \& Marchman, V. A. (2021). \emph{Variability and consistency in early language learning: The wordbank project}. MIT Press.
+Floccia, C., Sambrook, T. D., Delle Luche, C., Kwok, R., Goslin, J.,
+White, L., Cattani, A., Sullivan, E., Abbot‐Smith, K., Krott, A., Mills,
+D., Rowland, C., Gervain, J., \& Plunkett, K. (2018). I: introduction.
+\emph{Monographs of the Society for Research in Child Development},
+\emph{83}(1), 7--29. \url{https://doi.org/10.1111/mono.12348}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fourtassi2020growth}{}}%
+Fourtassi, A., Bian, Y., \& Frank, M. C. (2020). The growth of
+children's semantic and phonological networks: Insight from 10
+languages. \emph{Cognitive Science}, \emph{44}(7), e12847.
+\url{https://doi.org/10.1111/cogs.12847}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gampe2021does}{}}%
+Gampe, A., Quick, A. E., \& Daum, M. M. (2021). Does linguistic
+similarity affect early simultaneous bilingual language acquisition?
+\emph{Journal of Language Contact}, \emph{13}(3), 482--500.
+
+\leavevmode\vadjust pre{\hypertarget{ref-garcia-castro2023bvq}{}}%
+Garcia-Castro, G., Ávila-Varela, D. S., \& Sebastian-Galles, N. (2023).
+\emph{Bvq: Barcelona vocabulary questionnaire database and helper
+functions}. \url{https://gongcastro.github.io/bvq}
 
 \leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
-Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gelman1992inference}{}}%
-Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation using multiple sequences. \emph{Statistical Science}, \emph{7}(4), 457--472.
-
-\leavevmode\vadjust pre{\hypertarget{ref-goldfield1990early}{}}%
-Goldfield, B. A., \& Reznick, J. S. (1990). Early lexical acquisition: Rate, content, and the vocabulary spurt. \emph{Journal of Child Language}, \emph{17}(1), 171--183.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gonzalez2020bilingual}{}}%
-Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020). Bilingual adjusted vocabulary: A developmentally-informed bilingual vocabulary measure.
+Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other
+stories}. Cambridge University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-giguere2022bilingual}{}}%
+Giguere, D., \& Hoff, E. (2022). Bilingual development in the receptive
+and expressive domains: They differ. \emph{International Journal of
+Bilingual Education and Bilingualism}, \emph{25}(10), 3849--3858.
+\url{https://doi.org/10.1080/13670050.2022.2087039}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gillen2021tapping}{}}%
+Gillen, N. A., Siow, S., Lepadatu, I., Sucevic, J., Plunkett, K., \&
+Duta, M. (2021). \emph{Tapping into the potential of remote
+developmental research: Introducing the {OxfordBabylab} app}.
+{PsyArXiv}. \url{https://doi.org/10.31234/osf.io/kxhmw}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gimeno-martinez2021crosslinguistic}{}}%
+Gimeno-Martínez, M., Mädebach, A., \& Baus, C. (2021). Cross-linguistic
+interactions across modalities: Effects of the oral language on sign
+production. \emph{Bilingualism: Language and Cognition}, \emph{24}(4),
+779--790. \url{https://doi.org/10.1017/S1366728921000171}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gonzalez-barrero2020bilingual}{}}%
+Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020).
+\emph{Bilingual adjusted vocabulary: A developmentally-informed
+bilingual vocabulary measure}. {PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/x7s4u}
+
+\leavevmode\vadjust pre{\hypertarget{ref-grosjean2021extent}{}}%
+Grosjean, F. (2021). The extent of bilingualism. \emph{Life as a
+Bilingual}, 27--39.
 
 \leavevmode\vadjust pre{\hypertarget{ref-hamilton2000infant}{}}%
-Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary development assessed with a british communicative development inventory. \emph{Journal of Child Language}, \emph{27}(3), 689--705.
+Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary
+development assessed with a british communicative development inventory.
+\emph{Journal of Child Language}, \emph{27}(3), 689--705.
+
+\leavevmode\vadjust pre{\hypertarget{ref-havy2016phonetic}{}}%
+Havy, M., Bouchon, C., \& Nazzi, T. (2016). Phonetic processing when
+learning words: The case of bilingual infants. \emph{International
+Journal of Behavioral Development}, \emph{40}(1), 41--52.
+\url{https://doi.org/10.1177/0165025415570646}
+
+\leavevmode\vadjust pre{\hypertarget{ref-heeringa2003norwegian}{}}%
+Heeringa, W., \& Gooskens, C. (2003). Norwegian dialects examined
+perceptually and acoustically. \emph{Computers and the Humanities},
+\emph{37}(3), 293--315. \url{https://doi.org/10.1023/A:1025087115665}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanhell2008sentence}{}}%
+Hell, J. G. van, \& Groot, A. M. B. de. (2008). Sentence context
+modulates visual word recognition and translation in bilinguals.
+\emph{Acta Psychologica}, \emph{128}(3), 431--451.
+\url{https://doi.org/10.1016/j.actpsy.2008.03.010}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hidaka2013computational}{}}%
+Hidaka, S. (2013). A computational model associating learning process,
+word attributes, and age of acquisition. \emph{{PLOS} {ONE}},
+\emph{8}(11), e76242. \url{https://doi.org/10.1371/journal.pone.0076242}
 
 \leavevmode\vadjust pre{\hypertarget{ref-hoff2012dual}{}}%
-Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M. (2012). Dual language exposure and early bilingual development. \emph{Journal of Child Language}, \emph{39}(1), 1.
+Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M.
+(2012). Dual language exposure and early bilingual development*.
+\emph{Journal of Child Language}, \emph{39}(1), 1--27.
+\url{https://doi.org/10.1017/S0305000910000759}
 
 \leavevmode\vadjust pre{\hypertarget{ref-hoshino2008cognate}{}}%
-Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming: Does cross-language activation survive a change of script? \emph{Cognition}, \emph{106}(1), 501--511.
-
-\leavevmode\vadjust pre{\hypertarget{ref-jones2019children}{}}%
-Jones, S., \& Brandt, S. (2019). Do children really acquire dense neighbourhoods? \emph{Journal of Child Language}, \emph{46}(6), 1260--1273.
-
-\leavevmode\vadjust pre{\hypertarget{ref-jusczyk1995infants}{}}%
-Jusczyk, P. W., \& Aslin, R. N. (1995). Infants' detection of the sound patterns of words in fluent speech. \emph{Cognitive Psychology}, \emph{29}(1), 1--23.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kachergis2021toward}{}}%
-Kachergis, G., Marchman, V. A., \& Frank, M. C. (2021). Toward a {``standard model''} of early language learning.
-
-\leavevmode\vadjust pre{\hypertarget{ref-R-tidybayes}{}}%
-Kay, M. (2020). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
+Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming:
+Does cross-language activation survive a change of script?
+\emph{Cognition}, \emph{106}(1), 501--511.
+\url{https://doi.org/10.1016/j.cognition.2007.02.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-houston-price2007discrepancy}{}}%
+Houston-Price, C., Mather, E., \& Sakkalou, E. (2007). Discrepancy
+between parental reports of infants' receptive vocabulary and infants'
+behaviour in a preferential looking task. \emph{Journal of Child
+Language}, \emph{34}(4), 701--724.
+\url{https://doi.org/10.1017/S0305000907008124}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hustad2021speech}{}}%
+Hustad, K. C., Mahr, T. J., Natzke, P., \& Rathouz, P. J. (2021). Speech
+development between 30 and 119 months in typical children i:
+Intelligibility growth curves for single-word and multiword productions.
+\emph{Journal of Speech, Language, and Hearing Research}, \emph{64}(10),
+3707--3719. \url{https://doi.org/10.1044/2021_JSLHR-21-00142}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hustad2020development}{}}%
+Hustad, K. C., Mahr, T., Natzke, P. E. M., \& Rathouz, P. J. (2020).
+Development of speech intelligibility between 30 and 47 months in
+typically developing children: A cross-sectional study of growth.
+\emph{Journal of Speech, Language, and Hearing Research}, \emph{63}(6),
+1675--1687. \url{https://doi.org/10.1044/2020_JSLHR-20-00008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kachergis2022standard}{}}%
+Kachergis, G., Marchman, V. A., \& Frank, M. C. (2022). Toward a
+{``standard model''} of early language learning. \emph{Current
+Directions in Psychological Science}, \emph{31}(1), 20--27.
+\url{https://doi.org/10.1177/09637214211057836}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kay2021tidybayes}{}}%
+Kay, M. (2021). \emph{Tidybayes: Tidy data and geoms for bayesian
+models}. \url{http://mjskay.github.io/tidybayes/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-killing2008move}{}}%
+Killing, S. E. A., \& Bishop, D. V. M. (2008). Move it! Visual feedback
+enhances validity of preferential looking as a measure of individual
+differences in vocabulary in toddlers. \emph{Developmental Science},
+\emph{11}(4), 525--530.
+\url{https://doi.org/10.1111/j.1467-7687.2008.00698.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroll2017bilingual}{}}%
+Kroll, J. F., \& Ma, F. (2017). The bilingual lexicon. \emph{The
+Handbook of Psycholinguistics}, 294--319.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kruschke2018bayesian}{}}%
+Kruschke, J. K., \& Liddell, T. M. (2018). The bayesian new statistics:
+Hypothesis testing, estimation, meta-analysis, and planning from a
+bayesian perspective. \emph{Psychonomic Bulletin \&Review}, \emph{25},
+178--206. \url{https://doi.org/10.3758/s13423-016-1221-4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-laing2022phonological}{}}%
+Laing, C. E. (2022). \emph{Phonological networks and systematicity in
+early lexical acquisition}. {PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/z8pyg}
 
 \leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
-Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. In \emph{Soviet physics doklady} (Vol. 10, pp. 707--710). Soviet Union.
-
-\leavevmode\vadjust pre{\hypertarget{ref-magnusson2019bayesian}{}}%
-Magnusson, M., Andersen, M., Jonasson, J., \& Vehtari, A. (2019). Bayesian leave-one-out cross-validation for large data. In \emph{International conference on machine learning} (pp. 4244--4253). PMLR.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mayor2011statistical}{}}%
-Mayor, J., \& Plunkett, K. (2011). A statistical estimate of infant and toddler vocabulary size from CDI analysis. \emph{Developmental Science}, \emph{14}(4), 769--785.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcelreath2018statistical}{}}%
-McElreath, R. (2018). \emph{Statistical rethinking: A bayesian course with examples in r and stan}. Chapman; Hall/CRC.
-
-\leavevmode\vadjust pre{\hypertarget{ref-nelson1973structure}{}}%
-Nelson, K. (1973). Structure and strategy in learning to talk. \emph{Monographs of the Society for Research in Child Development}, 1--135.
+Levenshtein, V. I. (1966). Binary codes capable of correcting deletions,
+insertions, and reversals. \emph{Soviet Physics-Doklady}, \emph{10},
+707--710.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanderloo2014stringdist}{}}%
+Loo, M. P. J. van der. (2014). The stringdist package for approximate
+string matching. \emph{The R Journal}, \emph{6}(1), 111--122.
+\url{https://doi.org/10.32614/RJ-2014-011}
+
+\leavevmode\vadjust pre{\hypertarget{ref-macwhinney2000childes}{}}%
+MacWhinney, B. (2000). \emph{The {CHILDES} project: The database} (Vol.
+2). Psychology Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mani2011does}{}}%
+Mani, N., \& Plunkett, K. (2011). Does size matter? Subsegmental cues to
+vowel mispronunciation detection*. \emph{Journal of Child Language},
+\emph{38}(3), 606--627. \url{https://doi.org/10.1017/S0305000910000243}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mattock2010first}{}}%
+Mattock, K., Polka, L., Rvachew, S., \& Krehm, M. (2010). The first
+steps in word learning are easier when the shoes fit: Comparing
+monolingual and bilingual infants. \emph{Developmental Science},
+\emph{13}(1), 229--243.
+\url{https://doi.org/10.1111/j.1467-7687.2009.00891.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcmurray2007defusing}{}}%
+McMurray, B. (2007). Defusing the childhood vocabulary explosion.
+\emph{Science}, \emph{317}(5838), 631--631.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mitchell2022cognates}{}}%
+Mitchell, L., Tsui, R. K. Y., \& Byers-Heinlein, K. (2022).
+\emph{Cognates are advantaged in early bilingual expressive vocabulary
+development}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/daktp}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mollica2017how}{}}%
+Mollica, F., \& Piantadosi, S. T. (2017). How data drive early word
+learning: A cross-linguistic waiting time analysis. \emph{Open Mind},
+\emph{1}(2), 67--77. \url{https://doi.org/10.1162/OPMI_a_00006}
+
+\leavevmode\vadjust pre{\hypertarget{ref-morford2011when}{}}%
+Morford, J. P., Wilkinson, E., Villwock, A., Piñar, P., \& Kroll, J. F.
+(2011). When deaf signers read english: Do written words activate their
+sign translations? \emph{Cognition}, \emph{118}(2), 286--292.
+\url{https://doi.org/10.1016/j.cognition.2010.11.006}
 
 \leavevmode\vadjust pre{\hypertarget{ref-oller2002language}{}}%
-Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in bilingual children} (Vol. 2). Multilingual Matters.
+Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in
+bilingual children}. Multilingual Matters.
 
 \leavevmode\vadjust pre{\hypertarget{ref-patterson2004comparing}{}}%
-Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers' expressive vocabulary size.
+Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers'
+expressive vocabulary size. \emph{Journal of Speech, Language, and
+Hearing Research}, \emph{47}(5), 1213--1215.
+\url{https://doi.org/10.1044/1092-4388(2004/089)}
 
 \leavevmode\vadjust pre{\hypertarget{ref-patterson2004bilingual}{}}%
-Patterson, J., Pearson, B., \& Goldstein, B. (2004). Bilingual language development and disorders in spanish--english speakers.
+Patterson, J. L., \& Pearson, B. Z. (2004). Bilingual lexical
+development: Influences, contexts, and processes. In \emph{Bilingual
+language development and disorders in spanish-english speakers} (pp.
+77--104). Paul H. Brookes Publishing Co.
 
 \leavevmode\vadjust pre{\hypertarget{ref-pearson1994patterns}{}}%
-Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in the lexical growth in two languages of bilingual infants and toddlers. \emph{Language Learning}, \emph{44}(4), 617--653.
+Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in
+the lexical growth in two languages of bilingual infants and toddlers.
+\emph{Language Learning}, \emph{44}(4), 617--653.
+\url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}
 
 \leavevmode\vadjust pre{\hypertarget{ref-pearson1993lexical}{}}%
-Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical development in bilingual infants and toddlers: Comparison to monolingual norms. \emph{Language Learning}, \emph{43}(1), 93--120.
+Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical
+development in bilingual infants and toddlers: Comparison to monolingual
+norms. \emph{Language Learning}, \emph{43}(1), 93--120.
+\url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}
 
 \leavevmode\vadjust pre{\hypertarget{ref-petitto2001bilingual}{}}%
-Petitto, L.
-A., Katerelos, M., Levy, B.
-G., Gauna, K., Tétreault, K., \& Ferraro, V.
-(2001).
-Bilingual signed and spoken language acquisition from birth: Implications for the mechanisms underlying early bilingual language acquisition.
-\emph{Journal of Child Language}, \emph{28}(2), 453--496.
-
-\leavevmode\vadjust pre{\hypertarget{ref-poulin2013lexical}{}}%
+Petitto, L. A., Katerelos, M., Levy, B. G., Gauna, K., Tétreault, K., \&
+Ferraro, V. (2001). Bilingual signed and spoken language acquisition
+from birth: Implications for the mechanisms underlying early bilingual
+language acquisition. \emph{Journal of Child Language}, \emph{28}(2),
+453--496. \url{https://doi.org/10.1017/S0305000901004718}
+
+\leavevmode\vadjust pre{\hypertarget{ref-poarch2012crosslanguage}{}}%
+Poarch, G. J., \& Hell, J. G. van. (2012). Cross-language activation in
+children's speech production: Evidence from second language learners,
+bilinguals, and trilinguals. \emph{Journal of Experimental Child
+Psychology}, \emph{111}(3), 419--438.
+\url{https://doi.org/10.1016/j.jecp.2011.09.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-poulin-dubois2013lexical}{}}%
 Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J.
-(2013).
-Lexical access and vocabulary development in very young bilinguals.
-\emph{International Journal of Bilingualism}, \emph{17}(1), 57--70.
-
-\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
-R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schad2020capitalize}{}}%
-Schad, D.
-J., Vasishth, S., Hohenstein, S., \& Kliegl, R.
-(2020).
-How to capitalize on a priori contrasts in linear (mixed) models: A tutorial.
-\emph{Journal of Memory and Language}, \emph{110}, 104038.
+(2013). Lexical access and vocabulary development in very young
+bilinguals. \emph{International Journal of Bilingualism}, \emph{17}(1),
+57--70.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rcoreteam2013language}{}}%
+R Core Team. (2013). \emph{R: A language and environment for statistical
+computing}. R Foundation for Statistical Computing.
+\url{http://www.R-project.org/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2010are}{}}%
+Ramon-Casas, M., \& Bosch, L. (2010). Are non-cognate words
+phonologically better specified than cognates in the early lexicon of
+bilingual children. \emph{Selected Proceedings of the 4th Conference on
+Laboratory Approaches to Spanish Phonology}, 31--36.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2017minimalpair}{}}%
+Ramon-Casas, M., Fennell, C. T., \& Bosch, L. (2017). Minimal-pair word
+learning by bilingual toddlers: The catalan /e/-/ɛ/ contrast revisited.
+\emph{Bilingualism: Language and Cognition}, \emph{20}(3), 649--656.
+\url{https://doi.org/10.1017/S1366728916001115}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2009vowel}{}}%
+Ramon-Casas, M., Swingley, D., Sebastián-Gallés, N., \& Bosch, L.
+(2009). Vowel categorization during word recognition in bilingual
+toddlers. \emph{Cognitive Psychology}, \emph{59}(1), 96--121.
+\url{https://doi.org/10.1016/j.cogpsych.2009.02.002}
+
+\leavevmode\vadjust pre{\hypertarget{ref-samuelson2021precision}{}}%
+Samuelson, L. K. (2021). Toward a precision science of word learning:
+Understanding individual vocabulary pathways. \emph{Child Development
+Perspectives}, \emph{15}(2), 117--124.
+\url{https://doi.org/10.1111/cdep.12408}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sanchez2019childesdb}{}}%
+Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
+D., \& Frank, M. C. (2019). Childes-db: A flexible and reproducible
+interface to the child language data exchange system. \emph{Behavior
+Research Methods}, \emph{51}(4), 1928--1941.
 
 \leavevmode\vadjust pre{\hypertarget{ref-schelletter2002effect}{}}%
-Schelletter, C.
-(2002).
-The effect of form similarity on bilingual children's lexical development.
-\emph{Bilingualism: Language and Cognition}, \emph{5}(2), 93--107.
-\url{https://doi.org/10.1017/S1366728902000214}
+Schelletter, C. (2002). The effect of form similarity on bilingual
+children's lexical development. \emph{Bilingualism: Language and
+Cognition}, \emph{5}(2), 93--107.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schepens2012distributions}{}}%
+Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of
+cognates in europe as based on levenshtein distance. \emph{Bilingualism:
+Language and Cognition}, \emph{15}(1), 157--166.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schroter2016orthographic}{}}%
+Schröter, P., \& Schroeder, S. (2016). Orthographic processing in
+balanced bilingual children: Cross-language evidence from cognates and
+false friends. \emph{Journal of Experimental Child Psychology},
+\emph{141}, 239--246. \url{https://doi.org/10.1016/j.jecp.2015.09.005}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schwartz2007reading}{}}%
+Schwartz, A. I., Kroll, J. F., \& Diaz, M. (2007). Reading words in
+spanish and english: Mapping orthography to phonology in two languages.
+\emph{Language and Cognitive Processes}, \emph{22}(1), 106--129.
+\url{https://doi.org/10.1080/01690960500463920}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sebastian-galles2020bilingual}{}}%
+Sebastian-Galles, N., \& Santolin, C. (2020). Bilingual acquisition: The
+early steps. \emph{Annual Review of Developmental Psychology},
+\emph{2}(1), 47--68.
+\url{https://doi.org/10.1146/annurev-devpsych-013119-023724}
 
 \leavevmode\vadjust pre{\hypertarget{ref-smithson2014bilingualism}{}}%
-Smithson, L., Paradis, J., \& Nicoladis, E.
-(2014).
-Bilingualism and receptive vocabulary achievement: Could sociocultural context make a difference?
-\emph{Bilingualism: Language and Cognition}, \emph{17}(4), 810--821.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tardif2008baby}{}}%
-Tardif, T., Fletcher, P., Liang, W., Zhang, Z., Kaciroti, N., \& Marchman, V. A. (2008). Baby's first 10 words. \emph{Developmental Psychology}, \emph{44}(4), 929.
-
-\leavevmode\vadjust pre{\hypertarget{ref-thierry2007brain}{}}%
-Thierry, G., \& Wu, Y.
-J.
-(2007).
-Brain potentials reveal unconscious translation during foreign-language comprehension.
-\emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535.
-
-\leavevmode\vadjust pre{\hypertarget{ref-thordardottir2006bilingual}{}}%
-Thordardottir, E., Rothenberg, A., Rivard, M.-E., \& Naves, R.
-(2006).
-Bilingual assessment: Can overall proficiency be estimated from separate measurement of two languages?
-\emph{Journal of Multilingual Communication Disorders}, \emph{4}(1), 1--21.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999some}{}}%
-Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word comprehension in 6-month-olds. \emph{Psychological Science}, \emph{10}(2), 172--175.
-
-\leavevmode\vadjust pre{\hypertarget{ref-van2014subtlex}{}}%
-Van Heuven, W.
-J., Mandera, P., Keuleers, E., \& Brysbaert, M.
-(2014).
-SUBTLEX-UK: A new and improved word frequency database for british english.
-\emph{Quarterly Journal of Experimental Psychology}, \emph{67}(6), 1176--1190.
-
-\leavevmode\vadjust pre{\hypertarget{ref-vehtari2017practical}{}}%
-Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model evaluation using leave-one-out cross-validation and WAIC. \emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
-
-\leavevmode\vadjust pre{\hypertarget{ref-von2012language}{}}%
-Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access in bilingual toddlers. \emph{Journal of Experimental Child Psychology}, \emph{113}(4), 569--586.
-
-\leavevmode\vadjust pre{\hypertarget{ref-R-tidyverse}{}}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
-D., François, R., \ldots{} Yutani, H.
-(2019).
-Welcome to the {tidyverse}.
-\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
-\url{https://doi.org/10.21105/joss.01686}
+Smithson, L., Paradis, J., \& Nicoladis, E. (2014). Bilingualism and
+receptive vocabulary achievement: Could sociocultural context make a
+difference? \emph{Bilingualism: Language and Cognition}, \emph{17}(4),
+810--821. \url{https://doi.org/10.1017/S1366728913000813}
+
+\leavevmode\vadjust pre{\hypertarget{ref-spivey1999cross}{}}%
+Spivey, M. J., \& Marian, V. (1999). Cross talk between native and
+second languages: Partial activation of an irrelevant lexicon.
+\emph{Psychological Science}, \emph{10}(3), 281--284.
+
+\leavevmode\vadjust pre{\hypertarget{ref-swingley200511montholds}{}}%
+Swingley, D. (2005). 11-month-olds' knowledge of how familiar words
+sound. \emph{Developmental Science}, \emph{8}(5), 432--443.
+\url{https://doi.org/10.1111/j.1467-7687.2005.00432.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-swingley2000spoken}{}}%
+Swingley, D., \& Aslin, R. N. (2000). Spoken word recognition and
+lexical representation in very young children. \emph{Cognition},
+\emph{76}(2), 147--166.
+\url{https://doi.org/10.1016/S0010-0277(00)00081-0}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tamasi2017pupillometry}{}}%
+Tamási, K., McKean, C., Gafos, A., Fritzsche, T., \& Höhle, B. (2017).
+Pupillometry registers toddlers' sensitivity to degrees of
+mispronunciation. \emph{Journal of Experimental Child Psychology},
+\emph{153}, 140--148. \url{https://doi.org/10.1016/j.jecp.2016.07.014}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999beginnings}{}}%
+Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word
+comprehension in 6-month-olds. \emph{Psychological Science},
+\emph{10}(2), 172--175. \url{https://doi.org/10.1111/1467-9280.00127}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanheuven2014subtlexuk}{}}%
+Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
+{SUBTLEX}-{UK}: A new and improved word frequency database for british
+english. \emph{Quarterly Journal of Experimental Psychology},
+\emph{67}(6), 1176--1190.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonholzen2019impact}{}}%
+Von Holzen, K., Fennell, C. T., \& Mani, N. (2019). The impact of
+cross-language phonological overlap on bilingual and monolingual
+toddlers' word recognition. \emph{Bilingualism: Language and Cognition},
+\emph{22}(3), 476--499. \url{https://doi.org/10.1017/S1366728918000597}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonholzen2012language}{}}%
+Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access
+in bilingual toddlers. \emph{Journal of Experimental Child Psychology},
+\emph{113}(4), 569--586.
+\url{https://doi.org/10.1016/j.jecp.2012.08.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wells1995computercoding}{}}%
+Wells, J. C. (1995). \emph{Computer-coding the {IPA}: A proposed
+extension of {SAMPA}}. \emph{4}(28), 1995.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wewalaarachchi2017vowels}{}}%
+Wewalaarachchi, T. D., Wong, L. H., \& Singh, L. (2017). Vowels,
+consonants, and lexical tones: Sensitivity to phonological variation in
+monolingual mandarin and bilingual english--mandarin toddlers.
+\emph{Journal of Experimental Child Psychology}, \emph{159}, 16--33.
+\url{https://doi.org/10.1016/j.jecp.2017.01.009}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham2019welcome}{}}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
+Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
+Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
+Welcome to the tidyverse. \emph{Journal of Open Source Software},
+\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
 
 \leavevmode\vadjust pre{\hypertarget{ref-zipf1949human}{}}%
-Zipf, G. K. (1949). Human behavior and the principle of least effort: An introd. To human ecology.
+Zipf, G. K. (1949). \emph{Human behavior and the principle of least
+effort}. Addison-Wesley Press.
 
@@
 
-\endgroup
-
 \end{document}

### 2023-06-28 docs/index.qmd
Update PDF format

@@
 
-## Participants
+# Data
 
-```{r}
-#| label: fig-participants-time
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 10 
-#| fig-cap: "Date at which responses were collected"
-participants %>% 
-    group_by(time_stamp) %>% 
-    summarise(n = n(), .groups = "drop") %>% 
-    mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n)) +
-    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
-             angle = 90, vjust = 1.5, hjust = 1) +
-    geom_line(size = 1, colour = clrs[1]) +
-    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
-    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
-    theme(legend.position = "top",
-          legend.title = element_blank(),
-          panel.grid.minor = element_blank(),
-          axis.text.x = element_text(angle = 20, hjust = 1),
-          axis.title.x = element_blank()) +
+# Outputs
 
-    participants %>% 
-    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
-    group_by(time_stamp, age) %>% 
-    summarise(n = n(), .groups = "drop") %>% 
-    group_by(age) %>% 
-    mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n, colour = age)) +
-    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    geom_line(size = 1) +
-    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
-    scale_y_continuous(breaks = seq(0, 500, 50)) +
-    scale_colour_manual(values = clrs) +
-
-    plot_layout(ncol = 1) &
-
-    scale_x_date(date_breaks = "3 month", date_labels = "%B %Y") &
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 20, hjust = 1),
-        axis.title.x = element_blank(),
-    )
-```
-
-### Age
-
-Number of months elapsed between participants' birth date and questionnaire completion.
-
-```{r}
-#| label: fig-participants-age
-#| fig-width: 7
-#| fig-height: 3
-#| fig-cap: "Distribution of participants' ages"
-participants %>% 
-    mutate(age = floor(age)) %>% 
-    count(age) %>% 
-    ggplot() +
-    aes(x = age, y = n) +
-    geom_col(fill = "grey60") +
-    geom_text(aes(label = n), size = 3.5, vjust = -1) +
-    labs(x = "Age (months)",
-         y = "# participants") +
-    scale_x_continuous(breaks = seq(0, 40, 2)) +
-    scale_y_continuous(limits = c(0, 55),
-                       breaks = seq(0, 55, 10)) +
-    theme(
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
-
-### Language exposure (Degree of Exposure)
-
-Participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to.
-For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.
-
-```{r fig-participants}
-#| label: tab-participants
-#| message: false
-#| warning: false
-#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan.
-Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
-In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
-A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
-A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
-The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
-For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins.
-Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
-participants |>
-    mutate_at(vars(starts_with("doe_")), 
-              (\(x) floor(x * 10) / 10)) |> 
-    mutate(
-        age = cut(floor(age),
-                  c(10, 14, 18, 22, 26, 30, 34, 36),
-                  include.lowest = TRUE),
-        doe_catalan = cut(
-            doe_catalan,
-            c(0, 0.25, 0.5, 0.75, 1),
-            labels = c("0-25%",
-                       "25-50%", 
-                       "50-75%",
-                       "75-100%"),
-            include.lowest = TRUE
-        ) |>
-            as.character()
-    ) |>
-    count(age, doe_catalan) |>
-    arrange(desc(doe_catalan)) |>
-    pivot_wider(names_from = age, values_from = n) |>
-    gt() |>
-    sub_missing(everything(), everything(), missing_text = "--") |>
-    cols_label(doe_catalan = "Catalan exposure") |>
-    tab_spanner("Age (months)", 2:7) |>
-    tab_style(cell_text(align = "left"),
-              list(cells_body(doe_catalan),
-                   cells_column_labels(doe_catalan)))
-```
-
-### SES/parental education
-
-```{r}
-#| label: tbl-participants-edu
-#| tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)."
-participants %>%
-    mutate(id = as.character(id)) |> 
-    left_join(bvq_data$logs,
-              by = join_by(id, time, age, lp, doe_catalan, doe_spanish, edu_parent)) %>%
-    mutate(
-        edu_parent = edu_parent |> 
-            as.factor() |> 
-            fct_na_value_to_level(),
-        doe_catalan = cut(doe_catalan,
-                          breaks = seq(0, 1, 0.2),
-                          include.lowest = TRUE)
-    ) %>%
-    count(doe_catalan, edu_parent) %>%
-    right_join(
-        expand_grid(
-            doe_catalan = levels(.$doe_catalan),
-            edu_parent = unique(bvq_data$logs$edu_parent)
-        ),
-        by = join_by(doe_catalan, edu_parent)
-    ) %>%
-    rename(total = n) %>%
-    pivot_wider(names_from = edu_parent,
-                values_from = total,
-                values_fill = 0) %>%
-    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>%
-    rename(Total = total) %>%
-    select(-"NA") %>%
-    gt() |> 
-    cols_label(doe_catalan = "DoE Catalan") %>%
-    sub_missing(columns = everything(),
-                missing_text = "--") %>%
-    tab_spanner("Educational attainment", 
-                Secondary:Primary) %>%
-    tab_style(
-        style = list(cell_borders(weight = px(2),
-                                  sides = "left", 
-                                  color = "grey50")),
-        locations = cells_body(columns = 8,
-                               rows = 1:5)
-    ) %>%
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(columns = 2:7)
-    ) %>%
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = c(1, 8))
-    ) %>%
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(spanners = "Educational attainment")
-    )
-```
-
-## Items
-
-```{r}
-#| label: tbl-items-summary
-#| tbl-cap: "Lexical frequencies.
-Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
-items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))
-
-items %>%
-    unnest(list) %>%
-    summarise(n = n(),
-              across(c(freq, n_phon, lv),
-                     lst(mean, sd, min, max)),
-              .by = c(language, list)) |> 
-    gt(groupname_col = "language", 
-       rowname_col = "list") %>%
-    fmt_number(-c(list, n),
-               decimals = 1) %>%
-    tab_spanner("Frequency (Zipf)",
-                columns = starts_with("freq_")) %>%
-    tab_spanner("Length (Phonemes)",
-                columns = starts_with("n_phon_")) %>%
-    tab_spanner("Levenshtein (cognateness)", 
-                columns = starts_with("lv_")) %>%
-    cols_merge_uncert(col_val = freq_mean,
-                      col_uncert = freq_sd) %>%
-    cols_merge_uncert(col_val = n_phon_mean,
-                      col_uncert = n_phon_sd) %>%
-    cols_merge_uncert(col_val = lv_mean, 
-                      col_uncert = lv_sd) %>%
-    cols_merge_range(col_begin = freq_min, 
-                     col_end = freq_max) %>%
-    cols_merge_range(col_begin = n_phon_min, 
-                     col_end = n_phon_max) %>%
-    cols_merge_range(col_begin = lv_min,
-                     col_end = lv_max) %>%
-    cols_label(list = "",
-               n = md("*N*"),
-               freq_mean = md("Mean ± SD"),
-               freq_min = md("Range"),
-               n_phon_mean = md("Mean ± SD"),
-               n_phon_min = md("Range"),
-               lv_mean = md("Mean ± SD"),
-               lv_min = md("Range")) %>%
-    tab_style(cell_text(style = "italic"),
-              cells_column_labels(everything())) %>%
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners(everything()))
-```
-
-### Lexical frequency
-
-Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
-Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [`childesr`](https://langcog.github.io/childes-db-website/) R package.
-Available words in the corpora were then mapped to their Catalan and Spanish translations.
-Responses to words with missing lexical frequencies were excluded from analyses.
-
-```{r}
-#| label: fig-items-freq
-#| message: false
-#| warning: false
-#| fig-width: 6
-#| fig-height: 3
-#| fig-cap: "Distribution of lexical frequency across items.
-Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations.
-Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency.
-The black dot and interval indicate the mean and standard deviation."
-mean_sd <- function(x, ...) {
-    m <- mean(x, ...)
-    s <- sd(x, ...)
-    data.frame(mean = m, ymin = m-s, ymax = m+s)
-}
-
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    ggplot(aes(x = freq)) +
-    geom_histogram(fill = "grey60",
-                   colour = "white",
-                   bins = 30)  +
-    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        legend.title = element_blank()
-    )
-```
-
-### Word length
-
-Word length ($Length$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.
-
-```{r}
-#| label: fig-items-nphon
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Distribution of number of phonemes across items"
-items %>%
-    count(language, n_phon) %>%
-    ggplot(aes(x = n_phon, y = n)) +
-    facet_wrap(~language, ncol = 1) +
-    geom_col(fill = "grey60", colour = "white") +
-    geom_text(aes(label = n),
-              position = position_nudge(y = 5),
-              colour = "black") +
-    labs(x = "Phonemes",
-         y = "Number of words",
-         fill = "Language",
-         shape = "Language",
-         colour = "Language") +
-    scale_x_continuous(breaks = 1:15) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    theme(legend.position = "none",
-          panel.grid.major.x = element_blank(),
-          panel.grid.minor = element_blank(),
-          axis.ticks.x = element_line(),
-          legend.title = element_blank())
-```
-
-### Levenshtein (cognateness)
-
-Phonological similarity between translation equivalents was computed as the normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary].
-This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1.
-This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018vocabulary; @fourtassi2020growth; @laing2022phonological for similar approaches].
-
-```{r}
-#| label: fig-items-lv
-#| message: false
-#| warning: false
-#| fig-width: 9
-#| fig-cap: "Distribution of Levenshtein similarities across items"
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
-    count(lv) %>% 
-    mutate(prop = n/sum(n),
-           labels = paste0(
-               n, " (",
-               scales::percent(prop, accuracy = 0.1), 
-               ")")) %>% 
-    ggplot(aes(x = lv, n)) +
-    geom_col(fill = "grey60", colour = "white") +
-    geom_text(aes(label = labels),
-              position = position_nudge(y = 3), 
-              size = 3.5) +
-    labs(x = "Levenshtein distance", y = "Number of words") +
-    theme(
-        panel.grid.major.x = element_blank(),
-        legend.position = "top"
-    )
-```
-
-### Frequency by word length
-
-```{r}
-#| label: fig-items-frequency-phonemes
-#| message: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 4
-#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
-items %>% 
-    drop_na(freq) %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    count(n_phon, freq) %>% 
-    ggplot(aes(n_phon, freq)) +
-    geom_point(size = 2,
-               alpha = 0.5,
-               colour = "grey60", 
-               shape = 1, 
-               stroke = 1,
-               position = position_jitter(width = 0.1)) +
-    geom_smooth(method = "lm",
-                formula = "y ~ x" ,
-                colour = "black") +
-    labs(
-        x = "# Phonemes", 
-        y = "Lexical frequency (Zipf score)", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    ) 
-```
-
-### Frequency by Levenshtein
-
-```{r}
-#| label: fig-items-frequency-lv
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Association betwen lexical frequency and phonological similarity"
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    drop_na(freq) %>% 
-    count(freq, lv) %>% 
-    ggplot(aes(freq, lv)) +
-    geom_point(size = 2,
-               shape = 1, 
-               colour = "grey60",
-               stroke = 1) +
-    geom_smooth(method = "lm", 
-                formula = "y ~ x", 
-                colour ="black") +
-    labs(y = "Levenshtein similarity", 
-         x = "Lexical frequency (Zipf score)", 
-         colour = "Language", 
-         fill = "Language") +
-    scale_y_continuous(labels = scales::percent) +
-    theme(
-        legend.position = "none",
-        panel.grid.major.y = element_blank()
-    )
-```
-
-### Word length by Levenshtein
-
-```{r}
-#| label: fig-items-nphon-lv
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Association between number of phonemes and phonological similarity"
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    drop_na(freq) %>% 
-    count(n_phon, lv) %>% 
-    ggplot(aes(n_phon, lv, size = n)) +
-    geom_point(shape = 1,
-               stroke = 1,
-               colour = "grey60") +
-    geom_smooth(method = "lm",
-                formula = "y ~ x",
-                size = 1,
-                colour = "black") +
-    labs(x = "# Phonemes", 
-         y = "Levenshtein similarity", 
-         colour = "Language", 
-         fill = "Language") +
-    scale_y_continuous(labels = scales::percent) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(legend.position = "none",
-          panel.grid.minor = element_blank(),
-          panel.grid.major.x = element_blank()) 
-```
-
-### Word exposure index by lexical frequency
-
-We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
-the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion) (see @eq-exposure).
-
-$$
-\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
-$$ {#eq-exposure}
-
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`) would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)` (see @eq-exposure-example):
-
-$$
-\begin{aligned}
-\textbf{Exposure}_{\textbf{cavall, Catalan}} &= \textbf{Frequency}_{cavall} \times \textbf{DoE}_{Catalan} = 6.14 \times 0.8 = 4.91 \\
-\textbf{Exposure}_{\textbf{caballo, Spanish}} &= \textbf{Frequency}_{caballo} \times \textbf{DoE}_{Spanish} = 6.14 \times 0.2 = 1.23
-\end{aligned}
-$$ {#eq-exposure-example}
-
-```{r}
-#| label: fig-exposure-distribution
-#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
-#| warnings: false
-#| fig-height: 7
-#| fig-width: 13
-responses %>%
-    mutate(
-        exposure = cut(exposure,
-                       breaks = seq(0, 9, 1),
-                       include.lowest = TRUE)) %>%
-    count(exposure, language) %>%
-    ggplot(aes(exposure, n)) +
-    facet_wrap( ~ language) +
-    geom_col(colour = "white",
-             fill = "grey60") +
-    labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
-         y = "Number of observations",
-         fill = "Language") +
-
-    responses %>%
-    mutate(exposure = cut(exposure,
-                          breaks = seq(0, 9, 1),
-                          include.lowest = TRUE),
-           freq = cut(freq, 
-                      breaks = seq(0, 9, 0.25), 
-                      include.lowest = TRUE)) %>%
-    count(exposure, freq, language) %>%
-    ggplot(aes(x = freq,
-               y = exposure,
-               size = n)) +
-    facet_wrap( ~ language) +
-    geom_point(colour = "grey60",
-               alpha = 1/2) +
-    labs(x = "Frequency (Zipf score)",
-         y = "Exposure-weighted frequency\\
-         (Frequency \u00d7 DoE)",
-         colour = "Language",
-         fill = "Language",
-         size = "Number of observations") +
-    theme(panel.grid = element_blank()) +
-    plot_layout(ncol = 1) &
-    scale_colour_manual(values = clrs[c(1, 4)]) &
-    scale_fill_manual(values = clrs[c(1, 4)]) &
-    guides(fill = "none", colour = "none") &
-    theme(panel.grid.major.x = element_blank(),
-          panel.grid.minor.y = element_blank(),
-          legend.position = "none")
-```
-
-## Data analysis
-
-We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function.
-The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood (see @fig-likelihood for a graphical representation).
-
-$$
-\begin{aligned}
-\textbf{Likelihood:} \\
-y_{ij} &\sim \text{Cumulative}(p_{k})
-\end{aligned}
-$$ {#eq-likelihood}
-
-where:
-
-- $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$)
-- $i$ is the participant index
-- $j$ is the translation equivalent (TE) index
-- $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$)) between two response categories in the latent distribution
-$p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.
-
-```{r}
-#| label: fig-likelihood
-#| fig-cap: "Graphical representation of the assumed latent distribution that generates observed responses in the questionnaire.
-Cumulative ordinal regression assumes that observed categorical responses (*No*, *Understands* and *Understands and Says* are generated from a common latent (i.e., unobserved) distribution.
-For threotical reasons, we refer to the dimension captured by such distribution as *Familiarity*.
-*Familiarity* describes a given participant's familiarity with a given word.
-When familiarity is low (below some threshold 1), the reulting response is *No*.
-When familiarity is higher than such threshold 1, but lower than some threshold 2, the resulting result is *Understands* (i.e., comprehension).
-When familiarity exceeds threshold 2, the resulting response is *Understands and Says* (i.e., comprehension and production).
-These thresholds are estimated in the cumulative regression model as intercepts 1, and 2.
-The rest of the coefficients in the regression model shift the response thresholds, making comprehension or production more likely or less likely.
-See @burkner2019ordinal for a detailed description of cumulative ordinal regression and its implementation."
-#| fig-width: 7
-#| fig-height: 3.5
-dist_df <- tribble(~mean, ~sd, 0,   1)
-
-thresholds <- data.frame(
-    ymin = c(-Inf, -1.5, 1.5),
-    ymax = c(-1.5, 1.5, Inf),
-    xmin = -Inf,
-    xmax = Inf,
-    .category = str_wrap(c(
-        "No", 
-        "Understands",
-        "Understands and Says"
-    ),
-    width = 12)
-)
-
-thresholds %>%
-    ggplot(aes(ymin = ymin, ymax = ymax,
-               xmin = xmin, xmax = xmax,
-               fill = .category)) +
-    geom_rect(alpha = 0.5) +
-    stat_dist_slab(
-        data = dist_df, 
-        aes(dist = "norm", arg1 = mean, arg2 = sd),
-        position = "dodge", inherit.aes = FALSE,
-        fill = NA, colour = "black"
-    ) +
-    geom_hline(yintercept = c(-1.5, 1.5), 
-               linetype = "dotted",
-               linewidth = 1) +
-    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
-              vjust = 1, hjust = 0.5) +
-    annotate(
-        geom = "text",
-        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
-        y = c(-1.5, 2), parse = TRUE,
-        x = 0.5, angle = 90, vjust = -0.5
-    ) +
-    labs(y = "Hypothetical response-generating latent distribution",
-         x = "Response probability",
-         fill = "Response type",
-         colour = "Response type") +
-    coord_flip() +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_x_continuous(labels = scales::percent) +
-    scale_y_continuous(limits = c(-4, 4)) +
-    theme(panel.grid = element_blank(),
-          legend.position = "none")
-```
-
-To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), number of phonemes ($Phonemes$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$.
-We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset--each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
-For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].
-
-$$
-\begin{aligned}
-\textbf{Linear model:} \\
-logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
-& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
-& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Phonemes}_{ij} + & \\
-& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
-& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
-& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
-& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
-& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
-& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
-\end{aligned}
-$$ {#eq-linear}
-
-where:
-
-- $i$ and $j$ index the participant and translation equivalent (TE)
-- $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$
-- $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively
-- $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest
-- $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively
-
-We used the Bayesian framework to estimate the parameters in our model.
-This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical].
-This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations.
-We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].
-
-$$
-\begin{aligned}
-\\
-\textbf{Prior:} \\
-\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
-\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
-\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
-u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
-w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
-&&\mbox{[Participant-level coefficient variability]} \\ \\
-\Bigg(\begin{smallmatrix}
-u_{k_{0}} \\ 
-u_{1_{i}} \\ 
-\vdots \\ 
-u_{8_{i}} 
-\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
-\Bigg(\Bigg(\begin{smallmatrix}0 \\
-0 \\ 
-\vdots \\
-0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
-\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
-\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
-\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
-\vdots & \vdots & \vdots & \vdots \\
-\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
-\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
-\rho_{u} &\sim LKJcorr(2) \\
-\\
-&&\mbox{[TE-level coefficient variability]} \\ \\
-\Bigg(\begin{smallmatrix}
-w_{k_{0}}\\ 
-w_{1_{j}} \\ 
-\vdots \\ 
-w_{3_{j}} 
-\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
-0\\ 
-0 \\ 
-\vdots \\
-0 
-\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
-\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
-\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
-\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
-\vdots & \vdots & \vdots & \vdots \\
-\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
-\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
-\rho_{w_{0-3}} &\sim LKJcorr(2)
-\end{aligned}
-$$ {#eq-prior}
-
-where:
-
-- $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively
-- $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively
-- $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].
-
-::: {.panel-tabset}
-
-#### Prior-predicted mean
-
-```{r fig_prior-mean}
-#| label: fig-prior-mean
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 3
-#| out-width: 90%
-#| fig-cap: "Expected prior-predicted mean."
-nd <- expand.grid(n_phon_std = 0,
-                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
-                  exposure_std = 0,
-                  lv_std = 0)
-
-prior_epreds <- add_epred_draws(nd, 
-                                model_fit_prior,
-                                ndraws = 50, 
-                                re_formula = NA) |> 
-    filter(.category != "No") |> 
-    pivot_wider(names_from = ".category", values_from = ".epred") |> 
-    mutate(Understands = Understands + `Understands and Says`) |> 
-    pivot_longer(c(Understands, `Understands and Says`),
-                 names_to = ".category",
-                 values_to = ".epred") |> 
-    mutate(.category = factor(.category,
-                              levels = c("Understands",
-                                         "Understands and Says"),
-                              labels = c("Comprehension\n(Understands)",
-                                         "Production\n(Understands and Says)")))
-
-prior_epreds |> 
-    ggplot(aes(x = age_std, y = .epred)) +
-    facet_wrap(~ .category) +
-    geom_line(linewidth = 3/4,
-              alpha = 1/2,
-              aes(group = .draw),
-              colour = "grey50") + 
-    labs(x = "Age (months)", 
-         y = "Probability of acquisition",
-         colour = "Credible interval", 
-         fill = "Credible interval") +
-    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
-                                      mean(responses$age), 
-                                      sd(responses$age))[, 1],
-                       labels = seq(10, 36, 4)) +
-    theme(legend.position = "top",
-          axis.ticks.x = element_line(colour = "black"),
-          panel.grid = element_blank())
-```
-
-#### Prior-predictions
-
-```{r}
-#| label: fig-prior-predictions
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected prior predictions"
-nd <- expand.grid(
-    n_phon_std = 0,
-    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
-    exposure_std = 0,
-    lv_std = 0
-)
-
-m <- add_predicted_draws(nd, model_fit_prior, ndraws = NULL, re_formula = NA)
-
-m %>% 
-    count(age_std, .prediction) %>% 
-    pivot_wider(names_from = .prediction, values_from = n) %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(c(No, Understands, `Understands and Says`),
-                 names_to = ".prediction", 
-                 values_to = "n") %>%
-    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
-    geom_col(position = position_fill()) +
-    labs(
-        x = "Age (months)", 
-        y = "Probability of acquisition",
-        colour = "Category", 
-        fill = "Category"
-    ) +
-    scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
-        labels = seq(10, 36, 4)
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(labels = scales::percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        strip.background = element_rect(fill = "grey", colour = NA),
-        plot.background = element_rect(fill = "white", colour = NA)
-    )
-```
-
-:::
-
-### Model settings
-
--   OS: `r sessionInfo()$running`, `r sessionInfo()$platform`
--   R version: `r sessionInfo()$R.version$version.string`
--   Algorithm: Hamiltonian Monte Carlo (No U-turn Sampler)
--   Engine: Stan (`brms` interface, `r sessionInfo()$otherPkgs$brms$Version`)
--   `brms` backend: CmdStanR (`cmdstanr` `r sessionInfo()$loadedOnly$cmdstanr$Version`)
--   Chains: `r dim(model_fit$fit)[2]`
--   Cores: `r dim(model_fit$fit)[2]`
--   Iterations: `r format(dim(model_fit$fit)[1], big.mark = ",")` (`r format(dim(model_fit$fit)[1]/2, big.mark = ",")`)
-
-More details in Appendix: Session Info.
-
-### R code (`brms`)
-
-```{r}
-print(model_fit$formula)
-```
-:::
-
-# Results
-
-## Descriptive statistics
-
-::: panel-tabset
-
-### By number of phonemes
-
-```{r}
-#| lab: fig-responses-age-phonemes
-#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
-responses %>% 
-    mutate(
-        age = floor(age),
-        n_phon = cut_interval(n_phon, n = 4)
-    ) %>%
-    count(age, n_phon, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~n_phon) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(labels = scales::percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-
-### By exposure
-
-```{r}
-#| lab: fig-responses-age-exposure
-#| fig-cap: "Proportion of responses to each category by age and exposure."
-responses %>% 
-    mutate(
-        age = floor(age),
-        exposure = cut_interval(exposure, n = 4)
-    ) %>%
-    count(age, exposure, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~exposure) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
-    scale_y_continuous(labels = scales::percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-
-### By Levenshtein
-
-```{r}
-#| lab: fig-responses-age-lv
-#| fig-cap: "Proportion of responses to each category by age and Levenhstein similarity"
-responses %>%
-    mutate(
-        age = floor(age),
-        exposure = cut_interval(exposure, n = 4, labels = paste0("Q", 1:4)),
-        lv = cut_interval(lv, n = 4)
-    ) %>%
-    group_by(te, lv, exposure, response) %>%
-    count(lv, exposure, response) %>%
-    pivot_wider(names_from = response,
-                values_from = n,
-                values_fill = 0) %>%
-    mutate(
-        `Understands` = `Understands` + `Understands and Says`,
-        n_total = sum(No, `Understands`, `Understands and Says`)
-    ) %>%
-    pivot_longer(
-        c(`No`, `Understands`, `Understands and Says`),
-        names_to = "response",
-        values_to = "n"
-    ) %>%
-    filter(response != "No") %>%
-    mutate(prop = n / n_total) %>%
-    # group_by(age, lv, exposure, response) %>%
-    # summarise(prop_mean = mean(prop), .groups = "drop") %>%
-    ggplot(aes(exposure, prop, colour = lv, fill = lv)) +
-    facet_wrap(~ response, ncol = 1) +
-    geom_hline(yintercept = 0.5, linetype = "dashed") +
-    stat_dots(aes(group = lv),
-              position = position_dodge(width = 1),
-              layout = "swarm",
-              side = "both",
-              colour = "grey80",
-              fill = "grey80"
-    ) +
-    stat_summary(geom = "errorbar",
-                 fun.data = "mean_se",
-                 position = position_dodge(width = 1),
-                 width = 0.25) +
-    stat_summary(geom = "point",
-                 fun = "mean",
-                 position = position_dodge(width = 1),
-                 size = 2) +
-    labs(y = "Proportion of responses", 
-         x = "Exposure",
-         fill = "Levenshtein", 
-         colour = "Levenshtein") +
-    scale_colour_manual(values = clrs[c(1, 4, 3, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 4, 3, 5)]) +
-
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-:::
-
-## Fixed effects
-
-```{r tbl-coefs}
-#| label: tbl-coefs
-#| tbl-cap: "Posterior distribution of regression coefficients."
-#| tbl-note: "Median: median of the posterior distribution in the probability scale.
-95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.
-On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE).
-This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
-Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
-#| message: false
-#| warning: false
-#| out-width: 80%
-# summarise posterior draws
-intercept_label <- "Intercepts (at 22 months)"
-
-model_summary |>
-    select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
-    mutate(across(.median:.upper,
-                  \(x) ifelse(.type==intercept_label, plogis(x), x/4))) |> 
-    gt(groupname_col = ".type", 
-       rowname_col = ".variable_name") |>
-    fmt_percent(.rope_overlap, decimals = 3) |> 
-    fmt_number(c(.median:.upper), decimals = 3) |>
-    cols_merge(c(.lower, .upper),
-               pattern = "[{1}, {2}]") |>
-    cols_label(
-        .variable_name = "Parameter",
-        .median = "Median",
-        .lower = md("95% HDI"),
-        .rope_overlap = md("*p*(H0)")
-    ) |>
-    tab_style(cell_text(weight = "bold"),
-              cells_column_labels(columns = 1:5)) |>
-    tab_style(cell_text(align = "left"),
-              cells_title(groups = "subtitle")) |> 
-    tab_style(cell_text(size = "small"), 
-              list(cells_body(),
-                   cells_row_groups(),
-                   cells_stub())) |> 
-    tab_style(cell_text(style = "italic"),
-              cells_row_groups())
-```
-
-```{r}
-#| label: fig-results-fixed
-#| fig-width: 9
-#| fig-heigh: 7
-#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model.
-Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis).
-Intervals show the 95% highest density interval (*HDI*) of each distribution.
-This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-the mean and HDI limits are indicated below each distribution.
-Coefficients were transformed from the logit scale to the probability scale for interpretability.
-We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e.
-the derivative of the logistic function in its mid-point."
-
-model_draws %>%
-    ggplot(aes(.value, .variable_name)) +
-    annotate(geom = "rect", ymin = -Inf, ymax = Inf,
-             xmin = rope_interval["lower"],
-             xmax = rope_interval["upper"],
-             colour = NA, alpha = 0.2, fill = clrs[5],) +
-    geom_vline(xintercept = 0, size = 1, colour = clrs[5]) +
-    stat_slab(
-        aes(fill = stat(abs(x) < rope_interval["upper"]),
-            colour = stat(abs(x) < rope_interval["upper"]),),
-        size = 1, 
-        position = position_nudge(y = 0.15)
-    ) +
-    geom_errorbar(
-        data = model_summary,
-        aes(xmin = .lower, 
-            xmax = .upper, 
-            x = .median),
-        width = 0.15
-    ) +
-    geom_point(data = model_summary, 
-               aes(x = .median),
-               size = 1.5) +
-    geom_text(
-        data = model_summary,
-        aes(x = .median, 
-            y = .variable_name,
-            label = glue::glue(
-                "{round(.median, 1)}",
-                "[{round(.lower, 1)}, {round(.upper, 1)}]",
-                .sep = " ")),
-        colour = "black",
-        position = position_nudge(y = -0.3), size = 3
-    ) +
-    labs(
-        x = "Logistic regression coefficient estimate (Logit scale)",
-        y = "Variable", 
-        fill = "Overlaps with ROPE",
-        colour = "Overlaps with ROPE"
-    ) +
-    scale_x_continuous(breaks = seq(-3, 3, 0.2)) +
-    scale_fill_manual(
-        values = clrs[c(1, 4)],
-        labels = c("No", "Yes")
-    ) +
-    scale_colour_manual(
-        values = clrs[c(1, 4)],
-        labels = c("No", "Yes")
-    ) +
-    theme(
-        legend.position = "top",
-        legend.justification = c(1, 1),
-        axis.title.y = element_blank(),
-        panel.grid.major.x = element_line(
-            colour = "grey85",
-            linetype = "dotted"
-        ),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
-
-## Marginal effects
-
-```{r fig-marginal}
-#| label: fig-marginal
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 4.5
-#| out-width: 100%
-#| eval: true
-#| fig-cap: "Posterior marginal effects.
-Thin lines correspond to 50 mean predictions.
-Thicker lines indicate the median of those predictions.
-Different colours indicate different levels of cognateness (phonological similarity).
-Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
-Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom.
-In-sample predictions lie inside the grey rectangles."
-clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")
-
-model_epreds |> 
-    mutate(age = (age_std * sd(responses$age)) + mean(responses$age),
-           exposure = factor(exposure_std,
-                             levels = unique(exposure_std),
-                             labels = c("Low exposure (-1 SD)",
-                                        "Mean exposure",
-                                        "High exposure (+1 SD)"),
-                             ordered = TRUE),
-           lv = factor(lv_std,
-                       levels = unique(lv_std),
-                       labels = paste0(c(0, 50, 100), "%"),
-                       ordered = TRUE),
-           .category = ifelse(.category=="Understands",
-                              "Comprehension",
-                              "Production")) |> 
-    ggplot(aes(age, .value, colour = lv, fill = lv)) +
-    facet_grid(.category~exposure) +
-    annotate(geom = "rect",
-             xmin = min(responses$age),
-             xmax = max(responses$age),
-             ymin = -Inf,
-             ymax = Inf,
-             fill = "grey95",
-             colour = NA) +
-    annotate(geom = "text",
-             x = mean(responses$age),
-             y = 0,
-             hjust = -0.1,
-             vjust = 0,
-             label = "Observed age range",
-             size = 2) +
-    geom_vline(xintercept = mean(responses$age), 
-               linetype = "dotted") +
-    geom_hline(yintercept = 0.5,
-               linetype = "dotted") +
-    geom_line(aes(group = interaction(lv, .draw)),
-              linewidth = 0.5,
-              alpha = 1/10) +
-    stat_summary(fun = median, geom = "line", linewidth = 0.75) +
-    labs(x = "Age (months)",
-         y = "Probability of acquisition",
-         linetype = "Cognateness (phonological similarity)",
-         colour = "Cognateness (phonological similarity)",
-         fill = "Cognateness (phonological similarity)") +
-    guides(colour = guide_legend(override.aes = list(linewidth = 2))) +
-    scale_colour_manual(values = clrs_reorder[c(4, 5, 1)]) +
-    scale_fill_manual(values = clrs_reorder[c(4, 5, 1)]) +
-    scale_y_continuous(breaks = seq(0, 1, 0.25),
-                       limits = c(0, 1)) +
-    scale_x_continuous(breaks = seq(0, 50, 5)) +
-    theme(legend.position = "top",
-          legend.key.size = unit(1, "cm"),
-          legend.key.height = unit(1, "cm"),
-          axis.text.x = element_text(size = 9),
-          panel.grid = element_blank(),
-          panel.border = element_rect(fill = NA, 
-                                      colour = "black", 
-                                      linewidth = 0.75))
-```
-
-## Marginal effects
-
-In order to examine the impact of the word exposure index on the effect of phonological similarity we used the [`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects/index.html) R package to compute the average marginal effect of the Levenshtein distance on the probability of acquisition (comprehension or production) for the range of observed values of the word exposure index in our dataset.
-
-```{r}
-#| label: fig-marginal-effect
-#| fig-cap: "Average marginal effect of phonological similarity (Levenshtein distance) on the probability of comprehension and production across the range of values of exposure.
-The observed range of values of the word exposure index is indicated in the X-axis, and the estimated average marginal effect of phonological similarity (Levenshtein distance) is indicated in the Y-axis.
-This quantity is generated by averaging the derivative of the regression coefficient of the predictor Levenshtein across the range of values of all predictors but Exposure.
-We represent the uncertainty associated with the resulting estimate by generating multiple values of this marginal effect by drawing multiple samples from the posterior distribution of the model coefficients, calculating the average marginal effect for each posterior draw, and plotting the result as a thin grey line.
-Closer lines indicate lower uncertainty over the true value of the average marginal effect of Levenshein, while more distant lines indicate higher uncertainty."
-nd <- marginaleffects::datagrid(
-    model = model_fit,
-    n_phon_std = 0,
-    lv_std = 0,
-    exposure_std = seq(-1.5, 1.5, length.out = 100),
-    age_std = 0,
-    id = NA,
-    te = NA
-)
-
-marg_effects <- marginaleffects::marginaleffects(
-    model_fit,
-    newdata = nd, 
-    re_formula = NA,
-    ndraws = 50,
-    vcov = FALSE,
-    type = "link"
-)
-
-marg_effects %>% 
-    marginaleffects::posteriordraws() %>%
-    as_tibble() %>% 
-    filter(term=="lv_std") %>% 
-    mutate(draw = draw/4) %>% 
-    ggplot(aes(exposure_std, draw)) +
-    geom_hline(yintercept = 0, size = 1, colour = "grey") +
-    geom_line(aes(group = drawid), 
-              alpha = 1, 
-              size = 0.75,
-              colour = "grey") +
-    # stat_summary(fun.data = median_hdi, geom = "ribbon",
-    #              colour = NA, alpha = 0.5, fill = clrs[1]) +
-    stat_summary(fun = mean, 
-                 geom = "line", 
-                 size = 1, 
-                 color = clrs[1]) +
-    labs(
-        x = "Word exposure index",
-        y = "Marginal effect of phonological similarity\n(Levenshtein distance)"
-    ) +
-    scale_x_continuous(
-        labels = scales::label_number(style_positive = "plus", suffix = " SD"),
-        breaks = seq(-2, 2, 0.5)
-    ) +
-    scale_y_continuous(
-        labels = scales::label_percent(style_positive = "plus", ),
-        breaks = seq(-1, 1, 0.05)
-    ) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
-    )
-
-```
-
-## Random effects
-
-### Participant-level effects
-
-::: panel-tabset
-
-#### Intercept
-
-```{r}
-#| label: fig-results-random-id-intercept
-#| fig-width: 7
-#| fig-cap: "Posterior probability of acquisition by participant"
-re_id <- ranef(model_fit)$id[,,1] %>% 
-    as_tibble() %>% 
-    rownames_to_column("id") %>%
-    left_join(distinct(model_fit$data, id),
-              by = join_by(id)) %>% 
-    janitor::clean_names() %>% 
-    mutate(
-        across(
-            .cols = estimate:q97_5,
-            .fns = list(
-                Comprehension = ~plogis(. + fixef(model_fit)[1]),
-                Production = ~plogis(. + fixef(model_fit)[2])
-            ), 
-            .names = "{.col}__{.fn}"
-        )
-    ) %>% 
-    select(id, ends_with("Comprehension"), ends_with("Production")) %>% 
-    pivot_longer(
-        -id, 
-        names_to = c(".value", "type"),
-        names_sep = "__"
-    )
-
-re_id %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate), 
-        xmin = q2_5, xmax = q97_5,
-        fill = type, colour = type) +
-    facet_wrap(~type) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "Participant", 
-        colour = "Type"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age slope
-
-```{r}
-#| label: fig-results-random-id-age
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_age <- ranef(model_fit)$id[,,"age_std"]
-
-re_age %>% 
-    as_tibble() %>% 
-    janitor::clean_names() %>% 
-    mutate(id = unique(model_fit$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### # Phonemes slope
-
-```{r}
-#| label: fig-results-random-id-nphon
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the number of phonemes by participant"
-re_n_phon <- ranef(model_fit)$id[,,"n_phon_std"]
-
-re_n_phon %>% 
-    as_tibble() %>% 
-    janitor::clean_names() %>% 
-    mutate(id = unique(model_fit$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(
-        x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5
-    ) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Exposure slope
-
-```{r}
-#| label: fig-results-random-id-exposure
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of lexical frequency by participant"
-re_freq <- ranef(model_fit)$id[,,"exposure_std"]
-
-re_freq %>% 
-    as_tibble() %>% 
-    janitor::clean_names() %>% 
-    mutate(id = unique(model_fit$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-lv
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of phonological similarity by participant"
-re_lv <- ranef(model_fit)$id[,,"lv_std"]
-
-re_lv %>%
-    as_tibble() %>%
-    janitor::clean_names() %>%
-    mutate(id = unique(model_fit$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Doe-by-Levenshtein slope
-
-```{r}
-# | label: fig-results-random-id-exposure-lv
-# | message: false
-# | warning: false
-# | fig-width: 7
-# | fig-cap: "Posterior random slope of age by participant"
-re_exposure_lv <- ranef(model_fit)$id[,,"exposure_std:lv_std"]
-
-re_lv %>%
-    as_tibble() %>%
-    janitor::clean_names() %>%
-    mutate(id = unique(model_fit$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-DoE slope
-
-```{r}
-#| label: fig-results-random-id-age-doe
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
-re_age_exposure <- ranef(model_fit)$id[,,"age_std:exposure_std"]
-
-re_age_exposure %>%
-    as_tibble() %>%
-    janitor::clean_names() %>%
-    mutate(id = unique(model_fit$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-age-lv
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
-re_age_lv <- ranef(model_fit)$id[,,"age_std:lv_std"]
-
-re_age_lv %>%
-    as_tibble() %>%
-    janitor::clean_names() %>%
-    mutate(id = unique(model_fit$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-DoE-by-Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-age-exposure-lv
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
-re_age_exposure_lv <- ranef(model_fit)$id[,,"age_std:exposure_std:lv_std"]
-
-re_age_exposure_lv %>%
-    as_tibble() %>%
-    janitor::clean_names() %>%
-    mutate(id = unique(model_fit$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-:::
-
-### Item-level effects
-
-::: panel-tabset
-
-#### Intercept
-
-```{r}
-#| label: fig-results-random-te
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior probability of acquisition by translation equivalent"
-re_te <- ranef(model_fit)$te[,,"Intercept"] %>% 
-    as_tibble() %>% 
-    rownames_to_column("te") %>% 
-    janitor::clean_names() %>% 
-    mutate(
-        across(
-            .cols = where(is.numeric),
-            .fns = list(
-                Comprehension = ~plogis(. + fixef(model_fit)[1]),
-                Production = ~plogis(. + fixef(model_fit)[2])
-            ), .names = "{.col}__{.fn}"
-        )
-    ) %>% 
-    select(-c(estimate:q97_5)) %>% 
-    pivot_longer(
-        where(is.numeric), 
-        names_to = c(".value", "type"),
-        names_sep = "__"
-    ) %>% 
-    mutate(te = as.numeric(te))
-
-re_te %>% 
-    ggplot() +
-    aes(x = estimate,
-        y = reorder(te, estimate), 
-        xmin = q2_5,
-        xmax = q97_5,
-        colour = type, 
-        fill = type) +
-    facet_wrap(~type) +
-    geom_ribbon(aes(group = 1), 
-                alpha = 0.5, 
-                colour = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0.5, 
-               colour = "black", 
-               linetype = "dotted") +
-    labs(x = "Posterior probability of acquisition",
-         y = "TE", 
-         colour = "Type") +
-    scale_x_continuous(limits = c(0, 1),
-                       labels = scales::percent) +
-    scale_fill_manual(values = clrs[c(1, 4)]) +
-    scale_colour_manual(values = clrs[c(1, 4)]) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age slope
-
-```{r}
-#| label: fig-results-random-te-age
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by translation equivalent"
-re_age <- ranef(model_fit)$te[,,"age_std"]
-
-re_age %>% 
-    as_tibble() %>% 
-    janitor::clean_names() %>% 
-    mutate(te = unique(model_fit$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### # Phonemes slope
-
-```{r}
-#| label: fig-results-random-te-nphon
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
-re_n_phon <- ranef(model_fit)$te[,,"n_phon_std"]
-
-re_n_phon %>% 
-    as_tibble() %>% 
-    janitor::clean_names() %>% 
-    mutate(te = unique(model_fit$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Exosure slope
-
-```{r}
-#| label: fig-results-random-te-exposure
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_doe <- ranef(model_fit)$te[,,"exposure_std"]
-
-re_doe %>%
-    as_tibble() %>%
-    janitor::clean_names() %>%
-    mutate(te = unique(model_fit$data$te)) %>%
-    relocate(te) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-exposure slope
-
-```{r}
-#| label: fig-results-random-te-age-exposure
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_age_doe <- ranef(model_fit)$te[,,"age_std:exposure_std"]
-
-re_age_doe %>% 
-    as_tibble() %>% 
-    janitor::clean_names() %>% 
-    mutate(te = unique(model_fit$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-:::
-
-## Model diagnostics
-
-### Traceplots
-
-```{r}
-#| label: fig-diagnostics-traceplots
-#| message: false
-#| warning: false
-#| fig-width: 12
-gather_draws(model_fit, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
-    mutate(.chain = paste("Chain ", .chain)) %>%
-    ggplot(aes(.iteration, .value, colour = .chain)) +
-    facet_wrap(~.variable, scales = "free_y") +
-    annotate(geom = "rect", colour = NA,
-             xmin = 0, xmax = dim(model_fit$fit)[1]/2,
-             ymin = -Inf, ymax = Inf,
-             alpha = 0.25) +
-    geom_line() +
-    labs(
-        x = "Iteration",
-        y = "Sample value",
-        colour = "Chain"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 3, 4, 5)]) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank(),
-        axis.text.x = element_text(size = 7)
-    )
-```
-
-### Model convergence
-
-```{r fig-rhats-neffs}
-#| label: fig-rhats-neffs
-#| fig-cap: "MCMC convergence diagnostic of all parameters in the model. A: distribution of the Gelman-Rubin (R-hat) scores. B: distribution of the ratio of effective sa"
-#| message: false
-#| warning: false
-#| fig-height: 4
-#| fig-width: 7
-#| out-width: 100%
-model_convergence |> 
-    ggplot(aes(.rhat)) +
-    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
-    geom_vline(xintercept = 1.01, linetype = "dashed") +
-    labs(x = "R-hat",
-         y = "MCMC samples") +
-
-    model_convergence |> 
-    ggplot(aes(.neff)) +
-    geom_histogram(binwidth = 0.1, colour = "white", fill = "grey60") +
-    geom_vline(xintercept = 1.01, linetype = "dashed") +
-    labs(x = "Effective sample size ratio",
-         y = "MCMC samples") +
-
-    plot_layout(ncol = 1) &
-    plot_annotation(tag_levels = "A") &
-    theme(legend.position = "none",
-          panel.grid = element_blank())
-```
-
-# Appendix
-
-## Session info
+# Session info
 

### 2023-06-28 manuscript/appendix.qmd
Update PDF format

@@
 
-## Appendix A: model predictors {.appendix .unnumbered}
-
-{{< include _appendix/_model-predictors.qmd >}}
-
-{{< pagebreak >}}
-
-## Appendix B: model details {.appendix .unnumbered}
-
-{{< include _appendix/_model-details.qmd >}}
-
-{{< pagebreak >}}
-
-## Appendix C: convergence diagnostics {.appendix .unnumbered}
-
-{{< include _appendix/_model-diagnostics.qmd >}}
-
-{{< pagebreak >}}
-
-## Appendix D: frequency and language exposure as separate predictors {.appendix .unnumbered}
+## Appendix A: frequency and language exposure as separate predictors {.appendix .unnumbered}
 
@@
 
-## Appendix E: frequency and language exposure as separate predictors {.appendix .unnumbered}
+## Appendix B: frequency and language exposure as separate predictors {.appendix .unnumbered}
 
@@
 {{< pagebreak >}}
-
-## References

### 2023-06-29 manuscript/manuscript.qmd
Modify template

@@
   arxiv-pdf:
-    keep-tex: true  
+    keep-tex: false  
     linenumbers: true
@@
 
+thanks: |
+    The authors declare no conflicts of interest with regard to the funding source of this study.
+This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
+Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
+We are grateful to Chiara Santolin, Ege E.
+Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
+We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support, and Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
+We also thank all families and infants who participated in the experiments.
 fig-dpi: 1000
@@
 
-# Introduction[^author-note]
-
-[^author-note]: The authors declare no conflicts of interest with regard to the funding source of this study.
-This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
-Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
-We are grateful to Chiara Santolin, Ege E.
-Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
-We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support, and Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
-We also thank all families and infants who participated in the experiments.
+# Introduction
 
@@
 
-\renewcommand{\arraystretch}{1.25}
-
 ```{r results-values}

### 2023-07-03 manuscript/_appendix/_syllable-frequency.qmd
Add comments by DAV

@@
 
-We define syllable frequency as the rate of appearance of individual syllables in the word-forms included in the Barcelona Vocabulary Questionnaire (BVQ).
-Each item corresponds to a Catalan or Spanish word, and has an associated phonological transcription in X-SAMPA format.
+We define syllable frequency as the rate of appearance of individual syllables in the word-forms included in the Barcelona Vocabulary Questionnaire (BVQ) [@garcia-castro2023bvq].
+Each item corresponds to a Catalan or Spanish word, and has an associated phonological transcription in X-SAMPA format [@wells1995computercoding].
 These transcriptions are syllabified.
@@
 No words had more than five syllables (see @fig-syll-number).
-We extracted lexical frequencies from the English corpora in the CHILDES database.
+We extracted lexical frequencies from the English corpora in the CHILDES database [@macwhinney2000childes; sanchez2019childesdb].
 Using the Catalan and Spanish corpora was not possible due to the low number of children and tokens included in the corpora.
@@
 
-To estimate the association between word-level syllabic frequency and cognateness, while controlling for the number of syllables in the word, as words are expected to necessarily increase the syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ (Levenshtein similarity between a word and its translation equivalent) as predictors.
+To estimate the association between word-level syllabic frequency and cognateness, while controlling for the number of syllables in the word, as words are expected to necessarily increase the syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ [Levenshtein similarity between a word and its translation equivalent, @levenshtein1966binary] as predictors.
 We added translation equivalent-level random effects for the intercept and the main effect of $Syllables$ (some translation pairs had a different number of syllables in each language).
@@
 We used a weakly informative prior for all parameters involved in the model (see @eq-model-syllables for a formal equation of this model and its prior).
+We conducted statistical inference by evaluating the proportion of the 95% highest density interval (HDI) of the posterior posterior distribution of each coefficient that fell into the region of practical equivalence [ROPE, see the main manuscript for a more detailed explanation, @kruschke2018bayesian].
 
@@
 # summarise posterior draws
-post_syllables |> 
+tbl_data <- post_syllables |>
     select(.variable_name, .median, .lower, .upper, .rope_overlap) |>
-    gt(rowname_col = ".variable_name") |>
-    fmt_percent(.rope_overlap, decimals = 2) |> 
-    fmt_number(c(.median:.upper), decimals = 2) |>
-    cols_merge(c(.lower, .upper),
-               pattern = "[{1}, {2}]") |>
-    cols_label(
-        .variable_name = "Parameter",
-        .median = "Median",
-        .lower = md("95% HDI"),
-        .rope_overlap = md("ROPE prob.")
+    mutate(
+        .variable_name = gsub("×", '$\\\\times$', .variable_name),
+        .hdi = paste0("[", round(.lower, 3),
+                      ", ", round(.upper, 3), "]"),
+        .rope_overlap = sub("^(-?)0.", "\\1.",
+                            sprintf("%.3f", .rope_overlap))) |> 
+    select(.variable_name, .median, .lower, .upper, .hdi, .rope_overlap)
+
+tbl_data |> 
+    mutate(viz = "") |> 
+    select(-c(.lower, .upper)) |> 
+    knitr::kable(
+        format = "latex", 
+        escape = FALSE,
+        col.names = c("",
+                      "$\\beta$",
+                      "95\\% HDI",
+                      "$p(\\text{ROPE})$",
+                      ""),
+        digits = 3,
+        booktabs = TRUE,
+        align = "lclrc",
+        linesep = ""
     ) |> 
-    tab_style(cell_text(weight = "bold"),
-              cells_column_labels(columns = 1:5)) |>
-    tab_style(cell_text(align = "left"),
-              cells_title(groups = "subtitle")) |> 
-    tab_style(cell_text(size = "small"), 
-              list(cells_body(),
-                   cells_row_groups(),
-                   cells_stub())) |> 
-    tab_style(cell_text(style = "italic"),
-              cells_row_groups())
+    kableExtra::kable_styling() |> 
+    kableExtra::column_spec(
+        5,
+        background = "#f2f2f2",
+        image = kableExtra::spec_pointrange(
+            x = tbl_data$.median, 
+            xmin = tbl_data$.lower, 
+            xmax = tbl_data$.upper, 
+            vline = 0,
+            xaxt = "n",
+            col = "#000000",
+            res = 200)
+    )
 ```
@@
                                scale = sd(syllables_data$n_syll))[, 1]) |> 
-    add_epred_draws(model_fit_syllables, ndraws = 50,
+    add_epred_draws(model_fit_syllables,
                     re_formula = NA) |>
+    median_hdci() |> 
     mutate(
@@
     ggplot(aes(lv_std, .epred, 
-               colour = n_syll_std)) +
-    geom_point(data = filter(syllables_data, n_syll < 5),
+               shape = n_syll_std)) +
+    geom_point(data = filter(syllables_data, n_syll < 5) |> 
+                   mutate(n_syll_std = as.factor(n_syll)),
                aes(x = lv,
                    y = freq_syll,
-                   colour = as.factor(n_syll),
-                   shape = as.factor(n_syll)),
+                   shape = n_syll_std),
                alpha = 1/2) +
-    geom_line(aes(group = interaction(.draw, n_syll_std)),
-              alpha = 1/10) +
+    # geom_ribbon(aes(ymin = .lower,
+    #                 ymax = .upper),
+    #             alpha = 1/2,
+    #             colour = NA,
+    #             show.legend = FALSE) +
+    geom_line(aes(linetype = n_syll_std),
+              linewidth = 0.5) +
     labs(x = "Cognateness (Levenshtein similarity)",
          y = "Syllabic frequency",
+         linetype = "Number of syllables",
          shape = "Number of syllables",
          colour = "Number of syllables") +
-    scale_colour_grey(start = 0, end = 0.7) +
     scale_y_continuous(n.breaks = 8) +

### 2023-07-03 manuscript/manuscript.qmd
Add comments by DAV

@@
         name: Universidade de Lisboa
-        department: Facultade de Letras
-        city: Lisboa, 1600-214
+        department: Linguistics Center (CLUL), School of Arts and Humanities
+        city: Lisbon, 1600-214
   - name: Ignacio Castillejo
@@
   arxiv-pdf:
-    keep-tex: false  
+    keep-tex: true  
     linenumbers: true
-    doublespacing: true
+    doublespacing: false
     runninghead: "Cognate beginnings to bilingual lexical acquistion"
@@
     The authors declare no conflicts of interest with regard to the funding source of this study.
-This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award].
-Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019).
+GGC was supported by the This research was supported by a FPI 2020 research contract (PRE2019-088165) in a project funded by the Spanish Ministry of Science, Innovation, and Universities (PGC2018-101831-B-I00).
+DAV was supported by the European Union's Horizon 2023 research and innovation program under Marie Skłodowska–Curie Grant (765556) and a postdoctoral fellowship from the Foundation for Science and Technology of Portugal (UIDB/00214/2020).
+NSG was supported by Academia 2019 award from the Catalan Institution for Research and Advanced Studies (ICREA).
 We are grateful to Chiara Santolin, Ege E.
@@
 The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
-The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh.
+The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh.
 The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of cognateness between each pair of languages.
@@
 For example, English-Dutch bilinguals (languages with a high phonological overlap), were able to produce more Dutch words than English-Mandarin bilinguals (languages with a low phonological overlap) were able to produce in Mandarin.
-@blom2020crosslanguage, @bosma2019longitudinal, and @gampe2021does reported similar results, providing converging evidence of a facilitatory effect of language distance on vocabulary size.
+@blom2020crosslanguage, @bosma2019longitudinal, and @gampe2021does reported similar results, providing converging evidence of a facilitatory effect of a lower language distance (i.e., higher degre of cognateness) on vocabulary size.
 
@@
 Parallel activation of bilinguals’ lexicons has been proposed as the underlying mechanism for such facilitatory effect [e.g., @floccia2018introduction; @mitchell2022cognates].
-The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during language processing, even in fully monolingual contexts Evidence with adult bilinguals supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; see @kroll2017bilingual for review].
+The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during language processing, even in fully monolingual contexts.
+Evidence with adult bilinguals supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; see @kroll2017bilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 We rather consider that a learning instance of a word is so to the degree of the strength of activation of its phonological representation.
-This activation may be the result of the infant being exposed to the word-form, or the result of activation spreading through phonological or semantic links across lexicl representation, as in the case of parallel activation.].
+This activation may result from the infant being exposed to the word-form, or the result of activation spreading through phonological or semantic links across lexical representation, as in the case of parallel activation.].
 The strength of this co-activation is proportional to the phonological similarity between the two translation equivalents; given that cognates share higher phonological similarity than non-cognates, the former should be co-activated more frequently than the latter.
 This should lead to a faster accumulation of learning instances for cognates, compared to non-cognates.
-Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across languages: words from the lower-exposure language would receive additional activation through parallel activation more often than words from the higher-exposure language.
+Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across languages: words from the lower-exposure language (i.e., the language to which the infants are exposed to more often) would receive additional activation through parallel activation more frequently than words from the higher-exposure language.
 Therefore, the acquisition of words from the lower-exposure language would benefit more strongly from their cognate status than words from the higher-exposure language.
@@
 Therefore, this exposure will count as a learning instance for both co-activated forms.
-However, for the case of the non-cognate translation equivalent /\textipa{"gos}/, the low phonological similarity will result in a weak activation of /\textipa{"pe.ro}/ resulting in such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for /\textipa{"pe.ro}/.
+The case of the non-cognate translation equivalent /\textipa{"gos-"pe.ro}/ would be different.
+Given the low phonological similarity between both word-forms, an exposure to \textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/ leading to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for /\textipa{"pe.ro}/.
 While /\textipa{"gat}-\textipa{"ga.to}/ will benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not.
@@
 
-To test these predictions, we collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study.
+To test these predictions, we collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using the Barcelona Vocabulary Questionnaire (BVQ) [@garcia-castro2023bvq], a dual language online vocabulary checklist designed for the present study.
 We adopted a Bayesian explanatory item response theory approach [IRT, see @kachergis2022standard, for a similar approach] to model the probability of acquisition of `r nrow(distinct(items, language, item))` Catalan and Spanish nouns included in the vocabulary checklist.
@@
 As described in the methods section, rate of exposure was a composite measure taking into account participant’ language exposure and word’s lexical frequency.
-We predicted an interaction between cognate status and word exposure rate  in which the probability of comprehension is higher for low-exposure cognate words, but not for high-exposure words cognate words.
+We predicted an interaction between cognate status and word exposure rate  in which the probability of comprehension is higher for low-exposure cognate words, but not for high-exposure cognate words.
 
@@
 #| warning: false
-items |> 
+col_names <- c("Semantic category", 
+               "List A",
+               "List B",
+               "List C",
+               "List D",
+               "Examples")
+
+tbl_data <- items |> 
     distinct(te, .keep_all = TRUE) |> 
@@
     arrange(desc(total)) |> 
-    relocate(a:d, total, items) |> 
-    mutate(group = "Semantic category") |> 
-    select(-total) |> 
-    gt(groupname_col = "group",
-       rowname_col = "semantic_category") |> 
-    fmt_percent(matches("_prop"), decimals = 1) |> 
-    cols_label(a = "List A",
-               b = "List B",
-               c = "List C",
-               d = "List D",
-               items = "Examples") |> 
-    summary_rows(groups = "Semantic category",
-                 columns = a:d,
-                 fns = list("Total" = ~sum(., na.rm = TRUE))
-    ) |> 
-    tab_style(cell_text(size = "small"),
-              list(cells_body(),
-                   cells_summary(),
-                   cells_stub(),
-                   cells_column_labels(),
-                   cells_column_spanners(),
-                   cells_stub_summary(),
-                   cells_row_groups())) |> 
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) |> 
-    tab_style(cell_text(style = "italic"),
-              cells_column_labels()) |> 
-    tab_options(data_row.padding = px(2),
-                row_group.padding = px(1.2)) |> 
-    tab_style(cell_text(size = "x-small", align = "left"),
-              cells_body(items))
+    relocate(semantic_category, a:d, total, items) |> 
+    select(-total)
 
+sum_rows <- function(x) {
+    if (is.numeric(x)) sum(x) 
+    else if(is.list(x)) ""
+    else "N"
+}
+
+sums_row <- as.data.frame(lapply(tbl_data, sum_rows),
+                          col.names = colnames(tbl_data))
+tbl_data <- rbind(tbl_data, sums_row)
+
+knitr::kable(tbl_data,
+             format = "latex", 
+             digits = 1, 
+             booktabs = TRUE,
+             col.names = col_names,
+             align = "lccccl")  |> 
+    kableExtra::row_spec(nrow(tbl_data)-1, 
+                         hline_after = TRUE)
 ```
@@
 Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`, and were always allocated to the same questionnaire list (A, B, C, or D).
-Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra) and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants) for a more detailed description of the sample).
+Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra) and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample).
 In total, `r n_trilingual$n[2]` participants (`r scales::percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed to a third language other than Catalan and Spanish.
@@
 #| tbl-cap: "Participant sample size by age and degree of exposure to Catalan."
-#| tbl-note: "Information about language degree of exposure (DoE) was provided by families before filling the questionnaire.
-In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish.
-A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish.
-A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each.
-The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.
-For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4-month-wide bins.
-Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
-ftn <- "For participants exposed to Catalan and Spanish exclusively this proportion is complementary to the degree of exposue to Spanish.
-For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10%."
-participants |>
+#| tbl-note: "For participants exposed to Catalan and Spanish exclusively, the proportion of Catalan exposure shown in the table is complementary to the degree of exposue to Spanish.
+For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10%.
+Participants with 100\\%, 0\\%, and 50\\% Catalan exposure would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively."
+ftn <- ""
+
+col_names <- c("Catalan exposure",
+               "[10-14]",
+               "(14, 18]",
+               "(18, 22]", 
+               "(22-26]", 
+               "(26-30]", 
+               "(30-34]")
+
+options(knitr.kable.NA = 0)
+
+tbl_data <- participants |>
     mutate_at(vars(starts_with("doe_")), 
@@
     arrange(desc(doe_catalan)) |>
-    pivot_wider(names_from = age, values_from = n) |>
-    gt() |>
-    sub_missing(everything(), everything(), missing_text = "-") |>
-    cols_label(doe_catalan = "Catalan exposure") |>
-    tab_spanner("Age (months)", 2:7) |>
-    tab_style(cell_text(align = "left"),
-              list(cells_body(doe_catalan),
-                   cells_column_labels(doe_catalan))) |>
-    tab_footnote(ftn, location = cells_column_labels(doe_catalan))
+    pivot_wider(names_from = age, 
+                values_from = n,
+                names_repair = janitor::make_clean_names) |> 
+    mutate(across(where(is.numeric), \(x) ifelse(is.na(x), 0, x)))
+
+sum_rows <- function(x) if (is.numeric(x)) sum(x) else "N"
+sums_row <- as.data.frame(lapply(tbl_data, sum_rows),
+                          col.names = colnames(tbl_data))
+tbl_data <- rbind(tbl_data, sums_row)
+
+knitr::kable(tbl_data,
+             format = "latex", 
+             digits = 1, 
+             booktabs = TRUE,
+             col.names = col_names,
+             align = "lcccccc")  |> 
+    kableExtra::add_header_above(c(" " = 1,
+                                   "Age (months)" = ncol(tbl_data)-1)) |>
+    kableExtra::row_spec(nrow(tbl_data)-1, 
+                         hline_after = TRUE)
 ```
@@
 
-We collected responses to `r format(nrow(n_items), big.mark = ",")` items.
+### Data processing
+
+We collected `r format(n_obs, big.mark = ",")` responses to `r format(nrow(n_items), big.mark = ",")` items.
 We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, see @fourtassi2020growth for a similar approach].
@@
 Finally, we removed items without a translation in the other language (*n* = `r nrow(n_items_multi)-nrow(n_items_duplicated)`).
-This resulted in a final list of `r sum(n_items_final)` items, corresponding to words `r n_items_final["Catalan"]` Catalan words and their `r n_items_final["Spanish"]` Spanish translations (`r n_te` translation equivalents).
+This resulted in a final list of `r sum(n_items_final)` items, corresponding to `r n_items_final["Catalan"]` Catalan words and their `r n_items_final["Spanish"]` Spanish translations (`r n_te` translation equivalents).
 After collecting participants' responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item.
@@
 
+### Modelling approach
+
 We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, multilevel, ordinal regression model.
-This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest: *Age* (number of months elapsed between participants' birth date and questionnaire completion), *Length* (number of phonemes in the X-SAMPA phonological transcription of the word-form), *Exposure* (a language exposure-weighted lexical frequency), and *Cognateness* (defined as the phonological similarity between translation equivalentsFor each translation equivalent).
+This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest: *Age* (number of months elapsed between participants' birth date and questionnaire completion), *Length* (number of phonemes in the X-SAMPA phonological transcription of the word-form), *Exposure* (a language exposure-weighted lexical frequency), and *Cognateness* (defined as the phonological similarity between translation equivalents).
 We added these variables as main effects, together with the two-way and three-way interactions between *Age*, *Exposure*, and *Cognateness*.
-
-We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-forms lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
-First, we extracted the child-directed lexical frequency of each word from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
-Using the corresponding lexical frequencies directly from Catalan and Spanish was not possible due to the low number of Catalan participants and tokens available in their corresponding CHILDES corpora, so they were extracted from the English corpora instead.
-We then mapped the lexical frequencies of the English words to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed them to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human].
-We then multiplied the resulting lexical frequencies by the reported degree of exposure of the child to Catalan or Spanish.
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure score to the Catalan word *cavall* [*horse*]---with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`---would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.
-
-We defined *Cognateness* as the phonological similarity between each word-form and its translation.
-For each translation equivalent, we used the `stringdist` [@vanderloo2014stringdist] R package to calculate the Levenshtein distance between the Catalan and the Spanish phonological transcriptions of the word-forms.
-The Levenshtein distance measures the number of editions (insertions, deletions, or substitutions) that one string of phonemes/characters must go through to become identical to the other [@levenshtein1966binary].
-We divided the Levnehstein distance of each translation equivalent by the length of the longest word-form to correct for word length (longer strings are likely to show a larger number of mismatches).
-Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
-This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /'taw.l5-'me.sa/, *table*), and one indicates that the two word-forms are identical (e.g., /'mar-'mar/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
-
-Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
 Participant-level and item-level random intercepts and slopes were included where appropriate, according to the structure of the data [@barr2013random].
@@
 
+### Model predictors
+
+We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-forms lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
+First, we extracted the child-directed lexical frequency of each word from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
+Using the corresponding lexical frequencies directly from Catalan and Spanish was not possible due to the low number of Catalan participants and tokens available in their corresponding CHILDES corpora, so they were extracted from the English corpora instead.
+We then mapped the lexical frequencies of the English words to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed them to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human].
+We then multiplied the resulting lexical frequencies by the reported degree of exposure of the child to Catalan or Spanish.
+For instance, for a child whose degree of exposure is 80% for Catalan and 20% for Spanish, the expected *Exposure* score to the Catalan word *cavall* [*horse*]---with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`---would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.
+
+We defined *Cognateness* as the phonological similarity between each word-form and its translation.
+For each translation equivalent, we used the `stringdist` [@vanderloo2014stringdist] R package to calculate the Levenshtein distance between the Catalan and the Spanish phonological transcriptions of the word-forms.
+The Levenshtein distance measures the number of editions (insertions, deletions, or substitutions) that one string of phonemes/characters must go through to become identical to the other [@levenshtein1966binary].
+We divided the Levenshtein distance of each translation equivalent by the length of the longest word-form to correct for word length (longer strings are likely to show a larger number of mismatches).
+Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
+This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}-\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}-\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
+
+Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
+
+### Statistical inference
+
 We assessed the practical relevance of the estimated regression coefficients of the model model following @kruschke2018bayesian.
@@
 Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, noted as $p(\text{ROPE})$, which indicates the probability that the true value of the regression coefficient falls into the ROPE (and therefore should be considered equivalent to zero).
-For example, a 80% overlap between the HDI of a coefficient's posterior distribution and the  indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero.
+For example, $p(\text{ROPE})=.80$ indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero.
 
@@
 We implemented the model using `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
-We ran four MCMC  chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
+We ran four iteration chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
 Model posterior draws and predictions were handled using the `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
@@
 ```
-
-Model posterior showed adequate MCMC convergence diagnostics for all parameters ($\hat{R} \le 1.06$, $0.022 \le N_{eff}/N < 1.1001$) and little evidence of multicollinearity.
+The model posterior showed adequate chain convergence diagnostics for all parameters ($\hat{R} \le 1.06$, $0.022 \le N_{eff}/N < 1.1001$) and little evidence of multicollinearity.
 @tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients, and their degree of overlap with the ROPE.
@@
 intercept_label <- "Intercepts (at 22 months)"
-model_summary |>
+
+tbl_data <- model_summary |>
     select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
@@
         .hdi = paste0("[", round(.lower, 3),
-                      ", ", round(.upper, 3), "]")) |> 
-    select(.variable_name, .median, .hdi, .rope_overlap) |> 
-    knitr::kable(format = "latex", 
+                      ", ", round(.upper, 3), "]"),
+        .rope_overlap = sub("^(-?)0.", "\\1.",
+                            sprintf("%.3f", .rope_overlap))) |> 
+    select(.variable_name, .median, .lower, .upper, .hdi, .rope_overlap)
+
+tbl_data |> 
+    mutate(viz = "") |> 
+    select(-c(.lower, .upper)) |> 
+    knitr::kable(
+        format = "latex", 
         escape = FALSE,
         col.names = c("",
-                               "Median",
+                      "$\\beta$",
                       "95\\% HDI",
-                               "$p(\\text{ROPE})$"),
-                 digits = 4,
+                      "$p(\\text{ROPE})$",
+                      ""),
+        digits = 3,
         booktabs = TRUE,
+        align = "lclrc",
         linesep = ""
@@
     kableExtra::kable_styling() |> 
-    kableExtra::pack_rows("Intercepts (at 22 months)", start_row = 1, end_row = 2) |> 
-    kableExtra::pack_rows("Slopes (upper bound)", start_row = 3, end_row = 10)
+    kableExtra::pack_rows("Intercepts (at 22 months)",
+                          start_row = 1, 
+                          end_row = 2) |> 
+    kableExtra::pack_rows("Slopes (upper bound)",
+                          start_row = 3,
+                          end_row = 10) |> 
+    kableExtra::column_spec(
+        5,
+        background = "#f2f2f2",
+        image = kableExtra::spec_pointrange(
+            x = tbl_data$.median, 
+            xmin = tbl_data$.lower, 
+            xmax = tbl_data$.upper, 
+            vline = 0,
+            xaxt = "n",
+            col = "#000000",
+            res = 200)
+    )
 
@@
 For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.median)/sd(responses$n_phon), 3)` less likely to know it.
-The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word exposure index differed across ages: older children were more sensitive likely to acquire words with higher exposure rate than younger children.
+The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children.
 
@@
 Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
-This effect was negligible when for words with median or high exposure (+1 SD) (see @fig-marginal).
+This effect was negligible for words with median or high exposure (+1 SD) (see @fig-marginal).
 
@@
 #| fig-cap: "Posterior marginal effects.
-Lines correspond to 100 posterior-predicted medians.
+Lines and error bands correspond to the mean and 95% credible interval of the posterior-predicted means.
 Different colours indicate different levels of cognateness (phonological similarity).
@@
 In-sample predictions lie inside the grey rectangles."
-model_epreds |> 
+fig_data <- model_epreds |> 
+    summarise(.median = mean(.value),
+              .lower = qi(.value)[,1],
+              .upper = qi(.value)[,2],
+              .by = c(ends_with("_std"), .category)) |> 
     mutate(age = (age_std * sd(responses$age)) + mean(responses$age),
@@
                               "Comprehension",
-                              "Production")) |> 
-    ggplot(aes(age, .value, 
+                              "Production")) 
+fig_data |> 
+    ggplot(aes(age, .median, 
+               ymin = .lower,
+               ymax = .upper,
                colour = lv, 
-               fill = lv)) +
+               fill = lv,
+               linetype = lv)) +
     facet_grid(.category~exposure) +
@@
              ymax = Inf,
-             fill = "grey98",
+             fill = "grey95",
              linetype = "dotted") +
@@
                linetype = "dotted") +
-    geom_line(aes(group = interaction(lv, .draw)),
-              linewidth = 2/4,
-              alpha = 1/2) +
-    # stat_summary(geom = "line", fun = mean,
-    #              linewidth = 2/4,
-    #              colour = "black",
-    #              aes(linetype = lv,
-    #                  group = interaction(lv, exposure))) +
+    geom_ribbon(alpha = 3/4,
+                colour = NA) +
+    geom_line(linewidth = 2/3) +
     labs(x = "Age (months)",
@@
          colour = "Cognateness (phonological similarity)",
-         fill = "Cognateness (phonological similarity)\n") +
-    scale_color_grey(start = 0.9, end = 0.2,
-                     guide = guide_legend(
-                         override.aes = list(
-                             linewidth = 5,
-                             alpha = 1
-                         )
-                     )) +
+         fill = "Cognateness (phonological similarity)",
+         linetype = "Cognateness (phonological similarity)") +
+    scale_color_grey(start = 0.80, end = 0.1) +
+    scale_fill_grey(start = 0.80, end = 0.1) +
+    scale_linetype_manual(values = rev(c("solid", "dashed", "dotdash"))) +
     scale_y_continuous(breaks = seq(0, 1, 0.25),
@@
 
-This study investigated the impact of cognateness on the early bilingual lexicon.
+This study investigated the impact of cognateness (i.e., phonological similarity between each word-form and its translation) on the early bilingual lexicon.
 We used Bayesian Item Response Theory to model the acquisition trajectories of a large sample of Catalan and Spanish words, estimating the effect of cognateness on the probability of acquisition.
@@
 First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (e.g.
-shared many cognates, like., English-German) showed larger vocabulary sizes than those learning typologically distant languages (e.g.
+shared many cognates, like., English-Dutch) showed larger vocabulary sizes than those learning typologically distant languages (e.g.
 shared fewer cognates, like English-Mandarin).
@@
 Our study shares similar methodological limitations with previous work using vocabulary reports provided by caregivers.
-Such reports can be subject to measurement error induced by caregivers who may sometimes overestimate or underestimate participants' true probability of acquisition of words [e.g., @houston-price2007discrepancy].
+Such reports can be subject to measurement error induced by caregivers who may sometimes overestimate or underestimate participants' true probability of word acquisition [e.g., @houston-price2007discrepancy].
 In the case of bilingual research additional biases may be in place.

### 2023-07-03 manuscript/manuscript.tex
Add comments by DAV

@@
 \usepackage{booktabs}
-\usepackage{caption}
 \usepackage{longtable}
@@
 \linenumbers
-\usepackage{setspace}
-\doublespacing
 \newcommand{\runninghead}{A Preprint }
@@
 acquistion }
-\title{Cognate beginnings to bilingual lexical acquisition}
+\title{Cognate beginnings to bilingual lexical acquisition\thanks{The
+authors declare no conflicts of interest with regard to the funding
+source of this study. GGC was supported by the This research was
+supported by a FPI 2020 research contract (PRE2019-088165) in a project
+funded by the Spanish Ministry of Science, Innovation, and Universities
+(PGC2018-101831-B-I00). DAV was supported by the European Union's
+Horizon 2023 research and innovation program under Marie
+Skłodowska--Curie Grant (765556) and a postdoctoral fellowship from the
+Foundation for Science and Technology of Portugal (UIDB/00214/2020). NSG
+was supported by Academia 2019 award from the Catalan Institution for
+Research and Advanced Studies (ICREA). We are grateful to Chiara
+Santolin, Ege E. Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada,
+and the rest of the Speech Acquisition and Perception research group for
+their helpful feedback. We thank Xavier Mayoral, Silvia Blanch, and
+Cristina Cuadrado for their technical support, and Cristina Dominguez
+and Katia Pistrin for their efforts in recruiting infants. We also thank
+all families and infants who participated in the experiments.}}
 \author{
@@
 for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}\\\\\\
+08005\\\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}\And
 \textbf{Daniela S. Avila-Varela}~\orcidlink{0000-0002-3518-8117}\\Center
 for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}\\\\\\
+08005\\\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}\And
 \textbf{Ignacio
@@
 Psicología\\Universidad Autónoma de Madrid\\Madrid,
-28049\\\href{mailto:jignaciocastillejo@gmail.com}{jignaciocastillejo@gmail.com}\\\\\\
+28049\\\href{mailto:jignaciocastillejo@gmail.com}{jignaciocastillejo@gmail.com}\And
 \textbf{Nuria Sebastian-Galles}~\orcidlink{0000-0001-6938-2498}\\Center
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, frame hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
-
-\hypertarget{introductionauthor-note}{%
-\section[Introduction]{\texorpdfstring{Introduction\footnote{The authors
-  declare no conflicts of interest with regard to the funding source of
-  this study. This research was supported by grants from the Spanish
-  Ministerio de Ciencia, Innovación y Universidades
-  (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government
-  {[}ICREA (Catalan Institution for Research and Advanced Studies)
-  Academia 2019 award{]}. Gonzalo Garcia-Castro was supported by a
-  fellowship of the Spanish Ministerio de Ciencia, Innovación y
-  Universidades (FPI 2019). We are grateful to Chiara Santolin, Ege E.
-  Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest
-  of the Speech Acquisition and Perception research group for their
-  helpful feedback. We thank Xavier Mayoral, Silvia Blanch, and Cristina
-  Cuadrado for their technical support, and Cristina Dominguez and Katia
-  Pistrin for their efforts in recruiting infants. We also thank all
-  families and infants who participated in the experiments.}}{Introduction}}\label{introductionauthor-note}}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, sharp corners, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu,
-Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The
+Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh. The
 authors calculated the average phonological overlap between the words in
@@
 (\protect\hyperlink{ref-gampe2021does}{2021}) reported similar results,
-providing converging evidence of a facilitatory effect of language
-distance on vocabulary size.
+providing converging evidence of a facilitatory effect of a lower
+language distance (i.e., higher degre of cognateness) on vocabulary
+size.
 
@@
 languages simultaneously during language processing, even in fully
-monolingual contexts Evidence with adult bilinguals supporting the
+monolingual contexts. Evidence with adult bilinguals supporting the
 language-non selective account of lexical access has been reported for
@@
   is so to the degree of the strength of activation of its phonological
-  representation. This activation may be the result of the infant being
+  representation. This activation may result from the infant being
   exposed to the word-form, or the result of activation spreading
-  through phonological or semantic links across lexicl representation,
+  through phonological or semantic links across lexical representation,
   as in the case of parallel activation.}. The strength of this
@@
 asymmetric across languages: words from the lower-exposure language
+(i.e., the language to which the infants are exposed to more often)
 would receive additional activation through parallel activation more
-often than words from the higher-exposure language. Therefore, the
+frequently than words from the higher-exposure language. Therefore, the
 acquisition of words from the lower-exposure language would benefit more
@@
 parallel. Therefore, this exposure will count as a learning instance for
-both co-activated forms. However, for the case of the non-cognate
-translation equivalent /\textipa{"gos}/, the low phonological similarity
-will result in a weak activation of /\textipa{"pe.ro}/ resulting in such
-exposure counting as a learning instance for /\textipa{"gos}/ (which the
-child was exposed to), but not for /\textipa{"pe.ro}/. While
-/\textipa{"gat}-\textipa{"ga.to}/ will benefit from parallel activation,
-/\textipa{"gos}-\textipa{"pe.ro}/ will not. If the child receives
-linguistic input from one of the languages more often than from the
-other, this effect might affect each form of the cognate translation
-equivalent differently. For instance, if the child receives a larger
-amount of Catalan input than Spanish input, they will encounter the
-Catalan form /\textipa{"gat}/ more frequently than the Spanish form
+both co-activated forms. The case of the non-cognate translation
+equivalent /\textipa{"gos-"pe.ro}/ would be different. Given the low
+phonological similarity between both word-forms, an exposure to
+\textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/
+leading to such exposure counting as a learning instance for
+/\textipa{"gos}/ (which the child was exposed to), but not for
+/\textipa{"pe.ro}/. While /\textipa{"gat}-\textipa{"ga.to}/ will benefit
+from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not. If
+the child receives linguistic input from one of the languages more often
+than from the other, this effect might affect each form of the cognate
+translation equivalent differently. For instance, if the child receives
+a larger amount of Catalan input than Spanish input, they will encounter
+the Catalan form /\textipa{"gat}/ more frequently than the Spanish form
 /\textipa{"ga.to}/. Through parallel activation, /\textipa{"gat}/ will
@@
 and comprehension from a large sample of bilingual Catalan-Spanish
-children using an online vocabulary checklist designed for the present
-study. We adopted a Bayesian explanatory item response theory approach
-(IRT, see \protect\hyperlink{ref-kachergis2022standard}{Kachergis et
-al., 2022}, for a similar approach) to model the probability of
-acquisition of 604 Catalan and Spanish nouns included in the vocabulary
-checklist. Words were considered as acquired if caregivers reported such
-word to be understood (comprehension) or understood and said
-(production) by their child. We estimated the impact of several
-predictors of interest on the probability of acquisition, including
-participants' age and rate of exposure to the word-form, and the cognate
-status of the word-form. As described in the methods section, rate of
-exposure was a composite measure taking into account participant'
-language exposure and word's lexical frequency. We predicted an
-interaction between cognate status and word exposure rate in which the
-probability of comprehension is higher for low-exposure cognate words,
-but not for high-exposure words cognate words.
+children using the Barcelona Vocabulary Questionnaire (BVQ)
+(\protect\hyperlink{ref-garcia-castro2023bvq}{Garcia-Castro et al.,
+2023}), a dual language online vocabulary checklist designed for the
+present study. We adopted a Bayesian explanatory item response theory
+approach (IRT, see
+\protect\hyperlink{ref-kachergis2022standard}{Kachergis et al., 2022},
+for a similar approach) to model the probability of acquisition of 604
+Catalan and Spanish nouns included in the vocabulary checklist. Words
+were considered as acquired if caregivers reported such word to be
+understood (comprehension) or understood and said (production) by their
+child. We estimated the impact of several predictors of interest on the
+probability of acquisition, including participants' age and rate of
+exposure to the word-form, and the cognate status of the word-form. As
+described in the methods section, rate of exposure was a composite
+measure taking into account participant' language exposure and word's
+lexical frequency. We predicted an interaction between cognate status
+and word exposure rate in which the probability of comprehension is
+higher for low-exposure cognate words, but not for high-exposure cognate
+words.
 
@@
 \hypertarget{tbl-items}{}
-\begin{longtable}{l|rrrrc}
+\begin{table}
 \caption{\label{tbl-items}Summary of the items included in the final analyses. }\tabularnewline
 
+\centering
+\begin{tabular}{lccccl}
 \toprule
-\multicolumn{1}{l}{} & List A & List B & List C & List D & Examples \\ 
+Semantic category & List A & List B & List C & List D & Examples\\
 \midrule
-\multicolumn{6}{l}{Semantic category} \\ 
+Household items & 31 & 26 & 30 & 25 & clock, video\\
+Food and drink & 29 & 26 & 23 & 27 & sausage, yogurt\\
+Animals & 26 & 23 & 19 & 25 & panther, tiger\\
+Outside & 14 & 13 & 13 & 15 & farm , stone\\
+Body parts & 14 & 12 & 11 & 11 & face  , finger\\
+\addlinespace
+Toys & 11 & 11 & 12 & 13 & piano , racket\\
+Clothes & 12 & 12 & 10 & 10 & zipper, sandal\\
+Vehicles & 9 & 10 & 11 & 10 & helicopter, tractor\\
+Colours & 6 & 6 & 6 & 6 & red  , green\\
+People & 7 & 4 & 6 & 6 & police    , babysitter\\
+\addlinespace
+Furniture and rooms & 4 & 4 & 4 & 4 & corridor, terrace\\
+Time & 2 & 2 & 2 & 2 & day  , night\\
+Adventures & 1 & 1 & 1 & 1 & witch\\
+Parts of things & 1 & 1 & 1 & 1 & wheel\\
 \midrule
-Household items & 31 & 26 & 30 & 25 & clock, video \\ 
-Food and drink & 29 & 26 & 23 & 27 & sausage, yogurt \\ 
-Animals & 26 & 23 & 19 & 25 & panther, tiger \\ 
-Outside & 14 & 13 & 13 & 15 & farm, stone \\ 
-Body parts & 14 & 12 & 11 & 11 & face, finger \\ 
-Toys & 11 & 11 & 12 & 13 & piano, racket \\ 
-Clothes & 12 & 12 & 10 & 10 & zipper, sandal \\ 
-Vehicles & 9 & 10 & 11 & 10 & helicopter, tractor \\ 
-Colours & 6 & 6 & 6 & 6 & red, green \\ 
-People & 7 & 4 & 6 & 6 & police, babysitter \\ 
-Furniture and rooms & 4 & 4 & 4 & 4 & corridor, terrace \\ 
-Time & 2 & 2 & 2 & 2 & day, night \\ 
-Adventures & 1 & 1 & 1 & 1 & witch \\ 
-Parts of things & 1 & 1 & 1 & 1 & wheel \\ 
-\midrule 
-Total & 167 & 151 & 149 & 156 & — \\ 
+N & 167 & 151 & 149 & 156 & \\
 \bottomrule
-\end{longtable}
+\end{tabular}
+\end{table}
 
@@
 than 10\% of the time to a language other than Spanish or Catalan (see
-Table~\ref{tbl-participants}) for a more detailed description of the
+Table~\ref{tbl-participants} for a more detailed description of the
 sample). In total, 70 participants (16.06\%) participants were reported
@@
 \hypertarget{tbl-participants}{}
-\setlength{\LTpost}{0mm}
-\begin{longtable}{rrrrrrr}
+\begin{table}
 \caption{\label{tbl-participants}Participant sample size by age and degree of exposure to Catalan. }\tabularnewline
 
+\centering
+\begin{tabular}{lcccccc}
 \toprule
- & \multicolumn{6}{c}{Age (months)} \\ 
-\cmidrule(lr){2-7}
-Catalan exposure\textsuperscript{\textit{1}} & [10,14] & (14,18] & (18,22] & (22,26] & (26,30] & (30,34] \\ 
+\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Age (months)} \\
+\cmidrule(l{3pt}r{3pt}){2-7}
+Catalan exposure & {}[10-14] & (14, 18] & (18, 22] & (22-26] & (26-30] & (30-34]\\
 \midrule
-75-100\% & 18 & 23 & 36 & 38 & 20 & 7 \\ 
-50-75\% & 8 & 13 & 30 & 41 & 18 & 1 \\ 
-25-50\% & 10 & 17 & 45 & 29 & 17 & - \\ 
-0-25\% & 7 & 11 & 21 & 17 & 8 & 1 \\ 
+75-100\% & 18 & 23 & 36 & 38 & 20 & 7\\
+50-75\% & 8 & 13 & 30 & 41 & 18 & 1\\
+25-50\% & 10 & 17 & 45 & 29 & 17 & 0\\
+0-25\% & 7 & 11 & 21 & 17 & 8 & 1\\
+\midrule
+N & 43 & 64 & 132 & 125 & 63 & 9\\
 \bottomrule
-\end{longtable}
-\begin{minipage}{\linewidth}
-\textsuperscript{\textit{1}}For participants exposed to Catalan and Spanish exclusively this proportion is complementary to the degree of exposue to Spanish.
-For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10\%.\\
-\end{minipage}
+\end{tabular}
+\end{table}
 
@@
 
-We collected responses to 1,590 items. We restricted the analyses to
-responses to nouns (628 items corresponding to other grammatical classes
-were excluded, see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
-et al., 2020} for a similar approach). We then excluded items with
-missing lexical frequency scores (\emph{n} = 269), items that included
-more than one lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} =
-48), multi-word items or phrases (e.g., \emph{barrita de cereales}
-{[}cereal bar{]}, \emph{n} = 9). Finally, we removed items without a
-translation in the other language (\emph{n} = 32). This resulted in a
-final list of 604 items, corresponding to words 302 Catalan words and
-their 302 Spanish translations (302 translation equivalents). After
-collecting participants' responses, the final dataset consisted of
-138,078 observations, each corresponding to a single response of one
-participant to one item. Each translation equivalent received a median
-of 234 responses (\emph{Min} = 106, \emph{Max} = 872) from participants,
-both languages pooled together.
+\hypertarget{data-processing}{%
+\subsubsection{Data processing}\label{data-processing}}
+
+We collected 138,078 responses to 1,590 items. We restricted the
+analyses to responses to nouns (628 items corresponding to other
+grammatical classes were excluded, see
+\protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020} for
+a similar approach). We then excluded items with missing lexical
+frequency scores (\emph{n} = 269), items that included more than one
+lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} = 48), multi-word
+items or phrases (e.g., \emph{barrita de cereales} {[}cereal bar{]},
+\emph{n} = 9). Finally, we removed items without a translation in the
+other language (\emph{n} = 32). This resulted in a final list of 604
+items, corresponding to 302 Catalan words and their 302 Spanish
+translations (302 translation equivalents). After collecting
+participants' responses, the final dataset consisted of 138,078
+observations, each corresponding to a single response of one participant
+to one item. Each translation equivalent received a median of 234
+responses (\emph{Min} = 106, \emph{Max} = 872) from participants, both
+languages pooled together.
+
+\hypertarget{modelling-approach}{%
+\subsubsection{Modelling approach}\label{modelling-approach}}
 
@@
 exposure-weighted lexical frequency), and \emph{Cognateness} (defined as
-the phonological similarity between translation equivalentsFor each
-translation equivalent). We added these variables as main effects,
-together with the two-way and three-way interactions between \emph{Age},
-\emph{Exposure}, and \emph{Cognateness}.
-
-We developed the \emph{Exposure} predictor to account for the fact that
-bilinguals' exposure to a given word-form is not only a function of the
-word-forms lexical frequency, but also of the quantitative input they
-receive from the language such word-form belongs to, we expressed
-lexical frequencies as the product between both variables. First, we
-extracted the child-directed lexical frequency of each word from the
-CHILDES database
-(\protect\hyperlink{ref-macwhinney2000childes}{MacWhinney, 2000};
-\protect\hyperlink{ref-sanchez2019childesdb}{Sanchez et al., 2019}).
-Using the corresponding lexical frequencies directly from Catalan and
-Spanish was not possible due to the low number of Catalan participants
-and tokens available in their corresponding CHILDES corpora, so they
-were extracted from the English corpora instead. We then mapped the
-lexical frequencies of the English words to their Catalan and Spanish
-translations (see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
-et al., 2020} for a similar approach), and transformed them to Zipf
-scores (\protect\hyperlink{ref-vanheuven2014subtlexuk}{Van Heuven et
-al., 2014}; \protect\hyperlink{ref-zipf1949human}{Zipf, 1949}). We then
-multiplied the resulting lexical frequencies by the reported degree of
-exposure of the child to Catalan or Spanish. For instance, for a child
-who is reportedly exposure to Catalan 80\% of the time, and to Spanish
-20\% of the time, the expected exposure score to the Catalan word
-\emph{cavall} {[}\emph{horse}{]}---with a lexical frequency of ---would
-be , while that of its translation to Spanish \emph{caballo} would be .
-
-We defined \emph{Cognateness} as the phonological similarity between
-each word-form and its translation. For each translation equivalent, we
-used the \texttt{stringdist}
-(\protect\hyperlink{ref-vanderloo2014stringdist}{Loo, 2014}) R package
-to calculate the Levenshtein distance between the Catalan and the
-Spanish phonological transcriptions of the word-forms. The Levenshtein
-distance measures the number of editions (insertions, deletions, or
-substitutions) that one string of phonemes/characters must go through to
-become identical to the other
-(\protect\hyperlink{ref-levenshtein1966binary}{Levenshtein, 1966}). We
-divided the Levnehstein distance of each translation equivalent by the
-length of the longest word-form to correct for word length (longer
-strings are likely to show a larger number of mismatches). Finally, we
-subtracted the result from one so that it could be interpreted in terms
-of phonological similarity, instead of phonological distance. This led
-to a distance metric that ranged from zero to one, where zero indicates
-that both word-forms are completely different (e.g., /'taw.l5-'me.sa/,
-\emph{table}), and one indicates that the two word-forms are identical
-(e.g., /'mar-'mar/, \emph{sea}) (see
-\protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
-\protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020};
-\protect\hyperlink{ref-heeringa2003norwegian}{Heeringa \& Gooskens,
-2003}; \protect\hyperlink{ref-laing2022phonological}{Laing, 2022} for
-similar approaches;
-\protect\hyperlink{ref-schepens2012distributions}{Schepens et al.,
-2012}).
-
-Predictors were standardised before entering the model by subtracting
-the mean of the predictor from each value and dividing the result by the
-standard deviation of the predictor. Participant-level and item-level
-random intercepts and slopes were included where appropriate, according
-to the structure of the data
-(\protect\hyperlink{ref-barr2013random}{Barr et al., 2013}). We
-specified a weakly informative prior around the parameters of the model.
-Equation~\ref{eq-linear} shows a detailed description of the model.
+the phonological similarity between translation equivalents). We added
+these variables as main effects, together with the two-way and three-way
+interactions between \emph{Age}, \emph{Exposure}, and
+\emph{Cognateness}. Participant-level and item-level random intercepts
+and slopes were included where appropriate, according to the structure
+of the data (\protect\hyperlink{ref-barr2013random}{Barr et al., 2013}).
+We specified a weakly informative prior around the parameters of the
+model. Equation~\ref{eq-linear} shows a detailed description of the
+model.
 
@@
 
+\hypertarget{model-predictors}{%
+\subsubsection{Model predictors}\label{model-predictors}}
+
+We developed the \emph{Exposure} predictor to account for the fact that
+bilinguals' exposure to a given word-form is not only a function of the
+word-forms lexical frequency, but also of the quantitative input they
+receive from the language such word-form belongs to, we expressed
+lexical frequencies as the product between both variables. First, we
+extracted the child-directed lexical frequency of each word from the
+CHILDES database
+(\protect\hyperlink{ref-macwhinney2000childes}{MacWhinney, 2000};
+\protect\hyperlink{ref-sanchez2019childesdb}{Sanchez et al., 2019}).
+Using the corresponding lexical frequencies directly from Catalan and
+Spanish was not possible due to the low number of Catalan participants
+and tokens available in their corresponding CHILDES corpora, so they
+were extracted from the English corpora instead. We then mapped the
+lexical frequencies of the English words to their Catalan and Spanish
+translations (see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
+et al., 2020} for a similar approach), and transformed them to Zipf
+scores (\protect\hyperlink{ref-vanheuven2014subtlexuk}{Van Heuven et
+al., 2014}; \protect\hyperlink{ref-zipf1949human}{Zipf, 1949}). We then
+multiplied the resulting lexical frequencies by the reported degree of
+exposure of the child to Catalan or Spanish. For instance, for a child
+whose degree of exposure is 80\% for Catalan and 20\% for Spanish, the
+expected \emph{Exposure} score to the Catalan word \emph{cavall}
+{[}\emph{horse}{]}---with a lexical frequency of ---would be , while
+that of its translation to Spanish \emph{caballo} would be .
+
+We defined \emph{Cognateness} as the phonological similarity between
+each word-form and its translation. For each translation equivalent, we
+used the \texttt{stringdist}
+(\protect\hyperlink{ref-vanderloo2014stringdist}{Loo, 2014}) R package
+to calculate the Levenshtein distance between the Catalan and the
+Spanish phonological transcriptions of the word-forms. The Levenshtein
+distance measures the number of editions (insertions, deletions, or
+substitutions) that one string of phonemes/characters must go through to
+become identical to the other
+(\protect\hyperlink{ref-levenshtein1966binary}{Levenshtein, 1966}). We
+divided the Levenshtein distance of each translation equivalent by the
+length of the longest word-form to correct for word length (longer
+strings are likely to show a larger number of mismatches). Finally, we
+subtracted the result from one so that it could be interpreted in terms
+of phonological similarity, instead of phonological distance. This led
+to a distance metric that ranged from zero to one, where zero indicates
+that both word-forms are completely different (e.g.,
+/\textipa{"taw.l5}-\textipa{"me.sa}/, \emph{table}), and one indicates
+that the two word-forms are identical (e.g.,
+/\textipa{"mar}-\textipa{"mar}/, \emph{sea}) (see
+\protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
+\protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020};
+\protect\hyperlink{ref-heeringa2003norwegian}{Heeringa \& Gooskens,
+2003}; \protect\hyperlink{ref-laing2022phonological}{Laing, 2022} for
+similar approaches;
+\protect\hyperlink{ref-schepens2012distributions}{Schepens et al.,
+2012}).
+
+Predictors were standardised before entering the model by subtracting
+the mean of the predictor from each value and dividing the result by the
+standard deviation of the predictor.
+
+\hypertarget{statistical-inference}{%
+\subsubsection{Statistical inference}\label{statistical-inference}}
+
 We assessed the practical relevance of the estimated regression
@@
 the ROPE (and therefore should be considered equivalent to zero). For
-example, a 80\% overlap between the HDI of a coefficient's posterior
-distribution and the indicates that, given our data, there is a 80\%
-probability that the true value of the coefficient falls within the
-ROPE, and can therefore be considered equivalent to zero.
+example, \(p(\text{ROPE})=.80\) indicates that, given our data, there is
+a 80\% probability that the true value of the coefficient falls within
+the ROPE, and can therefore be considered equivalent to zero.
 
@@
 (\protect\hyperlink{ref-carpenter2017stan}{Carpenter et al., 2017}). We
-ran four MCMC chains with 1,000 iterations each and an additional 1,000
-warm-up iterations per chain. Model posterior draws and predictions were
-handled using the \texttt{tidybayes}
+ran four iteration chains with 1,000 iterations each and an additional
+1,000 warm-up iterations per chain. Model posterior draws and
+predictions were handled using the \texttt{tidybayes}
 (\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) and
@@
 
-\renewcommand{\arraystretch}{1.25}
-
-Model posterior showed adequate MCMC convergence diagnostics for all
-parameters (\(\hat{R} \le 1.06\), \(0.022 \le N_{eff}/N < 1.1001\)) and
-little evidence of multicollinearity. Table~\ref{tbl-coefs} shows the
-summary of the posterior distribution of the fixed regression
+The model posterior showed adequate chain convergence diagnostics for
+all parameters (\(\hat{R} \le 1.06\), \(0.022 \le N_{eff}/N < 1.1001\))
+and little evidence of multicollinearity. Table~\ref{tbl-coefs} shows
+the summary of the posterior distribution of the fixed regression
 coefficients, and their degree of overlap with the ROPE. For
@@
 \centering
-\begin{tabular}{lrlr}
+\begin{tabular}{lclr>{}c}
 \toprule
- & Median & 95\% HDI & $p(\text{ROPE})$\\
+ & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
 \midrule
 \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Intercepts (at 22 months)}}\\
-\hspace{1em}Comprehension and Production & 0.4378 & {}[0.379, 0.496] & 0.0884\\
-\hspace{1em}Comprehension & 0.9363 & {}[0.92, 0.949] & 0.0000\\
+\multicolumn{5}{l}{\textbf{Intercepts (at 22 months)}}\\
+\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & .088 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d082c12c1b.pdf}\\
+\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d08785c6c0e.pdf}\\
 \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Slopes (upper bound)}}\\
-\hspace{1em}Age (+1 SD, 4.87, months) & 0.4049 & {}[0.357, 0.451] & 0.0000\\
-\hspace{1em}Exposure (+1 SD, 1.81) & 0.2333 & {}[0.201, 0.268] & 0.0000\\
-\hspace{1em}Cognateness (+1 SD, 0.26) & 0.0584 & {}[0.014, 0.104] & 0.0371\\
-\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.0620 & {}[-0.086, -0.036] & 0.0000\\
-\hspace{1em}Age $\times$ Exposure & 0.0714 & {}[0.039, 0.104] & 0.0000\\
-\hspace{1em}Age $\times$ Cognateness & 0.0141 & {}[0, 0.026] & 0.9847\\
-\hspace{1em}Exposure $\times$ Cognateness & -0.0569 & {}[-0.069, -0.046] & 0.0000\\
-\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.0180 & {}[-0.027, -0.01] & 0.9747\\
+\multicolumn{5}{l}{\textbf{Slopes (upper bound)}}\\
+\hspace{1em}Age (+1 SD, 4.87, months) & 0.405 & {}[0.357, 0.451] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d087e6a2c55.pdf}\\
+\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d086eae2508.pdf}\\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .037 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d084c7d3767.pdf}\\
+\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d0815d12d78.pdf}\\
+\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d082a9a1cf2.pdf}\\
+\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .985 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d083e54715a.pdf}\\
+\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d087de616c.pdf}\\
+\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .975 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d08238a53bc.pdf}\\
 \bottomrule
@@
 the effect of the word exposure index differed across ages: older
-children were more sensitive likely to acquire words with higher
-exposure rate than younger children.
+children were more likely to acquire words with a higher exposure rate
+than younger children.
 
@@
 when exposure rate was low (e.g., -1 SD), cognateness increased the
-probability of acquisition substantially. This effect was negligible
-when for words with median or high exposure (+1 SD) (see
+probability of acquisition substantially. This effect was negligible for
+words with median or high exposure (+1 SD) (see
 Figure~\ref{fig-marginal}).
@@
 
-\caption{\label{fig-marginal}Posterior marginal effects. Lines
-correspond to 100 posterior-predicted medians. Different colours
-indicate different levels of cognateness (phonological similarity).
-Predictions are presented separately for different degrees of word
-exposure index: little exposure to the word, mean exposure, and high
-exposure. Predictions for Comprehension are shown on top and predictions
-for Comprehension and Production are shown on the bottom. In-sample
-predictions lie inside the grey rectangles.}
+\caption{\label{fig-marginal}Posterior marginal effects. Lines and error
+bands correspond to the mean and 95\% credible interval of the
+posterior-predicted means. Different colours indicate different levels
+of cognateness (phonological similarity). Predictions are presented
+separately for different degrees of word exposure index: little exposure
+to the word, mean exposure, and high exposure. Predictions for
+Comprehension are shown on top and predictions for Comprehension and
+Production are shown on the bottom. In-sample predictions lie inside the
+grey rectangles.}
 
@@
 
-This study investigated the impact of cognateness on the early bilingual
-lexicon. We used Bayesian Item Response Theory to model the acquisition
-trajectories of a large sample of Catalan and Spanish words, estimating
-the effect of cognateness on the probability of acquisition. This model
-corrected for participants' age, word-form length (number of phonemes),
-and a novel measure of participants' exposure rate to each word.
-Exposure rates were calculated as a language exposure-weighted lexical
-frequency score in which each word-form's lexical frequency was
+This study investigated the impact of cognateness (i.e., phonological
+similarity between each word-form and its translation) on the early
+bilingual lexicon. We used Bayesian Item Response Theory to model the
+acquisition trajectories of a large sample of Catalan and Spanish words,
+estimating the effect of cognateness on the probability of acquisition.
+This model corrected for participants' age, word-form length (number of
+phonemes), and a novel measure of participants' exposure rate to each
+word. Exposure rates were calculated as a language exposure-weighted
+lexical frequency score in which each word-form's lexical frequency was
 corrected by the degree to which the participant was exposed to each
@@
 bilingual toddlers learning two typologically close languages
-(e.g.~shared many cognates, like., English-German) showed larger
+(e.g.~shared many cognates, like., English-Dutch) showed larger
 vocabulary sizes than those learning typologically distant languages
@@
 subject to measurement error induced by caregivers who may sometimes
-overestimate or underestimate participants' true probability of
-acquisition of words (e.g.,
+overestimate or underestimate participants' true probability of word
+acquisition (e.g.,
 \protect\hyperlink{ref-houston-price2007discrepancy}{Houston-Price et

### 2023-07-03 manuscript/_appendix/_model-doe.qmd
Add comments by DAV

--- 2023-05-19 manuscript/_appendix/_model-doe.qmd
@@
+As a robustness check, we fit a model similar to the one described in the main manuscript, but including lexical frequency and language degree of exposure as separate predictors, instead of the composite measure *Exposure*.
+Language degree of exposure (*DoE*) was included in interaction with *Age* and *Cognateness*, while lexical frequency (*Frequency*) was included as a main effect.
+
 ```{r fig-coefs-doe}
 #| label: fig-coefs-doe
-#| fig-cap: "Posterior distribution of regression coefficients in the probability scale"
-#| fig-note: "Median: median of the posterior distribution in the logit scale.
-95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data.
-On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE).
-This proportion indicates the probability of the true value of the coefficient being equivalent to zero.
-Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
+#| fig-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency and degree of exposure (DoE) separately.
+Dots and error bars indicate the median and 95% HDI of the posterior distribution of each coefficient.
+Coefficients of the *Exposure* model are indicated in grey with triangles.
+Coefficients of the *Frequency + DoE* model are indicated in black with circless.
+The region of practical equivalence (ROPE) around zero is indicated with a grey area and a grey line."
+#| fig-note: ""
 #| echo: false
@@
              xmax = +0.1/4,
-             fill = "grey70") +
-    geom_errorbar(linewidth = 1, width = 0.25) +
-    geom_point(size = 2.5) +
+             fill = "grey90") +
+    geom_vline(xintercept = 0, 
+               linetype = "dashed",
+               linewidth = 3/4,
+               colour = "grey") +
+    geom_errorbar(linewidth = 3/4, 
+                  position = position_dodge(width = 0.5),
+                  width = 0.3) +
+    geom_point(size = 2.5,
+               position = position_dodge(width = 0.5)) +
     labs(x = "Median and 95% HDI of the posterior distribution",

### 2023-07-03 manuscript/_extensions/mikemahoney218/arxiv/partials/title.tex
Add comments by DAV

$-- Provides configuration of document metadata for writing the title block.
$-- Note that in addition to these templates and partials, Quarto will also make normalized authors and affiliations data available to the template,
$-- making is easy to write custom title blocks against a standard schema.
$--
$-- %%%% TODO %%%%%
$-- Customize is needed, like below for printing the authors. Otherwise remove this partials to use Quarto default one.
$-- %%%%%%%%%%%%%%%%
$if(linenumbers)$
\usepackage{lineno}
\linenumbers
$endif$
$if(doublespacing)$
\usepackage{setspace}
\doublespacing
$endif$
$if(date)$
\renewcommand{\today}{$date$}
$endif$
\newcommand{\runninghead}{A Preprint }
$if(runninghead)$
\renewcommand{\runninghead}{$runninghead$ }
$endif$
$if(title)$
\title{$title$$if(thanks)$\thanks{$thanks$}$endif$}
$endif$
$if(subtitle)$
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{$subtitle$}
$endif$
\author{${ by-author:_authors.tex()[\And] }}
\date{$date$}

### 2023-07-06 docs/index.qmd
Fix appendix

@@
 
-abstract: |
-    Language non-selectivity is prominent feature of the bilingual lexicon.
-An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates.
-The specific mechanisms behind such facilitation are unclear.
-We present an account of early bilingual vocabulary growth in which cognateness interacts with lexical frequency and language exposure to facilitate the word acquisition.
-We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals toddlers.
-We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler.
-We found cognateness facilitation effect only in low-frequency words.
-Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.
-
 fig-dpi: 1000
-
 ---
@@
 # load objects -----------------------------------------------------------------
-
-library(patchwork)
-library(gt)
-library(ggplot2)
-
 targets::tar_config_set(store = here::here("_targets"),
@@
 
-targets::tar_load_globals()
+yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
+title <- yaml_metadata$title
+abstract <- yaml_metadata$abstract
+thanks <- yaml_metadata$thanks
 
-targets::tar_load(c(bvq_data, childes, items, 
-                    participants, responses))
-
-targets::tar_load(c(model_fit, model_summary, model_epreds,
-                    model_convergence, model_fit_prior, 
-                    model_ppcs, posterior_doe_summary,
-                    model_draws, model_summary,
-                    posterior_doe_draws))
-
-targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))
-
-# process objects
-items <- bvq_data$pool  |>  
-    mutate(item = str_remove(item, "cat_|spa_")) |> 
-    select(item, language, semantic_category, class) |> 
-    right_join(items, by = c("item", "language"))
-
-rope_interval <- c(lower = -0.1, upper = 0.1)
-
-# set ggplot theme and colour palette ------------------------------------------
-
-my_theme <- theme_minimal() +
-    theme(panel.grid = element_line(colour = "grey",
-                                    linetype = "dotted"),
-          axis.line = element_line(colour = "black"),
-          text = element_text(size = 12, colour = "black"),
-          axis.text = element_text(colour = "black"))
-
-theme_set(my_theme)
-
-bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))
-
-clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
-set.seed(888)
-
-options(ggplot2.ordinal.fill = clrs,
-        ggplot2.ordinal.colour = clrs,
-        ggplot2.discrete.fill = clrs,
-        ggplot2.discrete.colour = clrs,
-        ggplot2.continuous.fill = scale_color_gradient,
-        ggplot2.continuous.colour = scale_color_gradient,
-        knitr.graphics.error = FALSE) 
 ```
 
-::: {.callout-tip}
-Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
+::: callout-tip
+Data and materials are available at [OSF](https://osf.io/hy984/), and
+code is available at
+[GitHub](https://github.com/gongcastro/trajectories).
 :::
 
-# Data
+`r abstract`
 
-# Outputs
+`r thanks`
+
+# Links
+
+![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height=100px}
+
+![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height=100px}
+
+![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height=100px}
+
+{{< include ../README.md >}}
 

### 2023-07-06 manuscript/_appendix/_syllable-frequency.qmd
Fix appendix

@@
 #| message: false
-items |> 
+tbl_data <- items |> 
     distinct(te, item, language, meaning, xsampa, n_syll) |> 
@@
     sample_n(15) |> 
-    gt() |> 
-    cols_hide(c(te, item_catalan, item_spanish)) |> 
-    tab_spanner("Catalan", matches("catalan")) |> 
-    tab_spanner("Spanish", matches("spanish")) |> 
-    cols_label(meaning = "Translation",
-               label_catalan = "Item",
-               xsampa_catalan = "X-SAMPA",
-               n_syll_catalan = "Syllables",
-               label_spanish = "Item",
-               xsampa_spanish = "X-SAMPA",
-               n_syll_spanish = "Syllables") |> 
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) |> 
-    tab_style(cell_text(style = "italic"),
-              cells_column_labels())
+    select(-c(te, item_catalan, item_spanish)) |> 
+    relocate(meaning, matches("catalan"), matches("spanish"))
+
+col_names <- c("Translation", "Item", "X-SAMPA", "Syllables",
+               "Item", "X-SAMPA", "Syllables")
+
+tbl_data |> 
+    kable(format = "latex", 
+          digits = 3,
+          booktabs = TRUE,
+          col.names = col_names, 
+          align = "lllcllc") 
 ```
@@
 No words had more than five syllables (see @fig-syll-number).
-We extracted lexical frequencies from the English corpora in the CHILDES database [@macwhinney2000childes; sanchez2019childesdb].
+We extracted lexical frequencies from the English corpora in the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
 Using the Catalan and Spanish corpora was not possible due to the low number of children and tokens included in the corpora.

### 2023-07-06 manuscript/_extensions/mikemahoney218/arxiv/partials/_authors.tex
Fix appendix

$-- You can use as many custom partials as you need. Convention is to prefix name with '_'
$-- It can be useful to use such template to split some template parts in smaller pieces, which is easier to reuse.
$-- This '_custom.tex' is used on 'title.tex' as example.
$-- See other existing format in quarto-journals/ organisation.
$-- %%%% TODO %%%%%
$-- Use it if you need to insert content at this specific place of the main Pandoc's template. Otherwise, remove it.
$-- Here we are using it to format the authors part of the template.
$-- %%%%%%%%%%%%%%%

\textbf{$it.name.literal$}$if(it.orcid)$~\orcidlink{$it.orcid$}$endif$$for(it.affiliations/first)$\\$it.department$\\$it.name$\\$if(it.city)$$it.city$$if(it.postal-code)$,\ $it.postal-code$$endif$$endif$$endfor$\\$if(it.email)$\texttt{\href{mailto:$it.email$}{$it.email$}}$endif$

### 2023-07-06 manuscript/appendix.qmd
Add comments by DAV
Fix appendix

@@
 ---
-title: "Appendix"
 csl: "../assets/apa7.csl"
@@
 format:
-  pdf:
-    documentclass: article
-    papersize: letter
-    fontsize: 12pt
-    linestretch: 1.5
-    header-includes:
-    - \usepackage{tipa}
-    geometry:
-    - top=1in
-    - bottom=1in
-    - left=1in
-    - right=1in
-    colorlinks: true
-    linkcolor: DodgerBlue
-    urlcolor: DodgerBlue
-    citecolor: DodgerBlue
-    link-citations: true
-    biblio-config: false
-    block-headings: false
-    indent: true
+  arxiv-pdf:
+    keep-tex: true  
+    linenumbers: true
+    doublespacing: false
+    runninghead: "Cognate beginnings to bilingual lexical acquistion (Appendix)"
+    include-in-header: 
+      text: |
+        \usepackage{tipa}
+        \usepackage{booktabs}
 ---
@@
 # load objects -----------------------------------------------------------------
-
-library(patchwork)
-library(gt)
+library(knitr)
+library(kableExtra)
 library(ggplot2)
@@
 
-targets::tar_load(c(bvq_data, childes, items, 
+targets::tar_load(c(bvq_data, childes, items, model_summary,
                     participants, responses))
 
-targets::tar_load(c(model_fit, model_summary, model_epreds,
-                    model_convergence, model_fit_prior, 
-                    model_ppcs, posterior_doe_summary,
-                    posterior_doe_draws))
+targets::tar_load(c(posterior_doe_summary, posterior_doe_draws))
 
@@
 
-# process objects
-items <- bvq_data$pool  |>  
-    mutate(item = str_remove(item, "cat_|spa_")) |> 
-    select(item, language, semantic_category, class) |> 
-    right_join(items, by = c("item", "language"))
-
-# set ggplot theme and colour palette ------------------------------------------
-
-my_theme <- function() {
-    theme_ggdist() +
+my_theme <- theme_minimal() +
     theme(panel.grid = element_line(colour = "grey",
@@
           axis.text = element_text(colour = "black"))
-}
 
-theme_set(my_theme())
-
-bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))
-
-set.seed(888)
-
-options(knitr.graphics.error = FALSE)
+theme_set(my_theme)
 ```

### 2023-07-06 manuscript/_appendix/_model-doe.qmd
Fix appendix

@@
 Language degree of exposure (*DoE*) was included in interaction with *Age* and *Cognateness*, while lexical frequency (*Frequency*) was included as a main effect.
+@tbl-coefs-doe shows a comparison between the posterior distribution of the regression coefficients of both models.
+Overall, results are equivalent.
 
-```{r fig-coefs-doe}
-#| label: fig-coefs-doe
-#| fig-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency and degree of exposure (DoE) separately.
-Dots and error bars indicate the median and 95% HDI of the posterior distribution of each coefficient.
-Coefficients of the *Exposure* model are indicated in grey with triangles.
-Coefficients of the *Frequency + DoE* model are indicated in black with circless.
-The region of practical equivalence (ROPE) around zero is indicated with a grey area and a grey line."
-#| fig-note: ""
+```{r tbl-coefs-doe}
+#| label: tbl-coefs-doe
+#| fig-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency (*Frequency*) and degree of exposure (*DoE*) separately.
+Median: median of the posterior distribution in the probability scale.
+95\\% HDI: 95\\% highest density interval of the distribution.
+p(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
+In the right column, we show a graphical representation of the median and 95% HDI of each coefficient, with a reference dotted line indicating the location of zero."
 #| echo: false
@@
 #| warning: false
-#| out-width: 100%
-#| fig-height: 4
-#| fig-width: 7
 # summarise posterior draws
-bind_rows("Exposure model" = model_summary,
+tbl_data <- bind_rows("Exposure model" = model_summary,
                       "DoE-Frequency model" = posterior_doe_summary,
-          .id = "model") |> 
+                      .id = ".model") |> 
     mutate(
@@
     filter(.type == "Slopes") |> 
-    ggplot(aes(.median, .variable_name,
-               xmin = .lower, xmax = .upper,
-               colour = model,
-               shape = model)) +
-    annotate(geom = "rect",
-             ymin = -Inf,
-             ymax = Inf,
-             xmin = -0.1/4,
-             xmax = +0.1/4,
-             fill = "grey90") +
-    geom_vline(xintercept = 0, 
-               linetype = "dashed",
-               linewidth = 3/4,
-               colour = "grey") +
-    geom_errorbar(linewidth = 3/4, 
-                  position = position_dodge(width = 0.5),
-                  width = 0.3) +
-    geom_point(size = 2.5,
-               position = position_dodge(width = 0.5)) +
-    labs(x = "Median and 95% HDI of the posterior distribution",
-         y = "Predictor",
-         colour = "Model") +
-    scale_colour_grey(start = 0, end = 0.6) +
-    theme(legend.position = "none",
-          axis.title.y = element_blank(),
-          panel.grid.major.y = element_blank())
+    select(.model, .variable_name, .median, .lower, .upper, .rope_overlap) |> 
+    mutate(
+        across(.median:.upper, plogis),
+        .variable_name = gsub("×", '$\\\\times$', .variable_name),
+        .hdi = paste0("[", round(.lower, 3),
+                      ", ", round(.upper, 3), "]"),
+        .rope_overlap = sub("^(-?)0.", "\\1.",
+                            sprintf("%.3f", .rope_overlap))) |> 
+    select(.variable_name, .median, .lower, .upper, .hdi, .rope_overlap)
+
+tbl_data |> 
+    mutate(viz = "") |> 
+    select(-c(.lower, .upper)) |> 
+    knitr::kable(
+        format = "latex", 
+        escape = FALSE,
+        col.names = c("",
+                      "$\\beta$",
+                      "95\\% HDI",
+                      "$p(\\text{ROPE})$",
+                      ""),
+        digits = 3,
+        booktabs = TRUE,
+        align = "lclr",
+        linesep = ""
+    ) |> 
+    kableExtra::pack_rows("Model: Exposure",
+                          start_row = 1,
+                          end_row = 8) |> 
+    kableExtra::pack_rows("Model: Frequency & DoE",
+                          start_row = 9,
+                          end_row = 17)
 ```

### 2023-07-06 manuscript/appendix.tex
Fix appendix

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Latin Modern Roman}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{tipa}
\usepackage{booktabs}
\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\usepackage{lineno}
\linenumbers
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Cognate beginnings to bilingual lexical
acquistion (Appendix) }
\author{}
\date{}
\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{appendix-a-frequency-and-language-exposure-as-separate-predictors}{%
\section*{Appendix A: frequency and language exposure as separate
predictors}\label{appendix-a-frequency-and-language-exposure-as-separate-predictors}}
\addcontentsline{toc}{section}{Appendix A: frequency and language
exposure as separate predictors}

As a robustness check, we fit a model similar to the one described in
the main manuscript, but including lexical frequency and language degree
of exposure as separate predictors, instead of the composite measure
\emph{Exposure}. Language degree of exposure (\emph{DoE}) was included
in interaction with \emph{Age} and \emph{Cognateness}, while lexical
frequency (\emph{Frequency}) was included as a main effect.
Table~\ref{tbl-coefs-doe} shows a comparison between the posterior
distribution of the regression coefficients of both models. Overall,
results are equivalent.

\begin{table}

\caption{\label{tbl-coefs-doe}\textbf{?(caption)}}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-coefs-doe-1}}

{\centering

\begin{tabular}{lclrl}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Model: Exposure}}\\
\hspace{1em}Age (+1 SD, 4.87, months) & 0.600 & {}[0.588, 0.611] & .000 & \\
\hspace{1em}Exposure (+1 SD, 1.81) & 0.558 & {}[0.55, 0.567] & .000 & \\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.515 & {}[0.504, 0.526] & .037 & \\
\hspace{1em}Length (+1 SD, 1.56 phonemes) & 0.485 & {}[0.478, 0.491] & .000 & \\
\hspace{1em}Age $\times$ Exposure & 0.518 & {}[0.51, 0.526] & .000 & \\
\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.5, 0.506] & .985 & \\
\hspace{1em}Exposure $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .975 & \\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Model: Frequency \& DoE}}\\
\hspace{1em}Age (+1 SD, 4.87, months) & 0.601 & {}[0.59, 0.612] & .000 & \\
\hspace{1em}DoE (+1 SD, 0.3) & 0.558 & {}[0.549, 0.566] & .000 & \\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.516 & {}[0.506, 0.528] & .016 & \\
\hspace{1em}Phonemes (+1 SD, 1.56 phonemes) & 0.486 & {}[0.479, 0.492] & .000 & \\
\hspace{1em}Frequency (+1 SD, 0.19) & 0.527 & {}[0.516, 0.54] & .000 & \\
\hspace{1em}Age $\times$ DoE & 0.518 & {}[0.51, 0.527] & .000 & \\
\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.501, 0.507] & .954 & \\
\hspace{1em}DoE $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
\hspace{1em}Age $\times$ DoE $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .962 & \\
\bottomrule
\end{tabular}

}

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}

{\centering

Posterior distribution of regression coefficients of the model including
the \emph{Exposure} composite predictor, and of the model including
lexical frequency (\emph{Frequency}) and degree of exposure (\emph{DoE})
separately. Median: median of the posterior distribution in the
probability scale. 95\% HDI: 95\% highest density interval of the
distribution. p(ROPE): overlap between the 95\% HDI and the ROPE,
indicating the posterior probability that the true value of the
coefficient is equivalent to zero. In the right column, we show a
graphical representation of the median and 95\% HDI of each coefficient,
with a reference dotted line indicating the location of zero.

}

\end{minipage}%

\end{table}

\newpage{}

\hypertarget{appendix-b-frequency-and-language-exposure-as-separate-predictors}{%
\section*{Appendix B: frequency and language exposure as separate
predictors}\label{appendix-b-frequency-and-language-exposure-as-separate-predictors}}
\addcontentsline{toc}{section}{Appendix B: frequency and language
exposure as separate predictors}

We define syllable frequency as the rate of appearance of individual
syllables in the word-forms included in the Barcelona Vocabulary
Questionnaire (BVQ)
(\protect\hyperlink{ref-garcia-castro2023bvq}{Garcia-Castro et al.,
2023}). Each item corresponds to a Catalan or Spanish word, and has an
associated phonological transcription in X-SAMPA format
(\protect\hyperlink{ref-wells1995computercoding}{Wells, 1995}). These
transcriptions are syllabified. Some examples:

\hypertarget{tbl-syll-items}{}
\begin{table}
\caption{\label{tbl-syll-items}Sample of items included in the BVQ questionnaire and their syllabified
SAMPA transcriptions in Catalan and Spanish }\tabularnewline

\centering
\begin{tabular}{lllcllc}
\toprule
Translation & Item & X-SAMPA & Syllables & Item & X-SAMPA & Syllables\\
\midrule
ladybug & marieta & m@.4i"E.t@ & 4 & mariquita & ma.4i"ki.ta & 4\\
beans & mongetes & muJ"ZE.t@s & 3 & alubias & a"lu.bjas & 3\\
tongue & llengua & "LeN.gw@ & 2 & lengua & "len.gwa & 2\\
kitchen & cuina & "kuj.n@ & 2 & cocina & ko"Ti.na & 3\\
cheek & galta & "ga5.t@ & 2 & mejilla & me"xi.La & 3\\
\addlinespace
purse & bossa & "bo.s@ & 2 & bolso & "bol.so & 2\\
farm & granja & g4aJ.Z@ & 2 & granja & "g4an.xa & 2\\
sausage & salsitxa & s@5"si.t͡S@ & 3 & salchicha & sal"tSi.tSa & 3\\
cereal & cereals & s@.4e"a5s & 3 & cereales & Te.4e"a.les & 4\\
bottle & ampolla & @m"po.L@ & 3 & botella & bo"te.La & 3\\
\addlinespace
boat & vaixell & b@"SeL & 2 & barco & "ba4.ko & 2\\
family & família & fa"mi.lia & 3 & familia & fa"mi.lia & 3\\
balloon & globus & "g5O.Bus & 2 & globo & "glo.Bo & 2\\
bee & abella & @"BE.L@ & 3 & abeja & a"Be.xa & 3\\
earring & arracada / arracades & @.r@"ka.D@ & 4 & pendientes & pen"Djen.tes & 3\\
\bottomrule
\end{tabular}
\end{table}

Most Catalan and Spanish words had two syllables, with Spanish words
having three and four syllables more often than Catalan words. Less than
1\% of the words included in the analyses presented in the main body of
the manuscripts had five syllables. No words had more than five
syllables (see Figure~\ref{fig-syll-number}). We extracted lexical
frequencies from the English corpora in the CHILDES database
(\protect\hyperlink{ref-macwhinney2000childes}{MacWhinney, 2000};
\protect\hyperlink{ref-sanchez2019childesdb}{Sanchez et al., 2019}).
Using the Catalan and Spanish corpora was not possible due to the low
number of children and tokens included in the corpora.

\begin{figure}

{\centering \includegraphics{appendix_files/figure-pdf/fig-syll-number-1.pdf}

}

\caption{\label{fig-syll-number}Distribution of the number of syllables
in Catalan and Spanish}

\end{figure}

We now present how syllable frequencies were calculated. Every exposure
to a word-form also counts as a exposure to each of the syllables that
make up such word. Every time a child hears the word \emph{casa}
{[}house{]}, they are exposed to the syllables \emph{ca} and \emph{sa}.
Syllables that appear embedded in words with higher lexical frequency
will also be more frequent. To compute the relative frequency of each
syllable in Catalan and Spanish (i.e., how many times the syllables
appears in every million words in Catalan or Spanish speech), we summed
the relative lexical frequency in CHILDES of every word that contains
such syllable in the corresponding language. Figure~\ref{fig-syll-freq}
shows the distribution of frequencies across syllables in Catalan and
Spanish. In the log10 scale, syllable frequencies in Catalan and Spanish
followed a slightly asymmetric distribution, with most syllables scoring
around 1,000 counts per million, and a longer tail to the right of the
distribution.

\begin{figure}

{\centering \includegraphics{appendix_files/figure-pdf/fig-syll-freq-1.pdf}

}

\caption{\label{fig-syll-freq}Distribution of apositional syllable
frequencies in Spanish and Catalan}

\end{figure}

To estimate the association between word-level syllabic frequency and
cognateness, while controlling for the number of syllables in the word,
as words are expected to necessarily increase the syllabic frequency of
the word), we fit a multilevel, Bayesian linear regression model with
syllabic frequency (the sum of the syllabic frequency of the syllables
in a word) as response variable, and the main effect of the number of
syllables (\(Syllables\)) and \(Cognateness\) (Levenshtein similarity
between a word and its translation equivalent,
\protect\hyperlink{ref-levenshtein1966binary}{Levenshtein, 1966}) as
predictors. We added translation equivalent-level random effects for the
intercept and the main effect of \(Syllables\) (some translation pairs
had a different number of syllables in each language). We used a
Gaussian distribution to model syllabic frequency scores after
standardising this variable and the predictors. We used a weakly
informative prior for all parameters involved in the model (see
Equation~\ref{eq-model-syllables} for a formal equation of this model
and its prior). We conducted statistical inference by evaluating the
proportion of the 95\% highest density interval (HDI) of the posterior
posterior distribution of each coefficient that fell into the region of
practical equivalence (ROPE, see the main manuscript for a more detailed
explanation, \protect\hyperlink{ref-kruschke2018bayesian}{Kruschke \&
Liddell, 2018}).

\begin{equation}\protect\hypertarget{eq-model-syllables}{}{
\begin{aligned}
\text{Syllable frequency} &\sim \mathcal{N}(\mu, \sigma)\\
\mu &= (\beta_0 + u_{0_{i}}) + (\beta_1 + u_{1_{i}}) \text{Syllables} + \beta_2 \text{Cognateness} \\
\beta_{0-3} &\sim \mathcal{N}(0, 10) \\
u_{{0-1}_{i}} &\sim \mathcal{N}(0, \sigma_{u_i}) \\
\sigma_y & \sim \text{Exponential}(2)\\
\Big(\begin{smallmatrix}
u_{0_{i}} \\ 
u_{1_{i}} \\ 
\end{smallmatrix}\Big) &\sim \mathcal{N} 
\Big(\Big(\begin{smallmatrix}0 \\
0\end{smallmatrix}\Big), \Sigma_{u}\Big) \\
\Sigma_{u} &= \Big(\begin{smallmatrix} \\
\sigma_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{u_{1}} \\ 
\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \sigma_{u_{1}} \end{smallmatrix}\Big) \\
\sigma_{u_{0-1}} &\sim \mathcal{N_{+}}(1, 0.1) \\
\rho_{u} &\sim LKJcorr(2) \\
\end{aligned}
}\label{eq-model-syllables}\end{equation}

We fit this model running 4 sampling chains with 1,000 iterations each.
Table~\ref{tbl-syll-coefs} shows a summary of the posterior distribution
of the fixed effects in the model. As expected, words with more
syllables scored higher in syllabic frequency: all posterior draws for
the regression coefficient of the main effect of this predictor fell
outside the ROPE defined between -0.5 and +0.5 (\(\beta\) = 5.64, 95\%
HDI = {[}5.58, 5.71{]}). Keeping the number of syllables constant, the
effect of cognateness was negligible: all of the posterior distributions
of this predictor fell within the ROPE, providing evidence that the true
value of the increment in syllabic frequency for every increase in
cognateness is equivalent to zero (\(\beta\) = 0.01, 95\% HDI =
{[}-0.06, 0.07{]}).

\hypertarget{tbl-syll-coefs}{}
\begin{table}
\caption{\label{tbl-syll-coefs}Posterior distribution of regression coefficients. }\tabularnewline

\centering
\begin{tabular}{lclr>{}c}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
Intercept & 16.089 & {}[16.023, 16.163] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e8078876ee9.pdf}\\
Syllables (+1 SD, 0.802) & 5.642 & {}[5.577, 5.711] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e807c1b3ef.pdf}\\
Cognateness (+1 SD, 0.24) & 0.007 & {}[-0.059, 0.075] & 1.000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e8043796412.pdf}\\
\bottomrule
\end{tabular}
\end{table}

Figure~\ref{fig-syll-marginal} shows the median posterior-predicted
syllabic frequencies for words with one to four syllables, for the whole
range of cognateness values. Overall, cognate words' syllabic frequency
is equivalent to that of non-cognates. This suggests that the cognate
facilitation effect in word acquisition reported in the present study is
not the result from an association between cognateness and higher
syllabic frequencies.

\begin{figure}

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-syll-marginal-1.pdf}

}

\caption{\label{fig-syll-marginal}Posterior-predictions of the syllabic
frequency model. Thicker lines indicate the median of the posterior
predictions, and thinner lines indicate individual posterior
predictions.}

\end{figure}

\newpage{}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-garcia-castro2023bvq}{}}%
Garcia-Castro, G., Ávila-Varela, D. S., \& Sebastian-Galles, N. (2023).
\emph{Bvq: Barcelona vocabulary questionnaire database and helper
functions}. \url{https://gongcastro.github.io/bvq}

\leavevmode\vadjust pre{\hypertarget{ref-kruschke2018bayesian}{}}%
Kruschke, J. K., \& Liddell, T. M. (2018). The bayesian new statistics:
Hypothesis testing, estimation, meta-analysis, and planning from a
bayesian perspective. \emph{Psychonomic Bulletin \&Review}, \emph{25},
178--206. \url{https://doi.org/10.3758/s13423-016-1221-4}

\leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
Levenshtein, V. I. (1966). Binary codes capable of correcting deletions,
insertions, and reversals. \emph{Soviet Physics-Doklady}, \emph{10},
707--710.

\leavevmode\vadjust pre{\hypertarget{ref-macwhinney2000childes}{}}%
MacWhinney, B. (2000). \emph{The {CHILDES} project: The database} (Vol.
2). Psychology Press.

\leavevmode\vadjust pre{\hypertarget{ref-sanchez2019childesdb}{}}%
Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
D., \& Frank, M. C. (2019). Childes-db: A flexible and reproducible
interface to the child language data exchange system. \emph{Behavior
Research Methods}, \emph{51}(4), 1928--1941.

\leavevmode\vadjust pre{\hypertarget{ref-wells1995computercoding}{}}%
Wells, J. C. (1995). \emph{Computer-coding the {IPA}: A proposed
extension of {SAMPA}}. \emph{4}(28), 1995.

\end{CSLReferences}

\end{document}

### 2023-07-06 manuscript/manuscript.qmd
Fix appendix

@@
 ---
-title: "Cognate beginnings to bilingual lexical acquisition"
-short-title: "Cognate beginnings"
-author:
-  - name: Gonzalo Garcia-Castro
-    corresponding: true
-    orcid: 0000-0002-8553-4209
-    email: gonzalo.garciadecastro@upf.edu
-    affiliations:
-      - id: cbc
-        name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-        city: Barcelona, 08005
-  - name: Daniela S. Avila-Varela
-    orcid: 0000-0002-3518-8117
-    email: avila.varela.daniela@gmail.com
-    affiliations: 
-      - ref: cbc
-      - id: ul
-        name: Universidade de Lisboa
-        department: Linguistics Center (CLUL), School of Arts and Humanities
-        city: Lisbon, 1600-214
-  - name: Ignacio Castillejo
-    orcid: 0000-0001-7445-0416
-    email: jignaciocastillejo@gmail.com
-    affiliations: 
-      - id: uam 
-        name: Universidad Autónoma de Madrid
-        department: Departamento de Psicología
-        city: Madrid, 28049
-  - name: Nuria Sebastian-Galles
-    orcid: 0000-0001-6938-2498
-    email: nuria.sebastian@upf.edu
-    affiliations: 
-      - ref: cbc
-keywords:
-  - cognate
-  - word acquisition
-  - vocabulary
-  - bilingualism
-  - item response theory
-  - bayesian
-csl: "../assets/apa7.csl"
-bibliography: "../assets/references.bib"
-link-citations: true
-editor: source
-lang: en
-
-format:
-  arxiv-pdf:
-    keep-tex: true  
-    linenumbers: true
-    doublespacing: false
-    runninghead: "Cognate beginnings to bilingual lexical acquistion"
-    include-in-header: 
-      text: |
-        \usepackage{tipa}
-        \usepackage{booktabs}
 abstract: |
@@
     The authors declare no conflicts of interest with regard to the funding source of this study.
-GGC was supported by the This research was supported by a FPI 2020 research contract (PRE2019-088165) in a project funded by the Spanish Ministry of Science, Innovation, and Universities (PGC2018-101831-B-I00).
+GGC was supported by the a FPI 2020 research contract (PRE2019-088165) in a project funded by the Spanish Ministry of Science, Innovation, and Universities (PGC2018-101831-B-I00).
 DAV was supported by the European Union's Horizon 2023 research and innovation program under Marie Skłodowska–Curie Grant (765556) and a postdoctoral fellowship from the Foundation for Science and Technology of Portugal (UIDB/00214/2020).
@@
 We also thank all families and infants who participated in the experiments.
-fig-dpi: 1000
 ---
@@
 
-library(gt)
+library(knitr)
+library(kableExtra)
 library(ggplot2)
-library(patchwork)
 
@@
 
-Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [cat], that  are phonologically very similar.
+Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [*cat*], that  are phonologically very similar.
 When the child is exposed to /\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in parallel.
 Therefore, this exposure will count as a learning instance for both co-activated forms.
-The case of the non-cognate translation equivalent /\textipa{"gos-"pe.ro}/ would be different.
+The case of the non-cognate translation equivalent /\textipa{"gos-"pe.ro}/ [*dog*] would be different.
 Given the low phonological similarity between both word-forms, an exposure to \textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/ leading to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for /\textipa{"pe.ro}/.
@@
               .by = semantic_category) |> 
-    mutate(total = rowSums(across(a:d)),
-           items = map(items, function(x) tail(x, 2))) |>  
+    mutate(
+        total = rowSums(across(a:d)),
+        items = map(items, function(x) {
+            paste(tail(x, 2),
+                  collapse = ", ")
+        })) |>  
+    unnest_longer(items) |> 
     arrange(desc(total)) |> 
@@
 
-sums_row <- as.data.frame(lapply(tbl_data, sum_rows),
-                          col.names = colnames(tbl_data))
+sums_row <- tbl_data |> 
+    summarise(semantic_category = "N",
+              across(a:d, sum),
+              items = "-")
+
 tbl_data <- rbind(tbl_data, sums_row)
@@
 stan_version <- "2.32.1"
-
 ```
@@
 
-We collected `r format(n_obs, big.mark = ",")` responses to `r format(nrow(n_items), big.mark = ",")` items.
+We collected data for `r format(nrow(n_items), big.mark = ",")` words.
 We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, see @fourtassi2020growth for a similar approach].
@@
 Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
+
+Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
 
@@
 We specified a weakly informative prior around the parameters of the model.
-@eq-linear shows a detailed description of the model.
+@eq-model shows a detailed description of the model.
 
@@
 \end{aligned}
-$$ {#eq-linear}
+$$ {#eq-model}
 
@@
 We then multiplied the resulting lexical frequencies by the reported degree of exposure of the child to Catalan or Spanish.
-For instance, for a child whose degree of exposure is 80% for Catalan and 20% for Spanish, the expected *Exposure* score to the Catalan word *cavall* [*horse*]---with a lexical frequency of `r round(items$freq[items$item=="cat_cavall"], 2)`---would be `r round(items$freq[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq[items$item=="cat_cavall"]*0.20, 2)`.
+For instance, for a child whose degree of exposure is 80% for Catalan and 20% for Spanish, the expected *Exposure* score to the Catalan word *cotxe* [*car*]---with a lexical frequency of `r round(items$freq[items$item=="cotxe"], 2)`---would be `r round(items$freq[items$item=="cotxe"]*0.80, 2)`, while that of its translation to Spanish *coche* would be `r round(items$freq[items$item=="coche"]*0.20, 2)`.
 
@@
 
-Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
 We implemented the model using `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
-We ran four iteration chains with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
+We ran four iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
 Model posterior draws and predictions were handled using the `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
@@
 ```
-The model posterior showed adequate chain convergence diagnostics for all parameters ($\hat{R} \le 1.06$, $0.022 \le N_{eff}/N < 1.1001$) and little evidence of multicollinearity.
+
+Chain convergence diagnostics showed adequate values, and the model posterior showed little evidence of correlation, as indicated by negligible pairwise correlation between the marginal posterior distribution of coefficients.
 @tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients, and their degree of overlap with the ROPE.
@@
 #| label: tbl-coefs
-#| tbl-cap: "Posterior distribution of regression coefficients."
-#| tbl-note: "Median: median of the posterior distribution in the probability scale.
+#| tbl-cap: "Posterior distribution of regression coefficients.
+Median: median of the posterior distribution in the probability scale.
 95\\% HDI: 95\\% highest density interval of the distribution.
-p(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero."
+p(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
+In the right column, we show a graphical representation of the median and 95% HDI of each coefficient, with a reference dotted line indicating the location of zero."
 #| echo: false
@@
 fig_data <- model_epreds |> 
-    summarise(.median = mean(.value),
-              .lower = qi(.value)[,1],
-              .upper = qi(.value)[,2],
+    reframe(.median = mean(.value),
+            .lower = qi(.value, .width = 0.95)[,1],
+            .upper = qi(.value, .width = 0.95)[,2],
             .by = c(ends_with("_std"), .category)) |> 
@@
                linetype = "dotted") +
-    geom_ribbon(alpha = 3/4,
-                colour = NA) +
-    geom_line(linewidth = 2/3) +
+    geom_lineribbon(aes(order = lv),
+                    linewidth = 0.5) +
     labs(x = "Age (months)",
@@
          linetype = "Cognateness (phonological similarity)") +
-    scale_color_grey(start = 0.80, end = 0.1) +
-    scale_fill_grey(start = 0.80, end = 0.1) +
     scale_linetype_manual(values = rev(c("solid", "dashed", "dotdash"))) +
+    scale_colour_grey(start = 0, end = 0) +
+    scale_fill_grey(start = 0.85, end = 0.35) +
     scale_y_continuous(breaks = seq(0, 1, 0.25),
@@
 Capitalising on accumulator models of language acquisition we put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
+
+# References

### 2023-07-06 manuscript/manuscript.tex
Fix appendix

@@
 authors declare no conflicts of interest with regard to the funding
-source of this study. GGC was supported by the This research was
-supported by a FPI 2020 research contract (PRE2019-088165) in a project
-funded by the Spanish Ministry of Science, Innovation, and Universities
-(PGC2018-101831-B-I00). DAV was supported by the European Union's
-Horizon 2023 research and innovation program under Marie
-Skłodowska--Curie Grant (765556) and a postdoctoral fellowship from the
-Foundation for Science and Technology of Portugal (UIDB/00214/2020). NSG
-was supported by Academia 2019 award from the Catalan Institution for
-Research and Advanced Studies (ICREA). We are grateful to Chiara
-Santolin, Ege E. Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada,
-and the rest of the Speech Acquisition and Perception research group for
-their helpful feedback. We thank Xavier Mayoral, Silvia Blanch, and
-Cristina Cuadrado for their technical support, and Cristina Dominguez
-and Katia Pistrin for their efforts in recruiting infants. We also thank
-all families and infants who participated in the experiments.}}
+source of this study. GGC was supported by the a FPI 2020 research
+contract (PRE2019-088165) in a project funded by the Spanish Ministry of
+Science, Innovation, and Universities (PGC2018-101831-B-I00). DAV was
+supported by the European Union's Horizon 2023 research and innovation
+program under Marie Skłodowska--Curie Grant (765556) and a postdoctoral
+fellowship from the Foundation for Science and Technology of Portugal
+(UIDB/00214/2020). NSG was supported by Academia 2019 award from the
+Catalan Institution for Research and Advanced Studies (ICREA). We are
+grateful to Chiara Santolin, Ege E. Özer, Alicia Franco-Martínez,
+Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and
+Perception research group for their helpful feedback. We thank Xavier
+Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical
+support, and Cristina Dominguez and Katia Pistrin for their efforts in
+recruiting infants. We also thank all families and infants who
+participated in the experiments.}}
 \author{
@@
 for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}\And
+08005\\\texttt{\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}}\And
 \textbf{Daniela S. Avila-Varela}~\orcidlink{0000-0002-3518-8117}\\Center
 for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}\And
+08005\\\texttt{\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}}\And
 \textbf{Ignacio
@@
 Psicología\\Universidad Autónoma de Madrid\\Madrid,
-28049\\\href{mailto:jignaciocastillejo@gmail.com}{jignaciocastillejo@gmail.com}\And
+28049\\\texttt{\href{mailto:jignaciocastillejo@gmail.com}{jignaciocastillejo@gmail.com}}\And
 \textbf{Nuria Sebastian-Galles}~\orcidlink{0000-0001-6938-2498}\\Center
 for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}
+08005\\\texttt{\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}}
 \date{}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, sharp corners, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 Consider the example of the Catalan-Spanish cognate translation
-equivalent /\textipa{"gat}-\textipa{"ga.to}/ {[}cat{]}, that are
+equivalent /\textipa{"gat}-\textipa{"ga.to}/ {[}\emph{cat}{]}, that are
 phonologically very similar. When the child is exposed to
@@
 both co-activated forms. The case of the non-cognate translation
-equivalent /\textipa{"gos-"pe.ro}/ would be different. Given the low
-phonological similarity between both word-forms, an exposure to
-\textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/
-leading to such exposure counting as a learning instance for
-/\textipa{"gos}/ (which the child was exposed to), but not for
-/\textipa{"pe.ro}/. While /\textipa{"gat}-\textipa{"ga.to}/ will benefit
-from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not. If
-the child receives linguistic input from one of the languages more often
-than from the other, this effect might affect each form of the cognate
-translation equivalent differently. For instance, if the child receives
-a larger amount of Catalan input than Spanish input, they will encounter
-the Catalan form /\textipa{"gat}/ more frequently than the Spanish form
-/\textipa{"ga.to}/. Through parallel activation, /\textipa{"gat}/ will
-activate /\textipa{"ga.to}/ more often than vice versa. Ultimately,
-/\textipa{"ga.to}/ will benefit more strongly from its cognate status
-than /\textipa{"gat}/, as it receives additional learning instances from
-its translation equivalent more often than /\textipa{"gat}/.
+equivalent /\textipa{"gos-"pe.ro}/ {[}\emph{dog}{]} would be different.
+Given the low phonological similarity between both word-forms, an
+exposure to \textipa{"gos} will result in a weak activation of
+/\textipa{"pe.ro}/ leading to such exposure counting as a learning
+instance for /\textipa{"gos}/ (which the child was exposed to), but not
+for /\textipa{"pe.ro}/. While /\textipa{"gat}-\textipa{"ga.to}/ will
+benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will
+not. If the child receives linguistic input from one of the languages
+more often than from the other, this effect might affect each form of
+the cognate translation equivalent differently. For instance, if the
+child receives a larger amount of Catalan input than Spanish input, they
+will encounter the Catalan form /\textipa{"gat}/ more frequently than
+the Spanish form /\textipa{"ga.to}/. Through parallel activation,
+/\textipa{"gat}/ will activate /\textipa{"ga.to}/ more often than vice
+versa. Ultimately, /\textipa{"ga.to}/ will benefit more strongly from
+its cognate status than /\textipa{"gat}/, as it receives additional
+learning instances from its translation equivalent more often than
+/\textipa{"gat}/.
 
@@
 Animals & 26 & 23 & 19 & 25 & panther, tiger\\
-Outside & 14 & 13 & 13 & 15 & farm , stone\\
-Body parts & 14 & 12 & 11 & 11 & face  , finger\\
+Outside & 14 & 13 & 13 & 15 & farm, stone\\
+Body parts & 14 & 12 & 11 & 11 & face, finger\\
 \addlinespace
-Toys & 11 & 11 & 12 & 13 & piano , racket\\
+Toys & 11 & 11 & 12 & 13 & piano, racket\\
 Clothes & 12 & 12 & 10 & 10 & zipper, sandal\\
 Vehicles & 9 & 10 & 11 & 10 & helicopter, tractor\\
-Colours & 6 & 6 & 6 & 6 & red  , green\\
-People & 7 & 4 & 6 & 6 & police    , babysitter\\
+Colours & 6 & 6 & 6 & 6 & red, green\\
+People & 7 & 4 & 6 & 6 & police, babysitter\\
 \addlinespace
 Furniture and rooms & 4 & 4 & 4 & 4 & corridor, terrace\\
-Time & 2 & 2 & 2 & 2 & day  , night\\
+Time & 2 & 2 & 2 & 2 & day, night\\
 Adventures & 1 & 1 & 1 & 1 & witch\\
@@
 \midrule
-N & 167 & 151 & 149 & 156 & \\
+N & 167 & 151 & 149 & 156 & -\\
 \bottomrule
@@
 
-We collected 138,078 responses to 1,590 items. We restricted the
-analyses to responses to nouns (628 items corresponding to other
-grammatical classes were excluded, see
-\protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020} for
-a similar approach). We then excluded items with missing lexical
-frequency scores (\emph{n} = 269), items that included more than one
-lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} = 48), multi-word
-items or phrases (e.g., \emph{barrita de cereales} {[}cereal bar{]},
-\emph{n} = 9). Finally, we removed items without a translation in the
-other language (\emph{n} = 32). This resulted in a final list of 604
-items, corresponding to 302 Catalan words and their 302 Spanish
-translations (302 translation equivalents). After collecting
+We collected data for 1,590 words. We restricted the analyses to
+responses to nouns (628 items corresponding to other grammatical classes
+were excluded, see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
+et al., 2020} for a similar approach). We then excluded items with
+missing lexical frequency scores (\emph{n} = 269), items that included
+more than one lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} =
+48), multi-word items or phrases (e.g., \emph{barrita de cereales}
+{[}cereal bar{]}, \emph{n} = 9). Finally, we removed items without a
+translation in the other language (\emph{n} = 32). This resulted in a
+final list of 604 items, corresponding to 302 Catalan words and their
+302 Spanish translations (302 translation equivalents). After collecting
 participants' responses, the final dataset consisted of 138,078
@@
 languages pooled together.
+
+Data processing and visualisation was done in R
+(\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
+version 4.2.2) using the Tidyverse family of packages
+(\protect\hyperlink{ref-wickham2019welcome}{Wickham et al., 2019}).
 
@@
 We specified a weakly informative prior around the parameters of the
-model. Equation~\ref{eq-linear} shows a detailed description of the
+model. Equation~\ref{eq-model} shows a detailed description of the
 model.
 
-\begin{equation}\protect\hypertarget{eq-linear}{}{
+\begin{equation}\protect\hypertarget{eq-model}{}{
 \begin{aligned}
@@
 \end{aligned}
-}\label{eq-linear}\end{equation}
+}\label{eq-model}\end{equation}
 
@@
 whose degree of exposure is 80\% for Catalan and 20\% for Spanish, the
-expected \emph{Exposure} score to the Catalan word \emph{cavall}
-{[}\emph{horse}{]}---with a lexical frequency of ---would be , while
-that of its translation to Spanish \emph{caballo} would be .
+expected \emph{Exposure} score to the Catalan word \emph{cotxe}
+{[}\emph{car}{]}---with a lexical frequency of 6.33---would be 5.06,
+while that of its translation to Spanish \emph{coche} would be 1.27.
 
@@
 
-Data processing and visualisation was done in R
-(\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
-version 4.2.2) using the Tidyverse family of packages
-(\protect\hyperlink{ref-wickham2019welcome}{Wickham et al., 2019}). We
-implemented the model using \texttt{brms}
+We implemented the model using \texttt{brms}
 (\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), a R interface
@@
 (\protect\hyperlink{ref-carpenter2017stan}{Carpenter et al., 2017}). We
-ran four iteration chains with 1,000 iterations each and an additional
-1,000 warm-up iterations per chain. Model posterior draws and
-predictions were handled using the \texttt{tidybayes}
+ran four iteration chains using the by-default No U-Turn Sampler
+algorithm with 1,000 iterations each and an additional 1,000 warm-up
+iterations per chain. Model posterior draws and predictions were handled
+using the \texttt{tidybayes}
 (\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) and
@@
 
-The model posterior showed adequate chain convergence diagnostics for
-all parameters (\(\hat{R} \le 1.06\), \(0.022 \le N_{eff}/N < 1.1001\))
-and little evidence of multicollinearity. Table~\ref{tbl-coefs} shows
-the summary of the posterior distribution of the fixed regression
-coefficients, and their degree of overlap with the ROPE. For
-interpretability, we report the estimated regression coefficients
-transformed to the probability scale\footnote{The logit and probability
-  scales relate non-linearly. This means that one logit difference is
-  not necessarily translated to a unique value in the probability scale.
-  For example, the probability of acquisition of a given word might
-  increase in 5\% when age increases from 22 to 23 months, the
-  probability of acquisition of the same word might only increase in
-  0.2\% when age increases from 31 to 32 months. The linear growth of
-  the probability of acquisition differs along the logistic curve, and
-  therefore deciding the age point at which to report the estimates of
-  the regression coefficients in the probability scale is not trivial.
-  Following Gelman et al.
+Chain convergence diagnostics showed adequate values, and the model
+posterior showed little evidence of correlation, as indicated by
+negligible pairwise correlation between the marginal posterior
+distribution of coefficients. Table~\ref{tbl-coefs} shows the summary of
+the posterior distribution of the fixed regression coefficients, and
+their degree of overlap with the ROPE. For interpretability, we report
+the estimated regression coefficients transformed to the probability
+scale\footnote{The logit and probability scales relate non-linearly.
+  This means that one logit difference is not necessarily translated to
+  a unique value in the probability scale. For example, the probability
+  of acquisition of a given word might increase in 5\% when age
+  increases from 22 to 23 months, the probability of acquisition of the
+  same word might only increase in 0.2\% when age increases from 31 to
+  32 months. The linear growth of the probability of acquisition differs
+  along the logistic curve, and therefore deciding the age point at
+  which to report the estimates of the regression coefficients in the
+  probability scale is not trivial. Following Gelman et al.
   (\protect\hyperlink{ref-gelman2020regression}{2020}), we report the
@@
 \begin{table}
-\caption{\label{tbl-coefs}Posterior distribution of regression coefficients. }\tabularnewline
+\caption{\label{tbl-coefs}Posterior distribution of regression coefficients. Median: median of the
+posterior distribution in the probability scale. 95\% HDI: 95\% highest
+density interval of the distribution. p(ROPE): overlap between the 95\%
+HDI and the ROPE, indicating the posterior probability that the true
+value of the coefficient is equivalent to zero. In the right column, we
+show a graphical representation of the median and 95\% HDI of each
+coefficient, with a reference dotted line indicating the location of
+zero. }\tabularnewline
 
@@
 \multicolumn{5}{l}{\textbf{Intercepts (at 22 months)}}\\
-\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & .088 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d082c12c1b.pdf}\\
-\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d08785c6c0e.pdf}\\
+\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & .088 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_2718113c70f2.pdf}\\
+\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_2718cf068d3.pdf}\\
 \addlinespace[0.3em]
 \multicolumn{5}{l}{\textbf{Slopes (upper bound)}}\\
-\hspace{1em}Age (+1 SD, 4.87, months) & 0.405 & {}[0.357, 0.451] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d087e6a2c55.pdf}\\
-\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d086eae2508.pdf}\\
-\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .037 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d084c7d3767.pdf}\\
-\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d0815d12d78.pdf}\\
-\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d082a9a1cf2.pdf}\\
-\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .985 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d083e54715a.pdf}\\
-\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d087de616c.pdf}\\
-\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .975 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d08238a53bc.pdf}\\
+\hspace{1em}Age (+1 SD, 4.87, months) & 0.405 & {}[0.357, 0.451] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27186be06e3b.pdf}\\
+\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27186d8d1719.pdf}\\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .037 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271852a346a4.pdf}\\
+\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27181f2174e.pdf}\\
+\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271847bb40f3.pdf}\\
+\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .985 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27184e7d1f57.pdf}\\
+\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271812b03ccf.pdf}\\
+\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .975 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27185801653e.pdf}\\
 \bottomrule
@@
 
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
 \hypertarget{refs}{}
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-arel-bundock2022marginaleffects}{}}%
+Arel-Bundock, V. (2022). \emph{Marginaleffects: Marginal effects,
+marginal means, predictions, and contrasts}.
+\url{https://CRAN.R-project.org/package=marginaleffects}
+
+\leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%
+Arslan, R. C., Walther, M. P., \& Tata, C. S. (2020). Formr: A study
+framework allowing for automated feedback generation and complex
+longitudinal experience-sampling studies using r. \emph{Behavior
+Research Methods}, \emph{52}(1), 376--387.
+\url{https://doi.org/10.3758/s13428-019-01236-y}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bailey2002phonological}{}}%
+Bailey, T. M., \& Plunkett, K. (2002). Phonological specificity in early
+words. \emph{Cognitive Development}, \emph{17}(2), 1265--1282.
+\url{https://doi.org/10.1016/S0885-2014(02)00116-8}
+
+\leavevmode\vadjust pre{\hypertarget{ref-barr2013random}{}}%
+Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random
+effects structure for confirmatory hypothesis testing: Keep it maximal.
+\emph{Journal of Memory and Language}, \emph{68}(3), 255--278.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bergelson2012months}{}}%
+Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants
+know the meanings of many common nouns. \emph{Proceedings of the
+National Academy of Sciences}, \emph{109}(9), 3253--3258.
+\url{https://doi.org/10.1073/pnas.1113380109}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bergelson2015early}{}}%
+Bergelson, E., \& Swingley, D. (2015). Early word comprehension in
+infants: Replication and extension. \emph{Language Learning and
+Development}, \emph{11}(4), 369--380.
+\url{https://doi.org/10.1080/15475441.2014.979387}
+
+\leavevmode\vadjust pre{\hypertarget{ref-blom2020crosslanguage}{}}%
+Blom, E., Boerma, T., Bosma, E., Cornips, L., Heuij, K. van den, \&
+Timmermeister, M. (2020). Cross-language distance influences receptive
+vocabulary outcomes of bilingual children. \emph{First Language},
+\emph{40}(2), 151--171. \url{https://doi.org/10.1177/0142723719892794}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosch2014first}{}}%
+Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in
+bilingual toddlers' expressive vocabulary: Does form similarity matter?
+\emph{International Journal of Behavioral Development}, \emph{38}(4),
+317--322. \url{https://doi.org/10.1177/0165025414532559}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosma2019longitudinal}{}}%
+Bosma, E., Blom, E., Hoekstra, E., \& Versloot, A. (2019). A
+longitudinal study on the gradual cognate facilitation effect in
+bilingual children's frisian receptive vocabulary. \emph{International
+Journal of Bilingual Education and Bilingualism}, \emph{22}(4),
+371--385. \url{https://doi.org/10.1080/13670050.2016.1254152}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosma2020cognate}{}}%
+Bosma, E., \& Nota, N. (2020). Cognate facilitation in frisian--dutch
+bilingual children's sentence reading: An eye-tracking study.
+\emph{Journal of Experimental Child Psychology}, \emph{189}, 104699.
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
+Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
+using stan. \emph{Journal of Statistical Software}, \emph{80}, 1--28.
+\url{https://doi.org/10.18637/jss.v080.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
+Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B.,
+Betancourt, M., Brubaker, M., Guo, J., Li, P., \& Riddell, A. (2017).
+Stan : A probabilistic programming language. \emph{Journal of
+Statistical Software}, \emph{76}(1).
+\url{https://doi.org/10.18637/jss.v076.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
+Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate
+facilitation effect: Implications for models of lexical access.
+\emph{Journal of Experimental Psychology: Learning, Memory, and
+Cognition}, \emph{26}, 1283--1296.
+\url{https://doi.org/10.1037/0278-7393.26.5.1283}
+
+\leavevmode\vadjust pre{\hypertarget{ref-2018els}{}}%
+\emph{Els usos lingüístics de la població de catalunya}. (2018).
+Generalitat de Catalunya.
+\url{https://llengua.gencat.cat/web/.content/documents/dadesestudis/altres/arxius/dossier-eulp-2018.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-feldman2005concurrent}{}}%
+Feldman, H. M., Dale, P. S., Campbell, T. F., Colborn, D. K.,
+Kurs-Lasky, M., Rockette, H. E., \& Paradise, J. L. (2005). Concurrent
+and predictive validity of parent reports of child language at ages 2
+and 3 years. \emph{Child Development}, \emph{76}(4), 856--868.
+\url{https://doi.org/10.1111/j.1467-8624.2005.00882.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fenson2007macarthurbates}{}}%
+Fenson, L. et al. (2007). \emph{{MacArthur}-bates communicative
+development inventories}. Paul H. Brookes Publishing Company Baltimore,
+{MD}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
+Floccia, C., Sambrook, T. D., Delle Luche, C., Kwok, R., Goslin, J.,
+White, L., Cattani, A., Sullivan, E., Abbot‐Smith, K., Krott, A., Mills,
+D., Rowland, C., Gervain, J., \& Plunkett, K. (2018). I: introduction.
+\emph{Monographs of the Society for Research in Child Development},
+\emph{83}(1), 7--29. \url{https://doi.org/10.1111/mono.12348}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fourtassi2020growth}{}}%
+Fourtassi, A., Bian, Y., \& Frank, M. C. (2020). The growth of
+children's semantic and phonological networks: Insight from 10
+languages. \emph{Cognitive Science}, \emph{44}(7), e12847.
+\url{https://doi.org/10.1111/cogs.12847}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gampe2021does}{}}%
+Gampe, A., Quick, A. E., \& Daum, M. M. (2021). Does linguistic
+similarity affect early simultaneous bilingual language acquisition?
+\emph{Journal of Language Contact}, \emph{13}(3), 482--500.
+
+\leavevmode\vadjust pre{\hypertarget{ref-garcia-castro2023bvq}{}}%
+Garcia-Castro, G., Ávila-Varela, D. S., \& Sebastian-Galles, N. (2023).
+\emph{Bvq: Barcelona vocabulary questionnaire database and helper
+functions}. \url{https://gongcastro.github.io/bvq}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
+Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other
+stories}. Cambridge University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-giguere2022bilingual}{}}%
+Giguere, D., \& Hoff, E. (2022). Bilingual development in the receptive
+and expressive domains: They differ. \emph{International Journal of
+Bilingual Education and Bilingualism}, \emph{25}(10), 3849--3858.
+\url{https://doi.org/10.1080/13670050.2022.2087039}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gillen2021tapping}{}}%
+Gillen, N. A., Siow, S., Lepadatu, I., Sucevic, J., Plunkett, K., \&
+Duta, M. (2021). \emph{Tapping into the potential of remote
+developmental research: Introducing the {OxfordBabylab} app}.
+{PsyArXiv}. \url{https://doi.org/10.31234/osf.io/kxhmw}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gimeno-martinez2021crosslinguistic}{}}%
+Gimeno-Martínez, M., Mädebach, A., \& Baus, C. (2021). Cross-linguistic
+interactions across modalities: Effects of the oral language on sign
+production. \emph{Bilingualism: Language and Cognition}, \emph{24}(4),
+779--790. \url{https://doi.org/10.1017/S1366728921000171}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gonzalez-barrero2020bilingual}{}}%
+Gonzalez-Barrero, A. M., Schott, E., \& Byers-Heinlein, K. (2020).
+\emph{Bilingual adjusted vocabulary: A developmentally-informed
+bilingual vocabulary measure}. {PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/x7s4u}
+
+\leavevmode\vadjust pre{\hypertarget{ref-grosjean2021extent}{}}%
+Grosjean, F. (2021). The extent of bilingualism. \emph{Life as a
+Bilingual}, 27--39.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hamilton2000infant}{}}%
+Hamilton, A., Plunkett, K., \& Schafer, G. (2000). Infant vocabulary
+development assessed with a british communicative development inventory.
+\emph{Journal of Child Language}, \emph{27}(3), 689--705.
+
+\leavevmode\vadjust pre{\hypertarget{ref-havy2016phonetic}{}}%
+Havy, M., Bouchon, C., \& Nazzi, T. (2016). Phonetic processing when
+learning words: The case of bilingual infants. \emph{International
+Journal of Behavioral Development}, \emph{40}(1), 41--52.
+\url{https://doi.org/10.1177/0165025415570646}
+
+\leavevmode\vadjust pre{\hypertarget{ref-heeringa2003norwegian}{}}%
+Heeringa, W., \& Gooskens, C. (2003). Norwegian dialects examined
+perceptually and acoustically. \emph{Computers and the Humanities},
+\emph{37}(3), 293--315. \url{https://doi.org/10.1023/A:1025087115665}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanhell2008sentence}{}}%
+Hell, J. G. van, \& Groot, A. M. B. de. (2008). Sentence context
+modulates visual word recognition and translation in bilinguals.
+\emph{Acta Psychologica}, \emph{128}(3), 431--451.
+\url{https://doi.org/10.1016/j.actpsy.2008.03.010}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hidaka2013computational}{}}%
+Hidaka, S. (2013). A computational model associating learning process,
+word attributes, and age of acquisition. \emph{{PLOS} {ONE}},
+\emph{8}(11), e76242. \url{https://doi.org/10.1371/journal.pone.0076242}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hoff2012dual}{}}%
+Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M.
+(2012). Dual language exposure and early bilingual development*.
+\emph{Journal of Child Language}, \emph{39}(1), 1--27.
+\url{https://doi.org/10.1017/S0305000910000759}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hoshino2008cognate}{}}%
+Hoshino, N., \& Kroll, J. F. (2008). Cognate effects in picture naming:
+Does cross-language activation survive a change of script?
+\emph{Cognition}, \emph{106}(1), 501--511.
+\url{https://doi.org/10.1016/j.cognition.2007.02.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-houston-price2007discrepancy}{}}%
+Houston-Price, C., Mather, E., \& Sakkalou, E. (2007). Discrepancy
+between parental reports of infants' receptive vocabulary and infants'
+behaviour in a preferential looking task. \emph{Journal of Child
+Language}, \emph{34}(4), 701--724.
+\url{https://doi.org/10.1017/S0305000907008124}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hustad2021speech}{}}%
+Hustad, K. C., Mahr, T. J., Natzke, P., \& Rathouz, P. J. (2021). Speech
+development between 30 and 119 months in typical children i:
+Intelligibility growth curves for single-word and multiword productions.
+\emph{Journal of Speech, Language, and Hearing Research}, \emph{64}(10),
+3707--3719. \url{https://doi.org/10.1044/2021_JSLHR-21-00142}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hustad2020development}{}}%
+Hustad, K. C., Mahr, T., Natzke, P. E. M., \& Rathouz, P. J. (2020).
+Development of speech intelligibility between 30 and 47 months in
+typically developing children: A cross-sectional study of growth.
+\emph{Journal of Speech, Language, and Hearing Research}, \emph{63}(6),
+1675--1687. \url{https://doi.org/10.1044/2020_JSLHR-20-00008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kachergis2022standard}{}}%
+Kachergis, G., Marchman, V. A., \& Frank, M. C. (2022). Toward a
+{``standard model''} of early language learning. \emph{Current
+Directions in Psychological Science}, \emph{31}(1), 20--27.
+\url{https://doi.org/10.1177/09637214211057836}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kay2021tidybayes}{}}%
+Kay, M. (2021). \emph{Tidybayes: Tidy data and geoms for bayesian
+models}. \url{http://mjskay.github.io/tidybayes/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-killing2008move}{}}%
+Killing, S. E. A., \& Bishop, D. V. M. (2008). Move it! Visual feedback
+enhances validity of preferential looking as a measure of individual
+differences in vocabulary in toddlers. \emph{Developmental Science},
+\emph{11}(4), 525--530.
+\url{https://doi.org/10.1111/j.1467-7687.2008.00698.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroll2017bilingual}{}}%
+Kroll, J. F., \& Ma, F. (2017). The bilingual lexicon. \emph{The
+Handbook of Psycholinguistics}, 294--319.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kruschke2018bayesian}{}}%
+Kruschke, J. K., \& Liddell, T. M. (2018). The bayesian new statistics:
+Hypothesis testing, estimation, meta-analysis, and planning from a
+bayesian perspective. \emph{Psychonomic Bulletin \&Review}, \emph{25},
+178--206. \url{https://doi.org/10.3758/s13423-016-1221-4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-laing2022phonological}{}}%
+Laing, C. E. (2022). \emph{Phonological networks and systematicity in
+early lexical acquisition}. {PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/z8pyg}
+
+\leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
+Levenshtein, V. I. (1966). Binary codes capable of correcting deletions,
+insertions, and reversals. \emph{Soviet Physics-Doklady}, \emph{10},
+707--710.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanderloo2014stringdist}{}}%
+Loo, M. P. J. van der. (2014). The stringdist package for approximate
+string matching. \emph{The R Journal}, \emph{6}(1), 111--122.
+\url{https://doi.org/10.32614/RJ-2014-011}
+
+\leavevmode\vadjust pre{\hypertarget{ref-macwhinney2000childes}{}}%
+MacWhinney, B. (2000). \emph{The {CHILDES} project: The database} (Vol.
+2). Psychology Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mani2011does}{}}%
+Mani, N., \& Plunkett, K. (2011). Does size matter? Subsegmental cues to
+vowel mispronunciation detection*. \emph{Journal of Child Language},
+\emph{38}(3), 606--627. \url{https://doi.org/10.1017/S0305000910000243}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mattock2010first}{}}%
+Mattock, K., Polka, L., Rvachew, S., \& Krehm, M. (2010). The first
+steps in word learning are easier when the shoes fit: Comparing
+monolingual and bilingual infants. \emph{Developmental Science},
+\emph{13}(1), 229--243.
+\url{https://doi.org/10.1111/j.1467-7687.2009.00891.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcmurray2007defusing}{}}%
+McMurray, B. (2007). Defusing the childhood vocabulary explosion.
+\emph{Science}, \emph{317}(5838), 631--631.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mitchell2022cognates}{}}%
+Mitchell, L., Tsui, R. K. Y., \& Byers-Heinlein, K. (2022).
+\emph{Cognates are advantaged in early bilingual expressive vocabulary
+development}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/daktp}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mollica2017how}{}}%
+Mollica, F., \& Piantadosi, S. T. (2017). How data drive early word
+learning: A cross-linguistic waiting time analysis. \emph{Open Mind},
+\emph{1}(2), 67--77. \url{https://doi.org/10.1162/OPMI_a_00006}
+
+\leavevmode\vadjust pre{\hypertarget{ref-morford2011when}{}}%
+Morford, J. P., Wilkinson, E., Villwock, A., Piñar, P., \& Kroll, J. F.
+(2011). When deaf signers read english: Do written words activate their
+sign translations? \emph{Cognition}, \emph{118}(2), 286--292.
+\url{https://doi.org/10.1016/j.cognition.2010.11.006}
+
+\leavevmode\vadjust pre{\hypertarget{ref-oller2002language}{}}%
+Oller, D. K., \& Eilers, R. E. (2002). \emph{Language and literacy in
+bilingual children}. Multilingual Matters.
+
+\leavevmode\vadjust pre{\hypertarget{ref-patterson2004comparing}{}}%
+Patterson, J. L. (2004). Comparing bilingual and monolingual toddlers'
+expressive vocabulary size. \emph{Journal of Speech, Language, and
+Hearing Research}, \emph{47}(5), 1213--1215.
+\url{https://doi.org/10.1044/1092-4388(2004/089)}
+
+\leavevmode\vadjust pre{\hypertarget{ref-patterson2004bilingual}{}}%
+Patterson, J. L., \& Pearson, B. Z. (2004). Bilingual lexical
+development: Influences, contexts, and processes. In \emph{Bilingual
+language development and disorders in spanish-english speakers} (pp.
+77--104). Paul H. Brookes Publishing Co.
+
+\leavevmode\vadjust pre{\hypertarget{ref-pearson1994patterns}{}}%
+Pearson, B. Z., \& Fernández, S. C. (1994). Patterns of interaction in
+the lexical growth in two languages of bilingual infants and toddlers.
+\emph{Language Learning}, \emph{44}(4), 617--653.
+\url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-pearson1993lexical}{}}%
+Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical
+development in bilingual infants and toddlers: Comparison to monolingual
+norms. \emph{Language Learning}, \emph{43}(1), 93--120.
+\url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-petitto2001bilingual}{}}%
+Petitto, L. A., Katerelos, M., Levy, B. G., Gauna, K., Tétreault, K., \&
+Ferraro, V. (2001). Bilingual signed and spoken language acquisition
+from birth: Implications for the mechanisms underlying early bilingual
+language acquisition. \emph{Journal of Child Language}, \emph{28}(2),
+453--496. \url{https://doi.org/10.1017/S0305000901004718}
+
+\leavevmode\vadjust pre{\hypertarget{ref-poarch2012crosslanguage}{}}%
+Poarch, G. J., \& Hell, J. G. van. (2012). Cross-language activation in
+children's speech production: Evidence from second language learners,
+bilinguals, and trilinguals. \emph{Journal of Experimental Child
+Psychology}, \emph{111}(3), 419--438.
+\url{https://doi.org/10.1016/j.jecp.2011.09.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-poulin-dubois2013lexical}{}}%
+Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J.
+(2013). Lexical access and vocabulary development in very young
+bilinguals. \emph{International Journal of Bilingualism}, \emph{17}(1),
+57--70.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rcoreteam2013language}{}}%
+R Core Team. (2013). \emph{R: A language and environment for statistical
+computing}. R Foundation for Statistical Computing.
+\url{http://www.R-project.org/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2010are}{}}%
+Ramon-Casas, M., \& Bosch, L. (2010). Are non-cognate words
+phonologically better specified than cognates in the early lexicon of
+bilingual children. \emph{Selected Proceedings of the 4th Conference on
+Laboratory Approaches to Spanish Phonology}, 31--36.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2017minimalpair}{}}%
+Ramon-Casas, M., Fennell, C. T., \& Bosch, L. (2017). Minimal-pair word
+learning by bilingual toddlers: The catalan /e/-/ɛ/ contrast revisited.
+\emph{Bilingualism: Language and Cognition}, \emph{20}(3), 649--656.
+\url{https://doi.org/10.1017/S1366728916001115}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ramon-casas2009vowel}{}}%
+Ramon-Casas, M., Swingley, D., Sebastián-Gallés, N., \& Bosch, L.
+(2009). Vowel categorization during word recognition in bilingual
+toddlers. \emph{Cognitive Psychology}, \emph{59}(1), 96--121.
+\url{https://doi.org/10.1016/j.cogpsych.2009.02.002}
+
+\leavevmode\vadjust pre{\hypertarget{ref-samuelson2021precision}{}}%
+Samuelson, L. K. (2021). Toward a precision science of word learning:
+Understanding individual vocabulary pathways. \emph{Child Development
+Perspectives}, \emph{15}(2), 117--124.
+\url{https://doi.org/10.1111/cdep.12408}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sanchez2019childesdb}{}}%
+Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
+D., \& Frank, M. C. (2019). Childes-db: A flexible and reproducible
+interface to the child language data exchange system. \emph{Behavior
+Research Methods}, \emph{51}(4), 1928--1941.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schelletter2002effect}{}}%
+Schelletter, C. (2002). The effect of form similarity on bilingual
+children's lexical development. \emph{Bilingualism: Language and
+Cognition}, \emph{5}(2), 93--107.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schepens2012distributions}{}}%
+Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of
+cognates in europe as based on levenshtein distance. \emph{Bilingualism:
+Language and Cognition}, \emph{15}(1), 157--166.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schroter2016orthographic}{}}%
+Schröter, P., \& Schroeder, S. (2016). Orthographic processing in
+balanced bilingual children: Cross-language evidence from cognates and
+false friends. \emph{Journal of Experimental Child Psychology},
+\emph{141}, 239--246. \url{https://doi.org/10.1016/j.jecp.2015.09.005}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schwartz2007reading}{}}%
+Schwartz, A. I., Kroll, J. F., \& Diaz, M. (2007). Reading words in
+spanish and english: Mapping orthography to phonology in two languages.
+\emph{Language and Cognitive Processes}, \emph{22}(1), 106--129.
+\url{https://doi.org/10.1080/01690960500463920}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sebastian-galles2020bilingual}{}}%
+Sebastian-Galles, N., \& Santolin, C. (2020). Bilingual acquisition: The
+early steps. \emph{Annual Review of Developmental Psychology},
+\emph{2}(1), 47--68.
+\url{https://doi.org/10.1146/annurev-devpsych-013119-023724}
+
+\leavevmode\vadjust pre{\hypertarget{ref-smithson2014bilingualism}{}}%
+Smithson, L., Paradis, J., \& Nicoladis, E. (2014). Bilingualism and
+receptive vocabulary achievement: Could sociocultural context make a
+difference? \emph{Bilingualism: Language and Cognition}, \emph{17}(4),
+810--821. \url{https://doi.org/10.1017/S1366728913000813}
+
+\leavevmode\vadjust pre{\hypertarget{ref-spivey1999cross}{}}%
+Spivey, M. J., \& Marian, V. (1999). Cross talk between native and
+second languages: Partial activation of an irrelevant lexicon.
+\emph{Psychological Science}, \emph{10}(3), 281--284.
+
+\leavevmode\vadjust pre{\hypertarget{ref-swingley200511montholds}{}}%
+Swingley, D. (2005). 11-month-olds' knowledge of how familiar words
+sound. \emph{Developmental Science}, \emph{8}(5), 432--443.
+\url{https://doi.org/10.1111/j.1467-7687.2005.00432.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-swingley2000spoken}{}}%
+Swingley, D., \& Aslin, R. N. (2000). Spoken word recognition and
+lexical representation in very young children. \emph{Cognition},
+\emph{76}(2), 147--166.
+\url{https://doi.org/10.1016/S0010-0277(00)00081-0}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tamasi2017pupillometry}{}}%
+Tamási, K., McKean, C., Gafos, A., Fritzsche, T., \& Höhle, B. (2017).
+Pupillometry registers toddlers' sensitivity to degrees of
+mispronunciation. \emph{Journal of Experimental Child Psychology},
+\emph{153}, 140--148. \url{https://doi.org/10.1016/j.jecp.2016.07.014}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tincoff1999beginnings}{}}%
+Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word
+comprehension in 6-month-olds. \emph{Psychological Science},
+\emph{10}(2), 172--175. \url{https://doi.org/10.1111/1467-9280.00127}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanheuven2014subtlexuk}{}}%
+Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
+{SUBTLEX}-{UK}: A new and improved word frequency database for british
+english. \emph{Quarterly Journal of Experimental Psychology},
+\emph{67}(6), 1176--1190.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonholzen2019impact}{}}%
+Von Holzen, K., Fennell, C. T., \& Mani, N. (2019). The impact of
+cross-language phonological overlap on bilingual and monolingual
+toddlers' word recognition. \emph{Bilingualism: Language and Cognition},
+\emph{22}(3), 476--499. \url{https://doi.org/10.1017/S1366728918000597}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vonholzen2012language}{}}%
+Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access
+in bilingual toddlers. \emph{Journal of Experimental Child Psychology},
+\emph{113}(4), 569--586.
+\url{https://doi.org/10.1016/j.jecp.2012.08.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wells1995computercoding}{}}%
+Wells, J. C. (1995). \emph{Computer-coding the {IPA}: A proposed
+extension of {SAMPA}}. \emph{4}(28), 1995.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wewalaarachchi2017vowels}{}}%
+Wewalaarachchi, T. D., Wong, L. H., \& Singh, L. (2017). Vowels,
+consonants, and lexical tones: Sensitivity to phonological variation in
+monolingual mandarin and bilingual english--mandarin toddlers.
+\emph{Journal of Experimental Child Psychology}, \emph{159}, 16--33.
+\url{https://doi.org/10.1016/j.jecp.2017.01.009}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham2019welcome}{}}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
+Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
+Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
+Welcome to the tidyverse. \emph{Journal of Open Source Software},
+\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
+
+\leavevmode\vadjust pre{\hypertarget{ref-zipf1949human}{}}%
+Zipf, G. K. (1949). \emph{Human behavior and the principle of least
+effort}. Addison-Wesley Press.
+
 \end{CSLReferences}

### 2023-07-06 _targets_packages.R

@@ -13,7 +13,8 @@ library(future)
 library(ggplot2)
 library(graphics)
 library(grDevices)
-library(gt)
+library(kableExtra)
+library(knitr)
 library(lubridate)
 library(markdown)
 library(methods)

### 2023-07-23 manuscript/_appendix/_syllable-frequency.qmd
Simplify dependencies

@@
            syll = lapply(syll, function(x) x[x != ""]),
-           n_syll = map_int(syll, length),
-           item = str_remove(item, "cat_|spa_"))
+           n_syll = purrr::map_int(syll, length),
+           item = gsub("cat_|spa_", "", item))
 
@@
            syll = lapply(syll, function(x) x[x != ""]),
-           n_syll = map_int(syll, length),
-           item = str_remove(item, "cat_|spa_"))
+           n_syll = purrr::map_int(syll, length),
+           item = gsub("cat_|spa_", "", item)) 
 ```
@@
     unnest_longer(syll) |> 
-    group_by(language, syll) |> 
     summarise(freq_syll = sum(freq_lex),
-              .groups = "drop") |> 
+              .by = c(language, syll)) |> 
     arrange(language, desc(freq_syll))
@@
 # tidy predictor names
-str_repl <- c(
-    "b_Intercept" = "Intercept",
-    "b_n_syll_std" = glue::glue("Syllables (+1 SD, {round(sd(syllables_data$n_syll), 3)})"),
-    "b_lv_std" = glue::glue("Cognateness (+1 SD, {round(sd(syllables_data$lv), 2)})")
-)
 
-post_syllables <- model_fit_syllables |>
-    bayestestR::describe_posterior(test = "rope",
-                                   rope_range = rope_interval,
-                                   ci_method = "HDI") |>
-    as_tibble() |>
-    janitor::clean_names() |>
-    mutate(.variable_name = factor(
-        parameter,
-        levels = names(str_repl),
-        labels = str_repl) |>
-            as.character(),
-        parameter = ifelse(grepl("Intercept", parameter),
-                           stringr::str_remove_all(parameter, "Intercept \\(|\\)"),
-                           parameter)) |>
-    select(.variable = parameter,
-           .variable_name,
-           .median = median,
-           .upper = ci_high,
-           .lower = ci_low,
-           .rope_overlap = rope_percentage) |>
-    ungroup()
-
-post_draws_list <- post_syllables |> 
-    split(post_syllables$.variable)
+post_draws_list <- posterior_syllables_summary |> 
+    split(posterior_syllables_summary$.variable)
 
@@
 @tbl-syll-coefs shows a summary of the posterior distribution of the fixed effects in the model.
-As expected, words with more syllables scored higher in syllabic frequency: all posterior draws for the regression coefficient of the main effect of this predictor fell outside the ROPE defined between -0.5 and +0.5 ($\beta$ = `r round(post_draws_list[["b_n_syll_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_syll_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_syll_std"]]$.upper, 2)`]).
-Keeping the number of syllables constant, the effect of cognateness was negligible: all of the posterior distributions of this predictor fell within the ROPE, providing evidence that the true value of the increment in syllabic frequency for every increase in cognateness is equivalent to zero ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).
+As expected, words with more syllables scored higher in syllabic frequency: all posterior draws for the regression coefficient of the main effect of this predictor fell outside the ROPE defined between -0.5 and +0.5 ($\beta$ = `r round(post_draws_list[["b_n_syll_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_syll_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_syll_std"]]$.upper, 2)`]).
+Keeping the number of syllables constant, the effect of cognateness was negligible: all of the posterior distributions of this predictor fell within the ROPE, providing evidence that the true value of the increment in syllabic frequency for every increase in cognateness is equivalent to zero ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).
 
@@
 # summarise posterior draws
-tbl_data <- post_syllables |>
-    select(.variable_name, .median, .lower, .upper, .rope_overlap) |>
+tbl_data <- posterior_syllables_summary |>
+    select(.variable_name, .value, .lower, .upper, .rope) |>
     mutate(
@@
                       ", ", round(.upper, 3), "]"),
-        .rope_overlap = sub("^(-?)0.", "\\1.",
-                            sprintf("%.3f", .rope_overlap))) |> 
-    select(.variable_name, .median, .lower, .upper, .hdi, .rope_overlap)
+        .rope = sub("^(-?)0.", "\\1.",
+                    sprintf("%.3f", .rope))) |> 
+    select(.variable_name, .value, .lower, .upper, .hdi, .rope)
 
@@
                       "95\\% HDI",
-                      "$p(\\text{ROPE})$",
-                      ""),
+                      "$p(\\text{ROPE})$"),
         digits = 3,
         booktabs = TRUE,
-        align = "lclrc",
+        align = "lclr",
         linesep = ""
     ) |> 
-    kableExtra::kable_styling() |> 
-    kableExtra::column_spec(
-        5,
-        background = "#f2f2f2",
-        image = kableExtra::spec_pointrange(
-            x = tbl_data$.median, 
-            xmin = tbl_data$.lower, 
-            xmax = tbl_data$.upper, 
-            vline = 0,
-            xaxt = "n",
-            col = "#000000",
-            res = 200)
-    )
+    kableExtra::kable_styling()
 ```
@@
                                scale = sd(syllables_data$n_syll))[, 1]) |> 
-    add_epred_draws(model_fit_syllables,
+    tidybayes::add_epred_draws(model_fit_syllables,
                                re_formula = NA) |>
-    median_hdci() |> 
+    tidybayes::median_hdci() |> 
     mutate(

### 2023-07-23 manuscript/appendix.tex
Simplify dependencies

@@
 \begin{document}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, breakable, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 \multicolumn{5}{l}{\textbf{Model: Exposure}}\\
-\hspace{1em}Age (+1 SD, 4.87, months) & 0.600 & {}[0.588, 0.611] & .000 & \\
-\hspace{1em}Exposure (+1 SD, 1.81) & 0.558 & {}[0.55, 0.567] & .000 & \\
-\hspace{1em}Cognateness (+1 SD, 0.26) & 0.515 & {}[0.504, 0.526] & .037 & \\
+\hspace{1em}Age (+1 SD, 4.86 months) & 0.600 & {}[0.588, 0.611] & .000 & \\
+\hspace{1em}Age $\times$ Exposure & 0.518 & {}[0.51, 0.526] & .000 & \\
+\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .907 & \\
+\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.5, 0.506] & .974 & \\
+\hspace{1em}Exposure (+1 SD, 1.81) & 0.558 & {}[0.55, 0.566] & .000 & \\
+\hspace{1em}Exposure $\times$ Cognateness & 0.486 & {}[0.483, 0.489] & .000 & \\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.515 & {}[0.504, 0.527] & .089 & \\
 \hspace{1em}Length (+1 SD, 1.56 phonemes) & 0.485 & {}[0.478, 0.491] & .000 & \\
-\hspace{1em}Age $\times$ Exposure & 0.518 & {}[0.51, 0.526] & .000 & \\
-\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.5, 0.506] & .985 & \\
-\hspace{1em}Exposure $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
-\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .975 & \\
 \addlinespace[0.3em]
 \multicolumn{5}{l}{\textbf{Model: Frequency \& DoE}}\\
-\hspace{1em}Age (+1 SD, 4.87, months) & 0.601 & {}[0.59, 0.612] & .000 & \\
+\hspace{1em}Age (+1 SD, 4.86, months) & 0.601 & {}[0.59, 0.612] & .000 & \\
+\hspace{1em}Age $\times$ DoE & 0.518 & {}[0.51, 0.527] & .000 & \\
+\hspace{1em}Age $\times$ DoE $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .901 & \\
+\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.501, 0.507] & .898 & \\
 \hspace{1em}DoE (+1 SD, 0.3) & 0.558 & {}[0.549, 0.566] & .000 & \\
-\hspace{1em}Cognateness (+1 SD, 0.26) & 0.516 & {}[0.506, 0.528] & .016 & \\
+\hspace{1em}DoE $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
+\hspace{1em}Frequency (+1 SD, 0.19) & 0.527 & {}[0.516, 0.54] & .000 & \\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.516 & {}[0.505, 0.527] & .043 & \\
 \hspace{1em}Phonemes (+1 SD, 1.56 phonemes) & 0.486 & {}[0.479, 0.492] & .000 & \\
-\hspace{1em}Frequency (+1 SD, 0.19) & 0.527 & {}[0.516, 0.54] & .000 & \\
-\hspace{1em}Age $\times$ DoE & 0.518 & {}[0.51, 0.527] & .000 & \\
-\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.501, 0.507] & .954 & \\
-\hspace{1em}DoE $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
-\hspace{1em}Age $\times$ DoE $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .962 & \\
 \bottomrule
@@
 \midrule
-ladybug & marieta & m@.4i"E.t@ & 4 & mariquita & ma.4i"ki.ta & 4\\
-beans & mongetes & muJ"ZE.t@s & 3 & alubias & a"lu.bjas & 3\\
+drink (beverage) & beguda & b@"Gu.D@ & 3 & bebida & be"Bi.Da & 3\\
+snail & cargol & k@r"gOl & 2 & caracol & ka.4a"kol & 3\\
+red & vermell & b@r"mEL & 2 & rojo & "ro.xo & 2\\
+book & conte & "kon.t@ & 2 & cuento & "kwen.to & 2\\
+hand & mà / mans & m@ & 1 & manos & "ma.nos & 2\\
+\addlinespace
+egg & ou & Ow & 1 & huevo & "we.Bo & 2\\
+pool & piscina & pi"si.n@ & 3 & piscina & pis"Ti.na & 3\\
+fox & guineu & gi"nEw & 2 & zorro & "To.ro & 2\\
+potato & patata & p@"ta.t@ & 3 & patata & pa"ta.ta & 3\\
+arm & braç / braços & b4as & 1 & brazos & "b4a.Tos & 2\\
+\addlinespace
+cow & vaca & "ba.k@ & 2 & vaca & "ba.ka & 2\\
+sausage & salsitxa & s@5"si.t͡S@ & 3 & salchicha & sal"tSi.tSa & 3\\
 tongue & llengua & "LeN.gw@ & 2 & lengua & "len.gwa & 2\\
-kitchen & cuina & "kuj.n@ & 2 & cocina & ko"Ti.na & 3\\
-cheek & galta & "ga5.t@ & 2 & mejilla & me"xi.La & 3\\
-\addlinespace
-purse & bossa & "bo.s@ & 2 & bolso & "bol.so & 2\\
-farm & granja & g4aJ.Z@ & 2 & granja & "g4an.xa & 2\\
-sausage & salsitxa & s@5"si.t͡S@ & 3 & salchicha & sal"tSi.tSa & 3\\
-cereal & cereals & s@.4e"a5s & 3 & cereales & Te.4e"a.les & 4\\
-bottle & ampolla & @m"po.L@ & 3 & botella & bo"te.La & 3\\
-\addlinespace
-boat & vaixell & b@"SeL & 2 & barco & "ba4.ko & 2\\
-family & família & fa"mi.lia & 3 & familia & fa"mi.lia & 3\\
-balloon & globus & "g5O.Bus & 2 & globo & "glo.Bo & 2\\
-bee & abella & @"BE.L@ & 3 & abeja & a"Be.xa & 3\\
-earring & arracada / arracades & @.r@"ka.D@ & 4 & pendientes & pen"Djen.tes & 3\\
+park & parc & park & 1 & corralito & ko.ra"li.to & 4\\
+black & negre & "nE.G4@ & 2 & negro & "ne.G4o & 2\\
 \bottomrule
@@
 outside the ROPE defined between -0.5 and +0.5 (\(\beta\) = 5.64, 95\%
-HDI = {[}5.58, 5.71{]}). Keeping the number of syllables constant, the
+HDI = {[}5.57, 5.71{]}). Keeping the number of syllables constant, the
 effect of cognateness was negligible: all of the posterior distributions
@@
 \midrule
-Intercept & 16.089 & {}[16.023, 16.163] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e8078876ee9.pdf}\\
-Syllables (+1 SD, 0.802) & 5.642 & {}[5.577, 5.711] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e807c1b3ef.pdf}\\
-Cognateness (+1 SD, 0.24) & 0.007 & {}[-0.059, 0.075] & 1.000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e8043796412.pdf}\\
+Intercept & 16.089 & {}[16.022, 16.166] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/gonza/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_b043454131.pdf}\\
+Cognateness (+1 SD, 0.24) & 0.008 & {}[-0.057, 0.074] & 1.000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/gonza/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_b0479223d59.pdf}\\
+Syllables (+1 SD, 0.802) & 5.642 & {}[5.574, 5.709] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/gonza/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_b042ac2624.pdf}\\
 \bottomrule

### 2023-07-23 manuscript/manuscript.tex
Simplify dependencies

@@
 authors declare no conflicts of interest with regard to the funding
-source of this study. GGC was supported by the a FPI 2020 research
-contract (PRE2019-088165) in a project funded by the Spanish Ministry of
-Science, Innovation, and Universities (PGC2018-101831-B-I00). DAV was
-supported by the European Union's Horizon 2023 research and innovation
-program under Marie Skłodowska--Curie Grant (765556) and a postdoctoral
+source of this study. GGC was supported by a FPI 2020 research contract
+(PRE2019-088165) in a project funded by the Spanish Ministry of Science,
+Innovation, and Universities (PGC2018-101831-B-I00). DAV was supported
+by the European Union's Horizon 2023 research and innovation program
+under Marie Skłodowska--Curie Grant (765556) and a postdoctoral
 fellowship from the Foundation for Science and Technology of Portugal
-(UIDB/00214/2020). NSG was supported by Academia 2019 award from the
-Catalan Institution for Research and Advanced Studies (ICREA). We are
-grateful to Chiara Santolin, Ege E. Özer, Alicia Franco-Martínez,
-Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and
-Perception research group for their helpful feedback. We thank Xavier
-Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical
-support, and Cristina Dominguez and Katia Pistrin for their efforts in
-recruiting infants. We also thank all families and infants who
-participated in the experiments.}}
+(UIDB/00214/2020). IC was supported by the Investigo program funded by
+the European Union's NextGenerationEU (NGEU) recovery plan. NSG was
+supported by Academia 2019 award from the Catalan Institution for
+Research and Advanced Studies (ICREA). We are grateful to Chiara
+Santolin, Ege E. Özer, and the rest of the Speech Acquisition and
+Perception research group, and to Alicia Franco-Martínez and Cristina
+Rodríguez-Prada, for their helpful feedback. We thank Xavier Mayoral,
+Silvia Blanch, and Cristina Cuadrado for their technical support, and
+Cristina Dominguez and Katia Pistrin for their efforts in recruiting
+infants. We also thank all families and infants who participated in the
+experiments.}}
 \author{
@@
 \begin{abstract}
-Bilingual infants show equivalent developmental trajectories of lexical
-acquisition compared to their monolingual peers. This is remarkable,
-given the increased complexity of their linguistic input. Recent studies
-suggested that bilingual vocabulary growth is boosted by the number of
-cognates shared by the pair of languages being learned, and that this
-facilitation effect is driven by a stronger parallel activation of
-cognates during linguistic exposure, compared to non-cognates. The
-mechanisms behind this facilitation are still unclear. In this study, we
-capitalise on accumulator models of language acquisition to propose an
-account of bilingual lexical acquisition in which parallel activation
-increases the rate at which children accumulate learning instances for
-words in both languages, even in fully monolingual situations. Under
-this hypothesis, we predicted a stronger cognate facilitation for words
-to which children were exposed less frequently (low-exposure words), as
-they are co-activated by their translation more often than high-exposure
-words do. We developed an extensive online vocabulary checklist, the
-Barcelona Vocabulary questionnaire (BVQ) to collect vocabulary data from
-366 Catalan-Spanish bilingual toddlers aged 12 to 32 months. We then
-used Bayesian explanatory item response theory to model the acquisition
-trajectories of 302 Catalan/Spanish translation equivalents. We found an
-interaction between exposure and cognateness, which pointed to
-cognateness facilitating the aquisition of low-exposure words, but not
-of mean exposure or high-exposure words. Overall, our findings suggest
-that cognateness plays a key role in bilingual lexical acquisition, and
+Bilingual infants' developmental trajectories of lexical acquisition are
+equivalent to those shown by their monolingual peers. This is
+remarkable, given the increased complexity of their linguistic input.
+Recent studies suggested that bilingual vocabulary growth is boosted by
+the number of cognates shared by the pair of languages being learned,
+and that this facilitation effect is driven by a stronger parallel
+activation of cognates during linguistic exposure, compared to
+non-cognates. The mechanisms behind this facilitation are still unclear.
+In this study, we capitalised on accumulator models of language
+acquisition to propose an account of bilingual lexical acquisition in
+which parallel activation increases the rate at which children
+accumulate learning instances for words in both languages, even in fully
+monolingual situations. Under this hypothesis, we predicted a stronger
+cognate facilitation for words to which children were exposed less
+frequently (low-exposure words), as they are co-activated by their
+translation more often than high-exposure words do. We developed an
+extensive online vocabulary checklist, the Barcelona Vocabulary
+questionnaire (BVQ), to collect vocabulary data from 366 Catalan-Spanish
+bilingual toddlers aged 12 to 32 months. We then used Bayesian
+explanatory item response theory to model the acquisition trajectories
+of 302 Catalan/Spanish translation equivalents. We found an interaction
+between exposure and cognateness, which pointed to cognateness
+facilitating the aquisition of low-exposure words, but not of mean
+exposure or high-exposure words. Overall, our findings suggest that
+cognateness plays a key role in bilingual lexical acquisition, and
 provides evidence for a frequency-mediated facilitation effect driven by
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 et al., 2014}). A more detailed analysis of the words in bilinguals'
-lexicon show some interesting patterns.
+lexicon shows some interesting patterns.
 
@@
 non-cognate word. The differences in the percentages of cognate and
-non-cognate words between two languages is related to historic reasons:
-languages typologically close (like Dutch and English or Italian and
-Spanish) share more cognates than languages typologically distant (like
-English and Chinese, or Urdu and Spanish). The conclusion that cognate
-words are easier to learn is based on two types of evidence: studies
-investigating vocabulary sizes in children learning language pairs with
-different percentages of cognates (that is, differing in their
-typological distance) and studies comparing the number of cognate and
-non-cognate words children know in a specific language pair.
+non-cognate words between two languages is related to historical
+reasons: languages typologically close (like Dutch and English or
+Italian and Spanish) share more cognates than languages typologically
+distant (like English and Chinese, or Urdu and Spanish). The conclusion
+that cognate words are easier to learn is based on two types of
+evidence: studies investigating vocabulary sizes in children learning
+language pairs with different percentages of cognates (that is,
+differing in their typological distance) and studies comparing the
+number of cognate and non-cognate words children know in a specific
+language pair.
 
@@
 Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh. The
-authors calculated the average phonological overlap between the words in
-each of these additional languages and their translation equivalents in
+authors calculated the average lexical overlap between the words in each
+of these additional languages and their translation equivalents in
 English, which was taken as a proxy of the degree of cognateness between
@@
 
-A second set of studies concluded that cognates are overrepresented in
+A second set of studies suggested that cognates are overrepresented in
 bilinguals' early lexicon. Bosch \& Ramon-Casas
@@
 English-German bilingual child produced cognates earlier than
-non-cognates, on average. Mitchell et al.
+non-cognates, on average. But the high proportion of cognates in the
+vocabulary of the participants in these two studies may not necessarily
+evidence of a facilitation effect of cognateness, but rather of simply
+the high proportion of cognates present in the pair of languages being
+learned. For instance, if two given languages share a high proportion of
+cognates like \(\sim 70\)\%, the vocabulary contents of children
+learning both languages should, in principle, approximate such
+proportion of cognates, even in the absence of a cognateness
+facilitation effect. More recently, Mitchell et al.
 (\protect\hyperlink{ref-mitchell2022cognates}{2022}) addressed this
-issue in a larger sample longitudinal study. These authors collected
-expressive vocabulary data of 47 16- to 30-month-old French-English
-bilinguals living in Canada, in both languages. They created two lists
-of translation equivalents: one made of 131 cognates, and one made of
-406 non-cognates. The proportion of words that children were reportedly
-able to produce was higher in the cognate lists than in the non-cognate
-list across ages, even when both lists were matched by semantic category
-(furniture, animals, food were similarity represented in both lists) and
-age-of-acquisition norms (an index of word difficulty). Taken together,
-the results of these two lines of research provide a strong support to
-the hypothesis that phonological similarity (as reflected in
-cognateness) plays a facilitation role in bilingual word acquisition.
+issue in a longitudinal study with a larger sample. The authors
+collected expressive vocabulary data of 47 16- to 30-month-old
+French-English bilinguals living in Canada, in both languages. They
+created two lists of translation equivalents: one made of 131 cognates,
+and one made of 406 non-cognates. The proportion of words that children
+were reportedly able to produce was higher in the cognate lists than in
+the non-cognate list across ages. Critically, this difference persisted
+after both lists were matched in size, controlling their semantic
+category (i.e., furniture, animals, food were similarity represented in
+both lists) and age-of-acquisition norms (an index of word difficulty).
+Taken together, the results of these two lines of research have been
+considered as providing support to the hypothesis that phonological
+similarity (as reflected in cognateness) plays a facilitation role in
+bilingual word acquisition.
 
@@
   representation. This activation may result from the infant being
-  exposed to the word-form, or the result of activation spreading
+  exposed to the actual word-form, or the result of activation spreading
   through phonological or semantic links across lexical representation,
@@
 phonological similarity than non-cognates, the former should be
-co-activated more frequently than the latter. This should lead to a
-faster accumulation of learning instances for cognates, compared to
+co-activated more strongly than the latter. This should lead to a faster
+accumulation of learning instances for cognates, compared to
 non-cognates. Parallel activation would allow bilingual children to
@@
 asymmetric across languages: words from the lower-exposure language
-(i.e., the language to which the infants are exposed to more often)
-would receive additional activation through parallel activation more
-frequently than words from the higher-exposure language. Therefore, the
-acquisition of words from the lower-exposure language would benefit more
-strongly from their cognate status than words from the higher-exposure
-language.
+would receive stronger activation from words in the higher-exposure
+language than vice versa. Therefore, the acquisition of words from the
+lower-exposure language would benefit more strongly from their cognate
+status than words from the higher-exposure language. This asymmetric
+cross-language activation would be consistent with previous reports of
+larger priming effects from the dominant to the non-dominant language,
 
@@
 
-We collected 436 responses to the questionnaire from 366 distinct
+We collected 440 responses to the questionnaire from 369 distinct
 children from the Metropolitan Area of Barcelona between the 30th of
-March, 2020 and the 31th of October, 2022: 312 of those participants
-participated once, 42 twice, 8 three times, and 4 four times. Recurrent
+marzo, 2020 and the 31th of octubre, 2022: 314 of those participants
+participated once, 43 twice, 8 three times, and 4 four times. Recurrent
 participants provided responses with a minimum of 25 days between
@@
 Table~\ref{tbl-participants} for a more detailed description of the
-sample). In total, 70 participants (16.06\%) participants were reported
+sample). In total, 70 participants (15.91\%) participants were reported
 to be exposed to a third language other than Catalan and Spanish. All
@@
 \midrule
-75-100\% & 18 & 23 & 36 & 38 & 20 & 7\\
+75-100\% & 18 & 23 & 37 & 38 & 20 & 7\\
 50-75\% & 8 & 13 & 30 & 41 & 18 & 1\\
-25-50\% & 10 & 17 & 45 & 29 & 17 & 0\\
+25-50\% & 10 & 18 & 46 & 30 & 17 & 0\\
 0-25\% & 7 & 11 & 21 & 17 & 8 & 1\\
 \midrule
-N & 43 & 64 & 132 & 125 & 63 & 9\\
+N & 43 & 65 & 134 & 126 & 63 & 9\\
 \bottomrule
@@
 \emph{educació universitària/educación universitaria} {[}university
-degree{]}. Most families reported university studies (356, 82\%),
+degree{]}. Most families reported university studies (358, 82\%),
 followed by families were the highest educational attainment were
-vocational studies (59, 14\%), secondary education (8, 2\%),
+vocational studies (61, 14\%), secondary education (8, 2\%),
 complementary studies (6, 1\%), primary education (1, \textless1\%), and
@@
 302 Spanish translations (302 translation equivalents). After collecting
-participants' responses, the final dataset consisted of 138,078
+participants' responses, the final dataset consisted of 139,326
 observations, each corresponding to a single response of one participant
 to one item. Each translation equivalent received a median of 234
-responses (\emph{Min} = 106, \emph{Max} = 872) from participants, both
+responses (\emph{Min} = 108, \emph{Max} = 880) from participants, both
 languages pooled together.
@@
 bilinguals' exposure to a given word-form is not only a function of the
-word-forms lexical frequency, but also of the quantitative input they
+word-form's lexical frequency, but also of the quantitative input they
 receive from the language such word-form belongs to, we expressed
@@
 and tokens available in their corresponding CHILDES corpora, so they
-were extracted from the English corpora instead. We then mapped the
-lexical frequencies of the English words to their Catalan and Spanish
+were extracted from the English corpora instead. We mapped the lexical
+frequencies of the English words to their Catalan and Spanish
 translations (see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
@@
 scores (\protect\hyperlink{ref-vanheuven2014subtlexuk}{Van Heuven et
-al., 2014}; \protect\hyperlink{ref-zipf1949human}{Zipf, 1949}). We then
+al., 2014}; \protect\hyperlink{ref-zipf1949human}{Zipf, 1949}). We
 multiplied the resulting lexical frequencies by the reported degree of
@@
 that both word-forms are completely different (e.g.,
-/\textipa{"taw.l5}-\textipa{"me.sa}/, \emph{table}), and one indicates
-that the two word-forms are identical (e.g.,
-/\textipa{"mar}-\textipa{"mar}/, \emph{sea}) (see
+/\textipa{"taw.l5}/--/\textipa{"me.sa}/, \emph{table}), and one
+indicates that the two word-forms are identical (e.g.,
+/\textipa{"mar}/--/\textipa{"mar}/, \emph{sea}) (see
 \protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
@@
 (\protect\hyperlink{ref-kruschke2018bayesian}{2018}). First, we
-specified a region of practical equivalence (ROPE) from -2.5\% to
-+2.5\%, in the probability scale. This region indicates the range of
+specified a region of practical equivalence (ROPE) from -0.025 to
++0.025, in the probability scale. This region indicates the range of
 values that we considered equivalent to zero. We then summarised the
@@
 using the \texttt{tidybayes}
-(\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) and
-\texttt{marginaleffects}
-(\protect\hyperlink{ref-arel-bundock2022marginaleffects}{Arel-Bundock,
-2022}) R packages.
+(\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) R package.
 
@@
 posterior distribution in the probability scale. 95\% HDI: 95\% highest
-density interval of the distribution. p(ROPE): overlap between the 95\%
-HDI and the ROPE, indicating the posterior probability that the true
-value of the coefficient is equivalent to zero. In the right column, we
-show a graphical representation of the median and 95\% HDI of each
-coefficient, with a reference dotted line indicating the location of
-zero. }\tabularnewline
+density interval of the distribution. \emph{p}(ROPE): overlap between
+the 95\% HDI and the ROPE, indicating the posterior probability that the
+true value of the coefficient is equivalent to zero. In the right
+column, we show a graphical representation of the median and 95\% HDI of
+each coefficient, with a reference dotted line indicating the location
+of zero. }\tabularnewline
 
 \centering
-\begin{tabular}{lclr>{}c}
+\begin{tabular}{lclr}
 \toprule
- & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
+ & $\beta$ & 95\% HDI & $p(\text{ROPE})$\\
 \midrule
 \addlinespace[0.3em]
-\multicolumn{5}{l}{\textbf{Intercepts (at 22 months)}}\\
-\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & .088 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_2718113c70f2.pdf}\\
-\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_2718cf068d3.pdf}\\
+\multicolumn{4}{l}{\textbf{Intercepts (at 22 months)}}\\
+\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & -\\
+\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & -\\
 \addlinespace[0.3em]
-\multicolumn{5}{l}{\textbf{Slopes (upper bound)}}\\
-\hspace{1em}Age (+1 SD, 4.87, months) & 0.405 & {}[0.357, 0.451] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27186be06e3b.pdf}\\
-\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27186d8d1719.pdf}\\
-\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .037 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271852a346a4.pdf}\\
-\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27181f2174e.pdf}\\
-\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271847bb40f3.pdf}\\
-\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .985 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27184e7d1f57.pdf}\\
-\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271812b03ccf.pdf}\\
-\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .975 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27185801653e.pdf}\\
+\multicolumn{4}{l}{\textbf{Slopes (upper bound)}}\\
+\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000\\
+\hspace{1em}Age (+1 SD, 4.86 months) & 0.405 & {}[0.357, 0.451] & .000\\
+\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000\\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .120\\
+\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000\\
+\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000\\
+\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .974\\
+\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .907\\
 \bottomrule
@@
 
-The main effect of \emph{Age} showed the strongest association with the
+The coefficient of \emph{Age} showed the strongest association with the
 probability of acquisition (\(\beta\) = 0.405, 95\% HDI = {[}0.357,
-0.451{]}), with all posterior samples falling out of the ROPE. A
+1.806{]}), with all posterior samples falling out of the ROPE. A
 one-month increment in age increased a maximum of 0.08 the probability
@@
 a strong effect on the probability of acquisition (\(\beta\) = 0.233,
-95\% HDI = {[}0.201, 0.268{]}). All of the posterior samples of this
+95\% HDI = {[}0.201, 1.074{]}). All of the posterior samples of this
 regression coefficient excluded the ROPE. The impact of this predictor
@@
 probability of acquisition (\(\beta\) = -0.062, 95\% HDI = {[}-0.086,
--0.036{]}). For every phoneme in the word-form, participants were -0.04
+-0.143{]}). For every phoneme in the word-form, participants were -0.04
 less likely to know it. The 95\% HDI of the regression coefficient of
 the \emph{Age} \(\times\) \emph{Exposure} interaction also excluded the
-ROPE (\(\beta\) = 0.071, 95\% HDI = {[}0.039, 0.104{]}), showing that
+ROPE (\(\beta\) = 0.071, 95\% HDI = {[}0.039, 0.414{]}), showing that
 the effect of the word exposure index differed across ages: older
@@
 The posterior distribution of the main effect of cognateness excluded
-the ROPE completely (\(\beta\) = 0.058, 95\% HDI = {[}0.014, 0.104{]}).
+the ROPE completely (\(\beta\) = 0.058, 95\% HDI = {[}0.014, 0.416{]}).
 For every 10\% increment in cognateness, the acquisition of a word
@@
 interaction excluded the ROPE entirely (\(\beta\) = -0.06, 95\% HDI =
-{[}-0.069, -0.046{]}), suggesting that the effect of cognateness on a
+{[}-0.069, -0.184{]}), suggesting that the effect of cognateness on a
 word's probability of acquisition changed depending on participants'
@@
 parallel activation of the two languages plays a central role during the
-consolidation of early representations in the bilingual lexicon, and in
+acquisition of early representations in the bilingual lexicon, and in
 which the dynamics of co-activation between translation equivalents
@@
 acquisition and word learning are, to a large extent, dissociated from
-the mechanisms proposed by previous work on word processing. Accumulator
-models of language acquisition may provide a convenient theoretical
-framework to narrow this gap.
+the mechanisms proposed by previous work on word processing. The notion
+of \emph{accumulator}, as conceptualised by accumulator models of
+language acquisition, may provide a convenient theoretical framework to
+narrow this gap.
 
@@
 accounts (\protect\hyperlink{ref-hidaka2013computational}{Hidaka, 2013};
-\protect\hyperlink{ref-mcmurray2007defusing}{McMurray, 2007}).
-Capistalising on accumulator models, we extended this account to the
-bilingual case, suggesting that the cognate facilitation effect on
-bilingual word acquisition is the result of cognate words being
-activated more strongly by their translation than non-cognates,
-therefore accumulating learning instances at a faster rate. We
-hypothesised that when a bilingual child is exposed to a word-form, they
-activate not only its corresponding lexical representation, but also the
-lexical representation of its translation. This cross-language
-activation occurs at the phonological level, and the amount of
-co-activation that spreads from the spoken word to its translation is
-proportional to the amount of phonological similarity between both
-word-forms. Cognates would receive more activation from their
-translation than non-cognates, leading children to accumulate learning
-instances with cognate words at a faster rate than with non-cognate
-words. As a result, lexical representations of cognate words would
-consolidate at earlier ages than those of non-cognate words.
+\protect\hyperlink{ref-mcmurray2007defusing}{McMurray, 2007}). Using the
+notion of accumulator, we extended this type of account to the bilingual
+case. We suggested that the cognate facilitation effect on bilingual
+word acquisition is the result of cognate words being activated more
+strongly by their translation than non-cognates. This would lead cognate
+words to accumulate learning instances at a faster rate than non-cognate
+words. When a bilingual child is exposed to a word-form, they activate
+not only its corresponding lexical representation, but also the lexical
+representation of its translation. The amount of co-activation that
+spreads from the spoken word to its translation is proportional to the
+amount of phonological similarity between both word-forms. Cognates
+would receive more activation from their translation than non-cognates,
+leading children to accumulate learning instances with cognate words at
+a faster rate than with non-cognate words. As a result, lexical
+representations of cognate words would consolidate at earlier ages than
+those of non-cognate words.
 
@@
 the former scenario predicts that every learning instance for a given
-word-form contributes to the consolidation of the representation of such
-word in the lexicon, while the consolidation of its translation remains
+word-form contributes to the acquisition of the representation of such
+word in the lexicon, while the acquisition of its translation remains
 unaffected by such experience. In the latter scenario, a learning
 instance to the same word-form would contribute not only to the
-consolidation of the representation of such word, but also, to some
-extent, to the consolidation of its translation. Our findings provide
+acquisition of the representation of such word, but also, to some
+extent, to the acquisition of its translation. Our findings provide
 strong support for an account of bilingual vocabulary growth in which
@@
 sounding word-forms in equivalent referential contexts (e.g., hearing
-/\textipa{"gat}\}/ and /\textipa{"ga.to}\}/ in the presence of a cat),
-it is possible that infants classify both are as acceptable variations
-of the same word, therefore treating them as a single lexical item. This
-would lead to a faster increase in cumulative learning instances, and to
+/\textipa{"gat}/ and /\textipa{"ga.to}/ in the same situations), it is
+possible that infants classify both are as acceptable variations of the
+same word, therefore treating them as a single lexical item. This would
+lead to a faster increase in cumulative learning instances, and to
 earlier ages of acquisition for cognate translation equivalents (for
@@
 \& Santolin, 2020}). Bilinguals differ across multiple dimensions. Such
-differences span from exclusively linguistic factors such as the amount
+differences span from exclusively linguistic factors; such as the amount
 of overlap between the phonemic inventories of the two languages being
 learned (e.g., low, like the case of English and Mandarin, or high like
-the case of Spanish and Greek), to higher-level factors like the
+the case of Spanish and Greek), to extralinguistic factors like the
 socio-linguistic situation in which the two languages co-exist (e.g., in
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-arel-bundock2022marginaleffects}{}}%
-Arel-Bundock, V. (2022). \emph{Marginaleffects: Marginal effects,
-marginal means, predictions, and contrasts}.
-\url{https://CRAN.R-project.org/package=marginaleffects}
-
 \leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-arel-bundock2022marginaleffects}{}}%
-Arel-Bundock, V. (2022). \emph{Marginaleffects: Marginal effects,
-marginal means, predictions, and contrasts}.
-\url{https://CRAN.R-project.org/package=marginaleffects}
-
 \leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%

### 2023-07-23 R/items.R

@@ -54,16 +54,14 @@ get_childes_frequencies <- function(collection = "Eng-NA",
                              "corpus_id",
                              "speaker_id", 
                              "transcript_id")) |>
-            mutate(
-                id = as.character(id),
-                age_months = target_child_age,
-                age_bin = as.integer(floor(age_months / 6) * 6),
-                token = tolower(gloss)
+            mutate(id = as.character(id),
+                   age_months = target_child_age,
+                   age_bin = as.integer(floor(age_months / 6) * 6),
+                   token = tolower(gloss)
             ) |>
-            summarise(
-                transcript_count = sum(count),
-                transcript_num_tokens = sum(num_tokens),
-                .by = c(age_bin, token, target_child_id, transcript_id)
+            summarise(transcript_count = sum(count),
+                      transcript_num_tokens = sum(num_tokens),
+                      .by = c(age_bin, token, target_child_id, transcript_id)
             ) |>
             dplyr::filter(between(age_bin,
                                   age_range[1],
@@ -87,7 +85,7 @@ get_childes_frequencies <- function(collection = "Eng-NA",
 
 #' Get item data
 #' 
-#' Levenshtein similarityties are computed using the [stringdist::stringsim()] function (see ?[stringdist::`stringdist-package`]).
+#' Levenshtein similarities are computed using the [stringdist::stringsim()] function (see ?[stringdist::`stringdist-package`]).
 #'
 #' @param bvq_data A named list resulting from calling [get_bvq()]
 #' @param childes A data frame with lexical frequencies extracted from CHILDES, as returned by the [get_childes_frequencies()]
@@ -124,7 +122,7 @@ get_items <- function(bvq_data, childes, class = "Noun")
         distinct(language, te, .keep_all = TRUE) |> 
         mutate(xsampa_flat = bvq::flatten_xsampa(xsampa),
                syll = bvq::syllabify_xsampa(xsampa),
-               n_syll = map_int(syll, length),
+               n_syll = purrr::map_int(syll, length),
                item = gsub("cat_|spa_", "", item)) 
     
     syll_freq <- pool_tmp |> 
@@ -178,7 +176,7 @@ get_items <- function(bvq_data, childes, class = "Noun")
 #' Get CHILDES lexical frequencies
 #'
 #' @param collection CHILDES corpora from where to fetch transcriptions. Takes "Eng-NA" (North American English by default). See [CHILDES Index to corpora(https://childes.talkbank.org/access/) to see options
[…]

### 2023-07-23 R/models.R

@@ -4,63 +4,44 @@
 #' @param ... Arguments to be passed to [brms::brm()]
 fit_model <- function(name, ...) 
 {
-    fit <- brm(...,
-               iter = 2000,
-               chains = 4,
-               init = 0.5,
-               seed = 888,
-               threads = threading(4),
-               backend = "cmdstanr",
-               file = glue::glue("results/fits/{name}.rds"),
-               control = list(adapt_delta = 0.9,
-                              max_treedepth = 15),
-               save_model = glue::glue("stan/{name}.stan"))
+    fit <- brms::brm(...,
+                     iter = 2000,
+                     chains = 4,
+                     init = 0.5,
+                     seed = 888,
+                     backend = "cmdstanr",
+                     file = glue::glue("results/fits/{name}.rds"),
+                     file_refit = "never",
+                     control = list(adapt_delta = 0.9,
+                                    max_treedepth = 15),
+                     save_model = glue::glue("stan/{name}.stan"))
     
     return(fit)
-}
-
-#' Dictionary for predictor names
-#' 
-#' @param data Dataset of responses as generated by the [get_responses()] function
-#' @returns A named character vector
-get_vars_dict <- function(responses) 
-{
-    c(
-        "b_Intercept[1]" = "Comprehension and Production",
-        "b_Intercept[2]" = "Comprehension",
-        "b_age_std" = glue::glue("Age (+1 SD, {round(sd(responses$age), 2)}, months)"),
-        "b_n_phon_std" = glue::glue("Length (+1 SD, {round(sd(responses$n_phon), 2)} phonemes)"),
-        "b_exposure_std" = glue::glue("Exposure (+1 SD, {round(sd(responses$exposure), 2)})"),
-        "b_lv_std" = glue::glue("Cognateness (+1 SD, {round(sd(responses$lv), 2)})"),
-        "b_exposure_std:lv_std" = "Exposure \u00d7 Cognateness",
-        "b_age_std:exposure_std" = "Age \u00d7 Exposure",
-        "b_age_std:lv_std" = "Age \u00d7 Cognateness",
-        "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Cognateness"
-    )
 }
 
 #' Extract posterior draws of fixed effect coefficients from brmsfit model via [tidybayes::gather_draws()].
 #' 
 #' @param model A brmsfit object
[…]

### 2023-07-23 R/participants.R

@@ -25,11 +25,11 @@ get_participants <- function(bvq_data,
             # make sure that degrees of exposure are between 0 and 1
             between(doe_spanish, 0, 1),
             between(doe_catalan, 0, 1),
-            between(doe_others, 0, 1)
-        ) |>
+            between(doe_others, 0, 1),
+        ) |> 
         mutate(time = as.integer(time)) |> 
         # see ?bvq::get_longitudinal
-        bvq::get_longitudinal(longitudinal = longitudinal) |>
+        bvq::get_longitudinal(longitudinal = longitudinal) |> 
         select(id, time, time_stamp = date_finished, list = version,
                age, lp, doe_catalan, doe_spanish, edu_parent) |> 
         arrange(id)

### 2023-07-23 _targets.R

@@ -4,12 +4,12 @@ suppressPackageStartupMessages({
         library(targets)
         library(tarchetypes)
         # data handling, cleaning, and testing
-        library(tidyverse)
+        library(dplyr)
+        library(tidyr)
         library(testthat)
         # modelling
         library(brms)
         library(cmdstanr)
-        library(tidybayes)
         library(collapse)
         # reporting
         library(quarto)
@@ -125,31 +125,30 @@ list(
     tar_target(model_draws,
                get_posterior_draws(model_fit,
                                    data = responses,
-                                   model_vars_dict)),
+                                   vars_dict = model_vars_dict)),
     
     # get summary of posterior draws for population-level effects
     tar_target(model_summary,
                get_posterior_summary(model_fit, 
                                      data = responses,
-                                     model_vars_dict)),
+                                     vars_dict = model_vars_dict)),
     
-    # get posterior draws for group-level effects
-    tar_target(model_draws_re,
-               get_posterior_draws_re(model_fit)),
     
     ## marginal effects --------------------------------------------------------
+    
     tar_target(model_epreds,
-               posterior_predictions(model = model_fit,
-                                     responses, 
-                                     ndraws = 100,
-                                     age_std = scale(seq(7, 40, length.out = 100),
-                                                     mean(responses$age),
-                                                     sd(responses$age)),
-                                     exposure_std = c(-1, 0, 1),
-                                     lv_std = scale(seq(0, 1, 0.5), 
-                                                    mean(responses$lv),
-                                                    sd(responses$lv)),
-                                     n_phon_std = 0)),
+               posterior_predictions(
+                   model = model_fit,
+                   responses, 
+                   age_std = scale(seq(7, 40, length.out = 100),
+                                   mean(responses$age),
+                                   sd(responses$age))[, 1],
+                   exposure_std = c(-1, 0, 1),
+                   lv_std = scale(seq(0, 1, 0.5), 
+                                  mean(responses$lv),
+                                  sd(responses$lv))[, 1],
+                   n_phon_std = 0)
[…]

### 2023-07-23 docs/index.qmd
Simplify dependencies

@@
 ---
-title: "Cognate beginnings to bilingual lexical acquisition"
-subtitle: "Lab notes"
-short-title: "Cognate beginnings"
-author:
-  - name: Gonzalo Garcia-Castro
-    corresponding: true
-    orcid: 0000-0002-8553-4209
-    email: gonzalo.garciadecastro@upf.edu
-    affiliations:
-      - id: cbc
-        name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-        address: Ramon Trias Fargas 25-27
-        city: Barcelona
-        region: Spain
-        postal-code: "08005"
-  - name: Daniela S. Avila-Varela
-    orcid: 0000-0002-3518-8117
-    affiliations: 
-      - ref: cbc
-      - id: ul
-        name: Universidade de Lisboa
-        department: Facultade de Letras
-        address: Alameda da Universidade
-        city: Lisboa
-        region: Portugal
-        postal-code: "1600-214"
-  - name: Ignacio Castillejo
-    orcid: 0000-0001-7445-0416
-    affiliations: 
-      - id: uam 
-        name: Universidad Autónoma de Madrid
-        department: Departamento de Psicología
-        address: Iván Pavlov, 6
-        city: Madrid
-        region: Spain
-        postal-code: "28049"
-  - name: Nuria Sebastian-Galles
-    orcid: 0000-0001-6938-2498
-    affiliations: 
-      - ref: cbc
-date: "May 10th, 2023"
-code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
-keywords:
-  - lexical acquisition
-  - vocabulary
-  - bilingualism
-  - item response theory
-  - bayesian
-csl: "../assets/apa7.csl"
-bibliography: "../assets/references.bib"
-link-citations: true
-editor: source
-
-code-fold: true
-warning: false
-
-format:
-  html: default
-
-fig-dpi: 1000
 ---
@@
 
-yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
+yaml_metadata <- rmarkdown::yaml_front_matter(
+    here::here("manuscript", "manuscript.qmd")
+)
+
 title <- yaml_metadata$title
@@
 thanks <- yaml_metadata$thanks
-
 ```

### 2023-07-23 manuscript/_appendix/_model-doe.qmd
Simplify dependencies

@@
 #| label: tbl-coefs-doe
-#| fig-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency (*Frequency*) and degree of exposure (*DoE*) separately.
+#| tbl-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency (*Frequency*) and degree of exposure (*DoE*) separately.
 Median: median of the posterior distribution in the probability scale.
 95\\% HDI: 95\\% highest density interval of the distribution.
-p(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
+*p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
 In the right column, we show a graphical representation of the median and 95% HDI of each coefficient, with a reference dotted line indicating the location of zero."
@@
 # summarise posterior draws
-tbl_data <- bind_rows("Exposure model" = model_summary,
-                      "DoE-Frequency model" = posterior_doe_summary,
+tbl_data <- list("Exposure model" = model_summary,
+                 "DoE-Frequency model" = posterior_doe_summary) |> 
+    purrr::map_df(\(x) mutate(x, across(.variable_name, as.character)),
                   .id = ".model") |> 
     mutate(
-        across(c(.median, .lower, .upper),
+        across(c(.value, .lower, .upper),
                function(x) {
@@
         )) |>
-    filter(.type == "Slopes") |> 
-    select(.model, .variable_name, .median, .lower, .upper, .rope_overlap) |> 
+    filter(.type=="Slopes") |> 
+    select(.model, .variable_name, .value, .lower, .upper, .rope) |> 
     mutate(
-        across(.median:.upper, plogis),
+        across(.value:.upper, plogis),
         .variable_name = gsub("×", '$\\\\times$', .variable_name),

### 2023-07-23 manuscript/appendix.qmd
Simplify dependencies

@@
 ---
-csl: "../assets/apa7.csl"
-bibliography: "../assets/references.bib"
+csl: "_assets/apa7.csl"
+bibliography: "_assets/references.bib"
 link-citations: true
@@
 
-targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))
+targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables,
+                    posterior_syllables_summary))
 

### 2023-07-23 manuscript/manuscript.qmd
Simplify dependencies

@@
 abstract: |
-    Bilingual infants show equivalent developmental trajectories of lexical acquisition compared to their monolingual peers.
+    Bilingual infants' developmental trajectories of lexical acquisition are equivalent to those shown by their monolingual peers.
 This is remarkable, given the increased complexity of their linguistic input.
@@
 The mechanisms behind this facilitation are still unclear.
-In this study, we capitalise on accumulator models of language acquisition to propose an account of bilingual lexical acquisition in which parallel activation increases the rate at which children accumulate learning instances for words in both languages, even in fully monolingual situations.
+In this study, we capitalised on accumulator models of language acquisition to propose an account of bilingual lexical acquisition in which parallel activation increases the rate at which children accumulate learning instances for words in both languages, even in fully monolingual situations.
 Under this hypothesis, we predicted a stronger cognate facilitation for words to which children were exposed less frequently (low-exposure words), as they are co-activated by their translation more often than high-exposure words do.
-We developed an extensive online vocabulary checklist, the Barcelona Vocabulary questionnaire (BVQ) to collect vocabulary data from 366 Catalan-Spanish bilingual toddlers aged 12 to 32 months.
+We developed an extensive online vocabulary checklist, the Barcelona Vocabulary questionnaire (BVQ), to collect vocabulary data from 366 Catalan-Spanish bilingual toddlers aged 12 to 32 months.
 We then used Bayesian explanatory item response theory to model the acquisition trajectories of 302 Catalan/Spanish translation equivalents.
@@
     The authors declare no conflicts of interest with regard to the funding source of this study.
-GGC was supported by the a FPI 2020 research contract (PRE2019-088165) in a project funded by the Spanish Ministry of Science, Innovation, and Universities (PGC2018-101831-B-I00).
+GGC was supported by a FPI 2020 research contract (PRE2019-088165) in a project funded by the Spanish Ministry of Science, Innovation, and Universities (PGC2018-101831-B-I00).
 DAV was supported by the European Union's Horizon 2023 research and innovation program under Marie Skłodowska–Curie Grant (765556) and a postdoctoral fellowship from the Foundation for Science and Technology of Portugal (UIDB/00214/2020).
+IC was supported by the Investigo program funded by the European Union's NextGenerationEU (NGEU) recovery plan.
 NSG was supported by Academia 2019 award from the Catalan Institution for Research and Advanced Studies (ICREA).
 We are grateful to Chiara Santolin, Ege E.
-Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and Perception research group for their helpful feedback.
+Özer, and the rest of the Speech Acquisition and Perception research group, and to Alicia Franco-Martínez and Cristina Rodríguez-Prada, for their helpful feedback.
 We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support, and Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants.
@@
 library(ggplot2)
+
+options(knitr.kable.NA = "-")
 
@@
 Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-A more detailed analysis of the words in bilinguals’ lexicon show some interesting patterns.
+A more detailed analysis of the words in bilinguals’ lexicon shows some interesting patterns.
 
@@
 For instance, the Spanish translation equivalent of *cat* is *gato*, a cognate word; the translation equivalent of *dog* is *perro*, a non-cognate word.
-The differences in the percentages of cognate and non-cognate words between two languages is related to historic reasons: languages typologically close (like Dutch and English or Italian and Spanish) share more cognates than languages typologically distant (like English and Chinese, or Urdu and Spanish).
+The differences in the percentages of cognate and non-cognate words between two languages is related to historical reasons: languages typologically close (like Dutch and English or Italian and Spanish) share more cognates than languages typologically distant (like English and Chinese, or Urdu and Spanish).
 The conclusion that cognate words are easier to learn is based on two types of evidence: studies investigating vocabulary sizes in children learning language pairs with different percentages of cognates (that is, differing in their typological distance) and studies comparing the number of cognate and non-cognate words children know in a specific language pair.
@@
 The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh.
-The authors calculated the average phonological overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of cognateness between each pair of languages.
+The authors calculated the average lexical overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of cognateness between each pair of languages.
 Floccia and co-workers reported an increase in vocabulary size in the additional language (i.e., not English) associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
@@
 
-A second set of studies concluded that cognates are overrepresented in bilinguals’ early lexicon.
+A second set of studies suggested that cognates are overrepresented in bilinguals’ early lexicon.
 @bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months and found that cognates represented a larger proportion of participant’s vocabulary than non-cognates.
 @schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual child produced cognates earlier than non-cognates, on average.
-@mitchell2022cognates addressed this issue in a larger sample longitudinal study.
-These authors collected expressive vocabulary data of 47 16- to 30-month-old French-English bilinguals living in Canada, in both languages.
+But the high proportion of cognates in the vocabulary of the participants in these two studies may not necessarily evidence of a facilitation effect of cognateness, but rather of simply the high proportion of cognates present in the pair of languages being learned.
+For instance, if two given languages share a high proportion of cognates like $\sim 70$%, the vocabulary contents of children learning both languages should, in principle, approximate such proportion of cognates, even in the absence of a cognateness facilitation effect.
+More recently, @mitchell2022cognates addressed this issue in a longitudinal study with a larger sample.
+The authors collected expressive vocabulary data of 47 16- to 30-month-old French-English bilinguals living in Canada, in both languages.
 They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
-The proportion of words that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age-of-acquisition norms (an index of word difficulty).
-Taken together, the results of these two lines of research provide a strong support to the hypothesis that phonological similarity (as reflected in cognateness) plays a facilitation role in bilingual word acquisition.
+The proportion of words that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages.
+Critically, this difference persisted after both lists were matched in size, controlling their semantic category (i.e., furniture, animals, food were similarity represented in both lists) and age-of-acquisition norms (an index of word difficulty).
+Taken together, the results of these two lines of research have been considered as providing support to the hypothesis that phonological similarity (as reflected in cognateness) plays a facilitation role in bilingual word acquisition.
 
@@
 We rather consider that a learning instance of a word is so to the degree of the strength of activation of its phonological representation.
-This activation may result from the infant being exposed to the word-form, or the result of activation spreading through phonological or semantic links across lexical representation, as in the case of parallel activation.].
-The strength of this co-activation is proportional to the phonological similarity between the two translation equivalents; given that cognates share higher phonological similarity than non-cognates, the former should be co-activated more frequently than the latter.
+This activation may result from the infant being exposed to the actual word-form, or the result of activation spreading through phonological or semantic links across lexical representation, as in the case of parallel activation.].
+The strength of this co-activation is proportional to the phonological similarity between the two translation equivalents; given that cognates share higher phonological similarity than non-cognates, the former should be co-activated more strongly than the latter.
 This should lead to a faster accumulation of learning instances for cognates, compared to non-cognates.
-Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across languages: words from the lower-exposure language (i.e., the language to which the infants are exposed to more often) would receive additional activation through parallel activation more frequently than words from the higher-exposure language.
+Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across languages: words from the lower-exposure language would receive stronger activation from words in the higher-exposure language than vice versa.
 Therefore, the acquisition of words from the lower-exposure language would benefit more strongly from their cognate status than words from the higher-exposure language.
+This asymmetric cross-language activation would be consistent with previous reports of larger priming effects from the dominant to the non-dominant language,
 
@@
     pull(n) |>
-    set_names(c("catalan", "spanish"))
+    purrr::set_names(c("catalan", "spanish"))
 
@@
     pivot_wider(names_from = language,
-                values_from = n) |> 
-    janitor::clean_names() |> 
+                values_from = n,
+                names_repair = janitor::make_clean_names) |> 
     mutate(version = tolower(version))
@@
         total = rowSums(across(a:d)),
-        items = map(items, function(x) {
+        items = purrr::map(items, function(x) {
             paste(tail(x, 2),
@@
 
-knitr::kable(tbl_data,
+kable(tbl_data,
       format = "latex", 
@@
       align = "lccccl")  |> 
-    kableExtra::row_spec(nrow(tbl_data)-1, 
+    row_spec(nrow(tbl_data)-1, 
              hline_after = TRUE)
@@
                   by = c("id", "age")) |> 
-        group_by(id) |> 
+        group_by() |> 
         arrange(id, time_stamp) |> 
-        mutate(diff_lag = time_stamp-dplyr::lag(time_stamp, default = NA)) |>
+        mutate(diff_lag = time_stamp-dplyr::lag(time_stamp, default = NA),
+               .by = id) |>
         summarise(diff_min = min(diff_lag, na.rm = TRUE),
                   diff_max = max(diff_lag, na.rm = TRUE),
-                  .groups = "drop") |> 
+                  .by = id) |> 
         dplyr::filter(!(diff_min %in% c(-Inf, Inf))) |>
@@
 
-dates <- participants$time_stamp |> 
-    range(na.rm = TRUE) |> 
-    lubridate::as_date()
+dates <- as.Date(range(participants$time_stamp, na.rm = TRUE))
 
@@
 
-options(knitr.kable.NA = 0)
-
 tbl_data <- participants |>
-    mutate_at(vars(starts_with("doe_")), 
-              (\(x) floor(x * 10) / 10)) |> 
+    mutate(across(starts_with("doe_"), 
+                  \(x) floor(x * 10) / 10)) |> 
     mutate(
@@
                   include.lowest = TRUE),
-        doe_catalan = cut(
-            doe_catalan,
+        doe_catalan = as.character(
+            cut(doe_catalan,
                 c(0, 0.25, 0.5, 0.75, 1),
@@
                 include.lowest = TRUE
-        ) |>
-            as.character()
+            ))
     ) |>
@@
                           col.names = colnames(tbl_data))
+
 tbl_data <- rbind(tbl_data, sums_row)
 
-knitr::kable(tbl_data,
+kable(tbl_data,
       format = "latex", 
@@
       align = "lcccccc")  |> 
-    kableExtra::add_header_above(c(" " = 1,
+    add_header_above(c(" " = 1,
                        "Age (months)" = ncol(tbl_data)-1)) |>
-    kableExtra::row_spec(nrow(tbl_data)-1, 
+    row_spec(nrow(tbl_data)-1, 
              hline_after = TRUE)
@@
     group_split(language) |> 
-    set_names(c("Catalan", "Spanish")) |>
-    map_int("n")
+    purrr::set_names(c("Catalan", "Spanish")) |>
+    purrr::map_int("n")
 
@@
 
-rope_interval <- scales::percent(c(lower = -0.1/4,
-                                   upper = 0.1/4),
-                                 accuracy = 0.1)
+rope_interval <- c(lower = -0.1/4, upper = 0.1/4, accuracy = 0.1)
 
@@
 
-We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-forms lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
+We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-form's lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
 First, we extracted the child-directed lexical frequency of each word from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
 Using the corresponding lexical frequencies directly from Catalan and Spanish was not possible due to the low number of Catalan participants and tokens available in their corresponding CHILDES corpora, so they were extracted from the English corpora instead.
-We then mapped the lexical frequencies of the English words to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed them to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human].
-We then multiplied the resulting lexical frequencies by the reported degree of exposure of the child to Catalan or Spanish.
+We mapped the lexical frequencies of the English words to their Catalan and Spanish translations [see @fourtassi2020growth for a similar approach], and transformed them to Zipf scores [@vanheuven2014subtlexuk; @zipf1949human].
+We multiplied the resulting lexical frequencies by the reported degree of exposure of the child to Catalan or Spanish.
 For instance, for a child whose degree of exposure is 80% for Catalan and 20% for Spanish, the expected *Exposure* score to the Catalan word *cotxe* [*car*]---with a lexical frequency of `r round(items$freq[items$item=="cotxe"], 2)`---would be `r round(items$freq[items$item=="cotxe"]*0.80, 2)`, while that of its translation to Spanish *coche* would be `r round(items$freq[items$item=="coche"]*0.20, 2)`.
@@
 Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
-This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}-\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}-\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
+This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}/--/\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}/--/\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
 
@@
 We ran four iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
-Model posterior draws and predictions were handled using the `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.
+Model posterior draws and predictions were handled using the `tidybayes` [@kay2021tidybayes] R package.
 
@@
 post_draws_list <- model_summary |> 
-    mutate(across(.median:.lower,
+    mutate(across(.value:.lower,
                   \(x) ifelse(grepl("Intercept", .type),
@@
 95\\% HDI: 95\\% highest density interval of the distribution.
-p(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
+*p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
 In the right column, we show a graphical representation of the median and 95% HDI of each coefficient, with a reference dotted line indicating the location of zero."
@@
 tbl_data <- model_summary |>
-    select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
+    select(.variable_name, .value, .lower, .upper, .rope, .type) |>
     mutate(
-        across(.median:.upper,
+        across(.value:.upper,
                \(x) ifelse(.type==intercept_label, plogis(x), x/4)),
@@
                       ", ", round(.upper, 3), "]"),
-        .rope_overlap = sub("^(-?)0.", "\\1.",
-                            sprintf("%.3f", .rope_overlap))) |> 
-    select(.variable_name, .median, .lower, .upper, .hdi, .rope_overlap)
+        .rope = ifelse(is.na(.rope),
+                       "-",
+                       sub("^(-?)0.", "\\1.", sprintf("%.3f", .rope)))
+    ) |> 
+    select(.variable_name, .value, .lower, .upper, .hdi, .rope)
+
+col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")
 
 tbl_data |> 
-    mutate(viz = "") |> 
     select(-c(.lower, .upper)) |> 
-    knitr::kable(
-        format = "latex", 
+    kable(format = "latex", 
           escape = FALSE,
-        col.names = c("",
-                      "$\\beta$",
-                      "95\\% HDI",
-                      "$p(\\text{ROPE})$",
-                      ""),
+          col.names = col_names,
           digits = 3,
           booktabs = TRUE,
-        align = "lclrc",
+          align = "lclr",
           linesep = ""
     ) |> 
-    kableExtra::kable_styling() |> 
-    kableExtra::pack_rows("Intercepts (at 22 months)",
+    kable_styling() |> 
+    pack_rows("Intercepts (at 22 months)",
               start_row = 1, 
               end_row = 2) |> 
-    kableExtra::pack_rows("Slopes (upper bound)",
+    pack_rows("Slopes (upper bound)",
               start_row = 3,
-                          end_row = 10) |> 
-    kableExtra::column_spec(
-        5,
-        background = "#f2f2f2",
-        image = kableExtra::spec_pointrange(
-            x = tbl_data$.median, 
-            xmin = tbl_data$.lower, 
-            xmax = tbl_data$.upper, 
-            vline = 0,
-            xaxt = "n",
-            col = "#000000",
-            res = 200)
-    )
-
+              end_row = 10) 
 ```
 
-The main effect of *Age* showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std"]]$.upper, 3)`]), with all posterior samples falling out of the ROPE.
-A one-month increment in age increased a maximum of `r round(post_draws_list[["b_age_std"]]$.median/sd(responses$age), 2)` the probability of acquisition.
-Similarly, the word exposure index (*Exposure*) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 3)`]).
+The coefficient of *Age* showed the strongest association with the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_age_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std"]]$.upper, 3)`]), with all posterior samples falling out of the ROPE.
+A one-month increment in age increased a maximum of `r round(post_draws_list[["b_age_std"]]$.value/sd(responses$age), 2)` the probability of acquisition.
+Similarly, the word exposure index (*Exposure*) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 3)`]).
 All of the posterior samples of this regression coefficient excluded the ROPE.
-The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), 3)` more likely to acquire it.
-Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r round(post_draws_list[["b_n_phon_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_n_phon_std"]]$.lower, 3)`, `r round(post_draws_list[["b_n_phon_std"]]$.upper, 3)`]).
-For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.median)/sd(responses$n_phon), 3)` less likely to know it.
-The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children.
+The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.value/sd(responses$exposure), 3)` more likely to acquire it.
+Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r round(post_draws_list[["b_n_phon_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_n_phon_std"]]$.lower, 3)`, `r round(post_draws_list[["b_n_phon_std"]]$.upper, 3)`]).
+For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.value)/sd(responses$n_phon), 3)` less likely to know it.
+The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children.
 
-The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 3)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 3)`]).
-For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.median/(40*sd(responses$lv)), 3)`.
-The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
+The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 3)`]).
+For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.value/(40*sd(responses$lv)), 3)`.
+The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
 Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
@@
 fig_data <- model_epreds |> 
+    ungroup() |> 
     reframe(.median = mean(.value),
-            .lower = qi(.value, .width = 0.95)[,1],
-            .upper = qi(.value, .width = 0.95)[,2],
+            .lower = tidybayes::qi(.value, .width = 0.95)[, 1],
+            .upper = tidybayes::qi(.value, .width = 0.95)[, 2],
             .by = c(ends_with("_std"), .category)) |> 
@@
                linetype = "dotted") +
-    geom_lineribbon(aes(order = lv),
-                    linewidth = 0.5) +
+    ggdist::geom_lineribbon(aes(order = lv),
+                               size = 0.5) +
     labs(x = "Age (months)",
@@
 Capitalising on accumulator models of language acquisition, we provide a theoretical account of bilingual lexical acquisition.
-In this account, parallel activation of the two languages plays a central role during the consolidation of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
+In this account, parallel activation of the two languages plays a central role during the acquisition of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
 
@@
 But the underpinnings of such effect were unclear: while parallel activation has been extensively described in experimental studies, current paradigms of bilingual word acquisition and word learning are, to a large extent, dissociated from the mechanisms proposed by previous work on word processing.
-Accumulator models of language acquisition may provide a convenient theoretical framework to narrow this gap.
+The notion of *accumulator*, as conceptualised by accumulator models of language acquisition, may provide a convenient theoretical framework to narrow this gap.
 
@@
 Through statistical inference, formalised accumulator models provide meaningful information about parameters of interest like the aforementioned predictors [@mollica2017how; @kachergis2022standard], and allow to generate quantitative predictions about age-of-acquisition and vocabulary growth under competing theoretical accounts [@mcmurray2007defusing; @hidaka2013computational].
-Capistalising on accumulator models, we extended this account to the bilingual case, suggesting that the cognate facilitation effect on bilingual word acquisition is the result of cognate words being activated more strongly by their translation than non-cognates, therefore accumulating learning instances at a faster rate.
-We hypothesised that when a bilingual child is exposed to a word-form, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
-This cross-language activation occurs at the phonological level, and the amount of co-activation that spreads from the spoken word to its translation is proportional to the amount of phonological similarity between both word-forms.
+Using the notion of accumulator, we extended this type of account to the bilingual case.
+We suggested that the cognate facilitation effect on bilingual word acquisition is the result of cognate words being activated more strongly by their translation than non-cognates.
+This would lead cognate words to accumulate learning instances at a faster rate than non-cognate words.
+When a bilingual child is exposed to a word-form, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
+The amount of co-activation that spreads from the spoken word to its translation is proportional to the amount of phonological similarity between both word-forms.
 Cognates would receive more activation from their translation than non-cognates, leading children to accumulate learning instances with cognate words at a faster rate than with non-cognate words.
@@
 These predictions address a critical subject in bilingualism research: do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
-In the context of lexical acquisition, the former scenario predicts that every learning instance for a given word-form contributes to the consolidation of the representation of such word in the lexicon, while the consolidation of its translation remains unaffected by such experience.
-In the latter scenario, a learning instance to the same word-form would contribute not only to the consolidation of the representation of such word, but also, to some extent, to the consolidation of its translation.
+In the context of lexical acquisition, the former scenario predicts that every learning instance for a given word-form contributes to the acquisition of the representation of such word in the lexicon, while the acquisition of its translation remains unaffected by such experience.
+In the latter scenario, a learning instance to the same word-form would contribute not only to the acquisition of the representation of such word, but also, to some extent, to the acquisition of its translation.
 Our findings provide strong support for an account of bilingual vocabulary growth in which the experience and learning outcomes accumulated by the child in one language impact those in the other language.
@@
 It might be argued that our results reflect the fact that cognate translation equivalents are represented in the initial bilingual lexicon as the *same* lexical entry.
-Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}}/ and /\textipa{"ga.to}}/ in the presence of a cat), it is possible that infants classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
+Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the same situations), it is possible that infants classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
 This would lead to a faster increase in cumulative learning instances, and to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
@@
 Bilinguals differ across multiple dimensions.
-Such differences span from exclusively linguistic factors such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to higher-level factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
+Such differences span from exclusively linguistic factors; such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to extralinguistic factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
 This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.

### 2023-07-23 R/predictions.R

@@ -11,21 +11,25 @@
 #' @param levels If group is `"te"` or `"id"`, specific levels (translation equivalents or participants) can be specified. If NULL (default), posterior predictions are generated for all levels of the grouping variable.
 #' @inheritParams marginaleffects::datagrid
 #' 
-generate_newdata <- function(model,
-                             data,
-                             group = NULL,
-                             levels = NULL,
-                             ...) {
-    # some checks
+generate_newdata <- function(model, group = NULL, levels = NULL, ...) {
+    
+    # validate data
+    is.data.provided <- length(list(...)) > 0
+    if (!is.data.provided) {
+        cli::cli_abort("Predictor values must be provided")
+    }
+    
+    # validate group variable names
+    grouping.vars <- names(brms::ranef(model))
     if (!is.null(group)) {
-        if (!(group %in% c("te", "id")) | length(group) > 1) {
-            cli::cli_abort("group must be one of 'te' or 'id")
+        if (!(group %in% grouping.vars) | length(group) > 1) {
+            cli::cli_abort("group must be one of {grouping.vars}")
         }
     }
     
     # make sure that all levels exist in the dataset
     if (!is.null(levels)) {
-        levels_in_data <- levels %in% data[[group]]
+        levels_in_data <- levels %in% model[["data"]][[group]]
         if (!all(levels_in_data)) {
             missing_levels <- paste0(levels[!levels_in_data], collapse = ", ")
             cli::cli_abort("Level {missing_levels} in {.field {group}} is missing")
@@ -34,26 +38,18 @@ generate_newdata <- function(model,
     
     # data frame with prediction combinations
     if (is.null(group)) {
-        nd <- marginaleffects::datagrid(model = model, te = NA, id = NA, ...)
+        nd <- expand_grid(te = NA, id = NA, ...)
     } else {
         # expand predictor levels in data frame to generate predictions
         if (group == "te") {
             # if group is "te", generate predictions for each level of `te`
-            nd <- marginaleffects::datagrid(
-                model = model, 
-                te = levels, 
-                id = NA, 
-                ...) |>
+            nd <- expand_grid(te = levels,  id = NA, ...) |>
                 select(-lv_std) |>
                 left_join(distinct(data, te, lv_std))
         }
         if (group == "id") {
             # expand predictor levels in data frame to generate predictions
-            nd <- marginaleffects::datagrid(
-                model = model, 
-                id = levels,
-                te = NA, ...
-            ) |>
[…]

### 2023-07-23 manuscript/_appendix/_model-doe.qmd
Fix bug in appendix

@@
 tbl_data |> 
-    mutate(viz = "") |> 
     select(-c(.lower, .upper)) |> 
@@
                       "95\\% HDI",
-                      "$p(\\text{ROPE})$",
-                      ""),
+                      "$p(\\text{ROPE})$"),
         digits = 3,

### 2023-07-23 manuscript/appendix.qmd
Fix bug in appendix

@@
   arxiv-pdf:
-    keep-tex: true  
     linenumbers: true

### 2023-07-24 manuscript/manuscript.qmd
Add new paragraph to discussion

@@
 Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
-
 Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
@@
 Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
-This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}/--/\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}/--/\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological for similar approaches].
-
+This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}/--/\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}/--/\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological; for similar approaches].
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
@@
 95\\% HDI: 95\\% highest density interval of the distribution.
-*p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
-In the right column, we show a graphical representation of the median and 95% HDI of each coefficient, with a reference dotted line indicating the location of zero."
+*p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero."
 #| echo: false
@@
 
+The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in adults.
+For  instance, unbalanced (or low-proficiency) bilingual adults benefit more strongly from forward (dominance to non-dominant) cross-language phonological priming during word processing than from backward phonological priming (non-dominant to dominant) [e.g., @grainger1998masked; @voga2007cognate; @de1991lexical; @dimitropoulou2011masked].
+On the contrary, more balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
+Similarly, the infants in the present study benefited from the cognate status of words in their less dominant language, but not in their more dominant language.
+It is possible that in both cases, the underlying cause behind the asymmetric effect of parallel activation lies in lexical representations in the non-dominant language receiving stronger activation from their translations in the dominant language than vice versa.
+
+This dominance-mediated difference in connection strength is contemplated by previous models of bilingual lexical access, such as the Revised Hierarchical Model [RHM; @kroll1994category], but the specific implementation of the source of such difference is still unclear.
+A more recent proposal, the Ontogenic Model of L2 Lexical Representation [@bordag2022ontogenesis; @cook2016fuzzy], points to a 'fuzzier'---or more poorly phonological specification---of lexical representations in the non-dominant second language, compared to those in the dominant language.
+This sub-optimal specification may be due to a mismatch between the phonological inventory of both languages [@cutler2006asymmetric; @sebastian2005influence; @sebastian2006first] or, more generally, an imprecise encoding of the features of the representation in any domain (e.g., form-meaning mapping, orthographical, phonological).
+Although this account seems convenient to explain cross-language asymmetries in language processing, it should be taken into account that most (if not all) of the proposals aforementioned are aimed at explaining results in adults, and not in children.
+Language dominance and proficiency are frequently defined in both populations using different dimensions.
+For instance, the low proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood.
+This might have changed substantially the developmental trajectory of language acquisition they followed, compared to the bilinguals in the present study, who were exposed to both languages from early.
+Moreover, our findings can in principle be explained in the absence of a poorer specification of representations in the non-dominant language.
+As suggested in the introduction, it is possible that simply the fact that infants are exposed more frequently to words in the dominant language leads to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language.
+This explanation does not necessarily compete with words in the dominant language activating more strongly their translations than vice versa, and further research is needed in order to clarify this issue.
+
 It might be argued that our results reflect the fact that cognate translation equivalents are represented in the initial bilingual lexicon as the *same* lexical entry.

### 2023-07-24 manuscript/manuscript.tex
Simplify dependencies
Add new paragraph to discussion

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, breakable, boxrule=0pt, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 responses (\emph{Min} = 108, \emph{Max} = 880) from participants, both
-languages pooled together.
-
-Data processing and visualisation was done in R
-(\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
+languages pooled together. Data processing and visualisation was done in
+R (\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
 version 4.2.2) using the Tidyverse family of packages
@@
 indicates that the two word-forms are identical (e.g.,
-/\textipa{"mar}/--/\textipa{"mar}/, \emph{sea}) (see
-\protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
-\protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020};
-\protect\hyperlink{ref-heeringa2003norwegian}{Heeringa \& Gooskens,
-2003}; \protect\hyperlink{ref-laing2022phonological}{Laing, 2022} for
-similar approaches;
-\protect\hyperlink{ref-schepens2012distributions}{Schepens et al.,
-2012}).
-
-Predictors were standardised before entering the model by subtracting
-the mean of the predictor from each value and dividing the result by the
-standard deviation of the predictor.
+/\textipa{"mar}/--/\textipa{"mar}/, \emph{sea}) {[}see Floccia et al.
+(\protect\hyperlink{ref-floccia2018introduction}{2018}); Heeringa \&
+Gooskens (\protect\hyperlink{ref-heeringa2003norwegian}{2003}); Schepens
+et al. (\protect\hyperlink{ref-schepens2012distributions}{2012});
+Fourtassi et al. (\protect\hyperlink{ref-fourtassi2020growth}{2020});
+Laing (\protect\hyperlink{ref-laing2022phonological}{2022}); for similar
+approaches{]}. Predictors were standardised before entering the model by
+subtracting the mean of the predictor from each value and dividing the
+result by the standard deviation of the predictor.
 
@@
 the 95\% HDI and the ROPE, indicating the posterior probability that the
-true value of the coefficient is equivalent to zero. In the right
-column, we show a graphical representation of the median and 95\% HDI of
-each coefficient, with a reference dotted line indicating the location
-of zero. }\tabularnewline
+true value of the coefficient is equivalent to zero. }\tabularnewline
 
@@
 sharing more cognates with English.
+
+The asymmetric facilitation of cognateness on word acquisition reported
+in the present study parallels previous findings in adults. For
+instance, unbalanced (or low-proficiency) bilingual adults benefit more
+strongly from forward (dominance to non-dominant) cross-language
+phonological priming during word processing than from backward
+phonological priming (non-dominant to dominant)
+(\protect\hyperlink{ref-de1991lexical}{De Groot \& Nas, 1991};
+\protect\hyperlink{ref-dimitropoulou2011masked}{Dimitropoulou et al.,
+2011}; e.g., \protect\hyperlink{ref-grainger1998masked}{Grainger, 1998};
+\protect\hyperlink{ref-voga2007cognate}{Voga \& Grainger, 2007}). On the
+contrary, more balanced (or high-proficiency bilinguals) show an
+equivalent priming facilitation in both directions
+(\protect\hyperlink{ref-basnight2007differences}{Basnight-Brown \&
+Altarriba, 2007};
+\protect\hyperlink{ref-dunabeitia2009masked}{Duñabeitia et al., 2009}).
+Similarly, the infants in the present study benefited from the cognate
+status of words in their less dominant language, but not in their more
+dominant language. It is possible that in both cases, the underlying
+cause behind the asymmetric effect of parallel activation lies in
+lexical representations in the non-dominant language receiving stronger
+activation from their translations in the dominant language than vice
+versa.
+
+This dominance-mediated difference in connection strength is
+contemplated by previous models of bilingual lexical access, such as the
+Revised Hierarchical Model {[}RHM; Kroll \& Stewart
+(\protect\hyperlink{ref-kroll1994category}{1994}){]}, but the specific
+implementation of the source of such difference is still unclear. A more
+recent proposal, the Ontogenic Model of L2 Lexical Representation
+(\protect\hyperlink{ref-bordag2022ontogenesis}{Bordag et al., 2022};
+\protect\hyperlink{ref-cook2016fuzzy}{Cook et al., 2016}), points to a
+`fuzzier'---or more poorly phonological specification---of lexical
+representations in the non-dominant second language, compared to those
+in the dominant language. This sub-optimal specification may be due to a
+mismatch between the phonological inventory of both languages
+(\protect\hyperlink{ref-cutler2006asymmetric}{Cutler et al., 2006};
+\protect\hyperlink{ref-sebastian2006first}{Sebastian-Gallés et al.,
+2006}; \protect\hyperlink{ref-sebastian2005influence}{Sebastián-Gallés
+et al., 2005}) or, more generally, an imprecise encoding of the features
+of the representation in any domain (e.g., form-meaning mapping,
+orthographical, phonological). Although this account seems convenient to
+explain cross-language asymmetries in language processing, it should be
+taken into account that most (if not all) of the proposals
+aforementioned are aimed at explaining results in adults, and not in
+children. Language dominance and proficiency are frequently defined in
+both populations using different dimensions. For instance, the low
+proficiency bilinguals in many of the aforementioned studies acquired
+their second language years after their toddlerhood. This might have
+changed substantially the developmental trajectory of language
+acquisition they followed, compared to the bilinguals in the present
+study, who were exposed to both languages from early. Moreover, our
+findings can in principle be explained in the absence of a poorer
+specification of representations in the non-dominant language. As
+suggested in the introduction, it is possible that simply the fact that
+infants are exposed more frequently to words in the dominant language
+leads to words in the non-dominant language receiving additional
+parallel activation, compared to words in the dominant language. This
+explanation does not necessarily compete with words in the dominant
+language activating more strongly their translations than vice versa,
+and further research is needed in order to clarify this issue.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-basnight2007differences}{}}%
+Basnight-Brown, D. M., \& Altarriba, J. (2007). Differences in semantic
+and translation priming across languages: The role of language direction
+and language dominance. \emph{Memory \& Cognition}, \emph{35}, 953--965.
+
 \leavevmode\vadjust pre{\hypertarget{ref-bergelson2012months}{}}%
@@
 \emph{40}(2), 151--171. \url{https://doi.org/10.1177/0142723719892794}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bordag2022ontogenesis}{}}%
+Bordag, D., Gor, K., \& Opitz, A. (2022). Ontogenesis model of the L2
+lexical representation. \emph{Bilingualism: Language and Cognition},
+\emph{25}(2), 185--201.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cook2016fuzzy}{}}%
+Cook, S. V., Pandža, N. B., Lancaster, A. K., \& Gor, K. (2016). Fuzzy
+nonnative phonolexical representations lead to fuzzy form-to-meaning
+mappings. \emph{Frontiers in Psychology}, \emph{7}, 1345.
+
 \leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cutler2006asymmetric}{}}%
+Cutler, A., Weber, A., \& Otake, T. (2006). Asymmetric mapping from
+phonetic to lexical representations in second-language listening.
+\emph{Journal of Phonetics}, \emph{34}(2), 269--284.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de1991lexical}{}}%
+De Groot, A. M., \& Nas, G. L. (1991). Lexical representation of
+cognates and noncognates in compound bilinguals. \emph{Journal of Memory
+and Language}, \emph{30}(1), 90--123.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dimitropoulou2011masked}{}}%
+Dimitropoulou, M., Duñabeitia, J. A., \& Carreiras, M. (2011). Masked
+translation priming effects with low proficient bilinguals. \emph{Memory
+\& Cognition}, \emph{39}, 260--275.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dunabeitia2009masked}{}}%
+Duñabeitia, J. A., Perea, M., \& Carreiras, M. (2009). Masked
+translation priming effects with highly proficient simultaneous
+bilinguals. \emph{Experimental Psychology}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-2018els}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-grainger1998masked}{}}%
+Grainger, J. (1998). Masked priming by translation equivalents in
+proficient bilinguals. \emph{Language and Cognitive Processes},
+\emph{13}(6), 601--623.
+
 \leavevmode\vadjust pre{\hypertarget{ref-grosjean2021extent}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kroll1994category}{}}%
+Kroll, J. F., \& Stewart, E. (1994). Category interference in
+translation and picture naming: Evidence for asymmetric connection
+between bilingual memory representations. \emph{Journal of Memory and
+Language}, \emph{33}(2), 149--174.
+\url{https://doi.org/10.1006/jmla.1994.1008}
+
 \leavevmode\vadjust pre{\hypertarget{ref-kruschke2018bayesian}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-sebastian2006first}{}}%
+Sebastian-Gallés, N., Rodrı́guez-Fornells, A., Diego-Balaguer, R. de, \&
+Dı́az, B. (2006). First-and second-language phonological representations
+in the mental lexicon. \emph{Journal of Cognitive Neuroscience},
+\emph{18}(8), 1277--1291.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sebastian2005influence}{}}%
+Sebastián-Gallés, N., Echeverrı́a, S., \& Bosch, L. (2005). The influence
+of initial exposure on lexical representation: Comparing early and
+simultaneous bilinguals. \emph{Journal of Memory and Language},
+\emph{52}(2), 240--255.
+
 \leavevmode\vadjust pre{\hypertarget{ref-smithson2014bilingualism}{}}%
@@
 \emph{67}(6), 1176--1190.
+
+\leavevmode\vadjust pre{\hypertarget{ref-voga2007cognate}{}}%
+Voga, M., \& Grainger, J. (2007). Cognate status and cross-script
+translation priming. \emph{Memory \& Cognition}, \emph{35}, 938--952.
 

### 2023-07-24 _targets.R

@@ -137,9 +137,8 @@ list(
     ## marginal effects --------------------------------------------------------
     
     tar_target(model_epreds,
-               posterior_predictions(
+               posterior_epreds(
                    model = model_fit,
-                   responses, 
                    age_std = scale(seq(7, 40, length.out = 100),
                                    mean(responses$age),
                                    sd(responses$age))[, 1],
@@ -249,5 +248,5 @@ list(
                    file_path <- here::here("_targets.yaml")
                    if (file.exists(file_path)) file.remove(file_path)
                }))
-        
+    
 )

### 2023-07-25 manuscript/manuscript.qmd
Fix Makefile and change Docker folder

@@
 The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in adults.
-For  instance, unbalanced (or low-proficiency) bilingual adults benefit more strongly from forward (dominance to non-dominant) cross-language phonological priming during word processing than from backward phonological priming (non-dominant to dominant) [e.g., @grainger1998masked; @voga2007cognate; @de1991lexical; @dimitropoulou2011masked].
+For  instance, unbalanced (or low-proficiency) bilingual adults benefit more strongly from forward (dominant to non-dominant) cross-language phonological priming during word processing than from backward (non-dominant to dominant) phonological priming [e.g.; @grainger1998masked; @voga2007cognate; @de1991lexical; @dimitropoulou2011masked].
 On the contrary, more balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
 Similarly, the infants in the present study benefited from the cognate status of words in their less dominant language, but not in their more dominant language.
-It is possible that in both cases, the underlying cause behind the asymmetric effect of parallel activation lies in lexical representations in the non-dominant language receiving stronger activation from their translations in the dominant language than vice versa.
+It is possible that, in both cases, that such an asymmetric effect of parallel activation is the result of lexical representations in the non-dominant language receiving stronger activation from their translations in the dominant language than vice versa.
+This dominance-mediated difference in connection strength is contemplated by previous models of bilingual lexical access, but these models differ in the specific implementation of the source of the asymmetry.
+Some models hard-code such asymmetries, so that the connections between word representation sand other words or their concepts in one language are stronger than those of word representations in the other language [e.g., the Revised Hierarchical Model; @kroll1994category].
+In other models, like the Multilink model [@dijkstra2019multilink], lexical representations in the non-dominant language have a lower resting activation due to the lower frequency at which their word-forms are used.
+Some other models implement a combination of the former.
+For instance, the BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture] models implemented a combination of (1) lower frequency---in a similar fashion than Multilink---, together with (2) a stronger top-down inhibition of word representations in the non-dominant language.
 
-This dominance-mediated difference in connection strength is contemplated by previous models of bilingual lexical access, such as the Revised Hierarchical Model [RHM; @kroll1994category], but the specific implementation of the source of such difference is still unclear.
+More recent proposals capture some developmental features of the bilingual lexicon, and suggest that structural and function asymmetries between the dominant and the non-dominant language may arise from simply a difference in the frequency of the input, the age of acquisition of the non-dominant language.
+For instance, simulations of subsequent bilinguals from the DevLex-II model [@zhao2010bilingual; @zhao2013simulating] result in an asymmetric cross-language priming, in which words from the dominant (earlier acquired) language primed more strongly the recognition of words in the non-dominant language (later acquired) than in the other direction.
+In addition, the simulated lexica showed sparser connectivity between words in the non-dominant language, compared to the connectivity between words in the dominant language.
+
 A more recent proposal, the Ontogenic Model of L2 Lexical Representation [@bordag2022ontogenesis; @cook2016fuzzy], points to a 'fuzzier'---or more poorly phonological specification---of lexical representations in the non-dominant second language, compared to those in the dominant language.

### 2023-07-25 manuscript/manuscript.qmd
Update Dockerfile

--- 2023-07-23 manuscript/manuscript.qmd
@@
 The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during language processing, even in fully monolingual contexts.
-Evidence with adult bilinguals supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; see @kroll2017bilingual for review].
+Evidence with adult bilinguals supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; @shook2012bimodal; see @kroll2017bilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 These results showed that bilinguals’ Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-Evidence of parallel activation has been reported in bilingual toddlers and children [@vonholzen2012language; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic; @poulin-dubois2013lexical].
+Evidence of parallel activation has been reported in bilingual toddlers and children [@vonholzen2012language; @floccia2020translation; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic; @poulin-dubois2013lexical].
 
@@
 
+The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in toddlers and adults.
+For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [for toddlers see @vonholzen2019impact; @singh2014one; for adults see @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked; but see @jardak2019labels].
+One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language; but see @duyck2009translation; @schoonbaert2009semantic].
+More balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
+These results have been taken as evidence for an asymmetry in the strength of forward and backward connections in the unbalanced bilingual lexicon.
+
+Although implemented in different ways, or found under different assumptions, a dominance-mediated asymmetry is accounted for by previous models of adult bilingual lexical access, like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective into how lexical representations in a second language are formed, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
+This seems to provide a convenient account for the interaction between language dominance and cognateness found in the present study.
+Following a similar pattern to those predicted by the aforementioned models, infants benefited from the cognate status of only words in their low-dominance language, and especially if their exposure to both languages was unbalanced.
+
+These models, however, are aimed at explaining results in adults, and their predictions should be taken with a grain of salt when extended to early language acquisition.
+Language dominance and proficiency are frequently defined in both populations using dimensions other than degree of exposure, which is the most common practice in infant research [@marian2021measuring; @rocha2023defining].
+For instance, the low-proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood.
+This has three important consequences for their assumed developmental trajectories of language acquisition.
+First, the acquisition of the phonological system of the new language must be negotiated with the already acquired phoneme inventory of the first language [e.g., @cutler2006asymmetric; @sebastian2005influence; @sebastian2006first], which is in place around the first year of life [see @werker2015critical for review].
+Second, adults acquiring a second language already possess a system of word-meaning mappings, while simulatenous bilingual infants must generate a lexicon in two languages in the absence of clear word-meaning mappings.
+Third, subjects are assumed to be literate and possess an orthographic system in place, which may shape how new words are integrated in the lexicon and processed during experimental tasks [e.g., @thierry2007brain].
+In this scenario, the acquisition of a second language may take place in a substantially different way compared to how bilingual infants acquire two languages from birth.
+
+The DevLex-II model [@zhao2010bilingual] captures some of these features of the early bilingual lexicon, and considers the case of bilinguals in the process of stimultaneously acquiring their two languages.
+In line with the aforementioned models, DevLex-II predicts asymmetries between the dominant and the non-dominant language that arise from a difference in the frequency of the input, or the order of acquisition of the non-dominant language (even delayed by only some months).
+For instance, simulations from DevLex-II result in an asymmetric cross-language priming, in which words from the dominant (acquired acquired) language primed more strongly the recognition of words in the non-dominant language (later acquired) than in the other direction [@zhao2013simulating].
+In addition, the simulated lexica showed sparser connectivity between words in the non-dominant language, compared to the connectivity between words in the dominant language.
+
+ As suggested in the introduction, it is possible that the asymmetric effect of cognateness found in the present study is simply the result of infants being exposed more frequently to words in the dominant language.
+This would lead to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language, and therefore benefiting more strongly from their cognate status.
+This explanation does not necessarily compete with words in the dominant language activating more *strongly* their translations than *vice versa*.
+Further research is needed in order to clarify this issue.
+
 It might be argued that our results reflect the fact that cognate translation equivalents are represented in the initial bilingual lexicon as the *same* lexical entry.

### 2023-07-25 manuscript/manuscript.qmd
Update Dockerfile

@@
 The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in toddlers and adults.
-For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [for toddlers see @vonholzen2019impact; @singh2014one; for adults see @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked; but see @jardak2019labels].
-One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language; but see @duyck2009translation; @schoonbaert2009semantic].
+For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [for toddlers see @vonholzen2019impact; @singh2014one, for adults see @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked, but see @jardak2019labels].
+One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language, but see @duyck2009translation; @schoonbaert2009semantic].
 More balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
-These results have been taken as evidence for an asymmetry in the strength of forward and backward connections in the unbalanced bilingual lexicon.
+These results have been taken as evidence for an asymmetry in the strength of forward and backward connections in the unbalanced bilingual lexicon Although implemented in different ways, or found under different assumptions, a dominance-mediated asymmetry is accounted for by previous models of adult bilingual lexical access, like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective into how lexical representations in a second language are formed, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
+This seems to provide a convenient account for the interaction between language dominance and cognateness found in the present study, where we found a cognate advantage only for words in their low-dominance language, and more strongly when exposure to both languages was unbalanced.
 
-Although implemented in different ways, or found under different assumptions, a dominance-mediated asymmetry is accounted for by previous models of adult bilingual lexical access, like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective into how lexical representations in a second language are formed, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
-This seems to provide a convenient account for the interaction between language dominance and cognateness found in the present study.
-Following a similar pattern to those predicted by the aforementioned models, infants benefited from the cognate status of only words in their low-dominance language, and especially if their exposure to both languages was unbalanced.
-
-These models, however, are aimed at explaining results in adults, and their predictions should be taken with a grain of salt when extended to early language acquisition.
+These models, however, are aimed at explaining results in adults, and their predictions should be taken with caution when extended to early language acquisition.
 Language dominance and proficiency are frequently defined in both populations using dimensions other than degree of exposure, which is the most common practice in infant research [@marian2021measuring; @rocha2023defining].
 For instance, the low-proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood.
-This has three important consequences for their assumed developmental trajectories of language acquisition.
-First, the acquisition of the phonological system of the new language must be negotiated with the already acquired phoneme inventory of the first language [e.g., @cutler2006asymmetric; @sebastian2005influence; @sebastian2006first], which is in place around the first year of life [see @werker2015critical for review].
-Second, adults acquiring a second language already possess a system of word-meaning mappings, while simulatenous bilingual infants must generate a lexicon in two languages in the absence of clear word-meaning mappings.
+We identify three critical ways in which this prevents a clear comparison between our results and those from studies on second language acuquisition in adults.
+First, in adult second language acquisition, the acquisition of the phonological system of the new language must be negotiated with the already acquired phoneme inventory of the first language [e.g., @cutler2006asymmetric; @sebastian2005influence; @sebastian2006first], which is in place around the first year of life [see @werker2015critical for review].
+Second, adults acquiring a second language already possess a system of form-meaning mappings, while simulatenous bilingual infants must build a lexicon in two languages in the absence of clear form-meaning mappings.
 Third, subjects are assumed to be literate and possess an orthographic system in place, which may shape how new words are integrated in the lexicon and processed during experimental tasks [e.g., @thierry2007brain].
 In this scenario, the acquisition of a second language may take place in a substantially different way compared to how bilingual infants acquire two languages from birth.
-
-The DevLex-II model [@zhao2010bilingual] captures some of these features of the early bilingual lexicon, and considers the case of bilinguals in the process of stimultaneously acquiring their two languages.
-In line with the aforementioned models, DevLex-II predicts asymmetries between the dominant and the non-dominant language that arise from a difference in the frequency of the input, or the order of acquisition of the non-dominant language (even delayed by only some months).
+A more similar case to ours is considered by the DevLex-II model [@zhao2010bilingual], which captures unique features of the early bilingual lexicon, and considers the case of bilinguals in the process of stimultaneously acquiring their two languages.
+In line with the aforementioned models, DevLex-II predicts asymmetries between word representation from the dominant and the non-dominant language, that arise from a difference in the frequency of the input and the particular order of acquisition of the languages.
 For instance, simulations from DevLex-II result in an asymmetric cross-language priming, in which words from the dominant (acquired acquired) language primed more strongly the recognition of words in the non-dominant language (later acquired) than in the other direction [@zhao2013simulating].

### 2023-07-25 manuscript/manuscript.qmd
Update Dockerfile

@@
 The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in toddlers and adults.
-For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [for toddlers see @vonholzen2019impact; @singh2014one, for adults see @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked, but see @jardak2019labels].
-One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language, but see @duyck2009translation; @schoonbaert2009semantic].
-More balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
-These results have been taken as evidence for an asymmetry in the strength of forward and backward connections in the unbalanced bilingual lexicon Although implemented in different ways, or found under different assumptions, a dominance-mediated asymmetry is accounted for by previous models of adult bilingual lexical access, like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective into how lexical representations in a second language are formed, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
-This seems to provide a convenient account for the interaction between language dominance and cognateness found in the present study, where we found a cognate advantage only for words in their low-dominance language, and more strongly when exposure to both languages was unbalanced.
+For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [@vonholzen2019impact; @singh2014one; @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked, but see @jardak2019labels].
+One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [@finkbeiner2004role; @hoshino2010erp; @midgley2009language, but see @duyck2009translation; @schoonbaert2009semantic].
+Balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
+These results have been taken as evidence for an asymmetry in the strength of forward and backward connections in the unbalanced bilingual lexicon.
+Although implemented in different ways, or found under different assumptions, a dominance-mediated asymmetry is accounted for by previous models of adult bilingual lexical access, like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
+Overall, this provides an apparently convenient account for the interaction between language dominance and cognateness found in the present study.
 
 These models, however, are aimed at explaining results in adults, and their predictions should be taken with caution when extended to early language acquisition.
-Language dominance and proficiency are frequently defined in both populations using dimensions other than degree of exposure, which is the most common practice in infant research [@marian2021measuring; @rocha2023defining].
-For instance, the low-proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood.
+In adult bilingual populations, language dominance and proficiency are frequently defined using dimensions other than degree of exposure, which is a more common practice in infant research [@marian2021measuring; @rocha2023defining].
+For instance, low-proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood.
 We identify three critical ways in which this prevents a clear comparison between our results and those from studies on second language acuquisition in adults.
-First, in adult second language acquisition, the acquisition of the phonological system of the new language must be negotiated with the already acquired phoneme inventory of the first language [e.g., @cutler2006asymmetric; @sebastian2005influence; @sebastian2006first], which is in place around the first year of life [see @werker2015critical for review].
-Second, adults acquiring a second language already possess a system of form-meaning mappings, while simulatenous bilingual infants must build a lexicon in two languages in the absence of clear form-meaning mappings.
-Third, subjects are assumed to be literate and possess an orthographic system in place, which may shape how new words are integrated in the lexicon and processed during experimental tasks [e.g., @thierry2007brain].
+First, in adult second language acquisition, the acquisition of the phonology of the new language must be negotiated with the already acquired phoneme inventory of the first language [e.g., @cutler2006asymmetric; @sebastian2006first], in place around the first year of life [see @werker2015critical for review].
+Second, adults acquiring a second language already possess a system of form-meaning mappings, whereas simulatenous bilingual infants must build a lexicon for two languages in the absence of clear form-meaning mappings.
+Third, adults are assumed to be literate and to possess an orthographic system in place, which may shape how new words are integrated in the lexicon and processed during experimental tasks [e.g., @thierry2007brain].
 In this scenario, the acquisition of a second language may take place in a substantially different way compared to how bilingual infants acquire two languages from birth.
-A more similar case to ours is considered by the DevLex-II model [@zhao2010bilingual], which captures unique features of the early bilingual lexicon, and considers the case of bilinguals in the process of stimultaneously acquiring their two languages.
-In line with the aforementioned models, DevLex-II predicts asymmetries between word representation from the dominant and the non-dominant language, that arise from a difference in the frequency of the input and the particular order of acquisition of the languages.
-For instance, simulations from DevLex-II result in an asymmetric cross-language priming, in which words from the dominant (acquired acquired) language primed more strongly the recognition of words in the non-dominant language (later acquired) than in the other direction [@zhao2013simulating].
-In addition, the simulated lexica showed sparser connectivity between words in the non-dominant language, compared to the connectivity between words in the dominant language.
+A more similar case to the one presented in this study is considered by the DevLex-II model [@zhao2010bilingual], which captures unique features of the early bilingual lexicon, and considers the case of infants stimultaneously acquiring their two language.
+In line with the adult models, DevLex-II predicts asymmetries between word representation from the dominant and the non-dominant language.
+Simulations from DevLex-II result in an asymmetric cross-language priming, in which words from the dominant (acquired acquired) language primed more strongly the recognition of words in the non-dominant language (later acquired) than in the other direction [@zhao2013simulating].
 
+In conclusion, there is a compelling case for attributing aymmetric effects of parallel activation to differences in activation strength between forward and backward connections.
 As suggested in the introduction, it is possible that the asymmetric effect of cognateness found in the present study is simply the result of infants being exposed more frequently to words in the dominant language.

### 2023-07-25 manuscript/manuscript.qmd
Install Quarto 1.3.0 in Docker image

@@
 The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in toddlers and adults.
-For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [@vonholzen2019impact; @singh2014one; @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked, but see @jardak2019labels].
-One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [@finkbeiner2004role; @hoshino2010erp; @midgley2009language, but see @duyck2009translation; @schoonbaert2009semantic].
+For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [@vonholzen2019impact; @singh2014one; @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked; but see @jardak2019labels].
+One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language; but see @duyck2009translation; @schoonbaert2009semantic].
 Balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
@@
 Overall, this provides an apparently convenient account for the interaction between language dominance and cognateness found in the present study.
+These models are aimed at explaining results in adults, and their predictions should be taken with caution when extended to early language acquisition.
 
-These models, however, are aimed at explaining results in adults, and their predictions should be taken with caution when extended to early language acquisition.
 In adult bilingual populations, language dominance and proficiency are frequently defined using dimensions other than degree of exposure, which is a more common practice in infant research [@marian2021measuring; @rocha2023defining].
@@
 In this scenario, the acquisition of a second language may take place in a substantially different way compared to how bilingual infants acquire two languages from birth.
-A more similar case to the one presented in this study is considered by the DevLex-II model [@zhao2010bilingual], which captures unique features of the early bilingual lexicon, and considers the case of infants stimultaneously acquiring their two language.
+A more similar case to the one concerning the present study is considered by the DevLex-II model [@zhao2010bilingual], which captures unique features of the early bilingual lexicon, and considers the case of infants stimultaneously acquiring their two language.
 In line with the adult models, DevLex-II predicts asymmetries between word representation from the dominant and the non-dominant language.
@@
 
-In conclusion, there is a compelling case for attributing aymmetric effects of parallel activation to differences in activation strength between forward and backward connections.
-As suggested in the introduction, it is possible that the asymmetric effect of cognateness found in the present study is simply the result of infants being exposed more frequently to words in the dominant language.
-This would lead to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language, and therefore benefiting more strongly from their cognate status.
-This explanation does not necessarily compete with words in the dominant language activating more *strongly* their translations than *vice versa*.
+In summary, there is a compelling case for attributing aymmetric effects of parallel activation to differences in activation strength between forward and backward connections.
+It is nonethless possible that, as argued in the introduction, the asymmetric effect of cognateness found in the present study is simply the result of infants being exposed more frequently to words in the dominant language than to words in the non-dominant language.
+This would lead to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language, and therefore benefitting more strongly from their cognate status.
+These two accounts are not mutually exclusive, as words in the dominant language may active more *strongly* their translations than *vice versa*, on top of such activation being more frequent.
 Further research is needed in order to clarify this issue.

### 2023-07-25 docs/index.qmd
Update Dockerfile

@@
----
----
-
 ```{r setup}
@@
 ::: callout-tip
-Data and materials are available at [OSF](https://osf.io/hy984/), and
-code is available at
-[GitHub](https://github.com/gongcastro/trajectories).
+Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
 :::
@@
 
-![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height=100px}
+![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height="100px"}
 
-![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height=100px}
+![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height="100px"}
 
-![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height=100px}
+![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height="100px"}
 
 {{< include ../README.md >}}
+
+# Data dictionaries
+
+```{r}
+#' Create data dictionary for a dataframe and save as CSV in the docs/ folder
+#' 
+#' @param x Data frame for which to create a data dictionary
+data_dict <- function(x) {
+    if (!(is.data.frame(x) | tibble::is_tibble(x))) {
+        cli::cli_abort("x must be a dataframe or tibble")
+    }
+
+    nmissing <- function(x) sum(is.na(x))
+    nlevels <- function(x) length(unique(x))
+
+    # describe character variables
+    x_character <- x |> 
+        summarise(across(where(\(x) is.character(x) | is.factor(x)),
+                         tibble::lst(nmissing, nlevels),
+                         .names = "{.col}.{.fn}"))
+
+    # describe numeric variables
+    numeric.fns <- tibble::lst(nmissing, mean, median, sd, min, max)
+    x_numeric <- x |> 
+        summarise(across(where(is.numeric), 
+                         numeric.fns,
+                         .names = "{.col}.{.fn}"))
+
+    # join
+    out <- bind_cols(x_character, x_numeric) |> 
+        pivot_longer(everything(),
+                     names_to = c("variable", "measure"),
+                     values_to = "value",
+                     names_sep = "\\.", ) |> 
+        pivot_wider(id_cols = variable,
+                    names_from = "measure",
+                    values_from = "value",
+                    values_fn = \(x) ifelse(is.character(x),
+                                            paste0(x, collapse = ","),
+                                            unlist(x)))
+
+    return(out)
+}
+```
+
+::: {.callout-info}
+
+## Items
+
+:::
 

### 2023-07-25 manuscript/manuscript.qmd
Update Dockerfile

@@
 For instance, low-proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood.
-We identify three critical ways in which this prevents a clear comparison between our results and those from studies on second language acuquisition in adults.
+We identify three critical ways in which this prevents a clear comparison between our results and those from studies on second language acquisition in adults.
 First, in adult second language acquisition, the acquisition of the phonology of the new language must be negotiated with the already acquired phoneme inventory of the first language [e.g., @cutler2006asymmetric; @sebastian2006first], in place around the first year of life [see @werker2015critical for review].
-Second, adults acquiring a second language already possess a system of form-meaning mappings, whereas simulatenous bilingual infants must build a lexicon for two languages in the absence of clear form-meaning mappings.
+Second, adults acquiring a second language already possess a system of form-meaning mappings, whereas simultaneous bilingual infants must build a lexicon for two languages in the absence of clear form-meaning mappings.
 Third, adults are assumed to be literate and to possess an orthographic system in place, which may shape how new words are integrated in the lexicon and processed during experimental tasks [e.g., @thierry2007brain].
 In this scenario, the acquisition of a second language may take place in a substantially different way compared to how bilingual infants acquire two languages from birth.
-A more similar case to the one concerning the present study is considered by the DevLex-II model [@zhao2010bilingual], which captures unique features of the early bilingual lexicon, and considers the case of infants stimultaneously acquiring their two language.
+A more similar case to the one concerning the present study is considered by the DevLex-II model [@zhao2010bilingual], which captures unique features of the early bilingual lexicon, and considers the case of infants simultaneously acquiring their two language.
 In line with the adult models, DevLex-II predicts asymmetries between word representation from the dominant and the non-dominant language.
@@
 
-In summary, there is a compelling case for attributing aymmetric effects of parallel activation to differences in activation strength between forward and backward connections.
-It is nonethless possible that, as argued in the introduction, the asymmetric effect of cognateness found in the present study is simply the result of infants being exposed more frequently to words in the dominant language than to words in the non-dominant language.
-This would lead to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language, and therefore benefitting more strongly from their cognate status.
+In summary, there is a compelling case for attributing asymmetric effects of parallel activation to differences in activation strength between forward and backward connections.
+It is nonetheless possible that, as argued in the introduction, the asymmetric effect of cognateness found in the present study is simply the result of infants being exposed more frequently to words in the dominant language than to words in the non-dominant language.
+This would lead to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language, and therefore benefiting more strongly from their cognate status.
 These two accounts are not mutually exclusive, as words in the dominant language may active more *strongly* their translations than *vice versa*, on top of such activation being more frequent.

### 2023-07-25 docs/index.qmd
Update Dockerfile

@@
 
+    x <- flatten_columns(x)
+
     nmissing <- function(x) sum(is.na(x))
@@
 
+    # describe date variables
+    date.fns <- tibble::lst(nmissing, min, max)
+    x_date <- x |> 
+        summarise(across(where(is.Date), 
+                         date.fns,
+                         .names = "{.col}.{.fn}"))
+
     # join
-    out <- bind_cols(x_character, x_numeric) |> 
+    out <- bind_cols(x_character, x_numeric, x_date) |> 
         pivot_longer(everything(),
@@
                                             paste0(x, collapse = ","),
-                                            unlist(x)))
+                                            unlist(x))) |> 
+        mutate(varclass = purrr::map_chr(x, class)) |> 
+        relocate(varclass, .after = variable)
 
@@
 
+```{r items-dict}
+tar_load(items)
+
+items_dict <- data_dict(items) |> 
+    mutate(comments = c(
+        "Concept associated to the word-form, uniquely identifies every translation equivalent",
+        "Language to which the word-form belongs to. Can be Catalan or Spanish",
+        "Uniquely identifies the word-form in the questionnare",
+        "Phonological transcription of the word-form in the IPA format",
+        "Phonological transcription of the word-form in X-SAMPA format",
+        "Syllables in the word-form",
+        "Questionnaire sub-list in which the word-form appears",
+        "Uniquely identifies every translation equivalent",
+        "Levenshtein distance between the word-form and its translation",
+        "Number of phonemes in the word-form",
+        "Number of syllables in the word-form",
+        "Lexical frequency in Zipf scores, from the English CHILDES corpora",
+        "Sum of the frequency of the syllables in the word-form, expressed as counts per million tokens"
+    ))
+options(knitr.kable.NA = '--')
+
+kable(items_dict,
+      col.names = c(
+          "Variable",
+          "Type",
+          "# missing",
+          "# levels",
+          "Mean",
+          "Median",
+          "SD",
+          "Min",
+          "Max",
+          "Description"
+      ))
+```
+
+## Participants
+
+```{r participants-dict}
+tar_load(participants)
+
+participants_dict <- data_dict(participants) |> 
+    mutate(comments = c(
+        "Concept associated to the word-form, uniquely identifies every translation equivalent",
+        "Language to which the word-form belongs to. Can be Catalan or Spanish",
+        "Uniquely identifies the word-form in the questionnare",
+        "Phonological transcription of the word-form in the IPA format",
+        "Phonological transcription of the word-form in X-SAMPA format",
+        "Syllables in the word-form",
+        "Questionnaire sub-list in which the word-form appears",
+        "Uniquely identifies every translation equivalent",
+        "Levenshtein distance between the word-form and its translation",
+        "Number of phonemes in the word-form",
+        "Number of syllables in the word-form",
+        "Lexical frequency in Zipf scores, from the English CHILDES corpora",
+        "Sum of the frequency of the syllables in the word-form, expressed as counts per million tokens"
+    ))
+options(knitr.kable.NA = '--')
+
+kable(items_dict,
+      col.names = c(
+          "Variable",
+          "Type",
+          "# missing",
+          "# levels",
+          "Mean",
+          "Median",
+          "SD",
+          "Min",
+          "Max",
+          "Description"
+      ))
 :::

### 2023-07-25 docs/index.qmd
Update Dockerfile

--- 2023-07-23 docs/index.qmd
@@
 
-![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height=100px}
+![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height="100px"}
 
-![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height=100px}
+![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height="100px"}
 
-![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height=100px}
+![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height="100px"}
 
 {{< include ../README.md >}}
+
+# Data dictionaries
+
+::: {.callout-info}
+
+## Items
+
+This data frame contains information about the word-forms included in the questionnaire, together with some identifiers used to relate participants' responses to the information of the word-forms they responded to.
+
+```{r tbl-items}
+items |> 
+    head(15) |> 
+    glimpse()
+```
+
+* `te`: integer that uniquely labels a translation equivalent, and is only repeated across the word-forms from Catalan and Spanish that are part of the same translation equivalent
+* `meaning`: character string that uniquely labels the concept associated to the word-form, and is only repeated across the word-forms from Catalan and Spanish that are part of the same translation equivalent
+* `language`: character string indicating the language (Catalan or Spanish) to which the word-form belongs to 
+* `item`: character string that uniquely identifies the word-form in the questionnaire, and links it to the formr item
+* `ipa`: phonological transcription of the word-form in IPA format, generated from the X-SAMPA transcription of the word-form using the `ipa::xsampa()` function.
+* `xsampa`: phonological transcription of the word-form in X-SAMPA format
+* `lv`: numeric value indicating the normalised inverse of the Levenshtein distance between the X-SAMPA phonological transcriptions of the word-form and of its translation equivalent, calculated using the `stringdist::stringsim()` (see the Methods section in the main manuscript for more details)
+* `n_phon`: integer indicating the number of phonemes included in the X-SAMPA phonological transcription of the word-form
+* `n_syll`: integer indicating the number of syllables included in the X-SAMPA phonological transcription of the word-form
+* `syll`: list of character strings in which each element is a syllable included in the word-form
+* `freq`: numeric values indicating the lexical frequency in Zipf scores, from the English CHILDES corpora
+* `freq_syll`: numeric value indicating the um of the frequency of the syllables in the word-form, expressed as counts per million tokens
+* `list`: characters string indicating the questionnaire sub-list(s) in which the word-form appears
+
+## Participants
+
+This data frame contains demographic and linguistic information about participants participants, together with some identifiers used to relate participants' responses to their corresponding information.
+
+```{r tbl-items}
+participants |> 
+    head(15) |> 
+    glimpse()
+```
+
+* `id`: integer that uniquely labels participant, and is only repeated across responses to the questionnaire from the same participant
+* `time`: integer indicating the cumulative number of times the participant has provided a valid response to the questionnaire
+* `time_stamp`: date at which the response to the questionnaire was recorded (last item responded)
+* `list`: character string indicating the questionnaire sub-list to which the participant responded, which is virtually always the same for the same participant
+* `age`: numeric value indicating the age of the participant when their response to the questionnaire was recorded, calculated as the difference in months between such date and the birth date of the participant
+* `lp`: character string indicating the language profile of the participant (Monolingual or Bilingual), calculated from `doe_catalan` and `doe_spanish` (`"Monolingual"` if >=80% DoE to Catalan or Spanish, `"Bilingual"` otherwise)
+* `doe_catalan`: numeric value indicating participant's degree of exposure (DoE) to Catalan, as reported by their caregivers
+* `doe_spanish`: numeric value indicating participant's degree of exposure (DoE) to Spanish, as reported by their caregivers
+* `edu_parent`: factor indicating the caregivers maximum educational attainment
+
+## Responses
+
+This data frame is the one used to fit the main model, and the model included in Appendix A.
+It contains participants' responses to each item included in the sub-list of the questionnaire they responded to, together with participant- and word-level predictors of interest.
+
+```{r tbl-items}
+responses |> 
+    head(15) |> 
+    glimpse()
+```
+
+* `id`: integer that uniquely labels participant, and is only repeated across responses to the questionnaire from the same participant
+* `time`: integer indicating the cumulative number of times the participant has provided a valid response to the questionnaire
+* `age`: numeric value indicating the age of the participant when their response to the questionnaire was recorded, calculated as the difference in months between such date and the birth date of the participant
+* `age_std`: numeric value indicating the participant's standardised `age`
+* `te`: integer that uniquely labels a translation equivalent, and is only repeated across the word-forms from Catalan and Spanish that are part of the same translation equivalent
+* `language`: character string indicating the language (Catalan or Spanish) to which the word-form belongs to 
+* `meaning`: character string that uniquely labels the concept associated to the word-form, and is only repeated across the word-forms from Catalan and Spanish that are part of the same translation equivalent
+* `item`: character string that uniquely identifies the word-form in the questionnaire, and links it to the formr item
+* `response`: ordered factor indicating participant's response to the item, which can take `"No"`, `"Understands"`, or `"Understands and Says"` as values
+* `lv`: numeric value indicating the normalised inverse of the Levenshtein distance between the X-SAMPA phonological transcriptions of the word-form and of its translation equivalent, calculated using the `stringdist::stringsim()` (see the Methods section in the main manuscript for more details)
+* `lv_std`: numeric value indicating the word-form's standardised `lv`
+* `freq`: numeric values indicating the lexical frequency in Zipf scores, from the English CHILDES corpora
+* `freq_std`: numeric value indicating the word-form's standardised `freq`
+* `n_phon`: integer indicating the number of phonemes included in the X-SAMPA phonological transcription of the word-form
+* `n_phon_std`: numeric value indicating the word-form's standardised `n_phon`
+* `doe`: numeric value indicating participant's degree of exposure (DoE) to the `language` the `item` belongs to
+* `doe_std`: numeric value indicating the participant's standardised `doe` to the `language` the `item` belongs to
+* `exposure`: numeric value indicating the participants' exposure score to the word-form, calculated as the product of `freq` and `doe`
+* `doe_std`: numeric value indicating the participant's standardised `exposure` score for the `item` and its `freq` (see the Methods section in the main Manuscript for more details)
 

### 2023-07-26 docs/index.qmd
Update Dockerfile

@@
 # load objects -----------------------------------------------------------------
-targets::tar_config_set(store = here::here("_targets"),
-                        script = here::here("_targets.R"))
+targets::tar_config_set(
+    store = here::here("_targets"),
+    script = here::here("_targets.R")
+)
 
@@
 thanks <- yaml_metadata$thanks
+
+targets::tar_load(items)
+targets::tar_load(participants)
+targets::tar_load(responses)
+
+library(dplyr)
+library(tidyr)
+library(knitr)
+
 ```
@@
 
-```{r tbl-items}
+```{r tbl-participants}
 participants |>
@@
 
-```{r tbl-items}
+```{r tbl-responses}
 responses |>

### 2023-07-26 docs/index.qmd
Update Dockerfile

@@
+# Cognate beginnings to bilingual lexical acquisition
+
 ```{r setup}
@@
 #| warning: false
+#| echo: false
 # load objects -----------------------------------------------------------------
@@
 
+## Abstract
+
 `r abstract`
 
+## Acknowledgements
+
 `r thanks`
-
-# Links
-
-![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height="100px"}
-
-![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height="100px"}
-
-![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height="100px"}
 

### 2023-07-26 docs/index.qmd
Update Dockerfile

@@
-# Cognate beginnings to bilingual lexical acquisition
+---
+title: "Cognate beginnings to bilingual lexical acquisition"
+subtitle: "Lab notes"
+short-title: "Cognate beginnings"
+author:
+  - name: Gonzalo Garcia-Castro
+    corresponding: true
+    orcid: 0000-0002-8553-4209
+    email: gonzalo.garciadecastro@upf.edu
+    affiliations:
+      - id: cbc
+        name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+        address: Ramon Trias Fargas 25-27
+        city: Barcelona
+        region: Spain
+        postal-code: "08005"
+  - name: Daniela S. Avila-Varela
+    orcid: 0000-0002-3518-8117
+    affiliations: 
+      - ref: cbc
+      - id: ul
+        name: Universidade de Lisboa
+        department: Facultade de Letras
+        address: Alameda da Universidade
+        city: Lisboa
+        region: Portugal
+        postal-code: "1600-214"
+  - name: Ignacio Castillejo
+    orcid: 0000-0001-7445-0416
+    affiliations: 
+      - id: uam 
+        name: Universidad Autónoma de Madrid
+        department: Departamento de Psicología
+        address: Iván Pavlov, 6
+        city: Madrid
+        region: Spain
+        postal-code: "28049"
+  - name: Nuria Sebastian-Galles
+    orcid: 0000-0001-6938-2498
+    affiliations: 
+      - ref: cbc
+date: "May 10th, 2023"
+code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
+keywords:
+  - lexical acquisition
+  - vocabulary
+  - bilingualism
+  - item response theory
+  - bayesian
+link-citations: true
+editor: source
+code-fold: true
+warning: false
+error: false
+format:
+  html: default
+toc: true
+fig-dpi: 1000
+---
 

### 2023-07-26 docs/index.qmd
Update Dockerfile

--- 2023-07-26 docs/index.qmd
@@
-# Cognate beginnings to bilingual lexical acquisition
+---
+title: "Cognate beginnings to bilingual lexical acquisition"
+subtitle: "Lab notes"
+short-title: "Cognate beginnings"
+author:
+  - name: Gonzalo Garcia-Castro
+    corresponding: true
+    orcid: 0000-0002-8553-4209
+    email: gonzalo.garciadecastro@upf.edu
+    affiliations:
+      - id: cbc
+        name: Universitat Pompeu Fabra
+        department: Center for Brain and Cognition
+        address: Ramon Trias Fargas 25-27
+        city: Barcelona
+        region: Spain
+        postal-code: "08005"
+  - name: Daniela S. Avila-Varela
+    orcid: 0000-0002-3518-8117
+    affiliations: 
+      - ref: cbc
+      - id: ul
+        name: Universidade de Lisboa
+        department: Facultade de Letras
+        address: Alameda da Universidade
+        city: Lisboa
+        region: Portugal
+        postal-code: "1600-214"
+  - name: Ignacio Castillejo
+    orcid: 0000-0001-7445-0416
+    affiliations: 
+      - id: uam 
+        name: Universidad Autónoma de Madrid
+        department: Departamento de Psicología
+        address: Iván Pavlov, 6
+        city: Madrid
+        region: Spain
+        postal-code: "28049"
+  - name: Nuria Sebastian-Galles
+    orcid: 0000-0001-6938-2498
+    affiliations: 
+      - ref: cbc
+date: "last-updated"
+code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
+keywords:
+  - lexical acquisition
+  - vocabulary
+  - bilingualism
+  - item response theory
+  - bayesian
+link-citations: true
+editor: source
+code-fold: true
+warning: false
+error: false
+format:
+  html: default
+toc: true
+fig-dpi: 1000
+---
 
@@
 ::: callout-tip
-Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
+Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories), and [Docker](https://hub.docker.com/r/gongcastro/cognate-beginnings).
 :::
@@
 
-::: {.callout-info}
+::: {.tabset-panel}
 
@@
 
-```{r tbl-items}
+```{r items}
 items |>
-    head(15) |>
-    glimpse()
+    head(15)
 ```
@@
 
-```{r tbl-participants}
+```{r participants}
 participants |>
-    head(15) |>
-    glimpse()
+    head(15)
 ```
@@
 
-```{r tbl-responses}
+```{r responses}
 responses |>
-    head(15) |>
-    glimpse()
+    head(15)
 ```
@@
 
+:::
+
 # Session info

### 2023-07-26 docs/index.qmd
Update Dockerfile

@@
       - ref: cbc
-date: "last-updated"
+date: last-updated
 code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
@@
 format:
-  html: default
+  html:
+    theme: _assets/style.scss
+
 toc: true

### 2023-07-26 _targets.R
Create website

@@ -34,58 +34,66 @@ invisible({
 
 # define global options --------------------------------------------------------
 
-options(repos = c(Stan = "https://mc-stan.org/r-packages/",
-                  gongcastro = "https://gongcastro.r-universe.dev", 
-                  CRAN = "https://cloud.r-project.org"),
-        mc.cores = parallel::detectCores(),
-        brms.backend = "cmdstanr",
-        brms.file_refit = "on_change",
-        tidyverse.quiet = TRUE,
-        knitr.duplicate.label = "allow",
-        loo.cores = 1,
-        knitr.graphics.error = FALSE)
+options(
+    repos = c(
+        Stan = "https://mc-stan.org/r-packages/",
+        gongcastro = "https://gongcastro.r-universe.dev",
+        CRAN = "https://cloud.r-project.org"
+    ),
+    mc.cores = parallel::detectCores(),
+    brms.backend = "cmdstanr",
+    brms.file_refit = "on_change",
+    tidyverse.quiet = TRUE,
+    knitr.duplicate.label = "allow",
+    loo.cores = 1,
+    knitr.graphics.error = FALSE
+)
 
 list(
     ## import data -------------------------------------------------------------
-    
+
     tar_target(bvq_data_file, "data-raw/bvq.rds", format = "rds"),
     tar_target(bvq_data, readRDS(bvq_data_file)),
-    
+
     # get CHILDES frequencies
     tar_target(childes, get_childes_frequencies(age_range = c(12, 32))),
-    
+
     # items
-    tar_target(items, get_items(bvq_data = bvq_data, 
-                                childes = childes)),
+    tar_target(items, get_items(
+        bvq_data = bvq_data,
+        childes = childes
+    )),
     tar_target(items_test, test_items(items)),
-    
+
     # participants
-    tar_target(participants,
-               get_participants(bvq_data,
-                                age = c(12, 32),
-                                lp = c("Monolingual", "Bilingual"),
-                                other_threshold = 0.1)),
-    
+    tar_target(
+        participants,
+        get_participants(bvq_data,
+            age = c(12, 32),
+            lp = c("Monolingual", "Bilingual"),
+            other_threshold = 0.1
+        )
+    ),
     tar_target(participants_test, test_participants(participants)),
-    
+
     # responses
     tar_target(responses, get_responses(bvq_data, items, participants)),
-    
     tar_target(responses_test, test_responses(responses)),
-    
+
[…]

### 2023-07-26 _data-dictionary.qmd
Built site for gh-pages

--- 2023-07-26 docs/index.qmd
@@
 ---
-title: "Cognate beginnings to bilingual lexical acquisition"
-subtitle: "Lab notes"
-short-title: "Cognate beginnings"
-author:
-  - name: Gonzalo Garcia-Castro
-    corresponding: true
-    orcid: 0000-0002-8553-4209
-    email: gonzalo.garciadecastro@upf.edu
-    affiliations:
-      - id: cbc
-        name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-        address: Ramon Trias Fargas 25-27
-        city: Barcelona
-        region: Spain
-        postal-code: "08005"
-  - name: Daniela S. Avila-Varela
-    orcid: 0000-0002-3518-8117
-    affiliations: 
-      - ref: cbc
-      - id: ul
-        name: Universidade de Lisboa
-        department: Facultade de Letras
-        address: Alameda da Universidade
-        city: Lisboa
-        region: Portugal
-        postal-code: "1600-214"
-  - name: Ignacio Castillejo
-    orcid: 0000-0001-7445-0416
-    affiliations: 
-      - id: uam 
-        name: Universidad Autónoma de Madrid
-        department: Departamento de Psicología
-        address: Iván Pavlov, 6
-        city: Madrid
-        region: Spain
-        postal-code: "28049"
-  - name: Nuria Sebastian-Galles
-    orcid: 0000-0001-6938-2498
-    affiliations: 
-      - ref: cbc
-date: "May 10th, 2023"
-code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
-keywords:
-  - lexical acquisition
-  - vocabulary
-  - bilingualism
-  - item response theory
-  - bayesian
-link-citations: true
-editor: source
-code-fold: true
-warning: false
-error: false
-format:
-  html: default
-toc: true
-fig-dpi: 1000
+title: Data dictionary
 ---
@@
 
-::: callout-tip
-Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
-:::
-
-## Abstract
-
-`r abstract`
-
-## Acknowledgements
-
-`r thanks`
-
-{{< include ../README.md >}}
-
-# Data dictionaries
-
-::: {.callout-info}
+::: {.panel-tabset}
 
@@
 
-```{r tbl-items}
+```{r items}
 items |>
-    head(15) |>
-    glimpse()
+    head(10) |>
+    knitr::kable(digits = 2)
 ```
@@
 
-```{r tbl-participants}
+```{r participants}
 participants |>
-    head(15) |>
-    glimpse()
+    head(10) |>
+    knitr::kable(digits = 2)
 ```
@@
 
-```{r tbl-responses}
+```{r responses}
 responses |>
-    head(15) |>
-    glimpse()
+    head(10) |>
+    knitr::kable(digits = 2)
 ```
@@
 
-# Session info
-
-```{r}
-sessionInfo()
-```
+:::

### 2023-07-26 _repro.qmd
Built site for gh-pages

---
---

We recommend two ways of reproducing or inspecting the code of this repository. One option is to clone the [Docker image](https://hub.docker.com/r/gongcastro/cognate-beginnings) we have pused to Dockerhub. A Docker image is just a virtual machine that contains all dependencies, code, and files necessary to run the code. Everything is already installed and setup, you will only have to open [https://localhost:8787](https://localhost:8787) in your browser, and a RStudio session will open for you. This option is more robust (dependencies are difficult to get right when using your own computer), but it also requires Docker to be installed.

A second option is the more traditional approach of downloading the repository (or cloning it, if you know your way through Git/GitHub), installing the necessary dependencies (we recommend using renv for R dependencies, other system dependencies are listed below), and running the code (we recommend using the targets workflow).

### 2023-07-26 docs/_repro-docker.qmd
Update README

---
title: Docker {{< fa brands docker >}}
---

1. Download and install Docker
2. Pull the gongcastro/cognate-beginnings image. Two ways of doing this:
    a) Using the Docker UI: go the the "Images" tab, click on "Search images to run", search for "gongcastro/cognate-beginnings", and click "Pull".
    b) Using the Docker CLI: open your console/terminal/command prompt, and run:
    ```bash 
    docker pull gongcastro/cognate-beginnings
    ```
3. Run the image in a Docker container.  Two ways of doing this:
    a) Using the Docker UI: in the "Images" tab, look for the "gongcastro/cognate-beginnings" image and click "Run".
    b) Using the Docker CLI:
    ```bash 
    docker run --rm -ti -e ROOT=true -e PASSWORD=rstudio -p 8787:8787 --name rstudio gongcastro/cognate-beginnings:latest
    ```

4. Open [https://localhost:8787](https://localhost:8787) in your browser. Log in using `rstudio` as users, and `rstudio` as password. An RStudio session should open, with all files, code and dependencies installed and ready. Check `targets::tar_visnetwork()` to examine the targets workflow. This will list the objects created by the R code, and how they relate to each other. To retrieve and explore a given object, run `targets::tar_load(target_name)`, where `target_name` is the name of the rarget you want to explore.

> 💡 The RStudio session opened by Docker does not have root permissions (you will not be able to run `targets::tar_make()` to run the code from the R console). Instead, open the "Terminal" tab, and run `sudo Rscript -e "targets::tar_make()"`. The targets workflow should now be triggered.

> 💡 The targets workflow is run to completion before your Rstudio session is open in Docker. This means that even if you trigger `targets::tar_make()`, virtually all targets will be skipped, as they are completed. If you want to run a specific target, delete it first (or any other target it depends on), and the targets workflow will be run, now generating all deleted targets again. Be aware that this might take some time, especially when fitting the models anew.

### 2023-07-26 docs/_repro-github.qmd
Update README

---
title: GitHub {{< fa brands github >}}
---

## Clone the repository and run the scripts

Follow these steps (see below for details):

0.  Download this repository
1.  Install software dependencies
2.  Install package dependencies with renv
3.  Running the code with targets
4.  Repository structure and files

### 0. Download this repository ⬇️

First, you will need to download this repository to your local machine.
We suggest three different ways:

a)  **Git terminal**: download and install
    [Git](https://git-scm.com/downloads) with default settings and clone
    this repository locally running the following command in your
    console:

``` bash
git clone https://github.com/gongcastro/cognate-beginnings.git
```

> 💡 You don't need to run this command from the Git console. Just open
> Command Prompt or Power Shell in Windows (just make sure that Git is
> included in your [PATH
> variables](https://www.delftstack.com/howto/git/add-git-to-path-on-windows/),
> which should be already done after the recommended Git installation),
> Terminal in MacOS, or command line in Linux and run this command.

b)  **GitHub releases**: download the latest
    [release](https://github.com/gongcastro/cognate-beginnings/releases)
    of the repository.

c)  **Direct download**: click the [green
    button](https://github.com/gongcastro/cognate-beginnings/archive/refs/heads/main.zip)
    "Code" in this page (upper-right corner), click "Download ZIP", and
    unzipping the downloaded repository.

### 1. Install software dependencies 💻

Here's a list of programs that you might need to install in your machine
to make the project work:

-   [R](https://www.r-project.org/) (4.2.2 or greater)

-   [RStudio](https://posit.co/download/rstudio-desktop/): although not
    strictly necessary, we encourage the use of Rstudio for reproducing
    the present project. We have not tested the reproducibility of the
    project in other IDEs.

-   [Quarto](https://quarto.org/docs/get-started/) (1.3.340 or greater):
    we use this software to generate the manuscript and lab notes. We
    recommend installing version 1.3.340 or higher, although previous
    version might work as well.

-   [Rtools](https://cran.r-project.org/bin/windows/Rtools/) (4.3 or
    greater).

-   [CmdStan](https://mc-stan.org/users/interfaces/cmdstan) (2.31.0 or
    greater): we use the CmdStan backend in brms to fit our Bayesian
    models. We recommend installing CmdStan using its R interface
    [CmdStanR](https://mc-stan.org/cmdstanr/) following the [*Getting
    started with
    CmdStanR*](https://mc-stan.org/cmdstanr/articles/cmdstanr.html)
    vignette. You can also install it following the [CmdStan
    installation guide](https://mc-stan.org/users/interfaces/), and then
    letting CmdStanR know the path to the folder using
    `cmdstanr::set_cmdstan_path()`.

    <div>

    > 💡 **Installing CmdStan** is sometimes inconvenient, and might
    > require some time to set up, specially if it's your first time
    > doing so. Have patience and follow the CmdStan installation guide
    > closely. If you encounter problems and just don't have the time,
    > you can download the models from the [Open Science Framework
    > repository](https://osf.io/hy984/) and skip the model-fitting
    > steps.

    </div>

### 2. Install package dependencies with renv 📦

Open the downloaded repository folder and click on the
[cognate-beginnings.Rproj](cognate-beginnings.Rproj) to open an RStudio
session (recommended) or just open an R session and set the working
directory to the downloaded folder. Once open run:

``` r
install.packages("renv") # in case you need to install renv
renv::restore()
```

This might take a couple of minutes, depending on the number of packages
that need installation or an update.

> 💡 **What does this step do?** The code in this repository needs some
> packages to run, many of which might not have been already installed
> in your local machine. Instead of having to install them yourself one
> by one or updating already installed ones (which might change how they
> behave, possibly breaking some of your code in other projects) this
> repository uses the R package
> [renv](https://rstudio.github.io/renv/articles/renv.html) to deal with
> package libraries. The command `renv::restore()` will install all
> necessary packages in the appropriate version (listed in the file
> [renv.lock](renv.lock) in a self-contained R package library. This
> process will not affect the packages that you had already installed in
> your machine.
>
> 💡 **Something went wrong with renv. Now what?** Although renv
> facilitates the reproducibility of the repository, it is not
> guaranteed that things will work perfectly. If you encounter trouble
> installing the packages using renv, try installing them individually
> using (`install.packages()` or `renv::install()`) and try fixing
> whatever issues arise (they might depend on your particular setup).
> Some packages are used by RStudio itself, and might refiuse to update
> them while you're using RStudio. Try installing them from the R
> console outside of RStudio. Open an
> [issue](https://github.com/gongcastro/cognate-beginnings/issues) on
> this repository if you need further assistance! We'll get back to you
> ASAP. :smile:

### 3. Running the code with targets 🚀

Once the package dependencies have been solved with renv, run the
following command:

``` r
targets::tar_make()
```

Mind that this process might take some time. Refitting the brms models
might take very long (days even, depending on your set-up). If you want
to skip this step, download the "results" folder in the [Open Science
Framework repository](https://osf.io/hy984/) and replace it in the
repository.

> 💡**What does this step do?** This repository's workflow is based on
> [targets](https://books.ropensci.org/targets/). This means that the
> code is run in the appropriate order according to its internal
> dependencies. The code is organised as *targets*, defined in the
> `_targets.R` file. Sometimes targets operate with the outcomes of
> other targets. For instance, one target might run a function that
> takes the outcome of a different target as argument. {targets} makes
> sure that the functions and objects needed to run each target have
> already been previously defined. You can visually explore this
> repository's targets and its dependency structure by running
> `targets::tar_visnetwork()`. You can explore the contents of the
> executed targets by running `targets::tar_load_everything()`, which
> will load all defined targets into your workspace.

### 2023-07-27 index.qmd
Update website

--- 2022-12-01 index.qmd
@@
     affiliations: 
-      - name: Universidad Atónoma de Madrid
+      - id: uam 
+        name: Universidad Autónoma de Madrid
         department: Departamento de Psicología
+        address: Iván Pavlov, 6
+        city: Madrid
+        region: Spain
+        postal-code: "28049"
   - name: Nuria Sebastian-Galles
@@
     affiliations: 
-      - name: Universitat Pompeu Fabra
-        department: Center for Brain and Cognition
-abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies.
-Floccia et al.
-(2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch).
-The mechanisms underlying this effect remain unknown.
-We explore two compatible scenarios.
-First, we test whether cognates are learnt earlier than non-cognates.
-This would account for the difference in vocabulary size associated to the amount of shared cognates across languages.
-Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap.
-This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals.
-We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
+      - ref: cbc
+keywords:
+  - lexical acquisition
+  - vocabulary
+  - bilingualism
+  - item response theory
+  - bayesian
+link-citations: true
+editor: source
+warning: false
+error: false
 format:
   html:
+    theme: _assets/style.scss
+    page-layout: article
+    code-fold: true
+    smooth-scroll: true
 toc: true
-        link-external-newwindow: true
-        number-sections: true
-        number-offset: 0
-        code-fold: true
-        code-overflow: scroll
-        code-line-numbers: true
-        code-copy: hover
 fig-dpi: 1000
-        reference-location: margin
-        execute-dir: "C:/Users/U155880/Documents/trajectories/"
-    pdf:
-        toc: true
-        number-sections: true
-        colorlinks: true
-        fig-dpi: 1000
-bibliography: "references.bib"
-csl: "manuscript/apa6.csl"
+sidebar: true
 ---
 
-```{r}
+```{r setup}
 #| label: setup
-#| include: false
-tar_load_globals() # load packages and options
-tar_load_everything() # load targets as R objects
-
-# resolve namespace conflicts
-conflict_prefer("filter", "dplyr")
-conflict_prefer("select", "dplyr")
-
-# set ggplot theme and colour palette
-theme_set(theme_custom()) # set custom ggplot theme
-clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
-
-options(
-    knitr.kable.NA = "-",
-    knitr.duplicate.label = "allow",
-    ggplot2.ordinal.fill = clrs,
-    ggplot2.ordinal.colour = clrs,
-    ggplot2.discrete.fill = clrs,
-    ggplot2.discrete.colour = clrs,
-    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
-    ggplot2.continuous.colour = ggplot2::scale_color_gradient
-)
-```
-
-::: {.callout-tip}
-Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
-:::
-
-# Methods
-
-Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
-
-## Questionnaires
-
-```{r}
-#| label: questionnaire
-#| echo: true
 #| message: false
 #| warning: false
-data("pool", package = "bvqdev")
-n_categories <- distinct(pool, semantic_category) %>% nrow()
-n_items <- length(unique(pool$item))
-n_item_language <- count(pool, language) %>% 
-    pull(n) %>%
-    set_names(c("catalan", "spanish"))
-```
-
-The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted of two lists of words: one in Catalan and one in Spanish.
-The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]` items.
-Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* \[dog\] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping.
-When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* \[*to finish*\] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* \[*monkey*\] and Catalan *mono/mico*.
-We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1).
-For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.
-
-For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\~400 translation equivalents, \~400 items in Catalan, \~400 items in Spanish).
-These last families were randomly assigned to one of four different subsets of the complete list of items.
-These lists were designed so that each contained a representative subsample of the items from the complete list.
-Semantic/functional categories with less than 16 items were not divided in the short version of the questionnaire to preserve a minimum of four items per list for each category: all of their items were included in the four lists.
-Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.
-Table 2 in Appendix 1 shows the distribution of items across questionnaire versions.
-We excluded from the analysis multi-word items (e.g., *barrita de cereales* \[cereal bar\]) and items that included more than one word-form (e.g., *mono / mico*).
-Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.
-
-## Participants
-
-```{r}
-#| label: fig-participants-time
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 10 
-#| fig-cap: "Date at which responses were collected"
-participants %>% 
-    group_by(time_stamp) %>% 
-    summarise(n = n(), .groups = "drop") %>% 
-    mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n)) +
-    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
-             angle = 90, vjust = 1.5, hjust = 1) +
-    geom_line(size = 1, colour = clrs[1]) +
-    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
-    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 20, hjust = 1),
-        axis.title.x = element_blank(),
-    ) +
-
-    participants %>% 
-    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
-    group_by(time_stamp, age) %>% 
-    summarise(n = n(), .groups = "drop") %>% 
-    group_by(age) %>% 
-    mutate(n = cumsum(n)) %>% 
-    ggplot(aes(time_stamp, n, colour = age)) +
-    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
-    geom_line(size = 1) +
-    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
-    scale_y_continuous(breaks = seq(0, 500, 50)) +
-    scale_colour_manual(values = clrs) +
-
-    plot_layout(ncol = 1) &
-
-    scale_x_date(date_breaks = "3 month", date_labels = "%B %Y") &
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 20, hjust = 1),
-        axis.title.x = element_blank(),
-    )
-```
-
-### Age
-
-Number of months elapsed between participants' birth date and questionnaire completion.
-
-```{r}
-#| label: fig-participants-age
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Distribution of participants' ages"
-participants %>% 
-    mutate(age = floor(age)) %>% 
-    count(age) %>% 
-    ggplot() +
-    aes(x = age, y = n) +
-    geom_col(fill = clrs[1]) +
-    geom_text(aes(label = n), size = 3.5, vjust = -1) +
-    labs(
-        x = "Age (months)",
-        y = "# participants"
-    ) +
-    scale_x_continuous(breaks = seq(0, 40, 2)) +
-    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
-    theme(
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
-
-### Language exposure (Degree of Exposure)
-
-Participant degree of exposure to the words' language ($DoE$): percentage of exposure to the language the word belonged to.
-For example, for a participant with 90% exposure to Catalan, and 10% to Spanish, the DoE of the Catalan word *taula* would be 90%, and the DoE for the Spanish word *mesa* would be 10%.
-
-```{r}
-#| label: fig-participants-lp
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
-participants %>% 
-    pivot_longer(
-        c(doe_catalan, doe_spanish),
-        names_to = "language", values_to = "doe", 
-        names_transform = ~str_to_sentence(str_remove(., "doe_"))
-    ) %>% 
-    mutate(
-        doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
-        id = paste0(id, " (", time, ")")
-    ) %>% 
-    count(language, doe) %>% 
-    ggplot(aes(doe, y = n, label = n, fill = language)) +
-    geom_col(colour = "white", size = 1) +
-    geom_text(position = position_nudge(y = 5)) +
-    facet_wrap(~language, ncol = 1) + 
-    labs(
-        x = "Degree of exposure to Catalan",
-        y = "Number of participants",
-        colour = "Language",
-        fill = "Language"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    theme(
-        legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor = element_blank()
-    )
-```
-
-### SES/parental education
-
-Maximum educational attainment of the participant parent(s).
-
-```{r}
-#| label: tbl-participants-edu
-#| tbl-cap: Participant exposure to the language (*DoE*) and socio-economic status (*SES*)
-#| echo: true
-#| message: false
-#| warning: false
-participants %>%
-    left_join(bvq_data$logs) %>% 
-    mutate(
-        edu_parent = fct_explicit_na(as.factor(edu_parent)),
-        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE)
-    ) %>% 
-    count(doe_catalan, edu_parent) %>% 
-    right_join(
-        expand_grid(
-            doe_catalan = levels(.$doe_catalan),
-            edu_parent = unique(bvq_data$logs$edu_parent)
-        )
-    ) %>% 
-    rename(total = n) %>% 
-    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
-    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
-    rename(Total = total) %>% 
-    select(-"NA") %>% 
-    gt() %>% 
-    summary_rows(
-        columns = Complementary:Primary, 
-        fns = list(Total = ~sum(., na.rm = TRUE)),
-        formatter = fmt_integer
-    ) %>% 
-    cols_label(doe_catalan = "DoE Catalan") %>%
-    sub_missing(columns = 2:8, missing_text = "--") %>% 
-    tab_spanner("Educational attainment", Secondary:Primary) %>% 
-    tab_style(
-        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
-        locations = cells_body(columns = 9, rows = 1:5)
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(columns = 2:7)
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = c(1, 8))
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(spanners = "Educational attainment")
-    )
-```
-
-## Items
-
-```{r}
-#| label: tbl-items-summary
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| tbl-cap: "Lexical frequencies.
-Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
-items <- left_join(items, select(bvqdev::pool, te, item, language, class, semantic_category))
-
-items %>%
-    unnest(list) %>% 
-    group_by(language, list) %>% 
-    summarise_at(
-        vars(freq_zipf, n_phon, lv), 
-        list(
-            n = ~sum(!is.na(.)),
-            mean = ~mean(., na.rm = TRUE),
-            sd = ~sd(., na.rm = TRUE),
-            min = ~min(., na.rm = TRUE),
-            max = ~max(., na.rm = TRUE)
-        )
-    ) %>% 
-    select(-c(n_phon_n, lv_n)) %>% 
-    rename(n = freq_zipf_n) %>% 
-    gt(groupname_col = "language", rowname_col = "list") %>% 
-    fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>% 
-    fmt_percent(matches("lv_"), decimals = 1) %>% 
-    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
-    tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>% 
-    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
-    cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>% 
-    cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>% 
-    cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>% 
-    cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>% 
-    cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>% 
-    cols_merge_range(col_begin = lv_min, col_end = lv_max) %>% 
-    cols_label(
-        list = "",
-        n = md("*N*"),
-        freq_zipf_mean = md("Mean ± SD"),
-        freq_zipf_min = md("Range"),
-        n_phon_mean = md("Mean ± SD"),
-        n_phon_min = md("Range"),
-        lv_mean = md("Mean ± SD"),
-        lv_min = md("Range")
-    ) %>% 
-    tab_style(
-        cell_text(style = "italic"),
-        cells_column_labels(everything())
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_spanners(everything())
-    ) %>% 
-    summary_rows(
-        fns = list("Grand mean" = ~mean(.)),
-        columns = ends_with("_mean")
-    ) %>% 
-    summary_rows(
-        fns = list("N" = ~sum(.)),
-        columns = n,
-        formatter = fmt_integer
-    )
-```
-
-### Lexical frequency
-
-Word lexical frequency ($Frequency$): lexical frequency of the word in its corresponding language, expressed as Zipf scores [@van2014subtlex; @zipf1949human].
-This variable ranges from 0 to 7, and follows and approximates a normal distribution, with most values in a corpus ranging from 3 to 5 points.
-Lexical frequencies were extracted from the English corpora of the [CHILDES](https://childes.talkbank.org/) database using the [{childesr}](https://langcog.github.io/childes-db-website/) R package.
-Available words in the corpora were then mapped to their Catalan and Spanish translations.
-Responses to words with missing lexical frequency score were excluded from analyses of the total number of items).
-
-```{r}
-#| label: fig-items-freq
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Distribution of lexical frequency across items"
-items %>% 
-    ggplot() +
-    aes(
-        x = language,
-        y = freq_zipf,
-        fill = language,
-        colour = language
-    ) +
-    stat_slab(slab_colour = "white", slab_size = 1, alpha = 0.5, scale = 0.5) +
-    stat_dots(side = "left", scale = 0.5, position = position_nudge(x = -0.05)) + 
-    stat_pointinterval(
-        position = position_dodge(width = 0.1),
-        point_size = 2.5,
-        colour = "black"
-    ) +
-    labs(
-        x = "Frequency (Zipf)",
-        y = "Probability density",
-        fill = "Language",
-        shape = "Language",
-        colour = "Language"
-    ) +
-    theme(
-        legend.position = "none",
-        legend.title = element_blank()
-    )
-```
-
-### \# Phonemes
-
-Word length ($Phonemes$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.
-
-```{r}
-#| label: fig-items-nphon
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Distribution of number of phonemes across items"
-items %>%
-    count(language, n_phon) %>% 
-    ggplot() +
-    aes(
-        x = n_phon,
-        y = n,
-        fill = language,
-        colour = language,
-        shape = language
-    ) +
-    facet_wrap(~language, ncol = 1) +
-    geom_col() +
-    geom_text(aes(label = n), position = position_nudge(y = 5), colour = "black") +
-    labs(
-        x = "# Phonemes",
-        y = "Number of words",
-        fill = "Language",
-        shape = "Language",
-        colour = "Language"
-    ) +
-    scale_x_continuous(breaks = 1:15) +
-    theme(
-        legend.position = "none",
-        axis.ticks.x = element_line(),
-        legend.title = element_blank()
-    )
-```
-
-### Phonological similarity (Levenshtein)
-
-Phonological similarity between translation equivalents was computed as the normalised Levenshtein similarity between the word-form and its translation in the other language [@levenshtein1966binary].
-This score is calculated by first calculating the Levenshtein distance between the two transcriptions (number of insertions, deletions or replacements needed for the shortest transcription to become identical to the longer transcription), then dividing the resulting value by the length of the longest transcription, and finally subtracting this value from 1.
-This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches].
-
-```{r}
-#| label: fig-items-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 9
-#| fig-cap: "Distribution of Levenshtein similarities across items"
-items %>% 
-    distinct(te, .keep_all = TRUE) %>% 
-    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
-    count(lv) %>% 
-    mutate(prop = n/sum(n)) %>% 
-    ggplot(aes(x = lv, n)) +
-    geom_col(fill = clrs[1], colour = "white") +
-    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
-              position = position_nudge(y = 3), size = 3.5) +
-    labs(x = "Levenshtein distance", y = "Probability density") +
-    theme(
-        panel.grid.major.x = element_blank(),
-        legend.position = "top"
-    )
-```
-
-### Frequency by \# Phonemes
-
-```{r}
-#| label: fig-items-frequency-phonemes
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 4
-#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
-items %>% 
-    drop_na(freq_zipf) %>% 
-    count(n_phon, freq_zipf, language) %>% 
-    ggplot(aes(n_phon, freq_zipf, colour = language, fill = language)) +
-    facet_wrap(~language) +
-    geom_point(size = 2, alpha = 0.25) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
-    labs(
-        x = "# Phonemes", 
-        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    ) 
-```
-
-### Frequency by Levenshtein
-
-```{r}
-#| label: fig-items-frequency-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Association betwen lexical frequency and phonological similarity"
-items %>% 
-    left_join(select(bvqdev::pool, item, language)) %>% 
-    drop_na(freq_zipf) %>% 
-    count(freq_zipf, lv, language) %>% 
-    ggplot(aes(freq_zipf, lv, colour = language, fill = language, size = n)) +
-    facet_wrap(~language) +
-    geom_point(size = 2, alpha = 0.25) +
-    geom_smooth(method = "lm", formula = "y ~ x") +
-    labs(
-        y = "Levenshtein similarity", 
-        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "none",
-        panel.grid.major.y = element_blank()
-    )
-```
-
-### \# Phonemes by Levenshtein
-
-```{r}
-#| label: fig-items-nphon-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Association betwen number of phonemes and phonological similarity"
-items %>% 
-    left_join(select(bvqdev::pool, item, language)) %>% 
-    drop_na(freq_zipf) %>% 
-    count(n_phon, lv, language) %>% 
-    ggplot(aes(n_phon, lv, colour = language, fill = language, size = n)) +
-    facet_wrap(~language) +
-    geom_point(shape = 1, stroke = 1) +
-    geom_smooth(method = "lm", formula = "y ~ x", size = 1) +
-    labs(
-        x = "# Phonemes", 
-        y = "Levenshtein similarity", 
-        colour = "Language", 
-        fill = "Language"
-    ) +
-    scale_y_continuous(labels = percent) +
-    scale_x_continuous(breaks = seq(0, 16, 1)) +
-    theme(
-        legend.position = "none",
-        panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank()
-    ) 
-```
-
-### Language exposure-weighted lexical frequency
-
-We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to.
-the exposure of the $i$-th child to the $j$-th word measure is the product of the lexical frequency of the word (Zipf score) and the child’s degree of exposure to the corresponding language (a proportion)
-
-$$
-\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
-$$
-
-For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)`:
-
-$$
-\begin{aligned}
-\textbf{Exposure}_{\textbf{cavall, Catalan}} &= \textbf{Frequency}_{cavall} \times \textbf{DoE}_{Catalan} = 6.14 \times 0.8 = 4.91 \\
-\textbf{Exposure}_{\textbf{caballo, Spanish}} &= \textbf{Frequency}_{caballo} \times \textbf{DoE}_{Spanish} = 6.14 \times 0.2 = 1.23
-\end{aligned}
-$$
-
-```{r}
-#| label: fig-exposure-distribution
-#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
-#| warnings: false
-#| fig-height: 7
-#| fig-width: 13
-df %>% 
-    mutate(
-        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
-        exposure = cut(exposure, breaks = seq(0, 9, 1), include.lowest = TRUE)
-    ) %>% 
-    count(exposure, language) %>% 
-    ggplot(aes(exposure, n, fill = language)) +
-    facet_wrap(~language) +
-    geom_col(colour = "white") +
-    labs(
-        x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
-        y = "Number of observations",
-        fill ="Language"
-    ) +
-
-    df %>% 
-    mutate(
-        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
-        exposure = cut(exposure, breaks = seq(0, 9, 1), include.lowest = TRUE),
-        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
-    ) %>% 
-    count(exposure, freq, language) %>%    
-    ggplot(aes(x = freq, y = exposure, colour = language, size = n)) +
-    facet_wrap(~language) + 
-    geom_point() +
-    labs(
-        x = "Frequency (Zipf score)",
-        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
-        colour = "Language",
-        fill = "Language",
-        size = "Number of observations"
-    ) +
-    theme(panel.grid = element_blank()) +
-
-    plot_layout(ncol = 1) &
-    guides(fill = "none", colour = "none") &
-    theme(panel.grid.major.x = element_blank(),
-          panel.grid.minor.y = element_blank())
-```
-
-## Data analysis
-
-### Model comparison
-
-We extracted four variables of interest for each observation (describe in previous sections):
-
-* Word lexical frequency ($Frequency$)
-* Word length ($Phonemes$)
-* Phonological similarity ($Levenshtein$)
-* Participant age ($Age$)
-* Word exposure index ($Exposure$)
-
-We fitted a series of nested Bayesian ordinal multilevel models to sequentially test the effects of the independent variables (number of phonemes, exposure, age, and Levenshtein) over the three levels of word learning (no, production, comprehension).
-In all these models participants’ and items’ individual differences were controlled by estimating different parameters for each one (different intercepts and/or slopes).
-This is similar to the control conducted by ANCOVAs.
-Thus, independent variable effects account for collective differences in word learning (unexplained by participants’ or items’ individual differences).
-
-We initially fitted a base model (`model_fit_0`) that only included the predictors $Age$ (`age_std`), $Phonemes$ (`n_phon_std`), along with random intercepts and slopes of all predictors by $Participant$ (`id`) and $Item$ (`te`), and all correlation parameters.
-We then expanded this model by adding the main effect of $Exposure$ (`exposure`) (`model_fit_1`), and then $Levenshtein$ (`lv_std`) (`model_fit_2`) and their corresponding random slopes by $Participant$.
-Then, we added the $Exposure \times Levenshtein$ (`exposure_std:lv_std`) two-way interaction (`model_fit_3`) and its random slope by $Participant$.
-Finally, we added the $Age \times Exposure \times Levenshtein$ (`age_std:exposure_std:lv_std`) three-way interaction with its random slope by $Participant$, along with a random slope for the $Age \times Exposure$ (`age_std:exposure_std`) two-way interaction.
-
-We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package.
-LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set.
-Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.
-
-### Model equation
-
-::: column-body-outset
-$$\begin{aligned}
-g[\mathrm{Pr}(y_{ij} \le a_s)] = \tau_s - \nu_{ij}
-\end{aligned}$$
-:::
-
-$s$ is the s-th response category, $a_1 < a_2 < a_3$ are ordered response categories, $\mathrm{Pr}(y_i \le a_3) = 1$ and $\tau_s$ are threshold parameters where $\tau_1 < \tau_2 < \tau_3$.
-
-::: column-body-outset
-
-$$\begin{aligned}
-\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2\mathrm{Phon}_{j} + \beta_3\mathrm{Exposure}_{ij} + \beta_4 \mathrm{Levenshtein}_{j} + & [\mbox{Main effects}]\\
-&\beta_5\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_6\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_7\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Two-way interactions}]\\
-&\beta_8{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Three-way interaction}]\\
-
-&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} + \theta_{2j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{4j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) + & [\mbox{Item-level effects}]\\
-
-&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +   \theta_{2j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{5j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) & [\mbox{Participant-level effects}]
-\end{aligned}$$
-:::
-
-### Model prior
-
-::: column-body-outset
-$$\begin{aligned}\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
-\beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
-\beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
-\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]\end{aligned}$$
-:::
-
-In addition, we specified a $LKJ$ prior--$\sim LKJ(2)$ for the correlations between group-level coefficients.
-
-::: {.panel-tabset}
-
-#### Prior-predicted mean
-
-```{r}
-#| label: fig-prior-mean
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected prior predictions"
-nd <- expand.grid(
-    n_phon_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
-    exposure_std = 0,
-    lv_std = 0
+#| echo: false
+# load objects -----------------------------------------------------------------
+targets::tar_config_set(
+    store = here::here("_targets"),
+    script = here::here("_targets.R")
 )
 
-m <- add_epred_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA) %>% 
-    filter(.category != "No") %>% 
-    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(
-        c(Understands, `Understands and Says`),
-        names_to = ".category",
-        values_to = ".epred"
-    ) %>% 
-    mutate(
-        .category = factor(
-            .category,
-            levels = c("Understands and Says", "Understands"),
-            labels = c("Production\n(Understands and Says))", "Comprehension\n(Understands)")
-        )
+yaml_metadata <- rmarkdown::yaml_front_matter(
+    here::here("manuscript", "manuscript.qmd")
 )
 
-m %>% 
-    ggplot(aes(x = age_std, y = .epred)) +
-    facet_wrap(~.category) +
-    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
-    stat_summary(fun = "mean", geom = "line", size = 0.75) +
-    scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 4)
-    ) +
-    scale_fill_manual(
-        values = rev(clrs), 
-        labels = c("95%", "89%", "67%", "50%")
-    ) +
-    labs(
-        x = "Age (months)", 
-        y = "Probability of acquisition",
-        colour = "Credible interval", 
-        fill = "Credible interval"
-    ) +
-    scale_y_continuous(labels = percent, limits = c(0, 1)) +
-    theme(
-        legend.position = "top",
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid = element_blank()
-    )
-```
+title <- yaml_metadata$title
+abstract <- yaml_metadata$abstract
+thanks <- yaml_metadata$thanks
 
-#### Prior-predictions
+targets::tar_load(items)
+targets::tar_load(participants)
+targets::tar_load(responses)
 
-```{r}
-#| label: fig-prior-predictions
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected prior predictions"
-nd <- expand.grid(
-    n_phon_std = 0,
-    age_std = (10:36-mean(df$age))/(sd(df$age)),
-    exposure_std = 0,
-    lv_std = 0
-)
-
-m <- add_predicted_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA)
-
-m %>% 
-    count(age_std, .prediction) %>% 
-    pivot_wider(names_from = .prediction, values_from = n) %>% 
-    mutate(Understands = Understands + `Understands and Says`) %>% 
-    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
-    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
-    geom_col(position = position_fill()) +
-    labs(
-        x = "Age (months)", 
-        y = "Probability of acquisition",
-        colour = "Category", 
-        fill = "Category"
-    ) +
-    scale_x_continuous(
-        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 36, 4)
-    ) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        axis.ticks.x = element_line(colour = "black"),
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
-        strip.background = element_rect(fill = "grey", colour = NA),
-        plot.background = element_rect(fill = "white", colour = NA)
-    )
-```
-
-:::
-
-### Model settings
-
--   OS: `r sessionInfo()$running`, `r sessionInfo()$platform`
--   R version: `r sessionInfo()$R.version$version.string`
--   Algorithm: Hamiltonian Montecarlo (No U-turn Sampler, *NUTS*)
--   Engine: Stan (brms interface, `r sessionInfo()$otherPkgs$brms$Version`)
--   brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
--   Chains: `r dim(model_fit_4$fit)[2]`
--   Cores: `r dim(model_fit_4$fit)[2]`
--   Iterations: `r format(dim(model_fit_4$fit)[1], big.mark = ",")` (`r format(dim(model_fit_4$fit)[1]/2, big.mark = ",")`)
-
-More details in Appendix: Session Info.
-
-### R code (`brms`)
-
-::: panel-tabset
-
-#### Model 0
-
-```{r}
-print(model_fit_0$formula)
-```
-
-#### Model 1
-
-```{r}
-print(model_fit_1$formula)
-```
-
-#### Model 2
-
-```{r}
-print(model_fit_2$formula)
-```
-
-#### Model 3
-
-```{r}
-print(model_fit_3$formula)
-```
-
-#### Model 4
-
-```{r}
-print(model_fit_4$formula)
-```
-:::
-
-# Results
-
-## Descriptive statistics
-
-::: panel-tabset
-
-### By number of phonemes
-
-```{r}
-#| lab: fig-responses-age-phonemes
-#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
-df %>% 
-    mutate(
-        age = floor(age),
-        n_phon = cut_interval(n_phon, n = 4)
-    ) %>%
-    count(age, n_phon, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~n_phon) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-
-### By exposure
-
-```{r}
-#| lab: fig-responses-age-exposure
-#| fig-cap: "Proportion of responses to each category by age and exposure."
-df %>% 
-    mutate(
-        age = floor(age),
-        exposure = cut_interval(exposure, n = 4)
-    ) %>%
-    count(age, exposure, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~exposure) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    scale_y_continuous(labels = percent) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-
-### By Levenshtein
-
-```{r}
-#| lab: fig-responses-age-lv
-#| fig-cap: "Proportion of responses to each category by age and Levenhstein similarity"
-df %>% 
-    mutate(
-        age = floor(age),
-        lv = cut_interval(lv, n = 4)
-    ) %>%
-    count(age, lv, response) %>% 
-    ggplot(aes(age, n, fill = response)) +
-    facet_wrap(~lv) +
-    geom_col(position = position_fill()) +
-    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
-    scale_y_continuous(labels = percent) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-:::
-
-## Model selection
-
-@tbl-results-loos shows the outputs of the *LOO-CV*.
-The extended model (`model_fit_3`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference.
-This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.
-
-```{r}
-#| label: tbl-results-loos
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV).
-This procedure involves estimating the likelihood of each observation given a model fitted without such observation.
-The resulting likelihoods of each model were summed to compute the expected log-predicted density (*ELPD*) of the model, and correcting it for the number of parameters in the model to penalised for its the complexity.
-*ELPD* values closer to zero indicate better predictive acuracy.
-We report the *ELPD*, effective number of parameters (*p*), and information criterion (*LOO-IC*) of each model along their associated standard errors in paratheses.
-The first column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_4`)."
-model_loos %>% 
-    loo_compare() %>% 
-    as.data.frame() %>%   
-    rownames_to_column("model") %>% 
-    relocate(
-        model, 
-        matches("diff"),
-        matches("elpd_loo"),
-        matches("p_loo"), 
-        matches("looic")
-    ) %>%
-    gt() %>% 
-    fmt_number(2:9, decimals = 1) %>% 
-    cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>% 
-    cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>% 
-    cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>% 
-    cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>% 
-    cols_label(
-        model = "Model",
-        elpd_loo = md("*ELPD*"),
-        p_loo = md("*p*"),
-        looic = md("*LOO-IC*"),
-        elpd_diff = md("Difference")
-    ) %>% 
-    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels()
-    ) %>% 
-    tab_style(
-        cell_fill(color = "grey", alpha = 0.5),
-        cells_body(matches("diff"))
-    )
-```
-
-### Leave-one-out diagnostics (Pareto-*k* diagnostics)
-
-::: panel-tabset
-
-#### Model 0
-
-```{r}
-#| label: fig-loo-diagnostics-0
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
-get_loo_plot <- function(x){
-    x$pointwise[,5] %>% 
-        data.frame(x = 1:length(.), y = .) %>% 
-        ggplot(aes(x, y)) +
-        geom_point(
-            size = 0.5, alpha = 0.1, colour = clrs[5], 
-            shape = 1
-        ) +
-        geom_hline(yintercept = 0.5, linetype = "dotted") +
-        labs(
-            x = "Data point",
-            y = "Pareto-k value",
-            title = "PSIS diagnostic plot"
-        ) +
-        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
-        scale_y_continuous(
-            limits = c(-0.6, 0.6)
-        )
-}
-
-get_loo_plot(model_loos$model_fit_0)
-```
-
-#### Model 1
-
-```{r}
-#| label: fig-loo-diagnostics-1
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
-get_loo_plot(model_loos$model_fit_1)
-```
-
-#### Model 2
-
-```{r}
-#| label: fig-loo-diagnostics-2
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
-get_loo_plot(model_loos$model_fit_2)
-```
-
-#### Model 3
-
-```{r}
-#| label: fig-loo-diagnostics-3
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
-get_loo_plot(model_loos$model_fit_3)
-```
-
-#### Model 4
-
-```{r}
-#| label: fig-loo-diagnostics-4
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
-get_loo_plot(model_loos$model_fit_4)
-```
-
-:::
-
-## Fixed effects
-
-```{r}
-#| label: results-posterior-fixed
-#| echo: true
-#| message: false
-#| warning: false
-post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)
-
-# tidy predictor names
-str_repl <- c(
-    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
-    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
-    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
-    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
-    "b_exposure_std" = paste0("Exposure (+1 SD, ",  percent(sd(df$exposure)), ")"),
-    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
-    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
-    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
-    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
-    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
-)
+library(dplyr)
+library(tidyr)
+library(knitr)
 
@@
 
-```{r}
-#| label: tbl-results-fixed
-#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-intercept_footnote <- paste0(
-    "Probability of comprehension or production at the grand age of the sample (", 
-    round(mean(df$age), 2), " months)"
-)
+# Cognate beginnings to bilingual lexical acquisition
 
-# summarise posterior draws
-posterior_description %>% 
-    mutate(parameter = factor(parameter, levels = names(str_repl), labels = str_repl)) %>%  
-    select(parameter, median, ci_high, ci_low, rope_percentage) %>% 
-    gt(rowname_col = "parameter") %>% 
-    fmt_number(c(1:4), decimals = 2) %>% 
-    fmt_percent(5, decimals = 2) %>% 
-    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>% 
-    cols_label(
-        median = "Median",
-        ci_low = md("95% *HDI*"),
-        rope_percentage = md("*ROPE* prob.")
-    ) %>% 
-    tab_style(
-        cell_text(weight = "bold"),
-        cells_column_labels(columns = 1:5)
-    )
+Gonzalo Garcia-Castro [:bird:](https://twitter.com/gongcastro) [:elephant:](https://fediscience.org/@gongcastro) • Universitat Pompeu Fabra
 
-post_draws_list <- split(posterior_description, posterior_description$parameter)
-```
+Daniela S. Avila Varela • Universidade de Lisboa
 
-We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian].
-We decided to draw this interval around zero so that it covered values smaller than those considered as negligible by similar previous studies: after consulting the regression tables of [ADD STUDIES], we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: 3%.
-consequently, we defined our *ROPE* from -3% to +3%.
-We report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*).
-This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data.
-Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the -3%--3% *ROPE* interval we previously defined.
-For those coefficients that did not exclude completely the *ROPE* we also report a Bayes Factor indicating how many times is such coefficient more likely to be bigger than the upper limit of the *ROPE*, or smaller than the lower limit of the *ROPE*, given our data.
+Ignacio Castillejo [:bird:](https://twitter.com/IndaloNacho) • Universidad Autónoma de Madrid
 
-For interpretability, we report the estimated regression coefficients transformed from the probability scale.
-The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly.
-This means that one logit difference is not necessarily translated to a unique value in the probability scale.
-For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months.
-The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial.
-We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e.
-derivative) of the logistic curve at the age at which most participants were acquiring a given word.
-This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].
+Nuria Sebastian-Galles • Universitat Pompeu Fabra
 
-@fig-results-fixed shows the outputs of the model.
-**Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r printnum(post_draws_list[["b_age_std"]]$median, digits = 2)`, 95% *HDI* = [`r printnum(post_draws_list[["b_age_std"]]$ci_low, digits = 2)`, `r printnum(post_draws_list[["b_age_std"]]$ci_high, digits = 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month.
+-------------------------------------------------
 
-The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]).
-For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.
+| Link                                                             | Contents                                                       |
+|------------------------------------------------------------------|----------------------------------------------------------------|
+| [Website](https://gongcastro.github.io/cognate.beginnings)       | Instructions for reproducibility, data dictionaries, lab notes |
+| PsyArxiv                                                         | Preprint and figures                                           |
+| [GitHub](https://github.com/gongcastro/cognate.beginnings)       | Code, preprint and figures                                     |
+| [Open Science Framework](https://osf.io/hy984/)                  | Code, preprint, and results (model outputs)                    |
+| [Docker](https://hub.docker.com/r/gongcastro/cognate-beginnings) | Docker image with reproducible RStudio session                 |
 
-**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]).
-All of the posterior samples of his regression excluded the *ROPE*.
-The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/(sd(df$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.
+## Abstract
 
-The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.
+`r abstract`
 
-However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word.
-Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially.
-This effect was negligible when exposure to both languages was balanced (e.g.
-50-50%).
-Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.
+## Acknowledgements
 
-The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*.
-This shows that the effect of exposure on the speed at which participants acquired words across ages was inconclusive.
-Given our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.
-
-The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.
-
-Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.
-
-```{r}
-#| label: fig-results-fixed
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 9
-#| fig-heigh: 7
-#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model.
-Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis).
-Intervals show the 95% highest density interval (*HDI*) of each distribution.
-This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data.
-the mean and HDI limits are indicated below each distribution.
-Coefficients were transformed from the logit scale to the probability scale for interpretability.
-We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e.
-the derivative of the logistic function in its mid-point."
-post <- post_draws %>% 
-    mutate(
-        .variable_name = factor(
-            .variable, 
-            levels = names(str_repl),
-            labels = str_repl,
-            ordered = TRUE
-        )
-    )
-
-post_summary <- post %>% 
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    median_hdi(.exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")) %>% 
-    mutate(
-        .variable_name = factor(
-            .variable, 
-            levels = names(str_repl),
-            labels = str_repl,
-            ordered = TRUE
-        )
-    )
-
-post %>%
-    filter(!grepl("Intercept|sd", .variable)) %>%
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    annotate(geom = "rect", ymin = -Inf, ymax = Inf,
-             xmin = rope_interval["lower"], xmax = rope_interval["upper"],
-             colour = NA, alpha = 0.2, fill = clrs[5],) +
-    geom_vline(xintercept = 0, size = 1, colour = clrs[5]) +
-    stat_slab(
-        aes(fill = stat(abs(x) < rope_interval["upper"]),
-            colour = stat(abs(x) < rope_interval["upper"]),),
-        size = 1, position = position_nudge(y = 0.15)
-    ) +
-    geom_errorbar(
-        data = post_summary,
-        aes(xmin = .lower, xmax = .upper, x = .value),
-        width = 0.15
-    ) +
-    geom_text(
-        data = post_summary,
-        aes(
-            x = .value, 
-            y = .variable_name,
-            label = paste0(
-                round(.value, 1),
-                " [", round(.lower, 1), ", ", round(.upper, 1), "]")
-        ),
-        colour = "black", position = position_nudge(y = -0.3), size = 3
-    ) +
-    labs(
-        x = "Logistic regression coefficient estimate (Logit scale)",
-        y = "Variable", 
-        fill = "Overlaps with ROPE",
-        colour = "Overlaps with ROPE"
-    ) +
-    scale_x_continuous(breaks = seq(-2, 2, 0.2)) +
-    scale_fill_manual(
-        values = clrs[c(1, 4)],
-        labels = c("No", "Yes")
-    ) +
-    scale_colour_manual(
-        values = clrs[c(1, 4)],
-        labels = c("No", "Yes")
-    ) +
-    theme(
-        legend.position = "top",
-        legend.justification = c(1, 1),
-        axis.title.y = element_blank(),
-        panel.grid.major.x = element_line(
-            colour = "grey85",
-            linetype = "dotted"
-        ),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.x = element_blank()
-    )
-```
-
-## Marginal predictions
-
-```{r}
-#| label: fig-results-marginal
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Expected posterior predictions.
-Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient.
-These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model.
-Therefore, we generated 4,000 predictions for a series of combination of levels of interest.
-Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels.
-The X-axis indicates the age (in months) for which the prediction is generated.
-The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*).
-Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents).
-Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual).
-Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
-
-nd <- datagrid(
-    model = model_fit_4,
-    age_std = seq_range(model_fit_4$data$age_std, n = 100),
-    lv_std = (seq(0, 1, 0.5)-mean(df$lv))/(sd(df$lv)),
-    exposure_std = c(-1, 0, 1),
-    te = NA,
-    id = NA
-)
-
-marg_epreds <- predictions(
-    model_fit_4,
-    newdata = nd,
-    ndraws = NULL, 
-    re_formula = NA
-)
-
-marg_epreds %>% 
-    posteriordraws() %>% 
-    as_tibble() %>% 
-    select(drawid, group, draw, age_std, lv_std, exposure_std) %>%
-    filter(group != "No") %>% 
-    pivot_wider(
-        id_cols = c(drawid, age_std, lv_std, exposure_std), 
-        names_from = group, 
-        values_from = draw
-    ) %>% 
-    mutate(`Understands` = `Understands and Says` + `Understands`) %>% 
-    pivot_longer(c(`Understands`, `Understands and Says`), names_to = "group", values_to = "draw") %>% 
-    mutate(
-        exposure_std = factor(
-            exposure_std, 
-            levels = unique(nd$exposure_std), 
-            labels = c("Exposure: -1 SD", "Exposure: mean", "Exposure: +1 SD")
-        ),
-        lv_std = factor(
-            lv_std,
-            levels = unique(nd$lv_std), 
-            labels = paste0(c(0, 50, 100), "% similarity")
-        )
-    ) %>% 
-    ggplot(aes(age_std, draw, colour = lv_std, fill = lv_std)) +
-    facet_grid(group~exposure_std) +
-    stat_summary(fun.data = median_hdi, geom = "ribbon",
-                 alpha = 0.5, colour = NA) +
-    # geom_line(aes(group = interaction(lv_std, drawid)), 
-    #           size = 0.75, alpha = 0.15) +
-    stat_summary(aes(linetype = lv_std), fun = median, geom = "line", size = 1) +
-    labs(
-        x = "Age (months)",
-        y = "Probability of acquisition",
-        linetype = "Phonological similarity (Levenshtein)",
-        colour = "Phonological similarity (Levenshtein)",
-        fill = "Phonological similarity (Levenshtein)"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    scale_y_continuous(
-        breaks = seq(0, 1, 0.25),
-        limits = c(0, 1),
-        labels = percent
-    ) +
-    scale_x_continuous(
-        breaks = (seq(10, 32, 4)-mean(df$age))/(sd(df$age)),
-        labels = seq(10, 32, 4)
-    ) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
-    )
-
-```
-
-## Marginal effects
-
-In order to examine the impact of the word exposure index on the effect of phonological similarity we used the [`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects/index.html) R package to compute the average marginal effect of the Levenshtein distance on the probability of acquisition (comprehension or production) for the range of observed values of the word exposure index in our dataset.
-
-```{r}
-#| label: fig-marginal-effect
-#| fig-cap: "Average marginal effect of phonological similarity (Levenshtein distance) on the probability of comprehension and production across the range of values of exposure.
-The observed range of values of the word exposure index is indicated in the X-axis, and the estimated average marginal effect of phonological similarity (Levenshtein distance) is indicated in the Y-axis.
-This quantity is generated by averaging the derivative of the regression coefficient of the predictor Levenshtein across the range of values of all predictors but Exposure.
-We represent the uncertainty associated with the resulting estimate by generating multiple values of this marginal effect by drawing multiple samples from the posterior distribution of the model coefficients, calculating the average marginal effect for each posterior draw, and plotting the result as a thin grey line.
-Closer lines indicate lower uncertainty over the true value of the average marginal effect of Levenshein, while more distant lines indicate higher uncertainty."
-nd <- datagrid(
-    model = model_fit_4,
-    n_phon_std = 0,
-    lv_std = 0,
-    exposure_std = seq(-1.5, 1.5, length.out = 100),
-    age_std = 0,
-    id = NA,
-    te = NA
-)
-
-marg_effects <- marginaleffects(
-    model_fit_4,
-    newdata = nd, 
-    re_formula = NA,
-    ndraws = NULL,
-    vcov = FALSE,
-    type = "link"
-)
-
-marg_effects %>% 
-    posteriordraws() %>%
-    as_tibble() %>% 
-    filter(term=="lv_std") %>% 
-    mutate(draw = draw/4) %>% 
-    ggplot(aes(exposure_std, draw)) +
-    geom_hline(yintercept = 0, size = 1, colour = "grey") +
-    # geom_line(aes(group = drawid), alpha = 0.25, size = 0.75, colour = clrs[1]) +
-    stat_summary(fun.data = median_hdi, geom = "ribbon",
-                 colour = NA, alpha = 0.5, fill = clrs[1]) +
-    stat_summary(fun = median, geom = "line", size = 1, color = clrs[1]) +
-    labs(
-        x = "Word exposure index",
-        y = "Marginal effect of phonological similarity\n(Levenshtein distance)"
-    ) +
-    scale_x_continuous(
-        labels = label_number(style_positive = "plus", suffix = " SD"),
-        breaks = seq(-2, 2, 0.5)
-    ) +
-    scale_y_continuous(
-        labels = label_percent(style_positive = "plus", ),
-        breaks = seq(-1, 1, 0.05)
-    ) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "black", size = 0.75)
-    )
-
-```
-
-## Random effects
-
-### Participant-level effects
-
-::: panel-tabset
-
-#### Intercept
-
-```{r}
-#| label: fig-results-random-id-intercept
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior probability of acquisition by participant"
-re_id <- ranef(model_fit_4)$id[,,1] %>% 
-    as_tibble() %>% 
-    rownames_to_column("id") %>% 
-    left_join(distinct(model_fit_4$data, id)) %>% 
-    clean_names() %>% 
-    mutate(
-        across(
-            .cols = where(is.numeric),
-            .fns = list(
-                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
-                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
-            ), .names = "{.col}__{.fn}"
-        )
-    ) %>% 
-    select(-c(estimate:q97_5)) %>% 
-    pivot_longer(
-        where(is.numeric), 
-        names_to = c(".value", "type"),
-        names_sep = "__"
-    )
-
-re_id %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate), 
-        xmin = q2_5, xmax = q97_5,
-        fill = type, colour = type) +
-    facet_wrap(~type) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "Participant", 
-        colour = "Type"
-    ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age slope
-
-```{r}
-#| label: fig-results-random-id-age
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_age <- ranef(model_fit_4)$id[,,"age_std"]
-
-re_age %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(id = unique(model_fit_4$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### # Phonemes slope
-
-```{r}
-#| label: fig-results-random-id-nphon
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the number of phonemes by participant"
-re_n_phon <- ranef(model_fit_4)$id[,,"n_phon_std"]
-
-re_n_phon %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(id = unique(model_fit_4$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(
-        x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5
-    ) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Exposure slope
-
-```{r}
-#| label: fig-results-random-id-exposure
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of lexical frequency by participant"
-re_freq <- ranef(model_fit_4)$id[,,"exposure_std"]
-
-re_freq %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(id = unique(model_fit_4$data$id)) %>% 
-    relocate(id) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of phonological similarity by participant"
-re_lv <- ranef(model_fit_4)$id[,,"lv_std"]
-
-re_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Doe-by-Levenshtein slope
-
-```{r}
-# | label: fig-results-random-id-exposure-lv
-# | echo: true
-# | message: false
-# | warning: false
-# | fig-width: 7
-# | fig-cap: "Posterior random slope of age by participant"
-re_exposure_lv <- ranef(model_fit_4)$id[,,"exposure_std:lv_std"]
-
-re_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-DoE slope
-
-```{r}
-#| label: fig-results-random-id-age-doe
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
-re_age_exposure <- ranef(model_fit_4)$id[,,"age_std:exposure_std"]
-
-re_age_exposure %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-age-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
-re_age_lv <- ranef(model_fit_4)$id[,,"age_std:lv_std"]
-
-re_age_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_4$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-DoE-by-Levenshtein slope
-
-```{r}
-#| label: fig-results-random-id-age-exposure-lv
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
-re_age_exposure_lv <- ranef(model_fit_4)$id[,,"age_std:exposure_std:lv_std"]
-
-re_age_exposure_lv %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(id = unique(model_fit_3$data$id)) %>%
-    relocate(id) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(id, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Correlations
-
-```{r}
-#| label: fig-results-random-id-correlations
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior correlations between random effects by participant"
-posterior_description_re %>% 
-    select(parameter, median) %>% 
-    filter(str_detect(parameter, "cor_id__")) %>% 
-    mutate(parameter = str_remove_all(parameter, "cor_id__")) %>%  
-    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
-    mutate(
-        across(
-            c(var_1, var_2), 
-            ~factor(
-                .,
-                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std", "exposure_std:lv_std", "age_std:exposure_std:lv_std"),
-                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein", "Exposure \u00d7 Levenshtein", "Age \u00d7 Exposure \u00d7 Levenshtein")
-            )
-        )
-    ) %>%
-    filter(var_1 != var_2) %>% 
-    ggplot(aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1))) +
-    geom_point(shape = 15, size = 20) +
-    geom_text(colour = "black") +
-    scale_colour_gradientn(colours = clrs) +
-    labs(colour = "Correlation", size = "Correlation") +
-    theme(
-        axis.title = element_blank(),
-        axis.text.x = element_text(angle = 35, hjust = 1),
-        panel.grid = element_blank()
-    )
-```
-:::
-
-### Item-level effects
-
-::: panel-tabset
-
-#### Intercept
-
-```{r}
-#| label: fig-results-random-te
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior probability of acquisition by translation equivalent"
-re_te <- ranef(model_fit_4)$te[,,"Intercept"] %>% 
-    as_tibble() %>% 
-    rownames_to_column("te") %>% 
-    clean_names() %>% 
-    mutate(
-        across(
-            .cols = where(is.numeric),
-            .fns = list(
-                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
-                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
-            ), .names = "{.col}__{.fn}"
-        )
-    ) %>% 
-    select(-c(estimate:q97_5)) %>% 
-    pivot_longer(
-        where(is.numeric), 
-        names_to = c(".value", "type"),
-        names_sep = "__"
-    ) %>% 
-    mutate(te = as.numeric(te))
-
-re_te %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate), 
-        xmin = q2_5, xmax = q97_5,
-        colour = type, fill = type) +
-    facet_wrap(~type) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior probability of acquisition",
-        y = "TE", 
-        colour = "Type"
-    ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age slope
-
-```{r}
-#| label: fig-results-random-te-age
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by translation equivalent"
-re_age <- ranef(model_fit_4)$te[,,"age_std"]
-
-re_age %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_4$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### # Phonemes slope
-
-```{r}
-#| label: fig-results-random-te-nphon
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
-re_n_phon <- ranef(model_fit_3)$te[,,"n_phon_std"]
-
-re_n_phon %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_3$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Exosure slope
-
-```{r}
-#| label: fig-results-random-te-exposure
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_doe <- ranef(model_fit_4)$te[,,"exposure_std"]
-
-re_doe %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(te = unique(model_fit_4$data$te)) %>%
-    relocate(te) %>%
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA, fill = clrs[1]) +
-    geom_line(aes(group = 1), size = 1, colour = clrs[1]) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant",
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Age-by-exposure slope
-
-```{r}
-#| label: fig-results-random-te-age-exposure
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior random slope of age by participant"
-re_age_doe <- ranef(model_fit_4)$te[,,"age_std:exposure_std"]
-
-re_age_doe %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(te = unique(model_fit_4$data$te)) %>% 
-    relocate(te) %>% 
-    ggplot() +
-    aes(x = estimate, y = reorder(te, estimate),
-        xmin = q2_5, xmax = q97_5) +
-    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
-    geom_line(aes(group = 1), size = 1) +
-    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
-    labs(
-        x = "Posterior random slope estimate",
-        y = "Participant", 
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.grid = element_line(linetype = "dotted"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.y = element_blank()
-    )
-```
-
-#### Correlations
-
-```{r}
-#| label: fig-results-random-te-correlations
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-cap: "Posterior correlations between random effects by translation equivalent"
-posterior_description_re %>% 
-    select(parameter, median) %>% 
-    filter(str_detect(parameter, "cor_te__")) %>% 
-    mutate(parameter = str_remove_all(parameter, "cor_te__")) %>%  
-    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
-    mutate(
-        across(
-            c(var_1, var_2), 
-            ~factor(
-                .,
-                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std"),
-                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein")
-            )
-        )
-    ) %>%
-    filter(var_1 != var_2) %>% 
-    ggplot() +
-    aes(x = var_1, y = fct_rev(var_2), colour = median, label = percent(median, accuracy = 0.1)) +
-    geom_point(shape = 15, size = 40) +
-    geom_text(colour = "black") +
-    scale_colour_gradientn(colours = clrs) +
-    labs(
-        colour = "Correlation",
-        size = "Correlation"
-    ) +
-    theme(
-        axis.title = element_blank(),
-        panel.grid = element_blank()
-    )
-```
-:::
-
-## Model diagnostics
-
-### Simulation-based calibration
-
-```{r}
-#| label: fig-simulation-based-calibration
-#| fig-cap: "Posterior predictive checks.
-We simulated 500 datasets from our model by drawing random samples from its posterior distribution.
-We then computed the proportion of 'No', 'Understands', and 'Understands and Says' responses in each dataset.
-These simulated proportions are depicted in the plot as solid horizontal lines.
-Their variation indicate the uncertainy of the model about what proportion of each type of response would be observe in a newly collected dataset.
-We also plotted the proportion of each type of response in the observed dataset (with a ± 5% interval for comparison)."
-predicted_dfs <- add_predicted_draws(
-    object = model_fit_4, 
-    newdata = df,
-    re_formula = NULL,
-    ndraws = 500,
-    seed = 888
-)
-
-observed_prop <- df %>% 
-    count(response) %>%
-    mutate(prop = n/sum(n)) %>% 
-    rename(.prediction = response)
-
-predicted_dfs %>% 
-    ungroup() %>% 
-    select(.prediction, .draw) %>% 
-    count(.draw, .prediction) %>%
-    group_by(.draw) %>% 
-    mutate(
-        prop = n/sum(n),
-        .prediction = factor(
-            .prediction,
-            levels = levels(df$response), 
-            ordered = TRUE
-        )
-    ) %>% 
-    ggplot(aes(.draw, prop, fill = .prediction, colour = .prediction)) +
-    geom_line(aes(group = .prediction)) +
-    geom_rect(
-        data = observed_prop,
-        aes(xmin = 0, xmax = max(predicted_dfs$.draw),
-            ymin = prop-0.025, ymax = prop+0.025,
-            colour = .prediction, fill = .prediction),
-        alpha = 0.25, colour = NA, inherit.aes = FALSE
-    ) +
-    # geom_segment(
-    #     data = observed_prop,
-    #     aes(x = -Inf, xend = Inf, y = prop, yend = prop,
-    #         colour = .prediction),
-    #     size = 1, inherit.aes = FALSE
-    # ) +
-    labs(
-        x = "Iteration",
-        y = "Proportion of responses",
-        colour = "Response",
-        fill = "Response"
-    ) +
-    scale_colour_d3() +
-    scale_fill_d3() +
-    scale_x_continuous(labels = label_number(big.mark = ","), 
-                       breaks = seq(0, max(predicted_dfs$.draw), 100)) +
-    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.05)) +
-    theme(
-        legend.position = "top",
-        panel.grid = element_blank()
-    )
-
-```
-
-### Traceplots
-
-```{r}
-#| label: fig-diagnostics-traceplots
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 12
-gather_draws(model_fit_4, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
-    mutate(.chain = paste("Chain ", .chain)) %>%
-    ggplot(aes(.iteration, .value, colour = .chain)) +
-    facet_wrap(~.variable, scales = "free_y") +
-    annotate(geom = "rect", colour = NA,
-             xmin = 0, xmax = dim(model_fit_4$fit)[1]/2,
-             ymin = -Inf, ymax = Inf,
-             alpha = 0.25) +
-    geom_line() +
-    labs(
-        x = "Iteration",
-        y = "Sample value",
-        colour = "Chain"
-    ) +
-    scale_colour_manual(values = clrs[c(1, 3, 4, 5)]) +
-    theme(
-        legend.position = "top",
-        legend.title = element_blank(),
-        panel.grid = element_blank(),
-        axis.text.x = element_text(size = 7)
-    )
-```
-
-### Gelman-Rubin diagnostic (R-hat)
-
-```{r}
-#| label: fig-diagnostics-rhat
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| eval: false
-mcmc_rhat_data(model_rhats$model_fit_4) %>% 
-    ggplot() +
-    aes(x = value) +
-    geom_histogram(bins = 100, na.rm = TRUE, fill = clrs[1], colour = "white") +
-    geom_rug(colour = clrs[1]) +
-    labs(
-        x = "R-hat",
-        y = "Number of samples",
-        colour = "Description",
-        fill = "Description"
-    ) +
-    theme(
-        legend.position = "top"
-    )
-```
-
-### Effective sample size
-
-```{r}
-#| label: fig-diagnostics-neff
-#| echo: true
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| eval: false
-mcmc_neff_data(model_neffs$model_fit_4) %>% 
-    ggplot() +
-    aes(x = value, fill = description, colour = description) +
-    geom_histogram(bins = 100, na.rm = TRUE, colour = "white") +
-    geom_rug() +
-    geom_vline(xintercept = 1, colour = "black", size = 0.75) +
-    labs(
-        x = "Effective sample size ratio",
-        y = "Number of samples",
-        colour = "Rating",
-        fill = "Rating"
-    ) +
-    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
-    scale_colour_manual(values = clrs[c(1, 3, 5)]) +
-    theme(
-        legend.position = "top"
-    )
-```
-
-# Discussion
-
-## Summary of the study
-
-## Summary of the results
-
-## Results in contrast with the literature
-
-## Mechanistic explanations
-
-## Limitations
-
-## Further steps
-
-# Appendix
-
-## Session info
-
-```{r}
-sessioninfo::session_info()
-```
+`r thanks`

### 2023-07-27 index.qmd
Update website

@@
 ---
-title: "Cognate beginnings to bilingual lexical acquisition"
-subtitle: "Lab notes"
-short-title: "Cognate beginnings"
 author:

### 2023-07-27 index.qmd
Update website

@@
 
-# Cognate beginnings to bilingual lexical acquisition
-
-Gonzalo Garcia-Castro [:bird:](https://twitter.com/gongcastro) [:elephant:](https://fediscience.org/@gongcastro) • Universitat Pompeu Fabra
-
-Daniela S. Avila Varela • Universidade de Lisboa
-
-Ignacio Castillejo [:bird:](https://twitter.com/IndaloNacho) • Universidad Autónoma de Madrid
-
-Nuria Sebastian-Galles • Universitat Pompeu Fabra
-
--------------------------------------------------
-
-| Link                                                             | Contents                                                       |
-|------------------------------------------------------------------|----------------------------------------------------------------|
-| [Website](https://gongcastro.github.io/cognate.beginnings)       | Instructions for reproducibility, data dictionaries, lab notes |
-| PsyArxiv                                                         | Preprint and figures                                           |
-| [GitHub](https://github.com/gongcastro/cognate.beginnings)       | Code, preprint and figures                                     |
-| [Open Science Framework](https://osf.io/hy984/)                  | Code, preprint, and results (model outputs)                    |
-| [Docker](https://hub.docker.com/r/gongcastro/cognate-beginnings) | Docker image with reproducible RStudio session                 |
+{{< include README.md >}}
 

### 2023-07-27 _repro-docker.qmd
Built site for gh-pages

--- 2023-07-26 docs/_repro-docker.qmd
@@
 1. Download and install Docker
-2. Pull the gongcastro/cognate-beginnings image. Two ways of doing this:
+2. Pull the [gongcastro/cognate-beginnings image](https://hub.docker.com/r/gongcastro/cognate-beginnings). Two ways of doing this:
     a) Using the Docker UI: go the the "Images" tab, click on "Search images to run", search for "gongcastro/cognate-beginnings", and click "Pull".
@@
 3. Run the image in a Docker container. Two ways of doing this:
-    a) Using the Docker UI: in the "Images" tab, look for the "gongcastro/cognate-beginnings" image and click "Run".
+    a) Using the Docker UI: in the "Images" tab, look for the `gongcastro/cognate-beginnings` image and click "Run".
     b) Using the Docker CLI:
@@
 Open [https://localhost:8787](https://localhost:8787) in your browser.
-Log in using `rstudio` as users, and `rstudio` as password.
+Log in using `rstudio` as user, and `rstudio` as password.
 An RStudio session should open, with all files, code and dependencies installed and ready.
-Check `targets::tar_visnetwork()` to examine the targets workflow.
-This will list the objects created by the R code, and how they relate to each other.
-To retrieve and explore a given object, run `targets::tar_load(target_name)`, where `target_name` is the name of the rarget you want to explore.
+To retrieve and explore a given object, run `targets::tar_load(target_name)`, where `target_name` is the name of the rarget you want to explore (e.g., `targets::tar_load(items)`).
 

### 2023-08-11 _repro.qmd
Built site for gh-pages

@@
 A second option is the more traditional approach of downloading the repository (or cloning it, if you know your way through Git/GitHub), installing the necessary dependencies (we recommend using renv for R dependencies, other system dependencies are listed below), and running the code (we recommend using the targets workflow).
+
+### Docker {{< fa brands docker >}}
+---
+
+1. Download and install Docker
+2. Pull the [gongcastro/cognate-beginnings image](https://hub.docker.com/r/gongcastro/cognate-beginnings). Two ways of doing this:
+    a) Using the Docker UI: go the the "Images" tab, click on "Search images to run", search for "gongcastro/cognate-beginnings", and click "Pull".
+    b) Using the Docker CLI: open your console/terminal/command prompt, and run:
+    ```bash 
+    docker pull gongcastro/cognate-beginnings
+    ```
+3. Run the image in a Docker container. Two ways of doing this:
+    a) Using the Docker UI: in the "Images" tab, look for the `gongcastro/cognate-beginnings` image and click "Run".
+    b) Using the Docker CLI:
+    ```bash 
+    docker run --rm -ti -e ROOT=true -e PASSWORD=rstudio -p 8787:8787 --name rstudio gongcastro/cognate-beginnings:latest
+    ```
+
+4.
+Open [https://localhost:8787](https://localhost:8787) in your browser.
+Log in using `rstudio` as user, and `rstudio` as password.
+An RStudio session should open, with all files, code and dependencies installed and ready.
+To retrieve and explore a given object, run `targets::tar_load(target_name)`, where `target_name` is the name of the rarget you want to explore (e.g., `targets::tar_load(items)`).
+
+> 💡 The RStudio session opened by Docker does not have root permissions (you will not be able to run `targets::tar_make()` to run the code from the R console).
+Instead, open the "Terminal" tab, and run `sudo Rscript -e "targets::tar_make()"`.
+The targets workflow should now be triggered.
+
+> 💡 The targets workflow is run to completion before your Rstudio session is open in Docker.
+This means that even if you trigger `targets::tar_make()`, virtually all targets will be skipped, as they are completed.
+If you want to run a specific target, delete it first (or any other target it depends on), and the targets workflow will be run, now generating all deleted targets again.
+Be aware that this might take some time, especially when fitting the models anew.
+
+## GitHub {{< fa brands github >}}
+
+0.  Download this repository
+1.  Install software dependencies
+2.  Install package dependencies with renv
+3.  Running the code with targets
+4.  Repository structure and files
+
+### 0. Download this repository ⬇️
+
+First, you will need to download this repository to your local machine.
+We suggest three different ways:
+
+a)  **Git terminal**: download and install
+    [Git](https://git-scm.com/downloads) with default settings and clone
+    this repository locally running the following command in your
+    console:
+
+``` bash
+git clone https://github.com/gongcastro/cognate-beginnings.git
+```
+
+> 💡 You don't need to run this command from the Git console. Just open
+> Command Prompt or Power Shell in Windows (just make sure that Git is
+> included in your [PATH
+> variables](https://www.delftstack.com/howto/git/add-git-to-path-on-windows/),
+> which should be already done after the recommended Git installation),
+> Terminal in MacOS, or command line in Linux and run this command.
+
+b)  **GitHub releases**: download the latest
+    [release](https://github.com/gongcastro/cognate-beginnings/releases)
+    of the repository.
+
+c)  **Direct download**: click the [green
+    button](https://github.com/gongcastro/cognate-beginnings/archive/refs/heads/main.zip)
+    "Code" in this page (upper-right corner), click "Download ZIP", and
+    unzipping the downloaded repository.
+
+### 1. Install software dependencies 💻
+
+Here's a list of programs that you might need to install in your machine
+to make the project work:
+
+-   [R](https://www.r-project.org/) (4.2.2 or greater)
+
+-   [RStudio](https://posit.co/download/rstudio-desktop/): although not
+    strictly necessary, we encourage the use of Rstudio for reproducing
+    the present project. We have not tested the reproducibility of the
+    project in other IDEs.
+
+-   [Quarto](https://quarto.org/docs/get-started/) (1.3.340 or greater):
+    we use this software to generate the manuscript and lab notes. We
+    recommend installing version 1.3.340 or higher, although previous
+    version might work as well.
+
+-   [Rtools](https://cran.r-project.org/bin/windows/Rtools/) (4.3 or
+    greater).
+
+-   [CmdStan](https://mc-stan.org/users/interfaces/cmdstan) (2.31.0 or
+    greater): we use the CmdStan backend in brms to fit our Bayesian
+    models. We recommend installing CmdStan using its R interface
+    [CmdStanR](https://mc-stan.org/cmdstanr/) following the [*Getting
+    started with
+    CmdStanR*](https://mc-stan.org/cmdstanr/articles/cmdstanr.html)
+    vignette. You can also install it following the [CmdStan
+    installation guide](https://mc-stan.org/users/interfaces/), and then
+    letting CmdStanR know the path to the folder using
+    `cmdstanr::set_cmdstan_path()`.
+
+    <div>
+
+    > 💡 **Installing CmdStan** is sometimes inconvenient, and might
+    > require some time to set up, specially if it's your first time
+    > doing so. Have patience and follow the CmdStan installation guide
+    > closely. If you encounter problems and just don't have the time,
+    > you can download the models from the [Open Science Framework
+    > repository](https://osf.io/hy984/) and skip the model-fitting
+    > steps.
+
+    </div>
+
+### 2. Install package dependencies with renv 📦
+
+Open the downloaded repository folder and click on the
+[cognate-beginnings.Rproj](cognate-beginnings.Rproj) to open an RStudio
+session (recommended) or just open an R session and set the working
+directory to the downloaded folder. Once open run:
+
+``` r
+install.packages("renv") # in case you need to install renv
+renv::restore()
+```
+
+This might take a couple of minutes, depending on the number of packages
+that need installation or an update.
+
+> 💡 **What does this step do?** The code in this repository needs some
+> packages to run, many of which might not have been already installed
+> in your local machine. Instead of having to install them yourself one
+> by one or updating already installed ones (which might change how they
+> behave, possibly breaking some of your code in other projects) this
+> repository uses the R package
+> [renv](https://rstudio.github.io/renv/articles/renv.html) to deal with
+> package libraries. The command `renv::restore()` will install all
+> necessary packages in the appropriate version (listed in the file
+> [renv.lock](renv.lock) in a self-contained R package library. This
+> process will not affect the packages that you had already installed in
+> your machine.
+>
+> 💡 **Something went wrong with renv. Now what?** Although renv
+> facilitates the reproducibility of the repository, it is not
+> guaranteed that things will work perfectly. If you encounter trouble
+> installing the packages using renv, try installing them individually
+> using (`install.packages()` or `renv::install()`) and try fixing
+> whatever issues arise (they might depend on your particular setup).
+> Some packages are used by RStudio itself, and might refiuse to update
+> them while you're using RStudio. Try installing them from the R
+> console outside of RStudio. Open an
+> [issue](https://github.com/gongcastro/cognate-beginnings/issues) on
+> this repository if you need further assistance! We'll get back to you
+> ASAP. :smile:
+
+### 3. Running the code with targets 🚀
+
+Once the package dependencies have been solved with renv, run the
+following command:
+
+``` r
+targets::tar_make()
+```
+
+Mind that this process might take some time. Refitting the brms models
+might take very long (days even, depending on your set-up). If you want
+to skip this step, download the "results" folder in the [Open Science
+Framework repository](https://osf.io/hy984/) and replace it in the
+repository.
+
+> 💡**What does this step do?** This repository's workflow is based on
+> [targets](https://books.ropensci.org/targets/). This means that the
+> code is run in the appropriate order according to its internal
+> dependencies. The code is organised as *targets*, defined in the
+> `_targets.R` file. Sometimes targets operate with the outcomes of
+> other targets. For instance, one target might run a function that
+> takes the outcome of a different target as argument. {targets} makes
+> sure that the functions and objects needed to run each target have
+> already been previously defined. You can visually explore this
+> repository's targets and its dependency structure by running
+> `targets::tar_visnetwork()`. You can explore the contents of the
+> executed targets by running `targets::tar_load_everything()`, which
+> will load all defined targets into your workspace.

### 2023-08-11 _repro.qmd
Built site for gh-pages

@@
 
-1. Download and install Docker
+1.
+Download and install Docker ([Windows](https://docs.docker.com/desktop/install/windows-install/), [Mac](https://docs.docker.com/desktop/install/mac-install/), [Linux](https://docs.docker.com/desktop/install/linux-install/))
 2. Pull the [gongcastro/cognate-beginnings image](https://hub.docker.com/r/gongcastro/cognate-beginnings). Two ways of doing this:
-    a) Using the Docker UI: go the the "Images" tab, click on "Search images to run", search for "gongcastro/cognate-beginnings", and click "Pull".
-    b) Using the Docker CLI: open your console/terminal/command prompt, and run:
+    a) Using the Docker Desktop app: go the "Images" tab, click on "Search images to run", search for "gongcastro/cognate-beginnings", and click "Pull".
     ```bash 
@@
     ```
+    b) Using the Docker CLI: open your console/terminal/command prompt, and run:
+
 3. Run the image in a Docker container. Two ways of doing this:

### 2023-08-11 _repro.qmd
Built site for gh-pages

@@
 2. Pull the [gongcastro/cognate-beginnings image](https://hub.docker.com/r/gongcastro/cognate-beginnings). Two ways of doing this:
-    a) Using the Docker Desktop app: go the "Images" tab, click on "Search images to run", search for "gongcastro/cognate-beginnings", and click "Pull".
+    a) Using the Docker CLI: open your console/terminal/command prompt, and run:
     ```bash 
@@
     ```
-    b) Using the Docker CLI: open your console/terminal/command prompt, and run:
-
+    b) Using the Docker Desktop app: go the "Images" tab, click on "Search images to run", search for "gongcastro/cognate-beginnings", and click "Pull".
 3. Run the image in a Docker container. Two ways of doing this:
-    a) Using the Docker UI: in the "Images" tab, look for the `gongcastro/cognate-beginnings` image and click "Run".
+    a) Using the Docker Desktop app: in the "Images" tab, look for the `gongcastro/cognate-beginnings` image and click "Run".
     b) Using the Docker CLI:
@@
 4.
-Open [https://localhost:8787](https://localhost:8787) in your browser.
+Open [http://localhost:8787](http://localhost:8787) in your browser.
 Log in using `rstudio` as user, and `rstudio` as password.

### 2023-08-11 manuscript/manuscript.qmd
Add comments by IC and NSG, fix links in site

@@
 abstract: |
-    Bilingual infants' developmental trajectories of lexical acquisition are equivalent to those shown by their monolingual peers.
-This is remarkable, given the increased complexity of their linguistic input.
-Recent studies suggested that bilingual vocabulary growth is boosted by the number of cognates shared by the pair of languages being learned, and that this facilitation effect is driven by a stronger parallel activation of cognates during linguistic exposure, compared to non-cognates.
+    Bilingual infants' developmental trajectories of lexical acquisition are equivalent to their monolingual peers'.
+This is remarkable, given the complexity of their linguistic input.
+Recent studies suggest that bilingual vocabulary growth is boosted by the number of cognates (form-similar translation equivalents) shared by the pair of languages being learned, and that this cognateness facilitation effect is driven by a stronger parallel activation of cognates during linguistic exposure, compared to non-cognates.
 The mechanisms behind this facilitation are still unclear.
-In this study, we capitalised on accumulator models of language acquisition to propose an account of bilingual lexical acquisition in which parallel activation increases the rate at which children accumulate learning instances for words in both languages, even in fully monolingual situations.
-Under this hypothesis, we predicted a stronger cognate facilitation for words to which children were exposed less frequently (low-exposure words), as they are co-activated by their translation more often than high-exposure words do.
+In this study, we propose an account of bilingual lexical acquisition in which parallel activation increases the rate at which children accumulate learning instances for words in both languages, even in fully monolingual situations.
+We predicted a stronger cognate facilitation for words to which children were exposed less frequently (low-exposure words), as they are co-activated by their translation more often than high-exposure words.
 We developed an extensive online vocabulary checklist, the Barcelona Vocabulary questionnaire (BVQ), to collect vocabulary data from 366 Catalan-Spanish bilingual toddlers aged 12 to 32 months.
-We then used Bayesian explanatory item response theory to model the acquisition trajectories of 302 Catalan/Spanish translation equivalents.
-We found an interaction between exposure and cognateness, which pointed to cognateness facilitating the aquisition of low-exposure words, but not of mean exposure or high-exposure words.
+We used Bayesian explanatory item response theory to model the acquisition trajectories of 604 Catalan and Spanish words.
+We found an interaction between exposure and cognateness, suggesting that cognateness facilitates the aquisition of low-exposure words, but not of mean exposure or high-exposure words.
 Overall, our findings suggest that cognateness plays a key role in bilingual lexical acquisition, and provides evidence for a frequency-mediated facilitation effect driven by parallel activation.
@@
     The authors declare no conflicts of interest with regard to the funding source of this study.
-GGC was supported by a FPI 2020 research contract (PRE2019-088165) in a project funded by the Spanish Ministry of Science, Innovation, and Universities (PGC2018-101831-B-I00).
+This study was supported by the Spanish Ministry for Science and Innovation and State Research Agency (Project PID2021- 123416NB-I00 financed by MCIN/ AEI/ 10.13039/501100011033 / FEDER, UE) and the Economic and Social Research Council (ESRC) (ES/S010947/1, UK).
+GGC was supported by a FPI research contract (PRE2019-088165).
 DAV was supported by the European Union's Horizon 2023 research and innovation program under Marie Skłodowska–Curie Grant (765556) and a postdoctoral fellowship from the Foundation for Science and Technology of Portugal (UIDB/00214/2020).
 IC was supported by the Investigo program funded by the European Union's NextGenerationEU (NGEU) recovery plan.
-NSG was supported by Academia 2019 award from the Catalan Institution for Research and Advanced Studies (ICREA).
+NSG was supported by an ICREA Academia award from the Catalan Institution for Research and Advanced Studies (ICREA).
 We are grateful to Chiara Santolin, Ege E.
@@
 
-targets::tar_config_set(store = here::here('_targets'),
-                        script = here::here('_targets.R'))
+targets::tar_config_set(
+    store = here::here("_targets"),
+    script = here::here("_targets.R")
+)
 
@@
 
-targets::tar_load(c(bvq_data, childes, items,
-                    participants, responses))
+targets::tar_load(c(
+    bvq_data, childes, items,
+    participants, responses
+))
 
-targets::tar_load(c(model_fit, model_summary, 
-                    model_epreds, model_convergence))
+targets::tar_load(c(
+    model_fit, model_summary, model_summary_re,
+    model_epreds, model_convergence
+))
 
@@
 my_theme <- theme_minimal() +
-    theme(panel.grid = element_line(colour = "grey",
-                                    linetype = "dotted"),
+    theme(
+        panel.grid = element_line(
+            colour = "grey",
+            linetype = "dotted"
+        ),
         axis.line = element_line(colour = "black"),
         text = element_text(size = 12, colour = "black"),
-          axis.text = element_text(colour = "black"))
+        axis.text = element_text(colour = "black")
+    )
 
@@
 
-The foundations of word learning are in place early in age.
-At six months, infants start directing their gaze to some objects when hearing their labels [@tincoff1999beginnings; @bergelson2012months; @bergelson2015early], and shortly after caregivers start reporting some words as acquired by their infant in vocabulary checklists [e.g., @fenson2007macarthurbates; @samuelson2021precision].
+The foundations of word learning are in place at an early age.
+At six months, infants start directing their gaze to objects when hearing their labels [@tincoff1999beginnings; @bergelson2012months; @bergelson2015early], and shortly after caregivers start reporting some words as acquired by their infant in vocabulary checklists [e.g., @fenson2007macarthurbates; @samuelson2021precision].
 Most research on early word acquisition relies extensively on data from monolingual children, and is oblivious to the fact that a substantial proportion of the world population acquires more than one language from early ages [@grosjean2021extent].
@@
 Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
-A more detailed analysis of the words in bilinguals’ lexicon shows some interesting patterns.
+A more detailed analysis of the words in bilinguals’ lexicons shows some interesting patterns.
 
@@
 The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh.
-The authors calculated the average lexical overlap between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of cognateness between each pair of languages.
+The authors calculated the average lexical overlap between the words in each of these additional languages and their translation equivalents in English.
+Lexical overlap was calculated in terms of phonological similarity (described below) and it was taken as a proxy of the degree of cognateness between each pair of languages.
 Floccia and co-workers reported an increase in vocabulary size in the additional language (i.e., not English) associated with an increase in the average phonological similarity between the translation equivalents of each language pair.
@@
 A second set of studies suggested that cognates are overrepresented in bilinguals’ early lexicon.
-@bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months and found that cognates represented a larger proportion of participant’s vocabulary than non-cognates.
+@bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months and found that cognates represented a larger proportion of vocabulary than non-cognates.
 @schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual child produced cognates earlier than non-cognates, on average.
 But the high proportion of cognates in the vocabulary of the participants in these two studies may not necessarily evidence of a facilitation effect of cognateness, but rather of simply the high proportion of cognates present in the pair of languages being learned.
-For instance, if two given languages share a high proportion of cognates like $\sim 70$%, the vocabulary contents of children learning both languages should, in principle, approximate such proportion of cognates, even in the absence of a cognateness facilitation effect.
-More recently, @mitchell2022cognates addressed this issue in a longitudinal study with a larger sample.
+For instance, if two given languages share a high proportion of cognates like 70%, the vocabulary contents of children learning both languages should, in principle, approximate such proportion of cognates, even in the absence of a cognateness facilitation effect.
+More recently, @mitchell2022cognates addressed this issue in a longitudinal study.
 The authors collected expressive vocabulary data of 47 16- to 30-month-old French-English bilinguals living in Canada, in both languages.
@@
 Critically, this difference persisted after both lists were matched in size, controlling their semantic category (i.e., furniture, animals, food were similarity represented in both lists) and age-of-acquisition norms (an index of word difficulty).
-Taken together, the results of these two lines of research have been considered as providing support to the hypothesis that phonological similarity (as reflected in cognateness) plays a facilitation role in bilingual word acquisition.
+Taken together, the results of these two lines of research have been considered as supporting to the hypothesis that phonological similarity (as reflected in cognateness) plays a facilitation role in bilingual word acquisition.
 
@@
 
-We propose an account in which a learning instance for a word may also represent a learning instance for its translation equivalent, to the extent that such translation equivalent is co-activated^[We use the term *learning instance* in the fashion of accumulator models of language acquisition; as an exposure to a word-form that constitutes an opportunity for the child to accumulate information about the word.
-We do not make strong assumptions about whether a learning instance is a discrete or a continuous unit of accumulation of information.
-We rather consider that a learning instance of a word is so to the degree of the strength of activation of its phonological representation.
-This activation may result from the infant being exposed to the actual word-form, or the result of activation spreading through phonological or semantic links across lexical representation, as in the case of parallel activation.].
+We propose an account in which a learning instance for a word may also represent a learning instance for its translation equivalent, to the extent that such translation equivalent is co-activated.
+We use the term *learning instance* in the fashion of accumulator models of language acquisition; as an exposure to a word-form that constitutes an opportunity for the child to accumulate information about the word.
+We do not assume if a learning instance is a discrete or a continuous unit of accumulation of information.
+We consider that a learning instance of a word consists of an exposure to the its phonological form if the resulting strength of activation of its phonological representation in the child's lexicon reaches some thoretical threshold that leads to word-form recognition.
+This activation may result from the infant being exposed to the actual word-form, or the result of activation spreading through phonological or semantic links across lexical representation, as in the case of parallel activation.
 The strength of this co-activation is proportional to the phonological similarity between the two translation equivalents; given that cognates share higher phonological similarity than non-cognates, the former should be co-activated more strongly than the latter.
@@
 Therefore, the acquisition of words from the lower-exposure language would benefit more strongly from their cognate status than words from the higher-exposure language.
-This asymmetric cross-language activation would be consistent with previous reports of larger priming effects from the dominant to the non-dominant language,
+This asymmetric cross-language activation would be consistent with previous reports of larger priming effects from the dominant to the non-dominant language.
 
-Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [*cat*], that  are phonologically very similar.
-When the child is exposed to /\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in parallel.
+Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [*cat*], which are phonologically very similar.
+When the child listens to /\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in parallel.
+If the child has already formed a form-meaning association for both word-forms, parallel activation may result from the activation of their common concept or from activation spreading throught phonological similarity.
+We assume semantic co-activation to be constant across cognate and non-cognate translation equivalents, and focus on phonological co-activation as an additional source of activation that affects cognates more strongly than non-cognates.
 Therefore, this exposure will count as a learning instance for both co-activated forms.
@@
 Given the low phonological similarity between both word-forms, an exposure to \textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/ leading to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for /\textipa{"pe.ro}/.
-While /\textipa{"gat}-\textipa{"ga.to}/ will benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not.
+While /\textipa{"gat}-\textipa{"ga.to}/ will benefit from phonological co-activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not.
 If the child receives linguistic input from one of the languages more often than from the other, this effect might affect each form of the cognate translation equivalent differently.
@@
 As described in the methods section, rate of exposure was a composite measure taking into account participant’ language exposure and word’s lexical frequency.
-We predicted an interaction between cognate status and word exposure rate  in which the probability of comprehension is higher for low-exposure cognate words, but not for high-exposure cognate words.
+We predicted an interaction between cognate status and word-form exposure rate  in which the probability of comprehension is higher for low-exposure cognate words, but not for high-exposure cognate words.
 
@@
     count(version, language) |>
-    pivot_wider(names_from = language,
+    pivot_wider(
+        names_from = language,
         values_from = n,
-                names_repair = janitor::make_clean_names) |> 
+        names_repair = janitor::make_clean_names
+    ) |>
     mutate(version = tolower(version))
@@
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
-Items in one language were translation equivalents of the items in the other language (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for *dog*), roughly following a one-to-one mapping^[Although for some of the included words in Catalan did not have a clear translation or had more than one possible translation in Spanish, and vice versa, therefore the unequal number of words included in the Catalan and Spanish lists] (see @tbl-items for a summary of the questionnaire items).
+Items in one language were translation equivalents of the items in the other language (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for *dog*), roughly following a one-to-one mapping.
+Some of the words in Catalan did not have a clear translation or had more than one possible translation in Spanish, and vice versa, therefore the unequal number of words included in the two lists (see @tbl-items for a summary of the questionnaire items).
 
@@
 #| warning: false
-col_names <- c("Semantic category", 
+col_names <- c(
+    "Semantic category",
     "List A",
@@
     "List D",
-               "Examples")
+    "Examples"
+)
 
@@
     arrange(language, list) |>
-    mutate(a = grepl("A", list),
+    mutate(
+        a = grepl("A", list),
         b = grepl("B", list),
         c = grepl("C", list),
-           d = grepl("D", list)) |> 
+        d = grepl("D", list)
+    ) |>
     summarise(across(a:d, sum),
         items = list(meaning),
-              .by = semantic_category) |> 
+        .by = semantic_category
+    ) |>
     mutate(
@@
             paste(tail(x, 2),
-                  collapse = ", ")
-        })) |>  
+                collapse = ", "
+            )
+        })
+    ) |>
     unnest_longer(items) |>
@@
 sum_rows <- function(x) {
-    if (is.numeric(x)) sum(x) 
-    else if(is.list(x)) ""
-    else "N"
+    if (is.numeric(x)) {
+        sum(x)
+    } else if (is.list(x)) {
+        ""
+    } else {
+        "N"
+    }
 }
@@
 sums_row <- tbl_data |>
-    summarise(semantic_category = "N",
+    summarise(
+        semantic_category = "N",
         across(a:d, sum),
-              items = "-")
+        items = "-"
+    )
 
@@
     col.names = col_names,
-      align = "lccccl")  |> 
-    row_spec(nrow(tbl_data)-1, 
-             hline_after = TRUE)
+    align = "lccccl"
+) |>
+    row_spec(nrow(tbl_data) - 1,
+        hline_after = TRUE
+    )
 ```
@@
         left_join(distinct(participants, id, time_stamp, age),
-                  by = c("id", "age")) |> 
+            by = c("id", "age")
+        ) |>
         group_by() |>
         arrange(id, time_stamp) |>
-        mutate(diff_lag = time_stamp-dplyr::lag(time_stamp, default = NA),
-               .by = id) |>
-        summarise(diff_min = min(diff_lag, na.rm = TRUE),
+        mutate(
+            diff_lag = time_stamp - dplyr::lag(time_stamp, default = NA),
+            .by = id
+        ) |>
+        summarise(
+            diff_min = min(diff_lag, na.rm = TRUE),
             diff_max = max(diff_lag, na.rm = TRUE),
-                  .by = id) |> 
+            .by = id
+        ) |>
         dplyr::filter(!(diff_min %in% c(-Inf, Inf))) |>
-        summarise(diff_min = min(diff_min, na.rm = TRUE),
-                  diff_max = max(diff_max, na.rm = TRUE))
+        summarise(
+            diff_min = min(diff_min, na.rm = TRUE),
+            diff_max = max(diff_max, na.rm = TRUE)
+        )
 })
@@
     select(starts_with("doe_")) |>
-    mutate(doe_total = rowSums(across(everything())),
-           is_trilingual = doe_total < 1) |> 
+    mutate(
+        doe_total = rowSums(across(everything())),
+        is_trilingual = doe_total < 1
+    ) |>
     add_count(name = "n_total") |>
     count(is_trilingual, n_total) |>
-    mutate(prop = n/n_total)
+    mutate(prop = n / n_total)
 
@@
 
-col_names <- c("Catalan exposure",
+col_names <- c(
+    "Catalan exposure",
     "[10-14]",
@@
     "(26-30]",
-               "(30-34]")
+    "(30-34]"
+)
 
 tbl_data <- participants |>
-    mutate(across(starts_with("doe_"), 
-                  \(x) floor(x * 10) / 10)) |> 
+    mutate(across(
+        starts_with("doe_"),
+        \(x) floor(x * 10) / 10
+    )) |>
     mutate(
@@
             c(10, 14, 18, 22, 26, 30, 34, 36),
-                  include.lowest = TRUE),
+            include.lowest = TRUE
+        ),
         doe_catalan = as.character(
@@
                 c(0, 0.25, 0.5, 0.75, 1),
-                labels = c("0-25%",
+                labels = c(
+                    "0-25%",
                     "25-50%",
                     "50-75%",
-                           "75-100%"),
+                    "75-100%"
+                ),
                 include.lowest = TRUE
-            ))
+            )
+        )
     ) |>
@@
     arrange(desc(doe_catalan)) |>
-    pivot_wider(names_from = age, 
+    pivot_wider(
+        names_from = age,
         values_from = n,
-                names_repair = janitor::make_clean_names) |> 
+        names_repair = janitor::make_clean_names
+    ) |>
     mutate(across(where(is.numeric), \(x) ifelse(is.na(x), 0, x)))
@@
 sums_row <- as.data.frame(lapply(tbl_data, sum_rows),
-                          col.names = colnames(tbl_data))
+    col.names = colnames(tbl_data)
+)
 
@@
     col.names = col_names,
-      align = "lcccccc")  |> 
-    add_header_above(c(" " = 1,
-                       "Age (months)" = ncol(tbl_data)-1)) |>
-    row_spec(nrow(tbl_data)-1, 
-             hline_after = TRUE)
+    align = "lcccccc"
+) |>
+    add_header_above(c(
+        " " = 1,
+        "Age (months)" = ncol(tbl_data) - 1
+    )) |>
+    row_spec(nrow(tbl_data) - 1,
+        hline_after = TRUE
+    )
 ```
@@
     summarise(across(n, lst(median, min, max),
-                     .names = "{.fn}"))
+        .names = "{.fn}"
+    ))
 
@@
 n_items_nouns <- n_items |>
-    dplyr::filter(class=="Noun")
+    dplyr::filter(class == "Noun")
 
@@
     left_join(childes,
-              by = c("childes_lemma" = "token")) |> 
+        by = c("childes_lemma" = "token")
+    ) |>
     drop_na(freq_zipf, wordbank_lemma)
@@
 
-rope_interval <- c(lower = -0.1/4, upper = 0.1/4, accuracy = 0.1)
+rope_interval <- c(lower = -0.1 / 4, upper = 0.1 / 4, accuracy = 0.1)
 
@@
     r_info$R.version$minor,
-                        .sep = ".")
+    .sep = "."
+)
 stan_version <- "2.32.1"
@@
 We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, see @fourtassi2020growth for a similar approach].
-We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
+We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`, see @sec-predictors), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
 Finally, we removed items without a translation in the other language (*n* = `r nrow(n_items_multi)-nrow(n_items_duplicated)`).
@@
 
-### Model predictors
+### Model predictors {#sec-predictors}
 
 We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-form's lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
-First, we extracted the child-directed lexical frequency of each word from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
+First, we extracted the child-directed lexical frequency of each word-form from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
 Using the corresponding lexical frequencies directly from Catalan and Spanish was not possible due to the low number of Catalan participants and tokens available in their corresponding CHILDES corpora, so they were extracted from the English corpora instead.
@@
 We multiplied the resulting lexical frequencies by the reported degree of exposure of the child to Catalan or Spanish.
-For instance, for a child whose degree of exposure is 80% for Catalan and 20% for Spanish, the expected *Exposure* score to the Catalan word *cotxe* [*car*]---with a lexical frequency of `r round(items$freq[items$item=="cotxe"], 2)`---would be `r round(items$freq[items$item=="cotxe"]*0.80, 2)`, while that of its translation to Spanish *coche* would be `r round(items$freq[items$item=="coche"]*0.20, 2)`.
+For instance, for a child whose degree of exposure is 80% for Catalan and 20% for Spanish, the expected *Exposure* score to the Catalan word-form *cotxe* [*car*]---with a lexical frequency of `r round(items$freq[items$item=="cotxe"], 2)`---would be `r round(items$freq[items$item=="cotxe"]*0.80, 2)`, while that of its translation to Spanish *coche* would be `r round(items$freq[items$item=="coche"]*0.20, 2)`.
 
@@
 The Levenshtein distance measures the number of editions (insertions, deletions, or substitutions) that one string of phonemes/characters must go through to become identical to the other [@levenshtein1966binary].
-We divided the Levenshtein distance of each translation equivalent by the length of the longest word-form to correct for word length (longer strings are likely to show a larger number of mismatches).
+We divided the Levenshtein distance of each translation equivalent by the length of the longest word-form to correct for word-form length (longer strings are likely to show a larger number of mismatches).
 Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
@@
 post_draws_list <- model_summary |>
-    mutate(across(.value:.lower,
+    mutate(across(
+        .value:.lower,
         \(x) ifelse(grepl("Intercept", .type),
-                              plogis(x), x/4))) |> 
+            plogis(x), x / 4
+        )
+    )) |>
     split(model_summary$.variable)
+
+re_data <- model_summary_re |>
+    filter(.variable %in% c(
+        "b_lv_std",
+        "b_exposure_std:lv_std"
+    ))
+
+re_data_prop <- re_data |>
+    mutate(
+        is_positive = .value > 0,
+        is_negative = .value < 0
+    ) |>
+    summarise(across(matches("is_"), mean),
+        .by = .variable_name
+    ) |>
+    mutate(prop = ifelse(is_positive > is_negative,
+        is_positive,
+        is_negative
+    ))
 ```
@@
     mutate(
-        across(.value:.upper,
-               \(x) ifelse(.type==intercept_label, plogis(x), x/4)),
-        .variable_name = gsub("×", '$\\\\times$', .variable_name),
-        .hdi = paste0("[", round(.lower, 3),
-                      ", ", round(.upper, 3), "]"),
+        across(
+            .value:.upper,
+            \(x) ifelse(.type == intercept_label, plogis(x), x / 4)
+        ),
+        .variable_name = gsub("×", "$\\\\times$", .variable_name),
+        .hdi = paste0(
+            "[", round(.lower, 3),
+            ", ", round(.upper, 3), "]"
+        ),
         .rope = ifelse(is.na(.rope),
             "-",
-                       sub("^(-?)0.", "\\1.", sprintf("%.3f", .rope)))
+            sub("^(-?)0.", "\\1.", sprintf("%.3f", .rope))
+        )
     ) |>
@@
     select(-c(.lower, .upper)) |>
-    kable(format = "latex", 
+    kable(
+        format = "latex",
         escape = FALSE,
@@
         start_row = 1,
-              end_row = 2) |> 
+        end_row = 2
+    ) |>
     pack_rows("Slopes (upper bound)",
         start_row = 3,
-              end_row = 10) 
+        end_row = 10
+    )
 ```
@@
 A one-month increment in age increased a maximum of `r round(post_draws_list[["b_age_std"]]$.value/sd(responses$age), 2)` the probability of acquisition.
-Similarly, the word exposure index (*Exposure*) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 3)`]).
+Similarly, the word-form exposure index (*Exposure*) had a strong effect on the probability of acquisition ($\beta$ = `r round(post_draws_list[["b_exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std"]]$.upper, 3)`]).
 All of the posterior samples of this regression coefficient excluded the ROPE.
@@
 For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.value)/sd(responses$n_phon), 3)` less likely to know it.
-The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children.
+The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word-form exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children.
 
@@
 For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.value/(40*sd(responses$lv)), 3)`.
-The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to word.
+The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to the word-form.
 Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
@@
 Different colours indicate different levels of cognateness (phonological similarity).
-Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure.
+Predictions are presented separately for different degrees of word-form exposure index: little exposure to the word-form, mean exposure, and high exposure).
 Predictions for Comprehension are shown on top and predictions for Comprehension and Production are shown on the bottom.
@@
     ungroup() |>
-    reframe(.median = mean(.value),
+    reframe(
+        .median = mean(.value),
         .lower = tidybayes::qi(.value, .width = 0.95)[, 1],
         .upper = tidybayes::qi(.value, .width = 0.95)[, 2],
-            .by = c(ends_with("_std"), .category)) |> 
-    mutate(age = (age_std * sd(responses$age)) + mean(responses$age),
+        .by = c(ends_with("_std"), .category)
+    ) |>
+    mutate(
+        age = (age_std * sd(responses$age)) + mean(responses$age),
         exposure = factor(exposure_std,
             levels = unique(exposure_std),
-                             labels = c("Low exposure (-1 SD)",
+            labels = c(
+                "Low exposure (-1 SD)",
                 "Mean exposure",
-                                        "High exposure (+1 SD)"),
-                             ordered = TRUE),
+                "High exposure (+1 SD)"
+            ),
+            ordered = TRUE
+        ),
         lv = factor(lv_std,
@@
             labels = paste0(c(0, 50, 100), "%"),
-                       ordered = TRUE),
-           .category = ifelse(.category=="Understands",
+            ordered = TRUE
+        ),
+        .category = ifelse(.category == "Understands",
             "Comprehension",
-                              "Production")) 
+            "Production"
+        )
+    )
 fig_data |>
@@
         fill = lv,
-               linetype = lv)) +
-    facet_grid(.category~exposure) +
-    annotate(geom = "rect",
+        linetype = lv
+    )) +
+    facet_grid(.category ~ exposure) +
+    annotate(
+        geom = "rect",
         xmin = min(responses$age),
@@
         ymax = Inf,
-             fill = "grey95",
-             linetype = "dotted") +
-    annotate(geom = "text",
+        fill = "orange",
+        alpha = 1 / 4,
+        linetype = "dotted"
+    ) +
+    annotate(
+        geom = "text",
         x = min(responses$age),
@@
         label = "Observed\nage range",
-             size = 2) +
-    geom_vline(xintercept = mean(responses$age), 
-               linewidth = 1/4,
-               linetype = "dotted") +
-    geom_hline(yintercept = 0.5, 
-               linewidth = 1/4,
-               linetype = "dotted") +
+        size = 2
+    ) +
+    geom_vline(
+        xintercept = mean(responses$age),
+        linewidth = 1 / 4,
+        linetype = "dotted"
+    ) +
+    geom_hline(
+        yintercept = 0.5,
+        linewidth = 1 / 4,
+        linetype = "dotted"
+    ) +
     ggdist::geom_lineribbon(aes(order = lv),
-                               size = 0.5) +
-    labs(x = "Age (months)",
+        size = 0.5
+    ) +
+    labs(
+        x = "Age (months)",
         y = "Probability of acquisition",
@@
         fill = "Cognateness (phonological similarity)",
-         linetype = "Cognateness (phonological similarity)") +
+        linetype = "Cognateness (phonological similarity)"
+    ) +
     scale_linetype_manual(values = rev(c("solid", "dashed", "dotdash"))) +
@@
     scale_fill_grey(start = 0.85, end = 0.35) +
-    scale_y_continuous(breaks = seq(0, 1, 0.25),
-                       limits = c(0, 1)) +
+    scale_y_continuous(
+        breaks = seq(0, 1, 0.25),
+        limits = c(0, 1)
+    ) +
     scale_x_continuous(breaks = seq(0, 50, 5)) +
-    theme(legend.position = "top",
+    theme(
+        legend.position = "top",
         legend.box = "vertical",
@@
         panel.grid = element_blank(),
-          panel.border = element_rect(fill = NA, 
+        panel.border = element_rect(
+            fill = NA,
             colour = "black",
-                                      linewidth = 0.75))
+            linewidth = 0.75
+        )
+    )
 ```
@@
 The resulting value provided an estimate of the number of times the syllable appears in child-directed speech, embedded within different words.
-Finally, for each word, we summed the frequency of its syllables, as an estimate of the syllabic frequency of the word.
+Finally, for each word-form, we summed the frequency of its syllables, as an estimate of the syllabic frequency of the word-form.
 We fit a Bayesian model with *Cognateness* as response variable, and the main effects of syllable frequency and number of syllables (to control for the fact words with more syllables are more likely to score higher in syllabic frequency) as predictors.
@@
         include.lowest = TRUE,
-                        labels = c("0-10%", "10-20%", "20-30%", "30-40%",
+        labels = c(
+            "0-10%", "10-20%", "20-30%", "30-40%",
             "40-50%", "50-60%", "60-70%", "70-80%",
-                                   "80-90%", "90-100%"))) |>
+            "80-90%", "90-100%"
+        )
+    )) |>
     add_count(name = "n_total") |>
     count(lv_bin, n_total) |>
-    mutate(prop = n/n_total)
+    mutate(prop = n / n_total)
 
 E_e <- bvq_data$pool |>
-    mutate(item = gsub("cat_|spa_", "", item),
-           xsampa = gsub("\\.", "", xsampa)) |>
-    dplyr::filter(te %in% items$te, 
-                  item %in% items$item) |>
+    mutate(
+        item = gsub("cat_|spa_", "", item),
+        xsampa = gsub("\\.", "", xsampa)
+    ) |>
+    dplyr::filter(
+        te %in% items$te,
+        item %in% items$item
+    ) |>
     select(te, language, xsampa) |>
-    pivot_wider(names_from = language,
+    pivot_wider(
+        names_from = language,
         values_from = xsampa,
         id_cols = te,
-                names_repair = janitor::make_clean_names) |>
-    dplyr::filter(grepl("E", catalan), 
-                  grepl("e", spanish)) |>
+        names_repair = janitor::make_clean_names
+    ) |>
+    dplyr::filter(
+        grepl("E", catalan),
+        grepl("e", spanish)
+    ) |>
     left_join(distinct(items, te, lv),
-              by = "te")
+        by = "te"
+    )
 ```
 
-This study investigated the impact of cognateness (i.e., phonological similarity between each word-form and its translation) on the early bilingual lexicon.
+This study investigated the impact of cognateness (i.e., phonological similarity between translation equivalents) on the early bilingual lexicon.
 We used Bayesian Item Response Theory to model the acquisition trajectories of a large sample of Catalan and Spanish words, estimating the effect of cognateness on the probability of acquisition.
-This model corrected for participants' age, word-form length (number of phonemes), and a novel measure of participants' exposure rate to each word.
+This model corrected for participants' age, word-form length (number of phonemes), and a novel measure of participants' exposure rate to each word-form.
 Exposure rates were calculated as a language exposure-weighted lexical frequency score in which each word-form's lexical frequency was corrected by the degree to which the participant was exposed to each language.
@@
 This effect was mediated by exposure rate: low-exposure word-forms benefited from their cognate status, whereas high-exposure word-forms did not.
-Capitalising on accumulator models of language acquisition, we provide a theoretical account of bilingual lexical acquisition.
-In this account, parallel activation of the two languages plays a central role during the acquisition of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
+Using the concept of accumulator [see @kachergis2022standard for review], we provide a theoretical account of bilingual lexical acquisition.
+In the present account, parallel activation of the two languages plays a central role during the acquisition of early representations in the bilingual lexicon, and in which the dynamics of co-activation between translation equivalents results in an earlier age-of-acquisition.
 
@@
 When a bilingual child is exposed to a word-form, they activate not only its corresponding lexical representation, but also the lexical representation of its translation.
-The amount of co-activation that spreads from the spoken word to its translation is proportional to the amount of phonological similarity between both word-forms.
+The amount of co-activation that spreads from the spoken word-form to its translation is proportional to the amount of phonological similarity between both word-forms.
 Cognates would receive more activation from their translation than non-cognates, leading children to accumulate learning instances with cognate words at a faster rate than with non-cognate words.
@@
 In the context of lexical acquisition, the former scenario predicts that every learning instance for a given word-form contributes to the acquisition of the representation of such word in the lexicon, while the acquisition of its translation remains unaffected by such experience.
-In the latter scenario, a learning instance to the same word-form would contribute not only to the acquisition of the representation of such word, but also, to some extent, to the acquisition of its translation.
-Our findings provide strong support for an account of bilingual vocabulary growth in which the experience and learning outcomes accumulated by the child in one language impact those in the other language.
-Such a facilitatory cross-language mechanism might be an important piece in the puzzle of bilingual language acquisition.
+In the latter scenario, a learning instance to the same word-form would contribute not only to the acquisition of the representation of the word, but also, to some extent, to the acquisition of its translation.
+Our findings provide strong support for an account of bilingual vocabulary growth in which the experience and learning outcomes accumulated by the child in one language impact those in the other language through cross-language phonological associations.
+Such a facilitatory mechanism might be an important piece in the puzzle of bilingual language acquisition.
 In particular, it may shed some light on why bilingual infants do not show relevant delays in language acquisition milestones compared to their monolingual peers, while receiving a reduced quantity of speech input in each of their languages.
-Our results provide some insights into this issue: infants benefited more strongly from the cognateness facilitation effect when acquiring words from the language of lower exposure than in the language of higher exposure.
-
-We suggest that this asymmetry is the result of children' unbalanced exposure to their languages.
-A bilingual child's frequency of exposure to a given word-form is mostly determined by two factors: the word-form's lexical frequency and the child's amount of language exposure to the language such word belongs to.
-A dual linguistic input means lower exposure to each of the languages, unless one makes the---arguably implausible---assumption that bilinguals are exposed on average to twice the amount of linguistic input than monolinguals.
-Because of this difference in exposure, words lower-exposure language might receive activation from their translation in the higher-exposure language more often than words from the higher-exposure language receive activation from their translation in the lower-exposure language.
-As a result, the cognateness facilitation effect should be stronger in words from the lower-exposure language.
+Infants in the present study benefited more strongly from the cognateness facilitation effect when acquiring words from the language of lower exposure than in the language of higher exposure.
 
@@
 The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in toddlers and adults.
-For instance, unbalanced (or low-proficiency) bilinguals benefit from forward priming (dominant to non-dominant) cross-language phonological priming during word processing [@vonholzen2019impact; @singh2014one; @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked; but see @jardak2019labels].
-One the other hand, backward priming (non-dominant to dominant) seems less robust and a more challenging effect to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language; but see @duyck2009translation; @schoonbaert2009semantic].
+For instance, unbalanced (or low-proficiency) bilinguals benefit from cross-language forward priming (dominant to non-dominant) during word processing [@vonholzen2019impact; @singh2014one; @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked; but see @jardak2019labels].
+One the other hand, backward priming (non-dominant to dominant) seems less robust and more challenging to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language; but see @duyck2009translation; @schoonbaert2009semantic].
 Balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
 These results have been taken as evidence for an asymmetry in the strength of forward and backward connections in the unbalanced bilingual lexicon.
-Although implemented in different ways, or found under different assumptions, a dominance-mediated asymmetry is accounted for by previous models of adult bilingual lexical access, like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
+Although implemented in different ways, or found under different assumptions, such a dominance-mediated asymmetry is accounted for by multiple models of lexical processing like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
 Overall, this provides an apparently convenient account for the interaction between language dominance and cognateness found in the present study.
@@
 It might be argued that our results reflect the fact that cognate translation equivalents are represented in the initial bilingual lexicon as the *same* lexical entry.
-Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the same situations), it is possible that infants classify both are as acceptable variations of the same word, therefore treating them as a single lexical item.
+Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the same situations), it is possible that infants classify both are as acceptable variations of the same word-form, therefore treating them as a single lexical item.
 This would lead to a faster increase in cumulative learning instances, and to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
@@
 To conclude, our study provides novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
-We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word's phonological similarity with its translation in the other language.
-Capitalising on accumulator models of language acquisition we put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
+We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word-forms's phonological similarity with its translation in the other language.
+Capitalising on the notion of *accumulator* of linguistic input, we put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.
 

### 2023-08-11 _targets.R

@@ -51,37 +51,37 @@ options(
 
 list(
     ## import data -------------------------------------------------------------
-    
+
     tar_target(bvq_data_file, "data-raw/bvq.rds", format = "rds"),
     tar_target(bvq_data, readRDS(bvq_data_file)),
-    
+
     # get CHILDES frequencies
     tar_target(childes, get_childes_frequencies(age_range = c(12, 32))),
-    
+
     # items
     tar_target(items, get_items(
         bvq_data = bvq_data,
         childes = childes
     )),
     tar_target(items_test, test_items(items)),
-    
+
     # participants
     tar_target(
         participants,
         get_participants(bvq_data,
-                         age = c(12, 32),
-                         lp = c("Monolingual", "Bilingual"),
-                         other_threshold = 0.1
+            age = c(12, 32),
+            lp = c("Monolingual", "Bilingual"),
+            other_threshold = 0.1
         )
     ),
     tar_target(participants_test, test_participants(participants)),
-    
+
     # responses
     tar_target(responses, get_responses(bvq_data, items, participants)),
     tar_target(responses_test, test_responses(responses)),
-    
+
     # fit models ---------------------------------------------------------------
-    
+
     # model prior
     tar_target(
         model_prior,
@@ -93,7 +93,7 @@ list(
             prior(lkj(2), class = "cor")
         )
     ),
-    
+
     # multilevel model with crossed random effects (participants an items)
     # responses are generated from a categorical distribution:
     #   - https://journals.sagepub.com/doi/full/10.1177/2515245918823199
@@ -101,7 +101,7 @@ list(
     #   - https://bookdown.org/content/3686/ordinal-predicted-variable.html
     # the probability of each response category is adjusted by age (population-level effect)
     # and adjusted for each individual participant and item (group-level effects)
-    
+
     # only intercepts (category boundaries)
     tar_target(
         model_formula,
@@ -112,7 +112,7 @@ list(
             family = cumulative("logit")
         )
     ),
-    
+
     # add age:exposure interaction
     tar_target(
         model_fit,
@@ -124,7 +124,7 @@ list(
             sample_prior = "yes"
         )
     ),
-    
+
     # model with only prior samples
     tar_target(
         model_fit_prior,
@@ -136,38 +136,31 @@ list(
             sample_prior = "only"
         )
     ),
-    
+
     ## describe models ---------------------------------------------------------
-    
+
     tar_target(model_vars_dict, get_vars_dict(responses)),
-    
+
[…]

### 2023-08-11 docs/_repro.qmd
Fix URLs

--- 2023-08-11 _repro.qmd
@@
 A Docker image is just a virtual machine that contains all dependencies, code, and files necessary to run the code.
-Everything is already installed and setup, you will only have to open [https://localhost:8787](https://localhost:8787) in your browser, and a RStudio session will open for you.
+Everything is already installed and setup, you will only have to open [http://localhost:8787](http://localhost:8787) in your browser, and a RStudio session will open for you.
 This option is more robust (dependencies are difficult to get right when using your own computer), but it also requires Docker to be installed.

### 2023-08-12 manuscript/manuscript.qmd
Add favicon to site

@@
 @hoff2012dual found that English-Spanish bilingual toddlers in South Florida (United States) knew less words in English than monolinguals did, but both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together.
-Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual].
+Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @gonzalez-barrero2020bilingual].
 A more detailed analysis of the words in bilinguals’ lexicons shows some interesting patterns.
@@
 The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language.
-The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh.
+The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi or Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh.
 The authors calculated the average lexical overlap between the words in each of these additional languages and their translation equivalents in English.
@@
 The parallel activation hypothesis stems from the language non-selective account of lexical access, which suggests that bilinguals activate both languages simultaneously during language processing, even in fully monolingual contexts.
-Evidence with adult bilinguals supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @vanhell2008sentence; @hoshino2008cognate; @morford2011when; @schwartz2007reading; @spivey1999cross; @shook2012bimodal; see @kroll2017bilingual for review].
+Evidence with adult bilinguals supporting the language-non selective account of lexical access has been reported for language comprehension and production, across the auditory and visual (reading and signing) modalities [@gimeno-martinez2021crosslinguistic; @hoshino2008cognate; @morford2011when; @spivey1999cross; @shook2012bimodal; see @kroll2017bilingual for review].
 One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate.
@@
 These results showed that bilinguals’ Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-Evidence of parallel activation has been reported in bilingual toddlers and children [@vonholzen2012language; @floccia2020translation; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @schroter2016orthographic; @poulin-dubois2013lexical].
+Evidence of parallel activation has been reported in bilingual toddlers and children [@vonholzen2012language; @floccia2020translation; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @poulin-dubois2013lexical].
 
@@
 
-Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}-\textipa{"ga.to}/ [*cat*], which are phonologically very similar.
+Consider the example of the Catalan-Spanish cognate translation equivalent /\textipa{"gat}/--/\textipa{"ga.to}/ [*cat*], which are phonologically very similar.
 When the child listens to /\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in parallel.
@@
 Given the low phonological similarity between both word-forms, an exposure to \textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/ leading to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for /\textipa{"pe.ro}/.
-While /\textipa{"gat}-\textipa{"ga.to}/ will benefit from phonological co-activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not.
+While /\textipa{"gat}/--/\textipa{"ga.to}/ will benefit from phonological co-activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not.
 If the child receives linguistic input from one of the languages more often than from the other, this effect might affect each form of the cognate translation equivalent differently.
@@
 
-To test these predictions, we collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using the Barcelona Vocabulary Questionnaire (BVQ) [@garcia-castro2023bvq], a dual language online vocabulary checklist designed for the present study.
-We adopted a Bayesian explanatory item response theory approach [IRT, see @kachergis2022standard, for a similar approach] to model the probability of acquisition of `r nrow(distinct(items, language, item))` Catalan and Spanish nouns included in the vocabulary checklist.
+To test these predictions, we collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children.
+We adopted a Bayesian explanatory item response theory approach to model the probability of acquisition of `r nrow(distinct(items, language, item))` Catalan and Spanish nouns included in the vocabulary checklist.
 Words were considered as acquired if caregivers reported such word to be understood (comprehension) or understood and said (production) by their child.
@@
 Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (OCDI) [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
-Both lists included items from a diverse sample of `r n_categories` semantic/functional categories.
+Both lists included items from a diverse sample of `r n_categories` semantic or functional categories.
 The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`.
@@
 Each subset contained a random but representative sub-sample of the items from the complete list (see @tbl-items).
-Semantic/functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Semantic or functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
 Items that were part of the trial lists of some ongoing experiments in the lab were also included in all versions.
@@
 Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
-Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the Tidyverse family of packages [@wickham2019welcome].
+Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the *tidyverse* family of packages [@wickham2019welcome].
 
@@
 For each translation equivalent, we used the `stringdist` [@vanderloo2014stringdist] R package to calculate the Levenshtein distance between the Catalan and the Spanish phonological transcriptions of the word-forms.
-The Levenshtein distance measures the number of editions (insertions, deletions, or substitutions) that one string of phonemes/characters must go through to become identical to the other [@levenshtein1966binary].
+The Levenshtein distance measures the number of editions (insertions, deletions, or substitutions) that one string of characters must go through to become identical to the other [@levenshtein1966binary].
 We divided the Levenshtein distance of each translation equivalent by the length of the longest word-form to correct for word-form length (longer strings are likely to show a larger number of mismatches).
 Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
-This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}/--/\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}/--/\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; @laing2022phonological; for similar approaches].
+This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}/--/\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}/--/\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; for similar approaches].
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
@@
 We ran four iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
-Model posterior draws and predictions were handled using the `tidybayes` [@kay2021tidybayes] R package.
 
 # Results {#sec-results}
-
 ```{r results-values}
@@
 The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in toddlers and adults.
-For instance, unbalanced (or low-proficiency) bilinguals benefit from cross-language forward priming (dominant to non-dominant) during word processing [@vonholzen2019impact; @singh2014one; @grainger1998masked; @voga2007cognate; @de1991lexical; @shook2019covert; @dimitropoulou2011masked; but see @jardak2019labels].
-One the other hand, backward priming (non-dominant to dominant) seems less robust and more challenging to detect [e.g., @finkbeiner2004role; @hoshino2010erp; @midgley2009language; but see @duyck2009translation; @schoonbaert2009semantic].
+For instance, unbalanced (or low-proficiency) bilinguals benefit from cross-language forward priming (dominant to non-dominant) during word processing [@vonholzen2019impact; @singh2014one; @grainger1998masked; @de1991lexical; @shook2019covert; but see @jardak2019labels].
+One the other hand, backward priming (non-dominant to dominant) seems less robust and more challenging to detect [e.g., @hoshino2010erp; @midgley2009language; but see @duyck2009translation].
 Balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked].
 These results have been taken as evidence for an asymmetry in the strength of forward and backward connections in the unbalanced bilingual lexicon.
-Although implemented in different ways, or found under different assumptions, such a dominance-mediated asymmetry is accounted for by multiple models of lexical processing like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
+Although implemented in different ways, or found under different assumptions, such a dominance-mediated asymmetry is accounted for by multiple models of lexical processing like the Revised Hierarchical Model [@kroll1994category], BIA/BIA+ [@dijkstra2013bia], BLINCS [@shook2013bilingual], or Multilink [@dijkstra2019multilink], and also by models providing a more development-oriented perspective, like the Ontogenic Model [@bordag2022ontogenesis; @cook2016fuzzy], and BIA-d [@grainger2010re].
 Overall, this provides an apparently convenient account for the interaction between language dominance and cognateness found in the present study.
@@
 This bias would especially affect similar-sounding words, i.e., cognates.
-Production estimates may be more prompt to such biases, in part because of the slower pace at which infants’ articulatory abilities develop, compared to their word recognition abilities [@hustad2020development; @hustad2021speech].
+Production estimates may be more prompt to such biases, in part because of the slower pace at which infants’ articulatory abilities develop, compared to their word recognition abilities [@hustad2021speech].
 This gap between comprehension and production is even larger in the less dominant language of bilingual children [@giguere2022bilingual].
 For this reason, caregivers may be more uncertain about what words can be counted as *acquired* in this modality.
-Despite such potential biases, vocabulary checklist filled by parents show strong evidence of concurrent validity with other estimates of vocabulary size or lexical processing [@killing2008move; @feldman2005concurrent; @gillen2021tapping; but see @houston-price2007discrepancy].
+Despite such potential biases, vocabulary checklist filled by parents show strong evidence of concurrent validity with other estimates of vocabulary size or lexical processing [@feldman2005concurrent; @gillen2021tapping; but see @houston-price2007discrepancy].
 

### 2023-08-12 manuscript/manuscript.qmd
Fix coefficients table

@@
 For instance, the Spanish translation equivalent of *cat* is *gato*, a cognate word; the translation equivalent of *dog* is *perro*, a non-cognate word.
-The differences in the percentages of cognate and non-cognate words between two languages is related to historical reasons: languages typologically close (like Dutch and English or Italian and Spanish) share more cognates than languages typologically distant (like English and Chinese, or Urdu and Spanish).
+For historical reasons, some pairs of languages share more cognates than others: languages typologically close (like Dutch and English or Italian and Spanish) share more cognates than languages typologically distant (like English and Chinese, or Urdu and Spanish).
 The conclusion that cognate words are easier to learn is based on two types of evidence: studies investigating vocabulary sizes in children learning language pairs with different percentages of cognates (that is, differing in their typological distance) and studies comparing the number of cognate and non-cognate words children know in a specific language pair.
@@
 In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish.
-In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of *table*), obviously, such distinction was only relevant for bilinguals.
+In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of *table*).
+Obviously, such distinction was only relevant for bilinguals.
 Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items.
@@
 These results showed that bilinguals’ Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-Evidence of parallel activation has been reported in bilingual toddlers and children [@vonholzen2012language; @floccia2020translation; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @poulin-dubois2013lexical].
-
+More recntly, evidence of parallel activation has been reported in bilingual toddlers and children too [@vonholzen2012language; @floccia2020translation; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @poulin-dubois2013lexical].
 Although there is a consensus on the role of parallel activation in bilinguals' lexical processing and acquisition, previous studies do not address its influence on the learning trajectories of words.
@@
 We do not assume if a learning instance is a discrete or a continuous unit of accumulation of information.
-We consider that a learning instance of a word consists of an exposure to the its phonological form if the resulting strength of activation of its phonological representation in the child's lexicon reaches some thoretical threshold that leads to word-form recognition.
+We consider that a learning instance of a word is an exposure to its (phonological) form if the resulting strength of activation of its representation in the child's lexicon reaches some thoretical threshold that leads to word-form recognition.
 This activation may result from the infant being exposed to the actual word-form, or the result of activation spreading through phonological or semantic links across lexical representation, as in the case of parallel activation.
@@
 To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire (BVQ) [@garcia-castro2023bvq].
-This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formr platform [@arslan2020formr].
+This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the `formr` platform [@arslan2020formr].
 This questionnaire is structured in three blocks: (1) a language exposure questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory (OCDI) [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
 Both lists included items from a diverse sample of `r n_categories` semantic or functional categories.
@@
 For each word included in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items.
+Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items: A, B, C, and D.
 Each subset contained a random but representative sub-sample of the items from the complete list (see @tbl-items).
-Semantic or functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists.
+Semantic or functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided: all of their items were included in the four lists.
 Items that were part of the trial lists of some ongoing experiments in the lab were also included in all versions.
@@
 
-We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between the `r format(dates[1], "%dth of %B, %Y")` and the `r format(dates[2], "%dth of %B, %Y")`: `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
-Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`, and were always allocated to the same questionnaire list (A, B, C, or D).
+We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between Marth 30th, 2020 and October 31th, 2022: `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
+Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)` days, and were always allocated to the same questionnaire list (A, B, C, or D).
 Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra) and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample).
-In total, `r n_trilingual$n[2]` participants (`r scales::percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed to a third language other than Catalan and Spanish.
+In total, `r n_trilingual$n[2]` participants (`r scales::percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed in less than 10% to a third language other than Catalan and Spanish.
 All families provided informed consent before participating.
@@
 Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together.
-Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`] using the *tidyverse* family of packages [@wickham2019welcome].
+Data processing and visualisation was done in R [@rcoreteam2013language, version `r r_version`].
 
@@
 
+We implemented the model using `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
+We ran four iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
+
 ### Model predictors {#sec-predictors}
 
-We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-form's lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to, we expressed lexical frequencies as the product between both variables.
+We developed the *Exposure* predictor to account for the fact that bilinguals' exposure to a given word-form is not only a function of the word-form's lexical frequency, but also of the quantitative input they receive from the language such word-form belongs to.
+We expressed lexical frequencies as the product between both variables.
 First, we extracted the child-directed lexical frequency of each word-form from the CHILDES database [@macwhinney2000childes; @sanchez2019childesdb].
@@
 Finally, we subtracted the result from one so that it could be interpreted in terms of phonological similarity, instead of phonological distance.
-This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}/--/\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}/--/\textipa{"mar}/, *sea*) [see @floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth; for similar approaches].
+This led to a distance metric that ranged from zero to one, where zero indicates that both word-forms are completely different (e.g., /\textipa{"taw.l5}/--/\textipa{"me.sa}/, *table*), and one indicates that the two word-forms are identical (e.g., /\textipa{"mar}/--/\textipa{"mar}/, *sea*) [@floccia2018introduction; @heeringa2003norwegian; @schepens2012distributions; @fourtassi2020growth].
 Predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor.
@@
 
-We implemented the model using `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
-We ran four iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
-
 # Results {#sec-results}
@@
     mutate(across(
-        .value:.lower,
+        .value:.upper,
         \(x) ifelse(grepl("Intercept", .type),
@@
 All of the posterior samples of this regression coefficient excluded the ROPE.
-The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.value/sd(responses$exposure), 3)` more likely to acquire it.
+The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r round(post_draws_list[["b_exposure_std"]]$.value/sd(responses$exposure), 3)` probability points more likely to acquire it.
 Word-form length also showed a significant association with probability of acquisition ($\beta$ = `r round(post_draws_list[["b_n_phon_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_n_phon_std"]]$.lower, 3)`, `r round(post_draws_list[["b_n_phon_std"]]$.upper, 3)`]).
-For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.value)/sd(responses$n_phon), 3)` less likely to know it.
+For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.value)/sd(responses$n_phon), 3)` probability points less likely to know it.
 The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word-form exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children.
 
-The posterior distribution of the main effect of cognateness excluded the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 3)`]).
-For every 10% increment in cognateness, the acquisition of a word increased in `r round(post_draws_list[["b_lv_std"]]$.value/(40*sd(responses$lv)), 3)`.
+Around `r scales::percent(post_draws_list[["b_lv_std"]]$.rope, accuracy = 0.01)` of the posterior samples of the main effect of *Cognateness* overlapped with the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 3)`]).
+For every 10% increment in cognateness, the probability of word acquisition increased in `r round(post_draws_list[["b_lv_std"]]$.value/(40*sd(responses$lv)), 3)`.
 The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to the word-form.
-Follow-up analyses on this interaction showed that when exposure rate was low (e.g., -1 SD), cognateness increased the probability of acquisition substantially.
-This effect was negligible for words with median or high exposure (+1 SD) (see @fig-marginal).
+When *Exposure* was low (e.g., -1 SD), *Cognateness* increased the probability of acquisition substantially, while this effect was negligible for words with median or high exposure (+1 SD) (see @fig-marginal).
 
@@
 This study investigated the impact of cognateness (i.e., phonological similarity between translation equivalents) on the early bilingual lexicon.
-We used Bayesian Item Response Theory to model the acquisition trajectories of a large sample of Catalan and Spanish words, estimating the effect of cognateness on the probability of acquisition.
+We used Bayesian item response theory to model the acquisition trajectories of a large sample of Catalan and Spanish words, estimating the effect of cognateness on the probability of acquisition.
 This model corrected for participants' age, word-form length (number of phonemes), and a novel measure of participants' exposure rate to each word-form.

### 2023-08-13 manuscript/_appendix/_model-diagnostics.qmd
Fix IC's email, add again the model diagnostics appendix, and add notes in formula

@@
 #| label: fig-rhats-neffs
-#| fig-cap: "MCMC convergence diagnostic of all parameters in the model. A: distribution of the Gelman-Rubin (R-hat) scores. B: distribution of the ratio of effective sa"
+#| fig-cap: "MCMC convergence diagnostic of all parameters in the model.
+Each dot represents the score of one parameter.
+(A) Distribution of the Gelman-Rubin (R-hat) scores.
+(B) Distribution of the ratio of effective sample size."
 #| echo: false
@@
     ggplot(aes(.rhat)) +
-    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
+    geom_dots(colour = "black") +
     geom_vline(xintercept = 1.01, linetype = "dashed") +
-    labs(x = "R-hat",
-         y = "MCMC samples") +
-
+    labs(
+        x = "R-hat",
+        y = "MCMC samples"
+    ) +
     model_convergence |> 
     ggplot(aes(.neff)) +
-    geom_histogram(binwidth = 0.1, colour = "white", fill = "grey60") +
+    geom_dots(colour = "black") +
     geom_vline(xintercept = 1.01, linetype = "dashed") +
-    labs(x = "Effective sample size ratio",
-         y = "MCMC samples") +
-
+    labs(
+        x = "Effective sample size ratio",
+        y = "MCMC samples"
+    ) +    
     plot_layout(ncol = 1) &
     plot_annotation(tag_levels = "A") &
-    theme(legend.position = "none",
-          panel.grid = element_blank())
+    theme(
+        legend.position = "none",
+        panel.grid = element_blank()
+    )
 ```
@@
 #| message: false
-titles <- c("Comprehension",
+titles <- c(
+    "Comprehension",
     "Production",
@@
     "Exposure × Cognateness",
-            "Age × Exposure × Cognateness")
+    "Age × Exposure × Cognateness"
+)
 
@@
     ggplot(data = data, mapping = mapping) +
-        stat_slab(colour = "white",
-                  fill = "grey15")
+        stat_slab(
+            colour = "white",
+            fill = "grey15"
+        )
 }
@@
             geom = "raster",
-                        contour = FALSE) +
-        scale_fill_distiller(palette = "Greys",
-                             direction = 1) 
+            contour = FALSE
+        ) +
+        scale_fill_distiller(
+            palette = "Greys",
+            direction = 1
+        )
 }
 
-posterior %>%
+posterior |> 
     GGally::ggpairs(
@@
         lower = NULL,
-        columnLabels = c("Intercept\n(Comprehension)",
+        columnLabels = c(
+            "Intercept\n(Comprehension)",
             "Intercept\n(Production)",
@@
             "Age × Levenshtein", "Exposure × Leveshtein",
-                         "Age × Exposure ×\nLevenshtein"),
+            "Age × Exposure ×\nLevenshtein"
+        ),
         axisLabels = "none"
     ) +
-    theme(strip.text = element_text(size = 5))
+    theme(
+        strip.text = element_text(size = 5),
+        panel.grid = element_blank(),
+        panel.border = element_blank()
+    )
 
@@
 #| out-height: 80%
-y_int <- as.integer(as.integer(responses$response))
+y_int <- as.integer(as.integer(model_fit$data$response))
 bayesplot::color_scheme_set(c(rep("grey60", 2), rep("black", 4)))
@@
 ) +
-    scale_x_continuous(breaks = 1:3,
-                       labels = c("No",
+    scale_x_continuous(
+        breaks = 1:3,
+        labels = c(
+            "No",
             "Understands",
-                                  "Understands and Says")) +
-    theme(legend.position = "none",
-          axis.text = element_text(size = 11))
+            "Understands and Says"
+        )
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text = element_text(size = 11),
+        panel.grid = element_blank()
+    )
 ```

### 2023-08-13 manuscript/appendix.qmd
Fix IC's email, add again the model diagnostics appendix, and add notes in formula

@@
 library(ggplot2)
+library(ggdist)
+library(patchwork)
 
-targets::tar_config_set(store = here::here("_targets"),
-                        script = here::here("_targets.R"))
+targets::tar_config_set(
+    store = here::here("_targets"),
+    script = here::here("_targets.R")
+)
 
@@
 
-targets::tar_load(c(bvq_data, childes, items, model_summary,
-                    participants, responses))
+targets::tar_load(c(
+    bvq_data, childes, items, model_summary,
+    participants, responses
+))
+
+targets::tar_load(c(model_fit, model_convergence, model_ppcs))
 
@@
 
-targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables,
-                    posterior_syllables_summary))
+targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables, posterior_syllables_summary))
 
 my_theme <- theme_minimal() +
-    theme(panel.grid = element_line(colour = "grey",
-                                    linetype = "dotted"),
+    theme(
+        panel.grid = element_line(
+            colour = "grey",
+            linetype = "dotted"
+        ),
         axis.line = element_line(colour = "black"),
         text = element_text(size = 12, colour = "black"),
-          axis.text = element_text(colour = "black"))
+        axis.text = element_text(colour = "black")
+    )
 
@@
 
-## Appendix A: frequency and language exposure as separate predictors {.appendix .unnumbered}
+## Appendix A: model diagnostics {.appendix .unnumbered}
+
+{{< include _appendix/_model-diagnostics.qmd >}}
+
+{{< pagebreak >}}
+
+## Appendix B: frequency and language exposure as separate predictors {.appendix .unnumbered}
 
@@
 
-## Appendix B: frequency and language exposure as separate predictors {.appendix .unnumbered}
+## Appendix C: frequency and language exposure as separate predictors {.appendix .unnumbered}
 

### 2023-08-13 manuscript/manuscript.qmd
Fix IC's email, add again the model diagnostics appendix, and add notes in formula

@@
 We found an interaction between exposure and cognateness, suggesting that cognateness facilitates the aquisition of low-exposure words, but not of mean exposure or high-exposure words.
-Overall, our findings suggest that cognateness plays a key role in bilingual lexical acquisition, and provides evidence for a frequency-mediated facilitation effect driven by parallel activation.
+Overall, our findings suggest that cognateness plays a key role in bilingual lexical acquisition, and provide evidence for a frequency-mediated facilitation effect driven by parallel activation.
 
@@
 &\text{where:}\\
-& u_{1-8_{i}}~\text{: participant-level adjustments} \\
-& w_{1-3_{j}}~\text{: TE-level adjustments} \\
+& \beta_{1-8}\text{: fixed effects}\\
+& u_{1-8_{i}}\text{: participant-level random effects} \\
+& w_{1-3_{j}}\text{: TE-level random effects} \\
 &\textbf{Prior} \\
@@
 \sigma_{u_{0-8}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2) \\
-&\text{where }~\rho_{u_{0-8}, w_{0-3}}~\text{are the correlations between group-level adjustments}\\
+&\text{where:}\\
+& \sigma_{u_{0-8}, w_{0-3}}\text{: are the standard deviations of }u\text{ and }w\\
+& \rho_{u_{0-8}, w_{0-3}}~\text{are the correlations between random effects in }u\text{ and }w\\
 \end{aligned}
@@
 
-Chain convergence diagnostics showed adequate values, and the model posterior showed little evidence of correlation, as indicated by negligible pairwise correlation between the marginal posterior distribution of coefficients.
+The model posterior showed adequate chain convergence diagnostics and posterior predictive checks (see Appendix A).
 @tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients, and their degree of overlap with the ROPE.
@@
 
-An additional analysis including lexical frequency and language exposure as separate predictors (instead of the composite *Exposure* measure) showed equivalent results (see Appendix A).
+An additional analysis including lexical frequency and language exposure as separate predictors (instead of the composite *Exposure* measure) showed equivalent results (see Appendix B).
 To rule out the possibility that cognateness facilitation effect we found was due to cognateness comprising more frequent syllables than non-cognates---and therefore not because of their cognate status itself---, we compared the syllabic frequency of cognates and non-cognates included in our analyses.
@@
 We fit a Bayesian model with *Cognateness* as response variable, and the main effects of syllable frequency and number of syllables (to control for the fact words with more syllables are more likely to score higher in syllabic frequency) as predictors.
-This model provided strong evidence for the association between cognateness and syllabic frequency being equivalent to zero (see Appendix B).
+This model provided strong evidence for the association between cognateness and syllabic frequency being equivalent to zero (see Appendix C).
 

### 2023-08-15 manuscript/manuscript.qmd
Fix affiliations and figure caption

@@
 Lines and error bands correspond to the mean and 95% credible interval of the posterior-predicted means.
-Different colours indicate different levels of cognateness (phonological similarity).
+Different shades of grey indicate different levels of cognateness (phonological similarity).
 Predictions are presented separately for different degrees of word-form exposure index: little exposure to the word-form, mean exposure, and high exposure).
 Predictions for Comprehension are shown on top and predictions for Comprehension and Production are shown on the bottom.
-In-sample predictions lie inside the grey rectangles."
+In-sample predictions lie inside the yellow rectangles.
+For reference, we indicate the 50% chance level of word acquisition (horizontal yellow lines), and the mean age of the sample (vertical yellow lines)."
 fig_data <- model_epreds |>
@@
         xintercept = mean(responses$age),
-        linewidth = 1 / 4,
-        linetype = "dotted"
+        linewidth = 2 / 4,
+        alpha = 0.5,
+        colour = "orange"
     ) +
@@
         yintercept = 0.5,
-        linewidth = 1 / 4,
-        linetype = "dotted"
+        linewidth = 2 / 4,
+        alpha = 0.5,
+        colour = "orange"
     ) +

### 2023-09-19 manuscript/manuscript.tex
Update models with 'on_change'

--- 2023-07-23 manuscript/manuscript.tex
@@
 authors declare no conflicts of interest with regard to the funding
-source of this study. GGC was supported by a FPI 2020 research contract
-(PRE2019-088165) in a project funded by the Spanish Ministry of Science,
-Innovation, and Universities (PGC2018-101831-B-I00). DAV was supported
-by the European Union's Horizon 2023 research and innovation program
-under Marie Skłodowska--Curie Grant (765556) and a postdoctoral
+source of this study. This study was supported by the Spanish Ministry
+for Science and Innovation and State Research Agency (Project PID2021-
+123416NB-I00 financed by MCIN/ AEI/ 10.13039/501100011033 / FEDER, UE)
+and the Economic and Social Research Council (ESRC) (ES/S010947/1, UK).
+GGC was supported by a FPI research contract (PRE2019-088165). DAV was
+supported by the European Union's Horizon 2023 research and innovation
+program under Marie Skłodowska--Curie Grant (765556) and a postdoctoral
 fellowship from the Foundation for Science and Technology of Portugal
@@
 the European Union's NextGenerationEU (NGEU) recovery plan. NSG was
-supported by Academia 2019 award from the Catalan Institution for
+supported by an ICREA Academia award from the Catalan Institution for
 Research and Advanced Studies (ICREA). We are grateful to Chiara
@@
 Bilingual infants' developmental trajectories of lexical acquisition are
-equivalent to those shown by their monolingual peers. This is
-remarkable, given the increased complexity of their linguistic input.
-Recent studies suggested that bilingual vocabulary growth is boosted by
-the number of cognates shared by the pair of languages being learned,
-and that this facilitation effect is driven by a stronger parallel
-activation of cognates during linguistic exposure, compared to
-non-cognates. The mechanisms behind this facilitation are still unclear.
-In this study, we capitalised on accumulator models of language
-acquisition to propose an account of bilingual lexical acquisition in
-which parallel activation increases the rate at which children
-accumulate learning instances for words in both languages, even in fully
-monolingual situations. Under this hypothesis, we predicted a stronger
+equivalent to their monolingual peers'. This is remarkable, given the
+complexity of their linguistic input. Recent studies suggest that
+bilingual vocabulary growth is boosted by the number of cognates
+(form-similar translation equivalents) shared by the pair of languages
+being learned, and that this cognateness facilitation effect is driven
+by a stronger parallel activation of cognates during linguistic
+exposure, compared to non-cognates. The mechanisms behind this
+facilitation are still unclear. In this study, we propose an account of
+bilingual lexical acquisition in which parallel activation increases the
+rate at which children accumulate learning instances for words in both
+languages, even in fully monolingual situations. We predicted a stronger
 cognate facilitation for words to which children were exposed less
 frequently (low-exposure words), as they are co-activated by their
-translation more often than high-exposure words do. We developed an
+translation more often than high-exposure words. We developed an
 extensive online vocabulary checklist, the Barcelona Vocabulary
 questionnaire (BVQ), to collect vocabulary data from 366 Catalan-Spanish
-bilingual toddlers aged 12 to 32 months. We then used Bayesian
-explanatory item response theory to model the acquisition trajectories
-of 302 Catalan/Spanish translation equivalents. We found an interaction
-between exposure and cognateness, which pointed to cognateness
-facilitating the aquisition of low-exposure words, but not of mean
-exposure or high-exposure words. Overall, our findings suggest that
-cognateness plays a key role in bilingual lexical acquisition, and
-provides evidence for a frequency-mediated facilitation effect driven by
-parallel activation.
+bilingual toddlers aged 12 to 32 months. We used Bayesian explanatory
+item response theory to model the acquisition trajectories of 604
+Catalan and Spanish words. We found an interaction between exposure and
+cognateness, suggesting that cognateness facilitates the aquisition of
+low-exposure words, but not of mean exposure or high-exposure words.
+Overall, our findings suggest that cognateness plays a key role in
+bilingual lexical acquisition, and provide evidence for a
+frequency-mediated facilitation effect driven by parallel activation.
 \end{abstract}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-The foundations of word learning are in place early in age. At six
-months, infants start directing their gaze to some objects when hearing
-their labels (\protect\hyperlink{ref-bergelson2012months}{Bergelson \&
+The foundations of word learning are in place at an early age. At six
+months, infants start directing their gaze to objects when hearing their
+labels (\protect\hyperlink{ref-bergelson2012months}{Bergelson \&
 Swingley, 2012}, \protect\hyperlink{ref-bergelson2015early}{2015};
@@
 2004}; \protect\hyperlink{ref-patterson2004bilingual}{Patterson \&
-Pearson, 2004}; \protect\hyperlink{ref-pearson1993lexical}{Pearson et
-al., 1993}; \protect\hyperlink{ref-pearson1994patterns}{Pearson \&
-Fernández, 1994}; \protect\hyperlink{ref-petitto2001bilingual}{Petitto
-et al., 2001}; \protect\hyperlink{ref-smithson2014bilingualism}{Smithson
-et al., 2014}). A more detailed analysis of the words in bilinguals'
-lexicon shows some interesting patterns.
+Pearson, 2004}; \protect\hyperlink{ref-pearson1994patterns}{Pearson \&
+Fernández, 1994};
+\protect\hyperlink{ref-smithson2014bilingualism}{Smithson et al.,
+2014}). A more detailed analysis of the words in bilinguals' lexicons
+shows some interesting patterns.
 
@@
 word; the translation equivalent of \emph{dog} is \emph{perro}, a
-non-cognate word. The differences in the percentages of cognate and
-non-cognate words between two languages is related to historical
-reasons: languages typologically close (like Dutch and English or
-Italian and Spanish) share more cognates than languages typologically
-distant (like English and Chinese, or Urdu and Spanish). The conclusion
-that cognate words are easier to learn is based on two types of
-evidence: studies investigating vocabulary sizes in children learning
-language pairs with different percentages of cognates (that is,
+non-cognate word. For historical reasons, some pairs of languages share
+more cognates than others: languages typologically close (like Dutch and
+English or Italian and Spanish) share more cognates than languages
+typologically distant (like English and Chinese, or Urdu and Spanish).
+The conclusion that cognate words are easier to learn is based on two
+types of evidence: studies investigating vocabulary sizes in children
+learning language pairs with different percentages of cognates (that is,
 differing in their typological distance) and studies comparing the
@@
 additional language was one of 13 typologically diverse languages:
-Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu,
+Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi or Urdu,
 Italian, Mandarin Chinese, Polish, Portuguese, Spanish, and Welsh. The
@@
 of these additional languages and their translation equivalents in
-English, which was taken as a proxy of the degree of cognateness between
-each pair of languages. Floccia and co-workers reported an increase in
-vocabulary size in the additional language (i.e., not English)
-associated with an increase in the average phonological similarity
-between the translation equivalents of each language pair. For example,
-English-Dutch bilinguals (languages with a high phonological overlap),
-were able to produce more Dutch words than English-Mandarin bilinguals
-(languages with a low phonological overlap) were able to produce in
-Mandarin. Blom et al.
+English. Lexical overlap was calculated in terms of phonological
+similarity (described below) and it was taken as a proxy of the degree
+of cognateness between each pair of languages. Floccia and co-workers
+reported an increase in vocabulary size in the additional language
+(i.e., not English) associated with an increase in the average
+phonological similarity between the translation equivalents of each
+language pair. For example, English-Dutch bilinguals (languages with a
+high phonological overlap), were able to produce more Dutch words than
+English-Mandarin bilinguals (languages with a low phonological overlap)
+were able to produce in Mandarin. Blom et al.
 (\protect\hyperlink{ref-blom2020crosslanguage}{2020}), Bosma et al.
@@
 18 months and found that cognates represented a larger proportion of
-participant's vocabulary than non-cognates. Schelletter
+vocabulary than non-cognates. Schelletter
 (\protect\hyperlink{ref-schelletter2002effect}{2002}) provided
@@
 learned. For instance, if two given languages share a high proportion of
-cognates like \(\sim 70\)\%, the vocabulary contents of children
-learning both languages should, in principle, approximate such
-proportion of cognates, even in the absence of a cognateness
-facilitation effect. More recently, Mitchell et al.
-(\protect\hyperlink{ref-mitchell2022cognates}{2022}) addressed this
-issue in a longitudinal study with a larger sample. The authors
-collected expressive vocabulary data of 47 16- to 30-month-old
-French-English bilinguals living in Canada, in both languages. They
-created two lists of translation equivalents: one made of 131 cognates,
-and one made of 406 non-cognates. The proportion of words that children
-were reportedly able to produce was higher in the cognate lists than in
-the non-cognate list across ages. Critically, this difference persisted
-after both lists were matched in size, controlling their semantic
-category (i.e., furniture, animals, food were similarity represented in
-both lists) and age-of-acquisition norms (an index of word difficulty).
-Taken together, the results of these two lines of research have been
-considered as providing support to the hypothesis that phonological
-similarity (as reflected in cognateness) plays a facilitation role in
-bilingual word acquisition.
+cognates like 70\%, the vocabulary contents of children learning both
+languages should, in principle, approximate such proportion of cognates,
+even in the absence of a cognateness facilitation effect. More recently,
+Mitchell et al. (\protect\hyperlink{ref-mitchell2022cognates}{2022})
+addressed this issue in a longitudinal study. The authors collected
+expressive vocabulary data of 47 16- to 30-month-old French-English
+bilinguals living in Canada, in both languages. They created two lists
+of translation equivalents: one made of 131 cognates, and one made of
+406 non-cognates. The proportion of words that children were reportedly
+able to produce was higher in the cognate lists than in the non-cognate
+list across ages. Critically, this difference persisted after both lists
+were matched in size, controlling their semantic category (i.e.,
+furniture, animals, food were similarity represented in both lists) and
+age-of-acquisition norms (an index of word difficulty). Taken together,
+the results of these two lines of research have been considered as
+supporting to the hypothesis that phonological similarity (as reflected
+in cognateness) plays a facilitation role in bilingual word acquisition.
 
@@
 (\protect\hyperlink{ref-gimeno-martinez2021crosslinguistic}{Gimeno-Martínez
-et al., 2021}; \protect\hyperlink{ref-vanhell2008sentence}{Hell \&
-Groot, 2008}; \protect\hyperlink{ref-hoshino2008cognate}{Hoshino \&
+et al., 2021}; \protect\hyperlink{ref-hoshino2008cognate}{Hoshino \&
 Kroll, 2008}; \protect\hyperlink{ref-morford2011when}{Morford et al.,
-2011}; \protect\hyperlink{ref-schwartz2007reading}{Schwartz et al.,
-2007}; \protect\hyperlink{ref-spivey1999cross}{Spivey \& Marian, 1999};
-see \protect\hyperlink{ref-kroll2017bilingual}{Kroll \& Ma, 2017} for
+2011}; \protect\hyperlink{ref-shook2012bimodal}{Shook \& Marian, 2012};
+\protect\hyperlink{ref-spivey1999cross}{Spivey \& Marian, 1999}; see
+\protect\hyperlink{ref-kroll2017bilingual}{Kroll \& Ma, 2017} for
 review). One of the clearest pieces of evidence of parallel activation
@@
 labels were non-cognates (\emph{mesa}-\emph{taula}, translations of
-\emph{table}), obviously, such distinction was only relevant for
+\emph{table}). Obviously, such distinction was only relevant for
 bilinguals. Bilinguals named cognate pictures faster than non-cognate
@@
 bilinguals' Catalan phonology was activated during the production of
-Spanish words, facilitating the naming of cognate pictures. Evidence of
-parallel activation has been reported in bilingual toddlers and children
+Spanish words, facilitating the naming of cognate pictures. More
+recntly, evidence of parallel activation has been reported in bilingual
+toddlers and children too
 (\protect\hyperlink{ref-bosma2020cognate}{Bosma \& Nota, 2020};
+\protect\hyperlink{ref-floccia2020translation}{Floccia et al., 2020};
 \protect\hyperlink{ref-poarch2012crosslanguage}{Poarch \& Hell, 2012};
 \protect\hyperlink{ref-poulin-dubois2013lexical}{Poulin-Dubois et al.,
-2013}; \protect\hyperlink{ref-schroter2016orthographic}{Schröter \&
-Schroeder, 2016}; \protect\hyperlink{ref-vonholzen2019impact}{Von Holzen
-et al., 2019}; \protect\hyperlink{ref-vonholzen2012language}{Von Holzen
-\& Mani, 2012}).
-
-Although there is a consensus on the role of parallel activation in
-bilinguals' lexical processing and acquisition, previous studies do not
-address its influence on the learning trajectories of words. Results are
-aggregated across words and provide no information about the specific
-dynamics of how parallel activation influences word learning. This is
-the goal of the present research.
+2013}; \protect\hyperlink{ref-vonholzen2019impact}{Von Holzen et al.,
+2019}; \protect\hyperlink{ref-vonholzen2012language}{Von Holzen \& Mani,
+2012}). Although there is a consensus on the role of parallel activation
+in bilinguals' lexical processing and acquisition, previous studies do
+not address its influence on the learning trajectories of words. Results
+are aggregated across words and provide no information about the
+specific dynamics of how parallel activation influences word learning.
+This is the goal of the present research.
 
@@
 represent a learning instance for its translation equivalent, to the
-extent that such translation equivalent is co-activated\footnote{We use
-  the term \emph{learning instance} in the fashion of accumulator models
-  of language acquisition; as an exposure to a word-form that
-  constitutes an opportunity for the child to accumulate information
-  about the word. We do not make strong assumptions about whether a
-  learning instance is a discrete or a continuous unit of accumulation
-  of information. We rather consider that a learning instance of a word
-  is so to the degree of the strength of activation of its phonological
-  representation. This activation may result from the infant being
-  exposed to the actual word-form, or the result of activation spreading
-  through phonological or semantic links across lexical representation,
-  as in the case of parallel activation.}. The strength of this
-co-activation is proportional to the phonological similarity between the
-two translation equivalents; given that cognates share higher
-phonological similarity than non-cognates, the former should be
-co-activated more strongly than the latter. This should lead to a faster
-accumulation of learning instances for cognates, compared to
-non-cognates. Parallel activation would allow bilingual children to
-accumulate learning instances for words in both languages even during
-fully monolingual situations, but the impact of this mechanism would be
-asymmetric across languages: words from the lower-exposure language
-would receive stronger activation from words in the higher-exposure
-language than vice versa. Therefore, the acquisition of words from the
-lower-exposure language would benefit more strongly from their cognate
-status than words from the higher-exposure language. This asymmetric
-cross-language activation would be consistent with previous reports of
-larger priming effects from the dominant to the non-dominant language,
+extent that such translation equivalent is co-activated. We use the term
+\emph{learning instance} in the fashion of accumulator models of
+language acquisition; as an exposure to a word-form that constitutes an
+opportunity for the child to accumulate information about the word. We
+do not assume if a learning instance is a discrete or a continuous unit
+of accumulation of information. We consider that a learning instance of
+a word is an exposure to its (phonological) form if the resulting
+strength of activation of its representation in the child's lexicon
+reaches some thoretical threshold that leads to word-form recognition.
+This activation may result from the infant being exposed to the actual
+word-form, or the result of activation spreading through phonological or
+semantic links across lexical representation, as in the case of parallel
+activation. The strength of this co-activation is proportional to the
+phonological similarity between the two translation equivalents; given
+that cognates share higher phonological similarity than non-cognates,
+the former should be co-activated more strongly than the latter. This
+should lead to a faster accumulation of learning instances for cognates,
+compared to non-cognates. Parallel activation would allow bilingual
+children to accumulate learning instances for words in both languages
+even during fully monolingual situations, but the impact of this
+mechanism would be asymmetric across languages: words from the
+lower-exposure language would receive stronger activation from words in
+the higher-exposure language than vice versa. Therefore, the acquisition
+of words from the lower-exposure language would benefit more strongly
+from their cognate status than words from the higher-exposure language.
+This asymmetric cross-language activation would be consistent with
+previous reports of larger priming effects from the dominant to the
+non-dominant language.
 
 Consider the example of the Catalan-Spanish cognate translation
-equivalent /\textipa{"gat}-\textipa{"ga.to}/ {[}\emph{cat}{]}, that are
-phonologically very similar. When the child is exposed to
+equivalent /\textipa{"gat}/--/\textipa{"ga.to}/ {[}\emph{cat}{]}, which
+are phonologically very similar. When the child listens to
 /\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in
-parallel. Therefore, this exposure will count as a learning instance for
-both co-activated forms. The case of the non-cognate translation
-equivalent /\textipa{"gos-"pe.ro}/ {[}\emph{dog}{]} would be different.
-Given the low phonological similarity between both word-forms, an
-exposure to \textipa{"gos} will result in a weak activation of
-/\textipa{"pe.ro}/ leading to such exposure counting as a learning
-instance for /\textipa{"gos}/ (which the child was exposed to), but not
-for /\textipa{"pe.ro}/. While /\textipa{"gat}-\textipa{"ga.to}/ will
-benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will
-not. If the child receives linguistic input from one of the languages
-more often than from the other, this effect might affect each form of
-the cognate translation equivalent differently. For instance, if the
-child receives a larger amount of Catalan input than Spanish input, they
-will encounter the Catalan form /\textipa{"gat}/ more frequently than
-the Spanish form /\textipa{"ga.to}/. Through parallel activation,
-/\textipa{"gat}/ will activate /\textipa{"ga.to}/ more often than vice
-versa. Ultimately, /\textipa{"ga.to}/ will benefit more strongly from
-its cognate status than /\textipa{"gat}/, as it receives additional
-learning instances from its translation equivalent more often than
-/\textipa{"gat}/.
+parallel. If the child has already formed a form-meaning association for
+both word-forms, parallel activation may result from the activation of
+their common concept or from activation spreading throught phonological
+similarity. We assume semantic co-activation to be constant across
+cognate and non-cognate translation equivalents, and focus on
+phonological co-activation as an additional source of activation that
+affects cognates more strongly than non-cognates. Therefore, this
+exposure will count as a learning instance for both co-activated forms.
+The case of the non-cognate translation equivalent
+/\textipa{"gos-"pe.ro}/ {[}\emph{dog}{]} would be different. Given the
+low phonological similarity between both word-forms, an exposure to
+\textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/
+leading to such exposure counting as a learning instance for
+/\textipa{"gos}/ (which the child was exposed to), but not for
+/\textipa{"pe.ro}/. While /\textipa{"gat}/--/\textipa{"ga.to}/ will
+benefit from phonological co-activation,
+/\textipa{"gos}-\textipa{"pe.ro}/ will not. If the child receives
+linguistic input from one of the languages more often than from the
+other, this effect might affect each form of the cognate translation
+equivalent differently. For instance, if the child receives a larger
+amount of Catalan input than Spanish input, they will encounter the
+Catalan form /\textipa{"gat}/ more frequently than the Spanish form
+/\textipa{"ga.to}/. Through parallel activation, /\textipa{"gat}/ will
+activate /\textipa{"ga.to}/ more often than vice versa. Ultimately,
+/\textipa{"ga.to}/ will benefit more strongly from its cognate status
+than /\textipa{"gat}/, as it receives additional learning instances from
+its translation equivalent more often than /\textipa{"gat}/.
 
@@
 and comprehension from a large sample of bilingual Catalan-Spanish
-children using the Barcelona Vocabulary Questionnaire (BVQ)
-(\protect\hyperlink{ref-garcia-castro2023bvq}{Garcia-Castro et al.,
-2023}), a dual language online vocabulary checklist designed for the
-present study. We adopted a Bayesian explanatory item response theory
-approach (IRT, see
-\protect\hyperlink{ref-kachergis2022standard}{Kachergis et al., 2022},
-for a similar approach) to model the probability of acquisition of 604
-Catalan and Spanish nouns included in the vocabulary checklist. Words
-were considered as acquired if caregivers reported such word to be
-understood (comprehension) or understood and said (production) by their
-child. We estimated the impact of several predictors of interest on the
+children. We adopted a Bayesian explanatory item response theory
+approach to model the probability of acquisition of 604 Catalan and
+Spanish nouns included in the vocabulary checklist. Words were
+considered as acquired if caregivers reported such word to be understood
+(comprehension) or understood and said (production) by their child. We
+estimated the impact of several predictors of interest on the
 probability of acquisition, including participants' age and rate of
@@
 lexical frequency. We predicted an interaction between cognate status
-and word exposure rate in which the probability of comprehension is
+and word-form exposure rate in which the probability of comprehension is
 higher for low-exposure cognate words, but not for high-exposure cognate
@@
 and its adaptations to other languages, and was implemented on-line
-using the formr platform (\protect\hyperlink{ref-arslan2020formr}{Arslan
-et al., 2020}). This questionnaire is structured in three blocks: (1) a
-language exposure questionnaire, (2) a demographic survey, and (3) two
-vocabulary checklists. Vocabulary checklists followed a similar
-structure as the Oxford Communicative Developmental Inventory (OCDI)
+using the \texttt{formr} platform
+(\protect\hyperlink{ref-arslan2020formr}{Arslan et al., 2020}). This
+questionnaire is structured in three blocks: (1) a language exposure
+questionnaire, (2) a demographic survey, and (3) two vocabulary
+checklists. Vocabulary checklists followed a similar structure as the
+Oxford Communicative Developmental Inventory
 (\protect\hyperlink{ref-hamilton2000infant}{Hamilton et al., 2000}) and
 consisted in two lists of words: one in Catalan and one in Spanish. Both
-lists included items from a diverse sample of 26 semantic/functional
+lists included items from a diverse sample of 26 semantic or functional
 categories. The Catalan checklist contained 793 items and the Spanish
@@
 participant responded to both \emph{gos} and \emph{perro}, Catalan and
-Spanish for \emph{dog}), roughly following a one-to-one
-mapping\footnote{Although for some of the included words in Catalan did
-  not have a clear translation or had more than one possible translation
-  in Spanish, and vice versa, therefore the unequal number of words
-  included in the Catalan and Spanish lists} (see Table~\ref{tbl-items}
-for a summary of the questionnaire items).
+Spanish for \emph{dog}), roughly following a one-to-one mapping. Some of
+the words in Catalan did not have a clear translation or had more than
+one possible translation in Spanish, and vice versa, therefore the
+unequal number of words included in the two lists (see
+Table~\ref{tbl-items} for a summary of the questionnaire items).
 
@@
 default). Given the large number of words in the vocabulary checklists,
-we created four different subsets of the complete list of items. Each
-subset contained a random but representative sub-sample of the items
-from the complete list (see Table~\ref{tbl-items}). Semantic/functional
-categories with less than 16 items---thus resulting in less than four
-items after dividing it in four lists---were not divided in the short
-version of the questionnaire: all of their items were included in the
-four lists. Items that were part of the trial lists of some ongoing
-experiments in the lab were also included in all versions. The resulting
-reduced list contained between 343 and 349 Catalan words, and between
-349 and 371 Spanish words. Participants were randomly allocated into one
-of the four subsets.
+we created four different subsets of the complete list of items: A, B,
+C, and D. Each subset contained a random but representative sub-sample
+of the items from the complete list (see Table~\ref{tbl-items}).
+Semantic or functional categories with less than 16 items---thus
+resulting in less than four items after dividing it in four lists---were
+not divided: all of their items were included in the four lists. Items
+that were part of the trial lists of some ongoing experiments in the lab
+were also included in all versions. The resulting reduced list contained
+between 343 and 349 Catalan words, and between 349 and 371 Spanish
+words. Participants were randomly allocated into one of the four
+subsets.
 
@@
 We collected 440 responses to the questionnaire from 369 distinct
-children from the Metropolitan Area of Barcelona between the 30th of
-marzo, 2020 and the 31th of octubre, 2022: 314 of those participants
-participated once, 43 twice, 8 three times, and 4 four times. Recurrent
-participants provided responses with a minimum of 25 days between
-responses, and a maximum of 527, and were always allocated to the same
-questionnaire list (A, B, C, or D). Participants were part of the
-database of the Laboratori de Recerca en Infància (Universitat Pompeu
-Fabra) and were contacted by e-mail or phone if their child was aged
-between 12 and 32 months, and had not been reported to be exposed more
-than 10\% of the time to a language other than Spanish or Catalan (see
+children from the Metropolitan Area of Barcelona between Marth 30th,
+2020 and October 31th, 2022: 314 of those participants participated
+once, 43 twice, 8 three times, and 4 four times. Recurrent participants
+provided responses with a minimum of 25 days between responses, and a
+maximum of 527 days, and were always allocated to the same questionnaire
+list (A, B, C, or D). Participants were part of the database of the
+Laboratori de Recerca en Infància (Universitat Pompeu Fabra) and were
+contacted by e-mail or phone if their child was aged between 12 and 32
+months, and had not been reported to be exposed more than 10\% of the
+time to a language other than Spanish or Catalan (see
 Table~\ref{tbl-participants} for a more detailed description of the
 sample). In total, 70 participants (15.91\%) participants were reported
-to be exposed to a third language other than Catalan and Spanish. All
-families provided informed consent before participating. Upon consent,
-families were sent a link to the questionnaire via e-mail, which they
-filled from a computer, laptop, or mobile device. Filling the
-questionnaire took 30 minutes approximately. After completion, families
-were rewarded with a token of appreciation.
+to be exposed in less than 10\% to a third language other than Catalan
+and Spanish. All families provided informed consent before
+participating. Upon consent, families were sent a link to the
+questionnaire via e-mail, which they filled from a computer, laptop, or
+mobile device. Filling the questionnaire took 30 minutes approximately.
+After completion, families were rewarded with a token of appreciation.
 
@@
 et al., 2020} for a similar approach). We then excluded items with
-missing lexical frequency scores (\emph{n} = 269), items that included
-more than one lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} =
-48), multi-word items or phrases (e.g., \emph{barrita de cereales}
-{[}cereal bar{]}, \emph{n} = 9). Finally, we removed items without a
-translation in the other language (\emph{n} = 32). This resulted in a
-final list of 604 items, corresponding to 302 Catalan words and their
-302 Spanish translations (302 translation equivalents). After collecting
-participants' responses, the final dataset consisted of 139,326
-observations, each corresponding to a single response of one participant
-to one item. Each translation equivalent received a median of 234
-responses (\emph{Min} = 108, \emph{Max} = 880) from participants, both
-languages pooled together.
-
+missing lexical frequency scores (\emph{n} = 269, see
+Section~\ref{sec-predictors}), items that included more than one lemma
+(e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} = 48), multi-word items
+or phrases (e.g., \emph{barrita de cereales} {[}cereal bar{]}, \emph{n}
+= 9). Finally, we removed items without a translation in the other
+language (\emph{n} = 32). This resulted in a final list of 604 items,
+corresponding to 302 Catalan words and their 302 Spanish translations
+(302 translation equivalents). After collecting participants' responses,
+the final dataset consisted of 139,326 observations, each corresponding
+to a single response of one participant to one item. Each translation
+equivalent received a median of 234 responses (\emph{Min} = 108,
+\emph{Max} = 880) from participants, both languages pooled together.
 Data processing and visualisation was done in R
 (\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
-version 4.2.2) using the Tidyverse family of packages
-(\protect\hyperlink{ref-wickham2019welcome}{Wickham et al., 2019}).
+version 4.2.2).
 
@@
 
-\hypertarget{model-predictors}{%
-\subsubsection{Model predictors}\label{model-predictors}}
+We implemented the model using \texttt{brms}
+(\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), a R interface
+to the Stan probabilistic language (2.32.1)
+(\protect\hyperlink{ref-carpenter2017stan}{Carpenter et al., 2017}). We
+ran four iteration chains using the by-default No U-Turn Sampler
+algorithm with 1,000 iterations each and an additional 1,000 warm-up
+iterations per chain.
+
+\hypertarget{sec-predictors}{%
+\subsubsection{Model predictors}\label{sec-predictors}}
 
@@
 word-form's lexical frequency, but also of the quantitative input they
-receive from the language such word-form belongs to, we expressed
+receive from the language such word-form belongs to. We expressed
 lexical frequencies as the product between both variables. First, we
-extracted the child-directed lexical frequency of each word from the
-CHILDES database
+extracted the child-directed lexical frequency of each word-form from
+the CHILDES database
 (\protect\hyperlink{ref-macwhinney2000childes}{MacWhinney, 2000};
@@
 whose degree of exposure is 80\% for Catalan and 20\% for Spanish, the
-expected \emph{Exposure} score to the Catalan word \emph{cotxe}
+expected \emph{Exposure} score to the Catalan word-form \emph{cotxe}
 {[}\emph{car}{]}---with a lexical frequency of 6.33---would be 5.06,
@@
 distance measures the number of editions (insertions, deletions, or
-substitutions) that one string of phonemes/characters must go through to
-become identical to the other
+substitutions) that one string of characters must go through to become
+identical to the other
 (\protect\hyperlink{ref-levenshtein1966binary}{Levenshtein, 1966}). We
 divided the Levenshtein distance of each translation equivalent by the
-length of the longest word-form to correct for word length (longer
+length of the longest word-form to correct for word-form length (longer
 strings are likely to show a larger number of mismatches). Finally, we
@@
 indicates that the two word-forms are identical (e.g.,
-/\textipa{"mar}/--/\textipa{"mar}/, \emph{sea}) (see
-\protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
+/\textipa{"mar}/--/\textipa{"mar}/, \emph{sea})
+(\protect\hyperlink{ref-floccia2018introduction}{Floccia et al., 2018};
 \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020};
 \protect\hyperlink{ref-heeringa2003norwegian}{Heeringa \& Gooskens,
-2003}; \protect\hyperlink{ref-laing2022phonological}{Laing, 2022} for
-similar approaches;
-\protect\hyperlink{ref-schepens2012distributions}{Schepens et al.,
-2012}).
-
-Predictors were standardised before entering the model by subtracting
-the mean of the predictor from each value and dividing the result by the
-standard deviation of the predictor.
+2003}; \protect\hyperlink{ref-schepens2012distributions}{Schepens et
+al., 2012}). Predictors were standardised before entering the model by
+subtracting the mean of the predictor from each value and dividing the
+result by the standard deviation of the predictor.
 
@@
 
-We implemented the model using \texttt{brms}
-(\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), a R interface
-to the Stan probabilistic language (2.32.1)
-(\protect\hyperlink{ref-carpenter2017stan}{Carpenter et al., 2017}). We
-ran four iteration chains using the by-default No U-Turn Sampler
-algorithm with 1,000 iterations each and an additional 1,000 warm-up
-iterations per chain. Model posterior draws and predictions were handled
-using the \texttt{tidybayes}
-(\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) R package.
-
 \hypertarget{sec-results}{%
@@
 
-Chain convergence diagnostics showed adequate values, and the model
-posterior showed little evidence of correlation, as indicated by
-negligible pairwise correlation between the marginal posterior
-distribution of coefficients. Table~\ref{tbl-coefs} shows the summary of
-the posterior distribution of the fixed regression coefficients, and
-their degree of overlap with the ROPE. For interpretability, we report
-the estimated regression coefficients transformed to the probability
-scale\footnote{The logit and probability scales relate non-linearly.
-  This means that one logit difference is not necessarily translated to
-  a unique value in the probability scale. For example, the probability
-  of acquisition of a given word might increase in 5\% when age
-  increases from 22 to 23 months, the probability of acquisition of the
-  same word might only increase in 0.2\% when age increases from 31 to
-  32 months. The linear growth of the probability of acquisition differs
-  along the logistic curve, and therefore deciding the age point at
-  which to report the estimates of the regression coefficients in the
-  probability scale is not trivial. Following Gelman et al.
+The model posterior showed adequate chain convergence diagnostics and
+posterior predictive checks (see Appendix A). Table~\ref{tbl-coefs}
+shows the summary of the posterior distribution of the fixed regression
+coefficients, and their degree of overlap with the ROPE. For
+interpretability, we report the estimated regression coefficients
+transformed to the probability scale\footnote{The logit and probability
+  scales relate non-linearly. This means that one logit difference is
+  not necessarily translated to a unique value in the probability scale.
+  For example, the probability of acquisition of a given word might
+  increase in 5\% when age increases from 22 to 23 months, the
+  probability of acquisition of the same word might only increase in
+  0.2\% when age increases from 31 to 32 months. The linear growth of
+  the probability of acquisition differs along the logistic curve, and
+  therefore deciding the age point at which to report the estimates of
+  the regression coefficients in the probability scale is not trivial.
+  Following Gelman et al.
   (\protect\hyperlink{ref-gelman2020regression}{2020}), we report the
@@
 \multicolumn{4}{l}{\textbf{Intercepts (at 22 months)}}\\
-\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & -\\
-\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & -\\
+\hspace{1em}Comprehension and Production & 0.435 & {}[0.383, 0.492] & -\\
+\hspace{1em}Comprehension & 0.936 & {}[0.921, 0.948] & -\\
 \addlinespace[0.3em]
 \multicolumn{4}{l}{\textbf{Slopes (upper bound)}}\\
-\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000\\
-\hspace{1em}Age (+1 SD, 4.86 months) & 0.405 & {}[0.357, 0.451] & .000\\
-\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000\\
-\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .120\\
-\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000\\
-\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000\\
-\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .974\\
-\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .907\\
+\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.089, -0.038] & .000\\
+\hspace{1em}Age (+1 SD, 4.86 months) & 0.406 & {}[0.359, 0.455] & .000\\
+\hspace{1em}Exposure (+1 SD, 1.81) & 0.231 & {}[0.197, 0.265] & .000\\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.056 & {}[0.007, 0.099] & .194\\
+\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.068, -0.046] & .000\\
+\hspace{1em}Age $\times$ Exposure & 0.072 & {}[0.037, 0.104] & .000\\
+\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0.001, 0.027] & .936\\
+\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.019 & {}[-0.027, -0.011] & .874\\
 \bottomrule
@@
 The coefficient of \emph{Age} showed the strongest association with the
-probability of acquisition (\(\beta\) = 0.405, 95\% HDI = {[}0.357,
-1.806{]}), with all posterior samples falling out of the ROPE. A
+probability of acquisition (\(\beta\) = 0.406, 95\% HDI = {[}0.359,
+0.455{]}), with all posterior samples falling out of the ROPE. A
 one-month increment in age increased a maximum of 0.08 the probability
-of acquisition. Similarly, the word exposure index (\emph{Exposure}) had
-a strong effect on the probability of acquisition (\(\beta\) = 0.233,
-95\% HDI = {[}0.201, 1.074{]}). All of the posterior samples of this
-regression coefficient excluded the ROPE. The impact of this predictor
-on the probability of acquisition was positive: for every standard
-deviation increase in exposure, the participant was 0.129 more likely to
-acquire it. Word-form length also showed a significant association with
-probability of acquisition (\(\beta\) = -0.062, 95\% HDI = {[}-0.086,
--0.143{]}). For every phoneme in the word-form, participants were -0.04
-less likely to know it. The 95\% HDI of the regression coefficient of
-the \emph{Age} \(\times\) \emph{Exposure} interaction also excluded the
-ROPE (\(\beta\) = 0.071, 95\% HDI = {[}0.039, 0.414{]}), showing that
-the effect of the word exposure index differed across ages: older
-children were more likely to acquire words with a higher exposure rate
-than younger children.
-
-The posterior distribution of the main effect of cognateness excluded
-the ROPE completely (\(\beta\) = 0.058, 95\% HDI = {[}0.014, 0.416{]}).
-For every 10\% increment in cognateness, the acquisition of a word
-increased in 0.006. The effect of \emph{Cognateness} interacted with
-that of \emph{Exposure}: the 95\% HDI of the regression coefficient of
-interaction excluded the ROPE entirely (\(\beta\) = -0.06, 95\% HDI =
-{[}-0.069, -0.184{]}), suggesting that the effect of cognateness on a
-word's probability of acquisition changed depending on participants'
-exposure to word. Follow-up analyses on this interaction showed that
-when exposure rate was low (e.g., -1 SD), cognateness increased the
-probability of acquisition substantially. This effect was negligible for
-words with median or high exposure (+1 SD) (see
-Figure~\ref{fig-marginal}).
+of acquisition. Similarly, the word-form exposure index
+(\emph{Exposure}) had a strong effect on the probability of acquisition
+(\(\beta\) = 0.231, 95\% HDI = {[}0.197, 0.265{]}). All of the posterior
+samples of this regression coefficient excluded the ROPE. The impact of
+this predictor on the probability of acquisition was positive: for every
+standard deviation increase in exposure, the participant was 0.128
+probability points more likely to acquire it. Word-form length also
+showed a significant association with probability of acquisition
+(\(\beta\) = -0.062, 95\% HDI = {[}-0.089, -0.038{]}). For every phoneme
+in the word-form, participants were -0.04 probability points less likely
+to know it. The 95\% HDI of the regression coefficient of the \emph{Age}
+\(\times\) \emph{Exposure} interaction also excluded the ROPE (\(\beta\)
+= 0.072, 95\% HDI = {[}0.037, 0.104{]}), showing that the effect of the
+word-form exposure index differed across ages: older children were more
+likely to acquire words with a higher exposure rate than younger
+children.
+
+Around 19.44\% of the posterior samples of the main effect of
+\emph{Cognateness} overlapped with the ROPE completely (\(\beta\) =
+0.056, 95\% HDI = {[}0.007, 0.099{]}). For every 10\% increment in
+cognateness, the probability of word acquisition increased in 0.005. The
+effect of \emph{Cognateness} interacted with that of \emph{Exposure}:
+the 95\% HDI of the regression coefficient of interaction excluded the
+ROPE entirely (\(\beta\) = -0.06, 95\% HDI = {[}-0.068, -0.046{]}),
+suggesting that the effect of cognateness on a word's probability of
+acquisition changed depending on participants' exposure to the
+word-form. When \emph{Exposure} was low (e.g., -1 SD),
+\emph{Cognateness} increased the probability of acquisition
+substantially, while this effect was negligible for words with median or
+high exposure (+1 SD) (see Figure~\ref{fig-marginal}).
 
@@
 bands correspond to the mean and 95\% credible interval of the
-posterior-predicted means. Different colours indicate different levels
-of cognateness (phonological similarity). Predictions are presented
-separately for different degrees of word exposure index: little exposure
-to the word, mean exposure, and high exposure. Predictions for
-Comprehension are shown on top and predictions for Comprehension and
-Production are shown on the bottom. In-sample predictions lie inside the
-grey rectangles.}
+posterior-predicted means. Different shades of grey indicate different
+levels of cognateness (phonological similarity). Predictions are
+presented separately for different degrees of word-form exposure index:
+little exposure to the word-form, mean exposure, and high exposure).
+Predictions for Comprehension are shown on top and predictions for
+Comprehension and Production are shown on the bottom. In-sample
+predictions lie inside the yellow rectangles. For reference, we indicate
+the 50\% chance level of word acquisition (horizontal yellow lines), and
+the mean age of the sample (vertical yellow lines).}
 
@@
 as separate predictors (instead of the composite \emph{Exposure}
-measure) showed equivalent results (see Appendix A). To rule out the
+measure) showed equivalent results (see Appendix B). To rule out the
 possibility that cognateness facilitation effect we found was due to
@@
 syllable appears in child-directed speech, embedded within different
-words. Finally, for each word, we summed the frequency of its syllables,
-as an estimate of the syllabic frequency of the word. We fit a Bayesian
-model with \emph{Cognateness} as response variable, and the main effects
-of syllable frequency and number of syllables (to control for the fact
-words with more syllables are more likely to score higher in syllabic
-frequency) as predictors. This model provided strong evidence for the
-association between cognateness and syllabic frequency being equivalent
-to zero (see Appendix B).
+words. Finally, for each word-form, we summed the frequency of its
+syllables, as an estimate of the syllabic frequency of the word-form. We
+fit a Bayesian model with \emph{Cognateness} as response variable, and
+the main effects of syllable frequency and number of syllables (to
+control for the fact words with more syllables are more likely to score
+higher in syllabic frequency) as predictors. This model provided strong
+evidence for the association between cognateness and syllabic frequency
+being equivalent to zero (see Appendix C).
 
@@
 This study investigated the impact of cognateness (i.e., phonological
-similarity between each word-form and its translation) on the early
-bilingual lexicon. We used Bayesian Item Response Theory to model the
-acquisition trajectories of a large sample of Catalan and Spanish words,
-estimating the effect of cognateness on the probability of acquisition.
-This model corrected for participants' age, word-form length (number of
-phonemes), and a novel measure of participants' exposure rate to each
-word. Exposure rates were calculated as a language exposure-weighted
-lexical frequency score in which each word-form's lexical frequency was
+similarity between translation equivalents) on the early bilingual
+lexicon. We used Bayesian item response theory to model the acquisition
+trajectories of a large sample of Catalan and Spanish words, estimating
+the effect of cognateness on the probability of acquisition. This model
+corrected for participants' age, word-form length (number of phonemes),
+and a novel measure of participants' exposure rate to each word-form.
+Exposure rates were calculated as a language exposure-weighted lexical
+frequency score in which each word-form's lexical frequency was
 corrected by the degree to which the participant was exposed to each
@@
 from their cognate status, whereas high-exposure word-forms did not.
-Capitalising on accumulator models of language acquisition, we provide a
-theoretical account of bilingual lexical acquisition. In this account,
-parallel activation of the two languages plays a central role during the
-acquisition of early representations in the bilingual lexicon, and in
-which the dynamics of co-activation between translation equivalents
-results in an earlier age-of-acquisition.
+Using the concept of accumulator (see
+\protect\hyperlink{ref-kachergis2022standard}{Kachergis et al., 2022}
+for review), we provide a theoretical account of bilingual lexical
+acquisition. In the present account, parallel activation of the two
+languages plays a central role during the acquisition of early
+representations in the bilingual lexicon, and in which the dynamics of
+co-activation between translation equivalents results in an earlier
+age-of-acquisition.
 
@@
 representation of its translation. The amount of co-activation that
-spreads from the spoken word to its translation is proportional to the
-amount of phonological similarity between both word-forms. Cognates
+spreads from the spoken word-form to its translation is proportional to
+the amount of phonological similarity between both word-forms. Cognates
 would receive more activation from their translation than non-cognates,
@@
 instance to the same word-form would contribute not only to the
-acquisition of the representation of such word, but also, to some
-extent, to the acquisition of its translation. Our findings provide
-strong support for an account of bilingual vocabulary growth in which
-the experience and learning outcomes accumulated by the child in one
-language impact those in the other language. Such a facilitatory
-cross-language mechanism might be an important piece in the puzzle of
-bilingual language acquisition. In particular, it may shed some light on
-why bilingual infants do not show relevant delays in language
-acquisition milestones compared to their monolingual peers, while
-receiving a reduced quantity of speech input in each of their languages.
-Our results provide some insights into this issue: infants benefited
-more strongly from the cognateness facilitation effect when acquiring
-words from the language of lower exposure than in the language of higher
+acquisition of the representation of the word, but also, to some extent,
+to the acquisition of its translation. Our findings provide strong
+support for an account of bilingual vocabulary growth in which the
+experience and learning outcomes accumulated by the child in one
+language impact those in the other language through cross-language
+phonological associations. Such a facilitatory mechanism might be an
+important piece in the puzzle of bilingual language acquisition. In
+particular, it may shed some light on why bilingual infants do not show
+relevant delays in language acquisition milestones compared to their
+monolingual peers, while receiving a reduced quantity of speech input in
+each of their languages. Infants in the present study benefited more
+strongly from the cognateness facilitation effect when acquiring words
+from the language of lower exposure than in the language of higher
 exposure.
-
-We suggest that this asymmetry is the result of children' unbalanced
-exposure to their languages. A bilingual child's frequency of exposure
-to a given word-form is mostly determined by two factors: the
-word-form's lexical frequency and the child's amount of language
-exposure to the language such word belongs to. A dual linguistic input
-means lower exposure to each of the languages, unless one makes
-the---arguably implausible---assumption that bilinguals are exposed on
-average to twice the amount of linguistic input than monolinguals.
-Because of this difference in exposure, words lower-exposure language
-might receive activation from their translation in the higher-exposure
-language more often than words from the higher-exposure language receive
-activation from their translation in the lower-exposure language. As a
-result, the cognateness facilitation effect should be stronger in words
-from the lower-exposure language.
 
@@
 
+The asymmetric facilitation of cognateness on word acquisition reported
+in the present study parallels previous findings in toddlers and adults.
+For instance, unbalanced (or low-proficiency) bilinguals benefit from
+cross-language forward priming (dominant to non-dominant) during word
+processing (\protect\hyperlink{ref-de1991lexical}{De Groot \& Nas,
+1991}; \protect\hyperlink{ref-grainger1998masked}{Grainger, 1998};
+\protect\hyperlink{ref-shook2019covert}{Shook \& Marian, 2019};
+\protect\hyperlink{ref-singh2014one}{Singh, 2014};
+\protect\hyperlink{ref-vonholzen2019impact}{Von Holzen et al., 2019};
+but see \protect\hyperlink{ref-jardak2019labels}{Jardak \&
+Byers-Heinlein, 2019}). One the other hand, backward priming
+(non-dominant to dominant) seems less robust and more challenging to
+detect (e.g., \protect\hyperlink{ref-hoshino2010erp}{Hoshino et al.,
+2010}; \protect\hyperlink{ref-midgley2009language}{Midgley et al.,
+2009}; but see \protect\hyperlink{ref-duyck2009translation}{Duyck \&
+Warlop, 2009}). Balanced (or high-proficiency bilinguals) show an
+equivalent priming facilitation in both directions
+(\protect\hyperlink{ref-basnight2007differences}{Basnight-Brown \&
+Altarriba, 2007};
+\protect\hyperlink{ref-dunabeitia2009masked}{Duñabeitia et al., 2009}).
+These results have been taken as evidence for an asymmetry in the
+strength of forward and backward connections in the unbalanced bilingual
+lexicon. Although implemented in different ways, or found under
+different assumptions, such a dominance-mediated asymmetry is accounted
+for by multiple models of lexical processing like the Revised
+Hierarchical Model (\protect\hyperlink{ref-kroll1994category}{Kroll \&
+Stewart, 1994}), BIA/BIA+
+(\protect\hyperlink{ref-dijkstra2013bia}{Dijkstra \& Van Heuven, 2013}),
+BLINCS (\protect\hyperlink{ref-shook2013bilingual}{Shook \& Marian,
+2013}), or Multilink
+(\protect\hyperlink{ref-dijkstra2019multilink}{Dijkstra et al., 2019}),
+and also by models providing a more development-oriented perspective,
+like the Ontogenic Model
+(\protect\hyperlink{ref-bordag2022ontogenesis}{Bordag et al., 2022};
+\protect\hyperlink{ref-cook2016fuzzy}{Cook et al., 2016}), and BIA-d
+(\protect\hyperlink{ref-grainger2010re}{Grainger et al., 2010}).
+Overall, this provides an apparently convenient account for the
+interaction between language dominance and cognateness found in the
+present study. These models are aimed at explaining results in adults,
+and their predictions should be taken with caution when extended to
+early language acquisition.
+
+In adult bilingual populations, language dominance and proficiency are
+frequently defined using dimensions other than degree of exposure, which
+is a more common practice in infant research
+(\protect\hyperlink{ref-marian2021measuring}{Marian \& Hayakawa, 2021};
+\protect\hyperlink{ref-rocha2023defining}{Rocha-Hidalgo \& Barr, 2023}).
+For instance, low-proficiency bilinguals in many of the aforementioned
+studies acquired their second language years after their toddlerhood. We
+identify three critical ways in which this prevents a clear comparison
+between our results and those from studies on second language
+acquisition in adults. First, in adult second language acquisition, the
+acquisition of the phonology of the new language must be negotiated with
+the already acquired phoneme inventory of the first language (e.g.,
+\protect\hyperlink{ref-cutler2006asymmetric}{Cutler et al., 2006};
+\protect\hyperlink{ref-sebastian2006first}{Sebastian-Gallés et al.,
+2006}), in place around the first year of life (see
+\protect\hyperlink{ref-werker2015critical}{Werker \& Hensch, 2015} for
+review). Second, adults acquiring a second language already possess a
+system of form-meaning mappings, whereas simultaneous bilingual infants
+must build a lexicon for two languages in the absence of clear
+form-meaning mappings. Third, adults are assumed to be literate and to
+possess an orthographic system in place, which may shape how new words
+are integrated in the lexicon and processed during experimental tasks
+(e.g., \protect\hyperlink{ref-thierry2007brain}{Thierry \& Wu, 2007}).
+In this scenario, the acquisition of a second language may take place in
+a substantially different way compared to how bilingual infants acquire
+two languages from birth. A more similar case to the one concerning the
+present study is considered by the DevLex-II model
+(\protect\hyperlink{ref-zhao2010bilingual}{Zhao \& Li, 2010}), which
+captures unique features of the early bilingual lexicon, and considers
+the case of infants simultaneously acquiring their two language. In line
+with the adult models, DevLex-II predicts asymmetries between word
+representation from the dominant and the non-dominant language.
+Simulations from DevLex-II result in an asymmetric cross-language
+priming, in which words from the dominant (acquired acquired) language
+primed more strongly the recognition of words in the non-dominant
+language (later acquired) than in the other direction
+(\protect\hyperlink{ref-zhao2013simulating}{Zhao \& Li, 2013}).
+
+In summary, there is a compelling case for attributing asymmetric
+effects of parallel activation to differences in activation strength
+between forward and backward connections. It is nonetheless possible
+that, as argued in the introduction, the asymmetric effect of
+cognateness found in the present study is simply the result of infants
+being exposed more frequently to words in the dominant language than to
+words in the non-dominant language. This would lead to words in the
+non-dominant language receiving additional parallel activation, compared
+to words in the dominant language, and therefore benefiting more
+strongly from their cognate status. These two accounts are not mutually
+exclusive, as words in the dominant language may active more
+\emph{strongly} their translations than \emph{vice versa}, on top of
+such activation being more frequent. Further research is needed in order
+to clarify this issue.
+
 It might be argued that our results reflect the fact that cognate
@@
 possible that infants classify both are as acceptable variations of the
-same word, therefore treating them as a single lexical item. This would
-lead to a faster increase in cumulative learning instances, and to
+same word-form, therefore treating them as a single lexical item. This
+would lead to a faster increase in cumulative learning instances, and to
 earlier ages of acquisition for cognate translation equivalents (for
@@
 articulatory abilities develop, compared to their word recognition
-abilities (\protect\hyperlink{ref-hustad2020development}{Hustad et al.,
-2020}, \protect\hyperlink{ref-hustad2021speech}{2021}). This gap between
-comprehension and production is even larger in the less dominant
-language of bilingual children
+abilities (\protect\hyperlink{ref-hustad2021speech}{Hustad et al.,
+2021}). This gap between comprehension and production is even larger in
+the less dominant language of bilingual children
 (\protect\hyperlink{ref-giguere2022bilingual}{Giguere \& Hoff, 2022}).
@@
 al., 2005}; \protect\hyperlink{ref-gillen2021tapping}{Gillen et al.,
-2021}; \protect\hyperlink{ref-killing2008move}{Killing \& Bishop, 2008};
-but see
+2021}; but see
 \protect\hyperlink{ref-houston-price2007discrepancy}{Houston-Price et
@@
 that during the acquisition of low frequency words, bilingual children
-seem to benefit more strongly from the word's phonological similarity
-with its translation in the other language. Capitalising on accumulator
-models of language acquisition we put forward a theoretical account of
-bilingual word learning, in which cognateness interacts with lexical
-frequency and language exposure to boost the acquisition of translation
-equivalents.
+seem to benefit more strongly from the word-forms's phonological
+similarity with its translation in the other language. Capitalising on
+the notion of \emph{accumulator} of linguistic input, we put forward a
+theoretical account of bilingual word learning, in which cognateness
+interacts with lexical frequency and language exposure to boost the
+acquisition of translation equivalents.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cutler2006asymmetric}{}}%
+Cutler, A., Weber, A., \& Otake, T. (2006). Asymmetric mapping from
+phonetic to lexical representations in second-language listening.
+\emph{Journal of Phonetics}, \emph{34}(2), 269--284.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de1991lexical}{}}%
+De Groot, A. M., \& Nas, G. L. (1991). Lexical representation of
+cognates and noncognates in compound bilinguals. \emph{Journal of Memory
+and Language}, \emph{30}(1), 90--123.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dijkstra2013bia}{}}%
+Dijkstra, T., \& Van Heuven, W. J. (2013). The BIA model and bilingual
+word recognition. In \emph{Localist connectionist approaches to human
+cognition} (pp. 189--225). Psychology Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dijkstra2019multilink}{}}%
+Dijkstra, T., Wahl, A., Buytenhuijs, F., Halem, N. V., Al-Jibouri, Z.,
+Korte, M. D., \& Rekké, S. (2019). Multilink: A computational model for
+bilingual word recognition and word translation. \emph{Bilingualism:
+Language and Cognition}, \emph{22}(4), 657--679.
+\url{https://doi.org/10.1017/S1366728918000287}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dunabeitia2009masked}{}}%
+Duñabeitia, J. A., Perea, M., \& Carreiras, M. (2009). Masked
+translation priming effects with highly proficient simultaneous
+bilinguals. \emph{Experimental Psychology}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-duyck2009translation}{}}%
+Duyck, W., \& Warlop, N. (2009). Translation priming between the native
+language and a second language: New evidence from dutch-french
+bilinguals. \emph{Experimental Psychology}, \emph{56}(3), 173--179.
+
 \leavevmode\vadjust pre{\hypertarget{ref-2018els}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-floccia2020translation}{}}%
+Floccia, C., Delle Luche, C., Lepadatu, I., Chow, J., Ratnage, P., \&
+Plunkett, K. (2020). Translation equivalent and cross-language semantic
+priming in bilingual toddlers. \emph{Journal of Memory and Language},
+\emph{112}, 104086.
+
 \leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-grainger1998masked}{}}%
+Grainger, J. (1998). Masked priming by translation equivalents in
+proficient bilinguals. \emph{Language and Cognitive Processes},
+\emph{13}(6), 601--623.
+
+\leavevmode\vadjust pre{\hypertarget{ref-grainger2010re}{}}%
+Grainger, J., Midgley, K., \& Holcomb, P. J. (2010). Re-thinking the
+bilingual interactive-activation model from a developmental perspective
+(BIA-d). \emph{Language Acquisition Across Linguistic and Cognitive
+Systems}, \emph{52}, 267--283.
+
 \leavevmode\vadjust pre{\hypertarget{ref-grosjean2021extent}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-vanhell2008sentence}{}}%
-Hell, J. G. van, \& Groot, A. M. B. de. (2008). Sentence context
-modulates visual word recognition and translation in bilinguals.
-\emph{Acta Psychologica}, \emph{128}(3), 431--451.
-\url{https://doi.org/10.1016/j.actpsy.2008.03.010}
-
 \leavevmode\vadjust pre{\hypertarget{ref-hidaka2013computational}{}}%
@@
 \url{https://doi.org/10.1016/j.cognition.2007.02.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hoshino2010erp}{}}%
+Hoshino, N., Midgley, K. J., Holcomb, P. J., \& Grainger, J. (2010). An
+ERP investigation of masked cross-script translation priming.
+\emph{Brain Research}, \emph{1344}, 159--172.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hustad2020development}{}}%
-Hustad, K. C., Mahr, T., Natzke, P. E. M., \& Rathouz, P. J. (2020).
-Development of speech intelligibility between 30 and 47 months in
-typically developing children: A cross-sectional study of growth.
-\emph{Journal of Speech, Language, and Hearing Research}, \emph{63}(6),
-1675--1687. \url{https://doi.org/10.1044/2020_JSLHR-20-00008}
+\leavevmode\vadjust pre{\hypertarget{ref-jardak2019labels}{}}%
+Jardak, A., \& Byers-Heinlein, K. (2019). Labels or concepts? The
+development of semantic networks in bilingual two-year-olds. \emph{Child
+Development}, \emph{90}(2), e212--e229.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kay2021tidybayes}{}}%
-Kay, M. (2021). \emph{Tidybayes: Tidy data and geoms for bayesian
-models}. \url{http://mjskay.github.io/tidybayes/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-killing2008move}{}}%
-Killing, S. E. A., \& Bishop, D. V. M. (2008). Move it! Visual feedback
-enhances validity of preferential looking as a measure of individual
-differences in vocabulary in toddlers. \emph{Developmental Science},
-\emph{11}(4), 525--530.
-\url{https://doi.org/10.1111/j.1467-7687.2008.00698.x}
-
 \leavevmode\vadjust pre{\hypertarget{ref-kroll2017bilingual}{}}%
@@
 Handbook of Psycholinguistics}, 294--319.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroll1994category}{}}%
+Kroll, J. F., \& Stewart, E. (1994). Category interference in
+translation and picture naming: Evidence for asymmetric connection
+between bilingual memory representations. \emph{Journal of Memory and
+Language}, \emph{33}(2), 149--174.
+\url{https://doi.org/10.1006/jmla.1994.1008}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-laing2022phonological}{}}%
-Laing, C. E. (2022). \emph{Phonological networks and systematicity in
-early lexical acquisition}. {PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/z8pyg}
-
 \leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
@@
 \emph{38}(3), 606--627. \url{https://doi.org/10.1017/S0305000910000243}
+
+\leavevmode\vadjust pre{\hypertarget{ref-marian2021measuring}{}}%
+Marian, V., \& Hayakawa, S. (2021). Measuring bilingualism: The quest
+for a {``bilingualism quotient.''} \emph{Applied Psycholinguistics},
+\emph{42}(2), 527--548.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-midgley2009language}{}}%
+Midgley, K. J., Holcomb, P. J., \& Grainger, J. (2009). Language effects
+in second language learners and proficient bilinguals investigated with
+event-related potentials. \emph{Journal of Neurolinguistics},
+\emph{22}(3), 281--300.
+
 \leavevmode\vadjust pre{\hypertarget{ref-mitchell2022cognates}{}}%
@@
 \url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pearson1993lexical}{}}%
-Pearson, B. Z., Fernández, S. C., \& Oller, D. K. (1993). Lexical
-development in bilingual infants and toddlers: Comparison to monolingual
-norms. \emph{Language Learning}, \emph{43}(1), 93--120.
-\url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-petitto2001bilingual}{}}%
-Petitto, L. A., Katerelos, M., Levy, B. G., Gauna, K., Tétreault, K., \&
-Ferraro, V. (2001). Bilingual signed and spoken language acquisition
-from birth: Implications for the mechanisms underlying early bilingual
-language acquisition. \emph{Journal of Child Language}, \emph{28}(2),
-453--496. \url{https://doi.org/10.1017/S0305000901004718}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-rocha2023defining}{}}%
+Rocha-Hidalgo, J., \& Barr, R. (2023). Defining bilingualism in infancy
+and toddlerhood: A scoping review. \emph{International Journal of
+Bilingualism}, \emph{27}(3), 253--274.
+
 \leavevmode\vadjust pre{\hypertarget{ref-samuelson2021precision}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-schroter2016orthographic}{}}%
-Schröter, P., \& Schroeder, S. (2016). Orthographic processing in
-balanced bilingual children: Cross-language evidence from cognates and
-false friends. \emph{Journal of Experimental Child Psychology},
-\emph{141}, 239--246. \url{https://doi.org/10.1016/j.jecp.2015.09.005}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schwartz2007reading}{}}%
-Schwartz, A. I., Kroll, J. F., \& Diaz, M. (2007). Reading words in
-spanish and english: Mapping orthography to phonology in two languages.
-\emph{Language and Cognitive Processes}, \emph{22}(1), 106--129.
-\url{https://doi.org/10.1080/01690960500463920}
-
 \leavevmode\vadjust pre{\hypertarget{ref-sebastian-galles2020bilingual}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-sebastian2006first}{}}%
+Sebastian-Gallés, N., Rodrı́guez-Fornells, A., Diego-Balaguer, R. de, \&
+Dı́az, B. (2006). First-and second-language phonological representations
+in the mental lexicon. \emph{Journal of Cognitive Neuroscience},
+\emph{18}(8), 1277--1291.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shook2012bimodal}{}}%
+Shook, A., \& Marian, V. (2012). Bimodal bilinguals co-activate both
+languages during spoken comprehension. \emph{Cognition}, \emph{124}(3),
+314--324.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shook2013bilingual}{}}%
+Shook, A., \& Marian, V. (2013). The bilingual language interaction
+network for comprehension of speech*. \emph{Bilingualism: Language and
+Cognition}, \emph{16}(2), 304--324.
+\url{https://doi.org/10.1017/S1366728912000466}
+
+\leavevmode\vadjust pre{\hypertarget{ref-shook2019covert}{}}%
+Shook, A., \& Marian, V. (2019). Covert co-activation of bilinguals'
+non-target language: Phonological competition from translations.
+\emph{Linguistic Approaches to Bilingualism}, \emph{9}(2), 228--252.
+
+\leavevmode\vadjust pre{\hypertarget{ref-singh2014one}{}}%
+Singh, L. (2014). One world, two languages: Cross-language semantic
+priming in bilingual toddlers. \emph{Child Development}, \emph{85}(2),
+755--766.
+
 \leavevmode\vadjust pre{\hypertarget{ref-smithson2014bilingualism}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-thierry2007brain}{}}%
+Thierry, G., \& Wu, Y. J. (2007). Brain potentials reveal unconscious
+translation during foreign-language comprehension. \emph{Proceedings of
+the National Academy of Sciences}, \emph{104}(30), 12530--12535.
+\url{https://doi.org/10.1073/pnas.0609927104}
+
 \leavevmode\vadjust pre{\hypertarget{ref-tincoff1999beginnings}{}}%
@@
 extension of {SAMPA}}. \emph{4}(28), 1995.
+
+\leavevmode\vadjust pre{\hypertarget{ref-werker2015critical}{}}%
+Werker, J. F., \& Hensch, T. K. (2015). Critical periods in speech
+perception: New directions. \emph{Annual Review of Psychology},
+\emph{66}, 173--196.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wickham2019welcome}{}}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
-François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
-Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
-Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
-Welcome to the tidyverse. \emph{Journal of Open Source Software},
-\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
+\leavevmode\vadjust pre{\hypertarget{ref-zhao2010bilingual}{}}%
+Zhao, X., \& Li, P. (2010). Bilingual lexical interactions in an
+unsupervised neural network model. \emph{International Journal of
+Bilingual Education and Bilingualism}, \emph{13}(5), 505--524.
+
+\leavevmode\vadjust pre{\hypertarget{ref-zhao2013simulating}{}}%
+Zhao, X., \& Li, P. (2013). Simulating cross-language priming with a
+dynamic computational model of the lexicon. \emph{Bilingualism: Language
+and Cognition}, \emph{16}(2), 288--303.
 

### 2023-09-21 manuscript/manuscript.qmd

@@
     split(model_summary$.variable)
-
-re_data <- model_summary_re |>
-    filter(.variable %in% c(
-        "b_lv_std",
-        "b_exposure_std:lv_std"
-    ))
-
-re_data_prop <- re_data |>
-    mutate(
-        is_positive = .value > 0,
-        is_negative = .value < 0
-    ) |>
-    summarise(across(matches("is_"), mean),
-        .by = .variable_name
-    ) |>
-    mutate(prop = ifelse(is_positive > is_negative,
-        is_positive,
-        is_negative
-    ))
 ```

### 2023-10-03 manuscript/_appendix/_model-details.qmd
Ignore .tiff, .eps, and .svg

--- 2023-05-24 manuscript/_appendix/_model-details.qmd
@@
-We used multilevel ordinal regression to model the cumulative probability of a *No* response, a *Understands* response, or a *Understands and Says* response [@burkner2019ordinal] using the *logit* link function.
-The likelihood of the cumulative probability distribution of the responses is defined by @eq-likelihood.
+### Model structure and priors
 
-$$
-\begin{aligned}
-\textbf{Likelihood:} \\
-y_{ij} &\sim \text{Cumulative}(p_{k})
-\end{aligned}
-$$ {#eq-likelihood}
+We used Stan [@carpenter2017stan] as the probabilistic language behind the estimation of our Bayesian models in this study, with `brms` as its R interface [@burkner2017brms].
+This language implements the Markov Chain Monte Carlo (MCMC) algorithm using the Hamiltonian Monte Carlo method (HMC) to explore the posterior distribution of the model.
+Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined.
+We run `r english::words(dim(model_fit$fit)[2])` MCMC chains, each `r format(dim(model_fit$fit)[1], big.mark = ",")` iterations long each.
 
-where: $y$ is an observed response ($y \in \{\text{No, Understands, Understands and Says}\}$), $i$ is the participant index, $j$ is the translation equivalent (TE) index, $p_{k}$ is a probability ($p \in (0, 1)$) that indicates the threshold $k$ ($k \in (1, 2)$) between two response categories in the latent distribution $p_{k}$ is then estimated using a logit regression model as indicated in @eq-linear.
+### Considerations on statistical power and sample size
 
-To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($\text{Age}$), word-form length ($\text{Length}$), word exposure index ($\text{Exposure}$), and of phonological similarity ($\text{Cognateness}$), the two-way interactions $\text{Age} \times \text{Exposure}$, $\text{Age} \times \text{Cognateness}$, and $\text{Exposure} \times \text{Cognateness}$, and the three-way interaction $\text{Age} \times \text{Exposure} \times \text{Cognateness}$.
-We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression].
-For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random], see (@eq-linear).
+There is little consensus about what approach is adequate for calculating the statistical power of a complex Bayesian model like the one in the present study, for several reasons.
+A first pitfall, shared with frequentist analysis, is that a closed solution for statistical power calculation is not possible or cannot be computed within reasonable time constraints.
+This rules out the use of many available pieces of software that are commonly offered for power analysis, as they commonly only consider the case of simpler models like t-tests, ANOVA, Pearson correlation, or regression (with only fixed effects), or trivial derivations of thereof.
+The more complicated case of multilevel models is usually not covered, not to mention those with a Bayesian approach.
 
-$$
-\begin{aligned}
-\textbf{Linear model:} \\
-\text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
-& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + & \\
-& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + & \\
-& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
-& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{ij} + & \\
-& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
-& (\beta_{6} + u_{6_{i}}) \cdot (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
-& (\beta_{7} + u_{7_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
-& (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij})
-\end{aligned}
-$$ {#eq-linear}
+An alternative way of estimating the statistical power of statistical test is simulation.
+This consists on simulating multiple datasets in which the hypothesised effect size is present, and fitting multiple instances of the model.
+The statistical power is derived from the proportion of contrasts that result in the rejection of the null hypothesis across datasets.
+Although this approach permits the estimation of statistical power in the case of more complex models, it involves costly computations.
+In the case of Bayesian models, and particularly the one in the present study, such cost can be infeasible.
+Sampling the posterior of our model took approximately seven days.
+Running this model, or an equivalent one, across 100 datasets (100 may even be considered too few by many) would take more than a year.
 
-where $i$ and $j$ index the participant and translation equivalent (TE), $\beta_{0_k}$ is the fixed coefficient of the regression model for the intercept of threshold $k$, $u_{0_{i}}$ and $w_{0_{j}}$ are the by-participant and by-TE adjustments for $\beta_{0_{k}}$ (i.e., random intercepts), respectively, $\beta_{1-8}$ are the fixed coefficients of the regression model for the predictors of interest, $u_{1-8_{i}}$ and $w_{1-3_{j}}$ are the by-participant and by-TE adjustments for$\beta_{1-8}$ (i.e., random slopes), respectively.
+Following @kruschke2014doing, we considered the precision of our estimates as a proxy to statistical power.
+In particular, we compared the width of the 95% HDI of the critical regression coefficient (*Exposure* $\times$ *Cognateness*) against some nominal interval width.
+We decided to use the half the width of the ROPE in the logit scale [-0.025, +0.025], that is, 0.05 as the reference interval width.
+The width of the fixed regression coefficient of *Exposure* $\times$ *Cognateness* ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.value/4, 3)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower/4, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper/4, 3)`]) was `r abs(round(post_draws_list[["b_exposure_std:lv_std"]]$.lower/4 - post_draws_list[["b_exposure_std:lv_std"]]$.upper/4, 3))`, around `r  round(0.05 /  abs(post_draws_list[["b_exposure_std:lv_std"]]$.lower/4 - post_draws_list[["b_exposure_std:lv_std"]]$.upper/4), 3)` times narrower than the reference interval.
+This indicates that the precision of the posterior 95% HDI of the critical parameter in the model is larger than required.
 
-We used the Bayesian framework to estimate the parameters in our model.
-This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical].
-This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations.
-We used a weakly informative prior for our parameters, with the exception of the main effect of $\text{Age}$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words (see @eq-prior).
+### Model diagnostics
 
-$$
-\begin{aligned}
-\\
-\textbf{Prior:} \\
-\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
-\beta_{1} &\sim \mathcal{N}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
-\beta_{2-8} &\sim \mathcal{N}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
-u_{0-8_{i}} &\sim \mathcal{N}(0, \sigma_{u_{0-8_{i}}}) & [\mbox{Participant-level coefficient variability}] \\
-w_{0-3_{j}} &\sim \mathcal{N}(0, \sigma_{w_{0-3_{j}}}) & [\mbox{TE-level coefficient variability}] \\\\
-&&\mbox{[Participant-level coefficient variability]} \\ \\
-\Bigg(\begin{smallmatrix}
-u_{k_{0}} \\ 
-u_{1_{i}} \\ 
-\vdots \\ 
-u_{8_{i}} 
-\end{smallmatrix}\Bigg) &\sim \mathcal{N} 
-\Bigg(\Bigg(\begin{smallmatrix}0 \\
-0 \\ 
-\vdots \\
-0\end{smallmatrix}\Bigg), \Sigma_{u}\Bigg) \\
-\Sigma_{u} &= \Bigg(\begin{smallmatrix} \\
-\rho_{u_{0}} & \rho_{u_{0}} \sigma_{u_{0_{k}}} \sigma_{u_{1}} & \dots & \rho_{u_{0}} \sigma_{u_{0}} \sigma_{w_{8}}\\ 
-\rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{0}} & \rho_{u_{1}} & \dots & \rho_{u_{1}} \sigma_{u_{1}} \sigma_{u_{8}}\\ 
-\vdots & \vdots & \vdots & \vdots \\
-\rho_{8} \sigma_{u_{8}} \sigma_{u_{0_{k}}} & \dots & \dots & \rho_{u_{8}} \end{smallmatrix}\Bigg) \\
-\sigma_{u_{0-8}} &\sim \mathcal{N_{+}}(1, 0.1) \\
-\rho_{u} &\sim LKJcorr(2) \\
-\\
-&&\mbox{[TE-level coefficient variability]} \\ \\
-\Bigg(\begin{smallmatrix}
-w_{k_{0}}\\ 
-w_{1_{j}} \\ 
-\vdots \\ 
-w_{3_{j}} 
-\end{smallmatrix}\Bigg) &\sim \mathcal{N} \Bigg(\Bigg(\begin{smallmatrix}
-0\\ 
-0 \\ 
-\vdots \\
-0 
-\end{smallmatrix}\Bigg), \Sigma_{w}\Bigg) \\
-\Sigma_{w} &= \Bigg(\begin{smallmatrix} \\
-\rho_{w_{0}} & \rho_{w_{0}} \sigma_{w_{0_{k}}} \sigma_{w_{1}} & \dots & \rho_{w_{0}} \sigma_{w_{0}} \sigma_{w_{3}}\\ 
-\rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{0}} & \rho_{w_{1}} & \dots & \rho_{w_{1}} \sigma_{w_{1}} \sigma_{w_{3}}\\ 
-\vdots & \vdots & \vdots & \vdots \\
-\rho_{3} \sigma_{w_{3}} \sigma_{w_{0_{k}}} & \dots & \dots & \rho_{w_{3}} \end{smallmatrix}\Bigg) \\
-\sigma_{w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.1) \\
-\rho_{w_{0-3}} &\sim LKJcorr(2)
-\end{aligned}
-$$ {#eq-prior}
+One way to diagnose the behaviour of HMC is to inspect whether the different MCMC chains (if more than one) have converged to a similar region of the posterior.
+The Gelman-Rubin diagnostic [$\hat{R}$ or R-hat @gelman1992inference] provides a measure of chain convergence by comparing the variance within each chain *versus* the variance between each chain.
+Both are expected to be identical when chains have perfectly converged, so that $\hat{R} = 1$.
+Values lower than 1.01 are recommended, while values higher than 1.05 indicate that chains might have trouble converging and therefore the estimated parameters must be taken with caution.
+@fig-rhats-neffs (A) shows the distribution of $\hat{R}$ values for the coefficients of the fixed effect of our models, which we used for statistical inference.
+Most values are lower than 1.01, and never higher than 1.05, which provides evidence of successful MCMC convergence.
 
-where: $\rho_{u_{0-8}}$ and $\rho_{w_{0-3}}$ indicate the correlation parameters between the by-participant and by-TE adjustments, respectively, $\sigma_{u_{0-8}}^2$ and $\sigma_{w_{0-3}}^2$ indicate the variance of the by-participant and by-TE variance of the adjustments, respectively, and $\mathcal{N}$ indicates a normal distribution, $\mathcal{N}_{+}$ indicates a truncated normal distribution with only positive values, and $LKJcorr$ indicates a [LKJ correlation distribution](https://mc-stan.org/docs/2_22/functions-reference/lkj-correlation.html) [@lewandowski2009generating].
+Another diagnostic of good MCMC converge is the ratio of effective sample size to total sample size ($N_{eff}/N$), which indicates the proportion of samples in the chain that resulted from a non-divergent transition.
+Values closer to 1 are ideal, as they indicate that all posterior samples from the MCMC were used to estimate the posterior distribution of the parameter.
+Values larger than 0.1 are recommended.
+@fig-rhats-neffs (B) shows the distribution of the effective sample sizes of the coefficients of the fixed effects in our models.
+Most values are larger than 0.1, although model 0 ($\mathcal{M}_0$) accumulates most effective sample sizes close to 0.1.
 
-```{r apafg}
-#| label: fig-prior-mean
+```{r fig-rhats-neffs}
+#| label: fig-rhats-neffs
+#| fig-cap: "MCMC convergence diagnostic of all parameters in the model.
+Each dot represents the score of one parameter.
+(A) Distribution of the Gelman-Rubin (R-hat) scores.
+(B) Distribution of the ratio of effective sample size."
 #| echo: false
@@
 #| warning: false
-#| fig-width: 7
 #| fig-height: 3
-#| out-width: 90%
-#| fig-cap: "Expected prior-predicted mean."
-nd <- expand.grid(n_phon_std = 0,
-                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
-                  exposure_std = 0,
-                  lv_std = 0)
+#| fig-width: 6
+#| out-width: 100%
+model_convergence |> 
+    ggplot(aes(.rhat)) +
+    geom_dots(colour = "black") +
+    geom_vline(xintercept = 1.01, linetype = "dashed") +
+    labs(
+        x = "R-hat",
+        y = "MCMC samples"
+    ) +
+    model_convergence |> 
+    ggplot(aes(.neff)) +
+    geom_dots(colour = "black", na.rm = TRUE) +
+    geom_vline(xintercept = 1.01, linetype = "dashed") +
+    labs(
+        x = "Effective sample size ratio",
+        y = "MCMC samples"
+    ) +    
+    plot_layout(ncol = 1) &
+    plot_annotation(tag_levels = "A") &
+    theme(
+        legend.position = "none",
+        panel.grid = element_blank()
+    )
 
-prior_epreds <- add_epred_draws(nd, 
-                                model_fit_prior,
-                                ndraws = 50, 
-                                re_formula = NA) |> 
-    filter(.category != "No") |> 
-    pivot_wider(names_from = ".category", values_from = ".epred") |> 
-    mutate(Understands = Understands + `Understands and Says`) |> 
-    pivot_longer(c(Understands, `Understands and Says`),
-                 names_to = ".category",
-                 values_to = ".epred") |> 
-    mutate(.category = factor(.category,
-                              levels = c("Understands",
-                                         "Understands and Says"),
-                              labels = c("Comprehension\n(Understands)",
-                                         "Production\n(Understands and Says)")))
-
-prior_epreds |> 
-    ggplot(aes(x = age_std, y = .epred)) +
-    facet_wrap(~ .category) +
-    geom_line(linewidth = 3/4,
-              alpha = 1/2,
-              aes(group = .draw),
-              colour = "grey50") + 
-    labs(x = "Age (months)", 
-         y = "Probability of acquisition",
-         colour = "Credible interval", 
-         fill = "Credible interval") +
-    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
-                                      mean(responses$age), 
-                                      sd(responses$age))[, 1],
-                       labels = seq(10, 36, 4)) +
-    theme(legend.position = "top",
-          axis.ticks.x = element_line(colour = "black"),
-          panel.grid = element_blank())
+save_image("_assets/img/Figure_1_SuppInfo", 
+           dpi = 1e3L, width = 6, height = 3)
 ```
 
-{{< pagebreak >}}
+Another way of assessing the behaviour of the HMC algorithm is to visualise the joint posterior distribution for pairs of parameters using bi-variate scatter plots.
+In @fig-model-pairs we show the pair-wise distribution of posterior samples.
+Broadly, posterior samples of two parameters should not be correlated.
+This is the case for all pairs of parameters but for the two intercepts.
+This is expected behaviour, given that these two parameters correspond to the thresholds between categories in the ordinal regression model, and the distance between both thresholds is fixed in the particular parametrisation of the model.
+
+```{r}
+#| label: fig-model-pairs
+#| fig-cap: "Marginal distribution and bi-variate scatterplot of posterior samples for the fixed regression coefficients in Model 3."
+#| fig-height: 8
+#| fig-width: 10
+#| out-width: 100%
+#| out-height: 100%
+#| echo: false
+#| warning: false
+#| message: false
+titles <- c("Comprehension",
+            "Production",
+            "Age",
+            "Length",
+            "Exposure",
+            "Cognateness",
+            "Age × Exposure",
+            "Age × Cognateness",
+            "Exposure × Cognateness",
+            "Age × Exposure × Cognateness")
+
+posterior <- as_draws_df(model_fit) |>
+    select(`b_Intercept[1]`:`b_age_std:exposure_std:lv_std`)
+
+my_diag <- function(data, mapping, ...) {
+    ggplot(data = data, mapping = mapping) +
+        stat_slab(
+            colour = "white",
+            fill = "grey15"
+        )
+}
+
+my_upper <- function(data, mapping, ...) {
+    ggplot(data = data, mapping = mapping) +
+        stat_density_2d(aes(fill = ..density..),
+                        geom = "raster",
+                        contour = FALSE
+        ) +
+        scale_fill_distiller(
+            palette = "Greys",
+            direction = 1
+        )
+}
+
+posterior |> 
+    GGally::ggpairs(
+        upper = list(continuous = my_upper),
+        diag = list(continuous = my_diag),
+        lower = NULL,
+        columnLabels = c("Intercept\n(Comprehension)",
+                         "Intercept\n(Production)",
+                         "Age", "Length", "Exposure",
+                         "Levenshtein", "Age × Exposure",
+                         "Age × Levenshtein", "Exposure × Leveshtein",
+                         "Age × Exposure ×\nLevenshtein"
+        ),
+        axisLabels = "none"
+    ) +
+    theme(strip.text = element_text(size = 5),
+          panel.grid = element_blank(),
+          panel.border = element_blank()
+    )
+
+save_image("_assets/img/Figure_2_SuppInfo",
+           dpi = 1e3L, width = 10, height = 8)
+
+```
+
+Finally, we also assessed the predictive performance of the model by doing posterior-predictive checks (PPCs).
+this involves simulating new datasets from the model and the posterior distribution of its parameters, and checking that, overall, the distribution of the response variable across the simulated datasets is equivalent to the one in the observed dataset.
+@fig-ppc shows the PPCs of the model.
+
+```{r fig-ppc}
+#| label: fig-ppc
+#| fig-cap: "Model posterior predictive checks (PPC).
+Bars indicate the observed proportion of responses to each category (No, Understands, and Understands and Says).
+Blue dots and error bars represent the mean proportion of responses simulated from the posterior for each category, and its 95\\% interval."
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 6
+#| fig-height: 3
+#| out-width: 80%
+#| out-height: 80%
+y_int <- as.integer(as.integer(model_fit$data$response))
+bayesplot::color_scheme_set(c(rep("grey60", 2), rep("black", 4)))
+
+bayesplot::ppc_bars(
+    y_int,
+    model_ppcs,
+    prob = 0.95,
+    linewidth = 1,
+    size = 1.25,
+    fatten = 2,
+    freq = FALSE
+) +
+    scale_x_continuous(
+        breaks = 1:3,
+        labels = c(
+            "No",
+            "Understands",
+            "Understands and Says"
+        )
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text = element_text(size = 11),
+        panel.grid = element_blank()
+    )
+
+save_image("_assets/img/Figure_3_SuppInfo",
+           dpi = 1e3L, width = 6, height = 3)
+```

### 2023-10-03 manuscript/_appendix/_model-doe.qmd
Ignore .tiff, .eps, and .svg

--- 2023-07-06 manuscript/_appendix/_model-doe.qmd
@@
 #| label: tbl-coefs-doe
-#| fig-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency (*Frequency*) and degree of exposure (*DoE*) separately.
-Median: median of the posterior distribution in the probability scale.
-95\\% HDI: 95\\% highest density interval of the distribution.
-p(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.
-In the right column, we show a graphical representation of the median and 95% HDI of each coefficient, with a reference dotted line indicating the location of zero."
+#| tbl-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency (*Frequency*) and degree of exposure (*DoE*) separately."
 #| echo: false
@@
                       ", ", round(.upper, 3), "]"),
-        .rope_overlap = sub("^(-?)0.", "\\1.",
-                            sprintf("%.3f", .rope_overlap))) |> 
-    select(.variable_name, .median, .lower, .upper, .hdi, .rope_overlap)
+        .rope = sub("^(-?)0.", "\\1.",
+                    sprintf("%.3f", .rope))) |> 
+    select(.variable_name, .value, .lower, .upper, .hdi, .rope)
+
+col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")
 
 tbl_data |> 
-    mutate(viz = "") |> 
     select(-c(.lower, .upper)) |> 
@@
         escape = FALSE,
-        col.names = c("",
-                      "$\\beta$",
-                      "95\\% HDI",
-                      "$p(\\text{ROPE})$",
-                      ""),
+        col.names = col_names,
         digits = 3,
@@
                           start_row = 9,
-                          end_row = 17)
+                          end_row = 17) |> 
+    add_footnote("$\\beta$: median of the posterior distribution in the probability scale.
+95\\% HDI: 95\\% highest density interval of the distribution.
+$p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none", escape = FALSE)
 ```

### 2023-10-03 manuscript/_appendix/_syllable-frequency.qmd
Ignore .tiff, .eps, and .svg

@@
 items <- items |> 
-    mutate(xsampa = ipa::ipa(ipa)) |> 
-    mutate(syll = strsplit(xsampa, 
+    mutate(xsampa = ipa::ipa(ipa),
+           syll = strsplit(xsampa, 
                            split = paste0(.sep = c("\\.", "\\\""), 
@@
 #| label: tbl-syll-items
-#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
+#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish.
 #| echo: false
@@
 tbl_data |> 
-    kable(format = "latex", 
-          digits = 3,
+    kable(digits = 3,
+          format = "latex",
           booktabs = TRUE,
@@
     facet_wrap(~ language) +
-    geom_col(colour = "white", fill = "grey70") +
+    geom_col(colour = "white", fill = "black") +
     geom_text(aes(label = scales::percent(prop)),
@@
     theme(legend.position = "none",
-          panel.grid.minor = element_blank(),
-          panel.grid.major.x = element_blank())
+          panel.grid = element_blank())
 ```
@@
                    bins = 20,
-                   fill = "grey70",
+                   fill = "black",
                    position = position_nudge(y = 0.1)) +
@@
 
-To estimate the association between word-level syllabic frequency and cognateness, while controlling for the number of syllables in the word, as words are expected to necessarily increase the syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ [Levenshtein similarity between a word and its translation equivalent, @levenshtein1966binary] as predictors.
+To estimate the association between word-level syllabic frequency and cognateness, while controlling for the number of syllables in the word, (an increment in the number of syllables necessarily increases the summed syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ [Levenshtein similarity between a word and its translation equivalent, @levenshtein1966binary] as predictors.
 We added translation equivalent-level random effects for the intercept and the main effect of $Syllables$ (some translation pairs had a different number of syllables in each language).
@@
 rope_interval <- c(lower = -0.5, upper = 0.5)
-# tidy predictor names
 
+# tidy up predictor names
 post_draws_list <- posterior_syllables_summary |> 
@@
         .rope = sub("^(-?)0.", "\\1.",
-                    sprintf("%.3f", .rope))) |> 
+                    sprintf("%.3f", .rope)),
+        .rope = ifelse(is.na(.rope), "-", .rope)) |> 
     select(.variable_name, .value, .lower, .upper, .hdi, .rope)
+
+col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")
 
@@
         escape = FALSE,
-        col.names = c("",
-                      "$\\beta$",
-                      "95\\% HDI",
-                      "$p(\\text{ROPE})$"),
+        col.names = col_names,
         digits = 3,
@@
     ) |> 
-    kableExtra::kable_styling()
+    kableExtra::kable_styling() |> 
+    add_footnote("$\\beta$: median of the posterior distribution in the probability scale.
+95\\% HDI: 95\\% highest density interval of the distribution.
+$p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none")
 ```
@@
 #| fig-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
-#| fig-width: 7
-#| fig-height: 5
+#| fig-width: 9
+#| fig-height: 2.5
 #| out-width: 80%
@@
 #| warning: false
-expand_grid(lv_std = scale(seq(0, 1, 0.1),
+syll_plot_data <- expand_grid(
+    lv_std = scale(seq(0, 1, 0.1),
                    center = mean(syllables_data$lv),
@@
         n_syll_std = (n_syll_std * sd(syllables_data$n_syll)) + mean(syllables_data$n_syll),
-        n_syll_std = as.factor(n_syll_std)) |> 
-    ggplot(aes(lv_std, .epred, 
-               shape = n_syll_std)) +
+        n_syll_std = as.factor(n_syll_std)
+    )
+
+ggplot(syll_plot_data, aes(lv_std, .epred)) +
+    facet_grid(~n_syll_std) +
     geom_point(data = filter(syllables_data, n_syll < 5) |>
@@
                aes(x = lv,
-                   y = freq_syll,
-                   shape = n_syll_std),
-               alpha = 1/2) +
-    # geom_ribbon(aes(ymin = .lower,
-    #                 ymax = .upper),
-    #             alpha = 1/2,
-    #             colour = NA,
-    #             show.legend = FALSE) +
-    geom_line(aes(linetype = n_syll_std),
-              linewidth = 0.5) +
+                   y = freq_syll),
+               alpha = 1/5,
+               shape = 1) +
+    geom_ribbon(aes(ymin = .lower,
+                    ymax = .upper),
+                alpha = 1/2,
+                colour = NA,
+                show.legend = FALSE) +
+    geom_line(linewidth = 0.5) +
     labs(x = "Cognateness (Levenshtein similarity)",
@@
           legend.position = "top")
+
+suppressMessages({
+    invisible({
+        devices <- c("png", "pdf", "svg", "tiff", "eps")
+        paths <- paste("_assets/img/Figure_6_SuppInfo", devices, sep = ".")
+        purrr::map(paths, \(x) ggsave(x, dpi = 1e3L, width = 9, height = 2.5, units = "in"))
+    })
+})
 ```

### 2023-10-03 manuscript/appendix.qmd
Ignore .tiff, .eps, and .svg

@@
 targets::tar_load(c(
-    bvq_data, childes, items, model_summary,
-    participants, responses
+    bvq_data, childes, 
+    items,
+    model_summary,
+    participants, 
+    responses
 ))
 
-targets::tar_load(c(model_fit, model_convergence, model_ppcs))
+targets::tar_load(c(model_fit, 
+                    model_convergence,
+                    model_ppcs))
 
-targets::tar_load(c(posterior_doe_summary, posterior_doe_draws))
+targets::tar_load(c(posterior_doe_summary, 
+                    posterior_doe_draws))
 
-targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables, posterior_syllables_summary))
+targets::tar_load(c(syllables_data,
+                    model_fit_syllables, 
+                    post_syllables,
+                    posterior_syllables_summary))
 
 my_theme <- theme_minimal() +
-    theme(
-        panel.grid = element_line(
+    theme(panel.grid = element_line(
         colour = "grey",
@@
 theme_set(my_theme)
+
+save_image <- function(file, ...) {
+    suppressWarnings({
+        invisible({
+            devices <- c("png", "pdf", "svg", "tiff", "eps")
+            paths <- file.path(paste(file, devices, sep = "."))
+            purrr::map(paths,
+                       \(x) ggsave(x, ...),
+                       .progress = TRUE)
+        })
+    })
+}
+
+post_draws_list <- model_summary |>
+    mutate(across(
+        .value:.upper,
+        \(x) ifelse(grepl("Intercept", .type),
+                    plogis(x), x / 4
+        )
+    )) |>
+    split(model_summary$.variable)
 ```
 
-## Appendix A: model diagnostics {.appendix .unnumbered}
+## Appendix A: model details {.appendix .unnumbered}
 
-{{< include _appendix/_model-diagnostics.qmd >}}
+{{< include _appendix/_model-details.qmd >}}
 

### 2023-10-03 manuscript/manuscript.qmd
Ignore .tiff, .eps, and .svg

@@
 We predicted a stronger cognate facilitation for words to which children were exposed less frequently (low-exposure words), as they are co-activated by their translation more often than high-exposure words.
-We developed an extensive online vocabulary checklist, the Barcelona Vocabulary questionnaire (BVQ), to collect vocabulary data from 366 Catalan-Spanish bilingual toddlers aged 12 to 32 months.
+We developed an extensive online vocabulary checklist, the Barcelona Vocabulary questionnaire (BVQ), to collect vocabulary data from 369 Catalan-Spanish bilingual toddlers aged 12 to 32 months.
 We used Bayesian explanatory item response theory to model the acquisition trajectories of 604 Catalan and Spanish words.
@@
 GGC was supported by a FPI research contract (PRE2019-088165).
-DAV was supported by the European Union's Horizon 2023 research and innovation program under Marie Skłodowska–Curie Grant (765556) and a postdoctoral fellowship from the Foundation for Science and Technology of Portugal (UIDB/00214/2020).
+DAV was supported by the European Union’s Horizon2020 research and innovation program under Marie Skłodowska–Curie Grant (765556) and partially supported by Portuguese national funds through the Foundation for Science and Technology, under project UIDB/00214/2020, awarded to the Center of Linguistics of the University of Lisbon.
 IC was supported by the Investigo program funded by the European Union's NextGenerationEU (NGEU) recovery plan.
-NSG was supported by an ICREA Academia award from the Catalan Institution for Research and Advanced Studies (ICREA).
+NSG was supported by an ICREA Academia 2019 award from the Catalan Institution for Research and Advanced Studies (ICREA).
 We are grateful to Chiara Santolin, Ege E.
@@
 
+clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
+
 set.seed(888)
@@
 For example, English-Dutch bilinguals (languages with a high phonological overlap), were able to produce more Dutch words than English-Mandarin bilinguals (languages with a low phonological overlap) were able to produce in Mandarin.
-@blom2020crosslanguage, @bosma2019longitudinal, and @gampe2021does reported similar results, providing converging evidence of a facilitatory effect of a lower language distance (i.e., higher degre of cognateness) on vocabulary size.
+@blom2020crosslanguage, @bosma2019longitudinal, and @gampe2021does reported similar results, providing converging evidence of a facilitatory effect of a lower language distance (i.e., higher degree of cognateness) on vocabulary size.
 
@@
 The authors collected expressive vocabulary data of 47 16- to 30-month-old French-English bilinguals living in Canada, in both languages.
-They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates.
-The proportion of words that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages.
+They created two lists of translation equivalents; one made of 131 cognates, and one made of 406 non-cognates.
+Across ages, the proportion of words that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list.
 Critically, this difference persisted after both lists were matched in size, controlling their semantic category (i.e., furniture, animals, food were similarity represented in both lists) and age-of-acquisition norms (an index of word difficulty).
-Taken together, the results of these two lines of research have been considered as supporting to the hypothesis that phonological similarity (as reflected in cognateness) plays a facilitation role in bilingual word acquisition.
+Taken together, the results of these two lines of research support the hypothesis that phonological similarity (as reflected in cognateness) plays a facilitation role in bilingual word learning.
 
@@
 These results showed that bilinguals’ Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.
-More recntly, evidence of parallel activation has been reported in bilingual toddlers and children too [@vonholzen2012language; @floccia2020translation; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @poulin-dubois2013lexical].
-Although there is a consensus on the role of parallel activation in bilinguals' lexical processing and acquisition, previous studies do not address its influence on the learning trajectories of words.
+More recently, evidence of parallel activation has been reported in bilingual toddlers and children too [@vonholzen2012language; @floccia2020translation; @vonholzen2019impact; @bosma2020cognate; @poarch2012crosslanguage; @poulin-dubois2013lexical].
+Although there is a consensus on the role of parallel activation in bilinguals' lexical processing and acquisition, previous studies do not address its influence on the learning trajectories of individual words.
 Results are aggregated across words and provide no information about the specific dynamics of how parallel activation influences word learning.
@@
 Parallel activation would allow bilingual children to accumulate learning instances for words in both languages even during fully monolingual situations, but the impact of this mechanism would be asymmetric across languages: words from the lower-exposure language would receive stronger activation from words in the higher-exposure language than vice versa.
-Therefore, the acquisition of words from the lower-exposure language would benefit more strongly from their cognate status than words from the higher-exposure language.
-This asymmetric cross-language activation would be consistent with previous reports of larger priming effects from the dominant to the non-dominant language.
+Therefore, the acquisition of words from the lower-exposure language would benefit more strongly from their cognate status than the acquisition of words from the higher-exposure language.
+This asymmetric cross-language activation would be consistent with previous reports of larger priming effects from the dominant to the non-dominant language [e.g., @grainger1998masked].
 
@@
 Therefore, this exposure will count as a learning instance for both co-activated forms.
-The case of the non-cognate translation equivalent /\textipa{"gos-"pe.ro}/ [*dog*] would be different.
+The case of the non-cognate translation equivalent /\textipa{"gos}/--/\textipa{"pe.ro}/ [*dog*] would be different.
 Given the low phonological similarity between both word-forms, an exposure to \textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/ leading to such exposure counting as a learning instance for /\textipa{"gos}/ (which the child was exposed to), but not for /\textipa{"pe.ro}/.
-While /\textipa{"gat}/--/\textipa{"ga.to}/ will benefit from phonological co-activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not.
+While /\textipa{"gat}/--/\textipa{"ga.to}/ will benefit from phonological co-activation, /\textipa{"gos}--\textipa{"pe.ro}/ will not.
 If the child receives linguistic input from one of the languages more often than from the other, this effect might affect each form of the cognate translation equivalent differently.
@@
 We estimated the impact of several predictors of interest on the probability of acquisition, including participants’ age and rate of exposure to the word-form, and the cognate status of the word-form.
-As described in the methods section, rate of exposure was a composite measure taking into account participant’ language exposure and word’s lexical frequency.
+As described in the Methods section, rate of exposure was a composite measure taking into account participant’ language exposure and word’s lexical frequency.
 We predicted an interaction between cognate status and word-form exposure rate in which the probability of comprehension is higher for low-exposure cognate words, but not for high-exposure cognate words.
@@
 All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/hy984/)) and GitHub ([https://github.com/gongcastro/cognate-beginnings](https://github.com/gongcastro/cognate-beginnings)) repositories.
+For reproducibility, a Docker image of the RStudio session is available on DockerHub([https://hub.docker.com/repository/docker/gongcastro/cognate-beginnings/](https://hub.docker.com/repository/docker/gongcastro/cognate-beginnings/)).
 This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.
@@
     mutate(version = tolower(version))
-
 ```
@@
 This questionnaire is structured in three blocks: (1) a language exposure questionnaire, (2) a demographic survey, and (3) two vocabulary checklists.
-Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish.
+Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted of two lists of words, one in Catalan and one in Spanish.
 Both lists included items from a diverse sample of `r n_categories` semantic or functional categories.
@@
 Items in one language were translation equivalents of the items in the other language (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for *dog*), roughly following a one-to-one mapping.
-Some of the words in Catalan did not have a clear translation or had more than one possible translation in Spanish, and vice versa, therefore the unequal number of words included in the two lists (see @tbl-items for a summary of the questionnaire items).
+Some of the words in Catalan did not have a clear translation or had more than one possible translation in Spanish, and vice versa, therefore the unequal number of words included in the two lists.
 
@@
     } else {
-        "N"
+        "*N*"
     }
@@
     summarise(
-        semantic_category = "N",
+        semantic_category = "*N*",
         across(a:d, sum),
@@
     booktabs = TRUE,
+    escape = FALSE,
     col.names = col_names,
@@
 For each word included in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default).
-Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items: A, B, C, and D.
-Each subset contained a random but representative sub-sample of the items from the complete list (see @tbl-items).
-Semantic or functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided: all of their items were included in the four lists.
+Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items (A, B, C, and D) Each subset contained a random but representative sub-sample of the items from the complete list (see @tbl-items).
+Semantic or functional categories with less than 16 items---thus resulting in less than four items after dividing it in four subsets---were not divided: all of their items were included in the four subsets.
 Items that were part of the trial lists of some ongoing experiments in the lab were also included in all versions.
 The resulting reduced list contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words.
-Participants were randomly allocated into one of the four subsets.
 
@@
 participants_edu <- table(participants$edu_parent)
+
+participants_age <- summarise(participants,
+                              across(age, lst(mean, median, sd, min, max)))
+
+participants_sex <- c(table(distinct(participants, id, sex)$sex),
+                      missing = sum(is.na(distinct(participants, id, sex)$sex)))
 ```
 
-We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between Marth 30th, 2020 and October 31th, 2022: `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times.
-Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)` days, and were always allocated to the same questionnaire list (A, B, C, or D).
-Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra) and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample).
+We collected `r n_responses` questionnaire responses from `r sum(n_participants)` distinct children (`r participants_sex[["female"]]` female, `r participants_sex[["male"]]` male, `r participants_sex[["missing"]]` not reported, mainly White).
+Participants were aged 12-32 months (*M* = `r round(participants_age$age_mean, 2)`, *SD* = `r round(participants_age$age_sd, 2)`, *Range* = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`).
+Of those participants, `r english::words(n_participants[4])` participated four times, `r english::words(n_participants[3])` participated three times, `r n_participants[2]` participated twice, and `r n_participants[1]` participated once.
+Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)` days.
+Participants were randomly allocated into one of the four questionnaire subsets (A, B, C, or D).
+Each participant was always allocated to the same subset.
+
+Participants were residents in the Metropolitan Area of Barcelona (Catalonia, Spain).
+Data collection took place between March 30th, 2020 and October 31th, 2022.
+Participants were part of the database of the Laboratori de Recerca en Infància at the Universitat Pompeu Fabra, and were contacted by e-mail or phone if their child was aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample).
 In total, `r n_trilingual$n[2]` participants (`r scales::percent(n_trilingual$prop[2], accuracy = 0.01)`) participants were reported to be exposed in less than 10% to a third language other than Catalan and Spanish.
@@
 #| tbl-cap: "Participant sample size by age and degree of exposure to Catalan."
-#| tbl-note: "For participants exposed to Catalan and Spanish exclusively, the proportion of Catalan exposure shown in the table is complementary to the degree of exposue to Spanish.
-For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10%.
-Participants with 100\\%, 0\\%, and 50\\% Catalan exposure would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively."
 ftn <- ""
@@
         hline_after = TRUE
-    )
+    ) |>
+    add_footnote("For participants exposed to Catalan and Spanish exclusively, the proportion of Catalan exposure shown in the table is complementary to the degree of exposue to Spanish.
+For participants exposed to a third languages this proportion is not complementary unless one adds the degree of exposed to the third language, which never exceeded 10\\%.
+Participants with 100\\%, 0\\%, and 50\\% Catalan exposure would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively.", notation = "none", escape = FALSE)
 ```
@@
 This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: *sense escolaritzar/sin escolarizar* [no education], *educació primària/educación primaria* [primary school], *educació secundària/educación secundaria* [secondary school], *batxillerat/bachillerato* [complementary studies/high school], *cicles formatius/ciclos formativos* [vocational training], and *educació universitària/educación universitaria* [university degree].
-Most families reported university studies (`r participants_edu["University"]`, `r scales::percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r scales::percent(participants_edu["Vocational"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r scales::percent(participants_edu["Secondary"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r scales::percent(participants_edu["Complementary"]/sum(participants_edu))`), primary education (`r participants_edu["Primary"]`, <1%), and no formal education (`r participants_edu["No education"]`, <1%).
+Most families reported university studies (`r participants_edu["University"]`, `r scales::percent(participants_edu["University"]/sum(participants_edu))`), followed by families where the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r scales::percent(participants_edu["Vocational"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r scales::percent(participants_edu["Secondary"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r scales::percent(participants_edu["Complementary"]/sum(participants_edu))`), primary education (`r participants_edu["Primary"]`, <1%), and no formal education (`r participants_edu["No education"]`, <1%).
 
@@
 We collected data for `r format(nrow(n_items), big.mark = ",")` words.
-We restricted the analyses to responses to nouns [`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded, see @fourtassi2020growth for a similar approach].
-We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`, see @sec-predictors), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
+We restricted the analyses to responses to nouns (`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded).
+We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`, see Model predictors section), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`).
 Finally, we removed items without a translation in the other language (*n* = `r nrow(n_items_multi)-nrow(n_items_duplicated)`).
@@
 We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, multilevel, ordinal regression model.
-This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest: *Age* (number of months elapsed between participants' birth date and questionnaire completion), *Length* (number of phonemes in the X-SAMPA phonological transcription of the word-form), *Exposure* (a language exposure-weighted lexical frequency), and *Cognateness* (defined as the phonological similarity between translation equivalents).
+This model allowed us to estimate both item and participant word-acquisition trajectories, while estimating the effect of our variables of interest: *Age* (number of months elapsed between participants' birth date and questionnaire completion), *Length* (number of phonemes in the X-SAMPA phonological transcription of the word-form), *Exposure* (a language exposure-weighted lexical frequency score), and *Cognateness* (defined as the phonological similarity between translation equivalents).
+A more detailed descriptions of these predictors is provided in the Model preedictors section.
 We added these variables as main effects, together with the two-way and three-way interactions between *Age*, *Exposure*, and *Cognateness*.
@@
 @eq-model shows a detailed description of the model.
+See Appendix A for a detailed description of the model and its diagnostics.
 
@@
 
-We implemented the model using `brms` [@burkner2017brms], a R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
-We ran four iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
+We implemented the model using `brms` [@burkner2017brms], an R interface to the Stan probabilistic language (`r stan_version`) [@carpenter2017stan].
+We ran `r english::words(dim(model_fit$fit)[2])` iteration chains using the by-default No U-Turn Sampler algorithm with `r format(niterations(model_fit), big.mark = ",")` iterations each and an additional `r format(niterations(model_fit), big.mark = ",")` warm-up iterations per chain.
 
@@
 
-We defined *Cognateness* as the phonological similarity between each word-form and its translation.
+Following Floccia et al., we defined *Cognateness* as the phonological similarity between each word-form and its translation.
 For each translation equivalent, we used the `stringdist` [@vanderloo2014stringdist] R package to calculate the Levenshtein distance between the Catalan and the Spanish phonological transcriptions of the word-forms.
@@
 
-We assessed the practical relevance of the estimated regression coefficients of the model model following @kruschke2018bayesian.
+We assessed the practical relevance of the estimated regression coefficients of the model following @kruschke2018bayesian.
 First, we specified a region of practical equivalence (ROPE) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`, in the probability scale.
@@
 For example, $p(\text{ROPE})=.80$ indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero.
+See Appendix A for considerations about statistical power and sample size.
 
@@
 The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor.
+
+We now present our confirmatory analysis, and report a summary of the regression coefficients of interest.
+@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients, and their degree of overlap with the ROPE.
+For interpretability, we report the estimated regression coefficients transformed to the probability scale.
+The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) equivalent to a one standard deviation change in each predictor.
 
@@
         end_row = 10
-    )
+    ) |>
+    add_footnote("$\\beta$: median of the posterior distribution in the probability scale.
+95\\% HDI: 95\\% highest density interval of the distribution.
+$p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none", escape = FALSE)
 ```
@@
 For every phoneme in the word-form, participants were `r round((post_draws_list[["b_n_phon_std"]]$.value)/sd(responses$n_phon), 3)` probability points less likely to know it.
-The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word-form exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children.
+The 95% HDI of the regression coefficient of the *Age* $\times$ *Exposure* interaction also excluded the ROPE ($\beta$ = `r round(post_draws_list[["b_age_std:exposure_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_age_std:exposure_std"]]$.lower, 3)`, `r round(post_draws_list[["b_age_std:exposure_std"]]$.upper, 3)`]), showing that the effect of the word-form exposure index differed across ages: older children were more likely to acquire words with a higher exposure rate than younger children were.
 
-Around `r scales::percent(post_draws_list[["b_lv_std"]]$.rope, accuracy = 0.01)` of the posterior samples of the main effect of *Cognateness* overlapped with the ROPE completely ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 3)`]).
+Around `r scales::percent(post_draws_list[["b_lv_std"]]$.rope, accuracy = 0.01)` of the posterior samples of the main effect of *Cognateness* overlapped with the ROPE ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 3)`]).
 For every 10% increment in cognateness, the probability of word acquisition increased in `r round(post_draws_list[["b_lv_std"]]$.value/(40*sd(responses$lv)), 3)`.
-The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to the word-form.
+The effect of *Cognateness* interacted with that of *Exposure*: the 95% HDI of the regression coefficient of interaction excluded the ROPE entirely ($\beta$ = `r round(post_draws_list[["b_exposure_std:lv_std"]]$.value, 3)`, 95% HDI = [`r round(post_draws_list[["b_exposure_std:lv_std"]]$.lower, 3)`, `r round(post_draws_list[["b_exposure_std:lv_std"]]$.upper, 3)`]), suggesting that the effect of cognateness on a word's probability of acquisition changed depending on participants' exposure to the word-form.
 When *Exposure* was low (e.g., -1 SD), *Cognateness* increased the probability of acquisition substantially, while this effect was negligible for words with median or high exposure (+1 SD) (see @fig-marginal).
@@
         ymax = Inf,
-        fill = "orange",
+        fill = "grey",
         alpha = 1 / 4,
@@
         alpha = 0.5,
-        colour = "orange"
+        colour = "grey"
     ) +
@@
         alpha = 0.5,
-        colour = "orange"
+        colour = "grey"
     ) +
-    ggdist::geom_lineribbon(aes(order = lv),
-        size = 0.5
-    ) +
+    geom_ribbon(linewidth = 0,
+                alpha = 3/5) +
+    geom_line(linewidth = 2/4,
+              colour = "black") +
+
+    # ggdist::geom_lineribbon(aes(order = lv),
+    #                         size = 0
+    # ) +
+
     labs(
@@
     scale_linetype_manual(values = rev(c("solid", "dashed", "dotdash"))) +
-    scale_colour_grey(start = 0, end = 0) +
-    scale_fill_grey(start = 0.85, end = 0.35) +
+    scale_colour_manual(values = clrs[c(1, 4, 5)]) +
+    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
     scale_y_continuous(
@@
     )
+
+suppressMessages({
+    invisible({
+        devices <- c("png", "pdf", "svg", "tiff", "eps")
+        paths <- paste("_assets/img/Figure_1", devices, sep = ".")
+        purrr::map(paths, \(x) ggsave(x, dpi = 1e3L, width = 18, units = "cm"))
+    })
+})
 ```
@@
 Overall, we found that cognates (i.e., phonologically similar translation equivalents) were acquired earlier than non-cognates.
-This effect was mediated by exposure rate: low-exposure word-forms benefited from their cognate status, whereas high-exposure word-forms did not.
+This effect was mediated by exposure rate.
+Low-exposure word-forms benefited from their cognate status, whereas high-exposure word-forms did not.
 Using the concept of accumulator [see @kachergis2022standard for review], we provide a theoretical account of bilingual lexical acquisition.
@@
 First, @floccia2018introduction reported that bilingual toddlers learning two typologically close languages (e.g.
-shared many cognates, like., English-Dutch) showed larger vocabulary sizes than those learning typologically distant languages (e.g.
-shared fewer cognates, like English-Mandarin).
+shared many cognates, like English-Dutch) showed larger vocabulary sizes than those learning typologically distant languages (e.g., shared fewer cognates, like English-Mandarin).
 Second, @mitchell2022cognates found an earlier age-of-acquisition for cognates, compared to non-cognates.
-The outcomes of both studies pointed to cognateness facilitating word acquisition through parallel activation.
-But the underpinnings of such effect were unclear: while parallel activation has been extensively described in experimental studies, current paradigms of bilingual word acquisition and word learning are, to a large extent, dissociated from the mechanisms proposed by previous work on word processing.
+The outcomes of both studies pointed to cognateness facilitating word acquisition through parallel activation, but the underpinnings of such effect were unclear.
+While parallel activation has been extensively described in experimental studies, current paradigms of bilingual word acquisition and word learning are, to a large extent, dissociated from the mechanisms proposed by previous work on word processing.
 The notion of *accumulator*, as conceptualised by accumulator models of language acquisition, may provide a convenient theoretical framework to narrow this gap.
@@
 
-These predictions address a critical subject in bilingualism research: do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
+These predictions address a critical subject in bilingualism research.
+Do bilingual infants accumulate learning experiences in both languages independently, or does exposure to one language impact the acquisition trajectory of the other language?
 In the context of lexical acquisition, the former scenario predicts that every learning instance for a given word-form contributes to the acquisition of the representation of such word in the lexicon, while the acquisition of its translation remains unaffected by such experience.
@@
 This mechanism might be extended to provide a plausible explanation for the language similarity facilitation reported by Floccia et al.
-The authors observed a facilitation in the additional (non-English) language: children learning two typologically close languages knew more words in the additional language than those learning two typologically more distant languages.
+The authors observed a facilitation in the additional (non-English) language.
+Children learning two typologically close languages knew more words in the additional language than those learning two typologically more distant languages.
 In their sample, the additional language was consistently also the lower-exposure language for most children, while English was the higher-exposure language.
@@
 Because cognates correspond to similar sounding word-forms in equivalent referential contexts (e.g., hearing /\textipa{"gat}/ and /\textipa{"ga.to}/ in the same situations), it is possible that infants classify both are as acceptable variations of the same word-form, therefore treating them as a single lexical item.
-This would lead to a faster increase in cumulative learning instances, and to earlier ages of acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
+This would lead to a faster increase in cumulative learning instances, and to an earlier age-of-acquisition for cognate translation equivalents (for which listening to each word-form contributes to the acquisition of its shared representation), compared to non-cognates (for which listening to each word-form contributes to the acquisition of a separate representation).
 This mechanism could potentially explain the earlier age-of-acquisition effect of cognates found in the present study, without the need of parallel activation playing any relevant role.
-@mitchell2022cognates discuss this possibility as a possible explanation of the cognate facilitation effect, in which bilinguals only need to map one word-form to the referent in the case of cognates, while mapping two distinct word-forms in the case of non-cognates.
+@mitchell2022cognates discuss this possibility as a candidate explanation of the cognate facilitation effect, in which bilinguals only need to map one word-form to the referent in the case of cognates, while mapping two distinct word-forms in the case of non-cognates.
 However, previous work on mispronunciation perception and learning of minimal pair words points in a different direction.
@@
 Bilinguals differ across multiple dimensions.
-Such differences span from exclusively linguistic factors; such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high like the case of Spanish and Greek), to extralinguistic factors like the socio-linguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages hardly has any societal presence, i.e., heritage languages).
+Such differences span from exclusively linguistic factors; such as the amount of overlap between the phonemic inventories of the two languages being learned (e.g., low, like the case of English and Mandarin, or high, like the case of Spanish and Greek), to extralinguistic factors like the sociolinguistic situation in which the two languages co-exist (e.g., in some regions both languages are co-official and used in similar contexts, while in others, one of the languages has a smaller societal presence, i.e., heritage languages).
 This diversity of situations in which bilingual toddlers acquire language calls for special consideration of the generalisability of results in bilingualism research.
-Our sample, although homogeneous (e.g., similar parental educational level across), represents a particular bilingual sociolinguistic environment: the languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages, both languages are used in fairly similar contexts, and both languages are known by the majority of the population.
+Our sample, although homogeneous (e.g., similar parental educational level across participants), represents a particular bilingual sociolinguistic environment.
+The languages involved in the present investigation, Catalan and Spanish, co-exist in Catalonia as official languages, both languages are used in fairly similar contexts, and both languages are known by the majority of the population.
 In 2018, more than 81.2% of a representative sample of 8,780 adults aged 15 years or older living in Catalonia reported being able to speak Catalan, and more than 99.5% of the same population reported being able to speak Spanish [@2018els].
@@
 To conclude, our study provides novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon.
-We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word-forms's phonological similarity with its translation in the other language.
+We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word-form's phonological similarity with its translation in the other language.
 Capitalising on the notion of *accumulator* of linguistic input, we put forward a theoretical account of bilingual word learning, in which cognateness interacts with lexical frequency and language exposure to boost the acquisition of translation equivalents.

### 2023-10-03 manuscript/manuscript.tex
Ignore .tiff, .eps, and .svg

@@
 GGC was supported by a FPI research contract (PRE2019-088165). DAV was
-supported by the European Union's Horizon 2023 research and innovation
-program under Marie Skłodowska--Curie Grant (765556) and a postdoctoral
-fellowship from the Foundation for Science and Technology of Portugal
-(UIDB/00214/2020). IC was supported by the Investigo program funded by
-the European Union's NextGenerationEU (NGEU) recovery plan. NSG was
-supported by an ICREA Academia award from the Catalan Institution for
-Research and Advanced Studies (ICREA). We are grateful to Chiara
-Santolin, Ege E. Özer, and the rest of the Speech Acquisition and
-Perception research group, and to Alicia Franco-Martínez and Cristina
-Rodríguez-Prada, for their helpful feedback. We thank Xavier Mayoral,
-Silvia Blanch, and Cristina Cuadrado for their technical support, and
-Cristina Dominguez and Katia Pistrin for their efforts in recruiting
-infants. We also thank all families and infants who participated in the
-experiments.}}
+supported by the European Union's Horizon2020 research and innovation
+program under Marie Skłodowska--Curie Grant (765556) and partially
+supported by Portuguese national funds through the Foundation for
+Science and Technology, under project UIDB/00214/2020, awarded to the
+Center of Linguistics of the University of Lisbon. IC was supported by
+the Investigo program funded by the European Union's NextGenerationEU
+(NGEU) recovery plan. NSG was supported by an ICREA Academia 2019 award
+from the Catalan Institution for Research and Advanced Studies (ICREA).
+We are grateful to Chiara Santolin, Ege E. Özer, and the rest of the
+Speech Acquisition and Perception research group, and to Alicia
+Franco-Martínez and Cristina Rodríguez-Prada, for their helpful
+feedback. We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado
+for their technical support, and Cristina Dominguez and Katia Pistrin
+for their efforts in recruiting infants. We also thank all families and
+infants who participated in the experiments.}}
 \author{
@@
 for Brain and Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\texttt{\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}}\And
+08005\\\texttt{\href{mailto:daniela.avila@upf.edu}{daniela.avila@upf.edu}}\And
 \textbf{Ignacio
@@
 extensive online vocabulary checklist, the Barcelona Vocabulary
-questionnaire (BVQ), to collect vocabulary data from 366 Catalan-Spanish
+questionnaire (BVQ), to collect vocabulary data from 369 Catalan-Spanish
 bilingual toddlers aged 12 to 32 months. We used Bayesian explanatory
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
 
@@
 providing converging evidence of a facilitatory effect of a lower
-language distance (i.e., higher degre of cognateness) on vocabulary
+language distance (i.e., higher degree of cognateness) on vocabulary
 size.
@@
 bilinguals living in Canada, in both languages. They created two lists
-of translation equivalents: one made of 131 cognates, and one made of
-406 non-cognates. The proportion of words that children were reportedly
-able to produce was higher in the cognate lists than in the non-cognate
-list across ages. Critically, this difference persisted after both lists
-were matched in size, controlling their semantic category (i.e.,
+of translation equivalents; one made of 131 cognates, and one made of
+406 non-cognates. Across ages, the proportion of words that children
+were reportedly able to produce was higher in the cognate lists than in
+the non-cognate list. Critically, this difference persisted after both
+lists were matched in size, controlling their semantic category (i.e.,
 furniture, animals, food were similarity represented in both lists) and
 age-of-acquisition norms (an index of word difficulty). Taken together,
-the results of these two lines of research have been considered as
-supporting to the hypothesis that phonological similarity (as reflected
-in cognateness) plays a facilitation role in bilingual word acquisition.
+the results of these two lines of research support the hypothesis that
+phonological similarity (as reflected in cognateness) plays a
+facilitation role in bilingual word learning.
 
@@
 Spanish words, facilitating the naming of cognate pictures. More
-recntly, evidence of parallel activation has been reported in bilingual
+recently, evidence of parallel activation has been reported in bilingual
 toddlers and children too
@@
 in bilinguals' lexical processing and acquisition, previous studies do
-not address its influence on the learning trajectories of words. Results
-are aggregated across words and provide no information about the
-specific dynamics of how parallel activation influences word learning.
-This is the goal of the present research.
+not address its influence on the learning trajectories of individual
+words. Results are aggregated across words and provide no information
+about the specific dynamics of how parallel activation influences word
+learning. This is the goal of the present research.
 
@@
 of words from the lower-exposure language would benefit more strongly
-from their cognate status than words from the higher-exposure language.
-This asymmetric cross-language activation would be consistent with
-previous reports of larger priming effects from the dominant to the
-non-dominant language.
+from their cognate status than the acquisition of words from the
+higher-exposure language. This asymmetric cross-language activation
+would be consistent with previous reports of larger priming effects from
+the dominant to the non-dominant language (e.g.,
+\protect\hyperlink{ref-grainger1998masked}{Grainger, 1998}).
 
@@
 The case of the non-cognate translation equivalent
-/\textipa{"gos-"pe.ro}/ {[}\emph{dog}{]} would be different. Given the
-low phonological similarity between both word-forms, an exposure to
-\textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/
-leading to such exposure counting as a learning instance for
-/\textipa{"gos}/ (which the child was exposed to), but not for
-/\textipa{"pe.ro}/. While /\textipa{"gat}/--/\textipa{"ga.to}/ will
-benefit from phonological co-activation,
-/\textipa{"gos}-\textipa{"pe.ro}/ will not. If the child receives
-linguistic input from one of the languages more often than from the
-other, this effect might affect each form of the cognate translation
+/\textipa{"gos}/--/\textipa{"pe.ro}/ {[}\emph{dog}{]} would be
+different. Given the low phonological similarity between both
+word-forms, an exposure to \textipa{"gos} will result in a weak
+activation of /\textipa{"pe.ro}/ leading to such exposure counting as a
+learning instance for /\textipa{"gos}/ (which the child was exposed to),
+but not for /\textipa{"pe.ro}/. While
+/\textipa{"gat}/--/\textipa{"ga.to}/ will benefit from phonological
+co-activation, /\textipa{"gos}--\textipa{"pe.ro}/ will not. If the child
+receives linguistic input from one of the languages more often than from
+the other, this effect might affect each form of the cognate translation
 equivalent differently. For instance, if the child receives a larger
@@
 exposure to the word-form, and the cognate status of the word-form. As
-described in the methods section, rate of exposure was a composite
+described in the Methods section, rate of exposure was a composite
 measure taking into account participant' language exposure and word's
@@
 (\url{https://github.com/gongcastro/cognate-beginnings}) repositories.
+For reproducibility, a Docker image of the RStudio session is available
+on
+DockerHub(\url{https://hub.docker.com/repository/docker/gongcastro/cognate-beginnings/}).
 This study was conducted according to guidelines laid down in the
@@
 (\protect\hyperlink{ref-hamilton2000infant}{Hamilton et al., 2000}) and
-consisted in two lists of words: one in Catalan and one in Spanish. Both
+consisted of two lists of words, one in Catalan and one in Spanish. Both
 lists included items from a diverse sample of 26 semantic or functional
@@
 one possible translation in Spanish, and vice versa, therefore the
-unequal number of words included in the two lists (see
-Table~\ref{tbl-items} for a summary of the questionnaire items).
+unequal number of words included in the two lists.
 
@@
 \midrule
-N & 167 & 151 & 149 & 156 & -\\
+*N* & 167 & 151 & 149 & 156 & -\\
 \bottomrule
@@
 default). Given the large number of words in the vocabulary checklists,
-we created four different subsets of the complete list of items: A, B,
-C, and D. Each subset contained a random but representative sub-sample
+we created four different subsets of the complete list of items (A, B,
+C, and D) Each subset contained a random but representative sub-sample
 of the items from the complete list (see Table~\ref{tbl-items}).
 Semantic or functional categories with less than 16 items---thus
-resulting in less than four items after dividing it in four lists---were
-not divided: all of their items were included in the four lists. Items
-that were part of the trial lists of some ongoing experiments in the lab
-were also included in all versions. The resulting reduced list contained
-between 343 and 349 Catalan words, and between 349 and 371 Spanish
-words. Participants were randomly allocated into one of the four
-subsets.
+resulting in less than four items after dividing it in four
+subsets---were not divided: all of their items were included in the four
+subsets. Items that were part of the trial lists of some ongoing
+experiments in the lab were also included in all versions. The resulting
+reduced list contained between 343 and 349 Catalan words, and between
+349 and 371 Spanish words.
 
@@
 
-We collected 440 responses to the questionnaire from 369 distinct
-children from the Metropolitan Area of Barcelona between Marth 30th,
-2020 and October 31th, 2022: 314 of those participants participated
-once, 43 twice, 8 three times, and 4 four times. Recurrent participants
-provided responses with a minimum of 25 days between responses, and a
-maximum of 527 days, and were always allocated to the same questionnaire
-list (A, B, C, or D). Participants were part of the database of the
-Laboratori de Recerca en Infància (Universitat Pompeu Fabra) and were
-contacted by e-mail or phone if their child was aged between 12 and 32
-months, and had not been reported to be exposed more than 10\% of the
+We collected 436 questionnaire responses from 366 distinct children (175
+female, 179 male, 12 not reported, mainly White). Participants were aged
+12-32 months (\emph{M} = 22.23, \emph{SD} = 4.88, \emph{Range} =
+12.06-31.93). Of those participants, four participated four times, eight
+participated three times, 42 participated twice, and 312 participated
+once. Recurrent participants provided responses with a minimum of 25
+days between responses, and a maximum of 527 days. Participants were
+randomly allocated into one of the four questionnaire subsets (A, B, C,
+or D). Each participant was always allocated to the same subset.
+
+Participants were residents in the Metropolitan Area of Barcelona
+(Catalonia, Spain). Data collection took place between March 30th, 2020
+and October 31th, 2022. Participants were part of the database of the
+Laboratori de Recerca en Infància at the Universitat Pompeu Fabra, and
+were contacted by e-mail or phone if their child was aged between 12 and
+32 months, and had not been reported to be exposed more than 10\% of the
 time to a language other than Spanish or Catalan (see
 Table~\ref{tbl-participants} for a more detailed description of the
-sample). In total, 70 participants (15.91\%) participants were reported
+sample). In total, 70 participants (16.06\%) participants were reported
 to be exposed in less than 10\% to a third language other than Catalan
@@
 \midrule
-75-100\% & 18 & 23 & 37 & 38 & 20 & 7\\
+75-100\% & 18 & 23 & 36 & 38 & 20 & 7\\
 50-75\% & 8 & 13 & 30 & 41 & 18 & 1\\
-25-50\% & 10 & 18 & 46 & 30 & 17 & 0\\
+25-50\% & 10 & 17 & 45 & 29 & 17 & 0\\
 0-25\% & 7 & 11 & 21 & 17 & 8 & 1\\
 \midrule
-N & 43 & 65 & 134 & 126 & 63 & 9\\
+N & 43 & 64 & 132 & 125 & 63 & 9\\
 \bottomrule
+\multicolumn{7}{l}{\textsuperscript{} For participants exposed to Catalan and Spanish exclusively, the proportion of Catalan exposure shown in}\\
+\multicolumn{7}{l}{the table is complementary to the degree of exposue to Spanish. For participants exposed to a third}\\
+\multicolumn{7}{l}{languages this proportion is not complementary unless one adds the degree of exposed to the third}\\
+\multicolumn{7}{l}{language, which never exceeded 10\%. Participants with 100\%, 0\%, and 50\% Catalan exposure would be}\\
+\multicolumn{7}{l}{traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual,}\\
+\multicolumn{7}{l}{respectively.}\\
 \end{tabular}
@@
 \emph{educació universitària/educación universitaria} {[}university
-degree{]}. Most families reported university studies (358, 82\%),
-followed by families were the highest educational attainment were
-vocational studies (61, 14\%), secondary education (8, 2\%),
+degree{]}. Most families reported university studies (356, 82\%),
+followed by families where the highest educational attainment were
+vocational studies (59, 14\%), secondary education (8, 2\%),
 complementary studies (6, 1\%), primary education (1, \textless1\%), and
@@
 responses to nouns (628 items corresponding to other grammatical classes
-were excluded, see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
-et al., 2020} for a similar approach). We then excluded items with
-missing lexical frequency scores (\emph{n} = 269, see
-Section~\ref{sec-predictors}), items that included more than one lemma
-(e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} = 48), multi-word items
-or phrases (e.g., \emph{barrita de cereales} {[}cereal bar{]}, \emph{n}
-= 9). Finally, we removed items without a translation in the other
-language (\emph{n} = 32). This resulted in a final list of 604 items,
-corresponding to 302 Catalan words and their 302 Spanish translations
-(302 translation equivalents). After collecting participants' responses,
-the final dataset consisted of 139,326 observations, each corresponding
-to a single response of one participant to one item. Each translation
-equivalent received a median of 234 responses (\emph{Min} = 108,
-\emph{Max} = 880) from participants, both languages pooled together.
-Data processing and visualisation was done in R
-(\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
-version 4.2.2).
+were excluded). We then excluded items with missing lexical frequency
+scores (\emph{n} = 269, see Model predictors section), items that
+included more than one lemma (e.g., \emph{mono/mico} {[}monkey{]},
+\emph{n} = 48), multi-word items or phrases (e.g., \emph{barrita de
+cereales} {[}cereal bar{]}, \emph{n} = 9). Finally, we removed items
+without a translation in the other language (\emph{n} = 32). This
+resulted in a final list of 604 items, corresponding to 302 Catalan
+words and their 302 Spanish translations (302 translation equivalents).
+After collecting participants' responses, the final dataset consisted of
+138,078 observations, each corresponding to a single response of one
+participant to one item. Each translation equivalent received a median
+of 234 responses (\emph{Min} = 106, \emph{Max} = 872) from participants,
+both languages pooled together. Data processing and visualisation was
+done in R (\protect\hyperlink{ref-rcoreteam2013language}{R Core Team,
+2013}, version 4.2.2).
 
@@
 transcription of the word-form), \emph{Exposure} (a language
-exposure-weighted lexical frequency), and \emph{Cognateness} (defined as
-the phonological similarity between translation equivalents). We added
-these variables as main effects, together with the two-way and three-way
-interactions between \emph{Age}, \emph{Exposure}, and
-\emph{Cognateness}. Participant-level and item-level random intercepts
-and slopes were included where appropriate, according to the structure
-of the data (\protect\hyperlink{ref-barr2013random}{Barr et al., 2013}).
-We specified a weakly informative prior around the parameters of the
-model. Equation~\ref{eq-model} shows a detailed description of the
-model.
+exposure-weighted lexical frequency score), and \emph{Cognateness}
+(defined as the phonological similarity between translation
+equivalents). A more detailed descriptions of these predictors is
+provided in the Model preedictors section. We added these variables as
+main effects, together with the two-way and three-way interactions
+between \emph{Age}, \emph{Exposure}, and \emph{Cognateness}.
+Participant-level and item-level random intercepts and slopes were
+included where appropriate, according to the structure of the data
+(\protect\hyperlink{ref-barr2013random}{Barr et al., 2013}). We
+specified a weakly informative prior around the parameters of the model.
+Equation~\ref{eq-model} shows a detailed description of the model. See
+Appendix A for a detailed description of the model and its diagnostics.
 
@@
 We implemented the model using \texttt{brms}
-(\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), a R interface
+(\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), an R interface
 to the Stan probabilistic language (2.32.1)
@@
 
-We defined \emph{Cognateness} as the phonological similarity between
-each word-form and its translation. For each translation equivalent, we
-used the \texttt{stringdist}
+Following Floccia et al., we defined \emph{Cognateness} as the
+phonological similarity between each word-form and its translation. For
+each translation equivalent, we used the \texttt{stringdist}
 (\protect\hyperlink{ref-vanderloo2014stringdist}{Loo, 2014}) R package
@@
 We assessed the practical relevance of the estimated regression
-coefficients of the model model following Kruschke \& Liddell
+coefficients of the model following Kruschke \& Liddell
 (\protect\hyperlink{ref-kruschke2018bayesian}{2018}). First, we
@@
 a 80\% probability that the true value of the coefficient falls within
-the ROPE, and can therefore be considered equivalent to zero.
+the ROPE, and can therefore be considered equivalent to zero. See
+Appendix A for considerations about statistical power and sample size.
 
@@
 
+We now present our confirmatory analysis, and report a summary of the
+regression coefficients of interest. Table~\ref{tbl-coefs} shows the
+summary of the posterior distribution of the fixed regression
+coefficients, and their degree of overlap with the ROPE. For
+interpretability, we report the estimated regression coefficients
+transformed to the probability scale. The resulting values correspond to
+the maximum difference in probability of acquisition
+(\emph{Comprehension} or \emph{Comprehension and Production}) equivalent
+to a one standard deviation change in each predictor.
+
 \hypertarget{tbl-coefs}{}
@@
 \multicolumn{4}{l}{\textbf{Intercepts (at 22 months)}}\\
-\hspace{1em}Comprehension and Production & 0.435 & {}[0.383, 0.492] & -\\
-\hspace{1em}Comprehension & 0.936 & {}[0.921, 0.948] & -\\
+\hspace{1em}Comprehension and Production & 0.434 & {}[0.378, 0.489] & -\\
+\hspace{1em}Comprehension & 0.935 & {}[0.92, 0.948] & -\\
 \addlinespace[0.3em]
 \multicolumn{4}{l}{\textbf{Slopes (upper bound)}}\\
-\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.089, -0.038] & .000\\
-\hspace{1em}Age (+1 SD, 4.86 months) & 0.406 & {}[0.359, 0.455] & .000\\
-\hspace{1em}Exposure (+1 SD, 1.81) & 0.231 & {}[0.197, 0.265] & .000\\
-\hspace{1em}Cognateness (+1 SD, 0.26) & 0.056 & {}[0.007, 0.099] & .194\\
-\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.068, -0.046] & .000\\
-\hspace{1em}Age $\times$ Exposure & 0.072 & {}[0.037, 0.104] & .000\\
-\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0.001, 0.027] & .936\\
-\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.019 & {}[-0.027, -0.011] & .874\\
+\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.087, -0.036] & .000\\
+\hspace{1em}Age (+1 SD, 4.87 months) & 0.405 & {}[0.357, 0.451] & .000\\
+\hspace{1em}Exposure (+1 SD, 1.81) & 0.234 & {}[0.198, 0.264] & .000\\
+\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.016, 0.103] & .102\\
+\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.068, -0.045] & .000\\
+\hspace{1em}Age $\times$ Exposure & 0.073 & {}[0.039, 0.106] & .000\\
+\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0.001, 0.027] & .928\\
+\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.026, -0.01] & .908\\
 \bottomrule
+\multicolumn{4}{l}{\textsuperscript{} $\beta$: median of the posterior distribution in the probability scale. 95\% HDI: 95\% highest density}\\
+\multicolumn{4}{l}{interval of the distribution. $p(\text{ROPE})$: overlap between the 95\% HDI and the ROPE, indicating the}\\
+\multicolumn{4}{l}{posterior probability that the true value of the coefficient is equivalent to zero.}\\
 \end{tabular}
@@
 The coefficient of \emph{Age} showed the strongest association with the
-probability of acquisition (\(\beta\) = 0.406, 95\% HDI = {[}0.359,
-0.455{]}), with all posterior samples falling out of the ROPE. A
+probability of acquisition (\(\beta\) = 0.405, 95\% HDI = {[}0.357,
+0.451{]}), with all posterior samples falling out of the ROPE. A
 one-month increment in age increased a maximum of 0.08 the probability
@@
 (\emph{Exposure}) had a strong effect on the probability of acquisition
-(\(\beta\) = 0.231, 95\% HDI = {[}0.197, 0.265{]}). All of the posterior
+(\(\beta\) = 0.234, 95\% HDI = {[}0.198, 0.264{]}). All of the posterior
 samples of this regression coefficient excluded the ROPE. The impact of
 this predictor on the probability of acquisition was positive: for every
-standard deviation increase in exposure, the participant was 0.128
+standard deviation increase in exposure, the participant was 0.129
 probability points more likely to acquire it. Word-form length also
 showed a significant association with probability of acquisition
-(\(\beta\) = -0.062, 95\% HDI = {[}-0.089, -0.038{]}). For every phoneme
+(\(\beta\) = -0.062, 95\% HDI = {[}-0.087, -0.036{]}). For every phoneme
 in the word-form, participants were -0.04 probability points less likely
@@
 \(\times\) \emph{Exposure} interaction also excluded the ROPE (\(\beta\)
-= 0.072, 95\% HDI = {[}0.037, 0.104{]}), showing that the effect of the
+= 0.073, 95\% HDI = {[}0.039, 0.106{]}), showing that the effect of the
 word-form exposure index differed across ages: older children were more
 likely to acquire words with a higher exposure rate than younger
-children.
-
-Around 19.44\% of the posterior samples of the main effect of
-\emph{Cognateness} overlapped with the ROPE completely (\(\beta\) =
-0.056, 95\% HDI = {[}0.007, 0.099{]}). For every 10\% increment in
-cognateness, the probability of word acquisition increased in 0.005. The
-effect of \emph{Cognateness} interacted with that of \emph{Exposure}:
-the 95\% HDI of the regression coefficient of interaction excluded the
-ROPE entirely (\(\beta\) = -0.06, 95\% HDI = {[}-0.068, -0.046{]}),
-suggesting that the effect of cognateness on a word's probability of
-acquisition changed depending on participants' exposure to the
-word-form. When \emph{Exposure} was low (e.g., -1 SD),
-\emph{Cognateness} increased the probability of acquisition
-substantially, while this effect was negligible for words with median or
-high exposure (+1 SD) (see Figure~\ref{fig-marginal}).
+children were.
+
+Around 10.20\% of the posterior samples of the main effect of
+\emph{Cognateness} overlapped with the ROPE (\(\beta\) = 0.058, 95\% HDI
+= {[}0.016, 0.103{]}). For every 10\% increment in cognateness, the
+probability of word acquisition increased in 0.006. The effect of
+\emph{Cognateness} interacted with that of \emph{Exposure}: the 95\% HDI
+of the regression coefficient of interaction excluded the ROPE entirely
+(\(\beta\) = -0.057, 95\% HDI = {[}-0.068, -0.045{]}), suggesting that
+the effect of cognateness on a word's probability of acquisition changed
+depending on participants' exposure to the word-form. When
+\emph{Exposure} was low (e.g., -1 SD), \emph{Cognateness} increased the
+probability of acquisition substantially, while this effect was
+negligible for words with median or high exposure (+1 SD) (see
+Figure~\ref{fig-marginal}).
 
@@
 translation equivalents) were acquired earlier than non-cognates. This
-effect was mediated by exposure rate: low-exposure word-forms benefited
+effect was mediated by exposure rate. Low-exposure word-forms benefited
 from their cognate status, whereas high-exposure word-forms did not.
@@
 bilingual toddlers learning two typologically close languages
-(e.g.~shared many cognates, like., English-Dutch) showed larger
-vocabulary sizes than those learning typologically distant languages
-(e.g.~shared fewer cognates, like English-Mandarin). Second, Mitchell et
-al. (\protect\hyperlink{ref-mitchell2022cognates}{2022}) found an
-earlier age-of-acquisition for cognates, compared to non-cognates. The
-outcomes of both studies pointed to cognateness facilitating word
-acquisition through parallel activation. But the underpinnings of such
-effect were unclear: while parallel activation has been extensively
-described in experimental studies, current paradigms of bilingual word
-acquisition and word learning are, to a large extent, dissociated from
-the mechanisms proposed by previous work on word processing. The notion
-of \emph{accumulator}, as conceptualised by accumulator models of
-language acquisition, may provide a convenient theoretical framework to
-narrow this gap.
+(e.g.~shared many cognates, like English-Dutch) showed larger vocabulary
+sizes than those learning typologically distant languages (e.g., shared
+fewer cognates, like English-Mandarin). Second, Mitchell et al.
+(\protect\hyperlink{ref-mitchell2022cognates}{2022}) found an earlier
+age-of-acquisition for cognates, compared to non-cognates. The outcomes
+of both studies pointed to cognateness facilitating word acquisition
+through parallel activation, but the underpinnings of such effect were
+unclear. While parallel activation has been extensively described in
+experimental studies, current paradigms of bilingual word acquisition
+and word learning are, to a large extent, dissociated from the
+mechanisms proposed by previous work on word processing. The notion of
+\emph{accumulator}, as conceptualised by accumulator models of language
+acquisition, may provide a convenient theoretical framework to narrow
+this gap.
 
@@
 
-These predictions address a critical subject in bilingualism research:
-do bilingual infants accumulate learning experiences in both languages
+These predictions address a critical subject in bilingualism research.
+Do bilingual infants accumulate learning experiences in both languages
 independently, or does exposure to one language impact the acquisition
@@
 authors observed a facilitation in the additional (non-English)
-language: children learning two typologically close languages knew more
+language. Children learning two typologically close languages knew more
 words in the additional language than those learning two typologically
@@
 would lead to a faster increase in cumulative learning instances, and to
-earlier ages of acquisition for cognate translation equivalents (for
+an earlier age-of-acquisition for cognate translation equivalents (for
 which listening to each word-form contributes to the acquisition of its
@@
 Mitchell et al. (\protect\hyperlink{ref-mitchell2022cognates}{2022})
-discuss this possibility as a possible explanation of the cognate
+discuss this possibility as a candidate explanation of the cognate
 facilitation effect, in which bilinguals only need to map one word-form
@@
 of overlap between the phonemic inventories of the two languages being
-learned (e.g., low, like the case of English and Mandarin, or high like
+learned (e.g., low, like the case of English and Mandarin, or high, like
 the case of Spanish and Greek), to extralinguistic factors like the
-socio-linguistic situation in which the two languages co-exist (e.g., in
+sociolinguistic situation in which the two languages co-exist (e.g., in
 some regions both languages are co-official and used in similar
-contexts, while in others, one of the languages hardly has any societal
+contexts, while in others, one of the languages has a smaller societal
 presence, i.e., heritage languages). This diversity of situations in
@@
 research. Our sample, although homogeneous (e.g., similar parental
-educational level across), represents a particular bilingual
-sociolinguistic environment: the languages involved in the present
-investigation, Catalan and Spanish, co-exist in Catalonia as official
-languages, both languages are used in fairly similar contexts, and both
-languages are known by the majority of the population. In 2018, more
-than 81.2\% of a representative sample of 8,780 adults aged 15 years or
-older living in Catalonia reported being able to speak Catalan, and more
-than 99.5\% of the same population reported being able to speak Spanish
-(\protect\hyperlink{ref-2018els}{\emph{Els Usos Lingüístics de La
-Població de Catalunya}, 2018}). In addition, Catalan and Spanish are
+educational level across participants), represents a particular
+bilingual sociolinguistic environment. The languages involved in the
+present investigation, Catalan and Spanish, co-exist in Catalonia as
+official languages, both languages are used in fairly similar contexts,
+and both languages are known by the majority of the population. In 2018,
+more than 81.2\% of a representative sample of 8,780 adults aged 15
+years or older living in Catalonia reported being able to speak Catalan,
+and more than 99.5\% of the same population reported being able to speak
+Spanish (\protect\hyperlink{ref-2018els}{\emph{Els Usos Lingüístics de
+La Població de Catalunya}, 2018}). In addition, Catalan and Spanish are
 Romance languages and share a considerable amount of cognates. Extending
@@
 that during the acquisition of low frequency words, bilingual children
-seem to benefit more strongly from the word-forms's phonological
+seem to benefit more strongly from the word-form's phonological
 similarity with its translation in the other language. Capitalising on

### 2023-10-03 tests/testthat/test-participants.R

@@ -7,6 +7,7 @@ test_participants <- function(participants) {
                 "time_stamp",
                 "list",
                 "age",
+                "sex",
                 "lp",
                 "doe_catalan",
                 "doe_spanish",
@@ -22,6 +23,7 @@ test_participants <- function(participants) {
         expect_type(participants$list, "character")
         expect_type(participants$age, "double")
         expect_type(participants$lp, "character")
+        expect_type(participants$sex, "character")
         expect_type(participants$doe_catalan, "double")
         expect_type(participants$doe_spanish, "double")
         expect_type(participants$edu_parent, "integer")

### 2023-10-03 manuscript/manuscript.qmd
Fix table caption

@@
     } else {
-        "*N*"
+        "N"
     }
@@
     summarise(
-        semantic_category = "*N*",
+        semantic_category = "N",
         across(a:d, sum),

### 2023-10-03 manuscript/manuscript.tex
Fix table caption

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 \midrule
-*N* & 167 & 151 & 149 & 156 & -\\
+N & 167 & 151 & 149 & 156 & -\\
 \bottomrule

### 2024-01-15 manuscript/manuscript.tex
Backup

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
 

### 2024-01-15 R/models.R

@@ -99,11 +99,11 @@ get_posterior_summary <- function(model,
             .value, .lower, .upper, .rope
         ) |>
         ungroup()
+    
+    save_files(posterior_summary, folder = "results/posterior", formats = "csv")
 
     return(posterior_summary)
 }
-
-
 
 #' Get model R-hat scores
 #'
@@ -117,11 +117,15 @@ get_posterior_summary <- function(model,
 #' [bayesplot::rhat()].
 #' * .neff: effective sample size, as returned by [bayesplot::neff_ratio()]
 get_model_convergence <- function(model, ...) {
-    tibble::tibble(
+    out <- tibble::tibble(
         .variable = variables(model),
         .rhat = rhat(model, ...),
         .neff = neff_ratio(model, ...)
     )
+    
+    save_files(out, folder = "results/posterior", formats = "csv")
+    
+    return(out)
 }
 
 #' Get model posterior predictive checks

### 2024-01-15 manuscript/manuscript.tex

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
 
